### 2020-06-28 skills/math/linear_algebra_summary.tex
linear_algebra

\documentclass[11pt]{article}

\usepackage{fancybox}

\usepackage{color}
\usepackage{url}
\usepackage[margin=1in]{geometry}

\setlength\parindent{0pt}
\renewcommand{\textfraction}{0.0}
\renewcommand{\topfraction}{1.0}
%\renewcommand{\textfloatsep}{5mm}

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
%\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\mathtoolsset{showonlyrefs=true}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\setcounter{figure}{0}   
\setcounter{table}{0}

\newcommand{\cmt}[1]{{\leavevmode\color{red}{#1}}}

\usepackage{dsfont}
\usepackage{multirow}

\DeclareMathOperator*{\minimize}{minimize}

\usepackage{mathtools}
\mathtoolsset{showonlyrefs}
\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}

\input macros.tex

\title{\textbf{Linear Algebra}\\A summary for MIT 18.06SC}

\date{\today}
\author{%
Jiaxin Hu
}

\begin{document}

% Makes the title and author information appear.

\maketitle

% Abstracts are required.
\section{Matrices \& Spaces}
\subsection{Basic concepts}
\begin{itemize}
    \item Given vectors $v_1 ,...,v_n$ and scalars $c_1, ..., c_n$, the sum $c_1 v_1 + \dots + c_n v_n$ is called the \textit{linear combination} of $v_1,...,v_n$.
    \item The vectors  $v_1 ,...,v_n$  are \textit{linearly independent}(or just \textit{indepedent}) if $c_1 v_1 + \dots + c_n v_n = 0$ holds only when all  $c_1,...,c_n = 0$. If the vectors  $v_1 ,...,v_n$  are \textit{dependent}, there exist scalars $c_1, ..., c_n$ which are not all equal to 0 and satisfy $c_1 v_1 + \dots + c_n v_n = 0$.
	\item Given a matrix $\mA \in \bbR^{m \times n}$ and a vector $x \in \bbR^{n}$,  the multiplication $\mA x$ is a linear combination of the columns of $\mA$ and $x^T \mA$ is a linear combination of the rows of $\mA$.
	\item Matrix multiplication is not communicative, i.e. $\mA \mB \neq \mB \mA.$ 
	\item Suppose $\mA$ is a square matrix. The matrix $\mA$ is \textit{invertible} or \textit{non-singular} if there exists a $\mA^{-1}$ such that $\mA^{-1} \mA = \mA \mA^{-1} = I$. Otherwise, the matrix $\mA$ is singular, i.e. its determinant is 0 and does not have inverse matrix. 
	\item The inverse of a matrix product $\mA\mB$ is $(\mA \mB)^{-1} = \mB^{-1} \mA^{-1}$. The product of invertible matrices is still invertible.
	\item The transpose of a matrix product  $\mA\mB$ is $(\mA \mB)^{T} = \mB^{T} \mA^{T}$. For any invertible matrix $\mA$, $(\mA^T)^{-1} = (\mA^{-1})^T$.
	\item A matrix $\mQ$ is orthogonal if $\mQ^T = \mQ^{-1}$. A matrix $\mQ$ is unitary if $\mQ^* = \mQ^{-1}$, where $\mQ^*$ is the conjugate transpose of $\mQ$.
\end{itemize}

\subsection{Permutation of matrices}
For any matrix $\mA$, we can swap its rows by multiplying a \textit{permutation matrix } $\mP$ on the left of $\mA$. For example,
\begin{align}
	\mP \mA = \begin{bmatrix}
		0 &0&1\\1 & 0 & 0\\ 0 & 1 & 0 
	\end{bmatrix}  \begin{bmatrix}
		a_1\\a_2\\ a_3
	\end{bmatrix}  = \begin{bmatrix}
		a_3\\a_1\\ a_2
	\end{bmatrix}
\end{align}
where $a_k$ refers to the $k$-th row of $\mA$. The inverse of permutation matrix $\mP$ is $\mP^{-1} = \mP^T$, which implies the orthogonality of permutation matrix. For an $n \times m$ matrix, there are $n!$ different row permutation matrix and these permutation matrices form a \textit{multiplicative group}.

\vspace{.2cm}
Similarly, we can also swap the columns of the matrix $\mA$ by multiplying a permutation matrix on the right of $\mA$.

\subsection{Elimination of matrices}
Elimination is an important technique in linear algebra. We eliminate the matrix by multiplications and subtractions. Take a 3-by-3 matrix $\mA$ as example.
\begin{align}
	\mA = \begin{bmatrix}
		1 &2&1\\3 & 8 & 1\\ 0 & 4 & 1 
	\end{bmatrix}
	\xrightarrow{\text{step 1}} \begin{bmatrix}
		1 &2&1\\0 & 2 & -2\\ 0 & 4 & 1 
	\end{bmatrix}
	\xrightarrow{\text{step 2}} \mU =  \begin{bmatrix}
		1 &2&1\\0 & 2 & -2\\ 0 & 0 & 5
	\end{bmatrix}
\end{align}
In step 1, we choose the number 1 in row 1 column 1 as a \textit{pivot}, then we recopy the first row and multiply an appropriate number(in this case, 3) and subtract those values from the numbers in the second row. We have thus eliminated 3 in row 2 column 1. Similarly, in step 2, we choose 2 in row 2 column 2 as a pivot and eliminate the number 4 in row 3 column 2.  The number 5 in row 3 column 3 is also a pivot. The matrix $\mU$ is an upper traingular matrix.

\vspace{.2cm}
The \textit{elimination matrix} used to eliminate the entry in row $m$ column $n$ is denoted $\mE_{mn}$. In previous example, 
\begin{align}
\mE_{21} \mA = 
\begin{bmatrix}
		1 &0&0\\-3 & 1 & 0\\ 0 & 0 & 1 
	\end{bmatrix}
	\begin{bmatrix}
		1 &2&1\\3 & 8 & 1\\ 0 & 4 & 1 
	\end{bmatrix} = \begin{bmatrix}
		1 &2&1\\0 & 2 & -2\\ 0 & 4 & 1 
	\end{bmatrix};\quad \mE_{32} (\mE_{21} \mA) = \mU.
\end{align}

Pivots can not be 0. If there is a 0 in the pivot position, we must exchange the row with one below to get a non-zero value in pivot position. If there is not non-zero value below the 0 pivot, then we skip this column and find a pivot in next column.

\vspace{.2cm}
Since matrix multiplication is associative, we can write $\mE_{32} (\mE_{21} \mA) = (\mE_{32} \mE_{21} ) \mA = \mU$. Let $\mE$ denote the product of all elimination matrices. If we need to permute the rows during the process, we  have $\mE \mP \mA = \mU$, where $\mP$ is the product of all the needed permutation matrices.

\vspace{.2cm}
We also prove the invertibility of the elimination matrix.
\begin{lem}[Invertiblity of elimination matrix]\label{lem:elim}
 Suppose there is an elimination matrix $\mE_{ij} \in \bbR^{n \times n}$ that means multiplying a scalar $-c$ to the  $j$-th row and subtracting the row from $i$-th row, where $i \neq j$. Then, $\mE_{ij}$ is invertible.
\end{lem}
\begin{proof}
	The elimination matrix can be written as:
	\[ \mE_{ij} = \mI_n + c e_i e_{j}^T, \]
	where $e_i \in \bbR^n$ is identity vector with value 1 on the $i$-th entry and value 0 on the other entries. Note that $e^T_i e_{j} = 0$ because $i \neq j$. Therefore, we have 
	\[ (\mI_n + c e_i e_{j}^T)(\mI_n - c e_i e_{j}^T)  = \mI_n - c^2 e_i e_{j}^T e_i e_{j}^T = \mI_n; \quad (\mI_n - c e_i e_{j}^T)(\mI_n + c e_i e_{j}^T)= \mI_n \]
	Thus, $\mI_n - c e_i e_{j}^T$ is the inverse of $\mE_{ij}$ and $\mE_{ij}$  is invertible.
\end{proof}

\subsection{Gauss-Jordan Elimination}
Consider an invertible matrix $\mA \in \bbR^{n \times n}$, one of the effective ways to find the inverse of $\mA$ uses elimination.

\vspace{0.2cm}
 The inverse of $\mA$, $\mA^{-1}$ , satisfies $\mA \mA^{-1} = \mI_n$. Suppose there is an elimination $\mE$ such that $\mE \mA = \mI_n$. Multiplying $\mE$ on the both side of the equation, we have $\mE  \mA \mA^{-1}= \mA^{-1} = \mE$. To obtain an such $\mE$, we do elimination to the \textit{augmented matrix} $[\mA|\mI_n]$ until $\mA$ becomes $\mI_n$.

 \vspace{0.2cm}
 We call the elimination process of finding $\mE$ as \textit{Gauss-Jordan Elimination}.

\subsection{Factorization of matrices}
By elimination, for any square matrix $\mA$, we have $\mE \mP \mA = \mU$ where $\mU$ is an upper triangular matrix. By "canceling" the elimination matrix $\mE$, we get $\mP \mA = \mE^{-1} \mU$. Because $\mE$ is invertible, the inverse $\mE^{-1}$ exists. Note that $\mE$ is lower triangular matrix, the inverse of $\mE$ is also a lower triangular matrix. We use $\mL$ to denote $\mE^{-1}$. Therefore, we can decompose an arbitrary square matrix $\mA$ as:
\[ \mP \mA = \mL \mU, \]
where $\mU$ is an upper triangular matrix with pivots on the diagonal, $\mL$ is lower triangular matrix with ones on the diagonal, and $\mP$ is a permutation matrix. Note that, there may exist other ways to decompose $\mP\mA$ into $\mL \mU$, where $\mU$ and $\mL$ have different settings.

\subsection{Time complexity of elimination}
For an $n$-by-$n$ matrix, multiplying one row and then subtracting it from another row require $2n$ operations. There are $n$ rows in the matrix, so the total number of operations used in elimination the first column is $2n^2$. The second row and column are shorter, which may cost $2(n-1)^2$ operations  and so on. Therefore, the time complexity to factorize $\mA$ into $\mL \mU$ is about $\tO(n^3)$: \[ 1^2 + 2^2 + \cdots +(n-1)^2 + n^2 = \sum_{i}^n i^2 \approx \int_{0}^n x^2 dx = \frac{1}{3} n^3.\]

\subsection{Reduced row echelon form of matrices}
By continuing to use the method of elimination, we can convert $\mU$ to a matrix $\mR$ in reduced row echelon form, with pivots equal to 1 and zeros above and below the pivots. The matrix $\mR$ is called the \textit{reduced row echelon form of matrices}(RREF) of $\mA$. In previous example,
\begin{align}
	\mU = \begin{bmatrix}
		1 &2&1\\0 & 2 & -2\\ 0 & 0 & 5
	\end{bmatrix} \xrightarrow{\text{make pivots = 1}} \begin{bmatrix}
		1 &2&1\\0 & 1 & -1\\ 0 & 0 & 1
	\end{bmatrix} 
	\xrightarrow{\text{0 above and below pivots}} \mR =  \begin{bmatrix}
		1 &0&0\\0 & 1 & 0\\ 0 & 0 & 1
	\end{bmatrix} .
\end{align}

For an another example,
\begin{align}
	\mU = \begin{bmatrix}
		1 &2&2&2\\0 &0 & 2 & 4\\ 0 & 0 & 0&0
	\end{bmatrix} \xrightarrow{\text{make pivots = 1}} \begin{bmatrix}
		1 &2&2&2\\0 &0 & 1 & 2\\ 0 & 0 & 0&0
	\end{bmatrix} 
	\xrightarrow{\text{0 above and below pivots}} \mR =  \begin{bmatrix}
		1 &2&0&-2\\0 &0 & 1 & 2\\ 0 & 0 & 0&0
	\end{bmatrix} .
\end{align}

With proper permutation, the matrix $\mR$ can be written in form $[\mI \quad \mF]$, $\begin{bmatrix} \mI&\mF\\0 &0	\end{bmatrix}$, $\begin{bmatrix} \mI \\0	\end{bmatrix}$ , or just $\mI$, where $\mF$ can be arbitrary matrix in proper dimension. The columns in $\mA$ which correspond to the identity matrix $\mI$ are called \textit{pivot columns} and the other columns are \textit{free columns}.

\subsection{Vector space, Subspace and Column space}
\begin{itemize}
	\item \textit{Vector space} is a collection of vectors that are closed under linear combination(addition and multiplication by any real number); i.e. for any vectors in the collection, all the combinations of these vectors are still in the collection.
	\item \textit{Subspaces of the vector space} is a vector space that is contained inside of another vector space.
\end{itemize}

Note that any vector space or subspace must include an origin. For a vector space $\tA$, the subspace of $\tA$ can be $\tA$ itself or can only contain a zero vector.

\begin{itemize}
	\item Vectors $v_1 ,...,v_n$ \textit{span} a space that consists all the combination of those vectors. 
	\item \textit{Column space} of matrix $\mA$ is the space spanned by the columns of $\mA$. Let $C(\mA)$ denote the column space of $\mA$.
\end{itemize}

Note that if  $v_1 ,...,v_n$ span a space $\tS$, then $\tS$ is the smallest space that contain those vectors.

\begin{itemize}
	\item \textit{Basis} of a vector space is a sequence of vectors $v_1 ,...,v_n$ satisfies: (1) $v_1 ,...,v_n$ are independent; (2) $v_1 ,...,v_n$ span the space.
	\item \textit{Dimension} of the space is the number of vectors in a basis of the space.
\end{itemize}

\subsection{Matrix rank}
The \textit{rank} of a matrix $\mA$ is defined as the dimension of the columns space of $\mA$. Rank is also equal to the number of pivot columns of $\mA$. That means:
\[ \text{rank}(\mA) = \# \text{ of pivot columns of }\mA = \text{dimension of } C(\mA).\]
We use $r$ to denote the rank of $\mA$. If $\mA\in \bbR^{m \times n}$, then we have $r \leq m, r \leq n$. We say the matrix is \textit{full rank} if $r = n$ or $r = m$.

\vspace{.2cm}
The rank of a square matrix is closely related to its invertibility. 
\begin{lem}[Full rankness and invertibility]\label{lem:fullrank} A square matrix $\mA \in \bbR^{n \times n}$ is full rank, if and only if $\mA$ is an invertible matrix.
\end{lem}

\begin{proof}
	First, assume $\mA$ is full rank, we prove that $\mA$ has an inverse.
	We can get find a RREF($\mA$) by eliminations and permutations. There are $\mE$ and $\mP$ such that
	\[ \mE \mP \mA =  \mR. \]
	Since $\mA$ is full rank, $\mA$ have $n$ pivots columns and that implies $\mR = \mI_n$. By lemma \ref{lem:elim}, $\mE$ is invertible. The permutation matrix $\mP$ is also invertible. Therefore, $\mE \mP$ is invertible and $\mA \mE \mP= \mI_n$. This implies $\mA$ is invertible.

	Second, assume $\mA$ has an inverse, we prove that $\mA$ is full rank. 
	We show this by contradiction. Assume $\mA$ has an inverse $\mA^{-1}$ and $\mA$ is not full rank. Because the rank of $\mA$ is equal to the dimension of $C(\mA)$, the columns of $\mA$ are linearly dependent without full rankness. So, there exist a non-zero vector $v$ such that \[ \mA v = 0.\] Multiplying $\mA^{-1}$ on the both sides of the equation, we have \[ v = \mA^{-1} 0 = 0.\] However, it contradicts to  non-zeroness of $v$. Therefore, $\mA$ must be full rank.
\end{proof}

The rank of $\mA$ also effects the number of solutions to the system $\mA x = b$. We will discuss it in next section.

\section{Solving $Ax = b$}

Here we discuss the solution situation of the linear system $Ax = b$. Without specific explanation, the matrix $\mA \in \bbR^{m \times n}$ and the vector $x \in \bbR^n$, $b \in \bbR^m$.

\subsection{Solving $Ax = 0$: Nullspace}
The \textit{nullspace} of matrix $\mA$ is the collection of all solutions $x$ to the system $\mA x = 0$. Let $N(\mA)$ denote the nullspace of $\mA$.

\vspace{.2cm}

\begin{lem}[Nullspace] 
The \textit{nullspace} of matrix $\mA$ is a vector space. 
\end{lem}

\begin{proof}
To show the $N(\mA)$ is a vector space, we need to show $N(\mA)$ is close to linear combination. For any integer $k$, take arbitrary vectors $v_1 ,...,v_k \in N(\mA)$ and arbitrary scalars $c_1, ..., c_k$. We have,
\[ \mA(c_1 v_1 + \cdots + c_k v_k) = c_1 \mA v_1 + \cdots + c_k \mA v_k = 0. \]
Therefore, the linear combination $(c_1 v_1 + \cdots + c_k v_k) \in N(\mA)$. Then $N(\mA)$ is a vector space.
\end{proof}

\begin{lem}[The rank of nullspace]\label{lem:ranknull}
Suppose the rank of $\mA$ is $r$, then the rank of $N(\mA)$ is $n - r$.
\end{lem}

\begin{proof} Let $\mR$ denote the RREF($\mA$). The matrix $\mR$ can be written in form $\mR = \begin{bmatrix} \mI_r & \mF \\ 0 & 0	\end{bmatrix}$ where $\mF \in \bbR^{r \times (n-r)}$ and $0$ are zero matrices with proper dimensions.
Let $\mX = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix}$. Then we have 
\begin{align}
	\mR \mX = \begin{bmatrix} \mI_r & \mF \\ 0 & 0	\end{bmatrix} \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix} = 0
\end{align}
Therefore, each column $\mX$ is a special solution to the system $\mA x = 0$. Next, we want to show other solutions to $\mA x = 0$ are linear combinations of those special solutions.

Suppose there is a $x = (x_1, x_2) \in N(\mA)$. Then 
\begin{align}
	\mR x = \begin{bmatrix} \mI_r & \mF \\ 0 & 0	\end{bmatrix} \begin{bmatrix} x_1 \\ x_2	\end{bmatrix} = \begin{bmatrix} x_1 + \mF x_2 \\  0 &\end{bmatrix} = 0.
\end{align}
That implies $x_1  = -\mF x_2$ and $x = \begin{bmatrix} x_1 \\ x_2	\end{bmatrix} = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix} x_2 = \mX x_2$. Therefore, arbitrary $x \in N(\mA)$ is a linear combination of special solutions, i.e. $C(\mX) = N(\mA)$.

Since $\mX$ contains an identity matrix, the special solutions are independent and the rank of $C(\mX)$ is $n -r$. Therefore, the rank of $N(\mA)$ is $n-r$. 
\end{proof}

Recall that the columns in $\mA$ correspond to the $\mI_r$ in $\mR$ are called pivot columns and other columns are free columns. In $\mA x = b$, the variables in $x$ that correspond to pivot columns are called \textit{pivot variables} and others are \textit{free variables}. As the proof shows, the way to find special solutions of $\mA x = 0$ is to let one of the free variables is 1 while other free variables are 0 and then solve the equation $\mA x = 0$. There are total $n-r$ special solutions.

\subsection{Solving $Ax = b$: complete solutions}

\begin{lem}[Solvability of $\mA x = b$]\label{lem:solve} The system $\mA x = b$ is solvable only when $b \in C(\mA)$.	
\end{lem}

\begin{proof}
	For any $x$, $\mA x \in C(\mA)$. Therefore, to satisfy $\mA x = b$, $b \in C(\mA)$. 
\end{proof}

\begin{lem}[Complete solution]\label{lem:compsolu} The complete solution of $\mA x = b$ is given by $x_{comp} = x_p + x_n$, where $x_p$ is a particular solution that $\mA x_p = b$ and $x_n \in N(\mA)$. 
\end{lem}

\begin{proof}
	Suppose $x = x_p + x_0$ is an arbitrary solution to $\mA x= b$. Then we have 
	\begin{align}
		\mA x - \mA x_p = \mA (x- x_p) = \mA x_0 = 0.
	\end{align}
	Then $x_0 \in N(\mA)$.
\end{proof}

One way to find a particular solution is to let all the free variables be 0 and solve the equations.

\vspace{.2cm}
Here is a summary table that discusses the rank of $\mA$, the form of $\mR$ and the situation about the solutions.

\begin{center}
\begin{tabular}{ c|c |c| c |c } 
 \hline
  & $r = m=n$ & $r = n < m$ & $r = m < n$ & $r<m,r<n$ \\ \hline 
  $\mR$& $\mI$ & $\begin{bmatrix} \mI \\ 0	\end{bmatrix}$ & $\begin{bmatrix} \mI & \mF \end{bmatrix}$ & $\begin{bmatrix} \mI & \mF \\ 0 &0	\end{bmatrix}$\\
  dimension of $N(\mA)$& 0 &0& $n -r$ & $n-r$ \\
  \# solutions to $\mA X = b$ & 1 & 0 or 1 & infinitely many & 0 or infinitely many\\
 \hline
\end{tabular}
\end{center}

\vspace{.2cm}

\end{document}

### 2020-06-30 skills/math/linear_algebra_summary.tex
update

@@
 
-\mathtoolsset{showonlyrefs=true}
-
 \theoremstyle{plain}
@@
 
-\usepackage{mathtools}
 \mathtoolsset{showonlyrefs}
@@
 \begin{itemize}
-    \item Given vectors $v_1 ,...,v_n$ and scalars $c_1, ..., c_n$, the sum $c_1 v_1 + \dots + c_n v_n$ is called the \textit{linear combination} of $v_1,...,v_n$.
-    \item The vectors  $v_1 ,...,v_n$  are \textit{linearly independent}(or just \textit{indepedent}) if $c_1 v_1 + \dots + c_n v_n = 0$ holds only when all  $c_1,...,c_n = 0$.
-If the vectors  $v_1 ,...,v_n$  are \textit{dependent}, there exist scalars $c_1, ..., c_n$ which are not all equal to 0 and satisfy $c_1 v_1 + \dots + c_n v_n = 0$.
+    \item Given vectors $v_1 ,...,v_n$ and scalars $c_1, ..., c_n$, the sum $c_1 v_1 + \dots + c_n v_n$ is called a \textit{linear combination} of $v_1,...,v_n$.
+    \item The vectors  $v_1 ,...,v_n$  are \textit{linearly independent} (or just \textit{indepedent}) if $c_1 v_1 + \dots + c_n v_n = 0$ holds only when all  $c_1,...,c_n = 0$.
+If the vectors  $v_1 ,...,v_n$  are \textit{dependent}, there exist scalars $c_1, ..., c_n$ which are not all equal to 0 such that $c_1 v_1 + \dots + c_n v_n = 0$.
 	\item Given a matrix $\mA \in \bbR^{m \times n}$ and a vector $x \in \bbR^{n}$,  the multiplication $\mA x$ is a linear combination of the columns of $\mA$ and $x^T \mA$ is a linear combination of the rows of $\mA$.
-	\item Matrix multiplication is not communicative, i.e. $\mA \mB \neq \mB \mA.$ 
+	\item Matrix multiplication is typically not communicative, i.e. $\mA \mB \neq \mB \mA$. Lemma~\ref{lem:matrixcomm} describes a special case where matrix multiplication is communicative.
 	\item Suppose $\mA$ is a square matrix.
 The matrix $\mA$ is \textit{invertible} or \textit{non-singular} if there exists a $\mA^{-1}$ such that $\mA^{-1} \mA = \mA \mA^{-1} = I$.
-Otherwise, the matrix $\mA$ is singular, i.e.
-its determinant is 0 and does not have inverse matrix.
+Otherwise, the matrix $\mA$ is \textit{singular}, and the determinant of $\mA$ is 0.
 	\item The inverse of a matrix product $\mA\mB$ is $(\mA \mB)^{-1} = \mB^{-1} \mA^{-1}$. The product of invertible matrices is still invertible.
@@
 \end{itemize}
+
+\begin{lem}[Communicative matrix multiplication]\label{lem:matrixcomm} For matrices $\mA \in \bbR^{n \times n}$ and $\mB \in \bbR^{n \times n}$, the matrix multiplication of $\mA$ and $\mB$ is communicative, i.e.
+$\mA\mB  = \mB \mA$, if all the normalized eigenvectors of $\mA$ and $\mB$ are identical.
+\end{lem}
+
+\begin{proof}
+	If all the normalized eigenvectors of $\mA$ and $\mB$ are identical, there exists a matrix $\mQ \in \bbR^{n \times n}$ such that
+	\[ \mA = \mQ \mD_A \mQ^{-1}, \quad \mB = \mQ \mD_B \mQ^{-1},\]
+	where columns of $\mQ$ are normalized eigenvectors of $\mA$ and $\mB$, and $\mD_A\in\bbR^{n \times n},  \mD_B\in\bbR^{n \times n} $ are diagonal matrices whose diagonal elements are corresponding eigenvalues of $\mA$ and $\mB$.
+Because matrix multiplication is communicative for two diagonal matrices with same dimensions, we have
+	\[  \mA \mB = \mQ \mD_A \mQ^{-1} \mQ \mD_B \mQ^{-1} = \mQ \mD_A \mD_B \mQ^{-1}  =\mQ \mD_B \mD_A \mQ^{-1} = \mQ \mD_B \mQ^{-1} \mQ \mD_A \mQ^{-1}  = \mB \mA .\]
+	Therefore, the matrix multiplication of $\mA$ and $\mB$ is communicative.
+\end{proof}
 
@@
 The inverse of permutation matrix $\mP$ is $\mP^{-1} = \mP^T$, which implies the orthogonality of permutation matrix.
-For an $n \times m$ matrix, there are $n!$ different row permutation matrix and these permutation matrices form a \textit{multiplicative group}.
+For an $n \times m$ matrix, there are $n!$ different row permutation matrices, which form a \textit{multiplicative group}.
 
@@
 \subsection{Elimination of matrices}
-Elimination is an important technique in linear algebra. We eliminate the matrix by multiplications and subtractions. Take a 3-by-3 matrix $\mA$ as example.
+Elimination is an important technique in linear algebra. We eliminate the matrix by multiplications and subtractions. Take a 3-by-3 matrix $\mA$ as an example.
 \begin{align}
@@
 \end{align}
-In step 1, we choose the number 1 in row 1 column 1 as a \textit{pivot}, then we recopy the first row and multiply an appropriate number(in this case, 3) and subtract those values from the numbers in the second row.
+In step 1, we choose the number 1 in row 1 column 1 as a \textit{pivot}, then we recopy the first row, multiply an appropriate number (in this case, 3), and subtract those values from the numbers in the second row.
 We have thus eliminated 3 in row 2 column 1.
-Similarly, in step 2, we choose 2 in row 2 column 2 as a pivot and eliminate the number 4 in row 3 column 2.
+Similarly, in step 2, we choose 2 in row 2 column 2 as a pivot, and eliminate the number 4 in row 3 column 2.
 The number 5 in row 3 column 3 is also a pivot.
@@
 \vspace{.2cm}
-The \textit{elimination matrix} used to eliminate the entry in row $m$ column $n$ is denoted $\mE_{mn}$. In previous example, 
+The \textit{elimination matrix} used to eliminate the entry in row $m$ column $n$ is denoted $\mE_{mn}$. In the previous example, 
 \begin{align}
@@
 
-Pivots can not be 0.
+Pivots are non-zero.
 If there is a 0 in the pivot position, we must exchange the row with one below to get a non-zero value in pivot position.
-If there is not non-zero value below the 0 pivot, then we skip this column and find a pivot in next column.
+If there is not non-zero value below the 0 pivot, we skip this column, and find a pivot in next column.
 
 \vspace{.2cm}
-Since matrix multiplication is associative, we can write $\mE_{32} (\mE_{21} \mA) = (\mE_{32} \mE_{21} ) \mA = \mU$.
+We write $\mE_{32} (\mE_{21} \mA) = (\mE_{32} \mE_{21} ) \mA = \mU$ because matrix multiplication is associative.
 Let $\mE$ denote the product of all elimination matrices.
-If we need to permute the rows during the process, we  have $\mE \mP \mA = \mU$, where $\mP$ is the product of all the needed permutation matrices.
+If we need to permute the rows during the process, we multiply a permutation matrix on the left of $\mA$.
+Therefore, the elimination process of $\mA$ is
+\begin{equation}\label{eq:factor}
+	\mE \mP \mA = \mU,
+\end{equation}
+ where  $\mU$ is an upper triangular matrix.
 
 \vspace{.2cm}
-We also prove the invertibility of the elimination matrix.
+Next, we prove the invertibility of the elimination matrix.
 \begin{lem}[Invertiblity of elimination matrix]\label{lem:elim}
- Suppose there is an elimination matrix $\mE_{ij} \in \bbR^{n \times n}$ that means multiplying a scalar $-c$ to the  $j$-th row and subtracting the row from $i$-th row, where $i \neq j$.
-Then, $\mE_{ij}$ is invertible.
+ Suppose there is an elimination matrix $\mE_{ij} \in \bbR^{n \times n}$ that multiplies a scalar $-c$ to the  $j$-th row, and subtracts the row from $i$-th row, where $i \neq j$.
+The matrix $\mE_{ij}$ is invertible.
 \end{lem}
 \begin{proof}
-	The elimination matrix can be written as:
+	We write the elimination matrix as
 	\[ \mE_{ij} = \mI_n + c e_i e_{j}^T, \]
-	where $e_i \in \bbR^n$ is identity vector with value 1 on the $i$-th entry and value 0 on the other entries. Note that $e^T_i e_{j} = 0$ because $i \neq j$. Therefore, we have 
-	\[ (\mI_n + c e_i e_{j}^T)(\mI_n - c e_i e_{j}^T)  = \mI_n - c^2 e_i e_{j}^T e_i e_{j}^T = \mI_n; \quad (\mI_n - c e_i e_{j}^T)(\mI_n + c e_i e_{j}^T)= \mI_n \]
-	Thus, $\mI_n - c e_i e_{j}^T$ is the inverse of $\mE_{ij}$ and $\mE_{ij}$  is invertible.
+	where $e_i \in \bbR^n$ denotes the vector with value 1 in the $i$-th entry and value 0 elsewhere. Note that $e^T_i e_{j} = 0$ because $i \neq j$. We have 
+	\[ (\mI_n + c e_i e_{j}^T)(\mI_n - c e_i e_{j}^T)  = \mI_n - c^2 e_i e_{j}^T e_i e_{j}^T = \mI_n; \quad (\mI_n - c e_i e_{j}^T)(\mI_n + c e_i e_{j}^T)= \mI_n. \]
+	Therefore, $\mI_n - c e_i e_{j}^T$ is the inverse of $\mE_{ij}$. The elimination matrix $\mE_{ij}$  is invertible.
+\end{proof}
+
+\begin{cor}[Inverse of elimination matrix]\label{cor:invelim} 
+Suppose the elimination matrix $\mE_{ij}$ in lemma~\ref{lem:elim} is a lower/upper-triangular matrix. The inverse $\mE_{ij}^{-1}$ is also a lower/upper-triangular matrix.
+\end{cor}
+\begin{proof}
+	By the proof of lemma~\ref{lem:elim}, the matrix $\mE_{ij}$ and its inverse are written as
+	\[ \mE_{ij} = \mI_n + c e_i e_{j}^T, \quad \mE^{-1}_{ij} = \mI_n - c e_i e_{j}^T.\]
+	WLOG, we assume $\mE_{ij}$ is a lower-triangular matrix. Then $c e_i e_{j}^T$ and  $-c e_i e_{j}^T$ are also lower-triangular matrices. Therefore, $\mE_{ij}^{-1}$ is a lower-triangular matrix.
 \end{proof}
@@
 \subsection{Gauss-Jordan Elimination}
-Consider an invertible matrix $\mA \in \bbR^{n \times n}$, one of the effective ways to find the inverse of $\mA$ uses elimination.
+We also use elimination to find the inverse of any invertible matrix.
 
 \vspace{0.2cm}
- The inverse of $\mA$, $\mA^{-1}$ , satisfies $\mA \mA^{-1} = \mI_n$.
+ Suppose $\mA \in \bbR^{n \times n}$ is an invertible matrix. The inverse of $\mA$, $\mA^{-1}$ , satisfies 
+ \begin{equation}\label{eq:inv}
+ 	\mA \mA^{-1} = \mI_n.
+ \end{equation}
  Suppose there is an elimination $\mE$ such that $\mE \mA = \mI_n$.
-Multiplying $\mE$ on the both side of the equation, we have $\mE  \mA \mA^{-1}= \mA^{-1} = \mE$.
-To obtain an such $\mE$, we do elimination to the \textit{augmented matrix} $[\mA|\mI_n]$ until $\mA$ becomes $\mI_n$.
+Multiplying $\mE$ on the both side of the equation~\eqref{eq:inv}, we have $\mE  \mA \mA^{-1}= \mA^{-1} = \mE$.
+To obtain an such $\mE$, we eliminate the \textit{augmented matrix} $[\mA|\mI_n]$ until $\mA$ becomes $\mI_n$.
+Then, the augmented matrix becomes $\mE[\mA|\mI_n] = [\mI_n|\mE]$, where $\mE$ is the inverse of $\mA$.
 
  \vspace{0.2cm}
- We call the elimination process of finding $\mE$ as \textit{Gauss-Jordan Elimination}.
+We call this elimination process of finding $\mE$ as \textit{Gauss-Jordan Elimination}.
 
 \subsection{Factorization of matrices}
-By elimination, for any square matrix $\mA$, we have $\mE \mP \mA = \mU$ where $\mU$ is an upper triangular matrix.
-By "canceling" the elimination matrix $\mE$, we get $\mP \mA = \mE^{-1} \mU$.
-Because $\mE$ is invertible, the inverse $\mE^{-1}$ exists.
-Note that $\mE$ is lower triangular matrix, the inverse of $\mE$ is also a lower triangular matrix.
-We use $\mL$ to denote $\mE^{-1}$.
-Therefore, we can decompose an arbitrary square matrix $\mA$ as:
-\[ \mP \mA = \mL \mU, \]
+By elimination, for any square matrix $\mA$, we have equation~\eqref{eq:factor}.
+ By lemma~\ref{lem:elim}, $\mE$ is invertible. We multiply $\mE^{-1}$ on both sides of equation~\eqref{eq:factor}. We have, 
+\[ \mP \mA = \mE^{-1} \mU. \]
+ Note that $\mE$ is a lower-triangular matrix.
+By corollary~\ref{cor:invelim}, $\mE^{-1}$ is also a lower-triangular matrix.
+Let $\mL$ denote $\mE^{-1}$, where the letter $L$ refers to ``lower triangular".
+Therefore, any square matrix $\mA$ has a factorization:
+\begin{equation}\label{eq:LUfact}
+	\mP \mA = \mL \mU,
+\end{equation}
 where $\mU$ is an upper triangular matrix with pivots on the diagonal, $\mL$ is lower triangular matrix with ones on the diagonal, and $\mP$ is a permutation matrix.
-Note that, there may exist other ways to decompose $\mP\mA$ into $\mL \mU$, where $\mU$ and $\mL$ have different settings.
+However, the equation~\eqref{eq:LUfact} is not the unique factorization of $\mA$.
+For example, $c\mL$ and $c^{-1}\mU$ also factorize $\mA$, where $c$ is a non-zero scalar.
 
 \subsection{Time complexity of elimination}
-For an $n$-by-$n$ matrix, multiplying one row and then subtracting it from another row require $2n$ operations.
-There are $n$ rows in the matrix, so the total number of operations used in elimination the first column is $2n^2$.
-The second row and column are shorter, which may cost $2(n-1)^2$ operations  and so on.
-Therefore, the time complexity to factorize $\mA$ into $\mL \mU$ is about $\tO(n^3)$: \[ 1^2 + 2^2 + \cdots +(n-1)^2 + n^2 = \sum_{i}^n i^2 \approx \int_{0}^n x^2 dx = \frac{1}{3} n^3.\]
+For an $n$-by-$n$ matrix, a single elimination multiplies a selected row and subtracts the selected row from another row.
+A single elimination requires $\tO(n)$ operations.
+To eliminate the elements below the first diagonal element, we need repeat single eliminations $(n-1)$ times,  and thus require $\tO(n^2)$ operations.
+Similarly, we require $\tO((n-1)^2)$  to eliminate the elements below the second diagonal element.
+Repeat the elimination until we meet the $n$-th diagonal element.
+Therefore, we require $\tO(n^3)$ operations to obtain an upper-triangular matrix by elimination:
+ \[ 1^2 + 2^2 + \cdots +(n)^2 = \sum_{i}^{n} i^2 \approx \int_{0}^{n} x^2 dx = \frac{1}{3} n^3 = \tO(n^3).\]
 
 \subsection{Reduced row echelon form of matrices}
-By continuing to use the method of elimination, we can convert $\mU$ to a matrix $\mR$ in reduced row echelon form, with pivots equal to 1 and zeros above and below the pivots.
-The matrix $\mR$ is called the \textit{reduced row echelon form of matrices}(RREF) of $\mA$.
-In previous example,
+In previous sections, we convert any matrix $\mA$ to an upper triangular matrix $\mU$.
+Next, we convert $\mU$ into \textit{reduced row echelon form} (RREF), which is a simpler form than upper triangle.
+We use $\mR = RREF(\mA)$ to denote the reduced row echelon form of $\mA$.
+In $\mR$, the pivots are equal to 1, and the elements above and below the pivots are eliminated to 0.
+In the previous example,
 \begin{align}
@@
 
-For an another example,
+There is another example,
 \begin{align}
@@
 
-With proper permutation, the matrix $\mR$ can be written in form $[\mI \quad \mF]$, $\begin{bmatrix} \mI&\mF\\0 &0	\end{bmatrix}$, $\begin{bmatrix} \mI \\0	\end{bmatrix}$ , or just $\mI$, where $\mF$ can be arbitrary matrix in proper dimension.
-The columns in $\mA$ which correspond to the identity matrix $\mI$ are called \textit{pivot columns} and the other columns are \textit{free columns}.
+Assume there are $r$ pivots in $\mA \in \bbR^{m \times n}$.
+With proper permutation, the matrix $\mR$ is in form $\begin{bmatrix} \mI_r&\mF\\0 &0	\end{bmatrix}$, where $\mF \in \bbR^{ r \times (n-r)}$ is an arbitrary matrix.
+The columns in $\mA$ which correspond to the identity matrix $\mI_r$ are called \textit{pivot columns}.
+The other columns are \textit{free columns}.
 
@@
 \begin{itemize}
-	\item \textit{Vector space} is a collection of vectors that are closed under linear combination(addition and multiplication by any real number); i.e.
+	\item \textit{Vector space} is a collection of vectors that is closed under linear combination (addition and multiplication by any real number); i.e.
 for any vectors in the collection, all the combinations of these vectors are still in the collection.
@@
 
-Note that any vector space or subspace must include an origin. For a vector space $\tA$, the subspace of $\tA$ can be $\tA$ itself or can only contain a zero vector.
+Note that any vector space or subspace must include an origin. For a vector space $\tA$, the subspace of $\tA$ can be $\tA$ itself or a set that contains only a zero vector.
 
@@
 
-Note that if  $v_1 ,...,v_n$ span a space $\tS$, then $\tS$ is the smallest space that contain those vectors.
+If  $v_1 ,...,v_n$ span a space $\tS$, then $\tS$ is the smallest space that contain those vectors.
 
 \begin{itemize}
-	\item \textit{Basis} of a vector space is a sequence of vectors $v_1 ,...,v_n$ satisfies: (1) $v_1 ,...,v_n$ are independent; (2) $v_1 ,...,v_n$ span the space.
-	\item \textit{Dimension} of the space is the number of vectors in a basis of the space.
+	\item \textit{Basis} of a vector space is a sequence of vectors $v_1 ,...,v_n$ that satisfy: (1) $v_1 ,...,v_n$ are independent; (2) $v_1 ,...,v_n$ span the space.
+	\item \textit{Dimension} of the space is the number of vectors in a basis of the space. Let $dim(\tA)$ denote the dimension of space $\tA$.
 \end{itemize}
@@
 \subsection{Matrix rank}
-The \textit{rank} of a matrix $\mA$ is defined as the dimension of the columns space of $\mA$. Rank is also equal to the number of pivot columns of $\mA$. That means:
-\[ \text{rank}(\mA) = \# \text{ of pivot columns of }\mA = \text{dimension of } C(\mA).\]
-We use $r$ to denote the rank of $\mA$. If $\mA\in \bbR^{m \times n}$, then we have $r \leq m, r \leq n$. We say the matrix is \textit{full rank} if $r = n$ or $r = m$.
+The \textit{rank} of a matrix $\mA$ is defined as the dimension of the columns space of $\mA$.
+Rank of matrix $\mA$ is also equal to the number of pivot columns of $\mA$.
+Let $rank(\mA)$ denote the rank of matrix $\mA$.
+We have
+\begin{equation}\label{eq:rank}
+	rank(\mA) = \# \text{ of pivot columns of }\mA = dim(C(\mA)).
+\end{equation}
+If $\mA\in \bbR^{m \times n}$ and $rank(\mA) = r$ , we have $r \leq \min\{m,n\} $. We say the matrix is \textit{full rank} if $r = \min\{m,n\}$.
 
@@
 \begin{proof}
-	First, assume $\mA$ is full rank, we prove that $\mA$ has an inverse.
-	We can get find a RREF($\mA$) by eliminations and permutations. There are $\mE$ and $\mP$ such that
+	First, we assume $\mA$ is full rank and prove the invertibility of $\mA$.
+
+	Consider the RREF form of $\mA$, $\mR = RREF(\mA)$. There are elimination matrix $\mE$ and permutation matrix $\mP$ such that
 	\[ \mE \mP \mA =  \mR. \]
-	Since $\mA$ is full rank, $\mA$ have $n$ pivots columns and that implies $\mR = \mI_n$.
+	By the full rankness of $\mA$,  $\mA$ has $n$ pivot columns, and thus $\mR = \mI_n$.
 By lemma \ref{lem:elim}, $\mE$ is invertible.
 The permutation matrix $\mP$ is also invertible.
-Therefore, $\mE \mP$ is invertible and $\mA \mE \mP= \mI_n$.
-This implies $\mA$ is invertible.
+Then, the matrix product $\mE \mP$ is invertible and $\mA$ is the inverse of $\mE \mP$.
+Therefore, $\mA$ is invertible.
 
-	Second, assume $\mA$ has an inverse, we prove that $\mA$ is full rank. 
-	We show this by contradiction.
-Assume $\mA$ has an inverse $\mA^{-1}$ and $\mA$ is not full rank.
-Because the rank of $\mA$ is equal to the dimension of $C(\mA)$, the columns of $\mA$ are linearly dependent without full rankness.
-So, there exist a non-zero vector $v$ such that \[ \mA v = 0.\] Multiplying $\mA^{-1}$ on the both sides of the equation, we have \[ v = \mA^{-1} 0 = 0.\] However, it contradicts to  non-zeroness of $v$.
-Therefore, $\mA$ must be full rank.
+	Second, we assume $\mA$ is invertible and prove the full rankness of $\mA$ by contradiction.
+
+  Assume $\mA$ has an inverse $\mA^{-1}$ and $rank(\mA) < n$.
+By equation~\eqref{eq:rank}, $dim(C(\mA)) = rank(\mA)<n$, which implies that the columns of $\mA$ are linearly dependent.
+Then, there exists a non-zero vector $v$ such that
+  \begin{equation}\label{eq:profull}
+  	\mA v = 0.
+  \end{equation}
+Multiplying $\mA^{-1}$ on both sides of equation~\eqref{eq:profull}, we have \[ v = \mA^{-1} 0 = 0.\] However, it contradicts the fact that $v$ is non-zero. Therefore, $\mA$ is full rank.
 \end{proof}
@@
 
-Here we discuss the solution situation of the linear system $Ax = b$. Without specific explanation, the matrix $\mA \in \bbR^{m \times n}$ and the vector $x \in \bbR^n$, $b \in \bbR^m$.
+Here we discuss the solutions of the linear system $Ax = b$, where $\mA \in \bbR^{m \times n}$ is a matrix, and $x \in \bbR^n$, $b \in \bbR^m$ are vectors.
 
@@
 \begin{proof}
-To show the $N(\mA)$ is a vector space, we need to show $N(\mA)$ is close to linear combination.
-For any integer $k$, take arbitrary vectors $v_1 ,...,v_k \in N(\mA)$ and arbitrary scalars $c_1, ..., c_k$.
-We have,
-\[ \mA(c_1 v_1 + \cdots + c_k v_k) = c_1 \mA v_1 + \cdots + c_k \mA v_k = 0. \]
-Therefore, the linear combination $(c_1 v_1 + \cdots + c_k v_k) \in N(\mA)$. Then $N(\mA)$ is a vector space.
+We need to show $N(\mA)$ is closed to linear combination to prove that $N(\mA)$ is a vector space. For $\forall v_1, v_2 \in N(\mA)$, we have,
+\begin{equation}\label{eq:pronull}
+	\mA(c_1 v_1 + c_2 v_2) = c_1 \mA v_1 +  c_2 \mA v_2 = 0, \quad \forall  c_1, c_2 \in \bbR.
+\end{equation}
+The equation~\eqref{eq:pronull} implies that any linear combination of vectors in $N(\mA)$ is also a vector in $N(\mA)$. Therefore, $N(\mA)$ is closed to linear combination.
 \end{proof}
@@
 \begin{lem}[The rank of nullspace]\label{lem:ranknull}
-Suppose the rank of $\mA$ is $r$, then the rank of $N(\mA)$ is $n - r$.
+If  $rank(\mA) = r$, the rank of nullspace $rank(N(\mA)) = n-r$ .
 \end{lem}
 
-\begin{proof} Let $\mR$ denote the RREF($\mA$).
-The matrix $\mR$ can be written in form $\mR = \begin{bmatrix} \mI_r & \mF \\ 0 & 0	\end{bmatrix}$ where $\mF \in \bbR^{r \times (n-r)}$ and $0$ are zero matrices with proper dimensions.
-Let $\mX = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix}$. Then we have 
+\begin{proof} Let $\mR$ denote the RREF($\mA$).  We write $\mR$  in form $\mR = \begin{bmatrix} \mI_r & \mF \\ 0 & 0	\end{bmatrix}$, where $\mF \in \bbR^{ r \times (n-r)}$ is arbitrary matrix.
+Let $\mX = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix}$.  We have 
 \begin{align}
-	\mR \mX = \begin{bmatrix} \mI_r & \mF \\ 0 & 0	\end{bmatrix} \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix} = 0
+	\mR \mX = \begin{bmatrix} \mI_r & \mF \\ 0 & 0	\end{bmatrix} \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix} = 0.
 \end{align}
-Therefore, each column $\mX$ is a special solution to the system $\mA x = 0$. Next, we want to show other solutions to $\mA x = 0$ are linear combinations of those special solutions.
+Therefore, each column of $\mX$ is a special solution to the system $\mA x = 0$. Next, we show other solutions to $\mA x = 0$ are linear combinations of those special solutions.
 
-Suppose there is a $x = (x_1, x_2) \in N(\mA)$. Then 
+Suppose there is a solution $x = (x_1, x_2) \in N(\mA)$, where $x_1 \in \bbR^{r}$ and $x_2 \in \bbR^{n-r}$. We have 
 \begin{align}
@@
 That implies $x_1  = -\mF x_2$ and $x = \begin{bmatrix} x_1 \\ x_2	\end{bmatrix} = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix} x_2 = \mX x_2$.
-Therefore, arbitrary $x \in N(\mA)$ is a linear combination of special solutions, i.e.
+Any vector  in $ N(\mA)$ is a linear combination of special solutions, i.e.
 $C(\mX) = N(\mA)$.
-
-Since $\mX$ contains an identity matrix, the special solutions are independent and the rank of $C(\mX)$ is $n -r$. Therefore, the rank of $N(\mA)$ is $n-r$. 
+Therefore, the rank of nullspace is $rank(N(\mA)) = dim(C(\mX)) = n-r$.
 \end{proof}
 
-Recall that the columns in $\mA$ correspond to the $\mI_r$ in $\mR$ are called pivot columns and other columns are free columns.
+Recall the definitions of pivot column and free column.
 In $\mA x = b$, the variables in $x$ that correspond to pivot columns are called \textit{pivot variables} and others are \textit{free variables}.
-As the proof shows, the way to find special solutions of $\mA x = 0$ is to let one of the free variables is 1 while other free variables are 0 and then solve the equation $\mA x = 0$.
-There are total $n-r$ special solutions.
+If $rank(\mA) = r$, there are $n-r$ free variables, which coincides with $rank(N(\mA))$.
+
+In the proof of lemma~\ref{lem:ranknull}, the columns of $\mX = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix}$ are special solutions that compose the basis of $N(\mA)$.
+Practically, we find those special solutions by assigning 1 to a free variable and 0 to other free variables, and then solving the system $\mA x = 0$.
 
@@
 \begin{proof}
-	For any $x$, $\mA x \in C(\mA)$. Therefore, to satisfy $\mA x = b$, $b \in C(\mA)$. 
+	 If $\mA x = b$ is solvable, there exists a $x$ such that $\mA x = b$. For any $x$, $\mA x \in C(\mA)$. Therefore, $b \in C(\mA)$.
 \end{proof}
 
-\begin{lem}[Complete solution]\label{lem:compsolu} The complete solution of $\mA x = b$ is given by $x_{comp} = x_p + x_n$, where $x_p$ is a particular solution that $\mA x_p = b$ and $x_n \in N(\mA)$.
+\begin{lem}[Complete solution]\label{lem:compsolu} The complete solution of $\mA x = b$ is given by $x_{comp} = x_p + x_n$, where $x_p$ is a particular solution such that $\mA x_p = b$, and $x_n \in N(\mA)$.
 \end{lem}
@@
 \begin{proof}
-	Suppose $x = x_p + x_0$ is an arbitrary solution to $\mA x= b$. Then we have 
+	Suppose $x = x_p + x_0$ is an arbitrary solution to $\mA x= b$. We have 
 	\begin{align}
@@
 	\end{align}
-	Then $x_0 \in N(\mA)$.
+	Therefore, $x_0 \in N(\mA)$.
 \end{proof}
 
-One way to find a particular solution is to let all the free variables be 0 and solve the equations.
+Usually, we find a particular solution by assigning 0 to free variables, and solving the system $\mA x = b$.
 
 \vspace{.2cm}
-Here is a summary table that discusses the rank of $\mA$, the form of $\mR$ and the situation about the solutions.
+The following table discusses the rank of $\mA$, the form of $\mR$, dimension of nullspace $N(\mA)$, and solutions of $\mA x = b$.
 
@@
   $\mR$& $\mI$ & $\begin{bmatrix} \mI \\ 0	\end{bmatrix}$ & $\begin{bmatrix} \mI & \mF \end{bmatrix}$ & $\begin{bmatrix} \mI & \mF \\ 0 &0	\end{bmatrix}$\\
-  dimension of $N(\mA)$& 0 &0& $n -r$ & $n-r$ \\
+   $dim(N(\mA))$& 0 &0& $n -r$ & $n-r$ \\
   \# solutions to $\mA X = b$ & 1 & 0 or 1 & infinitely many & 0 or infinitely many\\

### 2020-06-30 skills/math/linear_algebra_summary.tex
update

@@
 If the vectors  $v_1 ,...,v_n$  are \textit{dependent}, there exist scalars $c_1, ..., c_n$ which are not all equal to 0 such that $c_1 v_1 + \dots + c_n v_n = 0$.
-	\item Given a matrix $\mA \in \bbR^{m \times n}$ and a vector $x \in \bbR^{n}$,  the multiplication $\mA x$ is a linear combination of the columns of $\mA$ and $x^T \mA$ is a linear combination of the rows of $\mA$.
+	\item Given a matrix $\mA \in \bbR^{m \times n}$ and a vector $x \in \bbR^{n}$  the multiplication $\mA x$ is a linear combination of the columns of $\mA$, and $x^T \mA$ is a linear combination of the rows of $\mA$.
 	\item Matrix multiplication is typically not communicative, i.e. $\mA \mB \neq \mB \mA$. Lemma~\ref{lem:matrixcomm} describes a special case where matrix multiplication is communicative.

### 2020-06-30 skills/math/linear_algebra_summary.tex
update

@@
 \subsection{Permutation of matrices}
-For any matrix $\mA$, we can swap its rows by multiplying a \textit{permutation matrix } $\mP$ on the left of $\mA$. For example,
+For any matrix $\mA$, we swap its rows by multiplying a \textit{permutation matrix } $\mP$ on the left of $\mA$. For example,
 \begin{align}
@@
 \vspace{.2cm}
-Similarly, we can also swap the columns of the matrix $\mA$ by multiplying a permutation matrix on the right of $\mA$.
+Similarly, we also swap the columns of the matrix $\mA$ by multiplying a permutation matrix on the right of $\mA$.
 

### 2020-06-30 skills/math/linear_algebra_summary.tex

@@
 The permutation matrix $\mP$ is also invertible.
-Then, the matrix product $\mE \mP$ is invertible and $\mA$ is the inverse of $\mE \mP$.
+Then, the matrix product $\mE \mP$ is invertible, and $\mA$ is the inverse of $\mE \mP$.
 Therefore, $\mA$ is invertible.
@@
 \end{align}
-That implies $x_1  = -\mF x_2$ and $x = \begin{bmatrix} x_1 \\ x_2	\end{bmatrix} = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix} x_2 = \mX x_2$.
+That implies $x_1  = -\mF x_2$, and $x = \begin{bmatrix} x_1 \\ x_2	\end{bmatrix} = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix} x_2 = \mX x_2$.
 Any vector  in $ N(\mA)$ is a linear combination of special solutions, i.e.
@@
 Recall the definitions of pivot column and free column.
-In $\mA x = b$, the variables in $x$ that correspond to pivot columns are called \textit{pivot variables} and others are \textit{free variables}.
+In $\mA x = b$, the variables in $x$ that correspond to pivot columns are called \textit{pivot variables}, and others are \textit{free variables}.
 If $rank(\mA) = r$, there are $n-r$ free variables, which coincides with $rank(N(\mA))$.

### 2020-06-30 skills/math/linear_algebra_summary.tex
update

@@
 If the vectors  $v_1 ,...,v_n$  are \textit{dependent}, there exist scalars $c_1, ..., c_n$ which are not all equal to 0 such that $c_1 v_1 + \dots + c_n v_n = 0$.
-	\item Given a matrix $\mA \in \bbR^{m \times n}$ and a vector $x \in \bbR^{n}$  the multiplication $\mA x$ is a linear combination of the columns of $\mA$, and $x^T \mA$ is a linear combination of the rows of $\mA$.
+	\item Given a matrix $\mA \in \bbR^{m \times n}$ and a vector $x \in \bbR^{n}$,  the multiplication $\mA x$ is a linear combination of the columns of $\mA$, and $x^T \mA$ is a linear combination of the rows of $\mA$.
 	\item Matrix multiplication is typically not communicative, i.e. $\mA \mB \neq \mB \mA$. Lemma~\ref{lem:matrixcomm} describes a special case where matrix multiplication is communicative.
 	\item Suppose $\mA$ is a square matrix.
-The matrix $\mA$ is \textit{invertible} or \textit{non-singular} if there exists a $\mA^{-1}$ such that $\mA^{-1} \mA = \mA \mA^{-1} = I$.
+The matrix $\mA$ is \textit{invertible} or \textit{non-singular} if there exists an $\mA^{-1}$ such that $\mA^{-1} \mA = \mA \mA^{-1} = I$.
 Otherwise, the matrix $\mA$ is \textit{singular}, and the determinant of $\mA$ is 0.
@@
 \begin{lem}[Communicative matrix multiplication]\label{lem:matrixcomm} For matrices $\mA \in \bbR^{n \times n}$ and $\mB \in \bbR^{n \times n}$, the matrix multiplication of $\mA$ and $\mB$ is communicative, i.e.
-$\mA\mB  = \mB \mA$, if all the normalized eigenvectors of $\mA$ and $\mB$ are identical.
+$\mA\mB  = \mB \mA$, if $\mA$ and $\mB$ have the same set of eigenvectors corresponding to their non-zero eigenvalues.
 \end{lem}
@@
 \begin{proof}
-	If all the normalized eigenvectors of $\mA$ and $\mB$ are identical, there exists a matrix $\mQ \in \bbR^{n \times n}$ such that
+	If $\mA$ and $\mB$ have the same set of eigenvectors corresponding to their non-zero eigenvalues, there exists a matrix $\mQ \in \bbR^{n \times n}$ such that
 	\[ \mA = \mQ \mD_A \mQ^{-1}, \quad \mB = \mQ \mD_B \mQ^{-1},\]
-	where columns of $\mQ$ are normalized eigenvectors of $\mA$ and $\mB$, and $\mD_A\in\bbR^{n \times n},  \mD_B\in\bbR^{n \times n} $ are diagonal matrices whose diagonal elements are corresponding eigenvalues of $\mA$ and $\mB$.
+	where columns of $\mQ$ are eigenvectors of $\mA$ and $\mB$, and $\mD_A\in\bbR^{n \times n},  \mD_B\in\bbR^{n \times n} $ are diagonal matrices whose diagonal elements are eigenvalues of $\mA$ and $\mB$, respectively.
 Because matrix multiplication is communicative for two diagonal matrices with same dimensions, we have
@@
 where $a_k$ refers to the $k$-th row of $\mA$.
-The inverse of permutation matrix $\mP$ is $\mP^{-1} = \mP^T$, which implies the orthogonality of permutation matrix.
+The inverse of the permutation matrix $\mP$ is $\mP^{-1} = \mP^T$, which implies the orthogonality of permutation matrix.
 For an $n \times m$ matrix, there are $n!$ different row permutation matrices, which form a \textit{multiplicative group}.
@@
 \end{align}
-In step 1, we choose the number 1 in row 1 column 1 as a \textit{pivot}, then we recopy the first row, multiply an appropriate number (in this case, 3), and subtract those values from the numbers in the second row.
+In step 1, we choose the number 1 in row 1 column 1 as a \textit{pivot}, then we recopy the first row, multiply an appropriate number (in this case, 3) and subtract these values from the numbers in the second row.
 We have thus eliminated 3 in row 2 column 1.
-Similarly, in step 2, we choose 2 in row 2 column 2 as a pivot, and eliminate the number 4 in row 3 column 2.
+Similarly, in step 2, we choose 2 in row 2 column 2 as a pivot and eliminate the number 4 in row 3 column 2.
 The number 5 in row 3 column 3 is also a pivot.
@@
 \vspace{.2cm}
-The \textit{elimination matrix} used to eliminate the entry in row $m$ column $n$ is denoted $\mE_{mn}$. In the previous example, 
+The \textit{elimination matrix} used to eliminate the entry in row $m$ column $n$ is denoted as $\mE_{mn}$. In the previous example, 
 \begin{align}
@@
 If there is a 0 in the pivot position, we must exchange the row with one below to get a non-zero value in pivot position.
-If there is not non-zero value below the 0 pivot, we skip this column, and find a pivot in next column.
+If there is not non-zero value below the 0 pivot, we skip this column and find a pivot in the next column.
 
 \vspace{.2cm}
-We write $\mE_{32} (\mE_{21} \mA) = (\mE_{32} \mE_{21} ) \mA = \mU$ because matrix multiplication is associative.
+We write $\mE_{32} (\mE_{21} \mA) = (\mE_{32} \mE_{21} ) \mA = \mU$, because matrix multiplication is associative.
 Let $\mE$ denote the product of all elimination matrices.
@@
 \begin{lem}[Invertiblity of elimination matrix]\label{lem:elim}
- Suppose there is an elimination matrix $\mE_{ij} \in \bbR^{n \times n}$ that multiplies a scalar $-c$ to the  $j$-th row, and subtracts the row from $i$-th row, where $i \neq j$.
+ Suppose there is an elimination matrix $\mE_{ij} \in \bbR^{n \times n}$ that multiplies a scalar $-c$ to the  $j$-th row and subtracts the row from $i$-th row, where $i \neq j$.
 The matrix $\mE_{ij}$ is invertible.
@@
 	\[ \mE_{ij} = \mI_n + c e_i e_{j}^T, \]
-	where $e_i \in \bbR^n$ denotes the vector with value 1 in the $i$-th entry and value 0 elsewhere. Note that $e^T_i e_{j} = 0$ because $i \neq j$. We have 
+	where $e_i \in \bbR^n$ denotes the vector with value 1 in the $i$-th entry and value 0 elsewhere.  Because $i \neq j$, $e^T_i e_{j} = 0$.  We have 
 	\[ (\mI_n + c e_i e_{j}^T)(\mI_n - c e_i e_{j}^T)  = \mI_n - c^2 e_i e_{j}^T e_i e_{j}^T = \mI_n; \quad (\mI_n - c e_i e_{j}^T)(\mI_n + c e_i e_{j}^T)= \mI_n. \]
@@
 	\[ \mE_{ij} = \mI_n + c e_i e_{j}^T, \quad \mE^{-1}_{ij} = \mI_n - c e_i e_{j}^T.\]
-	WLOG, we assume $\mE_{ij}$ is a lower-triangular matrix. Then $c e_i e_{j}^T$ and  $-c e_i e_{j}^T$ are also lower-triangular matrices. Therefore, $\mE_{ij}^{-1}$ is a lower-triangular matrix.
+	WLOG, we assume that $\mE_{ij}$ is a lower-triangular matrix. Then $c e_i e_{j}^T$ and  $-c e_i e_{j}^T$ are also lower-triangular matrices. Therefore, $\mE_{ij}^{-1}$ is a lower-triangular matrix.
 \end{proof}
@@
 \subsection{Gauss-Jordan Elimination}
-We also use elimination to find the inverse of any invertible matrix.
+We also use elimination to find the inverse of an invertible matrix.
 
@@
 Multiplying $\mE$ on the both side of the equation~\eqref{eq:inv}, we have $\mE  \mA \mA^{-1}= \mA^{-1} = \mE$.
-To obtain an such $\mE$, we eliminate the \textit{augmented matrix} $[\mA|\mI_n]$ until $\mA$ becomes $\mI_n$.
+To obtain a such $\mE$, we eliminate the \textit{augmented matrix} $[\mA|\mI_n]$ until $\mA$ becomes $\mI_n$.
 Then, the augmented matrix becomes $\mE[\mA|\mI_n] = [\mI_n|\mE]$, where $\mE$ is the inverse of $\mA$.
@@
 \end{equation}
-where $\mU$ is an upper triangular matrix with pivots on the diagonal, $\mL$ is lower triangular matrix with ones on the diagonal, and $\mP$ is a permutation matrix.
+where $\mU$ is an upper triangular matrix with pivots on the diagonal, $\mL$ is a lower triangular matrix with ones on the diagonal, and $\mP$ is a permutation matrix.
 However, the equation~\eqref{eq:LUfact} is not the unique factorization of $\mA$.
@@
 A single elimination requires $\tO(n)$ operations.
-To eliminate the elements below the first diagonal element, we need repeat single eliminations $(n-1)$ times,  and thus require $\tO(n^2)$ operations.
-Similarly, we require $\tO((n-1)^2)$  to eliminate the elements below the second diagonal element.
+To eliminate the elements below the first diagonal element, we need to repeat single elimination $(n-1)$ times and thus require $\tO(n^2)$ operations.
+Similarly, we require $\tO((n-1)^2)$ operations to eliminate the elements below the second diagonal element.
 Repeat the elimination until we meet the $n$-th diagonal element.
-Therefore, we require $\tO(n^3)$ operations to obtain an upper-triangular matrix by elimination:
+Therefore, we require total $\tO(n^3)$ operations to obtain an upper-triangular matrix by elimination:
  \[ 1^2 + 2^2 + \cdots +(n)^2 = \sum_{i}^{n} i^2 \approx \int_{0}^{n} x^2 dx = \frac{1}{3} n^3 = \tO(n^3).\]
@@
 \subsection{Reduced row echelon form of matrices}
-In previous sections, we convert any matrix $\mA$ to an upper triangular matrix $\mU$.
-Next, we convert $\mU$ into \textit{reduced row echelon form} (RREF), which is a simpler form than upper triangle.
+In previous sections, we convert a matrix $\mA$ to an upper triangular matrix $\mU$.
+Next, we convert $\mU$ into the \textit{reduced row echelon form} (RREF), which is a simpler form than upper triangle.
 We use $\mR = RREF(\mA)$ to denote the reduced row echelon form of $\mA$.
@@
 
-Assume there are $r$ pivots in $\mA \in \bbR^{m \times n}$.
+Assume that there are $r$ pivots in $\mA \in \bbR^{m \times n}$.
 With proper permutation, the matrix $\mR$ is in form $\begin{bmatrix} \mI_r&\mF\\0 &0	\end{bmatrix}$, where $\mF \in \bbR^{ r \times (n-r)}$ is an arbitrary matrix.
@@
 \begin{itemize}
-	\item \textit{Vector space} is a collection of vectors that is closed under linear combination (addition and multiplication by any real number); i.e.
+	\item \textit{Vector space} is a collection of vectors that is close under linear combination (addition and multiplication by any real number); i.e.
 for any vectors in the collection, all the combinations of these vectors are still in the collection.
@@
 \begin{itemize}
-	\item Vectors $v_1 ,...,v_n$ \textit{span} a space that consists all the combination of those vectors. 
-	\item \textit{Column space} of matrix $\mA$ is the space spanned by the columns of $\mA$. Let $C(\mA)$ denote the column space of $\mA$.
+	\item Vectors $v_1 ,...,v_n$ \textit{span} a space that consists all the combination of these vectors. 
+	\item \textit{Column space} of a matrix $\mA$ is the space spanned by the columns of $\mA$. Let $C(\mA)$ denote the column space of $\mA$.
 \end{itemize}
 
-If  $v_1 ,...,v_n$ span a space $\tS$, then $\tS$ is the smallest space that contain those vectors.
+If  $v_1 ,...,v_n$ span a space $\tS$, then $\tS$ is the smallest space that contain these vectors.
 
@@
 	\item \textit{Basis} of a vector space is a sequence of vectors $v_1 ,...,v_n$ that satisfy: (1) $v_1 ,...,v_n$ are independent; (2) $v_1 ,...,v_n$ span the space.
-	\item \textit{Dimension} of the space is the number of vectors in a basis of the space. Let $dim(\tA)$ denote the dimension of space $\tA$.
+	\item \textit{Dimension} of a space is the number of vectors in a basis of this space. Let $dim(\tA)$ denote the dimension of space $\tA$.
 \end{itemize}
@@
 \subsection{Matrix rank}
-The \textit{rank} of a matrix $\mA$ is defined as the dimension of the columns space of $\mA$.
-Rank of matrix $\mA$ is also equal to the number of pivot columns of $\mA$.
-Let $rank(\mA)$ denote the rank of matrix $\mA$.
-We have
+Let $rank(\mA)$ denote the \textit{rank} of matrix $\mA$. We have
 \begin{equation}\label{eq:rank}
-	rank(\mA) = \# \text{ of pivot columns of }\mA = dim(C(\mA)).
+	rank(\mA) \overset{\Delta}{=} dim(C(\mA)) = \# \text{ of pivot columns of }\mA .
 \end{equation}
-If $\mA\in \bbR^{m \times n}$ and $rank(\mA) = r$ , we have $r \leq \min\{m,n\} $. We say the matrix is \textit{full rank} if $r = \min\{m,n\}$.
+If $\mA\in \bbR^{m \times n}$ and $rank(\mA) = r$ , we have $r \leq \min\{m,n\} $. The  matrix is \textit{full rank} if $r = \min\{m,n\}$.
 
 \vspace{.2cm}
-The rank of a square matrix is closely related to its invertibility. 
+
+The rank of a square matrix is closely related to the invertibility. 
 \begin{lem}[Full rankness and invertibility]\label{lem:fullrank} A square matrix $\mA \in \bbR^{n \times n}$ is full rank, if and only if $\mA$ is an invertible matrix.
@@
 \begin{proof}
-	First, we assume $\mA$ is full rank and prove the invertibility of $\mA$.
+	$(\Rightarrow)$ Suppose $\mA$ is full rank.
 
-	Consider the RREF form of $\mA$, $\mR = RREF(\mA)$. There are elimination matrix $\mE$ and permutation matrix $\mP$ such that
+	Let $\mR$ denote the RREF form of $\mA$. There exist an elimination matrix $\mE$ and a permutation matrix $\mP$ such that
 	\[ \mE \mP \mA =  \mR. \]
@@
 
-	Second, we assume $\mA$ is invertible and prove the full rankness of $\mA$ by contradiction.
+	$(\Leftarrow)$ Suppose $\mA$ is invertible.  We prove the full rankness of $\mA$ by contradiction.
 
-  Assume $\mA$ has an inverse $\mA^{-1}$ and $rank(\mA) < n$.
+  Assume that $rank(\mA) < n$.
 By equation~\eqref{eq:rank}, $dim(C(\mA)) = rank(\mA)<n$, which implies that the columns of $\mA$ are linearly dependent.
@@
   \end{equation}
-Multiplying $\mA^{-1}$ on both sides of equation~\eqref{eq:profull}, we have \[ v = \mA^{-1} 0 = 0.\] However, it contradicts the fact that $v$ is non-zero. Therefore, $\mA$ is full rank.
+By assumption, the inverse $\mA^{-1}$ exists. Multiplying $\mA^{-1}$ on both sides of equation~\eqref{eq:profull}, we have 
+ \begin{equation}\label{eq:profull1}
+  	v = \mA^{-1} 0 = 0.
+  \end{equation}
+However, equation~\eqref{eq:profull1} contradicts the fact that $v$ is a non-zero vector. Therefore, $\mA$ is full rank.
 \end{proof}
 
-The rank of $\mA$ also effects the number of solutions to the system $\mA x = b$. We will discuss it in next section.
+The rank of $\mA$ also affects the number of solutions to the system $\mA x = b$. We will discuss the relationship between matrix rank and the solutions in next section.
 
@@
 
-Here we discuss the solutions of the linear system $Ax = b$, where $\mA \in \bbR^{m \times n}$ is a matrix, and $x \in \bbR^n$, $b \in \bbR^m$ are vectors.
+In this section, we discuss the solutions to the linear system $\mA x = b$, where $\mA \in \bbR^{m \times n}$ is a matrix, and $x \in \bbR^n$, $b \in \bbR^m$ are vectors.
 
@@
 \begin{proof}
-We need to show $N(\mA)$ is closed to linear combination to prove that $N(\mA)$ is a vector space. For $\forall v_1, v_2 \in N(\mA)$, we have,
+We only need to prove that $N(\mA)$ is close under linear combination. For $\forall v_1, v_2 \in N(\mA)$, we have,
 \begin{equation}\label{eq:pronull}
@@
 \end{equation}
-The equation~\eqref{eq:pronull} implies that any linear combination of vectors in $N(\mA)$ is also a vector in $N(\mA)$. Therefore, $N(\mA)$ is closed to linear combination.
+The equation~\eqref{eq:pronull} implies that  $N(\mA)$ is close under linear combination. Therefore, $N(\mA)$ is a vector space. 
 \end{proof}
@@
 \end{align}
-Therefore, each column of $\mX$ is a special solution to the system $\mA x = 0$. Next, we show other solutions to $\mA x = 0$ are linear combinations of those special solutions.
+Therefore, each column of $\mX$ is a special solution to the system $\mA x = 0$. Next, we show that any vector $ x \in N(\mA)$ is a linear combination of the columns of $\mX$.
 
@@
 \end{align}
-That implies $x_1  = -\mF x_2$, and $x = \begin{bmatrix} x_1 \\ x_2	\end{bmatrix} = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix} x_2 = \mX x_2$.
-Any vector  in $ N(\mA)$ is a linear combination of special solutions, i.e.
+This implies that $x_1  = -\mF x_2$, and $x = \begin{bmatrix} x_1 \\ x_2	\end{bmatrix} = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix} x_2 = \mX x_2$.
+Any vector  in $ N(\mA)$ is a linear combination of the columns of $\mX$, i.e.
 $C(\mX) = N(\mA)$.
-Therefore, the rank of nullspace is $rank(N(\mA)) = dim(C(\mX)) = n-r$.
+Therefore, the rank of nullspace $rank(N(\mA)) = dim(C(\mX)) = n-r$.
 \end{proof}
 
-Recall the definitions of pivot column and free column.
+Recall the definitions of pivot columns and free columns.
 In $\mA x = b$, the variables in $x$ that correspond to pivot columns are called \textit{pivot variables}, and others are \textit{free variables}.
-If $rank(\mA) = r$, there are $n-r$ free variables, which coincides with $rank(N(\mA))$.
+If $rank(\mA) = r$, there are $n-r$ free variables.
 
 In the proof of lemma~\ref{lem:ranknull}, the columns of $\mX = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix}$ are special solutions that compose the basis of $N(\mA)$.
-Practically, we find those special solutions by assigning 1 to a free variable and 0 to other free variables, and then solving the system $\mA x = 0$.
+Practically, we find these special solutions by assigning 1 to a free variable and 0 to other free variables, and then we solve the system $\mA x = 0$.
 
@@
 
-Usually, we find a particular solution by assigning 0 to free variables, and solving the system $\mA x = b$.
+Usually, we find a particular solution by assigning 0 to free variables, and we solve the system $\mA x = b$.
 
 \vspace{.2cm}
-The following table discusses the rank of $\mA$, the form of $\mR$, dimension of nullspace $N(\mA)$, and solutions of $\mA x = b$.
+The following table discusses the rank of $\mA$, the form of $\mR$, the dimension of nullspace $N(\mA)$, and the number of solutions to $\mA x = b$.
 

### 2020-07-06 skills/math/linear_algebra/linear_algebra_summary.tex
upload

--- 2020-06-30 skills/math/linear_algebra_summary.tex
@@
     \item Given vectors $v_1 ,...,v_n$ and scalars $c_1, ..., c_n$, the sum $c_1 v_1 + \dots + c_n v_n$ is called a \textit{linear combination} of $v_1,...,v_n$.
-    \item The vectors  $v_1 ,...,v_n$  are \textit{linearly independent} (or just \textit{indepedent}) if $c_1 v_1 + \dots + c_n v_n = 0$ holds only when all  $c_1,...,c_n = 0$.
-If the vectors  $v_1 ,...,v_n$  are \textit{dependent}, there exist scalars $c_1, ..., c_n$ which are not all equal to 0 such that $c_1 v_1 + \dots + c_n v_n = 0$.
+    \item The vectors  $v_1 ,...,v_n$  are \textit{linearly independent} (or just \textit{indepedent}) if $c_1 v_1 + \dots + c_n v_n = 0$ holds only when $c_1 = ...=c_n = 0$.
+If the vectors  $v_1 ,...,v_n$  are \textit{dependent}, there exists a non-zero vector $(c_1, ..., c_n)$, such that $c_1 v_1 + \dots + c_n v_n = 0$.
 	\item Given a matrix $\mA \in \bbR^{m \times n}$ and a vector $x \in \bbR^{n}$,  the multiplication $\mA x$ is a linear combination of the columns of $\mA$, and $x^T \mA$ is a linear combination of the rows of $\mA$.
@@
 \subsection{Permutation of matrices}
-For any matrix $\mA$, we swap its rows by multiplying a \textit{permutation matrix } $\mP$ on the left of $\mA$. For example,
+Let $\mA$ be a matrix, we swap the rows of $\mA$ by multiplying a \textit{permutation matrix} $\mP$ on the left of $\mA$. For example,
 \begin{align}
@@
 \end{align}
-In step 1, we choose the number 1 in row 1 column 1 as a \textit{pivot}, then we recopy the first row, multiply an appropriate number (in this case, 3) and subtract these values from the numbers in the second row.
+In step 1, we choose the number 1 in row 1 column 1 as a \textit{pivot}.
+Then we recopy the first row, multiply an appropriate number (in this case, 3) and subtract these values from the numbers in the second row.
 We have thus eliminated 3 in row 2 column 1.
@@
 Pivots are non-zero.
-If there is a 0 in the pivot position, we must exchange the row with one below to get a non-zero value in pivot position.
-If there is not non-zero value below the 0 pivot, we skip this column and find a pivot in the next column.
+If there exists a 0 in the pivot position, we exchange the row with one below to get a non-zero value in pivot position.
+If all numbers below the pivots are 0, we skip this column and find a pivot in the next column.
 
@@
 	\[ \mE_{ij} = \mI_n + c e_i e_{j}^T, \]
-	where $e_i \in \bbR^n$ denotes the vector with value 1 in the $i$-th entry and value 0 elsewhere.  Because $i \neq j$, $e^T_i e_{j} = 0$.  We have 
+	where $e_i \in \bbR^n$ denotes the vector with 1 in the $i$-th entry and 0 elsewhere.  Because $i \neq j$, $e^T_i e_{j} = 0$.  We have 
 	\[ (\mI_n + c e_i e_{j}^T)(\mI_n - c e_i e_{j}^T)  = \mI_n - c^2 e_i e_{j}^T e_i e_{j}^T = \mI_n; \quad (\mI_n - c e_i e_{j}^T)(\mI_n + c e_i e_{j}^T)= \mI_n. \]
@@
 	\[ \mE_{ij} = \mI_n + c e_i e_{j}^T, \quad \mE^{-1}_{ij} = \mI_n - c e_i e_{j}^T.\]
-	WLOG, we assume that $\mE_{ij}$ is a lower-triangular matrix. Then $c e_i e_{j}^T$ and  $-c e_i e_{j}^T$ are also lower-triangular matrices. Therefore, $\mE_{ij}^{-1}$ is a lower-triangular matrix.
+	Without the loss of generality (WLOG), we assume that $\mE_{ij}$ is a lower-triangular matrix.
+Then $c e_i e_{j}^T$ and  $-c e_i e_{j}^T$ are also lower-triangular matrices.
+Therefore, $\mE_{ij}^{-1}$ is a lower-triangular matrix.
 \end{proof}
@@
 \vspace{0.2cm}
- Suppose $\mA \in \bbR^{n \times n}$ is an invertible matrix. The inverse of $\mA$, $\mA^{-1}$ , satisfies 
+ Suppose $\mA \in \bbR^{n \times n}$ is an invertible matrix. The inverse of $\mA$, denoted $\mA^{-1}$ , satisfies 
  \begin{equation}\label{eq:inv}
@@
  \end{equation}
- Suppose there is an elimination $\mE$ such that $\mE \mA = \mI_n$.
-Multiplying $\mE$ on the both side of the equation~\eqref{eq:inv}, we have $\mE  \mA \mA^{-1}= \mA^{-1} = \mE$.
+ Suppose there exists an elimination matrix $\mE$ such that $\mE \mA = \mI_n$.
+Multiplying $\mE$ on the both sides of the equation~\eqref{eq:inv}, we have $\mE  \mA \mA^{-1}= \mA^{-1} = \mE$.
 To obtain a such $\mE$, we eliminate the \textit{augmented matrix} $[\mA|\mI_n]$ until $\mA$ becomes $\mI_n$.
@@
 \subsection{Factorization of matrices}
-By elimination, for any square matrix $\mA$, we have equation~\eqref{eq:factor}.
- By lemma~\ref{lem:elim}, $\mE$ is invertible. We multiply $\mE^{-1}$ on both sides of equation~\eqref{eq:factor}. We have, 
+By elimination, any square matrix $\mA$ has the factorization~\eqref{eq:factor}.
+ By lemma~\ref{lem:elim}, $\mE$ is invertible. Multiplying $\mE^{-1}$ on both sides of equation~\eqref{eq:factor} yields, 
 \[ \mP \mA = \mE^{-1} \mU. \]
@@
 \subsection{Time complexity of elimination}
-For an $n$-by-$n$ matrix, a single elimination multiplies a selected row and subtracts the selected row from another row.
-A single elimination requires $\tO(n)$ operations.
+For an $n$-by-$n$ matrix, a single elimination step multiplies a selected row and subtracts the selected row from another row.
+A single elimination step requires $\tO(n)$ operations.
 To eliminate the elements below the first diagonal element, we need to repeat single elimination $(n-1)$ times and thus require $\tO(n^2)$ operations.
@@
 Repeat the elimination until we meet the $n$-th diagonal element.
-Therefore, we require total $\tO(n^3)$ operations to obtain an upper-triangular matrix by elimination:
+Therefore, we require in total $\tO(n^3)$ operations to obtain an upper-triangular matrix by elimination:
  \[ 1^2 + 2^2 + \cdots +(n)^2 = \sum_{i}^{n} i^2 \approx \int_{0}^{n} x^2 dx = \frac{1}{3} n^3 = \tO(n^3).\]
@@
 We use $\mR = RREF(\mA)$ to denote the reduced row echelon form of $\mA$.
-In $\mR$, the pivots are equal to 1, and the elements above and below the pivots are eliminated to 0.
+In the matrix $\mR$, the pivots are equal to 1, and the elements above and below the pivots are eliminated to 0.
 In the previous example,
@@
 Assume that there are $r$ pivots in $\mA \in \bbR^{m \times n}$.
-With proper permutation, the matrix $\mR$ is in form $\begin{bmatrix} \mI_r&\mF\\0 &0	\end{bmatrix}$, where $\mF \in \bbR^{ r \times (n-r)}$ is an arbitrary matrix.
+With proper permutation, the matrix $\mR$ is in the form of $\begin{bmatrix} \mI_r&\mF\\0 &0	\end{bmatrix}$, where $\mF \in \bbR^{ r \times (n-r)}$ is an arbitrary matrix.
 The columns in $\mA$ which correspond to the identity matrix $\mI_r$ are called \textit{pivot columns}.
@@
 
-\subsection{Vector space, Subspace and Column space}
+\subsection{Vector space, Subspace, and Column space}
 \begin{itemize}
-	\item \textit{Vector space} is a collection of vectors that is close under linear combination (addition and multiplication by any real number); i.e.
+	\item \textit{Vector space} is a collection of vectors that is closed under linear combination (addition and multiplication by any real number); i.e.
 for any vectors in the collection, all the combinations of these vectors are still in the collection.
@@
 \begin{itemize}
-	\item Vectors $v_1 ,...,v_n$ \textit{span} a space that consists all the combination of these vectors. 
+	\item Vectors $v_1 ,...,v_n$ \textit{span} a space that consists of all the linear combination of these vectors. 
 	\item \textit{Column space} of a matrix $\mA$ is the space spanned by the columns of $\mA$. Let $C(\mA)$ denote the column space of $\mA$.
@@
 
-If  $v_1 ,...,v_n$ span a space $\tS$, then $\tS$ is the smallest space that contain these vectors.
+If  $v_1 ,...,v_n$ span a space $\tS$, then $\tS$ is the smallest space that contains these vectors.
 
 \begin{itemize}
-	\item \textit{Basis} of a vector space is a sequence of vectors $v_1 ,...,v_n$ that satisfy: (1) $v_1 ,...,v_n$ are independent; (2) $v_1 ,...,v_n$ span the space.
+	\item \textit{Basis} of a vector space is a set of vectors $v_1 ,...,v_n$ that satisfy: (1) $v_1 ,...,v_n$ are independent; (2) $v_1 ,...,v_n$ span the space.
 	\item \textit{Dimension} of a space is the number of vectors in a basis of this space. Let $dim(\tA)$ denote the dimension of space $\tA$.
@@
 
-\subsection{Solving $Ax = 0$: Nullspace}
+\subsection{Solving $Ax = 0$: nullspace}
 The \textit{nullspace} of matrix $\mA$ is the collection of all solutions $x$ to the system $\mA x = 0$. Let $N(\mA)$ denote the nullspace of $\mA$.
@@
 \begin{proof}
-We only need to prove that $N(\mA)$ is close under linear combination. For $\forall v_1, v_2 \in N(\mA)$, we have,
+We only need to prove that $N(\mA)$ is closed under linear combination. For any $ v_1, v_2 \in N(\mA)$, we have,
 \begin{equation}\label{eq:pronull}
@@
 \end{equation}
-The equation~\eqref{eq:pronull} implies that  $N(\mA)$ is close under linear combination. Therefore, $N(\mA)$ is a vector space. 
+The equation~\eqref{eq:pronull} implies that  $N(\mA)$ is closed under linear combination. Therefore, $N(\mA)$ is a vector space. 
 \end{proof}
@@
 
-\begin{proof} Let $\mR$ denote the RREF($\mA$).  We write $\mR$  in form $\mR = \begin{bmatrix} \mI_r & \mF \\ 0 & 0	\end{bmatrix}$, where $\mF \in \bbR^{ r \times (n-r)}$ is arbitrary matrix.
+\begin{proof} Let $\mR$ denote the RREF($\mA$).  We write $\mR$  as $\mR = \begin{bmatrix} \mI_r & \mF \\ 0 & 0	\end{bmatrix}$, where $\mF \in \bbR^{ r \times (n-r)}$ is an arbitrary matrix.
 Let $\mX = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix}$.  We have 
@@
 \end{align}
-Therefore, each column of $\mX$ is a special solution to the system $\mA x = 0$. Next, we show that any vector $ x \in N(\mA)$ is a linear combination of the columns of $\mX$.
+Therefore,  the columns of $\mX$ are independent solutions to the system $\mA x = 0$ Next, we show that any vector $ x \in N(\mA)$ is a linear combination of the columns of $\mX$.
 
@@
 This implies that $x_1  = -\mF x_2$, and $x = \begin{bmatrix} x_1 \\ x_2	\end{bmatrix} = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix} x_2 = \mX x_2$.
-Any vector  in $ N(\mA)$ is a linear combination of the columns of $\mX$, i.e.
-$C(\mX) = N(\mA)$.
+Any vector  in $ N(\mA)$ is a linear combination of the columns of $\mX$, i.e.\  $C(\mX) = N(\mA)$.
 Therefore, the rank of nullspace $rank(N(\mA)) = dim(C(\mX)) = n-r$.
@@
 
-In the proof of lemma~\ref{lem:ranknull}, the columns of $\mX = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix}$ are special solutions that compose the basis of $N(\mA)$.
-Practically, we find these special solutions by assigning 1 to a free variable and 0 to other free variables, and then we solve the system $\mA x = 0$.
+In the proof of lemma~\ref{lem:ranknull}, the columns of $\mX = \begin{bmatrix} -\mF \\ \mI_{n-r}	\end{bmatrix}$  compose the basis of $N(\mA)$.
+Practically, we calculate the columns by assigning 1 to a free variable and 0 to other free variables, and then we solve the system $\mA x = 0$.
 
@@
 \begin{proof}
-	 If $\mA x = b$ is solvable, there exists a $x$ such that $\mA x = b$. For any $x$, $\mA x \in C(\mA)$. Therefore, $b \in C(\mA)$.
+	 If $\mA x = b$ is solvable, there exists a vector $x$ such that $\mA x = b$. For any $x$, $\mA x \in C(\mA)$. Therefore, $b \in C(\mA)$.
 \end{proof}
 
-\begin{lem}[Complete solution]\label{lem:compsolu} The complete solution of $\mA x = b$ is given by $x_{comp} = x_p + x_n$, where $x_p$ is a particular solution such that $\mA x_p = b$, and $x_n \in N(\mA)$.
+\begin{lem}[Complete solution]\label{lem:compsolu} The complete solution of $\mA x = b$ is given by $x_{comp} = x_p + x_n$, where $x_p$ is a solution such that $\mA x_p = b$, and $x_n \in N(\mA)$. 
 \end{lem}
@@
 \begin{proof}
-	Suppose $x = x_p + x_0$ is an arbitrary solution to $\mA x= b$. We have 
+	Suppose $x = x_p + x_0$ is a solution to $\mA x= b$. We have 
 	\begin{align}
@@
 
-Usually, we find a particular solution by assigning 0 to free variables, and we solve the system $\mA x = b$.
+Usually, we find a solution $x_p$ by assigning 0 to free variables, and we solve the system $\mA x = b$.
 

### 2020-07-06 skills/math/probability_theory/probability.tex
upload

\documentclass[11pt]{article}

\usepackage{fancybox}

\usepackage{color}
\usepackage{url}
\usepackage[margin=1in]{geometry}

\setlength\parindent{0pt}
\renewcommand{\textfraction}{0.0}
\renewcommand{\topfraction}{1.0}
%\renewcommand{\textfloatsep}{5mm}

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{extarrows}
\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
%\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\mathtoolsset{showonlyrefs=true}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\setcounter{figure}{0}   
\setcounter{table}{0}

\newcommand{\cmt}[1]{{\leavevmode\color{red}{#1}}}

\usepackage{dsfont}
\usepackage{multirow}

\DeclareMathOperator*{\minimize}{minimize}

\usepackage{mathtools}
\mathtoolsset{showonlyrefs}
\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}

\input macros.tex

\title{Summary for Probability Theory}

\date{\today}
\author{%
Jiaxin Hu
}

\begin{document}

% Makes the title and author information appear.

\maketitle

% Abstracts are required.
\section{Preliminary}
\begin{itemize}
	\item \textit{DeMorgan's Laws:} Let $\{A_i \}_{i=1}^{\infty}$ be a collection of set. Then, $(\bigcup_{i = 1}^{\infty} A_i)^c = \bigcap_{i=1}^{\infty} A_i^c$ and   $(\bigcap_{i = 1}^{\infty} A_i)^c = \bigcup_{i=1}^{\infty} A_i^c$.
	\item \textit{Some set operations:} Suppose $A,B$ are two sets. (1) $A-B = A \cap B^c$; (2) $\bigcup_{i= 1}^{\infty} A_i = \{x: x\in A_i \text{ for some }i\}$,  $\bigcap_{i= 1}^{\infty} A_i = \{x: x\in A_i \text{ for all }i\}$.
\end{itemize}

\section{Single variable}
\subsection{Probability and conditional probability}
\begin{defn}[\textit{Sample space}]\label{def:samplespace}
     The set $S$ containing all possible outcomes is called the sample space.
\end{defn}

\begin{defn}[$\sigma$-\textit{field}]\label{def:sfield} 
	A collection $\tF$ of subsets of a sample space $S$ is called a $\sigma$-field (or $\sigma$-algebra) if and only if (\textbf{iff}) it has the following properties: \\(1) The empty set $\varnothing \in \tF$; \\(2) If $A \in \tF$, then the complement $A^c \in \tF$;\\ (3) If $A_i \in \tF, i = 1,2,...$, then their union $\cup_i A_i \in \tF$.\\
	If $A \in \tF$, then $A$ is called an \textit{event}.
\end{defn}

\begin{defn}[\textit{Measure and probability}]\label{def:measure}
	A set function $v$ defined on a $\sigma$-field $\tF$ is called a measure \textbf{iff} it has the following properties:\\
	(1) $0 \leq v(A) \leq \infty$ for any $A \in \tF$;
	\\ (2) $v(\varnothing) = 0$;
	\\(3) If $A_i \in \tF,i = 1,2,...$ and $A_i$'s  are disjoint, i.e.\ $A_i \cap A_j = \varnothing, \forall i\neq j$, then 
	\[ v(\bigcup_{i=1}^{\infty} A_i ) = \sum_{i=1}^{\infty} v(A_i). \]
	If $v(\tF) = 1$, then $v$ is a probability defined on $\tF$ and we use notation $P$ instead of $v$. 
\end{defn}

\begin{thm}[Probability]\label{thm:probability}
	Let the sample space $S = \{s_1, s_2,...\}$ and $\tF$ be all subsets of $S$. Let $p_1, p_2,...$ be non-negative numbers that $\sum_i p_i = 1 $. The following defines a probability on $\tF$
	\[ P(A) = \sum_{i: s_i \in A} p_i, \quad A \in \tF. \]
\end{thm}

\begin{thm}[Properties of probability]
Let $P$ be a probability, $A,B$ be events and $\{A_i\}_{i=1}^{\infty}$ be a collection of event. Let $\{C_i\}_{i=1}^{\infty}$ be a partition of sample space $S$, i.e.\ $C_i \cap C_j, \forall i\neq j$ and $\bigcup_{i=1}^{\infty} C_i = S$.   Then, 
\begin{align}
&1. P(A) \leq 1; P(A^c) = 1 - P(A); P(A) = P(A \cap B) + P(A \cap B^c);\\
&2. \text{If } A \subset B\text{, then }P(A) \leq P(B); \\
&3. \text{(General addition formula)}\\
&P(\bigcup_{i = 1}^n A_i) = \sum_{i = 1}^n P(A_i) - \sum_{i < j} P(A_i \cap A_j) + \sum_{i<j<k} P(A_i \cap A_j \cap A_k) + \cdots +(-1)^{n-1} P(A_1 \cap \cdots \cap A_n);\\	
&4. P(A) = \sum_{i=1}^{\infty} P(A \cap C_i); \\
&5. \text{(Boole's inequality) } P(\bigcup_{i=1}^{\infty} A_i) \leq \sum_{i=1}^{\infty} P(A_i);\\
&6. \text{(Bonferroni's inequality) } P(\bigcap_{i=1}^n A_i) \geq \sum_{i=1}^n P(A_i) - (n-1).
\end{align}
\end{thm}

\begin{defn}[\textit{Conditional Probability}]\label{def:condprobability}
If $A$ and $B$ are events with $P(B) > 0$, then the conditional probability of $A$ given $B$ is \[ P(A | B) = \frac{P(A\cap B)}{P(B)}. \]
For convenience, we define $P(A|B) = 0$ when $P(B) = 0$.
\end{defn}

\begin{thm}[Useful formulas for conditional probability]\label{thm:condprobability}
Let $A,\{A_i\}_{i=1}^{\infty}, B, \{B_i\}_{i=1}^{\infty},C$ be events. Then we have:
\begin{align}
	&1.  P(A|B) = \frac{P(A)P(B|A)}{P(B)}; \\
	&2. P(A^c|B) = 1- P(A|B); P(A \cup C|B) = P(A|B) + P(C|B) - P(A \cap C|B);\\
	&3. P(\bigcap_{i=1}^{n} A_i) = P(A_1) P(A_2|A_1)\cdots P(A_n|\bigcap_{i=1}^{n-1} A_i);\\
	&4. \text{If } \{B_i\}_{i=1}^{\infty} \text{is a partition of }S, P(A) = \sum_{i=1}^{\infty} P(B_i) P(A|B_i).
\end{align}
\end{thm}

\begin{thm}[Bayes formula]\label{thm:bayesformula}
Let $A$ be an event and $\{ B_i\}_{i=1}^{\infty}$ be a partition of $S$. Then,
\[  P(B_i|A) = \frac{P(A|B_i) P(A)}{\sum_{j=1}^{\infty} P(A|B_i) P(B_i) }.  \]
\end{thm}

\begin{defn}[\textit{Independence}]\label{def:indep}
Two events $A,B$ are independent \textbf{iff} 
\[ P(A\cap B) = P(A) P(B) \quad \text{or} \quad P(A|B) = P(A) \quad \text{or} \quad  P(B|A) = P(B).\]
\end{defn}
If $A,B$ are independent, then the following pairs are also independent: $A$ and $B^c$,  $A^c$ and $B$, $A^c$ and $B^c$.

\begin{defn}[\textit{Mutual  and pairwise independence}]\label{def:mindep} A collection of events $A_1,...,A_n$ are mutually independent \textbf{iff} for any sub-collection $A_{i_1},...,A_{i_k},$
\[ P(A_{i_1} \cap \cdots \cap A_{i_k}) = P(A_{i_1})\cdots P(A_{i_K}). \]
The events $A_1,...,A_n$ are pairwise independent \textbf{iff} $A_i$ and $A_j$ are independent for all $i\neq j$.

Mutual independence is stronger than pairwise independence.
\end{defn}

\begin{defn}[\textit{Conditional independence}]\label{def:cindep}
Events $A$ and $B$ are conditionally independent given event $C$ \textbf{iff} 
\[ P(A\cap B|C) = P(A|C)P(B|C).\]
\end{defn}

Let $A,B,C$ are events. Independence does not imply conditional independence:
\[ P(A\cap B) = P(A) P(B) \nRightarrow P(A\cap B|C) = P(A|C) P(B|C). \]
Conditional independence does not imply independence:
\[ P(A\cap B|C) = P(A|C) P(B|C) \nRightarrow P(A\cap B) = P(A) P(B).  \]
Mutual independence implies conditional independence:
\[ A,B,C\text{ mutually independent} \nRightarrow P(A\cap B|C) = P(A|C) P(B|C). \]

\subsection{Random variable and distribution}
\begin{defn}[\textit{(Random variable and distribution)}]\label{def:rv}
 A random variable  $X$ is a function from $S$ to $\bbR$ such that, for any Borel set $\tB \subset \bbR$,    \[ \{X \in \tB\} = \{\omega \in S: X(\omega) \in \tB \}. \]
The induced probability of $X$ is \[ P_X(\tB) = P(X \in \tB) = P(\omega \in \{\omega \in S: X(\omega) \in \tB \}).\]
The probability $P_X$ is called the distribution of $X$.
\end{defn}

\begin{defn}[\textit{Cumulative distribution function(cdf)}]\label{def:cdf}
 The cdf of a random variable $X$, denoted by $F_X(x)$, is defined as
\[ F_X(x) = P(X \leq x), \quad x \in \bbR.\]	
\end{defn}

\begin{thm}[Cdf]\label{thm:cdf} The function $F(x)$ is a cdf \textbf{iff}
\begin{align}
	&1. \lim_{x \rightarrow -\infty} F(x) = 0 \text{ and } \lim_{x \rightarrow \infty} F(x) = 1;\\
	&2. F(x) \text{ is non-decreasing in }x;\\
	&3. F(x) \text{ is right-continuous}: \lim_{\epsilon>0, \epsilon \rightarrow 0} F(x + \epsilon) = F(x), \quad \forall x \in \bbR.
\end{align}

\end{thm}

\begin{defn}[\textit{Continuity of random variable}]\label{def:contrv}
A random variable\ $X$ is continuous if $F_X(x)$ is continuous in $x$. A random variable\ $x$ is discrete if $F_X(x)$ is a step function of $x$.
\end{defn}

Note that the continuity of a random variable depends on the cdf rather than pdf or pmf. There are random variable's that are mixtures of these two types.

\begin{defn}[\textit{Probability mass function(pmf)}]\label{def:pmf} The pmf of a discrete random variable\ $X$ is \[ f_X(x) = P(X = x), \quad x \in \bbR.\]
The cdf of $X$, $F_X(x) = P(X \leq x) = \sum_{k \leq x} f_X(k)$.
\end{defn}

\begin{defn}[\textit{Probability density function(pdf)}] The pdf of a continuous random variable\ $X$ is the function $f_X(x)$ such that
\[ F_X(x) = \int_{-\infty}^x f_X(t) dt, \quad x \in \bbR, \]
if $f_X(x)$ exists.  The continuous random variable $X$ has a pdf \textbf{iff} $F_X$ is absolutely continuous. If $f$ is a pdf, the set $\{x: f(x) >0\}$ is called its support.
\end{defn}

If $F_X$ is differentiable, then $f_X(x) = \frac{d}{dx} F_X(x)$.

\begin{thm}[Pdf]\label{thm:pdf} A function $f(x)$ is a pdf \textbf{iff:}
\begin{align}
	&1. f(x) \geq 0, \quad \forall x \in \bbR;\\
	&2. \int_{-\infty}^{\infty} f(x) dx = 1.
\end{align}
\end{thm}

How to find pdf given cdf? (1) $f_X(x) = F'_X(x)$ for $x$ at which $F_X$ is differentiable; (2) $f_X(x)$ can be any $c \geq 0$ for $x$ at which $F_X$ is not differentiable.

\subsection{Transformation}
Let $X$ be a random variable and $Y= g(X)$, where $g$ is function $\bbR \mapsto \tY$ and $\tY$ is the domain of $Y$. For any $A \in \tY$,
\[  P(Y \in A) = P(g(X) \in A) = P(X \in g^{-1}(A)), \quad \text{where } g^{-1}(A) = \{ x: g(x) \in A \}. \]

Given $F_X$ or $f_X$, we want to obtain $f_Y(y)$. 
If $X$ is discrete, then
\[ f_Y(y) = \sum_{x\in g^{-1}(\{y\}) } P(X =x)  \]

If $f_X$ is continuous and $g$ is a continuously differentiable monotone function, then
\begin{align}
	f_Y(y) = \begin{cases}
		f_X(g^{-1}(y)) \vert \frac{d}{dy} g^{-1}(y)\vert & y \in \tY\\
		0 & \text{otherwise}
	\end{cases}
\end{align}

\begin{thm}[Transformation for continuous random variable]\label{thm:transf}
	Let $X$ be a continuous random variable\ with pdf $f_X$. Suppose there are disjoint $\{A_i\}_{i=1}^{k}$ such that $P(X \in \bigcup_{i = 1}^k A_i) = 1$, and $f_X$ is continuous on each $A_i, i \in [k]$. There are functions $g_1(x),...,g_k(x)$ defined on $A_i, i \in [k]$  respectively, satisfying

	1. $g(x) = g_i(x), \forall x \in A_i$;\\
	2. $g_i(x)$ is strictly monotone on $A_i$;\\
	3. The set $\tY = \{ y: y = g_i(x)  \text{ for some } x \in A_i\}$ is the same for each $i$;\\
	4. $g^{-t}_i (y)$ has a continuous derivative on $\tY$ for each $i$.\\
	Then 
	\begin{align}
		f_Y(y) = \begin{cases} \sum_{i = 1}^k 
		f_X(g_i^{-1}(y)) \vert \frac{d}{dy} g_i^{-1}(y)\vert & y \in \tY\\
		0 & \text{otherwise}
	\end{cases}
	\end{align}
\end{thm}

\begin{example} Suppose a random variable\ $X \sim N(0,1)$. Obtain the distribution of $Y = |X|$.
\end{example}

\begin{proof}
	Let $A_1 = (-\infty, 0)$, $A_2 =(0, +\infty )$ and $\tY = (0,+\infty)$. On $A_1$, $g_1(x) = x$ and $g_1^{-1}(x) = x$. On $A_2$, $g_2(x) = -x$ and $g_2^{-1}(x) = -x$.

	By theorem~\ref{thm:transf}, 
	\[  f_Y(y) = \frac{1}{\sqrt{2\pi}} e^{-x^2/2} + \frac{1}{\sqrt{2\pi}} e^{-x^2/2} =  \sqrt{ \frac{2}{\pi}} e^{-x^2/2}.  \]
\end{proof}

\subsection{Expectation}
\begin{defn}[\textit{Expectation}]\label{def:expectation}
The expected value or mean of a random variable\ $g(x)$ is 
\begin{align}
	\mathbb{E}[g(x)] = \begin{cases} 
 	\int_{-\infty}^{+\infty} g(x) f_X(x) dx \xlongequal{Y = g(x)} \int_{-\infty}^{+\infty} y f_Y(y) dx  & \text{if } X \text{has pdf }f_X\\
 	\sum_x g(x) f_X(x)\xlongequal{Y = g(x)} \sum_x y f_Y(y)  & \text{if } X \text{has pmf }f_X
 \end{cases}
\end{align} 
	provided that $\mathbb{E}[|g(x)|] < \infty$. Otherwise, the expected value of $g(X)$ does not exist.
\end{defn}

\begin{thm}[Properties of expectation]\label{thm:propexpect}
Let $X,Y$ be random variables whose expectations exist. Let $a,b,c$ be constants.\\
1. $\mathbb{E}(aX + bY + c) = a \mathbb{E}(X) + b \mathbb{E}(Y) + c$;\\
2. If $X \geq Y$, then $\mathbb{E} (X) \geq \mathbb{E} (Y) $.
\end{thm}

\begin{thm}[Relationship between expectation and cdf]\label{thm:expectcdf}
	Let $F$ be the cdf of a random variable $X$. If $X$ has a pdf or pmf, then,
	\[ \mathbb{E}[|X|] = \int_{0}^{+\infty} [1 - F(x)] dx + \int_{-\infty}^0 F(x) dx,  \]
	and $\mathbb{E}[|X|] < +\infty $ \textbf{iff} both integrals are finite. In case where $\mathbb{E}[|X|] < +\infty $ , 
	\[   \mathbb{E}[X] = \int_{0}^{+\infty} [1 - F(x)] dx - \int_{-\infty}^0 F(x) dx.  \]
\end{thm}
\begin{proof}
	Without the loss of generality, suppose random variable $X$ has a pdf. Let $f$ be the pdf of $X$. We have
	\begin{align}
		\mathbb{E}[|X|] &= \int_{0}^{+\infty} x f(x) dx - \int_{-\infty}^0 x f(x) dx\\
		&= \int_{0}^{+\infty} \int_{0}^{x}  f(x) dt dx + \int_{-\infty}^0 \int_{x}^0 f(x)dt dx \\
		&=  \int_{0}^{+\infty} \int_{t}^{+\infty}  f(x) dx dt + \int_{-\infty}^0 \int_{-\infty}^{t} f(x) dx dt\\
		&= \int_{0}^{+\infty} [1 - F(t)] dt + \int_{-\infty}^0 F(t) dt.
	\end{align}
		Therefore, $\mathbb{E}[|X|]< +\infty$ \textbf{iff} the two integrals are finite. Simiarly,
		\begin{align}
			\mathbb{E}[X] &= \int_{0}^{+\infty} x f(x) dx + \int_{-\infty}^0 x f(x) dx\\
		&= \int_{0}^{+\infty} [1 - F(t)] dt - \int_{-\infty}^0 F(t) dt.
		\end{align}
\end{proof}

\begin{cor}[Relationship between expectation and cdf]\label{cor:expectcdf}
Let $F$ be the cdf of a random variable $X$. If $X$ has a pdf or pmf, then,
\[  \mathbb{E}[|X|] =  \int_{0}^{+\infty} P(X > x) + P(-X \leq x) dx, \] 
	and \[  \sum_{n=1}^{\infty} P(|X| \geq n) \leq \mathbb{E}[|X|] \leq 1 + \sum_{n=1}^{\infty} P(|X| \geq n). \]
\end{cor}
\begin{proof}
	By the proof of theorem~\ref{thm:expectcdf}, 
	\begin{align}
		\mathbb{E}[|X|] &= \int_{0}^{+\infty} [1 - F(t)] dt + \int_{-\infty}^0 F(t) dt\\
		&= \int_{0}^{+\infty} [1 - F(t)] dt + \int_{0}^{+\infty} F(-t) dt\\
		&= \int_{0}^{+\infty}  P( X > t ) + P( -X \leq t )  dt
	\end{align}
	To show $ \sum_{n=1}^{\infty} P(|X| \geq n) \leq \mathbb{E}[|X|]$, we have
	\begin{align}
		\mathbb{E}[|X|] &= \int_{0}^{+\infty}  P( X > t ) + P( -X \leq t )  dt\\
		&\geq \int_{0}^{+\infty}  P(| X |> t )  dt\\
		&= \sum_{n = 0}^{+\infty} \int_{n}^{(n+1)}  P(| X |> t )  dt\\
		&\geq \sum_{n = 0}^{+\infty} \int_{n}^{(n+1)}  P(| X |\geq (n+1) )  dt\\
		&= \sum_{n=1}^{+\infty}  P(| X |\geq n ) 
	\end{align} 
	To show $\mathbb{E}[|X|] \leq 1 + \sum_{n=1}^{\infty} P(|X| \geq n)$, we have
	\begin{align}
		\mathbb{E}[|X|] &= \int_{0}^{+\infty}  P( X > t ) + P( -X \leq t )  dt\\
		&\leq \int_{0}^{+\infty}  P(| X |\geq  t )  dt\\
		&= \sum_{n = 0}^{+\infty} \int_{n}^{(n+1)}  P(| X |\geq t )  dt\\
		&\leq \sum_{n = 0}^{+\infty} \int_{n}^{(n+1)}  P(| X |\geq n )  dt\\
		&\leq 1+ \sum_{n = 1}^{+\infty}  P(| X |\geq n ) .
	\end{align}
\end{proof}

\end{document}

### 2020-07-07 skills/math/linear_algebra/linear_algebra_summary_p2.tex
linear algebra

\documentclass[11pt]{article}

\usepackage{fancybox}

\usepackage{color}
\usepackage{url}
\usepackage[margin=1in]{geometry}

\setlength\parindent{0pt}
\renewcommand{\textfraction}{0.0}
\renewcommand{\topfraction}{1.0}
%\renewcommand{\textfloatsep}{5mm}

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
%\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\setcounter{figure}{0}   
\setcounter{table}{0}

\newcommand{\cmt}[1]{{\leavevmode\color{red}{#1}}}

\usepackage{dsfont}
\usepackage{multirow}

\DeclareMathOperator*{\minimize}{minimize}

\mathtoolsset{showonlyrefs}
\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}

\input macros.tex

\title{\textbf{Linear Algebra -- Part II}\\A summary for MIT 18.06SC}

\date{\today}
\author{%
Jiaxin Hu
}

\begin{document}

% Makes the title and author information appear.

\maketitle

% Abstracts are required.
\section{Orthogonality}

\subsection{Orthogonal vectors and subspaces}
\begin{defn}[\textit{Orthogonal vectors}]\label{def:vortho}
	Suppose two vectors $x,y \in \bbR^{n}$. The vectors $x$ and $y$ are orthogonal \textbf{iff} $x^T y = y^T x = 0$, denoted $x \perp y$.
\end{defn}

\begin{defn}[\textit{Orthogonal subspaces}]\label{def:sortho}
	Suppose two subspaces $S,T$. The subspaces $S$ and $T$ are orthogonal \textbf{iff} for any $s \in S$ and for any $ t \in T$, $s^T t = t^T s = 0$, denoted $S \perp T$.
\end{defn}

Given a matrix $\mA \in \bbR^{m \times n}$, there are four subspaces related to $\mA$: column space $C(\mA)$, row space $C(\mA^T)$, nullspace $N(\mA)$, and left nullspace $N(\mA^T)$. Suppose the matrix rank of $\mA$ is $rank(\mA) = r$, the dimensions of these subspaces are: 
\[ dim(C(\mA)) = dim(C(\mA^T)) = r,\quad  dim(N(A)) = n-r,\quad dim(N(\mA^T)) = m-r. \]

\begin{thm}[Orthogonality of matrix subspaces]\label{thm:ortho}
	Suppose a matrix $\mA \in \bbR^{m \times n}$. The row space $C(\mA^T)$ and the nullspace $N(\mA)$ are orthogonal; the column space $C(\mA)$ and the left nullspace $N(\mA^T)$ are orthogonal i.e.\
	\[ C(\mA^T) \perp N(\mA); \quad C(\mA) \perp N(\mA^T).  \]
\end{thm}

\begin{proof}
	Consider the matrix $\mA \in \bbR^{m \times n}$. For any vector $x \in N(\mA)$, we have $\mA x = 0$. 
	\begin{align}
		\mA x  = \begin{bmatrix}
			a_1^T x \\ \vdots \\ a_m^T x 
		\end{bmatrix} = \begin{bmatrix}
			0\\ \vdots \\0
		\end{bmatrix},
	\end{align}
	where $a_i \in \bbR^n, i \in [m]$ is the $i$-th  row of $\mA$. By the definition~\ref{def:vortho}, $x$ is orthogonal with the rows in matrix $\mA$. For any vector $v \in C(\mA^T)$, $v$ is a linear combination of the rows, i.e.\ $v = c_1 a_1 + ...+c_m a_m$, where $c_i, i\in [m]$ are constants. Multiplying vectors $v$ and $x$,
	\[  v^T x =  c_1 a_1^T x + \cdots + c_m a_m^T x = 0.  \]
	Therefore, $v \perp x$, and $N(\mA) \perp C(\mA^T)$.

	Similarly, for any $x \in N(\mA^T)$, we have $\mA^T x = 0$, which implies $N(\mA^T) \perp C(\mA)$. 
\end{proof}

\begin{thm}[Relationship between $\mA^T \mA$ and $\mA$]\label{thm:aa}
	Consider a matrix $\mA \in \bbR^{m \times}$. We have
	\begin{equation}
		N(\mA^T \mA) = N(\mA) ; \quad rank(\mA^T \mA) = rank(\mA).
	\end{equation}
\end{thm}
\begin{proof}
	Consider the matrix $\mA \in \bbR^{m \times n}$.

	If $x \in N(\mA)$, then $\mA x = 0 \Rightarrow \mA^T 0 = 0$, which implies that for any $x \in N(\mA)$, the vector $x \in N(\mA^T \mA)$. Therefore, we only need to prove that for any $x \in N(\mA^T \mA)$, the vector $x \in N(\mA)$. We prove this by contradiction.

	Suppose a vector $x \in N(\mA^T \mA^T)$ but $x \notin N(\mA)$. We have
	\[ \mA x = b \neq 0, \quad \mA^T \mA x = 0 \quad \Rightarrow \quad \mA^T b = 0. \]
	By the first equation $b \in C(\mA)$, and by the third equation $ b \in N(\mA^T)$ . This contradicts the theorem~\ref{thm:ortho}, i.e.\ $C(\mA) \perp N(\mA^T)$.

	Next, given $N(\mA^T \mA) = N(\mA)$, the rank of matrix $rank(\mA^T \mA) = n - dim(N(\mA^T \mA)) = n - dim(N(\mA)) = rank(\mA)$.
\end{proof}

\begin{cor}[Invertibility of $\mA^T \mA$]\label{cor:invert}
	If $\mA$ has independent columns, then $\mA^T \mA$ is invertible.
\end{cor}
\begin{proof}
	If $\mA \in \bbR^{m \times n}$ has independent columns, then $rank(\mA) = n$. By the theorem~\ref{thm:aa}, $rank(\mA^T \mA) = rank(\mA) = n$. Since $\mA^T \mA \in \bbR^{n \times n}$ is a square matrix, $\mA^T \mA$ is invertible.
\end{proof}

\subsection{Projections onto subspaces}
\begin{defn}[\textit{Projection and projection matrix}]\label{def:project}
	Consider a vector $x \in \bbR^{m}$ and a matrix $\mA^{m \times n}$ that has independent columns. Suppose a vector $p \in C(\mA)$, such that

	\begin{equation}\label{eq:project}
		(x-p) \perp C(\mA). 
	\end{equation}

	The vector $p$ is the projection of vector $x$ onto the space $C(\mA)$. Since $p \in C(\mA)$, there exists a vector $\hat x$ such that $p = \mA \hat x$. By equation~\eqref{eq:project}, we have
	\[ \mA^T (x - p) = \mA^T(x - \mA \hat x) \quad  \Rightarrow\quad  \hat x = (\mA^T \mA)^{-1} \mA^T x,\quad p = \mA(\mA^T \mA)^{-1} \mA^T x.  \] 
	The matrix $\mP =  \mA(\mA^T \mA)^{-1} \mA^T$ is called projection matrix.
\end{defn}

\begin{thm}[Properties of projection matrix]\label{thm:projectm}
	Consider a projection matrix $\mP$ to the column space $C(\mA)$, where $\mA \in \bbR^{m \times n}$ is a matrix. Then,
	\begin{equation}\label{eq:projectm}
		\mP^T = \mP;\quad  \mP^2 = \mP.
	\end{equation}
\end{thm}
\begin{proof}
	Since the projection matrix $\mP =   \mA(\mA^T \mA)^{-1} \mA^T$ and $\mA^T \mA$ is symmetric, then
	\[  \mP^T = (\mA(\mA^T \mA)^{-1} \mA^T )^T = \mA (\mA^T \mA)^{-1} \mA^T = \mP. \]
	\[  \mP^2 = \mP^T \mP =  \mA (\mA^T \mA)^{-1} \mA^T  \mA (\mA^T \mA)^{-1} \mA^T =  \mA (\mA^T \mA)^{-1} \mA^T = \mP. \]
\end{proof}

\begin{cor}[Projection onto $N(\mA^T)$]\label{cor:iprojectm}
	Suppose $\mP$ is a projection matrix in theorem~\ref{thm:projectm} , then $I-\mP$ is also a projection matrix to the left nullspace $N(\mA^T)$. 
\end{cor}

\begin{proof}
	For any $x \in \bbR^m$, we have $x - \mP x \perp C(\mA)$
	\[ x - \mP x \perp C(\mA) \quad \Rightarrow \quad (I-\mP)x \perp C(\mA) \quad \Rightarrow \quad  (I-\mP)x \in N(\mA^T) \text{ and } (x - (I-\mP)x) \perp N(\mA^T). \]
	Therefore, $I-\mP$ is a projection matrix to the left nullspace $N(\mA^T)$.
\end{proof}

\subsection{Projection matrices and least squares}
Given observation vector $y \in \bbR^n$ and the design matrix $\mX \in \bbR^{n \times (k+1)}$, we propose the linear regression model
\[ y = \mX \beta +\epsilon,\]
where $\beta = (\beta_0,\beta_1,...,\beta_k)$ are coefficients of our interests and $\epsilon$ is the noise. The least square estimate of $\beta$ minimizes the loss
\[  \hat \beta_{LS} = \argmin_{\beta \in \bbR^{k+1}} \onorm{y - \mX \beta}^2, \]
where $\onorm{\cdot}$ is the euclidean norm. The vector $\mX 
\hat \beta_{LS}$ can be considered as a the projection of $y$ onto the column space of $\mX$. Therefore, we may use projection tools to solve the least square estimate.

The projection $\mX \hat \beta_{LS}$ satisfies
\[ \mX^T( y - \mX \hat \beta_{LS} ) = 0 \quad \Rightarrow \quad \hat \beta_{LS} = (\mX^T \mX)^{-1} \mX^T y. \]

The estimate $\hat \beta_{LS}$ is identical to the estimates solved by other methods  using the derivative.

\subsection{Orthogonal matrices and Gram-Schimidt}
\begin{defn}[\textit{Orthonormal vectors}]\label{def:othronov}
	The vectors $q_1,...,q_n$ are orthonormal if
	\begin{align}
		q_i^T q_j = \begin{cases}
			0& \text{if } i\neq j\\
			1 & \text{if } i =  j
		\end{cases}.
	\end{align}
\end{defn}

Orthonormal vectors are always independent.

\begin{defn}[Orthonormal matrix and orthogonal matrix]\label{def:orthonom}
	Consider a matrix $\mQ \in \bbR^{m \times n}$. If the columns of $\mQ$ are orthonormal, the matrix $\mQ$ is an orthonormal matrix. If $ m= n$, the square matrix $\mQ$ is a orthogonal matrix.
\end{defn}

If $\mQ \in \bbR^{m \times n}$ is an orthonormal matrix, $\mQ^T \mQ = \mI_n$.  If $\mQ$ is an orthogonal matrix, $\mQ^T = \mQ^{-1}$.
For orthonormal matrix $\mQ$, the projection matrix to $C(\mP)$  becomes $\mP = \mI_m$.

\begin{defn}[Gram-Schimidt Process and QR decomposition]\label{gram-schimidt}
Consider a matrix $\mA \in \bbR^{m \times n}$ with $rank(\mA) = n$. Gram-Schimidt process finds the orthonormal basis for $C(\mA)$. Let $a_i \in \bbR^{m}, i \in [m]$ be the columns of matrix $\mA$. 
\begin{align}
	&u_1 = a_1, &&\quad e_1 = \frac{u_1}{\onorm{u_1}}\\
	&u_2 = a_2 - \frac{u_1^T a_2}{u_1^T u_1} u_1 ,&&\quad  e_2 = \frac{u_2}{\onorm{u_2}}\\
	&u_3 = a_3 - \frac{u_1^T a_3}{u_1^T u_1} u_1 - \frac{u_2^T a_3}{u_2^T u_2} u_2,&&\quad  e_2 = \frac{u_3}{\onorm{u_3}}\\
	\vdots
\end{align}
 The vectors $e_1,...,e_n$ are orthonormal basis of the $C(\mA)$. By matrix operations, we obtain a decomposition of matrix $\mA$
 \begin{equation}\label{eq:qr}
 	\mA = [a_1,...,a_n] = [e_1,....,e_2] \begin{bmatrix}
 		e_1^T a_1 & e_1^T a_2 & \cdots & e_1^T a_n\\
 		0&e_2^T a_2 & \cdots & e_2^T a_n\\
 		\vdots &\vdots & & \vdots \\
 		0 &0&\cdots & e_n^T a_n
 	\end{bmatrix} = \mQ \mR
 \end{equation}
	where $\mQ \in \bbR^{m \times n}$ is an orthonormal matrix and $\mR \in \bbR^{n \times n}$ is a upper triangular matrix. We call the matrix decomposition as equation~\eqref{eq:qr} QR decomposition.
\end{defn}

\end{document}

### 2020-07-07 skills/math/tail_bound/tailbound_solution.tex
solution

\documentclass[11pt]{article}

\usepackage{fancybox}

\usepackage{color}
\usepackage{url}
\usepackage[margin=1in]{geometry}

\setlength\parindent{0pt}
\renewcommand{\textfraction}{0.0}
\renewcommand{\topfraction}{1.0}
%\renewcommand{\textfloatsep}{5mm}

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{extarrows}
\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
%\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\mathtoolsset{showonlyrefs=true}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\setcounter{figure}{0}   
\setcounter{table}{0}

\newcommand{\cmt}[1]{{\leavevmode\color{red}{#1}}}

\usepackage{dsfont}
\usepackage{multirow}

\DeclareMathOperator*{\minimize}{minimize}

\usepackage{mathtools}
\mathtoolsset{showonlyrefs}
\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}

\input macros.tex

\title{Solution to ``Chapter 2: Basic tail and concentration bounds"}

\date{\today}
\author{%
Jiaxin Hu
}

\begin{document}

% Makes the title and author information appear.

\maketitle

\section{Summary}
\begin{thm}[Markov's inequality]\label{thm:markov}
	Suppose $X \geq 0$ is a random variable  with finite mean, we have
	\begin{equation}
		\bbP(X\geq t) \leq \frac{E[X]}{t},\quad \forall \ t > 0.
	\end{equation}  
\end{thm}

\begin{thm}[Chebyshev's inequality]\label{thm:cheb}
	Suppose $X \geq 0$ is a random variable  with finite mean $\mu$ and finite variance, we have
	\begin{equation}
		\bbP(|X - \mu |\geq t) \leq \frac{var(X)}{t^2},\quad \forall \ t > 0.
	\end{equation}  
\end{thm}

\begin{thm}[Markov's inequality for polynomial moments]\label{thm:polymoment}
	Suppose the random variable $X$ has a central moment of order $k$. Applying Markov's inequality to the random variable $|X - \mu|^k$ yields
	\begin{equation}
		\bbP(|X - \mu |\geq t) \leq \frac{\bbE[|X-\mu|^k]}{t^k},\quad \forall \ t > 0.
	\end{equation}  
\end{thm}

\begin{thm}[Chernoff bound]\label{thm:chernoff}
	Suppose the random variable $X$ has a moment generating function in the neighborhood of 0, i.e.\ $\varphi_X(\lambda) = \bbE[e^{\lambda X}] < +\infty, \forall \lambda \in (-b,b), b>0$. Applying Markov's inequality to the random variable $Y = e^{\lambda(X- \mu)}$ yields
	\begin{equation}
		\bbP((X - \mu) \geq t) \leq \frac{\bbE[e^{\lambda X}]}{e^{\lambda t}}.
	\end{equation}
	Optimizing the choice of $\lambda$ for the tightest bound yields the Chernoff bound
	\begin{equation}
		\bbP((X - \mu) \geq t) \leq \inf_{\lambda \in [0,b)} \frac{\bbE[e^{\lambda X}]}{e^{\lambda t}}.
	\end{equation}
\end{thm}

\section{Exercises}

\paragraph{Exercise 2.1} (Tightness of inequalities.) The Markov and Chebyshev's inequalities can not be improved in general. 
\begin{enumerate}
	\item[(a)] Provide a random variable $X \geq 0$ for which Markov's inequality~\eqref{thm:markov} is met with equality.  
	\item[(b)] Provide a random variable $Y$ for which Chebyshev's inequality~\eqref{thm:cheb} is met with equality.  
\end{enumerate}

\paragraph{Solution:} 
\begin{enumerate}
	\item[(a)] Recall the proof of Markov's inequality. For any $t >0$,
	\[ \bbE[X] = \int_{0}^t x f_X(x) dx + \int_{t}^{+\infty} x f_X(x) dx \geq \int_{t}^{+\infty} x f_X(x)\geq t \int_{t}^{+\infty} f_X(x) = t \bbP( X \geq t).\]
	 If Markov's inequality meets the equality, the inequalities above should meet equality.

	 Consider a variable $X$ with distribution $P(X = 0) = 1$. For any $t > 0$, the variable $X$ satisfies 
	 \[ \int_{0}^t x f_X(x) dx = 0 \text{ and } \int_{t}^{+\infty} x f_X(x) dx =\int_{t}^{+\infty} t f_X(x) dx. \]
	 Therefore, for variable $X$, the Markov's inequality is met with equality.
	 \item[(b)] Chebyshev's inequality follows by applying Markov's inequality to the non-negative random variable $Y = \bbE{(X - \bbE[X])^2}$. Let the distribution of $Y$ be $\bbP(Y = 0) = 1$. Then the Markov's inequality for $Y$ and the Chebyshev's inequality for $X$ meet the equalities. By transformation, the distribution of random variable $X$ is $\bbP(X = \bbE[X]) = 1$. Therefore, for any random variable $X$ with distribution $\bbP(X = c) = 1, c \in \bbR$, the Chebyshev's inequality is met with equality.
\end{enumerate}

\paragraph{Exercise 2.2}
\begin{lem}[Standard normal distribution]\label{lem:n01}
	Let $\phi(z)$ be the density function of a standard normal $Z \sim N(0,1)$ variable. Then, 
	\begin{equation}\label{eq:e22a}
		\phi'(z) + z \phi(z) = 0,
	\end{equation}
	and \begin{equation}\label{eq:e22b}
		\phi(z)(\frac{1}{z} - \frac{1}{z^3} ) \leq \bbP(Z \geq z) \leq \phi(z)(\frac{1}{z} - \frac{1}{z^3}+ \frac{3}{z^5} ), \quad \text{for all }z>0.
	\end{equation}
\end{lem}
\begin{proof}First, we prove the equation~\eqref{eq:e22a}. The pdf of standard normal distribution $\phi(z)$ satisfies
\[  \phi(z) = \frac{1}{\sqrt{2\pi}} \exp(-\frac{z^2}{2}  ); \quad \phi'(z) = -z \frac{1}{\sqrt{2\pi}} \exp(-\frac{z^2}{2} ) = -z \phi(z). \]
	Next, we prove the equation~\eqref{eq:e22b}. Using equation~\eqref{eq:e22a}, we have
	\begin{align}
		\bbP(Z \geq z) &= \int_{z}^{+\infty} \phi(t) dt = \int_{z}^{+\infty} - \frac{1}{t} \phi'(t) dt = \frac{1}{z} \phi(z) - \int_{z}^{+\infty} \frac{1}{t^2} \phi(t) dt\\
		&= \frac{1}{z} \phi(z) + \int_{z}^{+\infty} \frac{1}{t^3} \phi'(t) dt =  \frac{1}{z} \phi(z) - \frac{1}{z^3} \phi(z) + \int_{z}^{+\infty} \frac{3}{t^4} \phi(t) dt
	\end{align}
	Since $\frac{3}{t^4} \phi(t) \geq 0$, therefore $\bbP(Z \geq z) \geq \phi(z)(\frac{1}{z}  - \frac{1}{z^3} )$. On the other hand, 
	\[ \int_{z}^{+\infty} \frac{3}{t^4} \phi(t) dt = \int_{z}^{+\infty} - \frac{3}{t^5} \phi'(t) dt =\frac{3}{ z^5} \phi(z) -  \int_{z}^{+\infty} \frac{15}{t^6} \phi(t) dt \leq \frac{3}{ z^5} \phi(z). \]
	Therefore, $\bbP(Z \geq z) \leq \phi(z)(\frac{1}{z}  - \frac{1}{z^3} + \frac{3}{z^5} ).$
\end{proof}

\paragraph{Exercise 2.3}

\begin{lem}[Polynomial bound and Chernoff bound]\label{lem:momcher}
	Suppose $X \geq 0$, and that the moment generating function of $X$ exists in the neighborhood of 0. Given some $\delta > 0$ and integer $ k \in \mathbb{Z_+}$, we have
	\[  \inf_{k \in \mathbb{Z_+} } \frac{\bbE[|X|^k]}{\delta^k} \leq \inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}}. \]
	Consequently, an optimized bound based on polynomial moments is always at least as good as the Chernoff upper bound.
\end{lem}

\begin{proof}
	By power series, we have
	\begin{equation}\label{eq:power}
		 e^{\lambda X} = \sum_{k = 0}^{+\infty } \frac{X^k \lambda^k }{k!}, \quad \forall \lambda \in \bbR
	\end{equation}
	Since the moment generating function $\varphi_X(\lambda)$ exists in the neighbor hood of 0, there exists a constant $b > 0$ such that 
	\[  \inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}} = \inf_{\lambda \in (0,b)} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}} < +\infty. \]
	Taking the expectation on both sides of the power series~\eqref{eq:power} yields
	\[   \bbE [e^{\lambda X} ]= \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} < +\infty,\quad \forall \lambda \in (0,b). \]
	Therefore, the moment $\bbE[|X|^k] < +\infty,\  \forall k \in \mathbb{Z}_+$ exists. Applying the power serious to $e^{\lambda \delta}$, we obtain the result
	\[ \inf_{k \in \mathbb{Z_+} } \frac{\bbE[|X|^k]}{\delta^k} \leq \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k] }{\delta^k} = \sum_{k = 0}^{+\infty } \frac{\frac{ \bbE[|X|^k]  \lambda^k }{k!} }{\frac{\lambda^k \delta^k }{k!}  } = \inf_{\lambda > 0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}}.  \]
\end{proof}

\end{document}

### 2020-07-08 skills/math/linear_algebra/linear_algebra_summary_p2.tex
linear_algebra

@@
 
+\section{Determinants}
+The \textit{determinant} is a number associated with any square matrix. For a square matrix $\mA$, let $\det(\mA)$ or $|\mA|$ denote the determinant of matrix $\mA$.
+
+\subsection{Properties of determinants}
+We give ten properties of determinants. The last seven properties are deduced by first three basic properties
+\begin{enumerate}
+	\item\label{prop1} The determinant of identity matrix is equal to 1, i.e.\ $\det(\mI) = 1$.
+	\item\label{prop2} Exchanging two rows of a matrix reverses the sign of the matrix's determinant.
+Hence, an odd number of row exchanges reverse the sign of the determinant while an even number of row exchanges maintain the sign of the determinant.
+	\item \begin{enumerate} 
+	\item[(a)]\label{prop3a} If one row of the matrix is multiplied by a constant $t$, the determinant of the matrix is multiplied by $t$.
+	\begin{align}
+		\begin{vmatrix}
+			ta&tb\\c &d 
+		\end{vmatrix} = t \begin{vmatrix}
+			a&b\\c &d
+			\end{vmatrix}
+	\end{align} 
+	\item[(b)]\label{prop3b} The determinant is linearly additive on the rows of the matrix.
+	\begin{align}
+		\begin{vmatrix}
+			a+a'&b+b'\\c &d 
+		\end{vmatrix} =  \begin{vmatrix}
+			a&b\\c &d
+			\end{vmatrix} +  \begin{vmatrix}
+			a'&b'\\c &d
+			\end{vmatrix} 
+	\end{align}
+	\end{enumerate}
+
+	\item \label{propidenticalrow} If two rows of a matrix are equal, the determinant of the matrix is 0.  
+	\begin{proof}
+		Suppose $\mA$ has two identical rows $a_i, a_j$.
+Exchanging $a_i, a_j$, the exchanged matrix $\mA' = \mA$.
+By property~\ref{prop2}, $\det(\mA) = -\det(\mA') = -\det(\mA)$.
+Therefore, $\det(\mA) = 0$.
+	\end{proof}
+
+	\item \label{propelim} For a square matrix $\mA \in \bbR^{n \times n}$, subtracting $t$ times of the $i$-th row from the $j$-th row does not change the determinant of $\mA$, where $i \neq j$.
+	\begin{proof}
+		By property~\ref{prop3a}, property~\ref{prop3b}, and property~\ref{propidenticalrow}, 
+	\begin{align}
+		 \begin{vmatrix}
+			a & b \\ c - ta&d -tb
+		\end{vmatrix} = \begin{vmatrix}
+			a & b \\ c &d
+		\end{vmatrix} -t \begin{vmatrix}
+			a & b \\ a& b
+		\end{vmatrix}  = \begin{vmatrix}
+			a & b \\ c &d.
+		\end{vmatrix}
+	\end{align}
+	The proof for higher-dimension matrix is similar.
+	\end{proof}
+
+	\item\label{prop0row} If the matrix $\mA$ has a row that is all zeros, the determinant of $\mA$ is 0.
+	\begin{proof}
+		By property~\ref{prop3a}, letting $t = 0$ leads to property~\ref{prop0row}.
+	\end{proof}
+
+	\item\label{propdiag} Suppose $\mA \in \bbR^{n \times n}$ is a triangular matrix with diagonal elements $d_1,...,d_n$. The determinant of $\mA$ is equal to the product $d_1 ...d_n$.
+
+	\begin{proof}
+		By property~\ref{propelim}, eliminating the triangular matrix $\mA$ to a diagonal matrix does not change the determinant of $\mA$.
+By property~\ref{prop3a} and \ref{prop1}, $\det(\mA) = d_1 ...d_n \det(\mI_n) = d_1 ...d_n $.
+	\end{proof}
+
+	\item\label{propsingular} If the square matrix $\mA$ is singular, the determinant of $\mA$ is 0.
+
+	\begin{proof}
+		If $\mA$ is a singular matrix, the reduced row echelon form of $\mA$, RREF($\mA$), has a row with all 0 elements. By property~\ref{propelim} and \ref{prop0row}, the determinant of $\mA$ is 0.
+	\end{proof}
+
+	\item\label{propproduct} For square matrices $\mA, \mB \in \bbR^{n \times n}$, the determinant of the product satisfies $\det(\mA \mB) = \det(\mA) \det(\mB).$
+
+	\begin{proof}
+		If at least one of $\mA, \mB$ is singular, the product $\mA \mB$ is also singular.
+Then, $\det(\mA \mB) = \det(\mA) \det(\mB) = 0$.
+If both $\mA, \mB$ are invertible, $\mA, \mB$ can be eliminated to diagonal matrices $\mQ_{A}, \mQ_{B}$ by QR decomposition.
+Suppose $a_1,...,a_n$ and $b_1,...,b_n$ are diagonal elements for $\mQ_A$ and $\mQ_B$ respectively.
+By property~\ref{propdiag}, $\det(AB) = \prod_{i=1}^n a_i b_i = \det(\mA) \det(\mB)$.
+	\end{proof}
+
+	\item\label{proptrans} For a square matrix $\mA$, the determinant $\det(\mA^T) = \det(\mA)$.
+
+	\begin{proof}
+		By LU decomposition, the matrix $\mA = \mL \mU$, where $\mL$ is a lower-triangular matrix and $\mU$ is an upper-triangular matrix.
+The transport $\mL^T ,\mU^T$ have the same diagonal elements with $\mL, \mU$ respectively.
+By property~\ref{propdiag}, $\det(\mL^T)= \det(\mL),  \det(\mU^T) = \det(\mU)$.
+Therefore, by property~\ref{propproduct}, $\det(\mA) = \det(\mL) \det(\mU) = \det(\mU^T) \det(\mL^T) = \det(\mU^T \mL^T) = \det(\mA^T)$.
+			\end{proof}
+\end{enumerate}
+
+\subsection{Determinant computation}
+\begin{prop}[Big formula for computing determinant]
+	Consider a square matrix $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$. The big formula for computing the determinant of $\mA$ states below.
+	\[ \det(\mA) =  \sum_{(\alpha_1,...,\alpha_n) \in \tP} (-1)^{N(\alpha_1,...,\alpha_n)} a_{1,\alpha_1} ... a_{n \alpha_n},  \]
+	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$ and $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of element exchanges from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
+\end{prop}
+
+\begin{proof}
+	For a square matrix $\mA = \entry{a_ij} \in \bbR^{n \times n}$, we decompose the  matrix $\mA$ by the summation of $n^n$ matrices, in which there is only one entry comes from $\mA$ in each row and other entries are 0.
+By property~\ref{prop3b}, the determinant of $\mA$ is equal to the summation of the determinants of these $n^n$ matrices.
+Take a 3-by-3 square matrix as an example.
+	\begin{align}
+		\begin{vmatrix}
+			a_{11} & a_{12} & a_{13}\\
+			a_{21} & a_{22} & a_{23}\\
+			a_{31} & a_{32} & a_{33}
+		\end{vmatrix} = \begin{vmatrix}
+			a_{11} & 0& 0\\
+			a_{21} & 0 & 0\\
+			a_{31} &0 & 0
+		\end{vmatrix} + \begin{vmatrix}
+			a_{11} & 0& 0\\
+			0 & a_{22} & 0\\
+			a_{31} &0 & 0
+		\end{vmatrix} + \cdots +\begin{vmatrix}
+			0 & 0& a_{13}\\
+			0& 0 & a_{23}\\
+			 0&0 & a_{33}
+		\end{vmatrix} \overset{\Delta}{=} \mA_1 + \mA_2 +\cdots + \mA_{27}.
+	\end{align}
+	For the square matrix $\mA$,  there are $n!$ non-singular matrices among the $n^n$ matrices.
+The number of non-singular matrices coincides with the size of $\tP$.
+Intuitively, there are $n$ ways to choose an element from the first row, after which there are only $n-1$ ways to choose an element from the second row to avoid the zero determinant.
+Therefore, we have $n \times (n-1) \times (n-2) \times \cdots \times 2 = n!$ non-singular matrices.
+
+	By property~\ref{propdiag} and \ref{prop2}, the determinant for the non-singular matrix follows the formula \[(-1)^{N(\alpha_1,...,\alpha_n)}  a_{1,\alpha_1} ... a_{n \alpha_n},\]
+	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$ and $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of element exchanges from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
+\end{proof}
+
+\begin{defn}[Cofactors, cofactor matrix, and cofactor formula]
+	Consider a square matrix $\mA=  \entry{a_{ij}} \in \bbR^{n \times n}$.
+Let $\mA_{-i,-j}$ be the submatrix of $\mA$ after removing the $i$-th row and $j$-th column.
+The cofactor associated with $a_{ij}$ is \[C_{ij} = (-1)^{i+j}\det(\mA_{-i,-j}).\]
+	The matrix $\mC = \entry{C_{ij}} \in \bbR^{n \times n}$ is called cofactor matrix.  The cofactor formula of $\det(\mA)$ is 
+	\[ \det(\mA) = \sum_{j}^n a_{ij} C_{ij} = \sum_{j}^n a_{ji} C_{ji} ,\quad \forall i \in [n].  \]
+\end{defn}
+
+\begin{example}[Tridiagonal matrix]
+	One example usage of cofactor formula is computing the determinant of \textit{tridiagonal matrix}.
+A tridiagonal matrix is a matrix in which only non-zero elements lie on or adjacent to the diagonal.
+Let $\mT_n \in \bbR^{n \times n}$ denote the tridiagonal matrix of 1's, i.e.\
+\begin{align}
+	\mT_n = \begin{bmatrix}
+		1 &1&0 &\cdots & 0 &0\\
+		1&1&1&\cdots &0&0\\
+		0&1&1&\cdots &0&0\\
+		\vdots &\vdots &\vdots &&\vdots &\vdots \\
+		0&0&0&\cdots&1&1
+	\end{bmatrix}_{n \times n}
+\end{align}
+
+Let $\mT_{n, -i,-j}$ be the submatrix of  $\mT_n$ after removing the $i$-th row and the $j$-th column.   By cofactor formula, 
+\[  \det(\mT_n) = 1 \times \det(\mT_{n, -1,-1}) - 1 \times \det(\mT_{n,-1,-2}). \]
+By the definition of tridiagonal matrix, $\mT_{n,-1,-1} = \mT_{n-1}$ and $\det( \mA_{n,-1,-2}) = 1 \times \det(\mT_{n-2}) $. Therefore, 
+\[ \det(\mT_n) = \det(\mT_{n-1}) - \det(\mT_{n-2}).  \] 
+\end{example}
+
+\subsection{Inverse matrices}
+Previously, we use Gauss-Jordan elimination to obtain the inverse matrix of the invertible matrix. Here, we apply cofactor formula for computing the inverse matrix.
+
+\begin{thm}[Inverse matrix by cofactors]\label{thm:inver}
+Let $\mA \in \bbR^{n \times n}$ be an invertible matrix and $\mC$ be the cofactor matrix of $\mA$. The inverse matrix of $\mA$ satisfies
+\[  \mA^{-1} = \frac{1}{\det(\mA)} \mC^T. \]	
+\end{thm}
+
+\begin{proof}
+	It is equivalent to  show that $\mA \mC^T = \det(\mA) \mI_n$. By the definition of cofactor matrix,
+	\begin{align}
+		\mA  \mC^T = \begin{bmatrix}
+			a_{11} & a_{12} &\cdots &a_{1n}\\
+			a_{21} & a_{22} &\cdots &a_{2n}\\
+			\vdots & \vdots & & \vdots\\
+			a_{n1} & a_{n2} &\cdots &a_{nn}\\
+		\end{bmatrix} \begin{bmatrix}
+			C_{11} & C_{21} &\cdots &C_{n1}\\
+			C_{12} & C_{22} &\cdots &C_{n2}\\
+			\vdots & \vdots & & \vdots\\
+			C_{1n} & C_{2n} &\cdots &C_{nn}.\\
+		\end{bmatrix}
+	\end{align}
+
+	Therefore, the diagonal element of $\mA \mC^T$ satisfies $(\mA \mC^T)_{ii} = \sum_k^n a_{ik} C_{ik} = \det(\mA)$.
+Consider the off-diagonal entries $(\mA \mC^T)_{ij} = \sum_{k}^n a_{ik} C_{jk}$, where $i \neq j$.
+By cofactor formula, the summation   $ \sum_{k}^n a_{ik} C_{jk}$ is equal to the determinant of a matrix whose $i$-th row and $j$-th row are identical.
+By property~\ref{prop0row}, the off-diagonals of $\mA \mC^T$ are 0.
+Then, we have $\mA \mC^T = \det(\mA) \mI_n$.
+\end{proof}
+
+\begin{defn}[Cramer's rule]
+Consider the linear system $\mA x=b$, where $\mA\in\bbR^{n \times n}$ is an invertible matrix, and $ x, b \in \bbR^n$ are vectors.
+Applying theorem~\ref{thm:inver}, we obtain the Cramer's rule of $x = \mA^{-1} b$.
+\[  x = \frac{1}{\det(\mA)} \mC^T b; \quad x_j = \frac{\det(\mB_j)}{\mA_j}, \] 
+	where $B_j$ the matrix that replaces the $j$-th column of $\mA$ by $b$.
+\end{defn}
+
+\begin{defn}[Parallelepiped]
+	The parallelepiped determined by $n$ vectors $v_1,...,v_n \in \bbR^n$ is the subset 
+	\[ P = \{ a_1 v_1 + \cdots + a_n v_n : 0 \leq a_1,...,a_n \leq 1 \}. \]
+	 We use $vol(P)$ to denote the volume of the parallelepiped  $P$. 
+\end{defn}
+
+\begin{prop}[Determinants and volumes]
+The absolute determinant of a square matrix $\mA$ is the volume of parallelepiped determined by the rows of $\mA$, i.e.\
+	\[ |\det(\mA)| = vol(P), \]
+	where $P$ is the parallelepiped determined by the rows of $\mA$.
+\end{prop}
+
+\begin{example}
+	The area of triangle with vertices at $(x_1,y_1), (x_2,y_2), (x_3,y_2)$ is
+	\begin{align}
+		\frac{1}{2} \begin{vmatrix}
+			x_1 & y_1 & 1\\
+			x_2 & y_2 & 1\\
+			x_3 & y_3 & 1
+		\end{vmatrix}.
+	\end{align}
+
+	\begin{proof}
+		The two edges of the triangle are $v_1 = (x_2 - x_1, y_2 -y_1)$ and $v_2 = (x_3 - x_1, y_3 -y_1)$.
+The area of the triangle is half of the area of parallelepiped determined by $v_1, v_2$.
+Therefore, the area of triangle is
+		\begin{align}
+			\frac{1}{2} \begin{vmatrix}
+				x_2 - x_1 &  y_2 - y_1\\
+				x_3 - x_1& y_3 -y_1
+			\end{vmatrix} =  \frac{1}{2} \left(  \begin{vmatrix}
+				x_2  &  y_2 \\
+				x_3 & y_3
+			\end{vmatrix} + \begin{vmatrix}
+				x_1 &  y_1\\
+				x_2& y_2
+			\end{vmatrix}  - \begin{vmatrix}
+				x_1 &  y_1\\
+				x_3 & y_3 
+			\end{vmatrix}    \right) = \frac{1}{2} \begin{vmatrix}
+			x_1 & y_1 & 1\\
+			x_2 & y_2 & 1\\
+			x_3 & y_3 & 1
+		\end{vmatrix}.
+		\end{align}
+	\end{proof}
+\end{example}
+
 \end{document}

### 2020-07-09 skills/math/tail_bound/tailbound_solution.tex
tail_bound

@@
 
+\begin{thm}[Hoeffding bound for bounded variable]\label{thm:boundvar}
+Consider a random variable $X$ with mean $\mu = \bbE(X)$, and such that $X \in [a,b]$ almost surely, where $a, b$ are two constants. Then, for any $\lambda \in \bbR$, it holds
+\begin{equation}
+	\bbE[e^{\lambda X}] \leq e^{\frac{s(b-a)^2}{8}}.
+\end{equation}
+Particularly, the variable $X \sim subG(\frac{(b-a)^2}{4})$.
+\end{thm}
+
+\begin{proof}
+	See Exercise~\ref{pro:boundvar}.
+\end{proof}
+
 \section{Exercises}
 
-\paragraph{Exercise 2.1} (Tightness of inequalities.) The Markov and Chebyshev's inequalities can not be improved in general. 
+\subsection{Exercise 2.1} (Tightness of inequalities.) The Markov and Chebyshev's inequalities can not be improved in general. 
 \begin{enumerate}
@@
 
-\paragraph{Exercise 2.2}
+\subsection{Exercise 2.2}
 \begin{lem}[Standard normal distribution]\label{lem:n01}
@@
 
-\paragraph{Exercise 2.3}
+\subsection{Exercise 2.3}
 
@@
 
+\subsection{Exercise 2.4}\label{pro:boundvar}
+In Exercise~\ref{pro:boundvar}, we prove theorem~\ref{thm:boundvar}, Hoeffding bound for bounded variable.
+Consider a random variable $X$ with mean $\mu = \bbR[X]$ and such that $a \leq X \leq b$ almost surely.
+Define the function
+\[ \varphi(\lambda) = \log \bbE[e^{\lambda X }]. \]
+We apply Taylor Expansion of $\varphi(\lambda)$ at 0.
+\begin{equation}\label{eq:taylor}
+	\varphi(\lambda) =  \varphi(0) + \varphi'(0) \lambda  + \frac{\varphi^{''}(\lambda_0) }{2} \lambda^2, \quad \lambda_0 = t \lambda, \text{ for some } t \in [0,1]. 
+\end{equation}
+
+In equation~\eqref{eq:taylor}, the term $\varphi(0) =\log \bbE[e^{0}] = 0$. For the first-order derivative $\varphi'(\lambda)$, we apply the power series and have
+\begin{align}
+	\varphi'(\lambda ) &= \left( \log \bbE \left[ \sum_{k = 0}^n \frac{\lambda^k X^k}{k!} \right]   \right)' = \left( \log \sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]  \right)' \\
+	&= {\sum_{k = 1}^n \frac{k \lambda^k }{k!} \bbE[X^k] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]} = {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+1)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]} \\
+	&= \frac{\bbE[X e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]}
+\end{align}
+Therefore, $\varphi'(0) = \bbE[X] = \mu$. For the second-order derivative $\varphi^{''}(\lambda)$,
+\begin{align}
+	\varphi^{''}(\lambda) &= \left( {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+1)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]} \right)^{'} \\
+	&= {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+2)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]} - \left( {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+1)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]}  \right)^2\\
+	&= \frac{\bbE[X^2 e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]} - \left( \frac{\bbE[X e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]} \right)^2\\
+	& \overset{\Delta}{=} \bbE_{\lambda} [X^2] - (\bbE_{\lambda} [X])^2
+\end{align}
+Therefore, the second-order derivative $\varphi''(\lambda)$ can be interpreted as the variance of $X$ with a re-weighted distribution $d P' = \frac{e^{\lambda X}}{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$.
+Since
+\[ \int_{-\infty}^{+\infty} d P' = \int_{-\infty}^{+\infty}  \frac{e^{\lambda X}}{\bbE[ e^{\lambda X} ]} d P_X = 1, \]
+the function $P'$ is indeed a distribution. Under any distribution, we always have
+\[ var(X) = var(X - \frac{a+b}{2}) \leq \bbE[(X - \frac{a+b}{2} )^2] = \frac{(b-a)^2}{4}. \]
+Back to the equation~\eqref{eq:taylor},
+\[\varphi(\lambda) =  \varphi(0) + \varphi'(0) \lambda  + \frac{\varphi^{''}(\lambda_0) }{2} \lambda^2 \leq 0 + \lambda \mu+ \frac{(b-a)^2}{8} \lambda^2 \]
+Taking exponential on both sides of the inequality, we have
+\begin{equation}\label{eq:boundvar}
+	\bbE[e^{\lambda X}] = \exp(\varphi( \lambda)) \leq e^{\mu \lambda + \frac{(b-a)^2}{8} \lambda^2}.
+\end{equation}
+The equation~\eqref{eq:boundvar} implies that $X$ is a sub-Gaussian variable with at most $\sigma = \frac{(b-a)}{2}$.
+
+\subsection{Exercise 2.5}
+\begin{lem}[Sub-Gaussian bounds and means/variance] Consider a random variable $X$ such that
+\begin{equation}\label{eq:subg}
+	 \bbE[e^{\lambda X}] \leq e^{\frac{\lambda^2 \sigma^2}{2} + \mu \lambda}, \quad \forall \lambda \in \bbR. 
+\end{equation}
+	Then, $\bbE[X] = \mu$ and $var(X) \leq \sigma^2$.
+\end{lem}
+
+\begin{proof} First, by equation~\eqref{eq:subg}, the moment generating function of $X$ exists, and thus the mean and variance of $X$ exist.
+Applying power series on both sides of equation~\eqref{eq:subg},
+\begin{equation}\label{eq:taylor2}
+	 \lambda \bbE[X] + \frac{\lambda^2}{2} \bbE[X^2] + o(\lambda^2) \leq  \mu \lambda + \frac{\lambda^2 \sigma^2 + \lambda^2 \mu^2}{2} + o(\lambda^2).
+\end{equation}
+Dividing by $\lambda >0$ on both sides of equation~\eqref{eq:taylor2} and letting $\lambda \rightarrow 0^{+}$, we have $\bbE(X) \leq \mu$ ; dividing dividing by $\lambda <0$ on both sides of equation~\eqref{eq:taylor2} and letting $\lambda \rightarrow 0^{-}$, we have $\bbE(X) \geq \mu$.
+Therefore, the mean $\bbE[X] = \mu$.
+Similarly, we subtract $\bbE[X]\lambda$ and $\mu \lambda$ and divide $\frac{2}{\lambda^2}$ on both sides of equation~\eqref{eq:taylor2}.
+We have $ \bbE[X^2] \leq \sigma^2 + \mu^2$, and thus $var(X) \leq \bbE[X^2]  - (\bbE[X])^2 = \sigma^2$.
+\end{proof}
+
+\paragraph{Question:} Suppose the smallest possible $\sigma$ satisfying the inequality~\eqref{eq:subg} is chosen. Is it true that $var(X) = \sigma^2$?
+
+\paragraph{Solution:} The statement that $var(X) = \sigma^2$ is not always true.
+Recall the function $\varphi(\lambda)$ in exercise~\ref{pro:boundvar}.
+By the results in exercise~\ref{pro:boundvar},  the equation~\eqref{eq:subg} is equal to
+\begin{equation}\label{eq:question}
+	\varphi^{''}(\lambda) \leq  \sigma^2, \quad \forall \lambda \in \bbR,
+\end{equation}
+where $\varphi^{''} (\lambda)$ is the variance of $X$ with with a re-weighted distribution $d P' = \frac{e^{\lambda X}}{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$.
+If the statement that $var(X) = \sigma^2$ is true, then $\max_{\lambda}\varphi^{''}(\lambda) = \varphi^{''}(0)$, which is not always true.
+A counter example is below.
+
+Consider a random variable $X \sim Ber(1/3)$ with $var(X) = 2/9$. Let $\lambda = 1$. The re-weighted distribution $dP'$ is
+\[ P'(X = 0) = \frac{2}{3\bbE[ e^{X} ]};\quad  P'(X = 1) = \frac{e}{3\bbE[ e^{X} ]}, \quad \text{where } \bbE[ e^{X} ] = \frac{2}{3} + \frac{e}{3}.  \]
+Therefore, the variance of $X$ with $dP'$ is $\frac{2}{3\bbE[ e^{X} ]} \times  \frac{e}{3\bbE[ e^{X} ]} = 0.2442 > 2/9$.
+Therefore, the smallest possible $\sigma^2$ is strictly larger than $var(X)$ in this case.
 \end{document}

### 2020-07-10 skills/math/linear_algebra/linear_algebra_summary_p2.tex
linear algebra

@@
 
-\begin{defn}[Orthonormal matrix and orthogonal matrix]\label{def:orthonom}
+\begin{defn}[\textit{Orthonormal matrix and orthogonal matrix}]\label{def:orthonom}
 	Consider a matrix $\mQ \in \bbR^{m \times n}$. If the columns of $\mQ$ are orthonormal, the matrix $\mQ$ is an orthonormal matrix. If $ m= n$, the square matrix $\mQ$ is a orthogonal matrix.
@@
 
-\begin{defn}[Gram-Schimidt Process and QR decomposition]\label{gram-schimidt}
+\begin{defn}[\textit{Gram-Schimidt Process and QR decomposition}]\label{gram-schimidt}
 Consider a matrix $\mA \in \bbR^{m \times n}$ with $rank(\mA) = n$. Gram-Schimidt process finds the orthonormal basis for $C(\mA)$. Let $a_i \in \bbR^{m}, i \in [m]$ be the columns of matrix $\mA$. 
@@
 
-\begin{defn}[Cofactors, cofactor matrix, and cofactor formula]
+\begin{defn}[\textit{Cofactors, cofactor matrix, and cofactor formula}]
 	Consider a square matrix $\mA=  \entry{a_{ij}} \in \bbR^{n \times n}$.
@@
 
-\begin{defn}[Cramer's rule]
+\begin{defn}[\textit{Cramer's rule}]
 Consider the linear system $\mA x=b$, where $\mA\in\bbR^{n \times n}$ is an invertible matrix, and $ x, b \in \bbR^n$ are vectors.
@@
 
-\begin{defn}[Parallelepiped]
+\begin{defn}[\textit{Parallelepiped}]
 	The parallelepiped determined by $n$ vectors $v_1,...,v_n \in \bbR^n$ is the subset 
@@
 
+\section{Eigenvalues and eigenvectors}
+\begin{defn}[\textit{Eigenvalues and eigenvectors}]\label{def:eigen}
+	Consider a square matrix $\mA$, an eigenvector of $\mA$ is a non-zero vector such that:
+	\[ \mA x = \lambda x, \quad \text{for some } \lambda \in \bbR. \]
+	The value $\lambda$ is called the eigenvalue of $\mA$, and $x$ is the eigenvector associated with eigenvalue $\mA$.
+Usually, eigenvectors are normalized, i.e.\ $\onorm{x} = 1$, where $\onorm{\cdot}$ is the euclidean distance.
+
+	Particularly, the vectors associated with eigenvalue 0 consists of the nullspace of $\mA$.
+\end{defn}
+
+\begin{defn}[\textit{Trace of square matrix}]\label{def:trace}
+Consider a square matrix $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$. The trace of $\mA$ is defined as
+\[ tr(\mA) = \sum_{i}^n a_{ii}.\] 
+\end{defn}
+
+\begin{defn}[\textit{Characteristic polynomial}]\label{def:charac}
+	Let $\lambda$ denote the eigenvalue of the matrix $\mA\in \bbR^{n \times n}$. Then the matrix $\mA - \lambda \mI$ is a singular matrix. Therefore,
+	\[ \det(\mA - \lambda I_n) = 0 \quad  \Leftrightarrow \quad (-1)^n \lambda^n + tr(\mA)\lambda^{n-1} + \cdots + (-1)^n \det(\mA) \overset{\Delta}{=} P(\lambda) = 0.   \]
+	 The polynomial $P(\lambda)$ is called the characteristic equation of $\mA$. 
+\end{defn}
+
+Since the characteristic polynomial $P(\lambda)$ is of degree $n$, there are $n$ solutions to the equation $P(\lambda) = 0$.
+The solutions $\lambda_1,...,\lambda_n$ are $n$ eigenvalues of $\mA$, which may be complex and may not be distinct with each other.
+Note that the complex eigenvalues come in conjugate pairs, because $\overline{ P(\lambda)} = P(\overline{ \lambda})$.
+
+\begin{thm}[Summation and production of eigenvalues]\label{thm:speigen}
+	For an square matrix $\mA \in \bbR^{n \times n}$, let $\lambda_1,...\lambda_n$ denote the $n$ eigenvalues of $\mA$. Then 
+	\begin{equation}
+		\sum_{i}^n \lambda_i = tr(\mA); \quad \prod_{i=1}^n \lambda_i = \det(\mA).
+	\end{equation} 
+\end{thm}
+
+\begin{proof}
+	First, we re-write characteristic polynomial $P(\lambda)$ of $\mA$ as
+	\begin{equation}\label{eq:poly}
+		P(\lambda) = (-1)^n (\lambda - \lambda_1)\cdots (\lambda - \lambda_n), 
+	\end{equation}   
+	where $\lambda_i, i \in [n]$ are $n$ eigenvalues of $\mA$.
+By the equation~\eqref{eq:poly}, the coefficient for $\lambda^{n-1}$ is equal to $(-1)^n\sum_{i=1}^n \lambda_i$.
+Compared with definition~\ref{def:charac}, we have $tr(\mA) = \sum_{i=1}^n \lambda_i$.
+Similarly, the constant term in equation~\eqref{eq:poly} is $(-1)^n \prod_{i=1}^n \lambda_1 \cdots \lambda_n $.
+Therefore, $\prod_{i=1}^n = \det(\mA)$.
+\end{proof}
+
+\begin{thm}[Independence of eigenvectors]\label{thm:indepeigen} 
+Consider a square matrix $\mA$ with at least two different eigenvalues. The eigenvectors associated with different eigenvalues are independent.
+\end{thm}
+
+\begin{proof}
+	Suppose the square matrix has $n$ different eigenvalues $\lambda_1,...,\lambda_n$ with associated eigenvectors  $v_1,...,v_n$.
+Let the linear combination $c_1 v_1 + ...
++ c_n v_n = 0$.
+We need to show $c_1 =\cdots = c_n = 0$.
+
+	Then we have
+	\begin{equation}\label{eq:indep1}
+		\mA(c_1 v_1 + ... + c_n v_n ) = c_1 \lambda_1 v_1 + ... + c_n \lambda_n v_n = 0;  
+		\end{equation}
+		\begin{equation}\label{eq:indep2}
+		\lambda_1 (c_1 v_1 + ... + c_n v_n ) = c_1 \lambda_1 v_1 + ... + c_n \lambda_1 v_n = 0. 
+	\end{equation}
+	Subtracting the equations~\eqref{eq:indep1} from \eqref{eq:indep2}, 
+	\begin{equation}\label{eq:indep3}
+		 c_2 (\lambda_1 - \lambda_2) v_2 + \cdots + c_n (\lambda_1 - \lambda_n) v_n  = 0
+	\end{equation}
+	Repeating the steps from equation~\eqref{eq:indep1} to equation~\eqref{eq:indep3}, finally we have 
+	\[ c_n(\lambda_1 - \lambda_n)(\lambda_2 - \lambda_n)\cdots (\lambda_{n-1} - \lambda_n) v_n = 0.    \]
+	 Since $\lambda_1 ,..., \lambda_n$ are distinct, and $v_n$ is a non-zero vector, we have $c_n = 0$.
+Similarly, plugging $c_n = 0$ back to previous steps, we have $c_1 = \cdots = c_n = 0$.
+Therefore, $v_1,...,v_n$ are independent.
+\end{proof}
+
+\begin{thm}[Eigenvalues for symmetric and antisymmetric matrices]\label{def:symmeign}
+	Symmetric square matrices  with all real entries have real eigenvalues.
+For antisymmetric matrices,  i.e.\ $\mA^T = - \mA$, all eigenvalues are imaginary $\lambda = b i$, where $b \in \bbR, i = \sqrt{-1}$.
+\end{thm}
+\begin{proof}
+	First, we prove that symmetric square matrices have real eigenvalues by contradiction.
+
+	Suppose a real and symmetric matrix $\mA$ has a complex eigenvalue $\lambda$ with eigenvector $x$.
+Since $\mA$ is real, then the eigenvector $x$ is also complex.
+Let $\overline \lambda$ and $\overline x$ denote the conjugate eigenvalue and eigenvector, respectively.
+Then we have
+	\begin{equation}\label{eq:symm}
+		\mA x = \lambda x;\quad \mA \overline x = \overline \lambda \overline x.
+	\end{equation} 
+	Premultiply $\overline x^T$ on both sides of the first equation in \eqref{eq:symm} and premultiply $x^T$ on both sides of the second equation in \eqref{eq:symm}. By the symmetry of $\mA$,
+	\[ \overline x^T \mA x  = x^T \mA^T \overline x = x^T \mA \overline x .\]
+	Subtract the two equations in \eqref{eq:symm}, we have
+	\[  0 = ( \overline x^T \mA x -  x^T \mA \overline x ) = (\lambda - \overline \lambda) x^T \overline x.  \]
+	By the fact that $x^T \overline x$ is a real number, the coefficient for $i$ in $\lambda$ is equal to 0, which contradicts that $\lambda$ is complex.
+
+	Next, we prove that the eigenvalues for antisymmetric matrices are imaginary by the similar procedures above. 
+	Suppose $\mA$ is an antisymmetric matrix with complex eigenvalue $\lambda$ and complex eigenvector $x$, we have
+	\[ \overline x^T \mA x  = x^T \mA^T \overline x  =  -x^T \mA \overline x .  \] 
+	Therefore, we have
+	\[  0 = ( \overline x^T \mA x +  x^T \mA \overline x ) = (\lambda + \overline \lambda) x^T \overline x.   \]
+	Since $x^T \overline x$ is a real number, the real part in $\lambda$ is equal to 0. Therefore, eigenvalues for antisymmetric matrices are imaginary.
+\end{proof}
+
+\begin{thm}[Eigenvalues for triangular matrices]\label{thm:trianeigen}
+	For triangular matrices, the eigenvalues are the entries on the diagonal.
+\end{thm}
+
+\begin{proof}
+Suppose $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ is a square matrix. By the property of determinant~\ref{propdiag}, 
+\[ \det(\mA) = \prod_{i=1}^n a_{ii}  ; \quad \det(\mA - \lambda \mI_n) = \prod_{i=1}^n (a_{ii}-\lambda). \]
+To let $\det(\mA - \lambda \mI_n) = 0$, $\lambda = a_{ii}, \forall i\in[n]$. Therefore, the entries on the diagonal are the eigenvalues of a triangular matrix.
+\end{proof}
+
+\begin{thm}[Eigenvalues for powers of the matrix]
+	Consider a matrix $\mA$ with a eigenvalue $\lambda$ and the associated eigenvector $x$. For any polynomial $P$, 
+	\[ P(\mA) x = P(\lambda) x. \]
+\end{thm}
+\begin{proof}
+	For any integer $k \geq 0$ and constant $c \in \bbR$,
+	\[ c\mA^k x = c\mA^{k-1} \mA x= c\lambda \mA^{k-1} x = \cdots = c\lambda^k x.\]
+	Therefore, for any polynomial $P(\mA)x = P(\lambda)x$.
+\end{proof}
+
+\begin{thm}[Eigenvalues for the inverse]
+Suppose $\mA \in \bbR^{n \times n}$ is an invertible matrix.
+Let $\lambda_1,...,\lambda_n$ denote the non-zero eigenvalue of $\mA$, and $x_1,...x_n$eigenvectors corresponds to $\lambda_i$.
+Then  $\lambda_1^{-1},...,\lambda_m^{-1}, \lambda_{m+1},...,\lambda_n $ and $x_1,...x_n$ are eigenvalues and associated eigenvectors of $\mA^{-1}$.
+\end{thm}
+
+\begin{proof}
+	Since the nullspace of $\mA$ contains only a zero vector, $\lambda_i >0 \forall i \in [n]$.
+For any $\lambda_i$ and $x_i$, multiplying $x_i$ on both sides of the equation $\mA^{-1} \mA = \mI_n$, we have
+	\[ \mA^{-1} \mA x_i = x_i \quad \Rightarrow\quad  \mA^{-1} x_i = {\lambda}^{-1}.  \]
+	Therefore, $\mA^{-1}$ share the same eigenvectors with $\mA$, and the reciprocal of the eigenvalues of $\mA$ are the eigenvalues of $\mA^{-1}$.
+\end{proof}
+
 \end{document}

### 2020-07-10 skills/math/tail_bound/tailbound_solution.tex
tail_boung

@@
 	\item[(a)] Recall the proof of Markov's inequality. For any $t >0$,
-	\[ \bbE[X] = \int_{0}^t x f_X(x) dx + \int_{t}^{+\infty} x f_X(x) dx \geq \int_{t}^{+\infty} x f_X(x)\geq t \int_{t}^{+\infty} f_X(x) = t \bbP( X \geq t).\]
-	 If Markov's inequality meets the equality, the inequalities above should meet equality.
+	\begin{equation}\label{eq:markov}
+		\bbE[X] = \int_{0}^t x f_X(x) dx + \int_{t}^{+\infty} x f_X(x) dx \geq \int_{t}^{+\infty} x f_X(x)\geq t \int_{t}^{+\infty} f_X(x) = t \bbP( X \geq t).
+	\end{equation}
+	 If Markov's inequality meets the equality, the inequality~\eqref{eq:markov} meets the equality.
 

### 2020-07-10 skills/math/linear_algebra/linear_algebra_summary_p2.tex
linear_algebra

@@
 \begin{defn}[\textit{Orthogonal vectors}]\label{def:vortho}
-	Suppose two vectors $x,y \in \bbR^{n}$. The vectors $x$ and $y$ are orthogonal \textbf{iff} $x^T y = y^T x = 0$, denoted $x \perp y$.
+	Consider two vectors $x,y \in \bbR^{n}$. The vectors $x$ and $y$ are orthogonal,  denoted $x \perp y$, \textbf{iff} $x^T y = y^T x = 0$.
 \end{defn}
@@
 \begin{defn}[\textit{Orthogonal subspaces}]\label{def:sortho}
-	Suppose two subspaces $S,T$. The subspaces $S$ and $T$ are orthogonal \textbf{iff} for any $s \in S$ and for any $ t \in T$, $s^T t = t^T s = 0$, denoted $S \perp T$.
+	Consider two subspaces $S,T$. The subspaces $S$ and $T$ are orthogonal,  denoted $S \perp T$, \textbf{iff} for any $s \in S$ and $ t \in T$, $s^T t = t^T s = 0$.
 \end{defn}
@@
 Given a matrix $\mA \in \bbR^{m \times n}$, there are four subspaces related to $\mA$: column space $C(\mA)$, row space $C(\mA^T)$, nullspace $N(\mA)$, and left nullspace $N(\mA^T)$.
-Suppose the matrix rank of $\mA$ is $rank(\mA) = r$, the dimensions of these subspaces are:
+Suppose the rank of $\mA$ is $rank(\mA) = r$, the dimensions of these subspaces are:
 \[ dim(C(\mA)) = dim(C(\mA^T)) = r,\quad  dim(N(A)) = n-r,\quad dim(N(\mA^T)) = m-r. \]
@@
 \begin{thm}[Orthogonality of matrix subspaces]\label{thm:ortho}
-	Suppose a matrix $\mA \in \bbR^{m \times n}$. The row space $C(\mA^T)$ and the nullspace $N(\mA)$ are orthogonal; the column space $C(\mA)$ and the left nullspace $N(\mA^T)$ are orthogonal i.e.\
+	Consider a matrix $\mA \in \bbR^{m \times n}$. The row space $C(\mA^T)$ and the nullspace $N(\mA)$ are orthogonal; the column space $C(\mA)$ and the left nullspace $N(\mA^T)$ are orthogonal, i.e.\
 	\[ C(\mA^T) \perp N(\mA); \quad C(\mA) \perp N(\mA^T).  \]
@@
 	where $a_i \in \bbR^n, i \in [m]$ is the $i$-th  row of $\mA$.
-By the definition~\ref{def:vortho}, $x$ is orthogonal with the rows in matrix $\mA$.
-For any vector $v \in C(\mA^T)$, $v$ is a linear combination of the rows, i.e.\ $v = c_1 a_1 + ...+c_m a_m$, where $c_i, i\in [m]$ are constants.
-Multiplying vectors $v$ and $x$,
+By definition~\ref{def:vortho}, $x$ is orthogonal to the rows in matrix $\mA$.
+For any vector $v \in C(\mA^T)$, $v$ is a linear combination of the rows $v = c_1 a_1 + ...+c_m a_m$, where $c_i, i\in [m]$ are constants.
+Multiplying vectors $v$ and $x$, we have
 	\[  v^T x =  c_1 a_1^T x + \cdots + c_m a_m^T x = 0.  \]
@@
 \begin{thm}[Relationship between $\mA^T \mA$ and $\mA$]\label{thm:aa}
-	Consider a matrix $\mA \in \bbR^{m \times}$. We have
+	Consider a matrix $\mA \in \bbR^{m \times n}$. We have
 	\begin{equation}
@@
 \begin{proof}
-	Consider the matrix $\mA \in \bbR^{m \times n}$.
+	Consider a matrix $\mA \in \bbR^{m \times n}$. First, we prove that $N(\mA^T \mA) = N(\mA)$.
 
-	If $x \in N(\mA)$, then $\mA x = 0 \Rightarrow \mA^T 0 = 0$, which implies that for any $x \in N(\mA)$, the vector $x \in N(\mA^T \mA)$.
-Therefore, we only need to prove that for any $x \in N(\mA^T \mA)$, the vector $x \in N(\mA)$.
-We prove this by contradiction.
+	On one hand, if $x \in N(\mA)$, then $\mA x = 0 \Rightarrow \mA^T  \mA x = \mA^T 0 = 0$. Therefore, for any $x \in N(\mA)$, the vector $x \in N(\mA^T \mA)$.
 
-	Suppose a vector $x \in N(\mA^T \mA^T)$ but $x \notin N(\mA)$. We have
-	\[ \mA x = b \neq 0, \quad \mA^T \mA x = 0 \quad \Rightarrow \quad \mA^T b = 0. \]
-	By the first equation $b \in C(\mA)$, and by the third equation $ b \in N(\mA^T)$ . This contradicts the theorem~\ref{thm:ortho}, i.e.\ $C(\mA) \perp N(\mA^T)$.
+	On the other hand, we  prove that for any $x \in N(\mA^T \mA)$ the vector $x \in N(\mA)$ by contradiction.
 
-	Next, given $N(\mA^T \mA) = N(\mA)$, the rank of matrix $rank(\mA^T \mA) = n - dim(N(\mA^T \mA)) = n - dim(N(\mA)) = rank(\mA)$.
+	Suppose there is a vector $x \in N(\mA^T \mA^T)$ but $x \notin N(\mA)$. We have
+	\begin{equation}\label{eq:aaa}
+		\mA x = b \neq 0, \quad \mA^T \mA x = 0 \quad \Rightarrow \quad \mA^T b = 0.
+	\end{equation} 
+	By the first equality in \eqref{eq:aaa}, $b \in C(\mA)$, and by the third equation in \eqref{eq:aaa}, $ b \in N(\mA^T)$.
+This contradicts the theorem~\ref{thm:ortho}, $C(\mA) \perp N(\mA^T)$.
+Therefore, for any $x \in N(\mA^T \mA)$, the vector $x \in N(\mA)$.
+
+	Next, given $N(\mA^T \mA) = N(\mA)$, we have $rank(\mA^T \mA) = n - dim(N(\mA^T \mA)) = n - dim(N(\mA)) = rank(\mA)$.
 \end{proof}
@@
 	If $\mA \in \bbR^{m \times n}$ has independent columns, then $rank(\mA) = n$.
-By the theorem~\ref{thm:aa}, $rank(\mA^T \mA) = rank(\mA) = n$.
+By theorem~\ref{thm:aa}, $rank(\mA^T \mA) = rank(\mA) = n$.
 Since $\mA^T \mA \in \bbR^{n \times n}$ is a square matrix, $\mA^T \mA$ is invertible.
@@
 \begin{defn}[\textit{Projection and projection matrix}]\label{def:project}
-	Consider a vector $x \in \bbR^{m}$ and a matrix $\mA^{m \times n}$ that has independent columns. Suppose a vector $p \in C(\mA)$, such that
-
+	Given a vector $x \in \bbR^{n}$ and a matrix $\mA^{m \times n}$ with independent columns, the vector $p \in C(\mA)$, such that
 	\begin{equation}\label{eq:project}
-		(x-p) \perp C(\mA). 
+		(x-p) \perp C(\mA),
 	\end{equation}
 
-	The vector $p$ is the projection of vector $x$ onto the space $C(\mA)$. Since $p \in C(\mA)$, there exists a vector $\hat x$ such that $p = \mA \hat x$. By equation~\eqref{eq:project}, we have
-	\[ \mA^T (x - p) = \mA^T(x - \mA \hat x) \quad  \Rightarrow\quad  \hat x = (\mA^T \mA)^{-1} \mA^T x,\quad p = \mA(\mA^T \mA)^{-1} \mA^T x.  \] 
-	The matrix $\mP =  \mA(\mA^T \mA)^{-1} \mA^T$ is called projection matrix.
+	is the projection of vector $x$ onto the space $C(\mA)$. Since $p \in C(\mA)$, there exists a vector $\hat x$ such that $p = \mA \hat x$. By equation~\eqref{eq:project}, we have
+	\[ \mA^T (x - p) = \mA^T(x - \mA \hat x) = 0  \quad  \Rightarrow\quad  \hat x = (\mA^T \mA)^{-1} \mA^T x,\quad p = \mA(\mA^T \mA)^{-1} \mA^T x.  \] 
+	The matrix $\mP \overset{\Delta}{=}   \mA(\mA^T \mA)^{-1} \mA^T$ is called a projection matrix.
 \end{defn}
@@
 \begin{cor}[Projection onto $N(\mA^T)$]\label{cor:iprojectm}
-	Suppose $\mP$ is a projection matrix in theorem~\ref{thm:projectm} , then $I-\mP$ is also a projection matrix to the left nullspace $N(\mA^T)$. 
+	Suppose $\mP$ is a projection matrix in theorem~\ref{thm:projectm}, then $\mI-\mP$ is also a projection matrix to the left nullspace $N(\mA^T)$. 
 \end{cor}
@@
 \begin{proof}
-	For any $x \in \bbR^m$, we have $x - \mP x \perp C(\mA)$
-	\[ x - \mP x \perp C(\mA) \quad \Rightarrow \quad (I-\mP)x \perp C(\mA) \quad \Rightarrow \quad  (I-\mP)x \in N(\mA^T) \text{ and } (x - (I-\mP)x) \perp N(\mA^T). \]
-	Therefore, $I-\mP$ is a projection matrix to the left nullspace $N(\mA^T)$.
+	For any $x \in \bbR^n$, we have
+	\[ x - \mP x \perp C(\mA) \quad \Rightarrow \quad (I-\mP)x \perp C(\mA) \quad \Rightarrow \quad  (\mI-\mP)x \in N(\mA^T) \text{ and } (x - (\mI-\mP)x) \perp N(\mA^T). \]
+	Therefore, $\mI-\mP$ is a projection matrix to the left nullspace $N(\mA^T)$.
 \end{proof}
@@
 \subsection{Projection matrices and least squares}
-Given observation vector $y \in \bbR^n$ and the design matrix $\mX \in \bbR^{n \times (k+1)}$, we propose the linear regression model
+Given an observation vector $y \in \bbR^n$ and a design matrix $\mX \in \bbR^{n \times (k+1)}$, we propose the linear regression model
 \[ y = \mX \beta +\epsilon,\]
-where $\beta = (\beta_0,\beta_1,...,\beta_k)$ are coefficients of our interests and $\epsilon$ is the noise. The least square estimate of $\beta$ minimizes the loss
+where $\beta = (\beta_0,\beta_1,...,\beta_k)$ are regression coefficients and $\epsilon$ is the noise. The least square estimate of $\beta$ is the minimizer of the loss, i.e.\
 \[  \hat \beta_{LS} = \argmin_{\beta \in \bbR^{k+1}} \onorm{y - \mX \beta}^2, \]
 where $\onorm{\cdot}$ is the euclidean norm. The vector $\mX 
-\hat \beta_{LS}$ can be considered as a the projection of $y$ onto the column space of $\mX$. Therefore, we may use projection tools to solve the least square estimate.
+\hat \beta_{LS}$ can be considered as a the projection of $y$ onto the column space of $\mX$, which minimizes the distance from $y$ to the column space $C(\mX)$.
+Therefore, we may use projection tools to solve the least square estimate.
 
@@
 
-The estimate $\hat \beta_{LS}$ is identical to the estimates solved by other methods  using the derivative.
+The estimate $\hat \beta_{LS}$ consists with the estimates solved by using the derivative.
 
@@
 \begin{defn}[\textit{Orthonormal matrix and orthogonal matrix}]\label{def:orthonom}
-	Consider a matrix $\mQ \in \bbR^{m \times n}$. If the columns of $\mQ$ are orthonormal, the matrix $\mQ$ is an orthonormal matrix. If $ m= n$, the square matrix $\mQ$ is a orthogonal matrix.
+	Consider a matrix $\mQ \in \bbR^{m \times n}$. If the columns of $\mQ$ are orthonormal, the matrix $\mQ$ is an orthonormal matrix. If $ m= n$, the square matrix $\mQ$ is an orthogonal matrix.
 \end{defn}
@@
  \end{equation}
-	where $\mQ \in \bbR^{m \times n}$ is an orthonormal matrix and $\mR \in \bbR^{n \times n}$ is a upper triangular matrix. We call the matrix decomposition as equation~\eqref{eq:qr} QR decomposition.
+	where $\mQ \in \bbR^{m \times n}$ is an orthonormal matrix and $\mR \in \bbR^{n \times n}$ is an upper triangular matrix. We call the matrix decomposition in equation~\eqref{eq:qr} QR decomposition.
 \end{defn}
@@
 \subsection{Properties of determinants}
-We give ten properties of determinants. The last seven properties are deduced by first three basic properties
+We give ten properties of determinants. The last seven properties are deduced by first three basic properties.
 \begin{enumerate}
@@
 Hence, an odd number of row exchanges reverse the sign of the determinant while an even number of row exchanges maintain the sign of the determinant.
-	\item \begin{enumerate} 
-	\item[(a)]\label{prop3a} If one row of the matrix is multiplied by a constant $t$, the determinant of the matrix is multiplied by $t$.
+	\item\label{prop3} \begin{enumerate} 
+	\item[(a)] If one row of the matrix is multiplied by a constant $t$, the determinant of the matrix is multiplied by $t$.
 	\begin{align}
@@
 	\end{align} 
-	\item[(b)]\label{prop3b} The determinant is linearly additive on the rows of the matrix.
+	\item[(b)] The determinant is linearly additive on the rows of the matrix.
 	\begin{align}
@@
 		Suppose $\mA$ has two identical rows $a_i, a_j$.
-Exchanging $a_i, a_j$, the exchanged matrix $\mA' = \mA$.
+The matrix after exchanging $a_i, a_j$, $\mA'$, is the same as the original matrix $\mA$.
 By property~\ref{prop2}, $\det(\mA) = -\det(\mA') = -\det(\mA)$.
@@
 	\begin{proof}
-		By property~\ref{prop3a}, property~\ref{prop3b}, and property~\ref{propidenticalrow}, 
+		By property~\ref{prop3} and property~\ref{propidenticalrow}, 
 	\begin{align}
@@
 	\begin{proof}
-		By property~\ref{prop3a}, letting $t = 0$ leads to property~\ref{prop0row}.
+		By property~\ref{prop3}, letting $t = 0$ leads to property~\ref{prop0row}.
 	\end{proof}
 
-	\item\label{propdiag} Suppose $\mA \in \bbR^{n \times n}$ is a triangular matrix with diagonal elements $d_1,...,d_n$. The determinant of $\mA$ is equal to the product $d_1 ...d_n$.
-
+	\item\label{propdiag} Suppose $\mA \in \bbR^{n \times n}$ is a triangular matrix with diagonal elements $d_1,...,d_n$. Then, the determinant $\det(\mA) = \prod_{i = 1}^n d_i$. 
 	\begin{proof}
 		By property~\ref{propelim}, eliminating the triangular matrix $\mA$ to a diagonal matrix does not change the determinant of $\mA$.
-By property~\ref{prop3a} and \ref{prop1}, $\det(\mA) = d_1 ...d_n \det(\mI_n) = d_1 ...d_n $.
+By property~\ref{prop3} and \ref{prop1}, $\det(\mA) = d_1 ...d_n \det(\mI_n) = d_1 ...d_n $.
 	\end{proof}
@@
 	\begin{proof}
-		If $\mA$ is a singular matrix, the reduced row echelon form of $\mA$, RREF($\mA$), has a row with all 0 elements. By property~\ref{propelim} and \ref{prop0row}, the determinant of $\mA$ is 0.
+		If $\mA$ is a singular matrix, the reduced row echelon form of $\mA$, RREF($\mA$), has a row with all 0 entries. By property~\ref{propelim} and \ref{prop0row}, the determinant of $\mA$ is 0.
 	\end{proof}
@@
 Then, $\det(\mA \mB) = \det(\mA) \det(\mB) = 0$.
-If both $\mA, \mB$ are invertible, $\mA, \mB$ can be eliminated to diagonal matrices $\mQ_{A}, \mQ_{B}$ by QR decomposition.
-Suppose $a_1,...,a_n$ and $b_1,...,b_n$ are diagonal elements for $\mQ_A$ and $\mQ_B$ respectively.
+If both $\mA, \mB$ are invertible, $\mA, \mB$ can be eliminated to diagonal matrices $\mD_{A}, \mD_{B}$ by elimination.
+Suppose $a_1,...,a_n$ and $b_1,...,b_n$ are diagonal elements for $\mD_A$ and $\mD_B$ respectively.
 By property~\ref{propdiag}, $\det(AB) = \prod_{i=1}^n a_i b_i = \det(\mA) \det(\mB)$.
@@
 \begin{prop}[Big formula for computing determinant]
-	Consider a square matrix $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$. The big formula for computing the determinant of $\mA$ states below.
+	Consider a square matrix $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$. The big formula for computing the determinant of $\mA$ is
 	\[ \det(\mA) =  \sum_{(\alpha_1,...,\alpha_n) \in \tP} (-1)^{N(\alpha_1,...,\alpha_n)} a_{1,\alpha_1} ... a_{n \alpha_n},  \]
-	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$ and $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of element exchanges from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
+	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$, $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of necessary exchanges from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
 \end{prop}
@@
 	For a square matrix $\mA = \entry{a_ij} \in \bbR^{n \times n}$, we decompose the  matrix $\mA$ by the summation of $n^n$ matrices, in which there is only one entry comes from $\mA$ in each row and other entries are 0.
-By property~\ref{prop3b}, the determinant of $\mA$ is equal to the summation of the determinants of these $n^n$ matrices.
+By property~\ref{prop3}, the determinant of $\mA$ is equal to the summation of the determinants of these $n^n$ matrices.
 Take a 3-by-3 square matrix as an example.
@@
 	By property~\ref{propdiag} and \ref{prop2}, the determinant for the non-singular matrix follows the formula \[(-1)^{N(\alpha_1,...,\alpha_n)}  a_{1,\alpha_1} ... a_{n \alpha_n},\]
-	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$ and $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of element exchanges from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
+	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$, $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of necessary exchanges from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
 \end{proof}
@@
 \begin{example}[Tridiagonal matrix]
-	One example usage of cofactor formula is computing the determinant of \textit{tridiagonal matrix}.
+	One example of cofactor formula is computing the determinant of \textit{tridiagonal matrix}.
 A tridiagonal matrix is a matrix in which only non-zero elements lie on or adjacent to the diagonal.
@@
 \subsection{Inverse matrices}
-Previously, we use Gauss-Jordan elimination to obtain the inverse matrix of the invertible matrix. Here, we apply cofactor formula for computing the inverse matrix.
+Previously, we use Gauss-Jordan elimination to obtain the inverse matrix of an invertible matrix. Here, we apply cofactor formula to compute the inverse matrix.
 
@@
 
-	Therefore, the diagonal element of $\mA \mC^T$ satisfies $(\mA \mC^T)_{ii} = \sum_k^n a_{ik} C_{ik} = \det(\mA)$.
+	Therefore, the diagonal elements of $\mA \mC^T$ are $(\mA \mC^T)_{ii} = \sum_k^n a_{ik} C_{ik} = \det(\mA), \forall i \in [n]$.
 Consider the off-diagonal entries $(\mA \mC^T)_{ij} = \sum_{k}^n a_{ik} C_{jk}$, where $i \neq j$.
@@
 The absolute determinant of a square matrix $\mA$ is the volume of parallelepiped determined by the rows of $\mA$, i.e.\
-	\[ |\det(\mA)| = vol(P), \]
-	where $P$ is the parallelepiped determined by the rows of $\mA$.
+	\[ |\det(\mA)| = vol(P(\mA)), \]
+	where $P(\mA)$ is the parallelepiped determined by the rows of $\mA$.
 \end{prop}
@@
 \begin{example}
-	The area of triangle with vertices at $(x_1,y_1), (x_2,y_2), (x_3,y_2)$ is
+	The area of a triangle with vertices at $(x_1,y_1), (x_2,y_2), (x_3,y_2)$ is
 	\begin{align}
@@
 		The two edges of the triangle are $v_1 = (x_2 - x_1, y_2 -y_1)$ and $v_2 = (x_3 - x_1, y_3 -y_1)$.
-The area of the triangle is half of the area of parallelepiped determined by $v_1, v_2$.
-Therefore, the area of triangle is
+The area of the triangle is a half of the area of the parallelepiped determined by $v_1, v_2$.
+Therefore, the area of the triangle is
 		\begin{align}
@@
 	Consider a square matrix $\mA$, an eigenvector of $\mA$ is a non-zero vector such that:
-	\[ \mA x = \lambda x, \quad \text{for some } \lambda \in \bbR. \]
-	The value $\lambda$ is called the eigenvalue of $\mA$, and $x$ is the eigenvector associated with eigenvalue $\mA$.
+	\[ \mA x = \lambda x, \quad \text{for some } \lambda \in \mathbb{C}. \]
+	The value $\lambda$ is called the eigenvalue of $\mA$, and $x$ is the eigenvector associated with eigenvalue $\lambda$.
 Usually, eigenvectors are normalized, i.e.\ $\onorm{x} = 1$, where $\onorm{\cdot}$ is the euclidean distance.
-
-	Particularly, the vectors associated with eigenvalue 0 consists of the nullspace of $\mA$.
+Particularly, the eigenvectors associated with eigenvalue 0 span the nullspace of $\mA$.
 \end{defn}
@@
 \begin{defn}[\textit{Characteristic polynomial}]\label{def:charac}
-	Let $\lambda$ denote the eigenvalue of the matrix $\mA\in \bbR^{n \times n}$. Then the matrix $\mA - \lambda \mI$ is a singular matrix. Therefore,
+	Let $\lambda$ denote an eigenvalue of the matrix $\mA\in \bbR^{n \times n}$. Then the matrix $\mA - \lambda \mI$ is a singular matrix. Therefore,
 	\[ \det(\mA - \lambda I_n) = 0 \quad  \Leftrightarrow \quad (-1)^n \lambda^n + tr(\mA)\lambda^{n-1} + \cdots + (-1)^n \det(\mA) \overset{\Delta}{=} P(\lambda) = 0.   \]
@@
 	\begin{equation}
-		\sum_{i}^n \lambda_i = tr(\mA); \quad \prod_{i=1}^n \lambda_i = \det(\mA).
+		\sum_{i=1}^n \lambda_i = tr(\mA); \quad \prod_{i=1}^n \lambda_i = \det(\mA).
 	\end{equation} 
@@
 \begin{proof}
-	First, we re-write characteristic polynomial $P(\lambda)$ of $\mA$ as
+	First, we re-write the characteristic polynomial $P(\lambda)$ of $\mA$ as
 	\begin{equation}\label{eq:poly}
@@
 Similarly, the constant term in equation~\eqref{eq:poly} is $(-1)^n \prod_{i=1}^n \lambda_1 \cdots \lambda_n $.
-Therefore, $\prod_{i=1}^n = \det(\mA)$.
+Therefore, $\prod_{i=1}^n \lambda_i = \det(\mA)$.
 \end{proof}
@@
 \begin{proof}
-	Suppose the square matrix has $n$ different eigenvalues $\lambda_1,...,\lambda_n$ with associated eigenvectors  $v_1,...,v_n$.
-Let the linear combination $c_1 v_1 + ...
-+ c_n v_n = 0$.
-We need to show $c_1 =\cdots = c_n = 0$.
-
+	Suppose the square matrix $\mA$ has $n$ different eigenvalues $\lambda_1,...,\lambda_n$ with associated eigenvectors  $v_1,...,v_n$. Let the linear combination $c_1 v_1 + ... + c_n v_n = 0$.	
 	Then we have
@@
 	 Since $\lambda_1 ,..., \lambda_n$ are distinct, and $v_n$ is a non-zero vector, we have $c_n = 0$.
-Similarly, plugging $c_n = 0$ back to previous steps, we have $c_1 = \cdots = c_n = 0$.
+Similarly, plugging $c_n = 0$ back to previous steps leads to that $c_1 = \cdots = c_n = 0$.
 Therefore, $v_1,...,v_n$ are independent.
@@
 	Symmetric square matrices  with all real entries have real eigenvalues.
-For antisymmetric matrices,  i.e.\ $\mA^T = - \mA$, all eigenvalues are imaginary $\lambda = b i$, where $b \in \bbR, i = \sqrt{-1}$.
+For antisymmetric matrices which satisfy $\mA^T = - \mA$, all eigenvalues are imaginary, i.e.\ $\lambda = b i$, where $b \in \bbR, i = \sqrt{-1}$.
 \end{thm}
 \begin{proof}
-	First, we prove that symmetric square matrices have real eigenvalues by contradiction.
+	First, we prove that symmetric square matrices with all real entries have real eigenvalues by contradiction.
 
@@
 
-	Next, we prove that the eigenvalues for antisymmetric matrices are imaginary by the similar procedures above. 
-	Suppose $\mA$ is an antisymmetric matrix with complex eigenvalue $\lambda$ and complex eigenvector $x$, we have
+	Next, we prove that the eigenvalues for antisymmetric matrices are imaginary. 
+	Suppose $\mA$ is an antisymmetric matrix with complex eigenvalue $\lambda$ and complex eigenvector $x$.  By similar procedures for symmetric matrices, we have
 	\[ \overline x^T \mA x  = x^T \mA^T \overline x  =  -x^T \mA \overline x .  \] 
-	Therefore, we have
+	Then,
 	\[  0 = ( \overline x^T \mA x +  x^T \mA \overline x ) = (\lambda + \overline \lambda) x^T \overline x.   \]
@@
 \begin{proof}
-Suppose $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ is a square matrix. By the property of determinant~\ref{propdiag}, 
+Suppose $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ is a triangular matrix. By the property of determinant~\ref{propdiag}, 
 \[ \det(\mA) = \prod_{i=1}^n a_{ii}  ; \quad \det(\mA - \lambda \mI_n) = \prod_{i=1}^n (a_{ii}-\lambda). \]
@@
 Suppose $\mA \in \bbR^{n \times n}$ is an invertible matrix.
-Let $\lambda_1,...,\lambda_n$ denote the non-zero eigenvalue of $\mA$, and $x_1,...x_n$eigenvectors corresponds to $\lambda_i$.
-Then  $\lambda_1^{-1},...,\lambda_m^{-1}, \lambda_{m+1},...,\lambda_n $ and $x_1,...x_n$ are eigenvalues and associated eigenvectors of $\mA^{-1}$.
+Let $\lambda_1,...,\lambda_n$ denote the eigenvalue of $\mA$, and $x_1,...x_n$ denote the associated eigenvectors.
+Then,  $\lambda_1^{-1},...,\lambda_n^{-1}$ and $x_1,...x_n$ are the eigenvalues and associated eigenvectors of $\mA^{-1}$, respectively.
 \end{thm}
@@
 \begin{proof}
-	Since the nullspace of $\mA$ contains only a zero vector, $\lambda_i >0 \forall i \in [n]$.
+	Since the nullspace of $\mA$ contains only a zero vector, the eigenvaluest $\lambda_i >0,  \forall i \in [n]$.
 For any $\lambda_i$ and $x_i$, multiplying $x_i$ on both sides of the equation $\mA^{-1} \mA = \mI_n$, we have
-	\[ \mA^{-1} \mA x_i = x_i \quad \Rightarrow\quad  \mA^{-1} x_i = {\lambda}^{-1}.  \]
-	Therefore, $\mA^{-1}$ share the same eigenvectors with $\mA$, and the reciprocal of the eigenvalues of $\mA$ are the eigenvalues of $\mA^{-1}$.
+	\[ \mA^{-1} \mA x_i = x_i \quad \Rightarrow\quad  \mA^{-1} x_i = {\lambda_i}^{-1}.  \]
+	Therefore, $\mA^{-1}$ share the same eigenvectors set with $\mA$, and the reciprocal of the eigenvalues of $\mA$ are the eigenvalues of $\mA^{-1}$.
 \end{proof}

### 2020-07-11 skills/math/linear_algebra/linear_algebra_summary_p2.tex
linear algebra

@@
 \begin{defn}[\textit{Orthogonal vectors}]\label{def:vortho}
-	Consider two vectors $x,y \in \bbR^{n}$. The vectors $x$ and $y$ are orthogonal,  denoted $x \perp y$, \textbf{iff} $x^T y = y^T x = 0$.
+	Consider two vectors $x,y \in \bbR^{n}$. The vectors $x$ and $y$ are orthogonal,  denoted $x \perp y$, if and only if $x^T y = y^T x = 0$.
 \end{defn}
@@
 \begin{defn}[\textit{Orthogonal subspaces}]\label{def:sortho}
-	Consider two subspaces $S,T$. The subspaces $S$ and $T$ are orthogonal,  denoted $S \perp T$, \textbf{iff} for any $s \in S$ and $ t \in T$, $s^T t = t^T s = 0$.
+	Consider two subspaces $S,T$. The subspaces $S$ and $T$ are orthogonal,  denoted $S \perp T$, if and only if for any $s \in S$ and $ t \in T$, $s^T t = t^T s = 0$.
 \end{defn}
 
-Given a matrix $\mA \in \bbR^{m \times n}$, there are four subspaces related to $\mA$: column space $C(\mA)$, row space $C(\mA^T)$, nullspace $N(\mA)$, and left nullspace $N(\mA^T)$.
+Given a matrix $\mA \in \bbR^{m \times n}$, there are four subspaces related to $\mA$\colon column space $C(\mA)$, row space $C(\mA^T)$, nullspace $N(\mA)$, and left nullspace $N(\mA^T)$.
 Suppose the rank of $\mA$ is $rank(\mA) = r$, the dimensions of these subspaces are:
@@
 	By the first equality in \eqref{eq:aaa}, $b \in C(\mA)$, and by the third equation in \eqref{eq:aaa}, $ b \in N(\mA^T)$.
-This contradicts the theorem~\ref{thm:ortho}, $C(\mA) \perp N(\mA^T)$.
+This contradicts the Theorem~\ref{thm:ortho}, $C(\mA) \perp N(\mA^T)$.
 Therefore, for any $x \in N(\mA^T \mA)$, the vector $x \in N(\mA)$.
@@
 	If $\mA \in \bbR^{m \times n}$ has independent columns, then $rank(\mA) = n$.
-By theorem~\ref{thm:aa}, $rank(\mA^T \mA) = rank(\mA) = n$.
+By Theorem~\ref{thm:aa}, $rank(\mA^T \mA) = rank(\mA) = n$.
 Since $\mA^T \mA \in \bbR^{n \times n}$ is a square matrix, $\mA^T \mA$ is invertible.
@@
 \begin{cor}[Projection onto $N(\mA^T)$]\label{cor:iprojectm}
-	Suppose $\mP$ is a projection matrix in theorem~\ref{thm:projectm}, then $\mI-\mP$ is also a projection matrix to the left nullspace $N(\mA^T)$. 
+	Suppose $\mP$ is a projection matrix in Theorem~\ref{thm:projectm}, then $\mI-\mP$ is also a projection matrix to the left nullspace $N(\mA^T)$. 
 \end{cor}
@@
 	\end{align}
-	For the square matrix $\mA$,  there are $n!$ non-singular matrices among the $n^n$ matrices.
-The number of non-singular matrices coincides with the size of $\tP$.
+	For the square matrix $\mA$,  there are $n!$ nonsingular matrices among the $n^n$ matrices.
+The number of nonsingular matrices coincides with the size of $\tP$.
 Intuitively, there are $n$ ways to choose an element from the first row, after which there are only $n-1$ ways to choose an element from the second row to avoid the zero determinant.
-Therefore, we have $n \times (n-1) \times (n-2) \times \cdots \times 2 = n!$ non-singular matrices.
+Therefore, we have $n \times (n-1) \times (n-2) \times \cdots \times 2 = n!$ nonsingular matrices.
 
-	By property~\ref{propdiag} and \ref{prop2}, the determinant for the non-singular matrix follows the formula \[(-1)^{N(\alpha_1,...,\alpha_n)}  a_{1,\alpha_1} ... a_{n \alpha_n},\]
+	By property~\ref{propdiag} and \ref{prop2}, the determinant for the nonsingular matrix follows the formula \[(-1)^{N(\alpha_1,...,\alpha_n)}  a_{1,\alpha_1} ... a_{n \alpha_n},\]
 	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$, $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of necessary exchanges from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
@@
 	One example of cofactor formula is computing the determinant of \textit{tridiagonal matrix}.
-A tridiagonal matrix is a matrix in which only non-zero elements lie on or adjacent to the diagonal.
+A tridiagonal matrix is a matrix in which only nonzero elements lie on or adjacent to the diagonal.
 Let $\mT_n \in \bbR^{n \times n}$ denote the tridiagonal matrix of 1's, i.e.\
@@
 Consider the linear system $\mA x=b$, where $\mA\in\bbR^{n \times n}$ is an invertible matrix, and $ x, b \in \bbR^n$ are vectors.
-Applying theorem~\ref{thm:inver}, we obtain the Cramer's rule of $x = \mA^{-1} b$.
+Applying Theorem~\ref{thm:inver}, we obtain the Cramer's rule of $x = \mA^{-1} b$.
 \[  x = \frac{1}{\det(\mA)} \mC^T b; \quad x_j = \frac{\det(\mB_j)}{\mA_j}, \] 
@@
 	The parallelepiped determined by $n$ vectors $v_1,...,v_n \in \bbR^n$ is the subset 
-	\[ P = \{ a_1 v_1 + \cdots + a_n v_n : 0 \leq a_1,...,a_n \leq 1 \}. \]
+	\[ P = \{ a_1 v_1 + \cdots + a_n v_n \colon 0 \leq a_1,...,a_n \leq 1 \}. \]
 	 We use $vol(P)$ to denote the volume of the parallelepiped  $P$. 
@@
 \begin{defn}[\textit{Eigenvalues and eigenvectors}]\label{def:eigen}
-	Consider a square matrix $\mA$, an eigenvector of $\mA$ is a non-zero vector such that:
+	Consider a square matrix $\mA$, an eigenvector of $\mA$ is a nonzero vector such that:
 	\[ \mA x = \lambda x, \quad \text{for some } \lambda \in \mathbb{C}. \]
@@
 Usually, eigenvectors are normalized, i.e.\ $\onorm{x} = 1$, where $\onorm{\cdot}$ is the euclidean distance.
-Particularly, the eigenvectors associated with eigenvalue 0 span the nullspace of $\mA$.
+For simplicity, all the eigenvectors mentioned below are normalized.
+Besides, the eigenvectors associated with eigenvalue 0 span the nullspace of $\mA$.
 \end{defn}
@@
 
+\subsection{Properties for eigenvectors and eigenvalues}
+
 \begin{thm}[Summation and production of eigenvalues]\label{thm:speigen}
-	For an square matrix $\mA \in \bbR^{n \times n}$, let $\lambda_1,...\lambda_n$ denote the $n$ eigenvalues of $\mA$. Then 
+	For a square matrix $\mA \in \bbR^{n \times n}$, let $\lambda_1,...\lambda_n$ denote the $n$ eigenvalues of $\mA$. Then 
 	\begin{equation}
@@
 	where $\lambda_i, i \in [n]$ are $n$ eigenvalues of $\mA$.
-By the equation~\eqref{eq:poly}, the coefficient for $\lambda^{n-1}$ is equal to $(-1)^n\sum_{i=1}^n \lambda_i$.
+By equation~\eqref{eq:poly}, the coefficient for $\lambda^{n-1}$ is equal to $(-1)^n\sum_{i=1}^n \lambda_i$.
 Compared with definition~\ref{def:charac}, we have $tr(\mA) = \sum_{i=1}^n \lambda_i$.
@@
 
-\begin{thm}[Independence of eigenvectors]\label{thm:indepeigen} 
-Consider a square matrix $\mA$ with at least two different eigenvalues. The eigenvectors associated with different eigenvalues are independent.
+\begin{defn}[\textit{Similar matrices}]\label{def:similar}
+	Consider a square matrix $\mA \in \bbR^{n \times n}$. For any invertible matrix $\mB \in \bbR^{n \times n}$, the matrix product $\mB \mA \mB^{-1}$ is called the similar matrix of $\mA$.
+\end{defn}
+
+\begin{thm}[Eigenvalues for similar matrices]\label{thm:similar}
+	Consider a sqaure matrix $\mA$. The similar matrices of $\mA$ share the same eigenvalues of $\mA$.
+\end{thm}
+\begin{proof}
+	Let $\mB \mA \mB^{-1}$ be the similar matrix of $\mA$, and $\lambda_1,...,\lambda_n$ be the eigenvalues of $\mA$. Then, for any $\lambda_i$, there is an eigenvector, such that, 
+	\[ \lambda x = \mA x \quad \Leftrightarrow \quad \lambda \mB x = \mB \mA \mB^{-1} \mB x.\]
+	Since $x$ is a nonzero vector, and $\mB$ is invertible, the vector $\mB x$ is also a nonzero vector. Therefore, $\lambda$ is an eigenvalue of $\mB \mA \mB^{-1}$ with associated eigenvector $\mB x$.
+\end{proof}
+
+\begin{thm}[Eigenvalues for powers of the matrix]
+	Consider a matrix $\mA$ with a eigenvalue $\lambda$ and the associated eigenvector $x$. For any polynomial $P$, 
+	\[ P(\mA) x = P(\lambda) x. \]
+\end{thm}
+\begin{proof}
+	For any integer $k \geq 0$ and constant $c \in \bbR$,
+	\[ c\mA^k x = c\mA^{k-1} \mA x= c\lambda \mA^{k-1} x = \cdots = c\lambda^k x.\]
+	Therefore, for any polynomial $P$, $P(\mA)x = P(\lambda)x$.
+\end{proof}
+
+\begin{thm}[Eigenvalues for the inverse]
+Suppose $\mA \in \bbR^{n \times n}$ is an invertible matrix.
+Let $\lambda_1,...,\lambda_n$ denote the eigenvalue of $\mA$, and $x_1,...x_n$ denote the associated eigenvectors.
+Then,  $\lambda_1^{-1},...,\lambda_n^{-1}$ and $x_1,...x_n$ are the eigenvalues and associated eigenvectors of $\mA^{-1}$, respectively.
 \end{thm}
@@
 \begin{proof}
-	Suppose the square matrix $\mA$ has $n$ different eigenvalues $\lambda_1,...,\lambda_n$ with associated eigenvectors  $v_1,...,v_n$. Let the linear combination $c_1 v_1 + ... + c_n v_n = 0$.	
+	Since the nullspace of $\mA$ contains only a zero vector, the eigenvaluest $\lambda_i >0,  \forall i \in [n]$.
+For any $\lambda_i$ and $x_i$, multiplying $x_i$ on both sides of the equation $\mA^{-1} \mA = \mI_n$, we have
+	\[ \mA^{-1} \mA x_i = x_i \quad \Rightarrow\quad  \mA^{-1} x_i = {\lambda_i}^{-1} x_i.  \]
+	Therefore, $\mA^{-1}$ share the same eigenvectors set with $\mA$, and the reciprocal of the eigenvalues of $\mA$ are the eigenvalues of $\mA^{-1}$.
+\end{proof}
+
+\begin{thm}[Independence of eigenvectors]\label{thm:indepeigen} 
+Suppose the square matrix $\mA$ has at least two different eigenvalues. The eigenvectors associated with different eigenvalues are independent.
+\end{thm}
+
+\begin{proof}
+	Suppose the square matrix $\mA \in \bbR^{n \times n}$ has $m$ different eigenvalues $\lambda_1,...,\lambda_m$ with associated eigenvectors  $v_1,...,v_m$.
+Let the linear combination $c_1 v_1 + ...
++ c_m v_m = 0$.
 	Then we have
 	\begin{equation}\label{eq:indep1}
-		\mA(c_1 v_1 + ... + c_n v_n ) = c_1 \lambda_1 v_1 + ... + c_n \lambda_n v_n = 0;  
+		\mA(c_1 v_1 + ... + c_m v_m ) = c_1 \lambda_1 v_1 + ... + c_m \lambda_m v_m = 0;  
 		\end{equation}
 		\begin{equation}\label{eq:indep2}
-		\lambda_1 (c_1 v_1 + ... + c_n v_n ) = c_1 \lambda_1 v_1 + ... + c_n \lambda_1 v_n = 0. 
+		\lambda_1 (c_1 v_1 + ... + c_m v_m ) = c_1 \lambda_1 v_1 + ... + c_m \lambda_1 v_m = 0. 
 	\end{equation}
@@
 	\begin{equation}\label{eq:indep3}
-		 c_2 (\lambda_1 - \lambda_2) v_2 + \cdots + c_n (\lambda_1 - \lambda_n) v_n  = 0
+		 c_2 (\lambda_1 - \lambda_2) v_2 + \cdots + c_m (\lambda_1 - \lambda_m) v_m  = 0
 	\end{equation}
 	Repeating the steps from equation~\eqref{eq:indep1} to equation~\eqref{eq:indep3}, finally we have 
-	\[ c_n(\lambda_1 - \lambda_n)(\lambda_2 - \lambda_n)\cdots (\lambda_{n-1} - \lambda_n) v_n = 0.    \]
-	 Since $\lambda_1 ,..., \lambda_n$ are distinct, and $v_n$ is a non-zero vector, we have $c_n = 0$.
-Similarly, plugging $c_n = 0$ back to previous steps leads to that $c_1 = \cdots = c_n = 0$.
-Therefore, $v_1,...,v_n$ are independent.
-\end{proof}
-
-\begin{thm}[Eigenvalues for symmetric and antisymmetric matrices]\label{def:symmeign}
-	Symmetric square matrices  with all real entries have real eigenvalues.
-For antisymmetric matrices which satisfy $\mA^T = - \mA$, all eigenvalues are imaginary, i.e.\ $\lambda = b i$, where $b \in \bbR, i = \sqrt{-1}$.
-\end{thm}
-\begin{proof}
-	First, we prove that symmetric square matrices with all real entries have real eigenvalues by contradiction.
-
-	Suppose a real and symmetric matrix $\mA$ has a complex eigenvalue $\lambda$ with eigenvector $x$.
-Since $\mA$ is real, then the eigenvector $x$ is also complex.
-Let $\overline \lambda$ and $\overline x$ denote the conjugate eigenvalue and eigenvector, respectively.
-Then we have
-	\begin{equation}\label{eq:symm}
-		\mA x = \lambda x;\quad \mA \overline x = \overline \lambda \overline x.
-	\end{equation} 
-	Premultiply $\overline x^T$ on both sides of the first equation in \eqref{eq:symm} and premultiply $x^T$ on both sides of the second equation in \eqref{eq:symm}. By the symmetry of $\mA$,
-	\[ \overline x^T \mA x  = x^T \mA^T \overline x = x^T \mA \overline x .\]
-	Subtract the two equations in \eqref{eq:symm}, we have
-	\[  0 = ( \overline x^T \mA x -  x^T \mA \overline x ) = (\lambda - \overline \lambda) x^T \overline x.  \]
-	By the fact that $x^T \overline x$ is a real number, the coefficient for $i$ in $\lambda$ is equal to 0, which contradicts that $\lambda$ is complex.
-
-	Next, we prove that the eigenvalues for antisymmetric matrices are imaginary. 
-	Suppose $\mA$ is an antisymmetric matrix with complex eigenvalue $\lambda$ and complex eigenvector $x$.  By similar procedures for symmetric matrices, we have
-	\[ \overline x^T \mA x  = x^T \mA^T \overline x  =  -x^T \mA \overline x .  \] 
-	Then,
-	\[  0 = ( \overline x^T \mA x +  x^T \mA \overline x ) = (\lambda + \overline \lambda) x^T \overline x.   \]
-	Since $x^T \overline x$ is a real number, the real part in $\lambda$ is equal to 0. Therefore, eigenvalues for antisymmetric matrices are imaginary.
+	\[ c_m(\lambda_1 - \lambda_m)(\lambda_2 - \lambda_m)\cdots (\lambda_{m-1} - \lambda_m) v_m = 0.    \]
+	 Since $\lambda_1 ,..., \lambda_m$ are distinct, and $v_m$ is a nonzero vector, we have $c_m = 0$.
+Plugging $c_m = 0$ back to previous steps leads to that $c_1 = \cdots = c_m = 0$.
+Therefore, $v_1,...,v_m$ are independent.
 \end{proof}
@@
 
-\begin{thm}[Eigenvalues for powers of the matrix]
-	Consider a matrix $\mA$ with a eigenvalue $\lambda$ and the associated eigenvector $x$. For any polynomial $P$, 
-	\[ P(\mA) x = P(\lambda) x. \]
+\subsection{Eigenvalues and eigenvectors for symmetric matrices}
+
+\begin{thm}[Eigenvalues for symmetric and antisymmetric matrices]\label{thm:symmeign}
+	Symmetric square matrices  with all real entries have real eigenvalues.
+For antisymmetric matrices which satisfy $\mA^T = - \mA$, all eigenvalues are imaginary, i.e.\ $\lambda = b i$, where $b \in \bbR, i = \sqrt{-1}$.
 \end{thm}
 \begin{proof}
-	For any integer $k \geq 0$ and constant $c \in \bbR$,
-	\[ c\mA^k x = c\mA^{k-1} \mA x= c\lambda \mA^{k-1} x = \cdots = c\lambda^k x.\]
-	Therefore, for any polynomial $P(\mA)x = P(\lambda)x$.
+	First, we prove that symmetric square matrices with all real entries have real eigenvalues by contradiction.
+
+	Suppose a symmetric real matrix $\mA$ has a complex eigenvalue $\lambda$ with eigenvector $x$.
+Since $\mA$ is real, then the eigenvector $x$ is also complex.
+Let $\overline \lambda$ and $\overline x$ denote the conjugate eigenvalue and eigenvector, respectively.
+Then we have
+	\begin{equation}\label{eq:symm}
+		\mA x = \lambda x;\quad \mA \overline x = \overline \lambda \overline x.
+	\end{equation} 
+	Premultiply $\overline x^T$ on both sides of the first equation in \eqref{eq:symm} and premultiply $x^T$ on both sides of the second equation in \eqref{eq:symm}. By the symmetry of $\mA$,
+	\[ \overline x^T \mA x  = x^T \mA^T \overline x = x^T \mA \overline x .\]
+	Subtracting the two equations in \eqref{eq:symm}, we have
+	\[  0 = ( \overline x^T \mA x -  x^T \mA \overline x ) = (\lambda - \overline \lambda) x^T \overline x.  \]
+	By the fact that $x^T \overline x$ is a real number, the imaginary part of $\lambda$ is equal to 0, which contradicts that $\lambda$ is complex.
+
+	Next, we prove that the eigenvalues for antisymmetric matrices are imaginary. 
+	Suppose $\mA$ is an antisymmetric matrix with a complex eigenvalue $\lambda$ and complex eigenvector $x$.  By similar procedures for symmetric matrices, we have
+	\[ \overline x^T \mA x  = x^T \mA^T \overline x  =  -x^T \mA \overline x .  \] 
+	Then,
+	\[  0 = ( \overline x^T \mA x +  x^T \mA \overline x ) = (\lambda + \overline \lambda) x^T \overline x.   \]
+	Since $x^T \overline x$ is a real number, the real part of $\lambda$ is equal to 0. Therefore, eigenvalues for antisymmetric matrices are imaginary.
 \end{proof}
 
-\begin{thm}[Eigenvalues for the inverse]
-Suppose $\mA \in \bbR^{n \times n}$ is an invertible matrix.
-Let $\lambda_1,...,\lambda_n$ denote the eigenvalue of $\mA$, and $x_1,...x_n$ denote the associated eigenvectors.
-Then,  $\lambda_1^{-1},...,\lambda_n^{-1}$ and $x_1,...x_n$ are the eigenvalues and associated eigenvectors of $\mA^{-1}$, respectively.
+\begin{thm}[Orthogonality of eigenvectors for symmetric matrices]\label{thm:symmortho}
+The eigenvectors of a symmetric matrix $\mA \in \bbR^{n \times n}$ associated with different eigenvalues are orthogonal.
 \end{thm}
@@
 \begin{proof}
-	Since the nullspace of $\mA$ contains only a zero vector, the eigenvaluest $\lambda_i >0,  \forall i \in [n]$.
-For any $\lambda_i$ and $x_i$, multiplying $x_i$ on both sides of the equation $\mA^{-1} \mA = \mI_n$, we have
-	\[ \mA^{-1} \mA x_i = x_i \quad \Rightarrow\quad  \mA^{-1} x_i = {\lambda_i}^{-1}.  \]
-	Therefore, $\mA^{-1}$ share the same eigenvectors set with $\mA$, and the reciprocal of the eigenvalues of $\mA$ are the eigenvalues of $\mA^{-1}$.
+	Suppose $\lambda_1$ and $\lambda_2$ are two different eigenvalues of $\mA$ with associated eigenvector $x_1, x_2$, respectively. By the symmetry of $\mA$, we have
+	\[ x_2^T \mA x_1 = x_1^T \mA^T x_2  = x_1^T \mA x_2,\]
+	then 
+	\[  0 = ( x_2^T \mA x_1 -  x_1^T \mA x_2 ) = (\lambda_1 - \lambda_2) x_2^T x_1. \]
+	Since $\lambda_1, \lambda_2$ are different, we have $\lambda_1 - \lambda_2 \neq 0$ and $x_2^T x_1 = 0$. Therefore, $x_1$ and $x_2$ are orthogonal. 
+\end{proof}
+
+\begin{thm}[Eigenvectors for repeated eigenvalues of symmetric matrices]\label{thm:symmrepeat} Consider a symmetric matrix $\mA \in \bbR^{n \times n}$.
+If $\lambda_0$ is a repeated eigenvalue of $\mA$ with multiplicity $2 \leq m \leq n$, then there exist $m$ orthonormal eigenvectors corresponding to $\lambda_0$.
+\end{thm}
+
+\begin{proof}
+	First, there is at least one nonzero eigenvector $x_0$ for $\lambda_0$.
+For any $x_0 \in \bbR^{n}$, there are $n-1$ additional orthogonal vectors $y_1,...,y_{n-1}$, such that $y_j \perp x_0, \forall j \in [(n-1)]$ and $(x_0, y_1,...,y_{n-1})$ forms the basis of $\bbR^n$.
+Let $\mY = [y_1,...y_{n-1}]$ and $\mX = [x_0,Y]$.
+Since $\mA$ is a symmetric matrix, $x_0^T \mA = \lambda_0 x_0^T $.
+We have
+	\begin{align}\label{eq:symmrepeat}
+		\mX^T \mA \mX =  \begin{bmatrix}
+			x_0^T \mA x_0 & x_0^T \mA \mY\\
+			\mY^T \mA x_0 & \mY^T \mA \mY 
+		\end{bmatrix} =  \begin{bmatrix}
+			\lambda_0 & 0\\
+			0 & \mY^T \mA \mY 
+		\end{bmatrix} .
+	\end{align}
+	Note that $\mX$ is an orthogonal matrix which satisfies $\mX^T = \mX^{-1}$.
+By Theorem~\ref{thm:similar}, $\mX^T \mA \mX$ is a similar matrix of $\mA$, and $\mX^T \mA \mX$ share the same eigenvalues of $\mA$.
+Then, $\lambda_0$ is also a repeated eigenvalue of $\mX^T \mA \mX$.
+The determinant of $\mX^T \mA \mX - \lambda \mI_n$ is
+	\[  \det(\mX^T \mA \mX - \lambda \mI_n) = (\lambda_0 - \lambda) \det(\mY^T \mA \mY - \lambda \mI_{n-1}). \]
+	Since $\lambda_0$ has multiplicity $m \geq 2$, then $\det(\mY^T \mA \mY - \lambda_0 \mI_{n-1}) = 0$.
+Therefore, the dimension of the nullspace of $\mX^T \mA \mX - \lambda_0 \mI_n$ is larger than 2.
+Let $v_1,v_2$ be the two orthogonal vectors in the nullspace of $\mX^T \mA \mX - \lambda_0 \mI_n$.
+Then $v_1, v_2$ are two eigenvectors of $\mX^T \mA \mX$, and $\mX v_1, \mX v_2$ are two orthogonal eigenvectors of $\mA$ associated with $\lambda_0$.
+
+	Replace $x_0$ by two orthogonal vectors $x_1 = \mX v_1, x_2 = \mX v_2$ and repeat the steps above. Finally, we will have $m$ orthogonal eigenvectors of $\mA$ associated with $\lambda_0$.
+\end{proof}
+
+\subsection{Diagonalization}
+
+\begin{defn}[Diagonalization]
+	Suppose the square matrix $\mA \in \bbR^{n \times n}$ has $n$ independent eigenvectors. Then the matrix $\mA$ can be diagonalized as
+	\[ \mA = \mS^{-1} \Lambda \mS, \]
+	where $\mS \in \bbR{n \times n}$ is a matrix whose columns are $n$ independent eigenvectors of $\mA$ and $\Lambda \in \bbR^{n \times n}$ is a diagonal matrix whose entries are $n$ eigenvalues of $\mA$.
+\end{defn}
+
+\begin{thm}[Diagonalization of symmetric matrices]
+Suppose $\mA$ is a symmetric matrix, then $\mA$ can be diagonalized.
+\end{thm}
+
+\begin{proof}
+	By Theorems~\ref{thm:symmortho} and Theorem~\ref{thm:symmrepeat}, for any symmetric matrix $\mA \in \bbR^{n \times n}$, $\mA$ have $n$ orthogonal eigenvectors, even though there may have repeated eigenvalues.
+\end{proof}
+
+\begin{thm}[Descent of matrix powers]
+	If a square matrix $\mA \in \bbR^{n \times n}$ has $n$ independent eigenvectors with eigenvalues $\lambda_1,...,\lambda_n$, then $\mA^k \rightarrow 0$ as $k \rightarrow +\infty$ if and only if all $|\lambda_i| < 1$.
+\end{thm}
+
+\begin{proof}
+	We diagonalize the matrix $\mA$ as $\mA =\mS^{-1} \Lambda \mS$. For any positive integer $k$, $\mA^k = \mS^{-1} \Lambda^{k} \mS$. Then 
+	\[ \lim_{k \rightarrow +\infty} \mA^k = \lim_{k \rightarrow +\infty} \mS^{-1} \Lambda^{k} \mS  =  \lim_{k \rightarrow +\infty} \mS^{-1} \begin{bmatrix}
+		|\lambda_i|^k & \cdots &0\\
+		\vdots & &\vdots\\
+		0 & \cdots & |\lambda_n|^k
+	\end{bmatrix} \mS  = 0 \quad \Leftrightarrow \quad  |\lambda_i| < 1, \forall i \in [n]. \]
 \end{proof}

### 2020-07-13 skills/math/tail_bound/tailbound_solution.tex
solution

--- 2020-07-07 skills/math/tail_bound/tailbound_solution.tex
@@
 	Suppose $X \geq 0$ is a random variable  with finite mean, we have
-	\begin{equation}
+	\begin{equation}\label{eq:thmmarkov}
 		\bbP(X\geq t) \leq \frac{E[X]}{t},\quad \forall \ t > 0.
@@
 	Suppose $X \geq 0$ is a random variable  with finite mean $\mu$ and finite variance, we have
-	\begin{equation}
+	\begin{equation}\label{eq:cheby}
 		\bbP(|X - \mu |\geq t) \leq \frac{var(X)}{t^2},\quad \forall \ t > 0.
@@
 	Suppose the random variable $X$ has a central moment of order $k$. Applying Markov's inequality to the random variable $|X - \mu|^k$ yields
-	\begin{equation}
+	\begin{equation}\label{eq:poly}
 		\bbP(|X - \mu |\geq t) \leq \frac{\bbE[|X-\mu|^k]}{t^k},\quad \forall \ t > 0.
@@
 \begin{thm}[Chernoff bound]\label{thm:chernoff}
-	Suppose the random variable $X$ has a moment generating function in the neighborhood of 0, i.e.\ $\varphi_X(\lambda) = \bbE[e^{\lambda X}] < +\infty, \forall \lambda \in (-b,b), b>0$.
+	Suppose the random variable $X$ has a moment generating function in the neighborhood of 0; i.e.\, $\varphi_X(\lambda) = \bbE[e^{\lambda X}] < +\infty,$ for all $\lambda \in (-b,b)$ with some $b>0$.
 Applying Markov's inequality to the random variable $Y = e^{\lambda(X- \mu)}$ yields
-	\begin{equation}
-		\bbP((X - \mu) \geq t) \leq \frac{\bbE[e^{\lambda X}]}{e^{\lambda t}}.
+	\begin{equation}\label{eq:chernoff1}
+		\bbP((X - \mu) \geq t) \leq \frac{\bbE[e^{\lambda (X-\mu)}]}{e^{\lambda t}}, \quad \forall \lambda\in (-b,b).
 	\end{equation}
 	Optimizing the choice of $\lambda$ for the tightest bound yields the Chernoff bound
-	\begin{equation}
-		\bbP((X - \mu) \geq t) \leq \inf_{\lambda \in [0,b)} \frac{\bbE[e^{\lambda X}]}{e^{\lambda t}}.
+	\begin{equation}\label{eq:chernoff2}
+		\bbP((X - \mu) \geq t) \leq \inf_{\lambda \in [0,b)} \frac{\bbE[e^{\lambda (X-\mu)}]}{e^{\lambda t}}.
 	\end{equation}
@@
 
+\begin{thm}[Hoeffding bound for bounded variable]\label{thm:boundvar}
+Consider a random variable $X$ with mean $\mu = \bbE(X)$. Assume that $X$ is bounded and $X \in [a,b]$ almost surely, where $a, b$ are two constants. Then, for any $\lambda \in \bbR$, we have
+\begin{equation}
+	\bbE[e^{\lambda X}] \leq e^{\frac{s(b-a)^2}{8}}.
+\end{equation}
+Particularly, the variable $X \sim subG(\frac{(b-a)^2}{4})$.
+\end{thm}
+
+\begin{proof}
+	See Exercise~\ref{pro:boundvar}.
+\end{proof}
+
 \section{Exercises}
 
-\paragraph{Exercise 2.1} (Tightness of inequalities.) The Markov and Chebyshev's inequalities can not be improved in general. 
+\subsection{Exercise 2.1} (Tightness of inequalities.) The Markov's and Chebyshev's inequalities can not be improved in general. 
 \begin{enumerate}
-	\item[(a)] Provide a random variable $X \geq 0$ for which Markov's inequality~\eqref{thm:markov} is met with equality.  
-	\item[(b)] Provide a random variable $Y$ for which Chebyshev's inequality~\eqref{thm:cheb} is met with equality.  
+	\item[(a)] Provide a random variable $X \geq 0$ that attains the equality in Markov's inequality~\eqref{eq:thmmarkov}.  
+	\item[(b)] Provide a random variable $Y$ that attains the equality in Chebyshev's inequality~\eqref{eq:cheby}.  
 \end{enumerate}
@@
 	\item[(a)] Recall the proof of Markov's inequality. For any $t >0$,
-	\[ \bbE[X] = \int_{0}^t x f_X(x) dx + \int_{t}^{+\infty} x f_X(x) dx \geq \int_{t}^{+\infty} x f_X(x)\geq t \int_{t}^{+\infty} f_X(x) = t \bbP( X \geq t).\]
-	 If Markov's inequality meets the equality, the inequalities above should meet equality.
+	\begin{equation}\label{eq:markov}
+		\bbE[X] = \int_{0}^t x f_X(x) dx + \int_{t}^{+\infty} x f_X(x) dx \geq \int_{t}^{+\infty} x f_X(x)\geq t \int_{t}^{+\infty} f_X(x) = t \bbP( X \geq t).
+	\end{equation}
+	Below, given a constant $t > 0$, we construct a random variable that attains the equality in each line of equation~\eqref{eq:markov}.
 
-	 Consider a variable $X$ with distribution $P(X = 0) = 1$. For any $t > 0$, the variable $X$ satisfies 
+	 Consider a variable $X$ with distribution $P(X = t) = 1$. The variable $X$ satisfies 
 	 \[ \int_{0}^t x f_X(x) dx = 0 \text{ and } \int_{t}^{+\infty} x f_X(x) dx =\int_{t}^{+\infty} t f_X(x) dx. \]
-	 Therefore, for variable $X$, the Markov's inequality is met with equality.
-	 \item[(b)] Chebyshev's inequality follows by applying Markov's inequality to the non-negative random variable $Y = \bbE{(X - \bbE[X])^2}$.
-Let the distribution of $Y$ be $\bbP(Y = 0) = 1$.
+	 Therefore, for given $t$, the variable $X$ attains the equality of Markov's inequality.
+
+	 \item[(b)] Chebyshev's inequality follows by applying Markov's inequality to the non-negative random variable $Y = {(X - \bbE[X])^2}$.
+Let the distribution of $Y$ be $\bbP(Y = t^2) = 1$.
 Then the Markov's inequality for $Y$ and the Chebyshev's inequality for $X$ meet the equalities.
-By transformation, the distribution of random variable $X$ is $\bbP(X = \bbE[X]) = 1$.
-Therefore, for any random variable $X$ with distribution $\bbP(X = c) = 1, c \in \bbR$, the Chebyshev's inequality is met with equality.
+Consider the variable $Z$ with distribution $\bbP(Z = c+t) = \bbP(Z = c-t)  = 1/2 $ for any $ c \in \bbR$.
+For $Z$, the transformation probability satisfies $\bbP((Z - \bbE[Z])^2= t^2) = 1$.
+Therefore, the variable $Z$ attains the equality of Chebyshev's inequality.
 \end{enumerate}
 
-\paragraph{Exercise 2.2}
+\begin{rmk}[Tightness of Markov's inequality]
+	Only few variables can attain the equalities in Markov's and Chebyshev's inequalities. In research, we should be careful if there is a concentration bound that is tighter than Markov's inequality.
+\end{rmk}
+
+\subsection{Exercise 2.2}
 \begin{lem}[Standard normal distribution]\label{lem:n01}
-	Let $\phi(z)$ be the density function of a standard normal $Z \sim N(0,1)$ variable. Then, 
+	Let $\phi(z)$ be the density function of a standard normal variable $Z \sim N(0,1)$. Then, 
 	\begin{equation}\label{eq:e22a}
@@
 \end{lem}
-\begin{proof}First, we prove the equation~\eqref{eq:e22a}. The pdf of standard normal distribution $\phi(z)$ satisfies
-\[  \phi(z) = \frac{1}{\sqrt{2\pi}} \exp(-\frac{z^2}{2}  ); \quad \phi'(z) = -z \frac{1}{\sqrt{2\pi}} \exp(-\frac{z^2}{2} ) = -z \phi(z). \]
-	Next, we prove the equation~\eqref{eq:e22b}. Using equation~\eqref{eq:e22a}, we have
-	\begin{align}
-		\bbP(Z \geq z) &= \int_{z}^{+\infty} \phi(t) dt = \int_{z}^{+\infty} - \frac{1}{t} \phi'(t) dt = \frac{1}{z} \phi(z) - \int_{z}^{+\infty} \frac{1}{t^2} \phi(t) dt\\
-		&= \frac{1}{z} \phi(z) + \int_{z}^{+\infty} \frac{1}{t^3} \phi'(t) dt =  \frac{1}{z} \phi(z) - \frac{1}{z^3} \phi(z) + \int_{z}^{+\infty} \frac{3}{t^4} \phi(t) dt
-	\end{align}
-	Since $\frac{3}{t^4} \phi(t) \geq 0$, therefore $\bbP(Z \geq z) \geq \phi(z)(\frac{1}{z}  - \frac{1}{z^3} )$. On the other hand, 
-	\[ \int_{z}^{+\infty} \frac{3}{t^4} \phi(t) dt = \int_{z}^{+\infty} - \frac{3}{t^5} \phi'(t) dt =\frac{3}{ z^5} \phi(z) -  \int_{z}^{+\infty} \frac{15}{t^6} \phi(t) dt \leq \frac{3}{ z^5} \phi(z). \]
-	Therefore, $\bbP(Z \geq z) \leq \phi(z)(\frac{1}{z}  - \frac{1}{z^3} + \frac{3}{z^5} ).$
+\begin{proof}First, we prove the equation~\eqref{eq:e22a}.
+
+The pdf of standard normal distribution is
+\[  \phi(z) = \frac{1}{\sqrt{2\pi}} \exp(-\frac{z^2}{2}  ).\]
+
+The equation~\eqref{eq:e22a} follows by taking the derivative of $\phi(z)$.
+\[\quad \phi'(z) = -z \frac{1}{\sqrt{2\pi}} \exp(-\frac{z^2}{2} ) = -z \phi(z). \] 
+	Next, we prove the equation~\eqref{eq:e22b}.
+
+	We write the upper tail probability of the standard normal variable as:
+	\begin{equation}\label{eq:e22pro1}
+		\bbP(Z \geq z) = \int_{z}^{+\infty} \phi(t) dt = \int_{z}^{+\infty} - \frac{1}{t} \phi'(t) dt = \frac{1}{z} \phi(z) - \int_{z}^{+\infty} \frac{1}{t^2} \phi(t) dt,  
+	\end{equation}
+	where the second equality follows by the equation~\eqref{eq:e22b}. Applying the equation~\eqref{eq:e22b} to the last term in equation~\eqref{eq:e22pro1} yields
+	\begin{equation}\label{eq:e22pro2}
+	\int_{z}^{+\infty} \frac{1}{t^2} \phi(t) dt = \int_{z}^{+\infty} \frac{1}{t^3} \phi'(t) dt = - \frac{1}{z^3} \phi(z) + \int_{z}^{+\infty} \frac{3}{t^4} \phi(t) dt \geq  - \frac{1}{z^3} \phi(z)
+	\end{equation}
+	Plugging the equation~\eqref{eq:e22pro2} into the equation~\eqref{eq:e22pro1} , we have $\bbP(Z \geq z) \geq \phi(z)(\frac{1}{z}  - \frac{1}{z^3} )$.
+On the other hand, applying the equation~\eqref{eq:e22b} again to the equation~\eqref{eq:e22pro2} yields
+	\begin{equation}\label{eq:e22pro3}
+		\int_{z}^{+\infty} \frac{3}{t^4} \phi(t) dt = \int_{z}^{+\infty} - \frac{3}{t^5} \phi'(t) dt =\frac{3}{ z^5} \phi(z) -  \int_{z}^{+\infty} \frac{15}{t^6} \phi(t) dt \leq \frac{3}{ z^5} \phi(z).
+	\end{equation}
+
+	Combing with the equations~\eqref{eq:e22pro1}, \eqref{eq:e22pro2}, and  \eqref{eq:e22pro3}, we have $\bbP(Z \geq z) \leq \phi(z)(\frac{1}{z}  - \frac{1}{z^3} + \frac{3}{z^5} ).$
 \end{proof}
 
-\paragraph{Exercise 2.3}
+\begin{rmk}
+	Direct calculation of tail probability for a univariate normal variable is hard.
+Equation~\eqref{eq:e22b} provides a numerical approximation to the tail probability.
+Particularly, the tail probability decays at the rate of $z^{-1}e^{-z^2/2}$ as $z \rightarrow +\infty$.
+The decay rate is faster than polynomial rate $\tO(z^{-\alpha})$, for any $\alpha \geq 1$.
+\end{rmk}
+
+\subsection{Exercise 2.3}
 
 \begin{lem}[Polynomial bound and Chernoff bound]\label{lem:momcher}
-	Suppose $X \geq 0$, and that the moment generating function of $X$ exists in the neighborhood of 0. Given some $\delta > 0$ and integer $ k \in \mathbb{Z_+}$, we have
-	\[  \inf_{k \in \mathbb{Z_+} } \frac{\bbE[|X|^k]}{\delta^k} \leq \inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}}. \]
+	Consider a non-negative variable $X \geq 0$.
+Suppose that the moment generating function of $X$, $\varphi_{X}(\lambda)$, exists in the neighborhood of $\lambda = 0$.
+Given some $\delta > 0$, for any integer $ k \in \mathbb{Z_+}$, we have
+	\begin{equation}\label{eq:lempoly}
+		\inf_{k \in \mathbb{Z_+} } \frac{\bbE[|X|^k]}{\delta^k} \leq \inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}}. 
+	\end{equation}
+
 	Consequently, an optimized bound based on polynomial moments is always at least as good as the Chernoff upper bound.
@@
 	\begin{equation}\label{eq:power}
-		 e^{\lambda X} = \sum_{k = 0}^{+\infty } \frac{X^k \lambda^k }{k!}, \quad \forall \lambda \in \bbR
+		 e^{\lambda X} = \sum_{k = 0}^{+\infty } \frac{X^k \lambda^k }{k!}, \quad \forall \lambda \in \bbR.
 	\end{equation}
-	Since the moment generating function $\varphi_X(\lambda)$ exists in the neighbor hood of 0, there exists a constant $b > 0$ such that 
-	\[  \inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}} = \inf_{\lambda \in (0,b)} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}} < +\infty. \]
-	Taking the expectation on both sides of the power series~\eqref{eq:power} yields
+	Since the moment generating function $\varphi_X(\lambda)$ exists in the neighborhood of $\lambda = 0$, there exists a constant $b > 0$ such that 
 		\[   \bbE [e^{\lambda X} ]= \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} < +\infty,\quad \forall \lambda \in (0,b). \]
-	Therefore, the moment $\bbE[|X|^k] < +\infty,\  \forall k \in \mathbb{Z}_+$ exists. Applying the power serious to $e^{\lambda \delta}$, we obtain the result
-	\[ \inf_{k \in \mathbb{Z_+} } \frac{\bbE[|X|^k]}{\delta^k} \leq \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k] }{\delta^k} = \sum_{k = 0}^{+\infty } \frac{\frac{ \bbE[|X|^k]  \lambda^k }{k!} }{\frac{\lambda^k \delta^k }{k!}  } = \inf_{\lambda > 0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}}.
-\]
+	Therefore, the moment  $\bbE[|X|^k]$ exists for any $ k \in \mathbb{Z_+}$. Noted that  $\bbE[|X|^k] \geq 0$, we have
+	\begin{equation}\label{eq:single}
+		 \bbE[|X|^k] \leq \sum_{k = 0}^{+\infty}  \bbE[|X|^k], \quad \forall k \in \mathbb{Z_+}. 
+	\end{equation}
+	Applying power series~\eqref{eq:power} to the right hand side of equation~\eqref{eq:lempoly} yields
+	\begin{equation}\label{eq:powertopoly}
+		\inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}} = \sum_{k = 0}^{+\infty } \frac{\frac{ \bbE[|X|^k]  \lambda^k }{k!} }{\frac{\lambda^k \delta^k }{k!}  } =\sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k] }{\delta^k}.
+	\end{equation}
+	Combing with equation~\eqref{eq:single}, we have
+	\[  \inf_{k \in \mathbb{Z_+} } \frac{\bbE[|X|^k]}{\delta^k} \leq \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k] }{\delta^k}  = \inf_{\lambda > 0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}}. \]
 \end{proof}
 
+\begin{rmk}
+	Applying different functions $g(X)$ to the Markov's inequality leads to distinct bounds for the tail probability of variable $X$.
+Equation~\eqref{eq:lempoly} implies that optimized polynomial bound is tighter than the Chernoff bound, provided that the moment generating function of $X$ exsits.
+Actually, the optimization on polynomial bound can be released.
+Then, any polynomial bound is better than Chernoff bound.
+\end{rmk}
+
+\subsection{Exercise 2.4}\label{pro:boundvar}
+In Exercise~\ref{pro:boundvar}, we prove the theorem~\ref{thm:boundvar},  the Hoeffding bound for bounded variable.
+
+\begin{proof}
+
+Suppose $X$ is a bounded random variable that $X \in [a,b]$ almost surely. Let $\mu = \bbE[X]$. Define the function 
+\[ g(\lambda) = \log \bbE[e^{\lambda X }], \quad  \forall \lambda \in \bbR. \]
+Applying Taylor Expansion to $g(\lambda)$ at 0, we have
+\begin{equation}\label{eq:taylor}
+	g(\lambda) =  g(0) + g'(0) \lambda  + \frac{g^{''}(\lambda_0) }{2} \lambda^2, \quad \text{where } \lambda_0 = t \lambda, \text{ for some } t \in [0,1]. 
+\end{equation}
+
+In equation~\eqref{eq:taylor}, the term $g(0) =\log \bbE[e^{0}] = 0$. By power series \eqref{eq:power}, we calculate the first derivative $g'(\lambda)$ as following.
+\begin{align}
+	g'(\lambda ) & = \left( \log \sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]  \right)' \\
+	&= {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+1)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]}  \label{eq:firstde} \\
+	&= \frac{\bbE[X e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]}
+\end{align}
+Therefore, $g'(0) = \bbE[X] = \mu$. Taking the derivative to equation~\eqref{eq:firstde}, we calculate the second-order derivative $g^{''}(\lambda)$ as following.
+\begin{align}
+	g^{''}(\lambda)
+	&= {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+2)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]} - \left( {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+1)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]}  \right)^2\\
+	&= \frac{\bbE[X^2 e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]} - \left( \frac{\bbE[X e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]} \right)^2\\
+	%& \overset{\Delta}{=} \bbE_{\lambda} [X^2] - (\bbE_{\lambda} [X])^2
+\end{align}
+The second-order derivative $g''(\lambda)$ can be interpreted as the variance of $X$ with a re-weighted distribution $d P' = {e^{\lambda X}}/{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$.
+Taking integral to $dP'$, we have
+\begin{equation}
+	\int_{-\infty}^{+\infty} d P' = \int_{-\infty}^{+\infty}  \frac{e^{\lambda X}}{\bbE[ e^{\lambda X} ]} d P_X = 1,
+\end{equation} 
+which implies that the function $P'$ is indeed a distribution. Under any distribution, there is always an upper bound for the variance of random variable $X$:
+\[ var(X) = var(X - \frac{a+b}{2}) \leq \bbE[(X - \frac{a+b}{2} )^2] \leq \frac{(b-a)^2}{4}. \]
+Hence, the second-order derivative $g''(\lambda) \leq \frac{(b-a)^2}{4}$. We plug the results of $g'$ and $g^{''}$ into the equation~\eqref{eq:taylor}. Then,
+\begin{equation}\label{eq:e24}
+	g(\lambda) =  g(0) + g'(0) \lambda  + \frac{g^{''}(\lambda_0) }{2} \lambda^2 \leq 0 + \lambda \mu+ \frac{(b-a)^2}{8} \lambda^2. 
+\end{equation}
+Taking the exponential on both sides of the inequality~\eqref{eq:e24}, we have
+\begin{equation}\label{eq:boundvar}
+	\bbE[e^{\lambda X}] = \exp(g( \lambda)) \leq e^{\mu \lambda + \frac{(b-a)^2}{8} \lambda^2}.
+\end{equation}
+The equation~\eqref{eq:boundvar} implies that $X$ is a sub-Gaussian variable with at most $\sigma = \frac{(b-a)}{2}$.
+\end{proof}
+
+\begin{rmk}
+	For any bounded random variable $X$ with domain $[a,b]$, $X$ is a sub-gaussian variable with parameter at most $\sigma^2 = (b-a)^2/4$.
+Any propositions for sub-Gaussian variables can be applied to bounded variables.
+\end{rmk}
+
+\subsection{Exercise 2.5}
+\begin{lem}[Sub-Gaussian bounds and means/variance] Consider a random variable $X$ such that
+\begin{equation}\label{eq:subg}
+	 \bbE[e^{\lambda X}] \leq e^{\frac{\lambda^2 \sigma^2}{2} + \mu \lambda}, \quad \forall \lambda \in \bbR. 
+\end{equation}
+	Then, $\bbE[X] = \mu$ and $var(X) \leq \sigma^2$.
+\end{lem}
+
+\begin{proof} By equation~\eqref{eq:subg}, the moment generating function of $X$, $\varphi_X(\lambda)$ , exists in the neighborhood of $\lambda = 0$.
+Hence, the mean and variance of $X$ exist.
+Applying power series on both sides of equation~\eqref{eq:subg} yields
+\begin{equation}\label{eq:taylor2}
+	 \lambda \bbE[X] + \frac{\lambda^2}{2} \bbE[X^2] + o(\lambda^2) \leq  \mu \lambda + \frac{\lambda^2 \sigma^2 + \lambda^2 \mu^2}{2} + o(\lambda^2).
+\end{equation}
+Dividing by $\lambda >0$ on both sides of equation~\eqref{eq:taylor2} and letting $\lambda \rightarrow 0^{+}$, we have $\bbE(X) \leq \mu$ ; dividing by $\lambda <0$ on both sides of equation~\eqref{eq:taylor2} and letting $\lambda \rightarrow 0^{-}$, we have $\bbE(X) \geq \mu$.
+Therefore, the mean $\bbE[X] = \mu$.
+Then, we divide ${2}/{\lambda^2}$ on both sides of equation~\eqref{eq:taylor2}.
+The term $\bbE[X]\lambda$ and  $\mu \lambda$ are cancelled.
+We have $ \bbE[X^2] \leq \sigma^2 + \mu^2$, and thus $var(X) \leq \bbE[X^2]  - (\bbE[X])^2 = \sigma^2$.
+\end{proof}
+
+\paragraph{Question:} Let $\sigma^2_{min}$ denote the smallest possible $\sigma$ satisfying the inequality~\eqref{eq:subg}. Is it true that $var(X) = \sigma_{min}^2$?
+
+\paragraph{Solution:} The statement that $var(X) = \sigma_{min}^2$ is not always true.
+Recall the function $g(\lambda)$ in exercise~\ref{pro:boundvar}.
+By the results in exercise~\ref{pro:boundvar},  the equation~\eqref{eq:subg} is equal to
+\begin{equation}\label{eq:question}
+	g^{''}(\lambda) \leq  \sigma^2, \quad \forall \lambda \in \bbR,
+\end{equation}
+where $g^{''} (\lambda)$ is the variance of $X$ with with a re-weighted distribution $d P' = {e^{\lambda X}}/{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$ and $g^{''} (0) = var(X)$.
+Therefore, $\max_{\lambda}g^{''}(\lambda) = \sigma_{min}^2$.
+To satisfy the equality $var(X) = \sigma_{min}^2$, we need to show that $\max_{\lambda}g^{''}(\lambda) = g^{''}(0)$ holds for $X$.
+
+\vspace{0.2cm}
+However, the statement $\max_{\lambda}g^{''}(\lambda) = g^{''}(0)$ is not always true.  A counter example is below. 
+Consider a random variable $X \sim Ber(1/3)$. The variance of $X$ is $var(X) = 2/9$. Let $\lambda = 1$. The re-weighted distribution $dP'$ is
+\[ P'(X = 0) = \frac{2}{3\bbE[ e^{X} ]};\quad  P'(X = 1) = \frac{e}{3\bbE[ e^{X} ]}, \quad \text{where } \bbE[ e^{X} ] = \frac{2}{3} + \frac{e}{3}.  \]
+The variance of $X$ with $dP'$ is ${2}/{3\bbE[ e^{X} ]} \times  {e}/{3\bbE[ e^{X} ]} = 0.2442 > 2/9$. Therefore, for this variable $X$, $g^{''}(0) <  g^{''}(1) \leq \max_{\lambda}g^{''}(\lambda)$. 
+\begin{rmk}
+	Parameters of a sub-Gaussian distribution provide the exact value of the mean, $ \bbE[X] = \mu$, and an upper bound of the variance, $var(X) \leq \sigma^2$.
+For any variable $X$ whose moment generating function exists, the tail distribution of $X$ can be bounded by a sub-Gaussian distribution with a proper choice of $\sigma^2$.
+\end{rmk}
+
+\iffalse
+
+\subsection{Exercise 2.6}
+\begin{lem}[Lower bounds on squared sub-Gaussians] Let $\{ X_i \}_{i=1}^n$ be i.i.d.\ sequence of zero-mean sub-Gaussian variables with parameter $\sigma$.
+The normalized summation $Z_n = \frac{1}{n} \sum_{i=1}^n X_i^2$ satisfies
+\begin{equation}\label{lem:lowersquare}
+	\bbP[Z_n - \bbE[Z_n] \leq \sigma^2 \delta ] \leq e^{-n\delta^2 /16}, \quad \forall \delta \geq 0.
+\end{equation}
+	The equation~\eqref{lem:lowersquare} implies that the lower tail of the summation of squared sub-Gaussian variables behave in a sub-Gaussian way.
+\end{lem}
+
+\begin{proof}
+	Given a sub-Gaussian vector $(X_1,...,X_n)$, the moment generating function for $Z_n = \frac{1}{n} \sum_i X_i^2$, $\varphi_{Z_n}(\lambda)$, exists in the neighborhood of $ \lambda = 0$.
+By the Chernoff bound~\eqref{eq:chernoff1}, there exists a constant $b > 0$ such that
+	\begin{equation}
+		\bbP[Z_n - \bbE[Z_n] \geq t]  \leq \inf_{\lambda \in (-b,b)} \frac{\bbE[e^{\lambda (Z_n - \bbE[Z_n])}]}{e^{\lambda t}}.
+	\end{equation}
+	Hence, analysis to the expectation $\bbE[e^{\lambda (Z_n - \bbE[Z_n])}]$ is the key of the proof. Since $X_i$ are i.i.d.\, we re-write the expectation as following
+	 \[\bbE[e^{\lambda (Z_n - \bbE[Z_n])}] = (\bbE[ e^{ \frac{\lambda}{n} (X_1^2 - \bbE[X_1]^2)} ])^n. \]
+	Let $Z = X_1^2$. Next, we analyze the expectation $\bbE[e^{\lambda (Z - \bbE[Z])}]$.
+\end{proof}
+
+\fi
+
 \end{document}

### 2020-07-16 skills/math/linear_algebra/linear_algebra_summary_p2.tex
update

@@
 \begin{defn}[\textit{Orthogonal vectors}]\label{def:vortho}
-	Consider two vectors $x,y \in \bbR^{n}$. The vectors $x$ and $y$ are orthogonal,  denoted $x \perp y$, if and only if $x^T y = y^T x = 0$.
+	Let  $x,y \in \bbR^{n}$ be two vectors. The vectors $x$ and $y$ are orthogonal,  denoted $x \perp y$, if and only if $x^T y = y^T x = 0$.
 \end{defn}
@@
 \begin{defn}[\textit{Orthogonal subspaces}]\label{def:sortho}
-	Consider two subspaces $S,T$. The subspaces $S$ and $T$ are orthogonal,  denoted $S \perp T$, if and only if for any $s \in S$ and $ t \in T$, $s^T t = t^T s = 0$.
+	Let  $S,T$ be two subspaces $S,T$. The subspaces $S$ and $T$ are orthogonal,  denoted $S \perp T$, if and only if for any $s \in S$ and $ t \in T$, $s^T t = t^T s = 0$.
 \end{defn}
 
-Given a matrix $\mA \in \bbR^{m \times n}$, there are four subspaces related to $\mA$\colon column space $C(\mA)$, row space $C(\mA^T)$, nullspace $N(\mA)$, and left nullspace $N(\mA^T)$.
+Let $\mA \in \bbR^{m \times n}$ be a matrix.
+There are four subspaces related to $\mA\colon$ column space $C(\mA)$, row space $C(\mA^T)$, nullspace $N(\mA)$, and left nullspace $N(\mA^T)$.
 Suppose the rank of $\mA$ is $rank(\mA) = r$, the dimensions of these subspaces are:
@@
 \begin{thm}[Orthogonality of matrix subspaces]\label{thm:ortho}
-	Consider a matrix $\mA \in \bbR^{m \times n}$. The row space $C(\mA^T)$ and the nullspace $N(\mA)$ are orthogonal; the column space $C(\mA)$ and the left nullspace $N(\mA^T)$ are orthogonal, i.e.\
-	\[ C(\mA^T) \perp N(\mA); \quad C(\mA) \perp N(\mA^T).  \]
+	Let  $\mA \in \bbR^{m \times n}$ be a matrix. The row space $C(\mA^T)$ and the nullspace $N(\mA)$ are orthogonal. The column space $C(\mA)$ and the left nullspace $N(\mA^T)$ are orthogonal; i.e.,
+	\[ C(\mA^T) \perp N(\mA)\quad  \text{and} \quad C(\mA) \perp N(\mA^T).  \]
 \end{thm}
@@
 \begin{proof}
-	Consider the matrix $\mA \in \bbR^{m \times n}$. For any vector $x \in N(\mA)$, we have $\mA x = 0$. 
+	For any vector $x \in N(\mA)$, we have $\mA x = 0$. Specifically,
 	\begin{align}
@@
 	\end{align}
-	where $a_i \in \bbR^n, i \in [m]$ is the $i$-th  row of $\mA$.
-By definition~\ref{def:vortho}, $x$ is orthogonal to the rows in matrix $\mA$.
-For any vector $v \in C(\mA^T)$, $v$ is a linear combination of the rows $v = c_1 a_1 + ...+c_m a_m$, where $c_i, i\in [m]$ are constants.
-Multiplying vectors $v$ and $x$, we have
+	where $a_i \in \bbR^n$ is the $i$-th  row of $\mA$, for all $ i \in [m]$.
+By Definition~\ref{def:vortho}, $x$ is orthogonal to the rows in matrix $\mA$.
+For any vector $v \in C(\mA^T)$, $v$ is a linear combination of the rows $v = c_1 a_1 + ...+c_m a_m$, where $c_i \in \bbR$, for all $i\in [m]$.
+Taking inner product between vectors $v$ and $x$, we have
 	\[  v^T x =  c_1 a_1^T x + \cdots + c_m a_m^T x = 0.  \]
@@
 
-\begin{thm}[Relationship between $\mA^T \mA$ and $\mA$]\label{thm:aa}
-	Consider a matrix $\mA \in \bbR^{m \times n}$. We have
+\begin{thm}[Relationship between $\mA^T \mA$ and $\mA$]\label{thm:aa}Let $\mA \in \bbR^{m \times n}$ be a matrix. We have
 	\begin{equation}
@@
 \begin{proof}
-	Consider a matrix $\mA \in \bbR^{m \times n}$. First, we prove that $N(\mA^T \mA) = N(\mA)$.
+	Fxirst, we prove that $N(\mA^T \mA) = N(\mA)$.
 
-	On one hand, if $x \in N(\mA)$, then $\mA x = 0 \Rightarrow \mA^T  \mA x = \mA^T 0 = 0$. Therefore, for any $x \in N(\mA)$, the vector $x \in N(\mA^T \mA)$.
+	On one hand, if $x \in N(\mA)$, then $\mA x = 0$, which implies that $ \mA^T  \mA x = \mA^T 0 = 0$. Therefore, for any $x \in N(\mA)$, the vector $x \in N(\mA^T \mA)$.
 
-	On the other hand, we  prove that for any $x \in N(\mA^T \mA)$ the vector $x \in N(\mA)$ by contradiction.
+	On the other hand, we  prove  by contradiction that for any $x \in N(\mA^T \mA)$ the vector $x \in N(\mA)$.
 
@@
 	By the first equality in \eqref{eq:aaa}, $b \in C(\mA)$, and by the third equation in \eqref{eq:aaa}, $ b \in N(\mA^T)$.
-This contradicts the Theorem~\ref{thm:ortho}, $C(\mA) \perp N(\mA^T)$.
+This contradicts the Theorem~\ref{thm:ortho} that $C(\mA) \perp N(\mA^T)$.
 Therefore, for any $x \in N(\mA^T \mA)$, the vector $x \in N(\mA)$.
@@
 \begin{proof}
-	If $\mA \in \bbR^{m \times n}$ has independent columns, then $rank(\mA) = n$.
+Suppose $\mA \in \bbR^{m \times n}$ is a matrix with independent columns; i.e., $rank(\mA) = n$.
 By Theorem~\ref{thm:aa}, $rank(\mA^T \mA) = rank(\mA) = n$.
@@
 \subsection{Projections onto subspaces}
-\begin{defn}[\textit{Projection and projection matrix}]\label{def:project}
-	Given a vector $x \in \bbR^{n}$ and a matrix $\mA^{m \times n}$ with independent columns, the vector $p \in C(\mA)$, such that
+\begin{defn}[\textit{Projection and projection matrix}]\label{def:project} Let  $x \in \bbR^{n}$ be a vector and $\mA^{n \times m}$ be a matrix with independent columns.
+The vector $p \in C(\mA)$ that satisfies
 	\begin{equation}\label{eq:project}
@@
 	\end{equation} 
-
-	is the projection of vector $x$ onto the space $C(\mA)$. Since $p \in C(\mA)$, there exists a vector $\hat x$ such that $p = \mA \hat x$. By equation~\eqref{eq:project}, we have
-	\[ \mA^T (x - p) = \mA^T(x - \mA \hat x) = 0  \quad  \Rightarrow\quad  \hat x = (\mA^T \mA)^{-1} \mA^T x,\quad p = \mA(\mA^T \mA)^{-1} \mA^T x.  \] 
-	The matrix $\mP \overset{\Delta}{=}   \mA(\mA^T \mA)^{-1} \mA^T$ is called a projection matrix.
+	is called the projection of vector $x$ onto the column space $C(\mA)$. For all $x \in \bbR^{n}$ and the corresponding projection $p$, the matrix $\mP \in \bbR^{m \times m}$ that satisfies
+	\begin{equation}
+		p = \mP x,
+	\end{equation}
+	is called the projection matrix of $\mA$ from $\bbR^n$ onto the column space  $C(\mA)$.
 \end{defn}
 
+\begin{prop}[Projection matrix of $C(\mA)$]\label{prop:project}
+	Let  $x \in \bbR^{n}$ be a vector and $\mA^{m \times n}$ be a matrix with independent columns. The projection matrix of $\mA$ from $\bbR^n$ onto the column space  $C(\mA)$ is 
+	\begin{equation}
+		\mP =  \mA (\mA^T \mA)^{-1} \mA^T.
+	\end{equation}
+\end{prop}
+
+\begin{proof}
+	By Definition~\ref{def:project}, the projection $p \in C(\mA)$.
+Then, there exists a  vector $\hat x \in \bbR^m$ such that $p = \mA \hat x$.
+By equation~\eqref{eq:project}, for all $x \in \bbR^n$, we have
+\[ A^T (x - \mA \hat x) = 0 \quad \Rightarrow  \quad \hat x = (\mA^T \mA)^{-1} \mA^T x \quad \Rightarrow \quad p = \mA (\mA^T \mA)^{-1} \mA^T  x. \]
+Therefore, the matrix $\mP =  \mA (\mA^T \mA)^{-1} \mA^T$ is the projection matrix of $\mA$ from $\bbR^n$ onto the column space $C(\mA)$.
+\end{proof}
+
 \begin{thm}[Properties of projection matrix]\label{thm:projectm}
-	Consider a projection matrix $\mP$ to the column space $C(\mA)$, where $\mA \in \bbR^{m \times n}$ is a matrix. Then,
+Let $\mA \in \bbR^{m \times n}$ be a matrix, and $\mP$ be the projection matrix of $\mA$ onto the column space. Then,
 	\begin{equation}\label{eq:projectm}
@@
 \begin{proof}
-	Since the projection matrix $\mP =   \mA(\mA^T \mA)^{-1} \mA^T$ and $\mA^T \mA$ is symmetric, then
+	By Proposition~\ref{prop:project}, the projection matrix is $\mP =   \mA(\mA^T \mA)^{-1} \mA^T$. Then,
 	\[  \mP^T = (\mA(\mA^T \mA)^{-1} \mA^T )^T = \mA (\mA^T \mA)^{-1} \mA^T = \mP. \]
@@
 \begin{cor}[Projection onto $N(\mA^T)$]\label{cor:iprojectm}
-	Suppose $\mP$ is a projection matrix in Theorem~\ref{thm:projectm}, then $\mI-\mP$ is also a projection matrix to the left nullspace $N(\mA^T)$. 
+	Suppose $\mP$ is a projection matrix in Theorem~\ref{thm:projectm}. Then $\mI-\mP$ is also a projection matrix of $\mA$ from $\bbR^n$ onto the left nullspace $N(\mA^T)$. 
 \end{cor}
@@
 \begin{proof}
-	For any $x \in \bbR^n$, we have
-	\[ x - \mP x \perp C(\mA) \quad \Rightarrow \quad (I-\mP)x \perp C(\mA) \quad \Rightarrow \quad  (\mI-\mP)x \in N(\mA^T) \text{ and } (x - (\mI-\mP)x) \perp N(\mA^T). \]
-	Therefore, $\mI-\mP$ is a projection matrix to the left nullspace $N(\mA^T)$.
+	By Definition~\ref{def:project}, for any $x \in \bbR^n$, we have
+	\[ x - \mP x \perp C(\mA) \quad \Rightarrow \quad (I-\mP)x \perp C(\mA). \]
+	By Theorem~\ref{thm:ortho}, the column space of $\mA$ is orthogonal to the left null space of $\mA$. Then,
+	\[ (\mI-\mP)x \in N(\mA^T) \quad \text{ and }\quad  (x - (\mI-\mP)x) \perp N(\mA^T). \]
+	Therefore, $\mI-\mP$ is a projection matrix of $\mA$ from $\bbR^n$ onto the left nullspace $N(\mA^T)$.
 \end{proof}
@@
 \subsection{Projection matrices and least squares}
-Given an observation vector $y \in \bbR^n$ and a design matrix $\mX \in \bbR^{n \times (k+1)}$, we propose the linear regression model
+Let $y \in \bbR^n$ be a vector and $\mX \in \bbR^{n \times (k+1)}$ be a design matrix. We propose the linear regression model as
 \[ y = \mX \beta +\epsilon,\]
-where $\beta = (\beta_0,\beta_1,...,\beta_k)$ are regression coefficients and $\epsilon$ is the noise. The least square estimate of $\beta$ is the minimizer of the loss, i.e.\
+where $\beta = (\beta_0,\beta_1,...,\beta_k)$ are regression coefficients and $\epsilon$ is the noise. The least square estimate of $\beta$ is the minimizer of the loss; i.e.,
 \[  \hat \beta_{LS} = \argmin_{\beta \in \bbR^{k+1}} \onorm{y - \mX \beta}^2, \]
-where $\onorm{\cdot}$ is the euclidean norm. The vector $\mX 
-\hat \beta_{LS}$ can be considered as a the projection of $y$ onto the column space of $\mX$, which minimizes the distance from $y$ to the column space $C(\mX)$.
-Therefore, we may use projection tools to solve the least square estimate.
+where $\onorm{\cdot}$ is the euclidean norm. We consider the vector $\mX 
+\hat \beta_{LS}$ as the projection of $y$ onto the column space of $\mX$, which minimizes the distance from $y$ to the column space $C(\mX)$.
+Therefore, we may use projection to solve the minimization problem.
 
-The projection $\mX \hat \beta_{LS}$ satisfies
+By Definition~\ref{def:project}, the projection $\mX \hat \beta_{LS}$ satisfies
 \[ \mX^T( y - \mX \hat \beta_{LS} ) = 0 \quad \Rightarrow \quad \hat \beta_{LS} = (\mX^T \mX)^{-1} \mX^T y. \]
 
-The estimate $\hat \beta_{LS}$ consists with the estimates solved by using the derivative.
+The estimate $\hat \beta_{LS}$ is consistent with the estimates solved by using the derivative.
 
@@
 \begin{defn}[\textit{Orthonormal matrix and orthogonal matrix}]\label{def:orthonom}
-	Consider a matrix $\mQ \in \bbR^{m \times n}$. If the columns of $\mQ$ are orthonormal, the matrix $\mQ$ is an orthonormal matrix. If $ m= n$, the square matrix $\mQ$ is an orthogonal matrix.
+	Let  $\mQ \in \bbR^{m \times n}$ be a matrix. The matrix $\mQ$ is an orthonormal matrix, if the columns of $\mQ$ are orthonormal. When $ m= n$, the square matrix $\mQ$ is an orthogonal matrix.
 \end{defn}
 
-If $\mQ \in \bbR^{m \times n}$ is an orthonormal matrix, $\mQ^T \mQ = \mI_n$.  If $\mQ$ is an orthogonal matrix, $\mQ^T = \mQ^{-1}$.
-For orthonormal matrix $\mQ$, the projection matrix to $C(\mP)$  becomes $\mP = \mI_m$.
+Suppose $\mQ \in \bbR^{m \times n}$ is an orthonormal matrix.
+Then, we have $\mQ^T \mQ = \mI_n$.
+When $m = n$, $\mQ$ is an orthogonal matrix and $\mQ^T = \mQ^{-1}$.
+The projection matrix of $\mQ$ onto the column space $C(\mQ)$, denoted $\mP$,  becomes $\mP = \mI_m$.
 
 \begin{defn}[\textit{Gram-Schimidt Process and QR decomposition}]\label{gram-schimidt}
-Consider a matrix $\mA \in \bbR^{m \times n}$ with $rank(\mA) = n$. Gram-Schimidt process finds the orthonormal basis for $C(\mA)$. Let $a_i \in \bbR^{m}, i \in [m]$ be the columns of matrix $\mA$. 
+Let $\mA \in \bbR^{m \times n}$ be a matrix with $rank(\mA) = n$, and  $a_i \in \bbR^{m}$ be the column of matrix $\mA$, for all $i \in [n]$.
+Gram-Schimidt process finds the orthonormal basis for $C(\mA)$ as following,
 \begin{align}
-	&u_1 = a_1, &&\quad e_1 = \frac{u_1}{\onorm{u_1}}\\
-	&u_2 = a_2 - \frac{u_1^T a_2}{u_1^T u_1} u_1 ,&&\quad  e_2 = \frac{u_2}{\onorm{u_2}}\\
-	&u_3 = a_3 - \frac{u_1^T a_3}{u_1^T u_1} u_1 - \frac{u_2^T a_3}{u_2^T u_2} u_2,&&\quad  e_2 = \frac{u_3}{\onorm{u_3}}\\
-	\vdots
+	&u_1 = a_1, &&\quad e_1 = \frac{u_1}{\onorm{u_1}};\\
+	&u_2 = a_2 - \frac{u_1^T a_2}{u_1^T u_1} u_1 ,&&\quad  e_2 = \frac{u_2}{\onorm{u_2}};\\
+	&u_3 = a_3 - \frac{u_1^T a_3}{u_1^T u_1} u_1 - \frac{u_2^T a_3}{u_2^T u_2} u_2,&&\quad  e_2 = \frac{u_3}{\onorm{u_3}};\\
+	\vdots\\
+	&u_n = a_n - \sum_{k = 1}^{n-1} \frac{u_k^T a_k}{u_k^T u_k} u_k, && \quad e_n = \frac{u_n}{\onorm{u_n}}.
 \end{align}
- The vectors $e_1,...,e_n$ are orthonormal basis of the $C(\mA)$. By matrix operations, we obtain a decomposition of matrix $\mA$
+ The vectors $e_1,...,e_n$ are orthonormal basis of the $C(\mA)$. Based on the basis obtained by Gram-Schimidit, we decompose the matrix $\mA$ as following,
  \begin{equation}\label{eq:qr}
@@
  		0 &0&\cdots & e_n^T a_n
- 	\end{bmatrix} = \mQ \mR
+ 	\end{bmatrix} = \mQ \mR,
  \end{equation}
-	where $\mQ \in \bbR^{m \times n}$ is an orthonormal matrix and $\mR \in \bbR^{n \times n}$ is an upper triangular matrix. We call the matrix decomposition in equation~\eqref{eq:qr} QR decomposition.
+	where $\mQ \in \bbR^{m \times n}$ is an orthonormal matrix and $\mR \in \bbR^{n \times n}$ is an upper triangular matrix.
+The matrix decomposition in equation~\eqref{eq:qr} is called QR decomposition.
 \end{defn}
@@
 \subsection{Properties of determinants}
-We give ten properties of determinants. The last seven properties are deduced by first three basic properties.
+We give ten properties of determinants. The last seven properties are deduced by the first three basic properties.
 \begin{enumerate}
-	\item\label{prop1} The determinant of identity matrix is equal to 1, i.e.\ $\det(\mI) = 1$.
+	\item\label{prop1} The determinant of identity matrix is equal to 1; i.e., $\det(\mI) = 1$.
 	\item\label{prop2} Exchanging two rows of a matrix reverses the sign of the matrix's determinant.
@@
 
-	\item \label{propidenticalrow} If two rows of a matrix are equal, the determinant of the matrix is 0.  
+	\item \label{propidenticalrow} If the matrix contains two identical rows, the determinant of the matrix is 0.  
 	\begin{proof}
 		Suppose $\mA$ has two identical rows $a_i, a_j$.
-The matrix after exchanging $a_i, a_j$, $\mA'$, is the same as the original matrix $\mA$.
-By property~\ref{prop2}, $\det(\mA) = -\det(\mA') = -\det(\mA)$.
+The matrix after exchanging $a_i, a_j$, denoted $\mA'$, is the same as the original matrix $\mA$.
+By property~\ref{prop2}, we have $\det(\mA) = -\det(\mA') = -\det(\mA)$.
 Therefore, $\det(\mA) = 0$.
@@
 
-	\item \label{propelim} For a square matrix $\mA \in \bbR^{n \times n}$, subtracting $t$ times of the $i$-th row from the $j$-th row does not change the determinant of $\mA$, where $i \neq j$.
+	\item \label{propelim} Subtracting $t$ times of the $i$-th row from the $j$-th row does not change the determinant of the matrix, where $i \neq j$.
 	\begin{proof}
-		By property~\ref{prop3} and property~\ref{propidenticalrow}, 
+		Take a two-by-two matrix as an example. By property~\ref{prop3} and property~\ref{propidenticalrow}, 
 	\begin{align}
@@
 		\end{vmatrix}  = \begin{vmatrix}
-			a & b \\ c &d.
-		\end{vmatrix}
+			a & b \\ c &d
+		\end{vmatrix}.
 	\end{align}
@@
 
-	\item\label{prop0row} If the matrix $\mA$ has a row that is all zeros, the determinant of $\mA$ is 0.
+	\item\label{prop0row} If the matrix has a row that is all zeros, the determinant of the matrix is 0.
 	\begin{proof}
@@
 
-	\item\label{propdiag} Suppose $\mA \in \bbR^{n \times n}$ is a triangular matrix with diagonal elements $d_1,...,d_n$. Then, the determinant $\det(\mA) = \prod_{i = 1}^n d_i$. 
+	\item\label{propdiag} Let $\mA \in \bbR^{n \times n}$ be a triangular matrix with diagonal elements $d_1,...,d_n$. The determinant $\det(\mA) = \prod_{i = 1}^n d_i$. 
 	\begin{proof}
-		By property~\ref{propelim}, eliminating the triangular matrix $\mA$ to a diagonal matrix does not change the determinant of $\mA$.
-By property~\ref{prop3} and \ref{prop1}, $\det(\mA) = d_1 ...d_n \det(\mI_n) = d_1 ...d_n $.
+	We eliminate the matrix $\mA$ to a diagonal matrix, denoted $\mA'$.
+By property~\ref{propelim}, $\det{\mA} = \det{\mA'}$.
+Since $\mA$ is triangular, the diagonal elements of $\mA'$ are still $d_1,...,d_n$.
+ 		 By property~\ref{prop3} and \ref{prop1}, $\det(\mA') = \prod_{i = 1}^n d_i\det(\mI_n) =\prod_{i = 1}^n d_i$. Therefore, $\det(\mA) = \prod_{i = 1}^n d_i$.
 	\end{proof}
 
-	\item\label{propsingular} If the square matrix $\mA$ is singular, the determinant of $\mA$ is 0.
+	\item\label{propsingular} If the square matrix is singular, the determinant of the matrix is 0.
 
 	\begin{proof}
-		If $\mA$ is a singular matrix, the reduced row echelon form of $\mA$, RREF($\mA$), has a row with all 0 entries. By property~\ref{propelim} and \ref{prop0row}, the determinant of $\mA$ is 0.
+		If $\mA$ is a singular matrix, the reduced row echelon form of $\mA$, denoted $RREF(\mA)$, has a row with all 0 entries.
+By and \ref{prop0row}, $\det(RREF(\mA)) = 0$.
+Therefore, by property~\ref{propelim}, we have $\det(\mA) =\det(RREF(\mA)) =  0$.
 	\end{proof}
 
-	\item\label{propproduct} For square matrices $\mA, \mB \in \bbR^{n \times n}$, the determinant of the product satisfies $\det(\mA \mB) = \det(\mA) \det(\mB).$
+	\item\label{propproduct} Let  $\mA, \mB \in \bbR^{n \times n}$ be two square matrices, the determinant of the matrix product is $\det(\mA \mB) = \det(\mA) \det(\mB).$
 
 	\begin{proof}
-		If at least one of $\mA, \mB$ is singular, the product $\mA \mB$ is also singular.
-Then, $\det(\mA \mB) = \det(\mA) \det(\mB) = 0$.
-If both $\mA, \mB$ are invertible, $\mA, \mB$ can be eliminated to diagonal matrices $\mD_{A}, \mD_{B}$ by elimination.
-Suppose $a_1,...,a_n$ and $b_1,...,b_n$ are diagonal elements for $\mD_A$ and $\mD_B$ respectively.
-By property~\ref{propdiag}, $\det(AB) = \prod_{i=1}^n a_i b_i = \det(\mA) \det(\mB)$.
+		First, if at least one of $\mA, \mB$ is singular, the product $\mA \mB$ is also singular. Then, $\det(\mA \mB) = \det(\mA) \det(\mB) = 0$.
+
+		Second, let $\mE$ be an elimination matrix.
+The product $\mE \mB$ means implementing a linear operation on the rows of $\mB$.
+By property~\ref{propelim}, linear operation does not change the determinant of $\mB$.
+Therefore, we have
+		\begin{equation}\label{eq:producteli}
+			|\mE \mB| = |\mE| |\mB|
+		\end{equation}
+
+		Last, suppose that both $\mA, \mB$ are invertible. By elimination, there exist a sequence of elimination matrix $\{ \mE_i \}_{i = 1}^n$ such that 
+		\[ \mE_n \cdots \mE_1 \mA = \mI.  \]
+		By Lemma 2 in Linear Algebra-Part I, the inverse of an elimination matrix is still an elimination matrix. Let $\mE'_k$ denote the inverse $\mE^{-1}_k$, for all $k \in [n]$. We rewrite $\mA$ as
+		\[ \mA = \mE'_n \cdots \mE'_1. \]
+		Combing equation~\eqref{eq:producteli}, we have
+		\begin{equation}\label{eq:prod1}
+			|\mA \mB| = | \mE'_n \cdots \mE'_1 \mB | = |\mE'_n| |\mE'_{n-1} \cdots \mE'_1 \mB| = \cdots = |\mE'_n|\cdots |\mE'_1||\mB|. 
+		\end{equation}
+		Applying the equation~\eqref{eq:producteli} again to the term $|\mE'_n|\cdots |\mE'_1| $, we have
+		\begin{equation}\label{eq:prod2}
+			|\mE'_n|\cdots |\mE'_1|  =|\mE'_n|\cdots |\mE'_3| |\mE'_2 \mE'_1| = \cdots = | \mE'_n \cdots \mE'_1|.
+		\end{equation}
+		Hence, combing equation~\eqref{eq:prod1} with equation~\eqref{eq:prod2}, we obtain the result 
+		\[ |\mA \mB|  = | \mE'_n \cdots \mE'_1| |\mB| = |\mA| |\mB|.\]
 	\end{proof}
 
-	\item\label{proptrans} For a square matrix $\mA$, the determinant $\det(\mA^T) = \det(\mA)$.
+	\item\label{proptrans} Let $\mA$ be a square matrix, the determinant $\det(\mA^T) = \det(\mA)$.
 
 	\begin{proof}
-		By LU decomposition, the matrix $\mA = \mL \mU$, where $\mL$ is a lower-triangular matrix and $\mU$ is an upper-triangular matrix.
-The transport $\mL^T ,\mU^T$ have the same diagonal elements with $\mL, \mU$ respectively.
+		By LU decomposition, we rewrite the matrix as $\mA = \mL \mU$, where $\mL$ is a lower-triangular matrix and $\mU$ is an upper-triangular matrix.
+Note that the transports $\mL^T,\mU^T$ have the same diagonal elements with $\mL, \mU$ respectively.
 By property~\ref{propdiag}, $\det(\mL^T)= \det(\mL),  \det(\mU^T) = \det(\mU)$.
-Therefore, by property~\ref{propproduct}, $\det(\mA) = \det(\mL) \det(\mU) = \det(\mU^T) \det(\mL^T) = \det(\mU^T \mL^T) = \det(\mA^T)$.
+Therefore, combing property~\ref{propproduct}, we have  $\det(\mA) = \det(\mL) \det(\mU) = \det(\mU^T) \det(\mL^T) = \det(\mU^T \mL^T) = \det(\mA^T)$.
 			\end{proof}
@@
 \begin{prop}[Big formula for computing determinant]
-	Consider a square matrix $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$. The big formula for computing the determinant of $\mA$ is
+	Let  $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ be a square matrix. The big formula for computing the determinant of $\mA$ is
 	\[ \det(\mA) =  \sum_{(\alpha_1,...,\alpha_n) \in \tP} (-1)^{N(\alpha_1,...,\alpha_n)} a_{1,\alpha_1} ... a_{n \alpha_n},  \]
@@
 \begin{proof}
-	For a square matrix $\mA = \entry{a_ij} \in \bbR^{n \times n}$, we decompose the  matrix $\mA$ by the summation of $n^n$ matrices, in which there is only one entry comes from $\mA$ in each row and other entries are 0.
-By property~\ref{prop3}, the determinant of $\mA$ is equal to the summation of the determinants of these $n^n$ matrices.
+	Let $\mA = \entry{a_ij} \in \bbR^{n \times n}$ be a square matrix.
+We decompose the  matrix $\mA$ by the summation of $n^n$ matrices, denoted $\mA'_i$, where $i \in [n^n]$.
+For any $\mA'_i$, there is only one entry comes from $\mA$ in each row and other entries are 0.
+By property~\ref{prop3}, the determinant $\det(\mA) = \sum_{i=1}^{n^n} \det(\mA'_i)$.
 Take a 3-by-3 square matrix as an example.
@@
 			 0&0 & a_{33}
-		\end{vmatrix} \overset{\Delta}{=} \mA_1 + \mA_2 +\cdots + \mA_{27}.
+		\end{vmatrix} = \mA'_1 + \mA'_2 +\cdots + \mA'_{27}.
 	\end{align}
-	For the square matrix $\mA$,  there are $n!$ nonsingular matrices among the $n^n$ matrices.
+	There are $n!$ nonsingular matrices among matrices $\mA'_i$.
 The number of nonsingular matrices coincides with the size of $\tP$.
@@
 
-	By property~\ref{propdiag} and \ref{prop2}, the determinant for the nonsingular matrix follows the formula \[(-1)^{N(\alpha_1,...,\alpha_n)}  a_{1,\alpha_1} ... a_{n \alpha_n},\]
+	By property~\ref{propdiag} and \ref{prop2}, the determinant of a nonsingular $\mA'_i$ follows the formula \[(-1)^{N(\alpha_1,...,\alpha_n)}  a_{1,\alpha_1} ... a_{n \alpha_n},\]
 	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$, $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of necessary exchanges from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
@@
 \begin{defn}[\textit{Cofactors, cofactor matrix, and cofactor formula}]
-	Consider a square matrix $\mA=  \entry{a_{ij}} \in \bbR^{n \times n}$.
-Let $\mA_{-i,-j}$ be the submatrix of $\mA$ after removing the $i$-th row and $j$-th column.
-The cofactor associated with $a_{ij}$ is \[C_{ij} = (-1)^{i+j}\det(\mA_{-i,-j}).\]
+	Let $\mA=  \entry{a_{ij}} \in \bbR^{n \times n}$ be a square matrix, and $\mA_{-i,-j}$ be the submatrix of $\mA$ after removing the $i$-th row and $j$-th column.
+The cofactor associated with $a_{ij}$ is defined as \[C_{ij} = (-1)^{i+j}\det(\mA_{-i,-j}).\]
 	The matrix $\mC = \entry{C_{ij}} \in \bbR^{n \times n}$ is called cofactor matrix.  The cofactor formula of $\det(\mA)$ is
-	\[ \det(\mA) = \sum_{j}^n a_{ij} C_{ij} = \sum_{j}^n a_{ji} C_{ji} ,\quad \forall i \in [n].  \]
+	\[ \det(\mA) = \sum_{j}^n a_{ij} C_{ij} = \sum_{j}^n a_{ji} C_{ji} ,\quad \text{ for all } i \in [n].  \]
 \end{defn}
@@
 \begin{example}[Tridiagonal matrix]
-	One example of cofactor formula is computing the determinant of \textit{tridiagonal matrix}.
-A tridiagonal matrix is a matrix in which only nonzero elements lie on or adjacent to the diagonal.
-Let $\mT_n \in \bbR^{n \times n}$ denote the tridiagonal matrix of 1's, i.e.\
+	One example of using cofactor formula is computing the determinant of \textit{tridiagonal matrix}.
+The tridiagonal matrix is a matrix in which only nonzero elements lie on or adjacent to the diagonal.
+Let $\mT_n \in \bbR^{n \times n}$ denote the tridiagonal matrix of 1's; i.e.,
 \begin{align}
@@
 		0&0&0&\cdots&1&1
-	\end{bmatrix}_{n \times n}
+	\end{bmatrix}_{n \times n}.
 \end{align}
 
-Let $\mT_{n, -i,-j}$ be the submatrix of  $\mT_n$ after removing the $i$-th row and the $j$-th column.   By cofactor formula, 
+Let $\mT_{n, -i,-j}$ be the submatrix of  $\mT_n$ after removing the $i$-th row and the $j$-th column.   By cofactor formula, we have
 \[  \det(\mT_n) = 1 \times \det(\mT_{n, -1,-1}) - 1 \times \det(\mT_{n,-1,-2}). \]
-By the definition of tridiagonal matrix, $\mT_{n,-1,-1} = \mT_{n-1}$ and $\det( \mA_{n,-1,-2}) = 1 \times \det(\mT_{n-2}) $. Therefore, 
+By the definition of tridiagonal matrix, $\mT_{n,-1,-1} = \mT_{n-1}$, and $\det( \mA_{n,-1,-2}) = 1 \times \det(\mT_{n-2}) $. Therefore, 
 \[ \det(\mT_n) = \det(\mT_{n-1}) - \det(\mT_{n-2}).  \] 
@@
 \begin{thm}[Inverse matrix by cofactors]\label{thm:inver}
-Let $\mA \in \bbR^{n \times n}$ be an invertible matrix and $\mC$ be the cofactor matrix of $\mA$. The inverse matrix of $\mA$ satisfies
+Let $\mA \in \bbR^{n \times n}$ be an invertible matrix, and $\mC$ be the cofactor matrix of $\mA$. The inverse matrix of $\mA$ satisfies
 \[  \mA^{-1} = \frac{1}{\det(\mA)} \mC^T. \]	
@@
 \begin{proof}
-	It is equivalent to  show that $\mA \mC^T = \det(\mA) \mI_n$. By the definition of cofactor matrix,
+	It is equivalent to  show that $\mA \mC^T = \det(\mA) \mI_n$.
+
+	Let $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ be an invertible matrix, and $C_{ij}$ be the cofactor of $a_{ij}$. By the definition of cofactor matrix, we have
 	\begin{align}
@@
 			\vdots & \vdots & & \vdots\\
-			C_{1n} & C_{2n} &\cdots &C_{nn}.\\
-		\end{bmatrix}
+			C_{1n} & C_{2n} &\cdots &C_{nn}\\
+		\end{bmatrix},
 	\end{align}
 
-	Therefore, the diagonal elements of $\mA \mC^T$ are $(\mA \mC^T)_{ii} = \sum_k^n a_{ik} C_{ik} = \det(\mA), \forall i \in [n]$.
+	Therefore, the diagonal elements of $\mA \mC^T$ are $(\mA \mC^T)_{ii} = \sum_k^n a_{ik} C_{ik} = \det(\mA)$, for all $ i \in [n]$.
 Consider the off-diagonal entries $(\mA \mC^T)_{ij} = \sum_{k}^n a_{ik} C_{jk}$, where $i \neq j$.
 By cofactor formula, the summation   $ \sum_{k}^n a_{ik} C_{jk}$ is equal to the determinant of a matrix whose $i$-th row and $j$-th row are identical.
-By property~\ref{prop0row}, the off-diagonals of $\mA \mC^T$ are 0.
+Combining property~\ref{prop0row}, the off-diagonals of $\mA \mC^T$ are 0.
 Then, we have $\mA \mC^T = \det(\mA) \mI_n$.
@@
 \begin{defn}[\textit{Cramer's rule}]
-Consider the linear system $\mA x=b$, where $\mA\in\bbR^{n \times n}$ is an invertible matrix, and $ x, b \in \bbR^n$ are vectors.
-Applying Theorem~\ref{thm:inver}, we obtain the Cramer's rule of $x = \mA^{-1} b$.
-\[  x = \frac{1}{\det(\mA)} \mC^T b; \quad x_j = \frac{\det(\mB_j)}{\mA_j}, \] 
-	where $B_j$ the matrix that replaces the $j$-th column of $\mA$ by $b$.
+Let  $\mA\in\bbR^{n \times n}$  be  an invertible matrix, $ x, b \in \bbR^n$ be vectors.
+Applying Theorem~\ref{thm:inver} to the linear system $\mA x=b$, we obtain the Cramer's rule of $x = \mA^{-1} b$,
+\[  x = \frac{1}{\det(\mA)} \mC^T b\quad \text{ and } \quad x_j = \frac{\det(\mB_j)}{\mA_j}, \] 
+	where $\mB_j$ is the matrix $\mA$ after replacing the $j$-th column by $b$.
 \end{defn}
@@
 \begin{defn}[\textit{Parallelepiped}]
-	The parallelepiped determined by $n$ vectors $v_1,...,v_n \in \bbR^n$ is the subset 
+	The parallelepiped determined by $n$ vectors $v_1,...,v_n \in \bbR^n$ is defined as the following subset,
 	\[ P = \{ a_1 v_1 + \cdots + a_n v_n \colon 0 \leq a_1,...,a_n \leq 1 \}. \]
@@
 \begin{prop}[Determinants and volumes]
-The absolute determinant of a square matrix $\mA$ is the volume of parallelepiped determined by the rows of $\mA$, i.e.\
+The absolute determinant of a square matrix $\mA$ is the volume of parallelepiped determined by the rows of $\mA$; i.e.,
 	\[ |\det(\mA)| = vol(P(\mA)), \]
@@
 \begin{defn}[\textit{Eigenvalues and eigenvectors}]\label{def:eigen}
-	Consider a square matrix $\mA$, an eigenvector of $\mA$ is a nonzero vector such that:
+	Let $\mA \in \bbR^{n \times m}$ be a square matrix. Suppose there is a nonzero vector $x \in \bbR^{m}$ such that
 	\[ \mA x = \lambda x, \quad \text{for some } \lambda \in \mathbb{C}. \]
-	The value $\lambda$ is called the eigenvalue of $\mA$, and $x$ is the eigenvector associated with eigenvalue $\lambda$.
-Usually, eigenvectors are normalized, i.e.\ $\onorm{x} = 1$, where $\onorm{\cdot}$ is the euclidean distance.
+	The vector $x$ is called the eigenvector of $\mA$. The value $\lambda$ is called the eigenvalue of $\mA$, and $x$ is the eigenvector associated with eigenvalue $\lambda$.
+\end{defn}
+
+ Usually, eigenvectors are normalized; i.e., $\onorm{x} = 1$, where $\onorm{\cdot}$ is the euclidean distance.
 For simplicity, all the eigenvectors mentioned below are normalized.
 Besides, the eigenvectors associated with eigenvalue 0 span the nullspace of $\mA$.
-\end{defn}
 
 \begin{defn}[\textit{Trace of square matrix}]\label{def:trace}
-Consider a square matrix $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$. The trace of $\mA$ is defined as
+Let $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ be a square matrix. The trace of $\mA$ is defined as
 \[ tr(\mA) = \sum_{i}^n a_{ii}.\] 
@@
 \begin{defn}[\textit{Characteristic polynomial}]\label{def:charac}
-	Let $\lambda$ denote an eigenvalue of the matrix $\mA\in \bbR^{n \times n}$. Then the matrix $\mA - \lambda \mI$ is a singular matrix. Therefore,
-	\[ \det(\mA - \lambda I_n) = 0 \quad  \Leftrightarrow \quad (-1)^n \lambda^n + tr(\mA)\lambda^{n-1} + \cdots + (-1)^n \det(\mA) \overset{\Delta}{=} P(\lambda) = 0.   \]
-	 The polynomial $P(\lambda)$ is called the characteristic equation of $\mA$. 
+	Let $\lambda$ denote an eigenvalue of the matrix $\mA\in \bbR^{n \times n}$.
+	The determinant of $\mA - \lambda \mI_n$ is a polynomial of $\lambda$, denoted $P(\lambda)$. We call the polynomial $P(\lambda)$ as the characteristic polynomial of $\mA$. Specifically,
+	\begin{equation}\label{def:charapoly}
+		 P(\lambda)  = (-1)^n \lambda^n +  (-1)^n  tr(\mA)\lambda^{n-1} + \cdots + (-1)^n \det(\mA). 
+	\end{equation}
 \end{defn}
@@
 Since the characteristic polynomial $P(\lambda)$ is of degree $n$, there are $n$ solutions to the equation $P(\lambda) = 0$.
-The solutions $\lambda_1,...,\lambda_n$ are $n$ eigenvalues of $\mA$, which may be complex and may not be distinct with each other.
-Note that the complex eigenvalues come in conjugate pairs, because $\overline{ P(\lambda)} = P(\overline{ \lambda})$.
+The solutions to $P(\lambda) = 0$, denoted $\lambda_1,...,\lambda_n$, are $n$ eigenvalues of $\mA$, which may be complex.
+If the complex eigenvalues exist, the complex eigenvalues come in conjugate pairs, because $P(\overline{ \lambda}) = \overline{ P(\lambda)} $ is also equal to 0, where $\overline{ \lambda}$ is the conjugate of $\lambda$.
+Note that $n$ eigenvalues are not necessarily distinct with each other.
 
@@
 \begin{thm}[Summation and production of eigenvalues]\label{thm:speigen}
-	For a square matrix $\mA \in \bbR^{n \times n}$, let $\lambda_1,...\lambda_n$ denote the $n$ eigenvalues of $\mA$. Then 
+	Let  $\mA \in \bbR^{n \times n}$ be a square matrix, and $\lambda_1,...\lambda_n$ be the $n$ eigenvalues of $\mA$. Then,
 	\begin{equation}
@@
 	\end{equation}   
-	where $\lambda_i, i \in [n]$ are $n$ eigenvalues of $\mA$.
+	where $\lambda_i$ are $n$ eigenvalues of $\mA$, for all $ i \in [n]$.
 By equation~\eqref{eq:poly}, the coefficient for $\lambda^{n-1}$ is equal to $(-1)^n\sum_{i=1}^n \lambda_i$.
-Compared with definition~\ref{def:charac}, we have $tr(\mA) = \sum_{i=1}^n \lambda_i$.
-Similarly, the constant term in equation~\eqref{eq:poly} is $(-1)^n \prod_{i=1}^n \lambda_1 \cdots \lambda_n $.
-Therefore, $\prod_{i=1}^n \lambda_i = \det(\mA)$.
+Compared with Definition~\ref{def:charac}, we have $tr(\mA) = \sum_{i=1}^n \lambda_i$.
+Similarly, the constant term in equation~\eqref{eq:poly} is $(-1)^n \prod_{i=1}^n \lambda_1 \cdots \lambda_n $ while the constant term in Defintion~\ref{def:charac} is $(-1)^n \det(\mA)$.
+Therefore, we have $\prod_{i=1}^n \lambda_i = \det(\mA)$.
 \end{proof}
@@
 \begin{defn}[\textit{Similar matrices}]\label{def:similar}
-	Consider a square matrix $\mA \in \bbR^{n \times n}$. For any invertible matrix $\mB \in \bbR^{n \times n}$, the matrix product $\mB \mA \mB^{-1}$ is called the similar matrix of $\mA$.
+	Let $\mA \in \bbR^{n \times n}$ be a square matrix, and $\mB \in \bbR^{n \times n}$ be a invertible matrix. The matrix product $\mB \mA \mB^{-1}$ is called the similar matrix of $\mA$.
 \end{defn}
@@
 \begin{thm}[Eigenvalues for similar matrices]\label{thm:similar}
-	Consider a sqaure matrix $\mA$. The similar matrices of $\mA$ share the same eigenvalues of $\mA$.
+	Let $\mA$ be a square matrix. Any similar matrices of $\mA$ share the same eigenvalues of $\mA$.
 \end{thm}
 \begin{proof}
-	Let $\mB \mA \mB^{-1}$ be the similar matrix of $\mA$, and $\lambda_1,...,\lambda_n$ be the eigenvalues of $\mA$. Then, for any $\lambda_i$, there is an eigenvector, such that, 
-	\[ \lambda x = \mA x \quad \Leftrightarrow \quad \lambda \mB x = \mB \mA \mB^{-1} \mB x.\]
-	Since $x$ is a nonzero vector, and $\mB$ is invertible, the vector $\mB x$ is also a nonzero vector. Therefore, $\lambda$ is an eigenvalue of $\mB \mA \mB^{-1}$ with associated eigenvector $\mB x$.
+	Let $\mB \mA \mB^{-1}$ be a similar matrix of $\mA$, and $\lambda_1,...,\lambda_n$ be the eigenvalues of $\mA$. Then, for all $\lambda_i$, there is an eigenvector $x_i$ such that
+	\[ \lambda_i x_i = \mA x_i \quad \Leftrightarrow \quad \lambda_i \mB x_i = \mB \mA \mB^{-1} \mB x_i.\]
+	Since $x_i$ is a nonzero vector, and $\mB$ is invertible, the vector $\mB x_i$ is also a nonzero vector.
+Therefore, $\lambda_i$ is an eigenvalue of $\mB \mA \mB^{-1}$ with associated eigenvector $\mB x_i$, for all $i \in [n]$.
 \end{proof}
@@
 \begin{thm}[Eigenvalues for powers of the matrix]
-	Consider a matrix $\mA$ with a eigenvalue $\lambda$ and the associated eigenvector $x$. For any polynomial $P$, 
+	Let $\mA$ be a matrix, $\lambda$  be a eigenvalue of $\mA$, and $x$ be the eigenvector associated with $\lambda$. For any polynomial $P$, we have
 	\[ P(\mA) x = P(\lambda) x. \]
@@
 \begin{proof}
-	For any integer $k \geq 0$ and constant $c \in \bbR$,
+	For all integer $k \geq 0$ and constant $c \in \bbR$, we have
 	\[ c\mA^k x = c\mA^{k-1} \mA x= c\lambda \mA^{k-1} x = \cdots = c\lambda^k x.\]
-	Therefore, for any polynomial $P$, $P(\mA)x = P(\lambda)x$.
+	Therefore, $P(\mA)x = P(\lambda)x$, for any polynomial $P$.
 \end{proof}
@@
 \begin{thm}[Eigenvalues for the inverse]
-Suppose $\mA \in \bbR^{n \times n}$ is an invertible matrix.
-Let $\lambda_1,...,\lambda_n$ denote the eigenvalue of $\mA$, and $x_1,...x_n$ denote the associated eigenvectors.
-Then,  $\lambda_1^{-1},...,\lambda_n^{-1}$ and $x_1,...x_n$ are the eigenvalues and associated eigenvectors of $\mA^{-1}$, respectively.
+Let $\mA \in \bbR^{n \times n}$ be an invertible matrix, $\lambda_1,...,\lambda_n$ be the eigenvalues of $\mA$, and $x_1,...x_n$ be the eigenvectors associated with $\lambda_1,...,\lambda_n$.
+Then,  $\lambda_1^{-1},...,\lambda_n^{-1}$  are the eigenvalues of $\mA^{-1}$, and $x_1,...x_n$ are eigenvectors associated with  $\lambda_1^{-1},...,\lambda_n^{-1}$.
 \end{thm}
@@
 \begin{proof}
-	Since the nullspace of $\mA$ contains only a zero vector, the eigenvaluest $\lambda_i >0,  \forall i \in [n]$.
-For any $\lambda_i$ and $x_i$, multiplying $x_i$ on both sides of the equation $\mA^{-1} \mA = \mI_n$, we have
+	Since the nullspace of $\mA$ contains only a zero vector, the eigenvalue $\lambda_i >0$, for all $  i \in [n]$.
+For all $\lambda_i$ and $x_i$, multiplying $x_i$ on both sides of the equation $\mA^{-1} \mA = \mI_n$ yields
 	\[ \mA^{-1} \mA x_i = x_i \quad \Rightarrow\quad  \mA^{-1} x_i = {\lambda_i}^{-1} x_i.  \]
-	Therefore, $\mA^{-1}$ share the same eigenvectors set with $\mA$, and the reciprocal of the eigenvalues of $\mA$ are the eigenvalues of $\mA^{-1}$.
+	Therefore, $\lambda_i^{-1}$ is the eigenvalue of $\mA^{-1}$, and $x_i$ is the eigenvector associated with $\lambda^{-1}_i$, for all $i \in [n]$.
 \end{proof}
@@
 \begin{thm}[Independence of eigenvectors]\label{thm:indepeigen} 
-Suppose the square matrix $\mA$ has at least two different eigenvalues. The eigenvectors associated with different eigenvalues are independent.
+Let $\mA$ be a square matrix with at least two different eigenvalues. The eigenvectors associated with different eigenvalues are independent.
 \end{thm}
@@
 	Suppose the square matrix $\mA \in \bbR^{n \times n}$ has $m$ different eigenvalues $\lambda_1,...,\lambda_m$ with associated eigenvectors  $v_1,...,v_m$.
-Let the linear combination $c_1 v_1 + ...
-+ c_m v_m = 0$.
+Consider a linear combination $c_1 v_1 + ...
++ c_m v_m = 0$, where $c_i \in \bbR$, for all $i \in [m]$.
 	Then we have
 	\begin{equation}\label{eq:indep1}
-		\mA(c_1 v_1 + ... + c_m v_m ) = c_1 \lambda_1 v_1 + ... + c_m \lambda_m v_m = 0;  
+		\mA(c_1 v_1 + ... + c_m v_m ) = c_1 \lambda_1 v_1 + ... + c_m \lambda_m v_m = 0,
 		\end{equation}
+		and
 		\begin{equation}\label{eq:indep2}
@@
 	\end{equation}
-	Subtracting the equations~\eqref{eq:indep1} from \eqref{eq:indep2}, 
+	Subtracting the equations~\eqref{eq:indep1} from \eqref{eq:indep2} yields
 	\begin{equation}\label{eq:indep3}
-		 c_2 (\lambda_1 - \lambda_2) v_2 + \cdots + c_m (\lambda_1 - \lambda_m) v_m  = 0
+		 c_2 (\lambda_1 - \lambda_2) v_2 + \cdots + c_m (\lambda_1 - \lambda_m) v_m  = 0.
 	\end{equation}
-	Repeating the steps from equation~\eqref{eq:indep1} to equation~\eqref{eq:indep3}, finally we have 
+	Consider the new linear combination $ c_2 (\lambda_1 - \lambda_2) v_2 + \cdots + c_m (\lambda_1 - \lambda_m) v_m = 0$ and repeat the steps from equation~\eqref{eq:indep1} to equation~\eqref{eq:indep3}.
+After $m$ iterations, we have
 	\[ c_m(\lambda_1 - \lambda_m)(\lambda_2 - \lambda_m)\cdots (\lambda_{m-1} - \lambda_m) v_m = 0.    \]
-	 Since $\lambda_1 ,..., \lambda_m$ are distinct, and $v_m$ is a nonzero vector, we have $c_m = 0$.
-Plugging $c_m = 0$ back to previous steps leads to that $c_1 = \cdots = c_m = 0$.
+	 Since $\lambda_1 ,..., \lambda_m$ are distinct, and $v_m$ is a nonzero vector, the coefficient $c_m = 0$.
+Plugging $c_m = 0$ back to previous steps, we have $c_1 = \cdots = c_m = 0$.
 Therefore, $v_1,...,v_m$ are independent.
@@
 \begin{thm}[Eigenvalues for triangular matrices]\label{thm:trianeigen}
-	For triangular matrices, the eigenvalues are the entries on the diagonal.
+	The eigenvalues for a triangular matrix are the entries on the diagonal.
 \end{thm}
@@
 \begin{proof}
-Suppose $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ is a triangular matrix. By the property of determinant~\ref{propdiag}, 
+Let $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ be a triangular matrix. By the property of determinant~\ref{propdiag}, we have
 \[ \det(\mA) = \prod_{i=1}^n a_{ii}  ; \quad \det(\mA - \lambda \mI_n) = \prod_{i=1}^n (a_{ii}-\lambda). \]
-To let $\det(\mA - \lambda \mI_n) = 0$, $\lambda = a_{ii}, \forall i\in[n]$. Therefore, the entries on the diagonal are the eigenvalues of a triangular matrix.
+To let $\det(\mA - \lambda \mI_n) = 0$, we have $\lambda = a_{ii}$, for all $ i\in[n]$. Therefore, the entries on the diagonal are the eigenvalues of a triangular matrix.
 \end{proof}
@@
 
+\begin{defn}[Antisymmetric matrices]
+	The matrix $\mA$ is an antisymmetric matrix if $\mA$ satisfies
+	\[\mA^T = - \mA. \] 
+\end{defn}
+
 \begin{thm}[Eigenvalues for symmetric and antisymmetric matrices]\label{thm:symmeign}
-	Symmetric square matrices  with all real entries have real eigenvalues.
-For antisymmetric matrices which satisfy $\mA^T = - \mA$, all eigenvalues are imaginary, i.e.\ $\lambda = b i$, where $b \in \bbR, i = \sqrt{-1}$.
+All the eigenvalues of a symmetric square matrix with only real entries are real.
+All the eigenvalues of an antisymmetric square matrix are imaginary; i.e., $\lambda = b i$, where $b \in \bbR, i = \sqrt{-1}$.
 \end{thm}
 \begin{proof}
-	First, we prove that symmetric square matrices with all real entries have real eigenvalues by contradiction.
+	First, we prove by contradiction that symmetric square matrices with all real entries have real eigenvalues.
 
-	Suppose a symmetric real matrix $\mA$ has a complex eigenvalue $\lambda$ with eigenvector $x$.
-Since $\mA$ is real, then the eigenvector $x$ is also complex.
+	Let  $\mA$  be a symmetric  matrix with only real entries.
+Suppose $\lambda$  is a complex eigenvalue of $\mA$, and $x$ is the eigenvector associated with $\lambda$.
+Since all the entries of $\mA$ are real, the eigenvector $x$ is also complex.
 Let $\overline \lambda$ and $\overline x$ denote the conjugate eigenvalue and eigenvector, respectively.
@@
 	\begin{equation}\label{eq:symm}
-		\mA x = \lambda x;\quad \mA \overline x = \overline \lambda \overline x.
+		\overline x^T \mA x = \overline x^T \lambda x,
 	\end{equation} 
-	Premultiply $\overline x^T$ on both sides of the first equation in \eqref{eq:symm} and premultiply $x^T$ on both sides of the second equation in \eqref{eq:symm}. By the symmetry of $\mA$,
+	and
+	\begin{equation}\label{eq:symm2}
+		x^T \mA \overline x = x^T \overline \lambda \overline x.
+	\end{equation}
+	 By the symmetry of $\mA$, we have
 	\[ \overline x^T \mA x  = x^T \mA^T \overline x = x^T \mA \overline x .\] 
-	Subtracting the two equations in \eqref{eq:symm}, we have
-	\[  0 = ( \overline x^T \mA x -  x^T \mA \overline x ) = (\lambda - \overline \lambda) x^T \overline x.  \]
-	By the fact that $x^T \overline x$ is a real number, the imaginary part of $\lambda$ is equal to 0, which contradicts that $\lambda$ is complex.
+	 Subtracting the equation~\eqref{eq:symm} from equation~\eqref{eq:symm2} yields	\[  0 = ( \overline x^T \mA x -  x^T \mA \overline x ) = (\lambda - \overline \lambda) x^T \overline x.  \]
+	Since that $x^T \overline x$ is a real number, the imaginary part of $\lambda$ is equal to 0, which contradicts the assumption that $\lambda$ is complex.
 
 	Next, we prove that the eigenvalues for antisymmetric matrices are imaginary. 
-	Suppose $\mA$ is an antisymmetric matrix with a complex eigenvalue $\lambda$ and complex eigenvector $x$.  By similar procedures for symmetric matrices, we have
+	Let $\mA$ be an antisymmetric matrix,  $\lambda$ be a complex eigenvalue of $\mA$, and $x$ be the complex eigenvector associated with $\lambda$. By the antisymmetry of $\mA$, we have
 	\[ \overline x^T \mA x  = x^T \mA^T \overline x  =  -x^T \mA \overline x .  \] 
@@
 	\[  0 = ( \overline x^T \mA x +  x^T \mA \overline x ) = (\lambda + \overline \lambda) x^T \overline x.   \]
-	Since $x^T \overline x$ is a real number, the real part of $\lambda$ is equal to 0. Therefore, eigenvalues for antisymmetric matrices are imaginary.
+	Since that $x^T \overline x$ is a real number, the real part of $\lambda$ is equal to 0. Therefore, the eigenvalues for an antisymmetric matrix are imaginary.
 \end{proof}
@@
 \begin{proof}
-	Suppose $\lambda_1$ and $\lambda_2$ are two different eigenvalues of $\mA$ with associated eigenvector $x_1, x_2$, respectively. By the symmetry of $\mA$, we have
-	\[ x_2^T \mA x_1 = x_1^T \mA^T x_2  = x_1^T \mA x_2,\]
-	then 
+	Let $\lambda_1$ and $\lambda_2$ be two different eigenvalues of $\mA$, and  $x_1, x_2$ are two eigenvectors associated with $\lambda_1$ and $\lambda_2$, respectively.
+By the symmetry of $\mA$, we have
+	\[ x_2^T \mA x_1 = x_1^T \mA^T x_2  = x_1^T \mA x_2.\]
+	Then,
 	\[  0 = ( x_2^T \mA x_1 -  x_1^T \mA x_2 ) = (\lambda_1 - \lambda_2) x_2^T x_1. \]
@@
 
-\begin{thm}[Eigenvectors for repeated eigenvalues of symmetric matrices]\label{thm:symmrepeat} Consider a symmetric matrix $\mA \in \bbR^{n \times n}$.
-If $\lambda_0$ is a repeated eigenvalue of $\mA$ with multiplicity $2 \leq m \leq n$, then there exist $m$ orthonormal eigenvectors corresponding to $\lambda_0$.
+\begin{thm}[Eigenvectors for repeated eigenvalues of symmetric matrices]\label{thm:symmrepeat} Let  $\mA \in \bbR^{n \times n}$ be a symmetric matrix.
+Suppose $\lambda_0$ is a repeated eigenvalue of $\mA$ with multiplicity $m$, where $2 \leq m \leq n$.
+There exist $m$ orthonormal eigenvectors associated with $\lambda_0$.
 \end{thm}
@@
 \begin{proof}
-	First, there is at least one nonzero eigenvector $x_0$ for $\lambda_0$.
-For any $x_0 \in \bbR^{n}$, there are $n-1$ additional orthogonal vectors $y_1,...,y_{n-1}$, such that $y_j \perp x_0, \forall j \in [(n-1)]$ and $(x_0, y_1,...,y_{n-1})$ forms the basis of $\bbR^n$.
+	First, let $x_0$ be a nonzero eigenvector associated with $\lambda_0$.
+For any $x_0 \in \bbR^{n}$, there are $n-1$ additional orthogonal vectors $y_1,...,y_{n-1}$ such that $y_j \perp x_0$, for all $j \in [(n-1)]$,  and $(x_0, y_1,...,y_{n-1})$ forms the basis of $\bbR^n$.
 Let $\mY = [y_1,...y_{n-1}]$ and $\mX = [x_0,Y]$.
-Since $\mA$ is a symmetric matrix, $x_0^T \mA = \lambda_0 x_0^T $.
-We have
+Since $\mA$ is a symmetric matrix, we have $x_0^T \mA = \lambda_0 x_0^T $.
+Consider the following matrix product,
 	\begin{align}\label{eq:symmrepeat}
@@
 	Note that $\mX$ is an orthogonal matrix which satisfies $\mX^T = \mX^{-1}$.
-By Theorem~\ref{thm:similar}, $\mX^T \mA \mX$ is a similar matrix of $\mA$, and $\mX^T \mA \mX$ share the same eigenvalues of $\mA$.
-Then, $\lambda_0$ is also a repeated eigenvalue of $\mX^T \mA \mX$.
-The determinant of $\mX^T \mA \mX - \lambda \mI_n$ is
-	\[  \det(\mX^T \mA \mX - \lambda \mI_n) = (\lambda_0 - \lambda) \det(\mY^T \mA \mY - \lambda \mI_{n-1}). \]
-	Since $\lambda_0$ has multiplicity $m \geq 2$, then $\det(\mY^T \mA \mY - \lambda_0 \mI_{n-1}) = 0$.
+By Theorem~\ref{thm:similar}, $\mX^T \mA \mX$ is a similar matrix of $\mA$.
+Hence, $\lambda_0$ is also a repeated eigenvalue of $\mX^T \mA \mX$.
+The characteristic  polynomial of $\mX^T \mA \mX - \lambda \mI_n$ is
+	\[  P(\lambda) = \det(\mX^T \mA \mX - \lambda \mI_n) = (\lambda_0 - \lambda) \det(\mY^T \mA \mY - \lambda \mI_{n-1}). \]
+	Since $\lambda_0$ has multiplicity $m \geq 2$, the term $\det(\mY^T \mA \mY - \lambda_0 \mI_{n-1}) = 0$.
 Therefore, the dimension of the nullspace of $\mX^T \mA \mX - \lambda_0 \mI_n$ is larger than 2.
@@
 
-	Replace $x_0$ by two orthogonal vectors $x_1 = \mX v_1, x_2 = \mX v_2$ and repeat the steps above. Finally, we will have $m$ orthogonal eigenvectors of $\mA$ associated with $\lambda_0$.
+	Replace $x_0$ by two orthogonal vectors $x_1 = \mX v_1, x_2 = \mX v_2$ and repeat the above steps. Finally, we will have $m$ orthogonal eigenvectors of $\mA$ associated with $\lambda_0$.
 \end{proof}
@@
 \begin{defn}[Diagonalization]
-	Suppose the square matrix $\mA \in \bbR^{n \times n}$ has $n$ independent eigenvectors. Then the matrix $\mA$ can be diagonalized as
-	\[ \mA = \mS^{-1} \Lambda \mS, \]
-	where $\mS \in \bbR{n \times n}$ is a matrix whose columns are $n$ independent eigenvectors of $\mA$ and $\Lambda \in \bbR^{n \times n}$ is a diagonal matrix whose entries are $n$ eigenvalues of $\mA$.
+	Let  $\mA \in \bbR^{n \times n}$ be a square matrix with $n$ independent eigenvectors. We diagonalize the matrix $\mA$ as
+	\begin{equation}\label{eq:defdiag}
+		\mA = \mS^{-1} \Lambda \mS,
+	\end{equation}
+	where $\Lambda \in \bbR^{n \times n}$ is a diagonal matrix whose entries are $n$ eigenvalues of $\mA$ and  $\mS \in \bbR^{n \times n}$ is a matrix whose columns are independent eigenvectors of $\mA$ associated with eigenvalues in $\Lambda$.
 \end{defn}
@@
 \begin{thm}[Diagonalization of symmetric matrices]
-Suppose $\mA$ is a symmetric matrix, then $\mA$ can be diagonalized.
+Let $\mA$ be a symmetric matrix. The matrix $\mA$ can be diagonalized.
 \end{thm}
@@
 	By Theorems~\ref{thm:symmortho} and Theorem~\ref{thm:symmrepeat}, for any symmetric matrix $\mA \in \bbR^{n \times n}$, $\mA$ have $n$ orthogonal eigenvectors, even though there may have repeated eigenvalues.
+Therefore, the matrix $\mA$ can be diagonalized as equation~\eqref{eq:defdiag}.
 \end{proof}
@@
 \begin{thm}[Descent of matrix powers]
-	If a square matrix $\mA \in \bbR^{n \times n}$ has $n$ independent eigenvectors with eigenvalues $\lambda_1,...,\lambda_n$, then $\mA^k \rightarrow 0$ as $k \rightarrow +\infty$ if and only if all $|\lambda_i| < 1$.
+	Let $\mA \in \bbR^{n \times n}$ be  a square matrix with eigenvalues $\lambda_1,...,\lambda_n$, and  $n$ independent eigenvectors associated with the eigenvalues.
+The power of matrix $\mA^k \rightarrow 0$ as $k \rightarrow +\infty$ if and only if $|\lambda_i| < 1$, for all $i \in [n]$.
 \end{thm}
@@
 \begin{proof}
-	We diagonalize the matrix $\mA$ as $\mA =\mS^{-1} \Lambda \mS$. For any positive integer $k$, $\mA^k = \mS^{-1} \Lambda^{k} \mS$. Then 
+	We diagonalize the matrix $\mA$ as $\mA =\mS^{-1} \Lambda \mS$. For all integer $k\geq 0$, we have $\mA^k = \mS^{-1} \Lambda^{k} \mS$. Then,
 	\[ \lim_{k \rightarrow +\infty} \mA^k = \lim_{k \rightarrow +\infty} \mS^{-1} \Lambda^{k} \mS  =  \lim_{k \rightarrow +\infty} \mS^{-1} \begin{bmatrix}
@@
 		0 & \cdots & |\lambda_n|^k
-	\end{bmatrix} \mS  = 0 \quad \Leftrightarrow \quad  |\lambda_i| < 1, \forall i \in [n]. \]
+	\end{bmatrix} \mS  = 0 \quad \Leftrightarrow \quad  |\lambda_i| < 1, \text{ for all } i \in [n]. \]
 \end{proof}

### 2020-07-16 skills/math/tail_bound/tailbound_solution.tex
update

@@
 	\begin{equation}\label{eq:thmmarkov}
-		\bbP(X\geq t) \leq \frac{E[X]}{t},\quad \forall \ t > 0.
+		\bbP(X\geq t) \leq \frac{E[X]}{t},\quad \text{ for all } \ t > 0.
 	\end{equation}  
@@
 	\begin{equation}\label{eq:cheby}
-		\bbP(|X - \mu |\geq t) \leq \frac{var(X)}{t^2},\quad \forall \ t > 0.
+		\bbP(|X - \mu |\geq t) \leq \frac{var(X)}{t^2},\quad \text{ for all } \ t > 0.
 	\end{equation}  
@@
 	\begin{equation}\label{eq:poly}
-		\bbP(|X - \mu |\geq t) \leq \frac{\bbE[|X-\mu|^k]}{t^k},\quad \forall \ t > 0.
+		\bbP(|X - \mu |\geq t) \leq \frac{\bbE[|X-\mu|^k]}{t^k},\quad \text{ for all } \ t > 0.
 	\end{equation}  
@@
 	\begin{equation}\label{eq:chernoff1}
-		\bbP((X - \mu) \geq t) \leq \frac{\bbE[e^{\lambda (X-\mu)}]}{e^{\lambda t}}, \quad \forall \lambda\in (-b,b).
+		\bbP((X - \mu) \geq t) \leq \frac{\bbE[e^{\lambda (X-\mu)}]}{e^{\lambda t}}, \quad \text{ for all } \lambda\in (-b,b).
 	\end{equation}
@@
 
+\begin{thm}[Moment of sub-Gaussian variable]\label{thm:momsubg}
+Let $X \sim sG(\sigma^2)$. For all integer $k \geq 1$, we have
+\begin{equation}\label{eq:thmmomsubg}
+	\bbE[|X|^k] \leq  k 2^{k/2} \sigma^k \Gamma(\frac{k}{2} ),
+\end{equation} 
+where the Gamma function is defined as $\Gamma(x) = \int_{0}^{+\infty} t^{x -1} e^{-t} dt$. 
+\end{thm}
+
+\begin{thm}[One-sided Bernstein’s inequality]\label{thm:oneber}
+ Let $X$ be a random variable. If $X \leq b$ almost surely, then 
+\begin{equation}\label{eq:thmoneber1}
+	\bbE[e^{\lambda (X-\bbE[X])}] \leq exp\left\{ \frac{\lambda^2 \bbE[X^2]/2}{1 - b\lambda /3}  \right\}, \quad \text{for all } \lambda \in [0, 3/b).
+\end{equation}
+Consequently, suppose there are $n$ independent variables $X_i \leq b$ almost surely, for all $i \in [n]$. We have
+\begin{equation}\label{eq:thmoneber2}
+	\bbP[\sum_{i = 1}^n (X_i - \bbE[X_i] ) \geq n\delta ] \leq exp\left\{ - \frac{n\delta^2}{\sum_{i = 1}^n \bbE[X_i^2] /n + b \delta /3} \right\},\quad \text{ for all } \delta \geq 0.
+\end{equation}  
+Particularly, suppose there are $n$ independent non-negative variables $Y_i \geq 0$, for all $i \in [n]$. The equation~\eqref{eq:thmoneber2} becomes
+\begin{equation}\label{eq:thmonber0}
+	\bbP[\sum_{i = 1}^n (Y_i - \bbE[Y_i] ) \leq n\delta ] \leq exp\left\{ - \frac{n\delta^2}{\sum_{i = 1}^n \bbE[Y_i^2] /n} \right\},\quad \text{ for all } \delta \geq 0.
+\end{equation} 
+\end{thm}
+
 \section{Exercises}
@@
 \begin{enumerate}
+
+\item[(a)] For a given constant $t >0$, we define a variable $Y_{t} = X - t\mI_{[X\geq t]}$, where $\mI$ is the indicator function.
+Noted that $Y_{t}$ is a nonnegative variable, the Markov's inequality follows by taking expectation to $Y$,
+\begin{equation}
+	\bbE[Y_t] = \bbE[X] - t \bbP[X \geq t] \geq 0.
+\end{equation}
+Therefore, Markov's inequality meets the equality if and only if the expectation $\bbE[Y_t] = 0$.
+Since $Y_t$ is nonnegative, we have $\bbP(Y_t = 0) = 1$.
+Note that $Y_t = 0$ if and only if $X = 0$ or $X =t$.
+
+Hence, for the given constant $t >0$, the nonnegative variable $X$ with distribution $\bbP(X \in \{ 0,t\}) = 1$ attains the equality of Markov's inequality.
+
+\item[(b)]  Chebyshev's inequality follows by applying Markov's inequality to the non-negative random variable $Z = {(X - \bbE[X])^2}$.
+Similarly as part (a), given a constant $t > 0$, the variable $Z = {(X - \bbE[X])^2}$ with distribution $\bbP(Z \in \{ 0,t^2 \}) = 1$ attains the equality of the Markov's inequality for $Z$ and the Chebyshev's inequality for $X$.
+By transformation, the distribution of $X$ satisfies the followings formula,
+\begin{align}
+\bbP(X = x) = 
+	\begin{cases}
+		p & \text{if } x = c,\\
+		\frac{1-p}{2} & \text{if } x = c-t \text{ or } x = c+t, \\
+		0 & \text{ otherwise },
+	\end{cases}
+	\end{align}
+where $c \in \bbR$ is a constant and $p \in [0,1]$.
+
+\iffalse
 	\item[(a)] Recall the proof of Markov's inequality. For any $t >0$,
@@
 Therefore, the variable $Z$ attains the equality of Chebyshev's inequality.
+
+\fi
+
 \end{enumerate}
@@
 \begin{rmk}[Tightness of Markov's inequality]
-	Only few variables can attain the equalities in Markov's and Chebyshev's inequalities. In research, we should be careful if there is a concentration bound that is tighter than Markov's inequality.
+	Only a few variables can attain the equalities in Markov's and Chebyshev's inequalities. In research, we should pay attention to the concentration bounds tighter than Markov's inequality.
 \end{rmk}
@@
 
-	We write the upper tail probability of the standard normal variable as:
+	We write the upper tail probability of the standard normal variable as
 	\begin{equation}\label{eq:e22pro1}
@@
 
-	Combing with the equations~\eqref{eq:e22pro1}, \eqref{eq:e22pro2}, and  \eqref{eq:e22pro3}, we have $\bbP(Z \geq z) \leq \phi(z)(\frac{1}{z}  - \frac{1}{z^3} + \frac{3}{z^5} ).$
+	Combing equations~\eqref{eq:e22pro1}, \eqref{eq:e22pro2}, and  \eqref{eq:e22pro3}, we have $\bbP(Z \geq z) \leq \phi(z)(\frac{1}{z}  - \frac{1}{z^3} + \frac{3}{z^5} ).$
 \end{proof}
@@
 \begin{lem}[Polynomial bound and Chernoff bound]\label{lem:momcher}
-	Consider a non-negative variable $X \geq 0$.
-Suppose that the moment generating function of $X$, $\varphi_{X}(\lambda)$, exists in the neighborhood of $\lambda = 0$.
-Given some $\delta > 0$, for any integer $ k \in \mathbb{Z_+}$, we have
+	Let $X \geq 0$ be a non-negative variable.
+Suppose that the moment generating function of $X$, denoted $\varphi_{X}(\lambda)$, exists in the neighborhood of $\lambda = 0$.
+Given some $\delta > 0$, we have
 	\begin{equation}\label{eq:lempoly}
@@
 	\begin{equation}\label{eq:power}
-		 e^{\lambda X} = \sum_{k = 0}^{+\infty } \frac{X^k \lambda^k }{k!}, \quad \forall \lambda \in \bbR.
+		 e^{\lambda X} = \sum_{k = 0}^{+\infty } \frac{X^k \lambda^k }{k!}, \quad \text{ for all } \lambda \in \bbR.
 	\end{equation}
 	Since the moment generating function $\varphi_X(\lambda)$ exists in the neighborhood of $\lambda = 0$, there exists a constant $b > 0$ such that 
-		\[   \bbE [e^{\lambda X} ]= \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} < +\infty,\quad \forall \lambda \in (0,b). \]
-	Therefore, the moment  $\bbE[|X|^k]$ exists for any $ k \in \mathbb{Z_+}$. Noted that  $\bbE[|X|^k] \geq 0$, we have
-	\begin{equation}\label{eq:single}
-		 \bbE[|X|^k] \leq \sum_{k = 0}^{+\infty}  \bbE[|X|^k], \quad \forall k \in \mathbb{Z_+}. 
+		\[   \bbE [e^{\lambda X} ]= \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} < +\infty,\quad \text{ for all } \lambda \in (0,b). \]
+	Therefore, the moment  $\bbE[|X|^k]$ exists for all $ k \in \mathbb{Z_+}$. Applying power series~\eqref{eq:power} to the right hand side of equation~\eqref{eq:lempoly} yields
+	\begin{equation}\label{eq:powertopoly}
+		\inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}} = \frac{ \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} }{ \sum_{k = 0}^{+\infty } \frac{\lambda^k \delta^k }{k!}  }.	\end{equation}
+	By Cauchy's third inequality, we have
+	\begin{equation}\label{eq:cauchy3}
+		\frac{ \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} }{ \sum_{k = 0}^{+\infty } \frac{\lambda^k \delta^k }{k!}  } \geq \inf_{k \in \mathbb{Z}_{+} } \frac{\bbE[|X|^k]}{\delta^k}  
 	\end{equation}
-	Applying power series~\eqref{eq:power} to the right hand side of equation~\eqref{eq:lempoly} yields
-	\begin{equation}\label{eq:powertopoly}
-		\inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}} = \sum_{k = 0}^{+\infty } \frac{\frac{ \bbE[|X|^k]  \lambda^k }{k!} }{\frac{\lambda^k \delta^k }{k!}  } =\sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k] }{\delta^k}.
-	\end{equation}
-	Combing with equation~\eqref{eq:single}, we have
-	\[  \inf_{k \in \mathbb{Z_+} } \frac{\bbE[|X|^k]}{\delta^k} \leq \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k] }{\delta^k}  = \inf_{\lambda > 0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}}. \]
+	Therefore, the equation~\eqref{eq:lempoly} follows by combining the equation~\eqref{eq:powertopoly} with equation~\eqref{eq:cauchy3}.	
 \end{proof}
@@
 \begin{rmk}
-	Applying different functions $g(X)$ to the Markov's inequality leads to distinct bounds for the tail probability of variable $X$.
+	Applying different functions $g(X)$ to the Markov's inequality leads to different bounds for the tail probability of variable $X$.
 Equation~\eqref{eq:lempoly} implies that optimized polynomial bound is tighter than the Chernoff bound, provided that the moment generating function of $X$ exsits.
-Actually, the optimization on polynomial bound can be released.
-Then, any polynomial bound is better than Chernoff bound.
 \end{rmk}
@@
 \subsection{Exercise 2.4}\label{pro:boundvar}
-In Exercise~\ref{pro:boundvar}, we prove the theorem~\ref{thm:boundvar},  the Hoeffding bound for bounded variable.
+In Exercise~\ref{pro:boundvar}, we prove the Theorem~\ref{thm:boundvar},  the Hoeffding bound for a bounded variable.
 
@@
 
-Suppose $X$ is a bounded random variable that $X \in [a,b]$ almost surely. Let $\mu = \bbE[X]$. Define the function 
-\[ g(\lambda) = \log \bbE[e^{\lambda X }], \quad  \forall \lambda \in \bbR. \]
+Suppose $X$ is a bounded random variable, where $X \in [a,b]$ almost surely, and $a 
+\leq b \in \bbR$ are two constants. Let $\mu = \bbE[X]$. Define the function 
+\[ g(\lambda) = \log \bbE[e^{\lambda X }], \quad  \text{ for all } \lambda \in \bbR. \]
 Applying Taylor Expansion to $g(\lambda)$ at 0, we have
 \begin{equation}\label{eq:taylor}
-	g(\lambda) =  g(0) + g'(0) \lambda  + \frac{g^{''}(\lambda_0) }{2} \lambda^2, \quad \text{where } \lambda_0 = t \lambda, \text{ for some } t \in [0,1]. 
+	g(\lambda) =  g(0) + g'(0) \lambda  + \frac{g''(\lambda_0) }{2} \lambda^2, \quad \text{where } \lambda_0 = t \lambda, \text{ for some } t \in [0,1]. 
 \end{equation}
 
-In equation~\eqref{eq:taylor}, the term $g(0) =\log \bbE[e^{0}] = 0$. By power series \eqref{eq:power}, we calculate the first derivative $g'(\lambda)$ as following.
+In equation~\eqref{eq:taylor}, the term $g(0) =\log \bbE[e^{0}] = 0$. By power series \eqref{eq:power}, we calculate the first derivative $g'(\lambda)$ as following,
 \begin{align}
@@
 	&= {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+1)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]}  \label{eq:firstde} \\
-	&= \frac{\bbE[X e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]}
+	&= \frac{\bbE[X e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]}.
 \end{align}
-Therefore, $g'(0) = \bbE[X] = \mu$. Taking the derivative to equation~\eqref{eq:firstde}, we calculate the second-order derivative $g^{''}(\lambda)$ as following.
+Therefore, $g'(0) = \bbE[X] = \mu$. Taking the derivative to equation~\eqref{eq:firstde}, we calculate the second-order derivative $g''(\lambda)$ as following,
 \begin{align}
-	g^{''}(\lambda)
+	g''(\lambda)
 	&= {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+2)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]} - \left( {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+1)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]}  \right)^2\\
-	&= \frac{\bbE[X^2 e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]} - \left( \frac{\bbE[X e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]} \right)^2\\
+	&= \frac{\bbE[X^2 e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]} - \left( \frac{\bbE[X e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]} \right)^2.\\
 	%& \overset{\Delta}{=} \bbE_{\lambda} [X^2] - (\bbE_{\lambda} [X])^2
 \end{align}
-The second-order derivative $g''(\lambda)$ can be interpreted as the variance of $X$ with a re-weighted distribution $d P' = {e^{\lambda X}}/{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$.
-Taking integral to $dP'$, we have
+We interpret the second-order derivative $g''(\lambda)$ interpreted as the variance of $X$ with the re-weighted distribution $d P' = {e^{\lambda X}}/{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$.
+Taking integral of 1 with respect to $dP'$, we have
 \begin{equation}
@@
 \end{equation} 
-which implies that the function $P'$ is indeed a distribution. Under any distribution, there is always an upper bound for the variance of random variable $X$:
-\[ var(X) = var(X - \frac{a+b}{2}) \leq \bbE[(X - \frac{a+b}{2} )^2] \leq \frac{(b-a)^2}{4}. \]
-Hence, the second-order derivative $g''(\lambda) \leq \frac{(b-a)^2}{4}$. We plug the results of $g'$ and $g^{''}$ into the equation~\eqref{eq:taylor}. Then,
+which implies that the function $P'$ is indeed a distribution. Under all possible re-weighted distributions, the variance of $X$ is upper bounded as following,
+\[ var(X) = var(X - \frac{a+b}{2}) \leq \bbE[(X - \frac{a+b}{2} )^2] \leq \frac{(b-a)^2}{4}, \]
+where the term $\frac{(b-a)^2}{4}$ follows by letting $X$ distribute only on the boundary $a$ and $b$.
+Hence, the second-order derivative $g''(\lambda) \leq \frac{(b-a)^2}{4}$.
+We plug the results of $g'$ and $g''$ into the equation~\eqref{eq:taylor}.
+Then,
 \begin{equation}\label{eq:e24}
-	g(\lambda) =  g(0) + g'(0) \lambda  + \frac{g^{''}(\lambda_0) }{2} \lambda^2 \leq 0 + \lambda \mu+ \frac{(b-a)^2}{8} \lambda^2. 
+	g(\lambda) =  g(0) + g'(0) \lambda  + \frac{g''(\lambda_0) }{2} \lambda^2 \leq 0 + \lambda \mu+ \frac{(b-a)^2}{8} \lambda^2. 
 \end{equation}
@@
 \begin{rmk}
-	For any bounded random variable $X$ with domain $[a,b]$, $X$ is a sub-gaussian variable with parameter at most $\sigma^2 = (b-a)^2/4$.
-Any propositions for sub-Gaussian variables can be applied to bounded variables.
+	For any bounded random variable $X$ supported on $[a,b]$, $X$ is a sub-gaussian variable with parameter at most $\sigma^2 = (b-a)^2/4$.
+All the properties for sub-Gaussian variables apply to the bounded variables.
 \end{rmk}
@@
 \subsection{Exercise 2.5}
-\begin{lem}[Sub-Gaussian bounds and means/variance] Consider a random variable $X$ such that
+\begin{lem}[Sub-Gaussian bounds and means/variance] Let $X$ be a random variable  such that
 \begin{equation}\label{eq:subg}
-	 \bbE[e^{\lambda X}] \leq e^{\frac{\lambda^2 \sigma^2}{2} + \mu \lambda}, \quad \forall \lambda \in \bbR. 
+	 \bbE[e^{\lambda X}] \leq e^{\frac{\lambda^2 \sigma^2}{2} + \mu \lambda}, \quad \text{ for all } \lambda \in \bbR. 
 \end{equation}
@@
 
-\begin{proof} By equation~\eqref{eq:subg}, the moment generating function of $X$, $\varphi_X(\lambda)$ , exists in the neighborhood of $\lambda = 0$.
+\begin{proof} By equation~\eqref{eq:subg}, the moment generating function of $X$, denoted $\varphi_X(\lambda)$ , exists in the neighborhood of $\lambda = 0$.
 Hence, the mean and variance of $X$ exist.
-Applying power series on both sides of equation~\eqref{eq:subg} yields
+For all $\lambda$ in the neighborhood of $\lambda = 0$, applying power series on both sides of equation~\eqref{eq:subg} yields
 \begin{equation}\label{eq:taylor2}
@@
 \end{equation}
-Dividing by $\lambda >0$ on both sides of equation~\eqref{eq:taylor2} and letting $\lambda \rightarrow 0^{+}$, we have $\bbE(X) \leq \mu$ ; dividing by $\lambda <0$ on both sides of equation~\eqref{eq:taylor2} and letting $\lambda \rightarrow 0^{-}$, we have $\bbE(X) \geq \mu$.
+Dividing by $\lambda >0$ on both sides of equation~\eqref{eq:taylor2} and letting $\lambda \rightarrow 0^{+}$, we have $\bbE(X) \leq \mu$.
+Dividing by $\lambda <0$ on both sides of equation~\eqref{eq:taylor2} and letting $\lambda \rightarrow 0^{-}$, we have $\bbE(X) \geq \mu$.
 Therefore, the mean $\bbE[X] = \mu$.
@@
 
-\paragraph{Solution:} The statement that $var(X) = \sigma_{min}^2$ is not always true.
+\paragraph{Solution:} The statement that $var(X) = \sigma_{min}^2$ is not necessarily true.
 Recall the function $g(\lambda)$ in exercise~\ref{pro:boundvar}.
@@
 \begin{equation}\label{eq:question}
-	g^{''}(\lambda) \leq  \sigma^2, \quad \forall \lambda \in \bbR,
+	g''(\lambda) \leq  \sigma^2, \quad \text{ for all } \lambda \in \bbR,
 \end{equation}
-where $g^{''} (\lambda)$ is the variance of $X$ with with a re-weighted distribution $d P' = {e^{\lambda X}}/{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$ and $g^{''} (0) = var(X)$.
-Therefore, $\max_{\lambda}g^{''}(\lambda) = \sigma_{min}^2$.
-To satisfy the equality $var(X) = \sigma_{min}^2$, we need to show that $\max_{\lambda}g^{''}(\lambda) = g^{''}(0)$ holds for $X$.
+where $g'' (\lambda)$ is the variance of $X$ with with the re-weighted distribution $d P' = {e^{\lambda X}}/{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$ and $g'' (0) = var(X)$.
+Therefore, $\max_{\lambda}g''(\lambda) = \sigma_{min}^2$.
+To let the equality $var(X) = \sigma_{min}^2$ hold, we need to show that $\max_{\lambda}g''(\lambda) = g''(0)$ holds for $X$.
 
 \vspace{0.2cm}
-However, the statement $\max_{\lambda}g^{''}(\lambda) = g^{''}(0)$ is not always true.  A counter example is below. 
-Consider a random variable $X \sim Ber(1/3)$. The variance of $X$ is $var(X) = 2/9$. Let $\lambda = 1$. The re-weighted distribution $dP'$ is
-\[ P'(X = 0) = \frac{2}{3\bbE[ e^{X} ]};\quad  P'(X = 1) = \frac{e}{3\bbE[ e^{X} ]}, \quad \text{where } \bbE[ e^{X} ] = \frac{2}{3} + \frac{e}{3}.  \]
-The variance of $X$ with $dP'$ is ${2}/{3\bbE[ e^{X} ]} \times  {e}/{3\bbE[ e^{X} ]} = 0.2442 > 2/9$. Therefore, for this variable $X$, $g^{''}(0) <  g^{''}(1) \leq \max_{\lambda}g^{''}(\lambda)$. 
+However, the statement $\max_{\lambda}g''(\lambda) = g''(0)$ is not always true.  A counter example is below. 
+Consider a random variable $Y \sim Ber(1/3)$. The variance of $Y$ is $var(Y) = 2/9$. Let $\lambda = 1$. The re-weighted distribution $dP'$ is
+\[ P'(Y = 0) = \frac{2}{3\bbE[ e^{Y} ]}\quad  \text{and}\quad  P'(Y = 1) = \frac{e}{3 \bbE[ e^{ Y } ]}, \quad \text{where } \bbE[ e^{Y} ] = \frac{2}{3} + \frac{e}{3}.  \]
+The variance of $Y$ with $dP'$ is ${2}/{3\bbE[ e^{Y} ]} \times  {e}/{3\bbE[ e^{Y} ]} = 0.2442 > 2/9$.
+Therefore, for this variable $Y$, we have $var(Y) <  g''(1) \leq \max_{\lambda}g''(\lambda) = \sigma^2_{min}$.
 \begin{rmk}
@@
 
-\iffalse
+\subsection{Exercise 2.6}
 
-\subsection{Exercise 2.6}
-\begin{lem}[Lower bounds on squared sub-Gaussians] Let $\{ X_i \}_{i=1}^n$ be i.i.d.\ sequence of zero-mean sub-Gaussian variables with parameter $\sigma$.
+\begin{lem}[Lower bounds on squared sub-Gaussians]\label{lem:lowersquare} Let $\{ X_i \}_{i=1}^n$ be an i.i.d.\ sequence of zero-mean sub-Gaussian variables with parameter $\sigma$.
 The normalized summation $Z_n = \frac{1}{n} \sum_{i=1}^n X_i^2$ satisfies
-\begin{equation}\label{lem:lowersquare}
-	\bbP[Z_n - \bbE[Z_n] \leq \sigma^2 \delta ] \leq e^{-n\delta^2 /16}, \quad \forall \delta \geq 0.
+\begin{equation}\label{eq:lowersquare}
+	\bbP[Z_n - \bbE[Z_n] \leq \sigma^2 \delta ] \leq e^{-n\delta^2 /16}, \quad \text{ for all } \delta \geq 0.
 \end{equation}
-	The equation~\eqref{lem:lowersquare} implies that the lower tail of the summation of squared sub-Gaussian variables behave in a sub-Gaussian way.
+	The equation~\eqref{eq:lowersquare} implies that the lower tail of the summation of squared sub-Gaussian variables behave in a sub-Gaussian way.
 \end{lem}
@@
 \begin{proof}
-	Given a sub-Gaussian vector $(X_1,...,X_n)$, the moment generating function for $Z_n = \frac{1}{n} \sum_i X_i^2$, $\varphi_{Z_n}(\lambda)$, exists in the neighborhood of $ \lambda = 0$.
-By the Chernoff bound~\eqref{eq:chernoff1}, there exists a constant $b > 0$ such that
+	Since $X_i^2$ are i.i.d.\ nonnegative variables, we apply the equation~\eqref{eq:thmonber0} in Theorem~\ref{thm:oneber} to the variables $X_i^2$, where $i \in [n]$. Then,
+	\begin{equation}\label{eq:e26a}
+		\bbP[\sum_{i=1}^n (X_i^2 -\bbE[X_i^2]) \leq n \sigma^2 \delta] \leq exp\left\{ - \frac{n\delta^2 \sigma^4}{ \bbE[X_1^4]} \right\},  \quad \text{for all }\delta \geq 0.
+	\end{equation}
+	By the equation~\eqref{eq:thmmomsubg} in Theorem~\ref{thm:momsubg}, we have
+	\begin{equation}\label{eq:e26x4}
+		\bbE[X_1^4] \leq 16 \sigma^4. 
+	\end{equation}
+	Combing equations~\eqref{eq:e26a}, \eqref{eq:e26x4}, and the definition of $Z_n$, we have
 	\begin{equation}
-		\bbP[Z_n - \bbE[Z_n] \geq t]  \leq \inf_{\lambda \in (-b,b)} \frac{\bbE[e^{\lambda (Z_n - \bbE[Z_n])}]}{e^{\lambda t}}.
+		\bbP[Z_n - \bbE[Z_n] \leq \sigma^2 \delta ]  \leq exp\left\{ - \frac{n\delta^2}{16} \right\},  \quad \text{for all }\delta \geq 0.
 	\end{equation}
-	Hence, analysis to the expectation $\bbE[e^{\lambda (Z_n - \bbE[Z_n])}]$ is the key of the proof. Since $X_i$ are i.i.d.\, we re-write the expectation as following
-	 \[\bbE[e^{\lambda (Z_n - \bbE[Z_n])}] = (\bbE[ e^{ \frac{\lambda}{n} (X_1^2 - \bbE[X_1]^2)} ])^n. \]
-	Let $Z = X_1^2$. Next, we analyze the expectation $\bbE[e^{\lambda (Z - \bbE[Z])}]$.
 \end{proof}
 
-\fi
+\begin{rmk}
+	Equation~\eqref{eq:lowersquare} implies that the lower tail of the summation of squared sub-Gaussian variables behave in a sub-Gaussian way.
+In following sections, we will show that the variable $Z_n - \bbE[Z_n]$ in Lemma~\ref{lem:lowersquare} is a sub-exponential variable.
+\end{rmk}
 

### 2020-07-21 skills/math/linear_algebra/linear_algebra_summary_p2.tex
linear algebra

@@
 \begin{defn}[\textit{Orthogonal subspaces}]\label{def:sortho}
-	Let  $S,T$ be two subspaces $S,T$. The subspaces $S$ and $T$ are orthogonal,  denoted $S \perp T$, if and only if for any $s \in S$ and $ t \in T$, $s^T t = t^T s = 0$.
+	Let  $S,T$ be two subspaces. The subspaces $S$ and $T$ are orthogonal,  denoted $S \perp T$, if and only if for any $s \in S$ and any $ t \in T$, $s^T t = t^T s = 0$.
 \end{defn}
@@
 There are four subspaces related to $\mA\colon$ column space $C(\mA)$, row space $C(\mA^T)$, nullspace $N(\mA)$, and left nullspace $N(\mA^T)$.
-Suppose the rank of $\mA$ is $rank(\mA) = r$, the dimensions of these subspaces are:
+Suppose the rank of $\mA$ is $rank(\mA) = r$.
+The dimensions of these subspaces are:
 \[ dim(C(\mA)) = dim(C(\mA^T)) = r,\quad  dim(N(A)) = n-r,\quad dim(N(\mA^T)) = m-r. \]
@@
 \begin{thm}[Orthogonality of matrix subspaces]\label{thm:ortho}
-	Let  $\mA \in \bbR^{m \times n}$ be a matrix. The row space $C(\mA^T)$ and the nullspace $N(\mA)$ are orthogonal. The column space $C(\mA)$ and the left nullspace $N(\mA^T)$ are orthogonal; i.e.,
+	Let  $\mA \in \bbR^{m \times n}$ be a matrix. The row space $C(\mA^T)$ and the nullspace $N(\mA)$ are orthogonal. The column space $C(\mA)$ and the left nullspace $N(\mA^T)$ are orthogonal. That is
 	\[ C(\mA^T) \perp N(\mA)\quad  \text{and} \quad C(\mA) \perp N(\mA^T).  \]
@@
 
-	Similarly, for any $x \in N(\mA^T)$, we have $\mA^T x = 0$, which implies $N(\mA^T) \perp C(\mA)$. 
+	Similarly, for any $x \in N(\mA^T)$, we have $\mA^T x = 0$, which implies  that $N(\mA^T) \perp C(\mA)$. 
 \end{proof}
@@
 	\begin{equation}
-		N(\mA^T \mA) = N(\mA) ; \quad rank(\mA^T \mA) = rank(\mA).
+		N(\mA^T \mA) = N(\mA) \quad \text{and} \quad rank(\mA^T \mA) = rank(\mA).
 	\end{equation}
@@
 \begin{proof}
-	Fxirst, we prove that $N(\mA^T \mA) = N(\mA)$.
+	First, we prove that $N(\mA^T \mA) = N(\mA)$.
 
-	On one hand, if $x \in N(\mA)$, then $\mA x = 0$, which implies that $ \mA^T  \mA x = \mA^T 0 = 0$. Therefore, for any $x \in N(\mA)$, the vector $x \in N(\mA^T \mA)$.
+	On one hand, if the vector $x \in N(\mA)$, then $\mA x = 0$.
+Multiplying the matrix $\mA^T$ on both sides of the equation, we have $ \mA^T  \mA x = 0$.
+Therefore, for any $x \in N(\mA)$, the vector $x \in N(\mA^T \mA)$.
+	\iffalse
 
@@
 
-	Next, given $N(\mA^T \mA) = N(\mA)$, we have $rank(\mA^T \mA) = n - dim(N(\mA^T \mA)) = n - dim(N(\mA)) = rank(\mA)$.
+	\fi
+
+	On the other hand, if the vector $x \in N(\mA^T \mA^T)$, we have
+	\[ \mA^T \mA x = 0.\]
+	Then, the vector $\mA x \in N(\mA^T)$.
+By the definition of matrix-to-vector multiplication, we also have $\mA x \in C(\mA)$.
+By the Theorem~\ref{thm:ortho}, the subspaces $N(\mA^T) \perp C(\mA)$.
+Therefore, $\mA x = 0$, and thus $x \in N(\mA)$.
+	\vspace{0.2cm}
+
+	Next, we prove that $rank(\mA^T \mA) = rank(\mA)$.
+
+	Given $N(\mA^T \mA) = N(\mA)$, we have $rank(\mA^T \mA) = n - dim(N(\mA^T \mA)) = n - dim(N(\mA)) = rank(\mA)$.
 \end{proof}
@@
 \subsection{Projections onto subspaces}
-\begin{defn}[\textit{Projection and projection matrix}]\label{def:project} Let  $x \in \bbR^{n}$ be a vector and $\mA^{n \times m}$ be a matrix with independent columns.
-The vector $p \in C(\mA)$ that satisfies
+\begin{defn}[\textit{Projection and projection matrix}]\label{def:project} Let  $x \in \bbR^{x}$ be a vector and $\mA^{n \times m}$ be a matrix with independent columns.
+The vector $p_{C(\mA)}(x) \in C(\mA)$ that satisfies
 	\begin{equation}\label{eq:project}
-		(x-p) \perp C(\mA),
+		(x-p_{C(\mA)}(x)) \perp C(\mA),
 	\end{equation} 
-	is called the projection of vector $x$ onto the column space $C(\mA)$. For all $x \in \bbR^{n}$ and the corresponding projection $p$, the matrix $\mP \in \bbR^{m \times m}$ that satisfies
-	\begin{equation}
-		p = \mP x,
+	is called the projection of vector $x$ onto the column space $C(\mA)$. For all $x \in \bbR^n$, the matrix  $\mP_{C(\mA)} \in \bbR^{n \times n}$  that satisfies	\begin{equation}
+		p_{C(\mA)}(x) =\mP_{C(\mA)} x,
 	\end{equation}
 	is called the projection matrix of $\mA$ from $\bbR^n$ onto the column space  $C(\mA)$.
+Note that the projection $p_{C(\mA)}(x)$ depends on the vector $x$ and the space $C(\mA)$, while the projection matrix $\mP_{C(\mA)}$ depends only on the space $C(\mA)$.
 \end{defn}
@@
 \begin{prop}[Projection matrix of $C(\mA)$]\label{prop:project}
-	Let  $x \in \bbR^{n}$ be a vector and $\mA^{m \times n}$ be a matrix with independent columns. The projection matrix of $\mA$ from $\bbR^n$ onto the column space  $C(\mA)$ is 
+	Let  $x \in \bbR^{n}$ be a vector and $\mA^{n \times m}$ be a matrix with independent columns. The projection matrix of $\mA$ from $\bbR^n$ onto the column space  $C(\mA)$ is 
 	\begin{equation}
-		\mP =  \mA (\mA^T \mA)^{-1} \mA^T.
+		\mP_{C(\mA)} =  \mA (\mA^T \mA)^{-1} \mA^T.
 	\end{equation}
@@
 \begin{proof}
-	By Definition~\ref{def:project}, the projection $p \in C(\mA)$.
-Then, there exists a  vector $\hat x \in \bbR^m$ such that $p = \mA \hat x$.
-By equation~\eqref{eq:project}, for all $x \in \bbR^n$, we have
-\[ A^T (x - \mA \hat x) = 0 \quad \Rightarrow  \quad \hat x = (\mA^T \mA)^{-1} \mA^T x \quad \Rightarrow \quad p = \mA (\mA^T \mA)^{-1} \mA^T  x. \]
-Therefore, the matrix $\mP =  \mA (\mA^T \mA)^{-1} \mA^T$ is the projection matrix of $\mA$ from $\bbR^n$ onto the column space $C(\mA)$.
+Let $x_0$ be an arbitrary vector of $\bbR^n$.
+By Definition~\ref{def:project}, the projection $p_{C(\mA)}(x_0) \in C(\mA)$.
+Then, there exists a vector $v(x_0) \in \bbR^m$ such that $p_{C(\mA)}(x_0) = \mA  v(x_0)$.
+By equation~\eqref{eq:project}, we have
+\begin{equation}\label{eq:projectca}
+	A^T (x_0 - \mA v(x_0)) = 0 \quad \Rightarrow  \quad v(x_0) = (\mA^T \mA)^{-1} \mA^T x_0 \quad \Rightarrow \quad p_{C(\mA)}(x_0)  = \mA (\mA^T \mA)^{-1} \mA^T  x_0. 
+\end{equation}
+Since $x_0$ is arbitrary, the equation~\eqref{eq:projectca} still holds after replacing all the $x_0$ by $x$, for all $x \in \bbR^n$.
+Therefore, the matrix $\mP_{C(\mA)} =  \mA (\mA^T \mA)^{-1} \mA^T$ is the projection matrix of $\mA$ from $\bbR^n$ onto the column space $C(\mA)$.
 \end{proof}
@@
 \begin{thm}[Properties of projection matrix]\label{thm:projectm}
-Let $\mA \in \bbR^{m \times n}$ be a matrix, and $\mP$ be the projection matrix of $\mA$ onto the column space. Then,
+Let $\mA \in \bbR^{m \times n}$ be a matrix, and $mP_{C(\mA)}$ be the projection matrix of $\mA$ onto the column space. Then,
 	\begin{equation}\label{eq:projectm}
-		\mP^T = \mP;\quad  \mP^2 = \mP.
+		\mP_{C(\mA)}^T = \mP_{C(\mA)}\quad \text{and}\quad  \mP_{C(\mA)}^2 = \mP_{C(\mA)}.
 	\end{equation}
@@
 \begin{proof}
-	By Proposition~\ref{prop:project}, the projection matrix is $\mP =   \mA(\mA^T \mA)^{-1} \mA^T$. Then,
-	\[  \mP^T = (\mA(\mA^T \mA)^{-1} \mA^T )^T = \mA (\mA^T \mA)^{-1} \mA^T = \mP. \]
-	\[  \mP^2 = \mP^T \mP =  \mA (\mA^T \mA)^{-1} \mA^T  \mA (\mA^T \mA)^{-1} \mA^T =  \mA (\mA^T \mA)^{-1} \mA^T = \mP. \]
+	By Proposition~\ref{prop:project}, the projection matrix is $\mP_{C(\mA)} =   \mA(\mA^T \mA)^{-1} \mA^T$. Then,
+	\[  \mP_{C(\mA)}^T = (\mA(\mA^T \mA)^{-1} \mA^T )^T = \mA (\mA^T \mA)^{-1} \mA^T = \mP_{C(\mA)}. \]
+Similarly, we have
+	\[  \mP_{C(\mA)}^2 = \mP_{C(\mA)}^T \mP_{C(\mA)} =  \mA (\mA^T \mA)^{-1} \mA^T  \mA (\mA^T \mA)^{-1} \mA^T =  \mA (\mA^T \mA)^{-1} \mA^T = \mP_{C(\mA)}. \]
 \end{proof}
@@
 \begin{cor}[Projection onto $N(\mA^T)$]\label{cor:iprojectm}
-	Suppose $\mP$ is a projection matrix in Theorem~\ref{thm:projectm}. Then $\mI-\mP$ is also a projection matrix of $\mA$ from $\bbR^n$ onto the left nullspace $N(\mA^T)$. 
+	Suppose $\mP_{C(\mA)}$ is the projection matrix of $\mA$ from $\bbR^n$ onto the column space $C(\mA)$.
+Then $\mI-\mP_{C(\mA)}$ is the projection matrix of $\mA$ from $\bbR^n$ onto the left nullspace $N(\mA^T)$.
 \end{cor}
@@
 \begin{proof}
-	By Definition~\ref{def:project}, for any $x \in \bbR^n$, we have
-	\[ x - \mP x \perp C(\mA) \quad \Rightarrow \quad (I-\mP)x \perp C(\mA). \]
-	By Theorem~\ref{thm:ortho}, the column space of $\mA$ is orthogonal to the left null space of $\mA$. Then,
-	\[ (\mI-\mP)x \in N(\mA^T) \quad \text{ and }\quad  (x - (\mI-\mP)x) \perp N(\mA^T). \]
-	Therefore, $\mI-\mP$ is a projection matrix of $\mA$ from $\bbR^n$ onto the left nullspace $N(\mA^T)$.
+	By Definition~\ref{def:project}, for all $x \in \bbR^n$, $\mP_{C(\mA)} x$ is the projection of $x$ from $\bbR^n$ to $C(\mA)$. Then,
+	\[ x - \mP_{C(\mA)} x \perp C(\mA) \quad \Rightarrow \quad (\mI-\mP_{C(\mA)})x \perp C(\mA). \]
+	By Theorem~\ref{thm:ortho}, we have $C(\mA) \perp N(\mA^T)$. Since $rank(C(\mA)) + rank(N(\mA^T)) = n$, $C(\mA)  \cup N(\mA^T) = \bbR^n$. Then, we have
+	\[ (\mI-\mP_{C(\mA)})x \in N(\mA^T).\]
+	Since $\mP_{C(\mA)} x$ is the projection of $x$ from $\bbR^n$ to $C(\mA)$,  the vector $\mP_{C(\mA)} x \in C(\mA)$, for all  $x \in \bbR^n$. We have,
+	\[ x -  (\mI-\mP_{C(\mA)})x = \mP_{C(\mA)} x \in C(\mA) \quad \Rightarrow \quad  x -  (\mI-\mP_{C(\mA)})x \perp N(\mA^T).\]
+	Therefore, for all $x \in \bbR^n$, $(\mI-\mP_{C(\mA)})x$ is the projection of $x$ from $\bbR^n$ onto the left nullspace $N(\mA^T)$, and thus $\mI-\mP_{C(\mA)}$ is the projection matrix of $\mA$ from $\bbR^n$ onto $N(\mA^T)$.
 \end{proof}
@@
 \[ y = \mX \beta +\epsilon,\]
-where $\beta = (\beta_0,\beta_1,...,\beta_k)$ are regression coefficients and $\epsilon$ is the noise. The least square estimate of $\beta$ is the minimizer of the loss; i.e.,
+where $\beta = (\beta_0,\beta_1,...,\beta_k)^T$ are regression coefficients and $\epsilon$ is the noise. The least-squares estimate of $\beta$ is the minimizer of the squared loss; i.e.,
 \[  \hat \beta_{LS} = \argmin_{\beta \in \bbR^{k+1}} \onorm{y - \mX \beta}^2, \]
-where $\onorm{\cdot}$ is the euclidean norm. We consider the vector $\mX 
-\hat \beta_{LS}$ as the projection of $y$ onto the column space of $\mX$, which minimizes the distance from $y$ to the column space $C(\mX)$.
-Therefore, we may use projection to solve the minimization problem.
+where $\onorm{\cdot}$ is the Euclidean norm.  The projection of $y$ onto $C(\mA)$ minimizes the Euclidean distance from $y$ to the column space $C(\mX)$. Hence,   we consider the vector $\mX 
+\hat \beta_{LS}$ as the projection of $y$ onto the column space $C(\mX)$. Therefore, we may use projection to solve the minimization problem.
 
@@
 
-The estimate $\hat \beta_{LS}$ is consistent with the estimates solved by using the derivative.
+The estimate $\hat \beta_{LS}$ is consistent with the estimates solved using the first-order condition for the least-squares optimization.
 
@@
 \begin{defn}[\textit{Gram-Schimidt Process and QR decomposition}]\label{gram-schimidt}
-Let $\mA \in \bbR^{m \times n}$ be a matrix with $rank(\mA) = n$, and  $a_i \in \bbR^{m}$ be the column of matrix $\mA$, for all $i \in [n]$.
-Gram-Schimidt process finds the orthonormal basis for $C(\mA)$ as following,
+Let $\mA \in \bbR^{m \times n}$ be a matrix with $rank(\mA) = n$, and  $a_i \in \bbR^{m}$ be the $i$-th column of matrix $\mA$, for all $i \in [n]$.
+Gram-Schimidt process finds the orthonormal basis for $C(\mA)$ as follows,
 \begin{align}
@@
 \end{align}
- The vectors $e_1,...,e_n$ are orthonormal basis of the $C(\mA)$. Based on the basis obtained by Gram-Schimidit, we decompose the matrix $\mA$ as following,
+ The vectors $e_1,...,e_n$ are orthonormal basis of the $C(\mA)$. Based on the basis obtained by Gram-Schimidit, we decompose the matrix $\mA$ as follows,
  \begin{equation}\label{eq:qr}
@@
 \section{Determinants}
-The \textit{determinant} is a number associated with any square matrix. For a square matrix $\mA$, let $\det(\mA)$ or $|\mA|$ denote the determinant of matrix $\mA$.
+The \textit{determinant} is a number associated with a square matrix. For a square matrix $\mA$, let $\det(\mA)$ or $|\mA|$ denote the determinant of matrix $\mA$.
 
 \subsection{Properties of determinants}
-We give ten properties of determinants. The last seven properties are deduced by the first three basic properties.
+We give ten properties of determinants. The last seven properties are deduced from the first three basic properties.
 \begin{enumerate}
-	\item\label{prop1} The determinant of identity matrix is equal to 1; i.e., $\det(\mI) = 1$.
-	\item\label{prop2} Exchanging two rows of a matrix reverses the sign of the matrix's determinant.
-Hence, an odd number of row exchanges reverse the sign of the determinant while an even number of row exchanges maintain the sign of the determinant.
+	\item\label{prop1} The determinant of an identity matrix is equal to 1; i.e., $\det(\mI) = 1$.
+	\item\label{prop2} Exchanging two rows of a matrix reverses the sign of the determinant.
+Hence, an odd number of row exchanges reverse the sign of the determinant, while an even number of row exchanges maintain the sign of the determinant.
 	\item\label{prop3} \begin{enumerate} 
@@
 			a&b\\c &d
-			\end{vmatrix}
+			\end{vmatrix}.
 	\end{align} 
-	\item[(b)] The determinant is linearly additive on the rows of the matrix.
+	\item[(b)] The determinant is linearly additive by row.
 	\begin{align}
@@
 			a'&b'\\c &d
-			\end{vmatrix} 
+			\end{vmatrix}. 
 	\end{align}
@@
 
-	\item \label{propidenticalrow} If the matrix contains two identical rows, the determinant of the matrix is 0.  
+	\item \label{propidenticalrow} A matrix with two identical rows has the determinant 0.  
 	\begin{proof}
@@
 
-	\item \label{propelim} Subtracting $t$ times of the $i$-th row from the $j$-th row does not change the determinant of the matrix, where $i \neq j$.
+	\item \label{propelim} Subtracting a multiplier of the $i$-th row from the $j$-th row does not change the determinant of the matrix, where $i \neq j$.
 	\begin{proof}
-		Take a two-by-two matrix as an example. By property~\ref{prop3} and property~\ref{propidenticalrow}, 
+		Take a two-by-two matrix as an example. By properties~\ref{prop3} and~\ref{propidenticalrow}, for all constant $t \in \bbR$,
 	\begin{align}
@@
 	\end{align}
-	The proof for higher-dimension matrix is similar.
+	The proof for higher-dimension matrices is similar.
 	\end{proof}
 
-	\item\label{prop0row} If the matrix has a row that is all zeros, the determinant of the matrix is 0.
+	\item\label{prop0row} A matrix with a row that is all zeros has the determinant 0.
 	\begin{proof}
@@
 	We eliminate the matrix $\mA$ to a diagonal matrix, denoted $\mA'$.
-By property~\ref{propelim}, $\det{\mA} = \det{\mA'}$.
+By property~\ref{propelim}, $\det(\mA) = \det(\mA')$.
 Since $\mA$ is triangular, the diagonal elements of $\mA'$ are still $d_1,...,d_n$.
- 		 By property~\ref{prop3} and \ref{prop1}, $\det(\mA') = \prod_{i = 1}^n d_i\det(\mI_n) =\prod_{i = 1}^n d_i$. Therefore, $\det(\mA) = \prod_{i = 1}^n d_i$.
+ 		 By properties~\ref{prop3} and \ref{prop1}, $\det(\mA') = \prod_{i = 1}^n d_i\det(\mI_n) =\prod_{i = 1}^n d_i$. Therefore, $\det(\mA) = \prod_{i = 1}^n d_i$.
 	\end{proof}
 
-	\item\label{propsingular} If the square matrix is singular, the determinant of the matrix is 0.
+	\item\label{propsingular} A matrix is singular if and only if its determinant is  0.
 
 	\begin{proof}
-		If $\mA$ is a singular matrix, the reduced row echelon form of $\mA$, denoted $RREF(\mA)$, has a row with all 0 entries.
-By and \ref{prop0row}, $\det(RREF(\mA)) = 0$.
+		($\Rightarrow$)If $\mA$ is a singular matrix, the reduced row echelon form of $\mA$, denoted $RREF(\mA)$, has a row with all 0 entries.
+By property~\ref{prop0row}, the determinant $\det(RREF(\mA)) = 0$.
 Therefore, by property~\ref{propelim}, we have $\det(\mA) =\det(RREF(\mA)) =  0$.
+
+		($\Leftarrow$) If the determinant $\det(\mA) = 0$, the determinant of $RREF(\mA)$ has the determinant 0.
+Since the reduced row echelon form  of a matrix contains only distinct nonzero rows, the determinant 0 implies that at least one zero row exists in $RREF(\mA)$.
+Therefore, the matrix $\mA$ is singular.
 	\end{proof}
 
-	\item\label{propproduct} Let  $\mA, \mB \in \bbR^{n \times n}$ be two square matrices, the determinant of the matrix product is $\det(\mA \mB) = \det(\mA) \det(\mB).$
+	\item\label{propproduct} Let  $\mA, \mB \in \bbR^{n \times n}$ be two matrices. The determinant of the matrix product is $\det(\mA \mB) = \det(\mA) \det(\mB).$
 
@@
 The product $\mE \mB$ means implementing a linear operation on the rows of $\mB$.
-By property~\ref{propelim}, linear operation does not change the determinant of $\mB$.
+By property~\ref{propelim}, linear operation maintains the determinant of $\mB$.
+Note that the determinants of elimination matrices are equal to 1.
 Therefore, we have
 		\begin{equation}\label{eq:producteli}
-			|\mE \mB| = |\mE| |\mB|
+			|\mE \mB| = |\mE| |\mB|.
 		\end{equation}
 
-		Last, suppose that both $\mA, \mB$ are invertible. By elimination, there exist a sequence of elimination matrix $\{ \mE_i \}_{i = 1}^n$ such that 
+		Last, suppose that both $\mA, \mB$ are invertible. By elimination, there exist a sequence of elimination matrices $\{ \mE_i \}_{i = 1}^n$ such that 
 		\[ \mE_n \cdots \mE_1 \mA = \mI.  \]
@@
 		\[ \mA = \mE'_n \cdots \mE'_1. \]
-		Combing equation~\eqref{eq:producteli}, we have
+		By equation~\eqref{eq:producteli}, we have
 		\begin{equation}\label{eq:prod1}
@@
 
-	\item\label{proptrans} Let $\mA$ be a square matrix, the determinant $\det(\mA^T) = \det(\mA)$.
+	\item\label{proptrans} Let $\mA$ be a square matrix. The determinant $\det(\mA^T) = \det(\mA)$.
 
@@
 By property~\ref{propdiag}, $\det(\mL^T)= \det(\mL),  \det(\mU^T) = \det(\mU)$.
-Therefore, combing property~\ref{propproduct}, we have  $\det(\mA) = \det(\mL) \det(\mU) = \det(\mU^T) \det(\mL^T) = \det(\mU^T \mL^T) = \det(\mA^T)$.
+Therefore, by property~\ref{propproduct}, we have  $\det(\mA) = \det(\mL) \det(\mU) = \det(\mU^T) \det(\mL^T) = \det(\mU^T \mL^T) = \det(\mA^T)$.
 			\end{proof}
@@
 	Let  $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ be a square matrix. The big formula for computing the determinant of $\mA$ is
-	\[ \det(\mA) =  \sum_{(\alpha_1,...,\alpha_n) \in \tP} (-1)^{N(\alpha_1,...,\alpha_n)} a_{1,\alpha_1} ... a_{n \alpha_n},  \]
-	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$, $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of necessary exchanges from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
+	\[ \det(\mA) =  \sum_{(\alpha_1,...,\alpha_n) \in \tP} (-1)^{N(\alpha_1,...,\alpha_n)} a_{1\alpha_1} ... a_{n \alpha_n},  \]
+	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$, $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of swaps from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
 \end{prop}
@@
 \begin{proof}
-	Let $\mA = \entry{a_ij} \in \bbR^{n \times n}$ be a square matrix.
-We decompose the  matrix $\mA$ by the summation of $n^n$ matrices, denoted $\mA'_i$, where $i \in [n^n]$.
-For any $\mA'_i$, there is only one entry comes from $\mA$ in each row and other entries are 0.
+	Let $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ be a square matrix.
+We decompose the  matrix $\mA$ into the sum of $n^n$ matrices, denoted $\mA'_i$, where $i \in [n^n]$.
+For any $\mA'_i$, only one entry of each row comes from $\mA$, and other entries of $\mA'_i $ are 0.
 By property~\ref{prop3}, the determinant $\det(\mA) = \sum_{i=1}^{n^n} \det(\mA'_i)$.
@@
 			 0&0 & a_{33}
-		\end{vmatrix} = \mA'_1 + \mA'_2 +\cdots + \mA'_{27}.
+		\end{vmatrix} = \det(\mA'_1)  +\cdots + \det(\mA'_{27}).
 	\end{align}
@@
 The number of nonsingular matrices coincides with the size of $\tP$.
-Intuitively, there are $n$ ways to choose an element from the first row, after which there are only $n-1$ ways to choose an element from the second row to avoid the zero determinant.
+Specifically, there are $n$ ways to choose an element from the first row, after which there are only $n-1$ ways to choose an element from the second row to avoid the zero determinant.
 Therefore, we have $n \times (n-1) \times (n-2) \times \cdots \times 2 = n!$ nonsingular matrices.
 
-	By property~\ref{propdiag} and \ref{prop2}, the determinant of a nonsingular $\mA'_i$ follows the formula \[(-1)^{N(\alpha_1,...,\alpha_n)}  a_{1,\alpha_1} ... a_{n \alpha_n},\]
-	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$, $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of necessary exchanges from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
+	By properties~\ref{propdiag} and \ref{prop2}, the determinant of a nonsingular $\mA'_i$ follows the formula \[(-1)^{N(\alpha_1,...,\alpha_n)}  a_{1\alpha_1} ... a_{n \alpha_n},\]
+	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$, $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of swaps from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
+The determinant of $\mA$ follows by the summation of these nonsingular matrices' determinants.
 \end{proof}
@@
 The cofactor associated with $a_{ij}$ is defined as \[C_{ij} = (-1)^{i+j}\det(\mA_{-i,-j}).\]
-	The matrix $\mC = \entry{C_{ij}} \in \bbR^{n \times n}$ is called cofactor matrix.  The cofactor formula of $\det(\mA)$ is
-	\[ \det(\mA) = \sum_{j}^n a_{ij} C_{ij} = \sum_{j}^n a_{ji} C_{ji} ,\quad \text{ for all } i \in [n].  \]
+	The matrix $\mC = \entry{C_{ij}} \in \bbR^{n \times n}$ is called a cofactor matrix.  The cofactor formula of $\det(\mA)$ is
+	\[ \det(\mA) = \sum_{j=1}^n a_{ij} C_{ij} = \sum_{j=1}^n a_{ji} C_{ji} ,\quad \text{ for all } i \in [n].  \]
 \end{defn}
@@
 \begin{example}[Tridiagonal matrix]
-	One example of using cofactor formula is computing the determinant of \textit{tridiagonal matrix}.
-The tridiagonal matrix is a matrix in which only nonzero elements lie on or adjacent to the diagonal.
+	One example of using the cofactor formula is computing the determinant of a \textit{tridiagonal matrix}.
+The tridiagonal matrix is a matrix in which nonzero elements lie only on or adjacent to the diagonal.
 Let $\mT_n \in \bbR^{n \times n}$ denote the tridiagonal matrix of 1's; i.e.,
@@
 \subsection{Inverse matrices}
-Previously, we use Gauss-Jordan elimination to obtain the inverse matrix of an invertible matrix. Here, we apply cofactor formula to compute the inverse matrix.
+Previously, we use Gauss-Jordan elimination to obtain the inverse matrix of an invertible matrix. Here, we apply the cofactor formula to compute the inverse matrix.
 
@@
 
-	Therefore, the diagonal elements of $\mA \mC^T$ are $(\mA \mC^T)_{ii} = \sum_k^n a_{ik} C_{ik} = \det(\mA)$, for all $ i \in [n]$.
-Consider the off-diagonal entries $(\mA \mC^T)_{ij} = \sum_{k}^n a_{ik} C_{jk}$, where $i \neq j$.
-By cofactor formula, the summation   $ \sum_{k}^n a_{ik} C_{jk}$ is equal to the determinant of a matrix whose $i$-th row and $j$-th row are identical.
-Combining property~\ref{prop0row}, the off-diagonals of $\mA \mC^T$ are 0.
+	Therefore, the diagonal elements of $\mA \mC^T$ are $(\mA \mC^T)_{ii} = \sum_{k=1}^n a_{ik} C_{ik} = \det(\mA)$, for all $ i \in [n]$.
+Consider the off-diagonal entries $(\mA \mC^T)_{ij} = \sum_{k =1}^n a_{ik} C_{jk}$, where $i \neq j$.
+By the cofactor formula, the sum   $ \sum_{k =1}^n a_{ik} C_{jk}$ is equal to the determinant of a matrix whose $i$-th row and $j$-th row are identical.
+By property~\ref{prop0row}, the off-diagonals of $\mA \mC^T$ are 0.
 Then, we have $\mA \mC^T = \det(\mA) \mI_n$.
@@
 \begin{defn}[\textit{Cramer's rule}]
-Let  $\mA\in\bbR^{n \times n}$  be  an invertible matrix, $ x, b \in \bbR^n$ be vectors.
+Let  $\mA\in\bbR^{n \times n}$  be  an invertible matrix, and $ x, b \in \bbR^n$ be two vectors.
 Applying Theorem~\ref{thm:inver} to the linear system $\mA x=b$, we obtain the Cramer's rule of $x = \mA^{-1} b$,
-\[  x = \frac{1}{\det(\mA)} \mC^T b\quad \text{ and } \quad x_j = \frac{\det(\mB_j)}{\mA_j}, \] 
-	where $\mB_j$ is the matrix $\mA$ after replacing the $j$-th column by $b$.
+\[  x = \frac{1}{\det(\mA)} \mC^T b\quad \text{ and } \quad x_j = \frac{\det(\mB_j)}{\det(\mA)}, \quad \text{for all }j \in [n], \] 
+	where $x_j$ is $j$-th entry of $x$, and $\mB_j$ is the matrix $\mA$ after replacing the $j$-th column by $b$,  for all $j \in [n]$.
 \end{defn}
@@
 \begin{defn}[\textit{Eigenvalues and eigenvectors}]\label{def:eigen}
-	Let $\mA \in \bbR^{n \times m}$ be a square matrix. Suppose there is a nonzero vector $x \in \bbR^{m}$ such that
+	Let $\mA \in \bbR^{n \times m}$ be a square matrix. Suppose there exists a nonzero vector $x \in \bbR^{m}$ such that
 	\[ \mA x = \lambda x, \quad \text{for some } \lambda \in \mathbb{C}. \]
-	The vector $x$ is called the eigenvector of $\mA$. The value $\lambda$ is called the eigenvalue of $\mA$, and $x$ is the eigenvector associated with eigenvalue $\lambda$.
+	TThe value $\lambda$ is called the eigenvalue of $\mA$, and $x$ is called the eigenvector of $\mA$ associated with eigenvalue $\lambda$.
 \end{defn}
 
- Usually, eigenvectors are normalized; i.e., $\onorm{x} = 1$, where $\onorm{\cdot}$ is the euclidean distance.
+ Usually, eigenvectors are normalized; i.e., $\onorm{x} = 1$, where $\onorm{\cdot}$ is the Euclidean norm.
 For simplicity, all the eigenvectors mentioned below are normalized.
@@
 Let $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ be a square matrix. The trace of $\mA$ is defined as
-\[ tr(\mA) = \sum_{i}^n a_{ii}.\] 
+\[ tr(\mA) = \sum_{i = 1}^n a_{ii}.\] 
 \end{defn}
@@
 
-Since the characteristic polynomial $P(\lambda)$ is of degree $n$, there are $n$ solutions to the equation $P(\lambda) = 0$.
-The solutions to $P(\lambda) = 0$, denoted $\lambda_1,...,\lambda_n$, are $n$ eigenvalues of $\mA$, which may be complex.
-If the complex eigenvalues exist, the complex eigenvalues come in conjugate pairs, because $P(\overline{ \lambda}) = \overline{ P(\lambda)} $ is also equal to 0, where $\overline{ \lambda}$ is the conjugate of $\lambda$.
-Note that $n$ eigenvalues are not necessarily distinct with each other.
+Since the characteristic polynomial $P(\lambda)$ is of degree $n$, there exist $n$ solutions to the equation $P(\lambda) = 0$.
+The solutions to $P(\lambda) = 0$, denoted $\lambda_1,...,\lambda_n$, are eigenvalues of $\mA$.
+
+\vspace{0.2cm}
+The eigenvalues may be complex, and  complex eigenvalues come in conjugate pairs.
+Let $\lambda$ be a complex eigenvalue, $\overline \lambda$ be the conjugate of $\lambda$, and $\overline{P(\cdot)}$ be the conjugate of the characteristic polynomial.
+Because the characteristic polynomial satisfies $P(\overline{ \lambda}) = \overline{ P(\lambda)}$, the conjugate eigenvalue $\overline \lambda$ is also solution of $P(\lambda) = 0$.
+
+\vspace{0.2cm}
+The $n$ eigenvalues are not necessarily distinct with each other.
 
@@
 \begin{thm}[Summation and production of eigenvalues]\label{thm:speigen}
-	Let  $\mA \in \bbR^{n \times n}$ be a square matrix, and $\lambda_1,...\lambda_n$ be the $n$ eigenvalues of $\mA$. Then,
+	Let  $\mA \in \bbR^{n \times n}$ be a square matrix, and $\lambda_1,...\lambda_n$ be the eigenvalues of $\mA$. Then,
 	\begin{equation}
-		\sum_{i=1}^n \lambda_i = tr(\mA); \quad \prod_{i=1}^n \lambda_i = \det(\mA).
+		\sum_{i=1}^n \lambda_i = tr(\mA)\quad \text{and} \quad \prod_{i=1}^n \lambda_i = \det(\mA).
 	\end{equation} 
@@
 \begin{proof}
-	First, we re-write the characteristic polynomial $P(\lambda)$ of $\mA$ as
+	First, we re-write the characteristic polynomial $P(\lambda)$ of $\mA$ in the factor formula,
 	\begin{equation}\label{eq:poly}
@@
 	\end{equation}   
-	where $\lambda_i$ are $n$ eigenvalues of $\mA$, for all $ i \in [n]$.
+	where $\lambda_i$ are the eigenvalues of $\mA$, for all $ i \in [n]$.
 By equation~\eqref{eq:poly}, the coefficient for $\lambda^{n-1}$ is equal to $(-1)^n\sum_{i=1}^n \lambda_i$.
-Compared with Definition~\ref{def:charac}, we have $tr(\mA) = \sum_{i=1}^n \lambda_i$.
-Similarly, the constant term in equation~\eqref{eq:poly} is $(-1)^n \prod_{i=1}^n \lambda_1 \cdots \lambda_n $ while the constant term in Defintion~\ref{def:charac} is $(-1)^n \det(\mA)$.
+Comparing the equation~\eqref{eq:poly} with Definition~\ref{def:charac}, we have $tr(\mA) = \sum_{i=1}^n \lambda_i$.
+Similarly, the constant term in equation~\eqref{eq:poly} is $(-1)^n \prod_{i=1}^n \lambda_i $, while the constant term in Defintion~\ref{def:charac} is $(-1)^n \det(\mA)$.
 Therefore, we have $\prod_{i=1}^n \lambda_i = \det(\mA)$.
@@
 \begin{defn}[\textit{Similar matrices}]\label{def:similar}
-	Let $\mA \in \bbR^{n \times n}$ be a square matrix, and $\mB \in \bbR^{n \times n}$ be a invertible matrix. The matrix product $\mB \mA \mB^{-1}$ is called the similar matrix of $\mA$.
+	Let $\mA \in \bbR^{n \times n}$ be a square matrix, and $\mB \in \bbR^{n \times n}$ be an invertible matrix. The matrix product $\mB \mA \mB^{-1}$ is called the similar matrix of $\mA$.
 \end{defn}
@@
 \begin{thm}[Eigenvalues for similar matrices]\label{thm:similar}
-	Let $\mA$ be a square matrix. Any similar matrices of $\mA$ share the same eigenvalues of $\mA$.
+	Let $\mA$ be a square matrix. The similar matrices of $\mA$ share the same eigenvalues of $\mA$.
 \end{thm}
 \begin{proof}
-	Let $\mB \mA \mB^{-1}$ be a similar matrix of $\mA$, and $\lambda_1,...,\lambda_n$ be the eigenvalues of $\mA$. Then, for all $\lambda_i$, there is an eigenvector $x_i$ such that
-	\[ \lambda_i x_i = \mA x_i \quad \Leftrightarrow \quad \lambda_i \mB x_i = \mB \mA \mB^{-1} \mB x_i.\]
-	Since $x_i$ is a nonzero vector, and $\mB$ is invertible, the vector $\mB x_i$ is also a nonzero vector.
-Therefore, $\lambda_i$ is an eigenvalue of $\mB \mA \mB^{-1}$ with associated eigenvector $\mB x_i$, for all $i \in [n]$.
+	Let $\mB \mA \mB^{-1}$ be a similar matrix of $\mA$, and $(\lambda_i, x_i)$ be the $i$-th eigenvalue-eigenvector pair of $\mA$, for all $i \in [n]$. Then, we have
+	\[ \lambda_i x_i = \mA x_i \quad \Rightarrow \quad \lambda_i \mB x_i = \mB \mA \mB^{-1} \mB x_i, \quad \text{for all }i \in [n]. \]
+	Since that $x_i$ is a nonzero vector, and $\mB$ is invertible, the vector $\mB x_i$ is also a nonzero vector.
+Therefore, $(\lambda_i, \mB x_i)$ is the $i$-th eigenvalue-eigenvector pair of $\mB \mA \mB^{-1}$.
 \end{proof}
@@
 \begin{thm}[Eigenvalues for powers of the matrix]
-	Let $\mA$ be a matrix, $\lambda$  be a eigenvalue of $\mA$, and $x$ be the eigenvector associated with $\lambda$. For any polynomial $P$, we have
+	Let $\mA$ be a matrix, $(\lambda,x)$  be an eigenvalue-eigenvector pair of $\mA$. For any polynomial $P$, we have
 	\[ P(\mA) x = P(\lambda) x. \]
@@
 \begin{proof}
-	For all integer $k \geq 0$ and constant $c \in \bbR$, we have
+	For all integer $k \geq 0$ and any constant $c \in \bbR$, we have
 	\[ c\mA^k x = c\mA^{k-1} \mA x= c\lambda \mA^{k-1} x = \cdots = c\lambda^k x.\]
@@
 \begin{thm}[Eigenvalues for the inverse]
-Let $\mA \in \bbR^{n \times n}$ be an invertible matrix, $\lambda_1,...,\lambda_n$ be the eigenvalues of $\mA$, and $x_1,...x_n$ be the eigenvectors associated with $\lambda_1,...,\lambda_n$.
-Then,  $\lambda_1^{-1},...,\lambda_n^{-1}$  are the eigenvalues of $\mA^{-1}$, and $x_1,...x_n$ are eigenvectors associated with  $\lambda_1^{-1},...,\lambda_n^{-1}$.
+Let $\mA \in \bbR^{n \times n}$ be an invertible matrix, and $(\lambda_i, x_i)$ be the $i$-th eigenvalue-eigenvector pair of $\mA$, for all $i \in [n]$.
+Then, $(\lambda^{-1}_i, x_i)$ is the $i$-th eigenvalue-eigenvector pair of the inverse $\mA^{-1}$, for all $i \in [n]$.
 \end{thm}
@@
 \begin{proof}
-	Since the nullspace of $\mA$ contains only a zero vector, the eigenvalue $\lambda_i >0$, for all $  i \in [n]$.
-For all $\lambda_i$ and $x_i$, multiplying $x_i$ on both sides of the equation $\mA^{-1} \mA = \mI_n$ yields
-	\[ \mA^{-1} \mA x_i = x_i \quad \Rightarrow\quad  \mA^{-1} x_i = {\lambda_i}^{-1} x_i.  \]
-	Therefore, $\lambda_i^{-1}$ is the eigenvalue of $\mA^{-1}$, and $x_i$ is the eigenvector associated with $\lambda^{-1}_i$, for all $i \in [n]$.
+	Since $\mA$ is invertible, the eigenvalue $\lambda_i \neq 0$, for all $  i \in [n]$. Multiplying $x_i$ on both sides of the equation $\mA^{-1} \mA = \mI_n$ yields
+	\[ \mA^{-1} \mA x_i = x_i \quad \Rightarrow\quad  \mA^{-1} x_i = {\lambda_i}^{-1} x_i, \quad \text{for all } i \in [n]. \]
+Therefore, $(\lambda_i^{-1},x_i)$ is the eigenvalue-eigenvector pair of $\mA^{-1}$, for all $i \in [n]$.
 \end{proof}
@@
 \begin{thm}[Independence of eigenvectors]\label{thm:indepeigen} 
-Let $\mA$ be a square matrix with at least two different eigenvalues. The eigenvectors associated with different eigenvalues are independent.
+Let $\mA$ be a square matrix with at least two distinct eigenvalues. The eigenvectors associated with distinct eigenvalues are independent.
 \end{thm}
@@
 \begin{proof}
-	Suppose the square matrix $\mA \in \bbR^{n \times n}$ has $m$ different eigenvalues $\lambda_1,...,\lambda_m$ with associated eigenvectors  $v_1,...,v_m$.
+	Suppose the square matrix $\mA \in \bbR^{n \times n}$ has $m$ distinct eigenvalues $\lambda_1,...,\lambda_m$ with associated eigenvectors  $v_1,...,v_m$.
 Consider a linear combination $c_1 v_1 + ...
 + c_m v_m = 0$, where $c_i \in \bbR$, for all $i \in [m]$.
-	Then we have
+	Then, we have
 	\begin{equation}\label{eq:indep1}
@@
 	\end{equation}
-	Consider the new linear combination $ c_2 (\lambda_1 - \lambda_2) v_2 + \cdots + c_m (\lambda_1 - \lambda_m) v_m = 0$ and repeat the steps from equation~\eqref{eq:indep1} to equation~\eqref{eq:indep3}.
+	Consider the new linear combination $ c_2 (\lambda_1 - \lambda_2) v_2 + \cdots + c_m (\lambda_1 - \lambda_m) v_m = 0$, and repeat the steps from equation~\eqref{eq:indep1} to equation~\eqref{eq:indep3}.
 After $m$ iterations, we have
@@
 \begin{thm}[Eigenvalues for triangular matrices]\label{thm:trianeigen}
-	The eigenvalues for a triangular matrix are the entries on the diagonal.
+	The eigenvalues for a triangular matrix are the diagonal elements.
 \end{thm}
@@
 Let $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ be a triangular matrix. By the property of determinant~\ref{propdiag}, we have
-\[ \det(\mA) = \prod_{i=1}^n a_{ii}  ; \quad \det(\mA - \lambda \mI_n) = \prod_{i=1}^n (a_{ii}-\lambda). \]
-To let $\det(\mA - \lambda \mI_n) = 0$, we have $\lambda = a_{ii}$, for all $ i\in[n]$. Therefore, the entries on the diagonal are the eigenvalues of a triangular matrix.
+\[ \det(\mA) = \prod_{i=1}^n a_{ii}  \quad \text{and} \quad \det(\mA - \lambda \mI_n) = \prod_{i=1}^n (a_{ii}-\lambda). \]
+To let $\det(\mA - \lambda \mI_n) = 0$, we obtain $\lambda = a_{ii}$, for all $ i\in[n]$. Therefore, the diagonal elements are the eigenvalues of a triangular matrix.
 \end{proof}
@@
 \begin{defn}[Antisymmetric matrices]
-	The matrix $\mA$ is an antisymmetric matrix if $\mA$ satisfies
+	The matrix $\mA$ is an antisymmetric matrix if 
 	\[\mA^T = - \mA. \] 
@@
 \begin{thm}[Eigenvalues for symmetric and antisymmetric matrices]\label{thm:symmeign}
-All the eigenvalues of a symmetric square matrix with only real entries are real.
-All the eigenvalues of an antisymmetric square matrix are imaginary; i.e., $\lambda = b i$, where $b \in \bbR, i = \sqrt{-1}$.
+All the eigenvalues of a symmetric matrix with real entries only are real. All the eigenvalues of an antisymmetric matrix are imaginary; i.e., $\lambda = b i$, where $b \in \bbR$ and $ i = \sqrt{-1}$.
 \end{thm}
 \begin{proof}
-	First, we prove by contradiction that symmetric square matrices with all real entries have real eigenvalues.
+	First, we prove by contradiction that a symmetric matrix with real entries only has real eigenvalues.
 
-	Let  $\mA$  be a symmetric  matrix with only real entries.
+	Let  $\mA$  be a symmetric  matrix with real entries only.
 Suppose $\lambda$  is a complex eigenvalue of $\mA$, and $x$ is the eigenvector associated with $\lambda$.
@@
 Let $\overline \lambda$ and $\overline x$ denote the conjugate eigenvalue and eigenvector, respectively.
-Then we have
+Then,  we have
 	\begin{equation}\label{eq:symm}
@@
 	\[ \overline x^T \mA x  = x^T \mA^T \overline x = x^T \mA \overline x .\] 
-	 Subtracting the equation~\eqref{eq:symm} from equation~\eqref{eq:symm2} yields	\[  0 = ( \overline x^T \mA x -  x^T \mA \overline x ) = (\lambda - \overline \lambda) x^T \overline x.  \]
-	Since that $x^T \overline x$ is a real number, the imaginary part of $\lambda$ is equal to 0, which contradicts the assumption that $\lambda$ is complex.
+	 Subtracting the equation~\eqref{eq:symm2} from equation~\eqref{eq:symm} yields	\[  0 = ( \overline x^T \mA x -  x^T \mA \overline x ) = (\lambda - \overline \lambda) x^T \overline x.  \]
+	Since that $x^T \overline x$ is a real number, the imaginary part of $\lambda$ is equal to 0.
+Therefore, the eigenvalue $\lambda$ is a real number, which contradicts the assumption that $\lambda$ is complex.
 
@@
 \begin{thm}[Orthogonality of eigenvectors for symmetric matrices]\label{thm:symmortho}
-The eigenvectors of a symmetric matrix $\mA \in \bbR^{n \times n}$ associated with different eigenvalues are orthogonal.
+The eigenvectors of a symmetric matrix $\mA \in \bbR^{n \times n}$ associated with distinct eigenvalues are orthogonal.
 \end{thm}
@@
 \begin{proof}
-	Let $\lambda_1$ and $\lambda_2$ be two different eigenvalues of $\mA$, and  $x_1, x_2$ are two eigenvectors associated with $\lambda_1$ and $\lambda_2$, respectively.
-By the symmetry of $\mA$, we have
+	Let $\lambda_1$ and $\lambda_2$ be two distinct eigenvalues of $\mA$, and  $x_1, x_2$ are two eigenvectors associated with $\lambda_1$ and $\lambda_2$, respectively. By the symmetry of $\mA$, we have
 	\[ x_2^T \mA x_1 = x_1^T \mA^T x_2  = x_1^T \mA x_2.\]
-	Then,
+	Then, we have
 	\[  0 = ( x_2^T \mA x_1 -  x_1^T \mA x_2 ) = (\lambda_1 - \lambda_2) x_2^T x_1. \]
-	Since $\lambda_1, \lambda_2$ are different, we have $\lambda_1 - \lambda_2 \neq 0$ and $x_2^T x_1 = 0$. Therefore, $x_1$ and $x_2$ are orthogonal. 
+	Since $\lambda_1, \lambda_2$ are distinct, we have $\lambda_1 - \lambda_2 \neq 0$ and $x_2^T x_1 = 0$. Therefore, $x_1$ and $x_2$ are orthogonal. 
 \end{proof}
@@
 \begin{proof}
-	First, let $x_0$ be a nonzero eigenvector associated with $\lambda_0$.
-For any $x_0 \in \bbR^{n}$, there are $n-1$ additional orthogonal vectors $y_1,...,y_{n-1}$ such that $y_j \perp x_0$, for all $j \in [(n-1)]$,  and $(x_0, y_1,...,y_{n-1})$ forms the basis of $\bbR^n$.
-Let $\mY = [y_1,...y_{n-1}]$ and $\mX = [x_0,Y]$.
+	We prove the existence by construction.
+Let $x_0$ be a nonzero eigenvector associated with $\lambda_0$.
+For any $x_0 \in \bbR^{n}$, there are $n-1$ additional orthogonal vectors $y_1,...,y_{n-1}$ such that $(x_0, y_1,...,y_{n-1})$ forms an orthonormal basis of $\bbR^n$.
+Let $\mY = [y_1,...,y_{n-1}]$ and $\mX = [x_0,Y]$.
 Since $\mA$ is a symmetric matrix, we have $x_0^T \mA = \lambda_0 x_0^T $.
@@
 	\end{align}
-	Note that $\mX$ is an orthogonal matrix which satisfies $\mX^T = \mX^{-1}$.
+	Note that $\mX$ is an orthogonal matrix satisfying $\mX^T = \mX^{-1}$.
 By Theorem~\ref{thm:similar}, $\mX^T \mA \mX$ is a similar matrix of $\mA$.
@@
 The characteristic  polynomial of $\mX^T \mA \mX - \lambda \mI_n$ is
-	\[  P(\lambda) = \det(\mX^T \mA \mX - \lambda \mI_n) = (\lambda_0 - \lambda) \det(\mY^T \mA \mY - \lambda \mI_{n-1}). \]
+	\begin{equation}\label{eq:symmploy}
+		P(\lambda) = \det(\mX^T \mA \mX - \lambda \mI_n) = (\lambda_0 - \lambda) \det(\mY^T \mA \mY - \lambda \mI_{n-1}). 
+	\end{equation}
 	Since $\lambda_0$ has multiplicity $m \geq 2$, the term $\det(\mY^T \mA \mY - \lambda_0 \mI_{n-1}) = 0$.
@@
 Let $v_1,v_2$ be the two orthogonal vectors in the nullspace of $\mX^T \mA \mX - \lambda_0 \mI_n$.
-Then $v_1, v_2$ are two eigenvectors of $\mX^T \mA \mX$, and $\mX v_1, \mX v_2$ are two orthogonal eigenvectors of $\mA$ associated with $\lambda_0$.
+Then, $v_1, v_2$ are two eigenvectors of $\mX^T \mA \mX$, and $\mX v_1, \mX v_2$ are two orthogonal eigenvectors of $\mA$ associated with $\lambda_0$.
+Therefore, we have obtained two orthonormal  eigenvectors of $\mA$ associated with $\lambda_0$.
 
-	Replace $x_0$ by two orthogonal vectors $x_1 = \mX v_1, x_2 = \mX v_2$ and repeat the above steps. Finally, we will have $m$ orthogonal eigenvectors of $\mA$ associated with $\lambda_0$.
+	Now, replace $x_0$ by two orthogonal vectors $x_1 = \mX v_1, x_2 = \mX v_2$ and repeat the above steps.
+
+	After $(m-1)$ times iterations, we have obtained $m$ orthogonal eigenvectors of $\mA$ associated with $\lambda_0$.
+Let $\mX^{(m)}, \mY^{(m)}$ denote the matrix $\mX, \mY$ in the $m$-th iteration, respectively.
+The polynomial~\eqref{eq:symmploy} becomes
+	\[ P(\lambda) = \det(\mX^{(m),T} \mA \mX^{(m)} - \lambda \mI_n) = (\lambda_0 - \lambda)^m \det(\mY^{(m),T} \mA \mY^{(m)} - \lambda \mI_{n-m}). \]
+	Since $\lambda_0$ has multiplicity $m$, the determinant $\det(\mY^{(m),T} \mA \mY^{(Mm)} - \lambda \mI_{n-m}) \neq 0$.
+The dimension of the nullspace of $\mX^{(m),T} \mA \mX^{(m)} - \lambda_0 \mI_n$ is $m$.
+Therefore, we are unable to find the $(m+1)$-th orthonormal eigenvector associated with $\lambda_0$, and the iteration stops.
 \end{proof}
@@
 
-\begin{defn}[Diagonalization]
-	Let  $\mA \in \bbR^{n \times n}$ be a square matrix with $n$ independent eigenvectors. We diagonalize the matrix $\mA$ as
+\begin{defn}[Diagonalization]\label{def:diagonalization}
+	Let  $\mA \in \bbR^{n \times n}$ be a square matrix with $n$ independent eigenvectors. We are able to  diagonalize the matrix $\mA$ as
 	\begin{equation}\label{eq:defdiag}
@@
 	\end{equation}
-	where $\Lambda \in \bbR^{n \times n}$ is a diagonal matrix whose entries are $n$ eigenvalues of $\mA$ and  $\mS \in \bbR^{n \times n}$ is a matrix whose columns are independent eigenvectors of $\mA$ associated with eigenvalues in $\Lambda$.
+	where $\Lambda \in \bbR^{n \times n}$ is a diagonal matrix consisting of $n$ eigenvalues of $\mA$, and  $\mS \in \bbR^{n \times n}$ is a matrix whose columns are independent eigenvectors of $\mA$ ordered in the same as the eigenvalues in $\Lambda$.
 \end{defn}
@@
 \begin{thm}[Diagonalization of symmetric matrices]
-Let $\mA$ be a symmetric matrix. The matrix $\mA$ can be diagonalized.
+Let $\mA$ be a symmetric matrix. The matrix $\mA$ is diagonalizable.
 \end{thm}
@@
 \begin{proof}
-	By Theorems~\ref{thm:symmortho} and Theorem~\ref{thm:symmrepeat}, for any symmetric matrix $\mA \in \bbR^{n \times n}$, $\mA$ have $n$ orthogonal eigenvectors, even though there may have repeated eigenvalues.
-Therefore, the matrix $\mA$ can be diagonalized as equation~\eqref{eq:defdiag}.
+	By Theorems~\ref{thm:symmortho} and Theorem~\ref{thm:symmrepeat}, for any symmetric matrix $\mA \in \bbR^{n \times n}$, $\mA$ have $n$ orthogonal eigenvectors with possibly repeated eigenvalues.
+Therefore, the matrix $\mA$   is able to be diagonalized as equation~\eqref{eq:defdiag}.
 \end{proof}
@@
 \begin{proof}
-	We diagonalize the matrix $\mA$ as $\mA =\mS^{-1} \Lambda \mS$. For all integer $k\geq 0$, we have $\mA^k = \mS^{-1} \Lambda^{k} \mS$. Then,
+Since $\mA$ has $n$ independent eigenvectors, the matrix $\mA$ is diagonalizable.
+	Let $\mA =\mS^{-1} \Lambda \mS$, where $\Lambda, \mS$ are of the same definitions in Definition~\ref{def:diagonalization}. For all integer $k\geq 0$, we have $\mA^k = \mS^{-1} \Lambda^{k} \mS$. Then,
 	\[ \lim_{k \rightarrow +\infty} \mA^k = \lim_{k \rightarrow +\infty} \mS^{-1} \Lambda^{k} \mS  =  \lim_{k \rightarrow +\infty} \mS^{-1} \begin{bmatrix}

### 2020-07-22 skills/math/linear_algebra/linear_algebra_summary_p2.tex
linear algebra

@@
 \subsection{Projections onto subspaces}
-\begin{defn}[\textit{Projection and projection matrix}]\label{def:project} Let  $x \in \bbR^{x}$ be a vector and $\mA^{n \times m}$ be a matrix with independent columns.
-The vector $p_{C(\mA)}(x) \in C(\mA)$ that satisfies
+\begin{defn}[\textit{Projection and projection matrix}]\label{def:project} Let  $x \in \bbR^{x}$ be a vector and $\mA^{n \times m}$ be a matrix. The vector $p_{C(\mA)}(x) \in C(\mA)$ that satisfies
 	\begin{equation}\label{eq:project}
@@
 \begin{thm}[Properties of projection matrix]\label{thm:projectm}
-Let $\mA \in \bbR^{m \times n}$ be a matrix, and $mP_{C(\mA)}$ be the projection matrix of $\mA$ onto the column space. Then,
+Let $\mA \in \bbR^{m \times n}$ be a matrix, and $\mP_{C(\mA)}$ be the projection matrix of $\mA$ onto the column space $C(\mA)$. Then,
 	\begin{equation}\label{eq:projectm}
@@
 \[  \hat \beta_{LS} = \argmin_{\beta \in \bbR^{k+1}} \onorm{y - \mX \beta}^2, \]
-where $\onorm{\cdot}$ is the Euclidean norm.  The projection of $y$ onto $C(\mA)$ minimizes the Euclidean distance from $y$ to the column space $C(\mX)$. Hence,   we consider the vector $\mX 
+where $\onorm{\cdot}$ is the Euclidean norm.  The projection of $y$ onto $C(\mX)$ minimizes the Euclidean distance from $y$ to the column space $C(\mX)$. Hence,   we consider the vector $\mX 
 \hat \beta_{LS}$ as the projection of $y$ onto the column space $C(\mX)$. Therefore, we may use projection to solve the minimization problem.
@@
 When $m = n$, $\mQ$ is an orthogonal matrix and $\mQ^T = \mQ^{-1}$.
-The projection matrix of $\mQ$ onto the column space $C(\mQ)$, denoted $\mP$,  becomes $\mP = \mI_m$.
+The projection matrix of $\mQ$ onto the column space $C(\mQ)$, denoted $\mP_{C(\mQ)}$,  becomes $\mP_{C(\mQ)} = \mI_m$.
 
@@
 \end{align}
- The vectors $e_1,...,e_n$ are orthonormal basis of the $C(\mA)$. Based on the basis obtained by Gram-Schimidit, we decompose the matrix $\mA$ as follows,
+ The vectors $e_1,...,e_n$ are orthonormal basis of $C(\mA)$. Based on the basis obtained by Gram-Schimidit, we decompose the matrix $\mA$ as follows,
  \begin{equation}\label{eq:qr}
@@
 
-	\item \label{propelim} Subtracting a multiplier of the $i$-th row from the $j$-th row does not change the determinant of the matrix, where $i \neq j$.
+	\item \label{propelim} Subtracting a multiplier of the $i$-th row from the $j$-th row maintains the determinant of the matrix, where $i \neq j$.
 	\begin{proof}
-		Take a two-by-two matrix as an example. By properties~\ref{prop3} and~\ref{propidenticalrow}, for all constant $t \in \bbR$,
+		Take a 2-by-2 matrix as an example. By properties~\ref{prop3} and~\ref{propidenticalrow}, for any constant $t \in \bbR$, we have 
 	\begin{align}
@@
 
-		($\Leftarrow$) If the determinant $\det(\mA) = 0$, the determinant of $RREF(\mA)$ has the determinant 0.
-Since the reduced row echelon form  of a matrix contains only distinct nonzero rows, the determinant 0 implies that at least one zero row exists in $RREF(\mA)$.
+		($\Leftarrow$) If the determinant $\det(\mA) = 0$, the determinant of $RREF(\mA)$ is 0.
+Since the reduced row echelon form  of a matrix contains only distinct nonzero rows, the zero determinant implies that there exists at least one zero row in $RREF(\mA)$.
 Therefore, the matrix $\mA$ is singular.
@@
 	\begin{proof}
-		First, if at least one of $\mA, \mB$ is singular, the product $\mA \mB$ is also singular. Then, $\det(\mA \mB) = \det(\mA) \det(\mB) = 0$.
+		First, suppose that at least one of $\mA, \mB$ is singular. The product $\mA \mB$ is also singular. Then, $\det(\mA \mB) = \det(\mA) \det(\mB) = 0$.
 
@@
 By property~\ref{propelim}, linear operation maintains the determinant of $\mB$.
-Note that the determinants of elimination matrices are equal to 1.
+Note that the determinant of an elimination matrix is 1.
 Therefore, we have
@@
 		By LU decomposition, we rewrite the matrix as $\mA = \mL \mU$, where $\mL$ is a lower-triangular matrix and $\mU$ is an upper-triangular matrix.
-Note that the transports $\mL^T,\mU^T$ have the same diagonal elements with $\mL, \mU$ respectively.
-By property~\ref{propdiag}, $\det(\mL^T)= \det(\mL),  \det(\mU^T) = \det(\mU)$.
-Therefore, by property~\ref{propproduct}, we have  $\det(\mA) = \det(\mL) \det(\mU) = \det(\mU^T) \det(\mL^T) = \det(\mU^T \mL^T) = \det(\mA^T)$.
+Note that the transports $\mL^T,\mU^T$ have the same diagonal elements with $\mL, \mU$, respectively.
+By property~\ref{propdiag}, we have $\det(\mL^T)= \det(\mL)$ and $  \det(\mU^T) = \det(\mU)$.
+Therefore, by property~\ref{propproduct}, we obtain  $\det(\mA) = \det(\mL) \det(\mU) = \det(\mU^T) \det(\mL^T) = \det(\mU^T \mL^T) = \det(\mA^T)$.
 			\end{proof}
@@
 We decompose the  matrix $\mA$ into the sum of $n^n$ matrices, denoted $\mA'_i$, where $i \in [n^n]$.
-For any $\mA'_i$, only one entry of each row comes from $\mA$, and other entries of $\mA'_i $ are 0.
+For each $\mA'_i$, only one entry in each row comes from $\mA$, and the other entries of $\mA'_i $ are 0.
 By property~\ref{prop3}, the determinant $\det(\mA) = \sum_{i=1}^{n^n} \det(\mA'_i)$.
@@
 	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$, $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of swaps from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
-The determinant of $\mA$ follows by the summation of these nonsingular matrices' determinants.
+The determinant of $\mA$ follows by the summation of the nonsingular matrices' determinants.
 \end{proof}
@@
 The cofactor associated with $a_{ij}$ is defined as \[C_{ij} = (-1)^{i+j}\det(\mA_{-i,-j}).\]
-	The matrix $\mC = \entry{C_{ij}} \in \bbR^{n \times n}$ is called a cofactor matrix.  The cofactor formula of $\det(\mA)$ is
+	The matrix $\mC = \entry{C_{ij}} \in \bbR^{n \times n}$ is called the cofactor matrix of $\mA$.  The cofactor formula of $\det(\mA)$ is
 	\[ \det(\mA) = \sum_{j=1}^n a_{ij} C_{ij} = \sum_{j=1}^n a_{ji} C_{ji} ,\quad \text{ for all } i \in [n].  \]
@@
 
-Let $\mT_{n, -i,-j}$ be the submatrix of  $\mT_n$ after removing the $i$-th row and the $j$-th column.   By cofactor formula, we have
+Let $\mT_{n, -i,-j}$ be the submatrix of  $\mT_n$ after removing the $i$-th row and the $j$-th column.   By the cofactor formula, we have
 \[  \det(\mT_n) = 1 \times \det(\mT_{n, -1,-1}) - 1 \times \det(\mT_{n,-1,-2}). \]
-By the definition of tridiagonal matrix, $\mT_{n,-1,-1} = \mT_{n-1}$, and $\det( \mA_{n,-1,-2}) = 1 \times \det(\mT_{n-2}) $. Therefore, 
+By the definition of tridiagonal matrix, we have $\mT_{n,-1,-1} = \mT_{n-1}$, and $\det( \mA_{n,-1,-2}) = 1 \times \det(\mT_{n-2}) $. Therefore, 
 \[ \det(\mT_n) = \det(\mT_{n-1}) - \det(\mT_{n-2}).  \] 
@@
 	Therefore, the diagonal elements of $\mA \mC^T$ are $(\mA \mC^T)_{ii} = \sum_{k=1}^n a_{ik} C_{ik} = \det(\mA)$, for all $ i \in [n]$.
-Consider the off-diagonal entries $(\mA \mC^T)_{ij} = \sum_{k =1}^n a_{ik} C_{jk}$, where $i \neq j$.
+Consider the off-diagonal entries $(\mA \mC^T)_{ij} = \sum_{k =1}^n a_{ik} C_{jk}$, for all  $i \neq j$ and $i, j \in [n]$.
 By the cofactor formula, the sum   $ \sum_{k =1}^n a_{ik} C_{jk}$ is equal to the determinant of a matrix whose $i$-th row and $j$-th row are identical.
 By property~\ref{prop0row}, the off-diagonals of $\mA \mC^T$ are 0.
-Then, we have $\mA \mC^T = \det(\mA) \mI_n$.
+Then, we obtain $\mA \mC^T = \det(\mA) \mI_n$.
 \end{proof}
@@
 \[  x = \frac{1}{\det(\mA)} \mC^T b\quad \text{ and } \quad x_j = \frac{\det(\mB_j)}{\det(\mA)}, \quad \text{for all }j \in [n], \] 
-	where $x_j$ is $j$-th entry of $x$, and $\mB_j$ is the matrix $\mA$ after replacing the $j$-th column by $b$,  for all $j \in [n]$.
+	where $x_j$ is $j$-th entry of $x$, and $\mB_j$ is the matrix $\mA$ after replacing the $j$-th column by $b$.
 \end{defn}
@@
 		The two edges of the triangle are $v_1 = (x_2 - x_1, y_2 -y_1)$ and $v_2 = (x_3 - x_1, y_3 -y_1)$.
-The area of the triangle is a half of the area of the parallelepiped determined by $v_1, v_2$.
+The area of the triangle is half of the area of the parallelepiped determined by $v_1, v_2$.
 Therefore, the area of the triangle is
@@
 	\[ \mA x = \lambda x, \quad \text{for some } \lambda \in \mathbb{C}. \]
-	TThe value $\lambda$ is called the eigenvalue of $\mA$, and $x$ is called the eigenvector of $\mA$ associated with eigenvalue $\lambda$.
+	The value $\lambda$ is called the eigenvalue of $\mA$, and the vector $x$ is called the eigenvector of $\mA$ associated with eigenvalue $\lambda$.
 \end{defn}
@@
 \vspace{0.2cm}
-The eigenvalues may be complex, and  complex eigenvalues come in conjugate pairs.
+The eigenvalues may be complex, and the complex eigenvalues come in conjugate pairs.
 Let $\lambda$ be a complex eigenvalue, $\overline \lambda$ be the conjugate of $\lambda$, and $\overline{P(\cdot)}$ be the conjugate of the characteristic polynomial.
-Because the characteristic polynomial satisfies $P(\overline{ \lambda}) = \overline{ P(\lambda)}$, the conjugate eigenvalue $\overline \lambda$ is also solution of $P(\lambda) = 0$.
+Because the characteristic polynomial satisfies $P(\overline{ \lambda}) = \overline{ P(\lambda)}$, the conjugate eigenvalue $\overline \lambda$ is also a solution to $P(\lambda) = 0$.
 
 \vspace{0.2cm}
-The $n$ eigenvalues are not necessarily distinct with each other.
+The eigenvalues of a matrix are not necessarily distinct with each other.
 
@@
 \begin{proof}
-	First, we re-write the characteristic polynomial $P(\lambda)$ of $\mA$ in the factor formula,
+	First, we re-write the characteristic polynomial $P(\lambda)$ of $\mA$ as a factor formula,
 	\begin{equation}\label{eq:poly}
@@
 	where $\lambda_i$ are the eigenvalues of $\mA$, for all $ i \in [n]$.
-By equation~\eqref{eq:poly}, the coefficient for $\lambda^{n-1}$ is equal to $(-1)^n\sum_{i=1}^n \lambda_i$.
-Comparing the equation~\eqref{eq:poly} with Definition~\ref{def:charac}, we have $tr(\mA) = \sum_{i=1}^n \lambda_i$.
+In equation~\eqref{eq:poly}, the coefficient for $\lambda^{n-1}$ is equal to $(-1)^n\sum_{i=1}^n \lambda_i$.
+Comparing the equation~\eqref{eq:poly} with Definition~\ref{def:charac}, we obtain $tr(\mA) = \sum_{i=1}^n \lambda_i$.
 Similarly, the constant term in equation~\eqref{eq:poly} is $(-1)^n \prod_{i=1}^n \lambda_i $, while the constant term in Defintion~\ref{def:charac} is $(-1)^n \det(\mA)$.
-Therefore, we have $\prod_{i=1}^n \lambda_i = \det(\mA)$.
+Therefore, we obtain $\prod_{i=1}^n \lambda_i = \det(\mA)$.
 \end{proof}
@@
 	Since that $x_i$ is a nonzero vector, and $\mB$ is invertible, the vector $\mB x_i$ is also a nonzero vector.
-Therefore, $(\lambda_i, \mB x_i)$ is the $i$-th eigenvalue-eigenvector pair of $\mB \mA \mB^{-1}$.
+Therefore, $(\lambda_i, \mB x_i)$ is the $i$-th eigenvalue-eigenvector pair of $\mB \mA \mB^{-1}$, for all $i \in [n]$.
 \end{proof}
@@
 \begin{thm}[Eigenvalues for powers of the matrix]
-	Let $\mA$ be a matrix, $(\lambda,x)$  be an eigenvalue-eigenvector pair of $\mA$. For any polynomial $P$, we have
+	Let $\mA$ be a matrix, and $(\lambda,x)$  be an eigenvalue-eigenvector pair of $\mA$. For any polynomial $P$, we have
 	\[ P(\mA) x = P(\lambda) x. \]
@@
 	\[ c\mA^k x = c\mA^{k-1} \mA x= c\lambda \mA^{k-1} x = \cdots = c\lambda^k x.\]
-	Therefore, $P(\mA)x = P(\lambda)x$, for any polynomial $P$.
+	Therefore, $P(\mA)x = P(\lambda)x$ holds for any polynomial $P$.
 \end{proof}
@@
 \begin{thm}[Independence of eigenvectors]\label{thm:indepeigen} 
-Let $\mA$ be a square matrix with at least two distinct eigenvalues. The eigenvectors associated with distinct eigenvalues are independent.
+ The eigenvectors of a matrix associated with distinct eigenvalues are independent.
 \end{thm}
@@
 \begin{thm}[Eigenvalues for triangular matrices]\label{thm:trianeigen}
-	The eigenvalues for a triangular matrix are the diagonal elements.
+	The eigenvalues of a triangular matrix are the diagonal elements.
 \end{thm}
@@
 \[ \det(\mA) = \prod_{i=1}^n a_{ii}  \quad \text{and} \quad \det(\mA - \lambda \mI_n) = \prod_{i=1}^n (a_{ii}-\lambda). \]
-To let $\det(\mA - \lambda \mI_n) = 0$, we obtain $\lambda = a_{ii}$, for all $ i\in[n]$. Therefore, the diagonal elements are the eigenvalues of a triangular matrix.
+To let $\det(\mA - \lambda \mI_n) = 0$, we obtain $\lambda = a_{ii}$, for all $ i\in[n]$. Therefore, the diagonal elements are the eigenvalues of the triangular matrix.
 \end{proof}
@@
 \begin{thm}[Eigenvalues for symmetric and antisymmetric matrices]\label{thm:symmeign}
-All the eigenvalues of a symmetric matrix with real entries only are real. All the eigenvalues of an antisymmetric matrix are imaginary; i.e., $\lambda = b i$, where $b \in \bbR$ and $ i = \sqrt{-1}$.
+All the eigenvalues of a symmetric matrix with real entries only are real. All the eigenvalues of an antisymmetric matrix are imaginary.
 \end{thm}
@@
 	Let  $\mA$  be a symmetric  matrix with real entries only.
-Suppose $\lambda$  is a complex eigenvalue of $\mA$, and $x$ is the eigenvector associated with $\lambda$.
-Since all the entries of $\mA$ are real, the eigenvector $x$ is also complex.
+Suppose $\lambda$  is a complex eigenvalue of $\mA$ with nonzero imaginary part, and $x$ is the eigenvector associated with $\lambda$.
+Since all the entries of $\mA$ are real, the eigenvector $x$ is also a complex vector with nonzero imaginary part.
 Let $\overline \lambda$ and $\overline x$ denote the conjugate eigenvalue and eigenvector, respectively.
@@
 	 Subtracting the equation~\eqref{eq:symm2} from equation~\eqref{eq:symm} yields	\[  0 = ( \overline x^T \mA x -  x^T \mA \overline x ) = (\lambda - \overline \lambda) x^T \overline x.  \]
-	Since that $x^T \overline x$ is a real number, the imaginary part of $\lambda$ is equal to 0.
-Therefore, the eigenvalue $\lambda$ is a real number, which contradicts the assumption that $\lambda$ is complex.
+	Since that $x^T \overline x$ is a nonzero real number, the imaginary part of $\lambda$ is equal to 0, which contradicts the assumption that $\lambda$ has nonzero imaginary part.
+Therefore, the eigenvalues of $\mA$ are real.
 
+	\vspace{0.2cm}
 	Next, we prove that the eigenvalues for antisymmetric matrices are imaginary. 
@@
 	\[  0 = ( \overline x^T \mA x +  x^T \mA \overline x ) = (\lambda + \overline \lambda) x^T \overline x.   \]
-	Since that $x^T \overline x$ is a real number, the real part of $\lambda$ is equal to 0. Therefore, the eigenvalues for an antisymmetric matrix are imaginary.
+	Since that $x^T \overline x$ is a nonzero real number, the real part of $\lambda$ is equal to 0. Therefore, the eigenvalues for an antisymmetric matrix are imaginary.
 \end{proof}
@@
 \begin{thm}[Orthogonality of eigenvectors for symmetric matrices]\label{thm:symmortho}
-The eigenvectors of a symmetric matrix $\mA \in \bbR^{n \times n}$ associated with distinct eigenvalues are orthogonal.
+The eigenvectors of a symmetric matrix associated with distinct eigenvalues are orthogonal.
 \end{thm}
@@
 	We prove the existence by construction.
-Let $x_0$ be a nonzero eigenvector associated with $\lambda_0$.
-For any $x_0 \in \bbR^{n}$, there are $n-1$ additional orthogonal vectors $y_1,...,y_{n-1}$ such that $(x_0, y_1,...,y_{n-1})$ forms an orthonormal basis of $\bbR^n$.
+Let $x_0$ be an eigenvector associated with $\lambda_0$.
+Then, there exist $n-1$ additional orthogonal vectors $y_1,...,y_{n-1} \in \bbR^n$ such that $(x_0, y_1,...,y_{n-1})$ forms an orthonormal basis of $\bbR^n$.
 Let $\mY = [y_1,...,y_{n-1}]$ and $\mX = [x_0,Y]$.
@@
 	Note that $\mX$ is an orthogonal matrix satisfying $\mX^T = \mX^{-1}$.
-By Theorem~\ref{thm:similar}, $\mX^T \mA \mX$ is a similar matrix of $\mA$.
-Hence, $\lambda_0$ is also a repeated eigenvalue of $\mX^T \mA \mX$.
-The characteristic  polynomial of $\mX^T \mA \mX - \lambda \mI_n$ is
+The product $\mX^T \mA \mX$ is a similar matrix of $\mA$.
+By Theorem~\ref{thm:similar},  $\lambda_0$ is also a repeated eigenvalue of $\mX^T \mA \mX$ with multiplicity $m$.
+The characteristic  polynomial of $\mX^T \mA \mX$ is
 	\begin{equation}\label{eq:symmploy}
@@
 
+	\vspace{0.2cm}
 	Now, replace $x_0$ by two orthogonal vectors $x_1 = \mX v_1, x_2 = \mX v_2$ and repeat the above steps.
 
-	After $(m-1)$ times iterations, we have obtained $m$ orthogonal eigenvectors of $\mA$ associated with $\lambda_0$.
-Let $\mX^{(m)}, \mY^{(m)}$ denote the matrix $\mX, \mY$ in the $m$-th iteration, respectively.
+	\vspace{0.2cm}
+	After $(m-1)$ repetitions, we have obtained $m$ orthogonal eigenvectors of $\mA$ associated with $\lambda_0$.
+Let $\mX^{(m)}, \mY^{(m)}$ denote the matrix $\mX, \mY$ after the $(m-1)$-th repetition, respectively.
 The polynomial~\eqref{eq:symmploy} becomes
@@
 The dimension of the nullspace of $\mX^{(m),T} \mA \mX^{(m)} - \lambda_0 \mI_n$ is $m$.
-Therefore, we are unable to find the $(m+1)$-th orthonormal eigenvector associated with $\lambda_0$, and the iteration stops.
+Therefore, we are unable to find the $(m+1)$-th orthonormal eigenvector associated with $\lambda_0$, and the repetition stops.
 \end{proof}
@@
 \begin{defn}[Diagonalization]\label{def:diagonalization}
-	Let  $\mA \in \bbR^{n \times n}$ be a square matrix with $n$ independent eigenvectors. We are able to  diagonalize the matrix $\mA$ as
+	Let  $\mA \in \bbR^{n \times n}$ be a square matrix with $n$ independent eigenvectors. We diagonalize the matrix $\mA$ as
 	\begin{equation}\label{eq:defdiag}
@@
 \begin{thm}[Diagonalization of symmetric matrices]
-Let $\mA$ be a symmetric matrix. The matrix $\mA$ is diagonalizable.
+Symmetric matrices are diagonalizable.
 \end{thm}
@@
 \begin{proof}
-	By Theorems~\ref{thm:symmortho} and Theorem~\ref{thm:symmrepeat}, for any symmetric matrix $\mA \in \bbR^{n \times n}$, $\mA$ have $n$ orthogonal eigenvectors with possibly repeated eigenvalues.
+	By Theorems~\ref{thm:symmortho} and \ref{thm:symmrepeat}, for any symmetric matrix $\mA \in \bbR^{n \times n}$, $\mA$ have $n$ orthogonal eigenvectors with possibly repeated eigenvalues.
 Therefore, the matrix $\mA$   is able to be diagonalized as equation~\eqref{eq:defdiag}.
@@
 	Let $\mA \in \bbR^{n \times n}$ be  a square matrix with eigenvalues $\lambda_1,...,\lambda_n$, and  $n$ independent eigenvectors associated with the eigenvalues.
-The power of matrix $\mA^k \rightarrow 0$ as $k \rightarrow +\infty$ if and only if $|\lambda_i| < 1$, for all $i \in [n]$.
+The power of the matrix $\mA^k \rightarrow 0$ as $k \rightarrow +\infty$ if and only if $|\lambda_i| < 1$, for all $i \in [n]$.
 \end{thm}
@@
 Since $\mA$ has $n$ independent eigenvectors, the matrix $\mA$ is diagonalizable.
-	Let $\mA =\mS^{-1} \Lambda \mS$, where $\Lambda, \mS$ are of the same definitions in Definition~\ref{def:diagonalization}. For all integer $k\geq 0$, we have $\mA^k = \mS^{-1} \Lambda^{k} \mS$. Then,
+	Let $\mA =\mS^{-1} \Lambda \mS$, where the definitions of $\Lambda, \mS$ are the same as that in Definition~\ref{def:diagonalization}.
+For all integer $k\geq 0$, we have $\mA^k = \mS^{-1} \Lambda^{k} \mS$.
+Then,
 	\[ \lim_{k \rightarrow +\infty} \mA^k = \lim_{k \rightarrow +\infty} \mS^{-1} \Lambda^{k} \mS  =  \lim_{k \rightarrow +\infty} \mS^{-1} \begin{bmatrix}

### 2020-07-23 skills/math/linear_algebra/linear_algebra_summary_p2.tex
upload

@@
 \algnewcommand\OUTPUT{\item[\algorithmicoutput]}
+
+\DeclareMathOperator{\var}{var}
+\DeclareMathOperator{\rank}{rank}
+\DeclareMathOperator{\vol}{vol}
+\DeclareMathOperator{\tr}{tr}
 
@@
 There are four subspaces related to $\mA\colon$ column space $C(\mA)$, row space $C(\mA^T)$, nullspace $N(\mA)$, and left nullspace $N(\mA^T)$.
-Suppose the rank of $\mA$ is $rank(\mA) = r$.
+Suppose the rank of $\mA$ is $\rank(\mA) = r$.
 The dimensions of these subspaces are:
-\[ dim(C(\mA)) = dim(C(\mA^T)) = r,\quad  dim(N(A)) = n-r,\quad dim(N(\mA^T)) = m-r. \]
+\[ \dim(C(\mA)) = \dim(C(\mA^T)) = r,\quad  \dim(N(A)) = n-r,\quad \dim(N(\mA^T)) = m-r. \]
 
@@
 
+For notational convenience, we use the shorthands $\mA$ for the column space $C(\mA)$ and $\mA^{\perp}$ for the nullspace $N(\mA)$ in the rest of this section.
+Though $\mA$ usually denotes a matrix, the meaning of $\mA$ should be clear given the context.
+For example, $\rank(\mA)$ refers the rank of matrix $\mA$, while $\dim(\mA)$ refers to the dimension of the column space $C(\mA)$.
+Specifically, for a vector $v$, $v \in \mA$ means that the vector $v$ belongs to the column space $C(\mA)$.
+
+\vspace{0.2cm}
+Given a matrix $\mA \in \bbR^{m \times n}$, we also rewrite the Theorem~\ref{thm:ortho} as
+\[ \mA^T \perp \mA^{\perp} \quad \text{and} \quad \mA \perp \mA^{T,\perp}. \]
+
 \begin{thm}[Relationship between $\mA^T \mA$ and $\mA$]\label{thm:aa}Let $\mA \in \bbR^{m \times n}$ be a matrix. We have
 	\begin{equation}
-		N(\mA^T \mA) = N(\mA) \quad \text{and} \quad rank(\mA^T \mA) = rank(\mA).
+		(\mA^T \mA)^{\perp} = \mA^{\perp} \quad \text{and} \quad \rank(\mA^T \mA) = \rank(\mA).
 	\end{equation}
@@
 \begin{proof}
-	First, we prove that $N(\mA^T \mA) = N(\mA)$.
+	First, we prove that $(\mA^T \mA)^{\perp} = \mA^{\perp}$.
 
-	On one hand, if the vector $x \in N(\mA)$, then $\mA x = 0$.
+	On one hand, if the vector $x \in \mA^{\perp}$, then $\mA x = 0$.
 Multiplying the matrix $\mA^T$ on both sides of the equation, we have $ \mA^T  \mA x = 0$.
-Therefore, for any $x \in N(\mA)$, the vector $x \in N(\mA^T \mA)$.
+Therefore, for any $x \in \mA^{\perp}$, the vector $x \in (\mA^T \mA)^{\perp}$.
 	\iffalse
 
-	On the other hand, we  prove  by contradiction that for any $x \in N(\mA^T \mA)$ the vector $x \in N(\mA)$.
+	On the other hand, we  prove  by contradiction that for any $x \in (\mA^T \mA)^{\perp}$ the vector $x \in \mA^{\perp}$.
 
-	Suppose there is a vector $x \in N(\mA^T \mA^T)$ but $x \notin N(\mA)$. We have
+	Suppose there is a vector $x \in (\mA^T \mA^T)^{\perp}$ but $x \notin \mA^{\perp}$. We have
 	\begin{equation}\label{eq:aaa}
@@
 	\end{equation} 
-	By the first equality in \eqref{eq:aaa}, $b \in C(\mA)$, and by the third equation in \eqref{eq:aaa}, $ b \in N(\mA^T)$.
-This contradicts the Theorem~\ref{thm:ortho} that $C(\mA) \perp N(\mA^T)$.
-Therefore, for any $x \in N(\mA^T \mA)$, the vector $x \in N(\mA)$.
+	By the first equality in \eqref{eq:aaa}, $b \in \mA$, and by the third equation in \eqref{eq:aaa}, $ b \in \mA^{T,\perp}$.
+This contradicts the Theorem~\ref{thm:ortho} that $\mA \perp \mA^{T,\perp}$.
+Therefore, for any $x \in (\mA^T \mA)^{\perp}$, the vector $x \in \mA^{\perp}$.
 
@@
 
-	On the other hand, if the vector $x \in N(\mA^T \mA^T)$, we have
+	On the other hand, if the vector $x \in (\mA^T \mA)^{\perp}$, we have
 	\[ \mA^T \mA x = 0.\]
-	Then, the vector $\mA x \in N(\mA^T)$.
-By the definition of matrix-to-vector multiplication, we also have $\mA x \in C(\mA)$.
-By the Theorem~\ref{thm:ortho}, the subspaces $N(\mA^T) \perp C(\mA)$.
-Therefore, $\mA x = 0$, and thus $x \in N(\mA)$.
+	Then, the vector $\mA x \in  \mA^{T,\perp}$.
+By the definition of matrix-to-vector multiplication, we also have $\mA x \in \mA$.
+By the Theorem~\ref{thm:ortho}, the subspaces $\mA^{T,\perp} \perp \mA$.
+Therefore, $\mA x = 0$, and we obtain that $x \in \mA^{\perp}$.
 	\vspace{0.2cm}
 
-	Next, we prove that $rank(\mA^T \mA) = rank(\mA)$.
+	Next, we prove that $\rank(\mA^T \mA) = \rank(\mA)$.
 
-	Given $N(\mA^T \mA) = N(\mA)$, we have $rank(\mA^T \mA) = n - dim(N(\mA^T \mA)) = n - dim(N(\mA)) = rank(\mA)$.
+	Given $(\mA^T \mA)^{\perp} = \mA^{\perp}$, we have $\rank(\mA^T \mA) = n - \dim((\mA^T \mA)^{\perp}) = n - \dim(\mA^{\perp}) = \rank(\mA)$.
 \end{proof}
@@
 \begin{proof}
-Suppose $\mA \in \bbR^{m \times n}$ is a matrix with independent columns; i.e., $rank(\mA) = n$.
-By Theorem~\ref{thm:aa}, $rank(\mA^T \mA) = rank(\mA) = n$.
+Suppose $\mA \in \bbR^{m \times n}$ is a matrix with independent columns; i.e., $\rank(\mA) = n$.
+By Theorem~\ref{thm:aa}, $\rank(\mA^T \mA) = \rank(\mA) = n$.
 Since $\mA^T \mA \in \bbR^{n \times n}$ is a square matrix, $\mA^T \mA$ is invertible.
@@
 \subsection{Projections onto subspaces}
-\begin{defn}[\textit{Projection and projection matrix}]\label{def:project} Let  $x \in \bbR^{x}$ be a vector and $\mA^{n \times m}$ be a matrix. The vector $p_{C(\mA)}(x) \in C(\mA)$ that satisfies
+\begin{defn}[\textit{Projection and projection matrix}]\label{def:project} Let  $x \in \bbR^{x}$ be a vector and $\mA^{n \times m}$ be a matrix. The vector $p_{\mA}(x) \in \mA$ that satisfies
 	\begin{equation}\label{eq:project}
-		(x-p_{C(\mA)}(x)) \perp C(\mA),
+		(x-p_{\mA}(x)) \perp \mA,
 	\end{equation} 
-	is called the projection of vector $x$ onto the column space $C(\mA)$. For all $x \in \bbR^n$, the matrix  $\mP_{C(\mA)} \in \bbR^{n \times n}$  that satisfies	\begin{equation}
-		p_{C(\mA)}(x) =\mP_{C(\mA)} x,
+	is called the projection of vector $x$ onto the column space $\mA$. For all $x \in \bbR^n$, the matrix  $\mP_{\mA} \in \bbR^{n \times n}$  that satisfies	\begin{equation}
+		p_{\mA}(x) =\mP_{\mA} x,
 	\end{equation}
-	is called the projection matrix of $\mA$ from $\bbR^n$ onto the column space  $C(\mA)$.
-Note that the projection $p_{C(\mA)}(x)$ depends on the vector $x$ and the space $C(\mA)$, while the projection matrix $\mP_{C(\mA)}$ depends only on the space $C(\mA)$.
+	is called the projection matrix of $\mA$ from $\bbR^n$ onto the column space $\mA$.
+Note that the projection $p_{\mA}(x)$ depends on the vector $x$ and the space $\mA$, while the projection matrix $\mP_{\mA}$ depends on the space $\mA$ only.
 \end{defn}
 
-\begin{prop}[Projection matrix of $C(\mA)$]\label{prop:project}
-	Let  $x \in \bbR^{n}$ be a vector and $\mA^{n \times m}$ be a matrix with independent columns. The projection matrix of $\mA$ from $\bbR^n$ onto the column space  $C(\mA)$ is 
+\begin{prop}[Projection matrix of $\mA$]\label{prop:project}
+	Let  $x \in \bbR^{n}$ be a vector and $\mA^{n \times m}$ be a matrix with independent columns. The projection matrix of $\mA$ from $\bbR^n$ onto the column space  $\mA$ is 
 	\begin{equation}
-		\mP_{C(\mA)} =  \mA (\mA^T \mA)^{-1} \mA^T.
+		\mP_{\mA} =  \mA (\mA^T \mA)^{-1} \mA^T.
 	\end{equation}
@@
 Let $x_0$ be an arbitrary vector of $\bbR^n$.
-By Definition~\ref{def:project}, the projection $p_{C(\mA)}(x_0) \in C(\mA)$.
-Then, there exists a vector $v(x_0) \in \bbR^m$ such that $p_{C(\mA)}(x_0) = \mA  v(x_0)$.
+By Definition~\ref{def:project}, the projection $p_{\mA}(x_0) \in \mA$.
+Then, there exists a vector $v(x_0) \in \bbR^m$ such that $p_{\mA}(x_0) = \mA  v(x_0)$.
 By equation~\eqref{eq:project}, we have
 \begin{equation}\label{eq:projectca}
-	A^T (x_0 - \mA v(x_0)) = 0 \quad \Rightarrow  \quad v(x_0) = (\mA^T \mA)^{-1} \mA^T x_0 \quad \Rightarrow \quad p_{C(\mA)}(x_0)  = \mA (\mA^T \mA)^{-1} \mA^T  x_0. 
+	A^T (x_0 - \mA v(x_0)) = 0 \quad \Rightarrow  \quad v(x_0) = (\mA^T \mA)^{-1} \mA^T x_0 \quad \Rightarrow \quad p_{\mA}(x_0)  = \mA (\mA^T \mA)^{-1} \mA^T  x_0. 
 \end{equation}
 Since $x_0$ is arbitrary, the equation~\eqref{eq:projectca} still holds after replacing all the $x_0$ by $x$, for all $x \in \bbR^n$.
-Therefore, the matrix $\mP_{C(\mA)} =  \mA (\mA^T \mA)^{-1} \mA^T$ is the projection matrix of $\mA$ from $\bbR^n$ onto the column space $C(\mA)$.
+Therefore, the matrix $\mP_{\mA} =  \mA (\mA^T \mA)^{-1} \mA^T$ is the projection matrix of $\mA$ from $\bbR^n$ onto the column space $\mA$.
 \end{proof}
@@
 \begin{thm}[Properties of projection matrix]\label{thm:projectm}
-Let $\mA \in \bbR^{m \times n}$ be a matrix, and $\mP_{C(\mA)}$ be the projection matrix of $\mA$ onto the column space $C(\mA)$. Then,
+Let $\mA \in \bbR^{m \times n}$ be a matrix, and $\mP_{\mA}$ be the projection matrix of $\mA$ onto the column space $\mA$. Then,
 	\begin{equation}\label{eq:projectm}
-		\mP_{C(\mA)}^T = \mP_{C(\mA)}\quad \text{and}\quad  \mP_{C(\mA)}^2 = \mP_{C(\mA)}.
+		\mP_{\mA}^T = \mP_{\mA}\quad \text{and}\quad  \mP_{\mA}^2 = \mP_{\mA}.
 	\end{equation}
@@
 \begin{proof}
-	By Proposition~\ref{prop:project}, the projection matrix is $\mP_{C(\mA)} =   \mA(\mA^T \mA)^{-1} \mA^T$. Then,
-	\[  \mP_{C(\mA)}^T = (\mA(\mA^T \mA)^{-1} \mA^T )^T = \mA (\mA^T \mA)^{-1} \mA^T = \mP_{C(\mA)}. \]
+	By Proposition~\ref{prop:project}, the projection matrix is $\mP_{\mA} =   \mA(\mA^T \mA)^{-1} \mA^T$. Then,
+	\[  \mP_{\mA}^T = (\mA(\mA^T \mA)^{-1} \mA^T )^T = \mA (\mA^T \mA)^{-1} \mA^T = \mP_{\mA}. \]
 Similarly, we have
-	\[  \mP_{C(\mA)}^2 = \mP_{C(\mA)}^T \mP_{C(\mA)} =  \mA (\mA^T \mA)^{-1} \mA^T  \mA (\mA^T \mA)^{-1} \mA^T =  \mA (\mA^T \mA)^{-1} \mA^T = \mP_{C(\mA)}. \]
+	\[  \mP_{\mA}^2 = \mP_{\mA}^T \mP_{\mA} =  \mA (\mA^T \mA)^{-1} \mA^T  \mA (\mA^T \mA)^{-1} \mA^T =  \mA (\mA^T \mA)^{-1} \mA^T = \mP_{\mA}. \]
 \end{proof}
 
-\begin{cor}[Projection onto $N(\mA^T)$]\label{cor:iprojectm}
-	Suppose $\mP_{C(\mA)}$ is the projection matrix of $\mA$ from $\bbR^n$ onto the column space $C(\mA)$.
-Then $\mI-\mP_{C(\mA)}$ is the projection matrix of $\mA$ from $\bbR^n$ onto the left nullspace $N(\mA^T)$.
+\begin{cor}[Projection onto $\mA^{T,\perp}$]\label{cor:iprojectm}
+	Suppose $\mP_{\mA}$ is the projection matrix of $\mA$ from $\bbR^n$ onto the column space $\mA$.
+Then $\mI-\mP_{\mA}$ is the projection matrix of $\mA$ from $\bbR^n$ onto the left nullspace $\mA^{T,\perp}$.
 \end{cor}
@@
 \begin{proof}
-	By Definition~\ref{def:project}, for all $x \in \bbR^n$, $\mP_{C(\mA)} x$ is the projection of $x$ from $\bbR^n$ to $C(\mA)$. Then,
-	\[ x - \mP_{C(\mA)} x \perp C(\mA) \quad \Rightarrow \quad (\mI-\mP_{C(\mA)})x \perp C(\mA). \]
-	By Theorem~\ref{thm:ortho}, we have $C(\mA) \perp N(\mA^T)$. Since $rank(C(\mA)) + rank(N(\mA^T)) = n$, $C(\mA)  \cup N(\mA^T) = \bbR^n$. Then, we have
-	\[ (\mI-\mP_{C(\mA)})x \in N(\mA^T).\]
-	Since $\mP_{C(\mA)} x$ is the projection of $x$ from $\bbR^n$ to $C(\mA)$,  the vector $\mP_{C(\mA)} x \in C(\mA)$, for all  $x \in \bbR^n$. We have,
-	\[ x -  (\mI-\mP_{C(\mA)})x = \mP_{C(\mA)} x \in C(\mA) \quad \Rightarrow \quad  x -  (\mI-\mP_{C(\mA)})x \perp N(\mA^T).\]
-	Therefore, for all $x \in \bbR^n$, $(\mI-\mP_{C(\mA)})x$ is the projection of $x$ from $\bbR^n$ onto the left nullspace $N(\mA^T)$, and thus $\mI-\mP_{C(\mA)}$ is the projection matrix of $\mA$ from $\bbR^n$ onto $N(\mA^T)$.
+	By Definition~\ref{def:project}, for all $x \in \bbR^n$, $\mP_{\mA} x$ is the projection of $x$ from $\bbR^n$ onto $\mA$. Then,
+	\[ x - \mP_{\mA} x \perp \mA \quad \Rightarrow \quad (\mI-\mP_{\mA})x \perp \mA. \]
+	By Theorem~\ref{thm:ortho}, we have $\mA \perp \mA^{T,\perp}$. Since $\dim(\mA) + \dim(\mA^{T,\perp}) = n$, we have $\mA  \cup \mA^{T,\perp} = \bbR^n$. Then, we have
+	\[ (\mI-\mP_{\mA})x \in  \mA^{T,\perp}.\]
+	Since $\mP_{\mA} x$ is the projection of $x$ from $\bbR^n$ to $\mA$,  the vector $\mP_{\mA} x \in \mA$, for all  $x \in \bbR^n$. We have,
+	\[ x -  (\mI-\mP_{\mA})x = \mP_{\mA} x \in \mA \quad \Rightarrow \quad  x -  (\mI-\mP_{\mA})x \perp \mA^{T,\perp}.\]
+	Therefore, for all $x \in \bbR^n$, $(\mI-\mP_{\mA})x$ is the projection of $x$ from $\bbR^n$ onto the left nullspace $\mA^{T,\perp}$, and $\mI-\mP_{\mA}$ is the projection matrix of $\mA$ from $\bbR^n$ onto $\mA^{T,\perp}$.
 \end{proof}
@@
 \[  \hat \beta_{LS} = \argmin_{\beta \in \bbR^{k+1}} \onorm{y - \mX \beta}^2, \]
-where $\onorm{\cdot}$ is the Euclidean norm.  The projection of $y$ onto $C(\mX)$ minimizes the Euclidean distance from $y$ to the column space $C(\mX)$. Hence,   we consider the vector $\mX 
-\hat \beta_{LS}$ as the projection of $y$ onto the column space $C(\mX)$. Therefore, we may use projection to solve the minimization problem.
+where $\onorm{\cdot}$ is the Euclidean norm.  The projection of $y$ onto $\mX$ minimizes the Euclidean distance from $y$ to the column space $\mX$. Hence,   we consider the vector $\mX 
+\hat \beta_{LS}$ as the projection of $y$ onto the column space $\mX$. Therefore, we may use projection to solve the minimization problem.
 
@@
 When $m = n$, $\mQ$ is an orthogonal matrix and $\mQ^T = \mQ^{-1}$.
-The projection matrix of $\mQ$ onto the column space $C(\mQ)$, denoted $\mP_{C(\mQ)}$,  becomes $\mP_{C(\mQ)} = \mI_m$.
+The projection matrix of $\mQ$ onto the column space $C(\mQ)$, denoted $\mP_{\mQ}$,  becomes $\mP_{\mQ} = \mI_m$.
 
 \begin{defn}[\textit{Gram-Schimidt Process and QR decomposition}]\label{gram-schimidt}
-Let $\mA \in \bbR^{m \times n}$ be a matrix with $rank(\mA) = n$, and  $a_i \in \bbR^{m}$ be the $i$-th column of matrix $\mA$, for all $i \in [n]$.
-Gram-Schimidt process finds the orthonormal basis for $C(\mA)$ as follows,
+Let $\mA \in \bbR^{m \times n}$ be a matrix with $\rank(\mA) = n$, and  $a_i \in \bbR^{m}$ be the $i$-th column of matrix $\mA$, for all $i \in [n]$.
+Gram-Schimidt process finds the orthonormal basis for $\mA$ as follows,
 \begin{align}
@@
 \end{align}
- The vectors $e_1,...,e_n$ are orthonormal basis of $C(\mA)$. Based on the basis obtained by Gram-Schimidit, we decompose the matrix $\mA$ as follows,
+ The vectors $e_1,...,e_n$ are orthonormal basis of $\mA$. Based on the basis obtained by Gram-Schimidit, we decompose the matrix $\mA$ as follows,
  \begin{equation}\label{eq:qr}
@@
 	\[ P = \{ a_1 v_1 + \cdots + a_n v_n \colon 0 \leq a_1,...,a_n \leq 1 \}. \]
-	 We use $vol(P)$ to denote the volume of the parallelepiped  $P$. 
+	 We use $\vol(P)$ to denote the volume of the parallelepiped  $P$. 
 \end{defn}
@@
 The absolute determinant of a square matrix $\mA$ is the volume of parallelepiped determined by the rows of $\mA$; i.e.,
-	\[ |\det(\mA)| = vol(P(\mA)), \]
+	\[ |\det(\mA)| = \vol(P(\mA)), \]
 	where $P(\mA)$ is the parallelepiped determined by the rows of $\mA$.
@@
 Let $\mA = \entry{a_{ij}} \in \bbR^{n \times n}$ be a square matrix. The trace of $\mA$ is defined as
-\[ tr(\mA) = \sum_{i = 1}^n a_{ii}.\] 
+\[ \tr(\mA) = \sum_{i = 1}^n a_{ii}.\] 
 \end{defn}
@@
 	\begin{equation}\label{def:charapoly}
-		 P(\lambda)  = (-1)^n \lambda^n +  (-1)^n  tr(\mA)\lambda^{n-1} + \cdots + (-1)^n \det(\mA). 
+		 P(\lambda)  = (-1)^n \lambda^n +  (-1)^n  \tr(\mA)\lambda^{n-1} + \cdots + (-1)^n \det(\mA). 
 	\end{equation}
@@
 	\begin{equation}
-		\sum_{i=1}^n \lambda_i = tr(\mA)\quad \text{and} \quad \prod_{i=1}^n \lambda_i = \det(\mA).
+		\sum_{i=1}^n \lambda_i = \tr(\mA)\quad \text{and} \quad \prod_{i=1}^n \lambda_i = \det(\mA).
 	\end{equation} 
@@
 In equation~\eqref{eq:poly}, the coefficient for $\lambda^{n-1}$ is equal to $(-1)^n\sum_{i=1}^n \lambda_i$.
-Comparing the equation~\eqref{eq:poly} with Definition~\ref{def:charac}, we obtain $tr(\mA) = \sum_{i=1}^n \lambda_i$.
+Comparing the equation~\eqref{eq:poly} with Definition~\ref{def:charac}, we obtain $\tr(\mA) = \sum_{i=1}^n \lambda_i$.
 Similarly, the constant term in equation~\eqref{eq:poly} is $(-1)^n \prod_{i=1}^n \lambda_i $, while the constant term in Defintion~\ref{def:charac} is $(-1)^n \det(\mA)$.

### 2020-07-23 skills/math/tail_bound/tailbound_solution.tex
upload

@@
 \algnewcommand\OUTPUT{\item[\algorithmicoutput]}
+\usepackage{dsfont}
+\usepackage{colonequals}
 
 \mathtoolsset{showonlyrefs=true}
+
+\DeclareMathOperator{\var}{var}
+\DeclareMathOperator{\subG}{subG}
 
@@
 \begin{thm}[Markov's inequality]\label{thm:markov}
-	Suppose $X \geq 0$ is a random variable  with finite mean, we have
+	Let $X \geq 0$ be a random variable  with finite mean. We have
 	\begin{equation}\label{eq:thmmarkov}
-		\bbP(X\geq t) \leq \frac{E[X]}{t},\quad \text{ for all } \ t > 0.
+		\bbP(X\geq t) \leq \frac{\bbE[X]}{t},\quad \text{ for all } \ t > 0.
 	\end{equation}  
@@
 \begin{thm}[Chebyshev's inequality]\label{thm:cheb}
-	Suppose $X \geq 0$ is a random variable  with finite mean $\mu$ and finite variance, we have
+	Let $X \geq 0$ be a random variable  with finite mean $\mu$ and finite variance. We have
 	\begin{equation}\label{eq:cheby}
-		\bbP(|X - \mu |\geq t) \leq \frac{var(X)}{t^2},\quad \text{ for all } \ t > 0.
+		\bbP(|X - \mu |\geq t) \leq \frac{\var(X)}{t^2},\quad \text{ for all } \ t > 0.
 	\end{equation}  
@@
 \begin{thm}[Markov's inequality for polynomial moments]\label{thm:polymoment}
-	Suppose the random variable $X$ has a central moment of order $k$. Applying Markov's inequality to the random variable $|X - \mu|^k$ yields
+	Let $X$ be a random variable. Suppose that the order $k$ central moment of $X$ exists. Applying Markov's inequality to the random variable $|X - \mu|^k$ yields
 	\begin{equation}\label{eq:poly}
-		\bbP(|X - \mu |\geq t) \leq \frac{\bbE[|X-\mu|^k]}{t^k},\quad \text{ for all } \ t > 0.
+		\bbP(|X - \mu |\geq t) \leq \frac{\bbE\left[|X-\mu|^k\right]}{t^k},\quad \text{ for all } \ t > 0.
 	\end{equation}  
@@
 \begin{thm}[Chernoff bound]\label{thm:chernoff}
-	Suppose the random variable $X$ has a moment generating function in the neighborhood of 0; i.e.\, $\varphi_X(\lambda) = \bbE[e^{\lambda X}] < +\infty,$ for all $\lambda \in (-b,b)$ with some $b>0$.
+ Let $X$ be a random variable.
+Suppose that the moment generating function of $X$, denoted $\varphi_X(\lambda)$, exists in the neighborhood of 0; i.e., $\varphi_X(\lambda) = \bbE[e^{\lambda X}] < +\infty,$ for all $\lambda \in (-b,b)$ with some $b>0$.
 Applying Markov's inequality to the random variable $Y = e^{\lambda(X- \mu)}$ yields
 	\begin{equation}\label{eq:chernoff1}
-		\bbP((X - \mu) \geq t) \leq \frac{\bbE[e^{\lambda (X-\mu)}]}{e^{\lambda t}}, \quad \text{ for all } \lambda\in (-b,b).
+		\bbP((X - \mu) \geq t) \leq \frac{\bbE\left[e^{\lambda (X-\mu)}\right]}{e^{\lambda t}}, \quad \text{ for all } \lambda\in (-b,b).
 	\end{equation}
-	Optimizing the choice of $\lambda$ for the tightest bound yields the Chernoff bound
+	Optimizing the choice of $\lambda$ for the tightest bound, we obtain the Chernoff bound
 	\begin{equation}\label{eq:chernoff2}
-		\bbP((X - \mu) \geq t) \leq \inf_{\lambda \in [0,b)} \frac{\bbE[e^{\lambda (X-\mu)}]}{e^{\lambda t}}.
+		\bbP((X - \mu) \geq t) \leq \inf_{\lambda \in [0,b)} \frac{\bbE\left[e^{\lambda (X-\mu)}\right]}{e^{\lambda t}}.
 	\end{equation}
@@
 \begin{thm}[Hoeffding bound for bounded variable]\label{thm:boundvar}
-Consider a random variable $X$ with mean $\mu = \bbE(X)$. Assume that $X$ is bounded and $X \in [a,b]$ almost surely, where $a, b$ are two constants. Then, for any $\lambda \in \bbR$, we have
+Let $X$ be a random variable with mean $\mu = \bbE(X)$. Suppose that $X \in [a,b]$ almost surely, where $a\leq b\in\bbR$ are two constants. Then, we have
 \begin{equation}
-	\bbE[e^{\lambda X}] \leq e^{\frac{s(b-a)^2}{8}}.
+	\bbE[e^{\lambda X}] \leq e^{\frac{s(b-a)^2}{8}}, \quad \text{for all } \lambda \in \bbR.
 \end{equation}
-Particularly, the variable $X \sim subG(\frac{(b-a)^2}{4})$.
+Consequently, the variable $X \sim \subG\left(\frac{(b-a)^2}{4}\right)$.
 \end{thm}
@@
 \begin{thm}[Moment of sub-Gaussian variable]\label{thm:momsubg}
-Let $X \sim sG(\sigma^2)$. For all integer $k \geq 1$, we have
+Let $X \sim \subG(\sigma^2)$. For all integer $k \geq 1$, we have
 \begin{equation}\label{eq:thmmomsubg}
@@
 \begin{thm}[One-sided Bernstein’s inequality]\label{thm:oneber}
- Let $X$ be a random variable. If $X \leq b$ almost surely, then 
+ Let $X$ be a random variable. Suppose $X \leq b$ almost surely. We have 
 \begin{equation}\label{eq:thmoneber1}
-	\bbE[e^{\lambda (X-\bbE[X])}] \leq exp\left\{ \frac{\lambda^2 \bbE[X^2]/2}{1 - b\lambda /3}  \right\}, \quad \text{for all } \lambda \in [0, 3/b).
+	\bbE\left[e^{\lambda (X-\bbE[X])}\right] \leq exp\left\{ \frac{\lambda^2 \bbE[X^2]/2}{1 - b\lambda /3}  \right\}, \quad \text{for all } \lambda \in [0, 3/b).
 \end{equation}
-Consequently, suppose there are $n$ independent variables $X_i \leq b$ almost surely, for all $i \in [n]$. We have
+Consequently, let $X_i$ be independent variables, and $X_i \leq b$ almost surely, for all $i \in [n]$. We have
 \begin{equation}\label{eq:thmoneber2}
-	\bbP[\sum_{i = 1}^n (X_i - \bbE[X_i] ) \geq n\delta ] \leq exp\left\{ - \frac{n\delta^2}{\sum_{i = 1}^n \bbE[X_i^2] /n + b \delta /3} \right\},\quad \text{ for all } \delta \geq 0.
+	\bbP\left[\sum_{i = 1}^n (X_i - \bbE[X_i] ) \geq n\delta \right] \leq exp\left\{ - \frac{n\delta^2}{\sum_{i = 1}^n \bbE[X_i^2] /n + b \delta /3} \right\},\quad \text{ for all } \delta \geq 0.
 \end{equation}  
-Particularly, suppose there are $n$ independent non-negative variables $Y_i \geq 0$, for all $i \in [n]$. The equation~\eqref{eq:thmoneber2} becomes
+Particularly, let $X_i$ be independent nonnegative variables, for all $i \in [n]$. The equation~\eqref{eq:thmoneber2} becomes
 \begin{equation}\label{eq:thmonber0}
-	\bbP[\sum_{i = 1}^n (Y_i - \bbE[Y_i] ) \leq n\delta ] \leq exp\left\{ - \frac{n\delta^2}{\sum_{i = 1}^n \bbE[Y_i^2] /n} \right\},\quad \text{ for all } \delta \geq 0.
+	\bbP\left[\sum_{i = 1}^n (Y_i - \bbE[Y_i] ) \leq n\delta \right] \leq exp\left\{ - \frac{n\delta^2}{\sum_{i = 1}^n \bbE[Y_i^2] /n} \right\},\quad \text{ for all } \delta \geq 0.
 \end{equation} 
+\end{thm}
+
+\begin{defn}[Bernstein's condition]
+Let $X$ be a random variable with mean $\mu = \bbE[X]$ and variance $ \sigma^2 = \var(X)$. We say $X$ satisfies the Bernstein's condition with parameter $b$ if
+\begin{equation}
+	\left| \bbE [(X -\mu)^k ]\right| \leq \frac{1}{2} k! \sigma^2 b^{k-2}, \quad \text{for } k =3, 4,....
+\end{equation} 
+\end{defn}
+Note that every bounded random variable $X$ satisfies the Bernstein's condition.
+
+\begin{thm}[Bernstein-type bound] For any variable $X$ satisfying the Bernstein's condition, we have
+\begin{equation}
+	\bbE\left[e^{\lambda (X-\mu)}\right] \leq \exp\left\{ \frac{\lambda^2 \sigma^2}{2(1 - b |\lambda|)} \right\},\quad \text{for all } |\lambda| \leq \frac{1}{b}, 
+\end{equation}
+and the concentration inequality
+\begin{equation}
+	\bbP\left[|X - \mu| \geq t\right] \leq 2 \exp \left\{ -\frac{t^2}{2(\sigma^2 + bt)} \right\}, \quad \text{for all }t \geq 0.
+\end{equation}
+
 \end{thm}
@@
 
-\subsection{Exercise 2.1} (Tightness of inequalities.) The Markov's and Chebyshev's inequalities can not be improved in general. 
+\subsection{Exercise 2.1} (Tightness of inequalities.) The Markov's and Chebyshev's inequalities are not able to be improved in general. 
 \begin{enumerate}
@@
 
-\item[(a)] For a given constant $t >0$, we define a variable $Y_{t} = X - t\mI_{[X\geq t]}$, where $\mI$ is the indicator function.
-Noted that $Y_{t}$ is a nonnegative variable, the Markov's inequality follows by taking expectation to $Y$,
+\item[(a)] For a given constant $t >0$, we define a variable $Y_{t} = X - t \mathds{1}{[X\geq t]}$, where $\mathds{1}$ is the indicator function.
+Note that $Y_{t}$ is a nonnegative variable.
+The Markov's inequality follows by taking the expectation to $Y_t$,
 \begin{equation}
@@
 
-\item[(b)]  Chebyshev's inequality follows by applying Markov's inequality to the non-negative random variable $Z = {(X - \bbE[X])^2}$.
-Similarly as part (a), given a constant $t > 0$, the variable $Z = {(X - \bbE[X])^2}$ with distribution $\bbP(Z \in \{ 0,t^2 \}) = 1$ attains the equality of the Markov's inequality for $Z$ and the Chebyshev's inequality for $X$.
+\item[(b)]  Chebyshev's inequality follows by applying Markov's inequality to the nonnegative random variable $Z = {(X - \bbE[X])^2}$.
+As in part (a), given a constant $t > 0$, the variable $Z = {(X - \bbE[X])^2}$ with distribution $\bbP(Z \in \{ 0,t^2 \}) = 1$ attains the equality of the Markov's inequality for $Z$.
+Consequently, the variable $X$ attains the equality of the Chebyshev's inequality for $X$.
 By transformation, the distribution of $X$ satisfies the followings formula,
@@
 	\begin{cases}
-		p & \text{if } x = c,\\
-		\frac{1-p}{2} & \text{if } x = c-t \text{ or } x = c+t, \\
-		0 & \text{ otherwise },
+		\quad p & \text{if } x = c,\\
+		\quad \frac{1-p}{2} & \text{if } x = c-t \text{ or } x = c+t, \\
+		\quad 0 & \text{ otherwise },
 	\end{cases}
@@
 \begin{rmk}[Tightness of Markov's inequality]
-	Only a few variables can attain the equalities in Markov's and Chebyshev's inequalities. In research, we should pay attention to the concentration bounds tighter than Markov's inequality.
+	Only a few variables attain the equalities in Markov's and Chebyshev's inequalities. In research, we should pay attention to the concentration bounds tighter than Markov's inequality.
 \end{rmk}
@@
 	and \begin{equation}\label{eq:e22b}
-		\phi(z)(\frac{1}{z} - \frac{1}{z^3} ) \leq \bbP(Z \geq z) \leq \phi(z)(\frac{1}{z} - \frac{1}{z^3}+ \frac{3}{z^5} ), \quad \text{for all }z>0.
+		\phi(z)\left(\frac{1}{z} - \frac{1}{z^3} \right) \leq \bbP(Z \geq z) \leq \phi(z)\left(\frac{1}{z} - \frac{1}{z^3}+ \frac{3}{z^5} \right), \quad \text{for all }z>0.
 	\end{equation}
@@
 
-The pdf of standard normal distribution is
-\[  \phi(z) = \frac{1}{\sqrt{2\pi}} \exp(-\frac{z^2}{2}  ).\]
+The pdf of the standard normal distribution is
+\[  \phi(z) = \frac{1}{\sqrt{2\pi}} \exp\left(-\frac{z^2}{2}  \right).\]
 
-The equation~\eqref{eq:e22a} follows by taking the derivative of $\phi(z)$.
-\[\quad \phi'(z) = -z \frac{1}{\sqrt{2\pi}} \exp(-\frac{z^2}{2} ) = -z \phi(z). \] 
+The equation~\eqref{eq:e22a} follows by taking the derivative of $\phi(z)$. Specifically,
+\[\quad \phi'(z) = -z \frac{1}{\sqrt{2\pi}} \exp\left(-\frac{z^2}{2} \right) = -z \phi(z). \] 
 	Next, we prove the equation~\eqref{eq:e22b}.
@@
 	\end{equation}
-	where the second equality follows by the equation~\eqref{eq:e22b}. Applying the equation~\eqref{eq:e22b} to the last term in equation~\eqref{eq:e22pro1} yields
+	where the second equality follows by the equation~\eqref{eq:e22a}. Applying the equation~\eqref{eq:e22a} to the last term in equation~\eqref{eq:e22pro1} yields
 	\begin{equation}\label{eq:e22pro2}
@@
 	\end{equation}
-	Plugging the equation~\eqref{eq:e22pro2} into the equation~\eqref{eq:e22pro1} , we have $\bbP(Z \geq z) \geq \phi(z)(\frac{1}{z}  - \frac{1}{z^3} )$.
-On the other hand, applying the equation~\eqref{eq:e22b} again to the equation~\eqref{eq:e22pro2} yields
+	Plugging the equation~\eqref{eq:e22pro2} into the equation~\eqref{eq:e22pro1}, we obtain $\bbP(Z \geq z) \geq \phi(z)\left(\frac{1}{z}  - \frac{1}{z^3} \right)$.
+Applying the equation~\eqref{eq:e22a} again to the equation~\eqref{eq:e22pro2} yields
 	\begin{equation}\label{eq:e22pro3}
@@
 
-	Combing equations~\eqref{eq:e22pro1}, \eqref{eq:e22pro2}, and  \eqref{eq:e22pro3}, we have $\bbP(Z \geq z) \leq \phi(z)(\frac{1}{z}  - \frac{1}{z^3} + \frac{3}{z^5} ).$
+	Combing equations~\eqref{eq:e22pro1}, \eqref{eq:e22pro2} and  \eqref{eq:e22pro3}, we obtain $\bbP(Z \geq z) \leq \phi(z)\left(\frac{1}{z}  - \frac{1}{z^3} + \frac{3}{z^5} \right).$
 \end{proof}
@@
 \begin{lem}[Polynomial bound and Chernoff bound]\label{lem:momcher}
-	Let $X \geq 0$ be a non-negative variable.
+	Let $X \geq 0$ be a nonnegative variable.
 Suppose that the moment generating function of $X$, denoted $\varphi_{X}(\lambda)$, exists in the neighborhood of $\lambda = 0$.
@@
 		\[   \bbE [e^{\lambda X} ]= \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} < +\infty,\quad \text{ for all } \lambda \in (0,b). \]
-	Therefore, the moment  $\bbE[|X|^k]$ exists for all $ k \in \mathbb{Z_+}$. Applying power series~\eqref{eq:power} to the right hand side of equation~\eqref{eq:lempoly} yields
+	 Hence, the moment  $\bbE[|X|^k]$ exists, for all $ k \in \mathbb{Z_+}$. Applying power series~\eqref{eq:power} to the right hand side of equation~\eqref{eq:lempoly} yields
 	\begin{equation}\label{eq:powertopoly}
@@
 	\end{equation}
-	Therefore, the equation~\eqref{eq:lempoly} follows by combining the equation~\eqref{eq:powertopoly} with equation~\eqref{eq:cauchy3}.	
+	Therefore, we obtain the equation~\eqref{eq:lempoly} by combining the equation~\eqref{eq:powertopoly} with equation~\eqref{eq:cauchy3}.	
 \end{proof}
@@
 	Applying different functions $g(X)$ to the Markov's inequality leads to different bounds for the tail probability of variable $X$.
-Equation~\eqref{eq:lempoly} implies that optimized polynomial bound is tighter than the Chernoff bound, provided that the moment generating function of $X$ exsits.
+Equation~\eqref{eq:lempoly} implies that optimized polynomial bound is at least as tight as the Chernoff bound, provided that the moment generating function of $X$ exsits.
 \end{rmk}
@@
 \subsection{Exercise 2.4}\label{pro:boundvar}
-In Exercise~\ref{pro:boundvar}, we prove the Theorem~\ref{thm:boundvar},  the Hoeffding bound for a bounded variable.
+In Exercise~\ref{pro:boundvar}, we prove Theorem~\ref{thm:boundvar},  the Hoeffding bound for a bounded variable.
 
@@
 
-Suppose $X$ is a bounded random variable, where $X \in [a,b]$ almost surely, and $a 
+Let $X$ be a bounded random variable, and $X \in [a,b]$ almost surely, where $a 
 \leq b \in \bbR$ are two constants. Let $\mu = \bbE[X]$. Define the function 
@@
 
-In equation~\eqref{eq:taylor}, the term $g(0) =\log \bbE[e^{0}] = 0$. By power series \eqref{eq:power}, we calculate the first derivative $g'(\lambda)$ as following,
+In equation~\eqref{eq:taylor}, the term $g(0) =\log \bbE[e^{0}] = 0$. By power series \eqref{eq:power}, we obtain the first derivative $g'(\lambda)$ as follows,
 \begin{align}
@@
 \end{align}
-Therefore, $g'(0) = \bbE[X] = \mu$. Taking the derivative to equation~\eqref{eq:firstde}, we calculate the second-order derivative $g''(\lambda)$ as following,
+Therefore, $g'(0) = \bbE[X] = \mu$. Taking the derivative to equation~\eqref{eq:firstde}, we obtain the second-order derivative $g''(\lambda)$ as follows,
 \begin{align}
@@
 \end{align}
-We interpret the second-order derivative $g''(\lambda)$ interpreted as the variance of $X$ with the re-weighted distribution $d P' = {e^{\lambda X}}/{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$.
+We interpret the second-order derivative $g''(\lambda)$ as the variance of $X$ with the re-weighted distribution $d P' = {e^{\lambda X}}/{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$.
 Taking integral of 1 with respect to $dP'$, we have
@@
 \end{equation} 
-which implies that the function $P'$ is indeed a distribution. Under all possible re-weighted distributions, the variance of $X$ is upper bounded as following,
-\[ var(X) = var(X - \frac{a+b}{2}) \leq \bbE[(X - \frac{a+b}{2} )^2] \leq \frac{(b-a)^2}{4}, \]
-where the term $\frac{(b-a)^2}{4}$ follows by letting $X$ distribute only on the boundary $a$ and $b$.
+which implies that the function $P'$ is a valid probability distribution. Under all possible re-weighted distributions, the variance of $X$ is upper bounded as follows,
+\[ \var(X) = \var(X - \frac{a+b}{2}) \leq \bbE[(X - \frac{a+b}{2} )^2] \leq \frac{(b-a)^2}{4}, \]
+where the term $\frac{(b-a)^2}{4}$ follows by letting $X$ supported on the boundaries $a$ and $b$ only.
 Hence, the second-order derivative $g''(\lambda) \leq \frac{(b-a)^2}{4}$.
@@
 \end{equation}
-Taking the exponential on both sides of the inequality~\eqref{eq:e24}, we have
+Taking the exponentiation on both sides of the inequality~\eqref{eq:e24}, we have
 \begin{equation}\label{eq:boundvar}
@@
 \end{equation}
-	Then, $\bbE[X] = \mu$ and $var(X) \leq \sigma^2$.
+	Then, $\bbE[X] = \mu$ and $\var(X) \leq \sigma^2$.
 \end{lem}
 
-\begin{proof} By equation~\eqref{eq:subg}, the moment generating function of $X$, denoted $\varphi_X(\lambda)$ , exists in the neighborhood of $\lambda = 0$.
+\begin{proof} By equation~\eqref{eq:subg}, the moment generating function of $X$, denoted $\varphi_X(\lambda)$, exists in the neighborhood of $\lambda = 0$.
 Hence, the mean and variance of $X$ exist.
@@
 Dividing by $\lambda <0$ on both sides of equation~\eqref{eq:taylor2} and letting $\lambda \rightarrow 0^{-}$, we have $\bbE(X) \geq \mu$.
-Therefore, the mean $\bbE[X] = \mu$.
-Then, we divide ${2}/{\lambda^2}$ on both sides of equation~\eqref{eq:taylor2}.
+Therefore, we obtain the mean $\bbE[X] = \mu$.
+Then, we divide ${2}/{\lambda^2}$ on both sides of equation~\eqref{eq:taylor2}, for $\lambda \neq 0$.
 The term $\bbE[X]\lambda$ and  $\mu \lambda$ are cancelled.
-We have $ \bbE[X^2] \leq \sigma^2 + \mu^2$, and thus $var(X) \leq \bbE[X^2]  - (\bbE[X])^2 = \sigma^2$.
+We have $ \bbE[X^2] \leq \sigma^2 + \mu^2$, and thus the $\var(X) \leq \bbE[X^2]  - (\bbE[X])^2 = \sigma^2$.
 \end{proof}
 
-\paragraph{Question:} Let $\sigma^2_{min}$ denote the smallest possible $\sigma$ satisfying the inequality~\eqref{eq:subg}. Is it true that $var(X) = \sigma_{min}^2$?
+\paragraph{Question:} Let $\sigma^2_{min}$ denote the smallest possible $\sigma$ satisfying the inequality~\eqref{eq:subg}. Is it true that $\var(X) = \sigma_{min}^2$?
 
-\paragraph{Solution:} The statement that $var(X) = \sigma_{min}^2$ is not necessarily true.
-Recall the function $g(\lambda)$ in exercise~\ref{pro:boundvar}.
-By the results in exercise~\ref{pro:boundvar},  the equation~\eqref{eq:subg} is equal to
+\paragraph{Solution:} The statement that $\var(X) = \sigma_{min}^2$ is not necessarily true.
+Recall the function $g(\lambda)$ in Exercise~\ref{pro:boundvar}.
+By the results in Exercise~\ref{pro:boundvar},  the equation~\eqref{eq:subg} is equal to
 \begin{equation}\label{eq:question}
@@
 \end{equation}
-where $g'' (\lambda)$ is the variance of $X$ with with the re-weighted distribution $d P' = {e^{\lambda X}}/{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$ and $g'' (0) = var(X)$.
-Therefore, $\max_{\lambda}g''(\lambda) = \sigma_{min}^2$.
-To let the equality $var(X) = \sigma_{min}^2$ hold, we need to show that $\max_{\lambda}g''(\lambda) = g''(0)$ holds for $X$.
+where $g'' (\lambda)$ is the variance of $X$ with the re-weighted distribution defined in Exercise~\ref{pro:boundvar}.
+Therefore, we have $\max_{\lambda}g''(\lambda) = \sigma_{min}^2$.
+Note that $g''(0) = \var(X)$.
+To let the equality $\var(X) = \sigma_{min}^2$ hold, we need to show that $\max_{\lambda}g''(\lambda) = g''(0)$ holds for $X$.
 
 \vspace{0.2cm}
-However, the statement $\max_{\lambda}g''(\lambda) = g''(0)$ is not always true.  A counter example is below. 
-Consider a random variable $Y \sim Ber(1/3)$. The variance of $Y$ is $var(Y) = 2/9$. Let $\lambda = 1$. The re-weighted distribution $dP'$ is
+However, the statement $\max_{\lambda}g''(\lambda) = g''(0)$ is not necessarily true.  A counter example is below. 
+Consider a random variable $Y \sim Ber(1/3)$. The variance of $Y$ is $\var(Y) = 2/9$. Let $\lambda = 1$. The re-weighted distribution $dP'$ is
 \[ P'(Y = 0) = \frac{2}{3\bbE[ e^{Y} ]}\quad  \text{and}\quad  P'(Y = 1) = \frac{e}{3 \bbE[ e^{ Y } ]}, \quad \text{where } \bbE[ e^{Y} ] = \frac{2}{3} + \frac{e}{3}.  \]
 The variance of $Y$ with $dP'$ is ${2}/{3\bbE[ e^{Y} ]} \times  {e}/{3\bbE[ e^{Y} ]} = 0.2442 > 2/9$.
-Therefore, for this variable $Y$, we have $var(Y) <  g''(1) \leq \max_{\lambda}g''(\lambda) = \sigma^2_{min}$.
+Therefore, we have $\var(Y) <  g''(1) \leq \max_{\lambda}g''(\lambda) = \sigma^2_{min}$.
+The statement $\max_{\lambda}g''(\lambda) = g''(0)$ is not true for this variable $Y$.
 \begin{rmk}
-	Parameters of a sub-Gaussian distribution provide the exact value of the mean, $ \bbE[X] = \mu$, and an upper bound of the variance, $var(X) \leq \sigma^2$.
-For any variable $X$ whose moment generating function exists, the tail distribution of $X$ can be bounded by a sub-Gaussian distribution with a proper choice of $\sigma^2$.
+	Parameters of a sub-Gaussian distribution provide the exact value of the mean and an upper bound of the variance; i.e., $ \bbE[X] = \mu$ and $\var(X) \leq \sigma^2$.
+Suppose the moment generating function of variable  $X$ exists over the entire real interval.
+The tail distribution of $X$ is always bounded by a sub-Gaussian distribution with a proper choice of $\sigma^2$.
 \end{rmk}
@@
 \begin{lem}[Lower bounds on squared sub-Gaussians]\label{lem:lowersquare} Let $\{ X_i \}_{i=1}^n$ be an i.i.d.\ sequence of zero-mean sub-Gaussian variables with parameter $\sigma$.
-The normalized summation $Z_n = \frac{1}{n} \sum_{i=1}^n X_i^2$ satisfies
+The normalized sum $Z_n = \frac{1}{n} \sum_{i=1}^n X_i^2$ satisfies
 \begin{equation}\label{eq:lowersquare}
@@
 \end{equation}
-	The equation~\eqref{eq:lowersquare} implies that the lower tail of the summation of squared sub-Gaussian variables behave in a sub-Gaussian way.
+	The equation~\eqref{eq:lowersquare} implies that the lower tail of the sum of squared sub-Gaussian variables behaves in a sub-Gaussian way.
 \end{lem}
@@
 \begin{proof}
-	Since $X_i^2$ are i.i.d.\ nonnegative variables, we apply the equation~\eqref{eq:thmonber0} in Theorem~\ref{thm:oneber} to the variables $X_i^2$, where $i \in [n]$. Then,
+	Since $X_i^2$ are i.i.d.\ nonnegative variables, we apply the equation~\eqref{eq:thmonber0} to the variables $\{X_i^2\}_{i=1}^n$. Then, we have
 	\begin{equation}\label{eq:e26a}
-		\bbP[\sum_{i=1}^n (X_i^2 -\bbE[X_i^2]) \leq n \sigma^2 \delta] \leq exp\left\{ - \frac{n\delta^2 \sigma^4}{ \bbE[X_1^4]} \right\},  \quad \text{for all }\delta \geq 0.
+		\bbP\left[\sum_{i=1}^n (X_i^2 -\bbE[X_i^2]) \leq n \sigma^2 \delta\right] \leq \exp\left\{ - \frac{n\delta^2 \sigma^4}{ \bbE[X_1^4]} \right\},  \quad \text{for all }\delta \geq 0.
 	\end{equation}
-	By the equation~\eqref{eq:thmmomsubg} in Theorem~\ref{thm:momsubg}, we have
+	By equation~\eqref{eq:thmmomsubg}, we have
 	\begin{equation}\label{eq:e26x4}
@@
 	\end{equation}
-	Combing equations~\eqref{eq:e26a}, \eqref{eq:e26x4}, and the definition of $Z_n$, we have
+	Combing equations~\eqref{eq:e26a}, \eqref{eq:e26x4} and the definition of $Z_n$, we obtain 
 	\begin{equation}
-		\bbP[Z_n - \bbE[Z_n] \leq \sigma^2 \delta ]  \leq exp\left\{ - \frac{n\delta^2}{16} \right\},  \quad \text{for all }\delta \geq 0.
+		\bbP[Z_n - \bbE[Z_n] \leq \sigma^2 \delta ]  \leq \exp\left\{ - \frac{n\delta^2}{16} \right\},  \quad \text{for all }\delta \geq 0.
 	\end{equation}
@@
 \begin{rmk}
-	Equation~\eqref{eq:lowersquare} implies that the lower tail of the summation of squared sub-Gaussian variables behave in a sub-Gaussian way.
+	Equation~\eqref{eq:lowersquare} implies that the lower tail of the sum of squared sub-Gaussian variables behaves in a sub-Gaussian way.
 In following sections, we will show that the variable $Z_n - \bbE[Z_n]$ in Lemma~\ref{lem:lowersquare} is a sub-exponential variable.
@@
 
+\subsection{Exercise 2.7}
+\begin{lem}[Bennett’s inequality]\label{lem:bennett}
+	Let $X_1,...,X_n$ be a sequence of independent zero-mean random variables with $|X_i| \leq b$ and $\var(X_i) = \sigma_i^2$, for all $i \in [n]$. Then, we have Bennett's inequality
+	\begin{equation}\label{eq:bennett}
+		\mathbb{P}\left[\sum_{i=1}^{n} X_{i} \geq n \delta\right] \leq \exp \left\{-\frac{n \sigma^{2}}{b^{2}} h\left(\frac{b \delta}{\sigma^{2}}\right)\right\}, \quad \text{for all }\delta \geq 0,
+	\end{equation} 
+	where $\sigma^2 = \frac{1}{n} \sum_{i=1}^n \sigma_i^2$ and $h(t) \colonequals (1+t)\log (1+t) - t$ supported on $t \geq 0$.
+\end{lem}
+
+\begin{proof}
+	First, we consider the moment generating function of $X_i$, for all $i \in [n]$.
+
+	\vspace{0.2cm}
+	By the power series, for all $i \in [n]$, we have
+	\begin{align}\label{eq:e27power}
+		\bbE\left[e^{\lambda X_i}\right] = \sum_{k = 0}^{+\infty} \frac{\lambda^k \bbE[X_i^k]}{k!} 
+		= 1 + 0 + \sum_{k = 2}^{+\infty} \frac{\lambda^k \bbE[X_i^k]}{k!} \leq \exp\left\{ \sum_{k = 2}^{+\infty} \frac{\lambda^k \bbE[X_i^k]}{k!} \right\},
+	\end{align}
+	where the 0 is deduced from $\bbE[X_i] = 0$, and the last inequality follows by $1 + x \leq e^x$. By $|X_i|<b$, we obtain an upper bound of the sum term in equation~\eqref{eq:e27power} as follows
+	\begin{equation}\label{eq:e27upper}
+		\sum_{k = 2}^{+\infty} \frac{\lambda^k \bbE[X_i^k]}{k!}  \leq \sum_{k = 2}^{+\infty} \frac{\lambda^k \bbE[X_i^2 |X_i|^{k-2}] }{k!} \leq  \sum_{k = 2}^{+\infty} \frac{\lambda^k \sigma^2_i b^{k-2}}{k!} = \sigma_i^2  \left(  \frac{e^{\lambda b} - 1 -\lambda b}{b^2}   \right).
+	\end{equation}
+	Combing the equation~\eqref{eq:e27power} with equation~\eqref{eq:e27upper}, we find an upper bound of the moment generating function of $\sum_{i=1}^n X_i$. That is
+	\begin{equation}\label{eq:e27uppert}
+		\bbE\left[e^{\lambda \sum_{i = 1}^n X_i }\right] = \prod_{i = 1}^n \bbE\left[e^{\lambda X_i}\right] \leq \exp\left\{ n \sigma^2 \left( \frac{e^{\lambda b} - 1 -\lambda b}{b^2}   \right)  \right\},
+	\end{equation}
+	where $\sigma^2 = \frac{1}{n} \sum_{i=1}^n \sigma_i^2$. Combing the Chernoff bound with equation~\eqref{eq:e27uppert}, the upper tail of $\sum_{i=1}^n X_i$ follows 
+	\begin{align}
+		\mathbb{P}\left[\sum_{i=1}^{n} X_{i} \geq n \delta\right] &\leq  \exp\left\{  n \sigma^2 \left(  \frac{e^{\lambda b} - 1 -\lambda b}{b^2}   \right)  - 
+		\lambda n \delta \right\} \\
+		&= \exp \left\{ \frac{n\sigma^2}{b^2} \left( e^{\lambda b} - \lambda b- \lambda \frac{\delta b^2}{\sigma^2} - 1 \right)  \right\}, \quad \text{for all } \delta \geq 0. \label{eq:e27exp}
+	\end{align} 
+	By the first-order condition of minimization, the $\lambda$ that satisfies $\lambda b= \log\left(1 + \frac{\delta b}{\sigma^2} \right)$ minimizes the term $ \left( e^{\lambda b} - \lambda b- \lambda \frac{\delta b^2}{\sigma^2} - 1 \right) $.
+Plugging $\lambda b= \log\left(1 + \frac{\delta b}{\sigma^2} \right)$ into the equation~\eqref{eq:e27exp}, we obtain the Bennett's inequality
+	\begin{equation}\label{eq:e27bennett}
+		\mathbb{P}\left[\sum_{i=1}^{n} X_{i} \geq n \delta\right] \leq  \exp \left\{-\frac{n \sigma^{2}}{b^{2}} h\left(\frac{b \delta}{\sigma^{2}}\right)\right\}, \quad \text{for all }\delta \geq 0,
+	\end{equation}
+	where $h(t) \colonequals (1+t)\log (1+t) - t$ supported on $t \geq 0$.
+
+	\vspace{0.2cm}
+	Further, we show that the Bennett's inequality is at least as good as the Bernstein’s inequality.
+
+	\vspace{0.2cm}
+	The Bernstein's inequality for $\sum_{i=1}^n X_i$ is
+	\begin{align}\label{eq:e27bern}
+			\mathbb{P}\left[\sum_{i=1}^{n} X_{i} \geq n \delta \right] \leq \exp \left\{ \frac{-3n\delta^2}{(2 b \delta + 6 \sigma^2)} \right\}  %=\exp \left\{ -\frac{n \sigma^{2}}{ b^{2}} \frac{3b^2\delta^2}{2\sigma^2(b\delta + 6 \sigma^2)} \right\}
+			= \exp \left\{ -\frac{n \sigma^{2}}{ b^{2}} g\left(\frac{b\delta}{\sigma^2}\right) \right\}.
+	\end{align}
+	where $g(t) \colonequals  \frac{3t^2}{2t + 6} $, for all $ t \geq 0$.
+Since $g(t)\leq h(t)$ holds for all $t \geq 0$, we conclude that the Bennett's inequality~\eqref{eq:e27bennett} is at least as good as Bernstein's inequality~\eqref{eq:e27bern}.
+\end{proof}
+
+\begin{rmk} 
+	So far, we have three inequalities that control the tail of bounded variables: Hoeffding's inequality, Bernstein's inequality, and Bennett's inequality.
+Particularly, Hoeffding's inequality implies the sub-Gaussianity of bounded variables.
+As the proof for Lemma~\ref{lem:bennett} shows, Bennett's inequality is at least as good as the Bernstein's inequality, for all bounded variables.
+\end{rmk}
+
 \end{document}

### 2020-07-23 skills/math/linear_algebra/linear_algebra_summary_p2.tex
upload

@@
 	Let  $\mA$  be a symmetric  matrix with real entries only.
-Suppose $\lambda$  is a complex eigenvalue of $\mA$ with nonzero imaginary part, and $x$ is the eigenvector associated with $\lambda$.
-Since all the entries of $\mA$ are real, the eigenvector $x$ is also a complex vector with nonzero imaginary part.
+Suppose $\lambda$  is a complex eigenvalue of $\mA$ with  a nonzero imaginary part, and $x$ is the eigenvector associated with $\lambda$.
+Since all the entries of $\mA$ are real, the eigenvector $x$ is also a complex vector with a nonzero imaginary part.
 Let $\overline \lambda$ and $\overline x$ denote the conjugate eigenvalue and eigenvector, respectively.
@@
 	 Subtracting the equation~\eqref{eq:symm2} from equation~\eqref{eq:symm} yields	\[  0 = ( \overline x^T \mA x -  x^T \mA \overline x ) = (\lambda - \overline \lambda) x^T \overline x.  \]
-	Since that $x^T \overline x$ is a nonzero real number, the imaginary part of $\lambda$ is equal to 0, which contradicts the assumption that $\lambda$ has nonzero imaginary part.
+	Since that $x^T \overline x$ is a nonzero real number, the imaginary part of $\lambda$ is equal to 0, which contradicts the assumption that $\lambda$ has a nonzero imaginary part.
 Therefore, the eigenvalues of $\mA$ are real.
@@
 Since $\mA$ has $n$ independent eigenvectors, the matrix $\mA$ is diagonalizable.
-	Let $\mA =\mS^{-1} \Lambda \mS$, where the definitions of $\Lambda, \mS$ are the same as that in Definition~\ref{def:diagonalization}.
-For all integer $k\geq 0$, we have $\mA^k = \mS^{-1} \Lambda^{k} \mS$.
-Then,
+	Let $\mA =\mS^{-1} \Lambda \mS$, where $\Lambda, \mS$ are defined the same as Definition~\ref{def:diagonalization}. For all integer $k\geq 0$, we have $\mA^k = \mS^{-1} \Lambda^{k} \mS$. Then,
 	\[ \lim_{k \rightarrow +\infty} \mA^k = \lim_{k \rightarrow +\infty} \mS^{-1} \Lambda^{k} \mS  =  \lim_{k \rightarrow +\infty} \mS^{-1} \begin{bmatrix}

### 2020-07-23 skills/math/linear_algebra/linear_algebra_summary_p2.tex
revision

@@
 Given a matrix $\mA \in \bbR^{m \times n}$, we also rewrite the Theorem~\ref{thm:ortho} as
-\[ \mA^T \perp \mA^{\perp} \quad \text{and} \quad \mA \perp \mA^{T,\perp}. \]
+\[ \mA^T \perp \mA^{\perp} \quad \text{and} \quad \mA \perp (\mA^{T})^{\perp}. \]
 
@@
 
-\begin{cor}[Projection onto $\mA^{T,\perp}$]\label{cor:iprojectm}
+\begin{cor}[Projection onto $(\mA^{T})^{\perp}$]\label{cor:iprojectm}
 	Suppose $\mP_{\mA}$ is the projection matrix of $\mA$ from $\bbR^n$ onto the column space $\mA$.
-Then $\mI-\mP_{\mA}$ is the projection matrix of $\mA$ from $\bbR^n$ onto the left nullspace $\mA^{T,\perp}$.
+Then $\mI-\mP_{\mA}$ is the projection matrix of $\mA$ from $\bbR^n$ onto the left nullspace $(\mA^{T})^{\perp}$.
 \end{cor}
@@
 \begin{proof}
-	By Definition~\ref{def:project}, for all $x \in \bbR^n$, $\mP_{\mA} x$ is the projection of $x$ from $\bbR^n$ onto $\mA$. Then,
+	Let $x \in \bbR^n$ be a vector. By Definition~\ref{def:project}, $\mP_{\mA} x$ is the projection of $x$ from $\bbR^n$ onto $\mA$. Then,
 	\[ x - \mP_{\mA} x \perp \mA \quad \Rightarrow \quad (\mI-\mP_{\mA})x \perp \mA. \]
-	By Theorem~\ref{thm:ortho}, we have $\mA \perp \mA^{T,\perp}$. Since $\dim(\mA) + \dim(\mA^{T,\perp}) = n$, we have $\mA  \cup \mA^{T,\perp} = \bbR^n$. Then, we have
-	\[ (\mI-\mP_{\mA})x \in  \mA^{T,\perp}.\]
-	Since $\mP_{\mA} x$ is the projection of $x$ from $\bbR^n$ to $\mA$,  the vector $\mP_{\mA} x \in \mA$, for all  $x \in \bbR^n$. We have,
-	\[ x -  (\mI-\mP_{\mA})x = \mP_{\mA} x \in \mA \quad \Rightarrow \quad  x -  (\mI-\mP_{\mA})x \perp \mA^{T,\perp}.\]
-	Therefore, for all $x \in \bbR^n$, $(\mI-\mP_{\mA})x$ is the projection of $x$ from $\bbR^n$ onto the left nullspace $\mA^{T,\perp}$, and $\mI-\mP_{\mA}$ is the projection matrix of $\mA$ from $\bbR^n$ onto $\mA^{T,\perp}$.
+	By Theorem~\ref{thm:ortho}, we have $\mA \perp (\mA^{T})^{\perp}$. Since $\dim(\mA) + \dim((\mA^{T})^{\perp}) = n$, we have $\mA  \cup (\mA^{T})^{\perp} = \bbR^n$. Then, we have
+	\[ (\mI-\mP_{\mA})x \in  (\mA^{T})^{\perp}.\]
+	Since $\mP_{\mA} x$ is the projection of $x$ from $\bbR^n$ to $\mA$,  the vector $\mP_{\mA} x \in \mA$. We have,
+	\[ x -  (\mI-\mP_{\mA})x = \mP_{\mA} x \in \mA \quad \Rightarrow \quad  x -  (\mI-\mP_{\mA})x \perp (\mA^{T})^{\perp}.\]
+	Therefore, $(\mI-\mP_{\mA})x$ is the projection of $x$ from $\bbR^n$ onto the left nullspace $\mA^{T,\perp}$.
+Since the above conclusion holds for all $x \in \bbR^n$, $\mI-\mP_{\mA}$ is the projection matrix of $\mA$ from $\bbR^n$ onto $\mA^{T,\perp}$.
 \end{proof}
@@
 \[  \hat \beta_{LS} = \argmin_{\beta \in \bbR^{k+1}} \onorm{y - \mX \beta}^2, \]
-where $\onorm{\cdot}$ is the Euclidean norm.  The projection of $y$ onto $\mX$ minimizes the Euclidean distance from $y$ to the column space $\mX$. Hence,   we consider the vector $\mX 
-\hat \beta_{LS}$ as the projection of $y$ onto the column space $\mX$. Therefore, we may use projection to solve the minimization problem.
+where $\onorm{\cdot}$ is the Euclidean norm.
+The projection of $y$ onto $\mX$ minimizes the Euclidean distance from $y$ to the column space $\mX$.
+Hence, the vector $\mX \hat \beta_{LS}$ is the projection of $y$ onto the column space $\mX$.
+Therefore, we may use projection to solve the least squares problem.
 
@@
 
-The estimate $\hat \beta_{LS}$ is consistent with the estimates solved using the first-order condition for the least-squares optimization.
+The estimate $\hat \beta_{LS}$ is the same as the estimates solved using the first-order condition for the least-squares optimization.
 
@@
 
-Orthonormal vectors are always independent.
+Orthonormal vectors are independent.
 
@@
 
-	\item \label{propelim} Subtracting a multiplier of the $i$-th row from the $j$-th row maintains the determinant of the matrix, where $i \neq j$.
+	\item \label{propelim} Subtracting a multiplier of the $i$-th row from the $j$-th row preserves the determinant of the matrix, where $i \neq j$.
 	\begin{proof}
@@
 The product $\mE \mB$ means implementing a linear operation on the rows of $\mB$.
-By property~\ref{propelim}, linear operation maintains the determinant of $\mB$.
+By property~\ref{propelim}, linear operation preserves the determinant of $\mB$.
 Note that the determinant of an elimination matrix is 1.
@@
 	One example of using the cofactor formula is computing the determinant of a \textit{tridiagonal matrix}.
-The tridiagonal matrix is a matrix in which nonzero elements lie only on or adjacent to the diagonal.
+The tridiagonal matrix is a matrix in which nonzero elements locate only on or adjacent to the diagonal.
 Let $\mT_n \in \bbR^{n \times n}$ denote the tridiagonal matrix of 1's; i.e.,
@@
 \begin{proof}
-	First, we re-write the characteristic polynomial $P(\lambda)$ of $\mA$ as a factor formula,
+	First, we re-write the characteristic polynomial $P(\lambda)$ of $\mA$ using the factorization form,
 	\begin{equation}\label{eq:poly}
@@
 \begin{thm}[Descent of matrix powers]
-	Let $\mA \in \bbR^{n \times n}$ be  a square matrix with eigenvalues $\lambda_1,...,\lambda_n$, and  $n$ independent eigenvectors associated with the eigenvalues.
+	Let $\mA \in \bbR^{n \times n}$ be  a diagonalizable matrix with eigenvalues $\lambda_1,...,\lambda_n$.
 The power of the matrix $\mA^k \rightarrow 0$ as $k \rightarrow +\infty$ if and only if $|\lambda_i| < 1$, for all $i \in [n]$.

### 2020-07-23 skills/math/tail_bound/tailbound_solution.tex
revision

@@
 \item[(b)]  Chebyshev's inequality follows by applying Markov's inequality to the nonnegative random variable $Z = {(X - \bbE[X])^2}$.
-As in part (a), given a constant $t > 0$, the variable $Z = {(X - \bbE[X])^2}$ with distribution $\bbP(Z \in \{ 0,t^2 \}) = 1$ attains the equality of the Markov's inequality for $Z$.
+Simialrly as in part (a), given a constant $t > 0$, the variable $Z = {(X - \bbE[X])^2}$ with distribution $\bbP(Z \in \{ 0,t^2 \}) = 1$ attains the equality of the Markov's inequality for $Z$.
 Consequently, the variable $X$ attains the equality of the Chebyshev's inequality for $X$.
@@
 Suppose the moment generating function of variable  $X$ exists over the entire real interval.
-The tail distribution of $X$ is always bounded by a sub-Gaussian distribution with a proper choice of $\sigma^2$.
+Then, the tail distribution of $X$ is bounded by a sub-Gaussian distribution with a proper choice of $\sigma^2$.
 \end{rmk}
@@
 	\end{equation} 
-	where $\sigma^2 = \frac{1}{n} \sum_{i=1}^n \sigma_i^2$ and $h(t) \colonequals (1+t)\log (1+t) - t$ supported on $t \geq 0$.
+	where $\sigma^2 = \frac{1}{n} \sum_{i=1}^n \sigma_i^2$ and $h(t) \colonequals (1+t)\log (1+t) - t$ for $t \geq 0$.
 \end{lem}
@@
 	\vspace{0.2cm}
-	By the power series, for all $i \in [n]$, we have
+	By power series, for all $i \in [n]$, we have
 	\begin{align}\label{eq:e27power}
@@
 	\end{align}
-	where the 0 is deduced from $\bbE[X_i] = 0$, and the last inequality follows by $1 + x \leq e^x$. By $|X_i|<b$, we obtain an upper bound of the sum term in equation~\eqref{eq:e27power} as follows
+	where the 0 is comes from the fact that $\bbE[X_i] = 0$, and the last inequality follows from $1 + x \leq e^x$. By $|X_i|<b$, we bound the last term in equation~\eqref{eq:e27power} as follows
 	\begin{equation}\label{eq:e27upper}
@@
 	\end{equation}
-	Combing the equation~\eqref{eq:e27power} with equation~\eqref{eq:e27upper}, we find an upper bound of the moment generating function of $\sum_{i=1}^n X_i$. That is
+	Combing the equation~\eqref{eq:e27power} with equation~\eqref{eq:e27upper}, we obtain the following upper bound of the moment generating function of $\sum_{i=1}^n X_i$.
 	\begin{equation}\label{eq:e27uppert}
@@
 	\end{align} 
-	By the first-order condition of minimization, the $\lambda$ that satisfies $\lambda b= \log\left(1 + \frac{\delta b}{\sigma^2} \right)$ minimizes the term $ \left( e^{\lambda b} - \lambda b- \lambda \frac{\delta b^2}{\sigma^2} - 1 \right) $.
-Plugging $\lambda b= \log\left(1 + \frac{\delta b}{\sigma^2} \right)$ into the equation~\eqref{eq:e27exp}, we obtain the Bennett's inequality
+	The upper bound \eqref{eq:e27exp} achieves the minimum when $\lambda = b^{-1}\log\left(1 + \frac{\delta b}{\sigma^2} \right)$ by the first-order condition of minimization.
+Plugging $\lambda =b^{-1} \log\left(1 + \frac{\delta b}{\sigma^2} \right)$ into the equation~\eqref{eq:e27exp}, we obtain the Bennett's inequality
 	\begin{equation}\label{eq:e27bennett}
@@
 	\end{equation}
-	where $h(t) \colonequals (1+t)\log (1+t) - t$ supported on $t \geq 0$.
+	where $h(t) \colonequals (1+t)\log (1+t) - t$ for $t \geq 0$.
 
@@
 Particularly, Hoeffding's inequality implies the sub-Gaussianity of bounded variables.
-As the proof for Lemma~\ref{lem:bennett} shows, Bennett's inequality is at least as good as the Bernstein's inequality, for all bounded variables.
+As the proof for Lemma~\ref{lem:bennett} shows, Bennett's inequality is at least as good as the Bernstein's inequality, for bounded random variables.
 \end{rmk}

### 2020-07-24 skills/math/linear_algebra/linear_algebra_summary_p2.tex
revision

@@
 Though $\mA$ usually denotes a matrix, the meaning of $\mA$ should be clear given the context.
-For example, $\rank(\mA)$ refers the rank of matrix $\mA$, while $\dim(\mA)$ refers to the dimension of the column space $C(\mA)$.
+For example, $\rank(\mA)$ refers to the rank of matrix $\mA$, while $\dim(\mA)$ refers to the dimension of the column space $C(\mA)$.
 Specifically, for a vector $v$, $v \in \mA$ means that the vector $v$ belongs to the column space $C(\mA)$.
@@
 	\[ \mA^T \mA x = 0.\]
-	Then, the vector $\mA x \in  \mA^{T,\perp}$.
+	Then, the vector $\mA x \in  (\mA^{T})^{\perp}$.
 By the definition of matrix-to-vector multiplication, we also have $\mA x \in \mA$.
-By the Theorem~\ref{thm:ortho}, the subspaces $\mA^{T,\perp} \perp \mA$.
+By Theorem~\ref{thm:ortho}, the subspaces $(\mA^{T})^{\perp} \perp \mA$.
 Therefore, $\mA x = 0$, and we obtain that $x \in \mA^{\perp}$.
@@
 Hence, the vector $\mX \hat \beta_{LS}$ is the projection of $y$ onto the column space $\mX$.
-Therefore, we may use projection to solve the least squares problem.
+Therefore, we may use projection to solve the least-squares problem.
 
@@
 	where $(\alpha_1, \alpha_2, ...,\alpha_n)$ is the permutation of $(1,2,...,n)$, $\tP$ is the collection of all possible  $(\alpha_1,...,\alpha_n)$, and $N(\alpha_1,...,\alpha_n)$ is the number of swaps from $(\alpha_1,\alpha_2,...,\alpha_n)$  to $(1,2,...,n)$.
-The determinant of $\mA$ follows by the summation of the nonsingular matrices' determinants.
+The determinant of $\mA$ is the sum of the nonsingular matrices' determinants.
 \end{proof}
@@
 In equation~\eqref{eq:poly}, the coefficient for $\lambda^{n-1}$ is equal to $(-1)^n\sum_{i=1}^n \lambda_i$.
-Comparing the equation~\eqref{eq:poly} with Definition~\ref{def:charac}, we obtain $\tr(\mA) = \sum_{i=1}^n \lambda_i$.
+Comparing equation~\eqref{eq:poly} with Definition~\ref{def:charac}, we obtain $\tr(\mA) = \sum_{i=1}^n \lambda_i$.
 Similarly, the constant term in equation~\eqref{eq:poly} is $(-1)^n \prod_{i=1}^n \lambda_i $, while the constant term in Defintion~\ref{def:charac} is $(-1)^n \det(\mA)$.
@@
 	\[ \mA^{-1} \mA x_i = x_i \quad \Rightarrow\quad  \mA^{-1} x_i = {\lambda_i}^{-1} x_i, \quad \text{for all } i \in [n]. \]
-Therefore, $(\lambda_i^{-1},x_i)$ is the eigenvalue-eigenvector pair of $\mA^{-1}$, for all $i \in [n]$.
+Therefore, $(\lambda_i^{-1},x_i)$ is the $i$-th eigenvalue-eigenvector pair of $\mA^{-1}$, for all $i \in [n]$.
 \end{proof}
@@
 	\end{equation}
-	Subtracting the equations~\eqref{eq:indep1} from \eqref{eq:indep2} yields
+	Subtracting equations~\eqref{eq:indep1} from \eqref{eq:indep2} yields
 	\begin{equation}\label{eq:indep3}
@@
 \begin{defn}[Antisymmetric matrices]
-	The matrix $\mA$ is an antisymmetric matrix if 
+	A matrix $\mA$ is antisymmetric if 
 	\[\mA^T = - \mA. \] 
@@
 Suppose $\lambda$  is a complex eigenvalue of $\mA$ with  a nonzero imaginary part, and $x$ is the eigenvector associated with $\lambda$.
-Since all the entries of $\mA$ are real, the eigenvector $x$ is also a complex vector with a nonzero imaginary part.
+Since all the entries of $\mA$ are real, the eigenvector $x$ is also a complex vector.
 Let $\overline \lambda$ and $\overline x$ denote the conjugate eigenvalue and eigenvector, respectively.
@@
 	\[ \overline x^T \mA x  = x^T \mA^T \overline x = x^T \mA \overline x .\] 
-	 Subtracting the equation~\eqref{eq:symm2} from equation~\eqref{eq:symm} yields	\[  0 = ( \overline x^T \mA x -  x^T \mA \overline x ) = (\lambda - \overline \lambda) x^T \overline x.  \]
+	 Subtracting equation~\eqref{eq:symm2} from equation~\eqref{eq:symm} yields	\[  0 = ( \overline x^T \mA x -  x^T \mA \overline x ) = (\lambda - \overline \lambda) x^T \overline x.  \]
 	Since that $x^T \overline x$ is a nonzero real number, the imaginary part of $\lambda$ is equal to 0, which contradicts the assumption that $\lambda$ has a nonzero imaginary part.
@@
 \begin{proof}
-	By Theorems~\ref{thm:symmortho} and \ref{thm:symmrepeat}, for any symmetric matrix $\mA \in \bbR^{n \times n}$, $\mA$ have $n$ orthogonal eigenvectors with possibly repeated eigenvalues.
+	By Theorems~\ref{thm:symmortho} and \ref{thm:symmrepeat}, for any symmetric matrix $\mA \in \bbR^{n \times n}$, $\mA$ has $n$ orthogonal eigenvectors with possibly repeated eigenvalues.
 Therefore, the matrix $\mA$   is able to be diagonalized as equation~\eqref{eq:defdiag}.
@@
 \begin{proof}
-Since $\mA$ has $n$ independent eigenvectors, the matrix $\mA$ is diagonalizable.
-	Let $\mA =\mS^{-1} \Lambda \mS$, where $\Lambda, \mS$ are defined the same as Definition~\ref{def:diagonalization}. For all integer $k\geq 0$, we have $\mA^k = \mS^{-1} \Lambda^{k} \mS$. Then,
+Since $\mA$ is diagonalizable, let $\mA =\mS^{-1} \Lambda \mS$, where $\Lambda, \mS$ are defined the same as Definition~\ref{def:diagonalization}.
+For all integer $k\geq 0$, we have $\mA^k = \mS^{-1} \Lambda^{k} \mS$.
+Then,
 	\[ \lim_{k \rightarrow +\infty} \mA^k = \lim_{k \rightarrow +\infty} \mS^{-1} \Lambda^{k} \mS  =  \lim_{k \rightarrow +\infty} \mS^{-1} \begin{bmatrix}

### 2020-07-24 skills/math/tail_bound/tailbound_solution.tex
revision

@@
 \begin{thm}[Markov's inequality]\label{thm:markov}
-	Let $X \geq 0$ be a random variable  with finite mean. We have
+	Let $X \geq 0$ be a random variable  with a finite mean. We have
 	\begin{equation}\label{eq:thmmarkov}
@@
 \begin{thm}[Chebyshev's inequality]\label{thm:cheb}
-	Let $X \geq 0$ be a random variable  with finite mean $\mu$ and finite variance. We have
+	Let $X \geq 0$ be a random variable  with a finite mean $\mu$ and a finite variance. We have
 	\begin{equation}\label{eq:cheby}
@@
 \begin{thm}[Hoeffding bound for bounded variable]\label{thm:boundvar}
-Let $X$ be a random variable with mean $\mu = \bbE(X)$. Suppose that $X \in [a,b]$ almost surely, where $a\leq b\in\bbR$ are two constants. Then, we have
+Let $X$ be a random variable with $\mu = \bbE(X)$. Suppose that $X \in [a,b]$ almost surely, where $a\leq b\in\bbR$ are two constants. Then, we have
 \begin{equation}
@@
 \end{defn}
-Note that every bounded random variable $X$ satisfies the Bernstein's condition.
+Note that bounded random variables satisfy the Bernstein's condition.
 
@@
 	Applying different functions $g(X)$ to the Markov's inequality leads to different bounds for the tail probability of variable $X$.
-Equation~\eqref{eq:lempoly} implies that optimized polynomial bound is at least as tight as the Chernoff bound, provided that the moment generating function of $X$ exsits.
+Equation~\eqref{eq:lempoly} implies that the optimized polynomial bound is at least as tight as the Chernoff bound, provided that the moment generating function of $X$ exsits in the neighborhood of 0.
 \end{rmk}
@@
 We interpret the second-order derivative $g''(\lambda)$ as the variance of $X$ with the re-weighted distribution $d P' = {e^{\lambda X}}/{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$.
-Taking integral of 1 with respect to $dP'$, we have
+Taking the integral of 1 with respect to $dP'$, we have
 \begin{equation}
@@
 \begin{lem}[Bennett’s inequality]\label{lem:bennett}
-	Let $X_1,...,X_n$ be a sequence of independent zero-mean random variables with $|X_i| \leq b$ and $\var(X_i) = \sigma_i^2$, for all $i \in [n]$. Then, we have Bennett's inequality
+	Let $X_1,...,X_n$ be a sequence of independent zero-mean random variables with $|X_i| \leq b$ and $\var(X_i) = \sigma_i^2$, for all $i \in [n]$. Then, we have the Bennett's inequality
 	\begin{equation}\label{eq:bennett}
@@
 	\end{align}
-	where the 0 is comes from the fact that $\bbE[X_i] = 0$, and the last inequality follows from $1 + x \leq e^x$. By $|X_i|<b$, we bound the last term in equation~\eqref{eq:e27power} as follows
+	where the 0 comes from the fact that $\bbE[X_i] = 0$, and the last inequality follows from $1 + x \leq e^x$. By $|X_i|<b$, we bound the last term in equation~\eqref{eq:e27power} as follows
 	\begin{equation}\label{eq:e27upper}
@@
 			\mathbb{P}\left[\sum_{i=1}^{n} X_{i} \geq n \delta \right] \leq \exp \left\{ \frac{-3n\delta^2}{(2 b \delta + 6 \sigma^2)} \right\}  %=\exp \left\{ -\frac{n \sigma^{2}}{ b^{2}} \frac{3b^2\delta^2}{2\sigma^2(b\delta + 6 \sigma^2)} \right\}
-			= \exp \left\{ -\frac{n \sigma^{2}}{ b^{2}} g\left(\frac{b\delta}{\sigma^2}\right) \right\}.
+			= \exp \left\{ -\frac{n \sigma^{2}}{ b^{2}} g\left(\frac{b\delta}{\sigma^2}\right) \right\}, \quad \text{for all } \delta \geq 0,
 	\end{align}
-	where $g(t) \colonequals  \frac{3t^2}{2t + 6} $, for all $ t \geq 0$.
+	where $g(t) \colonequals  \frac{3t^2}{2t + 6} $ for $ t \geq 0$.
 Since $g(t)\leq h(t)$ holds for all $t \geq 0$, we conclude that the Bennett's inequality~\eqref{eq:e27bennett} is at least as good as Bernstein's inequality~\eqref{eq:e27bern}.
@@
 \begin{rmk} 
-	So far, we have three inequalities that control the tail of bounded variables: Hoeffding's inequality, Bernstein's inequality, and Bennett's inequality.
+	So far, we have three inequalities controlling the tail of bounded variables: Hoeffding's inequality, Bernstein's inequality, and Bennett's inequality.
 Particularly, Hoeffding's inequality implies the sub-Gaussianity of bounded variables.

### 2020-07-25 skills/math/tail_bound/tailbound_solution.tex
tail bound

@@
 \DeclareMathOperator{\subG}{subG}
+
+\newcommand{\of}[1]{\left(#1\right)}
+\newcommand{\off}[1]{\left[#1\right]}
+\newcommand{\offf}[1]{\left\{#1\right\}}
+\newcommand{\aabs}[1]{\left|#1\right|}
 
@@
 Let $X$ be a random variable with mean $\mu = \bbE[X]$ and variance $ \sigma^2 = \var(X)$. We say $X$ satisfies the Bernstein's condition with parameter $b$ if
-\begin{equation}
+\begin{equation}\label{eq:berncond}
 	\left| \bbE [(X -\mu)^k ]\right| \leq \frac{1}{2} k! \sigma^2 b^{k-2}, \quad \text{for } k =3, 4,....
@@
 and the concentration inequality
-\begin{equation}
+\begin{equation}\label{eq:thmbernconc}
 	\bbP\left[|X - \mu| \geq t\right] \leq 2 \exp \left\{ -\frac{t^2}{2(\sigma^2 + bt)} \right\}, \quad \text{for all }t \geq 0.
@@
 
+\subsection{Exercise 2.8}
+\begin{lem}[Bernstein and expectation] Let $Z$ be a nonnegative random variable satisfying the following concentration inequality
+\begin{equation}\label{eq:berncond1}
+	\bbP[Z \geq t] \leq C e^{- \frac{t^2}{2(\nu^2 +B t)}}, \quad \text{for all } t \geq 0,
+\end{equation}
+	where $(\nu, B)$ are two positive constants and $C \geq 1$. Then, the expectation of $Z$ satisfies 
+	\begin{equation}\label{eq:bernexpec}
+		\bbE[Z] \leq 2 \nu (\sqrt{\pi} + \sqrt{\log C}) + 4B (1 + \log C).
+	\end{equation}
+	Further, let $\{ X_i\}_{i=1}^n$ be a sequence of i.i.d.\ zero-mean variables satisfying the Bernstein condition \eqref{eq:berncond}. The sample mean of $\{ X_i\}_{i=1}^n$  satisfies
+	\begin{equation}\label{eq:e28sample}
+		\bbE \off{ \aabs{\frac{1}{n} \sum_{i=1}^n X_i }} \leq 2 \sigma (\sqrt{\pi} + \sqrt{\log 2}) + 4b (1 + \log 2).
+	\end{equation}
+\end{lem}
+
+\begin{proof}
+	First, we prove the equation~\eqref{eq:bernexpec}.
+
+	\vspace{0.2cm}
+	By equation~\eqref{eq:berncond1}, we have
+	\begin{align}
+		\bbP[Z \geq t] &\leq C \max \offf{\exp \of{-\frac{t^2}{4 \nu^2}}, \exp \of{-\frac{t^2}{4 Bt}} }\\
+		&\leq C \exp \of{-\frac{t^2}{4 \nu^2}} + C \exp \of{-\frac{t^2}{4 Bt}}.\label{eq:e28tail}
+	\end{align}
+
+	Plugging the inequality~\eqref{eq:e28tail} to $\bbE[Z] = \int_{0}^{+\infty} \bbP[Z \geq t] dt$, we have
+	\begin{equation}
+		\bbE[Z] = \int_{0}^{+ \infty} \min \offf{1, C \exp \of{-\frac{t^2}{4 \nu^2}}} dt + \int_{0}^{+ \infty} \min\offf{1, C \exp \of{-\frac{t^2}{4 Bt}}} dt   \colonequals I_1 + I_2.
+	\end{equation}
+
+	To evaluate $I_1$, we spilt the integral to avoid the minimization. Solving $1 = C \exp \of{-\frac{t^2}{4 \nu^2}} $, the minimization term becomes
+	\begin{align}
+		\min \offf{1, C \exp \of{-\frac{t^2}{4 \nu^2}}} = \begin{cases}
+			\quad 1 & \text{when } t \leq 2\nu \sqrt{\log C},\\
+			\quad C \exp \of{-\frac{t^2}{4 \nu^2}} & \text{when } t \geq 2\nu \sqrt{\log C}.
+		\end{cases}
+	\end{align}
+	Therefore, we evaluate $I_1$ as follows,
+	\begin{align}
+		I_1 &= \int_{0}^{ 2\nu \sqrt{\log C}} 1 dt + \int_{2\nu \sqrt{\log C}}^{+\infty} C \exp \of{-\frac{t^2}{4 \nu^2}} dt\\
+		(\text{let}\quad y = \frac{t}{2\nu} - \sqrt{\log C})\quad &=  2\nu \sqrt{\log C} + 2\nu\int_{0}^{+\infty} \exp \of{-y^2 - 2y \sqrt{\log C}} dy\\
+		&\leq  2\nu \sqrt{\log C} + 2\nu\int_{0}^{+\infty} \exp \of{-y^2} dy.
+	\end{align}
+	By Gaussian integral $\int_{0}^{+\infty} e^{-x^2} dx = \frac{\sqrt{\pi}}{2}$, we obtain $I_1 \leq 2\nu (\sqrt{\pi} + \sqrt{\log C})$. Similarly, we evaluate $I_2$ as follows 
+	\begin{align}
+		I_2 &= \int_{0}^{4B \log C} 1 dt + \int_{4B \log C}^{+\infty} C \exp \of{-\frac{t}{4B}} dt\\
+		 &= 4B (\log C + 1).
+	\end{align}
+	Hence, we obtain the expectation of $Z$,
+	\begin{equation}
+		\bbE[Z] = I_1 + I_2 \leq  2\nu (\sqrt{\pi} + \sqrt{\log C}) + 4B (\log C + 1).
+	\end{equation}
+
+	\vspace{0.2cm}
+	Next, we prove the equation~\eqref{eq:e28sample}.
+
+	\vspace{0.2cm}
+	For all $i \in [n]$, since $X_i$ satisfies the Bernstein condition with parameter $(\sigma,b)$, the variable $X_i$ satisfies the concentration bound~\eqref{eq:thmbernconc},
+	\begin{equation}
+		\bbP[|X_i| \geq t] \leq 2 \exp \offf{- \frac{t^2}{2(\sigma^2 + bt)}}, \quad \text{for all }t \geq 0. 
+	\end{equation}
+	By equation~\eqref{eq:bernexpec}, we have
+	\begin{equation}
+		\bbE[|X_i|] \leq  2 \sigma (\sqrt{\pi} + \sqrt{\log 2}) + 4b (1 + \log 2).
+	\end{equation}
+
+	Therefore, the expectation of the sample mean satisfies
+	\begin{equation}
+		\bbE \off{ \aabs{\frac{1}{n} \sum_{i=1}^n X_i }} \leq \frac{1}{n} \sum_{i=1}^n \bbE[|X_i|] \leq 2 \sigma (\sqrt{\pi} + \sqrt{\log 2}) + 4b (1 + \log 2).
+	\end{equation}
+
+\end{proof}
+
+\begin{rmk}
+	For all nonnegative random variables satisfying the Bernstein-type inequality \eqref{eq:berncond1} with parameters $(\nu,B,C)$, the expectations of the variables are upper bounded by a function of $(\nu,B,C)$.
+Particularly, for a zero-mean random variable $X$ satisfying the Bernstein condition with parameter $b$, $|X|$ satisfies the inequality \eqref{eq:berncond1} with parameters $(\sigma,b,2)$, where $\sigma^2 = \var(X)$.
+The expectation of the absolute variable $|X|$ depends on the variance $\sigma^2$ and the parameter $b$.
+\end{rmk}
+
 \end{document}

### 2020-07-25 skills/math/tail_bound/tailbound_solution.tex
tail bound

@@
 
+\subsection{Exercise 2.9}
+\begin{lem}
+	Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.\ Bernoulli variables with parameter $\alpha \in (0,1/2]$.
+Consider the binomial random variable $Z_n = \sum_{i=1}^n X_i$.
+The tail probability of $Z_n$ is upper bounded as
+	\begin{equation}\label{eq:lembinupper}
+		\bbP\off{Z_n \leq \delta n} \leq e^{-n D\of{\delta \Vert \alpha}}, \quad \text{for all }\delta \in (0,\alpha),
+	\end{equation}
+	where the quantity 
+	\begin{equation}
+		D(\delta \Vert \alpha) \colonequals \delta \log \frac{\delta}{\alpha} + (1-\delta) \log \frac{1-\delta}{1-\alpha}
+	\end{equation}
+	is the KL divergence between the Bernoulli distributions with parameters $\delta$ and $\alpha$, respectively.
+
+	Further, the upper bound~\eqref{eq:lembinupper} is strictly tighter than the Hoeffding bound for all $\delta \in (0,\alpha)$.
+\end{lem}
+
+\begin{proof}
+    Applying the Chernoff inequality to the random variable $Z_n$, we have
+    \begin{equation}\label{eq:e29upper}
+    	\bbP[Z_n \leq \delta n] \leq \frac{\bbE\off{e^{\lambda Z_n}}}{e^{\lambda \delta n}} = \offf {\exp \offf{\log \of{1 - \alpha + \alpha e^{\lambda})}- \lambda \delta }}^n,\quad \text{for all } \lambda \in \bbR,
+    \end{equation}
+    where the second equality follows from the moment generating function of $Z_n$
+    \[\bbE\off{e^{\lambda Z_n}} = \of{1 - \alpha + \alpha e^{\lambda}}^n. \]
+    The upper bound~\eqref{eq:e29upper} achieves the minimum when $\lambda = \log {\frac{\delta(1-\alpha)}{\alpha(1 - \delta)}}$ by the first-order condition.
+Plugging the minimizer into equation~\eqref{eq:e29upper}, we obtain the result
+    \[ \bbP\off{Z_n \leq \delta n}  \leq e^{-n D\of{\delta \Vert \alpha}}, \quad \text{for all }\delta \in (0,\alpha).  \]
+
+    \vspace{0.2cm}
+    Next, we show that equation~\eqref{eq:lembinupper} is strictly better than Hoeffding bound for all $\delta \in (0,\alpha)$.
+
+    \vspace{0.2cm}
+    Hoeffding bound for random variable $Z_n$ is 
+    \begin{equation}\label{eq:e29hoeff}
+    	\bbP\off{Z_n \leq \delta n} \leq e^{ {-2n(\delta - \alpha)^2}}, \quad \text{for all }\delta \in (0,\alpha).
+    \end{equation}
+    Note that parameter $\alpha$ is fixed.
+Consider the function $g(\delta) = 2(\delta - \alpha)^2 - D\of{\delta \Vert \alpha}$ for $\delta \in (0,\alpha)$.
+The first-order derivative and  second-order derivative of $g$ are
+    \[ g'(\delta) = 4(\delta - \alpha) - \log \frac{\delta}{\alpha} + \log \frac{1-\delta}{1-\alpha} \quad \text{and} \quad g''(\delta) = 4 - \frac{1}{\delta} - \frac{1}{1-\delta}. \] 
+    Note that $ g(\alpha) = g'(\alpha) = 0$, and $g''(\delta) < 0$ for all $0 < \delta < 1/2$.
+We have $g(\delta) < 0$, for all $\delta \in (0,\alpha)$.
+Hence, the upper bound~\eqref{eq:lembinupper} is strictly better than Hoeffding bound~\eqref{eq:e29hoeff}.
+
+\end{proof}
+
+\begin{rmk}
+	The upper bound on a binomial tail may be a function of the KL divergence between Bernoulli distributions with distinct parameters.
+The bound using KL divergence is strictly tighter than the Hoeffding bound.
+The underperformance of Hoeffding bound may attribute to the utilization of the sub-Gaussian parameter $\sigma^2$.
+The sub-Gaussian parameter $\sigma^2$ is not necessarily the optimal choice to describe the tail performance of a random variable with good properties.
+\end{rmk}
 \end{document}

### 2020-07-31 skills/math/tail_bound/tailbound_solution.tex
tailbound

@@
 \DeclareMathOperator{\subG}{subG}
+
+\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
+\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
 
@@
 		\min \offf{1, C \exp \of{-\frac{t^2}{4 \nu^2}}} = \begin{cases}
-			\quad 1 & \text{when } t \leq 2\nu \sqrt{\log C},\\
+			\quad 1 & \text{when } t < 2\nu \sqrt{\log C},\\
 			\quad C \exp \of{-\frac{t^2}{4 \nu^2}} & \text{when } t \geq 2\nu \sqrt{\log C}.
@@
 \begin{rmk}
-	For all nonnegative random variables satisfying the Bernstein-type inequality \eqref{eq:berncond1} with parameters $(\nu,B,C)$, the expectations of the variables are upper bounded by a function of $(\nu,B,C)$.
+	For a nonnegative random variables satisfying the Bernstein-type inequality \eqref{eq:berncond1}, the expectation of the variable is upper bounded by a function of the parameter $(\nu,B,C)$.
 Particularly, for a zero-mean random variable $X$ satisfying the Bernstein condition with parameter $b$, $|X|$ satisfies the inequality \eqref{eq:berncond1} with parameters $(\sigma,b,2)$, where $\sigma^2 = \var(X)$.
-The expectation of the absolute variable $|X|$ depends on the variance $\sigma^2$ and the parameter $b$.
+Then, the expectation of the absolute variable $|X|$ is upper bounded by a function of $(\sigma^2, b)$.
 \end{rmk}
@@
 \subsection{Exercise 2.9}
-\begin{lem}
+\begin{lem}[Sharp upper bounds on binomial tails]
 	Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.\ Bernoulli variables with parameter $\alpha \in (0,1/2]$.
@@
 	where the quantity 
-	\begin{equation}
+	\begin{equation}\label{eq:kldiv}
 		D(\delta \Vert \alpha) \colonequals \delta \log \frac{\delta}{\alpha} + (1-\delta) \log \frac{1-\delta}{1-\alpha}
@@
     \end{equation}
-    where the second equality follows from the moment generating function of $Z_n$
+    where the second equality follows from the moment generating function of $Z_n$,
     \[\bbE\off{e^{\lambda Z_n}} = \of{1 - \alpha + \alpha e^{\lambda}}^n. \]
@@
     \vspace{0.2cm}
-    Next, we show that equation~\eqref{eq:lembinupper} is strictly better than Hoeffding bound for all $\delta \in (0,\alpha)$.
+    Next, we show that equation~\eqref{eq:lembinupper} is strictly tighter than the Hoeffding bound of $Z_n$, for all $\delta \in (0,\alpha)$.
 
     \vspace{0.2cm}
-    Hoeffding bound for random variable $Z_n$ is 
+    Hoeffding bound of random variable $Z_n$ is 
     \begin{equation}\label{eq:e29hoeff}
@@
     Note that $ g(\alpha) = g'(\alpha) = 0$, and $g''(\delta) < 0$ for all $0 < \delta < 1/2$.
-We have $g(\delta) < 0$, for all $\delta \in (0,\alpha)$.
-Hence, the upper bound~\eqref{eq:lembinupper} is strictly better than Hoeffding bound~\eqref{eq:e29hoeff}.
+Then, we have $g(\delta) < 0$, for all $\delta \in (0,\alpha)$.
+Hence, the upper bound~\eqref{eq:lembinupper} is strictly tighter than Hoeffding bound~\eqref{eq:e29hoeff}.
 
@@
 \begin{rmk}
-	The upper bound on a binomial tail may be a function of the KL divergence between Bernoulli distributions with distinct parameters.
+	The upper bound on a binomial tail is a function of the KL divergence between Bernoulli distributions with distinct parameters.
 The bound using KL divergence is strictly tighter than the Hoeffding bound.
@@
 \end{rmk}
+
+\subsection{Exercise 2.10}
+\begin{lem}[Lower bounds on binomial tails]
+	Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.\ Bernoulli variables with parameter $\alpha \in (0,1/2]$.
+Consider the binomial random variable $Z_n \colonequals \sum_{i=1}^n X_i$.
+For some fixed $\delta \in (0,\alpha)$, let $m = \floor{n\delta}$; i.e., $m$ is the largest integer less or equal to $n \delta$.
+Let $ \tilde \delta = {m}/{n}$.
+We have
+	\begin{equation}\label{eq:e2101}
+		\frac{1}{n} \log \bbP \off{Z_n \leq n\delta} \geq \frac{1}{n} \log \begin{pmatrix}
+			n \\ m	\end{pmatrix} + \tilde \delta \log \alpha + (1 - \tilde \delta ) \log (1-\alpha).
+	\end{equation}
+
+	Further, the binomial coefficient satisfies
+	\begin{equation}\label{eq:e2102}
+		\frac{1}{n} \log \begin{pmatrix}
+			n \\ m	\end{pmatrix} \geq \phi(\tilde \delta ) - \frac{\log (n+1)}{n}.
+	\end{equation}
+
+	Consequently, the lower tail of binomial variable $Z_n$ satisfies
+	\begin{equation}\label{eq:e2103}
+		\bbP[Z_n \leq n \delta]  \geq \frac{1}{n+1} e^{-nD(\delta \Vert \alpha)},
+	\end{equation}
+	where $D(\delta \Vert \alpha)$ is the KL divergence defined in equation~\eqref{eq:kldiv}.
+\end{lem}
+
+\begin{proof}
+	First, we prove equation~\eqref{eq:e2101}.
+
+	\vspace{0.2cm}
+	Since $Z_n$ is a binomial variable with size $n$ and probability $\alpha$, we have
+	\begin{align}
+		\bbP[Z_n \leq n\delta ] &= \sum_{k = 1}^m \begin{pmatrix}
+			n \\ k
+		\end{pmatrix} \alpha^k (1-\alpha)^{n - k}\\
+		& \geq \begin{pmatrix}
+			n \\ m
+		\end{pmatrix} \alpha^m (1-\alpha)^{n - m}. \label{eq:e210pro1}
+	\end{align}
+	Taking the log and dividing by $n$ on the both sides of the inequality~\eqref{eq:e210pro1} , we obtain
+	\[ \frac{1}{n} \log \bbP[Z_n \leq n\delta ] \geq \frac{1}{n} \log \begin{pmatrix}
+			n \\ m		\end{pmatrix} + \tilde \delta \log \alpha + (1 - \tilde \delta ) \log (1-\alpha). \]
+
+	\vspace{0.2cm}
+	Next, we prove equation~\eqref{eq:e2102}.
+
+	\vspace{0.2cm}
+	Consider a binomial variable $Y \sim Bin(n,\tilde \delta)$. For all $k = 0,1,...,(n-1)$, the ratio between $\bbP[Y = k]$ and $\bbP[Y = k+1]$ is
+	\begin{equation}
+		\frac{\bbP[Y = k+1]}{\bbP[Y = k]} = \frac{\begin{pmatrix}
+			n \\ k+1
+		\end{pmatrix} \tilde \delta^{k+1} (1-\tilde \delta)^{n - k -1} }{\begin{pmatrix}
+			n \\ k
+		\end{pmatrix} \tilde \delta^{k} (1-\tilde \delta)^{n - k }} = \frac{(n-k) \tilde \delta}{(k+1) (1- \tilde \delta)  }.
+	\end{equation}
+	To let the ratio $	\frac{\bbP[Y = k+1]}{\bbP[Y = k]} \geq 1$, we need $(k+1)\leq (n+1) \tilde \delta $.
+Thus, the probability $\bbP[Y = l]$ achieves the maximum when $l = n \tilde \delta$.
+Consequently, we have
+	\begin{equation}\label{eq:e210pro2}
+		(n+1) \bbP[Y = n \tilde \delta] \geq 1 \quad \Leftrightarrow \quad  \bbP[Y = n \tilde \delta] \geq \frac{1}{n+1}.
+	\end{equation}
+	Taking the log and dividing by $n$ on the both sides of the inequality~\eqref{eq:e210pro2}, we obtain
+	\begin{equation}
+		 \frac{1}{n} \log \begin{pmatrix}
+			n \\ m		\end{pmatrix} \geq \tilde \delta \log(\alpha) - (1 - \tilde \delta)\log(1 - \alpha) - \frac{\log(n+1)}{n} = \phi(\tilde \delta)  - \frac{\log(n+1)}{n}.
+	\end{equation}
+
+	\vspace{0.2cm}
+	Last, we prove the lower bound equation~\eqref{eq:e2103}.
+
+	\vspace{0.2cm}
+	Plugging equation~\eqref{eq:e2102} into equation~\eqref{eq:e2101}, we have
+	\begin{equation}\label{eq:e210pro3}
+		\frac{1}{n} \log \bbP \off{Z_n \leq n\delta} \geq \phi(\tilde \delta) + \tilde \delta \log \alpha + (1 - \tilde \delta ) \log (1-\alpha) - \frac{\log(n+1)}{n} \geq -D(\delta \Vert \alpha ) - \frac{\log(n+1)}{n}.
+	\end{equation}
+	Multiplying $n$ and exponentiating on the both sides of the inequality~\eqref{eq:e210pro3}, we obtain the result
+	\begin{equation}
+		\bbP \off{Z_n \leq n\delta} \geq \frac{1}{n+1} e^{-nD(\delta \Vert \alpha)}.
+	\end{equation}
+\end{proof}
+
+\begin{rmk}
+	The lower bound on a binomial tail is also a function of the KL divergence between Bernoulli distributions with distinct parameters.
+Combining the upper bound \eqref{eq:lembinupper} and lower bound \eqref{eq:e2103}, we conclude that the binomial tail is (upper and lower) bounded by the functions of  KL divergence between Bernoulli distributions.
+\end{rmk}
+
+\subsection{Exercise 2.11}
+\begin{lem}[Gaussian maxima]
+Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.
+normal random variables following $N(0,\sigma^2)$.
+Consider the random variable $Z_n \colonequals \max_{i\in[n]} |X_i|$.
+Since the tail bound $\bbP[U \geq t] \leq \sqrt{\frac{2}{\pi}} \frac{1}{t} e^{-t^2 /2}$ holds for all standard normal random variable $U$, the expectation of $Z_n$ is upper bounded as follows
+\begin{equation}\label{eq:e2111}
+	\bbE[Z_n] \leq \sqrt{2 \sigma^2 \log n} + \frac{4\sigma}{\sqrt{2 \log n}}, \quad \text{for all } n\geq 2.
+\end{equation}
+\iffalse
+The lower bound of $\bbE[Z_n] $ is
+\begin{equation}\label{eq:e2112}
+	\bbE[Z_n] \geq (1- 1/e) \sqrt{2\sigma^2 \log n} , \quad \text{for all }n \geq 5.
+\end{equation}
+\fi
+
+\end{lem}
+
+\begin{proof}
+	%First, we prove the equation~\eqref{eq:e2111}.
+
+	\vspace{0.2cm}
+	Consider the tail probability of $Z_n$. We have
+	\begin{equation}
+		\bbP[Z_n \geq t] = \bbP[\max_{i \in [n]} |X_i| \geq t] = 1 - \bbP[\max_{i \in [n]} |X_i| < t] = 1 - (1- \bbP[|X_1| \geq t])^n,
+	\end{equation}
+	where the last equality follows from the independence of $\{X_i\}_{i=1}^n$. By Bernoulli's inequality, we have
+	\begin{equation}
+		(1- \bbP[|X_1| \geq t])^n \geq 1 - n \bbP[|X_1| \geq t], \quad \text{for all } t >0.
+	\end{equation} 
+	Therefore, the expectation of $Z_n$ follows
+	\begin{equation}\label{eq:e211pro1}
+		\bbE[Z_n] = \int_{0}^{+\infty} \bbP[Z_n \geq t] dt \leq  c +  \int_{c}^{+\infty} n\bbP[|X_1| \geq t] dt, \quad \text{for all } c > 0.
+	\end{equation}
+	 Since the tail bound $\bbP[U \geq t] \leq \sqrt{\frac{2}{\pi}} \frac{1}{t} e^{-t^2 /2}$ holds for all standard normal random variable $U$, the tail bound of $|X_1|$ satisfies
+	 \begin{equation}\label{eq:e211proupper}
+	 	\bbP[|X_1| \geq t] \leq 2 \sqrt{\frac{2}{\pi}} \frac{\sigma}{t} e^{-\frac{t^2}{2\sigma^2}}.
+	 \end{equation}
+	 Hence, the last integral in equation~\eqref{eq:e211pro1} follows
+	 \begin{align}
+	 	 \int_{c}^{+\infty} n\bbP[|X_1| \geq t] dt &\leq \frac{2n \sigma}{c} \sqrt{\frac{2}{\pi}} \int_{c}^{+\infty} e^{-\frac{t^2}{2\sigma^2} }dt \\ (\text{let } u = \frac{ t}{\sigma}) \quad  
+	 	 %& = \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}   \int_{\frac{c}{\sigma}}^{+\infty} e^{-\frac{u^2}{2}} dt \\
+	 	 & \leq \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}   \int_{\frac{c}{\sigma}}^{+\infty} u e^{-\frac{u^2}{2}} dt \\
+	 	 & = \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}  e^{-\frac{c^2}{2\sigma^2}}. \label{eq:e211pro2}
+	 \end{align}
+	 For all $n \geq 2$, let $c = \sqrt{2\sigma^2 \log n}$ and plug the $c$ into equations~\eqref{eq:e211pro1} and \eqref{eq:e211pro2}. Then, we obtain the result
+	 \begin{equation}
+	 	\bbE[Z_n] \leq \sqrt{2\sigma^2 \log n} + \frac{2\sigma}{\sqrt{2\log n}}   \sqrt{\frac{2}{\pi}}  \leq \sqrt{2\sigma^2 \log n} + \frac{4\sigma}{\sqrt{2\log n}}, \quad \text{for all }n \geq 2.
+	 \end{equation}
+
+	 \iffalse
+	 \vspace{0.2cm}
+	 Next, we prove the equation~\eqref{eq:e2112}.
+
+	 \vspace{0.2cm}
+
+	 By Markov's inequality, the expectation of $Z_n$ satisfies
+	 \begin{equation}
+	 	\bbE[Z_n] \geq t \bbP[Z_n \geq t],\quad \text{for all }t \geq 0.
+	 \end{equation}
+	 By the independence of $\{X_i\}_{i=1}^n$, the tail probability of $Z_n$ is
+	 \begin{equation}
+	 	\bbP[Z_n \geq t] = 1 - (1 - \bbP(|X_1| \geq t))^n.
+	 \end{equation}
+
+	 Let $t = \sqrt{2\sigma^2 \log n}$. By equation~\eqref{eq:e211proupper}, the tail probability of $X_1$ satisfies
+	 \begin{equation}
+	 	\bbP(|X_1| \geq \sqrt{2\sigma^2 \log n}) \leq \frac{2}{\sqrt{ \pi \log n}} \frac{1}{n} \leq \frac{1}{n},\quad \text{for all } n \geq 5.
+	 \end{equation}	 
+	 Hence, the tail probability of $Z_n$ with $t = \sqrt{2\sigma^2 \log n} $satisfies
+	 \begin{equation}
+	 	\bbP[Z_n \geq t] \geq 1 - 
+	 \end{equation}
+	 \fi
+\end{proof}
+
 \end{document}

### 2020-08-07 skills/math/tail_bound/tailbound_solution.tex
tail bound

@@
 \DeclareMathOperator{\subG}{subG}
+\DeclareMathOperator{\tr}{Trace}
 
@@
 	\bbP\left[|X - \mu| \geq t\right] \leq 2 \exp \left\{ -\frac{t^2}{2(\sigma^2 + bt)} \right\}, \quad \text{for all }t \geq 0.
+\end{equation}
+
+\end{thm}
+
+\begin{defn}[Bounded difference property] Let $f\colon \bbR^n \mapsto \bbR$ be a function. The function $f$ satisfies the bounded difference property with parameter $(L_1,...,L_n)$ if we have
+	\begin{equation}\label{eq:defbounded}
+		\aabs{f(x^{(k)}) - f(x^{'(k)})} \leq L_k,
+	\end{equation}
+	 for all $k \in [n]$ and for all $x^{(k)} = (x_1,...,x_k,...,x_n), x^{'(k)} = (x_1,...,x'_k,...,x_n)\in\bbR^n$.
+\end{defn}
+
+\begin{thm}[Bounded differences inequality] \label{thm:bounddiff}
+Let $f\colon \bbR^n \mapsto \bbR$ be a function satisfies the bounded difference property~\eqref{eq:defbounded}, and the random variable $X = (X_1,...,X_n)$ has independent components. Then,
+\begin{equation}
+	\bbP\off{\aabs{f(X) -\bbE[f(X)] } \geq t} \leq 2 e^{-\frac{2t^2}{\sum_{i=1}^n L_i^2}}, \quad \text{for all }t \geq 0.
 \end{equation}
@@
 
+\subsection{Exercise 2.12}
+\begin{lem}[Sharp upper bounds for sub-Gaussian maxima]
+	Let $\{X_i\}_{i = 1}^n$ be a sequence of zero-mean sub-Gaussian variables with parameter $\sigma$. Then, we have
+	\begin{equation}
+		\bbE \off{\max_{i \in [n]} X_i} \leq \sqrt{2\sigma^2 \log n}, \quad \text{for all } n \geq 1.
+	\end{equation}
+	Note that independence assumptions are unnecessary.
+\end{lem}
+
+\begin{proof}
+	For all $t >0$, the function $f(x) = \exp \of{tx}$ is a convex function. Apply the function $f(x)$ to $\bbE \off{\max_{i \in [n]} X_i} $. By Jensen's inequality, we have
+	\begin{equation}\label{eq:e212}
+		\exp \offf{t \bbE \off{\max_{i \in [n]} X_i} } \leq \bbE\off{\exp \of{t \max_{i \in [n]} X_i }} \leq \sum_{i= 1}^n \bbE\off{\exp \of{t X_i }} \leq n e^{\frac{t^2 \sigma^2}{2}}, \quad \text{for all }t > 0,
+	\end{equation}
+	where the last inequality follows from the sub-Gaussianity of $X_i$s. Take the log of both sides of equation~\eqref{eq:e212}. We have
+	\begin{equation}\label{eq:e212b}
+		\bbE \off{\max_{i \in [n]} X_i}  \leq \frac{\log n}{t} + \frac{t\sigma^2}{2}.
+	\end{equation}
+	Plugging $t = \frac{\sqrt{2\log n} }{\sigma}$ into the equation~\eqref{eq:e212b}, we obtain
+	\[ \bbE \off{\max_{i \in [n]} X_i}  \leq \sqrt{2 \sigma^2 \log n}.\]
+\end{proof}
+
+\subsection{Exercise 2.13}
+\begin{lem}[Operations on sub-Gaussian variables]
+	Let $X_1$ and $X_2$ be two zero-mean sub-Gaussian variables with parameters $\sigma_1$ and $\sigma_2$ respectively. 
+	\begin{enumerate}
+		\item[(a).] If $X_1$ and $X_2$ are independent, then random variable $X_1 + X_2$ is sub-Gaussian with parameter $\sqrt{\sigma_1^2 + \sigma_2^2}$.
+		\item[(b).] Without independence assumptions, then random variable $X_1 + X_2$ is sub-Gaussian with parameter at most 
+		%$2 \sqrt{\sigma_1^2 +\sigma_2^2 }$.
+		$\sigma_1 + \sigma_2$.
+		\item[(c).] If $X_1$ and $X_2$ are independent, then random variable $X_1 X_2$ is sub-exponential with parameter $(\nu, b) = \of{\sqrt{2}\sigma_1 \sigma_2, \frac{1}{\sqrt{2}\sigma_1 \sigma_2}}$.
+	\end{enumerate}
+\end{lem}
+
+\begin{proof}
+	First, we prove $(a)$.
+
+	\vspace{0.2cm}
+	Since $X_1\sim \subG(\sigma_1)$ and $X_2 \sim \subG(\sigma_2)$ are independent, we have
+	\begin{equation}
+		\bbE\off{e^{t(X_1 + X_2)}} = \bbE\off{e^{tX_1}} \bbE\off{e^{tX_2}} \leq e^{\frac{t^2(\sigma_1^2 + \sigma_2^2)}{2}},\quad \text{for all }t \in \bbR.
+	\end{equation}
+	Therefore, $X_1 +X_2$ is sub-Gaussian with parameter $\sqrt{\sigma_1^2 + \sigma_2^2}$.
+
+	\vspace{0.2cm}
+Next, we prove $(b)$.
+
+\vspace{0.2cm}
+For all $t \in \bbR$, we have
+\begin{equation}\label{eq:e213b}
+	\bbE\off{e^{t(X_1 + X_2)}} = \bbE \off{e^{t X_1} e^{t X_2}}.
+\end{equation}
+Introduce $p,q \geq 1$ such that $\frac{1}{p} + \frac{1}{q} = 1$. By Holder's inequality, the equation~\eqref{eq:e213b} becomes
+\begin{equation}\label{eq:e213b2}
+	\bbE\off{e^{t(X_1 + X_2)}} \leq \of{\bbE \off{e^{ p t X_1}  } }^{\frac{1}{p}}
+	 \of{\bbE \off{e^{ pt X_2}  }}^{\frac{1}{q}} \leq e^{  \frac{pt^2 \sigma_1^2 + qt^2 \sigma_2^2}{2} },
+\end{equation}
+where the last inequality follows from the sub-Gaussianity of $X_1$ and $X_2$.
+The inequality~\eqref{eq:e213b2} achieves its minimum when $p = \sigma_2/\sigma_1 +1$.
+Plugging the minimizer to inequality~\eqref{eq:e213b2}, we obtain
+\begin{equation}
+	\bbE\off{e^{t(X_1 + X_2)}} \leq e^{\frac{t^2(\sigma_1 + \sigma_2
+	)^2}{2}}.
+\end{equation}
+Therefore, the random variable $X_1 + X_2$ is sub-Gaussian with parameter at most $\sigma_1 + \sigma_2$.
+
+\vspace{0.2cm}
+
+Last, we prove $(c)$.
+
+\vspace{0.2cm}
+
+\end{proof}
+
+\subsection{Exercise 2.14}
+\begin{lem}
+	Let $X$ be a scalar random variable. Suppose there exist two positive constant $c_1, c_2$ such that
+	\begin{equation}\label{eq:leme214}
+		\bbP \off{|X - \bbE[X]| \geq t} \leq c_1 e^{-c_2 t^2},\quad \text{for all }t\geq 0.
+	\end{equation}
+	Then, the variance $\var(X) \leq {c_1}/{c_2}$.
+
+	\vspace{0.2cm}
+	Let $m_X$ be the median of $X$; i.e., $\bbP[X \geq m_X] \geq 1/2$ and  $\bbP[X \leq m_X] \geq 1/2$.
+Note that the median $m_X$ is not necessarily unique.
+Given mean concentration~\eqref{eq:leme214}, for all median $m_X$, we have
+	\begin{equation}\label{eq:leme214b}
+		\bbP[|X - m_X| \geq t] \leq c_3 e^{-c_4t^2}, \quad \text{for all }t \geq 0,
+	\end{equation} 
+	where $c_3 \colonequals 4c_1$ and $c_4 \colonequals c_2/8$.
+
+	 Conversely, if the equation~\eqref{eq:leme214b} holds, the mean concentration~\eqref{eq:leme214} also holds with parameter $c_1 = 2c_3$ and $c_2 = c_4/4$.
+\end{lem}
+
+\begin{proof}
+	First, we show the variance  $\var(X) \leq {c_1}/{c_2}$.
+
+	\vspace{0.2cm}
+
+	By the definition of variance, we have
+	\begin{equation}\label{eq:e2141}
+		\var(X) = \bbE[(X-\bbE[X])^2] = \int_{0}^{+\infty} \bbP \off{|X - \bbE[X]| \geq \sqrt{t} }dt.
+	\end{equation}
+	Since $X$ satisfies the mean concentration equation~\eqref{eq:leme214}, for all $t > 0$, we have
+	\begin{equation}\label{eq:e2142}
+		\bbP \off{|X - \bbE[X]| \geq \sqrt{t} } \leq c_1 e^{-c_2 t}.
+	\end{equation}
+	Plugging the equation~\eqref{eq:e2142} into equation~\eqref{eq:e2141}, we obtain
+	\[ \var(X) \leq \frac{c_1}{c_2}. \]
+
+	\vspace{0.2cm}
+
+	Next, we show by an example that the median of $X$, $m_X$, is not necessarily unique.
+
+	\vspace{0.2cm}
+	Consider a random variable $Y \sim Bin(5,1/2)$.
+Then, we have $\bbP[Y \geq 3] = 1/2$ and $\bbP[Y \leq 3] > 1/2$.
+Meanwhile, we also have $\bbP[Y \geq 2] > 1/2$ and $\bbP[Y \leq 2] = 1/2$.
+By the definition of median, the number $2$ and $3$ both are the median of $Y$.
+Therefore, the median of a random variable is not necessarily unique.
+
+	\vspace{0.2cm}
+
+	Then, we prove the median concentration~\eqref{eq:leme214b}, provided that random variable $X$ satisfies the mean concentration~\eqref{eq:leme214}.
+
+	\vspace{0.2cm}
+
+	Let $\Delta = |\bbE[X] - m_X|$. Consider the following two cases.
+	\begin{enumerate}
+		\item[1.] Case 1: Suppose $t \geq 2\Delta$.
+
+		\vspace{0.2cm}
+		Note the triangle inequality $|X - \bbE[X]| \geq |X - m_X| - \Delta$ and the assumption $\frac{t}{2} \geq \Delta$. We have
+		\begin{equation}
+			\bbP[|X - m_X| \geq t] \leq \bbP \off{|X - m_X| \geq \frac{t}{2} + \Delta} \leq   \bbP \off{|X - \bbE[X]|  \geq \frac{t}{2}}.
+		\end{equation}
+		By equation~\eqref{eq:leme214}, we obtain
+		\begin{equation}\label{eq:e214case1}
+			\bbP[|X - m_X| \geq t] \leq c_1 e^{-\frac{c_2t^2}{4}}.
+		\end{equation}
+
+		\item[2.] Case 2:   Suppose $t < 2\Delta$.
+
+		\vspace{0.2cm}
+		By the definition of median, we have
+		\begin{equation}\label{eq:e214ca2a}
+			\bbP[|X - \bbE[X]| \geq \Delta] \geq \bbP[X \geq m_X] \geq \frac{1}{2}.
+		\end{equation}
+		Meanwhile, by equation~\eqref{eq:leme214}, we have
+		\begin{equation}\label{eq:e214ca2b}
+			\bbP[|X - \bbE[X]| \geq \Delta] \leq c_1 e^{-c_2 \Delta^2}.
+		\end{equation}
+
+		Combing the assumption $\Delta > \frac{t}{2}$  and inequalities~\eqref{eq:e214ca2a} \eqref{eq:e214ca2b}, we obtain
+		\begin{equation}\label{eq:e214case2}
+			\bbP[|X - m_X| \geq t]\leq 1\leq 2c_1  e^{-c_2 \Delta^2} \leq 2c_1  e^{-\frac{c_2 t^2}{4}}. 
+		\end{equation}
+
+	\end{enumerate}
+
+	Hence, combing the inequality~\eqref{eq:e214case1} and inequality~\eqref{eq:e214case2}, we conclude that
+	\begin{equation}
+		\bbP[|X - m_X| \geq t] \leq c_3 e^{-c_4 t^2}, 
+	\end{equation}
+	where $c_3 = 2c_1$ and $c_4 = c_2/4$. The upper bound also holds when $c_3 = 4 c_1$ and $c_4 = c_2/8$.
+
+	\vspace{0.2cm}
+
+	Last, we prove the mean concentration~\eqref{eq:leme214}, provided that the random variable $X$ satisfies the median concentration~\eqref{eq:leme214b}.
+
+	\vspace{0.2cm}
+
+	As the proof for median concentration~\eqref{eq:leme214b}, we also consider the following two cases.
+	\begin{enumerate}
+		\item[1.] Case 1: Suppose $t \geq 2\Delta$.
+
+		\vspace{0.2cm}
+		Note the triangle inequality $|X - m_X| + \Delta \geq |X - \bbE[X]|$ and the assumption $\frac{t}{2} \geq \Delta$. We have
+		\begin{equation}
+			\bbP[|X - \bbE[X]| \geq t ] \leq \bbP \off{|X - m_X| \geq \frac{t}{2}} \leq c_3 e^{-\frac{c_4 t^2}{4}},
+		\end{equation}
+		where the last inequality follows from the median concentration~\eqref{eq:leme214b}.
+
+		\item[2.] Case 2: Suppose $t < 2\Delta$.
+	\end{enumerate}
+	\end{proof}
+
+	\subsection{Exercise 2.15}
+	\begin{lem}
+		Let $\{X_i\}_{i=1}^n$ be i.i.d.\ sample of random variables with density $f$ on the real line. A standard estimate of $f$ is the kernel density estimate 
+		\begin{equation}
+			\hat f_n (x) = \frac{1}{nh} \sum_{i=1}^n  K\of{ \frac{x- X_i}{h} },
+		\end{equation}
+		where $K \colon \bbR \mapsto [0,+\infty )$ is a kernel function satisfying $\int_{-\infty}^{+\infty} K(x) dx = 1$, and $h > 0$ is a bandwidth parameter.
+Suppose we assess the estimate $\hat f_n (x)$ using the $L^1$ norm; i.e.,
+		\[ \onorm{\hat f_n - f}_1 = \int_{-\infty}^{+\infty} |\hat f_n(t) - f(t)|dt. \]
+		The upper tail probability of the $L^1$ norm satisfies
+		\begin{equation}
+			\bbP\off{\onorm{\hat f_n - f}_1 - \bbE \off{\onorm{\hat f_n - f}_1}\geq \delta} \leq e^{-\frac{n\delta^2}{8}},\quad \text{for all } \delta \geq 0.
+		\end{equation}
+	\end{lem}
+
+	\begin{proof}
+		Note that $\onorm{\hat f_n - f}_1$ is a function of $X = (X_1,...,X_n)$, denoted $g(X) = g(X_1,...,X_n) = \onorm{\hat f_n - f}_1$ and $g\colon \bbR^n \mapsto \bbR$.
+Let $x^{(k)} = (x_1,...,x_k,...,x_n) \in \bbR^n$ and $x^{'(k)} = (x_1,...,x'_k,...,x_n)\in \bbR^n$ be two vectors.
+The absolute difference between the $g(x^{(k)})$ and $g(x^{'(k)})$ is
+		\begin{align}
+			|g(x^{(k)}) - g(x^{'(k)})|
+			&= \aabs{\int_{-\infty}^{+\infty} \aabs{ \frac{1}{nh} \sum_{i = 1}^n K \of{\frac{t-x_i}{h}} -f(t)  } -\aabs{ \frac{1}{nh} \sum_{i \neq k}^n K \of{\frac{t-x_i}{h}} -f(t) + \frac{1}{nh} K\of{\frac{t-x'_k}{h}}   } dt} \\
+			&\leq \int_{-\infty}^{+\infty} \aabs{\frac{1}{nh} \of{K\of{\frac{t-x_k}{h}} - K\of{\frac{t-x'_k}{h}}  } }dt\\
+			&\leq \frac{1}{nh} \of{ \int_{-\infty}^{+\infty} K\of{\frac{t-x_k}{h}}dt + \int_{-\infty}^{+\infty} K\of{\frac{t-x'_k}{h}}dt }.  \label{eq:e215int}
+		\end{align}
+
+		Let $t_1 = \frac{t-x_k}{h}$. By the definition of $K$, the first integral in equation~\eqref{eq:e215int} are
+		\begin{equation}
+			\int_{-\infty}^{+\infty} K\of{\frac{t-x_k}{h}}dt = h\int_{-\infty}^{+\infty} K\of{t_1}dt_1 = h.
+		\end{equation}
+		Similarly, the second integral $\int_{-\infty}^{+\infty} K\of{\frac{t-x'_k}{h}}dt = h$. Then, we conclude the function $g$ satisfies
+		\begin{equation}\label{eq:e215diff}
+			|g(x^{(k)}) - g(x^{'(k)})|  \leq \frac{2}{n}.
+		\end{equation}
+		The inequality~\eqref{eq:e215diff} also holds for all $k \in [n]$ and for all $x^{(k)},x^{'(k)} \in \bbR^n$.
+Therefore, $g(X)$ satisfies the bounded difference property~\eqref{eq:defbounded} with parameters $(\frac{2}{n},..., \frac{2}{n})$.
+By Theorem~\ref{thm:bounddiff}, we obtain
+		\begin{equation}
+			\bbP\off{\onorm{\hat f_n - f}_1 - \bbE \off{\onorm{\hat f_n - f}_1} \geq \delta} \leq e^{-\frac{n\delta^2}{2}} \leq e^{-\frac{n\delta^2}{8}} .
+		\end{equation}
+
+	\end{proof}
+
+    \subsection{Exercise 2.16}
+    \begin{lem}
+    	Let $\{X_i\}_{i=1}^n$ be a sequence of independent variables taking values in a Hilbert space $\mathbb{H}$.
+Suppose that $\onorm{X_i}_{\mathbb{H}} \leq b_i$ almost surely, for all $i \in [n]$.
+Consider the real valued random variable $S_n = \onorm{\sum_{i=1}^n X_i}_{\mathbb{H}}$.
+The concentration bound of $S_n$ is
+    	\begin{equation}\label{eq:e216a}
+    		\bbP\off{|S_n - \bbE[S_n]| \geq n\delta} \leq 2e^{-\frac{n\delta^2}{8b^2}}, \quad \text{for all } \delta \geq 0,
+    	\end{equation}
+    	where $b^2 = \frac{1}{n} \sum_{i=1}^n b_i^2$. The upper tail probability bound of $S_n$ is
+    	\begin{equation}\label{eq:e216b}
+    		\bbP \off{\frac{S_n}{n} \geq a + \delta} \leq e^{-\frac{n\delta^2}{8b^2}}, \quad \text{for all }\delta \geq 0,
+    	\end{equation}
+    	where $a = \sqrt{\frac{1}{n} \sum_{i=1}^n \bbE \off{\onorm{X_i}^2_{\mathbb{H}}} }$.
+    \end{lem}
+
+    \begin{proof} First, we prove the concentration bound~\eqref{eq:e216a}.
+
+    \vspace{0.2cm}
+    	Note that $S_n$ is a function of $X=(X_1,...,X_n)$, denoted $S_n(X) = S_n(X_1,...,X_n)$, and $S_n(X) \colon \bbR^n \mapsto \bbR$.
+Let $x^{(k)} = (x_1,...,x_k,...,x_n) \in \bbR^n$ and $x^{'(k)} = (x_1,...,x'_k,...,x_n)\in \bbR^n$ be two vectors.
+The absolute difference between $S_n(x^{(k)})$ and $S_n(x^{'(k)})$ is
+    	\begin{align}
+    		\aabs{S_n(x^{(k)}) - S_n(x^{'(k)})} &= \aabs{\onorm{x_1 + \cdots +x_k + \cdots + x_n}_{\mathbb{H}} - \onorm{x_1 + \cdots +x'_k + \cdots + x_n}_{\mathbb{H}} }\\
+    		& \leq \onorm{x_k - x'_k}_{\mathbb{H}}\\
+    		& \leq 2b_k, \label{eq:e216diff}
+    	\end{align}
+    	where the last inequality follows by the boundedness $\onorm{X_k}_{\mathbb{H}} \leq b_k$, for all $k \in [n]$.
+The inequality~\eqref{eq:e216diff} also holds for all $k \in [n]$ and for all $x^{(k)},x^{'(k)} \in \bbR^n$.
+Therefore, $S_n(X)$ satisfies the bounded property~\eqref{eq:defbounded} with parameters $(2b_1,...,2b_n)$.
+By Theorem~\ref{thm:bounddiff}, we obtain
+    	\begin{equation}
+    		\bbP\off{|S_n - \bbE[S_n]| \geq n\delta} \leq 2 e^{- \frac{n \delta^2}{2b^2}} \leq 2 e^{- \frac{n \delta^2}{8b^2}}, \quad \text{for all }\delta \geq 0.
+    	\end{equation}
+
+    	\vspace{0.2cm}
+    	Next, we prove the upper tail probability bound~\eqref{eq:e216b}.
+
+    	\vspace{0.2cm}
+    	The expectation of $S_n(X) = \bbE \off{ \sum_{i=1}^n  \onorm{X_i}_{\mathbb{H}}}$ satisfies
+    	\begin{equation}\label{eq:e216ineq}
+    		 \bbE \off{ \onorm{ \sum_{i=1}^n  X_i}_{\mathbb{H}} } \leq \sqrt{ \bbE \off{ \onorm{\sum_{i=1}^n X_i }_{\mathbb{H}}^2  }} = \sqrt{ \sum_{i=1}^n \bbE \off{ \onorm{X_i}^2_{\mathbb{H}} } } = na,
+    	\end{equation}
+    	where the first inequality follows by the fact that $\bbE[X^2] \geq (\bbE[X] )^2$ and the second equation follows by the independence of $\{X_i\}_{i=1}^n$.
+Therefore, combing the inequality~\eqref{eq:e216ineq} with the concentration bound~\eqref{eq:e216a}, we obtain
+    	\begin{equation}
+    		\bbP \off{\frac{S_n}{n} \geq a + \delta} = \bbP \off{S_n - na \geq n\delta} \leq \bbP \off{S_n - \bbE[S_n] \geq n\delta } \leq e^{-\frac{n\delta^2}{8 b^2}}.
+    	\end{equation}
+    \end{proof}
+
+    \subsection{Exercise 2.17}
+    \begin{lem}[Hanson-Wright Inequality]
+   Let $\mQ = \entry{\mQ_{ij}}  \in \bbR^{n \times n} $ be a positive semi-definite matrix, and $\{X_i\}_{i = 1}^n$ be i.i.d.\ random variables with mean zero, variance 1 and $\sigma-$sub-Gaussian.
+Consider the random variable $Z = \sum_{i= 1}^n \sum_{j = 1}^n \mQ_{ij} X_i X_j$.
+The Hanson-Wright inequality for $Z$ is
+   \begin{equation}
+   	\bbP \off{Z \geq \tr (\mQ) + \sigma t } \leq 2 e^{- \min \offf{\frac{c_1 t}{\onorm{\mQ}_{op}  } , \frac{c_2 t^2}{\onorm{\mQ}^2_F} }}, \quad \text{for all } t \geq 0,
+   \end{equation}
+   where $\onorm{\mQ}_{op}$ is the operator norm of matrix $\mQ$, $\onorm{\mQ}_F$ is the Frobenius norm of matrix $\mQ$, and $c_1,c_2$ are two universal constants independent with $n$ and $t$.
+    \end{lem}
+
+    The proof for Hanson-Wright Inequality under the special case that $X_i \stackrel{i.i.d.}{\sim} N(0,1)$ is below. Let $\sigma  = 1$.
+
+    \begin{proof}
+    	Since the random variables ${X_i}_{i=1}^n$ are independent and $\var(X_i) = 1$ for all $i \in [n]$, the expectation $\bbE[Z] = \sum_{i = 1}^{n} \mQ_{ii} = \tr(\mQ)$.
+Then, the upper tail probability of $Z$ becomes
+    	\begin{equation}
+    		\bbP \off{ Z \geq \tr(\mQ) +  t } = \bbP \off{ \of{  \sum_{i = 1}^{n} \mQ_{ii} (X^2_i - 1) + \sum_{i \neq j} \mQ_{ij} X_i X_j }\geq t   }
+    	\end{equation}
+    	For two random variable $X,Y$ and constant $t$, we have the inequality $ \bbP[X+Y \geq t] \leq \bbP[X \geq t/2] + \bbP[Y \geq t/2]$. Hence, the tail probability is upper bounded as
+    	\begin{equation}
+    		\bbP \off{ Z \geq \tr(\mQ) +  t } \leq \bbP\off{ \sum_{i =1 }^n \mQ_{ii}(X_i^2 - 1) \geq \frac{t}{2}  } + \bbP \off{ \sum_{i \neq j} \mQ_{ii} X_i X_j  \geq \frac{t}{2}} \equalscolon p_1 + p_2.
+    	\end{equation}
+
+    	Consider $p_1$.
+For a standard normal variable $X$,  the squared variable $X^2 \sim \chi^2_1$ and $X^2$ is sub-exponential with parameter $(\nu, b) = (2,4)$.
+Given a constant $c$, $cX^2$ is still sub-exponential with parameters $(2 c, 4 c^2)$.
+Hence, for all $i \in [n]$, the variable $\mQ_{ii} X_i^2$ is sub-exponential with parameters $(2 \mQ_{ii}, 4 \mQ^2_{ii} )$.
+By the Bernstein-type inequalities(non-asymptotic 5.16), the probability $p_1$ is upper bounded as following,
+    	\begin{equation}
+    		p_1 = \bbP \off{\sum_{i=1}^n { \of{ \mQ_{ii}X^2_i - \mQ_{ii}\bbE[X^2_i]   } \geq \frac{t}{2} }} \leq \exp \of{ - \min \offf{  \frac{c t^2}{\sum_{i}^n \mQ^2_{ii} }, \frac{c't}{\max_{i \in [n]  }|\mQ_{ii}| } }},
+    	\end{equation}
+    	where $c$ and $c'$ are two constants independent with $n$ and $t$.
+
+    	\vspace{0.2cm}
+
+    	Consider $p_2$.
+For two i.i.d.\ standard normal variables $X,Y$, the product $XY = \frac{1}{4}(X+Y)^2 - \frac{1}{4}(X-Y )^2$, where $ \frac{1}{2} (X+Y)^2$ and $\frac{1}{2} (X-Y)^2$ follow the chi-square distribution $\chi^2_1$ independently.
+Then, $XY$ is the subtraction of two independent sub-exponential variables.
+Let $D_1$ and $D_2$ be two identical and independent sub-exponential variables with parameters $(\nu, b)$.
+We have
+    	\[ \bbE[e^{\lambda (D_1-D_2)}] = \bbE[e^{\lambda D_1}] \bbE[e^{-\lambda D_2}] \leq e^{v^2 \lambda^2}, \quad \text{for all } |\lambda | < \frac{1}{b}.  \]
+    	Therefore, the subtraction of two identical and independent sub-exponential variables $D_1 -D_2$ is still sub-exponential with parameters $(\sqrt{2}\nu, b)$.
+
+    	\vspace{0.2cm}
+    	Back to $\mQ_{ij} X_i X_j$, where $i \neq j$.
+By the discussion above, the variable $\mQ_{ij} X_i X_j$ is sub-exp onential with parameters $(\sqrt{2} \mQ_{ij}, \mQ_{ij}^2)$.
+By the Bernstein-type inequalities, the probability $p_2$ is upper bounded as following,
+    	\begin{equation}
+    		p_2 = \bbP \off{ \sum_{i \neq j} \mQ_{ii} X_i X_j  \geq \frac{t}{2}} \leq \exp \of{ - \min \offf{  \frac{c'' t^2}{\sum_{i\neq j} \mQ^2_{ij} }, \frac{c'''t}{\max_{i\neq j   }|\mQ_{ij}| } }},
+    	\end{equation}
+    	where $c''$ and $c'''$ are two constants independent with $n$ and $t$.
+
+    	\vspace{0.2cm}
+
+    	Therefore, by the definition of Frobenius norm and the fact that $\max_{i,j\in[n]}|\mQ_{ij}| \leq \onorm{\mQ}_{op} $, we obtain the goal inequality
+    	\begin{equation}
+    		\bbP \off{ Z \geq \tr(\mQ) +  t } \leq p_1 + p_2 \leq 2\exp \of{ - \min \offf{  \frac{c_2 t^2}{\onorm{\mQ}^2_F}, \frac{c_1t}{\onorm{\mQ}_{op} }} },
+    	\end{equation}
+    	for some universal constants $c_1,c_2$ independent with $n$ and $t$.
+    \end{proof}
+
 \end{document}

### 2020-08-09 skills/math/tail_bound/tailbound_solution.tex
TAILBOUND

@@
 
-\subsection{Exercise 2.9}
+\subsection{Exercise 2.9}\label{exe:29}
 \begin{lem}[Sharp upper bounds on binomial tails]
@@
 
-\subsection{Exercise 2.10}
+\subsection{Exercise 2.10}\label{exe:210}
 \begin{lem}[Lower bounds on binomial tails]
@@
     \begin{proof}
-    	Since the random variables ${X_i}_{i=1}^n$ are independent and $\var(X_i) = 1$ for all $i \in [n]$, the expectation $\bbE[Z] = \sum_{i = 1}^{n} \mQ_{ii} = \tr(\mQ)$.
+    	Since the random variables $\{X_i\}_{i=1}^n$ are independent and $\var(X_i) = 1$ for all $i \in [n]$, the expectation $\bbE[Z] = \sum_{i = 1}^{n} \mQ_{ii} = \tr(\mQ)$.
 Then, the upper tail probability of $Z$ becomes
@@
     	\end{equation}
-    	For two random variable $X,Y$ and constant $t$, we have the inequality $ \bbP[X+Y \geq t] \leq \bbP[X \geq t/2] + \bbP[Y \geq t/2]$. Hence, the tail probability is upper bounded as
+    	For two random variables $X,Y$ and a constant $t$, we have the inequality $ \bbP[X+Y \geq t] \leq \bbP[X \geq t/2] + \bbP[Y \geq t/2]$. Hence, the tail probability is upper bounded as
     	\begin{equation}
@@
     	Consider $p_1$.
-For a standard normal variable $X$,  the squared variable $X^2 \sim \chi^2_1$ and $X^2$ is sub-exponential with parameter $(\nu, b) = (2,4)$.
+For a standard normal variable $X$,  the squared variable $X^2 \sim \chi^2_1$ and thus $X^2$ is sub-exponential with parameter $(\nu, b) = (2,4)$.
 Given a constant $c$, $cX^2$ is still sub-exponential with parameters $(2 c, 4 c^2)$.
@@
     	Back to $\mQ_{ij} X_i X_j$, where $i \neq j$.
-By the discussion above, the variable $\mQ_{ij} X_i X_j$ is sub-exp onential with parameters $(\sqrt{2} \mQ_{ij}, \mQ_{ij}^2)$.
+By the discussion above, for all $i\neq j$, the variable $\mQ_{ij} X_i X_j$ is sub-exponential with parameters $(\sqrt{2} \mQ_{ij}, \mQ_{ij}^2)$.
 By the Bernstein-type inequalities, the probability $p_2$ is upper bounded as following,
@@
 
+    \subsection{Exercise 2.18}
+
+    \begin{defn}[Orlicz Norms]\label{def:orlicz}
+     Let $\psi \colon \bbR_+ \mapsto \bbR_+$ be a strictly increasing convex function satisfying $\psi(0) = 0$. The $\psi$-Orlicz norm of a random variable $X$ is defined as
+    \begin{equation}
+    	\onorm{X}_{\psi} \colonequals \inf \offf{t > 0\colon \bbE\off{\psi \of{t^{-1} |X| } } \leq 1  }.
+    \end{equation}
+    The norm $\onorm{X}_{\psi}$ is infinite if there is no finite $t$ such that the expectation $ \bbE\off{\psi \of{t^{-1} |X| } }$ exists.
+    \end{defn}
+
+     Given the function $\psi_q(u) = u^{q}$ for some $q \in [1,+\infty)$, the Orlicz norm is simply the $l_q$-norm $\onorm{X}_q = \of{ \bbE[|X|^q] }^{1/q}$.
+
+    \begin{lem}
+    Let $\psi_q(u) = \exp(u^q) - 1$ be a function for $u \in \bbR_+$ with positive parameter $q \in [1,+\infty)$, and $\onorm{\cdot}_{\psi_q}$ denote the $\psi_q$-Orlicz norm. Then,
+    \begin{enumerate}
+    	\item[(a.)] If $\onorm{X}_{\psi_q} < +\infty$, there exists two positive constants $c_1,c_2$ such that
+    	\begin{equation}\label{eq:lemorlicztail}
+    		\bbP[|X| > t] \leq c_1 \exp \of{-c_2 t^q},\quad \text{for all } t > 0.
+    	\end{equation}
+    	In particular, the constants $c_1 = 2$ and $c_2 = \onorm{X}_{\psi_q}$.
+
+    	\item[(b.)] If the random variable $X$ satisfies the tail bound~\eqref{eq:lemorlicztail}, the Orlicz norm $\onorm{X}_{\psi_q} < +\infty$.
+    \end{enumerate}
+    \end{lem}
+    \begin{proof}
+    	First, we prove $(a)$.
+
+    	\vspace{0.2cm}
+
+    	Let $t_0$ denote the Orlicz norm $\onorm{X}_{\psi_q}$.
+Since $t_0 < +\infty$, the expectation $ \bbE \off{ \exp \of{\lambda |X|^q} }$ exists for $\lambda \in  [0,t^{-q}_0]$.
+Applying Chernoff's bound to $|X|^q$, the upper tail bound of $|X|$ becomes
+    	\begin{equation}\label{eq:e218tailbound}
+    		\bbP [|X|>t] = \bbP \off{\frac{|X|^q}{t^q_0} > \frac{t^q}{t^q_0}} \leq e^{-\lambda t^q} \bbE \off{ \exp \of{\lambda |X|^q} }, \quad \text{for all } t\geq 0, \lambda\in [0,t^{-q}_0].
+    	\end{equation}
+    	 By the definition of Orlicz norm, the expectation satisfies
+    	\begin{equation}\label{eq:e218ep}
+    		\bbE \off{ \exp \of{ t^{-q}_0|X|^q} } = \bbE \off{ \psi( t^{-1}_0 |X| )} + 1 \leq 2.
+    	\end{equation}
+    	Let $\lambda = t^{-q}_0$ and plug the expectation~\eqref{eq:e218ep} to the inequality~\eqref{eq:e218tailbound}. Then, we obtain the tail bound
+    	\begin{equation}
+    		\bbP [|X|>t] \leq 2 e^{- \onorm{X}^{-q}_{\psi_q} t^q },\quad 
+    		\text{for all }  t \geq 0.
+    	\end{equation}
+
+    	\vspace{0.2cm}
+    	Next, we prove $(b)$.
+
+    	\vspace{0.2cm}
+     Let $t_1$ be a positive constant. 	Note the fact that $\bbE[X] = \int_{-\infty}^{+\infty} \bbP(X >t) dt  $. The expectation of $\bbE[ \exp(t^{-q}_1 |X| )  ]$ is
+    	\begin{align}
+    		\bbE[ \exp(t^{-q}_1 |X|^q )  ] &= \int_{0}^{+\infty} \bbP \off{ \exp \of{\frac{|X|^q}{t^q_1}} > \exp \of{\frac{t^q}{t^q_1}} } dt \\
+    		& = \int_{0}^{+\infty} \bbP \off{ |X| > t } dt \\
+    		& \leq  \int_{0}^{+\infty} c_1 \exp(-c_2 t^q) dt\label{eq:e218integral},
+       	\end{align}
+       	where the last inequality follows from the tail bound~\eqref{eq:lemorlicztail}. Let $u = c_2 t^q$. The integral in equation~\eqref{eq:e218integral} becomes
+       	\[  \int_{0}^{+\infty} c_1 \exp(-c_2 t^q) dt = \frac{c_1}{q c_2^{1/q}} \int_{0}^{+\infty} u^{1/q-1}e^{-u} du =  \frac{c_1}{q c_2^{1/q}} \Gamma(1/q) < +\infty.  \]
+       	Therefore, the expectation 
+       	\begin{equation}
+       		\bbE \off{ \psi(t_1^{-1} |X| )} =  \bbE[ \exp(t^{-q}_1 |X|^q )  ] - 1 < +\infty, \quad \text{for all } t_1 > 0,
+       	\end{equation}
+       	which implies the Orlicz norm $\onorm{X}_{\psi_q}$ is definite.
+    \end{proof}
+
+\subsection{Exercise 2.19}
+\begin{lem}[Maxima of Orlicz variables] Recall the definition of Orlicz norm~\ref{def:orlicz}.
+Let $\{X\}_{i=1}^n$ be a sequence of i.i.d.\ zero-mean random variables.
+Given a positive, increasing and convex function $\psi \colon \bbR_+ \mapsto \bbR_+$, the random variable $X_i$ has a finite Orlicz norm $\sigma = \onorm{X_i}_{\psi}$, for all $i \in [n]$.
+Then, the maxima of $X_i$ satisfies
+\begin{equation}
+	\bbE \off{ \max_{i\in[n]} |X_i| } < \sigma \psi^{-1}(n).
+\end{equation}
+
+\end{lem}
+
+\begin{proof}
+	Since $\psi$ is a convex function, by Jensen's inequality, we have
+	\begin{equation}\label{eq:e219max}
+		\psi \of{\bbE \off{\max_{i \in [n]} \frac{|X_i|}{\sigma}  } } \leq \bbE \off{ \max_{i\in[n]} \psi \of{\frac{|X_i|}{\sigma}}   } \leq n \bbE \off{ \psi \of{\frac{|X_1|}{\sigma}} }.
+	\end{equation}
+	By the definition of Orlicz norm, we have $\bbE \off{ \psi \of{\frac{|X_1|}{\sigma}} } \leq 1$.
+Therefore, taking the inverse function $\psi^{-1}$ on both sides of inequality~\eqref{eq:e219max}, we obtain
+	\[  \bbE \off{\max_{i \in [n]} {|X_i|} } \leq \sigma \psi^{-1}(n).  \]
+\end{proof}
+
+\subsection{Exercise 2.20}
+
+\begin{lem}
+	Let $\{X_i\}_{i=1}^n$ be a sequence of independent zero-mean random variables. Suppose that for some fixed integer $m \geq 1$, the random variables satisfy the following moment bound,
+\begin{equation}\label{eq:e220mom}
+	\onorm{X_i}_{2m} \colonequals \of{\bbE[X^{2m}_i] }^{\frac{1}{2m}} \leq C_m, \quad \text{for all } i \in [n].
+\end{equation}
+Then, we have the tail bound
+\begin{equation}
+	\bbP \off{\aabs{ \frac{1}{n}  \sum_{i=1}^n X_i } \geq \delta } \leq B_m \of{\frac{1}{\sqrt{n} \delta}}^{2m}, \quad \text{for all } \delta >0,
+\end{equation}
+where $B_m$ is a universal constant depending on $C_m$ and $m$ only.
+\end{lem}
+
+\textit{Hint:} Let $\{X_i\}_{i=1}^n$ be a sequence of independent zero-mean random variables. The Rosenthal’s inequality is
+\begin{equation}\label{eq:rosenthal}
+	\bbE \off{\of{\sum_{i=1}^n X_i}^{2m}} \leq R_m \offf{\sum_{i=1}^n \bbE \off{X^{2m}_i} + \of{\sum_{i=1}^n \bbE [X^2_i]}^m },
+\end{equation}
+ where $R_m$ is a universal constant depending on $m$ only.
+
+ \begin{proof}
+ 	Applying the $2m$-order central moment version of the Markov's inequality to $\aabs{ \frac{1}{n}  \sum_{i=1}^n X_i } $, we have the tail bound
+ 	\begin{equation}\label{eq:e220tail}
+ 		\bbP \off{\aabs{ \frac{1}{n}  \sum_{i=1}^n X_i } \geq \delta } = \bbP \off{\aabs{  \sum_{i=1}^n X_i } \geq n \delta } \leq (n\delta)^{-2m} \bbE \off{ \of{\sum_{i=1}^n X_i}^{2m} }.
+ 	\end{equation}
+
+ 	By the moment bound, for all $i\in [n]$, we have
+ 	\begin{equation}\label{eq:e220inequ}
+ 		\bbE \off{X^2_i}  \leq \bbE \off{X^{2m}_i} \leq C^{2m}_m.
+ 	\end{equation}
+ 	Plugging the inequality~\eqref{eq:e220inequ} to the Rosenthal's inequality~\eqref{eq:rosenthal}, we obtain
+ 	\begin{equation}\label{eq:e220exp}
+ 		\bbE \off{\of{\sum_{i=1}^n X_i}^{2m}} \leq R_m\of{nC^{2m}_m  + n^m C^{2m^2}_m  }. 
+ 	\end{equation}
+ 	Combining the inequality~\eqref{eq:e220exp} with tai bound~\eqref{eq:e220tail}, we have
+ 	\begin{equation}
+ 		\bbP \off{\aabs{ \frac{1}{n}  \sum_{i=1}^n X_i } \geq \delta } \leq \frac{R_m\of{nC^{2m}_m  + n^m C^{2m^2}_m  } }{n^{2m} \delta^{2m}} = B_m\of{ \frac{1}{\sqrt{n} \delta}}^{2m}, \quad \text{for all }\delta > 0,
+ 	\end{equation}
+ 	where $B_m$ is a universal constant depending on $C_m$ and $m$ only.
+ \end{proof}
+
+ \subsection{Exercise 2.21}
+ \begin{defn}
+ 	Let $X=(X_1,...,X_n)$ be a sequence of i.i.d.\ Bernoulli variables with parameter $1/2$.
+Represent X using a codebook of $N$ binary vectors, denoted $\{z^1,...,z^N\}$, where $z^j = \entry{z^j_i} \in \bbR^n$ is a vector, for all $j \in [N]$.
+The rescaled Hamming distortion is defined as
+ 	\begin{equation}
+ 		\delta \colonequals \bbE \off{ \min_{j \in [N]} \rho_H(X, z^j) } = \bbE \off{ \min_{j \in [N]} \frac{1}{n} \sum_{i = 1}^n  \mathds{1} \off{X_i \neq z^j_i} }. 
+ 	\end{equation}
+ \end{defn}
+
+ The goal of lossy data compression is to find a representation of $X$ with rescaled Hamming distortion as small as possible.
+When we use a codebook with $N = 2^n$ codewords, we always achieve the 0 distortion.
+However, obtaining the distortion with $N = 2^n$ codewords is computationally expensive.
+Our goal is find a proper rate  $R < 1$ such that using a codebook with $N = 2^{Rn}$ codewords achieves a good compression performance.
+
+ \begin{lem}
+
+ 	Let $X=(X_1,...,X_n)$ be a sequence of i.i.d.\ Bernoulli variables with parameter $1/2$. Let $\delta \in (0,1)$. Suppose that the rate $R$ is upper bounded as
+ 	\begin{equation}
+ 		R < D_2(\delta || 1/2 ) = \delta \log_2 \frac{\delta}{1/2} + (1-\delta) \log_2 \frac{1-\delta}{1/2}.
+ 	\end{equation}
+
+ 	Let  $\Delta_R \colonequals R - D_2(\delta || 1/2 )$. If $\Delta_R < 0$, 
+ 	then the probability of achieving distortion $\delta$ goes to 0 as $n$ goes to the infinity, for all codebook $\{z^1,...,z^N\}$.
+If $\Delta_R > 0$, the probability of achieving distortion $\delta$ goes to 1 as $n$ goes to the infinity, for all codebook $\{z^1,...,z^N\}$.
+
+ \end{lem}
+
+ \begin{proof} First, we prove the probability of achieving the distortion under $\Delta_R < 0$.
+
+ \vspace{0.2cm}
+
+ 	Consider the random variables $V^j = \mathds{1}\off{\rho_H(X,z^j) \leq \delta}  $, for all $j \in [N]$.
+Given the codeword $z^j$, the variable $\mathds{1} \off{X_i - z^j_i}$ is still a Bernoulli random variable.
+Recalling the tail bound~\eqref{eq:lembinupper} in Exercise \ref{exe:29}, we have
+ 	\begin{equation}
+ 		\bbP \off{V^j = 1} = \bbP \off{\sum_{i=1}^n \mathds{1} \off{X_i - z^j_i} \leq n\delta} \leq e^{-nD(\delta || 1/2)}, \quad \text{for all }j\in [N].
+ 	\end{equation}
+ 	Since the sum of $V^j$, denoted $V = \sum_{j=1}^N V^j$, follows the Binomial distribution $Bin(N, \bbP \off{V^j = 1} )$, the probability of not achieving the distortion $\delta$ is
+ 	\begin{equation}\label{eq:e221av0}
+ 		\bbP[V = 0] \geq \of{1 - e^{-nD(\delta || 1/2)}}^N.
+ 	\end{equation}
+  Let $R_0 = D_2(\delta || 1/2)$ and $N_0 = 2^{R_0 n} = (2\delta)^{n\delta}(2(1-\delta))^{n(1-\delta)}$.
+By the definition of KL divergence, we have $e^{-nD(\delta || 1/2)} = N^{-1}_0$.
+Note that the ratio $\frac{N}{N_0} = 2^{(\Delta_R)n}$ goes to 0 as $n$ goes to the infinity.
+Hence, the limit of the probability~\eqref{eq:e221av0} is
+  \begin{equation}
+  	\lim_{n \rightarrow +\infty} \bbP[V = 0] \geq \lim_{n \rightarrow +\infty} \of{\of{1 - \frac{1}{N_0}}^{N_0} }^{\frac{N}{N_0}} = 1.
+  \end{equation}
+
+  Therefore, when $\Delta_R < 0$, the probability of achieving distortion $\delta$ goes to 0 as $n$ goes to the infinity.
+
+ 	\vspace{0.2cm}
+ 	Next,  we prove the probability of achieving the distortion under $\Delta_R > 0$.
+
+ 	\vspace{0.2cm}
+ 	Recall the tail bound~\eqref{eq:e2103} in Exercise~\ref{exe:210}. We have the upper bound of $\bbP \off{V^j = 1}$, where
+ 	\begin{equation}\label{eq:e221prob}
+ 		\bbP \off{V^j = 1} = \bbP \off{\sum_{i=1}^n \mathds{1} \off{X_i - z^j_i} \leq n\delta} \geq \frac{1}{n+1}e^{-nD(\delta || 1/2)}, \quad \text{for all }j\in [N].
+ 	\end{equation}
+ 	Then the probability of not achieving the distortion $\delta$ is
+ 	\begin{equation}
+ 		\bbP[V = 0] \leq \of{1 - \frac{1}{n+1} e^{-nD(\delta || 1/2)}}^N.
+ 	\end{equation}
+
+ 	Since the difference $\Delta_R > 0$, then the ratio $\frac{N}{N_0(n+1)} = \frac{2^{(\Delta_R) n} }{n+1}$ goes to infinity as $n$ goes to infinity.
+
+ 	 Hence, the limit of the probability of not achieving the distortion $\delta$ is
+ 	 \begin{equation}
+ 	 	\lim_{n \rightarrow +\infty} \bbP[V = 0] \leq \lim_{n \rightarrow +\infty} \of{1 - \frac{1}{n+1} e^{-nD(\delta || 1/2)}}^N = \lim_{n \rightarrow +\infty}\of{ \of{1 - \frac{1}{(n+1)N_0}}^{(n+1)N_0}}^{\frac{N}{(n+1)N_0}} = 0.
+ 	 \end{equation}
+
+ 	 Therefore, when $\Delta_R > 0$, the probability of achieving distortion $\delta$ goes to 1 as $n$ goes to the infinity.
+
+ 	 \vspace{0.2cm}
+
+ 	 An alternative version of the proof when $\Delta_R > 0$ is below.
+
+ 	 \vspace{0.2cm}
+ 	 Let $p = \bbP \off{V^j = 1}$. By the Bernoulli's inequality, the probability of not achieving the distortion $\delta$ satisfies
+ 	 \begin{equation}
+ 	 	\bbP \off{V = 0} = \of{1 - p}^N \leq \frac{1}{1 - Np} = \frac{Np(1-p)}{Np(1-p) + N^2 p^2} = \frac{(\bbE[V])^2}{\bbE \off{V^2}}.
+ 	 \end{equation}
+ 	  Therefore, by inequality~\eqref{eq:e221prob} the limit of the probability of achieving the distortion $\delta$ is
+ 	  \begin{equation}
+ 	  	\lim_{n \rightarrow +\infty}  \bbP \off{V \geq 1} \geq \lim_{n \rightarrow +\infty}   1- \frac{1}{N p} \geq \lim_{n \rightarrow +\infty}  1 - \frac{n+1}{N N_0 } = 1.
+ 	  \end{equation}
+ \end{proof}
+
+ \subsection{Exercise 2.22}
+
 \end{document}

### 2020-08-12 tensor_regress/unsolved_review.tex
unsolved

\documentclass[11pt]{article}

\usepackage{fancybox}

\usepackage{color}
\usepackage{url}
\usepackage[margin=1in]{geometry}

\setlength\parindent{0pt}
\renewcommand{\textfraction}{0.0}
\renewcommand{\topfraction}{1.0}
%\renewcommand{\textfloatsep}{5mm}

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
%\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\mathtoolsset{showonlyrefs=true}

\newcommand{\of}[1]{\left(#1\right)}
\newcommand{\off}[1]{\left[#1\right]}
\newcommand{\offf}[1]{\left\{#1\right\}}
\newcommand{\aabs}[1]{\left|#1\right|}
\usepackage{colonequals}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\setcounter{figure}{0}   
\setcounter{table}{0}

\newcommand{\cmt}[1]{{\leavevmode\color{red}{#1}}}

\usepackage{dsfont}
\usepackage{multirow}

\DeclareMathOperator*{\minimize}{minimize}

\usepackage{mathtools}
\mathtoolsset{showonlyrefs}
\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}

\input macros.tex

\title{Future Improvement}

\date{\today}
\author{%
Jiaxin Hu
}

\begin{document}

% Makes the title and author information appear.

\maketitle

% Abstracts are required.
\section{Unsolved Review Summary}

\subsection{Comparisons}
\begin{enumerate}
	\item Add comparison on tensor-on-tensor regression~\citep{lock2018tensor,llosa2018tensor,gahrooei2020multiple,raskutti2015convex}. {\bf NIPS R3}.
	\item Add comparison on scalar-on-tensor regression~\citep{chen2019non, zhou2013tensor}. {\bf ICML R1}.
	\item Add comparison on HOLRR~\citep{rabusseau2016low}. Though HOLRR is a special case of the proposed model, it is helpful to show the benefit of the incorporation of covariate on distinct modes, when all covariates are vectorized in a long vector in HOLRR. {\bf AISTAT R3}.
	\item Add comparison on HOPLS~\citep{zhao2012higher} and TPG \citep{yu2016learning} from the perspective of the model. {\bf AISTAT R1}.
	\item Add comparison on paper~\citep{smith2015positive}. {\bf NIPS R4}.
\end{enumerate}

\subsection{Experiments}
\begin{enumerate}
	\item Add comparisons on other scalar-on-tensor and tensor-on-tensor regressions in both simulations and real data analysis. {\bf NIPS R2, R3}.
	\item Add numerical results on large-scale or higher-order response. {\bf NIPS R2, AISTAT R3}.
	\item Show how many iterations and time are needed to converge. {\bf NIPS R2}.
\end{enumerate}

\subsection{Explanations}
\begin{enumerate}
	\item Add explanations on hyperparameter selection for our method and other methods (e.g.\ HOLRR, TPG, HOPLS) we want to compare. {\bf NIPS R2,R4}.
	\item Add explanations on rank selection, including relative supplements. Explain why not use the proposed rank selection strategy in simulations. Explain the computational issues of grid search. {\bf NIPS R3, ICML R1, R3}.
	\item Discuss the existence of $\tB_{true}$ and discuss the mismatching loss $\sum \hat \tB_{ijk} \neq {\tB_{true}}_{ijk}$. {\bf NIPS R4}.
	\item Compare the computational complexity with other methods. {\bf NIPS R2}.
	\item Explain how the 136 subjects in HCP are selected from thousands of subjects in the original data. {\bf NIPS R3}.
\end{enumerate}

\section{Model Equivalence}

\subsection{Our method}
Let $\tY = \entry{\tY_{i_1,...,i_k}} \in \bbR^{d_1,...,d_K}$ be the tensor observation, $\mX_k \in \bbR^{d_k \times p_k}, k \in [K]$ be the matrix covariates. Our tensor regression model is of form
\begin{equation}\label{ourmodel}
	g \of{ \bbE[\tY |\mX_1 ,...,\mX_K ] } = \tB \times_1 \mX_1 \times_2 \cdots \times_K \mX_K,
\end{equation}
where $g(\cdot)$ is the inverse link function, $\tB \in \bbR^{p_1 \times \cdots \times p_K}$ is the order-K tensor-valued coefficient. Reformulating the model~\eqref{ourmodel} into $\prod_{i=1}^K d_k$ univariate regressions, we have
\begin{equation}\label{ourmodel1}
	g \of{ \bbE[\tY_{i_1,...,i_K} |\mX_1 ,...,\mX_K ] }  = \langle \tB, \mX_{1i_1} \circ \cdots \circ \mX_{K i_K}  \rangle, \quad \text{for } i_k \in [d_k], k\in[K],
\end{equation}
where $\mX_{k i_k}$ refers to the $i_k$-th row of $\mX_k$, for all $k \in [K]$, $\langle \cdot, \cdot \rangle$ refers to the inner product of two tensors, and $\circ$ refers to the outer product of two vectors.

\subsection{Scalar-on-tensor regression}
Let $y \in \bbR$ denote the scalar response, $\tX \in \bbR^{p_1 \times \cdots \times p_K}$ denote the order-K tensor predictor, and $\mZ \in \bbR^{p_0}$ denote the vector-valued covariate. Given the i.i.d.\ sample set $\{y^{(i)}, \tX^{(i)}, \mZ^{(i)}\}_{i=1}^n$,  the scalar-on-tensor regression model is
\begin{equation}\label{scalar}
	g \of{\bbE[y^{(i)} | \tX^{(i)}]}  = \alpha + \gamma^T \mZ + \langle \tB, \tX^{(i)}  \rangle, \quad \text{for } i  = 1 ,...,n
\end{equation}

where $g(\cdot)$ is the inverse link function, $\alpha \in \bbR$ is the scalar coefficient, $\gamma \in \bbR^{p_0}$ is the vector-valued coefficient, and $\tB \in \bbR^{p_1 \times \cdots \times p_K}$ is the order-K tensor-valued coefficient.

Ignoring the scalar parameter $\alpha$ and the vector-valued covariate $\mZ$, our model~\eqref{ourmodel1} is equal to the scalar-on-tensor model~\eqref{scalar} by defining
\begin{equation}
	\bbE[y^{(i)} | \tX^{(i)}] \colonequals \bbE[\tY_{i_1,...,i_K} |\mX_1 ,...,\mX_K ] \quad \text{and} \quad \tX^{(i)} \colonequals \mX_{1i_1} \circ \cdots \circ \mX_{K i_K}, \quad \text{for } i = 1,..., n,
\end{equation}
where $n = \prod_{i=1}^K d_k$.

\subsection{Tensor-on-tensor regression}
Let $\mathbb{Y} \in \bbR^{ d_{K+1} \times \cdots \times d_{N}}$ be the tensor observation, $\mathbb{X} \in \bbR^{ d_1 \times \cdots \times d_K}$ be the tensor predictor. Given the i.i.d.\ sample set $\{\mathbb{Y}^{(i)}, \mathbb{X}^{(i)} \}_{i=1}^n$,  the tensor-on-tensor regression is of form
\begin{equation}\label{tensor}
	g \of{\bbE[\mathbb{Y}^{(i)} | \mathbb{X}^{(i)}]} = \langle \mathbb{X}^{(i)}, \mathbb{B}   \rangle_K, \quad \text{for } i = 1,...,n,
\end{equation}

where $g(\cdot)$ is the inverse link function, $\mathbb{B} \in \bbR^{d_1 \times \cdots \times d_N}$ is the order-K tensor-valued coefficient, $\langle \cdot, \cdot \rangle_K$ refers to the contracted tensor product; i.e., for two tensors $\mathbb{P} \in \bbR^{I_1 \times \cdots \times I_L \times d_1 \times \cdots \times d_K}$ and $\mathbb{Q} \in \bbR^{d_1 \times \cdots \times d_K \times J_1 \times \cdots \times J_M}$, the contracted tensor product $\langle \mathbb{P}, \mathbb{Q} \rangle_K \in \bbR^{I_1\times \cdots \times I_L \times J_1 \times \cdots \times J_M}$ and the $(i_1,...,i_L, j_1,...,j_M)$-the entry is
\begin{equation}
	(\langle \mathbb{P}, \mathbb{Q} \rangle_K)_{i_1,...,i_L, j_1,...,j_M} = \sum_{a_1,...,a_K}^{d_1,...,d_K} \mathbb{P}_{i_1,...,i_L, a_1,...,a_K} \mathbb{Q}_{a_1,...,a_K, j_1,...,j_M}.
\end{equation}

Let $N = K$. Then, the contracted product becomes usual inner product, and $\mathbb{Y}^{(i)}$ becomes a scalar for $i \in [n]$. Consequently, the model~\eqref{tensor} degenerates to the scalar-on-tensor regression~\eqref{scalar}, which is equal to our model.

\section{Iteration time table}

See Table~\ref{tab:time}.
\begin{table}[h]
\centering
%\resizebox{1\textwidth}{!}{
\begin{tabular}{|c|c|c|c|}
\hline
Setting &Gaussian& Bernoulli & Poisson  \\ \hline
$d = 30, r = 6$& 2.6 &  6.0& 5.7\\
$d = 30, r = 3$&  0.5 & 1.1&1.3\\
$d = 25, r = 6$&1.4 & 3.7& 4.0\\
$d = 25, r = 3$&0.3& 0.6& 0.6\\
\hline
\end{tabular}
%}
\caption{ Iteration time of Algorithm 1. Numerical values are the running time (seconds) required for one iteration in Algorithm 1 under different settings.  }\label{tab:time}
\end{table}

\bibliography{tensor_wang}
\bibliographystyle{apalike}

\end{document}

### 2020-08-13 skills/math/tail_bound/tailbound_solution.tex
tailbound

@@
 	\begin{equation}\label{eq:chernoff1}
-		\bbP((X - \mu) \geq t) \leq \frac{\bbE\left[e^{\lambda (X-\mu)}\right]}{e^{\lambda t}}, \quad \text{ for all } \lambda\in (-b,b).
+		\bbP((X - \mu) \geq t) \leq \frac{\bbE\left[e^{\lambda (X-\mu)}\right]}{e^{\lambda t}}, \quad \text{ for all } \lambda\in (0,b).
 	\end{equation}
@@
 	\begin{equation}
-		\bbE[Z] = \int_{0}^{+ \infty} \min \offf{1, C \exp \of{-\frac{t^2}{4 \nu^2}}} dt + \int_{0}^{+ \infty} \min\offf{1, C \exp \of{-\frac{t^2}{4 Bt}}} dt   \colonequals I_1 + I_2.
-	\end{equation}
-
-	To evaluate $I_1$, we spilt the integral to avoid the minimization. Solving $1 = C \exp \of{-\frac{t^2}{4 \nu^2}} $, the minimization term becomes
+		\bbE[Z] = \int_{0}^{+ \infty} \min \offf{1, C \exp \of{-\frac{t^2}{4 \nu^2}}} dt + \int_{0}^{+ \infty} \min\offf{1, C \exp \of{-\frac{t^2}{4 Bt}}} dt   \equalscolon I_1 + I_2.
+	\end{equation}
+
+	To evaluate $I_1$, we solve $1 = C \exp \of{-\frac{t^2}{4 \nu^2}} $, and the minimization term becomes
 	\begin{align}
@@
 	\end{align}
-	Therefore, we evaluate $I_1$ as follows,
+	Therefore, let $ y = \frac{t}{2\nu} - \sqrt{\log C}$, we obtain $I_1$ ,
 	\begin{align}
 		I_1 &= \int_{0}^{ 2\nu \sqrt{\log C}} 1 dt + \int_{2\nu \sqrt{\log C}}^{+\infty} C \exp \of{-\frac{t^2}{4 \nu^2}} dt\\
-		(\text{let}\quad y = \frac{t}{2\nu} - \sqrt{\log C})\quad &=  2\nu \sqrt{\log C} + 2\nu\int_{0}^{+\infty} \exp \of{-y^2 - 2y \sqrt{\log C}} dy\\
+		&=  2\nu \sqrt{\log C} + 2\nu\int_{0}^{+\infty} \exp \of{-y^2 - 2y \sqrt{\log C}} dy\\
 		&\leq  2\nu \sqrt{\log C} + 2\nu\int_{0}^{+\infty} \exp \of{-y^2} dy.
@@
 \subsection{Exercise 2.9}\label{exe:29}
-\begin{lem}[Sharp upper bounds on binomial tails]
+\begin{lem}[Sharp upper bounds on binomial tails]\label{lem:sharpbin}
 	Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.\ Bernoulli variables with parameter $\alpha \in (0,1/2]$.
 Consider the binomial random variable $Z_n = \sum_{i=1}^n X_i$.
-The tail probability of $Z_n$ is upper bounded as
+The left tail probability of $Z_n$ is upper bounded as
 	\begin{equation}\label{eq:lembinupper}
@@
 \begin{proof}
-    Applying the Chernoff inequality to the random variable $Z_n$, we have
+    Applying the Chernoff inequality to the random variable $-Z_n$, we have
+
+    \begin{equation}
+    	\bbP[-Z_n \geq -\delta n] =  \bbP[Z_n \leq \delta n]  \leq  \frac{\bbE\off{e^{-\lambda Z_n}}}{e^{-\lambda \delta n}}, \quad \text{for all } \lambda >0.
+    \end{equation}
+    Therefore, the left tail probability of random variable $Z_n$ satisfies
     \begin{equation}\label{eq:e29upper}
-    	\bbP[Z_n \leq \delta n] \leq \frac{\bbE\off{e^{\lambda Z_n}}}{e^{\lambda \delta n}} = \offf {\exp \offf{\log \of{1 - \alpha + \alpha e^{\lambda})}- \lambda \delta }}^n,\quad \text{for all } \lambda \in \bbR,
+    	\bbP[Z_n \leq \delta n] \leq \frac{\bbE\off{e^{-\lambda Z_n}}}{e^{-\lambda \delta n}} = \offf {\exp \offf{\log \of{1 - \alpha + \alpha e^{-\lambda})}+ \lambda \delta }}^n,\quad \text{for all } \lambda >0,
     \end{equation}
     where the second equality follows from the moment generating function of $Z_n$,
-    \[\bbE\off{e^{\lambda Z_n}} = \of{1 - \alpha + \alpha e^{\lambda}}^n. \]
-    The upper bound~\eqref{eq:e29upper} achieves the minimum when $\lambda = \log {\frac{\delta(1-\alpha)}{\alpha(1 - \delta)}}$ by the first-order condition.
+    \[\bbE\off{e^{-\lambda Z_n}} = \of{1 - \alpha + \alpha e^{-\lambda}}^n. \]
+    The upper bound~\eqref{eq:e29upper} achieves the minimum when $\lambda = -\log {\frac{\delta(1-\alpha)}{\alpha(1 - \delta)}}$ by the first-order condition.
 Plugging the minimizer into equation~\eqref{eq:e29upper}, we obtain the result
@@
 Hence, the upper bound~\eqref{eq:lembinupper} is strictly tighter than Hoeffding bound~\eqref{eq:e29hoeff}.
+
+\end{proof}
+
+\begin{cor}[Sharp upper bounds on binomial right tails] Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.\ Bernoulli variables with parameter $\alpha \in [1/2,1)$.
+Consider the binomial random variable $Z_n = \sum_{i=1}^n X_i$.
+The left tail probability of $Z_n$ is upper bounded as
+	\begin{equation}\label{eq:lembinupper}
+		\bbP\off{Z_n \geq \delta n} \leq e^{-n D\of{\delta \Vert \alpha}}, \quad \text{for all }\delta \in (\alpha,1),
+	\end{equation}
+	where $ D\of{\delta \Vert \alpha}$ is the KL divergence defined in Lemma~\ref{lem:sharpbin}.
+\end{cor}
+
+\begin{proof}
+	Applying the Chernoff inequality directly to the random variable $Z_n$, we have
+	 \begin{equation}
+    	\bbP[Z_n \geq \delta n] \leq \frac{\bbE\off{e^{\lambda Z_n}}}{e^{\lambda \delta n}} = \offf {\exp \offf{\log \of{1 - \alpha + \alpha e^{\lambda})}- \lambda \delta }}^n,\quad \text{for all } \lambda >0.
+    \end{equation}
+
+    Plugging the minimizer $\lambda = \log {\frac{\delta(1-\alpha)}{\alpha(1 - \delta)}}$ into the above inequality, we obtain the results
+    \[ \bbP[Z_n \geq \delta n]  \leq e^{-n D\of{\delta \Vert \alpha}}, \quad \text{for all }\delta \in (\alpha,1).\]
 
@@
 	\end{align}
-	Taking the log and dividing by $n$ on the both sides of the inequality~\eqref{eq:e210pro1} , we obtain
+	Taking the log and dividing by $n$ on the both sides of the inequality~\eqref{eq:e210pro1}, we obtain
 	\[ \frac{1}{n} \log \bbP[Z_n \leq n\delta ] \geq \frac{1}{n} \log \begin{pmatrix}
@@
 	\end{equation}
-	Multiplying $n$ and exponentiating on the both sides of the inequality~\eqref{eq:e210pro3}, we obtain the result
+	Multiplying by $n$ and take exponential on both sides of the inequality~\eqref{eq:e210pro3}, we obtain the result
 	\begin{equation}
@@
 
-\subsection{Exercise 2.11}
+\subsection{Exercise 2.11}\label{Gausmax}
 \begin{lem}[Gaussian maxima]
-Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.
-normal random variables following $N(0,\sigma^2)$.
+Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.\ normal random variables following $N(0,\sigma^2)$.
 Consider the random variable $Z_n \colonequals \max_{i\in[n]} |X_i|$.
-Since the tail bound $\bbP[U \geq t] \leq \sqrt{\frac{2}{\pi}} \frac{1}{t} e^{-t^2 /2}$ holds for all standard normal random variable $U$, the expectation of $Z_n$ is upper bounded as follows
+The expectation of $Z_n$ is upper bounded as follows
 \begin{equation}\label{eq:e2111}
@@
 \end{equation}
-\iffalse
+
 The lower bound of $\bbE[Z_n] $ is
 \begin{equation}\label{eq:e2112}
-	\bbE[Z_n] \geq (1- 1/e) \sqrt{2\sigma^2 \log n} , \quad \text{for all }n \geq 5.
-\end{equation}
-\fi
-
-\end{lem}
+	\bbE[Z_n] \geq (1- 1/e) \sqrt{\sigma^2 \log n} , \quad \text{for all }n \geq 6.
+\end{equation}
+\end{lem}
+
+The lower bound~\eqref{eq:e2112} is different to the original Exercise up to the constant.
 
@@
 	 \end{equation}
-	 Hence, the last integral in equation~\eqref{eq:e211pro1} follows
+	 Hence, let $ u = \frac{ t}{\sigma}$, the last integral in equation~\eqref{eq:e211pro1} is upper bounded as
 	 \begin{align}
-	 	 \int_{c}^{+\infty} n\bbP[|X_1| \geq t] dt &\leq \frac{2n \sigma}{c} \sqrt{\frac{2}{\pi}} \int_{c}^{+\infty} e^{-\frac{t^2}{2\sigma^2} }dt \\ (\text{let } u = \frac{ t}{\sigma}) \quad  
+	 	 \int_{c}^{+\infty} n\bbP[|X_1| \geq t] dt &\leq \frac{2n \sigma}{c} \sqrt{\frac{2}{\pi}} \int_{c}^{+\infty} e^{-\frac{t^2}{2\sigma^2} }dt \\  
 	 	 %& = \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}   \int_{\frac{c}{\sigma}}^{+\infty} e^{-\frac{u^2}{2}} dt \\
@@
 
+	 \vspace{0.2cm}
+	 Next, we prove the equation~\eqref{eq:e2112}.
+
+	 \vspace{0.2cm}
+
+	 The expectation of $Z_n$ is lower bounded as
+	 \begin{equation}
+	 	\bbE \off{Z_n} \geq \sigma \sqrt{\log n} \bbP \off{Z_n > \sigma \sqrt{\log n}}  = \sigma \sqrt{\log n} \off{1 - \of{1 - \bbP \off{|X_i| > \sigma \sqrt{\log n}}   }^n}.
+	 \end{equation}
+
+	 Since $|X_i|$ follows the half-normal distribution, the upper tail probability of $|X_i|$ is
+	 \begin{equation}
+	 	\bbP \off{ |X_i| > \sigma \sqrt{\log n}} = 1 - \text{erf}\of{\frac{\sqrt{\log n}}{\sqrt{2} } } \geq 1 - \sqrt{1 - n^{-\frac{2}{\pi}}},
+	 \end{equation} 
+	 where $\text{erf}(\cdot)$ is the CDF function of half-normal distribution, and the last inequality follows from the fact that $\text{erf}(x) \leq \sqrt{1 - \exp \of{\frac{4}{\pi} x^2 }}$.
+When $n \geq 6$, we have $1 - \sqrt{1 - n^{-\frac{2}{\pi}}} \geq \frac{1}{n}$, and thus the expectation satisfies
+	 \begin{equation}
+	 	\bbE \off{Z_n} \geq \sigma \sqrt{\log n} \off{1 - \of{1 - \frac{1}{n}   }^n} \geq \of{1 - \frac{1}{e}} \sigma \sqrt{\log n}, \quad \text{for all }n \geq 6.
+	 \end{equation}
+
 	 \iffalse
-	 \vspace{0.2cm}
-	 Next, we prove the equation~\eqref{eq:e2112}.
-
-	 \vspace{0.2cm}
-
 	 By Markov's inequality, the expectation of $Z_n$ satisfies
@@
 \end{proof}
+
+\begin{rmk}
+	The absolute Gaussian maxima $Z_n$ concentrates around $\sigma \sqrt{\log n}$ for large $n$.
+Standard Gaussian variables concentrate around 0, while the maxima of the $n$ standard Gaussian variables is no longer a Gaussian variables.
+The scale of Gaussian maxima scales as $\sqrt{n}$, however, the sum of $n$ i.i.d.\ Gaussian variables scales as $n$.
+\end{rmk}
 
@@
 \end{proof}
+
+\begin{rmk}
+	The maxima for $n$ i.i.d.\ sub-Gaussian variables also share the similar concentration property as Gaussian variables in Exercise~\ref{Gausmax}.
+The maxima $\max_{i \in [n]} X_i$ concentrates around the sub-Gaussian parameter $\sigma$.
+\end{rmk}
 
@@
 
+\begin{rmk}
+	Without the independence assumption, the sum of sub-Gaussian variables is still a sub-Gaussian variable.
+Under the independence assumption, the product of two sub-Gaussian variables is a sub-exponential variable.
+\end{rmk}
+
 \subsection{Exercise 2.14}
@@
 
+	\iffalse
 	\vspace{0.2cm}
@@
 	\end{enumerate}
-	\end{proof}
+
+	\fi
+	\end{proof}
+
+	\begin{rmk}
+		The concentration around mean implies the concentration around median for $n$ variables, vice versa. 
+	\end{rmk}
 
@@
 
+	\begin{rmk}
+		Kernel density estimator satisfies the bounded difference property and possesses a concentration property around the mean.
+	\end{rmk}
+
     \subsection{Exercise 2.16}
@@
     	\end{equation}
-    	where $a = \sqrt{\frac{1}{n} \sum_{i=1}^n \bbE \off{\onorm{X_i}^2_{\mathbb{H}}} }$.
+    	where $a = \sqrt{\frac{1}{n^2} \sum_{i=1}^n \bbE \off{\onorm{X_i}^2_{\mathbb{H}}} }$.
     \end{lem}
@@
 
+    \begin{rmk}
+    	The concentration inequalities also apply to the Hilbert space. 
+    \end{rmk}
+
     \subsection{Exercise 2.17}
@@
 
-    The proof for Hanson-Wright Inequality under the special case that $X_i \stackrel{i.i.d.}{\sim} N(0,1)$ is below. Let $\sigma  = 1$.
+    The proof for Hanson-Wright Inequality under the special case  $X_i \stackrel{i.i.d.}{\sim} N(0,1)$ is below. Let $\sigma  = 1$.
 
@@
 Hence, for all $i \in [n]$, the variable $\mQ_{ii} X_i^2$ is sub-exponential with parameters $(2 \mQ_{ii}, 4 \mQ^2_{ii} )$.
-By the Bernstein-type inequalities(non-asymptotic 5.16), the probability $p_1$ is upper bounded as following,
+By the Bernstein-type inequalities (see Step 1 in reference), the probability $p_1$ is upper bounded as following,
     	\begin{equation}
@@
 We have
-    	\[ \bbE[e^{\lambda (D_1-D_2)}] = \bbE[e^{\lambda D_1}] \bbE[e^{-\lambda D_2}] \leq e^{v^2 \lambda^2}, \quad \text{for all } |\lambda | < \frac{1}{b}.  \]
+    	\[ \bbE[e^{\lambda (D_1-D_2)}] = \bbE[e^{\lambda D_1}] \bbE[e^{-\lambda D_2}] \leq e^{2v^2 \lambda^2}, \quad \text{for all } |\lambda | < \frac{1}{b}.  \]
     	Therefore, the subtraction of two identical and independent sub-exponential variables $D_1 -D_2$ is still sub-exponential with parameters $(\sqrt{2}\nu, b)$.
@@
     	\vspace{0.2cm}
-    	Back to $\mQ_{ij} X_i X_j$, where $i \neq j$.
+    	Back to $\mQ_{ij} X_i X_j$.
 By the discussion above, for all $i\neq j$, the variable $\mQ_{ij} X_i X_j$ is sub-exponential with parameters $(\sqrt{2} \mQ_{ij}, \mQ_{ij}^2)$.
@@
     \end{proof}
+
+    Reference: http://www-personal.umich.edu/~rudelson/papers/rv-Hanson-Wright.pdf.
+
+    \begin{rmk}
+    	For a random vector $X$, Hanson-Wright Inequality implies the concentration property for the matrix product $X^T \mQ X$, where $\mQ$ is a positive semi-definite matrix.
+    \end{rmk}
 
@@
     \begin{enumerate}
-    	\item[(a.)] If $\onorm{X}_{\psi_q} < +\infty$, there exists two positive constants $c_1,c_2$ such that
+    	\item[(a.)] if $\onorm{X}_{\psi_q} < +\infty$, there exist two positive constants $c_1,c_2$ such that
     	\begin{equation}\label{eq:lemorlicztail}
@@
 
-    	\item[(b.)] If the random variable $X$ satisfies the tail bound~\eqref{eq:lemorlicztail}, the Orlicz norm $\onorm{X}_{\psi_q} < +\infty$.
+    	\item[(b.)] if the random variable $X$ satisfies the tail bound~\eqref{eq:lemorlicztail}, the Orlicz norm $\onorm{X}_{\psi_q} < +\infty$.
     \end{enumerate}
@@
 Since $t_0 < +\infty$, the expectation $ \bbE \off{ \exp \of{\lambda |X|^q} }$ exists for $\lambda \in  [0,t^{-q}_0]$.
-Applying Chernoff's bound to $|X|^q$, the upper tail bound of $|X|$ becomes
+Applying Chernoff's bound to $|X|^q$, the upper tail bound of $|X|$ is
     	\begin{equation}\label{eq:e218tailbound}
@@
 
+    \begin{rmk}
+    	For a random variable $X$, the existence of $X$'s Orlicz norm implies the boundedness of the tail probability of $X$.
+    \end{rmk}
+
 \subsection{Exercise 2.19}
@@
 
+\begin{rmk}
+	The maxima of Orlicz variables concentrates around the Orlicz norm $\sigma$ and scales as $\psi^{-1}(n)$.
+In previous exercises, the maxima of sub-Gaussian variables concentrates around the sub-Gaussian variable $\sigma$ and scales at $\log (n)$.
+\end{rmk}
+
 \subsection{Exercise 2.20}
@@
 \begin{lem}
-	Let $\{X_i\}_{i=1}^n$ be a sequence of independent zero-mean random variables. Suppose that for some fixed integer $m \geq 1$, the random variables satisfy the following moment bound,
+	Let $\{X_i\}_{i=1}^n$ be a sequence of independent zero-mean random variables. Suppose that for some fixed integer $m \geq 1$ the random variables satisfy the following moment bound,
 \begin{equation}\label{eq:e220mom}
@@
  \end{proof}
+
+ \begin{rmk}
+ 	For $n$ independent variables $X_i$, the moment condition for $X_i$ implies the boundedness of the tail probability for the sample mean $\frac{1}{n} \sum_{i=1}^n X_i$.
+ \end{rmk}
 
@@
 However, obtaining the distortion with $N = 2^n$ codewords is computationally expensive.
-Our goal is find a proper rate  $R < 1$ such that using a codebook with $N = 2^{Rn}$ codewords achieves a good compression performance.
+Our goal is to find a proper rate  $R < 1$ such that using a codebook with $N = 2^{Rn}$ codewords achieves a good compression performance.
 
@@
 
- 	 An alternative version of the proof when $\Delta_R > 0$ is below.
+ 	 An alternative version of the proof under $\Delta_R > 0$ is below.
 
@@
  	 \begin{equation}
- 	 	\bbP \off{V = 0} = \of{1 - p}^N \leq \frac{1}{1 - Np} = \frac{Np(1-p)}{Np(1-p) + N^2 p^2} = \frac{(\bbE[V])^2}{\bbE \off{V^2}}.
- 	 \end{equation}
- 	  Therefore, by inequality~\eqref{eq:e221prob} the limit of the probability of achieving the distortion $\delta$ is
- 	  \begin{equation}
- 	  	\lim_{n \rightarrow +\infty}  \bbP \off{V \geq 1} \geq \lim_{n \rightarrow +\infty}   1- \frac{1}{N p} \geq \lim_{n \rightarrow +\infty}  1 - \frac{n+1}{N N_0 } = 1.
- 	  \end{equation}
- \end{proof}
+ 	 	\bbP \off{V = 0} = \of{1 - p}^N \leq \frac{1}{1 - Np} 
+ 	 	%= \frac{Np(1-p)}{Np(1-p) + N^2 p^2} = \frac{(\bbE[V])^2}{\bbE \off{V^2}}.
+ 	 \end{equation}
+ 	  Therefore, by inequality~\eqref{eq:e221prob}, the limit of the probability of achieving the distortion $\delta$ is
+ 	  \begin{equation}
+ 	  	\lim_{n \rightarrow +\infty}  \bbP \off{V \geq 1} \geq \lim_{n \rightarrow +\infty}   1- \frac{1}{1- N p} \geq \lim_{n \rightarrow +\infty}  1 - \frac{N_0(n+1)}{N_0(n+1) - N } = 1.
+ 	  \end{equation}
+ \end{proof}
+
+ \begin{rmk}
+ 	To compress a random vector with Bernoulli entries and achieve a desired rescaled Hamming distortion $\delta$, the number of necessary codewords, $N$, should be larger than  $2^{D(\delta ||1/2) n}$, where $D(\delta ||1/2)$ is the KL divergence between the two Bernoulli distributions with parameter $\delta$ and $1/2$ and $n$ is the size of the random vector.
+ \end{rmk}
 

### 2020-08-27 tensor_regress/jasa-latexsource/tensor-jasa.tex
jasa-version

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL

\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{algorithm}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{1}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{-.3in}%
\addtolength{\topmargin}{-.8in}%

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}
\newtheorem{cor}{Corollary}[section]

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\usepackage{appendix}
\usepackage{wrapfig}
\mathtoolsset{showonlyrefs}

\input macros.tex

\usepackage[english]{babel}

\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}

\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\begin{document}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if1\blind
{
  \title{\bf Exponential Family Tensor Regression with Covariates on Multiple Modes}
  \author{Jiaxin Hu, Zhuoyan Xu, and Miaoyan Wang\\
    Department of Statistics, University of Wisconsin-Madison}
  \maketitle
} \fi

\if0\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Exponential Family Tensor Regression with Covariates on Multiple Modes}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
We consider the problem of tensor-response regression given covariates on multiple modes. Such data problems arise frequently in applications such as neuroimaging, network analysis, and spatial-temporal modeling. We propose a new family of tensor response regression models that incorporate covariates, and establish the theoretical accuracy guarantees. Unlike earlier methods, our estimation allows high-dimensionality in both the tensor response and the covariate matrices on multiple modes. An efficient alternating updating algorithm is further developed. Our proposal handles a broad range of data types, including continuous, count, and binary observations. Through simulation and applications to two real datasets, we demonstrate the outperformance of our approach over the state-of-art.
\end{abstract}

\noindent%
{\it Keywords:}  Tensor-response models, Multiple covariates, Dimension reduction, Reduced-rank regression, Generalized linear models 
\vfill

\newpage
\spacingset{1.5} % DON'T change the spacing!
\section{Introduction}
\label{sec:intro}

Many contemporary scientific and engineering studies collect multi-way array data, a.k.a.\ tensors, accompanied by additional covariates. One example is neuroimaging analysis~\cite{sun2017store,zhou2013tensor}, in which the brain connectivity networks are collected from a sample of individuals. Researchers are often interested in identifying connection edges that are affected by individual characteristics such as age, gender, and disease status (see Figure~\ref{fig:intro1}a). Another example is in the field of network analysis~\cite{baldin2018optimal,hoff2005bilinear}. A typical social network consists of nodes that represent people and edges that represent friendships. In addition, features on nodes and edges are often available, such as people's personality and demographic location. It is of keen scientific interest to identify the variation in the connection patterns (e.g., transitivity, community) that can be attributable to the node features.

{\bf Our contributions.} This paper presents a general treatment to these seemingly different problems. We formulate the learning task as a regression problem, with tensor observation serving as a response, and the node features and/or their interactions forming the predictor. Figure~\ref{fig:intro1}b illustrates the general set-up we consider. The regression approach allows the identification of variation in the data tensor that is explained by the covariates. Our model greatly improves the classical tensor regression~\cite{rabusseau2016low,yu2016learning} by incorporating covariates from multiple modes and the interactions thereof. The statistical convergence of our estimator is established, and we quantify the gain in predictive power.

A related contribution is that our method allows a broad range of tensor types, including continuous, count, and binary observations. While previous tensor regression methods~\cite{zhao2012higher,yu2016learning} are able to analyze Gaussian responses, none of them is suitable for exponential distribution family of tensors. We develop a generalized tensor regression framework, and as a by product, our models allows heteroscedasticity by relating the variance of tensor entry to its mean.  This flexibility is particularly important in practice, because social network, brain imaging, or gene expression datasets are often non-Gaussian.

\begin{figure*}[t]
\begin{center}
\includegraphics[width=16cm]{demo.pdf}
\end{center}
\caption{Examples of tensor response regression model with covariates on multiple modes. (a) Network population model. (b) Spatial-temporal growth model. }\label{fig:intro1}
\end{figure*}

{\bf Related work.} Our work is closely related to but also clearly distinctive from several lines of previous work. The first is a class of \emph{unsupervised} tensor decomposition such as Tucker and CP decomposition~\cite{de2000multilinear,hong2018generalized,zhang2018tensor}. Tucker decomposition is an unsupervised method that finds a low-rank representation of a data tensor. In contrast, our model is a {\it supervised} tensor model that identifies the association between a data tensor and multiple covariates. The low-rank structure is determined jointly by the tensor response and matrix covariates. The second line of work studies the network-response model~\cite{rabusseau2016low, li2017parsimonious}. Earlier development of this model focuses mostly on binary data in the presence of dyadic covariates~\cite{hoff2005bilinear}. We will demonstrate the enhanced accuracy as the order of data grows, and establish the general theory for exponential family which is arguably better suited to various data types.

\section{Preliminaries}

We begin by reviewing the basic properties about tensors~\citep{kolda2009tensor}. We use $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ to denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor. The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as
\begin{equation}
\tY \times_1\mX_1\ldots \times_K \mX_K=\entry{\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}x_{j_1,i_1}^{(1)}\ldots x_{j_K,i_K}^{(K)}},
\end{equation}
which results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor. For ease of presentation, we use shorthand notion $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product. For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as $\langle \tY, \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}$. The Frobenius norm of tensor $\tY$ is defined as $\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}$. A higher-order tensor can be reshaped into a lower-order object~\citep{wang2017operator}. We use $\text{vec}(\cdot)$ to denote the operation that reshapes the tensor into a vector, and $\text{Unfold}_k(\cdot)$ the operation that reshapes the tensor along mode-$k$ into a matrix of size $d_k$-by-$\prod_{i\neq k}d_i$. The Tucker rank of an order-$K$ tensor $\tY$ is defined as a length-$K$ vector $\mr=(r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\text{Unfold}_k(\tY)$, $k=1,\ldots,K$. We use lower-case letters (e.g.,\ $a,b,c$) for scalars/vectors, upper-case boldface letters (e.g.,\ $\mA,\mB,\mC,$) for matrices, and calligraphy letters (e.g.,\ $\tA, \tB, \tC,$) for tensors of order three or greater. We let $\mI_d$ denote the $d \times d$ identity matrix, $[d]$ denote the $d$-set $\{1,\ldots,d\}$, and allow an $\mathbb{R}\to \mathbb{R}$ function to be applied to tensors in an element-wise manner.

\section{Motivation and model}\label{sec:model}
Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor. Suppose we observe covariates on some of the $K$ modes. Let $\mX_k\in\mathbb{R}^{d_k\times p_k}$ denote the available covariates on the mode $k$, where $p_k\leq d_k$. We propose a multilinear structure on the conditional expectation of the tensor. Specifically, 
\begin{align}\label{eq:tensormodel}
&\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)=f(\Theta),\ \text{with}\\
&\Theta =\tB\times\{\mX_1,\ldots,\mX_K\} ,
\end{align}
where  $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the parameter tensor of interest, $f(\cdot)$ applied to $\Theta$ entrywise is a known link function in generalized linear model, and $\times$ denotes the tensor Tucker product. The choice of link function depends on the distribution of the response data. Some common choices are identity link for Gaussian tensor, logistic link for binary tensor, and $\exp(\cdot)$ link for Poisson tensor (see Table~\ref{table:link}).

\begin{table}[htb]
\centering
\begin{tabular}{c|ccc}
Data type &Gaussian & Poisson& Bernoulli\\
\hline
Domain $\mathbb{Y}$& $\mathbb{R}$&$\mathbb{N}$&$\{0,1\}$\\
 $b(\theta)$&$\theta^2/2$& $\exp(\theta)$&$\log (1+\exp(\theta))$\\
 link $f(\theta)$&$\theta$&$\exp(\theta)$&$(1+\exp(-\theta))^{-1}$
\end{tabular}
\caption{Canonical links for common distributions.}\label{table:link}
\end{table}

We give three concrete examples of tensor regression that arise in practice. 
\begin{example}[Spatio-temporal growth model]
Let $\tY=\entry{y_{ijk}}\in\mathbb{R}^{d \times m\times n}$ denote the pH measurements of $d$ lakes at $m$ levels of depth and for $n$ time points. Suppose the sampled lakes belong to $q$ types, with $p$ lakes in each type. Let $\{\ell_j\}_{j\in[m]}$ denote the sampled depth levels and $\{t_k\}_{k\in[n]}$ the time points. Assume that the expected pH trend in depth is a polynomial of order $r$ and that the expected trend in time is a polynomial of order $s$. Then, the spatio-temporal growth model can be represented as
\begin{equation}\label{eq:time}
\mathbb{E}(\tY|\mX_1,\mX_2,\mX_3)=\tB\times\{\mX_1,\mX_2,\mX_3\},
\end{equation}
where $\tB\in\mathbb{R}^{p\times (r+1)\times (s+1)}$ is the coefficient tensor of interest, $\mX_1=\text{blockdiag}\{\mathbf{1}_p,\ldots,\mathbf{1}_p\}\in \{0,1\}^{d\times q}$ is the design matrix for lake types, 
\[
\mX_2=
\begin{pmatrix}
1 & \ell_1&\cdots &\ell^{r}_1\\
1 & \ell_2&\cdots &\ell^{r}_2\\
\vdots &\vdots&\ddots&\vdots\\
1&\ell_{m}&\cdots&\ell^{r}_{m}
\end{pmatrix},\
\mX_3=
\begin{pmatrix}
1 & t_1&\cdots &t^{s}_1\\
1 & t_2&\cdots &t^{s}_2\\
\vdots &\vdots&\ddots&\vdots\\
1&t_{n}&\cdots&t^{s}_{n}
\end{pmatrix}
\]
are the design matrices for spatial and temporal effects, respectively. The model~\eqref{eq:time} is a higher-order extension of the ``growth curve'' model originally proposed for matrix data~\citep{gabriel1998generalised,potthoff1964generalized,srivastava2008models}. Clearly, the spatial-temporal model is a special case of our tensor regression model, with covariates available on each of the three modes.
\end{example}

\begin{example}[Network population model] 
Network response model is recently developed in the context of neuroimanig analysis. The goal is to study the relationship between network-valued response and the individual covariates. Suppose we observe $n$ i.i.d.\ observations $\{(\mY_i, \mx_i): i=1,\ldots,n\}$, where $\mY_i\in\{0,1\}^{d\times d}$ is the brain connectivity network on the $i$-th individual, and $\mx_i\in\mathbb{R}^p$ is the individual covariate such as age, gender, cognition, etc. The network-response model~\citep{rabusseau2016low, zhang2018network} has the form
\begin{equation}\label{eq:network}
\text{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n
\end{equation}
where $\tB\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor of interest.

The model~\eqref{eq:network} is a special case of our tensor-response model, with covariates on the last mode of the tensor. Specifically, stacking $\{\mY_i\}$ together
yields an order-3 response tensor $\tY\in\{0,1\}^{d\times d\times n}$, along with covariate matrix $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$. Then, the model~\eqref{eq:network} can be written as
\[
\text{logit}(\mathbb{E}(\tY|\mX))=\tB\times_3 \mX=\tB\times\{\mI_d, \mI_d, \mX\}.
\]
 \end{example}

 \begin{example}[Dyadic data with node attributes] Dyadic dataset consists of measurements on pairs of objects or under a pair of conditions. Common examples include networks and graphs. Let $\tG=(V,E)$ denote a network, where $V=[d]$ is the node set of the graph, and $E\subset V\times V$ is the edge set. Suppose that we also observe covariate $\mx_i\in\mathbb{R}^p$ associated to each $i\in V$. A probabilistic model on the graph $\tG=(V,E)$ can be described by the following matrix regression. The edge connects the two vertices $i$ and $j$ independently of other pairs, and the probability of connection is modeled as
\begin{equation}\label{eq:edge}
 \text{logit}(\mathbb{P}((i,j)\in E)=\mx^T_i\mB\mx_j=\langle \mB, \mx^T_i\mx_j\rangle.
 \end{equation}
The above model has demonstrated its success in modeling transitivity, balance, and communities in the networks~\citep{hoff2005bilinear}. We show that our tensor regression model~\eqref{eq:tensormodel} also incorporates the graph model as a special case. Let $\tY=\entry{y_{ij}}$ be a binary matrix where $y_{ij}=\mathds{1}_{(i,j)\in E}$. Define $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$. Then, the graph model~\eqref{eq:edge} can be expressed as
 \[
 \text{logit}(\mathbb{E}(\mY|\mX))=\tB\times\{\mX,\mX\}.
 \]
\end{example}

In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by the covariates. The regression coefficient $\tB$ in our model model~\eqref{eq:tensormodel} serves this goal by collecting the effects of covariates and the interaction thereof.  
To encourage the sharing among effects, we assume that the coefficient tensor $\tB$ lies in a low-dimensional parameter space:
\begin{equation}\label{eq:rank}
\small \tP_{r_1,\ldots,r_K}=\{\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}: r_k(\tB)\leq r_k \text{ for all } k\in[K]\},
\end{equation}
where $r_k(\tB)\leq p_k$ is the Tucker rank at mode $k$ of the tensor. The low-rank assumption is plausible in many scientific applications. In brain imaging analysis, for instance, it is often believed that the brain nodes can be grouped into fewer communities, and the numbers of communities are much smaller than the number of nodes. The low-rank structure encourages the shared information across tensor entries, thereby greatly improving the estimation stability. When no confusion arises, we drop the subscript $(r_1,\ldots,r_K)$ and write $\tP$ for simplicity.

Our tensor regression model is able to incorporate covariates on any subset of modes, whenever available. Without loss of generality, we denote by $\tX=\{\mX_1,\ldots,\mX_K\}$ the covariates in all modes and treat $\mX_k=\mI_{d_k}$ if the mode-$k$ has no (informative) covariate. Then, the final form of our tensor regression model can be written as:
\begin{align}\label{eq:tensormodel2}
&\mathbb{E}(\tY|\tX)=f(\Theta),\quad \Theta =\tB\times\{\mX_1,\ldots,\mX_K\},\notag\\
&\text{where}\ \text{rank}(\tB)\leq (r_1,\ldots,r_K),
\end{align}
where the entries of $\tY$ are independent r.v.'s conditional on $\tX$, and $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$ is the low-rank coefficient tensor of interest. We comment that other forms of tensor low-rankness are also possible, and here we choose Tucker rank just for parsimony. Similar models can be derived using various notions of low-rankness based on CP decomposition~\citep{hitchcock1927expression} and train decomposition~\citep{oseledets2011tensor}.

\section{Rank-constrained likelihood-based estimation}
We develop a likelihood-based procedure to estimate the coefficient tensor $\tB$ in~\eqref{eq:tensormodel2}. We adopt the exponential family as a flexible framework for different data types. In a classical generalized linear model (GLM) with a scalar response $y$ and covariate $\mx$, the density is expressed as:
\[
p(y|\mx, \boldsymbol{\beta})=c(y,\phi)\exp\left({y\theta- b(\theta) \over \phi}\right)\ \text{with}\ \theta=\boldsymbol{\beta}^T\mx,
\]
where $b(\cdot)$ is a known function, $\theta$ is the linear predictor, $\phi>0$ is the dispersion parameter, and $c(\cdot)$ is a known normalizing function. The choice of link functions depends on the data types and on the observation domain of $y$, denoted $\mathbb{Y}$. For example, the observation domain is $\mathbb{Y}=\mathbb{R}$ for continuous data, $\mathbb{Y}=\mathbb{N}$ for count data, and  $\mathbb{Y}=\{0,1\}$ for binary data. 
Note that the canonical link function $f$ is chosen to be $f(\cdot)=b'(\cdot)$. Table~1 summarizes the canonical link functions for common types of distributions.

In our context, we model the the entries in the response tensor $y_{ijk}$ conditional on $\theta_{ijk}$ as independent draws from an exponential family. The quasi log-likelihood of~\eqref{eq:tensormodel2} is equal (ignoring constant) to Bregman distance between $\tY$ and $b'(\Theta)$:
\begin{align}
\tL_{\tY}(\tB)&=\langle \tY, \Theta \rangle - \sum_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K}),\\
\text{where}\ \Theta&=\tB\times\{\mX_1,\ldots,\mX_K\}.
\end{align}
We assume that we have an additional information on an upper bound $\alpha>0$ such that $\mnormSize{}{\Theta}\leq \alpha$. This is the case for many applications we have in mind such as brain network analysis where fiber connections are bounded. We propose a constrained maximum likelihood estimator (MLE) for the coefficient tensor:
\begin{align} \label{eq:MLE} 
	\hat \tB &=\argmax_{\hat \tB \colon \text{rank}(\tB)\leq \mr, \mnormSize{}{\Theta(\tB)}\leq \alpha  } \tL_{\tY}(\tB).
\end{align}
In the following theoretical analysis, we assume the rank $\mr=(r_1,\ldots,r_K)$ is known and fixed. The adaptation of unknown $\mr$ will be addressed in Section~\ref{sec:tuning}.

\subsection{Statistical properties}\label{subsec:statprob}
We assess the estimation accuracy using the deviation in the Frobenius norm. For the true coefficient tensor $\trueB$ and its estimator $\hat \tB$, define
\[
\text{Loss}(\trueB,\ \hat \tB)=\FnormSize{}{\trueB- \hat \tB}^2.
\]

In modern applications, the response tensor and covariates are often large-scale. We are particularly interested in the high-dimensional region in which both $d_k$ and $p_k$ diverge; i.e.\ $d_k\to \infty$ and $p_k\to\infty$, while ${p_k\over d_k} \to \gamma_k \in[0,1)$. As the size of problem grows, and so does the number of unknown parameters. As such, the classical MLE theory does not directly apply. We leverage the recent development in random tensor theory and high-dimensional statistics to establish the error bounds of the estimation.

\begin{assumption}\label{ass}We make the following assumptions:
\begin{enumerate}
\item [A1.] There exist two positive constants $c_1, c_2>0$ such that $c_1\leq \sigma_{\min}(\mX_k)\leq  \sigma_{\max}(\mX_k)\leq c_2$ for all $k\in[K]$. Here $\sigma_{\text{min}}(\cdot)$ and $\sigma_{\text{max}}(\cdot)$ denotes the smallest and largest singular values, respectively.
\item [A2.] There exist two positive constants $L,\ U>0$ such that $L\phi \leq \text{Var}(y_{i_1,\ldots,i_K}|\theta_{i_1,\ldots,i_K})\leq U\phi $ for all $|\theta_{i_1,\ldots,i_K}|\leq \alpha$.
\item[A2'.] Equivalently, there exists two positive constants $L,\ U>0$ such that $L\leq b''(\theta) \leq U$ for all $|\theta|\leq \alpha$, where $\alpha$ is the upper bound of the linear predictor.

\end{enumerate}
\end{assumption}
The assumptions are fairly mild. Assumption A1 guarantees the non-singularity of the covariates, and Assumption A2 ensures the log-likelihood $\tY(\Theta)$ is strictly concave in the linear predictor $\Theta$. Assumption A2 and A2' are equivalent, because $\text{Var}(y_{i_1,\ldots,i_K}|\tX,\tB)=\phi b''(\theta_{i_1,\ldots,i_K})$ when $y_{i_1,\ldots,i_K}$ belongs to an exponential family~\citep{mccullagh1989generalized}.

\begin{thm}[Statistical convergence]\label{thm:main}
Consider a generalized tensor regression model with covariates on multiple modes $\tX=\{\mX_1,\ldots,\mX_K\}$. Suppose the entries in $\tY$ are independent realizations of an exponential family distribution, and $\mathbb{E}(\tY|\tX)$ follows the low-rank tensor regression model~\eqref{eq:tensormodel2}. Under Assumption~\ref{ass}, there exist two constants $C_1, C_2>0$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$, 
\begin{equation}\label{eq:bound}
\text{Loss}(\trueB,\ \hat \tB) \leq \frac{ C_{2}\prod_kr_k}{\max_k r_k}\sum_k p_k.
\end{equation}
Here, $C_2=C_2(\mr,\alpha, K)>0$ is a constant that does not depend on the dimensions $\{d_k\}$ and $\{p_k\}$. 
\end{thm}

Theorem~\ref{thm:main} establishes the statistical convergence for the estimator~\eqref{eq:MLE}. Actually, the proof in section~\ref{sec:appedix} shows that the statistically optimal rate holds, not only for the MLE~\eqref{eq:MLE},  but also any local estimators $\hat \tB$ in the level set $\{\hat \tB \in \tP \colon \tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true}) \}$ satisfies the upper bound~\ref{thm:main}. In Section~\ref{subsec:alg}, we implement empirical studies to verify the algorithmic stability to find a local optimum.

To gain further insight on the bound~\eqref{eq:bound}, we consider a special case when tensor dimensions are equal at each of the modes, i.e., $d_k=d$, $p_k=\gamma d$, $\gamma\in [0,1)$ for all $k\in[K]$, and the covariates $\mX_k$ are Gaussian design matrices with i.i.d.\ $N(0,1)$ entries. To put the context in the framework of Theorem~\ref{thm:main}, we rescale the covariates into $\check \mX_k={1\over \sqrt{d}}\mX_k$ so that the singular values of $\check \mX_k$ are bounded by $1\pm \sqrt{\gamma}$. The result in~\eqref{eq:bound} implies that the estimated coefficient has a convergence rate $\tO({p\over d^K})$ in the scale of the original covariates $\{\mX_k\}$. Therefore, our estimation is consistent as the dimension grows, and the convergence becomes especially favorably as the order of tensor data increases.

As immediate applications, we obtain the convergence rate for the three examples mentioned in Section~\ref{sec:model}. Without loss of generality, we assume that the singular values of the $d_k$-by-$p_k$ covariate matrix $\mX_k$ are bounded by $\sqrt{d_k}$.

\begin{cor}[Spatio-temporal growth model] The estimated type-by-time-by-space coefficient tensor converges at the rate $\tO\left({p+r+s\over dmn}\right)$ where $p\leq d$, $r\leq m$ and $s\leq n$. The estimation achieves consistency as long as the dimension grows in either of the three modes.
\end{cor}

\begin{cor} [Network population model] The estimated node-by-node-by-covariate tensor converges at the rate $\tO\left({2d+p\over d^2n}\right)$ where $p\leq n$. The estimation achieves consistency as the number of individuals or the number of nodes grows. 
\end{cor}

\begin{cor} [Dyadic data with node attributes] The estimated covariate-by-covariate matrix converges at the rate $\tO\left({p\over d^2}\right)$ where $p\leq d$. Again, our estimation achieves consistency as the number of nodes grows. 
\end{cor}

We conclude this section by providing the prediction accuracy, measured in KL divergence, for the response distribution.

\begin{thm}[Prediction error]\label{thm:KL}
Assume the same set-up as in Theorem~\ref{thm:main}. Let $\mathbb{P}_{\tY_{\text{true}}}$ and $\mathbb{P}_{\hat \tY}$ denote the distributions of $\tY$ given the true parameter $\trueB$ and estimated parameter $\hat \tB$, respectively. Then, we have, with probability at least $1-\exp(C_1\sum_k p_k)$,
\[
\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY})\leq \frac{ C_{4}\prod_kr_k}{\max_k r_k} \sum_k p_k,
\]
where $C_4=C_4(\mr,\alpha, K)>0$ is a constant that do not depend on the dimensions $\{d_k\}$ and $\{p_k\}$.
\end{thm}

\section{Numerical implementation}
\subsection{Alternating optimization} \label{subsec:alg}
In this section, we introduce an efficient algorithm to solve~\eqref{eq:MLE}. The objective function $\tL_{\tY}(\tB)$ is concave in $\tB$ when the link $f$ is the canonical link function. We utilize a Tucker factor representation of the coefficient tensor $\tB$ and turn the optimization into a block-wise convex problem.

Specifically, write the rank-$\mr$ decomposition of coefficient tensor $\tB$ as
\begin{equation}\label{eq:tucker}
\tB=\tC\times \{\mM_1,\ldots,\mM_K\},
\end{equation}
where $\tC\in\mathbb{R}^{r_1\times\cdots\times r_K}$ is a full-rank core tensor, $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices with orthogonal columns. Estimating $\tB$ amounts to finding both the core tensor $\tC$ and the factor matrices $\mM_k$'s. The optimization~\eqref{eq:MLE} can be written as $(\hat \tC, \{\hat \mM_k\})=\arg\max \tL_{\tY}(\tC, \mM_1,\ldots,\mM_K)$, where
\begin{align}
\tL_{\tY}(\tC, \mM_1,\ldots,\mM_K )&=\langle \tY, \Theta\rangle -\sum_{i_1,\ldots,i_K}b(\theta_{i_1,\ldots,i_K}),\\
\text{with } \Theta&=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
\end{align}
The decision variables in the above objective function consist of $K+1$ blocks of variables, one for the core tensor $\tC$ and $K$ for the factor matrices $\mM_k$'s. We notice that, if any $K$ out of the $K+1$ blocks of variables are known, then the optimization with respect to the last block of variables reduced to a simple GLM. This observation suggests that we can iteratively update one block at a time while keeping others fixed. After each iteration, we rescale the core tensor $\tC^{(t+1)}$ subject to the maximum norm constraint. This post-processing in principle may not guarantee the monotonic increase of the objective, but we found that in our experiment this simple step appears robust for obtaining a desirable solution. The full algorithm is described in Algorithm~\ref{alg:B}.

\begin{algorithm}[t]
\caption{Generalized tensor response regression with covariates on multiple modes}\label{alg:B}
\begin{algorithmic}[1]
\INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, covariate matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,\ldots,K$, target Tucker rank $\mr=(r_1,\ldots,r_K)$, link function $f$, maximum norm bound $\alpha$
\OUTPUT Low-rank estimation for the coefficient tensor $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$. 
\State Calculate $\check \tB=\tY\times_1 \left[(\mX_1^T\mX_1)^{-1}\mX^T_1\right] \times_2\cdots\times_K\left[(\mX_K^T\mX_K)^{-1}\mX^T_K\right] $.
\State Initialize the iteration index $t=0$. Initialize the core tensor $\tC^{(0)}$ and factor matrices $\mM^{(0)}_k\in\mathbb{R}^{p_k\times r_k}$ via rank-$\mr$ Tucker approximation of $\check\tB$, in the least-square sense. 
  \While {the relative increase in objective function $\tL_\tY(\tB)$ is less than the tolerance}
\State Update iteration index $t \leftarrow t+1$.
\For { $k=1$ to $K$}
\State Obtain the factor matrix $\tilde \mM_k^{(t+1)}\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$. 
\State Perform QR factorization on $\tilde \mM_k^{(t+1)}=\mQ\mR$, where $\mQ\in\mathbb{R}^{p_k\times r_k}$ has orthogonal columns. 
\State Update $\mM_k^{(t+1)}\leftarrow \mQ$ and core tensor $\tC^{(t+1)}\leftarrow \tC^{(t+1)}\times_k \mR$.
\EndFor
\State Obtain the core tensor $\tC^{(t+1)}\in\mathbb{R}^{r_1\times\cdots\times r_K}$ by solving a GLM with $\text{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM^{(t)}_k]$ as covariates, and $f$ as link function. Here $\otimes$ denotes the kronecker product of matrices. 
\State Rescale the core tensor subject to the maximum norm constraint. 
\State Update $\tB^{(t+1)}\leftarrow \tC^{(t+1)}\times_1\mM_1^{(t+1)}\times_2\cdots\times_K\mM_K^{(t+1)}$.
\EndWhile
\end{algorithmic}
\end{algorithm}

Note that the feasible set $\tP$ is a non-convex set. Therefore, the optimization~\eqref{eq:MLE} is a non-convex problem, and the Algorithm~\ref{alg:B} usually not theoretically possesses the global optimality. However, as mentioned in Section~\ref{subsec:statprob}, the desired convergence rate~\eqref{eq:bound} holds for the valid estimators satisfying $\tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true})$, which indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective of local optimums are large enough. Fortunately, we find the  Algorithm~\ref{alg:B}  often gives a satisfactory convergence point $\hat \tB$, when we initialize the parameters via continuous-valued Tucker decomposition. Figure~\ref{fig:loglike} shows the trajectory of the objective function for order-3 tensors under the balanced setting, where $\alpha = 10, d_k = d, p_k = 0.4d, r_k = r$ for $ k = 1,2,3, p \in \{25,30\}$  and $r \in \{3, 6\}$. We consider the inputs with Gaussian, Bernoulli, and Poisson entries. Under all combinations of the dimension $d$, rank $r$, and type of the entries, Algorithm~\ref{alg:B} converges quickly in a few iterations, and the objective value at convergent points are close to or larger than the value at true parameters.

\begin{figure}[t]
\centering
\includegraphics[width=15cm]{loglike}
\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c)Poisson models. The dashed line represents the objective value at true parameter $\tL_{\tY}(\tB_{true})$. A micro-iteration refers to the update for one of the $K$ blocks. Four micro-iterations consist of a complete ``while'' iteration in Algorithm~\ref{alg:B}.}\label{fig:loglike}
\end{figure}

\subsection{Rank selection}\label{sec:tuning}
Algorithm~\ref{alg:B} takes the rank $\mr$ as an input. Estimating an appropriate rank given the data is of practical importance. We propose to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
\begin{align}\label{eq:BIC}
\hat \mr&=\argmin_{\mr=(r_1,\ldots,r_K)} \text{BIC}(\mr)\\
&=\argmin_{\mr=(r_1,\ldots,r_K)}\left[-2\tL_{\tY}(\hat \tB)+p_e(\mr)\log \left(\prod_k d_k\right) \right],
\end{align}
where $p_e(\mr)\stackrel{\text{def}}{=}\sum_k (p_k-r_k-1)r_k+\prod_k r_k$ is the effective number of parameters in the model. We choose $\hat \mr$ that minimizes $\text{BIC}(\mr)$ via grid search. Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model. We test its empirical performance in Section~\ref{sec:simulation}.

\subsection{Time complexity}
The computational complexity of our tensor regression model is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the response tensor. More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor. The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$-covariate matrix. Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$.

\section{Simulation}\label{sec:simulation}
We evaluate the empirical performance of our generalized tensor regression through simulations. We consider order-3 tensors with a range of distribution types. The coefficient tensor $\tB$ is generated using the factorization form~\eqref{eq:tucker} where both the core and factor matrices are drawn i.i.d.\ from Uniform[-1,1]. The linear predictor is then simulated from $\tU=\tB\times\{\mX_1,\mX_2,\mX_3\}$, where $\mX_k$ is either an identity matrix (i.e.\ no covariate available) or Gaussian random matrix with i.i.d.\ entries from $N(0,\sigma_k^2)$. We set $\sigma_k=d_k^{-1/2}$ to ensure the singular values of $\mX_k$ are bounded as $d_k$ increases. The $\tU$ is scaled such that $\mnormSize{}{\tU}=1$. Conditional on the linear predictor $\tU=\entry{u_{ijk}}$, the entries in the tensor $\tY=\entry{y_{ijk}}$ are drawn independently according to one of the following three probabilistic models:

\begin{enumerate}
\item[(a)] (Gaussian). Continuous data $y_{ijk}\sim N\left(\alpha u_{ijk}, 1\right)$.
\item[(b)] (Poisson). Count data $y_{ijk}\sim\text{Poi}\left( e^{\alpha u_{ijk}}\right)$.
\item[(c)] (Bernoulli). Binary data $y_{ijk}\sim \text{Ber}\left( {e^{\alpha u_{ijk}} \over 1+e^{\alpha u_{ijk}}}\right)$.
\end{enumerate}
Here $\alpha>0$ is a scalar controlling the magnitude of the effect size. In each simulation study, we report the mean squared error (MSE) for the coefficient tensor averaged across $n_{\text{sim}}=30$ replications.

\subsection{Finite-sample Performance}

The first experiment assesses the selection accuracy of our BIC criterion~\eqref{eq:BIC}. We consider the balanced situation where $d_k=d$, $p_k=0.4d_k$ for $k=1,2,3$. We set $\alpha=10$ and consider various combinations of dimension $d$ and rank $\mr=(r_1,r_2,r_3)$. For each combination, we simulate tensor data following Gaussian, Bernoulli, and Poisson models. We then minimize BIC using a grid search over three dimensions. The hyper-parameter $\alpha$ is set to infinity in the fitting, which essentially imposes no prior on the coefficient magnitude. Table~\ref{tab:rank} reports the selected rank averaged over $n_{\text{sim}}=30$ replicates for Gaussian and Poisson models. We found that when $d=20$, the selected rank is slightly smaller than the true rank, and the accuracy improves immediately when the dimension increases to $d=40$. This agrees with our expectation, as in tensor regression, the sample size is related to the number of entries. A larger $d$ implies a larger sample size, so the BIC selection becomes more accurate.

\begin{table}[tb]
\centering
%\resizebox{1\textwidth}{!}{
\begin{tabular}{|c|cc|cc|}
\hline
True Rank &\multicolumn{2}{c|}{Dimension (Gaussian tensors)}&\multicolumn{2}{c|}{Dimension (Poisson tensors)}\\
$\mr$& $d=20$  &$d=40$&$d=20$ &$d=40$\\
\hline
$(3,\ 3,\ 3)$&$(2.1,\ 2.0,\ 2.0)$&$({\bf 3},\ {\bf 3},\ {\bf 3})$&$(2.0,\ 2.2,\ 2.1)$&$({\bf 3},\ {\bf 3},\ {\bf 3})$\\
$(4,\ 4,\ 6)$&$(3.2,\ 3.1,\ 5.0)$&$({\bf 4},\ {\bf 4},\ {\bf 6})$&$({\bf 4.0},\ {\bf 4.0},\ 5.2)$&$({\bf 4},\ {\bf 4},\ {\bf 6})$\\
$(6,\ 8,\ 8)$&$(5.1,\ 7.0,\ 6.9)$&$({\bf 6},\ {\bf 8},\ {\bf 8})$&$(5.0,\ 6.1,\ 7.1)$&$({\bf 6},\ {\bf 8},\ {\bf 8})$\\
\hline
\end{tabular}
%}
\caption{Rank selection via BIC. Bold number indicates no significant difference between the estimate and the ground truth, based on a $z$-test with a level 0.05.}\label{tab:rank}
\end{table}

The second experiment evaluates the accuracy when covariates are available on all modes. We set $\alpha=10, d_k=d, p_k=0.4d_k, r_k=r\in\{2,4,6\}$ and increase $d$ from 25 to 50. Our theoretical analysis suggests that $\hat \tB$ has a convergence rate $\tO(d^{-2})$ in this setting. Figure~1 plots the estimation error versus the ``effective sample size'', $d^2$, under three different distribution models. We found that the empirical MSE decreases roughly at the rate of $1/d^2$, which is consistent with our theoretical ascertainment. We also observed that, tensors with higher ranks tend to yield higher estimation errors, as reflected by the upward shift of the curves as $r$ increases. Indeed, a larger $r$ implies a higher model complexity and thus greater difficulty in the estimation. Similar behaviors can be observed in the non-Gaussian data in Figure~\ref{fig:dim}b-c.

\begin{figure}[t]
\centering
\includegraphics[width=15cm]{dimension.pdf}
\caption{Estimation error against effective sample size. The three panels depict the MSE when the response tensors are generated form (a) Gaussian (b) Poisson and (c) Bernoulli models. The dashed curves correspond to $\tO({1/d^2})$.}\label{fig:dim}
\end{figure}

The third experiment investigates our model's ability in handling correlation among coefficients. We mimic the scenario of brain imaging analysis. A sample of $d_3=50$ networks are simulated, one for each individual. Each network measures the connections between $d_1=d_2=20$ brain nodes. We simulate $p=5$ covariates for the each of the 50 individuals. These covariates may represent, for example, age, gender, cognitive score, etc. Recent study~\citep{robinson2015dynamic} has suggested that brain connectivity networks often exhibit community structure represented as a collection of subnetworks, and each subnetwork is comprised of a set of spatially distributed brain nodes. To accommodate this structure, we utilize the stochastic block model~\citep{abbe2017community} to generate the effect size. Specifically, we partition the nodes into $r$ blocks by assigning each node to a block with uniform probability. Edges within a same block are assumed to share the same covariate effects, where the effects are drawn i.i.d.\ from $N(0,1)$. We then apply our tensor regression model to the network data using the BIC-selected rank. Note that in this case, the true model rank is unknown; the rank of a $r$-block matrix is not necessarily equal to~$r$~\citep{zeng2019multiway}.

\begin{figure}[tb]
\centering
\includegraphics[width=15cm]{comparison.pdf}
\caption{Performance comparison when the networks have block structure. The three panels depict the MSE when the response tensors are generated form (a) Gaussian (b) Poisson and (c) Bernoulli models. The $x$-axis represents the number of blocks in the networks. }\label{fig:glm}
\end{figure}

Figure~\ref{fig:glm} compares the MSE of our method with a classical GLM approach. A classical GLM is to regress the dyadic edges, one at a time, on the covariates, and this model is repeatedly fitted for each edge. This repeated approach, however, does not account for the correlation among the edges, and may suffer from overfitting. As we can see in Figure~\ref{fig:glm}, out tensor regression method achieves significant error reduction in all three models considered. The outer-performance is significant in the presence of large communities, and even in the less structured case ($\sim 20/15=1.33$ nodes per block), our method still outer-performs GLM. This is because the low-rankness in our modeling automatically identifies the shared information across entries. By selecting the rank in a data-driven way, our method is able to achieve accurate estimation with improved interpretability.

\subsection{Comparison with alternative methods}

We compare our generalized tensor regression ({\bf GTR}) with three other supervised tensor methods:
\begin{itemize}
\item  Higher-order low-rank regression ({\bf HOLRR},~\citep{rabusseau2016low}) is a least-square based tensor regression that allows covariates on a single mode. 
\item Higher-order partial least square ({\bf HOPLS},~\citep{zhao2012higher}) is a dimension-reduction method that jointly models a tensor response and a tensor covariate. 
\item Subsampled tensor projected gradient ({\bf TPG},~\citep{yu2016learning}) tackles the same question as {\bf HOLRR} but instead uses a different algorithm to solve the problem. 
\end{itemize}
These three methods are the closest algorithms to ours, in that they relate a tensor response to covariates using a low-rank structure. All the three methods allow only Gaussian data, whereas ours is applicable to any exponential family distribution including Gaussian, Bernoulli, Multinomial, etc. For fair comparison, we consider only Gaussian response in the simulation. We measure the accuracy using mean squared prediction error, $\text{MSPE}=\sqrt{\sum_kd_k}\FnormSize{}{\hat \tY-\mathbb{E}(\tY|\tX)}$, where $\hat \tY$ is the fitted value from each of the methods.

The comparison was assessed from three aspects: (a) benefit of incorporating covariates from multiple modes; (b) prediction error with respect to sample size; (c) sensitivity of accuracy with respect to model complexity. We use similar simulation setups as in our experiment II, but consider combinations of rank ($\mr=(3,3,3)$ vs.\ $(4,5,6)$), noise ($\sigma = 1/2$ vs. $1/4$), and dimension ($d$ ranging from 20 to 100 for modes with covariates, $d = 20$ for modes without covariates).

Figure~\ref{fig:compare} shows the averaged prediction error across 30 replicates. We see that our {\bf GTR} outperforms others, especially in the high-rank high-noise setting. As the number of informative modes (i.e. modes with available covariates) increases, the {\bf GTR} exhibits a reduction in error whereas others have increased errors. This showcases the benefit toward prediction via incorporation of multiple covariates. Note that our method {\bf GTR} is most comparable to {\bf HOLRR} when there is only a single informative mode. In such a case, both methods share a same cost function but have different algorithms. {\bf GTR} alternates between informative and non-informative modes, whereas {\bf HOLRR} approximates the non-informative modes via unfolded response alone. The accuracy gain in Figure~\ref{fig:compare} demonstrates the benefit of alternating algorithm -- having informative modes also improves the estimation along non-informative modes.

\begin{figure}[ht]
\centering
\includegraphics[width=12cm]{merge.pdf}

\caption{Comparison of MSPE versus the number of modes with covariates. We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and noise $\sigma=1/2$ (high), $\sigma=1/4$ (low).}~\label{fig:compare}

\end{figure}

Figure~\ref{fig:compare2} compares the prediction error with respect to sample size. The sample size is the total number of entries in the tensor. In the low-rank setting, our method has similar performance as {\bf HOLRR}, and the improvement becomes more pronounced when the rank increases. Neither {\bf HOPLS} nor {\bf TPG} has satisfactory performance in high-rank or high-noise settings. One possible reason is that a higher rank implies a higher inter-mode complexity, and our {\bf GTR} method lends itself well to this context.

\begin{figure}[ht]
\centering
\includegraphics[width=12cm]{merge2.pdf}

\caption{Comparison of MSPE versus sample size. We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and noise $\sigma=1/2$ (high), $\sigma=1/4$ (low). }~\label{fig:compare2}

\end{figure}

\section{Data analysis}
We apply our tensor regression model to two real datasets. The first application concerns the brain network modeling in response to individual attributes (i.e.\ covariate on one mode), and the second application focuses on multi-relational network analysis with dyadic attributes (i.e.\ covariates on two modes). 
\subsection{Human Connectome Project (HCP)}
The Human connectome project (HCP~\citep{HCP}) aims to build a ``network map'' that characterizes the anatomical and functional connectivity within healthy human brains. We take a subset of HCP data that consists of 136 brain structural networks, one for each individual. Each brain network is represented as a 68-by-68 binary matrix, where the entries encode the presence or absence of fiber connections between the 68 brain regions. We consider four individual-covariates: gender (65 females vs.\ 71 males), age 22-25 ($n=35$), age 26-30 ($n=58$), and age 31+ ($n=43$). The goal is to identify the connection edges that are affected by the individual covariates. A key challenge in brain network is that the edges are correlated; for example, two edges may stem out from a same brain region, and it is of importance to take into account the within-dyad dependence.

\begin{figure}[hbt]
\centering
\includegraphics[width=10cm]{HCP.pdf}
\caption{Top edges with large effects. Red edges represent relatively strong connections and blue edges represent relatively weak connections. (a) Global effect; (b) Female effect; (c) Age 22-25; (d) Age 31+.}\label{fig:brain}
\end{figure}

We fit the tensor regression model to the HCP data. The response is a binary tensor $\tY\in\{0,1\}^{68\times 68\times 136}$ and the covariates are of dimension 4 along the 3$^{\text{rd}}$ mode. The BIC selection suggests a rank $\mr=(10,10,4)$ with log-likelihood $\tL_{\tY}=-174654.7$. Figure~\ref{fig:brain} shows the top edges with high effect size, overlaid on the Desikan atlas brain template~\citep{desikan2006automated,xia2013brainnet}. We utilize the sum-to-zero contrasts in the effects coding and depicted only the top 3\% edges whose connections are non-constant across the sample. It is observed that the global connection exhibits clear spatial separation, and that the nodes within each hemisphere are more densely connected with each other (Figure~\ref{fig:brain}a). In particular, the superior-temproal (\emph{SupT}), middle-temporal (\emph{MT}) and Insula are the top three popular nodes in the network. Interestingly, female brains display higher inter-hemispheric connectivity, especially in the frontal, parental and temporal lobes (Figure~\ref{fig:brain}b). This is in agreement with a recent study showing that female brains are optimized for inter-hemispheric communication~\citep{ingalhalikar2014sex}. We also found several edges with declined connection in the group Age 31+. Notably, those edges involve Frontal-pole (\emph{Fploe}), superior-frontal (\emph{SupF}) and Cuneus nodes. The Frontal-pole region has long been known for its importance in memory and cognition, and the detected decline with age further highlights its biological importance.

\subsection{Nations data}

The second application concerns the multi-relational network analysis with node-level attributes. We consider \emph{Nations} dataset~\citep{nickel2011three} which records 56 relations among 14 countries between 1950 and 1965. The multi-relational networks can be organized into a $14 \times 14 \times 56$ binary tensor, with each entry indicating the presence or absence of a connection, such as ``sending tourist to'', ``export'', ``import'', between countries. The 56 relations span the fields of politics, economics, military, religion, and so on. In addition, country-level attributes are also available, and we focus on the following six covariates: \emph{constitutional, catholics, lawngos, politicalleadership, geographyx}, and \emph{medicinengo}. The goal is to identify the variation in connections due to country-level attributes and interactions thereof. One of the key features is that the 56 relations are correlated, and we would like to take that into account in assessing the covariate effects.

We applied our tensor regression model to the \emph{Nations} data. The multi-relational network $\tY\in\{0,1\}^{14\times 14\times 56}$ was treated as the response tensor, and the country attributes $\mM\in\mathbb{R}^{14\times 6}$ were treated as covariates on both the 1st and 2nd modes. The BIC criterion suggests a rank $\mr=(4,4,4)$ for the coefficient tensor $\tB\in\mathbb{R}^{6\times 6 \times 56}$. Table~\sref{fig:s1} shows the $K$-mean clustering of the 56 relations based on the 3$^\text{rd}$ mode factor $\mM_3\in\mathbb{R}^{56\times 4}$. We found that the relations reflecting the similar aspects of international affairs are grouped together. In particular, Cluster I consists of political relations such as \emph{officialvisits, intergovorgs}, and \emph{militaryactions}; Clusters II and III capture the economical relations such as \emph{economicaid, booktranslations, tourism}; and Cluster IV represents the Cold War alliance blocs. The similarity among entities in each cluster suggests the plausibility of our dimension reduction.

\begin{figure}[t]
\centering
\includegraphics[width=10cm]{coef.pdf}
\caption{Effect estimation in the \emph{Nations} data. Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warnning}, \emph{economicaid},\emph{intergovorg}, and \emph{commonblock}, respectively. }\label{fig:est}

\end{figure}

To investigate the effects of dyadic attributes towards connections, we depicted the estimated coefficients $\hat \tB=\entry{\hat b_{ijk}}$ for several relation types (Figure~\ref{fig:est}). Note that entries $\hat b_{ijk}$ can be interpreted as the contribution, at the logit scale, of covariate pair $(i,j)$ ($i$th covariate for the ``sender'' country and $j$th covariate for the ``receiver'' country) towards the connection of relation $k$.  Several interesting findings emerge from the observation. We found that relations belonging to a same cluster tend to have similar covariate effects. For example, the relations \emph{warnings} and \emph{ecnomicaid} are classified into Cluster II, and both exhibit similar covariate pattern (Figure~\ref{fig:est}a-b). Moreover, the majority of the diagonal entries $\hat \tB(i,i,k)$ positively contribute to the connection. This suggests that countries with coherent attributes tend to interact more often than others. We also found that the \emph{constitutional} attribute is an important predictor for the \emph{commonbloc} relation, whereas the effect is weaker for other relations (Figure~\ref{fig:est}d). This is not surprising, as the block partition during Cold War is associated with the \emph{constitutional} attribute.

\section{Conclusion}

 We have developed a generalized tensor regression with covariates on multiple modes. A fundamental feature of tensor-valued data is the statistical interdependence among entries. Our proposed rank-constrained estimation achieves high accuracy with sound theoretical guarantees. The estimation accuracy is quantified via deviation in the Frobenius norm and K-L divergence. Other measures of accuracy may also be desirable, such as the spectral norm or the maximum norm of the deviation. Exploiting the properties and benefits of different error quantification warrants future research.

 \newpage

\bigskip
\begin{center}
{\large\bf SUPPLEMENTARY MATERIAL}
\end{center}

\appendix

\section{Proofs}\label{sec:appedix}
\begin{proof}[Proof of Theorem~\ref{thm:main}]
Define $\ell(\tB)=\mathbb{E}(\tL_{\tY}(\tB))$, where the expectation is taken with respect to $\tY\sim \trueB$ under the model with true parameter $\trueB$. We first prove the following two conclusions:
\begin{enumerate}
\item[C1.] There exists two positive constants $C_1$, $C_2>0$, such that, with probability at least $1-\exp(-C_1\log K\sum_k p_k)$, the stochastic deviation, $\tL_{\tY}(\tB)-\ell(\tB)$, satisfies
\[
|\tL_{\tY}(\tB)-\ell(\tB)|=|\langle \tE,\ \tB\times_1\mX_1\times_2\cdots\times_K \mX_K\rangle| \leq C_2\FnormSize{}{\tB} \sqrt{{\prod_k r_k \over \max_k r_k} \sum_k p_k}.
\]
\item[C2.] The inequality $\ell(\hat \tB) - \ell(\trueB) \leq  -{L\over 2}\FnormSize{}{\hat \Theta-\trueT}^2$ holds, where $L>0$ is the lower bound for $\min_{|\theta|\leq \alpha}|b''(\theta)|$. 
\end{enumerate}

To prove C1, we note that the stochastic deviation can be written as:
\begin{align}\label{bound}
\tL_{\tY}(\tB)-\ell(\tB)&=\langle \tY-\mathbb{E}(\tY|\tX),\ \Theta(\tB)\rangle\notag \\
&= \langle \tY- b'(\trueT),\ \Theta\rangle \notag \\
&= \langle \tE\times_1\mX^T_1\times_2\cdots\times_K\mX^T_K,\ \tB\rangle,
\end{align}
where $\tE\stackrel{\text{def}}{=}\tY-b'(\trueT)$, and the second line uses the property of exponential family that $\mathbb{E}(\tY|\tX)=b'(\trueT)$. Based on Proposition~\ref{prop}, the boundedness of $b''(\cdot)$ implies that $\tE$ is a sub-Gaussian-$(\phi U)$ tensor. Let $\check\tE\stackrel{\text{def}}{=}\tE\times_1\mX^T_1\times_2\cdots\times_K\mX^T_K$. By Proposition~\ref{prop:sub}, $\check \tE$ is a $(p_1,\ldots,p_K)$-dimensional sub-Gaussian tensor with parameter bounded by $C=\phi Uc^{K}_2$. Here $c_2>0$ is the upper bound of $\sigma_{\max}(\mX_k)$. Applying Cauchy-Schwarz inequality to~\eqref{bound} yields
\begin{equation}\label{eq:bound2}
|\tL_{\tY}(\tB)-\ell(\tB)|\leq \norm{\check \tE} \nnorm{\tB},
\end{equation}
where $\norm{\cdot}$ denotes the tensor spectral norm and $\nnorm{\cdot}$ denotes the tensor nuclear norm. The nuclear norm $\nnorm{\tB}$ is  bounded by $\nnorm{\tB}\leq \sqrt{{\prod_k r_k \over \max_k r_k}}\FnormSize{}{\tB}$ (c.f.~\citep{wang2018learning,wang2017operator}). The spectral norm $\norm{\check \tE}$ is bounded by $\norm{\check \tE}\leq C_2 \sqrt{\sum_k p_k}$ with probability at least $1-\exp(-C_1\log K \sum_kp_k)$ (c.f.~\citep{wang2018learning,tomioka2014spectral}). Combining these two bounds with~\eqref{eq:bound2}, we have, with probability at least $1-\exp(-C_1\log K \sum_kp_k)$, 
\[
|\tL_{\tY}(\tB)-\ell(\tB)|\leq C_2\FnormSize{}{\tB}  \sqrt{ {\prod_k r_k\over \max_k r_k}\sum_k p_k},
\]
where $C_2>0$ is a constant absorbing all factors that do not depend on $\{p_k\}$ and $\{r_k\}$.

Next we prove C2. Applying Taylor expansion to $\tL_{\tY}(\tB)$ around $\trueB$ yields
\begin{equation}\label{eq:logbefore}
\tL_{\tY}(\tB)=\tL_{\tY}(\trueB)+\Big\langle {\partial\tL_{\tY}(\tB) \over \tB}\big|_{\tB=\trueB},\tB-\trueB \Big\rangle + {1\over 2}\text{vec}(\tB-\trueB)^T \tH(\check \tB)\text{vec}(\tB-\trueB),
\end{equation}
where $\tH_{\tY}(\check \tB)$ is the (non-random) Hession of ${\partial \tL_{\tY}^2 (\tB)\over\partial^2 \tB}$ evaluated at $\check \tB =\alpha \text{vec}(\alpha \tB+(1-\alpha)\trueB)$ for some $\alpha\in[0,1]$. Note that we have $\mathbb{E}\left({\partial\tL_{\tY}(\tB) \over \partial \tB}\big|_{\tB=\trueB}\right)=0$. We take expectation with respect to $\tY\sim \trueB$ on both sides of~\eqref{eq:log} and obtain
\begin{equation}\label{eq:log}
\ell(\tB)=\ell(\trueB)+{1\over 2}\text{vec}(\tB-\trueB)^T \tH(\check \tB)\text{vec}(\tB-\trueB).
\end{equation}
By the fact ${\partial \tL_{\tY}^2(\Theta)\over \partial^2 \Theta}=-b''(\Theta)$ and chain rule over $\Theta=\Theta(\tB)=\tB\times_1\mX_1\cdots\times_K\mX_K$, the equation~\eqref{eq:log} implies that 
\begin{equation}\label{eq:logB}
\ell(\tB)-\ell(\trueB)=-{1\over 2}\sum_{i_1,\ldots,i_K}b''(\check \theta_{i_1,\ldots,i_K}) (\theta_{i_1,\ldots,i_K}-\theta_{\text{true},i_1,\ldots,i_K})^2 \leq -{L \over 2}\FnormSize{}{\Theta-\trueT}^2,
\end{equation}
holds for all $\tB\in\tP$, provided that $\min_{|\theta|\leq \alpha}|b''(\theta)|\geq L>0$. In particular, the inequality~\eqref{eq:log} also applies to the constrained MLE $\hat \tB$. So we have
\begin{equation}\label{upper}
\ell(\hat \tB)-\ell(\trueB)\leq -{L \over 2}\FnormSize{}{\hat \Theta-\trueT}^2.
\end{equation}
Now we have proved both C1 and C2. Note that $\tL_{\tY}(\hat \tB)- \tL_{\tY}(\trueB)\geq 0$ by the definition of $\hat \tB$, This implies that
\begin{align}
0&\leq \tL_{\tY}(\hat \tB)- \tL_{\tY}(\trueB) \\
&\leq \left(\tL_{\tY}(\hat \tB)-\ell(\hat \tB)\right)-\left( \tL_{\tY}(\trueB)-\ell(\trueB)\right)+\left(\ell(\hat \tB)-\ell(\trueB)\right)\\
&\leq \langle\tE,\ \Theta-\trueT    \rangle -{L\over 2}\FnormSize{}{\hat \Theta-\trueT}^2,
\end{align}
where the second line follows from~\eqref{upper}. Therefore,
\begin{align}\label{eq:1}
\FnormSize{}{\hat \Theta-\trueT}&\leq {2\over L}\big\langle \tE,\ {\hat \Theta -\trueT \over \FnormSize{}{\hat \Theta-\trueT}} \big\rangle\notag \\
&\leq {2\over L}\sup_{\Theta: \FnormSize{}{\Theta}=1, \Theta=\tB\times_1\mX_1\times_2\cdots\times_K \mX_K}\langle \tE,\ \Theta \rangle\notag \\
&\leq {2\over L}\sup_{\tB\in\tP: \FnormSize{}{\tB}\leq \prod_k \sigma^{-1}_{\min}(\mX_k)} \langle \tE,\ \ \tB\times_1\mX_1\times_2\cdots\times_K\mX_K\rangle.
\end{align}
Combining~\eqref{eq:1} with C1 yields 
\[
\FnormSize{}{\hat \Theta-\trueT} \leq {2C_2\over L} \prod_k \sigma^{-1}_{\min}(\mX_k)\sqrt{{\prod_k r_k \over \max r_k}\sum_k p_k}.
\]
Therefore, the final conclusion follows by noting that 
\[
\FnormSize{}{\hat \tB-\trueB}\leq \FnormSize{}{\hat \Theta-\trueT}\prod_{k}\sigma^{-1}_{\min}(\mX_k)\leq C\sqrt{\sum_k p_k},
\]
where $C=C(\mr, \alpha,K,c_1, c_2)>0$ is a constant that does not depend on the dimensions $\{d_k\}$ and $\{p_k\}$. 
\end{proof}

\begin{prop}[sub-Gaussian tensors]\label{prop:sub}
Let $\tS$ be a sub-Gaussian-$(\sigma)$ tensor of dimension $(d_1,\ldots,d_K)$, and $\mX_k\in\mathbb{R}^{p_k\times d_k}$ be non-random matrices for all $k\in[K]$. Then $\tE=\tS \times_1  \mX_1\times_2\cdots\times_K  \mX_K$ is a sub-Gaussian-$(\sigma')$ tensor of dimension $(p_1,\ldots,p_K)$, where $\sigma'\leq \sigma \prod_k\sigma_{\max}(\mX_k)$. Here $\sigma_{\max}(\cdot)$ denotes the largest singular value of the matrix. 
\end{prop}

\begin{proof}
To show $\tE$ is a sub-Guassian tensor, it suffices to show that the $\tE\times_1 \bmu^T_1\times_2\cdots\times_K \bmu^T_K$ is a sub-Gaussian scalar with parameter $\sigma'$, for any unit-1 vector $\bmu_k\in\mathbb{R}^{p_k}$, $k\in[K]$.

Note that, 
\begin{align}
\tE\times_1 \bmu^T_1\times_2\cdots\times_K \bmu^T_K&=\tS\times_1 (\bmu^T_1\mX_1) \times_2\cdots\times_K (\bmu^T_K\mX_K)\\
&= \left(\prod_k \vnormSize{}{\bmu^T_k\mX_k}\right) \KeepStyleUnderBrace{\left[\tS\times_1 { (\bmu^T_1\mX_1)\over \vnormSize{}{(\bmu^T_1\mX_1)}}\times_2\cdots\times_K { (\bmu^T_K\mX_K)\over \vnormSize{}{(\bmu^T_K\mX_K)}}\right]}_{\text{sub-Gaussian-$\sigma$ scalar}}.
\end{align}
Because $\vnormSize{}{(\bmu^T_k\mX_k)}\leq \sigma_{\max}(\mX^T_k)\vnormSize{}{\bmu_k}=\sigma_{\max}(\mX_k)$, we conclude that $\tE\times_1 \bmu^T_1\times_2\cdots\times_K \bmu^T_K$ is a sub-Gaussian tensor with parameter $\sigma \prod_k \sigma_{\max}(\mX_k)$. 
\end{proof}

\begin{prop}[sub-Gaussian residuals]\label{prop}
Define the residual tensor $\tE=\entry{\varepsilon_{i_1,\ldots,i_K}}=\tY-b'(\Theta)\in\mathbb{R}^{d_1\times \cdots \times d_K}$. Under the Assumption A2, $\varepsilon_{i_1,\ldots,i_K}$ is a sub-Gaussian random variable with sub-Gaussian parameter bounded by $\phi U$, for all $(i_1,\ldots,i_K)\in[d_1]\times\cdots\times[d_K]$.
\end{prop}
\begin{proof} The proof is similar to Lemma 3 in~\citep{fan2019generalized}. For ease of presentation, we drop the subscript $(i_1,\ldots,i_K)$ and simply write $\varepsilon$ ($=y-b'(\theta)$). For any given $t\in\mathbb{R}$, we have
\begin{align}
\mathbb{E}(\exp(t\varepsilon|\theta)&=\int c(x) \exp\left({\theta x - b(\theta)\over \phi}   \right)\exp \left(t(x-b'(\theta))\right)dx\\
&=\int c(x)\exp \left( {(\theta + \phi t)x - b (\theta+\phi t)+b(\theta+\phi t)-b(\theta)-\phi t b'(\theta) \over \phi}\right)dx\\
&=\exp\left( {b(\theta+\phi t)-b(\theta)-\phi t b'(\theta) \over \phi} \right)\\
&\leq \exp\left(\phi U t^2\over 2 \right),
\end{align}
where $c(\cdot)$ and $b(\cdot)$ are known functions in the exponential family corresponding to $y$. 
Therefore, $\varepsilon$ is sub-Gaussian-$(\phi U)$. 
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:KL}]
The proof is similar to~\citep{baldin2018optimal}. We sketch the main steps here for completeness. Recall that $\ell(\tB)=\mathbb{E}(\tL_{\tY}(\tB))$. By the definition of KL divergence, we have that,
\begin{align}
\ell(\hat \tB)&=\ell(\trueB)-\sum_{(i_1,\ldots,i_K)} KL(\theta_{\text{true}, i_1,\ldots, i_K}, \hat \theta_{i_1,\ldots,i_K})\\
&=\ell(\trueB)-\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY}),
\end{align}
where $\mathbb{P}_{\tY_{\text{true}}}$ denotes the distribution of $\tY|\tX$ with true parameter $\trueB$, and $\mathbb{P}_{\hat \tY}$ denotes the distribution with estimated parameter $\hat \tB$. Therefore
\begin{align}
\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY}) &= \ell(\trueB)-\ell(\hat \tB)\\
&={1\over 2}\sum_{i_1\ldots,i_K}b''(\check \theta_{i_1,\ldots,i_K})(\theta_{i_1,\ldots,i_K}-\theta_{\text{true},i_1,\ldots,i_K})^2\\
&\leq {U\over 2} \FnormSize{}{\Theta-\trueT}^2\\
&\leq {U \over 2}c^{2K}_2\FnormSize{}{\tB-\trueB}^2,
\end{align}
where the second line comes from~\eqref{eq:log}, and $c_2>0$ is the upper bound for the $\sigma_{\max}(\mX_k)$. The result then follows from Theorem~\ref{thm:main}.
\end{proof}

\section{Additional results for real data analysis}

\subsection{HCP data analysis}

Figure~\ref{fig:s1} compares the estimated coefficients from our method (tensor regression) with those from classical GLM approach. A classical GLM is to regress the brain edges, one at a time, on the individual-level covariates, and this logistic model is repeatedly fitted for every edge $\in [68]\times [68]$. As we can see in the figure, our tensor regression shrinkages the coefficients towards center, thereby enforcing the sharing between coefficient entries.

\begin{figure}[t]
\includegraphics[width=17cm]{compare.pdf}
\caption{Comparison of coefficient estimation in the HCP data.}\label{fig:s1}
\end{figure}

\subsection{Nations data analysis}
Table~\ref{tab:s1} summarizes the $K$-means clustering of the 56 relations based on the $3^{\text{rd}}$ mode factor $\mM_3\in\mathbb{R}^{56\times 4}$ in the tensor regression model.

\begin{table}[H]
\begin{tabular}{c|l}
\hline
\multirow{4}{*}{Cluster I}  &officialvisits, intergovorgs,  militaryactions, violentactions, duration, \\
&negativebehavior, boycottembargo, aidenemy, negativecomm, accusation, \\
& protestsunoffialacts,    nonviolentbehavior, emigrants, relexports,   \\
& timesincewar, commonbloc2, rintergovorgs3, relintergovorgs\\
\hline
\multirow{4}{*}{Cluster II}& economicaid, booktranslations, tourism, relbooktranslations, releconomicaid, \\
& conferences, severdiplomatic, expeldiplomats, attackembassy, unweightedunvote,\\
& reltourism, tourism3, relemigrants, emigrants3, students, relstudents, \\
&exports, exports3, lostterritory, dependent, militaryalliance, warning  \\
\hline
\multirow{2}{*}{Cluster III} & treaties, reltreaties, exportbooks, relexportbooks, weightedunvote, ngo, \\
&relngo, ngoorgs3, embassy, reldiplomacy, timesinceally, independence, commonbloc1\\
\hline
Cluster IV & commonbloc0, blockpositionindex\\
\end{tabular}
\caption{$K$-means clustering of relations based on factor matrix in the coefficient tensor.}\label{tab:s1}
\end{table}

\section*{Acknowledgements}
This research was supported by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation. The authors would like to thank Chen Zhang from Wisconsin State Lab of Hygiene for help with figure 4.

\bibliography{tensor_wang}
\bibliographystyle{apalike}
\end{document}

### 2020-08-27 tensor_regress/jasa-latexsource/tensor-jasa.tex
jasa

@@
 
-Many contemporary scientific and engineering studies collect multi-way array data, a.k.a.\ tensors, accompanied by additional covariates.
-One example is neuroimaging analysis~\cite{sun2017store,zhou2013tensor}, in which the brain connectivity networks are collected from a sample of individuals.
-Researchers are often interested in identifying connection edges that are affected by individual characteristics such as age, gender, and disease status (see Figure~\ref{fig:intro1}a).
-Another example is in the field of network analysis~\cite{baldin2018optimal,hoff2005bilinear}.
-A typical social network consists of nodes that represent people and edges that represent friendships.
-In addition, features on nodes and edges are often available, such as people's personality and demographic location.
-It is of keen scientific interest to identify the variation in the connection patterns (e.g., transitivity, community) that can be attributable to the node features.
+Nowadays, multi-dimensional arrays, a.k.a.\ tensors, are collected with additional covariates on multiple modes in many modern scientific and engineering studies.
+One example is neuroimaging~\citep{sun2017store,zhou2013tensor}.
+The brain connectivity networks are collected from a sample of individuals, accompanied by individual characteristics such as age, gender, and diseases status (see Figure~\ref{fig:intro1}a).
+Another example is in the field of network analysis~\citep{baldin2018optimal,hoff2005bilinear}.
+A typical social network consists of nodes that represent people and edges that represent the friendships.
+Other covariates such as people’s demographic information and the closeness of a friendship are often available.
+In both examples, it is of keen scientific interest to identify the variation in the response (e.g., brain connectivity, social community patterns) that may be affected by additional covariates.
+These seemingly different problems can be formulated as a regression problem in the tensor sense.
 
-{\bf Our contributions.} This paper presents a general treatment to these seemingly different problems.
-We formulate the learning task as a regression problem, with tensor observation serving as a response, and the node features and/or their interactions forming the predictor.
-Figure~\ref{fig:intro1}b illustrates the general set-up we consider.
-The regression approach allows the identification of variation in the data tensor that is explained by the covariates.
-Our model greatly improves the classical tensor regression~\cite{rabusseau2016low,yu2016learning} by incorporating covariates from multiple modes and the interactions thereof.
-The statistical convergence of our estimator is established, and we quantify the gain in predictive power.
-
-A related contribution is that our method allows a broad range of tensor types, including continuous, count, and binary observations.
-While previous tensor regression methods~\cite{zhao2012higher,yu2016learning} are able to analyze Gaussian responses, none of them is suitable for exponential distribution family of tensors.
-We develop a generalized tensor regression framework, and as a by product, our models allows heteroscedasticity by relating the variance of tensor entry to its mean.
-This flexibility is particularly important in practice, because social network, brain imaging, or gene expression datasets are often non-Gaussian.
+Emerging instances of the non-Gaussian datasets also require the effective tools applicable to a wide range of data, such as continuous, binary and count data.
+Examples include the Political Social science data \citep{hu2015scalable} which is a count dataset, and the brain connectivity network data \citep{wang2019common} which is usually encoded as binary datasets.
+Easily treating these datasets as Gaussian-valued data will yield unstable estimations and lead the predictions fall outside the valid range.
+A number of previous methods have been proposed \citep{yu2016learning, llosa2018tensor, raskutti2015convex, zhao2012higher} to address the tensor regression problem in various forms (e.g.\ scalar-to-tensor regression, tensor-response regression) from distinct perspectives.
+However, these methods assume the Gaussianity in the model, treating the input as Gaussian valued entries.
+Novel methods for exponential family tensor regression are necessary this time.
 
@@
 {\bf Related work.} Our work is closely related to but also clearly distinctive from several lines of previous work.
-The first is a class of \emph{unsupervised} tensor decomposition such as Tucker and CP decomposition~\cite{de2000multilinear,hong2018generalized,zhang2018tensor}.
-Tucker decomposition is an unsupervised method that finds a low-rank representation of a data tensor.
-In contrast, our model is a {\it supervised} tensor model that identifies the association between a data tensor and multiple covariates.
+
+The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor,zhang2018tensor, hong2020generalized} that aims to find the best low-rank representation of data tensor.
+In contrast, our model can be viewed a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and covariates.
 The low-rank structure is determined jointly by the tensor response and matrix covariates.
-The second line of work studies the network-response model~\cite{rabusseau2016low, li2017parsimonious}.
-Earlier development of this model focuses mostly on binary data in the presence of dyadic covariates~\cite{hoff2005bilinear}.
-We will demonstrate the enhanced accuracy as the order of data grows, and establish the general theory for exponential family which is arguably better suited to various data types.
+
+The second line of work studies the scalar-to-tensor regression models, in which the response is a scalar and the \textit{predictor} is a tensor~\citep{zhou2013tensor,chen2019non}.
+Our proposal is orthogonal to theirs because we treat the tensor as a \textit{response}.
+The distinct model settings yield distinct focuses and interpretations.
+Take neuroimaging analysis as an example.
+The scalar-to-tensor regression aims to study the changes of the prediction outcomes when the brain connectivity network varies, while tensor-response model focuses on understanding the variation in the brain connectivity that may attribute to the individual features such as age, gender and disease status.
+Since scalar-to-tensor models do not utilize all the information in the response, they are not optimal when there exist correlations among response variables.
+In contrast, our model does not assume the independence among response entries, which allow the model to capture more structural information.
+
+The third line of work studies the tensor-on-tensor regression models, in which both the response and predictor are tensors~\citep{raskutti2015convex, lock2018tensor, llosa2018tensor, gahrooei2020multiple}.
+Though our model actually tackles the similar problem as tensor-on-tensor regression, our work differs with previous work in several ways.
+First, previous methods mainly develop the model with Gaussian error.
+Their models are not suitable for general cases.
+Second, our study differs from previous work in the scope of results.
+\citep{gahrooei2020multiple} verifies the performance of estimation mainly via simulations, while our study provides a theoretical evidence on estimation.
+\citep{lock2018tensor, llosa2018tensor} perform solid inferences on parameter estimation, however, they do not assess the estimation and prediction errors, which are studied in our paper.
+Risk bounds are well studied in~\citep{raskutti2015convex} and our convergence rate coincides with their results in the order-3 low-rank case.
+Whereas, the convexity assumption of the regularization in their model yields a convex optimization problem.
+In contrast,  our model tackles a non-convex optimization problem, and employee distinct techniques.
+
+{\bf Our contribution.} In this article, we present a general model and the associated theory for exponential family tensor regression.
+Exponential family tensor observation serves as the response, and the node features and/or the interactions form the matrix predictor.
+Figure~\ref{fig:intro1}b illustrates the general set-up we consider.
+Due to the generalized framework, our models allows heteroscedasticity by relating the variance of tensor entry to its mean.
+This ﬂexibility is particularly important in practice.
+The regression approach allows the identiﬁcation of variation in the data tensor that is explained by the covariates.
+We utilize a low-rank constraint in the regression coeﬃcient to encourage the sharing among tensor entries.
+The statistical convergence of our estimator is established, and we quantify the gain in predictive power.
 

### 2020-08-30 skills/math/tail_bound/tailbound_solution.tex
tail_bound
tail bound

@@
 	\end{align}
-	Therefore, let $ y = \frac{t}{2\nu} - \sqrt{\log C}$, we obtain $I_1$ ,
+	Therefore, let $ y = \frac{t}{2\nu} - \sqrt{\log C}$, we have 
 	\begin{align}
@@
 \end{equation}
+
+Moreover, we have
+\begin{equation}\label{eq:e2113}
+	\frac{\bbE[Z_n]}{\sqrt{2 \sigma^2 \log n}} \rightarrow 1, \quad \text{as $n \rightarrow +\infty$}.
+\end{equation}
+
 \end{lem}
@@
 \begin{proof}
-	%First, we prove the equation~\eqref{eq:e2111}.
+	First, we prove the equation~\eqref{eq:e2111}.
 
@@
 	 \end{equation}
-	 Hence, let $ u = \frac{ t}{\sigma}$, the last integral in equation~\eqref{eq:e211pro1} is upper bounded as
+	Let $ u = \frac{ t}{\sigma}$. Hence, for $c > \sigma$, the last integral in equation~\eqref{eq:e211pro1} is upper bounded as
 	 \begin{align}
@@
 
+	 \textbf{Alternative Way:} An alternative proof for equation~\eqref{eq:e2111} is below.
+
+	 Since $\{X_i\}_{i=1}^n$ are i.i.d.\ random normal variables following $N(0, \sigma^2)$, by Exercise~\ref{ex:12}, we have 
+	 \begin{equation}
+	 	\bbE[\max_{i \in [n]} X_i] \leq \sqrt{2 \sigma^2 \log n}, \quad \text{for all } n \geq 1.
+	 \end{equation}
+	 Note that $\max_{i \in [n]} |X_i| = \max_{i \in [2n]} X_i$, where $X_{n + i} = - X_i $, for all $i\in[n]$. Then, we have $\max_{i \in [n]}|X_i| \leq \sqrt{2 \sigma^2 \log(2n)}$. Note that
+	 \begin{equation}
+	 	 \of{\sqrt{2 \sigma^2 \log n} + \frac{4\sigma}{\sqrt{2 \log n}}}^2 = 2 \sigma^2 \log n + 8\sigma^2 \of{1 + \frac{1}{\log n}} \geq 2 \sigma^2 \log n + 2 \sigma^2 \log 2 = (\sqrt{2 \sigma^2 \log 2n} )^2.
+ 	 \end{equation}
+ 	 Therefore, we obtain the upper bound
+ 	 \begin{equation}
+ 	 	\bbE[Z_n] \leq \sqrt{2 \sigma^2 \log(2n)} \leq \sqrt{2 \sigma^2 \log n} + \frac{4\sigma}{\sqrt{2 \log n}}.
+ 	 \end{equation}
+
 	 \vspace{0.2cm}
@@
 
+	 \vspace{0.2cm}
+	 Last, we prove the convergence property~\eqref{eq:e2113}.
+
+	\textbf{My thoughts:} To prove $\lim_{n \rightarrow +\infty}\frac{ \bbE[Z_n]}{\sqrt{2\sigma^2 \log n}}  = 1$, we need to prove $\limsup_{n \rightarrow +\infty}\frac{ \bbE[Z_n]}{\sqrt{2\sigma^2 \log n}}  = \liminf_{n \rightarrow +\infty}\frac{ \bbE[Z_n]}{\sqrt{2\sigma^2 \log n}} = 1$.
+The $\limsup$ part is easy because of the upper bound~\eqref{eq:e2111}.
+Then, we only need to prove the $\liminf$ part.
+
+	Note that 
+	\[\bbE[Z_n] \geq \sqrt{2 \sigma^2 \log n}  \bbP \off{Z_n > \sigma \sqrt{2 \log n}} =  \sigma \sqrt{2\log n} \off{1 - \of{1 - \bbP \off{|X_i| > \sigma \sqrt{2\log n}}   }^n}. \]
+
+	If I can prove that $\lim_{n \rightarrow +\infty}\of{1 - \bbP \off{|X_i| > \sigma \sqrt{2\log n}}   }^n = 0 $, the problem is done.
+
+	Note that $\bbP \off{|X_i| > \sigma \sqrt{2\log n}}  = 2 \bbP \off{X_i > \sigma \sqrt{2\log n}} \leq 2 \sqrt{\frac{2}{\pi}} \frac{\sigma}{\sqrt{2 \sigma^2 \log n}} e^{-\log n} = \frac{2}{\pi n \sqrt{\log n}}$.
+Then, the limit is upper bounded
+	 \[\lim_{n \rightarrow +\infty}\of{1 - \bbP \off{|X_i| > \sigma \sqrt{2\log n}}   }^n \leq \lim_{n \rightarrow +\infty} \of{1 -  \frac{2}{\pi n \sqrt{\log n}} }^n = 1. \]
+
+	 I can not get the result I want.
+
 	 \iffalse
@@
 	The absolute Gaussian maxima $Z_n$ concentrates around $\sigma \sqrt{\log n}$ for large $n$.
-Standard Gaussian variables concentrate around 0, while the maxima of the $n$ standard Gaussian variables is no longer a Gaussian variables.
+Standard Gaussian variables concentrate around 0, while the maxima of the $n$ standard Gaussian variables is no longer a Gaussian variable.
 The scale of Gaussian maxima scales as $\sqrt{n}$, however, the sum of $n$ i.i.d.\ Gaussian variables scales as $n$.
@@
 
-\subsection{Exercise 2.12}
+\subsection{Exercise 2.12}\label{ex:12}
 \begin{lem}[Sharp upper bounds for sub-Gaussian maxima]
@@
 \begin{rmk}
-	The maxima for $n$ i.i.d.\ sub-Gaussian variables also share the similar concentration property as Gaussian variables in Exercise~\ref{Gausmax}.
-The maxima $\max_{i \in [n]} X_i$ concentrates around the sub-Gaussian parameter $\sigma$.
+	The maxima for $n$ i.i.d.\ sub-Gaussian variables is upper bounded by $\sigma \sqrt{\log n}$. The independence assumption is not necessary for the upper bound of sub-Gaussian maxima.
 \end{rmk}
@@
 \vspace{0.2cm}
+\textbf{My thoughts:} Note that $X_1 X_2 = \frac{1}{4}\of{(X_1 + X_2)^2 - (X_1 - X_2)^2}$.
+By part $(a)$, we have $X_1 + X_2$ and $X_1 - X_2$ are sub-Gaussian with the parameter $\sqrt{\sigma_1^2 + \sigma^2_2}$.
+Note that for a variable $X \sim \subG(\sigma^2)$, the variable $Z = X^2 - \bbE[X^2] \sim \text{subE}(16\sigma^2, 16\sigma^2)$.
+\vspace{0.2cm}
+
+The characteristic function of $X_1 X_2$ is 
+\begin{equation}\label{eq:e13c1}
+	\bbE[e^{t X_1 X_2}] = \bbE[e^{\frac{t}{4} \of{(X_1 + X_2)^2 - (X_1 - X_2)^2} }] = \bbE[ e^{\frac{t}{4} (X_1 +X_2 )^2} e^{-\frac{t}{4} (X_1 - X_2)^2 } ], \quad \text{for all } t \in \bbR.
+\end{equation} 
+By Cauchy-Schwartz inequality, the expectation~\eqref{eq:e13c1} satisfies
+\begin{equation}\label{eq:e13c2}
+	\bbE[ e^{\frac{t}{4} (X_1 +X_2 )^2} e^{-\frac{t}{4} (X_1 - X_2)^2 } ] \leq \of{\bbE[ e^{\frac{t}{2} (X_1 +X_2 )^2} ]}^{1/2} \of{\bbE[ e^{-\frac{t}{2} (X_1 -X_2 )^2} ]}^{1/2}.
+\end{equation}
+
+Let $Z_1 = (X_1 + X_2)^2 - \bbE[(X_1 + X_2)^2]$.
+Note that $\bbE[X^2_i] = \var{(X_i)} \leq \sigma^2_i $, for $ i = 1,2$, and the variable $Z_1 \sim \text{subE}(16(\sigma^2_1 + \sigma^2_2), 16(\sigma^2_1 + \sigma^2_2))$.
+We have
+\begin{equation}\label{eq:e13c3}
+	\bbE[ e^{\frac{t}{2} (X_1 +X_2 )^2} ] = \bbE[e^{\frac{t}{2} Z_1 + \frac{t}{2} \bbE[(X_1 + X_2)^2] }] \leq \bbE[e^{\frac{t}{2} Z_1}] e^{ \frac{t}{2} (\sigma_1^2 + \sigma_2^2) } \leq e^{32 t^2 (\sigma^2_1 + \sigma_2^2)^2 + \frac{t}{2} (\sigma^2_1 + \sigma_2^2) },
+\end{equation}
+for all  $|t| \leq \frac{1}{8 (\sigma^2_1 + \sigma_2^2)}$. Similarly, we have
+\begin{equation}\label{eq:e13c4}
+	\bbE[ e^{-\frac{t}{2} (X_1 -X_2 )^2} ]  \leq  e^{32 t^2 (\sigma^2_1 + \sigma_2^2)^2 - \frac{t}{2} (\sigma^2_1 + \sigma_2^2) },
+\end{equation}
+for all  $|t| \leq \frac{1}{8 (\sigma^2_1 + \sigma_2^2)}$. Plug the inequalities \eqref{eq:e13c3} and \eqref{eq:e13c4} into the inequality~\eqref{eq:e13c2}. We have
+\begin{equation}
+		\bbE[ e^{\frac{t}{4} (X_1 +X_2 )^2} e^{-\frac{t}{4} (X_1 - X_2)^2 } ]  \leq  e^{32 t^2 (\sigma^2_1 + \sigma_2^2)^2}, \quad \text{for all } |t| \leq \frac{1}{8 (\sigma^2_1 + \sigma_2^2)}.
+\end{equation}
+Therefore, the variable $X_1 X_2 \sim \text{subE}(8(\sigma^2_1 + \sigma^2_2),8(\sigma^2_1 + \sigma^2_2) )$.
 
@@
 	\begin{rmk}
-		The concentration around mean implies the concentration around median for $n$ variables, vice versa. 
+	If a variable concentrates around mean, the variable also concentrates around the median, vice versa.
 	\end{rmk}
@@
 	\begin{rmk}
-		Kernel density estimator satisfies the bounded difference property and possesses a concentration property around the mean.
+		Kernel density estimator satisfies the bounded difference property and possesses a concentration property.
+Kernel density estimator converges to its mean in probability at the speed of $e^{-n}$ when $n$ goes to infinity.
 	\end{rmk}
@@
     \begin{rmk}
-    	The concentration inequalities also apply to the Hilbert space. 
+    	The sample mean also for $n$ variables in a Hilbert space converge to its expectation in probability at the speed of $e^{-n}$.
     \end{rmk}
@@
     \begin{rmk}
-    	For a random vector $X$, Hanson-Wright Inequality implies the concentration property for the matrix product $X^T \mQ X$, where $\mQ$ is a positive semi-definite matrix.
+    	For a random vector $X$, Hanson-Wright Inequality implies the tail probability of matrix product $X^T \mQ X$ is upper bounded at the leve of $e^{-\min\{ \onorm{\mQ}^{-2}_F, \onorm{\mQ}^{-1}_{op} \}}$, where $\mQ$ is a positive semi-definite matrix.
     \end{rmk}
@@
     \begin{rmk}
-    	For a random variable $X$, the existence of $X$'s Orlicz norm implies the boundedness of the tail probability of $X$.
+    	For a random variable $X$, the existence of $X$'s Orlicz norm implies the tail probability of $X$ is upper bounded at the level of $e^{-\onorm{X}^{-q}_{\varphi} }$. 
     \end{rmk}
@@
 \begin{rmk}
-	The maxima of Orlicz variables concentrates around the Orlicz norm $\sigma$ and scales as $\psi^{-1}(n)$.
+	The maxima of Orlicz variables is upper bounded by the Orlicz norm $\sigma$ and scales as $\psi^{-1}(n)$.
 In previous exercises, the maxima of sub-Gaussian variables concentrates around the sub-Gaussian variable $\sigma$ and scales at $\log (n)$.
@@
  \begin{rmk}
- 	For $n$ independent variables $X_i$, the moment condition for $X_i$ implies the boundedness of the tail probability for the sample mean $\frac{1}{n} \sum_{i=1}^n X_i$.
+ 	For $n$ independent zero-mean variables $X_i$, the order $2m$ moment condition for $X_i$ implies that the sample mean $\frac{1}{n} \sum_{i=1}^n X_i$ converges to 0 at the speed of $n^{-m}$.
  \end{rmk}
@@
 
- \subsection{Exercise 2.22}
-
 \end{document}

### 2020-09-02 tensor_regress/jasa-latexsource/tensor-jasa.tex
diff
update

@@
 {
-  \title{\bf Exponential Family Tensor Regression with Covariates on Multiple Modes}
+  \title{\bf Supervised Tensor Decomposition with Covariates on Multiple Modes}
   \author{Jiaxin Hu, Zhuoyan Xu, and Miaoyan Wang\\
@@
 \begin{abstract}
-We consider the problem of tensor-response regression given covariates on multiple modes.
+We consider the problem of supervised tensor decomposition given covariates on multiple modes.
 Such data problems arise frequently in applications such as neuroimaging, network analysis, and spatial-temporal modeling.
@@
 \begin{center}
-\includegraphics[width=16cm]{demo.pdf}
+\includegraphics[width=16cm]{new_introfigure1.pdf}
 \end{center}

### 2020-09-02 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
 {
-  \title{\bf Supervised Tensor Decomposition with Covariates on Multiple Modes}
+  \title{\bf Supervised Tensor Decomposition with Side-information on Multiple Modes}
   \author{Jiaxin Hu, Zhuoyan Xu, and Miaoyan Wang\\
@@
 \begin{abstract}
-We consider the problem of supervised tensor decomposition given covariates on multiple modes.
+We consider the problem of supervised tensor decomposition with exponential-family entries given side information on multiple modes.
 Such data problems arise frequently in applications such as neuroimaging, network analysis, and spatial-temporal modeling.
-We propose a new family of tensor response regression models that incorporate covariates, and establish the theoretical accuracy guarantees.
-Unlike earlier methods, our estimation allows high-dimensionality in both the tensor response and the covariate matrices on multiple modes.
+We propose a new family of tensor response regression that treats side-information as incorporate covariates and we establish the accuracy guarantees for our proposed estimation.
+Unlike earlier methods on unsupervised tensor decomposition, our estimation allows high-dimensionality in both the tensor data and the covariate matrix data on multiple modes.
 An efficient alternating updating algorithm is further developed.

### 2020-09-02 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
 \begin{abstract}
-We consider the problem of supervised tensor decomposition with exponential-family entries given side information on multiple modes.
+We consider the problem of decomposing a high-order tensor with exponential-family entries, given side information on multiple modes.
 Such data problems arise frequently in applications such as neuroimaging, network analysis, and spatial-temporal modeling.
 We propose a new family of tensor response regression that treats side-information as incorporate covariates and we establish the accuracy guarantees for our proposed estimation.
-Unlike earlier methods on unsupervised tensor decomposition, our estimation allows high-dimensionality in both the tensor data and the covariate matrix data on multiple modes.
+Unlike earlier methods on unsupervised tensor decomposition, our supervised approach allows high-dimensionality in both the tensor data and the covariate matrices on multiple modes.
 An efficient alternating updating algorithm is further developed.

### 2020-09-04 tensor_regress/jasa-latexsource/tensor-jasa.tex
upload

--- 2020-09-02 tensor_regress/jasa-latexsource/tensor-jasa.tex
@@
 {
-  \title{\bf Supervised Tensor Decomposition with Covariates on Multiple Modes}
+  \title{\bf Supervised Tensor Decomposition with Side Information on Multiple Modes}
   \author{Jiaxin Hu, Zhuoyan Xu, and Miaoyan Wang\\
@@
   \begin{center}
-    {\LARGE\bf Exponential Family Tensor Regression with Covariates on Multiple Modes}
+    {\LARGE\bf General Tensor Regression with Covariates on Multiple Modes}
 \end{center}
@@
 \begin{abstract}
-We consider the problem of supervised tensor decomposition given covariates on multiple modes.
-Such data problems arise frequently in applications such as neuroimaging, network analysis, and spatial-temporal modeling.
-We propose a new family of tensor response regression models that incorporate covariates, and establish the theoretical accuracy guarantees.
-Unlike earlier methods, our estimation allows high-dimensionality in both the tensor response and the covariate matrices on multiple modes.
+Higher-order tensors have recently received increasing attention across science and engineering.
+While most tensor decomposition methods are developed for a single tensor observation, scientific applications often generate side information, including node features and interactions thereof, together with the tensor data.
+Identifying the relationship between a high-dimensional tensor and multi-mode features is an important yet challenging problem in neuroimaging, network analysis, and spatial-temporal modeling. 
+Here, we develop a regression approach that incorporates side information as structured covariates, and present several case studies to show the benefit of proposed approach.  
+Unlike unsupervised tensor decomposition, our supervised tensor factors capture the most effective dimension reduction of the data tensor confined to the feature spaces. 
+%supervised tensor decomposition method that incorporates side information on multiple modes. 
+%We consider the problem of decomposing a higher-order tensor with exponential-family entries, given side information on multiple modes. 
+%Such data problems arise frequently in applications such as neuroimaging, network analysis, and spatial-temporal modeling. 
+%We develop a regression approach that incorporates side information as structured covariates along multiple sides. 
+%Unlike unsupervised tensor decomposition, the supervised approach finds the optimal low-rank factor matrices confined to the feature spaces. 
 An efficient alternating updating algorithm is further developed.
 Our proposal handles a broad range of data types, including continuous, count, and binary observations.
-Through simulation and applications to two real datasets, we demonstrate the outperformance of our approach over the state-of-art.
+We apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
+Our approach identifies the key global connectivity pattern and pinpoints the local regions that are associated with covariates.
+
 \end{abstract}
@@
 \noindent%
-{\it Keywords:}  Tensor-response models, Multiple covariates, Dimension reduction, Reduced-rank regression, Generalized linear models 
+{\it Keywords:} Applications and case studies, Supervised dimension reduction, Exponential family observation, Reduced-rank tensor regression
 \vfill
@@
 
-Nowadays, multi-dimensional arrays, a.k.a.\ tensors, are collected with additional covariates on multiple modes in many modern scientific and engineering studies.
-One example is neuroimaging~\citep{sun2017store,zhou2013tensor}.
+Multi-dimensional arrays, known as tensors, are collected with additional covariates on multiple modes in many modern scientific and engineering studies.
+A popular example is the filed of neuroimaging~\citep{sun2017store,zhou2013tensor}.
 The brain connectivity networks are collected from a sample of individuals, accompanied by individual characteristics such as age, gender, and diseases status (see Figure~\ref{fig:intro1}a).
@@
 A typical social network consists of nodes that represent people and edges that represent the friendships.
-Other covariates such as people’s demographic information and the closeness of a friendship are often available.
-In both examples, it is of keen scientific interest to identify the variation in the response (e.g., brain connectivity, social community patterns) that may be affected by additional covariates.
+Covariates such as people’s demographic information and the closeness of a friendship are often available.
+In both examples, it is of keen scientific interest to identify the variation in the array data (e.g., brain connectivity, social community patterns) that is affected by additional covariates.
 These seemingly different problems can be formulated as a regression problem in the tensor sense.
@@
 \centering
-\includegraphics[width=15cm]{loglike}
-\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c)Poisson models.
+\includegraphics[width=15cm]{new_log.pdf}
+\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c) Poisson models.
 The dashed line represents the objective value at true parameter $\tL_{\tY}(\tB_{true})$.
@@
 
+\begin{table}[H]
+\begin{tabular}{c|l}
+\hline
+\multirow{4}{*}{Cluster I}  &officialvisits, intergovorgs,  militaryactions, violentactions, duration, \\
+&negativebehavior, boycottembargo, aidenemy, negativecomm, accusation, \\
+& protestsunoffialacts,    nonviolentbehavior, emigrants, relexports,   \\
+& timesincewar, commonbloc2, rintergovorgs3, relintergovorgs\\
+\hline
+\multirow{4}{*}{Cluster II}& economicaid, booktranslations, tourism, relbooktranslations, releconomicaid, \\
+& conferences, severdiplomatic, expeldiplomats, attackembassy, unweightedunvote,\\
+& reltourism, tourism3, relemigrants, emigrants3, students, relstudents, \\
+&exports, exports3, lostterritory, dependent, militaryalliance, warning  \\
+\hline
+\multirow{2}{*}{Cluster III} & treaties, reltreaties, exportbooks, relexportbooks, weightedunvote, ngo, \\
+&relngo, ngoorgs3, embassy, reldiplomacy, timesinceally, independence, commonbloc1\\
+\hline
+Cluster IV & commonbloc0, blockpositionindex\\
+\end{tabular}
+\caption{$K$-means clustering of relations based on factor matrix in the coefficient tensor.}\label{tab:s1}
+\end{table}
+
 \section{Conclusion}
@@
 
-\subsection{Nations data analysis}
-Table~\ref{tab:s1} summarizes the $K$-means clustering of the 56 relations based on the $3^{\text{rd}}$ mode factor $\mM_3\in\mathbb{R}^{56\times 4}$ in the tensor regression model.
-
-\begin{table}[H]
-\begin{tabular}{c|l}
-\hline
-\multirow{4}{*}{Cluster I}  &officialvisits, intergovorgs,  militaryactions, violentactions, duration, \\
-&negativebehavior, boycottembargo, aidenemy, negativecomm, accusation, \\
-& protestsunoffialacts,    nonviolentbehavior, emigrants, relexports,   \\
-& timesincewar, commonbloc2, rintergovorgs3, relintergovorgs\\
-\hline
-\multirow{4}{*}{Cluster II}& economicaid, booktranslations, tourism, relbooktranslations, releconomicaid, \\
-& conferences, severdiplomatic, expeldiplomats, attackembassy, unweightedunvote,\\
-& reltourism, tourism3, relemigrants, emigrants3, students, relstudents, \\
-&exports, exports3, lostterritory, dependent, militaryalliance, warning  \\
-\hline
-\multirow{2}{*}{Cluster III} & treaties, reltreaties, exportbooks, relexportbooks, weightedunvote, ngo, \\
-&relngo, ngoorgs3, embassy, reldiplomacy, timesinceally, independence, commonbloc1\\
-\hline
-Cluster IV & commonbloc0, blockpositionindex\\
-\end{tabular}
-\caption{$K$-means clustering of relations based on factor matrix in the coefficient tensor.}\label{tab:s1}
-\end{table}
-
 \section*{Acknowledgements}

### 2020-09-05 tensor_regress/jasa-latexsource/tensor-jasa.tex
revise
update

@@
 \usepackage{dsfont,multirow,hyperref,setspace,enumerate}
-\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}
+\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue}}
 
@@
 {
-  \title{\bf Supervised Tensor Decomposition with Side Information on Multiple Modes}
+  \title{\bf Supervised Tensor Decomposition with Multiple Side Information}
   \author{Jiaxin Hu, Zhuoyan Xu, and Miaoyan Wang\\
@@
 \begin{abstract}
-Higher-order tensors have recently received increasing attention across science and engineering.
-While most tensor decomposition methods are developed for a single tensor observation, scientific applications often generate side information, including node features and interactions thereof, together with the tensor data.
-Identifying the relationship between a high-dimensional tensor and multi-mode features is an important yet challenging problem in neuroimaging, network analysis, and spatial-temporal modeling. 
-Here, we develop a regression approach that incorporates side information as structured covariates, and present several case studies to show the benefit of proposed approach.  
-Unlike unsupervised tensor decomposition, our supervised tensor factors capture the most effective dimension reduction of the data tensor confined to the feature spaces. 
+Higher-order tensors have received increased attention across science and engineering.
+While most tensor decomposition methods are developed for a single tensor observation, scientific studies often collect side information, in the form of node features and interactions thereof, together with the tensor data.
+Such data problems are common in neuroimaging, network analysis, and spatial-temporal modeling.
+Identifying the relationship between a high-dimensional tensor and side information is important yet challenging.
+Here, we develop a tensor decomposition method that incorporates multiple side information as structured covariates, and present case studies to show the benefit of proposed approach.
+Unlike unsupervised tensor decomposition, our supervised decomposition captures the most effective dimension reduction of the data tensor confined to the feature space on each mode.
 %supervised tensor decomposition method that incorporates side information on multiple modes. 
@@
 We apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
-Our approach identifies the key global connectivity pattern and pinpoints the local regions that are associated with covariates.
+We identify the key global connectivity pattern and pinpoints the local regions that are associated with covariates.
+Our method will help the practitioners efficiently analyze tensor datasets in various areas.
+Toward this end, the package and data used are available at ~\url{https://CRAN.R-project.org/package=tensorregress}.
 
@@
 \noindent%
-{\it Keywords:} Applications and case studies, Supervised dimension reduction, Exponential family observation, Reduced-rank tensor regression
+{\it Keywords:} Applications and case studies, Supervised dimension reduction, Exponential family observation, Tensor decomposition
 \vfill
@@
 
-Multi-dimensional arrays, known as tensors, are collected with additional covariates on multiple modes in many modern scientific and engineering studies.
-A popular example is the filed of neuroimaging~\citep{sun2017store,zhou2013tensor}.
+Multi-dimensional arrays, known as tensors, are often collected with side information on multiple modes in modern scientific and engineering studies.
+A popular example is in neuroimaging~\citep{sun2017store,zhou2013tensor}.
 The brain connectivity networks are collected from a sample of individuals, accompanied by individual characteristics such as age, gender, and diseases status (see Figure~\ref{fig:intro1}a).
@@
 A typical social network consists of nodes that represent people and edges that represent the friendships.
-Covariates such as people’s demographic information and the closeness of a friendship are often available.
-In both examples, it is of keen scientific interest to identify the variation in the array data (e.g., brain connectivity, social community patterns) that is affected by additional covariates.
-These seemingly different problems can be formulated as a regression problem in the tensor sense.
+Side information such as people’s demographic information and the closeness of a friendship are often available.
+In both examples, it is of keen scientific interest to identify the variation in the tensor observation (e.g., brain connectivities, social community patterns) that is affected by available feature data.
+These seemingly different scenarios pose a common yet challenging problem for tensor modeling.
 
-Emerging instances of the non-Gaussian datasets also require the effective tools applicable to a wide range of data, such as continuous, binary and count data.
-Examples include the Political Social science data \citep{hu2015scalable} which is a count dataset, and the brain connectivity network data \citep{wang2019common} which is usually encoded as binary datasets.
-Easily treating these datasets as Gaussian-valued data will yield unstable estimations and lead the predictions fall outside the valid range.
-A number of previous methods have been proposed \citep{yu2016learning, llosa2018tensor, raskutti2015convex, zhao2012higher} to address the tensor regression problem in various forms (e.g.\ scalar-to-tensor regression, tensor-response regression) from distinct perspectives.
-However, these methods assume the Gaussianity in the model, treating the input as Gaussian valued entries.
-Novel methods for exponential family tensor regression are necessary this time.
+In addition to the aforementioned challenges, many instances of tensor observation consist of non-Gaussian measurements.
+Examples include the political interaction dataset \citep{hu2015scalable} which measures action counts between countries for various relationship, and the brain connectivity network data \citep{wang2019common} which is encoded as binary adjacency matrices.
+Classical tensor decomposition methods are based on minimizing the Frobenious norm of the reconstruction error, leading to suboptimal predictions for binary or count responses.
+A number of supervised tensor methods have been proposed \citep{yu2016learning, raskutti2015convex, zhao2012higher} to address the tensor regression problem in various forms (e.g.\ scalar-to-tensor regression, tensor-response regression).
+These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the covariate matrices, both of which are less suitable for applications of our interest.
+The gap in theory and practice suggests a huge opportunity for improved modeling paradigms that better capture the complexity in tensor data with side information.
 
@@
 
-{\bf Related work.} Our work is closely related to but also clearly distinctive from several lines of previous work.
+We present a modeling paradigm and associated method for analyzing exponential family tensors with multiple side information. 
+We formulate the learning task as a structured regression problem, with tensor observation serving as the response, and the side information forming multiple covariate matrices.
+Figure~\ref{fig:intro1}b illustrates our model in the special case of order-3 tensors.
+A low-rank structure is imposed to the mean tensor $\mathbb{E}(\tY)\in\mathbb{R}^{d_1\times d_2\times d_3}$, where unlike classical decomposition, the tensor factors are assumed to belong to the subspace spanned by known features $\mX_k\in\mathbb{R}^{p_k\times d_k}$ for $k=1,2,3$.
+The unknown matrix $\mM_k\in\mathbb{R}^{p_k\times r_k}$ link the data tensor to each of the feature matrices, thus allowing the identification the variation in the tensor entries attributable to the side information
 
+In addition, we use ideas from generalized linear model (GLM) to allow heteroscedacity due to mean-variance relationship across tensor entries.
+This flexibility is important in practice.
+In classical GLM, the sample size $n$ and feature dimension $d$ are well defined; however, in the tensor data analysis, we observe only one single realization of tensor response but up to $K$ feature matrices, where $K$ is the number of tensor modes.
+Both the sample size and feature dimension grow exponentially in $K$.
+We overcome the high-dimensionality using a possibly non-linear, low-rank constraint in the mean tensor, thereby encouraging the sharing among entries in the data tensor.
+The statistical convergence of our estimator is established, and we quantify the gain in predictive power through several case studies.
+
+Our work is closely related to but also clearly distinctive from several lines of previous work.
 The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor,zhang2018tensor, hong2020generalized} that aims to find the best low-rank representation of data tensor.
@@
 The second line of work studies the scalar-to-tensor regression models, in which the response is a scalar and the \textit{predictor} is a tensor~\citep{zhou2013tensor,chen2019non}.
-Our proposal is orthogonal to theirs because we treat the tensor as a \textit{response}.
+Our proposal is orthogonal to theirs because we treat the tensor as random variables.
 The distinct model settings yield distinct focuses and interpretations.
 Take neuroimaging analysis as an example.
-The scalar-to-tensor regression aims to study the changes of the prediction outcomes when the brain connectivity network varies, while tensor-response model focuses on understanding the variation in the brain connectivity that may attribute to the individual features such as age, gender and disease status.
-Since scalar-to-tensor models do not utilize all the information in the response, they are not optimal when there exist correlations among response variables.
-In contrast, our model does not assume the independence among response entries, which allow the model to capture more structural information.
+The scalar-to-tensor regression aims to study the changes of the disease status when the brain connectivity network varies, while tensor-response model focuses on understanding the variation in the brain connectivity that is attribute to the individual features such as age, gender and disease status.
+Since scalar-to-tensor models do not model the stochastic noise in the tensor, they are not suited for dimension reduction for tensor data.
 
-The third line of work studies the tensor-on-tensor regression models, in which both the response and predictor are tensors~\citep{raskutti2015convex, lock2018tensor, llosa2018tensor, gahrooei2020multiple}.
+The third line of work studies the tensor-on-tensor regression models, in which both the response and predictor are tensors~\citep{raskutti2015convex, lock2018tensor, gahrooei2020multiple}.
+
 Though our model actually tackles the similar problem as tensor-on-tensor regression, our work differs with previous work in several ways.
-First, previous methods mainly develop the model with Gaussian error.
+Previous methods mainly develop the model with Gaussian error.
 Their models are not suitable for general cases.
@@
 \citep{gahrooei2020multiple} verifies the performance of estimation mainly via simulations, while our study provides a theoretical evidence on estimation.
-\citep{lock2018tensor, llosa2018tensor} perform solid inferences on parameter estimation, however, they do not assess the estimation and prediction errors, which are studied in our paper.
+\citep{lock2018tensor} perform Bayesian inferences on parameter estimation, however, they do not assess the estimation and prediction errors, which are studied in our paper.
 Risk bounds are well studied in~\citep{raskutti2015convex} and our convergence rate coincides with their results in the order-3 low-rank case.
@@
 
-{\bf Our contribution.} In this article, we present a general model and the associated theory for exponential family tensor regression.
-Exponential family tensor observation serves as the response, and the node features and/or the interactions form the matrix predictor.
-Figure~\ref{fig:intro1}b illustrates the general set-up we consider.
-Due to the generalized framework, our models allows heteroscedasticity by relating the variance of tensor entry to its mean.
-This ﬂexibility is particularly important in practice.
-The regression approach allows the identiﬁcation of variation in the data tensor that is explained by the covariates.
-We utilize a low-rank constraint in the regression coeﬃcient to encourage the sharing among tensor entries.
-The statistical convergence of our estimator is established, and we quantify the gain in predictive power.
-
 \section{Preliminaries}
 
-We begin by reviewing the basic properties about tensors~\citep{kolda2009tensor}.
+We summarize the basic tensor properties used in the paper; more details can be found in~\citet{kolda2009tensor}.
 We use $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ to denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
@@
 \section{Motivation and model}\label{sec:model}
+We start with general model formulation and then provide concrete examples that motivate the model.
 Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor.
@@
 Let $\mX_k\in\mathbb{R}^{d_k\times p_k}$ denote the available covariates on the mode $k$, where $p_k\leq d_k$.
-We propose a multilinear structure on the conditional expectation of the tensor.
+We consider a multilinear structure on the conditional expectation of the tensor.
 Specifically,
@@
 We assume that we have an additional information on an upper bound $\alpha>0$ such that $\mnormSize{}{\Theta}\leq \alpha$.
-This is the case for many applications we have in mind such as brain network analysis where fiber connections are bounded.
 We propose a constrained maximum likelihood estimator (MLE) for the coefficient tensor:
@@
 
-\begin{cor}[Spatio-temporal growth model] The estimated type-by-time-by-space coefficient tensor converges at the rate $\tO\left({p+r+s\over dmn}\right)$ where $p\leq d$, $r\leq m$ and $s\leq n$.
+\begin{example}[Spatio-temporal growth model] The estimated type-by-time-by-space coefficient tensor converges at the rate $\tO\left({p+r+s\over dmn}\right)$ where $p\leq d$, $r\leq m$ and $s\leq n$.
 The estimation achieves consistency as long as the dimension grows in either of the three modes.
-\end{cor}
+\end{example}
 
-\begin{cor} [Network population model] The estimated node-by-node-by-covariate tensor converges at the rate $\tO\left({2d+p\over d^2n}\right)$ where $p\leq n$.
+\begin{example} [Network population model] The estimated node-by-node-by-covariate tensor converges at the rate $\tO\left({2d+p\over d^2n}\right)$ where $p\leq n$.
 The estimation achieves consistency as the number of individuals or the number of nodes grows.
-\end{cor}
+\end{example}
 
-\begin{cor} [Dyadic data with node attributes] The estimated covariate-by-covariate matrix converges at the rate $\tO\left({p\over d^2}\right)$ where $p\leq d$.
+\begin{example} [Dyadic data with node attributes] The estimated covariate-by-covariate matrix converges at the rate $\tO\left({p\over d^2}\right)$ where $p\leq d$.
 Again, our estimation achieves consistency as the number of nodes grows.
-\end{cor}
+\end{example}
 
@@
 Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$.
+
+\subsection{Missing values}
+.....
 
@@
 \section{Data analysis}
-We apply our tensor regression model to two real datasets.
+We apply our tensor regression model to two datasets.
 The first application concerns the brain network modeling in response to individual attributes (i.e.\ covariate on one mode), and the second application focuses on multi-relational network analysis with dyadic attributes (i.e.\ covariates on two modes).
-\subsection{Human Connectome Project (HCP)}
-The Human connectome project (HCP~\citep{HCP}) aims to build a ``network map'' that characterizes the anatomical and functional connectivity within healthy human brains.
+
+\subsection{Analysis of human brain connection data}
+The Human Connectome Project (HCP) aims to build a ``network map'' that characterizes the anatomical and functional connectivity within healthy human brains~\citep{HCP}.
 We take a subset of HCP data that consists of 136 brain structural networks, one for each individual.
@@
 
-\subsection{Nations data}
+Figure~\ref{fig:s1} compares the estimated coefficients from our method (tensor regression) with those from classical GLM approach.
+A classical GLM is to regress the brain edges, one at a time, on the individual-level covariates, and this logistic model is repeatedly fitted for every edge $\in [68]\times [68]$.
+As we can see in the figure, our tensor regression shrinkages the coefficients towards center, thereby enforcing the sharing between coefficient entries.
+
+\subsection{Analysis of political relational data}
 
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=10cm]{coef.pdf}
+\caption{Effect estimation in the \emph{Nations} data.
+Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warnning}, \emph{economicaid},\emph{intergovorg}, and \emph{commonblock}, respectively.
+}\label{fig:est}
+
+\end{figure}
+
 We applied our tensor regression model to the \emph{Nations} data.
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=10cm]{coef.pdf}
-\caption{Effect estimation in the \emph{Nations} data.
-Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warnning}, \emph{economicaid},\emph{intergovorg}, and \emph{commonblock}, respectively.
-}\label{fig:est}
-
-\end{figure}
+\begin{table}[H]
+\begin{tabular}{c|c|l}
+\hline
+\multirow{4}{*}{Cluster I}  & Government &officialvisits, intergovorgs,  militaryactions, violentactions, duration, \\
+&&negativebehavior, boycottembargo, aidenemy, negativecomm, accusation, \\
+& &protestsunoffialacts,    nonviolentbehavior, emigrants, relexports,   \\
+& &timesincewar, commonbloc2, rintergovorgs3, relintergovorgs\\
+\hline
+\multirow{4}{*}{Cluster II}& Economics &economicaid, booktranslations, tourism, relbooktranslations, releconomicaid, \\
+&& conferences, severdiplomatic, expeldiplomats, attackembassy, unweightedunvote,\\
+&& reltourism, tourism3, relemigrants, emigrants3, students, relstudents, \\
+&&exports, exports3, lostterritory, dependent, militaryalliance, warning  \\
+\hline
+\multirow{2}{*}{Cluster III} & treaties &treaties, reltreaties, exportbooks, relexportbooks, weightedunvote, ngo, \\
+&&relngo, ngoorgs3, embassy, reldiplomacy, timesinceally, independence, commonbloc1\\
+\hline
+Cluster IV & Politics &commonbloc0, blockpositionindex\\
+\end{tabular}
+\caption{$K$-means clustering of relations based on factor matrix in the coefficient tensor.}\label{tab:s1}
+\end{table}
 
@@
 This is not surprising, as the block partition during Cold War is associated with the \emph{constitutional} attribute.
-
-\begin{table}[H]
-\begin{tabular}{c|l}
-\hline
-\multirow{4}{*}{Cluster I}  &officialvisits, intergovorgs,  militaryactions, violentactions, duration, \\
-&negativebehavior, boycottembargo, aidenemy, negativecomm, accusation, \\
-& protestsunoffialacts,    nonviolentbehavior, emigrants, relexports,   \\
-& timesincewar, commonbloc2, rintergovorgs3, relintergovorgs\\
-\hline
-\multirow{4}{*}{Cluster II}& economicaid, booktranslations, tourism, relbooktranslations, releconomicaid, \\
-& conferences, severdiplomatic, expeldiplomats, attackembassy, unweightedunvote,\\
-& reltourism, tourism3, relemigrants, emigrants3, students, relstudents, \\
-&exports, exports3, lostterritory, dependent, militaryalliance, warning  \\
-\hline
-\multirow{2}{*}{Cluster III} & treaties, reltreaties, exportbooks, relexportbooks, weightedunvote, ngo, \\
-&relngo, ngoorgs3, embassy, reldiplomacy, timesinceally, independence, commonbloc1\\
-\hline
-Cluster IV & commonbloc0, blockpositionindex\\
-\end{tabular}
-\caption{$K$-means clustering of relations based on factor matrix in the coefficient tensor.}\label{tab:s1}
-\end{table}
 
@@
 
-\section{Additional results for real data analysis}
-
-\subsection{HCP data analysis}
-
-Figure~\ref{fig:s1} compares the estimated coefficients from our method (tensor regression) with those from classical GLM approach.
-A classical GLM is to regress the brain edges, one at a time, on the individual-level covariates, and this logistic model is repeatedly fitted for every edge $\in [68]\times [68]$.
-As we can see in the figure, our tensor regression shrinkages the coefficients towards center, thereby enforcing the sharing between coefficient entries.
-
-\begin{figure}[t]
+\begin{figure}[H]
 \includegraphics[width=17cm]{compare.pdf}

### 2020-09-05 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
 \end{center}
-\caption{Examples of tensor response regression model with covariates on multiple modes. (a) Network population model. (b) Spatial-temporal growth model. }\label{fig:intro1}
+\caption{Examples of supervised tensor decomposition with multiple side information. (a) Network population model. (b) Spatial-temporal growth model. }\label{fig:intro1}
 \end{figure*}
 
-We present a modeling paradigm and associated method for analyzing exponential family tensors with multiple side information. 
+We present a general model and associated method for decomposing exponential family tensors with multiple side information. 
 We formulate the learning task as a structured regression problem, with tensor observation serving as the response, and the side information forming multiple covariate matrices.
@@
 A low-rank structure is imposed to the mean tensor $\mathbb{E}(\tY)\in\mathbb{R}^{d_1\times d_2\times d_3}$, where unlike classical decomposition, the tensor factors are assumed to belong to the subspace spanned by known features $\mX_k\in\mathbb{R}^{p_k\times d_k}$ for $k=1,2,3$.
-The unknown matrix $\mM_k\in\mathbb{R}^{p_k\times r_k}$ link the data tensor to each of the feature matrices, thus allowing the identification the variation in the tensor entries attributable to the side information
+The unknown ``transformation'' matrix $\mM_k\in\mathbb{R}^{p_k\times r_k}$ link the data tensor to each of the feature matrices, thus allowing the identification the variation in the tensor entries attributable to the side information
 
@@
 In contrast, our model can be viewed a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and covariates.
-The low-rank structure is determined jointly by the tensor response and matrix covariates.
+The low-rank structure is determined jointly by the tensor data and feature matrices on multiple sides.
 
@@
 Our proposal is orthogonal to theirs because we treat the tensor as random variables.
-The distinct model settings yield distinct focuses and interpretations.
-Take neuroimaging analysis as an example.
-The scalar-to-tensor regression aims to study the changes of the disease status when the brain connectivity network varies, while tensor-response model focuses on understanding the variation in the brain connectivity that is attribute to the individual features such as age, gender and disease status.
-Since scalar-to-tensor models do not model the stochastic noise in the tensor, they are not suited for dimension reduction for tensor data.
+The difference in model settings yields different focuses and interpretations.
+Consider the neuroimaging analysis as an example.
+The scalar-to-tensor regression aims to predict the individual response based on brain connectivity network, while tensor-to-scalar model focuses on understanding the variation in the brain connectivity that is attribute to the individual features such as age, gender and disease status.
+The scalar-to-tensor model is not suitable for tensor dimension reduction because it lacks the modeling of stochastic noise in the data tensor.
 
 The third line of work studies the tensor-on-tensor regression models, in which both the response and predictor are tensors~\citep{raskutti2015convex, lock2018tensor, gahrooei2020multiple}.
+Our model shares a common ground but is complementary to earlier approaches.
+As we show in Section~\ref{subsec:statprob}, the supervised tensor decomposition
+has an interesting interpretation as a special tensor-on-tensor regression.
+Nevertheless, our work differs from previous work in scope and application domains.
+Previous methods~\citep{gahrooei2020multiple,lock2018tensor} mainly focus on Gaussian tensor.
+The Frobenius norm used in the objective is statistical suboptimal for general exponential family tensors.
+Maximum likelihood estimator (MLE) is studied in~\cite{raskutti2015convex} and a convex relaxation algorithm is proposed to solve for low-rank tensor coefficients.
+However, in the tensor case, convex MLE suffers from both computational intractability and statistical suboptimality.
+We advocate the non-convex approach and provide strong evidence for its success in our setting.
+Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2015convex,gahrooei2020multiple}, and we focuses on finding optimal transformation matrices $\mM_k$ in order to pinpoint important feature subspaces towards prediction (see Figure~\ref{fig:intro1}b).
+The latter approach greatly improves the \emph{interpretability} in prediction.
+In this regards, our method opens up new opportunities for tensor data analysis in a wider range of possible applications.
 
-Though our model actually tackles the similar problem as tensor-on-tensor regression, our work differs with previous work in several ways.
-Previous methods mainly develop the model with Gaussian error.
-Their models are not suitable for general cases.
-Second, our study differs from previous work in the scope of results.
-\citep{gahrooei2020multiple} verifies the performance of estimation mainly via simulations, while our study provides a theoretical evidence on estimation.
-\citep{lock2018tensor} perform Bayesian inferences on parameter estimation, however, they do not assess the estimation and prediction errors, which are studied in our paper.
-Risk bounds are well studied in~\citep{raskutti2015convex} and our convergence rate coincides with their results in the order-3 low-rank case.
-Whereas, the convexity assumption of the regularization in their model yields a convex optimization problem.
-In contrast,  our model tackles a non-convex optimization problem, and employee distinct techniques.
+Section~\eqref{}
 
@@
 
-We summarize the basic tensor properties used in the paper; more details can be found in~\citet{kolda2009tensor}.
-We use $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ to denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
+We introduce the basic tensor properties used in the paper; more details can be found in~\citet{kolda2009tensor}.
+We use $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ to denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor, where $K$ corresponds to the number of modes of $\tY$ and is called the order.
 The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as
@@
 For ease of presentation, we use shorthand notion $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product.
-For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as $\langle \tY, \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}$.
-The Frobenius norm of tensor $\tY$ is defined as $\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}$.
-A higher-order tensor can be reshaped into a lower-order object~\citep{wang2017operator}.
+For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as
+\[
+\langle \tY, \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.
+\] 
+The Frobenius norm of $\tY$ is defined as 
+\[
+\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}.
+\]
+A higher-order tensor can be reshaped into a lower-order object.
 We use $\text{vec}(\cdot)$ to denote the operation that reshapes the tensor into a vector, and $\text{Unfold}_k(\cdot)$ the operation that reshapes the tensor along mode-$k$ into a matrix of size $d_k$-by-$\prod_{i\neq k}d_i$.
@@
 \section{Motivation and model}\label{sec:model}
-We start with general model formulation and then provide concrete examples that motivate the model.
+We begin with general model formulation and then provide motivating examples.
 Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor.
-Suppose we observe covariates on some of the $K$ modes.
-Let $\mX_k\in\mathbb{R}^{d_k\times p_k}$ denote the available covariates on the mode $k$, where $p_k\leq d_k$.
+Suppose we collect side information on some of the $K$ modes.
+Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the available feature matrix on the mode $k$, where $x_{ij}$ denotes the value of $j$-th feature for the $i$-th tensor entity, $p_k\leq d_k$.
 We consider a multilinear structure on the conditional expectation of the tensor.

### 2020-09-05 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
 Unlike unsupervised tensor decomposition, our supervised decomposition captures the most effective dimension reduction of the data tensor confined to the feature space on each mode.
-%supervised tensor decomposition method that incorporates side information on multiple modes. 
-%We consider the problem of decomposing a higher-order tensor with exponential-family entries, given side information on multiple modes. 
-%Such data problems arise frequently in applications such as neuroimaging, network analysis, and spatial-temporal modeling. 
-%We develop a regression approach that incorporates side information as structured covariates along multiple sides. 
-%Unlike unsupervised tensor decomposition, the supervised approach finds the optimal low-rank factor matrices confined to the feature spaces. 
 An efficient alternating updating algorithm is further developed.
@@
 
-Section~\eqref{}
+The remainder of this paper is organized as follows.
+Section~\ref{sec:pre} introduces tensor notation and preliminaries.
+Section~\ref{sec:model} presents the main model and three motivating examples for supervised tensor decomposition.
+Section~\ref{sec:est} describes a rank-constrained log-likelihood approach for tensor estimation and its statistical properties.
+We next develop an alternating optimization algorithm in Section~\ref{sec:algorithm} and present the simulation results in Section~\ref{sec:simulation}.
+In Section~\ref{sec:data}, we apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
+We conclude in Section~\ref{sec:con} with discussions about our findings and avenues of future work.
 
-\section{Preliminaries}
+\section{Preliminaries}\label{sec:pre}
 
-We introduce the basic tensor properties used in the paper; more details can be found in~\citet{kolda2009tensor}.
+We introduce the basic tensor properties used in the paper; more details on tensor notation can be found in~\citet{kolda2009tensor}.
 We use $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ to denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor, where $K$ corresponds to the number of modes of $\tY$ and is called the order.
@@
 
-\section{Rank-constrained likelihood-based estimation}
+\section{Rank-constrained likelihood-based estimation}\label{sec:est}
 We develop a likelihood-based procedure to estimate the coefficient tensor $\tB$ in~\eqref{eq:tensormodel2}.
@@
 
-\section{Numerical implementation}
+\section{Numerical implementation}\label{sec:algorithm}
 \subsection{Alternating optimization} \label{subsec:alg}
@@
 
-\section{Data analysis}
+\section{Data analysis}\label{sec:data}
 We apply our tensor regression model to two datasets.
@@
 
-\section{Conclusion}
+\section{Conclusion}\label{sec:con}
 

### 2020-09-05 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
 Figure~\ref{fig:intro1}b illustrates our model in the special case of order-3 tensors.
-A low-rank structure is imposed to the mean tensor $\mathbb{E}(\tY)\in\mathbb{R}^{d_1\times d_2\times d_3}$, where unlike classical decomposition, the tensor factors are assumed to belong to the subspace spanned by known features $\mX_k\in\mathbb{R}^{p_k\times d_k}$ for $k=1,2,3$.
-The unknown ``transformation'' matrix $\mM_k\in\mathbb{R}^{p_k\times r_k}$ link the data tensor to each of the feature matrices, thus allowing the identification the variation in the tensor entries attributable to the side information
+A low-rank structure is imposed to the mean tensor $\mathbb{E}(\tY)\in\mathbb{R}^{d_1\times d_2\times d_3}$, where unlike classical decomposition, the tensor factors are assumed to belong to the space spanned by known features $\mX_k\in\mathbb{R}^{p_k\times d_k}$ for $k=1,2,3$.
+The unknown matrix (referred to as ``dimension reduction matrix'') $\mM_k\in\mathbb{R}^{r_k\times p_k}$ link the tensor factor to the feature space, thus allowing the identification the variation in the tensor entries attributable to the side information.
 
@@
 In contrast, our model can be viewed a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and covariates.
-The low-rank structure is determined jointly by the tensor data and feature matrices on multiple sides.
+The low-rank factorization is determined jointly by the tensor data and feature matrices on multiple sides.
 
@@
 We advocate the non-convex approach and provide strong evidence for its success in our setting.
-Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2015convex,gahrooei2020multiple}, and we focuses on finding optimal transformation matrices $\mM_k$ in order to pinpoint important feature subspaces towards prediction (see Figure~\ref{fig:intro1}b).
+Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2015convex,gahrooei2020multiple}, and we go step further by finding the supervised factors $\mM_k\mX_k$ in order to pinpoint important features for prediction (see Figure~\ref{fig:intro1}b).
 The latter approach greatly improves the \emph{interpretability} in prediction.
@@
 Section~\ref{sec:model} presents the main model and three motivating examples for supervised tensor decomposition.
-Section~\ref{sec:est} describes a rank-constrained log-likelihood approach for tensor estimation and its statistical properties.
-We next develop an alternating optimization algorithm in Section~\ref{sec:algorithm} and present the simulation results in Section~\ref{sec:simulation}.
+Section~\ref{sec:est} describes a rank-constrained likelihood-based estimation and associated alternating optimization algorithm.
+In Section~\ref{sec:simulation}, we present simulation results and performance comparison to alternative methods.
 In Section~\ref{sec:data}, we apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
@@
 \[
-\langle \tY, \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.
+\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.
 \] 
@@
 We use $\text{vec}(\cdot)$ to denote the operation that reshapes the tensor into a vector, and $\text{Unfold}_k(\cdot)$ the operation that reshapes the tensor along mode-$k$ into a matrix of size $d_k$-by-$\prod_{i\neq k}d_i$.
-The Tucker rank of an order-$K$ tensor $\tY$ is defined as a length-$K$ vector $\mr=(r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\text{Unfold}_k(\tY)$, $k=1,\ldots,K$.
-We use lower-case letters (e.g.,\ $a,b,c$) for scalars/vectors, upper-case boldface letters (e.g.,\ $\mA,\mB,\mC,$) for matrices, and calligraphy letters (e.g.,\ $\tA, \tB, \tC,$) for tensors of order three or greater.
-We let $\mI_d$ denote the $d \times d$ identity matrix, $[d]$ denote the $d$-set $\{1,\ldots,d\}$, and allow an $\mathbb{R}\to \mathbb{R}$ function to be applied to tensors in an element-wise manner.
+The multilinear rank of an order-$K$ tensor $\tY$ is defined as a length-$K$ vector $\mr=(r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\text{Unfold}_k(\tY)$, $k=1,\ldots,K$.
+Given a matrix $\mM\in\mathbb{R}^{d\times r}$, we use $\text{Span}(\mM)$ to denote the space spanned by columns of $\mM$.
+We use lower-case letters (e.g.,\ $a,b,c$) for scalars/vectors, upper-case boldface letters (e.g.,\ $\mA,\ \mB,\ \mC$) for matrices, and calligraphy letters (e.g.,\ $\tA,\ \tB, \ \tC$) for tensors of order three or greater.
+We let $\mI_d$ denote the $d \times d$ identity matrix and $[d]$ denote the $d$-set $\{1,\ldots,d\}$.
+For ease of notation, we allow the basic arithmetic operators ($+$, $-$, $\geq $) and functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner.
 
 \section{Motivation and model}\label{sec:model}
-We begin with general model formulation and then provide motivating examples.
+\subsection{General framework for supervised tensor decomposition}
+We first present a general framework for supervised tensor decomposition and then discuss its implication in three concrete examples.
 Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor.
-Suppose we collect side information on some of the $K$ modes.
-Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the available feature matrix on the mode $k$, where $x_{ij}$ denotes the value of $j$-th feature for the $i$-th tensor entity, $p_k\leq d_k$.
-We consider a multilinear structure on the conditional expectation of the tensor.
-Specifically,
+Suppose the side information is available on each of the $K$ modes.
+Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the feature matrix on the mode $k\in[K]$, where $x_{ij}$ denotes the $j$-th feature value for the $i$-th tensor entity, for $(i,p)\in[d_k]\times[p_k]$, $p_k\leq d_k$.
+We model the relationship between the data tensor and feature matrices using an exponential family, multilinear mean model.
+Specifically, given features $\mX_k$, the entries of tensor $\tY$ are assumed to be independently distributed from an exponential family with conditional mean
 \begin{align}\label{eq:tensormodel}
-&\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)=f(\Theta),\ \text{with}\\
-&\Theta =\tB\times\{\mX_1,\ldots,\mX_K\} ,
+&\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)=f(\Theta),\quad \text{with}\quad \Theta =\tB\times\{\mX_1,\ \ldots,\ \mX_K\} ,
 \end{align}
-where  $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the parameter tensor of interest, $f(\cdot)$ applied to $\Theta$ entrywise is a known link function in generalized linear model, and $\times$ denotes the tensor Tucker product.
-The choice of link function depends on the distribution of the response data.
-Some common choices are identity link for Gaussian tensor, logistic link for binary tensor, and $\exp(\cdot)$ link for Poisson tensor (see Table~\ref{table:link}).
-
+where $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the unknown parameter tensor, $f(\cdot)$ is a known link function according to the data type of $\tY$, and $\times$ denotes the tensor Tucker product.
+The choice of link function depends on the assumed distribution family of tensor entries.
+Table~\ref{table:link} lists common choices of link functions, such as identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution.
+In general, dispersion parameters can also be included in the model.
+Because our main focus is the tensor decomposition under the mean model, we omit the dispersion parameter for notational simplicity.
 \begin{table}[htb]
@@
 
-We give three concrete examples of tensor regression that arise in practice. 
+In classical tensor decomposition, tensor factorization is performed on either data tensor $\tY$ or mean tensor $\mathbb{E}(\tY)$.
+In the context of supervised tensor decomposition, we propose to factorize the latent parameter tensor $\tB$,
+\begin{equation}\label{eq:factor}
+\tB=\tC\times\{\mM_1,\ \ldots,\ \mM_K\},
+\end{equation}
+where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, and $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices with orthonormal columns, for all $k\in[K]$.
+By the definition of multilinear rank, model~\eqref{eq:factor} implies that the latent tensor $\tB$ has multilinear-rank $\mr=(r_1,\ldots,r_K)\leq (p_1,\ldots,p_K)$.
+We assume the rank $\mr$ is known in our main analysis; the adaptation to unknown rank will be discussed in Section~\ref{sec:tuning}.
+
+Combining~\eqref{eq:tensormodel} and~\eqref{eq:factor} yields our final tensor decomposition model,
+\begin{equation}\label{eq:decomp}
+\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) = f\left(\tC\times\{\mX_1\mM_1,\ \ldots,\ \mX_K\mM_K\}\right),
+\end{equation}
+where the parameters of interest are $\mM_k$ and $\tC$.
+As illustrated in Figure~\ref{fig:intro1}b, the columns of $\mX_k\mM_k$ are linear combinations of features that explain the variation in the mean tensor.
+We will call $\mX_k\mM_k$ the ``supervised tensor factors'' and $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects of supervised tensor factors across multiple modes.
+Our goal is to find $\mM_k$ and $\tC$.
+Note that  $\mM_k$ and $\tC$ are identifiable only up to orthonormal transformations.
+
+\subsection{Three examples}
+We give three concrete examples of supervised tensor decomposition model~\eqref{eq:decomp} that arises in practice. 
 \begin{example}[Spatio-temporal growth model]
@@
 \begin{equation}\label{eq:time}
-\mathbb{E}(\tY|\mX_1,\mX_2,\mX_3)=\tB\times\{\mX_1,\mX_2,\mX_3\},
+\mathbb{E}(\tY|\mX_1,\mX_2,\mX_3)=\tC\times\{\mX_1\mM_1,\ \mX_2\mM_2,\ \mX_3\mM_3\},
 \end{equation}
-where $\tB\in\mathbb{R}^{p\times (r+1)\times (s+1)}$ is the coefficient tensor of interest, $\mX_1=\text{blockdiag}\{\mathbf{1}_p,\ldots,\mathbf{1}_p\}\in \{0,1\}^{d\times q}$ is the design matrix for lake types,
+where $\mX_1=\text{blockdiag}\{\mathbf{1}_p,\ldots,\mathbf{1}_p\}\in \{0,1\}^{d\times q}$ is the design matrix for lake types, and
 \[
@@
 1&\ell_{m}&\cdots&\ell^{r}_{m}
-\end{pmatrix},\
+\end{pmatrix},\quad
 \mX_3=
@@
 \]
-are the design matrices for spatial and temporal effects, respectively.
-The model~\eqref{eq:time} is a higher-order extension of the ``growth curve'' model originally proposed for matrix data~\citep{gabriel1998generalised,potthoff1964generalized,srivastava2008models}.
-Clearly, the spatial-temporal model is a special case of our tensor regression model, with covariates available on each of the three modes.
+are the design matrices for spatial and temporal effects, respectively, $\tC\in\mathbb{R}^{r_1\times r_2\times r_3}$ is the unknown core tensor, and $\mM_k$ are unknown dimension reduction matrices of coherent sizes.
+The model~\eqref{eq:time} is a higher-order extension of the bilinear ``growth curve'' model originally proposed for matrix data~\citep{gabriel1998generalised,potthoff1964generalized,srivastava2008models}.
+Clearly, the spatial-temporal model is a special case of our supervised tensor decomposition model, with features available on each of the three modes.
 \end{example}
@@
 \end{equation}
-where $\tB\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor of interest.
+where $\tB\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor. {\color{red}One often impose rank structure on $\tB$.... }
 
-The model~\eqref{eq:network} is a special case of our tensor-response model, with covariates on the last mode of the tensor. Specifically, stacking $\{\mY_i\}$ together
-yields an order-3 response tensor $\tY\in\{0,1\}^{d\times d\times n}$, along with covariate matrix $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$.
+The model~\eqref{eq:network} is a special case of our supervised tensor decomposition, with feature matrix on the last mode of the tensor.
+Specifically, we stack the network observations $\{\mY_i\}$ together and obtain an order-3 response tensor $\tY\in\{0,1\}^{d\times d\times n}$.
+Define a feature matrix $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$.
 Then, the model~\eqref{eq:network} can be written as
 \[
-\text{logit}(\mathbb{E}(\tY|\mX))=\tB\times_3 \mX=\tB\times\{\mI_d, \mI_d, \mX\}.
+\text{logit}(\mathbb{E}(\tY|\mX))=\tB\times\{\mM_1,\ \mM_2,\ \mX\},
 \]
+where we have implicitly $\mM_3=\mI_p$.
  \end{example}
@@
 \begin{equation}\label{eq:edge}
- \text{logit}(\mathbb{P}((i,j)\in E)=\mx^T_i\mB\mx_j=\langle \mB, \mx^T_i\mx_j\rangle.
+ \text{logit}\left\{\mathbb{P}\left[(i,j)\in E\right]\right\}=\mx^T_i\mB\mx_j=\langle \mB,\ \mx^T_i\mx_j\rangle.
  \end{equation}
@@
  \[
- \text{logit}(\mathbb{E}(\mY|\mX))=\tB\times\{\mX,\mX\}.
+ \text{logit}(\mathbb{E}(\mY|\mX))=\mC\times\{\mM\mX,\ \mM\mX\}, \quad \text{where}\quad \mC\in\mathbb{R}^{p\times p}.
  \]
@@
 \end{equation}
-where $r_k(\tB)\leq p_k$ is the Tucker rank at mode $k$ of the tensor.
+where $r_k(\tB)\leq p_k$ is the rank at mode $k$ of the tensor.
 The low-rank assumption is plausible in many scientific applications.
@@
 where the entries of $\tY$ are independent r.v.'s conditional on $\tX$, and $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$ is the low-rank coefficient tensor of interest.
-We comment that other forms of tensor low-rankness are also possible, and here we choose Tucker rank just for parsimony.
+
+%We comment that other forms of tensor low-rankness are also possible, and here we choose Tucker rank just for parsimony.
 Similar models can be derived using various notions of low-rankness based on CP decomposition~\citep{hitchcock1927expression} and train decomposition~\citep{oseledets2011tensor}.
 
-\section{Rank-constrained likelihood-based estimation}\label{sec:est}
+\subsection{A different interpretation as univariate regression in tensor project feature space}
+
+Multiple side information. ..
+
+Let $\tX=\mX_1\otimes \cdots \otimes \mX_K \in \mathbb{R}^{\prod_k d_k \times \prod_k p_k}$ be the tensor project of feature spaces.
+Denote $\text{Unfold}(\tX) \in\mathbb{R}^{d_{\text{total}}\times p_{\text{total}}}$, where $d_{\text{total}}=\prod_k d_k$ and $p_{\text{total}}=\prod_k p_k$.
+\[
+f(\mathbb{E}(y)}) =\beta_{111}\mx^1\my^1\mz^1+\beta_{121}\mx^1\my^2\mz^1+\beta_{112}\mx^1\my^1\mz^2+\cdots+\beta_{222}\mx^2\my^2\mz^2.
+\]
+where each $[\mx^1,\mx^2]=\mM_1\mX$, $[\my^1,\my^2]=\mM_2\mY$, $[\mz^1,\mz^2]=\mM_3\mZ$ are principal features. To single index model, when ....Full interaction model with rank-1 features.
+
+\section{Estimation}\label{sec:est}
+
+\subsection{Rank-constrained likelihood estimator}
 We develop a likelihood-based procedure to estimate the coefficient tensor $\tB$ in~\eqref{eq:tensormodel2}.
@@
 In the following theoretical analysis, we assume the rank $\mr=(r_1,\ldots,r_K)$ is known and fixed. The adaptation of unknown $\mr$ will be addressed in Section~\ref{sec:tuning}.
+
+\subsection{Alternating optimization} \label{subsec:alg}
+In this section, we introduce an efficient algorithm to solve~\eqref{eq:MLE}.
+The objective function $\tL_{\tY}(\tB)$ is concave in $\tB$ when the link $f$ is the canonical link function.
+We utilize a Tucker factor representation of the coefficient tensor $\tB$ and turn the optimization into a block-wise convex problem.
+
+Specifically, write the rank-$\mr$ decomposition of coefficient tensor $\tB$ as
+\begin{equation}\label{eq:tucker}
+\tB=\tC\times \{\mM_1,\ldots,\mM_K\},
+\end{equation}
+where $\tC\in\mathbb{R}^{r_1\times\cdots\times r_K}$ is a full-rank core tensor, $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices with orthogonal columns.
+Estimating $\tB$ amounts to finding both the core tensor $\tC$ and the factor matrices $\mM_k$'s.
+The optimization~\eqref{eq:MLE} can be written as $(\hat \tC, \{\hat \mM_k\})=\arg\max \tL_{\tY}(\tC, \mM_1,\ldots,\mM_K)$, where
+\begin{align}
+\tL_{\tY}(\tC, \mM_1,\ldots,\mM_K )&=\langle \tY, \Theta\rangle -\sum_{i_1,\ldots,i_K}b(\theta_{i_1,\ldots,i_K}),\\
+\text{with } \Theta&=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
+\end{align}
+The decision variables in the above objective function consist of $K+1$ blocks of variables, one for the core tensor $\tC$ and $K$ for the factor matrices $\mM_k$'s.
+We notice that, if any $K$ out of the $K+1$ blocks of variables are known, then the optimization with respect to the last block of variables reduced to a simple GLM.
+This observation suggests that we can iteratively update one block at a time while keeping others fixed.
+After each iteration, we rescale the core tensor $\tC^{(t+1)}$ subject to the maximum norm constraint.
+This post-processing in principle may not guarantee the monotonic increase of the objective, but we found that in our experiment this simple step appears robust for obtaining a desirable solution.
+The full algorithm is described in Algorithm~\ref{alg:B}.
+
+\begin{algorithm}[t]
+\caption{Generalized tensor response regression with covariates on multiple modes}\label{alg:B}
+\begin{algorithmic}[1]
+\INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, covariate matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,\ldots,K$, target Tucker rank $\mr=(r_1,\ldots,r_K)$, link function $f$, maximum norm bound $\alpha$
+\OUTPUT Low-rank estimation for the coefficient tensor $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$. 
+\State Calculate $\check \tB=\tY\times_1 \left[(\mX_1^T\mX_1)^{-1}\mX^T_1\right] \times_2\cdots\times_K\left[(\mX_K^T\mX_K)^{-1}\mX^T_K\right] $.
+\State Initialize the iteration index $t=0$.
+Initialize the core tensor $\tC^{(0)}$ and factor matrices $\mM^{(0)}_k\in\mathbb{R}^{p_k\times r_k}$ via rank-$\mr$ Tucker approximation of $\check\tB$, in the least-square sense.
+  \While {the relative increase in objective function $\tL_\tY(\tB)$ is less than the tolerance}
+\State Update iteration index $t \leftarrow t+1$.
+\For { $k=1$ to $K$}
+\State Obtain the factor matrix $\tilde \mM_k^{(t+1)}\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$. 
+\State Perform QR factorization on $\tilde \mM_k^{(t+1)}=\mQ\mR$, where $\mQ\in\mathbb{R}^{p_k\times r_k}$ has orthogonal columns. 
+\State Update $\mM_k^{(t+1)}\leftarrow \mQ$ and core tensor $\tC^{(t+1)}\leftarrow \tC^{(t+1)}\times_k \mR$.
+\EndFor
+\State Obtain the core tensor $\tC^{(t+1)}\in\mathbb{R}^{r_1\times\cdots\times r_K}$ by solving a GLM with $\text{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM^{(t)}_k]$ as covariates, and $f$ as link function.
+Here $\otimes$ denotes the kronecker product of matrices.
+\State Rescale the core tensor subject to the maximum norm constraint. 
+\State Update $\tB^{(t+1)}\leftarrow \tC^{(t+1)}\times_1\mM_1^{(t+1)}\times_2\cdots\times_K\mM_K^{(t+1)}$.
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
+
+Note that the feasible set $\tP$ is a non-convex set.
+Therefore, the optimization~\eqref{eq:MLE} is a non-convex problem, and the Algorithm~\ref{alg:B} usually not theoretically possesses the global optimality.
+However, as mentioned in Section~\ref{subsec:statprob}, the desired convergence rate~\eqref{eq:bound} holds for the valid estimators satisfying $\tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true})$, which indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective of local optimums are large enough.
+Fortunately, we find the  Algorithm~\ref{alg:B}  often gives a satisfactory convergence point $\hat \tB$, when we initialize the parameters via continuous-valued Tucker decomposition.
+Figure~\ref{fig:loglike} shows the trajectory of the objective function for order-3 tensors under the balanced setting, where $\alpha = 10, d_k = d, p_k = 0.4d, r_k = r$ for $ k = 1,2,3, p \in \{25,30\}$  and $r \in \{3, 6\}$.
+We consider the inputs with Gaussian, Bernoulli, and Poisson entries.
+Under all combinations of the dimension $d$, rank $r$, and type of the entries, Algorithm~\ref{alg:B} converges quickly in a few iterations, and the objective value at convergent points are close to or larger than the value at true parameters.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=15cm]{new_log.pdf}
+\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c) Poisson models.
+The dashed line represents the objective value at true parameter $\tL_{\tY}(\tB_{true})$.
+A micro-iteration refers to the update for one of the $K$ blocks.
+Four micro-iterations consist of a complete ``while'' iteration in Algorithm~\ref{alg:B}.}\label{fig:loglike}
+\end{figure}
+
+\subsection{Rank selection and time complexity}\label{sec:tuning}
+Algorithm~\ref{alg:B} takes the rank $\mr$ as an input.
+Estimating an appropriate rank given the data is of practical importance.
+We propose to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
+\begin{align}\label{eq:BIC}
+\hat \mr&=\argmin_{\mr=(r_1,\ldots,r_K)} \text{BIC}(\mr)\\
+&=\argmin_{\mr=(r_1,\ldots,r_K)}\left[-2\tL_{\tY}(\hat \tB)+p_e(\mr)\log \left(\prod_k d_k\right) \right],
+\end{align}
+where $p_e(\mr)\stackrel{\text{def}}{=}\sum_k (p_k-r_k-1)r_k+\prod_k r_k$ is the effective number of parameters in the model.
+We choose $\hat \mr$ that minimizes $\text{BIC}(\mr)$ via grid search.
+Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model.
+We test its empirical performance in Section~\ref{sec:simulation}.
+
+The computational complexity of our tensor regression model is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the response tensor.
+More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor.
+The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$-covariate matrix.
+Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$.
 
@@
 \end{thm}
-
-\section{Numerical implementation}\label{sec:algorithm}
-\subsection{Alternating optimization} \label{subsec:alg}
-In this section, we introduce an efficient algorithm to solve~\eqref{eq:MLE}.
-The objective function $\tL_{\tY}(\tB)$ is concave in $\tB$ when the link $f$ is the canonical link function.
-We utilize a Tucker factor representation of the coefficient tensor $\tB$ and turn the optimization into a block-wise convex problem.
-
-Specifically, write the rank-$\mr$ decomposition of coefficient tensor $\tB$ as
-\begin{equation}\label{eq:tucker}
-\tB=\tC\times \{\mM_1,\ldots,\mM_K\},
-\end{equation}
-where $\tC\in\mathbb{R}^{r_1\times\cdots\times r_K}$ is a full-rank core tensor, $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices with orthogonal columns.
-Estimating $\tB$ amounts to finding both the core tensor $\tC$ and the factor matrices $\mM_k$'s.
-The optimization~\eqref{eq:MLE} can be written as $(\hat \tC, \{\hat \mM_k\})=\arg\max \tL_{\tY}(\tC, \mM_1,\ldots,\mM_K)$, where
-\begin{align}
-\tL_{\tY}(\tC, \mM_1,\ldots,\mM_K )&=\langle \tY, \Theta\rangle -\sum_{i_1,\ldots,i_K}b(\theta_{i_1,\ldots,i_K}),\\
-\text{with } \Theta&=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
-\end{align}
-The decision variables in the above objective function consist of $K+1$ blocks of variables, one for the core tensor $\tC$ and $K$ for the factor matrices $\mM_k$'s.
-We notice that, if any $K$ out of the $K+1$ blocks of variables are known, then the optimization with respect to the last block of variables reduced to a simple GLM.
-This observation suggests that we can iteratively update one block at a time while keeping others fixed.
-After each iteration, we rescale the core tensor $\tC^{(t+1)}$ subject to the maximum norm constraint.
-This post-processing in principle may not guarantee the monotonic increase of the objective, but we found that in our experiment this simple step appears robust for obtaining a desirable solution.
-The full algorithm is described in Algorithm~\ref{alg:B}.
-
-\begin{algorithm}[t]
-\caption{Generalized tensor response regression with covariates on multiple modes}\label{alg:B}
-\begin{algorithmic}[1]
-\INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, covariate matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,\ldots,K$, target Tucker rank $\mr=(r_1,\ldots,r_K)$, link function $f$, maximum norm bound $\alpha$
-\OUTPUT Low-rank estimation for the coefficient tensor $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$. 
-\State Calculate $\check \tB=\tY\times_1 \left[(\mX_1^T\mX_1)^{-1}\mX^T_1\right] \times_2\cdots\times_K\left[(\mX_K^T\mX_K)^{-1}\mX^T_K\right] $.
-\State Initialize the iteration index $t=0$.
-Initialize the core tensor $\tC^{(0)}$ and factor matrices $\mM^{(0)}_k\in\mathbb{R}^{p_k\times r_k}$ via rank-$\mr$ Tucker approximation of $\check\tB$, in the least-square sense.
-  \While {the relative increase in objective function $\tL_\tY(\tB)$ is less than the tolerance}
-\State Update iteration index $t \leftarrow t+1$.
-\For { $k=1$ to $K$}
-\State Obtain the factor matrix $\tilde \mM_k^{(t+1)}\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$. 
-\State Perform QR factorization on $\tilde \mM_k^{(t+1)}=\mQ\mR$, where $\mQ\in\mathbb{R}^{p_k\times r_k}$ has orthogonal columns. 
-\State Update $\mM_k^{(t+1)}\leftarrow \mQ$ and core tensor $\tC^{(t+1)}\leftarrow \tC^{(t+1)}\times_k \mR$.
-\EndFor
-\State Obtain the core tensor $\tC^{(t+1)}\in\mathbb{R}^{r_1\times\cdots\times r_K}$ by solving a GLM with $\text{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM^{(t)}_k]$ as covariates, and $f$ as link function.
-Here $\otimes$ denotes the kronecker product of matrices.
-\State Rescale the core tensor subject to the maximum norm constraint. 
-\State Update $\tB^{(t+1)}\leftarrow \tC^{(t+1)}\times_1\mM_1^{(t+1)}\times_2\cdots\times_K\mM_K^{(t+1)}$.
-\EndWhile
-\end{algorithmic}
-\end{algorithm}
-
-Note that the feasible set $\tP$ is a non-convex set.
-Therefore, the optimization~\eqref{eq:MLE} is a non-convex problem, and the Algorithm~\ref{alg:B} usually not theoretically possesses the global optimality.
-However, as mentioned in Section~\ref{subsec:statprob}, the desired convergence rate~\eqref{eq:bound} holds for the valid estimators satisfying $\tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true})$, which indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective of local optimums are large enough.
-Fortunately, we find the  Algorithm~\ref{alg:B}  often gives a satisfactory convergence point $\hat \tB$, when we initialize the parameters via continuous-valued Tucker decomposition.
-Figure~\ref{fig:loglike} shows the trajectory of the objective function for order-3 tensors under the balanced setting, where $\alpha = 10, d_k = d, p_k = 0.4d, r_k = r$ for $ k = 1,2,3, p \in \{25,30\}$  and $r \in \{3, 6\}$.
-We consider the inputs with Gaussian, Bernoulli, and Poisson entries.
-Under all combinations of the dimension $d$, rank $r$, and type of the entries, Algorithm~\ref{alg:B} converges quickly in a few iterations, and the objective value at convergent points are close to or larger than the value at true parameters.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=15cm]{new_log.pdf}
-\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c) Poisson models.
-The dashed line represents the objective value at true parameter $\tL_{\tY}(\tB_{true})$.
-A micro-iteration refers to the update for one of the $K$ blocks.
-Four micro-iterations consist of a complete ``while'' iteration in Algorithm~\ref{alg:B}.}\label{fig:loglike}
-\end{figure}
-
-\subsection{Rank selection}\label{sec:tuning}
-Algorithm~\ref{alg:B} takes the rank $\mr$ as an input.
-Estimating an appropriate rank given the data is of practical importance.
-We propose to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
-\begin{align}\label{eq:BIC}
-\hat \mr&=\argmin_{\mr=(r_1,\ldots,r_K)} \text{BIC}(\mr)\\
-&=\argmin_{\mr=(r_1,\ldots,r_K)}\left[-2\tL_{\tY}(\hat \tB)+p_e(\mr)\log \left(\prod_k d_k\right) \right],
-\end{align}
-where $p_e(\mr)\stackrel{\text{def}}{=}\sum_k (p_k-r_k-1)r_k+\prod_k r_k$ is the effective number of parameters in the model.
-We choose $\hat \mr$ that minimizes $\text{BIC}(\mr)$ via grid search.
-Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model.
-We test its empirical performance in Section~\ref{sec:simulation}.
-
-\subsection{Time complexity}
-The computational complexity of our tensor regression model is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the response tensor.
-More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor.
-The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$-covariate matrix.
-Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$.
-
-\subsection{Missing values}
-.....
 
@@
 
+Our method leverages on the alternating updating algorithm for the optimization.
+Although a non-convex optimization procedure such as Algorithm~\ref{alg:binary} has no guarantee on global optimality, our numerical experiments have suggested that, upon random initializations, the convergence point $\hat\Theta$ is often satisfactory, in that the corresponding objective value $\tL_{\tY}(\hat\Theta)$ is close to the objective value $\tL_{\tY}(\trueT)$.
+When starting from random initializations, there could be multiple estimates, whose objective values are all greater than $\tL_{\tY}(\trueT)$.
+In theory, any of those choices perform equally well in estimating $\trueT$.
+In this sense, local optimality is not necessarily a severe concern in our context.
+On the other hand, characterizing global optimality for non-convex optimization problem of this type is itself of great interest.
+There has been recent progress investigating the landscape of non-convex optimization involving tensors~\citep{anandkumar2014tensor, richard2014statistical, ge2017optimization}.
+The problem is challenging, as the geometry can  depend on multiple factors including the tensor rank, dimension, and factorization form.
+In some special cases such as rank-1 or orthogonally decomposable tensors, one may further obtain the required asymptotical number of initializations, however, at the cost of more stringent assumptions on the target tensor~\citep{anandkumar2014tensor, richard2014statistical}.
+We leave the pursuit of optimization landscape as future research.
+
+Another an..
+is the (conditional) independence assumption on the tensor entries.
+Instead of glm, we may introduce a more general tensor-variate distribution on the $\tY$.
+For the case of normal distribution.
+For random effects.....
+
+Inference. ... uncertainly quantification... bootstrap ....
+
  \newpage

### 2020-09-06 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
   \begin{center}
-    {\LARGE\bf General Tensor Regression with Covariates on Multiple Modes}
+    {\LARGE\bf Supervised Tensor Decomposition with Multiple Side Information}
 \end{center}
@@
 Identifying the relationship between a high-dimensional tensor and side information is important yet challenging.
-Here, we develop a tensor decomposition method that incorporates multiple side information as structured covariates, and present case studies to show the benefit of proposed approach.
-Unlike unsupervised tensor decomposition, our supervised decomposition captures the most effective dimension reduction of the data tensor confined to the feature space on each mode.
+Here, we develop a tensor decomposition method that incorporates multiple side information as structured features, and present case studies to show the benefit of proposed approach.
+Unlike unsupervised tensor decomposition, our supervised decomposition captures the effective dimension reduction of the data tensor confined to the feature space on each mode.
 An efficient alternating updating algorithm is further developed.
@@
 We apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
-We identify the key global connectivity pattern and pinpoints the local regions that are associated with covariates.
+We identify the key global connectivity pattern and pinpoints the local regions that are associated with available features.
 Our method will help the practitioners efficiently analyze tensor datasets in various areas.
@@
 A number of supervised tensor methods have been proposed \citep{yu2016learning, raskutti2015convex, zhao2012higher} to address the tensor regression problem in various forms (e.g.\ scalar-to-tensor regression, tensor-response regression).
-These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the covariate matrices, both of which are less suitable for applications of our interest.
+These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest.
 The gap in theory and practice suggests a huge opportunity for improved modeling paradigms that better capture the complexity in tensor data with side information.
@@
 We present a general model and associated method for decomposing exponential family tensors with multiple side information. 
-We formulate the learning task as a structured regression problem, with tensor observation serving as the response, and the side information forming multiple covariate matrices.
+We formulate the learning task as a structured regression problem, with tensor observation serving as the response, and the side information forming multiple feature matrices.
 Figure~\ref{fig:intro1}b illustrates our model in the special case of order-3 tensors.
-A low-rank structure is imposed to the mean tensor $\mathbb{E}(\tY)\in\mathbb{R}^{d_1\times d_2\times d_3}$, where unlike classical decomposition, the tensor factors are assumed to belong to the space spanned by known features $\mX_k\in\mathbb{R}^{p_k\times d_k}$ for $k=1,2,3$.
-The unknown matrix (referred to as ``dimension reduction matrix'') $\mM_k\in\mathbb{R}^{r_k\times p_k}$ link the tensor factor to the feature space, thus allowing the identification the variation in the tensor entries attributable to the side information.
+A low-rank structure is imposed to the mean tensor $\mathbb{E}(\tY)\in\mathbb{R}^{d_1\times d_2\times d_3}$, where unlike classical decomposition, the tensor factors are assumed to belong to the space spanned by features $\mX_k\in\mathbb{R}^{p_k\times d_k}$ for $k=1,2,3$.
+The unknown matrix of interest $\mM_k\in\mathbb{R}^{r_k\times p_k}$ (referred to as ``dimension reduction matrix'') link the mean tensor to the feature space, thus allowing the identification of variations across tensor entries that can be explained by side information.
 
@@
 The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor,zhang2018tensor, hong2020generalized} that aims to find the best low-rank representation of data tensor.
-In contrast, our model can be viewed a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and covariates.
+In contrast, our model can be viewed a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and features.
 The low-rank factorization is determined jointly by the tensor data and feature matrices on multiple sides.
@@
 We advocate the non-convex approach and provide strong evidence for its success in our setting.
-Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2015convex,gahrooei2020multiple}, and we go step further by finding the supervised factors $\mM_k\mX_k$ in order to pinpoint important features for prediction (see Figure~\ref{fig:intro1}b).
+Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2015convex,gahrooei2020multiple}, and we go step further by finding the sufficient dimension reduction~\citep{adragni2009sufficient}, $\text{Span}(\mM_k)$, in order to pinpoint effective features in prediction (see Figure~\ref{fig:intro1}b).
 The latter approach greatly improves the \emph{interpretability} in prediction.
@@
 We let $\mI_d$ denote the $d \times d$ identity matrix and $[d]$ denote the $d$-set $\{1,\ldots,d\}$.
-For ease of notation, we allow the basic arithmetic operators ($+$, $-$, $\geq $) and functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner.
+For ease of notation, we allow the basic arithmetic operators ($+$, $-$, $\geq $) and univariate
+functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner.
 
@@
 We model the relationship between the data tensor and feature matrices using an exponential family, multilinear mean model.
-Specifically, given features $\mX_k$, the entries of tensor $\tY$ are assumed to be independently distributed from an exponential family with conditional mean
+Specifically, assumed that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent random variables from an exponential family, and the conditional mean tensor has the form
 \begin{align}\label{eq:tensormodel}
@@
 where $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the unknown parameter tensor, $f(\cdot)$ is a known link function according to the data type of $\tY$, and $\times$ denotes the tensor Tucker product.
-The choice of link function depends on the assumed distribution family of tensor entries.
-Table~\ref{table:link} lists common choices of link functions, such as identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution.
+The choice of link function depends on the specific distribution family of tensor entries.
+Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution.
 In general, dispersion parameters can also be included in the model.
 Because our main focus is the tensor decomposition under the mean model, we omit the dispersion parameter for notational simplicity.
-\begin{table}[htb]
-\centering
-\begin{tabular}{c|ccc}
-Data type &Gaussian & Poisson& Bernoulli\\
-\hline
-Domain $\mathbb{Y}$& $\mathbb{R}$&$\mathbb{N}$&$\{0,1\}$\\
- $b(\theta)$&$\theta^2/2$& $\exp(\theta)$&$\log (1+\exp(\theta))$\\
- link $f(\theta)$&$\theta$&$\exp(\theta)$&$(1+\exp(-\theta))^{-1}$
-\end{tabular}
-\caption{Canonical links for common distributions.}\label{table:link}
-\end{table}
 
@@
 By the definition of multilinear rank, model~\eqref{eq:factor} implies that the latent tensor $\tB$ has multilinear-rank $\mr=(r_1,\ldots,r_K)\leq (p_1,\ldots,p_K)$.
-We assume the rank $\mr$ is known in our main analysis; the adaptation to unknown rank will be discussed in Section~\ref{sec:tuning}.
-
 Combining~\eqref{eq:tensormodel} and~\eqref{eq:factor} yields our final tensor decomposition model,
-\begin{equation}\label{eq:decomp}
-\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) = f\left(\tC\times\{\mX_1\mM_1,\ \ldots,\ \mX_K\mM_K\}\right),
-\end{equation}
+\begin{align}\label{eq:decomp}
+\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1,\ \ldots,\ \mX_K\mM_K\}\right),\\
+\text{where} \ &\ \mM_k\mM^T_k=\mI_{p_k},\quad \text{for all }k=1,\ldots,K
+\end{align}
 where the parameters of interest are $\mM_k$ and $\tC$.
-As illustrated in Figure~\ref{fig:intro1}b, the columns of $\mX_k\mM_k$ are linear combinations of features that explain the variation in the mean tensor.
-We will call $\mX_k\mM_k$ the ``supervised tensor factors'' and $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects of supervised tensor factors across multiple modes.
-Our goal is to find $\mM_k$ and $\tC$.
+As illustrated in Figure~\ref{fig:intro1}b, the columns of $\mX_k\mM_k$ are linear combinations of features that are relevant in the mean tensor.
+We call $\mX_k\mM_k$ the ``supervised tensor factors'' or ``sufficient features''~\citep{adragni2009sufficient}, and call $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects between sufficient features across multiple modes.
+Our goal is to find $\mM_k$ and corresponding $\tC$.
 Note that $\mM_k$ and $\tC$ are identifiable only up to orthonormal transformations.
@@
 Let $\{\ell_j\}_{j\in[m]}$ denote the sampled depth levels and $\{t_k\}_{k\in[n]}$ the time points.
-Assume that the expected pH trend in depth is a polynomial of order $r$ and that the expected trend in time is a polynomial of order $s$.
-Then, the spatio-temporal growth model can be represented as
+Assume that the expected pH trend in depth is a polynomial of order at most $r$ and that the expected trend in time is a polynomial of order $s$.
+Then, the conditional mean model for the spatio-temporal growth model can be represented as
 \begin{equation}\label{eq:time}
@@
 The model~\eqref{eq:time} is a higher-order extension of the bilinear ``growth curve'' model originally proposed for matrix data~\citep{gabriel1998generalised,potthoff1964generalized,srivastava2008models}.
+The factors $\mX_k\mM_k$ are sufficient features for the mean model~\eqref{eq:time}.
 Clearly, the spatial-temporal model is a special case of our supervised tensor decomposition model, with features available on each of the three modes.
@@
 Network response model is recently developed in the context of neuroimanig analysis.
-The goal is to study the relationship between network-valued response and the individual covariates.
-Suppose we observe $n$ i.i.d.\ observations $\{(\mY_i, \mx_i): i=1,\ldots,n\}$, where $\mY_i\in\{0,1\}^{d\times d}$ is the brain connectivity network on the $i$-th individual, and $\mx_i\in\mathbb{R}^p$ is the individual covariate such as age, gender, cognition, etc.
-The network-response model~\citep{rabusseau2016low, zhang2018network} has the form
+The goal is to study the relationship between brain network connectivity pattern and the individual features.
+Suppose we have a sample of $n$ observations, $\{(\mY_i, \mx_i)\colon i=1,\ldots,n\}$, where for each individual $i\in[n]$, $\mY_i\in\{0,1\}^{d\times d}$ is the symmetric adjacency matrix whose entries indicate presences/absences of connectivities between $d$ brain nodes, and $\mx_i\in\mathbb{R}^p$ is the individual's feature such as age, gender, cognition score, etc.
+The network-response model~\citep{rabusseau2016low, zhang2018network} has the conditional mean
 \begin{equation}\label{eq:network}
@@
 \end{equation}
-where $\tB\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor. {\color{red}One often impose rank structure on $\tB$.... }
+where $\tB\in \mathbb{R}^{d\times d\times p}$ is a coefficient tensor of rank $r(\tB)=(r,r,r')$.
 
@@
 \[
-\text{logit}(\mathbb{E}(\tY|\mX))=\tB\times\{\mM_1,\ \mM_2,\ \mX\},
+\text{logit}(\mathbb{E}(\tY|\mX))=\tC\times\{\mM,\ \mM,\ \mX\mM'\},
 \]
-where we have implicitly $\mM_3=\mI_p$.
+where $\tC\in\mathbb{R}^{r\times r\times r'}$ is the core tensor, $\mM\in\mathbb{R}^{d\times r}$ is the dimension reduction matrix at the first two modes, and $\mM'\in\mathbb{R}^{p\times r'}$ is for the last mode.
 \end{example}
@@
  \begin{example}[Dyadic data with node attributes] Dyadic dataset consists of measurements on pairs of objects or under a pair of conditions.
-Common examples include networks and graphs.
-Let $\tG=(V,E)$ denote a network, where $V=[d]$ is the node set of the graph, and $E\subset V\times V$ is the edge set.
-Suppose that we also observe covariate $\mx_i\in\mathbb{R}^p$ associated to each $i\in V$.
+Common examples include graphs and networks.
+Let $\tG=(V,E)$ denote a graph, where $V=[d]$ is the node set of the graph, and $E\subset V\times V$ is the edge set.
+Suppose that we also observe feature vector $\mx_i\in\mathbb{R}^p$ associated to each $i\in V$.
 A probabilistic model on the graph $\tG=(V,E)$ can be described by the following matrix regression.
@@
 \begin{equation}\label{eq:edge}
- \text{logit}\left\{\mathbb{P}\left[(i,j)\in E\right]\right\}=\mx^T_i\mB\mx_j=\langle \mB,\ \mx^T_i\mx_j\rangle.
+ \text{logit}\left(\mathbb{P}\left((i,j)\in E\right)\right)=\mx^T_i\mB\mx_j=\langle \mB,\ \mx^T_i\mx_j\rangle,
  \end{equation}
-The above model has demonstrated its success in modeling transitivity, balance, and communities in the networks~\citep{hoff2005bilinear}.
+ where $\mB\in\mathbb{R}^{p\times p}$ is a symmetric, low rank matrix.
+The low-rankness in $\mB$ has demonstrated its success in modeling transitivity, balance, and communities in the networks~\citep{hoff2005bilinear}.
 We show that our tensor regression model~\eqref{eq:tensormodel} also incorporates the graph model as a special case.
@@
  \[
- \text{logit}(\mathbb{E}(\mY|\mX))=\mC\times\{\mM\mX,\ \mM\mX\}, \quad \text{where}\quad \mC\in\mathbb{R}^{p\times p}.
+ \text{logit}(\mathbb{E}(\mY|\mX))=\mC\times\{\mX\mM,\ \mX\mM\}, 
   \]
+  where $\mC\in\mathbb{R}^{r\times r}$ and $\mM\in\mathbb{R}^{p\times r}$ can be taken as the singular value decomposition of $\mB=\mM\mC\mM^T$. 
   \end{example}
 
-In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by the covariates.
-The regression coefficient $\tB$ in our model model~\eqref{eq:tensormodel} serves this goal by collecting the effects of covariates and the interaction thereof.
+{\color{red}revise from here....}
+In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by the features.
+The regression coefficient $\tB$ in our model model~\eqref{eq:tensormodel} serves this goal by collecting the feature effects and the interaction thereof.
 To encourage the sharing among effects, we assume that the coefficient tensor $\tB$ lies in a low-dimensional parameter space:
-\begin{equation}\label{eq:rank}
-\small \tP_{r_1,\ldots,r_K}=\{\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}: r_k(\tB)\leq r_k \text{ for all } k\in[K]\},
-\end{equation}
-where $r_k(\tB)\leq p_k$ is the rank at mode $k$ of the tensor.
+%\begin{equation}\label{eq:rank}
+%\small \tP_{r_1,\ldots,r_K}=\{\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}: r_k(\tB)\leq r_k \text{ for all } k\in[K]\},
+%\end{equation}
+%where $r_k(\tB)\leq p_k$ is the rank at mode $k$ of the tensor. 
 The low-rank assumption is plausible in many scientific applications.
@@
 The low-rank structure encourages the shared information across tensor entries, thereby greatly improving the estimation stability.
-When no confusion arises, we drop the subscript $(r_1,\ldots,r_K)$ and write $\tP$ for simplicity.
+%When no confusion arises, we drop the subscript $(r_1,\ldots,r_K)$ and write $\tP$ for simplicity.
 
-Our tensor regression model is able to incorporate covariates on any subset of modes, whenever available.
-Without loss of generality, we denote by $\tX=\{\mX_1,\ldots,\mX_K\}$ the covariates in all modes and treat $\mX_k=\mI_{d_k}$ if the mode-$k$ has no (informative) covariate.
-Then, the final form of our tensor regression model can be written as:
-\begin{align}\label{eq:tensormodel2}
-&\mathbb{E}(\tY|\tX)=f(\Theta),\quad \Theta =\tB\times\{\mX_1,\ldots,\mX_K\},\notag\\
-&\text{where}\ \text{rank}(\tB)\leq (r_1,\ldots,r_K),
-\end{align}
-where the entries of $\tY$ are independent r.v.'s conditional on $\tX$, and $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$ is the low-rank coefficient tensor of interest.
-
+Our tensor regression model is able to incorporate an arbitrary number of feature matrices along the tensor modes.
+When the side information is unavailable for certain mode $k$, we set $\mX_k=\mI_{d_k}$.
+In particular, our model~\eqref{eq:decomp} reduces to classical unsupervised tensor decomposition model~\citep{de2000multilinear,hong2018generalized} when no side information is available, i.e., $\mX_k=\mI_{d_k}$ for all $k\in[K]$.
+%Then, the final form of our tensor regression model can be written as:
+%\begin{align}\label{eq:tensormodel2}
+%&\mathbb{E}(\tY|\tX)=f(\Theta),\quad \Theta =\tC\times\{\mX_1\mM_k,\ldots,\mX_k\mK_k\},\notag\\
+%&\text{where}\ \mM^T_k\mM_k=1
+%\end{align}
+%where the entries of $\tY$ are independent random variables conditional on $\tX$, and $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$ is the low-rank coefficient tensor of interest. 
 %We comment that other forms of tensor low-rankness are also possible, and here we choose Tucker rank just for parsimony.
@@
 
-\subsection{A different interpretation as univariate regression in tensor project feature space}
+\subsection{Connection to sufficient dimension reduction and tensor-on-tensor regression}
+Joint dimension reduction of tensor and features....
 
-Multiple side information. ..
-
-Let $\tX=\mX_1\otimes \cdots \otimes \mX_K \in \mathbb{R}^{\prod_k d_k \times \prod_k p_k}$ be the tensor project of feature spaces.
-Denote $\text{Unfold}(\tX) \in\mathbb{R}^{d_{\text{total}}\times p_{\text{total}}}$, where $d_{\text{total}}=\prod_k d_k$ and $p_{\text{total}}=\prod_k p_k$.
+Denote $d_{\text{total}}=\prod_k d_k$ and $p_{\text{total}}=\prod_k p_k$.
+Let $\tX=\mX_1\otimes \cdots \otimes \mX_K \in \mathbb{R}^{d_{\text{total}} \times p_{\text{total}}}$ be the tensor product of feature spaces, and $\text{Unfold}(\tX) \in\mathbb{R}^{d_{\text{total}}\times p_{\text{total}}}$.
 \[
-f(\mathbb{E}(y)}) =\beta_{111}\mx^1\my^1\mz^1+\beta_{121}\mx^1\my^2\mz^1+\beta_{112}\mx^1\my^1\mz^2+\cdots+\beta_{222}\mx^2\my^2\mz^2.
+\mathbb{E}(\text{vec}(\tY)|\mX,\mZ,\mW) =\beta_{111}\mx^1\mz^1\mw^1+\beta_{121}\mx^1\mz^2\mw^1+\beta_{112}\mz^1\mz^1\mw^2+\cdots+\beta_{222}\mx^2\mz^2\mw^2.
 \]
-where each $[\mx^1,\mx^2]=\mM_1\mX$, $[\my^1,\my^2]=\mM_2\mY$, $[\mz^1,\mz^2]=\mM_3\mZ$ are principal features. To single index model, when ....Full interaction model with rank-1 features.
+where each $\mM_1\mX=[\mx^1,\mx^2]$, $\mM_2\mZ=[\mz^1,\mz^2]$, $\mM_3\mW=[\mw^1,\mw^2]$ are sufficient features. when ....Full interaction model with rank-1 features.
 
@@
 \subsection{Rank-constrained likelihood estimator}
-We develop a likelihood-based procedure to estimate the coefficient tensor $\tB$ in~\eqref{eq:tensormodel2}.
+We develop a likelihood-based procedure to estimate $\tC$ and $\mM_k$ in~\eqref{eq:decomp}.
 We adopt the exponential family as a flexible framework for different data types.
-In a classical generalized linear model (GLM) with a scalar response $y$ and covariate $\mx$, the density is expressed as:
+In a classical generalized linear model (GLM) with a scalar response $y$ and feature $\mx$, the density is expressed as:
 \[
@@
 
+\begin{table}[htb]
+\centering
+\begin{tabular}{c|ccc}
+Data type &Gaussian & Poisson& Bernoulli\\
+\hline
+Domain $\mathbb{Y}$& $\mathbb{R}$&$\mathbb{N}$&$\{0,1\}$\\
+ $b(\theta)$&$\theta^2/2$& $\exp(\theta)$&$\log (1+\exp(\theta))$\\
+ link $f(\theta)$&$\theta$&$\exp(\theta)$&$(1+\exp(-\theta))^{-1}$
+\end{tabular}
+\caption{Canonical links for common distributions.}\label{table:link}
+\end{table}
+
 In our context, we model the the entries in the response tensor $y_{ijk}$ conditional on $\theta_{ijk}$ as independent draws from an exponential family.
-The quasi log-likelihood of~\eqref{eq:tensormodel2} is equal (ignoring constant) to Bregman distance between $\tY$ and $b'(\Theta)$:
+The quasi log-likelihood of~\eqref{eq:decomp} is equal (ignoring constant) to Bregman distance between $\tY$ and $b'(\Theta)$:
 \begin{align}
-\tL_{\tY}(\tB)&=\langle \tY, \Theta \rangle - \sum_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K}),\\
-\text{where}\ \Theta&=\tB\times\{\mX_1,\ldots,\mX_K\}.
+\tL_{\tY}(\tC,\mM_k)&=\langle \tY, \Theta \rangle - \sum_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K}),\\
+\text{where}\ \Theta&=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
 \end{align}
-We assume that we have an additional information on an upper bound $\alpha>0$ such that $\mnormSize{}{\Theta}\leq \alpha$.
-We propose a constrained maximum likelihood estimator (MLE) for the coefficient tensor:
+We assume that we have an additional information on an upper bound $\alpha>0$ such that $\mnormSize{}{\Theta}\leq \alpha$. We propose a constrained maximum likelihood estimator (MLE)
+%\begin{align} \label{eq:MLE} 
+%	\hat \tB &=\argmax_{\hat \tB \colon \text{rank}(\tB)\leq \mr, \mnormSize{}{\Theta(\tB)}\leq \alpha  } \tL_{\tY}(\tB).
+%\end{align}
+
 \begin{align} \label{eq:MLE} 
-	\hat \tB &=\argmax_{\hat \tB \colon \text{rank}(\tB)\leq \mr, \mnormSize{}{\Theta(\tB)}\leq \alpha  } \tL_{\tY}(\tB).
+(\hat \tC, \hat \mM_1,\ldots,\hat \mM_k) &=\argmax_{(\tC,\mM_k)\in \tP_{r_1,\ldots,r_K}} \tL_{\tY}(\tC,\mM_k)\\
+\text{where } \tP_{r_1,\ldots,r_K}&=\left\{(\tC,\mM_k)\colon \mM_k\mM^T_k=\mathbf{1}_{p_k},\ \mnormSize{}{\Theta(\tC,\mM_k)}\leq \alpha, \right.\\
+&\quad \quad\quad \quad  \quad \quad\quad \quad \left.\mK_k\in\mathbb{R}^{p_k\times r_k},\ \tC\in\mathbb{R}^{r_1\times \cdots\times r_K}\right\}.
 \end{align}
+
 In the following theoretical analysis, we assume the rank $\mr=(r_1,\ldots,r_K)$ is known and fixed. The adaptation of unknown $\mr$ will be addressed in Section~\ref{sec:tuning}.
@@
 \subsection{Alternating optimization} \label{subsec:alg}
-In this section, we introduce an efficient algorithm to solve~\eqref{eq:MLE}.
-The objective function $\tL_{\tY}(\tB)$ is concave in $\tB$ when the link $f$ is the canonical link function.
+We propose an alternating optimization algorithm to solve~\eqref{eq:MLE}. 
+%The objective function $\tL_{\tY}(\tB)$ is concave in $\tB$ when the link $f$ is the canonical link function.
 We utilize a Tucker factor representation of the coefficient tensor $\tB$ and turn the optimization into a block-wise convex problem.
-
-Specifically, write the rank-$\mr$ decomposition of coefficient tensor $\tB$ as
-\begin{equation}\label{eq:tucker}
-\tB=\tC\times \{\mM_1,\ldots,\mM_K\},
-\end{equation}
-where $\tC\in\mathbb{R}^{r_1\times\cdots\times r_K}$ is a full-rank core tensor, $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices with orthogonal columns.
+%Specifically, write the rank-$\mr$ decomposition of coefficient tensor $\tB$ as
+%\begin{equation}\label{eq:tucker}
+%\tB=\tC\times \{\mM_1,\ldots,\mM_K\},
+%\end{equation}
+%where $\tC\in\mathbb{R}^{r_1\times\cdots\times r_K}$ is a full-rank core tensor, $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices with orthogonal columns.
 Estimating $\tB$ amounts to finding both the core tensor $\tC$ and the factor matrices $\mM_k$'s.
-The optimization~\eqref{eq:MLE} can be written as $(\hat \tC, \{\hat \mM_k\})=\arg\max \tL_{\tY}(\tC, \mM_1,\ldots,\mM_K)$, where
-\begin{align}
-\tL_{\tY}(\tC, \mM_1,\ldots,\mM_K )&=\langle \tY, \Theta\rangle -\sum_{i_1,\ldots,i_K}b(\theta_{i_1,\ldots,i_K}),\\
-\text{with } \Theta&=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
-\end{align}
-The decision variables in the above objective function consist of $K+1$ blocks of variables, one for the core tensor $\tC$ and $K$ for the factor matrices $\mM_k$'s.
+The decision variables in the objective function~\eqref{eq:MLE} consist of $K+1$ blocks of variables, one for the core tensor $\tC$ and $K$ for the factor matrices $\mM_k$'s.
 We notice that, if any $K$ out of the $K+1$ blocks of variables are known, then the optimization with respect to the last block of variables reduced to a simple GLM.
@@
 \begin{algorithm}[t]
-\caption{Generalized tensor response regression with covariates on multiple modes}\label{alg:B}
+\caption{Supervised Tensor Decomposition with Multiple Side Information}\label{alg:B}
 \begin{algorithmic}[1]
-\INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, covariate matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,\ldots,K$, target Tucker rank $\mr=(r_1,\ldots,r_K)$, link function $f$, maximum norm bound $\alpha$
+\INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, feature matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,\ldots,K$, target Tucker rank $\mr=(r_1,\ldots,r_K)$, link function $f$, maximum norm bound $\alpha$
 \OUTPUT Low-rank estimation for the coefficient tensor $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$. 
@@
 \EndFor
-\State Obtain the core tensor $\tC^{(t+1)}\in\mathbb{R}^{r_1\times\cdots\times r_K}$ by solving a GLM with $\text{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM^{(t)}_k]$ as covariates, and $f$ as link function.
+\State Obtain the core tensor $\tC^{(t+1)}\in\mathbb{R}^{r_1\times\cdots\times r_K}$ by solving a GLM with $\text{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM^{(t)}_k]$ as features, and $f$ as link function.
 Here $\otimes$ denotes the kronecker product of matrices.
@@
 More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor.
-The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$-covariate matrix.
+The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$ feature matrix.
 Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$.
@@
 
-In modern applications, the response tensor and covariates are often large-scale.
+In modern applications, the response tensor and features are often large-scale.
 We are particularly interested in the high-dimensional region in which both $d_k$ and $p_k$ diverge; i.e.\ $d_k\to \infty$ and $p_k\to\infty$, while ${p_k\over d_k} \to \gamma_k \in[0,1)$.
@@
 The assumptions are fairly mild.
-Assumption A1 guarantees the non-singularity of the covariates, and Assumption A2 ensures the log-likelihood $\tY(\Theta)$ is strictly concave in the linear predictor $\Theta$.
+Assumption A1 guarantees the non-singularity of the feature matrices, and Assumption A2 ensures the log-likelihood $\tY(\Theta)$ is strictly concave in the linear predictor $\Theta$.
 Assumption A2 and A2' are equivalent, because $\text{Var}(y_{i_1,\ldots,i_K}|\tX,\tB)=\phi b''(\theta_{i_1,\ldots,i_K})$ when $y_{i_1,\ldots,i_K}$ belongs to an exponential family~\citep{mccullagh1989generalized}.
@@
 \begin{thm}[Statistical convergence]\label{thm:main}
-Consider a generalized tensor regression model with covariates on multiple modes $\tX=\{\mX_1,\ldots,\mX_K\}$.
+Consider a generalized tensor regression model with features on multiple modes $\tX=\{\mX_1,\ldots,\mX_K\}$.
 Suppose the entries in $\tY$ are independent realizations of an exponential family distribution, and $\mathbb{E}(\tY|\tX)$ follows the low-rank tensor regression model~\eqref{eq:tensormodel2}.
@@
 
-To gain further insight on the bound~\eqref{eq:bound}, we consider a special case when tensor dimensions are equal at each of the modes, i.e., $d_k=d$, $p_k=\gamma d$, $\gamma\in [0,1)$ for all $k\in[K]$, and the covariates $\mX_k$ are Gaussian design matrices with i.i.d.\ $N(0,1)$ entries.
-To put the context in the framework of Theorem~\ref{thm:main}, we rescale the covariates into $\check \mX_k={1\over \sqrt{d}}\mX_k$ so that the singular values of $\check \mX_k$ are bounded by $1\pm \sqrt{\gamma}$.
-The result in~\eqref{eq:bound} implies that the estimated coefficient has a convergence rate $\tO({p\over d^K})$ in the scale of the original covariates $\{\mX_k\}$.
+To gain further insight on the bound~\eqref{eq:bound}, we consider a special case when tensor dimensions are equal at each of the modes, i.e., $d_k=d$, $p_k=\gamma d$, $\gamma\in [0,1)$ for all $k\in[K]$, and the feature matrices $\mX_k$ are Gaussian design with i.i.d.\ $N(0,1)$ entries.
+To put the context in the framework of Theorem~\ref{thm:main}, we rescale the feature matrices into $\check \mX_k={1\over \sqrt{d}}\mX_k$ so that the singular values of $\check \mX_k$ are bounded by $1\pm \sqrt{\gamma}$.
+The result in~\eqref{eq:bound} implies that the estimated coefficient has a convergence rate $\tO({p\over d^K})$ in the scale of the original feature $\{\mX_k\}$.
 Therefore, our estimation is consistent as the dimension grows, and the convergence becomes especially favorably as the order of tensor data increases.
@@
 As immediate applications, we obtain the convergence rate for the three examples mentioned in Section~\ref{sec:model}.
-Without loss of generality, we assume that the singular values of the $d_k$-by-$p_k$ covariate matrix $\mX_k$ are bounded by $\sqrt{d_k}$.
+Without loss of generality, we assume that the singular values of the $d_k$-by-$p_k$ feature matrix $\mX_k$ are bounded by $\sqrt{d_k}$.
 
@@
 
-\begin{example} [Network population model] The estimated node-by-node-by-covariate tensor converges at the rate $\tO\left({2d+p\over d^2n}\right)$ where $p\leq n$.
+\begin{example} [Network population model] The estimated node-by-node-by-feature tensor converges at the rate $\tO\left({2d+p\over d^2n}\right)$ where $p\leq n$.
 The estimation achieves consistency as the number of individuals or the number of nodes grows.
@@
 
-\begin{example} [Dyadic data with node attributes] The estimated covariate-by-covariate matrix converges at the rate $\tO\left({p\over d^2}\right)$ where $p\leq d$.
+\begin{example} [Dyadic data with node attributes] The estimated feature-by-feature matrix converges at the rate $\tO\left({p\over d^2}\right)$ where $p\leq d$.
 Again, our estimation achieves consistency as the number of nodes grows.
@@
 The coefficient tensor $\tB$ is generated using the factorization form~\eqref{eq:tucker} where both the core and factor matrices are drawn i.i.d.\ from Uniform[-1,1].
-The linear predictor is then simulated from $\tU=\tB\times\{\mX_1,\mX_2,\mX_3\}$, where $\mX_k$ is either an identity matrix (i.e.\ no covariate available) or Gaussian random matrix with i.i.d.\ entries from $N(0,\sigma_k^2)$.
+The linear predictor is then simulated from $\tU=\tB\times\{\mX_1,\mX_2,\mX_3\}$, where $\mX_k$ is either an identity matrix (i.e.\ no feature available) or Gaussian random matrix with i.i.d.\ entries from $N(0,\sigma_k^2)$.
 We set $\sigma_k=d_k^{-1/2}$ to ensure the singular values of $\mX_k$ are bounded as $d_k$ increases.
@@
 
-The second experiment evaluates the accuracy when covariates are available on all modes.
+The second experiment evaluates the accuracy when features are available on all modes.
 We set $\alpha=10, d_k=d, p_k=0.4d_k, r_k=r\in\{2,4,6\}$ and increase $d$ from 25 to 50.
@@
 Each network measures the connections between $d_1=d_2=20$ brain nodes.
-We simulate $p=5$ covariates for the each of the 50 individuals.
-These covariates may represent, for example, age, gender, cognitive score, etc.
+We simulate $p=5$ features for the each of the 50 individuals.
+These features may represent, for example, age, gender, cognitive score, etc.
 Recent study~\citep{robinson2015dynamic} has suggested that brain connectivity networks often exhibit community structure represented as a collection of subnetworks, and each subnetwork is comprised of a set of spatially distributed brain nodes.
@@
 Specifically, we partition the nodes into $r$ blocks by assigning each node to a block with uniform probability.
-Edges within a same block are assumed to share the same covariate effects, where the effects are drawn i.i.d.\ from $N(0,1)$.
+Edges within a same block are assumed to share the same feature effects, where the effects are drawn i.i.d.\ from $N(0,1)$.
 We then apply our tensor regression model to the network data using the BIC-selected rank.
@@
 Figure~\ref{fig:glm} compares the MSE of our method with a classical GLM approach.
-A classical GLM is to regress the dyadic edges, one at a time, on the covariates, and this model is repeatedly fitted for each edge.
+A classical GLM is to regress the dyadic edges, one at a time, on the features, and this model is repeatedly fitted for each edge.
 This repeated approach, however, does not account for the correlation among the edges, and may suffer from overfitting.
@@
 \begin{itemize}
-\item  Higher-order low-rank regression ({\bf HOLRR},~\citep{rabusseau2016low}) is a least-square based tensor regression that allows covariates on a single mode. 
-\item Higher-order partial least square ({\bf HOPLS},~\citep{zhao2012higher}) is a dimension-reduction method that jointly models a tensor response and a tensor covariate. 
+\item  Higher-order low-rank regression ({\bf HOLRR},~\citep{rabusseau2016low}) is a least-square based tensor regression that allows features on a single mode. 
+\item Higher-order partial least square ({\bf HOPLS},~\citep{zhao2012higher}) is a dimension-reduction method that jointly models a tensor response and a tensor feature. 
 \item Subsampled tensor projected gradient ({\bf TPG},~\citep{yu2016learning}) tackles the same question as {\bf HOLRR} but instead uses a different algorithm to solve the problem. 
 \end{itemize}
-These three methods are the closest algorithms to ours, in that they relate a tensor response to covariates using a low-rank structure.
+These three methods are the closest algorithms to ours, in that they relate a tensor response to features using a low-rank structure.
 All the three methods allow only Gaussian data, whereas ours is applicable to any exponential family distribution including Gaussian, Bernoulli, Multinomial, etc.
@@
 
-The comparison was assessed from three aspects: (a) benefit of incorporating covariates from multiple modes; (b) prediction error with respect to sample size; (c) sensitivity of accuracy with respect to model complexity.
+The comparison was assessed from three aspects: (a) benefit of incorporating features from multiple modes; (b) prediction error with respect to sample size; (c) sensitivity of accuracy with respect to model complexity.
 We use similar simulation setups as in our experiment II, but consider combinations of rank ($\mr=(3,3,3)$ vs.\ $(4,5,6)$), noise ($\sigma = 1/2$ vs.
-$1/4$), and dimension ($d$ ranging from 20 to 100 for modes with covariates, $d = 20$ for modes without covariates).
+$1/4$), and dimension ($d$ ranging from 20 to 100 for modes with features, $d = 20$ for modes without features).
 
@@
 As the number of informative modes (i.e.
-modes with available covariates) increases, the {\bf GTR} exhibits a reduction in error whereas others have increased errors.
-This showcases the benefit toward prediction via incorporation of multiple covariates.
+modes with available features) increases, the {\bf GTR} exhibits a reduction in error whereas others have increased errors.
+This showcases the benefit toward prediction via incorporation of multiple features.
 Note that our method {\bf GTR} is most comparable to {\bf HOLRR} when there is only a single informative mode.
@@
 
-\caption{Comparison of MSPE versus the number of modes with covariates.
+\caption{Comparison of MSPE versus the number of modes with features.
 We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and noise $\sigma=1/2$ (high), $\sigma=1/4$ (low).}~\label{fig:compare}
@@
 We apply our tensor regression model to two datasets.
-The first application concerns the brain network modeling in response to individual attributes (i.e.\ covariate on one mode), and the second application focuses on multi-relational network analysis with dyadic attributes (i.e.\ covariates on two modes).
+The first application concerns the brain network modeling in response to individual attributes (i.e.\ feature on one mode), and the second application focuses on multi-relational network analysis with dyadic attributes (i.e.\ features on two modes).
 
@@
 Each brain network is represented as a 68-by-68 binary matrix, where the entries encode the presence or absence of fiber connections between the 68 brain regions.
-We consider four individual-covariates: gender (65 females vs.\ 71 males), age 22-25 ($n=35$), age 26-30 ($n=58$), and age 31+ ($n=43$).
-The goal is to identify the connection edges that are affected by the individual covariates.
+We consider four individual features: gender (65 females vs.\ 71 males), age 22-25 ($n=35$), age 26-30 ($n=58$), and age 31+ ($n=43$).
+The goal is to identify the connection edges that are affected by the individual feature.
 A key challenge in brain network is that the edges are correlated; for example, two edges may stem out from a same brain region, and it is of importance to take into account the within-dyad dependence.
@@
 We fit the tensor regression model to the HCP data.
-The response is a binary tensor $\tY\in\{0,1\}^{68\times 68\times 136}$ and the covariates are of dimension 4 along the 3$^{\text{rd}}$ mode.
+The response is a binary tensor $\tY\in\{0,1\}^{68\times 68\times 136}$ and the features are of dimension 4 along the 3$^{\text{rd}}$ mode.
 The BIC selection suggests a rank $\mr=(10,10,4)$ with log-likelihood $\tL_{\tY}=-174654.7$.
@@
 Figure~\ref{fig:s1} compares the estimated coefficients from our method (tensor regression) with those from classical GLM approach.
-A classical GLM is to regress the brain edges, one at a time, on the individual-level covariates, and this logistic model is repeatedly fitted for every edge $\in [68]\times [68]$.
+A classical GLM is to regress the brain edges, one at a time, on the individual-level features, and this logistic model is repeatedly fitted for every edge $\in [68]\times [68]$.
 As we can see in the figure, our tensor regression shrinkages the coefficients towards center, thereby enforcing the sharing between coefficient entries.
@@
 The 56 relations span the fields of politics, economics, military, religion, and so on.
-In addition, country-level attributes are also available, and we focus on the following six covariates: \emph{constitutional, catholics, lawngos, politicalleadership, geographyx}, and \emph{medicinengo}.
+In addition, country-level attributes are also available, and we focus on the following six features: \emph{constitutional, catholics, lawngos, politicalleadership, geographyx}, and \emph{medicinengo}.
 The goal is to identify the variation in connections due to country-level attributes and interactions thereof.
-One of the key features is that the 56 relations are correlated, and we would like to take that into account in assessing the covariate effects.
+One of the key features is that the 56 relations are correlated, and we would like to take that into account in assessing the feature effects.
 
@@
 We applied our tensor regression model to the \emph{Nations} data.
-The multi-relational network $\tY\in\{0,1\}^{14\times 14\times 56}$ was treated as the response tensor, and the country attributes $\mM\in\mathbb{R}^{14\times 6}$ were treated as covariates on both the 1st and 2nd modes.
+The multi-relational network $\tY\in\{0,1\}^{14\times 14\times 56}$ was treated as the response tensor, and the country attributes $\mM\in\mathbb{R}^{14\times 6}$ were treated as features on both the 1st and 2nd modes.
 The BIC criterion suggests a rank $\mr=(4,4,4)$ for the coefficient tensor $\tB\in\mathbb{R}^{6\times 6 \times 56}$.
@@
 To investigate the effects of dyadic attributes towards connections, we depicted the estimated coefficients $\hat \tB=\entry{\hat b_{ijk}}$ for several relation types (Figure~\ref{fig:est}).
-Note that entries $\hat b_{ijk}$ can be interpreted as the contribution, at the logit scale, of covariate pair $(i,j)$ ($i$th covariate for the ``sender'' country and $j$th covariate for the ``receiver'' country) towards the connection of relation $k$.
+Note that entries $\hat b_{ijk}$ can be interpreted as the contribution, at the logit scale, of feature pair $(i,j)$ ($i$th feature for the ``sender'' country and $j$th feature for the ``receiver'' country) towards the connection of relation $k$.
 Several interesting findings emerge from the observation.
-We found that relations belonging to a same cluster tend to have similar covariate effects.
-For example, the relations \emph{warnings} and \emph{ecnomicaid} are classified into Cluster II, and both exhibit similar covariate pattern (Figure~\ref{fig:est}a-b).
+We found that relations belonging to a same cluster tend to have similar feature effects.
+For example, the relations \emph{warnings} and \emph{ecnomicaid} are classified into Cluster II, and both exhibit similar feature pattern (Figure~\ref{fig:est}a-b).
 Moreover, the majority of the diagonal entries $\hat \tB(i,i,k)$ positively contribute to the connection.
@@
 
-We have developed a generalized tensor regression with covariates on multiple modes.
+We have developed a generalized tensor regression with features on multiple modes.
 A fundamental feature of tensor-valued data is the statistical interdependence among entries.

### 2020-09-07 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
 \usepackage{natbib}
-\usepackage{url} % not crucial - just used below for the URL
+\usepackage{url}
 
@@
 \usepackage{dsfont,multirow,hyperref,setspace,enumerate}
-\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue}}
+\hypersetup{colorlinks,linkcolor={black},citecolor={blue},urlcolor={blue}}
 
-%\pdfminorversion=4
-% NOTE: To produce blinded version, replace "0" with "1" below.
 \newcommand{\blind}{1}
 
-% DON'T change margins - should be 1 inch all around.
 \addtolength{\oddsidemargin}{-.5in}%
@@
 
+\def\ci{\perp\!\!\!\perp}
+
+\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}
 \begin{document}
-
-%\bibliographystyle{natbib}
 
@@
 {
-  \title{\bf Supervised Tensor Decomposition with Multiple Side Information}
-  \author{Jiaxin Hu, Zhuoyan Xu, and Miaoyan Wang\\
-    Department of Statistics, University of Wisconsin-Madison}
+  \title{\bf Supervised Tensor Decomposition with Interactive Side Information}
+  %\author{Jiaxin Hu, Zhuoyan Xu, and Miaoyan Wang\\
+   % Department of Statistics, University of Wisconsin-Madison}
   \maketitle
@@
   \begin{center}
-    {\LARGE\bf Supervised Tensor Decomposition with Multiple Side Information}
+    {\LARGE\bf Supervised Tensor Decomposition with Interactive Side Information}
 \end{center}
@@
 Identifying the relationship between a high-dimensional tensor and side information is important yet challenging.
-Here, we develop a tensor decomposition method that incorporates multiple side information as structured features, and present case studies to show the benefit of proposed approach.
-Unlike unsupervised tensor decomposition, our supervised decomposition captures the effective dimension reduction of the data tensor confined to the feature space on each mode.
+Here, we develop a tensor decomposition method that incorporates multiple side information as interactive features.
+Unlike unsupervised tensor decomposition, our supervised decomposition captures the effective dimension reduction of the data tensor confined to feature space on each mode.
 An efficient alternating updating algorithm is further developed.
@@
 Our method will help the practitioners efficiently analyze tensor datasets in various areas.
-Toward this end, the package and data used are available at ~\url{https://CRAN.R-project.org/package=tensorregress}.
+Toward this end, the package and data used are available at~\url{https://CRAN.R-project.org/package=tensorregress}.
 
@@
 \noindent%
-{\it Keywords:} Applications and case studies, Supervised dimension reduction, Exponential family observation, Tensor decomposition
+{\it Keywords:} Applications and case studies, Tensor data analysis, Supervised dimension reduction, Exponential family distribution, Generalized multilinear model
 \vfill
@@
 A typical social network consists of nodes that represent people and edges that represent the friendships.
-Side information such as people’s demographic information and the closeness of a friendship are often available.
-In both examples, it is of keen scientific interest to identify the variation in the tensor observation (e.g., brain connectivities, social community patterns) that is affected by available feature data.
-These seemingly different scenarios pose a common yet challenging problem for tensor modeling.
+Side information such as people’s demographic information and friendship types are often available.
+In both examples, it is of keen scientific interest to identify the variation in the tensor data (e.g., brain connectivities, social community patterns) that is affected by available features.
+These seemingly different scenarios pose a common yet challenging problem for tensor data modeling.
 
-In addition to the aforementioned challenges, many instances of tensor observation consist of non-Gaussian measurements.
-Examples include the political interaction dataset \citep{hu2015scalable} which measures action counts between countries for various relationship, and the brain connectivity network data \citep{wang2019common} which is encoded as binary adjacency matrices.
-Classical tensor decomposition methods are based on minimizing the Frobenious norm of the reconstruction error, leading to suboptimal predictions for binary or count responses.
-A number of supervised tensor methods have been proposed \citep{yu2016learning, raskutti2015convex, zhao2012higher} to address the tensor regression problem in various forms (e.g.\ scalar-to-tensor regression, tensor-response regression).
+In addition to the aforementioned challenges, many instances of tensor datasets consist of non-Gaussian measurements.
+Examples include the political interaction dataset \citep{hu2015scalable} which measures action counts between countries under various relationships, and the brain connectivity network dataset \citep{wang2019common} which is a collection of binary adjacency matrices.
+Classical tensor decomposition methods are based on minimizing the Frobenious norm of the reconstruction error, leading to suboptimal predictions for binary- or count-valued response variables.
+A number of supervised tensor methods have been proposed \citep{narita2012tensor, zhao2012higher, yu2016learning} to address the tensor regression problem in various forms (e.g.\ scalar-to-tensor regression, tensor-response regression).
 These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest.
-The gap in theory and practice suggests a huge opportunity for improved modeling paradigms that better capture the complexity in tensor data with side information.
+The gap in theory and practice suggests a great opportunity for improved modeling paradigms that better capture the complexity in tensor data with side information.
 
@@
 \end{center}
-\caption{Examples of supervised tensor decomposition with multiple side information. (a) Network population model. (b) Spatial-temporal growth model. }\label{fig:intro1}
+\caption{Examples of supervised tensor decomposition with interactive side information. (a) Network population model. (b) Spatial-temporal growth model. }\label{fig:intro1}
 \end{figure*}
 
-We present a general model and associated method for decomposing exponential family tensors with multiple side information. 
-We formulate the learning task as a structured regression problem, with tensor observation serving as the response, and the side information forming multiple feature matrices.
+We present a general model and associated method for decomposing a data tensor with exponential family entries and interactive side information. 
+We formulate the learning task as a structured regression problem, with tensor observation serving as the response, and the multiple side information as interactive features.
 Figure~\ref{fig:intro1}b illustrates our model in the special case of order-3 tensors.
-A low-rank structure is imposed to the mean tensor $\mathbb{E}(\tY)\in\mathbb{R}^{d_1\times d_2\times d_3}$, where unlike classical decomposition, the tensor factors are assumed to belong to the space spanned by features $\mX_k\in\mathbb{R}^{p_k\times d_k}$ for $k=1,2,3$.
-The unknown matrix of interest $\mM_k\in\mathbb{R}^{r_k\times p_k}$ (referred to as ``dimension reduction matrix'') link the mean tensor to the feature space, thus allowing the identification of variations across tensor entries that can be explained by side information.
+A low-rank structure is imposed to the conditional mean tensor, where unlike classical decomposition, the tensor factors $\mX_k\mM_k\in\mathbb{R}^{d_k\times r_k}$ are assumed to belong to the space spanned by features $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,2,3$.
+The unknown matrices $\mM_k\in\mathbb{R}^{p_k\times r_k}$ (referred to as ``dimension reduction matrices'') link the conditional mean to the feature spaces, thereby allowing the identification of variations in the tensor data attributable to the side information.
 
-In addition, we use ideas from generalized linear model (GLM) to allow heteroscedacity due to mean-variance relationship across tensor entries.
+In addition, we use tools from generalized linear model (GLM) to allow heteroscedacity due to the mean-variance relationship in the non-Gaussian data.
 This flexibility is important in practice.
-In classical GLM, the sample size $n$ and feature dimension $d$ are well defined; however, in the tensor data analysis, we observe only one single realization of tensor response but up to $K$ feature matrices, where $K$ is the number of tensor modes.
+In classical GLM, the sample size and feature dimension are well defined; however, in the tensor data analysis, we observe only one single realization of tensor response but up to $K$ interactive feature matrices, where $K$ is the number of tensor modes.
 Both the sample size and feature dimension grow exponentially in $K$.
-We overcome the high-dimensionality using a possibly non-linear, low-rank constraint in the mean tensor, thereby encouraging the sharing among entries in the data tensor.
+We use a low-rank constraint in the conditional mean tensor (with a suitable transformation) to mitigate the curse of high dimensionality.
 The statistical convergence of our estimator is established, and we quantify the gain in predictive power through several case studies.
@@
 Our work is closely related to but also clearly distinctive from several lines of previous work.
-The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor,zhang2018tensor, hong2020generalized} that aims to find the best low-rank representation of data tensor.
-In contrast, our model can be viewed a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and features.
+The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor,zhang2018tensor, hong2020generalized} that aims to find the best low-rank representation of a data tensor.
+In contrast, our model is a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and multiple features.
 The low-rank factorization is determined jointly by the tensor data and feature matrices on multiple sides.
 
-The second line of work studies the scalar-to-tensor regression models, in which the response is a scalar and the \textit{predictor} is a tensor~\citep{zhou2013tensor,chen2019non}.
-Our proposal is orthogonal to theirs because we treat the tensor as random variables.
+The second line of work studies the scalar-to-tensor regression, in which the response is a scalar and the predictor is a tensor~\citep{zhou2013tensor,chen2019non}.
+Our proposal is orthogonal to theirs because we treat the data tensor as random variables.
 The difference in model settings yields different focuses and interpretations.
 Consider the neuroimaging analysis as an example.
-The scalar-to-tensor regression aims to predict the individual response based on brain connectivity network, while tensor-to-scalar model focuses on understanding the variation in the brain connectivity that is attribute to the individual features such as age, gender and disease status.
-The scalar-to-tensor model is not suitable for tensor dimension reduction because it lacks the modeling of stochastic noise in the data tensor.
+The scalar-to-tensor regression aims to predict the individual response based on brain connectivity network, while tensor-to-scalar model focuses on understanding the variation in the brain connectivity attributable to the individual features such as age, gender, and disease status.
+The scalar-to-tensor model is insufficient for tensor dimension reduction because it lacks the modeling of stochastic noise in the data tensor.
 
-The third line of work studies the tensor-on-tensor regression models, in which both the response and predictor are tensors~\citep{raskutti2015convex, lock2018tensor, gahrooei2020multiple}.
-Our model shares a common ground but is complementary to earlier approaches.
+The third line of work studies the tensor-on-tensor regression, in which both the response and predictor are tensors~\citep{raskutti2015convex, lock2018tensor, gahrooei2020multiple}.
+Our model shares a common ground but provides a more efficient context-specific solutions than earlier appraoches.
 As we show in Section~\ref{subsec:statprob}, the supervised tensor decomposition
 has an interesting interpretation as a special tensor-on-tensor regression.
-Nevertheless, our work differs from previous work in scope and application domains.
+Nevertheless, our work improves from previous work in scope and applicability.
 Previous methods~\citep{gahrooei2020multiple,lock2018tensor} mainly focus on Gaussian tensor.
@@
 However, in the tensor case, convex MLE suffers from both computational intractability and statistical suboptimality.
-We advocate the non-convex approach and provide strong evidence for its success in our setting.
-Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2015convex,gahrooei2020multiple}, and we go step further by finding the sufficient dimension reduction~\citep{adragni2009sufficient}, $\text{Span}(\mM_k)$, in order to pinpoint effective features in prediction (see Figure~\ref{fig:intro1}b).
+We advocate a non-convex approach and provide strong evidence for its success in our setting.
+Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2015convex,gahrooei2020multiple}, and we go step further by finding the sufficient dimension reduction~\citep{adragni2009sufficient}, $\text{Span}(\mM_k)$, that facilitates the identification of effective features in prediction (see Figure~\ref{fig:intro1}b).
 The latter approach greatly improves the \emph{interpretability} in prediction.
 In this regards, our method opens up new opportunities for tensor data analysis in a wider range of possible applications.
+
+\fixme{Miaoyan}{One more paragraph. Current literature on tensor decomposition with auxiliary side information...}
 
@@
 Section~\ref{sec:est} describes a rank-constrained likelihood-based estimation and associated alternating optimization algorithm.
-In Section~\ref{sec:simulation}, we present simulation results and performance comparison to alternative methods.
+In Section~\ref{sec:simulation}, we present numerical experiments and assess the performance in comparison to alternative methods.
 In Section~\ref{sec:data}, we apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
@@
 which results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor.
-For ease of presentation, we use shorthand notion $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product.
+For ease of presentation, we use the shorthand $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product.
 For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as
@@
 \] 
-The Frobenius norm of $\tY$ is defined as 
+The Frobenius norm and maximum norm of $\tY$ are defined as 
 \[
-\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}.
+\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}\quad \text{and}\quad \mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.
 \]
@@
 Given a matrix $\mM\in\mathbb{R}^{d\times r}$, we use $\text{Span}(\mM)$ to denote the space spanned by columns of $\mM$.
-We use lower-case letters (e.g.,\ $a,b,c$) for scalars/vectors, upper-case boldface letters (e.g.,\ $\mA,\ \mB,\ \mC$) for matrices, and calligraphy letters (e.g.,\ $\tA,\ \tB, \ \tC$) for tensors of order three or greater.
+We use lower-case letters (e.g.,\ $a,\ b,\ c$) for scalars and vectors, upper-case boldface letters (e.g.,\ $\mA,\ \mB,\ \mC$) for matrices, and calligraphy letters (e.g.,\ $\tA,\ \tB,\  \tC$) for tensors of order three or greater.
 We let $\mI_d$ denote the $d \times d$ identity matrix and $[d]$ denote the $d$-set $\{1,\ldots,d\}$.
-For ease of notation, we allow the basic arithmetic operators ($+$, $-$, $\geq $) and univariate
+For ease of notation, we allow the basic arithmetic operators (e.g.,\ $+$, $-$, $\geq $) and univariate
 functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner.
@@
 \subsection{General framework for supervised tensor decomposition}
-We first present a general framework for supervised tensor decomposition and then discuss its implication in three concrete examples.
+We begin with a general framework for supervised tensor decomposition and then discuss its implication in three concrete examples.
 Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor.
 Suppose the side information is available on each of the $K$ modes.
-Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the feature matrix on the mode $k\in[K]$, where $x_{ij}$ denotes the $j$-th feature value for the $i$-th tensor entity, for $(i,p)\in[d_k]\times[p_k]$, $p_k\leq d_k$.
-We model the relationship between the data tensor and feature matrices using an exponential family, multilinear mean model.
-Specifically, assumed that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent random variables from an exponential family, and the conditional mean tensor has the form
+Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the feature matrix on the mode $k\in[K]$, where $x_{ij}$ denotes the $j$-th feature value for the $i$-th tensor entity, for $(i,j)\in[d_k]\times[p_k]$, $p_k\leq d_k$.
+We propose a multilinear conditional mean model between the data tensor and feature matrices.
+Assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent random variables following an exponential family distribution, and the conditional mean admits the form
 \begin{align}\label{eq:tensormodel}
@@
 \end{align}
-where $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the unknown parameter tensor, $f(\cdot)$ is a known link function according to the data type of $\tY$, and $\times$ denotes the tensor Tucker product.
-The choice of link function depends on the specific distribution family of tensor entries.
+where $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the unknown parameter tensor, $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$, and $\times$ denotes the tensor-by-matrix product.
+The choice of link function is based on the assumed distribution family of tensor entries.
 Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution.
 In general, dispersion parameters can also be included in the model.
-Because our main focus is the tensor decomposition under the mean model, we omit the dispersion parameter for notational simplicity.
+Because our main focus is the tensor decomposition under the mean model, we omit the dispersion parameter in this section for ease of presentation.
 
@@
 \end{equation}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, and $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices with orthonormal columns, for all $k\in[K]$.
-By the definition of multilinear rank, model~\eqref{eq:factor} implies that the latent tensor $\tB$ has multilinear-rank $\mr=(r_1,\ldots,r_K)\leq (p_1,\ldots,p_K)$.
-Combining~\eqref{eq:tensormodel} and~\eqref{eq:factor} yields our final tensor decomposition model,
+where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, and $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices consisting of orthonormal columns, where $r_k\leq p_k$ for all $k\in[K]$.
+By the definition of multilinear rank, model equations~\eqref{eq:tensormodel} and~\eqref{eq:factor} imply the low-rankness $\mr=(r_1,\ldots,r_K)$ of the conditional mean tensor under the link function.
+We now reach our final model for supervised tensor decomposition,
 \begin{align}\label{eq:decomp}
-\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1,\ \ldots,\ \mX_K\mM_K\}\right),\\
-\text{where} \ &\ \mM_k\mM^T_k=\mI_{p_k},\quad \text{for all }k=1,\ldots,K
+\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1,\ \ldots,\ \mX_K\mM_K\}\right),\notag\\
+\text{where} \ &\ \mM_k\mM^T_k=\mI_{p_k},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\ \text{for all }k=1,\ldots,K,
 \end{align}
 where the parameters of interest are $\mM_k$ and $\tC$.
-As illustrated in Figure~\ref{fig:intro1}b, the columns of $\mX_k\mM_k$ are linear combinations of features that are relevant in the mean tensor.
-We call $\mX_k\mM_k$ the ``supervised tensor factors'' or ``sufficient features''~\citep{adragni2009sufficient}, and call $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects between sufficient features across multiple modes.
-Our goal is to find $\mM_k$ and corresponding $\tC$.
+The model formulation~\eqref{eq:decomp} assumes a fixed, known rank $\mr=(r_1,\ldots,r_K)$; the adaptation to unknown rank will be described in Section~\ref{sec:tuning}.
+Figure~\ref{fig:intro1}b provides a schematic illustration of our model.
+Note that the features $\mX_k$ affect the distribution of tensor $\tY$ through the form $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$.
+We call $\mX_k\mM_k$ the ``supervised tensor factors'' or ``sufficient features''~\citep{adragni2009sufficient}, and call $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes.
+Our goal is to find $\mM_k$ and the corresponding $\tC$.
 Note that $\mM_k$ and $\tC$ are identifiable only up to orthonormal transformations.
@@
 \begin{equation}\label{eq:network}
-\text{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n
+\text{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n,
 \end{equation}
@@
  \end{equation}
- where $\mB\in\mathbb{R}^{p\times p}$ is a symmetric, low rank matrix.
+ where $\mB\in\mathbb{R}^{p\times p}$ is a symmetric rank-$r$ matrix.
 The low-rankness in $\mB$ has demonstrated its success in modeling transitivity, balance, and communities in the networks~\citep{hoff2005bilinear}.
-We show that our tensor regression model~\eqref{eq:tensormodel} also incorporates the graph model as a special case.
+We show that our supervised tensor decompostiion~\eqref{eq:tensormodel} also incorporates the graph model as a special case.
 Let $\tY=\entry{y_{ij}}$ be a binary matrix where $y_{ij}=\mathds{1}_{(i,j)\in E}$.
@@
 
-{\color{red}revise from here....}
-In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by the features.
+In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by available features.
 The regression coefficient $\tB$ in our model model~\eqref{eq:tensormodel} serves this goal by collecting the feature effects and the interaction thereof.
@@
 
-Our tensor regression model is able to incorporate an arbitrary number of feature matrices along the tensor modes.
-When the side information is unavailable for certain mode $k$, we set $\mX_k=\mI_{d_k}$.
-In particular, our model~\eqref{eq:decomp} reduces to classical unsupervised tensor decomposition model~\citep{de2000multilinear,hong2018generalized} when no side information is available, i.e., $\mX_k=\mI_{d_k}$ for all $k\in[K]$.
+Our supervised tensor decomposition~\eqref{eq:decomp} is able to incorporate any arbitrary number of feature matrices.
+When certain mode $k$ has no available side information, we set $\mX_k=\mI_{d_k}$.
+In particular, our model~\eqref{eq:decomp} reduces to classical unsupervised tensor decomposition~\citep{de2000multilinear,hong2018generalized} when no side information is available at all; i.e., $\mX_k=\mI_{d_k}$ for all $k\in[K]$.
 %Then, the final form of our tensor regression model can be written as:
@@
 %\end{align}
-%where the entries of $\tY$ are independent random variables conditional on $\tX$, and $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$ is the low-rank coefficient tensor of interest. 
+%where the entriesof $\tY$ are independent random variables conditional on $\tX$, and $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$ is the low-rank coefficient tensor of interest. 
 %We comment that other forms of tensor low-rankness are also possible, and here we choose Tucker rank just for parsimony.
@@
 \subsection{Connection to sufficient dimension reduction and tensor-on-tensor regression}
-Joint dimension reduction of tensor and features....
+An important feature of our method is that we allow high-dimensionality in both tensor dimension $d_k$ and feature dimension $p_k$.
+In most applications, the rank is much smaller than tensor dimension $d_k$ and the feature dimension $p_k$.
+In this case, the matrices $\mM_k$ serve the role of simultaneous dimension reduction of the data tensor and features.
+The key implications of the sufficient features $\mX_k\mM_k$ can be seen in the following two conditional independent assumptions in the model~\eqref{eq:decomp},
+\begin{align}
+\tY \ci \{\mX_k\}&\ \big|\ \{\mX_k\mM_k\} \ \ \text{\small (independence between the tensor and multiple features)},\\
+y_\omega \ci y_{\omega'} & \ \big|\ \{\mX_k\mM_k\}  \ \ \text{\small (independence within the tensor)},
+\end{align}
+where the second line holds for all $\omega\neq \omega' \in[d_1]\times\cdots\times[d_K]$, and $\ci$ denotes the independence.
+The first property highlights the ``decorrelation'' role of $\mM_k$ akin to the sufficient dimension reduction in univariate supervised learning~\citep{adragni2009sufficient}, whereas the second property highlights the tensor dimension reduction in the usual unsupervised sense (consider, for example, $\mX_k=\mathbf{I}_{d_k}$ for all $k\in[K]$).
 
 Denote $d_{\text{total}}=\prod_k d_k$ and $p_{\text{total}}=\prod_k p_k$.
-Let $\tX=\mX_1\otimes \cdots \otimes \mX_K \in \mathbb{R}^{d_{\text{total}} \times p_{\text{total}}}$ be the tensor product of feature spaces, and $\text{Unfold}(\tX) \in\mathbb{R}^{d_{\text{total}}\times p_{\text{total}}}$.
+Let $\tX=\mX_1\otimes \cdots \otimes \mX_K \in \mathbb{R}^{d_{\text{total}} \times p_{\text{total}}}$ be the tensor product of feature spaces.
+....
+$\text{Unfold}(\tX) \in\mathbb{R}^{d_{\text{total}}\times p_{\text{total}}}$.
 \[
-\mathbb{E}(\text{vec}(\tY)|\mX,\mZ,\mW) =\beta_{111}\mx^1\mz^1\mw^1+\beta_{121}\mx^1\mz^2\mw^1+\beta_{112}\mz^1\mz^1\mw^2+\cdots+\beta_{222}\mx^2\mz^2\mw^2.
+\mathbb{E}(\text{vec}(\tY)|\mX,\mZ,\mW) =\beta_{111}\mx^1\mz^1\mw^1+\beta_{121}\mx^1\mz^2\mw^1+\beta_{112}\mx^1\mz^1\mw^2+\cdots+\beta_{222}\mx^2\mz^2\mw^2.
 \]
-where each $\mM_1\mX=[\mx^1,\mx^2]$, $\mM_2\mZ=[\mz^1,\mz^2]$, $\mM_3\mW=[\mw^1,\mw^2]$ are sufficient features. when ....Full interaction model with rank-1 features.
+where each $\mM_1\mX=[\mx^1,\mx^2]$, $\mM_2\mZ=[\mz^1,\mz^2]$, $\mM_3\mW=[\mw^1,\mw^2]$ are sufficient features.  ....Full interaction model with rank-1 features.
 
@@
 
-\subsection{Rank-constrained likelihood estimator}
+\subsection{Rank-constrained M-estimator}
 We develop a likelihood-based procedure to estimate $\tC$ and $\mM_k$ in~\eqref{eq:decomp}.
@@
 \end{align}
-We assume that we have an additional information on an upper bound $\alpha>0$ such that $\mnormSize{}{\Theta}\leq \alpha$. We propose a constrained maximum likelihood estimator (MLE)
+
+{\color{red}We assume that we have an additional information on an upper bound $\alpha>0$ such that $\mnormSize{}{\Theta}\leq \alpha$.} We propose a constrained maximum likelihood estimator (M-estimator)
 %\begin{align} \label{eq:MLE} 
@@
 \begin{align} \label{eq:MLE} 
-(\hat \tC, \hat \mM_1,\ldots,\hat \mM_k) &=\argmax_{(\tC,\mM_k)\in \tP_{r_1,\ldots,r_K}} \tL_{\tY}(\tC,\mM_k)\\
-\text{where } \tP_{r_1,\ldots,r_K}&=\left\{(\tC,\mM_k)\colon \mM_k\mM^T_k=\mathbf{1}_{p_k},\ \mnormSize{}{\Theta(\tC,\mM_k)}\leq \alpha, \right.\\
-&\quad \quad\quad \quad  \quad \quad\quad \quad \left.\mK_k\in\mathbb{R}^{p_k\times r_k},\ \tC\in\mathbb{R}^{r_1\times \cdots\times r_K}\right\}.
+(\hat \tC, \hat \mM_1,\ldots,\hat \mM_k) &=\argmax_{(\tC,\mM_1,\ldots,\mM_K)\in \tP} \ \tL_{\tY}(\tC,\mM_1,\ldots,\mM_K),
 \end{align}
+where the parameter space $\tP$ is described as
+\[
+\tP=\left\{\tC\in\mathbb{R}^{r_1\times \cdots\times r_K},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\text{ for all }k\in[K]\ \Big| \ \mM_k\mM^T_k=\mathbf{I}_{p_k}, \mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha \right\}.
+\]
 
@@
 \begin{algorithm}[t]
-\caption{Supervised Tensor Decomposition with Multiple Side Information}\label{alg:B}
+\caption{Supervised Tensor Decomposition with Interactive Side Information}\label{alg:B}
 \begin{algorithmic}[1]
@@
 Therefore, the optimization~\eqref{eq:MLE} is a non-convex problem, and the Algorithm~\ref{alg:B} usually not theoretically possesses the global optimality.
-However, as mentioned in Section~\ref{subsec:statprob}, the desired convergence rate~\eqref{eq:bound} holds for the valid estimators satisfying $\tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true})$, which indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective of local optimums are large enough.
+\fixme{Miaoyan}{This is wrong.
+Proof of Theorem 4.1 is based on global optimality condition.
+Please relax the proof.} However, as mentioned in Section~\ref{subsec:statprob}, the desired convergence rate~\eqref{eq:bound} holds for the valid estimators satisfying $\tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true})$, which indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective of local optimums are large enough.
 Fortunately, we find the  Algorithm~\ref{alg:B}  often gives a satisfactory convergence point $\hat \tB$, when we initialize the parameters via continuous-valued Tucker decomposition.
@@
 
-The computational complexity of our tensor regression model is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the response tensor.
+The computational complexity of our alternating optimization algorithm is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the response tensor.
 More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor.
@@
 \begin{thm}[Statistical convergence]\label{thm:main}
-Consider a generalized tensor regression model with features on multiple modes $\tX=\{\mX_1,\ldots,\mX_K\}$.
-Suppose the entries in $\tY$ are independent realizations of an exponential family distribution, and $\mathbb{E}(\tY|\tX)$ follows the low-rank tensor regression model~\eqref{eq:tensormodel2}.
+Consider an order-$K$ dimensional-$(d_1,\ldots,d_K)$ tensor $\tY$ and multiple feature matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$, for $k\in[K]$.
+Suppose that, conditional on the features, the entries in $\tY$ are independent random variables following an exponential family distribution with conditional mean model~\eqref{eq:decomp}.
+Let $r_\text{total}=\prod_k r_k$, $r_{\max}=\max_k r_k$, and $\lambda_{\min}(\mM_k)$ be the minimal non-zero singular value of $\mM_k$.
+Assume that $\lambda_{\min}(\mM_k)>0$ and $\FnormSize{}{\mX_k} \asymp \Omega(\sqrt{d_k})$, for all $k\in[K]$.
 Under Assumption~\ref{ass}, there exist two constants $C_1, C_2>0$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$,
 \begin{equation}\label{eq:bound}
-\text{Loss}(\trueB,\ \hat \tB) \leq \frac{ C_{2}\prod_kr_k}{\max_k r_k}\sum_k p_k.
+\textup{sin}^2 \Theta(\mM_k,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\lambda_{\min}(\mM_k)}\frac{\sum_k p_k}{\prod_k d_k}, \quad\text{for all }k\in[K],
 \end{equation}
-Here, $C_2=C_2(\mr,\alpha, K)>0$ is a constant that does not depend on the dimensions $\{d_k\}$ and $\{p_k\}$. 
+and
+\[
+\FnormSize{}{\trueB- \hat \tB}^2\leq \frac{ C_{2} r_{\textup{total}}}{r_{\max}}\frac{\sum_k p_k}{\prod_k d_k}.
+\]
+Here, $C_2=C_2(\mr,\alpha, K)>0$ is a constant independent of the dimensions $\{d_k\}$ and $\{p_k\}$. 
 \end{thm}
@@
 Theorem~\ref{thm:main} establishes the statistical convergence for the estimator~\eqref{eq:MLE}.
-Actually, the proof in section~\ref{sec:appedix} shows that the statistically optimal rate holds, not only for the MLE~\eqref{eq:MLE},  but also any local estimators $\hat \tB$ in the level set $\{\hat \tB \in \tP \colon \tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true}) \}$ satisfies the upper bound~\ref{thm:main}.
+\fixme{Miaoyan}{Wrong statement based on current proof}Actually, the proof in section~\ref{sec:appedix} shows that the statistically optimal rate holds, not only for the MLE~\eqref{eq:MLE},  but also any local estimators $\hat \tB$ in the level set $\{\hat \tB \in \tP \colon \tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true}) \}$ satisfies the upper bound~\ref{thm:main}.
 In Section~\ref{subsec:alg}, we implement empirical studies to verify the algorithmic stability to find a local optimum.
@@
 \[
-\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY})\leq \frac{ C_{4}\prod_kr_k}{\max_k r_k} \sum_k p_k,
+\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY})\leq \frac{ C_{4}r_{\textup{total}}}{r_{\max}} {\sum_k p_k\over \prod_k d_k},
 \]
-where $C_4=C_4(\mr,\alpha, K)>0$ is a constant that do not depend on the dimensions $\{d_k\}$ and $\{p_k\}$.
+where $C_4=C_4(\mr,\alpha, K)>0$ is a constant that independent of dimensions $\{d_k\}$ and $\{p_k\}$.
 \end{thm}
 
-\section{Simulation}\label{sec:simulation}
+\section{Numerical experiments}\label{sec:simulation}
+{\color{red} Simulation needs to be re-done, since code has been changed. }
+
 We evaluate the empirical performance of our generalized tensor regression through simulations.
@@
 modes with available features) increases, the {\bf GTR} exhibits a reduction in error whereas others have increased errors.
-This showcases the benefit toward prediction via incorporation of multiple features.
+This showcases the benefit toward prediction via incorporation of interactive features.
 Note that our method {\bf GTR} is most comparable to {\bf HOLRR} when there is only a single informative mode.
@@
 
-\begin{figure}[hbt]
+\begin{figure}[H]
 \centering
-\includegraphics[width=10cm]{HCP.pdf}
+\includegraphics[width=13cm]{HCP.pdf}
 \caption{Top edges with large effects.
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=16cm]{compare.pdf}
+\caption{Comparison of coefficient estimation in the HCP data.}\label{fig:s1}
+\end{figure}
+
 \subsection{Analysis of political relational data}
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=10cm]{coef.pdf}
-\caption{Effect estimation in the \emph{Nations} data.
-Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warnning}, \emph{economicaid},\emph{intergovorg}, and \emph{commonblock}, respectively.
-}\label{fig:est}
-
-\end{figure}
-
 We applied our tensor regression model to the \emph{Nations} data.
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[width=14cm]{coef.pdf}
+\caption{Effect estimation in the \emph{Nations} data.
+Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warnning}, \emph{economicaid},\emph{intergovorg}, and \emph{commonblock}, respectively.
+}\label{fig:est}
+\end{figure}
+
 \begin{table}[H]
@@
 \hline
-\multirow{4}{*}{Cluster I}  & Government &officialvisits, intergovorgs,  militaryactions, violentactions, duration, \\
-&&negativebehavior, boycottembargo, aidenemy, negativecomm, accusation, \\
-& &protestsunoffialacts,    nonviolentbehavior, emigrants, relexports,   \\
-& &timesincewar, commonbloc2, rintergovorgs3, relintergovorgs\\
+\multirow{5}{*}{Cluster I}  & Government &officialvisits, intergovorgs,  militaryactions,  \\
+&&negativebehavior, boycottembargo, aidenemy, negativecomm, \\
+&&protestsunoffialacts,    nonviolentbehavior, emigrants,  \\
+&&timesincewar, commonbloc2, rintergovorgs3, relintergovorgs\\
+&&violentactions, duration,accusation,relexports\\
 \hline
-\multirow{4}{*}{Cluster II}& Economics &economicaid, booktranslations, tourism, relbooktranslations, releconomicaid, \\
-&& conferences, severdiplomatic, expeldiplomats, attackembassy, unweightedunvote,\\
-&& reltourism, tourism3, relemigrants, emigrants3, students, relstudents, \\
-&&exports, exports3, lostterritory, dependent, militaryalliance, warning  \\
+\multirow{6}{*}{Cluster II}& Economics &economicaid, booktranslations, tourism,  \\
+&& conferences, severdiplomatic, expeldiplomats, attackembassy, \\
+&& reltourism, tourism3, relemigrants, emigrants3, students, \\
+&&exports, exports3, lostterritory, dependent, militaryalliance  \\
+&&relbooktranslations, releconomicaid, unweightedunvote, \\
+&& warning, relstudents,\\
 \hline
-\multirow{2}{*}{Cluster III} & treaties &treaties, reltreaties, exportbooks, relexportbooks, weightedunvote, ngo, \\
-&&relngo, ngoorgs3, embassy, reldiplomacy, timesinceally, independence, commonbloc1\\
+\multirow{3}{*}{Cluster III} & treaties &treaties, reltreaties, exportbooks, relexportbooks,  \\
+&&relngo, ngoorgs3, embassy, reldiplomacy, timesinceally, \\
+&&independence, commonbloc1,weightedunvote, ngo,\\
 \hline
 Cluster IV & Politics &commonbloc0, blockpositionindex\\
+\hline
 \end{tabular}
@@
 
-Our method leverages on the alternating updating algorithm for the optimization.
+ Add three more paragraphs:
+
+ 1. Optimization 
+%Our method leverages on the alternating updating algorithm for the optimization.
 Although a non-convex optimization procedure such as Algorithm~\ref{alg:binary} has no guarantee on global optimality, our numerical experiments have suggested that, upon random initializations, the convergence point $\hat\Theta$ is often satisfactory, in that the corresponding objective value $\tL_{\tY}(\hat\Theta)$ is close to the objective value $\tL_{\tY}(\trueT)$.
@@
 
-Another an..
+%Another an..
 is the (conditional) independence assumption on the tensor entries.
@@
 
-Inference. ... uncertainly quantification... bootstrap ....
+2. Assumption on (conditional) independence assumption on the tensor entries.
+
+3. Inference. ... uncertainly quantification... bootstrap ....  Testing for Interaction + main effects.
+
+\section*{Acknowledgements}
+This research was supported by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
+%The authors would like to thank Chen Zhang from Wisconsin State Lab of Hygiene for help with figure 4.
+
+\bibliography{tensor_wang}
+\bibliographystyle{apalike}
 
 \newpage
-
-\bigskip
 \begin{center}
@@
 
-\begin{figure}[H]
-\includegraphics[width=17cm]{compare.pdf}
-\caption{Comparison of coefficient estimation in the HCP data.}\label{fig:s1}
-\end{figure}
+\section{Code}
+The source code and data used in the paper are available at \url{https://CRAN.R-project.org/package=tensorregress.}
 
-\section*{Acknowledgements}
-This research was supported by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
-The authors would like to thank Chen Zhang from Wisconsin State Lab of Hygiene for help with figure 4.
-
-\bibliography{tensor_wang}
-\bibliographystyle{apalike}
 \end{document}

### 2020-09-08 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
 Unlike unsupervised tensor decomposition, our supervised decomposition captures the effective dimension reduction of the data tensor confined to feature space on each mode.
-An efficient alternating updating algorithm is further developed.
+An efficient alternating optimization algorithm is further developed.
 Our proposal handles a broad range of data types, including continuous, count, and binary observations.
@@
 Figure~\ref{fig:intro1}b illustrates our model in the special case of order-3 tensors.
-A low-rank structure is imposed to the conditional mean tensor, where unlike classical decomposition, the tensor factors $\mX_k\mM_k\in\mathbb{R}^{d_k\times r_k}$ are assumed to belong to the space spanned by features $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,2,3$.
+A low-rank structure is imposed to the conditional mean tensor, where unlike classical decomposition, the tensor factors $\mX_k\mM_k\in\mathbb{R}^{d_k\times r_k}$ belong to the space spanned by features $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,2,3$.
 The unknown matrices $\mM_k\in\mathbb{R}^{p_k\times r_k}$ (referred to as ``dimension reduction matrices'') link the conditional mean to the feature spaces, thereby allowing the identification of variations in the tensor data attributable to the side information.
@@
 We use a low-rank constraint in the conditional mean tensor (with a suitable transformation) to mitigate the curse of high dimensionality.
-The statistical convergence of our estimator is established, and we quantify the gain in predictive power through several case studies.
+The statistical convergence of our estimator is established, and we quantify the gain in prediction through several case studies.
 
@@
 In contrast, our model is a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and multiple features.
-The low-rank factorization is determined jointly by the tensor data and feature matrices on multiple sides.
+The low-rank factorization is determined jointly by the tensor data and feature matrices.
 
 The second line of work studies the scalar-to-tensor regression, in which the response is a scalar and the predictor is a tensor~\citep{zhou2013tensor,chen2019non}.
-Our proposal is orthogonal to theirs because we treat the data tensor as random variables.
+Our proposal is orthogonal to this line because we treat the data tensor as random variables.
 The difference in model settings yields different focuses and interpretations.
@@
 The third line of work studies the tensor-on-tensor regression, in which both the response and predictor are tensors~\citep{raskutti2015convex, lock2018tensor, gahrooei2020multiple}.
-Our model shares a common ground but provides a more efficient context-specific solutions than earlier appraoches.
+Our model shares a common ground but provides a more efficient context-specific solutions than earlier approaches.
 As we show in Section~\ref{subsec:statprob}, the supervised tensor decomposition
@@
 Nevertheless, our work improves from previous work in scope and applicability.
-Previous methods~\citep{gahrooei2020multiple,lock2018tensor} mainly focus on Gaussian tensor.
-The Frobenius norm used in the objective is statistical suboptimal for general exponential family tensors.
+Previous methods~\citep{gahrooei2020multiple,lock2018tensor} mainly focus on Gaussian tensors.
+The Frobenius norm used in the objective function is statistical suboptimal for general exponential family tensors.
 Maximum likelihood estimator (MLE) is studied in~\cite{raskutti2015convex} and a convex relaxation algorithm is proposed to solve for low-rank tensor coefficients.
@@
 We introduce the basic tensor properties used in the paper; more details on tensor notation can be found in~\citet{kolda2009tensor}.
-We use $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ to denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor, where $K$ corresponds to the number of modes of $\tY$ and is called the order.
+We use $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$\fixme{Miaoyan}{perhaps use $y_\omega$} to denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor, where $K$ corresponds to the number of modes of $\tY$ and is called the order.
 The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as
@@
 We propose a multilinear conditional mean model between the data tensor and feature matrices.
-Assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent random variables following an exponential family distribution, and the conditional mean admits the form
+Assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent realizations from an exponential family distribution, and the conditional mean tensor admits the form
 \begin{align}\label{eq:tensormodel}
@@
 where the parameters of interest are $\mM_k$ and $\tC$.
-The model formulation~\eqref{eq:decomp} assumes a fixed, known rank $\mr=(r_1,\ldots,r_K)$; the adaptation to unknown rank will be described in Section~\ref{sec:tuning}.
+Note that model~\eqref{eq:decomp} assumes a fixed, known rank $\mr=(r_1,\ldots,r_K)$; the adaptation to unknown rank will be described in Section~\ref{sec:tuning}.
 Figure~\ref{fig:intro1}b provides a schematic illustration of our model.
-Note that the features $\mX_k$ affect the distribution of tensor $\tY$ through the form $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$.
+The features $\mX_k$ affect the distribution of tensor entries in $\tY$ through the form $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$.
 We call $\mX_k\mM_k$ the ``supervised tensor factors'' or ``sufficient features''~\citep{adragni2009sufficient}, and call $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes.
@@
 \begin{example}[Spatio-temporal growth model]
+The growth curve model~\citep{gabriel1998generalised,potthoff1964generalized,srivastava2008models} was originally proposed as an example of bilinear model for matrix data, and we adopt its higher-order extension here.
 Let $\tY=\entry{y_{ijk}}\in\mathbb{R}^{d \times m\times n}$ denote the pH measurements of $d$ lakes at $m$ levels of depth and for $n$ time points.
@@
 are the design matrices for spatial and temporal effects, respectively, $\tC\in\mathbb{R}^{r_1\times r_2\times r_3}$ is the unknown core tensor, and $\mM_k$ are unknown dimension reduction matrices of coherent sizes.
-The model~\eqref{eq:time} is a higher-order extension of the bilinear ``growth curve'' model originally proposed for matrix data~\citep{gabriel1998generalised,potthoff1964generalized,srivastava2008models}.
 The factors $\mX_k\mM_k$ are sufficient features for the mean model~\eqref{eq:time}.
-Clearly, the spatial-temporal model is a special case of our supervised tensor decomposition model, with features available on each of the three modes.
+The spatial-temporal model is a special case of our supervised tensor decomposition model, with features available on each of the three modes.
 \end{example}
 
-\begin{example}[Network population model] 
-Network response model is recently developed in the context of neuroimanig analysis.
-The goal is to study the relationship between brain network connectivity pattern and the individual features.
+\begin{example}[Network population model]\label{example:brain}
+Network response model~\citep{rabusseau2016low, zhang2018network} is recently developed for neuroimanig analysis.
+The goal is to study the relationship between brain network connectivity pattern and features of individuals.
 Suppose we have a sample of $n$ observations, $\{(\mY_i, \mx_i)\colon i=1,\ldots,n\}$, where for each individual $i\in[n]$, $\mY_i\in\{0,1\}^{d\times d}$ is the symmetric adjacency matrix whose entries indicate presences/absences of connectivities between $d$ brain nodes, and $\mx_i\in\mathbb{R}^p$ is the individual's feature such as age, gender, cognition score, etc.
-The network-response model~\citep{rabusseau2016low, zhang2018network} has the conditional mean
+The network-response model  has the conditional mean
 \begin{equation}\label{eq:network}
@@
   \]
-  where $\mC\in\mathbb{R}^{r\times r}$ and $\mM\in\mathbb{R}^{p\times r}$ can be taken as the singular value decomposition of $\mB=\mM\mC\mM^T$. 
+  where $\mC\in\mathbb{R}^{r\times r}$ and $\mM\in\mathbb{R}^{p\times r}$ are based on the singular value decomposition of $\mB=\mM\mC\mM^T$. 
   \end{example}
@@
 The regression coefficient $\tB$ in our model model~\eqref{eq:tensormodel} serves this goal by collecting the feature effects and the interaction thereof.
-To encourage the sharing among effects, we assume that the coefficient tensor $\tB$ lies in a low-dimensional parameter space:
+To encourage the sharing among effects, decomposition~\eqref{eq:factor} assumes that the coefficient tensor $\tB$ lies in a low-dimensional parameter space.
 %\begin{equation}\label{eq:rank}
@@
 The low-rank assumption is plausible in many scientific applications.
-In brain imaging analysis, for instance, it is often believed that the brain nodes can be grouped into fewer communities, and the numbers of communities are much smaller than the number of nodes.
-The low-rank structure encourages the shared information across tensor entries, thereby greatly improving the estimation stability.
+In the Example~\ref{example:brain} of brain imaging analysis, for instance, it is often believed that the brain nodes can be grouped into fewer communities, and the numbers of communities are much smaller than the number of nodes.
+{\color{red}The low-rank structure encourages the shared information across tensor entries, thereby greatly improving the estimation stability.
+}
 %When no confusion arises, we drop the subscript $(r_1,\ldots,r_K)$ and write $\tP$ for simplicity.
 
-Our supervised tensor decomposition~\eqref{eq:decomp} is able to incorporate any arbitrary number of feature matrices.
-When certain mode $k$ has no available side information, we set $\mX_k=\mI_{d_k}$.
-In particular, our model~\eqref{eq:decomp} reduces to classical unsupervised tensor decomposition~\citep{de2000multilinear,hong2018generalized} when no side information is available at all; i.e., $\mX_k=\mI_{d_k}$ for all $k\in[K]$.
+Our supervised tensor decomposition~\eqref{eq:decomp} is able to incorporate arbitrary numbers of feature matrices.
+We set $\mX_k=\mI_{d_k}$ if certain mode $k$ has no available side information.
+In particular, our model~\eqref{eq:decomp} reduces to classical unsupervised tensor decomposition~\citep{de2000multilinear,hong2018generalized} if no side information is available; i.e., $\mX_k=\mI_{d_k}$ for all $k\in[K]$.
 %Then, the final form of our tensor regression model can be written as:
@@
 In most applications, the rank is much smaller than tensor dimension $d_k$ and the feature dimension $p_k$.
-In this case, the matrices $\mM_k$ serve the role of simultaneous dimension reduction of the data tensor and features.
-The key implications of the sufficient features $\mX_k\mM_k$ can be seen in the following two conditional independent assumptions in the model~\eqref{eq:decomp},
+In such a case, the matrices $\mM_k$ serve the role of simultaneous dimension reduction of the data tensor and features.
+The implications of the sufficient features $\mX_k\mM_k$ can be seen in the following two conditional independence assumptions in model~\eqref{eq:decomp},
 \begin{align}
-\tY \ci \{\mX_k\}&\ \big|\ \{\mX_k\mM_k\} \ \ \text{\small (independence between the tensor and multiple features)},\\
-y_\omega \ci y_{\omega'} & \ \big|\ \{\mX_k\mM_k\}  \ \ \text{\small (independence within the tensor)},
+\tY \ci \{\mX_k\}&\ \big|\ \{\mX_k\mM_k\} \ \ \text{(independence between the tensor and multiple features)},\\
+y_\omega \ci y_{\omega'} & \ \big|\ \{\mX_k\mM_k\}  \ \ \text{(independence within the tensor)},
 \end{align}
 where the second line holds for all $\omega\neq \omega' \in[d_1]\times\cdots\times[d_K]$, and $\ci$ denotes the independence.
-The first property highlights the ``decorrelation'' role of $\mM_k$ akin to the sufficient dimension reduction in univariate supervised learning~\citep{adragni2009sufficient}, whereas the second property highlights the tensor dimension reduction in the usual unsupervised sense (consider, for example, $\mX_k=\mathbf{I}_{d_k}$ for all $k\in[K]$).
-
-Denote $d_{\text{total}}=\prod_k d_k$ and $p_{\text{total}}=\prod_k p_k$.
-Let $\tX=\mX_1\otimes \cdots \otimes \mX_K \in \mathbb{R}^{d_{\text{total}} \times p_{\text{total}}}$ be the tensor product of feature spaces.
-....
-$\text{Unfold}(\tX) \in\mathbb{R}^{d_{\text{total}}\times p_{\text{total}}}$.
-\[
-\mathbb{E}(\text{vec}(\tY)|\mX,\mZ,\mW) =\beta_{111}\mx^1\mz^1\mw^1+\beta_{121}\mx^1\mz^2\mw^1+\beta_{112}\mx^1\mz^1\mw^2+\cdots+\beta_{222}\mx^2\mz^2\mw^2.
-\]
-where each $\mM_1\mX=[\mx^1,\mx^2]$, $\mM_2\mZ=[\mz^1,\mz^2]$, $\mM_3\mW=[\mw^1,\mw^2]$ are sufficient features.  ....Full interaction model with rank-1 features.
+The first property highlights the ``decorrelation'' role of $\mM_k$ akin to the sufficient dimension reduction~\citep{adragni2009sufficient} in high-dimensional supervised learning, whereas the second property highlights the tensor dimension reduction in the usual unsupervised sense (consider, for example, $\mX_k=\mathbf{I}_{d_k}$ for all $k\in[K]$).
+
+Our model also has a close connection with tensor-on-tensor regression.
+Model~\eqref{eq:decomp} can be viewed as a multivariate regression model, where the response is vectorized tensor entries and the covariates are interactions between all sufficient features across modes.
+We give an Gaussian model on order-3 tensors for illustration.
+Let $\mX,\mZ,\mW$ denote the feature matrix on mode $k=1, 2, 3$, respectively.
+Suppose each mode has two-dimensional sufficient features, denoted $\mM_1\mX=[\mx_1,\mx_2]$, $\mM_2\mZ=[\mz_1,\mz_2]$, $\mM_3\mW=[\mw_1,\mw_2]$.
+Here $\mx_1,\mx_2,\ldots,\mw_1, \mw_2$ are column vectors.
+Then the model~\eqref{eq:decomp} can be written as
+\[
+\mathbb{E}(y_{ijk}|\mX,\mZ,\mW)=c_{111}\mx_{1i}\mz_{1j}\mw_{1k}+c_{121}\mx_{i1}\mz_{j2}\mw_{k1}+\cdots+c_{221}\mx_{2i}\mz_{2j}\mw_{1k}+c_{222}\mx_{2i}\mz_{2j}\mw_{2k},
+\]
+where $\entry{c_{ijk}} \in \mathbb{R}^{2\times 2\times 3}$ are unknown regression parameters, and $\mx_{1i}$ denotes the $i$-th entry in the vector $\mx_1$, and similar notation for other covariates.
+In this perspective, our supervised tensor decomposition is useful for identifying across-mode interactions between sufficient features.
+
+%All in all, these two connection interprete.
+our method from two perspectives, one from the residual structure conditional on $\{\mX_k\mM_k\}$ and the interaction between $\mX_k\mM_k$ across $k\in[K]$.
+These two perspetive, in total, give rise the predictive prediction in our model.
 
@@
 \subsection{Statistical properties}\label{subsec:statprob}
-We assess the estimation accuracy using the deviation in the Frobenius norm. For the true coefficient tensor $\trueB$ and its estimator $\hat \tB$, define
+We assess the estimation accuracy using the deviation in the Frobenius norm.
+Add angle deviation in dimension reduction matrices.
+Add identifiability condition.
+For the true coefficient tensor $\trueB$ and its estimator $\hat \tB$, define
 \[
@@
 Consider an order-$K$ dimensional-$(d_1,\ldots,d_K)$ tensor $\tY$ and multiple feature matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$, for $k\in[K]$.
-Suppose that, conditional on the features, the entries in $\tY$ are independent random variables following an exponential family distribution with conditional mean model~\eqref{eq:decomp}.
-Let $r_\text{total}=\prod_k r_k$, $r_{\max}=\max_k r_k$, and $\lambda_{\min}(\mM_k)$ be the minimal non-zero singular value of $\mM_k$.
-Assume that $\lambda_{\min}(\mM_k)>0$ and $\FnormSize{}{\mX_k} \asymp \Omega(\sqrt{d_k})$, for all $k\in[K]$.
+Suppose that, conditional on the features, the entries in $\tY$ are independent realizations from an exponential family distribution with conditional mean model~\eqref{eq:decomp}.
+Let $r_\text{total}=\prod_k r_k$, $r_{\max}=\max_k r_k$, and $\sigma_{\min}(\mM_k)$ be the minimal non-zero singular value of $\mM_k$.
+Assume that $\sigma_{\min}(\mM_k)>0$ and $\FnormSize{}{\mX_k} \asymp \Omega(\sqrt{d_k})$, for all $k\in[K]$.
 Under Assumption~\ref{ass}, there exist two constants $C_1, C_2>0$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$,

### 2020-09-08 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
 Define a feature matrix $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$.
-Then, the model~\eqref{eq:network} can be written as
+Then, the model~\eqref{eq:network} has the equivalent representation of supervised tensor decomposition,
 \[
@@
 Our model also has a close connection with tensor-on-tensor regression.
-Model~\eqref{eq:decomp} can be viewed as a multivariate regression model, where the response is vectorized tensor entries and the covariates are interactions between all sufficient features across modes.
-We give an Gaussian model on order-3 tensors for illustration.
+Specifically, model~\eqref{eq:decomp} can be viewed as a multivariate regression model, where the response is vectorized tensor entries and the covariates are interactions between all sufficient features across modes.
+We take an order-3 tensor under the Gaussian model for illustration.
 Let $\mX,\mZ,\mW$ denote the feature matrix on mode $k=1, 2, 3$, respectively.
@@
 Here $\mx_1,\mx_2,\ldots,\mw_1, \mw_2$ are column vectors.
-Then the model~\eqref{eq:decomp} can be written as
-\[
+Then the model~\eqref{eq:decomp} can be written as a regression model with across-mode interactions,
+\begin{equation}\label{eq:interaction}
 \mathbb{E}(y_{ijk}|\mX,\mZ,\mW)=c_{111}\mx_{1i}\mz_{1j}\mw_{1k}+c_{121}\mx_{i1}\mz_{j2}\mw_{k1}+\cdots+c_{221}\mx_{2i}\mz_{2j}\mw_{1k}+c_{222}\mx_{2i}\mz_{2j}\mw_{2k},
-\]
-where $\entry{c_{ijk}} \in \mathbb{R}^{2\times 2\times 3}$ are unknown regression parameters, and $\mx_{1i}$ denotes the $i$-th entry in the vector $\mx_1$, and similar notation for other covariates.
-In this perspective, our supervised tensor decomposition is useful for identifying across-mode interactions between sufficient features.
-
-%All in all, these two connection interprete.
-our method from two perspectives, one from the residual structure conditional on $\{\mX_k\mM_k\}$ and the interaction between $\mX_k\mM_k$ across $k\in[K]$.
-These two perspetive, in total, give rise the predictive prediction in our model.
+\end{equation}
+where $\entry{c_{ijk}} \in \mathbb{R}^{2\times 2\times 3}$ are unknown interaction effects, $\mx_{1i}$ denotes the $i$-th entry in the feature vector $\mx_1$, and similar notations apply to other features.
+Model~\eqref{eq:interaction} suggests the use of our supervised tensor decomposition for identifying across-mode interactions.
+Note that lower-order interactions are naturally incorporated in~\eqref{eq:interaction} if we include an intercept column in the sufficient feature matrices.
 

### 2020-09-08 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
 Let $\mX,\mZ,\mW$ denote the feature matrix on mode $k=1, 2, 3$, respectively.
-Suppose each mode has two-dimensional sufficient features, denoted $\mM_1\mX=[\mx_1,\mx_2]$, $\mM_2\mZ=[\mz_1,\mz_2]$, $\mM_3\mW=[\mw_1,\mw_2]$.
+Suppose that each mode has two-dimensional sufficient features, denoted $\mM_1\mX=[\mx_1,\mx_2]$, $\mM_2\mZ=[\mz_1,\mz_2]$, $\mM_3\mW=[\mw_1,\mw_2]$.
 Here $\mx_1,\mx_2,\ldots,\mw_1, \mw_2$ are column vectors.
@@
 Note that lower-order interactions are naturally incorporated in~\eqref{eq:interaction} if we include an intercept column in the sufficient feature matrices.
+
+\fixme{Miaoyan}{do we make use of the above two implications in the real data analysis or simulations?}
 
@@
 \subsection{Statistical properties}\label{subsec:statprob}
-We assess the estimation accuracy using the deviation in the Frobenius norm.
-Add angle deviation in dimension reduction matrices.
-Add identifiability condition.
-For the true coefficient tensor $\trueB$ and its estimator $\hat \tB$, define
+We assess the estimation accuracy using the deviation in the Frobenius norm. \fixme{Miaoyan}{needs revision...} For the true coefficient tensor $\trueB$ and its estimator $\hat \tB$, define
 \[
@@
 Let $r_\text{total}=\prod_k r_k$, $r_{\max}=\max_k r_k$, and $\sigma_{\min}(\mM_k)$ be the minimal non-zero singular value of $\mM_k$.
-Assume that $\sigma_{\min}(\mM_k)>0$ and $\FnormSize{}{\mX_k} \asymp \Omega(\sqrt{d_k})$, for all $k\in[K]$.
+Assume that $\sigma_{\min}(\mM_k)>0$ and $\FnormSize{}{\mX_k} \asymp \Omega(\sqrt{d_k})$ as $d_k\to\infty$, for all $k\in[K]$.
 Under Assumption~\ref{ass}, there exist two constants $C_1, C_2>0$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$,
@@
 Theorem~\ref{thm:main} establishes the statistical convergence for the estimator~\eqref{eq:MLE}.
-\fixme{Miaoyan}{Wrong statement based on current proof}Actually, the proof in section~\ref{sec:appedix} shows that the statistically optimal rate holds, not only for the MLE~\eqref{eq:MLE},  but also any local estimators $\hat \tB$ in the level set $\{\hat \tB \in \tP \colon \tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true}) \}$ satisfies the upper bound~\ref{thm:main}.
+Actually, the proof in section~\ref{sec:appedix} shows that the statistically optimal rate holds, not only for the MLE~\eqref{eq:MLE}, but also any local estimators $\hat \tB$ in the level set $\{\hat \tB \in \tP \colon \tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true}) \}$ satisfies the upper bound~\ref{thm:main}\fixme{Miaoyan}{Wrong statement based on current proof}.
 In Section~\ref{subsec:alg}, we implement empirical studies to verify the algorithmic stability to find a local optimum.
@@
 \section{Numerical experiments}\label{sec:simulation}
-{\color{red} Simulation needs to be re-done, since code has been changed. }
+{\color{red} Simulation/figures needs to be re-run once more, since code has been changed. }
 
@@
 
-\section{Data analysis}\label{sec:data}
+\section{Data analyses}\label{sec:data}
 We apply our tensor regression model to two datasets.
@@
 \section{Conclusion}\label{sec:con}
-
+\fixme{Miaoyan}{No more than 35 pages. Reserve at least 1.5 pages for Conclusion}
 We have developed a generalized tensor regression with features on multiple modes.
@@
 
- Add three more paragraphs:
-
- 1. Optimization 
+ \fixme{Miaoyan}{Add two more paragraphs: 
  %Our method leverages on the alternating updating algorithm for the optimization.
@@
 We leave the pursuit of optimization landscape as future research.
-
 %Another an..
@@
 For random effects.....
-
-2. Assumption on (conditional) independence assumption on the tensor entries.
-
+2.
+Assumption on (conditional) independence assumption on the tensor entries.
+Mixed effect model, tensor-variate normal distribution..Current approach is to use fixed factor; or a combination of fixed factor + random effects.
 3. Inference. ... uncertainly quantification... bootstrap ....  Testing for Interaction + main effects. 
+}
 

### 2020-09-08 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
 Consider the neuroimaging analysis as an example.
-The scalar-to-tensor regression aims to predict the individual response based on brain connectivity network, while tensor-to-scalar model focuses on understanding the variation in the brain connectivity attributable to the individual features such as age, gender, and disease status.
+The scalar-to-tensor regression aims to predict the people's disease status based on brain connectivity network, while tensor-to-scalar model focuses on understanding the variation in the brain connectivity attributable to the individual features such as age, gender, and disease status.
 The scalar-to-tensor model is insufficient for tensor dimension reduction because it lacks the modeling of stochastic noise in the data tensor.
@@
 where the parameters of interest are $\mM_k$ and $\tC$.
-Note that model~\eqref{eq:decomp} assumes a fixed, known rank $\mr=(r_1,\ldots,r_K)$; the adaptation to unknown rank will be described in Section~\ref{sec:tuning}.
+Note that model~\eqref{eq:decomp} assumes a fixed, known rank $\mr=(r_1,\ldots,r_K)$; the adaptation to unknown rank will be addressed in Section~\ref{sec:tuning}.
 Figure~\ref{fig:intro1}b provides a schematic illustration of our model.
@@
 
-In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by available features.
-The regression coefficient $\tB$ in our model model~\eqref{eq:tensormodel} serves this goal by collecting the feature effects and the interaction thereof.
-To encourage the sharing among effects, decomposition~\eqref{eq:factor} assumes that the coefficient tensor $\tB$ lies in a low-dimensional parameter space.
+In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by features. 
+%The core tensor $\tC$ in our model model~\eqref{eq:tensormodel} serves this goal by collecting the sufficient feature effects and the interaction thereof.  
+\fixme{Miaoyan}{Unclear... To encourage the sharing among effects, decomposition~\eqref{eq:factor} assumes that the feature effects $\tB$ lie in a low-dimensional parameter space.
 %\begin{equation}\label{eq:rank}
@@
 The low-rank assumption is plausible in many scientific applications.
-In the Example~\ref{example:brain} of brain imaging analysis, for instance, it is often believed that the brain nodes can be grouped into fewer communities, and the numbers of communities are much smaller than the number of nodes.
-{\color{red}The low-rank structure encourages the shared information across tensor entries, thereby greatly improving the estimation stability.
+In the Example~\ref{example:brain} of brain imaging analysis, for instance, it is often believed that the brain nodes can be partitioned into several functional and structural regions, and the person's attribute effect the connectivity in a similar way within a region.
 }
+
+%{\color{red}The low-rank structure encourages the shared information across tensor entries, thereby greatly improving the estimation stability. }
 %When no confusion arises, we drop the subscript $(r_1,\ldots,r_K)$ and write $\tP$ for simplicity.
@@
 An important feature of our method is that we allow high-dimensionality in both tensor dimension $d_k$ and feature dimension $p_k$.
-In most applications, the rank is much smaller than tensor dimension $d_k$ and the feature dimension $p_k$.
+The model rank is typically smaller than tensor dimension $d_k$ and the feature dimension $p_k$.
 In such a case, the matrices $\mM_k$ serve the role of simultaneous dimension reduction of the data tensor and features.
@@
 Our model also has a close connection with tensor-on-tensor regression.
-Specifically, model~\eqref{eq:decomp} can be viewed as a multivariate regression model, where the response is vectorized tensor entries and the covariates are interactions between all sufficient features across modes.
+Specifically, model~\eqref{eq:decomp} can be viewed as a multivariate regression model, where the response is vectorized tensor and the covariates are interactions between all sufficient features across modes.
 We take an order-3 tensor under the Gaussian model for illustration.
@@
 
-{\color{red}We assume that we have an additional information on an upper bound $\alpha>0$ such that $\mnormSize{}{\Theta}\leq \alpha$.} We propose a constrained maximum likelihood estimator (M-estimator)
+We propose a constrained maximum likelihood estimator (M-estimator)
 %\begin{align} \label{eq:MLE} 
@@
 %\end{align}
-
 \begin{align} \label{eq:MLE} 
@@
 \]
-
-In the following theoretical analysis, we assume the rank $\mr=(r_1,\ldots,r_K)$ is known and fixed. The adaptation of unknown $\mr$ will be addressed in Section~\ref{sec:tuning}.
+The entry-wise bound assumption is a technical condition that avoids the divergence in non-Gaussian variance.
+
+%In the following theoretical analysis, we assume the rank $\mr=(r_1,\ldots,r_K)$ is known and fixed. The adaptation of unknown $\mr$ will be addressed in Section~\ref{sec:tuning}.
 
@@
 The decision variables in the objective function~\eqref{eq:MLE} consist of $K+1$ blocks of variables, one for the core tensor $\tC$ and $K$ for the factor matrices $\mM_k$'s.
-We notice that, if any $K$ out of the $K+1$ blocks of variables are known, then the optimization with respect to the last block of variables reduced to a simple GLM.
-This observation suggests that we can iteratively update one block at a time while keeping others fixed.
+We notice that, if any $K$ out of the $K+1$ blocks of variables are known, then the optimization reduces to a simple GLM with respect to the last block of variables.
+This observation suggests an iterative updating scheme for one block at a time while keeping others fixed.
 After each iteration, we rescale the core tensor $\tC^{(t+1)}$ subject to the maximum norm constraint.
@@
 
-Note that the feasible set $\tP$ is a non-convex set.
-Therefore, the optimization~\eqref{eq:MLE} is a non-convex problem, and the Algorithm~\ref{alg:B} usually not theoretically possesses the global optimality.
+The optimization~\eqref{eq:MLE} is a non-convex problem due to the non-convexity in the feasible set $\tP$.
+Global optimality of non-convex optimization of this type is often challenging to obtain.
+However, as mentioned in Section~\ref{subsec:statprob}, the desired convergence rate~\eqref{eq:bound} holds for the valid estimators satisfying $\tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true})$, which indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective of local optimums are large enough.
 \fixme{Miaoyan}{This is wrong.
 Proof of Theorem 4.1 is based on global optimality condition.
-Please relax the proof.} However, as mentioned in Section~\ref{subsec:statprob}, the desired convergence rate~\eqref{eq:bound} holds for the valid estimators satisfying $\tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true})$, which indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective of local optimums are large enough.
-Fortunately, we find the  Algorithm~\ref{alg:B}  often gives a satisfactory convergence point $\hat \tB$, when we initialize the parameters via continuous-valued Tucker decomposition.
+Please relax the proof.} Fortunately, we find the  Algorithm~\ref{alg:B}  often gives a satisfactory convergence point $\hat \tB$, when we initialize the parameters via continuous-valued Tucker decomposition.
 Figure~\ref{fig:loglike} shows the trajectory of the objective function for order-3 tensors under the balanced setting, where $\alpha = 10, d_k = d, p_k = 0.4d, r_k = r$ for $ k = 1,2,3, p \in \{25,30\}$  and $r \in \{3, 6\}$.
@@
 We consider order-3 tensors with a range of distribution types.
-The coefficient tensor $\tB$ is generated using the factorization form~\eqref{eq:tucker} where both the core and factor matrices are drawn i.i.d.\ from Uniform[-1,1].
+The coefficient tensor $\tB$ is generated using the factorization form~\eqref{eq:factor} where both the core and factor matrices are drawn i.i.d.\ from Uniform[-1,1].
 The linear predictor is then simulated from $\tU=\tB\times\{\mX_1,\mX_2,\mX_3\}$, where $\mX_k$ is either an identity matrix (i.e.\ no feature available) or Gaussian random matrix with i.i.d.\ entries from $N(0,\sigma_k^2)$.
@@
 
+\fixme{miaoyan}{references need update..}
+
 \bibliography{tensor_wang}

### 2020-09-10 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
 
-\fixme{Miaoyan}{do we make use of the above two implications in the real data analysis or simulations?}
+\fixme{Miaoyan}{The above two aspects above two aspects in the simulation and real data analysis.}
 
@@
 
-We propose a constrained maximum likelihood estimator (M-estimator)
-%\begin{align} \label{eq:MLE} 
-%	\hat \tB &=\argmax_{\hat \tB \colon \text{rank}(\tB)\leq \mr, \mnormSize{}{\Theta(\tB)}\leq \alpha  } \tL_{\tY}(\tB).
-%\end{align}
+We consider a constrained maximum likelihood estimator (M-estimator),
 \begin{align} \label{eq:MLE} 
@@
 \INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, feature matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,\ldots,K$, target Tucker rank $\mr=(r_1,\ldots,r_K)$, link function $f$, maximum norm bound $\alpha$
-\OUTPUT Low-rank estimation for the coefficient tensor $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$. 
-\State Calculate $\check \tB=\tY\times_1 \left[(\mX_1^T\mX_1)^{-1}\mX^T_1\right] \times_2\cdots\times_K\left[(\mX_K^T\mX_K)^{-1}\mX^T_K\right] $.
-\State Initialize the iteration index $t=0$.
+\OUTPUT Estimated core tensor $\hat \tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ and factor matrices $\hat \mM_k\in\mathbb{R}^{p_k\times r_k}$. 
+%\State Calculate $\check \tB=\tY\times_1 \left[(\mX_1^T\mX_1)^{-1}\mX^T_1\right] \times_2\cdots\times_K\left[(\mX_K^T\mX_K)^{-1}\mX^T_K\right] $.
+%\State Initialize the iteration index $t=0$.
 Initialize the core tensor $\tC^{(0)}$ and factor matrices $\mM^{(0)}_k\in\mathbb{R}^{p_k\times r_k}$ via rank-$\mr$ Tucker approximation of $\check\tB$, in the least-square sense.
-  \While {the relative increase in objective function $\tL_\tY(\tB)$ is less than the tolerance}
-\State Update iteration index $t \leftarrow t+1$.
+\State Random initialization of the core tensor $\tC$ and factor matrices $\mM_k$. 
+ % \While {the relative increase in objective function $\tL_\tY(\tB)$ is less than the tolerance}
+\While{Do until convergence}
+%\State Update iteration index $t \leftarrow t+1$.
 \For { $k=1$ to $K$}
-\State Obtain the factor matrix $\tilde \mM_k^{(t+1)}\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$. 
-\State Perform QR factorization on $\tilde \mM_k^{(t+1)}=\mQ\mR$, where $\mQ\in\mathbb{R}^{p_k\times r_k}$ has orthogonal columns. 
-\State Update $\mM_k^{(t+1)}\leftarrow \mQ$ and core tensor $\tC^{(t+1)}\leftarrow \tC^{(t+1)}\times_k \mR$.
+\State Obtain the factor matrix $\tilde \mM_k\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$. 
+\State Perform QR factorization $\tilde \mM_k=\mQ\mR$, where $\mQ\in\mathbb{R}^{p_k\times r_k}$ consists of orthogonal columns. 
+\State Update $\mM_k\leftarrow \mQ$ and core tensor $\tC\leftarrow \tC\times_k \mR$.
 \EndFor
-\State Obtain the core tensor $\tC^{(t+1)}\in\mathbb{R}^{r_1\times\cdots\times r_K}$ by solving a GLM with $\text{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM^{(t)}_k]$ as features, and $f$ as link function.
+\State Update the core tensor $\tC$ by solving a GLM with $\text{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM_k]$ as features, and $f$ as link function.
 Here $\otimes$ denotes the kronecker product of matrices.
 \State Rescale the core tensor subject to the maximum norm constraint. 
-\State Update $\tB^{(t+1)}\leftarrow \tC^{(t+1)}\times_1\mM_1^{(t+1)}\times_2\cdots\times_K\mM_K^{(t+1)}$.
+%\State Update $\tB^{(t+1)}\leftarrow \tC^{(t+1)}\times_1\mM_1^{(t+1)}\times_2\cdots\times_K\mM_K^{(t+1)}$.
 \EndWhile
@@
 Please relax the proof.} Fortunately, we find the Algorithm~\ref{alg:B}  often gives a satisfactory convergence point $\hat \tB$, when we initialize the parameters via continuous-valued Tucker decomposition.
-Figure~\ref{fig:loglike} shows the trajectory of the objective function for order-3 tensors under the balanced setting, where $\alpha = 10, d_k = d, p_k = 0.4d, r_k = r$ for $ k = 1,2,3, p \in \{25,30\}$  and $r \in \{3, 6\}$.
+Figure~\ref{fig:loglike} shows typical examples of objective trajectories for order-3 tensors under the model~\eqref{eq:decomp}, where $d_k = d \in\{20,30\}$, $p_k = 0.4d$, $r_k = r\in\{3,6\}$ for $ k = 1,2,3$.
 We consider the inputs with Gaussian, Bernoulli, and Poisson entries.
@@
 Suppose that, conditional on the features, the entries in $\tY$ are independent realizations from an exponential family distribution with conditional mean model~\eqref{eq:decomp}.
-Let $r_\text{total}=\prod_k r_k$, $r_{\max}=\max_k r_k$, and $\sigma_{\min}(\mM_k)$ be the minimal non-zero singular value of $\mM_k$.
+Let $r_\text{total}=\prod_k r_k$, $r_{\max}=\max_k r_k$, and $\sigma_{\min}(\mM_k)$ be the smallest singular value of $\mM_k$.
 Assume that $\sigma_{\min}(\mM_k)>0$ and $\FnormSize{}{\mX_k} \asymp \Omega(\sqrt{d_k})$ as $d_k\to\infty$, for all $k\in[K]$.
@@
 \begin{equation}\label{eq:bound}
-\textup{sin}^2 \Theta(\mM_k,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\lambda_{\min}(\mM_k)}\frac{\sum_k p_k}{\prod_k d_k}, \quad\text{for all }k\in[K],
+\textup{sin}^2 \Theta(\mM_k,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\sigma^2_{\min}(\mM_k)}\frac{\sum_k p_k}{\prod_k d_k}, \quad\text{for all }k\in[K],
 \end{equation}
@@
 \section{Numerical experiments}\label{sec:simulation}
-{\color{red} Simulation/figures needs to be re-run once more, since code has been changed. }
-
-We evaluate the empirical performance of our generalized tensor regression through simulations.
-We consider order-3 tensors with a range of distribution types.
-The coefficient tensor $\tB$ is generated using the factorization form~\eqref{eq:factor} where both the core and factor matrices are drawn i.i.d.\ from Uniform[-1,1].
-The linear predictor is then simulated from $\tU=\tB\times\{\mX_1,\mX_2,\mX_3\}$, where $\mX_k$ is either an identity matrix (i.e.\ no feature available) or Gaussian random matrix with i.i.d.\ entries from $N(0,\sigma_k^2)$.
+We evaluate the empirical performance of our generalized tensor regression through simulations. We consider order-3 tensors with a range of distribution types. 
+Unless otherwise specified, the conditional mean tensor is generated form model~\eqref{eq:decomp}, where the core tensor entries are i.i.d.\ drawn from Uniform[-1,1], the factor matrix $\mM_k$ is uniformly sampled (with respect to Haar measure) from matrices with orthonormal columns, and the feature matrix $\mX_k$ is either an identity matrix (i.e.\ no feature available) or Gaussian random matrix with i.i.d.\ entries from $N(0,\sigma_k^2)$.
 We set $\sigma_k=d_k^{-1/2}$ to ensure the singular values of $\mX_k$ are bounded as $d_k$ increases.
-The $\tU$ is scaled such that $\mnormSize{}{\tU}=1$.
-Conditional on the linear predictor $\tU=\entry{u_{ijk}}$, the entries in the tensor $\tY=\entry{y_{ijk}}$ are drawn independently according to one of the following three probabilistic models:
+The linear predictor $\Theta=\tC\times\{\mM_1\mX_1,\ldots,\mM_3\mX_3\}$ is scaled such that $\mnormSize{}{\Theta}=1$.
+Conditional on the linear predictor $\Theta=\entry{\theta_{ijk}}$, the entries in the tensor $\tY=\entry{y_{ijk}}$ are drawn independently according to one of the following three probabilistic models:
 
 \begin{enumerate}
-\item[(a)] (Gaussian). Continuous data $y_{ijk}\sim N\left(\alpha u_{ijk}, 1\right)$.
-\item[(b)] (Poisson). Count data $y_{ijk}\sim\text{Poi}\left( e^{\alpha u_{ijk}}\right)$.
-\item[(c)] (Bernoulli). Binary data $y_{ijk}\sim \text{Ber}\left( {e^{\alpha u_{ijk}} \over 1+e^{\alpha u_{ijk}}}\right)$.
+\item[(a)] (Gaussian). Continuous data $y_{ijk}\sim N\left(\alpha \theta_{ijk}, 1\right)$, for all $(i,j,k)\in[d_1]\times[d_2]\times[d_3]$.
+\item[(b)] (Poisson). Count data $y_{ijk}\sim\text{Poi}\left( e^{\alpha \theta_{ijk}}\right)$, for all $(i,j,k)\in[d_1]\times[d_2]\times[d_3]$.
+\item[(c)] (Bernoulli). Binary data $y_{ijk}\sim \text{Ber}\left( {e^{\alpha \theta_{ijk}} \over 1+e^{\alpha \theta_{ijk}}}\right)$, for all $(i,j,k)\in[d_1]\times[d_2]\times[d_3]$.
 \end{enumerate}
 Here $\alpha>0$ is a scalar controlling the magnitude of the effect size.
-In each simulation study, we report the mean squared error (MSE) for the coefficient tensor averaged across $n_{\text{sim}}=30$ replications.
+In each simulation study, we report the mean squared error (MSE) for the coefficient tensor averaged across $n_{\text{sim}}=20$ replications.
 
@@
 We consider the balanced situation where $d_k=d$, $p_k=0.4d_k$ for $k=1,2,3$.
-We set $\alpha=10$ and consider various combinations of dimension $d$ and rank $\mr=(r_1,r_2,r_3)$.
-For each combination, we simulate tensor data following Gaussian, Bernoulli, and Poisson models.
+We set $\alpha=4$ and consider various combinations of dimension $d$ and rank $\mr=(r_1,r_2,r_3)$.
+For each combination, we simulate tensor data following Gaussian and Poisson models.
 We then minimize BIC using a grid search over three dimensions.
 The hyper-parameter $\alpha$ is set to infinity in the fitting, which essentially imposes no prior on the coefficient magnitude.
-Table~\ref{tab:rank} reports the selected rank averaged over $n_{\text{sim}}=30$ replicates for Gaussian and Poisson models.
+Table~\ref{tab:rank} reports the selected rank averaged over $n_{\text{sim}}=20$ replicates for Gaussian and Poisson models.
 We found that when $d=20$, the selected rank is slightly smaller than the true rank, and the accuracy improves immediately when the dimension increases to $d=40$.
@@
 \centering
-%\resizebox{1\textwidth}{!}{
 \begin{tabular}{|c|cc|cc|}
@@
 \hline
-$(3,\ 3,\ 3)$&$(2.1,\ 2.0,\ 2.0)$&$({\bf 3},\ {\bf 3},\ {\bf 3})$&$(2.0,\ 2.2,\ 2.1)$&$({\bf 3},\ {\bf 3},\ {\bf 3})$\\
-$(4,\ 4,\ 6)$&$(3.2,\ 3.1,\ 5.0)$&$({\bf 4},\ {\bf 4},\ {\bf 6})$&$({\bf 4.0},\ {\bf 4.0},\ 5.2)$&$({\bf 4},\ {\bf 4},\ {\bf 6})$\\
-$(6,\ 8,\ 8)$&$(5.1,\ 7.0,\ 6.9)$&$({\bf 6},\ {\bf 8},\ {\bf 8})$&$(5.0,\ 6.1,\ 7.1)$&$({\bf 6},\ {\bf 8},\ {\bf 8})$\\
+$(3,\ 3,\ 3)$&$({\bf 3},\ 3,\ {\bf 3})$&$({\bf 3},\ {\bf 3},\ {\bf 3})$&$(2.0,\ 2.2,\ 2.1)$&$({\bf 3},\ {\bf 3},\ {\bf 3})$\\
+$(4,\ 4,\ 6)$&$(3,\ 3,\ {\bf 6})$&$({\bf 4},\ {\bf 4},\ 5)$&$({\bf 4.0},\ {\bf 4.0},\ 5.2)$&$({\bf 4},\ {\bf 4},\ {\bf 6})$\\
+$(6,\ 8,\ 8)$&$(5,\ 7,\ 7)$&$({\bf 6},\ {\bf 8},\ {\bf 8})$&$(5.0,\ 6.1,\ 7.1)$&$({\bf 6},\ {\bf 8},\ {\bf 8})$\\
 \hline
 \end{tabular}
-%}
 \caption{Rank selection via BIC. Bold number indicates no significant difference between the estimate and the ground truth, based on a $z$-test with a level 0.05.}\label{tab:rank}
@@
 The second experiment evaluates the accuracy when features are available on all modes.
-We set $\alpha=10, d_k=d, p_k=0.4d_k, r_k=r\in\{2,4,6\}$ and increase $d$ from 25 to 50.
+We set $\alpha=10, d_k=d, p_k=0.4d_k, r_k=r\in\{2,4,6\}$ and increase $d$ from 30 to 60.
 Our theoretical analysis suggests that $\hat \tB$ has a convergence rate $\tO(d^{-2})$ in this setting.
@@
 A classical GLM is to regress the dyadic edges, one at a time, on the features, and this model is repeatedly fitted for each edge.
-This repeated approach, however, does not account for the correlation among the edges, and may suffer from overfitting.
-As we can see in Figure~\ref{fig:glm}, out tensor regression method achieves significant error reduction in all three models considered.
+As we find in Figure~\ref{fig:glm}, out tensor regression method achieves significant error reduction in all three models considered.
 The outer-performance is significant in the presence of large communities, and even in the less structured case ($\sim 20/15=1.33$ nodes per block), our method still outer-performs GLM.
-This is because the low-rankness in our modeling automatically identifies the shared information across entries.
+The possible reason is that the massive GLM approach does not account for the correlation among the edges, and suffer from overfitting.
+In contrast, the low-rankness in our modeling automatically identifies the shared information across entries.
 By selecting the rank in a data-driven way, our method is able to achieve accurate estimation with improved interpretability.
@@
 \subsection{Comparison with alternative methods}
-
 We compare our generalized tensor regression ({\bf GTR}) with three other supervised tensor methods:
@@
 
-Figure~\ref{fig:compare} shows the averaged prediction error across 30 replicates.
+Figure~\ref{fig:compare} shows the averaged prediction error across 20 replicates.
 We see that our {\bf GTR} outperforms others, especially in the high-rank high-noise setting.

### 2020-09-11 tensor_regress/jasa-latexsource/tensor-jasa.tex
update

@@
 \newcommand{\blind}{1}
+
+Have people asked about the inverse question; that is, what kind of minimal distribution assumptions do we have to impose in order to have vanishing confidence interval lenth?
 
@@
 \begin{equation}\label{eq:logbefore}
-\tL_{\tY}(\tB)=\tL_{\tY}(\trueB)+\Big\langle {\partial\tL_{\tY}(\tB) \over \tB}\big|_{\tB=\trueB},\tB-\trueB \Big\rangle + {1\over 2}\text{vec}(\tB-\trueB)^T \tH(\check \tB)\text{vec}(\tB-\trueB),
-\end{equation}
-where $\tH_{\tY}(\check \tB)$ is the (non-random) Hession of ${\partial \tL_{\tY}^2 (\tB)\over\partial^2 \tB}$ evaluated at $\check \tB =\alpha \text{vec}(\alpha \tB+(1-\alpha)\trueB)$ for some $\alpha\in[0,1]$.
+\tL_{\tY}(\tB)=\tL_{\tY}(\trueB)+\Big\langle {\partial\tL_{\tY}(\tB) \over \partial
+\tB}\big|_{\tB=\trueB},\tB-\trueB \Big\rangle + {1\over 2}\text{vec}(\tB-\trueB)^T \tH(\check \tB)\text{vec}(\tB-\trueB),
+\end{equation}
+where $\tH_{\tY}(\check \tB)$ is the (non-random) Hession of ${\partial \tL_{\tY}^2 (\tB)\over\partial^2 \tB}$ evaluated at $\check \tB = \text{vec}(\alpha \tB+(1-\alpha)\trueB)$ for some $\alpha\in[0,1]$.
 Note that we have $\mathbb{E}\left({\partial\tL_{\tY}(\tB) \over \partial \tB}\big|_{\tB=\trueB}\right)=0$.
@@
 \begin{align}
-\mathbb{E}(\exp(t\varepsilon|\theta)&=\int c(x) \exp\left({\theta x - b(\theta)\over \phi}   \right)\exp \left(t(x-b'(\theta))\right)dx\\
+\mathbb{E}(\exp(t\varepsilon|\theta))&=\int c(x) \exp\left({\theta x - b(\theta)\over \phi}   \right)\exp \left(t(x-b'(\theta))\right)dx\\
 &=\int c(x)\exp \left( {(\theta + \phi t)x - b (\theta+\phi t)+b(\theta+\phi t)-b(\theta)-\phi t b'(\theta) \over \phi}\right)dx\\

### 2020-09-11 tensor_regress/code_JASA/Figure2.R

@@ -1,0 +1,213 @@
+#### code for Figure 2. Log-likelihood trajectory for three models under different (d,r) combinations
+install.packages("tensorregress.tar.gz",repos = NULL, type="source")
+source("simulation.R")
+library(tensorregress)
+library(ggplot2)
+library(patchwork)
+
+
+set.seed(0)## set seed for the entire simulation
+model = c("normal","binary","poisson")
+dimen = c(30,25)
+rank = c(6,3)
+
+################### simulation ##################
+
+lglk_sum = list()
+true_lglk_sum = list()
+time_sum = list()
+
+
+for (i in 1:3) {## three data types
+  lglk = list()
+  true_lglk = c()
+  time = c()
+  k = 1
+  for (j in 1:2) {## two possbiel dimensions
+      for (r in 1:2) {## two possible ranks
+        if(i==1) signal=3
+        else if(i==2) signal=10
+        else if(i==3) signal=6
+      my_data = sim_data(seed = NA, rep(dimen[j],3), rep(rank[r],3), p = rep(floor(0.4*dimen[j]) ,3),dist = model[i], dup = 1,signal = signal)
+      ptm = proc.time()
+      my_reg =  tensor_regress(tsr = my_data$tsr[[1]], X_covar1 = my_data$X_covar1, X_covar2 = my_data$X_covar2,X_covar3 = my_data$X_covar3,rep(rank[r],3), Nsim = 20, cons = "non",dist = model[i],traj_long=TRUE)
+      time1 = proc.time() - ptm
+      time[k] = time1[3] ## running time
+      true_lglk[k] = loglike(my_data$tsr[[1]], my_data$U,dist = model[i])
+      lglk[[k]] = my_reg$lglk
+      k = k+1
+    }
+  }
+  lglk_sum[[i]] = lglk
+  true_lglk_sum[[i]] = true_lglk
+  time_sum[[i]] = time
+  
+}
+
+############ ggplot ##################
+
+#### normal model with d = 30, r = 6,3
+lglk_normal = c(lglk_sum[[1]][[1]] ,lglk_sum[[1]][[2]])
+case_normal=  c(rep("d3r6",33), rep("d3r3",33))
+iter = rep(1:33,times = 2)
+olglk1 = true_lglk_sum[[1]][1]
+olglk2 = true_lglk_sum[[1]][2]
[…]

### 2020-09-11 tensor_regress/code_JASA/Figure3.R

@@ -1,0 +1,102 @@
+#### code for Figure 3. Convergence of MSE with respect to dimension for three models under different ranks
+install.packages("tensorregress.tar.gz",repos = NULL, type="source")
+source("simulation.R")
+library(tensorregress)
+library(ggplot2)
+
+### covariates on all three modes
+d_pre=c(30,35,40,45,50,55,60)
+p_pre=round(0.4*d_pre)
+c_pre=c(2,4,6)
+d_range=cbind(d_pre,d_pre,d_pre)
+p_range=cbind(p_pre,p_pre,p_pre)
+c_range=cbind(c_pre,c_pre,c_pre)
+
+############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line #83 to run the code ##########
+
+#################### plot MSE for normal model ###############
+load("presaved/F1_normal.RData")
+table=F1_normal
+
+MSE_matrix=apply(table[[1]],c(1,2),mean)
+sd_matrix=apply(table[[1]],c(1,2),sd)
+dl=length(d_pre)
+cl=length(c_pre)
+res=cbind(rep(d_pre,cl),rep(p_pre,cl),rep(c_pre,rep(dl,cl)),c(MSE_matrix),c(sd_matrix))
+res=data.frame(res)
+colnames(res)=c("d","p","r","MSE","sd")
+
+fun.1 <- function(x) max(1.2*res[,4]*res[,1]^2)/x
+
+pdf("F1_normal.pdf",width=5,height=4)
+figure=ggplot(res, aes(x =d^3/d, y = MSE)) + geom_line(aes(color = as.factor(r)),size = 1)+geom_point(size=1)+theme(axis.text=element_text(size=12),axis.title=element_text(size=14))+labs(x="effective sample size",y="mean sqaured error (MSE)")
+
+figure = figure + geom_errorbar(aes(ymin=MSE-sd,ymax=MSE+sd),width=0.001,position=position_dodge(0.01))+coord_cartesian(ylim = c(0,0.6))+stat_function(fun = fun.1,linetype = 2)
+figure
+
+
+################## plot MSE for binary model ###############
+load("presaved/F1_binary.RData")
+table=F1_binary
+
+MSE_matrix=apply(table[[1]],c(1,2),mean)
+sd_matrix=apply(table[[1]],c(1,2),sd)
+dl=length(d_pre)
+cl=length(c_pre)
[…]

### 2020-09-11 tensor_regress/code_JASA/Figure4.R

@@ -1,0 +1,113 @@
+##Code for figure 4. MSE comparison between tensor method vs. massive GLMs.
+install.packages("tensorregress.tar.gz",repos = NULL, type="source")
+source("simulation.R")
+library(tensorregress)
+library(ggplot2)
+
+############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line #60 to run the code ##########
+
+#################### plot MSE for normal model ###############
+load("presaved/comparison_normal.RData")
+mean=c(apply(table,1,mean),apply(table_naive,1,mean))
+sd=c(apply(table,1,sd),apply(table_naive,1,sd))
+data=cbind(round(mean,2),sd,c(1,1,1,2,2,2),c(1,2,3,1,2,3))
+data=data.frame(data)
+names(data)=c("MSE","sd","method","rank")
+
+pdf("comparison_normal.pdf",width=5,height=4)
+p=ggplot(data=data, aes(x=as.factor(rank), y=MSE, fill=as.factor(method))) +
+geom_bar(stat="identity", position=position_dodge())+
+geom_errorbar(aes(ymin=MSE-sd, ymax=MSE+sd), width=.2,
+position=position_dodge(.9))+scale_fill_manual(values=c('#069AA0','#C0C0C0'))+labs(x="number of blocks",y="mean sqaured error (MSE)")+coord_cartesian(ylim = c(0, 1.15))
+p
+dev.off()
+
+
+#################### plot MSE for poisson model ###############
+load("presaved/comparison_poisson.RData")
+mean=c(apply(table,1,mean),apply(table_naive,1,mean))
+sd=c(apply(table,1,sd),apply(table_naive,1,sd))
+data=cbind(round(mean,2),sd,c(1,1,1,2,2,2),c(1,2,3,1,2,3))
+data=data.frame(data)
+names(data)=c("MSE","sd","method","rank")
+
+pdf("comparison_poisson.pdf",width=5,height=4)
+p=ggplot(data=data, aes(x=as.factor(rank), y=MSE, fill=as.factor(method))) +
+geom_bar(stat="identity", position=position_dodge())+
+geom_errorbar(aes(ymin=MSE-sd, ymax=MSE+sd), width=.2,
[…]

### 2020-09-11 tensor_regress/code_JASA/Figure8.R

@@ -1,0 +1,43 @@
+### Code for Figure 8. Comparison between classical GLM and tensor methods in the HCP data anaysis
+source("simulation.R")
+data=load("HCP.RData")
+############ Method 1: tensor method  ##############
+load("output_HCP.RData")
+table(attr[,5])
+## age group: 22-25 26-30 31+
+## number of individuals: 35  58    43
+X=attr[,4:5]
+levels(X[,2])=c("22-25","26-30","31+","31+") ## three age groups
+contrasts(X[,1]) <- contr.sum
+contrasts(X[,2]) <- contr.sum
+### intercept, F 1, M,-1; 22-25 [1,0]; 26-30 [0 1]; 30+
+coef=array(0,dim=c(68,68,5))
+coef[,,1:4]=result$C_ts[,,1:4]##intercept, gender, age 22-25, age 26-30
+coef[,,5]=-coef[,,3]-coef[,,4] ## age 30+
+
+
+############## Method 2: Classical GLM  ##############
+tsr=tensor
+X_covar3=model.matrix(~as.factor(X[,1])+as.factor(X[,2])) ## baseline female and age 22-25
+massive=massive_glm(tsr,X_covar3,"binary")
+mass=array(0,dim=c(68,68,5))
+mass[,,1]=massive[,,1]
+mass[,,2]=massive[,,2]
+mass[,,3]=massive[,,3]
+mass[,,4]=massive[,,4]
+mass[,,5]=-massive[,,4]-massive[,,3]
+
+
+### Plot histogram to compare two methods###
+pdf("compare.pdf",width=15,height=6)
+par(mfrow=c(2,4))
+hist(c(mass[,,1]),nclass=40,xlab="Effect size",main="Intercept (Classical GLM)")
+hist(c(mass[,,2]),nclass=40,xlab="Effect size",main="Gender (Classical GLM)")
+hist(c(mass[,,4]),nclass=40,xlab="Effect size",main="Age 26-30 (Classical GLM)")
+hist(c(mass[,,5]),nclass=40,xlab="Effect size",main="Age 31+ (Classical GLM)")
+
+hist(c(coef[,,1]),nclass=40,xlab="Effect size",main="Intercept (Tensor regression)")
+hist(c(coef[,,2]),nclass=40,xlab="Effect size",main="Gender (Tensor regression)")
[…]

### 2020-09-11 tensor_regress/code_JASA/HOLRR.R

@@ -1,0 +1,38 @@
+## This part follows algorithm 1 in Guillaume et al. 2016
+## HOLRR
+library(rTensor)
+HOLRR = function(X,tsr,core_shape, gamma = 0){
+  # core_shape = (d0, d1, ... dp)
+  # X_shape =  N , d0
+  # tsr_shape = N, d1, d2, ...,dp
+  p = length(core_shape) - 1
+  tsr = as.tensor(tsr)
+  Y_1 = k_unfold(tsr, m = 1)@data
+  I = diag(dim(X)[2])
+  A = solve(t(X)%*%X + gamma*I)%*%t(X)%*%Y_1%*%t(Y_1)%*%X
+  if(core_shape[1] == dim(X)[2]){
+    U0 = diag(core_shape[1])
+  }else{
+    U0 = eigen(A)$vectors[,1:core_shape[1]]
+  }
+  U = list(); U[[1]] = U0
+  for(i in seq(2,length(core_shape))){
+    if(core_shape[i] == dim(tsr)[i]){
+      U[[i]] = diag(core_shape[i])
+    }else{
+      Y_i = k_unfold(tsr, m = i)@data
+      U[[i]] = eigen(Y_i%*%t(Y_i))$vectors[,1:core_shape[i]]
+    }
+  }
+  M = solve(t(U0)%*%(t(X)%*%X + gamma*I)%*%U0)%*%t(U0)%*%t(X)
+  
+  U_M = lapply(seq(length(core_shape)), function(x) t(U[[x]]))
+  U_M[[1]] = M
+  
+  G = ttl(tsr, U_M, ms = seq(p+1))
+  C_ts = ttl(G, U, ms = seq(p+1))
+  pre=ttm(C_ts,X,1)
+  return(list(G = G@data, U = U, C_ts = C_ts@data,pre=pre))
+}
+
+  

### 2020-09-11 tensor_regress/code_JASA/HOPLS.R

@@ -1,0 +1,122 @@
+## This part follows algorithm 1+2 --> 3 in Q. Zhao et al. 2012
+## HOPLS: for N-order tensor, N \geq 3
+library(rTensor)
+## X is a matrix
+HOPLS = function(X,tsr, R, Kn, tol){
+  # R: number of latent vectors
+  # Kn: vector, contains number of loadings in tsr (see fig 2 in Q. Zhao et al. 2012)
+  # tol: tolerance, epsilon in algorithm
+  E = list(); Fr = list() ## denote residuals of X, tsr
+  t = list() ## latent vectors
+  Q = list() ## loadings
+  D = list() ## tensor of tsr (see fig 2 in Q. Zhao et al. 2012)
+  E[[1]] = X; Fr[[1]] = as.tensor(tsr)
+  for(r in 1:R){
+    if(sum(E[[r]]^2) > tol & sum(Fr[[r]]@data^2) > tol){
+      Cr = ttm(Fr[[r]],t(E[[r]]), m = 1)
+      tckr = tucker(Cr, ranks = c(1,Kn))
+      t_inte = E[[r]]%*%tckr$U[[1]]  ## pr = tckr$U[[1]]  vector
+      t[[r]] = t_inte/sqrt(sum(t_inte^2))
+      Q[[r]] = tckr$U ; Q[[r]][[1]] = t[[r]]
+      Qrt = lapply(1:(1+length(Kn)), function(x) t(Q[[r]][[x]]))  ## get t(Q[[r]])
+      #D_inte = ttm(Fr[[r]],t(t[[r]]), m = 1)
+      #D[[r]] = ttl(D_inte,Qrt,ms = seq(2,length(Kn)+1))
+      D[[r]] = ttl(Fr[[r]],Qrt,ms = seq(1,length(Kn)+1))
+      E[[r+1]] = E[[r]] - t[[r]]%*%t(tckr$U[[1]])
+      #inte = ttm(D[[r]], t[[r]], m = 1)
+      #Fr[[r+1]] = Fr[[r]] - ttl(inte, Q[[r]], m = seq(2,length(Kn)+1))
+      Fr[[r+1]] = Fr[[r]] - ttl(D[[r]], Q[[r]], m = seq(1,length(Kn)+1))
+    }else break
+  }
+  pre = 0
+  for(r in 1:R){
+      inte = ttl(D[[r]], Q[[r]], m = seq(1,length(Kn)+1))
[…]

### 2020-09-11 tensor_regress/code_JASA/Table2.R

@@ -1,0 +1,90 @@
+#### code for Table 2. Accuracy of BIC selection for normal and Possion models under different (d,r) combinations
+install.packages("tensorregress.tar.gz",repos = NULL, type="source")
+source("simulation.R")
+library(tensorregress)
+
+set.seed(0) ## set seed for this simulation
+seed=0
+
+### all covariates included ##
+d=c(20,40)
+rank_range=array(0,dim=c(3,7,3))
+rank_range[1,,]=rbind(c(3,3,3),c(3,3,2),c(3,2,2),c(2,2,2),c(3,3,4),c(3,4,4),c(4,4,4))
+rank_range[2,,]=rbind(c(4,4,6),c(4,4,5),c(3,3,6),c(4,5,7),c(3,4,6),c(4,5,6),c(5,5,5))
+rank_range[3,,]=rbind(c(6,8,8),c(6,7,8),c(5,8,8),c(7,8,8),c(5,7,7),c(5,7,8),c(6,7,7))
+core_shape=rbind(c(3,3,3),c(4,4,6),c(6,8,8))
+dup=30
+
+
+########################## normal model ##########################
+dist="normal"
+final=array(0,dim=c(3,2,dup,3))
+for(s in 1:3){ ## three possible ranks
+    for(i in 1:2){ ## two possible dimensions
+        whole_shape = c(d[i],d[i],d[i])
+        p=0.4*whole_shape
+        data=sim_data(seed, whole_shape = whole_shape, core_shape=core_shape[s,],p=p,dist=dist, dup=dup, signal=4)
+        table=NULL
+        for(j in 1:dup){ ## simulation replicates
+            res=sele_rank(data$tsr[[j]],X_covar1 = data$X_covar1, X_covar2 = data$X_covar2,X_covar3 = data$X_covar3,rank_range=rank_range[s,,],Nsim=10,cons = 'non',dist=dist)
+            table=rbind(table,res$rank)
+        }
+        final[s,i,,]=table
+    }
+}
+
+save(final,file="normal_rank.RData")
+
+## d = 20
+apply(final[1,1,,],2,mean) ## ground truth (3,3,3)
+apply(final[2,1,,],2,mean) ## ground truth (4,4,6)
+apply(final[3,1,,],2,mean) ## ground truth (6,8,8)
+apply(final[1,1,,],2,sd) ## ground truth (3,3,3)
+apply(final[2,1,,],2,sd) ## ground truth (4,4,6)
[…]

### 2020-09-11 tensor_regress/code_JASA/Yu.R

@@ -1,0 +1,81 @@
+
+
+##------------------------   TPG + ITP
+library(rTensor)
+## This part follows algorithm 2 in Rose Yu et al. 2016
+## ITP: for N-order tensor, N = 3
+ITP = function(W, R, U){
+    # R: scalar constraining tucker rank
+    # W: Coefficient tensor: d1, d2, T
+    # U: list containing factor matrices
+    W = as.tensor(W)
+    U1 = U[[1]]; U2 = U[[2]]; U3 = U[[3]]
+    i = 1
+    while(i <= R){
+        u10 = U1[,i]; u20 = U2[,i]; u30 = U3[,i]
+        for(j in 1:100){  ##  power iteration
+            u11 = ttl(W,list(t(as.matrix(u20)), t(as.matrix(u30))), ms = c(2,3))@data
+            u11 = as.vector(u11); u11 = u11/sum(u11^2)
+            u21 = ttl(W,list(t(as.matrix(u11)), t(as.matrix(u30))), ms = c(1,3))@data
+            u21 = as.vector(u21); u21 = u21/sum(u21^2)
+            u31 = ttl(W,list(t(as.matrix(u11)), t(as.matrix(u21))), ms = c(1,2))@data
+            u31 = as.vector(u31); u31 = u31/sum(u31^2)
+            u10 = u11; u20 = u21; u30 = u31
+            if((sum((u11 - u10)^2) + sum((u21 - u20)^2) + sum((u31 - u30)^2)) <= 0.05) break
+        }
+        U1[,i] = u11;  U2[,i] = u21;  U3[,i] = u31 
+        i = i + 1
+    }
+    U = list(U1,U2,U3)
+    UUT = lapply(1:3, function(x) U[[x]]%*%t(U[[x]]))
+    W = ttl(W, UUT, m = c(1,2,3))@data
+    return(list(W = W, U = U))
+}
+
+sketch = function(x){
+    i = sample(1:length(x),1)
+    x[i] = sample(c(1,-1),1,prob = c(0.5,0.5))
+    return(x)
+}
+
+## This part follows algorithm 1 in Rose Yu et al. 2016
+## TPG: for N-order tensor, N = 3
+TPG = function(X, tsr, R, eta, iter, tol){
+    # tsr: d3, d2, T
+    # X: d3, d1
+    # W: Coefficient tensor: d1, d2, T
+    # eta: step size
[…]

### 2020-09-11 tensor_regress/code_JASA/simulation.R

@@ -1,0 +1,85 @@
+######### massive GLM to estimate the covariate effects ###
+massive_glm=function(response,X,dist){
+    d1=dim(response)[1]
+    d2=dim(response)[2]
+    coe=array(0,dim=c(d1,d2,dim(X)[2]))
+    if(dist=="normal"){
+        for(i in 1:d1){
+            for(j in 1:d2){
+                fit=lm(response[i,j,]~-1+X)
+                coe[i,j,]=coef(fit)
+            }
+        }
+    }
+    else if(dist=="binary"){
+        for(i in 1:d1){
+            for(j in 1:d2){
+                fit=suppressWarnings(glm(response[i,j,]~-1+X,family=binomial("logit")))
+                coe[i,j,]=coef(fit)
+            }
+        }
+    }
+    else if(dist=="poisson"){
+        for(i in 1:d1){
+            for(j in 1:d2){
+                fit=suppressWarnings(glm(response[i,j,]~-1+X,family=poisson("log")))
+                coe[i,j,]=coef(fit)
+            }
+        }
+    }
+    return(coe)
+}
+
+### compute loglikelihood under different models
+loglike=function(data,linearpre,dist){
+    if(dist=="binary"){
+        p=plogis(linearpre)## log-likelihood
+        L=dbinom(data,1,p,log=TRUE)
+        return(sum(L))
+    }
+    else if (dist=="normal"){
+        sigma_est=mean((data-linearpre)^2)
+        L=dnorm(data,linearpre,sqrt(sigma_est),log=TRUE)
+        sum(L)##log-likelihood
+    }
+    else if (dist=="poisson"){
+        lambda=exp(linearpre)
+        L=dpois(data,lambda, log = TRUE) ## log-likelihood for Poisson
+        return(sum(L))
+    }
+}
+
+#### wrapped simulation function to estimate covariate effects using tensor methods or massive glm methods
+conv_rate = function(seed,signal=10,Nsim=20,cons="no",lambda = 1,alpha=10,solver ="GC",c_range,dist="binary",dup=10,d_range,p_range,naive=TRUE){
+    #cons can be "non","vanilla","penalty"
+    n=m=nrow(d_range)
+    s=nrow(c_range)
+    error=error_naive=array(0,dim=c(n,s,dup))
+    for(i in 1:n){
+        for(k in 1:s){
+            data = sim_data(seed, d_range[i,], c_range[k,],p_range[i,],dist, dup, signal)
[…]

### 2020-09-11 tensor_regress/code_JASA/readme.c
update

@@ -1,0 +1,35 @@
+1. Nations (Nickel et al., 2011, dnations.mat): This is a 14 × 14 × 56 binary tensor consisting of 56 political relations of 14 countries between 1950 and 1965. The tensor entry indicates the presence or absence of a political action, such as “treaties”, “sends tourists to”, between the nations. Please set the diagonal elements Y(i,i,k) = 0 in the analysis.
+
+I have pre-selected ~10 covariates for each nation. These covariates describe a few important country attributes, e.g. whether a nation is actively involved in medicine NGO, law NGO, or belongs to a catholic nation, etc. 
+
+2. HCP (Wang et al., 2017a, Count_VSPLOT.mat): This is a 68 × 68 × 136 binary tensor consisting of structural connectivity patterns among 68 brain regions for 136 individuals from Human Connectome Project (HCP). All the individual images were preprocessed following a standard pipeline (Zhang et al., 2018), and the brain was parcellated to 68 regions-of-interest following the Desikan atlas (Desikan et al., 2006). The tensor entries encode the presence or absence of fiber connections between those 68 brain regions for each of the 136 individuals.
+
+There are 573 covariates available for each individual. I have pre-selected ~20 important covariates. Please see the listed below. You may start with them first and then add a few new covariates of your interest. 
+The full list of covariates can be found at:
+https://wiki.humanconnectome.org/display/PublicData/HCP+Data+Dictionary+Public-+Updated+for+the+1200+Subject+Release?preview=/53444663/113377285/HCP_S1200_DataDictionary_April_20_2018.xlsx#HCPDataDictionaryPublic-Updatedforthe1200SubjectRelease-Excelversion:
+
+
+Below is the dictinary for the important covariates:
+Gender: binary, M or F
+Age: categorical, 22-25, 26-30, 31-35, >35
+MMSE_Score: numerical, Mini Mental Status Exam Total Score
+PSQI_Score: numerical, Pittsburgh Sleep Questionnaire Total Score
+PicSeq_Unadj: Measure of Episoidc Memory. NIH Toolbox Picture Sequence Memory Test: Unadjusted Scale Score, 
[…]

### 2020-09-14 tensor_regress/JASA_submission/paper_source/regress.tex
clean

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url}
\usepackage{enumitem}

\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{algorithm}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={black},citecolor={black},urlcolor={black}}

\newcommand{\blind}{1}

\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{-.3in}%
\addtolength{\topmargin}{-.8in}%

\theoremstyle{definition}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{cor}{Corollary}[section]

\theoremstyle{definition}
\newtheorem{assumption}{Assumption}
\newtheorem{defn}{Definition}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\newtheorem{innercustomgeneric}{\customgenericname}
\providecommand{\customgenericname}{}
\newcommand{\newcustomtheorem}[2]{%
  \newenvironment{#1}[1]
  {%
   \renewcommand\customgenericname{#2}%
   \renewcommand\theinnercustomgeneric{##1}%
   \innercustomgeneric
  }
  {\endinnercustomgeneric}
}

\newcustomtheorem{customexample}{Example}

\usepackage{appendix}
\usepackage{wrapfig}
\mathtoolsset{showonlyrefs}

\input macros.tex

\usepackage[english]{babel}

\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}

\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\def\ci{\perp\!\!\!\perp}

\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}
\usepackage{booktabs}
\newcommand\doubleRule{\toprule\toprule}
\allowdisplaybreaks

\usepackage{xr}
\externaldocument{supp}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if1\blind
{
  \title{\bf Supervised Tensor Decomposition with Interactive Side Information}
  % blind version
  \author{Jiaxin Hu, Zhuoyan Xu, Chanwoo Lee, and Miaoyan Wang\\
    Department of Statistics, University of Wisconsin-Madison}
  \maketitle
} \fi

\if0\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Supervised Tensor Decomposition with Interactive Side Information}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Higher-order tensors have received increased attention across science and engineering. While most tensor decomposition methods are developed for a single tensor observation, scientific studies often collect side information, in the form of node features and interactions thereof, together with the tensor data. Such data problems are common in neuroimaging, network analysis, and spatial-temporal modeling. Identifying the relationship between a high-dimensional tensor and side information is important yet challenging. Here, we develop a tensor decomposition method that incorporates multiple side information as interactive features. Unlike unsupervised tensor decomposition, our supervised decomposition captures the effective dimension reduction of the data tensor confined to feature space on each mode. An efficient alternating optimization algorithm is further developed. Our proposal handles a broad range of data types, including continuous, count, and binary observations. We apply the method to diffusion tensor imaging data from human connectome project and multi-relational political network data. We identify the key global connectivity pattern and pinpoint the local regions that are associated with available features. Our method will help the practitioners efficiently analyze tensor datasets in various areas. Toward this end, the package and data used are available at~\url{https://CRAN.R-project.org/package=tensorregress}.

\end{abstract}

\noindent%
{\it Keywords:} Applications and case studies, Tensor data analysis, Supervised dimension reduction, Exponential family distribution, Generalized multilinear model
\vfill

\newpage
\spacingset{1.5} % DON'T change the spacing!
\section{Introduction}
\label{sec:intro}

Multi-dimensional arrays, known as tensors, are often collected with side information on multiple modes in modern scientific and engineering studies. A popular example is in neuroimaging~\citep{zhou2013tensor,wang2019common}. The brain connectivity networks are collected from a sample of individuals, accompanied by individual characteristics such as age, gender, and diseases status (see Figure~\ref{fig:intro1}a). Another example is in network analysis~\citep{pmlr-v108-berthet20a,hoff2005bilinear}. A typical social network consists of nodes that represent people and edges that represent  friendships. Side information such as people’s demographic information and friendship types are often available. In both examples, it is of keen scientific interest to identify the variation in the data tensor (e.g., brain connectivities, social community patterns) that is affected by available features. These seemingly different scenarios pose a common yet challenging problem for tensor data modeling.

In addition to the aforementioned challenges, many tensor datasets consist of non-Gaussian measurements. Examples include the political interaction dataset \citep{hu2015scalable} which measures action counts between countries under various relation types, and the brain connectivity network dataset \citep{zhang2018mapping} which is a collection of binary adjacency matrices. Classical tensor decomposition methods are based on minimizing the Frobenious norm of the reconstruction error, leading to suboptimal predictions for binary- or count-valued response variables. A number of supervised tensor methods have been proposed \citep{narita2012tensor, zhao2012higher, yu2016learning} to address the tensor regression problem in various forms, such as scalar-to-tensor regression and tensor-response regression. These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest. The gap between theory and practice means a great opportunity to improve modeling paradigms to better capture the complexity in tensor data.

\begin{figure*}[t]
\begin{center}
\includegraphics[width=16cm]{demo.pdf}
\end{center}
\caption{Examples of supervised tensor decomposition with interactive side information. (a) Network population model. (b) Schematic diagram of supervised tensor decomposition for an order-3 tensor $\tY$ with feature matrices $\mX_k$ on each of the three modes.} \label{fig:intro1}
\end{figure*}

We present a general model and associated method for decomposing a data tensor whose entries are from exponential family with interactive side information. We formulate the learning task as a structured regression problem, with tensor observation serving as the response, and the multiple side information as interactive features. Figure~\ref{fig:intro1}b illustrates our model in the special case of order-3 tensors. A low-rank structure is imposed to the conditional mean of tensor observation, where unlike classical decomposition, the tensor factors $\mX_k\mM_k\in\mathbb{R}^{d_k\times r_k}$ belong to the space spanned by features $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,2,3$. The unknown matrices $\mM_k\in\mathbb{R}^{p_k\times r_k}$ (referred to as ``dimension reduction matrices'') link the conditional mean to the feature spaces, thereby allowing the identification of variations in the tensor data attributable to the side information.

Our proposal blends the modeling power of generalized linear model (GLM) and the exploratory capability of tensor dimension reduction in order to take the best out of both worlds. We leverage GLM to allow heteroscedacity due to the mean-variance relationship in the non-Gaussian data. This flexibility is important in practice. Furthermore, our low-rank model on the (transformed) conditional mean tensor effectively mitigates the curse of high dimensionality. In classical GLM, the sample size and feature dimension are well defined; however, in the tensor data analysis, we observe only one realization of an order-$K$ tensor and up to $K$ interactive feature matrices. Both the number of tensor entries and feature dimension grow exponentially in $K$. Dimension reduction is therefore crucial for prediction and interpretability. We establish the statistical convergence of our estimator, and we quantify the gain in prediction through simulations and case studies.

Our work is closely related to but also clearly distinctive from several lines of previous work. The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor, hong2020generalized} that aims to find the best low-rank representation of a data tensor. In contrast, our model is a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and multiple features. The low-rank factorization is determined jointly by the tensor data and feature matrices in our model.

The second line of work studies the tensor-on-tensor regression~\citep{raskutti2019convex, lock2018tensor, gahrooei2020multiple}. Our model shares a common ground with earlier approaches, but we provide more efficient solutions to new settings that have more practical significance. As we show in Section~\ref{sec:connection}, the supervised tensor decomposition has an interesting connection with tensor-on-tensor regression. Previous methods~\citep{gahrooei2020multiple,lock2018tensor} mainly focus on Gaussian tensors. The Frobenius norm used in the objective function is statistical suboptimal for general exponential family tensors. Maximum likelihood estimator (MLE) is studied in~\citep{raskutti2019convex} and a convex relaxation algorithm is proposed to solve for low-rank tensor coefficients. However, in the tensor case, convex MLE suffers from both computational intractability and statistical suboptimality. We advocate a non-convex approach and provide strong evidence for its success in our setting. Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2019convex,gahrooei2020multiple}, and we go step further by finding the sufficient dimension reduction~\citep{adragni2009sufficient}, $\textup{Span}(\mM_k)$, that facilitates the identification of interaction effects in features (see Figure~\ref{fig:intro1}b). The latter approach greatly improves the \emph{interpretability} in prediction. In this regards, our method opens up new opportunities for tensor data analysis in a wider range of applications.

The third line of work uses side information for various tensor learning tasks, such as for completion~\citep{narita2012tensor,cao2016semi} and for recommendation system~\citep{ioannidis2019coupled,farias2019learning}. These methods also study tensors with side information, same as we do, but they take regularization approaches to penalize predictions that are distant from side information. One important difference is that their goal is prediction but not parameter estimation. The effects of features and their interactions are not estimated in these data-driven approaches. In contrast, our goal is interpretable prediction, and we estimate factor matrices $\mM_k$ using a model-based approach. Estimating $\mM_k$ allows us to identify sufficient features and the interactions thereof. We numerically compare these two approaches in Section~\ref{sec:simulation}.

The remainder of this paper is organized as follows. Section~\ref{sec:pre} introduces tensor notation and preliminaries. Section~\ref{sec:model} presents the main model and three motivating examples for supervised tensor decomposition. Section~\ref{sec:est} describes the rank-constrained quasi-likelihood estimation and an alternating optimization algorithm. In Section~\ref{sec:simulation}, we present numerical experiments and assess the performance in comparison to alternative methods. In Section~\ref{sec:data}, we apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data. We conclude in Section~\ref{sec:con} with discussions about our findings and avenues of future work.

\section{Preliminaries}\label{sec:pre}
We introduce the basic tensor properties used in the paper; more details on tensor notation can be found in~\cite{kolda2009tensor}. We use $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ to denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor, where $K$ corresponds to the number of modes of $\tY$ and is called the order. The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as
\begin{equation}
\tY \times_1\mX_1\ldots \times_K \mX_K=\entry{\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}x_{j_1,i_1}^{(1)}\ldots x_{j_K,i_K}^{(K)}},
\end{equation}
which results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor. For ease of presentation, we use the shorthand $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product. For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as 
\[
\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.
\] 
The Frobenius norm and maximum norm of $\tY$ are defined as 
\[
\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}\quad \text{and}\quad \mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.
\]
When $a$ is a vector, we use $\vnormSize{}{a}=\langle a,a \rangle^{1/2}$ to denote the vector 2-norm.

A higher-order tensor can be reshaped into a lower-order object. We use $\textup{vec}(\cdot)$ to denote the operation that reshapes the tensor into a vector, and $\text{Unfold}_k(\cdot)$ to denote the operation that reshapes the tensor along mode $k$ into a matrix of size $d_k$-by-$\prod_{i\neq k}d_i$. The multilinear rank of an order-$K$ tensor $\tY$ is defined as a length-$K$ vector $\mr=(r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\textup{Unfold}_k(\tY)$, $k=1,\ldots,K$.

We use lower-case letters (e.g.,\ $a,b,c$) for scalars and vectors, upper-case boldface letters (e.g.,\ $\mA, \mB, \mC$) for matrices, and calligraphy letters (e.g.,\ $\tA, \tB, \tC$) for tensors of order three or greater. We let $\mI_d$ denote the $d \times d$ identity matrix and $[d]$ denote the $d$-set $\{1,\ldots,d\}$. For ease of notation, we allow the basic arithmetic operators (e.g.,\ $+, -, \geq $) and univariate functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner. We use $f'(\cdot)$ to denote the first-order derivative and $f''(\cdot)$ the second-order derivative of function $f(\cdot)$.

\section{Motivation and model}\label{sec:model}
\subsection{General framework for supervised tensor decomposition}
We begin with a general framework for supervised tensor decomposition and then discuss its implication in three concrete examples. Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor. Suppose the side information is available on each of the $K$ modes. Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the feature matrix on the mode $k\in[K]$, where $x_{ij}$ denotes the $j$-th feature value for the $i$-th tensor entity, for $(i,j)\in[d_k]\times[p_k]$, $p_k\leq d_k$. We propose a multilinear conditional mean model between the data tensor and feature matrices. Assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent realizations from an exponential family distribution, and the conditional mean tensor admits the form
\begin{align}\label{eq:tensormodel}
&\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)=f(\Theta),\quad \text{with}\quad \Theta =\tB\times\{\mX_1,\ \ldots,\ \mX_K\} ,
\end{align}
where $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the unknown parameter tensor, $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$, and $\times$ denotes the tensor-by-matrix product. The choice of link function is based on the assumed distribution family of tensor entries. Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution. In general, dispersion parameters can also be included in the model. Because our main focus is the tensor decomposition under the mean model, we omit the dispersion parameter in this section for ease of presentation.

In classical tensor decomposition, tensor factorization is performed on either data tensor $\tY$ or mean tensor $\mathbb{E}(\tY)$. In the context of supervised tensor decomposition, we propose to factorize the latent parameter tensor $\tB$,
\begin{equation}\label{eq:factor}
\tB=\tC\times\{\mM_1,\ \ldots,\ \mM_K\},
\end{equation}
where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, and $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices consisting of orthonormal columns, where $r_k\leq p_k$ for all $k\in[K]$. By the definition of multilinear rank, model equations~\eqref{eq:tensormodel} and~\eqref{eq:factor} imply the low-rankness $\mr=(r_1,\ldots,r_K)$ of the conditional mean tensor under the link function. We now reach our final model for supervised tensor decomposition,
\begin{align}\label{eq:decomp}
\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1,\ \ldots,\ \mX_K\mM_K\}\right),\notag\\
\text{with} \ &\ \mM_k^T\mM_k = \mI_{r_k},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\quad \text{for all }k=1,\ldots,K,
\end{align}
where the parameters of interest are $\mM_k$ and $\tC$. Note that model~\eqref{eq:decomp} assumes a fixed, known rank $\mr=(r_1,\ldots,r_K)$; the adaptation to unknown rank will be addressed in Section~\ref{sec:tuning}. Figure~\ref{fig:intro1}b provides a schematic illustration of our model. The features $\mX_k$ affect the distribution of tensor entries in $\tY$ through the form $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$. We call $\mX_k\mM_k$ the ``supervised tensor factors'' or ``sufficient features''~\citep{adragni2009sufficient}, and call $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes. Our goal is to find $\mM_k$ and the corresponding $\tC$. Note that $\mM_k$ and $\tC$ are identifiable only up to orthonormal transformations.

\subsection{Three examples}
We give three seemingly different examples that can all be formulated as our supervised tensor decomposition model~\eqref{eq:decomp}.
\begin{example}[Spatio-temporal growth model]
The growth curve model~\citep{gabriel1998generalised,srivastava2008models} was originally proposed as an example of bilinear model for matrix data, and we adopt its higher-order extension here. Let $\tY=\entry{y_{ijk}}\in\mathbb{R}^{d \times m\times n}$ denote the pH measurements of $d$ lakes at $m$ levels of depth and for $n$ time points. Suppose the sampled lakes belong to $q$ types, with $p$ lakes in each type. Let $\{\ell_j\}_{j\in[m]}$ denote the sampled depth levels and $\{t_k\}_{k\in[n]}$ the time points. Assume that the expected pH trend in depth is a polynomial of order at most $r$ and that the expected trend in time is a polynomial of order $s$. Then, the conditional mean model for the spatio-temporal growth can be represented as
\begin{equation}\label{eq:time}
\mathbb{E}(\tY|\mX_1,\mX_2,\mX_3)=\tC\times\{\mX_1\mM_1,\ \mX_2\mM_2,\ \mX_3\mM_3\},
\end{equation}
where $\mX_1=\text{blockdiag}\{\mathbf{1}_p,\ldots,\mathbf{1}_p\}\in \{0,1\}^{d\times q}$ is the design matrix for lake types, and
\[
\mX_2=
\begin{pmatrix}
1 & \ell_1&\cdots &\ell^{r}_1\\
1 & \ell_2&\cdots &\ell^{r}_2\\
\vdots &\vdots&\ddots&\vdots\\
1&\ell_{m}&\cdots&\ell^{r}_{m}
\end{pmatrix},\quad
\mX_3=
\begin{pmatrix}
1 & t_1&\cdots &t^{s}_1\\
1 & t_2&\cdots &t^{s}_2\\
\vdots &\vdots&\ddots&\vdots\\
1&t_{n}&\cdots&t^{s}_{n}
\end{pmatrix}
\]
are the design matrices for spatial and temporal effects, respectively, $\tC\in\mathbb{R}^{r_1\times r_2\times r_3}$ is the unknown core tensor, and $\mM_k$ are unknown dimension reduction matrices on each mode. The factors $\mX_k\mM_k$ are sufficient features in the mean model~\eqref{eq:time}. The spatial-temporal model is a special case of our supervised tensor decomposition model~\eqref{eq:decomp}, with features available on each of the three modes.
\end{example}

\begin{example}[Network population model]\label{example:brain}
Network response model~\citep{rabusseau2016low, zhang2018network} is recently developed for neuroimanig analysis. The goal is to study the relationship between brain network connectivity pattern and features of individuals. Suppose we have a sample of $n$ observations, $\{(\mY_i, \mx_i)\colon i=1,\ldots,n\}$, where for each individual $i\in[n]$, $\mY_i\in\{0,1\}^{d\times d}$ is the undirected adjacency matrix whose entries indicate presences/absences of connectivities between $d$ brain nodes, and $\mx_i\in\mathbb{R}^p$ is the individual's feature such as age, gender, cognition score, etc. The network-response model  has the conditional mean
\begin{equation}\label{eq:network}
\textup{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n,
\end{equation}
where $\tB\in \mathbb{R}^{d\times d\times p}$ is a rank-$(r_1,r_1,r_2)$ coefficient tensor, and $\tB$ is assumed to be symmetric in the first two modes.

The model~\eqref{eq:network} is a special case of our supervised tensor decomposition, with feature matrix on the last mode of the tensor. Specifically, we stack the network observations $\{\mY_i\}$ together and obtain an order-3 response tensor $\tY\in\{0,1\}^{d\times d\times n}$. Define a feature matrix $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$. Then, the model~\eqref{eq:network} has the equivalent representation of supervised tensor decomposition,
\[
\textup{logit}(\mathbb{E}(\tY|\mX))=\tC\times\{\mM,\ \mM,\ \mX\mM'\},
\]
where $\tC\in\mathbb{R}^{r_1\times r_1\times r_2}$ is the core tensor, $\mM\in\mathbb{R}^{d\times r}$ is the dimension reduction matrix on the first two modes, and $\mM'\in\mathbb{R}^{p\times r_2}$ is for the last mode.  \end{example}

 \begin{example}[Dyadic data with node attributes] Dyadic dataset consists of measurements on pairs of objects. Common examples include graphs and networks. Let $\tG=(V,E)$ denote a graph, where $V=[d]$ is the node set of the graph, and $E\subset V\times V$ is the edge set. Suppose that we also observe feature vector $\mx_i\in\mathbb{R}^p$ associated to each node $i\in V$. A probabilistic model on the graph $\tG=(V,E)$ can be described by the following matrix regression. The edge connects the two vertices $i$ and $j$ independently of other pairs, and the probability of connection is modeled as
\begin{equation}\label{eq:edge}
 \textup{logit}\left(\mathbb{P}\left((i,j)\in E\right)\right)=\mx^T_i\mB\mx_j=\langle \mB,\ \mx^T_i\mx_j\rangle,
 \end{equation}
 where $\mB\in\mathbb{R}^{p\times p}$ is a symmetric rank-$r$ matrix. The low-rankness in $\mB$ has demonstrated its success in modeling transitivity, balance, and communities in  networks~\citep{hoff2005bilinear}. We show that our supervised tensor decompostiion~\eqref{eq:decomp} also incorporates the graph model as a special case. Let $\tY=\entry{y_{ij}}$ be a binary matrix where $y_{ij}=\mathds{1}_{(i,j)\in E}$. Define $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$. Then, the graph model~\eqref{eq:edge} can be expressed as
 \[
 \textup{logit}(\mathbb{E}(\mY|\mX))=\mC\times\{\mX\mM,\ \mX\mM\}, 
  \]
  where $\mC\in\mathbb{R}^{r\times r}$ and $\mM\in\mathbb{R}^{p\times r}$ are based on the singular value decomposition of $\mB=\mM\mC\mM^T$. 
  \end{example}

In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by features. The coefficient tensor $\tB$ in model~\eqref{eq:tensormodel} serves this goal by collecting the feature effects. To encourage the sharing among effects, decomposition~\eqref{eq:factor} assumes that the feature effects $\tB$ lie in a low-dimensional parameter space. The low-rank assumption is plausible in many scientific applications. In the Example~\ref{example:brain} of brain imaging analysis, for instance, it is often believed that the brain nodes can be partitioned into several regions, and the person's attributes affect the connectivity in a similar way within a region.

Our supervised tensor decomposition~\eqref{eq:decomp} is able to incorporate arbitrary numbers of feature matrices. When certain mode $k$ has no available side information, we set $\mX_k=\mI_{d_k}$ in the model~\eqref{eq:decomp}. In particular, our model~\eqref{eq:decomp} reduces to classical unsupervised tensor decomposition~\citep{de2000multilinear,hong2020generalized} when no side information is available; i.e., $\mX_k=\mI_{d_k}$ for all $k\in[K]$.

\subsection{Connection to sufficient dimension reduction and tensor-on-tensor regression}\label{sec:connection}
One important implication of our method is that we allow high-dimensionality in both tensor dimension $d_k$ and feature dimension $p_k$. Recall that the model rank $r_k$ is typically smaller than $d_k$ and $p_k$. In such a case, the matrices $\mM_k$ serve the role of simultaneous dimension reduction of the data tensor and features. The role of the sufficient features $\mX_k\mM_k$ can be seen in the following two conditional independence assumptions in model~\eqref{eq:decomp},
\begin{align}
\tY \ci \{\mX_k\}&\ \big|\ \{\mX_k\mM_k\} \ \ \text{(independence between the tensor and multiple features)},\\
y_{i_1,\ldots,i_K} \ci y_{i'_1,\ldots,i'_K} & \ \big|\ \{\mX_k\mM_k\}  \ \ \text{(independence within the tensor)},
\end{align}
where the second line holds for all $(i_1,\ldots,i_K)\neq(i'_1,\ldots,i'_K) \in[d_1]\times\cdots\times[d_K]$, and $\ci$ denotes the independence. The first property highlights the ``decorrelation'' role of $\mM_k$, in the same spirit as the sufficient dimension reduction~\citep{adragni2009sufficient} in high-dimensional supervised learning, whereas the second property highlights the tensor dimension reduction in the usual unsupervised sense (consider, for example, $\mX_k=\mathbf{I}_{d_k}$ for all $k\in[K]$).

Our model also has a close connection with tensor-on-tensor regression~\citep{raskutti2019convex, lock2018tensor,han2020optimal}. Specifically, model~\eqref{eq:decomp} can be viewed as a multivariate regression model, where the response is the vectorized tensor and the covariates are interactions between sufficient features across modes. We take an order-3 tensor under and Gaussian model for illustration. Let $\mX,\mZ,\mW$ denote the feature matrix on mode $k=1, 2, 3$, respectively. Suppose that each mode has two-dimensional sufficient features, denoted $\mM_1\mX=[\mx_1,\mx_2]$, $\mM_2\mZ=[\mz_1,\mz_2]$, $\mM_3\mW=[\mw_1,\mw_2]$. Here $\mx_1,\mx_2,\ldots,\mw_1, \mw_2$ are column vectors. Then the model~\eqref{eq:decomp} is a regression model with across-mode interactions,
\begin{equation}\label{eq:interaction}
\mathbb{E}(y_{ijk}|\mX,\mZ,\mW)=c_{111}\mx_{1i}\mz_{1j}\mw_{1k}+c_{121}\mx_{i1}\mz_{j2}\mw_{k1}+\cdots+c_{221}\mx_{2i}\mz_{2j}\mw_{1k}+c_{222}\mx_{2i}\mz_{2j}\mw_{2k},
\end{equation}
where $\entry{c_{ijk}} \in \mathbb{R}^{2\times 2\times 3}$ are unknown interaction effects, $\mx_{1i}$ denotes the $i$-th entry in the feature vector $\mx_1$, and similar notations apply to other features. Note that lower-order interactions are naturally incorporated in~\eqref{eq:interaction} if we include an intercept column in the sufficient feature matrices. Model~\eqref{eq:interaction} shows the benefit of our supervised tensor decomposition for identifying across-mode interactions.

\section{Estimation}\label{sec:est}

\subsection{Rank-constrained M-estimator}
We develop a likelihood-based procedure to estimate $\tC$ and $\mM_k$ in~\eqref{eq:decomp}. We adopt the exponential family as a flexible framework for different data types. In a classical generalized linear model with a scalar response $y$ and feature $\mx$, the density is expressed as
\[
p(y|\mx, \boldsymbol{\beta})=c(y,\phi)\exp\left(\frac{y\theta- b(\theta)}{\phi}\right)\ \text{with}\ \theta=\boldsymbol{\beta}^T\mx,
\]
where $b(\cdot)$ is a known function, $\theta$ is the linear predictor, $\phi>0$ is the dispersion parameter, and $c(\cdot)$ is a known normalizing function. The choice of link functions depends on the data types and on the observation domain of $y$, denoted $\mathbb{Y}$. For example, the observation domain is $\mathbb{Y}=\mathbb{R}$ for continuous data, $\mathbb{Y}=\mathbb{N}$ for count data, and  $\mathbb{Y}=\{0,1\}$ for binary data. The canonical link function $f$ is chosen to be $f(\cdot)=b'(\cdot)$. Table~1 summarizes the canonical link functions for common types of distributions.

\begin{table}[htb]
\centering
\begin{tabular}{c|ccc}
Data type &Gaussian & Poisson& Bernoulli\\
\hline
Domain $\mathbb{Y}$& $\mathbb{R}$&$\mathbb{N}$&$\{0,1\}$\\
 $b(\theta)$&$\theta^2/2$& $\exp(\theta)$&$\log (1+\exp(\theta))$\\
 link $f(\theta)$&$\theta$&$\exp(\theta)$&$(1+\exp(-\theta))^{-1}$
\end{tabular}
\caption{Canonical links for common distributions.}\label{table:link}
\end{table}

In our context, we model the tensor entries $y_{ijk}$, conditional on $\theta_{ijk}$, as independent draws from an exponential family. Ignoring constants that do not depend on $\Theta$, the quasi log-likelihood of~\eqref{eq:decomp} is equal to Bregman distance between $\tY$ and $b'(\Theta)$:
\begin{align}\label{eq:loglikelihood}
\tL_{\tY}(\tC,\mM_1,\ldots,\mM_K)&=\langle \tY, \Theta \rangle - \sum_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K}),\notag\\
\text{where}\ \Theta&=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
\end{align}

We propose a constrained maximum quasi-likelihood estimator (M-estimator),
\begin{align} \label{eq:MLE} 
(\hat \tC, \hat \mM_1,\ldots,\hat \mM_k) &=\argmax_{(\tC,\mM_1,\ldots,\mM_K)\in \tP} \ \tL_{\tY}(\tC,\mM_1,\ldots,\mM_K),
\end{align}
where the parameter space $\tP$ is
\[
\tP=\left\{(\tC, \mM_1,\ldots,\mM_K) \ \Big| \ \mM_k^T\mM_k=\mathbf{I}_{r_k},\ \mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha \right\}.
\]
The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to avoid the divergence in the non-Gaussian variance.

\subsection{Alternating optimization} \label{subsec:alg}
We propose an alternating optimization algorithm to solve~\eqref{eq:MLE}. 
The decision variables in the objective function~\eqref{eq:MLE} consist of $K+1$ blocks of variables, one for the core tensor $\tC$ and $K$ for the factor matrices $\mM_k$. We notice that, if any $K$ out of the $K+1$ blocks of variables are known, then the optimization reduces to a simple GLM with respect to the last block of variables. This observation leads to an iterative updating scheme for one block at a time while keeping others fixed. After each iteration, we rescale the core tensor $\tC^{(t+1)}$ subject to the maximum norm constraint. The full algorithm is described in Algorithm~\ref{alg:B}.

\begin{algorithm}[t]
\caption{Supervised Tensor Decomposition with Interactive Side Information}\label{alg:B}
\begin{algorithmic}[1]
\INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, feature matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,\ldots,K$, target Tucker rank $\mr=(r_1,\ldots,r_K)$, link function $f$, maximum norm bound $\alpha$
\OUTPUT Estimated core tensor $\hat \tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ and factor matrices $\hat \mM_k\in\mathbb{R}^{p_k\times r_k}$. 
\State Random initialization of the core tensor $\tC$ and factor matrices $\mM_k$. 
\While{Do until convergence}
\For { $k=1$ to $K$}
\State Obtain the factor matrix $\tilde \mM_k\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$.
\State Perform QR factorization $\tilde \mM_k=\mQ\mR$, where $\mQ\in\mathbb{R}^{p_k\times r_k}$ consists of orthonormal columns. 
\State Update $\mM_k\leftarrow \mQ$ and core tensor $\tC\leftarrow \tC\times_k \mR$.
\EndFor
\State Update the core tensor $\tC$ by solving a GLM with $\textup{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM_k]$ as features, and $f$ as link function. Here $\otimes$ denotes the Kronecker product of matrices. 
\State Rescale the core tensor $\tC$ subject to the maximum norm constraint. 
\EndWhile
\end{algorithmic}
\end{algorithm}

The optimization~\eqref{eq:MLE} is a non-convex problem due to the non-convexity in the feasible set $\tP$. Under mild conditions, our algorithm enjoys global convergence; i.e.\ any sequence of iterates generated by the alternating algorithm converges to a stationary point of $\tL_\tY(\cdot)$ module orthogonal transformation (see Supplementary Notes, Section~\ref{sec:SAlgorithm}). In theory, global optimality of non-convex optimization is often challenging to obtain. Fortunately, we will show in Section~\ref{subsec:statprob} that the desired statistical performance holds for all local optimizers satisfying $\tL_{\tY}(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)\geq \tL_{\tY} (\tC_{\textup{true}},\mM_{1,\textup{true}},\ldots,\mM_{K,\textup{true}})$, where the subscript ``true'' denotes the true parameter in model~\eqref{eq:decomp}. This result indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective is large enough. In the experiments we considered, we find that Algorithm~\ref{alg:B} typically gives satisfactory convergent points upon random initialization. Figure~\ref{fig:loglike} shows the trajectories of objective function for order-3 tensors based on model~\eqref{eq:decomp}, where $d_k = d \in\{25,30\}$, $p_k = 0.4d$, $r_k = r\in\{3,6\}$ for $ k = 1,2,3$. We consider input tensors with Gaussian, Bernoulli, and Poisson entries. Under all combinations of the dimension $d$, rank $r$, and type of the entries, Algorithm~\ref{alg:B} converges quickly in a few iterations, and the objective values at convergent points are close to or larger than the value at true parameters.

\begin{figure}[t]
\centering
\includegraphics[width=15cm]{loglike.pdf}
\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c) Poisson models. The dashed line represents the objective value at true parameters. }\label{fig:loglike}
\end{figure}

\subsection{Rank selection and time complexity}\label{sec:tuning}
Algorithm~\ref{alg:B} assumes the rank $\mr$ is given. In practice, the rank is often unknown and must be determined from the data. We propose to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
\begin{align}\label{eq:BIC}
\hat \mr&=\argmin_{\mr=(r_1,\ldots,r_K)} \textup{BIC}(\mr)\notag\\
&=\argmin_{\mr=(r_1,\ldots,r_K)}\left[-2\tL_{\tY}(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)+p_e(\mr)\log \left(\prod_k d_k\right) \right],
\end{align}
where $p_e(\mr)\stackrel{\textup{def}}{=}\sum_k (p_k-r_k-1)r_k+\prod_k r_k$ is the effective number of parameters in the model. We choose $\hat \mr$ that minimizes $\textup{BIC}(\mr)$ via grid search. Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model. We evaluate the empirical performance of BIC in Section~\ref{sec:simulation}.

The computational complexity of our Algorithm~\eqref{alg:B} is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the data tensor. More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor. The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$ feature matrix. Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$. This complexity in tensor dimension matches with the classical tensor decomposition~\citep{kolda2009tensor}.

\subsection{Statistical properties}\label{subsec:statprob}
In this section, we provide the accuracy guarantee for the proposed M-estimator~\eqref{eq:MLE}. Note that the factor matrices $\mM_k$ are identifiable only up to orthogonal transformation. Therefore, we use angle distance to assess the accuracy in estimating the column space, $\textup{Span}(\mM_k)$. For any two matrices $\mA,\mB$ with orthonormal columns of same dimension, the angle distance is defined as
\[
\sin \Theta(\mA,\mB)=\snormSize{}{\mA^T \mB^{\perp}}=\max\left\{ \frac{\langle \mx, \my\rangle}{\vnormSize{}{\mx}\vnormSize{}{\my}}\colon \ \mx\in \textup{Span}(\mA),\ \my\in \textup{Span}(\mB)\right\}.
\]

In modern applications, the tensor data and features are often large-scale. We are particularly interested in the high-dimensional regime in which both $d_k$ and $p_k$ diverge; i.e.\ $d_k\to \infty$ and $p_k\to\infty$, while $p_k/d_k \to \gamma_k \in[0,1)$. As the size of problem grows, and so does the number of unknown parameters. The classical MLE theory does not directly apply. We leverage the recent development in random tensor theory and high-dimensional statistics to establish the error bounds of the estimation.

\begin{assumption}\label{ass}We make the following assumptions:
\begin{enumerate}
\item [A1.] There exist two positive constants $c_1, c_2>0$ such that $c_1\leq \sigma_{\min}(\mX_k)\leq  \sigma_{\max}(\mX_k)\leq c_2$ for all $k\in[K]$. Here $\sigma_{min}(\cdot)$ and $\sigma_{max}(\cdot)$ denote the smallest and largest matrix singular values.
\item [A1'.] The feature matrices $\mX_k$ are Gaussian designs with i.i.d.\ $N(0,1)$ entries.
\item [A2.] There exist two positive constants $L, U>0$, such that $L\phi \leq \textup{Var}(y_{i_1,\ldots,i_K}|\theta_{i_1,\ldots,i_K})\leq U\phi $ for all $|\theta_{i_1,\ldots,i_K}|\leq \alpha$. Equivalently, $L\leq b''(\theta_{i_1,\ldots,i_K}) \leq U$ for all $|\theta_{i_1,\ldots,i_K}|\leq \alpha$. Here $\alpha$ is the upper bound of the linear predictor in~\eqref{eq:MLE}.

\end{enumerate}
\end{assumption}
The assumptions are fairly mild. Assumptions A1 and A1' consider two different scenarios about feature matrices. We study the cases A1 and A1' separately but have the same theoretical results. Assumption A1 is applicable when feature matrix is asymptotically non-singular and has bounded spectral norm, whereas Assumption A1' imposes the commonly-used Gaussian design in high-dimensional statistics~\citep{chen2019non}. Assumption A2 ensures the log-likelihood $\tL_\tY(\Theta)$ is strictly concave in the linear predictor $\Theta$. The two equivalent statements in Assumption A2 are based on the fact that $\textup{Var}(y_{i_1,\ldots,i_K}|\tX,\tB)=\phi b''(\theta_{i_1,\ldots,i_K})$ when $y_{i_1,\ldots,i_K}$ belongs to an exponential family~\citep{mccullagh1989generalized}.

\begin{thm}[Statistical convergence]\label{thm:main}
Consider a data tensor generated from model~\eqref{eq:decomp}, where the entries are conditionally independent realizations from an exponential family. Let $(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)$ be the M-estimator in~\eqref{eq:MLE} and $\hat \tB=\hat \tC\times \hat \mM_1\times\cdots \times \hat \mM_K$. Define $r_\textup{total}=\prod_k r_k$ and $r_{\max}=\max_k r_k$. Under Assumptions A1 and A2 with scaled feature matrices $\check \mX_k= \sqrt{d_k}\mX_k$, or under Assumptions A1' and A2 with original feature matrices, there exist two positive constants $C_1=C_1(\alpha,K), C_2=C_2(\alpha, K)>0$ independent of dimensions $\{d_k\}$ and $\{p_k\}$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$, 
\begin{equation}\label{eq:bound}
    \FnormSize{}{\trueB- \hat \tB}^2\leq \frac{ C_{2} r_{\textup{total}}}{r_{\max}}\frac{\sum_k p_k}{\prod_k d_k}.
\end{equation}
Furthermore, if the unfolded core tensor has non-degenerate singular values at mode $k\in[K]$, i.e., $\sigma_{\min}(\textup{Unfold}_k(\tC_\textup{true})) \geq c>0$ for some constant $c$, then
\begin{equation}\label{eq:sinebound}
\textup{sin}^2 \Theta(\trueM,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\sigma^2_{\min}(\textup{Unfold}_k(\tC_\textup{true}))}\frac{\sum_k p_k}{\prod_k d_k}.
\end{equation}
\end{thm}

Theorem~\ref{thm:main} establishes the statistical convergence for the estimator~\eqref{eq:MLE}. The proof is provided in Appendix~\ref{sec:appedix}. In fact, the proof shows that the desired convergence rate holds not only for the M-estimator, but also for any local optimizers satisfying $\tL_\tY(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K) \geq \tL_\tY(\tC_{\textup{true}},  \mM_{1,\textup{true}},\ldots,\mM_{K,\textup{true}})$. Consider a special case when tensor dimensions are equal on each of the modes, i.e., $d_k=d$ for all $k\in[K]$, and feature dimension grows with tensor dimension, $p_k=\gamma d$, $\gamma\in [0,1)$, for $k\in[K]$. The result in~\eqref{eq:bound} implies that the estimation has a convergence rate $\tO(d^{-(K-1)})$. Therefore, our estimation is consistent in high dimensional regimes, and the convergence becomes especially favorable as the order of tensor data increases.

As immediate applications, we obtain the convergence rate for the three examples mentioned in Section~\ref{sec:model}.

\begin{customexample}{1}[Spatio-temporal growth model] The estimated type-by-time-by-space coefficient tensor converges at the rate $\tO\left(\frac{p+r+s}{dmn}\right)$ where $p\leq d$, $r\leq m$ and $s\leq n$. The estimation achieves consistency as long as the dimension grows along either of the three modes.
\end{customexample}

\begin{customexample}{2} [Network population model] The estimated node-by-node-by-feature tensor converges at the rate $\tO\left(\frac{2d+p}{d^2n}\right)$ where $p\leq n$. The estimation achieves consistency as the number of individuals or the number of nodes grows. 
\end{customexample}

\begin{customexample}{3} [Dyadic data with node attributes] The estimated feature-by-feature matrix converges at the rate $\tO\left(\frac{p}{d^2}\right)$ where $p\leq d$. Again, our estimation achieves consistency as the number of nodes grows. 
\end{customexample}

We conclude this section by providing the prediction accuracy, measured in KL divergence, for the response distribution.

\begin{thm}[Prediction error]\label{thm:KL}
Assume the same set-up as in Theorem~\ref{thm:main}. Let $\mathbb{P}_{\tY_{\textup{true}}}$ and $\mathbb{P}_{\hat \tY}$ denote the distributions of $\tY$ given the true parameter $\trueB$ and estimated parameter $\hat \tB$, respectively. Then, we have, with probability at least $1-\exp(C_1\sum_k p_k)$,
\[
\textup{KL}(\mathbb{P}_{\tY_{\textup{true}}},\ \mathbb{P}_{\hat \tY})\leq \frac{C_3r_{\textup{total}}}{r_{\max}}\sum_k p_k,
\]
where $C_3=C_3(\alpha, K)>0$ is a constant independent of dimensions $\{d_k\}$ and $\{p_k\}$.
\end{thm}

\section{Numerical experiments}\label{sec:simulation}
We evaluate the empirical performance of our supervised tensor decomposition (STD) through simulations. We consider order-3 tensors with a range of distribution types. Unless otherwise specified, the conditional mean tensor is generated form model~\eqref{eq:decomp}, where the core tensor entries are i.i.d.\ drawn from Uniform[-1,1], the factor matrix $\mM_k$ is uniformly sampled with respect to Haar measure from matrices with orthonormal columns. The feature matrix $\mX_k$ is either an identity matrix (i.e.,\ no feature  available) or Gaussian random matrix with i.i.d.\ entries from $N(0,1)$. The linear predictor $\Theta=\tC\times\{\mM_1\mX_1,\mM_2\mX_3,\mM_3\mX_3\}$ is scaled such that $\mnormSize{}{\Theta}=1$. Conditional on the linear predictor $\Theta=\entry{\theta_{ijk}}$, the entries in the tensor $\tY=\entry{y_{ijk}}$ are drawn independently according to one of the following three probabilistic models:

\begin{enumerate}
\item[(a)] (Gaussian). Continuous entries $y_{ijk}\sim N\left(\alpha \theta_{ijk}, 1\right)$, for all $(i,j,k)\in[d_1]\times[d_2]\times[d_3]$.
\item[(b)] (Poisson). Count entries $y_{ijk}\sim\textup{Poisson}\left( e^{\alpha \theta_{ijk}}\right)$, for all $(i,j,k)\in[d_1]\times[d_2]\times[d_3]$.
\item[(c)] (Bernoulli). Binary entries $y_{ijk}\sim \textup{Bernoulli}\left( \frac{e^{\alpha \theta_{ijk}}}{1+e^{\alpha \theta_{ijk}}}\right)$, for all $(i,j,k)\in[d_1]\times[d_2]\times[d_3]$.
\end{enumerate}
Here $\alpha>0$ is a scalar controlling the magnitude of the effect size. In each simulation study, we report the mean squared error (MSE), $\FnormSize{}{\tB_{\textup{true}}-\hat \tB}^2$, for the coefficient tensor averaged across $n_{\textup{sim}}=30$ replications.

\subsection{Finite-sample performance}
The first experiment assesses the selection accuracy of our BIC criterion~\eqref{eq:BIC}. We consider the balanced situation where $d_k=d$, $p_k=0.4d_k$ for $k=1,2,3$. We set $\alpha=4$ and consider various combinations of dimension $d$ and rank $\mr=(r_1,r_2,r_3)$. For each combination, we simulate tensor data following Gaussian and Poisson models. We then minimize BIC using a grid search over three modes. The hyper-parameter $\alpha$ is set to infinity in the fitting, which essentially imposes no constraint on the tensor magnitude. Table~\ref{tab:rank} reports the selected rank averaged over $n_{\textup{sim}}=30$ replicates. We found that in the high-rank setting with $d=20$, the selected rank slightly underestimates the true rank, and the accuracy immediately improves when either the dimension increases to $d = 40$ or the rank reduces to $\mr = (3,3,3)$. This agrees with our expectation, because in the tensor decomposition, the sample size is related to the number of tensor entries. A larger $d$ implies a larger sample size, so the BIC selection becomes more accurate.

\begin{table}[tb]
\centering
\begin{tabular}{|c|cc|cc|}
\hline
True Rank &\multicolumn{2}{c|}{Dimension (Gaussian tensors)}&\multicolumn{2}{c|}{Dimension (Poisson tensors)}\\
$\mr$& $d=20$  &$d=40$&$d=20$ &$d=40$\\
\hline
$(3,\ 3,\ 3)$&$({\bf 3},\ {\bf 3},\ {\bf 3})$&$({\bf 3},\ {\bf 3},\ {\bf 3})$& $({\bf 3},\ {\bf 3},\ {\bf 3})$ & $({\bf 3},\ {\bf 3},\ {\bf 3})$\\
$(4,\ 4,\ 6)$&$(3,\ 3,\ {\bf 4.6})$&$({\bf 4},\ {\bf 4},\ {\bf 5.3})$&$(3,\ 3,\ {\bf 5.3})$&$({\bf 4},\ {\bf 4},\ {\bf 5.6})$\\
$(6,\ 8,\ 8)$&$(5,\ 5,\ 5)$&$({\bf 6},\ {\bf 8},\ {\bf 8})$&$(5,\ 5,\ 6.7)$&$({\bf 6},\ {\bf 8},\ {\bf 8})$\\
\hline
\end{tabular}
\caption{Rank selection via BIC. The estimated ranks are averaged across 30 simulation. Bold number indicates the ground truth is within two standard deviations of the estimate.}\label{tab:rank}
\end{table}

The second experiment evaluates the accuracy when features are available on all modes. We set $\alpha=10, d_k=d, p_k=0.4d_k, r_k=r\in\{2,4,6\}$ and increase $d$ from 30 to 60. Our theoretical analysis suggests that $\hat \tB$ has a convergence rate $\tO(d^{-2})$ in this setting. Figure~1 plots the estimation error versus the ``effective sample size'', $d^2$, under three different distribution models. We find that the empirical MSE decreases roughly at the rate of $1/d^2$, which is consistent with our theoretical results. We also observe that, tensors with higher rank tend to yield higher estimation errors, as reflected by the upward shift of the curves as $r$ increases. Indeed, a larger $r$ implies a higher model complexity and thus greater difficulty in the estimation. Similar behaviors are observed in the non-Gaussian tensors in Figures~\ref{fig:dim}b-c.

\begin{figure}[ht]
\centering
\includegraphics[width=15cm]{dimension.pdf}
\caption{Estimation error against effective sample size. The three panels plot the MSE when the response tensors are generated from (a) Gaussian (b) Poisson and (c) Bernoulli models. The dashed curves correspond to $\tO({1/d^2})$.}\label{fig:dim}
\end{figure}

\subsection{Comparison with GLMs under stochastic block models}
The third experiment investigates the performance of our model under correlated feature effects. We mimic the scenario of brain imaging analysis. A sample of $d_3=50$ networks are simulated, one for each individual. Each network measures the connections between $d_1=d_2=20$ brain nodes. We simulate $p=5$ features for the each of the 50 individuals. These features may represent, for example, age, gender, cognitive score, etc. Recent study has suggested that brain connectivity networks often exhibit community structure represented as a collection of subnetworks, and each subnetwork is comprised of a set of spatially distributed brain nodes. To accommodate this structure, we utilize the stochastic block model~\citep{abbe2017community} to generate the effect size. Specifically, we partition the nodes into $r$ blocks by assigning each node to a block with uniform probability. Edges within a same block are assumed to share the same feature effects, where the effects are drawn i.i.d.\ from $N(0,1)$. We then apply our tensor regression model to the network data using the BIC-selected rank. Note that in this case, the true model rank is unknown; the rank of a $r$-block network is not necessarily equal to matrix rank $r$~\citep{zeng2019multiway}.

\begin{figure}[ht]
\centering
\includegraphics[width=16cm]{compare_GLM.pdf}
\caption{Performance comparison under stochastic block models. The three panels plot the MSE when the response tensors are generated from (a) Gaussian (b) Poisson and (c) Bernoulli models. The $x$-axis represents the number of blocks in the networks. }\label{fig:glm}
\end{figure}

Figure~\ref{fig:glm} compares the MSE of our method with a multiple-response GLM approach. The multiple-response GLM is to regress the dyadic edges, one at a time, on the features, and this model is repeatedly fitted for each edge. As we find in Figure~\ref{fig:glm}, our tensor regression method achieves significant error reduction in all three data types considered. The outperformance is substantial in the presence of large communities; even in the less structured case ($\sim 20/15=1.33$ nodes per block), our method still outer-performs GLM. The possible reason is that the multiple-response GLM approach does not account for the correlation among the edges, and suffers from overfitting. In contrast, the low-rankness in our modeling incorporates the shared information across entries. By selecting the rank in a data-driven way, our method achieves accurate estimation in a wide range of settings.

\subsection{Comparison with other tensor methods}
We compare our supervised tensor decomposition with three other tensor methods:
\begin{itemize}
\item  Higher-order low-rank regression ({\bf HOLRR},~\citet{rabusseau2016low}) is a least-square based tensor regression that allows features on a single mode. 
\item Higher-order partial least square ({\bf HOPLS},~\citet{zhao2012higher}) is a dimension-reduction method that jointly models a tensor response and a tensor feature. 
\item Subsampled tensor projected gradient ({\bf TPG},~\citet{yu2016learning}) considers the same objective as {\bf HOLRR} but instead uses a different algorithm to solve the problem. 
\end{itemize}
These three methods are the closest algorithms to ours, in that they all relate a data tensor to features using a low-rank structure. The three existing methods allow only Gaussian data, whereas ours is applicable to any exponential family distribution including Gaussian, Bernoulli, Poisson, etc. For fair comparison, we consider Gaussian tensors in the experiment. Because not every method returns the effect estimate $\hat \tB$ as outputs, we measure the accuracy using mean squared prediction error, $\text{MSPE}=(\prod_kd_k)^{-1}\FnormSize{}{\hat \tY-f(\Theta)}^2$, where $f(\Theta)$ is the conditional mean of the tensor, and $\hat \tY$ is the fitted tensor from each method.

The comparison is assessed from three aspects: (a) benefit of incorporating features from multiple modes; (b) prediction error with respect to sample size; (c) sensitivity of accuracy with respect to model complexity. We use similar simulation setups as in our experiment II, but consider combinations of rank ($\mr=(3,3,3)$ vs.\ $(4,5,6)$), signal ($\alpha = 3$ vs.\ $6$), and dimension ($d$ ranging from 20 to 100 for modes with features, $d = 20$ for modes without features). Two methods ({\bf STD} and {\bf HOLRR}) require the tensor rank as inputs. For fair comparison, we provide both algorithms the true rank. For algorithms ({\bf HOPLS} and {\bf TPG}) that use different notions of model rank, we use a grid search to set the hyperparameter that gives the best mean square prediction error.

\begin{figure}[ht]
\centering
\includegraphics[width=16cm]{compare_alternative.pdf} 
\caption{Comparison between different tensor methods. Panels (a) and (b) plot MSPE versus the number of modes with available features. Panels (c) and (d) plot MSPE versus the effective sample size $d^2$. We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and signal $\alpha =3 $ (low), $\alpha=6$ (high).}~\label{fig:compare}
\end{figure}

Figures~\ref{fig:compare}a-b show the averaged prediction error across 30 replicates. We see that our {\bf STD} outperforms others, especially in the low-signal, high-rank setting. As the number of informative modes (i.e., modes with available features) increases, the {\bf STD} exhibits a substantial reduction in error whereas others remain unchanged (Figure~\ref{fig:compare}b). This showcases the benefit of incorporation of multiple features. Note that our method {\bf STD} is most comparable to {\bf HOLRR} when there is only a single informative mode. In such a case, both methods share a same cost function but have different algorithms. {\bf STD} alternates between informative and non-informative modes, whereas {\bf HOLRR} approximates the non-informative modes without alternating. The accuracy gain in Figure~\ref{fig:compare} demonstrates the benefit of alternating algorithm -- incorporation of informative modes also improves the estimation in the non-informative modes.

Figures~\ref{fig:compare}c-d compare the prediction error with respect to effective sample size when only one mode has side information. In the high-signal low-rank setting, our method has similar performance as {\bf HOLRR}, and the improvement becomes more pronounced in the low-signal high-rank setting. Note that the latter setting is a harder learning task because of the higher inter-mode complexity, and our {\bf STD} method shows the advantage in addressing this challenge.

\section{Data analyses}\label{sec:data}
We apply our supervised tensor decomposition to two datasets. The first application studies the brain networks in response to individual attributes (i.e.,\ feature on one mode), and the second application focuses on multi-relational network analysis with dyadic attributes (i.e.,\ features on two modes).

\subsection{Analyses of human brain connection data}
The Human Connectome Project (HCP) aims to build a network map that characterizes the anatomical and functional connectivity within healthy human brains~\citep{HCP}. We analyze the structural connectivity patterns among 68 brain regions for 136 individuals from HCP. We follow the preprocessing procedure as in~\cite{zhang2018mapping} and parcellate the brain into 68 regions of interest~\citep{desikan2006automated}. The preprocessed dataset is released in our R package \texttt{tensorregress}.

The dataset consists of 136 brain structural networks, one for each individual. Each brain network is represented as a 68-by-68 binary matrix, where the entries encode the presence or absence of fiber connections between the 68 brain regions. We consider four individual features: gender (65 females vs.\ 71 males), age 22-25 ($n=35$), age 26-30 ($n=58$), and age 31+ ($n=43$). The goal is to identify the connection edges that are affected by individual features. A key challenge in brain network is that the edges are correlated; for example, the nodes in edges may be from a same brain region, and it is of importance to take into account the within-dyad dependence.

\begin{figure}[!h]
\centering
\includegraphics[width=12cm]{HCP.pdf}
\caption{Top edges with large effects. (a) Global effect; (b) Female effect; (c) Age 22-25; (d) Age 31+. Red edges represent positive effects and blue edges represent negative effects. The edge-width is proportional to the magnitude of the effect size. }\label{fig:brain}
\end{figure}

We perform the supervised tensor decomposition to the HCP data. The data tensor is binary, $\tY\in\{0,1\}^{68\times 68\times 136}$, and the features are of dimension 4 on the 3$^{\text{rd}}$ mode. The BIC selection suggests a rank $\mr=(10,10,4)$ with quasi log-likelihood $\tL_{\tY}=-174654.7$. We utilize the sum-to-zero contrasts in coding the feature effects, and depict only the top 3\% edges whose connections are non-constant across the sample. Figure~\ref{fig:brain} shows the top edges with high effect size, overlaid on the Desikan atlas brain template~\citep{desikan2006automated,xia2013brainnet}. We find that the global connection exhibits clear spatial separation, and that the nodes within each hemisphere are more densely connected with each other (Figure~\ref{fig:brain}a). In particular, the superior-temproal (\emph{SupT}), middle-temporal (\emph{MT}) and Insula are the top three popular nodes in the network. Interestingly, female brains display higher inter-hemispheric connectivity, especially in the frontal, parental and temporal lobes (Figure~\ref{fig:brain}b). This is in agreement with a recent study showing that female brains are optimized for inter-hemispheric communication~\citep{ingalhalikar2014sex}. We also found several edges with declined connection in the group Age 31+. Those edges involve Frontal-pole (\emph{Fploe}), superior-frontal (\emph{SupF}) and Cuneus nodes. The Frontal-pole region has long been known for its importance in memory and cognition, and the detected decline with age further highlights its biological importance.

\begin{figure}[!h]
\centering
\includegraphics[width=16cm]{compare_HCP.pdf}
\caption{Comparison of feature effect estimation in the HCP data.}\label{fig:s1}
\end{figure}

Figure~\ref{fig:s1} compares the estimated coefficients from our method with those from multiple-response GLM approach. The multiple-response GLM is to regress the brain edges, one at a time, on the individual-level features, and this logistic model is repeatedly fitted for every edge $\in [68]\times [68]$. As we can see in the figure, our tensor decomposition method shrinkages the coefficients towards center, thereby implicitly enforcing the sharing between feature effects.

\subsection{Application to political relation data}

The second application studies the multi-relational networks with node-level attributes. We consider \emph{Nations} dataset~\citep{nickel2011three} which records 56 relations among 14 countries between 1950 and 1965 (post-World War II). The multi-relational networks can be organized into a $14 \times 14 \times 56$ binary tensor, with each entry indicating the presence or absence of an action, such as ``sending tourist to'', ``export'', ``import'', between countries. The 56 relations span the fields of politics, economics, military, religion, and so on. In addition, country-level attributes are also available, and we focus on the following six features: \emph{constitutional, catholics, law ngo, political leadership, geography}, and \emph{medicine ngo}. The goal is to identify the variation in connections due to country-level attributes and interactions thereof. One of the key features is that the 56 relations are correlated, and we would like to take it into account in assessing the feature effects.

\begin{table}[H]
\begin{tabular}{c|c|l}
Cluster & Category & \multicolumn{1}{c}{Relations}\\
\hline
\hline
\multirow{3}{*}{Cluster I}  & \multirow{3}{*}{Military} & warning,           violentactions, militaryactions, duration,\\ &&negativebehavior, severdiplomatic, expeldiplomats, protests,\\
&&timesincewar, commonbloc0, blockpositionindex, commonbloc1\\   
\hline
\multirow{4}{*}{Cluster II}& \multirow{4}{*}{Emigrant} & emigrants, relemigrants, emigrants3, accusation, \\      
&&nonviolentbehavior, students, relstudents, economicaid\\
&&releconomicaid, negativecomm, relintergovorgs,relngo\\
&&intergovorgs3, ngoorgs3, militaryalliance, commonbloc2\\    
\hline
\multirow{6}{*}{Cluster III} & \multirow{6}{*}{Economy} & treaties, reltreaties, oficialvisits, conferences, exportbooks\\
&& relexportbooks, booktranslations, relbooktranslations\\
&&boycottembargo, weightedunvote, unweightedunvote, tourism\\
&&reltourism, tourism3, exports, relexports, exports3, \\
&&intergovorgs, ngo, embassy, reldiplomacy,timesinceally,\\
&& dependent, independence\\
\hline
Cluster IV & Territory &aidenemy, lostterritory, unoffialacts, attackembassy\\
\hline
\end{tabular}
\caption{$K$-means clustering of relations based on dimension reduction on the 3$^\text{rd}$ mode. }\label{tab:s1}
\end{table}

We apply our tensor model to the \emph{Nations} data. The multi-relational network $\tY\in\{0,1\}^{14\times 14\times 56}$ is a binary data tensor, and the country attributes $\mX\in\mathbb{R}^{14\times 6}$ are features on both the 1$^\text{st}$ and 2$^\text{nd}$ modes. The BIC criterion suggests a rank $\mr=(4,4,4)$ for the coefficient tensor $\tB\in\mathbb{R}^{6\times 6 \times 56}$. 
We perform the supervised tensor decomposition and obtain the dimension reduction matrices $\hat \mM_k$ from the model. Then we apply $K$-mean clustering to dimension reduction matrix on each of the modes. Table~\ref{fig:s1} shows the K-means clustering of the 56 relations based on the dimension reduction matrix on the 3$^\text{rd}$ mode. We find that the relations reflecting the similar aspects of actions are grouped together. In particular, Cluster I consists of military relations such as \emph{violentactions}, \emph{warnings} and \emph{militaryactions}; Clusters II and III capture the economical relations such as \emph{economicaid, booktranslations, tourism}; and Cluster IV represents the political alliance and territory relations.

\begin{figure}[!h]
\centering
\includegraphics[width=12cm]{coef.pdf}
\caption{Estimated feature effects in the \emph{Nations} data analysis. Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warnning}, \emph{violentactions},\emph{treaties}, and \emph{aidenemy}, respectively. }\label{fig:est}
\end{figure}

To investigate the effects of dyadic attributes towards connections, we depict the estimated coefficients $\hat \tB=\entry{\hat b_{ijk}}$ for several relation types (Figure~\ref{fig:est}). Note that the each entry $\hat b_{ijk}$ estimates the contribution, at the logit scale, of feature pair $(i,j)$ ($i$th feature for the ``sender'' country and $j$th feature for the ``receiver'' country) towards the connection of relation $k$. Several interesting findings emerge from the observation. We find that relations belonging to a same cluster tend to have similar feature effects. For example, the relations ``warning'' and ''violentactions'' are classified into Cluster I, and both exhibit similar effect patterns (Figures~\ref{fig:est}a-b). Moreover, the feature \emph{constitutional} has a strong effect in the relation ``violentactions'' and ``warning'', whereas the effect is weaker in the relation ``treaties''. The result is plausible because the constitutional attributes affect political actions more often than economical actions. The entries in $\tB$ are useful for revealing interaction effects in a context-specific way. For a given relation $k$, the off-diagonal entries in the matrix $\hat \tB(\colon,\colon,k)$ represent the pairwise interactions between features. From Figure~\ref{fig:est}, we find a strong interaction between \emph{geography} and \emph{political leadership} in the relation ``warning'', and a strong interaction between \emph{geogrphy} and \emph{medicine ngo} in the relation ``aidenemy''. The relation-specific effect pattern showcases the applicability of our method in revealing complex interactions.

\section{Discussion and future work}\label{sec:con}
We have developed a supervised tensor decomposition method with side information on multiple modes. One important challenge of tensor data analysis is the complex interdependence among tensor entries and between multiple features. Our approach incorporates side information as feature matrices in the conditional mean tensor. The empirical results demonstrate the improved interpretability and accuracy over previous approaches. Applications to the brain connection and political relationship datasets yield conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.

We comment on a number of possible extensions. We have provided accuracy guarantees for parameter estimation in the supervised tensor model. Statistical inference based on tensor decomposition is an important future direction. Measures of uncertainty, such as confidence envelope for space estimation, would be useful. One possible approach would be performing parametric bootstrap~\citep{efron1994boot} to assess the uncertainty in the estimation. For example, one can simulate tensors from the fitted low-rank model based on the estimates, and then assess the empirical distribution of the estimates. While being simple, bootstrap approach is often computationally expensive for large-scale data. Another possibility is to leverage recent development in debiased inference with distributional characterization~\citep{chen2019inference,xia2019confidence}. This approach has led to fruitful results for matrix data analysis. Uncertainly quantification involving tensors are generally harder than matrices, and establishing distribution theory for tensor estimation remains an open problem.

One assumption made by our method is that tensor entries are conditionally independent given the linear predictor $\Theta$. This assumption can be extended by introducing a more general mixed-effect tensor model. For example, in the special case of Gaussian model, we can model the first two moments of data tensor using
\begin{align}\label{eq:extention}
\mathbb{E}(\tY|\mX_1,\ldots,\mM_K)&=\tC\times\mM_1\times \cdots \times \mM_K, \notag \\
\text{Var}(\tY|\mX_1,\ldots,\mM_K)&=\mPhi_1\otimes \cdots \otimes \mPhi_K,
\end{align}
where $\mPhi_k\in\mathbb{R}^{d_k\times d_k}$ is the unknown covariance matrix on the mode $k\in[K]$. For general exponential family, an additional mean-variance relationship should also be considered. The joint estimation of mean model $\Theta$ and variance model $\mPhi_k$ will lead to more efficient estimation in the presence of unmeasured confounding effects. The introduction of unknown covariance matrices $\mPhi_k$ dramatically increases the number of parameters in the problem. Suitable regularizatiosn such as graphical lasso or specially-structured covariance assumptions should be considered. The extension of tensor modeling with heterogeneous mixed-effects will be an important future direction.

Although we have presented the supervised tensor decomposition in the context of order-3 data tensors, the framework applies to more general multi-way datasets. One possible extension is the integrative analysis of omics data, in which multiple types of omics measurements (gene expression, DNA methylation, microRNA) are collected in the same set of individuals~\citep{lock2013joint,wang2019three}. In such a case, the observation is a stack of data matrices, which may not be of the same dimension. One possible extension is to impose linked dimension reduction among different data matrices, and incorporate side information as features in the joint decomposition. Other applications include time-series tensor data with multiple side information. Exploiting the benefits and properties of the supervised tensor decomposition in specialized task will boost scientific discoveries.

% for blind version
\section*{Acknowledgements}
This research was supported by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.

\bigskip
\begin{center}
{\large\bf SUPPLEMENTARY MATERIALS}
\end{center}

\begin{description}

\item[Supplementary notes:]  technical proofs and additional simulation results.

\item[Data and software:] R-package \texttt{tensorregress} performs our supervised tensor decomposition method. The source code and data used in the paper are available at \url{https://CRAN.R-project.org/package=tensorregress.} We also provide R scripts for reproducing all numbers in Figures 2-8 and Tables 2-3 in the paper.

\end{description}

\bibliography{tensor_wang}
\bibliographystyle{apalike}

\end{document}

### 2020-09-14 tensor_regress/JASA_submission/paper_source/supp.tex
clean

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url}
\usepackage{enumitem}

\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{algorithm}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={black},citecolor={black},urlcolor={black}}

\newcommand{\blind}{1}

\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{-.3in}%
\addtolength{\topmargin}{-.8in}%

\theoremstyle{definition}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{cor}{Corollary}[section]

\theoremstyle{definition}
\newtheorem{assumption}{Assumption}
\newtheorem{defn}{Definition}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\newtheorem{innercustomgeneric}{\customgenericname}
\providecommand{\customgenericname}{}
\newcommand{\newcustomtheorem}[2]{%
  \newenvironment{#1}[1]
  {%
   \renewcommand\customgenericname{#2}%
   \renewcommand\theinnercustomgeneric{##1}%
   \innercustomgeneric
  }
  {\endinnercustomgeneric}
}

\newcustomtheorem{customexample}{Example}

\usepackage{appendix}
\usepackage{wrapfig}
\mathtoolsset{showonlyrefs}

\usepackage{xr}
\externaldocument{regress}

\input macros.tex

\usepackage[english]{babel}

\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}

\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\def\ci{\perp\!\!\!\perp}

\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}
\usepackage{booktabs}
\newcommand\doubleRule{\toprule\toprule}
\allowdisplaybreaks

\usepackage[compact]{titlesec}

\usepackage{sectsty}
\sectionfont{\fontsize{12}{10}\selectfont}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
{\large\bf Supplementary Notes to ``Supervised Tensor Decomposition \\
with Interactive Side Information''}
\end{center}

\appendix
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\section{Proofs}\label{sec:appedix}
\begin{proof}[Proof of Theorem~\ref{thm:main}]
Let $\sigma_{\text{min}}^{(k)} = \sigma_{\text{min}}(\mX_k)$ and  $\sigma_{\text{max}}^{(k)} = \sigma_{\text{max}}(\mX_k)$. First we prove \eqref{eq:bound}. 
Define $\ell(\tB)=\mathbb{E}(\tL_{\tY}(\tB))$, where the expectation is taken with respect to $\tY\sim \trueB$ under the model with true parameter $\trueB$. We prove the following two conclusions:
\begin{enumerate}
\item[C1.] There exist two positive constants $C_1$, $C_2>0$, such that, with probability at least $1-\exp(-C_1\log K\sum_k p_k)$, the stochastic deviation, $\tL_{\tY}(\tB)-\ell(\tB)$, satisfies
\[
|\tL_{\tY}(\tB)-\ell(\tB)|=|\langle \tE,\ \tB\times_1\mX_1\times_2\cdots\times_K \mX_K\rangle| \leq C_2\FnormSize{}{\tB} \left(\prod_k \sigma_{\text{max}}^{(k)}\right)\sqrt{{\prod_k r_k \over \max_k r_k} \sum_k p_k}.
\]
\item[C2.] The inequality $\ell(\hat \tB) - \ell(\trueB) \leq  -{L\over 2}\FnormSize{}{\hat \Theta-\trueT}^2$ holds, where $L>0$ is the lower bound for $\min_{|\theta|\leq \alpha}|b''(\theta)|$. 
\end{enumerate}

To prove C1, we note that the stochastic deviation based on~\eqref{eq:loglikelihood} can be written as:
\begin{align}\label{bound}
\tL_{\tY}(\tB)-\ell(\tB)&=\langle \tY-\mathbb{E}(\tY|\trueT),\ \Theta(\tB)\rangle\notag \\
&= \langle \tY- b'(\trueT),\ \Theta\rangle \notag \\
&= \langle \tE\times_1\mX^T_1\times_2\cdots\times_K\mX^T_K,\ \tB\rangle,
\end{align}
where $\tE\stackrel{\text{def}}{=}\tY-b'(\trueT)$, and the second line uses the property of exponential family that $\mathbb{E}(\tY|\tX)=b'(\trueT)$. Based on Proposition~\ref{prop}, the boundedness of $b''(\cdot)$ implies that $\tE$ is a sub-Gaussian-$(\phi U)$ tensor. Let $\check\tE\stackrel{\text{def}}{=}\tE\times_1\mX^T_1\times_2\cdots\times_K\mX^T_K$. By Proposition~\ref{prop:sub}, $\check \tE$ is a $(p_1,\ldots,p_K)$-dimensional sub-Gaussian tensor with parameter bounded by $C=\phi U\prod_k \sigma_{\text{max}}^{(k)}$. Applying Cauchy-Schwarz inequality to~\eqref{bound} yields
\begin{equation}\label{eq:bound2}
|\tL_{\tY}(\tB)-\ell(\tB)|\leq \norm{\check \tE} \nnorm{\tB},
\end{equation}
where $\norm{\cdot}$ denotes the tensor spectral norm and $\nnorm{\cdot}$ denotes the tensor nuclear norm. The nuclear norm $\nnorm{\tB}$ is  bounded by $\nnorm{\tB}\leq \sqrt{{\prod_k r_k \over \max_k r_k}}\FnormSize{}{\tB}$~\citep{wang2017operator,wang2018learning}. The spectral norm $\norm{\check \tE}$ is bounded by $\norm{\check \tE}\leq C_2\prod_k \sigma_{\text{max}}^{(k)} \sqrt{\sum_k p_k}$ with probability at least $1-\exp(-C_1\log K \sum_kp_k)$ ~\citep{tomioka2014spectral}. Combining these two bounds with~\eqref{eq:bound2}, we have, with probability at least $1-\exp(-C_1\log K \sum_kp_k)$, 
\[
|\tL_{\tY}(\tB)-\ell(\tB)|\leq C_2\FnormSize{}{\tB}\left(\prod_k \sigma_{\text{max}}^{(k)}\right)  \sqrt{ {\prod_k r_k\over \max_k r_k}\sum_k p_k},
\]
where $C_2>0$ is a constant absorbing all factors that do not depend on $\{p_k\}$ and $\{r_k\}$.

Next we prove C2. Applying Taylor expansion to $\tL_{\tY}(\tB)$ around $\trueB$ yields
\begin{equation}\label{eq:logbefore}
\tL_{\tY}(\tB)=\tL_{\tY}(\trueB)+\Big\langle {\partial\tL_{\tY}(\tB) \over \partial
\tB}\big|_{\tB=\trueB},\tB-\trueB \Big\rangle + {1\over 2}\text{vec}(\tB-\trueB)^T \tH(\check \tB)\text{vec}(\tB-\trueB),
\end{equation}
where $\tH_{\tY}(\check \tB)$ is the (non-random) Hession of ${\partial \tL_{\tY}^2 (\tB)\over\partial^2 \tB}$ evaluated at $\check \tB = \text{vec}(\alpha \tB+(1-\alpha)\trueB)$ for some $\alpha\in[0,1]$. Note that we have used the fact that $\mathbb{E}\left({\partial\tL_{\tY}(\tB) \over \partial \tB}\big|_{\tB=\trueB}\right)=0$. This is because $\tL_{\tY}(\Theta) $ is defined as $\langle \tY,\Theta\rangle - \sum_{i_1,\ldots,i_K}b(\Theta_{i_1,\ldots,i_K})$ and
\begin{align}
{\partial \tL_\tY(\theta_{i_1,\ldots,i_K})\over \partial \theta_{i_1,\ldots,i_K}}\big|_{\Theta=\trueT}& = y_{i_1,\ldots,i_K} - {\partial b(\theta_{i_1,\ldots,i_K})\over\partial \theta_{i_1,\ldots,i_K}}\big|_{\Theta=\trueT} \\
&= y_{i_1,\ldots,i_K} - \mathbb{E}(y_{i_1,\ldots,i_K}|\theta_{i_1,\ldots,i_K\text{true}}),
\end{align}
where the last equality has used the fact that $b'(\theta) = \mathbb{E}(y|\theta)$. Taking derivative with respect to $\tB$ has the same result because of the chain rule.

We take expectation with respect to $\tY\sim \trueB$ on both sides of~\eqref{eq:log} and obtain
\begin{equation}\label{eq:log}
\ell(\tB)=\ell(\trueB)+{1\over 2}\text{vec}(\tB-\trueB)^T \tH(\check \tB)\text{vec}(\tB-\trueB).
\end{equation}
By the fact ${\partial \tL_{\tY}^2(\Theta)\over \partial^2 \Theta}=-b''(\Theta)$ and chain rule over $\Theta=\Theta(\tB)=\tB\times_1\mX_1\cdots\times_K\mX_K$, the equation~\eqref{eq:log} implies that 
\begin{equation}\label{eq:logB}
\ell(\tB)-\ell(\trueB)=-{1\over 2}\sum_{i_1,\ldots,i_K}b''(\check \theta_{i_1,\ldots,i_K}) (\theta_{i_1,\ldots,i_K}-\theta_{\text{true},i_1,\ldots,i_K})^2 \leq -{L \over 2}\FnormSize{}{\Theta-\trueT}^2,
\end{equation}
holds for all $\tB\in\tP$, provided that $\min_{|\theta|\leq \alpha}|b''(\theta)|\geq L>0$. In particular, the inequality~\eqref{eq:log} also applies to the constrained MLE $\hat \tB$. So we have
\begin{equation}\label{upper}
\ell(\hat \tB)-\ell(\trueB)\leq -{L \over 2}\FnormSize{}{\hat \Theta-\trueT}^2.
\end{equation}
Now we have proved both C1 and C2. Note that $\tL_{\tY}(\hat \tB)- \tL_{\tY}(\trueB)\geq 0$ by the definition of $\hat \tB$, This implies that
\begin{align}
0&\leq \tL_{\tY}(\hat \tB)- \tL_{\tY}(\trueB) \\
&\leq \left(\tL_{\tY}(\hat \tB)-\ell(\hat \tB)\right)-\left( \tL_{\tY}(\trueB)-\ell(\trueB)\right)+\left(\ell(\hat \tB)-\ell(\trueB)\right)\\
&\leq \langle\tE,\ \Theta-\trueT    \rangle -{L\over 2}\FnormSize{}{\hat \Theta-\trueT}^2,
\end{align}
where the second line follows from~\eqref{upper}. Therefore,
\begin{align}\label{eq:1}
\FnormSize{}{\hat \Theta-\trueT}&\leq {2\over L}\big\langle \tE,\ {\hat \Theta -\trueT \over \FnormSize{}{\hat \Theta-\trueT}} \big\rangle\notag \\
&\leq {2\over L}\sup_{\Theta: \FnormSize{}{\Theta}=1, \Theta=\tB\times_1\mX_1\times_2\cdots\times_K \mX_K}\langle \tE,\ \Theta \rangle\notag \\
&\leq {2\over L}\sup_{\tB\in\tP: \FnormSize{}{\tB}\leq \left(\prod_k \sigma_{\min}^{(k)}\right)^{-1}} \langle \tE,\ \ \tB\times_1\mX_1\times_2\cdots\times_K\mX_K\rangle.
\end{align}
Combining~\eqref{eq:1} with C1 yields 
\begin{align}\label{eq:theta}
\FnormSize{}{\hat \Theta-\trueT} \leq {2C_2\over L} \frac{\prod_k \sigma_{\max}^{(k)}}{\prod_k \sigma_{\min}^{(k)}}\sqrt{{\prod_k r_k \over \max r_k}\sum_k p_k}.
\end{align}
Therefore,
\begin{equation}\label{eq:Bbound}
\FnormSize{}{\hat \tB-\trueB}\leq \FnormSize{}{\hat \Theta-\trueT}\left(\prod_k \sigma_{\min}^{(k)}\right)^{-1}\leq {2C_2\over L} \frac{\prod_k \sigma_{\max}^{(k)}}{\left(\prod_k \sigma_{\min}^{(k)}\right)^2}\sqrt{{\prod_k r_k \over \max r_k}\sum_k p_k}.
\end{equation}
We now consider the two cases of assumptions on the feature matrices. 
\begin{itemize}[leftmargin=0cm]
\item[] [Case 1] Under Assumption A1 with scaled feature matrices, we have the singular value
\begin{equation}\label{eq:case1}
\frac{\prod_k \sigma_{\max}^{(k)}}{\left(\prod_k \sigma_{\min}^{(k)}\right)^2} = {\prod_k c_2\sqrt{d_k}\over \left(\prod_k c_1\sqrt{d_k}\right)^2}.
\end{equation}
\item[] [Case 2] Under Assumption A1', we have asymptotic behavior of extreme singular values  \citep{rudelson2010non} as 
\[
\sigma_{\text{min}}^{(k)} \sim \sqrt{d_k}-\sqrt{p_k} \text{ and  } \sigma_{\text{max}}^{(k)} \sim \sqrt{d_k}+\sqrt{p_k}.
\]
In this case, we obtain 
\begin{equation}\label{eq:case2}
\frac{\prod_k \sigma_{\max}^{(k)}}{\left(\prod_k \sigma_{\min}^{(k)}\right)^2} = {\prod_k (\sqrt{d_k}+\sqrt{p_k})\over \prod_k (\sqrt{d_k}-\sqrt{p_k})^2} =\frac{\prod_k(1+\sqrt{\gamma_k})\sqrt{d_k}}{\prod_k(1-\sqrt{\gamma_k})^2d_k}.
\end{equation}
\end{itemize}
Combining~\eqref{eq:Bbound} with either~\eqref{eq:case1} or~\eqref{eq:case2}, we obtain the same conclusion in both cases,
\[
\FnormSize{}{\hat \tB-\trueB}^2\leq \frac{ C_{2} r_{\textup{total}}}{r_{\max}}\frac{\sum_k p_k}{\prod_k d_k},
\]
where $C_2=C_2(\alpha,K,c_1, c_2)>0$ in the first case and $C_2=C_2(\alpha,K,\gamma)>0$ in the second case, both of which are constants that do not depend on the dimensions $\{d_k\}$ and $\{p_k\}$.

Now we prove \eqref{eq:sinebound}.
We unfold tensors $\trueB$ and $\hat\tB$ along the mode $k$ and obtain $\text{Unfold}_k(\trueB)$ and $\text{Unfold}_k(\hat\tB)$. Notice that 
\begin{align}
    \text{Unfold}_k(\trueB) &= \mM_k \text{Unfold}_k(\trueC)\left(\mM_{k+1}\otimes\mM_{k+2}\otimes \cdots\otimes \mM_1\otimes \cdots\mM_{k-1}\right)^T,\\
    \text{Unfold}_k(\hat\tB) &= \hat\mM_k \text{Unfold}_k(\hat\tC)\left(\hat\mM_{k+1}\otimes\hat\mM_{k+2}\otimes \cdots\otimes \hat\mM_1\otimes \cdots\hat\mM_{k-1}\right)^T,
\end{align} where $\otimes$ denotes the Kronecker product of matrices.
Notice that $\mM_k$ and $\hat\mM_k$ have the same image of left singular matrices of $\text{Unfold}_k(\trueB)$ and $\text{Unfold}_k(\hat\tB)$ respectively.  Applying Proposition \ref{prop:sinebound}, we have
\begin{align}
    \text{sin}\Theta^2(\mM_k,\hat\mM_k)\leq \frac{\FnormSize{}{\text{Unfold}_k(\hat\tB)-\text{Unfold}_k(\trueB)}}{\sigma_{\text{min}}(\text{Unfold}_k(\trueB))}= \frac{\FnormSize{}{\hat\tB-\trueB}}{\sigma_{\text{min}}(\text{Unfold}_k(\trueC))},
\end{align} where $\sigma_{\text{min}}(\text{Unfold}_k(\trueB)) = \sigma_{\text{min}}(\text{Unfold}_k(\trueC))$ holds based on the orthonormality of factor matrices. We finally prove \eqref{eq:sinebound} combining \eqref{eq:bound}.
\end{proof}

\begin{prop}[sub-Gaussian tensors]\label{prop:sub}
Let $\tS$ be a sub-Gaussian-$(\sigma)$ tensor of dimension $(d_1,\ldots,d_K)$, and $\mX_k\in\mathbb{R}^{p_k\times d_k}$ be non-random matrices for all $k\in[K]$. Then $\tE=\tS \times_1  \mX_1\times_2\cdots\times_K  \mX_K$ is a sub-Gaussian-$(\sigma')$ tensor of dimension $(p_1,\ldots,p_K)$, where $\sigma'\leq \sigma \prod_k\sigma_{\max}(\mX_k)$. Here $\sigma_{\max}(\cdot)$ denotes the largest singular value of the matrix. 
\end{prop}

\begin{proof}
To show $\tE$ is a sub-Guassian tensor, it suffices to show that the $\tE\times_1 \bmu^T_1\times_2\cdots\times_K \bmu^T_K$ is a sub-Gaussian scalar with parameter $\sigma'$, for any unit-1 vector $\bmu_k\in\mathbb{R}^{p_k}$, $k\in[K]$.

Note that, 
\begin{align}
\tE\times_1 \bmu^T_1\times_2\cdots\times_K \bmu^T_K&=\tS\times_1 (\bmu^T_1\mX_1) \times_2\cdots\times_K (\bmu^T_K\mX_K)\\
&= \left(\prod_k \vnormSize{}{\bmu^T_k\mX_k}\right) \KeepStyleUnderBrace{\left[\tS\times_1 { (\bmu^T_1\mX_1)\over \vnormSize{}{(\bmu^T_1\mX_1)}}\times_2\cdots\times_K { (\bmu^T_K\mX_K)\over \vnormSize{}{(\bmu^T_K\mX_K)}}\right]}_{\text{sub-Gaussian-$\sigma$ scalar}}.
\end{align}
Because $\vnormSize{}{(\bmu^T_k\mX_k)}\leq \sigma_{\max}(\mX^T_k)\vnormSize{}{\bmu_k}=\sigma_{\max}(\mX_k)$, we conclude that $\tE\times_1 \bmu^T_1\times_2\cdots\times_K \bmu^T_K$ is a sub-Gaussian tensor with parameter $\sigma \prod_k \sigma_{\max}(\mX_k)$. 
\end{proof}

\begin{prop}[sub-Gaussian residuals]\label{prop}
Define the residual tensor $\tE=\entry{\varepsilon_{i_1,\ldots,i_K}}=\tY-b'(\Theta)\in\mathbb{R}^{d_1\times \cdots \times d_K}$. Under the Assumption A2, $\varepsilon_{i_1,\ldots,i_K}$ is a sub-Gaussian random variable with sub-Gaussian parameter bounded by $\phi U$, for all $(i_1,\ldots,i_K)\in[d_1]\times\cdots\times[d_K]$.
\end{prop}
\begin{proof} The proof is similar to~\citet[Lemma 3]{fan2019generalized}. For ease of presentation, we drop the subscript $(i_1,\ldots,i_K)$ and simply write $\varepsilon$ ($=y-b'(\theta)$). For any given $t\in\mathbb{R}$, we have
\begin{align}
\mathbb{E}(\exp(t\varepsilon|\theta))&=\int c(x) \exp\left({\theta x - b(\theta)\over \phi}   \right)\exp \left(t(x-b'(\theta))\right)dx\\
&=\int c(x)\exp \left( {(\theta + \phi t)x - b (\theta+\phi t)+b(\theta+\phi t)-b(\theta)-\phi t b'(\theta) \over \phi}\right)dx\\
&=\exp\left( {b(\theta+\phi t)-b(\theta)-\phi t b'(\theta) \over \phi} \right)\\
&\leq \exp\left(\phi U t^2\over 2 \right),
\end{align}
where $c(\cdot)$ and $b(\cdot)$ are known functions in the exponential family corresponding to $y$. 
Therefore, $\varepsilon$ is sub-Gaussian-$(\phi U)$. \end{proof}

\begin{prop}[Wedin's $\sin\Theta$ Theorem]\label{prop:sinebound}
Let $\mB$ and $\hat\mB$ be two $m\times n$ real or complex with SVDs $\mB = \mU\Sigma\mV^T$ and $\hat\mB = \hat\mU\hat\mSigma\hat\mV^T.$ If $\sigma_{\text{min}}(\mB)>0$ and $\FnormSize{}{\hat\mB-\mB}\ll \sigma_{\text{min}}(\mB)$, then
\begin{align}\label{eq:sine}
    \text{sin}\Theta(\mU,\hat\mU) \leq \frac{C\FnormSize{}{\hat\mB-\mB}}{\sigma_{\text{min}}(\mB)},
\end{align}
for some constant $C>0$.
\end{prop}
\begin{proof}
From Theorem 6.1 in \cite{wang2017tensor}, we  obtain the following bound 
\begin{align}
    \max\left\{\snormSize{}{\text{sin}\Theta(\mU,\hat\mU)},\snormSize{}{\text{sin}\Theta(\mV,\hat\mV)}\right\}\leq \frac{\max\left\{\snormSize{}{\hat\mB\mV-\mU\mSigma},\snormSize{}{\hat\mB^T\mU-\mV\mSigma}\right\}}{\sigma_{\text{min}}(\mB)}.
\end{align}
Notice that
\begin{align}
    \snormSize{}{\hat\mB\mV-\mU\mSigma} &= \snormSize{}{\hat\mB\mV-\mB\mV} = \snormSize{}{\hat\mB-\mB} \leq \FnormSize{}{\hat\mB-\mB},\\
    \snormSize{}{\hat\mB^T\mU-\mV\mSigma} &= \snormSize{}{\hat\mB^T\mU-\mB^T\mU} = \snormSize{}{\hat\mB-\mB} \leq \FnormSize{}{\hat\mB-\mB}.
\end{align}
Therefore,we prove \eqref{eq:sine}.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:KL}]
The proof is similar to~\citet{pmlr-v108-berthet20a}. We sketch the main steps here for completeness. Recall that $\ell(\tB)=\mathbb{E}(\tL_{\tY}(\tB))$. By the definition of KL divergence, we have that,
\begin{align}
\ell(\hat \tB)&=\ell(\trueB)-\sum_{(i_1,\ldots,i_K)} KL(\theta_{\text{true}, i_1,\ldots, i_K}, \hat \theta_{i_1,\ldots,i_K})\\
&=\ell(\trueB)-\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY}),
\end{align}
where $\mathbb{P}_{\tY_{\text{true}}}$ denotes the distribution of $\tY|\tX$ with true parameter $\trueB$, and $\mathbb{P}_{\hat \tY}$ denotes the distribution with estimated parameter $\hat \tB$. Therefore
\begin{align}
\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY}) &= \ell(\trueB)-\ell(\hat \tB)\\
&={1\over 2}\sum_{i_1\ldots,i_K}b''(\check \theta_{i_1,\ldots,i_K})(\theta_{i_1,\ldots,i_K}-\theta_{\text{true},i_1,\ldots,i_K})^2\\
&\leq {U\over 2} \FnormSize{}{\Theta-\trueT}^2\\
&\leq C_4{\prod_k r_k \over \max r_k}\sum_k p_k,
\end{align}
where the second line comes from~\eqref{eq:log}, and the last line is derived from~\eqref{eq:theta}. Notice that $C_4 = C(\alpha,K,U,c_1,c_2)>0$ in Assumption 1 and $C_4(\alpha,K,U,\gamma)>0$ in Assumption 1' are constants that do not depend on the dimension $\{d_k\}$ and $\{p_k\}$. 
\end{proof}

\section{Algorithm properties}\label{sec:SAlgorithm}
In this section, we provide the convergence properties of Algorithm~\ref{alg:B}.
For notational convenience, we drop the subscript $\tY$ from the objective $\tL_\tY(\cdot)$ and simply write as $\tL(\cdot)$. Let $\tA = (\tC, \mM_1,\ldots,\mM_K) \in \mathbb{R}^{d_{\text{total}}}$ denote the collection of decision variables in the alternating optimization, where $d_{\text{total}} = \prod_k r_k + \sum_k r_kd_k$. We introduce the equivalent relationship induced by orthogonal transformation. Let $\mathbb{O}_{d,r}$ be the collection of all $d$-by-$r$ matrices with orthogonal columns, $\mathbb{O}_{d,r}:=\{\mP\in\mathbb{R}^{d\times r}\colon \mP^T\mP = \mathbf{1}_r\}$, where $\mathbf{1}_r$ is the $r$-by-$r$ identity matrix.

\begin{defn}[Equivalance relation]\label{eq:equivalance}
Two parameters $\tA'=(\tC', \mM'_1,\ldots,\mM'_k)$ and $\tA=(\tC, \mM_1,\ldots,\mM_k)$ are called equivalent, denoted $\tA\sim \tA'$, if and only if there exist a set of orthogonal matrices $\mP_k \in \mathbb{O}_{d_k,r_k}$ such that
\begin{align}
	\mM'_k \mP^T_k = \mM_k ,\ \forall k \in [K], \quad \text{and}\quad \tC' \times_1 \mP_1 \times_2  \dots \times_K \mP_K = \tC.
\end{align}
\end{defn}
Equivalently, two parameters $\tA,\ \tA'$ are equivalent if the corresponding Tucker tensors are the same, $\tB(\tA)=\tB'(\tA')$.

\begin{prop}[Global convergence]
Assume the set $\{\tA\ |\ \tL(\tA) \geq \tL(\tA^{(0)}) \}$ is compact and the stationary points of $\tL(\tA)$ are isolated module the equivalence defined in~\eqref{eq:equivalance}. Furthermore, assume that $\alpha = \infty$; i.e., we impose no entrywise bound constrains on the parameter space. Then any sequence $\tA^{(t)}$ generated by alternating algorithm converges to a stationary point of $\tL(\tA)$ module equivalence class.
\end{prop}

\begin{proof}
Pick an arbitrary iterate $\tA^{(t)}$.  Because of the compactness of set $\{\tA\colon  \tL(\tA) \geq \tL(\tA^{(0)}) \}$ and the boundedness of the decision domain, there exist a sub-sequence of $\tA^{(t)}$ that converges. Let $\tA^*$ denote one of the limiting points of $\tA^{(t)}$. Let $\tS = \{\tA^*\}$ denote the set of all the limiting points of $\tA^{(t)}$. We have $\tS \subset \{\tA\colon  \tL(\tA) \geq \tL(\tA^{(0)}) \}$ and thus $\tS$ is a compact set. By~\citet[Propositions 8.2.1 and 13.4.2]{Lange:2012:NAS:2432073}, $\tS$ is also connected. Note that all points in $\tS$ are also stationary points of $\tL(\cdot)$, because of the monotonic increase of $\tL(\tA^{(t)})$ as $t \rightarrow \infty$.

 Consider the equivalence of Tucker tensor representation of elements in $\tS$. We define an enlarged set $\tE_S$ induced by the equivalent class of elements in $\tS$,
\begin{equation}
 \tE_S = \left\{\tA\colon \tA\sim \tA^* \text{ for some }\tA^* \in \tS\right\}.
\end{equation}
The enlarged set $\tE_S$ satisfies the two properties below:
 \begin{enumerate}
\item [1.] [Union of stationary points] The set $\tE_S$ is an union of equivalent  classes generated by the limiting points in $\tS$.
\item [2.] [Connectedness module the equivalence]  The set $\tE_S$ is connected module the equivalence relationship. That property is obtained by the connectedness of $S$.
 \end{enumerate}
Now, note that the isolation of stationary points and Property 1 imply that $\tE_S$ contains only finite number of equivalent classes. Otherwise, there is a subsequence of non-equivalent stationary points whose limit is not isolated, which contradicts the isolation assumption. Combining the finiteness with Property 2, we conclude that $\tE_S$ contains only a single equivalent class; i.e. $\tE_S = \tE_{\{\tA^*\}}$, where $\tA^*$ is a stationary point of $\tL(\tA)$. Therefore, all the convergent sub-sequences of $\tA^{(t)}$ converge to one stationary point $\tA^*$ up to equivalence. We conclude that, any iterate $\tA^{(t)}$ generated by Algorithm 1 converges to a stationary point of $\tL(\tA)$ up to equivalence.
 \end{proof}

\newpage
\section{Additional simulation results}
Section~\ref{sec:simulation} in the main text has provided simulation results for two settings: low-signal, high-rank setting and high-signal, low-rank setting. Here, we perform the simulations for the full combinations of rank $\mr=(3,3,3), (4,5,6)$ and signal $\alpha=3, 6$. Figures~\ref{fig:S1} and~\ref{fig:S2} confirm the outperformance of the supervised tensor method in a range of model complexities.

\begin{figure}[!h]
\centering
\includegraphics[width=16cm]{Supp_Figure1.pdf} 
\caption{Comparison of MSPE versus the number of modes with features. We consider full combinations of rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and signal $\alpha=3$ (low), $\alpha=6$ (high).}~\label{fig:S1}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=16cm]{Supp_Figure2.pdf} 
\caption{Comparison of MSPE versus effective sample size. We consider full combinations rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and signal $\alpha=3$ (low), $\alpha=6$ (high). }~\label{fig:S2}
\end{figure}

\bibliography{tensor_wang}
\bibliographystyle{apalike}

\end{document}

### 2020-09-14 tensor_regress/oldcode/new_manuscript/tensor_new.tex
clean

--- 2020-09-08 tensor_regress/jasa-latexsource/tensor-jasa.tex
@@
-\documentclass[12pt]{article}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{enumerate}
+\documentclass[11pt]{article}
+
+\usepackage{fancybox}
+\usepackage{color}
+\usepackage{url}
+\usepackage[margin=1in]{geometry}
+\newcommand{\theHalgorithm}{\arabic{algorithm}}
+\usepackage{comment}
+\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
 \usepackage{natbib}
-\usepackage{url}
+\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
+\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}
+\usepackage{algorithm}
+%\usepackage{algorithmic}
+\usepackage{makecell}
 
-\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
-\usepackage{algorithm}
-\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
-\hypersetup{colorlinks,linkcolor={black},citecolor={blue},urlcolor={blue}}
-
-\newcommand{\blind}{1}
-
-\addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
-\addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{-.3in}%
-\addtolength{\topmargin}{-.8in}%
+\setlength\parindent{0pt}
+\usepackage[parfill]{parskip}
 
@@
 
-\usepackage{appendix}
+\usepackage{pifont}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+
 \usepackage{wrapfig}
 \mathtoolsset{showonlyrefs}
+
+%\linespread{1.5}
+
+\usepackage{enumitem}
+\usepackage{graphicx}
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{subfig}
+%\usepackage{algorithmic}
+\usepackage{appendix}
 
@@
 
-\def\ci{\perp\!\!\!\perp}
+%\externaldocument{tensor_regression_supp}
 
-\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}
+ %%% %%% %%% %%% %%% %%% %%% %%% %%% %%% %%% %%% %%% %%% %%% %%% %%% %%% %%% %%%
 \begin{document}
 
-\def\spacingset#1{\renewcommand{\baselinestretch}%
-{#1}\small\normalsize} \spacingset{1}
+\begin{center}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+{\Large \vspace*{.3cm}
+Exponential Family Tensor Regression with Covariates on Multiple Modes}
 
-\if1\blind
-{
-  \title{\bf Supervised Tensor Decomposition with Interactive Side Information}
-  %\author{Jiaxin Hu, Zhuoyan Xu, and Miaoyan Wang\\
-   % Department of Statistics, University of Wisconsin-Madison}
-  \maketitle
-} \fi
+\vspace*{1cm}
+Jiaxin Hu, Zhuoyan Xu, and Miaoyan Wang\footnote{Miaoyan Wang is Assistant Professor, Department of Statistics, University of Wisconsin-Madison, Madison, WI 53706, E-mail: miaoyan.wang@wisc.edu.
+Jiaxin Hu is a Ph.D.
+student in Statistics, E-mail: jhu267@wisc.edu.
+Zhuoyan Xu is a Ph.D.
+student in Statistics, E-mail: zxu444@wisc.edu.
+}
 
-\if0\blind
-{
-  \bigskip
-  \bigskip
-  \bigskip
-  \begin{center}
-    {\LARGE\bf Supervised Tensor Decomposition with Interactive Side Information}
+\vspace*{5mm}
+Department of Statistics, University of Wisconsin-Madison
 \end{center}
-  \medskip
-} \fi
 
-\bigskip
+\vspace*{.5cm}
+
 \begin{abstract}
-Higher-order tensors have received increased attention across science and engineering.
-While most tensor decomposition methods are developed for a single tensor observation, scientific studies often collect side information, in the form of node features and interactions thereof, together with the tensor data.
-Such data problems are common in neuroimaging, network analysis, and spatial-temporal modeling.
-Identifying the relationship between a high-dimensional tensor and side information is important yet challenging.
-Here, we develop a tensor decomposition method that incorporates multiple side information as interactive features.
-Unlike unsupervised tensor decomposition, our supervised decomposition captures the effective dimension reduction of the data tensor confined to feature space on each mode.
-An efficient alternating optimization algorithm is further developed.
+We consider the problem of tensor-response regression given covariates on multiple modes.
+Such data problems arise frequently in applications such as neuroimaging, network analysis, and spatial-temporal modeling.
+We propose a new family of tensor response regression models that incorporate covariates, and establish the theoretical accuracy guarantees.
+Unlike earlier methods, our estimation allows high-dimensionality in both the tensor response and the covariate matrices on multiple modes.
+An efficient alternating updating algorithm is further developed.
 Our proposal handles a broad range of data types, including continuous, count, and binary observations.
-We apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
-We identify the key global connectivity pattern and pinpoints the local regions that are associated with available features.
-Our method will help the practitioners efficiently analyze tensor datasets in various areas.
-Toward this end, the package and data used are available at~\url{https://CRAN.R-project.org/package=tensorregress}.
-
+Through simulation and applications to two real datasets, we demonstrate the outperformance of our approach over the state-of-art.
 \end{abstract}
 
-\noindent%
-{\it Keywords:} Applications and case studies, Tensor data analysis, Supervised dimension reduction, Exponential family distribution, Generalized multilinear model
-\vfill
+{\bf Keywords:} Tensor-response models, Multiple covariates, Dimension reduction, Reduced-rank regression, Generalized linear models
 
-\newpage
-\spacingset{1.5} % DON'T change the spacing!
 \section{Introduction}
-\label{sec:intro}
 
-Multi-dimensional arrays, known as tensors, are often collected with side information on multiple modes in modern scientific and engineering studies.
-A popular example is in neuroimaging~\citep{sun2017store,zhou2013tensor}.
+Nowadays, multi-dimensional arrays, a.k.a.\ tensors, are collected with additional covariates on multiple modes in many modern scientific and engineering studies.
+One example is neuroimaging~\citep{sun2017store,zhou2013tensor}.
 The brain connectivity networks are collected from a sample of individuals, accompanied by individual characteristics such as age, gender, and diseases status (see Figure~\ref{fig:intro1}a).
@@
 A typical social network consists of nodes that represent people and edges that represent the friendships.
-Side information such as people’s demographic information and friendship types are often available.
-In both examples, it is of keen scientific interest to identify the variation in the tensor data (e.g., brain connectivities, social community patterns) that is affected by available features.
-These seemingly different scenarios pose a common yet challenging problem for tensor data modeling.
+Other covariates such as people’s demographic information and the closeness of a friendship are often available.
+In both examples, it is of keen scientific interest to identify the variation in the response (e.g., brain connectivity, social community patterns) that may be affected by additional covariates.
+These seemingly different problems can be formulated as a regression problem in the tensor sense.
 
-In addition to the aforementioned challenges, many instances of tensor datasets consist of non-Gaussian measurements.
-Examples include the political interaction dataset \citep{hu2015scalable} which measures action counts between countries under various relationships, and the brain connectivity network dataset \citep{wang2019common} which is a collection of binary adjacency matrices.
-Classical tensor decomposition methods are based on minimizing the Frobenious norm of the reconstruction error, leading to suboptimal predictions for binary- or count-valued response variables.
-A number of supervised tensor methods have been proposed \citep{narita2012tensor, zhao2012higher, yu2016learning} to address the tensor regression problem in various forms (e.g.\ scalar-to-tensor regression, tensor-response regression).
-These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest.
-The gap in theory and practice suggests a great opportunity for improved modeling paradigms that better capture the complexity in tensor data with side information.
+Emerging instances of the non-Gaussian datasets also require the effective tools applicable to a wide range of data, such as continuous, binary and count data.
+Examples include the Political Social science data \citep{hu2015scalable} which is a count dataset, and the brain connectivity network data \citep{wang2019common} which is usually encoded as binary datasets.
+Easily treating these datasets as Gaussian-valued data will yield unstable estimations and lead the predictions fall outside the valid range.
+A number of previous methods have been proposed \citep{yu2016learning, llosa2018tensor, raskutti2015convex, zhao2012higher} to address the tensor regression problem in various forms (e.g.\ scalar-to-tensor regression, tensor-response regression) from distinct perspectives.
+However, these methods assume the Gaussianity in the model, treating the input as Gaussian valued entries.
+Novel methods for exponential family tensor regression are necessary this time.
 
@@
 \begin{center}
-\includegraphics[width=16cm]{new_introfigure1.pdf}
+\includegraphics[width=16cm]{demo.pdf}
 \end{center}
-\caption{Examples of supervised tensor decomposition with interactive side information. (a) Network population model. (b) Spatial-temporal growth model. }\label{fig:intro1}
+\caption{Examples of tensor response regression model with covariates on multiple modes. (a) Network population model. (b) Spatial-temporal growth model. }\label{fig:intro1}
 \end{figure*}
 
-We present a general model and associated method for decomposing a data tensor with exponential family entries and interactive side information. 
-We formulate the learning task as a structured regression problem, with tensor observation serving as the response, and the multiple side information as interactive features.
-Figure~\ref{fig:intro1}b illustrates our model in the special case of order-3 tensors.
-A low-rank structure is imposed to the conditional mean tensor, where unlike classical decomposition, the tensor factors $\mX_k\mM_k\in\mathbb{R}^{d_k\times r_k}$ belong to the space spanned by features $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,2,3$.
-The unknown matrices $\mM_k\in\mathbb{R}^{p_k\times r_k}$ (referred to as ``dimension reduction matrices'') link the conditional mean to the feature spaces, thereby allowing the identification of variations in the tensor data attributable to the side information.
+{\bf Related work.} Our work is closely related to but also clearly distinctive from several lines of previous work.
 
-In addition, we use tools from generalized linear model (GLM) to allow heteroscedacity due to the mean-variance relationship in the non-Gaussian data.
-This flexibility is important in practice.
-In classical GLM, the sample size and feature dimension are well defined; however, in the tensor data analysis, we observe only one single realization of tensor response but up to $K$ interactive feature matrices, where $K$ is the number of tensor modes.
-Both the sample size and feature dimension grow exponentially in $K$.
-We use a low-rank constraint in the conditional mean tensor (with a suitable transformation) to mitigate the curse of high dimensionality.
-The statistical convergence of our estimator is established, and we quantify the gain in prediction through several case studies.
+The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor,zhang2018tensor, hong2020generalized} that aims to find the best low-rank representation of data tensor.
+In contrast, our model can be viewed a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and covariates.
+The low-rank structure is determined jointly by the tensor response and matrix covariates.
 
-Our work is closely related to but also clearly distinctive from several lines of previous work.
-The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor,zhang2018tensor, hong2020generalized} that aims to find the best low-rank representation of a data tensor.
-In contrast, our model is a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and multiple features.
-The low-rank factorization is determined jointly by the tensor data and feature matrices.
+The second line of work studies the scalar-to-tensor regression models, in which the response is a scalar and the \textit{predictor} is a tensor~\citep{zhou2013tensor,chen2019non}.
+Our proposal is orthogonal to theirs because we treat the tensor as a \textit{response}.
+The distinct model settings yield distinct focuses and interpretations.
+Take neuroimaging analysis as an example.
+The scalar-to-tensor regression aims to study the changes of the prediction outcomes when the brain connectivity network varies, while tensor-response model focuses on understanding the variation in the brain connectivity that may attribute to the individual features such as age, gender and disease status.
+Since scalar-to-tensor models do not utilize all the information in the response, they are not optimal when there exist correlations among response variables.
+In contrast, our model does not assume the independence among response entries, which allow the model to capture more structural information.
 
-The second line of work studies the scalar-to-tensor regression, in which the response is a scalar and the predictor is a tensor~\citep{zhou2013tensor,chen2019non}.
-Our proposal is orthogonal to this line because we treat the data tensor as random variables.
-The difference in model settings yields different focuses and interpretations.
-Consider the neuroimaging analysis as an example.
-The scalar-to-tensor regression aims to predict the people's disease status based on brain connectivity network, while tensor-to-scalar model focuses on understanding the variation in the brain connectivity attributable to the individual features such as age, gender, and disease status.
-The scalar-to-tensor model is insufficient for tensor dimension reduction because it lacks the modeling of stochastic noise in the data tensor.
+The third line of work studies the tensor-on-tensor regression models, in which both the response and predictor are tensors~\citep{raskutti2015convex, lock2018tensor, llosa2018tensor, gahrooei2020multiple}.
+Though our model actually tackles the similar problem as tensor-on-tensor regression, our work differs with previous work in several ways.
+First, previous methods mainly develop the model with Gaussian error.
+These methods may not be applicable for other exponential family data, and their theoretical results may not be extended to generalized case trivially.
+Second, our study differs from previous work in the scope of results.
+\citep{gahrooei2020multiple} verifies the performance of estimation mainly via simulations, while our study provides a theoretical evidence on estimation.
+\citep{lock2018tensor, llosa2018tensor} perform solid inferences on parameter estimations, however, they do not assess the distance between estimation and true parameters and the prediction error, which are studied in our paper.
+Risk bounds are well studied in~\citep{raskutti2015convex} and our convergence rate coincides with their results in the order-3 low-rank case.
+Whereas, the convexity assumption of the regularization in their model yields a convex optimization problem.
+In contrast,  our model tackles a non-convex optimization problem, and employee distinct techniques.
 
-The third line of work studies the tensor-on-tensor regression, in which both the response and predictor are tensors~\citep{raskutti2015convex, lock2018tensor, gahrooei2020multiple}.
-Our model shares a common ground but provides a more efficient context-specific solutions than earlier approaches.
-As we show in Section~\ref{subsec:statprob}, the supervised tensor decomposition
-has an interesting interpretation as a special tensor-on-tensor regression.
-Nevertheless, our work improves from previous work in scope and applicability.
-Previous methods~\citep{gahrooei2020multiple,lock2018tensor} mainly focus on Gaussian tensors.
-The Frobenius norm used in the objective function is statistical suboptimal for general exponential family tensors.
-Maximum likelihood estimator (MLE) is studied in~\cite{raskutti2015convex} and a convex relaxation algorithm is proposed to solve for low-rank tensor coefficients.
-However, in the tensor case, convex MLE suffers from both computational intractability and statistical suboptimality.
-We advocate a non-convex approach and provide strong evidence for its success in our setting.
-Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2015convex,gahrooei2020multiple}, and we go step further by finding the sufficient dimension reduction~\citep{adragni2009sufficient}, $\text{Span}(\mM_k)$, that facilitates the identification of effective features in prediction (see Figure~\ref{fig:intro1}b).
-The latter approach greatly improves the \emph{interpretability} in prediction.
-In this regards, our method opens up new opportunities for tensor data analysis in a wider range of possible applications.
+{\bf Our contribution.} In this article, we present a general model and the associated theory for exponential family tensor regression.
+Exponential family tensor observation serves as a response, and the node features and/or the interactions forms the matrix predictor.
+Figure~\ref{fig:intro1}b illustrates the general set-up we consider.
+Due to the generalized framework, our models allows heteroscedasticity by relating the variance of tensor entry to its mean.
+This ﬂexibility is particularly important in practice.
+The regression approach allows the identiﬁcation of variation in the data tensor that is explained by the covariates.
+We utilize a low-rank constraint in the regression coeﬃcient to encourage the sharing among tensor entries.
+The statistical convergence of our estimator is established, and we quantify the gain in predictive power.
 
-\fixme{Miaoyan}{One more paragraph. Current literature on tensor decomposition with auxiliary side information...}
+\iffalse
+Many contemporary scientific and engineering studies collect multi-way array data, a.k.a.\ tensors, accompanied by additional covariates.
+One example is neuroimaging  analysis~\citep{sun2017store,zhou2013tensor}, in which the brain connectivity networks are collected from a sample of individuals.
+Researchers are often interested in identifying connection edges that are affected by individual characteristics such as age, gender, and disease status (see Figure~\ref{fig:intro1}a).
+Another example is in the field of network analysis~\citep{wu2017generalized,baldin2018optimal,hoff2005bilinear}.
+A typical social network consists of nodes that represent people and edges that represent friendships.
+In addition, features on nodes and edges are often available, such as people's personality and demographic location.
+It is of keen scientific interest to identify the variation in the connection patterns (e.g., transitivity, community) that can be attributable to the node features.
 
-The remainder of this paper is organized as follows.
-Section~\ref{sec:pre} introduces tensor notation and preliminaries.
-Section~\ref{sec:model} presents the main model and three motivating examples for supervised tensor decomposition.
-Section~\ref{sec:est} describes a rank-constrained likelihood-based estimation and associated alternating optimization algorithm.
-In Section~\ref{sec:simulation}, we present numerical experiments and assess the performance in comparison to alternative methods.
-In Section~\ref{sec:data}, we apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
-We conclude in Section~\ref{sec:con} with discussions about our findings and avenues of future work.
+This paper presents a general treatment to these seemingly different problems.
+We formulate the learning task as a regression problem, with tensor observation serving as a response, and the node features and/or their interactions forming the predictor.
+Figure~\ref{fig:intro1}b illustrates the general set-up we consider.
+The regression approach allows the identification of variation in the data tensor that is explained by the covariates.
+In contrast to earlier work~\citep{zhang2018network,wu2017generalized}, our method allows the contribution of covariates from multiple modes, whenever available.
+We utilize a low-rank constraint in the regression coefficient to encourage the sharing among tensor entries.
+The statistical convergence of our estimator is established, and we quantify the gain in accuracy compared to classical multivariate regression approach.
 
-\section{Preliminaries}\label{sec:pre}
+{\bf Related work.} Our work is closely related to but also clearly distinctive from several lines of previous work.
+The first is a class of \emph{unsupervised} tensor decomposition~\citep{de2000multilinear, kolda2009tensor,zhang2018tensor} that aims to find a low-rank representation of a data tensor.
+In contrast, our model can be viewed a \emph{supervised} tensor learning, which aims to identify the association between a data tensor and covariates.
+The second related line~\citep{zhou2013tensor,chen2019non} tackles tensor regression where the response is a scalar and the \emph{predictor} is a tensor.
+Our proposal is orthogonal to theirs because we treat the tensor as a \emph{response}.
+The tensor-response model is appealing for high-dimensional analysis when both the response and the covariate dimensions grow.
+The last line of work studies the network-response model~\citep{rabusseau2016low, li2017parsimonious}.
+The earlier development of this model focuses mostly on binary data in the presence of dyadic covariates~\citep{hoff2005bilinear,wu2017generalized}.
+We will demonstrate the enhanced accuracy as the order of data grows, and establish the general theory for exponential family which is arguably better suited to various data types.
+\fi
 
-We introduce the basic tensor properties used in the paper; more details on tensor notation can be found in~\citet{kolda2009tensor}.
-We use $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$\fixme{Miaoyan}{perhaps use $y_\omega$} to denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor, where $K$ corresponds to the number of modes of $\tY$ and is called the order.
+\vspace{-.1cm}
+\section{Preliminaries}
+\vspace{-.1cm}
+We begin by reviewing the basic properties about tensors~\citep{kolda2009tensor}.
+We use $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ to denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
 The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as
@@
 which results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor.
-For ease of presentation, we use the shorthand $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product.
-For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as
-\[
-\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.
-\] 
-The Frobenius norm and maximum norm of $\tY$ are defined as 
-\[
-\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}\quad \text{and}\quad \mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.
-\]
- A higher-order tensor can be reshaped into a lower-order object.
+For ease of presentation, we use shorthand notion $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product.
+For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as $\langle \tY, \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}$.
+The Frobenius norm of tensor $\tY$ is defined as $\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}$.
+A higher-order tensor can be reshaped into a lower-order object~\citep{wang2017operator}.
 We use $\text{vec}(\cdot)$ to denote the operation that reshapes the tensor into a vector, and $\text{Unfold}_k(\cdot)$ the operation that reshapes the tensor along mode-$k$ into a matrix of size $d_k$-by-$\prod_{i\neq k}d_i$.
-The multilinear rank of an order-$K$ tensor $\tY$ is defined as a length-$K$ vector $\mr=(r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\text{Unfold}_k(\tY)$, $k=1,\ldots,K$.
-Given a matrix $\mM\in\mathbb{R}^{d\times r}$, we use $\text{Span}(\mM)$ to denote the space spanned by columns of $\mM$.
-We use lower-case letters (e.g.,\ $a,\ b,\ c$) for scalars and vectors, upper-case boldface letters (e.g.,\ $\mA,\ \mB,\ \mC$) for matrices, and calligraphy letters (e.g.,\ $\tA,\ \tB,\  \tC$) for tensors of order three or greater.
-We let $\mI_d$ denote the $d \times d$ identity matrix and $[d]$ denote the $d$-set $\{1,\ldots,d\}$.
-For ease of notation, we allow the basic arithmetic operators (e.g.,\ $+$, $-$, $\geq $) and univariate
-functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner.
+The Tucker rank of an order-$K$ tensor $\tY$ is defined as a length-$K$ vector $\mr=(r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\text{Unfold}_k(\tY)$, $k=1,\ldots,K$.
+We use lower-case letters (e.g.,\ $a,b,c$) for scalars/vectors, upper-case boldface letters (e.g.,\ $\mA,\mB,\mC,$) for matrices, and calligraphy letters (e.g.,\ $\tA, \tB, \tC,$) for tensors of order three or greater.
+We let $\mI_d$ denote the $d \times d$ identity matrix, $[d]$ denote the $d$-set $\{1,\ldots,d\}$, and allow an $\mathbb{R}\to \mathbb{R}$ function to be applied to tensors in an element-wise manner.
 
 \section{Motivation and model}\label{sec:model}
-\subsection{General framework for supervised tensor decomposition}
-We begin with a general framework for supervised tensor decomposition and then discuss its implication in three concrete examples.
 Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor.
-Suppose the side information is available on each of the $K$ modes.
-Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the feature matrix on the mode $k\in[K]$, where $x_{ij}$ denotes the $j$-th feature value for the $i$-th tensor entity, for $(i,j)\in[d_k]\times[p_k]$, $p_k\leq d_k$.
-We propose a multilinear conditional mean model between the data tensor and feature matrices.
-Assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent realizations from an exponential family distribution, and the conditional mean tensor admits the form
+Suppose we observe covariates on some of the $K$ modes.
+Let $\mX_k\in\mathbb{R}^{d_k\times p_k}$ denote the available covariates on the mode $k$, where $p_k\leq d_k$.
+We propose a multilinear structure on the conditional expectation of the tensor.
+Specifically,
 \begin{align}\label{eq:tensormodel}
-&\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)=f(\Theta),\quad \text{with}\quad \Theta =\tB\times\{\mX_1,\ \ldots,\ \mX_K\} ,
+&\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)=f(\Theta),\ \text{with}\\
+&\Theta =\tB\times\{\mX_1,\ldots,\mX_K\} ,
 \end{align}
-where $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the unknown parameter tensor, $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$, and $\times$ denotes the tensor-by-matrix product.
-The choice of link function is based on the assumed distribution family of tensor entries.
-Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution.
-In general, dispersion parameters can also be included in the model.
-Because our main focus is the tensor decomposition under the mean model, we omit the dispersion parameter in this section for ease of presentation.
-
-In classical tensor decomposition, tensor factorization is performed on either data tensor $\tY$ or mean tensor $\mathbb{E}(\tY)$.
-In the context of supervised tensor decomposition, we propose to factorize the latent parameter tensor $\tB$,
-\begin{equation}\label{eq:factor}
-\tB=\tC\times\{\mM_1,\ \ldots,\ \mM_K\},
-\end{equation}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, and $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices consisting of orthonormal columns, where $r_k\leq p_k$ for all $k\in[K]$.
-By the definition of multilinear rank, model equations~\eqref{eq:tensormodel} and~\eqref{eq:factor} imply the low-rankness $\mr=(r_1,\ldots,r_K)$ of the conditional mean tensor under the link function.
-We now reach our final model for supervised tensor decomposition,
-\begin{align}\label{eq:decomp}
-\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1,\ \ldots,\ \mX_K\mM_K\}\right),\notag\\
-\text{where} \ &\ \mM_k\mM^T_k=\mI_{p_k},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\ \text{for all }k=1,\ldots,K,
-\end{align}
-where the parameters of interest are $\mM_k$ and $\tC$.
-Note that model~\eqref{eq:decomp} assumes a fixed, known rank $\mr=(r_1,\ldots,r_K)$; the adaptation to unknown rank will be addressed in Section~\ref{sec:tuning}.
-Figure~\ref{fig:intro1}b provides a schematic illustration of our model.
-The features $\mX_k$ affect the distribution of tensor entries in $\tY$ through the form $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$.
-We call $\mX_k\mM_k$ the ``supervised tensor factors'' or ``sufficient features''~\citep{adragni2009sufficient}, and call $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes.
-Our goal is to find $\mM_k$ and the corresponding $\tC$.
-Note that $\mM_k$ and $\tC$ are identifiable only up to orthonormal transformations.
-
-\subsection{Three examples}
-We give three concrete examples of supervised tensor decomposition model~\eqref{eq:decomp} that arises in practice. 
-\begin{example}[Spatio-temporal growth model]
-The growth curve model~\citep{gabriel1998generalised,potthoff1964generalized,srivastava2008models} was originally proposed as an example of bilinear model for matrix data, and we adopt its higher-order extension here.
-Let $\tY=\entry{y_{ijk}}\in\mathbb{R}^{d \times m\times n}$ denote the pH measurements of $d$ lakes at $m$ levels of depth and for $n$ time points.
-Suppose the sampled lakes belong to $q$ types, with $p$ lakes in each type.
-Let $\{\ell_j\}_{j\in[m]}$ denote the sampled depth levels and $\{t_k\}_{k\in[n]}$ the time points.
-Assume that the expected pH trend in depth is a polynomial of order at most $r$ and that the expected trend in time is a polynomial of order $s$.
-Then, the conditional mean model for the spatio-temporal growth model can be represented as
-\begin{equation}\label{eq:time}
-\mathbb{E}(\tY|\mX_1,\mX_2,\mX_3)=\tC\times\{\mX_1\mM_1,\ \mX_2\mM_2,\ \mX_3\mM_3\},
-\end{equation}
-where $\mX_1=\text{blockdiag}\{\mathbf{1}_p,\ldots,\mathbf{1}_p\}\in \{0,1\}^{d\times q}$ is the design matrix for lake types, and
-\[
-\mX_2=
-\begin{pmatrix}
-1 & \ell_1&\cdots &\ell^{r}_1\\
-1 & \ell_2&\cdots &\ell^{r}_2\\
-\vdots &\vdots&\ddots&\vdots\\
-1&\ell_{m}&\cdots&\ell^{r}_{m}
-\end{pmatrix},\quad
-\mX_3=
-\begin{pmatrix}
-1 & t_1&\cdots &t^{s}_1\\
-1 & t_2&\cdots &t^{s}_2\\
-\vdots &\vdots&\ddots&\vdots\\
-1&t_{n}&\cdots&t^{s}_{n}
-\end{pmatrix}
-\]
-are the design matrices for spatial and temporal effects, respectively, $\tC\in\mathbb{R}^{r_1\times r_2\times r_3}$ is the unknown core tensor, and $\mM_k$ are unknown dimension reduction matrices of coherent sizes.
-The factors $\mX_k\mM_k$ are sufficient features for the mean model~\eqref{eq:time}.
-The spatial-temporal model is a special case of our supervised tensor decomposition model, with features available on each of the three modes.
-\end{example}
-
-\begin{example}[Network population model]\label{example:brain}
-Network response model~\citep{rabusseau2016low, zhang2018network} is recently developed for neuroimanig analysis.
-The goal is to study the relationship between brain network connectivity pattern and features of individuals.
-Suppose we have a sample of $n$ observations, $\{(\mY_i, \mx_i)\colon i=1,\ldots,n\}$, where for each individual $i\in[n]$, $\mY_i\in\{0,1\}^{d\times d}$ is the symmetric adjacency matrix whose entries indicate presences/absences of connectivities between $d$ brain nodes, and $\mx_i\in\mathbb{R}^p$ is the individual's feature such as age, gender, cognition score, etc.
-The network-response model  has the conditional mean
-\begin{equation}\label{eq:network}
-\text{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n,
-\end{equation}
-where $\tB\in \mathbb{R}^{d\times d\times p}$ is a coefficient tensor of rank $r(\tB)=(r,r,r')$.
-
-The model~\eqref{eq:network} is a special case of our supervised tensor decomposition, with feature matrix on the last mode of the tensor.
-Specifically, we stack the network observations $\{\mY_i\}$ together and obtain an order-3 response tensor $\tY\in\{0,1\}^{d\times d\times n}$.
-Define a feature matrix $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$.
-Then, the model~\eqref{eq:network} has the equivalent representation of supervised tensor decomposition,
-\[
-\text{logit}(\mathbb{E}(\tY|\mX))=\tC\times\{\mM,\ \mM,\ \mX\mM'\},
-\]
-where $\tC\in\mathbb{R}^{r\times r\times r'}$ is the core tensor, $\mM\in\mathbb{R}^{d\times r}$ is the dimension reduction matrix at the first two modes, and $\mM'\in\mathbb{R}^{p\times r'}$ is for the last mode.
-\end{example}
-
- \begin{example}[Dyadic data with node attributes] Dyadic dataset consists of measurements on pairs of objects or under a pair of conditions.
-Common examples include graphs and networks.
-Let $\tG=(V,E)$ denote a graph, where $V=[d]$ is the node set of the graph, and $E\subset V\times V$ is the edge set.
-Suppose that we also observe feature vector $\mx_i\in\mathbb{R}^p$ associated to each $i\in V$.
-A probabilistic model on the graph $\tG=(V,E)$ can be described by the following matrix regression.
-The edge connects the two vertices $i$ and $j$ independently of other pairs, and the probability of connection is modeled as
-\begin{equation}\label{eq:edge}
- \text{logit}\left(\mathbb{P}\left((i,j)\in E\right)\right)=\mx^T_i\mB\mx_j=\langle \mB,\ \mx^T_i\mx_j\rangle,
- \end{equation}
- where $\mB\in\mathbb{R}^{p\times p}$ is a symmetric rank-$r$ matrix.
-The low-rankness in $\mB$ has demonstrated its success in modeling transitivity, balance, and communities in the networks~\citep{hoff2005bilinear}.
-We show that our supervised tensor decompostiion~\eqref{eq:tensormodel} also incorporates the graph model as a special case.
-Let $\tY=\entry{y_{ij}}$ be a binary matrix where $y_{ij}=\mathds{1}_{(i,j)\in E}$.
-Define $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$.
-Then, the graph model~\eqref{eq:edge} can be expressed as
- \[
- \text{logit}(\mathbb{E}(\mY|\mX))=\mC\times\{\mX\mM,\ \mX\mM\}, 
-  \]
-  where $\mC\in\mathbb{R}^{r\times r}$ and $\mM\in\mathbb{R}^{p\times r}$ are based on the singular value decomposition of $\mB=\mM\mC\mM^T$. 
-  \end{example}
-
-In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by features. 
-%The core tensor $\tC$ in our model model~\eqref{eq:tensormodel} serves this goal by collecting the sufficient feature effects and the interaction thereof.  
-\fixme{Miaoyan}{Unclear... To encourage the sharing among effects, decomposition~\eqref{eq:factor} assumes that the feature effects $\tB$ lie in a low-dimensional parameter space.
-%\begin{equation}\label{eq:rank}
-%\small \tP_{r_1,\ldots,r_K}=\{\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}: r_k(\tB)\leq r_k \text{ for all } k\in[K]\},
-%\end{equation}
-%where $r_k(\tB)\leq p_k$ is the rank at mode $k$ of the tensor. 
-The low-rank assumption is plausible in many scientific applications.
-In the Example~\ref{example:brain} of brain imaging analysis, for instance, it is often believed that the brain nodes can be partitioned into several functional and structural regions, and the person's attribute effect the connectivity in a similar way within a region.
-}
-
-%{\color{red}The low-rank structure encourages the shared information across tensor entries, thereby greatly improving the estimation stability. }
-%When no confusion arises, we drop the subscript $(r_1,\ldots,r_K)$ and write $\tP$ for simplicity.
-
-Our supervised tensor decomposition~\eqref{eq:decomp} is able to incorporate arbitrary numbers of feature matrices.
-We set $\mX_k=\mI_{d_k}$ if certain mode $k$ has no available side information.
-In particular, our model~\eqref{eq:decomp} reduces to classical unsupervised tensor decomposition~\citep{de2000multilinear,hong2018generalized} if no side information is available; i.e., $\mX_k=\mI_{d_k}$ for all $k\in[K]$.
-%Then, the final form of our tensor regression model can be written as:
-%\begin{align}\label{eq:tensormodel2}
-%&\mathbb{E}(\tY|\tX)=f(\Theta),\quad \Theta =\tC\times\{\mX_1\mM_k,\ldots,\mX_k\mK_k\},\notag\\
-%&\text{where}\ \mM^T_k\mM_k=1
-%\end{align}
-%where the entriesof $\tY$ are independent random variables conditional on $\tX$, and $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$ is the low-rank coefficient tensor of interest. 
-%We comment that other forms of tensor low-rankness are also possible, and here we choose Tucker rank just for parsimony.
-Similar models can be derived using various notions of low-rankness based on CP decomposition~\citep{hitchcock1927expression} and train decomposition~\citep{oseledets2011tensor}.
-
-\subsection{Connection to sufficient dimension reduction and tensor-on-tensor regression}
-An important feature of our method is that we allow high-dimensionality in both tensor dimension $d_k$ and feature dimension $p_k$.
-The model rank is typically smaller than tensor dimension $d_k$ and the feature dimension $p_k$.
-In such a case, the matrices $\mM_k$ serve the role of simultaneous dimension reduction of the data tensor and features.
-The implications of the sufficient features $\mX_k\mM_k$ can be seen in the following two conditional independence assumptions in model~\eqref{eq:decomp},
-\begin{align}
-\tY \ci \{\mX_k\}&\ \big|\ \{\mX_k\mM_k\} \ \ \text{(independence between the tensor and multiple features)},\\
-y_\omega \ci y_{\omega'} & \ \big|\ \{\mX_k\mM_k\}  \ \ \text{(independence within the tensor)},
-\end{align}
-where the second line holds for all $\omega\neq \omega' \in[d_1]\times\cdots\times[d_K]$, and $\ci$ denotes the independence.
-The first property highlights the ``decorrelation'' role of $\mM_k$ akin to the sufficient dimension reduction~\citep{adragni2009sufficient} in high-dimensional supervised learning, whereas the second property highlights the tensor dimension reduction in the usual unsupervised sense (consider, for example, $\mX_k=\mathbf{I}_{d_k}$ for all $k\in[K]$).
-
-Our model also has a close connection with tensor-on-tensor regression.
-Specifically, model~\eqref{eq:decomp} can be viewed as a multivariate regression model, where the response is vectorized tensor and the covariates are interactions between all sufficient features across modes.
-We take an order-3 tensor under the Gaussian model for illustration.
-Let $\mX,\mZ,\mW$ denote the feature matrix on mode $k=1, 2, 3$, respectively.
-Suppose that each mode has two-dimensional sufficient features, denoted $\mM_1\mX=[\mx_1,\mx_2]$, $\mM_2\mZ=[\mz_1,\mz_2]$, $\mM_3\mW=[\mw_1,\mw_2]$.
-Here $\mx_1,\mx_2,\ldots,\mw_1, \mw_2$ are column vectors.
-Then the model~\eqref{eq:decomp} can be written as a regression model with across-mode interactions,
-\begin{equation}\label{eq:interaction}
-\mathbb{E}(y_{ijk}|\mX,\mZ,\mW)=c_{111}\mx_{1i}\mz_{1j}\mw_{1k}+c_{121}\mx_{i1}\mz_{j2}\mw_{k1}+\cdots+c_{221}\mx_{2i}\mz_{2j}\mw_{1k}+c_{222}\mx_{2i}\mz_{2j}\mw_{2k},
-\end{equation}
-where $\entry{c_{ijk}} \in \mathbb{R}^{2\times 2\times 3}$ are unknown interaction effects, $\mx_{1i}$ denotes the $i$-th entry in the feature vector $\mx_1$, and similar notations apply to other features.
-Model~\eqref{eq:interaction} suggests the use of our supervised tensor decomposition for identifying across-mode interactions.
-Note that lower-order interactions are naturally incorporated in~\eqref{eq:interaction} if we include an intercept column in the sufficient feature matrices.
-
-\fixme{Miaoyan}{do we make use of the above two implications in the real data analysis or simulations?}
-
-\section{Estimation}\label{sec:est}
-
-\subsection{Rank-constrained M-estimator}
-We develop a likelihood-based procedure to estimate $\tC$ and $\mM_k$ in~\eqref{eq:decomp}.
-We adopt the exponential family as a flexible framework for different data types.
-In a classical generalized linear model (GLM) with a scalar response $y$ and feature $\mx$, the density is expressed as:
-\[
-p(y|\mx, \boldsymbol{\beta})=c(y,\phi)\exp\left({y\theta- b(\theta) \over \phi}\right)\ \text{with}\ \theta=\boldsymbol{\beta}^T\mx,
-\]
-where $b(\cdot)$ is a known function, $\theta$ is the linear predictor, $\phi>0$ is the dispersion parameter, and $c(\cdot)$ is a known normalizing function.
-The choice of link functions depends on the data types and on the observation domain of $y$, denoted $\mathbb{Y}$.
-For example, the observation domain is $\mathbb{Y}=\mathbb{R}$ for continuous data, $\mathbb{Y}=\mathbb{N}$ for count data, and  $\mathbb{Y}=\{0,1\}$ for binary data.
-Note that the canonical link function $f$ is chosen to be $f(\cdot)=b'(\cdot)$. Table~1 summarizes the canonical link functions for common types of distributions.
+where  $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the parameter tensor of interest, $f(\cdot)$ applied to $\Theta$ entrywise is a known link function in generalized linear model, and $\times$ denotes the tensor Tucker product.
+The choice of link function depends on the distribution of the response data.
+Some common choices are identity link for Gaussian tensor, logistic link for binary tensor, and $\exp(\cdot)$ link for Poisson tensor (see Table~\ref{table:link}).
 
@@
 
+We give three concrete examples of tensor regression that arise in practice. 
+\begin{example}[Spatio-temporal growth model]
+Let $\tY=\entry{y_{ijk}}\in\mathbb{R}^{d \times m\times n}$ denote the pH measurements of $d$ lakes at $m$ levels of depth and for $n$ time points.
+Suppose the sampled lakes belong to $q$ types, with $p$ lakes in each type.
+Let $\{\ell_j\}_{j\in[m]}$ denote the sampled depth levels and $\{t_k\}_{k\in[n]}$ the time points.
+Assume that the expected pH trend in depth is a polynomial of order $r$ and that the expected trend in time is a polynomial of order $s$.
+Then, the spatio-temporal growth model can be represented as
+\begin{equation}\label{eq:time}
+\mathbb{E}(\tY|\mX_1,\mX_2,\mX_3)=\tB\times\{\mX_1,\mX_2,\mX_3\},
+\end{equation}
+where $\tB\in\mathbb{R}^{p\times (r+1)\times (s+1)}$ is the coefficient tensor of interest, $\mX_1=\text{blockdiag}\{\mathbf{1}_p,\ldots,\mathbf{1}_p\}\in \{0,1\}^{d\times q}$ is the design matrix for lake types,
+\[
+\mX_2=
+\begin{pmatrix}
+1 & \ell_1&\cdots &\ell^{r}_1\\
+1 & \ell_2&\cdots &\ell^{r}_2\\
+\vdots &\vdots&\ddots&\vdots\\
+1&\ell_{m}&\cdots&\ell^{r}_{m}
+\end{pmatrix},\
+\mX_3=
+\begin{pmatrix}
+1 & t_1&\cdots &t^{s}_1\\
+1 & t_2&\cdots &t^{s}_2\\
+\vdots &\vdots&\ddots&\vdots\\
+1&t_{n}&\cdots&t^{s}_{n}
+\end{pmatrix}
+\]
+are the design matrices for spatial and temporal effects, respectively.
+The model~\eqref{eq:time} is a higher-order extension of the ``growth curve'' model originally proposed for matrix data~\citep{gabriel1998generalised,potthoff1964generalized,srivastava2008models}.
+Clearly, the spatial-temporal model is a special case of our tensor regression model, with covariates available on each of the three modes.
+
+\end{example}
+\begin{example}[Network population model] 
+Network response model is recently developed in the context of neuroimanig analysis.
+The goal is to study the relationship between network-valued response and the individual covariates.
+Suppose we observe $n$ i.i.d.\ observations $\{(\mY_i, \mx_i): i=1,\ldots,n\}$, where $\mY_i\in\{0,1\}^{d\times d}$ is the brain connectivity network on the $i$-th individual, and $\mx_i\in\mathbb{R}^p$ is the individual covariate such as age, gender, cognition, etc.
+The network-response model~\citep{rabusseau2016low, zhang2018network} has the form
+\begin{equation}\label{eq:network}
+\text{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n
+\end{equation}
+where $\tB\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor of interest.
+
+The model~\eqref{eq:network} is a special case of our tensor-response model, with covariates on the last mode of the tensor. Specifically, stacking $\{\mY_i\}$ together
+yields an order-3 response tensor $\tY\in\{0,1\}^{d\times d\times n}$, along with covariate matrix $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$.
+Then, the model~\eqref{eq:network} can be written as
+\[
+\text{logit}(\mathbb{E}(\tY|\mX))=\tB\times_3 \mX=\tB\times\{\mI_d, \mI_d, \mX\}.
+\]
+ \end{example}
+
+ \begin{example}[Dyadic data with node attributes] Dyadic dataset consists of measurements on pairs of objects or under a pair of conditions.
+Common examples include networks and graphs.
+Let $\tG=(V,E)$ denote a network, where $V=[d]$ is the node set of the graph, and $E\subset V\times V$ is the edge set.
+Suppose that we also observe covariate $\mx_i\in\mathbb{R}^p$ associated to each $i\in V$.
+A probabilistic model on the graph $\tG=(V,E)$ can be described by the following matrix regression.
+The edge connects the two vertices $i$ and $j$ independently of other pairs, and the probability of connection is modeled as
+\begin{equation}\label{eq:edge}
+ \text{logit}(\mathbb{P}((i,j)\in E)=\mx^T_i\mB\mx_j=\langle \mB, \mx^T_i\mx_j\rangle.
+ \end{equation}
+The above model has demonstrated its success in modeling transitivity, balance, and communities in the networks~\citep{hoff2005bilinear}.
+We show that our tensor regression model~\eqref{eq:tensormodel} also incorporates the graph model as a special case.
+Let $\tY=\entry{y_{ij}}$ be a binary matrix where $y_{ij}=\mathds{1}_{(i,j)\in E}$.
+Define $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$.
+Then, the graph model~\eqref{eq:edge} can be expressed as
+ \[
+ \text{logit}(\mathbb{E}(\mY|\mX))=\tB\times\{\mX,\mX\}.
+ \]
+\end{example}
+
+In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by the covariates.
+The regression coefficient $\tB$ in our model model~\eqref{eq:tensormodel} serves this goal by collecting the effects of covariates and the interaction thereof.
+To encourage the sharing among effects, we assume that the coefficient tensor $\tB$ lies in a low-dimensional parameter space:
+\begin{equation}\label{eq:rank}
+\small \tP_{r_1,\ldots,r_K}=\{\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}: r_k(\tB)\leq r_k \text{ for all } k\in[K]\},
+\end{equation}
+where $r_k(\tB)\leq p_k$ is the Tucker rank at mode $k$ of the tensor.
+The low-rank assumption is plausible in many scientific applications.
+In brain imaging analysis, for instance, it is often believed that the brain nodes can be grouped into fewer communities, and the numbers of communities are much smaller than the number of nodes.
+The low-rank structure encourages the shared information across tensor entries, thereby greatly improving the estimation stability.
+When no confusion arises, we drop the subscript $(r_1,\ldots,r_K)$ and write $\tP$ for simplicity.
+
+Our tensor regression model is able to incorporate covariates on any subset of modes, whenever available.
+Without loss of generality, we denote by $\tX=\{\mX_1,\ldots,\mX_K\}$ the covariates in all modes and treat $\mX_k=\mI_{d_k}$ if the mode-$k$ has no (informative) covariate.
+Then, the final form of our tensor regression model can be written as:
+\begin{align}\label{eq:tensormodel2}
+&\mathbb{E}(\tY|\tX)=f(\Theta),\quad \Theta =\tB\times\{\mX_1,\ldots,\mX_K\},\notag\\
+&\text{where}\ \text{rank}(\tB)\leq (r_1,\ldots,r_K),
+\end{align}
+where the entries of $\tY$ are independent r.v.'s conditional on $\tX$, and $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$ is the low-rank coefficient tensor of interest.
+We comment that other forms of tensor low-rankness are also possible, and here we choose Tucker rank just for parsimony.
+Similar models can be derived using various notions of low-rankness based on CP decomposition~\citep{hitchcock1927expression} and train decomposition~\citep{oseledets2011tensor}.
+
+\section{Rank-constrained likelihood-based estimation}
+We develop a likelihood-based procedure to estimate the coefficient tensor $\tB$ in~\eqref{eq:tensormodel2}.
+We adopt the exponential family as a flexible framework for different data types.
+In a classical generalized linear model (GLM) with a scalar response $y$ and covariate $\mx$, the density is expressed as:
+\[
+p(y|\mx, \boldsymbol{\beta})=c(y,\phi)\exp\left({y\theta- b(\theta) \over \phi}\right)\ \text{with}\ \theta=\boldsymbol{\beta}^T\mx,
+\]
+where $b(\cdot)$ is a known function, $\theta$ is the linear predictor, $\phi>0$ is the dispersion parameter, and $c(\cdot)$ is a known normalizing function.
+The choice of link functions depends on the data types and on the observation domain of $y$, denoted $\mathbb{Y}$.
+For example, the observation domain is $\mathbb{Y}=\mathbb{R}$ for continuous data, $\mathbb{Y}=\mathbb{N}$ for count data, and  $\mathbb{Y}=\{0,1\}$ for binary data.
+Note that the canonical link function $f$ is chosen to be $f(\cdot)=b'(\cdot)$. Table~1 summarizes the canonical link functions for common types of distributions.
+
 In our context, we model the the entries in the response tensor $y_{ijk}$ conditional on $\theta_{ijk}$ as independent draws from an exponential family.
-The quasi log-likelihood of~\eqref{eq:decomp} is equal (ignoring constant) to Bregman distance between $\tY$ and $b'(\Theta)$:
+The quasi log-likelihood of~\eqref{eq:tensormodel2} is equal (ignoring constant) to Bregman distance between $\tY$ and $b'(\Theta)$:
 \begin{align}
-\tL_{\tY}(\tC,\mM_k)&=\langle \tY, \Theta \rangle - \sum_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K}),\\
-\text{where}\ \Theta&=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
+\tL_{\tY}(\tB)&=\langle \tY, \Theta \rangle - \sum_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K}),\\
+\text{where}\ \Theta&=\tB\times\{\mX_1,\ldots,\mX_K\}.
 \end{align}
-
-We propose a constrained maximum likelihood estimator (M-estimator)
-%\begin{align} \label{eq:MLE} 
-%	\hat \tB &=\argmax_{\hat \tB \colon \text{rank}(\tB)\leq \mr, \mnormSize{}{\Theta(\tB)}\leq \alpha  } \tL_{\tY}(\tB).
-%\end{align}
+We assume that we have an additional information on an upper bound $\alpha>0$ such that $\mnormSize{}{\Theta}\leq \alpha$.
+This is the case for many applications we have in mind such as brain network analysis where fiber connections are bounded.
+We propose a constrained maximum likelihood estimator (MLE) for the coefficient tensor:
 \begin{align} \label{eq:MLE} 
-(\hat \tC, \hat \mM_1,\ldots,\hat \mM_k) &=\argmax_{(\tC,\mM_1,\ldots,\mM_K)\in \tP} \ \tL_{\tY}(\tC,\mM_1,\ldots,\mM_K),
+	\hat \tB &=\argmax_{\hat \tB \colon \text{rank}(\tB)\leq \mr, \mnormSize{}{\Theta(\tB)}\leq \alpha  } \tL_{\tY}(\tB).
 \end{align}
-where the parameter space $\tP$ is described as
-\[
-\tP=\left\{\tC\in\mathbb{R}^{r_1\times \cdots\times r_K},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\text{ for all }k\in[K]\ \Big| \ \mM_k\mM^T_k=\mathbf{I}_{p_k}, \mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha \right\}.
-\]
-The entry-wise bound assumption is a technical condition that avoids the divergence in non-Gaussian variance.
-
-%In the following theoretical analysis, we assume the rank $\mr=(r_1,\ldots,r_K)$ is known and fixed. The adaptation of unknown $\mr$ will be addressed in Section~\ref{sec:tuning}.
-
-\subsection{Alternating optimization} \label{subsec:alg}
-We propose an alternating optimization algorithm to solve~\eqref{eq:MLE}. 
-%The objective function $\tL_{\tY}(\tB)$ is concave in $\tB$ when the link $f$ is the canonical link function.
-We utilize a Tucker factor representation of the coefficient tensor $\tB$ and turn the optimization into a block-wise convex problem.
-%Specifically, write the rank-$\mr$ decomposition of coefficient tensor $\tB$ as
-%\begin{equation}\label{eq:tucker}
-%\tB=\tC\times \{\mM_1,\ldots,\mM_K\},
-%\end{equation}
-%where $\tC\in\mathbb{R}^{r_1\times\cdots\times r_K}$ is a full-rank core tensor, $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices with orthogonal columns.
-Estimating $\tB$ amounts to finding both the core tensor $\tC$ and the factor matrices $\mM_k$'s.
-The decision variables in the objective function~\eqref{eq:MLE} consist of $K+1$ blocks of variables, one for the core tensor $\tC$ and $K$ for the factor matrices $\mM_k$'s.
-We notice that, if any $K$ out of the $K+1$ blocks of variables are known, then the optimization reduces to a simple GLM with respect to the last block of variables.
-This observation suggests an iterative updating scheme for one block at a time while keeping others fixed.
-After each iteration, we rescale the core tensor $\tC^{(t+1)}$ subject to the maximum norm constraint.
-This post-processing in principle may not guarantee the monotonic increase of the objective, but we found that in our experiment this simple step appears robust for obtaining a desirable solution.
-The full algorithm is described in Algorithm~\ref{alg:B}.
-
-\begin{algorithm}[t]
-\caption{Supervised Tensor Decomposition with Interactive Side Information}\label{alg:B}
-\begin{algorithmic}[1]
-\INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, feature matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,\ldots,K$, target Tucker rank $\mr=(r_1,\ldots,r_K)$, link function $f$, maximum norm bound $\alpha$
-\OUTPUT Low-rank estimation for the coefficient tensor $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$. 
-\State Calculate $\check \tB=\tY\times_1 \left[(\mX_1^T\mX_1)^{-1}\mX^T_1\right] \times_2\cdots\times_K\left[(\mX_K^T\mX_K)^{-1}\mX^T_K\right] $.
-\State Initialize the iteration index $t=0$.
-Initialize the core tensor $\tC^{(0)}$ and factor matrices $\mM^{(0)}_k\in\mathbb{R}^{p_k\times r_k}$ via rank-$\mr$ Tucker approximation of $\check\tB$, in the least-square sense.
-  \While {the relative increase in objective function $\tL_\tY(\tB)$ is less than the tolerance}
-\State Update iteration index $t \leftarrow t+1$.
-\For { $k=1$ to $K$}
-\State Obtain the factor matrix $\tilde \mM_k^{(t+1)}\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$. 
-\State Perform QR factorization on $\tilde \mM_k^{(t+1)}=\mQ\mR$, where $\mQ\in\mathbb{R}^{p_k\times r_k}$ has orthogonal columns. 
-\State Update $\mM_k^{(t+1)}\leftarrow \mQ$ and core tensor $\tC^{(t+1)}\leftarrow \tC^{(t+1)}\times_k \mR$.
-\EndFor
-\State Obtain the core tensor $\tC^{(t+1)}\in\mathbb{R}^{r_1\times\cdots\times r_K}$ by solving a GLM with $\text{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM^{(t)}_k]$ as features, and $f$ as link function.
-Here $\otimes$ denotes the kronecker product of matrices.
-\State Rescale the core tensor subject to the maximum norm constraint. 
-\State Update $\tB^{(t+1)}\leftarrow \tC^{(t+1)}\times_1\mM_1^{(t+1)}\times_2\cdots\times_K\mM_K^{(t+1)}$.
-\EndWhile
-\end{algorithmic}
-\end{algorithm}
-
-The optimization~\eqref{eq:MLE} is a non-convex problem due to the non-convexity in the feasible set $\tP$.
-Global optimality of non-convex optimization of this type is often challenging to obtain.
-However, as mentioned in Section~\ref{subsec:statprob}, the desired convergence rate~\eqref{eq:bound} holds for the valid estimators satisfying $\tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true})$, which indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective of local optimums are large enough.
-\fixme{Miaoyan}{This is wrong.
-Proof of Theorem 4.1 is based on global optimality condition.
-Please relax the proof.} Fortunately, we find the  Algorithm~\ref{alg:B}  often gives a satisfactory convergence point $\hat \tB$, when we initialize the parameters via continuous-valued Tucker decomposition.
-Figure~\ref{fig:loglike} shows the trajectory of the objective function for order-3 tensors under the balanced setting, where $\alpha = 10, d_k = d, p_k = 0.4d, r_k = r$ for $ k = 1,2,3, p \in \{25,30\}$  and $r \in \{3, 6\}$.
-We consider the inputs with Gaussian, Bernoulli, and Poisson entries.
-Under all combinations of the dimension $d$, rank $r$, and type of the entries, Algorithm~\ref{alg:B} converges quickly in a few iterations, and the objective value at convergent points are close to or larger than the value at true parameters.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=15cm]{new_log.pdf}
-\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c) Poisson models.
-The dashed line represents the objective value at true parameter $\tL_{\tY}(\tB_{true})$.
-A micro-iteration refers to the update for one of the $K$ blocks.
-Four micro-iterations consist of a complete ``while'' iteration in Algorithm~\ref{alg:B}.}\label{fig:loglike}
-\end{figure}
-
-\subsection{Rank selection and time complexity}\label{sec:tuning}
-Algorithm~\ref{alg:B} takes the rank $\mr$ as an input.
-Estimating an appropriate rank given the data is of practical importance.
-We propose to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
-\begin{align}\label{eq:BIC}
-\hat \mr&=\argmin_{\mr=(r_1,\ldots,r_K)} \text{BIC}(\mr)\\
-&=\argmin_{\mr=(r_1,\ldots,r_K)}\left[-2\tL_{\tY}(\hat \tB)+p_e(\mr)\log \left(\prod_k d_k\right) \right],
-\end{align}
-where $p_e(\mr)\stackrel{\text{def}}{=}\sum_k (p_k-r_k-1)r_k+\prod_k r_k$ is the effective number of parameters in the model.
-We choose $\hat \mr$ that minimizes $\text{BIC}(\mr)$ via grid search.
-Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model.
-We test its empirical performance in Section~\ref{sec:simulation}.
-
-The computational complexity of our alternating optimization algorithm is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the response tensor.
-More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor.
-The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$ feature matrix.
-Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$.
+In the following theoretical analysis, we assume the rank $\mr=(r_1,\ldots,r_K)$ is known and fixed. The adaptation of unknown $\mr$ will be addressed in Section~\ref{sec:tuning}.
 
 \subsection{Statistical properties}\label{subsec:statprob}
-We assess the estimation accuracy using the deviation in the Frobenius norm. \fixme{Miaoyan}{needs revision...} For the true coefficient tensor $\trueB$ and its estimator $\hat \tB$, define
+We assess the estimation accuracy using the deviation in the Frobenius norm. For the true coefficient tensor $\trueB$ and its estimator $\hat \tB$, define
 \[
@@
 
-In modern applications, the response tensor and features are often large-scale.
+In modern applications, the response tensor and covariates are often large-scale.
 We are particularly interested in the high-dimensional region in which both $d_k$ and $p_k$ diverge; i.e.\ $d_k\to \infty$ and $p_k\to\infty$, while ${p_k\over d_k} \to \gamma_k \in[0,1)$.
@@
 The assumptions are fairly mild.
-Assumption A1 guarantees the non-singularity of the feature matrices, and Assumption A2 ensures the log-likelihood $\tY(\Theta)$ is strictly concave in the linear predictor $\Theta$.
+Assumption A1 guarantees the non-singularity of the covariates, and Assumption A2 ensures the log-likelihood $\tY(\Theta)$ is strictly concave in the linear predictor $\Theta$.
 Assumption A2 and A2' are equivalent, because $\text{Var}(y_{i_1,\ldots,i_K}|\tX,\tB)=\phi b''(\theta_{i_1,\ldots,i_K})$ when $y_{i_1,\ldots,i_K}$ belongs to an exponential family~\citep{mccullagh1989generalized}.
@@
 \begin{thm}[Statistical convergence]\label{thm:main}
-Consider an order-$K$ dimensional-$(d_1,\ldots,d_K)$ tensor $\tY$ and multiple feature matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$, for $k\in[K]$.
-Suppose that, conditional on the features, the entries in $\tY$ are independent realizations from an exponential family distribution with conditional mean model~\eqref{eq:decomp}.
-Let $r_\text{total}=\prod_k r_k$, $r_{\max}=\max_k r_k$, and $\sigma_{\min}(\mM_k)$ be the minimal non-zero singular value of $\mM_k$.
-Assume that $\sigma_{\min}(\mM_k)>0$ and $\FnormSize{}{\mX_k} \asymp \Omega(\sqrt{d_k})$ as $d_k\to\infty$, for all $k\in[K]$.
+Consider a generalized tensor regression model with covariates on multiple modes $\tX=\{\mX_1,\ldots,\mX_K\}$.
+Suppose the entries in $\tY$ are independent realizations of an exponential family distribution, and $\mathbb{E}(\tY|\tX)$ follows the low-rank tensor regression model~\eqref{eq:tensormodel2}.
 Under Assumption~\ref{ass}, there exist two constants $C_1, C_2>0$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$,
 \begin{equation}\label{eq:bound}
-\textup{sin}^2 \Theta(\mM_k,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\lambda_{\min}(\mM_k)}\frac{\sum_k p_k}{\prod_k d_k}, \quad\text{for all }k\in[K],
+\text{Loss}(\trueB,\ \hat \tB) \leq \frac{ C_{2}\prod_kr_k}{\max_k r_k}\sum_k p_k.
 \end{equation}
-and
-\[
-\FnormSize{}{\trueB- \hat \tB}^2\leq \frac{ C_{2} r_{\textup{total}}}{r_{\max}}\frac{\sum_k p_k}{\prod_k d_k}.
-\]
-Here, $C_2=C_2(\mr,\alpha, K)>0$ is a constant independent of the dimensions $\{d_k\}$ and $\{p_k\}$. 
+Here, $C_2=C_2(\mr,\alpha, K)>0$ is a constant that does not depend on the dimensions $\{d_k\}$ and $\{p_k\}$. 
 \end{thm}
@@
 Theorem~\ref{thm:main} establishes the statistical convergence for the estimator~\eqref{eq:MLE}.
-Actually, the proof in section~\ref{sec:appedix} shows that the statistically optimal rate holds, not only for the MLE~\eqref{eq:MLE}, but also any local estimators $\hat \tB$ in the level set $\{\hat \tB \in \tP \colon \tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true}) \}$ satisfies the upper bound~\ref{thm:main}\fixme{Miaoyan}{Wrong statement based on current proof}.
+Actually, the proof in section~\ref{sec:appedix} shows that the statistically optimal rate holds, not only for the MLE~\eqref{eq:MLE},  but also any local estimators $\hat \tB$ in the level set $\{\hat \tB \in \tP \colon \tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true}) \}$ satisfies the upper bound~\ref{thm:main}.
 In Section~\ref{subsec:alg}, we implement empirical studies to verify the algorithmic stability to find a local optimum.
 
-To gain further insight on the bound~\eqref{eq:bound}, we consider a special case when tensor dimensions are equal at each of the modes, i.e., $d_k=d$, $p_k=\gamma d$, $\gamma\in [0,1)$ for all $k\in[K]$, and the feature matrices $\mX_k$ are Gaussian design with i.i.d.\ $N(0,1)$ entries.
-To put the context in the framework of Theorem~\ref{thm:main}, we rescale the feature matrices into $\check \mX_k={1\over \sqrt{d}}\mX_k$ so that the singular values of $\check \mX_k$ are bounded by $1\pm \sqrt{\gamma}$.
-The result in~\eqref{eq:bound} implies that the estimated coefficient has a convergence rate $\tO({p\over d^K})$ in the scale of the original feature $\{\mX_k\}$.
+To gain further insight on the bound~\eqref{eq:bound}, we consider a special case when tensor dimensions are equal at each of the modes, i.e., $d_k=d$, $p_k=\gamma d$, $\gamma\in [0,1)$ for all $k\in[K]$, and the covariates $\mX_k$ are Gaussian design matrices with i.i.d.\ $N(0,1)$ entries.
+To put the context in the framework of Theorem~\ref{thm:main}, we rescale the covariates into $\check \mX_k={1\over \sqrt{d}}\mX_k$ so that the singular values of $\check \mX_k$ are bounded by $1\pm \sqrt{\gamma}$.
+The result in~\eqref{eq:bound} implies that the estimated coefficient has a convergence rate $\tO({p\over d^K})$ in the scale of the original covariates $\{\mX_k\}$.
 Therefore, our estimation is consistent as the dimension grows, and the convergence becomes especially favorably as the order of tensor data increases.
@@
 As immediate applications, we obtain the convergence rate for the three examples mentioned in Section~\ref{sec:model}.
-Without loss of generality, we assume that the singular values of the $d_k$-by-$p_k$ feature matrix $\mX_k$ are bounded by $\sqrt{d_k}$.
+Without loss of generality, we assume that the singular values of the $d_k$-by-$p_k$ covariate matrix $\mX_k$ are bounded by $\sqrt{d_k}$.
 
-\begin{example}[Spatio-temporal growth model] The estimated type-by-time-by-space coefficient tensor converges at the rate $\tO\left({p+r+s\over dmn}\right)$ where $p\leq d$, $r\leq m$ and $s\leq n$.
+\begin{cor}[Spatio-temporal growth model] The estimated type-by-time-by-space coefficient tensor converges at the rate $\tO\left({p+r+s\over dmn}\right)$ where $p\leq d$, $r\leq m$ and $s\leq n$.
 The estimation achieves consistency as long as the dimension grows in either of the three modes.
-\end{example}
+\end{cor}
 
-\begin{example} [Network population model] The estimated node-by-node-by-feature tensor converges at the rate $\tO\left({2d+p\over d^2n}\right)$ where $p\leq n$.
+\begin{cor} [Network population model] The estimated node-by-node-by-covariate tensor converges at the rate $\tO\left({2d+p\over d^2n}\right)$ where $p\leq n$.
 The estimation achieves consistency as the number of individuals or the number of nodes grows.
-\end{example}
+\end{cor}
 
-\begin{example} [Dyadic data with node attributes] The estimated feature-by-feature matrix converges at the rate $\tO\left({p\over d^2}\right)$ where $p\leq d$.
+\begin{cor} [Dyadic data with node attributes] The estimated covariate-by-covariate matrix converges at the rate $\tO\left({p\over d^2}\right)$ where $p\leq d$.
 Again, our estimation achieves consistency as the number of nodes grows.
-\end{example}
+\end{cor}
 
@@
 \[
-\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY})\leq \frac{ C_{4}r_{\textup{total}}}{r_{\max}} {\sum_k p_k\over \prod_k d_k},
+\text{KL}(\mathbb{P}_{\tY_{\text{true}}},\ \mathbb{P}_{\hat \tY})\leq \frac{ C_{4}\prod_kr_k}{\max_k r_k} \sum_k p_k,
 \]
-where $C_4=C_4(\mr,\alpha, K)>0$ is a constant that independent of dimensions $\{d_k\}$ and $\{p_k\}$.
+where $C_4=C_4(\mr,\alpha, K)>0$ is a constant that do not depend on the dimensions $\{d_k\}$ and $\{p_k\}$.
 \end{thm}
 
-\section{Numerical experiments}\label{sec:simulation}
-{\color{red} Simulation/figures needs to be re-run once more, since code has been changed. }
+\section{Numerical implementation}
+\subsection{Alternating optimization} \label{subsec:alg}
+In this section, we introduce an efficient algorithm to solve~\eqref{eq:MLE}.
+The objective function $\tL_{\tY}(\tB)$ is concave in $\tB$ when the link $f$ is the canonical link function.
+We utilize a Tucker factor representation of the coefficient tensor $\tB$ and turn the optimization into a block-wise convex problem.
 
+Specifically, write the rank-$\mr$ decomposition of coefficient tensor $\tB$ as
+\begin{equation}\label{eq:tucker}
+\tB=\tC\times \{\mM_1,\ldots,\mM_K\},
+\end{equation}
+where $\tC\in\mathbb{R}^{r_1\times\cdots\times r_K}$ is a full-rank core tensor, $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices with orthogonal columns.
+Estimating $\tB$ amounts to finding both the core tensor $\tC$ and the factor matrices $\mM_k$'s.
+The optimization~\eqref{eq:MLE} can be written as $(\hat \tC, \{\hat \mM_k\})=\arg\max \tL_{\tY}(\tC, \mM_1,\ldots,\mM_K)$, where
+\begin{align}
+\tL_{\tY}(\tC, \mM_1,\ldots,\mM_K )&=\langle \tY, \Theta\rangle -\sum_{i_1,\ldots,i_K}b(\theta_{i_1,\ldots,i_K}),\\
+\text{with } \Theta&=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
+\end{align}
+The decision variables in the above objective function consist of $K+1$ blocks of variables, one for the core tensor $\tC$ and $K$ for the factor matrices $\mM_k$'s.
+We notice that, if any $K$ out of the $K+1$ blocks of variables are known, then the optimization with respect to the last block of variables reduced to a simple GLM.
+This observation suggests that we can iteratively update one block at a time while keeping others fixed.
+After each iteration, we rescale the core tensor $\tC^{(t+1)}$ subject to the maximum norm constraint.
+This post-processing in principle may not guarantee the monotonic increase of the objective, but we found that in our experiment this simple step appears robust for obtaining a desirable solution.
+The full algorithm is described in Algorithm~\ref{alg:B}.
+
+\begin{algorithm}[t]
+\caption{Generalized tensor response regression with covariates on multiple modes}\label{alg:B}
+\begin{algorithmic}[1]
+\INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, covariate matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,\ldots,K$, target Tucker rank $\mr=(r_1,\ldots,r_K)$, link function $f$, maximum norm bound $\alpha$
+\OUTPUT Low-rank estimation for the coefficient tensor $\tB\in\mathbb{R}^{p_1\times \cdots\times p_K}$. 
+\State Calculate $\check \tB=\tY\times_1 \left[(\mX_1^T\mX_1)^{-1}\mX^T_1\right] \times_2\cdots\times_K\left[(\mX_K^T\mX_K)^{-1}\mX^T_K\right] $.
+\State Initialize the iteration index $t=0$.
+Initialize the core tensor $\tC^{(0)}$ and factor matrices $\mM^{(0)}_k\in\mathbb{R}^{p_k\times r_k}$ via rank-$\mr$ Tucker approximation of $\check\tB$, in the least-square sense.
+  \While {the relative increase in objective function $\tL_\tY(\tB)$ is less than the tolerance}
+\State Update iteration index $t \leftarrow t+1$.
+\For { $k=1$ to $K$}
+\State Obtain the factor matrix $\tilde \mM_k^{(t+1)}\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$. 
+\State Perform QR factorization on $\tilde \mM_k^{(t+1)}=\mQ\mR$, where $\mQ\in\mathbb{R}^{p_k\times r_k}$ has orthogonal columns. 
+\State Update $\mM_k^{(t+1)}\leftarrow \mQ$ and core tensor $\tC^{(t+1)}\leftarrow \tC^{(t+1)}\times_k \mR$.
+\EndFor
+\State Obtain the core tensor $\tC^{(t+1)}\in\mathbb{R}^{r_1\times\cdots\times r_K}$ by solving a GLM with $\text{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM^{(t)}_k]$ as covariates, and $f$ as link function.
+Here $\otimes$ denotes the kronecker product of matrices.
+\State Rescale the core tensor subject to the maximum norm constraint. 
+\State Update $\tB^{(t+1)}\leftarrow \tC^{(t+1)}\times_1\mM_1^{(t+1)}\times_2\cdots\times_K\mM_K^{(t+1)}$.
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
+
+Note that the feasible set $\tP$ is a non-convex set.
+Therefore, the optimization~\eqref{eq:MLE} is a non-convex problem, and the Algorithm~\ref{alg:B} usually not theoretically possesses the global optimality.
+However, as mentioned in Section~\ref{subsec:statprob}, the desired convergence rate~\eqref{eq:bound} holds for the valid estimators satisfying $\tL_{\tY} (\hat \tB) \geq \tL_{\tY} (\tB_{true})$, which indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective of local optimums are large enough.
+Fortunately, we find the  Algorithm~\ref{alg:B}  often gives a satisfactory convergence point $\hat \tB$, when we initialize the parameters via continuous-valued Tucker decomposition.
+Figure~\ref{fig:loglike} shows the trajectory of the objective function for order-3 tensors under the balanced setting, where $\alpha = 10, d_k = d, p_k = 0.4d, r_k = r$ for $ k = 1,2,3, p \in \{25,30\}$  and $r \in \{3, 6\}$.
+We consider the inputs with Gaussian, Bernoulli, and Poisson entries.
+Under all combinations of the dimension $d$, rank $r$, and type of the entries, Algorithm~\ref{alg:B} converges quickly in a few iterations, and the objective value at convergent points are close to or larger than the value at true parameters.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=15cm]{loglike}
+\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c)Poisson models.
+The dashed line represents the objective value at true parameter $\tL_{\tY}(\tB_{true})$.
+A micro-iteration refers to the update for one of the $K$ blocks.
+Four micro-iterations consist of a complete ``while'' iteration in Algorithm~\ref{alg:B}.}\label{fig:loglike}
+\end{figure}
+
+\subsection{Rank selection}\label{sec:tuning}
+Algorithm~\ref{alg:B} takes the rank $\mr$ as an input.
+Estimating an appropriate rank given the data is of practical importance.
+We propose to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
+\begin{align}\label{eq:BIC}
+\hat \mr&=\argmin_{\mr=(r_1,\ldots,r_K)} \text{BIC}(\mr)\\
+&=\argmin_{\mr=(r_1,\ldots,r_K)}\left[-2\tL_{\tY}(\hat \tB)+p_e(\mr)\log \left(\prod_k d_k\right) \right],
+\end{align}
+where $p_e(\mr)\stackrel{\text{def}}{=}\sum_k (p_k-r_k-1)r_k+\prod_k r_k$ is the effective number of parameters in the model.
+We choose $\hat \mr$ that minimizes $\text{BIC}(\mr)$ via grid search.
+Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model.
+We test its empirical performance in Section~\ref{sec:simulation}.
+
+\section{Simulation}\label{sec:simulation}
 We evaluate the empirical performance of our generalized tensor regression through simulations.
 We consider order-3 tensors with a range of distribution types.
-The coefficient tensor $\tB$ is generated using the factorization form~\eqref{eq:factor} where both the core and factor matrices are drawn i.i.d.\ from Uniform[-1,1].
-The linear predictor is then simulated from $\tU=\tB\times\{\mX_1,\mX_2,\mX_3\}$, where $\mX_k$ is either an identity matrix (i.e.\ no feature available) or Gaussian random matrix with i.i.d.\ entries from $N(0,\sigma_k^2)$.
+The coefficient tensor $\tB$ is generated using the factorization form~\eqref{eq:tucker} where both the core and factor matrices are drawn i.i.d.\ from Uniform[-1,1].
+The linear predictor is then simulated from $\tU=\tB\times\{\mX_1,\mX_2,\mX_3\}$, where $\mX_k$ is either an identity matrix (i.e.\ no covariate available) or Gaussian random matrix with i.i.d.\ entries from $N(0,\sigma_k^2)$.
 We set $\sigma_k=d_k^{-1/2}$ to ensure the singular values of $\mX_k$ are bounded as $d_k$ increases.
@@
 
-\begin{enumerate}
+\begin{enumerate}[itemsep=0pt,topsep=0pt]
 \item[(a)] (Gaussian). Continuous data $y_{ijk}\sim N\left(\alpha u_{ijk}, 1\right)$.
@@
 
-\begin{table}[tb]
+\begin{table}[htb]
 \centering
@@
 
-The second experiment evaluates the accuracy when features are available on all modes.
+The second experiment evaluates the accuracy when covariates are available on all modes.
 We set $\alpha=10, d_k=d, p_k=0.4d_k, r_k=r\in\{2,4,6\}$ and increase $d$ from 25 to 50.
@@
 Each network measures the connections between $d_1=d_2=20$ brain nodes.
-We simulate $p=5$ features for the each of the 50 individuals.
-These features may represent, for example, age, gender, cognitive score, etc.
+We simulate $p=5$ covariates for the each of the 50 individuals.
+These covariates may represent, for example, age, gender, cognitive score, etc.
 Recent study~\citep{robinson2015dynamic} has suggested that brain connectivity networks often exhibit community structure represented as a collection of subnetworks, and each subnetwork is comprised of a set of spatially distributed brain nodes.
@@
 Specifically, we partition the nodes into $r$ blocks by assigning each node to a block with uniform probability.
-Edges within a same block are assumed to share the same feature effects, where the effects are drawn i.i.d.\ from $N(0,1)$.
+Edges within a same block are assumed to share the same covariate effects, where the effects are drawn i.i.d.\ from $N(0,1)$.
 We then apply our tensor regression model to the network data using the BIC-selected rank.
@@
 
-\begin{figure}[tb]
+\begin{figure}[htb]
 \centering
@@
 Figure~\ref{fig:glm} compares the MSE of our method with a classical GLM approach.
-A classical GLM is to regress the dyadic edges, one at a time, on the features, and this model is repeatedly fitted for each edge.
+A classical GLM is to regress the dyadic edges, one at a time, on the covariates, and this model is repeatedly fitted for each edge.
 This repeated approach, however, does not account for the correlation among the edges, and may suffer from overfitting.
@@
 We compare our generalized tensor regression ({\bf GTR}) with three other supervised tensor methods:
-\begin{itemize}
-\item  Higher-order low-rank regression ({\bf HOLRR},~\citep{rabusseau2016low}) is a least-square based tensor regression that allows features on a single mode. 
-\item Higher-order partial least square ({\bf HOPLS},~\citep{zhao2012higher}) is a dimension-reduction method that jointly models a tensor response and a tensor feature. 
-\item Subsampled tensor projected gradient ({\bf TPG},~\citep{yu2016learning}) tackles the same question as {\bf HOLRR} but instead uses a different algorithm to solve the problem. 
+\begin{itemize}[itemsep=-1pt,topsep=-1pt,leftmargin=*,partopsep=-1pt]
+\item  Higher-order low-rank regression ({\bf HOLRR},~\citet{rabusseau2016low}) is a least-square based tensor regression that allows covariates on a single mode. 
+\item Higher-order partial least square ({\bf HOPLS},~\citet{zhao2012higher}) is a dimension-reduction method that jointly models a tensor response and a tensor covariate. 
+\item Subsampled tensor projected gradient ({\bf TPG},~\citet{yu2016learning}) tackles the same question as {\bf HOLRR} but instead uses a different algorithm to solve the problem. 
 \end{itemize}
-These three methods are the closest algorithms to ours, in that they relate a tensor response to features using a low-rank structure.
+These three methods are the closest algorithms to ours, in that they relate a tensor response to covariates using a low-rank structure.
 All the three methods allow only Gaussian data, whereas ours is applicable to any exponential family distribution including Gaussian, Bernoulli, Multinomial, etc.
@@
 
-The comparison was assessed from three aspects: (a) benefit of incorporating features from multiple modes; (b) prediction error with respect to sample size; (c) sensitivity of accuracy with respect to model complexity.
+The comparison was assessed from three aspects: (a) benefit of incorporating covariates from multiple modes; (b) prediction error with respect to sample size; (c) sensitivity of accuracy with respect to model complexity.
 We use similar simulation setups as in our experiment II, but consider combinations of rank ($\mr=(3,3,3)$ vs.\ $(4,5,6)$), noise ($\sigma = 1/2$ vs.
-$1/4$), and dimension ($d$ ranging from 20 to 100 for modes with features, $d = 20$ for modes without features).
+$1/4$), and dimension ($d$ ranging from 20 to 100 for modes with covariates, $d = 20$ for modes without covariates).
 
@@
 As the number of informative modes (i.e.
-modes with available features) increases, the {\bf GTR} exhibits a reduction in error whereas others have increased errors.
-This showcases the benefit toward prediction via incorporation of interactive features.
+modes with available covariates) increases, the {\bf GTR} exhibits a reduction in error whereas others have increased errors.
+This showcases the benefit toward prediction via incorporation of multiple covariates.
 Note that our method {\bf GTR} is most comparable to {\bf HOLRR} when there is only a single informative mode.
@@
 
-\caption{Comparison of MSPE versus the number of modes with features.
+\caption{Comparison of MSPE versus the number of modes with covariates.
 We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and noise $\sigma=1/2$ (high), $\sigma=1/4$ (low).}~\label{fig:compare}
@@
 
-\section{Data analyses}\label{sec:data}
-We apply our tensor regression model to two datasets.
-The first application concerns the brain network modeling in response to individual attributes (i.e.\ feature on one mode), and the second application focuses on multi-relational network analysis with dyadic attributes (i.e.\ features on two modes).
-
-\subsection{Analysis of human brain connection data}
-The Human Connectome Project (HCP) aims to build a ``network map'' that characterizes the anatomical and functional connectivity within healthy human brains~\citep{HCP}.
+\section{Data analysis}
+We apply our tensor regression model to two real datasets.
+The first application concerns the brain network modeling in response to individual attributes (i.e.\ covariate on one mode), and the second application focuses on multi-relational network analysis with dyadic attributes (i.e.\ covariates on two modes).
+\subsection{Human Connectome Project (HCP)}
+The Human connectome project (HCP~\citep{HCP}) aims to build a ``network map'' that characterizes the anatomical and functional connectivity within healthy human brains.
 We take a subset of HCP data that consists of 136 brain structural networks, one for each individual.
 Each brain network is represented as a 68-by-68 binary matrix, where the entries encode the presence or absence of fiber connections between the 68 brain regions.
-We consider four individual features: gender (65 females vs.\ 71 males), age 22-25 ($n=35$), age 26-30 ($n=58$), and age 31+ ($n=43$).
-The goal is to identify the connection edges that are affected by the individual feature.
+We consider four individual-covariates: gender (65 females vs.\ 71 males), age 22-25 ($n=35$), age 26-30 ($n=58$), and age 31+ ($n=43$).
+The goal is to identify the connection edges that are affected by the individual covariates.
 A key challenge in brain network is that the edges are correlated; for example, two edges may stem out from a same brain region, and it is of importance to take into account the within-dyad dependence.
 
-\begin{figure}[H]
+\begin{figure}[t]
 \centering
-\includegraphics[width=13cm]{HCP.pdf}
+\includegraphics[width=12cm]{HCP.pdf}
 \caption{Top edges with large effects.
@@
 We fit the tensor regression model to the HCP data.
-The response is a binary tensor $\tY\in\{0,1\}^{68\times 68\times 136}$ and the features are of dimension 4 along the 3$^{\text{rd}}$ mode.
+The response is a binary tensor $\tY\in\{0,1\}^{68\times 68\times 136}$ and the covariates are of dimension 4 along the 3$^{\text{rd}}$ mode.
 The BIC selection suggests a rank $\mr=(10,10,4)$ with log-likelihood $\tL_{\tY}=-174654.7$.
@@
 
-Figure~\ref{fig:s1} compares the estimated coefficients from our method (tensor regression) with those from classical GLM approach.
-A classical GLM is to regress the brain edges, one at a time, on the individual-level features, and this logistic model is repeatedly fitted for every edge $\in [68]\times [68]$.
-As we can see in the figure, our tensor regression shrinkages the coefficients towards center, thereby enforcing the sharing between coefficient entries.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=16cm]{compare.pdf}
-\caption{Comparison of coefficient estimation in the HCP data.}\label{fig:s1}
-\end{figure}
-
-\subsection{Analysis of political relational data}
+\subsection{Nations data}
 
@@
 The 56 relations span the fields of politics, economics, military, religion, and so on.
-In addition, country-level attributes are also available, and we focus on the following six features: \emph{constitutional, catholics, lawngos, politicalleadership, geographyx}, and \emph{medicinengo}.
+In addition, country-level attributes are also available, and we focus on the following six covariates: \emph{constitutional, catholics, lawngos, politicalleadership, geographyx}, and \emph{medicinengo}.
 The goal is to identify the variation in connections due to country-level attributes and interactions thereof.
-One of the key features is that the 56 relations are correlated, and we would like to take that into account in assessing the feature effects.
+One of the key features is that the 56 relations are correlated, and we would like to take that into account in assessing the covariate effects.
 
 We applied our tensor regression model to the \emph{Nations} data.
-The multi-relational network $\tY\in\{0,1\}^{14\times 14\times 56}$ was treated as the response tensor, and the country attributes $\mM\in\mathbb{R}^{14\times 6}$ were treated as features on both the 1st and 2nd modes.
+The multi-relational network $\tY\in\{0,1\}^{14\times 14\times 56}$ was treated as the response tensor, and the country attributes $\mM\in\mathbb{R}^{14\times 6}$ were treated as covariates on both the 1st and 2nd modes.
 The BIC criterion suggests a rank $\mr=(4,4,4)$ for the coefficient tensor $\tB\in\mathbb{R}^{6\times 6 \times 56}$.
@@
 
-\begin{figure}[H]
+\begin{figure}[t]
 \centering
-\includegraphics[width=14cm]{coef.pdf}
+\includegraphics[width=12cm]{coef.pdf}
 \caption{Effect estimation in the \emph{Nations} data.
@@
 }\label{fig:est}
+
 \end{figure}
 
-\begin{table}[H]
-\begin{tabular}{c|c|l}
-\hline
-\multirow{5}{*}{Cluster I}  & Government &officialvisits, intergovorgs,  militaryactions,  \\
-&&negativebehavior, boycottembargo, aidenemy, negativecomm, \\
-&&protestsunoffialacts,    nonviolentbehavior, emigrants,  \\
-&&timesincewar, commonbloc2, rintergovorgs3, relintergovorgs\\
-&&violentactions, duration,accusation,relexports\\
-\hline
-\multirow{6}{*}{Cluster II}& Economics &economicaid, booktranslations, tourism,  \\
-&& conferences, severdiplomatic, expeldiplomats, attackembassy, \\
-&& reltourism, tourism3, relemigrants, emigrants3, students, \\
-&&exports, exports3, lostterritory, dependent, militaryalliance  \\
-&&relbooktranslations, releconomicaid, unweightedunvote, \\
-&& warning, relstudents,\\
-\hline
-\multirow{3}{*}{Cluster III} & treaties &treaties, reltreaties, exportbooks, relexportbooks,  \\
-&&relngo, ngoorgs3, embassy, reldiplomacy, timesinceally, \\
-&&independence, commonbloc1,weightedunvote, ngo,\\
-\hline
-Cluster IV & Politics &commonbloc0, blockpositionindex\\
-\hline
-\end{tabular}
-\caption{$K$-means clustering of relations based on factor matrix in the coefficient tensor.}\label{tab:s1}
-\end{table}
-
 To investigate the effects of dyadic attributes towards connections, we depicted the estimated coefficients $\hat \tB=\entry{\hat b_{ijk}}$ for several relation types (Figure~\ref{fig:est}).
-Note that entries $\hat b_{ijk}$ can be interpreted as the contribution, at the logit scale, of feature pair $(i,j)$ ($i$th feature for the ``sender'' country and $j$th feature for the ``receiver'' country) towards the connection of relation $k$.
+Note that entries $\hat b_{ijk}$ can be interpreted as the contribution, at the logit scale, of covariate pair $(i,j)$ ($i$th covariate for the ``sender'' country and $j$th covariate for the ``receiver'' country) towards the connection of relation $k$.
 Several interesting findings emerge from the observation.
-We found that relations belonging to a same cluster tend to have similar feature effects.
-For example, the relations \emph{warnings} and \emph{ecnomicaid} are classified into Cluster II, and both exhibit similar feature pattern (Figure~\ref{fig:est}a-b).
+We found that relations belonging to a same cluster tend to have similar covariate effects.
+For example, the relations \emph{warnings} and \emph{ecnomicaid} are classified into Cluster II, and both exhibit similar covariate pattern (Figure~\ref{fig:est}a-b).
 Moreover, the majority of the diagonal entries $\hat \tB(i,i,k)$ positively contribute to the connection.
@@
 
-\section{Conclusion}\label{sec:con}
-\fixme{Miaoyan}{No more than 35 pages. Reserve at least 1.5 pages for Conclusion}
-We have developed a generalized tensor regression with features on multiple modes.
+\section{Conclusion}
+
+ We have developed a generalized tensor regression with covariates on multiple modes.
 A fundamental feature of tensor-valued data is the statistical interdependence among entries.
@@
 
- \fixme{Miaoyan}{Add two more paragraphs: 
- %Our method leverages on the alternating updating algorithm for the optimization.
-Although a non-convex optimization procedure such as Algorithm~\ref{alg:binary} has no guarantee on global optimality, our numerical experiments have suggested that, upon random initializations, the convergence point $\hat\Theta$ is often satisfactory, in that the corresponding objective value $\tL_{\tY}(\hat\Theta)$ is close to the objective value $\tL_{\tY}(\trueT)$.
-When starting from random initializations, there could be multiple estimates, whose objective values are all greater than $\tL_{\tY}(\trueT)$.
-In theory, any of those choices perform equally well in estimating $\trueT$.
-In this sense, local optimality is not necessarily a severe concern in our context.
-On the other hand, characterizing global optimality for non-convex optimization problem of this type is itself of great interest.
-There has been recent progress investigating the landscape of non-convex optimization involving tensors~\citep{anandkumar2014tensor, richard2014statistical, ge2017optimization}.
-The problem is challenging, as the geometry can  depend on multiple factors including the tensor rank, dimension, and factorization form.
-In some special cases such as rank-1 or orthogonally decomposable tensors, one may further obtain the required asymptotical number of initializations, however, at the cost of more stringent assumptions on the target tensor~\citep{anandkumar2014tensor, richard2014statistical}.
-We leave the pursuit of optimization landscape as future research.
-%Another an..
-is the (conditional) independence assumption on the tensor entries.
-Instead of glm, we may introduce a more general tensor-variate distribution on the $\tY$.
-For the case of normal distribution.
-For random effects.....
-2.
-Assumption on (conditional) independence assumption on the tensor entries.
-Mixed effect model, tensor-variate normal distribution..Current approach is to use fixed factor; or a combination of fixed factor + random effects.
-3. Inference. ... uncertainly quantification... bootstrap ....  Testing for Interaction + main effects. 
-}
-
-\section*{Acknowledgements}
-This research was supported by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
-%The authors would like to thank Chen Zhang from Wisconsin State Lab of Hygiene for help with figure 4.
-
-\fixme{miaoyan}{references need update..}
-
-\bibliography{tensor_wang}
-\bibliographystyle{apalike}
-
-\newpage
-\begin{center}
-{\large\bf SUPPLEMENTARY MATERIAL}
-\end{center}
-
 \appendix
-
 \section{Proofs}\label{sec:appedix}
@@
 
-\section{Code}
-The source code and data used in the paper are available at \url{https://CRAN.R-project.org/package=tensorregress.}
+\section{Time complexity}
+The computational complexity of our tensor regression model is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the response tensor.
+More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor.
+The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$-covariate matrix.
+Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$.
+
+\section{Additional results for real data analysis}
+Here we provide additional results for the real data analysis.
+
+\subsection{HCP data analysis}
+
+Figure~\ref{fig:s1} compares the estimated coefficients from our method (tensor regression) with those from classical GLM approach.
+A classical GLM is to regress the brain edges, one at a time, on the individual-level covariates, and this logistic model is repeatedly fitted for every edge $\in [68]\times [68]$.
+As we can see in the figure, our tensor regression shrinkages the coefficients towards center, thereby enforcing the sharing between coefficient entries.
+
+\begin{figure}[t]
+\includegraphics[width=17cm]{compare.pdf}
+\caption{Comparison of coefficient estimation in the HCP data.}\label{fig:s1}
+\end{figure}
+
+\subsection{Nations data analysis}
+Table~\ref{tab:s1} summarizes the $K$-means clustering of the 56 relations based on the $3^{\text{rd}}$ mode factor $\mM_3\in\mathbb{R}^{56\times 4}$ in the tensor regression model.
+
+\begin{table}[H]
+\begin{tabular}{c|l}
+\hline
+\multirow{4}{*}{Cluster I}  &officialvisits, intergovorgs,  militaryactions, violentactions, duration, \\
+&negativebehavior, boycottembargo, aidenemy, negativecomm, accusation, \\
+& protestsunoffialacts,    nonviolentbehavior, emigrants, relexports,   \\
+& timesincewar, commonbloc2, rintergovorgs3, relintergovorgs\\
+\hline
+\multirow{4}{*}{Cluster II}& economicaid, booktranslations, tourism, relbooktranslations, releconomicaid, \\
+& conferences, severdiplomatic, expeldiplomats, attackembassy, unweightedunvote,\\
+& reltourism, tourism3, relemigrants, emigrants3, students, relstudents, \\
+&exports, exports3, lostterritory, dependent, militaryalliance, warning  \\
+\hline
+\multirow{2}{*}{Cluster III} & treaties, reltreaties, exportbooks, relexportbooks, weightedunvote, ngo, \\
+&relngo, ngoorgs3, embassy, reldiplomacy, timesinceally, independence, commonbloc1\\
+\hline
+Cluster IV & commonbloc0, blockpositionindex\\
+\end{tabular}
+\caption{$K$-means clustering of relations based on factor matrix in the coefficient tensor.}\label{tab:s1}
+\end{table}
+
+\section*{Acknowledgements}
+This research was supported by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
+The authors would like to thank Chen Zhang from Wisconsin State Lab of Hygiene for help with figure 4.
+
+%\bibliographystyle{unsrt}
+\bibliography{tensor_wang}
+\bibliographystyle{apalike}
 

### 2020-09-14 code/sbm_gao/get_sbm_data.R

@@ -1,45 +1,0 @@
-## get data function for SBM in Gao
-
-source("/Users/March/Desktop/myresearch/code/sbm_gao/bricks_sbm.R")
-
-get_sbm_data = function(n,k,a,b){
-  # n is the matrix demsion
-  # k is the # of clusters
-  # a is the signal level of diagonal elements 
-  # b is the signal level of off-diagonal elements
-  
-  # n = 50; k = 4
-  # a = 25; b = 25
-  
-  prob = matrix(runif(k^2),nrow = k)
-  # symmetric
-  prob[lower.tri(prob)] = t(prob)[lower.tri(prob)]  
-  # signal contrain
-  # min Bii > a/n; max Bij < b/n
-  diag(prob)[which(diag(prob) <= a/n)] = a/n
-  index_mat = prob >= b/n
-  diag(index_mat) = FALSE
-  prob[index_mat]  = b/n 
-  
-  if (k == n) truthC = 1:n
-  else if (k == 1) truthC = rep(1,n)
-  else {truthC = sort(ReNumber(sample(1:k,n,replace=TRUE)) )}
-  
-  x = matrix(1,nrow = n,ncol = n)
-  truthX =  matrix(0,nrow = n,ncol = n)
-  for(i in 1:max(truthC)){
-    for(j in 1:max(truthC)){
-        x[truthC==i, truthC==j] = rbinom(sum(x[truthC==i, truthC==j]),1,prob[i,j])
-        truthX[truthC==i, truthC==j] =  prob[i,j]
-    }
-  }
-  
-  result = list("x"=x,"truthX"=truthX,"truthC"=truthC,"prob"=prob)
-  # x is the observation
-  # truthX is true connection
-  # truthC is true partition
-  # prob is true adjacency
-  return(result)
-}
-
-

### 2020-09-14 code/sbm_gao/greedy_initial_clustering.R

@@ -1,74 +1,0 @@
-## initialization in sbm
-library(pracma)
-
-source("/Users/March/Desktop/myresearch/code/sbm_gao/bricks_sbm.R")
-
-greedy_initial_clustering = function(A,k,trim = TRUE, tau = Inf,mu){
-  # USC
-  #A is adjacency
-  #k is # of clusters
-  #tau is trimming para
-  #r is cluster radius
-  
-  n = dim(A)[1]
-  r = mu*sqrt(k/n)
-  
-  if(trim == TRUE){
-    dvec = colSums(t(A))
-    index_trim = dvec >= tau
-    A[index_trim,] = 0
-    A[,index_trim] = 0
-  }
-  
-  U = eigen(A)$vectors[,1:k]
-  
-  # U = matrix(1:20,ncol = 2)
-  # n = 10
-  # k = 3
-  # r = 5*sqrt(2/10)
-  
-  S = 1:n
-  
-  est_C = rep(0,n)
-  # generate norm matrix
-  norm_mat = gene_norm_mat(U)
-  norm_mat0 = as.matrix(norm_mat)
-  
-  for (i in 1:k) {
-    
-    index_norm_mat = norm_mat0 < r
-    if(class(index_norm_mat) == "numeric") {rsum = norm_mat0
-    } else {rsum = colSums(index_norm_mat)}
-    ti = which(rsum == max(rsum))[1]
-    Ci = which((index_norm_mat[,ti]) == TRUE)
-    est_C[S[Ci]]= i
-    
-    S = S[-Ci]
-    
-    if(i < k & length(S) == 0){
-      print("please take smaller mu")
-      return(est_C)
-    }
-    
-    norm_mat0 = norm_mat0[-Ci,]
-    if(class(norm_mat0) == "numeric"){norm_mat0 = matrix(norm_mat0,nrow = 1)}
-    norm_mat0 = (norm_mat0)[,-Ci]
-  }
-  
-  est_C
-  
-  if(length(S) != 0){
-    for (i in 1:length(S)) {
-      usum = rep(0,k)
-      for (j in 1:k) {
-        index_c = which(est_C == j)
-        usum[j] = sum(norm_mat[index_c,S[i]])/length(index_c)
-      }
-      est_C[S[i]] = which(usum == min(usum))
-    }
-  }
-  est_C
-  return(est_C)
-}
-
-

### 2020-09-14 code/sbm_gao/sbm_main.R

@@ -1,131 +1,0 @@
-#### Algorithm for SBM in Gao ####
-
-library(plot.matrix)
-
-source("/Users/March/Desktop/myresearch/code/sbm_gao/bricks_sbm.R")
-source("/Users/March/Desktop/myresearch/code/sbm_gao/get_sbm_data.R")
-source("/Users/March/Desktop/myresearch/code/sbm_gao/greedy_initial_clustering.R")
-source("/Users/March/Desktop/myresearch/code/sbm_gao/refinement_sbm.R")
-
-
-## generate data
-set.seed(1001)
-test_data = get_sbm_data(50,3,20,10)
-plot(test_data$x)
-plot(test_data$truthX)
-plot(test_data$prob)
-test_data$truthC
-
-
-## initialization clustering \sigma^0
-test_ini = greedy_initial_clustering(test_data$x,3,mu = 0.5)
-test_ini
-
-svd(test_data$truthX)$d
-# when lambda k of truth x is larger, the test ini is more accurate.
-
-## refinement
-test_refine = refinement_sbm(test_data$x,3,mu = 0.5)
-test_refine$sigma_est
-test_data$truthC
-
-est_x = matrix(0,nrow = 50, ncol = 50)
-for (i in 1:3) {
-  for (j in 1:3) {
-    est_x[test_refine$sigma_est == i,test_refine$sigma_est == j] = test_refine$B_est[i,j]
-  }
-}
-plot(est_x)
-plot(test_refine$B_est)
-
-# plot x
-axis = 1:50
-data1 = expand.grid(X=axis, Y=axis)
-gridx = test_data$x
-dim(gridx) = c(2500,1)
-data1$gridx = gridx
-
-gridtruth = test_data$truthX
-dim(gridtruth) = c(2500,1)
-data1$gridtruth = gridtruth
-
-gridjudge = est_x
-dim(gridjudge) = c(2500,1)
-data1$gridjudge = gridjudge
-
-sbmx = ggplot(data1, aes(X, Y, fill= gridx)) + 
-  geom_tile() +
-  guides( fill = guide_legend(title = "x")) 
-
-sbmt = ggplot(data1, aes(X, Y, fill= gridtruth)) + 
-  geom_tile() +
[…]

### 2020-09-14 code_sbm/sbm_gao/bricks_sbm.R

@@ -1,0 +1,90 @@
+### bricks for sbm
+ReNumber = function (Cs) 
+{
+  newCs <- rep(NA, length(Cs))
+  uniq <- sort(unique(Cs))
+  for (i in 1:length(uniq)) {
+    newCs[Cs == uniq[i]] <- i
+  }
+  return(newCs)
+}
+
+
+norm_cal = function(v1,v2){
+  return(Norm(v1-v2))
+}
+
+gene_norm_mat = function(U){
+  # U \in R^n \times k
+  # cal every || Ui. - Uj. ||
+  
+  #U = matrix(1:10,nrow = 5,ncol = 2)
+  n = dim(U)[1]
+  
+  norm_mat = matrix(0,nrow = n,ncol = n)
+  for (i in 1:(n-2)) {
+    norm_mat[(i+1):n,i] = apply(U[(i+1):n,], 1, norm_cal,v2 = U[i,])
+  }
+  norm_mat[n,(n-1)] = norm_cal(U[n,],U[(n-1)])
+  
+  norm_mat[upper.tri(norm_mat)] = t(norm_mat)[upper.tri(norm_mat)]
+  #norm_mat is a symmetric matrix
+  return(norm_mat)
+}
+
+## cal loss
+
+gene_conf_mat_one = function(r,true_assign, est_assign){
+  #assigns are vectors
+  
+  conf_mat = matrix(0,nrow = r,ncol = r)
+  d = length(true_assign)
+  
+  for (i in 1:d) {
+    
+    x1 = true_assign[i] 
+    x2 = est_assign[i] 
+    conf_mat[x1,x2] = conf_mat[x1,x2] + 1
+    
+  }
+  
+  return(conf_mat)
+}
+
+cal_loss_sbm = function(k,true_assign,est_assign){
+  conf_mat = gene_conf_mat_one(k,true_assign,est_assign)
+  #conf_mat =  matrix(c(10,8,11,1,5,4,2,9,5,7,4,9,5,6,8,6),nrow = 4)
+  
+  conf_mat0 = conf_mat
+  S = 1:k
+  for (i in 1:(k-1)) {
+    
+    index_max = which(conf_mat0 == max(conf_mat0),arr.ind = T)[1,]
+    
+    if(index_max[1] != index_max[2]){
+      indr = S[index_max[1]]; indc = S[index_max[2]]
+      #swap colmuns indr and indc
+      est_col = conf_mat[,indc]
[…]

### 2020-09-14 code_sbm/sbm_gao/get_sbm_data.R

@@ -1,0 +1,45 @@
+## get data function for SBM in Gao
+
+source("/Users/March/Desktop/myresearch/code/sbm_gao/bricks_sbm.R")
+
+get_sbm_data = function(n,k,a,b){
+  # n is the matrix demsion
+  # k is the # of clusters
+  # a is the signal level of diagonal elements 
+  # b is the signal level of off-diagonal elements
+  
+  # n = 50; k = 4
+  # a = 25; b = 25
+  
+  prob = matrix(runif(k^2),nrow = k)
+  # symmetric
+  prob[lower.tri(prob)] = t(prob)[lower.tri(prob)]  
+  # signal contrain
+  # min Bii > a/n; max Bij < b/n
+  diag(prob)[which(diag(prob) <= a/n)] = a/n
+  index_mat = prob >= b/n
+  diag(index_mat) = FALSE
+  prob[index_mat]  = b/n 
+  
+  if (k == n) truthC = 1:n
+  else if (k == 1) truthC = rep(1,n)
+  else {truthC = sort(ReNumber(sample(1:k,n,replace=TRUE)) )}
+  
+  x = matrix(1,nrow = n,ncol = n)
+  truthX =  matrix(0,nrow = n,ncol = n)
+  for(i in 1:max(truthC)){
+    for(j in 1:max(truthC)){
+        x[truthC==i, truthC==j] = rbinom(sum(x[truthC==i, truthC==j]),1,prob[i,j])
+        truthX[truthC==i, truthC==j] =  prob[i,j]
+    }
+  }
+  
+  result = list("x"=x,"truthX"=truthX,"truthC"=truthC,"prob"=prob)
+  # x is the observation
+  # truthX is true connection
+  # truthC is true partition
+  # prob is true adjacency
+  return(result)
+}
+
+

### 2020-09-14 code_sbm/sbm_gao/sbm_main.R

@@ -1,0 +1,131 @@
+#### Algorithm for SBM in Gao ####
+
+library(plot.matrix)
+
+source("/Users/March/Desktop/myresearch/code/sbm_gao/bricks_sbm.R")
+source("/Users/March/Desktop/myresearch/code/sbm_gao/get_sbm_data.R")
+source("/Users/March/Desktop/myresearch/code/sbm_gao/greedy_initial_clustering.R")
+source("/Users/March/Desktop/myresearch/code/sbm_gao/refinement_sbm.R")
+
+
+## generate data
+set.seed(1001)
+test_data = get_sbm_data(50,3,20,10)
+plot(test_data$x)
+plot(test_data$truthX)
+plot(test_data$prob)
+test_data$truthC
+
+
+## initialization clustering \sigma^0
+test_ini = greedy_initial_clustering(test_data$x,3,mu = 0.5)
+test_ini
+
+svd(test_data$truthX)$d
+# when lambda k of truth x is larger, the test ini is more accurate.
+
+## refinement
+test_refine = refinement_sbm(test_data$x,3,mu = 0.5)
+test_refine$sigma_est
+test_data$truthC
+
+est_x = matrix(0,nrow = 50, ncol = 50)
+for (i in 1:3) {
+  for (j in 1:3) {
+    est_x[test_refine$sigma_est == i,test_refine$sigma_est == j] = test_refine$B_est[i,j]
+  }
+}
+plot(est_x)
+plot(test_refine$B_est)
+
+# plot x
+axis = 1:50
+data1 = expand.grid(X=axis, Y=axis)
+gridx = test_data$x
+dim(gridx) = c(2500,1)
+data1$gridx = gridx
+
+gridtruth = test_data$truthX
+dim(gridtruth) = c(2500,1)
+data1$gridtruth = gridtruth
+
+gridjudge = est_x
+dim(gridjudge) = c(2500,1)
+data1$gridjudge = gridjudge
+
+sbmx = ggplot(data1, aes(X, Y, fill= gridx)) + 
+  geom_tile() +
+  guides( fill = guide_legend(title = "x")) 
+
+sbmt = ggplot(data1, aes(X, Y, fill= gridtruth)) + 
+  geom_tile() +
[…]

### 2020-09-14 tensor_regress/JASA_submission/code/Figure4.R

@@ -1,0 +1,126 @@
+##Code for figure 4. MSE comparison between tensor method vs. massive GLMs.
+install.packages("tensorregress.tar.gz",repos = NULL, type="source")
+source("simulation.R")
+library(tensorregress)
+library(ggplot2)
+
+############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line #60 to run the code ##########
+
+#################### plot MSE for normal model ###############
+load("presaved/Figure4_normal.RData")
+mean=c(apply(table,1,mean),apply(table_naive,1,mean))
+sd=c(apply(table,1,sd),apply(table_naive,1,sd))
+data=cbind(round(mean,2),sd,c(1,1,1,2,2,2),c(5,10,15,5,10,15))
+data=data.frame(data)
+data[,3]=as.factor(data[,3])
+levels(data[,3])=c("Ours","GLM")
+names(data)=c("MSE","sd","method","rank")
+
+
+pdf("Figure4_normal.pdf",width=6,height=4)
+p=ggplot(data=data, aes(x=as.factor(rank), y=MSE, fill=as.factor(method))) +
+geom_bar(stat="identity", position=position_dodge())+
+geom_errorbar(aes(ymin=MSE-sd, ymax=MSE+sd), width=.2,
+position=position_dodge(.9))+scale_fill_manual(values=c('#069AA0','#C0C0C0'))+labs(x="number of blocks",y="MSE")+coord_cartesian(ylim = c(0, 1.3))+ theme(text = element_text(size = 15))
+p
+dev.off()
+
+
+#################### plot MSE for poisson model ###############
+load("presaved/Figure4_poisson.RData")
+mean=c(apply(table,1,mean),apply(table_naive,1,mean))
+sd=c(apply(table,1,sd),apply(table_naive,1,sd))
+data=cbind(round(mean,2),sd,c(1,1,1,2,2,2),c(5,10,15,5,10,15))
+data=data.frame(data)
+data[,3]=as.factor(data[,3])
+levels(data[,3])=c("Ours","GLM")
+names(data)=c("MSE","sd","method","rank")
+
+pdf("Figure4_poisson.pdf",width=6,height=4)
+p=ggplot(data=data, aes(x=as.factor(rank), y=MSE, fill=as.factor(method))) +
+geom_bar(stat="identity", position=position_dodge())+
[…]

### 2020-09-14 tensor_regress/JASA_submission/code/Figure5.R

@@ -1,0 +1,203 @@
+## Code for Figure 5. Comparison of mean sqaured prediction error between our method and the other three previous methods
+install.packages("tensorregress.tar.gz",repos = NULL, type="source")
+source("simulation.R")
+source("comparison.R")
+library(tensorregress)
+library(ggplot2)
+
+############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line #40 to run the code ##########
+
+######## plot MSPE for three methods under different (signal, rank) setting #######
+
+load("presaved/Figure5.RData")
+final = final_mode
+finalsd = final_sd_mode
+
+library(ggplot2)
+library(patchwork)
+new_color = c("#069AA0","#CCC591","#BCA455","#D6CFC4")
+
+signal_range=c(3,6)
+core_range=rbind(c(3,3,3),c(4,5,6))
+pdf("Figure5.pdf",width=12,height=6)
+
+s=1;r=2;
+data=data.frame(PMSE=c(final[s,r,1,],final[s,r,2,],final[s,r,3,])/signal_range[s],sd=c(finalsd[s,r,1,],finalsd[s,r,2,],finalsd[s,r,3,])/signal_range[s],Method=rep(c('STD (Our method)','HOLRR (Low-rank regression)','HOPLS (Partial least square)','TPG (Projection gradient)'),3),Category=c(rep(1,4),rep(2,4),rep(3,4)))
+data[,3]=factor(data[,3],levels=c("STD (Our method)","HOLRR (Low-rank regression)","HOPLS (Partial least square)","TPG (Projection gradient)"))
+p2=ggplot(data=data, aes(x=as.factor(Category),y=PMSE, fill=Method))+geom_bar(stat="identity", position=position_dodge())+geom_errorbar(aes(ymin=PMSE-sd, ymax=PMSE+sd), width=.2,position=position_dodge(.9))+labs(x="Number of modes with available features",y="MSPE")+coord_cartesian(ylim = c(0, 0.1)) +  labs(title = "Low Signal, High Rank",size = 5) +theme(plot.title = element_text(hjust = 0.5,size = 11))+
+scale_fill_manual(values=new_color)+theme(axis.text=element_text(size=12),axis.title=element_text(size=10))
+
+s=2;r=1;
[…]

### 2020-09-14 tensor_regress/JASA_submission/code/Figure6.R

@@ -1,0 +1,18 @@
+### Code for Figure 6. HCP data analysis
+install.packages("tensorregress.tar.gz",repos = NULL, type="source")
+library(tensorregress)
+
+data=load("presaved/HCP.RData")
+tsr=tensor
+X=attr[,4:5]
+levels(X[,2])=c("22-25","26-30","31+","31+") ## three groups
+contrasts(X[,1]) <- contr.sum
+contrasts(X[,2]) <- contr.sum
+X_covar3=model.matrix(~as.factor(X[,1])+as.factor(X[,2]))
+
+core_shape=c(10,10,4)
+## the following line takes quite a while to finish. We do not sugget to run on personal laptop. We implement it in the server.
+result=tensor_regress(tsr,X_covar1=NULL,X_covar2=NULL,X_covar3,core_shape,Nsim=50,cons="non",lambda=0.1,alpha=10,solver="CG",dist="binary")
+
+save(result,file="output_HCP.RData")
+### the matrix for plotting brain connectivity is saved in presaved/data_for_Figure6

### 2020-09-14 tensor_regress/JASA_submission/code/SFigure1.R

@@ -1,0 +1,114 @@
+## Code for Supplementary Figure 1. Comparison of mean sqaured prediction error between our method and the other three previous methods
+install.packages("tensorregress.tar.gz",repos = NULL, type="source")
+source("simulation.R")
+source("comparison.R")
+library(tensorregress)
+library(ggplot2)
+
+############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line #40 to run the code ##########
+
+######## plot MSPE for three methods under different (signal, rank) setting #######
+load("presaved/SFigure1.RData")
+library(ggplot2)
+library(patchwork)
+new_color = c("#069AA0","#CCC591","#BCA455","#D6CFC4")
+
+signal_range=c(3,6)
+core_range=rbind(c(3,3,3),c(4,5,6))
+pdf("SFigure1.pdf",width=20.5,height=3)
+s=r=1;
+data=data.frame(PMSE=c(final[s,r,1,],final[s,r,2,],final[s,r,3,])/signal_range[s],sd=c(finalsd[s,r,1,],finalsd[s,r,2,],finalsd[s,r,3,])/signal_range[s],Method=rep(c('STD (Our method)','HOLRR (Low-rank regression)','HOPLS (Partial least square)','TPG (Projection gradient)'),3),Category=c(rep(1,4),rep(2,4),rep(3,4)))
+data[,3]=factor(data[,3],levels=c("STD (Our method)","HOLRR (Low-rank regression)","HOPLS (Partial least square)","TPG (Projection gradient)"))
+p1=ggplot(data=data, aes(x=as.factor(Category),y=PMSE, fill=Method))+geom_bar(stat="identity", position=position_dodge())+geom_errorbar(aes(ymin=PMSE-sd, ymax=PMSE+sd), width=.2,position=position_dodge(.9))+labs(x="Number of modes with available features",y="MSPE")+coord_cartesian(ylim = c(0, max(data[,1])+0.05))+  labs(title = "Low Signal, Low Rank",size = 5) +theme(plot.title = element_text(hjust = 0.5,size = 11))+
+    scale_fill_manual(values=new_color)+theme(axis.text=element_text(size=12),axis.title=element_text(size=10))
+
+s=1;r=2;
[…]

### 2020-09-14 tensor_regress/JASA_submission/code/SFigure2.R

@@ -1,0 +1,123 @@
+## Code for Supplementary Figure2. Comparison of mean sqaured prediction error between our method and the other three previous methods
+install.packages("tensorregress.tar.gz",repos = NULL, type="source")
+source("simulation.R")
+source("comparison.R")
+library(tensorregress)
+
+############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line #40 to run the code ##########
+
+######## plot MSPE for three methods under different (signal, rank) setting #######
+load("presaved/SFigure2.RData")
+library(ggplot2)
+library(patchwork)
+d=c(20,40,60,80,100)
+signal_range=c(3,6)
+core_range=rbind(c(3,3,3),c(4,5,6))
+
+
+signal_range=c(3,6)
+new_color = c("#069AA0","#CCC591","#BCA455","#D6CFC4")
+pdf("SFigure2.pdf",width=21,height=3)
+s=r=1;
+data = data.frame(d = rep(d,4), Method = c(rep('STD (Our method)',length(d)),rep('HOLRR (Low-rank regression)',length(d)),rep('HOPLS (Partial least square)',length(d)),rep('TPG (Projection gradient)',length(d))),PMSE = c(final[s,r,,1],final[s,r,,2],final[s,r,,3],final[s,r,,4])/signal_range[s],sd = c(finalsd[s,r,,1],finalsd[s,r,,2],finalsd[s,r,,3],finalsd[s,r,,4])/signal_range[s])
+data[,2]=factor(data[,2],levels=c("STD (Our method)","HOLRR (Low-rank regression)","HOPLS (Partial least square)","TPG (Projection gradient)"))
+p1=ggplot(data, aes(x = d*400, y = PMSE)) + geom_line(aes(color = Method),size = 1.2) + scale_colour_manual(values = new_color) +
+        geom_point(size = 2,aes(shape = Method)) + scale_shape_manual(values = c(16,5,17,4)) + theme(axis.text=element_text(size=12), axis.title=element_text(size=10))+ xlab('Sample Size') + labs(title="Low Signal, Low Rank") + ylab('MSPE')+ theme(plot.title = element_text(hjust = 0.5,size = 11))+coord_cartesian(ylim = c(0, max(data[,3])+0.03))
+
[…]

### 2020-09-14 tensor_regress/JASA_submission/code/simulation.R

@@ -1,0 +1,99 @@
+######### massive GLM to estimate the covariate effects ###
+massive_glm=function(response,X,dist){
+    d1=dim(response)[1]
+    d2=dim(response)[2]
+    coe=array(0,dim=c(d1,d2,dim(X)[2]))
+    if(dist=="normal"){
+        for(i in 1:d1){
+            for(j in 1:d2){
+                fit=lm(response[i,j,]~-1+X)
+                coe[i,j,]=coef(fit)
+            }
+        }
+    }
+    else if(dist=="binary"){
+        for(i in 1:d1){
+            for(j in 1:d2){
+                fit=suppressWarnings(glm(response[i,j,]~-1+X,family=binomial("logit")))
+                coe[i,j,]=coef(fit)
+            }
+        }
+    }
+    else if(dist=="poisson"){
+        for(i in 1:d1){
+            for(j in 1:d2){
+                fit=suppressWarnings(glm(response[i,j,]~-1+X,family=poisson("log")))
+                coe[i,j,]=coef(fit)
+            }
+        }
+    }
+    return(coe)
+}
+
+#### valid rank range for core tensor ## the maximum rank should be no smaller than the product of other.
+## Input: rank_range = expand.grid(...)
+valid_rank=function(rank_range){
+    N=dim(rank_range)[1]
+    output=NULL
+    for(i in 1:N){
+        rank=rank_range[i,]
+        rank=sort(rank)
+        if(rank[1]*rank[2]>=rank[3])
+        output=rbind(output,rank_range[i,])
+    }
+    return(output)
+}
+
+### compute loglikelihood under different models
+loglike=function(data,linearpre,dist){
+    if(dist=="binary"){
+        p=plogis(linearpre)## log-likelihood
+        L=dbinom(data,1,p,log=TRUE)
+        return(sum(L))
+    }
+    else if (dist=="normal"){
+        sigma_est=mean((data-linearpre)^2)
+        L=dnorm(data,linearpre,sqrt(sigma_est),log=TRUE)
+        sum(L)##log-likelihood
+    }
+    else if (dist=="poisson"){
+        lambda=exp(linearpre)
+        L=dpois(data,lambda, log = TRUE) ## log-likelihood for Poisson
+        return(sum(L))
+    }
+}
+
+#### wrapped simulation function to estimate covariate effects using tensor methods or massive glm methods
[…]

### 2020-09-14 tensor_regress/JASA_submission/code/tensorregress/R/tensor_regress.R

@@ -1,0 +1,447 @@
+#' Generalized Tensor Regression
+#'
+#' Tensor-response regression given covariates on multiple modes. Main function in the package. The function takes a response tensor, multiple covariate matrices, and a desired Tucker rank as input. The output is a constrained
+#' MLE for the coefficient tensor.
+#'
+#' @param tsr    response tensor with 3 modes
+#' @param X_covar1    covariate on first mode
+#' @param X_covar2    covariate on second mode
+#' @param X_covar3    covariate on third mode
+#' @param core_shape  the Tucker rank of the regression coefficients
+#' @param Nsim        max number of iterations if update does not convergence
+#' @param cons        the constraint method, "non" for without constraint, "vanilla" for global scale down at each iteration, "penalty" for adding log-barrier penalty to object function
+#' @param lambda      penalty coefficient for "penalty" constraint
+#' @param alpha       max norm constraint on linear predictor
+#' @param solver      solver for solving object function when using "penalty" constraint, see "details"
+#' @param dist        distribution of the response tensor, see "details"
+#' @return     a list containing the following:
+#'
+#'                  \code{W} {a list of orthogonal coefficient matrices - one for each mode, with the number of columns given by \code{core_shape}}
+#'
+#'                  \code{G}  {an array, core tensor with the size specified by \code{core_shape}}
+#'
+#'                  \code{C_ts}  {an array, coefficient tensor, Tucker product of \code{G},\code{A},\code{B},\code{C}}
+#'
+#'                  \code{U} {linear predictor,i.e. Tucker product of \code{C_ts},\code{X_covar1},\code{X_covar2},\code{X_covar3}}
+#'
+#'                  \code{lglk} {a vector containing loglikelihood at convergence}
+#'
[…]

### 2020-09-14 tensor_regress/code_JASA/Figure7.R

@@ -1,43 +1,0 @@
-### Code for Figure 8. Comparison between classical GLM and tensor methods in the HCP data anaysis
-source("simulation.R")
-data=load("presaved/HCP.RData")
-############ Method 1: tensor method  ##############
-load("presaved/output_HCP.RData")
-table(attr[,5])
-## age group: 22-25 26-30 31+
-## number of individuals: 35  58    43
-X=attr[,4:5]
-levels(X[,2])=c("22-25","26-30","31+","31+") ## three age groups
-contrasts(X[,1]) <- contr.sum
-contrasts(X[,2]) <- contr.sum
-### intercept, F 1, M,-1; 22-25 [1,0]; 26-30 [0 1]; 30+
-coef=array(0,dim=c(68,68,5))
-coef[,,1:4]=result$C_ts[,,1:4]##intercept, gender, age 22-25, age 26-30
-coef[,,5]=-coef[,,3]-coef[,,4] ## age 30+
-
-
-############## Method 2: Classical GLM  ##############
-tsr=tensor
-X_covar3=model.matrix(~as.factor(X[,1])+as.factor(X[,2])) ## baseline female and age 22-25
-massive=massive_glm(tsr,X_covar3,"binary")
-mass=array(0,dim=c(68,68,5))
-mass[,,1]=massive[,,1]
-mass[,,2]=massive[,,2]
-mass[,,3]=massive[,,3]
-mass[,,4]=massive[,,4]
-mass[,,5]=-massive[,,4]-massive[,,3]
-
-
-### Plot histogram to compare two methods###
-pdf("compare.pdf",width=15,height=6)
-par(mfrow=c(2,4))
-hist(c(mass[,,1]),nclass=40,xlab="Effect size",main="Intercept (Classical GLM)")
-hist(c(mass[,,2]),nclass=40,xlab="Effect size",main="Gender (Classical GLM)")
-hist(c(mass[,,4]),nclass=40,xlab="Effect size",main="Age 26-30 (Classical GLM)")
-hist(c(mass[,,5]),nclass=40,xlab="Effect size",main="Age 31+ (Classical GLM)")
-
-hist(c(coef[,,1]),nclass=40,xlab="Effect size",main="Intercept (Tensor regression)")
-hist(c(coef[,,2]),nclass=40,xlab="Effect size",main="Gender (Tensor regression)")
[…]

### 2020-09-14 tensor_regress/code_JASA/comparison.R

@@ -1,221 +1,0 @@
-## The script implements three alternative methods of supervised tensor decomposition
-## This part follows algorithm 1 in Guillaume et al. 2016
-## HOLRR
-library(rTensor)
-HOLRR = function(X,tsr,core_shape, gamma = 0){
-  # core_shape = (d0, d1, ... dp)
-  # X_shape =  N , d0
-  # tsr_shape = N, d1, d2, ...,dp
-  p = length(core_shape) - 1
-  tsr = as.tensor(tsr)
-  Y_1 = k_unfold(tsr, m = 1)@data
-  I = diag(dim(X)[2])
-  A = solve(t(X)%*%X + gamma*I)%*%t(X)%*%Y_1%*%t(Y_1)%*%X
-  if(core_shape[1] == dim(X)[2]){
-    U0 = diag(core_shape[1])
-  }else{
-    U0 = eigen(A)$vectors[,1:core_shape[1]]
-  }
-  U = list(); U[[1]] = U0
-  for(i in seq(2,length(core_shape))){
-    if(core_shape[i] == dim(tsr)[i]){
-      U[[i]] = diag(core_shape[i])
-    }else{
-      Y_i = k_unfold(tsr, m = i)@data
-      U[[i]] = eigen(Y_i%*%t(Y_i))$vectors[,1:core_shape[i]]
-    }
-  }
-  M = solve(t(U0)%*%(t(X)%*%X + gamma*I)%*%U0)%*%t(U0)%*%t(X)
-  
-  U_M = lapply(seq(length(core_shape)), function(x) t(U[[x]]))
-  U_M[[1]] = M
-  
-  G = ttl(tsr, U_M, ms = seq(p+1))
-  C_ts = ttl(G, U, ms = seq(p+1))
-  pre=ttm(C_ts,X,1)
-  return(list(G = G@data, U = U, C_ts = C_ts@data,pre=pre))
-}
-
-  
-## This part follows algorithm 1+2 --> 3 in Q. Zhao et al. 2012
-## HOPLS: for N-order tensor, N \geq 3
-## Input: X is a matrix
-HOPLS = function(X,tsr, R, Kn, tol){
-    # R: number of latent vectors
-    # Kn: vector, contains number of loadings in tsr (see fig 2 in Q. Zhao et al. 2012)
-    # tol: tolerance, epsilon in algorithm
-    E = list(); Fr = list() ## denote residuals of X, tsr
-    t = list() ## latent vectors
-    Q = list() ## loadings
[…]

### 2020-09-14 tensor_regress/code_server/run_F3.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+# untar your R installation. Make sure you are using the right version!
+tar -xzf R351.tar.gz
+# (optional) if you have a set of packages (created in Part 1), untar them also
+tar -xzf packages.tar.gz
+
+# make sure the script will use your R installation,
+# and the working directory as its home location
+export PATH=$PWD/R/bin:$PATH
+export RHOME=$PWD/R
+export R_LIBS=$PWD/packages
+
+# run your script
+Rscript server_F3.R

### 2020-09-14 tensor_regress/code_server/run_F5.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+# untar your R installation. Make sure you are using the right version!
+tar -xzf R351.tar.gz
+# (optional) if you have a set of packages (created in Part 1), untar them also
+tar -xzf packages.tar.gz
+
+# make sure the script will use your R installation,
+# and the working directory as its home location
+export PATH=$PWD/R/bin:$PATH
+export RHOME=$PWD/R
+export R_LIBS=$PWD/packages
+
+# run your script
+Rscript server_F5.R

### 2020-09-14 tensor_regress/code_server/server_F3.R

@@ -1,0 +1,25 @@
+source("simulation.R")
+library(tensorregress)
+set.seed(0) ## set seed for this simulation
+seed=0
+
+### all covariates
+dup=30
+d_pre=c(30,35,40,45,50,55,60)
+p_pre=round(0.4*d_pre)
+c_pre=c(2,4,6)
+d_range=cbind(d_pre,d_pre,d_pre)
+p_range=cbind(p_pre,p_pre,p_pre)
+c_range=cbind(c_pre,c_pre,c_pre)
+
+
+F1_normal=conv_rate(seed=NA,signal=10,cons="non",c_range=c_range,dist="normal",alpha=10,dup=dup,d_range=d_range,p_range=p_range,naive=FALSE)
+save(F1_normal,file="F1_normal.RData")
+
+F1_binary=conv_rate(seed=NA,signal=10,cons="non",c_range=c_range,dist="binary",alpha=10,dup=dup,d_range=d_range,p_range=p_range,naive=FALSE)
+save(F1_binary,file="F1_binary.RData")
+
+F1_poisson=conv_rate(seed=NA,signal=10,cons="non",c_range=c_range,dist="poisson",alpha=10,dup=dup,d_range=d_range,p_range=p_range,naive=FALSE)
+save(F1_poisson,file="F1_poisson.RData")
+
+

### 2020-09-14 tensor_regress/code_server/server_F3_11.R

@@ -1,0 +1,25 @@
+source("simulation.R")
+library(tensorregress)
+set.seed(0) ## set seed for this simulation
+seed=0
+
+### all covariates
+dup=30
+d_pre=c(30,35,40,45,50,55,60)
+p_pre=round(0.4*d_pre)
+c_pre=c(2,4,6)
+d_range=cbind(d_pre,d_pre,d_pre)
+p_range=cbind(p_pre,p_pre,p_pre)
+c_range=cbind(c_pre,c_pre,c_pre)
+
+
+F1_normal=conv_rate(seed=NA,signal=10,cons="non",c_range=c_range,dist="normal",alpha=10,dup=dup,d_range=d_range,p_range=p_range,naive=FALSE)
+save(F1_normal,file="F1_normal_new.RData")
+
+F1_binary=conv_rate(seed=NA,signal=10,cons="non",c_range=c_range,dist="binary",alpha=10,dup=dup,d_range=d_range,p_range=p_range,naive=FALSE)
+save(F1_binary,file="F1_binary_new.RData")
+
+F1_poisson=conv_rate(seed=NA,signal=10,cons="non",c_range=c_range,dist="poisson",alpha=10,dup=dup,d_range=d_range,p_range=p_range,naive=FALSE)
+save(F1_poisson,file="F1_poisson_new.RData")
+
+

### 2020-09-14 tensor_regress/code_server/server_F4.R

@@ -1,0 +1,56 @@
+source("simulation.R")
+library(tensorregress)
+
+set.seed(0)
+seed=0
+## one-side covariates
+c_range=c(5,5,5) ## selected rank
+c_range=rbind(c_range,c(10,10,5),c(15,15,5))
+optimal_rank=c(5,5,5)
+optimal_rank=rbind(optimal_rank,c(10,10,5),c(14,14,5)) ## to be determined
+dup=30
+
+### simulate data from normal model
+table=table_naive=matrix(0,nrow=3,ncol=dup)
+##############################
+for(s in 1:3){
+    data=sim_data(whole_shape=c(20,20,50),core_shape=c_range[s,],p=c(0,0,5),dist="normal",dup=dup,signal=5,block=c(TRUE,TRUE,FALSE))
+    for(i in 1:dup){
+        result=tensor_regress(tsr=data$tsr[[i]],X_covar1 =data$X_covar1, X_covar2 = data$X_covar2,X_covar3= data$X_covar3, core_shape=optimal_rank[s,], Nsim=10, cons="no", lambda = 0.1, alpha = 1, solver ="CG",dist="normal")
+        table[s,i]=mean((result$C_ts-data$C_ts)^2) ## tensor method
+        naive_C=massive_glm(data$tsr[[i]],data$X_covar3,dist="normal")
+        table_naive[s,i]=mean((naive_C-data$C_ts)^2) ## GLM method
+    }
+}
+save(table,table_naive,file="comparison_normal.RData")
+
+### simulate data from poisson model
+table=table_naive=matrix(0,nrow=3,ncol=dup)
+##############################
+for(s in 1:3){
+    data=sim_data(whole_shape=c(20,20,50),core_shape=c_range[s,],p=c(0,0,5),dist="poisson",dup=dup,signal=5,block=c(TRUE,TRUE,FALSE))
+    for(i in 1:dup){
+        result=tensor_regress(tsr=data$tsr[[i]],X_covar1 =data$X_covar1, X_covar2 = data$X_covar2,X_covar3= data$X_covar3, core_shape=optimal_rank[s,], Nsim=10, cons="no", lambda = 0.1, alpha = 1, solver ="CG",dist="poisson") ## tensor method
+        table[s,i]=mean((result$C_ts-data$C_ts)^2)
+        naive_C=massive_glm(data$tsr[[i]],data$X_covar3,dist="poisson") ## glm method
[…]

### 2020-09-14 tensor_regress/code_server/server_F5.R

@@ -1,0 +1,71 @@
+source("simulation.R")
+source("comparison.R")
+library(tensorregress)
+
+set.seed(0)
+seed=0
+## combinations of signal and rank
+signal_range=c(3,6)
+core_range=rbind(c(3,3,3),c(4,5,6))
+##
+dup=30;
+d=20;
+whole_shape = rep(d,3);
+final=array(0,dim=c(2,2,3,4))
+finalsd=array(0,dim=c(2,2,3,4))
+err_res=err_holrr=err_pls=err_tpg=array(0,dim=c(2,2,3,dup))
+dist="normal";
+
+
+for(s in 1:2){ ## signal level
+    for(r in 1:2){ ## rank
+        for(i in 1:3){ ## number of informative covariates
+            for(n in 1:dup){ ## simulation replicates
+                core_shape=core_range[r,]
+                signal=signal_range[s]
+                if(i==1){ ## one mode covaraite
+                    data=sim_data(whole_shape = whole_shape, core_shape=core_shape,p=c(0.4*d,0,0),dist=dist, dup=dup, signal=signal)
+                    res = tensor_regress(data$tsr[[n]],X_covar1 = data$X_covar1,core_shape=core_shape,Nsim=10, cons = 'non', dist = dist)### tensor_regress
+                }else if(i==2){ ## two mode covaraite
+                    data=sim_data(whole_shape = whole_shape, core_shape=core_shape,p=c(0.4*d,0.4*d,0),dist=dist, dup=dup, signal=signal)
+                    res = tensor_regress(data$tsr[[n]],X_covar1 = data$X_covar1,X_covar2 = data$X_covar2,core_shape=core_shape,Nsim=10, cons = 'non', dist = dist)### tensor_regress
+                }else if (i==3){## three model covariate
+                    data=sim_data(whole_shape = whole_shape, core_shape=core_shape,p=c(0.4*d,0.4*d,0.4*d),dist=dist, dup=dup, signal=signal)
+                    res = tensor_regress(data$tsr[[n]],X_covar1 = data$X_covar1,X_covar2 = data$X_covar2,X_covar3 = data$X_covar3,core_shape=core_shape,Nsim=10, cons = 'non', dist = dist)### tensor_regress
+                }
+                err_res[s,r,i,n]= mean((res$U - data$U)^2)
+                
+                ### holrr
[…]

### 2020-09-14 tensor_regress/code_server/server_T2.R

@@ -1,0 +1,72 @@
+source("simulation.R")
+library(tensorregress)
+
+set.seed(0) ## set seed for this simulation
+seed=0
+
+### all covariates included ##
+d=c(20,40)
+rank_range=array(0,dim=c(3,10,3))
+rank_range[1,,]=rbind(c(3,3,3),c(3,3,2),c(3,2,2),c(2,2,2),c(3,3,4),c(3,4,4),c(4,4,4),c(2,2,4),c(3,2,3),c(2,3,2))
+rank_range[2,,]=rbind(c(4,4,6),c(4,4,5),c(3,3,6),c(4,5,7),c(3,4,6),c(4,5,6),c(5,5,5),c(3,3,5),c(3,3,4),c(3,4,7))
+rank_range[3,,]=rbind(c(6,8,8),c(6,7,8),c(5,8,8),c(7,8,8),c(5,7,7),c(5,7,8),c(6,7,7),c(5,6,7),c(5,6,6),c(5,5,5))
+core_shape=rbind(c(3,3,3),c(4,4,6),c(6,8,8))
+
+########################## normal model ##########################
+dist="normal"
+dup=30
+final=array(0,dim=c(3,2,dup,3))
+for(s in 1:3){
+    for(i in 1:2){
+        whole_shape = c(d[i],d[i],d[i])
+        p=0.4*whole_shape
+        data=sim_data(seed, whole_shape = whole_shape, core_shape=core_shape[s,],p=p,dist=dist, dup=dup, signal=4)
+        table=NULL
+        
+        for(j in 1:dup){
+            res=sele_rank(data$tsr[[j]],X_covar1 = data$X_covar1, X_covar2 = data$X_covar2,X_covar3 = data$X_covar3,rank_range=rank_range[s,,],Nsim=10,cons = 'non',dist=dist)
+            table=rbind(table,res$rank)
+        }
+        
+        final[s,i,,]=table
+    }
+}
+
+save(final,file="normal_rank.RData")
+
+
+########################## poisson model ##########################
+dist="poisson"
+final=array(0,dim=c(3,2,dup,3))
+for(s in 1:3){
+    for(i in 1:2){
+        whole_shape = c(d[i],d[i],d[i])
+        p=0.4*whole_shape
+        data=sim_data(seed, whole_shape = whole_shape, core_shape=core_shape[s,],p=p,dist=dist, dup=dup, signal=4)
+        table=NULL
+        
[…]

### 2020-09-14 tensor_regress/oldcode/Figure5.R

@@ -1,115 +1,0 @@
-## Code for Figure 5. Comparison of mean sqaured prediction error between our method and the other three previous methods
-install.packages("tensorregress.tar.gz",repos = NULL, type="source")
-source("simulation.R")
-source("comparison.R")
-library(tensorregress)
-library(ggplot2)
-
-############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line #40 to run the code ##########
-
-######## plot MSPE for three methods under different (signal, rank) setting #######
-load("presaved/Figure5.RData")
-library(ggplot2)
-library(wesanderson)
-library(patchwork)
-new_color = c("#069AA0","#CCC591","#BCA455","#D6CFC4")
-
-signal_range=c(3,6)
-core_range=rbind(c(3,3,3),c(4,5,6))
-pdf("Figure5.pdf",width=20.5,height=3)
-s=r=1;
-data=data.frame(PMSE=c(final[s,r,1,],final[s,r,2,],final[s,r,3,])/signal_range[s],sd=c(finalsd[s,r,1,],finalsd[s,r,2,],finalsd[s,r,3,])/signal_range[s],Method=rep(c('GTD (Our method)','HOLRR (Low-rank regression)','HOPLS (Partial least square)','TPG (Projection gradient)'),3),Category=c(rep(1,4),rep(2,4),rep(3,4)))
-data[,3]=factor(data[,3],levels=c("STD (Our method)","HOLRR (Low-rank regression)","HOPLS (Partial least square)","TPG (Projection gradient)"))
-p1=ggplot(data=data, aes(x=as.factor(Category),y=PMSE, fill=Method))+geom_bar(stat="identity", position=position_dodge())+geom_errorbar(aes(ymin=PMSE-sd, ymax=PMSE+sd), width=.2,position=position_dodge(.9))+labs(x="Number of modes with available features",y="MSPE")+coord_cartesian(ylim = c(0, max(data[,1])+0.05))+  labs(title = "Low Signal, Low Rank",size = 5) +theme(plot.title = element_text(hjust = 0.5,size = 11))+
-    scale_fill_manual(values=new_color)+theme(axis.text=element_text(size=12),axis.title=element_text(size=10))
-
-s=1;r=2;
[…]

### 2020-09-14 tensor_regress/oldcode/Figure5and6.R

@@ -1,60 +1,0 @@
-## Code for Figure 5. Comparison of mean sqaured prediction error between our method and the other three previous methods
-install.packages("tensorregress.tar.gz",repos = NULL, type="source")
-source("simulation.R")
-source("comparison.R")
-library(tensorregress)
-library(ggplot2)
-
-############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line #40 to run the code ##########
-
-######## plot MSPE for three methods under different (signal, rank) setting #######
-load("presaved/Figure5.RData")
-library(ggplot2)
-library(wesanderson)
-library(patchwork)
-new_color = c("#069AA0","#CCC591","#BCA455","#D6CFC4")
-
-signal_range=c(3,6)
-core_range=rbind(c(3,3,3),c(4,5,6))
-pdf("Figure.pdf",width=12,height=6)
-
-s=1;r=2;
-data=data.frame(PMSE=c(final[s,r,1,],final[s,r,2,],final[s,r,3,])/signal_range[s],sd=c(finalsd[s,r,1,],finalsd[s,r,2,],finalsd[s,r,3,])/signal_range[s],Method=rep(c('STD (Our method)','HOLRR (Low-rank regression)','HOPLS (Partial least square)','TPG (Projection gradient)'),3),Category=c(rep(1,4),rep(2,4),rep(3,4)))
-data[,3]=factor(data[,3],levels=c("STD (Our method)","HOLRR (Low-rank regression)","HOPLS (Partial least square)","TPG (Projection gradient)"))
-p2=ggplot(data=data, aes(x=as.factor(Category),y=PMSE, fill=Method))+geom_bar(stat="identity", position=position_dodge())+geom_errorbar(aes(ymin=PMSE-sd, ymax=PMSE+sd), width=.2,position=position_dodge(.9))+labs(x="Number of modes with available features",y="MSPE")+coord_cartesian(ylim = c(0, 0.1)) +  labs(title = "Low Signal, High Rank",size = 5) +theme(plot.title = element_text(hjust = 0.5,size = 11))+
-    scale_fill_manual(values=new_color)+theme(axis.text=element_text(size=12),axis.title=element_text(size=10))
-
-s=2;r=1;
[…]

### 2020-09-14 tensor_regress/oldcode/Figure6.R

@@ -1,124 +1,0 @@
-## Code for Figure 6. Comparison of mean sqaured prediction error between our method and the other three previous methods
-install.packages("tensorregress.tar.gz",repos = NULL, type="source")
-source("simulation.R")
-source("comparison.R")
-library(tensorregress)
-
-############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line #40 to run the code ##########
-
-######## plot MSPE for three methods under different (signal, rank) setting #######
-load("presaved/Figure6.RData")
-library(ggplot2)
-library(wesanderson)
-library(patchwork)
-d=c(20,40,60,80,100)
-signal_range=c(3,6)
-core_range=rbind(c(3,3,3),c(4,5,6))
-
-
-signal_range=c(3,6)
-new_color = c("#069AA0","#CCC591","#BCA455","#D6CFC4")
-pdf("Figure6.pdf",width=21,height=3)
-s=r=1;
-data = data.frame(d = rep(d,4), Method = c(rep('STD (Our method)',length(d)),rep('HOLRR (Low-rank regression)',length(d)),rep('HOPLS (Partial least square)',length(d)),rep('TPG (Projection gradient)',length(d))),PMSE = c(final[s,r,,1],final[s,r,,2],final[s,r,,3],final[s,r,,4])/signal_range[s],sd = c(finalsd[s,r,,1],finalsd[s,r,,2],finalsd[s,r,,3],finalsd[s,r,,4])/signal_range[s])
-data[,2]=factor(data[,2],levels=c("STD (Our method)","HOLRR (Low-rank regression)","HOPLS (Partial least square)","TPG (Projection gradient)"))
-p1=ggplot(data, aes(x = d*400, y = PMSE)) + geom_line(aes(color = Method),size = 1.2) + scale_colour_manual(values = new_color) +
-        geom_point(size = 2,aes(shape = Method)) + scale_shape_manual(values = c(16,5,17,4)) + theme(axis.text=element_text(size=12), axis.title=element_text(size=10))+ xlab('Sample Size') + labs(title="Low Signal, Low Rank") + ylab('MSPE')+ theme(plot.title = element_text(hjust = 0.5,size = 11))+coord_cartesian(ylim = c(0, max(data[,3])+0.03))
-
[…]

### 2020-09-14 tensor_regress/oldcode/new_graph/new_graph.R

@@ -1,0 +1,232 @@
+#### code for new graph
+set.seed(0) ## set seed for the entire simulation
+library(tensorregress)
+library(ggplot2)
+#devtools::install_github("thomasp85/patchwork")
+library(patchwork)
+source("new_graph_bricks.R")
+source("sim_data.R")
+
+model = c("normal","binary","poisson")
+dimen = c(30,25)
+rank = c(6,3)
+#seeds = c(250,612,900)
+
+lglk_sum = list()
+true_lglk_sum = list()
+time_sum = list()
+
+
+
+for (i in 1:3) {
+  lglk = list()
+  true_lglk = c()
+  time = c()
+  
+  k = 1
+  for (j in 1:2) {
+    for (r in 1:2) {
+        if(i==1) signal=3
+        else if(i==2) signal=10
+        else if(i==3) signal=6
+      my_data = sim_data(seed = NA, rep(dimen[j],3), rep(rank[r],3), p = rep(floor(0.4*dimen[j]) ,3),dist = model[i], dup = 1,signal = signal)
+      ptm = proc.time()
+      my_reg =  tensor_regress1(tsr = my_data$tsr[[1]], X_covar1 = my_data$X_covar1, X_covar2 = my_data$X_covar2,X_covar3 = my_data$X_covar3,rep(rank[r],3), Nsim = 20, cons = "non",dist = model[i])
+      time1 = proc.time() - ptm
+      
+      time[k] = time1[3]
+      true_lglk[k] = loglike(my_data$tsr[[1]], my_data$U,dist = model[i])
+      lglk[[k]] = my_reg$lglk
+      
+      k = k+1
+    }
+  }
+  
+  lglk_sum[[i]] = lglk
+  true_lglk_sum[[i]] = true_lglk
+  time_sum[[i]] = time
+  
+}
+
+###### ggplot ######
+
+# do not use loop here, every subfigure has somewhere special
+
+## normal d = 30, r = 6,3
+
+lglk_normal = c(lglk_sum[[1]][[1]] ,lglk_sum[[1]][[2]])
+case_normal=  c(rep("d3r6",33), rep("d3r3",33))
+iter = rep(1:33,times = 2)
+olglk1 = true_lglk_sum[[1]][1]
+olglk2 = true_lglk_sum[[1]][2]
+plotdata = data.frame(lglk = lglk_normal,case = case_normal,iter = iter)
+
[…]

### 2020-09-14 tensor_regress/oldcode/new_graph/new_graph_bricks.R

@@ -1,0 +1,339 @@
+### bricks for new graph
+
+#library(rTensor)
+library(pracma)
+#library(gtools)
+library(MASS)
+library(speedglm)
+#library(fastglm)
+#library(rgl)
+#library("RColorBrewer")
+
+
+
+#######################
+
+######----------   Tensor regression
+
+## This function is indeed the same as tensor_regress in the package tensorregress
+## except the iteration number is forced to be larger than 8
+## even though the algorithm has already converged.
+## This modification is made for better visualization.
+
+tensor_regress1 = function(tsr,X_covar1 = NULL, X_covar2 = NULL,X_covar3 = NULL, core_shape, Nsim=20, cons = c("non","vanilla","penalty"), lambda = 0.1, alpha = 1, solver ="CG",dist = c("binary", "poisson","normal")){
+  
+  
+  tsr = as.tensor(tsr)
+  Y_1 = unfold(tsr, row_idx = 1, col_idx = c(2,3))@data
+  Y_2 = unfold(tsr, row_idx = 2, col_idx = c(1,3))@data
+  Y_3 = unfold(tsr, row_idx = 3, col_idx = c(1,2))@data
+  d1 = dim(tsr)[1] ; d2 = dim(tsr)[2] ; d3 = dim(tsr)[3]
+  r1 = core_shape[1] ; r2 = core_shape[2] ; r3 = core_shape[3]
+  
+  ###  check whether unsupervised on each mode
+  un_m1 = FALSE ; un_m2 = FALSE ; un_m3 = FALSE
+  if(is.null(X_covar1)|(identical(X_covar1,diag(d1)))) {X_covar1 = diag(d1) ; un_m1 = TRUE}
+  if(is.null(X_covar2)|(identical(X_covar2,diag(d2)))) {X_covar2 = diag(d2) ; un_m2 = TRUE}
+  if(is.null(X_covar3)|(identical(X_covar3,diag(d3)))) {X_covar3 = diag(d3) ; un_m3 = TRUE}
+  p1 = dim(X_covar1)[2] ; p2 = dim(X_covar2)[2] ; p3 = dim(X_covar3)[2]
+  
+  
+  if(dist=="binary"){
+    tsr.transform=as.tensor(2*tsr@data-1)
+  }else if(dist=="poisson"){
[…]

### 2020-09-17 tensor_regress/JASA_submission/paper_source/regress.tex
update

@@
 Multi-dimensional arrays, known as tensors, are often collected with side information on multiple modes in modern scientific and engineering studies.
-A popular example is in neuroimaging~\citep{zhou2013tensor,wang2019common}.
+A popular example is in neuroimaging~\citep{zhou2013tensor}.
 The brain connectivity networks are collected from a sample of individuals, accompanied by individual characteristics such as age, gender, and diseases status (see Figure~\ref{fig:intro1}a).
@@
 In addition to the aforementioned challenges, many tensor datasets consist of non-Gaussian measurements.
-Examples include the political interaction dataset \citep{hu2015scalable} which measures action counts between countries under various relation types, and the brain connectivity network dataset \citep{zhang2018mapping} which is a collection of binary adjacency matrices.
-Classical tensor decomposition methods are based on minimizing the Frobenious norm of the reconstruction error, leading to suboptimal predictions for binary- or count-valued response variables.
-A number of supervised tensor methods have been proposed \citep{narita2012tensor, zhao2012higher, yu2016learning} to address the tensor regression problem in various forms, such as scalar-to-tensor regression and tensor-response regression.
+Examples include the political interaction dataset \citep{nickel2011three} which measures action counts between countries under various relations, and the brain connectivity network dataset \citep{zhang2018mapping} which is a collection of binary adjacency matrices.
+Classical tensor decomposition methods are based on minimizing the Frobenious norm of deviation, leading to suboptimal predictions for binary- or count-valued response variables.
+A number of supervised tensor methods have been proposed \citep{narita2012tensor, zhao2012higher, yu2016learning,lock2018supervised} to address the tensor regression problem in various forms, such as scalar-to-tensor regression and tensor-response regression.
 These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest.
-The gap between theory and practice means a great opportunity to improve modeling paradigms to better capture the complexity in tensor data.
+The gap between theory and practice means a great opportunity to modeling paradigms and better capture the complexity in tensor data.
 
@@
 \includegraphics[width=16cm]{demo.pdf}
+\vspace{-.5cm}
 \end{center}
-\caption{Examples of supervised tensor decomposition with interactive side information.
-(a) Network population model.
-(b) Schematic diagram of supervised tensor decomposition for an order-3 tensor $\tY$ with feature matrices $\mX_k$ on each of the three modes.} \label{fig:intro1}
+\caption{Examples of supervised tensor decomposition with interactive side information. (a) Network population model. (b) Spatio-temporal growth model.} \label{fig:intro1}
+\vspace{-.3cm}
 \end{figure*}
@@
 
-The second line of work studies the tensor-on-tensor regression~\citep{raskutti2019convex, lock2018tensor, gahrooei2020multiple}.
+The second line of work studies the tensor-on-tensor regression~\citep{raskutti2019convex, lock2018tensor,gahrooei2020multiple}.
 Our model shares a common ground with earlier approaches, but we provide more efficient solutions to new settings that have more practical significance.
 As we show in Section~\ref{sec:connection}, the supervised tensor decomposition has an interesting connection with tensor-on-tensor regression.
-Previous methods~\citep{gahrooei2020multiple,lock2018tensor} mainly focus on Gaussian tensors.
+Previous methods~\citep{lock2018tensor} mainly focus on Gaussian tensors.
 The Frobenius norm used in the objective function is statistical suboptimal for general exponential family tensors.
@@
 We advocate a non-convex approach and provide strong evidence for its success in our setting.
-Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2019convex,gahrooei2020multiple}, and we go step further by finding the sufficient dimension reduction~\citep{adragni2009sufficient}, $\textup{Span}(\mM_k)$, that facilitates the identification of interaction effects in features (see Figure~\ref{fig:intro1}b).
+Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2019convex}, and we go step further by finding the sufficient dimension reduction~\citep{adragni2009sufficient}, $\textup{Span}(\mM_k)$, that facilitates the identification of interaction effects in features (see Figure~\ref{fig:intro1}b).
 The latter approach greatly improves the \emph{interpretability} in prediction.
@@
 
-The third line of work uses side information for various tensor learning tasks, such as for completion~\citep{narita2012tensor,cao2016semi} and for recommendation system~\citep{ioannidis2019coupled,farias2019learning}.
-These methods also study tensors with side information, same as we do, but they take regularization approaches to penalize predictions that are distant from side information.
+The third line of work uses side information for various tensor learning tasks, such as for completion~\citep{narita2012tensor} and for recommendation system~\citep{ioannidis2019coupled,farias2019learning}.
+These methods also study tensors with side information, but they take regularization approaches to penalize predictions that are distant from side information~\cite{lock2018supervised}.
 One important difference is that their goal is prediction but not parameter estimation.
@@
 
-The remainder of this paper is organized as follows.
-Section~\ref{sec:pre} introduces tensor notation and preliminaries.
+The paper is organized as follows.
+Section~\ref{sec:pre} introduces tensor preliminaries.
 Section~\ref{sec:model} presents the main model and three motivating examples for supervised tensor decomposition.
-Section~\ref{sec:est} describes the rank-constrained quasi-likelihood estimation and an alternating optimization algorithm.
-In Section~\ref{sec:simulation}, we present numerical experiments and assess the performance in comparison to alternative methods.
+We describe the quasi-likelihood estimation in Section~\ref{sec:est} and provide numerical experiments in Section~\ref{sec:simulation}.
+In Section~\ref{sec:data}, we present two data applications, one on diffusion tensor imaging data and the other on multi-relational network data.
+%and applications to diffusion tensor imaging data and multi-relational social network data in Section~\ref{sec:data}. 
+%In Section~\ref{sec:simulation}, we present numerical experiments and assess the performance in comparison to alternative methods.
 In Section~\ref{sec:data}, we apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
@@
 \section{Preliminaries}\label{sec:pre}
-We introduce the basic tensor properties used in the paper; more details on tensor notation can be found in~\cite{kolda2009tensor}.
-We use $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ to denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor, where $K$ corresponds to the number of modes of $\tY$ and is called the order.
-The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as
-\begin{equation}
-\tY \times_1\mX_1\ldots \times_K \mX_K=\entry{\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}x_{j_1,i_1}^{(1)}\ldots x_{j_K,i_K}^{(K)}},
-\end{equation}
+We introduce the basic tensor properties~\citep{kolda2009tensor} used in the paper.
+We use lower-case letters (e.g.,\ $a,b,c$) for scalars and vectors, upper-case boldface letters (e.g.,\ $\mA, \mB, \mC$) for matrices, and calligraphy letters (e.g.,\ $\tA, \tB, \tC$) for tensors of order three or greater.
+Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
+The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as $\tY \times_1\mX_1\times \cdots \times_K \mX_K=\entry{\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}x_{j_1,i_1}^{(1)}\cdots x_{j_K,i_K}^{(K)}}$,
 which results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor.
 For ease of presentation, we use the shorthand $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product.
-For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as
-\[
-\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.
-\] 
-The Frobenius norm and maximum norm of $\tY$ are defined as 
-\[
-\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}\quad \text{and}\quad \mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.
-\]
+For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as $\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.$
+The tensor Frobenius norm is defined as $\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}$, and the maximum norm is defined $\mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.$
 When $a$ is a vector, we use $\vnormSize{}{a}=\langle a,a \rangle^{1/2}$ to denote the vector 2-norm.
-
 A higher-order tensor can be reshaped into a lower-order object.
@@
 The multilinear rank of an order-$K$ tensor $\tY$ is defined as a length-$K$ vector $\mr=(r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\textup{Unfold}_k(\tY)$, $k=1,\ldots,K$.
-
-We use lower-case letters (e.g.,\ $a,b,c$) for scalars and vectors, upper-case boldface letters (e.g.,\ $\mA, \mB, \mC$) for matrices, and calligraphy letters (e.g.,\ $\tA, \tB, \tC$) for tensors of order three or greater.
+We use $\sigma_{\min}(\cdot)$ and $\sigma_{\max}(\cdot)$ denote the smallest and largest matrix singular values.
 We let $\mI_d$ denote the $d \times d$ identity matrix and $[d]$ denote the $d$-set $\{1,\ldots,d\}$.
 For ease of notation, we allow the basic arithmetic operators (e.g.,\ $+, -, \geq $) and univariate functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner.
-We use $f'(\cdot)$ to denote the first-order derivative and $f''(\cdot)$ the second-order derivative of function $f(\cdot)$.
+%We use $f'(\cdot)$, $f''(\cdot)$ to denote the first-order and second-order derivative of a function $f(\cdot)$.
 
@@
 \begin{align}\label{eq:tensormodel}
-&\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)=f(\Theta),\quad \text{with}\quad \Theta =\tB\times\{\mX_1,\ \ldots,\ \mX_K\} ,
+&\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)=f(\Theta),\quad \text{with}\quad \Theta =\tB\times\{\mX_1,\ldots, \mX_K\} ,
 \end{align}
@@
 \begin{example}[Spatio-temporal growth model]
-The growth curve model~\citep{gabriel1998generalised,srivastava2008models} was originally proposed as an example of bilinear model for matrix data, and we adopt its higher-order extension here.
+The growth curve model~\citep{gabriel1998generalised} was originally proposed as an example of bilinear model for matrix data, and we adopt its higher-order extension here.
 Let $\tY=\entry{y_{ijk}}\in\mathbb{R}^{d \times m\times n}$ denote the pH measurements of $d$ lakes at $m$ levels of depth and for $n$ time points.
@@
 \begin{example}[Network population model]\label{example:brain}
-Network response model~\citep{rabusseau2016low, zhang2018network} is recently developed for neuroimanig analysis.
+Network response model~\citep{rabusseau2016low} is recently developed for neuroimanig analysis.
 The goal is to study the relationship between brain network connectivity pattern and features of individuals.
@@
   \]
-  where $\mC\in\mathbb{R}^{r\times r}$ and $\mM\in\mathbb{R}^{p\times r}$ are based on the singular value decomposition of $\mB=\mM\mC\mM^T$. 
+  where $\mC\in\mathbb{R}^{r\times r}, \mM\in\mathbb{R}^{p\times r}$ are from the singular value decomposition of $\mB=\mM\mC\mM^T$. 
   \end{example}
 
-In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by features.
+%In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by features.
 The coefficient tensor $\tB$ in model~\eqref{eq:tensormodel} serves this goal by collecting the feature effects.
@@
 
-Our supervised tensor decomposition~\eqref{eq:decomp} is able to incorporate arbitrary numbers of feature matrices.
-When certain mode $k$ has no available side information, we set $\mX_k=\mI_{d_k}$ in the model~\eqref{eq:decomp}.
+In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by features.
+Our supervised tensor decomposition~\eqref{eq:decomp} allows arbitrary numbers of feature matrices.
+When certain mode $k$ has no side information, we set $\mX_k=\mI_{d_k}$ in the model~\eqref{eq:decomp}.
 In particular, our model~\eqref{eq:decomp} reduces to classical unsupervised tensor decomposition~\citep{de2000multilinear,hong2020generalized} when no side information is available; i.e., $\mX_k=\mI_{d_k}$ for all $k\in[K]$.
@@
 The role of the sufficient features $\mX_k\mM_k$ can be seen in the following two conditional independence assumptions in model~\eqref{eq:decomp},
+
 \begin{align}
-\tY \ci \{\mX_k\}&\ \big|\ \{\mX_k\mM_k\} \ \ \text{(independence between the tensor and multiple features)},\\
+\tY \ci \{\mX_k\}&\ \big|\ \{\mX_k\mM_k\} \ \ \text{(independence between the tensor and features)},\\
 y_{i_1,\ldots,i_K} \ci y_{i'_1,\ldots,i'_K} & \ \big|\ \{\mX_k\mM_k\}  \ \ \text{(independence within the tensor)},
@@
 where the second line holds for all $(i_1,\ldots,i_K)\neq(i'_1,\ldots,i'_K) \in[d_1]\times\cdots\times[d_K]$, and $\ci$ denotes the independence.
-The first property highlights the ``decorrelation'' role of $\mM_k$, in the same spirit as the sufficient dimension reduction~\citep{adragni2009sufficient} in high-dimensional supervised learning, whereas the second property highlights the tensor dimension reduction in the usual unsupervised sense (consider, for example, $\mX_k=\mathbf{I}_{d_k}$ for all $k\in[K]$).
+The first property highlights the ``decorrelation'' role of $\mM_k$, in the same spirit as the sufficient dimension reduction~\citep{adragni2009sufficient} in supervised learning, whereas the second property highlights the tensor dimension reduction in the usual unsupervised sense (consider, for example, $\mX_k=\mathbf{I}_{d_k}$ for all $k\in[K]$).
 
-Our model also has a close connection with tensor-on-tensor regression~\citep{raskutti2019convex, lock2018tensor,han2020optimal}.
+Our model also has a close connection with tensor-on-tensor regression~\citep{raskutti2019convex, lock2018tensor,gahrooei2020multiple}.
 Specifically, model~\eqref{eq:decomp} can be viewed as a multivariate regression model, where the response is the vectorized tensor and the covariates are interactions between sufficient features across modes.
-We take an order-3 tensor under and Gaussian model for illustration.
+We take an order-3 tensor under the Gaussian model for illustration.
 Let $\mX,\mZ,\mW$ denote the feature matrix on mode $k=1, 2, 3$, respectively.
@@
 For example, the observation domain is $\mathbb{Y}=\mathbb{R}$ for continuous data, $\mathbb{Y}=\mathbb{N}$ for count data, and  $\mathbb{Y}=\{0,1\}$ for binary data.
-The canonical link function $f$ is chosen to be $f(\cdot)=b'(\cdot)$.
+The canonical link function $f$ is chosen to be $f(\cdot)=b'(\cdot)$, the first-order derivative of $b(\cdot)$.
 Table~1 summarizes the canonical link functions for common types of distributions.
@@
 \end{tabular}
+\vspace{-.2cm}
 \caption{Canonical links for common distributions.}\label{table:link}
@@
 
-In our context, we model the tensor entries $y_{ijk}$, conditional on $\theta_{ijk}$, as independent draws from an exponential family.
+In our context, we model the tensor entries $y_{i_1,\ldots,i_K}$, conditional on $\theta_{i_1,\ldots,i_K}$, as independent draws from an exponential family.
 Ignoring constants that do not depend on $\Theta$, the quasi log-likelihood of~\eqref{eq:decomp} is equal to Bregman distance between $\tY$ and $b'(\Theta)$:
-\begin{align}\label{eq:loglikelihood}
-\tL_{\tY}(\tC,\mM_1,\ldots,\mM_K)&=\langle \tY, \Theta \rangle - \sum_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K}),\notag\\
-\text{where}\ \Theta&=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
-\end{align}
-
+\begin{equation}\label{eq:loglikelihood}
+\small
+\tL_{\tY}(\tC,\mM_1,\ldots,\mM_K)=\langle \tY, \Theta \rangle - \sum\nolimits_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K})\ \text{with}\ \Theta=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
+\end{equation}
 We propose a constrained maximum quasi-likelihood estimator (M-estimator),
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[!h]
 \caption{Supervised Tensor Decomposition with Interactive Side Information}\label{alg:B}
@@
 \State Obtain the factor matrix $\tilde \mM_k\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$.
-\State Perform QR factorization $\tilde \mM_k=\mQ\mR$, where $\mQ\in\mathbb{R}^{p_k\times r_k}$ consists of orthonormal columns. 
+\State Perform QR factorization $\tilde \mM_k=\mQ\mR$.
 \State Update $\mM_k\leftarrow \mQ$ and core tensor $\tC\leftarrow \tC\times_k \mR$.
@@
 Under all combinations of the dimension $d$, rank $r$, and type of the entries, Algorithm~\ref{alg:B} converges quickly in a few iterations, and the objective values at convergent points are close to or larger than the value at true parameters.
+The time complexity is provided in Supplementary Notes.
 
@@
 \includegraphics[width=15cm]{loglike.pdf}
+\vspace{-.25cm}
 \caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c) Poisson models.
@@
 }\label{fig:loglike}
+\vspace{-.3cm}
 \end{figure}
 
-\subsection{Rank selection and time complexity}\label{sec:tuning}
+\subsection{Rank selection}\label{sec:tuning}
 Algorithm~\ref{alg:B} assumes the rank $\mr$ is given.
 In practice, the rank is often unknown and must be determined from the data.
-We propose to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
-\begin{align}\label{eq:BIC}
-\hat \mr&=\argmin_{\mr=(r_1,\ldots,r_K)} \textup{BIC}(\mr)\notag\\
-&=\argmin_{\mr=(r_1,\ldots,r_K)}\left[-2\tL_{\tY}(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)+p_e(\mr)\log \left(\prod_k d_k\right) \right],
-\end{align}
-where $p_e(\mr)\stackrel{\textup{def}}{=}\sum_k (p_k-r_k-1)r_k+\prod_k r_k$ is the effective number of parameters in the model.
+We propose to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC, where
+\begin{equation}\label{eq:BIC}
+%\hat \mr&=\argmin_{\mr=(r_1,\ldots,r_K)} \textup{BIC}(\mr)\notag\\
+\textup{BIC}(\mr)=-2\tL_{\tY}(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)+p_e(\mr)\log (\prod\nolimits_k d_k).
+\end{equation}
+Here, $p_e(\mr)\stackrel{\textup{def}}{=}\sum_k (p_k-r_k-1)r_k+\prod_k r_k$ is the effective number of parameters in the model.
 We choose $\hat \mr$ that minimizes $\textup{BIC}(\mr)$ via grid search.
@@
 
-The computational complexity of our Algorithm~\eqref{alg:B} is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the data tensor.
+%The computational complexity of our Algorithm~\eqref{alg:B} is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the data tensor.
 More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor.
@@
 \begin{assumption}\label{ass}We make the following assumptions:
-\begin{enumerate}
+\begin{enumerate}[noitemsep,topsep=0pt]
 \item [A1.] There exist two positive constants $c_1, c_2>0$ such that $c_1\leq \sigma_{\min}(\mX_k)\leq  \sigma_{\max}(\mX_k)\leq c_2$ for all $k\in[K]$. 
-Here $\sigma_{min}(\cdot)$ and $\sigma_{max}(\cdot)$ denote the smallest and largest matrix singular values.
+%Here $\sigma_{min}(\cdot)$ and $\sigma_{max}(\cdot)$ denote the smallest and largest matrix singular values.
 \item [A1'.] The feature matrices $\mX_k$ are Gaussian designs with i.i.d.\ $N(0,1)$ entries.
-\item [A2.] There exist two positive constants $L, U>0$, such that $L\phi \leq \textup{Var}(y_{i_1,\ldots,i_K}|\theta_{i_1,\ldots,i_K})\leq U\phi $ for all $|\theta_{i_1,\ldots,i_K}|\leq \alpha$.
-Equivalently, $L\leq b''(\theta_{i_1,\ldots,i_K}) \leq U$ for all $|\theta_{i_1,\ldots,i_K}|\leq \alpha$.
-Here $\alpha$ is the upper bound of the linear predictor in~\eqref{eq:MLE}.
+\item [A2.] There exist two positive constants $L, U>0$, such that $L\phi \leq \textup{Var}(y_{i_1,\ldots,i_K}|\theta_{i_1,\ldots,i_K})\leq U\phi $, or equivalently, $L\leq b''(\theta_{i_1,\ldots,i_K}) \leq U$, for all $|\theta_{i_1,\ldots,i_K}|\leq \alpha$.
+Here $\alpha$ is the upper bound of the linear predictor in~\eqref{eq:MLE}, and $b''(\cdot)$ denotes the second-order derivative.
 
@@
 The assumptions are fairly mild.
-Assumptions A1 and A1' consider two different scenarios about feature matrices.
-We study the cases A1 and A1' separately but have the same theoretical results.
-Assumption A1 is applicable when feature matrix is asymptotically non-singular and has bounded spectral norm, whereas Assumption A1' imposes the commonly-used Gaussian design in high-dimensional statistics~\citep{chen2019non}.
+Assumptions A1 and A1' consider two separate scenarios about feature matrices.
+Assumption A1 is applicable when feature matrix is asymptotically non-singular and has bounded spectral norm, whereas Assumption A1' imposes the commonly-used Gaussian design~\citep{raskutti2019convex}.
 Assumption A2 ensures the log-likelihood $\tL_\tY(\Theta)$ is strictly concave in the linear predictor $\Theta$.
-The two equivalent statements in Assumption A2 are based on the fact that $\textup{Var}(y_{i_1,\ldots,i_K}|\tX,\tB)=\phi b''(\theta_{i_1,\ldots,i_K})$ when $y_{i_1,\ldots,i_K}$ belongs to an exponential family~\citep{mccullagh1989generalized}.
+%The two equivalent statements in Assumption A2 are based on the fact that $\textup{Var}(y_{i_1,\ldots,i_K}|\tX,\tB)=\phi b''(\theta_{i_1,\ldots,i_K})$ when $y_{i_1,\ldots,i_K}$ belongs to an exponential family~\citep{mccullagh1989generalized}.
 
@@
 Theorem~\ref{thm:main} establishes the statistical convergence for the estimator~\eqref{eq:MLE}.
-The proof is provided in Appendix~\ref{sec:appedix}.
-In fact, the proof shows that the desired convergence rate holds not only for the M-estimator, but also for any local optimizers satisfying $\tL_\tY(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K) \geq \tL_\tY(\tC_{\textup{true}},  \mM_{1,\textup{true}},\ldots,\mM_{K,\textup{true}})$.
+In fact, our proof shows that the desired convergence rate holds not only for the M-estimator, but also for any local optimizers satisfying $\tL_\tY(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K) \geq \tL_\tY(\tC_{\textup{true}},  \mM_{1,\textup{true}},\ldots,\mM_{K,\textup{true}})$.
 Consider a special case when tensor dimensions are equal on each of the modes, i.e., $d_k=d$ for all $k\in[K]$, and feature dimension grows with tensor dimension, $p_k=\gamma d$, $\gamma\in [0,1)$, for $k\in[K]$.
@@
 
-We conclude this section by providing the prediction accuracy, measured in KL divergence, for the response distribution.
+We conclude this section by providing the accuracy measured in KL divergence.
+%, for the response distribution.
 
-\begin{thm}[Prediction error]\label{thm:KL}
+\begin{cor}[Prediction error]\label{thm:KL}
 Assume the same set-up as in Theorem~\ref{thm:main}.
@@
 where $C_3=C_3(\alpha, K)>0$ is a constant independent of dimensions $\{d_k\}$ and $\{p_k\}$.
-\end{thm}
+\end{cor}
 
@@
 The linear predictor $\Theta=\tC\times\{\mM_1\mX_1,\mM_2\mX_3,\mM_3\mX_3\}$ is scaled such that $\mnormSize{}{\Theta}=1$.
-Conditional on the linear predictor $\Theta=\entry{\theta_{ijk}}$, the entries in the tensor $\tY=\entry{y_{ijk}}$ are drawn independently according to one of the following three probabilistic models:
+Conditional on the linear predictor $\Theta=\entry{\theta_{ijk}}$, the entries in the tensor $\tY=\entry{y_{ijk}}$ are drawn independently according to three probabilistic models:
 
-\begin{enumerate}
-\item[(a)] (Gaussian). Continuous entries $y_{ijk}\sim N\left(\alpha \theta_{ijk}, 1\right)$, for all $(i,j,k)\in[d_1]\times[d_2]\times[d_3]$.
-\item[(b)] (Poisson). Count entries $y_{ijk}\sim\textup{Poisson}\left( e^{\alpha \theta_{ijk}}\right)$, for all $(i,j,k)\in[d_1]\times[d_2]\times[d_3]$.
-\item[(c)] (Bernoulli). Binary entries $y_{ijk}\sim \textup{Bernoulli}\left( \frac{e^{\alpha \theta_{ijk}}}{1+e^{\alpha \theta_{ijk}}}\right)$, for all $(i,j,k)\in[d_1]\times[d_2]\times[d_3]$.
+\begin{enumerate}[noitemsep,topsep=0pt]
+\item[(a)] Gaussian model: continuous tensor entries $y_{ijk}\sim N\left(\alpha \theta_{ijk}, 1\right)$.
+\item[(b)] Poisson model: count tensor entries $y_{ijk}\sim\textup{Poisson}\left( e^{\alpha \theta_{ijk}}\right)$.
+\item[(c)] Bernoulli model: binary tensor entries $y_{ijk}\sim \textup{Bernoulli}\left( \frac{e^{\alpha \theta_{ijk}}}{1+e^{\alpha \theta_{ijk}}}\right)$.
 \end{enumerate}
 Here $\alpha>0$ is a scalar controlling the magnitude of the effect size.
-In each simulation study, we report the mean squared error (MSE), $\FnormSize{}{\tB_{\textup{true}}-\hat \tB}^2$, for the coefficient tensor averaged across $n_{\textup{sim}}=30$ replications.
+In each simulation study, we report the mean squared error (MSE), $\FnormSize{}{\tB_{\textup{true}}-\hat \tB}^2$, averaged across $30$ replications.
 
@@
 We set $\alpha=4$ and consider various combinations of dimension $d$ and rank $\mr=(r_1,r_2,r_3)$.
-For each combination, we simulate tensor data following Gaussian and Poisson models.
-We then minimize BIC using a grid search over three modes.
+For each combination, we minimize BIC using a grid search over three modes.
 The hyper-parameter $\alpha$ is set to infinity in the fitting, which essentially imposes no constraint on the tensor magnitude.
@@
 \begin{table}[tb]
+\resizebox{\textwidth}{!}{%
 \centering
-\begin{tabular}{|c|cc|cc|}
+\begin{tabular}{c|cc|cc}
+%True Rank $\mr$ &\multicolumn{2}{c|}{Dimension (Gaussian tensors)}&\multicolumn{2}{c|}{Dimension (Poisson tensors)}\\
+True rank $\mr$& $d=20$  (Gaussian) &$d=40$ (Gaussian) &$d=20$ (Poisson) &$d=40$ (Poisson)\\
 \hline
-True Rank &\multicolumn{2}{c|}{Dimension (Gaussian tensors)}&\multicolumn{2}{c|}{Dimension (Poisson tensors)}\\
-$\mr$& $d=20$  &$d=40$&$d=20$ &$d=40$\\
-\hline
-$(3,\ 3,\ 3)$&$({\bf 3},\ {\bf 3},\ {\bf 3})$&$({\bf 3},\ {\bf 3},\ {\bf 3})$& $({\bf 3},\ {\bf 3},\ {\bf 3})$ & $({\bf 3},\ {\bf 3},\ {\bf 3})$\\
-$(4,\ 4,\ 6)$&$(3,\ 3,\ {\bf 4.6})$&$({\bf 4},\ {\bf 4},\ {\bf 5.3})$&$(3,\ 3,\ {\bf 5.3})$&$({\bf 4},\ {\bf 4},\ {\bf 5.6})$\\
-$(6,\ 8,\ 8)$&$(5,\ 5,\ 5)$&$({\bf 6},\ {\bf 8},\ {\bf 8})$&$(5,\ 5,\ 6.7)$&$({\bf 6},\ {\bf 8},\ {\bf 8})$\\
-\hline
+$(3,\ 3,\ 3)$&$({\bf 3.0},\ {\bf 3.0},\ {\bf 3.0})$&$({\bf 3.0},\ {\bf 3.0},\ {\bf 3.0})$& $({\bf 3.0},\ {\bf 3.0},\ {\bf 3.0})$ & $({\bf 3.0},\ {\bf 3.0},\ {\bf 3.0})$\\
+$(4,\ 4,\ 6)$&$(3.0,\ 3.0,\ {\bf 4.6})$&$({\bf 4.0},\ {\bf 4.0},\ {\bf 5.3})$&$(3.0,\ 3.0,\ {\bf 5.3})$&$({\bf 4.0},\ {\bf 4.0},\ {\bf 5.6})$\\
+$(6,\ 8,\ 8)$&$(5.0,\ 5.0,\ 5.0)$&$({\bf 6.0},\ {\bf 8.0},\ {\bf 8.0})$&$(5.0,\ 5.0,\ 6.7)$&$({\bf 6.0},\ {\bf 8.0},\ {\bf 8.0})$\\
 \end{tabular}
+}
 \caption{Rank selection via BIC. The estimated ranks are averaged across 30 simulation. Bold number indicates the ground truth is within two standard deviations of the estimate.}\label{tab:rank}
+\vspace{-.1cm}
 \end{table}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=14cm]{dimension.pdf}
+\caption{Estimation error against effective sample size.
+The three panels plot the MSE when the response tensors are generated from (a) Gaussian (b) Poisson and (c) Bernoulli models.
+The dashed curves correspond to $\tO({1/d^2})$.}\label{fig:dim}
+\vspace{-.1cm}
+\end{figure}
 
@@
 Our theoretical analysis suggests that $\hat \tB$ has a convergence rate $\tO(d^{-2})$ in this setting.
-Figure~1 plots the estimation error versus the ``effective sample size'', $d^2$, under three different distribution models.
+Figure~\ref{fig:dim} plots the estimation error versus the ``effective sample size'', $d^2$, under three different distribution models.
 We find that the empirical MSE decreases roughly at the rate of $1/d^2$, which is consistent with our theoretical results.
@@
 Indeed, a larger $r$ implies a higher model complexity and thus greater difficulty in the estimation.
-Similar behaviors are observed in the non-Gaussian tensors in Figures~\ref{fig:dim}b-c.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=15cm]{dimension.pdf}
-\caption{Estimation error against effective sample size.
-The three panels plot the MSE when the response tensors are generated from (a) Gaussian (b) Poisson and (c) Bernoulli models.
-The dashed curves correspond to $\tO({1/d^2})$.}\label{fig:dim}
-\end{figure}
 
@@
 }\label{fig:glm}
+\vspace{-.2cm}
 \end{figure}
@@
 We compare our supervised tensor decomposition with three other tensor methods:
-\begin{itemize}
+\begin{itemize}[noitemsep,topsep=0pt]
 \item  Higher-order low-rank regression ({\bf HOLRR},~\citet{rabusseau2016low}) is a least-square based tensor regression that allows features on a single mode. 
@@
 The comparison is assessed from three aspects: (a) benefit of incorporating features from multiple modes; (b) prediction error with respect to sample size; (c) sensitivity of accuracy with respect to model complexity.
-We use similar simulation setups as in our experiment II, but consider combinations of rank ($\mr=(3,3,3)$ vs.\ $(4,5,6)$), signal ($\alpha = 3$ vs.\ $6$), and dimension ($d$ ranging from 20 to 100 for modes with features, $d = 20$ for modes without features).
+We use similar simulation setups as in our experiment II, but consider combinations of rank, $\mr=(3,3,3)$ (low) vs.\ $(4,5,6)$ (high), signal $\alpha = 3$ (low) vs.\ $6$ (high), dimension $d$ ranging from 20 to 100 for modes with features, and $d = 20$ for modes without features.
 Two methods ({\bf STD} and {\bf HOLRR}) require the tensor rank as inputs.
@@
 \centering
-\includegraphics[width=16cm]{compare_alternative.pdf} 
+\includegraphics[width=15cm]{compare_alternative.pdf} 
 \caption{Comparison between different tensor methods.
 Panels (a) and (b) plot MSPE versus the number of modes with available features.
-Panels (c) and (d) plot MSPE versus the effective sample size $d^2$.
-We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and signal $\alpha =3 $ (low), $\alpha=6$ (high).}~\label{fig:compare}
+Panels (c) and (d) plot MSPE versus the effective sample size $d^2$.}~\label{fig:compare}
+%We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and signal $\alpha =3 $ (low), $\alpha=6$ (high).}~\label{fig:compare}
+\vspace{-.4cm}
 \end{figure}
@@
 In the high-signal low-rank setting, our method has similar performance as {\bf HOLRR}, and the improvement becomes more pronounced in the low-signal high-rank setting.
-Note that the latter setting is a harder learning task because of the higher inter-mode complexity, and our {\bf STD} method shows the advantage in addressing this challenge.
+The latter setting is harder because of the higher inter-mode complexity, and our {\bf STD} method shows the advantage in addressing this challenge.
 
@@
 
-\subsection{Analyses of human brain connection data}
+\subsection{Application to human brain connection data}
 The Human Connectome Project (HCP) aims to build a network map that characterizes the anatomical and functional connectivity within healthy human brains~\citep{HCP}. 
-We analyze the structural connectivity patterns among 68 brain regions for 136 individuals from HCP.
+%We analyze the structural connectivity patterns among 68 brain regions for 136 individuals from HCP. 
 We follow the preprocessing procedure as in~\cite{zhang2018mapping} and parcellate the brain into 68 regions of interest~\citep{desikan2006automated}.
-The preprocessed dataset is released in our R package \texttt{tensorregress}.
-
 The dataset consists of 136 brain structural networks, one for each individual.
@@
 We consider four individual features: gender (65 females vs.\ 71 males), age 22-25 ($n=35$), age 26-30 ($n=58$), and age 31+ ($n=43$).
+The preprocessed dataset is released in our R package \texttt{tensorregress}.
 The goal is to identify the connection edges that are affected by individual features.
@@
 \centering
-\includegraphics[width=12cm]{HCP.pdf}
+\includegraphics[width=16cm]{HCP.pdf}
 \caption{Top edges with large effects.
@@
 }\label{fig:brain}
+\vspace{1cm}
+\includegraphics[width=16cm]{compare_HCP.pdf}
+\caption{Comparison of estimated feature effects the HCP data using (a) multi-response GLM and (b) supervised tensor decomposition (STD). }\label{fig:s1}
+\vspace{-.1cm}
 \end{figure}
 
-We perform the supervised tensor decomposition to the HCP data.
-The data tensor is binary, $\tY\in\{0,1\}^{68\times 68\times 136}$, and the features are of dimension 4 on the 3$^{\text{rd}}$ mode.
+We perform the supervised tensor decomposition to the HCP data. %The data tensor is binary, $\tY\in\{0,1\}^{68\times 68\times 136}$, and the features are of dimension 4 on the 3$^{\text{rd}}$ mode. 
 The BIC selection suggests a rank $\mr=(10,10,4)$ with quasi log-likelihood $\tL_{\tY}=-174654.7$.
 We utilize the sum-to-zero contrasts in coding the feature effects, and depict only the top 3\% edges whose connections are non-constant across the sample.
-Figure~\ref{fig:brain} shows the top edges with high effect size, overlaid on the Desikan atlas brain template~\citep{desikan2006automated,xia2013brainnet}.
+Figure~\ref{fig:brain} shows the top edges with high effect size, overlaid on the Desikan atlas brain template~\citep{desikan2006automated}.
 We find that the global connection exhibits clear spatial separation, and that the nodes within each hemisphere are more densely connected with each other (Figure~\ref{fig:brain}a).
@@
 This is in agreement with a recent study showing that female brains are optimized for inter-hemispheric communication~\citep{ingalhalikar2014sex}.
-We also found several edges with declined connection in the group Age 31+.
+We find several edges with declined connection in the group Age 31+.
 Those edges involve Frontal-pole (\emph{Fploe}), superior-frontal (\emph{SupF}) and Cuneus nodes.
-The Frontal-pole region has long been known for its importance in memory and cognition, and the detected decline with age further highlights its biological importance.
+The Frontal-pole region is known for its importance in memory and cognition, and the detected decline with age further highlights its biological importance.
 
-\begin{figure}[!h]
-\centering
-\includegraphics[width=16cm]{compare_HCP.pdf}
-\caption{Comparison of feature effect estimation in the HCP data.}\label{fig:s1}
-\end{figure}
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=16cm]{compare_HCP.pdf}
+%\caption{Comparison of feature effect estimation in the HCP data.}\label{fig:s1}
+%\end{figure}
 
 Figure~\ref{fig:s1} compares the estimated coefficients from our method with those from multiple-response GLM approach.
-The multiple-response GLM is to regress the brain edges, one at a time, on the individual-level features, and this logistic model is repeatedly fitted for every edge $\in [68]\times [68]$.
+The multiple-response GLM is to regress the brain edges, one at a time, on the individual-level features, and this model is repeatedly fitted for every edge $\in [68]\times [68]$.
 As we can see in the figure, our tensor decomposition method shrinkages the coefficients towards center, thereby implicitly enforcing the sharing between feature effects.
@@
 The second application studies the multi-relational networks with node-level attributes.
-We consider \emph{Nations} dataset~\citep{nickel2011three} which records 56 relations among 14 countries between 1950 and 1965 (post-World War II).
+We consider \emph{Nations} dataset~\citep{nickel2011three} which records 56 relations among 14 countries between 1950 and 1965.
 The multi-relational networks can be organized into a $14 \times 14 \times 56$ binary tensor, with each entry indicating the presence or absence of an action, such as ``sending tourist to'', ``export'', ``import'', between countries.
-The 56 relations span the fields of politics, economics, military, religion, and so on.
+The 56 relations span the fields of politics, economics, military, religion, etc.
 In addition, country-level attributes are also available, and we focus on the following six features: \emph{constitutional, catholics, law ngo, political leadership, geography}, and \emph{medicine ngo}.
-The goal is to identify the variation in connections due to country-level attributes and interactions thereof.
-One of the key features is that the 56 relations are correlated, and we would like to take it into account in assessing the feature effects.
-
-\begin{table}[H]
-\begin{tabular}{c|c|l}
-Cluster & Category & \multicolumn{1}{c}{Relations}\\
-\hline
-\hline
-\multirow{3}{*}{Cluster I}  & \multirow{3}{*}{Military} & warning,           violentactions, militaryactions, duration,\\ &&negativebehavior, severdiplomatic, expeldiplomats, protests,\\
-&&timesincewar, commonbloc0, blockpositionindex, commonbloc1\\   
-\hline
-\multirow{4}{*}{Cluster II}& \multirow{4}{*}{Emigrant} & emigrants, relemigrants, emigrants3, accusation, \\      
-&&nonviolentbehavior, students, relstudents, economicaid\\
-&&releconomicaid, negativecomm, relintergovorgs,relngo\\
-&&intergovorgs3, ngoorgs3, militaryalliance, commonbloc2\\    
-\hline
-\multirow{6}{*}{Cluster III} & \multirow{6}{*}{Economy} & treaties, reltreaties, oficialvisits, conferences, exportbooks\\
-&& relexportbooks, booktranslations, relbooktranslations\\
-&&boycottembargo, weightedunvote, unweightedunvote, tourism\\
-&&reltourism, tourism3, exports, relexports, exports3, \\
-&&intergovorgs, ngo, embassy, reldiplomacy,timesinceally,\\
-&& dependent, independence\\
-\hline
-Cluster IV & Territory &aidenemy, lostterritory, unoffialacts, attackembassy\\
-\hline
-\end{tabular}
-\caption{$K$-means clustering of relations based on dimension reduction on the 3$^\text{rd}$ mode. }\label{tab:s1}
-\end{table}
+The goal is to identify the variation in connections due to country-level attributes and their interactions.
+%One of the key features is that the 56 relations are correlated, and we would like to take it into account in assessing the feature effects.
 
 We apply our tensor model to the \emph{Nations} data.
-The multi-relational network $\tY\in\{0,1\}^{14\times 14\times 56}$ is a binary data tensor, and the country attributes $\mX\in\mathbb{R}^{14\times 6}$ are features on both the 1$^\text{st}$ and 2$^\text{nd}$ modes.
+The multi-relational network $\tY$ is a binary data tensor, and the country attributes $\mX\in\mathbb{R}^{14\times 6}$ are features on both the 1$^\text{st}$ and 2$^\text{nd}$ modes.
 The BIC criterion suggests a rank $\mr=(4,4,4)$ for the coefficient tensor $\tB\in\mathbb{R}^{6\times 6 \times 56}$.
@@
 We find that the relations reflecting the similar aspects of actions are grouped together.
-In particular, Cluster I consists of military relations such as \emph{violentactions}, \emph{warnings} and \emph{militaryactions}; Clusters II and III capture the economical relations such as \emph{economicaid, booktranslations, tourism}; and Cluster IV represents the political alliance and territory relations.
+In particular, Cluster I consists of military relations such as \emph{violentactions}, \emph{warnings} and \emph{militaryactions}; Clusters II and III capture the economic relations such as \emph{economicaid, booktranslations, tourism}; and Cluster IV represents the political alliance and territory relations.
 
@@
 \centering
-\includegraphics[width=12cm]{coef.pdf}
+\includegraphics[width=16cm]{coef.pdf}
 \caption{Estimated feature effects in the \emph{Nations} data analysis.
-Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warnning}, \emph{violentactions},\emph{treaties}, and \emph{aidenemy}, respectively.
+Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warnning}, \emph{violentactions}, \emph{treaties}, and \emph{aidenemy}, respectively.
 }\label{fig:est}
+\vspace{-.4cm}
 \end{figure}
+
+\begin{table}[!h]
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{c|c|l}
+ & Category & \multicolumn{1}{c}{Relations}\\
+\hline
+\multirow{2}{*}{ I}  & \multirow{2}{*}{Military} & warning, violentactions, militaryactions, duration, negativebehavior, protests, severdiplomatic \\
+&&timesincewar, commonbloc0, commonbloc1, blockpositionindex,  expeldiplomats\\   
+\hline
+\multirow{2}{*}{II}& \multirow{2}{*}{Emigrant} & emigrants, emigrants3, relemigrants, accusation, nonviolentbehavior, ngoorgs, commonbloc2, intergovorgs3\\      
+&&releconomicaid, relintergovorgs, relngo,  students, relstudents, economicaid, negativecomm, militaryalliance \\
+\hline
+\multirow{3}{*}{ III} & \multirow{3}{*}{Economics} & treaties, reltreaties, officialvisits, exportbooks, relexportbooks, booktranslations, relbooktranslations\\
+&&boycottembargo, weightedunvote, unweightedunvote, reltourism, tourism, tourism3, exports, exports3  \\
+&&relexports, intergovorgs, ngo, embassy, reldiplomacy, timesinceally, independence, conferences, dependent\\
+\hline
+IV & Territory &aidenemy, lostterritory, unoffialacts, attackembassy\\
+\hline
+\end{tabular}
+}
+\caption{$K$-means clustering of relations based on dimension reduction on the 3$^\text{rd}$ mode. }\label{tab:s1}
+\vspace{-.4cm}
+\end{table}
 
@@
 The entries in $\tB$ are useful for revealing interaction effects in a context-specific way.
-For a given relation $k$, the off-diagonal entries in the matrix $\hat \tB(\colon,\colon,k)$ represent the pairwise interactions between features.
+%For a given relation $k$, the off-diagonal entries in the matrix $\hat \tB(\colon,\colon,k)$ represent the pairwise interactions between features. 
 From Figure~\ref{fig:est}, we find a strong interaction between \emph{geography} and \emph{political leadership} in the relation ``warning'', and a strong interaction between \emph{geogrphy} and \emph{medicine ngo} in the relation ``aidenemy''.
@@
 
-We comment on a number of possible extensions.
+%We comment on a number of possible extensions. 
 We have provided accuracy guarantees for parameter estimation in the supervised tensor model.
@@
 Measures of uncertainty, such as confidence envelope for space estimation, would be useful.
-One possible approach would be performing parametric bootstrap~\citep{efron1994boot} to assess the uncertainty in the estimation.
-For example, one can simulate tensors from the fitted low-rank model based on the estimates, and then assess the empirical distribution of the estimates.
+One possible approach would be performing parametric bootstrap~\citep{tibshirani1993introduction} to assess the uncertainty in the estimation.
+%For example, one can simulate tensors from the fitted low-rank model based on the estimates, and then assess the empirical distribution of the estimates. 
 While being simple, bootstrap approach is often computationally expensive for large-scale data.
@@
 \begin{align}\label{eq:extention}
-\mathbb{E}(\tY|\mX_1,\ldots,\mM_K)&=\tC\times\mM_1\times \cdots \times \mM_K, \notag \\
-\text{Var}(\tY|\mX_1,\ldots,\mM_K)&=\mPhi_1\otimes \cdots \otimes \mPhi_K,
+\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)&=\tC\times\mM_1\times \cdots \times \mM_K\ \ \text{and}\ \ \text{Var}(\tY|\mX_1,\ldots,\mX_K)=\mPhi_1\otimes \cdots \otimes \mPhi_K,
 \end{align}
@@
 The joint estimation of mean model $\Theta$ and variance model $\mPhi_k$ will lead to more efficient estimation in the presence of unmeasured confounding effects.
-The introduction of unknown covariance matrices $\mPhi_k$ dramatically increases the number of parameters in the problem.
+However, the introduction of unknown covariance matrices $\mPhi_k$ dramatically increases the number of parameters in the problem.
 Suitable regularizatiosn such as graphical lasso or specially-structured covariance assumptions should be considered.
@@
 Although we have presented the supervised tensor decomposition in the context of order-3 data tensors, the framework applies to more general multi-way datasets.
-One possible extension is the integrative analysis of omics data, in which multiple types of omics measurements (gene expression, DNA methylation, microRNA) are collected in the same set of individuals~\citep{lock2013joint,wang2019three}.
+One possible extension is the integrative analysis of omics data, in which multiple types of omics measurements (gene expression, DNA methylation, microRNA) are collected in the same set of individuals~\citep{lock2013joint}.
 In such a case, the observation is a stack of data matrices, which may not be of the same dimension.
-One possible extension is to impose linked dimension reduction among different data matrices, and incorporate side information as features in the joint decomposition.
+%One possible extension is to impose linked dimension reduction among different data matrices, and incorporate side information as features in the joint decomposition. 
 Other applications include time-series tensor data with multiple side information.
@@
 % for blind version
-\section*{Acknowledgements}
-This research was supported by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
+%\section*{Acknowledgements}
+%This research was supported by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
 
@@
 \begin{center}
-{\large\bf SUPPLEMENTARY MATERIALS}
+{\large \bf SUPPLEMENTARY MATERIALS}
 \end{center}
@@
 \begin{description}
-
-\item[Supplementary notes:]  technical proofs and additional simulation results.
-
-\item[Data and software:] R-package \texttt{tensorregress} performs our supervised tensor decomposition method.
-The source code and data used in the paper are available at \url{https://CRAN.R-project.org/package=tensorregress.} We also provide R scripts for reproducing all numbers in Figures 2-8 and Tables 2-3 in the paper.
-
+%\item[Supplementary notes:]  technical proofs and additional simulation results.
+\item[Data and software:] %R-package \texttt{tensorregress} performs our supervised tensor decomposition method. 
+The simulation code, R-package \texttt{tensorregress}, and datasets used in the paper are available at \url{https://CRAN.R-project.org/package=tensorregress} 
+%We also provide R scripts for reproducing all numbers in Figures 2-8 and Tables 2-3 in the paper.
 \end{description}

### 2020-09-17 tensor_regress/JASA_submission/paper_source/supp.tex
update

@@
 
-\usepackage{sectsty}
-\sectionfont{\fontsize{12}{10}\selectfont}
-
 \begin{document}
@@
 \def\spacingset#1{\renewcommand{\baselinestretch}%
-{#1}\small\normalsize} \spacingset{1}
+{#1}\small\normalsize} \spacingset{1.5}
 
@@
 \begin{center}
-{\large\bf Supplementary Notes to ``Supervised Tensor Decomposition \\
+{\Large\bf Supplementary Notes to ``Supervised Tensor Decomposition
 with Interactive Side Information''}
@@
 \section{Proofs}\label{sec:appedix}
+We restate the Theorem 4.1 from the main text. 
+ \begin{thm}[Statistical convergence]\label{thm:main}
+Consider a data tensor generated from model~\eqref{eq:decomp}, where the entries are conditionally independent realizations from an exponential family.
+Let $(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)$ be the M-estimator in~\eqref{eq:MLE} and $\hat \tB=\hat \tC\times \hat \mM_1\times\cdots \times \hat \mM_K$.
+Define $r_\textup{total}=\prod_k r_k$ and $r_{\max}=\max_k r_k$.
+Under Assumptions A1 and A2 with scaled feature matrices $\check \mX_k= \sqrt{d_k}\mX_k$, or under Assumptions A1' and A2 with original feature matrices, there exist two positive constants $C_1=C_1(\alpha,K), C_2=C_2(\alpha, K)>0$ independent of dimensions $\{d_k\}$ and $\{p_k\}$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$,
+\begin{equation}\label{eq:bound}
+    \FnormSize{}{\trueB- \hat \tB}^2\leq \frac{ C_{2} r_{\textup{total}}}{r_{\max}}\frac{\sum_k p_k}{\prod_k d_k}.
+\end{equation}
+Furthermore, if the unfolded core tensor has non-degenerate singular values at mode $k\in[K]$, i.e., $\sigma_{\min}(\textup{Unfold}_k(\tC_\textup{true})) \geq c>0$ for some constant $c$, then
+\begin{equation}\label{eq:sinebound}
+\textup{sin}^2 \Theta(\trueM,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\sigma^2_{\min}(\textup{Unfold}_k(\tC_\textup{true}))}\frac{\sum_k p_k}{\prod_k d_k}.
+\end{equation}
+\end{thm}
+
 \begin{proof}[Proof of Theorem~\ref{thm:main}]
@@
 \end{equation}
-Now we have proved both C1 and C2. Note that $\tL_{\tY}(\hat \tB)- \tL_{\tY}(\trueB)\geq 0$ by the definition of $\hat \tB$, This implies that
-\begin{align}
+Now we have proved both C1 and C2. Note that $\tL_{\tY}(\hat \tB)- \tL_{\tY}(\trueB)\geq 0$ by the definition of $\hat \tB$. This implies that
+\begin{align}\label{eq:F-norm}
 0&\leq \tL_{\tY}(\hat \tB)- \tL_{\tY}(\trueB) \\
@@
 \end{align}
-where the second line follows from~\eqref{upper}. Therefore,
+where the second line follows from~\eqref{upper}.
+
+The inequality~\eqref{eq:F-norm} can be rewritten as
 \begin{align}\label{eq:1}
@@
 
-Now we prove \eqref{eq:sinebound}.
+Now we prove bound for sin$\Theta$ distance. 
 We unfold tensors $\trueB$ and $\hat\tB$ along the mode $k$ and obtain $\text{Unfold}_k(\trueB)$ and $\text{Unfold}_k(\hat\tB)$. Notice that 
@@
 Applying Proposition \ref{prop:sinebound}, we have
-\begin{align}
-    \text{sin}\Theta^2(\mM_k,\hat\mM_k)\leq \frac{\FnormSize{}{\text{Unfold}_k(\hat\tB)-\text{Unfold}_k(\trueB)}}{\sigma_{\text{min}}(\text{Unfold}_k(\trueB))}= \frac{\FnormSize{}{\hat\tB-\trueB}}{\sigma_{\text{min}}(\text{Unfold}_k(\trueC))},
+\begin{align}\label{eq:sintheta}
+    \text{sin}^2\Theta(\mM_k,\hat\mM_k)\leq \frac{\FnormSize{}{\text{Unfold}_k(\hat\tB)-\text{Unfold}_k(\trueB)}}{\sigma_{\text{min}}(\text{Unfold}_k(\trueB))}= \frac{\FnormSize{}{\hat\tB-\trueB}}{\sigma_{\text{min}}(\text{Unfold}_k(\trueC))},
 \end{align} where $\sigma_{\text{min}}(\text{Unfold}_k(\trueB)) = \sigma_{\text{min}}(\text{Unfold}_k(\trueC))$ holds based on the orthonormality of factor matrices.
-We finally prove \eqref{eq:sinebound} combining \eqref{eq:bound}.
+We finally prove the sin$\Theta$ distance by combining~\eqref{eq:sintheta} and \eqref{eq:bound}.
 \end{proof}
@@
 \begin{align}\label{eq:sine}
-    \text{sin}\Theta(\mU,\hat\mU) \leq \frac{C\FnormSize{}{\hat\mB-\mB}}{\sigma_{\text{min}}(\mB)},
+    \text{sin} \Theta(\mU,\hat\mU) \leq \frac{\FnormSize{}{\hat\mB-\mB}}{\sigma_{\text{min}}(\mB)}.
 \end{align}
-for some constant $C>0$.
+
 \end{prop}
@@
 \end{align}
-Therefore,we prove \eqref{eq:sine}.
+Therefore, we prove \eqref{eq:sine}.
 \end{proof}
@@
 
-\newpage
+\section{Computational complexity}
+The computational complexity of our Algorithm~\eqref{alg:B} is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the data tensor.
+More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor.
+The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$ feature matrix.
+Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$.
+This complexity in tensor dimension matches with the classical tensor decomposition~\citep{kolda2009tensor}.
+
 \section{Additional simulation results}
@@
 
-\begin{figure}[!h]
+\begin{figure}[H]
 \centering
-\includegraphics[width=16cm]{Supp_Figure1.pdf} 
+\includegraphics[width=15cm]{Supp_Figure1.pdf} 
 \caption{Comparison of MSPE versus the number of modes with features.
@@
 
-\begin{figure}[!h]
+\begin{figure}[H]
 \centering
-\includegraphics[width=16cm]{Supp_Figure2.pdf} 
+\includegraphics[width=15cm]{Supp_Figure2.pdf} 
 \caption{Comparison of MSPE versus effective sample size. We consider full combinations rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and signal $\alpha=3$ (low), $\alpha=6$ (high). }~\label{fig:S2}
@@
 
+\newpage
 \bibliography{tensor_wang}

### 2020-09-17 tensor_regress/JASA_submission/code/Figure7.R

@@ -29,15 +29,15 @@ mass[,,5]=-massive[,,4]-massive[,,3]
 
 
 ### Plot histogram to compare two methods###
-pdf("compare.pdf",width=15,height=6)
+pdf("compare.pdf",width=15,height=5.5)
 par(mfrow=c(2,4))
-hist(c(mass[,,1]),nclass=40,xlab="Effect size",main="Intercept (Classical GLM)")
-hist(c(mass[,,2]),nclass=40,xlab="Effect size",main="Gender (Classical GLM)")
-hist(c(mass[,,4]),nclass=40,xlab="Effect size",main="Age 26-30 (Classical GLM)")
-hist(c(mass[,,5]),nclass=40,xlab="Effect size",main="Age 31+ (Classical GLM)")
+hist(c(mass[,,1]),nclass=40,xlab="Effect size",main="Intercept (multi-response GLM)")
+hist(c(mass[,,2]),nclass=40,xlab="Effect size",main="Gender (multi-response GLM)")
+hist(c(mass[,,4]),nclass=40,xlab="Effect size",main="Age 26-30 (multi-response GLM)")
+hist(c(mass[,,5]),nclass=40,xlab="Effect size",main="Age 31+ (multi-response GLM)")
 
-hist(c(coef[,,1]),nclass=40,xlab="Effect size",main="Intercept (Tensor regression)")
-hist(c(coef[,,2]),nclass=40,xlab="Effect size",main="Gender (Tensor regression)")
-hist(c(coef[,,4]),nclass=40,xlab="Effect size",main="Age 26-30 (Tensor regression)")
-hist(c(coef[,,5]),xlab="Effect size",nclass=40,main="Age 31+ (Tensor regression)")
+hist(c(coef[,,1]),nclass=40,xlab="Effect size",main="Intercept (STD)")
+hist(c(coef[,,2]),nclass=40,xlab="Effect size",main="Gender (STD)")
+hist(c(coef[,,4]),nclass=40,xlab="Effect size",main="Age 26-30 (STD)")
+hist(c(coef[,,5]),xlab="Effect size",nclass=40,main="Age 31+ (STD)")
 dev.off()

### 2020-09-17 tensor_regress/JASA_submission/paper_source/regress.tex
page limits

@@
 Examples include the political interaction dataset \citep{nickel2011three} which measures action counts between countries under various relations, and the brain connectivity network dataset \citep{zhang2018mapping} which is a collection of binary adjacency matrices.
-Classical tensor decomposition methods are based on minimizing the Frobenious norm of deviation, leading to suboptimal predictions for binary- or count-valued response variables.
+Classical tensor decomposition methods are based on minimizing the Frobenius norm of deviation, leading to suboptimal predictions for binary- or count-valued response variables.
 A number of supervised tensor methods have been proposed \citep{narita2012tensor, zhao2012higher, yu2016learning,lock2018supervised} to address the tensor regression problem in various forms, such as scalar-to-tensor regression and tensor-response regression.
@@
 \begin{align}\label{eq:decomp}
-\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1,\ \ldots,\ \mX_K\mM_K\}\right),\notag\\
+\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1, \ldots, \mX_K\mM_K\}\right),\notag\\
 \text{with} \ &\ \mM_k^T\mM_k = \mI_{r_k},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\quad \text{for all }k=1,\ldots,K,
@@
 \begin{example}[Network population model]\label{example:brain}
-Network response model~\citep{rabusseau2016low} is recently developed for neuroimanig analysis.
+Network response model~\citep{rabusseau2016low} is recently developed for neuroimaging analysis.
 The goal is to study the relationship between brain network connectivity pattern and features of individuals.
@@
 The low-rankness in $\mB$ has demonstrated its success in modeling transitivity, balance, and communities in  networks~\citep{hoff2005bilinear}.
-We show that our supervised tensor decompostiion~\eqref{eq:decomp} also incorporates the graph model as a special case.
+We show that our supervised tensor decompostion~\eqref{eq:decomp} also incorporates the graph model as a special case.
 Let $\tY=\entry{y_{ij}}$ be a binary matrix where $y_{ij}=\mathds{1}_{(i,j)\in E}$.
@@
 \caption{Estimated feature effects in the \emph{Nations} data analysis.
-Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warnning}, \emph{violentactions}, \emph{treaties}, and \emph{aidenemy}, respectively.
+Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warning}, \emph{violentactions}, \emph{treaties}, and \emph{aidenemy}, respectively.
 }\label{fig:est}
@@
 However, the introduction of unknown covariance matrices $\mPhi_k$ dramatically increases the number of parameters in the problem.
-Suitable regularizatiosn such as graphical lasso or specially-structured covariance assumptions should be considered.
+Suitable regularization such as graphical lasso or specially-structured covariance assumptions should be considered.
 The extension of tensor modeling with heterogeneous mixed-effects will be an important future direction.
@@
 \item[Data and software:] %R-package \texttt{tensorregress} performs our supervised tensor decomposition method. 
-The simulation code, R-package \texttt{tensorregress}, and datasets used in the paper are available at \url{https://CRAN.R-project.org/package=tensorregress} 
+Our simulation code, R-package \texttt{tensorregress}, and datasets used in the paper are available at \url{https://CRAN.R-project.org/package=tensorregress} 
 %We also provide R scripts for reproducing all numbers in Figures 2-8 and Tables 2-3 in the paper.

### 2020-09-22 tensor_regress/JCGS_submission/jcgs-tensor/jcgs-tensor.tex
jcgs
update

--- 2020-09-17 tensor_regress/JASA_submission/paper_source/regress.tex
@@
 
-\newcommand{\blind}{1}
+%\usepackage{enumerate}
+\usepackage{natbib}
+\usepackage{url} % not crucial - just used below for the URL
 
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
+
+% DON'T change margins - should be 1 inch all around.
 \addtolength{\oddsidemargin}{-.5in}%
@@
 \addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{-.3in}%
+\addtolength{\textheight}{1.3in}%
 \addtolength{\topmargin}{-.8in}%
@@
 \newtheorem{lem}{Lemma}
-\newtheorem{prop}{Proposition}
+\newtheorem{prop}{Proposition}[section]
 \newtheorem{pro}{Property}
@@
 
+%\bibliographystyle{natbib}
+
 \def\spacingset#1{\renewcommand{\baselinestretch}%
@@
 
-\if1\blind
+\if0\blind
 {
   \title{\bf Supervised Tensor Decomposition with Interactive Side Information}
-  % blind version
-  \author{Jiaxin Hu, Zhuoyan Xu, Chanwoo Lee, and Miaoyan Wang\\
+  \author{Jiaxin Hu, Zhuoyan Xu, Chanwoo Lee, and Miaoyan Wang\thanks{The authors gratefully acknowledge NSF grant DMS-1915978 and funding from the Wisconsin Alumni Research Foundation.
+}\hspace{.2cm}\\
     Department of Statistics, University of Wisconsin-Madison}
   \maketitle
+
 } \fi
 
-\if0\blind
+\if1\blind
 {
@@
 \noindent%
-{\it Keywords:} Applications and case studies, Tensor data analysis, Supervised dimension reduction, Exponential family distribution, Generalized multilinear model
+{\it Keywords:}  Tensor data analysis, Supervised dimension reduction, Exponential family distribution, Generalized multilinear model, Alternating optimization
 \vfill
@@
 In addition to the aforementioned challenges, many tensor datasets consist of non-Gaussian measurements.
-Examples include the political interaction dataset \citep{nickel2011three} which measures action counts between countries under various relations, and the brain connectivity network dataset \citep{zhang2018mapping} which is a collection of binary adjacency matrices.
+Examples include the political interaction dataset \citep{nickel2011three,hu2015scalable} which measures action counts between countries under various relations, and the brain connectivity network dataset \citep{zhang2018mapping,wang2018learning,wang2019common} which is a collection of binary adjacency matrices.
 Classical tensor decomposition methods are based on minimizing the Frobenius norm of deviation, leading to suboptimal predictions for binary- or count-valued response variables.
-A number of supervised tensor methods have been proposed \citep{narita2012tensor, zhao2012higher, yu2016learning,lock2018supervised} to address the tensor regression problem in various forms, such as scalar-to-tensor regression and tensor-response regression.
+A number of supervised tensor methods have been proposed \citep{narita2012tensor, lock2018supervised,rai2014scalable} to address the tensor regression problem in various forms, such as scalar-to-tensor regression and tensor-response regression.
 These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest.
@@
 \includegraphics[width=16cm]{demo.pdf}
-\vspace{-.5cm}
 \end{center}
 \caption{Examples of supervised tensor decomposition with interactive side information. (a) Network population model. (b) Spatio-temporal growth model.} \label{fig:intro1}
-\vspace{-.3cm}
 \end{figure*}
@@
 Our work is closely related to but also clearly distinctive from several lines of previous work.
-The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor, hong2020generalized} that aims to find the best low-rank representation of a data tensor.
+The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor, hong2020generalized,wang2017tensor,bi2018recom,chi2012tensors} that aims to find the best low-rank representation of a data tensor.
 In contrast, our model is a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and multiple features.
@@
 As we show in Section~\ref{sec:connection}, the supervised tensor decomposition has an interesting connection with tensor-on-tensor regression.
-Previous methods~\citep{lock2018tensor} mainly focus on Gaussian tensors.
+Previous methods~\citep{lock2018tensor,lock2018supervised} mainly focus on Gaussian tensors.
 The Frobenius norm used in the objective function is statistical suboptimal for general exponential family tensors.
-Maximum likelihood estimator (MLE) is studied in~\citep{raskutti2019convex} and a convex relaxation algorithm is proposed to solve for low-rank tensor coefficients.
+Maximum likelihood estimator (MLE) is studied in~\cite{raskutti2019convex} and a convex relaxation algorithm is proposed to solve for low-rank tensor coefficients.
 However, in the tensor case, convex MLE suffers from both computational intractability and statistical suboptimality.
@@
 
-The third line of work uses side information for various tensor learning tasks, such as for completion~\citep{narita2012tensor} and for recommendation system~\citep{ioannidis2019coupled,farias2019learning}.
-These methods also study tensors with side information, but they take regularization approaches to penalize predictions that are distant from side information~\cite{lock2018supervised}.
+The third line of work uses side information for various tensor learning tasks, such as for completion~\citep{narita2012tensor,song2019tensor,cao2016semi} and for recommendation system~\citep{ioannidis2019coupled,farias2019learning}.
+These methods also study tensors with side information, but they take regularization approaches to penalize predictions that are distant from side information~\citep{cao2016semi,song2019tensor}.
 One important difference is that their goal is prediction but not parameter estimation.
@@
 
-The paper is organized as follows.
+The remainder of the paper is organized as follows.
 Section~\ref{sec:pre} introduces tensor preliminaries.
 Section~\ref{sec:model} presents the main model and three motivating examples for supervised tensor decomposition.
-We describe the quasi-likelihood estimation in Section~\ref{sec:est} and provide numerical experiments in Section~\ref{sec:simulation}.
-In Section~\ref{sec:data}, we present two data applications, one on diffusion tensor imaging data and the other on multi-relational network data.
-%and applications to diffusion tensor imaging data and multi-relational social network data in Section~\ref{sec:data}. 
-%In Section~\ref{sec:simulation}, we present numerical experiments and assess the performance in comparison to alternative methods.
+We describe the quasi-likelihood estimation and alternating optimization algorithm in Section~\ref{sec:est}.
+In Section~\ref{sec:simulation}, we present numerical experiments and assess the performance in comparison to alternative methods.
 In Section~\ref{sec:data}, we apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
 We conclude in Section~\ref{sec:con} with discussions about our findings and avenues of future work.
+All proofs are deferred to Supplementary Materials.
 
@@
 We use lower-case letters (e.g.,\ $a,b,c$) for scalars and vectors, upper-case boldface letters (e.g.,\ $\mA, \mB, \mC$) for matrices, and calligraphy letters (e.g.,\ $\tA, \tB, \tC$) for tensors of order three or greater.
-Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
-The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as $\tY \times_1\mX_1\times \cdots \times_K \mX_K=\entry{\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}x_{j_1,i_1}^{(1)}\cdots x_{j_K,i_K}^{(K)}}$,
+Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor, where $K$ is the number of modes and also called the order.
+The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as
+\[
+\tY \times_1\mX_1\times \cdots \times_K \mX_K=\entry{\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}x_{j_1,i_1}^{(1)}\cdots x_{j_K,i_K}^{(K)}},
+\]
 which results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor.
 For ease of presentation, we use the shorthand $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product.
-For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as $\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.$
-The tensor Frobenius norm is defined as $\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}$, and the maximum norm is defined $\mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.$
+For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as
+\[
+\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.
+\]
+The tensor Frobenius norm and maximum norm are defined as
+\[
+\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2},\quad \text{and}\quad \mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.
+\]
 When $a$ is a vector, we use $\vnormSize{}{a}=\langle a,a \rangle^{1/2}$ to denote the vector 2-norm.
+
 A higher-order tensor can be reshaped into a lower-order object.
@@
 We let $\mI_d$ denote the $d \times d$ identity matrix and $[d]$ denote the $d$-set $\{1,\ldots,d\}$.
+We use $\mathbb{O}_{d,r}$ to denote the collection of all $d$-by-$r$ matrices with orthogonal columns; i.e., $\mathbb{O}_{d,r}=\{\mP\in\mathbb{R}^{d\times r}\colon \mP^T\mP = \mathbf{1}_r\}$.
 For ease of notation, we allow the basic arithmetic operators (e.g.,\ $+, -, \geq $) and univariate functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner.
@@
 \begin{example}[Spatio-temporal growth model]
-The growth curve model~\citep{gabriel1998generalised} was originally proposed as an example of bilinear model for matrix data, and we adopt its higher-order extension here.
+The growth curve model~\citep{gabriel1998generalised,srivastava2008models} was originally proposed as an example of bilinear model for matrix data, and we adopt its higher-order extension here.
 Let $\tY=\entry{y_{ijk}}\in\mathbb{R}^{d \times m\times n}$ denote the pH measurements of $d$ lakes at $m$ levels of depth and for $n$ time points.
@@
 The role of the sufficient features $\mX_k\mM_k$ can be seen in the following two conditional independence assumptions in model~\eqref{eq:decomp},
-
 \begin{align}
@@
 
-Our model also has a close connection with tensor-on-tensor regression~\citep{raskutti2019convex, lock2018tensor,gahrooei2020multiple}.
+Our model also has a close connection with tensor-on-tensor regression~\citep{raskutti2019convex, lock2018tensor,gahrooei2020multiple,hao2020sparse}.
 Specifically, model~\eqref{eq:decomp} can be viewed as a multivariate regression model, where the response is the vectorized tensor and the covariates are interactions between sufficient features across modes.
@@
 \end{tabular}
-\vspace{-.2cm}
 \caption{Canonical links for common distributions.}\label{table:link}
@@
 Ignoring constants that do not depend on $\Theta$, the quasi log-likelihood of~\eqref{eq:decomp} is equal to Bregman distance between $\tY$ and $b'(\Theta)$:
-\begin{equation}\label{eq:loglikelihood}
-\small
-\tL_{\tY}(\tC,\mM_1,\ldots,\mM_K)=\langle \tY, \Theta \rangle - \sum\nolimits_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K})\ \text{with}\ \Theta=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
-\end{equation}
+\begin{align}\label{eq:loglikelihood}
+\tL_{\tY}(\tC,\mM_1,\ldots,\mM_K)&=\langle \tY, \Theta \rangle - \sum_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K}),\\
+ \text{where}\quad \Theta&=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
+\end{align}
 We propose a constrained maximum quasi-likelihood estimator (M-estimator),
@@
 \[
-\tP=\left\{(\tC, \mM_1,\ldots,\mM_K) \ \Big| \ \mM_k^T\mM_k=\mathbf{I}_{r_k},\ \mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha \right\}.
+\tP=\left\{(\tC, \mM_1,\ldots,\mM_K) \ \Big| \ \mM_k^T\mM_k=\mathbf{I}_{r_k}\ \text{for all }k\in[K],\  \mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha \right\}.
 \]
@@
 \State Obtain the factor matrix $\tilde \mM_k\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$.
-\State Perform QR factorization $\tilde \mM_k=\mQ\mR$.
+\State Perform QR factorization $\tilde \mM_k=\mQ\mR$, where $\mQ\in\mathbb{R}^{p_k\times r_k}$ consists of orthonormal columns.
 \State Update $\mM_k\leftarrow \mQ$ and core tensor $\tC\leftarrow \tC\times_k \mR$.
@@
 The optimization~\eqref{eq:MLE} is a non-convex problem due to the non-convexity in the feasible set $\tP$.
-Under mild conditions, our algorithm enjoys global convergence; i.e.\ any sequence of iterates generated by the alternating algorithm converges to a stationary point of $\tL_\tY(\cdot)$ module orthogonal transformation (see Supplementary Notes, Section~\ref{sec:SAlgorithm}).
+Under mild conditions, our algorithm enjoys global convergence; i.e.\ any sequence of iterates generated by the alternating algorithm converges to a stationary point of $\tL_\tY(\cdot)$ module orthogonal transformation.
+To establish the convergence properties of Algorithm~\ref{alg:B}, we first introduce the equivalent relationship induced by orthogonal transformation.
+
+\begin{defn}[Equivalence class]Let $\tA = (\tC, \mM_1,\ldots,\mM_K) \in \mathbb{R}^{d_{\text{total}}}$ denote the collection of decision variables in the alternating optimization, where $d_{\text{total}} = \prod_k r_k + \sum_k r_kd_k$.
+Two parameters $\tA'=(\tC', \mM'_1,\ldots,\mM'_k)$ and $\tA=(\tC, \mM_1,\ldots,\mM_k)$ are called equivalent, denoted $\tA\sim \tA'$, if and only if there exist orthogonal matrices $\mP_k \in \mathbb{O}_{d_k,r_k}$ such that
+\begin{align}\label{eq:equivalance}
+	\mM'_k \mP^T_k = \mM_k \quad \text{for all }k \in [K], \quad \text{and}\quad \tC' \times_1 \mP_1 \times_2  \dots \times_K \mP_K = \tC.
+\end{align}
+Equivalently, two parameters $\tA,\ \tA'$ are equivalent if the corresponding Tucker tensors~\eqref{eq:factor} are the same, $\tB(\tA)=\tB'(\tA')$. 
+\end{defn}
+
+\begin{prop}[Global convergence]\label{prop:alg}
+Assume the set $\{\tA\ |\ \tL_\tY(\tA) \geq \tL_\tY(\tA^{(0)}) \}$ is compact and the stationary points of $\tL_\tY(\tA)$ are isolated module the equivalence defined in~\eqref{eq:equivalance}.
+Furthermore, assume that $\alpha = \infty$; i.e., we impose no entrywise bound constrains on the parameter space.
+Then any sequence $\tA^{(t)}$ generated by Algorithm~\ref{alg:B} converges to a stationary point of $\tL_\tY(\tA)$ module equivalence class.
+\end{prop}
+
 In theory, global optimality of non-convex optimization is often challenging to obtain.
@@
 Under all combinations of the dimension $d$, rank $r$, and type of the entries, Algorithm~\ref{alg:B} converges quickly in a few iterations, and the objective values at convergent points are close to or larger than the value at true parameters.
-The time complexity is provided in Supplementary Notes.
 
@@
 \centering
-\includegraphics[width=15cm]{loglike.pdf}
-\vspace{-.25cm}
+\includegraphics[width=16cm]{loglike.pdf}
 \caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c) Poisson models.
@@
 }\label{fig:loglike}
-\vspace{-.3cm}
 \end{figure}
 
-\subsection{Rank selection}\label{sec:tuning}
+\subsection{Rank selection and computational complexity}\label{sec:tuning}
 Algorithm~\ref{alg:B} assumes the rank $\mr$ is given.
@@
 
-%The computational complexity of our Algorithm~\eqref{alg:B} is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the data tensor.
+The computational complexity of our Algorithm~\ref{alg:B} is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the data tensor.
 More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor.
@@
 Therefore, we use angle distance to assess the accuracy in estimating the column space, $\textup{Span}(\mM_k)$.
-For any two matrices $\mA,\mB$ with orthonormal columns of same dimension, the angle distance is defined as
+For any two column-orthonormal matrices $\mA,\mB\in\mathbb{O}(d,r)$ of same dimension, the angle distance is defined as
 \[
-\sin \Theta(\mA,\mB)=\snormSize{}{\mA^T \mB^{\perp}}=\max\left\{ \frac{\langle \mx, \my\rangle}{\vnormSize{}{\mx}\vnormSize{}{\my}}\colon \ \mx\in \textup{Span}(\mA),\ \my\in \textup{Span}(\mB)\right\}.
+\sin \Theta(\mA,\mB)=\snormSize{}{\mA^T \mB^{\perp}}=\max\left\{ \frac{\langle \mx, \my\rangle}{\vnormSize{}{\mx}\vnormSize{}{\my}}\colon \ \mx\in \textup{Span}(\mA),\ \my\in \textup{Span}(\mB^{\perp})\right\}.
 \]
@@
 
-We conclude this section by providing the accuracy measured in KL divergence.
-%, for the response distribution.
+We conclude this section by providing the accuracy, measured in KL divergence, for the response distribution.
 
@@
 \begin{tabular}{c|cc|cc}
-%True Rank $\mr$ &\multicolumn{2}{c|}{Dimension (Gaussian tensors)}&\multicolumn{2}{c|}{Dimension (Poisson tensors)}\\
-True rank $\mr$& $d=20$  (Gaussian) &$d=40$ (Gaussian) &$d=20$ (Poisson) &$d=40$ (Poisson)\\
+True Rank $\mr$ &\multicolumn{2}{c|}{Dimension (Gaussian tensors)}&\multicolumn{2}{c|}{Dimension (Poisson tensors)}\\
+%True rank $\mr$& $d=20$  (Gaussian) &$d=40$ (Gaussian) &$d=20$ (Poisson) &$d=40$ (Poisson)\\
 \hline
@@
 \caption{Rank selection via BIC. The estimated ranks are averaged across 30 simulation. Bold number indicates the ground truth is within two standard deviations of the estimate.}\label{tab:rank}
-\vspace{-.1cm}
 \end{table}
@@
 The dashed curves correspond to $\tO({1/d^2})$.}\label{fig:dim}
-\vspace{-.1cm}
 \end{figure}
@@
 }\label{fig:glm}
-\vspace{-.2cm}
 \end{figure}
@@
 Panels (a) and (b) plot MSPE versus the number of modes with available features.
-Panels (c) and (d) plot MSPE versus the effective sample size $d^2$.}~\label{fig:compare}
-%We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and signal $\alpha =3 $ (low), $\alpha=6$ (high).}~\label{fig:compare}
-\vspace{-.4cm}
+Panels (c) and (d) plot MSPE versus the effective sample size $d^2$.
+We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and signal $\alpha =3 $ (low), $\alpha=6$ (high).}~\label{fig:compare}
 \end{figure}
@@
 \caption{Comparison of estimated feature effects the HCP data using (a) multi-response GLM and (b) supervised tensor decomposition (STD). }\label{fig:s1}
-\vspace{-.1cm}
 \end{figure}
@@
 The goal is to identify the variation in connections due to country-level attributes and their interactions.
-%One of the key features is that the 56 relations are correlated, and we would like to take it into account in assessing the feature effects.
+One of the key features is that the 56 relations are correlated, and we would like to take it into account in assessing the feature effects.
 
@@
 }\label{fig:est}
-\vspace{-.4cm}
 \end{figure}
@@
 \caption{$K$-means clustering of relations based on dimension reduction on the 3$^\text{rd}$ mode. }\label{tab:s1}
-\vspace{-.4cm}
 \end{table}
@@
 
-%We comment on a number of possible extensions. 
+There are several possible extensions from the work.
 We have provided accuracy guarantees for parameter estimation in the supervised tensor model.
@@
 One possible approach would be performing parametric bootstrap~\citep{tibshirani1993introduction} to assess the uncertainty in the estimation.
-%For example, one can simulate tensors from the fitted low-rank model based on the estimates, and then assess the empirical distribution of the estimates. 
+For example, one can simulate tensors from the fitted low-rank model based on the estimates, and then assess the empirical distribution of the estimates. 
 While being simple, bootstrap approach is often computationally expensive for large-scale data.
@@
 \begin{align}\label{eq:extention}
-\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)&=\tC\times\mM_1\times \cdots \times \mM_K\ \ \text{and}\ \ \text{Var}(\tY|\mX_1,\ldots,\mX_K)=\mPhi_1\otimes \cdots \otimes \mPhi_K,
+\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)&=\tC\times\mM_1\times \cdots \times \mM_K,\\
+ \text{Var}(\tY|\mX_1,\ldots,\mX_K)&=\mPhi_1\otimes \cdots \otimes \mPhi_K,
 \end{align}
@@
 Although we have presented the supervised tensor decomposition in the context of order-3 data tensors, the framework applies to more general multi-way datasets.
-One possible extension is the integrative analysis of omics data, in which multiple types of omics measurements (gene expression, DNA methylation, microRNA) are collected in the same set of individuals~\citep{lock2013joint}.
+One possible extension is the integrative analysis of omics data, in which multiple types of omics measurements (gene expression, DNA methylation, microRNA) are collected in the same set of individuals~\citep{lock2013joint,wang2019three}.
 In such a case, the observation is a stack of data matrices, which may not be of the same dimension.
@@
 Exploiting the benefits and properties of the supervised tensor decomposition in specialized task will boost scientific discoveries.
-
-% for blind version
-%\section*{Acknowledgements}
-%This research was supported by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
 
@@
 \begin{description}
-%\item[Supplementary notes:]  technical proofs and additional simulation results.
+\item[Supplementary notes:]  technical proofs and additional simulation results.
 \item[Data and software:] %R-package \texttt{tensorregress} performs our supervised tensor decomposition method. 
 Our simulation code, R-package \texttt{tensorregress}, and datasets used in the paper are available at \url{https://CRAN.R-project.org/package=tensorregress} 
-%We also provide R scripts for reproducing all numbers in Figures 2-8 and Tables 2-3 in the paper.
+We also provide R scripts for reproducing all numbers in Figures 2-8 and Tables 2-3 in the paper.
 \end{description}

### 2020-09-22 tensor_regress/JCGS_submission/jcgs-tensor/supp.tex
jcgs
update

--- 2020-09-17 tensor_regress/JASA_submission/paper_source/supp.tex
@@
 
+% DON'T change margins - should be 1 inch all around.
 \addtolength{\oddsidemargin}{-.5in}%
@@
 \usepackage{xr}
-\externaldocument{regress}
+\externaldocument{jcgs-tensor}
 
@@
 \section{Proofs}\label{sec:appedix}
+
+\subsection{Proofs of Proposition~\ref{prop:alg}}\label{sec:SAlgorithm}
+For notational convenience, we drop the subscript $\tY$ from the objective $\tL_\tY(\cdot)$ and simply write as $\tL(\cdot)$.
+Let $\tA = (\tC, \mM_1,\ldots,\mM_K) \in \mathbb{R}^{d_{\text{total}}}$ denote the collection of decision variables in the alternating optimization, where $d_{\text{total}} = \prod_k r_k + \sum_k r_kd_k$.
+We recall the equivalent relationship induced by orthogonal transformation.
+
+\begin{defn}[Equivalence class]
+Two parameters $\tA'=(\tC', \mM'_1,\ldots,\mM'_k)$ and $\tA=(\tC, \mM_1,\ldots,\mM_k)$ are called equivalent, denoted $\tA\sim \tA'$, if and only if there exist a set of orthogonal matrices $\mP_k \in \mathbb{O}_{d_k,r_k}$ such that
+\begin{align}\label{eq:equivalance}
+	\mM'_k \mP^T_k = \mM_k ,\ \forall k \in [K], \quad \text{and}\quad \tC' \times_1 \mP_1 \times_2  \dots \times_K \mP_K = \tC.
+\end{align}
+\end{defn}
+
+\begin{prop}[Global convergence]
+Assume the set $\{\tA\ |\ \tL(\tA) \geq \tL(\tA^{(0)}) \}$ is compact and the stationary points of $\tL(\tA)$ are isolated module the equivalence defined in~\eqref{eq:equivalance}.
+Furthermore, assume that $\alpha = \infty$; i.e., we impose no entrywise bound constrains on the parameter space.
+Then any sequence $\tA^{(t)}$ generated by alternating algorithm converges to a stationary point of $\tL(\tA)$ module equivalence class.
+\end{prop}
+
+\begin{proof}
+Pick an arbitrary iterate $\tA^{(t)}$.
+Because of the compactness of set $\{\tA\colon  \tL(\tA) \geq \tL(\tA^{(0)}) \}$ and the boundedness of the decision domain, there exist a sub-sequence of $\tA^{(t)}$ that converges.
+Let $\tA^*$ denote one of the limiting points of $\tA^{(t)}$.
+Let $\tS = \{\tA^*\}$ denote the set of all the limiting points of $\tA^{(t)}$.
+We have $\tS \subset \{\tA\colon  \tL(\tA) \geq \tL(\tA^{(0)}) \}$ and thus $\tS$ is a compact set.
+By~\citet[Propositions 8.2.1 and 13.4.2]{Lange:2012:NAS:2432073}, $\tS$ is also connected.
+Note that all points in $\tS$ are also stationary points of $\tL(\cdot)$, because of the monotonic increase of $\tL(\tA^{(t)})$ as $t \rightarrow \infty$.
+
+ Consider the equivalence of Tucker tensor representation of elements in $\tS$. We define an enlarged set $\tE_S$ induced by the equivalent class of elements in $\tS$,
+\begin{equation}
+ \tE_S = \left\{\tA\colon \tA\sim \tA^* \text{ for some }\tA^* \in \tS\right\}.
+\end{equation}
+The enlarged set $\tE_S$ satisfies the two properties below:
+ \begin{enumerate}
+\item [1.] [Union of stationary points] The set $\tE_S$ is an union of equivalent  classes generated by the limiting points in $\tS$.
+\item [2.] [Connectedness module the equivalence]  The set $\tE_S$ is connected module the equivalence relationship. That property is obtained by the connectedness of $S$.
+ \end{enumerate}
+Now, note that the isolation of stationary points and Property 1 imply that $\tE_S$ contains only finite number of equivalent classes.
+Otherwise, there is a subsequence of non-equivalent stationary points whose limit is not isolated, which contradicts the isolation assumption.
+Combining the finiteness with Property 2, we conclude that $\tE_S$ contains only a single equivalent class; i.e.
+$\tE_S = \tE_{\{\tA^*\}}$, where $\tA^*$ is a stationary point of $\tL(\tA)$.
+Therefore, all the convergent sub-sequences of $\tA^{(t)}$ converge to one stationary point $\tA^*$ up to equivalence.
+We conclude that, any iterate $\tA^{(t)}$ generated by Algorithm 1 converges to a stationary point of $\tL(\tA)$ up to equivalence.
+ \end{proof}
+
+\subsection{Proof of Theorem 4.1}
 We restate the Theorem 4.1 from the main text. 
  \begin{thm}[Statistical convergence]\label{thm:main}
-Consider a data tensor generated from model~\eqref{eq:decomp}, where the entries are conditionally independent realizations from an exponential family.
-Let $(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)$ be the M-estimator in~\eqref{eq:MLE} and $\hat \tB=\hat \tC\times \hat \mM_1\times\cdots \times \hat \mM_K$.
+Consider a data tensor generated from supervised tensor model, where the entries are conditionally independent realizations from an exponential family.
+Let $(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)$ be the M-estimator and $\hat \tB=\hat \tC\times \hat \mM_1\times\cdots \times \hat \mM_K$.
 Define $r_\textup{total}=\prod_k r_k$ and $r_{\max}=\max_k r_k$.
@@
 
-\begin{proof}[Proof of Theorem~\ref{thm:KL}]
+\subsection{Proof of Corollary~\ref{thm:KL}}
+\begin{proof}[Proof of Corollary~\ref{thm:KL}]
 The proof is similar to~\citet{pmlr-v108-berthet20a}.
@@
 \end{proof}
-
-\section{Algorithm properties}\label{sec:SAlgorithm}
-In this section, we provide the convergence properties of Algorithm~\ref{alg:B}.
-For notational convenience, we drop the subscript $\tY$ from the objective $\tL_\tY(\cdot)$ and simply write as $\tL(\cdot)$.
-Let $\tA = (\tC, \mM_1,\ldots,\mM_K) \in \mathbb{R}^{d_{\text{total}}}$ denote the collection of decision variables in the alternating optimization, where $d_{\text{total}} = \prod_k r_k + \sum_k r_kd_k$.
-We introduce the equivalent relationship induced by orthogonal transformation.
-Let $\mathbb{O}_{d,r}$ be the collection of all $d$-by-$r$ matrices with orthogonal columns, $\mathbb{O}_{d,r}:=\{\mP\in\mathbb{R}^{d\times r}\colon \mP^T\mP = \mathbf{1}_r\}$, where $\mathbf{1}_r$ is the $r$-by-$r$ identity matrix.
-
-\begin{defn}[Equivalance relation]\label{eq:equivalance}
-Two parameters $\tA'=(\tC', \mM'_1,\ldots,\mM'_k)$ and $\tA=(\tC, \mM_1,\ldots,\mM_k)$ are called equivalent, denoted $\tA\sim \tA'$, if and only if there exist a set of orthogonal matrices $\mP_k \in \mathbb{O}_{d_k,r_k}$ such that
-\begin{align}
-	\mM'_k \mP^T_k = \mM_k ,\ \forall k \in [K], \quad \text{and}\quad \tC' \times_1 \mP_1 \times_2  \dots \times_K \mP_K = \tC.
-\end{align}
-\end{defn}
-Equivalently, two parameters $\tA,\ \tA'$ are equivalent if the corresponding Tucker tensors are the same, $\tB(\tA)=\tB'(\tA')$.
-
-\begin{prop}[Global convergence]
-Assume the set $\{\tA\ |\ \tL(\tA) \geq \tL(\tA^{(0)}) \}$ is compact and the stationary points of $\tL(\tA)$ are isolated module the equivalence defined in~\eqref{eq:equivalance}.
-Furthermore, assume that $\alpha = \infty$; i.e., we impose no entrywise bound constrains on the parameter space.
-Then any sequence $\tA^{(t)}$ generated by alternating algorithm converges to a stationary point of $\tL(\tA)$ module equivalence class.
-\end{prop}
-
-\begin{proof}
-Pick an arbitrary iterate $\tA^{(t)}$.
-Because of the compactness of set $\{\tA\colon  \tL(\tA) \geq \tL(\tA^{(0)}) \}$ and the boundedness of the decision domain, there exist a sub-sequence of $\tA^{(t)}$ that converges.
-Let $\tA^*$ denote one of the limiting points of $\tA^{(t)}$.
-Let $\tS = \{\tA^*\}$ denote the set of all the limiting points of $\tA^{(t)}$.
-We have $\tS \subset \{\tA\colon  \tL(\tA) \geq \tL(\tA^{(0)}) \}$ and thus $\tS$ is a compact set.
-By~\citet[Propositions 8.2.1 and 13.4.2]{Lange:2012:NAS:2432073}, $\tS$ is also connected.
-Note that all points in $\tS$ are also stationary points of $\tL(\cdot)$, because of the monotonic increase of $\tL(\tA^{(t)})$ as $t \rightarrow \infty$.
-
- Consider the equivalence of Tucker tensor representation of elements in $\tS$. We define an enlarged set $\tE_S$ induced by the equivalent class of elements in $\tS$,
-\begin{equation}
- \tE_S = \left\{\tA\colon \tA\sim \tA^* \text{ for some }\tA^* \in \tS\right\}.
-\end{equation}
-The enlarged set $\tE_S$ satisfies the two properties below:
- \begin{enumerate}
-\item [1.] [Union of stationary points] The set $\tE_S$ is an union of equivalent  classes generated by the limiting points in $\tS$.
-\item [2.] [Connectedness module the equivalence]  The set $\tE_S$ is connected module the equivalence relationship. That property is obtained by the connectedness of $S$.
- \end{enumerate}
-Now, note that the isolation of stationary points and Property 1 imply that $\tE_S$ contains only finite number of equivalent classes.
-Otherwise, there is a subsequence of non-equivalent stationary points whose limit is not isolated, which contradicts the isolation assumption.
-Combining the finiteness with Property 2, we conclude that $\tE_S$ contains only a single equivalent class; i.e.
-$\tE_S = \tE_{\{\tA^*\}}$, where $\tA^*$ is a stationary point of $\tL(\tA)$.
-Therefore, all the convergent sub-sequences of $\tA^{(t)}$ converge to one stationary point $\tA^*$ up to equivalence.
-We conclude that, any iterate $\tA^{(t)}$ generated by Algorithm 1 converges to a stationary point of $\tL(\tA)$ up to equivalence.
- \end{proof}
-
-\section{Computational complexity}
-The computational complexity of our Algorithm~\eqref{alg:B} is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the data tensor.
-More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor.
-The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$ feature matrix.
-Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$.
-This complexity in tensor dimension matches with the classical tensor decomposition~\citep{kolda2009tensor}.
 

### 2020-10-01 tensor_regress/NeurIPSw/neuripsw_tensor.tex
workshop

--- 2020-09-17 tensor_regress/JASA_submission/paper_source/regress.tex
@@
-\documentclass[12pt]{article}
+\documentclass{article}
+
+% if you need to pass options to natbib, use, e.g.:
+%     \PassOptionsToPackage{numbers, compress}{natbib}
+% before loading neurips_2020
+
+% ready for submission
+ \usepackage{neurips_2020}
+
+% to compile a preprint version, e.g., for submission to arXiv, add add the
+% [preprint] option:
+%     \usepackage[preprint]{neurips_2020}
+
+% to compile a camera-ready version, add the [final] option, e.g.:
+%     \usepackage[final]{neurips_2020}
+
+% to avoid loading the natbib package, add option nonatbib:
+%     \usepackage[nonatbib]{neurips_2020}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+
 \usepackage{amsmath}
@@
 \hypersetup{colorlinks,linkcolor={black},citecolor={black},urlcolor={black}}
-
-\newcommand{\blind}{1}
-
-\addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
-\addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{-.3in}%
-\addtolength{\topmargin}{-.8in}%
 
@@
 
-\newtheorem{innercustomgeneric}{\customgenericname}
-\providecommand{\customgenericname}{}
-\newcommand{\newcustomtheorem}[2]{%
-  \newenvironment{#1}[1]
-  {%
-   \renewcommand\customgenericname{#2}%
-   \renewcommand\theinnercustomgeneric{##1}%
-   \innercustomgeneric
-  }
-  {\endinnercustomgeneric}
-}
-
-\newcustomtheorem{customexample}{Example}
-
 \usepackage{appendix}
@@
 \mathtoolsset{showonlyrefs}
-
-\input macros.tex
-
-\usepackage[english]{babel}
-
-\newcommand*{\KeepStyleUnderBrace}[1]{%f
-  \mathop{%
-    \mathchoice
-    {\underbrace{\displaystyle#1}}%
-    {\underbrace{\textstyle#1}}%
-    {\underbrace{\scriptstyle#1}}%
-    {\underbrace{\scriptscriptstyle#1}}%
-  }\limits
-}
-\usepackage{xr}
 
@@
 
-\def\ci{\perp\!\!\!\perp}
+\usepackage{xr}
 
-\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}
-\usepackage{booktabs}
-\newcommand\doubleRule{\toprule\toprule}
-\allowdisplaybreaks
+\input macros.tex
 
-\usepackage{xr}
-\externaldocument{supp}
+\title{Learning Multiple Networks via Supervised Tensor Decomposition}
+
+% The \author macro works with any number of authors. There are two commands
+% used to separate the names and addresses of multiple authors: \And and \AND.
+%
+% Using \And between authors leaves it to LaTeX to determine where to break the
+% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
+% authors names on the first line, and the last on the second line, try using
+% \AND instead of \And before the third author name.
+
+%\author{%}
 
@@
 
-\def\spacingset#1{\renewcommand{\baselinestretch}%
-{#1}\small\normalsize} \spacingset{1}
+\maketitle
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{abstract}
+   We develop a tensor decomposition method that incorporates multiple side information as interactive features.
+Unlike unsupervised tensor decomposition, our supervised decomposition captures the effective dimension reduction of the data tensor confined to feature space on each mode.
+An efficient alternating optimization algorithm is further developed.
+Our proposal handles a broad range of data types, including continuous, count, and binary observations.
+We apply the method to diffusion tensor imaging data from human connectome project.
+We identify the key global brain connectivity pattern and pinpoint the local regions that are associated with available features.
+Our method will help the practitioners efficiently analyze tensor datasets in various areas.
+\end{abstract}
 
-\if1\blind
-{
-  \title{\bf Supervised Tensor Decomposition with Interactive Side Information}
-  % blind version
-  \author{Jiaxin Hu, Zhuoyan Xu, Chanwoo Lee, and Miaoyan Wang\\
-    Department of Statistics, University of Wisconsin-Madison}
-  \maketitle
-} \fi
-
-\if0\blind
-{
-  \bigskip
-  \bigskip
-  \bigskip
-  \begin{center}
-    {\LARGE\bf Supervised Tensor Decomposition with Interactive Side Information}
-\end{center}
-  \medskip
-} \fi
-
-\bigskip
-\begin{abstract}
+\section{Introduction}
  Higher-order tensors have received increased attention across science and engineering.
@@
 Such data problems are common in neuroimaging, network analysis, and spatial-temporal modeling.
-Identifying the relationship between a high-dimensional tensor and side information is important yet challenging.
-Here, we develop a tensor decomposition method that incorporates multiple side information as interactive features.
-Unlike unsupervised tensor decomposition, our supervised decomposition captures the effective dimension reduction of the data tensor confined to feature space on each mode.
-An efficient alternating optimization algorithm is further developed.
-Our proposal handles a broad range of data types, including continuous, count, and binary observations.
-We apply the method to diffusion tensor imaging data from human connectome project and multi-relational political network data.
-We identify the key global connectivity pattern and pinpoint the local regions that are associated with available features.
-Our method will help the practitioners efficiently analyze tensor datasets in various areas.
-Toward this end, the package and data used are available at~\url{https://CRAN.R-project.org/package=tensorregress}.
-
-\end{abstract}
-
-\noindent%
-{\it Keywords:} Applications and case studies, Tensor data analysis, Supervised dimension reduction, Exponential family distribution, Generalized multilinear model
-\vfill
-
-\newpage
-\spacingset{1.5} % DON'T change the spacing!
-\section{Introduction}
-\label{sec:intro}
-
-Multi-dimensional arrays, known as tensors, are often collected with side information on multiple modes in modern scientific and engineering studies.
 A popular example is in neuroimaging~\citep{zhou2013tensor}.
@@
 
-In addition to the aforementioned challenges, many tensor datasets consist of non-Gaussian measurements.
-Examples include the political interaction dataset \citep{nickel2011three} which measures action counts between countries under various relations, and the brain connectivity network dataset \citep{zhang2018mapping} which is a collection of binary adjacency matrices.
-Classical tensor decomposition methods are based on minimizing the Frobenius norm of deviation, leading to suboptimal predictions for binary- or count-valued response variables.
-A number of supervised tensor methods have been proposed \citep{narita2012tensor, zhao2012higher, yu2016learning,lock2018supervised} to address the tensor regression problem in various forms, such as scalar-to-tensor regression and tensor-response regression.
-These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest.
-The gap between theory and practice means a great opportunity to modeling paradigms and better capture the complexity in tensor data.
-
-\begin{figure*}[t]
+ \begin{figure}[hbt]
 \begin{center}
-\includegraphics[width=16cm]{demo.pdf}
+\includegraphics[width = 13.5cm]{demo.pdf}
 \vspace{-.5cm}
@@
 \vspace{-.3cm}
-\end{figure*}
+\end{figure}
+
+In addition to the aforementioned challenges, many tensor datasets consist of non-Gaussian measurements.
+Classical tensor decomposition methods are based on minimizing the Frobenius norm of deviation, leading to suboptimal predictions for binary- or count-valued response variables.
+A number of supervised tensor methods have been proposed \citep{narita2012tensor, zhao2012higher, yu2016learning,lock2018supervised}.
+These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest.
+The gap between theory and practice means a great opportunity to modeling paradigms and better capture the complexity in tensor data.
 
@@
 We formulate the learning task as a structured regression problem, with tensor observation serving as the response, and the multiple side information as interactive features.
-Figure~\ref{fig:intro1}b illustrates our model in the special case of order-3 tensors.
-A low-rank structure is imposed to the conditional mean of tensor observation, where unlike classical decomposition, the tensor factors $\mX_k\mM_k\in\mathbb{R}^{d_k\times r_k}$ belong to the space spanned by features $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,2,3$.
+We leverage generalized linear model (GLM)  to allow heteroscedacity due to the mean-variance relationship in the non-Gaussian data.
+The low-rank structure on the conditional mean tensor effectively mitigates the curse of high dimensionality.
+Our proposal blends the modeling power of GLM and the exploratory capability of tensor dimension reduction in order to take the best out of both worlds.
+%A low-rank structure is imposed to the conditional mean of tensor observation.where unlike classical decomposition, the tensor factors $\mX_k\mM_k\in\mathbb{R}^{d_k\times r_k}$ belong to the space spanned by features $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,2,3$.
 The unknown matrices $\mM_k\in\mathbb{R}^{p_k\times r_k}$ (referred to as ``dimension reduction matrices'') link the conditional mean to the feature spaces, thereby allowing the identification of variations in the tensor data attributable to the side information.
 
-Our proposal blends the modeling power of generalized linear model (GLM) and the exploratory capability of tensor dimension reduction in order to take the best out of both worlds.
-We leverage GLM to allow heteroscedacity due to the mean-variance relationship in the non-Gaussian data.
-This flexibility is important in practice.
+ % This flexibility is important in practice.
 Furthermore, our low-rank model on the (transformed) conditional mean tensor effectively mitigates the curse of high dimensionality.
@@
 
-Our work is closely related to but also clearly distinctive from several lines of previous work.
-The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor, hong2020generalized} that aims to find the best low-rank representation of a data tensor.
-In contrast, our model is a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and multiple features.
-The low-rank factorization is determined jointly by the tensor data and feature matrices in our model.
+\vspace{-0.3cm}
 
-The second line of work studies the tensor-on-tensor regression~\citep{raskutti2019convex, lock2018tensor,gahrooei2020multiple}.
-Our model shares a common ground with earlier approaches, but we provide more efficient solutions to new settings that have more practical significance.
-As we show in Section~\ref{sec:connection}, the supervised tensor decomposition has an interesting connection with tensor-on-tensor regression.
-Previous methods~\citep{lock2018tensor} mainly focus on Gaussian tensors.
-The Frobenius norm used in the objective function is statistical suboptimal for general exponential family tensors.
-Maximum likelihood estimator (MLE) is studied in~\citep{raskutti2019convex} and a convex relaxation algorithm is proposed to solve for low-rank tensor coefficients.
-However, in the tensor case, convex MLE suffers from both computational intractability and statistical suboptimality.
-We advocate a non-convex approach and provide strong evidence for its success in our setting.
-Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2019convex}, and we go step further by finding the sufficient dimension reduction~\citep{adragni2009sufficient}, $\textup{Span}(\mM_k)$, that facilitates the identification of interaction effects in features (see Figure~\ref{fig:intro1}b).
-The latter approach greatly improves the \emph{interpretability} in prediction.
-In this regards, our method opens up new opportunities for tensor data analysis in a wider range of applications.
+\section{Method}
 
-The third line of work uses side information for various tensor learning tasks, such as for completion~\citep{narita2012tensor} and for recommendation system~\citep{ioannidis2019coupled,farias2019learning}.
-These methods also study tensors with side information, but they take regularization approaches to penalize predictions that are distant from side information~\cite{lock2018supervised}.
-One important difference is that their goal is prediction but not parameter estimation.
-The effects of features and their interactions are not estimated in these data-driven approaches.
-In contrast, our goal is interpretable prediction, and we estimate factor matrices $\mM_k$ using a model-based approach.
-Estimating $\mM_k$ allows us to identify sufficient features and the interactions thereof.
-We numerically compare these two approaches in Section~\ref{sec:simulation}.
-
-The paper is organized as follows.
-Section~\ref{sec:pre} introduces tensor preliminaries.
-Section~\ref{sec:model} presents the main model and three motivating examples for supervised tensor decomposition.
-We describe the quasi-likelihood estimation in Section~\ref{sec:est} and provide numerical experiments in Section~\ref{sec:simulation}.
-In Section~\ref{sec:data}, we present two data applications, one on diffusion tensor imaging data and the other on multi-relational network data.
-%and applications to diffusion tensor imaging data and multi-relational social network data in Section~\ref{sec:data}. 
-%In Section~\ref{sec:simulation}, we present numerical experiments and assess the performance in comparison to alternative methods.
-In Section~\ref{sec:data}, we apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
-We conclude in Section~\ref{sec:con} with discussions about our findings and avenues of future work.
-
-\section{Preliminaries}\label{sec:pre}
-We introduce the basic tensor properties~\citep{kolda2009tensor} used in the paper.
+\subsection{Preliminary}
+We introduce the basic tensor properties used in the paper.
 We use lower-case letters (e.g.,\ $a,b,c$) for scalars and vectors, upper-case boldface letters (e.g.,\ $\mA, \mB, \mC$) for matrices, and calligraphy letters (e.g.,\ $\tA, \tB, \tC$) for tensors of order three or greater.
 Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
-The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as $\tY \times_1\mX_1\times \cdots \times_K \mX_K=\entry{\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}x_{j_1,i_1}^{(1)}\cdots x_{j_K,i_K}^{(K)}}$,
-which results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor.
-For ease of presentation, we use the shorthand $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product.
-For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as $\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.$
-The tensor Frobenius norm is defined as $\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}$, and the maximum norm is defined $\mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.$
-When $a$ is a vector, we use $\vnormSize{}{a}=\langle a,a \rangle^{1/2}$ to denote the vector 2-norm.
-A higher-order tensor can be reshaped into a lower-order object.
-We use $\textup{vec}(\cdot)$ to denote the operation that reshapes the tensor into a vector, and $\text{Unfold}_k(\cdot)$ to denote the operation that reshapes the tensor along mode $k$ into a matrix of size $d_k$-by-$\prod_{i\neq k}d_i$.
-The multilinear rank of an order-$K$ tensor $\tY$ is defined as a length-$K$ vector $\mr=(r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\textup{Unfold}_k(\tY)$, $k=1,\ldots,K$.
-We use $\sigma_{\min}(\cdot)$ and $\sigma_{\max}(\cdot)$ denote the smallest and largest matrix singular values.
-We let $\mI_d$ denote the $d \times d$ identity matrix and $[d]$ denote the $d$-set $\{1,\ldots,d\}$.
-For ease of notation, we allow the basic arithmetic operators (e.g.,\ $+, -, \geq $) and univariate functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner.
-%We use $f'(\cdot)$, $f''(\cdot)$ to denote the first-order and second-order derivative of a function $f(\cdot)$.
+The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as $\tY\times\{\mX_1,\ldots,\mX_K\} =\entry{\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}x_{j_1,i_1}^{(1)}\cdots x_{j_K,i_K}^{(K)}}$,
+which results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor.For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as $\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.$
+The tensor Frobenius norm is defined as $\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}$, and the maximum norm is defined $\mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.$  We let $\mI_d$ denote the $d \times d$ identity matrix and $[d]$ denote the $d$-set $\{1,\ldots,d\}$.
+%For any two matrices $\mA,\mB$ with orthonormal columns of same dimension, the angle distance is defined as
+%\[ \sin \Theta(\mA,\mB)=\snormSize{}{\mA^T \mB^{\perp}}=\max\left\{ \frac{\langle \mx, \my\rangle}{\vnormSize{}{\mx}\vnormSize{}{\my}}\colon \ \mx\in \textup{Span}(\mA),\ \my\in \textup{Span}(\mB^\perp)\right\}.
+\]
 
-\section{Motivation and model}\label{sec:model}
-\subsection{General framework for supervised tensor decomposition}
-We begin with a general framework for supervised tensor decomposition and then discuss its implication in three concrete examples.
+\subsection{General Model}
+
 Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor.
@@
 Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the feature matrix on the mode $k\in[K]$, where $x_{ij}$ denotes the $j$-th feature value for the $i$-th tensor entity, for $(i,j)\in[d_k]\times[p_k]$, $p_k\leq d_k$.
-We propose a multilinear conditional mean model between the data tensor and feature matrices.
 Assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent realizations from an exponential family distribution, and the conditional mean tensor admits the form
@@
 \end{align}
-where $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the unknown parameter tensor, $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$, and $\times$ denotes the tensor-by-matrix product.
+where $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the unknown parameter tensor, $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$.
 The choice of link function is based on the assumed distribution family of tensor entries.
-Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution.
-In general, dispersion parameters can also be included in the model.
-Because our main focus is the tensor decomposition under the mean model, we omit the dispersion parameter in this section for ease of presentation.
 
@@
 where the parameters of interest are $\mM_k$ and $\tC$.
-Note that model~\eqref{eq:decomp} assumes a fixed, known rank $\mr=(r_1,\ldots,r_K)$; the adaptation to unknown rank will be addressed in Section~\ref{sec:tuning}.
+Note that model~\eqref{eq:decomp} assumes a fixed, known rank $\mr=(r_1,\ldots,r_K)$.
 Figure~\ref{fig:intro1}b provides a schematic illustration of our model.
 The features $\mX_k$ affect the distribution of tensor entries in $\tY$ through the form $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$.
-We call $\mX_k\mM_k$ the ``supervised tensor factors'' or ``sufficient features''~\citep{adragni2009sufficient}, and call $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes.
+We call $\mX_k\mM_k$ the ``supervised tensor factors'' or ``sufficient features''~\citep{adragni2009sufficient}, and call $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes, which links the conditional mean to the feature spaces, and thereby allows the identification of variations in the tensor data attributable to the side information.
 Our goal is to find $\mM_k$ and the corresponding $\tC$.
@@
 
-\subsection{Three examples}
-We give three seemingly different examples that can all be formulated as our supervised tensor decomposition model~\eqref{eq:decomp}.
-\begin{example}[Spatio-temporal growth model]
-The growth curve model~\citep{gabriel1998generalised} was originally proposed as an example of bilinear model for matrix data, and we adopt its higher-order extension here.
-Let $\tY=\entry{y_{ijk}}\in\mathbb{R}^{d \times m\times n}$ denote the pH measurements of $d$ lakes at $m$ levels of depth and for $n$ time points.
-Suppose the sampled lakes belong to $q$ types, with $p$ lakes in each type.
-Let $\{\ell_j\}_{j\in[m]}$ denote the sampled depth levels and $\{t_k\}_{k\in[n]}$ the time points.
-Assume that the expected pH trend in depth is a polynomial of order at most $r$ and that the expected trend in time is a polynomial of order $s$.
-Then, the conditional mean model for the spatio-temporal growth can be represented as
-\begin{equation}\label{eq:time}
-\mathbb{E}(\tY|\mX_1,\mX_2,\mX_3)=\tC\times\{\mX_1\mM_1,\ \mX_2\mM_2,\ \mX_3\mM_3\},
-\end{equation}
-where $\mX_1=\text{blockdiag}\{\mathbf{1}_p,\ldots,\mathbf{1}_p\}\in \{0,1\}^{d\times q}$ is the design matrix for lake types, and
-\[
-\mX_2=
-\begin{pmatrix}
-1 & \ell_1&\cdots &\ell^{r}_1\\
-1 & \ell_2&\cdots &\ell^{r}_2\\
-\vdots &\vdots&\ddots&\vdots\\
-1&\ell_{m}&\cdots&\ell^{r}_{m}
-\end{pmatrix},\quad
-\mX_3=
-\begin{pmatrix}
-1 & t_1&\cdots &t^{s}_1\\
-1 & t_2&\cdots &t^{s}_2\\
-\vdots &\vdots&\ddots&\vdots\\
-1&t_{n}&\cdots&t^{s}_{n}
-\end{pmatrix}
-\]
-are the design matrices for spatial and temporal effects, respectively, $\tC\in\mathbb{R}^{r_1\times r_2\times r_3}$ is the unknown core tensor, and $\mM_k$ are unknown dimension reduction matrices on each mode.
-The factors $\mX_k\mM_k$ are sufficient features in the mean model~\eqref{eq:time}.
-The spatial-temporal model is a special case of our supervised tensor decomposition model~\eqref{eq:decomp}, with features available on each of the three modes.
-\end{example}
-
-\begin{example}[Network population model]\label{example:brain}
-Network response model~\citep{rabusseau2016low} is recently developed for neuroimaging analysis.
-The goal is to study the relationship between brain network connectivity pattern and features of individuals.
-Suppose we have a sample of $n$ observations, $\{(\mY_i, \mx_i)\colon i=1,\ldots,n\}$, where for each individual $i\in[n]$, $\mY_i\in\{0,1\}^{d\times d}$ is the undirected adjacency matrix whose entries indicate presences/absences of connectivities between $d$ brain nodes, and $\mx_i\in\mathbb{R}^p$ is the individual's feature such as age, gender, cognition score, etc.
-The network-response model  has the conditional mean
-\begin{equation}\label{eq:network}
-\textup{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n,
-\end{equation}
-where $\tB\in \mathbb{R}^{d\times d\times p}$ is a rank-$(r_1,r_1,r_2)$ coefficient tensor, and $\tB$ is assumed to be symmetric in the first two modes.
-
-The model~\eqref{eq:network} is a special case of our supervised tensor decomposition, with feature matrix on the last mode of the tensor.
-Specifically, we stack the network observations $\{\mY_i\}$ together and obtain an order-3 response tensor $\tY\in\{0,1\}^{d\times d\times n}$.
-Define a feature matrix $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$.
-Then, the model~\eqref{eq:network} has the equivalent representation of supervised tensor decomposition,
-\[
-\textup{logit}(\mathbb{E}(\tY|\mX))=\tC\times\{\mM,\ \mM,\ \mX\mM'\},
-\]
-where $\tC\in\mathbb{R}^{r_1\times r_1\times r_2}$ is the core tensor, $\mM\in\mathbb{R}^{d\times r}$ is the dimension reduction matrix on the first two modes, and $\mM'\in\mathbb{R}^{p\times r_2}$ is for the last mode.
-\end{example}
-
- \begin{example}[Dyadic data with node attributes] Dyadic dataset consists of measurements on pairs of objects.
-Common examples include graphs and networks.
-Let $\tG=(V,E)$ denote a graph, where $V=[d]$ is the node set of the graph, and $E\subset V\times V$ is the edge set.
-Suppose that we also observe feature vector $\mx_i\in\mathbb{R}^p$ associated to each node $i\in V$.
-A probabilistic model on the graph $\tG=(V,E)$ can be described by the following matrix regression.
-The edge connects the two vertices $i$ and $j$ independently of other pairs, and the probability of connection is modeled as
-\begin{equation}\label{eq:edge}
- \textup{logit}\left(\mathbb{P}\left((i,j)\in E\right)\right)=\mx^T_i\mB\mx_j=\langle \mB,\ \mx^T_i\mx_j\rangle,
- \end{equation}
- where $\mB\in\mathbb{R}^{p\times p}$ is a symmetric rank-$r$ matrix.
-The low-rankness in $\mB$ has demonstrated its success in modeling transitivity, balance, and communities in  networks~\citep{hoff2005bilinear}.
-We show that our supervised tensor decompostion~\eqref{eq:decomp} also incorporates the graph model as a special case.
-Let $\tY=\entry{y_{ij}}$ be a binary matrix where $y_{ij}=\mathds{1}_{(i,j)\in E}$.
-Define $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$.
-Then, the graph model~\eqref{eq:edge} can be expressed as
- \[
- \textup{logit}(\mathbb{E}(\mY|\mX))=\mC\times\{\mX\mM,\ \mX\mM\}, 
-  \]
-  where $\mC\in\mathbb{R}^{r\times r}, \mM\in\mathbb{R}^{p\times r}$ are from the singular value decomposition of $\mB=\mM\mC\mM^T$. 
-  \end{example}
-
-%In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by features.
-The coefficient tensor $\tB$ in model~\eqref{eq:tensormodel} serves this goal by collecting the feature effects.
-To encourage the sharing among effects, decomposition~\eqref{eq:factor} assumes that the feature effects $\tB$ lie in a low-dimensional parameter space.
-The low-rank assumption is plausible in many scientific applications.
-In the Example~\ref{example:brain} of brain imaging analysis, for instance, it is often believed that the brain nodes can be partitioned into several regions, and the person's attributes affect the connectivity in a similar way within a region.
-
-In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by features.
-Our supervised tensor decomposition~\eqref{eq:decomp} allows arbitrary numbers of feature matrices.
-When certain mode $k$ has no side information, we set $\mX_k=\mI_{d_k}$ in the model~\eqref{eq:decomp}.
-In particular, our model~\eqref{eq:decomp} reduces to classical unsupervised tensor decomposition~\citep{de2000multilinear,hong2020generalized} when no side information is available; i.e., $\mX_k=\mI_{d_k}$ for all $k\in[K]$.
-
-\subsection{Connection to sufficient dimension reduction and tensor-on-tensor regression}\label{sec:connection}
-One important implication of our method is that we allow high-dimensionality in both tensor dimension $d_k$ and feature dimension $p_k$.
-Recall that the model rank $r_k$ is typically smaller than $d_k$ and $p_k$.
-In such a case, the matrices $\mM_k$ serve the role of simultaneous dimension reduction of the data tensor and features.
-The role of the sufficient features $\mX_k\mM_k$ can be seen in the following two conditional independence assumptions in model~\eqref{eq:decomp},
-
-\begin{align}
-\tY \ci \{\mX_k\}&\ \big|\ \{\mX_k\mM_k\} \ \ \text{(independence between the tensor and features)},\\
-y_{i_1,\ldots,i_K} \ci y_{i'_1,\ldots,i'_K} & \ \big|\ \{\mX_k\mM_k\}  \ \ \text{(independence within the tensor)},
-\end{align}
-where the second line holds for all $(i_1,\ldots,i_K)\neq(i'_1,\ldots,i'_K) \in[d_1]\times\cdots\times[d_K]$, and $\ci$ denotes the independence.
-The first property highlights the ``decorrelation'' role of $\mM_k$, in the same spirit as the sufficient dimension reduction~\citep{adragni2009sufficient} in supervised learning, whereas the second property highlights the tensor dimension reduction in the usual unsupervised sense (consider, for example, $\mX_k=\mathbf{I}_{d_k}$ for all $k\in[K]$).
-
-Our model also has a close connection with tensor-on-tensor regression~\citep{raskutti2019convex, lock2018tensor,gahrooei2020multiple}.
-Specifically, model~\eqref{eq:decomp} can be viewed as a multivariate regression model, where the response is the vectorized tensor and the covariates are interactions between sufficient features across modes.
-We take an order-3 tensor under the Gaussian model for illustration.
-Let $\mX,\mZ,\mW$ denote the feature matrix on mode $k=1, 2, 3$, respectively.
-Suppose that each mode has two-dimensional sufficient features, denoted $\mM_1\mX=[\mx_1,\mx_2]$, $\mM_2\mZ=[\mz_1,\mz_2]$, $\mM_3\mW=[\mw_1,\mw_2]$.
-Here $\mx_1,\mx_2,\ldots,\mw_1, \mw_2$ are column vectors.
-Then the model~\eqref{eq:decomp} is a regression model with across-mode interactions,
-\begin{equation}\label{eq:interaction}
-\mathbb{E}(y_{ijk}|\mX,\mZ,\mW)=c_{111}\mx_{1i}\mz_{1j}\mw_{1k}+c_{121}\mx_{i1}\mz_{j2}\mw_{k1}+\cdots+c_{221}\mx_{2i}\mz_{2j}\mw_{1k}+c_{222}\mx_{2i}\mz_{2j}\mw_{2k},
-\end{equation}
-where $\entry{c_{ijk}} \in \mathbb{R}^{2\times 2\times 3}$ are unknown interaction effects, $\mx_{1i}$ denotes the $i$-th entry in the feature vector $\mx_1$, and similar notations apply to other features.
-Note that lower-order interactions are naturally incorporated in~\eqref{eq:interaction} if we include an intercept column in the sufficient feature matrices.
-Model~\eqref{eq:interaction} shows the benefit of our supervised tensor decomposition for identifying across-mode interactions.
-
-\section{Estimation}\label{sec:est}
-
+\vspace{-.3cm}
+\section{Estimation}
 \subsection{Rank-constrained M-estimator}
-We develop a likelihood-based procedure to estimate $\tC$ and $\mM_k$ in~\eqref{eq:decomp}.
-We adopt the exponential family as a flexible framework for different data types.
-In a classical generalized linear model with a scalar response $y$ and feature $\mx$, the density is expressed as
+ We adopt the exponential family as a flexible framework for different data types. In a classical generalized linear model with a scalar response $y$ and feature $\mx$, the density is expressed as
 \[
@@
 \]
-where $b(\cdot)$ is a known function, $\theta$ is the linear predictor, $\phi>0$ is the dispersion parameter, and $c(\cdot)$ is a known normalizing function.
-The choice of link functions depends on the data types and on the observation domain of $y$, denoted $\mathbb{Y}$.
-For example, the observation domain is $\mathbb{Y}=\mathbb{R}$ for continuous data, $\mathbb{Y}=\mathbb{N}$ for count data, and  $\mathbb{Y}=\{0,1\}$ for binary data.
-The canonical link function $f$ is chosen to be $f(\cdot)=b'(\cdot)$, the first-order derivative of $b(\cdot)$.
-Table~1 summarizes the canonical link functions for common types of distributions.
-
-\begin{table}[htb]
-\centering
-\begin{tabular}{c|ccc}
-Data type &Gaussian & Poisson& Bernoulli\\
-\hline
-Domain $\mathbb{Y}$& $\mathbb{R}$&$\mathbb{N}$&$\{0,1\}$\\
- $b(\theta)$&$\theta^2/2$& $\exp(\theta)$&$\log (1+\exp(\theta))$\\
- link $f(\theta)$&$\theta$&$\exp(\theta)$&$(1+\exp(-\theta))^{-1}$
-\end{tabular}
-\vspace{-.2cm}
-\caption{Canonical links for common distributions.}\label{table:link}
-\end{table}
-
+where $b(\cdot)$ is a known function depending on the data types, $\theta$ is the linear predictor, $\phi>0$ is the dispersion parameter, and $c(\cdot)$ is a known normalizing function.
 In our context, we model the tensor entries $y_{i_1,\ldots,i_K}$, conditional on $\theta_{i_1,\ldots,i_K}$, as independent draws from an exponential family.
@@
 \end{align}
-where the parameter space $\tP$ is
-\[
+where parameter space 
+$
 \tP=\left\{(\tC, \mM_1,\ldots,\mM_K) \ \Big| \ \mM_k^T\mM_k=\mathbf{I}_{r_k},\ \mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha \right\}.
-\]
+$
 The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to avoid the divergence in the non-Gaussian variance.
 
-\subsection{Alternating optimization} \label{subsec:alg}
-We propose an alternating optimization algorithm to solve~\eqref{eq:MLE}. 
+\subsection{Alternating optimization}
+
 The decision variables in the objective function~\eqref{eq:MLE} consist of $K+1$ blocks of variables, one for the core tensor $\tC$ and $K$ for the factor matrices $\mM_k$.
@@
 This observation leads to an iterative updating scheme for one block at a time while keeping others fixed.
-After each iteration, we rescale the core tensor $\tC^{(t+1)}$ subject to the maximum norm constraint.
-The full algorithm is described in Algorithm~\ref{alg:B}.
+A simplified version of the algorithm is described in Algorithm~\ref{alg:B}.
 
 \begin{algorithm}[!h]
-\caption{Supervised Tensor Decomposition with Interactive Side Information}\label{alg:B}
+\caption{Supervised Tensor Decomposition with Interactive Side Information (Simplified)}\label{alg:B}
 \begin{algorithmic}[1]
@@
 \While{Do until convergence}
-\For { $k=1$ to $K$}
-\State Obtain the factor matrix $\tilde \mM_k\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$.
-\State Perform QR factorization $\tilde \mM_k=\mQ\mR$.
-\State Update $\mM_k\leftarrow \mQ$ and core tensor $\tC\leftarrow \tC\times_k \mR$.
-\EndFor
-\State Update the core tensor $\tC$ by solving a GLM with $\textup{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM_k]$ as features, and $f$ as link function.
-Here $\otimes$ denotes the Kronecker product of matrices.
-\State Rescale the core tensor $\tC$ subject to the maximum norm constraint. 
+
+%\State Obtain the factor matrix $\tilde \mM_k\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$.
+%\State Perform QR factorization $\tilde \mM_k=\mQ\mR$.
+%\State Update $\mM_k\leftarrow \mQ$ and core tensor $\tC\leftarrow \tC\times_k \mR$.
+\State Obtain $\tilde \mM_k\in\mathbb{R}^{p_k\times r_k}$ by a GLM. Orthogonalize $\tilde \mM_k$ by QR factorization, for $k \in [K]$.
+
+\State Update the core tensor $\tC$ by solving a GLM. Rescale the core tensor $\tC$ such that $\maxnorm{\tC} \leq \alpha$. 
 \EndWhile
@@
 
-The optimization~\eqref{eq:MLE} is a non-convex problem due to the non-convexity in the feasible set $\tP$.
-Under mild conditions, our algorithm enjoys global convergence; i.e.\ any sequence of iterates generated by the alternating algorithm converges to a stationary point of $\tL_\tY(\cdot)$ module orthogonal transformation (see Supplementary Notes, Section~\ref{sec:SAlgorithm}).
-In theory, global optimality of non-convex optimization is often challenging to obtain.
-Fortunately, we will show in Section~\ref{subsec:statprob} that the desired statistical performance holds for all local optimizers satisfying $\tL_{\tY}(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)\geq \tL_{\tY} (\tC_{\textup{true}},\mM_{1,\textup{true}},\ldots,\mM_{K,\textup{true}})$, where the subscript ``true'' denotes the true parameter in model~\eqref{eq:decomp}.
-This result indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective is large enough.
-In the experiments we considered, we find that Algorithm~\ref{alg:B} typically gives satisfactory convergent points upon random initialization.
-Figure~\ref{fig:loglike} shows the trajectories of objective function for order-3 tensors based on model~\eqref{eq:decomp}, where $d_k = d \in\{25,30\}$, $p_k = 0.4d$, $r_k = r\in\{3,6\}$ for $ k = 1,2,3$.
-We consider input tensors with Gaussian, Bernoulli, and Poisson entries.
-Under all combinations of the dimension $d$, rank $r$, and type of the entries, Algorithm~\ref{alg:B} converges quickly in a few iterations, and the objective values at convergent points are close to or larger than the value at true parameters.
-The time complexity is provided in Supplementary Notes.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=15cm]{loglike.pdf}
-\vspace{-.25cm}
-\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c) Poisson models.
-The dashed line represents the objective value at true parameters.
-}\label{fig:loglike}
-\vspace{-.3cm}
-\end{figure}
-
-\subsection{Rank selection}\label{sec:tuning}
-Algorithm~\ref{alg:B} assumes the rank $\mr$ is given.
-In practice, the rank is often unknown and must be determined from the data.
-We propose to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC, where
-\begin{equation}\label{eq:BIC}
-%\hat \mr&=\argmin_{\mr=(r_1,\ldots,r_K)} \textup{BIC}(\mr)\notag\\
-\textup{BIC}(\mr)=-2\tL_{\tY}(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)+p_e(\mr)\log (\prod\nolimits_k d_k).
-\end{equation}
-Here, $p_e(\mr)\stackrel{\textup{def}}{=}\sum_k (p_k-r_k-1)r_k+\prod_k r_k$ is the effective number of parameters in the model.
-We choose $\hat \mr$ that minimizes $\textup{BIC}(\mr)$ via grid search.
-Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model.
-We evaluate the empirical performance of BIC in Section~\ref{sec:simulation}.
-
-%The computational complexity of our Algorithm~\eqref{alg:B} is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the data tensor.
-More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor.
-The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$ feature matrix.
-Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$.
-This complexity in tensor dimension matches with the classical tensor decomposition~\citep{kolda2009tensor}.
-
 \subsection{Statistical properties}\label{subsec:statprob}
-In this section, we provide the accuracy guarantee for the proposed M-estimator~\eqref{eq:MLE}.
-Note that the factor matrices $\mM_k$ are identifiable only up to orthogonal transformation.
-Therefore, we use angle distance to assess the accuracy in estimating the column space, $\textup{Span}(\mM_k)$.
-For any two matrices $\mA,\mB$ with orthonormal columns of same dimension, the angle distance is defined as
-\[
-\sin \Theta(\mA,\mB)=\snormSize{}{\mA^T \mB^{\perp}}=\max\left\{ \frac{\langle \mx, \my\rangle}{\vnormSize{}{\mx}\vnormSize{}{\my}}\colon \ \mx\in \textup{Span}(\mA),\ \my\in \textup{Span}(\mB)\right\}.
-\]
 
@@
 The classical MLE theory does not directly apply.
-We leverage the recent development in random tensor theory and high-dimensional statistics to establish the error bounds of the estimation.
-
-\begin{assumption}\label{ass}We make the following assumptions:
-\begin{enumerate}[noitemsep,topsep=0pt]
-\item [A1.] There exist two positive constants $c_1, c_2>0$ such that $c_1\leq \sigma_{\min}(\mX_k)\leq  \sigma_{\max}(\mX_k)\leq c_2$ for all $k\in[K]$. 
-%Here $\sigma_{min}(\cdot)$ and $\sigma_{max}(\cdot)$ denote the smallest and largest matrix singular values.
-\item [A1'.] The feature matrices $\mX_k$ are Gaussian designs with i.i.d.\ $N(0,1)$ entries.
-\item [A2.] There exist two positive constants $L, U>0$, such that $L\phi \leq \textup{Var}(y_{i_1,\ldots,i_K}|\theta_{i_1,\ldots,i_K})\leq U\phi $, or equivalently, $L\leq b''(\theta_{i_1,\ldots,i_K}) \leq U$, for all $|\theta_{i_1,\ldots,i_K}|\leq \alpha$.
-Here $\alpha$ is the upper bound of the linear predictor in~\eqref{eq:MLE}, and $b''(\cdot)$ denotes the second-order derivative.
-
-\end{enumerate}
-\end{assumption}
-The assumptions are fairly mild.
-Assumptions A1 and A1' consider two separate scenarios about feature matrices.
-Assumption A1 is applicable when feature matrix is asymptotically non-singular and has bounded spectral norm, whereas Assumption A1' imposes the commonly-used Gaussian design~\citep{raskutti2019convex}.
-Assumption A2 ensures the log-likelihood $\tL_\tY(\Theta)$ is strictly concave in the linear predictor $\Theta$.
-%The two equivalent statements in Assumption A2 are based on the fact that $\textup{Var}(y_{i_1,\ldots,i_K}|\tX,\tB)=\phi b''(\theta_{i_1,\ldots,i_K})$ when $y_{i_1,\ldots,i_K}$ belongs to an exponential family~\citep{mccullagh1989generalized}.
+We leverage the recent development in random tensor theory and high-dimensional statistics to establish the error bounds.
 
 \begin{thm}[Statistical convergence]\label{thm:main}
-Consider a data tensor generated from model~\eqref{eq:decomp}, where the entries are conditionally independent realizations from an exponential family.
+Consider a data tensor generated from model~\eqref{eq:decomp}.
 Let $(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)$ be the M-estimator in~\eqref{eq:MLE} and $\hat \tB=\hat \tC\times \hat \mM_1\times\cdots \times \hat \mM_K$.
 Define $r_\textup{total}=\prod_k r_k$ and $r_{\max}=\max_k r_k$.
-Under Assumptions A1 and A2 with scaled feature matrices $\check \mX_k= \sqrt{d_k}\mX_k$, or under Assumptions A1' and A2 with original feature matrices, there exist two positive constants $C_1=C_1(\alpha,K), C_2=C_2(\alpha, K)>0$ independent of dimensions $\{d_k\}$ and $\{p_k\}$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$,
+Under mild technical assumptions, there exist two positive constants $C_1=C_1(\alpha,K), C_2=C_2(\alpha, K)>0$ independent of dimensions $\{d_k\}$ and $\{p_k\}$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$,
 \begin{equation}\label{eq:bound}
@@
 \begin{equation}\label{eq:sinebound}
-\textup{sin}^2 \Theta(\trueM,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\sigma^2_{\min}(\textup{Unfold}_k(\tC_\textup{true}))}\frac{\sum_k p_k}{\prod_k d_k}.
+\textup{sin}^2 \Theta(\trueM,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\sigma^2_{\min}(\textup{Unfold}_k(\tC_\textup{true}))}\frac{\sum_k p_k}{\prod_k d_k},
 \end{equation}
+where $
+\sin \Theta(\trueM,\hat \mM_k)=\snormSize{}{\trueM^T \hat \mM_k^{\perp}}=\max\left\{ \frac{\langle \mx, \my\rangle}{\vnormSize{}{\mx}\vnormSize{}{\my}}\colon \ \mx\in \textup{Span}(\trueM),\ \my\in \textup{Span}({\hat \mM_k}^\perp)\right\}
+$ is the angle distance to assess the accuracy in estimating the column space, $\textup{Span}(\mM_k)$.
 \end{thm}
 
-Theorem~\ref{thm:main} establishes the statistical convergence for the estimator~\eqref{eq:MLE}.
-In fact, our proof shows that the desired convergence rate holds not only for the M-estimator, but also for any local optimizers satisfying $\tL_\tY(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K) \geq \tL_\tY(\tC_{\textup{true}},  \mM_{1,\textup{true}},\ldots,\mM_{K,\textup{true}})$.
-Consider a special case when tensor dimensions are equal on each of the modes, i.e., $d_k=d$ for all $k\in[K]$, and feature dimension grows with tensor dimension, $p_k=\gamma d$, $\gamma\in [0,1)$, for $k\in[K]$.
-The result in~\eqref{eq:bound} implies that the estimation has a convergence rate $\tO(d^{-(K-1)})$.
-Therefore, our estimation is consistent in high dimensional regimes, and the convergence becomes especially favorable as the order of tensor data increases.
+\section{Real data analysis}
 
-As immediate applications, we obtain the convergence rate for the three examples mentioned in Section~\ref{sec:model}.
-
-\begin{customexample}{1}[Spatio-temporal growth model] The estimated type-by-time-by-space coefficient tensor converges at the rate $\tO\left(\frac{p+r+s}{dmn}\right)$ where $p\leq d$, $r\leq m$ and $s\leq n$.
-The estimation achieves consistency as long as the dimension grows along either of the three modes.
-\end{customexample}
-
-\begin{customexample}{2} [Network population model] The estimated node-by-node-by-feature tensor converges at the rate $\tO\left(\frac{2d+p}{d^2n}\right)$ where $p\leq n$.
-The estimation achieves consistency as the number of individuals or the number of nodes grows.
-\end{customexample}
-
-\begin{customexample}{3} [Dyadic data with node attributes] The estimated feature-by-feature matrix converges at the rate $\tO\left(\frac{p}{d^2}\right)$ where $p\leq d$.
-Again, our estimation achieves consistency as the number of nodes grows.
-\end{customexample}
-
-We conclude this section by providing the accuracy measured in KL divergence.
-%, for the response distribution.
-
-\begin{cor}[Prediction error]\label{thm:KL}
-Assume the same set-up as in Theorem~\ref{thm:main}.
-Let $\mathbb{P}_{\tY_{\textup{true}}}$ and $\mathbb{P}_{\hat \tY}$ denote the distributions of $\tY$ given the true parameter $\trueB$ and estimated parameter $\hat \tB$, respectively.
-Then, we have, with probability at least $1-\exp(C_1\sum_k p_k)$,
-\[
-\textup{KL}(\mathbb{P}_{\tY_{\textup{true}}},\ \mathbb{P}_{\hat \tY})\leq \frac{C_3r_{\textup{total}}}{r_{\max}}\sum_k p_k,
-\]
-where $C_3=C_3(\alpha, K)>0$ is a constant independent of dimensions $\{d_k\}$ and $\{p_k\}$.
-\end{cor}
-
-\section{Numerical experiments}\label{sec:simulation}
-We evaluate the empirical performance of our supervised tensor decomposition (STD) through simulations.
-We consider order-3 tensors with a range of distribution types.
-Unless otherwise specified, the conditional mean tensor is generated form model~\eqref{eq:decomp}, where the core tensor entries are i.i.d.\ drawn from Uniform[-1,1], the factor matrix $\mM_k$ is uniformly sampled with respect to Haar measure from matrices with orthonormal columns.
-The feature matrix $\mX_k$ is either an identity matrix (i.e.,\ no feature  available) or Gaussian random matrix with i.i.d.\ entries from $N(0,1)$.
-The linear predictor $\Theta=\tC\times\{\mM_1\mX_1,\mM_2\mX_3,\mM_3\mX_3\}$ is scaled such that $\mnormSize{}{\Theta}=1$.
-Conditional on the linear predictor $\Theta=\entry{\theta_{ijk}}$, the entries in the tensor $\tY=\entry{y_{ijk}}$ are drawn independently according to three probabilistic models:
-
-\begin{enumerate}[noitemsep,topsep=0pt]
-\item[(a)] Gaussian model: continuous tensor entries $y_{ijk}\sim N\left(\alpha \theta_{ijk}, 1\right)$.
-\item[(b)] Poisson model: count tensor entries $y_{ijk}\sim\textup{Poisson}\left( e^{\alpha \theta_{ijk}}\right)$.
-\item[(c)] Bernoulli model: binary tensor entries $y_{ijk}\sim \textup{Bernoulli}\left( \frac{e^{\alpha \theta_{ijk}}}{1+e^{\alpha \theta_{ijk}}}\right)$.
-\end{enumerate}
-Here $\alpha>0$ is a scalar controlling the magnitude of the effect size.
-In each simulation study, we report the mean squared error (MSE), $\FnormSize{}{\tB_{\textup{true}}-\hat \tB}^2$, averaged across $30$ replications.
-
-\subsection{Finite-sample performance}
-The first experiment assesses the selection accuracy of our BIC criterion~\eqref{eq:BIC}.
-We consider the balanced situation where $d_k=d$, $p_k=0.4d_k$ for $k=1,2,3$.
-We set $\alpha=4$ and consider various combinations of dimension $d$ and rank $\mr=(r_1,r_2,r_3)$.
-For each combination, we minimize BIC using a grid search over three modes.
-The hyper-parameter $\alpha$ is set to infinity in the fitting, which essentially imposes no constraint on the tensor magnitude.
-Table~\ref{tab:rank} reports the selected rank averaged over $n_{\textup{sim}}=30$ replicates.
-We found that in the high-rank setting with $d=20$, the selected rank slightly underestimates the true rank, and the accuracy immediately improves when either the dimension increases to $d = 40$ or the rank reduces to $\mr = (3,3,3)$.
-This agrees with our expectation, because in the tensor decomposition, the sample size is related to the number of tensor entries.
-A larger $d$ implies a larger sample size, so the BIC selection becomes more accurate.
-
-\begin{table}[tb]
-\resizebox{\textwidth}{!}{%
-\centering
-\begin{tabular}{c|cc|cc}
-%True Rank $\mr$ &\multicolumn{2}{c|}{Dimension (Gaussian tensors)}&\multicolumn{2}{c|}{Dimension (Poisson tensors)}\\
-True rank $\mr$& $d=20$  (Gaussian) &$d=40$ (Gaussian) &$d=20$ (Poisson) &$d=40$ (Poisson)\\
-\hline
-$(3,\ 3,\ 3)$&$({\bf 3.0},\ {\bf 3.0},\ {\bf 3.0})$&$({\bf 3.0},\ {\bf 3.0},\ {\bf 3.0})$& $({\bf 3.0},\ {\bf 3.0},\ {\bf 3.0})$ & $({\bf 3.0},\ {\bf 3.0},\ {\bf 3.0})$\\
-$(4,\ 4,\ 6)$&$(3.0,\ 3.0,\ {\bf 4.6})$&$({\bf 4.0},\ {\bf 4.0},\ {\bf 5.3})$&$(3.0,\ 3.0,\ {\bf 5.3})$&$({\bf 4.0},\ {\bf 4.0},\ {\bf 5.6})$\\
-$(6,\ 8,\ 8)$&$(5.0,\ 5.0,\ 5.0)$&$({\bf 6.0},\ {\bf 8.0},\ {\bf 8.0})$&$(5.0,\ 5.0,\ 6.7)$&$({\bf 6.0},\ {\bf 8.0},\ {\bf 8.0})$\\
-\end{tabular}
-}
-\caption{Rank selection via BIC. The estimated ranks are averaged across 30 simulation. Bold number indicates the ground truth is within two standard deviations of the estimate.}\label{tab:rank}
-\vspace{-.1cm}
-\end{table}
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=14cm]{dimension.pdf}
-\caption{Estimation error against effective sample size.
-The three panels plot the MSE when the response tensors are generated from (a) Gaussian (b) Poisson and (c) Bernoulli models.
-The dashed curves correspond to $\tO({1/d^2})$.}\label{fig:dim}
-\vspace{-.1cm}
-\end{figure}
-
-The second experiment evaluates the accuracy when features are available on all modes.
-We set $\alpha=10, d_k=d, p_k=0.4d_k, r_k=r\in\{2,4,6\}$ and increase $d$ from 30 to 60.
-Our theoretical analysis suggests that $\hat \tB$ has a convergence rate $\tO(d^{-2})$ in this setting.
-Figure~\ref{fig:dim} plots the estimation error versus the ``effective sample size'', $d^2$, under three different distribution models.
-We find that the empirical MSE decreases roughly at the rate of $1/d^2$, which is consistent with our theoretical results.
-We also observe that, tensors with higher rank tend to yield higher estimation errors, as reflected by the upward shift of the curves as $r$ increases.
-Indeed, a larger $r$ implies a higher model complexity and thus greater difficulty in the estimation.
-
-\subsection{Comparison with GLMs under stochastic block models}
-The third experiment investigates the performance of our model under correlated feature effects.
-We mimic the scenario of brain imaging analysis.
-A sample of $d_3=50$ networks are simulated, one for each individual.
-Each network measures the connections between $d_1=d_2=20$ brain nodes.
-We simulate $p=5$ features for the each of the 50 individuals.
-These features may represent, for example, age, gender, cognitive score, etc.
-Recent study has suggested that brain connectivity networks often exhibit community structure represented as a collection of subnetworks, and each subnetwork is comprised of a set of spatially distributed brain nodes.
-To accommodate this structure, we utilize the stochastic block model~\citep{abbe2017community} to generate the effect size.
-Specifically, we partition the nodes into $r$ blocks by assigning each node to a block with uniform probability.
-Edges within a same block are assumed to share the same feature effects, where the effects are drawn i.i.d.\ from $N(0,1)$.
-We then apply our tensor regression model to the network data using the BIC-selected rank.
-Note that in this case, the true model rank is unknown; the rank of a $r$-block network is not necessarily equal to matrix rank $r$~\citep{zeng2019multiway}.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=16cm]{compare_GLM.pdf}
-\caption{Performance comparison under stochastic block models.
-The three panels plot the MSE when the response tensors are generated from (a) Gaussian (b) Poisson and (c) Bernoulli models.
-The $x$-axis represents the number of blocks in the networks.
-}\label{fig:glm}
-\vspace{-.2cm}
-\end{figure}
-
-Figure~\ref{fig:glm} compares the MSE of our method with a multiple-response GLM approach.
-The multiple-response GLM is to regress the dyadic edges, one at a time, on the features, and this model is repeatedly fitted for each edge.
-As we find in Figure~\ref{fig:glm}, our tensor regression method achieves significant error reduction in all three data types considered.
-The outperformance is substantial in the presence of large communities; even in the less structured case ($\sim 20/15=1.33$ nodes per block), our method still outer-performs GLM.
-The possible reason is that the multiple-response GLM approach does not account for the correlation among the edges, and suffers from overfitting.
-In contrast, the low-rankness in our modeling incorporates the shared information across entries.
-By selecting the rank in a data-driven way, our method achieves accurate estimation in a wide range of settings.
-
-\subsection{Comparison with other tensor methods}
-We compare our supervised tensor decomposition with three other tensor methods:
-\begin{itemize}[noitemsep,topsep=0pt]
-\item  Higher-order low-rank regression ({\bf HOLRR},~\citet{rabusseau2016low}) is a least-square based tensor regression that allows features on a single mode. 
-\item Higher-order partial least square ({\bf HOPLS},~\citet{zhao2012higher}) is a dimension-reduction method that jointly models a tensor response and a tensor feature. 
-\item Subsampled tensor projected gradient ({\bf TPG},~\citet{yu2016learning}) considers the same objective as {\bf HOLRR} but instead uses a different algorithm to solve the problem. 
-\end{itemize}
-These three methods are the closest algorithms to ours, in that they all relate a data tensor to features using a low-rank structure.
-The three existing methods allow only Gaussian data, whereas ours is applicable to any exponential family distribution including Gaussian, Bernoulli, Poisson, etc.
-For fair comparison, we consider Gaussian tensors in the experiment.
-Because not every method returns the effect estimate $\hat \tB$ as outputs, we measure the accuracy using mean squared prediction error, $\text{MSPE}=(\prod_kd_k)^{-1}\FnormSize{}{\hat \tY-f(\Theta)}^2$, where $f(\Theta)$ is the conditional mean of the tensor, and $\hat \tY$ is the fitted tensor from each method.
-
-The comparison is assessed from three aspects: (a) benefit of incorporating features from multiple modes; (b) prediction error with respect to sample size; (c) sensitivity of accuracy with respect to model complexity.
-We use similar simulation setups as in our experiment II, but consider combinations of rank, $\mr=(3,3,3)$ (low) vs.\ $(4,5,6)$ (high), signal $\alpha = 3$ (low) vs.\ $6$ (high), dimension $d$ ranging from 20 to 100 for modes with features, and $d = 20$ for modes without features.
-Two methods ({\bf STD} and {\bf HOLRR}) require the tensor rank as inputs.
-For fair comparison, we provide both algorithms the true rank.
-For algorithms ({\bf HOPLS} and {\bf TPG}) that use different notions of model rank, we use a grid search to set the hyperparameter that gives the best mean square prediction error.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=15cm]{compare_alternative.pdf} 
-\caption{Comparison between different tensor methods.
-Panels (a) and (b) plot MSPE versus the number of modes with available features.
-Panels (c) and (d) plot MSPE versus the effective sample size $d^2$.}~\label{fig:compare}
-%We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and signal $\alpha =3 $ (low), $\alpha=6$ (high).}~\label{fig:compare}
-\vspace{-.4cm}
-\end{figure}
-
-Figures~\ref{fig:compare}a-b show the averaged prediction error across 30 replicates.
-We see that our {\bf STD} outperforms others, especially in the low-signal, high-rank setting.
-As the number of informative modes (i.e., modes with available features) increases, the {\bf STD} exhibits a substantial reduction in error whereas others remain unchanged (Figure~\ref{fig:compare}b).
-This showcases the benefit of incorporation of multiple features.
-Note that our method {\bf STD} is most comparable to {\bf HOLRR} when there is only a single informative mode.
-In such a case, both methods share a same cost function but have different algorithms.
-{\bf STD} alternates between informative and non-informative modes, whereas {\bf HOLRR} approximates the non-informative modes without alternating.
-The accuracy gain in Figure~\ref{fig:compare} demonstrates the benefit of alternating algorithm -- incorporation of informative modes also improves the estimation in the non-informative modes.
-
-Figures~\ref{fig:compare}c-d compare the prediction error with respect to effective sample size when only one mode has side information.
-In the high-signal low-rank setting, our method has similar performance as {\bf HOLRR}, and the improvement becomes more pronounced in the low-signal high-rank setting.
-The latter setting is harder because of the higher inter-mode complexity, and our {\bf STD} method shows the advantage in addressing this challenge.
-
-\section{Data analyses}\label{sec:data}
-We apply our supervised tensor decomposition to two datasets.
-The first application studies the brain networks in response to individual attributes (i.e.,\ feature on one mode), and the second application focuses on multi-relational network analysis with dyadic attributes (i.e.,\ features on two modes).
-
-\subsection{Application to human brain connection data}
 The Human Connectome Project (HCP) aims to build a network map that characterizes the anatomical and functional connectivity within healthy human brains~\citep{HCP}. 
@@
 We consider four individual features: gender (65 females vs.\ 71 males), age 22-25 ($n=35$), age 26-30 ($n=58$), and age 31+ ($n=43$).
-The preprocessed dataset is released in our R package \texttt{tensorregress}.
 The goal is to identify the connection edges that are affected by individual features.
-A key challenge in brain network is that the edges are correlated; for example, the nodes in edges may be from a same brain region, and it is of importance to take into account the within-dyad dependence.
+%A key challenge in brain network is that the edges are correlated; for example, the nodes in edges may be from a same brain region, and it is of importance to take into account the within-dyad dependence.
 
@@
 \centering
-\includegraphics[width=16cm]{HCP.pdf}
+\includegraphics[width=14cm]{HCP.pdf}
 \caption{Top edges with large effects.
@@
 }\label{fig:brain}
-\vspace{1cm}
-\includegraphics[width=16cm]{compare_HCP.pdf}
-\caption{Comparison of estimated feature effects the HCP data using (a) multi-response GLM and (b) supervised tensor decomposition (STD). }\label{fig:s1}
-\vspace{-.1cm}
+%\vspace{1cm}
+%\includegraphics[width=14cm]{compare_HCP.pdf}
+%\caption{Comparison of estimated feature effects the HCP data using (a) multi-response GLM and (b) supervised tensor decomposition (STD). }\label{fig:s1}
+%\vspace{-.1cm}
 \end{figure}
@@
 
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=16cm]{compare_HCP.pdf}
-%\caption{Comparison of feature effect estimation in the HCP data.}\label{fig:s1}
-%\end{figure}
-
-Figure~\ref{fig:s1} compares the estimated coefficients from our method with those from multiple-response GLM approach.
-The multiple-response GLM is to regress the brain edges, one at a time, on the individual-level features, and this model is repeatedly fitted for every edge $\in [68]\times [68]$.
-As we can see in the figure, our tensor decomposition method shrinkages the coefficients towards center, thereby implicitly enforcing the sharing between feature effects.
-
-\subsection{Application to political relation data}
-
-The second application studies the multi-relational networks with node-level attributes.
-We consider \emph{Nations} dataset~\citep{nickel2011three} which records 56 relations among 14 countries between 1950 and 1965.
-The multi-relational networks can be organized into a $14 \times 14 \times 56$ binary tensor, with each entry indicating the presence or absence of an action, such as ``sending tourist to'', ``export'', ``import'', between countries.
-The 56 relations span the fields of politics, economics, military, religion, etc.
-In addition, country-level attributes are also available, and we focus on the following six features: \emph{constitutional, catholics, law ngo, political leadership, geography}, and \emph{medicine ngo}.
-The goal is to identify the variation in connections due to country-level attributes and their interactions.
-%One of the key features is that the 56 relations are correlated, and we would like to take it into account in assessing the feature effects.
-
-We apply our tensor model to the \emph{Nations} data.
-The multi-relational network $\tY$ is a binary data tensor, and the country attributes $\mX\in\mathbb{R}^{14\times 6}$ are features on both the 1$^\text{st}$ and 2$^\text{nd}$ modes.
-The BIC criterion suggests a rank $\mr=(4,4,4)$ for the coefficient tensor $\tB\in\mathbb{R}^{6\times 6 \times 56}$.
-We perform the supervised tensor decomposition and obtain the dimension reduction matrices $\hat \mM_k$ from the model.
-Then we apply $K$-mean clustering to dimension reduction matrix on each of the modes.
-Table~\ref{fig:s1} shows the K-means clustering of the 56 relations based on the dimension reduction matrix on the 3$^\text{rd}$ mode.
-We find that the relations reflecting the similar aspects of actions are grouped together.
-In particular, Cluster I consists of military relations such as \emph{violentactions}, \emph{warnings} and \emph{militaryactions}; Clusters II and III capture the economic relations such as \emph{economicaid, booktranslations, tourism}; and Cluster IV represents the political alliance and territory relations.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=16cm]{coef.pdf}
-\caption{Estimated feature effects in the \emph{Nations} data analysis.
-Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warning}, \emph{violentactions}, \emph{treaties}, and \emph{aidenemy}, respectively.
-}\label{fig:est}
-\vspace{-.4cm}
-\end{figure}
-
-\begin{table}[!h]
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{c|c|l}
- & Category & \multicolumn{1}{c}{Relations}\\
-\hline
-\multirow{2}{*}{ I}  & \multirow{2}{*}{Military} & warning, violentactions, militaryactions, duration, negativebehavior, protests, severdiplomatic \\
-&&timesincewar, commonbloc0, commonbloc1, blockpositionindex,  expeldiplomats\\   
-\hline
-\multirow{2}{*}{II}& \multirow{2}{*}{Emigrant} & emigrants, emigrants3, relemigrants, accusation, nonviolentbehavior, ngoorgs, commonbloc2, intergovorgs3\\      
-&&releconomicaid, relintergovorgs, relngo,  students, relstudents, economicaid, negativecomm, militaryalliance \\
-\hline
-\multirow{3}{*}{ III} & \multirow{3}{*}{Economics} & treaties, reltreaties, officialvisits, exportbooks, relexportbooks, booktranslations, relbooktranslations\\
-&&boycottembargo, weightedunvote, unweightedunvote, reltourism, tourism, tourism3, exports, exports3  \\
-&&relexports, intergovorgs, ngo, embassy, reldiplomacy, timesinceally, independence, conferences, dependent\\
-\hline
-IV & Territory &aidenemy, lostterritory, unoffialacts, attackembassy\\
-\hline
-\end{tabular}
-}
-\caption{$K$-means clustering of relations based on dimension reduction on the 3$^\text{rd}$ mode. }\label{tab:s1}
-\vspace{-.4cm}
-\end{table}
-
-To investigate the effects of dyadic attributes towards connections, we depict the estimated coefficients $\hat \tB=\entry{\hat b_{ijk}}$ for several relation types (Figure~\ref{fig:est}).
-Note that the each entry $\hat b_{ijk}$ estimates the contribution, at the logit scale, of feature pair $(i,j)$ ($i$th feature for the ``sender'' country and $j$th feature for the ``receiver'' country) towards the connection of relation $k$.
-Several interesting findings emerge from the observation.
-We find that relations belonging to a same cluster tend to have similar feature effects.
-For example, the relations ``warning'' and ''violentactions'' are classified into Cluster I, and both exhibit similar effect patterns (Figures~\ref{fig:est}a-b).
-Moreover, the feature \emph{constitutional} has a strong effect in the relation ``violentactions'' and ``warning'', whereas the effect is weaker in the relation ``treaties''.
-The result is plausible because the constitutional attributes affect political actions more often than economical actions.
-The entries in $\tB$ are useful for revealing interaction effects in a context-specific way.
-%For a given relation $k$, the off-diagonal entries in the matrix $\hat \tB(\colon,\colon,k)$ represent the pairwise interactions between features. 
-From Figure~\ref{fig:est}, we find a strong interaction between \emph{geography} and \emph{political leadership} in the relation ``warning'', and a strong interaction between \emph{geogrphy} and \emph{medicine ngo} in the relation ``aidenemy''.
-The relation-specific effect pattern showcases the applicability of our method in revealing complex interactions.
-
-\section{Discussion and future work}\label{sec:con}
-We have developed a supervised tensor decomposition method with side information on multiple modes.
-One important challenge of tensor data analysis is the complex interdependence among tensor entries and between multiple features.
-Our approach incorporates side information as feature matrices in the conditional mean tensor.
-The empirical results demonstrate the improved interpretability and accuracy over previous approaches.
-Applications to the brain connection and political relationship datasets yield conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.
-
-%We comment on a number of possible extensions. 
-We have provided accuracy guarantees for parameter estimation in the supervised tensor model.
-Statistical inference based on tensor decomposition is an important future direction.
-Measures of uncertainty, such as confidence envelope for space estimation, would be useful.
-One possible approach would be performing parametric bootstrap~\citep{tibshirani1993introduction} to assess the uncertainty in the estimation.
-%For example, one can simulate tensors from the fitted low-rank model based on the estimates, and then assess the empirical distribution of the estimates. 
-While being simple, bootstrap approach is often computationally expensive for large-scale data.
-Another possibility is to leverage recent development in debiased inference with distributional characterization~\citep{chen2019inference,xia2019confidence}.
-This approach has led to fruitful results for matrix data analysis.
-Uncertainly quantification involving tensors are generally harder than matrices, and establishing distribution theory for tensor estimation remains an open problem.
-
-One assumption made by our method is that tensor entries are conditionally independent given the linear predictor $\Theta$.
-This assumption can be extended by introducing a more general mixed-effect tensor model.
-For example, in the special case of Gaussian model, we can model the first two moments of data tensor using
-\begin{align}\label{eq:extention}
-\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)&=\tC\times\mM_1\times \cdots \times \mM_K\ \ \text{and}\ \ \text{Var}(\tY|\mX_1,\ldots,\mX_K)=\mPhi_1\otimes \cdots \otimes \mPhi_K,
-\end{align}
-where $\mPhi_k\in\mathbb{R}^{d_k\times d_k}$ is the unknown covariance matrix on the mode $k\in[K]$.
-For general exponential family, an additional mean-variance relationship should also be considered.
-The joint estimation of mean model $\Theta$ and variance model $\mPhi_k$ will lead to more efficient estimation in the presence of unmeasured confounding effects.
-However, the introduction of unknown covariance matrices $\mPhi_k$ dramatically increases the number of parameters in the problem.
-Suitable regularization such as graphical lasso or specially-structured covariance assumptions should be considered.
-The extension of tensor modeling with heterogeneous mixed-effects will be an important future direction.
-
-Although we have presented the supervised tensor decomposition in the context of order-3 data tensors, the framework applies to more general multi-way datasets.
-One possible extension is the integrative analysis of omics data, in which multiple types of omics measurements (gene expression, DNA methylation, microRNA) are collected in the same set of individuals~\citep{lock2013joint}.
-In such a case, the observation is a stack of data matrices, which may not be of the same dimension.
-%One possible extension is to impose linked dimension reduction among different data matrices, and incorporate side information as features in the joint decomposition. 
-Other applications include time-series tensor data with multiple side information.
-Exploiting the benefits and properties of the supervised tensor decomposition in specialized task will boost scientific discoveries.
-
-% for blind version
-%\section*{Acknowledgements}
-%This research was supported by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
-
-\bigskip
-\begin{center}
-{\large \bf SUPPLEMENTARY MATERIALS}
-\end{center}
-
-\begin{description}
-%\item[Supplementary notes:]  technical proofs and additional simulation results.
-\item[Data and software:] %R-package \texttt{tensorregress} performs our supervised tensor decomposition method. 
-Our simulation code, R-package \texttt{tensorregress}, and datasets used in the paper are available at \url{https://CRAN.R-project.org/package=tensorregress} 
-%We also provide R scripts for reproducing all numbers in Figures 2-8 and Tables 2-3 in the paper.
-\end{description}
-
 \bibliography{tensor_wang}

### 2020-10-02 tensor_regress/NeurIPSw/neuripsw_tensor.tex
workshop
revised abstract

@@
    We develop a tensor decomposition method that incorporates multiple side information as interactive features.
-Unlike unsupervised tensor decomposition, our supervised decomposition captures the effective dimension reduction of the data tensor confined to feature space on each mode.
+Unlike classical tensor decomposition, our supervised decomposition captures the effective dimension reduction of the data tensor confined to feature space on each mode.
 An efficient alternating optimization algorithm is further developed.
@@
 
-In addition to the aforementioned challenges, many tensor datasets consist of non-Gaussian measurements.
+In addition to the challenge of incorporating side information, many tensor datasets consist of non-Gaussian measurements.
 Classical tensor decomposition methods are based on minimizing the Frobenius norm of deviation, leading to suboptimal predictions for binary- or count-valued response variables.
@@
 We call $\mX_k\mM_k$ the ``supervised tensor factors'' or ``sufficient features''~\citep{adragni2009sufficient}, and call $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes, which links the conditional mean to the feature spaces, and thereby allows the identification of variations in the tensor data attributable to the side information.
-Our goal is to find $\mM_k$ and the corresponding $\tC$.
+Our goal is to find $\mM_k$ and the corresponding $\tC$,thereby allowing us to reveal the relationship between side information $\mX$ and the observed tensor $\tY$.
 Note that $\mM_k$ and $\tC$ are identifiable only up to orthonormal transformations.
@@
 
+\newpage 
+\section*{Broader Impact}
+
+Our supervised tensor decomposition method is widely applicable to network analysis, dyadic data analysis, spatial-temporal model, and recommendation systems.
+We have shown the improved predictive power and enhanced interpretability by incorporating the interactive side information in tensor decomposition method.
+The application to the brain connection dataset yields conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.
+We believe that our model enriches the research of tensor-based learning and is a powerful tool to boost scientific discoveries in various fields.
+
+%We have developed a supervised tensor decomposition method with side information on multiple modes.
+One important challenge of tensor data analysis is the complex interdependence among tensor entries and between multiple features.
+Our approach incorporates side information as feature matrices in the conditional mean tensor.
+Our empirical results demonstrate the improved interpretability and accuracy over previous approaches.
+The application to the brain connection dataset yields conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.
+We believe that our model fills the gap between theory and practice and is a powerful tool to boost scientific discoveries.
+
+%Our exponential family tensor regression  We have shown the improved predictive power and enhanced interpretability by incorporating interactions between multiple covariate matrices.
+We acknowledge the opportunities and challenges in the tensor regression modeling.
+We hope the method developed in this paper will open up new research directions towards richer tensor-based learning.
+All code and data are publicly available, and we encourage members of the machine learning community to participate in these exciting problems.
+
 \bibliography{tensor_wang}

### 2020-10-02 tensor_regress/NeurIPSw/neuripsw_tensor_v2.tex
update

--- 2020-10-02 tensor_regress/NeurIPSw/neuripsw_tensor.tex
@@
 \begin{abstract}
-   We develop a tensor decomposition method that incorporates multiple side information as interactive features.
-Unlike classical tensor decomposition, our supervised decomposition captures the effective dimension reduction of the data tensor confined to feature space on each mode.
+We consider the problem of tensor decomposition with multiple side information available as interactive features.
+Such problems are common in neuroimaging, network modeling, and spatial-temporal analysis.
+We develop a new family of exponential tensor decomposition models and establish the theoretical accuracy guarantees.
 An efficient alternating optimization algorithm is further developed.
-Our proposal handles a broad range of data types, including continuous, count, and binary observations.
-We apply the method to diffusion tensor imaging data from human connectome project.
-We identify the key global brain connectivity pattern and pinpoint the local regions that are associated with available features.
+Unlike earlier methods, our proposal handles a broad range of data types, including continuous, count, and binary observations, along with available features.
+We apply the method to diffusion tensor imaging data from human connectome project and identify the key brain connectivity patterns associated with available features.
 Our method will help the practitioners efficiently analyze tensor datasets in various areas.
+Toward this end, all data and code has been made available to the public.
 \end{abstract}
@@
 The brain connectivity networks are collected from a sample of individuals, accompanied by individual characteristics such as age, gender, and diseases status (see Figure~\ref{fig:intro1}a).
-Another example is in network analysis~\citep{pmlr-v108-berthet20a,hoff2005bilinear}.
-A typical social network consists of nodes that represent people and edges that represent  friendships.
+Another example is in network analysis~\citep{pmlr-v108-berthet20a}.
 Side information such as people’s demographic information and friendship types are often available.
-In both examples, it is of keen scientific interest to identify the variation in the data tensor (e.g., brain connectivities, social community patterns) that is affected by available features.
+In both examples, scientists are interested in identifing the variation in the data tensor (e.g., brain connectivities, social community patterns) that is affected by available features.
 These seemingly different scenarios pose a common yet challenging problem for tensor data modeling.
@@
  \begin{figure}[hbt]
+ \vspace{-.4cm}
 \begin{center}
 \includegraphics[width = 13.5cm]{demo.pdf}
-\vspace{-.5cm}
+\vspace{-.4cm}
 \end{center}
 \caption{Examples of supervised tensor decomposition with interactive side information. (a) Network population model. (b) Spatio-temporal growth model.} \label{fig:intro1}
-\vspace{-.3cm}
+\vspace{-.2cm}
 \end{figure}
@@
 
-We present a general model and associated method for decomposing a data tensor whose entries are from exponential family with interactive side information.
-We formulate the learning task as a structured regression problem, with tensor observation serving as the response, and the multiple side information as interactive features.
-We leverage generalized linear model (GLM)  to allow heteroscedacity due to the mean-variance relationship in the non-Gaussian data.
-The low-rank structure on the conditional mean tensor effectively mitigates the curse of high dimensionality.
-Our proposal blends the modeling power of GLM and the exploratory capability of tensor dimension reduction in order to take the best out of both worlds.
+{\bf Our contribution.} This paper presents a general model and associated method for decomposing a data tensor whose entries are from exponential family with interactive side information.
+We formulate the learning task as a low-rank tensor regression problem, with tensor observation serving as the response, and the multiple side information as interactive features.
+We blend the modeling power of generalized linear model (GLM) and the exploratory capability of tensor dimension reduction in order to take the best out of both worlds.
+Our methods greatly improves the classical tensor decomposition, and we quantify the gain in prediction through numerical experiments and data applications.
+
+{\bf Notation.} We follow the tensor notation as in~\cite{kolda2009tensor}.
+The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as $\tY\times\{\mX_1,\ldots,\mX_K\} =\entry{\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}x_{j_1,i_1}^{(1)}\cdots x_{j_K,i_K}^{(K)}}$,
+which results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor.
+The inner product between two tensors of equal size is defined as $\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}$.
+For ease of notation, we allow basic arithmetic operators (e.g.,\ $+, -$) and univariate functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner.
+
 %A low-rank structure is imposed to the conditional mean of tensor observation.where unlike classical decomposition, the tensor factors $\mX_k\mM_k\in\mathbb{R}^{d_k\times r_k}$ belong to the space spanned by features $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,2,3$.
@@
 
-\vspace{-0.3cm}
-
-\section{Method}
-
-\subsection{Preliminary}
-We introduce the basic tensor properties used in the paper.
-We use lower-case letters (e.g.,\ $a,b,c$) for scalars and vectors, upper-case boldface letters (e.g.,\ $\mA, \mB, \mC$) for matrices, and calligraphy letters (e.g.,\ $\tA, \tB, \tC$) for tensors of order three or greater.
-Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
-The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as $\tY\times\{\mX_1,\ldots,\mX_K\} =\entry{\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}x_{j_1,i_1}^{(1)}\cdots x_{j_K,i_K}^{(K)}}$,
-which results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor.For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as $\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.$
-The tensor Frobenius norm is defined as $\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}$, and the maximum norm is defined $\mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.$  We let $\mI_d$ denote the $d \times d$ identity matrix and $[d]$ denote the $d$-set $\{1,\ldots,d\}$.
+\section{Proposed models and motivating examples}
 %For any two matrices $\mA,\mB$ with orthonormal columns of same dimension, the angle distance is defined as
@@
 \]
-
-\subsection{General Model}
-
 Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor.
@@
 Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the feature matrix on the mode $k\in[K]$, where $x_{ij}$ denotes the $j$-th feature value for the $i$-th tensor entity, for $(i,j)\in[d_k]\times[p_k]$, $p_k\leq d_k$.
-Assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent realizations from an exponential family distribution, and the conditional mean tensor admits the form
-\begin{align}\label{eq:tensormodel}
-&\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)=f(\Theta),\quad \text{with}\quad \Theta =\tB\times\{\mX_1,\ldots, \mX_K\} ,
-\end{align}
-where $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the unknown parameter tensor, $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$.
-The choice of link function is based on the assumed distribution family of tensor entries.
-
-In classical tensor decomposition, tensor factorization is performed on either data tensor $\tY$ or mean tensor $\mathbb{E}(\tY)$.
-In the context of supervised tensor decomposition, we propose to factorize the latent parameter tensor $\tB$,
-\begin{equation}\label{eq:factor}
-\tB=\tC\times\{\mM_1,\ \ldots,\ \mM_K\},
-\end{equation}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, and $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices consisting of orthonormal columns, where $r_k\leq p_k$ for all $k\in[K]$.
-By the definition of multilinear rank, model equations~\eqref{eq:tensormodel} and~\eqref{eq:factor} imply the low-rankness $\mr=(r_1,\ldots,r_K)$ of the conditional mean tensor under the link function.
-We now reach our final model for supervised tensor decomposition,
+We assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent realizations from an exponential family distribution, and the conditional mean tensor admits the form
 \begin{align}\label{eq:decomp}
 \mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1, \ldots, \mX_K\mM_K\}\right),\notag\\
-\text{with} \ &\ \mM_k^T\mM_k = \mI_{r_k},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\quad \text{for all }k=1,\ldots,K,
+\text{with} \ &\ \mM_k^T\mM_k = \mI_{r_k},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\quad \text{for all }k=1,\ldots,K.
 \end{align}
-where the parameters of interest are $\mM_k$ and $\tC$.
-Note that model~\eqref{eq:decomp} assumes a fixed, known rank $\mr=(r_1,\ldots,r_K)$.
+where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, and $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices consisting of orthonormal columns, where $r_k\leq p_k$ for all $k\in[K]$, $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$.
+Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution.
+
 Figure~\ref{fig:intro1}b provides a schematic illustration of our model.
 The features $\mX_k$ affect the distribution of tensor entries in $\tY$ through the form $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$.
-We call $\mX_k\mM_k$ the ``supervised tensor factors'' or ``sufficient features''~\citep{adragni2009sufficient}, and call $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes, which links the conditional mean to the feature spaces, and thereby allows the identification of variations in the tensor data attributable to the side information.
-Our goal is to find $\mM_k$ and the corresponding $\tC$,thereby allowing us to reveal the relationship between side information $\mX$ and the observed tensor $\tY$.
+The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes, which links the conditional mean to the feature spaces, and thereby allows the identification of variations in the tensor data attributable to the side information.
+Our goal is to find $\mM_k$ and the corresponding $\tC$, thereby allowing us to reveal the relationship between side information $\mX_k$ and the observed tensor $\tY$.
 Note that $\mM_k$ and $\tC$ are identifiable only up to orthonormal transformations.
 
-\vspace{-.3cm}
-\section{Estimation}
-\subsection{Rank-constrained M-estimator}
- We adopt the exponential family as a flexible framework for different data types. In a classical generalized linear model with a scalar response $y$ and feature $\mx$, the density is expressed as
+We give two examples of supervised tensor decomposition models~\eqref{eq:decomp} that arise in practice.
+\begin{example}[Spatio-temporal growth model]
+The growth curve model~\citep{srivastava2008models} was originally proposed as an example of bilinear model for matrix data, and we extend it to higher-order cases.
+Let $\tY=\entry{y_{ijk}}\in\mathbb{R}^{d \times m\times n}$ denote the pH measurements of $d$ lakes at $m$ levels of depth and for $n$ time points.
+Suppose the sampled lakes belong to $q$ types, with $p$ lakes in each type.
+Let $\{\ell_j\}_{j\in[m]}$ denote the sampled depth levels and $\{t_k\}_{k\in[n]}$ the time points.
+Assume that the expected pH trend in depth is a polynomial of order at most $r$ and that the expected trend in time is a polynomial of order $s$.
+Then, the conditional mean model for the spatio-temporal growth is a special case of our model~\eqref{eq:decomp}, where $\mX_1=\text{blockdiag}\{\mathbf{1}_q,\ldots,\mathbf{1}_q\}\in \{0,1\}^{d\times p}$ is the design matrix for lake types,
 \[
-p(y|\mx, \boldsymbol{\beta})=c(y,\phi)\exp\left(\frac{y\theta- b(\theta)}{\phi}\right)\ \text{with}\ \theta=\boldsymbol{\beta}^T\mx,
+\mX_2=
+\begin{pmatrix}
+1 & \ell_1&\cdots &\ell^{r}_1\\
+1 & \ell_2&\cdots &\ell^{r}_2\\
+\vdots &\vdots&\ddots&\vdots\\
+1&\ell_{m}&\cdots&\ell^{r}_{m}
+\end{pmatrix},\
+\mX_3=
+\begin{pmatrix}
+1 & t_1&\cdots &t^{s}_1\\
+1 & t_2&\cdots &t^{s}_2\\
+\vdots &\vdots&\ddots&\vdots\\
+1&t_{n}&\cdots&t^{s}_{n}
+\end{pmatrix}
 \]
-where $b(\cdot)$ is a known function depending on the data types, $\theta$ is the linear predictor, $\phi>0$ is the dispersion parameter, and $c(\cdot)$ is a known normalizing function.
-In our context, we model the tensor entries $y_{i_1,\ldots,i_K}$, conditional on $\theta_{i_1,\ldots,i_K}$, as independent draws from an exponential family.
-Ignoring constants that do not depend on $\Theta$, the quasi log-likelihood of~\eqref{eq:decomp} is equal to Bregman distance between $\tY$ and $b'(\Theta)$:
+are the design matrices for spatial and temporal effects, respectively. The spatial-temporal mode has covariates available on each of the three modes.
+
+\begin{example}[Network population model] 
+Network response model is recently developed in the context of neuroimanig analysis.
+The goal is to study the relationship between network-valued response and the individual covariates.
+Suppose we observe $n$ i.i.d.\ observations $\{(\mY_i, \mx_i): i=1,\ldots,n\}$, where $\mY_i\in\{0,1\}^{d\times d}$ is the brain connectivity network on the $i$-th individual, and $\mx_i\in\mathbb{R}^p$ is the individual covariate such as age, gender, cognition, etc.
+The network-response model~\citep{rabusseau2016low} has the form
+\begin{equation}\label{eq:network}
+\text{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n
+\end{equation}
+where $\tB\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor of interest.
+The model~\eqref{eq:network} is also a special case of our tensor-response model, with covariates on the last mode of the tensor.
+
+\section{Estimation algorithms}
+We develop a likelihood-based procedure to estimate $\tC$ and $\mM_k$ in~\eqref{eq:decomp}. Ignoring constants that do not depend on $\Theta$, the quasi log-likelihood of~\eqref{eq:decomp} is equal to
 \begin{equation}\label{eq:loglikelihood}
 \small
-\tL_{\tY}(\tC,\mM_1,\ldots,\mM_K)=\langle \tY, \Theta \rangle - \sum\nolimits_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K})\ \text{with}\ \Theta=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
+\tL_{\tY}(\tC,\mM_1,\ldots,\mM_K)=\langle \tY, \Theta \rangle - \sum\nolimits_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K})\ \text{with}\ \Theta=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\},
 \end{equation}
+where $b(\theta)=\theta^2/2$ for Gaussian response, $b(\theta)=\exp(\theta)$ for Poisson response, and $b(\theta)=\log(1+\exp(\theta))$ for Bernoulli response.
 We propose a constrained maximum quasi-likelihood estimator (M-estimator),
@@
 $
-The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to avoid the divergence in the non-Gaussian variance.
-
-\subsection{Alternating optimization}
 
@@
 
-\subsection{Statistical properties}\label{subsec:statprob}
-
-In modern applications, the tensor data and features are often large-scale.
-We are particularly interested in the high-dimensional regime in which both $d_k$ and $p_k$ diverge; i.e.\ $d_k\to \infty$ and $p_k\to\infty$, while $p_k/d_k \to \gamma_k \in[0,1)$.
-As the size of problem grows, and so does the number of unknown parameters.
-The classical MLE theory does not directly apply.
-We leverage the recent development in random tensor theory and high-dimensional statistics to establish the error bounds.
-
-\begin{thm}[Statistical convergence]\label{thm:main}
-Consider a data tensor generated from model~\eqref{eq:decomp}.
+We provide the accuracy guarantee for the proposed M-estimator~\eqref{eq:MLE} by leveraging recent development in random tensor theory and high-dimensional statistics.
+\begin{thm}[Convergence]\label{thm:main}
 Let $(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)$ be the M-estimator in~\eqref{eq:MLE} and $\hat \tB=\hat \tC\times \hat \mM_1\times\cdots \times \hat \mM_K$.
 Define $r_\textup{total}=\prod_k r_k$ and $r_{\max}=\max_k r_k$.
-Under mild technical assumptions, there exist two positive constants $C_1=C_1(\alpha,K), C_2=C_2(\alpha, K)>0$ independent of dimensions $\{d_k\}$ and $\{p_k\}$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$,
+Under mild technical assumptions, there exist two positive constants $C_1, C_2>0$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$,
 \begin{equation}\label{eq:bound}
-    \FnormSize{}{\trueB- \hat \tB}^2\leq \frac{ C_{2} r_{\textup{total}}}{r_{\max}}\frac{\sum_k p_k}{\prod_k d_k}.
+    \FnormSize{}{\trueB- \hat \tB}^2\leq \frac{ C_{2} r_{\textup{total}}}{r_{\max}}\frac{\sum_k p_k}{\prod_k d_k},\quad \text{and}\quad
+\sin^2 \Theta(\trueM,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\sigma^2_{\min}(\textup{Unfold}_k(\tC_\textup{true}))}\frac{\sum_k p_k}{\prod_k d_k},
 \end{equation}
-Furthermore, if the unfolded core tensor has non-degenerate singular values at mode $k\in[K]$, i.e., $\sigma_{\min}(\textup{Unfold}_k(\tC_\textup{true})) \geq c>0$ for some constant $c$, then
-\begin{equation}\label{eq:sinebound}
-\textup{sin}^2 \Theta(\trueM,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\sigma^2_{\min}(\textup{Unfold}_k(\tC_\textup{true}))}\frac{\sum_k p_k}{\prod_k d_k},
-\end{equation}
-where $
-\sin \Theta(\trueM,\hat \mM_k)=\snormSize{}{\trueM^T \hat \mM_k^{\perp}}=\max\left\{ \frac{\langle \mx, \my\rangle}{\vnormSize{}{\mx}\vnormSize{}{\my}}\colon \ \mx\in \textup{Span}(\trueM),\ \my\in \textup{Span}({\hat \mM_k}^\perp)\right\}
-$ is the angle distance to assess the accuracy in estimating the column space, $\textup{Span}(\mM_k)$.
+where $\sin \Theta(\trueM,\hat \mM_k)=\snormSize{}{\trueM^T \hat \mM_k^{\perp}}$ is the angle distance between column spaces. 
 \end{thm}
+Theorem~\ref{thm:main} implies that the estimation has a convergence rate $\tO(d^{-(K-1)})$ in the special case when tensor dimensions are equal on each of the modes, i.e., $d_k=d$ for all $k\in[K]$, and feature dimension grows with tensor dimension, $p_k=\gamma d$, $\gamma\in [0,1)$, for $k\in[K]$.
+The convergence of our estimation becomes especially favorable as the order of tensor data increases.
 
-\section{Real data analysis}
+\section{Numerical experiments}
+We compare our supervised tensor decomposition ({\bf STD}) with three other supervised tensor methods: Higher-order low-rank regression ({\bf HOLRR}~\cite{rabusseau2016low}), Higher-order partial least square ({\bf HOPLS}~\cite{zhao2012higher}) and Subsampled tensor projected gradient ({\bf TPG}~\cite{yu2016learning}).
+Figure~\ref{fig:compare} shows that {\bf STD} outperforms others, especially in the low-signal, high-rank setting.
+As the number of informative modes (i.e., modes with available features) increases, the {\bf STD} exhibits a substantial reduction in error whereas others remain unchanged (Figure~\ref{fig:compare}b).
+This showcases the benefit of incorporation of multiple features.
+The accuracy gain in Figure~\ref{fig:compare} demonstrates the benefit of alternating algorithm -- incorporation of informative modes also improves the estimation in the non-informative modes.
 
-The Human Connectome Project (HCP) aims to build a network map that characterizes the anatomical and functional connectivity within healthy human brains~\citep{HCP}. 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=14cm]{compare_alternative_w.pdf} 
+\vspace{-.5cm}
+\caption{\small Comparison between different tensor methods.
+Panels (a) and (b) plot mean squared prediction error (MSPE) versus the number of modes with available features.
+Panels (c) and (d) plot MSPE versus the effective sample size $d^2$.
+We consider rank $\mr=(3,3,3)$ (low) vs (4,5,6)
+ (high), and signal $\alpha =3 $ (low) vs.\ 6 (high).}~\label{fig:compare}
+\vspace{-.5cm}
+\end{figure}
+
+We then apply our method to brain structural connectivity networks from Human Connectome Project (HCP)~\citep{HCP}.
 %We analyze the structural connectivity patterns among 68 brain regions for 136 individuals from HCP.
-We follow the preprocessing procedure as in~\cite{zhang2018mapping} and parcellate the brain into 68 regions of interest~\citep{desikan2006automated}.
 The dataset consists of 136 brain structural networks, one for each individual.
@@
 \includegraphics[width=14cm]{HCP.pdf}
-\caption{Top edges with large effects.
+\vspace{-.5cm}
+\caption{\small Top edges with large effects.
 (a) Global effect; (b) Female effect; (c) Age 22-25; (d) Age 31+.
-Red edges represent positive effects and blue edges represent negative effects.
-The edge-width is proportional to the magnitude of the effect size.
+Red (blue) edges represent positive (negative) effects.
+Edge-widths are proportional to the magnitudes of effect sizes.
 }\label{fig:brain}
-%\vspace{1cm}
+\vspace{-.2cm}
 %\includegraphics[width=14cm]{compare_HCP.pdf}
@@
 
-We perform the supervised tensor decomposition to the HCP data. %The data tensor is binary, $\tY\in\{0,1\}^{68\times 68\times 136}$, and the features are of dimension 4 on the 3$^{\text{rd}}$ mode. 
+We perform the supervised tensor decomposition to the HCP data.
 The BIC selection suggests a rank $\mr=(10,10,4)$ with quasi log-likelihood $\tL_{\tY}=-174654.7$.
-We utilize the sum-to-zero contrasts in coding the feature effects, and depict only the top 3\% edges whose connections are non-constant across the sample.
 Figure~\ref{fig:brain} shows the top edges with high effect size, overlaid on the Desikan atlas brain template~\citep{desikan2006automated}.
@@
 Those edges involve Frontal-pole (\emph{Fploe}), superior-frontal (\emph{SupF}) and Cuneus nodes.
-The Frontal-pole region is known for its importance in memory and cognition, and the detected decline with age further highlights its biological importance.
+Our results highlight the importance of Frontal-pole region, and the detected decline further suggests the age effects to brain connections.
+
+\section{Conclusion}
+\vspace{-.2cm}
+We have developed a supervised tensor decomposition method with side information on multiple modes.
+The empirical results demonstrate the improved interpretability and accuracy over previous approaches.
+Applications to the brain connection data yield conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.
 
@@
 The application to the brain connection dataset yields conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.
+Tensor learning is a clear challenge for further research.
 We believe that our model enriches the research of tensor-based learning and is a powerful tool to boost scientific discoveries in various fields.
+We hope the work opens up new inquiry that allows more machine learning researchers to contribute to this field.
 

### 2020-10-03 tensor_regress/NeurIPSw/neuripsw_tensor_v3.tex
v3

--- 2020-10-02 tensor_regress/NeurIPSw/neuripsw_tensor_v2.tex
@@
 \end{align}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, and $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices consisting of orthonormal columns, where $r_k\leq p_k$ for all $k\in[K]$, $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$.
+where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices consisting of orthonormal columns with $r_k\leq p_k$ for all $k\in[K]$, and $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$.
 Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution.
@@
 
+\end{example}
 \begin{example}[Network population model] 
@@
 The model~\eqref{eq:network} is also a special case of our tensor-response model, with covariates on the last mode of the tensor.
+\end{example}
 

### 2020-11-13 tensor_regress/NeurIPSw/camera_ready/neuripsw_tensor.tex
camera ready

--- 2020-10-03 tensor_regress/NeurIPSw/neuripsw_tensor_v3.tex
@@
 % ready for submission
- \usepackage{neurips_2020}
-
+% \usepackage{neurips_2020}
+\usepackage[final]{neurips_2020_ml4ps}
 % to compile a preprint version, e.g., for submission to arXiv, add add the
@@
 
+\author{%
+  Jiaxin Hu \\
+  Department of Statistics\\
+  University of Wisconsin-Madison\\
+  Madison, WI 53706\\
+  \texttt{jhu267@wisc.edu} \\
+
+  \And
+  Chanwoo Lee \\
+  Department of Statistics\\
+  University of Wisconsin-Madison\\
+  Madison, WI 53706\\
+  \texttt{chanwoo.lee@wisc.edu} \\
+
+  \And
+
+  Miaoyan Wang \\
+  Department of Statistics\\
+  University of Wisconsin-Madison\\
+  Madison, WI 53706\\
+  \texttt{miaoyan.wang@wisc.edu} \\
+
+}
+
 \begin{document}
@@
 Our method will help the practitioners efficiently analyze tensor datasets in various areas.
-Toward this end, all data and code has been made available to the public.
+Toward this end, all data and code have been made available to the public.
 \end{abstract}
@@
 For ease of notation, we allow basic arithmetic operators (e.g.,\ $+, -$) and univariate functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner.
+Besides, let $\otimes$ denote the Kronecker product of matrices.
 
-%A low-rank structure is imposed to the conditional mean of tensor observation.where unlike classical decomposition, the tensor factors $\mX_k\mM_k\in\mathbb{R}^{d_k\times r_k}$ belong to the space spanned by features $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,2,3$.
-The unknown matrices $\mM_k\in\mathbb{R}^{p_k\times r_k}$ (referred to as ``dimension reduction matrices'') link the conditional mean to the feature spaces, thereby allowing the identification of variations in the tensor data attributable to the side information.
-
- % This flexibility is important in practice.
-Furthermore, our low-rank model on the (transformed) conditional mean tensor effectively mitigates the curse of high dimensionality.
-In classical GLM, the sample size and feature dimension are well defined; however, in the tensor data analysis, we observe only one realization of an order-$K$ tensor and up to $K$ interactive feature matrices.
-Both the number of tensor entries and feature dimension grow exponentially in $K$.
-Dimension reduction is therefore crucial for prediction and interpretability.
-We establish the statistical convergence of our estimator, and we quantify the gain in prediction through simulations and case studies.
-
+\vspace{-0.2cm}
 \section{Proposed models and motivating examples}
+\vspace{-0.2cm}
 %For any two matrices $\mA,\mB$ with orthonormal columns of same dimension, the angle distance is defined as
@@
 
+\vspace{-0.2cm}
+
 \section{Estimation algorithms}
+
+\vspace{-0.2cm}
 We develop a likelihood-based procedure to estimate $\tC$ and $\mM_k$ in~\eqref{eq:decomp}. Ignoring constants that do not depend on $\Theta$, the quasi log-likelihood of~\eqref{eq:decomp} is equal to
@@
 $
-\tP=\left\{(\tC, \mM_1,\ldots,\mM_K) \ \Big| \ \mM_k^T\mM_k=\mathbf{I}_{r_k},\ \mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha \right\}.
-$
+\tP=\left\{(\tC, \mM_1,\ldots,\mM_K) \ \Big| \ \mM_k^T\mM_k=\mathbf{I}_{r_k},\ \mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha \right\}
+$, and $\alpha$ is a constant. The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to avoid the divergence in the non-Gaussian variance.
 
@@
 
-\begin{algorithm}[!h]
-\caption{Supervised Tensor Decomposition with Interactive Side Information (Simplified)}\label{alg:B}
+\begin{algorithm}[htb]
+\caption{Supervised Tensor Decomposition with Interactive Side Information}\label{alg:B}
 \begin{algorithmic}[1]
@@
 \While{Do until convergence}
-
-%\State Obtain the factor matrix $\tilde \mM_k\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$.
-%\State Perform QR factorization $\tilde \mM_k=\mQ\mR$.
-%\State Update $\mM_k\leftarrow \mQ$ and core tensor $\tC\leftarrow \tC\times_k \mR$.
-\State Obtain $\tilde \mM_k\in\mathbb{R}^{p_k\times r_k}$ by a GLM. Orthogonalize $\tilde \mM_k$ by QR factorization, for $k \in [K]$.
-
-\State Update the core tensor $\tC$ by solving a GLM. Rescale the core tensor $\tC$ such that $\maxnorm{\tC} \leq \alpha$. 
+\For { $k=1$ to $K$}
+\State Obtain the factor matrix $\tilde \mM_k\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$.
+\State Perform QR factorization $\tilde \mM_k=\mQ\mR$. Update $\mM_k\leftarrow \mQ$ and core tensor $\tC\leftarrow \tC\times_k \mR$.
+\EndFor
+\State Update the core tensor $\tC$ by solving a GLM with $\textup{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM_k]$ as features, with link function $f$.
+Rescale the core tensor $\tC$ such that $\mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha$. 
 \EndWhile
@@
 \section{Numerical experiments}
-We compare our supervised tensor decomposition ({\bf STD}) with three other supervised tensor methods: Higher-order low-rank regression ({\bf HOLRR}~\cite{rabusseau2016low}), Higher-order partial least square ({\bf HOPLS}~\cite{zhao2012higher}) and Subsampled tensor projected gradient ({\bf TPG}~\cite{yu2016learning}).
+
+We evaluate the empirical performance of our supervised tensor decomposition (STD) through simulations.
+We consider order-3 tensors, where the conditional mean tensor is generated form model~\eqref{eq:decomp}.
+% The core tensor entries are i.i.d.\ drawn from Uniform[-1,1], the factor matrix $\mM_k$ is uniformly sampled with respect to Haar measure from matrices with orthonormal columns.
+The feature matrix $\mX_k$ is either an identity matrix (i.e.,\ no feature  available) or Gaussian random matrix with i.i.d.\ entries from $N(0,1)$.
+The linear predictor $\Theta=\entry{\theta_{ijk}} = \tC\times\{\mM_1\mX_1,\mM_2\mX_3,\mM_3\mX_3\}$ is scaled such that $\mnormSize{}{\Theta}=1$.
+ Given the generated linear predictor $\Theta=\entry{\theta_{ijk}} = \tC\times\{\mM_1\mX_1,\mM_2\mX_3,\mM_3\mX_3\}$, the entries in the tensor $\tY=\entry{y_{ijk}}$ are drawn independently according to three probabilistic models: (a)  Gaussian model: $y_{ijk}\sim N\left( \theta_{ijk}, 1\right)$; (b)  Poisson model: $y_{ijk}\sim\textup{Poisson}\left( e^{ \theta_{ijk}}\right)$; (c) Bernoulli model:  $y_{ijk}\sim \textup{Bernoulli}\left( {e^{\theta_{ijk}}}/{1+e^{\theta_{ijk}}}\right)$.
+
+The experiment I evaluates the accuracy when covariates are available on all modes.
+We set $\alpha=10, d_k=d, p_k=0.4d_k, r_k=r\in\{2,4,6\}$ and increase $d$ from 25 to 50.
+Our theoretical analysis suggests that $\hat \tB$ has a convergence rate $\tO(d^{-2})$ in this setting.
+Figure~\ref{fig:dim} plots the estimation error versus the ``effective sample size'', $d^2$, under three different distribution models.
+We found that the empirical MSE decreases roughly at the rate of $1/d^2$, which is consistent with our theoretical ascertainment.
+%We also observed that, tensors with higher ranks tend to yield higher estimation errors, as reflected by the upward shift of the curves as $r$ increases.
+Indeed, a larger $r$ implies a higher model complexity and thus greater difficulty in the estimation.
+Similar behaviors can be observed in the non-Gaussian data in Figures~2b-c.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=14cm]{dim_alternative_w.pdf}
+\vspace{-.5cm}
+\caption{(a)-(c): Estimation error against effective sample size.
+The dashed curves correspond to $\tO({1/d^2})$.
+(d)-(f): Performance comparison under stochastic block models.
+The $x$-axis represents the number of blocks in the networks.
+The response tensors are generated from Gaussian (a, d), Poisson (b, e) and  Bernoulli  (d, f) models .
+}\label{fig:dim}
+\end{figure}
+
+The experiment II investigates the capability of our model in handling correlation among coefficients.
+We mimic the scenario of brain imaging analysis.
+A sample of $d_3=50$ networks are simulated, where each network measures the connections between $d_1=d_2=20$ brain nodes.
+We simulate $p=5$ features for the each of the 50 individuals.
+These features may represent, for example, age, gender, cognitive score, etc.
+Recent study has suggested that brain connectivity networks often exhibit community structure represented as a collection of subnetworks, and each subnetwork is comprised of a set of spatially distributed brain nodes.
+To accommodate this structure, we utilize the stochastic block model~\citep{abbe2017community} to generate the effect size.
+Specifically, we partition the nodes into $r$ blocks by assigning each node to a block with uniform probability.
+Edges within a same block are assumed to share the same feature effects, where the effects are drawn i.i.d.\ from $N(0,1)$.
+
+Figure~\ref{fig:dim}(d)-(f) compares the MSE of our method with a multiple-response GLM approach.
+The multiple-response GLM is to regress the dyadic edges, one at a time, on the features, and this model is repeatedly fitted for each edge.
+As we find in Figure~\ref{fig:dim}(d)-(f), our tensor regression method achieves significant error reduction in all three data types considered.
+The outperformance is substantial in the presence of large communities; even in the less structured case ($\sim 20/15=1.33$ nodes per block), our method still outer-performs GLM.
+The possible reason is that the multiple-response GLM approach does not account for the correlation among the edges, and suffers from overfitting.
+In contrast, the low-rankness in our modeling incorporates the shared information across entries.
+
+The experiment III compares {\bf STD} with three other supervised tensor methods: Higher-order low-rank regression ({\bf HOLRR}~\cite{rabusseau2016low}), Higher-order partial least square ({\bf HOPLS}~\cite{zhao2012higher}) and Subsampled tensor projected gradient ({\bf TPG}~\cite{yu2016learning}).
 Figure~\ref{fig:compare} shows that {\bf STD} outperforms others, especially in the low-signal, high-rank setting.

### 2020-11-14 tensor_regress/NeurIPSw/camera_ready/neuripsw_tensor_c_v2.tex
v2

--- 2020-11-13 tensor_regress/NeurIPSw/camera_ready/neuripsw_tensor.tex
@@
 An efficient alternating optimization algorithm is further developed.
-Unlike earlier methods, our proposal handles a broad range of data types, including continuous, count, and binary observations, along with available features.
+Unlike earlier methods, our proposal is able to handle a broad range of data types, including continuous, count, and binary observations.
 We apply the method to diffusion tensor imaging data from human connectome project and identify the key brain connectivity patterns associated with available features.
@@
 
-In addition to the challenge of incorporating side information, many tensor datasets consist of non-Gaussian measurements.
+In addition to the challenge of incorporating side information, another challenge is that many tensor datasets consist of non-Gaussian measurements.
 Classical tensor decomposition methods are based on minimizing the Frobenius norm of deviation, leading to suboptimal predictions for binary- or count-valued response variables.
@@
 These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest.
-The gap between theory and practice means a great opportunity to modeling paradigms and better capture the complexity in tensor data.
+We overcome those challenges from new modeling and narrow the gap between theory and practice.
 
@@
 We formulate the learning task as a low-rank tensor regression problem, with tensor observation serving as the response, and the multiple side information as interactive features.
-We blend the modeling power of generalized linear model (GLM) and the exploratory capability of tensor dimension reduction in order to take the best out of both worlds.
-Our methods greatly improves the classical tensor decomposition, and we quantify the gain in prediction through numerical experiments and data applications.
+We blend the modeling power of generalized linear model (GLM) and the exploratory capability of tensor dimension reduction in order to take the best out of both sides.
+Our method greatly improves the classical tensor decomposition, and we quantify the improvement in prediction through numerical experiments and data applications.
 
@@
 Figure~\ref{fig:intro1}b provides a schematic illustration of our model.
-The features $\mX_k$ affect the distribution of tensor entries in $\tY$ through the form $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$.
-The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes, which links the conditional mean to the feature spaces, and thereby allows the identification of variations in the tensor data attributable to the side information.
-Our goal is to find $\mM_k$ and the corresponding $\tC$, thereby allowing us to reveal the relationship between side information $\mX_k$ and the observed tensor $\tY$.
+The features $\mX_k$ affect the distribution of tensor entries in $\tY$ through the sufficient features of the form $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$.
+The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes, which links the conditional mean to the feature spaces, and thus allows the identification of variations in the tensor data attributable to the side information.
+Our goal is to find $\mM_k$ and the corresponding $\tC$ to reveal the relationship between side information $\mX_k$ and the observed tensor $\tY$.
 Note that $\mM_k$ and $\tC$ are identifiable only up to orthonormal transformations.
@@
 
-\section{Estimation algorithms}
+\section{Estimation methods and theoretical results}
 
@@
 Our theoretical analysis suggests that $\hat \tB$ has a convergence rate $\tO(d^{-2})$ in this setting.
-Figure~\ref{fig:dim} plots the estimation error versus the ``effective sample size'', $d^2$, under three different distribution models.
+Figure~\ref{fig:dim}a plots the estimation error versus the ``effective sample size ($d^2$)'', under three different distribution models.
 We found that the empirical MSE decreases roughly at the rate of $1/d^2$, which is consistent with our theoretical ascertainment.
@@
 
-Figure~\ref{fig:dim}(d)-(f) compares the MSE of our method with a multiple-response GLM approach.
+Figure~\ref{fig:dim}d-f compares the MSE of our method with a multiple-response GLM approach.
 The multiple-response GLM is to regress the dyadic edges, one at a time, on the features, and this model is repeatedly fitted for each edge.
-As we find in Figure~\ref{fig:dim}(d)-(f), our tensor regression method achieves significant error reduction in all three data types considered.
-The outperformance is substantial in the presence of large communities; even in the less structured case ($\sim 20/15=1.33$ nodes per block), our method still outer-performs GLM.
+As we find in Figure~\ref{fig:dim}d-f, our tensor regression method achieves significant error reduction in all three data types considered.
+The outperformance is substantial in the presence of large communities; even in the less structured case ($\sim 20/15=1.33$ nodes per block), our method still outperforms GLM.
 The possible reason is that the multiple-response GLM approach does not account for the correlation among the edges, and suffers from overfitting.
@@
 
-We perform the supervised tensor decomposition to the HCP data.
+We appy the supervised tensor decomposition to the HCP data.
 The BIC selection suggests a rank $\mr=(10,10,4)$ with quasi log-likelihood $\tL_{\tY}=-174654.7$.
@@
 Those edges involve Frontal-pole (\emph{Fploe}), superior-frontal (\emph{SupF}) and Cuneus nodes.
-Our results highlight the importance of Frontal-pole region, and the detected decline further suggests the age effects to brain connections.
+The Frontal-pole region is known for its importance in memory and cognition, and the detected decline further suggests the age effects to brain connections.
 
@@
 Our supervised tensor decomposition method is widely applicable to network analysis, dyadic data analysis, spatial-temporal model, and recommendation systems.
-We have shown the improved predictive power and enhanced interpretability by incorporating the interactive side information in tensor decomposition method.
-The application to the brain connection dataset yields conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.
+The new method improve the predictive power and enhance interpretability through incorporation of the interactive side information in tensor decomposition.
+% We have shown the improved predictive power and enhanced interpretability by incorporating the interactive side information in tensor decomposition method.  
+The application to  brain connection dataset  shows the practical utility of the proposed method.
+%The application to the brain connection dataset yields conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.
 Tensor learning is a clear challenge for further research.
-We believe that our model enriches the research of tensor-based learning and is a powerful tool to boost scientific discoveries in various fields.
-We hope the work opens up new inquiry that allows more machine learning researchers to contribute to this field.
+ We believe that our model enriches the research of tensor-based learning and becomes a powerful tool to boost scientific discoveries in various fields. 
+%We hope the work opens up new inquiry that allows more machine learning researchers to contribute to this field.
 

### 2020-11-15 tensor_regress/NeurIPSw/camera_ready/neuripsw_tensor_c_v3.tex
v3

--- 2020-11-14 tensor_regress/NeurIPSw/camera_ready/neuripsw_tensor_c_v2.tex
@@
 
+\vspace{-0.5cm}
 \begin{abstract}
@@
 Our method will help the practitioners efficiently analyze tensor datasets in various areas.
-Toward this end, all data and code have been made available to the public.
+Toward this end, all data and code are available  at~\url{https://CRAN.R-project.org/package=tensorregress}.
 \end{abstract}
 
+\vspace{-0.2cm}
 \section{Introduction}
+\vspace{-0.2cm}
  Higher-order tensors have received increased attention across science and engineering.
@@
 \end{center}
-\caption{Examples of supervised tensor decomposition with interactive side information. (a) Network population model. (b) Spatio-temporal growth model.} \label{fig:intro1}
+\caption{\small Examples of supervised tensor decomposition with interactive side information. (a) Network population model. (b) Spatio-temporal growth model.} \label{fig:intro1}
 \vspace{-.2cm}
@@
 
-In addition to the challenge of incorporating side information, another challenge is that many tensor datasets consist of non-Gaussian measurements.
+In addition to the challenge of incorporating side information, another challenge is that many tensor datasets consist of non-Gaussian measurements \citep{wang2020learning, lee2020tensor} .
 Classical tensor decomposition methods are based on minimizing the Frobenius norm of deviation, leading to suboptimal predictions for binary- or count-valued response variables.
@@
 These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest.
-We overcome those challenges from new modeling and narrow the gap between theory and practice.
 
+\vspace{-.2cm}
 {\bf Our contribution.} This paper presents a general model and associated method for decomposing a data tensor whose entries are from exponential family with interactive side information.
@@
 The inner product between two tensors of equal size is defined as $\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}$.
+We use $\FnormSize{}{\cdot}$, $\mnormSize{}{\cdot}$, and $\text{vec}(\cdot)$ to denote tensor F-norm, tensor infinity norm, and the operation that reshapes the tensor into a vector, respectively.
+We also use $\onorm{\cdot}_{\sigma}$ and $\otimes$ to denote the matrix spectral norm and Kronecker product of matrices, respectively.
 For ease of notation, we allow basic arithmetic operators (e.g.,\ $+, -$) and univariate functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner.
-Besides, let $\otimes$ denote the Kronecker product of matrices.
 
@@
 The features $\mX_k$ affect the distribution of tensor entries in $\tY$ through the sufficient features of the form $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$.
-The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes, which links the conditional mean to the feature spaces, and thus allows the identification of variations in the tensor data attributable to the side information.
+The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes, and thus allows the identification of variations in the tensor data attributable to the side information.
 Our goal is to find $\mM_k$ and the corresponding $\tC$ to reveal the relationship between side information $\mX_k$ and the observed tensor $\tY$.
@@
 \end{equation}
-where $\tB\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor of interest.
+where $\tB\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor of interest, and $\times_3$ denotes the tensor-by-matrix multiplication along the 3rd mode \citep{kolda2009tensor}.
 The model~\eqref{eq:network} is also a special case of our tensor-response model, with covariates on the last mode of the tensor.
@@
 
-\section{Estimation methods and theoretical results}
+\section{Estimation methods with accuracy guarantees}
 
@@
 \tP=\left\{(\tC, \mM_1,\ldots,\mM_K) \ \Big| \ \mM_k^T\mM_k=\mathbf{I}_{r_k},\ \mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha \right\}
-$, and $\alpha$ is a constant. The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to avoid the divergence in the non-Gaussian variance.
+$, and $\alpha$ is a constant. The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to avoid the divergence of the non-Gaussian variance.
 
@@
 We provide the accuracy guarantee for the proposed M-estimator~\eqref{eq:MLE} by leveraging recent development in random tensor theory and high-dimensional statistics.
-\begin{thm}[Convergence]\label{thm:main}
-Let $(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)$ be the M-estimator in~\eqref{eq:MLE} and $\hat \tB=\hat \tC\times \hat \mM_1\times\cdots \times \hat \mM_K$.
-Define $r_\textup{total}=\prod_k r_k$ and $r_{\max}=\max_k r_k$.
+\begin{thm}[Statistical Convergence]\label{thm:main}
+Let $(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)$ be the M-estimator in~\eqref{eq:MLE} and $\hat \tB=\hat \tC\times \{ \hat \mM_1,\cdots, \hat \mM_K\}$.
+Define $r_\textup{total}=\prod_k r_k$, $r_{\max}=\max_k r_k$, and $\tB_{\text{true}} =  \tC\times \{ \mM_1,\cdots,\mM_K\}$.
 Under mild technical assumptions, there exist two positive constants $C_1, C_2>0$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$,
@@
 \end{thm}
-Theorem~\ref{thm:main} implies that the estimation has a convergence rate $\tO(d^{-(K-1)})$ in the special case when tensor dimensions are equal on each of the modes, i.e., $d_k=d$ for all $k\in[K]$, and feature dimension grows with tensor dimension, $p_k=\gamma d$, $\gamma\in [0,1)$, for $k\in[K]$.
-The convergence of our estimation becomes especially favorable as the order of tensor data increases.
+\vspace{-0.1cm}
+Theorem~\ref{thm:main} implies that the estimation has a convergence rate $\tO(d^{-(K-1)})$ in the special case when tensor dimensions are equal on every mode, i.e., $d_k=d$ for all $k\in[K]$, and feature dimension grows with tensor dimension, $p_k=\gamma d$, $\gamma\in [0,1)$, for $k\in[K]$.
+The convergence of our estimation becomes favorable as the order of tensor data increases.
+The proof is provided in the \cite{xu2019generalized}.
 
+\vspace{-0.5cm}
 \section{Numerical experiments}
-
-We evaluate the empirical performance of our supervised tensor decomposition (STD) through simulations.
+\vspace{-0.2cm}
+We evaluate the empirical performance of our supervised tensor decomposition ({\bf STD}) through simulations.
 We consider order-3 tensors, where the conditional mean tensor is generated form model~\eqref{eq:decomp}.
@@
 We set $\alpha=10, d_k=d, p_k=0.4d_k, r_k=r\in\{2,4,6\}$ and increase $d$ from 25 to 50.
-Our theoretical analysis suggests that $\hat \tB$ has a convergence rate $\tO(d^{-2})$ in this setting.
-Figure~\ref{fig:dim}a plots the estimation error versus the ``effective sample size ($d^2$)'', under three different distribution models.
-We found that the empirical MSE decreases roughly at the rate of $1/d^2$, which is consistent with our theoretical ascertainment.
-%We also observed that, tensors with higher ranks tend to yield higher estimation errors, as reflected by the upward shift of the curves as $r$ increases.
-Indeed, a larger $r$ implies a higher model complexity and thus greater difficulty in the estimation.
-Similar behaviors can be observed in the non-Gaussian data in Figures~2b-c.
+Our theoretical analysis suggests that $\hat \tB = \hat\tC \times \{ \hat \mM_1,\hat \mM_2,\hat \mM_3 \}$ has a convergence rate $\tO(d^{-2})$ in this setting.
+Figure~\ref{fig:dim}a plots the estimation error versus the effective sample size ($d^2$), under three different distribution models.
+We found that the empirical mean squared error (MSE) decreases roughly at the rate of $1/d^2$, which is consistent with our theoretical ascertainment.
+Similar behaviors can be observed in the non-Gaussian data in Figure~2b-c.
 
@@
 \centering
+
 \includegraphics[width=14cm]{dim_alternative_w.pdf}
 \vspace{-.5cm}
-\caption{(a)-(c): Estimation error against effective sample size.
+\caption{\small (a)-(c): Estimation error against effective sample size.
 The dashed curves correspond to $\tO({1/d^2})$.
@@
 The $x$-axis represents the number of blocks in the networks.
-The response tensors are generated from Gaussian (a, d), Poisson (b, e) and  Bernoulli  (d, f) models .
+Response tensors are generated from Gaussian (a, d), Poisson (b, e) and  Bernoulli  (d, f) models .
 }\label{fig:dim}
 \end{figure}
-
+\vspace{-.2cm}
 The experiment II investigates the capability of our model in handling correlation among coefficients.
@@
 
-The experiment III compares {\bf STD} with three other supervised tensor methods: Higher-order low-rank regression ({\bf HOLRR}~\cite{rabusseau2016low}), Higher-order partial least square ({\bf HOPLS}~\cite{zhao2012higher}) and Subsampled tensor projected gradient ({\bf TPG}~\cite{yu2016learning}).
+The experiment III compares {\bf STD} with three other supervised tensor methods: Higher-order low-rank regression ({\bf HOLRR})~\citep{rabusseau2016low}, Higher-order partial least square ({\bf HOPLS})~\citep{zhao2012higher} and Subsampled tensor projected gradient ({\bf TPG})~\citep{yu2016learning}.
 Figure~\ref{fig:compare} shows that {\bf STD} outperforms others, especially in the low-signal, high-rank setting.
@@
 This showcases the benefit of incorporation of multiple features.
-The accuracy gain in Figure~\ref{fig:compare} demonstrates the benefit of alternating algorithm -- incorporation of informative modes also improves the estimation in the non-informative modes.
 
@@
 \centering
-\includegraphics[width=14cm]{compare_alternative_w.pdf} 
-\vspace{-.5cm}
+\includegraphics[width=13.5cm]{compare_alternative_w1.pdf} 
+\vspace{-.1cm}
 \caption{\small Comparison between different tensor methods.
@@
 \centering
-\includegraphics[width=14cm]{HCP.pdf}
-\vspace{-.5cm}
+\includegraphics[width=13.5cm]{HCP.pdf}
+\vspace{-.1cm}
 \caption{\small Top edges with large effects.
@@
 We find several edges with declined connection in the group Age 31+.
-Those edges involve Frontal-pole (\emph{Fploe}), superior-frontal (\emph{SupF}) and Cuneus nodes.
-The Frontal-pole region is known for its importance in memory and cognition, and the detected decline further suggests the age effects to brain connections.
+Those edges involve frontal-pole (\emph{Fploe}), superior-frontal (\emph{SupF}) and cuneus nodes.
+The frontal-pole region is known for its importance in memory and cognition, and the detected decline further suggests the age effects to brain connections.
 
+\vspace{-.2cm}
 \section{Conclusion}
@@
 Applications to the brain connection data yield conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.
+Further exploring the benefits of supervised tensor decomposition in specialized tasks will be necessary to boost the scientific discoveries.
+%Further exploring the benefits of supervised tensor decomposition in specialized tasks will be necessary to boost the scientific discoveries.
+We hope that released STD software will provide a systematic framework that allows more machine learning researchers to contribute to this field. 
 
@@
 Our supervised tensor decomposition method is widely applicable to network analysis, dyadic data analysis, spatial-temporal model, and recommendation systems.
-The new method improve the predictive power and enhance interpretability through incorporation of the interactive side information in tensor decomposition.
-% We have shown the improved predictive power and enhanced interpretability by incorporating the interactive side information in tensor decomposition method.  
+The new method improves the predictive power and enhances interpretability by incorporating of the interactive side information into tensor decomposition.
 The application to  brain connection dataset  shows the practical utility of the proposed method.
-%The application to the brain connection dataset yields conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.
-Tensor learning is a clear challenge for further research.
 We believe that our model enriches the research of tensor-based learning and becomes a powerful tool to boost scientific discoveries in various fields.
-%We hope the work opens up new inquiry that allows more machine learning researchers to contribute to this field.
 
-%We have developed a supervised tensor decomposition method with side information on multiple modes.
-One important challenge of tensor data analysis is the complex interdependence among tensor entries and between multiple features.
-Our approach incorporates side information as feature matrices in the conditional mean tensor.
-Our empirical results demonstrate the improved interpretability and accuracy over previous approaches.
-The application to the brain connection dataset yields conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.
-We believe that our model fills the gap between theory and practice and is a powerful tool to boost scientific discoveries.
-
-%Our exponential family tensor regression  We have shown the improved predictive power and enhanced interpretability by incorporating interactions between multiple covariate matrices.
-We acknowledge the opportunities and challenges in the tensor regression modeling.
-We hope the method developed in this paper will open up new research directions towards richer tensor-based learning.
-All code and data are publicly available, and we encourage members of the machine learning community to participate in these exciting problems.
+\section*{Acknowledgements}
+The research is supported in part by NSF grant DMS-1915978 and fundings from Wisconsin Alumni Research Foundation. We thank Zhuoyan Xu for the help with the software.
 

### 2020-11-15 tensor_regress/NeurIPSw/camera_ready/neuripsw_tensor_c_v3.tex
v3

@@
 The inner product between two tensors of equal size is defined as $\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}$.
-We use $\FnormSize{}{\cdot}$, $\mnormSize{}{\cdot}$, and $\text{vec}(\cdot)$ to denote tensor F-norm, tensor infinity norm, and the operation that reshapes the tensor into a vector, respectively.
+We use $\FnormSize{}{\cdot}$ and $\mnormSize{}{\cdot}$ to denote tensor F-norm and tensor infinity norm, respectively.
+We use $\textup{vec}(\cdot)$ to denote the operation that reshapes the tensor into a vector, and $\text{Unfold}_k(\cdot)$ to denote the operation that reshapes the tensor along mode $k$ into a matrix of size $d_k$-by-$\prod_{i\neq k}d_i$.
 We also use $\onorm{\cdot}_{\sigma}$ and $\otimes$ to denote the matrix spectral norm and Kronecker product of matrices, respectively.

### 2020-11-15 tensor_regress/NeurIPSw/camera_ready/submission/neuripsw_tensor_final.tex
submission

--- 2020-11-15 tensor_regress/NeurIPSw/camera_ready/neuripsw_tensor_c_v3.tex
@@
 
-In addition to the challenge of incorporating side information, another challenge is that many tensor datasets consist of non-Gaussian measurements \citep{wang2020learning, lee2020tensor} .
+In addition to the side information, another challenge is that many tensor datasets consist of non-Gaussian measurements \citep{wang2020learning, lee2020tensor} .
 Classical tensor decomposition methods are based on minimizing the Frobenius norm of deviation, leading to suboptimal predictions for binary- or count-valued response variables.
@@
 where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices consisting of orthonormal columns with $r_k\leq p_k$ for all $k\in[K]$, and $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$.
-Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution.
+Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and exponential link for Poisson distribution.
 
@@
 \tP=\left\{(\tC, \mM_1,\ldots,\mM_K) \ \Big| \ \mM_k^T\mM_k=\mathbf{I}_{r_k},\ \mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha \right\}
-$, and $\alpha$ is a constant. The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to avoid the divergence of the non-Gaussian variance.
+$, and $\alpha$ is a positive constant. The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to avoid the divergence of the non-Gaussian variance.
 
@@
 The convergence of our estimation becomes favorable as the order of tensor data increases.
-The proof is provided in the \cite{xu2019generalized}.
 
@@
 \centering
-\includegraphics[width=13.5cm]{HCP.pdf}
+\includegraphics[width=13.5cm]{HCP.jpg}
 \vspace{-.1cm}

### 2020-12-07 tensor_regress/NeurIPSw/submission/neuripsw_tensor.tex
neurips review

--- 2020-10-02 tensor_regress/NeurIPSw/neuripsw_tensor.tex
@@
 We call $\mX_k\mM_k$ the ``supervised tensor factors'' or ``sufficient features''~\citep{adragni2009sufficient}, and call $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes, which links the conditional mean to the feature spaces, and thereby allows the identification of variations in the tensor data attributable to the side information.
-Our goal is to find $\mM_k$ and the corresponding $\tC$,thereby allowing us to reveal the relationship between side information $\mX$ and the observed tensor $\tY$.
+Our goal is to find $\mM_k$ and the corresponding $\tC$,thereby allowing us to reveal the relationship between side information $\mX_k$ and the observed tensor $\tY$.
 Note that $\mM_k$ and $\tC$ are identifiable only up to orthonormal transformations.

### 2020-12-07 tensor_regress/NeurIPSw/NeurIPS_review.tex
Create NeurIPS_review.tex

--- 2020-08-09 skills/math/tail_bound/tailbound_solution.tex
@@
 \algnewcommand\OUTPUT{\item[\algorithmicoutput]}
-\usepackage{dsfont}
-\usepackage{colonequals}
 
 \mathtoolsset{showonlyrefs=true}
-
-\DeclareMathOperator{\var}{var}
-\DeclareMathOperator{\subG}{subG}
-\DeclareMathOperator{\tr}{Trace}
-
-\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
-\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
-
-\newcommand{\of}[1]{\left(#1\right)}
-\newcommand{\off}[1]{\left[#1\right]}
-\newcommand{\offf}[1]{\left\{#1\right\}}
-\newcommand{\aabs}[1]{\left|#1\right|}
 
@@
 \setcounter{table}{0}
+
+\newcommand{\of}[1]{\left(#1\right)}
+\newcommand{\off}[1]{\left[#1\right]}
+\newcommand{\offf}[1]{\left\{#1\right\}}
+\newcommand{\aabs}[1]{\left|#1\right|}
+
+%%%For Code
+\usepackage{listings}
+\usepackage{color}
+\definecolor{mygreen}{rgb}{0,0.6,0}
+\definecolor{mymauve}{rgb}{0.58,0,0.82}
+\lstset{ %
+  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}; should come as last argument
+  basicstyle=\scriptsize,        % the size of the fonts that are used for the code
+  breaklines=true,                 % sets automatic line breaking
+  captionpos=t,                    % sets the caption-position to bottom
+  frame=single,	                   % adds a frame around the code
+  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
+  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
+  numbersep=5pt,                   % how far the line-numbers are from the code
+  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
+  tabsize=2,	                   % sets default tabsize to 2 spaces
+  commentstyle=\color{mygreen},
+  stringstyle=\color{mymauve},
+}
+%%%
 
@@
 
-\title{Solution to ``Chapter 2: Basic tail and concentration bounds"}
+\title{Review for NeurIPS 2020}
 
@@
 \maketitle
+\section{Day 1: EXPO Day}
+EXPO presentations give us an opportunity to know more about the practical application of machining learning in the industry. I attend several talks of my interest.
+\begin{enumerate}
+	\item \textbf{Accelerating Deep Learning for Entertainment with Sony's Neural Network Libraries and Console.}
 
-\section{Summary}
-\begin{thm}[Markov's inequality]\label{thm:markov}
-	Let $X \geq 0$ be a random variable  with a finite mean. We have
-	\begin{equation}\label{eq:thmmarkov}
-		\bbP(X\geq t) \leq \frac{\bbE[X]}{t},\quad \text{ for all } \ t > 0.
-	\end{equation}  
-\end{thm}
+	\vspace{0.2cm}
+	This talk mainly introduces the application of Deep Learning (DL) techniques used in the Sony Entertainment with their own tool package and platform: Sony's Neural Network Libraries and Console.
+The presenter takes three examples related to the content creation.
 
-\begin{thm}[Chebyshev's inequality]\label{thm:cheb}
-	Let $X \geq 0$ be a random variable  with a finite mean $\mu$ and a finite variance. We have
-	\begin{equation}\label{eq:cheby}
-		\bbP(|X - \mu |\geq t) \leq \frac{\var(X)}{t^2},\quad \text{ for all } \ t > 0.
-	\end{equation}  
-\end{thm}
+	First one is the reference-based image colorization.
+The initial challenge is the temporal incoherence issue.
+Colorizing the video frame by frame may lead to color inconsistence when playing the whole video.
+To address this issue, the proposed approach predicts the input frame with the correspondence in the reference images.
+However, the correspondence may also lead incorrect colorization when multiple objects of the same class are present.
+To overcome this limitation, the team incorporate masking skills, instance tracking and dense tracking, to restrict the correspondence region.
+The proposed method outperforms the state-of-arts.
 
-\begin{thm}[Markov's inequality for polynomial moments]\label{thm:polymoment}
-	Let $X$ be a random variable. Suppose that the order $k$ central moment of $X$ exists. Applying Markov's inequality to the random variable $|X - \mu|^k$ yields
-	\begin{equation}\label{eq:poly}
-		\bbP(|X - \mu |\geq t) \leq \frac{\bbE\left[|X-\mu|^k\right]}{t^k},\quad \text{ for all } \ t > 0.
-	\end{equation}  
-\end{thm}
+	Second one is the music source separation.
+The task is to separate the music from different sources, such as voice, string instruments, and drums.
+A new proposed CNN, named D3Net is proposed for the task.
+In addition, adversarial attacks is an important topic in audio separation.
+The adversarial noise can be unnoticeable but interfere the separation severely, which may lead to separator malfunction or be used in content protection.
 
-\begin{thm}[Chernoff bound]\label{thm:chernoff}
- Let $X$ be a random variable.
-Suppose that the moment generating function of $X$, denoted $\varphi_X(\lambda)$, exists in the neighborhood of 0; i.e., $\varphi_X(\lambda) = \bbE[e^{\lambda X}] < +\infty,$ for all $\lambda \in (-b,b)$ with some $b>0$.
-Applying Markov's inequality to the random variable $Y = e^{\lambda(X- \mu)}$ yields
-	\begin{equation}\label{eq:chernoff1}
-		\bbP((X - \mu) \geq t) \leq \frac{\bbE\left[e^{\lambda (X-\mu)}\right]}{e^{\lambda t}}, \quad \text{ for all } \lambda\in (-b,b).
-	\end{equation}
-	Optimizing the choice of $\lambda$ for the tightest bound, we obtain the Chernoff bound
-	\begin{equation}\label{eq:chernoff2}
-		\bbP((X - \mu) \geq t) \leq \inf_{\lambda \in [0,b)} \frac{\bbE\left[e^{\lambda (X-\mu)}\right]}{e^{\lambda t}}.
-	\end{equation}
-\end{thm}
-
-\begin{thm}[Hoeffding bound for bounded variable]\label{thm:boundvar}
-Let $X$ be a random variable with $\mu = \bbE(X)$. Suppose that $X \in [a,b]$ almost surely, where $a\leq b\in\bbR$ are two constants. Then, we have
-\begin{equation}
-	\bbE[e^{\lambda X}] \leq e^{\frac{s(b-a)^2}{8}}, \quad \text{for all } \lambda \in \bbR.
-\end{equation}
-Consequently, the variable $X \sim \subG\left(\frac{(b-a)^2}{4}\right)$.
-\end{thm}
-
-\begin{proof}
-	See Exercise~\ref{pro:boundvar}.
-\end{proof}
-
-\begin{thm}[Moment of sub-Gaussian variable]\label{thm:momsubg}
-Let $X \sim \subG(\sigma^2)$. For all integer $k \geq 1$, we have
-\begin{equation}\label{eq:thmmomsubg}
-	\bbE[|X|^k] \leq  k 2^{k/2} \sigma^k \Gamma(\frac{k}{2} ),
-\end{equation} 
-where the Gamma function is defined as $\Gamma(x) = \int_{0}^{+\infty} t^{x -1} e^{-t} dt$. 
-\end{thm}
-
-\begin{thm}[One-sided Bernstein’s inequality]\label{thm:oneber}
- Let $X$ be a random variable. Suppose $X \leq b$ almost surely. We have 
-\begin{equation}\label{eq:thmoneber1}
-	\bbE\left[e^{\lambda (X-\bbE[X])}\right] \leq exp\left\{ \frac{\lambda^2 \bbE[X^2]/2}{1 - b\lambda /3}  \right\}, \quad \text{for all } \lambda \in [0, 3/b).
-\end{equation}
-Consequently, let $X_i$ be independent variables, and $X_i \leq b$ almost surely, for all $i \in [n]$. We have
-\begin{equation}\label{eq:thmoneber2}
-	\bbP\left[\sum_{i = 1}^n (X_i - \bbE[X_i] ) \geq n\delta \right] \leq exp\left\{ - \frac{n\delta^2}{\sum_{i = 1}^n \bbE[X_i^2] /n + b \delta /3} \right\},\quad \text{ for all } \delta \geq 0.
-\end{equation}  
-Particularly, let $X_i$ be independent nonnegative variables, for all $i \in [n]$. The equation~\eqref{eq:thmoneber2} becomes
-\begin{equation}\label{eq:thmonber0}
-	\bbP\left[\sum_{i = 1}^n (Y_i - \bbE[Y_i] ) \leq n\delta \right] \leq exp\left\{ - \frac{n\delta^2}{\sum_{i = 1}^n \bbE[Y_i^2] /n} \right\},\quad \text{ for all } \delta \geq 0.
-\end{equation} 
-\end{thm}
-
-\begin{defn}[Bernstein's condition]
-Let $X$ be a random variable with mean $\mu = \bbE[X]$ and variance $ \sigma^2 = \var(X)$. We say $X$ satisfies the Bernstein's condition with parameter $b$ if
-\begin{equation}\label{eq:berncond}
-	\left| \bbE [(X -\mu)^k ]\right| \leq \frac{1}{2} k! \sigma^2 b^{k-2}, \quad \text{for } k =3, 4,....
-\end{equation} 
-\end{defn}
-Note that bounded random variables satisfy the Bernstein's condition.
-
-\begin{thm}[Bernstein-type bound] For any variable $X$ satisfying the Bernstein's condition, we have
-\begin{equation}
-	\bbE\left[e^{\lambda (X-\mu)}\right] \leq \exp\left\{ \frac{\lambda^2 \sigma^2}{2(1 - b |\lambda|)} \right\},\quad \text{for all } |\lambda| \leq \frac{1}{b}, 
-\end{equation}
-and the concentration inequality
-\begin{equation}\label{eq:thmbernconc}
-	\bbP\left[|X - \mu| \geq t\right] \leq 2 \exp \left\{ -\frac{t^2}{2(\sigma^2 + bt)} \right\}, \quad \text{for all }t \geq 0.
-\end{equation}
-
-\end{thm}
-
-\begin{defn}[Bounded difference property] Let $f\colon \bbR^n \mapsto \bbR$ be a function. The function $f$ satisfies the bounded difference property with parameter $(L_1,...,L_n)$ if we have
-	\begin{equation}\label{eq:defbounded}
-		\aabs{f(x^{(k)}) - f(x^{'(k)})} \leq L_k,
-	\end{equation}
-	 for all $k \in [n]$ and for all $x^{(k)} = (x_1,...,x_k,...,x_n), x^{'(k)} = (x_1,...,x'_k,...,x_n)\in\bbR^n$.
-\end{defn}
-
-\begin{thm}[Bounded differences inequality] \label{thm:bounddiff}
-Let $f\colon \bbR^n \mapsto \bbR$ be a function satisfies the bounded difference property~\eqref{eq:defbounded}, and the random variable $X = (X_1,...,X_n)$ has independent components. Then,
-\begin{equation}
-	\bbP\off{\aabs{f(X) -\bbE[f(X)] } \geq t} \leq 2 e^{-\frac{2t^2}{\sum_{i=1}^n L_i^2}}, \quad \text{for all }t \geq 0.
-\end{equation}
-
-\end{thm}
-
-\section{Exercises}
-
-\subsection{Exercise 2.1} (Tightness of inequalities.) The Markov's and Chebyshev's inequalities are not able to be improved in general. 
-\begin{enumerate}
-	\item[(a)] Provide a random variable $X \geq 0$ that attains the equality in Markov's inequality~\eqref{eq:thmmarkov}.  
-	\item[(b)] Provide a random variable $Y$ that attains the equality in Chebyshev's inequality~\eqref{eq:cheby}.  
+	Third one is the CPU-GPU memory swapping for large-scale data training. The key idea is using CPU to expand the memory usage on GPU.
 \end{enumerate}
 
-\paragraph{Solution:} 
-\begin{enumerate}
-
-\item[(a)] For a given constant $t >0$, we define a variable $Y_{t} = X - t \mathds{1}{[X\geq t]}$, where $\mathds{1}$ is the indicator function.
-Note that $Y_{t}$ is a nonnegative variable.
-The Markov's inequality follows by taking the expectation to $Y_t$,
-\begin{equation}
-	\bbE[Y_t] = \bbE[X] - t \bbP[X \geq t] \geq 0.
-\end{equation}
-Therefore, Markov's inequality meets the equality if and only if the expectation $\bbE[Y_t] = 0$.
-Since $Y_t$ is nonnegative, we have $\bbP(Y_t = 0) = 1$.
-Note that $Y_t = 0$ if and only if $X = 0$ or $X =t$.
-
-Hence, for the given constant $t >0$, the nonnegative variable $X$ with distribution $\bbP(X \in \{ 0,t\}) = 1$ attains the equality of Markov's inequality.
-
-\item[(b)]  Chebyshev's inequality follows by applying Markov's inequality to the nonnegative random variable $Z = {(X - \bbE[X])^2}$.
-Simialrly as in part (a), given a constant $t > 0$, the variable $Z = {(X - \bbE[X])^2}$ with distribution $\bbP(Z \in \{ 0,t^2 \}) = 1$ attains the equality of the Markov's inequality for $Z$.
-Consequently, the variable $X$ attains the equality of the Chebyshev's inequality for $X$.
-By transformation, the distribution of $X$ satisfies the followings formula,
-\begin{align}
-\bbP(X = x) = 
-	\begin{cases}
-		\quad p & \text{if } x = c,\\
-		\quad \frac{1-p}{2} & \text{if } x = c-t \text{ or } x = c+t, \\
-		\quad 0 & \text{ otherwise },
-	\end{cases}
-	\end{align}
-where $c \in \bbR$ is a constant and $p \in [0,1]$.
-
-\iffalse
-	\item[(a)] Recall the proof of Markov's inequality. For any $t >0$,
-	\begin{equation}\label{eq:markov}
-		\bbE[X] = \int_{0}^t x f_X(x) dx + \int_{t}^{+\infty} x f_X(x) dx \geq \int_{t}^{+\infty} x f_X(x)\geq t \int_{t}^{+\infty} f_X(x) = t \bbP( X \geq t).
-	\end{equation}
-	Below, given a constant $t > 0$, we construct a random variable that attains the equality in each line of equation~\eqref{eq:markov}.
-
-	 Consider a variable $X$ with distribution $P(X = t) = 1$. The variable $X$ satisfies 
-	 \[ \int_{0}^t x f_X(x) dx = 0 \text{ and } \int_{t}^{+\infty} x f_X(x) dx =\int_{t}^{+\infty} t f_X(x) dx. \]
-	 Therefore, for given $t$, the variable $X$ attains the equality of Markov's inequality.
-
-	 \item[(b)] Chebyshev's inequality follows by applying Markov's inequality to the non-negative random variable $Y = {(X - \bbE[X])^2}$.
-Let the distribution of $Y$ be $\bbP(Y = t^2) = 1$.
-Then the Markov's inequality for $Y$ and the Chebyshev's inequality for $X$ meet the equalities.
-Consider the variable $Z$ with distribution $\bbP(Z = c+t) = \bbP(Z = c-t)  = 1/2 $ for any $ c \in \bbR$.
-For $Z$, the transformation probability satisfies $\bbP((Z - \bbE[Z])^2= t^2) = 1$.
-Therefore, the variable $Z$ attains the equality of Chebyshev's inequality.
-
-\fi
-
-\end{enumerate}
-
-\begin{rmk}[Tightness of Markov's inequality]
-	Only a few variables attain the equalities in Markov's and Chebyshev's inequalities. In research, we should pay attention to the concentration bounds tighter than Markov's inequality.
-\end{rmk}
-
-\subsection{Exercise 2.2}
-\begin{lem}[Standard normal distribution]\label{lem:n01}
-	Let $\phi(z)$ be the density function of a standard normal variable $Z \sim N(0,1)$. Then, 
-	\begin{equation}\label{eq:e22a}
-		\phi'(z) + z \phi(z) = 0,
-	\end{equation}
-	and \begin{equation}\label{eq:e22b}
-		\phi(z)\left(\frac{1}{z} - \frac{1}{z^3} \right) \leq \bbP(Z \geq z) \leq \phi(z)\left(\frac{1}{z} - \frac{1}{z^3}+ \frac{3}{z^5} \right), \quad \text{for all }z>0.
-	\end{equation}
-\end{lem}
-\begin{proof}First, we prove the equation~\eqref{eq:e22a}.
-
-The pdf of the standard normal distribution is
-\[  \phi(z) = \frac{1}{\sqrt{2\pi}} \exp\left(-\frac{z^2}{2}  \right).\]
-
-The equation~\eqref{eq:e22a} follows by taking the derivative of $\phi(z)$. Specifically,
-\[\quad \phi'(z) = -z \frac{1}{\sqrt{2\pi}} \exp\left(-\frac{z^2}{2} \right) = -z \phi(z). \] 
-	Next, we prove the equation~\eqref{eq:e22b}.
-
-	We write the upper tail probability of the standard normal variable as
-	\begin{equation}\label{eq:e22pro1}
-		\bbP(Z \geq z) = \int_{z}^{+\infty} \phi(t) dt = \int_{z}^{+\infty} - \frac{1}{t} \phi'(t) dt = \frac{1}{z} \phi(z) - \int_{z}^{+\infty} \frac{1}{t^2} \phi(t) dt,  
-	\end{equation}
-	where the second equality follows by the equation~\eqref{eq:e22a}. Applying the equation~\eqref{eq:e22a} to the last term in equation~\eqref{eq:e22pro1} yields
-	\begin{equation}\label{eq:e22pro2}
-	\int_{z}^{+\infty} \frac{1}{t^2} \phi(t) dt = \int_{z}^{+\infty} \frac{1}{t^3} \phi'(t) dt = - \frac{1}{z^3} \phi(z) + \int_{z}^{+\infty} \frac{3}{t^4} \phi(t) dt \geq  - \frac{1}{z^3} \phi(z)
-	\end{equation}
-	Plugging the equation~\eqref{eq:e22pro2} into the equation~\eqref{eq:e22pro1}, we obtain $\bbP(Z \geq z) \geq \phi(z)\left(\frac{1}{z}  - \frac{1}{z^3} \right)$.
-Applying the equation~\eqref{eq:e22a} again to the equation~\eqref{eq:e22pro2} yields
-	\begin{equation}\label{eq:e22pro3}
-		\int_{z}^{+\infty} \frac{3}{t^4} \phi(t) dt = \int_{z}^{+\infty} - \frac{3}{t^5} \phi'(t) dt =\frac{3}{ z^5} \phi(z) -  \int_{z}^{+\infty} \frac{15}{t^6} \phi(t) dt \leq \frac{3}{ z^5} \phi(z).
-	\end{equation}
-
-	Combing equations~\eqref{eq:e22pro1}, \eqref{eq:e22pro2} and  \eqref{eq:e22pro3}, we obtain $\bbP(Z \geq z) \leq \phi(z)\left(\frac{1}{z}  - \frac{1}{z^3} + \frac{3}{z^5} \right).$
-\end{proof}
-
-\begin{rmk}
-	Direct calculation of tail probability for a univariate normal variable is hard.
-Equation~\eqref{eq:e22b} provides a numerical approximation to the tail probability.
-Particularly, the tail probability decays at the rate of $z^{-1}e^{-z^2/2}$ as $z \rightarrow +\infty$.
-The decay rate is faster than polynomial rate $\tO(z^{-\alpha})$, for any $\alpha \geq 1$.
-\end{rmk}
-
-\subsection{Exercise 2.3}
-
-\begin{lem}[Polynomial bound and Chernoff bound]\label{lem:momcher}
-	Let $X \geq 0$ be a nonnegative variable.
-Suppose that the moment generating function of $X$, denoted $\varphi_{X}(\lambda)$, exists in the neighborhood of $\lambda = 0$.
-Given some $\delta > 0$, we have
-	\begin{equation}\label{eq:lempoly}
-		\inf_{k \in \mathbb{Z_+} } \frac{\bbE[|X|^k]}{\delta^k} \leq \inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}}. 
-	\end{equation}
-
-	Consequently, an optimized bound based on polynomial moments is always at least as good as the Chernoff upper bound.
-\end{lem}
-
-\begin{proof}
-	By power series, we have
-	\begin{equation}\label{eq:power}
-		 e^{\lambda X} = \sum_{k = 0}^{+\infty } \frac{X^k \lambda^k }{k!}, \quad \text{ for all } \lambda \in \bbR.
-	\end{equation}
-	Since the moment generating function $\varphi_X(\lambda)$ exists in the neighborhood of $\lambda = 0$, there exists a constant $b > 0$ such that 
-		\[   \bbE [e^{\lambda X} ]= \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} < +\infty,\quad \text{ for all } \lambda \in (0,b). \]
-	 Hence, the moment  $\bbE[|X|^k]$ exists, for all $ k \in \mathbb{Z_+}$. Applying power series~\eqref{eq:power} to the right hand side of equation~\eqref{eq:lempoly} yields
-	\begin{equation}\label{eq:powertopoly}
-		\inf_{\lambda >0} \frac{\bbE[e^{\lambda X}]}{e^{\lambda \delta}} = \frac{ \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} }{ \sum_{k = 0}^{+\infty } \frac{\lambda^k \delta^k }{k!}  }.	\end{equation}
-	By Cauchy's third inequality, we have
-	\begin{equation}\label{eq:cauchy3}
-		\frac{ \sum_{k = 0}^{+\infty } \frac{ \bbE[|X|^k]  \lambda^k }{k!} }{ \sum_{k = 0}^{+\infty } \frac{\lambda^k \delta^k }{k!}  } \geq \inf_{k \in \mathbb{Z}_{+} } \frac{\bbE[|X|^k]}{\delta^k}  
-	\end{equation}
-	Therefore, we obtain the equation~\eqref{eq:lempoly} by combining the equation~\eqref{eq:powertopoly} with equation~\eqref{eq:cauchy3}.	
-\end{proof}
-
-\begin{rmk}
-	Applying different functions $g(X)$ to the Markov's inequality leads to different bounds for the tail probability of variable $X$.
-Equation~\eqref{eq:lempoly} implies that the optimized polynomial bound is at least as tight as the Chernoff bound, provided that the moment generating function of $X$ exsits in the neighborhood of 0.
-\end{rmk}
-
-\subsection{Exercise 2.4}\label{pro:boundvar}
-In Exercise~\ref{pro:boundvar}, we prove Theorem~\ref{thm:boundvar},  the Hoeffding bound for a bounded variable.
-
-\begin{proof}
-
-Let $X$ be a bounded random variable, and $X \in [a,b]$ almost surely, where $a 
-\leq b \in \bbR$ are two constants. Let $\mu = \bbE[X]$. Define the function 
-\[ g(\lambda) = \log \bbE[e^{\lambda X }], \quad  \text{ for all } \lambda \in \bbR. \]
-Applying Taylor Expansion to $g(\lambda)$ at 0, we have
-\begin{equation}\label{eq:taylor}
-	g(\lambda) =  g(0) + g'(0) \lambda  + \frac{g''(\lambda_0) }{2} \lambda^2, \quad \text{where } \lambda_0 = t \lambda, \text{ for some } t \in [0,1]. 
-\end{equation}
-
-In equation~\eqref{eq:taylor}, the term $g(0) =\log \bbE[e^{0}] = 0$. By power series \eqref{eq:power}, we obtain the first derivative $g'(\lambda)$ as follows,
-\begin{align}
-	g'(\lambda ) & = \left( \log \sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]  \right)' \\
-	&= {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+1)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]}  \label{eq:firstde} \\
-	&= \frac{\bbE[X e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]}.
-\end{align}
-Therefore, $g'(0) = \bbE[X] = \mu$. Taking the derivative to equation~\eqref{eq:firstde}, we obtain the second-order derivative $g''(\lambda)$ as follows,
-\begin{align}
-	g''(\lambda)
-	&= {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+2)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]} - \left( {\sum_{k = 0}^n \frac{ \lambda^k }{k!} \bbE[X^{(k+1)}] }\Bigg/{\sum_{k = 0}^n \frac{\lambda^k }{k!} \bbE[X^k]}  \right)^2\\
-	&= \frac{\bbE[X^2 e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]} - \left( \frac{\bbE[X e^{\lambda X} ]}{\bbE[ e^{\lambda X} ]} \right)^2.\\
-	%& \overset{\Delta}{=} \bbE_{\lambda} [X^2] - (\bbE_{\lambda} [X])^2
-\end{align}
-We interpret the second-order derivative $g''(\lambda)$ as the variance of $X$ with the re-weighted distribution $d P' = {e^{\lambda X}}/{\bbE[ e^{\lambda X} ]} d P_X $, where $P_X$ is the distribution of $X$.
-Taking the integral of 1 with respect to $dP'$, we have
-\begin{equation}
-	\int_{-\infty}^{+\infty} d P' = \int_{-\infty}^{+\infty}  \frac{e^{\lambda X}}{\bbE[ e^{\lambda X} ]} d P_X = 1,
-\end{equation} 
-which implies that the function $P'$ is a valid probability distribution. Under all possible re-weighted distributions, the variance of $X$ is upper bounded as follows,
-\[ \var(X) = \var(X - \frac{a+b}{2}) \leq \bbE[(X - \frac{a+b}{2} )^2] \leq \frac{(b-a)^2}{4}, \]
-where the term $\frac{(b-a)^2}{4}$ follows by letting $X$ supported on the boundaries $a$ and $b$ only.
-Hence, the second-order derivative $g''(\lambda) \leq \frac{(b-a)^2}{4}$.
-We plug the results of $g'$ and $g''$ into the equation~\eqref{eq:taylor}.
-Then,
-\begin{equation}\label{eq:e24}
-	g(\lambda) =  g(0) + g'(0) \lambda  + \frac{g''(\lambda_0) }{2} \lambda^2 \leq 0 + \lambda \mu+ \frac{(b-a)^2}{8} \lambda^2. 
-\end{equation}
-Taking the exponentiation on both sides of the inequality~\eqref{eq:e24}, we have
-\begin{equation}\label{eq:boundvar}
-	\bbE[e^{\lambda X}] = \exp(g( \lambda)) \leq e^{\mu \lambda + \frac{(b-a)^2}{8} \lambda^2}.
-\end{equation}
-The equation~\eqref{eq:boundvar} implies that $X$ is a sub-Gaussian variable with at most $\sigma = \frac{(b-a)}{2}$.
-\end{proof}
-
-\begin{rmk}
-	For any bounded random variable $X$ supported on $[a,b]$, $X$ is a sub-gaussian variable with parameter at most $\sigma^2 = (b-a)^2/4$.
-All the properties for sub-Gaussian variables apply to the bounded variables.
-\end{rmk}
-
-\subsection{Exercise 2.5}
-\begin{lem}[Sub-Gaussian bounds and means/variance] Let $X$ be a random variable  such that
-\begin{equation}\label{eq:subg}
-	 \bbE[e^{\lambda X}] \leq e^{\frac{\lambda^2 \sigma^2}{2} + \mu \lambda}, \quad \text{ for all } \lambda \in \bbR. 
-\end{equation}
-	Then, $\bbE[X] = \mu$ and $\var(X) \leq \sigma^2$.
-\end{lem}
-
-\begin{proof} By equation~\eqref{eq:subg}, the moment generating function of $X$, denoted $\varphi_X(\lambda)$, exists in the neighborhood of $\lambda = 0$.
-Hence, the mean and variance of $X$ exist.
-For all $\lambda$ in the neighborhood of $\lambda = 0$, applying power series on both sides of equation~\eqref{eq:subg} yields
-\begin{equation}\label{eq:taylor2}
-	 \lambda \bbE[X] + \frac{\lambda^2}{2} \bbE[X^2] + o(\lambda^2) \leq  \mu \lambda + \frac{\lambda^2 \sigma^2 + \lambda^2 \mu^2}{2} + o(\lambda^2). 
-\end{equation}
-Dividing by $\lambda >0$ on both sides of equation~\eqref{eq:taylor2} and letting $\lambda \rightarrow 0^{+}$, we have $\bbE(X) \leq \mu$.
-Dividing by $\lambda <0$ on both sides of equation~\eqref{eq:taylor2} and letting $\lambda \rightarrow 0^{-}$, we have $\bbE(X) \geq \mu$.
-Therefore, we obtain the mean $\bbE[X] = \mu$.
-Then, we divide ${2}/{\lambda^2}$ on both sides of equation~\eqref{eq:taylor2}, for $\lambda \neq 0$.
-The term $\bbE[X]\lambda$ and  $\mu \lambda$ are cancelled.
-We have $ \bbE[X^2] \leq \sigma^2 + \mu^2$, and thus the $\var(X) \leq \bbE[X^2]  - (\bbE[X])^2 = \sigma^2$.
-\end{proof}
-
-\paragraph{Question:} Let $\sigma^2_{min}$ denote the smallest possible $\sigma$ satisfying the inequality~\eqref{eq:subg}. Is it true that $\var(X) = \sigma_{min}^2$?
-
-\paragraph{Solution:} The statement that $\var(X) = \sigma_{min}^2$ is not necessarily true.
-Recall the function $g(\lambda)$ in Exercise~\ref{pro:boundvar}.
-By the results in Exercise~\ref{pro:boundvar},  the equation~\eqref{eq:subg} is equal to
-\begin{equation}\label{eq:question}
-	g''(\lambda) \leq  \sigma^2, \quad \text{ for all } \lambda \in \bbR,
-\end{equation}
-where $g'' (\lambda)$ is the variance of $X$ with the re-weighted distribution defined in Exercise~\ref{pro:boundvar}.
-Therefore, we have $\max_{\lambda}g''(\lambda) = \sigma_{min}^2$.
-Note that $g''(0) = \var(X)$.
-To let the equality $\var(X) = \sigma_{min}^2$ hold, we need to show that $\max_{\lambda}g''(\lambda) = g''(0)$ holds for $X$.
-
-\vspace{0.2cm}
-However, the statement $\max_{\lambda}g''(\lambda) = g''(0)$ is not necessarily true.  A counter example is below. 
-Consider a random variable $Y \sim Ber(1/3)$. The variance of $Y$ is $\var(Y) = 2/9$. Let $\lambda = 1$. The re-weighted distribution $dP'$ is
-\[ P'(Y = 0) = \frac{2}{3\bbE[ e^{Y} ]}\quad  \text{and}\quad  P'(Y = 1) = \frac{e}{3 \bbE[ e^{ Y } ]}, \quad \text{where } \bbE[ e^{Y} ] = \frac{2}{3} + \frac{e}{3}.  \]
-The variance of $Y$ with $dP'$ is ${2}/{3\bbE[ e^{Y} ]} \times  {e}/{3\bbE[ e^{Y} ]} = 0.2442 > 2/9$.
-Therefore, we have $\var(Y) <  g''(1) \leq \max_{\lambda}g''(\lambda) = \sigma^2_{min}$.
-The statement $\max_{\lambda}g''(\lambda) = g''(0)$ is not true for this variable $Y$.
-\begin{rmk}
-	Parameters of a sub-Gaussian distribution provide the exact value of the mean and an upper bound of the variance; i.e., $ \bbE[X] = \mu$ and $\var(X) \leq \sigma^2$.
-Suppose the moment generating function of variable  $X$ exists over the entire real interval.
-Then, the tail distribution of $X$ is bounded by a sub-Gaussian distribution with a proper choice of $\sigma^2$.
-\end{rmk}
-
-\subsection{Exercise 2.6}
-
-\begin{lem}[Lower bounds on squared sub-Gaussians]\label{lem:lowersquare} Let $\{ X_i \}_{i=1}^n$ be an i.i.d.\ sequence of zero-mean sub-Gaussian variables with parameter $\sigma$.
-The normalized sum $Z_n = \frac{1}{n} \sum_{i=1}^n X_i^2$ satisfies
-\begin{equation}\label{eq:lowersquare}
-	\bbP[Z_n - \bbE[Z_n] \leq \sigma^2 \delta ] \leq e^{-n\delta^2 /16}, \quad \text{ for all } \delta \geq 0.
-\end{equation}
-	The equation~\eqref{eq:lowersquare} implies that the lower tail of the sum of squared sub-Gaussian variables behaves in a sub-Gaussian way.
-\end{lem}
-
-\begin{proof}
-	Since $X_i^2$ are i.i.d.\ nonnegative variables, we apply the equation~\eqref{eq:thmonber0} to the variables $\{X_i^2\}_{i=1}^n$. Then, we have
-	\begin{equation}\label{eq:e26a}
-		\bbP\left[\sum_{i=1}^n (X_i^2 -\bbE[X_i^2]) \leq n \sigma^2 \delta\right] \leq \exp\left\{ - \frac{n\delta^2 \sigma^4}{ \bbE[X_1^4]} \right\},  \quad \text{for all }\delta \geq 0.
-	\end{equation}
-	By equation~\eqref{eq:thmmomsubg}, we have
-	\begin{equation}\label{eq:e26x4}
-		\bbE[X_1^4] \leq 16 \sigma^4. 
-	\end{equation}
-	Combing equations~\eqref{eq:e26a}, \eqref{eq:e26x4} and the definition of $Z_n$, we obtain 
-	\begin{equation}
-		\bbP[Z_n - \bbE[Z_n] \leq \sigma^2 \delta ]  \leq \exp\left\{ - \frac{n\delta^2}{16} \right\},  \quad \text{for all }\delta \geq 0.
-	\end{equation}
-\end{proof}
-
-\begin{rmk}
-	Equation~\eqref{eq:lowersquare} implies that the lower tail of the sum of squared sub-Gaussian variables behaves in a sub-Gaussian way.
-In following sections, we will show that the variable $Z_n - \bbE[Z_n]$ in Lemma~\ref{lem:lowersquare} is a sub-exponential variable.
-\end{rmk}
-
-\subsection{Exercise 2.7}
-\begin{lem}[Bennett’s inequality]\label{lem:bennett}
-	Let $X_1,...,X_n$ be a sequence of independent zero-mean random variables with $|X_i| \leq b$ and $\var(X_i) = \sigma_i^2$, for all $i \in [n]$. Then, we have the Bennett's inequality
-	\begin{equation}\label{eq:bennett}
-		\mathbb{P}\left[\sum_{i=1}^{n} X_{i} \geq n \delta\right] \leq \exp \left\{-\frac{n \sigma^{2}}{b^{2}} h\left(\frac{b \delta}{\sigma^{2}}\right)\right\}, \quad \text{for all }\delta \geq 0,
-	\end{equation} 
-	where $\sigma^2 = \frac{1}{n} \sum_{i=1}^n \sigma_i^2$ and $h(t) \colonequals (1+t)\log (1+t) - t$ for $t \geq 0$.
-\end{lem}
-
-\begin{proof}
-	First, we consider the moment generating function of $X_i$, for all $i \in [n]$.
-
-	\vspace{0.2cm}
-	By power series, for all $i \in [n]$, we have
-	\begin{align}\label{eq:e27power}
-		\bbE\left[e^{\lambda X_i}\right] = \sum_{k = 0}^{+\infty} \frac{\lambda^k \bbE[X_i^k]}{k!} 
-		= 1 + 0 + \sum_{k = 2}^{+\infty} \frac{\lambda^k \bbE[X_i^k]}{k!} \leq \exp\left\{ \sum_{k = 2}^{+\infty} \frac{\lambda^k \bbE[X_i^k]}{k!} \right\},
-	\end{align}
-	where the 0 comes from the fact that $\bbE[X_i] = 0$, and the last inequality follows from $1 + x \leq e^x$. By $|X_i|<b$, we bound the last term in equation~\eqref{eq:e27power} as follows
-	\begin{equation}\label{eq:e27upper}
-		\sum_{k = 2}^{+\infty} \frac{\lambda^k \bbE[X_i^k]}{k!}  \leq \sum_{k = 2}^{+\infty} \frac{\lambda^k \bbE[X_i^2 |X_i|^{k-2}] }{k!} \leq  \sum_{k = 2}^{+\infty} \frac{\lambda^k \sigma^2_i b^{k-2}}{k!} = \sigma_i^2  \left(  \frac{e^{\lambda b} - 1 -\lambda b}{b^2}   \right).
-	\end{equation}
-	Combing the equation~\eqref{eq:e27power} with equation~\eqref{eq:e27upper}, we obtain the following upper bound of the moment generating function of $\sum_{i=1}^n X_i$.
-	\begin{equation}\label{eq:e27uppert}
-		\bbE\left[e^{\lambda \sum_{i = 1}^n X_i }\right] = \prod_{i = 1}^n \bbE\left[e^{\lambda X_i}\right] \leq \exp\left\{ n \sigma^2 \left( \frac{e^{\lambda b} - 1 -\lambda b}{b^2}   \right)  \right\},
-	\end{equation}
-	where $\sigma^2 = \frac{1}{n} \sum_{i=1}^n \sigma_i^2$. Combing the Chernoff bound with equation~\eqref{eq:e27uppert}, the upper tail of $\sum_{i=1}^n X_i$ follows 
-	\begin{align}
-		\mathbb{P}\left[\sum_{i=1}^{n} X_{i} \geq n \delta\right] &\leq  \exp\left\{  n \sigma^2 \left(  \frac{e^{\lambda b} - 1 -\lambda b}{b^2}   \right)  - 
-		\lambda n \delta \right\} \\
-		&= \exp \left\{ \frac{n\sigma^2}{b^2} \left( e^{\lambda b} - \lambda b- \lambda \frac{\delta b^2}{\sigma^2} - 1 \right)  \right\}, \quad \text{for all } \delta \geq 0. \label{eq:e27exp}
-	\end{align} 
-	The upper bound \eqref{eq:e27exp} achieves the minimum when $\lambda = b^{-1}\log\left(1 + \frac{\delta b}{\sigma^2} \right)$ by the first-order condition of minimization.
-Plugging $\lambda =b^{-1} \log\left(1 + \frac{\delta b}{\sigma^2} \right)$ into the equation~\eqref{eq:e27exp}, we obtain the Bennett's inequality
-	\begin{equation}\label{eq:e27bennett}
-		\mathbb{P}\left[\sum_{i=1}^{n} X_{i} \geq n \delta\right] \leq  \exp \left\{-\frac{n \sigma^{2}}{b^{2}} h\left(\frac{b \delta}{\sigma^{2}}\right)\right\}, \quad \text{for all }\delta \geq 0,
-	\end{equation}
-	where $h(t) \colonequals (1+t)\log (1+t) - t$ for $t \geq 0$.
-
-	\vspace{0.2cm}
-	Further, we show that the Bennett's inequality is at least as good as the Bernstein’s inequality.
-
-	\vspace{0.2cm}
-	The Bernstein's inequality for $\sum_{i=1}^n X_i$ is
-	\begin{align}\label{eq:e27bern}
-			\mathbb{P}\left[\sum_{i=1}^{n} X_{i} \geq n \delta \right] \leq \exp \left\{ \frac{-3n\delta^2}{(2 b \delta + 6 \sigma^2)} \right\}  %=\exp \left\{ -\frac{n \sigma^{2}}{ b^{2}} \frac{3b^2\delta^2}{2\sigma^2(b\delta + 6 \sigma^2)} \right\}
-			= \exp \left\{ -\frac{n \sigma^{2}}{ b^{2}} g\left(\frac{b\delta}{\sigma^2}\right) \right\}, \quad \text{for all } \delta \geq 0,
-	\end{align}
-	where $g(t) \colonequals  \frac{3t^2}{2t + 6} $ for $ t \geq 0$.
-Since $g(t)\leq h(t)$ holds for all $t \geq 0$, we conclude that the Bennett's inequality~\eqref{eq:e27bennett} is at least as good as Bernstein's inequality~\eqref{eq:e27bern}.
-\end{proof}
-
-\begin{rmk} 
-	So far, we have three inequalities controlling the tail of bounded variables: Hoeffding's inequality, Bernstein's inequality, and Bennett's inequality.
-Particularly, Hoeffding's inequality implies the sub-Gaussianity of bounded variables.
-As the proof for Lemma~\ref{lem:bennett} shows, Bennett's inequality is at least as good as the Bernstein's inequality, for bounded random variables.
-\end{rmk}
-
-\subsection{Exercise 2.8}
-\begin{lem}[Bernstein and expectation] Let $Z$ be a nonnegative random variable satisfying the following concentration inequality
-\begin{equation}\label{eq:berncond1}
-	\bbP[Z \geq t] \leq C e^{- \frac{t^2}{2(\nu^2 +B t)}}, \quad \text{for all } t \geq 0,
-\end{equation}
-	where $(\nu, B)$ are two positive constants and $C \geq 1$. Then, the expectation of $Z$ satisfies 
-	\begin{equation}\label{eq:bernexpec}
-		\bbE[Z] \leq 2 \nu (\sqrt{\pi} + \sqrt{\log C}) + 4B (1 + \log C).
-	\end{equation}
-	Further, let $\{ X_i\}_{i=1}^n$ be a sequence of i.i.d.\ zero-mean variables satisfying the Bernstein condition \eqref{eq:berncond}. The sample mean of $\{ X_i\}_{i=1}^n$  satisfies
-	\begin{equation}\label{eq:e28sample}
-		\bbE \off{ \aabs{\frac{1}{n} \sum_{i=1}^n X_i }} \leq 2 \sigma (\sqrt{\pi} + \sqrt{\log 2}) + 4b (1 + \log 2).
-	\end{equation}
-\end{lem}
-
-\begin{proof}
-	First, we prove the equation~\eqref{eq:bernexpec}.
-
-	\vspace{0.2cm}
-	By equation~\eqref{eq:berncond1}, we have
-	\begin{align}
-		\bbP[Z \geq t] &\leq C \max \offf{\exp \of{-\frac{t^2}{4 \nu^2}}, \exp \of{-\frac{t^2}{4 Bt}} }\\
-		&\leq C \exp \of{-\frac{t^2}{4 \nu^2}} + C \exp \of{-\frac{t^2}{4 Bt}}.\label{eq:e28tail}
-	\end{align}
-
-	Plugging the inequality~\eqref{eq:e28tail} to $\bbE[Z] = \int_{0}^{+\infty} \bbP[Z \geq t] dt$, we have
-	\begin{equation}
-		\bbE[Z] = \int_{0}^{+ \infty} \min \offf{1, C \exp \of{-\frac{t^2}{4 \nu^2}}} dt + \int_{0}^{+ \infty} \min\offf{1, C \exp \of{-\frac{t^2}{4 Bt}}} dt   \colonequals I_1 + I_2.
-	\end{equation}
-
-	To evaluate $I_1$, we spilt the integral to avoid the minimization. Solving $1 = C \exp \of{-\frac{t^2}{4 \nu^2}} $, the minimization term becomes
-	\begin{align}
-		\min \offf{1, C \exp \of{-\frac{t^2}{4 \nu^2}}} = \begin{cases}
-			\quad 1 & \text{when } t < 2\nu \sqrt{\log C},\\
-			\quad C \exp \of{-\frac{t^2}{4 \nu^2}} & \text{when } t \geq 2\nu \sqrt{\log C}.
-		\end{cases}
-	\end{align}
-	Therefore, we evaluate $I_1$ as follows,
-	\begin{align}
-		I_1 &= \int_{0}^{ 2\nu \sqrt{\log C}} 1 dt + \int_{2\nu \sqrt{\log C}}^{+\infty} C \exp \of{-\frac{t^2}{4 \nu^2}} dt\\
-		(\text{let}\quad y = \frac{t}{2\nu} - \sqrt{\log C})\quad &=  2\nu \sqrt{\log C} + 2\nu\int_{0}^{+\infty} \exp \of{-y^2 - 2y \sqrt{\log C}} dy\\
-		&\leq  2\nu \sqrt{\log C} + 2\nu\int_{0}^{+\infty} \exp \of{-y^2} dy.
-	\end{align}
-	By Gaussian integral $\int_{0}^{+\infty} e^{-x^2} dx = \frac{\sqrt{\pi}}{2}$, we obtain $I_1 \leq 2\nu (\sqrt{\pi} + \sqrt{\log C})$. Similarly, we evaluate $I_2$ as follows 
-	\begin{align}
-		I_2 &= \int_{0}^{4B \log C} 1 dt + \int_{4B \log C}^{+\infty} C \exp \of{-\frac{t}{4B}} dt\\
-		 &= 4B (\log C + 1).
-	\end{align}
-	Hence, we obtain the expectation of $Z$,
-	\begin{equation}
-		\bbE[Z] = I_1 + I_2 \leq  2\nu (\sqrt{\pi} + \sqrt{\log C}) + 4B (\log C + 1).
-	\end{equation}
-
-	\vspace{0.2cm}
-	Next, we prove the equation~\eqref{eq:e28sample}.
-
-	\vspace{0.2cm}
-	For all $i \in [n]$, since $X_i$ satisfies the Bernstein condition with parameter $(\sigma,b)$, the variable $X_i$ satisfies the concentration bound~\eqref{eq:thmbernconc},
-	\begin{equation}
-		\bbP[|X_i| \geq t] \leq 2 \exp \offf{- \frac{t^2}{2(\sigma^2 + bt)}}, \quad \text{for all }t \geq 0. 
-	\end{equation}
-	By equation~\eqref{eq:bernexpec}, we have
-	\begin{equation}
-		\bbE[|X_i|] \leq  2 \sigma (\sqrt{\pi} + \sqrt{\log 2}) + 4b (1 + \log 2).
-	\end{equation}
-
-	Therefore, the expectation of the sample mean satisfies
-	\begin{equation}
-		\bbE \off{ \aabs{\frac{1}{n} \sum_{i=1}^n X_i }} \leq \frac{1}{n} \sum_{i=1}^n \bbE[|X_i|] \leq 2 \sigma (\sqrt{\pi} + \sqrt{\log 2}) + 4b (1 + \log 2).
-	\end{equation}
-
-\end{proof}
-
-\begin{rmk}
-	For a nonnegative random variables satisfying the Bernstein-type inequality \eqref{eq:berncond1}, the expectation of the variable is upper bounded by a function of the parameter $(\nu,B,C)$.
-Particularly, for a zero-mean random variable $X$ satisfying the Bernstein condition with parameter $b$, $|X|$ satisfies the inequality \eqref{eq:berncond1} with parameters $(\sigma,b,2)$, where $\sigma^2 = \var(X)$.
-Then, the expectation of the absolute variable $|X|$ is upper bounded by a function of $(\sigma^2, b)$.
-\end{rmk}
-
-\subsection{Exercise 2.9}\label{exe:29}
-\begin{lem}[Sharp upper bounds on binomial tails]
-	Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.\ Bernoulli variables with parameter $\alpha \in (0,1/2]$.
-Consider the binomial random variable $Z_n = \sum_{i=1}^n X_i$.
-The tail probability of $Z_n$ is upper bounded as
-	\begin{equation}\label{eq:lembinupper}
-		\bbP\off{Z_n \leq \delta n} \leq e^{-n D\of{\delta \Vert \alpha}}, \quad \text{for all }\delta \in (0,\alpha),
-	\end{equation}
-	where the quantity 
-	\begin{equation}\label{eq:kldiv}
-		D(\delta \Vert \alpha) \colonequals \delta \log \frac{\delta}{\alpha} + (1-\delta) \log \frac{1-\delta}{1-\alpha}
-	\end{equation}
-	is the KL divergence between the Bernoulli distributions with parameters $\delta$ and $\alpha$, respectively.
-
-	Further, the upper bound~\eqref{eq:lembinupper} is strictly tighter than the Hoeffding bound for all $\delta \in (0,\alpha)$.
-\end{lem}
-
-\begin{proof}
-    Applying the Chernoff inequality to the random variable $Z_n$, we have
-    \begin{equation}\label{eq:e29upper}
-    	\bbP[Z_n \leq \delta n] \leq \frac{\bbE\off{e^{\lambda Z_n}}}{e^{\lambda \delta n}} = \offf {\exp \offf{\log \of{1 - \alpha + \alpha e^{\lambda})}- \lambda \delta }}^n,\quad \text{for all } \lambda \in \bbR,
-    \end{equation}
-    where the second equality follows from the moment generating function of $Z_n$,
-    \[\bbE\off{e^{\lambda Z_n}} = \of{1 - \alpha + \alpha e^{\lambda}}^n. \]
-    The upper bound~\eqref{eq:e29upper} achieves the minimum when $\lambda = \log {\frac{\delta(1-\alpha)}{\alpha(1 - \delta)}}$ by the first-order condition.
-Plugging the minimizer into equation~\eqref{eq:e29upper}, we obtain the result
-    \[ \bbP\off{Z_n \leq \delta n}  \leq e^{-n D\of{\delta \Vert \alpha}}, \quad \text{for all }\delta \in (0,\alpha).  \]
-
-    \vspace{0.2cm}
-    Next, we show that equation~\eqref{eq:lembinupper} is strictly tighter than the Hoeffding bound of $Z_n$, for all $\delta \in (0,\alpha)$.
-
-    \vspace{0.2cm}
-    Hoeffding bound of random variable $Z_n$ is 
-    \begin{equation}\label{eq:e29hoeff}
-    	\bbP\off{Z_n \leq \delta n} \leq e^{ {-2n(\delta - \alpha)^2}}, \quad \text{for all }\delta \in (0,\alpha).
-    \end{equation}
-    Note that parameter $\alpha$ is fixed.
-Consider the function $g(\delta) = 2(\delta - \alpha)^2 - D\of{\delta \Vert \alpha}$ for $\delta \in (0,\alpha)$.
-The first-order derivative and  second-order derivative of $g$ are
-    \[ g'(\delta) = 4(\delta - \alpha) - \log \frac{\delta}{\alpha} + \log \frac{1-\delta}{1-\alpha} \quad \text{and} \quad g''(\delta) = 4 - \frac{1}{\delta} - \frac{1}{1-\delta}. \] 
-    Note that $ g(\alpha) = g'(\alpha) = 0$, and $g''(\delta) < 0$ for all $0 < \delta < 1/2$.
-Then, we have $g(\delta) < 0$, for all $\delta \in (0,\alpha)$.
-Hence, the upper bound~\eqref{eq:lembinupper} is strictly tighter than Hoeffding bound~\eqref{eq:e29hoeff}.
-
-\end{proof}
-
-\begin{rmk}
-	The upper bound on a binomial tail is a function of the KL divergence between Bernoulli distributions with distinct parameters.
-The bound using KL divergence is strictly tighter than the Hoeffding bound.
-The underperformance of Hoeffding bound may attribute to the utilization of the sub-Gaussian parameter $\sigma^2$.
-The sub-Gaussian parameter $\sigma^2$ is not necessarily the optimal choice to describe the tail performance of a random variable with good properties.
-\end{rmk}
-
-\subsection{Exercise 2.10}\label{exe:210}
-\begin{lem}[Lower bounds on binomial tails]
-	Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.\ Bernoulli variables with parameter $\alpha \in (0,1/2]$.
-Consider the binomial random variable $Z_n \colonequals \sum_{i=1}^n X_i$.
-For some fixed $\delta \in (0,\alpha)$, let $m = \floor{n\delta}$; i.e., $m$ is the largest integer less or equal to $n \delta$.
-Let $ \tilde \delta = {m}/{n}$.
-We have
-	\begin{equation}\label{eq:e2101}
-		\frac{1}{n} \log \bbP \off{Z_n \leq n\delta} \geq \frac{1}{n} \log \begin{pmatrix}
-			n \\ m	\end{pmatrix} + \tilde \delta \log \alpha + (1 - \tilde \delta ) \log (1-\alpha).
-	\end{equation}
-
-	Further, the binomial coefficient satisfies
-	\begin{equation}\label{eq:e2102}
-		\frac{1}{n} \log \begin{pmatrix}
-			n \\ m	\end{pmatrix} \geq \phi(\tilde \delta ) - \frac{\log (n+1)}{n}.
-	\end{equation}
-
-	Consequently, the lower tail of binomial variable $Z_n$ satisfies
-	\begin{equation}\label{eq:e2103}
-		\bbP[Z_n \leq n \delta]  \geq \frac{1}{n+1} e^{-nD(\delta \Vert \alpha)},
-	\end{equation}
-	where $D(\delta \Vert \alpha)$ is the KL divergence defined in equation~\eqref{eq:kldiv}.
-\end{lem}
-
-\begin{proof}
-	First, we prove equation~\eqref{eq:e2101}.
-
-	\vspace{0.2cm}
-	Since $Z_n$ is a binomial variable with size $n$ and probability $\alpha$, we have
-	\begin{align}
-		\bbP[Z_n \leq n\delta ] &= \sum_{k = 1}^m \begin{pmatrix}
-			n \\ k
-		\end{pmatrix} \alpha^k (1-\alpha)^{n - k}\\
-		& \geq \begin{pmatrix}
-			n \\ m
-		\end{pmatrix} \alpha^m (1-\alpha)^{n - m}. \label{eq:e210pro1}
-	\end{align}
-	Taking the log and dividing by $n$ on the both sides of the inequality~\eqref{eq:e210pro1} , we obtain
-	\[ \frac{1}{n} \log \bbP[Z_n \leq n\delta ] \geq \frac{1}{n} \log \begin{pmatrix}
-			n \\ m		\end{pmatrix} + \tilde \delta \log \alpha + (1 - \tilde \delta ) \log (1-\alpha). \]
-
-	\vspace{0.2cm}
-	Next, we prove equation~\eqref{eq:e2102}.
-
-	\vspace{0.2cm}
-	Consider a binomial variable $Y \sim Bin(n,\tilde \delta)$. For all $k = 0,1,...,(n-1)$, the ratio between $\bbP[Y = k]$ and $\bbP[Y = k+1]$ is
-	\begin{equation}
-		\frac{\bbP[Y = k+1]}{\bbP[Y = k]} = \frac{\begin{pmatrix}
-			n \\ k+1
-		\end{pmatrix} \tilde \delta^{k+1} (1-\tilde \delta)^{n - k -1} }{\begin{pmatrix}
-			n \\ k
-		\end{pmatrix} \tilde \delta^{k} (1-\tilde \delta)^{n - k }} = \frac{(n-k) \tilde \delta}{(k+1) (1- \tilde \delta)  }.
-	\end{equation}
-	To let the ratio $	\frac{\bbP[Y = k+1]}{\bbP[Y = k]} \geq 1$, we need $(k+1)\leq (n+1) \tilde \delta $.
-Thus, the probability $\bbP[Y = l]$ achieves the maximum when $l = n \tilde \delta$.
-Consequently, we have
-	\begin{equation}\label{eq:e210pro2}
-		(n+1) \bbP[Y = n \tilde \delta] \geq 1 \quad \Leftrightarrow \quad  \bbP[Y = n \tilde \delta] \geq \frac{1}{n+1}.
-	\end{equation}
-	Taking the log and dividing by $n$ on the both sides of the inequality~\eqref{eq:e210pro2}, we obtain
-	\begin{equation}
-		 \frac{1}{n} \log \begin{pmatrix}
-			n \\ m		\end{pmatrix} \geq \tilde \delta \log(\alpha) - (1 - \tilde \delta)\log(1 - \alpha) - \frac{\log(n+1)}{n} = \phi(\tilde \delta)  - \frac{\log(n+1)}{n}.
-	\end{equation}
-
-	\vspace{0.2cm}
-	Last, we prove the lower bound equation~\eqref{eq:e2103}.
-
-	\vspace{0.2cm}
-	Plugging equation~\eqref{eq:e2102} into equation~\eqref{eq:e2101}, we have
-	\begin{equation}\label{eq:e210pro3}
-		\frac{1}{n} \log \bbP \off{Z_n \leq n\delta} \geq \phi(\tilde \delta) + \tilde \delta \log \alpha + (1 - \tilde \delta ) \log (1-\alpha) - \frac{\log(n+1)}{n} \geq -D(\delta \Vert \alpha ) - \frac{\log(n+1)}{n}.
-	\end{equation}
-	Multiplying $n$ and exponentiating on the both sides of the inequality~\eqref{eq:e210pro3}, we obtain the result
-	\begin{equation}
-		\bbP \off{Z_n \leq n\delta} \geq \frac{1}{n+1} e^{-nD(\delta \Vert \alpha)}.
-	\end{equation}
-\end{proof}
-
-\begin{rmk}
-	The lower bound on a binomial tail is also a function of the KL divergence between Bernoulli distributions with distinct parameters.
-Combining the upper bound \eqref{eq:lembinupper} and lower bound \eqref{eq:e2103}, we conclude that the binomial tail is (upper and lower) bounded by the functions of  KL divergence between Bernoulli distributions.
-\end{rmk}
-
-\subsection{Exercise 2.11}
-\begin{lem}[Gaussian maxima]
-Let $\{X_i\}_{i=1}^n$ be a sequence of i.i.d.
-normal random variables following $N(0,\sigma^2)$.
-Consider the random variable $Z_n \colonequals \max_{i\in[n]} |X_i|$.
-Since the tail bound $\bbP[U \geq t] \leq \sqrt{\frac{2}{\pi}} \frac{1}{t} e^{-t^2 /2}$ holds for all standard normal random variable $U$, the expectation of $Z_n$ is upper bounded as follows
-\begin{equation}\label{eq:e2111}
-	\bbE[Z_n] \leq \sqrt{2 \sigma^2 \log n} + \frac{4\sigma}{\sqrt{2 \log n}}, \quad \text{for all } n\geq 2.
-\end{equation}
-\iffalse
-The lower bound of $\bbE[Z_n] $ is
-\begin{equation}\label{eq:e2112}
-	\bbE[Z_n] \geq (1- 1/e) \sqrt{2\sigma^2 \log n} , \quad \text{for all }n \geq 5.
-\end{equation}
-\fi
-
-\end{lem}
-
-\begin{proof}
-	%First, we prove the equation~\eqref{eq:e2111}.
-
-	\vspace{0.2cm}
-	Consider the tail probability of $Z_n$. We have
-	\begin{equation}
-		\bbP[Z_n \geq t] = \bbP[\max_{i \in [n]} |X_i| \geq t] = 1 - \bbP[\max_{i \in [n]} |X_i| < t] = 1 - (1- \bbP[|X_1| \geq t])^n,
-	\end{equation}
-	where the last equality follows from the independence of $\{X_i\}_{i=1}^n$. By Bernoulli's inequality, we have
-	\begin{equation}
-		(1- \bbP[|X_1| \geq t])^n \geq 1 - n \bbP[|X_1| \geq t], \quad \text{for all } t >0.
-	\end{equation} 
-	Therefore, the expectation of $Z_n$ follows
-	\begin{equation}\label{eq:e211pro1}
-		\bbE[Z_n] = \int_{0}^{+\infty} \bbP[Z_n \geq t] dt \leq  c +  \int_{c}^{+\infty} n\bbP[|X_1| \geq t] dt, \quad \text{for all } c > 0.
-	\end{equation}
-	 Since the tail bound $\bbP[U \geq t] \leq \sqrt{\frac{2}{\pi}} \frac{1}{t} e^{-t^2 /2}$ holds for all standard normal random variable $U$, the tail bound of $|X_1|$ satisfies
-	 \begin{equation}\label{eq:e211proupper}
-	 	\bbP[|X_1| \geq t] \leq 2 \sqrt{\frac{2}{\pi}} \frac{\sigma}{t} e^{-\frac{t^2}{2\sigma^2}}.
-	 \end{equation}
-	 Hence, the last integral in equation~\eqref{eq:e211pro1} follows
-	 \begin{align}
-	 	 \int_{c}^{+\infty} n\bbP[|X_1| \geq t] dt &\leq \frac{2n \sigma}{c} \sqrt{\frac{2}{\pi}} \int_{c}^{+\infty} e^{-\frac{t^2}{2\sigma^2} }dt \\ (\text{let } u = \frac{ t}{\sigma}) \quad  
-	 	 %& = \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}   \int_{\frac{c}{\sigma}}^{+\infty} e^{-\frac{u^2}{2}} dt \\
-	 	 & \leq \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}   \int_{\frac{c}{\sigma}}^{+\infty} u e^{-\frac{u^2}{2}} dt \\
-	 	 & = \frac{2n \sigma^2}{c} \sqrt{\frac{2}{\pi}}  e^{-\frac{c^2}{2\sigma^2}}. \label{eq:e211pro2}
-	 \end{align}
-	 For all $n \geq 2$, let $c = \sqrt{2\sigma^2 \log n}$ and plug the $c$ into equations~\eqref{eq:e211pro1} and \eqref{eq:e211pro2}. Then, we obtain the result
-	 \begin{equation}
-	 	\bbE[Z_n] \leq \sqrt{2\sigma^2 \log n} + \frac{2\sigma}{\sqrt{2\log n}}   \sqrt{\frac{2}{\pi}}  \leq \sqrt{2\sigma^2 \log n} + \frac{4\sigma}{\sqrt{2\log n}}, \quad \text{for all }n \geq 2.
-	 \end{equation}
-
-	 \iffalse
-	 \vspace{0.2cm}
-	 Next, we prove the equation~\eqref{eq:e2112}.
-
-	 \vspace{0.2cm}
-
-	 By Markov's inequality, the expectation of $Z_n$ satisfies
-	 \begin{equation}
-	 	\bbE[Z_n] \geq t \bbP[Z_n \geq t],\quad \text{for all }t \geq 0.
-	 \end{equation}
-	 By the independence of $\{X_i\}_{i=1}^n$, the tail probability of $Z_n$ is
-	 \begin{equation}
-	 	\bbP[Z_n \geq t] = 1 - (1 - \bbP(|X_1| \geq t))^n.
-	 \end{equation}
-
-	 Let $t = \sqrt{2\sigma^2 \log n}$. By equation~\eqref{eq:e211proupper}, the tail probability of $X_1$ satisfies
-	 \begin{equation}
-	 	\bbP(|X_1| \geq \sqrt{2\sigma^2 \log n}) \leq \frac{2}{\sqrt{ \pi \log n}} \frac{1}{n} \leq \frac{1}{n},\quad \text{for all } n \geq 5.
-	 \end{equation}	 
-	 Hence, the tail probability of $Z_n$ with $t = \sqrt{2\sigma^2 \log n} $satisfies
-	 \begin{equation}
-	 	\bbP[Z_n \geq t] \geq 1 - 
-	 \end{equation}
-	 \fi
-\end{proof}
-
-\subsection{Exercise 2.12}
-\begin{lem}[Sharp upper bounds for sub-Gaussian maxima]
-	Let $\{X_i\}_{i = 1}^n$ be a sequence of zero-mean sub-Gaussian variables with parameter $\sigma$. Then, we have
-	\begin{equation}
-		\bbE \off{\max_{i \in [n]} X_i} \leq \sqrt{2\sigma^2 \log n}, \quad \text{for all } n \geq 1.
-	\end{equation}
-	Note that independence assumptions are unnecessary.
-\end{lem}
-
-\begin{proof}
-	For all $t >0$, the function $f(x) = \exp \of{tx}$ is a convex function. Apply the function $f(x)$ to $\bbE \off{\max_{i \in [n]} X_i} $. By Jensen's inequality, we have
-	\begin{equation}\label{eq:e212}
-		\exp \offf{t \bbE \off{\max_{i \in [n]} X_i} } \leq \bbE\off{\exp \of{t \max_{i \in [n]} X_i }} \leq \sum_{i= 1}^n \bbE\off{\exp \of{t X_i }} \leq n e^{\frac{t^2 \sigma^2}{2}}, \quad \text{for all }t > 0,
-	\end{equation}
-	where the last inequality follows from the sub-Gaussianity of $X_i$s. Take the log of both sides of equation~\eqref{eq:e212}. We have
-	\begin{equation}\label{eq:e212b}
-		\bbE \off{\max_{i \in [n]} X_i}  \leq \frac{\log n}{t} + \frac{t\sigma^2}{2}.
-	\end{equation}
-	Plugging $t = \frac{\sqrt{2\log n} }{\sigma}$ into the equation~\eqref{eq:e212b}, we obtain
-	\[ \bbE \off{\max_{i \in [n]} X_i}  \leq \sqrt{2 \sigma^2 \log n}.\]
-\end{proof}
-
-\subsection{Exercise 2.13}
-\begin{lem}[Operations on sub-Gaussian variables]
-	Let $X_1$ and $X_2$ be two zero-mean sub-Gaussian variables with parameters $\sigma_1$ and $\sigma_2$ respectively. 
-	\begin{enumerate}
-		\item[(a).] If $X_1$ and $X_2$ are independent, then random variable $X_1 + X_2$ is sub-Gaussian with parameter $\sqrt{\sigma_1^2 + \sigma_2^2}$.
-		\item[(b).] Without independence assumptions, then random variable $X_1 + X_2$ is sub-Gaussian with parameter at most 
-		%$2 \sqrt{\sigma_1^2 +\sigma_2^2 }$.
-		$\sigma_1 + \sigma_2$.
-		\item[(c).] If $X_1$ and $X_2$ are independent, then random variable $X_1 X_2$ is sub-exponential with parameter $(\nu, b) = \of{\sqrt{2}\sigma_1 \sigma_2, \frac{1}{\sqrt{2}\sigma_1 \sigma_2}}$.
-	\end{enumerate}
-\end{lem}
-
-\begin{proof}
-	First, we prove $(a)$.
-
-	\vspace{0.2cm}
-	Since $X_1\sim \subG(\sigma_1)$ and $X_2 \sim \subG(\sigma_2)$ are independent, we have
-	\begin{equation}
-		\bbE\off{e^{t(X_1 + X_2)}} = \bbE\off{e^{tX_1}} \bbE\off{e^{tX_2}} \leq e^{\frac{t^2(\sigma_1^2 + \sigma_2^2)}{2}},\quad \text{for all }t \in \bbR.
-	\end{equation}
-	Therefore, $X_1 +X_2$ is sub-Gaussian with parameter $\sqrt{\sigma_1^2 + \sigma_2^2}$.
-
-	\vspace{0.2cm}
-Next, we prove $(b)$.
-
-\vspace{0.2cm}
-For all $t \in \bbR$, we have
-\begin{equation}\label{eq:e213b}
-	\bbE\off{e^{t(X_1 + X_2)}} = \bbE \off{e^{t X_1} e^{t X_2}}.
-\end{equation}
-Introduce $p,q \geq 1$ such that $\frac{1}{p} + \frac{1}{q} = 1$. By Holder's inequality, the equation~\eqref{eq:e213b} becomes
-\begin{equation}\label{eq:e213b2}
-	\bbE\off{e^{t(X_1 + X_2)}} \leq \of{\bbE \off{e^{ p t X_1}  } }^{\frac{1}{p}}
-	 \of{\bbE \off{e^{ pt X_2}  }}^{\frac{1}{q}} \leq e^{  \frac{pt^2 \sigma_1^2 + qt^2 \sigma_2^2}{2} },
-\end{equation}
-where the last inequality follows from the sub-Gaussianity of $X_1$ and $X_2$.
-The inequality~\eqref{eq:e213b2} achieves its minimum when $p = \sigma_2/\sigma_1 +1$.
-Plugging the minimizer to inequality~\eqref{eq:e213b2}, we obtain
-\begin{equation}
-	\bbE\off{e^{t(X_1 + X_2)}} \leq e^{\frac{t^2(\sigma_1 + \sigma_2
-	)^2}{2}}.
-\end{equation}
-Therefore, the random variable $X_1 + X_2$ is sub-Gaussian with parameter at most $\sigma_1 + \sigma_2$.
-
-\vspace{0.2cm}
-
-Last, we prove $(c)$.
-
-\vspace{0.2cm}
-
-\end{proof}
-
-\subsection{Exercise 2.14}
-\begin{lem}
-	Let $X$ be a scalar random variable. Suppose there exist two positive constant $c_1, c_2$ such that
-	\begin{equation}\label{eq:leme214}
-		\bbP \off{|X - \bbE[X]| \geq t} \leq c_1 e^{-c_2 t^2},\quad \text{for all }t\geq 0.
-	\end{equation}
-	Then, the variance $\var(X) \leq {c_1}/{c_2}$.
-
-	\vspace{0.2cm}
-	Let $m_X$ be the median of $X$; i.e., $\bbP[X \geq m_X] \geq 1/2$ and  $\bbP[X \leq m_X] \geq 1/2$.
-Note that the median $m_X$ is not necessarily unique.
-Given mean concentration~\eqref{eq:leme214}, for all median $m_X$, we have
-	\begin{equation}\label{eq:leme214b}
-		\bbP[|X - m_X| \geq t] \leq c_3 e^{-c_4t^2}, \quad \text{for all }t \geq 0,
-	\end{equation} 
-	where $c_3 \colonequals 4c_1$ and $c_4 \colonequals c_2/8$.
-
-	 Conversely, if the equation~\eqref{eq:leme214b} holds, the mean concentration~\eqref{eq:leme214} also holds with parameter $c_1 = 2c_3$ and $c_2 = c_4/4$.
-\end{lem}
-
-\begin{proof}
-	First, we show the variance  $\var(X) \leq {c_1}/{c_2}$.
-
-	\vspace{0.2cm}
-
-	By the definition of variance, we have
-	\begin{equation}\label{eq:e2141}
-		\var(X) = \bbE[(X-\bbE[X])^2] = \int_{0}^{+\infty} \bbP \off{|X - \bbE[X]| \geq \sqrt{t} }dt.
-	\end{equation}
-	Since $X$ satisfies the mean concentration equation~\eqref{eq:leme214}, for all $t > 0$, we have
-	\begin{equation}\label{eq:e2142}
-		\bbP \off{|X - \bbE[X]| \geq \sqrt{t} } \leq c_1 e^{-c_2 t}.
-	\end{equation}
-	Plugging the equation~\eqref{eq:e2142} into equation~\eqref{eq:e2141}, we obtain
-	\[ \var(X) \leq \frac{c_1}{c_2}. \]
-
-	\vspace{0.2cm}
-
-	Next, we show by an example that the median of $X$, $m_X$, is not necessarily unique.
-
-	\vspace{0.2cm}
-	Consider a random variable $Y \sim Bin(5,1/2)$.
-Then, we have $\bbP[Y \geq 3] = 1/2$ and $\bbP[Y \leq 3] > 1/2$.
-Meanwhile, we also have $\bbP[Y \geq 2] > 1/2$ and $\bbP[Y \leq 2] = 1/2$.
-By the definition of median, the number $2$ and $3$ both are the median of $Y$.
-Therefore, the median of a random variable is not necessarily unique.
-
-	\vspace{0.2cm}
-
-	Then, we prove the median concentration~\eqref{eq:leme214b}, provided that random variable $X$ satisfies the mean concentration~\eqref{eq:leme214}.
-
-	\vspace{0.2cm}
-
-	Let $\Delta = |\bbE[X] - m_X|$. Consider the following two cases.
-	\begin{enumerate}
-		\item[1.] Case 1: Suppose $t \geq 2\Delta$.
-
-		\vspace{0.2cm}
-		Note the triangle inequality $|X - \bbE[X]| \geq |X - m_X| - \Delta$ and the assumption $\frac{t}{2} \geq \Delta$. We have
-		\begin{equation}
-			\bbP[|X - m_X| \geq t] \leq \bbP \off{|X - m_X| \geq \frac{t}{2} + \Delta} \leq   \bbP \off{|X - \bbE[X]|  \geq \frac{t}{2}}.
-		\end{equation}
-		By equation~\eqref{eq:leme214}, we obtain
-		\begin{equation}\label{eq:e214case1}
-			\bbP[|X - m_X| \geq t] \leq c_1 e^{-\frac{c_2t^2}{4}}.
-		\end{equation}
-
-		\item[2.] Case 2:   Suppose $t < 2\Delta$.
-
-		\vspace{0.2cm}
-		By the definition of median, we have
-		\begin{equation}\label{eq:e214ca2a}
-			\bbP[|X - \bbE[X]| \geq \Delta] \geq \bbP[X \geq m_X] \geq \frac{1}{2}.
-		\end{equation}
-		Meanwhile, by equation~\eqref{eq:leme214}, we have
-		\begin{equation}\label{eq:e214ca2b}
-			\bbP[|X - \bbE[X]| \geq \Delta] \leq c_1 e^{-c_2 \Delta^2}.
-		\end{equation}
-
-		Combing the assumption $\Delta > \frac{t}{2}$  and inequalities~\eqref{eq:e214ca2a} \eqref{eq:e214ca2b}, we obtain
-		\begin{equation}\label{eq:e214case2}
-			\bbP[|X - m_X| \geq t]\leq 1\leq 2c_1  e^{-c_2 \Delta^2} \leq 2c_1  e^{-\frac{c_2 t^2}{4}}. 
-		\end{equation}
-
-	\end{enumerate}
-
-	Hence, combing the inequality~\eqref{eq:e214case1} and inequality~\eqref{eq:e214case2}, we conclude that
-	\begin{equation}
-		\bbP[|X - m_X| \geq t] \leq c_3 e^{-c_4 t^2}, 
-	\end{equation}
-	where $c_3 = 2c_1$ and $c_4 = c_2/4$. The upper bound also holds when $c_3 = 4 c_1$ and $c_4 = c_2/8$.
-
-	\vspace{0.2cm}
-
-	Last, we prove the mean concentration~\eqref{eq:leme214}, provided that the random variable $X$ satisfies the median concentration~\eqref{eq:leme214b}.
-
-	\vspace{0.2cm}
-
-	As the proof for median concentration~\eqref{eq:leme214b}, we also consider the following two cases.
-	\begin{enumerate}
-		\item[1.] Case 1: Suppose $t \geq 2\Delta$.
-
-		\vspace{0.2cm}
-		Note the triangle inequality $|X - m_X| + \Delta \geq |X - \bbE[X]|$ and the assumption $\frac{t}{2} \geq \Delta$. We have
-		\begin{equation}
-			\bbP[|X - \bbE[X]| \geq t ] \leq \bbP \off{|X - m_X| \geq \frac{t}{2}} \leq c_3 e^{-\frac{c_4 t^2}{4}},
-		\end{equation}
-		where the last inequality follows from the median concentration~\eqref{eq:leme214b}.
-
-		\item[2.] Case 2: Suppose $t < 2\Delta$.
-	\end{enumerate}
-	\end{proof}
-
-	\subsection{Exercise 2.15}
-	\begin{lem}
-		Let $\{X_i\}_{i=1}^n$ be i.i.d.\ sample of random variables with density $f$ on the real line. A standard estimate of $f$ is the kernel density estimate 
-		\begin{equation}
-			\hat f_n (x) = \frac{1}{nh} \sum_{i=1}^n  K\of{ \frac{x- X_i}{h} },
-		\end{equation}
-		where $K \colon \bbR \mapsto [0,+\infty )$ is a kernel function satisfying $\int_{-\infty}^{+\infty} K(x) dx = 1$, and $h > 0$ is a bandwidth parameter.
-Suppose we assess the estimate $\hat f_n (x)$ using the $L^1$ norm; i.e.,
-		\[ \onorm{\hat f_n - f}_1 = \int_{-\infty}^{+\infty} |\hat f_n(t) - f(t)|dt. \]
-		The upper tail probability of the $L^1$ norm satisfies
-		\begin{equation}
-			\bbP\off{\onorm{\hat f_n - f}_1 - \bbE \off{\onorm{\hat f_n - f}_1}\geq \delta} \leq e^{-\frac{n\delta^2}{8}},\quad \text{for all } \delta \geq 0.
-		\end{equation}
-	\end{lem}
-
-	\begin{proof}
-		Note that $\onorm{\hat f_n - f}_1$ is a function of $X = (X_1,...,X_n)$, denoted $g(X) = g(X_1,...,X_n) = \onorm{\hat f_n - f}_1$ and $g\colon \bbR^n \mapsto \bbR$.
-Let $x^{(k)} = (x_1,...,x_k,...,x_n) \in \bbR^n$ and $x^{'(k)} = (x_1,...,x'_k,...,x_n)\in \bbR^n$ be two vectors.
-The absolute difference between the $g(x^{(k)})$ and $g(x^{'(k)})$ is
-		\begin{align}
-			|g(x^{(k)}) - g(x^{'(k)})|
-			&= \aabs{\int_{-\infty}^{+\infty} \aabs{ \frac{1}{nh} \sum_{i = 1}^n K \of{\frac{t-x_i}{h}} -f(t)  } -\aabs{ \frac{1}{nh} \sum_{i \neq k}^n K \of{\frac{t-x_i}{h}} -f(t) + \frac{1}{nh} K\of{\frac{t-x'_k}{h}}   } dt} \\
-			&\leq \int_{-\infty}^{+\infty} \aabs{\frac{1}{nh} \of{K\of{\frac{t-x_k}{h}} - K\of{\frac{t-x'_k}{h}}  } }dt\\
-			&\leq \frac{1}{nh} \of{ \int_{-\infty}^{+\infty} K\of{\frac{t-x_k}{h}}dt + \int_{-\infty}^{+\infty} K\of{\frac{t-x'_k}{h}}dt }.  \label{eq:e215int}
-		\end{align}
-
-		Let $t_1 = \frac{t-x_k}{h}$. By the definition of $K$, the first integral in equation~\eqref{eq:e215int} are
-		\begin{equation}
-			\int_{-\infty}^{+\infty} K\of{\frac{t-x_k}{h}}dt = h\int_{-\infty}^{+\infty} K\of{t_1}dt_1 = h.
-		\end{equation}
-		Similarly, the second integral $\int_{-\infty}^{+\infty} K\of{\frac{t-x'_k}{h}}dt = h$. Then, we conclude the function $g$ satisfies
-		\begin{equation}\label{eq:e215diff}
-			|g(x^{(k)}) - g(x^{'(k)})|  \leq \frac{2}{n}.
-		\end{equation}
-		The inequality~\eqref{eq:e215diff} also holds for all $k \in [n]$ and for all $x^{(k)},x^{'(k)} \in \bbR^n$.
-Therefore, $g(X)$ satisfies the bounded difference property~\eqref{eq:defbounded} with parameters $(\frac{2}{n},..., \frac{2}{n})$.
-By Theorem~\ref{thm:bounddiff}, we obtain
-		\begin{equation}
-			\bbP\off{\onorm{\hat f_n - f}_1 - \bbE \off{\onorm{\hat f_n - f}_1} \geq \delta} \leq e^{-\frac{n\delta^2}{2}} \leq e^{-\frac{n\delta^2}{8}} .
-		\end{equation}
-
-	\end{proof}
-
-    \subsection{Exercise 2.16}
-    \begin{lem}
-    	Let $\{X_i\}_{i=1}^n$ be a sequence of independent variables taking values in a Hilbert space $\mathbb{H}$.
-Suppose that $\onorm{X_i}_{\mathbb{H}} \leq b_i$ almost surely, for all $i \in [n]$.
-Consider the real valued random variable $S_n = \onorm{\sum_{i=1}^n X_i}_{\mathbb{H}}$.
-The concentration bound of $S_n$ is
-    	\begin{equation}\label{eq:e216a}
-    		\bbP\off{|S_n - \bbE[S_n]| \geq n\delta} \leq 2e^{-\frac{n\delta^2}{8b^2}}, \quad \text{for all } \delta \geq 0,
-    	\end{equation}
-    	where $b^2 = \frac{1}{n} \sum_{i=1}^n b_i^2$. The upper tail probability bound of $S_n$ is
-    	\begin{equation}\label{eq:e216b}
-    		\bbP \off{\frac{S_n}{n} \geq a + \delta} \leq e^{-\frac{n\delta^2}{8b^2}}, \quad \text{for all }\delta \geq 0,
-    	\end{equation}
-    	where $a = \sqrt{\frac{1}{n} \sum_{i=1}^n \bbE \off{\onorm{X_i}^2_{\mathbb{H}}} }$.
-    \end{lem}
-
-    \begin{proof} First, we prove the concentration bound~\eqref{eq:e216a}.
-
-    \vspace{0.2cm}
-    	Note that $S_n$ is a function of $X=(X_1,...,X_n)$, denoted $S_n(X) = S_n(X_1,...,X_n)$, and $S_n(X) \colon \bbR^n \mapsto \bbR$.
-Let $x^{(k)} = (x_1,...,x_k,...,x_n) \in \bbR^n$ and $x^{'(k)} = (x_1,...,x'_k,...,x_n)\in \bbR^n$ be two vectors.
-The absolute difference between $S_n(x^{(k)})$ and $S_n(x^{'(k)})$ is
-    	\begin{align}
-    		\aabs{S_n(x^{(k)}) - S_n(x^{'(k)})} &= \aabs{\onorm{x_1 + \cdots +x_k + \cdots + x_n}_{\mathbb{H}} - \onorm{x_1 + \cdots +x'_k + \cdots + x_n}_{\mathbb{H}} }\\
-    		& \leq \onorm{x_k - x'_k}_{\mathbb{H}}\\
-    		& \leq 2b_k, \label{eq:e216diff}
-    	\end{align}
-    	where the last inequality follows by the boundedness $\onorm{X_k}_{\mathbb{H}} \leq b_k$, for all $k \in [n]$.
-The inequality~\eqref{eq:e216diff} also holds for all $k \in [n]$ and for all $x^{(k)},x^{'(k)} \in \bbR^n$.
-Therefore, $S_n(X)$ satisfies the bounded property~\eqref{eq:defbounded} with parameters $(2b_1,...,2b_n)$.
-By Theorem~\ref{thm:bounddiff}, we obtain
-    	\begin{equation}
-    		\bbP\off{|S_n - \bbE[S_n]| \geq n\delta} \leq 2 e^{- \frac{n \delta^2}{2b^2}} \leq 2 e^{- \frac{n \delta^2}{8b^2}}, \quad \text{for all }\delta \geq 0.
-    	\end{equation}
-
-    	\vspace{0.2cm}
-    	Next, we prove the upper tail probability bound~\eqref{eq:e216b}.
-
-    	\vspace{0.2cm}
-    	The expectation of $S_n(X) = \bbE \off{ \sum_{i=1}^n  \onorm{X_i}_{\mathbb{H}}}$ satisfies
-    	\begin{equation}\label{eq:e216ineq}
-    		 \bbE \off{ \onorm{ \sum_{i=1}^n  X_i}_{\mathbb{H}} } \leq \sqrt{ \bbE \off{ \onorm{\sum_{i=1}^n X_i }_{\mathbb{H}}^2  }} = \sqrt{ \sum_{i=1}^n \bbE \off{ \onorm{X_i}^2_{\mathbb{H}} } } = na,
-    	\end{equation}
-    	where the first inequality follows by the fact that $\bbE[X^2] \geq (\bbE[X] )^2$ and the second equation follows by the independence of $\{X_i\}_{i=1}^n$.
-Therefore, combing the inequality~\eqref{eq:e216ineq} with the concentration bound~\eqref{eq:e216a}, we obtain
-    	\begin{equation}
-    		\bbP \off{\frac{S_n}{n} \geq a + \delta} = \bbP \off{S_n - na \geq n\delta} \leq \bbP \off{S_n - \bbE[S_n] \geq n\delta } \leq e^{-\frac{n\delta^2}{8 b^2}}.
-    	\end{equation}
-    \end{proof}
-
-    \subsection{Exercise 2.17}
-    \begin{lem}[Hanson-Wright Inequality]
-   Let $\mQ = \entry{\mQ_{ij}}  \in \bbR^{n \times n} $ be a positive semi-definite matrix, and $\{X_i\}_{i = 1}^n$ be i.i.d.\ random variables with mean zero, variance 1 and $\sigma-$sub-Gaussian.
-Consider the random variable $Z = \sum_{i= 1}^n \sum_{j = 1}^n \mQ_{ij} X_i X_j$.
-The Hanson-Wright inequality for $Z$ is
-   \begin{equation}
-   	\bbP \off{Z \geq \tr (\mQ) + \sigma t } \leq 2 e^{- \min \offf{\frac{c_1 t}{\onorm{\mQ}_{op}  } , \frac{c_2 t^2}{\onorm{\mQ}^2_F} }}, \quad \text{for all } t \geq 0,
-   \end{equation}
-   where $\onorm{\mQ}_{op}$ is the operator norm of matrix $\mQ$, $\onorm{\mQ}_F$ is the Frobenius norm of matrix $\mQ$, and $c_1,c_2$ are two universal constants independent with $n$ and $t$.
-    \end{lem}
-
-    The proof for Hanson-Wright Inequality under the special case that $X_i \stackrel{i.i.d.}{\sim} N(0,1)$ is below. Let $\sigma  = 1$.
-
-    \begin{proof}
-    	Since the random variables $\{X_i\}_{i=1}^n$ are independent and $\var(X_i) = 1$ for all $i \in [n]$, the expectation $\bbE[Z] = \sum_{i = 1}^{n} \mQ_{ii} = \tr(\mQ)$.
-Then, the upper tail probability of $Z$ becomes
-    	\begin{equation}
-    		\bbP \off{ Z \geq \tr(\mQ) +  t } = \bbP \off{ \of{  \sum_{i = 1}^{n} \mQ_{ii} (X^2_i - 1) + \sum_{i \neq j} \mQ_{ij} X_i X_j }\geq t   }
-    	\end{equation}
-    	For two random variables $X,Y$ and a constant $t$, we have the inequality $ \bbP[X+Y \geq t] \leq \bbP[X \geq t/2] + \bbP[Y \geq t/2]$. Hence, the tail probability is upper bounded as
-    	\begin{equation}
-    		\bbP \off{ Z \geq \tr(\mQ) +  t } \leq \bbP\off{ \sum_{i =1 }^n \mQ_{ii}(X_i^2 - 1) \geq \frac{t}{2}  } + \bbP \off{ \sum_{i \neq j} \mQ_{ii} X_i X_j  \geq \frac{t}{2}} \equalscolon p_1 + p_2.
-    	\end{equation}
-
-    	Consider $p_1$.
-For a standard normal variable $X$,  the squared variable $X^2 \sim \chi^2_1$ and thus $X^2$ is sub-exponential with parameter $(\nu, b) = (2,4)$.
-Given a constant $c$, $cX^2$ is still sub-exponential with parameters $(2 c, 4 c^2)$.
-Hence, for all $i \in [n]$, the variable $\mQ_{ii} X_i^2$ is sub-exponential with parameters $(2 \mQ_{ii}, 4 \mQ^2_{ii} )$.
-By the Bernstein-type inequalities(non-asymptotic 5.16), the probability $p_1$ is upper bounded as following,
-    	\begin{equation}
-    		p_1 = \bbP \off{\sum_{i=1}^n { \of{ \mQ_{ii}X^2_i - \mQ_{ii}\bbE[X^2_i]   } \geq \frac{t}{2} }} \leq \exp \of{ - \min \offf{  \frac{c t^2}{\sum_{i}^n \mQ^2_{ii} }, \frac{c't}{\max_{i \in [n]  }|\mQ_{ii}| } }},
-    	\end{equation}
-    	where $c$ and $c'$ are two constants independent with $n$ and $t$.
-
-    	\vspace{0.2cm}
-
-    	Consider $p_2$.
-For two i.i.d.\ standard normal variables $X,Y$, the product $XY = \frac{1}{4}(X+Y)^2 - \frac{1}{4}(X-Y )^2$, where $ \frac{1}{2} (X+Y)^2$ and $\frac{1}{2} (X-Y)^2$ follow the chi-square distribution $\chi^2_1$ independently.
-Then, $XY$ is the subtraction of two independent sub-exponential variables.
-Let $D_1$ and $D_2$ be two identical and independent sub-exponential variables with parameters $(\nu, b)$.
-We have
-    	\[ \bbE[e^{\lambda (D_1-D_2)}] = \bbE[e^{\lambda D_1}] \bbE[e^{-\lambda D_2}] \leq e^{v^2 \lambda^2}, \quad \text{for all } |\lambda | < \frac{1}{b}.  \]
-    	Therefore, the subtraction of two identical and independent sub-exponential variables $D_1 -D_2$ is still sub-exponential with parameters $(\sqrt{2}\nu, b)$.
-
-    	\vspace{0.2cm}
-    	Back to $\mQ_{ij} X_i X_j$, where $i \neq j$.
-By the discussion above, for all $i\neq j$, the variable $\mQ_{ij} X_i X_j$ is sub-exponential with parameters $(\sqrt{2} \mQ_{ij}, \mQ_{ij}^2)$.
-By the Bernstein-type inequalities, the probability $p_2$ is upper bounded as following,
-    	\begin{equation}
-    		p_2 = \bbP \off{ \sum_{i \neq j} \mQ_{ii} X_i X_j  \geq \frac{t}{2}} \leq \exp \of{ - \min \offf{  \frac{c'' t^2}{\sum_{i\neq j} \mQ^2_{ij} }, \frac{c'''t}{\max_{i\neq j   }|\mQ_{ij}| } }},
-    	\end{equation}
-    	where $c''$ and $c'''$ are two constants independent with $n$ and $t$.
-
-    	\vspace{0.2cm}
-
-    	Therefore, by the definition of Frobenius norm and the fact that $\max_{i,j\in[n]}|\mQ_{ij}| \leq \onorm{\mQ}_{op} $, we obtain the goal inequality
-    	\begin{equation}
-    		\bbP \off{ Z \geq \tr(\mQ) +  t } \leq p_1 + p_2 \leq 2\exp \of{ - \min \offf{  \frac{c_2 t^2}{\onorm{\mQ}^2_F}, \frac{c_1t}{\onorm{\mQ}_{op} }} },
-    	\end{equation}
-    	for some universal constants $c_1,c_2$ independent with $n$ and $t$.
-    \end{proof}
-
-    \subsection{Exercise 2.18}
-
-    \begin{defn}[Orlicz Norms]\label{def:orlicz}
-     Let $\psi \colon \bbR_+ \mapsto \bbR_+$ be a strictly increasing convex function satisfying $\psi(0) = 0$. The $\psi$-Orlicz norm of a random variable $X$ is defined as
-    \begin{equation}
-    	\onorm{X}_{\psi} \colonequals \inf \offf{t > 0\colon \bbE\off{\psi \of{t^{-1} |X| } } \leq 1  }.
-    \end{equation}
-    The norm $\onorm{X}_{\psi}$ is infinite if there is no finite $t$ such that the expectation $ \bbE\off{\psi \of{t^{-1} |X| } }$ exists.
-    \end{defn}
-
-     Given the function $\psi_q(u) = u^{q}$ for some $q \in [1,+\infty)$, the Orlicz norm is simply the $l_q$-norm $\onorm{X}_q = \of{ \bbE[|X|^q] }^{1/q}$.
-
-    \begin{lem}
-    Let $\psi_q(u) = \exp(u^q) - 1$ be a function for $u \in \bbR_+$ with positive parameter $q \in [1,+\infty)$, and $\onorm{\cdot}_{\psi_q}$ denote the $\psi_q$-Orlicz norm. Then,
-    \begin{enumerate}
-    	\item[(a.)] If $\onorm{X}_{\psi_q} < +\infty$, there exists two positive constants $c_1,c_2$ such that
-    	\begin{equation}\label{eq:lemorlicztail}
-    		\bbP[|X| > t] \leq c_1 \exp \of{-c_2 t^q},\quad \text{for all } t > 0.
-    	\end{equation}
-    	In particular, the constants $c_1 = 2$ and $c_2 = \onorm{X}_{\psi_q}$.
-
-    	\item[(b.)] If the random variable $X$ satisfies the tail bound~\eqref{eq:lemorlicztail}, the Orlicz norm $\onorm{X}_{\psi_q} < +\infty$.
-    \end{enumerate}
-    \end{lem}
-    \begin{proof}
-    	First, we prove $(a)$.
-
-    	\vspace{0.2cm}
-
-    	Let $t_0$ denote the Orlicz norm $\onorm{X}_{\psi_q}$.
-Since $t_0 < +\infty$, the expectation $ \bbE \off{ \exp \of{\lambda |X|^q} }$ exists for $\lambda \in  [0,t^{-q}_0]$.
-Applying Chernoff's bound to $|X|^q$, the upper tail bound of $|X|$ becomes
-    	\begin{equation}\label{eq:e218tailbound}
-    		\bbP [|X|>t] = \bbP \off{\frac{|X|^q}{t^q_0} > \frac{t^q}{t^q_0}} \leq e^{-\lambda t^q} \bbE \off{ \exp \of{\lambda |X|^q} }, \quad \text{for all } t\geq 0, \lambda\in [0,t^{-q}_0].
-    	\end{equation}
-    	 By the definition of Orlicz norm, the expectation satisfies
-    	\begin{equation}\label{eq:e218ep}
-    		\bbE \off{ \exp \of{ t^{-q}_0|X|^q} } = \bbE \off{ \psi( t^{-1}_0 |X| )} + 1 \leq 2.
-    	\end{equation}
-    	Let $\lambda = t^{-q}_0$ and plug the expectation~\eqref{eq:e218ep} to the inequality~\eqref{eq:e218tailbound}. Then, we obtain the tail bound
-    	\begin{equation}
-    		\bbP [|X|>t] \leq 2 e^{- \onorm{X}^{-q}_{\psi_q} t^q },\quad 
-    		\text{for all }  t \geq 0.
-    	\end{equation}
-
-    	\vspace{0.2cm}
-    	Next, we prove $(b)$.
-
-    	\vspace{0.2cm}
-     Let $t_1$ be a positive constant. 	Note the fact that $\bbE[X] = \int_{-\infty}^{+\infty} \bbP(X >t) dt  $. The expectation of $\bbE[ \exp(t^{-q}_1 |X| )  ]$ is
-    	\begin{align}
-    		\bbE[ \exp(t^{-q}_1 |X|^q )  ] &= \int_{0}^{+\infty} \bbP \off{ \exp \of{\frac{|X|^q}{t^q_1}} > \exp \of{\frac{t^q}{t^q_1}} } dt \\
-    		& = \int_{0}^{+\infty} \bbP \off{ |X| > t } dt \\
-    		& \leq  \int_{0}^{+\infty} c_1 \exp(-c_2 t^q) dt\label{eq:e218integral},
-       	\end{align}
-       	where the last inequality follows from the tail bound~\eqref{eq:lemorlicztail}. Let $u = c_2 t^q$. The integral in equation~\eqref{eq:e218integral} becomes
-       	\[  \int_{0}^{+\infty} c_1 \exp(-c_2 t^q) dt = \frac{c_1}{q c_2^{1/q}} \int_{0}^{+\infty} u^{1/q-1}e^{-u} du =  \frac{c_1}{q c_2^{1/q}} \Gamma(1/q) < +\infty.  \]
-       	Therefore, the expectation 
-       	\begin{equation}
-       		\bbE \off{ \psi(t_1^{-1} |X| )} =  \bbE[ \exp(t^{-q}_1 |X|^q )  ] - 1 < +\infty, \quad \text{for all } t_1 > 0,
-       	\end{equation}
-       	which implies the Orlicz norm $\onorm{X}_{\psi_q}$ is definite.
-    \end{proof}
-
-\subsection{Exercise 2.19}
-\begin{lem}[Maxima of Orlicz variables] Recall the definition of Orlicz norm~\ref{def:orlicz}.
-Let $\{X\}_{i=1}^n$ be a sequence of i.i.d.\ zero-mean random variables.
-Given a positive, increasing and convex function $\psi \colon \bbR_+ \mapsto \bbR_+$, the random variable $X_i$ has a finite Orlicz norm $\sigma = \onorm{X_i}_{\psi}$, for all $i \in [n]$.
-Then, the maxima of $X_i$ satisfies
-\begin{equation}
-	\bbE \off{ \max_{i\in[n]} |X_i| } < \sigma \psi^{-1}(n).
-\end{equation}
-
-\end{lem}
-
-\begin{proof}
-	Since $\psi$ is a convex function, by Jensen's inequality, we have
-	\begin{equation}\label{eq:e219max}
-		\psi \of{\bbE \off{\max_{i \in [n]} \frac{|X_i|}{\sigma}  } } \leq \bbE \off{ \max_{i\in[n]} \psi \of{\frac{|X_i|}{\sigma}}   } \leq n \bbE \off{ \psi \of{\frac{|X_1|}{\sigma}} }.
-	\end{equation}
-	By the definition of Orlicz norm, we have $\bbE \off{ \psi \of{\frac{|X_1|}{\sigma}} } \leq 1$.
-Therefore, taking the inverse function $\psi^{-1}$ on both sides of inequality~\eqref{eq:e219max}, we obtain
-	\[  \bbE \off{\max_{i \in [n]} {|X_i|} } \leq \sigma \psi^{-1}(n).  \]
-\end{proof}
-
-\subsection{Exercise 2.20}
-
-\begin{lem}
-	Let $\{X_i\}_{i=1}^n$ be a sequence of independent zero-mean random variables. Suppose that for some fixed integer $m \geq 1$, the random variables satisfy the following moment bound,
-\begin{equation}\label{eq:e220mom}
-	\onorm{X_i}_{2m} \colonequals \of{\bbE[X^{2m}_i] }^{\frac{1}{2m}} \leq C_m, \quad \text{for all } i \in [n].
-\end{equation}
-Then, we have the tail bound
-\begin{equation}
-	\bbP \off{\aabs{ \frac{1}{n}  \sum_{i=1}^n X_i } \geq \delta } \leq B_m \of{\frac{1}{\sqrt{n} \delta}}^{2m}, \quad \text{for all } \delta >0,
-\end{equation}
-where $B_m$ is a universal constant depending on $C_m$ and $m$ only.
-\end{lem}
-
-\textit{Hint:} Let $\{X_i\}_{i=1}^n$ be a sequence of independent zero-mean random variables. The Rosenthal’s inequality is
-\begin{equation}\label{eq:rosenthal}
-	\bbE \off{\of{\sum_{i=1}^n X_i}^{2m}} \leq R_m \offf{\sum_{i=1}^n \bbE \off{X^{2m}_i} + \of{\sum_{i=1}^n \bbE [X^2_i]}^m },
-\end{equation}
- where $R_m$ is a universal constant depending on $m$ only.
-
- \begin{proof}
- 	Applying the $2m$-order central moment version of the Markov's inequality to $\aabs{ \frac{1}{n}  \sum_{i=1}^n X_i } $, we have the tail bound
- 	\begin{equation}\label{eq:e220tail}
- 		\bbP \off{\aabs{ \frac{1}{n}  \sum_{i=1}^n X_i } \geq \delta } = \bbP \off{\aabs{  \sum_{i=1}^n X_i } \geq n \delta } \leq (n\delta)^{-2m} \bbE \off{ \of{\sum_{i=1}^n X_i}^{2m} }.
- 	\end{equation}
-
- 	By the moment bound, for all $i\in [n]$, we have
- 	\begin{equation}\label{eq:e220inequ}
- 		\bbE \off{X^2_i}  \leq \bbE \off{X^{2m}_i} \leq C^{2m}_m.
- 	\end{equation}
- 	Plugging the inequality~\eqref{eq:e220inequ} to the Rosenthal's inequality~\eqref{eq:rosenthal}, we obtain
- 	\begin{equation}\label{eq:e220exp}
- 		\bbE \off{\of{\sum_{i=1}^n X_i}^{2m}} \leq R_m\of{nC^{2m}_m  + n^m C^{2m^2}_m  }. 
- 	\end{equation}
- 	Combining the inequality~\eqref{eq:e220exp} with tai bound~\eqref{eq:e220tail}, we have
- 	\begin{equation}
- 		\bbP \off{\aabs{ \frac{1}{n}  \sum_{i=1}^n X_i } \geq \delta } \leq \frac{R_m\of{nC^{2m}_m  + n^m C^{2m^2}_m  } }{n^{2m} \delta^{2m}} = B_m\of{ \frac{1}{\sqrt{n} \delta}}^{2m}, \quad \text{for all }\delta > 0,
- 	\end{equation}
- 	where $B_m$ is a universal constant depending on $C_m$ and $m$ only.
- \end{proof}
-
- \subsection{Exercise 2.21}
- \begin{defn}
- 	Let $X=(X_1,...,X_n)$ be a sequence of i.i.d.\ Bernoulli variables with parameter $1/2$.
-Represent X using a codebook of $N$ binary vectors, denoted $\{z^1,...,z^N\}$, where $z^j = \entry{z^j_i} \in \bbR^n$ is a vector, for all $j \in [N]$.
-The rescaled Hamming distortion is defined as
- 	\begin{equation}
- 		\delta \colonequals \bbE \off{ \min_{j \in [N]} \rho_H(X, z^j) } = \bbE \off{ \min_{j \in [N]} \frac{1}{n} \sum_{i = 1}^n  \mathds{1} \off{X_i \neq z^j_i} }. 
- 	\end{equation}
- \end{defn}
-
- The goal of lossy data compression is to find a representation of $X$ with rescaled Hamming distortion as small as possible.
-When we use a codebook with $N = 2^n$ codewords, we always achieve the 0 distortion.
-However, obtaining the distortion with $N = 2^n$ codewords is computationally expensive.
-Our goal is find a proper rate  $R < 1$ such that using a codebook with $N = 2^{Rn}$ codewords achieves a good compression performance.
-
- \begin{lem}
-
- 	Let $X=(X_1,...,X_n)$ be a sequence of i.i.d.\ Bernoulli variables with parameter $1/2$. Let $\delta \in (0,1)$. Suppose that the rate $R$ is upper bounded as
- 	\begin{equation}
- 		R < D_2(\delta || 1/2 ) = \delta \log_2 \frac{\delta}{1/2} + (1-\delta) \log_2 \frac{1-\delta}{1/2}.
- 	\end{equation}
-
- 	Let  $\Delta_R \colonequals R - D_2(\delta || 1/2 )$. If $\Delta_R < 0$, 
- 	then the probability of achieving distortion $\delta$ goes to 0 as $n$ goes to the infinity, for all codebook $\{z^1,...,z^N\}$.
-If $\Delta_R > 0$, the probability of achieving distortion $\delta$ goes to 1 as $n$ goes to the infinity, for all codebook $\{z^1,...,z^N\}$.
-
- \end{lem}
-
- \begin{proof} First, we prove the probability of achieving the distortion under $\Delta_R < 0$.
-
- \vspace{0.2cm}
-
- 	Consider the random variables $V^j = \mathds{1}\off{\rho_H(X,z^j) \leq \delta}  $, for all $j \in [N]$.
-Given the codeword $z^j$, the variable $\mathds{1} \off{X_i - z^j_i}$ is still a Bernoulli random variable.
-Recalling the tail bound~\eqref{eq:lembinupper} in Exercise \ref{exe:29}, we have
- 	\begin{equation}
- 		\bbP \off{V^j = 1} = \bbP \off{\sum_{i=1}^n \mathds{1} \off{X_i - z^j_i} \leq n\delta} \leq e^{-nD(\delta || 1/2)}, \quad \text{for all }j\in [N].
- 	\end{equation}
- 	Since the sum of $V^j$, denoted $V = \sum_{j=1}^N V^j$, follows the Binomial distribution $Bin(N, \bbP \off{V^j = 1} )$, the probability of not achieving the distortion $\delta$ is
- 	\begin{equation}\label{eq:e221av0}
- 		\bbP[V = 0] \geq \of{1 - e^{-nD(\delta || 1/2)}}^N.
- 	\end{equation}
-  Let $R_0 = D_2(\delta || 1/2)$ and $N_0 = 2^{R_0 n} = (2\delta)^{n\delta}(2(1-\delta))^{n(1-\delta)}$.
-By the definition of KL divergence, we have $e^{-nD(\delta || 1/2)} = N^{-1}_0$.
-Note that the ratio $\frac{N}{N_0} = 2^{(\Delta_R)n}$ goes to 0 as $n$ goes to the infinity.
-Hence, the limit of the probability~\eqref{eq:e221av0} is
-  \begin{equation}
-  	\lim_{n \rightarrow +\infty} \bbP[V = 0] \geq \lim_{n \rightarrow +\infty} \of{\of{1 - \frac{1}{N_0}}^{N_0} }^{\frac{N}{N_0}} = 1.
-  \end{equation}
-
-  Therefore, when $\Delta_R < 0$, the probability of achieving distortion $\delta$ goes to 0 as $n$ goes to the infinity.
-
- 	\vspace{0.2cm}
- 	Next,  we prove the probability of achieving the distortion under $\Delta_R > 0$.
-
- 	\vspace{0.2cm}
- 	Recall the tail bound~\eqref{eq:e2103} in Exercise~\ref{exe:210}. We have the upper bound of $\bbP \off{V^j = 1}$, where
- 	\begin{equation}\label{eq:e221prob}
- 		\bbP \off{V^j = 1} = \bbP \off{\sum_{i=1}^n \mathds{1} \off{X_i - z^j_i} \leq n\delta} \geq \frac{1}{n+1}e^{-nD(\delta || 1/2)}, \quad \text{for all }j\in [N].
- 	\end{equation}
- 	Then the probability of not achieving the distortion $\delta$ is
- 	\begin{equation}
- 		\bbP[V = 0] \leq \of{1 - \frac{1}{n+1} e^{-nD(\delta || 1/2)}}^N.
- 	\end{equation}
-
- 	Since the difference $\Delta_R > 0$, then the ratio $\frac{N}{N_0(n+1)} = \frac{2^{(\Delta_R) n} }{n+1}$ goes to infinity as $n$ goes to infinity.
-
- 	 Hence, the limit of the probability of not achieving the distortion $\delta$ is
- 	 \begin{equation}
- 	 	\lim_{n \rightarrow +\infty} \bbP[V = 0] \leq \lim_{n \rightarrow +\infty} \of{1 - \frac{1}{n+1} e^{-nD(\delta || 1/2)}}^N = \lim_{n \rightarrow +\infty}\of{ \of{1 - \frac{1}{(n+1)N_0}}^{(n+1)N_0}}^{\frac{N}{(n+1)N_0}} = 0.
- 	 \end{equation}
-
- 	 Therefore, when $\Delta_R > 0$, the probability of achieving distortion $\delta$ goes to 1 as $n$ goes to the infinity.
-
- 	 \vspace{0.2cm}
-
- 	 An alternative version of the proof when $\Delta_R > 0$ is below.
-
- 	 \vspace{0.2cm}
- 	 Let $p = \bbP \off{V^j = 1}$. By the Bernoulli's inequality, the probability of not achieving the distortion $\delta$ satisfies
- 	 \begin{equation}
- 	 	\bbP \off{V = 0} = \of{1 - p}^N \leq \frac{1}{1 - Np} = \frac{Np(1-p)}{Np(1-p) + N^2 p^2} = \frac{(\bbE[V])^2}{\bbE \off{V^2}}.
- 	 \end{equation}
- 	  Therefore, by inequality~\eqref{eq:e221prob} the limit of the probability of achieving the distortion $\delta$ is
- 	  \begin{equation}
- 	  	\lim_{n \rightarrow +\infty}  \bbP \off{V \geq 1} \geq \lim_{n \rightarrow +\infty}   1- \frac{1}{N p} \geq \lim_{n \rightarrow +\infty}  1 - \frac{n+1}{N N_0 } = 1.
- 	  \end{equation}
- \end{proof}
-
- \subsection{Exercise 2.22}
-
 \end{document}

### 2021-04-02 tensor_regress/ICSA_student/latexresource/abstract.tex
submission

--- 2020-09-14 tensor_regress/JASA_submission/paper_source/regress.tex
@@
 \addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{-.3in}%
+\addtolength{\textheight}{1in}%
 \addtolength{\topmargin}{-.8in}%
@@
   % blind version
-  \author{Jiaxin Hu, Zhuoyan Xu, Chanwoo Lee, and Miaoyan Wang\\
-    Department of Statistics, University of Wisconsin-Madison}
+ \author{
+    Jiaxin Hu\\
+   \texttt{jhu267@wisc.edu}\\
+   \texttt{Tel:6089600989}
+  \and
+    Chanwoo Lee\\
+   \texttt{chanwoo.lee@wisc.edu}\\
+   \texttt{Tel:6085569906}
+    \and
+    Miaoyan Wang\\
+    \texttt{miaoyan.wang@wisc.edu}\\
+    \texttt{Tel:6082653990}\\
+    \\
+    Department of Statistics, University of Wisconsin-Madison\\
+    \texttt{1300 University Ave,Madison,WI 53706,USA}
+
+}
+
+  %\author{Jiaxin Hu, Chanwoo Lee, and Miaoyan Wang\\
+   % Department of Statistics, University of Wisconsin-Madison}
+   \date{}
   \maketitle
@@
 
-\bigskip
 \begin{abstract}
@@
 
-\newpage
-\spacingset{1.5} % DON'T change the spacing!
-\section{Introduction}
-\label{sec:intro}
-
-Multi-dimensional arrays, known as tensors, are often collected with side information on multiple modes in modern scientific and engineering studies.
-A popular example is in neuroimaging~\citep{zhou2013tensor,wang2019common}.
-The brain connectivity networks are collected from a sample of individuals, accompanied by individual characteristics such as age, gender, and diseases status (see Figure~\ref{fig:intro1}a).
-Another example is in network analysis~\citep{pmlr-v108-berthet20a,hoff2005bilinear}.
-A typical social network consists of nodes that represent people and edges that represent  friendships.
-Side information such as people’s demographic information and friendship types are often available.
-In both examples, it is of keen scientific interest to identify the variation in the data tensor (e.g., brain connectivities, social community patterns) that is affected by available features.
-These seemingly different scenarios pose a common yet challenging problem for tensor data modeling.
-
-In addition to the aforementioned challenges, many tensor datasets consist of non-Gaussian measurements.
-Examples include the political interaction dataset \citep{hu2015scalable} which measures action counts between countries under various relation types, and the brain connectivity network dataset \citep{zhang2018mapping} which is a collection of binary adjacency matrices.
-Classical tensor decomposition methods are based on minimizing the Frobenious norm of the reconstruction error, leading to suboptimal predictions for binary- or count-valued response variables.
-A number of supervised tensor methods have been proposed \citep{narita2012tensor, zhao2012higher, yu2016learning} to address the tensor regression problem in various forms, such as scalar-to-tensor regression and tensor-response regression.
-These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest.
-The gap between theory and practice means a great opportunity to improve modeling paradigms to better capture the complexity in tensor data.
-
-\begin{figure*}[t]
-\begin{center}
-\includegraphics[width=16cm]{demo.pdf}
-\end{center}
-\caption{Examples of supervised tensor decomposition with interactive side information.
-(a) Network population model.
-(b) Schematic diagram of supervised tensor decomposition for an order-3 tensor $\tY$ with feature matrices $\mX_k$ on each of the three modes.} \label{fig:intro1}
-\end{figure*}
-
-We present a general model and associated method for decomposing a data tensor whose entries are from exponential family with interactive side information.
-We formulate the learning task as a structured regression problem, with tensor observation serving as the response, and the multiple side information as interactive features.
-Figure~\ref{fig:intro1}b illustrates our model in the special case of order-3 tensors.
-A low-rank structure is imposed to the conditional mean of tensor observation, where unlike classical decomposition, the tensor factors $\mX_k\mM_k\in\mathbb{R}^{d_k\times r_k}$ belong to the space spanned by features $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,2,3$.
-The unknown matrices $\mM_k\in\mathbb{R}^{p_k\times r_k}$ (referred to as ``dimension reduction matrices'') link the conditional mean to the feature spaces, thereby allowing the identification of variations in the tensor data attributable to the side information.
-
-Our proposal blends the modeling power of generalized linear model (GLM) and the exploratory capability of tensor dimension reduction in order to take the best out of both worlds.
-We leverage GLM to allow heteroscedacity due to the mean-variance relationship in the non-Gaussian data.
-This flexibility is important in practice.
-Furthermore, our low-rank model on the (transformed) conditional mean tensor effectively mitigates the curse of high dimensionality.
-In classical GLM, the sample size and feature dimension are well defined; however, in the tensor data analysis, we observe only one realization of an order-$K$ tensor and up to $K$ interactive feature matrices.
-Both the number of tensor entries and feature dimension grow exponentially in $K$.
-Dimension reduction is therefore crucial for prediction and interpretability.
-We establish the statistical convergence of our estimator, and we quantify the gain in prediction through simulations and case studies.
-
-Our work is closely related to but also clearly distinctive from several lines of previous work.
-The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor, hong2020generalized} that aims to find the best low-rank representation of a data tensor.
-In contrast, our model is a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and multiple features.
-The low-rank factorization is determined jointly by the tensor data and feature matrices in our model.
-
-The second line of work studies the tensor-on-tensor regression~\citep{raskutti2019convex, lock2018tensor, gahrooei2020multiple}.
-Our model shares a common ground with earlier approaches, but we provide more efficient solutions to new settings that have more practical significance.
-As we show in Section~\ref{sec:connection}, the supervised tensor decomposition has an interesting connection with tensor-on-tensor regression.
-Previous methods~\citep{gahrooei2020multiple,lock2018tensor} mainly focus on Gaussian tensors.
-The Frobenius norm used in the objective function is statistical suboptimal for general exponential family tensors.
-Maximum likelihood estimator (MLE) is studied in~\citep{raskutti2019convex} and a convex relaxation algorithm is proposed to solve for low-rank tensor coefficients.
-However, in the tensor case, convex MLE suffers from both computational intractability and statistical suboptimality.
-We advocate a non-convex approach and provide strong evidence for its success in our setting.
-Most previous tensor regression focuses on prediction~\citep{lock2018tensor,raskutti2019convex,gahrooei2020multiple}, and we go step further by finding the sufficient dimension reduction~\citep{adragni2009sufficient}, $\textup{Span}(\mM_k)$, that facilitates the identification of interaction effects in features (see Figure~\ref{fig:intro1}b).
-The latter approach greatly improves the \emph{interpretability} in prediction.
-In this regards, our method opens up new opportunities for tensor data analysis in a wider range of applications.
-
-The third line of work uses side information for various tensor learning tasks, such as for completion~\citep{narita2012tensor,cao2016semi} and for recommendation system~\citep{ioannidis2019coupled,farias2019learning}.
-These methods also study tensors with side information, same as we do, but they take regularization approaches to penalize predictions that are distant from side information.
-One important difference is that their goal is prediction but not parameter estimation.
-The effects of features and their interactions are not estimated in these data-driven approaches.
-In contrast, our goal is interpretable prediction, and we estimate factor matrices $\mM_k$ using a model-based approach.
-Estimating $\mM_k$ allows us to identify sufficient features and the interactions thereof.
-We numerically compare these two approaches in Section~\ref{sec:simulation}.
-
-The remainder of this paper is organized as follows.
-Section~\ref{sec:pre} introduces tensor notation and preliminaries.
-Section~\ref{sec:model} presents the main model and three motivating examples for supervised tensor decomposition.
-Section~\ref{sec:est} describes the rank-constrained quasi-likelihood estimation and an alternating optimization algorithm.
-In Section~\ref{sec:simulation}, we present numerical experiments and assess the performance in comparison to alternative methods.
-In Section~\ref{sec:data}, we apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
-We conclude in Section~\ref{sec:con} with discussions about our findings and avenues of future work.
-
-\section{Preliminaries}\label{sec:pre}
-We introduce the basic tensor properties used in the paper; more details on tensor notation can be found in~\cite{kolda2009tensor}.
-We use $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ to denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor, where $K$ corresponds to the number of modes of $\tY$ and is called the order.
-The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as
-\begin{equation}
-\tY \times_1\mX_1\ldots \times_K \mX_K=\entry{\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}x_{j_1,i_1}^{(1)}\ldots x_{j_K,i_K}^{(K)}},
-\end{equation}
-which results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor.
-For ease of presentation, we use the shorthand $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product.
-For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as
-\[
-\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.
-\] 
-The Frobenius norm and maximum norm of $\tY$ are defined as 
-\[
-\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2}\quad \text{and}\quad \mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.
-\]
-When $a$ is a vector, we use $\vnormSize{}{a}=\langle a,a \rangle^{1/2}$ to denote the vector 2-norm.
-
-A higher-order tensor can be reshaped into a lower-order object.
-We use $\textup{vec}(\cdot)$ to denote the operation that reshapes the tensor into a vector, and $\text{Unfold}_k(\cdot)$ to denote the operation that reshapes the tensor along mode $k$ into a matrix of size $d_k$-by-$\prod_{i\neq k}d_i$.
-The multilinear rank of an order-$K$ tensor $\tY$ is defined as a length-$K$ vector $\mr=(r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\textup{Unfold}_k(\tY)$, $k=1,\ldots,K$.
-
-We use lower-case letters (e.g.,\ $a,b,c$) for scalars and vectors, upper-case boldface letters (e.g.,\ $\mA, \mB, \mC$) for matrices, and calligraphy letters (e.g.,\ $\tA, \tB, \tC$) for tensors of order three or greater.
-We let $\mI_d$ denote the $d \times d$ identity matrix and $[d]$ denote the $d$-set $\{1,\ldots,d\}$.
-For ease of notation, we allow the basic arithmetic operators (e.g.,\ $+, -, \geq $) and univariate functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner.
-We use $f'(\cdot)$ to denote the first-order derivative and $f''(\cdot)$ the second-order derivative of function $f(\cdot)$.
-
-\section{Motivation and model}\label{sec:model}
-\subsection{General framework for supervised tensor decomposition}
-We begin with a general framework for supervised tensor decomposition and then discuss its implication in three concrete examples.
-Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor.
-Suppose the side information is available on each of the $K$ modes.
-Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the feature matrix on the mode $k\in[K]$, where $x_{ij}$ denotes the $j$-th feature value for the $i$-th tensor entity, for $(i,j)\in[d_k]\times[p_k]$, $p_k\leq d_k$.
-We propose a multilinear conditional mean model between the data tensor and feature matrices.
-Assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent realizations from an exponential family distribution, and the conditional mean tensor admits the form
-\begin{align}\label{eq:tensormodel}
-&\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)=f(\Theta),\quad \text{with}\quad \Theta =\tB\times\{\mX_1,\ \ldots,\ \mX_K\} ,
-\end{align}
-where $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$ is the multilinear predictor, $\tB\in\mathbb{R}^{p_1\times \cdots \times p_K}$ is the unknown parameter tensor, $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$, and $\times$ denotes the tensor-by-matrix product.
-The choice of link function is based on the assumed distribution family of tensor entries.
-Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and $\exp(\cdot)$ link for Poisson distribution.
-In general, dispersion parameters can also be included in the model.
-Because our main focus is the tensor decomposition under the mean model, we omit the dispersion parameter in this section for ease of presentation.
-
-In classical tensor decomposition, tensor factorization is performed on either data tensor $\tY$ or mean tensor $\mathbb{E}(\tY)$.
-In the context of supervised tensor decomposition, we propose to factorize the latent parameter tensor $\tB$,
-\begin{equation}\label{eq:factor}
-\tB=\tC\times\{\mM_1,\ \ldots,\ \mM_K\},
-\end{equation}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, and $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices consisting of orthonormal columns, where $r_k\leq p_k$ for all $k\in[K]$.
-By the definition of multilinear rank, model equations~\eqref{eq:tensormodel} and~\eqref{eq:factor} imply the low-rankness $\mr=(r_1,\ldots,r_K)$ of the conditional mean tensor under the link function.
-We now reach our final model for supervised tensor decomposition,
-\begin{align}\label{eq:decomp}
-\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1,\ \ldots,\ \mX_K\mM_K\}\right),\notag\\
-\text{with} \ &\ \mM_k^T\mM_k = \mI_{r_k},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\quad \text{for all }k=1,\ldots,K,
-\end{align}
-where the parameters of interest are $\mM_k$ and $\tC$.
-Note that model~\eqref{eq:decomp} assumes a fixed, known rank $\mr=(r_1,\ldots,r_K)$; the adaptation to unknown rank will be addressed in Section~\ref{sec:tuning}.
-Figure~\ref{fig:intro1}b provides a schematic illustration of our model.
-The features $\mX_k$ affect the distribution of tensor entries in $\tY$ through the form $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$.
-We call $\mX_k\mM_k$ the ``supervised tensor factors'' or ``sufficient features''~\citep{adragni2009sufficient}, and call $\mM_k$ the ``dimension reduction matrix.'' The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes.
-Our goal is to find $\mM_k$ and the corresponding $\tC$.
-Note that $\mM_k$ and $\tC$ are identifiable only up to orthonormal transformations.
-
-\subsection{Three examples}
-We give three seemingly different examples that can all be formulated as our supervised tensor decomposition model~\eqref{eq:decomp}.
-\begin{example}[Spatio-temporal growth model]
-The growth curve model~\citep{gabriel1998generalised,srivastava2008models} was originally proposed as an example of bilinear model for matrix data, and we adopt its higher-order extension here.
-Let $\tY=\entry{y_{ijk}}\in\mathbb{R}^{d \times m\times n}$ denote the pH measurements of $d$ lakes at $m$ levels of depth and for $n$ time points.
-Suppose the sampled lakes belong to $q$ types, with $p$ lakes in each type.
-Let $\{\ell_j\}_{j\in[m]}$ denote the sampled depth levels and $\{t_k\}_{k\in[n]}$ the time points.
-Assume that the expected pH trend in depth is a polynomial of order at most $r$ and that the expected trend in time is a polynomial of order $s$.
-Then, the conditional mean model for the spatio-temporal growth can be represented as
-\begin{equation}\label{eq:time}
-\mathbb{E}(\tY|\mX_1,\mX_2,\mX_3)=\tC\times\{\mX_1\mM_1,\ \mX_2\mM_2,\ \mX_3\mM_3\},
-\end{equation}
-where $\mX_1=\text{blockdiag}\{\mathbf{1}_p,\ldots,\mathbf{1}_p\}\in \{0,1\}^{d\times q}$ is the design matrix for lake types, and
-\[
-\mX_2=
-\begin{pmatrix}
-1 & \ell_1&\cdots &\ell^{r}_1\\
-1 & \ell_2&\cdots &\ell^{r}_2\\
-\vdots &\vdots&\ddots&\vdots\\
-1&\ell_{m}&\cdots&\ell^{r}_{m}
-\end{pmatrix},\quad
-\mX_3=
-\begin{pmatrix}
-1 & t_1&\cdots &t^{s}_1\\
-1 & t_2&\cdots &t^{s}_2\\
-\vdots &\vdots&\ddots&\vdots\\
-1&t_{n}&\cdots&t^{s}_{n}
-\end{pmatrix}
-\]
-are the design matrices for spatial and temporal effects, respectively, $\tC\in\mathbb{R}^{r_1\times r_2\times r_3}$ is the unknown core tensor, and $\mM_k$ are unknown dimension reduction matrices on each mode.
-The factors $\mX_k\mM_k$ are sufficient features in the mean model~\eqref{eq:time}.
-The spatial-temporal model is a special case of our supervised tensor decomposition model~\eqref{eq:decomp}, with features available on each of the three modes.
-\end{example}
-
-\begin{example}[Network population model]\label{example:brain}
-Network response model~\citep{rabusseau2016low, zhang2018network} is recently developed for neuroimanig analysis.
-The goal is to study the relationship between brain network connectivity pattern and features of individuals.
-Suppose we have a sample of $n$ observations, $\{(\mY_i, \mx_i)\colon i=1,\ldots,n\}$, where for each individual $i\in[n]$, $\mY_i\in\{0,1\}^{d\times d}$ is the undirected adjacency matrix whose entries indicate presences/absences of connectivities between $d$ brain nodes, and $\mx_i\in\mathbb{R}^p$ is the individual's feature such as age, gender, cognition score, etc.
-The network-response model  has the conditional mean
-\begin{equation}\label{eq:network}
-\textup{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n,
-\end{equation}
-where $\tB\in \mathbb{R}^{d\times d\times p}$ is a rank-$(r_1,r_1,r_2)$ coefficient tensor, and $\tB$ is assumed to be symmetric in the first two modes.
-
-The model~\eqref{eq:network} is a special case of our supervised tensor decomposition, with feature matrix on the last mode of the tensor.
-Specifically, we stack the network observations $\{\mY_i\}$ together and obtain an order-3 response tensor $\tY\in\{0,1\}^{d\times d\times n}$.
-Define a feature matrix $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$.
-Then, the model~\eqref{eq:network} has the equivalent representation of supervised tensor decomposition,
-\[
-\textup{logit}(\mathbb{E}(\tY|\mX))=\tC\times\{\mM,\ \mM,\ \mX\mM'\},
-\]
-where $\tC\in\mathbb{R}^{r_1\times r_1\times r_2}$ is the core tensor, $\mM\in\mathbb{R}^{d\times r}$ is the dimension reduction matrix on the first two modes, and $\mM'\in\mathbb{R}^{p\times r_2}$ is for the last mode.
-\end{example}
-
- \begin{example}[Dyadic data with node attributes] Dyadic dataset consists of measurements on pairs of objects.
-Common examples include graphs and networks.
-Let $\tG=(V,E)$ denote a graph, where $V=[d]$ is the node set of the graph, and $E\subset V\times V$ is the edge set.
-Suppose that we also observe feature vector $\mx_i\in\mathbb{R}^p$ associated to each node $i\in V$.
-A probabilistic model on the graph $\tG=(V,E)$ can be described by the following matrix regression.
-The edge connects the two vertices $i$ and $j$ independently of other pairs, and the probability of connection is modeled as
-\begin{equation}\label{eq:edge}
- \textup{logit}\left(\mathbb{P}\left((i,j)\in E\right)\right)=\mx^T_i\mB\mx_j=\langle \mB,\ \mx^T_i\mx_j\rangle,
- \end{equation}
- where $\mB\in\mathbb{R}^{p\times p}$ is a symmetric rank-$r$ matrix.
-The low-rankness in $\mB$ has demonstrated its success in modeling transitivity, balance, and communities in  networks~\citep{hoff2005bilinear}.
-We show that our supervised tensor decompostiion~\eqref{eq:decomp} also incorporates the graph model as a special case.
-Let $\tY=\entry{y_{ij}}$ be a binary matrix where $y_{ij}=\mathds{1}_{(i,j)\in E}$.
-Define $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$.
-Then, the graph model~\eqref{eq:edge} can be expressed as
- \[
- \textup{logit}(\mathbb{E}(\mY|\mX))=\mC\times\{\mX\mM,\ \mX\mM\}, 
-  \]
-  where $\mC\in\mathbb{R}^{r\times r}$ and $\mM\in\mathbb{R}^{p\times r}$ are based on the singular value decomposition of $\mB=\mM\mC\mM^T$. 
-  \end{example}
-
-In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by features.
-The coefficient tensor $\tB$ in model~\eqref{eq:tensormodel} serves this goal by collecting the feature effects.
-To encourage the sharing among effects, decomposition~\eqref{eq:factor} assumes that the feature effects $\tB$ lie in a low-dimensional parameter space.
-The low-rank assumption is plausible in many scientific applications.
-In the Example~\ref{example:brain} of brain imaging analysis, for instance, it is often believed that the brain nodes can be partitioned into several regions, and the person's attributes affect the connectivity in a similar way within a region.
-
-Our supervised tensor decomposition~\eqref{eq:decomp} is able to incorporate arbitrary numbers of feature matrices.
-When certain mode $k$ has no available side information, we set $\mX_k=\mI_{d_k}$ in the model~\eqref{eq:decomp}.
-In particular, our model~\eqref{eq:decomp} reduces to classical unsupervised tensor decomposition~\citep{de2000multilinear,hong2020generalized} when no side information is available; i.e., $\mX_k=\mI_{d_k}$ for all $k\in[K]$.
-
-\subsection{Connection to sufficient dimension reduction and tensor-on-tensor regression}\label{sec:connection}
-One important implication of our method is that we allow high-dimensionality in both tensor dimension $d_k$ and feature dimension $p_k$.
-Recall that the model rank $r_k$ is typically smaller than $d_k$ and $p_k$.
-In such a case, the matrices $\mM_k$ serve the role of simultaneous dimension reduction of the data tensor and features.
-The role of the sufficient features $\mX_k\mM_k$ can be seen in the following two conditional independence assumptions in model~\eqref{eq:decomp},
-\begin{align}
-\tY \ci \{\mX_k\}&\ \big|\ \{\mX_k\mM_k\} \ \ \text{(independence between the tensor and multiple features)},\\
-y_{i_1,\ldots,i_K} \ci y_{i'_1,\ldots,i'_K} & \ \big|\ \{\mX_k\mM_k\}  \ \ \text{(independence within the tensor)},
-\end{align}
-where the second line holds for all $(i_1,\ldots,i_K)\neq(i'_1,\ldots,i'_K) \in[d_1]\times\cdots\times[d_K]$, and $\ci$ denotes the independence.
-The first property highlights the ``decorrelation'' role of $\mM_k$, in the same spirit as the sufficient dimension reduction~\citep{adragni2009sufficient} in high-dimensional supervised learning, whereas the second property highlights the tensor dimension reduction in the usual unsupervised sense (consider, for example, $\mX_k=\mathbf{I}_{d_k}$ for all $k\in[K]$).
-
-Our model also has a close connection with tensor-on-tensor regression~\citep{raskutti2019convex, lock2018tensor,han2020optimal}.
-Specifically, model~\eqref{eq:decomp} can be viewed as a multivariate regression model, where the response is the vectorized tensor and the covariates are interactions between sufficient features across modes.
-We take an order-3 tensor under and Gaussian model for illustration.
-Let $\mX,\mZ,\mW$ denote the feature matrix on mode $k=1, 2, 3$, respectively.
-Suppose that each mode has two-dimensional sufficient features, denoted $\mM_1\mX=[\mx_1,\mx_2]$, $\mM_2\mZ=[\mz_1,\mz_2]$, $\mM_3\mW=[\mw_1,\mw_2]$.
-Here $\mx_1,\mx_2,\ldots,\mw_1, \mw_2$ are column vectors.
-Then the model~\eqref{eq:decomp} is a regression model with across-mode interactions,
-\begin{equation}\label{eq:interaction}
-\mathbb{E}(y_{ijk}|\mX,\mZ,\mW)=c_{111}\mx_{1i}\mz_{1j}\mw_{1k}+c_{121}\mx_{i1}\mz_{j2}\mw_{k1}+\cdots+c_{221}\mx_{2i}\mz_{2j}\mw_{1k}+c_{222}\mx_{2i}\mz_{2j}\mw_{2k},
-\end{equation}
-where $\entry{c_{ijk}} \in \mathbb{R}^{2\times 2\times 3}$ are unknown interaction effects, $\mx_{1i}$ denotes the $i$-th entry in the feature vector $\mx_1$, and similar notations apply to other features.
-Note that lower-order interactions are naturally incorporated in~\eqref{eq:interaction} if we include an intercept column in the sufficient feature matrices.
-Model~\eqref{eq:interaction} shows the benefit of our supervised tensor decomposition for identifying across-mode interactions.
-
-\section{Estimation}\label{sec:est}
-
-\subsection{Rank-constrained M-estimator}
-We develop a likelihood-based procedure to estimate $\tC$ and $\mM_k$ in~\eqref{eq:decomp}.
-We adopt the exponential family as a flexible framework for different data types.
-In a classical generalized linear model with a scalar response $y$ and feature $\mx$, the density is expressed as
-\[
-p(y|\mx, \boldsymbol{\beta})=c(y,\phi)\exp\left(\frac{y\theta- b(\theta)}{\phi}\right)\ \text{with}\ \theta=\boldsymbol{\beta}^T\mx,
-\]
-where $b(\cdot)$ is a known function, $\theta$ is the linear predictor, $\phi>0$ is the dispersion parameter, and $c(\cdot)$ is a known normalizing function.
-The choice of link functions depends on the data types and on the observation domain of $y$, denoted $\mathbb{Y}$.
-For example, the observation domain is $\mathbb{Y}=\mathbb{R}$ for continuous data, $\mathbb{Y}=\mathbb{N}$ for count data, and  $\mathbb{Y}=\{0,1\}$ for binary data.
-The canonical link function $f$ is chosen to be $f(\cdot)=b'(\cdot)$.
-Table~1 summarizes the canonical link functions for common types of distributions.
-
-\begin{table}[htb]
-\centering
-\begin{tabular}{c|ccc}
-Data type &Gaussian & Poisson& Bernoulli\\
-\hline
-Domain $\mathbb{Y}$& $\mathbb{R}$&$\mathbb{N}$&$\{0,1\}$\\
- $b(\theta)$&$\theta^2/2$& $\exp(\theta)$&$\log (1+\exp(\theta))$\\
- link $f(\theta)$&$\theta$&$\exp(\theta)$&$(1+\exp(-\theta))^{-1}$
-\end{tabular}
-\caption{Canonical links for common distributions.}\label{table:link}
-\end{table}
-
-In our context, we model the tensor entries $y_{ijk}$, conditional on $\theta_{ijk}$, as independent draws from an exponential family.
-Ignoring constants that do not depend on $\Theta$, the quasi log-likelihood of~\eqref{eq:decomp} is equal to Bregman distance between $\tY$ and $b'(\Theta)$:
-\begin{align}\label{eq:loglikelihood}
-\tL_{\tY}(\tC,\mM_1,\ldots,\mM_K)&=\langle \tY, \Theta \rangle - \sum_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K}),\notag\\
-\text{where}\ \Theta&=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
-\end{align}
-
-We propose a constrained maximum quasi-likelihood estimator (M-estimator),
-\begin{align} \label{eq:MLE} 
-(\hat \tC, \hat \mM_1,\ldots,\hat \mM_k) &=\argmax_{(\tC,\mM_1,\ldots,\mM_K)\in \tP} \ \tL_{\tY}(\tC,\mM_1,\ldots,\mM_K),
-\end{align}
-where the parameter space $\tP$ is
-\[
-\tP=\left\{(\tC, \mM_1,\ldots,\mM_K) \ \Big| \ \mM_k^T\mM_k=\mathbf{I}_{r_k},\ \mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha \right\}.
-\]
-The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to avoid the divergence in the non-Gaussian variance.
-
-\subsection{Alternating optimization} \label{subsec:alg}
-We propose an alternating optimization algorithm to solve~\eqref{eq:MLE}. 
-The decision variables in the objective function~\eqref{eq:MLE} consist of $K+1$ blocks of variables, one for the core tensor $\tC$ and $K$ for the factor matrices $\mM_k$.
-We notice that, if any $K$ out of the $K+1$ blocks of variables are known, then the optimization reduces to a simple GLM with respect to the last block of variables.
-This observation leads to an iterative updating scheme for one block at a time while keeping others fixed.
-After each iteration, we rescale the core tensor $\tC^{(t+1)}$ subject to the maximum norm constraint.
-The full algorithm is described in Algorithm~\ref{alg:B}.
-
-\begin{algorithm}[t]
-\caption{Supervised Tensor Decomposition with Interactive Side Information}\label{alg:B}
-\begin{algorithmic}[1]
-\INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, feature matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,\ldots,K$, target Tucker rank $\mr=(r_1,\ldots,r_K)$, link function $f$, maximum norm bound $\alpha$
-\OUTPUT Estimated core tensor $\hat \tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ and factor matrices $\hat \mM_k\in\mathbb{R}^{p_k\times r_k}$. 
-\State Random initialization of the core tensor $\tC$ and factor matrices $\mM_k$. 
-\While{Do until convergence}
-\For { $k=1$ to $K$}
-\State Obtain the factor matrix $\tilde \mM_k\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$.
-\State Perform QR factorization $\tilde \mM_k=\mQ\mR$, where $\mQ\in\mathbb{R}^{p_k\times r_k}$ consists of orthonormal columns. 
-\State Update $\mM_k\leftarrow \mQ$ and core tensor $\tC\leftarrow \tC\times_k \mR$.
-\EndFor
-\State Update the core tensor $\tC$ by solving a GLM with $\textup{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM_k]$ as features, and $f$ as link function.
-Here $\otimes$ denotes the Kronecker product of matrices.
-\State Rescale the core tensor $\tC$ subject to the maximum norm constraint. 
-\EndWhile
-\end{algorithmic}
-\end{algorithm}
-
-The optimization~\eqref{eq:MLE} is a non-convex problem due to the non-convexity in the feasible set $\tP$.
-Under mild conditions, our algorithm enjoys global convergence; i.e.\ any sequence of iterates generated by the alternating algorithm converges to a stationary point of $\tL_\tY(\cdot)$ module orthogonal transformation (see Supplementary Notes, Section~\ref{sec:SAlgorithm}).
-In theory, global optimality of non-convex optimization is often challenging to obtain.
-Fortunately, we will show in Section~\ref{subsec:statprob} that the desired statistical performance holds for all local optimizers satisfying $\tL_{\tY}(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)\geq \tL_{\tY} (\tC_{\textup{true}},\mM_{1,\textup{true}},\ldots,\mM_{K,\textup{true}})$, where the subscript ``true'' denotes the true parameter in model~\eqref{eq:decomp}.
-This result indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective is large enough.
-In the experiments we considered, we find that Algorithm~\ref{alg:B} typically gives satisfactory convergent points upon random initialization.
-Figure~\ref{fig:loglike} shows the trajectories of objective function for order-3 tensors based on model~\eqref{eq:decomp}, where $d_k = d \in\{25,30\}$, $p_k = 0.4d$, $r_k = r\in\{3,6\}$ for $ k = 1,2,3$.
-We consider input tensors with Gaussian, Bernoulli, and Poisson entries.
-Under all combinations of the dimension $d$, rank $r$, and type of the entries, Algorithm~\ref{alg:B} converges quickly in a few iterations, and the objective values at convergent points are close to or larger than the value at true parameters.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=15cm]{loglike.pdf}
-\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c) Poisson models.
-The dashed line represents the objective value at true parameters.
-}\label{fig:loglike}
-\end{figure}
-
-\subsection{Rank selection and time complexity}\label{sec:tuning}
-Algorithm~\ref{alg:B} assumes the rank $\mr$ is given.
-In practice, the rank is often unknown and must be determined from the data.
-We propose to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
-\begin{align}\label{eq:BIC}
-\hat \mr&=\argmin_{\mr=(r_1,\ldots,r_K)} \textup{BIC}(\mr)\notag\\
-&=\argmin_{\mr=(r_1,\ldots,r_K)}\left[-2\tL_{\tY}(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)+p_e(\mr)\log \left(\prod_k d_k\right) \right],
-\end{align}
-where $p_e(\mr)\stackrel{\textup{def}}{=}\sum_k (p_k-r_k-1)r_k+\prod_k r_k$ is the effective number of parameters in the model.
-We choose $\hat \mr$ that minimizes $\textup{BIC}(\mr)$ via grid search.
-Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model.
-We evaluate the empirical performance of BIC in Section~\ref{sec:simulation}.
-
-The computational complexity of our Algorithm~\eqref{alg:B} is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the data tensor.
-More precisely, the update of core tensor costs $O(r^3d)$, where $r=\prod_k r_k$ is the total size of the core tensor.
-The update of each factor matrix $\mM_k$ involves a GLM with a $d$-length response, and $d$-by-$(r_kp_k)$ feature matrix.
-Solving such a GLM requires $O(dr^3_kp^3_k)$, and therefore the cost for updating $K$ factors in total is $O( d\sum_k r^3_k p_k^3)$.
-This complexity in tensor dimension matches with the classical tensor decomposition~\citep{kolda2009tensor}.
-
-\subsection{Statistical properties}\label{subsec:statprob}
-In this section, we provide the accuracy guarantee for the proposed M-estimator~\eqref{eq:MLE}.
-Note that the factor matrices $\mM_k$ are identifiable only up to orthogonal transformation.
-Therefore, we use angle distance to assess the accuracy in estimating the column space, $\textup{Span}(\mM_k)$.
-For any two matrices $\mA,\mB$ with orthonormal columns of same dimension, the angle distance is defined as
-\[
-\sin \Theta(\mA,\mB)=\snormSize{}{\mA^T \mB^{\perp}}=\max\left\{ \frac{\langle \mx, \my\rangle}{\vnormSize{}{\mx}\vnormSize{}{\my}}\colon \ \mx\in \textup{Span}(\mA),\ \my\in \textup{Span}(\mB)\right\}.
-\]
-
-In modern applications, the tensor data and features are often large-scale.
-We are particularly interested in the high-dimensional regime in which both $d_k$ and $p_k$ diverge; i.e.\ $d_k\to \infty$ and $p_k\to\infty$, while $p_k/d_k \to \gamma_k \in[0,1)$.
-As the size of problem grows, and so does the number of unknown parameters.
-The classical MLE theory does not directly apply.
-We leverage the recent development in random tensor theory and high-dimensional statistics to establish the error bounds of the estimation.
-
-\begin{assumption}\label{ass}We make the following assumptions:
-\begin{enumerate}
-\item [A1.] There exist two positive constants $c_1, c_2>0$ such that $c_1\leq \sigma_{\min}(\mX_k)\leq  \sigma_{\max}(\mX_k)\leq c_2$ for all $k\in[K]$.
-Here $\sigma_{min}(\cdot)$ and $\sigma_{max}(\cdot)$ denote the smallest and largest matrix singular values.
-\item [A1'.] The feature matrices $\mX_k$ are Gaussian designs with i.i.d.\ $N(0,1)$ entries.
-\item [A2.] There exist two positive constants $L, U>0$, such that $L\phi \leq \textup{Var}(y_{i_1,\ldots,i_K}|\theta_{i_1,\ldots,i_K})\leq U\phi $ for all $|\theta_{i_1,\ldots,i_K}|\leq \alpha$.
-Equivalently, $L\leq b''(\theta_{i_1,\ldots,i_K}) \leq U$ for all $|\theta_{i_1,\ldots,i_K}|\leq \alpha$.
-Here $\alpha$ is the upper bound of the linear predictor in~\eqref{eq:MLE}.
-
-\end{enumerate}
-\end{assumption}
-The assumptions are fairly mild.
-Assumptions A1 and A1' consider two different scenarios about feature matrices.
-We study the cases A1 and A1' separately but have the same theoretical results.
-Assumption A1 is applicable when feature matrix is asymptotically non-singular and has bounded spectral norm, whereas Assumption A1' imposes the commonly-used Gaussian design in high-dimensional statistics~\citep{chen2019non}.
-Assumption A2 ensures the log-likelihood $\tL_\tY(\Theta)$ is strictly concave in the linear predictor $\Theta$.
-The two equivalent statements in Assumption A2 are based on the fact that $\textup{Var}(y_{i_1,\ldots,i_K}|\tX,\tB)=\phi b''(\theta_{i_1,\ldots,i_K})$ when $y_{i_1,\ldots,i_K}$ belongs to an exponential family~\citep{mccullagh1989generalized}.
-
-\begin{thm}[Statistical convergence]\label{thm:main}
-Consider a data tensor generated from model~\eqref{eq:decomp}, where the entries are conditionally independent realizations from an exponential family.
-Let $(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)$ be the M-estimator in~\eqref{eq:MLE} and $\hat \tB=\hat \tC\times \hat \mM_1\times\cdots \times \hat \mM_K$.
-Define $r_\textup{total}=\prod_k r_k$ and $r_{\max}=\max_k r_k$.
-Under Assumptions A1 and A2 with scaled feature matrices $\check \mX_k= \sqrt{d_k}\mX_k$, or under Assumptions A1' and A2 with original feature matrices, there exist two positive constants $C_1=C_1(\alpha,K), C_2=C_2(\alpha, K)>0$ independent of dimensions $\{d_k\}$ and $\{p_k\}$, such that, with probability at least $1-\exp(-C_1\sum_k p_k)$,
-\begin{equation}\label{eq:bound}
-    \FnormSize{}{\trueB- \hat \tB}^2\leq \frac{ C_{2} r_{\textup{total}}}{r_{\max}}\frac{\sum_k p_k}{\prod_k d_k}.
-\end{equation}
-Furthermore, if the unfolded core tensor has non-degenerate singular values at mode $k\in[K]$, i.e., $\sigma_{\min}(\textup{Unfold}_k(\tC_\textup{true})) \geq c>0$ for some constant $c$, then
-\begin{equation}\label{eq:sinebound}
-\textup{sin}^2 \Theta(\trueM,\ \hat \mM_k) \leq  \frac{ C_{2}r_{\textup{total}}}{ r_{\max}\sigma^2_{\min}(\textup{Unfold}_k(\tC_\textup{true}))}\frac{\sum_k p_k}{\prod_k d_k}.
-\end{equation}
-\end{thm}
-
-Theorem~\ref{thm:main} establishes the statistical convergence for the estimator~\eqref{eq:MLE}.
-The proof is provided in Appendix~\ref{sec:appedix}.
-In fact, the proof shows that the desired convergence rate holds not only for the M-estimator, but also for any local optimizers satisfying $\tL_\tY(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K) \geq \tL_\tY(\tC_{\textup{true}},  \mM_{1,\textup{true}},\ldots,\mM_{K,\textup{true}})$.
-Consider a special case when tensor dimensions are equal on each of the modes, i.e., $d_k=d$ for all $k\in[K]$, and feature dimension grows with tensor dimension, $p_k=\gamma d$, $\gamma\in [0,1)$, for $k\in[K]$.
-The result in~\eqref{eq:bound} implies that the estimation has a convergence rate $\tO(d^{-(K-1)})$.
-Therefore, our estimation is consistent in high dimensional regimes, and the convergence becomes especially favorable as the order of tensor data increases.
-
-As immediate applications, we obtain the convergence rate for the three examples mentioned in Section~\ref{sec:model}.
-
-\begin{customexample}{1}[Spatio-temporal growth model] The estimated type-by-time-by-space coefficient tensor converges at the rate $\tO\left(\frac{p+r+s}{dmn}\right)$ where $p\leq d$, $r\leq m$ and $s\leq n$.
-The estimation achieves consistency as long as the dimension grows along either of the three modes.
-\end{customexample}
-
-\begin{customexample}{2} [Network population model] The estimated node-by-node-by-feature tensor converges at the rate $\tO\left(\frac{2d+p}{d^2n}\right)$ where $p\leq n$.
-The estimation achieves consistency as the number of individuals or the number of nodes grows.
-\end{customexample}
-
-\begin{customexample}{3} [Dyadic data with node attributes] The estimated feature-by-feature matrix converges at the rate $\tO\left(\frac{p}{d^2}\right)$ where $p\leq d$.
-Again, our estimation achieves consistency as the number of nodes grows.
-\end{customexample}
-
-We conclude this section by providing the prediction accuracy, measured in KL divergence, for the response distribution.
-
-\begin{thm}[Prediction error]\label{thm:KL}
-Assume the same set-up as in Theorem~\ref{thm:main}.
-Let $\mathbb{P}_{\tY_{\textup{true}}}$ and $\mathbb{P}_{\hat \tY}$ denote the distributions of $\tY$ given the true parameter $\trueB$ and estimated parameter $\hat \tB$, respectively.
-Then, we have, with probability at least $1-\exp(C_1\sum_k p_k)$,
-\[
-\textup{KL}(\mathbb{P}_{\tY_{\textup{true}}},\ \mathbb{P}_{\hat \tY})\leq \frac{C_3r_{\textup{total}}}{r_{\max}}\sum_k p_k,
-\]
-where $C_3=C_3(\alpha, K)>0$ is a constant independent of dimensions $\{d_k\}$ and $\{p_k\}$.
-\end{thm}
-
-\section{Numerical experiments}\label{sec:simulation}
-We evaluate the empirical performance of our supervised tensor decomposition (STD) through simulations.
-We consider order-3 tensors with a range of distribution types.
-Unless otherwise specified, the conditional mean tensor is generated form model~\eqref{eq:decomp}, where the core tensor entries are i.i.d.\ drawn from Uniform[-1,1], the factor matrix $\mM_k$ is uniformly sampled with respect to Haar measure from matrices with orthonormal columns.
-The feature matrix $\mX_k$ is either an identity matrix (i.e.,\ no feature  available) or Gaussian random matrix with i.i.d.\ entries from $N(0,1)$.
-The linear predictor $\Theta=\tC\times\{\mM_1\mX_1,\mM_2\mX_3,\mM_3\mX_3\}$ is scaled such that $\mnormSize{}{\Theta}=1$.
-Conditional on the linear predictor $\Theta=\entry{\theta_{ijk}}$, the entries in the tensor $\tY=\entry{y_{ijk}}$ are drawn independently according to one of the following three probabilistic models:
-
-\begin{enumerate}
-\item[(a)] (Gaussian). Continuous entries $y_{ijk}\sim N\left(\alpha \theta_{ijk}, 1\right)$, for all $(i,j,k)\in[d_1]\times[d_2]\times[d_3]$.
-\item[(b)] (Poisson). Count entries $y_{ijk}\sim\textup{Poisson}\left( e^{\alpha \theta_{ijk}}\right)$, for all $(i,j,k)\in[d_1]\times[d_2]\times[d_3]$.
-\item[(c)] (Bernoulli). Binary entries $y_{ijk}\sim \textup{Bernoulli}\left( \frac{e^{\alpha \theta_{ijk}}}{1+e^{\alpha \theta_{ijk}}}\right)$, for all $(i,j,k)\in[d_1]\times[d_2]\times[d_3]$.
-\end{enumerate}
-Here $\alpha>0$ is a scalar controlling the magnitude of the effect size.
-In each simulation study, we report the mean squared error (MSE), $\FnormSize{}{\tB_{\textup{true}}-\hat \tB}^2$, for the coefficient tensor averaged across $n_{\textup{sim}}=30$ replications.
-
-\subsection{Finite-sample performance}
-The first experiment assesses the selection accuracy of our BIC criterion~\eqref{eq:BIC}.
-We consider the balanced situation where $d_k=d$, $p_k=0.4d_k$ for $k=1,2,3$.
-We set $\alpha=4$ and consider various combinations of dimension $d$ and rank $\mr=(r_1,r_2,r_3)$.
-For each combination, we simulate tensor data following Gaussian and Poisson models.
-We then minimize BIC using a grid search over three modes.
-The hyper-parameter $\alpha$ is set to infinity in the fitting, which essentially imposes no constraint on the tensor magnitude.
-Table~\ref{tab:rank} reports the selected rank averaged over $n_{\textup{sim}}=30$ replicates.
-We found that in the high-rank setting with $d=20$, the selected rank slightly underestimates the true rank, and the accuracy immediately improves when either the dimension increases to $d = 40$ or the rank reduces to $\mr = (3,3,3)$.
-This agrees with our expectation, because in the tensor decomposition, the sample size is related to the number of tensor entries.
-A larger $d$ implies a larger sample size, so the BIC selection becomes more accurate.
-
-\begin{table}[tb]
-\centering
-\begin{tabular}{|c|cc|cc|}
-\hline
-True Rank &\multicolumn{2}{c|}{Dimension (Gaussian tensors)}&\multicolumn{2}{c|}{Dimension (Poisson tensors)}\\
-$\mr$& $d=20$  &$d=40$&$d=20$ &$d=40$\\
-\hline
-$(3,\ 3,\ 3)$&$({\bf 3},\ {\bf 3},\ {\bf 3})$&$({\bf 3},\ {\bf 3},\ {\bf 3})$& $({\bf 3},\ {\bf 3},\ {\bf 3})$ & $({\bf 3},\ {\bf 3},\ {\bf 3})$\\
-$(4,\ 4,\ 6)$&$(3,\ 3,\ {\bf 4.6})$&$({\bf 4},\ {\bf 4},\ {\bf 5.3})$&$(3,\ 3,\ {\bf 5.3})$&$({\bf 4},\ {\bf 4},\ {\bf 5.6})$\\
-$(6,\ 8,\ 8)$&$(5,\ 5,\ 5)$&$({\bf 6},\ {\bf 8},\ {\bf 8})$&$(5,\ 5,\ 6.7)$&$({\bf 6},\ {\bf 8},\ {\bf 8})$\\
-\hline
-\end{tabular}
-\caption{Rank selection via BIC. The estimated ranks are averaged across 30 simulation. Bold number indicates the ground truth is within two standard deviations of the estimate.}\label{tab:rank}
-\end{table}
-
-The second experiment evaluates the accuracy when features are available on all modes.
-We set $\alpha=10, d_k=d, p_k=0.4d_k, r_k=r\in\{2,4,6\}$ and increase $d$ from 30 to 60.
-Our theoretical analysis suggests that $\hat \tB$ has a convergence rate $\tO(d^{-2})$ in this setting.
-Figure~1 plots the estimation error versus the ``effective sample size'', $d^2$, under three different distribution models.
-We find that the empirical MSE decreases roughly at the rate of $1/d^2$, which is consistent with our theoretical results.
-We also observe that, tensors with higher rank tend to yield higher estimation errors, as reflected by the upward shift of the curves as $r$ increases.
-Indeed, a larger $r$ implies a higher model complexity and thus greater difficulty in the estimation.
-Similar behaviors are observed in the non-Gaussian tensors in Figures~\ref{fig:dim}b-c.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=15cm]{dimension.pdf}
-\caption{Estimation error against effective sample size.
-The three panels plot the MSE when the response tensors are generated from (a) Gaussian (b) Poisson and (c) Bernoulli models.
-The dashed curves correspond to $\tO({1/d^2})$.}\label{fig:dim}
-\end{figure}
-
-\subsection{Comparison with GLMs under stochastic block models}
-The third experiment investigates the performance of our model under correlated feature effects.
-We mimic the scenario of brain imaging analysis.
-A sample of $d_3=50$ networks are simulated, one for each individual.
-Each network measures the connections between $d_1=d_2=20$ brain nodes.
-We simulate $p=5$ features for the each of the 50 individuals.
-These features may represent, for example, age, gender, cognitive score, etc.
-Recent study has suggested that brain connectivity networks often exhibit community structure represented as a collection of subnetworks, and each subnetwork is comprised of a set of spatially distributed brain nodes.
-To accommodate this structure, we utilize the stochastic block model~\citep{abbe2017community} to generate the effect size.
-Specifically, we partition the nodes into $r$ blocks by assigning each node to a block with uniform probability.
-Edges within a same block are assumed to share the same feature effects, where the effects are drawn i.i.d.\ from $N(0,1)$.
-We then apply our tensor regression model to the network data using the BIC-selected rank.
-Note that in this case, the true model rank is unknown; the rank of a $r$-block network is not necessarily equal to matrix rank $r$~\citep{zeng2019multiway}.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=16cm]{compare_GLM.pdf}
-\caption{Performance comparison under stochastic block models.
-The three panels plot the MSE when the response tensors are generated from (a) Gaussian (b) Poisson and (c) Bernoulli models.
-The $x$-axis represents the number of blocks in the networks.
-}\label{fig:glm}
-\end{figure}
-
-Figure~\ref{fig:glm} compares the MSE of our method with a multiple-response GLM approach.
-The multiple-response GLM is to regress the dyadic edges, one at a time, on the features, and this model is repeatedly fitted for each edge.
-As we find in Figure~\ref{fig:glm}, our tensor regression method achieves significant error reduction in all three data types considered.
-The outperformance is substantial in the presence of large communities; even in the less structured case ($\sim 20/15=1.33$ nodes per block), our method still outer-performs GLM.
-The possible reason is that the multiple-response GLM approach does not account for the correlation among the edges, and suffers from overfitting.
-In contrast, the low-rankness in our modeling incorporates the shared information across entries.
-By selecting the rank in a data-driven way, our method achieves accurate estimation in a wide range of settings.
-
-\subsection{Comparison with other tensor methods}
-We compare our supervised tensor decomposition with three other tensor methods:
-\begin{itemize}
-\item  Higher-order low-rank regression ({\bf HOLRR},~\citet{rabusseau2016low}) is a least-square based tensor regression that allows features on a single mode. 
-\item Higher-order partial least square ({\bf HOPLS},~\citet{zhao2012higher}) is a dimension-reduction method that jointly models a tensor response and a tensor feature. 
-\item Subsampled tensor projected gradient ({\bf TPG},~\citet{yu2016learning}) considers the same objective as {\bf HOLRR} but instead uses a different algorithm to solve the problem. 
-\end{itemize}
-These three methods are the closest algorithms to ours, in that they all relate a data tensor to features using a low-rank structure.
-The three existing methods allow only Gaussian data, whereas ours is applicable to any exponential family distribution including Gaussian, Bernoulli, Poisson, etc.
-For fair comparison, we consider Gaussian tensors in the experiment.
-Because not every method returns the effect estimate $\hat \tB$ as outputs, we measure the accuracy using mean squared prediction error, $\text{MSPE}=(\prod_kd_k)^{-1}\FnormSize{}{\hat \tY-f(\Theta)}^2$, where $f(\Theta)$ is the conditional mean of the tensor, and $\hat \tY$ is the fitted tensor from each method.
-
-The comparison is assessed from three aspects: (a) benefit of incorporating features from multiple modes; (b) prediction error with respect to sample size; (c) sensitivity of accuracy with respect to model complexity.
-We use similar simulation setups as in our experiment II, but consider combinations of rank ($\mr=(3,3,3)$ vs.\ $(4,5,6)$), signal ($\alpha = 3$ vs.\ $6$), and dimension ($d$ ranging from 20 to 100 for modes with features, $d = 20$ for modes without features).
-Two methods ({\bf STD} and {\bf HOLRR}) require the tensor rank as inputs.
-For fair comparison, we provide both algorithms the true rank.
-For algorithms ({\bf HOPLS} and {\bf TPG}) that use different notions of model rank, we use a grid search to set the hyperparameter that gives the best mean square prediction error.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=16cm]{compare_alternative.pdf} 
-\caption{Comparison between different tensor methods.
-Panels (a) and (b) plot MSPE versus the number of modes with available features.
-Panels (c) and (d) plot MSPE versus the effective sample size $d^2$.
-We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and signal $\alpha =3 $ (low), $\alpha=6$ (high).}~\label{fig:compare}
-\end{figure}
-
-Figures~\ref{fig:compare}a-b show the averaged prediction error across 30 replicates.
-We see that our {\bf STD} outperforms others, especially in the low-signal, high-rank setting.
-As the number of informative modes (i.e., modes with available features) increases, the {\bf STD} exhibits a substantial reduction in error whereas others remain unchanged (Figure~\ref{fig:compare}b).
-This showcases the benefit of incorporation of multiple features.
-Note that our method {\bf STD} is most comparable to {\bf HOLRR} when there is only a single informative mode.
-In such a case, both methods share a same cost function but have different algorithms.
-{\bf STD} alternates between informative and non-informative modes, whereas {\bf HOLRR} approximates the non-informative modes without alternating.
-The accuracy gain in Figure~\ref{fig:compare} demonstrates the benefit of alternating algorithm -- incorporation of informative modes also improves the estimation in the non-informative modes.
-
-Figures~\ref{fig:compare}c-d compare the prediction error with respect to effective sample size when only one mode has side information.
-In the high-signal low-rank setting, our method has similar performance as {\bf HOLRR}, and the improvement becomes more pronounced in the low-signal high-rank setting.
-Note that the latter setting is a harder learning task because of the higher inter-mode complexity, and our {\bf STD} method shows the advantage in addressing this challenge.
-
-\section{Data analyses}\label{sec:data}
-We apply our supervised tensor decomposition to two datasets.
-The first application studies the brain networks in response to individual attributes (i.e.,\ feature on one mode), and the second application focuses on multi-relational network analysis with dyadic attributes (i.e.,\ features on two modes).
-
-\subsection{Analyses of human brain connection data}
-The Human Connectome Project (HCP) aims to build a network map that characterizes the anatomical and functional connectivity within healthy human brains~\citep{HCP}.
-We analyze the structural connectivity patterns among 68 brain regions for 136 individuals from HCP.
-We follow the preprocessing procedure as in~\cite{zhang2018mapping} and parcellate the brain into 68 regions of interest~\citep{desikan2006automated}.
-The preprocessed dataset is released in our R package \texttt{tensorregress}.
-
-The dataset consists of 136 brain structural networks, one for each individual.
-Each brain network is represented as a 68-by-68 binary matrix, where the entries encode the presence or absence of fiber connections between the 68 brain regions.
-We consider four individual features: gender (65 females vs.\ 71 males), age 22-25 ($n=35$), age 26-30 ($n=58$), and age 31+ ($n=43$).
-The goal is to identify the connection edges that are affected by individual features.
-A key challenge in brain network is that the edges are correlated; for example, the nodes in edges may be from a same brain region, and it is of importance to take into account the within-dyad dependence.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=12cm]{HCP.pdf}
-\caption{Top edges with large effects.
-(a) Global effect; (b) Female effect; (c) Age 22-25; (d) Age 31+.
-Red edges represent positive effects and blue edges represent negative effects.
-The edge-width is proportional to the magnitude of the effect size.
-}\label{fig:brain}
-\end{figure}
-
-We perform the supervised tensor decomposition to the HCP data.
-The data tensor is binary, $\tY\in\{0,1\}^{68\times 68\times 136}$, and the features are of dimension 4 on the 3$^{\text{rd}}$ mode.
-The BIC selection suggests a rank $\mr=(10,10,4)$ with quasi log-likelihood $\tL_{\tY}=-174654.7$.
-We utilize the sum-to-zero contrasts in coding the feature effects, and depict only the top 3\% edges whose connections are non-constant across the sample.
-Figure~\ref{fig:brain} shows the top edges with high effect size, overlaid on the Desikan atlas brain template~\citep{desikan2006automated,xia2013brainnet}.
-We find that the global connection exhibits clear spatial separation, and that the nodes within each hemisphere are more densely connected with each other (Figure~\ref{fig:brain}a).
-In particular, the superior-temproal (\emph{SupT}), middle-temporal (\emph{MT}) and Insula are the top three popular nodes in the network.
-Interestingly, female brains display higher inter-hemispheric connectivity, especially in the frontal, parental and temporal lobes (Figure~\ref{fig:brain}b).
-This is in agreement with a recent study showing that female brains are optimized for inter-hemispheric communication~\citep{ingalhalikar2014sex}.
-We also found several edges with declined connection in the group Age 31+.
-Those edges involve Frontal-pole (\emph{Fploe}), superior-frontal (\emph{SupF}) and Cuneus nodes.
-The Frontal-pole region has long been known for its importance in memory and cognition, and the detected decline with age further highlights its biological importance.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=16cm]{compare_HCP.pdf}
-\caption{Comparison of feature effect estimation in the HCP data.}\label{fig:s1}
-\end{figure}
-
-Figure~\ref{fig:s1} compares the estimated coefficients from our method with those from multiple-response GLM approach.
-The multiple-response GLM is to regress the brain edges, one at a time, on the individual-level features, and this logistic model is repeatedly fitted for every edge $\in [68]\times [68]$.
-As we can see in the figure, our tensor decomposition method shrinkages the coefficients towards center, thereby implicitly enforcing the sharing between feature effects.
-
-\subsection{Application to political relation data}
-
-The second application studies the multi-relational networks with node-level attributes.
-We consider \emph{Nations} dataset~\citep{nickel2011three} which records 56 relations among 14 countries between 1950 and 1965 (post-World War II).
-The multi-relational networks can be organized into a $14 \times 14 \times 56$ binary tensor, with each entry indicating the presence or absence of an action, such as ``sending tourist to'', ``export'', ``import'', between countries.
-The 56 relations span the fields of politics, economics, military, religion, and so on.
-In addition, country-level attributes are also available, and we focus on the following six features: \emph{constitutional, catholics, law ngo, political leadership, geography}, and \emph{medicine ngo}.
-The goal is to identify the variation in connections due to country-level attributes and interactions thereof.
-One of the key features is that the 56 relations are correlated, and we would like to take it into account in assessing the feature effects.
-
-\begin{table}[H]
-\begin{tabular}{c|c|l}
-Cluster & Category & \multicolumn{1}{c}{Relations}\\
-\hline
-\hline
-\multirow{3}{*}{Cluster I}  & \multirow{3}{*}{Military} & warning,           violentactions, militaryactions, duration,\\ &&negativebehavior, severdiplomatic, expeldiplomats, protests,\\
-&&timesincewar, commonbloc0, blockpositionindex, commonbloc1\\   
-\hline
-\multirow{4}{*}{Cluster II}& \multirow{4}{*}{Emigrant} & emigrants, relemigrants, emigrants3, accusation, \\      
-&&nonviolentbehavior, students, relstudents, economicaid\\
-&&releconomicaid, negativecomm, relintergovorgs,relngo\\
-&&intergovorgs3, ngoorgs3, militaryalliance, commonbloc2\\    
-\hline
-\multirow{6}{*}{Cluster III} & \multirow{6}{*}{Economy} & treaties, reltreaties, oficialvisits, conferences, exportbooks\\
-&& relexportbooks, booktranslations, relbooktranslations\\
-&&boycottembargo, weightedunvote, unweightedunvote, tourism\\
-&&reltourism, tourism3, exports, relexports, exports3, \\
-&&intergovorgs, ngo, embassy, reldiplomacy,timesinceally,\\
-&& dependent, independence\\
-\hline
-Cluster IV & Territory &aidenemy, lostterritory, unoffialacts, attackembassy\\
-\hline
-\end{tabular}
-\caption{$K$-means clustering of relations based on dimension reduction on the 3$^\text{rd}$ mode. }\label{tab:s1}
-\end{table}
-
-We apply our tensor model to the \emph{Nations} data.
-The multi-relational network $\tY\in\{0,1\}^{14\times 14\times 56}$ is a binary data tensor, and the country attributes $\mX\in\mathbb{R}^{14\times 6}$ are features on both the 1$^\text{st}$ and 2$^\text{nd}$ modes.
-The BIC criterion suggests a rank $\mr=(4,4,4)$ for the coefficient tensor $\tB\in\mathbb{R}^{6\times 6 \times 56}$.
-We perform the supervised tensor decomposition and obtain the dimension reduction matrices $\hat \mM_k$ from the model.
-Then we apply $K$-mean clustering to dimension reduction matrix on each of the modes.
-Table~\ref{fig:s1} shows the K-means clustering of the 56 relations based on the dimension reduction matrix on the 3$^\text{rd}$ mode.
-We find that the relations reflecting the similar aspects of actions are grouped together.
-In particular, Cluster I consists of military relations such as \emph{violentactions}, \emph{warnings} and \emph{militaryactions}; Clusters II and III capture the economical relations such as \emph{economicaid, booktranslations, tourism}; and Cluster IV represents the political alliance and territory relations.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=12cm]{coef.pdf}
-\caption{Estimated feature effects in the \emph{Nations} data analysis.
-Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warnning}, \emph{violentactions},\emph{treaties}, and \emph{aidenemy}, respectively.
-}\label{fig:est}
-\end{figure}
-
-To investigate the effects of dyadic attributes towards connections, we depict the estimated coefficients $\hat \tB=\entry{\hat b_{ijk}}$ for several relation types (Figure~\ref{fig:est}).
-Note that the each entry $\hat b_{ijk}$ estimates the contribution, at the logit scale, of feature pair $(i,j)$ ($i$th feature for the ``sender'' country and $j$th feature for the ``receiver'' country) towards the connection of relation $k$.
-Several interesting findings emerge from the observation.
-We find that relations belonging to a same cluster tend to have similar feature effects.
-For example, the relations ``warning'' and ''violentactions'' are classified into Cluster I, and both exhibit similar effect patterns (Figures~\ref{fig:est}a-b).
-Moreover, the feature \emph{constitutional} has a strong effect in the relation ``violentactions'' and ``warning'', whereas the effect is weaker in the relation ``treaties''.
-The result is plausible because the constitutional attributes affect political actions more often than economical actions.
-The entries in $\tB$ are useful for revealing interaction effects in a context-specific way.
-For a given relation $k$, the off-diagonal entries in the matrix $\hat \tB(\colon,\colon,k)$ represent the pairwise interactions between features.
-From Figure~\ref{fig:est}, we find a strong interaction between \emph{geography} and \emph{political leadership} in the relation ``warning'', and a strong interaction between \emph{geogrphy} and \emph{medicine ngo} in the relation ``aidenemy''.
-The relation-specific effect pattern showcases the applicability of our method in revealing complex interactions.
-
-\section{Discussion and future work}\label{sec:con}
-We have developed a supervised tensor decomposition method with side information on multiple modes.
-One important challenge of tensor data analysis is the complex interdependence among tensor entries and between multiple features.
-Our approach incorporates side information as feature matrices in the conditional mean tensor.
-The empirical results demonstrate the improved interpretability and accuracy over previous approaches.
-Applications to the brain connection and political relationship datasets yield conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.
-
-We comment on a number of possible extensions.
-We have provided accuracy guarantees for parameter estimation in the supervised tensor model.
-Statistical inference based on tensor decomposition is an important future direction.
-Measures of uncertainty, such as confidence envelope for space estimation, would be useful.
-One possible approach would be performing parametric bootstrap~\citep{efron1994boot} to assess the uncertainty in the estimation.
-For example, one can simulate tensors from the fitted low-rank model based on the estimates, and then assess the empirical distribution of the estimates.
-While being simple, bootstrap approach is often computationally expensive for large-scale data.
-Another possibility is to leverage recent development in debiased inference with distributional characterization~\citep{chen2019inference,xia2019confidence}.
-This approach has led to fruitful results for matrix data analysis.
-Uncertainly quantification involving tensors are generally harder than matrices, and establishing distribution theory for tensor estimation remains an open problem.
-
-One assumption made by our method is that tensor entries are conditionally independent given the linear predictor $\Theta$.
-This assumption can be extended by introducing a more general mixed-effect tensor model.
-For example, in the special case of Gaussian model, we can model the first two moments of data tensor using
-\begin{align}\label{eq:extention}
-\mathbb{E}(\tY|\mX_1,\ldots,\mM_K)&=\tC\times\mM_1\times \cdots \times \mM_K, \notag \\
-\text{Var}(\tY|\mX_1,\ldots,\mM_K)&=\mPhi_1\otimes \cdots \otimes \mPhi_K,
-\end{align}
-where $\mPhi_k\in\mathbb{R}^{d_k\times d_k}$ is the unknown covariance matrix on the mode $k\in[K]$.
-For general exponential family, an additional mean-variance relationship should also be considered.
-The joint estimation of mean model $\Theta$ and variance model $\mPhi_k$ will lead to more efficient estimation in the presence of unmeasured confounding effects.
-The introduction of unknown covariance matrices $\mPhi_k$ dramatically increases the number of parameters in the problem.
-Suitable regularizatiosn such as graphical lasso or specially-structured covariance assumptions should be considered.
-The extension of tensor modeling with heterogeneous mixed-effects will be an important future direction.
-
-Although we have presented the supervised tensor decomposition in the context of order-3 data tensors, the framework applies to more general multi-way datasets.
-One possible extension is the integrative analysis of omics data, in which multiple types of omics measurements (gene expression, DNA methylation, microRNA) are collected in the same set of individuals~\citep{lock2013joint,wang2019three}.
-In such a case, the observation is a stack of data matrices, which may not be of the same dimension.
-One possible extension is to impose linked dimension reduction among different data matrices, and incorporate side information as features in the joint decomposition.
-Other applications include time-series tensor data with multiple side information.
-Exploiting the benefits and properties of the supervised tensor decomposition in specialized task will boost scientific discoveries.
-
-% for blind version
-\section*{Acknowledgements}
-This research was supported by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
-
-\bigskip
-\begin{center}
-{\large\bf SUPPLEMENTARY MATERIALS}
-\end{center}
-
-\begin{description}
-
-\item[Supplementary notes:]  technical proofs and additional simulation results.
-
-\item[Data and software:] R-package \texttt{tensorregress} performs our supervised tensor decomposition method.
-The source code and data used in the paper are available at \url{https://CRAN.R-project.org/package=tensorregress.} We also provide R scripts for reproducing all numbers in Figures 2-8 and Tables 2-3 in the paper.
-
-\end{description}
-
-\bibliography{tensor_wang}
-\bibliographystyle{apalike}
-
 \end{document}

### 2021-04-02 tensor_regress/ICSA_student/latexresource/icsa.tex
submission

--- 2020-09-22 tensor_regress/JCGS_submission/jcgs-tensor/jcgs-tensor.tex
@@
 
-%\usepackage{enumerate}
-\usepackage{natbib}
-\usepackage{url} % not crucial - just used below for the URL
-
-%\pdfminorversion=4
-% NOTE: To produce blinded version, replace "0" with "1" below.
 \newcommand{\blind}{0}
 
-% DON'T change margins - should be 1 inch all around.
 \addtolength{\oddsidemargin}{-.5in}%
@@
 \addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{1.3in}%
+\addtolength{\textheight}{1in}%
 \addtolength{\topmargin}{-.8in}%
@@
 \newtheorem{lem}{Lemma}
-\newtheorem{prop}{Proposition}[section]
+\newtheorem{prop}{Proposition}
 \newtheorem{pro}{Property}
@@
 
-%\bibliographystyle{natbib}
-
 \def\spacingset#1{\renewcommand{\baselinestretch}%
@@
 
-\if0\blind
+\if1\blind
 {
   \title{\bf Supervised Tensor Decomposition with Interactive Side Information}
-  \author{Jiaxin Hu, Zhuoyan Xu, Chanwoo Lee, and Miaoyan Wang\thanks{The authors gratefully acknowledge NSF grant DMS-1915978 and funding from the Wisconsin Alumni Research Foundation.
-}\hspace{.2cm}\\
+  % blind version
+  \author{Jiaxin Hu, Zhuoyan Xu, Chanwoo Lee, and Miaoyan Wang\\
     Department of Statistics, University of Wisconsin-Madison}
   \maketitle
-
 } \fi
 
-\if1\blind
+\if0\blind
 {
@@
 Our method will help the practitioners efficiently analyze tensor datasets in various areas.
-Toward this end, the package and data used are available at~\url{https://CRAN.R-project.org/package=tensorregress}.
+%Toward this end, the package and data used are available at~\url{https://CRAN.R-project.org/package=tensorregress}.
 
@@
 \noindent%
-{\it Keywords:}  Tensor data analysis, Supervised dimension reduction, Exponential family distribution, Generalized multilinear model, Alternating optimization
+{\it Keywords:} Applications and case studies, Tensor data analysis, Supervised dimension reduction, Exponential family distribution, Generalized multilinear model
 \vfill
@@
 
+In addition to the side information, another challenge is that many tensor datasets consist of non-Gaussian measurements \citep{wang2020learning, lee2020tensor}.
+Classical tensor decomposition methods are based on minimizing the Frobenius norm of deviation, leading to suboptimal predictions for binary- or count-valued response variables.
+A number of supervised tensor methods have been proposed \citep{narita2012tensor, zhao2012higher, yu2016learning,lock2018supervised}.
+These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest.
+
+\iffalse
 In addition to the aforementioned challenges, many tensor datasets consist of non-Gaussian measurements.
@@
 The gap between theory and practice means a great opportunity to modeling paradigms and better capture the complexity in tensor data.
+\fi
 
-\begin{figure*}[t]
+\begin{figure*}[ht]
 \begin{center}
-\includegraphics[width=16cm]{demo.pdf}
+\includegraphics[width=15cm]{demo.pdf}
 \end{center}
@@
 Our work is closely related to but also clearly distinctive from several lines of previous work.
-The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor, hong2020generalized,wang2017tensor,bi2018recom,chi2012tensors} that aims to find the best low-rank representation of a data tensor.
+The first line is a class of \textit{unsupervised} tensor decomposition such as Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor, hong2020generalized,wang2017tensor,bi2018recom} that aims to find the best low-rank representation of a data tensor.
 In contrast, our model is a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and multiple features.
-The low-rank factorization is determined jointly by the tensor data and feature matrices in our model.
+The low-rank factorization is decided jointly by tensor data and feature matrices in model.
 
@@
 Our model shares a common ground with earlier approaches, but we provide more efficient solutions to new settings that have more practical significance.
-As we show in Section~\ref{sec:connection}, the supervised tensor decomposition has an interesting connection with tensor-on-tensor regression.
+%As we show in Section~\ref{sec:connection}, the supervised tensor decomposition has an interesting connection with tensor-on-tensor regression. 
 Previous methods~\citep{lock2018tensor,lock2018supervised} mainly focus on Gaussian tensors.
@@
 In contrast, our goal is interpretable prediction, and we estimate factor matrices $\mM_k$ using a model-based approach.
-Estimating $\mM_k$ allows us to identify sufficient features and the interactions thereof.
-We numerically compare these two approaches in Section~\ref{sec:simulation}.
+Estimating $\mM_k$ allows us to identify sufficient features and interactions thereof.
+%We numerically compare these two approaches in Section~\ref{sec:simulation}.
 
-The remainder of the paper is organized as follows.
-Section~\ref{sec:pre} introduces tensor preliminaries.
-Section~\ref{sec:model} presents the main model and three motivating examples for supervised tensor decomposition.
-We describe the quasi-likelihood estimation and alternating optimization algorithm in Section~\ref{sec:est}.
-In Section~\ref{sec:simulation}, we present numerical experiments and assess the performance in comparison to alternative methods.
-In Section~\ref{sec:data}, we apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data.
-We conclude in Section~\ref{sec:con} with discussions about our findings and avenues of future work.
-All proofs are deferred to Supplementary Materials.
-
+\vspace{-0.6cm}
 \section{Preliminaries}\label{sec:pre}
@@
 For ease of presentation, we use the shorthand $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product.
-For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as
-\[
-\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.
-\]
-The tensor Frobenius norm and maximum norm are defined as
-\[
-\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2},\quad \text{and}\quad \mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.
-\]
-When $a$ is a vector, we use $\vnormSize{}{a}=\langle a,a \rangle^{1/2}$ to denote the vector 2-norm.
+For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as  $\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.$ The tensor Frobenius norm and maximum norm are defined as $\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2},\quad \text{and}\quad \mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.$ When $a$ is a vector, we use $\vnormSize{}{a}=\langle a,a \rangle^{1/2}$ to denote the vector 2-norm.
 
@@
 
+\vspace{-0.5cm}
+
 \section{Motivation and model}\label{sec:model}
-\subsection{General framework for supervised tensor decomposition}
+%\subsection{General framework for supervised tensor decomposition}
+
+Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor.
+Suppose the side information is available on each of the $K$ modes.
+Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the feature matrix on the mode $k\in[K]$, where $x_{ij}$ denotes the $j$-th feature value for the $i$-th tensor entity, for $(i,j)\in[d_k]\times[p_k]$, $p_k\leq d_k$.
+We assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent realizations from an exponential family distribution, and the conditional mean tensor admits the form
+\begin{align}\label{eq:decomp}
+\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1, \ldots, \mX_K\mM_K\}\right),\notag\\
+\text{with} \ &\ \mM_k^T\mM_k = \mI_{r_k},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\quad \text{for all }k=1,\ldots,K.
+\end{align}
+where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a full-rank core tensor, $\mM_k\in\mathbb{R}^{p_k\times r_k}$ are factor matrices consisting of orthonormal columns with $r_k\leq p_k$ for all $k\in[K]$, and $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$.
+Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and exponential link for Poisson distribution.
+
+Figure~\ref{fig:intro1}b provides a schematic illustration of our model.
+The features $\mX_k$ affect the distribution of tensor entries in $\tY$ through the sufficient features of the form $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$.
+The core tensor $\tC$ collects the interaction effects between sufficient features across $K$ modes, and thus allows the identification of variations in the tensor data attributable to the side information.
+Our goal is to find $\mM_k$ and the corresponding $\tC$ to reveal the relationship between side information $\mX_k$ and the observed tensor $\tY$.
+Note that $\mM_k$ and $\tC$ are identifiable only up to orthonormal transformations.
+
+\iffalse
 We begin with a general framework for supervised tensor decomposition and then discuss its implication in three concrete examples.
@@
 
+\fi
+
+We give two seemingly different examples that can all be formulated as our supervised tensor decomposition model~\eqref{eq:decomp}.
+
+\begin{example}[Spatio-temporal growth model]
+The growth curve model~\citep{srivastava2008models} was originally proposed as an example of bilinear model for matrix data, and we extend it to higher-order cases.
+Let $\tY=\entry{y_{ijk}}\in\mathbb{R}^{d \times m\times n}$ denote the pH measurements of $d$ lakes at $m$ levels of depth and for $n$ time points.
+Suppose the sampled lakes belong to $q$ types, with $p$ lakes in each type.
+Let $\{\ell_j\}_{j\in[m]}$ denote the sampled depth levels and $\{t_k\}_{k\in[n]}$ the time points.
+Assume that the expected pH trend in depth is a polynomial of order at most $r$ and that the expected trend in time is a polynomial of order $s$.
+Then, the conditional mean model for the spatio-temporal growth is a special case of our model~\eqref{eq:decomp}, where $\mX_1=\text{blockdiag}\{\mathbf{1}_q,\ldots,\mathbf{1}_q\}\in \{0,1\}^{d\times p}$ is the design matrix for lake types,
+\[
+\mX_2=
+\begin{pmatrix}
+1 & \ell_1&\cdots &\ell^{r}_1\\
+1 & \ell_2&\cdots &\ell^{r}_2\\
+\vdots &\vdots&\ddots&\vdots\\
+1&\ell_{m}&\cdots&\ell^{r}_{m}
+\end{pmatrix},\
+\mX_3=
+\begin{pmatrix}
+1 & t_1&\cdots &t^{s}_1\\
+1 & t_2&\cdots &t^{s}_2\\
+\vdots &\vdots&\ddots&\vdots\\
+1&t_{n}&\cdots&t^{s}_{n}
+\end{pmatrix}
+\]
+are the design matrices for spatial and temporal effects, respectively. The spatial-temporal mode has covariates available on each of the three modes.
+
+\end{example}
+\begin{example}[Network population model] 
+Network response model is recently developed in the context of neuroimanig analysis.
+The goal is to study the relationship between network-valued response and the individual covariates.
+Suppose we observe $n$ i.i.d.\ observations $\{(\mY_i, \mx_i): i=1,\ldots,n\}$, where $\mY_i\in\{0,1\}^{d\times d}$ is the brain connectivity network on the $i$-th individual, and $\mx_i\in\mathbb{R}^p$ is the individual covariate such as age, gender, cognition, etc.
+The network-response model~\citep{rabusseau2016low} has the form
+\begin{equation}\label{eq:network}
+\text{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n
+\end{equation}
+where $\tB\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor of interest, and $\times_3$ denotes the tensor-by-matrix multiplication along the 3rd mode \citep{kolda2009tensor}.
+The model~\eqref{eq:network} is also a special case of our tensor-response model, with covariates on the last mode of the tensor.
+\end{example}
+
+\iffalse
 \subsection{Three examples}
-We give three seemingly different examples that can all be formulated as our supervised tensor decomposition model~\eqref{eq:decomp}.
+
 \begin{example}[Spatio-temporal growth model]
@@
   \end{example}
+\fi
 
@@
 
-In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by features.
+In the above two examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by features.
 Our supervised tensor decomposition~\eqref{eq:decomp} allows arbitrary numbers of feature matrices.
@@
 
+\iffalse
 \subsection{Connection to sufficient dimension reduction and tensor-on-tensor regression}\label{sec:connection}
@@
 Model~\eqref{eq:interaction} shows the benefit of our supervised tensor decomposition for identifying across-mode interactions.
-
+\fi
 \section{Estimation}\label{sec:est}
@@
 where $b(\cdot)$ is a known function, $\theta$ is the linear predictor, $\phi>0$ is the dispersion parameter, and $c(\cdot)$ is a known normalizing function.
-The choice of link functions depends on the data types and on the observation domain of $y$, denoted $\mathbb{Y}$.
+The choice of link functions depends on the data types and the observation domain of $y$, denoted $\mathbb{Y}$.
 For example, the observation domain is $\mathbb{Y}=\mathbb{R}$ for continuous data, $\mathbb{Y}=\mathbb{N}$ for count data, and  $\mathbb{Y}=\{0,1\}$ for binary data.
 The canonical link function $f$ is chosen to be $f(\cdot)=b'(\cdot)$, the first-order derivative of $b(\cdot)$.
+
+\iffalse
 Table~1 summarizes the canonical link functions for common types of distributions.
@@
 \end{table}
+\fi
 
@@
 Ignoring constants that do not depend on $\Theta$, the quasi log-likelihood of~\eqref{eq:decomp} is equal to Bregman distance between $\tY$ and $b'(\Theta)$:
-\begin{align}\label{eq:loglikelihood}
-\tL_{\tY}(\tC,\mM_1,\ldots,\mM_K)&=\langle \tY, \Theta \rangle - \sum_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K}),\\
- \text{where}\quad \Theta&=\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
-\end{align}
+\begin{equation}\label{eq:loglikelihood}
+\tL_{\tY}(\tC,\mM_1,\ldots,\mM_K) =\langle \tY, \Theta \rangle - \sum_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K}), \text{ with}\quad \Theta =\tC\times\{\mM_1\mX_1,\ldots,\mM_K\mX_K\}.
+\end{equation}
 We propose a constrained maximum quasi-likelihood estimator (M-estimator),
 \begin{align} \label{eq:MLE} 
+\vspace{-0.1cm}
 (\hat \tC, \hat \mM_1,\ldots,\hat \mM_k) &=\argmax_{(\tC,\mM_1,\ldots,\mM_K)\in \tP} \ \tL_{\tY}(\tC,\mM_1,\ldots,\mM_K),
 \end{align}
+\vspace{-0.1cm}
 where the parameter space $\tP$ is
-\[
+\begin{equation}
 	\tP=\left\{(\tC, \mM_1,\ldots,\mM_K) \ \Big| \ \mM_k^T\mM_k=\mathbf{I}_{r_k}\ \text{for all }k\in[K],\  \mnormSize{}{\Theta(\tC,\mM_1,\ldots,\mM_K)}\leq \alpha \right\}.
-\]
+\end{equation}
+\vspace{-0.1cm}
 The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to avoid the divergence in the non-Gaussian variance.
@@
 Under mild conditions, our algorithm enjoys global convergence; i.e.\ any sequence of iterates generated by the alternating algorithm converges to a stationary point of $\tL_\tY(\cdot)$ module orthogonal transformation.
+\iffalse
 To establish the convergence properties of Algorithm~\ref{alg:B}, we first introduce the equivalent relationship induced by orthogonal transformation.
@@
 
+\fi
 In theory, global optimality of non-convex optimization is often challenging to obtain.
@@
 In the experiments we considered, we find that Algorithm~\ref{alg:B} typically gives satisfactory convergent points upon random initialization.
-Figure~\ref{fig:loglike} shows the trajectories of objective function for order-3 tensors based on model~\eqref{eq:decomp}, where $d_k = d \in\{25,30\}$, $p_k = 0.4d$, $r_k = r\in\{3,6\}$ for $ k = 1,2,3$.
+
+%Figure~\ref{fig:loglike} shows the trajectories of objective function for order-3 tensors based on model~\eqref{eq:decomp}, where $d_k = d \in\{25,30\}$, $p_k = 0.4d$, $r_k = r\in\{3,6\}$ for $ k = 1,2,3$.
 We consider input tensors with Gaussian, Bernoulli, and Poisson entries.
@@
 
+\iffalse
 \begin{figure}[t]
@@
 
-\subsection{Rank selection and computational complexity}\label{sec:tuning}
+\fi
+
+%\subsection{Rank selection and computational complexity}\label{sec:tuning}
 Algorithm~\ref{alg:B} assumes the rank $\mr$ is given.
@@
 We choose $\hat \mr$ that minimizes $\textup{BIC}(\mr)$ via grid search.
-Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model.
-We evaluate the empirical performance of BIC in Section~\ref{sec:simulation}.
-
+%Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model. 
+More discussions and simulations about the rank selection are in Supplementary Note.
+%We evaluate the empirical performance of BIC in Section~\ref{sec:simulation}.
 The computational complexity of our Algorithm~\ref{alg:B} is $O\left(d \sum_k p^3_k\right)$ for each loop of iterations, where $d=\prod_k d_k$ is the total size of the data tensor.
@@
 This complexity in tensor dimension matches with the classical tensor decomposition~\citep{kolda2009tensor}.
+
+\vspace{-0.6cm}
 
@@
 Therefore, our estimation is consistent in high dimensional regimes, and the convergence becomes especially favorable as the order of tensor data increases.
-
-As immediate applications, we obtain the convergence rate for the three examples mentioned in Section~\ref{sec:model}.
+As immediate applications, we obtain the convergence rate for the two examples mentioned in Section~\ref{sec:model}.
 
@@
 
+\iffalse
 \begin{customexample}{3} [Dyadic data with node attributes] The estimated feature-by-feature matrix converges at the rate $\tO\left(\frac{p}{d^2}\right)$ where $p\leq d$.
@@
 \end{cor}
+\fi
 
@@
 The linear predictor $\Theta=\tC\times\{\mM_1\mX_1,\mM_2\mX_3,\mM_3\mX_3\}$ is scaled such that $\mnormSize{}{\Theta}=1$.
-Conditional on the linear predictor $\Theta=\entry{\theta_{ijk}}$, the entries in the tensor $\tY=\entry{y_{ijk}}$ are drawn independently according to three probabilistic models:
-
-\begin{enumerate}[noitemsep,topsep=0pt]
-\item[(a)] Gaussian model: continuous tensor entries $y_{ijk}\sim N\left(\alpha \theta_{ijk}, 1\right)$.
-\item[(b)] Poisson model: count tensor entries $y_{ijk}\sim\textup{Poisson}\left( e^{\alpha \theta_{ijk}}\right)$.
-\item[(c)] Bernoulli model: binary tensor entries $y_{ijk}\sim \textup{Bernoulli}\left( \frac{e^{\alpha \theta_{ijk}}}{1+e^{\alpha \theta_{ijk}}}\right)$.
-\end{enumerate}
+Conditional on the linear predictor $\Theta=\entry{\theta_{ijk}}$, the entries in the tensor $\tY=\entry{y_{ijk}}$ are drawn independently according to three probabilistic models: (a) Gaussian model: continuous tensor entries $y_{ijk}\sim N\left(\alpha \theta_{ijk}, 1\right)$; (b) Poisson model: count tensor entries $y_{ijk}\sim\textup{Poisson}\left( e^{\alpha \theta_{ijk}}\right)$; (c) Bernoulli model: binary tensor entries $y_{ijk}\sim \textup{Bernoulli}\left( \frac{e^{\alpha \theta_{ijk}}}{1+e^{\alpha \theta_{ijk}}}\right)$.
 Here $\alpha>0$ is a scalar controlling the magnitude of the effect size.
@@
 
+\vspace{-0.5cm}
 \subsection{Finite-sample performance}
+
+\iffalse
 The first experiment assesses the selection accuracy of our BIC criterion~\eqref{eq:BIC}.
@@
 
-\begin{figure}[!h]
-\centering
-\includegraphics[width=14cm]{dimension.pdf}
-\caption{Estimation error against effective sample size.
-The three panels plot the MSE when the response tensors are generated from (a) Gaussian (b) Poisson and (c) Bernoulli models.
-The dashed curves correspond to $\tO({1/d^2})$.}\label{fig:dim}
-\end{figure}
+\fi
 
-The second experiment evaluates the accuracy when features are available on all modes.
+The first experiment evaluates the accuracy when features are available on all modes.
 We set $\alpha=10, d_k=d, p_k=0.4d_k, r_k=r\in\{2,4,6\}$ and increase $d$ from 30 to 60.
@@
 
+\begin{figure}[!h]
+\centering
+\includegraphics[width=12cm]{dimension.pdf}
+\caption{Estimation error against effective sample size.
+The three panels plot the MSE when the response tensors are generated from (a) Gaussian (b) Poisson and (c) Bernoulli models.
+The dashed curves correspond to $\tO({1/d^2})$.}\label{fig:dim}
+\end{figure}
+
+\vspace{-0.6cm}
 \subsection{Comparison with GLMs under stochastic block models}
-The third experiment investigates the performance of our model under correlated feature effects.
+The second experiment investigates the performance of our model under correlated feature effects.
 We mimic the scenario of brain imaging analysis.
@@
 \centering
-\includegraphics[width=16cm]{compare_GLM.pdf}
+\includegraphics[width=13cm]{compare_GLM.pdf}
 \caption{Performance comparison under stochastic block models.
@@
 \subsection{Comparison with other tensor methods}
-We compare our supervised tensor decomposition with three other tensor methods:
+We compare our supervised tensor decomposition with three other tensor methods: Higher-order low-rank regression ({\bf HOLRR},~\citet{rabusseau2016low}),
+% is a least-square based tensor regression that allows features on a single mode.
+ Higher-order partial least square ({\bf HOPLS},~\citet{zhao2012higher}), 
+%is a dimension-reduction method that jointly models a tensor response and a tensor feature. 
+ and Subsampled tensor projected gradient ({\bf TPG},~\citet{yu2016learning}).
+% considers the same objective as {\bf HOLRR} but instead uses a different algorithm to solve the problem. 
+\iffalse
 \begin{itemize}[noitemsep,topsep=0pt]
@@
 \end{itemize}
+\fi 
 These three methods are the closest algorithms to ours, in that they all relate a data tensor to features using a low-rank structure.
@@
 The comparison is assessed from three aspects: (a) benefit of incorporating features from multiple modes; (b) prediction error with respect to sample size; (c) sensitivity of accuracy with respect to model complexity.
-We use similar simulation setups as in our experiment II, but consider combinations of rank, $\mr=(3,3,3)$ (low) vs.\ $(4,5,6)$ (high), signal $\alpha = 3$ (low) vs.\ $6$ (high), dimension $d$ ranging from 20 to 100 for modes with features, and $d = 20$ for modes without features.
+We use similar simulation setups as in our second experiment, but consider combinations of rank, $\mr=(3,3,3)$ (low) vs.\ $(4,5,6)$ (high), signal $\alpha = 3$ (low) vs.\ $6$ (high), dimension $d$ ranging from 20 to 100 for modes with features, and $d = 20$ for modes without features.
 Two methods ({\bf STD} and {\bf HOLRR}) require the tensor rank as inputs.
@@
 \centering
-\includegraphics[width=15cm]{compare_alternative.pdf} 
+\includegraphics[width=13cm]{compare_alternative.pdf} 
 \caption{Comparison between different tensor methods.
@@
 We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and signal $\alpha =3 $ (low), $\alpha=6$ (high).}~\label{fig:compare}
+\vspace{-0.5cm}
 \end{figure}
@@
 {\bf STD} alternates between informative and non-informative modes, whereas {\bf HOLRR} approximates the non-informative modes without alternating.
-The accuracy gain in Figure~\ref{fig:compare} demonstrates the benefit of alternating algorithm -- incorporation of informative modes also improves the estimation in the non-informative modes.
+
+%The accuracy gain in Figure~\ref{fig:compare} demonstrates the benefit of alternating algorithm -- incorporation of informative modes also improves the estimation in the non-informative modes.
 
@@
 
-\section{Data analyses}\label{sec:data}
-We apply our supervised tensor decomposition to two datasets.
+\vspace{-0.5cm}
+\section{Data analysis}\label{sec:data}
+%We apply our supervised tensor decomposition to two datasets.
 The first application studies the brain networks in response to individual attributes (i.e.,\ feature on one mode), and the second application focuses on multi-relational network analysis with dyadic attributes (i.e.,\ features on two modes).
 
-\subsection{Application to human brain connection data}
-The Human Connectome Project (HCP) aims to build a network map that characterizes the anatomical and functional connectivity within healthy human brains~\citep{HCP}. 
+%\subsection{Application to human brain connection data}
+We then apply our method to brain structural connectivity networks from Human Connectome Project (HCP)~\citep{HCP}.
+HCP aims to build a network map that characterizes the anatomical and functional connectivity within healthy human brains.
 %We analyze the structural connectivity patterns among 68 brain regions for 136 individuals from HCP. 
@@
 Each brain network is represented as a 68-by-68 binary matrix, where the entries encode the presence or absence of fiber connections between the 68 brain regions.
-We consider four individual features: gender (65 females vs.\ 71 males), age 22-25 ($n=35$), age 26-30 ($n=58$), and age 31+ ($n=43$).
-The preprocessed dataset is released in our R package \texttt{tensorregress}.
+We consider four individual features: gender (65 females vs.\ 71 males), age 22-25 ($n=35$), age 26-30 ($n=58$), and age 31+ ($n=43$).%The preprocessed dataset is released in our R package \texttt{tensorregress}.
 The goal is to identify the connection edges that are affected by individual features.
@@
 }\label{fig:brain}
-\vspace{1cm}
-\includegraphics[width=16cm]{compare_HCP.pdf}
-\caption{Comparison of estimated feature effects the HCP data using (a) multi-response GLM and (b) supervised tensor decomposition (STD). }\label{fig:s1}
+%\vspace{1cm}
+%\includegraphics[width=16cm]{compare_HCP.pdf}
+%\caption{Comparison of estimated feature effects the HCP data using (a) multi-response GLM and (b) supervised tensor decomposition (STD). }\label{fig:s1}
 \end{figure}
@@
 
-Figure~\ref{fig:s1} compares the estimated coefficients from our method with those from multiple-response GLM approach.
+%Figure~\ref{fig:s1} compares the estimated coefficients from our method with those from multiple-response GLM approach.
 The multiple-response GLM is to regress the brain edges, one at a time, on the individual-level features, and this model is repeatedly fitted for every edge $\in [68]\times [68]$.
 As we can see in the figure, our tensor decomposition method shrinkages the coefficients towards center, thereby implicitly enforcing the sharing between feature effects.
+
+\iffalse
 
@@
 The relation-specific effect pattern showcases the applicability of our method in revealing complex interactions.
+\fi
 
-\section{Discussion and future work}\label{sec:con}
+\vspace{-0.5cm}
+\section{Conclusion}\label{sec:con}
+
+We have developed a supervised tensor decomposition method with side information on multiple modes.
+The empirical results demonstrate the improved interpretability and accuracy over previous approaches.
+Applications to the brain connection data yield conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.
+Further exploring the benefits of supervised tensor decomposition in specialized tasks will be necessary to boost the scientific discoveries.
+\iffalse
 We have developed a supervised tensor decomposition method with side information on multiple modes.
@@
 
+\fi
+
+\spacingset{1.25} 
 \bibliography{tensor_wang}

### 2021-06-21 tensor_regress/JCGS/Response_submission/manuscript/jcgs-tensor.tex
new figure 2

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url}
\usepackage{enumitem}

\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{algorithm}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={black},citecolor={black},urlcolor={black}}
\usepackage{lscape}
\usepackage{afterpage}
%\usepackage{pifont}% http://ctan.org/pkg/pifont
%\newcommand{\cmark}{\ding{51}}%
%\newcommand{\xmark}{\ding{55}}%

\allowdisplaybreaks
%\usepackage{enumerate}
\usepackage{url} % not crucial - just used below for the URL

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\theoremstyle{definition}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}[section]
\newtheorem{pro}{Property}
\newtheorem{cor}{Corollary}[section]

\theoremstyle{definition}
\newtheorem{assumption}{Assumption}
\newtheorem{defn}{Definition}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}
\newtheorem{condition}{Condition}

\newtheorem{innercustomgeneric}{\customgenericname}
\providecommand{\customgenericname}{}
\newcommand{\newcustomtheorem}[2]{%
  \newenvironment{#1}[1]
  {%
   \renewcommand\customgenericname{#2}%
   \renewcommand\theinnercustomgeneric{##1}%
   \innercustomgeneric
  }
  {\endinnercustomgeneric}
}

\newcustomtheorem{customexample}{Example}

\usepackage{appendix}
\usepackage{wrapfig}
\mathtoolsset{showonlyrefs}

\input macros.tex

\usepackage[english]{babel}

\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}

\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\def\ci{\perp\!\!\!\perp}

\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}
\usepackage{booktabs}
\newcommand\doubleRule{\toprule\toprule}
\allowdisplaybreaks

\usepackage{xr}
% this is for overleaf cross reference
\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}
\myexternaldocument{supp}

\begin{document}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Supervised tensor decomposition with features on multiple modes}
  \author{Jiaxin Hu, Chanwoo Lee, and Miaoyan Wang\thanks{The authors gratefully acknowledge NSF grant DMS-1915978 and funding from the Wisconsin Alumni Research Foundation. }\hspace{.2cm}\\
    Department of Statistics, University of Wisconsin-Madison}
  \maketitle

} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Supervised tensor decomposition with features on multiple modes}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Higher-order tensors have received increased attention across science and engineering. While most tensor decomposition methods are developed for a single tensor observation, scientific studies often collect side information, in the form of node features and interactions thereof, together with the tensor data. Such data problems are common in neuroimaging, network analysis, and spatial-temporal modeling. Identifying the relationship between a high-dimensional tensor and side information is important yet challenging. Here, we develop a tensor decomposition method that incorporates multiple feature matrices as side information. Unlike unsupervised tensor decomposition, our supervised decomposition captures the effective dimension reduction of the data tensor confined to feature space of interest. An efficient alternating optimization algorithm with provable spectral initialization is further developed. Our proposal handles a broad range of data types, including continuous, count, and binary observations. We apply the method to diffusion tensor imaging data from human connectome project and multi-relational political network data. We identify the key global connectivity pattern and pinpoint the local regions that are associated with available features. The package and data used are available at~\url{https://CRAN.R-project.org/package=tensorregress}. Supplementary materials for this article are available online .

\end{abstract}

\noindent%
{\it Keywords:}  Tensor data analysis, Supervised dimension reduction, Exponential family distribution, Generalized multilinear model, Alternating optimization
\vfill

\newpage
\spacingset{1.5} % DON'T change the spacing!
\section{Introduction}
\label{sec:intro}
Multi-dimensional arrays, known as tensors, are often collected with side information on multiple modes in modern scientific and engineering studies. A popular example is in neuroimaging~\citep{zhou2013tensor}. The brain connectivity networks are collected from a sample of individuals, accompanied by individual characteristics such as age, gender, and diseases status (see Figure~\ref{fig:intro1}a). Another example is in network analysis~\citep{pmlr-v108-berthet20a,hoff2005bilinear}. A typical social network consists of nodes that represent people and edges that represent  friendships. Side information such as people’s demographic information and friendship types are often available. In both examples, we are interested in identifying the variation in the data tensor (e.g., brain connectivities, social community patterns) that is affected by available features. These seemingly different scenarios pose a common yet challenging problem for tensor data modeling.

In addition to the aforementioned challenges, many tensor datasets consist of non-Gaussian measurements. Examples include the political interaction dataset \citep{nickel2011three} which measures action counts between countries under various relations, and the brain connectivity network dataset \citep{zhou2013tensor,wang2020learning} which is a collection of binary adjacency matrices. Classical tensor decomposition methods are based on minimizing the Frobenius norm of deviation, leading to suboptimal predictions for binary- or count-valued response variables. A number of supervised tensor methods have been proposed \citep{lock2018supervised,li2017parsimonious,sun2017store,hao2019sparse,raskutti2019convex} to address the tensor regression problem in various forms, such as scalar-to-tensor regression and tensor-response regression. These methods often assume Gaussian distribution for the tensor entries, or impose random designs for the feature matrices, both of which are less suitable for applications of our interest. The gap between theory and practice means a great opportunity to model paradigms and better capture the complexity in tensor data.

\begin{figure*}[t]
\begin{center}
\includegraphics[width=16cm]{demo.pdf}
\end{center}
\caption{Examples of supervised tensor decomposition with side information. (a) Network population model. (b) Spatio-temporal growth model.} \label{fig:intro1}
\vspace{-.2cm}
\end{figure*}

We present a general model and associated method for decomposing a data tensor whose entries are from exponential family with side information. We formulate the learning task as a structured regression problem, with tensor observation serving as the response, and the multiple side information as features. Figure~\ref{fig:intro1}b illustrates our model in the special case of order-3 tensors. A low-rank structure is imposed to the conditional mean of tensor observation, where unlike classical decomposition, the tensor factors $\mX_k\mM_k\in\mathbb{R}^{d_k\times r_k}$ belong to the space spanned by features $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,2,3$. The unknown matrices $\mM_k\in\mathbb{R}^{p_k\times r_k}$ (referred to as ``dimension reduction matrices'') link the conditional mean to the feature spaces, thereby allowing the identification of variations in the tensor data attributable to the side information.

Our proposal blends the modeling power of generalized linear model (GLM) and the exploratory capability of tensor dimension reduction in order to take the best out of both worlds. We leverage GLM to allow heteroscedacity due to the mean-variance relationship in the non-Gaussian data. This flexibility is important in practice. Furthermore, our low-rank model on the (transformed) conditional mean tensor effectively mitigates the curse of high dimensionality. In classical GLM, the sample size and feature dimension are well defined; however, in the tensor data analysis, we observe only one realization of an order-$K$ tensor and up to $K$ feature matrices. Both the number of tensor entries and feature dimension grow exponentially in $K$. Dimension reduction is therefore crucial for prediction and interpretability. We establish the statistical and algorithmic convergences of our estimator, and we quantify the gain in accuracy through simulations and case studies.

Our work is closely related to but also clearly distinctive from several lines of previous work. {The first line is a class of}\label{change:firstline} \textit{unsupervised} tensor decomposition such as classical Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor} and generalized decomposition for non-Gaussian data \citep{chi2012tensors, tarzanagh2019regularized,hong2020generalized,  li2020generalized}. {\color{blue}Regardless of the implementation, the unsupervised methods aim to find the best low-rank representation of a data tensor alone. In contrast, our model is a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and multiple features. The low-rank factorization is determined jointly by the tensor data and feature matrices in our model.}

The second line of work studies the tensor-to-tensor regression. {\color{blue}This category is further divided into three scenarios, depending on whether tensor is treated as predictors~\citep{zhou2013tensor,raskutti2019convex,han2020optimal}, as responses~\citep{li2017parsimonious,sun2017store,zhang2018network,lock2018supervised,luo2018leveraging}, or both~\citep{lock2018tensor,gahrooei2020multiple}. As we show in Section~\ref{sec:connection}, our supervised tensor decomposition falls into this general category, and we provide a \emph{provable} solution in new settings that have broader practical significance. Earlier work in this vein~\citep{lock2018tensor,lock2018supervised,gahrooei2020multiple,li2020generalized} focuses on algorithm development, but not on the statistical accuracy. \cite{li2017parsimonious} introduces an envelope-based approach to identify sufficient dimension reduction~\citep{adragni2009sufficient}, but its theory is restricted to Gaussian data with one-sided feature matrix only. \cite{raskutti2019convex} establishes the statistical accuracy for convex relaxed maximum likelihood estimator (MLE) of tensor regression. However, convex relaxation for tensor optimizations suffers from computational intractability and statistical sub-optimality. Recent work has demonstrated the success of non-convex approaches in various tensor problems~\citep{sun2017store,zhang2018network,raskutti2019convex,han2020optimal}; we go step further by allowing multiple feature matrices with either fixed or random designs. In Sections~\ref{subsec:statprob}, we show that incorporating multiple feature matrices substantially improves the statistical accuracy. We provide a detailed comparison in Section~\ref{sec:connection}; see Table~\ref{table:comp_table}.}

The third line of work uses side information for various tensor learning tasks, such as for completion~\citep{song2019tensor} and for recommendation system~\citep{farias2019learning}. These methods also study tensors with side information, but they take data-mining approaches to penalize predictions that are distant from side information. One important difference is that their goal is prediction but not parameter estimation. The effects of features and their interactions are not estimated in these data-driven approaches. In contrast, our goal is interpretable prediction, and we estimate the low-rank decomposition using a model-based approach.
The model-based approaches benefits the interpretability in prediction. In this regards, our method opens up new opportunities for tensor data analysis in a wider range of applications.

The remainder of the paper is organized as follows. Section~\ref{sec:pre} introduces tensor preliminaries. Section~\ref{sec:model} presents the main model and three motivating examples for supervised tensor decomposition. We describe the likelihood estimation and alternating optimization algorithm with theoretical guarantees in Section~\ref{sec:est}. Connection with related work is provided in Section~\ref{sec:connection}. In Section~\ref{sec:simulation}, we present numerical experiments and assess the performance in comparison to alternative methods. In Section~\ref{sec:data}, we apply the method to diffusion tensor imaging data from human connectome project and multi-relational social network data. We conclude in Section~\ref{sec:con} with discussions about our findings and avenues of future work. All proofs are deferred to Supplementary Notes.

\section{Preliminaries}\label{sec:pre}
We introduce the basic tensor properties used in the paper. We use lower-case letters (e.g.,\ $a,b,c$) for scalars and vectors, upper-case boldface letters (e.g.,\ $\mA, \mB, \mC$) for matrices, and calligraphy letters (e.g.,\ $\tA, \tB, \tC$) for tensors of order three or greater. We use $\mI$ to denote the identity matrix whose dimension may vary from line by line given the contexts. Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in \mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor, where $K$ is the number of modes and also called the order. The multilinear multiplication of a tensor $\tY\in\mathbb{R}^{d_1\times \cdots\times d_K}$ by matrices $\mX_k=\entry{x_{i_k,j_k}^{(k)}}\in\mathbb{R}^{p_k\times d_k}$ is defined as 
\[
\tY \times_1\mX_1\times \cdots \times_K \mX_K=\entry{\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}x_{j_1,i_1}^{(1)}\cdots x_{j_K,i_K}^{(K)}},
\]
which results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor. For ease of presentation, we use the shorthand $\tY\times\{\mX_1,\ldots,\mX_K\}$ to denote the tensor-by-matrix product. For any two tensors $\tY=\entry{y_{i_1,\ldots,i_K}}$, $\tY'=\entry{y'_{i_1,\ldots,i_K}}$ of identical order and dimensions, their inner product is defined as 
\[
\langle \tY,\ \tY'\rangle =\sum_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}y'_{i_1,\ldots,i_K}.
\]
The tensor Frobenius norm and maximum norm are defined as
\[
\FnormSize{}{\tY}=\langle \tY,\ \tY \rangle^{1/2},\quad \text{and}\quad \mnormSize{}{\tY}=\max_{i_1,\ldots,i_K}y_{i_1,\ldots,i_K}.
\]
When $a$ is a vector, we use $\vnormSize{}{a}=\langle a,a \rangle^{1/2}$ to denote the vector 2-norm. We use $[d]$ to denote the $d$-set $[d]=\{1,\ldots,d\}$, and use $\mathbb{O}(d,r)$ to denote the collection of all $d$-by-$r$ matrices with orthonormal columns; i.e., $\mathbb{O}(d,r)=\{\mP\in\mathbb{R}^{d\times r}\colon \mP^T\mP = \mI\}$.

A higher-order tensor can be reshaped into a lower-order object. We use $\textup{vec}(\cdot)$ to denote the operation that reshapes the tensor into a vector, and $\text{Unfold}_k(\cdot)$ to denote the unfolding operation that reshapes the tensor along mode $k$ into a matrix of size $d_k$-by-$\prod_{i\neq k}d_i$. We use $\text{rank}(\tY)=\mr$ to denote the multilinear rank of an order-$K$ tensor $\tY$, where $\mr=(r_1,\ldots,r_K)$ is a length-$K$ vector and $r_k$ is the rank of matrix $\textup{Unfold}_k(\tY)$ for $k\in[K]$. For ease of notation, we allow the basic arithmetic operators (e.g.,\ $+, -, \geq $) and univariate functions $f\colon \mathbb{R}\to \mathbb{R}$ to be applied to tensors in an element-wise manner. For two positive sequences $\{a_n\}$ and $\{b_n\}$, we use $a_n\lesssim b_n$ or $a_n = \tO(b_n)$ to denote the fact that $a_n \leq Cb_n$ for some constant $C >0$.

\section{Motivation and model}\label{sec:model}
\subsection{General framework for tensor decomposition}\label{sec:mainmodel}
We begin with a general framework for supervised tensor decomposition and then discuss its implication in three concrete examples. Let $\tY=\entry{y_{i_1,\ldots,i_K}}\in\mathbb{R}^{d_1\times \cdots\times d_K}$ denote an order-$K$ data tensor. Suppose the side information is available on each of the $K$ modes. Let $\mX_k=\entry{x_{ij}}\in\mathbb{R}^{d_k\times p_k}$ denote the feature matrix on the mode $k\in[K]$, where $x_{ij}$ denotes the $j$-th feature value for the $i$-th tensor entity, for $(i,j)\in[d_k]\times[p_k]$.

We propose a multilinear conditional mean model between the data tensor and feature matrices. Assume that, conditional on the features $\mX_k$, the entries of tensor $\tY$ are independent realizations from an exponential family distribution. Further, the conditional mean tensor admits the rank-$\mr$ model with $\mr=(r_1,\ldots,r_K)$,
\begin{align}\label{eq:decomp}
\mathbb{E}(\tY|\mX_1,\ldots,\mX_K) &= f\left(\tC\times\{\mX_1\mM_1, \ldots, \mX_K\mM_K\}\right),\notag\\
\text{with} \ &\ \mM_k^T\mM_k = \mI_{r_k},\ \mM_k\in\mathbb{R}^{p_k\times r_k}\quad \text{for all }k=1,\ldots,K,
\end{align}
where $\tC \in \mathbb{R}^{r_1\times \cdots \times r_K}$ is an unknown full-rank core tensor, $\mM_k \in \mathbb{R}^{p_k\times r_k}$ are unknown factor matrices for all $k\in [K]$, $f(\cdot)$ is a known link function whose form depending on the data type of $\tY$, and $\times$ denotes the tensor-by-matrix product. The choice of link function is based on the assumed distribution family of tensor entries. Common choices of link functions include identity link for Gaussian distribution, logistic link for Bernoulli distribution, and exponential link for Poisson distribution. In general, dispersion parameters can also be included in the model. Because our main focus is the tensor decomposition under the mean model, we suppress the dispersion parameter in this section for ease of presentation.

Figure~\ref{fig:intro1}b provides a schematic illustration of our model. The features $\mX_k$ affect the distribution of tensor entries in $\tY$ through the reduced features $\mX_k\mM_k$, which are $r_k$ linear combinations of features on mode $k$. We call $\mM_k$ the ``dimension reduction matrix'' or ``tensor factors.'' The core tensor $\tC$ collects the interaction effects between reduced features across $K$ modes. We call $\tB=\tC\times\{\mM_1,\ldots,\mM_K\}$ the coefficient tensor, and $\Theta=\tB\times\{\mX_1,\cdots,\mX_K\}$ the linear predictor. By the definition of multilinear rank, the model~\eqref{eq:decomp} implies the linear predictor $\Theta$ and coefficient tensor $\tB$ are of rank-$\mr$. The conditional mean tensor $\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)$ is however often high rank, due to the nonlinearity of the link function~\citep{lee2021beyond}.

Our goal is to estimate the low-rank tensor $\tB$, or equivalently, the core tensor and factors $(\tC, \mM_1,\ldots,\mM_K)$, from our model~\eqref{eq:decomp}. {\color{blue}We make several remarks about model identifiability. First, the identifiability of $\tB$ requires the feature matrices $\mX_k$ are of full column rank with $p_k\leq d_k$. We impose this rank non-deficiency assumption to $\mX_k$; this is a mild condition common in literature~\citep{lock2018supervised, li2017parsimonious, li2020generalized}. In the presence of rank deficiency, we recommend to remove redundant features from $\mX_k$ before applying our method. Second, the decomposition $\tB = \tC\times\{\mM_1,\ldots,\mM_K\}$ are non-unique, as in standard tensor decomposition~\citep{kolda2009tensor}. For any invertible matrices $\mO_k\in\mathbb{R}^{r_k\times r_k}$, $\tB=\tC\times\{\mM_1,\ldots,\mM_K\}=\tC'\times\{\mM_1\mO_1,\ldots,\mM_K\mO_K\}$ are two equivalent parameterizations with $\tC'=\tC\times\{\mO^{-1}_1,\ldots,\mO^{-1}_K\}$ . To resolve this ambiguity, we impose orthonormality to $\mM_k\in\mathbb{O}(p_k,r_k)$ and assess the estimation error of $\mM_k$ using angle distance. The angle distance is invariant to orthogonal rotations due to its geometric definition. See Section~\ref{subsec:statprob} for more details. The orthonormality of $\mM_k$ is imposed purely for technical convenience. This normalization incurs no impacts in our statistical inference, but may help with numerical stability in empirical optimization~\citep{de2000multilinear, kolda2009tensor}.} Finally, the problem size is quantified by $p_k$ and $d_k$, where $p_k$ specifies the number of features and $d_k$ the number of samples at mode $k\in[K]$. Our theory treats the rank $r_k$ as known and fixed, whereas both $p_k$ and $d_k$ are allowed to increase. The adaptation to unknown rank in practice will be addressed in Section~\ref{sec:tuning}.

\subsection{Three examples}\label{sec:motivation}
We give three seemingly different examples that can all be formulated as our supervised tensor decomposition model~\eqref{eq:decomp}.
\begin{example}[Spatio-temporal growth model]
The growth curve model~\citep{gabriel1998generalised,srivastava2008models} was originally proposed as an example of bilinear model for matrix data, and we adopt its higher-order extension here. Let $\tY=\entry{y_{ijk}}\in\mathbb{R}^{d \times m\times n}$ denote the pH measurements of $d$ lakes at $m$ levels of depth and for $n$ time points. Suppose the sampled lakes belong to $q$ types, with $p$ lakes in each type. Let $\{\ell_j\}_{j\in[m]}$ denote the sampled depth levels and $\{t_k\}_{k\in[n]}$ the time points. Assume that the expected pH trend in depth is a polynomial of order at most $r$ and that the expected trend in time is a polynomial of order $s$. Then, the conditional mean model for the spatio-temporal growth can be represented as
\begin{equation}\label{eq:time}
\mathbb{E}(\tY|\mX_1,\mX_2,\mX_3)=\tC\times\{\mX_1\mM_1,\ \mX_2\mM_2,\ \mX_3\mM_3\},
\end{equation}
where $\mX_1=\text{blockdiag}\{\mathbf{1}_p,\ldots,\mathbf{1}_p\}\in \{0,1\}^{d\times q}$ is the design matrix for lake types, and
\[
\mX_2=
\begin{pmatrix}
1 & \ell_1&\cdots &\ell^{r}_1\\
1 & \ell_2&\cdots &\ell^{r}_2\\
\vdots &\vdots&\ddots&\vdots\\
1&\ell_{m}&\cdots&\ell^{r}_{m}
\end{pmatrix},\quad
\mX_3=
\begin{pmatrix}
1 & t_1&\cdots &t^{s}_1\\
1 & t_2&\cdots &t^{s}_2\\
\vdots &\vdots&\ddots&\vdots\\
1&t_{n}&\cdots&t^{s}_{n}
\end{pmatrix}
\]
are the design matrices for spatial and temporal effects, respectively, $\tC\in\mathbb{R}^{r_1\times r_2\times r_3}$ is the unknown core tensor, and $\mM_k$ are unknown dimension reduction matrices on each mode. The factors $\mX_k\mM_k$ are reduced features in the mean model~\eqref{eq:time}. The spatial-temporal model is a special case of our supervised tensor decomposition model~\eqref{eq:decomp}, with features available on each of the three modes.
\end{example}

\begin{example}[Network population model]\label{example:brain}
Network response model~\citep{rabusseau2016low} is recently developed for neuroimaging analysis. The goal is to study the relationship between brain network connectivity pattern and features of individuals. Suppose we have a sample of $n$ observations, $\{(\mY_i, \mx_i)\colon i=1,\ldots,n\}$, where for each individual $i\in[n]$, $\mY_i\in\{0,1\}^{d\times d}$ is the undirected adjacency matrix whose entries indicate presences/absences of connectivities between $d$ brain nodes, and $\mx_i\in\mathbb{R}^p$ is the individual's feature such as age, gender, cognition score, etc. The network-response model  has the conditional mean
\begin{equation}\label{eq:network}
\textup{logit}(\mathbb{E}(\mY_i|\mx_i))=\tB\times_3\mx_i, \quad \text{for }i=1,\ldots,n,
\end{equation}
where $\tB\in \mathbb{R}^{d\times d\times p}$ is a rank-$(r_1,r_1,r_2)$ coefficient tensor, and $\tB$ is assumed to be symmetric in the first two modes.

The model~\eqref{eq:network} is a special case of our supervised tensor decomposition, with feature matrix on the last mode of the tensor. Specifically, we stack the network observations $\{\mY_i\}$ together and obtain an order-3 response tensor $\tY\in\{0,1\}^{d\times d\times n}$. Define a feature matrix $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$. Then, the model~\eqref{eq:network} has the equivalent representation of supervised tensor decomposition,
\[
\textup{logit}(\mathbb{E}(\tY|\mX))=\tC\times\{\mM,\ \mM,\ \mX\mM'\},
\]
where $\tC\in\mathbb{R}^{r_1\times r_1\times r_2}$ is the core tensor, $\mM\in\mathbb{R}^{d\times r_1}$ is the dimension reduction matrix on the first two modes, and $\mM'\in\mathbb{R}^{p\times r_2}$ is for the last mode.  \end{example}

 \begin{example}[Dyadic data with node attributes] Dyadic dataset consists of measurements on pairs of objects. Common examples include graphs and networks. Let $\tG=(V,E)$ denote a graph, where $V=[d]$ is the node set of the graph, and $E\subset V\times V$ is the edge set. Suppose that we also observe feature vector $\mx_i\in\mathbb{R}^p$ associated to each node $i\in V$. A probabilistic model on the graph $\tG=(V,E)$ can be described by the following matrix regression. The edge connects the two vertices $i$ and $j$ independently of other pairs, and the probability of connection is modeled as
\begin{equation}\label{eq:edge}
 \textup{logit}\left(\mathbb{P}\left((i,j)\in E\right)\right)=\mx^T_i\mB\mx_j=\langle \mB,\ \mx^T_i\mx_j\rangle,
 \end{equation}
 where $\mB\in\mathbb{R}^{p\times p}$ is a symmetric rank-$r$ matrix. The low-rankness in $\mB$ has demonstrated its success in modeling transitivity, balance, and communities in  networks~\citep{hoff2005bilinear}. We show that our supervised tensor decompostion~\eqref{eq:decomp} also incorporates the graph model as a special case. Let $\tY=\entry{y_{ij}}$ be a binary matrix where $y_{ij}=\mathds{1}_{(i,j)\in E}$. Define $\mX=[\mx_1,\ldots,\mx_n]^T\in\mathbb{R}^{n\times p}$. Then, the graph model~\eqref{eq:edge} can be expressed as
 \[
 \textup{logit}(\mathbb{E}(\mY|\mX))=\mC\times\{\mX\mM,\ \mX\mM\}, 
  \]
  where $\mC\in\mathbb{R}^{r\times r}, \mM\in\mathbb{R}^{p\times r}$ are from the singular value decomposition of $\mB=\mM\mC\mM^T$. 
  \end{example}

In the above three examples and many other studies, researchers are interested in uncovering the variation in the data tensor that can be explained by features. Our supervised tensor decomposition~\eqref{eq:decomp} allows arbitrary numbers of feature matrices. When certain mode $k$ has no side information, we set $\mX_k=\mI$ in the model~\eqref{eq:decomp}. In particular, our model~\eqref{eq:decomp} reduces to classical unsupervised tensor decomposition~\citep{de2000multilinear,hong2020generalized} when no side information is available; i.e., $\mX_k=\mI$ for all $k\in[K]$.

\section{Estimation}\label{sec:est}

\subsection{Rank-constrained MLE}\label{sec:rankM}
We develop a likelihood-based procedure to estimate $\tC$ and $\mM_k$ in~\eqref{eq:decomp}. We adopt the exponential family as a flexible framework for different data types. In a classical generalized linear model with a scalar response $y$ and feature $\mx$, the density is expressed as
\[
p(y|\mx, \boldsymbol{\beta})=c(y,\phi)\exp\left(\frac{y\theta- b(\theta)}{\phi}\right)\ \text{with}\ \theta=\boldsymbol{\beta}^T\mx,
\]
where $b(\cdot)$ is a known function, $\theta$ is the linear predictor, $\phi>0$ is the dispersion parameter, and $c(\cdot)$ is a known normalizing function. The choice of link functions depends on the data types and on the observation domain of $y$, denoted $\mathbb{Y}$. For example, the observation domain is $\mathbb{Y}=\mathbb{R}$ for continuous data, $\mathbb{Y}=\mathbb{N}$ for count data, and  $\mathbb{Y}=\{0,1\}$ for binary data. The canonical link function $f$ is chosen to be $f(\cdot)=b'(\cdot)$, the first-order derivative of $b(\cdot)$. Table~\ref{table:link} summarizes the canonical link functions for common types of distributions.

\begin{table}[htb]
\centering
\begin{tabular}{c|cccc}
Data type &Gaussian & Poisson& Bernoulli &  \\
\hline
Domain $\mathbb{Y}$& $\mathbb{R}$&$\mathbb{N}$&$\{0,1\}$  \\
 $b(\theta)$&$\theta^2/2$& $\exp(\theta)$&$\log (1+\exp(\theta))$  \\
 link $f(\theta)$&$\theta$&$\exp(\theta)$&$(1+\exp(-\theta))^{-1}$ \\
\end{tabular}
\caption{Canonical links for common distributions.}\label{table:link}
\end{table}

In our context, we model the entries in data tensor $\tY$, conditional on linear predictor $\Theta$, as independent draws from an exponential family. Ignoring constants that do not depend on $\Theta$, the quasi log-likelihood of~\eqref{eq:decomp} is equal to Bregman distance between $\tY$ and $b'(\Theta)$:
\begin{align}\label{eq:loglikelihood}
\tL_{\tY}(\tC,\mM_1,\ldots,\mM_K)&=\langle \tY, \Theta \rangle - \sum_{i_1,\ldots,i_K} b(\theta_{i_1,\ldots,i_K}),\notag \\
 \text{where}\quad \Theta&=\tC\times\{\mX_1\mM_1,\ldots,\mX_K\mM_K\}.
\end{align}
 We propose the constrained maximum quasi-likelihood estimate (MLE),
\begin{align} \label{eq:MLE} 
(\hat \tC_{\text{MLE}}, \hat \mM_{1,\text{MLE}},\ldots,\hat \mM_{K,\text{MLE}}) =\argmax_{(\tC,\mM_1,\ldots,\mM_K)\in \tP(\mr)} \ \tL_{\tY}(\tC,\mM_1,\ldots,\mM_K),
\end{align}
where the parameter space $\tP(\mr)$ is defined by
\begin{equation}\label{eq:p}
\tP(\mr)=\left\{(\tC, \mM_1,\ldots,\mM_K) \ \Big| \ \mM_k\in\mathbb{O}(p_k,r_k)\ \text{for all }k\in[K],\  \mnormSize{}{\Theta}\leq \alpha \right\},
\end{equation}
with a large constant $\alpha>0$. Recall that $\tB=\tC\times\{\mX_1,\ldots,\mM_K\}$ by definition. Correspondingly, we estimate the coefficient tensor $\tB$ by
\[
\hat \tB_{\text{MLE}}=\hat \tC_{\text{MLE}}\times\{\hat \mM_{1,\text{MLE}},\ldots,\hat \mM_{K,\text{MLE}}\}.
\]

The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to ensures the existence (boundedness) of MLE. 
{\color{blue}The condition precludes the ill-defined MLE when the optimizer of~\eqref{eq:MLE} diverges to $\pm \infty$; this phenomenon may happen in logistic regression when the Bernoulli responses $\{0,1\}$ are perfectly separable by covariates~\citep{wang2020learning}. For Gaussian models, no maximum norm constraint is needed. In Section~\ref{subsec:statprob}, we show that setting $\alpha$ to an extremely large constant does not compromise the statistical rate in quantities of interest. In practice, the unbounded search is often indistinguishable from the bounded search, since the boundary constraint $\onorm{\Theta}_{\infty} \leq \alpha$ would likely never be active. Similar techniques are commonly used in high-dimensional non-Gaussian problems~\citep{wang2020learning,han2020optimal}.}

The optimization~\eqref{eq:MLE} is a non-convex problem with possibly local optimizers. We propose an alternating optimization algorithm to {\it approximately} solve~\eqref{eq:MLE}. 
The decision variables in the objective function~\eqref{eq:MLE} consist of $K+1$ blocks of variables, one for the core tensor $\tC$ and $K$ for the factor matrices $\mM_k$. We notice that, if any $K$ out of the $K+1$ blocks of variables are known, then the optimization reduces to a simple GLM with respect to the last block of variables. This observation leads to an iterative updating scheme for one block at a time while keeping others fixed. Given an initialization $(\hat\tC^{(0)},\hat\mM_1^{(0)}, \ldots, \hat\mM_K^{(0)})$ to be described in the next paragraph, the $t$-th iterate from the algorithm is denoted $(\hat\tC^{(t)},  \hat\mM^{(t)}_1,\ldots,\hat\mM^{(t)}_K)$ for $t=1, 2,3, \ldots$ The iteration scheme is detailed in Algorithm~\ref{alg:B}. 
\begin{algorithm}[!h]
\caption{Supervised Tensor Decomposition with Side Information}\label{alg:B}
\begin{algorithmic}[1]
\INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, feature matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$ for $k=1,\ldots,K$, target rank $\mr=(r_1,\ldots,r_K)$, link function $f$, initialization $(\hat\tC^{(0)},\hat \mM^{(0)}_1,\ldots,\hat\mM^{(0)}_K)$.
\For {$t=1,2,3,\ldots$}
\For { $k=1$ to $K$}
\State Obtain the factor matrix $\hat\mM^{(t)}_k\in\mathbb{R}^{p_k\times r_k}$ by a GLM with link function $f$.
\State Perform QR factorization $\hat\mM^{(t)}_k=\mQ_k\mR_k$, where $\mQ_k\in\mathbb{O}(p_k, r_k)$.
\State Update $\hat\mM^{(t)}_k\leftarrow \mQ_k$ and core tensor $\hat\tC^{(t)}\leftarrow \hat\tC^{(t)}\times_k \mR_k$.
\EndFor
\State Update the core tensor $\tC$ by solving a GLM with $\textup{vec}(\tY)$ as response, $\otimes_{k=1}^K[ \mX_k\mM_k]$ as features, and $f$ as link function. Here $\otimes$ denotes the Kronecker product of matrices. 
\EndFor
\OUTPUT factor estimate $(\hat\tC^{(t)}, \hat\mM^{(t)}_1,\ldots,\hat\mM^{(t)}_K)$ from the $t$-th iterate, and coefficient tensor estimate $\hat\tB^{(t)}=\hat\tC^{(t)}\times\{\hat\mM^{(t)}_1,\ldots, \hat\mM^{(t)}_K\}$.
\end{algorithmic}
\end{algorithm}

{\color{blue}We provide two initialization schemes, one with QR-adjusted spectral initialization (warm initialization), and the other with random initialization (cold initialization). The warm initialization is an extension of unsupervised spectral initialization~\citep{zhang2018tensor} to supervised setting with multiple feature matrices.} Specifically, we project normalized data tensor $\bar \tY$ to the normalized multilinear feature space and obtain an unconstrained coefficient tensor $\hat\tB^{(0)}$. We perform a rank-$\mr$ higher-order SVD (HOSVD) on $\bar \tB$, which yields the rank-constrained $\hat\tB^{(0)}$. The desired initialization is obtained by re-normalizing $\hat\tB^{(0)}$ back to the original scales of features. The initialization scheme is described in Algorithm~\ref{alg:A}.

{\color{blue}The warm initialization enjoys provable accuracy guarantees at a cost of extra technical assumptions (see Section~\ref{subsec:statprob}). The cold initialization, on the other hand, shows robust in practice but its theoretical guarantee remains an open challenge~\citep{luo2021low}. We incorporate both options in our software package to provide flexibility to practitioners.}

\begin{algorithm}[!h]
\caption{QR-adjusted spectral initialization}\label{alg:A}
\begin{algorithmic}[1]
\INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, feature matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$, Tucker rank $\mr$. 
\State Normalize date tensor $\bar \tY\leftarrow \tY$ for Gaussian model, $\bar \tY\leftarrow 2\tY-1$ for Bernoulli model, and $\bar \tY\leftarrow \log(\tY+0.5)$ for Poisson model. 
\State Normalize feature matrices via QR factorization $\mX_k=\mQ_k\mR_k$ for all $k\in[K]$.
\State Obtain $\bar \tB \leftarrow \bar \tY\times\{\mQ^T_1,\ldots,\mQ^T_K\}$ by projecting $\bar \tY$ to the multilinear feature space. 
\State Obtain $\hat\tB^{(0)}\leftarrow \textup{HOSVD}(\bar \tB,\mr)$.
\State Normalize representation $\{\hat\tC^{(0)}, \hat\mM^{(0)}_1,\ldots,\hat\mM^{(0)}_K\}$ such that $\hat\tC^{(0)}\times\{ \hat\mM^{(0)}_1,\ldots,\hat\mM^{(1)}_K\} = \hat\tB^{(0)}\times\{ \mR^{-1}_1,\ldots,\mR^{-1}_K\}$ and $\hat\mM^{(0)}_k\in\mathbb{O}(p,r)$ for all $k\in[K]$.
\OUTPUT Core tensor $\hat\tC^{(0)}$ and factors $\hat\mM^{(0)}_k$ for all $k\in[K]$.
\end{algorithmic}
\end{algorithm}

\subsection{Statistical accuracy}\label{subsec:statprob}
{\color{blue}This section presents the accuracy guarantees for both global and local optimizers of~\eqref{eq:MLE}. We first provide the statistical accuracy for the global MLE~\eqref{eq:MLE}. Then, we provide the convergence rate for the local optimizer from Algorithm~\ref{alg:B} with warm initialization. The rate reveals an interesting interplay between statistical and computational efficiency. We show that a polynomial number of iterations suffices to reach the desired accuracy under certain assumptions.} The empirical performance for cold initialization is also investigated.

For cleaner exposition, we present the results for balanced setting in this section, i.e., $p_1=\cdots=p_K=p$, $r_1=\cdots=r_K=r$, and $d_1=\cdots=d_K=d$. The general setting follows exactly the same framework and incurs only notational complexity. We are particularly interested in the high-dimensional regime in which both $d$ and $p$ grows while $p\leq d$. {\color{blue}The requirement $p\leq d$ is necessary to ensure rank non-deficiency of  feature matrices $\mX_k$.} The classical MLE theory is not directly applicable, because the number of unknown parameters grows with the size of data tensor. We leverage the recent development in random tensor theory and high-dimensional statistics to establish the error bounds of the estimation.

\begin{assumption}\label{ass}We make the following assumptions:
\begin{enumerate}[noitemsep,topsep=0pt]
\item [A1.] There exist two positive constants $c_1, c_2>0$ such that $c_1\leq \sigma_{\min}(\mX_k)\leq  \sigma_{\max}(\mX_k)\leq c_2$ for all $k\in[K]$. Here $\sigma_{\min}(\cdot)$ and $\sigma_{\max}(\cdot)$ denote the smallest and largest matrix singular values.
\item [A1'.] The feature matrices $\mX_k$ are Gaussian designs with i.i.d.\ $N(0,1)$ entries.
\item [A2.] There exist two positive constants $L, U>0$, such that $\min_{|\theta|\leq \alpha}b''(\theta)\geq \phi L$ and $\sup_{\theta\in \mathbb{R}}b''(\theta)\leq \phi U$. Here, $\alpha$ is the upper bound of the linear predictor in~\eqref{eq:MLE}, and $b''(\cdot)$ denotes the second-order derivative.
\end{enumerate}
\end{assumption}
The assumptions are fairly mild. Assumptions A1 and A1' consider two separate scenarios about feature matrices. Assumption A1 is applicable when feature matrix is asymptotically non-singular and has bounded spectral norm, whereas Assumption A1' imposes the commonly-used Gaussian design~\citep{raskutti2019convex}.  The Assumption 2 is essentially imposed to the response variance because of the identity $\text{Var}(y|\theta)=\phi b''(\theta)$~\citep{mccullagh1989generalized}. The lower bound  ensures the non-degeneracy of the variance in the feasible domain of $\theta$, whereas the upper bound ensures the finiteness of the variance in the entire family. In fact, except for Poisson responses, most members in the exponential family, e.g., Gaussian, Bernoulli, and binomial responses, satisfy this condition.

\subsubsection{Statistical accuracy for global optimizers}\label{sec:global}
We need some extra notation to state the results in full generality. Recall that the factor matrices $\mM_k$ are identifiable only up to orthogonal rotations. Therefore, we choose to use angle distance to assess the estimation accuracy of $\mM_k$. For any two column-orthonormal matrices $\mA,\mB\in\mathbb{O}(d,r)$ of same dimension, the angle distance is defined as
\[
\sin \Theta(\mA,\mB)=\max\left\{ \frac{\langle \mx, \my\rangle}{\vnormSize{}{\mx}\vnormSize{}{\my}}\colon \ \mx\in \textup{Span}(\mA),\ \my\in \textup{Span}(\mB^{\perp})\right\},
\]
where $\text{Span}(\cdot)$ represents the column space of the matrix. We use the superscript ``true'' to denote the true parameters from generic decision variables in optimization. For instance, $\trueB$ denotes the true coefficient tensor, whereas $\tB$ denotes a decision variable in~\eqref{eq:loglikelihood}.

Define the signal level $\lambda$ as the minimal singular value of the unfolded matrices obtained from $\trueB$,  
\[
\lambda=\min_{k\in[K]}\sigma_{\min}(\text{Unfold}_k(\tB_{\text{true}})).
\]
Intuitively, $\lambda$ quantifies the level of rank non-degeneracy for the true coefficient tensor $\trueB$.

\begin{thm}[Statistical rate for global optimizers]\label{thm:MLE}Consider generalized tensor models with multiple feature matrices. Under Assumptions A1 and A2 with scaled feature matrices $\bar \mX_k= \sqrt{d}\mX_k$, or Assumptions A1' and A2 with original feature matrices, we have
\begin{equation}\label{eq:bound}
\max_{k\in[K]}\sin^2\Theta(\trueM, \hat \mM_{k,\text{MLE}})\lesssim {\phi (r^K+Kpr)\over \lambda^2 d^K}, \quad \FnormSize{}{\tB_{\text{true}}-\hat \tB_{\text{MLE}} }^2\lesssim {\phi (r^K+Kpr) \over d^K},
\end{equation}
with probability at least $1-\exp(-p)$.
\end{thm}
Theorem~\ref{thm:MLE} establishes the statistical convergence for the global MLE~\eqref{eq:MLE}. The result in~\eqref{eq:bound} implies that the estimation has a convergence rate $\tO(Kp/d^{K})$ as $(p,d)\to \infty$. {\color{blue}This rate agrees with intuition, since in our setting, the number of parameters with $K$ feature matrices is of order $\tO(Kp)$, whereas the number of tensor entries $\tO(d^K)$ corresponds to the total sample size. Because $p\leq d$, our rate is faster than $\tO(d^{-(K-1)})$ obtained by tensor decomposition without features~\citep{wang2020learning}.}

Inspection of our proof (Supplementary Notes) shows that the desired convergence rate holds not only for the MLE, but also for all local optimizers satisfying $\tL_{\tY}(\tC, \mM_1,\ldots,\mM_K)\geq \tL_{\tY} (\tC_{\textup{true}},\mM_{1,\textup{true}},\ldots,\mM_{K,\textup{true}})$. The observation indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective is large enough. In next section, we will provide the statistical accuracy for \emph{local} optimizer with provable convergence guarantee, at a cost of extra signal requirement.

\subsubsection{Empirical accuracy for local optimizers}\label{sec:local}

The optimization~\eqref{eq:MLE} is a non-convex problem due to the low-rank constraint in the feasible set $\tP$. {\color{blue}Under mild conditions, our warm initialization enjoys stable performance, and the subsequent iterations further improve the accuracy via linear convergence; i.e.\ sequence of iterates generated by Algorithm~\ref{alg:B} converges to optimal solutions at a linear rate.

\begin{prop}[Polynomial-time angle estimation]\label{lem:ini} Consider Gaussian tensor models with $b(\theta)=\theta^2/2$ in the objective function~\eqref{eq:loglikelihood}. Suppose the signal-to-noise ratio $\lambda^2/\phi \geq C p^{K/2}d^{-K}$ for some sufficiently large universal constant $C>0$. 
Under Assumption A1 with scaled feature matrices $\bar \mX_k=\sqrt{d}\mX_k$, or Assumption A1' with original feature matrices, the outputs from initialization Algorithm~\ref{alg:A} and iteration Algorithm~\ref{alg:B} satisfy the following two properties.
\begin{enumerate}[label=(\alph*)]
\item With probability at least $1-\exp(-p)$. 
\begin{equation}\label{eq:warm}
\max_{k\in[K]}\sin^2\Theta(\trueM,\hat \mM_k^{(0)})\leq {1\over 4}. 
\end{equation}
\item Let $t=1,2,3,\ldots,$ denote the iteration. There exists a contraction parameter $\rho\in(0,1)$, such that, with probability at least $1-\exp(-p)$, 
\begin{equation}\label{eq:local}
\max_{k\in[K]}\textup{sin}^2 \Theta(\trueM, \hat \mM^{(t)}_k) \lesssim \KeepStyleUnderBrace{{\phi p \over \lambda^2 d^K}}_{\text{statistical error}}+\KeepStyleUnderBrace{\rho^{t}\max_{k\in[K]}\textup{sin} \Theta^2(\trueM, \hat \mM_k^{(0)})}_{\text{algorithmic error}}.
\end{equation}
\end{enumerate}
\end{prop}

Proposition~\ref{lem:ini} provides the estimation errors for algorithm outputs at initialization and at each of the subsequent iterations. The initialization bound~\eqref{eq:warm} demonstrates the stability of warm initialization under a mild SNR requirement $\lambda^2/\phi \gtrsim p^{K/2}d^{-K}$. We can think of $d$ as the sample size while $p$ the number of parameters at mode $K$. This threshold is less stringent than $d^{K/2}$ required for unsupervised tensor decomposition features~\citep{han2020optimal,zhang2018tensor}. The condition confirms that a higher sample size mitigates the required signal level. The iteration bound~\eqref{eq:local} consists of two terms: the first term is the statistical error, and the second is the algorithmic error. The algorithmic error decays exponentially with the number of iterations, whereas the statistical error remains the same as $t$ grows. The statistical error is unavoidable and also appears in the global MLE; see Theorem~\ref{thm:MLE}.

As a direct consequence, we find the optimal iteration $t$ after which the algorithmic error is negligible compared to statistical error. 
\begin{thm}[Statistical rate for local optimizers]\label{thm:local}Consider the same condition as in Proposition~\ref{lem:ini} and the outputs by combining algorithms 1 and 2. There exists a constant $C>0$, such that, after $t\gtrsim K\log_{1/\rho}p$ iterations, our algorithm outputs satisfies
\[
\max_{k\in[K]}\textup{sin}^2 \Theta(\trueM, \hat \mM^{(t)}_k) \lesssim {\phi p\over \lambda^2d^K}, \quad \FnormSize{}{\trueB- \hat \tB^{(t)}}^2 \lesssim {\phi(r^K+Kpr)\over d^K}.
\]
\end{thm}
}

In practice, the signal level $\lambda$ is unknown, so the assumption in Theorem~\ref{thm:local} is challenging to verify in practice. We supply the theory by providing an alternative scheme -- random initialization -- and investigate its empirical performance. Figure~\ref{fig:loglike} shows the trajectories of objective function for order-3 tensors based on model~\eqref{eq:decomp}, where $ d \in\{25,30\}$, $p = 0.4d$, $ r\in\{3,6\}$ at all three modes. We consider data tensors with Gaussian, Bernoulli, and Poisson entries. Under all combinations of the dimension $d$, rank $r$, and type of the entries, Algorithm~\ref{alg:B} converges quickly in a few iterations upon random initialization, and the objective values at convergent points are close to or larger than the value at true parameters. In the experiment we conduct, we find little difference in the final estimation errors between the two initialization schemes. Random initialization appears good enough for Algorithm~\ref{alg:B} to find a convergent point with desired statistical guarantees. In practice, we recommend to run both warm and cold initializations, and choose the one with better convergent objective values.

\begin{figure}[t]
\centering
\includegraphics[width=16cm]{loglike.pdf}
\caption{Trajectory of the objective function with various dimension $d$ and rank $r$ under (a) Gaussian (b) Bernoulli (c) Poisson models. The dashed line represents the objective value at true parameters. }\label{fig:loglike}
\vspace{-.2cm}
\end{figure}

We conclude this section by revisiting the three examples mentioned in Section~\ref{sec:model}.

\begin{customexample}{1}[Spatio-temporal growth model] The estimated type-by-time-by-space coefficient tensor converges at the rate $\tO\left((p+r+s)/(dmn)\right)$ with $(p,r,s)\leq (d,m,n)$. The estimation achieves consistency as the dimension grows along either of the three modes.
\end{customexample}

\begin{customexample}{2} [Network population model] The estimated node-by-node-by-feature tensor converges at the rate $\tO\left((2d+p)/(d^2n)\right)$ with $p\leq n$. The estimation achieves consistency as the number of individuals or the number of nodes grows. 
\end{customexample}

\begin{customexample}{3} [Dyadic data with node attributes] The estimated feature-by-feature matrix converges at the rate $\tO\left(p/d^2\right)$ with $p\leq d$. Again, our estimation achieves consistency as the number of nodes grows. 
\end{customexample}

\subsection{Rank selection and computational complexity}\label{sec:tuning}
Our algorithm assumes the rank $\mr$ is given. In practice, the rank is often unknown and must be determined from the data. We propose to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC, where
\begin{equation}\label{eq:BIC}
\textup{BIC}(\mr)=-2\tL_{\tY}(\hat \tC, \hat \mM_1,\ldots,\hat \mM_K)+p_e(\mr)\log (\prod\nolimits_k d_k).
\end{equation}
Here, $p_e(\mr)\stackrel{\textup{def}}{=}\sum_k (p_k-r_k)r_k+\prod_k r_k$ is the effective number of parameters in the model. We choose $\hat \mr$ that minimizes $\textup{BIC}(\mr)$ via grid search. Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model. We evaluate the empirical performance of BIC in Section~\ref{sec:simulation}.

The computational complexity of our Algorithm is $\tO\left(d \sum_k p^3_k\right)$ for each iteration, where $d=\prod_k d_k$ is the total size of the data tensor. The update of $K$ factor matrices is $\tO(d\sum_k r^3_k p_k^3)$ via standard GLM routines. Furthermore, we demonstrate that, under certain SNR conditions, a polynomial number of iterations suffices to reach the desired statistical accuracy. Therefore, the total computational cost is polynomial in $p$ and $d$.

\section{Connection to other tensor regression methods}\label{sec:connection}
{\color{blue}We compare our supervised tensor decomposition (\textbf{STD}) with recent 12 tensor methods in the literature. Table~\ref{table:comp_table} summarizes these methods with their properties from four aspects: i) model specification, ii) number of feature matrices allowed, (iii) capability of addressing non-Gaussian response, and (iv) capability of addressing non-independent noise. The four closet methods to our are {\bf SupCP}~\citep{lock2018supervised}, {\bf Envelope}~\citep{li2017parsimonious}, {\bf mRRR}~\citep{luo2018leveraging} and {\bf GLSNet}~\citep{zhang2018network}; these methods all relate a data tensor to feature matrices with low-rank structure on the coefficients. As seen from the table, our method is the only one that allows multiple feature matrices among the five. {\bf Envelope} and {\bf SupCP} are developed for Gaussian data, and the Gaussianity facilities flexible extension to non-independent noise. In particular, {\bf Envelope} allows noise correlation in Kronecker structured form, whereas {\bf SupCP} allows noise correlation implicitly through decomposing the latent factors into fixed effects (related to features) and random effects (unrelated to features). On the other hand, the other three methods ({\bf mRRR}, {\bf GLSNet} and {\bf STD}) are developed for exponential family distribution with possibly non-additive noise. The generality makes the full modeling of noise correlation computationally challenging. We will compare the numerical performance of these methods in Section~\ref{sec:simulation}.

\afterpage{
\begin{landscape}
 \renewcommand{\arraystretch}{1.5} % Default value: 1
\begin{table}[h!]
\centering
\color{blue}
\resizebox{\columnwidth}{!}{
\begin{tabular}{c||c|ccc}
\hline
Method & Model& No.\ of features & non-Gaussianity & Non-independence \\
\hline
	STD (Ours) & $\bbE\tY = f(\tB \times \{\mX_1,\mX_2,\mX_3\}),\ \tB = \tC \times \{\mM_1,\mM_2,\mM_3\}$ &$3$& $\surd$& $\times$\\
	GCP, CP-ARP, CORALS & $ \bbE\tY = f(\entry{ \mA_1,\mA_2,\mA_3})$ &0& $\surd$& $\times$\\
         DCOT & $\bbE\tY = f((\tC_1+\tC_2) \times \{\mM_1, \mM_2, \mM_3 \})$ &0& $\surd$ & $\times$\\
         LRT, CRT & $y_{n} = \langle \tB,\ \tX_{n} \rangle + \epsilon_{n}$, various structure on $\tB$&0& $\times$& $\times$\\
	STAR & $y_n = \sum_m \langle \tB_m, \tF_m(\tX_{ijk})\rangle + \epsilon_n$, sparse-CP $\tB_m$ &0& $\times$ & $\times$\\
	SupCP &   $\tY = \entry{ \mA_1,\mA_2,\mA_3} + \tE,\  \mA_1 = \mX \mB +\tE',\ \tE\perp \tE'$ &1& $\times$& $\surd$\\
	mRRR &  $\bbE \mY =  f(\mX \mB)$, low-rank $\mB$ &1& $\surd$& $\times$ \\
	Envelope & $\tY = \tB \times_3 \mX + \tE,\  \tB = \tC \times \{ \mM_1, \mM_2, \mI \},\  \tE \sim \mathcal{TN}(\mSigma_1, \mSigma_2,\mI)$ &1& $\times$ & $\surd$\\
	GLSNet& $\bbE\tY = f({\bf 1} \otimes {\bf \Theta} + \tB \times_3 \mX)$, low-rank ${\bf \Theta}$, sparse $\tB$&1& $\surd$& $\times$\\
	STORE & $\tY = \tB \times_3 \mX + \tE$,  sparse-CP $\tB$ &1& $\times$& $\times$ \\
	\hline
\end{tabular}
}
\caption{\footnotesize Comparison of tensor regression/factorization methods. We focus on order-3 tensors for illustration. Calligraphic letters denote tensors, bold capital letters denote matrices, and little letters denote scalars. The dimension of tensors and matrices can be identified from the contexts.\\
\vspace{-0.1cm}\\
- Data: tensor response $\tY$, feature matrices $\mX,\mX_k$, predictor tensor $\tX_n$, scalar response $y_n$, sample index $n$, tensor mode $k=1,2,3$.  \\
- Parameter: Tuckor factors $\mM_k$, CP factors $\mA_k$, CP decomposition $\entry{\mA_1, \mA_2, \mA_3}$, coefficient tensor and matrix $\tB, \tB_m, {\bf \Theta}, \mB$.\\
- Function: a known link function $f(\cdot)$, a known basis function $\tF_m(\cdot)$. \\
- Noise: Gaussian tensor with i.i.d.\ entries $\tE, \tE'$, Gaussian tensor with Kronecker covariance $\tE \sim \mathcal{TN}(\mSigma_1, \mSigma_2, \mI)$, meaning $\text{Cov}(\text{vec}(\tE))=\mSigma_1\otimes\mSigma_2\otimes \mI$.\\
\vspace{-0.1cm}\\
- GCP: Generalized canonical polyadic tensor decomposition~\citep{hong2020generalized};\\
- CP-APR: CP alternating Poisson regression~\citep{chi2012tensors};\\
- CORALS: Generalized co-clustering method~\citep{li2020generalized};\\
- DCOT: Double core tensor decomposition~\citep{tarzanagh2019regularized};\\
- SupCP: Supervised PARAFAC/CANDECOMP factorization~\citep{lock2018supervised};\\
- mRRR: Mixed-response reduced-rank regression~\citep{luo2018leveraging};\\
- Envelope: Parsimonious tensor response regression~\citep{li2017parsimonious};\\
- GLSNet: Generalized connectivity matrix response regression~\citep{zhang2018network};\\
- STORE: Sparse tensor response regression~\citep{sun2017store};\\
- LTR: Low-rank tensor regression~\citep{han2020optimal};\\
- CRT: Convex regularized multi-response tensor regression~\citep{raskutti2019convex};\\
- STAR: Sparse tensor additive regression~\citep{hao2019sparse}.
}\label{table:comp_table}
\end{table}
\end{landscape}
}
}

Our model also has a close connection to higher-order interaction model~\citep{hao2020sparse} and tensor-to-tensor regression~\citep{lock2018tensor}. Model~\eqref{eq:decomp} can be viewed as a regression model with across-mode interactions in the reduced feature space. We take an order-3 tensor under the Gaussian model for illustration. Let $\mX,\mZ,\mW$ denote the feature matrix on mode $k=1, 2, 3$, respectively. Suppose that each mode has two-dimensional reduced features, denoted $\mM_1\mX=[\mx_1,\mx_2]$, $\mM_2\mZ=[\mz_1,\mz_2]$, $\mM_3\mW=[\mw_1,\mw_2]$. Here $\mx_1,\mx_2,\ldots,\mw_1, \mw_2$ are column vectors. Then the model~\eqref{eq:decomp} is equivalent to a regression model with across-mode interactions
\begin{equation}
\mathbb{E}(y_{ijk}|\mX,\mZ,\mW)=c_{111}\mx_{1i}\mz_{1j}\mw_{1k}+c_{121}\mx_{i1}\mz_{j2}\mw_{k1}+\cdots+c_{221}\mx_{2i}\mz_{2j}\mw_{1k}+c_{222}\mx_{2i}\mz_{2j}\mw_{2k},
\end{equation}
where $\entry{c_{ijk}} \in \mathbb{R}^{2\times 2\times 3}$ are unknown interaction effects, $\mx_{1i}$ denotes the $i$-th entry in the feature vector $\mx_1$, and similar notations apply to other features. Note that lower-order interactions are naturally incorporated if we include an intercept column in the reduced feature matrices. The above example shows the connection of our supervised tensor decomposition to multivariate regressions.

\section{Numerical experiments}\label{sec:simulation}
We evaluate the empirical performance of our supervised tensor decomposition (\textbf{STD}) through simulations. We consider order-3 tensors with a range of distribution types. Unless otherwise specified, the conditional mean tensor is generated form model~\eqref{eq:decomp}, where the core tensor entries are i.i.d.\ drawn from Uniform[-1,1], the factor matrix $\mM_k$ is uniformly sampled with respect to Haar measure from matrices with orthonormal columns. The feature matrix $\mX_k$ is either an identity matrix (i.e.,\ no feature  available) or Gaussian random matrix with i.i.d.\ entries from $N(0,1)$. The linear predictor $\Theta=\tC\times\{\mM_1\mX_1,\mM_2\mX_2,\mM_3\mX_3\}$ is scaled such that $\mnormSize{}{\Theta}=1$. Conditional on the linear predictor $\Theta=\entry{\theta_{ijk}}$, the entries in the tensor $\tY=\entry{y_{ijk}}$ are drawn independently according to three probabilistic models:

\begin{enumerate}[noitemsep,topsep=0pt]
\item[(a)] Gaussian model: continuous tensor entries $y_{ijk}\sim N\left(\alpha \theta_{ijk}, 1\right)$.
\item[(b)] Poisson model: count tensor entries $y_{ijk}\sim\textup{Poisson}\left(e^{\alpha \theta_{ijk}}\right)$.
\item[(c)] Bernoulli model: binary tensor entries $y_{ijk}\sim \textup{Bernoulli}\left( \frac{e^{\alpha \theta_{ijk}}}{1+e^{\alpha \theta_{ijk}}}\right)$.
\end{enumerate}
Here $\alpha>0$ controls the magnitude of the effect size, which is also the maximum norm of coefficient tensor as in~\eqref{eq:p}.
In each experiment, we report the summary statistics averaged across $30$ simulation replications.

\subsection{Finite-sample performance}
The first experiment assesses the selection accuracy of our BIC criterion~\eqref{eq:BIC}. We consider the balanced situation where $d_k=d$, $p_k=0.4d_k$ for $k=1,2,3$. We set $\alpha=4$ and consider various combinations of dimension $d$ and rank $\mr=(r_1,r_2,r_3)$. For each combination, we minimize BIC using a grid search from $(r_1-3,r_2-3,r_3-3)$ to $(r_1+3,r_2+3,r_3+3)$. We remove invalid rank such as $r^2_{\max} \geq \prod_{k=1}^3 r_k$ and use parallel search to reduce the computational cost. Table~\ref{tab:rank} reports the selected rank averaged over $n_{\textup{sim}}=30$ replicates. We find that in the high-rank setting with $d=20$, the selected rank slightly underestimates the true rank, and the accuracy immediately improves when either the dimension increases to $d = 40$ or the rank reduces to $\mr = (3,3,3)$. This agrees with our expectation, because in the tensor decomposition, the sample size is related to the number of tensor entries. A larger $d$ implies a larger sample size, so the BIC selection becomes more accurate.

\begin{table}[tb]
\resizebox{\textwidth}{!}{%
\centering
\begin{tabular}{c|cc|cc}
%True Rank $\mr$ &\multicolumn{2}{c|}{Dimension (Gaussian tensors)}&\multicolumn{2}{c}{Dimension (Poisson tensors)}\\
True rank $\mr$& $d=20$  (Gaussian) &$d=40$ (Gaussian) &$d=20$ (Poisson) &$d=40$ (Poisson)\\
\hline
$(3,\ 3,\ 3)$&$({\bf 3.0},\ {\bf 3.0},\ {\bf 3.0})$&$({\bf 3.0},\ {\bf 3.0},\ {\bf 3.0})$& $({\bf 3.0},\ {\bf 3.0},\ {\bf 3.0})$ & $({\bf 3.0},\ {\bf 3.0},\ {\bf 3.0})$\\
$(4,\ 4,\ 6)$&$(3.0,\ 3.0,\ {\bf 4.6})$&$({\bf 4.0},\ {\bf 4.0},\ {\bf 5.3})$&$(3.0,\ 3.0,\ {\bf 5.3})$&$({\bf 4.0},\ {\bf 4.0},\ {\bf 5.6})$\\
$(6,\ 8,\ 8)$&$(5.0,\ 5.0,\ 5.0)$&$({\bf 6.0},\ {\bf 8.0},\ {\bf 8.0})$&$(5.0,\ 5.0,\ 6.7)$&$({\bf 6.0},\ {\bf 8.0},\ {\bf 8.0})$\\
\end{tabular}
}
\caption{Rank selection via BIC. The estimated ranks are averaged across 30 simulation. Bold number indicates the ground truth is within two standard deviations of the estimate.}\label{tab:rank}
\end{table}

\begin{figure}[!h]
\centering
\includegraphics[width=14cm]{dimension.pdf}
\caption{Estimation error against effective sample size. The three panels plot the MSE when the response tensors are generated from (i) Gaussian (ii) Poisson and (iii) Bernoulli models. The dashed curves correspond to $\tO({1/d^2})$.}\label{fig:dim}
\vspace{-.1cm}
\end{figure}

The second experiment evaluates the accuracy when features are available on all modes. We set $\alpha=10, d_k=d, p_k=0.4d_k, r_k=r\in\{2,4,6\}$ and increase $d$ from 30 to 60. Our theoretical analysis suggests that $\hat \tB$ has a convergence rate $\tO(d^{-2})$ in this setting. Figure~\ref{fig:dim} plots the mean squared error (MSE) $\FnormSize{}{\hat \tB-\tB_{\text{true}}}^2$ versus the effective sample size $d^2$ under three different distribution models. We find that the empirical MSE decreases roughly at the rate of $1/d^2$, which is consistent with our theoretical results. We also observe that, tensors with higher rank tend to yield higher estimation errors, as reflected by the upward shift of the curves as $r$ increases. Indeed, a larger $r$ implies a higher model complexity and thus greater difficulty in the estimation.

\subsection{Comparison with other tensor methods}
We compare our supervised tensor decomposition with three other tensor methods:

{\color{blue}\begin{itemize}[noitemsep,topsep=0pt]
\item Supervised PARAFAC/CANDECOMP factorization (\textbf{SupCP}, \citep{lock2018supervised}).
    \item Parsimonious tensor response regression (\textbf{Envelope}, \citep{li2017parsimonious});
    \item Mixed-response reduced-rank regression (\textbf{mRRR}, \citep{luo2018leveraging});
    \item Generalized connectivity matrix response regression (\textbf{GLSNet}, \citep{zhang2018network});
\end{itemize}
}

{\color{blue}These four methods are the closest methods to ours, in that they all relate a data tensor to feature matrices with low-rank structure on the coefficients.} We consider Gaussian and Bernoulli tensors in experiments. For methods not applicable for Bernoulli data ({\bf SupCP} and {\bf Envelope}), we provide the algorithm $\{-1,1\}$-valued tensors as inputs. Because {\bf mRRR} allows matrix response only, we provide the algorithm the unfolded matrix of response tensor as inputs. We measure the accuracy using the response error defined as $1-\text{Cor}(\hat \tY, f(\trueT))$, where $\hat \tY$ is the fitted tensor from each method, and $f(\trueT)$ is the true conditional mean of the tensor. Note that the response error is a scale-insensitive metric; a smaller error implies a better fit of the model.

{\color{blue}The comparison is assessed from three aspects: (i) benefit of incorporating features from multiple modes; (ii) prediction error with respect to sample size; (iii) robustness of model misspecification.} We use similar simulation setups as in our first experiment in last section. We consider rank $\mr=(3,3,3)$ (low) vs.\ $(4,5,6)$ (high), effect size $\alpha = 3$ (low) vs.\ $6$ (high), dimension $d$ ranging from 20 to 100 for modes with features, and $d = 20$ for modes without features. The method \textbf{Envelope} and \textbf{mRRR} require the tensor rank as inputs, respectively. For fairness, we provide both algorithms the true rank. The methods \textbf{SupCP} and \textbf{GLSNet} use different notions of model rank, and \textbf{GLSNet} takes sparsity as an input. We use a grid search to set the hyperparameters in \textbf{SupCP} and \textbf{GLSNet} that give the best performance.

\begin{figure}[t]
\centering
\includegraphics[width=12cm]{comp_final_normal.pdf} 
\caption{Comparison between tensor methods with Gaussian data. Panels (a) and (b) plot estimation error versus the number of modes with available features. Panels (c) and (d) plot ME versus the effective sample size $d^2$.
We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and effect size $\alpha =3 $ (low), $\alpha=6$ (high).}~\label{fig:comp}
\vspace{-.5cm}
\end{figure}

{\color{blue}Figure~\ref{fig:comp}a-b shows the impact of features to estimation error. We see that our \textbf{STD} outperforms others, especially in the low-effect high-rank setting. As the number of informative modes increases, the \textbf{STD} exhibits a reduction in error whereas others remain unchanged. The accuracy gain in Figure~\ref{fig:comp} demonstrates the benefit of incorporating informative features from multiple modes. In addition, we find that the relative performance among the competing methods reveals the benefits of low-rankness. 
The second best method is \textbf{SupCP} which imposes low-rankness on three modes; the next one is \textbf{Envelope} which imposes low-rankness on two modes; the less favorable one is \textbf{mRRR} which imposes low-rank structure on one mode only; the worst one is \textbf{GLSNet} which imposes sparsity but no low-rankness on the feature effects.

Figure~\ref{fig:comp}c-d compares the prediction error with respect to effective sample size $d^2$. For fair comparison, we consider the setting with feature matrix on one mode only. We find that our \textbf{STD} method has similar performance as \textbf{Envelope} and \textbf{SupCP} in the high-effect low-rank regime, whereas the improvement becomes more pronounced in the low-effect high-rank regime. The latter setting is notably harder, and our \textbf{STD} method shows advantage in addressing this challenge. Among other methods, \textbf{Envelope}, \textbf{SupCP}, and \textbf{mRRR} show decreasing errors as $d$ increases, implying the benefits of low-rankness methods. In contrast, \textbf{GLSNet} suffers from non-decreasing error and indicates the poor fit of sparsity methods in addressing low-rank data.

\begin{figure}[t]
\centering
\includegraphics[width=12cm]{comp_final_binary.pdf} 
\caption{Comparison between tensor methods with Binary data. The panel legends are the same as in Figure~\ref{fig:comp}.}~\label{fig:comp_b}
\vspace{-.5cm}
\end{figure}

We also evaluate the performance comparison with Bernoulli tensors. Figure~\ref{fig:comp_b} indicates the necessarity of generalized model in addressing non-Gaussian data. Indeed, methods that assume Gaussiannity (\textbf{Envelope} and \textbf{SucCP}) perform less favorably in Bernoulli setting (Figure~\ref{fig:comp_b}c) compared to Gaussian setting (Figure~\ref{fig:comp}c). Our method shows improved accuracy as the number of informative features increases (Figure~\ref{fig:comp_b}a-b). In the absence of multiple features, our method still performs favorably compared to others (Figure~\ref{fig:comp_b}c-d), for the same reasons we have argued in Gaussian data.

\begin{figure}[t]
\centering
\includegraphics[width=12cm]{noniid_final.pdf} 
\caption{Comparison between tensor methods under model misspecification. Panels (a)-(b) assess the noise correlation, and panels (d)-(d) assess the sparsity. }~\label{fig:noniid}
\vspace{-.5cm}
 \end{figure}

Lastly, we assess the performance of our method {\bf STD} under model misspecification. We consider two aspects: (i) non-independent noise, and (ii) sparse feature effects. Note that our method {\bf STD} imposes neither of these two assumptions, so the experiment allows us to assess the robustness. We select competing methods from Table~\ref{table:comp_table} that specifically addresses these two aspects. We use {\bf Envelope} and {\bf SupCP} as benchmark for noise correlation experiment, and {\bf GLSNet} for sparsity experiment.

Figure~\ref{fig:noniid}a-b assesses the impact of noise correlation to the estimation accuracy. The data is simulated from {\bf Envelope} model with envelope dimensions $r=(3,3)$ (low) and $(4,5)$ (high). The noise is generated from a zero-mean Gaussian tensor with Kronecker structured covariance; see Supplementary Notes for details. As expected, {\bf Envelope} shows the best performance in the high correlation setting. Remarkably, we find that our method {\bf STD} has comparable and sometimes better performance when noise correlation is moderate-to-low. In contrast, {\bf SupCP} appears less suitable in this setting. Although {\bf SupCP} allows noise correlation implicitly through latent random factors, the induced correlation may not belong to the Kronecker covariance structure in the simulation.

Figure~\ref{fig:noniid}c-d assesses the impact of sparsity to estimation performance. We generate data from {\bf GLSNet} model, except that we modify the coefficient tensor to be joint sparse and low-rank (the original {\bf GLSNet} model assumes full-rankness on the coefficient tensor). 
The sparsity level ($x$-axis in Figure~\ref{fig:noniid}c-d) quantifies the proportion of zero entries in the coefficient tensor. Since neither our method {\bf STD} nor {\bf GLSNet} follow the simulated model, this setting allows a fair comparison. We find that our method outperforms {\bf GLSNet} in the low-rank setting, whereas {\bf GLSNet} shows a better performance in the high-rank setting. This observation suggests the robustness of our method to sparsity when the tensor of interest is simultaneously low-rank and sparse. When sparsity is the only salient structure, then methods specifically addressing sparsity would provide a better fit. }

\section{Data applications}\label{sec:data}
We apply our supervised tensor decomposition to two datasets. The first application studies the brain networks in response to individual attributes (i.e.,\ feature on one mode), and the second application focuses on multi-relational network analysis with dyadic attributes (i.e.,\ features on two modes).

\subsection{Application to human brain connection data}

The Human Connectome Project (HCP) aims to build a network map that characterizes the anatomical and functional connectivity within healthy human brains~\citep{van2013wu}. We follow the preprocessing procedure as in~\citep{desikan2006automated} and parcellate the brain into 68 regions of interest. The dataset consists of 136 brain structural networks, one for each individual. Each brain network is represented as a 68-by-68 binary matrix, where the entries encode the presence or absence of fiber connections between the 68 brain regions. We consider four individual features: gender (65 females vs.\ 71 males), age 22-25 ($n=35$), age 26-30 ($n=58$), and age 31+ ($n=43$). The preprocessed dataset is released in our R package \texttt{tensorregress}. The goal is to identify the connection edges that are affected by individual features. A key challenge in brain network is that the edges are correlated; for example, the nodes in edges may be from a same brain region, and it is of importance to take into account the within-dyad dependence.

\begin{figure}[!h]
\centering
\includegraphics[width=16cm]{HCP.pdf}
\caption{Top edges with large effects. (a) Global effect; (b) Female effect; (c) Age 22-25; (d) Age 31+. Red edges represent positive effects and blue edges represent negative effects. The edge-width is proportional to the magnitude of the effect size.
}\label{fig:brain}\label{fig:s1}
\vspace{-.5cm}
\end{figure}

We perform the supervised tensor decomposition to the HCP data. %The data tensor is binary, $\tY\in\{0,1\}^{68\times 68\times 136}$, and the features are of dimension 4 on the 3$^{\text{rd}}$ mode. 
The BIC selection suggests a rank $\mr=(10,10,4)$ with quasi log-likelihood $\tL_{\tY}=-174654.7$. We utilize the sum-to-zero contrasts in coding the feature effects, and depict only the top 3\% edges whose connections are non-constant across the sample. Figure~\ref{fig:brain} shows the top edges with high effect size, overlaid on the Desikan atlas brain template~\citep{desikan2006automated}. We find that the global connection exhibits clear spatial separation, and that the nodes within each hemisphere are more densely connected with each other (Figure~\ref{fig:brain}a). In particular, the superior-temproal (\emph{SupT}), middle-temporal (\emph{MT}) and Insula are the top three popular nodes in the network. Interestingly, female brains display higher inter-hemispheric connectivity, especially in the frontal, parental and temporal lobes (Figure~\ref{fig:brain}b). This is in agreement with a recent study showing that female brains are optimized for inter-hemispheric communication~\citep{ingalhalikar2014sex}. We find several edges with declined connection in the group Age 31+. Those edges involve Frontal-pole (\emph{Fploe}), superior-frontal (\emph{SupF}) and Cuneus nodes. The Frontal-pole region is known for its importance in memory and cognition, and the detected decline with age further highlights its biological importance.

\subsection{Application to political relation data}

The second application studies the multi-relational networks with node-level attributes. We consider \emph{Nations} dataset~\citep{nickel2011three} which records 56 relations among 14 countries between 1950 and 1965. The multi-relational networks can be organized into a $14 \times 14 \times 56$ binary tensor, with each entry indicating the presence or absence of an action, such as ``sending tourist to'', ``export'', ``import'', between countries. The 56 relations span the fields of politics, economics, military, religion, etc. In addition, country-level attributes are also available, and we focus on the following six features: \emph{constitutional, catholics, law ngo, political leadership, geography}, and \emph{medicine ngo}. The goal is to identify the variation in connections due to country-level attributes and their interactions.

We apply our tensor model to the \emph{Nations} data. The multi-relational network $\tY$ is a binary data tensor, and the country attributes $\mX\in\mathbb{R}^{14\times 6}$ are features on both the 1$^\text{st}$ and 2$^\text{nd}$ modes. {\color{blue} We use BIC as guidance to select the rank of coefficient tensor $\tB$. Since several rank configurations give similar BIC values, we present here the most interpretable results with $\mr=(4,4,4)$. Detailed rank selection procedure is in Supplementary Notes.}
We perform the supervised tensor decomposition and obtain the dimension reduction matrices $\hat \mM_k$ from the model. Then we apply $K$-mean clustering to dimension reduction matrix on each of the modes. Supplementary Table~\ref{fig:s1} shows the K-means clustering of the 56 relations based on the dimension reduction matrix on the 3$^\text{rd}$ mode. We find that the relations reflecting the similar aspects of actions are grouped together. In particular, Cluster I consists of military relations such as \emph{violentactions}, \emph{warnings} and \emph{militaryactions}; Clusters II and III capture the economic relations such as \emph{economicaid, booktranslations, tourism}; and Cluster IV represents the political alliance and territory relations.

\begin{figure}[!h]
\centering
\includegraphics[width=16cm]{coef.pdf}
\caption{Estimated feature effects in the \emph{Nations} data analysis. Panels (a)-(d) represent the estimated effects of country-level attributes towards the connection probability, for relations \emph{warning}, \emph{violentactions}, \emph{treaties}, and \emph{aidenemy}, respectively. }\label{fig:est}
\end{figure}

To investigate the effects of dyadic attributes towards connections, we depict the estimated coefficients $\hat \tB=\entry{\hat b_{ijk}}$ for several relation types (Figure~\ref{fig:est}). {\color{blue}The entry $\hat b_{ijk}$ estimates the contribution, at the logit scale, of feature pair $(i,j)$ ($i$th feature for the ``sender'' country and $j$th feature for the ``receiver'' country) towards the connection of relation $k$.} Several interesting findings emerge from the observation. We find that relations belonging to a same cluster tend to have similar feature effects. For example, the relations ``warning'' and ''violentactions'' are classified into Cluster I, and both exhibit similar effect patterns (Figure~\ref{fig:est}a-b). Moreover, the feature \emph{constitutional} has a strong effect in the relation ``violentactions'' and ``warning'', whereas the effect is weaker in the relation ``treaties''. The result is plausible because the constitutional attributes affect political actions more often than economical actions. The entries in $\tB$ are useful for revealing interaction effects in a context-specific way. 
%For a given relation $k$, the off-diagonal entries in the matrix $\hat \tB(\colon,\colon,k)$ represent the pairwise interactions between features. 
From Figure~\ref{fig:est}, we find a strong interaction between \emph{geography} and \emph{political leadership} in the relation ``warning'', and a strong interaction between \emph{geogrphy} and \emph{medicine ngo} in the relation ``aidenemy''. The relation-specific effect pattern showcases the applicability of our method in revealing complex interactions.

\vspace{-0.5cm}

\section{Discussion and future work}\label{sec:con}
We have developed a supervised tensor decomposition method with side information on multiple modes. One important challenge of tensor data analysis is the complex interdependence among tensor entries and between multiple features. Our approach incorporates side information as feature matrices in the conditional mean tensor. The empirical results demonstrate the improved interpretability and accuracy over previous approaches. Applications to the brain connection and political relationship datasets yield conclusions with sensible interpretations, suggesting the practical utility of the proposed approach.

There are several possible extensions from the work. We have provided accuracy guarantees for parameter estimation in the supervised tensor model. Statistical inference based on tensor decomposition is an important future direction. Measures of uncertainty, such as confidence envelope for space estimation, would be useful. One possible approach would be performing parametric bootstrap~\citep{efron1994introduction} to assess the uncertainty in the estimation. 
For example, one can simulate tensors from the fitted low-rank model based on the estimates, and then assess the empirical distribution of the estimates. 
While being simple, bootstrap approach is often computationally expensive for large-scale data. Another possibility is to leverage recent development in debiased inference with distributional characterization~\citep{chen2019inference}. This approach has led to fruitful results for matrix data analysis. Uncertainly quantification involving tensors are generally harder, and establishing distribution theory for tensor estimation remains an open problem.

One assumption made by our method is that tensor entries are conditionally independent given the linear predictor $\Theta$. This assumption can be extended by introducing a more general mixed-effect tensor model. For example, in the special case of Gaussian model, we can model the first two moments of data tensor using
\begin{align}\label{eq:extention}
\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)&=\tC\times\mM_1\times \cdots \times \mM_K,\\
 \text{Var}(\tY|\mX_1,\ldots,\mX_K)&=\mPhi_1\otimes \cdots \otimes \mPhi_K,
\end{align}
where $\mPhi_k\in\mathbb{R}^{d_k\times d_k}$ is the unknown covariance matrix on the mode $k\in[K]$. For general exponential family, an additional mean-variance relationship should also be considered. The joint estimation of mean model $\Theta$ and variance model $\mPhi_k$ will lead to more efficient estimation in the presence of unmeasured confounding effects. However, the introduction of unknown covariance matrices $\mPhi_k$ dramatically increases the number of parameters in the problem. {\color{blue}Suitable regularization such as graphical lasso or specially-structured covariance~\citep{li2017parsimonious,lock2018supervised} should be considered.} The extension of tensor modeling with heterogeneous mixed-effects will be an important future direction.

Although we have presented the data applications in the context of order-3 data tensors, the framework of the supervised tensor decomposition applies to a variety of multi-way datasets. One possible application is the integrative analysis of omics data, in which multiple types of omics measurements (gene expression, DNA methylation, microRNA) are collected in the same set of individuals~\citep{lock2013joint,wang2019three}. Other applications include time-series tensor data with multiple side information. Exploiting the benefits and properties of  supervised tensor decomposition in specialized task will boost scientific discoveries.

\bigskip
\begin{center}
{\large \bf SUPPLEMENTARY MATERIALS}
\end{center}

\begin{description}
\item[Supplementary notes:]  technical proofs, additional simulation, and data analysis results.
\item[Data and software:] 
Our simulation code, R-package \texttt{tensorregress}, and datasets used in the paper are available at \url{https://CRAN.R-project.org/package=tensorregress}
\end{description}

\setstretch{1.25}
\bibliographystyle{apalike}

\bibliography{tensor_wang}

\end{document}

### 2021-06-21 tensor_regress/JCGS/Response_submission/manuscript/supp.tex
new figure 2

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url}
\usepackage{enumitem}
\usepackage[dvipsnames]{xcolor}

\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{algorithm}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={black},citecolor={black},urlcolor={black}}

\newcommand{\blind}{1}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\theoremstyle{definition}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}[section]
\newtheorem{prop}{Proposition}[section]
\newtheorem{pro}{Property}[section]
\newtheorem{cor}{Corollary}[section]

\newtheorem{assumption}{Assumption}[section]
\newtheorem{defn}{Definition}[section]
\newtheorem{example}{Example}[section]
\newtheorem{rmk}{Remark}[section]

\newtheorem{innercustomgeneric}{\customgenericname}
\providecommand{\customgenericname}{}
\newcommand{\newcustomtheorem}[2]{%
  \newenvironment{#1}[1]
  {%
   \renewcommand\customgenericname{#2}%
   \renewcommand\theinnercustomgeneric{##1}%
   \innercustomgeneric
  }
  {\endinnercustomgeneric}
}

\newcustomtheorem{customexample}{Example}

\usepackage{appendix}
\usepackage{wrapfig}
\mathtoolsset{showonlyrefs}

%This is for cross reference in overleaf
\usepackage{xr}
\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}
\myexternaldocument{jcgs-tensor}

\input macros.tex

\usepackage[english]{babel}

\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}

\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\def\ci{\perp\!\!\!\perp}

\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}
\usepackage{booktabs}
\newcommand\doubleRule{\toprule\toprule}
\allowdisplaybreaks

\usepackage[compact]{titlesec}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1.5}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
{\Large\bf Supplementary Notes to ``Supervised tensor decomposition with features on multiple modes''}

\vspace{.5cm}

{\large Jiaxin Hu, Chanwoo Lee, Miaoyan Wang}\\
University of Wisconsin-Madison

\vspace{.5cm}
\end{center}
\appendix
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}   
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

The supplementary note consists of proofs (Section~\ref{sec:appendix}), additional simulation results (Section~\ref{sec:asimulation}), and data applications (Section~\ref{sec:adata}).

\section{Proofs}\label{sec:appendix}
\subsection{Proof of Theorem~\ref{thm:MLE}}
We denote several quantities:
\begin{equation}\label{eq:quantities}
\underline \gamma = \prod_{k\in[K]} \sigma_{\min}(\mX_k),\quad \bar \gamma =\prod_{k\in[K]}\sigma_{\max}(\mX_k),  \quad \lambda =\min_{k\in[K]}\sigma_{\min}(\text{Unfold}_k(\tB_{\text{true}})),
\end{equation}
where $\underline{\gamma}$ quantifies the rank non-deficiency of feature matrices, $\bar \gamma$ quantifies the magnitude of feature matrices, and $\lambda$ is the smallest singular value of mode-$k$ unfolded matrices $\text{Unfold}_k(\trueB)$ for all possible $k\in[K]$. 
For notational convenience, we drop the subscript $\tY$ from the objective $\tL_\tY(\cdot)$ and simply write as $\tL(\cdot)$. We write $\tL(\tB)$ in place of $\tL(\tC, \mM_1,\ldots,\mM_K)$ when we want to emphasize the role of $\tB$.

\begin{prop}[sub-Gaussian residuals]\label{prop}
Define the residual tensor $\tE=\entry{\varepsilon_{i_1,\ldots,i_K}}=\tY-b'(\Theta)\in\mathbb{R}^{d_1\times \cdots \times d_K}$. Under the Assumption A2, $\varepsilon_{i_1,\ldots,i_K}$ is a sub-Gaussian random variable with sub-Gaussian parameter bounded by $\phi U$, for all $(i_1,\ldots,i_K)\in[d_1]\times\cdots\times[d_K]$.
\end{prop}

\begin{prop}[Properties of tensor GLM]\label{lem}Consider tensor GLMs under Assumption A2.
\begin{enumerate}[label=(\alph*)]
\item (Strong convexity) For all $\tB$ and all realized data tensor $\tY$,
\[
\tL(\trueB)\geq \tL(\tB)+ \langle \nabla \tL(\tB_{\text{true}}), \trueB-\tB \rangle+{1\over 2} \underline{\gamma}^2L\FnormSize{}{\trueB-\tB}^2,
\] 
where $\nabla L(\cdot)$ denotes the derivative of $\tL$ with respect to $\tB$.   
\item (Model complexity) 
Suppose $\tY$ follows generalized tensor model with parameter $\trueB$. Then, with probability at least $1-\exp(-p)$, 
\begin{equation}\label{eq:b}
\text{Err}_{\text{ideal}}(\mr):= \sup_{\substack{ \FnormSize{}{\tB}=1, \tB\in\tP(\mr)}}\langle \nabla \tL(\tB_{\text{true}}), \tB \rangle\lesssim \bar \gamma \sqrt{\phi U(r^K+Kpr)}. 
\end{equation}
\end{enumerate}
\end{prop}
The proofs of Propositions~\ref{prop}-\ref{lem} are in Section~\ref{sec:lemma}.

\begin{proof}[Proof of Theorem~\ref{thm:MLE}]
First we prove the error bound for $\MLEB$. By the definition of $\MLEB$, $\tL_{\tY}(\trueB)-\tL_{\tY}(\MLEB)\leq 0$. By the strong convexity in Proposition~\ref{lem},
\begin{align}\label{eq:F-norm}
0&\geq \tL_{\tY}(\trueB)- \tL_{\tY}(\MLEB)  \geq \langle \nabla \tL(\tB_{\text{true}}), \trueB-\MLEB \rangle+{1\over 2}\underline{\gamma}^2L\FnormSize{}{\trueB-\MLEB}^2.
\end{align}
Rearranging~\eqref{eq:F-norm} gives
\[
\FnormSize{}{\MLEB-\trueB}\leq {2\over \underline{\gamma}^2L} \left\langle \nabla \tL(\tB_{\text{true}}), {\MLEB-\trueB\over\FnormSize{}{\MLEB-\trueB}} \right\rangle\leq {2\over \underline{\gamma}^2L}  \text{Err}_{\text{ideal}}(2\mr),
\]
where the last inequality comes from the definition of $\text{Err}_{\text{ideal}}(2\mr)$ and the fact that $\text{rank}(\MLEB-\trueB)\leq \text{rank}(\MLEB)+\text{rank}(\trueB)\leq 2\mr$. By~\eqref{eq:b} in Proposition~\ref{lem}, we have 
\begin{equation}\label{eq:Bbound}
\FnormSize{}{\MLEB-\trueB} \lesssim {\bar \gamma \sqrt{\phi U}\over \underline{\gamma}^2 L}\sqrt{r^K+Kpr},
\end{equation}
with probability at least $1-\exp(-p)$.

Now, we specialize $\bar \gamma/\underline{\gamma}^2 $ in the following two cases of assumptions on feature matrices. 
\begin{itemize}[leftmargin=0cm]
\item[] [Case 1] Under Assumption A1 with scaled feature matrices, we have
\begin{equation}\label{eq:case1}
{\bar \gamma\over \underline{\gamma}^2} \leq {c^K_2d^{K/2} \over c^{2K}_1d^{K}} \lesssim \sqrt{1\over d^{K}}.
\end{equation}
\item[] [Case 2] Under Assumption A1' with original feature matrices, the asymptotic behavior of extreme singular values \citep{rudelson2010non} are
\[
\sigma_{\min}(\mX_k) \asymp \sqrt{d}-\sqrt{p} \text{ and  } \sigma_{\max}(\mX_k) \asymp \sqrt{d}+\sqrt{p},\quad \text{for all }k\in[K].
\]
In this case, we obtain 
\begin{equation}\label{eq:case2}
{\bar \gamma\over \underline{\gamma}^2} \asymp {(\sqrt{d}+\sqrt{p})^{K} \over (\sqrt{d}-\sqrt{p})^{2K}} \lesssim \sqrt{1\over d^K}.
\end{equation}
\end{itemize}
Combining~\eqref{eq:Bbound} with either~\eqref{eq:case1} or~\eqref{eq:case2}, in both cases we obtain the same conclusion
\begin{equation}\label{eq:bound2}
\FnormSize{}{\MLEB-\trueB}^2\lesssim {\phi(r^K+Kpr)\over d^K}.
\end{equation}
Now we prove the bound for sin$\Theta$ distance. 
We unfold tensors $\trueB$ and $\MLEB$ along the mode $k$ and obtain $\text{Unfold}_k(\trueB)$ and $\text{Unfold}_k(\MLEB)$. Notice that $\trueM$ and $\hat \mM_{k,\textup{MLE}}$ span the top-$r$ left singular spaces of $\text{Unfold}_k(\trueB)$ and $\text{Unfold}_k(\MLEB)$, respectively.  Applying Proposition~\ref{prop:sinebound} to this setting gives
\begin{align}\label{eq:sintheta}
    \text{sin}\Theta(\trueM,\hat\mM_{k,\textup{MLE}})\leq \frac{\FnormSize{}{\text{Unfold}_k(\MLEB)-\text{Unfold}_k(\trueB)}}{\sigma_{\text{min}}(\text{Unfold}_k(\trueB))}= \frac{\FnormSize{}{\MLEB-\trueB}}{\lambda}.
    \end{align}
The proof is complete by combining~\eqref{eq:bound2} and~\eqref{eq:sintheta}.
\end{proof}

\subsection{Proofs of Proposition~\ref{lem:ini} and Theorem~\ref{thm:local}}
\begin{proof}[Proof of Proposition~\ref{lem:ini}]

We express the Gaussian model as 
\begin{align}\label{eq:gaussian}
    \tY = \tB_{\text{true}}\times \{\mX_1,\ldots,\mX_K\} + \tE,
\end{align}
where $\tE$ is a noise tensor consisting of i.i.d.\ entries from $N(0,\sqrt{\phi}).$ 
By QR decomposition on feature matrices, $\mX_k = \mQ_k\mR_k$ for all $k\in[K]$, we have 
\begin{align}\label{eq:gaussian2}
    \bar\tY = \tB_{\text{true}}\times\{\mR_1,\ldots \mR_K\}+\bar\tE,
\end{align}
where $\bar\tY = \tY\times\{\mQ_1,\ldots,\mQ_K\}$ and  $\bar\tE = \tE\times\{\mQ_1,\ldots,\mQ_K\}$.
Notice that entries of $\bar\tE\in\mathbb{R}^{p\times\cdots\times p}$ are i.i.d drawn from $N(0,\sqrt{\phi})$ by the orthonormality of $\{\mQ_k\}_{k=1}^K$. Reparameterize the signal in \eqref{eq:gaussian2} as 
\begin{align}\label{eq:repara}
    \tS_{\text{true}} := \tB_{\text{true}}\times\{\mR_1,\ldots \mR_K\}&=  \tC_{\text{true}}\times\{\mR_1\mM_{1,\text{true}},\ldots \mR_K\mM_{K,\text{true}}\}\notag \\
    & = \tC'_{\text{true}}\times\{\mU_{1,\text{true}},
    \ldots,\mU_{K,\text{true}}\},
\end{align}
where $\mU_{k.\text{true}}\in\mathbb{O}(p_k, r_k)$ are orthonormal matrices and $\tC'_{\text{true}}\in\mathbb{R}^{r\times \cdots\times r}$ is a full rank core tensor.
By definition of quantities in \eqref{eq:quantities}, we have
\begin{align}\label{eq:lambdas}
    \lambda' := \min_{k\in[K]}\sigma_{\min}(\text{Unfold}_k(\tS_{\text{true}}))\in[  \lambda \underline \gamma,\ \lambda \bar \gamma].
\end{align}

Now our setup shares the same setting as in~\citet[Theorem 1]{zhang2018tensor}.
We summarize the relationships between our algorithm outputs and the ones in \cite{zhang2018tensor}. For all $k\in[K],$
\begin{enumerate}
    \item  $\mM_{k,\text{true}} = \text{SVD}_{r_k}\left(\mR_k^{-1}\mU_{k,\text{true}}\right):=$  the first $r_k$ left singular vectors of $\mR_k^{-1}\mU_{k,\text{true}}$;
   % \item $\hat\mM^{(0)}_k = \hat \mU_k^{(0)}$.
    \item $\hat\mM_{k}^{(t)} = \text{SVD}_{r_k}\left(\mR_k^{-1}\hat\mU_{k}^{(t)}\right)$ for all $t=0,1,2,\ldots$;
\end{enumerate}
where $\hat \mU^{(t)}_k$ denotes the $t$-th iteration output of Higher Order Orthogonal Iteration (HOOI) algorithm \citep{zhang2018tensor} with inputs $\bar \tY$. 
The first relationship is from~\eqref{eq:repara}, and second relationship is from induction by $t$. Briefly, $t=0$ holds because of the definition  $\hat\mM^{(0)}_k$ based on lines 4-5 of our initialization Algorithm~\ref{alg:A}. For $t\geq 1,\ldots$, notice that $\hat\mM_k^{(t)}$ is an optimizer of the objective \[\FnormSize{}{\bar\tY-\hat\tC^{(t-1)}\times\{\mR_1\hat\mM_1^{(t)},\ldots,\mR_{k-1}\hat\mM_{k-1}^{(t)},\mR_k\mM,\mR_{k+1}\hat\mM_{k+1}^{(t-1)},\ldots,\mR_K\hat\mM^{(t-1)}_K\}}^2,\]
from the line 3 of  Algorithm~\ref{alg:B}. By unfolding along the mode $k$, the optimizer $\mM_k^{(t)}$ must satisfy
\begin{align}\label{eq:HOOI}
    \text{Unfold}_k&\left(\bar\tY\times\left\{ (\hat\mM_1^{(t)})^T\mR_1^{-1},\ldots, (\hat\mM_{k-1}^{(t)})^T\mR_{k-1}^{-1},\ \mI_{p_k}, \ (\hat\mM_{k+1}^{(t-1)})^T\mR_{k+1}^{-1}, \ldots, (\hat\mM_{K}^{(t-1)})^T\mR_{K}^{-1}\right\} \right)\notag \\
    &=\mR_k\hat\mM_k^{(t)}\text{Unfold}_k\left(\hat\tC^{(t-1)}\right)\left(\mI_{r_K}\otimes\cdots\otimes\mI_{r_{k+1}}\otimes\mI_{r_{k-1}}\otimes \mI_{r_1}\right).
\end{align}
Notice that the first $r_k$ left singular vectors of the left side of~\eqref{eq:HOOI} is $\hat\mU^{(t)}_k$ in HOOI algorithm. Therefore, we prove the  the second relationship by induction.

Combination of Lemma~\ref{lem:sint} and the relationships between our algorithm outputs and the ones in \cite{zhang2018tensor} gives us
\begin{align}\label{eq:sinrel}
    \left(\frac{\underline\gamma}{\bar\gamma}\right)^2\max_{k\in[K]}\sin\Theta(\mU_{k,\text{true}},\hat\mU_k^{(t)})\leq
    \max_{k\in[K]}\sin\Theta(\mM_{k,\text{true}},\hat\mM_k^{(t)})\leq \left(\frac{\bar\gamma}{\underline\gamma}\right)^2\max_{k\in[K]}\sin\Theta(\mU_{k,\text{true}},\hat\mU_k^{(t)}).
\end{align}
Now, we prove the property (a) in Proposition~\ref{lem:ini}. Based on Lemma~\ref{lem:lts}(a), whenever $\lambda'/\sqrt{\phi}\geq C_{\text{gap}}p^{K/4}$, we have
\begin{align}\label{eq:inim}
    \max_{k\in[K]}\sin \Theta(\mU_{k,\text{true}},\hat\mU_k^{(0)}) \leq c\left(\frac{p^{K/2}}{(\lambda\underline\gamma)^2/\phi}\right),
\end{align}
 with probability at least $1-\exp(-p)$.
Notice that
\begin{align}\label{eq:cgap}
    \lambda'\stackrel{\eqref{eq:lambdas}}{\geq} \lambda \underline \gamma\gtrsim \lambda d^{K/2}\geq C_{\text{gap}}\sqrt{\phi} p^{K/4},
\end{align}
where the second inequality uses [Case 1] and [Case 2] in the proof of Theorem~\ref{thm:MLE}. 
The condition $\lambda/\sqrt{\phi}\geq C p^{K/4} d^{-K/2}$ guarantees  a sufficiently large $C_{\text{gap}}$ that satisfies $\lambda'/\sqrt{\phi}\geq C_{\text{gap}}p^{K/4}$.
Thus combining \eqref{eq:sinrel} and \eqref{eq:inim} yields
\begin{align}
    \max_{k\in[K]}\sin\Theta(\mM_{k,\text{true}},\hat\mM_k^{(0)}) &
    \leq \left(\frac{\bar\gamma}{\underline\gamma}\right)^2\left(\sqrt{\phi}p^{K/4}\over\lambda\underline\gamma\right)^2
    \\
    &\leq \frac{1}{2},
\end{align}
where the last inequality uses the fact that  $\underline\gamma\asymp d^{K/2}$ and $\bar\gamma/\underline\gamma$ is bounded by a constant in [Case 1] and [Case 2], and the condition $\lambda /\sqrt{\phi}\geq Cp^{K/4}d^{-K/2}$.

Now, we prove the property (b) in Proposition~\ref{lem:ini}. Based on Lemma~\ref{lem:lts}(b), we have
 \begin{align}\label{eq:lt2}
      \max_{k\in[K]}\sin\Theta(\mU_{k,\text{true}},\hat\mU_k^{(t)})\lesssim \frac{\sqrt{p\phi}}{\lambda\underline\gamma}+ \left(1\over 2\right)^t\max_{k\in[K]}\sin\Theta(\mU_{k,\text{true}},\hat\mU_k^{(0)}),
 \end{align}
  with probability at least $1-\exp(-p)$.
  Combining \eqref{eq:sinrel} with the above inequality yields
 \begin{align}
     \max_{k\in[K]}\sin\Theta(\mM_{k,\text{true}},\hat\mM_k^{(t)}) 
    &\lesssim \max_{k\in[K]}\sin\Theta(\mU_{k,\text{true}},\hat\mU_k^{(t)})\\&\lesssim \frac{\sqrt{p\phi}}{\lambda\underline\gamma}+ \left(1\over 2\right)^t\max_{k\in[K]}\sin\Theta(\mU_{k,\text{true}},\hat\mU_k^{(0)})\\&\lesssim\frac{\sqrt{p\phi}}{\lambda\underline\gamma}+ \left(1\over2\right)^t\max_{k\in[K]}\sin\Theta(\mM_{k,\text{true}},\hat\mM_k^{(0)}).
 \end{align}
Finally, the proof is completed applying $\underline\gamma\asymp d^{K/2}$ from [Case 1] and [Case 2].
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:local}]
Combining Proposition~\ref{lem:ini}(b) and \eqref{eq:inim}, we obtain
 \begin{align}
      \max_{k\in[K]}\sin\Theta(\mU_{k,\text{true}},\hat\mU_k^{(t)})\lesssim \frac{\sqrt{p\phi}}{\lambda\underline\gamma}+ \left(1\over 2\right)^t\left(\frac{p^{K/2}}{(\lambda\underline\gamma)^2/\phi}\right),
 \end{align}
 with probability at least $1-\exp(-p)$.
We set $t\gtrsim \log \frac{p^{(K-1)/2}}{\lambda\underline\gamma}$ to make the second term negligible.
 Therefore, the first part of proof is completed by noticing that \[ \frac{p^{(K-1)/2}}{\lambda\underline\gamma}\lesssim\log \frac{p^{(K-1)/2}}{\lambda d^{K/2}}\lesssim K\log  p,\] where the first inequality uses $\underline\gamma\asymp d^{K/2}$ from [Case 1] and [Case 2], and the last inequality is from the condition $\lambda/\sqrt{\phi}\geq Cp^{K/4} d^{-K/2}$.

 For the estimation error with respect to Frobenius norm, direct application of Lemma~\ref{lem:lts}(c) with $t\gtrsim K\log p\gtrsim \log\frac{p^{(K-1)/2}}{\lambda\underline\gamma}$ yields
 \begin{align}\label{eq:sbound}
     \FnormSize{}{\hat\tS^{(t)}-\tS_{\text{true}}}^2\lesssim \phi(r^K+Kpr),
 \end{align}
 with probability at least $1-\exp(-p).$
Notice that
\begin{align}\label{eq:sbbound}
    \FnormSize{}{\hat\tS^{(t)}-\tS_{\text{true}}}^2 &= \FnormSize{}{\left(\hat\tB^{(t)}-\tB_{\text{true}}\right)\times\{\mR_1,\ldots \mR_K\}}^2\nonumber\\&\geq{\underline{\gamma}}^2 \FnormSize{}{\hat\tB^{(t)}-\tB_{\text{true}}}^2\nonumber\\&\gtrsim d^{K}\FnormSize{}{\hat\tB^{(t)}-\tB_{\text{true}}}^2, \quad \text{from [Case 1] and [Case 2]}.
\end{align}
Combining \eqref{eq:sbound} and \eqref{eq:sbbound} completes the proof.
\end{proof}

\subsection{Auxiliary Lemmas}\label{sec:lemma}
\begin{proof}[Proof of Proposition~\ref{prop}]For ease of presentation, we drop the subscript $(i_1,\ldots,i_K)$ and simply write $\varepsilon$ ($=y-b'(\theta)$). For any given $t\in\mathbb{R}$, we have
\begin{align}
\mathbb{E}(\exp(t\varepsilon|\theta))&=\int c(x) \exp\left({\theta x - b(\theta)\over \phi}   \right)\exp \left(t(x-b'(\theta))\right)dx\\
&=\int c(x)\exp \left( {(\theta + \phi t)x - b (\theta+\phi t)+b(\theta+\phi t)-b(\theta)-\phi t b'(\theta) \over \phi}\right)dx\\
&=\exp\left( {b(\theta+\phi t)-b(\theta)-\phi t b'(\theta) \over \phi} \right)\\
&\leq \exp\left(\phi U t^2\over 2 \right),
\end{align}
where $c(\cdot)$ and $b(\cdot)$ are known functions in the exponential family corresponding to $y$, and the last line uses the fact that $\sup_{\theta\in\mathbb{R}}b''(\theta)\leq U$. 
Therefore, $\varepsilon$ is sub-Gaussian-$(\phi U)$. \end{proof}

\begin{defn}[$\alpha$-convexity] A real-valued function $f\colon \tS \to\mathbb{R}$ is called $\alpha$-convex, if
\[
f(x_1)\geq f(x_2)+\langle \nabla_xf(x_2),\ x_1-x_2\rangle+{\alpha}\FnormSize{}{x_1-x_2}^2,\ \text{for all }x_1,x_2\in\tS.
\]
\end{defn}
\begin{lem}[Convexity under linear transformation]\label{lem:convex}
Suppose $f\colon \mathbb{R}^{d\times \cdots\times d}\to \mathbb{R}$ is a $\alpha$-convex function. Define a function $g\colon \mathbb{R}^{p\times \cdots \times p}\to \mathbb{R}$ by
$g(\tB)=f(\tB\times \{\mX_1,\ldots,\mX_K\})$ for all $\tB\in\mathbb{R}^{p\times \cdots\times p}$. Then, $g$ is a $(\underline{\gamma}^2\alpha)$-convex function. 
\end{lem}
\begin{proof}[Proof of Lemma~\ref{lem:convex}]
By the definition of $\alpha$-convexity, we have
\begin{equation}\label{eq:f}
f(\Theta_1)\geq f(\Theta_2) + \langle \nabla_\Theta f(\Theta_2), \Theta_1-\Theta_2\rangle + \alpha\FnormSize{}{\Theta_1-\Theta_2}^2, \ \text{for all }\Theta_1,\Theta_2\in\mathbb{R}^{d\times \cdots\times d},
\end{equation}
where $\nabla_\Theta f(\cdot)$ denotes the derivative of $f$ with respect to $\Theta\in\mathbb{R}^{d\times \cdots\times d}$. For any $\tB_1, \tB_2\in\mathbb{R}^{p\times \cdots \times p}$, we notice that $\tB_i\times\{\mX_1,\ldots,\mX_K\}\in\mathbb{R}^{d\times \cdots \times d}$ for $i=1,2$. Applying~\eqref{eq:f} to this setting gives
\begin{align}\label{eq:1}
&f(\tB_1\times\{\mX_1,\ldots,\mX_K\}) \notag \\
\geq &\ f(\tB_2\times\{\mX_1,\ldots,\mX_K\})+\langle  \nabla_\Theta f(\tB_2\times\{\mX_1,\ldots,\mX_K\}), (\tB_1-\tB_2)\times\{\mX_1,\ldots,\mX_K\} \rangle \notag \\
&\ +  \alpha\FnormSize{}{(\tB_1-\tB_2)\times\{\mX_1,\ldots,\mX_K\}}^2\notag \\
\geq &\ f(\tB_2\times\{\mX_1,\ldots,\mX_K\})+\langle  \nabla_\Theta f(\tB_2\times\{\mX_1,\ldots,\mX_K\})\times\{\mX^T_1,\ldots,\mX^T_K\}, (\tB_1-\tB_2)\rangle \notag \\
&\ + \alpha \underline{\gamma}^2  \FnormSize{}{\tB_1-\tB_2}^2.
\end{align}
By the definition of $g$ and the linearity from $\tB$ to $\Theta$, we have
\begin{equation}\label{eq:2}
\nabla g_\tB(\tB_2)=\nabla f_\Theta (\tB_2\times\{\mX_1,\ldots,\mX_K\}) \times\{\mX^T_1,\ldots,\mX^T_K\}.
\end{equation}
The convexity of $g$ directly follows by plugging~\eqref{eq:2} into~\eqref{eq:1},
\[
g(\tB_1)\geq g(\tB_2)  +\langle \nabla g_\tB(\tB_2), \tB_1-\tB_2\rangle + \alpha \underline{\gamma}^2  \FnormSize{}{\tB_1-\tB_2}^2.
\]
\end{proof}

\begin{proof}[Proof of Proposition~\ref{lem}] We first prove the strong concavity by viewing the log-likelihood as a function of the linear predictor $\Theta$. Write
\[
\bar \tL(\Theta)=\langle \tY, \Theta \rangle -\sum_{i_1,\ldots,i_K}b(\theta_{i_1,\ldots,i_K}).
\]
Direct calculation shows that the Hession of $\bar \tL(\Theta)$ can be expressed as
\[
{\partial \bar \tL(\Theta) \over \partial \theta_{i_1,\ldots,i_K}\partial \theta_{j_1,\ldots,j_K}}=
\begin{cases}
 -b''(\theta_{i_1,\ldots,i_K})<-L<0,\quad \text{if } (i_1,\ldots,i_K)= (j_1,\ldots,j_K),\\
 0, \quad \text{otherwise},
 \end{cases}
\]
Therefore, the Hession matrix of $\bar \tL(\Theta)$ is strictly negative definite with eigenvalues upper bounded by $-L<0$. By Taylor expansion, $-\bar \tL(\Theta)$ is $L/2$-convex with respect to $\Theta$. Note that $\bar \tL(\Theta)=\tL(\tB)$ via the linear mapping $\Theta=\tB\times\{\mX_1,\ldots,\mX_K\}$. Therefore, by Lemma~\ref{lem:convex}, $\tL(\tB)$ is $(\gamma^2L/2)$-convex with respect to $\tB$.

To prove the second part of Proposition~\ref{lem}, we note
\[
\langle \nabla \tL(\trueB), \tB\rangle=\langle \nabla \bar\tL(\trueT)\times\{\mX^T_1,\ldots,\mX^T_K\},\ \tB \rangle =\langle \tY-b'(\trueT),\ \tB\times\{\mX_1,\dots,\mX_K\}\rangle.
 \]
By Proposition~\ref{prop}, $\tY-b'(\trueT)$ is a random tensor consisting of i.i.d.\ sub-Gaussian-($U\phi$) entries under Assumption 2. We write $\tE=\tY-b'(\trueT)$ and consider the sub-Gaussian maxima
\[
\text{Err}_{\textup{ideal}}(\mr)=\sup_{\FnormSize{}{\tB}=1, \tB\in\tP(r)}\langle \tE, \tB\times\{\mX_1,\ldots,\mX_K\}\rangle.
\]
The quantity $\text{Err}_{\textup{ideal}}(\mr)$ is closely related to the localized Gaussian width~\citep{chen2019non,han2020optimal} that measures the model complexity of $\tP(\mr)$. By adapting \citet[Lemma E.5]{han2020optimal} in our context, we have
\[
\text{Err}_{\textup{ideal}}(\mr) \lesssim  \sqrt{\phi U(r^K+Kpr)}\prod_{k\in[K]} \sigma_{\max}({\mX_k}) \leq \bar \gamma  \sqrt{\phi U(r^K+Kpr)},
\]
with probability at least $1-\exp(-p)$. 
\end{proof}

The following Lemma is adopted from~\citet[Theorem 6.1]{wang2017tensor} in our contexts. 
\begin{lem}[Wedin's $\sin\Theta$ Theorem]\label{prop:sinebound}
Let $\mB$ and $\hat\mB$ be two $m\times n$ real matrix SVDs $\mB = \mU\Sigma\mV^T$ and $\hat\mB = \hat\mU\hat\mSigma\hat\mV^T.$ If $\sigma_{\text{min}}(\mB)>0$ and $\FnormSize{}{\hat\mB-\mB}\ll \sigma_{\text{min}}(\mB)$, then
\begin{align}\label{eq:sine}
    \text{sin} \Theta(\mU,\hat\mU) \leq \frac{\sigma_{\max}(\hat\mB-\mB)}{\sigma_{\text{min}}(\mB)}\leq \frac{\FnormSize{}{\hat\mB-\mB}}{\sigma_{\text{min}}(\mB)}.
\end{align}
\end{lem}

The following theorem~\cite{zhang2018tensor} provides the statistical guarantees for unsupervised tensor decomposition based on alternating least square algorithm. For simplicity, we consider the balanced dimension $p_1 =\cdots=p_K =p$ and $r_1 =\cdots=r_K =r$. 
\begin{lem}[Theorem 1 in \cite{zhang2018tensor}]\label{lem:lts}
Consider the Gaussian tensor model
\begin{align}
    \tY  = \tS_{\text{true}} + \tE,
\end{align}
where $\tS_{\text{true}} = \tC_{\text{true}}\times\{\mU_{1,\text{true}},\ldots,\mU_{K,\text{true}}\}$ is an unknown signal tensor, $\tC_{\text{true}}\in\mathbb{R}^{r\times\cdots\times r}$ is a full rank core tensor, $\mU_{k,\text{true}}\in\mathbb{O}(p,r)$ are orthornomal matrices, and $\tE\in\mathbb{R}^{p\times \cdots \times p}$ is a Gaussian noise tensor consisting of i.i.d entries from $N(0,\sigma)$. Let $\lambda$ denote the smallest singular value of matrices $\text{Unfold}_k(\tS_{\text{true}})$ over all possible $k$, 
\[
\lambda' =\min_{k\in[K]}\sigma_{\min}(\text{Unfold}_k(\tS_{\text{true}})).
\]
Then, the following two properties hold whenever $\lambda'/\sigma\geq C_{\textup{gap}}p^{K/4}$ for some universal constant $C_{\textup{gap}}>0$.

\begin{enumerate}[label=(\alph*)]
\item  With probability at least $1- \exp(-p)$, the spectral initialization $\hat\mU_k^{(0)}$ has
\begin{align}
   \max_{k\in[K]}\sin \Theta(\mU_{k,\text{true}},\hat\mU_k^{(0)})\leq c \frac{p^{K/2}}{\lambda'^2/\sigma^2},
 \end{align} 
 for some constant $c>0.$

%  and 
%  \begin{align}
%      \FnormSize{}{\hat\tS^{(0)}-\tS_{\text{true}}}\lesssim \sigma\sqrt{r^K+Kpr}+r\sqrt{p}\left(\frac{\sqrt{p}}{\lambda'/\sigma}+\frac{p^{K/2}}{\lambda'^2/\sigma^2}\right).
%  \end{align}
  \item Let $t = 1,2, \ldots,$ denote the  iteration in HOOI algorithm.  With probability at least $1- \exp(-p)$, the alternating optimization $\hat\mU_k^{(t)}$ satisfies
 \begin{align}\label{eq:lt}
      \max_{k\in[K]}\sin \Theta(\mU_{k,\text{true}},\hat\mU_k^{(t)})\lesssim \frac{\sqrt{p}}{\lambda'/\sigma}+ \left({1\over 2}\right)^t\max_{k\in[K]}\sin \Theta(\mU_{k,\text{true}},\hat\mU_k^{(0)}),
 \end{align}
 \item When $t\gtrsim \log {p^{(K-1)/2}\over \lambda'}$, the tensor estimate $\hat \tS^{(t)}$ from HOOI satisfies
 \[
\FnormSize{}{\hat \tS^{(t)}-\tS_{\text{true}}}^2\lesssim  \sigma^2 (r^K+Kpr),
 \]
 with probability at least $1-\exp(-p)$. 
 \end{enumerate}
\end{lem}

% \begin{proof}
% Angle distance is directly from Theorem 1 \cite{zhang2018tensor}.
% We prove the estimation error with respect to Frobenius norm.
% Define $L_0 = \max_{k\in[K]}\sin(\Theta(\mU_{k,\text{true}},\hat\mU_k^{(0)}))$ and $L_t = \max_{k\in[K]}\sin(\Theta(\mU_{k,\text{true}},\hat\mU_k^{(t)}))$ for $t = 1,2,\ldots$
% We have the following decomposition, 
% \begin{align}\label{eq:decomp}
% \FnormSize{}{\hat\tS^{(t)}-\tS_{\text{true}}}&\leq \FnormSize{}{\tS_{\text{true}}\times \{\mP_{\hat\mU^{(t)}_1},\ldots \mP_{\hat\mU^{(t)}_K}\}-\tS_{\text{true}}}+\FnormSize{}{\tE\times \{\mP_{\hat\mU^{(t)}_1},\ldots \mP_{\hat\mU^{(t)}_K}\}}
% \end{align}

% By (37) in \cite{zhang2018tensor}, the first term is bounded by
% \begin{align}
%     \FnormSize{}{\tS_{\text{true}}\times \{\mP_{\hat\mU^{(t)}_1},\ldots \mP_{\hat\mU^{(t)}_K}\}-\tS_{\text{true}}}\leq C K(\sqrt{pr}+r\sqrt{p}L_t).
% \end{align}
% Slight modification on the proof of Proposition~\ref{lem}(b) yields
% \begin{align}
%     \FnormSize{}{\tE'\times \{\mP_{\hat\mU^{(t)}_1},\ldots \mP_{\hat\mU^{(t)}_K}\}} &\leq  \sup_{\|\tT\|_F = 1,\tT\in\tP(r)}\langle \tE,\tT\times \{\hat\mU^{(t)},\ldots,\hat\mU^{(t)}\}\rangle\\&\leq C\sigma \sqrt{r^{K}+ Kpr},
% \end{align}
% with probability at least $
% 1-\exp(-p).$
% Finally, combining the above two bounds into \eqref{eq:decomp} gives us the estimation bound in (b),
% \begin{align}\label{eq:ltf}
%     \FnormSize{}{\hat\tS^{(t)}-\tS_{\text{true}}}&\leq C K(\sqrt{pr}+r\sqrt{p}L_t)+C\sigma\sqrt{r^K+Kpr}\\&\stackrel{\eqref{eq:lt}}{\lesssim}  r\sqrt{p}\left(C_2\sqrt{rp}\over \lambda'/\sigma\right)^tL_0+\sigma\sqrt{r^K+Kpr}.
% \end{align}
% Plugging $t = 0$  into \eqref{eq:ltf} gives us the bound in (a).
% \end{proof}

\begin{lem}[Angle distance under linear transformation]\label{lem:sint}
Let $\mU$ and $\hat\mU$ be two $m\times n$ real matrices where $m>n$.  Let $\mR$ be an $m\times m$ invertible matrix. If $\sin\Theta(\mU,\hat\mU)\leq L$ for some constant $L\in[0,1]$, then
\begin{align}
    \sin \Theta(\mR\mU,\mR\hat\mU)\leq\left(\sigma_{\max}(\mR)\over \sigma_{\min}(\mR)\right)^2 L.
\end{align}
\end{lem}
\begin{proof}
Suppose that orthonormal basis of $\text{Span}(\mU)$  and $\text{Span}(\hat\mU^{\perp})$ are $\{\mu_1,\ldots,\mu_n\}$ and $\{\nu_{n+1},\ldots,\nu_{m}\}$ respectively. By definition, 
\begin{align}
    \sin \Theta(\mU,\hat\mU) = \max_{\sum_{i=1}^n a_i^2 = \sum_{j=n+1}^m b_j^2 = 1} \left\langle \sum_{i=1}^n a_i\mu_i,\sum_{j=n+1}^m b_j\nu_j\right\rangle \leq L.
\end{align}
We write $\mx = \mR\sum_{i=1}^n a_i\mu_i$ and $\my = \mR\sum_{j=n+1}^m b_j\nu_j$ for any $\mx\in \text{Span}(\mR\mU)$ and $\my\in\text{Span}((\mR\hat\mU)^{\perp})$.
Then,
\begin{align}
    {\langle \mx,\my\rangle\over \|\mx\|_2\|\my\|_2} &= {\langle  \mR\sum_{i=1}^n a_i\mu_i, \mR\sum_{j=n+1}^m b_j\nu_j\rangle\over \|\mR\sum_{i=1}^n a_i\mu_i\|_2\|\mR\sum_{j=n+1}^m b_j\nu_j\|_2 }\\&\leq {\sigma_{\max}(\mR^T\mR)\langle \sum_{i=1}^n a_i\mu_i,\sum_{j=n+1}^m b_j\nu_j\rangle\over \sigma_{\min}^2(\mR)\sqrt{ \sum_{i=1}^n a_i^2}\sqrt{\sum_{j=n+1}^m b_j^2} }\\&\leq \left(\sigma_{\max}(\mR)\over \sigma_{\min}(\mR)\right)^2 \sin\Theta(\mU,\hat\mU).
\end{align}

\end{proof}

\section{Additional simulation results}\label{sec:asimulation}

\subsection{Detailed simulation setup for Figure~\ref{fig:noniid}a-b}

We generate data from \textbf{Envelope} model \citep{li2017parsimonious} with slight modification. We simulate response tensor $\tY\in\mathbb{R}^{d\times d\times d}$ from the following model with envelope dimension $(u_1, u_2)$,
\begin{align}\label{eq:e}
    \tY|\mX &= \tB \times_3 \mX + \tE = \tC \times \{\mGamma_1, \mGamma_2,\mX\}+ \tE, \notag \\
    \text{with} \quad \tE &\sim \mathcal{TN}(\mSigma_1, \mSigma_2, \mI), \quad \mSigma_k = \mGamma_k \mOmega_k \mGamma_k^T + \mGamma_{0k} \mOmega_{0k} \mGamma_{0k}^T+\mI, \quad k = 1,2,
\end{align}
 where $\mX\in\mathbb{R}^{d\times p}$ is the feature matrix, $\tB=\tC\times\{\mGamma_1,\mGamma_2, \mI\}\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor, $\tC\in\mathbb{R}^{\mu_1\times \mu_2\times p}$ is a full-rank core tensor, $\mathcal{TN}(\cdot,\ \cdot,\ \cdot)$ represents zero-mean tensor normal distribution with Kronecker structured covariance, 
 $\mGamma_{k} \in \mathbb{O}(d,u_k)$ 
 consists of orthogonal columns, $\mGamma_{0k} \in \mathbb{O}(d,d-u_k)$ is the orthogonal complement of $\mGamma_k$, and  $\mOmega_k = \mA_k\mA_k^T$, $\mOmega_{0k} = \mA_{k0}\mA_{k0}^T$ with $\mA_k \in \bbR^{u_k \times u_k}$,  $\mA_{k0} \in \bbR^{(d - u_k) \times (d - u_k)}$.

The entries of $\mX$ are i.i.d.\ drawn from $\tN(0,1)$, the entries of $\mA_k$, $\mA_{k0}$ are i.i.d.\ drawn from $\textup{Uniform}[-\gamma, \gamma]$, and the entries of core tensor $\tC$ are i.i.d.\ drawn from $\textup{Uniform}[-3,  3]$. We call $\gamma$ the \textit{correlation level}. Note that the only distinction between model~\eqref{eq:e} and standard {\bf Envelope} model is the additional identity matrix $\mI$ in the expression of $\mSigma_k$. When $\gamma = 0$, the model~\eqref{eq:e} reduces to our \textbf{STD} model with rank $\mr = (u_1,u_2,p)$. We set $d=20, p=5$ in our simulation.

\subsection{Detailed simulation setup for Figure~\ref{fig:noniid}c-d}

We generate the data from \textbf{GLSNet} model \citep{zhang2018network} with slight modification. We simulate the binary response tensor $\tY \in \{0,1\}^{d \times d \times d}$ from the following model
\begin{align}
    \mathbb{E}[\tY|\mX] = f({\bf 1} \otimes {\bf \Theta} + \tB \times_3 \mX),
\end{align}
where $f(\cdot)$ is the logistic link, $\mX \in \mathbb{O}(d,p)$ is the feature matrix with orthonormal columns, ${\bf \Theta}=\mA \mA^T \in \bbR^{ d \times d }$ is a rank-$R$  intercept matrix, where the entries of $\mA \in \bbR^{d \times R}$ are simulated from i.i.d.\ standard normal. 
Unlike original \textbf{GLSNet} model, we generate joint sparse and low-rank structure to the coefficient tensor $\tB$ as follows.

To generate $\tB$, we firstly generate a low-rank tensor $\tB_0$ as 
\begin{align}
    \tB_0 = \tC \times \mM_1 \times \mM_2 \times \mM_3,
\end{align}
where $\tC \in \bbR^{R \times R \times R}$ is a full-rank core tensor, $\mM_1, \mM_2 \in \bbR^{d \times R}$ and $\mM_3 \in \bbR^{p \times R}$ are the factor matrices with orthonormal columns. We simulate i.i.d.\ uniform entries in $\tC$ and rescale the tensor $\tB_0$ such that $\onorm{\tB_0}_{\max} = 2$. Last, we obtain a sparse $\tB$ by randomly setting $sd^2p$ entries in $\tB_0$ to zero. We call $s$ the \textit{sparsity level} which quantifies the proportion of zero's in $\tB$. Hence, the generated tensor $\tB$ is of sparsity level $s$ and of low-rank $(R,R,R)$. We set $d = 20$, $p = 5$ and consider the combination of rank $R = 2$ (low), $4$ (high) and sparsity level $s = \{0, 0.3,0.5\}$ in the simulation.

\subsection{Comparison with GLMs under stochastic block models}
We investigate the performance of our model under correlated feature effects. We mimic the scenario of brain imaging analysis. A sample of $d_3=50$ networks are simulated, one for each individual. Each network measures the connections between $d_1=d_2=20$ brain nodes. We simulate $p=5$ features for the each of the 50 individuals. These features may represent, for example, age, gender, cognitive score, etc. Recent study has suggested that brain connectivity networks often exhibit community structure represented as a collection of subnetworks, and each subnetwork is comprised of a set of spatially distributed brain nodes. To accommodate this structure, we utilize the stochastic block model~\citep{abbe2017community} to generate the effect size. Specifically, we partition the nodes into $r$ blocks by assigning each node to a block with uniform probability. Edges within a same block are assumed to share the same feature effects, where the effects are i.i.d.\ drawn from $N(0,1)$. We then apply our tensor regression model to the network data using the BIC-selected rank. Note that in this case, the true model rank is unknown; the rank of a $r$-block network is not necessarily equal to matrix rank $r$~\citep{zeng2019multiway}.

\begin{figure}[ht]
\centering
\includegraphics[width=16cm]{compare_GLM.pdf}
\caption{Performance comparison under stochastic block models. The three panels plot the MSE when the response tensors are generated from (a) Gaussian (b) Poisson and (c) Bernoulli models. The $x$-axis represents the number of blocks in the networks. }\label{fig:glm}
\end{figure}

Figure~\ref{fig:glm} compares the MSE of our method with a multiple-response GLM approach. The multiple-response GLM is to regress the dyadic edges, one at a time, on the features, and this model is repeatedly fitted for each edge. As we find in Figure~\ref{fig:glm}, our tensor regression method achieves significant error reduction in all three data types considered. The outperformance is substantial in the presence of large communities; even in the less structured case ($\sim 20/15=1.33$ nodes per block), our method still outer-performs GLM. The possible reason is that the multiple-response GLM approach does not account for the correlation among the edges, and suffers from overfitting. In contrast, the low-rankness in our modeling incorporates the shared information across entries. By selecting the rank in a data-driven way, our method achieves accurate estimation in a wide range of settings.

\section{Additional results on data application}\label{sec:adata}
\subsection{Rank selection for Nations data}
Table~\ref{tab:rank_select} summarizes the BIC results in the grid search $\mr\in \{3,4,5\}^3$. We set $r_1=r_2$ due to the symmetry in the dataset. Table~\ref{tab:rank_select} shows that $(r_1,r_2)=(4,4)$ consistently provides the minimal BIC under a range of $r_3$. Because multiple values of $r_3$ give similar BIC, we choose $r_3$ based on the interpretability of the results. Tables~\ref{tab:s1_3}-\ref{tab:s1_5} compare the clustering results for $r_3=3,4,5$. For ease of visualisation, we list only the subset of relations for which the three configurations yield incoherent clustering. We find that the clustering with $r_3 = 4$ (Table~\ref{tab:s1_4}) provides the cleanest results. Table~\ref{tab:s1_3} with $r_3 = 3$ mixes the categories Economics with Organization and Military. Table~\ref{tab:s1_5} with $r_3 = 5$ mixes Economics with Organization, while splitting Military and Territory into different clusters. Therefore, we choose the rank $\mr=(4,4,4)$ in the main paper. The running time for the rank selection via grid search is 95 secs in total, on an iMac macOS High Sierra 10.13.6 with Intel Core i5 3.8 GHz CPU and 8 GB RAM. This indicates the BIC is feasible in the considered setting.

     \begingroup
     \setlength{\tabcolsep}{4.5pt} % Default value: 6pt
    \renewcommand{\arraystretch}{1} % Default value: 1
     \begin{table}[htb]
         \centering
         \begin{tabular}{c||ccc|ccc|ccc}
         $r_3$ & \multicolumn{3}{c|}{ $r_3 = 3$} & \multicolumn{3}{c|}{ $r_3 = 4$} & \multicolumn{3}{c}{ $r_3 = 5$}\\
         \hline
             $(r_1,r_2)$ & $(3,3)$ &  $(4,4)$ &  $(5,5)$ & $(3,3)$ &  $(4,4)$ &  $(5,5)$ &  $(3,3)$ &  $(4,4)$ &  $(5,5)$ \\
             \hline\hline
              % $\ell_{\tY}$& -4733& {\bf -4560}& -4715 & -4919& {\bf -4609} & -4703 & -7347 & {\bf -4700} &-7259 \\
              BIC& 11364 & {\bf 11194} & 11701 & 12275 & {\bf 11897} & 12365 & 17652 & {\bf 12666} & 18146\\
         \end{tabular}
         \caption{BIC results for {\it Nations} data under different tensor rank. Bold number indicates the minimal BIC with a certain $r_3$. }
         \label{tab:rank_select}
     \end{table}
     \endgroup

     \begin{table}[h!]
    \resizebox{1\textwidth}{!}{%
    \begin{tabular}{c|l}
     Cluster& \multicolumn{1}{c}{Relations}\\
    \hline
    \multirow{2}{*}{ I}  & \textcolor{YellowOrange}{exportbooks}, \textcolor{YellowOrange}{relexportbooks}, \textcolor{OliveGreen}{protests}, \textcolor{YellowOrange}{tourism}, \textcolor{YellowOrange}{reltourism}, \textcolor{Blue}{relintergovorgs}\\
    & \textcolor{Blue}{relngo}, \textcolor{Blue}{intergovorgs3}, \textcolor{Blue}{ngoorgs3}, \textcolor{OliveGreen}{militaryalliance},\textcolor{OliveGreen}{commonbloc1} \\
    \hline
    \multirow{2}{*}{ II}  & \textcolor{OliveGreen}{militaryactions}, \textcolor{OliveGreen}{severdiplomatic},\textcolor{OliveGreen}{expeldiplomats}, \textcolor{OliveGreen}{commonbloc0},
    \textcolor{BrickRed}{aidenemy} \\ & \textcolor{BrickRed}{attackembassy},  \textcolor{BrickRed}{lostterritory},    \textcolor{OliveGreen}{blockpositionindex} \\
    \hline
    \multirow{2}{*}{III} & \textcolor{YellowOrange}{tourism3}, \textcolor{YellowOrange}{exports}, \textcolor{YellowOrange}{relexports}, \textcolor{YellowOrange}{exports3}, \textcolor{Blue}{intergovorgs}, \\ & 
    \textcolor{Blue}{ngo} ,\textcolor{OliveGreen}{ embassy}, \textcolor{OliveGreen}{reldiplomacy}, \textcolor{OliveGreen}{commonbloc2} \\
    \hline
     \multicolumn{2}{l}{ \footnotesize \fcolorbox{black}{YellowOrange}{\hspace{2mm}} 
     \raisebox{-1mm}{Economics} \quad   \fcolorbox{black}{OliveGreen}{\hspace{2mm}}  \raisebox{-1mm}{ Military}\quad \fcolorbox{black}{Blue}{\hspace{2mm}} \raisebox{-1mm}{Organization} \quad  \fcolorbox{black}{BrickRed}{\hspace{2mm}} \raisebox{-1mm}{Territory}}
    \end{tabular}
    }
    \caption{$K$-mean relations clustering with $r_3 = 3$. For visualization purpose, only a subset of relations are presented. See texts for details.}\label{tab:s1_3}
    \end{table}

    \begin{table}[H]
    \resizebox{1\textwidth}{!}{%
    \begin{tabular}{c|l}
     Cluster& \multicolumn{1}{c}{Relations}\\
    \hline
    \multirow{1}{*}{ I} &\textcolor{BrickRed}{aidenemy}, \textcolor{BrickRed}{attackembassy}, \textcolor{BrickRed}{lostterritory}\\
    \hline
    \multirow{2}{*}{ II}  &\textcolor{OliveGreen}{militaryactions}, \textcolor{OliveGreen}{severdiplomatic}, \textcolor{OliveGreen}{expeldiplomats}, \textcolor{OliveGreen}{protests}, \\
    & \textcolor{OliveGreen}{commonbloc0}, \textcolor{OliveGreen}{blockpositionindex}, \textcolor{OliveGreen}{commonbloc1}\\
    \hline
    \multirow{1}{*}{III} & \textcolor{Blue}{relintergovorgs}, \textcolor{Blue}{relngo}, \textcolor{Blue}{intergovorgs3}, \textcolor{Blue}{ngoorgs3}, \textcolor{OliveGreen}{militaryalliance}, \textcolor{OliveGreen}{commonbloc2}\\
    \hline
    \multirow{2}{*}{IV} &\textcolor{YellowOrange}{exportbooks}, \textcolor{YellowOrange}{relexportbooks}, \textcolor{YellowOrange}{tourism}, \textcolor{YellowOrange}{reltourism}, \textcolor{YellowOrange}{tourism3} \\
    &\textcolor{YellowOrange}{exports}, \textcolor{YellowOrange}{relexports}, \textcolor{YellowOrange}{exports3}, \textcolor{Blue}{intergovorgs}, \textcolor{Blue}{ngo}, \textcolor{OliveGreen}{embassy}, \textcolor{OliveGreen}{reldiplomacy}\\
    \hline
    \multicolumn{2}{l}{ \footnotesize   \fcolorbox{black}{YellowOrange}{\hspace{2mm}} \raisebox{-1mm}{Economics} \quad 
 \fcolorbox{black}{OliveGreen}{\hspace{2mm}}    \raisebox{-1mm}{Military}\quad \fcolorbox{black}{Blue}{\hspace{2mm}}  \raisebox{-1mm}{Organization} \quad  \fcolorbox{black}{BrickRed}{\hspace{2mm}}  \raisebox{-1mm}{Territory}}
    \end{tabular}
    }
    \caption{$K$-mean relations clustering with $r_3 = 4$. For visualization purpose, only a subset of relations are presented. See texts for details.}\label{tab:s1_4}
    \end{table}

    \begin{table}[H]
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{c|l}
     Cluster& \multicolumn{1}{c}{Relations}\\
    \hline
    \multirow{2}{*}{ I} &\textcolor{YellowOrange}{exportbooks}, \textcolor{YellowOrange}{relexportbooks}, \textcolor{YellowOrange}{tourism}, \textcolor{YellowOrange}{reltourism}, \textcolor{YellowOrange}{tourism3}, \textcolor{YellowOrange}{exports}, \textcolor{YellowOrange}{relexports}, \textcolor{YellowOrange}{exports3}\\
    &\textcolor{Blue}{intergovorgs}, \textcolor{Blue}{relintergovorgs}, \textcolor{Blue}{ngo}, \textcolor{Blue}{relngo}, \textcolor{Blue}{intergovorgs3}, \textcolor{Blue}{ngoorgs3}, \textcolor{OliveGreen}{embassy}, \textcolor{OliveGreen}{reldiplomacy}\\
    \hline
    \multirow{1}{*}{ II}  &\textcolor{BrickRed}{attackembassy}\\
    \hline
    \multirow{1}{*}{III} & \textcolor{OliveGreen}{commonbloc0}, \textcolor{OliveGreen}{blockpositionindex}\\
    \hline
    \multirow{1}{*}{IV} & \textcolor{OliveGreen}{militaryalliance}, \textcolor{OliveGreen}{commonbloc2}\\
    \hline
    \multirow{2}{*}{V} & \textcolor{OliveGreen}{militaryactions}, \textcolor{OliveGreen}{severdiplomatic}, \textcolor{OliveGreen}{expeldiplomats}, \textcolor{BrickRed}{aidenemy}, \textcolor{BrickRed}{lostterritory},\\ & \textcolor{OliveGreen}{protests}, \textcolor{OliveGreen}{commonbloc1}\\
    \hline
    \multicolumn{2}{l}{ \footnotesize \fcolorbox{black}{YellowOrange}{\hspace{2mm}} \raisebox{-1mm}{Economics} \quad   \fcolorbox{black}{OliveGreen}{\hspace{2mm}}   \raisebox{-1mm}{Military}\quad \fcolorbox{black}{Blue}{\hspace{2mm}} \raisebox{-1mm}{Organization} \quad 
    \fcolorbox{black}{BrickRed}{\hspace{2mm}} \raisebox{-1mm}{Territory}}
    \end{tabular}
    }
    \caption{$K$-mean relations clustering with $r_3 = 5$. For visualization purpose, only a subset of relations are presented. See texts for details.}\label{tab:s1_5}
    \end{table}

\subsection{Comparison with unsupervised decomposition}\label{sec:compare1}

We compare the supervised vs.\ unsupervised decomposition in the {\it Nations} data analysis. Table~\ref{tab:unsup} shows the clustering results based on classical unsupervised Tucker decomposition without the feature matrices. Table~\ref{tab:s1} shows the clustering results based on supervised tensor decomposition ({\bf STD}). Compared with supervised decomposition, the unsupervised clustering loses some interpretation. Similar relations {\it exports} and {\it relexports}, {\it ngo} and {\it relngo} are separated into different clusters.

\begin{table}[!h]
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{c|l}
     Cluster& \multicolumn{1}{c}{Relations}\\
    \hline
    \multirow{2}{*}{ I}  &\textcolor{YellowOrange}{economicaid}, \textcolor{YellowOrange}{releconomicaid}, \textcolor{YellowOrange}{exportbooks}, \textcolor{YellowOrange}{relexportbooks}, \textcolor{OliveGreen}{weightedunvote}, \textcolor{OliveGreen}{unweightedunvote}, \\
    &\textcolor{YellowOrange}{tourism}, \textcolor{YellowOrange}{reltourism}, \textcolor{YellowOrange}{tourism3}, \textcolor{YellowOrange}{exports}, \textcolor{Blue}{intergovorgs}, \textcolor{Blue}{ngo}, \textcolor{OliveGreen}{militaryalliance}\\
    \hline
    \multirow{2}{*}{ II}  & \textcolor{OliveGreen}{warning}, \textcolor{OliveGreen}{violentactions}, \textcolor{OliveGreen}{militaryactions}, \textcolor{OliveGreen}{duration}, \textcolor{OliveGreen}{severdiplomatic}, \textcolor{OliveGreen}{expeldiplomats}, \textcolor{YellowOrange}{boycottembargo}, \textcolor{BrickRed}{aidenemy}, \\
    &\textcolor{OliveGreen}{negativecomm}, \textcolor{Blue}{accusation}, \textcolor{OliveGreen}{protests}, \textcolor{OliveGreen}{unoffialacts}, \textcolor{BrickRed}{attackembassy}, \textcolor{Blue}{relemigrants}, \textcolor{OliveGreen}{timesincewar}, \textcolor{BrickRed}{lostterritory}, \textcolor{OliveGreen}{dependent}\\
    \hline
    III & \textcolor{OliveGreen}{timesinceally}, \textcolor{OliveGreen}{independence}, \textcolor{OliveGreen}{commonbloc0}, \textcolor{OliveGreen}{blockpositionindex}\\
    \hline
    \multirow{3}{*}{IV}  &\textcolor{YellowOrange}{treaties}, \textcolor{YellowOrange}{reltreaties}, \textcolor{YellowOrange}{officialvisits}, \textcolor{YellowOrange}{conferences}, \textcolor{YellowOrange}{booktranslations}, \textcolor{YellowOrange}{relbooktranslations}\\
    &\textcolor{OliveGreen}{negativebehavior}, \textcolor{Blue}{nonviolentbehavior}, \textcolor{Blue}{emigrants}, \textcolor{Blue}{emigrants3}, \textcolor{Blue}{students}, \textcolor{Blue}{relstudents}, \textcolor{YellowOrange}{relexports}, \textcolor{YellowOrange}{exports3}\\
    & \textcolor{Blue}{relintergovorgs}, \textcolor{Blue}{relngo}, \textcolor{Blue}{intergovorgs3}, \textcolor{Blue}{ngoorgs3}, \textcolor{OliveGreen}{embassy}, \textcolor{OliveGreen}{reldiplomacy}, \textcolor{OliveGreen}{commonbloc1}, \textcolor{OliveGreen}{commonbloc2}\\
    \hline
    \multicolumn{2}{l}{ \footnotesize \fcolorbox{black}{YellowOrange}{\hspace{2mm}} \raisebox{-1mm}{Economics} \quad   \fcolorbox{black}{OliveGreen}{\hspace{2mm}}   \raisebox{-1mm}{Military}\quad \fcolorbox{black}{Blue}{\hspace{2mm}} \raisebox{-1mm}{Organization} \quad 
    \fcolorbox{black}{BrickRed}{\hspace{2mm}} \raisebox{-1mm}{Territory}}
    \end{tabular}
    }
    \caption{Clustering of relations based on unsupervised tensor decomposition. }\label{tab:unsup}
    \end{table}

\begin{table}[!h]
\resizebox{\textwidth}{!}{%
\begin{tabular}{c|l}
 Category & \multicolumn{1}{c}{Relations}\\
\hline
\multirow{2}{*}{ I}   & \textcolor{OliveGreen}{warning}, \textcolor{OliveGreen}{violentactions}, \textcolor{OliveGreen}{militaryactions}, \textcolor{OliveGreen}{duration}, \textcolor{OliveGreen}{negativebehavior}, \textcolor{OliveGreen}{protests}, \textcolor{OliveGreen}{severdiplomatic} \\
&\textcolor{OliveGreen}{timesincewar}, \textcolor{OliveGreen}{commonbloc0}, \textcolor{OliveGreen}{commonbloc1}, \textcolor{OliveGreen}{blockpositionindex},  \textcolor{OliveGreen}{expeldiplomats}\\   
\hline
\multirow{2}{*}{II}& \textcolor{Blue}{emigrants}, \textcolor{Blue}{emigrants3}, \textcolor{Blue}{relemigrants}, \textcolor{Blue}{accusation}, \textcolor{Blue}{nonviolentbehavior}, \textcolor{Blue}{ngoorgs3}, \textcolor{OliveGreen}{commonbloc2}, \textcolor{Blue}{intergovorgs3}\\      
&\textcolor{YellowOrange}{releconomicaid}, \textcolor{Blue}{relintergovorgs}, \textcolor{Blue}{relngo},  \textcolor{Blue}{students}, \textcolor{Blue}{relstudents}, \textcolor{YellowOrange}{economicaid}, \textcolor{OliveGreen}{negativecomm}, \textcolor{OliveGreen}{militaryalliance} \\
\hline
\multirow{3}{*}{ III} & \textcolor{YellowOrange}{treaties}, \textcolor{YellowOrange}{reltreaties}, \textcolor{YellowOrange}{officialvisits}, \textcolor{YellowOrange}{exportbooks}, \textcolor{YellowOrange}{relexportbooks}, \textcolor{YellowOrange}{booktranslations}, \textcolor{YellowOrange}{relbooktranslations}\\
&\textcolor{YellowOrange}{boycottembargo}, \textcolor{OliveGreen}{weightedunvote}, \textcolor{OliveGreen}{unweightedunvote}, \textcolor{YellowOrange}{reltourism}, \textcolor{YellowOrange}{tourism}, \textcolor{YellowOrange}{tourism3}, \textcolor{YellowOrange}{exports}, \textcolor{YellowOrange}{exports3}  \\
&\textcolor{YellowOrange}{relexports}, \textcolor{Blue}{intergovorgs}, \textcolor{Blue}{ngo}, \textcolor{OliveGreen}{embassy}, \textcolor{OliveGreen}{reldiplomacy}, \textcolor{OliveGreen}{timesinceally}, \textcolor{OliveGreen}{independence}, \textcolor{YellowOrange}{conferences}, \textcolor{OliveGreen}{dependent}\\
\hline
IV &\textcolor{BrickRed}{aidenemy}, \textcolor{BrickRed}{lostterritory}, \textcolor{OliveGreen}{unoffialacts}, \textcolor{BrickRed}{attackembassy}\\
\hline
\multicolumn{2}{l}{ \footnotesize \fcolorbox{black}{YellowOrange}{\hspace{2mm}} \raisebox{-1mm}{Economics} \quad   \fcolorbox{black}{OliveGreen}{\hspace{2mm}}   \raisebox{-1mm}{Military}\quad \fcolorbox{black}{Blue}{\hspace{2mm}} \raisebox{-1mm}{Organization} \quad 
    \fcolorbox{black}{BrickRed}{\hspace{2mm}} \raisebox{-1mm}{Territory}}
\end{tabular}
}
\caption{Clustering of relations based on supervised tensor decomposition. }\label{tab:s1}
\end{table}

\subsection{How different are supervised vs.\ unsupervised factors in general?}\label{sec:compare2}
It is helpful to realize that the unsupervised and methods address different aspects of the problem. The unsupervised decomposition identifies factors that explain most variation in the tensor, whereas the supervised decomposition identifies factors that are most attributable to side features.

We provide a simple example here for illustration.
\begin{example}
Consider the following data tensor $\tY$ and one-sided feature matrix $\mX$, 
\begin{align}
    \tY = \me_1\otimes \me_1 \otimes \me_1+10\me_2\otimes \me_2 \otimes \me_2, \quad \mX=\me_1,
\end{align}
where $\me_i=(0,\ldots,0, 1,0,\ldots,0)^T$ is the $i$th canonical basis vector in $\mathbb{R}^d$ for $i=1,2$. 
Now, consider the unsupervised vs.\ supervised decomposition of $\tY$ with rank $\mr = (1,1,1)$. 
Then, the top supervised and unsupervised factors are perpendicular to each other, 
\[
\mM_{\text{sup},k} \perp \mM_{\text{unsup},k}, \quad \text{for all } k=1,2,3,
\]
where $\mM_{\text{sup},k}$, $\mM_{\text{unsup},k}$ denote the mode-$k$ factors from supervised and unsupervised decompositions, respectively. 
\end{example}

\begin{rmk}
This example shows complementary information between factors from supervised vs.\ unsupervised decompositions. In general, one could construct examples such that these two methods return {\bf arbitrarily different} factors.   
\end{rmk}

\newpage
\bibliography{tensor_wang}
\bibliographystyle{apalike}

\end{document}

### 2021-06-21 tensor_regress/JCGS/Revision/Code/Figure2.R

@@ -1,234 +1,0 @@
-# Code for Figure 2. Log-likelihood trajectory for three models under different (d,r) combinations
-
-# load dependencies ---------------------------------------------------------------------------------------------
-# for plotting
-library(ggplot2)
-library(patchwork)
-
-# for simulation
-install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
-library(tensorregress)
-
-library(pracma)
-source("simulation.R")
-
-# start simulation ------------------------------------------------------------------------------------------------
-
-set.seed(0)
-model <- c("normal", "binary", "poisson")
-dimen <- c(30, 25)
-rank <- c(6, 3)
-
-lglk_sum <- list()
-true_lglk_sum <- list()
-time_sum <- list()
-
-
-for (i in 1:3) { # three data types
-  lglk <- list()
-  true_lglk <- c()
-  time <- c()
-  k <- 1
-  for (j in 1:2) { # two possible dimensions
-    for (r in 1:2) { # two possible ranks
-      if (i == 1) {
-        signal <- 3
-      } else if (i == 2) {
-        signal <- 10
-      } else if (i == 3) signal <- 6
-      my_data <- sim_data(seed = NA, rep(dimen[j], 3), rep(rank[r], 3), p = rep(floor(0.4 * dimen[j]), 3), dist = model[i], dup = 1, signal = signal)
-      ptm <- proc.time()
-      my_reg <- tensor_regress(tsr = my_data$tsr[[1]], X_covar1 = my_data$X_covar1, X_covar2 = my_data$X_covar2, X_covar3 = my_data$X_covar3, rep(rank[r], 3), niter = 20, cons = "non", dist = model[i], traj_long = TRUE, initial = "random")
-      time1 <- proc.time() - ptm
-      time[k] <- time1[3] # running time
-      true_lglk[k] <- loglike(my_data$tsr[[1]], my_data$U, dist = model[i])
-      lglk[[k]] <- my_reg$lglk
-      k <- k + 1
-    }
-  }
-  lglk_sum[[i]] <- lglk
-  true_lglk_sum[[i]] <- true_lglk
-  time_sum[[i]] <- time
-}
-
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Code/Figure3.R

@@ -1,133 +1,0 @@
-# Code for Figure 3. Convergence of MSE with respect to dimension for three models under different ranks
-
-# load dependencies ---------------------------------------------------------------------------------------------
-# for plotting
-library(ggplot2)
-
-# for simulation
-install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
-library(tensorregress)
-
-library(pracma)
-source("simulation.R")
-
-# start plotting with saved data from prior simulation ---------------------------------------------------------------
-# if a new run of simulation is desired, please go to line #110 to run the code --------------------------------------
-
-# parameters on all three modes
-d_pre <- c(30, 35, 40, 45, 50, 55, 60)
-p_pre <- round(0.4 * d_pre)
-c_pre <- c(2, 4, 6)
-d_range <- cbind(d_pre, d_pre, d_pre)
-p_range <- cbind(p_pre, p_pre, p_pre)
-c_range <- cbind(c_pre, c_pre, c_pre)
-
-# MSE vs Effective Sample Size in Normal model ------------------------------
-
-# load pre-saved data
-load("presaved/Figure3_normal.RData")
-table <- F1_normal
-
-MSE_matrix <- apply(table[[1]], c(1, 2), mean)
-sd_matrix <- apply(table[[1]], c(1, 2), sd)
-dl <- length(d_pre)
-cl <- length(c_pre)
-res <- cbind(rep(d_pre, cl), rep(p_pre, cl), rep(c_pre, rep(dl, cl)), c(MSE_matrix), c(sd_matrix))
-res <- data.frame(res)
-colnames(res) <- c("d", "p", "r", "MSE", "sd")
-
-fun.1 <- function(x) max(1.2 * res[, 4] * res[, 1]^2) / x
-
-pdf("Figure3_normal.pdf", width = 5, height = 4)
-figure <- ggplot(res, aes(x = d^3 / d, y = MSE)) +
-  geom_line(aes(color = as.factor(r)), size = 1) +
-  geom_point(size = 1) +
-  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
-  labs(x = "effective sample size", y = "mean squared error (MSE)")
-
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Code/Figure4.R

@@ -1,485 +1,0 @@
-# Code for Figure 4: ME comparison among STD, Envelope, GLSNet, mRRR, and SupCP
-
-# load dependencies ---------------------------------------------------------------------------------------------
-# for plotting
-library(ggplot2)
-library(patchwork)
-
-# for simulation
-install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
-library(tensorregress)
-
-library(TRES) 
-library(rrpack)
-library(rmatio)
-
-library(pracma)
-library(rTensor)
-source("software/netglm_code_adjust.R")
-source("simulation.R")
-
-# start plotting with saved data from prior simulation ---------------------------------------------------------------
-# if a new run of simulation is desired, please go to line #200 to run the code --------------------------------------
-
-new_color <- c("#069AA0", "#7B533E", "#BCA455", "#CCC591", "#D6CFC4")
-
-# ME vs Number of modes with available   features ----------------------------
-
-# load pre-saved data
-load("presaved/Figure4_mode.RData")
-
-final <- 1 - final_mode_cor[, , , c(1, 5, 2, 4, 3)]
-finalsd <- final_sd_mode_cor[, , , c(1, 5, 2, 4, 3)]
-
-s <- 1
-r <- 1
-data <- data.frame(PMSE = c(final[s, r, 1, ], final[s, r, 2, ], final[s, r, 3, ]), sd = c(finalsd[s, r, 1, ], finalsd[s, r, 2, ], finalsd[s, r, 3, ]) / sqrt(30), Method = rep(c("STD (Our method)", "SupCP", "Envelope", "mRRR", "GLSNet"), 3), Category = c(rep(1, 5), rep(2, 5), rep(3, 5)))
-data[, 3] <- factor(data[, 3], levels = c("STD (Our method)", "SupCP", "Envelope", "mRRR", "GLSNet"))
-p1 <- ggplot(data = data, aes(x = as.factor(Category), y = PMSE, fill = Method)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = PMSE - sd, ymax = PMSE + sd), width = .2, position = position_dodge(.9)) +
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Code/Figure5.R

@@ -1,493 +1,0 @@
-# Code for Figure 5: ME comparison among STD, Envelope, GLSNet, mRRR, and SupCP with binary data 
-
-# load dependencies ---------------------------------------------------------------------------------------------
-# for plotting
-library(ggplot2)
-library(patchwork)
-
-# for simulation
-install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
-library(tensorregress)
-
-library(TRES) 
-library(rrpack)
-library(rmatio)
-
-library(pracma)
-library(rTensor)
-source("software/netglm_code_adjust.R")
-source("simulation.R")
-
-
-# start plotting with saved data from prior simulation ---------------------------------------------------------------
-# if a new run of simulation is desired, please go to line #200 to run the code --------------------------------------
-
-new_color <- c("#069AA0", "#7B533E", "#BCA455", "#CCC591", "#D6CFC4")
-
-# ME vs Number of modes with available features ----------------------------
-
-# load pre-saved data
-load("presaved/Figure5_mode.RData")
-
-final <- 1 - final_mode_cor[, , , c(1, 5, 2, 4, 3)]
-finalsd <- final_sd_mode_cor[, , , c(1, 5, 2, 4, 3)]
-
-s <- 1
-r <- 1
-data <- data.frame(PMSE = c(final[s, r, 1, ], final[s, r, 2, ], final[s, r, 3, ]), sd = c(finalsd[s, r, 1, ], finalsd[s, r, 2, ], finalsd[s, r, 3, ]) / sqrt(30), Method = rep(c("STD (Our method)", "SupCP", "Envelope", "mRRR", "GLSNet"), 3), Category = c(rep(1, 5), rep(2, 5), rep(3, 5)))
-data[, 3] <- factor(data[, 3], levels = c("STD (Our method)", "SupCP", "Envelope", "mRRR", "GLSNet"))
-p1 <- ggplot(data = data, aes(x = as.factor(Category), y = PMSE, fill = Method)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = PMSE - sd, ymax = PMSE + sd), width = .2, position = position_dodge(.9)) +
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Code/Figure6.R

@@ -1,360 +1,0 @@
-# Code for Figure 6: ME comparison of STD and other methods under model misspecification
-# load dependencies ---------------------------------------------------------------------------------------------
-# for plotting
-library(ggplot2)
-library(patchwork)
-
-# for simulation
-install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
-library(tensorregress)
-
-library(TRES) 
-library(rrpack)
-library(rmatio)
-
-library(pracma)
-library(rTensor)
-source("software/netglm_code_adjust.R")
-source("simulation.R")
-
-
-
-# start plotting with saved data from prior simulation ---------------------------------------------------------------
-# if a new run of simulation is desired, please go to line #184 to run the code --------------------------------------
-
-# model misspecification with non-i.i.d. noise ----------------------------
-
-# load pre-saved data
-load("presaved/Figure6_noniid.RData")
-
-# ME vs Correlation_level
-new_color <- c("#069AA0", "#BCA455", "#7B533E")
-
-final <- 1 - final_c_cor
-finalsd <- final_sd_c_cor
-
-s <- 1 # rank
-r <- 1 # signal
-
-data <- data.frame(PMSE = c(final[s, r, 1, ], final[s, r, 2, ], final[s, r, 3, ], final[s, r, 4, ]), sd = c(finalsd[s, r, 1, ], finalsd[s, r, 2, ], finalsd[s, r, 3, ], finalsd[s, r, 4, ]) / sqrt(30), Method = rep(c("STD (Our method)", "Envelope", "SupCP"), 4), Category = c(rep(0, 3), rep(0.2, 3), rep(0.3, 3), rep(0.4, 3)))
-data[, 3] <- factor(data[, 3], levels = c("STD (Our method)", "Envelope", "SupCP"))
-p1 <- ggplot(data = data, aes(x = as.factor(Category), y = PMSE, fill = Method)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = PMSE - sd, ymax = PMSE + sd), width = .2, position = position_dodge(.9)) +
-  labs(x = "Correlation level", y = "Misalignment error") +
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Code/Figure8.R

@@ -1,66 +1,0 @@
-# Code for Figure 8. Nations data analysis
-
-# load dependencies ---------------------------------------------------------------------------------------------
-install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
-library(tensorregress)
-
-library(lattice)
-library(RColorBrewer)
-
-library(pracma)
-source("simulation.R")
-
-# load data -----------------------------------------------------------------------------------------------------
-data <- load("rawdata/nations.RData")
-
-# start analysis ------------------------------------------------------------------------------------------------
-set.seed(0)
-seed <- 0
-
-tsr <- R
-tsr[is.na(tsr)] <- 0
-X <- cov
-
-# If a new run of rank selection is desired, run the following two lines
-# rank_range=valid_rank(expand.grid(c(3:5),c(3:5),c(3:5)))
-# res=sele_rank(tsr,X_covar1=cov,X_covar2=cov,X_covar3=NULL,rank_range=rank_range,niter=10,cons = 'non',dist="binary",initial = "random")
-
-core_shape <- c(4, 4, 4)
-result <- tensor_regress(tsr, X_covar1 = cov, X_covar2 = cov, X_covar3 = NULL, core_shape, niter = 10, cons = "penalty", lambda = 0.1, alpha = 10, solver = "CG", dist = "binary", initial = "random")
-
-save(result, file = "presaved/output_nations.RData")
-clustering <- kmeans(result$W[[3]], 4, nstart = 5)
-relnames <- names(tsr[1, 1, ])
-relnames[clustering$cluster == 1]
-relnames[clustering$cluster == 2]
-relnames[clustering$cluster == 3]
-relnames[clustering$cluster == 4]
-
-
-# start plotting ------------------------------------------------------------------------------------------------
-my.palette <- brewer.pal(n = 11, name = "RdBu")
-myPanel <- function(x, y, z, ...) {
-  panel.levelplot(x, y, z, ...)
-  # use handy matrix indexing
-  panel.text(x, y, round(M[cbind(x, y)], 2)) 
-}
-
-pdf("Figure8_1.pdf", width = 6, height = 6)
-M <- result$C_ts[, , which(relnames == "warning")]
-levelplot(M, panel = myPanel, col.regions = my.palette, at = seq(-1.6, 1.6, length.out = 12), main = "Relation: warning")
-dev.off()
-
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Code/software/SupCP-master/Example.m

@@ -1,16 +1,0 @@
-%%Simple simulation for supervised parafac (cp) method
-
-Utrue=20*randn(30,2);
-Vtrue=randn(10,2);
-Vtrue=normc(Vtrue); 
-Wtrue=randn(10,2);
-Wtrue=normc(Wtrue);
-Ztrue=randn(10,2);
-Ztrue=normc(Ztrue);
-Xtrue=TensProd({Utrue,Vtrue,Wtrue,Ztrue});
-X=  Xtrue+5*randn(30,10,10,10); %%add noise
-Y=Utrue(:,1)+randn(30,1); %Y is vector related to one of the signals in U
-
-args = struct('convg_thres',10^(-5),'max_niter',2000,'AnnealIters',1000)
-[B,V,U,se2,Sf,rec]=SupParafacEM(Y,X,2,args);
-

### 2021-06-21 tensor_regress/JCGS/jcgs_submission/jcgs-tensor.tex
clean

--- 2020-09-22 tensor_regress/JCGS_submission/jcgs-tensor/jcgs-tensor.tex
@@
   \title{\bf Supervised Tensor Decomposition with Interactive Side Information}
-  \author{Jiaxin Hu, Zhuoyan Xu, Chanwoo Lee, and Miaoyan Wang\thanks{The authors gratefully acknowledge NSF grant DMS-1915978 and funding from the Wisconsin Alumni Research Foundation.
-}\hspace{.2cm}\\
+  \author{Jiaxin Hu, Chanwoo Lee, and Miaoyan Wang\thanks{The authors gratefully acknowledge NSF grant DMS-1915978 and funding from the Wisconsin Alumni Research Foundation. }\hspace{.2cm}\\
     Department of Statistics, University of Wisconsin-Madison}

### 2021-06-21 tensor_regress/JCGS/Revision/Others/Code/Figure4.R

@@ -1,384 +1,0 @@
-####### Code for Figure 4: ME comparison among STD, Envelope, GLSNet, mRRR, and SupCP ######
-
-############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line  to run the code ##########
-library(ggplot2)
-library(patchwork)
-
-new_color = c("#069AA0","#7B533E","#BCA455","#CCC591","#D6CFC4")
-
-######## plot 1 - Correlation for five methods under different (signal, rank) setting #######
-
-#### info number vs 1 - cor
-load("presaved/fig4_mode.RData")
-
-final = 1 - final_mode_cor[,,,c(1,5,2,4,3)]
-finalsd = final_sd_mode_cor[,,,c(1,5,2,4,3)]
-
-s=1;r=1;
-data=data.frame(PMSE=c(final[s,r,1,],final[s,r,2,],final[s,r,3,]),sd=c(finalsd[s,r,1,],finalsd[s,r,2,],finalsd[s,r,3,])/sqrt(30),Method=rep(c('STD (Our method)',"SupCP","Envelope","mRRR","GLSNet"),3),Category=c(rep(1,5),rep(2,5),rep(3,5)))
-data[,3]=factor(data[,3],levels=c('STD (Our method)',"SupCP","Envelope","mRRR","GLSNet"))
-p1=ggplot(data=data, aes(x=as.factor(Category),y=PMSE, fill=Method))+geom_bar(stat="identity", position=position_dodge())+geom_errorbar(aes(ymin=PMSE-sd, ymax=PMSE+sd), width=.2,position=position_dodge(.9))+labs(x="Number of modes with available features",y="Misalignment error")+coord_cartesian(ylim = c(0, 1)) +  labs(title = "Low Signal, Low Rank",size = 5) +theme(plot.title = element_text(hjust = 0.5,size = 11))+
-  scale_fill_manual(values=new_color)+theme(axis.text=element_text(size=12),axis.title=element_text(size=10)) 
-#p1
-
-s=1;r=2;
-data=data.frame(PMSE=c(final[s,r,1,],final[s,r,2,],final[s,r,3,]),sd=c(finalsd[s,r,1,],finalsd[s,r,2,],finalsd[s,r,3,])/sqrt(30),Method=rep(c('STD (Our method)',"SupCP","Envelope","mRRR","GLSNet"),3),Category=c(rep(1,5),rep(2,5),rep(3,5)))
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Others/Code/Figure5.R

@@ -1,388 +1,0 @@
-####### Code for Figure 5: ME comparison among STD, Envelope, GLSNet, mRRR, and SupCP with binary data ######
-
-############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line  to run the code ##########
-library(ggplot2)
-library(patchwork)
-
-new_color = c("#069AA0","#7B533E","#BCA455","#CCC591","#D6CFC4")
-
-######## plot 1 - Correlation for five methods under different (signal, rank) setting #######
-
-#### info number vs 1 - cor
-load("presaved/fig5_mode.RData")
-
-final = 1 - final_mode_cor[,,,c(1,5,2,4,3)]
-finalsd = final_sd_mode_cor[,,,c(1,5,2,4,3)]
-
-s=1;r=1;
-data=data.frame(PMSE=c(final[s,r,1,],final[s,r,2,],final[s,r,3,]),sd=c(finalsd[s,r,1,],finalsd[s,r,2,],finalsd[s,r,3,])/sqrt(30),Method=rep(c('STD (Our method)',"SupCP","Envelope","mRRR","GLSNet"),3),Category=c(rep(1,5),rep(2,5),rep(3,5)))
-data[,3]=factor(data[,3],levels=c('STD (Our method)',"SupCP","Envelope","mRRR","GLSNet"))
-p1=ggplot(data=data, aes(x=as.factor(Category),y=PMSE, fill=Method))+geom_bar(stat="identity", position=position_dodge())+geom_errorbar(aes(ymin=PMSE-sd, ymax=PMSE+sd), width=.2,position=position_dodge(.9))+labs(x="Number of modes with available features",y="Misalignment error")+coord_cartesian(ylim = c(0, 1)) +  labs(title = "Low Signal, Low Rank",size = 5) +theme(plot.title = element_text(hjust = 0.5,size = 11))+
-  scale_fill_manual(values=new_color)+theme(axis.text=element_text(size=12),axis.title=element_text(size=10)) 
-#p1
-
-s=1;r=2;
-data=data.frame(PMSE=c(final[s,r,1,],final[s,r,2,],final[s,r,3,]),sd=c(finalsd[s,r,1,],finalsd[s,r,2,],finalsd[s,r,3,])/sqrt(30),Method=rep(c('STD (Our method)',"SupCP","Envelope","mRRR","GLSNet"),3),Category=c(rep(1,5),rep(2,5),rep(3,5)))
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Others/Code/Figure6.R

@@ -1,152 +1,0 @@
-####### Code for Figure 6: Robustness to non-i.i.d. noise ######
-
-############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line  to run the code ##########
-library(ggplot2)
-library(patchwork)
-
-new_color = c("#069AA0","#BCA455","#7B533E")
-
-load("presave/fig6.RData")
-
-#### cor_level vs 1 - Cor
-final = 1 - final_c_cor
-finalsd = final_sd_c_cor
-
-s=1;r=1; # rank, signal
-data=data.frame(PMSE=c(final[s,r,1,],final[s,r,2,],final[s,r,3,]),sd=c(finalsd[s,r,1,],finalsd[s,r,2,],finalsd[s,r,3,])/sqrt(30),Method=rep(c('STD (Our method)',"Envelope","SupCP"),3),Category=c(rep(0,3),rep(0.3,3),rep(0.5,3)))
-data[,3]=factor(data[,3],levels=c('STD (Our method)',"Envelope","SupCP"))
-p1=ggplot(data=data, aes(x=as.factor(Category),y=PMSE, fill=Method))+geom_bar(stat="identity", position=position_dodge())+geom_errorbar(aes(ymin=PMSE-sd, ymax=PMSE+sd), width=.2,position=position_dodge(.9))+labs(x="Correlation level",y="Misallignment error")+coord_cartesian(ylim = c(0, 1)) +  labs(title = "Low Signal, Low Rank",size = 5) +theme(plot.title = element_text(hjust = 0.5,size = 11))+
-  scale_fill_manual(values=new_color)+theme(axis.text=element_text(size=12),axis.title=element_text(size=10))
-p1
-
-s=1;r=2;
-data=data.frame(PMSE=c(final[s,r,1,],final[s,r,2,],final[s,r,3,]),sd=c(finalsd[s,r,1,],finalsd[s,r,2,],finalsd[s,r,3,])/sqrt(30),Method=rep(c('STD (Our method)',"Envelope","SupCP"),3),Category=c(rep(0,3),rep(0.3,3),rep(0.5,3)))
-data[,3]=factor(data[,3],levels=c('STD (Our method)',"Envelope","SupCP"))
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Others/Code/function/SupCP-master/SupParafacEM.m

@@ -1,293 +1,0 @@
-function [B,V,U,se2,Sf,rec]=SupParafacEM(Y,X,R,args)
-% EM algorithm to fit the SupCP model:
-% X=U X V1 X ... X VK + E 
-% U=YB + F 
-%
-%
-% Input
-%       Y           n*q full column rank response matrix (necessarily n>=q)
-%                  
-%        
-%       X           n*p1*...*pK design array
-%      
-%       R          fixed rank of approximation, R<=min(n,p)   
-%
-%       args       struct parameter with optional additional parameters
-%                  args = struct('field1',values1, ...
-%                            'field2',values2, ...
-%                            'field3',values3) .
-%                  The following fields may be specified:
-%
-%           AnnealIters:  Annealing iterations (default =100)
-%
-%           fig: binary argument for whether log likelihood should be
-%           plotted at each iteration (default = 0)
-%
-%           ParafacStart:binary argument for whether to initialize with
-%           Parafac factorization (default = 0)
-%
-%           max_niter: maximum number of iterations (default = 1000)
-%
-%           convg_thres: convergence threshold for difference in log
-%           likelihood (default = 10^(-3))
-%
-%           Sf_diag: whether Sf is diagnal (default =1, diagonal)
-%
-% Output
-%       B           q*r coefficient matrix for U~Y, 
-%                   
-%       V           list of length K-1. V{k} is a pXr coefficient matrix 
-%                    with columns of norm 1
-%
-%       U           Conditional expectation of U: nXr
-%
-%       se2         scalar, var(E)
-%       Sf          r*r diagonal matrix, cov(F)
-%  
-%       rec        log likelihood for each iteration
-%      
-%
-% Created by: Eric Lock (elock@umn.edu) and Gen Li
-
-AnnealIters = 100; %%default to 100 annealing iterations
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Others/Code/supcp.m

@@ -1,251 +1,0 @@
-% supCP vs STD comparison
-
-addpath("function/SupCP-master");
-
-% PMSE/Cor vs info number
-rng(0);
-%load("/Volumes/GoogleDrive/My Drive/research/tensor_regress/jcgs_revision/jcgs_code/test/s_2_r_2_i_5_n_30_fig4_sample.mat");
-load("mat_data/mode_fig5_v1.mat");
-
-dup = 30;
-% store the errors
-err_sup = zeros(2,2,3,dup);
-cor_sup = zeros(2,2,3,dup);
-
-final_mode_sup = zeros(2,2,3);
-final_sd_mode_sup = zeros(2,2,3);
-final_mode_cor_sup = zeros(2,2,3);
-final_sd_mode_cor_sup = zeros(2,2,3);
-
-rank_range = [2,4,6,8,10];
-rank_choose = zeros(2,2,3,dup);
-rank_choose_cor = zeros(2,2,3,dup);
-
-args = struct('max_niter',2000,'AnnealIters',1000);
-for s=1:2 % signal alpha = 3,6
-    for r=1:2 % core shape = (3,3,3) or (4,5,6)
-        for i=1:3 % info number
-            for n=1:dup % dup
-                
-                %s = 2; r = 2; i = 3; n = 30;
-                disp([para{s,r}{i,n}'])
-                Y = y{s,r}{i,n};
-                X = tsr{s,r}{i,n};
-                X_true = x_true{s,r}{i,n};
-                
-                test = zeros(1,length(rank_range)); test_cor = zeros(1,length(rank_range));
-                for(k=1:length(rank_range))
-                    rank = rank_range(k);
-                    disp(['choose rank as ', num2str(rank)]);
-                    
-                    [B,V,U,se2,Sf,rec]=SupParafacEM(Y,X,rank,args);
-                    
-                    % calculate fitted value
-                    x_fit = TensProd([{U}, V]);
-                    test(k) = mean((x_fit - X_true).^2, 'all');
-                    correlation = corrcoef(x_fit,X_true);
-                    test_cor(k) = correlation(1,2);
-                end % rank choose
-                
-                [M,I] = min(test); [M_cor,I_cor] = max(test_cor);
-                rank_choose(s,r,i,n) = rank_range(I);
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Others/old_note_fig_code/code/Figure4.R

@@ -1,114 +1,0 @@
-###### Mreg vs STD comparison ######
-
-## R package rTensor has been removed from the R CRAN. 
-## Install the old version here.
-library(remotes)
-install_version("rTensor", "1.3")
-library("rTensor")
-
-## Mreg code
-source("netglm_code_adjust.R") # original data has a typo in the output of best_SymGLM() 
-library(tensorregress)
-
-
-###### run simulation. PMSE vs. number of informative modes
-set.seed(0)
-seed=0
-## combinations of signal and rank
-signal_range=c(3,6)
-core_range=rbind(c(3,3,3),c(4,5,6))
-##
-dup=10;
-d=20;
-whole_shape = rep(d,3);
-final_mode=array(0,dim=c(2,2,3,2)) # ours vs mreg
-final_sd_mode=array(0,dim=c(2,2,3,2))
-err_res=err_mreg=array(0,dim=c(2,2,3,dup))
-dist="binary";
-
-
-for(s in 1:2){ ## signal level
-  for(r in 1:2){ ## rank
-    for(i in 1:3){ ## number of informative covariates
-      for(n in 1:dup){ ## simulation replicates
-        core_shape=core_range[r,]
-        signal=signal_range[s]
-        cat("signal = ",signal, ", rank = ", core_shape, ", info number = ",i,", dup = ", n,"\n" )
-        if(i==1){ ## one mode covaraite
-          data=sim_data(whole_shape = whole_shape, core_shape=core_shape,p=c(0,0,0.4*d),dist=dist, dup=dup, signal=signal)
-          res = tensor_regress(data$tsr[[n]],X_covar3 = data$X_covar3,core_shape=core_shape,Nsim=10, cons = 'non', dist = dist)### tensor_regress
-        }else if(i==2){ ## two mode covaraite
-          data=sim_data(whole_shape = whole_shape, core_shape=core_shape,p=c(0,0.4*d,0.4*d),dist=dist, dup=dup, signal=signal)
-          res = tensor_regress(data$tsr[[n]],X_covar2 = data$X_covar2,X_covar3 = data$X_covar3,core_shape=core_shape,Nsim=10, cons = 'non', dist = dist)### tensor_regress
-        }else if (i==3){## three model covariate
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Others/old_note_fig_code/code/netglm_code_adjust.R

@@ -1,300 +1,0 @@
-#############################################
-########## this is the code #################
-########## for binary undirected  ###########
-#############################################
-
-### adjusted by Jiaxin Hu on May 8 
-### for the JCGS revision
-
-library(rTensor)
-library(MASS)
-library(gplots)
-library(glmnet)
-library(igraph)
-library(logistf)
-library(mgcv)
-library(RMTstat)
-library(reliaR)
-
-EPS = 1e-3
-
-logit<-function(x){
-	log(x/(1-x))
-}
-
-
-logistic<-function(x){
-	1-1/(1+exp(x))
-}
-
-
-norm_vec <- function(x) sqrt(sum(x^2))
-scale_vec <- function(x) x/sqrt(sum(x^2))
-
-
-
-mytruncate<-function(x,s){
-	xtruncate = rep(0, length(x))
-	xtruncate[ which(rank(-abs(x))<= s)] = x[which(rank(-abs(x))<= s)]
-	xtruncate
-}
-
-#######################################################################
-## This function estimates the low rank parameter only ################
-## gamma_ar is the step size and it needs tuning ######################
-## if gamma_ar is too large, then the algorithm will not converge #####
-## gamma_ar may set to decrease with iter number for faster convergence  
-#######################################################################
-
-mySymGLM0<-function(Z, X, R=null, sparsity=null, niter=50){
-	n<-dim(Z)[1]
-	N<-dim(Z)[3]
-	p<-dim(X)[2]
-    ### initialize ####		
-    Abar<-apply(Z@data,c(1,2),mean)
-    Abar[Abar<=0]<-EPS
-    Abar[Abar>=1]<-1-EPS	
-	c<-rep(1,N)
-    initial<-svd(logit(Abar))
-    w<-initial$d[1:R]*(sign(initial$u[1,1:R])*sign(initial$v[1,1:R]))
-    A<-initial$u[,1:R] 
-
-    iter<-1; cond<-FALSE; 
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/for_chtc/Figure4_chtc_mode.R

@@ -1,111 +1,0 @@
-######### Figure 4 (CHTC version) ########
-rm(list = ls())
-
-install.packages("tensorregress_4.0.tar.gz", repos = NULL, type="source")
-library(tensorregress)
-
-if(!require(TRES)){ # Envelope
-  install.packages(TRES)
-  library(TRES)
-} 
-
-if(!require(rrpack)){ # mRRR
-  install.packages(rrpack)
-  library(rrpack)
-} 
-
-if(!require(rmatio)){ # write .mat file for SupCP
-  install.packages(rmatio)
-  library(ramtio)
-}
-
-source("netglm_code_adjust.R") # GLSNet
-
-source("funcs.R") # useful functions
-
-
-########### run simulation. 1 - Correlation vs number of informative modes #######
-
-####### Gaussian #######
-
-args <- commandArgs(T) # signal, rank, info, dup
-s = as.numeric(args[1]); r = as.numeric(args[2]); 
-i = as.numeric(args[3]); n = as.numeric(args[4])
-
-#s = 1; r = 2; i = 3; n = 24
-seed = 10000*s + 1000*r + 100*i + n
-set.seed(seed)
-
-d=20; whole_shape = rep(d,3); dup = 30
-## combinations of signal and rank
-signal_range=c(3,6)
-core_range=rbind(c(3,3,3),c(4,5,6))
-
-err_res = err_env = err_rrr = err_mreg1 = c()
-cor_res = cor_env = cor_rrr = cor_mreg1 = c()
-
-dist="normal"
-
-
-core_shape=core_range[r,]
-signal=signal_range[s]
-
-if(i==1){ ## one mode covaraite
-  data=sim_data(seed,whole_shape = whole_shape, core_shape=core_shape,p=c(0,0,0.4*d),dist=dist, dup=dup, signal=signal,ortho = T)
-  res = tensor_regress(data$tsr[[n]],X_covar3 = data$X_covar3,core_shape=core_shape, cons = 'non', dist = dist, initial = "QR_tucker")### tensor_regress
-}else if(i==2){ ## two mode covaraite
-  data=sim_data(seed,whole_shape = whole_shape, core_shape=core_shape,p=c(0,0.4*d,0.4*d),dist=dist, dup=dup, signal=signal,ortho = T)
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/for_chtc/Figure4_chtc_sample.R

@@ -1,97 +1,0 @@
-######### Figure 4 (CHTC version) ########
-rm(list = ls())
-
-install.packages("tensorregress_4.0.tar.gz", repos = NULL, type="source")
-library(tensorregress)
-
-if(!require(TRES)){ # Envelope
-  install.packages(TRES)
-  library(TRES)
-} 
-
-if(!require(rrpack)){ # mRRR
-  install.packages(rrpack)
-  library(rrpack)
-} 
-
-if(!require(rmatio)){ # write .mat file for SupCP
-  install.packages(rmatio)
-  library(ramtio)
-}
-
-source("netglm_code_adjust.R") # GLSNet
-
-source("funcs.R") # useful functions
-
-########### run simulation. 1 - Correlation vs effective sample size #######
-
-####### Gaussian #######
-args <- commandArgs(T) # signal, rank, dim, dup
-s = as.numeric(args[1]); r = as.numeric(args[2]); 
-i = as.numeric(args[3]); n = as.numeric(args[4])
-
-seed = 10000*s + 1000*r + 100*i + n
-set.seed(seed)
-
-## combinations of signal and rank
-d=c(20,40,60,80,100); dup = 30
-signal_range=c(3,6)
-core_range=rbind(c(3,3,3),c(4,5,6))
-
-err_res=err_env=err_rrr=err_mreg1=array(0,dim=c(2,2,5,dup))
-
-cor_res=cor_env=cor_rrr=cor_mreg1 =array(0,dim=c(2,2,5,dup))
-
-dist="normal"
-
-whole_shape = c(20,20,d[i]); 
-core_shape=core_range[r,]
-signal=signal_range[s]
-
-data=sim_data(seed,whole_shape = whole_shape, core_shape=core_shape,p=c(0,0,0.4*d[i]),dist=dist, dup=dup, signal=signal, ortho = T)
-res = tensor_regress(data$tsr[[n]],X_covar3 = data$X_covar3,core_shape=core_shape,niter=10, cons = 'non', dist = dist, initial = "QR_tucker")### tensor_regress
-        
-# tensorregress
-err_res= mean((res$U - data$U)^2)
-cor_res= cor(as.vector(res$U), as.vector(data$U))
-
-# envelope
-env_fit = TRR.fit(t(data$X_covar3), data$tsr[[n]], core_shape[1:2], method = "FG")
-
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/for_chtc/Figure5_chtc_mode.R

@@ -1,106 +1,0 @@
-######### Figure 5 (CHTC version) ########
-rm(list = ls())
-
-install.packages("tensorregress_4.0.tar.gz", repos = NULL, type="source")
-library(tensorregress)
-
-if(!require(TRES)){ # Envelope
-  install.packages(TRES)
-  library(TRES)
-} 
-
-if(!require(rrpack)){ # mRRR
-  install.packages(rrpack)
-  library(rrpack)
-} 
-
-if(!require(rmatio)){ # write .mat file for SupCP
-  install.packages(rmatio)
-  library(ramtio)
-}
-
-source("netglm_code_adjust.R") # GLSNet
-
-source("funcs.R") # useful functions
-
-########### run simulation. 1 - Correlation vs number of informative modes #######
-
-
-####### Binary #######
-args <- commandArgs(T) # signal, rank, info, dup
-s = as.numeric(args[1]); r = as.numeric(args[2]); 
-i = as.numeric(args[3]); n = as.numeric(args[4])
-
-seed = 10000*s + 1000*r + 100*i + n
-set.seed(seed)
-
-d=20; whole_shape = rep(d,3); dup = 30
-## combinations of signal and rank
-signal_range=c(3,6)
-core_range=rbind(c(3,3,3),c(4,5,6))
-
-err_res = err_env = err_rrr = err_mreg1 = c()
-cor_res = cor_env = cor_rrr = cor_mreg1 = c()
-
-dist="binary"
-
-core_shape=core_range[r,]
-signal=signal_range[s]
-
-if(i==1){ ## one mode covaraite
-  data=sim_data(seed,whole_shape = whole_shape, core_shape=core_shape,p=c(0,0,0.4*d),dist=dist, dup=dup, signal=signal,ortho = T)
-  res = tensor_regress(data$tsr[[n]],X_covar3 = data$X_covar3,core_shape=core_shape, cons = 'non', dist = dist, initial = "QR_tucker")### tensor_regress
-}else if(i==2){ ## two mode covaraite
-  data=sim_data(seed,whole_shape = whole_shape, core_shape=core_shape,p=c(0,0.4*d,0.4*d),dist=dist, dup=dup, signal=signal,ortho = T)
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/for_chtc/Figure6_chtc.R

@@ -1,66 +1,0 @@
-######## Figure 6 (CHTC version) #######
-rm(list = ls())
-
-install.packages("tensorregress_4.0.tar.gz", repos = NULL, type="source")
-library(tensorregress)
-
-if(!require(TRES)){ # Envelope
-  install.packages("TRES")
-  library(TRES)
-} 
-
-if(!require(pracma)){ # Envelope
-  install.packages("pracma")
-  library(pracma)
-} 
-
-if(!require(rmatio)){ # write .mat file for SupCP
-  install.packages("rmatio")
-  library(ramtio)
-}
-
-source("funcs.R") # useful functions
-
-
-args <- commandArgs(T) # env_dim, signal_level, cor_level, dup
-i = as.numeric(args[1]); j = as.numeric(args[2]); 
-k = as.numeric(args[3]); m = as.numeric(args[4])
-
-seed = 10000*i + 1000*j + 100*k + m
-set.seed(seed)
-
-D = c(20,20); n = 20; p = 5 ;dup = 30
-u_range = rbind(c(3,3), c(4,5))
-signal_level_range = c(3,6)
-cor_level_range = c(0, 0.3 ,0.5) # 0 = i.i.d.
-
-err_res=err_env=c()
-cor_res=cor_env=c()
-
-u = u_range[i,]; signal_level = signal_level_range[j]; cor_level = cor_level_range[k]
-
-# generate data
-dat = env_sim(D, u, p, n, signal_level, cor_level, dup)
-
-### fit envelope 
-env_fit = TRR.fit(t(dat$X), dat$tsr[[m]]@data, u, method = "FG")
-err_env = mean((env_fit$fitted.values@data - dat$U@data)^2)
-cor_env = cor(as.vector(env_fit$fitted.values@data), as.vector(dat$U@data))
-
-#### fit tensorregress
-std_fit = tensor_regress(dat$tsr[[m]]@data, X_covar3 = dat$X, core_shape = c(u,p), niter = 10, cons = "non", dist = "normal", initial = "QR_tucker")
-err_res = mean((std_fit$U - dat$U@data)^2)
-cor_res = cor(as.vector(std_fit$U), as.vector(dat$U@data))
-
-result = data.frame(err = c(err_res,err_env), cor = c(cor_res,cor_env))
-save(result, file = paste0("i_",i,"_j_",j,"_k_",k,"_m_",m,"_fig6.RData") )
-
-# save data to matlab
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/for_chtc/Figure7_chtc.R

@@ -1,47 +1,0 @@
-######## Figure 7 (CHTC version) #######
-rm(list = ls())
-
-install.packages("tensorregress_4.0.tar.gz", repos = NULL, type="source")
-library(tensorregress)
-
-source("netglm_code_adjust.R") # GLSNet
-
-source("funcs.R") # useful functions
-
-
-args <- commandArgs(T) # signal, rank, sparsity, dup
-i = as.numeric(args[1]); j = as.numeric(args[2]); 
-k = as.numeric(args[3]); m = as.numeric(args[4])
-
-#i = 1; j = 2; k = 3; m = 30
-seed = 10000*i + 1000*j + 100*k + m
-set.seed(seed)
-
-d = 20; d3 = 20; p = 8; dup = 30
-signal_range = c(2,4); r_range = c(2,4); s_range = c(0.1,0.5,0.9)
-
-err_res=err_mreg=c()
-cor_res=cor_mreg=c()
-
-signal = signal_range[i]; R = r3 = r_range[j]; s = s_range[k]
-dat = sparse_sim(d, d3, p, R, r3, s, signal, dup)
-
-#### fit tensorregress
-std_fit = tensor_regress(dat$tsr[[m]], X_covar3 = dat$X, core_shape = c(R,R,r3), niter = 20, cons = "non", dist = "binary", initial = "QR_tucker")
-err_res = mean((std_fit$U - dat$U)^2)
-cor_res = cor(as.vector(std_fit$U), as.vector(dat$U))
-
-##### fit GLSNet
-mreg_res = mySymGLM(as.tensor(dat$tsr[[m]]), dat$X, R, sparsity=d*d*p*s, niter=50)
-
-Theta = mreg_res$A%*%diag(mreg_res$w)%*%t(mreg_res$A)
-mreg_B = ttm(mreg_res$BB,dat$X,m = 3)@data
-mreg_U = mreg_B
-for (l in 1:dim(mreg_B)[3]) {
-    mreg_U[,,l] = mreg_B[,,l] + Theta
-}
-err_mreg =  mean((mreg_U - dat$U)^2)
-cor_mreg = cor(as.vector(mreg_U), as.vector(dat$U))
-
-result = data.frame(err = c(err_res,err_mreg), cor = c(cor_res,cor_mreg))
-save(result, file = paste0("i_",i,"_j_",j,"_k_",k,"_m_",m,"_fig7.RData") )

### 2021-06-21 tensor_regress/JCGS/Revision/for_chtc/merge_sample.R

@@ -1,66 +1,0 @@
-#### merge the output file in form paste0("s_",s,"_r_",r,"_i_",i,"_n_",n,"_fig5_sample.RData") 
-final_sample=final_sd_sample=array(0,dim=c(2,2,5,5)) # signal, rank, info, method
-final_sample_cor=final_sd_sample_cor=array(0,dim=c(2,2,5,5)) 
-
-dup = 30
-for (s in 1:2) {
-    for (r in 1:2) {
-        for (i in 1:5){
-
-            #s = 1; r = 1; i = 3
-            err_table = cor_table = c() # std, env, mreg, rrr
-            for (n in 1:dup){
-               # s = 1; r = 1; i = 3; n = 26
-                file = paste0("home/jhu267/jcgs-code/Figure5/sample_output_v1/","s_",s,"_r_",r,"_i_",i,"_n_",n,"_fig5_sample.RData")
-                load(file)
-
-                err_table = rbind(err_table, result$err)
-                cor_table = rbind(cor_table, result$cor)
-            }
-
-            final_sample[s,r,i,1:4] = apply(err_table, 2, mean)
-            final_sd_sample[s,r,i,1:4] = apply(err_table, 2, sd)
-
-            final_sample_cor[s,r,i,1:4] = apply(cor_table, 2, mean)
-            final_sd_sample_cor[s,r,i,1:4] = apply(cor_table, 2, sd)
-        }
-    }
-}
-
-save(final_sample, final_sd_sample, final_sample_cor, final_sd_sample_cor, file = "sample_fig5_v1.RData")
-
-
-####### merge .mat file in form into one big file paste0("s_",s,"_r_",r,"_i_",i,"_n_",n,"_fig5_sample.mat")
-
-if(!require(rmatio)){ # write .mat file for SupCP
-  install.packages("rmatio")
-  library(rmatio)
-}
-
-tsr_list_s = list(); x_list_s = list(); u_list_s = list();para_list_s = list()
-for (s in 1:2) {
-    tsr_list_r = list(); x_list_r = list(); u_list_r = list();para_list_r = list()
-    for (r in 1:r) {
-        tsr_list_i = list(); x_list_i = list(); u_list_i = list();  para_list_i = list()
-        for (i in 1:5) {
-           tsr_list = list(); x_list = list(); u_list = list();  para_list = list()
-           for (n in 1:dup) {
[…]

### 2021-06-21 tensor_regress/code_chtc/Figure4_chtc_sample.R

@@ -1,0 +1,97 @@
+######### Figure 4 (CHTC version) ########
+rm(list = ls())
+
+install.packages("tensorregress_4.0.tar.gz", repos = NULL, type="source")
+library(tensorregress)
+
+if(!require(TRES)){ # Envelope
+  install.packages(TRES)
+  library(TRES)
+} 
+
+if(!require(rrpack)){ # mRRR
+  install.packages(rrpack)
+  library(rrpack)
+} 
+
+if(!require(rmatio)){ # write .mat file for SupCP
+  install.packages(rmatio)
+  library(ramtio)
+}
+
+source("netglm_code_adjust.R") # GLSNet
+
+source("funcs.R") # useful functions
+
+########### run simulation. 1 - Correlation vs effective sample size #######
+
+####### Gaussian #######
+args <- commandArgs(T) # signal, rank, dim, dup
+s = as.numeric(args[1]); r = as.numeric(args[2]); 
+i = as.numeric(args[3]); n = as.numeric(args[4])
+
+seed = 10000*s + 1000*r + 100*i + n
+set.seed(seed)
+
+## combinations of signal and rank
+d=c(20,40,60,80,100); dup = 30
+signal_range=c(3,6)
+core_range=rbind(c(3,3,3),c(4,5,6))
+
+err_res=err_env=err_rrr=err_mreg1=array(0,dim=c(2,2,5,dup))
+
+cor_res=cor_env=cor_rrr=cor_mreg1 =array(0,dim=c(2,2,5,dup))
+
+dist="normal"
+
+whole_shape = c(20,20,d[i]); 
+core_shape=core_range[r,]
+signal=signal_range[s]
+
+data=sim_data(seed,whole_shape = whole_shape, core_shape=core_shape,p=c(0,0,0.4*d[i]),dist=dist, dup=dup, signal=signal, ortho = T)
+res = tensor_regress(data$tsr[[n]],X_covar3 = data$X_covar3,core_shape=core_shape,niter=10, cons = 'non', dist = dist, initial = "QR_tucker")### tensor_regress
+        
+# tensorregress
+err_res= mean((res$U - data$U)^2)
+cor_res= cor(as.vector(res$U), as.vector(data$U))
+
+# envelope
+env_fit = TRR.fit(t(data$X_covar3), data$tsr[[n]], core_shape[1:2], method = "FG")
+
[…]

### 2021-06-21 tensor_regress/code_chtc/Figure5_chtc_mode.R

@@ -1,0 +1,106 @@
+######### Figure 5 (CHTC version) ########
+rm(list = ls())
+
+install.packages("tensorregress_4.0.tar.gz", repos = NULL, type="source")
+library(tensorregress)
+
+if(!require(TRES)){ # Envelope
+  install.packages(TRES)
+  library(TRES)
+} 
+
+if(!require(rrpack)){ # mRRR
+  install.packages(rrpack)
+  library(rrpack)
+} 
+
+if(!require(rmatio)){ # write .mat file for SupCP
+  install.packages(rmatio)
+  library(ramtio)
+}
+
+source("netglm_code_adjust.R") # GLSNet
+
+source("funcs.R") # useful functions
+
+########### run simulation. 1 - Correlation vs number of informative modes #######
+
+
+####### Binary #######
+args <- commandArgs(T) # signal, rank, info, dup
+s = as.numeric(args[1]); r = as.numeric(args[2]); 
+i = as.numeric(args[3]); n = as.numeric(args[4])
+
+seed = 10000*s + 1000*r + 100*i + n
+set.seed(seed)
+
+d=20; whole_shape = rep(d,3); dup = 30
+## combinations of signal and rank
+signal_range=c(3,6)
+core_range=rbind(c(3,3,3),c(4,5,6))
+
+err_res = err_env = err_rrr = err_mreg1 = c()
+cor_res = cor_env = cor_rrr = cor_mreg1 = c()
+
+dist="binary"
+
+core_shape=core_range[r,]
+signal=signal_range[s]
+
+if(i==1){ ## one mode covaraite
+  data=sim_data(seed,whole_shape = whole_shape, core_shape=core_shape,p=c(0,0,0.4*d),dist=dist, dup=dup, signal=signal,ortho = T)
+  res = tensor_regress(data$tsr[[n]],X_covar3 = data$X_covar3,core_shape=core_shape, cons = 'non', dist = dist, initial = "QR_tucker")### tensor_regress
+}else if(i==2){ ## two mode covaraite
+  data=sim_data(seed,whole_shape = whole_shape, core_shape=core_shape,p=c(0,0.4*d,0.4*d),dist=dist, dup=dup, signal=signal,ortho = T)
[…]

### 2021-06-21 tensor_regress/code_chtc/fig4_mode.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+
+# untar your R installation. Make sure you are using the right version!
+tar -xzf R402.tar.gz
+# (optional) if you have a set of packages (created in Part 1), untar them also
+tar -xzf packages.tar.gz
+
+# make sure the script will use your R installation, 
+# and the working directory as its home location
+export PATH=$PWD/R/bin:$PATH
+export RHOME=$PWD/R
+export R_LIBS=$PWD/packages
+
+# $1 signal; $2 rank; $3 info; $4 dup
+# run your script
+Rscript Figure4_chtc_mode.R $1 $2 $3 $4

### 2021-06-21 tensor_regress/code_chtc/fig4_sample.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+
+# untar your R installation. Make sure you are using the right version!
+tar -xzf R402.tar.gz
+# (optional) if you have a set of packages (created in Part 1), untar them also
+tar -xzf packages.tar.gz
+
+# make sure the script will use your R installation, 
+# and the working directory as its home location
+export PATH=$PWD/R/bin:$PATH
+export RHOME=$PWD/R
+export R_LIBS=$PWD/packages
+
+# $1 signal; $2 rank; $3 info; $4 dup
+# run your script
+Rscript Figure4_chtc_sample.R $1 $2 $3 $4

### 2021-06-21 tensor_regress/code_chtc/merge_sample.R

@@ -1,0 +1,66 @@
+#### merge the output file in form paste0("s_",s,"_r_",r,"_i_",i,"_n_",n,"_fig5_sample.RData") 
+final_sample=final_sd_sample=array(0,dim=c(2,2,5,5)) # signal, rank, info, method
+final_sample_cor=final_sd_sample_cor=array(0,dim=c(2,2,5,5)) 
+
+dup = 30
+for (s in 1:2) {
+    for (r in 1:2) {
+        for (i in 1:5){
+
+            #s = 1; r = 1; i = 3
+            err_table = cor_table = c() # std, env, mreg, rrr
+            for (n in 1:dup){
+               # s = 1; r = 1; i = 3; n = 26
+                file = paste0("home/jhu267/jcgs-code/Figure5/sample_output_v1/","s_",s,"_r_",r,"_i_",i,"_n_",n,"_fig5_sample.RData")
+                load(file)
+
+                err_table = rbind(err_table, result$err)
+                cor_table = rbind(cor_table, result$cor)
+            }
+
+            final_sample[s,r,i,1:4] = apply(err_table, 2, mean)
+            final_sd_sample[s,r,i,1:4] = apply(err_table, 2, sd)
+
+            final_sample_cor[s,r,i,1:4] = apply(cor_table, 2, mean)
+            final_sd_sample_cor[s,r,i,1:4] = apply(cor_table, 2, sd)
+        }
+    }
+}
+
+save(final_sample, final_sd_sample, final_sample_cor, final_sd_sample_cor, file = "sample_fig5_v1.RData")
+
+
+####### merge .mat file in form into one big file paste0("s_",s,"_r_",r,"_i_",i,"_n_",n,"_fig5_sample.mat")
+
+if(!require(rmatio)){ # write .mat file for SupCP
+  install.packages("rmatio")
+  library(rmatio)
+}
+
+tsr_list_s = list(); x_list_s = list(); u_list_s = list();para_list_s = list()
+for (s in 1:2) {
+    tsr_list_r = list(); x_list_r = list(); u_list_r = list();para_list_r = list()
+    for (r in 1:r) {
+        tsr_list_i = list(); x_list_i = list(); u_list_i = list();  para_list_i = list()
+        for (i in 1:5) {
+           tsr_list = list(); x_list = list(); u_list = list();  para_list = list()
+           for (n in 1:dup) {
[…]

### 2021-06-21 tensor_regress/code_server/run_F3.sh

@@ -1,15 +1,0 @@
-#!/bin/bash
-
-# untar your R installation. Make sure you are using the right version!
-tar -xzf R351.tar.gz
-# (optional) if you have a set of packages (created in Part 1), untar them also
-tar -xzf packages.tar.gz
-
-# make sure the script will use your R installation,
-# and the working directory as its home location
-export PATH=$PWD/R/bin:$PATH
-export RHOME=$PWD/R
-export R_LIBS=$PWD/packages
-
-# run your script
-Rscript server_F3.R

### 2021-06-21 tensor_regress/code_server/run_F3_11.sh

@@ -1,15 +1,0 @@
-#!/bin/bash
-
-# untar your R installation. Make sure you are using the right version!
-tar -xzf R351.tar.gz
-# (optional) if you have a set of packages (created in Part 1), untar them also
-tar -xzf packages.tar.gz
-
-# make sure the script will use your R installation,
-# and the working directory as its home location
-export PATH=$PWD/R/bin:$PATH
-export RHOME=$PWD/R
-export R_LIBS=$PWD/packages
-
-# run your script
-Rscript server_F3_11.R

### 2021-06-21 tensor_regress/code_server/run_F4.sh

@@ -1,15 +1,0 @@
-#!/bin/bash
-
-# untar your R installation. Make sure you are using the right version!
-tar -xzf R351.tar.gz
-# (optional) if you have a set of packages (created in Part 1), untar them also
-tar -xzf packages.tar.gz
-
-# make sure the script will use your R installation,
-# and the working directory as its home location
-export PATH=$PWD/R/bin:$PATH
-export RHOME=$PWD/R
-export R_LIBS=$PWD/packages
-
-# run your script
-Rscript server_F4.R

### 2021-06-21 tensor_regress/code_server/server_F3.R

@@ -1,25 +1,0 @@
-source("simulation.R")
-library(tensorregress)
-set.seed(0) ## set seed for this simulation
-seed=0
-
-### all covariates
-dup=30
-d_pre=c(30,35,40,45,50,55,60)
-p_pre=round(0.4*d_pre)
-c_pre=c(2,4,6)
-d_range=cbind(d_pre,d_pre,d_pre)
-p_range=cbind(p_pre,p_pre,p_pre)
-c_range=cbind(c_pre,c_pre,c_pre)
-
-
-F1_normal=conv_rate(seed=NA,signal=10,cons="non",c_range=c_range,dist="normal",alpha=10,dup=dup,d_range=d_range,p_range=p_range,naive=FALSE)
-save(F1_normal,file="F1_normal.RData")
-
-F1_binary=conv_rate(seed=NA,signal=10,cons="non",c_range=c_range,dist="binary",alpha=10,dup=dup,d_range=d_range,p_range=p_range,naive=FALSE)
-save(F1_binary,file="F1_binary.RData")
-
-F1_poisson=conv_rate(seed=NA,signal=10,cons="non",c_range=c_range,dist="poisson",alpha=10,dup=dup,d_range=d_range,p_range=p_range,naive=FALSE)
-save(F1_poisson,file="F1_poisson.RData")
-
-

### 2021-06-21 tensor_regress/code_server/server_F3_11.R

@@ -1,25 +1,0 @@
-source("simulation.R")
-library(tensorregress)
-set.seed(0) ## set seed for this simulation
-seed=0
-
-### all covariates
-dup=30
-d_pre=c(30,35,40,45,50,55,60)
-p_pre=round(0.4*d_pre)
-c_pre=c(2,4,6)
-d_range=cbind(d_pre,d_pre,d_pre)
-p_range=cbind(p_pre,p_pre,p_pre)
-c_range=cbind(c_pre,c_pre,c_pre)
-
-
-F1_normal=conv_rate(seed=NA,signal=10,cons="non",c_range=c_range,dist="normal",alpha=10,dup=dup,d_range=d_range,p_range=p_range,naive=FALSE)
-save(F1_normal,file="F1_normal_new.RData")
-
-F1_binary=conv_rate(seed=NA,signal=10,cons="non",c_range=c_range,dist="binary",alpha=10,dup=dup,d_range=d_range,p_range=p_range,naive=FALSE)
-save(F1_binary,file="F1_binary_new.RData")
-
-F1_poisson=conv_rate(seed=NA,signal=10,cons="non",c_range=c_range,dist="poisson",alpha=10,dup=dup,d_range=d_range,p_range=p_range,naive=FALSE)
-save(F1_poisson,file="F1_poisson_new.RData")
-
-

### 2021-06-21 tensor_regress/code_server/server_F4.R

@@ -1,56 +1,0 @@
-source("simulation.R")
-library(tensorregress)
-
-set.seed(0)
-seed=0
-## one-side covariates
-c_range=c(5,5,5) ## selected rank
-c_range=rbind(c_range,c(10,10,5),c(15,15,5))
-optimal_rank=c(5,5,5)
-optimal_rank=rbind(optimal_rank,c(10,10,5),c(14,14,5)) ## to be determined
-dup=30
-
-### simulate data from normal model
-table=table_naive=matrix(0,nrow=3,ncol=dup)
-##############################
-for(s in 1:3){
-    data=sim_data(whole_shape=c(20,20,50),core_shape=c_range[s,],p=c(0,0,5),dist="normal",dup=dup,signal=5,block=c(TRUE,TRUE,FALSE))
-    for(i in 1:dup){
-        result=tensor_regress(tsr=data$tsr[[i]],X_covar1 =data$X_covar1, X_covar2 = data$X_covar2,X_covar3= data$X_covar3, core_shape=optimal_rank[s,], Nsim=10, cons="no", lambda = 0.1, alpha = 1, solver ="CG",dist="normal")
-        table[s,i]=mean((result$C_ts-data$C_ts)^2) ## tensor method
-        naive_C=massive_glm(data$tsr[[i]],data$X_covar3,dist="normal")
-        table_naive[s,i]=mean((naive_C-data$C_ts)^2) ## GLM method
-    }
-}
-save(table,table_naive,file="comparison_normal.RData")
-
-### simulate data from poisson model
-table=table_naive=matrix(0,nrow=3,ncol=dup)
-##############################
-for(s in 1:3){
-    data=sim_data(whole_shape=c(20,20,50),core_shape=c_range[s,],p=c(0,0,5),dist="poisson",dup=dup,signal=5,block=c(TRUE,TRUE,FALSE))
-    for(i in 1:dup){
-        result=tensor_regress(tsr=data$tsr[[i]],X_covar1 =data$X_covar1, X_covar2 = data$X_covar2,X_covar3= data$X_covar3, core_shape=optimal_rank[s,], Nsim=10, cons="no", lambda = 0.1, alpha = 1, solver ="CG",dist="poisson") ## tensor method
-        table[s,i]=mean((result$C_ts-data$C_ts)^2)
-        naive_C=massive_glm(data$tsr[[i]],data$X_covar3,dist="poisson") ## glm method
[…]

### 2021-06-21 tensor_regress/oldcode/new_graph/new_graph.R

@@ -1,232 +1,0 @@
-#### code for new graph
-set.seed(0) ## set seed for the entire simulation
-library(tensorregress)
-library(ggplot2)
-#devtools::install_github("thomasp85/patchwork")
-library(patchwork)
-source("new_graph_bricks.R")
-source("sim_data.R")
-
-model = c("normal","binary","poisson")
-dimen = c(30,25)
-rank = c(6,3)
-#seeds = c(250,612,900)
-
-lglk_sum = list()
-true_lglk_sum = list()
-time_sum = list()
-
-
-
-for (i in 1:3) {
-  lglk = list()
-  true_lglk = c()
-  time = c()
-  
-  k = 1
-  for (j in 1:2) {
-    for (r in 1:2) {
-        if(i==1) signal=3
-        else if(i==2) signal=10
-        else if(i==3) signal=6
-      my_data = sim_data(seed = NA, rep(dimen[j],3), rep(rank[r],3), p = rep(floor(0.4*dimen[j]) ,3),dist = model[i], dup = 1,signal = signal)
-      ptm = proc.time()
-      my_reg =  tensor_regress1(tsr = my_data$tsr[[1]], X_covar1 = my_data$X_covar1, X_covar2 = my_data$X_covar2,X_covar3 = my_data$X_covar3,rep(rank[r],3), Nsim = 20, cons = "non",dist = model[i])
-      time1 = proc.time() - ptm
-      
-      time[k] = time1[3]
-      true_lglk[k] = loglike(my_data$tsr[[1]], my_data$U,dist = model[i])
-      lglk[[k]] = my_reg$lglk
-      
-      k = k+1
-    }
-  }
-  
-  lglk_sum[[i]] = lglk
-  true_lglk_sum[[i]] = true_lglk
-  time_sum[[i]] = time
-  
-}
-
-###### ggplot ######
-
-# do not use loop here, every subfigure has somewhere special
-
-## normal d = 30, r = 6,3
-
-lglk_normal = c(lglk_sum[[1]][[1]] ,lglk_sum[[1]][[2]])
-case_normal=  c(rep("d3r6",33), rep("d3r3",33))
-iter = rep(1:33,times = 2)
-olglk1 = true_lglk_sum[[1]][1]
-olglk2 = true_lglk_sum[[1]][2]
-plotdata = data.frame(lglk = lglk_normal,case = case_normal,iter = iter)
-
[…]

### 2021-06-21 tensor_regress/oldcode/new_graph/new_graph_bricks.R

@@ -1,339 +1,0 @@
-### bricks for new graph
-
-#library(rTensor)
-library(pracma)
-#library(gtools)
-library(MASS)
-library(speedglm)
-#library(fastglm)
-#library(rgl)
-#library("RColorBrewer")
-
-
-
-#######################
-
-######----------   Tensor regression
-
-## This function is indeed the same as tensor_regress in the package tensorregress
-## except the iteration number is forced to be larger than 8
-## even though the algorithm has already converged.
-## This modification is made for better visualization.
-
-tensor_regress1 = function(tsr,X_covar1 = NULL, X_covar2 = NULL,X_covar3 = NULL, core_shape, Nsim=20, cons = c("non","vanilla","penalty"), lambda = 0.1, alpha = 1, solver ="CG",dist = c("binary", "poisson","normal")){
-  
-  
-  tsr = as.tensor(tsr)
-  Y_1 = unfold(tsr, row_idx = 1, col_idx = c(2,3))@data
-  Y_2 = unfold(tsr, row_idx = 2, col_idx = c(1,3))@data
-  Y_3 = unfold(tsr, row_idx = 3, col_idx = c(1,2))@data
-  d1 = dim(tsr)[1] ; d2 = dim(tsr)[2] ; d3 = dim(tsr)[3]
-  r1 = core_shape[1] ; r2 = core_shape[2] ; r3 = core_shape[3]
-  
-  ###  check whether unsupervised on each mode
-  un_m1 = FALSE ; un_m2 = FALSE ; un_m3 = FALSE
-  if(is.null(X_covar1)|(identical(X_covar1,diag(d1)))) {X_covar1 = diag(d1) ; un_m1 = TRUE}
-  if(is.null(X_covar2)|(identical(X_covar2,diag(d2)))) {X_covar2 = diag(d2) ; un_m2 = TRUE}
-  if(is.null(X_covar3)|(identical(X_covar3,diag(d3)))) {X_covar3 = diag(d3) ; un_m3 = TRUE}
-  p1 = dim(X_covar1)[2] ; p2 = dim(X_covar2)[2] ; p3 = dim(X_covar3)[2]
-  
-  
-  if(dist=="binary"){
-    tsr.transform=as.tensor(2*tsr@data-1)
-  }else if(dist=="poisson"){
[…]

### 2021-06-21 tensor_regress/oldcode/new_graph/sim_data.R

@@ -1,125 +1,0 @@
-#' Simulation of tensor regression models
-#'
-#' Generate response tensors with multiple covariates under different simulation models, specifically for tensors with 3 modes
-#' @param seed    seed for generating data
-#' @param whole_shape  a vector containing dimension of the tensor
-#' @param core_shape   a vector containing Tucker rank of the coefficient tensor
-#' @param p            a vector containing numbers of covariates on each mode, see "details"
-#' @param dist         distribution of response tensor, see "details"
-#' @param dup          number of simulated tensors from the same linear predictor
-#' @param signal       a scalar controlling the max norm of the linear predictor
-#' @param block        a vector containing boolean variables, see "details"
-#' @return     a list containing the following:
-#'
-#' \code{tsr} {a list of simulated tensors, with the number of replicates specified by \code{dup}}
-#'
-#' \code{X_covar1}  {covariate on first mode}
-#'
-#' \code{X_covar2}  {covariate on second mode}
-#'
-#' \code{X_covar3}  {covariate on third mode}
-#'
-#' \code{W} {a list of orthogonal coefficient matrices - one for each mode, with the number of columns given by \code{core_shape}}
-#'
-#' \code{G}  {core tensor, with size specified by \code{core_shape}}
-#'
-#' \code{C_ts}  {coefficient tensor, Tucker product of \code{G},\code{A},\code{B},\code{C}}
-#'
-#' \code{U} {linear predictor,i.e. Tucker product of \code{C_ts},\code{X_covar1},\code{X_covar2},\code{X_covar3}}
-#'
-#' @details    By default non-positive entry in \code{p} indicates no covariate on the corresponding mode of the tensor.
-#'
-#'             \code{dist} specifies three distributions of response tensor: binary, poisson and normal distribution.
-#'
[…]

### 2021-06-21 tensor_regress/package/new_pack/tensorregress/R/tensor_regress.R

@@ -1,539 +1,0 @@
-#' Supervised Tensor Decomposition with Interactive Side Information
-#'
-#' Supervised tensor decomposition with interactive side information on multiple modes. Main function in the package. The function takes a response tensor, multiple side information matrices,
-#' and a desired Tucker rank as input. The output is a rank-constrained M-estimate of the core tensor and factor matrices.
-#'
-#'
-#' @param tsr    response tensor with 3 modes
-#' @param X_covar1    side information on first mode
-#' @param X_covar2    side information on second mode
-#' @param X_covar3    side information on third mode
-#' @param core_shape  the Tucker rank of the tensor decomposition
-#' @param niter        max number of iterations if update does not convergence
-#' @param cons        the constraint method, "non" for without constraint, "vanilla" for global scale down at each iteration, "penalty" for adding log-barrier penalty to object function
-#' @param lambda      penalty coefficient for "penalty" constraint
-#' @param alpha       max norm constraint on linear predictor
-#' @param solver      solver for solving object function when using "penalty" constraint, see "details"
-#' @param dist        distribution of the response tensor, see "details"
-#' @param traj_long   if "TRUE", set the minimal iteration number to 8; if "FALSE", set the minimal iteration number to 0
-#' @param initial     initialization of the alternating optimiation, "random" for random initialization, "QR_tucker" for deterministic initialization using tucker decomposition
-#' @return     a list containing the following:
-#'
-#'                  \code{W} {a list of orthogonal factor matrices - one for each mode, with the number of columns given by \code{core_shape}}
-#'
-#'                  \code{G}  {an array, core tensor with the size specified by \code{core_shape}}
-#'
[…]

### 2021-06-21 tensor_regress/JASA_submission/code/Figure2.R
clean up

@@ -1,213 +1,0 @@
-#### code for Figure 2. Log-likelihood trajectory for three models under different (d,r) combinations
-install.packages("tensorregress.tar.gz",repos = NULL, type="source")
-source("simulation.R")
-library(tensorregress)
-library(ggplot2)
-library(patchwork)
-
-
-set.seed(0)## set seed for the entire simulation
-model = c("normal","binary","poisson")
-dimen = c(30,25)
-rank = c(6,3)
-
-################### simulation ##################
-
-lglk_sum = list()
-true_lglk_sum = list()
-time_sum = list()
-
-
-for (i in 1:3) {## three data types
-  lglk = list()
-  true_lglk = c()
-  time = c()
-  k = 1
-  for (j in 1:2) {## two possbiel dimensions
-      for (r in 1:2) {## two possible ranks
-        if(i==1) signal=3
-        else if(i==2) signal=10
-        else if(i==3) signal=6
-      my_data = sim_data(seed = NA, rep(dimen[j],3), rep(rank[r],3), p = rep(floor(0.4*dimen[j]) ,3),dist = model[i], dup = 1,signal = signal)
-      ptm = proc.time()
-      my_reg =  tensor_regress(tsr = my_data$tsr[[1]], X_covar1 = my_data$X_covar1, X_covar2 = my_data$X_covar2,X_covar3 = my_data$X_covar3,rep(rank[r],3), Nsim = 20, cons = "non",dist = model[i],traj_long=TRUE)
-      time1 = proc.time() - ptm
-      time[k] = time1[3] ## running time
-      true_lglk[k] = loglike(my_data$tsr[[1]], my_data$U,dist = model[i])
-      lglk[[k]] = my_reg$lglk
-      k = k+1
-    }
-  }
-  lglk_sum[[i]] = lglk
-  true_lglk_sum[[i]] = true_lglk
-  time_sum[[i]] = time
-  
-}
-
-############ ggplot ##################
-
-#### normal model with d = 30, r = 6,3
-lglk_normal = c(lglk_sum[[1]][[1]] ,lglk_sum[[1]][[2]])
-case_normal=  c(rep("d3r6",33), rep("d3r3",33))
-iter = rep(1:33,times = 2)
-olglk1 = true_lglk_sum[[1]][1]
-olglk2 = true_lglk_sum[[1]][2]
[…]

### 2021-06-21 tensor_regress/JASA_submission/code/Figure6.R

@@ -1,18 +1,0 @@
-### Code for Figure 6. HCP data analysis
-install.packages("tensorregress.tar.gz",repos = NULL, type="source")
-library(tensorregress)
-
-data=load("presaved/HCP.RData")
-tsr=tensor
-X=attr[,4:5]
-levels(X[,2])=c("22-25","26-30","31+","31+") ## three groups
-contrasts(X[,1]) <- contr.sum
-contrasts(X[,2]) <- contr.sum
-X_covar3=model.matrix(~as.factor(X[,1])+as.factor(X[,2]))
-
-core_shape=c(10,10,4)
-## the following line takes quite a while to finish. We do not sugget to run on personal laptop. We implement it in the server.
-result=tensor_regress(tsr,X_covar1=NULL,X_covar2=NULL,X_covar3,core_shape,Nsim=50,cons="non",lambda=0.1,alpha=10,solver="CG",dist="binary")
-
-save(result,file="output_HCP.RData")
-### the matrix for plotting brain connectivity is saved in presaved/data_for_Figure6

### 2021-06-21 tensor_regress/JASA_submission/code/Figure7.R

@@ -1,43 +1,0 @@
-### Code for Figure 7. Comparison between classical GLM and tensor methods in the HCP data anaysis
-source("simulation.R")
-data=load("presaved/HCP.RData")
-############ Method 1: tensor method  ##############
-load("presaved/output_HCP.RData")
-table(attr[,5])
-## age group: 22-25 26-30 31+
-## number of individuals: 35  58    43
-X=attr[,4:5]
-levels(X[,2])=c("22-25","26-30","31+","31+") ## three age groups
-contrasts(X[,1]) <- contr.sum
-contrasts(X[,2]) <- contr.sum
-### intercept, F 1, M,-1; 22-25 [1,0]; 26-30 [0 1]; 30+
-coef=array(0,dim=c(68,68,5))
-coef[,,1:4]=result$C_ts[,,1:4]##intercept, gender, age 22-25, age 26-30
-coef[,,5]=-coef[,,3]-coef[,,4] ## age 30+
-
-
-############## Method 2: Classical GLM  ##############
-tsr=tensor
-X_covar3=model.matrix(~as.factor(X[,1])+as.factor(X[,2])) ## baseline female and age 22-25
-massive=massive_glm(tsr,X_covar3,"binary")
-mass=array(0,dim=c(68,68,5))
-mass[,,1]=massive[,,1]
-mass[,,2]=massive[,,2]
-mass[,,3]=massive[,,3]
-mass[,,4]=massive[,,4]
-mass[,,5]=-massive[,,4]-massive[,,3]
-
-
-### Plot histogram to compare two methods###
-pdf("compare.pdf",width=15,height=5.5)
-par(mfrow=c(2,4))
-hist(c(mass[,,1]),nclass=40,xlab="Effect size",main="Intercept (multi-response GLM)")
-hist(c(mass[,,2]),nclass=40,xlab="Effect size",main="Gender (multi-response GLM)")
-hist(c(mass[,,4]),nclass=40,xlab="Effect size",main="Age 26-30 (multi-response GLM)")
-hist(c(mass[,,5]),nclass=40,xlab="Effect size",main="Age 31+ (multi-response GLM)")
-
-hist(c(coef[,,1]),nclass=40,xlab="Effect size",main="Intercept (STD)")
-hist(c(coef[,,2]),nclass=40,xlab="Effect size",main="Gender (STD)")
[…]

### 2021-06-21 tensor_regress/JASA_submission/code/Figure8.R

@@ -1,55 +1,0 @@
-### Code for Figure 8. Nations data analysis
-install.packages("tensorregress.tar.gz",repos = NULL, type="source")
-library(tensorregress)
-library(lattice)
-library(RColorBrewer)
-my.palette <- brewer.pal(n = 12, name = "RdBu")
-source("simulation.R")
-set.seed(0)
-seed=0
-data=load("presaved/nations.RData")
-tsr=R
-tsr[is.na(tsr)]=0
-X=cov
-
-#rank_range=valid_rank(expand.grid(c(3:5),c(3:5),c(3:5)))
-#res=sele_rank(tsr,X_covar1=cov,X_covar2=cov,X_covar3=NULL,rank_range=rank_range,Nsim=10,cons = 'non',dist="binary")
-
-
-core_shape=c(4,4,4)
-result=tensor_regress(tsr,X_covar1=cov,X_covar2=cov,X_covar3=NULL,core_shape,Nsim=10,cons="penalty",lambda=0.1,alpha=10,solver="CG",dist="binary")
-
-save(result,file="output_nations.RData")
-library(lattice)
-clustering=kmeans(result$W[[3]],4,nstart=5)
-relnames=names(tsr[1,1,])
-relnames[clustering$cluster==1]
-relnames[clustering$cluster==2]
-relnames[clustering$cluster==3]
-relnames[clustering$cluster==4]
-
-
-myPanel <- function(x, y, z,...) {
-    panel.levelplot(x,y,z,...)
-    panel.text(x, y,  round(M[cbind(x,y)],2)) ## use handy matrix indexing
-}
-
-pdf("Figure8.pdf",width=6,height=6)
-M=result$C_ts[,,which(relnames=="warning")]
-levelplot(M,panel=myPanel,col.regions =my.palette ,at=seq(-1.6, 1.6, length.out=12),main="Relation: warning")
-
-
-M=result$C_ts[,,which(relnames=="violentactions")]
-levelplot(M,panel=myPanel,col.regions =my.palette ,at=seq(-1.6, 1.6, length.out=12),main="Relation: violentactions")
-
-
-M=result$C_ts[,,which(relnames=="treaties")]
-levelplot(M,panel=myPanel,col.regions =my.palette ,at=seq(-1.6, 1.6, length.out=12),main="Relation: treaties")
-
-
-
-M=result$C_ts[,,which(relnames=="aidenemy")]
[…]

### 2021-06-21 tensor_regress/JASA_submission/code/SFigure1.R

@@ -1,114 +1,0 @@
-## Code for Supplementary Figure 1. Comparison of mean sqaured prediction error between our method and the other three previous methods
-install.packages("tensorregress.tar.gz",repos = NULL, type="source")
-source("simulation.R")
-source("comparison.R")
-library(tensorregress)
-library(ggplot2)
-
-############# plot figures using saved data from prior simulation. If a new run of simulation is desired, please go to line #40 to run the code ##########
-
-######## plot MSPE for three methods under different (signal, rank) setting #######
-load("presaved/SFigure1.RData")
-library(ggplot2)
-library(patchwork)
-new_color = c("#069AA0","#CCC591","#BCA455","#D6CFC4")
-
-signal_range=c(3,6)
-core_range=rbind(c(3,3,3),c(4,5,6))
-pdf("SFigure1.pdf",width=20.5,height=3)
-s=r=1;
-data=data.frame(PMSE=c(final[s,r,1,],final[s,r,2,],final[s,r,3,])/signal_range[s],sd=c(finalsd[s,r,1,],finalsd[s,r,2,],finalsd[s,r,3,])/signal_range[s],Method=rep(c('STD (Our method)','HOLRR (Low-rank regression)','HOPLS (Partial least square)','TPG (Projection gradient)'),3),Category=c(rep(1,4),rep(2,4),rep(3,4)))
-data[,3]=factor(data[,3],levels=c("STD (Our method)","HOLRR (Low-rank regression)","HOPLS (Partial least square)","TPG (Projection gradient)"))
-p1=ggplot(data=data, aes(x=as.factor(Category),y=PMSE, fill=Method))+geom_bar(stat="identity", position=position_dodge())+geom_errorbar(aes(ymin=PMSE-sd, ymax=PMSE+sd), width=.2,position=position_dodge(.9))+labs(x="Number of modes with available features",y="MSPE")+coord_cartesian(ylim = c(0, max(data[,1])+0.05))+  labs(title = "Low Signal, Low Rank",size = 5) +theme(plot.title = element_text(hjust = 0.5,size = 11))+
-    scale_fill_manual(values=new_color)+theme(axis.text=element_text(size=12),axis.title=element_text(size=10))
-
-s=1;r=2;
[…]

### 2021-06-21 tensor_regress/JASA_submission/code/simulation.R

@@ -1,99 +1,0 @@
-######### massive GLM to estimate the covariate effects ###
-massive_glm=function(response,X,dist){
-    d1=dim(response)[1]
-    d2=dim(response)[2]
-    coe=array(0,dim=c(d1,d2,dim(X)[2]))
-    if(dist=="normal"){
-        for(i in 1:d1){
-            for(j in 1:d2){
-                fit=lm(response[i,j,]~-1+X)
-                coe[i,j,]=coef(fit)
-            }
-        }
-    }
-    else if(dist=="binary"){
-        for(i in 1:d1){
-            for(j in 1:d2){
-                fit=suppressWarnings(glm(response[i,j,]~-1+X,family=binomial("logit")))
-                coe[i,j,]=coef(fit)
-            }
-        }
-    }
-    else if(dist=="poisson"){
-        for(i in 1:d1){
-            for(j in 1:d2){
-                fit=suppressWarnings(glm(response[i,j,]~-1+X,family=poisson("log")))
-                coe[i,j,]=coef(fit)
-            }
-        }
-    }
-    return(coe)
-}
-
-#### valid rank range for core tensor ## the maximum rank should be no smaller than the product of other.
-## Input: rank_range = expand.grid(...)
-valid_rank=function(rank_range){
-    N=dim(rank_range)[1]
-    output=NULL
-    for(i in 1:N){
-        rank=rank_range[i,]
-        rank=sort(rank)
-        if(rank[1]*rank[2]>=rank[3])
-        output=rbind(output,rank_range[i,])
-    }
-    return(output)
-}
-
-### compute loglikelihood under different models
-loglike=function(data,linearpre,dist){
-    if(dist=="binary"){
-        p=plogis(linearpre)## log-likelihood
-        L=dbinom(data,1,p,log=TRUE)
-        return(sum(L))
-    }
-    else if (dist=="normal"){
-        sigma_est=mean((data-linearpre)^2)
-        L=dnorm(data,linearpre,sqrt(sigma_est),log=TRUE)
-        sum(L)##log-likelihood
-    }
-    else if (dist=="poisson"){
-        lambda=exp(linearpre)
-        L=dpois(data,lambda, log = TRUE) ## log-likelihood for Poisson
-        return(sum(L))
-    }
-}
-
-#### wrapped simulation function to estimate covariate effects using tensor methods or massive glm methods
[…]

### 2021-06-21 tensor_regress/JASA_submission/code/table3.R

@@ -1,24 +1,0 @@
-### Code for Table 3. Nations data analysis
-install.packages("tensorregress.tar.gz",repos = NULL, type="source")
-library(tensorregress)
-set.seed(0)
-seed=0
-data=load("presaved/nations.RData")
-tsr=R
-tsr[is.na(tsr)]=0
-X=cov
-
-#rank_range=valid_rank(expand.grid(c(3:5),c(3:5),c(3:5)))
-#res=sele_rank(tsr,X_covar1=cov,X_covar2=cov,X_covar3=NULL,rank_range=rank_range,Nsim=10,cons = 'non',dist="binary")
-
-
-core_shape=c(4,4,4)
-result=tensor_regress(tsr,X_covar1=cov,X_covar2=cov,X_covar3=NULL,core_shape,Nsim=10,cons="penalty",lambda=0.1,alpha=10,solver="CG",dist="binary")
-save(result,file="output_nations.RData")
-clustering=kmeans(result$W[[3]],4,nstart=5)
-relnames=names(tsr[1,1,])
-relnames[clustering$cluster==1]
-relnames[clustering$cluster==2]
-relnames[clustering$cluster==3]
-relnames[clustering$cluster==4]
-

### 2021-06-21 tensor_regress/JASA_submission/code/tensorregress/R/bricks.R

@@ -1,206 +1,0 @@
-#library(rTensor)
-library(pracma)
-#library(gtools)
-library(MASS)
-library(speedglm)
-#library(fastglm)
-#library(rgl)
-#library("RColorBrewer")
-
-
-
-#######################
-######----------   functions for update
-
-## This part used for penalty (Conjugate gradient method)
-
-loss = function(beta,y,X,lambda,alpha,dist){
-  U = X %*% beta
-  L= - loglike(y,U,dist)
-  if(max(abs(U))>alpha) return(Inf)
-  else{
-    L2 =L - lambda * sum(log(1 - (U/ alpha)^2))  ## object function
-    return(c(L2))
-  }
-}
-
-loglike=function(data,linearpre,dist){
-  if(dist=="binary"){
-    p=plogis(linearpre)## log-likelihood
-    L=dbinom(data,1,p,log=TRUE)
-    return(sum(L))
-  }
-  else if (dist=="normal"){
-    sigma_est=mean((data-linearpre)^2)
-    L=dnorm(data,linearpre,sqrt(sigma_est),log=TRUE)
-    sum(L)##log-likelihood
-  }
-  else if (dist=="poisson"){
-    lambda=exp(linearpre)
-    L=dpois(data,lambda, log = TRUE) ## log-likelihood for Poisson
-    return(sum(L))
-  }
-}
-
-loss_gr <- function(beta,y,X,lambda,alpha,dist){
-  U = X %*% beta
-  if(dist=="binary") p=plogis(U)
-  else if (dist=="normal") p=U
-  else if (dist=="poisson") p=exp(U)
-  L_g=t(X) %*% (p - y)
-  penal_g = 2 * lambda * t(X) %*% (U/(alpha^2 - U^2))
-  return(c(L_g) + c(penal_g))
-}
-
-
-glm_modify=function(y,x,dist){
-  if(dist=="binary"){
-    fit1 = tryCatch({
-        speedglm(y~-1+x,family=binomial(link="logit"))
-    }, warning = function(w) {
- fit1=suppressWarnings(speedglm(y~-1+x,family=binomial(link="logit")))
-    }, error = function(e) {
-        return_value="non-convergence"
-    })
-    
-    if(inherits(fit1,"character")==TRUE){
-        return(list(rep(0,dim(as.matrix(x))[2]),-Inf))
-    }else
-    return(list(coef(fit1), logLik(fit1)))
[…]

### 2021-06-21 tensor_regress/JCGS/Response_submission/Code/Figure2.R

@@ -1,234 +1,0 @@
-# Code for Figure 2. Log-likelihood trajectory for three models under different (d,r) combinations
-
-# load dependencies ---------------------------------------------------------------------------------------------
-# for plotting
-library(ggplot2)
-library(patchwork)
-
-# for simulation
-install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
-library(tensorregress)
-
-library(pracma)
-source("simulation.R")
-
-# start simulation ------------------------------------------------------------------------------------------------
-
-set.seed(0)
-model <- c("normal", "binary", "poisson")
-dimen <- c(30, 25)
-rank <- c(6, 3)
-
-lglk_sum <- list()
-true_lglk_sum <- list()
-time_sum <- list()
-
-
-for (i in 1:3) { # three data types
-  lglk <- list()
-  true_lglk <- c()
-  time <- c()
-  k <- 1
-  for (j in 1:2) { # two possible dimensions
-    for (r in 1:2) { # two possible ranks
-      if (i == 1) {
-        signal <- 3
-      } else if (i == 2) {
-        signal <- 10
-      } else if (i == 3) signal <- 6
-      my_data <- sim_data(seed = NA, rep(dimen[j], 3), rep(rank[r], 3), p = rep(floor(0.4 * dimen[j]), 3), dist = model[i], dup = 1, signal = signal)
-      ptm <- proc.time()
-      my_reg <- tensor_regress(tsr = my_data$tsr[[1]], X_covar1 = my_data$X_covar1, X_covar2 = my_data$X_covar2, X_covar3 = my_data$X_covar3, rep(rank[r], 3), niter = 20, cons = "non", dist = model[i], traj_long = TRUE, initial = "random")
-      time1 <- proc.time() - ptm
-      time[k] <- time1[3] # running time
-      true_lglk[k] <- loglike(my_data$tsr[[1]], my_data$U, dist = model[i])
-      lglk[[k]] <- my_reg$lglk
-      k <- k + 1
-    }
-  }
-  lglk_sum[[i]] <- lglk
-  true_lglk_sum[[i]] <- true_lglk
-  time_sum[[i]] <- time
-}
-
[…]

### 2021-06-21 tensor_regress/JCGS/Response_submission/Code/Figure3.R

@@ -1,133 +1,0 @@
-# Code for Figure 3. Convergence of MSE with respect to dimension for three models under different ranks
-
-# load dependencies ---------------------------------------------------------------------------------------------
-# for plotting
-library(ggplot2)
-
-# for simulation
-install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
-library(tensorregress)
-
-library(pracma)
-source("simulation.R")
-
-# start plotting with saved data from prior simulation ---------------------------------------------------------------
-# if a new run of simulation is desired, please go to line #110 to run the code --------------------------------------
-
-# parameters on all three modes
-d_pre <- c(30, 35, 40, 45, 50, 55, 60)
-p_pre <- round(0.4 * d_pre)
-c_pre <- c(2, 4, 6)
-d_range <- cbind(d_pre, d_pre, d_pre)
-p_range <- cbind(p_pre, p_pre, p_pre)
-c_range <- cbind(c_pre, c_pre, c_pre)
-
-# MSE vs Effective Sample Size in Normal model ------------------------------
-
-# load pre-saved data
-load("presaved/Figure3_normal.RData")
-table <- F1_normal
-
-MSE_matrix <- apply(table[[1]], c(1, 2), mean)
-sd_matrix <- apply(table[[1]], c(1, 2), sd)
-dl <- length(d_pre)
-cl <- length(c_pre)
-res <- cbind(rep(d_pre, cl), rep(p_pre, cl), rep(c_pre, rep(dl, cl)), c(MSE_matrix), c(sd_matrix))
-res <- data.frame(res)
-colnames(res) <- c("d", "p", "r", "MSE", "sd")
-
-fun.1 <- function(x) max(1.2 * res[, 4] * res[, 1]^2) / x
-
-pdf("Figures/Figure3_normal.pdf", width = 5, height = 4)
-figure <- ggplot(res, aes(x = d^3 / d, y = MSE)) +
-  geom_line(aes(color = as.factor(r)), size = 1) +
-  geom_point(size = 1) +
-  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
-  labs(x = "effective sample size", y = "mean squared error (MSE)")
-
[…]

### 2021-06-21 tensor_regress/JCGS/Response_submission/Code/Figure5.R

@@ -1,493 +1,0 @@
-# Code for Figure 5: ME comparison among STD, Envelope, GLSNet, mRRR, and SupCP with binary data 
-
-# load dependencies ---------------------------------------------------------------------------------------------
-# for plotting
-library(ggplot2)
-library(patchwork)
-
-# for simulation
-install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
-library(tensorregress)
-
-library(TRES) 
-library(rrpack)
-library(rmatio)
-
-library(pracma)
-library(rTensor)
-source("software/netglm_code_adjust.R")
-source("simulation.R")
-
-
-# start plotting with saved data from prior simulation ---------------------------------------------------------------
-# if a new run of simulation is desired, please go to line #200 to run the code --------------------------------------
-
-new_color <- c("#069AA0", "#7B533E", "#BCA455", "#CCC591", "#D6CFC4")
-
-# ME vs Number of modes with available features ----------------------------
-
-# load pre-saved data
-load("presaved/Figure5_mode.RData")
-
-final <- 1 - final_mode_cor[, , , c(1, 5, 2, 4, 3)]
-finalsd <- final_sd_mode_cor[, , , c(1, 5, 2, 4, 3)]
-
-s <- 1
-r <- 1
-data <- data.frame(PMSE = c(final[s, r, 1, ], final[s, r, 2, ], final[s, r, 3, ]), sd = c(finalsd[s, r, 1, ], finalsd[s, r, 2, ], finalsd[s, r, 3, ]) / sqrt(30), Method = rep(c("STD (Our method)", "SupCP", "Envelope", "mRRR", "GLSNet"), 3), Category = c(rep(1, 5), rep(2, 5), rep(3, 5)))
-data[, 3] <- factor(data[, 3], levels = c("STD (Our method)", "SupCP", "Envelope", "mRRR", "GLSNet"))
-p1 <- ggplot(data = data, aes(x = as.factor(Category), y = PMSE, fill = Method)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = PMSE - sd, ymax = PMSE + sd), width = .2, position = position_dodge(.9)) +
[…]

### 2021-06-21 tensor_regress/JCGS/Response_submission/Code/Figure6.R

@@ -1,360 +1,0 @@
-# Code for Figure 6: ME comparison of STD and other methods under model misspecification
-# load dependencies ---------------------------------------------------------------------------------------------
-# for plotting
-library(ggplot2)
-library(patchwork)
-
-# for simulation
-install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
-library(tensorregress)
-
-library(TRES) 
-library(rrpack)
-library(rmatio)
-
-library(pracma)
-library(rTensor)
-source("software/netglm_code_adjust.R")
-source("simulation.R")
-
-
-
-# start plotting with saved data from prior simulation ---------------------------------------------------------------
-# if a new run of simulation is desired, please go to line #184 to run the code --------------------------------------
-
-# model misspecification with non-i.i.d. noise ----------------------------
-
-# load pre-saved data
-load("presaved/Figure6_noniid.RData")
-
-# ME vs Correlation_level
-new_color <- c("#069AA0", "#BCA455", "#7B533E")
-
-final <- 1 - final_c_cor
-finalsd <- final_sd_c_cor
-
-s <- 1 # rank
-r <- 1 # signal
-
-data <- data.frame(PMSE = c(final[s, r, 1, ], final[s, r, 2, ], final[s, r, 3, ], final[s, r, 4, ]), sd = c(finalsd[s, r, 1, ], finalsd[s, r, 2, ], finalsd[s, r, 3, ], finalsd[s, r, 4, ]) / sqrt(30), Method = rep(c("STD (Our method)", "Envelope", "SupCP"), 4), Category = c(rep(0, 3), rep(0.2, 3), rep(0.3, 3), rep(0.4, 3)))
-data[, 3] <- factor(data[, 3], levels = c("STD (Our method)", "Envelope", "SupCP"))
-p1 <- ggplot(data = data, aes(x = as.factor(Category), y = PMSE, fill = Method)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = PMSE - sd, ymax = PMSE + sd), width = .2, position = position_dodge(.9)) +
-  labs(x = "Correlation level", y = "Misalignment error") +
[…]

### 2021-06-21 tensor_regress/JCGS/Response_submission/Code/Figure8.R

@@ -1,66 +1,0 @@
-# Code for Figure 8. Nations data analysis
-
-# load dependencies ---------------------------------------------------------------------------------------------
-install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
-library(tensorregress)
-
-library(lattice)
-library(RColorBrewer)
-
-library(pracma)
-source("simulation.R")
-
-# load data -----------------------------------------------------------------------------------------------------
-data <- load("rawdata/nations.RData")
-
-# start analysis ------------------------------------------------------------------------------------------------
-set.seed(0)
-seed <- 0
-
-tsr <- R
-tsr[is.na(tsr)] <- 0
-X <- cov
-
-# If a new run of rank selection is desired, run the following two lines
-# rank_range=valid_rank(expand.grid(c(3:5),c(3:5),c(3:5)))
-# res=sele_rank(tsr,X_covar1=cov,X_covar2=cov,X_covar3=NULL,rank_range=rank_range,niter=10,cons = 'non',dist="binary",initial = "random")
-
-core_shape <- c(4, 4, 4)
-result <- tensor_regress(tsr, X_covar1 = cov, X_covar2 = cov, X_covar3 = NULL, core_shape, niter = 10, cons = "penalty", lambda = 0.1, alpha = 10, solver = "CG", dist = "binary", initial = "random")
-
-save(result, file = "presaved/output_nations.RData")
-clustering <- kmeans(result$W[[3]], 4, nstart = 5)
-relnames <- names(tsr[1, 1, ])
-relnames[clustering$cluster == 1]
-relnames[clustering$cluster == 2]
-relnames[clustering$cluster == 3]
-relnames[clustering$cluster == 4]
-
-
-# start plotting ------------------------------------------------------------------------------------------------
-my.palette <- brewer.pal(n = 11, name = "RdBu")
-myPanel <- function(x, y, z, ...) {
-  panel.levelplot(x, y, z, ...)
-  # use handy matrix indexing
-  panel.text(x, y, round(M[cbind(x, y)], 2)) 
-}
-
-pdf("Figures/Figure8_1.pdf", width = 6, height = 6)
-M <- result$C_ts[, , which(relnames == "warning")]
-levelplot(M, panel = myPanel, col.regions = my.palette, at = seq(-1.6, 1.6, length.out = 12), main = "Relation: warning")
[…]

### 2021-06-21 tensor_regress/JCGS/Response_submission/Code/simulation.R

@@ -1,209 +1,0 @@
-# This file collect the bricks for the simulations
-library(pracma)
-
-# massive GLM to estimate the covariate effects
-massive_glm <- function(response, X, dist) {
-  d1 <- dim(response)[1]
-  d2 <- dim(response)[2]
-  coe <- array(0, dim = c(d1, d2, dim(X)[2]))
-  if (dist == "normal") {
-    for (i in 1:d1) {
-      for (j in 1:d2) {
-        fit <- lm(response[i, j, ] ~ -1 + X)
-        coe[i, j, ] <- coef(fit)
-      }
-    }
-  }
-  else if (dist == "binary") {
-    for (i in 1:d1) {
-      for (j in 1:d2) {
-        fit <- suppressWarnings(glm(response[i, j, ] ~ -1 + X, family = binomial("logit")))
-        coe[i, j, ] <- coef(fit)
-      }
-    }
-  }
-  else if (dist == "poisson") {
-    for (i in 1:d1) {
-      for (j in 1:d2) {
-        fit <- suppressWarnings(glm(response[i, j, ] ~ -1 + X, family = poisson("log")))
-        coe[i, j, ] <- coef(fit)
-      }
-    }
-  }
-  return(coe)
-}
-
-# valid rank range for core tensor ## the maximum rank should be no smaller than the product of other
-# Input: rank_range = expand.grid(...)
-valid_rank <- function(rank_range) {
-  N <- dim(rank_range)[1]
-  output <- NULL
-  for (i in 1:N) {
-    rank <- rank_range[i, ]
-    rank <- sort(rank)
-    if (rank[1] * rank[2] >= rank[3]) {
-      output <- rbind(output, rank_range[i, ])
-    }
-  }
-  return(output)
-}
-
-# compute loglikelihood under different models
-loglike <- function(data, linearpre, dist) {
-  if (dist == "binary") {
-    p <- plogis(linearpre) # log-likelihood
-    L <- dbinom(data, 1, p, log = TRUE)
-    return(sum(L))
-  }
-  else if (dist == "normal") {
-    sigma_est <- mean((data - linearpre)^2)
-    L <- dnorm(data, linearpre, sqrt(sigma_est), log = TRUE)
-    sum(L) # log-likelihood
-  }
[…]

### 2021-06-21 tensor_regress/JCGS/Response_submission/Code/software/tensorregress/R/HCP.R

@@ -1,15 +1,0 @@
-#' HCP data
-#'
-#' The array "tensor" is a 68 × 68 × 136 binary tensor consisting of structural connectivity patterns among 68 brain regions for 136 individuals from Human Connectome Project (HCP). All the individual images were preprocessed following a standard pipeline (Zhang et al., 2018), and the brain was parcellated to 68 regions-of-interest following the Desikan atlas (Desikan et al., 2006). The tensor entries encode the presence or absence of fiber connections between those 68 brain regions for each of the 136 individuals.
-#' The data frame "attr" is a 136 × 573 matrix consisting of 573 personal features for 136 individuals. The full list of covariates can be found at:
-#' https://wiki.humanconnectome.org/display/PublicData/
-#'
-#'@docType data
-#'@usage data(HCP)
-#'
-#'@format  A list. Includes a 68-68-136 binary array named "tensor" and a 136-573 data frame named "attr".
-#'
-#'@keywords datasets
-#'
-#'
-"HCP"

### 2021-06-21 tensor_regress/JCGS/Response_submission/Code/software/tensorregress/R/bricks.R

@@ -1,212 +1,0 @@
-#library(rTensor)
-library(pracma)
-#library(gtools)
-library(MASS)
-library(speedglm)
-#library(fastglm)
-#library(rgl)
-#library("RColorBrewer")
-
-
-
-#######################
-######----------   functions for update
-
-## This part used for penalty (Conjugate gradient method)
-
-loss = function(beta,y,X,lambda,alpha,dist){
-  U = X %*% beta
-  L= - loglike(y,U,dist)
-  if(max(abs(U))>alpha) return(Inf)
-  else{
-    L2 =L - lambda * sum(log(1 - (U/ alpha)^2))  ## object function
-    return(c(L2))
-  }
-}
-
-loglike=function(data,linearpre,dist){
-  if(dist=="binary"){
-    p=plogis(linearpre)## log-likelihood
-    L=dbinom(data,1,p,log=TRUE)
-    return(sum(L))
-  }
-  else if (dist=="normal"){
-    sigma_est=mean((data-linearpre)^2)
-    L=dnorm(data,linearpre,sqrt(sigma_est),log=TRUE)
-    sum(L)##log-likelihood
-  }
-  else if (dist=="poisson"){
-    lambda=exp(linearpre)
-    L=dpois(data,lambda, log = TRUE) ## log-likelihood for Poisson
-    return(sum(L))
-  }
-}
-
-loss_gr <- function(beta,y,X,lambda,alpha,dist){
-  U = X %*% beta
-  if(dist=="binary") p=plogis(U)
-  else if (dist=="normal") p=U
-  else if (dist=="poisson") p=exp(U)
-  L_g=t(X) %*% (p - y)
-  penal_g = 2 * lambda * t(X) %*% (U/(alpha^2 - U^2))
-  return(c(L_g) + c(penal_g))
-}
-
-
-glm_modify=function(y,x,dist){
-  if(dist=="binary"){
-    fit1 = tryCatch({
-        speedglm(y~-1+x,family=binomial(link="logit"))
-    }, warning = function(w) {
- fit1=suppressWarnings(speedglm(y~-1+x,family=binomial(link="logit")))
-    }, error = function(e) {
-        return_value="non-convergence"
-    })
-    
-    if(inherits(fit1,"character")==TRUE){
-        return(list(rep(0,dim(as.matrix(x))[2]),-Inf))
-    }else
[…]

### 2021-06-21 tensor_regress/JCGS/Response_submission/Code/software/tensorregress/R/nations.R

@@ -1,13 +1,0 @@
-#' nations data
-#'
-#' The array "R" is a 14 × 14 × 56 binary tensor consisting of 56 political relations of 14 countries between 1950 and 1965. The tensor entry indicates the presence or absence of a political action, such as “treaties”, “sends tourists to”, between the nations. Please set the diagonal elements Y(i,i,k) = 0 in the analysis.
-#' The matrix "cov" is a 14 × 6 matrix describing a few important country attributes, e.g. whether a nation is actively involved in medicine NGO, law NGO, or belongs to a catholic nation, etc.
-#'
-#'@docType data
-#'@usage data(nations)
-#'
-#'@format  A list. Includes a 14-14-56 binary array named "R" and a 14-6 matrix named "cov".
-#'@keywords datasets
-#'
-#'
-"nations"

### 2021-06-21 tensor_regress/JCGS/Revision/Code/Figure2.R

@@ -1,0 +1,234 @@
+# Code for Figure 2. Log-likelihood trajectory for three models under different (d,r) combinations
+
+# load dependencies ---------------------------------------------------------------------------------------------
+# for plotting
+library(ggplot2)
+library(patchwork)
+
+# for simulation
+install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
+library(tensorregress)
+
+library(pracma)
+source("simulation.R")
+
+# start simulation ------------------------------------------------------------------------------------------------
+
+set.seed(0)
+model <- c("normal", "binary", "poisson")
+dimen <- c(30, 25)
+rank <- c(6, 3)
+
+lglk_sum <- list()
+true_lglk_sum <- list()
+time_sum <- list()
+
+
+for (i in 1:3) { # three data types
+  lglk <- list()
+  true_lglk <- c()
+  time <- c()
+  k <- 1
+  for (j in 1:2) { # two possible dimensions
+    for (r in 1:2) { # two possible ranks
+      if (i == 1) {
+        signal <- 3
+      } else if (i == 2) {
+        signal <- 10
+      } else if (i == 3) signal <- 6
+      my_data <- sim_data(seed = NA, rep(dimen[j], 3), rep(rank[r], 3), p = rep(floor(0.4 * dimen[j]), 3), dist = model[i], dup = 1, signal = signal)
+      ptm <- proc.time()
+      my_reg <- tensor_regress(tsr = my_data$tsr[[1]], X_covar1 = my_data$X_covar1, X_covar2 = my_data$X_covar2, X_covar3 = my_data$X_covar3, rep(rank[r], 3), niter = 20, cons = "non", dist = model[i], traj_long = TRUE, initial = "random")
+      time1 <- proc.time() - ptm
+      time[k] <- time1[3] # running time
+      true_lglk[k] <- loglike(my_data$tsr[[1]], my_data$U, dist = model[i])
+      lglk[[k]] <- my_reg$lglk
+      k <- k + 1
+    }
+  }
+  lglk_sum[[i]] <- lglk
+  true_lglk_sum[[i]] <- true_lglk
+  time_sum[[i]] <- time
+}
+
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Code/Figure4.m

@@ -1,0 +1,133 @@
+% Code for Figure 4: ME comparison among STD, Envelope, GLSNet, mRRR, and SupCP
+
+%load software
+addpath("software/SupCP-master");
+
+% Error vs Info Number ----------------------------------------------------
+% load data
+rng(0);
+load("Figure4_mode_data.mat");
+
+dup = 30;
+
+err_sup = zeros(2,2,3,dup);
+cor_sup = zeros(2,2,3,dup);
+
+final_mode_sup = zeros(2,2,3);
+final_sd_mode_sup = zeros(2,2,3);
+final_mode_cor_sup = zeros(2,2,3);
+final_sd_mode_cor_sup = zeros(2,2,3);
+
+rank_range = [2,4,6,8,10];
+rank_choose = zeros(2,2,3,dup);
+rank_choose_cor = zeros(2,2,3,dup);
+
+args = struct('max_niter',2000,'AnnealIters',1000);
+for s = 1:2 % signal alpha = 3,6
+    for r = 1:2 % core shape = (3,3,3) or (4,5,6)
+        for i = 1:3 % info number
+            for n = 1:dup % dup
+                
+                disp([para{s,r}{i,n}'])
+                Y = y{s,r}{i,n};
+                X = tsr{s,r}{i,n};
+                X_true = x_true{s,r}{i,n};
+                
+                test = zeros(1,length(rank_range)); test_cor = zeros(1,length(rank_range));
+                for k = 1:length(rank_range)
+                    rank = rank_range(k);
+                    disp(['choose rank as ', num2str(rank)]);
+                    
+                    [B,V,U,se2,Sf,rec] = SupParafacEM(Y,X,rank,args);
+                    
+                    x_fit = TensProd([{U}, V]);
+                    test(k) = mean((x_fit - X_true).^2, 'all');
+                    correlation = corrcoef(x_fit,X_true);
+                    test_cor(k) = correlation(1,2);
+                end 
+                
+                [M,I] = min(test); 
+                [M_cor,I_cor] = max(test_cor);
+                rank_choose(s,r,i,n) = rank_range(I);
+                rank_choose_cor(s,r,i,n) = rank_range(I_cor);
+                
+                err_sup(s,r,i,n) = M;
+                cor_sup(s,r,i,n) = M_cor;
+            end 
+            
+            final_mode_sup(s,r,i) = mean(err_sup(s,r,i,:));
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Code/Figure6.m

@@ -1,0 +1,71 @@
+% Code for Figure 6: ME comparison of STD and other methods under model misspecifications
+
+% load software
+addpath("software/SupCP-master");
+
+% model misspecification with noniid noise --------------------------------
+% load data
+rng(10);
+load("Figure6_noniid_data.mat");
+
+dup = 30;
+
+err_sup = zeros(2,2,4,dup);
+cor_sup = zeros(2,2,4,dup);
+
+final_c_sup = zeros(2,2,4);
+final_sd_c_sup = zeros(2,2,4);
+final_c_cor_sup = zeros(2,2,4);
+final_sd_c_cor_sup = zeros(2,2,4);
+
+rank_range = [2,4,6,8,10];
+rank_choose = zeros(2,2,4,dup);
+rank_choose_cor = zeros(2,2,4,dup);
+
+args = struct('max_niter',2000,'AnnealIters',1000);
+for s = 1:2 % rank
+    for r = 1:2 % signal
+        for i = 1:4 % cor_level
+            for n = 1:dup % dup
+                
+                disp([para{s,r}{i,n}'])
+                Y = y{s,r}{i,n};
+                X = tsr{s,r}{i,n};
+                X_true = x_true{s,r}{i,n};
+                
+                test = zeros(1,length(rank_range)); 
+                test_cor = zeros(1,length(rank_range));
+                
+                for k = 1:length(rank_range)
+                    rank = rank_range(k);
+                    disp(['choose rank as ', num2str(rank)]);
+                    
+                    [B,V,U,se2,Sf,rec] = SupParafacEM(Y,X,rank,args);
+                    
+                    x_fit = TensProd([{U}, V]);
+                    test(k) = mean((x_fit - X_true).^2, 'all');
+                    correlation = corrcoef(x_fit,X_true);
+                    test_cor(k) = correlation(1,2);
+                end
+                
+                [M,I] = min(test); 
+                [M_cor,I_cor] = max(test_cor);
+                rank_choose(s,r,i,n) = rank_range(I);
+                rank_choose_cor(s,r,i,n) = rank_range(I_cor);
+                
+                err_sup(s,r,i,n) = M;
+                cor_sup(s,r,i,n) = M_cor;
+            end
+            
+            final_c_sup(s,r,i) = mean(err_sup(s,r,i,:));
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Code/Figure8.R

@@ -1,0 +1,66 @@
+# Code for Figure 8. Nations data analysis
+
+# load dependencies ---------------------------------------------------------------------------------------------
+install.packages("software/tensorregress.tar.gz", repos = NULL, type = "source")
+library(tensorregress)
+
+library(lattice)
+library(RColorBrewer)
+
+library(pracma)
+source("simulation.R")
+
+# load data -----------------------------------------------------------------------------------------------------
+data <- load("rawdata/nations.RData")
+
+# start analysis ------------------------------------------------------------------------------------------------
+set.seed(0)
+seed <- 0
+
+tsr <- R
+tsr[is.na(tsr)] <- 0
+X <- cov
+
+# If a new run of rank selection is desired, run the following two lines
+# rank_range=valid_rank(expand.grid(c(3:5),c(3:5),c(3:5)))
+# res=sele_rank(tsr,X_covar1=cov,X_covar2=cov,X_covar3=NULL,rank_range=rank_range,niter=10,cons = 'non',dist="binary",initial = "random")
+
+core_shape <- c(4, 4, 4)
+result <- tensor_regress(tsr, X_covar1 = cov, X_covar2 = cov, X_covar3 = NULL, core_shape, niter = 10, cons = "penalty", lambda = 0.1, alpha = 10, solver = "CG", dist = "binary", initial = "random")
+
+save(result, file = "presaved/output_nations.RData")
+clustering <- kmeans(result$W[[3]], 4, nstart = 5)
+relnames <- names(tsr[1, 1, ])
+relnames[clustering$cluster == 1]
+relnames[clustering$cluster == 2]
+relnames[clustering$cluster == 3]
+relnames[clustering$cluster == 4]
+
+
+# start plotting ------------------------------------------------------------------------------------------------
+my.palette <- brewer.pal(n = 11, name = "RdBu")
+myPanel <- function(x, y, z, ...) {
+  panel.levelplot(x, y, z, ...)
+  # use handy matrix indexing
+  panel.text(x, y, round(M[cbind(x, y)], 2)) 
+}
+
+pdf("Figures/Figure8_1.pdf", width = 6, height = 6)
+M <- result$C_ts[, , which(relnames == "warning")]
+levelplot(M, panel = myPanel, col.regions = my.palette, at = seq(-1.6, 1.6, length.out = 12), main = "Relation: warning")
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/Code/software/tensorregress/R/tensor_regress.R

@@ -1,0 +1,539 @@
+#' Supervised Tensor Decomposition with Interactive Side Information
+#'
+#' Supervised tensor decomposition with interactive side information on multiple modes. Main function in the package. The function takes a response tensor, multiple side information matrices,
+#' and a desired Tucker rank as input. The output is a rank-constrained M-estimate of the core tensor and factor matrices.
+#'
+#'
+#' @param tsr    response tensor with 3 modes
+#' @param X_covar1    side information on first mode
+#' @param X_covar2    side information on second mode
+#' @param X_covar3    side information on third mode
+#' @param core_shape  the Tucker rank of the tensor decomposition
+#' @param niter        max number of iterations if update does not convergence
+#' @param cons        the constraint method, "non" for without constraint, "vanilla" for global scale down at each iteration, "penalty" for adding log-barrier penalty to object function
+#' @param lambda      penalty coefficient for "penalty" constraint
+#' @param alpha       max norm constraint on linear predictor
+#' @param solver      solver for solving object function when using "penalty" constraint, see "details"
+#' @param dist        distribution of the response tensor, see "details"
+#' @param traj_long   if "TRUE", set the minimal iteration number to 8; if "FALSE", set the minimal iteration number to 0
+#' @param initial     initialization of the alternating optimiation, "random" for random initialization, "QR_tucker" for deterministic initialization using tucker decomposition
+#' @return     a list containing the following:
+#'
+#'                  \code{W} {a list of orthogonal factor matrices - one for each mode, with the number of columns given by \code{core_shape}}
+#'
+#'                  \code{G}  {an array, core tensor with the size specified by \code{core_shape}}
+#'
[…]

### 2021-06-21 tensor_regress/JCGS/Revision/manuscript/response.tex

\documentclass[11pt]{article}

\usepackage{fancybox}

\usepackage{color}
\usepackage{url}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{lscape}
\usepackage{afterpage}

\setlength{\parskip}{1em}%
\setlength\parindent{0pt}
\renewcommand{\textfraction}{0.0}
\renewcommand{\topfraction}{1.0}
%\renewcommand{\textfloatsep}{5mm}

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{extarrows}
\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
%\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\mathtoolsset{showonlyrefs=true}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\setcounter{figure}{0}   
\setcounter{table}{0}

\newcommand{\of}[1]{\left(#1\right)}
\newcommand{\off}[1]{\left[#1\right]}
\newcommand{\offf}[1]{\left\{#1\right\}}
\newcommand{\aabs}[1]{\left|#1\right|}

%\usepackage{xcolor}

\newcommand{\cmt}[1]{{\leavevmode\color{red}{#1}}}

\usepackage{dsfont}
\usepackage{multirow}

\DeclareMathOperator*{\minimize}{minimize}

\usepackage{mathtools}
\mathtoolsset{showonlyrefs}
\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}

% adjust table separation
%\setlength{\tabcolsep}{6pt} % Default value: 6pt
%\renewcommand{\arraystretch}{1.5}

\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}

\usepackage{xr}
\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother
\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}
\myexternaldocument{jcgs-tensor}
\myexternaldocument{supp}

\input macros.tex

\usepackage[english]{babel}
\usepackage[dvipsnames]{xcolor}

\usepackage{enumitem,kantlipsum}

\title{\textbf{Response Letter}}

\date{}
\author{}

%\setstretch{1.125}
\begin{document}

\maketitle
\vspace{-1.5cm}

We are grateful to the reviewers for their careful reading and thoughtful comments. Please note that the \textbf{Response to Major Comments Summarized by Editor} consists of general comments raised by both reviewers. We address these issues upfront {\bf in pages 1-7 of this letter}.

\begin{center}
    \textbf{Response to Major Comments Summarized by Editor}
\end{center}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
    \item \textit{Comparison with other related methods should be included as pointed out by a reviewer.}

    \textbf{Response:} We thank the editor and both reviewers for pointing out additional related literature. Some references \citep{chi2012tensors, hong2020generalized, raskutti2019convex} were already included in the first version, but we have added more comparisons in the revision. We have also added other relevant papers suggested by reviewers. The current comparison consists of 12 tensor regression/factorization methods. Moreover, we have revised our numerical experiments substantially to compare our method with more recent work. Specifically, 
    \begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
        \item we have added discussions about unsupervised model in Section~\ref{sec:intro}: 
        \begin{quote}
            ``...The first line is a class of \textit{unsupervised} tensor decomposition such as classical Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor, wang2017tensor} and generalized decomposition for non-Gaussian data \citep{chi2012tensors, tarzanagh2019regularized,hong2020generalized,  li2020generalized}. Regardless of the implementation, the unsupervised methods aim to find the best low-rank representation of a data tensor alone. In contrast, our model is a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and multiple features. The low-rank factorization is determined jointly by the tensor data and feature matrices in our model."
        \end{quote}

        \item we have added discussions about tensor-on-tensor models in Section~\ref{sec:intro}: 
        \begin{quote}
            ``...The second line of work studies the tensor-to-tensor regression. This category is further divided into three scenarios, depending on whether tensor is treated as predictors~\citep{zhou2013tensor,raskutti2019convex,han2020optimal}, as responses~\citep{li2017parsimonious,sun2017store,zhang2018network,lock2018supervised,luo2018leveraging}, or both~\citep{lock2018tensor,gahrooei2020multiple}. As we show in Section~\ref{sec:connection}, our supervised tensor decomposition falls into this general category, and we provide a \emph{provable} solution in new settings that have broader practical significance. Earlier work in this vein~\citep{lock2018tensor,lock2018supervised,gahrooei2020multiple,li2020generalized} focuses on algorithm development, but not on the statistical accuracy. \cite{li2017parsimonious} introduces an envelope-based approach to identify sufficient dimension reduction~\citep{adragni2009sufficient}, but its theory is restricted to Gaussian data with one-sided feature matrix only. \cite{raskutti2019convex} establishes the statistical accuracy for convex relaxed maximum likelihood estimator (MLE) of tensor regression. However, convex relaxation for tensor optimizations suffers from computational intractability and statistical sub-optimality. Recent work has demonstrated the success of non-convex approaches in various tensor problems~\citep{sun2017store,zhang2018network,raskutti2019convex,han2020optimal}; we go step further by allowing multiple feature matrices with either fixed or random designs. In Sections~\ref{subsec:statprob}, we show that incorporating multiple feature matrices substantially improves the statistical accuracy. We provide a detailed comparison in Section~\ref{sec:connection}; see Table~\ref{table:comp_table}."
        \end{quote}

 \item we have added a new Section~\ref{sec:connection}. \textbf{Connection to other tensor regression methods}:

        \begin{quote}
            ``...We compare our supervised tensor decomposition (\textbf{STD}) with recent 12 tensor methods in the literature. Table~\ref{table:comp_table} summarizes these methods with their properties from four aspects: i) model specification, ii) number of feature matrices allowed, (iii) capability of addressing non-Gaussian response, and (iv) capability of addressing non-independent noise. The four closet methods to our are {\bf SupCP}~\citep{lock2018supervised}, {\bf Envelope}~\citep{li2017parsimonious}, {\bf mRRR}~\citep{luo2018leveraging} and {\bf GLSNet}~\citep{zhang2018network}; these methods all relate a data tensor to feature matrices with low-rank structure on the coefficients. As seen from the table, our method is the only one that allows multiple feature matrices among the five. {\bf Envelope} and {\bf SupCP} are developed for Gaussian data, and the Gaussianity facilities flexible extension to non-independent noise. In particular, {\bf Envelope} allows correlation in Kronecker structured form, whereas {\bf SupCP} allows correlation implicitly through decomposing the latent factors into fixed effects (related to features) and random effects (unrelated to features). On the other hand, the other three methods ({\bf mRRR}, {\bf GLSNet} and {\bf STD}) are developed for exponential family distribution with possibly non-additive noises. The generality makes the full modeling of correlation computationally challenging. We will compare the numerical performance of these methods in Section~\ref{sec:simulation}.''

\afterpage{
\begin{landscape}
 \renewcommand{\arraystretch}{1.5} % Default value: 1
\begin{table}[h!]
\centering
\resizebox{\columnwidth}{!}{
\begin{tabular}{c||c|ccc}
\hline
Method & Model& No.\ of features & non-Gaussianity & Non-independence \\
\hline
	STD (Ours) & $\bbE\tY = f(\tB \times \{\mX_1,\mX_2,\mX_3\}),\ \tB = \tC \times \{\mM_1,\mM_2,\mM_3\}$ &$3$& $\surd$& $\times$\\
	GCP, CP-ARP, CORALS & $ \bbE\tY = f(\entry{ \mA_1,\mA_2,\mA_3})$ &0& $\surd$& $\times$\\
         DCOT & $\bbE\tY = f((\tC_1+\tC_2) \times \{\mM_1, \mM_2, \mM_3 \})$ &0& $\surd$ & $\times$\\
         LRT, CRT & $y_{n} = \langle \tB,\ \tX_{n} \rangle + \epsilon_{n}$, various structure on $\tB$&0& $\times$& $\times$\\
	STAR & $y_n = \sum_m \langle \tB_m, \tF_m(\tX_{ijk})\rangle + \epsilon_n$, sparse-CP $\tB_m$ &0& $\times$ & $\times$\\
	SupCP &   $\tY = \entry{ \mA_1,\mA_2,\mA_3} + \tE,\  \mA_1 = \mX \mB +\tE',\ \tE\perp \tE'$ &1& $\times$& $\surd$\\
	mRRR &  $\bbE \mY =  f(\mX \mB)$, low-rank $\mB$ &1& $\surd$& $\times$ \\
	Envelope & $\tY = \tB \times_3 \mX + \tE,\  \tB = \tC \times \{ \mM_1, \mM_2, \mI \},\  \tE \sim \mathcal{TN}(\mSigma_1, \mSigma_2,\mI)$ &1& $\times$ & $\surd$\\
	GLSNet& $\bbE\tY = f({\bf 1} \otimes {\bf \Theta} + \tB \times_3 \mX)$, low-rank ${\bf \Theta}$, sparse $\tB$&1& $\surd$& $\times$\\
	STORE & $\tY = \tB \times_3 \mX + \tE$,  sparse-CP $\tB$ &1& $\times$& $\times$ \\
	\hline
\end{tabular}
}
\caption{\footnotesize Comparison of tensor regression/factorization methods. We focus on order-3 tensors for illustration. Calligraphic letters denote tensors, bold capital letters denote matrices, and little letters denote scalars. The dimension of tensors and matrices can be identified from the contexts.\\
\vspace{-0.1cm}\\
- Data: tensor response $\tY$, feature matrices $\mX,\mX_k$, predictor tensor $\tX_n$, scalar response $y_n$, sample index $n$, tensor mode $k=1,2,3$.  \\
- Parameter: Tuckor factors $\mM_k$, CP factors $\mA_k$, CP decomposition $\entry{\mA_1, \mA_2, \mA_3}$, coefficient tensor and matrix $\tB, \tB_m, {\bf \Theta}, \mB$.\\
- Function: a known link function $f(\cdot)$, a known basis function $\tF_m(\cdot)$. \\
- Noise: Gaussian tensor with i.i.d.\ entries $\tE, \tE'$, Gaussian tensor with Kronecker covariance $\tE \sim \mathcal{TN}(\mSigma_1, \mSigma_2, \mI)$, meaning $\text{Cov}(\text{vec}(\tE))=\mSigma_1\otimes\mSigma_2\otimes \mI$.\\
\vspace{-0.1cm}\\
- GCP: Generalized canonical polyadic tensor decomposition~\citep{hong2020generalized};\\
- CP-APR: CP alternating Poisson regression~\citep{chi2012tensors};\\
- CORALS: Generalized co-clustering method~\citep{li2020generalized};\\
- DCOT: Double core tensor decomposition~\citep{tarzanagh2019regularized};\\
- SupCP: Supervised PARAFAC/CANDECOMP factorization~\citep{lock2018supervised};\\
- mRRR: Mixed-response reduced-rank regression~\citep{luo2018leveraging};\\
- Envelope: Parsimonious tensor response regression~\citep{li2017parsimonious};\\
- GLSNet: Generalized connectivity matrix response regression~\citep{zhang2018network};\\
- STORE: Sparse tensor response regression~\citep{sun2017store};\\
- LTR: Low-rank tensor regression~\citep{han2020optimal};\\
- CRT: Convex regularized multi-response tensor regression~\citep{raskutti2019convex};\\
- STAR: Sparse tensor additive regression~\citep{hao2019sparse}.
}\label{table:comp_table}
\end{table}
\end{landscape}
}

\end{quote}

        \item We have added new numerical experiments in Section~\ref{sec:simulation} comparing the more current and related methods with ours. The section now reads:

\begin{quote}
            ``...These four methods are the closest methods to ours, in that they all relate a data tensor to feature matrices with low-rank structure on the coefficients. We consider Gaussian and Bernoulli tensors in experiments. For methods not applicable for Bernoulli data ({\bf SupCP} and {\bf Envelope}), we provide the algorithm $\{-1,1\}$-valued tensors as inputs. Because {\bf mRRR} allows matrix response only, we provide the algorithm the unfolded matrix of response tensor as inputs. We measure the accuracy using the response error defined as $1-\text{Cor}(\hat \tY, f(\trueT))$, where $\hat \tY$ is the fitted tensor from each method, and $f(\trueT)$ is the true conditional mean of the tensor. Note that the response error is a scale-insensitive metric; a smaller error implies a better fit of the model.

The comparison is assessed from three aspects: (i) benefit of incorporating features from multiple modes; (ii) prediction error with respect to sample size; (iii) robustness of model misspecification. We use similar simulation setups as in our first experiment in last section. We consider rank $\mr=(3,3,3)$ (low) vs.\ $(4,5,6)$ (high), signal $\alpha = 3$ (low) vs.\ $6$ (high), dimension $d$ ranging from 20 to 100 for modes with features, and $d = 20$ for modes without features. The method \textbf{Envelope} and \textbf{mRRR} require the tensor rank as inputs, respectively. For fairness, we provide both algorithms the true rank. The methods \textbf{SupCP} and \textbf{GLSNet} use different notions of model rank, and \textbf{GLSNet} takes sparsity as an input. We use a grid search to set the hyperparameters in \textbf{SupCP} and \textbf{GLSNet} that give the best performance.

\begin{figure}[H]
\centering
\includegraphics[width=12cm]{comp_final_normal.pdf} 
\caption{Comparison between tensor methods with Gaussian data. Panels (a) and (b) plot estimation error versus the number of modes with available features. Panels (c) and (d) plot ME versus the effective sample size $d^2$.
We consider rank $\mr=(3,3,3)$ (low), $\mr=(4,5,6)$ (high), and signal $\alpha =3 $ (low), $\alpha=6$ (high).}~\label{fig:comp}
\end{figure}

Figure~\ref{fig:comp}a-b shows the impact of features to estimation error. We see that our \textbf{STD} outperforms others, especially in the low-signal high-rank setting. As the number of informative modes increases, the \textbf{STD} exhibits a reduction in error whereas others remain unchanged. The accuracy gain in Figure~\ref{fig:comp} demonstrates the benefit of incorporating informative features from multiple modes. In addition, we find that the relative performance among the competing methods reveals the benefits of low-rankness. 
The second best method is \textbf{SupCP} which imposes low-rankness on three modes; the next one is \textbf{Envelope} which imposes low-rankness on two modes; the less favorable one is \textbf{mRRR} which imposes low-rank structure on one mode only; the worst one is \textbf{GLSNet} which imposes sparsity but no low-rankness on the feature effects.

Figure~\ref{fig:comp}c-d compares the prediction error with respect to effective sample size $d^2$. For fair comparison, we consider the setting with feature matrix on one mode only. We find that our \textbf{STD} method has similar performance as \textbf{Envelope} and \textbf{SupCP} in the high-signal low-rank regime, whereas the improvement becomes more pronounced in the low-signal high-rank regime. The latter setting is notably harder, and our \textbf{STD} method shows advantage in addressing this challenge. Among other methods, \textbf{Envelope}, \textbf{SupCP}, and \textbf{mRRR} show decreasing errors as $d$ increases, implying the benefits of low-rankness methods. In contrast, \textbf{GLSNet} suffers from non-decreasing error and indicates the poor fit of sparsity methods in addressing low-rank data.

\begin{figure}[htb]
\centering
\includegraphics[width=12cm]{comp_final_binary.pdf} 
\caption{Comparison between tensor methods with Binary data. The panel legends are the same as in Figure~\ref{fig:comp}.}~\label{fig:comp_b}
\end{figure}

We also evaluate the performance comparison with Bernoulli tensors. Figure~\ref{fig:comp_b} indicates the necessarity of generalized model in addressing non-Gaussian data. Indeed, methods that assume Gaussiannity (\textbf{Envelope} and \textbf{SucCP}) perform less favorably in Bernoulli setting (Figure~\ref{fig:comp_b}c) compared to Gaussian setting (Figure~\ref{fig:comp}c). Our method shows improved accuracy as the number of informative features increases (Figure~\ref{fig:comp_b}a-b). In the absence of multiple features, our method still performs favorably compared to others (Figure~\ref{fig:comp_b}c-d), for the same reasons we have argued in Gaussian data. ''
  \end{quote}

        \item we have added a new experiment in Section~\ref{sec:simulation} to evaluate the robustness of our method with model misspecification:
        \begin{quote}
            ``Lastly, we assess the performance of our method {\bf STD} under model misspecification. We consider two aspects: (i) non-independent noise, and (ii) sparse feature effects. Note that our method {\bf STD} imposes neither of these two assumptions, so the experiment allows us to assess the robustness. We select competing methods from Table~\ref{table:comp_table} that specifically addresses these two aspects. We use {\bf Envelope} and {\bf SupCP} as benchmark for noise correlation experiment, and {\bf GLSNet} for sparsity experiment.

Figure~\ref{fig:noniid}a-b assesses the impact of noise correlation to the estimation accuracy. The data is simulated from {\bf Envelope} model with envelope dimensions $r=(3,3)$ (low) and $(4,5)$ (high). The noise is generated from a zero-mean Gaussian tensor with Kronecker structured covariance; see Supplementary Notes for details. As expected, {\bf Envelope} shows the best performance in the high correlation setting. Remarkably, we find that our method {\bf STD} has comparable and sometimes better performance when noise correlation is moderate-to-low. In contrast, {\bf SupCP} appears less suitable in this setting. Although {\bf SupCP} allows noise correlation implicitly through latent random factors, the induced correlation may not belong to the Kronecker covariance structure in the simulation.

Figure~\ref{fig:noniid}c-d assesses the impact of sparsity to estimation performance. We generate data from {\bf GLSNet} model, except that we modify the coefficient tensor to be joint sparse and low-rank (the original {\bf GLSNet} model assumes full-rankness on the coefficient tensor). 
The sparsity level ($x$-axis in Figure~\ref{fig:noniid}c-d) quantifies the proportion of zero entries in the coefficient tensor. Since neither our method {\bf STD} nor {\bf GLSNet} follow the simulated model, this setting allows a fair comparison. We find that our method outperforms {\bf GLSNet} in the low-rank setting, whereas {\bf GLSNet} shows a better performance in the high-rank setting. This observation suggests the robustness of our method to sparsity when the tensor of interest is simultaneously low-rank and sparse. When sparsity is the only salient structure, then methods specifically addressing sparsity would provide a better fit.''

\begin{figure}[htb]
\centering
\includegraphics[width=12cm]{noniid_final.pdf} 
\caption{Comparison between tensor methods under model misspecification. Panels (a)-(b) assess the error correlation, and panels (c)-(d) assess the sparsity. }~\label{fig:noniid}
 \end{figure}

 \end{quote}

 For self-consistency, we quote the detailed simulation setups for the Figure~\ref{fig:noniid} in the Supplementary Notes, Section \ref{sec:asimulation}:
\begin{quote}
    ``
    \textbf{B.1 Detailed simulation setup for Figure~\ref{fig:noniid}a-b}

    \vspace{0.2cm}

    We generate data from \textbf{Envelope} model \citep{li2017parsimonious} with slight modification. We simulate response tensor $\tY\in\mathbb{R}^{d\times d\times d}$ from the following model with envelope dimension $(u_1, u_2)$,
\begin{align}\label{eq:e}
    \tY|\mX &= \tB \times_3 \mX + \tE = \tC \times \{\mGamma_1, \mGamma_2,\mX\}+ \tE, \notag \\
    \text{with} \quad \tE &\sim \mathcal{TN}(\mSigma_1, \mSigma_2, \mI), \quad \mSigma_k = \mGamma_k \mOmega_k \mGamma_k^T + \mGamma_{0k} \mOmega_{0k} \mGamma_{0k}^T+\mI, \quad k = 1,2,
\end{align}
 where $\mX\in\mathbb{R}^{d\times p}$ is the feature matrix, $\tB=\tC\times\{\mGamma_1,\mGamma_2, \mI\}\in \mathbb{R}^{d\times d\times p}$ is the coefficient tensor, $\tC\in\mathbb{R}^{\mu_1\times \mu_2\times p}$ is a full-rank core tensor, $\mathcal{TN}(\cdot,\ \cdot,\ \cdot)$ represents zero-mean tensor normal distribution with Kronecker structured covariance, 
 $\mGamma_{k} \in \mathbb{O}(d,u_k)$ 
 consists of orthogonal columns, $\mGamma_{0k} \in \mathbb{O}(d,d-u_k)$ is the orthogonal complement of $\mGamma_k$, and  $\mOmega_k = \mA_k\mA_k^T$, $\mOmega_{0k} = \mA_{k0}\mA_{k0}^T$ with $\mA_k \in \bbR^{u_k \times u_k}$,  $\mA_{k0} \in \bbR^{(d - u_k) \times (d - u_k)}$.

 The entries of $\mX$ are i.i.d.\ drawn from $\tN(0,1)$, the entries of $\mA_k$, $\mA_{k0}$ are i.i.d.\ drawn from $\textup{Uniform}[-\gamma, \gamma]$, and the entries of core tensor $\tC$ are i.i.d.\ drawn from $\textup{Uniform}[-3,  3]$. We call $\gamma$ the \textit{correlation level}. Note that the only distinction between model~\eqref{eq:e} and standard {\bf Envelope} model is the additional identity matrix $\mI$ in the expression of $\mSigma_k$. When $\gamma = 0$, the model~\eqref{eq:e} reduces to our \textbf{STD} model with rank $\mr = (u_1,u_2,p)$. We set $d=20, p=5$ in our simulation.

 \vspace{0.2cm}
 \textbf{B.2 Detailed simulation setup for Figure~\ref{fig:noniid}c-d}

 \vspace{0.2cm}

We generate the data from  \textbf{GLSNet} model \citep{zhang2018network} with slight modification. We simulate the binary response tensor $\tY \in \{0,1\}^{d \times d \times d}$ from the following model
\begin{align}
    \mathbb{E}[\tY|\mX] = f({\bf 1} \otimes {\bf \Theta} + \tB \times_3 \mX),
\end{align}
where $f(\cdot)$ is the logistic link, $\mX \in \mathbb{O}(d,p)$ is the feature matrix with orthonormal columns, ${\bf \Theta}=\mA \mA^T \in \bbR^{ d \times d }$ is a rank-$R$  intercept matrix, where the entries of $\mA \in \bbR^{d \times R}$ are simulated from i.i.d.\ standard normal. 
Unlike original \textbf{GLSNet} model, we generate joint sparse and low-rank structure to the coefficient tensor $\tB$ as follows.

To generate $\tB$, we firstly generate a low-rank tensor $\tB_0$ as 
\begin{align}
    \tB_0 = \tC \times \mM_1 \times \mM_2 \times \mM_3,
\end{align}
where $\tC \in \bbR^{R \times R \times R}$ is a full-rank core tensor, $\mM_1, \mM_2 \in \bbR^{d \times R}$ and $\mM_3 \in \bbR^{p \times R}$ are the factor matrices with orthonormal columns. We simulate i.i.d.\ uniform entries in $\tC$ and rescale the tensor $\tB_0$ such that $\onorm{\tB_0}_{\max} = 2$. Last, we obtain a sparse $\tB$ by randomly setting $sd^2p$ entries in $\tB_0$ to zero. We call $s$ the \textit{sparsity level} which quantifies the proportion of zero's in $\tB$. Hence, the generated tensor $\tB$ is of sparsity level $s$ and of low-rank $(R,R,R)$. We set $d = 20$, $p = 5$ and consider the combination of rank $R = 2$ (low), $4$ (high) and sparsity level $s = \{0, 0.3,0.5\}$ in the simulation. "

\end{quote}

\end{enumerate}

    \item \textit{Additional discussion on side information should be included that involves i) how do we know in practice if side information is appropriate; ii) dimensionality of the side information.}

    \vspace{0.2cm}
    \textbf{Response:} See our response to \textbf{\# 5, Reviewer 2}.

    \item \textit{The authors should discuss initialization and how to obtain the initial error condition. Can something be said about the proposed random initialization in Algorithm 1?}

         \textbf{Response:} We have added theoretical justifications to two initializations. See our response {\bf \# Initialization, Reviewer 1.}

    \item \textit{There is a disconnect between the proposed algorithm and established theory. Can the authors establish statistical properties for the iterative algorithm? If not, what are the technical challenges that are not present in the related literature?}

     \textbf{Response:} We have added new theory to establish the accuracy for both (i) global MLE and (ii) local optimizer. See our response to {\bf \# Theoretical Studies, Reviewer 1.}

    \item \textit{Some discussion on the computationally tractable selection of ranks should be included. See comment from Reviewer 2.}

    \textbf{Response:} We have added more detailed discussions and reported the running time for rank selection. See our response to \textbf{\# 8, Reviewer 2}.
\end{enumerate}

\newpage

\begin{center}
    \textbf{Point-by-point Response to Reviewer 1}
\end{center}

We greatly appreciate your valuable comments and suggestions. We have carefully addressed all the questions. Note that general comments raised by both reviewers were addressed upfront in {\bf pages 1-7 of this response letter, Response to Major Comments Summarized by Editor}. We will make explicit reference to earlier section when needed.

\vspace{0.2cm}

\textit{This paper introduces a novel supervised tensor decomposition model when there are side information on multiple modes. This model can be viewed as an extension to tensor response regression. As compared to existing tensor response regression models \citep{rabusseau2016low,li2017parsimonious,sun2017store}, the proposed method has three advantages. First, it allows covariate information on multiple modes instead of one mode. Second, it allows high-dimensionality on both the tensor size $d_k$ and covariate size $p_k$. Third, it allows a ﬂexible data type of the tensor, including continuous, count, and binary observations. Data satisfying these properties are not rare, and the paper has provided three data examples. Therefore, I believe this is a useful and well-motivated work. The construction of the estimator leads to a non-convex optimization. Both algorithm and theoretical study are constructed. This work can be regarded as a non-trivial extension of these tensor response regression models as well as \cite{wang2020learning} on probabilistic tensor decomposition which decomposes a tensor with continuous, count, or binary observations. Overall, I think this is an interesting work that has potential usefulness and will contribute to the literature of tensor regression. I enjoy reading the paper and feel it would be suitable for JCGS if the following comments are addressed. I recommend a major revision.}

\vspace{0.2cm}
\textit{\textbf{Initialization:}}
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
    \item \textit{How to initialize to achieve the initial error condition, as stated in Proposition 4.1 (Global convergence)?}

    {\bf Response:} We recommend two schemes for initialization: random initialization (cold start) and spectral initialization (warm start). In the high signal-to-noise regime, we have added theoretical guarantees for the statistical accuracy of warm start within polynomial running time. In the low signal-to-noise regime, the statistical-computational gap arises in which case random initialization often provides better empirical results. {\bf We have added a new section for provable initialization and its implementation.} See our detailed response to  {\bf\# Theoretical Studies}.

    We have added the following paragraph in Section~\ref{sec:rankM}:
    \begin{quote}
    ``...We provide two initialization schemes, one with QR-adjusted spectral initialization (warm initialization), and the other with random initialization (cold initialization). The warm initialization is an extension of unsupervised spectral initialization~\citep{zhang2018tensor} to supervised setting with multiple feature matrices. ... The initialization scheme is described in Algorithm~\ref{alg:A}.

The warm initialization enjoys provable accuracy guarantees at a cost of extra technical assumptions (see Section~\ref{subsec:statprob}). The cold initialization, on the other hand, shows robust in practice but its theoretical guarantee remains an open challenge~\citep{luo2021low}. We incorporate both options in our software package to provide flexibility to practitioners...''
\end{quote}

\begin{algorithm}[!h]
\caption{QR-adjusted spectral initialization}\label{alg:A}
\begin{algorithmic}[1]
\INPUT Response tensor $\tY\in \mathbb{R}^{d_1\times \cdots \times d_K}$, feature matrices $\mX_k\in\mathbb{R}^{d_k\times p_k}$, Tucker rank $\mr$. 
\State Normalize date tensor $\bar \tY\leftarrow \tY$ for Gaussian model, $\bar \tY\leftarrow 2\tY-1$ for Bernoulli model, and $\bar \tY\leftarrow \log(\tY+0.5)$ for Poisson model. 
\State Normalize feature matrices via QR factorization $\mX_k=\mQ_k\mR_k$ for all $k\in[K]$.
\State Obtain $\bar \tB \leftarrow \bar \tY\times\{\mQ^T_1,\ldots,\mQ^T_K\}$ by projecting $\bar \tY$ to the multilinear feature space. 
\State Obtain $\hat\tB^{(0)}\leftarrow \textup{HOSVD}(\bar \tB,\mr)$.
\State Normalize representation $\{\hat\tC^{(0)}, \hat\mM^{(0)}_1,\ldots,\hat\mM^{(0)}_K\}$ such that $\hat\tC^{(0)}\times\{ \hat\mM^{(0)}_1,\ldots,\hat\mM^{(1)}_K\} = \hat\tB^{(0)}\times\{ \mR^{-1}_1,\ldots,\mR^{-1}_K\}$ and $\hat\mM^{(0)}_k\in\mathbb{O}(p,r)$ for all $k\in[K]$.
\OUTPUT Core tensor $\hat\tC^{(0)}$ and factors $\hat\mM^{(0)}_k$ for all $k\in[K]$.
\end{algorithmic}
\end{algorithm}

    \item  \textit{As the proposed estimator is obtained from a nonconvex optimization, the possibility of having multiple local optimizers is a problem. In Algorithm 1, a random initialization is used. Can the authors construct an empirical study of how serious the local optimal affects the statistical performance of the randomly initialized estimator? Does the random initialization always lead to the global optimum?}

     \textbf{Response:} The empirical performance of random initialization was investigated in Figure~\ref{fig:loglike} in the original paper and the remarks thereof. We have shown in Theorem~\ref{thm:MLE} that the same statistical rate holds, not only for the {\it global optimizer}, but also for every {\it local optimizer} with sufficiently large objective values. In this sense, local optimality is not necessarily a severe concern in our context. We have added the following sentences for clarification in Sections~\ref{sec:global} and Section~\ref{sec:local}:

     \begin{quote}
         ``...Inspection of our proof (Supplementary Notes) shows that the desired convergence rate holds not only for the MLE, but also for all local optimizers in the set $\tL_{\tY}(\tC, \mM_1,\ldots,\mM_K)\geq \tL_{\tY} (\tC_{\textup{true}},\mM_{1,\textup{true}},\ldots,\mM_{K,\textup{true}})$. The observation indicates the global optimality is not necessarily a serious concern in our context, as long as the convergent objective is large enough. In next section, we will provide the statistical accuracy for \emph{local} optimizer with provable convergence guarantee, at a cost of extra signal requirement...''
     \end{quote}

     \begin{quote}
     ``...We supply the theory by providing an alternative scheme -- random initialization -- and investigate its empirical performance. Figure~\ref{fig:loglike} shows the trajectories of objective function ... In the experiment we conduct, we find random initialization appears good enough for Algorithm~\ref{alg:B} to find a convergent point with desired statistical guarantees. In practice, we recommend to run both warm and cold initializations, and choose the one with better convergent objective values...''
     \end{quote}

     On a related note, characterizing the optimization landscape and impact of initialization are problems of interest. Inspired by the reviewer's comment, {\bf we have added new theory for provable initialization and per-iteration accuracy.} See our response to {\bf\# Theoretical Studies}.

     \begin{quote}

\end{quote}

\end{enumerate}

\vspace{0.2cm}
\textit{\textbf{Theoretical Studies:}}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
    \item \textit{The statistical properties established in Theorem 4.1 is for the M-estimator in (8). However, due to non-convexity of this optimization, the maximizer of (8) may not be achievable by practical algorithms. Therefore, it would be more convincing to prove statistical properties for the iterative estimator from the alternating optimization in Algorithm 1. Such theorem is pretty standard in recent tensor regression models with low-rank structure \citep{sun2017store, raskutti2019convex,hao2019sparse,han2020optimal}. If this proof is not possible, it would be helpful if the authors could discuss the technical challenges.}

    {\bf Response: }Thank you for the suggestion. Inspired by your comments, we have now proved the statistical properties for the iterative estimator from the alternating optimization in Algorithm 1. We summarize the new results below; the detailed revisions are in Section~\ref{subsec:statprob}.

     \begin{quote}
    ``...This section presents the accuracy guarantees for both global and local optimizers of~\eqref{eq:MLE}. We first provide the statistical accuracy for the global MLE~\eqref{eq:MLE}. Then, we provide the convergence rate for the local optimizer from Algorithm~\ref{alg:B} with warm initialization. The rate reveals an interesting interplay between statistical and computational efficiency. We show that a polynomial number of iterations suffices to reach the desired accuracy under certain assumptions...'' 
    \end{quote}

     \begin{quote}
``...Under mild conditions, our warm initialization enjoys stable performance, and the subsequent iterations further improve the accuracy via linear convergence; i.e.\ sequence of iterates generated by Algorithm~\ref{alg:B} converges to optimal solutions at a linear rate.

\begin{prop}[Polynomial-time angle estimation]\label{lem:ini} Consider Gaussian tensor models with $b(\theta)=\theta^2/2$ in the objective function~\eqref{eq:loglikelihood}. Suppose the signal-to-noise ratio $\lambda^2/\phi \geq C p^{K/2}d^{-K}$ for some sufficiently large universal constant $C>0$. 
Under Assumption A1 with scaled feature matrices $\bar \mX_k=\sqrt{d}\mX_k$, or Assumption A1' with original feature matrices, the outputs from initialization Algorithm~\ref{alg:A} and iteration Algorithm~\ref{alg:B} satisfy the following two properties.
\begin{enumerate}[label=(\alph*)]
\item With probability at least $1-\exp(-p)$. 
\begin{equation}\label{eq:warm}
\max_{k\in[K]}\sin^2\Theta(\trueM,\hat \mM_k^{(0)})\leq {1\over 4}. 
\end{equation}
\item Let $t=1,2,3,\ldots,$ denote the iteration. There exists a contraction parameter $\rho\in(0,1)$, such that, with probability at least $1-\exp(-p)$, 
\begin{equation}\label{eq:local}
\max_{k\in[K]}\textup{sin}^2 \Theta(\trueM, \hat \mM^{(t)}_k) \lesssim \KeepStyleUnderBrace{{\phi p \over \lambda^2 d^K}}_{\text{statistical error}}+\KeepStyleUnderBrace{\rho^{t}\max_{k\in[K]}\textup{sin} \Theta^2(\trueM, \hat \mM_k^{(0)})}_{\text{algorithmic error}}.
\end{equation}
\end{enumerate}
\end{prop}

Proposition~\ref{lem:ini} provides the estimation errors for algorithm outputs at initialization and at each of the subsequent iterations. The initialization bound~\eqref{eq:warm} demonstrates the stability of QR-adjusted spectral initialization under a mild SNR requirement $\lambda^2/\phi \gtrsim p^{K/2}d^{-K}$. We can think of $d$ as the sample size while $p$ the number of parameters at mode $K$. The condition confirms that a higher sample size mitigates the required signal level. The iteration bound~\eqref{eq:local} consists of two terms: the first term is the statistical error, and the second is the algorithmic error. The algorithmic error decays exponentially with the number of iterations, whereas the statistical error remains the same as $t$ grows. The statistical error is unavoidable and also appears in the global MLE; see Theorem~\ref{thm:MLE}.

As a direct consequence, we find the optimal iteration $t$ after which the algorithmic error is negligible compared to statistical error. 
\begin{thm}[Statistical rate for local optimizers]\label{thm:local}Consider the same condition as in Proposition~\ref{lem:ini} and the outputs by combining algorithms 1 and 2. There exists a constant $C>0$, such that, after $t\gtrsim K\log_{1/\rho}p$ iterations, our algorithm outputs satisfies
\[
\max_{k\in[K]}\textup{sin}^2 \Theta(\trueM, \hat \mM^{(t)}_k) \lesssim {\phi p\over \lambda^2d^K}, \quad \FnormSize{}{\trueB- \hat \tB^{(t)}}^2 \lesssim {\phi(r^K+Kpr)\over d^K}.
\]
\end{thm}

In practice, the signal level $\lambda$ is unknown, so the assumption in Theorem~\ref{thm:local} is challenging to verify in practice. We supply the theory by providing an alternative scheme -- random initialization -- and investigate its empirical performance...''

    \end{quote}

    \item \textit{As shown in Theorem 4.1, the rate of convergence is in the order $\sum_k p_k/ \prod_k d_k$ when other terms fixed. Here $p_k$ is the covariate size along $k$-th mode. I am curious if there is any insight on the optimality of this scaling. Since handling multiple covariates is one of the key contributions of this work, I think more discussion of the order of $p_k$ would be useful. For instance, it would be helpful to compare this rate with the optimal rate established in tensor regression \citep{han2020optimal} and probabilistic tensor decomposition \citep{wang2020learning}. }

    {\bf Response:} We have explained the intuition about our optimality and the benefit of incorporating multiple feature matrices. We added discussions to the scaling factors in Section~\ref{subsec:statprob}:

\begin{quote}
``...The result in~\eqref{eq:bound} implies that the estimation has a convergence rate $\tO(Kp/d^{K})$...This rate agrees with intuition, since in our setting, the number of parameters with $K$ feature matrices is of order $\tO(Kp)$, whereas the number of tensor entries $\tO(d^K)$ corresponds to the total sample size. Because $p\leq d$, our rate is faster than $\tO(d^{-(K-1)})$ obtained by tensor decomposition without features~\citep{wang2020learning}..''

``...The initialization bound~\eqref{eq:warm} demonstrates the stability of warm initialization under a mild SNR requirement $\lambda^2/\phi \gtrsim p^{K/2}d^{-K}$. We can think of $d$ as the sample size while $p$ the number of parameters at mode $K$. This threshold is less stringent than $d^{K/2}$ required for tensor decomposition without multiple features~\citep{han2020optimal,zhang2018tensor}...''
\end{quote}
\end{enumerate}

\vspace{0.2cm}
\textit{\textbf{Numerical comparison:}}
\begin{enumerate}
    \item \textit{In Section 5.2 (simulation) and Section 6.1 (HCP data), only the multiple-response GLM method is compared. This setting seems to be a special case of the network response regression \citep{zhang2018network}, therefore it would be interesting to add it in the comparison. If they are not directly comparable, it would be helpful to discuss it.}

    \textbf{Response:} We have added comparison with 12 methods, including GLSNet~\citep{zhang2018network}, in the new Section~\ref{sec:connection}. For alternative methods in numerical analysis, we have also replaced the earlier old methods by more recent methods, including~\cite{zhang2018network}, in Section~\ref{sec:simulation}. See our response to \textbf{\# 1, Major Comments Summarized by Editor} for full details.

    \item \textit{There are other recent tensor response regression models \citep{li2017parsimonious,sun2017store,raskutti2019convex}. Some of them are missing in the paper. It would be interesting to see the comparison with more recent tensor response regression models in Section 5.3. If they are not directly comparable, it would be helpful to discuss it.}

    \textbf{Response:} We have made changes as suggested. Comparisons are summarized in our response to \textbf{\# 1, Major Comments Summarized by Editor}. 
\end{enumerate}

\vspace{0.2cm}
\textit{\textbf{Minor issues and typos:}}

\begin{enumerate}
     \item \textit{The notation $\alpha$ is used for both the maximum norm constraint on page 13, and the signal level on page 19. Are they same? If not, it would be helpful to use different notations.}

         \textbf{Response:} 
        They are the same. The sentence now reads:
         \begin{quote}
             ``... Here $\alpha>0$ controls the magnitude of the effect size, which is also the maximum norm of coefficient tensor as in~\eqref{eq:p}...."
         \end{quote}

    \item \textit{In the optimization (8), the authors add a maximum norm constraint on the linear predictor to avoid the divergence in the non-Gaussian variance. Can authors provide more explanations on this constraint? }

    \textbf{Response:} We have added the following explanation in Section~\ref{sec:rankM}:
    \begin{quote}
       ``...The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to ensures the existence (boundedness) of MLE. 
The condition precludes the ill-defined MLE when the optimizer of~\eqref{eq:MLE} diverges to $\pm \infty$; this phenomenon may happen in logistic regression when the Bernoulli responses $\{0,1\}$ are perfectly separable by covariates~\citep{wang2020learning}. For Gaussian models, no maximum norm constraint is needed. In Section~\ref{subsec:statprob}, we show that setting $\alpha$ to an extremely large constant does not compromise the statistical rate in quantities of interest. In practice, the unbounded search is often indistinguishable from the bounded search, since the boundary constraint $\onorm{\Theta}_{\infty} \leq \alpha$ would likely never be active. Similar techniques are commonly used in high-dimensional non-Gaussian problems~\citep{wang2020learning,han2020optimal}...''
    \end{quote}

     \item \textit{What’s the tuning range for the ranks $r_1,...,r_K$ in the experiments?}

     \textbf{Response:} 
     For simulation, we have added the following details in Section~\ref{sec:simulation}:
     \begin{quote}
         ``...For each combination, we minimize BIC using a grid search from $(r_1-3, r_2-3, r_3-3)$ to $(r_1+3, r_2+3, r_3+3)$. We remove invalid rank combinations such as $r^2_{\max} \geq \prod_{k=1}^3 r_k$ and use parallel search to reduce the computational cost....'' 
     \end{quote}

 For data application, we have added the following clarification in Section~\ref{sec:data}. Detailed rank selection is provided in a new section {\bf Rank selection for Nations data, Supplementary Notes}. 
   \begin{quote}
   ``... We use BIC as guidance to select the rank of coefficient tensor $\tB$. Since several rank configurations give similar BIC values, we present here the most interpretable results with $\mr=(4,4,4)$. Detailed rank selection procedure is in Supplementary Notes...''
\end{quote}

\item \textit{Typos: 1) line -3 on page 18, $\mM_2 \mX_3$ should be $\mM_2\mX_2$; 2) Table 2, the column names $d = 20$ and $d = 40$ are missing.}

\textbf{Response:} We have made the correction as suggested.

\end{enumerate}

\clearpage
\newpage

\begin{center}
    \textbf{Point-by-point Response to Reviewer 2}
\end{center}

We greatly appreciate your valuable comments and suggestions. We have carefully addressed all the questions. Note that general comments raised by both reviewers were addressed upfront in {\bf pages 1-7 of this response letter, Response to Major Comments Summarized by Editor}. We will make explicit reference to earlier section when needed.

{\it The paper develops a Tucker decomposition technique for tensor data, which can consider side information in each mode. In particular, the model is built upon the exponential family to handle non-Gaussian data, and the method is formulated as a generalized tensor regression problem. An alternating algorithm is proposed to estimate the model parameters. The statistical convergence is derived for the proposed method. The applications to a brain imaging example and an international relation example reveal some interesting structures in data. Overall, the paper is well written and easy to read. The numerical results demonstrate the potential utilities of the proposed method. There are a few places where the paper can be further improved. }

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
    \item \textit{The term “interactive” is used in the title and throughout the paper. However, it is not very clear from the paper what it means to be interactive. It seems the side information is always assumed given and static. In that case, I would recommend removing the term. }

    \textbf{Response:} We have removed the term as suggested.

    \item \textit{There are quite a few non-Gaussian tensor models for various analyses in the literature \citep{tarzanagh2019regularized, li2020generalized, chi2012tensors, hong2020generalized}. A more thorough literature review is warranted. }

    \textbf{Response:} Thank you for pointing out the related literature. The methods \citep{chi2012tensors, hong2020generalized} were reviewed and compared in the first version. We have added detailed literature discussions with 12 more methods and the numerical comparisons. See our response to \textbf{\# 1, Major Comments Summarized by Editor} for the full comparison.

    \item \textit{The proposed method is closely related to the envelope model (especially for Gaussian data)\citep{li2017parsimonious}. It would be of interest to discuss the similarity and difference between the two.}

    \textbf{Response:} Per reviewer's suggestion, we have added more literature reviews and updated the numerical comparisons. See our response to, \textbf{\# 1, Major Comments Summarized by Editor}.

    \item \textit{The dimension requirements seem unnecessarily strong. In particular, the number of features in the side information ($p_k$) is required to be smaller than the dimension of the corresponding mode of the tensor data ($d_k$); the rank of the core tensor ($r_k$) needs to be smaller than $p_k$. In practice, how to handle a large number of features (i.e., $p > d$)? If the number of features is small (e.g., in the HCP example, there are only 2 individual features where age is further discretized as dummy variables), can the rank of the core tensor exceed the feature dimension? More discussions are needed.}

    \textbf{Response:} There are two questions here, one on $r\leq p$, and the other on $p\leq d$. We discuss them separately.

  {\bf Regarding the condition $r\leq p$.} This is not a requirement but rather a natural fact based on definition of tensor rank.  Notice that $p$ is the dimension of parameter tensor of interest, and $r$ is the rank of the parameter tensor. Numerically, one must have $r\leq p$; i.e, the rank of core tensor cannot exceed the feature dimension.

 From the modeling perspective, the fact $r\leq p$ comes from the goal of the problem: we aim to identify tensor factors that are most attributable to features. In the HCP example, our goal is to find the relationship between data tensor  $\tY\in\{0,1\}^{68\times 68\times 136}$ and the feature $\mX\in\mathbb{R}^{4\times 136}$. The fullest linear model has coefficient $\tB\in\mathbb{R}^{68\times 68\times 4}$, with rank upper bounded by 4 on the last mode.

{\bf Regarding the condition $p\leq d$.} This condition requires the feature to have full column rank. We view it as a mild condition, because same condition is also imposed in classical linear regression. In the presence of rank deficiency, we recommend to remove redundant features before applying our method.

We have added more discussions in Sections~\ref{sec:model} and~\ref{subsec:statprob}. 
\begin{quote}
        ``..the identifiability of $\tB$ requires the feature matrices $\mX_k$ are of full column rank with $p_k\leq d_k$. We impose this rank non-deficiency assumption to $\mX_k$; this is a mild condition common in literature~\citep{lock2018supervised, li2017parsimonious, li2020generalized}. In the presence of rank deficiency, we recommend to remove redundant features from $\mX_k$ before applying our method. ..."
\end{quote}
  \begin{quote}
  ``...the requirement $p\leq d$ is necessary to ensure rank non-deficiency of  feature matrices $\mX_k$..''
  \end{quote}

\item \textit{A related question to the above comment: if the side information is little (small $p_k$) or irrelevant to the underlying structure of the tensor data, would the proposed model structure (3) be overly restrictive rather than helpful? In other words, the “supervised tensor factor” $\mX_k\mM_k$ must reside in the column space of the features $\mX_k$. If $\mX_k$ is not well selected, would that lead to a biased decomposition of the original data? In practice, how do we know if a set of features are appropriate to use? The envelope model \citep{li2017parsimonious,lock2018supervised} may offer an alternative perspective by separating the data variation into a material part (related to $\mX_k$) and an immaterial part (irrelevant to $\mX_k$). It’s not clear whether the idea directly applies to the proposed method in this paper since the low-rank model is on the deterministic mean tensor. Some discussions would be helpful.}

    \textbf{Response:} We agree with the basic premise of the reviewer that the our method relies on the choice of feature matrices. In fact, this does not undermine our method's utility. A major point we want to make in response is that the goal of our problem is to estimate the {\it conditional pattern} of $\tY|\mX$, not the {\it overall pattern} of $\tY$. This motivation has guaranteed the validness (unbiasedness) of our factors, even when the $\mX$ is poorly related to $\tY$. The tensor factors we learned are not intended to explain the most variation in the $\tY$; instead, they are intended to capture the variation that is most attributable to features.

    There are methods that jointly impose data variation in conditional component $\tY|\mX$ and residual component $\tY-\tY|\mX$. These methods have the benefits of joint learning the conditional variation of $\mathbb{E}(\tY|\mX)$ and the variation of residuals. The referred methods, Envelope~\citep{li2017parsimonious} and SupCP~\citep{lock2018supervised}, fall into this category. We have added connection and comparison with these two methods in  Section~\ref{sec:connection} and in Section~\ref{sec:simulation}. See our response to {\bf \# 1, Major Comments Summarized by Editor}.

For reviewer's interest, we discuss the practical guidance about ``which method should we use.'' 
\begin{enumerate}
    \item In general, the answer depends on the goal and problem at hand. If the goal is to identify the relationship between tensor data $\tY$ and features $\mX$, then we would recommend our method. In the three examples  illustrated in Section~\ref{sec:motivation} and two data applications in Section~\ref{sec:data}, we are interested in identifying the relationship between data tensor ( brain connectivities, social community patterns) and available features (age, gender). In these cases, direct modeling $\tY|\mX$ yields a higher accuracy, especially when variation in $\tY|\mX$ is weak compared to the overall variation in $\tY$.

    \item If the goal is to identify unknown variation in the data tensor $\tY$, then we recommend Envelope~\citep{li2017parsimonious} and SupCP~\citep{lock2018supervised}. These two methods have the benefits of separating the variations of $\tY$ into explained part by $\mX$ and unexplained part due to unmeasured latent factors.

    \item If the goal is to identify the relationship between $\tY$ and $\mX$ while adjusting for latent factors, then we would recommend generalized mixed-effect tensor decomposition (MSTD). The MSTD combines the advantages of our STD method and the envelope method. We have described MSTD in the original version of Discussion, and we now elaborate the discussion in light of aforementioned work.

    \begin{quote}
    ``(our method)...can be extended by introducing a more general mixed-effect tensor model. For example, in the special case of Gaussian model, we can model the first two moments of data tensor using
\begin{align}\label{eq:extention}
\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)&=\tC\times\mM_1\times \cdots \times \mM_K,\\
 \text{Var}(\tY|\mX_1,\ldots,\mX_K)&=\mPhi_1\otimes \cdots \otimes \mPhi_K,
\end{align}
where $\mPhi_k\in\mathbb{R}^{d_k\times d_k}$ is the unknown covariance matrix on the mode $k\in[K]$. For general exponential family, an additional mean-variance relationship should also be considered. The joint estimation of mean model $\Theta$ and variance model $\mPhi_k$ will lead to more efficient estimation in the presence of unmeasured confounding effects....Suitable regularization such as .... specially-structured covariance~\citep{li2017parsimonious,lock2018supervised} should be considered...''
    \end{quote}
\end{enumerate}

     \item \textit{Identifiability. Is the proposed model (3) identifiable? It seems not from Definition 1 (Equivalence class). If it’s not identifiable, what benefit does the orthogonality constraint on $\mM_k$ induce? How to interpret the estimated parameters? }

     \textbf{Response:} Our method is to estimate the low-rank tensor $\tB$, or equivalently, the core tensor $\tC$ and factors $\mM_k$ that forms $\tB = \tC \times \{ \mM_1,...,\mM_K \}$. The tensor $\tB$ is identifiable while the factors $\{\tC, \mM_1,...,\mM_K \}$ are not. The non-unique factors incur no concern to the interpretation, because all our parameter estimation and interpretation are about $\tB$.

     The orthonormality of $\mM_k$ is imposed purely for technical convenience. This normalization incurs no impacts in our statistical inference, but may help with numerical stability in empirical optimization~\citep{de2000multilinear, kolda2009tensor}.

    We have added the following discussions in Section~\ref{sec:mainmodel} and Section~\ref{sec:global}:   
     \begin{quote}
         ``...We make several remarks about model identifiability. First, the identifiability of $\tB$ requires the feature matrices $\mX_k$ are of full column rank....Second, the decomposition $\tB = \tC\times\{\mM_1,\ldots,\mM_K\}$ are non-unique, as in standard tensor decomposition~\citep{kolda2009tensor}. For any invertible matrices $\mO_k\in\mathbb{R}^{r_k\times r_k}$, $\tB=\tC\times\{\mM_1,\ldots,\mM_K\}=\tC'\times\{\mM_1\mO_1,\ldots,\mM_K\mO_K\}$ are two equivalent parameterizations with $\tC'=\tC\times\{\mO^{-1}_1,\ldots,\mO^{-1}_K\}$ . To resolve this ambiguity, we impose orthonormality to $\mM_k\in\mathbb{O}(p_k,r_k)$ and assess the estimation error of $\mM_k$ using angle distance. The angle distance is invariant to orthogonal rotations due to its geometric definition. See Section~\ref{subsec:statprob} for more details. The orthonormality of $\mM_k$ is imposed purely for technical convenience. This normalization incurs no impacts in our statistical inference, but may help with numerical stability in empirical optimization~\citep{de2000multilinear, kolda2009tensor}.''
     \end{quote}

   \begin{quote}
       ``...Recall that the factor matrices $\mM_k$ are identifiable only up to orthogonal rotations. Therefore, we choose to use angle distance to assess the estimation accuracy of $\mM_k$...''
   \end{quote}

    \item \textit{The maximum norm constraint on the linear predictor Theta is a bit ad hoc. How to choose the upper bound $\alpha$ in practice? How does the value of alpha affect the final results? }

    \textbf{Response:} The maximum norm constraint is a technical condition purely for theoretical analysis. In practice, no choice of $\alpha$ is needed.

    We have added the following explanations in Section~\ref{sec:rankM}:
    \begin{quote}
      ``...The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to ensures the existence (boundedness) of MLE. The condition precludes the ill-defined MLE when the optimizer of~\eqref{eq:MLE} diverges to $\pm \infty$; this phenomenon may happen in logistic regression when the Bernoulli responses $\{0,1\}$ are perfectly separable by covariates~\citep{wang2020learning}. For Gaussian models, no maximum norm constraint is needed. In Section~\ref{subsec:statprob}, we show that setting $\alpha$ to an extremely large constant does not compromise the statistical rate in quantities of interest. In practice, the unbounded search is often indistinguishable from the bounded search, since the boundary constraint $\onorm{\Theta}_{\infty} \leq \alpha$ would likely never be active. Similar techniques are commonly used in high-dimensional non-Gaussian problems~\citep{wang2020learning,han2020optimal}....
    \end{quote}

    \item \textit{The authors propose to use BIC to select ranks via grid search. However, for a high-order tensor, one may have to search over a multi-dimensional grid to find the best rank combination. It seems computationally prohibitive (even for a 3-way tensor). Is there any practical guidance about how to more efficiently select the ranks? What are the computational costs for rank selection in the numerical examples? }

    \textbf{Response:} We utilize parallel computing in BIC search to reduce the computational cost. In the absence of parallel computation, one may speed up the rank search using uni-modal properties of marginal BIC. Specifically, we suggest alternating the search between modes and bracket the possible rank range in one mode while holding others fixed. Once the search region is reduced, a small-scale grid search is performed to identify the minimal BIC. We find such practice gives reasonable rank estimates in the two applications considered.

    We have added the following sentence in  Section~\ref{sec:simulation}:
    \begin{quote}
     ``(In simulation)...For each combination, we minimize BIC using a grid search from $(r_1-3, r_2-3, r_3-3)$ to $(r_1+3, r_2+3, r_3+3)$. We removed invalid rank combinations such as $r^2_{\max} \geq \prod_k r_k$ and used parallel search to reduce the computational cost...''
    \end{quote}

    The BIC selection on real data is detailed in Supplementary Notes, Section~\ref{sec:adata}:
    \begin{quote}
         ``(In real data)...The running time for the rank selection via grid search is 95 secs in total, on an iMac macOS High Sierra 10.13.6 with Intel Core i5 3.8 GHz CPU and 8 GB RAM. This indicates the BIC is feasible in the considered setting. "
    \end{quote}

    \item \textit{ In numerical studies, the authors compare the proposed method with a multiple-response GLM approach and comment that the latter is suboptimal because it cannot account for the correlation among responses. There are methods in the literature that can address the issue \citep{yee2003reduced, luo2018leveraging}. It would be interesting to see some updated comparisons. }

    \textbf{Response:} We have added new numerical comparisons using aforementioned references in the current version. See our response to {\bf \# 1, Major Comments Summarized by Editor}.

    \item \textit{In the HCP example, how are the networks constructed in Fig 6? Will the connectivity change with rotations in the equivalence class? In the Nations data example, what would the clustering results be without accounting for the national features (i.e., directly applying Tucker decomposition to the data)? More clarifications are warranted. }

    \textbf{Response:} Figure~\ref{fig:brain} plots top entries in $\tB_{..i}$ for $i=1,2,3,4$. The connectivity is invariant to the rotation of factor matrices. See our response to {\bf \# 6, Reviewer 1}.

    As suggested by reviewer, we have added comparison between supervised vs.\ unsupervised decomposition for  Nations data analysis. The full results are in Section~\ref{sec:compare1}-\ref{sec:compare2} Supplementary Notes:
    \begin{quote}
 ``...We compare the supervised vs.\ unsupervised decomposition in the {\it Nations} data analysis. Table~\ref{tab:unsup} shows the clustering results based on classical unsupervised Tucker decomposition without the feature matrices. Table~\ref{tab:s1} shows the clustering results based on supervised tensor decomposition ({\bf STD}). Compared with supervised decomposition, the unsupervised clustering loses some interpretation. Similar relations {\it exports} and {\it relexports}, {\it ngo} and {\it relngo} are separated into different clusters...''

\begin{table}[!h]
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{c|l}
     Cluster& \multicolumn{1}{c}{Relations}\\
    \hline
    \multirow{2}{*}{ I}  &\textcolor{YellowOrange}{economicaid}, \textcolor{YellowOrange}{releconomicaid}, \textcolor{YellowOrange}{exportbooks}, \textcolor{YellowOrange}{relexportbooks}, \textcolor{OliveGreen}{weightedunvote}, \textcolor{OliveGreen}{unweightedunvote}, \\
    &\textcolor{YellowOrange}{tourism}, \textcolor{YellowOrange}{reltourism}, \textcolor{YellowOrange}{tourism3}, \textcolor{YellowOrange}{exports}, \textcolor{Blue}{intergovorgs}, \textcolor{Blue}{ngo}, \textcolor{OliveGreen}{militaryalliance}\\
    \hline
    \multirow{2}{*}{ II}  & \textcolor{OliveGreen}{warning}, \textcolor{OliveGreen}{violentactions}, \textcolor{OliveGreen}{militaryactions}, \textcolor{OliveGreen}{duration}, \textcolor{OliveGreen}{severdiplomatic}, \textcolor{OliveGreen}{expeldiplomats}, \textcolor{YellowOrange}{boycottembargo}, \textcolor{BrickRed}{aidenemy}, \\
    &\textcolor{OliveGreen}{negativecomm}, \textcolor{Blue}{accusation}, \textcolor{OliveGreen}{protests}, \textcolor{OliveGreen}{unoffialacts}, \textcolor{BrickRed}{attackembassy}, \textcolor{Blue}{relemigrants}, \textcolor{OliveGreen}{timesincewar}, \textcolor{BrickRed}{lostterritory}, \textcolor{OliveGreen}{dependent}\\
    \hline
    III & \textcolor{OliveGreen}{timesinceally}, \textcolor{OliveGreen}{independence}, \textcolor{OliveGreen}{commonbloc0}, \textcolor{OliveGreen}{blockpositionindex}\\
    \hline
    \multirow{3}{*}{IV}  &\textcolor{YellowOrange}{treaties}, \textcolor{YellowOrange}{reltreaties}, \textcolor{YellowOrange}{officialvisits}, \textcolor{YellowOrange}{conferences}, \textcolor{YellowOrange}{booktranslations}, \textcolor{YellowOrange}{relbooktranslations}\\
    &\textcolor{OliveGreen}{negativebehavior}, \textcolor{Blue}{nonviolentbehavior}, \textcolor{Blue}{emigrants}, \textcolor{Blue}{emigrants3}, \textcolor{Blue}{students}, \textcolor{Blue}{relstudents}, \textcolor{YellowOrange}{relexports}, \textcolor{YellowOrange}{exports3}\\
    & \textcolor{Blue}{relintergovorgs}, \textcolor{Blue}{relngo}, \textcolor{Blue}{intergovorgs3}, \textcolor{Blue}{ngoorgs3}, \textcolor{OliveGreen}{embassy}, \textcolor{OliveGreen}{reldiplomacy}, \textcolor{OliveGreen}{commonbloc1}, \textcolor{OliveGreen}{commonbloc2}\\
    \hline
    \multicolumn{2}{l}{ \footnotesize \fcolorbox{black}{YellowOrange}{\hspace{2mm}} \raisebox{-1mm}{Economics} \quad   \fcolorbox{black}{OliveGreen}{\hspace{2mm}}   \raisebox{-1mm}{Military}\quad \fcolorbox{black}{Blue}{\hspace{2mm}} \raisebox{-1mm}{Organization} \quad 
    \fcolorbox{black}{BrickRed}{\hspace{2mm}} \raisebox{-1mm}{Territory}}
    \end{tabular}
    }
    \caption{Clustering of relations based on unsupervised tensor decomposition. }\label{tab:unsup}
    \end{table}

\begin{table}[!h]
\resizebox{\textwidth}{!}{%
\begin{tabular}{c|l}
 Category & \multicolumn{1}{c}{Relations}\\
\hline
\multirow{2}{*}{ I}   & \textcolor{OliveGreen}{warning}, \textcolor{OliveGreen}{violentactions}, \textcolor{OliveGreen}{militaryactions}, \textcolor{OliveGreen}{duration}, \textcolor{OliveGreen}{negativebehavior}, \textcolor{OliveGreen}{protests}, \textcolor{OliveGreen}{severdiplomatic} \\
&\textcolor{OliveGreen}{timesincewar}, \textcolor{OliveGreen}{commonbloc0}, \textcolor{OliveGreen}{commonbloc1}, \textcolor{OliveGreen}{blockpositionindex},  \textcolor{OliveGreen}{expeldiplomats}\\   
\hline
\multirow{2}{*}{II}& \textcolor{Blue}{emigrants}, \textcolor{Blue}{emigrants3}, \textcolor{Blue}{relemigrants}, \textcolor{Blue}{accusation}, \textcolor{Blue}{nonviolentbehavior}, \textcolor{Blue}{ngoorgs3}, \textcolor{OliveGreen}{commonbloc2}, \textcolor{Blue}{intergovorgs3}\\      
&\textcolor{YellowOrange}{releconomicaid}, \textcolor{Blue}{relintergovorgs}, \textcolor{Blue}{relngo},  \textcolor{Blue}{students}, \textcolor{Blue}{relstudents}, \textcolor{YellowOrange}{economicaid}, \textcolor{OliveGreen}{negativecomm}, \textcolor{OliveGreen}{militaryalliance} \\
\hline
\multirow{3}{*}{ III} & \textcolor{YellowOrange}{treaties}, \textcolor{YellowOrange}{reltreaties}, \textcolor{YellowOrange}{officialvisits}, \textcolor{YellowOrange}{exportbooks}, \textcolor{YellowOrange}{relexportbooks}, \textcolor{YellowOrange}{booktranslations}, \textcolor{YellowOrange}{relbooktranslations}\\
&\textcolor{YellowOrange}{boycottembargo}, \textcolor{OliveGreen}{weightedunvote}, \textcolor{OliveGreen}{unweightedunvote}, \textcolor{YellowOrange}{reltourism}, \textcolor{YellowOrange}{tourism}, \textcolor{YellowOrange}{tourism3}, \textcolor{YellowOrange}{exports}, \textcolor{YellowOrange}{exports3}  \\
&\textcolor{YellowOrange}{relexports}, \textcolor{Blue}{intergovorgs}, \textcolor{Blue}{ngo}, \textcolor{OliveGreen}{embassy}, \textcolor{OliveGreen}{reldiplomacy}, \textcolor{OliveGreen}{timesinceally}, \textcolor{OliveGreen}{independence}, \textcolor{YellowOrange}{conferences}, \textcolor{OliveGreen}{dependent}\\
\hline
IV &\textcolor{BrickRed}{aidenemy}, \textcolor{BrickRed}{lostterritory}, \textcolor{OliveGreen}{unoffialacts}, \textcolor{BrickRed}{attackembassy}\\
\hline
\multicolumn{2}{l}{ \footnotesize \fcolorbox{black}{YellowOrange}{\hspace{2mm}} \raisebox{-1mm}{Economics} \quad   \fcolorbox{black}{OliveGreen}{\hspace{2mm}}   \raisebox{-1mm}{Military}\quad \fcolorbox{black}{Blue}{\hspace{2mm}} \raisebox{-1mm}{Organization} \quad 
    \fcolorbox{black}{BrickRed}{\hspace{2mm}} \raisebox{-1mm}{Territory}}
\end{tabular}
}
\caption{Clustering of relations based on supervised tensor decomposition. }\label{tab:s1}
\end{table}

\end{quote}

We comment that, although both methods return clustering outputs, the unsupervised and supervised decomposition tackle different problems. The unsupervised decomposition identifies factors that explain most variation in the tensor, whereas the supervised decomposition identifies factors that are most attributable to side features. There is in general no particular pattern one could expect between these two methods. We provide a simple example here for illustration.
\begin{example}[Complementary information in supervised vs.\ unsupervised factors]
Consider the following data tensor $\tY$ and one-sided feature matrix $\mX$, 
\begin{align}
    \tY = \me_1\otimes \me_1 \otimes \me_1+10\me_2\otimes \me_2 \otimes \me_2, \quad \mX=\me_1,
\end{align}
where $\me_i=(0,\ldots,0, 1,0,\ldots,0)^T$ is the $i$th canonical basis vector in $\mathbb{R}^d$ for $i=1,2$. 
Now, consider the unsupervised vs.\ supervised decomposition of $\tY$ with rank $\mr = (1,1,1)$. 
Then, the top supervised and unsupervised factors are perpendicular to each other, 
\[
\mM_{\text{sup},k} \perp \mM_{\text{unsup},k}, \quad \text{for all } k=1,2,3,
\]
where $\mM_{\text{sup},k}$, $\mM_{\text{unsup},k}$ denote the mode-$k$ factors from supervised and unsupervised decompositions, respectively. 
\end{example}

\begin{rmk}
This example shows complementary information between factors from supervised vs.\ unsupervised decompositions. In general, one could construct examples such that these two methods return {\bf arbitrarily different} factors.   
\end{rmk}

\end{enumerate}

\bibliography{tensor_wang}
\bibliographystyle{apalike}

\end{document}

### 2021-08-31 tensor_regress/JCGS/publication/jcgs_main/jcgs-tensor.tex
update

--- 2021-06-21 tensor_regress/JCGS/Response_submission/manuscript/jcgs-tensor.tex
@@
 }
-\myexternaldocument{supp}
+
+%\myexternaldocument{supp}
 
@@
 {
-  \title{\bf Supervised tensor decomposition with features on multiple modes}
-  \author{Jiaxin Hu, Chanwoo Lee, and Miaoyan Wang\thanks{The authors gratefully acknowledge NSF grant DMS-1915978 and funding from the Wisconsin Alumni Research Foundation. }\hspace{.2cm}\\
+  \title{\bf Generalized tensor decomposition with features on multiple modes}
+  \author{Jiaxin Hu, Chanwoo Lee, and Miaoyan Wang\thanks{Corresponding author: Miaoyan Wang, miaoyan.wang@wisc.edu. }\hspace{.2cm}\\
     Department of Statistics, University of Wisconsin-Madison}
+
+    \date{}
   \maketitle
@@
   \begin{center}
-    {\LARGE\bf Supervised tensor decomposition with features on multiple modes}
+    {\LARGE\bf Generalized tensor decomposition with features on multiple modes}
 \end{center}
@@
 The package and data used are available at~\url{https://CRAN.R-project.org/package=tensorregress}.
-Supplementary materials for this article are available online .
+Supplementary materials for this article are available online.
 
@@
 {The first line is a class of}\label{change:firstline} \textit{unsupervised} tensor decomposition such as classical Tucker and CP decomposition~\citep{de2000multilinear, kolda2009tensor} and generalized decomposition for non-Gaussian data \citep{chi2012tensors, tarzanagh2019regularized,hong2020generalized,  li2020generalized}.
-{\color{blue}Regardless of the implementation, the unsupervised methods aim to find the best low-rank representation of a data tensor alone.
+{  Regardless of the implementation, the unsupervised methods aim to find the best low-rank representation of a data tensor alone.
 In contrast, our model is a \textit{supervised} tensor learning, which aims to identify the association between a data tensor and multiple features.
@@
 The second line of work studies the tensor-to-tensor regression.
-{\color{blue}This category is further divided into three scenarios, depending on whether tensor is treated as predictors~\citep{zhou2013tensor,raskutti2019convex,han2020optimal}, as responses~\citep{li2017parsimonious,sun2017store,zhang2018network,lock2018supervised,luo2018leveraging}, or both~\citep{lock2018tensor,gahrooei2020multiple}.
+{  This category is further divided into three scenarios, depending on whether tensor is treated as predictors~\citep{zhou2013tensor,raskutti2019convex,han2020optimal}, as responses~\citep{li2017parsimonious,sun2017store,zhang2018network,lock2018supervised,luo2018leveraging}, or both~\citep{lock2018tensor,gahrooei2020multiple}.
 As we show in Section~\ref{sec:connection}, our supervised tensor decomposition falls into this general category, and we provide a \emph{provable} solution in new settings that have broader practical significance.
@@
 Our goal is to estimate the low-rank tensor $\tB$, or equivalently, the core tensor and factors $(\tC, \mM_1,\ldots,\mM_K)$, from our model~\eqref{eq:decomp}.
-{\color{blue}We make several remarks about model identifiability.
+{  We make several remarks about model identifiability.
 First, the identifiability of $\tB$ requires the feature matrices $\mX_k$ are of full column rank with $p_k\leq d_k$.
@@
 The maximum norm constraint on the linear predictor $\Theta$ is a technical condition to ensures the existence (boundedness) of MLE. 
-{\color{blue}The condition precludes the ill-defined MLE when the optimizer of~\eqref{eq:MLE} diverges to $\pm \infty$; this phenomenon may happen in logistic regression when the Bernoulli responses $\{0,1\}$ are perfectly separable by covariates~\citep{wang2020learning}.
+{  The condition precludes the ill-defined MLE when the optimizer of~\eqref{eq:MLE} diverges to $\pm \infty$; this phenomenon may happen in logistic regression when the Bernoulli responses $\{0,1\}$ are perfectly separable by covariates~\citep{wang2020learning}.
 For Gaussian models, no maximum norm constraint is needed.
@@
 
-{\color{blue}We provide two initialization schemes, one with QR-adjusted spectral initialization (warm initialization), and the other with random initialization (cold initialization).
+{  We provide two initialization schemes, one with QR-adjusted spectral initialization (warm initialization), and the other with random initialization (cold initialization).
 The warm initialization is an extension of unsupervised spectral initialization~\citep{zhang2018tensor} to supervised setting with multiple feature matrices.} Specifically, we project normalized data tensor $\bar \tY$ to the normalized multilinear feature space and obtain an unconstrained coefficient tensor $\hat\tB^{(0)}$.
@@
 
-{\color{blue}The warm initialization enjoys provable accuracy guarantees at a cost of extra technical assumptions (see Section~\ref{subsec:statprob}).
+{  The warm initialization enjoys provable accuracy guarantees at a cost of extra technical assumptions (see Section~\ref{subsec:statprob}).
 The cold initialization, on the other hand, shows robust in practice but its theoretical guarantee remains an open challenge~\citep{luo2021low}.
@@
 \subsection{Statistical accuracy}\label{subsec:statprob}
-{\color{blue}This section presents the accuracy guarantees for both global and local optimizers of~\eqref{eq:MLE}.
+{  This section presents the accuracy guarantees for both global and local optimizers of~\eqref{eq:MLE}.
 We first provide the statistical accuracy for the global MLE~\eqref{eq:MLE}.
@@
 We are particularly interested in the high-dimensional regime in which both $d$ and $p$ grows while $p\leq d$.
-{\color{blue}The requirement $p\leq d$ is necessary to ensure rank non-deficiency of  feature matrices $\mX_k$.} The classical MLE theory is not directly applicable, because the number of unknown parameters grows with the size of data tensor.
+{  The requirement $p\leq d$ is necessary to ensure rank non-deficiency of  feature matrices $\mX_k$.} The classical MLE theory is not directly applicable, because the number of unknown parameters grows with the size of data tensor.
 We leverage the recent development in random tensor theory and high-dimensional statistics to establish the error bounds of the estimation.
@@
 \[
-\lambda=\min_{k\in[K]}\sigma_{\min}(\text{Unfold}_k(\tB_{\text{true}})).
+\lambda=\min_{k\in[K]}\sigma_{r}(\text{Unfold}_k(\tB_{\text{true}})).
 \]
@@
 The result in~\eqref{eq:bound} implies that the estimation has a convergence rate $\tO(Kp/d^{K})$ as $(p,d)\to \infty$.
-{\color{blue}This rate agrees with intuition, since in our setting, the number of parameters with $K$ feature matrices is of order $\tO(Kp)$, whereas the number of tensor entries $\tO(d^K)$ corresponds to the total sample size.
+{  This rate agrees with intuition, since in our setting, the number of parameters with $K$ feature matrices is of order $\tO(Kp)$, whereas the number of tensor entries $\tO(d^K)$ corresponds to the total sample size.
 Because $p\leq d$, our rate is faster than $\tO(d^{-(K-1)})$ obtained by tensor decomposition without features~\citep{wang2020learning}.}
@@
 The optimization~\eqref{eq:MLE} is a non-convex problem due to the low-rank constraint in the feasible set $\tP$.
-{\color{blue}Under mild conditions, our warm initialization enjoys stable performance, and the subsequent iterations further improve the accuracy via linear convergence; i.e.\ sequence of iterates generated by Algorithm~\ref{alg:B} converges to optimal solutions at a linear rate.
+{  Under mild conditions, our warm initialization enjoys stable performance, and the subsequent iterations further improve the accuracy via linear convergence; i.e.\ sequence of iterates generated by Algorithm~\ref{alg:B} converges to optimal solutions at a linear rate.
 
@@
 \section{Connection to other tensor regression methods}\label{sec:connection}
-{\color{blue}We compare our supervised tensor decomposition (\textbf{STD}) with recent 12 tensor methods in the literature.
+{  We compare our supervised tensor decomposition (\textbf{STD}) with recent 12 tensor methods in the literature.
 Table~\ref{table:comp_table} summarizes these methods with their properties from four aspects: i) model specification, ii) number of feature matrices allowed, (iii) capability of addressing non-Gaussian response, and (iv) capability of addressing non-independent noise.
@@
 \centering
-\color{blue}
+
 \resizebox{\columnwidth}{!}{
@@
 
-{\color{blue}\begin{itemize}[noitemsep,topsep=0pt]
+{  \begin{itemize}[noitemsep,topsep=0pt]
 \item Supervised PARAFAC/CANDECOMP factorization (\textbf{SupCP}, \citep{lock2018supervised}).
@@
 
-{\color{blue}These four methods are the closest methods to ours, in that they all relate a data tensor to feature matrices with low-rank structure on the coefficients.} We consider Gaussian and Bernoulli tensors in experiments.
+{  These four methods are the closest methods to ours, in that they all relate a data tensor to feature matrices with low-rank structure on the coefficients.} We consider Gaussian and Bernoulli tensors in experiments.
 For methods not applicable for Bernoulli data ({\bf SupCP} and {\bf Envelope}), we provide the algorithm $\{-1,1\}$-valued tensors as inputs.
@@
 
-{\color{blue}The comparison is assessed from three aspects: (i) benefit of incorporating features from multiple modes; (ii) prediction error with respect to sample size; (iii) robustness of model misspecification.} We use similar simulation setups as in our first experiment in last section.
+{  The comparison is assessed from three aspects: (i) benefit of incorporating features from multiple modes; (ii) prediction error with respect to sample size; (iii) robustness of model misspecification.} We use similar simulation setups as in our first experiment in last section.
 We consider rank $\mr=(3,3,3)$ (low) vs.\ $(4,5,6)$ (high), effect size $\alpha = 3$ (low) vs.\ $6$ (high), dimension $d$ ranging from 20 to 100 for modes with features, and $d = 20$ for modes without features.
@@
 
-{\color{blue}Figure~\ref{fig:comp}a-b shows the impact of features to estimation error.
+{  Figure~\ref{fig:comp}a-b shows the impact of features to estimation error.
 We see that our \textbf{STD} outperforms others, especially in the low-effect high-rank setting.
@@
 The multi-relational network $\tY$ is a binary data tensor, and the country attributes $\mX\in\mathbb{R}^{14\times 6}$ are features on both the 1$^\text{st}$ and 2$^\text{nd}$ modes.
-{\color{blue} We use BIC as guidance to select the rank of coefficient tensor $\tB$.
+{   We use BIC as guidance to select the rank of coefficient tensor $\tB$.
 Since several rank configurations give similar BIC values, we present here the most interpretable results with $\mr=(4,4,4)$.
@@
 Then we apply $K$-mean clustering to dimension reduction matrix on each of the modes.
-Supplementary Table~\ref{fig:s1} shows the K-means clustering of the 56 relations based on the dimension reduction matrix on the 3$^\text{rd}$ mode.
+Table S6 in Supplementary Notes shows the K-means clustering of the 56 relations based on the dimension reduction matrix on the 3$^\text{rd}$ mode.
 We find that the relations reflecting the similar aspects of actions are grouped together.
@@
 To investigate the effects of dyadic attributes towards connections, we depict the estimated coefficients $\hat \tB=\entry{\hat b_{ijk}}$ for several relation types (Figure~\ref{fig:est}).
-{\color{blue}The entry $\hat b_{ijk}$ estimates the contribution, at the logit scale, of feature pair $(i,j)$ ($i$th feature for the ``sender'' country and $j$th feature for the ``receiver'' country) towards the connection of relation $k$.} Several interesting findings emerge from the observation.
+{  The entry $\hat b_{ijk}$ estimates the contribution, at the logit scale, of feature pair $(i,j)$ ($i$th feature for the ``sender'' country and $j$th feature for the ``receiver'' country) towards the connection of relation $k$.} Several interesting findings emerge from the observation.
 We find that relations belonging to a same cluster tend to have similar feature effects.
@@
 \begin{align}\label{eq:extention}
-\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)&=\tC\times\mM_1\times \cdots \times \mM_K,\\
+\mathbb{E}(\tY|\mX_1,\ldots,\mX_K)&=\tC\times\{\mX_1\mM_1, \cdots, \mX_K \mM_K\},\\
  \text{Var}(\tY|\mX_1,\ldots,\mX_K)&=\mPhi_1\otimes \cdots \otimes \mPhi_K,
@@
 However, the introduction of unknown covariance matrices $\mPhi_k$ dramatically increases the number of parameters in the problem.
-{\color{blue}Suitable regularization such as graphical lasso or specially-structured covariance~\citep{li2017parsimonious,lock2018supervised} should be considered.} The extension of tensor modeling with heterogeneous mixed-effects will be an important future direction.
+{  Suitable regularization such as graphical lasso or specially-structured covariance~\citep{li2017parsimonious,lock2018supervised} should be considered.} The extension of tensor modeling with heterogeneous mixed-effects will be an important future direction.
 
@@
 
-\bigskip
+\begin{center}
+{\large \bf ACKNOWLEDGEMENT}
+\end{center}
+
+This research was supported by NSF grants DMS-1915978, DMS-2023239, and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
+We thank Zhuoyan Xu for the help with the software.
+
 \begin{center}
@@
 
-\setstretch{1.25}
-\bibliographystyle{apalike}
+\setstretch{1.2}
 
-\bibliography{tensor_wang}
+\input{jcgs-tensor.bbl}
 

### 2021-11-02 IFDS/ifds21Sra_proposal/main.tex
proposal

--- 2020-08-12 tensor_regress/unsolved_review.tex
@@
 \usepackage{multirow}
+\usepackage{extarrows}
 \usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
@@
 \mathtoolsset{showonlyrefs=true}
-
-\newcommand{\of}[1]{\left(#1\right)}
-\newcommand{\off}[1]{\left[#1\right]}
-\newcommand{\offf}[1]{\left\{#1\right\}}
-\newcommand{\aabs}[1]{\left|#1\right|}
-\usepackage{colonequals}
 
@@
 \setcounter{table}{0}
+
+\newcommand{\of}[1]{\left(#1\right)}
+\newcommand{\off}[1]{\left[#1\right]}
+\newcommand{\offf}[1]{\left\{#1\right\}}
+\newcommand{\aabs}[1]{\left|#1\right|}
 
@@
 
-\title{Future Improvement}
+\setlength{\parskip}{0.8em}
+
+\title{IFDS 2021 Spring RA research proposal}
 
@@
 
-% Abstracts are required.
-\section{Unsolved Review Summary}
+Multiway arrays are collected in many fields including social networks \citep{anandkumar2014tensor} and computer science \citep{koniusz2016sparse}, and tensors represent the multiway data efficiently.
+One example is \emph{hypergraph} network~\citep{ghoshdastidar2017uniform,ghoshdastidar2017consistency,ahn2019community,ke2019community} in social science.
+A $K$-uniform hypergraph can be naturally represented as an order-$K$ tensor, in which each entry indicates the presence of $K$-way hyperedge among nodes.
+Identifying the similarity among nodes is of scientific interests.
+We refer such clustering as \emph{higher-order clustering}.
+A huge amounts methods have been developed for matrix clustering while applying matrix methods to tensors results in sub-optimal performances.
+Unsupervised tensor models and tools for higher-order clustering have arisen recently~\citep{ wang2019multiway,chi2020provable,han2020exact}.
 
-\subsection{Comparisons}
-\begin{enumerate}
-	\item Add comparison on tensor-on-tensor regression~\citep{lock2018tensor,llosa2018tensor,gahrooei2020multiple,raskutti2015convex}. {\bf NIPS R3}.
-	\item Add comparison on scalar-on-tensor regression~\citep{chen2019non, zhou2013tensor}. {\bf ICML R1}.
-	\item Add comparison on HOLRR~\citep{rabusseau2016low}.
-Though HOLRR is a special case of the proposed model, it is helpful to show the benefit of the incorporation of covariate on distinct modes, when all covariates are vectorized in a long vector in HOLRR.
-{\bf AISTAT R3}.
-	\item Add comparison on HOPLS~\citep{zhao2012higher} and TPG \citep{yu2016learning} from the perspective of the model. {\bf AISTAT R1}.
-	\item Add comparison on paper~\citep{smith2015positive}. {\bf NIPS R4}.
-\end{enumerate}
+Meanwhile, semi-supervised learning has been widely applied to clustering problems~\citep{zhu2003semi,yang2014unified}.
+The potential useful prior information such as a small proportion of labeled data and auxiliary features are well-incorporated in the learning tasks, which is believed to improve clustering performance than unsupervised methods.
+Semi-supervised learning are also used in tensor decomposition~\citep{cao2016semi,hu2021generalized}.
+However, applying semi-supervised techniques in higher-order clustering is still new to the field.
+Several questions such as how to incorporate prior information in tensor models and how the prior information affects the clustering performance are waiting to be answered.
 
-\subsection{Experiments}
-\begin{enumerate}
-	\item Add comparisons on other scalar-on-tensor and tensor-on-tensor regressions in both simulations and real data analysis. {\bf NIPS R2, R3}.
-	\item Add numerical results on large-scale or higher-order response. {\bf NIPS R2, AISTAT R3}.
-	\item Show how many iterations and time are needed to converge. {\bf NIPS R2}.
-\end{enumerate}
+Therefore, this project will focus on the application of semi-supervised learning in higher-order clustering problem.
+The work will enrich the methodology in higher-order clustering literature and deepen the understanding of semi-supervised learning in higher-order regime, which tackles the main theme of IFDS.
 
-\subsection{Explanations}
-\begin{enumerate}
-	\item Add explanations on hyperparameter selection for our method and other methods (e.g.\ HOLRR, TPG, HOPLS) we want to compare. {\bf NIPS R2,R4}.
-	\item Add explanations on rank selection, including relative supplements.
-Explain why not use the proposed rank selection strategy in simulations.
-Explain the computational issues of grid search.
-{\bf NIPS R3, ICML R1, R3}.
-	\item Discuss the existence of $\tB_{true}$ and discuss the mismatching loss $\sum \hat \tB_{ijk} \neq {\tB_{true}}_{ijk}$. {\bf NIPS R4}.
-	\item Compare the computational complexity with other methods. {\bf NIPS R2}.
-	\item Explain how the 136 subjects in HCP are selected from thousands of subjects in the original data. {\bf NIPS R3}.
-\end{enumerate}
+\textbf{Bio:}
 
-\section{Model Equivalence}
+Jiaxin Hu is a second-year Ph.D.\ student in Statistics at the University of Wisconsin-Madison, advised by Prof.
+Miaoyan Wang.
+Her research mainly focuses on the statistical machine learning methods for matrix/tensor data analysis.
+She actively attended the weekly IFDS seminars and the summer school and workshop at IFDS in August, 2021.
+During the workshop, she gave a five minute lightening talk with title ``Supervised tensor decomposition with features on multiple modes".
 
-\subsection{Our method}
-Let $\tY = \entry{\tY_{i_1,...,i_k}} \in \bbR^{d_1,...,d_K}$ be the tensor observation, $\mX_k \in \bbR^{d_k \times p_k}, k \in [K]$ be the matrix covariates. Our tensor regression model is of form
-\begin{equation}\label{ourmodel}
-	g \of{ \bbE[\tY |\mX_1 ,...,\mX_K ] } = \tB \times_1 \mX_1 \times_2 \cdots \times_K \mX_K,
-\end{equation}
-where $g(\cdot)$ is the inverse link function, $\tB \in \bbR^{p_1 \times \cdots \times p_K}$ is the order-K tensor-valued coefficient.
-Reformulating the model~\eqref{ourmodel} into $\prod_{i=1}^K d_k$ univariate regressions, we have
-\begin{equation}\label{ourmodel1}
-	g \of{ \bbE[\tY_{i_1,...,i_K} |\mX_1 ,...,\mX_K ] }  = \langle \tB, \mX_{1i_1} \circ \cdots \circ \mX_{K i_K}  \rangle, \quad \text{for } i_k \in [d_k], k\in[K],
-\end{equation}
-where $\mX_{k i_k}$ refers to the $i_k$-th row of $\mX_k$, for all $k \in [K]$, $\langle \cdot, \cdot \rangle$ refers to the inner product of two tensors, and $\circ$ refers to the outer product of two vectors.
-
-\subsection{Scalar-on-tensor regression}
-Let $y \in \bbR$ denote the scalar response, $\tX \in \bbR^{p_1 \times \cdots \times p_K}$ denote the order-K tensor predictor, and $\mZ \in \bbR^{p_0}$ denote the vector-valued covariate.
-Given the i.i.d.\ sample set $\{y^{(i)}, \tX^{(i)}, \mZ^{(i)}\}_{i=1}^n$,  the scalar-on-tensor regression model is
-\begin{equation}\label{scalar}
-	g \of{\bbE[y^{(i)} | \tX^{(i)}]}  = \alpha + \gamma^T \mZ + \langle \tB, \tX^{(i)}  \rangle, \quad \text{for } i  = 1 ,...,n
-\end{equation}
-
-where $g(\cdot)$ is the inverse link function, $\alpha \in \bbR$ is the scalar coefficient, $\gamma \in \bbR^{p_0}$ is the vector-valued coefficient, and $\tB \in \bbR^{p_1 \times \cdots \times p_K}$ is the order-K tensor-valued coefficient.
-
-Ignoring the scalar parameter $\alpha$ and the vector-valued covariate $\mZ$, our model~\eqref{ourmodel1} is equal to the scalar-on-tensor model~\eqref{scalar} by defining
-\begin{equation}
-	\bbE[y^{(i)} | \tX^{(i)}] \colonequals \bbE[\tY_{i_1,...,i_K} |\mX_1 ,...,\mX_K ] \quad \text{and} \quad \tX^{(i)} \colonequals \mX_{1i_1} \circ \cdots \circ \mX_{K i_K}, \quad \text{for } i = 1,..., n,
-\end{equation}
-where $n = \prod_{i=1}^K d_k$.
-
-\subsection{Tensor-on-tensor regression}
-Let $\mathbb{Y} \in \bbR^{ d_{K+1} \times \cdots \times d_{N}}$ be the tensor observation, $\mathbb{X} \in \bbR^{ d_1 \times \cdots \times d_K}$ be the tensor predictor.
-Given the i.i.d.\ sample set $\{\mathbb{Y}^{(i)}, \mathbb{X}^{(i)} \}_{i=1}^n$,  the tensor-on-tensor regression is of form
-\begin{equation}\label{tensor}
-	g \of{\bbE[\mathbb{Y}^{(i)} | \mathbb{X}^{(i)}]} = \langle \mathbb{X}^{(i)}, \mathbb{B}   \rangle_K, \quad \text{for } i = 1,...,n,
-\end{equation}
-
-where $g(\cdot)$ is the inverse link function, $\mathbb{B} \in \bbR^{d_1 \times \cdots \times d_N}$ is the order-K tensor-valued coefficient, $\langle \cdot, \cdot \rangle_K$ refers to the contracted tensor product; i.e., for two tensors $\mathbb{P} \in \bbR^{I_1 \times \cdots \times I_L \times d_1 \times \cdots \times d_K}$ and $\mathbb{Q} \in \bbR^{d_1 \times \cdots \times d_K \times J_1 \times \cdots \times J_M}$, the contracted tensor product $\langle \mathbb{P}, \mathbb{Q} \rangle_K \in \bbR^{I_1\times \cdots \times I_L \times J_1 \times \cdots \times J_M}$ and the $(i_1,...,i_L, j_1,...,j_M)$-the entry is
-\begin{equation}
-	(\langle \mathbb{P}, \mathbb{Q} \rangle_K)_{i_1,...,i_L, j_1,...,j_M} = \sum_{a_1,...,a_K}^{d_1,...,d_K} \mathbb{P}_{i_1,...,i_L, a_1,...,a_K} \mathbb{Q}_{a_1,...,a_K, j_1,...,j_M}.
-\end{equation}
-
-Let $N = K$.
-Then, the contracted product becomes usual inner product, and $\mathbb{Y}^{(i)}$ becomes a scalar for $i \in [n]$.
-Consequently, the model~\eqref{tensor} degenerates to the scalar-on-tensor regression~\eqref{scalar}, which is equal to our model.
-
-\section{Iteration time table}
-
-See Table~\ref{tab:time}.
-\begin{table}[h]
-\centering
-%\resizebox{1\textwidth}{!}{
-\begin{tabular}{|c|c|c|c|}
-\hline
-Setting &Gaussian& Bernoulli & Poisson  \\ \hline
-$d = 30, r = 6$& 2.6 &  6.0& 5.7\\
-$d = 30, r = 3$&  0.5 & 1.1&1.3\\
-$d = 25, r = 6$&1.4 & 3.7& 4.0\\
-$d = 25, r = 3$&0.3& 0.6& 0.6\\
-\hline
-\end{tabular}
-%}
-\caption{ Iteration time of Algorithm 1. Numerical values are the running time (seconds) required for one iteration in Algorithm 1 under different settings.  }\label{tab:time}
-\end{table}
-
-\bibliography{tensor_wang}
+\bibliography{tensor_wang.bib}
 \bibliographystyle{apalike}

### 2021-11-02 IFDS/ifds21Sra_proposal/main.tex
revised proposal

@@
 
-Multiway arrays are collected in many fields including social networks \citep{anandkumar2014tensor} and computer science \citep{koniusz2016sparse}, and tensors represent the multiway data efficiently.
+My research interests lie in the statistical machine learning and matrix/tensor data analysis.
+I am currently a second-year Ph.D.\ student in Statistics.
+I am planning to work under the supervision of Prof.
+Miaoyan Wang (Ph.D.\ advisor) and Prof.
+Xiaojin (Jerry) Zhu (co-mentor).
+
+I have been working on statistical methods for tensor analysis, and I plan to continue the research on tensor methodology in the next semester.
+
+\textbf{I plan to develop a semi-supervised higher-order clustering method.} Higher-order clustering problem aims to identify the similarity among entities (e.g.\ people in the social network) via multiway arrays, which is a higher-order extension of one-dimensional clustering and matrix biclustering.
+Amount of unsupervised tensor models and tools for higher-order clustering have arisen recently.
+Meanwhile, semi-supervised learning method is widely employed in vector and matrix clustering with prior information.
+The potential useful prior information such as a small proportion of labeled data and auxiliary features is believed to improve clustering performance than unsupervised methods.
+However, applying semi-supervised techniques in higher-order clustering is still new to the field.
+I will propose a semi-supervised higher-order clustering method, and answer the two key questions: (1) how to incorporate prior information in tensor models; (2) how the prior information affects the clustering performance.
+I expect the proposed method will enrich the higher-order clustering methodology and deepen the understanding of semi-supervised learning in higher-order regime.
+
+\textbf{I also plan to actively participate the IFDS events and present my research work at IFDS next semester.} I have regularly attended IFDS seminars since 2019.
+I also attended IFDS summer school and presented my previous work on generalized tensor decomposition during the lighting talk session of the IFDS-MADLab workshop this August.
+I will keep interacting with IFDS faculty/members and contributing to the community.
+
+% Multiway arrays are collected in many fields including social networks \citep{anandkumar2014tensor} and computer science \citep{koniusz2016sparse}, and tensors represent the multiway data efficiently.
 One example is \emph{hypergraph} network~\citep{ghoshdastidar2017uniform,ghoshdastidar2017consistency,ahn2019community,ke2019community} in social science.
@@
 
-Meanwhile, semi-supervised learning has been widely applied to clustering problems~\citep{zhu2003semi,yang2014unified}.
+% Meanwhile, semi-supervised learning has been widely applied to clustering problems~\citep{zhu2003semi,yang2014unified}.
 The potential useful prior information such as a small proportion of labeled data and auxiliary features are well-incorporated in the learning tasks, which is believed to improve clustering performance than unsupervised methods.
@@
 
-Therefore, this project will focus on the application of semi-supervised learning in higher-order clustering problem.
+% Therefore, this project will focus on the application of semi-supervised learning in higher-order clustering problem.
 The work will enrich the methodology in higher-order clustering literature and deepen the understanding of semi-supervised learning in higher-order regime, which tackles the main theme of IFDS.
 
-\textbf{Bio:}
+% \textbf{Bio:}
 
-Jiaxin Hu is a second-year Ph.D.\ student in Statistics at the University of Wisconsin-Madison, advised by Prof.
+% Jiaxin Hu is a second-year Ph.D.\ student in Statistics at the University of Wisconsin-Madison, advised by Prof.
 Miaoyan Wang.
 Her research mainly focuses on the statistical machine learning methods for matrix/tensor data analysis.
-She actively attended the weekly IFDS seminars and the summer school and workshop at IFDS in August, 2021.
+She actively attended the weekly IFDS seminars and the IFDS summer school and IFDS-MADLab workshop this August, 2021.
 During the workshop, she gave a five minute lightening talk with title ``Supervised tensor decomposition with features on multiple modes".

### 2021-12-17 code_sbm/sbm_gao/refinement_sbm.R
organize

@@ -1,92 +1,0 @@
-### refinement in sbm
-source("/Users/March/Desktop/myresearch/code/sbm_gao/bricks_sbm.R")
-
-library(matrixcalc)
-
-
-refinement_sbm = function(A, k, trim = TRUE, tau = Inf,mu){
-  # A is adjacency
-  # k is # of clusters
-  # trim, tau, mu is for initialization
-  
-  # A = sbm_data$x
-  # plot(A)
-  # k = 3
-  
-  n = dim(A)[1]
-  
-  sigma_t = matrix(0,nrow = n,ncol = n)
-  
-  for (i in 1:n) {
-    
-    #get A-u
-    A_sub = A[-i,]; A_sub = A_sub[,-i]
-    
-    sigmau = rep(0,n)
-    #need to change
-    sigmau[-i] = greedy_initial_clustering(A_sub,k,trim, tau,mu)
-    
-    B_est = matrix(0,nrow = k,ncol = k)
-    for (j in 1:(k-1)) {
-      Cj = which(sigmau == j)
-      if(length(Cj) <= 1){
-        B_est[j,j] = A[Cj,Cj]
-      } else{
-        B_est[j,j] = sum(A[Cj,Cj])/(2*length(Cj)*(length(Cj) - 1))
-      }
-      
-      for (l in (j+1):k) {
-        Cl = which(sigmau == l)
-        B_est[l,j] = sum(A[Cl,Cj])/(length(Cj)*length(Cl))
-      }
-    }
-    
-    Ck = which(sigmau == k)
-    if(length(Ck) <= 1){
-      B_est[k,k]  = A[Ck,Ck]
-    } else{
-      B_est[k,k] = sum(A[Ck,Ck])/(2*length(Ck)*(length(Ck) - 1))
-    }
-    
-    # now B_est is a lower triangle matrix
-    B_est[upper.tri(B_est)] = t(B_est)[upper.tri(B_est)]
-    
-    a_est = n*min(diag(B_est))
-    #B_off_est = B_est; diag(B_off_est) = 0
-    b_est = n*max(B_est[lower.tri(B_est)])
-    
-    if(b_est >= a_est){
-      b_est = a_est - 0.01
-    }
-    
-    #calculate rho
-    tu = (1/2)*log((a_est*(1-b_est/n))/(b_est*(1-a_est/n))) 
[…]

### 2021-12-17 others/code_sbm/sbm_gao/refinement_sbm.R

@@ -1,0 +1,92 @@
+### refinement in sbm
+source("/Users/March/Desktop/myresearch/code/sbm_gao/bricks_sbm.R")
+
+library(matrixcalc)
+
+
+refinement_sbm = function(A, k, trim = TRUE, tau = Inf,mu){
+  # A is adjacency
+  # k is # of clusters
+  # trim, tau, mu is for initialization
+  
+  # A = sbm_data$x
+  # plot(A)
+  # k = 3
+  
+  n = dim(A)[1]
+  
+  sigma_t = matrix(0,nrow = n,ncol = n)
+  
+  for (i in 1:n) {
+    
+    #get A-u
+    A_sub = A[-i,]; A_sub = A_sub[,-i]
+    
+    sigmau = rep(0,n)
+    #need to change
+    sigmau[-i] = greedy_initial_clustering(A_sub,k,trim, tau,mu)
+    
+    B_est = matrix(0,nrow = k,ncol = k)
+    for (j in 1:(k-1)) {
+      Cj = which(sigmau == j)
+      if(length(Cj) <= 1){
+        B_est[j,j] = A[Cj,Cj]
+      } else{
+        B_est[j,j] = sum(A[Cj,Cj])/(2*length(Cj)*(length(Cj) - 1))
+      }
+      
+      for (l in (j+1):k) {
+        Cl = which(sigmau == l)
+        B_est[l,j] = sum(A[Cl,Cj])/(length(Cj)*length(Cl))
+      }
+    }
+    
+    Ck = which(sigmau == k)
+    if(length(Ck) <= 1){
+      B_est[k,k]  = A[Ck,Ck]
+    } else{
+      B_est[k,k] = sum(A[Ck,Ck])/(2*length(Ck)*(length(Ck) - 1))
+    }
+    
+    # now B_est is a lower triangle matrix
+    B_est[upper.tri(B_est)] = t(B_est)[upper.tri(B_est)]
+    
+    a_est = n*min(diag(B_est))
+    #B_off_est = B_est; diag(B_off_est) = 0
+    b_est = n*max(B_est[lower.tri(B_est)])
+    
+    if(b_est >= a_est){
+      b_est = a_est - 0.01
+    }
+    
+    #calculate rho
+    tu = (1/2)*log((a_est*(1-b_est/n))/(b_est*(1-a_est/n))) 
[…]

### 2021-12-17 skills/CHTC/my_rscript.R

@@ -1,383 +1,0 @@
-library(pracma)
-library(rTensor)
-library(quadprog)
-args <- (commandArgs(trailingOnly=TRUE))
-cat(args[1])
-if(length(args) == 1){
-    BATCH <- as.numeric(args[1])
-} else {
-    stop()
-}
-
-
-Makepositive = function(mat){
-  h = eigen(mat,symmetric = T)
-  nmat = (h$vectors)%*%diag(pmax(h$values,10^-4),nrow=nrow(mat))%*%t(h$vectors)
-  return(nmat)
-}
-
-
-Karray = function(X,kernel = function(X1,X2) t(X1)%*%X2, type="col"){
-  N = length(X);
-  if(type=="row"){
-    Xt=list();
-    for(i in 1:N){
-      Xt[[i]]=t(X[[i]])
-    }
-    X=Xt
-  }
-  
-  m= nrow(X[[1]]); n = ncol(X[[1]]);
-  K = array(dim = c(N,N,n,n))
-  for(i in 1:N){
-    for(j in 1:N){
-      K[i,j, , ] = kernel(X[[i]],X[[j]])
-    }
-  }
-  return(as.tensor(K))
-}
-
-# K = function(A,B,kernel = function(a,b) sum(a*b),type = c("row","col")){
-#   if(nrow(A) != nrow(B)) stop("two matrix sizes are different")
-#   if(ncol(A) != ncol(B)) stop("two matrix sizes are different")
-#   d1 = nrow(A); d2 = ncol(A)
-#   if (type == "row") {
-#     Kab = matrix(nrow= d1,ncol = d1)
-#     for(i in 1:d1){
-#       for(j in 1:d1){
-#         Kab[i,j] = kernel(A[i,],B[j,])
-#       }
-#     }
-#   }else{
-#     Kab = matrix(nrow= d2,ncol = d2)
-#     for(i in 1:d2){
-#       for(j in 1:d2){
-#         Kab[i,j] = kernel(A[,i],B[,j])
-#       }
-#     }
-#   }
-#   return(Kab)
-# }
-
-expkernel = function(Y,Z){
-  n = ncol(Y)
-  A = matrix(0,n,n)
-  for(i in 1:n){
-    for(j in 1:n){
-      A[i,j] = t(Y[,i]-Z[,j])%*%(Y[,i]-Z[,j])
-    }
-  }
-  return(exp(-A))
-}
-
[…]

### 2021-12-20 tensor_clustering/tensor_clustering_ieee/ieee.tex
ieee manuscript

\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
%\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
% updated with editorial comments 8/9/2021

\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage[authoryear]{natbib}
\usepackage{url}
\usepackage{enumitem}

\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{algorithm}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor=blue,citecolor=blue,urlcolor={black}}
\usepackage{lscape}
\usepackage{afterpage}
\usepackage{hyperref}

\usepackage{microtype}
\usepackage{wrapfig}
\allowdisplaybreaks
\usepackage{algorithm}
\usepackage{algpseudocode}

\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}[section]
\newtheorem{pro}{Property}
\newtheorem{cor}{Corollary}

\theoremstyle{definition}
\newtheorem{assumption}{Assumption}
\newtheorem{defn}{Definition}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}
\newtheorem{condition}{Condition}
\newtheorem{conjecture}{Conjecture}
\newtheorem{property}{Property}
\newtheorem{hypothesis}{Hypothesis}

\newtheorem{innercustomgeneric}{\customgenericname}
\providecommand{\customgenericname}{}
\newcommand{\newcustomtheorem}[2]{%
  \newenvironment{#1}[1]
  {%
   \renewcommand\customgenericname{#2}%
   \renewcommand\theinnercustomgeneric{##1}%
   \innercustomgeneric
  }
  {\endinnercustomgeneric}
}

\newcustomtheorem{customexample}{Example}

%\usepackage{appendix}
\usepackage{fullpage} 
\usepackage{microtype}
\usepackage{wrapfig}
\mathtoolsset{showonlyrefs}

\newcommand{\of}[1]{\left(#1\right)}
\newcommand{\off}[1]{\left[#1\right]}
\newcommand{\offf}[1]{\left\{#1\right\}}
\newcommand{\aabs}[1]{\left|#1\right|}
\newcommand{\ang}[1]{\left\langle#1\right\rangle}

\newcommand{\distgap}{D_{\ell_2}}
\newcommand{\Mat}{\text{Mat}}

\def\ci{\perp\!\!\!\perp}

\usepackage{booktabs}
\newcommand\doubleRule{\toprule\toprule}
\allowdisplaybreaks

\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}

\input macros.tex
\newcommand*{\Scale}[2][4]{\scalebox{#1}{$#2$}}%
\setlength\parindent{0pt}
\usepackage[parfill]{parskip}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsmath}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\newcommand\Algphase[1]{%
\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\columnwidth}{0.4pt}%
\Statex\hspace*{-\algorithmicindent}\textbf{#1}%
\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\columnwidth}{0.4pt}%
}
\usepackage{caption}
\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}

\begin{document}

\title{Multiway Spherical Clustering via \\ Degree-Corrected Tensor Block Models}

\author{ Jiaxin Hu and Miaoyan Wang
%IEEE Publication Technology,~\IEEEmembership{Staff,~IEEE,}
        % <-this % stops a space
%\thanks{This paper was produced by the IEEE Publication Technology Group. They are in Piscataway, NJ.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2021; revised August 16, 2021.}
}

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2021}%
%{Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}

%\IEEEpubid{0000--0000/00\$00.00~\copyright~2021 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

\maketitle

\begin{abstract}
We consider the problem of multiway clustering in the presence of unknown degree heterogeneity. Such data problems arise commonly in applications such as recommendation system, neuroimaging, community detection, and hypergraph partitions in social networks. The allowance of degree heterogeneity provides great flexibility in clustering models, but the extra complexity poses significant challenges in both statistics and computation. Here, we develop a degree-corrected tensor block model with estimation accuracy guarantees. We present the phase transition of clustering performance based on the notion of angle separability, and we characterize three signal-to-noise regimes corresponding to different statistical-computational behaviors. In particular, we demonstrate that an intrinsic statistical-to-computational gap emerges only for tensors of order three or greater. Further, we develop an efficient polynomial-time algorithm that provably achieves exact clustering under mild signal conditions. The efficacy of our procedure is demonstrated through two data applications, one on human brain connectome project, and another on Peru Legislation network datasets. 
\end{abstract}

\begin{IEEEkeywords}
tensor clustering, degree correction, statistical-computational efficiency, human brain connectome networks
\end{IEEEkeywords}

{
\color{red}

\textbf{Other conference paper to journal}
\begin{enumerate}
    \item Garvesh. Add more discussion about Theorem 5 and empirical simulations.
    %\item Eric Chi. Add one more remakr; One more section about the numerical issue; More comparison with other methods; add more future discussion.
    \item Anru. Add application in introduction; more remarks for algorithm and theorems; One more section about the application to a specific model; more experiments and discussions.
\end{enumerate}

}

\section{Introduction}\label{sec:intro}

\fixme{Jiaxin}{Change the wording: \emph{individual/block effects} $\rightarrow$ \emph{individual/block parameters}, \emph{degree heretogeneity} $\rightarrow$ \emph{non-equal degrees}.}

\IEEEPARstart{M}{ultiway} arrays have been widely collected in various fields including social networks \citep{anandkumar2014tensor}, neuroscience \citep{wang2017bayesian}, and computer science \citep{koniusz2016sparse}. Tensors effectively represent the multiway data and serve as the foundation in higher-order data analysis. One data example is from multi-tissue multi-individual gene expression study~\citep{wang2019three,hore2016tensor}, where the data tensor consists of expression measurements indexed by (gene, individual, tissue) triplets. Another example is \emph{hypergraph} network~\citep{ghoshdastidar2017uniform,ghoshdastidar2017consistency,ahn2019community,ke2019community} in social science. A $K$-uniform hypergraph can be naturally represented as an order-$K$ tensor, where each entry indicates the presence of $K$-way hyperedge among nodes (a.k.a.\ entities). In both examples, identifying the similarity among tensor entities is important for scientific discovery.

%To overcome the high-dimensionality, low-rankness is one of the popular structures utilized in tensor data analysis. Specifically, for an order-$K$ tensor $
%\tY$ with low-rank, we have the following decomposition:
%\begin{equation}\label{eq:low_rank}
 %   \tY = \tS \times_1 \mM_1 \cdots \times_K \mM_K, 
%\end{equation}
%where $\tS$ is the low-dimensional core tensor, $\mM_1, ..., \mM_K$ refer to the factor matrices, and $\times_k$ denotes the tensor-to-matrix product.

This paper studies the problem of multiway clustering based on a data tensor. The goal of multiway clustering is to identify a checkerboard structure from a noisy data tensor. Figure~\ref{fig:intro} illustrates the noisy tensor and the underlying structures discovered by multiway clustering methods. The checkerboard structure serves as a meta tool to many popular structures including the low-rankness~\citep{young2018universality}, latent space models~\citep{wang2018learning}, and isotonic models~\citep{pananjady2020isotonic}. In the hypergraph example, the multiway clustering aims to identify the underlying block partition of nodes based on their higher-order connectivities; therefore, we also refer to the clustering as \emph{higher-order clustering}. The most common model for higher-order clustering is called \emph{tensor block model} (TBM)~\citep{wang2019multiway}, which extends the usual matrix stochastic block model~\citep{abbe2017community} to tensors. 
%The analysis tools developed for matrices,
The matrix analysis tools, however, are sub-optimal for higher-order clustering. Developing tensor tools for solving block models has received increased interest recently~\citep{ wang2019multiway,chi2020provable,han2020exact}.

Classical tensor block model suffers from drawbacks to model real world data in spite of the popularity. The key underlying assumption of block model is that all nodes in the same community are exchangeable; i.e., the nodes have no individual effects apart from the block effects. However, the exchangeability assumption is often non-realistic. Each node may contribute to the data variation by its own multiplicative effect. Such degree heterogeneity appears commonly in social networks. For example, regular block model fails to model the Karate Club network~\citep{bickel2009nonparametric} without addressing
%accounting for 
degree heterogeneity. 
\fixme{Jiaxin}{Non-degree model ignoring the individual heterogeneity may seriously mislead the clustering results. Karate Club data analysis shows that clustering without heterogeneity tends to assign membership based on the degree of communication (two clusters for leaders and non-leaders) rather than the affiliation (two clusters for people from different clubs). }

The \emph{degree-corrected tensor block model} (dTBM) has been proposed recently to account for the degree heterogeneity~\citep{ke2019community}. The dTBM combines a higher-order checkerboard structure with degree parameter $\mtheta=(\mtheta(1),\ldots,\mtheta(p))^T$ to allow heterogeneity among $p$ nodes.  Figure~\ref{fig:intro} compares the underlying structures of TBM and dTBM with the same number of communities. The dTBM allows varying values within the same community, thereby allowing a richer structure.

{
\color{red}
We use the idea of \emph{spherical clustering} in initialization and iteration. Using angle-based similarity avoids the estimation of $\mtheta$.  
}

\begin{figure*}[t]
    \centering
    \includegraphics[width = .9\textwidth]{intro2_arxiv.pdf}
    \caption{Examples for order-3 tensor block model (TBM) with and without degree correction. Both TBM and dTBM have four communities on each mode, while dTBM allows a richer structure with degree heterogeneity.
    %\fixme{Miaoyan}{Use horizontal figure} 
    }
    \label{fig:intro}
\end{figure*}

{\bf Our contributions.} The primary goal of this paper is to provide both statistical and computational guarantees for dTBM. Our main contributions are summarized below.
\begin{itemize}[leftmargin=*]

 \item We develop a general dTBM and establish the identifiability for the uniqueness of clustering using the notion of angle seperability.

        \item  We present the phase transition of clustering performance with respect to three different statistical and computational behaviors.  We characterize, for the first time, the critical signal-to-noise (SNR) thresholds in dTBMs, revealing the intrinsic distinctions among (vector) one-dimensional clustering, (matrix) biclustering, and (tensor) higher-order clustering. Specific SNR thresholds and algorithm behaviours are depicted in  Figure~\ref{fig:phase_axis}.

    \item We provide an angle-based algorithm that achieves exact clustering \emph{in polynomial time} under mild conditions. Simulation and data studies demonstrate the outperformance of our algorithm compared with existing higher-order clustering algorithms. 
\end{itemize}
The last two contributions, to our best knowledge, are new to the literature of dTBMs. %Table~\ref{tab:comp} highlights the improvement of our methods over previous results. Our algorithm addresses higher-order tensor data with degree heterogeneity, allows either binary or continuous entries, and achieves exponentially fast rate in misclustering error.

\begin{figure*}[t]
    \centering
    \includegraphics[width = 17cm]{phase.pdf}
    \caption{SNR thresholds for statistical and computational limits in order-$K$ dTBM with dimension $(p,...,p)$ and $K \geq 2$. The SNR gap between statistical possibility and computational efficiency  exists only for tensors with $K \geq 3$. }
    \label{fig:phase_axis}
\end{figure*}

{\bf Related work.} %Our work is closely related to, but also distinct from several lines of existing research. 
We emphasize the comparisons that set our work apart from earlier literature from two perspectives---models and algorithms. From the model perspective, our work extends the previous degree-corrected model from matrices to tensors. There is a huge literature on degree-corrected matrix models; see the review paper~\citep{abbe2017community} and the references therein. The tensor counterparts, however, are relatively less understood. Table~\ref{tab:comp} summarizes  the most relevant models for higher-order clustering. Earlier tensor methods either fail to allow degree heterogeneity~\citep{chien2018community,han2020exact,zhang2021exact,ghoshdastidar2017uniform,kim2018stochastic}, or suffer from sub-optimal misclustering rates~\citep{ke2019community,ghoshdastidar2017consistency,chi2020provable}. In contrast, our method addresses the degree heterogeneity, allows discrete and continuous entries, and achieves exponentially fast rate in clustering tasks.

From the algorithm perspective, our methods generalize the recent global-to-local strategy for matrix learning~\citep{gao2018community,chi2019nonconvex,yun2016optimal} to tensors~\citep{han2020exact,ahn2018hypergraph,kim2018stochastic}. Despite the conceptual similarity, we address several fundamental challenges associated with this non-convex, non-continuous problem. We show the insufficiency of the conventional tensor HOSVD~\citep{kolda2009tensor}, and we develop a weighted higher-order initialization that relaxes the eigen-gap separation condition. Furthermore, our local iteration leverages the angle-based clustering in order to avoid explicit estimation of degree heterogeneity. Our bounds reveal the interesting interplay between the computational and statistical errors. We show that our final estimate provably achieves the exact clustering within only polynomial-time complexity.

{
\color{red}

Our work is closely related to, but also distinct from several lines of existing research:
\begin{enumerate}

\item \textbf{Degree-corrected model.}
    \begin{itemize}
        \item Matrix. (1) Previous work, survey paper (2) Unique technical challenges in Tensor (3) matrix does not show statistical-computational gap;
        \item Tensor. (Comparison with Ke) (1) Ke does not provide the SNR limits (2) Ke's algorithm is sub-optimal (3) Ke focuses on Binary data. 
    \end{itemize}

    \item \textbf{Higher-order clustering.} (1) Previous work fails to account the heterogeneity, which is not realistic; (2) The extension of theoretical results highly challenging.  Extra nuisance parameter makes it more difficult to establish the statistical and computational limits.

    \item \textbf{Spherical clustering.}

    \item \textbf{Global-to-local algorithm.} (1) Matrix global-to-local algorithm (2) Tensor global-to-local algorithm is used for non-degree algorithm. Extra nuisance parameter makes it necessary to consider angle-based similarity; angle-based clustering makes the Cartesian coordinates based analysis non-applicable for our setting. 
\end{enumerate}

}

\begin{table*}[t]
\resizebox{\textwidth}{!}{%
    \begin{tabular}{c|ccccc}
\hline
    & \cite{gao2018community}&  \cite{han2020exact}& \cite{ke2019community}&  \cite{ghoshdastidar2017consistency} & \textbf{Ours}\\
    \hline
        {\small Allow tensors}& $\times$ & $\surd$ & $\surd$ &$\surd$ & $\surd$  \\
        {\small Allow heterogeneity} &  $\surd$ & $\times$ & $\surd$ &$\surd$ & $\surd$ \\
        %Assortative assumption&  $\surd$ & $\times$ & $\times$ & $\times$\\
       {\small Allow various data} & $\times$ & $\surd$ & $\times$ & $\surd$  &$\surd$\\
        {\small Misclustering rate (for order 2$^*$)}& $\exp(-p)$ & $\exp(-p)$ & $p^{-1}$ & $p^{-1}$& $\exp(-p)$\\
        \hline
    \end{tabular}
    }
    \caption{Comparison between previous methods with our method. $^*$We list here only the result for order-2 tensors (i.e., matrix). Our results allow general tensors of arbitrary order $K$; See Theorem~\ref{thm:refinement}.}\label{tab:comp}
\end{table*}

{
\color{red}
Organization.
}

{\bf Notation.} We use lower-case letters (e.g., $a,b$) for scalars, lower-case boldface letters (e.g., $\ma,\mtheta$) for vectors, upper-case boldface letters (e.g., $\mX,\mY$) for matrices, and calligraphy letters (e.g., $\tX,\tY$) for tensors of order three or greater. We use $|\cdot|$ for the cardinality of a set and $\ind\{\cdot\}$ for the indicator function. For an integer $p\in\mathbb{N}_{+}$, we use the shorthand $[p]= \offf{1,2,...,p}$. For a length-$p$ vector $\ma$, we use $a(i)\in\mathbb{R}$ to denote the $i$-th entry of $\ma$, and use $\ma_{I}$ to denote the sub-vector by restricting the indices in the set $I\subset [p]$.  We use  $\onorm{\ma}=\sqrt{\sum_{i}a^2(i)}$ to denote the $\ell_2$-norm, $\onorm{\ma}_1=\sum_i |a_i|$ to denote the $\ell_1$ norm of $\ma$. For two vector $\ma, \mb$ of the same dimension, we denote the angle between $\ma, \mb$ by 
\begin{equation}
    \cos \of{\ma, \mb} = \frac{\ang{ \ma, \mb}}{ \onorm{\ma} \onorm{\mb} },
\end{equation}
where $\ang{\ma,\mb}$ is the inner product of two vectors and $\cos \of{\ma, \mb} \in [-1,1]$. We make the convention that $\cos \of{\ma, \mb} = \cos \of{\ma^T, \mb^T}$.

For a matrix $\mY$, we use $\mY_{i:}$ to  denote the $i$-th row of the matrix. Let $\tY  \in \bbR^{p_1 \times \cdots \times p_K}$ be an order-$K$ $(p_1,...,p_K)$-dimensional tensor. We use $\tY(i_1,\ldots,i_K)$ to denote the $(i_1,\ldots,i_K)$-th entry of $\tY$. The multilinear multiplication of a tensor $\tS\in \bbR^{r_1\times \cdots \times r_K}$ by matrices $\mM_k \in\mathbb{R}^{p_k\times r_k}$ results in an order-$d$ $(p_1,\ldots,p_K)$-dimensional tensor $\tX$, denoted
\[
\tX=\tS \times_1 \mM_1 \times \cdots \times_K \mM_K,
\]
where the entries of $\tX$ are defined by
\begin{align}
    &\tX(i_1,\ldots, i_K)\\
    &=\sum_{(j_1,\ldots,j_K)}\tS(j_1,\ldots,j_K)\mM_1(i_1,j_1)\cdots \mM_K(i_K,j_K).
\end{align} 
 We use $\text{Ave}(\cdot)$ to denote the operation of taking averages across elements and $\text{Mat}_k(\cdot)$ to denote the unfolding operation that reshapes the tensor along mode $k$ into a matrix. For a symmetric tensor $\tY\in\mathbb{R}^{p\times \cdots \times p}$, we omit the subscript and use $\text{Mat}(\tY)\in\mathbb{R}^{p\times p^{K-1}}$ to denote the unfolding. For two sequences $\{a_p\}, \{b_p\}$, we denote $a_p\lesssim b_p$ if $\lim_{p\to\infty} a_p/b_p \leq c$ and $a_p = \Omega(b_p)$ if $c b_p \leq a_p \leq C b_p$, for some constants $c, C\geq 0$. Throughout the paper, we use the terms ``community'' and ``clusters'' exchangeably. 
 %\fixme{Miaoyan}{\color{red} need to define $\cos(\ma,\mb)$. Be careful with signs.}

\section{Model formulation}\label{sec:model}

\subsection{Degree-corrected tensor block model}

Suppose we have an order-$K$ data tensor $\tY \in \bbR^{p \times \cdots \times p}$. For ease of notation, we focus on symmetric tensors in this section; our framework easily extends to general asymmetric tensors. Assume there exist $r \geq 2$  disjoint communities among the $p$ nodes. We represent the community assignment by a function $z \colon [p]\mapsto[r]$, where $z(i) = a$ for $i$-th node that belongs to the $a$-th community. Then, $z^{-1}(a)=\{i\in[p]\colon z(i)=a\}$ denotes the set of nodes that belong to the $a$-th community, and $|z^{-1}(a)|$ denotes the number of nodes in the $a$-th community.
%, and we let $p_a = \sum_{i \in [p]}\ind \offf{ z_i= a}$ denote the size of $a$-th community. 
Let $\mtheta=(\theta(1),\ldots,\theta(p))^T$ denote the degree heterogeneity for $p$ nodes. We consider the order-$K$ dTBM~\citep{ghoshdastidar2017consistency,ke2019community},
\begin{equation}\label{eq:model_margin}
    \tY(i_1,\ldots,i_K) = \tS(z(i_1),\ldots,z(i_K)) \prod_{k = 1}^K\theta_{i_k} + \tE(i_1,\ldots,i_K), 
\end{equation}
\normalsize
where $\tS \in \bbR^{r \times \cdots \times r}$ is an order-$K$ tensor collecting the block means among communities, and $\tE\in\mathbb{R}^{p\times \cdots \times p}$ is a noise tensor consisting of independent mean-zero sub-Gaussian entries with variance bounded by $\sigma^2$. The unknown parameters are $z$, $S$, and $\mtheta$. The dTBM can be equivalently written in a compact form of tensor-matrix product:
\begin{equation}\label{eq:model_tensor}
 \mathbb{E}\tY = \tS \times_1 \mTheta \mM \times_2 \cdots \times_K  \mTheta \mM,
\end{equation}
where $\mTheta = \text{diag}(\theta(1),...,\theta(p)) \in \bbR^{p \times p}$ is a diagonal matrix, $\mM \in \offf{0,1}^{p \times r}$ is the membership matrix associated with community assignment $z$ such that $\mM(i,j)=\ind\{z(i)=j\}$. By definition, each row of $\mM$ has one copy of 1's and 0's elsewhere. Note that the discrete nature of $\mM$ renders our model~\eqref{eq:model_tensor} more challenging than Tucker decomposition. %The Tucker rank of the signal tensor $\bbE \tY$ is bounded by, but not necessarily equal to, $(r,...,r)$~\citep{wang2019multiway}. 
We call a tensor $\tX$ an $r$-block tensor with degree $\mtheta$
%with the core tensor $\tS$ 
if $\tX$ admits dTBM~\eqref{eq:model_tensor}. Here, we give two special cases of dTBM.
\begin{example}[Gaussian TBM]
Let $\theta(i)= 1$ for all $i \in [p]$ and $\tE$ be a noise tensor with i.i.d.\ $N(0,\sigma^2)$ entries. Our dTBM reduces to a non-degree Gaussian TBM, which is widely used in previous clustering algorithms~\citep{wang2019multiway,chi2020provable,han2020exact}. The theoretical results in TBM serve as benchmarks for dTBM. 
\end{example}

\begin{example}[Binary dTBM]
Consider a $K$-uniform hypergraph $H = (V,E)$, where $V = [p]$ collects the nodes with $r$ disjoint communities and $E$ collects all the $K$-way hyperedges. Let $\tY \in \{0,1\}^{p \times \cdots \times p}$ denote the adjacency tensor, where the entries encode the presence or absence of hyperdeges among $p$ nodes. Specifically, let $\tY(i_1,...,i_K) = 1$ if  $(i_1,...,i_K) \in E$, otherwise, $\tY(i_1,...,i_K) = 0$, for all $(i_1,\ldots,i_K) \in [p]^K$. The equation~\eqref{eq:model_tensor} models $\mathbb{E}\tY$ with unknown degree heterogeneity and subgaussianity parameter $\sigma^2 = 1/4$. 
\end{example}

\begin{rmk}[Comparison with non-degree models]
Our dTBM uses fewer block parameters than TBM. Let the subscripts ``deg" and ``non" denote quantities in the models with and without degrees, respectively. Then, every $r_{\text{non}}$-block tensor can be represented by a degree-corrected $r_{\text{deg}}$-block tensor with $r_{\text{deg}}\leq r_{\text{non}}$. In particular, there exist tensors with $r_{\text{non}}=p$ but $r_{\text{deg}}=1$, so the reduction in $r$ can be dramatic from $p$ to 1. This fact highlights the benefits of introducing degree heterogeneity in higher-order clustering tasks.
\end{rmk}

% In addition, the following example highlights the benefits of degree correction in clustering tasks. 
% \begin{example}[Comparison to non-degree models]\label{ex:benefit}
% Every $r_{\text{non}}$-block tensor can be represented by degree-corrected $r_{\text{deg}}$-block tensor with $r_{\text{deg}}\leq r_{\text{non}}$. In particular, there exist tensors with $r_{\text{non}}=p$ but $r_{\text{deg}}=1$.
% \end{example}
% %In principle, every $(p,\ldots,p)$-dimensional tensor can be represented by a $p$-block tensor.
% The above example shows that our dTBM uses fewer block parameters than TBM, and the reduction in $r$ can be dramatic from $p$ to 1. This fact shows the advantages of introducing degree heterogeneity in high-dimensional clustering.

\subsection{Identifiability under angle gap condition}
The goal of clustering is to estimate the partition function $z$ from model~\eqref{eq:model_tensor}. We use $\tP$ to denote the following parameter space for $(z,\tS,\mtheta)$,
\footnotesize
\begin{align}\label{eq:family}
\tP&=\bigg\{(z,\tS,\mtheta)\colon  \ \mtheta\in\mathbb{R}^p_{+},\ 
{c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r},\\& c_3\leq \onormSize{}{\text{Mat}(\tS)_{a:}}\leq c_4, \onorm{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|, a\in[r]\bigg\},
\end{align}
\normalsize
where $c_i>0$'s are universal constants.  
%We briefly describe the rationale of the constraints in~\eqref{eq:family}. 
First, the entrywise positivity constraint on  $\mtheta\in\mathbb{R}^p_{+}$ is imposed to avoid sign ambiguity between entries in $\mtheta_{z^{-1}(a)}$ and $\tS$. This constraint can be achieved without sacrificing model flexibility, by using a slightly larger dimension of $\tS$ in the factorization~\eqref{eq:model_tensor}; see Example~\ref{ex:positive}. Second, 
%the quantity $|z^{-1}(a)|$ denotes the number of nodes in $a$-th community. T
the constants $c_1, c_2$ in the $|z^{-1}(a)|$ bound assume the roughly balanced size across $r$ communities. Third, the constants $c_3, c_4 $ in the magnitude of $\text{Mat}(\tS)_{a:}$ requires no purely zero slide in $\tS$, so the core tensor $\tS$ is not trivially reduced to lower rank. Lastly, the $\ell_1$ normalization $\onormSize{}{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|$ is imposed to avoid the scalar ambiguity between $\mtheta_{z^{-1}(a)}$ and $\tS$. This constraint, again, incurs no restriction to model flexibility but makes our presentation cleaner.

\begin{example}[Positivity of degree parameters]~\label{ex:positive}
Here we provide an example to show the positivity constraints on $\mtheta$ incurs no loss on the model flexibility. Without loss of generality, we consider the order-2 (i.e., matrix) block models. Consider a $4$-by-$4$ matrix with core matrix $\mS = 1$ and degree $\mtheta=(1,1,-1,-1)^T$. We have the mean matrix 
\begin{equation}
    \mX = \mS \times_1 \mTheta \mM \times_2 \mTheta\mM = \begin{bmatrix}
1 & 1 & -1& -1\\
1 & 1 & -1& -1\\
-1 & -1 & 1& 1\\
-1 & -1 & 1& 1
\end{bmatrix},
\end{equation}
$ \text{with}\quad \mTheta=\text{diag}(\mtheta)$. Note that $\mX$ is a 1-block matrix with mixed-signed degree $\mtheta$. 
Now, instead of original decomposition, we encode $\mX$ as a $2$-block matrix with positive-signed degree,
\begin{equation}
\mX= \mS' \times_1 \mTheta' \mM \times_2 \mTheta' \mM,\quad \text{with}\quad   \mS' = \begin{bmatrix} 1 & -1 \\
    -1 & 1
    \end{bmatrix}
\end{equation}
and $ \mTheta'=\text{diag}(\mtheta')=\text{diag}(1, 1, 1, 1)^T$, where the triplet $(z',\mS', \mtheta')$ lies in our parameter space~\eqref{eq:family}. In general, we can always reparameterize a block-$r$ tensor with mixed-signed degree using a block-$2r$ tensor with positive-signed degree. Since we assumes $r=\tO(1)$ throughout the paper, the splitting does not affect the error rates of our interest.
\end{example}

We first provide the identifiablity conditions for our model before estimation procedures. When $r=1$, the decomposition~\eqref{eq:model_tensor} is always unique (up to cluster label permutation) in $\tP$, because dTBM is equivalent to the rank-1 tensor family under this case. When $r\geq 2$, the Tucker rank of signal tensor $\mathbb{E}\tY$ in~\eqref{eq:model_tensor} is bounded by, but not necessarily equal to, the number of blocks $r$~\citep{wang2019multiway}. Therefore, one can not apply the classical identifiability conditions for low-rank tensors to dTBM. Here, we introduce a key separation condition on the core tensor.

\begin{assumption}[Angle gap] \label{assmp:min_gap}Let $\mS = \text{Mat}(\tS)$. Assume the minimal gap between normalized rows of $\mS$ is bounded away from zero, i.e.,
\begin{equation}\label{eq:minimal_gap}
    \Delta_{\min} := \min_{a \neq b \in [r]} \onorm{\frac{\mS_{a:}}{\onormSize{}{ \mS_{a:}}} - \frac{\mS_{b:}}{\onormSize{}{ \mS_{b:}}} }>0, \quad \text{for} \quad r \geq 2,
\end{equation}
and set the convention $\Delta_{\min} = 1$ for $r = 1$.
    Equivalently, none of the two rows in $\mS$ are parallel, i.e., when $r \geq 2$, $ \max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}}  = 1-\Delta^2_{\min}/2<1$.
\end{assumption}
The quantity $\Delta_{\min}$ characterizes the non-redundancy among clusters measured by angle separation. The definition~\eqref{eq:minimal_gap} is well posed because of the lower bound on $\onorm{\mS_{a:}}$ in~\eqref{eq:family}. The following theorem shows that the angle gap separation is sufficient and necessary for parameter identifiability under dTBM.

\begin{thm}[Identifiability]\label{thm:unique} Consider the dTBM with $r\geq 2$. The parameterization~\eqref{eq:model_tensor} is unique in $\tP$ up to cluster label permutations, if and only if Assumption~\ref{assmp:min_gap} holds.
\end{thm}

The identifiability guarantee for the dTBM is more appealing than classical Tucker model. In the Tucker model, the factor matrix $\mM$ is identifiable only up to orthogonal rotations. In contrast, our model does not suffer from rotational invariance. As we will show in Section~\ref{sec:alg}, each column of the membership matrix $\mM$ can be precisely recovered under our algorithm. This property benefits the interpretation of dTBM in practice.

\section{Statistical-computational gaps for higher-order tensors}\label{sec:limits}

In this section, we study the statistical and computational limits of dTBM. We reserve the term higher-order tensors for tensors of order $K\geq 3$. We propose signal-to-noise ratio~(SNR),
\begin{align}\label{eq:gamma}
  \text{SNR}:= \Delta^2_{\min}/\sigma^2 = p^{\gamma}, 
\end{align}
with varying $\gamma \in \bbR$ that quantifies different regimes of interest. We call $\gamma$ the \emph{signal exponent}. Intuitively, a larger SNR, or equivalently a larger $\gamma$, benefits the clustering in the presence of noise. With quantification~\eqref{eq:gamma}, we consider the following parameter space,
\begin{equation}\label{eq:gammafamily}
    \tP(\gamma) = \tP\cap\{\tS \text{ satisfies SNR condition~\eqref{eq:gamma} with $\gamma$} \}.
\end{equation}
Note that $1$-block dTBM does not belong to the space $\tP(\gamma)$ when $\gamma < 0$ by Assumption~\ref{assmp:min_gap}. Our goal is to characterize the clustering accuracy with respect to $\gamma$ when $r \geq 2$. Let $\hat z$ and $z$ be estimated and true clustering functions in the family~\eqref{eq:family}. Define the misclustering error by
\[
\ell(\hat z, z)={1\over p}\min_{\pi \in \Pi}\sum_{i\in[p]}\ind\{\hat z(i)\neq \pi \circ z(i)\},
\]
where $\pi: [r] \mapsto [r]$ is a permutation of cluster labels, $\circ$ denotes the composition operation, and $\Pi$ denotes the collection of all possible permutations. The infinitum over all permutations accounts for the ambiguity in cluster label permutation.

In Sections~\ref{sec:statlimit}-\ref{sec:complimit}, we provide the lower bounds of $\ell (\hat z, z)$ for general Gaussian dTBMs without symmetric assumptions. For general (asymmetric) dTBMs, we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode. For notational simplicity, we still use $z$ and $\tP(\gamma)$ for this general (asymmetric) model. All lower bounds should be interpreted as the worst-case results across $K$ modes.

%For technical simplicity, we will present our main theory with a focus on Gaussian tensors. Our algorithm and comparison extend to general sub-Gaussian models with little modification.

\subsection{Statistical critical values}\label{sec:statlimit}
Our first main result is to show the minimax lower bound of SNR for exact recovery in dTBM. 
%We call a dTBM general if no symmetric assumption is imposed to the signal tensor. 
\begin{thm}[Statistical lower bound]\label{thm:stats} Consider general Gaussian dTBMs under the parameter space $\tP(\gamma)$ with $K\geq 1$. Assume $r \lesssim p^{1/3}$. If the signal exponent satisfies $\gamma < -(K-1)$, then, every estimator $\hat z_{\text{stat}}$ obeys
\begin{equation}
    \sup_{(z, \tS, \mtheta) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
\end{equation}
\end{thm}
Theorem~\ref{thm:stats} demonstrates the impossibility of exact recovery of the assignment when $\gamma < -(K-1)$ in the high-dimensional regime $p\to \infty$ for fixed $r$. The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to, maximum likelihood estimation (MLE)~\citep{wang2019multiway,ke2019community} and trace maximization~\citep{ghoshdastidar2017uniform}. Our derived SNR threshold $-(K-1)$ is also a minimax upper bound, because MLE achieves exact recovery when $\gamma > -(K-1)$. Hence, the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the  critical value for statistical performance of dTBM.

\subsection{Computational critical values}\label{sec:complimit}
An important ingredient to establish the computational limits is the \emph{hypergraphic planted clique (HPC) conjecture} \citep{zhang2018tensor, brennan2020reducibility}. The HPC conjecture indicates the impossibility of fully recovering the planted cliques with polynomial-time algorithm when the clique size is less than the number of vertices in the hypergraph.
%\fixme{Miaoyan}{Is this condition used in our Theorem 3?}. 
The formal statement of HPC detection and conjecture can be found in Definition~\ref{def:HPC} and Conjecture~\ref{hypo:HPC} as following.

\begin{defn}[Hypergraphic planted clique (HPC) detection]~\label{def:HPC} Consider an order-$K$ hypergraph $H = (V,E)$ where $V = [p]$ collects vertices and $E$ collects all the $K$-way edges. Let $\tH_k(p, 1/2)$ denote the Erd\H{o}s-R\'{e}nyi $K$-hypergraph where the edge $(i_1,\ldots, i_K)$ belongs to $E$ with probability $1/2$. Further, we let $\tH_K(p, 1/2, \kappa)$ denote the hyhpergraph with planted cliques of size $\kappa$. Specifically, we generate a hypergraph from $\tH_k(p, 1/2)$, pick $\kappa$ vertices uniformly from $[p]$, denoted $K$, and then connect all the hyperedges with vertices in $K$. Note that the clique size $\kappa$ can be a function of $p$, denoted $\kappa_p$.

The order-$K$ HPC detection aims to identify whether there exists a planted clique hidden in a Erd\H{o}s-R\'{e}nyi $K$-hypergraph. We formulate HPC detection as the following hypothesis testing problem
\begin{equation}
    H_0:\ H \sim \tH_K(p,1/2) \quad \text{versus} \quad H_1: \ H \sim \tH_K(p,1/2, \kappa_p).
\end{equation}
\end{defn}

\begin{conjecture}[Hypergraphic planted clique (HPC) conjecture]\label{hypo:HPC} Consider the HPC detection problem in Definition~\ref{def:HPC}. Suppose the sequence $\{\kappa_p\}$ such that $\limsup_{p \rightarrow \infty} \log \kappa_p/ \log \sqrt{p} \leq (1 - \tau)$. Then, for any sequence of polynomial-times test $\{ \varphi_p\}: H \mapsto \{0,1\}$ we have 
\begin{equation}
    \liminf_{p \rightarrow \infty} \bbP_{H_0} \of{ \varphi_p(H) =1 } +  \bbP_{H_1} \of{ \varphi_p(H) =0} \geq \frac{1}{2}.
\end{equation}
\end{conjecture}

Under the HPC conjecture, we establish the SNR lower bound that is necessary for any \emph{polynomial-time} estimator to achieve exact clustering.

\begin{thm}[Computational lower bound]\label{thm:comp} Consider general Gaussian dTBMs under the parameter space $\tP(\gamma)$ with $K\geq 2$. Assume HPC conjecture holds. If the signal exponent $\gamma < -K/2$, then, every \emph{polynomial-time estimator} $\hat z_{\text{comp}}$ obeys
\begin{align}
   \liminf_{p\to \infty}\sup_{(z, \tS, \mtheta) \in \tP(\gamma)}  \bbE \left[ p\ell(\hat z_{\text{comp}}, z) \right]\geq 1.
\end{align}
\end{thm}
Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$.  Therefore, $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM.  In Section~\ref{sec:alg}, we will show the condition $\gamma> -K/2$ suffices for our proposed polynomial-time estimator. Thus, $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM.

\begin{rmk}[Statistical-computational gaps]
Now, we have established the phase transition of exact clustering under order-$K$ dTBM combing Theorems~\ref{thm:stats} and \ref{thm:comp}. Figure~\ref{fig:phase_axis} summarizes our results of critical SNRs when $K \geq 2$. Particularly, dTBM reduces to matrix degree-corrected model when $K =2$, and the statistical and computational bounds show the same critical value. When $K =1$, dTBM reduces to the degree-corrected sub-Gaussian mixture model (GMM) with model
\begin{equation}
    \mY = \mTheta \mM \mS + \mE,
\end{equation}
where $\mY \in \bbR^{p \times d}$ collects $n$ data points in $\bbR^d$, $\mS \in \bbR^{r \times d}$ collects the $d$-dimensional centroids for $r$ clusters, and $\mTheta \in \bbR^{p \times p}, \mM \in \{0,1\}^{p \times r}, \mE \in \bbR^{p \times d}$ have the same meaning as in dTBM. \cite{lu2016statistical} implies polynomial-times algorithms are able to achieve the statistical minimax lower bound in GMM. Therefore, we conclude that the statistical-to-computational gap emerges only for higher-order tensors with $K \geq 3$. The result reveals the intrinsic distinctions among (vector) one-dimensional clustering, (matrix) biclustering, and (tensor) higher-order clustering. 
\end{rmk}

%\fixme{Miaoyan}{What happens when $K=1$??} \textcolor{blue}{In our context, when $K=1$ the core tensor $\tS$ reduces to a vector and thus $\Delta_{\min}^2 = 0$. \fixme{Miaoyan}{Not correct. When $K=1$, our model reduces to Gaussian mixture models $\tY=\tS\mM\mTheta$ with $\tS\in\mathbb{R}^{r\times p}$, unknown membership matrix is $\mM\times\{0,1\}^{p\times n}$, and data $\tY\in\mathbb{R}^{n\times p}$. The goal is to cluster $n$ vectors into $r$ clusters, all in dimension $\bbR^p$. Add a remark. citep earlier work about (lack of) stat-comp gap.} } \textcolor{red}{\cite{lu2016statistical} implies there does not exist a statistical-computational bound for non-degree corrected gaussian mixture model. can replace this reference if find a reference for degree-corrected cmm.}

\begin{rmk}[Comparison with non-degree models]
We compare our results to non-degree tensor models. 
%Our parameter space of interest consists of discrete structure (for $z$) and continuous structure (for $\tS$ and $\mtheta$). 
The allowance of degree heterogeneity $\mtheta$ makes the model more flexible, but it incurs extra statistical and computational complexity. Fortunately, we find that the extra complexity does not render the estimation of $z$ qualitatively harder; see the comparison of our phase transition with non-degree TBM~\citep{han2020exact}. 
%In next section, we will supplement the general theory by providing an efficient algorithm that achieves empirical accuracy in the presence of unknown degree heterogeneity.
\end{rmk}

%\subsection{Comparison with non-degree models}

% We have established the phase transition of exact clustering under dTBM. Figure~\ref{fig:phase_axis} summarizes our results about the critical SNRs. In the weak SNR region $\gamma < -(K-1)$, no statistical estimator succeeds in degree-corrected higher-order clustering. In the strong SNR region $\gamma  > -K/2$, our proposed algorithm precisely recovers the clustering in polynomial time. In the moderate SNR regime, $-(K-1)\leq \gamma \leq -K/2$, the degree-corrected clustering problem is statistically easy but computationally hard. 
% We find that the statistical-to-computational gap emerges only for higher-order tensors with $K \geq 3$. The result reveals the intrinsic distinctions between  (matrix) biclustering and (tensor) higher-order clustering.

% We compare our results to non-degree tensor models. Our parameter space of interest consists of discrete structure (for $z$) and continuous structure (for $\tS$ and $\mtheta$). The allowance of degree heterogeneity $\mtheta$ makes the model more flexible, but it incurs extra statistical and computational complexity. Fortunately, we find that the extra complexity does not render the estimation of $z$ qualitatively harder; see the comparison of our phase transition with non-degree TBM~\citep{han2020exact}. In next section, we will supplement the general theory by providing an efficient algorithm that achieves empirical accuracy in the presence of unknown degree heterogeneity.

\section{Polynomial-time algorithm under mild SNR}\label{sec:alg}
We present a two-stage clustering algorithm. The procedure takes a global-to-local approach. See Figure~\ref{fig:demo} for illustration. The global step finds the basin of attraction with polynomial miclustering error, whereas the local iterations improve the initial clustering to exact recovery. Both steps are critical to obtain a satisfactory algorithm output.

\begin{figure}[t]
\centering
\includegraphics[width=\columnwidth]{alg_demo.pdf}
\caption{Illustration of our global-to-local algorithm.}\label{fig:demo}
\end{figure}

\subsection{Weighted higher-order initialization}

We start with weighted higher-order clustering algorithm as initialization. To gain insights, we use an order-3 symmetric tensor as a working example. Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$. 
By model~\eqref{eq:model_tensor}, for all $i \in [p]$, we have
\begin{equation}\label{eq:kmeans}
    \theta(i)^{-1} \mX_{i:} = \off{\text{Mat}( \tS \times_2 \mTheta \mM \times_3  \mTheta \mM )}_{z(i):}. 
\end{equation}
This implies that, all node $i$ belonging to $a$-th community (i.e., $z(i)=a$) share the same normalized mean vector $\theta(i)^{-1} \mX_{i:}$, and vice versa. Intuitively, one can apply $k$-means clustering to the vectors $\{ \theta(i)^{-1} \mX_{i:} \}_{i\in[p]}$, which leads to main idea of our Sub-algorithm~\hyperref[alg:main]{1}.

Specifically, our initialization consists of denoising step and clustering step. The denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) estimates $\tX$ from $\tY$ by a double projection spectral method. The double projection improves usual matrix spectral methods in order to alleviate the noise tensor. 
%The first projection performs HOSVD~\citep{de2000multilinear} via $\mU_{\text{pre}} = \text{SVD}_{r} \of{ \text{Mat}(\tY) }$, where $\text{SVD}_r(\cdot)$ returns the top-$r$ left singular vectors. The second projection performs HOSVD on the projected $\tY$ onto the multlilinear Kronecker space $\mU_{\text{pre}} \otimes  \mU_{\text{pre}}$; i.e.,
%\begin{equation}\label{eq:two-step_factor}
 %   \hat \mU = \text{SVD}_{r} \of{\text{Mat}\of{ \tY\times_1  \mU_{\text{pre}}^T \times_2  \mU_{\text{pre}}^T }}.
%\end{equation}
%The final denoised tensor $\hat \tX$ is defined by
%\begin{equation}\label{eq:two-step_est}
 %   \hat \tX = \tY \times_1 \hat \mU \hat 
%\mU^T \times_2 \hat \mU \hat \mU^T \times_3 \hat \mU \hat \mU^T. 
%\end{equation}
%Compared with conventional HOSVD, our double projection spectral method is shown to improve the estimation substantially for higher-order tensors~\citep{han2020exact}. 
The clustering step (lines 3-5 in Sub-algorithm~\hyperref[alg:main]{1}) performs the weighted $k$-means clustering. 
%We write $\hat \mX=\text{Mat}(\hat \tX)$, and normalize the rows into $\hat \mX^s_{i:}=\onormSize{}{\hat \mX_{i:}}^{-1}\hat \mX_{i:}$ as a surrogate of $\theta(i)^{-1} \mX_{i:}$. Then, a weighted $k$-means clustering is performed on the normalized rows with weights equal to $\onormSize{}{\hat \mX_{i:}}^2$. 
The choice of weights is to bound the $k$-means objective function by the Frobenius-norm accuracy of $\hat \tX$. 
Unlike existing clustering algorithm~\citep{ke2019community}, we apply the clustering on the unfolded tensor $\hat \mX$ rather than on the factors $\hat \mU$. This strategy relaxes the eigen-gap separation condition~\citep{gao2018community, han2020exact}. 
%We assign degenerate rows with purely zero entries to an arbitrarily random cluster; these nodes are negligible in high-dimensions because of the lower bound on $\onormSize{}{\text{Mat}(\tS)_{a:}}$ in~\eqref{eq:family}. The final result gives the initial clustering assignment $\hat z^{(0)}$. 
Full procedures are provided in Sub-algorithm~\hyperref[alg:main]{1}.

We now establish the misclustering error rate of initialization. We call $\mtheta$ is balanced, if the relative extent of heterogeneity is comparable across clusters in that
\begin{equation}\label{eq:degree}
{\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
\end{equation}
Note that, the assumption~\eqref{eq:degree} does not preclude degree heterogeneity. Indeed, within each of the clusters, the highest degree can be $\theta(i) = \Omega(p)$, whereas the lowest degree can be $\theta(i)=\tO(1)$. 
%instead, we assume the relative extent of heterogeneity is comparable across clusters.

\begin{algorithm}[h!]
\caption*{\bf Algorithm: Multiway spherical clustering for degree-corrected tensor block model }
\vspace{.15cm}
\begin{algorithmic}[1]
\Algphase{Sub-algorithm 1: Weighted higher-order initialization}
\INPUT Observation $\tY \in \bbR^{p\times \cdots \times p}$, number of cluster $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
\State Compute factor matrix $ \mU_{\text{pre}} = \text{SVD}_{r} (\text{Mat}(\tY))$ and the $(K-1)$-mode projection $
\tX_{\text{pre}} = \tY \times_1  \mU_{\text{pre}}^T \times_2 \cdots \times_{K-1} \mU_{\text{pre}}^T.$
\State Compute factor matrix $\hat \mU = \text{SVD}_{r}(\text{Mat}(\tX_{\text{pre}}))$ and denoised tensor
$\hat \tX = \tY \times_1 \hat \mU \hat \mU^T \times_2 \cdots \times_K \hat \mU \hat \mU^T.$
\State {Let $\hat \mX = \text{Mat}(\hat \tX)$ and $I_0=\{i \in [p]: \onormSize{}{\hat \mX_{i:}} = 0\}$. Set $\hat z(i)$ randomly in $[r]$ for $i \in I_0$.}
\State{For all $i\in I_0^c$, compute normalized rows
$\hat \mX_{i:}^s :=\onormSize{}{\hat \mX_{i:}}^{-1} \hat \mX_{i:}.$
}
\State {Solve the clustering $\hat z\colon [p]\to[r]$ and centroids $ (\hat \mx_j)_{j\in[r_k]}$ using weighted $k$-means, such that}
\begin{align}
    &\sum_{i \in I_0^c }  \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{\hat z(i)} }^2 
    \leq \\
    &\eta \min_{\substack{\bar \mx_j, j\in[r], \bar z(i),i\in I_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\bar z(i)}}^2.
\end{align}
\OUTPUT Initial clustering $z^{(0)} \leftarrow \hat z$.

\Algphase{Sub-algorithm 2: Angle-based iteration}
%\caption{Angle-based iteration}\label{alg:refinement}
\INPUT Observation $\tY \in \bbR^{p \times \cdots \times p}$, initialization $z^{(0)} \colon [p]\to[r]$ from Sub-algorithm 1, iteration number $T$.
\For {$t = 0$ to $T-1$}
\State Update the block tensor $\tS^{(t)}$ via
$\tS^{(t)} (i_1,...,i_K)= \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}(i_k) = j_k, k \in [K]\}.$
\State Calculate reduced tensor $\tY^{\text{d}} \in \bbR^{p \times r \times \cdots \times r}$ via
\begin{align}
    &\tY^{\text{d}}(i,a_2,\ldots,a_K) \\
    &= 
\text{Ave}\{\tY(i,i_2,\ldots,i_K): z^{(t)}(i_k) = a_k, k \neq 1 \}.
\end{align}

\State Let $\mY^{\text{d}} = \text{Mat}(\tY^{\text{d}})$ and $J_0 = \{ i\in[p]: \onorm{\mY^{\text{d}}_{i:}} = 0\}$. Set $z^{(t+1)}(i)$ randomly in $[r]$ for $i \in J_0$.

\State Let $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. For all $i \in J_0^c$ update the cluster assignment by
\begin{equation}
    z(i)^{(t+1)} = \argmax_{a \in [r]} \cos \left( \mY^{\text{d}}_{i:},\ \mS^{(t)}_{a:} \right).
\end{equation}

\EndFor

\OUTPUT Estimated clustering $z^{(T)}  \in [r]^{p}$.

\end{algorithmic}
\end{algorithm}

%\begin{figure*}[h]
 %   \centering
 %   \includegraphics[width=1\textwidth]{algorithm.pdf}
 %   \label{alg:main}
    %\caption{\color{red}\fixme{Jiaxin}{Should be $\argmax_{a \in [r] } \cos \of{ \mY_{i:}^d , \mS_{a:}^{(t)}}$. $\sin$ can not handle the obtuse angles; e.g.\ $\sin \pi = \sin 0= 0$, where $\pi$ is not the desired angle.} \fixme{Miaoyan}{Agree. I corrected it.}}
%\end{figure*}

\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general Gaussian dTBM under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}. 
%With probability at least $1 -C \exp(-cp)$, we have
%\begin{equation}
%\min_{\pi \in \Pi}\sum_{i\in[p]}\theta^2_i\ind\{ z^{(0)}(i)\neq \pi \circ z(j)\} \lesssim {Mr^K\sigma^2\over \Delta^2_{\min}}p^{-{(K-2)/2}},
%\end{equation}
%where $M>1$ is the relation constant in weighted angle $k$-means. 
Assume $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$. Let $ z^{(0)}$ denote the output of Sub-algorithm~\hyperref[alg:main]{1}. With probability going to 1, we have
\begin{equation}\label{eq:ini}
   \ell(z^{(0)}, z) \lesssim {r^K p^{-K/2}\over \text{SNR}}. 
\end{equation}
\end{thm}

\begin{rmk}[Comparison to previous results] For fixed SNR, our initialization error rate with $K=2$ agrees with the initialization error rate $\tO(p)$ in matrix models~\citep{gao2018community}. Furthermore, in the special case of non-degree TBMs, we achieve the same initial misclassification error $\tO(p^{-K/2})$ as in non-degree models~\citep{han2020exact}. Theorem~\ref{thm:initial} implies the advantage of our algorithm in achieving both accuracy and model flexibility. 
%The result demonstrates the advantage of our algorithm in achieving both accuracy and model flexibility. 
\end{rmk}

%Assume the number of clusters $r$ is fixed. To ensure the misclassification error $\text{MisClust}(z^{(0)},z)$ decays to 0 as the number of nodes $p \rightarrow \infty$, we need the signal level $\Delta_{\min}^2$ satisfying
%\begin{equation}
%    \Delta_{\min}^2/\sigma^2  \gtrsim p^{-K/2}.
%\end{equation}
%This implies that desired polynomial rate can be achieved with lower signal level as sample size increases. Moreover, to achieve exact clustering, we should ensure $\text{MisClust}(z^{(0)},z) < \frac{1}{p}$ and thus it requires $\Delta_{\min}^2 \gtrsim p^{2 - K/2}$, which is a fairly strong signal condition especially when $K$ is small. However, the following theorem indicates a weaker signal condition is enough to achieve exact clustering with proper refinement.
\begin{rmk}[Failure of conventional tensor HOSVD] If we use conventional HOSVD for tensor denoising; that is, we use $\mU_{\text{pre}}$ in place of $\hat \mU$ in line 2, then the misclustering rate becomes $\tO(p)$ for all $K\geq 2$. This rate is substantially worse than our current rate~\eqref{eq:ini}.
\end{rmk}

\subsection{Angle-based iteration}
%A polynomial-time local iteration refines the initial spectral clustering to achieve an exponential error rate in TBM~\citep{han2020exact}.  
We propose an angle-based local iteration to improve the outputs from Sub-algorithm~\hyperref[alg:main]{1}. To gain the intuition, consider an one-dimensional degree-corrected clustering problem with data vectors $\mx_i = \theta(i) \ms_{z(i)} + \mepsilon_i, i \in [p]$, where $\ms_i$'s are known cluster centroids, $\theta(i)$'s are unknown degrees, and $z\colon [p] \mapsto [r]$ is the clustering assignment of interest. The angle-based $k$-means algorithm estimates the assignment $z$ by minimizing the angle between data vectors and centroids; i.e., 
 \begin{equation}\label{eq:angle_kmeans}
     z(i) = \argmax_{a \in [r]} \cos ( \mx_i,\ \ms_{a} ),\ \text{ for all }i \in [p].
 \end{equation}
The classical Euclidean-distance based clustering~\citep{han2020exact} fails to recover $z$ in the presence of degree heterogeneity, even under noiseless case. In contrast, the proposed angle-based $k$-means achieves accurate recovery without explicit estimation of $\mtheta$. Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as angle-based $k$-means, except that we use estimated centroids $\ms^{(t)}_a$ in place of $\ms_a$ based on estimated assignment in previous iterations. 
 % Algorithm~\ref{alg:refinement} updates estimated core tensor and cluster assignments are updated in each iteration. For core tensor, we consider the following update strategy  
 %\[
%\tilde \tS(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z(i_k)=a_k, k\in[3]\}.
%\]
%Intuitively, $\tS^{(t)}$ becomes closer to the true core $\tS$ as $z^{(t)}$ is more precise.  we first aggregate the slices of $\tY$ and obtain a reduced tensor $\tY^{\text{d}}\in \bbR^{p \times r \times r}$ with given $z^{(t)}$, where
%\[
%\tY^{\text{d}}(i,a_2,a_3)=\text{Ave}\{\tY(i,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k \neq 1\}.
%\]
%The row $\text{Mat}(\tY^{\text{d}})_{i:}$ and $\text{Mat}(\tS^{(t)})_{a:}$ corresponds to the $\mx_i$ and $\ms_a$s in the one-dimensional clustering~\eqref{eq:angle_kmeans}. Then, we obtain the updated assignment as
%\[z(i)^{(t+1)} = \argmin_{a \in [r]} \sin \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} }, \text{ for all } i \in [p], \]
%where $z^{(t+1)}(i)$ is randomly assigned for degenerate rows. 
Full procedures for our angle-based iteration are described in Sub-algorithm~\hyperref[alg:main]{2}.

We then establish the misclustering error rate of iterations. Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\epsilon)=\{\bar z\colon \ell(\bar z, z)\leq \epsilon\}$. We define two cluster-size vectors for $\bar z$,
\begin{align}
    &\mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T,\\
    &\mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
\end{align}

We call the degree is $\varepsilon$-locally linearly stable if and only if 
\begin{equation}
    \sin(\mp(\bar z),\ \mp_{\mtheta}(\bar z))\lesssim \varepsilon \Delta_{\min} ,\quad \text{for all } \bar z\in \tN(z, \varepsilon).
\end{equation}

The local stability holds trivially for $\varepsilon=0$ based on the construction of parameter space~\eqref{eq:family}.

%  In Algorithm~\ref{alg:refinement}, the estimated core tensor and cluster assignments are updated in each iteration. For core tensor, notice that the estimator $\tilde S$ with
%   \[
% \tilde \tS(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z(i_k)=a_k, k\in[3]\}
% \]
% is an unbiased estimator satisfying $\bbE[\tilde \tS_{a_1,...,a_K}] = \tS_{a_1,...,a_K}$ with true assignment $z$. This leads to a reasonable core tensor update strategy 
%  \[
% \tS^{(t)}(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k\in[3]\}.
% \]
% Intuitively, the estimate $\tS^{(t)}$ becomes better as $z^{(t)}$ getting more precise.

% For cluster assignment, we reduce the higher-order problem to the one-dimensional clustering by matrization. Given $z^{(t)}$, we first aggregate the slices of $\tY$ and obtain a reduced tensor $\tY^{\text{d}}\in \bbR^{p \times r \times r}$, where
% \[
% \tY^{\text{d}}(i,a_2,a_3)=\text{Ave}\{\tY(i,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k \neq 1\}.
% \]
% Let $\mY^{\text{d}} = \text{Mat}(\tY^{\text{d}})$ and $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. Then, the row $\mY^{\text{d}}_{i:}$s and $\mS^{(t)}_{a:}$s serve as the data vectors $\mx_i$ and centroids $\ms_a$s in the one-dimensional clustering~\eqref{eq:angle_kmeans}. We obtain the updated assignment as
% \begin{equation}
%     z(i)^{(t+1)} = \argmin_{a \in [r]} \sin \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} }, \text{ for all } i \in [p].
% \end{equation}
% In particular, $z^{(t+1)}(i)$ is randomly assigned for degenerate row with $\onorm{\mY^{\text{d}}_{i:}} = 0$. The refinement outputs the final cluster assignment after $T$ iterations denoted $z^{(T)}$. The full procedure for general order-$K$ tensor is described in Algorithm~\ref{alg:refinement}

\begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the setup as in Theorem~\ref{thm:initial}. Assume the local linear stability of degree holds in all neighborhoods $\tN(z,\epsilon)$ for any $\epsilon \leq \log^{-1}p$. Suppose 
$r=\tO(1)$ and $\text{SNR} \gtrsim  p^{-K/2}\log p$. Let $z^{(t)}$ denote the $t$-th iteration output in Sub-algorithm~\hyperref[alg:main]{2} with initialization $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{1}. With probability going to 1, there exists a contraction parameter $\rho \in (0,1)$ such that 
\begin{align}\label{eq:final}
    \ell(z, \hat z^{(t+1)}) \lesssim &\ \KeepStyleUnderBrace{
   \text{SNR}^{-1}
    \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}}_{\substack{\text{statistical}\\ \text{error}} }+ \KeepStyleUnderBrace{ \rho^t \ell(z, z^{(0)}). }_{\substack{\text{computational}\\ \text{error}}}
\end{align}
\end{thm}
The iteration error is decomposed into two parts: statistical error and computational error. The statistical error is unavoidable with noisy data regardless $t$, whereas the computational error decays in an exponential rate as the number of iterations $t \rightarrow \infty$.

Theorem~\ref{thm:refinement} implies that, with probability going to 1, our estimate $z^{(T)}$ achieves exact recovery within polynomial iterations; more precisely,
\begin{equation}
     z^{(T)} = \pi \circ z, \quad \text{for all }T\gtrsim \log_{1/\rho} p,
\end{equation}
for some permutation $\pi \in \Pi$.

\begin{rmk}[Extension for general dTBMs]
Our two-stage algorithm is able to be extended for the general (asymmetric) dTBMs. Specifically, in the Sub-Algorithm~\hyperref[alg:main]{1}, we make the following changes: (1)
Replace the matrcization $\mat(\tY)$ by $\mat_k(\tY)$; (2) Repeat the Steps 1-5 with mode-specified number of cluster $r_k$; (3) Obtain the collection initialization $ \{z^{(0)}_k\}_{k=1}^K$. In the Sub-Algorithm~\hyperref[alg:main]{2}, we make the following changes: (1) Take the collection $\{z^{(0)}_k\}_{k=1}^K$ as input, and update the block tensor $\tS^{(t)}$ with the collection $\{ z_k^{(t)}\}_{k=1}^K$, $
        \tS^{(t)}(i_1,\ldots, i_K) =  \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}_k(i_k) = j_k, k \in [K]\}$;
(2) Calculate reduced tensor $\tY^{\text{d}}_k$ for each mode via
\begin{align}
    &\tY^{\text{d}}_k(a_1,\ldots,a_{k-1}, i ,a_{k+1},\ldots,a_K) \\
    &= \text{Ave}\{\tY(i_1,\ldots,i_{k-1},i,i_{k+1},\ldots,i_K): z^{(t)}(i_j) = a_j, j \neq k \}; 
\end{align}

(3) Repeat Step 8-10 with $\mat_k(\cdot), \tY_k^{\text{d}}$ for each $k \in [K]$ and obtain the collection $\{z^{(T)}_k\}_{k=1}^K$.

Correspondingly, Theorems~\ref{thm:initial} and \ref{thm:refinement} still hold with $\ell(z^{(0)}_k, z_k)$ and $\ell(z^{(t+1)}_k, z_k)$  for all $k \in [K]$.

\end{rmk}

{
\color{red}

\subsection{Bernoulli Case}

\fixme{Jiaxin}{My thoughts about the Bernoulli case}
\paragraph{Bernoulli Case of Proposition 1 in \cite{han2020exact}} Here I write my thoughts about the Bernoulli Proposition 1 following the notation in \cite{han2020exact}.

According to \cite{han2020exact}, we have 
\begin{align}
    \onorm{\hat \tX - \tX}_F \leq &\onorm{ \tX \times_1 \hat \mU_1  \hat \mU_1^T \times_2 \cdots \times_d  \hat \mU_d  \hat \mU_d^T - \tX}_F \\
    &+ \onorm{\tZ \times_1 \hat \mU_1^T \times_2 \cdots \times_d \mU_d^T}_F, 
\end{align}
where the second term 
\begin{equation}
    \onorm{\tZ \times_1 \hat \mU_1^T \times_2 \cdots \times_d \mU_d^T}_F \lesssim \sqrt{p},
\end{equation}
dose not require the noise entries to be identically distributed following Lemma 9. For the first term, 
\begin{align}
    &\onorm{ \tX \times_1 \hat \mU_1  \hat \mU_1^T \times_2 \cdots \times_d  \hat \mU_d  \hat \mU_d^T - \tX}_F \\
    &\leq  \sum_{k=1}^d \onorm{\hat \mU^T_{k, \perp} \mX_k}_F 
     \lesssim   \sum_{k=1}^d  \onorm{\hat \mU_{k,\perp}^{T}  \bar \mX_k}_F + \onorm{\mX_1 - \mX_1'}_F,
\end{align}
where the term $\onorm{\hat \mU_{k,\perp}^{T}  \bar \mX_k}_F \lesssim \sqrt{p}$ following Lemma 8 and the term $\onorm{\mX_1 - \mX_1'}_F \lesssim p^{d/4} + \sqrt{p}$ following the definition of $\mX_1'$. Particularly, $\mX'_1$ can be considered as the original signal tensor only with "strong" singular values, and the threshold for "strong" singular value is $p^{d/4} \vee \sqrt{p}$.

We choose $p^{d/4} \vee \sqrt{p}$ due to the accuracy of SVD. Let $r$ denote the rank of $\mX'_1$ and $r$ is related to the threshold. If the threshold is lower than $p^{d/4} \vee \sqrt{p}$, then the principle angle between the SVD singular space and the true rank $r$ singular space will inflate, i.e., the inequality between (48) (49) in \cite{han2020exact}
\begin{equation}
    \onorm{ \tilde \mU_{k, \perp}^T \mU_k' } \leq \frac{1}{\sqrt{2}},
\end{equation}
does not hold. Note that the above inequality is where the $p^{K/2}$ term comes from in the Proposition 1, and the inequality follows by Lemma 3. Lemma 3 follows the Lemma 4 in \cite{cai2018rate} which is only applicable for the i.i.d.\ sub-Gaussian noise and is the reason why the Proposition 1 requires i.i.d.

The Lemma 7 in \cite{gao2018community} provides the accuracy of SVD of the binary matrix, i.e., for a binary matrix $\mY \in \{0,1\}^{p_1 \times p_2}$ with $\bbE \mY = \mP$ and  $\hat \mP$ which is the rank $r$ SVD of $\mY$, we have
\begin{equation}
    \onorm{\hat \mP - \mP}_F \leq \sqrt{r (p_1 \vee p_2)}.
\end{equation}
However, Lemma 7 does not provide the result for the principle angle between the estimated and true singular spaces. So, I am not sure where can we apply Lemma 7 to the Bernoulli case of Proposition 1. In addition, both Han's algorithm and our algorithm do not matricize the tensor to a $p^{\ceil{K/2}}$-by-$p^{\floor{K/2}}$ matrix and apply SVD to the matrix. Thus, I am not sure why there will have the $p^{\ceil{K/2}}$ term in the Bernoulli case.

}

{
\color{red}
\subsection{Rank Selection}
\subsection{Computational Complexity}
}

%We now discuss the conditions for which exact recovery occurs, i.e., $\ell(z, \hat z^{(t+1)})=0$ for some $t\geq T=\text{poly}(p)$. Based on Theorem~\ref{thm:refinement}, we need
%\[
%T\geq \log p +\log \ell(z, \hat z^{(0)})
%\]
%given initial assignment $z^{(0)}$ from Algorithm~\ref{alg:initial} and iteration number $T \geq 2 \log p$, with probability going to 1, the output of our Algorithm~\ref{alg:refinement} achieves exact recovery, i.e., for some permutation $\pi \in \Pi$,
%\begin{equation}
 %    z^{(T)} = \pi \circ z, \quad \text{for all }T\geq 2\log p.
%\end{equation}

%\begin{algorithm}[h]

%\caption{Angle-based iteration}\label{alg:refinement}
%\begin{algorithmic}[1]

%\INPUT Observation $\tY \in \bbR^{p \times \cdots \times p}$, initialization $z^{(0)} \colon [p]\to[r]$, iteration number $T$.
%\For {$t = 0$ to $T-1$}
%\State Update the block tensor $\tS^{(t)}$ via
%\begin{align}
%    &\tS^{(t)} (i_1,...,i_K)\\
 %   & \quad = \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}(i_k) = j_k, k \in [K]\}
%\end{align}
%\State Calculate reduced tensor $\tY^{\text{d}} \in \bbR^{p \times r \times \cdots \times r}$
%\begin{align}
%    &\tY^{\text{d}}(i,a_2,\ldots,a_%K) \\
%    &= 
%\text{Ave}\{\tY(i,i_2,\ldots,i_K): z^{(t)}(i_k) = a_k, k \neq 1 \}
%\end{align}

%\State Let $\mY^{\text{d}} = \text{Mat}(\tY^{\text{d}})$ and $J_0 = \{ i\in[p]: \onorm{\mY^{\text{d}}_{i:}} = 0\}$. Set $z^{(t+1)}(i)$ randomly in $[r]$ for $i \in J_0$.

%\State Let $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. For all $i \in J_0^c$ update cluster assignment by
%\begin{equation}
%    z(i)^{(t+1)} = \argmin_{a \in [r]} \sin \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} }.
%\end{equation}

%\EndFor

%\OUTPUT Estimated clustering $z^{(T)}  \in [r]^{p}$.
%\end{algorithmic}
%\end{algorithm}

%We refer to the combination of Algorithms~\ref{alg:initial}+ \ref{alg:refinement} as provable dTBM algorithm.

%\fixme{Miaoyan}{add discussion as in my JMLR binary tensor paper (stat-comp dominating terms)}

\section{Numerical studies}\label{sec:simulation}

 We evaluate the performance of the weighted higher-order initialization and angle-based iteration in this section. We report average errors and standard deviations across 30 replications in each experiment. Clustering accuracy is assessed by clustering error rate (CER, i.e., one minus rand index). Note that CER between $(\hat z, z)$ is equivalent to misclustering error $\ell(\hat z, z)$ up to constant multiplications~\citep{meilua2012local}, and a lower CER indicates a better performance.

We generate order-3 tensors with \emph{assortative}~\citep{gao2018community} core tensors to control SNR; i.e., we set $\tS_{aaa} = s_1$ for $a \in [r]$ and others be $s_2$, where $s_1 > s_2 > 0$. Let $\alpha = s_1/s_2$. We set $\alpha$ close to 1 such that $1-\alpha=o(p)$. In particular, we have $\alpha = 1 + \Omega(p^{\gamma/2})$ %\fixme{Miaoyan}{My calculation shows $\text{SNR}=\Omega((1-\alpha)^2)\Rightarrow \alpha=1+\Omega(p^{\gamma/2})$. Double check.} 
  by Assumption~\ref{assmp:min_gap} and definition~\eqref{eq:gamma}.
  Hence, we easily adjust SNR via varying $\alpha$. 
 %We also set $\onorm{\text{Mat}(\tS)_{a:}} = c$ for some $c>0$ to exclude the effect of signal magnitude. 
 Note that the assortative setting is proposed for simulations, and our algorithm is applicable for general tensors in practice. The cluster assignment $z$ is randomly generated with equal probability across  $r$ clusters for each mode. Without further explanation, we generate degree heterogeneity $\mtheta$ from absolute normal distribution as $\theta(i) = |X_i| + 1 - 1/\sqrt{2\pi}$ with $|X_i| \stackrel{\text{i.i.d.}}\sim N(0,1), i \in [p]$ and normalize $\mtheta$ to satisfy \eqref{eq:family}. We set $\sigma^2 = 1$ for Gaussian data.

%  To control SNR, we generate ``assortative" core tensors for all experiments, as a higher-order analogy of the setting in \citep{gao2018community}; i.e.,
% \begin{align} \tS_{ijk} = 
%     \begin{cases}
%     s_1 & \text{if } i = j = k,\\
%     s_2 & \text{if $i,j,k$ are not identical, }
%     \end{cases}
% \end{align}
% where $s_1 > s_2 > 0$. Let $\alpha = s_1/s_2$ be the ratio between diagonal and off-diagonal elements. We have $\alpha = 1 + \Omega(p^{\gamma})$ by Assumption~\ref{assmp:min_gap} and definition~\eqref{eq:gamma} with fixed $\sigma^2$. This helps us to easily adjust SNR via varying $\alpha$. We also set $\onorm{\text{Mat}(\tS)_{a:}} = c$ for some $c>0$ to exclude the effect of signal magnitude. The ``assortative" setting is proposed for convenience. In practice, our algorithm is applicable for general core tensor.

% The cluster assignment $z$ is randomly generated with given integer $r$. Without specific explanation, we generate degree heterogeneity $\mtheta$ from absolute normal distribution as $\theta(i) = |X_i| + 1 - 1/\sqrt{2\pi}, i \in [p]$ with $|X_i| \sim_{i.i.d.} N(0,1)$ and normalize $\mtheta$ to satisfy \eqref{eq:family}. We generate the noisy tensor under Gaussian model with $\sigma^2 = 1$ and Bernoulli model with mean tensor  $\tS \times_1 \mTheta \mM \times_2 \mTheta \mM \times_3 \mTheta \mM$. We denote provable dTBM algorithm as \textbf{Ours} for simplicity.

\subsection{Verification of theoretical results}

The first experiment verifies  statistical-computational gap described in Section~\ref{sec:limits}. Consider the Gaussian model with $p = \{80, 100\}$, $r = 5$. We vary $\gamma $ in $ [-1.2, -0.4]$ and $[-2.1, -1.4]$ for matrix ($K=2$) and tensor $(K = 3)$ clustering, respectively. Note that finding MLE under dTBM is computationally intractable. We approximate MLE using an oracle estimator; i.e., the output of Sub-algorithm~\hyperref[alg:main]{2} initialized from true assignment. Figure~\ref{fig:phase}a shows that both our algorithm and oracle estimator start to decrease around the critical value $\gamma_{\text{stat}}  = \gamma_{\text{comp}}  = -1$ in matrix case. In contrast, Figure~\ref{fig:phase}b shows a significant gap in the phase transitions between the algorithm estimator and oracle estimator in tensor case. The oracle error rapidly decreases to 0 when $\gamma_{\text{stat}} = -2$, whereas the algorithm estimator tends to achieve exact clustering when $\gamma_{\text{comp}} = -1.5$. Figure~\ref{fig:phase} confirms the existence of the statistical-computational gap in our Theorems~\ref{thm:stats} and~\ref{thm:comp}.

%\fixme{Miaoyan}{\color{red} add spacing between two figure panels}
\begin{figure}[htb]
    \centering
    \includegraphics[width=\columnwidth]{phase_anno3.pdf}
    \caption{SNR phase transitions for clustering in dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.
    }
    \label{fig:phase}
\end{figure}

The second experiment verifies the performance guarantees of two algorithms: (i) weighted higher-order initialization; (ii) combined algorithm of weighted higher-order initialization and angle-based iteration. We consider both the Gaussian and Bernoulli models with $p = \{80, 100\}$, $r = 5$, $\gamma \in [-2.1, -1.4]$. 
%We set $c = 0.5$ and $c = 0.25$ for Gaussian and Bernoulli cases, respectively. 
Figure~\ref{fig:ini_re} shows the substantial improvement of combined algorithm over initialization, especially under weak and intermediate signals. This phenomenon agrees with the error rates in Theorems~\ref{thm:initial} and \ref{thm:refinement} 
and confirms the necessity of the local iterations.

\begin{figure}[htp!]
    \centering
     \includegraphics[width=\columnwidth]{ini_re_anno3.pdf}
    \caption{CER versus signal exponent $(\gamma)$ for initialization only and for combined algorithm. We set $p = \{80, 100\}, r = 5, \gamma \in [-2.1, -1.4]$ under (a) Gaussian models and (b) Bernoulli models. }
    \label{fig:ini_re}

\end{figure}

{
\color{red} 
More experiments:
\begin{enumerate}
    \item Initialization impact to Angle-based iteration.(Figure 4 in Han)
    \item Imbalanced cluster size.(Figure 9 in Han)
    \item Non-degree comparison. 
\end{enumerate}
}

\subsection{Comparison with other methods}\label{subsec:comp}

We compare our algorithm with following higher-order clustering methods:
\begin{itemize}[wide,topsep=-3pt,itemsep=0pt,parsep=1pt]
    \item \textbf{\small HOSVD }: HOSVD on data tensor and $k$-means on the rows of the factor matrix;
    \item \textbf{\small HOSVD+}: HOSVD on data tensor and $k$-means on the $\ell_2$-normalized rows of the factor matrix;
    \item \textbf{\small HLloyd}: High-order Lloyd algorithm and high-order spectral clustering~\citep{han2020exact};
    \item \textbf{\small SCORE}: Tensor-SCORE for clustering~\citep{ke2019community};
    %\item \textbf{\small SCORE-}: Incomplete tensor-SCORE without regularized higher-order orthogonal iteration.
\end{itemize}

 Among the four alternative algorithms, the \textbf{\small SCORE} is the closest method to ours.
%, since \textbf{\small SCORE} also The incomplete version \textbf{\small SCORE-} skips the key step, regularized HOOI, in \textbf{\small SCORE}. We involve \textbf{\small SCORE-} to study the critical components of \textbf{\small SCORE}. 
We set the tuning parameters of \textbf{\small SCORE} as in previous literature \citep{ke2019community}. The methods \textbf{\small SCORE} and \textbf{\small HOSVD+} are designed for degree models, whereas \textbf{\small HOSVD} and \textbf{\small HLloyd} are designed for non-degree models.
%which shares a similar idea of our spectral initialization but implements naive denoising and clustering steps.  We refer our algorithm as \textbf{\small dTBM}.
%, and the later one is expected to outperform than spectral methods under TBM. 
We conduct two experiments to assess the impacts of (i) signal strength and (ii) degree heterogeneity, based on Gaussian and Bernoulli models with $ p = 100, r = 5$. We refer to our algorithm as \textbf{\small dTBM} in the comparison. 
%Consider $ p = 100, r = 5, c = 0.5$ for both experiments.

We investigate the effects of signal to clustering performance by varying $\gamma \in [-1.5, -1.1]$. Figure~\ref{fig:comp_gamma} shows the consistent outperformance of our method \textbf{\small dTBM} among all algorithms. The sub-optimality of \textbf{\small SCORE} and \textbf{\small HOSVD+} indicates the necessity of local iterations on the clustering. Furthermore,  Figure~\ref{fig:comp_gamma} shows the inadequacy of non-degree algorithms in the presence of mild degree heterogeneity. 
The experiment demonstrates the benefits of addressing heterogeneity in higher-order clustering tasks.

\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{comp_gamma_anno3.pdf}
    \caption{CER versus signal exponent (denoted $\gamma$) for different methods. We set $p = 100, r = 5, \gamma \in [-1.5, -1.1]$ under (a) Gaussian and (b) Bernoulli models.}
    \label{fig:comp_gamma}
\end{figure}

The only exception in Figure~\ref{fig:comp_gamma} is the slightly better performance of \textbf{\small HLloyd} over \textbf{\small HOSVD+} under Gaussian model. However, we find the advantage of \textbf{\small HLloyd} disappears with higher degree heterogeneity. We perform extra simulations to verify the impact of degree effects. We use the same setting as in the first experiment in the Section~\ref{subsec:comp}, except that we now generate the degree heterogeneity $\mtheta$ from Pareto distribution with shape parameter $a$ prior to normalization. We consider the Gaussian model under low $(a = 6)$ and high $(a = 2)$ degree heterogeneity. Figure~\ref{fig:comp_gamma_theta} shows that the errors for non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) increases with degree heterogeneity. In addition, the advantage of \textbf{\small HLloyd} over \textbf{\small HOSVD+} disappears with higher degree heterogeneity.

%Figure~\ref{fig:comp_gamma} implies all methods perform better with larger signal. Our \textbf{\small dTBM} and \textbf{\small SCORE} show the best accuracy as expected. The sub-optimality of \textbf{\small SCORE-} and \textbf{\small HOSVD ++} implies the necessity of local refinement for spectral clustering. A surprising observation is that \textbf{\small HLloyd} beats  \textbf{\small SCORE-} and \textbf{\small HOSVD ++} under Gaussian case, which indicates the robustness of \textbf{\small HLloyd} for mild heterogeneity.

\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{comp_theta_anno3.pdf}
    \caption{CER versus shape parameter in degree (denoted $a\in[3,6]$) for different methods. We set $p = 100, r = 5, \gamma = -1.2$ under (a) Gaussian and (b) Bernoulli models.}
    \label{fig:comp_theta}
\end{figure}

\begin{figure}[htp!]
    \centering
    \includegraphics[width=\columnwidth]{comp_gamma_theta3.pdf}
    \caption{CER comparison versus signal exponent (denoted $\gamma$) under (a) low (shape parameter $a = 6$)  (b) high (shape parameter $a = 2$) degree heterogeneity. We set $p = 100, r = 5, \gamma \in [-1.5, -1.1]$ under Gaussian model.}
    \label{fig:comp_gamma_theta}
\end{figure}

The last experiment investigates the effects of degree heterogeneity to clustering performance. We fix the signal exponent $\gamma = -1.2$ and vary the extent of degree heterogeneity. In this experiment, we generate $\mtheta$ from Pareto distribution prior to normalization. The density function of Pareto distribution is $f(x|a,b) = a b^a x^{-(a+1)} \ind\{ x \geq b \}$, where $a$ is called \emph{shape} parameter. We vary $a \in [3,6]$ and choose $b$ such that $\bbE[X] = a(a - 1)^{-1}b = 1$ for $X$ following Pareto$(a,b)$. Note that a smaller $a$ leads to a larger variance in $\mtheta$ and hence a larger degree heterogeneity. Figure~\ref{fig:comp_theta} demonstrates the stability of degree-corrected algorithms (\textbf{\small dTBM}, \textbf{\small SCORE}, \textbf{\small HOSVD+}) over the entire range of degree heterogeneity under consideration. In contrast, non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) show poor performance with large heterogeneity, especially in Bernoulli cases. This experiment, again, highlights the benefit of addressing degree heterogeneity in higher-order clustering.

%We fix the signal with $\alpha = 2$ and set $c = 1$ and $c = 0.25$ for Gaussian and Bernoulli cases, respectively.   Figure~\ref{fig:comp_theta} demonstrates the outstanding performance of \textbf{\small dTBM} and \textbf{\small SCORE} under all range of degree heterogeneity. Specifically, \textbf{\small SCORE} slightly beats \textbf{\small dTBM} in Gaussian case. A possible reason for the outperformance is that \textbf{\small SCORE} purposes a relaxer constrain on $\mtheta$. The terrible performances of \textbf{\small HLloyd} and \textbf{\small HOSVD+} confirm the failure of non-degree algorithms  facing strong heterogeneity.\section{Real data analysis}

% \begin{table}[hbt]
%     \centering
%     \begin{tabular}{c |c c c}
%     \hline
%         Method & \textbf{\small Ours} &   \textbf{\small HOSVD} & \textbf{\small HOSVD+}\\
%          CER & \textbf{0.116} & 0.22 &0.213 \\
%          \hline
%          Method & \textbf{\small HLloyd} &\textbf{\small SCORE} & \textbf{\small SCORE-} \\
%          CER &0.149 &  0.199 & 0.228\\
%          \hline
%     \end{tabular}
%     \caption{CER of real data for different methods.}
%     \label{tab:peru}
% \end{table}

\subsection{Human brain connectome data analysis}

The Human Connectome Project (HCP) aims to construct the structural and functional neural connections in human brains~\citep{van2013wu}. We preprocess the original dataset following \cite{desikan2006automated} and partition the brain into 68 regions. The cleaned data includes brain networks for 136 individuals. Each brain network is represented by a 68-by-68 binary symmetric matrix, where the entries with value 1 refers to the presence of connection among 68 nodes while value 0 refers to the absence. We use $\tY \in \{0,1\}^{68 \times 68 \times 136}$ to denote the binary observation. Individual attributes such as gender and sex are recorded.

We apply our generalized Algorithm to the HCP data with the number of clusters on three modes $r_1 = r_2 = 4$ and $r_3 = 3$. The selection of $r_1$ and $r_2$ follows the human brain anatomy and the symmetry in the brain network, and the $r_3$ is chosen to be small following previous analysis~\citep{hu2021generalized}. The estimated brain node clustering results on the first and second mode are the same. Figure~\ref{fig:cluster_brain} indicates that brain connection exhibits a strong spatial separation structure. Specifically, the first cluster, named \emph{L.Hemis}, involves all the nodes in the left hemisphere. The nodes in the right hemisphere are further separated into three clusters led by the middle-part tissues in Temporal and Parietal lobes (\emph{R.Temporal}), the back-part tissues in Occipital lobe (\emph{R.Occipital}), and the front-part tissues in Frontal and Parietal lobes (\emph{R.Supra}). This clustering result is consistent with the common sense that the left and right hemispheres play different roles in human brain.

% \vspace{0.2cm}
% \begin{table}[h]
%  \centering
% %\resizebox{\textwidth}{!}{
% \begin{tabular}{c|c}
% \hline
%          Cluster & \multicolumn{1}{c}{Brain nodes} \\
%          \hline
%         \multirow{1}{*}{L.Hemis} & All brain nodes in left hemisphere\\
%         \hline
%         \multirow{2}{*}{R.Temporal} & SupF Insula SupF SupT SupT MT MT MT\\
%         &IT IT IT SupP IP SupF IsthmusC Precuneus SupT \\
%         \hline
%         \multirow{1}{*}{R.Occiptial} &  LO  LO Fpole Tpole  MOF Cuneus paraHippo Lingual\\
%         \hline
%         \multirow{2}{*}{R.Supra} & preCentral postCentral SupraM SupraM SupraM SupraM \\
%         &CaudMF parsTriangularis parsOpercularis \\
%         \hline
%     \end{tabular}
%     \caption{Brain node clustering results for HCP data.}
%     \label{tab:cluster}
% \end{table}

\begin{figure}[htb]
    \centering
    \includegraphics[width = \columnwidth]{brain_node_cluster.pdf}
    \caption{Illustration of brain node clustering results for HCP data with (a) top and (b) side views. }
    \label{fig:cluster_brain}
\end{figure}

\begin{figure*}[htb]
    \centering
    \includegraphics[width = 1\textwidth]{est_S_anno.pdf}
    \caption{Mode 3 slices of estimated core tensor $\hat \tS$. (a) Average estimated slice weighted by the group size; (b)-(d) Group-specified enrichment, i.e., the subtraction between each slice of $\hat \tS$ and the averaged slice. }
    \label{fig:ests}
\end{figure*}

\begin{figure*}[htb]
    \centering
    \includegraphics[width = 1\textwidth]{brain_connection.pdf}
    \caption{Observed brain connections in the population and each group of individuals.(a) Average brain network; (b)-(d) Group-specified brain networks enrichment in Groups 1-3. Red edges refer to the positive enrichment and blue edges refer the negative reduction.}
    \label{fig:brain_conn}
\end{figure*}

Figure~\ref{fig:ests} illustrates the estimated core tensor $\hat \tS$ with estimated clustering, and Figure~\ref{fig:brain_conn} visualizes the average observed brain connections and the connection enrichment with average observed networks in each group. In general, we find that the inner-hemisphere connection has stronger connection compared to inter-hemisphere connections (Figure~\ref{fig:ests}a). Also, the back and front parts (\emph{R.Occipital}, \emph{R.Supra}) are shown to have more interactions with temporal tissues than inner-cluster connections. In addition, the group 1 with 54\% females implies an enrichment on the inter-hemisphere connections (Figure~\ref{fig:ests}b) while group 4 with only 36\% females exhibits a reduction (Figure~\ref{fig:ests}d). This result agrees with previous findings in \cite{hu2021generalized}. The enrichment on the back-front connection is also recognized in group 3 (Figure~\ref{fig:ests}c). The interpretive pattern in results demonstrate the usefulness of our clustering methods in the human brain connectome data application.

\subsection{Peru Legislation data analysis}

We also apply our method to the legislation networks in the Congress of the Republic of Peru \citep{lee2017time}. Because of the frequent political power shifts in the Peruvian Congress during 2006-2011, we choose to focus on the data for the first half of 2006-2007 year. The dataset records the co-sponsorship of 116 legislators from top 5 parties and 802 bill proposals. We reconstruct legislation network as an order-3 binary tensor $\tY \in \{0,1\}^{116 \times 116 \times 116}$, where $\tY_{ijk} = 1$ if the legislators $(i,j,k)$ have sponsored the same bill, and $\tY_{ijk} = 0$ otherwise. The true party affiliations of legislators are provided and serve as ground truth. We apply various higher-order clustering methods to $\tY$ with $r = 5$. Table~\ref{tab:peru} shows that our \textbf{\small dTBM} achieves the best performance compared to others. The second best method is the two-stage algorithm \textbf{\small HLloyd}, followed by the spectral methods \textbf{\small SCORE} and \textbf{\small HOSVD+}. This result is consistent with our simulations under strong signal and moderate degree heterogeneity. The comparison suggests that our method \textbf{\small dTBM} is more appealing in real-world applications.

\begin{table}[ht]
    \centering
    \begin{tabular}{c |c  c c c}
    \hline
        Method & \textbf{\small dTBM} 
        %&\textbf{\small HOSVD}
        &\textbf{\small HOSVD+} & \textbf{\small HLloyd} &  \textbf{\small SCORE}\\
         CER & \textbf{0.116}
         %&  0.22 
         &0.213 & 0.149 &0.199\\
         \hline
    \end{tabular}
    \caption{Clustering errors (measured by CER) for various methods in the analysis of Peru Legislation dataset.}
    \label{tab:peru}
\end{table}

\section{Conclusion}
We have developed a general degree-corrected tensor block model with a two-step angle-based polynomial-times algorithm. We have, for the first time, characterized the statistical and computational behaviors of the degree-corrected tensor block model under different signal-to-noise ratio regimes. Simulations and Peru Legislation and Human brain connection data analysis confirm the potential of our method for practical applications.

{\color{red}
Future direction. 
\begin{enumerate}
    \item Missing data.
    \item Sub-exponential data.
    \item Imbalance in community size and heterogeneity.
\end{enumerate}
}

{
\color{red}

\section{Proof Sketches}

\begin{enumerate}
\item \textbf{Angle-based techniques} 
\begin{itemize}
    \item Padding of the vectors won't extremely change the angle gap. 
    \item Using geometry property to upper bound $I_{11}, J_{11}$. (have not found a good name for this technique yet)
\end{itemize}
    \item \textbf{Proof idea of Theorem 4}
    \item \textbf{Proof idea of Theorem 5}
\end{enumerate}

}

\bibliographystyle{apalike}
\bibliography{tensor_wang}

%\section*{Appendices}

\newpage
\appendix

We provide the proofs for all the theorems in our main paper. In each sub-section, we show the proof of main theorem and attach the useful lemmas in the end.

\section*{Notation}
Before the proofs, we first introduce the notations used throughout the following sections and the generalized dTBM without symmetric assumptions for the proofs of Theorem~\ref{thm:unique} and Theorem~\ref{thm:stats}. The parameter space and minimal gap assumption for are also extended the generalized dTBM. The conclusions of Theorem~\ref{thm:unique} and \ref{thm:stats} in the main paper are obtained by simply setting $z_k = z, \mS_k = \mS, \mtheta_k = \mtheta, k \in [K]$ for the generalized dTBM.

\paragraph{Notations.}
\begin{enumerate}
    \item For all $ k \in [K]$, denote the tensor matricizations as
    \begin{equation}
        \mY_k = \mat_k \of{ \tY }, \quad \mS_k = \mat_k \of{\tS}, \quad \mE_k = \mat_k \of{ \tE}, \quad \mX_k = \mat_k \of{\tX}.
    \end{equation}
    %We ignore the subscript $k$ when considering symmetric tensors. 
    \item For a vector $\ma$, let $\ma^{s} \coloneqq \ma/\onorm{\ma}$ denote the normalized vector. We make the convention that $\ma^s = 0$ if $\ma = 0$. 
    % Then we rewrite the angle gap~\eqref{eq:general_minial_gap} as
    % \begin{equation}
    %     \Delta_{\min} = \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onorm{ \mS_{k,a:}^s - \mS_{k, b:}^s }.
    % \end{equation}
    \item For a matrix $\mA \in \bbR^{n \times m} \in \bbR^{n^K \times m^K}$, let $\mA^{\otimes K}$ denotes the kronecker product of $K$ matrices $\mA \otimes \cdots \otimes \mA$,
    \item For two terms $a$ and $b$, let $a \asymp b$ if there exist two positive constants $c, C$ such that $cb \leq a\leq Cb$. 
\end{enumerate}

 \paragraph{Generalized dTBM.}

 The general order-$K$ $(p_1, \ldots, p_K)$-dimensional dTBM model with $r_k$ communities and degree heterogeneity $\mtheta_k = \entry{\theta_k(i)} \in \bbR_+^{p_k}$ is represented by
\begin{equation}\label{eq:general_dtbm}
    \tY = \tX+ \tE,\quad \text{where}\quad \tX=\tS \times_1 \mTheta_1 \mM_1 \times_2 \cdots \times_K \mTheta_K \mM_K,
\end{equation}
where $\tY \in \bbR^{p_1 \times \cdots \times p_K}$ is the data tensor, $\tX\in \mathbb{R}^{p_1\times \cdots \times p_K}$ is the mean tensor, $\tS \in \bbR^{r_1 \times \cdots \times r_K}$ is the core tensor, $\tE \in \bbR^{p_1 \times \cdots \times p_K}$ is the noise tensor consisting of independent mean-zero sub-Gaussian entries with variance bounded by $\sigma^2$, $\mTheta_k = \text{diag}(\mtheta_k)$, and $\mM_k\in \{0,1\}^{p_k \times r_k}$ is the membership matrix corresponding to the assignment $z_k: [p_k] \mapsto [r_k]$, for all $k \in [K]$.

For ease of notation, we use $\{z_k\}$ to denote the collection $\{z_k\}_{k=1}^K$, and $\{\mtheta_k\}$ to denote the collection $\{\mtheta_k\}_{k=1}^K$. Correspondingly, we consider the parameter space for the triplet $\of{\{z_k\}, \tS, \{\mtheta_k\}}$,

% \begin{equation}\label{eq:general_family}
%     \tP (r)= \offf{ \of{\{z_k\}, \tS, \{\mtheta_k\}}: \mtheta_k \in\mathbb{R}^p_{+}, {c_1p\over r}\leq |z_k^{-1}(a)| \leq {c_2 p\over r}, c_3\leq \onorm{\mat_k(\tS)}_{a:}\leq c_4 , \onormSize{}{\mtheta_{z_k^{-1}(a)}}_1=|z_k^{-1}(a)|, a \in [r], k\in[K]}.
% \end{equation}
\begin{align}
    \tP(\{r_k\}) = \Bigg\{ \of{\{z_k\}, \tS, \{\mtheta_k\}}:\  &\mtheta_k \in\mathbb{R}^p_{+}, {c_1 p_k\over r_k} |z_k^{-1}(a)| \leq {c_2 p_k\over r_k}, c_3 \leq  \onorm{\mS_{k,a:}} \leq c_4 ,\\
    &\onormSize{}{\mtheta_{k,z_k^{-1}(a)}}_1=|z_k^{-1}(a)|, \text{ for all } a \in [r_k], k\in[K]\Bigg\}. \label{eq:general_family}
\end{align}
% \begin{equation}
%     \tP = \offf{ \of{\{z_k\}, \tS, \{\mtheta_k\}}: \mtheta_k \in\mathbb{R}^p_{+}, |z_k^{-1}(a)| \asymp {p_k\over r_k},  \onorm{\mS_{k,a:}} \asymp 1 , \onormSize{}{\mtheta_{k,z_k^{-1}(a)}}_1=|z_k^{-1}(a)|, a \in [r], k\in[K]}.
% \end{equation}

We call the collection of degree heterogeneity $\{\mtheta_k\}$ is balanced if for all $k \in [K]$,
\begin{equation}\label{eq:general_balanced}
    {\min_{a\in[r]} \onormSize{}{\mtheta_{k, z_k^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{k,z_k^{-1}(a)}}}.
\end{equation}

We also consider the generalized Assumption~\ref{assmp:min_gap} on angle gap.
\begin{assumption}[Generalized angle gap]\label{assmp:general_minimal_gap} We assume the minimal gap between normalized rows of $\mS_k$ is bounded away from zero, i.e.,
\begin{equation}
     \Delta_{\min} \coloneqq \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onorm{ \mS_{k,a:}^s - \mS_{k, b:}^s } > 0
\end{equation}
for all $k \in [K]$. Similarly, let $\text{SNR} = \Delta_{\min}^2/\sigma^2$ with generalized minimal gap $\Delta_{\min}^2$ in Assumption~\ref{assmp:general_minimal_gap}. We define the regime
\begin{equation}
    \tP(\gamma) = \tP (\{r_k\}) \cap\{\tS \text{ satisfies $\text{SNR} = p^{\gamma}$ and $p_k \asymp p, k \in [K]$} \}.
\end{equation}
\end{assumption}
% by extending the definition~\eqref{eq:minimal_gap} to asymmetric setting, 
% \begin{equation}\label{eq:general_minial_gap}
%     \Delta_{\min} = \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onorm{ \frac{ \of{\mat_k(\tS)}_{a:} }{ \onorm{\of{\mat_k(\tS)}_{a:}}} - \frac{ \of{\mat_k(\tS)}_{b:} }{ \onorm{\of{\mat_k(\tS)}_{b:} }} } > 0.
% \end{equation}

\section*{Proof of Theorem~\ref{thm:unique}}

\begin{proof}[Proof of Theorem~\ref{thm:unique}] 
%Recall that we prove Theorem~\ref{thm:unique} under the generalized dTBM~\eqref{eq:general_dtbm}. The identifiability for the symmetric dTBM~\eqref{eq:model_tensor} is indicated by simply setting $z_k = z, \mS_k = \mS, \mtheta_k = \mtheta, k \in [K]$.

To study the identifiability, we consider the noiseless model with $\tE = 0$. Assume there exists two parameterizations satisfying
\begin{equation}\label{eq:another}
    \tX=\tS\times_1\Theta_1 \mM_1 \times_2 \cdots \times_K \Theta_K \mM'_K=\tS'\times_1\Theta'_1 \mM'_1 \times_2 \cdots \times_K \Theta'_K \mM'_K
\end{equation}
with $\of{ \{z_k\}, \tS, \{ \mtheta_k \} } \in \tP(\{r_k\})$ and $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} } \in \tP(\{r_k'\})$. We prove the sufficient and necessary conditions separately.

\begin{enumerate}
    \item[$(\Leftarrow)$] For the necessity, it is equivalent to show that there exists a triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ is not identical to $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation if the model~\eqref{eq:general_dtbm} violates Assumption~\ref{assmp:general_minimal_gap}. Without loss of generality, we assume $\onorm{ \mS_{1,1:}^s - \mS_{1,2:}^s } = 0$.

 If $\mS_{1,1:}$ is a zero vector, consider $\mtheta_1'$ such that $\mtheta'_{z_1^{-1}(1)} \neq \mtheta_{z_1^{-1}(1)}$. Let $\{z'_k\} = \{z_k\}$, $\tS' = \tS$, and $\mtheta'_k = \mtheta_k$ for all $k = 2, \ldots, K$. Then the triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ is not identical to $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation. We have a similar story when $\mS_{1,2:}$ is a zero vector.

If neither $\mS_{1,1:}$ nor $\mS_{1,2:}$ is a zero vector, there exists a positive constant $c$ such that $\mS_{1,1:} = c \mS_{1,2:}$. Thus, there exists a core tensor $\tS_0 \in \bbR^{r_1 -1 \times \cdots \times r_K}$ such that 
\begin{equation}
    \tS = \tS_0 \times_1 \mC \mR, \quad \text{where} \quad \mC = \text{diag}(1, c, 1,...,1) \in \bbR^{r_1 \times r_1}, \quad \mR = \begin{pmatrix}
    1& 0\\
    1&0 \\
    0 & \mone_{r_1-2}
    \end{pmatrix} \in \bbR^{r_1 \times (r_1 -1)}.
\end{equation}
Let $\mD = \text{diag}(1+c, 1,...,1) \in \bbR^{r_1 -1 \times r_1 -1}$. Consider the parameterization
\begin{equation}
    \mM'_1 = \mM_1 \mR, \quad \tS' = \tS_0 \times_1 \mD, \quad \theta'_{1}(i) = \begin{cases}
    \frac{1}{1+c} \theta_{1}(i) & i \in z_{1}^{-1}(1)\\
     \frac{c}{1+c} \theta_{1}(i) & i \in z_{1}^{-1}(2)\\
     \theta_{1}(i) & \text{ otherwise }
    \end{cases},
\end{equation}
and $\mM'_k = \mM_k, \mtheta'_k = \mtheta_k$ for all $k = 2, \ldots, K$. Then we have constructed a
triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ that is not identical to $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation.

\item[$(\Rightarrow)$] For the sufficiency, it is equivalent to show that all possible triplets $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ are identical to $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation if the model~\eqref{eq:general_dtbm} satisfies Assumption~\eqref{assmp:general_minimal_gap}. We show the uniqueness of the parameters separately.

%Suppose $(\tS', \mM', \Theta') $ satisfies assumption~\ref{eq:gap}. We want to show the decomposition~\eqref{eq:degree} is unique up to permutation.

First, we show the uniqueness of $\mM_k$ for all $k \in [K]$. Specifically, we show the uniqueness of the first mode membership matrix, i.e., $\mM'_1 = \mM_1\mP_1$ where $\mP_1$ is a permutation matrix. The uniqueness of $\mM_k$ for $k = 2,\ldots, K$ can be showed in the same way, and we omit the repeated procedures.

On one hand, consider the pair of nodes $(i,j)$ such that $z_1(i) = z_1(j)$. We have $\onormSize{}{\mX_{1, z_1(i):}^s - \mX_{1, z_1(j):}^s } = 0$ and thus $\onormSize{}{ (\mS')_{1, z'_1(i):}^{s} - (\mS')_{1, z'_1(j):}^{s} } = 0$ by Lemma~\ref{lem:angle}. Then, by Assumption~\eqref{assmp:general_minimal_gap}, we have $z'_1(i) = z'_1(j)$. On the other hand, consider the pair of nodes $(i,j)$ such that $z_1(i) \neq z_1(j)$. We have $ \onorm{\mX_{1,i:}^s - \mX_{1,j:}^s} \neq 0$ and thus $\onorm{ (\mS')_{1, z'_1(i):}^{s} - (\mS')_{1, z'_1(j):}^{s} } \neq 0$ by Lemma~\ref{lem:angle}. Hence, we have $z'_1(i) \neq z'_1(j)$. Therefore, we have proven that $z'_1$ is equal to $z_i$ up to label permutation.

Next, we show the uniqueness of $\mtheta_k$ for all $k \in [K]$ given that $z_k = z_k'$. Similarly, we show the detailed proof of the uniqueness of $\mtheta_1$, i.e., $\mtheta'_1 = \mtheta_1$, and omit the repeated procedures for $\mtheta_k, k = 2,\ldots, K$.

Consider an arbitrary $j \in [p_1]$ such that $z_1(j) = a$. Then for all the node $i \in  z_1^{-1}(a)$ in the same cluster of $j$, we have 
\begin{equation}
    \frac{\mX_{1,z_1(i):}}{\mX_{1,z_1(j):}} = \frac{\mX'_{1,z_1(i):}}{\mX'_{1,z_1(j):}}, \text{ which implies }  \frac{\theta_1(j)}{\theta_1(i)} = \frac{\theta'_1(j)}{\theta'_1(i)}.\label{eq:theta_uniq}
\end{equation}
Let $\theta'_1(j) = c\theta_1(j)$ for some constant $c$. By the equation~\eqref{eq:theta_uniq}, we have $\theta'_1(i) = c \theta_1(i)$ for all $ i \in  z_1^{-1}(a)$. Note that $(\{z_k\}, \tS', \{\mtheta'_k\}) \in \tP(\{r_k\})$. We have 
\begin{equation}
    \sum_{j \in z^{-1}(a)} \theta'_1(j) = c \sum_{j \in z^{-1}(a)} \theta_1(j) = 1,
\end{equation}
which implies $c = 1$. Hence, we have proven $\mtheta_1 = \mtheta'_1$ given that $z_1 = z'_1$.

Last, we show the uniqueness of $\tS$, i.e., $\tS'=\tS\times_1 \mP^{-1}_1\times_2\cdots \times_K \mP^{-1}_K$, where $\mP_k, k \in [K]$ are permutation matrices.  Given $z'_k = z_k, \mtheta'_k = \mtheta_k$, we have $\mM'_k = \mM_k \mP_k$ and $\mTheta'_k = \mTheta_k$ for all $k \in [K]$.

Let $\mD_k = \off{ (\mTheta'_k \mM'_k)^T (\mTheta'_k \mM'_k) }^{-1} (\mTheta'_k \mM'_k)^T, k \in [K]$. By the parameterization, we have 
\begin{align}
    \tS' &= \tX \times_1 \mD_1 \times_2 \cdots \times_K \mD_K \\
    &= \tS \times_1 \mD_1 \mTheta_1 \mM_1 \times_1 \cdots \times_K \mD_K \mTheta_K \mM_K \\
    &= \tS \times_1 \mP^{-1}_1 \times_2 \cdots \times_K \mP^{-1}_K.
\end{align}

\end{enumerate}

Therefore, we finish the proof of Theorem~\ref{thm:unique}.
\end{proof}

\paragraph{Useful Lemma for the Proof of Theorem~\ref{thm:unique}}

\begin{lem}[Motivation of angle-based clustering]\label{lem:angle} Consider the signal tensor $\tX$ in the generalized dTBM~\eqref{eq:general_dtbm} with $(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\{r_k\})$ and $r_k \geq 2$. Then, for any $k \in [K]$ and index pair $(i,j)\in[p_k]^2$, we have 
\begin{equation}
    \onorm{ \mS_{k,z_k(i):}^s -  \mS_{k,z_k(j):}^s } = 0 \quad \text{if and only if} \quad \onorm{  \mX_{k, z_k(i):}^s -  \mX_{k,z_k(j):}^s } = 0.
\end{equation}
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:angle}] For simplicity, we show the detailed proof for $k = 1$ and drop the subscript $k$ in $\mX_k, \mS_k$.The repeated proofs for $k = 2, \ldots, K$ are omitted.

By tensor matricization, we have
\begin{equation}
    \mX_{j:} = \theta_1(j) \mS_{z_1(j):} \off{\mTheta_2 \mM_2 \otimes \cdots \otimes \mTheta_K \mM_K}^T.
\end{equation}     
Let $\tilde \mM = \mTheta_2 \mM_2 \otimes \cdots \otimes \mTheta_K \mM_K$. Notice that for two vectors $\ma, \mb$ and two positive constants $c_1, c_2 >0$, we have
\begin{equation}
\onorm{\ma^s - \mb^s} = \onorm{(c_1 \ma)^s - (c_2\mb)^s}.
\end{equation}
Thus it is sufficient to show the following statement that for any index pair $(i,j)\in[p_1]^2$,
\begin{equation}
\onorm{ \mS_{z_1(i):}^s - \mS_{z_1(j):}^s} = 0 \quad \text{if and only if} \quad \onorm{ \off{\mS_{z_1(i):} \tilde \mM^T }^s - \off{\mS_{z_1(j):}\tilde \mM^T}^s} = 0.
\end{equation}
\begin{enumerate}
    \item[$(\Leftarrow)$] Suppose $\onorm{ \off{\mS_{z_1(i):} \tilde \mM^T }^s - \off{\mS_{z_1(j):}\tilde \mM^T}^s} = 0$. There exists a positive constant $c$ such that $\mS_{z_1(i):} \tilde \mM^T= c \mS_{z_1(j):} \tilde \mM^T$. Note that
\begin{equation}
    \mS_{z_1(i):} = \mS_{z_1(i):} \tilde \mM^T \off{ \tilde \mM \of{ \tilde \mM^T  \tilde \mM}^{-1}},
\end{equation}
where $ \tilde \mM^T  \tilde \mM$ is an invertiable diagonal matrix with positive diagonal elements. Thus, we have $ \mS_{z_1(i):} = c  \mS_{z_1(j):}$, which implies $ \onorm{  \mS_{z_1(i):}^s -  \mS_{z_1(j):}^s } = 0 $.

\item[$(\Rightarrow)$] Suppose $ \onorm{ \mS_{z_1(i):}^s - \mS_{z_1(j):}^s } = 0 $. There exists a positive constant $c$ such that $\mS_{z_1(i):} = c \mS_{z_1(j):}$, and thus $\mS_{z_1(i):} \tilde \mM^T = c \mS_{z_1(j):} \tilde \mM^T$, which implies $\onorm{\left[\mS_{z_1(i):} \tilde \mM^T\right]^s- \left[\mS_{z_1(j):} \tilde \mM^T\right]^s}=0$.
\end{enumerate}
Therefore, we finish the proof of Lemma~\ref{lem:angle}.
% \begin{equation}
%     \tilde \mM^T \tilde \mM = \off{ \mM^T \Theta^2 \mM }^{\otimes K-1}, \quad \text{and} \quad \mM^T \Theta^2 \mM  = \text{diag} \of{ \sum_{j \in z^{-1}(1)} \theta(j)^2, ...,  \sum_{j \in z^{-1}(r)} \theta(j)^2}.
% \end{equation}
\end{proof}

\section*{Proof of Theorem~\ref{thm:stats}}

\begin{proof}[Proof of Theorem~\ref{thm:stats}]
We will prove a more general conclusion than the main paper by allowing growing $r_k$'s. Consider the generalized dTBM in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$. Specifically, we will show that, under 
the assumptions $K\geq 1$, $r\lesssim p^{1/3}$ and SNR condition
\[
{\Delta^2_{\min}\over \sigma^2}\lesssim {r^{K-1}\over p^{K-1}},\quad \text{or equivalently}, \quad \gamma\leq -(K-1)(1+\log_p r),
\]
 the desired conclusion in Theorem~\ref{thm:stats} holds; i.e, for all $k \in [K]$, every estimator $\hat z_{k,\text{stat}}$ obeys
\begin{equation}\label{eq:max}
    \sup_{(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{k,\text{stat}}, z_k) \right]\geq 1.
\end{equation}

 Noticed that inequality~\eqref{eq:max} is a minimax lower bound, it suffices to show the inequality holds for a particular $(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)$. Specifically, we consider the estimation problem based on a particular parameter point $(\{z_k\}, \tS, \{\mtheta_k\})$ with the following three properties:
\begin{equation}\label{eq:construction}
    \text{(i)}\  \theta_k(i)=1 \text{ for all }i\in[p]; \  \text{(ii)}\ % \onorm{{\mS_1\over \onormSize{}{\mS_1}}-{\mS_2\over \onormSize{}{\mS_2}}}=
    \Delta_{\min}\lesssim  \left({p\over r}\right)^{-\frac{K-1}{2}}\sigma; \  \text{(iii)}\ |z^{-1}_k(a)|={p \over r} \in \mathbb{Z}_+ \text{ for all }a\in[r],
\end{equation}
for all $k \in [K]$.
% \begin{enumerate}[label=(\roman*)]\label{eq:construction}
% \item $\theta(i)=1 \text{ for all }i\in[p]$.
% \item $\onorm{{\mS_1\over \onormSize{}{\mS_1}}-{\mS_2\over \onormSize{}{\mS_2}}}=\Delta_{\min}\lesssim \left({p\over r}\right)^{-(K-1)/2}\sigma$.
% \item $|z(a)|={p \over r} \text{ for all }a\in[r]$.
% \end{enumerate}
Furthermore, we define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation. Based on \citet[Proof of Theorem 6]{han2020exact}, we consider the minimax rate over the restricted family of $\hat z_k$'s for which the following three conditions are satisfied:
\footnotesize
\begin{equation}
    \text{(iv)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \ \text{(v)} \ |T^c_k|\asymp {p\over r}; \ \text{(vi)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
\end{equation}
\normalsize
for all $k \in [K]$.
% \begin{enumerate}[label=(\roman*)]
% \setcounter{enumi}{3}
%     \item $\hat z(i)=z(i)$ for all $i\in T$.
%     \item $|T^c|\asymp {p\over r}$.
%     \item $\min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z(i) \neq \pi\circ z (i)\} = \sum_{i\in[p]}\ind\{\hat z(i) \neq  z (i)\}$.
% \end{enumerate}
The construction of $T$ is precisely the same as \citet[Proof of Theorem 6]{han2020exact}. 
Then, following the proof of~\citet[Theorem 2]{gao2018community}, for all $k \in [K]$, we have
\begin{equation}\label{eq:minimax}
\inf_{\hat z_k}\sup_{z_k}\bbE \ell(\hat z_k, z_k) \gtrsim {1\over r^3 |T^c_k|}\sum_{i\in T^c_k}\inf_{\hat z_k}\left\{\bbP[\hat z_k(i)=2|z_k(i)=1] + \bbP[\hat z_k(i)=1|z_k(i)=2] \right\},
\end{equation}
where $\hat z_k, z_k$ on the left hand side denote the generic clustering functions in $\tP(\gamma)$, $z_k$ on the right hand side denotes a particular parameter satisfying properties (i)-(vi), and the infimum on the right hand side is taken over the restricted family of $\hat z$ satisfying (iv)-(vi). Here, the factor $r^3=r\cdot r^2$ in~\eqref{eq:minimax} comes from two sources: $r^2\asymp {r\choose 2}$ comes from the multiple testing burden for all pairwise comparisons among $r$ clusters, and another $r$ comes from the number of elements $|T^c_k|\asymp {p\over r}$ to be clustered.

Next, we need to find the lower bound of the rightmost side in~\eqref{eq:minimax}. For simplicity, we only show the bound for the mode-1 case $k = 1$. We drop the subscripts $1$ in $z_1, T_1, \mS_1, \mtheta_1$ and omit the repeated procedures for the cases of $k = 2,\ldots, K$.

We consider the hypothesis test based on model~\eqref{eq:general_dtbm}. First, we reparameterize the model under the construction~\eqref{eq:construction}
\begin{equation}
    \mx_a = \off{\mat_1 \of{ \tS\times_2 \mM_2\times_3\cdots \times_K \mM_K  }}_{a:}, \quad \text{for all} \quad a \in [r],
\end{equation}
where $\mx_a$'s are centroids in $\bbR^{p^{K-1}}$. Without loss of generality, we consider the lower bound for the summand in~\eqref{eq:minimax} for $i=1$. The analysis for other $i\in T^c$ are similar. For notational simplicity, we suppressed the subscript $i$ and write $\my, \theta, z$ in place of $\my_1, \theta_1$ and $z(1)$, respectively. The equivalent vector problem for assessing the summand in~\eqref{eq:minimax} is
\begin{equation}\label{eq:z}
\my=\theta\mx_{z}+\me,
\end{equation}
where $\theta\in\bbR_{+}$ and $z\in \{1,2\}$ are unknown parameters, $\mx_1,\mx_2\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.  Then, we consider the hypothesis testing under the model~\eqref{eq:z}:
\begin{equation}\label{eq:test}
H_0\colon z=1, \quad \text{v.s.}\quad H_\alpha\colon z=2.
\end{equation}
Note that the profile log-likelihood with respect to $z$ is
\[
\tL(z,  \theta(z);\my)\propto - \inf_{\theta>0}\onormSize{}{\my- \theta\mx_{z}}^2 \propto  \cos^2(\my, \mx_{z}) \ind\{ \langle \my,\mx_{z}\rangle  >0 \}, 
\]
and the maximum likelihood estimators (MLE) of $\theta$ and $z$ are
\begin{equation}\label{eq:MLE}
    \hat \theta_{\text{MLE}}= \hat \theta(\hat z_{\text{MLE}})={\langle \my,\mx_{\hat z_{\text{MLE}}}\rangle \over \onorm{\mx_{\hat z_{\text{MLE}}}}^2}\vee 0 , \quad 
\hat z_{\text{MLE}}= \argmax_{a\in\{1,2\}} \offf{\cos(\my,\ \mx_a)\vee 0}. 
\end{equation}

Then, the decision rule $\hat z_{\text{MLE}} \in\{1,2\}$ based on profile log-likelihood ratio is defined as
\begin{equation}\label{eq:ratio}
\hat z_{\text{MLE}}=
\begin{cases}
1 & \text{if }\cos(\my,\mx_1)\geq  \cos(\my,\mx_2) \text{ and }\langle \my,\mx_1\rangle > 0, \\
2 & \text{if }\cos(\my,\mx_1)<\cos(\my,\mx_2) \text{ and }\langle \my,\mx_2\rangle > 0,\\
\text{1 or 2 with equal probability} & \text{otherwise.}
\end{cases}
\end{equation}

Neyman-Pearson Lemma implies
\begin{align}\label{eq:lower}
&\inf_{\hat z}\left\{\bbP[\hat z=2|z=1] + \bbP[\hat z=1|z=2]\right\}= \bbP[\hat z_{\text{MLE}}=1|z=2]+\bbP[\hat z_{\text{MLE}}=2|z=1].
\end{align}
By symmetric, it suffices to bound $\bbP[\hat z_{\text{MLE}}=1|z=2]$. Using~\eqref{eq:ratio}, we obtain
%\fixme{Jiaxin}{$\theta^* = \hat \theta_{\text{MLE}}$?}
\begin{align}\label{eq:bound}
\bbP[\hat z_{\text{MLE}}=1|z=2]&= \bbP\left[ \cos(\theta\mx_2+\me,\mx_1)\geq  \cos(\theta\mx_2+\me, \mx_2) \text{ and } \langle \theta\mx_2+\me,\mx_1\rangle > 0  \right] \notag \\
&\stackrel{(*)}{\geq} \bbP \left[\left\langle \me, {\mx^s_1 -\mx^s_2\over\onorm{\mx^s_1- \mx^s_2}}  \right\rangle  \geq {\theta \over 2}\onormSize{}{\mx_2} \onormSize{}{\mx^s_1-\mx^s_2}\rangle\right]-\notag \\
&\hspace{1cm} \bbP\left[\langle \me, \mx^s_1\rangle \leq -{\theta \over 2}\onorm{\mx_2}\left(2-\onormSize{}{\mx^s_1-\mx^s_2}^2\right)\right]\notag \\
&\stackrel{(**)}{=}\Phi\left({\theta \over 2}\onorm{\mx_2}(2-\onormSize{}{\mx^s_1-\mx^s_2}^2) \right)-\Phi\left({\theta\over 2}\onorm{\mx_2}\onormSize{}{\mx^s_1-\mx^s_2} \right),
\end{align}
where $\Phi(\cdot)$ denotes the CDF for standard normal distribution. Here step $(*)$ is based on the inequality $\bbP(A\cup B)\geq \bbP(A)-\bbP(B^c)$ and the identity $1-\langle \mx^s_1,\mx^s_2 \rangle={1\over 2}\onorm{\mx^s_1-\mx^s_2}^2$; and step $(**)$ is based on isotropic property of i.i.d.\ Gaussian distribution
\[
\left\langle\me, {\mx^s_1-\mx^s_2\over \onorm{\mx^s_1-\mx^s_2}} \right\rangle \sim N(0, \sigma^2),\quad \langle \me, \mx^s_1 \rangle \sim N(0,\sigma^2).
\]
By construction~\eqref{eq:construction} of $(\{z_k\}, \tS, \{\mtheta_k\})$ with three properties and lower bound $\min_{a\in[r]}\onorm{\mS_{a:}}\geq c_3$ in the definition of $\tP(\gamma)$, we have $\theta^*=1$, $\onorm{\mx_2}\geq \onorm{\mS_{2:}}\min_{a\in[r]}\onorm{\mtheta_{\mz^{-1}(a)}}^{K-1} \gtrsim \left(p \over r\right)^{(K-1)/2}$. Also, note that under the construction~\eqref{eq:construction}
\begin{equation}
    \cos(\mx_1, \mx_2) = \frac{\and{\mx_1, \mx_2}}{\onorm{\mx_1} \onorm{\mx_2} } = \frac{(p/r)^{K-1} \ang{\mS_{1:}, \mS_{2:}}}{ \sqrt{(p/r)^{K-1} \onorm{\mS_{1:}}^2} \sqrt{(p/r)^{K-1} \onorm{\mS_{2:}}^2} } = \cos(\mS_{1:}, \mS_{2:}),
\end{equation}
which implies $\onormSize{}{\mx^s_1-\mx^s_2} = \onormSize{}{\mS_{1:}^s-\mS_{2:}^s} = \Delta_{\min} \leq 1$. Therefore, the equation~\eqref{eq:bound} is lower bounded by
\begin{equation}\label{eq:normal}
\bbP[\hat z_{\text{MLE}}=1|z=2] \geq \bbP\left[ \left({p\over r}\right)^{(K-1)/2}\Delta_{\min}\lesssim N(0,1)\lesssim  \left({p\over r}\right)^{(K-1)/2} \right]\geq C>0,
\end{equation}
 where the existence of strictly positive constant $C$ is based on the SNR assumption~\eqref{eq:construction}. Combining~\eqref{eq:minimax}, \eqref{eq:lower} and~\eqref{eq:normal} yields
\[
\inf_{\hat z_1}\sup_{(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\gamma)}\bbE\ell(\hat z_1, z_1)\gtrsim C>0, 
\]
and henceforth for all $k \in [K]$
\[
\inf_{\hat z_k}\sup_{(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\gamma)}\bbE \left[p\ell(\hat z_k, z_k)\right]\geq 1.
\]

\end{proof}

\section*{Proof of Theorem~\ref{thm:comp}}

\begin{proof}[Proof of Theorem~\ref{thm:comp}]
The idea of proving computational hardness is to show the computational lower bound for a special class of degree-corrected tensor clustering model with $K\geq 2$. We construct the following special class of higher-order degree-corrected tensor clustering  model. For a given signal level $\gamma\in\bbR$ and noise variance $\sigma$, define a rank-2 symmetric tensor $\tS\in\bbR^{3\times \cdots\times 3}$ subject to
\begin{equation}\label{eq:S}
\tS=\tS(\gamma) = \begin{bmatrix}
		1\\
		1\\
		1
		\end{bmatrix}^{\otimes K}
+\sigma p^{-\gamma/2} \begin{bmatrix}
		1\\
		-1\\
		0
	\end{bmatrix}^{\otimes K}.
\end{equation}
Then, we consider the signal tensor family
\begin{align}
\tP_{\text{shifted}}(\gamma)=\{\tX\colon & \tX=\tS\times_1\mM_1\times_2 \cdots \times_K\mM_K,\  \text{$\mM_k\in\{0,1\}^{p\times 3}$ is a membership matrix that}
\\
& \text{satisfies $|\mM_k(\colon,i)|\asymp p$ for all $i\in[3]$ and $k\in[K]$}\}.
\end{align}
We claim that the constructed family satisfies the following two properties:
\begin{enumerate}[wide,label=(\roman*)]
    \item For every $\gamma\in \mathbb{R}$, $\tP_{\text{shifted}}(\gamma)\subset \tP(\gamma)$, where $\tP(\gamma)$ is the degree-corrected cluster tensor family~\eqref{eq:gammafamily}.
    \item For every $\gamma\in \mathbb{R}$, $\{\tX-1\colon \tX\in \tP_{\text{shifted}}(\gamma)\}\subset \tP_{\text{non-degree}}(\gamma)$, where $\tP_{\text{non-degree}}(\gamma)$ denotes the sub-family of rank-one tensor block model constructed in the proof of \citet[Theorem 7]{han2020exact}. 
\end{enumerate}
The verification of the above two properties is provided in the end of this proof.

Now, following the proof of~\citet[Theorem 7]{han2020exact}, when $\gamma<-K/2$, every polynomial-time algorithm estimator $(\hat \mM_k)_{k\in[K]}$ obeys
\begin{align}\label{eq:p}
\liminf_{p\to \infty} \sup_{\tX\in \tP_{\text{non-degree}}(\gamma)}\mathbb{P}(\exists k \in[K],\  \hat \mM_k \neq \mM_k)\geq 1/2,
\end{align}
under the HPC Conjecture~\ref{hypo:HPC}.
The inequality~\eqref{eq:p} implies
\[
\liminf_{p\to \infty} \sup_{\tX\in \tP_{\text{non-degree}}(\gamma)}\max_{k\in[K]}\mathbb{E}[p\ell(z_k, \hat z_k)]\geq 1.
\]
Based on properties (i)-(ii), we conclude that
\[
\liminf_{p\to \infty} \sup_{\tX\in \tP(\gamma)}\max_{k\in[K]}\mathbb{E}[p\ell(z_k, \hat z_k)]\geq 1.
\]
We complete the proof by verifying the properties (i)-(ii). For (i), we verify that the angle gap for the core tensor $\tS$ in~\eqref{eq:S} is on the order of $\sigma p^{-\gamma/2}$. Specifically, write $\mone=(1,1,1)$ and $\me=(1,-1,0)$. We have
\[
\Mat(\tS)=
\begin{bmatrix}
\text{Vec}(\mone^{\otimes K-1})+\sigma p^{-\gamma/2}\text{Vec}\left(\me^{\otimes (K-1)}\right) \\
\text{Vec}(\mone^{\otimes K-1})-\sigma p^{-\gamma/2}\text{Vec}\left(\me^{\otimes (K-1)}\right) \\
\text{Vec}(\mone^{\otimes K-1})
\end{bmatrix}.
\]
Based on the orthogonality $\langle \mone, \me\rangle=0$, the minimal angle gap among rows of $\Mat(\tS)$ is
\begin{align}
\Delta^2_{\min}(\tS)\asymp \tan^2(\Mat(\tS)_{1:}, \Mat(\tS)_{3:})
=\left(\onorm{\me}_2 \over \onorm{\mone}_2\right)^{2(K-1)} \sigma^2 d^{-\gamma}\asymp \sigma^2 d^{-\gamma}.
\end{align}
Therefore, we have shown that $\tP_{\text{shifited}}(\gamma)=\tP(\gamma)$. Finally, the property (ii) follows directly by comparing the definition of $\tS$ in~\eqref{eq:S} with that in the proof of \citet[Theorem 7]{han2020exact}. 
\end{proof}

%\fixme{Miaoyan}{To Jiaxin: I fixed the above proof for Thm 3. Please revise everything below to make the contexts consistent with our current proofs. }

\paragraph{Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:comp}} 
\begin{defn}[Principal angles] For two matrices $\mU, \hat \mU \in \mathbb{O}^{p \times r}$ %\fixme{Miaoyan}{do we need column organization  $\mathbb{O}(p,r)?$. Yes.}
, we define the principal angles between $\mU$ and $\hat \mU$ by the following diagonal matrix
\begin{equation}
    \text{Angle}(\mU, \hat \mU) = \text{diag}( \arccos(\sigma_1), \cdots, \arccos (\sigma_r) ),
\end{equation}
where $\sigma_1 \geq \cdots \geq \sigma_r>0$ are singular values of $\mU^T \hat \mU$. 
%\fixme{Miaoyan}{Geometry intuition. $\text{Angle}(\mU \mx, \hat \mU \mx) \leq \sigma_r$ for all $\mx$?} 
We define the norm of the principal angles as 
\begin{equation}
     \onorm{\sin(\mU, \hat \mU)}_{\sigma} = \max_{i \in [r]} \sin ( \arccos \of{ \sigma_i} ).
\end{equation}
%\fixme{Miayan}{Properties: $\onormSize{}{\sin(\mU, \hat \mU)}_\sigma= \onormSize{}{\mU^{\perp}\hat \mU}_{\sigma}$. $\max_{\mx\in \mathbb{R}^r}\sin \text{Angle}(\mU \mx, \hat \mU x)\leq \onormSize{}{\mU^{\perp}\hat \mU}_{\sigma}$}

%\[ J_{11}=\text{Angle}(\ma^T \mX \mV, \ma^T\mX\mV^{(t)}) \leq \text{Angle}(\mV^T\mX^T, \mV^{(t),T}\mX^T) \]

% The Schatten-$q$-$\sin$-norm is defined as
% \begin{equation}
%     \onorm{\sin(\mU, \hat \mU)}_q := \of{ \sum_{i=1}^r \sin^q \of{ \arccos \of{ \sigma_i}  } }^{1/q} = \of{ \sum_{i=1}^r \of{ 1 - \sigma_i^2 }^{q/2} }^{1/q},
% \end{equation}
% for $1 \leq q \leq \infty$. We omit the subscript using $\onorm{ \sin(\mU, \hat \mU)}$ when $q = +\infty$.
\end{defn}

% \begin{defn}[Hypergraphic planted clique (HPC) detection]~\label{def:HPC} Consider an order-$K$ hypergraph $H = (V,E)$ where $V = [p]$ collects vertices and $E$ collects all the $K$-way edges. Let $\tH_k(p, 1/2)$ denote the Erd\H{o}s-R\'{e}nyi $K$-hypergraph where the edge $(i_1,\ldots, i_K)$ belongs to $E$ with probability $1/2$. Further, we let $\tH_K(p, 1/2, \kappa)$ denote the hyhpergraph with planted cliques of size $\kappa$. Specifically, we generate a hypergraph from $\tH_k(p, 1/2)$, pick $\kappa$ vertices uniformly from $[p]$, denoted $K$, and then connect all the hyperedges with vertices in $K$. Note that the clique size $\kappa$ can be a function of $p$, denoted $\kappa_p$.

% The order-$K$ HPC detection aims to identify whether there exists a planted clique hidden in a Erd\H{o}s-R\'{e}nyi $K$-hypergraph. We formulate HPC detection as the following hypothesis testing problem
% \begin{equation}
%     H_0:\ H \sim \tH_K(p,1/2) \quad \text{versus} \quad H_1: \ H \sim \tH_K(p,1/2, \kappa_p).
% \end{equation}
% \end{defn}

% \begin{conjecture}[Hypergraphic planted clique (HPC) conjecture]\label{hypo:HPC} Consider the HPC detection problem in Definition~\ref{def:HPC}. Suppose the sequence $\{\kappa_p\}$ such that $\limsup_{p \rightarrow \infty} \log \kappa_p/ \log \sqrt{p} \leq (1 - \tau)$. Then, for any sequence of polynomial-times test $\{ \varphi_p\}: H \mapsto \{0,1\}$ we have 
% \begin{equation}
%     \liminf_{p \rightarrow \infty} \bbP_{H_0} \of{ \varphi_p(H) =1 } +  \bbP_{H_1} \of{ \varphi_p(H) =0} \geq \frac{1}{2}.
% \end{equation}
% \end{conjecture}

We restate the Lemma 3 from \cite{zhang2018tensor} for self-contentedness. 
\begin{lem}[Bounds for principal angle spectral norm.]\label{lem:equivalence_qnorm}
    For any $\mU_1, \mU_2 \in \mathbb{O}_{p,r}$ and all $1 \leq q \leq +\infty$, we have
    \begin{align}
        \frac{1}{4} \onorm{\mU_1 \mU_1^T - \mU_2 \mU_2^T}_{\sigma} \leq \onorm{\sin(\mU_1, \mU_2) }_{\sigma} \leq  \onorm{\mU_1 \mU_1^T - \mU_2 \mU_2^T}_{\sigma}.
    \end{align}
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:equivalence_qnorm}] See \citet[Proof of Lemma 3]{zhang2018tensor}.
\end{proof}

\begin{lem}[Computational lower bound of tensor decomposition. (Theorem 4 in \cite{zhang2018tensor})]\label{lem:hpc} Consider the tucker decomposition model 
\begin{equation}
    \tY = \tS \times_1 \mM_1 \times \cdots \times_K \mM_K + \tE,
\end{equation}
where $\tY \in \bbR^{p\otimes^K }, \tS \in \bbR^{r \otimes^K}$ with $r \geq 1$ is a full rank core tensor, $\mM_k \in \bbR^{p \times r}$ has orthonormal columns for all $k \in [K]$, and $\tE$ has independent entries following $N(0,\sigma^2)$. Let $\tX = \tS \times_1 \mM_1 \times \cdots \times_K \mM_K $. Define the singular signal $\lambda = \min_{k \in [K]}\lambda_{\min} (\Mat_k(\tX))$, where $\lambda_{\min}(\cdot)$ denotes the minimal non-zero singular value. Suppose Conjecture~\ref{hypo:HPC} holds for some $\tau \in (0,1)$. If $\lambda /\sigma \lesssim p^{K(1- \tau)/4}$, then for any polynomial times estimator $\hat \mM_k$, the following inequality holds
\begin{align}
    \liminf_{p \rightarrow \infty} \sup_{(\tS, \mM_1,...,\mM_K) \text{ has singular signal }\lambda} \bbE \onorm{ \sin(\hat \mM_k, \mM_k) }^2_{\sigma} \geq c,
\end{align}
%\fixme{Miaoyan}{what norm in Lemmas 7 and 8?}
where $c >0$ is a positive constant. 
%The term $\sin \Theta( \cdot, \cdot)$ denotes the angle distance between two matrices with $\Theta(\cdot, \cdot)$ referring to the principal angles between two matrices.
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:hpc}]
See \citet[Proof of Theorem 4]{zhang2018tensor}.
\end{proof}

\begin{lem}[Angle distance and misclassification error]\label{lem:angle_mcr} Consider two membership matrices $\mM, \hat \mM \in \bbR^{p \times r}$ with corresponding assignments $z, \hat z: [p] \mapsto [r]$. If the angle distance $\onorm{\sin (\mM, \hat \mM)}_{\sigma} \geq c$ for a constant $c >0$, then the misclassification error $\min_{\pi \in \Pi}p\ell(\hat z,\pi \circ z )\geq 1$.
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:angle_mcr}]
By Lemma~\ref{lem:equivalence_qnorm}, we know that for any permutation matrix $\mP \in \bbR^{r \times r}$,
\begin{equation}\label{eq:sin}
    \onorm{ \hat \mM \hat \mM^T  - \mM \mP \mP ^T\mM^T}_F \geq  \onorm{ \hat \mM \hat \mM^T  - \mM \mP \mP^T \mM^T}_{\sigma} \geq \onorm{\sin (\mM, \hat \mM)}_{\sigma} \geq c>0,
\end{equation}
where the second inequality follows from the invariance of $\sin$ distance to permutation. The inequality~\eqref{eq:sin} implies that there exists at least one disagreement in $z$ and $\hat z$; i.e., there exists a pair $(i,j)$ such that $z(i) = z(j)$ but $\hat z(i) \neq \hat z(j)$. Therefore, we have $\min_{\pi \in \Pi} p\ell(\hat z, \pi \circ z)\geq 1$.
\end{proof}
%\fixme{Miaoyan}{Effect of permutation $\pi$ in the proof?. } \textcolor{blue}{$\sin \Theta (\mM, \hat \mM)$ is invariant to the permutation.}
%\fixme{Miaoyan}{State Lemma 3 of Zhang and Xia. }

\section*{Proof of Theorem~\ref{thm:initial}}
\begin{proof}[Proof of Theorem 4] Recall that we prove the Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$. We drop the subscript $k$ in the matricizations $\mX_k, \mS_k, \mX_k$. For simplicity, let $\hat z$ denote the output of Sub-Algorithm~\hyperref[alg:main]{1} $\hat z^{(0)}$.

First, by Lemma~\ref{lem:angle_gap_x}, there exists a positive constant such that $\min_{z(i) \neq z(j)} \onorm{ \mX^s_{i:} - \mX^s_{j:} } \geq c_0  \Delta_{\min}$. Also, by balanced assumption on $\mtheta$ and Lemma~\ref{lem:upper_mis}, we have 
 \begin{equation}\label{eq:theta_bound}
          \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2  \leq  \sum_{i \in S_0} \theta(i)^2 + 4 \sum_{i \in S} \theta(i)^2 ,
    \end{equation}
    where 
    \begin{equation}
        S_0 = \{i: \hat z(i) = 0\}, \quad S = \{i \in S_0^c: \onorm{\hat \mx_{\hat z(i)} - \mX_{i:}^s } \geq c_0  \Delta_{\min}/2\}.
    \end{equation}
    On one hand, for any set $P \in [p]$ note that 
    \begin{align}
        \sum_{i \in P} \onorm{\mX_{i:}}^2 &= \sum_{i \in P} \onorm{ \theta(i) \mS_{z(i):} (\mTheta \mM)^{T, \otimes (K-1)} }^2 \\
        & \geq \sum_{i \in P} \theta(i)^2 \min_{a \in [r]} \onorm{\mS_{a:}}^2 \lambda_{r}^{2(K-1)} (\mTheta \mM)\\
        & \geq C \sum_{i \in P} \theta(i)^2 p^{K-1} r^{-(K-1)}, 
    \end{align}
    where the last inequality follows Lemma~\ref{lem:singular_thetam}, the assumption that $\min_{i \in [p]} \theta(i) \geq c >0$ and $\min_{a \in [r]} \onorm{\mS_{a:}} \geq c_3$ in the parameter space~\eqref{eq:family}. Thus, we have 
    \begin{equation}\label{eq:theta_p}
        \sum_{i \in P} \theta(i)^2 \lesssim \sum_{i \in P} \onorm{\mX_{i:}}^2 p^{-(K-1)} r^{K-1}.
    \end{equation}

    On the other hand, note that 
    \begin{align}
         \sum_{i \in S} \onorm{\mX_{i:}}^2 & \leq 2 \sum_{i \in S} \onorm{\hat \mX_{i:}}^2  +  2\sum_{i \in S} \onorm{\hat \mX_{i:} - \mX_{i:}}^2 \label{eq:tr1}\\
         & \leq \frac{8}{c_0^2 \Delta_{\min}^2  } \sum_{i \in S} \onorm{\hat \mX_{i:}}^2   \onorm{\hat \mx_{\hat z(i)} - \mX_{i:}^s }^2 +2 \onorm{\hat \tX - \tX}_F^2 \label{eq:defs} \\
         & \leq \frac{16}{c_0^2 \Delta_{\min}^2  }\sum_{i \in S} \onorm{\hat \mX_{i:}}^2  \off{ \onorm{\hat \mx_{\hat z(i)} - \hat \mX_{i:}^s }^2 + \onorm{\hat \mX_{i:}^s - \mX_{i:}^s }^2 } + 2\onorm{\hat \tX - \tX}_F^2 \label{eq:tr2} \\
         & \leq  \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 }\sum_{i \in S} \onorm{\hat \mX_{i:}}^2  \onorm{\hat \mX_{i:}^s - \mX_{i:}^s }^2  + 2\onorm{\hat \tX - \tX}_F^2 \label{eq:km} \\
         & \leq \of{ \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 } + 2 }\onorm{\hat \tX - \tX}_F^2 \label{eq:lem1}\\
         & \lesssim \of{ \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 } + 2 } \of{ p^{K/2}r + pr^2 + r^K } \sigma^2 \label{eq:lem_err},
    \end{align}
    where inequalities \eqref{eq:tr1} and \eqref{eq:tr2} follow the triangle inequality, \eqref{eq:defs} follows the definition of $S$, \eqref{eq:km} follows the update rule of $k$-means in Sub-Algorithm~\hyperref[alg:main]{1}, \eqref{eq:lem1} follows Lemma~\ref{lem:norm_diff}, and the last inequality~\eqref{eq:lem_err} follows Lemma~\ref{lem:two-step_esterror}. Also, note that 
    \begin{equation}\label{eq:s0}
         \sum_{i \in S_0} \onorm{\mX_{i:}}^2 =  \sum_{i \in S_0} \onorm{\hat \mX_{i:}- \mX_{i:}}^2 \leq \onorm{\hat \tX - \tX}_F^2 \lesssim \of{ p^{K/2}r + pr^2 + r^K } \sigma^2,
    \end{equation}
    where the equation follows by the definition of $S_0$. Therefore, combining the inequalities~\eqref{eq:theta_bound}, \eqref{eq:theta_p}, \eqref{eq:lem_err}, and \eqref{eq:s0}, we have 
    \begin{align}
        \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2  &\lesssim \of{\sum_{i \in S} \onorm{\mX_{i:}}^2 + \sum_{i \in S_0} \onorm{\mX_{i:}}^2  } p^{-(K-1)} r^{K-1} \\
        & \lesssim \frac{\sigma^2 r^{K-1}}{ \Delta_{\min}^2  p^{K-1} } \of{ p^{K/2}r + pr^2 + r^K }.
    \end{align}
    With the assumption that $\min_{i \in [p]} \theta(i) \geq c >0$, we finally obtain the result
    \begin{equation}
        \ell(z, z) \lesssim \frac{1}{p}\min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \lesssim \frac{r^K p^{-K/2} }{ \text{SNR} },
    \end{equation}
    where the last inequality follows the definition $\text{SNR} = \Delta_{\min}^2/\sigma^2$.
\end{proof}

\paragraph{Useful Corollary of Theorem~\ref{thm:initial}}

\begin{cor}\label{cor:L0} Suppose we have the initial assignment $z^{(0)}$ from Sub-Algorithm~\hyperref[alg:main]{1} and the identity permutation minimizes 
the misclassification error, i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} $ $ \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$. Suppose SNR $\gtrsim p^{-K/2} \log p$. Then, the misclassification loss for the initialization is upper bounded as 
\begin{equation}
    L^{(0)} \lesssim \frac{\Delta_{\min}^2}{r \log p}.
\end{equation}
\end{cor}

\begin{proof}
Note that 
\begin{align}
     \sum_{i \in [p]} \onorm{\mX_{i:}^s - \hat \mx_a}^2 &\leq  2\sum_{i \in [p]} \onorm{\mX_{i:}^s - \hat \mX_{i:}^s}^2 + \sum_{i \in [p]} \onorm{\hat \mX_{i:}^s - \hat \mx_a }^2 \\
    & \leq \frac{2}{\min_{i \in [p]} \onorm{\mX_{i:}}^2}  \off{ \sum_{i \in [p]} \onorm{\mX_{i:}}^2  \onorm{\mX_{i:}^s - \hat \mX_{i:}^s}^2 +  \sum_{i \in [p]}  \onorm{\mX_{i:}}^2  \onorm{\hat \mX_{i:}^s - \hat \mx_a }^2 }\\
    & \lesssim \frac{r^{K-1} (1 + \eta)}{p^{K-1}} \sum_{i \in [p]} \onorm{\mX_{i:}}^2  \onorm{\mX_{i:}^s - \hat \mX_{i:}^s}^2 \\
    & \lesssim  \frac{r^{K-1} (1 + \eta)}{p^{K-1}}  \onorm{\hat  \tX - \tX}_F^2 \\
    & \lesssim \frac{r^{K-1} (1 + \eta)}{p^{K-1}} p^{-K/2}.
\end{align}

Suppose $z^{(0)} = 1$.  Note that $\mX^s_{i:}$ only have $r$ different values. Let $\mX_{a}^s = \mX^s_{i:}$ with $z(i) = a, a \in [r]$. Note that 
\begin{equation}
    |z^{-1}(a) \cap (z^{(0)})^{-1} (a)| \geq |z^{-1}(a)| - p\ell(z^{(0)} - z) \gtrsim \frac{p}{r} - \frac{1}{\log p} \gtrsim \frac{p}{r},
\end{equation}
when $\Delta_{\min}^2 \geq p^{-K/2} \log p$. Then
\begin{align}
    \onorm{\mX_a^s - \hat \mx_a}^2 &= \frac{ \sum_{i \in z^{-1}(a) \cap (z^{(0)})^{-1} (a) } \onorm{ \mX_{i:}^s - \hat \mx_a}^2 }{|z^{-1}(a) \cap (z^{(0)})^{-1} (a) |} \\
    & \lesssim \frac{ r \sum_{i \in [p]} \onorm{\mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 }{ p}\\
    & \lesssim \frac{r^K(1 + \eta)}{p^{K}} p^{-K/2} 
\end{align}

Note that 
\begin{align}
    L^{(0)} &= \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \onorm{\mS_{z^{(0)}(i),:}^s - \mS_{z(i),:}^s}^2 \\
    & \lesssim \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \onorm{\mX^s_{z^{-1}(z^{(0)}):} - \mX^s_{i,:}}^2\\
    & \lesssim \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \off{ \onorm{ \mX^s_{z^{-1}(z^{(0)}):} - \hat \mx_{z^{(0)}(i)} }^2 + \onorm{ \hat \mx_{z^{(0)}(i)} - \mX^s_{i,:}}^2 }\\
    & \lesssim   \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \frac{ \onorm{\hat \mX_{i:} - \mX_{i:}}^2 }{ \onorm{ \mX_{i:}}^2 } +  \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i)  \frac{r^K(1 + \eta)}{p^{K}} p^{-K/2} \\
    & \lesssim \frac{r^K}{p^{K}} p^{-K/2} \\
    & \lesssim \frac{\Delta_{\min}^2}{r \log p}.
    %& \lesssim \frac{1}{p} \lambda^{-2(K-1)}_r(\mTheta \mM) \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \onorm{\mX^{(0)}_{i:}-\mX_{i:}}^2 \\
    %& \lesssim \frac{r^{K-1}}{p^K} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \off{ \onorm{\mX^{(0)}_{i:}-\hat \mX_{i:}}^2 +  \onorm{\hat \mX_{i:}-\mX_{i:}}^2 }
\end{align}
where $\frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \lesssim \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i)^2$ due to the boundedness.
\end{proof}

\paragraph{Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:initial}} 
\begin{lem}[Basic inequality]\label{lem:norm_diff} For any two nonzero vectors $\mv_1,\mv_2$ of same dimension, we have 
\[
\sin(\mv_1 , \mv_2) \leq \onorm{{\mv_1^s}-{\mv_2^s}}\leq {2\onorm{\mv_1-\mv_2}\over \max\left(\onorm{\mv_1},\onorm{\mv_2}\right)}.
\]
\end{lem}
\begin{proof}[Proof of Lemma~\ref{lem:norm_diff}]
For the first inequality, let $\alpha \in [0,\pi]$ denote the angle between $\mv_1$ and $\mv_2$. We have 
\begin{align}
     \onorm{{\mv_1^s}-{\mv_2^s}} = \sqrt{2(1 - \cos \alpha)} = 2 \sin \frac{\alpha}{2} \geq \sin \alpha,
\end{align}
where the equations follows by the properties of trigonometric function and the inequality follows by the fact the $\cos \frac{\alpha}{2} \leq 1$ and $\sin \alpha = 2 \sin \frac{\alpha}{2} \cos \frac{\alpha}{2} > 0$ for $\alpha \in [0, \pi]$.

For the second inequality, without loss of generality, we assume $\onorm{\mv_1}\geq \onorm{\mv_2}$. Then
\begin{align}
\onorm{{\mv_1^s}-{\mv_2^s}}&=\onorm{{\mv_1\over \onorm{\mv_1}}- {\mv_2\over \onorm{\mv_1}}+{\mv_2\over \onorm{\mv_1}}-{\mv_2\over \onorm{\mv_2}}}\\
&\leq {\onorm{\mv_1-\mv_2}\over \onorm{\mv_1}}+{\onorm{\mv_2}\onorm{\mv_1}-\onorm{\mv_2}\over \onorm{\mv_1}\onorm{\mv_2}}\\
&\leq {2\onorm{\mv_1-\mv_2}\over \onorm{\mv_2}}.
\end{align}

Therefore, Lemma~\ref{lem:norm_diff} is proved.
\end{proof}

\begin{defn}[Weighted padding vectors]\label{def:pad} For a vector $\ma = \entry{a_i} \in \bbR^d$, we define the padding vector of $\ma$ with the weight collection $\mw = \{\mw_i\}_{i = 1}^d$ and $\mw_i = \entry{w_{ik}} \in \bbR^{p_i}, i \in [d]$ as
\begin{equation}
    \pad_{\mw}(\ma) = [a_1 \circ \mw_1, \ldots, a_d \circ \mw_d], \quad \text{ where} \quad  a_i \circ \mw_i = [a_i w_{i1},\ldots, a_i w_{i p_i}], \text{ for all } i \in [d]
\end{equation}
and $\pad: \bbR^{d}\mapsto \bbR^{\sum_{i\in[d]}}p_i$ can be viewed as an operator, and by definition, we have $\pad (\ma) \in \bbR^{\sum_{i \in[d]} p_i}$. Then , we have the bounds of the weighted padding vector
\begin{equation}\label{eq:pad_bound}
     \min_{i \in [d]} \onorm{\mw_i}^2 \onorm{\ma}^2 \leq \onorm{\pad_{\mw}(\ma)}^2 \leq \max_{i \in [d]} \onorm{\mw_i}^2 \onorm{\ma}^2.
\end{equation}
Further, we define the inverse weighted padding operator $\pad^{-1}: \bbR^{\sum_{i\in[d]}p_i}\mapsto \bbR^d$ which satisfies 
\begin{equation}
    \pad^{-1}_{\mw}(\pad_{\mw}(\ma)) = \ma.
\end{equation}
\end{defn}

\begin{lem}[Angle for weighted padding vectors]\label{lem:pad} Suppose we have two non-zero vectors $\ma, \mb \in \bbR^d$. Given the weight collection $\mw$, we have 
\begin{equation}
   \frac{\min_{i \in [d]} \onorm{\mw_i}}{\max_{i \in [d]} \onorm{\mw_i} } \sin (\ma, \mb) \leq \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb)) \leq \frac{ \max_{i \in [d]} \onorm{\mw_i} }{\min_{i \in [d]} \onorm{\mw_i}} \sin (\ma, \mb).
\end{equation}
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:pad}] We prove the two inequalities separately with the similar idea.

First, we prove the inequality $\sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb)) \leq \frac{ \max_{i \in [d]} \onorm{\mw_i}^2 }{\min_{i \in [d]} \onorm{\mw_i}^2} \sin (\ma, \mb)$. Decomposing $\mb$, we have
\begin{equation}
    \mb = \cos (\ma, \mb) \frac{\onorm{\mb}}{\onorm{\ma}} \ma + \sin (\ma, \mb) \frac{\onorm{\mb}}{\onorm{\ma^{\perp}}}\ma^{\perp},
\end{equation}
where $\ma^{\perp} \in \bbR^d$ is the orthogonal vector of $\ma$. By the Definition~\ref{def:pad}, we have 
\begin{equation}\label{eq:pad_second_ineq}
    \pad_{\mw}(\mb) = \cos (\ma, \mb) \frac{\onorm{b}}{\onorm{a}} \pad_{\mw}(\ma) + \sin (\ma, \mb) \frac{\onorm{b}}{\onorm{a^{\perp}}} \pad_{\mw}(\ma^{\perp}).
\end{equation}
Note that $\pad_{\mw}(\ma^{\perp})$ is not necessary equal to the orthogonal vector of $\pad(\ma)$, i.e., $\pad_{\mw}(\ma^{\perp}) \neq (\pad_{\mw}(\ma))^{\perp}$. Then by the geometry property of trigonometric functions, we obtain
\begin{align}
    \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb))  &\leq \frac{  \onorm{\mb} \onorm{\pad_{\mw}(\ma^{\perp})} }{ \onorm{\ma^{\perp}} \onorm{ \pad_{\mw}(\mb)}} \sin (\ma, \mb)  \\
    & \leq  \frac{ \max_{i \in [d]} \onorm{\mw_i} }{\min_{i \in [d]} \onorm{\mw_i}} \sin (\ma, \mb),
\end{align}
where the second inequality follows by applying the property~\eqref{eq:pad_bound} to vectors $\mb, \ma^{\perp}$.

Next, we prove $\frac{\min_{i \in [d]} \onorm{\mw_i}^2}{\max_{i \in [d]} \onorm{\mw_i}^2 } \sin (\ma, \mb) \leq \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb))$. With the decomposition of $\pad_{\mw}(\mb)$ and the inverse weighted padding operator, we have 
\begin{align}
    \mb &=  \cos(\pad_{\mw}(\ma), \pad_{\mw}(\mb)) \frac{\onorm{\pad_{\mw}(\mb)}}{\onorm{\pad_{\mw}(\ma)}} \ma +\\
    & \quad \quad \quad \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb))\frac{\onorm{\pad_{\mw}(\mb)}}{\onorm{(\pad_{\mw}(\ma))^{\perp}}} \pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp}).
\end{align}

Therefore, we obtain 
\begin{align}
    \sin(\ma, \mb) &\leq \frac{ \onorm{\pad_{\mw}(\mb) } \onorm{\pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp})  } }{ \onorm{(\pad_{\mw}(\ma))^{\perp} } \onorm{\mb} }  \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb))\\
    & \leq \frac{ \max_{i \in [d]} \onorm{\mw_i} }{\min_{i \in [d]} \onorm{\mw_i}} \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb)),
\end{align}
where the second inequality follows by applying the property~\eqref{eq:pad_bound} to vectors $\mb, \pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp})$.

\end{proof}

\begin{lem}[Angle gap in $\tX$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor}. Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying~\eqref{eq:degree}. Then the angle gap in $\tX$ is approximate to angle gap in $\tS$, i.e.,
\begin{equation}
    \Delta_{\min} \lesssim  \min_{(i,j): z(i) \neq z(j)} \onorm{\mX_{i:}^s - \mX_{j:}^s}. 
\end{equation}
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:angle_gap_x}] 
Note that the vector $\mS_{z(i):}$ can be folded to a tensor $\tS' =\entry{\tS'_{a_2,\ldots,a_K}} \in \bbR^{r^{K-1}}$, i.e., $\text{vec}(\tS') = \mS_{z(i):}$. Let the weight vector $\mw_{a_2 \cdots a_K}$ correspond to the elements $\tS'_{a_2,\ldots,a_K}$ as
\begin{equation}
    \mw_{a_2 \cdots a_K} = [ \mtheta_{z^{-1} (a_2)} \otimes \cdots \otimes \mtheta_{z^{-1} (a_K)}] \in \bbR^{|z^{-1} (a_2)| \times \cdots \times |z^{-1} (a_K)|},
\end{equation}
for all $a_k \in [r], k = 2,\ldots, K$, where $\otimes$ denotes the Kronecker product. Therefore, we have  $\mX_{i:} = \theta(i) \pad_{\mw}(\mS_{z(i):})$ where $\mw = \{ \mw_{a_2 \cdots a_K}\}$. Specifically, we have $\onorm{\mw_{a_2,\ldots,a_K}}^2 = \prod_{k = 2}^K \onorm{ \mtheta_{z^{-1}(a_k)}}^2$, and by the balanced assumption we have 
\begin{equation}\label{eq:pad_balance}
    \max_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}}^2 = (1 + o(1))  \min_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}}^2.
\end{equation}

Consider the inner product of $\mX_{i:}$ and $\mX_{j:}$ for $z(i) \neq z(j)$. By the definition of weighted padding operator and the balanced assumption~\eqref{eq:pad_balance}, we have 
\begin{align}
    \ang{\mX_{i:},\mX_{j:}} &= \theta(i) \theta(j) \ang{ \pad_{\mw}(\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):} )}\\
    &= \theta(i) \theta(j) \min_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}}^2 \ang{\mS_{z(i):}, \mS_{z(j):}}(1 + o(1)).
\end{align}
Thus, when $p$ large enough, the inner product $\ang{\mX_{i:},\mX_{j:}} $ as the same sign with $\ang{\mS_{z(i):}, \mS_{z(j):}}$. Next, we discuss the angle between $\mX_{i:}$ and $\mX_{j:}$ by cases.

\begin{enumerate}
    \item \textbf{Case 1:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} \leq 0$. Then, we also have $\ang{\mX_{i:},\mX_{j:}} \leq 0$, which implies $\onorm{\mX_{i:}^s -  \mX_{j:}^s} \geq \sqrt{2}$. Note that $\onorm{\mS_{z(i):}^s - \mS_{z(j):}^s} \leq 2$. We have $\onorm{\mX_{i:}^s -  \mX_{j:}^s} \gtrsim \onorm{\mS_{z(i):}^s - \mS_{z(j):}^s}$.
    \item \textbf{Case 2:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} > 0$. Then, we have $\cos(\mS_{z(i):}, \mS_{z(j):}) > 0$. Note that the fact $\sqrt{1 - \cos \alpha} = 2 \sin \frac{\alpha}{2} \lesssim \sin \alpha$ for the angle $\alpha \in [0, \frac{\pi}{2})$. Then, we have 
    \begin{align}
        \onorm{\mS_{z(i):}^s - \mS_{z(j):}^s} &= \sqrt{ 1 - \cos(\mS_{z(i):}, \mS_{z(j):})}  \\
        & \lesssim \sin (\mS_{z(i):}, \mS_{z(j):})\\
        & \leq \frac{ \max_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}} }{  \min_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}} }  \sin ( \pad_{\mw} (\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):})) \\
        & \leq (1 + o(1)) \onorm{\mX_{i:}^s -  \mX_{j:}^s},
    \end{align}
    where the second inequality follows by Lemma~\ref{lem:pad}, and the last inequality follows by the balanced assumption~\eqref{eq:pad_balance} and Lemma~\ref{lem:norm_diff}.
\end{enumerate}

Hence, we conclude that 
\begin{equation}
   \min_{(i,j): z(i) \neq z(j)} \onorm{\mX_{i:}^s - \mX_{j:}^s} \gtrsim \min_{(i,j): z(i) \neq z(j)}  \onorm{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
\end{equation}

\end{proof}

% % generalized version
% \begin{lem}[Angle gap in $\tX$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:general_dtbm}. Suppose Assumption~\ref{assmp:general_minimal_gap} holds and $\{\mtheta_k\}$ is balanced satisfying~\eqref{eq:general_balanced}. Then the angle gap in $\tX$ is approximate to angle gap in $\tS$, i.e.,
% \begin{equation}
%     \min\{1, \Delta_{\min}\} \lesssim \min_{k \in [K]} \min_{(i,j): z_k(i) \neq z_k(j)} \onorm{\mX_{k,i:}^s - \mX_{k,j:}^s}. % \lesssim \Delta_{\min}.
% \end{equation}
% \end{lem}

% \begin{proof}[Proof of Lemma~\ref{lem:angle_gap_x}] 
% For simplicity, we consider the case with $k = 1$ and omit the subscript $k$ in the matricizations $\mX,\mS$.

% Note that the vector $\mS_{z_1(i):}$ can be folded to a tensor $\tS' =\entry{\tS'_{a_2,\ldots,a_K}} \in \bbR^{r_2 \times \cdots \times r_K}$, i.e., $\text{vec}(\tS') = \mS_{z_1(i):}$. Let the weight vector $\mw_{a_2 \cdots a_K}$ correspond to the elements $\tS'_{a_2,\ldots,a_K}$ as
% \begin{equation}
%     \mw_{a_2 \cdots a_K} = [ \mtheta_{2, z_2^{-1} (a_2)} \otimes \cdots \otimes \mtheta_{K, z_K^{-1} (a_K)}] \in \bbR^{|z_2^{-1} (a_2)| \times \cdots \times |z_K^{-1} (a_K)|},
% \end{equation}
% for all $a_k \in [r_k], k = 2,\ldots, K$, where $\otimes$ denotes the Kronecker product. Therefore, we have  $\mX_{i:} = \theta_1(i) \pad_{\mw}(\mS_{z_1(i):})$ where $\mw = \{ \mw_{a_2 \cdots a_K}\}$. Specifically, we have $\onorm{\mw_{a_2,\ldots,a_K}}^2 = \prod_{k = 2}^K \onorm{ \mtheta_{k,z_k^{-1}(a_k)}}^2$, and by the balanced assumption we have 
% \begin{equation}\label{eq:pad_balance}
%     \max_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}}^2 = (1 + o(1))  \min_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}}^2.
% \end{equation}

% Consider the inner product of $\mX_{i:}$ and $\mX_{j:}$ for $z_1(i) \neq z_1(j)$. By the definition of weighted padding operator and the balanced assumption~\eqref{eq:pad_balance}, we have 
% \begin{align}
%     \ang{\mX_{i:},\mX_{j:}} &= \theta_1(i) \theta_1(j) \ang{ \pad_{\mw}(\mS_{z_1(i):}), \pad_{\mw}(\mS_{z_1(j):} )}\\
%     &= \theta_1(i) \theta_1(j) \min_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}}^2 \ang{\mS_{z_1(i):}, \mS_{z_1(j):}}(1 + o(1)).
% \end{align}
% Thus, when $p$ large enough, the inner product $\ang{\mX_{i:},\mX_{j:}} $ as the same sign with $\ang{\mS_{z(i):}, \mS_{z(j):}}$. Next, we discuss the angle between $\mX_{k,i:}$ and $\mX_{k,j:}$ by cases.

% \begin{enumerate}
%     \item \textbf{Case 1:} Suppose $\ang{\mS_{z_1(i):}, \mS_{z_1(j):}} \leq 0$. Then, we also have $\ang{\mX_{i:},\mX_{j:}} \leq 0$, which implies $\onorm{\mX_{i:}^s -  \mX_{j:}^s} \geq \sqrt{2}$.
%     \item \textbf{Case 2:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} > 0$. Then, we have $\cos(\mS_{z(i):}, \mS_{z(j):}) > 0$. Note that the fact $\sqrt{1 - \cos \alpha} = 2 \sin \frac{\alpha}{2} \lesssim \sin \alpha$ for the angle $\alpha \in [0, \frac{\pi}{2})$. Then, we have 
%     \begin{align}
%         \onorm{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} &= \sqrt{ 1 - \cos(\mS_{z_1(i):}, \mS_{z_1(j):})}  \\
%         & \lesssim \sin (\mS_{z_1(i):}, \mS_{z_1(j):})\\
%         & \leq \frac{ \max_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}} }{  \min_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}} }  \sin ( \pad_{\mw} (\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):})) \\
%         & \leq (1 + o(1)) \onorm{\mX_{i:}^s -  \mX_{j:}^s},
%     \end{align}
%     where the second inequality follows by Lemma~\ref{lem:pad}, and the last inequality follows by the balanced assumption~\eqref{eq:pad_balance} and Lemma~\ref{lem:norm_diff}.
% \end{enumerate}

% Hence, we conclude that for all $i,j \in [p_1], z_1(i) \neq z_1(j)$ 
% \begin{equation}
%     \onorm{\mX_{i:}^s -  \mX_{j:}^s} \gtrsim \min \offf{ 1, \onorm{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s}}, 
% \end{equation}
% and the same conclusion for matricizations $\mX_k, \mS_k, k \in [K]$ can be obtained by the same procedures. Therefore, by the definition $\Delta_{\min} \coloneqq \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onorm{ \mS_{k,a:}^s - \mS_{k, b:}^s }$, we finally obtain that 
% \begin{equation}
%     \min_{k \in [K]} \min_{(i,j): z_k(i) \neq z_k(j)} \onorm{\mX_{k,i:}^s - \mX_{k,j:}^s} \gtrsim \min\{1, \Delta_{\min}\}.
% \end{equation}

% % Define an order-$(K-1)$  $(r,...,r)$-dimensional weight tensor $\tW$ with entries $\tW(a_2,...,a_K) \coloneqq \prod_{k=2}^K \onorm{\mtheta_{z^{-1}(a_k)}}^2$ and let $ \mw = \Vec(\tW)$. By Lemma~\ref{lem:weighted},
% % \begin{equation}
% %     \ang{ \mX_{k,i:},  \mX_{k,j:} } = \theta(i) \theta(j) \ang{ \mS_{k,z(i):}, \mS_{k,z(j):} }_{\mw} = \theta(i) \theta(j) \ang{ \mS_{k,z(i):}, \mS_{k,z(j):} } + o(1),
% % \end{equation}
% % and 
% % \begin{equation}
% %     \onorm{ \mX_{k,i:} } = \theta(i) \onorm{ \mS_{k,z(i):} }_{\mw} = \theta(i) \onorm{ \mS_{k,z(i):} } + o(1).
% % \end{equation}
% % Then, we have 
% % \begin{equation}
% %     \cos ( \mX_{k,i:},  \mX_{k,j:}  ) = \frac{ \ang{ \mX_{k,i:},  \mX_{k,j:} } }{\onorm{ \mX_{k,i:} } \onorm{ \mX_{k,j:} }} = \frac{\ang{ \mS_{k,z(i):}, \mS_{k,z(j):} } }{\onorm{ \mS_{k,z(i):} } \onorm{ \mS_{k,z(j):} }} + o(1) = \cos (\mS_{k,z(i):},\mS_{k,z(j):} ) + o(1).
% % \end{equation}
% % Hence, we finally obtain that 
% % \begin{align}
% %     \min_{k \in [K]} \min_{(i,j): z_k(i) \neq z_k(j)} \onorm{\mX_{k,i:}^s - \mX_{k,j:}^s}  &=  \min_{k \in [K]} \min_{(i,j): z_k(i) \neq z_k(j)} \sqrt{ 1 - \cos ( \mX_{k,i:},  \mX_{k,j:}  )}\\
% %     &= \min_{k \in [K]} \min_{(i,j): z_k(i) \neq z_k(j)} \sqrt{ 1 - \cos (\mS_{k,z(i):},\mS_{k,z(j):} ) }  + o(1) \\
% %     &\asymp \Delta_{\min},
% % \end{align}
% % where the last inequality follows by the minimal gap assumption.

% % \fixme{Jiaxin}{ Problem here: if $\Delta_{\min}^2 = O(p^{-K/2})$, then $o(1)$ is much larger than $\Delta_{\min}$...}
% \end{proof}

\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the assumption that $\min_{i \in [p]} \theta(i) \geq c_0 >0$, we have the minimal singular value of $\mTheta \mM$ bounded as 
\begin{equation}\label{eq:membership2}
     \sqrt{p/r} \leq \sqrt{ \min_{a \in [r]}\onorm{ \mtheta_{ z^{-1}(a)} }^2 }  \lesssim \lambda_{r}(\mTheta \mM) \leq \onorm{\mTheta \mM}_{\sigma} \leq\sqrt{ \max_{a \in [r]}\onorm{ \mtheta_{ z^{-1}(a)} }^2 } \lesssim  p /r .
\end{equation}
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:singular_thetam}] Note that 
\begin{equation}
    (\mTheta\mM)^T \mTheta \mM = \mD, \quad \text{with} \quad \mD = \text{diag}(D_1,\ldots, D_r), \ D_a = \onorm{\mtheta_{z^{-1}(a)}}^2, a \in [r].
\end{equation}
By the definition of singular value, we have 
\begin{equation}
     \sqrt{\min_{a \in [r]}\onorm{ \mtheta_{ z^{-1}(a)} }^2 } \leq \lambda_r(\mTheta \mM) \leq \onorm{\mTheta\mM}_{\sigma} \leq \sqrt{ \max_{a \in [r]}\onorm{ \mtheta_{ z^{-1}(a)} }^2}.
\end{equation}

Given that $\min_{i \in [p]}\theta(i) \geq c_0$, we have  
\begin{equation}
   \min_{a \in [r]} \onorm{ \mtheta_{ z^{-1}(a)} }^2 \geq c_0^2 \min_{a \in [r]} |z^{-1}(a)| \gtrsim \frac{p}{r},
\end{equation}
where the last inequality follows by the constraint in parameter space~\eqref{eq:family}. Also, notice that 
\begin{equation}
    \sqrt{\max_{a \in [r]}\onorm{ \mtheta_{ z^{-1}(a)} }^2 } \leq \max_{a \in [r]} \sqrt{ \onorm{ \mtheta_{ z^{-1}(a)} }^2_1} \lesssim \frac{p}{r}.
\end{equation}

 Therefore, we complete the proof of Lemma~\ref{lem:singular_thetam}.
\end{proof}

\begin{lem}[Singular-value-gap-free tensor estimation error bound]\label{lem:two-step_esterror} Let an order-$K$ tensor $\tA = \tX + \tZ \in \bbR^{p \times \cdots \times p}$, and $\tX$ has tucker rank $(r,...r)$ and $\tZ$ has independent sub-Gaussian entries with parameter $\sigma^2$. Let $\hat \tX$ denote the two-step estimated tensor in line 2 of Sub-Algorithm 1 in the main paper. Then with probability at least $1 - C \exp\of{- cp }$, we have
\begin{align}
    \onorm{\hat \tX - \tX}_F^2 \leq C \sigma^2 \of{ p^{K/2}r + pr^2 + r^K }.
\end{align}
\end{lem}

\begin{proof}
See Proposition 1 in \cite{han2020exact}.
\end{proof}

\begin{lem}[Upper bound of misclassification error]\label{lem:upper_mis} Let $z: [p] \mapsto [r]$ be a clustering assignment such that $|z^{-1}(a)| \asymp p/r$ for all $a \in [r]$. Let node $i$ corresponds to a vector $\mx_i  = \theta(i) \mv_{z(i)} \in \bbR^d$ where $\{\mv_a\}_{a = 1}^r$ are the cluster centers and $\mtheta = \entry{\theta(i)} \in \bbR^p_+$ is the positive degree heterogeneity. Assume that  $\mtheta$ satisfies the balanced assumption such that ${\max_{a \in [r]} \onorm{\mtheta_{z^{-1}(a)}}^2 \over \min_{a \in [r]} \onorm{\mtheta_{z^{-1}(a)}}^2 }= 1 + o(1)$. Consider an arbitrary estimate $\hat z$ with $\hat \mx_i = \hat \mv_{\hat z(i)}$. Then, if
\begin{equation}
    \min_{a \neq b \in [r]} \onorm{\mv_a - \mv_b} \geq 2c,
\end{equation}
 for some constant $c >0$, we have 
\begin{equation}
    \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \leq \sum_{i \in S_0} \theta(i)^2 + 4 \sum_{i \in S} \theta(i)^2,
\end{equation}
where 
\begin{equation}
    S_0 = \{i: \hat z(i) = 0\} \quad  \text{and}  \quad S = \{i \in S_0^c: \onorm{\hat \mx_i - \mx_i } \geq c \}.
\end{equation}

\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:upper_mis}]

For each cluster $u\in[r]$, we use $C_u$ to collect the subset of points for which the estimated and true positions $\hat \mx_i, \mx_i$ are within distance $c$. Specifically, define
For each cluster $u \in [r]$, we define

\begin{equation}
    C_u = \{ i \in z^{-1}(u) \cap S_0^c: \onorm{\hat \mx_i - \mx_i} < c \},
\end{equation}
and divide $[r]$ into three groups based on $C_u$ as 
\begin{align}
    R_1 &= \{ u \in [r]: C_u = \emptyset \},\\
    R_2 &= \{ u \in [r]: C_u \neq \emptyset, \text{ for all } i, j \in C_u, \hat z(i) = \hat z(j) \},\\
    R_3 &= \{ u \in [r]: C_u \neq \emptyset, \text{ there exist } i, j \in C_u, \hat z(i) \neq \hat z(j) \}.
\end{align}
Then, we have 
\begin{equation}
    \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \leq \sum_{i \in S_0} \theta(i)^2 + \sum_{i \in S} \theta(i)^2 + \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2.
\end{equation}
Thus, we only need to bound $\sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2$ to finish the proof. Note that we have $|R_3| \leq |R_1|$ by Lemma 6 in \citep{gao2018community}. We have 
\begin{align}
    \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2 &\leq |R_3| \max_{a \in [r]} \onorm{\mtheta_{z^{-1}(a)}}^2 \\
    & \leq |R_1| \max_{a \in [r]} \onorm{\mtheta_{z^{-1}(a)}}^2  \\
    & \leq {\max_{a \in [r]} \onorm{\mtheta_{z^{-1}(a)}}^2 \over \min_{a \in [r]} \onorm{\mtheta_{z^{-1}(a)}}^2 } \sum_{i \in \cup_{u \in R_1} z^{-1}(u)} \theta(i)^2\\
    & \leq 2 \sum_{i \in S} \theta(i)^2,
\end{align}
where the second inequality follows the conclusion $|R_3| \leq |R_1|$ from the proof of Lemma 6 in \citep{gao2018community}, and the last inequality holds by the balanced assumption on $\mtheta$ when $p$ is large enough, and the fact that $ \cup_{u \in R_1} z^{-1}(u) \subset S$.

\end{proof}

\section*{Proof of Theorem~\ref{thm:refinement}}

% Extra condition: 
% \begin{equation}\label{cond:extra}
%   c_0 < \min_{i \in [p]} \theta(i), \quad \Delta_{\min}^2 \gtrsim p^{-K/2}\log p,
% \end{equation}
% the second inequality requires the initialization error tends to 0 as $p \rightarrow \infty$. Otherwise, we can not handle $\eta$ in $A_{11}$ very well.

%\fixme{Miaoyan}{We claim the following %lower bound on $\alpha$
%\[
%\sin(\mp(\bar z), \mp_{\theta}(\bar %z))\lesssim \sin\alpha.
%\]
%Therefore, the condition on $\sin(\mp(\bar %z), \mp_{\theta}(\bar z))$ is necessary. %Construct $\bar z$ to be the ground-truth %clustering. Then $\tilde \mD=\mI_r$ and %$\mD=\text{diag}(\mp)^{-1}\text{diag}(\mp_%{\mtheta})$
%\[
%\sin \alpha=\sin(\mS\tilde \mD, \mS) \geq %\lambda_{\min}(\mD){\min_{a\in[r]}\FnormSi%ze{}{\mS_{a:}}\over %\max_{a\in[r]}\FnormSize{}{\mS_{a:}}}\gtrs%im ? \sin(\mp,\mp_{\mtheta})
%\]
%}

%  \begin{defn}[Locally linear stability of degrees] Let $z$ be the ground-truth clustering, and define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\epsilon)=\{\bar z\colon \ell(\bar z, z)\leq \epsilon\}$. We define two cluster-size vectors for $\bar z$,
%  \[
%  \mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T,\quad \mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
%  \]
% We call the degree is $\varepsilon$-locally linearly stable if and only if 
% \begin{equation}
%     \sin(\mp(\bar z),\ \mp_{\mtheta}(\bar z))\lesssim \varepsilon \Delta_{\min} ,\quad \text{for all } \bar z\in \tN(z, \varepsilon).
% \end{equation}

% \end{defn}

%  \begin{rmk}
%  The local stability holds trivially for $\varepsilon=0$ based on the construction of parameter space~\eqref{eq:family}.
%  \end{rmk}

%  \begin{assumption}[Degree stability]\label{assmp:stability}
%  Assume the local linear stability of degree holds in all neighborhoods $\tN(z,\epsilon)$ for any $\epsilon \leq \log^{-1}p$.
%  \end{assumption}

%\begin{rmk}[Comparison to~\cite{gao2018community}]
%We consider the special case when $r=2$, $z^{-1}(1)=\{1,\ldots,p\}$, $z^{-1}(2)=\{p+1,\ldots,2p\}$, $\sigma=1$, and 
%\[
%\mS=
%\begin{bmatrix}
%1 & s_1\\
%1 & s_2
%\end{bmatrix}.
%\]
%Under our model, the signal level is
%\[
%\Delta^2_{\min}\asymp 1- {1+\sqrt{(s^2_1-1)(s^2_2-1))}\over s_1s_2} \asymp 
%\]
%The assumption on $\theta$ is
%\begin{enumerate}
  %  \item $\theta_1+\cdots+\theta_{p}=p$ and $\theta_{p+1}+\cdots+\theta_{2p}=p$.
 %   \item $\mp(\bar z)=(p(1\pm \varepsilon),\ p(1\pm \varepsilon))^T$ and $\mp_{\mtheta}(\bar z)=(p\pm \sum_{i\in[I], |I|\leq p\varepsilon}\theta_i,\ p\pm \sum_{j\in[J], |J|\leq p\varepsilon}\theta_{p+j})^T$, 
 %   \[
 %   {\max_{|I|\leq p\varepsilon}\sum_{j\in[I]}\theta_{i} \over 
 %   \min_{|J|\leq p\varepsilon}\sum_{j\in[J]}\theta_{p+j}
 %   }\lesssim \varepsilon \sqrt{|1-s|}.
 %   \]
%\end{enumerate}
%This model can not be incorporated into~\cite{gao2018community}, because $\mS$ is not assortative.
% \end{rmk}

\begin{proof}[Proof of Theorem~\ref{thm:refinement}]
Recall that we prove Theorem~\ref{thm:refinement} under the dTBM~\ref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$. We drop the subscript $k$ in the matricizations $\mM_k, \mS_k, \mX_k$. Without loss of generality, we assume that the noise variance $\sigma = 1$ and the identity permutation minimizes the initial misclassification error, i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$.

%Here, we also prove Theorem~\ref{thm:refinement} under the symmetric case, i.e., assuming $p_k = p, r_k = r$, and $\tS \in \bbR^{r \times \cdots \times r}, \tX \in \bbR^{p \times \cdots \times p}$ are symmetric with $z_k = z, \mM_k = \mM, \mtheta_k = \mtheta$. Without loss of generality, we assume $\sigma = 1$ and $\pi^{(0)}$ is the identity mapping on $[r]$.

\paragraph{Step 1:} We first introduce additional notations and the necessary condition to complete the proof. We will verify that the conditions hold in our context under high probability in the last step of the proof.

\textbf{Notations.}
\begin{enumerate}

\item Norms. Let $\onorm{\mA}_\sigma$ denote the spectral norm of matrix $\mA$, which is equal to the maximal singular value of $\mA$; let $\onorm{\mA}_F$ denote the Frobenius norm of matrix $\mA$.
\item Projection. For a vector $\mv \in \bbR^d$, let $\text{Proj}(\mv) \in \bbR^{d \times d}$ denote the projection matrix to $\mv$. Then $\mI_d - \text{Proj}(\mv)$ is the projection matrix to the orthogonal complement $\mv^{\perp}$.

    \item Normalized membership matrix
    \begin{align}
        \mW = \mM \of{ \text{diag}(\mone_{p}^T \mM) }^{-1}, \quad \mW^{(t)} = \mM^{(t)} \of{ \text{diag}(\mone_{p}^T \mM^{(t)}) }^{-1},
    \end{align}
    and dual normalized membership matrix
    \begin{align}
         \mV &= \mW^{\otimes (K-1)}, \quad \mV^{(t)} =\of{ \mW^{(t)}}^{\otimes (K-1)}
    \end{align}
    \item Estimator of $\tS$ in the $t$-th iteration, denoted $\tS^{(t)}$ and the oracle estimator of $\tS$ given true assignment $z$, denoted $\tilde \tS$,
    \begin{align}
        \tS^{(t)} = \tY \times_1 \of{\mW^{(t)}}^T \times_2 \cdots \times_K \of{\mW^{(t)}}^T, \quad \tilde \tS &= \tY \times_1 \mW^T \times_2 \cdots \times_K \mW^T.
    \end{align}
    \item Matricization of tensors
    \begin{align}
        &\mS = \mat_1(\tS), \quad \mS^{(t)} = \mat_1(\tS^{(t)}), \quad \tilde \mS = \mat_1(\tilde \tS)\\
        &\mY = \mat_1(\tY), \quad \mX = \mat_1(\tX), \quad \mE = \mat_1(\tE).
    \end{align}
    \item The angle-based misclassification loss in the $t$-th iteration %\fixme{Jiaxin}{Need $\theta(i)$ in the definition}
    \begin{equation}
         L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
    \end{equation}
     and the oracle loss
    \begin{align}
         \xi & = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \Big\{\ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s} \\
         & \hspace{3cm}\leq - \frac{ \theta(i) m}{4} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 \Big\} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{ b:}  }^s  }^2,
    \end{align}
    where $m$ is a positive constant related to $c_3$ and is defined in~\eqref{eq:xv_lower}.
\end{enumerate}

\begin{condition}\label{cond:origin} Let $\mathbb{O}_{p, r}$ denote the collection of all the $p$-by-$r$ matrices with orthonormal columns. For all $a \in [r]$, we have 
\begin{equation}\label{eq:cond1}
    \onorm{\mE \mV}_\sigma \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}} \of{ p ^{1/2}+ r^{(K-1)/2}}, \quad \onorm{\mE \mV}_F \lesssim \sqrt{\frac{ r^{2(K-1)}}{p^{K-2}}},\quad \onorm{ \mW_{a:}^T \mE \mV} \lesssim \frac{r^K}{p^{K/2}},
\end{equation}
% \begin{equation}\label{eq:cond11}
%     , \quad \onorm{ \mW_{a:}^{(t),T} \mE \mV} \lesssim \frac{r^K}{p^{K/2}},
% \end{equation}
\begin{equation}\label{eq:cond2}
    \sup_{ \mU_k \in \mathbb{O}_{p, r}, k = 2,\ldots, K } \onorm{ \mE (\mU_{2} \otimes \cdots \otimes \mU_K)}_\sigma \lesssim \of{ \sqrt{r^{K-1}} + K\sqrt{pr}},
\end{equation}

\begin{equation}\label{eq:cond3}
    \sup_{ \mU_k \in \mathbb{O}_{p, r}, k = 2,\ldots, K } \onorm{ \mE (\mU_{2} \otimes \cdots \otimes \mU_K)}_F \lesssim \of{ \sqrt{pr^{K-1}} + K\sqrt{pr}},
\end{equation}

\begin{equation}\label{eq:cond_oracle}
    \xi \lesssim \exp\of{ - \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}},
\end{equation}

% \begin{equation}\label{eq:cond_intial}
%   \sum_{i: z^{(t)}(i) \neq z(i)} \theta(i)^2 \lesssim p^{-K/2 + 1} \Delta_{\min}^{-2}, \quad  L^{(t)} \lesssim \Delta_{\min}^2 / r, \quad \text{for} \quad t = 0, 1, \ldots, T.
% \end{equation}

\begin{equation}\label{eq:cond_intial}
    L^{(t)} \lesssim \frac{\Delta_{\min}^2}{r \log p}, \quad \text{for} \quad t = 0, 1, \ldots, T.
\end{equation}
% \fixme{Jiaxin}{See corollary 1}

 Particularly, the inequality~\eqref{eq:cond1} holds by replacing $\mV$ to $\mV^{(t)}$ and $\mW_{:a}$ to $\mW_{:a}^{(t),T}$ when initialization condition~\eqref{eq:cond_intial} holds.
\end{condition}

\paragraph{Step 2:} Next, we derive upper bound of $L^{(t+1)}$ for $t = 0 ,1, \ldots, T-1$. By Sub-Algorithm~\hyperref[alg:main]{2}, we update the assignment in $t$-th iteration as 
    \begin{equation}
        z^{(t+1)}(i) = \argmin_{a \in [r]} \onorm{ \off{ \mY_{i:} \mV^{(t)}  }^s - \off{\mS_{a:}^{(t)}}^s }^2,
    \end{equation}
    following the fact that $\mat_1(\tY^\text{d}) = \mY \mV^{(t)}$, where $\tY^\text{d}$ is the reduced tensor defined in Sub-Algorithm~\hyperref[alg:main]{2}. Then the event $z^{(t+1)}(i) = b$ implies
    \begin{equation}
        \onorm{ \off{ \mY_{j:} \mV^{(t)}  }^s - \off{\mS_{b:}^{(t)}}^s }^2 \leq \onorm{ \off{ \mY_{j:} \mV^{(t)}  }^s - \off{\mS_{z(j):}^{(t)}}^s }^2.
    \end{equation}
    Arranging the terms, we obtain the decomposition
    \begin{equation}\label{eq:decomp}
        2 \ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s } \leq \onorm{\mX_{i:} \mV^{(t)}} \of{ - \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 + G_{ib}^{(t)} + H_{ib}^{(t)} } +  F_{ib}^{(t)},
    \end{equation}
    where
    % \begin{equation}\label{eq:decomp}
    %     2 \ang{ \mE_{j:} \mV, \off{  \tilde \mS_{z(j):} }^s - \off{  \tilde \mS_{b:} }^s } \leq 2 \ang{ \mX_{j:} \mV^{(t)}, \off{  \mS_{b:}^{(t)}  }^s -\off{  \mS_{z(j):}^{(t)}  }^s  } + F_{jb}^{(t)},
    % \end{equation}
    \begin{align}
        F_{ib}^{(t)} &= 2 \ang{\mE_{i:} \mV^{(t)}, \of{ \off{  \tilde \mS_{z(i):} }^s  -  \off{  \mS_{z(i):}^{(t)} }^s  }  - \of{  \off{  \tilde \mS_{b:} }^s  -  \off{  \mS_{b:}^{(t)} }^s  }  } \\
        & \quad \quad \quad \quad \quad + 2 \ang{ \mE_{i:} \of{ \mV - \mV^{(t)} }, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s  },\\
        G_{ib}^{(t)} &=   \of{ \onorm{ \off{\mX_{i:} \mV^{(t)}}^s -  \off{  \mS_{z(i):}^{(t)}  }^s}^2 -  \onorm{ \off{\mX_{i:} \mV^{(t)}}^s -  \off{  \mW_{:z(i)}^T \mY \mV^{(t)} }^s}^2}   \\
        & \quad \quad \quad \quad \quad -   \of{ \onorm{ \off{\mX_{i:} \mV^{(t)}}^s -  \off{  \mS_{b:}^{(t)}  }^s}^2 -  \onorm{ \off{\mX_{i:} \mV^{(t)}}^s -  \off{  \mW_{:b}^T \mY \mV^{(t)} }^s}^2 }, \\
        H_{ib}^{(t)} &=   \onorm{ \off{\mX_{i:} \mV^{(t)}}^s -  \off{  \mW_{:z(i)}^T \mY \mV^{(t)} }^s}^2 - \onorm{ \off{\mX_{i:} \mV^{(t)}}^s -  \off{  \mW_{:b}^T \mY \mV^{(t)} }^s}^2  + \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 .
    \end{align}

  %\fixme{Miaoyan}{Add extra condition $\max_{a\in[R]}\onorm{\theta_{z^{-1}(a)}}\lesssim \sqrt{p}$ in order to guarantee $\onorm{\mX_{j:}\mV^{(t)}}<c$. Specifically, $\onorm{\mX_{j:}\mV^{(t)}} \leq \onorm{\mX_{j:}}\onorm{\mV^{(t)}}_{\sigma} \lesssim p^{-(K-1)/2}\max_{a\in[R]}\onorm{\theta_{z^{-1}(a)}}^{K-1} \lesssim c<\infty$. The second inequality is from Lemma 4 of Han et al. Add Lemma 4.}
    Then, we have the following upper bound 
    \begin{align}
         &\ind \offf{ z^{(t+1)}(j) = b  }\\ & \quad \quad \quad \leq \ind \offf{ z^{(t+1)}(j) = b ,   \ang{ \mE_{j:} \mV, \off{  \tilde \mS_{z(j):} }^s - \off{  \tilde \mS_{b:} }^s }  \leq - \frac{1}{4} \onorm{\mX_{i:} \mV^{(t)}} \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^2 } \\
         & \quad \quad \quad  \quad \quad + \ind \offf{z^{(t+1)}(j) = b, \frac{1}{2}\onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^2 \leq \onorm{\mX_{i:} \mV^{(t)}}^{-1} F_{jb}^{(t)} + G_{jb}^{(t)} + H_{jb}^{(t)} }. \label{eq:ind_b}
    \end{align}

   Note that 
    \begin{align}
        \onorm{\mX_{i:} \mV^{(t)}} &= \onorm{ \mX_{i:} \mW^{(t), \otimes^{K-1}}} \\
        &\geq \onorm{\mS_{i:} (\mTheta \mM)^{\otimes (K-1),T} } \lambda^{K-1}_r(\mW^{(t)}) \\
        & \geq \theta(i) \onorm{\mS_{z(i):}} \lambda_r^{K-1}(\mTheta \mM) \lambda^{K-1}_r(\mW^{(t)})\\
        & \geq\theta(i) m , \label{eq:xv_lower}
    \end{align}
    where the first two inequalities follow by the property of eigenvalues, the last inequality follows by Lemmas~\ref{lem:singular_thetam} and \ref{lem:membership}, the assumption that $\min_{a \in [r]} \onorm{\mS_{z(i):}} \geq c_3$, and $m >0$ is a positive constant related to $ c_3$. Plugging the lower bound of $ \onorm{\mX_{i:} \mV^{(t)}}$ \eqref{eq:xv_lower} into the inequality \ref{eq:ind_b}, we obtain that 
    \begin{equation}\label{eq:decomp1}
        \ind \offf{ z^{(t+1)}(i) = b  } \leq I_1 + I_2,
    \end{equation}
    where 
    \begin{align}
        A_{ib} &= \ind \offf{\ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s }  \leq - \frac{\theta(i) m}{4} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 },\\
        B_{ib} &= \ind \offf{z^{(t+1)}(i) = b, \frac{1}{2}\onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }.
    \end{align}
    On one hand, the event $A_{ib}$ is the critical component in $\xi$, where
    \begin{equation}
        \xi =   \frac{1}{p} \sum_{i \in [p] } \theta(i) \sum_{b \in [r]/z(i)}  \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 A_{ib}
    \end{equation}
    We will investigate $\bbP(A_{ib} = 1)$ when verifying the Condition~\ref{cond:origin}.

    On the other hand, the terms $ F_{ib}^{(t)}, G_{ib}^{(t)}$, and $H_{ib}^{(t)}$ in $B_{ib}$ describe the misclassification error in $t$-th iteration, which intuitively can be represented by $L^{(t)}$. Specifically, under Condition~\ref{cond:origin}, Lemma~\ref{lem:upper_fgh} and \citet[Step 4, Proof of Theorem 2]{han2020exact} imply that 
    %\fixme{Jiaxin}{Need to elaborate more. Step 4 is not completely the same as Han. Need to care about $\theta$}
    \begin{equation}
        \frac{1}{p} \sum_{i \in [p] } \theta(i) \sum_{b \in [r]/z(i)}  \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 B_{ib} \leq c' L^{(t+1)} + \rho L^{(t)},
    \end{equation}
    where $c', \rho \in (0,1)$ are two small positive constants. Hence, by the upper bound~\eqref{eq:decomp1} and the defintion of $L^{(t+1)}$, we obtain the upper bound for the misclassification loss in the $t+1$-th iteration
    \begin{equation}\label{eq:decomp_t}
        L^{(t+1)} \leq M \xi + \rho L^{(t)},
    \end{equation}
    where $M \geq 1$ and $\rho$ is called the contraction parameter.

    \paragraph{Step 3:} Last, we verify the Condition~\ref{cond:origin} under high probability to finish the proof. Note that the inequalities~\eqref{eq:cond1}, \eqref{eq:cond2}, and \eqref{eq:cond3} describes the property of the noise tensor $\tE$, and the readers can find the proof directly in \citet[Step 5, Proof of Theorem 2]{han2020exact}.

    Now, we verify the oracle loss condition~\eqref{eq:cond_oracle}. Recall the definition of $\xi$
    \begin{align}
        \xi = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \offf{  \ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s }  \leq - \frac{\theta(i) m}{4} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 }\\
        \cdot \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{ b:}  }^s  }^2,
    \end{align}

    Let $e_i = \mE_{i:} \mV$ denote the aggregated noise vector for $i \in [p]$, and $e_i$ are independent mean-zero sub-Gaussian vector in $\bbR^{r^{K-1}}$. Also, each coordinate in $e_i$ is also independent mean-zero sub-Gaussian variable with sub-Gaussian norm upper bounded by $C\sqrt{r^{K-1}/p^{K-1}}$ with some positive constant $C$.
    %\fixme{Miaoyan}{\color{red} explain $e_j$. See my comments at page 34. Please correct all similar issues throughout the appendix.} 
    Note that probability 
    \begin{align}
        \bbP \of{ \ang{ e_i, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s }  \leq - \frac{ m}{4} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2} \leq P_1 + P_2 + P_3,
    \end{align}
    where 
    \begin{align}
        P_1 &= \bbP \of{ \ang{e_i, \off{  \mS_{z(i):} }^s - \off{ \mS_{b:} }^s} \leq -\frac{m}{8}  \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 }, \\
        P_2 &= \bbP \of{ \ang{e_i, \off{ \tilde  \mS_{z(i):} }^s - \off{ \mS_{z(i):} }^s} \leq -\frac{m}{16}  \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 },\\
        P_3 &= \bbP \of{ \ang{e_i, \off{  \mS_{b:} }^s - \off{\tilde  \mS_{b:} }^s} \leq -\frac{m}{16}   \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 }.
    \end{align}

    For $P_1$, notice that the inner product $\ang{e_j,\mS_{z(j):}^s - \mS_{b:}^s} $ is a sub-Gaussian variable with sub-Gaussian norm bounded by $C \sqrt{ r^{K-1}/ p^{K-1} } \onorm{ \mS_{z(j):} ^s -  \mS_{b:}^s }$. Then, by Chernoff bound, we have  
    \begin{equation}\label{eq:p1_2}
         P_1 \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^2 }.
    \end{equation}

    For $P_2$ and $P_3$, we only need to derive the upper bound of $P_2$ due to the symmetry. By the law of total probability, we have 
    \begin{equation}\label{eq:p2}
        P_2 \leq P_{21} + P_{22},
    \end{equation}
    where with some positive constant $t$
    \begin{align}
        P_{21}& =  \bbP \of{ t <   \onorm{ \off{ \tilde  \mS_{z(i):} }^s - \off{ \mS_{z(i):} }^s }},\\
        P_{22} &= \bbP \of{ \left. \ang{e_i, \off{ \tilde  \mS_{z(i):} }^s - \off{ \mS_{z(i):} }^s} \leq -\frac{m}{16}  \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 \right| \onorm{ \off{ \tilde  \mS_{z(i):} }^s - \off{ \mS_{z(i):} }^s } < t  }.
    \end{align}

    For $P_{21}$, note that the term $\mW^T_{: z(i)} \mE \mV = \frac{ \sum_{ j \neq i, j \in [p]} \ind \{ z(j) = z(i) \} e_j }{\sum_{j \in [p]} \ind \{ z(j) = z(i) \} }$  is a sub-Gaussian vector with sub-Gaussian norm bounded by $C_1 \sqrt{r^K/ p^K}$. Then, we have 
    \begin{align}
        P_{21} &\leq \bbP\of{ t \onorm{\mS_{z(i):} } < \onorm{ \tilde  \mS_{z(i):}  -  \mS_{z(i):} }  }\\
        & \leq \bbP\of{c_3 t <   \onorm{ \mW^T_{: z(i)} \mE \mV  } } \\
        & \lesssim \exp\of{ - \frac{p^K t^2}{r^K} }, \label{eq:p21}
    \end{align}
    where the first inequality follows by the basic inequality in Lemma~\ref{lem:norm_diff}, the second inequality follows by the assumption that $ \min_{a \in [r]} \onorm{\mS_{z(i):}} \geq c_3$, and the last inequality follows by the Bernstein inequality.

    For $P_{22}$, the inner product $ \ang{e_i, \off{ \tilde  \mS_{z(i):} }^s - \off{ \mS_{z(i):} }^s}$ is also a sub-Gaussian variable with sub-Gaussian norm $ C \sqrt{ r^{K-1}/ p^{K-1} } t$, conditioned on the boundedness $\onorm{ \off{ \tilde  \mS_{z(j):} }^s - \off{ \mS_{z(j):} }^s } < t$. Then, by Chernoff bound, we have 
    \begin{equation}\label{eq:p22}
        P_{22} \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1} t^2}   \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^4 }. 
    \end{equation}
    Take $t =  \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }$ in $P_{21}$ and $P_{22}$, and plug the inequalities \eqref{eq:p21} and \eqref{eq:p22} into to the upper bound for $P_2$~\eqref{eq:p2}. We obtain that 
    \begin{equation}\label{eq:p2_2}
        P_2 \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^2 }.
    \end{equation}
    Therefore, combining the upper bounds~\eqref{eq:p1_2} and \eqref{eq:p2_2}, we have 
    \begin{equation}\label{eq:xi_p}
          \bbP \of{ \ang{ e_i, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s }  \leq - \frac{ m}{4} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2} \lesssim  \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^2 }.
    \end{equation}

    Hence, we have 
    \begin{align}
        \bbE[\xi] & =  \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \bbP \offf{  \ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s }  \leq - \frac{ m}{4} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 } \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{ b:}  }^s  }^2 \\
        &\lesssim \frac{1}{p} \sum_{i \in [p]}  \sqrt{p} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2   \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^2 } \\
        & \lesssim  \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 },
    \end{align}
    where the first inequality follows by the fact that $\max_{i \in [p]} \theta(i) \lesssim \sqrt{p}$.

    By Markov's inequality, we have 
    \begin{equation}
        \bbP\of{ \xi \lesssim \bbE[\xi] + \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } } \geq 1 -  C_2 \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 },
    \end{equation}
    and thus the condition~\eqref{eq:cond_oracle} holds with probability at least $1 -  C_2 \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }$ for some constant $C_2 > 0$.

    Finally, we verify the bounded loss condition~\eqref{eq:cond_intial} by induction. By Corollary~\ref{cor:L0}, we have 
    \begin{equation}
        L^{(0)} \lesssim \frac{\Delta_{\min}^2}{r \log p}
    \end{equation}
    when $p$ is large enough by Theorem~\ref{thm:initial}, and thus condition~\eqref{eq:cond_intial} holds for $t = 0$. Assume the condition~\eqref{eq:cond_intial} also holds for all $t \leq t_0$. Then, by the decomposition~\eqref{eq:decomp_t}, we have 
    \begin{align}
         L^{(t_0+1)} &\lesssim M \xi + \rho L^{(t_0)} \\
         & \lesssim  \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } + \frac{\Delta_{\min}^2}{r \log p} \\
         & \lesssim \frac{\Delta_{\min}^2}{r\log p} ,
    \end{align}
    where the second inequality follows by condition~\eqref{eq:cond_oracle} and the last inequality follows by the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2}$. Thus, the condition~\eqref{eq:cond_intial} holds for $t_0 + 1$, and the condition~\eqref{eq:cond_intial} is proved by induction.
\end{proof}

% \begin{lem}[Bernstein's type bound]\label{lem:bernstein}
%  Let $e_j = \mE_j \mV$. Let $\eta_j = \mW_{:z_j}^T \mE \mV$, i.e.,
%  \begin{align}
%      \eta_j = \frac{\sum_{l \in [p]} \mone\offf{ z_l = z_j} e_l }{\sum_{l \in [p]} \mone\offf{ z_l = z_j}}.
%  \end{align}
%  Then, we have
%  \begin{align}
%      \bbP\of{ \onorm{\eta_j}^2 \geq \frac{C r}{p} (r^{K-1} + 2 \sqrt{x r^{K-1}} + 2 x)  } \leq \exp \of{- \frac{x}{\mtheta_{\max}^K \alpha_2}}.
%  \end{align}
% \end{lem}

% \begin{proof}[Proof of Lemma~\ref{lem:bernstein}]
% Note that $\eta_j$ is still a mean-0 sub Gaussian random vector with independent entries where $\text{Var}( (\eta_j)_i ) \leq \frac{C r^K}{p^K} \mtheta_{\max}^K \alpha_2$. Then, we obtain the bound following the Bernstein inequality.
% \end{proof}

\paragraph{Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:refinement}}

\begin{lem}[Misclassification error and loss]\label{lem:mis} Define the misclustering error in the $t$-th iteration as $\ell^{(t)} = \ell(z^{(t)}, z)$. We have 
\begin{equation}
    \ell^{(t)} \lesssim \frac{1}{p}\sum_{i \in [p]} \theta(i) \ind\offf{ z^{(t)}(i) \neq z(i) } \leq \frac{L^{(t)}}{\Delta_{\min}^2}.
\end{equation}
%Define the misclustering error in the $t$-th iteration as $\ell^{(t)} = \ell(z^{(t)}, z)$, where $\ell(\cdot,\cdot)$ is the misclustering error in Section~\ref{sec:limits}. Then, we have $\ell^{(t)} \leq L^{(t)}/\Delta_{\min}^2$.
\end{lem}

\begin{lem}[Membership matrices]\label{lem:membership} Suppose the condition~\eqref{eq:cond_intial} holds. Then, for any $a \in [r]$, we have $|\of{z^{(t)}}^{-1}(a)| \asymp p/r$. Moreover, we have 
\begin{equation}\label{eq:membership1}
    \lambda_{r}(\mM) \asymp \onorm{\mM}_{\sigma} \asymp \sqrt{p/r}, \quad  \lambda_{r}(\mW) \asymp \onorm{\mW}_{\sigma} \asymp \sqrt{r/p}.
\end{equation}
The inequalities~\eqref{eq:membership1} also hold by replacing $\mM$ and $\mW$ to $\mM^{(t)}$ and $\mW^{(t)}$ respectively. 
Further, we have 
\begin{equation}\label{eq:WWsingular}
   \lambda_{\min}(\mW\mW^T) \asymp \onorm{\mW\mW^T}_{\sigma} \asymp r/p,
\end{equation}
which is also true for $\mW^{(t)}\mW^{(t),T}$.
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:membership}] See \citet[Proof of Lemma 4]{han2020exact} for inequality~\eqref{eq:membership1}.

For inequality~\eqref{eq:WWsingular}, note that 
\begin{align}
    \lambda(\mW\mW^T) = \sqrt{\text{eigen} (\mW \mW^T\mW\mW^T)} \asymp \sqrt{\frac{r}{p} \text{eigen} (\mW \mW^T)} = \sqrt{\frac{r}{p} \lambda^2(\mW)} \asymp \frac{r}{p},
\end{align}
where the first inequality follows the fact that $\mW^T\mW$ is a diagonal matrix with elements of order $\frac{r}{p}$, and the second equation follows by the definition of singular value.
\end{proof}

     \begin{lem}[Relationship between error and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin}, we have
    \begin{equation}\label{eq:inter1}
        \onormSize{}{\mV - \mV^{(t)}}_{\sigma} \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}}  \frac{r(K-1)}{\Delta_{\min}^2} L^{(t)},
    \end{equation}
    \begin{equation}\label{eq:inter2}
        \onormSize{}{\mE (\mV - \mV^{(t)})}_{\sigma} \lesssim \sqrt{ \frac{r^{K-1}(p r^{K-1} + p r)}{p^{K-1}}} 
 \frac{r(K-1)}{\Delta_{\min}^2} L^{(t)}.    
 \end{equation}
 \begin{equation}\label{eq:inter3}
     \max_{b \in [r]} \onormSize{}{ [ \mW_{:b}^T \mY \mV ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV  ]^s } \lesssim \frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{r^{2K} + pr^{K+1}}{p^{K}} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}
 \end{equation}
 \begin{equation}\label{eq:inter4}
     \max_{b \in [r]} \onormSize{}{  [  \mW_{:b}^{(t), T}  \mY \mV]^s - [  \mW_{:b}^{(t), T}  \mY \mV^{(t)}]^s } \lesssim  \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} +  \frac{r L^{(t)}}{\Delta_{\min}}
     \end{equation}
    \begin{equation}\label{eq:inter5}
         \max_{b \in [r]} \onormSize{}{ [\mW_{:b}^T \mY \mV^{(t)} ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s } \lesssim \frac{rL^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
    \end{equation}

    In addition, the inequality~\eqref{eq:inter4} also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$.
    \end{lem}

    \begin{proof}[Proof of Lemma~\ref{lem:intermediate}] We follow and use several intermediate conclusions in \citet[Proof of Lemma 5]{han2020exact}. We prove each inequality separately.
    \begin{enumerate}
    \item Inequality~\eqref{eq:inter1}. By \citet[Proof of Lemma 5]{han2020exact}, we have 
    \begin{equation}
         \onormSize{}{\mV - \mV^{(t)}}_{\sigma} \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}} r \ell^{(t)}.
    \end{equation}
    Then, we complete the proof of inequality~\eqref{eq:inter1} by applying Lemma~\ref{lem:mis} to the above inequality.
    \item Inequality~\eqref{eq:inter2}. By \citet[Proof of Lemma 5]{han2020exact}, we have 
    \begin{equation}
           \onormSize{}{\mE (\mV - \mV^{(t)})}_{\sigma} \lesssim \sqrt{ \frac{r^{K-1}(p r^{K-1} + p r)}{p^{K-1}}} r(K-1) \ell^{(t)}.
    \end{equation}
    Also, we complete the proof of inequality~\eqref{eq:inter1} by applying Lemma~\ref{lem:mis} to the above inequality.

        %\item Inequalities~\eqref{eq:inter1} and \eqref{eq:inter2}. We follow the \citet[Proof of Lemma 5]{han2020exact} except upper bounding their misclassification error ``$h_k^{(t)}$" by $L^{(t)}/\Delta_{\min}^2$ rather than ``$ l_k^{(t)}/\Delta_k^2$".
        %defined in their context. 
        %The technique that replacing ``$ l_k^{(t)}/\Delta_k^2$" by $L^{(t)}/\Delta_{\min}^2$ is also used in the proofs for other inequalities. 
    \item Inequality~\eqref{eq:inter3}. We upper bound the desired quantity by triangle inequality as
    \begin{equation}
        \onormSize{}{ [ \mW_{:b}^T \mY \mV ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV  ]^s } \leq I_1 + I_2 + I_3,
    \end{equation}
    where 
    \begin{align}
        I_1 &=  \onorm{ \frac{ \mW_{:b}^T \mY \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{ \mW_{:b}^{(t), T} \mY \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }}   },\\
        I_2 & = \onorm{\of{ \frac{1}{\onormSize{}{\mW_{:b}^T \mY \mV}} -  \frac{1}{\onormSize{}{\mW_{:b}^T \mX \mV}}  } \mW_{:b}^T \mY \mV },\\
        I_3 &= \onorm{\of{ \frac{1}{\onormSize{}{\mW_{:b}^{(t), T} \mY \mV  }} -  \frac{1}{\onormSize{}{\mW_{:b}^{(t), T} \mX \mV  }}  } \mW_{:b}^{(t), T} \mY \mV }.
    \end{align}
    Next, we upper bound the quantities $I_1, I_2, I_3$ separately.

    For $I_1$, we further bound $I_1$ by triangle inequality as
    %and Lemma~\ref{lem:norm_diff} as
    \begin{equation}
        I_1 \leq I_{11} + I_{12},
    \end{equation}
    where
    % \begin{equation}
    %     I_{11} = \onorm{ \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }}   } \quad \text{and} \quad I_{12} = \frac{1}{\onormSize{}{\mW_{:b}^T \mX \mV}}\onorm{ \of{\mW_{:b}^T - \mW_{:b}^{(t), T}} \mE \mV  }.
    % \end{equation}
     \begin{equation}
        I_{11} = \onorm{ \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }}   } \quad \text{and} \quad I_{12} = \onorm{ \frac{\mW_{:b}^T \mE \mV}{ \onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mE \mV}{ \onormSize{}{  \mW_{:b}^{(t),T} \mX \mV }} }
    \end{equation}
    Consider $I_{11}$. We first define the confusion matrix $\mD = \mM^T \mTheta^T \mW^{(t)} = \entry{D_{ab}} \in \bbR^{r \times r}$ where 
    \begin{equation}
        D_{ab} = \frac{\sum_{i \in [p]}  \theta(i) \ind \offf{ z(i) = a, z^{(t)}(i) = b } }{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }}, \text{  for all } a, b \in [r].
    \end{equation}
    By Lemma~\ref{lem:membership},  we have $\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } \gtrsim p/r$. Then, we have
    \begin{equation}\label{eq:dab_bound}
        \sum_{a \neq b, a,b \in [r]} D_{ab} \lesssim \frac{r}{p} \sum_{i \colon z^{(t)}(i) \neq z(i)} \theta(i) \lesssim\frac{L^{(t)}}{\Delta_{\min}^2}  \lesssim \frac{1}{\log p}, 
    \end{equation}
    and for all $b \in [r]$
    \begin{equation}\label{eq:dbb_bound}
        D_{bb} = \frac{\sum_{i \in [p]}  \theta(i) \ind \offf{ z(i) = z^{(t)}(i) = b } }{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }} \geq \frac{c_0(\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } - p\ell^{(t)})}{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }} \gtrsim 1 - \frac{1}{\log p},
    \end{equation}
    under the inequality \eqref{eq:cond_intial} in Condition~\ref{cond:origin}. By the definition of $\mW,\mW^{(t)},\mV$, we have 
    \begin{equation}
        \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} = \off{\mS_{b:}}^s, \quad \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }} = [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}]^s, 
    \end{equation}
    % \begin{equation}
    %     D_a = \frac{\sum_{i \in [p]}  \theta(i) \ind \offf{ z(i) = a, z^{(t)}(i) = b } }{\sum_{i \in [p]} \theta(i)  \ind\offf{  z^{(t)}(i) = b }}, \text{  for all } a \neq b, \quad D = \sum_{a \neq b, a \in [r]} D_a.
    % \end{equation}

    Let $\alpha$ denote the angle between $\mS_{b:}$ and $D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}$. To rough estimate the range of $\alpha$, we consider the inner product 
    \begin{align}
        \ang{\mS_{b:},  D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}}  &= D_{bb} \onorm{\mS_{b:}}^2 + \sum_{a \neq b} D_{ab}\ang{\mS_{b:}, \mS_{a:}}\\
        &\geq D_{bb} \onorm{\mS_{b:}}^2 -  \sum_{a \neq b, a \in [r]} D_{ab}   \onorm{\mS_{b:}} \max_{a \in [r]}  \onorm{\mS_{a:}} \\
        & \geq C,
    \end{align}
    where $C$ is a positive constant, and the last inequality holds when $p$ is large enough following the constraint of $\onorm{\mS_{b:}}$ in parameter space~\eqref{eq:family} and the bounds of $\mD$ \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}.

    The positive inner product between $\mS_{b:}$ and $D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}$ indicates $\alpha \in [0,  \pi/2)$, and thus $2\sin \frac{\alpha}{2} \leq \sqrt{2} \sin \alpha$. Then, by the geometry property of trigonometric function, we have 
    \begin{align}
       \onormSize{}{ [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}] \sin \alpha} &= \onormSize{}{(\mI_d - \text{Proj}(\mS_{b:})) \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}}\\
       & = \sum_{a \neq b, a \in [r]} D_{ab} \onorm{(\mI_d - \text{Proj}(\mS_{b:}))\mS_{a:}  }\\
       &= \sum_{a \neq b, a \in [r]} D_{ab} \onorm{ \mS_{a:} \sin (\mS_{b:}, \mS_{a:}) }\\
       & \leq \sum_{a \neq b, a \in [r]} D_{ab} \onorm{\mS_{a:}} \onorm{ \mS^s_{b:} - \mS^s_{a:}}, \label{eq:i11_sin_num}
    \end{align}
    where the last inequality follows by Lemma~\ref{lem:norm_diff}. Also, note that with bounds \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}, when $p$ is large enough we have 
    \begin{equation}\label{eq:i11_sin_dom}
           \onormSize{}{\mW_{:b}^{(t), T} \mX \mV } = \onormSize{}{ D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}}  \geq D_{bb} \onorm{\mS_{b:}} - \sum_{a \neq b, a \in [r]} D_{ab} \onorm{ \mS_{a:}} \geq C_1,
    \end{equation}
    for some positive constant $C_1$. Notice that $I_{11} = \sqrt{1 - \cos \alpha} = 2 \sin \frac{\alpha}{2}$. Therefore, we obtain
    \begin{align}
        I_{11} &\leq \sqrt{2} \sin \alpha \\
        & = \frac{  \onormSize{}{ [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}] \sin \alpha}}{  \onormSize{}{ D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}} } \\
        & \leq \frac{ 1}{C_1}  \sum_{a \neq b, a \in [r]} D_{ab} \onorm{\mS_{a:}} \onorm{ \mS^s_{b:} - \mS^s_{a:}} \\
        & \lesssim \frac{r}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind\offf{z^{(t)}(i) = b} \onorm{ \mS^s_{b:} - \mS^s_{a:}}\\
        & = \frac{r L^{(t)}}{\Delta_{\min}},\label{eq:i11}
    \end{align}
    where the second inequality follows by~\eqref{eq:i11_sin_num} and \eqref{eq:i11_sin_dom}, and the last inequality follows by the definition of $D_a$ and the constraint of $\onorm{\mS_{b:}}$ in parameter space~\eqref{eq:family}.

   % \fixme{Jiaxin}{Need to control $\frac{1}{p}\sum_{i \in [p]} \theta(i) \ind\{ z(i) \neq z^{(t)}(i)\} \leq \log^{-1} p$. Actually can be implied from $L^{(t)}$. Note that $\frac{1}{p} \sum \theta(i) \leq L^{(t)}/\Delta_{\min}^2$.}
    % Following the assumption in parameter space~\eqref{eq:family} that $c_3 \leq \onorm{\mS_{a:}} \leq c_4$ for all $a \in [r]$, we have 
    % \begin{equation}\label{eq:a11_norm}
    %     \onorm{(1 - \eta) \mS_{b:} + \sum_{a \neq b} \eta_a \mS_{a:}} \leq (1 - \eta) \onorm{\mS_{b:}} + \sum_{a \neq b, a\in [r]} \eta_a \onorm{\mS_{a:}} \leq \of{ 1 - \eta + \eta c_4 /c_3 } \onorm{\mS_{b:}},
    % \end{equation}
    % %\fixme{Jiaxin}{Definition of $\eta$ should include $\theta$. correct this in the following proof}
    % where the first inequality uses triangle inequality, and 
    % \begin{equation}\label{eq:a11_ang}
    %     \ang{ \mS_{b:}, \mS_{a:} } \geq - \onorm{\mS_{b:}} \onorm{\mS_{a:}} \geq -c_4/c_3 \onorm{\mS_{b:}}^2,
    % \end{equation}
    % where the first inequality uses the fact that $ \cos \of{\mS_{b:}, \mS_{a:} } \geq -1$. Let $\alpha \in [0, \pi]$ denote the angle between $\mS_{b:}$ and $ (1 - \eta) \mS_{b:} + \sum_{a \neq b} \eta_a \mS_{a:}$, we have 
    % \begin{align}
    %     \cos \alpha&= \frac{ \ang{ \mS_{b:},  (1 - \eta) \mS_{b:} + \sum_{a \neq b} \eta_a \mS_{a:}} }{ \onorm{ \mS_{b:} } \onorm{ (1 - \eta) \mS_{b:} + \sum_{a \neq b} \eta_a \mS_{a:} }}\\
    %     &\geq \frac{ (1- \eta) \onorm{\mS_{b:}}^2 + \sum_{a \neq b} \eta_a \ang{\mS_{b:}, \mS_{a:}} }{ \of{ 1 - \eta + \eta c_4 /c_3  } \onorm{\mS_{b:}}^2 }\\
    %     & \geq \frac{ (1- \eta - \eta c_4/c_3 )  \onorm{\mS_{b:}}^2  }{ \of{1 - \eta + \eta c_4 /c_3   } \onorm{\mS_{b:}}^2} \\
    %     &\geq 1 - 2 \frac{c_4}{c_3} \eta.\label{eq:cos_sb_sbmix}
    % \end{align}

    % Note that  $\sum_{i \in [p]} \ind\offf{  z^{(t)}(i) = b } \asymp \frac{p}{r}$ by Lemma~\ref{lem:membership}. Then 
    % \begin{equation}\label{eq:eta}
    %     \eta = \frac{ \sum_{z^{(t)}(i) = b, z(i) \neq b }\theta(i)}{\sum_{z^{(t)}(i) = b } \theta(i)}  \lesssim \frac{r}{c_0 p} \sum_{i: z^{(t)}(i) \neq z(i) } \theta(i) \lesssim \frac{1}{\log p}
    % \end{equation}
    % where the first inequality follows by the assumption that $\min_{i \in [p]}\theta(i) \geq c_0$, the second and third inequality follows by the inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin} and the assumption on $\Delta_{\min}^2$.

    % % \begin{align}
    % %     \eta &= \frac{ \sum_{z^{(t)}(i) = b, z(i) \neq b }\theta(i)  \sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b , z(i) \neq b} }{ \sum_{z^{(t)}(i) = b } \theta(i) \sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } } \\
    % %     & \leq \frac{ \sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b , z(i) \neq b} }{ \sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } } \\
    % %     &\lesssim \frac{r}{p} \sum_{i \in [p]} \ind \offf{ z(i) \neq z^{(t)}(i)} \lesssim \ell^{(t)}.
    % % \end{align}

    % % Also, by the bounded loss condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}, we have 
    % % \begin{equation}
    % %     \eta \lesssim \frac{L^{(t)}}{ \Delta_{\min}^2} = \frac{1}{r \log p} \leq \frac{c_3\sqrt{2}}{4c_4},
    % % \end{equation}
    % % where the last inequality holds when $p$ is large enough and thereof $\alpha \leq \frac{\pi}{4}$.

    % Then, when $p$ is large enough, we have $\eta \leq \frac{c_3\sqrt{2}}{4c_4}$ and $\alpha \leq \frac{\pi}{4}$. Note that by the property of trigonometric functions, we have
    % \begin{equation}\label{eq:trig}
    %      1- \cos \alpha = 2 \sin^2 \frac{\alpha}{2}, \quad \sin \frac{\alpha}{2} \leq \sin \alpha \leq \tan \alpha,
    % \end{equation}
    % and with triangle inequality
    % \begin{equation}\label{eq:sin}
    %     \onorm{ \off{(1 - \eta) \mS_{b:} + \sum_{a \neq b, a\in [r]} \eta_a \mS_{a:}} \sin \alpha } \leq \sum_{a \neq b, a\in [r]} \eta_a \onorm{\mS_{a:}} \sin \of{ \mS_{a:}, \mS_{b:} },  
    % \end{equation}
    % and 
    % \begin{equation}\label{eq:cos}
    %     \onorm{ \off{(1 - \eta) \mS_{b:} + \sum_{a \neq b, a\in [r]} \eta_a \mS_{a:}} \cos \alpha } \leq (1-\eta) \onorm{\mS_{b:}}.  
    % \end{equation}

    % Hence, we have the upper for $I_{11}$
    % \begin{align}
    %     I_{11} &\stackrel{\text{def }}{=}  2 \sin \frac{\alpha}{2}\\
    %     &\stackrel{\text{\eqref{eq:trig}}}{\leq}  2 \tan \alpha \\
    %     &\stackrel{\text{\eqref{eq:sin}, \eqref{eq:cos}}}{\leq} 2 \frac{ \sum_{a \neq b, a \in [r]} \eta_a \onorm{\mS_{a:}} \sin\of{ \mS_{a:}, \mS_{b:} } }{ (1- \eta) \onorm{\mS_{b:} } }\\
    %   & \stackrel{\text{\eqref{eq:family}, Lemma~\ref{lem:norm_diff}}}{\lesssim} \sum_{a \neq b, a \in [r]}\eta_a \onorm{ \off{\mS_{b:}}^s - \off{ \mS_{a;} }^s }\\
    %     &\stackrel{\text{ Lemma~\ref{lem:membership}}}{\lesssim} \frac{r}{p}  \sum_{i \in [p]} \theta(i) \sum_{a \neq b, a \in [r]} \ind \offf{ z(i) = b, z^{(i)} = a }  \onorm{ \off{\mS_{b:}}^s - \off{ \mS_{a:} }^s } \label{eq:i11_p} \\ 
    %     & \leq \frac{r L^{(t)}}{\Delta_{\min}}. \label{eq:i11}
    % \end{align}

    % Thus, we have $\sin \frac{\alpha}{2} \leq \sin \alpha \leq \tan \alpha$ by the property of trigonometric functions. We have 
    % \begin{align}
    %     I_{11}  & \leq 2 \tan \alpha\\
    %     & \leq 2 \frac{ \sum_{a \neq b, a \in [r]} \eta_a \onorm{\mS_{a:}} \sin\of{ \mS_{a:}, \mS_{b:} } }{ (1- \eta) \onorm{\mS_{b:} } }\\
    %     & \lesssim \sum_{a \neq b} \eta_a \onorm{ \off{\mS_{b:}}^s - \off{ \mS_{a;} }^s }\\
    %     & \lesssim \frac{r}{p} \theta(i) \sum_{i \in [p]} \sum_{a \neq b, a \in [r]} \ind \offf{ z(i) = b, z^{(i)} = a }  \onorm{ \off{\mS_{b:}}^s - \off{ \mS_{a:} }^s } \label{eq:i11_p} \\ 
    %     & \leq L^{(t)}/\Delta_{\min}, 
    % \end{align}
    % where the second inequality follows by the definition of $\tan(\cdot)$ and triangle inequality, the third inequality follows by the Lemma~\ref{lem:norm_diff}, the third inequality follows by the fact hat $\eta_a \leq \frac{ \sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b , z(i) = a} }{ \sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } } $, the last two inequalities follows by the definition of $L^{(t)}$.

    Consider $I_{12}$. By triangle inequality, we have 
    \begin{equation}
        I_{12} \leq \frac{1}{\onormSize{}{  \mW_{:b}^T \mX \mV }}\onormSize{}{ (\mW_{:b}^T - \mW_{:b}^{(t),T}) \mE \mV } + \frac{ \onormSize{}{(\mW_{:b}^T - \mW_{:b}^{(t),T}) \mX \mV} }{\onormSize{}{  \mW_{:b}^T \mX \mV } \onormSize{}{  \mW_{:b}^{(t),T} \mX \mV } }\onormSize{}{ \mW_{:b}^{(t),T} \mE \mV}.
    \end{equation}
    By \citet[Proof of Lemma 5]{han2020exact}, we have 
    \begin{equation}\label{eq:i12_wwe}
        \onormSize{}{ (\mW_{:b}^T - \mW_{:b}^{(t),T}) \mE \mV }  \lesssim \sqrt{\frac{ r^{2K} + p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}.
    \end{equation}
    Notice that 
    \begin{equation}\label{eq:i12_wwx}
        \onormSize{}{(\mW_{:b}^T - \mW_{:b}^{(t),T}) \mX \mV} \leq \onormSize{}{\mW_{:b}^T - \mW_{:b}^{(t),T}} \onorm{\mX\mV}_F \lesssim \frac{r^{3/2}L^{(t)}}{\sqrt{p}\Delta_{\min}^2} \onormSize{}{\mS} \onormSize{}{\mTheta \mM}_{\sigma} \lesssim \frac{ \sqrt{r L^{(t)}} }{\Delta_{\min}}, 
    \end{equation}
    where the second inequality follows by  \citet[Proof of Lemma 5]{han2020exact} and the last inequality follows by Lemma~\ref{lem:singular_thetam} and \eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
    % Also, by the definition of $\mW^{(t)}$ and Condition~\ref{cond:origin}, we have
    % \begin{equation}
    %     \onormSize{}{ \mW_{:b}^{(t),T} \mE \mV} \leq \onormSize{}{\mW_{:b}^{(t),T}} \onorm{\mE \mV}_{\sigma} \lesssim \sqrt{\frac{r^{K-1}}{p^K}} (\sqrt{p} + \sqrt{r^{K-1}}).
    % \end{equation}
     Note that $\onorm{\mW_{:b}^T \mX \mV} = \onorm{\mS_{b:}} \geq c_3$ and $\onormSize{}{  \mW_{:b}^{(t),T} \mX \mV } \geq C_1$ by inequality~\eqref{eq:i11_sin_dom}. Therefore, we have 
     \begin{align}
         I_{12} &\lesssim  \onormSize{}{ (\mW_{:b}^T - \mW_{:b}^{(t),T}) \mE \mV }  +   \onormSize{}{(\mW_{:b}^T - \mW_{:b}^{(t),T}) \mX \mV}   \onormSize{}{ \mW_{:b}^{(t),T} \mE \mV} \\
         &\lesssim \sqrt{\frac{ r^{2K} + p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} + \frac{ \sqrt{r L^{(t)}} }{\Delta_{\min}}\sqrt{\frac{r^{2K}}{p^K}} \\
         & \lesssim \sqrt{\frac{ r^{2K} + p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},\label{eq:i12}
     \end{align}
     where second inequality follows the inequalities~\eqref{eq:i12_wwe}, \eqref{eq:i12_wwx}, and \eqref{eq:cond1} in Condition~\ref{cond:origin}.

    Hence, combining inequalities~\eqref{eq:i11} and~\eqref{eq:i12},  we have 
    \begin{equation}\label{eq:i1}
        I_1 \lesssim \frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{\frac{ r^{2K} + p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}.
    \end{equation}
    % Therefore, we complete the proof of inequality~\eqref{eq:inter3}.

    % \begin{align}
    %      \aabs{ \frac{1}{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{1}{\onormSize{}{  \mW_{:b}^{(t),T} \mX \mV }} } \leq \frac{\onormSize{}{ \mW_{:b}^T \mX \mV} + \onormSize{}{ \mW_{:b}^{(t),T} \mX \mV} }{\onormSize{}{  \mW_{:b}^T \mX \mV }\onormSize{}{  \mW_{:b}^{(t),T} \mX \mV }}
    % \end{align}

    % Note that $\onorm{\mS_{b:}} = \onorm{\mW_{:b}^T \mX \mV}$. We have
    % \begin{equation}\label{eq:i12}
    %      I_{12} \lesssim  \onorm{ \of{\mW_{:b}^T - \mW_{:b}^{(t), T}} \mE \mV} \lesssim \sqrt{\frac{r^{2K} +p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
    % \end{equation}
    % where the first inequality follows by the constraint~\eqref{eq:family}, and the second inequality follows by \citet[Equation (121), Proof of Lemma 5]{han2020exact}.

    For $I_2$ and $I_3$, recall that $\onorm{\mW_{:b}^T \mX \mV} = \onorm{\mS_{b:}} \geq c_3$ and $\onormSize{}{  \mW_{:b}^{(t),T} \mX \mV } \geq C_1$ by inequality~\eqref{eq:i11_sin_dom}. By triangle inequality and \eqref{eq:cond1} in Condition~\ref{cond:origin}, we have 
    \begin{equation}\label{eq:i2}
         I_2 \leq \frac{ \onormSize{}{ \mW_{:b}^T \mE \mV } }{ \onormSize{}{\mW_{:b}^T \mX \mV}  } \lesssim  \onormSize{}{ \mW_{:b}^T \mE \mV }  \lesssim \frac{r^K}{p^{K/2}},
    \end{equation}
    and 
    \begin{equation}\label{eq:i3}
        I_3 \leq \frac{ \onormSize{}{ \mW_{:b}^{(t),T} \mE \mV } }{ \onormSize{}{\mW_{:b}^{(t),T} \mX \mV}  }  \lesssim \onormSize{}{ \mW_{:b}^{(t),T} \mE \mV } \lesssim \frac{r^K}{p^{K/2}}.
     \end{equation}

    % we have 
    % \begin{equation}\label{eq:i2}
    %     I_2 \leq \frac{ \onorm{ \mW_{:b}^T \mE \mV } }{ \onorm{\mW_{:b}^T \mX \mV}  } \lesssim  \onorm{ \mW_{:b}^T \mE \mV }  \lesssim \frac{r^K}{p^{K/2}},
    % \end{equation}
    % where the first inequality follows by triangle inequality, the second inequality follows~\eqref{eq:family}, and the last inequality follows by Condition~\ref{cond:origin}.

    % For $I_3$, note that 
    % \begin{align}
    %     \onorm{\mW_{:b}^{(t),T} \mX \mV} = \onorm{ (1 - \eta) \mS_{b:} + \sum_{a \neq b, a \in [r]} \eta_a \mS_{a:}} \geq (1 - 2\eta) c_3 \gtrsim 1,
    % \end{align}
    % where the second inequality holds by~\eqref{eq:eta} when $p$ is large enough. Then, we have 
    % \begin{equation}\label{eq:i3}
    %     I_3 \leq \frac{ \onorm{ \mW_{:b}^{(t),T} \mE \mV } }{ \onorm{\mW_{:b}^{(t),T} \mX \mV}  }  \lesssim \onorm{ \mW_{:b}^{(t),T} \mE \mV } \lesssim \frac{r^K}{p^{K/2}},
    % \end{equation}
    % where the last inequality follows by Lemma~\ref{lem:membership} and Condition~\ref{cond:origin}.

    Therefore, combining the inequalities~\eqref{eq:i1}, \eqref{eq:i2}, and \eqref{eq:i3}, we finish the proof of inequality~\eqref{eq:inter3}.

    \item Inequality~\eqref{eq:inter4}. Here we only show the detailed proof of inequality~\eqref{eq:inter4} with $\mW_{:b}^{(t)}$. The proof also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$, and we omit the repeated procedures.

    We upper bound the desired quantity by triangle inequality as
    \begin{equation}
         \onormSize{}{  [   \mW_{:b}^{(t), T}  \mY \mV]^s -  [  \mW_{:b}^{(t), T}  \mY \mV^{(t)}]^s }  \leq J_1 + J_2 + J_3,
    \end{equation}
    where 
    \begin{align}
        J_1 &= \onorm{ \frac{\mW_{:b}^{(t),T} \mY \mV }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mY \mV^{(t)} }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV^{(t)} }}  },\\
        J_2 & = \onorm{\of{ \frac{1}{\onormSize{}{\mW_{:b}^{(t),T} \mY \mV}} -  \frac{1}{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV}}  } \mW_{:b}^{(t),T} \mY \mV },\\
        J_3 &= \onorm{\of{ \frac{1}{\onormSize{}{\mW_{:b}^{(t),T} \mY \mV^{(t)}}} -  \frac{1}{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)}} }  } \mW_{:b}^{(t), T} \mY \mV^{(t)} }.
    \end{align}
    Next, we upper bound the quantities $J_1, J_2, J_3$ separately.

    For $J_1$, by triangle inequality and Lemma~\ref{lem:norm_diff}, we have 
    \begin{equation}
        J_1 \leq J_{11} + J_{12},
    \end{equation}
    where 
    \begin{equation}
        J_{11} =  \onorm{ \frac{\mW_{:b}^{(t),T} \mX \mV }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mX \mV^{(t)} }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV^{(t)} }}  }, \quad J_{12} =  \onorm{ \frac{\mW_{:b}^{(t),T}  \mE \mV}{ \onormSize{}{\mW_{:b}^{(t),T}  \mX \mV} } - \frac{\mW_{:b}^{(t),T}  \mE \mV^{(t)}}{ \onormSize{}{\mW_{:b}^{(t),T}  \mX \mV^{(t)}} }   }.
    \end{equation}

    % \fixme{Jiaxin}{Start $J_{11}$}

    Consider $J_{11}$.  Define the matrix $\mV^{k} \coloneqq \mW^{\otimes (k-1)} \otimes \mW^{(t), \otimes (K-k)}$ for $k = 2,\ldots, K-1$, and denote $\mV^1 = \mV^{(t)}, \mV^K = \mV$. Then, define the quantity
    \begin{equation}
        J_{11}^k = \onormSize{}{ [\mW_{:b}^{(t),T} \mX \mV^k ]^s - [\mW_{:b}^{(t),T} \mX \mV^{k+1} ]^s },
     \end{equation}
    for $k = 1,\ldots, K-1$. Let $\beta_k$ denote the angle between $\mW_{:b}^{(t),T} \mX \mV^k$ and $\mW_{:b}^{(t),T} \mX \mV^{k+1}$. With the same idea to prove $I_{11}$, we bound $J_{11}^k$ by considering the trigonometric function of $\beta_k$.

    To roughly estimate the range of $\beta_k$, we consider the inner product between $\mW_{:b}^{(t),T} \mX \mV^k$ and $\mW_{:b}^{(t),T} \mX \mV^{k+1}$. Before the specific derivation of the inner product, note that 
    \begin{equation}
        \mW_{:b}^{(t),T} \mX \mV^k = \mat_1(\tT_k), \quad  \mW_{:b}^{(t),T} \mX \mV^{k+1} = \mat_1(\tT_{k+1}),
    \end{equation}
    where 
    \begin{align}
        \tT_k &= \tX \times_1  \mW_{:b}^{(t),T} \times_2 \mW^T \times_3 \cdots \times_k \mW^T \times_{k+1} \mW^{(t),T} \times_{k+1} \cdots \times_K \mW^{(t),T}\\
        \tT_{k+1} &= \tX \times_1  \mW_{:b}^{(t),T} \times_2 \mW^T \times_3 \cdots \times_k \mW^T \times_{k+1} \mW^{T} \times_{k+1} \cdots \times_K \mW^{(t),T}. 
    \end{align}
    Recall the definition of confusion matrix $\mD = \mM^T \mTheta^T \mW^{(t)} = \entry{D_{ab}} \in \bbR^{r \times r}$.
    %, and let  $\mZ^k =  \mW^{\otimes (k-1)} \otimes \mW^{(t),\otimes (K-k-1)} \in \bbR^{ p^{K-2} \times r^{K-2} }$.  
    Then, we have 
    \begin{align}
         \ang{\mW_{:b}^{(t),T} \mX \mV^k , \mW_{:b}^{(t),T} \mX \mV^{k+1}} 
         %&= \ang{ \mat_1(\tT_k), \mat_1(\tT_{k+1}) } \\
         &  =\ang{ \mat_{k+1}(\tT_k), \mat_{k+1}(\tT_{k+1}) } \label{eq:j11_tensormat} \\
         %& = \ang{ \mW^{(t),T} \mX \mW_{:b}^{(t)} \otimes \mZ^k , \mW^T \mX \mW_{:b}^{(t)} \otimes \mZ^k} \\
         %& = \ang{ \mD^T \mS (\mTheta \mM)^{\otimes (K-1), T}\mW_{:b}^{(t)} \otimes \mZ^k,  \mS (\mTheta \mM)^{\otimes (K-1), T}\mW_{:b}^{(t)} \otimes \mZ^k   }\\
         & =  \ang{ \mD^T \mS \mZ^k ,  \mS \mZ^k   }\\
         & = \sum_{b \in [r]} \of{ D_{bb} \onormSize{}{\mS_{b:}\mZ^k}^2 + \sum_{a \neq b, a \in [r]} D_{ab} \ang{\mS_{a:} \mZ^k, \mS_{b:} \mZ^k}} \\
         & \gtrsim (1 - \log p^{-1}) \min_{a \in [r]}\onormSize{}{\mS_{a:}\mZ^k}^2 - \log p^{-1} \max_{a \in [r]} \onormSize{}{\mS_{a:}\mZ^k}^2, \label{eq:j11inner}
    \end{align}
    where $\mZ^k = \mD_{:b} \otimes \mI_r^{\otimes (k-1)} \otimes  \mD^{\otimes (K-k-1)}$, the equations follow by the tensor algebra and the definitions, and the last inequality follows by the bounds of $\mD$ \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}.

    Note that 
    \begin{equation}\label{eq:singular_Du}
        \onorm{\mD}_{\sigma} \leq \onorm{\mD}_{F} \leq  \sqrt{\sum_{b \in [r]} D_{bb}^2 + (\sum_{a \neq b, a ,b \in [r]} D_{ab} )^2} \lesssim \sqrt{r  + \log^2 p^{-1}} \lesssim 1,
    \end{equation}
    where the third inequality follows by the fact that for all $b \in [r]$ 
    \begin{equation}
        D_{bb} \lesssim \frac{r}{p} \sum_{i\colon z(i) = b} \theta(i) \lesssim 1.
    \end{equation}
    Also, we have 
    \begin{equation}\label{eq:singular_Dl}
        \lambda_r(\mD) \geq \lambda_r(\mW^{(t)}) \lambda_r(\mTheta \mM) \gtrsim 1,
    \end{equation}
    following the Lemma~\ref{lem:singular_thetam} and Lemma~\ref{lem:membership}. Then, for all $k \in [K]$, we have 
    \begin{equation}\label{eq:singular_z}
       1 \lesssim \onorm{\mD_{:b}} \lambda_r(\mD)^{K-k-1}  \leq \lambda_{r^{K-2}}(\mZ^{k}) \leq \onormSize{}{\mZ^k}_{\sigma} \leq \onorm{\mD_{:b}} \onorm{\mD}_{\sigma}^{K-k-1} \lesssim 1.
    \end{equation}
    Thus, we have bounds 
    \begin{equation}
         \max_{a \in [r]} \onormSize{}{\mS_{a:} \mZ^k} \leq \max_{a \in [r]} \onorm{\mS_{a:}} \onormSize{}{\mZ^k}_{\sigma} \lesssim 1, \quad \min_{a \in [r]} \onormSize{}{\mS_{a:}\mZ^k}  \geq \min_{a \in [r]} \onorm{\mS_{a:}} \lambda_{r^{K-2}}(\mZ^{k}) \gtrsim 1.
    \end{equation}
    Hence, when $p$ is large enough, the inner product~\eqref{eq:j11inner} is positive, which implies $\beta_k \in [0, \pi/2)$ and thus $2 \sin \frac{\beta_k}{2} \leq \sqrt{2} \sin \beta_k$.

    Next, we upper bound the angle $\sin \beta_k$. Note that
    \begin{align}
         \sin \beta_k &= \sin ({\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k} }, {\mD_{:b}^{T} \mS \mI_r^{\otimes k} \otimes \mD^{\otimes K-k-1} } )\\
         &\leq \sin \beta_{k1} + \sin \beta_{k2},
    \end{align}
    where
    \begin{align}
         \sin \beta_{k1} &= \sin ({\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k} }, {\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \tilde \mD \otimes \mD^{\otimes K-k-1} } ),\\
         \sin \beta_{k2} & = \sin ({\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \tilde \mD \otimes \mD^{\otimes K-k-1} } , {\mD_{:b}^{T} \mS \mI_r^{\otimes k} \otimes \mD^{\otimes K-k-1} }   ),
    \end{align}
    and $\tilde \mD$ is the normalized confusion matrix with entries $\tilde \mD_{ab} = \frac{\sum_{i \in [p]} \theta(i) \ind\{z^{(t)} = b, z(i) = a\}}{\sum_{i \in [p] }\theta(i) \ind\{z^{(t)} = b\}}$.

    On one hand, recall the definitions for any cluster assignment $\bar z$
    \begin{equation}
        \mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T,\quad \mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
    \end{equation}
    Then, we have $\text{diag}(\mp(z^{(t)})) \mD = \text{diag}(\mp_{\mtheta}(z^{(t)})) \tilde \mD$. By Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}, we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \lesssim \log^{-1}(p) $. Then, by the stability assumption, we have 
    \begin{equation}
         \sin (\mp(z^{(t)}), \mp_{\mtheta}(z^{(t)})) \lesssim \frac{L^{(t)}}{\Delta_{\min}}. 
    \end{equation}
    Note that $\sin(\ma,\mb) =\min_{c \in \bbR} \frac{ \onormSize{}{\ma - c\mb} }{\onormSize{}{\ma}}$ for vectors $\ma,\mb$ of same dimension, and let $c_0 = \argmin_{c \in \bbR} \frac{\onormSize{}{\mp(z^{(t)}) - c\mp_{\mtheta}(z^{(t)})}}{\onormSize{}{\mp(z^{(t)})}}$. We have
    \begin{align}
       \min_{c \in \bbR} \onormSize{}{ \mD - c \tilde \mD}_F &\leq  \onormSize{}{ \mI_r - c_0 \text{diag}(\mp(z^{(t)})) \text{diag}^{-1}(\mp_{\mtheta}(z^{(t)}))}_F \onormSize{}{\mD}_F\\
       &\lesssim \frac{ \onormSize{}{ \mp(z^{(t)}) - c_0\mp_{\mtheta}(z^{(t)})} }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  } \\
       & \lesssim \frac{\onormSize{}{ \mp(z^{(t)})} }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1 } \sin (\mp(z^{(t)}), \mp_{\mtheta}(z^{(t)}))\\
       & \lesssim \frac{L^{(t)}}{\Delta_{\min}},
    \end{align}
    where the last inequality follows Lemma~\ref{lem:membership} that $\onormSize{}{ \mp(z^{(t)})} \lesssim p$ and $\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  \gtrsim p$. By the geometry property of trigonometric function, we have
    \begin{align}
        \sin \beta_{k1} &= \min_{c \in \bbR} \frac{\onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes (\mD - c \tilde \mD) \otimes \mD^{\otimes K-k - 1} } }{ \onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k}} }\\
        &\leq \frac{ \onormSize{}{\mD_{:b}^{T} \mS} \onormSize{}{\mD - c_0 \tilde \mD}_{\sigma} \onormSize{}{\mD}_{\sigma}^{K-k-1} }{\onormSize{}{\mD_{:b}^{T} \mS} \lambda_r^{K-k}(\mD) }\\
        & \lesssim  \onormSize{}{\mD - c_0 \tilde \mD}_{F}\\
        & \lesssim \frac{L^{(t)}}{\Delta_{\min}},
    \end{align}
    where the second inequality follows by the singular property of $\mD$ in \eqref{eq:singular_Du} and \eqref{eq:singular_Dl}.

    On the other hand, let $\mC = \text{diag}(\{\onormSize{}{\mS_{a:}}\}_{a \in [r]})$. We have 
    \begin{align}
        \sin \beta_{k2} &\lesssim \frac{ \onorm{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes (\mI_r - \tilde \mD) \otimes \mD^{\otimes K-k-1} } }{\onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k} \otimes \mD^{\otimes K-k-1} }  }\\
        & \lesssim \frac{ \onormSize{}{(\mI_r - \tilde \mD^T) \mS \mZ^k}_F }{\onormSize{}{\mD_{:b}^{T} \mS} \lambda_r^{K-k-1}(\mD)} \\
        & \lesssim \onormSize{ }{(\mI_r - \tilde \mD^T) \mS \mC}_F \onormSize{}{\mC^{-1} \mZ^k}_{\sigma}\\
        & \lesssim \frac{r}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \onormSize{}{\mS_{b:}^s - \mS_{z(i):}^s} \\
        & \lesssim \frac{L^{(t)}}{\Delta_{\min}}, 
    \end{align}
    where the third inequality follows by the singular property of $\mD$ and the boundedness of $\mS$, and the fourth inequality follows by the definition of $\tilde \mD$, boundedness of $\mS$, lower bound of $\mtheta$, and the singular property of $\mZ^k$ in inequality~\eqref{eq:singular_z}.

    Therefore, we have shown the 
    \begin{equation}
        \sin \beta_k \leq \sin \beta_{k1} + \sin \beta_{k2} \lesssim \frac{L^{(t)}}{\Delta_{\min}}.
    \end{equation}

%     \fixme{Jiaxin}{Proof for $J_{11}$ starts here. Blue color is the latest proof.}

%     {\color{blue}
%     Consider $\mA \mx, \mB \mx$ where $\mB = \mD^T \mA$ and let $\alpha$ denote the angle between $\mA \mx, \mB \mx$. Specifically, $\mA = (\mV\mX)^T = \mI_r \mS^T$ and $\mB = \mD^T \mS^T$. Our goal is to upper bound $\sin \alpha$. Let $\tilde \mD$ is the normalized confusion matrix with $\tilde \mD_{ab} = \frac{\sum_{i \in [p]} \theta(i) \ind\{z^{(t)} = b, z(i) = a\}}{\sum_{i \in [p] }\theta(i) \ind\{z^{(t)} = b\}}$. Then we have 
%     \begin{equation}
%         \sin \alpha \leq \sin (\mA \mx, \tilde \mB\mx) +  \sin (\tilde \mB \mx,  \mB\mx),
%     \end{equation}
%     where $\tilde \mB  = \tilde \mD^T \mA$. For the first term, we have 
%     \begin{align}
%         \sin (\mA \mx, \tilde \mB\mx) \leq  \frac{ \onorm{\mA \mx - \tilde \mB \mx } }{\onorm{\mA \mx}} \lesssim \onorm{(\mI - \tilde \mD^T) \mS^s}_F \lesssim \frac{1}{c_0 p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \onorm{\mS_{b:}^s -\mS_{z(i):}^s} \ind\{z^{(t)}(i) = b\} = \frac{L^{(t)}}{\Delta_{\min}}.
%     \end{align}
%     For the second term, we have 
%     \begin{equation}
%         \sin (\tilde \mB \mx,  \mB\mx) \lesssim \onorm{\tilde \mB - \mB}_F \lesssim \max_{b \in [r]} \aabs{ 1 -  \frac{\sum_{i \in [p]} \theta(i)\ind \offf{ z^{(t)}(i) = b } }{\sum_{i \in [p]} \ind \offf{ z^{(t)}(i) = b } }} \lesssim \frac{L^{(t)}}{\Delta_{\min}^2}.
%     \end{equation}

% }

% \fixme{Miaoyan}{Merge the results of Lemma 14 into your proof. No need to write Lemma 14 as a stand-alone lemma.}
%     \begin{lem}[Angle Stability]\label{lem:stability} Assume the local linear stability of degree holds in the neighborhood $\tN(z,\log^{-1}p)$. Let $\tilde \mB$ be.... Then, we have
%     \[
%     \sin(\tilde \mB \mx, \mB \mx)\lesssim {L^{(t)}\over \Delta_{\min}}.
%     \]
%     \end{lem}

%     \begin{proof}[Proof of Lemma~\ref{lem:stability}]
%     %By assumption, there exists a constant $\eta=\eta(\bar z)\in\bbR$ ($\bbR_{+}$?) such that $
%   %\eta\mp_{\mtheta}(\bar z)= \mp(\bar z)$. Therefore, by the definition of angles, we have
%   % \[
%     %\sin(\tilde \mB \mx, \mB \mx)\leq \onormSize{}{\tilde \mB-\eta\mB}_F=0.
%     %\]

%     Note that 
%     \begin{equation}
%         \min_{c \in \bbR} \frac{ \onormSize{}{\mp(z^{(t)}) - c \mp_{\mtheta}(z^{(t)})}}{p} \asymp \sin (\mp(z^{(t)}) , \mp_{\mtheta}(z^{(t)})) \lesssim \varepsilon \Delta_{\min} \lesssim \frac{L^{(t)}}{\Delta_{\min}}.
%     \end{equation}

%     Note that $\text{diag}(\mp_{\mtheta}(\bar z)) \tilde \mD = \text{diag}(\mp(\bar z))  \mD$. Let $c_0 = \argmin_{c \in \bbR} \onormSize{}{\mp(z^{(t)}) - c \mp_{\mtheta}(z^{(t)})}$. We have
%     \begin{align}
%          \sin(\tilde \mB \mx, \mB \mx) & \leq  \frac{ \onormSize{}{\mD - c_0\tilde \mD}_F \onorm{\mA \mx} }{\lambda_r(\mD) \onorm{\mA \mx}}\\
%          &\leq  \frac{\onorm{\mI - c_0\text{diag}(\mp(\bar z)) \text{diag}(\mp_{\mtheta}(\bar z))^{-1}}_F \onorm{\mD}_F \onorm{\mA \mx}}{ \lambda_r(\mD) \onorm{\mA \mx} }\\
%          & \leq \onorm{ \text{diag} \of{ \frac{ \onormSize{}{\mtheta_{z^{(t), -1}(a)}}_1 - c_0 |z^{(t), -1}(a)| }{ \onormSize{}{\mtheta_{z^{(t), -1}(a)}}_1 }  }  }_F \\
%          & \leq \frac{1}{p}  \onormSize{}{\mp(z^{(t)}) - c_0 \mp_{\mtheta}(z^{(t)})}\\
%          & \leq \frac{1}{p} \sin (\mp(z^{(t)}) , \mp_{\mtheta}(z^{(t)}))\\
%          & \lesssim \frac{L^{(t)}}{\Delta_{\min}},
%     \end{align}
%     where the third inequality follows by the singular property of $\mD$ in inequalities~\eqref{eq:singular_Dl} and \eqref{eq:singular_Du}.
%     \end{proof}

    % \fixme{Jiaxin}{concatenation may work}

    % { \color{blue}

    % \begin{lem}[Concatenation] \label{lem:concatenation}
    % Consider two vectors $\mx = [\mx_1,\ldots, \mx_r]$ and $\mx' = [\mx'_1, \ldots, \mx'_r]$. Suppose $\onorm{\mx_a} \asymp \onorm{\mx'_a} \asymp 1$ for all $ a\in [r]$. Let $\alpha$ denote the angle between $\mx$ and $\mx'$, and $\alpha_a$ denote the angle between $\mx_a$ and $\mx'_a$. Then, we have 
    % \begin{equation}
    %     %\min_{a \in [r]} (1 - \cos \alpha_a)  \lesssim 
    %     \max_{a \in [r]} (1 - \cos \alpha_a) \lesssim  (1 - \cos\alpha ) \lesssim 1 - \min_{a \in [r]} \cos \alpha_a,
    % \end{equation}

    % \end{lem}

    % \begin{proof}[Proof of Lemma~\ref{lem:concatenation}]
    % For the first inequality, note that 
    % \begin{align}
    %     \cos \alpha = \frac{ \sum_{a \in [r]} \cos \alpha_a \onorm{\mx_a} \onorm{\mx'_a} }{ \sqrt{ \sum_{a \in [r]} \onorm{\mx_a}^2} \sqrt{\sum_{a \in [r]} \onorm{ \mx'_a}^2} } \leq \frac{ \sum_{a \in [r]} \cos \alpha_a \onorm{\mx_a} \onorm{\mx'_a}}{ \sum_{a \in [r]} \onorm{\mx_a}\onorm{ \mx'_a}  }.
    % \end{align}
    % Then, we have 
    % \begin{equation}
    %     1 - \cos \alpha \geq\sum_{a \in [r]}   \frac{ \onorm{\mx_a} \onorm{\mx'_a} }{\sum_{a \in [r]} \onorm{\mx_a}\onorm{ \mx'_a}   } (1 - \cos \alpha_b) \gtrsim \max_{a \in [r]} (1 - \cos \alpha_b),
    % \end{equation}
    % where the last inequality follows by the fact that $\onorm{\mx_a} \asymp \onorm{\mx'_a} \asymp 1$.

    % For the second inequality, note that 
    % \end{proof}

    % Consider the quantity
    % \begin{equation}
    %     J = \onorm{ \frac{\mW^{(t),T} \mX \mV^k}{\onormSize{}{\mW^{(t),T} \mX \mV^k}_F } - \frac{\mW^{(t),T} \mX \mV^{k+1}}{\onormSize{}{\mW^{(t),T} \mX \mV^{k+1}}_F } }_F.
    % \end{equation}
    % Let $\mx = [ \mx_1, \ldots, \mx_r]$ where $\mx_a = \mW_{:a}^{(t),T} \mX \mV^k$, and let $\mx' = [ \mx'_1, \ldots, \mx'_r]$ where $\mx'_a = \mW_{:a}^{(t),T} \mX \mV^{k+1}$. Also, let $\alpha$ denote the angle between $\mx$ and $\mx'$, and $\alpha_a$ denote the angle between $\mx_a, \mx'_a$. Note that $\onorm{\mx_a} \asymp \onorm{\mx'_a} \asymp 1$. Then, by Lemma~\ref{lem:concatenation}, we have 
    % \begin{equation}
    %     J_{11}^2 = (1 - \cos \alpha_b) \leq \max_{a \in [r]} (1 - \cos \alpha_a) \lesssim (1 - \cos \alpha) = J^2.
    % \end{equation}
    % Then, we only need to bound $J$. By matricization, we have 
    % \begin{equation}
    %     J = \onorm{ \frac{\mW^{T} \mX \mZ^k}{\onormSize{}{\mW^{T} \mX \mZ^k}_F } - \frac{\mW^{(t), T} \mX \mZ^k}{\onormSize{}{\mW^{(t), T} \mX \mZ^k}_F } }_F.
    % \end{equation}
    % Let $ \my = [\my_1,\ldots, \my_r]$ where $\my_a = \mW_{:a}^T \mX \mZ^k$ and $\my'= [\my'_1, \ldots, \my'_r]$ where $\my'_a = \mW_{:a}^{(t),T} \mX \mZ^k$. Let $\beta$ denote the angle between $\my, \my'$ and $\beta_a$ denote the angle between $\my_a, \my'_a$. Note that $\onorm{\my_a} \asymp \onorm{\my'_a} \asymp 1$. Also, note that for any $a \in [r]$
    % \begin{equation}
    %     1 - \cos \beta_a \lesssim (L^{(t)}/\Delta_{\min})^2,
    % \end{equation}
    % where the inequality may follow the procedures in $I_{11}$ and the padding lemma. Then, we have 
    % \begin{equation}
    %     (1 - \cos \alpha) = (1 - \cos \beta) \asymp \max_{b \in [r]} (1 - \cos \beta_a)  \lesssim (L^{(t)}/\Delta_{\min})^2.
    % \end{equation}
    % Thus we finish the proof.

    % }

%     \fixme{Jiaxin}{Trying to fix the above issue.}
%     Next, we derive the upper bound of $\sin \beta_k$. Recall the conclusion in the proof of $I_{11}$. For arbitrary $b \in [r]$ and let $\alpha$ the angle between $\mS_{b:}$ and $D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}$, we have 
%     \begin{equation}
%         \sin \alpha \lesssim \frac{r L^{(t)}}{\Delta_{\min}} \lesssim \frac{\Delta_{\min}}{\log p} \lesssim \frac{1}{\log p},
%     \end{equation}
%     under the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin} and the last inequality follows by the fact that $\Delta_{\min} = \tO(1)$. This indicates that when $p$ is large enough the cosine $\cos \alpha \geq \frac{1}{2}$. Then, by geometry, we have 
%     \begin{align}
%         \onormSize{}{ \mS_{b:} - [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}]} = \frac{  \onormSize{}{ [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}] \sin \alpha} }{ \cos \alpha}
%     \end{align}
% \fixme{Jiaxin}{This inequality seems not correct in geometry...}

%\fixme{Miaoyan}{The above bound has same form as (55). }

%   Also, let $\mD = \mM^T \mTheta^T \mW^{(t)} = \entry{D_{ab}} \in \bbR^{r \times r}$ be the confusion matrix, where 
%   \begin{equation}
%       D_{ab} = \frac{\sum_{i \in [p]} \theta(i) \ind \offf{ z(i) = a ,z^{(t)}(i) = b } }{\sum_{i \in [p]} \theta(i) \ind\offf{  z^{(t)}(i) = b} }.
%   \end{equation}
%   Since $\sum_{i \in [p]} \theta(i) \ind\offf{  z^{(t)}(i) = b}  \gtrsim p/r$ by Lemma~\ref{lem:membership} and the constraint $\min_{i \in [p]} \theta(i) \geq c_0$, we have 
%   \begin{equation}
%         \sum_{a \neq b} D_{ab} \lesssim \frac{r}{p} \sum_{i\colon z^{(t)}(i) \neq z(i)} \theta(i) \lesssim\frac{L^{(t)}}{\Delta_{\min}^2}  \lesssim \frac{1}{\log p}, \quad 
%   \end{equation}

    % Let $\mD = \mM^T \mTheta^T \mW^{(t)} \in \bbR^{r \times r}$ be the confusion matrix and $\mD_{ab} = \frac{\sum_{i \in [p]} \theta(i) \ind \offf{ z(i) = a ,z^{(t)}(i) = b } }{\sum_{i \in [p]} \theta(i) \ind\offf{  z^{(t)}(i) = b} }$. By Lemma~\ref{lem:membership}, we have $\sum_{i \in [p]} \ind\offf{ z^{(t)}(i) = a}  \asymp \frac{p}{r}$ for all $ a \in [r]$. Then we have
    % \begin{equation}\label{eq:d_property}
    %   \sum_{a \neq b} \mD_{ab} \asymp \frac{r}{p}  \sum_{a \in [r]} \sum_{b \neq a}  \ind \offf{ z(i) = a ,z^{(t)}(i) = b }  = \ell^{(t)}, \quad \text{and} \quad \sum_{a \in [r]} \mD_{aa} \asymp 1 -   \ell^{(t)}
    % \end{equation}  Let $\mZ^k =  \mW^{\otimes (k-1)} \otimes \mW^{(t),\otimes (K-k-1)} \in \bbR^{ p^{K-2} \times r^{K-2} }$, and by Lemma~\ref{lem:membership}
    % \begin{align}
    %     &\onorm{\mZ^k}_{\sigma} \lesssim  \onorm{\mW}^{(k-1)}_{\sigma} \onorm{\mW^{(t)}}_{\sigma}^{K-k-1} \lesssim \sqrt{\frac{r^{K-2}}{p^{K-2}}},\label{eq:z_singular1}\\
    %     &\lambda_{r^{K-2}} ( \mZ^k) \gtrsim   \lambda_{r}^{k-1} (\mW) \lambda_{r}^{K-k-1}(\mW^{(t)}) \gtrsim \sqrt{\frac{r^{K-2}}{p^{K-2}}}.\label{eq:z_singular2}
    % \end{align} We have
    % \begin{align}
    %     \ang{\mW_{:b}^{(t),T} \mX \mV^k , \mW_{:b}^{(t),T} \mX \mV^{k+1}} &= \ang{ \mW^{(t),T} \mX \mZ^k, \mW^T \mX \mZ^k} \\
    %     & = \ang{ \mD^T \mS (\mM^{\otimes K-1})^T \mW_{:b}^{(t)} \otimes \mZ^k, \mS (\mM^{\otimes K-1})^T \mW_{:b}^{(t)} \otimes \mZ^k } \\
    %     & = \sum_{a \in [r]}\of{ \mD_{aa} \onorm{\mS'_{a:}}^2 + \sum_{b \neq a}\mD_{ba} \ang{\mS_{a:}', \mS_{b:}'}} \\
    %     & \gtrsim (1 -   \ell^{(t)}) \min_{a \in [r]}  \onorm{\mS'_{a:}}^2  - \ell^{(t)} \max_{a \in [r]} \onorm{\mS'_{a:}}^2,
    % \end{align}
    % where $\mS'_{a:}$ refers to the $a$-th column of $\mS (\mM^{\otimes K-1})^T\mW_{:b}^{(t)} \otimes \mZ^k$ for $a \in [r]$, the first equation follows by tensor matricization, the second and third equations follows by the definition of $\mX$ and $\mD$, and the last inequality follows by the property~\eqref{eq:d_property}. By Lemma~\ref{lem:membership}, inequalities~\eqref{eq:z_singular1}, \eqref{eq:z_singular2}, and the constraint $c_3 \leq \min_{a \in [r]}\onorm{\mS_{a:}} \leq \max_{a \in [r]}\onorm{\mS_{a:}} \leq c_4 $ we have
    % \begin{align}
    %     \onorm{\mS'_{a:}} \leq \max_{a \in [r]}\onorm{\mS_{a:}} \onorm{\mM}_{\sigma}^{K-1} \onorm{\mW_{:b}^{(t)} \otimes}_F \onorm{\mZ^k}_{\sigma} \lesssim C_4 \\
    %     \onorm{\mS'_{a:}} \geq \min_{a \in [r]}\onorm{\mS_{a:}} \lambda_r^{K-1}(\mM) \onorm{\mW_{:b}^{(t)} \otimes}_F \lambda_{r^{K-2}} ( \mZ^k) \gtrsim C_3.
    % \end{align}
    % By the initialization condition that $\ell^{(t)} = o(\log^{-1} p)$, we have $\ang{\mW_{:b}^{(t),T} \mX \mV^k , \mW_{:b}^{(t),T} \mX \mV^{k+1}}  > 0$ when $p$ large enough. This implies the angle $\alpha_k \in [0,\pi/2]$, and the trigonometric functions satisfy $\sqrt{1 - \cos \alpha_k} = 2 \sin \frac{\alpha_k}{2} \lesssim \sin \alpha_k$ for the acute angle. 
    % \fixme{Jiaxin}{Below are previous proof.}
    % To upper bound $J_{11}^k$, we derive the upper bound for $\onorm{\mX \mV^k - \mX \mV^{k+1}}_F$. Note that 
    % \begin{align}\label{eq:a-b_f}
    %     \onorm{\mX \mV^k - \mX \mV^{k+1}}_F &= \onorm{ \mW^{(t),T} \mX( \mI\otimes \mZ^k) - \mW^T \mX( \mI\otimes \mZ^k) }_F\\
    %     & = \onorm{ \mS (\mM^{\otimes K-1})^T ( \mI\otimes \mZ^k)  - \mD \mS  (\mM^{\otimes K-1})^T ( \mI\otimes \mZ^k)  }_F \\
    %     & \leq \onorm{ \mC (\mI - \mD^T) \mS^s }_F \onorm{(\mM^{\otimes K-1})^T ( \mI\otimes \mZ^k)}_{\sigma}\\
    %     & \lesssim \onorm{ (\mI - \mD^T) \mS^s }_F \onorm{\mM}_{\sigma}^{K-1} \onorm{\mZ^k}^{K-1}\\
    %     & \lesssim \frac{r}{p} \sum_{a \in [r]} \sum_{b \neq a} \ind\offf{z(i) = a, z^{(t)}(i) = b} \onorm{ \mS^s_{a:} - \mS_{b:}^s}_F \sqrt{\frac{p}{k}}, 
    % \end{align}
    % \fixme{Jiaxin}{This inequality is not correct due to the definition of $\mD$.}
    % {\color{blue}
    % \begin{align}
    %     \onorm{(\mI -\mD^T)\mS^s}_F &\leq \sum_{b \in [r]}   \onormSize{}{ \mS_{b:}^s - [D_{bb} \mS_{b:}^s + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}^s]} \\
    %     & \leq \sum_{b \in [r]}   \onormSize{}{ \mS_{b:}^s - \sum_{a \in [r]} D_{ab} \mS_{b:}^s + \sum_{a \neq b, a \in [r]} D_{ab} (\mS_{b:}^s- \mS_{a:}^s)}  \\
    %     & \leq \sum_{b \in [r]} \of{ \onorm{(1 - \sum_{a \in [r]} D_{ab})\mS_{b:}^s }  + \sum_{a \neq b} D_{ab} \onorm{\mS_{b:}^s - \mS_{a:}^s}}\\
    %     & \lesssim \frac{L^{(t)}}{\Delta_{\min}^2} + \frac{L^{(t)}}{\Delta_{\min}}
    % \end{align}
    % }

   % \[
    %\onormSize{}{\mC(\mI-\mD^T)\mS^s}_F\leq \sum_b\onormSize{}{\mS^s_{b\colon}-[(\mI-\mD^T)\mS^s]_{b\colon}}_F
    %\]

    %\fixme{Miaoyan}{how about using spectral norm? When $\mS^s$ is identity matrix.}
   % \[
   % \onormSize{}{\mD^T\mD}_\sigma\gtrsim \min_{b}\left(D_{bb}-\sum_{a\in[r],a\neq  b}D_{ab}\right)^2\gtrsim 1-4\max_{b}\sum_{a\in[r],a\neq b}D_{ab}.
   % \]
  %\[
  %\onormSize{}{\mI-\mD^T}_\sigma \lesssim \max_{b}\sum_{a\in[r],a\neq b}D_{ab}.
  %\]
  %General $\mS^s$:
  %\[
  %\onormSize{}{(\mI-\mD^T)\mS^s}_\sigma
  %\]

    % where $\mC = \text{diag}(\onorm{\mS_{1:}}, \ldots, \onorm{\mS_{r:}})$ is a diagonal matrix with lower and upper bounded components, and $\mS^s \in \bbR^{r \times r^{K-1}}$ is the matrix with normalized rows $\mS_{a:}^s$, and the last inequality follows by the definition of $\mD$, triangle inequality and the singular property of $\mM$ and $\mZ^k$.

    % Also, note that 
    % \begin{align}\label{eq:wxv_bound}
    %     \onorm{\mW_{:b}^{(t),T} \mX \mV^{k+1}} &= \onorm{\mW_{:b}^{(t),T} \mM \mS (\mM^{\otimes K-1})^T \mV^{k+1} } \\
    %     &\geq \onorm{ \mW_{:b}^{(t),T} \mM \mS }_F \lambda_{r^{K-1}} \of{(\mM^{\otimes K-1})^T \mV^{k+1} } \\
    %     & \gtrsim \lambda_r^{K-1}(\mM)^T \lambda_r^{k+1}(\mW) \lambda_r^{K-k-2}(\mW^{(t)}) \\
    %     & \gtrsim c,
    % \end{align}
    % where $c$ is some positive constant, and the last inequality follows by the boundedness assumption on $\mS$ and the singular property in Lemma~\ref{lem:membership}.

    % For two vectors $\mv_1, \mv_2 \in \bbR^d$, let $\text{Proj}(\mv_1) \in \bbR^{d \times d}$ denote the projection matrix from $\bbR^d$ to the direction of $\mv_1$. We have the fact that 
    % \begin{align}\label{eq:proj_tri}
    %     \onorm{(\mI - \text{Proj}(\mv_1)) \mv_2} \leq  \onorm{\mv_1 - \mv_2}.
    % \end{align}

    % Hence, we have 
    % \begin{align}
    %     J_{11}^k  \stackrel{\sin \frac{\alpha}{2} \lesssim \sin \alpha, \text{ for } \alpha \in [0,\pi/2]}{\lesssim} &  \sin \alpha_k \\
    %     =  \quad \quad \quad \quad &\frac{ \onorm{(\mI - \text{Proj}(\mW_{:b}^{(t),T} \mX \mV^k )) \mW_{:b}^{(t),T} \mX \mV^{k+1}}  }{ \onorm{\mW_{:b}^{(t),T} \mX \mV^{k+1}} }\\
    %     \stackrel{\eqref{eq:proj_tri},\eqref{eq:wxv_bound}  }{\lesssim}\quad \quad \quad  &  \onorm{\mW_{:b}^{(t),T} \mX \mV^k  - \mW_{:b}^{(t),T} \mX \mV^{k+1} }\\
    %     \leq \quad \quad \quad \quad & \onorm{\mX \mV^k - \mX \mV^{k+1}}_F  \onorm{\mW_{:b}^{(t),T} }\\
    %     \stackrel{\eqref{eq:a-b_f}  }{\lesssim} \quad \quad \quad \quad &  \frac{r}{p} \sum_{a \in [r]} \sum_{b \neq a} \ind\offf{z(i) = a, z^{(t)}(i) = b} \onorm{ \mS^s_{a:} - \mS_{b:}^s}_F \\
    %   \leq  \quad \quad \quad \quad & \frac{r L^{(t)}}{\Delta_{\min}}.
    % \end{align}

    Finally, by triangle inequality, we obtain
    \begin{equation}\label{eq:j1}
        J_{11} \leq \sum_{k = 1}^{K-1} J_{11}^k \lesssim (K-1)\frac{r L^{(t)}}{\Delta_{\min}}.
    \end{equation}

    % \fixme{Jiaxin}{End $J_{11}$}

    Consider $J_{12}$. By triangle inequality, we have 
    \begin{align}
        J_{12} &\leq \frac{ 1}{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }} \onormSize{}{\mW_{:b}^{(t),T} \mE (\mV-\mV^{(t)})} \\ &\quad \quad \quad  +
        \frac{ \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})} }{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)} }} \onormSize{}{\mW_{:b}^{(t),T} \mE \mV^{(t)} }.
    \end{align}

     Note that 
    \begin{equation}\label{eq:j11_wxvt}
       \onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)}} = \onormSize{}{\mD^T \mS \mZ^1} \geq \lambda_r(\mD) \onorm{\mS} \lambda_{r^{K-2}}(\mZ^1) \gtrsim 1,
    \end{equation}
    where inequality follows by the bounds~\eqref{eq:singular_Dl} and \eqref{eq:singular_z}.

    By \citet[Proof of Lemma 5]{han2020exact}, we have 
    \begin{equation}\label{eq:j12_wev}
        \onormSize{}{\mW_{:b}^{(t),T} \mE (\mV-\mV^{(t)})} \lesssim \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{(K-1)\sqrt{L^{(t)}}}{\Delta_{\min}}.
    \end{equation}
    Notice that 
    \begin{align}
        \onormSize{}{\mX (\mV^k-\mV^{k+1})}_F &\leq \onormSize{}{(\mI - \mD^T) \mS (\mI_r^{\otimes (k-1)} \otimes  \mD^{\otimes (K-k-1)})}_F \\
        &\leq \onormSize{}{(\mW^T - \mW^{(t),T}) \mTheta \mM}_F \onorm{\mS}_F \onorm{ \mD}_{\sigma}^{K-k-1}\\
        & \lesssim \onormSize{}{ \mW^T - \mW^{(t),T} } \onorm{\mTheta \mM}_{\sigma} \\
        & \lesssim \frac{ \sqrt{r L^{(t)}}}{\Delta_{\min}},
    \end{align}
    where the first inequality follows by the tensor algebra in equation~\label{eq:j11_inner}, and the second inequality follows by the fact  that $\mI = \mW^T \mTheta \mM$, and the last inequality follows by \citet[Proof of Lemma 5]{han2020exact}. Thus, we have
    \begin{equation}
           \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})}\leq \onormSize{}{\mW_{:b}^{(t),T} }  \sum_{k = 1}^{K-1} \onormSize{}{\mX (\mV^k-\mV^{k+1})}_F \lesssim \frac{ (K-1) \sqrt{r L^{(t)}}}{\sqrt{p} \Delta_{\min}}.\label{eq:j12_wxv}
    \end{equation}
    Note that $\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }$ and $\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)} }$ are lower bounded by inequalities~\eqref{eq:i11_sin_dom} and \eqref{eq:j11_wxvt}, respectively. We have 
    \begin{align}
        J_{12} &\lesssim \onormSize{}{\mW_{:b}^{(t),T} \mE (\mV-\mV^{(t)})} + \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})}\onormSize{}{\mW_{:b}^{(t),T} \mE \mV^{(t)} }\\
        &\lesssim \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{(K-1)\sqrt{L^{(t)}}}{\Delta_{\min}} + \frac{ (K-1) \sqrt{r L^{(t)}}}{\sqrt{p} \Delta_{\min}} \sqrt{\frac{r^{2K}}{p^K}} \\
        & \lesssim (K-1) \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
    \end{align}
    where the second inequality follows by inequalities~\eqref{eq:j12_wev}, \eqref{eq:j12_wxv}, and the inequality \eqref{eq:cond1} in Condition~\ref{cond:origin}.

    % Consider $J_{12}$. Note that by Lemma~\ref{lem:membership}, $c_3 \leq \onorm{\mS_{b:}} = \onorm{\mW_{:b}^T \mX \mV}  $ and $\onorm{\mW_{:b}^{(t),T} \mX \mV}$ is bouneded. Also, by \citet[Inequality (123) (124) (125), Proof of Lemma 5]{han2020exact}, we have 
    % \begin{equation}\label{eq:j12}
    %     J_{12} \lesssim \onorm{\mW_{:b}^{(t),T}  \mE (\mV - \mV^{(t)})  } \lesssim \sqrt{ \frac{rr^{2K} + pr^{K+2}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}.
    % \end{equation}

    For $J_2$ and $J_3$, recall that $\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }$ and $\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)} }$ are lower bounded by inequalities~\eqref{eq:i11_sin_dom} and \eqref{eq:j11_wxvt}, respectively. By triangle inequality and inequality \eqref{eq:cond1} in Condition~\ref{cond:origin}, we have 
    \begin{equation}\label{eq:j2}
        J_2 \leq \frac{ \onormSize{}{\mW_{:b}^{(t),T} \mE \mV}  }{  \onormSize{}{\mW_{:b}^{(t),T} \mX \mV} } \lesssim \onormSize{}{\mW_{:b}^{(t),T} \mE \mV} \lesssim \frac{r^K}{p^{K/2}},
    \end{equation}
    and 
    \begin{equation}\label{eq:j3}
        J_3 \leq \frac{ \onormSize{}{\mW_{:b}^{(t),T} \mE \mV^{(t)}}  }{  \onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)}} } \lesssim \onormSize{}{\mW_{:b}^{(t),T} \mE \mV} \lesssim \frac{r^K}{p^{K/2}}.
    \end{equation}

    Therefore, combining the inequalities~\eqref{eq:j1}, \eqref{eq:j2}, and \eqref{eq:j3}, we finish the proof of inequality~\eqref{eq:inter4}.

    \item Inequality~\eqref{eq:inter5}. By triangle inequality, we upper bound the desired quantity 
    \begin{align}
        \onormSize{}{ [ \mW_{:b}^T \mY \mV^{(t)} ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s } &\leq \onormSize{}{[ \mW_{:b}^T \mY \mV^{(t)} ]^s - [ \mW_{:b}^T \mY \mV ]^s } \\
        & + \onormSize{}{[\mW_{:b}^T \mY \mV ]^s  - [ \mW_{:b}^{(t),T} \mY \mV ]^s } \\
        &+ \onormSize{}{[\mW_{:b}^{(t),T} \mY \mV ]^s  -  [\mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s }\\
        &\lesssim \frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
    \end{align}
    following the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4}. Therefore, we finish the proof of inequality~\eqref{eq:inter5}.
    % The proof follows the proof of inequality~\eqref{eq:inter3} except (1) the term $I_{11}$ now is upper bounded as 
    % \begin{equation}
    %     I_{11} \lesssim  \frac{r}{p} \sum_{i \in [p]} \sum_{a \neq b, a \in [r]} \ind \offf{ z(i) = b, z^{(i)} = a }  \onorm{ \off{\mW_{:b}^T \mX \mV^{(t)}}^s - \off{ \mW_{:a}^T \mX \mV^{(t)} }^s },
    % \end{equation}
    % and (2) the term $I_{12}$ has the upper bound in \citet[Inequality (127), Proof of Lemma 5]{han2020exact}. For $I_{11}$, following the procedures to derive $J_{11}$ \eqref{eq:inter4_newcenter} in the proof of inequality~\eqref{eq:inter4}, we also have 
    % \begin{equation}\label{eq:inter_wa_wb}
    %     \onorm{ \off{\mW_{:b}^T \mX \mV^{(t)}}^s - \off{ \mW_{:a}^T \mX \mV^{(t)} }^s } \lesssim  \onorm{ \off{\mS_{b:}}^s - \off{ \mS_{a:} }^s }.
    % \end{equation}
    % Therefore, we finish the proof of inequality~\eqref{eq:inter5}.
    \end{enumerate}

    \end{proof}

    \begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin}, we have 
    \begin{equation}\label{eq:fib}
        \max_{i \in [p]} \max_{b \neq z(i)} \frac{\of{F_{ib}^{(t)}}^2}{ \onormSize{}{ [  \mS_{z(i):} ]^s  -  [ \mS_{b:} ]^s }^2 } \lesssim  \frac{r L^{(t)}}{ \Delta_{\min}^2} \onormSize{}{\mE_{i:} \mV}^2 + \of{1 +  \frac{ r L^{(t)}}{ \Delta_{\min}^2}} \onormSize{}{\mE_{i:} (\mV - \mV^{(t)})}^2 ,
    \end{equation}
    \begin{equation}\label{eq:gib}
        \max_{i \in [p]} \max_{b \neq z(i)} \frac{ \of{  G_{ib}^{(t)} }^2  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2}  \lesssim \of{ \Delta_{\min}^2 +L^{(t)}},
    \end{equation}
    where $c$ is a small positive constant near to 0.
    \begin{equation}\label{eq:hib}
        \max_{i \in [p]} \max_{b \neq z(i)}  \frac{ \aabs{  H_{ib}^{(t)} }  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2} \leq \frac{1}{4}.
    \end{equation}
    \end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:upper_fgh}]
We prove Lemma~\ref{lem:upper_fgh} by the following order:
\begin{enumerate}
    \item Upper bound for $F_{ib}^{(t)}$~\eqref{eq:fib}. Recall the definition of $F_{ib}^{(t)}$ 
    \begin{align}
         F_{ib}^{(t)} &= 2 \ang{\mE_{i:} \mV^{(t)}, \of{ [ \tilde \mS_{z(i):} ]^s  - [\mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  - [ \mS_{b:}^{(t)} ]^s  }  } \\
        & \quad \quad \quad \quad \quad + 2 \ang{ \mE_{i:}(\mV - \mV^{(t)} ), [ \tilde \mS_{z(i):} ]^s - [ \tilde \mS_{b:} ]^s  }.
    \end{align}
    By Cauchy-Schwartz inequality, we have 
    \begin{align}
        \of{F_{ib}^{(t)}}^2 & \leq 8 \of{\ang{\mE_{i:} \mV^{(t)}, \of{ [  \tilde \mS_{z(i):} ]^s  -  [  \mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  -  [  \mS_{b:}^{(t)} ]^s  }  }  }^2 \\
         & \quad \quad \quad \quad \quad + 8\of{ \ang{ \mE_{i:} (\mV - \mV^{(t)} ), [  \tilde \mS_{z(i):} ]^s - [ \tilde  \mS_{b:} ]^s  }}^2 \\
         & \lesssim \of{ \onormSize{}{\mE_{i:} \mV}^2 + \onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2 } \max_{a \in [r]^s} \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  } \\
          & \quad \quad \quad \quad \quad +\onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2  \onormSize{}{  [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }.
    \end{align}
    Note that for any $a \in [r]$ 
    \begin{align}
        \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  }^2 &= \onormSize{}{ [ \mW_{:a}^T \mY \mV ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }^2 \\
        &\leq 2 \onormSize{}{ [ \mW_{:a}^T \mY \mV ]^s  - [ \mW_{:a}^{(t),T} \mY \mV ]^s  }^2  \\
        & \quad \quad \quad +2 \onormSize{}{[ \mW_{:a}^{(t),T} \mY \mV ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }^2\\
        & \lesssim \frac{ r^2 (L^{(t)})^2}{ \Delta_{\min}^2} + \frac{ r r^{2K} + p r^{K+2}}{p^K} \frac{L^{(t)}}{\Delta_{\min}^2}\\
        & \lesssim r L^{(t)}+  \frac{ r r^{2K} + p r^{K+2}}{p^K} \frac{L^{(t)}}{\Delta_{\min}^2}\\
        & \lesssim r L^{(t)},
    \end{align}
    where the second inequality follows the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4} in Lemma~\ref{lem:intermediate}, the third inequality follows the initialization condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}, and the last inequality follows the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.

    Also note that 
    \begin{align}
        \onormSize{}{  [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }^2 &=  \onormSize{}{  [  \tilde \mS_{z(i):} ]^s -[ \mS_{z(i):} ]^s +  [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s + [ \mS_{b:} ]^s -[  \tilde \mS_{b:} ]^s  }^2 \\
        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2 + \max_{a \in [r]} \onormSize{}{ [ \mS_{a:} ]^s -[  \tilde \mS_{a:} ]^s }^2 \\
        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2  + \max_{a \in [r]} \frac{1}{\onormSize{}{\mS_{a:} }^2} \onormSize{}{ \mW_{:a}^T \mE \mV}^2\\
        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2,
    \end{align}
    where the second inequality follows by Lemma~\ref{lem:norm_diff}, and the last inequality follows by the assumptions on $\onormSize{}{\mS_{a:}}$ in the parameter space~\eqref{eq:family}, the inequality~\eqref{eq:cond1} Condition~\ref{cond:origin} and the assumption $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.

    Therefore, we finish the proof of inequality~\eqref{eq:fib}.

    \item Upper bound for $G_{ib}^{(t)}$~\eqref{eq:gib}. Recall the definition of $G_{ib}^{(t)}$ and rearrange the terms
    \begin{align}
        G_{ib}^{(t)} &=   \of{ \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2}   \\
        & \quad \quad \quad \quad \quad -   \of{ \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{b:}^{(t)}  ]^s}^2 -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 } \\
        &= 2 \ang{  [\mX_{i:} \mV^{(t)}]^s , \of{[  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s } - \of{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s }}\\
        %& \leq 4 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [  \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }.
        &= G_1 + G_2 - G_3,
    \end{align}
    where 
    \begin{align}
        G_1 &= \onormSize{}{ [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 - \onormSize{}{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s }^2,\\
        G_2 &= 2 \ang{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s,   [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s},\\
        G_3 & = 2 \ang{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s,  [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s}.
    \end{align}

    For $G_1$, we have 
    \begin{align}
        |G_1|^2 &\leq \aabs{ \onormSize{}{ [ \mW_{:z(i)}^T \mY \mV^{(t)} ]^s - [ \mS_{z(i):}^{(t)}  ]^s}^2 - \onormSize{}{ [ \mW_{:b}^T \mY \mV^{(t)} ]^s -  [ \mS_{b:}^{(t)}  ]^s }^2 }^2 \\
        & \leq \max_{a \in [r]} \onormSize{}{ [ \mW_{:a}^T \mY \mV^{(t)} ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^4\\
        & \lesssim \frac{r^4}{\Delta_{\min}^4} (L^{(t)})^4 +  \frac{r^2 r^{4K} + p^2r^{2K+4}}{p^{2K}} \frac{(L^{(t)})^2}{\Delta_{\min}^4} \\
        & \leq c\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}}, \label{eq:g1}
    \end{align}
    where the third inequality follows by the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:membership} and the last inequality follows by the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.

    For $G_2$, noted that $[\mX_{i:} \mV^{(t)}]^s = [ \mW_{z(i):}^T\mX \mV^{(t)}]^s$, we have 
    \begin{align}
        |G_2|^2 &\leq 2\onormSize{}{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 \onormSize{}{ [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s }^2 \\
        & \leq \frac{2}{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2} \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
        & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} \of{ \frac{r^2}{\Delta_{\min}^2} (L^{(t)})^2 + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} }\\
        & \leq c \Delta_{\min}^2 L^{(t)},\label{eq:g2}
    \end{align}
    where the second inequality follows by Lemma~\ref{lem:norm_diff}, the third inequality follows by the bound~\eqref{eq:j11_wxvt}, inequality~\eqref{eq:cond2} in Condition~\ref{cond:origin}, inequality~\eqref{eq:inter5} in Lemma~\ref{lem:intermediate},  and the last inequality follows by the assumption $\Delta_{\min}^2 \geq p^{-K/2} \log p$.

    For $G_3$, note that by triangle inequality
    \begin{align}
        \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2 &\leq \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + 2\max_{a \in [r]}\onormSize{}{[\mW_{:a}^T\mX \mV^{(t)}]^s - [\mW_{:a}^T\mX \mV]^s }^2\\
        & \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2}, \label{eq:g3xv_wxv}
    \end{align}
    where the last inequality follows by the derivation of $J_{11}$ in the proof of Lemma~\ref{lem:intermediate}.

    % \begin{align}

    %     %& \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2},
    % \end{align}
   % where the last inequality follows by the inequality~\eqref{eq:inter4} in Lemma~\ref{lem:intermediate}. 
   Then we have 
    \begin{align}
        |G_3|^2 & \leq 2 \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2 \\
        & \leq 2 \of{ \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  +  \onormSize{}{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  } \\
        & \quad \cdot \max_{a \in [r]} \onorm{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
        & \lesssim \of{\onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} } \of{ \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2}}+ c \Delta_{\min}^2 L^{(t)} \\
        & \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 (\Delta_{\min}^2 + L^{(t)}) + c\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}},\label{eq:g3}
    \end{align}
    where the third inequality follows by the procedures to derive~\eqref{eq:g1} and \eqref{eq:g2}, the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:intermediate}, and the last inequality follows by the assumption $\Delta_{\min}^2 \geq p^{-K/2} \log p$.

    Combining the inequalities~\eqref{eq:g1}, \eqref{eq:g2}, and \eqref{eq:g3}, we finish the proof of inequality~\eqref{eq:gib}.

    % According to the inequality~\eqref{eq:inter_wa_wb} in the proof of Lemma~\ref{lem:intermediate}, we have
    % \begin{equation}
    %     \onorm{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  \lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2.
    % \end{equation}
    % Then, we have 
    % \begin{align}
    %     |G_3|^2 &\lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 \max_{a \in [r]} \onorm{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2 + |G_2|^2 \\
    %     &\lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 \of{ \frac{r^2}{\Delta_{\min}^2} \of{L^{(t)}}^2 + K  \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} } + |G_2|^2\\
    %     & \leq \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2  c \of{ \Delta_{\min}^2 + K L^{(t)} }+|G_2|^2, \label{eq:g3}
    % \end{align}
    % where the second inequality follows by Lemma~\ref{lem:intermediate}, and the last inequality follows by the assumption on $\Delta_{\min}^2$.

    % Therefore, combining inequalities~\eqref{eq:g1}, \eqref{eq:g2}, \eqref{eq:g3} and apply the inequality of arithmetic and geometric means, .

    \item Upper bound for $H_{ib}^{(t)}$~\eqref{eq:hib}. Recall the definition of $H_{ib}$ and rearrange the terms
    \begin{align}
        H_{ib} &= \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2\\
        & \quad \quad \quad \quad \quad + \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
        & =  \onormSize{}{ [ \mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 \\
        & \quad \quad \quad \quad \quad + \of{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }  } \\
        & \quad \quad \quad \quad \quad  - \of{\onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mY \mV^{(t)} ]^s } -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }   } \\
        & = H_1 + H_2 + H_3,
    \end{align}
    where 
    \begin{align}
        H_1 &= \onormSize{}{ [ \mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 - \onormSize{}{ [ \mW_{:b}^T \mX \mV^{(t)}]^s -  [ \mW_{:b}^T \mY \mV^{(t)}]^s}^2, \\
        H_2 &=  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }^2  , \\
        H_3 & = 2 \ang{  [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s,  [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }.
    \end{align}
    For $H_1$, we have 
    \begin{align}
        |H_1| \leq  \frac{4 \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 }{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2}
         \leq  \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} 
        \lesssim  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2\label{eq:h1},
    \end{align}
    following the derivation of $G_2$ in inequality~\eqref{eq:g2} and the assumption that $\Delta_{\min}^2 \geq p^{-K/2} \log p$.
    % \begin{align}
    %     |H_1| &\leq 2 \max_{a \in [r]}  \frac{1}{ \onormSize{}{  \mW_{:a}^T \mX \mV^{(t)}}^2} \onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2  \\
    %     & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}}\\
    %     & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K} \Delta_{\min}^2}  \onorm{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
    %     & \lesssim  \onorm{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 ,
    % \end{align}
    % where the first two inequalities follow the proof of $G_2$~\eqref{eq:g2} and the last inequality follows by the assumption on $\Delta_{\min}^2$.

    For $H_2$, by the inequality~\eqref{eq:g3xv_wxv}, we have 
    \begin{equation}\label{eq:h2}
        |H_2| \lesssim 2 \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
    \end{equation}
    where the last inequality follows by the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.

    % For $H_2$,  sccording to the inequality~\eqref{eq:inter_wa_wb} in the proof of Lemma~\ref{lem:intermediate}, we have
    % \begin{equation}
    %     \onorm{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  \lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
    % \end{equation}
    % and thus 
    % \begin{equation}\label{eq:h2}
    %     |H_2| \lesssim  \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2.
    % \end{equation}

    For $H_3$,  by Cauchy-Schwartz inequality, we have 
    \begin{equation}\label{eq:h3}
         |H_3| \lesssim \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
    \end{equation}
    following the inequalities~\eqref{eq:g3xv_wxv} and \eqref{eq:h1}.

    % For $H_3$, by Cauchy-Schwartz inequality, we have 
    % \begin{align}
    %     |H_3| &\lesssim    \onorm{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \\
    %     &\lesssim  \frac{ r^{K-1/2} + K p^{1/2} r^{K/2+1/2} }{p^{K/2}} \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }\\
    %     & \lesssim \Delta_{\min} \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }\\
    %     & \lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2, \label{eq:h3}
    % \end{align}
    % where the third inequality follows by the assumption on $\Delta_{\min}^2$.

    Therefore, combining inequalities~\eqref{eq:h1}, \eqref{eq:h2}, and \eqref{eq:h3}, we have finish the proof of inequality~\eqref{eq:hib}.

\end{enumerate}

\end{proof}

\end{document}

### 2022-01-02 tensor_clustering/tensor_clustering_ieee/ieee.tex
update

@@
-\documentclass[lettersize,journal]{IEEEtran}
+\documentclass[lettersize,onecolumn,journal]{IEEEtran}
 \usepackage{amsmath,amsfonts}
@@
 \usepackage{dsfont,multirow,hyperref,setspace,enumerate}
-\hypersetup{colorlinks,linkcolor=blue,citecolor=blue,urlcolor={black}}
+\hypersetup{colorlinks,linkcolor=black,citecolor=black,urlcolor={black}}
 \usepackage{lscape}
@@
 \usepackage{algpseudocode}
+
+\usepackage{tabularx,booktabs}
+\newcolumntype{Y}{>{\centering\arraybackslash}X}
 
@@
 
-{
-\color{red}
-
-\textbf{Other conference paper to journal}
-\begin{enumerate}
-    \item Garvesh. Add more discussion about Theorem 5 and empirical simulations.
-    %\item Eric Chi. Add one more remakr; One more section about the numerical issue; More comparison with other methods; add more future discussion.
-    \item Anru. Add application in introduction; more remarks for algorithm and theorems; One more section about the application to a specific model; more experiments and discussions.
-\end{enumerate}
-
-}
+% {
+% \color{red}
+
+% \textbf{Other conference paper to journal}
+% \begin{enumerate}
+%     \item Garvesh. Add more discussion about Theorem 5 and empirical simulations.
+%     %\item Eric Chi. Add one more remakr; One more section about the numerical issue; More comparison with other methods; add more future discussion.
+%     \item Anru. Add application in introduction; more remarks for algorithm and theorems; One more section about the application to a specific model; more experiments and discussions.
+
+%     \item Christina. Almost re-write the paper. 
+% \end{enumerate}
+% }
+
+% {
+% \color{blue}
+
+% \subsection{Our contribution.} 
+% \begin{itemize}
+%     \item We propose a generalized dTBM.
+%     \item Phase transition.
+%     \item We propose a multiway spherical clustering algorithm.
+% \end{itemize}
+
+% \subsection{Related work.} We emphasize the comparisons that set our work apart from earlier literature from two perspectives -- models and algorithms.
+
+% From the model perspective, our proposed model is related, but also distinct to following lines of works.
+% \begin{itemize}
+
+%     %\item \textit{Tensor block model.}
+
+%     \item \textit{Degree-corrected block model.}
+
+%     \item \textit{Degree-corrected block model for hypergraphs.}
+
+%     \item \textit{Other higher-order clustering models.}
+% \end{itemize}
+
+% From the algorithm perspective, our work is inspired by two key algorithmic techniques.
+
+% \begin{itemize}
+%     \item \textit{Spherical clustering.}
+
+%     \item \textit{Global-to-local strategy.}
+% \end{itemize}
+
+% \subsection{Binary multiway spherical clustering.}
+
+% \paragraph{Algorithm extension}
+
+% \paragraph{Theoretical guarantee}
+
+% }
 
@@
 
-\fixme{Jiaxin}{Change the wording: \emph{individual/block effects} $\rightarrow$ \emph{individual/block parameters}, \emph{degree heretogeneity} $\rightarrow$ \emph{non-equal degrees}.}
+%\fixme{Jiaxin}{Change the wording: \emph{individual/block effects} $\rightarrow$ \emph{individual/block parameters}, \emph{degree heretogeneity} $\rightarrow$ \emph{non-equal degrees}.}
 
@@
 Such degree heterogeneity appears commonly in social networks.
-For example, regular block model fails to model the Karate Club network~\citep{bickel2009nonparametric} without addressing
+{\color{blue} Ignoring the individual heterogeneity may seriously mislead the clustering results.} For example, regular block model fails to model {\color{blue} the member affiliation in} the Karate Club network~\citep{bickel2009nonparametric} without addressing
 %accounting for 
 degree heterogeneity. 
-\fixme{Jiaxin}{Non-degree model ignoring the individual heterogeneity may seriously mislead the clustering results.
+%\fixme{Jiaxin}{Non-degree model ignoring the individual heterogeneity may seriously mislead the clustering results.
 Karate Club data analysis shows that clustering without heterogeneity tends to assign membership based on the degree of communication (two clusters for leaders and non-leaders) rather than the affiliation (two clusters for people from different clubs).
@@
 The dTBM allows varying values within the same community, thereby allowing a richer structure.
-
-{
-\color{red}
-We use the idea of \emph{spherical clustering} in initialization and iteration. Using angle-based similarity avoids the estimation of $\mtheta$.  
+{\color{blue} To solve dTBM, we project clustering objects to a unit sphere and consider the angle similarity; detailed algorithms are in Section~\ref{sec:alg}.
+On one hand, the \textit{spherical} clustering avoids the estimation of nuisance degree heterogeneity.
+On the other hand, the usage of angle similarity brings new challenges to the derivations of theoretical results, and the polar coordinates based techniques are equipped in the proof.
 }
+
+% {
+% \color{red}
+% We use the idea of \emph{spherical clustering} in initialization and iteration. Using angle-based similarity avoids the estimation of $\mtheta$.  
+% }
 
@@
 
-{\bf Related work.} %Our work is closely related to, but also distinct from several lines of existing research. 
-We emphasize the comparisons that set our work apart from earlier literature from two perspectives---models and algorithms.
+{\bf Related work.} 
+{\color{blue}
+Our work is closely related to, but also distinct from several lines of existing research. 
+\begin{itemize}
+    \item \textit{Block model for clustering.} Block models such as stochastic block model (SBM) and degree-corrected SBM has been widely used for matrix clustering problems.
+The theoretical properties and algorithm performance for block models have been well-studied; see the review paper~\citep{abbe2017community} and the references therein.
+However, The tensor counterparts are relatively less understood.
+Table~\ref{tab:comp} summarizes the most relevant models.
+Specifically,
+    \begin{itemize}
+        \item the tensor block model (TBM, \cite{wang2019multiway, han2020exact}), a higher-order extension of SBM, fails to allow degree heterogeneity.
+The Cartesian coordinates based analysis in \cite{han2020exact} is also non-applicable to handle the extra flexibility brought from heterogeneity.
+In contrast, our model addresses the degree heterogeneity, and the polar coordinates based tools are adapted for the theoretical analysis;
+        \item the hypergraph degree-corrected block model (hDCBM) proposed by \cite{ke2019community, yuan2018testing} accounts for degree heterogeneity.
+Whereas, the hDCBM is designed only for binary observations, and the proposed spectral algorithm in \cite{ke2019community} achieves a sub-optimal clustering accuracy in higher-order scenarios.
+In constrast, our model allows discrete and continuous entries, and achieves exponentially fast rate in clustering tasks.
+More importantly, to our best knowledge, we are the first to provide the statistical and computational limits analyses for the degree-corrected block model in tensor clustering.
+    \end{itemize}
+
+    \item \textit{Global-to-local algorithm strategy.} Our methods generalize the recent global-to-local strategy for matrix learning~\citep{gao2018community,chi2019nonconvex,yun2016optimal} to tensors~\citep{han2020exact,ahn2018hypergraph,kim2018stochastic}.
+Despite the conceptual similarity, we address several fundamental challenges associated with this non-convex, non-continuous problem.
+We show the insufficiency of the conventional tensor HOSVD~\citep{kolda2009tensor}, and we develop a weighted higher-order initialization that relaxes the eigen-gap separation condition.
+Furthermore, our local iteration leverages the angle-based clustering in order to avoid explicit estimation of degree heterogeneity.
+Our bounds reveal the interesting interplay between the computational and statistical errors.
+We show that our final estimate provably achieves the exact clustering within only polynomial-time complexity.
+
+\end{itemize}
+}
+
+%We emphasize the comparisons that set our work apart from earlier literature from two perspectives---models and algorithms.
 From the model perspective, our work extends the previous degree-corrected model from matrices to tensors.
@@
 
-From the algorithm perspective, our methods generalize the recent global-to-local strategy for matrix learning~\citep{gao2018community,chi2019nonconvex,yun2016optimal} to tensors~\citep{han2020exact,ahn2018hypergraph,kim2018stochastic}.
+%From the algorithm perspective, our methods generalize the recent global-to-local strategy for matrix learning~\citep{gao2018community,chi2019nonconvex,yun2016optimal} to tensors~\citep{han2020exact,ahn2018hypergraph,kim2018stochastic}.
 Despite the conceptual similarity, we address several fundamental challenges associated with this non-convex, non-continuous problem.
@@
 
-{
-\color{red}
-
-Our work is closely related to, but also distinct from several lines of existing research:
-\begin{enumerate}
-
-\item \textbf{Degree-corrected model.}
-    \begin{itemize}
-        \item Matrix. (1) Previous work, survey paper (2) Unique technical challenges in Tensor (3) matrix does not show statistical-computational gap;
-        \item Tensor. (Comparison with Ke) (1) Ke does not provide the SNR limits (2) Ke's algorithm is sub-optimal (3) Ke focuses on Binary data. 
-    \end{itemize}
-
-    \item \textbf{Higher-order clustering.} (1) Previous work fails to account the heterogeneity, which is not realistic; (2) The extension of theoretical results highly challenging.
+% {
+% \color{red}
+
+% Our work is closely related to, but also distinct from several lines of existing research:
+% \begin{enumerate}
+
+% \item \textbf{Degree-corrected model.}
+%     \begin{itemize}
+%         \item Matrix. (1) Previous work, survey paper (2) Unique technical challenges in Tensor (3) matrix does not show statistical-computational gap;
+%         \item Tensor. (Comparison with Ke) (1) Ke does not provide the SNR limits (2) Ke's algorithm is sub-optimal (3) Ke focuses on Binary data. 
+%     \end{itemize}
+
+%     \item \textbf{Higher-order clustering.} (1) Previous work fails to account the heterogeneity, which is not realistic; (2) The extension of theoretical results highly challenging.
 Extra nuisance parameter makes it more difficult to establish the statistical and computational limits.
 
-    \item \textbf{Spherical clustering.}
-
-    \item \textbf{Global-to-local algorithm.} (1) Matrix global-to-local algorithm (2) Tensor global-to-local algorithm is used for non-degree algorithm.
+%     \item \textbf{Spherical clustering.}
+
+%     \item \textbf{Global-to-local algorithm.} (1) Matrix global-to-local algorithm (2) Tensor global-to-local algorithm is used for non-degree algorithm.
 Extra nuisance parameter makes it necessary to consider angle-based similarity; angle-based clustering makes the Cartesian coordinates based analysis non-applicable for our setting.
-\end{enumerate}
-
-}
+% \end{enumerate}
+
+% }
 
@@
 \resizebox{\textwidth}{!}{%
+\color{blue}
     \begin{tabular}{c|ccccc}
 \hline
-    & \cite{gao2018community}&  \cite{han2020exact}& \cite{ke2019community}&  \cite{ghoshdastidar2017consistency} & \textbf{Ours}\\
+    & \cite{gao2018community}&  \cite{han2020exact}&  \cite{ghoshdastidar2017consistency} &\cite{ke2019community} & \textbf{Ours}\\
     \hline
@@
         %Assortative assumption&  $\surd$ & $\times$ & $\times$ & $\times$\\
-       {\small Allow various data} & $\times$ & $\surd$ & $\times$ & $\surd$  &$\surd$\\
-        {\small Misclustering rate (for order 2$^*$)}& $\exp(-p)$ & $\exp(-p)$ & $p^{-1}$ & $p^{-1}$& $\exp(-p)$\\
+       %{\small Allow various data} & $\times$ & $\surd$ & $\times$ & $\surd$  &$\surd$\\
+        {\small Eigen free clustering} &  $\surd$ & $\surd$ & $\times$ & $\times$ &$\surd$ \\
+        %{\small Misclustering rate (for order 2$^*$)}& $\exp(-p)$ & $\exp(-p)$ & $p^{-1}$ & $p^{-1}$& $\exp(-p)$\\
+        {\small Misclustering rate (for order $K^*$)}& - & $\exp(-p^{K/2})$ & $p^{-1}$& $p^{-2}$ &  $\exp(-p^{K/2})$\\
         \hline
@@
     }
-    \caption{Comparison between previous methods with our method.
-$^*$We list here only the result for order-2 tensors (i.e., matrix).
-Our results allow general tensors of arbitrary order $K$; See Theorem~\ref{thm:refinement}.}\label{tab:comp}
+    \caption{Comparison between previous methods with our method. $^*$We list the result for order-K tensors with $K \geq 3$ and general number of communities $r = \tO(1)$.
+   % Our results allow general tensors of arbitrary order $K$; See Theorem~\ref{thm:refinement}.
+    }\label{tab:comp}
 \end{table*}
-
-{
-\color{red}
-Organization.
-}
 
@@
  %\fixme{Miaoyan}{\color{red} need to define $\cos(\ma,\mb)$. Be careful with signs.}
+
+{
+\color{blue}
+
+{\bf Organization.} The rest of this paper is organized as follows.
+Section~\ref{sec:model} introduces the degree-corrected tensor block model (dTBM) with three motivating examples and presents the identifiability of dTBM under the angle gap condition.
+We show the phase transition and the existence of statistical-computational gaps for the higher-order dTBM in Section~\ref{sec:limits}.
+In Section~\ref{sec:alg}, we provide a polynomial-time two-stage algorithm with misclustering rate guarantees.
+Numerical studies including the simulations to assess the theoretical results, comparison with other methods, and real data analysis on human brain connectome data and Peru legislation data are in Section~\ref{sec:simulation}.
+Last, we conclude our paper in Section~\ref{sec:con}.
+
+}
 
@@
 if $\tX$ admits dTBM~\eqref{eq:model_tensor}. Here, we give two special cases of dTBM.
-\begin{example}[Gaussian TBM]
-Let $\theta(i)= 1$ for all $i \in [p]$ and $\tE$ be a noise tensor with i.i.d.\ $N(0,\sigma^2)$ entries.
+
+{
+\color{blue}
+
+\subsection{Motivating Examples.} Here, we provide three applications to illustrate the practical necessity of dTBM.
+
+\vspace{0.2cm}
+\paragraph{Tensor block model} Consider the model~\eqref{eq:model_tensor}.
+Let $\theta(i)=1$ for all $ i \in [p]$.
+The model~\eqref{eq:model_tensor} reduces to the tensor block model, which is widely used in previous clustering algorithms~\citep{wang2019multiway,chi2020provable,han2020exact}.
+The theoretical results in TBM serve as benchmarks for dTBM.
+
+\vspace{0.2cm}
+
+\paragraph{Community detection in hypergraphs} Hypergraph network is a powerful tool to collect the complex entity relations with higher-order interactions~\citep{ke2019community}.
+A typical undirected hypergraph is denoted as $H = (V,E)$, where $V = [p]$ is the set of nodes and $E$ is the set of undirected hyperedges.
+Each hyperedge in $E$ is a subset of $V$, and we call the hyperedge an order-$K$ edge if the corresponding subset involves $K$ nodes.
+We call $H$ a $K$-uniform hypergraph if $E$ only contains order-$K$ edges.
+
+Similar with the adjacency matrix, it is natural to represent the $K$-uniform hypergraph by a binary order-$K$ adjacency tensor.
+Let $\tY \in \{0,1\}^{p \times \cdots \times p}$ denote the adjacency tensor, where the entries encode the presence or absence of hyperdeges among $p$ nodes.
+Specifically,
+\begin{equation}
+    \tY(i_1,...,i_K) =  \begin{cases}
+    1  & \text{if }  (i_1,...,i_K) \in E\\
+    0 & \text{if }  (i_1,...,i_K) \notin E
+    \end{cases},
+\end{equation}
+for all $(i_1,\ldots,i_K) \in [p]^K$.
+
+Assume there exist $r$ disjoint communities among $p$ nodes.
+The equation~\eqref{eq:model_tensor} models $\mathbb{E}\tY$ with unknown degree heterogeneity $\mtheta$ and subgaussianity parameter $\sigma^2 = 1/4$.
+
+\vspace{0.2cm}
+\paragraph{Multi-layer weighted network} 
+Multi-layer weighted network data consists of multiple networks with the same set of nodes.
+One representative example is the brain structural connectome data~\citep{zhang2019tensor}.
+The multi-layer weighted network $\tY$ has dimension of $p \times p \times L$, where $p$ denotes the number of brain regions in interest, and $L$ denotes the number of layer.
+Each of the $L$ networks describes one aspect of the brain connections, and the tensor $\tY$ include a mixture slices with continuous, binary, and count entries.
+
+% We use $G = (V, \{E_l\}_{l=1}^L)$ denote a $L$-layer undirected multi-layer network, where $E_l$ is the set of edges in the network of the $l$-th layer.
+Let $\mY_l \in \{0,1\}^{p\times p}$ denote the adjacency matrix for the $L$-th layer with nodes $V$ and edges $E_l$.
+Then, the multi-layer network is represented by an order-3 binary tensor $\tY \in \{0,1\}^{p \times p \times L}$, where
+% \begin{equation}
+%     \tY(i_1,i_2,l) = \mY_l(i_1,i_2),
+% \end{equation}
+% for all $(i_1,i_2) \in [p]^2$ and $l \in [L]$.
+
+Assume there exist $r$ disjoint communities among $p$ nodes and $r_l$ disjoint communities among the $L$ layers.
+The multi-layer network community detection is modeled by the generalized asymmetric dTBM model~\eqref{eq:model_tensor}
+\begin{equation}
+    \bbE[\tY] = \tS \times_1 \mTheta \mM \times_2 \mTheta \mM \times_3 \mTheta_l \mM_l, 
+\end{equation}
+where $(\mtheta \in \bbR^p, \mM \in \{0,1\}^{p \times r})$ and $ (\mtheta_l \in \bbR^L, \mM_l \in \{0,1\}^{L \times r_l})$ are the degree heterogeneity and membership matrices corresponding to the community structure for $p$ nodes and $L$ layers, respectively.
+\vspace{0.2cm}
+
+\paragraph{Gaussian higher-order clustering} Datasets in various fields such as medical image, genetics, and computer science are formulated as Gaussian tensors.
+One typical example is the multi-tissue gene expression dataset, which records the different gene expression in different individuals and different tissues.
+The dataset, denoted as $\tY \in \bbR^{p \times n \times t}$, consists of the expression data for $p$ genes of $n$ individuals in $t$ tissues.
+
+Assume there exist $r_1, r_2, r_3$ disjoint clusters for $p$ genes, $n$ individuals, and $t$ tissues, respectively. We apply the generalized asymmetric dTBM model~\eqref{eq:model_tensor} 
+\begin{equation}
+    \bbE[\tY] = \tS \times_1 \mTheta_1 \mM_1 \times_2 \mTheta_2 \mM_2 \times_3 \mTheta_3 \mM_3, 
+\end{equation}
+where $\{(\mtheta_k, \mM_k)\}_{k=1}^3$ refers to the heterogeneity and membership for genes, individuals, and tissues.
+
+}
+% \begin{example}[Gaussian TBM]
+% Let $\theta(i)= 1$ for all $i \in [p]$ and $\tE$ be a noise tensor with i.i.d.\ $N(0,\sigma^2)$ entries.
 Our dTBM reduces to a non-degree Gaussian TBM, which is widely used in previous clustering algorithms~\citep{wang2019multiway,chi2020provable,han2020exact}.
 The theoretical results in TBM serve as benchmarks for dTBM.
-\end{example}
-
-\begin{example}[Binary dTBM]
-Consider a $K$-uniform hypergraph $H = (V,E)$, where $V = [p]$ collects the nodes with $r$ disjoint communities and $E$ collects all the $K$-way hyperedges.
+% \end{example}
+
+% \begin{example}[Binary dTBM]
+% Consider a $K$-uniform hypergraph $H = (V,E)$, where $V = [p]$ collects the nodes with $r$ disjoint communities and $E$ collects all the $K$-way hyperedges.
 Let $\tY \in \{0,1\}^{p \times \cdots \times p}$ denote the adjacency tensor, where the entries encode the presence or absence of hyperdeges among $p$ nodes.
@@
 The equation~\eqref{eq:model_tensor} models $\mathbb{E}\tY$ with unknown degree heterogeneity and subgaussianity parameter $\sigma^2 = 1/4$.
-\end{example}
+% \end{example}
 
@@
 The goal of clustering is to estimate the partition function $z$ from model~\eqref{eq:model_tensor}. We use $\tP$ to denote the following parameter space for $(z,\tS,\mtheta)$,
-\footnotesize
 \begin{align}\label{eq:family}
-\tP&=\bigg\{(z,\tS,\mtheta)\colon  \ \mtheta\in\mathbb{R}^p_{+},\ 
-{c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r},\\& c_3\leq \onormSize{}{\text{Mat}(\tS)_{a:}}\leq c_4, \onorm{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|, a\in[r]\bigg\},
+\tP=\bigg\{(z,\tS,\mtheta)\colon  \ \mtheta\in\mathbb{R}^p_{+},\ 
+{c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r}, c_3\leq \onormSize{}{\text{Mat}(\tS)_{a:}}\leq c_4, \onorm{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|, a\in[r]\bigg\},
 \end{align}
@@
 %We briefly describe the rationale of the constraints in~\eqref{eq:family}. 
-First, the entrywise positivity constraint on  $\mtheta\in\mathbb{R}^p_{+}$ is imposed to avoid sign ambiguity between entries in $\mtheta_{z^{-1}(a)}$ and $\tS$.
+First, the entrywise positivity constraint on  $\mtheta\in\mathbb{R}^p_{+}$ is imposed to avoid sign ambiguity between entries in $\mtheta_{z^{-1}(a)}$ and $\tS$ {\color{blue} and thereof allow the trigonometric $\cos$ to describe the angle similarity in the following Assumption~\ref{assmp:min_gap} and Sub-algorithm~\hyperref[alg:main]{2} in Section~\ref{sec:alg}}.
 This constraint can be achieved without sacrificing model flexibility, by using a slightly larger dimension of $\tS$ in the factorization~\eqref{eq:model_tensor}; see Example~\ref{ex:positive}.
@@
 Here we provide an example to show the positivity constraints on $\mtheta$ incurs no loss on the model flexibility. 
-Without loss of generality, we consider the order-2 (i.e., matrix) block models.
-Consider a $4$-by-$4$ matrix with core matrix $\mS = 1$ and degree $\mtheta=(1,1,-1,-1)^T$.
-We have the mean matrix
-\begin{equation}
-    \mX = \mS \times_1 \mTheta \mM \times_2 \mTheta\mM = \begin{bmatrix}
-1 & 1 & -1& -1\\
-1 & 1 & -1& -1\\
--1 & -1 & 1& 1\\
--1 & -1 & 1& 1
+{\color{blue}
+We consider a order-3 dTBM with core tensor $\tS = 1$ and degree $\mtheta = (1,1,-1, -1)^T$. We have the mean tensor 
+\begin{equation}
+    \tX = \tS \times_1 \mTheta \mM \times_2 \mTheta \mM \times_3 \mTheta \mM,
+\end{equation}
+where $\mTheta = \text{diag}(\mtheta)$ and $\mM = (1,1,1,1)^T$.
+Note that $\tX \in \bbR^{4 \times 4 \times 4}$ is a 1-block tensor with mixed-signed degree $\mtheta$, and the mode-3 slices of $\tX$ are
+\begin{equation}
+    \tX_{::1} = \tX_{::2} = - \tX_{::3} = -\tX_{::4}  = \begin{bmatrix}
+    1 & 1& -1& -1 \\
+    1 & 1& -1& -1 \\
+    -1& -1&1 & 1\\
+     -1& -1&1 & 1
+    \end{bmatrix}.
+\end{equation}
+Now, instead of original decomposition, we encode $\tS$ as a 2-block tensor with positive-signed degree
+\begin{equation}
+     \tX = \tS' \times_1 \mTheta' \mM' \times_2 \mTheta' \mM'  \times_3 \mTheta' \mM' , 
+\end{equation}
+where $\tS' \in \bbR^{2 \times 2 \times 2}$ have mode-3 slices 
+\begin{equation}
+    \tS'_{::1} = -\tS'_{::2} = \begin{bmatrix}
+    1 & -1 \\
+    -1 & 1
+    \end{bmatrix}, \quad \mM' = \begin{bmatrix}
+    \mI_2 & {\bf 0}\\
+    {\bf 0} &  \mI_2
     \end{bmatrix},
 \end{equation}
-$ \text{with}\quad \mTheta=\text{diag}(\mtheta)$. Note that $\mX$ is a 1-block matrix with mixed-signed degree $\mtheta$. 
-Now, instead of original decomposition, we encode $\mX$ as a $2$-block matrix with positive-signed degree,
-\begin{equation}
-\mX= \mS' \times_1 \mTheta' \mM \times_2 \mTheta' \mM,\quad \text{with}\quad   \mS' = \begin{bmatrix} 1 & -1 \\
-    -1 & 1
-    \end{bmatrix}
-\end{equation}
-and $ \mTheta'=\text{diag}(\mtheta')=\text{diag}(1, 1, 1, 1)^T$, where the triplet $(z',\mS', \mtheta')$ lies in our parameter space~\eqref{eq:family}.
+and $\mTheta' = \text{diag}(\mtheta') = \text{diag}(1,1, 1, 1)$.
+The triplet $(z',\tS', \mtheta')$ lies in our parameter space~\eqref{eq:family}.
+In general, we can always reparameterize a block-$r$ tensor with mixed-signed degree using a block-$2r$ tensor with positive-signed degree.
+Since we assumes $r=\tO(1)$ throughout the paper, the splitting does not affect the error rates of our interest.
+}
+
+% Without loss of generality, we consider the order-2 (i.e., matrix) block models. Consider a $4$-by-$4$ matrix with core matrix $\mS = 1$ and degree $\mtheta=(1,1,-1,-1)^T$. We have the mean matrix 
+% \begin{equation}
+%     \mX = \mS \times_1 \mTheta \mM \times_2 \mTheta\mM = \begin{bmatrix}
+% 1 & 1 & -1& -1\\
+% 1 & 1 & -1& -1\\
+% -1 & -1 & 1& 1\\
+% -1 & -1 & 1& 1
+% \end{bmatrix},
+% \end{equation}
+% $ \text{with}\quad \mTheta=\text{diag}(\mtheta)$. Note that $\mX$ is a 1-block matrix with mixed-signed degree $\mtheta$. 
+% Now, instead of original decomposition, we encode $\mX$ as a $2$-block matrix with positive-signed degree,
+% \begin{equation}
+% \mX= \mS' \times_1 \mTheta' \mM \times_2 \mTheta' \mM,\quad \text{with}\quad   \mS' = \begin{bmatrix} 1 & -1 \\
+%     -1 & 1
+%     \end{bmatrix}
+% \end{equation}
+%and $ \mTheta'=\text{diag}(\mtheta')=\text{diag}(1, 1, 1, 1)^T$, where the triplet $(z',\mS', \mtheta')$ lies in our parameter space~\eqref{eq:family}.
 In general, we can always reparameterize a block-$r$ tensor with mixed-signed degree using a block-$2r$ tensor with positive-signed degree.
@@
 
-In Sections~\ref{sec:statlimit}-\ref{sec:complimit}, we provide the lower bounds of $\ell (\hat z, z)$ for general Gaussian dTBMs without symmetric assumptions.
+In Sections~\ref{sec:statlimit} and \ref{sec:complimit}, we provide the lower bounds of $\ell (\hat z, z)$ for general Gaussian dTBMs without symmetric assumptions.
 For general (asymmetric) dTBMs, we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode.
@@
 Theorem~\ref{thm:stats} demonstrates the impossibility of exact recovery of the assignment when $\gamma < -(K-1)$ in the high-dimensional regime $p\to \infty$ for fixed $r$.
-The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to, maximum likelihood estimation (MLE)~\citep{wang2019multiway,ke2019community} and trace maximization~\citep{ghoshdastidar2017uniform}.
+The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to, maximum likelihood estimation (MLE)~\citep{wang2019multiway} and trace maximization~\citep{ghoshdastidar2017uniform}.
 Our derived SNR threshold $-(K-1)$ is also a minimax upper bound, because MLE achieves exact recovery when $\gamma > -(K-1)$.
@@
 The global step finds the basin of attraction with polynomial miclustering error, whereas the local iterations improve the initial clustering to exact recovery.
-Both steps are critical to obtain a satisfactory algorithm output.
+Both steps are critical to obtain a satisfactory algorithm output.{\color{blue} In this section, we first use the symmetric tensor as a working example to describe the algorithm procedures to gain insight.
+Our theoretical analysis focuses on the noisy tensor with i.i.d.\ sub-Gaussian noise such as Gaussian and uniform observations.
+The extensions for asymmetric tensor and Bernoulli observation and other practical issues are in subsection~\ref{subsec:exten}.}
 
@@
 \centering
-\includegraphics[width=\columnwidth]{alg_demo.pdf}
+\includegraphics[width=\columnwidth]{alg_demo1.pdf}
 \caption{Illustration of our global-to-local algorithm.}\label{fig:demo}
@@
 We start with weighted higher-order clustering algorithm as initialization.
-To gain insights, we use an order-3 symmetric tensor as a working example.
-Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
+{\color{blue} We take the order-3 tensor in illustration for simplicity.}  Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
 By model~\eqref{eq:model_tensor}, for all $i \in [p]$, we have
@@
 The denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) estimates $\tX$ from $\tY$ by a double projection spectral method.
+{\color{blue}
+The first projection performs HOSVD~\citep{de2000multilinear} via $\mU_{\text{pre}} = \text{SVD}_{r} \of{ \text{Mat}(\tY) }$, where $\text{SVD}_r(\cdot)$ returns the top-$r$ left singular vectors.
+The second projection performs HOSVD on the projected $\tY$ onto the multlilinear Kronecker space $\mU_{\text{pre}} \otimes  \mU_{\text{pre}}$; i.e.,
+\begin{equation}\label{eq:two-step_factor}
+    \hat \mU = \text{SVD}_{r} \of{\text{Mat}\of{ \tY\times_1  \mU_{\text{pre}}^T \times_2  \mU_{\text{pre}}^T }}.
+\end{equation}
+The final denoised tensor $\hat \tX$ is defined by
+\begin{equation}\label{eq:two-step_est}
+    \hat \tX = \tY \times_1 \hat \mU \hat 
+\mU^T \times_2 \hat \mU \hat \mU^T \times_3 \hat \mU \hat \mU^T. 
+\end{equation}
+%Compared with conventional HOSVD, our double projection spectral method is shown to improve the estimation substantially for higher-order tensors~\citep{han2020exact}. 
 The double projection improves usual matrix spectral methods in order to alleviate the noise tensor.
-%The first projection performs HOSVD~\citep{de2000multilinear} via $\mU_{\text{pre}} = \text{SVD}_{r} \of{ \text{Mat}(\tY) }$, where $\text{SVD}_r(\cdot)$ returns the top-$r$ left singular vectors.
-The second projection performs HOSVD on the projected $\tY$ onto the multlilinear Kronecker space $\mU_{\text{pre}} \otimes  \mU_{\text{pre}}$; i.e.,
-%\begin{equation}\label{eq:two-step_factor}
- %   \hat \mU = \text{SVD}_{r} \of{\text{Mat}\of{ \tY\times_1  \mU_{\text{pre}}^T \times_2  \mU_{\text{pre}}^T }}.
-%\end{equation}
-%The final denoised tensor $\hat \tX$ is defined by
-%\begin{equation}\label{eq:two-step_est}
- %   \hat \tX = \tY \times_1 \hat \mU \hat 
-%\mU^T \times_2 \hat \mU \hat \mU^T \times_3 \hat \mU \hat \mU^T. 
-%\end{equation}
-%Compared with conventional HOSVD, our double projection spectral method is shown to improve the estimation substantially for higher-order tensors~\citep{han2020exact}. 
+
+}
+
 The clustering step (lines 3-5 in Sub-algorithm~\hyperref[alg:main]{1}) performs the weighted $k$-means clustering. 
-%We write $\hat \mX=\text{Mat}(\hat \tX)$, and normalize the rows into $\hat \mX^s_{i:}=\onormSize{}{\hat \mX_{i:}}^{-1}\hat \mX_{i:}$ as a surrogate of $\theta(i)^{-1} \mX_{i:}$.
+{
+\color{blue}
+We write $\hat \mX=\text{Mat}(\hat \tX)$, and normalize the rows into $\hat \mX^s_{i:}=\onormSize{}{\hat \mX_{i:}}^{-1}\hat \mX_{i:}$ as a surrogate of $\theta(i)^{-1} \mX_{i:}$.
 Then, a weighted $k$-means clustering is performed on the normalized rows with weights equal to $\onormSize{}{\hat \mX_{i:}}^2$.
+}
 The choice of weights is to bound the $k$-means objective function by the Frobenius-norm accuracy of $\hat \tX$. 
@@
 This strategy relaxes the eigen-gap separation condition~\citep{gao2018community, han2020exact}.
-%We assign degenerate rows with purely zero entries to an arbitrarily random cluster; these nodes are negligible in high-dimensions because of the lower bound on $\onormSize{}{\text{Mat}(\tS)_{a:}}$ in~\eqref{eq:family}.
+{
+\color{blue}
+We assign degenerate rows with purely zero entries to an arbitrarily random cluster; these nodes are negligible in high-dimensions because of the lower bound on $\onormSize{}{\text{Mat}(\tS)_{a:}}$ in~\eqref{eq:family}.
 The final result gives the initial clustering assignment $\hat z^{(0)}$.
+}
 Full procedures are provided in Sub-algorithm~\hyperref[alg:main]{1}.
@@
     &\sum_{i \in I_0^c }  \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{\hat z(i)} }^2 
-    \leq \\
-    &\eta \min_{\substack{\bar \mx_j, j\in[r], \bar z(i),i\in I_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\bar z(i)}}^2.
+    \leq 
+    \eta \min_{\substack{\bar \mx_j, j\in[r], \bar z(i),i\in I_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\bar z(i)}}^2.
 \end{align}
@@
 \begin{align}
-    &\tY^{\text{d}}(i,a_2,\ldots,a_K) \\
-    &= 
+    &\tY^{\text{d}}(i,a_2,\ldots,a_K) 
+    = 
 \text{Ave}\{\tY(i,i_2,\ldots,i_K): z^{(t)}(i_k) = a_k, k \neq 1 \}.
@@
 
-\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general Gaussian dTBM under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}. 
+\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general {\color{blue} sub-Gaussian} dTBM {\color{blue} with i.i.d.\ noise} under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
 %With probability at least $1 -C \exp(-cp)$, we have
@@
 However, the following theorem indicates a weaker signal condition is enough to achieve exact clustering with proper refinement.
+
 \begin{rmk}[Failure of conventional tensor HOSVD] If we use conventional HOSVD for tensor denoising; that is, we use $\mU_{\text{pre}}$ in place of $\hat \mU$ in line 2, then the misclustering rate becomes $\tO(p)$ for all $K\geq 2$.
@@
 
+{
+\color{blue}
+
+\begin{rmk}[Eigen free clustering] Note that our initialization works on the estimated mean tensor $\hat \tX$ directly rather than the leading tensor decomposition factors of $\tX$.
+On one hand,  clustering on $\hat \tX$ avoids the eigen-gap assumption in previous work~\cite{ke2019community}.
+On the other hand,  vanilla spectral methods working on decomposition factors suffers the non-identifiability of the eigenspace with orthogonal transformation when the number of blocks $r \geq 3$.
+Such ambiguity comes from the possible multiplicity of eigenvalue and causes trouble for effective clustering~\citep{abbe2020entrywise}.
+In contrast, our eigen free strategy working on $\hat \tX$ avoids such non-identifiability issue regardless the number of blocks.
+\end{rmk}
+
+}
+
 \subsection{Angle-based iteration}
@@
 We propose an angle-based local iteration to improve the outputs from Sub-algorithm~\hyperref[alg:main]{1}.
-To gain the intuition, consider an one-dimensional degree-corrected clustering problem with data vectors $\mx_i = \theta(i) \ms_{z(i)} + \mepsilon_i, i \in [p]$, where $\ms_i$'s are known cluster centroids, $\theta(i)$'s are unknown degrees, and $z\colon [p] \mapsto [r]$ is the clustering assignment of interest.
+To gain the intuition, consider an one-dimensional degree-corrected clustering problem with data vectors $\mx_i = \theta(i) \ms_{z(i)} + \mepsilon_i, i \in [p]$, where $\ms_i$'s are known cluster centroids, $\theta(i)$'s are unknown positive degrees, and $z\colon [p] \mapsto [r]$ is the clustering assignment of interest.
 The angle-based $k$-means algorithm estimates the assignment $z$ by minimizing the angle between data vectors and centroids; i.e.,
@@
 In contrast, the proposed angle-based $k$-means achieves accurate recovery without explicit estimation of $\mtheta$.
-Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as angle-based $k$-means, except that we use estimated centroids $\ms^{(t)}_a$ in place of $\ms_a$ based on estimated assignment in previous iterations.
- % Algorithm~\ref{alg:refinement} updates estimated core tensor and cluster assignments are updated in each iteration. For core tensor, we consider the following update strategy  
- %\[
-%\tilde \tS(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z(i_k)=a_k, k\in[3]\}.
-%\]
-%Intuitively, $\tS^{(t)}$ becomes closer to the true core $\tS$ as $z^{(t)}$ is more precise.
-we first aggregate the slices of $\tY$ and obtain a reduced tensor $\tY^{\text{d}}\in \bbR^{p \times r \times r}$ with given $z^{(t)}$, where
-%\[
-%\tY^{\text{d}}(i,a_2,a_3)=\text{Ave}\{\tY(i,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k \neq 1\}.
-%\]
-%The row $\text{Mat}(\tY^{\text{d}})_{i:}$ and $\text{Mat}(\tS^{(t)})_{a:}$ corresponds to the $\mx_i$ and $\ms_a$s in the one-dimensional clustering~\eqref{eq:angle_kmeans}.
+
+Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as angle-based $k$-means. {\color{blue} We still take the order-3 tensor for illustration.} 
+{
+\color{blue}
+Specifically, Sub-algorithm~\hyperref[alg:main]{2} updates estimated core tensor and cluster assignment in each iteration. For core tensor, we consider the following update strategy  
+ \[
+ \tS^{(t)}(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k\in[3]\}.
+\]
+Intuitively, $\tS^{(t)}$ becomes closer to the true core $\tS$ as $z^{(t)}$ is more precise.
+For cluster assignment, we first aggregate the slices of $\tY$ and obtain a reduced tensor $\tY^{\text{d}}\in \bbR^{p \times r \times r}$ with given $z^{(t)}$, where
+\[
+\tY^{\text{d}}(i,a_2,a_3)=\text{Ave}\{\tY(i,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k \neq 1\}.
+\]
+The row $\text{Mat}(\tY^{\text{d}})_{i:}$ and $\text{Mat}(\tS^{(t)})_{a:}$ corresponds to the $\mx_i$ and $\ms_a$s in the one-dimensional clustering~\eqref{eq:angle_kmeans}.
 Then, we obtain the updated assignment as
-%\[z(i)^{(t+1)} = \argmin_{a \in [r]} \sin \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} }, \text{ for all } i \in [p], \]
-%where $z^{(t+1)}(i)$ is randomly assigned for degenerate rows. 
+\[z(i)^{(t+1)} = \argmin_{a \in [r]} \sin \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} }, \text{ for all } i \in [p], \]
+where $z^{(t+1)}(i)$ is randomly assigned for degenerate rows. 
+}
+%except that we use estimated centroids $\ms^{(t)}_a$ in place of $\ms_a$ based on estimated assignment in previous iterations. 
 Full procedures for our angle-based iteration are described in Sub-algorithm~\hyperref[alg:main]{2}.
 
-We then establish the misclustering error rate of iterations.
-Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\epsilon)=\{\bar z\colon \ell(\bar z, z)\leq \epsilon\}$.
-We define two cluster-size vectors for $\bar z$,
-\begin{align}
-    &\mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T,\\
-    &\mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
+We then establish the misclustering error rate of iterations 
+{ \color{blue} under the stability assumption.
+
+\begin{defn}[Locally linear stability] 
+Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\epsilon)=\{\bar z\colon \ell(\bar z, z)\leq \epsilon\}$. We define two cluster-size vectors for $\bar z$,
+\begin{align}
+    &\mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T, \quad \mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
 \end{align}
@@
 \end{equation}
+\end{defn}
+}
 
 The local stability holds trivially for $\varepsilon=0$ based on the construction of parameter space~\eqref{eq:family}.
+{\color{blue} The locally linear stability avoids the concentration of entities with extremely large or small heterogeneity, when a good estimated assignment with a small misclustering error $\epsilon$ is given.
+}
 
@@
    \text{SNR}^{-1}
-    \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}}_{\substack{\text{statistical}\\ \text{error}} }+ \KeepStyleUnderBrace{ \rho^t \ell(z, z^{(0)}). }_{\substack{\text{computational}\\ \text{error}}}
+    \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}}_{\substack{\text{statistical error}} }+ \KeepStyleUnderBrace{ \rho^t \ell(z, z^{(0)}). }_{\substack{\text{computational error}}}
 \end{align}
@@
 \end{equation}
-for some permutation $\pi \in \Pi$.
-
-\begin{rmk}[Extension for general dTBMs]
+for some permutation $\pi \in \Pi$. {
+\color{blue} We call our algorithm \textit{computationally efficient} with SNR $\gtrsim p^{-K/2} \log p$.
+Note that the minimal SNR requirement, $p^{-K/2 }\log p$, coincides with the computational lower bound in Theorem~\ref{thm:comp} ignoring the logarithmic term.
+Therefore, our algorithm is optimal regarding the signal demand and lies in the most right ``computationally efficient" regime in Figure~\ref{fig:phase_axis}.
+}
+
+{
+\color{blue}
+\subsection{Extension and practical issue}~\label{subsec:exten}
+{\bf Extension for Bernoulli observations.} The main difficulty to establish the statistical guarantee for Bernoulli observations lies in the initialization Sub-Algorithm~\hyperref[alg:main]{1}.
+Theorem~\ref{thm:refinement} still holds for Bernoulli observations once the initialization accuracy satisfies the upper bound~\eqref{eq:ini} in Theorem~\ref{thm:initial}.
+
+Specifically, the derivation of Theorem~\ref{thm:initial} relies on the upper bound of the estimation error for the mean tensor in Lemma~\ref{lem:two-step_esterror}, i.e., with high probability
+\begin{equation}\label{eq:prop1}
+    \onormSize{}{\hat \tX - \tX}_F^2 \lesssim p^{K/2},
+\end{equation}
+where $\tX = \bbE[\tY]$ and $\hat \tX$ is defined in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}.
+Unfortunately, the inequality~\eqref{eq:prop1} only holds for i.i.d.\ sub-Gaussian observations while Bernoulli observations are not identically distributed.
+
+One possible remedy is to apply singular value decomposition to the unfolded Bernoulli observation $\tY$.
+Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{\floor{p^{K/2}} \times \ceil{p^{K/2}} }$ denote the square unfolded binary tensor.
+We have the estimate
+\begin{equation}
+    \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX)) -  \mat_{sq}(\tY)}_F^2.
+\end{equation}
+Following Lemma 7 in \cite{gao2018community}, with high probability, we have
+\begin{equation}
+    \onormSize{}{\hat \tX' - \tX}_F^2 \lesssim p^{\ceil{K/2}}.
+\end{equation} 
+Replacing the estimate $\hat \tX$ by $\hat \tX'$, the high probability upper bound for Bernoulli initialization is
+\begin{equation}\label{eq:ini_b}
+    \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}. 
+\end{equation}
+The Bernoulli bound~\eqref{eq:ini_b} is relatively looser than Gaussian bound~\eqref{eq:ini}, especially when $K$ is small.
+A tighter Bernoulli bound will be achieved once a low-rank binary tensor estimation scheme with better accuracy is provided in the future.
+
+{\bf Extension for general dTBMs.}
 Our two-stage algorithm is able to be extended for the general (asymmetric) dTBMs. Specifically, in the Sub-Algorithm~\hyperref[alg:main]{1}, we make the following changes: (1)
@@
 \begin{align}
-    &\tY^{\text{d}}_k(a_1,\ldots,a_{k-1}, i ,a_{k+1},\ldots,a_K) \\
-    &= \text{Ave}\{\tY(i_1,\ldots,i_{k-1},i,i_{k+1},\ldots,i_K): z^{(t)}(i_j) = a_j, j \neq k \}; 
+    &\tY^{\text{d}}_k(a_1,\ldots,a_{k-1}, i ,a_{k+1},\ldots,a_K) 
+    = \text{Ave}\{\tY(i_1,\ldots,i_{k-1},i,i_{k+1},\ldots,i_K): z^{(t)}(i_j) = a_j, j \neq k \}; 
 \end{align}
@@
 
-\end{rmk}
-
-{
-\color{red}
-
-\subsection{Bernoulli Case}
-
-\fixme{Jiaxin}{My thoughts about the Bernoulli case}
-\paragraph{Bernoulli Case of Proposition 1 in \cite{han2020exact}} Here I write my thoughts about the Bernoulli Proposition 1 following the notation in \cite{han2020exact}.
-
-According to \cite{han2020exact}, we have 
-\begin{align}
-    \onorm{\hat \tX - \tX}_F \leq &\onorm{ \tX \times_1 \hat \mU_1  \hat \mU_1^T \times_2 \cdots \times_d  \hat \mU_d  \hat \mU_d^T - \tX}_F \\
-    &+ \onorm{\tZ \times_1 \hat \mU_1^T \times_2 \cdots \times_d \mU_d^T}_F, 
-\end{align}
-where the second term 
-\begin{equation}
-    \onorm{\tZ \times_1 \hat \mU_1^T \times_2 \cdots \times_d \mU_d^T}_F \lesssim \sqrt{p},
-\end{equation}
-dose not require the noise entries to be identically distributed following Lemma 9. For the first term, 
-\begin{align}
-    &\onorm{ \tX \times_1 \hat \mU_1  \hat \mU_1^T \times_2 \cdots \times_d  \hat \mU_d  \hat \mU_d^T - \tX}_F \\
-    &\leq  \sum_{k=1}^d \onorm{\hat \mU^T_{k, \perp} \mX_k}_F 
-     \lesssim   \sum_{k=1}^d  \onorm{\hat \mU_{k,\perp}^{T}  \bar \mX_k}_F + \onorm{\mX_1 - \mX_1'}_F,
-\end{align}
-where the term $\onorm{\hat \mU_{k,\perp}^{T}  \bar \mX_k}_F \lesssim \sqrt{p}$ following Lemma 8 and the term $\onorm{\mX_1 - \mX_1'}_F \lesssim p^{d/4} + \sqrt{p}$ following the definition of $\mX_1'$.
-Particularly, $\mX'_1$ can be considered as the original signal tensor only with "strong" singular values, and the threshold for "strong" singular value is $p^{d/4} \vee \sqrt{p}$.
-
-We choose $p^{d/4} \vee \sqrt{p}$ due to the accuracy of SVD.
-Let $r$ denote the rank of $\mX'_1$ and $r$ is related to the threshold.
-If the threshold is lower than $p^{d/4} \vee \sqrt{p}$, then the principle angle between the SVD singular space and the true rank $r$ singular space will inflate, i.e., the inequality between (48) (49) in \cite{han2020exact}
-\begin{equation}
-    \onorm{ \tilde \mU_{k, \perp}^T \mU_k' } \leq \frac{1}{\sqrt{2}},
-\end{equation}
-does not hold.
-Note that the above inequality is where the $p^{K/2}$ term comes from in the Proposition 1, and the inequality follows by Lemma 3.
-Lemma 3 follows the Lemma 4 in \cite{cai2018rate} which is only applicable for the i.i.d.\ sub-Gaussian noise and is the reason why the Proposition 1 requires i.i.d.
-
-The Lemma 7 in \cite{gao2018community} provides the accuracy of SVD of the binary matrix, i.e., for a binary matrix $\mY \in \{0,1\}^{p_1 \times p_2}$ with $\bbE \mY = \mP$ and  $\hat \mP$ which is the rank $r$ SVD of $\mY$, we have
-\begin{equation}
-    \onorm{\hat \mP - \mP}_F \leq \sqrt{r (p_1 \vee p_2)}.
-\end{equation}
-However, Lemma 7 does not provide the result for the principle angle between the estimated and true singular spaces.
-So, I am not sure where can we apply Lemma 7 to the Bernoulli case of Proposition 1.
-In addition, both Han's algorithm and our algorithm do not matricize the tensor to a $p^{\ceil{K/2}}$-by-$p^{\floor{K/2}}$ matrix and apply SVD to the matrix.
-Thus, I am not sure why there will have the $p^{\ceil{K/2}}$ term in the Bernoulli case.
-
-}
-
-{
-\color{red}
-\subsection{Rank Selection}
-\subsection{Computational Complexity}
+{\bf Computational Complexity.} Our two-stage algorithm has a polynomial computational cost.
+Specifically, the complexity of Sub-Algorithm~\hyperref[alg:main]{1} is $\tO(K p^{K+1} + Krp^K)$, where the first term is contributed by the first step SVD and the calculation of $\hat \tX$, and the second term comes from normalization and the $k$-means.
+The cost of each update in Sub-Algorithm~\hyperref[alg:main]{2} is $\tO(p^K + pr^K)$, where $p^K$ comes from the calculation of $\tS^{(t)}$ and $\tY^{\text{d}}$ and $pr^K$ comes from the normalization of $\tY^{\text{d}}, \tS^{(t)}$ and the cluster assignment update in Step 10.
+
+%that comes from the addition operation and comparison in Step 7,8, and 10.
+Note that the number of iteration is a logarithmic function of $p$ to achieve exact clustering.
+We conclude that the two-stage algorithm leads to a polynomial complexity in total.
+
+{\bf Parameter Selection.} Note that we assume the number of clusters $r$ is given for the algorithm. We impose the Bayesian information criterion (BIC)
+and choose the number of cluster that minimizes BIC, i.e., under dTBM~\eqref{eq:model_tensor},
+\begin{equation}
+    \hat r = \argmin_{r \in \mathbb{Z}_+ } \of{ -p^K \tL_{\tY}(\hat z(r), \hat \tS(r), \hat \mtheta(r)) + p_e(r) K \log p},
+\end{equation}
+where the triplet $(\hat z(r), \hat \tS(r), \hat \mtheta(r))$ are estimated parameters with given number of cluster $r$, $\tL_{\tY}$ denote the log-likelihood with observations $\tY$, and $p_e = r^K + p(\log r + 1)$ is the effective number of parameters.
+Particularly, we obtain the estimate $\hat \mtheta(r)$ by first calculating the reduced tensor $\tY^{\text{d}}$ with $\hat z(r)$ and then normalizing the row norms $\onormSize{}{\mY_{i:}^{\text{d}}}$ to 1 in each cluster.
+With Gaussian observation $\tY$, we have
+\begin{equation}
+    \tL_{\tY} = - \log(\onormSize{}{\hat \tX - \tY}_F^2), \quad \hat \tX = \tS(r) \times_1 \hat \mTheta(r) \hat \mM(r) \times_2 \cdots \times_K \hat \mTheta(r) \hat \mM(r).
+\end{equation}
+We evaluate the performance of the BIC criterion in Section~\ref{subsec:num_theory}.
+
+% We here provide a simple rank selection criteria when $r$ is unknown. With a fixed $r$, calculate the estimated minimal gap
+% \begin{equation}
+%     \hat \Delta_{\min}(r) = \min_{a \neq b \in [r]} \onorm{ \frac{\hat \mS_{a:}(r)}{\onormSize{}{\hat \mS_{a:}(r)}} - \frac{\hat \mS_{b:}(r)}{\onormSize{}{\hat \mS_{b:}(r)}} },
+% \end{equation}
+% where $\mS(r) = \mat(\hat \tS(r))$ and $\hat \tS(r)$ is the estimated signal tensor obtained by the same procedure in Step 7 of Sub-Algorithm~\hyperref[alg:main]{2} with the estimated assignment $\hat z(r)$.
+With a given upper bound for the number of clusters, denoted $R$, we choose $r$ that maximizes the estimated minimal gap, i.e.,
+% \begin{equation}
+%     \hat r = \argmax_{r \in [R]} \hat \Delta_{\min}(r).
+% \end{equation}
+
 }
@@
 Without further explanation, we generate degree heterogeneity $\mtheta$ from absolute normal distribution as $\theta(i) = |X_i| + 1 - 1/\sqrt{2\pi}$ with $|X_i| \stackrel{\text{i.i.d.}}\sim N(0,1), i \in [p]$ and normalize $\mtheta$ to satisfy \eqref{eq:family}.
-We set $\sigma^2 = 1$ for Gaussian data.
+Also, we set $\sigma^2 = 1$ for Gaussian data without further specification.
 
@@
 
-\subsection{Verification of theoretical results}
+\subsection{Verification of theoretical results}\label{subsec:num_theory}
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{phase_anno3.pdf}
+    \includegraphics[width=.8\columnwidth]{phase_anno3.pdf}
     \caption{SNR phase transitions for clustering in dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.
@@
     \centering
-     \includegraphics[width=\columnwidth]{ini_re_anno3.pdf}
+     \includegraphics[width=.8\columnwidth]{ini_re_anno3.pdf}
     \caption{CER versus signal exponent $(\gamma)$ for initialization only and for combined algorithm.
@@
 {
-\color{red} 
-More experiments:
-\begin{enumerate}
-    \item Initialization impact to Angle-based iteration.(Figure 4 in Han)
-    \item Imbalanced cluster size.(Figure 9 in Han)
-    \item Non-degree comparison. 
-\end{enumerate}
+\color{blue}
+
+\begin{table}[hbt]
+\color{blue}
+    \centering
+    \begin{tabularx}{\textwidth}{c *{12}{Y}}
+    \toprule
+    Settings & \multicolumn{3}{c}{$p = 50, \sigma^2 = 0.25$} & \multicolumn{3}{c}{$p = 50, \sigma^2 = 1$} & \multicolumn{3}{c}{$p = 80, \sigma^2 = 0.25$} & \multicolumn{3}{c}{$p = 80, \sigma^2 = 1$}\\
+    \cmidrule(lr){2-4} \cmidrule(l){5-7} \cmidrule(l){8-10} \cmidrule(l){11-13}
+         True number of clusters $r$ & 2&3 & 4 & 2 & 3 & 4 & 2 & 3 & 4 & 2 & 3 & 4  \\
+         \midrule
+         Estimated number of clusters $\hat r$ &  2 & 3 & 3.9& 2 & 3 & 3.1 & 2 & 3 & 4 & 2 & 3& 3.9   \\
+     \bottomrule
+    \end{tabularx}
+    \caption{Estimated number of clusters given by BIC criterion under the small noise $(\sigma^2 = 0.25)$ and large noise $(\sigma^2 = 0.5)$ cases.
+We report the averaged estimated number of clusters $\hat r$ over 30 duplication.}
+    \label{tab:select}
+\end{table}
+
+The third experiment evaluates the empirical performance of the BIC criterion to select unknown number of cluster.
+We generate the data from an order-3 Gaussian model with $p = \{50,80\}$, $r = \{2,4\}$ and consider the noise $\sigma^2 \in \{ 0.25,1\}$ with $\alpha = 400$.
+Table~\ref{tab:select} implies that our BIC criterion exactly choose the true $r$ under all the settings with small $r = 2, 3$.
+Note that the BIC underestimates the large true number of clusters $(r = 4)$ with smaller SNR $(\sigma^2 = 1)$, and the accuracy immediately increases with larger dimension $p = 80$.
+The improvement follows by the fact that a larger dimension $p$ indicates a larger sample size in the block model.
+Therefore, we conclude that BIC criterion is a reasonable way to tune the number of clusters.
 }
+% {
+% \color{red} 
+% More experiments:
+% \begin{enumerate}
+%     \item Initialization impact to Angle-based iteration.(Figure 4 in Han)
+%     \item Imbalanced cluster size.(Figure 9 in Han)
+%     \item Non-degree comparison. 
+% \end{enumerate}
+% }
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{comp_gamma_anno3.pdf}
+    \includegraphics[width=.8\columnwidth]{comp_gamma_anno3.pdf}
     \caption{CER versus signal exponent (denoted $\gamma$) for different methods. We set $p = 100, r = 5, \gamma \in [-1.5, -1.1]$ under (a) Gaussian and (b) Bernoulli models.}
@@
     \centering
-    \includegraphics[width=\columnwidth]{comp_theta_anno3.pdf}
+    \includegraphics[width=.8\columnwidth]{comp_theta_anno3.pdf}
     \caption{CER versus shape parameter in degree (denoted $a\in[3,6]$) for different methods. We set $p = 100, r = 5, \gamma = -1.2$ under (a) Gaussian and (b) Bernoulli models.}
@@
     \centering
-    \includegraphics[width=\columnwidth]{comp_gamma_theta3.pdf}
+    \includegraphics[width=.8\columnwidth]{comp_gamma_theta3.pdf}
     \caption{CER comparison versus signal exponent (denoted $\gamma$) under (a) low (shape parameter $a = 6$)  (b) high (shape parameter $a = 2$) degree heterogeneity.
@@
     \centering
-    \includegraphics[width = \columnwidth]{brain_node_cluster.pdf}
+    \includegraphics[width = .8\columnwidth]{brain_node_cluster.pdf}
     \caption{Illustration of brain node clustering results for HCP data with (a) top and (b) side views. }
@@
     \centering
-    \begin{tabular}{c |c  c c c}
+    \begin{tabular}{c |c  c cc c}
     \hline
         Method & \textbf{\small dTBM} 
-        %&\textbf{\small HOSVD}
+        &\textbf{\small HOSVD}
         &\textbf{\small HOSVD+} & \textbf{\small HLloyd} &  \textbf{\small SCORE}\\
          CER & \textbf{0.116}
-         %&  0.22 
+         &  0.22 
          &0.213 & 0.149 &0.199\\
@@
 
-\section{Conclusion}
+\section{Conclusion}\label{sec:con}
 We have developed a general degree-corrected tensor block model with a two-step angle-based polynomial-times algorithm.
@@
 
-{\color{red}
-Future direction. 
-\begin{enumerate}
-    \item Missing data.
-    \item Sub-exponential data.
-    \item Imbalance in community size and heterogeneity.
-\end{enumerate}
-}
+% {\color{red}
+% Future direction. 
+% \begin{enumerate}
+%     \item Missing data.
+%     \item Sub-exponential data.
+%     \item Imbalance in community size and heterogeneity.
+% \end{enumerate}
+% }
 

### 2022-01-06 tensor_clustering/tensor_clustering_ieee/ieee.tex
update

@@
 %The Tucker rank of the signal tensor $\bbE \tY$ is bounded by, but not necessarily equal to, $(r,...,r)$~\citep{wang2019multiway}.
-We call a tensor $\tX$ an $r$-block tensor with degree $\mtheta$
+We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$
 %with the core tensor $\tS$ 
-if $\tX$ admits dTBM~\eqref{eq:model_tensor}. Here, we give two special cases of dTBM.
+if $\tY$ admits dTBM~\eqref{eq:model_tensor}. Here, we give two special cases of dTBM.
 
@@
 $\hat \tX = \tY \times_1 \hat \mU \hat \mU^T \times_2 \cdots \times_K \hat \mU \hat \mU^T.$
-\State {Let $\hat \mX = \text{Mat}(\hat \tX)$ and $I_0=\{i \in [p]: \onormSize{}{\hat \mX_{i:}} = 0\}$. Set $\hat z(i)$ randomly in $[r]$ for $i \in I_0$.}
-\State{For all $i\in I_0^c$, compute normalized rows
+\State {Let $\hat \mX = \text{Mat}(\hat \tX)$ and $S_0=\{i \in [p]: \onormSize{}{\hat \mX_{i:}} = 0\}$. Set $\hat z(i)$ randomly in $[r]$ for $i \in S_0$.}
+\State{For all $i\in S_0^c$, compute normalized rows
 $\hat \mX_{i:}^s :=\onormSize{}{\hat \mX_{i:}}^{-1} \hat \mX_{i:}.$
@@
 \begin{align}
-    &\sum_{i \in I_0^c }  \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{\hat z(i)} }^2 
+    &\sum_{i \in S_0^c }  \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{\hat z(i)} }^2 
     \leq 
-    \eta \min_{\substack{\bar \mx_j, j\in[r], \bar z(i),i\in I_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\bar z(i)}}^2.
+    \eta \min_{\substack{\bar \mx_j, j\in[r], \bar z(i),i\in S_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\bar z(i)}}^2.
 \end{align}
@@
 \begin{rmk}[Comparison to previous results] For fixed SNR, our initialization error rate with $K=2$ agrees with the initialization error rate $\tO(p)$ in matrix models~\citep{gao2018community}.
-Furthermore, in the special case of non-degree TBMs, we achieve the same initial misclassification error $\tO(p^{-K/2})$ as in non-degree models~\citep{han2020exact}.
+Furthermore, in the special case of non-degree TBMs, we achieve the same initial misclustering error $\tO(p^{-K/2})$ as in non-degree models~\citep{han2020exact}.
 Theorem~\ref{thm:initial} implies the advantage of our algorithm in achieving both accuracy and model flexibility.
@@
 
-\begin{defn}[Locally linear stability] 
+\begin{defn}[Locally linear stability] \label{def:stable}
 Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\epsilon)=\{\bar z\colon \ell(\bar z, z)\leq \epsilon\}$. We define two cluster-size vectors for $\bar z$,
@@
 \end{equation}
-where the triplet $(\hat z(r), \hat \tS(r), \hat \mtheta(r))$ are estimated parameters with given number of cluster $r$, $\tL_{\tY}$ denote the log-likelihood with observations $\tY$, and $p_e = r^K + p(\log r + 1)$ is the effective number of parameters.
-Particularly, we obtain the estimate $\hat \mtheta(r)$ by first calculating the reduced tensor $\tY^{\text{d}}$ with $\hat z(r)$ and then normalizing the row norms $\onormSize{}{\mY_{i:}^{\text{d}}}$ to 1 in each cluster.
+where the triplet $(\hat z(r), \hat \tS(r), \hat \mtheta(r))$ are estimated parameters with given number of cluster $r$, $\tL_{\tY}$ denote the log-likelihood with observations $\tY$, and $p_e = r^K + p(\log r + 1) - 1$ is the effective number of parameters.
+Particularly, we obtain the estimate $\hat \mtheta(r)$ by first calculating the reduced tensor $\tY^{\text{d}}$ with $\hat z(r)$ and then normalizing the row norms $\onormSize{}{\mY_{i:}^{\text{d}}}$ to 1 in each cluster; i.e.,
+\begin{equation}
+    \hat \theta(i,r) = \frac{\onormSize{}{\hat \mY^{\text{d}}(r)_{i:}}}{\sum_{j: \hat z(j,r) =  \hat z(i,r)} \onormSize{}{\hat \mY^{\text{d}}(r)_{j:}}},
+\end{equation}
+where $\hat \mY^{\text{d}}(r) = \mat(\hat \tY^{\text{d}}(r))$ and 
+\begin{equation}
+    \hat \tY^{\text{d}}(r)(i,a_2,\ldots,a_K) = 
+\text{Ave}\{\tY(i,i_2,\ldots,i_K): \hat z(i_k, r) = a_k, k \neq 1 \}.
+\end{equation}
+
 With Gaussian observation $\tY$, we have
@@
 {
-\color{red}
+\color{blue}
 
@@
 
-\begin{enumerate}
-\item \textbf{Angle-based techniques} 
-\begin{itemize}
-    \item Padding of the vectors won't extremely change the angle gap. 
-    \item Using geometry property to upper bound $I_{11}, J_{11}$. (have not found a good name for this technique yet)
-\end{itemize}
-    \item \textbf{Proof idea of Theorem 4}
-    \item \textbf{Proof idea of Theorem 5}
-\end{enumerate}
+In this section, we provide the proof sketches for Theorem~\ref{thm:initial} and Theorem~\ref{thm:refinement}. Detail proofs are provided in Appendix.
+
+\subsection{Proof sketch of Theorem~\ref{thm:initial}}
+The proof of Theorem~\ref{thm:initial} is mainly inspired by the proof idea of \citet[Lemma 1]{gao2018community}, and extra difficulties due to the angle gap assumption and tensor property are addressed in our proof.
+Specifically, we control the misclustering error by the estimation error of $\hat \tX$ calculated in Step 2.
+We prove the following inequality
+\begin{equation}\label{eq:proof_4}
+    \ell(z^{(0)},z) \lesssim \frac{1}{p}\min_{\pi \in \Pi} \sum_{i: z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2 \lesssim \frac{\sigma^2 r^{K-1}}{ \Delta_{\min}^2 p^K} \onormSize{}{\hat \tX - \tX}_F^2 \lesssim \frac{r^K p^{-K/2}}{\text{SNR}},
+\end{equation}
+where $\tX = \bbE[\tY]$ is the true mean.
+The first inequality in~\eqref{eq:proof_4} holds with the assumption $\min_{i \in [p]} \theta(i) \geq c$ in Theorem~\ref{thm:initial}.
+The second inequality relies on the intermediate conclusion that the angle gap of mean tensor $\tX$ is lower bounded by a function of core tensor minimal angle gap $\Delta_{\min}$ defined in Assumption~\ref{assmp:min_gap}.
+Let $\ma^{s} \coloneqq \ma/\onorm{\ma}$ denote the normalized vector and make the convention that $\ma^s = 0$ if $\ma = 0$.
+We show that
+\begin{equation}\label{eq:proof_4_gap}
+\min_{z(i) \neq z(j)} \onormSize{}{[\mX_{i:}]^s - [\mX_{j:}]^s} \gtrsim \Delta_{\min},
+\end{equation}
+where $\mX = \mat(\tX)$.
+The most challenging part in the proof of Theorem~\ref{thm:initial} lies in the derivation of inequality~\eqref{eq:proof_4_gap}, in which the proof of \cite{gao2018community} is no longer applicable due to different angle gap assumption in our dTBM.
+Extra padding technique in Lemma~\ref{lem:pad} and balance assumption~\eqref{eq:degree} are equipped to derive~\eqref{eq:proof_4_gap}.
+Last, we finish the proof of Theorem~\ref{thm:initial} by showing the third inequality with \citet[Proposition 1]{han2020exact}.
+
+\subsection{Proof sketch of Theorem 5}
+The proof of Theorem~\ref{thm:refinement} is mainly inspired by the proof idea of  \citet[Theorem 2]{han2020exact}, and extra polar coordinate based techniques are imposed due to the nuisance degree heterogeneity and angle gap assumption.
+We conduct the contraction property of the misclustering loss
+\begin{equation}
+     L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
+\end{equation}
+and introduce the oracle estimators for core tensor and assignment
+\begin{equation}
+    \tilde \tS = \tY \times_1 \mW^T \times_2 \cdots \times_K \mW^T, 
+    %\quad \tilde z(i) = \argmin_{a \in [r]} \onormSize{}{ { [\mat(\tY)_{i:} \mV]^s - [\mat(\tilde \tS)_{a:}]^s } }^2
+\end{equation}
+where $\mW = \mM \of{ \text{diag}(\ind_{p}^T \mM) }^{-1}$ is the weighted true membership matrix.
+Let $ \mV = \mW^{\otimes (K-1)}$ denote the Kronecker product of $K-1$ $\mW$ matrices, and similarly define $\mW^{(t)}, \mV^{(t)}$ with $\mM^{(t)}$ corresponding to $z^{(t)}$.
+To evaluate $L^{(t+1)}$, we consider the event
+\begin{equation}\label{eq:proof_5_event}
+    \ind \offf{ z^{(t+1)}(i) = b } = \ind \offf{       \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{b:}^{(t)}]^s }^2 \leq \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{z(i):}^{(t)}]^s }^2} \leq A_{ib} + B_{ib}
+\end{equation}
+where $\mY = \mat(\tY)$, $ \mS = \mat(\tS)$, $\mS^{(t)} = \mat(\tS^{(t)})$ and
+\begin{align}
+        A_{ib} &= \ind \offf{\ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s } \lesssim -  \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 },\\
+        B_{ib} &= \ind \offf{\onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 \lesssim F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }.
+\end{align}
+The terms $F_{ib}^{(t)}, G_{ib}^{(t)}, H_{ib}^{(t)}$ are controlled by $z^{(t)}, \tS^{(t)}$, and detail definitions are in \eqref{eq:f}, \eqref{eq:g}, \eqref{eq:h}.
+Note that the event $A_{ib}$ only involves the oracle estimator while all the terms related to the $t$-th iteration estimator are in $B_{ib}$.
+Thus, the inequality~\eqref{eq:proof_5_event} decomposes the misclustering loss in the $(t+1)$-th iteration by the oracle loss and the loss in $t$-th iteration.
+This decomposition leads to the separation of statistical error and computational error in the final upper bound of Theorem~\ref{thm:refinement}.
+
+Specifically, we prove the contraction inequality
+\begin{equation}\label{eq:proof_5_ineq}
+    L^{(t+1)} \lesssim \xi + \rho L^{(t)}, \quad \xi = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  A_{ib} \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
+\end{equation}
+where $\rho \in (0,1)$ is the contraction parameter, and we call $\xi$ the oracle loss.
+Controlling the probability of event $B_{ib}$ and obtaining the $\rho L^{(t)}$ term in the right hand side of~\eqref{eq:proof_5_ineq} is the most tricky and challenging part in the proof of Theorem~\ref{thm:refinement}.
+Note that the true and estimated core tensors are involved as normalized rows such as $\mS_{a:}^s, \tilde \mS_{a:}^s, [\mS^{(t)}_{a:}]^s$.
+The Cartesian coordinate based analysis in \cite{han2020exact} is no longer applicable in our case.
+Instead, we use the polar coordinate based analysis and the geometry property of trigonometric functions to derive the high probability upper bounds for $F_{ib}^{(t)}, G_{ib}^{(t)}, H_{ib}^{(t)}$.
+
+Further, by sub-Gaussian concentration, we prove the high probability upper bound for oracle loss
+\begin{equation}\label{eq:proof_5_xi}
+    \xi  \lesssim \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}.
+\end{equation}
+Combining the decomposition~\eqref{eq:proof_5_ineq} and the oracle bound~\eqref{eq:proof_5_xi}, we finish the proof of Theorem~\ref{thm:refinement}.
 
@@
 
-\paragraph{Notations.}
+{\bf Notations.}
 \begin{enumerate}
@@
     \item For a matrix $\mA \in \bbR^{n \times m} \in \bbR^{n^K \times m^K}$, let $\mA^{\otimes K}$ denotes the kronecker product of $K$ matrices $\mA \otimes \cdots \otimes \mA$,
+    \item For a matrix $\mA$, let $\onormSize{}{\mA}_\sigma$ denote the spectral norm of matrix $\mA$, which is equal to the maximal singular value of $\mA$; let $\lambda_k(\mA)$ denote the $k$-th largest singular value of $\mA$; let $\onormSize{}{\mA}_F$ denote the Frobenius norm of matrix $\mA$.
     \item For two terms $a$ and $b$, let $a \asymp b$ if there exist two positive constants $c, C$ such that $cb \leq a\leq Cb$. 
@@
 
- \paragraph{Generalized dTBM.}
+{\bf Generalized dTBM.}
 
@@
 % \end{equation}
-\begin{align}
-    \tP(\{r_k\}) = \Bigg\{ \of{\{z_k\}, \tS, \{\mtheta_k\}}:\  &\mtheta_k \in\mathbb{R}^p_{+}, {c_1 p_k\over r_k} |z_k^{-1}(a)| \leq {c_2 p_k\over r_k}, c_3 \leq  \onorm{\mS_{k,a:}} \leq c_4 ,\\
-    &\onormSize{}{\mtheta_{k,z_k^{-1}(a)}}_1=|z_k^{-1}(a)|, \text{ for all } a \in [r_k], k\in[K]\Bigg\}. \label{eq:general_family}
-\end{align}
+\footnotesize
+\begin{align}
+    \tP(\{r_k\}) = \Bigg\{ \of{\{z_k\}, \tS, \{\mtheta_k\}}:\  &\mtheta_k \in\mathbb{R}^p_{+}, {c_1 p_k\over r_k} |z_k^{-1}(a)| \leq {c_2 p_k\over r_k}, c_3 \leq  \onorm{\mS_{k,a:}} \leq c_4 ,\onormSize{}{\mtheta_{k,z_k^{-1}(a)}}_1=|z_k^{-1}(a)|, a \in [r_k], k\in[K]\Bigg\}.
+\label{eq:general_family}
+\end{align}
+\normalsize
 % \begin{equation}
@@
 
-\paragraph{Useful Lemma for the Proof of Theorem~\ref{thm:unique}}
+{\bf Useful Lemma for the Proof of Theorem~\ref{thm:unique}}
 
@@
 
-\paragraph{Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:comp}} 
+{\bf Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:comp}} 
 \begin{defn}[Principal angles] For two matrices $\mU, \hat \mU \in \mathbb{O}^{p \times r}$ %\fixme{Miaoyan}{do we need column organization  $\mathbb{O}(p,r)?$. Yes.}
@@
 
-\begin{lem}[Angle distance and misclassification error]\label{lem:angle_mcr} Consider two membership matrices $\mM, \hat \mM \in \bbR^{p \times r}$ with corresponding assignments $z, \hat z: [p] \mapsto [r]$.
-If the angle distance $\onorm{\sin (\mM, \hat \mM)}_{\sigma} \geq c$ for a constant $c >0$, then the misclassification error $\min_{\pi \in \Pi}p\ell(\hat z,\pi \circ z )\geq 1$.
+\begin{lem}[Angle distance and misclustering error]\label{lem:angle_mcr} Consider two membership matrices $\mM, \hat \mM \in \bbR^{p \times r}$ with corresponding assignments $z, \hat z: [p] \mapsto [r]$.
+If the angle distance $\onorm{\sin (\mM, \hat \mM)}_{\sigma} \geq c$ for a constant $c >0$, then the misclustering error $\min_{\pi \in \Pi}p\ell(\hat z,\pi \circ z )\geq 1$.
 \end{lem}
@@
 
+\newpage
+
 \section*{Proof of Theorem~\ref{thm:initial}}
-\begin{proof}[Proof of Theorem 4] Recall that we prove the Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
-We drop the subscript $k$ in the matricizations $\mX_k, \mS_k, \mX_k$.
-For simplicity, let $\hat z$ denote the output of Sub-Algorithm~\hyperref[alg:main]{1} $\hat z^{(0)}$.
+\begin{proof}[Proof of Theorem 4] We prove Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
+We drop the subscript $k$ in the matricizations $\mM_k, \mX_k, \mS_k$.
+For simplicity, let $\hat z$ denote the output of Sub-Algorithm~\hyperref[alg:main]{1}, $\hat z^{(0)}$.
 
 First, by Lemma~\ref{lem:angle_gap_x}, there exists a positive constant such that $\min_{z(i) \neq z(j)} \onorm{ \mX^s_{i:} - \mX^s_{j:} } \geq c_0  \Delta_{\min}$.
-Also, by balanced assumption on $\mtheta$ and Lemma~\ref{lem:upper_mis}, we have
+By the balance assumption on $\mtheta$ and Lemma~\ref{lem:upper_mis}, we have
  \begin{equation}\label{eq:theta_bound}
-          \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2  \leq  \sum_{i \in S_0} \theta(i)^2 + 4 \sum_{i \in S} \theta(i)^2 ,
+          \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2  \leq  \sum_{i \in S_I} \theta(i)^2 + 4 \sum_{i \in S} \theta(i)^2 ,
     \end{equation}
@@
     \begin{equation}
-        S_0 = \{i: \hat z(i) = 0\}, \quad S = \{i \in S_0^c: \onorm{\hat \mx_{\hat z(i)} - \mX_{i:}^s } \geq c_0  \Delta_{\min}/2\}.
-    \end{equation}
-    On one hand, for any set $P \in [p]$ note that 
-    \begin{align}
-        \sum_{i \in P} \onorm{\mX_{i:}}^2 &= \sum_{i \in P} \onorm{ \theta(i) \mS_{z(i):} (\mTheta \mM)^{T, \otimes (K-1)} }^2 \\
-        & \geq \sum_{i \in P} \theta(i)^2 \min_{a \in [r]} \onorm{\mS_{a:}}^2 \lambda_{r}^{2(K-1)} (\mTheta \mM)\\
-        & \geq C \sum_{i \in P} \theta(i)^2 p^{K-1} r^{-(K-1)}, 
-    \end{align}
-    where the last inequality follows Lemma~\ref{lem:singular_thetam}, the assumption that $\min_{i \in [p]} \theta(i) \geq c >0$ and $\min_{a \in [r]} \onorm{\mS_{a:}} \geq c_3$ in the parameter space~\eqref{eq:family}.
+        S_0 = \{i: \onormSize{}{\hat \mX_{i:}} = 0\}, \quad S = \{i \in S_0^c: \onormSize{}{\hat \mx_{\hat z(i)} - \mX_{i:}^s } \geq c_0  \Delta_{\min}/2\}.
+    \end{equation}
+    On one hand, note that for any set $P \in [p]$,
+    \begin{align}
+        \sum_{i \in P} \onormSize{}{\mX_{i:}}^2 &= \sum_{i \in P} \onormSize{}{ \theta(i) \mS_{z(i):} (\mTheta \mM)^{T, \otimes (K-1)} }^2 \\
+        & \geq \sum_{i \in P} \theta(i)^2 \min_{a \in [r]} \onormSize{}{\mS_{a:}}^2 \lambda_{r}^{2(K-1)} (\mTheta \mM)\\
+        & \gtrsim \sum_{i \in P} \theta(i)^2 p^{K-1} r^{-(K-1)}, 
+    \end{align}
+    where the last inequality follows Lemma~\ref{lem:singular_thetam}, the assumption that $\min_{i \in [p]} \theta(i) \geq c$, and the constraint $\min_{a \in [r]} \onormSize{}{\mS_{a:}} \geq c_3$ in the parameter space~\eqref{eq:family}.
 Thus, we have
     \begin{equation}\label{eq:theta_p}
-        \sum_{i \in P} \theta(i)^2 \lesssim \sum_{i \in P} \onorm{\mX_{i:}}^2 p^{-(K-1)} r^{K-1}.
+        \sum_{i \in P} \theta(i)^2 \lesssim \sum_{i \in P} \onormSize{}{\mX_{i:}}^2 p^{-(K-1)} r^{K-1}.
     \end{equation}
@@
     \begin{align}
-         \sum_{i \in S} \onorm{\mX_{i:}}^2 & \leq 2 \sum_{i \in S} \onorm{\hat \mX_{i:}}^2  +  2\sum_{i \in S} \onorm{\hat \mX_{i:} - \mX_{i:}}^2 \label{eq:tr1}\\
-         & \leq \frac{8}{c_0^2 \Delta_{\min}^2  } \sum_{i \in S} \onorm{\hat \mX_{i:}}^2   \onorm{\hat \mx_{\hat z(i)} - \mX_{i:}^s }^2 +2 \onorm{\hat \tX - \tX}_F^2 \label{eq:defs} \\
-         & \leq \frac{16}{c_0^2 \Delta_{\min}^2  }\sum_{i \in S} \onorm{\hat \mX_{i:}}^2  \off{ \onorm{\hat \mx_{\hat z(i)} - \hat \mX_{i:}^s }^2 + \onorm{\hat \mX_{i:}^s - \mX_{i:}^s }^2 } + 2\onorm{\hat \tX - \tX}_F^2 \label{eq:tr2} \\
-         & \leq  \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 }\sum_{i \in S} \onorm{\hat \mX_{i:}}^2  \onorm{\hat \mX_{i:}^s - \mX_{i:}^s }^2  + 2\onorm{\hat \tX - \tX}_F^2 \label{eq:km} \\
-         & \leq \of{ \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 } + 2 }\onorm{\hat \tX - \tX}_F^2 \label{eq:lem1}\\
+         \sum_{i \in S} \onormSize{}{\mX_{i:}}^2 & \leq 2 \sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  +  2\sum_{i \in S} \onormSize{}{\hat \mX_{i:} - \mX_{i:}}^2 \label{eq:tr1}\\
+         & \leq \frac{8}{c_0^2 \Delta_{\min}^2  } \sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2   \onormSize{}{\hat \mx_{\hat z(i)} - \mX_{i:}^s }^2 +2 \onormSize{}{\hat \tX - \tX}_F^2 \label{eq:defs} \\
+         & \leq \frac{16}{c_0^2 \Delta_{\min}^2  }\sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  \off{ \onormSize{}{\hat \mx_{\hat z(i)} - \hat \mX_{i:}^s }^2 + \onormSize{}{\hat \mX_{i:}^s - \mX_{i:}^s }^2 } + 2\onormSize{}{\hat \tX - \tX}_F^2 \label{eq:tr2} \\
+         & \leq  \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 }\sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  \onormSize{}{\hat \mX_{i:}^s - \mX_{i:}^s }^2  + 2\onormSize{}{\hat \tX - \tX}_F^2 \label{eq:km} \\
+         & \leq \of{ \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 } + 2 }\onormSize{}{\hat \tX - \tX}_F^2 \label{eq:lem1}\\
          & \lesssim \of{ \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 } + 2 } \of{ p^{K/2}r + pr^2 + r^K } \sigma^2 \label{eq:lem_err},
     \end{align}
-    where inequalities \eqref{eq:tr1} and \eqref{eq:tr2} follow the triangle inequality, \eqref{eq:defs} follows the definition of $S$, \eqref{eq:km} follows the update rule of $k$-means in Sub-Algorithm~\hyperref[alg:main]{1}, \eqref{eq:lem1} follows Lemma~\ref{lem:norm_diff}, and the last inequality~\eqref{eq:lem_err} follows Lemma~\ref{lem:two-step_esterror}.
+    where inequalities \eqref{eq:tr1} and \eqref{eq:tr2} follow the triangle inequality, \eqref{eq:defs} follows the definition of $S$, \eqref{eq:km} follows the update rule of $k$-means in Step 5 of Sub-Algorithm~\hyperref[alg:main]{1}, \eqref{eq:lem1} follows Lemma~\ref{lem:norm_diff}, and the last inequality~\eqref{eq:lem_err} follows Lemma~\ref{lem:two-step_esterror}.
 Also, note that
     \begin{equation}\label{eq:s0}
-         \sum_{i \in S_0} \onorm{\mX_{i:}}^2 =  \sum_{i \in S_0} \onorm{\hat \mX_{i:}- \mX_{i:}}^2 \leq \onorm{\hat \tX - \tX}_F^2 \lesssim \of{ p^{K/2}r + pr^2 + r^K } \sigma^2,
+         \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2 =  \sum_{i \in S_0} \onormSize{}{\hat \mX_{i:}- \mX_{i:}}^2 \leq \onormSize{}{\hat \tX - \tX}_F^2 \lesssim \of{ p^{K/2}r + pr^2 + r^K } \sigma^2,
     \end{equation}
@@
     \begin{align}
-        \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2  &\lesssim \of{\sum_{i \in S} \onorm{\mX_{i:}}^2 + \sum_{i \in S_0} \onorm{\mX_{i:}}^2  } p^{-(K-1)} r^{K-1} \\
+        \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2  &\lesssim \of{\sum_{i \in S} \onormSize{}{\mX_{i:}}^2 + \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2  } p^{-(K-1)} r^{K-1} \\
         & \lesssim \frac{\sigma^2 r^{K-1}}{ \Delta_{\min}^2  p^{K-1} } \of{ p^{K/2}r + pr^2 + r^K }.
     \end{align}
-    With the assumption that $\min_{i \in [p]} \theta(i) \geq c >0$, we finally obtain the result
+    With the assumption that $\min_{i \in [p]} \theta(i) \geq c$, we finally obtain the result
     \begin{equation}
@@
 
-\paragraph{Useful Corollary of Theorem~\ref{thm:initial}}
-
-\begin{cor}\label{cor:L0} Suppose we have the initial assignment $z^{(0)}$ from Sub-Algorithm~\hyperref[alg:main]{1} and the identity permutation minimizes 
-the misclassification error, i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} $ $ \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
-Suppose SNR $\gtrsim p^{-K/2} \log p$.
-Then, the misclassification loss for the initialization is upper bounded as
-\begin{equation}
-    L^{(0)} \lesssim \frac{\Delta_{\min}^2}{r \log p}.
-\end{equation}
+{\bf Useful Corollary of Theorem~\ref{thm:initial}}
+
+\begin{cor}\label{cor:L0} 
+%Suppose we have the initial assignment $z^{(0)}$ from Sub-Algorithm~\hyperref[alg:main]{1} and the identity permutation minimizes the misclustering error, i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} $ $ \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
+Suppose SNR $\gtrsim p^{-K/2} \log p$ and  $\sigma^2 = 1$. Then, the misclustering loss for the initialization is upper bounded as 
+\begin{equation}
+    L^{(0)} =  \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(0)}(i) = b } \onormSize{}{ [ \mS_{ \pi^{(0)}( z(i)):}  ]^s - [ \mS_{b:}  ]^s  }^2 \lesssim \frac{\Delta_{\min}^2}{r \log p},
+\end{equation}
+where $\pi^{(0)}$ minimizes the initial misclustering error, i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} $ $ \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$.
 \end{cor}
@@
 \begin{proof}
-Note that 
-\begin{align}
-     \sum_{i \in [p]} \onorm{\mX_{i:}^s - \hat \mx_a}^2 &\leq  2\sum_{i \in [p]} \onorm{\mX_{i:}^s - \hat \mX_{i:}^s}^2 + \sum_{i \in [p]} \onorm{\hat \mX_{i:}^s - \hat \mx_a }^2 \\
-    & \leq \frac{2}{\min_{i \in [p]} \onorm{\mX_{i:}}^2}  \off{ \sum_{i \in [p]} \onorm{\mX_{i:}}^2  \onorm{\mX_{i:}^s - \hat \mX_{i:}^s}^2 +  \sum_{i \in [p]}  \onorm{\mX_{i:}}^2  \onorm{\hat \mX_{i:}^s - \hat \mx_a }^2 }\\
-    & \lesssim \frac{r^{K-1} (1 + \eta)}{p^{K-1}} \sum_{i \in [p]} \onorm{\mX_{i:}}^2  \onorm{\mX_{i:}^s - \hat \mX_{i:}^s}^2 \\
-    & \lesssim  \frac{r^{K-1} (1 + \eta)}{p^{K-1}}  \onorm{\hat  \tX - \tX}_F^2 \\
-    & \lesssim \frac{r^{K-1} (1 + \eta)}{p^{K-1}} p^{-K/2}.
-\end{align}
-
-Suppose $z^{(0)} = 1$.  Note that $\mX^s_{i:}$ only have $r$ different values. Let $\mX_{a}^s = \mX^s_{i:}$ with $z(i) = a, a \in [r]$. Note that 
-\begin{equation}
-    |z^{-1}(a) \cap (z^{(0)})^{-1} (a)| \geq |z^{-1}(a)| - p\ell(z^{(0)} - z) \gtrsim \frac{p}{r} - \frac{1}{\log p} \gtrsim \frac{p}{r},
-\end{equation}
-when $\Delta_{\min}^2 \geq p^{-K/2} \log p$. Then
-\begin{align}
-    \onorm{\mX_a^s - \hat \mx_a}^2 &= \frac{ \sum_{i \in z^{-1}(a) \cap (z^{(0)})^{-1} (a) } \onorm{ \mX_{i:}^s - \hat \mx_a}^2 }{|z^{-1}(a) \cap (z^{(0)})^{-1} (a) |} \\
-    & \lesssim \frac{ r \sum_{i \in [p]} \onorm{\mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 }{ p}\\
-    & \lesssim \frac{r^K(1 + \eta)}{p^{K}} p^{-K/2} 
-\end{align}
-
-Note that 
-\begin{align}
-    L^{(0)} &= \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \onorm{\mS_{z^{(0)}(i),:}^s - \mS_{z(i),:}^s}^2 \\
-    & \lesssim \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \onorm{\mX^s_{z^{-1}(z^{(0)}):} - \mX^s_{i,:}}^2\\
-    & \lesssim \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \off{ \onorm{ \mX^s_{z^{-1}(z^{(0)}):} - \hat \mx_{z^{(0)}(i)} }^2 + \onorm{ \hat \mx_{z^{(0)}(i)} - \mX^s_{i,:}}^2 }\\
-    & \lesssim   \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \frac{ \onorm{\hat \mX_{i:} - \mX_{i:}}^2 }{ \onorm{ \mX_{i:}}^2 } +  \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i)  \frac{r^K(1 + \eta)}{p^{K}} p^{-K/2} \\
-    & \lesssim \frac{r^K}{p^{K}} p^{-K/2} \\
-    & \lesssim \frac{\Delta_{\min}^2}{r \log p}.
-    %& \lesssim \frac{1}{p} \lambda^{-2(K-1)}_r(\mTheta \mM) \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \onorm{\mX^{(0)}_{i:}-\mX_{i:}}^2 \\
-    %& \lesssim \frac{r^{K-1}}{p^K} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \off{ \onorm{\mX^{(0)}_{i:}-\hat \mX_{i:}}^2 +  \onorm{\hat \mX_{i:}-\mX_{i:}}^2 }
-\end{align}
-where $\frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \lesssim \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i)^2$ due to the boundedness.
+
+{
+\color{blue}
+Without loss of generality, we assume $\pi^{(0)}$ is the identity mapping such that  $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
+Note that $\mX_{i:}^s$ only have $r$ different values.
+We let $\mX_a^s = \mX_{i:}^s$ for all $i$ such that $z(i) = a, a \in [r]$.
+
+Notice that 
+\begin{equation}
+    \onormSize{}{\mX_{i:}}^2 \gtrsim p^{K-1} r^{-(K-1)} \quad \text{and} \quad \onormSize{}{\mX_{i:} - \hat \mX_{i:}}^2 \leq \onormSize{}{\hat \tX - \tX}_F^2 \lesssim  p^{K/2}r + pr^2 + r^K .
+\end{equation}
+Thus, when $p$ is large enough, we have 
+\begin{align}
+    \sum_{i \in [p]} \onormSize{}{\mX_{i:}}^2 \onormSize{}{\hat \mX_i^s - \hat \mx_{z^{(0)}(i)}}^2 &\lesssim \sum_{i \in [p]} \of{\onormSize{}{\mX_{i:}}^2 - \onormSize{}{\mX_{i:} - \hat \mX_{i:}}^2} \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2\\
+   & \lesssim \sum_{i \in [p]} \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 \\
+   & \lesssim \eta  \sum_{i \in [p]} \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \mX_{i:}^s}^2 \\
+   & \lesssim \onormSize{}{\hat \tX - \tX}_F^2 \\
+   & \lesssim p^{K/2}r + pr^2 + r^K .\label{eq:cor_1}
+\end{align}
+
+Hence, we have 
+\begin{align}
+    \sum_{i \in [p]} \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 &\lesssim \sum_{i \in [p]} \theta(i)^2 \onormSize{}{\hat \mX_i^s - \hat \mx_{z^{(0)}(i)}}^2 \\
+    &\lesssim \frac{r^{K-1}}{p^{K-1}} \sum_{i \in [p]} \onormSize{}{\mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2\\
+    & \lesssim \frac{r^{K-1}}{p^{K-1}} \of{ p^{K/2}r + pr^2 + r^K },\label{eq:cor_2}
+\end{align}
+where the first inequality follows by the assumption $\min_{i \in [p]} \theta(i) \geq c$, the second inequality follows by the inequality~\eqref{eq:theta_p}, and the last inequality comes from the inequality~\eqref{eq:cor_1}.
+
+Next, we consider the following quantity,
+\begin{align}
+    \sum_{i \in [p]} \theta(i) \onormSize{}{ \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 &\lesssim \sum_{i \in [p]} \theta(i)^2 \onormSize{}{\mX_{i:}^s - \hat \mX_{i:}^s}^2 + \sum_{i \in [p]} \theta(i)^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)} }^2\\
+    & \lesssim \sum_{i \in [p]} \frac{ \theta(i)^2}{ \onormSize{}{\mX_{i:}}^2} \onormSize{}{\mX_{i:}  - \hat \mX_{i:}}^2 +  \sum_{i \in [p]} \theta(i)^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)} }^2\\
+    & \lesssim \frac{r^{K-1}}{p^{K-1}} \of{ p^{K/2}r + pr^2 + r^K} ,\label{eq:cor_3}
+\end{align}
+where the first inequality follows by the assumption of $\theta(i)$ and triangle inequality, the second inequality follows by Lemma~\ref{lem:norm_diff}, and the last inequality follows by~\eqref{eq:cor_2}.
+In addition, with Theorem~\ref{thm:initial} and the condition SNR $\gtrsim p^{-K/2} \log p$, for any $a \in [r]$
+\begin{equation}
+    |z^{-1}(a) \cap (z^{(0)})^{-1} (a)| \geq |z^{-1}(a)| - p\ell(z^{(0)} ,  z) \gtrsim \frac{p}{r} - \frac{p}{\log p} \gtrsim \frac{p}{r},
+\end{equation}
+with $p$ large enough. Then, for any $a \in [r]$, we have 
+\begin{align}
+       \onormSize{}{ \hat \mx_{a} - \mX_{a}^s  }^2  &= \frac{ \sum_{i \in z^{-1}(a) \cap (z^{(0)})^{-1} (a) } \onorm{ \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 }{|z^{-1}(a) \cap (z^{(0)})^{-1} (a) |}  \\
+       & \lesssim \frac{r}{p} \of{  \sum_{i \in [p]} \onormSize{}{ \mX_{i:}^s - \hat \mX_{i:}^s }^2 + \sum_{i \in [p]} \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 } \\
+       & \lesssim  \frac{r^{K}}{p^{K}}\of{ p^{K/2}r + pr^2 + r^K}, \label{eq:cor_4}
+\end{align}
+where the last inequality follows the inequality~\eqref{eq:cor_2}.
+
+Therefore, we finally obtain 
+\begin{align}
+     L^{(0)} &= \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(0)}(i) = b } \onormSize{}{ [ \mS_{  z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
+     & \lesssim \frac{1}{p}  \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i)  \onormSize{}{ \mX_{i:}^s -  \mX_{z^{(0)}(i)}^s  }^2\\
+     & \lesssim \frac{1}{p}  \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \of{ \onormSize{}{\mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 + \onormSize{}{ \hat \mx_{z^{(0)}(i)} - \mX_{z^{(0)}(i)}^s  }^2 }, \\
+     & \lesssim \frac{r^{K}}{p^{K}} \of{ p^{K/2}r + pr^2 + r^K}, \\
+     & \lesssim \frac{\Delta_{\min}^2}{r \log p}
+\end{align}
+where the first inequality follows by Lemma~\ref{lem:angle_gap_x}, and the third inequality follows inequalities~\eqref{eq:cor_3} and \eqref{eq:cor_4}.
+
+}
+
+% Note that 
+% \begin{align}
+%      \sum_{i \in [p]} \onorm{\mX_{i:}^s - \hat \mx_a}^2 &\leq  2\sum_{i \in [p]} \onorm{\mX_{i:}^s - \hat \mX_{i:}^s}^2 + 2\sum_{i \in [p]} \onorm{\hat \mX_{i:}^s - \hat \mx_a }^2 \\
+%     & \leq \frac{2}{\min_{i \in [p]} \onorm{\mX_{i:}}^2}  \off{ \sum_{i \in [p]} \onorm{\mX_{i:}}^2  \onorm{\mX_{i:}^s - \hat \mX_{i:}^s}^2 +  \sum_{i \in [p]}  \onorm{\mX_{i:}}^2  \onorm{\hat \mX_{i:}^s - \hat \mx_a }^2 }\\
+%     & \lesssim \frac{r^{K-1} (1 + \eta)}{p^{K-1}} \sum_{i \in [p]} \onorm{\mX_{i:}}^2  \onorm{\mX_{i:}^s - \hat \mX_{i:}^s}^2 \\
+%     & \lesssim  \frac{r^{K-1} (1 + \eta)}{p^{K-1}}  \onorm{\hat  \tX - \tX}_F^2 \\
+%     & \lesssim \frac{r^{K-1} (1 + \eta)}{p^{K-1}} p^{-K/2}.
+% \end{align}
+
+% Suppose $z^{(0)} = 1$.  Note that $\mX^s_{i:}$ only have $r$ different values. Let $\mX_{a}^s = \mX^s_{i:}$ with $z(i) = a, a \in [r]$. Note that 
+% \begin{equation}
+%     |z^{-1}(a) \cap (z^{(0)})^{-1} (a)| \geq |z^{-1}(a)| - p\ell(z^{(0)} ,  z) \gtrsim \frac{p}{r} - \frac{1}{\log p} \gtrsim \frac{p}{r},
+% \end{equation}
+% when $\Delta_{\min}^2 \geq p^{-K/2} \log p$. Then
+% \begin{align}
+%     \onorm{\mX_a^s - \hat \mx_a}^2 &= \frac{ \sum_{i \in z^{-1}(a) \cap (z^{(0)})^{-1} (a) } \onorm{ \mX_{i:}^s - \hat \mx_a}^2 }{|z^{-1}(a) \cap (z^{(0)})^{-1} (a) |} \\
+%     & \lesssim \frac{ r \sum_{i \in [p]} \onorm{\mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 }{ p}\\
+%     & \lesssim \frac{r^K(1 + \eta)}{p^{K}} p^{-K/2} 
+% \end{align}
+
+% Note that 
+% \begin{align}
+%     L^{(0)} &= \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \onorm{\mS_{z^{(0)}(i),:}^s - \mS_{z(i),:}^s}^2 \\
+%     & \lesssim \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \onorm{\mX^s_{z^{-1}(z^{(0)}):} - \mX^s_{i,:}}^2\\
+%     & \lesssim \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \off{ \onorm{ \mX^s_{z^{-1}(z^{(0)}):} - \hat \mx_{z^{(0)}(i)} }^2 + \onorm{ \hat \mx_{z^{(0)}(i)} - \mX^s_{i,:}}^2 }\\
+%     & \lesssim   \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \frac{ \onorm{\hat \mX_{i:} - \mX_{i:}}^2 }{ \onorm{ \mX_{i:}}^2 } +  \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i)  \frac{r^K(1 + \eta)}{p^{K}} p^{-K/2} \\
+%     & \lesssim \frac{r^K}{p^{K}} p^{-K/2} \\
+%     & \lesssim \frac{\Delta_{\min}^2}{r \log p}.
+%     %& \lesssim \frac{1}{p} \lambda^{-2(K-1)}_r(\mTheta \mM) \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \onorm{\mX^{(0)}_{i:}-\mX_{i:}}^2 \\
+%     %& \lesssim \frac{r^{K-1}}{p^K} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \off{ \onorm{\mX^{(0)}_{i:}-\hat \mX_{i:}}^2 +  \onorm{\hat \mX_{i:}-\mX_{i:}}^2 }
+% \end{align}
+% where $\frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \lesssim \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i)^2$ due to the boundedness.
 \end{proof}
 
-\paragraph{Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:initial}} 
+{\bf Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:initial}} 
 \begin{lem}[Basic inequality]\label{lem:norm_diff} For any two nonzero vectors $\mv_1,\mv_2$ of same dimension, we have 
@@
 \begin{equation}
-    \pad_{\mw}(\ma) = [a_1 \circ \mw_1, \ldots, a_d \circ \mw_d], \quad \text{ where} \quad  a_i \circ \mw_i = [a_i w_{i1},\ldots, a_i w_{i p_i}], \text{ for all } i \in [d]
-\end{equation}
-and $\pad: \bbR^{d}\mapsto \bbR^{\sum_{i\in[d]}}p_i$ can be viewed as an operator, and by definition, we have $\pad (\ma) \in \bbR^{\sum_{i \in[d]} p_i}$.
-Then , we have the bounds of the weighted padding vector
+    \pad_{\mw}(\ma) = [a_1 \circ \mw_1, \ldots, a_d \circ \mw_d]^T, \quad \text{ where} \quad  a_i \circ \mw_i = [a_i w_{i1},\ldots, a_i w_{i p_i}]^T, \text{ for all } i \in [d]
+\end{equation}
+and $\pad: \bbR^{d}\mapsto \bbR^{\sum_{i\in[d]}p_i}$ can be viewed as an operator.
+%, and by definition we have $\pad (\ma) \in \bbR^{\sum_{i \in[d]} p_i}$. 
+We have the bounds of the weighted padding vector
 \begin{equation}\label{eq:pad_bound}
-     \min_{i \in [d]} \onorm{\mw_i}^2 \onorm{\ma}^2 \leq \onorm{\pad_{\mw}(\ma)}^2 \leq \max_{i \in [d]} \onorm{\mw_i}^2 \onorm{\ma}^2.
+     \min_{i \in [d]} \onormSize{}{\mw_i}^2 \onormSize{}{\ma}^2 \leq \onormSize{}{\pad_{\mw}(\ma)}^2 \leq \max_{i \in [d]} \onormSize{}{\mw_i}^2 \onormSize{}{\ma}^2.
 \end{equation}
@@
 \begin{equation}
-   \frac{\min_{i \in [d]} \onorm{\mw_i}}{\max_{i \in [d]} \onorm{\mw_i} } \sin (\ma, \mb) \leq \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb)) \leq \frac{ \max_{i \in [d]} \onorm{\mw_i} }{\min_{i \in [d]} \onorm{\mw_i}} \sin (\ma, \mb).
+   \frac{\min_{i \in [d]} \onormSize{}{\mw_i}}{\max_{i \in [d]} \onormSize{}{\mw_i} } \sin (\ma, \mb) \leq \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb)) \leq \frac{ \max_{i \in [d]} \onormSize{}{\mw_i} }{\min_{i \in [d]} \onormSize{}{\mw_i}} \sin (\ma, \mb).
 \end{equation}
@@
 
-\begin{proof}[Proof of Lemma~\ref{lem:pad}] We prove the two inequalities separately with the similar idea.
-
-First, we prove the inequality $\sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb)) \leq \frac{ \max_{i \in [d]} \onorm{\mw_i}^2 }{\min_{i \in [d]} \onorm{\mw_i}^2} \sin (\ma, \mb)$. Decomposing $\mb$, we have
-\begin{equation}
-    \mb = \cos (\ma, \mb) \frac{\onorm{\mb}}{\onorm{\ma}} \ma + \sin (\ma, \mb) \frac{\onorm{\mb}}{\onorm{\ma^{\perp}}}\ma^{\perp},
+\begin{proof}[Proof of Lemma~\ref{lem:pad}] We prove the two inequalities separately with similar ideas.
+
+First, we prove the inequality $\sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb)) \leq \frac{ \max_{i \in [d]} \onormSize{}{\mw_i}^2 }{\min_{i \in [d]} \onormSize{}{\mw_i}^2} \sin (\ma, \mb)$.
+Decomposing $\mb$, we have
+\begin{equation}
+    \mb = \cos (\ma, \mb) \frac{\onormSize{}{\mb}}{\onormSize{}{\ma}} \ma + \sin (\ma, \mb) \frac{\onormSize{}{\mb}}{\onormSize{}{\ma^{\perp}}}\ma^{\perp},
 \end{equation}
@@
 \begin{equation}\label{eq:pad_second_ineq}
-    \pad_{\mw}(\mb) = \cos (\ma, \mb) \frac{\onorm{b}}{\onorm{a}} \pad_{\mw}(\ma) + \sin (\ma, \mb) \frac{\onorm{b}}{\onorm{a^{\perp}}} \pad_{\mw}(\ma^{\perp}).
+    \pad_{\mw}(\mb) = \cos (\ma, \mb) \frac{\onormSize{}{b}}{\onormSize{}{a}} \pad_{\mw}(\ma) + \sin (\ma, \mb) \frac{\onormSize{}{b}}{\onormSize{}{a^{\perp}}} \pad_{\mw}(\ma^{\perp}).
 \end{equation}
@@
 \begin{align}
-    \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb))  &\leq \frac{  \onorm{\mb} \onorm{\pad_{\mw}(\ma^{\perp})} }{ \onorm{\ma^{\perp}} \onorm{ \pad_{\mw}(\mb)}} \sin (\ma, \mb)  \\
-    & \leq  \frac{ \max_{i \in [d]} \onorm{\mw_i} }{\min_{i \in [d]} \onorm{\mw_i}} \sin (\ma, \mb),
+    \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb))  &\leq \frac{  \onormSize{}{\mb} \onormSize{}{\pad_{\mw}(\ma^{\perp})} }{ \onormSize{}{\ma^{\perp}} \onormSize{}{ \pad_{\mw}(\mb)}} \sin (\ma, \mb)  \\
+    & \leq  \frac{ \max_{i \in [d]} \onormSize{}{\mw_i} }{\min_{i \in [d]} \onormSize{}{\mw_i}} \sin (\ma, \mb),
 \end{align}
@@
 
-Next, we prove $\frac{\min_{i \in [d]} \onorm{\mw_i}^2}{\max_{i \in [d]} \onorm{\mw_i}^2 } \sin (\ma, \mb) \leq \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb))$.
+Next, we prove $\frac{\min_{i \in [d]} \onormSize{}{\mw_i}^2}{\max_{i \in [d]} \onormSize{}{\mw_i}^2 } \sin (\ma, \mb) \leq \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb))$.
 With the decomposition of $\pad_{\mw}(\mb)$ and the inverse weighted padding operator, we have
 \begin{align}
-    \mb &=  \cos(\pad_{\mw}(\ma), \pad_{\mw}(\mb)) \frac{\onorm{\pad_{\mw}(\mb)}}{\onorm{\pad_{\mw}(\ma)}} \ma +\\
-    & \quad \quad \quad \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb))\frac{\onorm{\pad_{\mw}(\mb)}}{\onorm{(\pad_{\mw}(\ma))^{\perp}}} \pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp}).
+    \mb &=  \cos(\pad_{\mw}(\ma), \pad_{\mw}(\mb)) \frac{\onormSize{}{\pad_{\mw}(\mb)}}{\onormSize{}{\pad_{\mw}(\ma)}} \ma +  \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb))\frac{\onormSize{}{\pad_{\mw}(\mb)}}{\onormSize{}{(\pad_{\mw}(\ma))^{\perp}}} \pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp}).
 \end{align}
@@
 \begin{align}
-    \sin(\ma, \mb) &\leq \frac{ \onorm{\pad_{\mw}(\mb) } \onorm{\pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp})  } }{ \onorm{(\pad_{\mw}(\ma))^{\perp} } \onorm{\mb} }  \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb))\\
-    & \leq \frac{ \max_{i \in [d]} \onorm{\mw_i} }{\min_{i \in [d]} \onorm{\mw_i}} \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb)),
+    \sin(\ma, \mb) &\leq \frac{ \onormSize{}{\pad_{\mw}(\mb) } \onormSize{}{\pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp})  } }{ \onormSize{}{(\pad_{\mw}(\ma))^{\perp} } \onormSize{}{\mb} }  \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb))\\
+    & \leq \frac{ \max_{i \in [d]} \onormSize{}{\mw_i} }{\min_{i \in [d]} \onormSize{}{\mw_i}} \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb)),
 \end{align}
 where the second inequality follows by applying the property~\eqref{eq:pad_bound} to vectors $\mb, \pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp})$.
-
 \end{proof}
@@
 Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying~\eqref{eq:degree}.
-Then the angle gap in $\tX$ is approximate to angle gap in $\tS$, i.e.,
-\begin{equation}
-    \Delta_{\min} \lesssim  \min_{(i,j): z(i) \neq z(j)} \onorm{\mX_{i:}^s - \mX_{j:}^s}. 
-\end{equation}
+Then the angle gap in $\tX$ is approximate to angle gap in $\tS$, i.e., for all $i,j$ such that $z(i) \neq z(j)$
+% \begin{equation}
+%     \Delta_{\min} \leq \min_{(i,j): z(i) \neq z(j)}  \onormSize{}Size{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s}  \lesssim \min_{(i,j): z(i) \neq z(j)} \onormSize{}Size{}{\mX_{i:}^s - \mX_{j:}^s}. 
+% \end{equation}
+\begin{equation}
+    \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \gtrsim  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
+\end{equation}
+
 \end{lem}
@@
 Note that the vector $\mS_{z(i):}$ can be folded to a tensor $\tS' =\entry{\tS'_{a_2,\ldots,a_K}} \in \bbR^{r^{K-1}}$, i.e., $\text{vec}(\tS') = \mS_{z(i):}$.
-Let the weight vector $\mw_{a_2 \cdots a_K}$ correspond to the elements $\tS'_{a_2,\ldots,a_K}$ as
-\begin{equation}
-    \mw_{a_2 \cdots a_K} = [ \mtheta_{z^{-1} (a_2)} \otimes \cdots \otimes \mtheta_{z^{-1} (a_K)}] \in \bbR^{|z^{-1} (a_2)| \times \cdots \times |z^{-1} (a_K)|},
+Let the weight vectors $\mw_{a_2, \cdots, a_K}$ correspond to the elements in $\tS'_{a_2,\ldots,a_K}$ as
+\begin{equation}
+    \mw_{a_2 \cdots a_K} = [ \mtheta_{z^{-1} (a_2)}^T \otimes \cdots \otimes \mtheta_{z^{-1} (a_K)}^T] \in \bbR^{|z^{-1} (a_2)| \times \cdots \times |z^{-1} (a_K)|},
 \end{equation}
 for all $a_k \in [r], k = 2,\ldots, K$, where $\otimes$ denotes the Kronecker product.
-Therefore, we have  $\mX_{i:} = \theta(i) \pad_{\mw}(\mS_{z(i):})$ where $\mw = \{ \mw_{a_2 \cdots a_K}\}$.
-Specifically, we have $\onorm{\mw_{a_2,\ldots,a_K}}^2 = \prod_{k = 2}^K \onorm{ \mtheta_{z^{-1}(a_k)}}^2$, and by the balanced assumption we have
+Therefore, we have  $\mX_{i:} = \theta(i) \pad_{\mw}(\mS_{z(i):})$ where $\mw = \{ \mw_{a_2, \cdots, a_K}\}_{a_k \in [r], k\in [K]/\{1\}}$.
+Specifically, we have $\onormSize{}{\mw_{a_2, \ldots, a_K}}^2 = \prod_{k = 2}^K \onormSize{}{ \mtheta_{z^{-1}(a_k)}}^2$, and by the balanced assumption~\eqref{eq:degree} we have
 \begin{equation}\label{eq:pad_balance}
-    \max_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}}^2 = (1 + o(1))  \min_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}}^2.
+    \max_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2 = (1 + o(1))  \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2.
 \end{equation}
@@
     \ang{\mX_{i:},\mX_{j:}} &= \theta(i) \theta(j) \ang{ \pad_{\mw}(\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):} )}\\
-    &= \theta(i) \theta(j) \min_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}}^2 \ang{\mS_{z(i):}, \mS_{z(j):}}(1 + o(1)).
+    &= \theta(i) \theta(j) \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2 \ang{\mS_{z(i):}, \mS_{z(j):}}(1 + o(1)).
 \end{align}
@@
     \item \textbf{Case 1:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} \leq 0$.
-Then, we also have $\ang{\mX_{i:},\mX_{j:}} \leq 0$, which implies $\onorm{\mX_{i:}^s -  \mX_{j:}^s} \geq \sqrt{2}$.
-Note that $\onorm{\mS_{z(i):}^s - \mS_{z(j):}^s} \leq 2$.
-We have $\onorm{\mX_{i:}^s -  \mX_{j:}^s} \gtrsim \onorm{\mS_{z(i):}^s - \mS_{z(j):}^s}$.
+Then, we also have $\ang{\mX_{i:},\mX_{j:}} \leq 0$, which implies $\onormSize{}{\mX_{i:}^s -  \mX_{j:}^s} \geq \sqrt{2}$.
+Note that $\onormSize{}{\mS_{z(i):}^s - \mS_{z(j):}^s} \leq 2$ by the definition of angle gap.
+We have $\onormSize{}{\mX_{i:}^s -  \mX_{j:}^s} \gtrsim \onormSize{}{\mS_{z(i):}^s - \mS_{z(j):}^s}$.
     \item \textbf{Case 2:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} > 0$.
@@
     \begin{align}
-        \onorm{\mS_{z(i):}^s - \mS_{z(j):}^s} &= \sqrt{ 1 - \cos(\mS_{z(i):}, \mS_{z(j):})}  \\
+        \onormSize{}{\mS_{z(i):}^s - \mS_{z(j):}^s} &= \sqrt{ 1 - \cos(\mS_{z(i):}, \mS_{z(j):})}  \\
         & \lesssim \sin (\mS_{z(i):}, \mS_{z(j):})\\
-        & \leq \frac{ \max_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}} }{  \min_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}} }  \sin ( \pad_{\mw} (\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):})) \\
-        & \leq (1 + o(1)) \onorm{\mX_{i:}^s -  \mX_{j:}^s},
-    \end{align}
-    where the second inequality follows by Lemma~\ref{lem:pad}, and the last inequality follows by the balanced assumption~\eqref{eq:pad_balance} and Lemma~\ref{lem:norm_diff}.
+        & \leq \frac{ \max_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}} }{  \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}} }  \sin ( \pad_{\mw} (\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):})) \\
+        & \leq (1 + o(1)) \onormSize{}{\mX_{i:}^s -  \mX_{j:}^s},
+    \end{align}
+    where the second inequality follows by Lemma~\ref{lem:pad}, and the last inequality follows by the balanced weight~\eqref{eq:pad_balance} and Lemma~\ref{lem:norm_diff}.
 \end{enumerate}
 
-Hence, we conclude that 
-\begin{equation}
-   \min_{(i,j): z(i) \neq z(j)} \onorm{\mX_{i:}^s - \mX_{j:}^s} \gtrsim \min_{(i,j): z(i) \neq z(j)}  \onorm{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
-\end{equation}
-
+Hence, we conclude that for all $i,j$ such that $z(i) \neq z(j)$
+% \begin{equation}
+%   \min_{(i,j): z(i) \neq z(j)} \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \gtrsim \min_{(i,j): z(i) \neq z(j)}  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
+% \end{equation}
+\begin{equation}
+    \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \gtrsim  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
+\end{equation}
 \end{proof}
@@
 
-\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the assumption that $\min_{i \in [p]} \theta(i) \geq c_0 >0$, we have the minimal singular value of $\mTheta \mM$ bounded as
+\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the assumption that $\min_{i \in [p]} \theta(i) \geq c$, we have the singular values of $\mTheta \mM$ are bounded as
 \begin{equation}\label{eq:membership2}
-     \sqrt{p/r} \leq \sqrt{ \min_{a \in [r]}\onorm{ \mtheta_{ z^{-1}(a)} }^2 }  \lesssim \lambda_{r}(\mTheta \mM) \leq \onorm{\mTheta \mM}_{\sigma} \leq\sqrt{ \max_{a \in [r]}\onorm{ \mtheta_{ z^{-1}(a)} }^2 } \lesssim  p /r .
+     \sqrt{p/r} \lesssim \sqrt{ \min_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 }  \leq \lambda_{r}(\mTheta \mM) \leq \onormSize{}{\mTheta \mM}_{\sigma} \leq\sqrt{ \max_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 } \lesssim  p /r .
 \end{equation}
@@
 \begin{equation}
-    (\mTheta\mM)^T \mTheta \mM = \mD, \quad \text{with} \quad \mD = \text{diag}(D_1,\ldots, D_r), \ D_a = \onorm{\mtheta_{z^{-1}(a)}}^2, a \in [r].
+    (\mTheta\mM)^T \mTheta \mM = \mD, \quad \text{with} \quad \mD = \text{diag}(D_1,\ldots, D_r), \ D_a = \onormSize{}{\mtheta_{z^{-1}(a)}}^2, a \in [r].
 \end{equation}
@@
 \begin{equation}
-     \sqrt{\min_{a \in [r]}\onorm{ \mtheta_{ z^{-1}(a)} }^2 } \leq \lambda_r(\mTheta \mM) \leq \onorm{\mTheta\mM}_{\sigma} \leq \sqrt{ \max_{a \in [r]}\onorm{ \mtheta_{ z^{-1}(a)} }^2}.
-\end{equation}
-
-Given that $\min_{i \in [p]}\theta(i) \geq c_0$, we have  
-\begin{equation}
-   \min_{a \in [r]} \onorm{ \mtheta_{ z^{-1}(a)} }^2 \geq c_0^2 \min_{a \in [r]} |z^{-1}(a)| \gtrsim \frac{p}{r},
+     \sqrt{\min_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 } \leq \lambda_r(\mTheta \mM) \leq \onormSize{}{\mTheta\mM}_{\sigma} \leq \sqrt{ \max_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2}.
+\end{equation}
+
+Given that $\min_{i \in [p]}\theta(i) \geq c$, we have  
+\begin{equation}
+   \min_{a \in [r]} \onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 \geq c^2 \min_{a \in [r]} |z^{-1}(a)| \gtrsim \frac{p}{r},
 \end{equation}
@@
 \begin{equation}
-    \sqrt{\max_{a \in [r]}\onorm{ \mtheta_{ z^{-1}(a)} }^2 } \leq \max_{a \in [r]} \sqrt{ \onorm{ \mtheta_{ z^{-1}(a)} }^2_1} \lesssim \frac{p}{r}.
+    \sqrt{\max_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 } \leq \max_{a \in [r]} \sqrt{ \onormSize{}{ \mtheta_{ z^{-1}(a)} }^2_1} \lesssim \frac{p}{r}.
 \end{equation}
@@
 \begin{lem}[Singular-value-gap-free tensor estimation error bound]\label{lem:two-step_esterror} Let an order-$K$ tensor $\tA = \tX + \tZ \in \bbR^{p \times \cdots \times p}$, and $\tX$ has tucker rank $(r,...r)$ and $\tZ$ has independent sub-Gaussian entries with parameter $\sigma^2$.
-Let $\hat \tX$ denote the two-step estimated tensor in line 2 of Sub-Algorithm 1 in the main paper.
+Let $\hat \tX$ denote the two-step estimated tensor in Step 2 of Sub-Algorithm~\hyperref[alg:main]{1} in the main paper.
 Then with probability at least $1 - C \exp\of{- cp }$, we have
 \begin{align}
-    \onorm{\hat \tX - \tX}_F^2 \leq C \sigma^2 \of{ p^{K/2}r + pr^2 + r^K }.
-\end{align}
+    \onormSize{}{\hat \tX - \tX}_F^2 \leq C \sigma^2 \of{ p^{K/2}r + pr^2 + r^K },
+\end{align}
+where $C, c$ are some positive constants.
 \end{lem}
@@
 \begin{proof}
-See Proposition 1 in \cite{han2020exact}.
+See \citet[Proposition 1]{han2020exact}.
 \end{proof}
 
-\begin{lem}[Upper bound of misclassification error]\label{lem:upper_mis} Let $z: [p] \mapsto [r]$ be a clustering assignment such that $|z^{-1}(a)| \asymp p/r$ for all $a \in [r]$.
+\begin{lem}[Upper bound of misclustering error]\label{lem:upper_mis} Let $z: [p] \mapsto [r]$ be a clustering assignment such that $|z^{-1}(a)| \asymp p/r$ for all $a \in [r]$.
 Let node $i$ corresponds to a vector $\mx_i  = \theta(i) \mv_{z(i)} \in \bbR^d$ where $\{\mv_a\}_{a = 1}^r$ are the cluster centers and $\mtheta = \entry{\theta(i)} \in \bbR^p_+$ is the positive degree heterogeneity.
-Assume that  $\mtheta$ satisfies the balanced assumption such that ${\max_{a \in [r]} \onorm{\mtheta_{z^{-1}(a)}}^2 \over \min_{a \in [r]} \onorm{\mtheta_{z^{-1}(a)}}^2 }= 1 + o(1)$.
-Consider an arbitrary estimate $\hat z$ with $\hat \mx_i = \hat \mv_{\hat z(i)}$.
+Assume that  $\mtheta$ satisfies the balanced assumption~\eqref{eq:degree} such that ${\max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 \over \min_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 }= 1 + o(1)$.
+Consider an arbitrary estimate $\hat z$ with $\hat \mx_i = \hat \mv_{\hat z(i)}$ for all $ i \in S$.
 Then, if
 \begin{equation}
-    \min_{a \neq b \in [r]} \onorm{\mv_a - \mv_b} \geq 2c,
+    \min_{a \neq b \in [r]} \onormSize{}{\mv_a - \mv_b} \geq 2c,
 \end{equation}
@@
 \end{equation}
-where 
-\begin{equation}
-    S_0 = \{i: \hat z(i) = 0\} \quad  \text{and}  \quad S = \{i \in S_0^c: \onorm{\hat \mx_i - \mx_i } \geq c \}.
+where $S_0$ is defined in Step 3 of Sub-Algorithm~\hyperref[alg:main]{1} and
+\begin{equation}
+   S = \{i \in S_0^c: \onormSize{}{\hat \mx_i - \mx_i } \geq c \}.
 \end{equation}
@@
 For each cluster $u\in[r]$, we use $C_u$ to collect the subset of points for which the estimated and true positions $\hat \mx_i, \mx_i$ are within distance $c$. Specifically, define
-For each cluster $u \in [r]$, we define
-
-\begin{equation}
-    C_u = \{ i \in z^{-1}(u) \cap S_0^c: \onorm{\hat \mx_i - \mx_i} < c \},
+\begin{equation}
+    C_u = \{ i \in z^{-1}(u) \cap S_0^c: \onormSize{}{\hat \mx_i - \mx_i} < c \},
 \end{equation}
@@
 \end{align}
-Then, we have 
-\begin{equation}
-    \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \leq \sum_{i \in S_0} \theta(i)^2 + \sum_{i \in S} \theta(i)^2 + \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2.
-\end{equation}
-Thus, we only need to bound $\sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2$ to finish the proof. Note that we have $|R_3| \leq |R_1|$ by Lemma 6 in \citep{gao2018community}. We have 
-\begin{align}
-    \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2 &\leq |R_3| \max_{a \in [r]} \onorm{\mtheta_{z^{-1}(a)}}^2 \\
-    & \leq |R_1| \max_{a \in [r]} \onorm{\mtheta_{z^{-1}(a)}}^2  \\
-    & \leq {\max_{a \in [r]} \onorm{\mtheta_{z^{-1}(a)}}^2 \over \min_{a \in [r]} \onorm{\mtheta_{z^{-1}(a)}}^2 } \sum_{i \in \cup_{u \in R_1} z^{-1}(u)} \theta(i)^2\\
+Following \citet[Lemma 6]{gao2018community}, we have 
+\begin{equation}
+    \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \leq \sum_{i \in S_0} \theta(i)^2 + \sum_{i \in S} \theta(i)^2 + \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2,
+\end{equation}
+and $|R_3| \leq |R_1|$. We only need to bound $\sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2$ to finish the proof. Note that 
+\begin{align}
+    \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2 &\leq |R_3| \max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 \\
+    & \leq |R_1| \max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2  \\
+    & \leq {\max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 \over \min_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 } \sum_{i \in \cup_{u \in R_1} z^{-1}(u)} \theta(i)^2\\
     & \leq 2 \sum_{i \in S} \theta(i)^2,
 \end{align}
-where the second inequality follows the conclusion $|R_3| \leq |R_1|$ from the proof of Lemma 6 in \citep{gao2018community}, and the last inequality holds by the balanced assumption on $\mtheta$ when $p$ is large enough, and the fact that $ \cup_{u \in R_1} z^{-1}(u) \subset S$.
-
+where the last inequality holds by the balanced assumption on $\mtheta$ when $p$ is large enough, and the fact that $ \cup_{u \in R_1} z^{-1}(u) \subset S$.
 \end{proof}
+
+\newpage
 
@@
 \begin{proof}[Proof of Theorem~\ref{thm:refinement}]
-Recall that we prove Theorem~\ref{thm:refinement} under the dTBM~\ref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
+We prove Theorem~\ref{thm:refinement} under the dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
 We drop the subscript $k$ in the matricizations $\mM_k, \mS_k, \mX_k$.
-Without loss of generality, we assume that the noise variance $\sigma = 1$ and the identity permutation minimizes the initial misclassification error, i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
+Without loss of generality, we assume that the variance $\sigma = 1$ and the identity permutation minimizes the initial misclustering error, i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
 
@@
 
-\paragraph{Step 1:} We first introduce additional notations and the necessary condition to complete the proof.
+{\bf Step 1 (Notations and conditions):} We first introduce additional notations and the necessary condition to complete the proof.
 We will verify that the conditions hold in our context under high probability in the last step of the proof.
@@
 
-\item Norms. Let $\onorm{\mA}_\sigma$ denote the spectral norm of matrix $\mA$, which is equal to the maximal singular value of $\mA$; let $\onorm{\mA}_F$ denote the Frobenius norm of matrix $\mA$.
+%\item Norms. 
 \item Projection.
@@
 
-    \item Normalized membership matrix
+    \item Normalized membership matrices
     \begin{align}
@@
     \end{align}
-    and dual normalized membership matrix
-    \begin{align}
-         \mV &= \mW^{\otimes (K-1)}, \quad \mV^{(t)} =\of{ \mW^{(t)}}^{\otimes (K-1)}
-    \end{align}
-    \item Estimator of $\tS$ in the $t$-th iteration, denoted $\tS^{(t)}$ and the oracle estimator of $\tS$ given true assignment $z$, denoted $\tilde \tS$,
+    and the dual normalized membership matrices
+    \begin{align}
+         \mV &= \mW^{\otimes (K-1)}, \quad \mV^{(t)} =\of{ \mW^{(t)}}^{\otimes (K-1)}.
+    \end{align}
+    \item Estimator of $\tS$ in the $t$-th iteration, $\tS^{(t)}$, and the oracle estimator of $\tS$ given true assignment $z$, $\tilde \tS$,
     \begin{align}
@@
     \begin{align}
-        &\mS = \mat_1(\tS), \quad \mS^{(t)} = \mat_1(\tS^{(t)}), \quad \tilde \mS = \mat_1(\tilde \tS)\\
-        &\mY = \mat_1(\tY), \quad \mX = \mat_1(\tX), \quad \mE = \mat_1(\tE).
-    \end{align}
-    \item The angle-based misclassification loss in the $t$-th iteration %\fixme{Jiaxin}{Need $\theta(i)$ in the definition}
-    \begin{equation}
-         L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
+        &\mS = \mat(\tS), \quad \mS^{(t)} = \mat(\tS^{(t)}), \quad \tilde \mS = \mat(\tilde \tS)\\
+        &\mY = \mat(\tY), \quad \mX = \mat(\tX), \quad \mE = \mat(\tE).
+    \end{align}
+    \item The angle-based misclustering loss in the $t$-th iteration %\fixme{Jiaxin}{Need $\theta(i)$ in the definition}
+    \begin{equation}
+         L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2,
     \end{equation}
@@
     \begin{align}
-         \xi & = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \Big\{\ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s} \\
-         & \hspace{3cm}\leq - \frac{ \theta(i) m}{4} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 \Big\} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{ b:}  }^s  }^2,
-    \end{align}
-    where $m$ is a positive constant related to $c_3$ and is defined in~\eqref{eq:xv_lower}.
+         \xi & = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \Big\{\ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s}  \leq - \frac{ \theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \Big\} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2,
+    \end{align}
+    where $m$ is a positive universal constant defined in~\eqref{eq:xv_lower}.
 \end{enumerate}
@@
 \begin{equation}\label{eq:cond1}
-    \onorm{\mE \mV}_\sigma \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}} \of{ p ^{1/2}+ r^{(K-1)/2}}, \quad \onorm{\mE \mV}_F \lesssim \sqrt{\frac{ r^{2(K-1)}}{p^{K-2}}},\quad \onorm{ \mW_{a:}^T \mE \mV} \lesssim \frac{r^K}{p^{K/2}},
+    \onormSize{}{\mE \mV}_\sigma \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}} \of{ p ^{1/2}+ r^{(K-1)/2}}, \quad \onormSize{}{\mE \mV}_F \lesssim \sqrt{\frac{ r^{2(K-1)}}{p^{K-2}}},\quad \onormSize{}{ \mW_{a:}^T \mE \mV} \lesssim \frac{r^K}{p^{K/2}},
 \end{equation}
 % \begin{equation}\label{eq:cond11}
-%     , \quad \onorm{ \mW_{a:}^{(t),T} \mE \mV} \lesssim \frac{r^K}{p^{K/2}},
+%     , \quad \onormSize{}{ \mW_{a:}^{(t),T} \mE \mV} \lesssim \frac{r^K}{p^{K/2}},
 % \end{equation}
 \begin{equation}\label{eq:cond2}
-    \sup_{ \mU_k \in \mathbb{O}_{p, r}, k = 2,\ldots, K } \onorm{ \mE (\mU_{2} \otimes \cdots \otimes \mU_K)}_\sigma \lesssim \of{ \sqrt{r^{K-1}} + K\sqrt{pr}},
+    \sup_{ \mU_k \in \mathbb{O}_{p, r}, k = 2,\ldots, K } \onormSize{}{ \mE (\mU_{2} \otimes \cdots \otimes \mU_K)}_\sigma \lesssim \of{ \sqrt{r^{K-1}} + K\sqrt{pr}},
 \end{equation}
@@
 \begin{equation}\label{eq:cond3}
-    \sup_{ \mU_k \in \mathbb{O}_{p, r}, k = 2,\ldots, K } \onorm{ \mE (\mU_{2} \otimes \cdots \otimes \mU_K)}_F \lesssim \of{ \sqrt{pr^{K-1}} + K\sqrt{pr}},
+    \sup_{ \mU_k \in \mathbb{O}_{p, r}, k = 2,\ldots, K } \onormSize{}{ \mE (\mU_{2} \otimes \cdots \otimes \mU_K)}_F \lesssim \of{ \sqrt{pr^{K-1}} + K\sqrt{pr}},
 \end{equation}
@@
 
- Particularly, the inequality~\eqref{eq:cond1} holds by replacing $\mV$ to $\mV^{(t)}$ and $\mW_{:a}$ to $\mW_{:a}^{(t),T}$ when initialization condition~\eqref{eq:cond_intial} holds.
+Further, inequality~\eqref{eq:cond1} holds by replacing $\mV$ to $\mV^{(t)}$ and $\mW_{:a}$ to $\mW_{:a}^{(t),T}$ when initialization condition~\eqref{eq:cond_intial} holds.
 \end{condition}
 
-\paragraph{Step 2:} Next, we derive upper bound of $L^{(t+1)}$ for $t = 0 ,1, \ldots, T-1$. By Sub-Algorithm~\hyperref[alg:main]{2}, we update the assignment in $t$-th iteration as 
-    \begin{equation}
-        z^{(t+1)}(i) = \argmin_{a \in [r]} \onorm{ \off{ \mY_{i:} \mV^{(t)}  }^s - \off{\mS_{a:}^{(t)}}^s }^2,
-    \end{equation}
-    following the fact that $\mat_1(\tY^\text{d}) = \mY \mV^{(t)}$, where $\tY^\text{d}$ is the reduced tensor defined in Sub-Algorithm~\hyperref[alg:main]{2}.
+{\bf Step 2 (Misclustering loss decomposition):} Next, we derive the upper bound of $L^{(t+1)}$ for $t = 0 ,1, \ldots, T-1$.
+By Sub-Algorithm~\hyperref[alg:main]{2}, we update the assignment in $t$-th iteration via
+    \begin{equation}
+        z^{(t+1)}(i) = \argmin_{a \in [r]} \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{a:}^{(t)}]^s }^2,
+    \end{equation}
+    following the facts that $\onormSize{}{\ma^s - \mb^s}^2 = 1 - \cos(\ma, \mb)$ for vectors $\ma,\mb$ of same dimension and $\mat(\tY^\text{d}) = \mY \mV^{(t)}$ where $\tY^\text{d}$ is the reduced tensor defined in Step 8 of Sub-Algorithm~\hyperref[alg:main]{2}.
 Then the event $z^{(t+1)}(i) = b$ implies
-    \begin{equation}
-        \onorm{ \off{ \mY_{j:} \mV^{(t)}  }^s - \off{\mS_{b:}^{(t)}}^s }^2 \leq \onorm{ \off{ \mY_{j:} \mV^{(t)}  }^s - \off{\mS_{z(j):}^{(t)}}^s }^2.
-    \end{equation}
-    Arranging the terms, we obtain the decomposition
+    \begin{equation}\label{eq:event}
+        \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{b:}^{(t)}]^s }^2 \leq \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{z(i):}^{(t)}]^s }^2.
+    \end{equation}
+    {\color{blue} Note that the event~\eqref{eq:event} also holds for the degenerate entity $i$ with $\onormSize{}{\mY_{i:} \mV^{(t)}} = 0$ due to the convention that $\ma^s = 0$ if $\ma = 0$. }
+    Arranging the terms in \eqref{eq:event}, we obtain the decomposition
     \begin{equation}\label{eq:decomp}
-        2 \ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s } \leq \onorm{\mX_{i:} \mV^{(t)}} \of{ - \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 + G_{ib}^{(t)} + H_{ib}^{(t)} } +  F_{ib}^{(t)},
+        2 \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s } \leq \onormSize{}{\mX_{i:} \mV^{(t)}} \of{ - \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 + G_{ib}^{(t)} + H_{ib}^{(t)} } +  F_{ib}^{(t)},
     \end{equation}
@@
     \begin{align}
-        F_{ib}^{(t)} &= 2 \ang{\mE_{i:} \mV^{(t)}, \of{ \off{  \tilde \mS_{z(i):} }^s  -  \off{  \mS_{z(i):}^{(t)} }^s  }  - \of{  \off{  \tilde \mS_{b:} }^s  -  \off{  \mS_{b:}^{(t)} }^s  }  } \\
-        & \quad \quad \quad \quad \quad + 2 \ang{ \mE_{i:} \of{ \mV - \mV^{(t)} }, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s  },\\
-        G_{ib}^{(t)} &=   \of{ \onorm{ \off{\mX_{i:} \mV^{(t)}}^s -  \off{  \mS_{z(i):}^{(t)}  }^s}^2 -  \onorm{ \off{\mX_{i:} \mV^{(t)}}^s -  \off{  \mW_{:z(i)}^T \mY \mV^{(t)} }^s}^2}   \\
-        & \quad \quad \quad \quad \quad -   \of{ \onorm{ \off{\mX_{i:} \mV^{(t)}}^s -  \off{  \mS_{b:}^{(t)}  }^s}^2 -  \onorm{ \off{\mX_{i:} \mV^{(t)}}^s -  \off{  \mW_{:b}^T \mY \mV^{(t)} }^s}^2 }, \\
-        H_{ib}^{(t)} &=   \onorm{ \off{\mX_{i:} \mV^{(t)}}^s -  \off{  \mW_{:z(i)}^T \mY \mV^{(t)} }^s}^2 - \onorm{ \off{\mX_{i:} \mV^{(t)}}^s -  \off{  \mW_{:b}^T \mY \mV^{(t)} }^s}^2  + \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 .
+        F_{ib}^{(t)} &= 2 \ang{\mE_{i:} \mV^{(t)}, \of{ [  \tilde \mS_{z(i):} ]^s  -  [  \mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  -  [  \mS_{b:}^{(t)} ]^s  }  }  + 2 \ang{ \mE_{i:} \of{ \mV - \mV^{(t)} }, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }, \label{eq:f}\\
+        G_{ib}^{(t)} &=   \of{ \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2}   \\
+        & \quad \quad \quad \quad \quad -   \of{ \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{b:}^{(t)}  ]^s}^2 -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 }, \label{eq:g} \\
+        H_{ib}^{(t)} &=   \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2  + \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 .
+\label{eq:h}
     \end{align}
@@
 Add Lemma 4.}
-    Then, we have the following upper bound 
-    \begin{align}
-         &\ind \offf{ z^{(t+1)}(j) = b  }\\ & \quad \quad \quad \leq \ind \offf{ z^{(t+1)}(j) = b ,   \ang{ \mE_{j:} \mV, \off{  \tilde \mS_{z(j):} }^s - \off{  \tilde \mS_{b:} }^s }  \leq - \frac{1}{4} \onorm{\mX_{i:} \mV^{(t)}} \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^2 } \\
-         & \quad \quad \quad  \quad \quad + \ind \offf{z^{(t+1)}(j) = b, \frac{1}{2}\onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^2 \leq \onorm{\mX_{i:} \mV^{(t)}}^{-1} F_{jb}^{(t)} + G_{jb}^{(t)} + H_{jb}^{(t)} }.
+    Then, we upper bound the event
+    \begin{align}
+         \ind \offf{ z^{(t+1)}(i) = b  }& \leq \ind \offf{ z^{(t+1)}(i) = b ,   \ang{ \mE_{j:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{1}{4} \onormSize{}{\mX_{i:} \mV^{(t)}} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \\
+         & + \ind \offf{z^{(t+1)}(i) = b, \frac{1}{2}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq \onormSize{}{\mX_{i:} \mV^{(t)}}^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }.
 \label{eq:ind_b}
     \end{align}
-
    Note that 
     \begin{align}
-        \onorm{\mX_{i:} \mV^{(t)}} &= \onorm{ \mX_{i:} \mW^{(t), \otimes^{K-1}}} \\
-        &\geq \onorm{\mS_{i:} (\mTheta \mM)^{\otimes (K-1),T} } \lambda^{K-1}_r(\mW^{(t)}) \\
-        & \geq \theta(i) \onorm{\mS_{z(i):}} \lambda_r^{K-1}(\mTheta \mM) \lambda^{K-1}_r(\mW^{(t)})\\
-        & \geq\theta(i) m , \label{eq:xv_lower}
-    \end{align}
-    where the first two inequalities follow by the property of eigenvalues, the last inequality follows by Lemmas~\ref{lem:singular_thetam} and \ref{lem:membership}, the assumption that $\min_{a \in [r]} \onorm{\mS_{z(i):}} \geq c_3$, and $m >0$ is a positive constant related to $ c_3$.
-Plugging the lower bound of $ \onorm{\mX_{i:} \mV^{(t)}}$ \eqref{eq:xv_lower} into the inequality \ref{eq:ind_b}, we obtain that
+     \onormSize{}{\mX_{i:} \mV^{(t)}}  = \theta(i) \onormSize{}{\mS_{i:} (\mTheta \mM)^{\otimes (K-1),T} \mW^{(t), \otimes^{K-1}}}
+       % &\geq \theta(i) \onormSize{}{\mS_{i:} (\mTheta \mM)^{\otimes (K-1),T} } \lambda^{K-1}_r(\mW^{(t)}) \\
+         \geq \theta(i) \onormSize{}{\mS_{z(i):}} \lambda_r^{K-1}(\mTheta \mM) \lambda^{K-1}_r(\mW^{(t)})
+         \geq\theta(i) m , \label{eq:xv_lower}
+    \end{align}
+    where the first inequality follows by the property of eigenvalues, the last inequality follows by Lemmas~\ref{lem:singular_thetam}, \ref{lem:membership}, the assumption that $\min_{a \in [r]} \onormSize{}{\mS_{z(i):}} \geq c_3$, and $m >0$ is a positive constant related to $ c_3$.
+Plugging the lower bound of $ \onormSize{}{\mX_{i:} \mV^{(t)}}$ \eqref{eq:xv_lower} into the inequality \eqref{eq:ind_b}, we obtain that
     \begin{equation}\label{eq:decomp1}
-        \ind \offf{ z^{(t+1)}(i) = b  } \leq I_1 + I_2,
+        \ind \offf{ z^{(t+1)}(i) = b  } \leq A_{ib} + B_{ib},
     \end{equation}
@@
     \begin{align}
-        A_{ib} &= \ind \offf{\ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s }  \leq - \frac{\theta(i) m}{4} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 },\\
-        B_{ib} &= \ind \offf{z^{(t+1)}(i) = b, \frac{1}{2}\onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }.
-    \end{align}
-    On one hand, the event $A_{ib}$ is the critical component in $\xi$, where
-    \begin{equation}
-        \xi =   \frac{1}{p} \sum_{i \in [p] } \theta(i) \sum_{b \in [r]/z(i)}  \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 A_{ib}
-    \end{equation}
-    We will investigate $\bbP(A_{ib} = 1)$ when verifying the Condition~\ref{cond:origin}.
-
-    On the other hand, the terms $ F_{ib}^{(t)}, G_{ib}^{(t)}$, and $H_{ib}^{(t)}$ in $B_{ib}$ describe the misclassification error in $t$-th iteration, which intuitively can be represented by $L^{(t)}$.
-Specifically, under Condition~\ref{cond:origin}, Lemma~\ref{lem:upper_fgh} and \citet[Step 4, Proof of Theorem 2]{han2020exact} imply that
-    %\fixme{Jiaxin}{Need to elaborate more. Step 4 is not completely the same as Han. Need to care about $\theta$}
-    \begin{equation}
-        \frac{1}{p} \sum_{i \in [p] } \theta(i) \sum_{b \in [r]/z(i)}  \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 B_{ib} \leq c' L^{(t+1)} + \rho L^{(t)},
-    \end{equation}
-    where $c', \rho \in (0,1)$ are two small positive constants.
-Hence, by the upper bound~\eqref{eq:decomp1} and the defintion of $L^{(t+1)}$, we obtain the upper bound for the misclassification loss in the $t+1$-th iteration
+        A_{ib} &= \ind \offf{z^{(t+1)}(i) = b, \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 },\\
+        B_{ib} &= \ind \offf{z^{(t+1)}(i) = b, \frac{1}{2}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }.
+    \end{align}
+
+    {
+    \color{blue} Take the weighted summation of~\eqref{eq:decomp1} over $i \in [p]$. We have 
+    \begin{equation}
+        L^{(t+1)} \leq \xi + \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}^{(t)}, 
+    \end{equation}
+    where $\xi$ is the oracle loss such that 
+    \begin{equation}\label{eq:decomp2}
+        \xi =   \frac{1}{p} \sum_{i \in [p] } \theta(i) \sum_{b \in [r]/z(i)} A_{ib} \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2.
+    \end{equation}
+    and we define
+    \begin{equation}
+        \zeta_{ib}^{(t)} =   \theta(i) B_{ib} \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2.
+    \end{equation}
+
+    {\bf Step 3 (Derivation of contraction inequality):} In this step we derive the upper bound of $\zeta_{ib}$ and thereof obtain the contraction inequality. Note that 
+    \begin{align}
+        \zeta_{ib}^{(t)} &= \theta(i) \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \ind \offf{z^{(t+1)}(i) = b, \frac{1}{2}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }\\
+        & \leq \theta(i) \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \ind \offf{z^{(t+1)}(i) = b, \frac{1}{4}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} } \\
+        & \lesssim \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \theta(i)m^2\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    }\\
+        &\leq C \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \theta(i)m^2\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    },\label{eq:zeta}
+    \end{align}
+    where the first inequality follows by the inequality~\eqref{eq:hib} in Lemma~\ref{lem:upper_fgh}, and the last inequality follows by the assumption that $\min_{i \in [p]} \theta(i) \geq c$.
+Following \citet[Step 4, Proof of Theorem 2]{han2020exact} and Lemma~\ref{lem:upper_fgh}, we have
+    \begin{align}\label{eq:f_sum}
+          \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{(F_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \leq C_1 L^{(t)},
+    \end{align}
+    and 
+    \begin{align}
+         \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}  &\lesssim \frac{1}{p} \sum_{i \in [p]}\theta(i) \sum_{b \in [r]/z(i)}   \ind \offf{ z^{(t+1)}(i) = b}  (\Delta_{\min}^2 + L^{(t)}) \\
+         & \leq C_2 (L^{(t+1)} + L^{(t)}), \label{eq:g_sum}
+    \end{align}
+    where the last inequality follows by the definition of $L^{(t)}$ and the constraint of $\mtheta$ in parameter space~\eqref{eq:family}.
+Choosing the constants $C, C_1, C_2$ in~\eqref{eq:zeta}~\eqref{eq:f_sum}~\eqref{eq:g_sum} carefully, we have
+    \begin{equation}\label{eq:zeta_upper}
+        \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}^{(t)} \leq  C_4 L^{(t+1)}  + C_5 L^{(t)},
+    \end{equation}
+    for some small positive constants $C_4, C_5$. Plugging the inequality~\eqref{eq:zeta_upper} into the decomposition~\eqref{eq:decomp2}, we obtain the contraction inequality 
     \begin{equation}\label{eq:decomp_t}
@@
     \end{equation}
-    where $M \geq 1$ and $\rho$ is called the contraction parameter.
-
-    \paragraph{Step 3:} Last, we verify the Condition~\ref{cond:origin} under high probability to finish the proof.
-Note that the inequalities~\eqref{eq:cond1}, \eqref{eq:cond2}, and \eqref{eq:cond3} describes the property of the noise tensor $\tE$, and the readers can find the proof directly in \citet[Step 5, Proof of Theorem 2]{han2020exact}.
+     where $M \geq 1$ and $\rho \in [0,1)$ is the contraction parameter.
+
+     Combining the inequality~\eqref{eq:cond_oracle} in Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}, we have proven Theorem~\ref{thm:refinement} with inequality~\eqref{eq:decomp_t}.
+    }
+
+    % On one hand, the event $A_{ib}$ is a critical component of $\xi$, where
+
+    % %We will investigate $\bbP(A_{ib} = 1)$ when verifying the Condition~\ref{cond:origin}.
+
+    % On the other hand, the terms $ F_{ib}^{(t)}, G_{ib}^{(t)}$, and $H_{ib}^{(t)}$ in $B_{ib}$ describe the misclustering error in that the $t$-th iteration, which intuitively can be represented by $L^{(t)}$.
+Specifically, under Condition~\ref{cond:origin}, Lemma~\ref{lem:upper_fgh} and \citet[Step 4, Proof of Theorem 2]{han2020exact} imply that
+    % %\fixme{Jiaxin}{Need to elaborate more. Step 4 is not completely the same as Han. Need to care about $\theta$}
+    % \begin{equation}
+    %     \frac{1}{p} \sum_{i \in [p] } \theta(i) \sum_{b \in [r]/z(i)} B_{ib} \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2  \leq c' L^{(t+1)} + \rho L^{(t)},
+    % \end{equation}
+    % where $c', \rho \in (0,1)$ are two small positive constants.
+Hence, by the upper bound~\eqref{eq:decomp1} and the defintion of $L^{(t+1)}$, we obtain the upper bound for the misclustering loss in the $t+1$-th iteration
+    % \begin{equation}\label{eq:decomp_t}
+    %     L^{(t+1)} \leq M \xi + \rho L^{(t)},
+    % \end{equation}
+    % where $M \geq 1$ and $\rho$ is called the contraction parameter.
+
+    {\bf Step 4 (Verification of Condition~\ref{cond:origin}):} Last, we verify the Condition~\ref{cond:origin} under high probability to finish the proof.
+Note that the inequalities~\eqref{eq:cond1}, \eqref{eq:cond2}, and \eqref{eq:cond3} describes the property of the sub-Gaussian noise tensor $\tE$, and the readers can find the proof directly in \citet[Step 5, Proof of Theorem 2]{han2020exact}.
+Here, we only include the verification of inequalities~\eqref{eq:cond_oracle} and \eqref{eq:cond_intial}.
 
@@
     \begin{align}
-        \xi = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \offf{  \ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s }  \leq - \frac{\theta(i) m}{4} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 }\\
-        \cdot \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{ b:}  }^s  }^2,
-    \end{align}
-
-    Let $e_i = \mE_{i:} \mV$ denote the aggregated noise vector for $i \in [p]$, and $e_i$ are independent mean-zero sub-Gaussian vector in $\bbR^{r^{K-1}}$.
-Also, each coordinate in $e_i$ is also independent mean-zero sub-Gaussian variable with sub-Gaussian norm upper bounded by $C\sqrt{r^{K-1}/p^{K-1}}$ with some positive constant $C$.
+        \xi = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \offf{  \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2,
+    \end{align}
+
+    Let $e_i = \mE_{i:} \mV$ denote the aggregated noise vector for all $i \in [p]$, and $e_i$ are independent mean-zero sub-Gaussian vector in $\bbR^{r^{K-1}}$.
+Each coordinate in $e_i$ is a independent mean-zero sub-Gaussian variable with sub-Gaussian norm upper bounded by $C\sqrt{r^{K-1}/p^{K-1}}$ with some positive constant $C$.
     %\fixme{Miaoyan}{\color{red} explain $e_j$. See my comments at page 34. Please correct all similar issues throughout the appendix.} 
@@
     \begin{align}
-        \bbP \of{ \ang{ e_i, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s }  \leq - \frac{ m}{4} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2} \leq P_1 + P_2 + P_3,
+        \bbP \of{ \ang{ e_i, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{ \theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \leq P_1 + P_2 + P_3,
     \end{align}
@@
     \begin{align}
-        P_1 &= \bbP \of{ \ang{e_i, \off{  \mS_{z(i):} }^s - \off{ \mS_{b:} }^s} \leq -\frac{m}{8}  \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 }, \\
-        P_2 &= \bbP \of{ \ang{e_i, \off{ \tilde  \mS_{z(i):} }^s - \off{ \mS_{z(i):} }^s} \leq -\frac{m}{16}  \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 },\\
-        P_3 &= \bbP \of{ \ang{e_i, \off{  \mS_{b:} }^s - \off{\tilde  \mS_{b:} }^s} \leq -\frac{m}{16}   \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 }.
-    \end{align}
-
-    For $P_1$, notice that the inner product $\ang{e_j,\mS_{z(j):}^s - \mS_{b:}^s} $ is a sub-Gaussian variable with sub-Gaussian norm bounded by $C \sqrt{ r^{K-1}/ p^{K-1} } \onorm{ \mS_{z(j):} ^s -  \mS_{b:}^s }$.
+        P_1 &= \bbP \of{ \ang{e_i, [  \mS_{z(i):} ]^s - [ \mS_{b:} ]^s} \leq -\frac{\theta(i)m}{8}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 }, \\
+        P_2 &= \bbP \of{ \ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s} \leq -\frac{\theta(i)m}{16}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 },\\
+        P_3 &= \bbP \of{ \ang{e_i, [  \mS_{b:} ]^s - [\tilde  \mS_{b:} ]^s} \leq -\frac{\theta(i)m}{16}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 }.
+    \end{align}
+
+    For $P_1$, notice that the inner product $\ang{e_j,\mS_{z(j):}^s - \mS_{b:}^s} $ is a sub-Gaussian variable with sub-Gaussian norm bounded by $C \sqrt{ r^{K-1}/ p^{K-1} } \onormSize{}{ \mS_{z(i):}^s -  \mS_{b:}^s }$.
 Then, by Chernoff bound, we have
     \begin{equation}\label{eq:p1_2}
-         P_1 \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^2 }.
+         P_1 \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(j):}  ]^s - [ \mS_{b:}  ]^s  }^2 }.
     \end{equation}
@@
     \begin{align}
-        P_{21}& =  \bbP \of{ t <   \onorm{ \off{ \tilde  \mS_{z(i):} }^s - \off{ \mS_{z(i):} }^s }},\\
+        P_{21}& =  \bbP \of{ t \leq   \onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s }},\\
         P_{22} &= \bbP \of{ \left.
-\ang{e_i, \off{ \tilde  \mS_{z(i):} }^s - \off{ \mS_{z(i):} }^s} \leq -\frac{m}{16}  \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 \right| \onorm{ \off{ \tilde  \mS_{z(i):} }^s - \off{ \mS_{z(i):} }^s } < t  }.
+\ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s} \leq -\frac{\theta(i)m}{16}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \right| \onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s } < t  }.
     \end{align}
@@
     For $P_{21}$, note that the term $\mW^T_{: z(i)} \mE \mV = \frac{ \sum_{ j \neq i, j \in [p]} \ind \{ z(j) = z(i) \} e_j }{\sum_{j \in [p]} \ind \{ z(j) = z(i) \} }$  is a sub-Gaussian vector with sub-Gaussian norm bounded by $C_1 \sqrt{r^K/ p^K}$.
-Then, we have
-    \begin{align}
-        P_{21} &\leq \bbP\of{ t \onorm{\mS_{z(i):} } < \onorm{ \tilde  \mS_{z(i):}  -  \mS_{z(i):} }  }\\
-        & \leq \bbP\of{c_3 t <   \onorm{ \mW^T_{: z(i)} \mE \mV  } } \\
-        & \lesssim \exp\of{ - \frac{p^K t^2}{r^K} }, \label{eq:p21}
-    \end{align}
-    where the first inequality follows by the basic inequality in Lemma~\ref{lem:norm_diff}, the second inequality follows by the assumption that $ \min_{a \in [r]} \onorm{\mS_{z(i):}} \geq c_3$, and the last inequality follows by the Bernstein inequality.
-
-    For $P_{22}$, the inner product $ \ang{e_i, \off{ \tilde  \mS_{z(i):} }^s - \off{ \mS_{z(i):} }^s}$ is also a sub-Gaussian variable with sub-Gaussian norm $ C \sqrt{ r^{K-1}/ p^{K-1} } t$, conditioned on the boundedness $\onorm{ \off{ \tilde  \mS_{z(j):} }^s - \off{ \mS_{z(j):} }^s } < t$.
+We have
+    \begin{align}
+        P_{21} \leq \bbP\of{ t \onormSize{}{\mS_{z(i):} } \leq \onormSize{}{ \tilde  \mS_{z(i):}  -  \mS_{z(i):} }  } \leq \bbP\of{c_3 t \leq  \onormSize{}{ \mW^T_{: z(i)} \mE \mV  } } 
+        \lesssim \exp\of{ - \frac{p^K t^2}{r^K} }, \label{eq:p21}
+    \end{align}
+    where the first inequality follows by the basic inequality in Lemma~\ref{lem:norm_diff}, the second inequality follows by the assumption that $ \min_{a \in [r]} \onormSize{}{\mS_{z(i):}} \geq c_3$ in~\eqref{eq:family}, and the last inequality follows by the Bernstein inequality.
+
+    For $P_{22}$, the inner product $ \ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s}$ is also a sub-Gaussian variable with sub-Gaussian norm $ C \sqrt{ r^{K-1}/ p^{K-1} } t$, conditioned on $\onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s } < t$.
 Then, by Chernoff bound, we have
     \begin{equation}\label{eq:p22}
-        P_{22} \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1} t^2}   \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^4 }. 
-    \end{equation}
-    Take $t =  \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }$ in $P_{21}$ and $P_{22}$, and plug the inequalities \eqref{eq:p21} and \eqref{eq:p22} into to the upper bound for $P_2$~\eqref{eq:p2}.
+        P_{22} \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1} t^2}   \onormSize{}{ [ \mS_{z(j):}  ]^s - [ \mS_{b:}  ]^s  }^4 }. 
+    \end{equation}
+    Take $t =  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }$ in $P_{21}$ and $P_{22}$, and plug the inequalities \eqref{eq:p21} and \eqref{eq:p22} into to the upper bound for $P_2$~\eqref{eq:p2}.
 We obtain that
     \begin{equation}\label{eq:p2_2}
-        P_2 \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^2 }.
+        P_2 \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 }.
     \end{equation}
@@
     \begin{equation}\label{eq:xi_p}
-          \bbP \of{ \ang{ e_i, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s }  \leq - \frac{ m}{4} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2} \lesssim  \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^2 }.
+          \bbP \of{ \ang{ e_i, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \lesssim  \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 }.
     \end{equation}
@@
     \begin{align}
-        \bbE[\xi] & =  \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \bbP \offf{  \ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s }  \leq - \frac{ m}{4} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 } \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{ b:}  }^s  }^2 \\
-        &\lesssim \frac{1}{p} \sum_{i \in [p]}  \sqrt{p} \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2   \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onorm{ \off{ \mS_{z(j):}  }^s - \off{ \mS_{b:}  }^s  }^2 } \\
+        \bbE[\xi] & =  \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \bbP \offf{  \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2 \\
+        &\lesssim \frac{1}{p} \sum_{i \in [p]}  \theta(i) \max_{i \in [p], b \in [r]}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2   \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \\
         & \lesssim  \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 },
     \end{align}
-    where the first inequality follows by the fact that $\max_{i \in [p]} \theta(i) \lesssim \sqrt{p}$.
+    where the first inequality follows by the constraint that $\sum_{i \in [p]} \theta(i) = p$.
 
@@
 
-    Finally, we verify the bounded loss condition~\eqref{eq:cond_intial} by induction. By Corollary~\ref{cor:L0}, we have 
+    Finally, we verify the bounded loss condition~\eqref{eq:cond_intial} by induction.
+With output $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{2} and the assumption SNR $\gtrsim p^{-K/2} \log p$, by Corollary~\ref{cor:L0}, we have
     \begin{equation}
@@
     \end{equation}
-    when $p$ is large enough by Theorem~\ref{thm:initial}, and thus condition~\eqref{eq:cond_intial} holds for $t = 0$.
+    when $p$ is large enough, and thus condition~\eqref{eq:cond_intial} holds for $t = 0$.
 Assume the condition~\eqref{eq:cond_intial} also holds for all $t \leq t_0$.
@@
     \end{align}
-    where the second inequality follows by condition~\eqref{eq:cond_oracle} and the last inequality follows by the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2}$.
+    where the second inequality follows by condition~\eqref{eq:cond_oracle} and the last inequality follows by the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2} \log p$.
 Thus, the condition~\eqref{eq:cond_intial} holds for $t_0 + 1$, and the condition~\eqref{eq:cond_intial} is proved by induction.
@@
 
-\paragraph{Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:refinement}}
-
-\begin{lem}[Misclassification error and loss]\label{lem:mis} Define the misclustering error in the $t$-th iteration as $\ell^{(t)} = \ell(z^{(t)}, z)$. We have 
+{\bf Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:refinement}}
+
+\begin{lem}[Misclustering error and loss]\label{lem:mis} Define the misclustering error in the $t$-th iteration as $\ell^{(t)} = \ell(z^{(t)}, z)$. We have 
 \begin{equation}
@@
 
-\begin{lem}[Membership matrices]\label{lem:membership} Suppose the condition~\eqref{eq:cond_intial} holds. Then, for any $a \in [r]$, we have $|\of{z^{(t)}}^{-1}(a)| \asymp p/r$. Moreover, we have 
+\begin{proof}[Proof of Lemma~\ref{lem:mis}]
+By the definition of minimal gap in Assumption~\ref{assmp:min_gap}, we have 
+\begin{equation}
+         L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \geq \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \Delta_{\min}^2 \geq c \ell^{(t)} \Delta_{\min}^2,
+    \end{equation}
+    where the last inequality follows by the assumption $\min_{i \in [p]} \theta(i) \geq c$.
+\end{proof}
+
+\begin{lem}[Singular property of membership matrices]\label{lem:membership} Suppose the condition~\eqref{eq:cond_intial} holds.
+Then, for any $a \in [r]$, we have $|\of{z^{(t)}}^{-1}(a)| \asymp p/r$.
+Moreover, we have
 \begin{equation}\label{eq:membership1}
@@
 \begin{equation}\label{eq:WWsingular}
-   \lambda_{\min}(\mW\mW^T) \asymp \onorm{\mW\mW^T}_{\sigma} \asymp r/p,
+   \lambda_{r}(\mW\mW^T) \asymp \onorm{\mW\mW^T}_{\sigma} \asymp r/p,
 \end{equation}
@@
 
-For inequality~\eqref{eq:WWsingular}, note that 
-\begin{align}
-    \lambda(\mW\mW^T) = \sqrt{\text{eigen} (\mW \mW^T\mW\mW^T)} \asymp \sqrt{\frac{r}{p} \text{eigen} (\mW \mW^T)} = \sqrt{\frac{r}{p} \lambda^2(\mW)} \asymp \frac{r}{p},
-\end{align}
-where the first inequality follows the fact that $\mW^T\mW$ is a diagonal matrix with elements of order $\frac{r}{p}$, and the second equation follows by the definition of singular value.
+For inequality~\eqref{eq:WWsingular}, note that for all $k \in [r]$
+\begin{align}
+    \lambda_k(\mW\mW^T) = \sqrt{\text{eigen}_k (\mW \mW^T\mW\mW^T)} \asymp \sqrt{\frac{r}{p} \text{eigen}_k (\mW \mW^T)} = \sqrt{\frac{r}{p} \lambda^2_k(\mW)} \asymp \frac{r}{p},
+\end{align}
+where $\text{eigen}_k(\mA)$ denotes the $k$-th largest eigenvalue of the square matrix $\mA$, the first inequality follows the fact that $\mW^T\mW$ is a diagonal matrix with elements of order $r/p$, and the second equation follows by the definition of singular value.
 \end{proof}
 
-     \begin{lem}[Relationship between error and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin}, we have
+\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin}, we have 
+    \begin{equation}\label{eq:fib}
+        \max_{i \in [p]} \max_{b \neq z(i)} \frac{\of{F_{ib}^{(t)}}^2}{ \onormSize{}{ [  \mS_{z(i):} ]^s  -  [ \mS_{b:} ]^s }^2 } \lesssim  \frac{r L^{(t)}}{ \Delta_{\min}^2} \onormSize{}{\mE_{i:} \mV}^2 + \of{1 +  \frac{ r L^{(t)}}{ \Delta_{\min}^2}} \onormSize{}{\mE_{i:} (\mV - \mV^{(t)})}^2 ,
+    \end{equation}
+    \begin{equation}\label{eq:gib}
+        \max_{i \in [p]} \max_{b \neq z(i)} \frac{ \of{  G_{ib}^{(t)} }^2  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2}  \lesssim \of{ \Delta_{\min}^2 +L^{(t)}},
+    \end{equation}
+   % where $c$ is a small positive constant near to 0.
+    \begin{equation}\label{eq:hib}
+        \max_{i \in [p]} \max_{b \neq z(i)}  \frac{ \aabs{  H_{ib}^{(t)} }  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2} \leq \frac{1}{4}.
+    \end{equation}
+    \end{lem}
+
+\begin{proof}[Proof of Lemma~\ref{lem:upper_fgh}]
+We prove the inequalities in Lemma~\ref{lem:upper_fgh} separately.
+\begin{enumerate}
+    \item Upper bound for $F_{ib}^{(t)}$, inequality \eqref{eq:fib}. Recall the definition of $F_{ib}^{(t)}$ 
+    \begin{align}
+         F_{ib}^{(t)} = 2 \ang{\mE_{i:} \mV^{(t)}, \of{ [ \tilde \mS_{z(i):} ]^s  - [\mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  - [ \mS_{b:}^{(t)} ]^s  }  }  + 2 \ang{ \mE_{i:}(\mV - \mV^{(t)} ), [ \tilde \mS_{z(i):} ]^s - [ \tilde \mS_{b:} ]^s  }.
+    \end{align}
+    By Cauchy-Schwartz inequality, we have 
+    \begin{align}
+        \of{F_{ib}^{(t)}}^2 & \leq 8 \of{\ang{\mE_{i:} \mV^{(t)}, \of{ [  \tilde \mS_{z(i):} ]^s  -  [  \mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  -  [  \mS_{b:}^{(t)} ]^s  }  }  }^2 \\
+         & \quad \quad \quad \quad \quad + 8\of{ \ang{ \mE_{i:} (\mV - \mV^{(t)} ), [  \tilde \mS_{z(i):} ]^s - [ \tilde  \mS_{b:} ]^s  }}^2 \\
+         & \lesssim \of{ \onormSize{}{\mE_{i:} \mV}^2 + \onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2 } \max_{a \in [r]^s} \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  } \\
+          & \quad \quad \quad \quad \quad +\onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2  \onormSize{}{  [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }. \label{eq:fib_decomp}
+    \end{align}
+    Note that for any $a \in [r]$ 
+    \begin{align}
+        \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  }^2 &= \onormSize{}{ [ \mW_{:a}^T \mY \mV ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }^2 \\
+        &\leq 2 \onormSize{}{ [ \mW_{:a}^T \mY \mV ]^s  - [ \mW_{:a}^{(t),T} \mY \mV ]^s  }^2 + 2 \onormSize{}{[ \mW_{:a}^{(t),T} \mY \mV ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }^2\\
+        & \lesssim \frac{ r^2 (L^{(t)})^2}{ \Delta_{\min}^2} + \frac{ r r^{2K} + p r^{K+2}}{p^K} \frac{L^{(t)}}{\Delta_{\min}^2}\\
+        & \lesssim r L^{(t)}+  \frac{ r r^{2K} + p r^{K+2}}{p^K} \frac{L^{(t)}}{\Delta_{\min}^2}\\
+        & \lesssim r L^{(t)},\label{eq:fib_1}
+    \end{align}
+    where the second inequality follows the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4} in Lemma~\ref{lem:intermediate}, the third inequality follows the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}, and the last inequality follows the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.
+
+    Also note that 
+    \begin{align}
+        \onormSize{}{  [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }^2 &=  \onormSize{}{  [  \tilde \mS_{z(i):} ]^s -[ \mS_{z(i):} ]^s +  [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s + [ \mS_{b:} ]^s -[  \tilde \mS_{b:} ]^s  }^2 \\
+        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2 + \max_{a \in [r]} \onormSize{}{ [ \mS_{a:} ]^s -[  \tilde \mS_{a:} ]^s }^2 \\
+        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2  + \max_{a \in [r]} \frac{1}{\onormSize{}{\mS_{a:} }^2} \onormSize{}{ \mW_{:a}^T \mE \mV}^2\\
+        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2, \label{eq:fib_2}
+    \end{align}
+    where the second inequality follows by Lemma~\ref{lem:norm_diff}, and the last inequality follows by the assumptions on $\onormSize{}{\mS_{a:}}$ in the parameter space~\eqref{eq:family}, the inequality~\eqref{eq:cond1} Condition~\ref{cond:origin} and the assumption $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.
+
+    Therefore, we finish the proof of inequality~\eqref{eq:fib} by plugging the inequalities~\eqref{eq:fib_1} and \eqref{eq:fib_2} into the upper bound~\eqref{eq:fib_decomp}.
+
+    \item Upper bound for $G_{ib}^{(t)}$, inequality~\eqref{eq:gib}. Recall the definition of $G_{ib}^{(t)}$ and rearrange the terms
+    \begin{align}
+        G_{ib}^{(t)} &=   \of{ \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2}   \\
+        & \quad \quad \quad \quad \quad -   \of{ \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{b:}^{(t)}  ]^s}^2 -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 } \\
+        &= 2 \ang{  [\mX_{i:} \mV^{(t)}]^s , \of{[  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s } - \of{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s }}\\
+        %& \leq 4 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [  \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }.
+        &= G_1 + G_2 - G_3, \label{eq:gib_decomp}
+    \end{align}
+    where 
+    \begin{align}
+        G_1 &= \onormSize{}{ [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 - \onormSize{}{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s }^2,\\
+        G_2 &= 2 \ang{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s,   [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s},\\
+        G_3 & = 2 \ang{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s,  [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s}.
+    \end{align}
+
+    For $G_1$, we have 
+    \begin{align}
+        |G_1|^2 &\leq \aabs{ \onormSize{}{ [ \mW_{:z(i)}^T \mY \mV^{(t)} ]^s - [ \mS_{z(i):}^{(t)}  ]^s}^2 - \onormSize{}{ [ \mW_{:b}^T \mY \mV^{(t)} ]^s -  [ \mS_{b:}^{(t)}  ]^s }^2 }^2 \\
+        & \leq \max_{a \in [r]} \onormSize{}{ [ \mW_{:a}^T \mY \mV^{(t)} ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^4\\
+        & \lesssim \frac{r^4}{\Delta_{\min}^4} (L^{(t)})^4 +  \frac{r^2 r^{4K} + p^2r^{2K+4}}{p^{2K}} \frac{(L^{(t)})^2}{\Delta_{\min}^4} \\
+        & \leq c\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}}, \label{eq:g1}
+    \end{align}
+    where the third inequality follows by the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:membership} and the last inequality follows by the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
+
+    For $G_2$, noted that $[\mX_{i:} \mV^{(t)}]^s = [ \mW_{z(i):}^T\mX \mV^{(t)}]^s$, we have 
+    \begin{align}
+        |G_2|^2 &\leq 2\onormSize{}{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 \onormSize{}{ [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s }^2 \\
+        & \leq \frac{2}{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2} \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
+        & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} \of{ \frac{r^2}{\Delta_{\min}^2} (L^{(t)})^2 + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} }\\
+        & \leq c \Delta_{\min}^2 L^{(t)},\label{eq:g2}
+    \end{align}
+    where the second inequality follows by Lemma~\ref{lem:norm_diff}, the third inequality follows by the inequality~\eqref{eq:cond2} in Condition~\ref{cond:origin}, and inequalities~\eqref{eq:inter5}, \eqref{eq:j11_wxvt} in the proof of Lemma~\ref{lem:intermediate},  and the last inequality follows by the assumption $\Delta_{\min}^2 \geq p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
+
+    For $G_3$, note that by triangle inequality
+    \begin{align}
+        \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2 &\leq \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + 2\max_{a \in [r]}\onormSize{}{[\mW_{:a}^T\mX \mV^{(t)}]^s - [\mW_{:a}^T\mX \mV]^s }^2\\
+        & \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2}, \label{eq:g3xv_wxv}
+    \end{align}
+    where the last inequality follows by the inequality~\eqref{eq:j1} in the proof of Lemma~\ref{lem:intermediate}.
+    % \begin{align}
+    %     %& \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2},
+    % \end{align}
+   % where the last inequality follows by the inequality~\eqref{eq:inter4} in Lemma~\ref{lem:intermediate}. 
+   Then we have 
+    \begin{align}
+        |G_3|^2 & \leq 2 \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2 \\
+        & \leq 2 \of{ \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  +  \onormSize{}{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  } \\
+        & \quad \cdot \max_{a \in [r]} \onorm{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
+        & \lesssim \of{\onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} } \of{ \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2}}+ c \Delta_{\min}^2 L^{(t)} \\
+        & \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 (\Delta_{\min}^2 + L^{(t)}) + c\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}},\label{eq:g3}
+    \end{align}
+    where the third inequality follows by the procedures to derive~\eqref{eq:g1} and \eqref{eq:g2}
+    %, the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:intermediate},
+    , and the last inequality follows by the assumption $\Delta_{\min}^2 \geq p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
+
+    Plugging the inequalities~\eqref{eq:g1}, \eqref{eq:g2}, and \eqref{eq:g3} into the upper bound~\eqref{eq:gib_decomp}, we finish the proof of inequality~\eqref{eq:gib}.
+
+    % According to the inequality~\eqref{eq:inter_wa_wb} in the proof of Lemma~\ref{lem:intermediate}, we have
+    % \begin{equation}
+    %     \onorm{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  \lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2.
+    % \end{equation}
+    % Then, we have 
+    % \begin{align}
+    %     |G_3|^2 &\lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 \max_{a \in [r]} \onorm{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2 + |G_2|^2 \\
+    %     &\lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 \of{ \frac{r^2}{\Delta_{\min}^2} \of{L^{(t)}}^2 + K  \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} } + |G_2|^2\\
+    %     & \leq \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2  c \of{ \Delta_{\min}^2 + K L^{(t)} }+|G_2|^2, \label{eq:g3}
+    % \end{align}
+    % where the second inequality follows by Lemma~\ref{lem:intermediate}, and the last inequality follows by the assumption on $\Delta_{\min}^2$.
+
+    % Therefore, combining inequalities~\eqref{eq:g1}, \eqref{eq:g2}, \eqref{eq:g3} and apply the inequality of arithmetic and geometric means, .
+
+    \item Upper bound for $H_{ib}^{(t)}$, inequality~\eqref{eq:hib}. Recall the definition of $H_{ib}$ and rearrange the terms
+    \begin{align}
+        H_{ib} &= \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 + \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
+        & =  \onormSize{}{ [ \mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 \\
+        & \quad \quad \quad \quad \quad + \of{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }  } \\
+        & \quad \quad \quad \quad \quad  - \of{\onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mY \mV^{(t)} ]^s } -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }   } \\
+        & = H_1 + H_2 + H_3,
+    \end{align}
+    where 
+    \begin{align}
+        H_1 &= \onormSize{}{ [ \mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 - \onormSize{}{ [ \mW_{:b}^T \mX \mV^{(t)}]^s -  [ \mW_{:b}^T \mY \mV^{(t)}]^s}^2, \\
+        H_2 &=  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }^2  , \\
+        H_3 & = 2 \ang{  [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s,  [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }.
+    \end{align}
+    For $H_1$, we have 
+    \begin{align}
+        |H_1| \leq  \frac{4 \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 }{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2}
+         \leq  \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} 
+        \lesssim  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2\label{eq:h1},
+    \end{align}
+    following the derivation of $G_2$ in inequality~\eqref{eq:g2} and the assumption that $\Delta_{\min}^2 \geq p^{-K/2} \log p$.
+    % \begin{align}
+    %     |H_1| &\leq 2 \max_{a \in [r]}  \frac{1}{ \onormSize{}{  \mW_{:a}^T \mX \mV^{(t)}}^2} \onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2  \\
+    %     & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}}\\
+    %     & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K} \Delta_{\min}^2}  \onorm{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
+    %     & \lesssim  \onorm{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 ,
+    % \end{align}
+    % where the first two inequalities follow the proof of $G_2$~\eqref{eq:g2} and the last inequality follows by the assumption on $\Delta_{\min}^2$.
+
+    For $H_2$, by the inequality~\eqref{eq:g3xv_wxv}, we have 
+    \begin{equation}\label{eq:h2}
+        |H_2| \lesssim 2 \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
+    \end{equation}
+    where the last inequality follows by the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
+
+    % For $H_2$,  sccording to the inequality~\eqref{eq:inter_wa_wb} in the proof of Lemma~\ref{lem:intermediate}, we have
+    % \begin{equation}
+    %     \onorm{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  \lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
+    % \end{equation}
+    % and thus 
+    % \begin{equation}\label{eq:h2}
+    %     |H_2| \lesssim  \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2.
+    % \end{equation}
+
+    For $H_3$,  by Cauchy-Schwartz inequality, we have 
+    \begin{equation}\label{eq:h3}
+         |H_3| \lesssim \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
+    \end{equation}
+    following the inequalities~\eqref{eq:g3xv_wxv} and \eqref{eq:h1}.
+
+    % For $H_3$, by Cauchy-Schwartz inequality, we have 
+    % \begin{align}
+    %     |H_3| &\lesssim    \onorm{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \\
+    %     &\lesssim  \frac{ r^{K-1/2} + K p^{1/2} r^{K/2+1/2} }{p^{K/2}} \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }\\
+    %     & \lesssim \Delta_{\min} \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }\\
+    %     & \lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2, \label{eq:h3}
+    % \end{align}
+    % where the third inequality follows by the assumption on $\Delta_{\min}^2$.
+
+    Therefore, combining inequalities~\eqref{eq:h1}, \eqref{eq:h2}, and \eqref{eq:h3}, we have finish the proof of inequality~\eqref{eq:hib}.
+
+\end{enumerate}
+
+\end{proof}
+
+     \begin{lem}[Relationship between misclustering loss and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin}, we have
     \begin{equation}\label{eq:inter1}
-        \onormSize{}{\mV - \mV^{(t)}}_{\sigma} \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}}  \frac{r(K-1)}{\Delta_{\min}^2} L^{(t)},
+        \onormSize{}{\mV - \mV^{(t)}}_{\sigma} \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}}  \frac{r}{\Delta_{\min}^2} L^{(t)},
     \end{equation}
@@
         \onormSize{}{\mE (\mV - \mV^{(t)})}_{\sigma} \lesssim \sqrt{ \frac{r^{K-1}(p r^{K-1} + p r)}{p^{K-1}}} 
- \frac{r(K-1)}{\Delta_{\min}^2} L^{(t)}.    
+ \frac{r}{\Delta_{\min}^2} L^{(t)}.    
  \end{equation}
@@
     \begin{equation}
-           \onormSize{}{\mE (\mV - \mV^{(t)})}_{\sigma} \lesssim \sqrt{ \frac{r^{K-1}(p r^{K-1} + p r)}{p^{K-1}}} r(K-1) \ell^{(t)}.
+           \onormSize{}{\mE (\mV - \mV^{(t)})}_{\sigma} \lesssim \sqrt{ \frac{r^{K-1}(p r^{K-1} + p r)}{p^{K-1}}} r \ell^{(t)}.
     \end{equation}
@@
      \begin{equation}
-        I_{11} = \onorm{ \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }}   } \quad \text{and} \quad I_{12} = \onorm{ \frac{\mW_{:b}^T \mE \mV}{ \onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mE \mV}{ \onormSize{}{  \mW_{:b}^{(t),T} \mX \mV }} }
+        I_{11} = \onorm{ \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }}   } \quad \text{and} \quad I_{12} = \onorm{ \frac{\mW_{:b}^T \mE \mV}{ \onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mE \mV}{ \onormSize{}{  \mW_{:b}^{(t),T} \mX \mV }} }.
     \end{equation}
@@
     \begin{equation}\label{eq:dbb_bound}
-        D_{bb} = \frac{\sum_{i \in [p]}  \theta(i) \ind \offf{ z(i) = z^{(t)}(i) = b } }{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }} \geq \frac{c_0(\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } - p\ell^{(t)})}{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }} \gtrsim 1 - \frac{1}{\log p},
+        D_{bb} = \frac{\sum_{i \in [p]}  \theta(i) \ind \offf{ z(i) = z^{(t)}(i) = b } }{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }} \geq \frac{c(\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } - p\ell^{(t)})}{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }} \gtrsim 1 - \frac{1}{\log p},
     \end{equation}
@@
 
-    Let $\alpha$ denote the angle between $\mS_{b:}$ and $D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}$. To rough estimate the range of $\alpha$, we consider the inner product 
+    Let $\alpha$ denote the angle between $\mS_{b:}$ and $D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}$. To roughly estimate the range of $\alpha$, we consider the inner product 
     \begin{align}
@@
     \end{align}
-    where $C$ is a positive constant, and the last inequality holds when $p$ is large enough following the constraint of $\onorm{\mS_{b:}}$ in parameter space~\eqref{eq:family} and the bounds of $\mD$ \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}.
+    where $C$ is a positive constant, and the last inequality holds when $p$ is large enough following the constraint of $\onorm{\mS_{b:}}$ in parameter space~\eqref{eq:family} and the bounds of $\mD$ in \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}.
 
@@
        \onormSize{}{ [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}] \sin \alpha} &= \onormSize{}{(\mI_d - \text{Proj}(\mS_{b:})) \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}}\\
-       & = \sum_{a \neq b, a \in [r]} D_{ab} \onorm{(\mI_d - \text{Proj}(\mS_{b:}))\mS_{a:}  }\\
+       &\leq \sum_{a \neq b, a \in [r]} D_{ab} \onorm{(\mI_d - \text{Proj}(\mS_{b:}))\mS_{a:}  }\\
        &= \sum_{a \neq b, a \in [r]} D_{ab} \onorm{ \mS_{a:} \sin (\mS_{b:}, \mS_{a:}) }\\
@@
     \end{align}
-    where the last inequality follows by Lemma~\ref{lem:norm_diff}. Also, note that with bounds \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}, when $p$ is large enough we have 
+    where the first inequality follows by triangle inequality, and the last inequality follows by Lemma~\ref{lem:norm_diff}.
+Also, note that with bounds \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}, when $p$ is large enough we have
     \begin{equation}\label{eq:i11_sin_dom}
@@
         & \lesssim \frac{r}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind\offf{z^{(t)}(i) = b} \onorm{ \mS^s_{b:} - \mS^s_{a:}}\\
-        & = \frac{r L^{(t)}}{\Delta_{\min}},\label{eq:i11}
-    \end{align}
-    where the second inequality follows by~\eqref{eq:i11_sin_num} and \eqref{eq:i11_sin_dom}, and the last inequality follows by the definition of $D_a$ and the constraint of $\onorm{\mS_{b:}}$ in parameter space~\eqref{eq:family}.
+        & \leq \frac{r L^{(t)}}{\Delta_{\min}},\label{eq:i11}
+    \end{align}
+    where the second inequality follows by~\eqref{eq:i11_sin_num} and \eqref{eq:i11_sin_dom}, and the last two inequalities follow by the definition of $D_a, L^{(t)}$ and the constraint of $\onorm{\mS_{b:}}$ in parameter space~\eqref{eq:family}.
 
@@
     \end{equation}
-    where the second inequality follows by  \citet[Proof of Lemma 5]{han2020exact} and the last inequality follows by Lemma~\ref{lem:singular_thetam} and \eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
+    where the second inequality follows by  \citet[Inequality (121), Proof of Lemma 5]{han2020exact} and the last inequality follows by Lemma~\ref{lem:singular_thetam} and \eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
     % Also, by the definition of $\mW^{(t)}$ and Condition~\ref{cond:origin}, we have
@@
 
-    For $J_1$, by triangle inequality and Lemma~\ref{lem:norm_diff}, we have 
+    For $J_1$, by triangle inequality 
+    %and Lemma~\ref{lem:norm_diff}
+    , we have 
     \begin{equation}
@@
     \begin{equation}
-        J_{11} =  \onorm{ \frac{\mW_{:b}^{(t),T} \mX \mV }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mX \mV^{(t)} }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV^{(t)} }}  }, \quad J_{12} =  \onorm{ \frac{\mW_{:b}^{(t),T}  \mE \mV}{ \onormSize{}{\mW_{:b}^{(t),T}  \mX \mV} } - \frac{\mW_{:b}^{(t),T}  \mE \mV^{(t)}}{ \onormSize{}{\mW_{:b}^{(t),T}  \mX \mV^{(t)}} }   }.
+        J_{11} =  \onorm{ \frac{\mW_{:b}^{(t),T} \mX \mV }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mX \mV^{(t)} }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV^{(t)} }}  } \quad \text{and} \quad \quad J_{12} =  \onorm{ \frac{\mW_{:b}^{(t),T}  \mE \mV}{ \onormSize{}{\mW_{:b}^{(t),T}  \mX \mV} } - \frac{\mW_{:b}^{(t),T}  \mE \mV^{(t)}}{ \onormSize{}{\mW_{:b}^{(t),T}  \mX \mV^{(t)}} }   }.
     \end{equation}
@@
 Let $\beta_k$ denote the angle between $\mW_{:b}^{(t),T} \mX \mV^k$ and $\mW_{:b}^{(t),T} \mX \mV^{k+1}$.
-With the same idea to prove $I_{11}$, we bound $J_{11}^k$ by considering the trigonometric function of $\beta_k$.
+With the same idea to prove $I_{11}$ in inequality~\eqref{eq:i11}, we bound $J_{11}^k$ by the trigonometric function of $\beta_k$.
 
@@
     %, and let  $\mZ^k =  \mW^{\otimes (k-1)} \otimes \mW^{(t),\otimes (K-k-1)} \in \bbR^{ p^{K-2} \times r^{K-2} }$.  
-    Then, we have 
+    We have 
     \begin{align}
@@
     \end{align}
-    where $\mZ^k = \mD_{:b} \otimes \mI_r^{\otimes (k-1)} \otimes  \mD^{\otimes (K-k-1)}$, the equations follow by the tensor algebra and the definitions, and the last inequality follows by the bounds of $\mD$ \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}.
+    where $\mZ^k = \mD_{:b} \otimes \mI_r^{\otimes (k-1)} \otimes  \mD^{\otimes (K-k-1)}$, the equations follow by the tensor algebra and the definitions, and the last inequality follows by the bounds of $\mD$ in \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}.
 
@@
     \end{equation}
-    where the third inequality follows by the fact that for all $b \in [r]$ 
+    where the second inequality follows by inequality~\eqref{eq:dab_bound}, and the fact that for all $b \in [r]$ 
     \begin{equation}
@@
 
-    Next, we upper bound the angle $\sin \beta_k$. Note that
+    Next, we upper bound the trigonometric function $\sin \beta_k$. Note that
     \begin{align}
@@
 
-    On one hand, recall the definitions for any cluster assignment $\bar z$
+    On one hand, recall Definition~\ref{def:stable} that for any cluster assignment $\bar z$
     \begin{equation}
@@
     \end{equation}
-    Then, we have $\text{diag}(\mp(z^{(t)})) \mD = \text{diag}(\mp_{\mtheta}(z^{(t)})) \tilde \mD$.
-By Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}, we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \lesssim \log^{-1}(p) $.
-Then, by the stability assumption, we have
+    Note that we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \lesssim \log^{-1}(p) $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}. With stability assumption, we have 
     \begin{equation}
@@
     \end{equation}
-    Note that $\sin(\ma,\mb) =\min_{c \in \bbR} \frac{ \onormSize{}{\ma - c\mb} }{\onormSize{}{\ma}}$ for vectors $\ma,\mb$ of same dimension, and let $c_0 = \argmin_{c \in \bbR} \frac{\onormSize{}{\mp(z^{(t)}) - c\mp_{\mtheta}(z^{(t)})}}{\onormSize{}{\mp(z^{(t)})}}$.
-We have
+    Note that $\text{diag}(\mp(z^{(t)})) \mD = \text{diag}(\mp_{\mtheta}(z^{(t)})) \tilde \mD$.
+Also, note that $\sin(\ma,\mb) =\min_{c \in \bbR} \frac{ \onormSize{}{\ma - c\mb} }{\onormSize{}{\ma}}$ for vectors $\ma,\mb$ of same dimension, and let $c_0 = \argmin_{c \in \bbR} \frac{\onormSize{}{\mp(z^{(t)}) - c\mp_{\mtheta}(z^{(t)})}}{\onormSize{}{\mp(z^{(t)})}}$.
+Then, we have
     \begin{align}
@@
        &\lesssim \frac{ \onormSize{}{ \mp(z^{(t)}) - c_0\mp_{\mtheta}(z^{(t)})} }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  } \\
-       & \lesssim \frac{\onormSize{}{ \mp(z^{(t)})} }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1 } \sin (\mp(z^{(t)}), \mp_{\mtheta}(z^{(t)}))\\
+       & = \frac{\onormSize{}{ \mp(z^{(t)})} }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1 } \sin (\mp(z^{(t)}), \mp_{\mtheta}(z^{(t)}))\\
        & \lesssim \frac{L^{(t)}}{\Delta_{\min}},
     \end{align}
-    where the last inequality follows Lemma~\ref{lem:membership} that $\onormSize{}{ \mp(z^{(t)})} \lesssim p$ and $\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  \gtrsim p$.
+    where the last inequality follows Lemma~\ref{lem:membership} and constraint $\min_{i \in [p]}\theta(i) \geq c$ and thereof $\onormSize{}{ \mp(z^{(t)})} \lesssim p$ and $\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  \gtrsim p$.
 By the geometry property of trigonometric function, we have
@@
     \end{align}
-    where the second inequality follows by the singular property of $\mD$ in \eqref{eq:singular_Du} and \eqref{eq:singular_Dl}.
+    where the second inequality follows by the singular property of $\mD$ in \eqref{eq:singular_Du}, \eqref{eq:singular_Dl} and the constraint of $\mS$ in \eqref{eq:family}.
 
@@
         & \lesssim \frac{ \onormSize{}{(\mI_r - \tilde \mD^T) \mS \mZ^k}_F }{\onormSize{}{\mD_{:b}^{T} \mS} \lambda_r^{K-k-1}(\mD)} \\
-        & \lesssim \onormSize{ }{(\mI_r - \tilde \mD^T) \mS \mC}_F \onormSize{}{\mC^{-1} \mZ^k}_{\sigma}\\
+        & \lesssim \onormSize{ }{(\mI_r - \tilde \mD^T) \mS \mC^{-1}}_F \onormSize{}{\mC \mZ^k}_{\sigma}\\
         & \lesssim \frac{r}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \onormSize{}{\mS_{b:}^s - \mS_{z(i):}^s} \\
@@
     \begin{equation}\label{eq:j1}
-        J_{11} \leq \sum_{k = 1}^{K-1} J_{11}^k \lesssim (K-1)\frac{r L^{(t)}}{\Delta_{\min}}.
+        J_{11} \leq \sum_{k = 1}^{K-1} J_{11}^k \lesssim  \sum_{k = 1}^{K-1} \sin \beta_k  \lesssim (K-1)\frac{r L^{(t)}}{\Delta_{\min}}.
     \end{equation}
@@
     \begin{align}
-        J_{12} &\leq \frac{ 1}{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }} \onormSize{}{\mW_{:b}^{(t),T} \mE (\mV-\mV^{(t)})} \\ &\quad \quad \quad  +
+        J_{12} &\leq \frac{ 1}{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }} \onormSize{}{\mW_{:b}^{(t),T} \mE (\mV-\mV^{(t)})}   +
         \frac{ \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})} }{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)} }} \onormSize{}{\mW_{:b}^{(t),T} \mE \mV^{(t)} }.
@@
     \end{align}
-    where the first inequality follows by the tensor algebra in equation~\label{eq:j11_inner}, and the second inequality follows by the fact  that $\mI = \mW^T \mTheta \mM$, and the last inequality follows by \citet[Proof of Lemma 5]{han2020exact}.
+    where the first inequality follows by the tensor algebra in 
+    inequality~\eqref{eq:j11inner}, and the second inequality follows by the fact  that $\mI = \mW^T \mTheta \mM$, and the last inequality follows by \citet[Proof of Lemma 5]{han2020exact}.
 Thus, we have
     \begin{equation}
-           \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})}\leq \onormSize{}{\mW_{:b}^{(t),T} }  \sum_{k = 1}^{K-1} \onormSize{}{\mX (\mV^k-\mV^{k+1})}_F \lesssim \frac{ (K-1) \sqrt{r L^{(t)}}}{\sqrt{p} \Delta_{\min}}.\label{eq:j12_wxv}
+           \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})}\leq \onormSize{}{\mW_{:b}^{(t),T} }  \sum_{k = 1}^{K-1} \onormSize{}{\mX (\mV^k-\mV^{k+1})}_F \lesssim \frac{ \sqrt{r L^{(t)}}}{\sqrt{p} \Delta_{\min}}.\label{eq:j12_wxv}
     \end{equation}
@@
         J_{12} &\lesssim \onormSize{}{\mW_{:b}^{(t),T} \mE (\mV-\mV^{(t)})} + \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})}\onormSize{}{\mW_{:b}^{(t),T} \mE \mV^{(t)} }\\
-        &\lesssim \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{(K-1)\sqrt{L^{(t)}}}{\Delta_{\min}} + \frac{ (K-1) \sqrt{r L^{(t)}}}{\sqrt{p} \Delta_{\min}} \sqrt{\frac{r^{2K}}{p^K}} \\
-        & \lesssim (K-1) \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
+        &\lesssim \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} + \frac{ \sqrt{r L^{(t)}}}{\sqrt{p} \Delta_{\min}} \sqrt{\frac{r^{2K}}{p^K}} \\
+        & \lesssim  \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
     \end{align}
@@
     \begin{align}
-        \onormSize{}{ [ \mW_{:b}^T \mY \mV^{(t)} ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s } &\leq \onormSize{}{[ \mW_{:b}^T \mY \mV^{(t)} ]^s - [ \mW_{:b}^T \mY \mV ]^s } \\
-        & + \onormSize{}{[\mW_{:b}^T \mY \mV ]^s  - [ \mW_{:b}^{(t),T} \mY \mV ]^s } \\
-        &+ \onormSize{}{[\mW_{:b}^{(t),T} \mY \mV ]^s  -  [\mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s }\\
+        \onormSize{}{ [ \mW_{:b}^T \mY \mV^{(t)} ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s } &\leq \onormSize{}{[ \mW_{:b}^T \mY \mV^{(t)} ]^s - [ \mW_{:b}^T \mY \mV ]^s }  + \onormSize{}{[\mW_{:b}^T \mY \mV ]^s  - [ \mW_{:b}^{(t),T} \mY \mV ]^s } \\
+        & \quad \quad \quad\quad\quad+ \onormSize{}{[\mW_{:b}^{(t),T} \mY \mV ]^s  -  [\mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s }\\
         &\lesssim \frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
@@
 
-    \begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin}, we have 
-    \begin{equation}\label{eq:fib}
-        \max_{i \in [p]} \max_{b \neq z(i)} \frac{\of{F_{ib}^{(t)}}^2}{ \onormSize{}{ [  \mS_{z(i):} ]^s  -  [ \mS_{b:} ]^s }^2 } \lesssim  \frac{r L^{(t)}}{ \Delta_{\min}^2} \onormSize{}{\mE_{i:} \mV}^2 + \of{1 +  \frac{ r L^{(t)}}{ \Delta_{\min}^2}} \onormSize{}{\mE_{i:} (\mV - \mV^{(t)})}^2 ,
-    \end{equation}
-    \begin{equation}\label{eq:gib}
-        \max_{i \in [p]} \max_{b \neq z(i)} \frac{ \of{  G_{ib}^{(t)} }^2  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2}  \lesssim \of{ \Delta_{\min}^2 +L^{(t)}},
-    \end{equation}
-    where $c$ is a small positive constant near to 0.
-    \begin{equation}\label{eq:hib}
-        \max_{i \in [p]} \max_{b \neq z(i)}  \frac{ \aabs{  H_{ib}^{(t)} }  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2} \leq \frac{1}{4}.
-    \end{equation}
-    \end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:upper_fgh}]
-We prove Lemma~\ref{lem:upper_fgh} by the following order:
-\begin{enumerate}
-    \item Upper bound for $F_{ib}^{(t)}$~\eqref{eq:fib}. Recall the definition of $F_{ib}^{(t)}$ 
-    \begin{align}
-         F_{ib}^{(t)} &= 2 \ang{\mE_{i:} \mV^{(t)}, \of{ [ \tilde \mS_{z(i):} ]^s  - [\mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  - [ \mS_{b:}^{(t)} ]^s  }  } \\
-        & \quad \quad \quad \quad \quad + 2 \ang{ \mE_{i:}(\mV - \mV^{(t)} ), [ \tilde \mS_{z(i):} ]^s - [ \tilde \mS_{b:} ]^s  }.
-    \end{align}
-    By Cauchy-Schwartz inequality, we have 
-    \begin{align}
-        \of{F_{ib}^{(t)}}^2 & \leq 8 \of{\ang{\mE_{i:} \mV^{(t)}, \of{ [  \tilde \mS_{z(i):} ]^s  -  [  \mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  -  [  \mS_{b:}^{(t)} ]^s  }  }  }^2 \\
-         & \quad \quad \quad \quad \quad + 8\of{ \ang{ \mE_{i:} (\mV - \mV^{(t)} ), [  \tilde \mS_{z(i):} ]^s - [ \tilde  \mS_{b:} ]^s  }}^2 \\
-         & \lesssim \of{ \onormSize{}{\mE_{i:} \mV}^2 + \onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2 } \max_{a \in [r]^s} \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  } \\
-          & \quad \quad \quad \quad \quad +\onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2  \onormSize{}{  [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }.
-    \end{align}
-    Note that for any $a \in [r]$ 
-    \begin{align}
-        \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  }^2 &= \onormSize{}{ [ \mW_{:a}^T \mY \mV ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }^2 \\
-        &\leq 2 \onormSize{}{ [ \mW_{:a}^T \mY \mV ]^s  - [ \mW_{:a}^{(t),T} \mY \mV ]^s  }^2  \\
-        & \quad \quad \quad +2 \onormSize{}{[ \mW_{:a}^{(t),T} \mY \mV ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }^2\\
-        & \lesssim \frac{ r^2 (L^{(t)})^2}{ \Delta_{\min}^2} + \frac{ r r^{2K} + p r^{K+2}}{p^K} \frac{L^{(t)}}{\Delta_{\min}^2}\\
-        & \lesssim r L^{(t)}+  \frac{ r r^{2K} + p r^{K+2}}{p^K} \frac{L^{(t)}}{\Delta_{\min}^2}\\
-        & \lesssim r L^{(t)},
-    \end{align}
-    where the second inequality follows the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4} in Lemma~\ref{lem:intermediate}, the third inequality follows the initialization condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}, and the last inequality follows the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.
-
-    Also note that 
-    \begin{align}
-        \onormSize{}{  [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }^2 &=  \onormSize{}{  [  \tilde \mS_{z(i):} ]^s -[ \mS_{z(i):} ]^s +  [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s + [ \mS_{b:} ]^s -[  \tilde \mS_{b:} ]^s  }^2 \\
-        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2 + \max_{a \in [r]} \onormSize{}{ [ \mS_{a:} ]^s -[  \tilde \mS_{a:} ]^s }^2 \\
-        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2  + \max_{a \in [r]} \frac{1}{\onormSize{}{\mS_{a:} }^2} \onormSize{}{ \mW_{:a}^T \mE \mV}^2\\
-        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2,
-    \end{align}
-    where the second inequality follows by Lemma~\ref{lem:norm_diff}, and the last inequality follows by the assumptions on $\onormSize{}{\mS_{a:}}$ in the parameter space~\eqref{eq:family}, the inequality~\eqref{eq:cond1} Condition~\ref{cond:origin} and the assumption $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.
-
-    Therefore, we finish the proof of inequality~\eqref{eq:fib}.
-
-    \item Upper bound for $G_{ib}^{(t)}$~\eqref{eq:gib}. Recall the definition of $G_{ib}^{(t)}$ and rearrange the terms
-    \begin{align}
-        G_{ib}^{(t)} &=   \of{ \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2}   \\
-        & \quad \quad \quad \quad \quad -   \of{ \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{b:}^{(t)}  ]^s}^2 -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 } \\
-        &= 2 \ang{  [\mX_{i:} \mV^{(t)}]^s , \of{[  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s } - \of{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s }}\\
-        %& \leq 4 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [  \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }.
-        &= G_1 + G_2 - G_3,
-    \end{align}
-    where 
-    \begin{align}
-        G_1 &= \onormSize{}{ [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 - \onormSize{}{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s }^2,\\
-        G_2 &= 2 \ang{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s,   [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s},\\
-        G_3 & = 2 \ang{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s,  [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s}.
-    \end{align}
-
-    For $G_1$, we have 
-    \begin{align}
-        |G_1|^2 &\leq \aabs{ \onormSize{}{ [ \mW_{:z(i)}^T \mY \mV^{(t)} ]^s - [ \mS_{z(i):}^{(t)}  ]^s}^2 - \onormSize{}{ [ \mW_{:b}^T \mY \mV^{(t)} ]^s -  [ \mS_{b:}^{(t)}  ]^s }^2 }^2 \\
-        & \leq \max_{a \in [r]} \onormSize{}{ [ \mW_{:a}^T \mY \mV^{(t)} ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^4\\
-        & \lesssim \frac{r^4}{\Delta_{\min}^4} (L^{(t)})^4 +  \frac{r^2 r^{4K} + p^2r^{2K+4}}{p^{2K}} \frac{(L^{(t)})^2}{\Delta_{\min}^4} \\
-        & \leq c\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}}, \label{eq:g1}
-    \end{align}
-    where the third inequality follows by the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:membership} and the last inequality follows by the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.
-
-    For $G_2$, noted that $[\mX_{i:} \mV^{(t)}]^s = [ \mW_{z(i):}^T\mX \mV^{(t)}]^s$, we have 
-    \begin{align}
-        |G_2|^2 &\leq 2\onormSize{}{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 \onormSize{}{ [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s }^2 \\
-        & \leq \frac{2}{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2} \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
-        & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} \of{ \frac{r^2}{\Delta_{\min}^2} (L^{(t)})^2 + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} }\\
-        & \leq c \Delta_{\min}^2 L^{(t)},\label{eq:g2}
-    \end{align}
-    where the second inequality follows by Lemma~\ref{lem:norm_diff}, the third inequality follows by the bound~\eqref{eq:j11_wxvt}, inequality~\eqref{eq:cond2} in Condition~\ref{cond:origin}, inequality~\eqref{eq:inter5} in Lemma~\ref{lem:intermediate},  and the last inequality follows by the assumption $\Delta_{\min}^2 \geq p^{-K/2} \log p$.
-
-    For $G_3$, note that by triangle inequality
-    \begin{align}
-        \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2 &\leq \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + 2\max_{a \in [r]}\onormSize{}{[\mW_{:a}^T\mX \mV^{(t)}]^s - [\mW_{:a}^T\mX \mV]^s }^2\\
-        & \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2}, \label{eq:g3xv_wxv}
-    \end{align}
-    where the last inequality follows by the derivation of $J_{11}$ in the proof of Lemma~\ref{lem:intermediate}.
-
-    % \begin{align}
-
-    %     %& \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2},
-    % \end{align}
-   % where the last inequality follows by the inequality~\eqref{eq:inter4} in Lemma~\ref{lem:intermediate}. 
-   Then we have 
-    \begin{align}
-        |G_3|^2 & \leq 2 \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2 \\
-        & \leq 2 \of{ \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  +  \onormSize{}{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  } \\
-        & \quad \cdot \max_{a \in [r]} \onorm{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
-        & \lesssim \of{\onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} } \of{ \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2}}+ c \Delta_{\min}^2 L^{(t)} \\
-        & \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 (\Delta_{\min}^2 + L^{(t)}) + c\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}},\label{eq:g3}
-    \end{align}
-    where the third inequality follows by the procedures to derive~\eqref{eq:g1} and \eqref{eq:g2}, the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:intermediate}, and the last inequality follows by the assumption $\Delta_{\min}^2 \geq p^{-K/2} \log p$.
-
-    Combining the inequalities~\eqref{eq:g1}, \eqref{eq:g2}, and \eqref{eq:g3}, we finish the proof of inequality~\eqref{eq:gib}.
-
-    % According to the inequality~\eqref{eq:inter_wa_wb} in the proof of Lemma~\ref{lem:intermediate}, we have
-    % \begin{equation}
-    %     \onorm{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  \lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2.
-    % \end{equation}
-    % Then, we have 
-    % \begin{align}
-    %     |G_3|^2 &\lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 \max_{a \in [r]} \onorm{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2 + |G_2|^2 \\
-    %     &\lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 \of{ \frac{r^2}{\Delta_{\min}^2} \of{L^{(t)}}^2 + K  \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} } + |G_2|^2\\
-    %     & \leq \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2  c \of{ \Delta_{\min}^2 + K L^{(t)} }+|G_2|^2, \label{eq:g3}
-    % \end{align}
-    % where the second inequality follows by Lemma~\ref{lem:intermediate}, and the last inequality follows by the assumption on $\Delta_{\min}^2$.
-
-    % Therefore, combining inequalities~\eqref{eq:g1}, \eqref{eq:g2}, \eqref{eq:g3} and apply the inequality of arithmetic and geometric means, .
-
-    \item Upper bound for $H_{ib}^{(t)}$~\eqref{eq:hib}. Recall the definition of $H_{ib}$ and rearrange the terms
-    \begin{align}
-        H_{ib} &= \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2\\
-        & \quad \quad \quad \quad \quad + \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
-        & =  \onormSize{}{ [ \mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 \\
-        & \quad \quad \quad \quad \quad + \of{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }  } \\
-        & \quad \quad \quad \quad \quad  - \of{\onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mY \mV^{(t)} ]^s } -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }   } \\
-        & = H_1 + H_2 + H_3,
-    \end{align}
-    where 
-    \begin{align}
-        H_1 &= \onormSize{}{ [ \mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 - \onormSize{}{ [ \mW_{:b}^T \mX \mV^{(t)}]^s -  [ \mW_{:b}^T \mY \mV^{(t)}]^s}^2, \\
-        H_2 &=  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }^2  , \\
-        H_3 & = 2 \ang{  [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s,  [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }.
-    \end{align}
-    For $H_1$, we have 
-    \begin{align}
-        |H_1| \leq  \frac{4 \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 }{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2}
-         \leq  \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} 
-        \lesssim  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2\label{eq:h1},
-    \end{align}
-    following the derivation of $G_2$ in inequality~\eqref{eq:g2} and the assumption that $\Delta_{\min}^2 \geq p^{-K/2} \log p$.
-    % \begin{align}
-    %     |H_1| &\leq 2 \max_{a \in [r]}  \frac{1}{ \onormSize{}{  \mW_{:a}^T \mX \mV^{(t)}}^2} \onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2  \\
-    %     & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}}\\
-    %     & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K} \Delta_{\min}^2}  \onorm{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
-    %     & \lesssim  \onorm{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 ,
-    % \end{align}
-    % where the first two inequalities follow the proof of $G_2$~\eqref{eq:g2} and the last inequality follows by the assumption on $\Delta_{\min}^2$.
-
-    For $H_2$, by the inequality~\eqref{eq:g3xv_wxv}, we have 
-    \begin{equation}\label{eq:h2}
-        |H_2| \lesssim 2 \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
-    \end{equation}
-    where the last inequality follows by the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
-
-    % For $H_2$,  sccording to the inequality~\eqref{eq:inter_wa_wb} in the proof of Lemma~\ref{lem:intermediate}, we have
-    % \begin{equation}
-    %     \onorm{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  \lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
-    % \end{equation}
-    % and thus 
-    % \begin{equation}\label{eq:h2}
-    %     |H_2| \lesssim  \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2.
-    % \end{equation}
-
-    For $H_3$,  by Cauchy-Schwartz inequality, we have 
-    \begin{equation}\label{eq:h3}
-         |H_3| \lesssim \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
-    \end{equation}
-    following the inequalities~\eqref{eq:g3xv_wxv} and \eqref{eq:h1}.
-
-    % For $H_3$, by Cauchy-Schwartz inequality, we have 
-    % \begin{align}
-    %     |H_3| &\lesssim    \onorm{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \\
-    %     &\lesssim  \frac{ r^{K-1/2} + K p^{1/2} r^{K/2+1/2} }{p^{K/2}} \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }\\
-    %     & \lesssim \Delta_{\min} \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }\\
-    %     & \lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2, \label{eq:h3}
-    % \end{align}
-    % where the third inequality follows by the assumption on $\Delta_{\min}^2$.
-
-    Therefore, combining inequalities~\eqref{eq:h1}, \eqref{eq:h2}, and \eqref{eq:h3}, we have finish the proof of inequality~\eqref{eq:hib}.
-
-\end{enumerate}
-
-\end{proof}
-
 \end{document}

### 2022-01-17 tensor_clustering/tensor_clustering_ieee/ieee_v1.tex
update

--- 2022-01-06 tensor_clustering/tensor_clustering_ieee/ieee.tex
@@
 \end{equation}
-where the triplet $(\hat z(r), \hat \tS(r), \hat \mtheta(r))$ are estimated parameters with given number of cluster $r$, $\tL_{\tY}$ denote the log-likelihood with observations $\tY$, and $p_e = r^K + p(\log r + 1) - 1$ is the effective number of parameters.
+where the triplet $(\hat z(r), \hat \tS(r), \hat \mtheta(r))$ are estimated parameters with given number of cluster $r$, $\tL_{\tY}$ denote the log-likelihood with observations $\tY$, and $p_e = r^K + p(\log r + 1) - r$ is the effective number of parameters.
 Particularly, we obtain the estimate $\hat \mtheta(r)$ by first calculating the reduced tensor $\tY^{\text{d}}$ with $\hat z(r)$ and then normalizing the row norms $\onormSize{}{\mY_{i:}^{\text{d}}}$ to 1 in each cluster; i.e.,
@@
 
+% \begin{table}[hbt]
+% \color{blue}
+%     \centering
+
+%     \begin{tabularx}{\textwidth}{c *{12}{Y}}
+%     \toprule
+%     Settings & \multicolumn{3}{c}{$p = 50, \sigma^2 = 0.25$} & \multicolumn{3}{c}{$p = 50, \sigma^2 = 1$} & \multicolumn{3}{c}{$p = 80, \sigma^2 = 0.25$} & \multicolumn{3}{c}{$p = 80, \sigma^2 = 1$}\\
+%     \cmidrule(lr){2-4} \cmidrule(l){5-7} \cmidrule(l){8-10} \cmidrule(l){11-13}
+%          True number of clusters $r$ & 2&3 & 4 & 2 & 3 & 4 & 2 & 3 & 4 & 2 & 3 & 4  \\
+%          \midrule
+%          Estimated number of clusters $\hat r$ &  2 & 3 & 3.9(0.3)& 2 & 3  & 3.1(0.5) & 2  & 3  & 4  & 2  & 3 & 3.9(0.31)   \\
+%      \bottomrule
+%     \end{tabularx}
+%     \caption{Estimated number of clusters given by BIC criterion under the small noise $(\sigma^2 = 0.25)$ and large noise $(\sigma^2 = 0.5)$ cases.
+Numbers in parentheses are standard deviations of $\hat r$ over 30 replications.}
+%     \label{tab:select}
+% \end{table}
+
 \begin{table}[hbt]
@@
     \centering
-    \begin{tabularx}{\textwidth}{c *{12}{Y}}
+
+    \begin{tabularx}{\textwidth}{c *{8}{Y}}
     \toprule
-    Settings & \multicolumn{3}{c}{$p = 50, \sigma^2 = 0.25$} & \multicolumn{3}{c}{$p = 50, \sigma^2 = 1$} & \multicolumn{3}{c}{$p = 80, \sigma^2 = 0.25$} & \multicolumn{3}{c}{$p = 80, \sigma^2 = 1$}\\
-    \cmidrule(lr){2-4} \cmidrule(l){5-7} \cmidrule(l){8-10} \cmidrule(l){11-13}
-         True number of clusters $r$ & 2&3 & 4 & 2 & 3 & 4 & 2 & 3 & 4 & 2 & 3 & 4  \\
+    Settings & \multicolumn{2}{c}{$p = 50, \sigma^2 = 0.25$} & \multicolumn{2}{c}{$p = 50, \sigma^2 = 1$} & \multicolumn{2}{c}{$p = 80, \sigma^2 = 0.25$} & \multicolumn{2}{c}{$p = 80, \sigma^2 = 1$}\\
+    \cmidrule(lr){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9}
+         True number of clusters $r$ & 2 & 4 & 2  & 4 & 2  & 4 & 2 & 4  \\
          \midrule
-         Estimated number of clusters $\hat r$ &  2 & 3 & 3.9& 2 & 3 & 3.1 & 2 & 3 & 4 & 2 & 3& 3.9   \\
+         Estimated number of clusters $\hat r$ &  2(0)  & 3.9(0.25)& 2(0)    & 3.1(0.52) & 2(0)    & 4(0)   & 2(0)    & 3.9(0.31)   \\
      \bottomrule
@@
     \caption{Estimated number of clusters given by BIC criterion under the small noise $(\sigma^2 = 0.25)$ and large noise $(\sigma^2 = 0.5)$ cases.
-We report the averaged estimated number of clusters $\hat r$ over 30 duplication.}
+Numbers in parentheses are standard deviations of $\hat r$ over 30 replications.}
     \label{tab:select}
@@
 We generate the data from an order-3 Gaussian model with $p = \{50,80\}$, $r = \{2,4\}$ and consider the noise $\sigma^2 \in \{ 0.25,1\}$ with $\alpha = 400$.
-Table~\ref{tab:select} implies that our BIC criterion exactly choose the true $r$ under all the settings with small $r = 2, 3$.
+Table~\ref{tab:select} implies that our BIC criterion exactly choose the true $r$ under all the settings with small $r = 2$.
 Note that the BIC underestimates the large true number of clusters $(r = 4)$ with smaller SNR $(\sigma^2 = 1)$, and the accuracy immediately increases with larger dimension $p = 80$.
@@
         & \lesssim \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \theta(i)m^2\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    }\\
-        &\leq C \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \theta(i)m^2\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    },\label{eq:zeta}
+        &\leq C \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    },\label{eq:zeta}
     \end{align}

### 2022-01-18 tensor_clustering/tensor_clustering_ieee/ieee_v3.tex
v3

--- 2022-01-06 tensor_clustering/tensor_clustering_ieee/ieee.tex
@@
 \usepackage{amsmath,amsfonts}
-%\usepackage{algorithmic}
 \usepackage{algorithm}
@@
 
-%\usepackage{appendix}
 \usepackage{fullpage} 
@@
 
-\newcommand{\distgap}{D_{\ell_2}}
 \newcommand{\Mat}{\text{Mat}}
@@
 
-\author{ Jiaxin Hu and Miaoyan Wang
+\author{ Jiaxin Hu and Miaoyan Wang\\
+University of Wisconsin - Madison
 %IEEE Publication Technology,~\IEEEmembership{Staff,~IEEE,}
@@
 Further, we develop an efficient polynomial-time algorithm that provably achieves exact clustering under mild signal conditions.
-The efficacy of our procedure is demonstrated through two data applications, one on human brain connectome project, and another on Peru Legislation network datasets.
+The efficacy of our procedure is demonstrated through two data applications, one on human brain connectome project, and another on Peru Legislation network dataset.
 \end{abstract}
@@
 
-% {
-% \color{red}
-
-% \textbf{Other conference paper to journal}
-% \begin{enumerate}
-%     \item Garvesh. Add more discussion about Theorem 5 and empirical simulations.
-%     %\item Eric Chi. Add one more remakr; One more section about the numerical issue; More comparison with other methods; add more future discussion.
-%     \item Anru. Add application in introduction; more remarks for algorithm and theorems; One more section about the application to a specific model; more experiments and discussions.
-
-%     \item Christina. Almost re-write the paper. 
-% \end{enumerate}
-% }
-
-% {
-% \color{blue}
-
-% \subsection{Our contribution.} 
-% \begin{itemize}
-%     \item We propose a generalized dTBM.
-%     \item Phase transition.
-%     \item We propose a multiway spherical clustering algorithm.
-% \end{itemize}
-
-% \subsection{Related work.} We emphasize the comparisons that set our work apart from earlier literature from two perspectives -- models and algorithms.
-
-% From the model perspective, our proposed model is related, but also distinct to following lines of works.
-% \begin{itemize}
-
-%     %\item \textit{Tensor block model.}
-
-%     \item \textit{Degree-corrected block model.}
-
-%     \item \textit{Degree-corrected block model for hypergraphs.}
-
-%     \item \textit{Other higher-order clustering models.}
-% \end{itemize}
-
-% From the algorithm perspective, our work is inspired by two key algorithmic techniques.
-
-% \begin{itemize}
-%     \item \textit{Spherical clustering.}
-
-%     \item \textit{Global-to-local strategy.}
-% \end{itemize}
-
-% \subsection{Binary multiway spherical clustering.}
-
-% \paragraph{Algorithm extension}
-
-% \paragraph{Theoretical guarantee}
-
-% }
-
 \section{Introduction}\label{sec:intro}
-
-%\fixme{Jiaxin}{Change the wording: \emph{individual/block effects} $\rightarrow$ \emph{individual/block parameters}, \emph{degree heretogeneity} $\rightarrow$ \emph{non-equal degrees}.}
-
 \IEEEPARstart{M}{ultiway} arrays have been widely collected in various fields including social networks \citep{anandkumar2014tensor}, neuroscience \citep{wang2017bayesian}, and computer science \citep{koniusz2016sparse}.
@@
 
-%To overcome the high-dimensionality, low-rankness is one of the popular structures utilized in tensor data analysis. Specifically, for an order-$K$ tensor $
-%\tY$ with low-rank, we have the following decomposition:
-%\begin{equation}\label{eq:low_rank}
- %   \tY = \tS \times_1 \mM_1 \cdots \times_K \mM_K, 
-%\end{equation}
-%where $\tS$ is the low-dimensional core tensor, $\mM_1, ..., \mM_K$ refer to the factor matrices, and $\times_k$ denotes the tensor-to-matrix product.
-
-This paper studies the problem of multiway clustering based on a data tensor.
+We study the problem of multiway clustering based on a data tensor.
 The goal of multiway clustering is to identify a checkerboard structure from a noisy data tensor.
-Figure~\ref{fig:intro} illustrates the noisy tensor and the underlying structures discovered by multiway clustering methods.
-The checkerboard structure serves as a meta tool to many popular structures including the low-rankness~\citep{young2018universality}, latent space models~\citep{wang2018learning}, and isotonic models~\citep{pananjady2020isotonic}.
+Figure~\ref{fig:intro} illustrates the noisy tensor and the underlying checkerboard structures discovered by multiway clustering methods.
 In the hypergraph example, the multiway clustering aims to identify the underlying block partition of nodes based on their higher-order connectivities; therefore, we also refer to the clustering as \emph{higher-order clustering}.
 The most common model for higher-order clustering is called \emph{tensor block model} (TBM)~\citep{wang2019multiway}, which extends the usual matrix stochastic block model~\citep{abbe2017community} to tensors.
-%The analysis tools developed for matrices,
 The matrix analysis tools, however, are sub-optimal for higher-order clustering.
@@
 Such degree heterogeneity appears commonly in social networks.
-{\color{blue} Ignoring the individual heterogeneity may seriously mislead the clustering results.} For example, regular block model fails to model {\color{blue} the member affiliation in} the Karate Club network~\citep{bickel2009nonparametric} without addressing
-%accounting for 
-degree heterogeneity. 
-%\fixme{Jiaxin}{Non-degree model ignoring the individual heterogeneity may seriously mislead the clustering results.
-Karate Club data analysis shows that clustering without heterogeneity tends to assign membership based on the degree of communication (two clusters for leaders and non-leaders) rather than the affiliation (two clusters for people from different clubs).
-}
+Ignoring the degree heterogeneity may seriously mislead the clustering results.
+For example, regular block model fails to model the member affiliation in the Karate Club network~\citep{bickel2009nonparametric} without addressing degree heterogeneity.
 
@@
 The dTBM allows varying values within the same community, thereby allowing a richer structure.
-{\color{blue} To solve dTBM, we project clustering objects to a unit sphere and consider the angle similarity; detailed algorithms are in Section~\ref{sec:alg}.
-On one hand, the \textit{spherical} clustering avoids the estimation of nuisance degree heterogeneity.
-On the other hand, the usage of angle similarity brings new challenges to the derivations of theoretical results, and the polar coordinates based techniques are equipped in the proof.
-}
-
-% {
-% \color{red}
-% We use the idea of \emph{spherical clustering} in initialization and iteration. Using angle-based similarity avoids the estimation of $\mtheta$.  
-% }
+To solve dTBM, we project clustering objects to a unit sphere and perform iterative clustering based on angle similarity.
+We refer to the algorithm as the \textit{spherical} clustering; detailed procedures are in Section~\ref{sec:alg}.
+The spherical clustering avoids the estimation of nuisance degree heterogeneity.
+The usage of angle similarity brings new challenges to the theoretical results, and we develop new polar-coordinate based techniques in the proofs.
 
@@
 Both TBM and dTBM have four communities on each mode, while dTBM allows a richer structure with degree heterogeneity.
-    %\fixme{Miaoyan}{Use horizontal figure} 
     }
@@
 
- \item We develop a general dTBM and establish the identifiability for the uniqueness of clustering using the notion of angle seperability.
+ \item We develop a general dTBM and establish the identifiability for the uniqueness of clustering using the notion of angle separability.
 
@@
 We characterize, for the first time, the critical signal-to-noise (SNR) thresholds in dTBMs, revealing the intrinsic distinctions among (vector) one-dimensional clustering, (matrix) biclustering, and (tensor) higher-order clustering.
-Specific SNR thresholds and algorithm behaviours are depicted in  Figure~\ref{fig:phase_axis}.
+Specific SNR thresholds and algorithm behaviors are depicted in  Figure~\ref{fig:phase_axis}.
 
@@
 The last two contributions, to our best knowledge, are new to the literature of dTBMs.
-%Table~\ref{tab:comp} highlights the improvement of our methods over previous results.
-Our algorithm addresses higher-order tensor data with degree heterogeneity, allows either binary or continuous entries, and achieves exponentially fast rate in misclustering error.
 
@@
 {\bf Related work.} 
-{\color{blue}
-Our work is closely related to, but also distinct from several lines of existing research. 
+Our work is closely related to but also distinct from several lines of existing research. Table~\ref{tab:comp} summarizes the most relevant models.
+
 \begin{itemize}
-    \item \textit{Block model for clustering.} Block models such as stochastic block model (SBM) and degree-corrected SBM has been widely used for matrix clustering problems.
-The theoretical properties and algorithm performance for block models have been well-studied; see the review paper~\citep{abbe2017community} and the references therein.
+    \item \textit{Block model for clustering.} Block models such as stochastic block model (SBM) and degree-corrected SBM have been widely used for matrix clustering problems.
+The theoretical properties and algorithm performance for matrix block models have been well-studied~\citep{gao2018community}; see the review paper~\citep{abbe2017community} and the references therein.
 However, The tensor counterparts are relatively less understood.
-Table~\ref{tab:comp} summarizes the most relevant models.
-Specifically,
-    \begin{itemize}
-        \item the tensor block model (TBM, \cite{wang2019multiway, han2020exact}), a higher-order extension of SBM, fails to allow degree heterogeneity.
-The Cartesian coordinates based analysis in \cite{han2020exact} is also non-applicable to handle the extra flexibility brought from heterogeneity.
-In contrast, our model addresses the degree heterogeneity, and the polar coordinates based tools are adapted for the theoretical analysis;
-        \item the hypergraph degree-corrected block model (hDCBM) proposed by \cite{ke2019community, yuan2018testing} accounts for degree heterogeneity.
-Whereas, the hDCBM is designed only for binary observations, and the proposed spectral algorithm in \cite{ke2019community} achieves a sub-optimal clustering accuracy in higher-order scenarios.
-In constrast, our model allows discrete and continuous entries, and achieves exponentially fast rate in clustering tasks.
+
+    \item  \textit{Tensor block model.} The tensor block model (TBM, \cite{wang2019multiway, han2020exact, ghoshdastidar2017consistency}) is a higher-order extension of SBM, but it fails to allow degree heterogeneity.
+The Cartesian coordinates based analysis in \cite{han2020exact} is non-applicable to handle the extra complexity brought by degree heterogeneity.
+In contrast, our model addresses the degree heterogeneity, and we develop polar-coordinate based tools for the theoretical analysis.
+
+    \item \textit{Degree-corrected block model.} The hypergraph degree-corrected block model (hDCBM) proposed by \cite{ke2019community, yuan2018testing} accounts for degree heterogeneity.
+However, the hDCBM is designed only for binary observations, and the proposed spectral algorithm in \cite{ke2019community} achieves sub-optimal polynomial clustering rate in higher-order scenarios.
+In constrast, our model allows discrete and continuous entries, and achieves \emph{exponentially} fast rate in clustering tasks.
 More importantly, to our best knowledge, we are the first to provide the statistical and computational limits analyses for the degree-corrected block model in tensor clustering.
-    \end{itemize}
+See Fig~\ref{fig:phase_axis} for overview of our results.
 
@@
 Despite the conceptual similarity, we address several fundamental challenges associated with this non-convex, non-continuous problem.
-We show the insufficiency of the conventional tensor HOSVD~\citep{kolda2009tensor}, and we develop a weighted higher-order initialization that relaxes the eigen-gap separation condition.
+We show the insufficiency of the conventional tensor HOSVD~\citep{de2000multilinear}, and we develop a weighted higher-order initialization that relaxes the singular-value gap separation condition.
 Furthermore, our local iteration leverages the angle-based clustering in order to avoid explicit estimation of degree heterogeneity.
 Our bounds reveal the interesting interplay between the computational and statistical errors.
-We show that our final estimate provably achieves the exact clustering within only polynomial-time complexity.
+We show that our final estimate \emph{provably} achieves the exact clustering within only polynomial-time complexity.
 
 \end{itemize}
-}
-
-%We emphasize the comparisons that set our work apart from earlier literature from two perspectives---models and algorithms.
-From the model perspective, our work extends the previous degree-corrected model from matrices to tensors.
-There is a huge literature on degree-corrected matrix models; see the review paper~\citep{abbe2017community} and the references therein.
-The tensor counterparts, however, are relatively less understood.
-Table~\ref{tab:comp} summarizes  the most relevant models for higher-order clustering.
-Earlier tensor methods either fail to allow degree heterogeneity~\citep{chien2018community,han2020exact,zhang2021exact,ghoshdastidar2017uniform,kim2018stochastic}, or suffer from sub-optimal misclustering rates~\citep{ke2019community,ghoshdastidar2017consistency,chi2020provable}.
-In contrast, our method addresses the degree heterogeneity, allows discrete and continuous entries, and achieves exponentially fast rate in clustering tasks.
-
-%From the algorithm perspective, our methods generalize the recent global-to-local strategy for matrix learning~\citep{gao2018community,chi2019nonconvex,yun2016optimal} to tensors~\citep{han2020exact,ahn2018hypergraph,kim2018stochastic}.
-Despite the conceptual similarity, we address several fundamental challenges associated with this non-convex, non-continuous problem.
-We show the insufficiency of the conventional tensor HOSVD~\citep{kolda2009tensor}, and we develop a weighted higher-order initialization that relaxes the eigen-gap separation condition.
-Furthermore, our local iteration leverages the angle-based clustering in order to avoid explicit estimation of degree heterogeneity.
-Our bounds reveal the interesting interplay between the computational and statistical errors.
-We show that our final estimate provably achieves the exact clustering within only polynomial-time complexity.
-
-% {
-% \color{red}
-
-% Our work is closely related to, but also distinct from several lines of existing research:
-% \begin{enumerate}
-
-% \item \textbf{Degree-corrected model.}
-%     \begin{itemize}
-%         \item Matrix. (1) Previous work, survey paper (2) Unique technical challenges in Tensor (3) matrix does not show statistical-computational gap;
-%         \item Tensor. (Comparison with Ke) (1) Ke does not provide the SNR limits (2) Ke's algorithm is sub-optimal (3) Ke focuses on Binary data. 
-%     \end{itemize}
-
-%     \item \textbf{Higher-order clustering.} (1) Previous work fails to account the heterogeneity, which is not realistic; (2) The extension of theoretical results highly challenging.
-Extra nuisance parameter makes it more difficult to establish the statistical and computational limits.
-
-%     \item \textbf{Spherical clustering.}
-
-%     \item \textbf{Global-to-local algorithm.} (1) Matrix global-to-local algorithm (2) Tensor global-to-local algorithm is used for non-degree algorithm.
-Extra nuisance parameter makes it necessary to consider angle-based similarity; angle-based clustering makes the Cartesian coordinates based analysis non-applicable for our setting.
-% \end{enumerate}
-
-% }
 
@@
 \resizebox{\textwidth}{!}{%
-\color{blue}
     \begin{tabular}{c|ccccc}
@@
     \hline
-        {\small Allow tensors}& $\times$ & $\surd$ & $\surd$ &$\surd$ & $\surd$  \\
-        {\small Allow heterogeneity} &  $\surd$ & $\times$ & $\surd$ &$\surd$ & $\surd$ \\
-        %Assortative assumption&  $\surd$ & $\times$ & $\times$ & $\times$\\
-       %{\small Allow various data} & $\times$ & $\surd$ & $\times$ & $\surd$  &$\surd$\\
-        {\small Eigen free clustering} &  $\surd$ & $\surd$ & $\times$ & $\times$ &$\surd$ \\
-        %{\small Misclustering rate (for order 2$^*$)}& $\exp(-p)$ & $\exp(-p)$ & $p^{-1}$ & $p^{-1}$& $\exp(-p)$\\
-        {\small Misclustering rate (for order $K^*$)}& - & $\exp(-p^{K/2})$ & $p^{-1}$& $p^{-2}$ &  $\exp(-p^{K/2})$\\
+      Allow tensors of arbitrary order & $\times$ & $\surd$ & $\surd$ &$\surd$ & $\surd$  \\
+        Allow degree heterogeneity &  $\surd$ & $\times$ & $\surd$ &$\surd$ & $\surd$ \\
+        Singular-value gap-free clustering &  $\surd$ & $\surd$ & $\times$ & $\times$ &$\surd$ \\
+        Misclustering rate (for order $K^*$)& - & $\exp(-p^{K/2})$ & $p^{-1}$& $p^{-2}$ &  $\exp(-p^{K/2})$\\
         \hline
@@
     }
-    \caption{Comparison between previous methods with our method. $^*$We list the result for order-K tensors with $K \geq 3$ and general number of communities $r = \tO(1)$.
-   % Our results allow general tensors of arbitrary order $K$; See Theorem~\ref{thm:refinement}.
-    }\label{tab:comp}
+    \caption{Comparison between previous methods with our method. $^*$We list the result for order-K tensors with $K \geq 3$ and general number of communities $r = \tO(1)$. }\label{tab:comp}
 \end{table*}
@@
 {\bf Notation.} We use lower-case letters (e.g., $a,b$) for scalars, lower-case boldface letters (e.g., $\ma,\mtheta$) for vectors, upper-case boldface letters (e.g., $\mX,\mY$) for matrices, and calligraphy letters (e.g., $\tX,\tY$) for tensors of order three or greater.
+We use $\mone_p$ to denote a vector of length $p$ with all entries to be 1.
 We use $|\cdot|$ for the cardinality of a set and $\ind\{\cdot\}$ for the indicator function.
@@
 
-For a matrix $\mY$, we use $\mY_{i:}$ to  denote the $i$-th row of the matrix.
 Let $\tY  \in \bbR^{p_1 \times \cdots \times p_K}$ be an order-$K$ $(p_1,...,p_K)$-dimensional tensor.
@@
 \begin{align}
-    &\tX(i_1,\ldots, i_K)\\
-    &=\sum_{(j_1,\ldots,j_K)}\tS(j_1,\ldots,j_K)\mM_1(i_1,j_1)\cdots \mM_K(i_K,j_K).
-\end{align} 
+    &\tX(i_1,\ldots, i_K)=\sum_{(j_1,\ldots,j_K)}\tS(j_1,\ldots,j_K)\mM_1(i_1,j_1)\cdots \mM_K(i_K,j_K).
+\end{align} 
+For a matrix $\mY$, we use $\mY_{i:}$ (respectively, $\mY_{:i}$) to denote the $i$-th row (respectively, $i$-th column) of the matrix.
+Similarly, for an order-3 tensor, we use $\tY_{::i}$ to denote the $i$-th matrix slide of the tensor.
 We use $\text{Ave}(\cdot)$ to denote the operation of taking averages across elements and $\text{Mat}_k(\cdot)$ to denote the unfolding operation that reshapes the tensor along mode $k$ into a matrix.
 For a symmetric tensor $\tY\in\mathbb{R}^{p\times \cdots \times p}$, we omit the subscript and use $\text{Mat}(\tY)\in\mathbb{R}^{p\times p^{K-1}}$ to denote the unfolding.
-For two sequences $\{a_p\}, \{b_p\}$, we denote $a_p\lesssim b_p$ if $\lim_{p\to\infty} a_p/b_p \leq c$ and $a_p = \Omega(b_p)$ if $c b_p \leq a_p \leq C b_p$, for some constants $c, C\geq 0$.
+For two sequences $\{a_p\}, \{b_p\}$, we denote $a_p\lesssim b_p$ or $a_p=\tO(b_p)$ if $\lim_{p\to\infty}a_p /b_p\leq c$ for some constant $c\geq 0$, $a_p=o(b_p)$ if $\lim_{p\to\infty}a_p/b_p =0$, and $a_p = \Omega(b_p)$ if both $b_p \lesssim a_p$ and $a_p\lesssim b_p$.
 Throughout the paper, we use the terms ``community'' and ``clusters'' exchangeably.
- %\fixme{Miaoyan}{\color{red} need to define $\cos(\ma,\mb)$. Be careful with signs.}
-
-{
-\color{blue}
 
@@
 In Section~\ref{sec:alg}, we provide a polynomial-time two-stage algorithm with misclustering rate guarantees.
-Numerical studies including the simulations to assess the theoretical results, comparison with other methods, and real data analysis on human brain connectome data and Peru legislation data are in Section~\ref{sec:simulation}.
-Last, we conclude our paper in Section~\ref{sec:con}.
-
-}
-
-\section{Model formulation}\label{sec:model}
+Numerical studies including the simulation, comparison with other methods, and two real dataset analyses are in Sections~\ref{sec:simulation}-\ref{sec:real}.
+The main technical ideas we develop for addressing main theorems are provided in Section~\ref{sec:mainproof}.
+Detailed proofs and extra theoretical results are provided in Appendix.
+
+\section{Model formulation and motivations}\label{sec:model}
 
@@
 Suppose we have an order-$K$ data tensor $\tY \in \bbR^{p \times \cdots \times p}$.
-For ease of notation, we focus on symmetric tensors in this section; our framework easily extends to general asymmetric tensors.
+For ease of notation, we focus on symmetric tensors in this section; the extension to general asymmetric tensors is provided in Section~\ref{subsec:exten}.
 Assume there exist $r \geq 2$  disjoint communities among the $p$ nodes.
@@
 Then, $z^{-1}(a)=\{i\in[p]\colon z(i)=a\}$ denotes the set of nodes that belong to the $a$-th community, and $|z^{-1}(a)|$ denotes the number of nodes in the $a$-th community.
-%, and we let $p_a = \sum_{i \in [p]}\ind \offf{ z_i= a}$ denote the size of $a$-th community. 
-Let $\mtheta=(\theta(1),\ldots,\theta(p))^T$ denote the degree heterogeneity for $p$ nodes. We consider the order-$K$ dTBM~\citep{ghoshdastidar2017consistency,ke2019community},
+Let $\mtheta=(\theta(1),\ldots,\theta(p))^T$ denote the degree heterogeneity for $p$ nodes.
+We consider the order-$K$ dTBM~\citep{ghoshdastidar2017consistency,ke2019community},
 \begin{equation}\label{eq:model_margin}
@@
 \normalsize
-where $\tS \in \bbR^{r \times \cdots \times r}$ is an order-$K$ tensor collecting the block means among communities, and $\tE\in\mathbb{R}^{p\times \cdots \times p}$ is a noise tensor consisting of independent mean-zero sub-Gaussian entries with variance bounded by $\sigma^2$.
+where $\tS \in \bbR^{r \times \cdots \times r}$ is an order-$K$ tensor collecting the block means among communities, and $\tE\in\mathbb{R}^{p\times \cdots \times p}$ is a noise tensor consisting of independent zero-mean sub-Gaussian entries with variance bounded by $\sigma^2$.
 The unknown parameters are $z$, $S$, and $\mtheta$.
@@
 \begin{equation}\label{eq:model_tensor}
- \mathbb{E}\tY = \tS \times_1 \mTheta \mM \times_2 \cdots \times_K  \mTheta \mM,
+\bbE\tY = \tS \times_1 \mTheta \mM \times_2 \cdots \times_K  \mTheta \mM,
 \end{equation}
@@
 Note that the discrete nature of $\mM$ renders our model~\eqref{eq:model_tensor} more challenging than Tucker decomposition.
-%The Tucker rank of the signal tensor $\bbE \tY$ is bounded by, but not necessarily equal to, $(r,...,r)$~\citep{wang2019multiway}.
-We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$
-%with the core tensor $\tS$ 
-if $\tY$ admits dTBM~\eqref{eq:model_tensor}. Here, we give two special cases of dTBM.
-
-{
-\color{blue}
-
-\subsection{Motivating Examples.} Here, we provide three applications to illustrate the practical necessity of dTBM.
+We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$ if $\tY$ admits dTBM~\eqref{eq:model_tensor}.
+The goal of clustering is to estimate $z$ from a single noisy tensor $\tY$.
+We are particularly interested in the high-dimensional regime where $p$ grows whereas $r=\tO(1)$.
+
+\subsection{Motivating examples} Here, we provide four applications to illustrate the practical necessity of dTBM.
 
@@
 
-\paragraph{Community detection in hypergraphs} Hypergraph network is a powerful tool to collect the complex entity relations with higher-order interactions~\citep{ke2019community}.
+\paragraph{Community detection in hypergraphs} Hypergraph network is a powerful tool to represent the complex entity relations with higher-order interactions~\citep{ke2019community}.
 A typical undirected hypergraph is denoted as $H = (V,E)$, where $V = [p]$ is the set of nodes and $E$ is the set of undirected hyperedges.
@@
 
-Similar with the adjacency matrix, it is natural to represent the $K$-uniform hypergraph by a binary order-$K$ adjacency tensor.
-Let $\tY \in \{0,1\}^{p \times \cdots \times p}$ denote the adjacency tensor, where the entries encode the presence or absence of hyperdeges among $p$ nodes.
-Specifically,
+It is natural to represent the $K$-uniform hypergraph using a binary order-$K$ adjacency tensor.
+Let $\tY \in \{0,1\}^{p \times \cdots \times p}$ denote the adjacency tensor, where the entries encode the presence or absence of order-$K$ edges among $p$ nodes.
+Specifically, for all $(i_1,\ldots,i_K) \in [p]^K$, we have
 \begin{equation}
     \tY(i_1,...,i_K) =  \begin{cases}
-    1  & \text{if }  (i_1,...,i_K) \in E\\
-    0 & \text{if }  (i_1,...,i_K) \notin E
-    \end{cases},
-\end{equation}
-for all $(i_1,\ldots,i_K) \in [p]^K$.
-
-Assume there exist $r$ disjoint communities among $p$ nodes.
-The equation~\eqref{eq:model_tensor} models $\mathbb{E}\tY$ with unknown degree heterogeneity $\mtheta$ and subgaussianity parameter $\sigma^2 = 1/4$.
+    1  & \text{if }  (i_1,...,i_K) \in E,\\
+    0 & \text{if }  (i_1,...,i_K) \notin E.
+    \end{cases}
+\end{equation}
+
+Assume there exist $r$ disjoint communities among $p$ nodes, and the connection probabilities depend on the community assignments and node effects. 
+Then, the equation~\eqref{eq:model_tensor} models $\bbE \tY$ with unknown degree heterogeneity $\mtheta$ and sub-Gaussianity parameter $\sigma^2 = 1/4$.
 
@@
 \paragraph{Multi-layer weighted network} 
-Multi-layer weighted network data consists of multiple networks with the same set of nodes.
-One representative example is the brain structural connectome data~\citep{zhang2019tensor}.
-The multi-layer weighted network $\tY$ has dimension of $p \times p \times L$, where $p$ denotes the number of brain regions in interest, and $L$ denotes the number of layer.
-Each of the $L$ networks describes one aspect of the brain connections, and the tensor $\tY$ include a mixture slices with continuous, binary, and count entries.
-
-% We use $G = (V, \{E_l\}_{l=1}^L)$ denote a $L$-layer undirected multi-layer network, where $E_l$ is the set of edges in the network of the $l$-th layer.
-Let $\mY_l \in \{0,1\}^{p\times p}$ denote the adjacency matrix for the $L$-th layer with nodes $V$ and edges $E_l$.
-Then, the multi-layer network is represented by an order-3 binary tensor $\tY \in \{0,1\}^{p \times p \times L}$, where
-% \begin{equation}
-%     \tY(i_1,i_2,l) = \mY_l(i_1,i_2),
-% \end{equation}
-% for all $(i_1,i_2) \in [p]^2$ and $l \in [L]$.
+Multi-layer weighted network data consists of multiple networks over the same set of nodes.
+One representative example is the brain connectome data~\citep{zhang2019tensor}.
+The multi-layer weighted network $\tY$ has dimension of $p \times p \times L$, where $p$ denotes the number of brain regions of interest, and $L$ denotes the number of layers (networks).
+Each of the $L$ networks describes one aspect of the brain connectivity, such as functional connectivity or structural connectivity.
+The resulting tensor $\tY$ consists of a mixture of slices with various data types.
 
@@
 \begin{equation}
-    \bbE[\tY] = \tS \times_1 \mTheta \mM \times_2 \mTheta \mM \times_3 \mTheta_l \mM_l, 
+    \bbE \tY = \tS \times_1 \mTheta \mM \times_2 \mTheta \mM \times_3 \mTheta_l \mM_l, 
 \end{equation}
@@
 \begin{equation}
-    \bbE[\tY] = \tS \times_1 \mTheta_1 \mM_1 \times_2 \mTheta_2 \mM_2 \times_3 \mTheta_3 \mM_3, 
-\end{equation}
-where $\{(\mtheta_k, \mM_k)\}_{k=1}^3$ refers to the heterogeneity and membership for genes, individuals, and tissues.
-
-}
-% \begin{example}[Gaussian TBM]
-% Let $\theta(i)= 1$ for all $i \in [p]$ and $\tE$ be a noise tensor with i.i.d.\ $N(0,\sigma^2)$ entries.
-Our dTBM reduces to a non-degree Gaussian TBM, which is widely used in previous clustering algorithms~\citep{wang2019multiway,chi2020provable,han2020exact}.
-The theoretical results in TBM serve as benchmarks for dTBM.
-% \end{example}
-
-% \begin{example}[Binary dTBM]
-% Consider a $K$-uniform hypergraph $H = (V,E)$, where $V = [p]$ collects the nodes with $r$ disjoint communities and $E$ collects all the $K$-way hyperedges.
-Let $\tY \in \{0,1\}^{p \times \cdots \times p}$ denote the adjacency tensor, where the entries encode the presence or absence of hyperdeges among $p$ nodes.
-Specifically, let $\tY(i_1,...,i_K) = 1$ if  $(i_1,...,i_K) \in E$, otherwise, $\tY(i_1,...,i_K) = 0$, for all $(i_1,\ldots,i_K) \in [p]^K$.
-The equation~\eqref{eq:model_tensor} models $\mathbb{E}\tY$ with unknown degree heterogeneity and subgaussianity parameter $\sigma^2 = 1/4$.
-% \end{example}
+    \bbE\tY = \tS \times_1 \mTheta_1 \mM_1 \times_2 \mTheta_2 \mM_2 \times_3 \mTheta_3 \mM_3, 
+\end{equation}
+where $\{(\mtheta_k, \mM_k)\}_{k=1}^3$ represents the heterogeneity and membership for genes, individuals, and tissues. \\
 
@@
 Our dTBM uses fewer block parameters than TBM.
-Let the subscripts ``deg" and ``non" denote quantities in the models with and without degrees, respectively.
-Then, every $r_{\text{non}}$-block tensor can be represented by a degree-corrected $r_{\text{deg}}$-block tensor with $r_{\text{deg}}\leq r_{\text{non}}$.
-In particular, there exist tensors with $r_{\text{non}}=p$ but $r_{\text{deg}}=1$, so the reduction in $r$ can be dramatic from $p$ to 1.
+In particular, every non-degree $r_1$-block tensor can be represented by a \emph{degree-corrected} $r_2$-block tensor with $r_2\leq r_1$.
+In particular, there exist tensors with $r_1=p$ but $r_2=1$, so the reduction in model complexity can be dramatic from $p$ to 1.
 This fact highlights the benefits of introducing degree heterogeneity in higher-order clustering tasks.
@@
 
-% In addition, the following example highlights the benefits of degree correction in clustering tasks. 
-% \begin{example}[Comparison to non-degree models]\label{ex:benefit}
-% Every $r_{\text{non}}$-block tensor can be represented by degree-corrected $r_{\text{deg}}$-block tensor with $r_{\text{deg}}\leq r_{\text{non}}$.
-In particular, there exist tensors with $r_{\text{non}}=p$ but $r_{\text{deg}}=1$.
-% \end{example}
-% %In principle, every $(p,\ldots,p)$-dimensional tensor can be represented by a $p$-block tensor.
-% The above example shows that our dTBM uses fewer block parameters than TBM, and the reduction in $r$ can be dramatic from $p$ to 1.
-This fact shows the advantages of introducing degree heterogeneity in high-dimensional clustering.
-
 \subsection{Identifiability under angle gap condition}
-The goal of clustering is to estimate the partition function $z$ from model~\eqref{eq:model_tensor}. We use $\tP$ to denote the following parameter space for $(z,\tS,\mtheta)$,
+The goal of clustering is to estimate the partition function $z$ from model~\eqref{eq:model_tensor}.
+For ease of notation, we focus on symmetric tensors; the extension to non-symmetric tensors are similar.
+We use $\tP$ to denote the following parameter space for $(z,\tS,\mtheta)$,
 \begin{align}\label{eq:family}
@@
 \normalsize
-where $c_i>0$'s are universal constants.  
-%We briefly describe the rationale of the constraints in~\eqref{eq:family}. 
-First, the entrywise positivity constraint on  $\mtheta\in\mathbb{R}^p_{+}$ is imposed to avoid sign ambiguity between entries in $\mtheta_{z^{-1}(a)}$ and $\tS$ {\color{blue} and thereof allow the trigonometric $\cos$ to describe the angle similarity in the following Assumption~\ref{assmp:min_gap} and Sub-algorithm~\hyperref[alg:main]{2} in Section~\ref{sec:alg}}.
-This constraint can be achieved without sacrificing model flexibility, by using a slightly larger dimension of $\tS$ in the factorization~\eqref{eq:model_tensor}; see Example~\ref{ex:positive}.
-Second,
-%the quantity $|z^{-1}(a)|$ denotes the number of nodes in $a$-th community. T
-the constants $c_1, c_2$ in the $|z^{-1}(a)|$ bound assume the roughly balanced size across $r$ communities.
-Third, the constants $c_3, c_4 $ in the magnitude of $\text{Mat}(\tS)_{a:}$ requires no purely zero slide in $\tS$, so the core tensor $\tS$ is not trivially reduced to lower rank.
+where $c_i>0$'s are universal constants. We briefly describe the rationale of the constraints in~\eqref{eq:family}. 
+First, the entrywise positivity constraint on  $\mtheta\in\mathbb{R}^p_{+}$ is imposed to avoid sign ambiguity between entries in $\mtheta_{z^{-1}(a)}$ and $\tS$.
+This constraint allows the trigonometric $\cos$ to describe the angle similarity in the Assumption~\ref{assmp:min_gap} below and Sub-algorithm~\hyperref[alg:main]{2} in Section~\ref{sec:alg}.
+Note that the positivity constraint can be achieved without sacrificing model flexibility, by using a slightly larger dimension of $\tS$ in the factorization~\eqref{eq:model_tensor}; see Example~\ref{ex:positive} below.
+Second, recall that the quantity $|z^{-1}(a)|$ denotes the number of nodes in $a$-th community.
+The constants $c_1, c_2$ in the $|z^{-1}(a)|$ bound assume the roughly balanced size across $r$ communities.
+Third, the constants $c_3, c_4 $ in the magnitude of $\text{Mat}(\tS)_{a:}$ requires no purely zero slide in $\tS$, so the core tensor $\tS$ is not trivially reduced to a lower rank.
 Lastly, the $\ell_1$ normalization $\onormSize{}{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|$ is imposed to avoid the scalar ambiguity between $\mtheta_{z^{-1}(a)}$ and $\tS$.
@@
 Here we provide an example to show the positivity constraints on $\mtheta$ incurs no loss on the model flexibility. 
-{\color{blue}
-We consider a order-3 dTBM with core tensor $\tS = 1$ and degree $\mtheta = (1,1,-1, -1)^T$. We have the mean tensor 
+Consider an order-3 dTBM with core tensor $\tS = 1$ and degree $\mtheta = (1,1,-1, -1)^T$. We have the mean tensor 
 \begin{equation}
@@
 where $\mTheta = \text{diag}(\mtheta)$ and $\mM = (1,1,1,1)^T$.
-Note that $\tX \in \bbR^{4 \times 4 \times 4}$ is a 1-block tensor with mixed-signed degree $\mtheta$, and the mode-3 slices of $\tX$ are
+Note that $\tX \in \bbR^{4 \times 4 \times 4}$ is a 1-block tensor with \emph{mixed-signed} degree $\mtheta$, and the mode-3 slices of $\tX$ are
 \begin{equation}
@@
 \end{equation}
-Now, instead of original decomposition, we encode $\tS$ as a 2-block tensor with positive-signed degree
+Now, instead of original decomposition, we encode $\tX$ as a 2-block tensor with \emph{positive-signed} degree. Specifically, we write
 \begin{equation}
@@
 \end{equation}
-where $\tS' \in \bbR^{2 \times 2 \times 2}$ have mode-3 slices 
+where $\mTheta' = \text{diag}(\mtheta') = \text{diag}(1,1, 1, 1)$, the core tensor $\tS' \in \bbR^{2 \times 2 \times 2}$ has mode-3 slices, and the membership matrix $\mM'\in\{0,1\}^{4\times 2}$ defines the clustering $z'\colon[4]\to[2]$,
 \begin{equation}
@@
     \end{bmatrix}, \quad \mM' = \begin{bmatrix}
-    \mI_2 & {\bf 0}\\
-    {\bf 0} &  \mI_2
-    \end{bmatrix},
-\end{equation}
-and $\mTheta' = \text{diag}(\mtheta') = \text{diag}(1,1, 1, 1)$.
+   1 & 0 \\
+   1 &0 \\
+    0 & 1 \\
+    0&1
+    \end{bmatrix}.
+\end{equation}
  The triplet $(z',\tS', \mtheta')$ lies in our parameter space~\eqref{eq:family}.
 In general, we can always reparameterize a block-$r$ tensor with mixed-signed degree using a block-$2r$ tensor with positive-signed degree.
-Since we assumes $r=\tO(1)$ throughout the paper, the splitting does not affect the error rates of our interest.
-}
-
-% Without loss of generality, we consider the order-2 (i.e., matrix) block models. Consider a $4$-by-$4$ matrix with core matrix $\mS = 1$ and degree $\mtheta=(1,1,-1,-1)^T$. We have the mean matrix 
-% \begin{equation}
-%     \mX = \mS \times_1 \mTheta \mM \times_2 \mTheta\mM = \begin{bmatrix}
-% 1 & 1 & -1& -1\\
-% 1 & 1 & -1& -1\\
-% -1 & -1 & 1& 1\\
-% -1 & -1 & 1& 1
-% \end{bmatrix},
-% \end{equation}
-% $ \text{with}\quad \mTheta=\text{diag}(\mtheta)$. Note that $\mX$ is a 1-block matrix with mixed-signed degree $\mtheta$. 
-% Now, instead of original decomposition, we encode $\mX$ as a $2$-block matrix with positive-signed degree,
-% \begin{equation}
-% \mX= \mS' \times_1 \mTheta' \mM \times_2 \mTheta' \mM,\quad \text{with}\quad   \mS' = \begin{bmatrix} 1 & -1 \\
-%     -1 & 1
-%     \end{bmatrix}
-% \end{equation}
-%and $ \mTheta'=\text{diag}(\mtheta')=\text{diag}(1, 1, 1, 1)^T$, where the triplet $(z',\mS', \mtheta')$ lies in our parameter space~\eqref{eq:family}.
-In general, we can always reparameterize a block-$r$ tensor with mixed-signed degree using a block-$2r$ tensor with positive-signed degree.
-Since we assumes $r=\tO(1)$ throughout the paper, the splitting does not affect the error rates of our interest.
-\end{example}
-
-We first provide the identifiablity conditions for our model before estimation procedures.
+Since we assume $r=\tO(1)$ throughout the paper, the splitting does not affect the error rates of our interest.\end{example}
+
+We now provide the identifiability conditions for our model before estimation procedures.
 When $r=1$, the decomposition~\eqref{eq:model_tensor} is always unique (up to cluster label permutation) in $\tP$, because dTBM is equivalent to the rank-1 tensor family under this case.
-When $r\geq 2$, the Tucker rank of signal tensor $\mathbb{E}\tY$ in~\eqref{eq:model_tensor} is bounded by, but not necessarily equal to, the number of blocks $r$~\citep{wang2019multiway}.
+When $r\geq 2$, the Tucker rank of signal tensor $\bbE\tY$ in~\eqref{eq:model_tensor} is bounded by, but not necessarily equal to, the number of blocks $r$~\citep{wang2019multiway}.
 Therefore, one can not apply the classical identifiability conditions for low-rank tensors to dTBM.
@@
 
-\begin{assumption}[Angle gap] \label{assmp:min_gap}Let $\mS = \text{Mat}(\tS)$. Assume the minimal gap between normalized rows of $\mS$ is bounded away from zero, i.e.,
+\begin{assumption}[Angle gap] \label{assmp:min_gap}Let $\mS = \text{Mat}(\tS)$. Assume the minimal gap between normalized rows of $\mS$ is bounded away from zero; i.e.,
 \begin{equation}\label{eq:minimal_gap}
-    \Delta_{\min} := \min_{a \neq b \in [r]} \onorm{\frac{\mS_{a:}}{\onormSize{}{ \mS_{a:}}} - \frac{\mS_{b:}}{\onormSize{}{ \mS_{b:}}} }>0, \quad \text{for} \quad r \geq 2,
-\end{equation}
-and set the convention $\Delta_{\min} = 1$ for $r = 1$.
-    Equivalently, none of the two rows in $\mS$ are parallel, i.e., when $r \geq 2$, $ \max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}}  = 1-\Delta^2_{\min}/2<1$.
+    \Delta_{\min} := \min_{a \neq b \in [r]} \onorm{\frac{\mS_{a:}}{\onormSize{}{ \mS_{a:}}} - \frac{\mS_{b:}}{\onormSize{}{ \mS_{b:}}} }>0, \quad \text{for} \quad r \geq 2.
+\end{equation}
 \end{assumption}
+We make the convention $\Delta_{\min} = 1$ for $r = 1$.
+Equivalently, \eqref{eq:minimal_gap} says that none of the two rows in $\mS$ are parallel; i.e., $\max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}}  = 1-\Delta^2_{\min}/2<1$.
 The quantity $\Delta_{\min}$ characterizes the non-redundancy among clusters measured by angle separation.
-The definition~\eqref{eq:minimal_gap} is well posed because of the lower bound on $\onorm{\mS_{a:}}$ in~\eqref{eq:family}.
-The following theorem shows that the angle gap separation is sufficient and necessary for parameter identifiability under dTBM.
-
-\begin{thm}[Identifiability]\label{thm:unique} Consider the dTBM with $r\geq 2$.
+The denominators involved in definition~\eqref{eq:minimal_gap} are well posed because of the lower bound on $\onorm{\mS_{a:}}$ in~\eqref{eq:family}.
+
+Our first main result is the following theorem showing the sufficiency and necessarily of the angle gap separation condition for the parameter identifiability under dTBM.
+
+\begin{thm}[Model identifiability]\label{thm:unique} Consider the dTBM with $r\geq 2$.
 The parameterization~\eqref{eq:model_tensor} is unique in $\tP$ up to cluster label permutations, if and only if Assumption~\ref{assmp:min_gap} holds.
@@
 
-\section{Statistical-computational gaps for higher-order tensors}\label{sec:limits}
-
-In this section, we study the statistical and computational limits of dTBM. We reserve the term higher-order tensors for tensors of order $K\geq 3$. We propose signal-to-noise ratio~(SNR),
+\section{Statistical-computational limits for higher-order tensors}\label{sec:limits}
+
+In this section, we study the statistical and computational limits of dTBM. We propose signal-to-noise ratio~(SNR),
 \begin{align}\label{eq:gamma}
@@
 Note that $1$-block dTBM does not belong to the space $\tP(\gamma)$ when $\gamma < 0$ by Assumption~\ref{assmp:min_gap}.
-Our goal is to characterize the clustering accuracy with respect to $\gamma$ when $r \geq 2$.
-Let $\hat z$ and $z$ be estimated and true clustering functions in the family~\eqref{eq:family}.
+Our goal is to characterize the clustering accuracy with respect to $\gamma$.
+Let $\hat z$ and $z$ be the estimated and true clustering functions in the family~\eqref{eq:family}.
 Define the misclustering error by
@@
 
-In Sections~\ref{sec:statlimit} and \ref{sec:complimit}, we provide the lower bounds of $\ell (\hat z, z)$ for general Gaussian dTBMs without symmetric assumptions.
-For general (asymmetric) dTBMs, we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode.
+In Sections~\ref{sec:statlimit} and \ref{sec:complimit}, we provide the lower bounds of $\ell (\hat z, z)$ for general Gaussian dTBMs~\eqref{eq:model_margin} without symmetric assumptions.
+For general (asymmetric) Gaussian dTBMs, we assume Gaussian noise $\tE(i_1,\ldots,i_K)\stackrel{\text{i.i.d.}}{\sim} N(0,\sigma^2)$, and we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode.
 For notational simplicity, we still use $z$ and $\tP(\gamma)$ for this general (asymmetric) model.
@@
 
-%For technical simplicity, we will present our main theory with a focus on Gaussian tensors. Our algorithm and comparison extend to general sub-Gaussian models with little modification.
-
 \subsection{Statistical critical values}\label{sec:statlimit}
-Our first main result is to show the minimax lower bound of SNR for exact recovery in dTBM. 
-%We call a dTBM general if no symmetric assumption is imposed to the signal tensor. 
+The statistical limit means the minimal SNR required for solving dTBMs with \emph{unlimited computational cost.} Our following result shows the minimax lower bound of SNR for exact recovery in dTBM.
+
 \begin{thm}[Statistical lower bound]\label{thm:stats} Consider general Gaussian dTBMs under the parameter space $\tP(\gamma)$ with $K\geq 1$.
@@
 The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to, maximum likelihood estimation (MLE)~\citep{wang2019multiway} and trace maximization~\citep{ghoshdastidar2017uniform}.
-Our derived SNR threshold $-(K-1)$ is also a minimax upper bound, because MLE achieves exact recovery when $\gamma > -(K-1)$.
+As we will show in Section~\ref{sec:alg}, the SNR threshold $-(K-1)$ is also a minimax upper bound, because MLE achieves exact recovery when $\gamma > -(K-1)$.
 Hence, the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM.
@@
 \subsection{Computational critical values}\label{sec:complimit}
+In this section, we derive the computational limits of dTBMs.
+The computational limit means the minimal SNR required for exactly recovery with \emph{polynomial-time} computational cost.
 An important ingredient to establish the computational limits is the \emph{hypergraphic planted clique (HPC) conjecture} \citep{zhang2018tensor, brennan2020reducibility}.
 The HPC conjecture indicates the impossibility of fully recovering the planted cliques with polynomial-time algorithm when the clique size is less than the number of vertices in the hypergraph.
-%\fixme{Miaoyan}{Is this condition used in our Theorem 3?}. 
-The formal statement of HPC detection and conjecture can be found in Definition~\ref{def:HPC} and Conjecture~\ref{hypo:HPC} as following.
-
-\begin{defn}[Hypergraphic planted clique (HPC) detection]~\label{def:HPC} Consider an order-$K$ hypergraph $H = (V,E)$ where $V = [p]$ collects vertices and $E$ collects all the $K$-way edges.
+The formal statement of HPC detection conjecture is provided in Definition~\ref{def:HPC} and Conjecture~\ref{hypo:HPC} as follows.
+
+\begin{defn}[Hypergraphic planted clique (HPC) detection]~\label{def:HPC} Consider an order-$K$ hypergraph $H = (V,E)$ where $V = [p]$ collects vertices and $E$ collects all the order-$K$ edges.
 Let $\tH_k(p, 1/2)$ denote the Erd\H{o}s-R\'{e}nyi $K$-hypergraph where the edge $(i_1,\ldots, i_K)$ belongs to $E$ with probability $1/2$.
@@
 Note that the clique size $\kappa$ can be a function of $p$, denoted $\kappa_p$.
-
-The order-$K$ HPC detection aims to identify whether there exists a planted clique hidden in a Erd\H{o}s-R\'{e}nyi $K$-hypergraph.
-We formulate HPC detection as the following hypothesis testing problem
+The order-$K$ HPC detection aims to identify whether there exists a planted clique hidden in an Erd\H{o}s-R\'{e}nyi $K$-hypergraph.
+The HPC detection is formulated as the following hypothesis testing problem
 \begin{equation}
@@
 
-\begin{conjecture}[Hypergraphic planted clique (HPC) conjecture]\label{hypo:HPC} Consider the HPC detection problem in Definition~\ref{def:HPC}.
+\begin{conjecture}[HPC conjecture]\label{hypo:HPC} Consider the HPC detection problem in Definition~\ref{def:HPC}.
 Suppose the sequence $\{\kappa_p\}$ such that $\limsup_{p \rightarrow \infty} \log \kappa_p/ \log \sqrt{p} \leq (1 - \tau)$.
-Then, for any sequence of polynomial-times test $\{ \varphi_p\}: H \mapsto \{0,1\}$ we have
+Then, for every sequence of polynomial-time test $\{ \varphi_p\}: H \mapsto \{0,1\}$ we have
 \begin{equation}
@@
 \begin{rmk}[Statistical-computational gaps]
-Now, we have established the phase transition of exact clustering under order-$K$ dTBM combing Theorems~\ref{thm:stats} and \ref{thm:comp}.
+Now, we have established the phase transition of exact clustering under order-$K$ dTBM by combing Theorems~\ref{thm:stats} and \ref{thm:comp}.
 Figure~\ref{fig:phase_axis} summarizes our results of critical SNRs when $K \geq 2$.
+In the weak SNR region $\gamma < -(K-1)$, no statistical estimator succeeds in degree-corrected higher-order clustering.
+In the strong SNR region $\gamma  > -K/2$, our proposed algorithm precisely recovers the clustering in polynomial time.
+In the moderate SNR regime, $-(K-1)\leq \gamma \leq -K/2$, the degree-corrected clustering problem is statistically easy but computationally hard.
 Particularly, dTBM reduces to matrix degree-corrected model when $K =2$, and the statistical and computational bounds show the same critical value.
@@
 where $\mY \in \bbR^{p \times d}$ collects $n$ data points in $\bbR^d$, $\mS \in \bbR^{r \times d}$ collects the $d$-dimensional centroids for $r$ clusters, and $\mTheta \in \bbR^{p \times p}, \mM \in \{0,1\}^{p \times r}, \mE \in \bbR^{p \times d}$ have the same meaning as in dTBM.
-\cite{lu2016statistical} implies polynomial-times algorithms are able to achieve the statistical minimax lower bound in GMM.
+\cite{lu2016statistical} implies that polynomial-time algorithms are able to achieve the statistical minimax lower bound in GMM.
 Therefore, we conclude that the statistical-to-computational gap emerges only for higher-order tensors with $K \geq 3$.
@@
 
-%\fixme{Miaoyan}{What happens when $K=1$??} \textcolor{blue}{In our context, when $K=1$ the core tensor $\tS$ reduces to a vector and thus $\Delta_{\min}^2 = 0$.
-\fixme{Miaoyan}{Not correct.
-When $K=1$, our model reduces to Gaussian mixture models $\tY=\tS\mM\mTheta$ with $\tS\in\mathbb{R}^{r\times p}$, unknown membership matrix is $\mM\times\{0,1\}^{p\times n}$, and data $\tY\in\mathbb{R}^{n\times p}$.
-The goal is to cluster $n$ vectors into $r$ clusters, all in dimension $\bbR^p$.
-Add a remark.
-citep earlier work about (lack of) stat-comp gap.} } \textcolor{red}{\cite{lu2016statistical} implies there does not exist a statistical-computational bound for non-degree corrected gaussian mixture model.
-can replace this reference if find a reference for degree-corrected cmm.}
-
 \begin{rmk}[Comparison with non-degree models]
 We compare our results to non-degree tensor models.
-%Our parameter space of interest consists of discrete structure (for $z$) and continuous structure (for $\tS$ and $\mtheta$). 
 The allowance of degree heterogeneity $\mtheta$ makes the model more flexible, but it incurs extra statistical and computational complexity.
 Fortunately, we find that the extra complexity does not render the estimation of $z$ qualitatively harder; see the comparison of our phase transition with non-degree TBM~\citep{han2020exact}.
-%In next section, we will supplement the general theory by providing an efficient algorithm that achieves empirical accuracy in the presence of unknown degree heterogeneity.
 \end{rmk}
 
-%\subsection{Comparison with non-degree models}
-
-% We have established the phase transition of exact clustering under dTBM.
-Figure~\ref{fig:phase_axis} summarizes our results about the critical SNRs.
-In the weak SNR region $\gamma < -(K-1)$, no statistical estimator succeeds in degree-corrected higher-order clustering.
-In the strong SNR region $\gamma  > -K/2$, our proposed algorithm precisely recovers the clustering in polynomial time.
-In the moderate SNR regime, $-(K-1)\leq \gamma \leq -K/2$, the degree-corrected clustering problem is statistically easy but computationally hard.
-% We find that the statistical-to-computational gap emerges only for higher-order tensors with $K \geq 3$.
-The result reveals the intrinsic distinctions between  (matrix) biclustering and (tensor) higher-order clustering.
-
-% We compare our results to non-degree tensor models.
-Our parameter space of interest consists of discrete structure (for $z$) and continuous structure (for $\tS$ and $\mtheta$).
-The allowance of degree heterogeneity $\mtheta$ makes the model more flexible, but it incurs extra statistical and computational complexity.
-Fortunately, we find that the extra complexity does not render the estimation of $z$ qualitatively harder; see the comparison of our phase transition with non-degree TBM~\citep{han2020exact}.
-In next section, we will supplement the general theory by providing an efficient algorithm that achieves empirical accuracy in the presence of unknown degree heterogeneity.
-
 \section{Polynomial-time algorithm under mild SNR}\label{sec:alg}
-We present a two-stage clustering algorithm.
+In this section, we present an efficient polynomial-time clustering algorithm under mild SNR.
 The procedure takes a global-to-local approach.
@@
 The global step finds the basin of attraction with polynomial miclustering error, whereas the local iterations improve the initial clustering to exact recovery.
-Both steps are critical to obtain a satisfactory algorithm output.{\color{blue} In this section, we first use the symmetric tensor as a working example to describe the algorithm procedures to gain insight.
+Both steps are critical to obtain a satisfactory algorithm output.
+In what follows, we first use the symmetric tensor as a working example to describe the algorithm procedures to gain insight.
 Our theoretical analysis focuses on the noisy tensor with i.i.d.\ sub-Gaussian noise such as Gaussian and uniform observations.
-The extensions for asymmetric tensor and Bernoulli observation and other practical issues are in subsection~\ref{subsec:exten}.}
-
-\begin{figure}[t]
+The extensions for asymmetric tensor and Bernoulli observation and other practical issues are in Section~\ref{subsec:exten}.
+
+\begin{figure}[ht!]
 \centering
@@
 We start with weighted higher-order clustering algorithm as initialization.
-{\color{blue} We take the order-3 tensor in illustration for simplicity.}  Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
+We take an order-3 symmetric tensor as illustration for insight.
+Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
 By model~\eqref{eq:model_tensor}, for all $i \in [p]$, we have
@@
 The denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) estimates $\tX$ from $\tY$ by a double projection spectral method.
-{\color{blue}
 The first projection performs HOSVD~\citep{de2000multilinear} via $\mU_{\text{pre}} = \text{SVD}_{r} \of{ \text{Mat}(\tY) }$, where $\text{SVD}_r(\cdot)$ returns the top-$r$ left singular vectors.
-The second projection performs HOSVD on the projected $\tY$ onto the multlilinear Kronecker space $\mU_{\text{pre}} \otimes  \mU_{\text{pre}}$; i.e.,
+The second projection performs HOSVD on the projected $\tY$ onto the multilinear Kronecker space $\mU_{\text{pre}} \otimes  \mU_{\text{pre}}$; i.e.,
 \begin{equation}\label{eq:two-step_factor}
-    \hat \mU = \text{SVD}_{r} \of{\text{Mat}\of{ \tY\times_1  \mU_{\text{pre}}^T \times_2  \mU_{\text{pre}}^T }}.
+    \hat \mU = \text{SVD}_{r} \of{\text{Mat}\of{ \tY\times_1  \mU_{\text{pre}} \mU_{\text{pre}}^T \times_2   \mU_{\text{pre}} \mU_{\text{pre}}^T }}.
 \end{equation}
@@
 \end{equation}
-%Compared with conventional HOSVD, our double projection spectral method is shown to improve the estimation substantially for higher-order tensors~\citep{han2020exact}. 
-The double projection improves usual matrix spectral methods in order to alleviate the noise tensor.
-
-}
+The double projection improves usual matrix spectral methods in order to alleviate the noise effects for $K\geq 3$~\citep{han2020exact}.
 
 The clustering step (lines 3-5 in Sub-algorithm~\hyperref[alg:main]{1}) performs the weighted $k$-means clustering. 
-{
-\color{blue}
 We write $\hat \mX=\text{Mat}(\hat \tX)$, and normalize the rows into $\hat \mX^s_{i:}=\onormSize{}{\hat \mX_{i:}}^{-1}\hat \mX_{i:}$ as a surrogate of $\theta(i)^{-1} \mX_{i:}$.
 Then, a weighted $k$-means clustering is performed on the normalized rows with weights equal to $\onormSize{}{\hat \mX_{i:}}^2$.
-}
 The choice of weights is to bound the $k$-means objective function by the Frobenius-norm accuracy of $\hat \tX$.
 Unlike existing clustering algorithm~\citep{ke2019community}, we apply the clustering on the unfolded tensor $\hat \mX$ rather than on the factors $\hat \mU$.
-This strategy relaxes the eigen-gap separation condition~\citep{gao2018community, han2020exact}.
-{
-\color{blue}
+This strategy relaxes the singular-value gap condition~\citep{gao2018community, han2020exact}.
 We assign degenerate rows with purely zero entries to an arbitrarily random cluster; these nodes are negligible in high-dimensions because of the lower bound on $\onormSize{}{\text{Mat}(\tS)_{a:}}$ in~\eqref{eq:family}.
-The final result gives the initial clustering assignment $\hat z^{(0)}$.
-}
+The final result gives the initial cluster assignment $\hat z^{(0)}$.
 Full procedures are provided in Sub-algorithm~\hyperref[alg:main]{1}.
-
-We now establish the misclustering error rate of initialization. We call $\mtheta$ is balanced, if the relative extent of heterogeneity is comparable across clusters in that
-\begin{equation}\label{eq:degree}
-{\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
-\end{equation}
-Note that, the assumption~\eqref{eq:degree} does not preclude degree heterogeneity.
-Indeed, within each of the clusters, the highest degree can be $\theta(i) = \Omega(p)$, whereas the lowest degree can be $\theta(i)=\tO(1)$.
-%instead, we assume the relative extent of heterogeneity is comparable across clusters.
 
@@
 \Algphase{Sub-algorithm 1: Weighted higher-order initialization}
-\INPUT Observation $\tY \in \bbR^{p\times \cdots \times p}$, number of cluster $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
+\INPUT Observation $\tY \in \bbR^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
 \State Compute factor matrix $ \mU_{\text{pre}} = \text{SVD}_{r} (\text{Mat}(\tY))$ and the $(K-1)$-mode projection $
-\tX_{\text{pre}} = \tY \times_1  \mU_{\text{pre}}^T \times_2 \cdots \times_{K-1} \mU_{\text{pre}}^T.$
+\tX_{\text{pre}} = \tY \times_1   \mU_{\text{pre}} \mU_{\text{pre}}^T \times_2 \cdots \times_{K-1}  \mU_{\text{pre}} \mU_{\text{pre}}^T.$
 \State Compute factor matrix $\hat \mU = \text{SVD}_{r}(\text{Mat}(\tX_{\text{pre}}))$ and denoised tensor
@@
 \Algphase{Sub-algorithm 2: Angle-based iteration}
-%\caption{Angle-based iteration}\label{alg:refinement}
 \INPUT Observation $\tY \in \bbR^{p \times \cdots \times p}$, initialization $z^{(0)} \colon [p]\to[r]$ from Sub-algorithm 1, iteration number $T$.
@@
 
-%\begin{figure*}[h]
- %   \centering
- %   \includegraphics[width=1\textwidth]{algorithm.pdf}
- %   \label{alg:main}
-    %\caption{\color{red}\fixme{Jiaxin}{Should be $\argmax_{a \in [r] } \cos \of{ \mY_{i:}^d , \mS_{a:}^{(t)}}$.
-$\sin$ can not handle the obtuse angles; e.g.\ $\sin \pi = \sin 0= 0$, where $\pi$ is not the desired angle.} \fixme{Miaoyan}{Agree.
-I corrected it.}}
-%\end{figure*}
-
-\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general {\color{blue} sub-Gaussian} dTBM {\color{blue} with i.i.d.\ noise} under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
-%With probability at least $1 -C \exp(-cp)$, we have
-%\begin{equation}
-%\min_{\pi \in \Pi}\sum_{i\in[p]}\theta^2_i\ind\{ z^{(0)}(i)\neq \pi \circ z(j)\} \lesssim {Mr^K\sigma^2\over \Delta^2_{\min}}p^{-{(K-2)/2}},
-%\end{equation}
-%where $M>1$ is the relation constant in weighted angle $k$-means. 
+We now establish the misclustering error rate of initialization. We call $\mtheta$ is balanced, if the relative extent of heterogeneity is comparable across clusters in that
+\begin{equation}\label{eq:degree}
+{\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
+\end{equation}
+Note that, the assumption~\eqref{eq:degree} does not preclude degree heterogeneity.
+Indeed, within each of the clusters, the highest degree can be $\theta(i) = \Omega(p)$, whereas the lowest degree can be $\theta(i)=\tO(1)$.
+
+\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with i.i.d.\ noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
 Assume $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$.
@@
 
-\begin{rmk}[Comparison to previous results] For fixed SNR, our initialization error rate with $K=2$ agrees with the initialization error rate $\tO(p)$ in matrix models~\citep{gao2018community}.
-Furthermore, in the special case of non-degree TBMs, we achieve the same initial misclustering error $\tO(p^{-K/2})$ as in non-degree models~\citep{han2020exact}.
+\begin{rmk}[Comparison to previous results] For fixed SNR, our initialization error rate with $K=2$ agrees with the initialization error rate $\tO(p^{-1})$ in matrix models~\citep{gao2018community}.
+Furthermore, in the special case of non-degree TBMs with $\theta_1=\cdots=\theta_p=1$, we achieve the same initial misclustering error $\tO(p^{-K/2})$ as in non-degree models~\citep{han2020exact}.
 Theorem~\ref{thm:initial} implies the advantage of our algorithm in achieving both accuracy and model flexibility.
-%The result demonstrates the advantage of our algorithm in achieving both accuracy and model flexibility. 
 \end{rmk}
 
-%Assume the number of clusters $r$ is fixed.
-To ensure the misclassification error $\text{MisClust}(z^{(0)},z)$ decays to 0 as the number of nodes $p \rightarrow \infty$, we need the signal level $\Delta_{\min}^2$ satisfying
-%\begin{equation}
-%    \Delta_{\min}^2/\sigma^2  \gtrsim p^{-K/2}.
-%\end{equation}
-%This implies that desired polynomial rate can be achieved with lower signal level as sample size increases.
-Moreover, to achieve exact clustering, we should ensure $\text{MisClust}(z^{(0)},z) < \frac{1}{p}$ and thus it requires $\Delta_{\min}^2 \gtrsim p^{2 - K/2}$, which is a fairly strong signal condition especially when $K$ is small.
-However, the following theorem indicates a weaker signal condition is enough to achieve exact clustering with proper refinement.
-
-\begin{rmk}[Failure of conventional tensor HOSVD] If we use conventional HOSVD for tensor denoising; that is, we use $\mU_{\text{pre}}$ in place of $\hat \mU$ in line 2, then the misclustering rate becomes $\tO(p)$ for all $K\geq 2$.
+\begin{rmk}[Failure of conventional tensor HOSVD] If we use conventional HOSVD for tensor denoising; that is, we use $\mU_{\text{pre}}$ in place of $\hat \mU$ in line 2, then the misclustering rate becomes $\tO(p^{-1})$ for all $K\geq 2$.
 This rate is substantially worse than our current rate~\eqref{eq:ini}.
@@
 
-{
-\color{blue}
-
-\begin{rmk}[Eigen free clustering] Note that our initialization works on the estimated mean tensor $\hat \tX$ directly rather than the leading tensor decomposition factors of $\tX$.
-On one hand,  clustering on $\hat \tX$ avoids the eigen-gap assumption in previous work~\cite{ke2019community}.
-On the other hand,  vanilla spectral methods working on decomposition factors suffers the non-identifiability of the eigenspace with orthogonal transformation when the number of blocks $r \geq 3$.
-Such ambiguity comes from the possible multiplicity of eigenvalue and causes trouble for effective clustering~\citep{abbe2020entrywise}.
-In contrast, our eigen free strategy working on $\hat \tX$ avoids such non-identifiability issue regardless the number of blocks.
+\begin{rmk}[Singular-value gap-free clustering] Note that our clustering directly applies to the estimated mean tensor $\hat \tX$ rather than the leading tensor factors $\hat \mU$.
+Applying clustering to the tensor factors suffers from the non-identifiability issue due to the infinitely many orthogonal rotations when the number of blocks $r \geq 3$ in the absence of singular-value gaps.
+Such ambiguity causes the trouble for effective clustering~\citep{abbe2020entrywise}.
+In contrast, our initialization algorithm applies the clustering to the overall mean tensor $\hat \tX$.
+This strategy avoids the non-identifiability issue regardless of the number of blocks and singular-value gaps.
 \end{rmk}
 
-}
-
 \subsection{Angle-based iteration}
-%A polynomial-time local iteration refines the initial spectral clustering to achieve an exponential error rate in TBM~\citep{han2020exact}.  
+Our Theorem~\ref{thm:initial} has shown the polynomially decaying error rate from our initialization.
+Now we improve the error rate to exponential decay using local iterations.
 We propose an angle-based local iteration to improve the outputs from Sub-algorithm~\hyperref[alg:main]{1}.
-To gain the intuition, consider an one-dimensional degree-corrected clustering problem with data vectors $\mx_i = \theta(i) \ms_{z(i)} + \mepsilon_i, i \in [p]$, where $\ms_i$'s are known cluster centroids, $\theta(i)$'s are unknown positive degrees, and $z\colon [p] \mapsto [r]$ is the clustering assignment of interest.
+To gain the intuition, consider an one-dimensional degree-corrected clustering problem with data vectors $\mx_i = \theta(i) \ms_{z(i)} + \mepsilon_i, i \in [p]$, where $\ms_i$'s are known cluster centroids, $\theta(i)$'s are unknown positive degrees, and $z\colon [p] \mapsto [r]$ is the cluster assignment of interest.
 The angle-based $k$-means algorithm estimates the assignment $z$ by minimizing the angle between data vectors and centroids; i.e.,
@@
 
-Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as angle-based $k$-means. {\color{blue} We still take the order-3 tensor for illustration.} 
-{
-\color{blue}
-Specifically, Sub-algorithm~\hyperref[alg:main]{2} updates estimated core tensor and cluster assignment in each iteration. For core tensor, we consider the following update strategy  
+Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as in angle-based $k$-means.
+We still take the order-3 tensor for illustration.
+Specifically, Sub-algorithm~\hyperref[alg:main]{2} updates estimated core tensor and cluster assignment in each iteration.
+We use superscript $\cdot^{(t)}$ to denote the estimate from $t$-th iteration, where $t=1,\ldots.$ For core tensor, we consider the following update strategy
  \[
@@
 \]
-The row $\text{Mat}(\tY^{\text{d}})_{i:}$ and $\text{Mat}(\tS^{(t)})_{a:}$ corresponds to the $\mx_i$ and $\ms_a$s in the one-dimensional clustering~\eqref{eq:angle_kmeans}.
-Then, we obtain the updated assignment as
-\[z(i)^{(t+1)} = \argmin_{a \in [r]} \sin \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} }, \text{ for all } i \in [p], \]
-where $z^{(t+1)}(i)$ is randomly assigned for degenerate rows. 
-}
-%except that we use estimated centroids $\ms^{(t)}_a$ in place of $\ms_a$ based on estimated assignment in previous iterations. 
+We use $\mY^d$ and $ \mS^{(t)}$ to denote the $\text{Mat}(\tY^{\text{d}})$ and $\text{Mat}(\tS^{(t)})$.
+The rows $\mY^d_{i:}$ and $\mS^{(t)}_{a:}$ correspond to the $\mx_i$ and $\ms_a$ in the one-dimensional clustering~\eqref{eq:angle_kmeans}.
+Then, we obtain the updated assignment by
+\[
+z(i)^{(t+1)} = \argmax_{a \in [r]} \cos \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} },\ \text{ for all } i \in [p],
+\]
+provided that $\mS^{(t)}_{a:}$ is a non-zero vector.
+Otherwise, if $\mS^{(t)}_{a:}$ is a zero vector, then we make the convention to assign $z^{(t+1)}(i)$ randomly in $[p]$.
 Full procedures for our angle-based iteration are described in Sub-algorithm~\hyperref[alg:main]{2}.
 
-We then establish the misclustering error rate of iterations 
-{ \color{blue} under the stability assumption.
+We now establish the misclustering error rate of iterations under the stability assumption.
 
 \begin{defn}[Locally linear stability] \label{def:stable}
-Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\epsilon)=\{\bar z\colon \ell(\bar z, z)\leq \epsilon\}$. We define two cluster-size vectors for $\bar z$,
+Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\epsilon)=\{\bar z\colon \ell(\bar z, z)\leq \epsilon\}$.
+Let $\bar z\colon[p]\to [r]$ be a clustering function.
+We define two vectors associated with $\bar z$,
 \begin{align}
@@
 We call the degree is $\varepsilon$-locally linearly stable if and only if 
-\begin{equation}
+\begin{equation}\label{eq:local}
     \sin(\mp(\bar z),\ \mp_{\mtheta}(\bar z))\lesssim \varepsilon \Delta_{\min} ,\quad \text{for all } \bar z\in \tN(z, \varepsilon).
@@
 \end{defn}
-}
-
+
+Roughly speaking, the vector $\mp(\bar z)$ represents the raw cluster sizes, and $\mp_{\theta}(\bar z)$ represents the relative cluster sizes weighted by degrees. 
 The local stability holds trivially for $\varepsilon=0$ based on the construction of parameter space~\eqref{eq:family}.
-{\color{blue} The locally linear stability avoids the concentration of entities with extremely large or small heterogeneity, when a good estimated assignment with a small misclustering error $\epsilon$ is given.
-}
-
-%  In Algorithm~\ref{alg:refinement}, the estimated core tensor and cluster assignments are updated in each iteration. For core tensor, notice that the estimator $\tilde S$ with
-%   \[
-% \tilde \tS(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z(i_k)=a_k, k\in[3]\}
-% \]
-% is an unbiased estimator satisfying $\bbE[\tilde \tS_{a_1,...,a_K}] = \tS_{a_1,...,a_K}$ with true assignment $z$. This leads to a reasonable core tensor update strategy 
-%  \[
-% \tS^{(t)}(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k\in[3]\}.
-% \]
-% Intuitively, the estimate $\tS^{(t)}$ becomes better as $z^{(t)}$ getting more precise.
-
-% For cluster assignment, we reduce the higher-order problem to the one-dimensional clustering by matrization.
-Given $z^{(t)}$, we first aggregate the slices of $\tY$ and obtain a reduced tensor $\tY^{\text{d}}\in \bbR^{p \times r \times r}$, where
-% \[
-% \tY^{\text{d}}(i,a_2,a_3)=\text{Ave}\{\tY(i,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k \neq 1\}.
-% \]
-% Let $\mY^{\text{d}} = \text{Mat}(\tY^{\text{d}})$ and $\mS^{(t)} = \text{Mat}(\tS^{(t)})$.
-Then, the row $\mY^{\text{d}}_{i:}$s and $\mS^{(t)}_{a:}$s serve as the data vectors $\mx_i$ and centroids $\ms_a$s in the one-dimensional clustering~\eqref{eq:angle_kmeans}.
-We obtain the updated assignment as
-% \begin{equation}
-%     z(i)^{(t+1)} = \argmin_{a \in [r]} \sin \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} }, \text{ for all } i \in [p].
-% \end{equation}
-% In particular, $z^{(t+1)}(i)$ is randomly assigned for degenerate row with $\onorm{\mY^{\text{d}}_{i:}} = 0$.
-The refinement outputs the final cluster assignment after $T$ iterations denoted $z^{(T)}$.
-The full procedure for general order-$K$ tensor is described in Algorithm~\ref{alg:refinement}
+The condition~\eqref{eq:local} controls the impact of node degree to the $\mp_{\theta}(\cdot)$ with respect to the misclassification rate $\varepsilon$ and angle gap.
 
 \begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the setup as in Theorem~\ref{thm:initial}.
-Assume the local linear stability of degree holds in all neighborhoods $\tN(z,\epsilon)$ for any $\epsilon \leq \log^{-1}p$.
+Assume the local linear stability of degree holds in the neighborhood $\tN(z,\epsilon)$ for $\epsilon = \log^{-1}p$.
 Suppose
-$r=\tO(1)$ and $\text{SNR} \gtrsim  p^{-K/2}\log p$.
+$r=\tO(1)$ and $\text{SNR} \geq \tilde C p^{-K/2}\log p$ for some large constant $\tilde C$.
 Let $z^{(t)}$ denote the $t$-th iteration output in Sub-algorithm~\hyperref[alg:main]{2} with initialization $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{1}.
@@
 \end{thm}
-The iteration error is decomposed into two parts: statistical error and computational error.
+From the conclusion~\eqref{eq:final}, we find that the iteration error is decomposed into two parts: statistical error and computational error.
 The statistical error is unavoidable with noisy data regardless $t$, whereas the computational error decays in an exponential rate as the number of iterations $t \rightarrow \infty$.
@@
 \end{equation}
-for some permutation $\pi \in \Pi$. {
-\color{blue} We call our algorithm \textit{computationally efficient} with SNR $\gtrsim p^{-K/2} \log p$.
-Note that the minimal SNR requirement, $p^{-K/2 }\log p$, coincides with the computational lower bound in Theorem~\ref{thm:comp} ignoring the logarithmic term.
-Therefore, our algorithm is optimal regarding the signal demand and lies in the most right ``computationally efficient" regime in Figure~\ref{fig:phase_axis}.
-}
-
-{
-\color{blue}
-\subsection{Extension and practical issue}~\label{subsec:exten}
-{\bf Extension for Bernoulli observations.} The main difficulty to establish the statistical guarantee for Bernoulli observations lies in the initialization Sub-Algorithm~\hyperref[alg:main]{1}.
+for some permutation $\pi \in \Pi$.
+Therefore, our combined algorithm is \textit{computationally efficient} as long as SNR $\gtrsim p^{-K/2} \log p$.
+Note that, ignoring the logarithmic term, the minimal SNR requirement, $p^{-K/2}$, coincides with the computational lower bound in Theorem~\ref{thm:comp}.
+Therefore, our algorithm is optimal regarding the signal requirement and lies in the sharpest \emph{computationally efficient} regime in Figure~\ref{fig:phase_axis}.
+
+\subsection{Extensions and practical issues}~\label{subsec:exten}
+\hspace{-.3cm} {\bf Extension for Bernoulli observations.} The main difficulty to establish the statistical guarantee for Bernoulli observations lies in the initialization Sub-algorithm~\hyperref[alg:main]{1}.
 Theorem~\ref{thm:refinement} still holds for Bernoulli observations once the initialization accuracy satisfies the upper bound~\eqref{eq:ini} in Theorem~\ref{thm:initial}.
 
-Specifically, the derivation of Theorem~\ref{thm:initial} relies on the upper bound of the estimation error for the mean tensor in Lemma~\ref{lem:two-step_esterror}, i.e., with high probability
+We now provide a high-level explanation for the technical difficulty when applying Theorem~\ref{thm:initial} to Bernoulli observations.
+The derivation of Theorem~\ref{thm:initial} relies on the upper bound of the estimation error for the mean tensor in Lemma~\ref{lem:two-step_esterror}; i.e., with high probability
 \begin{equation}\label{eq:prop1}
@@
 \end{equation}
-where $\tX = \bbE[\tY]$ and $\hat \tX$ is defined in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}.
-Unfortunately, the inequality~\eqref{eq:prop1} only holds for i.i.d.\ sub-Gaussian observations while Bernoulli observations are not identically distributed.
-
-One possible remedy is to apply singular value decomposition to the unfolded Bernoulli observation $\tY$.
-Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{\floor{p^{K/2}} \times \ceil{p^{K/2}} }$ denote the square unfolded binary tensor.
-We have the estimate
-\begin{equation}
-    \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX)) -  \mat_{sq}(\tY)}_F^2.
-\end{equation}
-Following Lemma 7 in \cite{gao2018community}, with high probability, we have
+where $\tX = \bbE\tY$ and $\hat \tX$ is defined in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}.
+Unfortunately, the inequality~\eqref{eq:prop1} holds only for i.i.d.\ sub-Gaussian observations while Bernoulli observations are generally not identically distributed.
+
+One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square} of Bernoulli tensor $\tY$.
+Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{\floor{p^{K/2}} \times \ceil{p^{K/2}} }$ denote the nearly square unfolded Bernoulli tensor.
+Define a new estimate
+\begin{equation}\label{eq:matrixsvd}
+    \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX) -  \mat_{sq}(\tY)}_F^2.
+\end{equation}
+The optimization~\eqref{eq:matrixsvd} is simply a matrix SVD problem. Following Lemma 7 in \cite{gao2018community}, with high probability, the new estimate satisfies
 \begin{equation}
@@
 \end{equation} 
-Replacing the estimate $\hat \tX$ by $\hat \tX'$, the high probability upper bound for Bernoulli initialization is
+Replacing the estimate $\hat \tX$ by $\hat \tX'$ in Theorem~\ref{thm:initial}, the high probability upper bound for Bernoulli initialization is
 \begin{equation}\label{eq:ini_b}
@@
 The Bernoulli bound~\eqref{eq:ini_b} is relatively looser than Gaussian bound~\eqref{eq:ini}, especially when $K$ is small.
-A tighter Bernoulli bound will be achieved once a low-rank binary tensor estimation scheme with better accuracy is provided in the future.
-
-{\bf Extension for general dTBMs.}
-Our two-stage algorithm is able to be extended for the general (asymmetric) dTBMs. Specifically, in the Sub-Algorithm~\hyperref[alg:main]{1}, we make the following changes: (1)
-Replace the matrcization $\mat(\tY)$ by $\mat_k(\tY)$; (2) Repeat the Steps 1-5 with mode-specified number of cluster $r_k$; (3) Obtain the collection initialization $ \{z^{(0)}_k\}_{k=1}^K$.
-In the Sub-Algorithm~\hyperref[alg:main]{2}, we make the following changes: (1) Take the collection $\{z^{(0)}_k\}_{k=1}^K$ as input, and update the block tensor $\tS^{(t)}$ with the collection $\{ z_k^{(t)}\}_{k=1}^K$, $
+Nevertheless, our bound~\eqref{eq:ini_b} is already tighter than the pervious work~\citep{ke2019community}.
+The investigation of the gap between upper bound $p^{- \floor{K/2}}$ and the lower bound $p^{- K/2}$ for Bernoulli tensors will be left as future work.
+
+{\bf Extension for general dTBMs.} Our two-stage algorithm is able to be extended for the general (asymmetric) dTBMs.
+Specifically, in the Sub-algorithm~\hyperref[alg:main]{1}, we make the following changes: (1)
+Replace the matrcization $\mat(\tY)$ by $\mat_k(\tY)$; (2) Repeat the Steps 1-5 with mode-specified cluster number $r_k$; (3) Obtain the collection initialization $ \{z^{(0)}_k\}_{k=1}^K$.
+In the Sub-algorithm~\hyperref[alg:main]{2}, we make the following changes: (1) Take the collection $\{z^{(0)}_k\}_{k=1}^K$ as input, and update the block tensor $\tS^{(t)}$ with the collection $\{ z_k^{(t)}\}_{k=1}^K$, $
         \tS^{(t)}(i_1,\ldots, i_K) =  \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}_k(i_k) = j_k, k \in [K]\}$;
@@
 Correspondingly, Theorems~\ref{thm:initial} and \ref{thm:refinement} still hold with $\ell(z^{(0)}_k, z_k)$ and $\ell(z^{(t+1)}_k, z_k)$  for all $k \in [K]$.
-
-{\bf Computational Complexity.} Our two-stage algorithm has a polynomial computational cost.
-Specifically, the complexity of Sub-Algorithm~\hyperref[alg:main]{1} is $\tO(K p^{K+1} + Krp^K)$, where the first term is contributed by the first step SVD and the calculation of $\hat \tX$, and the second term comes from normalization and the $k$-means.
-The cost of each update in Sub-Algorithm~\hyperref[alg:main]{2} is $\tO(p^K + pr^K)$, where $p^K$ comes from the calculation of $\tS^{(t)}$ and $\tY^{\text{d}}$ and $pr^K$ comes from the normalization of $\tY^{\text{d}}, \tS^{(t)}$ and the cluster assignment update in Step 10.
-
-%that comes from the addition operation and comparison in Step 7,8, and 10.
-Note that the number of iteration is a logarithmic function of $p$ to achieve exact clustering.
-We conclude that the two-stage algorithm leads to a polynomial complexity in total.
-
-{\bf Parameter Selection.} Note that we assume the number of clusters $r$ is given for the algorithm. We impose the Bayesian information criterion (BIC)
-and choose the number of cluster that minimizes BIC, i.e., under dTBM~\eqref{eq:model_tensor},
-\begin{equation}
-    \hat r = \argmin_{r \in \mathbb{Z}_+ } \of{ -p^K \tL_{\tY}(\hat z(r), \hat \tS(r), \hat \mtheta(r)) + p_e(r) K \log p},
-\end{equation}
-where the triplet $(\hat z(r), \hat \tS(r), \hat \mtheta(r))$ are estimated parameters with given number of cluster $r$, $\tL_{\tY}$ denote the log-likelihood with observations $\tY$, and $p_e = r^K + p(\log r + 1) - 1$ is the effective number of parameters.
-Particularly, we obtain the estimate $\hat \mtheta(r)$ by first calculating the reduced tensor $\tY^{\text{d}}$ with $\hat z(r)$ and then normalizing the row norms $\onormSize{}{\mY_{i:}^{\text{d}}}$ to 1 in each cluster; i.e.,
-\begin{equation}
-    \hat \theta(i,r) = \frac{\onormSize{}{\hat \mY^{\text{d}}(r)_{i:}}}{\sum_{j: \hat z(j,r) =  \hat z(i,r)} \onormSize{}{\hat \mY^{\text{d}}(r)_{j:}}},
-\end{equation}
-where $\hat \mY^{\text{d}}(r) = \mat(\hat \tY^{\text{d}}(r))$ and 
-\begin{equation}
-    \hat \tY^{\text{d}}(r)(i,a_2,\ldots,a_K) = 
-\text{Ave}\{\tY(i,i_2,\ldots,i_K): \hat z(i_k, r) = a_k, k \neq 1 \}.
-\end{equation}
-
-With Gaussian observation $\tY$, we have
-\begin{equation}
-    \tL_{\tY} = - \log(\onormSize{}{\hat \tX - \tY}_F^2), \quad \hat \tX = \tS(r) \times_1 \hat \mTheta(r) \hat \mM(r) \times_2 \cdots \times_K \hat \mTheta(r) \hat \mM(r).
-\end{equation}
+The detailed model extension for general asymmetric dTBMS can be found in Appendix.
+
+{\bf Computational complexity.} Our two-stage algorithm has a computational cost polynomial in tensor dimension $p$.
+Specifically, the complexity of Sub-algorithm~\hyperref[alg:main]{1} is $\tO(K p^{K+1} + Krp^K)$, where the first term is contributed by the double projection and the calculation of $\hat \tX$, and the second term comes from normalization and the $k$-means.
+The cost of each update in Sub-algorithm~\hyperref[alg:main]{2} is $\tO(p^K + pr^K)$, where $p^K$ comes from the calculation of $\tS^{(t)}$ and $\tY^{\text{d}}$, and $pr^K$ comes from the normalization of $\tY^{\text{d}}$, the calculation of $\tS^{(t)}$, and the cluster assignment update in Step 10.
+
+{\bf Hyper-parameter selection.} In our theoretical analysis, we have assumed the true cluster number $r$ is given to our algorithm.
+In practice, the cluster number $r$ is often unknown, and we now propose a method to choose $r$ from data.
+We impose the Bayesian information criterion (BIC) and choose the cluster number that minimizes BIC; i.e., under the symmetric Gaussian dTBM~\eqref{eq:model_tensor},
+\begin{align}\label{eq:BIC}
+ \hat r &= \argmin_{r \in \mathbb{Z}_+ } \of{p^K \log(\onormSize{}{\hat \tX - \tY}_F^2)  + p_e(r) K \log p},\\
+ \text{where}  \quad \hat \tX &= \hat \tS(r) \times_1 \hat \mTheta(r) \hat \mM(r) \times_2 \cdots \times_K \hat \mTheta(r) \hat \mM(r),
+\end{align}
+where the triplet $(\hat z(r), \hat \mS(r),\hat \mtheta(r))$ are estimated parameters with cluster number $r$, and $p_e (r)= r^K + p(\log r + 1) - r$ is the effective number of parameters.
+Note that we have added the argument $(r)$ to related quantities as functions of $r$.
+In particular, the estimate $\hat \mtheta(r)$ in~\eqref{eq:BIC} is obtained by first calculating the reduced tensor $\hat \tY^{\text{d}}$ with $\hat z(r)$, and then normalizing the row norms $\onormSize{}{\hat \mY_{i:}^{\text{d}}}$ to 1 in each cluster; i.e.,
+\begin{equation}
+ \hat \mtheta(r)=(\hat \theta(1,r),\ldots,\hat \theta(p,r))^T,\quad\text{where}\quad   \hat \theta(i,r) = \frac{\onormSize{}{\hat \mY^{\text{d}}(r)_{i:}}}{\sum_{j: \hat z(j,r) =  \hat z(i,r)} \onormSize{}{\hat \mY^{\text{d}}(r)_{j:}}},
+\end{equation}
+where $\hat \mY^{\text{d}}(r) = \mat(\hat \tY^{\text{d}}(r))$, $\hat \tY^{\text{d}}(r)(i,a_2,\ldots,a_K) =  \text{Ave}\{\tY(i,i_2,\ldots,i_K): \hat z(i_k, r) = a_k, k \neq 1 \}$, and $\hat z(i,r)$ denotes the community label for the $i$-th node with given cluster number $r$.
 We evaluate the performance of the BIC criterion in Section~\ref{subsec:num_theory}.
-
-% We here provide a simple rank selection criteria when $r$ is unknown. With a fixed $r$, calculate the estimated minimal gap
-% \begin{equation}
-%     \hat \Delta_{\min}(r) = \min_{a \neq b \in [r]} \onorm{ \frac{\hat \mS_{a:}(r)}{\onormSize{}{\hat \mS_{a:}(r)}} - \frac{\hat \mS_{b:}(r)}{\onormSize{}{\hat \mS_{b:}(r)}} },
-% \end{equation}
-% where $\mS(r) = \mat(\hat \tS(r))$ and $\hat \tS(r)$ is the estimated signal tensor obtained by the same procedure in Step 7 of Sub-Algorithm~\hyperref[alg:main]{2} with the estimated assignment $\hat z(r)$.
-With a given upper bound for the number of clusters, denoted $R$, we choose $r$ that maximizes the estimated minimal gap, i.e.,
-% \begin{equation}
-%     \hat r = \argmax_{r \in [R]} \hat \Delta_{\min}(r).
-% \end{equation}
-
-}
-
-%We now discuss the conditions for which exact recovery occurs, i.e., $\ell(z, \hat z^{(t+1)})=0$ for some $t\geq T=\text{poly}(p)$. Based on Theorem~\ref{thm:refinement}, we need
-%\[
-%T\geq \log p +\log \ell(z, \hat z^{(0)})
-%\]
-%given initial assignment $z^{(0)}$ from Algorithm~\ref{alg:initial} and iteration number $T \geq 2 \log p$, with probability going to 1, the output of our Algorithm~\ref{alg:refinement} achieves exact recovery, i.e., for some permutation $\pi \in \Pi$,
-%\begin{equation}
- %    z^{(T)} = \pi \circ z, \quad \text{for all }T\geq 2\log p.
-%\end{equation}
-
-%\begin{algorithm}[h]
-
-%\caption{Angle-based iteration}\label{alg:refinement}
-%\begin{algorithmic}[1]
-
-%\INPUT Observation $\tY \in \bbR^{p \times \cdots \times p}$, initialization $z^{(0)} \colon [p]\to[r]$, iteration number $T$.
-%\For {$t = 0$ to $T-1$}
-%\State Update the block tensor $\tS^{(t)}$ via
-%\begin{align}
-%    &\tS^{(t)} (i_1,...,i_K)\\
- %   & \quad = \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}(i_k) = j_k, k \in [K]\}
-%\end{align}
-%\State Calculate reduced tensor $\tY^{\text{d}} \in \bbR^{p \times r \times \cdots \times r}$
-%\begin{align}
-%    &\tY^{\text{d}}(i,a_2,\ldots,a_%K) \\
-%    &= 
-%\text{Ave}\{\tY(i,i_2,\ldots,i_K): z^{(t)}(i_k) = a_k, k \neq 1 \}
-%\end{align}
-
-%\State Let $\mY^{\text{d}} = \text{Mat}(\tY^{\text{d}})$ and $J_0 = \{ i\in[p]: \onorm{\mY^{\text{d}}_{i:}} = 0\}$. Set $z^{(t+1)}(i)$ randomly in $[r]$ for $i \in J_0$.
-
-%\State Let $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. For all $i \in J_0^c$ update cluster assignment by
-%\begin{equation}
-%    z(i)^{(t+1)} = \argmin_{a \in [r]} \sin \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} }.
-%\end{equation}
-
-%\EndFor
-
-%\OUTPUT Estimated clustering $z^{(T)}  \in [r]^{p}$.
-%\end{algorithmic}
-%\end{algorithm}
-
-%We refer to the combination of Algorithms~\ref{alg:initial}+ \ref{alg:refinement} as provable dTBM algorithm.
-
-%\fixme{Miaoyan}{add discussion as in my JMLR binary tensor paper (stat-comp dominating terms)}
 
@@
 We set $\alpha$ close to 1 such that $1-\alpha=o(p)$.
-In particular, we have $\alpha = 1 + \Omega(p^{\gamma/2})$ %\fixme{Miaoyan}{My calculation shows $\text{SNR}=\Omega((1-\alpha)^2)\Rightarrow \alpha=1+\Omega(p^{\gamma/2})$.
-Double check.}
-  by Assumption~\ref{assmp:min_gap} and definition~\eqref{eq:gamma}.
+In particular, we have $\alpha = 1 + \Omega(p^{\gamma/2})$ with $\gamma<0$ by Assumption~\ref{assmp:min_gap} and definition~\eqref{eq:gamma}.
 Hence, we easily adjust SNR via varying $\alpha$.
- %We also set $\onorm{\text{Mat}(\tS)_{a:}} = c$ for some $c>0$ to exclude the effect of signal magnitude. 
 Note that the assortative setting is proposed for simulations, and our algorithm is applicable for general tensors in practice.
 The cluster assignment $z$ is randomly generated with equal probability across  $r$ clusters for each mode.
-Without further explanation, we generate degree heterogeneity $\mtheta$ from absolute normal distribution as $\theta(i) = |X_i| + 1 - 1/\sqrt{2\pi}$ with $|X_i| \stackrel{\text{i.i.d.}}\sim N(0,1), i \in [p]$ and normalize $\mtheta$ to satisfy \eqref{eq:family}.
+Without further explanation, we generate degree heterogeneity $\mtheta$ from absolute normal distribution by $\theta(i) = |X_i| + 1 - 1/\sqrt{2\pi}$ with $|X_i| \stackrel{\text{i.i.d.}}\sim N(0,1), i \in [p]$ and normalize $\mtheta$ to satisfy \eqref{eq:family}.
 Also, we set $\sigma^2 = 1$ for Gaussian data without further specification.
-
-%  To control SNR, we generate ``assortative" core tensors for all experiments, as a higher-order analogy of the setting in \citep{gao2018community}; i.e.,
-% \begin{align} \tS_{ijk} = 
-%     \begin{cases}
-%     s_1 & \text{if } i = j = k,\\
-%     s_2 & \text{if $i,j,k$ are not identical, }
-%     \end{cases}
-% \end{align}
-% where $s_1 > s_2 > 0$.
-Let $\alpha = s_1/s_2$ be the ratio between diagonal and off-diagonal elements.
-We have $\alpha = 1 + \Omega(p^{\gamma})$ by Assumption~\ref{assmp:min_gap} and definition~\eqref{eq:gamma} with fixed $\sigma^2$.
-This helps us to easily adjust SNR via varying $\alpha$.
-We also set $\onorm{\text{Mat}(\tS)_{a:}} = c$ for some $c>0$ to exclude the effect of signal magnitude.
-The ``assortative" setting is proposed for convenience.
-In practice, our algorithm is applicable for general core tensor.
-
-% The cluster assignment $z$ is randomly generated with given integer $r$.
-Without specific explanation, we generate degree heterogeneity $\mtheta$ from absolute normal distribution as $\theta(i) = |X_i| + 1 - 1/\sqrt{2\pi}, i \in [p]$ with $|X_i| \sim_{i.i.d.} N(0,1)$ and normalize $\mtheta$ to satisfy \eqref{eq:family}.
-We generate the noisy tensor under Gaussian model with $\sigma^2 = 1$ and Bernoulli model with mean tensor  $\tS \times_1 \mTheta \mM \times_2 \mTheta \mM \times_3 \mTheta \mM$.
-We denote provable dTBM algorithm as \textbf{Ours} for simplicity.
 
@@
 Note that finding MLE under dTBM is computationally intractable.
-We approximate MLE using an oracle estimator; i.e., the output of Sub-algorithm~\hyperref[alg:main]{2} initialized from true assignment.
+We approximate MLE using an oracle estimator, i.e., the output of Sub-algorithm~\hyperref[alg:main]{2} initialized from true assignment.
 Figure~\ref{fig:phase}a shows that both our algorithm and oracle estimator start to decrease around the critical value $\gamma_{\text{stat}}  = \gamma_{\text{comp}}  = -1$ in matrix case.
@@
 
-%\fixme{Miaoyan}{\color{red} add spacing between two figure panels}
 \begin{figure}[htb]
@@
 We consider both the Gaussian and Bernoulli models with $p = \{80, 100\}$, $r = 5$, $\gamma \in [-2.1, -1.4]$.
-%We set $c = 0.5$ and $c = 0.25$ for Gaussian and Bernoulli cases, respectively. 
 Figure~\ref{fig:ini_re} shows the substantial improvement of combined algorithm over initialization, especially under weak and intermediate signals.
@@
 
-{
-\color{blue}
-
 \begin{table}[hbt]
-\color{blue}
 \centering
-    \begin{tabularx}{\textwidth}{c *{12}{Y}}
+    \begin{tabularx}{\textwidth}{c *{8}{Y}}
     \toprule
-    Settings & \multicolumn{3}{c}{$p = 50, \sigma^2 = 0.25$} & \multicolumn{3}{c}{$p = 50, \sigma^2 = 1$} & \multicolumn{3}{c}{$p = 80, \sigma^2 = 0.25$} & \multicolumn{3}{c}{$p = 80, \sigma^2 = 1$}\\
-    \cmidrule(lr){2-4} \cmidrule(l){5-7} \cmidrule(l){8-10} \cmidrule(l){11-13}
-         True number of clusters $r$ & 2&3 & 4 & 2 & 3 & 4 & 2 & 3 & 4 & 2 & 3 & 4  \\
+    Settings & \multicolumn{2}{c}{$p = 50, \sigma^2 = 0.25$} & \multicolumn{2}{c}{$p = 50, \sigma^2 = 1$} & \multicolumn{2}{c}{$p = 80, \sigma^2 = 0.25$} & \multicolumn{2}{c}{$p = 80, \sigma^2 = 1$}\\
+    \cmidrule(lr){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9}
+         True cluster number $r$ & 2 & 4 & 2  & 4 & 2  & 4 & 2 & 4  \\
          \midrule
-         Estimated number of clusters $\hat r$ &  2 & 3 & 3.9& 2 & 3 & 3.1 & 2 & 3 & 4 & 2 & 3& 3.9   \\
+         Estimated cluster number $\hat r$ &  2(0)  & 3.9(0.25)& 2(0)    & 3.1(0.52) & 2(0)    & 4(0)   & 2(0)    & 3.9(0.31)   \\
      \bottomrule
     \end{tabularx}
-    \caption{Estimated number of clusters given by BIC criterion under the small noise $(\sigma^2 = 0.25)$ and large noise $(\sigma^2 = 0.5)$ cases.
-We report the averaged estimated number of clusters $\hat r$ over 30 duplication.}
+    \caption{Estimated cluster number given by BIC criterion under the low noise level $(\sigma^2 = 0.25)$ and high noise level $(\sigma^2 = 0.5)$ settings.
+Numbers in parentheses are standard deviations of $\hat r$ over 30 replications.}
     \label{tab:select}
@@
 
-The third experiment evaluates the empirical performance of the BIC criterion to select unknown number of cluster.
-We generate the data from an order-3 Gaussian model with $p = \{50,80\}$, $r = \{2,4\}$ and consider the noise $\sigma^2 \in \{ 0.25,1\}$ with $\alpha = 400$.
-Table~\ref{tab:select} implies that our BIC criterion exactly choose the true $r$ under all the settings with small $r = 2, 3$.
-Note that the BIC underestimates the large true number of clusters $(r = 4)$ with smaller SNR $(\sigma^2 = 1)$, and the accuracy immediately increases with larger dimension $p = 80$.
-The improvement follows by the fact that a larger dimension $p$ indicates a larger sample size in the block model.
-Therefore, we conclude that BIC criterion is a reasonable way to tune the number of clusters.
-}
-% {
-% \color{red} 
-% More experiments:
-% \begin{enumerate}
-%     \item Initialization impact to Angle-based iteration.(Figure 4 in Han)
-%     \item Imbalanced cluster size.(Figure 9 in Han)
-%     \item Non-degree comparison. 
-% \end{enumerate}
-% }
+The third experiment evaluates the empirical performance of the BIC criterion to select unknown cluster number.
+We generate the data from an order-3 Gaussian model with $p = \{50,80\}$, $r = \{2,4\}$, and noise level $\sigma^2 \in \{ 0.25,1\}$.
+Table~\ref{tab:select} shows that our BIC criterion well chooses the true $r$ under most settings.
+Note that the BIC slightly underestimates the true cluster number $(r = 4)$ with smaller dimension and higher noise $(p = 50, \sigma=1)$, and the accuracy immediately increases with larger dimension $p = 80$.
+The improvement follows from the fact that a larger dimension $p$ indicates a larger sample size in the tensor block model.
+Therefore, we conclude that BIC criterion is a reasonable way to tune the cluster number.
 
@@
 \begin{itemize}[wide,topsep=-3pt,itemsep=0pt,parsep=1pt]
-    \item \textbf{\small HOSVD }: HOSVD on data tensor and $k$-means on the rows of the factor matrix;
+    \item \textbf{\small HOSVD}: HOSVD on data tensor and $k$-means on the rows of the factor matrix;
     \item \textbf{\small HOSVD+}: HOSVD on data tensor and $k$-means on the $\ell_2$-normalized rows of the factor matrix;
-    \item \textbf{\small HLloyd}: High-order Lloyd algorithm and high-order spectral clustering~\citep{han2020exact};
-    \item \textbf{\small SCORE}: Tensor-SCORE for clustering~\citep{ke2019community};
-    %\item \textbf{\small SCORE-}: Incomplete tensor-SCORE without regularized higher-order orthogonal iteration.
+    \item \textbf{\small HLloyd}~\citep{han2020exact}: High-order clustering algorithm developed for non-degree tensor block models;
+    \item \textbf{\small SCORE}~\citep{ke2019community}: Tensor-SCORE for clustering developed for binary tensors.
 \end{itemize}
@@
 Among the four alternative algorithms, the \textbf{\small SCORE} is the closest method to ours.
-%, since \textbf{\small SCORE} also The incomplete version \textbf{\small SCORE-} skips the key step, regularized HOOI, in \textbf{\small SCORE}.
-We involve \textbf{\small SCORE-} to study the critical components of \textbf{\small SCORE}.
 We set the tuning parameters of \textbf{\small SCORE} as in previous literature \citep{ke2019community}.
 The methods \textbf{\small SCORE} and \textbf{\small HOSVD+} are designed for degree models, whereas \textbf{\small HOSVD} and \textbf{\small HLloyd} are designed for non-degree models.
-%which shares a similar idea of our spectral initialization but implements naive denoising and clustering steps.  We refer our algorithm as \textbf{\small dTBM}.
-%, and the later one is expected to outperform than spectral methods under TBM. 
 We conduct two experiments to assess the impacts of (i) signal strength and (ii) degree heterogeneity, based on Gaussian and Bernoulli models with $ p = 100, r = 5$.
 We refer to our algorithm as \textbf{\small dTBM} in the comparison.
-%Consider $ p = 100, r = 5, c = 0.5$ for both experiments.
 
@@
 We perform extra simulations to verify the impact of degree effects.
-We use the same setting as in the first experiment in the Section~\ref{subsec:comp}, except that we now generate the degree heterogeneity $\mtheta$ from Pareto distribution with shape parameter $a$ prior to normalization.
+We use the same setting as in the first experiment in the Section~\ref{subsec:comp}, except that we now generate the degree heterogeneity $\mtheta$ from Pareto distribution prior to normalization.
+The density function of Pareto distribution is $f(x|a,b) = a b^a x^{-(a+1)} \ind\{ x \geq b \}$, where $a$ is called \emph{shape} parameter.
+We vary $a \in \{2,6\}$ and choose $b$ such that $\bbE X = a(a - 1)^{-1}b = 1$ for $X$ following Pareto$(a,b)$.
+Note that a smaller $a$ leads to a larger variance in $\mtheta$ and hence a larger degree heterogeneity.
 We consider the Gaussian model under low $(a = 6)$ and high $(a = 2)$ degree heterogeneity.
-Figure~\ref{fig:comp_gamma_theta} shows that the errors for non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) increases with degree heterogeneity.
+Figure~\ref{fig:comp_gamma_theta} shows that the errors for non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) increase with degree heterogeneity.
 In addition, the advantage of \textbf{\small HLloyd} over \textbf{\small HOSVD+} disappears with higher degree heterogeneity.
-
-%Figure~\ref{fig:comp_gamma} implies all methods perform better with larger signal.
-Our \textbf{\small dTBM} and \textbf{\small SCORE} show the best accuracy as expected.
-The sub-optimality of \textbf{\small SCORE-} and \textbf{\small HOSVD ++} implies the necessity of local refinement for spectral clustering.
-A surprising observation is that \textbf{\small HLloyd} beats  \textbf{\small SCORE-} and \textbf{\small HOSVD ++} under Gaussian case, which indicates the robustness of \textbf{\small HLloyd} for mild heterogeneity.
 
@@
 In this experiment, we generate $\mtheta$ from Pareto distribution prior to normalization.
-The density function of Pareto distribution is $f(x|a,b) = a b^a x^{-(a+1)} \ind\{ x \geq b \}$, where $a$ is called \emph{shape} parameter.
-We vary $a \in [3,6]$ and choose $b$ such that $\bbE[X] = a(a - 1)^{-1}b = 1$ for $X$ following Pareto$(a,b)$.
-Note that a smaller $a$ leads to a larger variance in $\mtheta$ and hence a larger degree heterogeneity.
+We vary the shape parameter $a \in [3,6]$ in the Pareto distribution to investigate a range of degree heterogeneities.
 Figure~\ref{fig:comp_theta} demonstrates the stability of degree-corrected algorithms (\textbf{\small dTBM}, \textbf{\small SCORE}, \textbf{\small HOSVD+}) over the entire range of degree heterogeneity under consideration.
@@
 
-%We fix the signal with $\alpha = 2$ and set $c = 1$ and $c = 0.25$ for Gaussian and Bernoulli cases, respectively.
-Figure~\ref{fig:comp_theta} demonstrates the outstanding performance of \textbf{\small dTBM} and \textbf{\small SCORE} under all range of degree heterogeneity.
-Specifically, \textbf{\small SCORE} slightly beats \textbf{\small dTBM} in Gaussian case.
-A possible reason for the outperformance is that \textbf{\small SCORE} purposes a relaxer constrain on $\mtheta$.
-The terrible performances of \textbf{\small HLloyd} and \textbf{\small HOSVD+} confirm the failure of non-degree algorithms  facing strong heterogeneity.\section{Real data analysis}
-
-% \begin{table}[hbt]
-%     \centering
-%     \begin{tabular}{c |c c c}
-%     \hline
-%         Method & \textbf{\small Ours} &   \textbf{\small HOSVD} & \textbf{\small HOSVD+}\\
-%          CER & \textbf{0.116} & 0.22 &0.213 \\
-%          \hline
-%          Method & \textbf{\small HLloyd} &\textbf{\small SCORE} & \textbf{\small SCORE-} \\
-%          CER &0.149 &  0.199 & 0.228\\
-%          \hline
-%     \end{tabular}
-%     \caption{CER of real data for different methods.}
-%     \label{tab:peru}
-% \end{table}
-
+\section{Real data applications}\label{sec:real}
 \subsection{Human brain connectome data analysis}
@@
 We preprocess the original dataset following \cite{desikan2006automated} and partition the brain into 68 regions.
-The cleaned data includes brain networks for 136 individuals.
-Each brain network is represented by a 68-by-68 binary symmetric matrix, where the entries with value 1 refers to the presence of connection among 68 nodes while value 0 refers to the absence.
-We use $\tY \in \{0,1\}^{68 \times 68 \times 136}$ to denote the binary observation.
+The cleaned dataset includes brain networks for 136 individuals.
+Each brain network is represented by a 68-by-68 binary symmetric matrix, where the entry with value 1 indicates the presence of connection between node pairs, while the value 0 indicates the absence.
+We use $\tY \in \{0,1\}^{68 \times 68 \times 136}$ to denote the binary tensor.
 Individual attributes such as gender and sex are recorded.
 
-We apply our generalized Algorithm to the HCP data with the number of clusters on three modes $r_1 = r_2 = 4$ and $r_3 = 3$.
-The selection of $r_1$ and $r_2$ follows the human brain anatomy and the symmetry in the brain network, and the $r_3$ is chosen to be small following previous analysis~\citep{hu2021generalized}.
-The estimated brain node clustering results on the first and second mode are the same.
-Figure~\ref{fig:cluster_brain} indicates that brain connection exhibits a strong spatial separation structure.
+We apply our generalized algorithm to the HCP data with the numbers of clusters on three modes $r_1 = r_2 = 4$ and $r_3 = 3$.
+The selection of $r_1$ and $r_2$ follows the human brain anatomy and the symmetry in the brain network, and the $r_3$ is specified following previous analysis~\citep{hu2021generalized}.
+Because of the symmetry in the data, the estimated brain node clustering results are the same on the first and second modes.
+Figure~\ref{fig:cluster_brain} shows that brain connection exhibits a strong spatial separation structure.
 Specifically, the first cluster, named \emph{L.Hemis}, involves all the nodes in the left hemisphere.
 The nodes in the right hemisphere are further separated into three clusters led by the middle-part tissues in Temporal and Parietal lobes (\emph{R.Temporal}), the back-part tissues in Occipital lobe (\emph{R.Occipital}), and the front-part tissues in Frontal and Parietal lobes (\emph{R.Supra}).
-This clustering result is consistent with the common sense that the left and right hemispheres play different roles in human brain.
-
-% \vspace{0.2cm}
-% \begin{table}[h]
-%  \centering
-% %\resizebox{\textwidth}{!}{
-% \begin{tabular}{c|c}
-% \hline
-%          Cluster & \multicolumn{1}{c}{Brain nodes} \\
-%          \hline
-%         \multirow{1}{*}{L.Hemis} & All brain nodes in left hemisphere\\
-%         \hline
-%         \multirow{2}{*}{R.Temporal} & SupF Insula SupF SupT SupT MT MT MT\\
-%         &IT IT IT SupP IP SupF IsthmusC Precuneus SupT \\
-%         \hline
-%         \multirow{1}{*}{R.Occiptial} &  LO  LO Fpole Tpole  MOF Cuneus paraHippo Lingual\\
-%         \hline
-%         \multirow{2}{*}{R.Supra} & preCentral postCentral SupraM SupraM SupraM SupraM \\
-%         &CaudMF parsTriangularis parsOpercularis \\
-%         \hline
-%     \end{tabular}
-%     \caption{Brain node clustering results for HCP data.}
-%     \label{tab:cluster}
-% \end{table}
+This clustering result is reasonable since the left and right hemispheres often play different roles in human brains.
 
@@
     \caption{Mode 3 slices of estimated core tensor $\hat \tS$.
-(a) Average estimated slice weighted by the group size; (b)-(d) Group-specified enrichment, i.e., the subtraction between each slice of $\hat \tS$ and the averaged slice.
+(a) Average estimated slice weighted by the group size; (b)-(d) Group-specified enrichment, i.e., the difference between each slice of $\hat \tS$ and the averaged slice.
 }
@@
     \includegraphics[width = 1\textwidth]{brain_connection.pdf}
-    \caption{Observed brain connections in the population and each group of individuals.(a) Average brain network; (b)-(d) Group-specified brain networks enrichment in Groups 1-3.
-Red edges refer to the positive enrichment and blue edges refer the negative reduction.}
+    \caption{Observed brain connections in the population and each group of individuals.
+(a) Average brain network; (b)-(d) Group-specified brain network enrichments in Groups 1-3.
+Red edges represent the positive enrichment and blue edges represent the negative enrichment.}
     \label{fig:brain_conn}
@@
 
-Figure~\ref{fig:ests} illustrates the estimated core tensor $\hat \tS$ with estimated clustering, and Figure~\ref{fig:brain_conn} visualizes the average observed brain connections and the connection enrichment with average observed networks in each group.
+Figure~\ref{fig:ests} illustrates the estimated core tensor $\hat \tS$ with estimated clustering, and Figure~\ref{fig:brain_conn} visualizes the average brain connections and the connection enrichment in contrast to average networks in each group.
 In general, we find that the inner-hemisphere connection has stronger connection compared to inter-hemisphere connections (Figure~\ref{fig:ests}a).
 Also, the back and front parts (\emph{R.Occipital}, \emph{R.Supra}) are shown to have more interactions with temporal tissues than inner-cluster connections.
-In addition, the group 1 with 54\% females implies an enrichment on the inter-hemisphere connections (Figure~\ref{fig:ests}b) while group 4 with only 36\% females exhibits a reduction (Figure~\ref{fig:ests}d).
+In addition, the group 1 with 54\% females shows an enrichment on the inter-hemisphere connections (Figure~\ref{fig:ests}b), while group 4 with only 36\% females exhibits a reduction (Figure~\ref{fig:ests}d).
 This result agrees with previous findings in \cite{hu2021generalized}.
-The enrichment on the back-front connection is also recognized in group 3 (Figure~\ref{fig:ests}c).
-The interpretive pattern in results demonstrate the usefulness of our clustering methods in the human brain connectome data application.
+The enrichment on the back-front connection is also recognized in group~3 (Figure~\ref{fig:ests}c).
+The interpretive patterns in our results demonstrate the usefulness of our clustering methods in the human brain connectome data application.
 
@@
 We reconstruct legislation network as an order-3 binary tensor $\tY \in \{0,1\}^{116 \times 116 \times 116}$, where $\tY_{ijk} = 1$ if the legislators $(i,j,k)$ have sponsored the same bill, and $\tY_{ijk} = 0$ otherwise.
-The true party affiliations of legislators are provided and serve as ground truth.
+The true party affiliations of legislators are provided and serve as the ground truth.
 We apply various higher-order clustering methods to $\tY$ with $r = 5$.
@@
 
-\section{Conclusion}\label{sec:con}
-We have developed a general degree-corrected tensor block model with a two-step angle-based polynomial-times algorithm.
-We have, for the first time, characterized the statistical and computational behaviors of the degree-corrected tensor block model under different signal-to-noise ratio regimes.
-Simulations and Peru Legislation and Human brain connection data analysis confirm the potential of our method for practical applications.
-
-% {\color{red}
-% Future direction. 
-% \begin{enumerate}
-%     \item Missing data.
-%     \item Sub-exponential data.
-%     \item Imbalance in community size and heterogeneity.
-% \end{enumerate}
-% }
-
-{
-\color{blue}
-
-\section{Proof Sketches}
-
-In this section, we provide the proof sketches for Theorem~\ref{thm:initial} and Theorem~\ref{thm:refinement}. Detail proofs are provided in Appendix.
+\section{Proof Sketches}\label{sec:mainproof}
+
+In this section, we provide the proof sketches for the main Theorems~\ref{thm:initial}-\ref{thm:refinement}. Detail proofs and extra theoretical results are provided in Appendix.
 
 \subsection{Proof sketch of Theorem~\ref{thm:initial}}
-The proof of Theorem~\ref{thm:initial} is mainly inspired by the proof idea of \citet[Lemma 1]{gao2018community}, and extra difficulties due to the angle gap assumption and tensor property are addressed in our proof.
-Specifically, we control the misclustering error by the estimation error of $\hat \tX$ calculated in Step 2.
+The proof of Theorem~\ref{thm:initial} is inspired by the proof idea of \citet[Lemma 1]{gao2018community}.
+The extra difficulties are the angle gap characterization and multilinear algebra property in tensors; we address both challenges in our proof.
+Specifically, we control the misclustering error by the estimation error of $\hat \tX$ calculated in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}.
 We prove the following inequality
@@
 \end{equation}
-where $\tX = \bbE[\tY]$ is the true mean.
-The first inequality in~\eqref{eq:proof_4} holds with the assumption $\min_{i \in [p]} \theta(i) \geq c$ in Theorem~\ref{thm:initial}.
-The second inequality relies on the intermediate conclusion that the angle gap of mean tensor $\tX$ is lower bounded by a function of core tensor minimal angle gap $\Delta_{\min}$ defined in Assumption~\ref{assmp:min_gap}.
-Let $\ma^{s} \coloneqq \ma/\onorm{\ma}$ denote the normalized vector and make the convention that $\ma^s = 0$ if $\ma = 0$.
-We show that
+where $\tX = \bbE \tY$ is the true mean.
+The first inequality in~\eqref{eq:proof_4} holds with the assumption $\min_{i \in [p]} \theta(i) \geq c>0$ in Theorem~\ref{thm:initial}.
+The second inequality relies on an important conclusion that the angle gap of mean tensor $\tX$ is lower bounded by that of core tensor $\tS$, i.e., the minimal angle gap $\Delta_{\min}$ defined in Assumption~\ref{assmp:min_gap}.
+Let $\ma^{s} \coloneqq \ma/\onorm{\ma}$ denote the normalized vector with the convention that $\ma^s = 0$ if $\ma = 0$.
+We want to show that
 \begin{equation}\label{eq:proof_4_gap}
@@
 The most challenging part in the proof of Theorem~\ref{thm:initial} lies in the derivation of inequality~\eqref{eq:proof_4_gap}, in which the proof of \cite{gao2018community} is no longer applicable due to different angle gap assumption in our dTBM.
-Extra padding technique in Lemma~\ref{lem:pad} and balance assumption~\eqref{eq:degree} are equipped to derive~\eqref{eq:proof_4_gap}.
-Last, we finish the proof of Theorem~\ref{thm:initial} by showing the third inequality with \citet[Proposition 1]{han2020exact}.
+We develop the extra padding technique in Lemma~\ref{lem:pad} and balance assumption~\eqref{eq:degree} to derive~\eqref{eq:proof_4_gap}.
+Last, we finish the proof of Theorem~\ref{thm:initial} by showing the third inequality of~\eqref{eq:proof_4} using \citet[Proposition 1]{han2020exact}.
 
 \subsection{Proof sketch of Theorem 5}
-The proof of Theorem~\ref{thm:refinement} is mainly inspired by the proof idea of  \citet[Theorem 2]{han2020exact}, and extra polar coordinate based techniques are imposed due to the nuisance degree heterogeneity and angle gap assumption.
-We conduct the contraction property of the misclustering loss
+The proof of Theorem~\ref{thm:refinement} is inspired by the proof idea of \citet[Theorem 2]{han2020exact}.
+We develop extra polar-coordinate based techniques with angle gap characterization to address the nuisance degree heterogeneity.
+We introduce an intermediate quantity called misclustering loss
 \begin{equation}
@@
 \end{equation}
-and introduce the oracle estimators for core tensor and assignment
+where the superscript $\cdot^{s}$ denotes the normalized vector; i.e., $\ma^s:=\ma/\onorm{\ma}$ if $\ma\neq 0$ and $\ma^{s}=0$ if $\ma=0$ for any vector $\ma$.
+We show that $L^{(t)}$ provides an upper bound for the misclassification error of interest via the inequality $\ell^{(t)}\leq {L^{(t)}\over \Delta^2_{\min}}$.
+Therefore, it suffices to control $L^{(t)}$.
+Further, we introduce the oracle estimators for core tensor under the true cluster assignment via
 \begin{equation}
     \tilde \tS = \tY \times_1 \mW^T \times_2 \cdots \times_K \mW^T, 
-    %\quad \tilde z(i) = \argmin_{a \in [r]} \onormSize{}{ { [\mat(\tY)_{i:} \mV]^s - [\mat(\tilde \tS)_{a:}]^s } }^2
-\end{equation}
-where $\mW = \mM \of{ \text{diag}(\ind_{p}^T \mM) }^{-1}$ is the weighted true membership matrix.
-Let $ \mV = \mW^{\otimes (K-1)}$ denote the Kronecker product of $K-1$ $\mW$ matrices, and similarly define $\mW^{(t)}, \mV^{(t)}$ with $\mM^{(t)}$ corresponding to $z^{(t)}$.
-To evaluate $L^{(t+1)}$, we consider the event
+\end{equation}
+where $\mW = \mM \of{ \text{diag}(\mone_{p}^T \mM) }^{-1}$ is the weighted true membership matrix.
+Let $ \mV = \mW^{\otimes (K-1)}$ denote the Kronecker product of $(K-1)$ copies of $\mW$ matrices, and we define the $t$-th iteration quantities $\mW^{(t)}, \mV^{(t)}$ corresponding to $\mM^{(t)}$ (or equivalently $z^{(t)}$).
+To evaluate $L^{(t+1)}$, we prove the bound
 \begin{equation}\label{eq:proof_5_event}
-    \ind \offf{ z^{(t+1)}(i) = b } = \ind \offf{       \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{b:}^{(t)}]^s }^2 \leq \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{z(i):}^{(t)}]^s }^2} \leq A_{ib} + B_{ib}
+    \ind \offf{ z^{(t+1)}(i) = b } = \ind \offf{       \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{b:}^{(t)}]^s }^2 \leq \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{z(i):}^{(t)}]^s }^2} \leq A_{ib} + B_{ib},
 \end{equation}
@@
 \end{align}
-The terms $F_{ib}^{(t)}, G_{ib}^{(t)}, H_{ib}^{(t)}$ are controlled by $z^{(t)}, \tS^{(t)}$, and detail definitions are in \eqref{eq:f}, \eqref{eq:g}, \eqref{eq:h}.
-Note that the event $A_{ib}$ only involves the oracle estimator while all the terms related to the $t$-th iteration estimator are in $B_{ib}$.
-Thus, the inequality~\eqref{eq:proof_5_event} decomposes the misclustering loss in the $(t+1)$-th iteration by the oracle loss and the loss in $t$-th iteration.
+The terms $F_{ib}^{(t)}, G_{ib}^{(t)}, H_{ib}^{(t)}$ are controlled by $z^{(t)}, \tS^{(t)}$; see the detailed definitions in \eqref{eq:f}, \eqref{eq:g}, \eqref{eq:h}.
+Note that the event $A_{ib}$ only involves the oracle estimator independent of $t$, while all the terms related to the $t$-th iteration are in $B_{ib}$.
+Thus, the inequality~\eqref{eq:proof_5_event} decomposes the misclustering loss in the $(t+1)$-th iteration into the oracle loss and the loss in $t$-th iteration.
 This decomposition leads to the separation of statistical error and computational error in the final upper bound of Theorem~\ref{thm:refinement}.
@@
 \begin{equation}\label{eq:proof_5_ineq}
-    L^{(t+1)} \lesssim \xi + \rho L^{(t)}, \quad \xi = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  A_{ib} \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
-\end{equation}
-where $\rho \in (0,1)$ is the contraction parameter, and we call $\xi$ the oracle loss.
-Controlling the probability of event $B_{ib}$ and obtaining the $\rho L^{(t)}$ term in the right hand side of~\eqref{eq:proof_5_ineq} is the most tricky and challenging part in the proof of Theorem~\ref{thm:refinement}.
-Note that the true and estimated core tensors are involved as normalized rows such as $\mS_{a:}^s, \tilde \mS_{a:}^s, [\mS^{(t)}_{a:}]^s$.
+    L^{(t+1)} \leq M \xi + \rho L^{(t)}, \quad \xi = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  A_{ib} \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
+\end{equation}
+where $M$ is a positive constant, $\rho \in (0,1)$ is the contraction parameter, and we call $\xi$ the oracle loss.
+Controlling the probability of event $B_{ib}$ and obtaining the $\rho L^{(t)}$ term in the right hand side of~\eqref{eq:proof_5_ineq} are the most challenging parts in the proof of Theorem~\ref{thm:refinement}.
+Note that the true and estimated core tensors are involved via their normalized rows such as $\mS_{a:}^s, \tilde \mS_{a:}^s, [\mS^{(t)}_{a:}]^s$.
 The Cartesian coordinate based analysis in \cite{han2020exact} is no longer applicable in our case.
-Instead, we use the polar coordinate based analysis and the geometry property of trigonometric functions to derive the high probability upper bounds for $F_{ib}^{(t)}, G_{ib}^{(t)}, H_{ib}^{(t)}$.
+Instead, we use the polar-coordinate based analysis and the geometry property of trigonometric functions to derive the high probability upper bounds for $F_{ib}^{(t)}, G_{ib}^{(t)}, H_{ib}^{(t)}$.
 
@@
 
-}
+\section*{Acknowledgments}
+This research is supported in part by NSF grants DMS-1915978, DMS-2023239, EF-2133740, and funding from the Wisconsin Alumni Research foundation.
+We thank Zheng Tracy Ke, Rungang Han, Yuetian Luo for helpful discussions and for sharing software packages.
 
@@
 
-%\section*{Appendices}
-
 \newpage
@@
 
-We provide the proofs for all the theorems in our main paper. In each sub-section, we show the proof of main theorem and attach the useful lemmas in the end.
+We provide the proofs for all the theorems in our main paper. In each sub-section, we first show the proof of main theorem and then collect the useful lemmas in the end.
 
 \section*{Notation}
-Before the proofs, we first introduce the notations used throughout the following sections and the generalized dTBM without symmetric assumptions for the proofs of Theorem~\ref{thm:unique} and Theorem~\ref{thm:stats}.
-The parameter space and minimal gap assumption for are also extended the generalized dTBM.
-The conclusions of Theorem~\ref{thm:unique} and \ref{thm:stats} in the main paper are obtained by simply setting $z_k = z, \mS_k = \mS, \mtheta_k = \mtheta, k \in [K]$ for the generalized dTBM.
-
-{\bf Notations.}
-\begin{enumerate}
-    \item For all $ k \in [K]$, denote the tensor matricizations as
+Before the proofs, we first introduce the notation used throughout the appendix and the generalized dTBM without symmetric assumptions.
+The parameter space and minimal gap assumption are also extended for the generalized dTBM.
+
+{\bf Preliminaries.}
+\begin{enumerate}[wide]
+    \item For mode $ k \in [K]$, denote the mode-$k$ tensor matricizations by
     \begin{equation}
@@
     \end{equation}
-    %We ignore the subscript $k$ when considering symmetric tensors. 
-    \item For a vector $\ma$, let $\ma^{s} \coloneqq \ma/\onorm{\ma}$ denote the normalized vector. We make the convention that $\ma^s = 0$ if $\ma = 0$. 
-    % Then we rewrite the angle gap~\eqref{eq:general_minial_gap} as
-    % \begin{equation}
-    %     \Delta_{\min} = \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onorm{ \mS_{k,a:}^s - \mS_{k, b:}^s }.
-    % \end{equation}
-    \item For a matrix $\mA \in \bbR^{n \times m} \in \bbR^{n^K \times m^K}$, let $\mA^{\otimes K}$ denotes the kronecker product of $K$ matrices $\mA \otimes \cdots \otimes \mA$,
+    \item For a vector $\ma$, let $\ma^{s} \coloneqq \ma/\onorm{\ma}$ denote the normalized vector. We make the convention that $\ma^s = {\bf 0}$ if $\ma = {\bf 0}$. 
+    \item For a matrix $\mA \in \bbR^{n \times m} $, let $\mA^{\otimes K}:=\mA\otimes \cdots \otimes \mA\in \bbR^{n^K \times m^K}$ denote the Kronecker product of $K$ copies of matrices $\mA $.
     \item For a matrix $\mA$, let $\onormSize{}{\mA}_\sigma$ denote the spectral norm of matrix $\mA$, which is equal to the maximal singular value of $\mA$; let $\lambda_k(\mA)$ denote the $k$-th largest singular value of $\mA$; let $\onormSize{}{\mA}_F$ denote the Frobenius norm of matrix $\mA$.
-    \item For two terms $a$ and $b$, let $a \asymp b$ if there exist two positive constants $c, C$ such that $cb \leq a\leq Cb$. 
+    \item For two sequence $a$ and $b$, let $a \asymp b$ if there exist two positive constants $c, C$ such that $cb \leq a\leq Cb$. 
 \end{enumerate}
 
-{\bf Generalized dTBM.}
+{\bf Model extension to generalized dTBM.}
 
@@
 \end{equation}
-where $\tY \in \bbR^{p_1 \times \cdots \times p_K}$ is the data tensor, $\tX\in \mathbb{R}^{p_1\times \cdots \times p_K}$ is the mean tensor, $\tS \in \bbR^{r_1 \times \cdots \times r_K}$ is the core tensor, $\tE \in \bbR^{p_1 \times \cdots \times p_K}$ is the noise tensor consisting of independent mean-zero sub-Gaussian entries with variance bounded by $\sigma^2$, $\mTheta_k = \text{diag}(\mtheta_k)$, and $\mM_k\in \{0,1\}^{p_k \times r_k}$ is the membership matrix corresponding to the assignment $z_k: [p_k] \mapsto [r_k]$, for all $k \in [K]$.
+where $\tY \in \bbR^{p_1 \times \cdots \times p_K}$ is the data tensor, $\tX\in \mathbb{R}^{p_1\times \cdots \times p_K}$ is the mean tensor, $\tS \in \bbR^{r_1 \times \cdots \times r_K}$ is the core tensor, $\tE \in \bbR^{p_1 \times \cdots \times p_K}$ is the noise tensor consisting of independent zero-mean sub-Gaussian entries with variance bounded by $\sigma^2$, $\mTheta_k = \text{diag}(\mtheta_k)$, and $\mM_k\in \{0,1\}^{p_k \times r_k}$ is the membership matrix corresponding to the assignment $z_k: [p_k] \mapsto [r_k]$, for all $k \in [K]$.
 
@@
 
-% \begin{equation}\label{eq:general_family}
-%     \tP (r)= \offf{ \of{\{z_k\}, \tS, \{\mtheta_k\}}: \mtheta_k \in\mathbb{R}^p_{+}, {c_1p\over r}\leq |z_k^{-1}(a)| \leq {c_2 p\over r}, c_3\leq \onorm{\mat_k(\tS)}_{a:}\leq c_4 , \onormSize{}{\mtheta_{z_k^{-1}(a)}}_1=|z_k^{-1}(a)|, a \in [r], k\in[K]}.
-% \end{equation}
-\footnotesize
-\begin{align}
-    \tP(\{r_k\}) = \Bigg\{ \of{\{z_k\}, \tS, \{\mtheta_k\}}:\  &\mtheta_k \in\mathbb{R}^p_{+}, {c_1 p_k\over r_k} |z_k^{-1}(a)| \leq {c_2 p_k\over r_k}, c_3 \leq  \onorm{\mS_{k,a:}} \leq c_4 ,\onormSize{}{\mtheta_{k,z_k^{-1}(a)}}_1=|z_k^{-1}(a)|, a \in [r_k], k\in[K]\Bigg\}.
+\begin{align}
+ \tP(\{r_k\}) = \Bigg\{ &\of{\{z_k\}, \tS, \{\mtheta_k\}}:\\
+   & \mtheta_k \in\mathbb{R}^p_{+}, {c_1 p_k\over r_k} |z_k^{-1}(a)| \leq {c_2 p_k\over r_k}, c_3 \leq  \onorm{\mS_{k,a:}} \leq c_4 ,\onormSize{}{\mtheta_{k,z_k^{-1}(a)}}_1=|z_k^{-1}(a)|, a \in [r_k], k\in[K]\Bigg\}.
 \label{eq:general_family}
 \end{align}
-\normalsize
-% \begin{equation}
-%     \tP = \offf{ \of{\{z_k\}, \tS, \{\mtheta_k\}}: \mtheta_k \in\mathbb{R}^p_{+}, |z_k^{-1}(a)| \asymp {p_k\over r_k},  \onorm{\mS_{k,a:}} \asymp 1 , \onormSize{}{\mtheta_{k,z_k^{-1}(a)}}_1=|z_k^{-1}(a)|, a \in [r], k\in[K]}.
-% \end{equation}
-
-We call the collection of degree heterogeneity $\{\mtheta_k\}$ is balanced if for all $k \in [K]$,
+
+We call the degree heterogeneity $\{\mtheta_k\}$ is balanced if for all $k \in [K]$,
 \begin{equation}\label{eq:general_balanced}
@@
 We also consider the generalized Assumption~\ref{assmp:min_gap} on angle gap.
-\begin{assumption}[Generalized angle gap]\label{assmp:general_minimal_gap} We assume the minimal gap between normalized rows of $\mS_k$ is bounded away from zero, i.e.,
-\begin{equation}
-     \Delta_{\min} \coloneqq \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onorm{ \mS_{k,a:}^s - \mS_{k, b:}^s } > 0
-\end{equation}
-for all $k \in [K]$. Similarly, let $\text{SNR} = \Delta_{\min}^2/\sigma^2$ with generalized minimal gap $\Delta_{\min}^2$ in Assumption~\ref{assmp:general_minimal_gap}. We define the regime
-\begin{equation}
-    \tP(\gamma) = \tP (\{r_k\}) \cap\{\tS \text{ satisfies $\text{SNR} = p^{\gamma}$ and $p_k \asymp p, k \in [K]$} \}.
+\begin{assumption}[Generalized angle gap]\label{assmp:general_minimal_gap} Recall $\mS_k=\mat_k(\tS)$.
+We assume the minimal gap between normalized rows of $\mS_k$ is bounded away from zero for all $k\in[K]$; i.e.,
+\begin{equation}
+     \Delta_{\min} \coloneqq \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onorm{ \mS_{k,a:}^s - \mS_{k, b:}^s } > 0.
 \end{equation}
 \end{assumption}
-% by extending the definition~\eqref{eq:minimal_gap} to asymmetric setting, 
-% \begin{equation}\label{eq:general_minial_gap}
-%     \Delta_{\min} = \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onorm{ \frac{ \of{\mat_k(\tS)}_{a:} }{ \onorm{\of{\mat_k(\tS)}_{a:}}} - \frac{ \of{\mat_k(\tS)}_{b:} }{ \onorm{\of{\mat_k(\tS)}_{b:} }} } > 0.
-% \end{equation}
+Similarly, let $\text{SNR} = \Delta_{\min}^2/\sigma^2$ with the generalized minimal gap $\Delta_{\min}^2$ defined in Assumption~\ref{assmp:general_minimal_gap}. We define the regime
+\begin{equation}
+    \tP(\gamma) = \tP (\{r_k\}) \cap\{\tS \text{ satisfies $\text{SNR} = p^{\gamma}$ and $p_k \asymp p, \text{for all }k \in [K]$} \}.
+\end{equation}
 
@@
 \begin{proof}[Proof of Theorem~\ref{thm:unique}]
-%Recall that we prove Theorem~\ref{thm:unique} under the generalized dTBM~\eqref{eq:general_dtbm}.
-The identifiability for the symmetric dTBM~\eqref{eq:model_tensor} is indicated by simply setting $z_k = z, \mS_k = \mS, \mtheta_k = \mtheta, k \in [K]$.
-
-To study the identifiability, we consider the noiseless model with $\tE = 0$. Assume there exists two parameterizations satisfying
+
+To study the identifiability, we consider the noiseless model with $\tE = 0$. Assume there exist two parameterizations satisfying
 \begin{equation}\label{eq:another}
-    \tX=\tS\times_1\Theta_1 \mM_1 \times_2 \cdots \times_K \Theta_K \mM'_K=\tS'\times_1\Theta'_1 \mM'_1 \times_2 \cdots \times_K \Theta'_K \mM'_K
-\end{equation}
-with $\of{ \{z_k\}, \tS, \{ \mtheta_k \} } \in \tP(\{r_k\})$ and $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} } \in \tP(\{r_k'\})$. We prove the sufficient and necessary conditions separately.
-
-\begin{enumerate}
-    \item[$(\Leftarrow)$] For the necessity, it is equivalent to show that there exists a triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ is not identical to $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation if the model~\eqref{eq:general_dtbm} violates Assumption~\ref{assmp:general_minimal_gap}.
+    \tX=\tS\times_1\Theta_1 \mM_1 \times_2 \cdots \times_K \Theta_K \mM'_K=\tS'\times_1\Theta'_1 \mM'_1 \times_2 \cdots \times_K \Theta'_K \mM'_K,
+\end{equation}
+where $\of{ \{z_k\}, \tS, \{ \mtheta_k \} } \in \tP(\{r_k\})$ and $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} } \in \tP(\{r_k'\})$ are two sets of parameters.
+We prove the sufficient and necessary conditions separately.
+
+\begin{enumerate}[wide]
+    \item[$(\Leftarrow)$] For the necessity, it suffices to construct two distinct parameters up to cluster label permutation, if the model~\eqref{eq:general_dtbm} violates Assumption~\ref{assmp:general_minimal_gap}.
 Without loss of generality, we assume $\onorm{ \mS_{1,1:}^s - \mS_{1,2:}^s } = 0$.
 
- If $\mS_{1,1:}$ is a zero vector, consider $\mtheta_1'$ such that $\mtheta'_{z_1^{-1}(1)} \neq \mtheta_{z_1^{-1}(1)}$.
+ If $\mS_{1,1:}$ is a zero vector, construct $\mtheta_1'$ such that $\mtheta'_{1,z_1^{-1}(1)} \neq \mtheta_{1,z_1^{-1}(1)}$.
 Let $\{z'_k\} = \{z_k\}$, $\tS' = \tS$, and $\mtheta'_k = \mtheta_k$ for all $k = 2, \ldots, K$.
-Then the triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ is not identical to $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation.
-We have a similar story when $\mS_{1,2:}$ is a zero vector.
+Then the triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ is distinct from $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation.
+Similar conclusion holds when $\mS_{1,2:}$ is a zero vector.
 
@@
     \mM'_1 = \mM_1 \mR, \quad \tS' = \tS_0 \times_1 \mD, \quad \theta'_{1}(i) = \begin{cases}
-    \frac{1}{1+c} \theta_{1}(i) & i \in z_{1}^{-1}(1)\\
-     \frac{c}{1+c} \theta_{1}(i) & i \in z_{1}^{-1}(2)\\
-     \theta_{1}(i) & \text{ otherwise }
-    \end{cases},
+    \frac{1}{1+c} \theta_{1}(i) & i \in z_{1}^{-1}(1),\\
+     \frac{c}{1+c} \theta_{1}(i) & i \in z_{1}^{-1}(2),\\
+     \theta_{1}(i) & \text{ otherwise},
+    \end{cases}
 \end{equation}
 and $\mM'_k = \mM_k, \mtheta'_k = \mtheta_k$ for all $k = 2, \ldots, K$. Then we have constructed a
-triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ that is not identical to $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation.
-
-\item[$(\Rightarrow)$] For the sufficiency, it is equivalent to show that all possible triplets $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ are identical to $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation if the model~\eqref{eq:general_dtbm} satisfies Assumption~\eqref{assmp:general_minimal_gap}.
-We show the uniqueness of the parameters separately.
-
-%Suppose $(\tS', \mM', \Theta') $ satisfies assumption~\ref{eq:gap}. We want to show the decomposition~\eqref{eq:degree} is unique up to permutation.
+triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ that is distinct from $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation.
+
+\item[$(\Rightarrow)$] For the sufficiency, it suffices to show that all possible triplets $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ are identical to $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation if the model~\eqref{eq:general_dtbm} satisfies Assumption~\eqref{assmp:general_minimal_gap}.
+We show the uniqueness of the three parameters, $\{\mM_k\}, \{\tS\}, \{\mtheta_k\}$ separately.
 
 First, we show the uniqueness of $\mM_k$ for all $k \in [K]$.
-Specifically, we show the uniqueness of the first mode membership matrix, i.e., $\mM'_1 = \mM_1\mP_1$ where $\mP_1$ is a permutation matrix.
-The uniqueness of $\mM_k$ for $k = 2,\ldots, K$ can be showed in the same way, and we omit the repeated procedures.
-
-On one hand, consider the pair of nodes $(i,j)$ such that $z_1(i) = z_1(j)$.
-We have $\onormSize{}{\mX_{1, z_1(i):}^s - \mX_{1, z_1(j):}^s } = 0$ and thus $\onormSize{}{ (\mS')_{1, z'_1(i):}^{s} - (\mS')_{1, z'_1(j):}^{s} } = 0$ by Lemma~\ref{lem:angle}.
+Without loss of generality, we consider $k=1$ and show the first mode membership matrix; i.e., $\mM'_1 = \mM_1\mP_1$ where $\mP_1$ is a permutation matrix.
+The conclusion for $k\geq 2$ can be showed similarly and thus omitted.
+
+Consider an arbitrary node pair $(i,j)$.
+If $z_1(i) = z_1(j)$, then we have $\onormSize{}{\mX_{1, z_1(i):}^s - \mX_{1, z_1(j):}^s } = 0$ and thus $\onormSize{}{ (\mS')_{1, z'_1(i):}^{s} - (\mS')_{1, z'_1(j):}^{s} } = 0$ by Lemma~\ref{lem:angle}.
 Then, by Assumption~\eqref{assmp:general_minimal_gap}, we have $z'_1(i) = z'_1(j)$.
-On the other hand, consider the pair of nodes $(i,j)$ such that $z_1(i) \neq z_1(j)$.
-We have $ \onorm{\mX_{1,i:}^s - \mX_{1,j:}^s} \neq 0$ and thus $\onorm{ (\mS')_{1, z'_1(i):}^{s} - (\mS')_{1, z'_1(j):}^{s} } \neq 0$ by Lemma~\ref{lem:angle}.
+Conversely, if $z_1(i) \neq z_1(j)$, then we have $ \onorm{\mX_{1,i:}^s - \mX_{1,j:}^s} \neq 0$ and thus $\onorm{ (\mS')_{1, z'_1(i):}^{s} - (\mS')_{1, z'_1(j):}^{s} } \neq 0$ by Lemma~\ref{lem:angle}.
 Hence, we have $z'_1(i) \neq z'_1(j)$.
-Therefore, we have proven that $z'_1$ is equal to $z_i$ up to label permutation.
-
-Next, we show the uniqueness of $\mtheta_k$ for all $k \in [K]$ given that $z_k = z_k'$.
-Similarly, we show the detailed proof of the uniqueness of $\mtheta_1$, i.e., $\mtheta'_1 = \mtheta_1$, and omit the repeated procedures for $\mtheta_k, k = 2,\ldots, K$.
-
-Consider an arbitrary $j \in [p_1]$ such that $z_1(j) = a$. Then for all the node $i \in  z_1^{-1}(a)$ in the same cluster of $j$, we have 
+Therefore, we have proven that $z'_1$ is identical $z_i$ up to label permutation.
+
+Next, we show the uniqueness of $\mtheta_k$ for all $k \in [K]$ provided that $z_k = z_k'$. Similarly, consider $k=1$ only, and omit the procedure for $k\geq 2$.
+
+Consider an arbitrary $j \in [p_1]$ such that $z_1(j) = a$. Then for all the nodes $i \in  z_1^{-1}(a)$ in the same cluster of $j$, we have 
 \begin{equation}
@@
 \end{equation}
-Let $\theta'_1(j) = c\theta_1(j)$ for some constant $c$.
-By the equation~\eqref{eq:theta_uniq}, we have $\theta'_1(i) = c \theta_1(i)$ for all $ i \in  z_1^{-1}(a)$.
-Note that $(\{z_k\}, \tS', \{\mtheta'_k\}) \in \tP(\{r_k\})$.
-We have
+Let $\theta'_1(j) = c\theta_1(j)$ for some positive constant $c$.
+By equation~\eqref{eq:theta_uniq}, we have $\theta'_1(i) = c \theta_1(i)$ for all $ i \in  z_1^{-1}(a)$.
+By the constraint $(\{z_k\}, \tS', \{\mtheta'_k\}) \in \tP(\{r_k\})$, we have
 \begin{equation}
@@
 \end{equation}
-which implies $c = 1$. Hence, we have proven $\mtheta_1 = \mtheta'_1$ given that $z_1 = z'_1$.
-
-Last, we show the uniqueness of $\tS$, i.e., $\tS'=\tS\times_1 \mP^{-1}_1\times_2\cdots \times_K \mP^{-1}_K$, where $\mP_k, k \in [K]$ are permutation matrices.
-Given $z'_k = z_k, \mtheta'_k = \mtheta_k$, we have $\mM'_k = \mM_k \mP_k$ and $\mTheta'_k = \mTheta_k$ for all $k \in [K]$.
-
-Let $\mD_k = \off{ (\mTheta'_k \mM'_k)^T (\mTheta'_k \mM'_k) }^{-1} (\mTheta'_k \mM'_k)^T, k \in [K]$. By the parameterization, we have 
+which implies $c = 1$. Hence, we have proven $\mtheta_1 = \mtheta'_1$ provided that $z_1 = z'_1$.
+
+Last, we show the uniqueness of $\tS$; i.e., $\tS'=\tS\times_1 \mP^{-1}_1\times_2\cdots \times_K \mP^{-1}_K$, where $\mP_k$'s are permutation matrices for all $k\in[K]$.
+Provided $z'_k = z_k, \mtheta'_k = \mtheta_k$, we have $\mM'_k = \mM_k \mP_k$ and $\mTheta'_k = \mTheta_k$ for all $k \in [K]$.
+
+Let $\mD_k = \off{ (\mTheta'_k \mM'_k)^T (\mTheta'_k \mM'_k) }^{-1} (\mTheta'_k \mM'_k)^T, k \in [K]$. By the parameterization~\eqref{eq:another}, we have 
 \begin{align}
@@
 
-\begin{proof}[Proof of Lemma~\ref{lem:angle}] For simplicity, we show the detailed proof for $k = 1$ and drop the subscript $k$ in $\mX_k, \mS_k$.The repeated proofs for $k = 2, \ldots, K$ are omitted.
-
+\begin{proof}[Proof of Lemma~\ref{lem:angle}] Without loss of generality, we prove $k = 1$ only and drop the subscript $k$ in $\mX_k, \mS_k$ for notational convenience. 
 By tensor matricization, we have
@@
 \end{equation}
-Thus it is sufficient to show the following statement that for any index pair $(i,j)\in[p_1]^2$,
+Thus it suffices to show the following statement holds for any index pair $(i,j)\in[p_1]^2$,
 \begin{equation}
@@
 \end{equation}
-\begin{enumerate}
+\begin{enumerate}[wide]
     \item[$(\Leftarrow)$] Suppose $\onorm{ \off{\mS_{z_1(i):} \tilde \mM^T }^s - \off{\mS_{z_1(j):}\tilde \mM^T}^s} = 0$.
@@
 Therefore, we finish the proof of Lemma~\ref{lem:angle}.
-% \begin{equation}
-%     \tilde \mM^T \tilde \mM = \off{ \mM^T \Theta^2 \mM }^{\otimes K-1}, \quad \text{and} \quad \mM^T \Theta^2 \mM  = \text{diag} \of{ \sum_{j \in z^{-1}(1)} \theta(j)^2, ...,  \sum_{j \in z^{-1}(r)} \theta(j)^2}.
-% \end{equation}
 \end{proof}
@@
 We will prove a more general conclusion than the main paper by allowing growing $r_k$'s.
-Consider the generalized dTBM in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$.
+Consider the generalized dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$.
 Specifically, we will show that, under
@@
 
- Noticed that inequality~\eqref{eq:max} is a minimax lower bound, it suffices to show the inequality holds for a particular $(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)$.
+Since the inequality~\eqref{eq:max} is a minimax lower bound, it suffices to show the inequality holds for a particular $(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)$.
 Specifically, we consider the estimation problem based on a particular parameter point $(\{z_k\}, \tS, \{\mtheta_k\})$ with the following three properties:
 \begin{equation}\label{eq:construction}
-    \text{(i)}\  \theta_k(i)=1 \text{ for all }i\in[p]; \  \text{(ii)}\ % \onorm{{\mS_1\over \onormSize{}{\mS_1}}-{\mS_2\over \onormSize{}{\mS_2}}}=
-    \Delta_{\min}\lesssim  \left({p\over r}\right)^{-\frac{K-1}{2}}\sigma; \  \text{(iii)}\ |z^{-1}_k(a)|={p \over r} \in \mathbb{Z}_+ \text{ for all }a\in[r],
+    \text{(i)}\  \theta_k(i)=1 \text{ for all }i\in[p]; \quad  \text{(ii)}\ 
+    \Delta_{\min}\lesssim  \left({p\over r}\right)^{-\frac{K-1}{2}}\sigma; \quad  \text{(iii)}\ |z^{-1}_k(a)|={p \over r} \in \mathbb{Z}_+ \text{ for all }a\in[r],
 \end{equation}
 for all $k \in [K]$.
-% \begin{enumerate}[label=(\roman*)]\label{eq:construction}
-% \item $\theta(i)=1 \text{ for all }i\in[p]$.
-% \item $\onorm{{\mS_1\over \onormSize{}{\mS_1}}-{\mS_2\over \onormSize{}{\mS_2}}}=\Delta_{\min}\lesssim \left({p\over r}\right)^{-(K-1)/2}\sigma$.
-% \item $|z(a)|={p \over r} \text{ for all }a\in[r]$.
-% \end{enumerate}
 Furthermore, we define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation.
 Based on \citet[Proof of Theorem 6]{han2020exact}, we consider the minimax rate over the restricted family of $\hat z_k$'s for which the following three conditions are satisfied:
-\footnotesize
-\begin{equation}
-    \text{(iv)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \ \text{(v)} \ |T^c_k|\asymp {p\over r}; \ \text{(vi)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
+\begin{equation}
+    \text{(iv)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \quad \text{(v)} \ |T^c_k|\asymp {p\over r}; \quad \text{(vi)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
 \end{equation}
@@
 for all $k \in [K]$.
-% \begin{enumerate}[label=(\roman*)]
-% \setcounter{enumi}{3}
-%     \item $\hat z(i)=z(i)$ for all $i\in T$.
-%     \item $|T^c|\asymp {p\over r}$.
-%     \item $\min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z(i) \neq \pi\circ z (i)\} = \sum_{i\in[p]}\ind\{\hat z(i) \neq  z (i)\}$.
-% \end{enumerate}
 The construction of $T$ is precisely the same as \citet[Proof of Theorem 6]{han2020exact}. 
@@
 \end{equation}
-where $\hat z_k, z_k$ on the left hand side denote the generic clustering functions in $\tP(\gamma)$, $z_k$ on the right hand side denotes a particular parameter satisfying properties (i)-(vi), and the infimum on the right hand side is taken over the restricted family of $\hat z$ satisfying (iv)-(vi).
-Here, the factor $r^3=r\cdot r^2$ in~\eqref{eq:minimax} comes from two sources: $r^2\asymp {r\choose 2}$ comes from the multiple testing burden for all pairwise comparisons among $r$ clusters, and another $r$ comes from the number of elements $|T^c_k|\asymp {p\over r}$ to be clustered.
+where $\hat z_k$ and $z_k$ on the left hand side denote the generic clustering functions in $\tP(\gamma)$, $z_k$ on the right hand side denotes a particular parameter satisfying properties (i)-(vi), and the infimum on the right hand side is taken over the restricted family of $\hat z$ satisfying (iv)-(vi).
+Here, the factor $r^3=r\cdot r^2$ in~\eqref{eq:minimax} comes from two sources: $r^2\asymp {r\choose 2}$ comes from the multiple testing burden for all pairwise comparisons among $r$ clusters; and another $r$ comes from the number of elements $|T^c_k|\asymp p/r$ to be clustered.
 
 Next, we need to find the lower bound of the rightmost side in~\eqref{eq:minimax}.
-For simplicity, we only show the bound for the mode-1 case $k = 1$.
+For simplicity, we show the bound for the mode-1 case $k = 1$ only.
 We drop the subscripts $1$ in $z_1, T_1, \mS_1, \mtheta_1$ and omit the repeated procedures for the cases of $k = 2,\ldots, K$.
@@
 \begin{equation}
-    \mx_a = \off{\mat_1 \of{ \tS\times_2 \mM_2\times_3\cdots \times_K \mM_K  }}_{a:}, \quad \text{for all} \quad a \in [r],
+    \mx_a = \off{\mat_1 \of{ \tS\times_2 \mM_2\times_3\cdots \times_K \mM_K  }}_{a:}, \quad \text{for all} \ a \in [r],
 \end{equation}
@@
 The analysis for other $i\in T^c$ are similar.
-For notational simplicity, we suppressed the subscript $i$ and write $\my, \theta, z$ in place of $\my_1, \theta_1$ and $z(1)$, respectively.
+For notational simplicity, we suppress the subscript $i$ and write $\my, \theta, z$ in place of $\my_1, \theta_1$ and $z(1)$, respectively.
 The equivalent vector problem for assessing the summand in~\eqref{eq:minimax} is
@@
 \]
-and the maximum likelihood estimators (MLE) of $\theta$ and $z$ are
+and the MLE's of $\theta$ and $z$ are
 \begin{equation}\label{eq:MLE}
@@
 
-Neyman-Pearson Lemma implies
+The Neyman-Pearson Lemma implies
 \begin{align}\label{eq:lower}
@@
 By symmetric, it suffices to bound $\bbP[\hat z_{\text{MLE}}=1|z=2]$. Using~\eqref{eq:ratio}, we obtain
-%\fixme{Jiaxin}{$\theta^* = \hat \theta_{\text{MLE}}$?}
 \begin{align}\label{eq:bound}
@@
 \[
-\liminf_{p\to \infty} \sup_{\tX\in \tP_{\text{non-degree}}(\gamma)}\max_{k\in[K]}\mathbb{E}[p\ell(z_k, \hat z_k)]\geq 1.
+\liminf_{p\to \infty} \sup_{\tX\in \tP_{\text{non-degree}}(\gamma)}\max_{k\in[K]}\bbE[p\ell(z_k, \hat z_k)]\geq 1.
 \]
@@
 \[
-\liminf_{p\to \infty} \sup_{\tX\in \tP(\gamma)}\max_{k\in[K]}\mathbb{E}[p\ell(z_k, \hat z_k)]\geq 1.
+\liminf_{p\to \infty} \sup_{\tX\in \tP(\gamma)}\max_{k\in[K]}\bbE[p\ell(z_k, \hat z_k)]\geq 1.
 \]
@@
 
-%\fixme{Miaoyan}{To Jiaxin: I fixed the above proof for Thm 3. Please revise everything below to make the contexts consistent with our current proofs. }
-
-{\bf Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:comp}} 
-\begin{defn}[Principal angles] For two matrices $\mU, \hat \mU \in \mathbb{O}^{p \times r}$ %\fixme{Miaoyan}{do we need column organization  $\mathbb{O}(p,r)?$. Yes.}
-, we define the principal angles between $\mU$ and $\hat \mU$ by the following diagonal matrix
-\begin{equation}
-    \text{Angle}(\mU, \hat \mU) = \text{diag}( \arccos(\sigma_1), \cdots, \arccos (\sigma_r) ),
-\end{equation}
-where $\sigma_1 \geq \cdots \geq \sigma_r>0$ are singular values of $\mU^T \hat \mU$. 
-%\fixme{Miaoyan}{Geometry intuition. $\text{Angle}(\mU \mx, \hat \mU \mx) \leq \sigma_r$ for all $\mx$?} 
-We define the norm of the principal angles as 
-\begin{equation}
-     \onorm{\sin(\mU, \hat \mU)}_{\sigma} = \max_{i \in [r]} \sin ( \arccos \of{ \sigma_i} ).
-\end{equation}
-%\fixme{Miayan}{Properties: $\onormSize{}{\sin(\mU, \hat \mU)}_\sigma= \onormSize{}{\mU^{\perp}\hat \mU}_{\sigma}$.
-$\max_{\mx\in \mathbb{R}^r}\sin \text{Angle}(\mU \mx, \hat \mU x)\leq \onormSize{}{\mU^{\perp}\hat \mU}_{\sigma}$}
-
-%\[ J_{11}=\text{Angle}(\ma^T \mX \mV, \ma^T\mX\mV^{(t)}) \leq \text{Angle}(\mV^T\mX^T, \mV^{(t),T}\mX^T) \]
-
-% The Schatten-$q$-$\sin$-norm is defined as
-% \begin{equation}
-%     \onorm{\sin(\mU, \hat \mU)}_q := \of{ \sum_{i=1}^r \sin^q \of{ \arccos \of{ \sigma_i}  } }^{1/q} = \of{ \sum_{i=1}^r \of{ 1 - \sigma_i^2 }^{q/2} }^{1/q},
-% \end{equation}
-% for $1 \leq q \leq \infty$. We omit the subscript using $\onorm{ \sin(\mU, \hat \mU)}$ when $q = +\infty$.
-\end{defn}
-
-% \begin{defn}[Hypergraphic planted clique (HPC) detection]~\label{def:HPC} Consider an order-$K$ hypergraph $H = (V,E)$ where $V = [p]$ collects vertices and $E$ collects all the $K$-way edges.
-Let $\tH_k(p, 1/2)$ denote the Erd\H{o}s-R\'{e}nyi $K$-hypergraph where the edge $(i_1,\ldots, i_K)$ belongs to $E$ with probability $1/2$.
-Further, we let $\tH_K(p, 1/2, \kappa)$ denote the hyhpergraph with planted cliques of size $\kappa$.
-Specifically, we generate a hypergraph from $\tH_k(p, 1/2)$, pick $\kappa$ vertices uniformly from $[p]$, denoted $K$, and then connect all the hyperedges with vertices in $K$.
-Note that the clique size $\kappa$ can be a function of $p$, denoted $\kappa_p$.
-
-% The order-$K$ HPC detection aims to identify whether there exists a planted clique hidden in a Erd\H{o}s-R\'{e}nyi $K$-hypergraph.
-We formulate HPC detection as the following hypothesis testing problem
-% \begin{equation}
-%     H_0:\ H \sim \tH_K(p,1/2) \quad \text{versus} \quad H_1: \ H \sim \tH_K(p,1/2, \kappa_p).
-% \end{equation}
-% \end{defn}
-
-% \begin{conjecture}[Hypergraphic planted clique (HPC) conjecture]\label{hypo:HPC} Consider the HPC detection problem in Definition~\ref{def:HPC}.
-Suppose the sequence $\{\kappa_p\}$ such that $\limsup_{p \rightarrow \infty} \log \kappa_p/ \log \sqrt{p} \leq (1 - \tau)$.
-Then, for any sequence of polynomial-times test $\{ \varphi_p\}: H \mapsto \{0,1\}$ we have
-% \begin{equation}
-%     \liminf_{p \rightarrow \infty} \bbP_{H_0} \of{ \varphi_p(H) =1 } +  \bbP_{H_1} \of{ \varphi_p(H) =0} \geq \frac{1}{2}.
-% \end{equation}
-% \end{conjecture}
-
-We restate the Lemma 3 from \cite{zhang2018tensor} for self-contentedness. 
-\begin{lem}[Bounds for principal angle spectral norm.]\label{lem:equivalence_qnorm}
-    For any $\mU_1, \mU_2 \in \mathbb{O}_{p,r}$ and all $1 \leq q \leq +\infty$, we have
-    \begin{align}
-        \frac{1}{4} \onorm{\mU_1 \mU_1^T - \mU_2 \mU_2^T}_{\sigma} \leq \onorm{\sin(\mU_1, \mU_2) }_{\sigma} \leq  \onorm{\mU_1 \mU_1^T - \mU_2 \mU_2^T}_{\sigma}.
-    \end{align}
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:equivalence_qnorm}] See \citet[Proof of Lemma 3]{zhang2018tensor}.
-\end{proof}
-
-\begin{lem}[Computational lower bound of tensor decomposition. (Theorem 4 in \cite{zhang2018tensor})]\label{lem:hpc} Consider the tucker decomposition model 
-\begin{equation}
-    \tY = \tS \times_1 \mM_1 \times \cdots \times_K \mM_K + \tE,
-\end{equation}
-where $\tY \in \bbR^{p\otimes^K }, \tS \in \bbR^{r \otimes^K}$ with $r \geq 1$ is a full rank core tensor, $\mM_k \in \bbR^{p \times r}$ has orthonormal columns for all $k \in [K]$, and $\tE$ has independent entries following $N(0,\sigma^2)$.
-Let $\tX = \tS \times_1 \mM_1 \times \cdots \times_K \mM_K $.
-Define the singular signal $\lambda = \min_{k \in [K]}\lambda_{\min} (\Mat_k(\tX))$, where $\lambda_{\min}(\cdot)$ denotes the minimal non-zero singular value.
-Suppose Conjecture~\ref{hypo:HPC} holds for some $\tau \in (0,1)$.
-If $\lambda /\sigma \lesssim p^{K(1- \tau)/4}$, then for any polynomial times estimator $\hat \mM_k$, the following inequality holds
-\begin{align}
-    \liminf_{p \rightarrow \infty} \sup_{(\tS, \mM_1,...,\mM_K) \text{ has singular signal }\lambda} \bbE \onorm{ \sin(\hat \mM_k, \mM_k) }^2_{\sigma} \geq c,
-\end{align}
-%\fixme{Miaoyan}{what norm in Lemmas 7 and 8?}
-where $c >0$ is a positive constant. 
-%The term $\sin \Theta( \cdot, \cdot)$ denotes the angle distance between two matrices with $\Theta(\cdot, \cdot)$ referring to the principal angles between two matrices.
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:hpc}]
-See \citet[Proof of Theorem 4]{zhang2018tensor}.
-\end{proof}
-
-\begin{lem}[Angle distance and misclustering error]\label{lem:angle_mcr} Consider two membership matrices $\mM, \hat \mM \in \bbR^{p \times r}$ with corresponding assignments $z, \hat z: [p] \mapsto [r]$.
-If the angle distance $\onorm{\sin (\mM, \hat \mM)}_{\sigma} \geq c$ for a constant $c >0$, then the misclustering error $\min_{\pi \in \Pi}p\ell(\hat z,\pi \circ z )\geq 1$.
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:angle_mcr}]
-By Lemma~\ref{lem:equivalence_qnorm}, we know that for any permutation matrix $\mP \in \bbR^{r \times r}$,
-\begin{equation}\label{eq:sin}
-    \onorm{ \hat \mM \hat \mM^T  - \mM \mP \mP ^T\mM^T}_F \geq  \onorm{ \hat \mM \hat \mM^T  - \mM \mP \mP^T \mM^T}_{\sigma} \geq \onorm{\sin (\mM, \hat \mM)}_{\sigma} \geq c>0,
-\end{equation}
-where the second inequality follows from the invariance of $\sin$ distance to permutation.
-The inequality~\eqref{eq:sin} implies that there exists at least one disagreement in $z$ and $\hat z$; i.e., there exists a pair $(i,j)$ such that $z(i) = z(j)$ but $\hat z(i) \neq \hat z(j)$.
-Therefore, we have $\min_{\pi \in \Pi} p\ell(\hat z, \pi \circ z)\geq 1$.
-\end{proof}
-%\fixme{Miaoyan}{Effect of permutation $\pi$ in the proof?. } \textcolor{blue}{$\sin \Theta (\mM, \hat \mM)$ is invariant to the permutation.}
-%\fixme{Miaoyan}{State Lemma 3 of Zhang and Xia. }
-
-\newpage
-
 \section*{Proof of Theorem~\ref{thm:initial}}
-\begin{proof}[Proof of Theorem 4] We prove Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
+\begin{proof}[Proof of Theorem 4] We prove Theorem~\ref{thm:initial} under the symmetric dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
 We drop the subscript $k$ in the matricizations $\mM_k, \mX_k, \mS_k$.
-For simplicity, let $\hat z$ denote the output of Sub-Algorithm~\hyperref[alg:main]{1}, $\hat z^{(0)}$.
+For simplicity, let $\hat z$ denote the output, $\hat z^{(0)}$, of Sub-algorithm~\hyperref[alg:main]{1}.
 
@@
     \end{align}
-    where inequalities \eqref{eq:tr1} and \eqref{eq:tr2} follow the triangle inequality, \eqref{eq:defs} follows the definition of $S$, \eqref{eq:km} follows the update rule of $k$-means in Step 5 of Sub-Algorithm~\hyperref[alg:main]{1}, \eqref{eq:lem1} follows Lemma~\ref{lem:norm_diff}, and the last inequality~\eqref{eq:lem_err} follows Lemma~\ref{lem:two-step_esterror}.
+    where inequalities \eqref{eq:tr1} and \eqref{eq:tr2} follow from the triangle inequality, \eqref{eq:defs} follows from the definition of $S$, \eqref{eq:km} follows from the update rule of $k$-means in Step 5 of Sub-algorithm~\hyperref[alg:main]{1}, \eqref{eq:lem1} follows from Lemma~\ref{lem:norm_diff}, and the last inequality~\eqref{eq:lem_err} follows from Lemma~\ref{lem:two-step_esterror}.
 Also, note that
@@
     \end{equation}
-    where the equation follows by the definition of $S_0$. Therefore, combining the inequalities~\eqref{eq:theta_bound}, \eqref{eq:theta_p}, \eqref{eq:lem_err}, and \eqref{eq:s0}, we have 
+    where the equation follows from the definition of $S_0$. Therefore, combining the inequalities~\eqref{eq:theta_bound}, \eqref{eq:theta_p}, \eqref{eq:lem_err}, and \eqref{eq:s0}, we have 
     \begin{align}
@@
     \end{equation}
-    where the last inequality follows the definition $\text{SNR} = \Delta_{\min}^2/\sigma^2$.
+    where the last inequality follows from the definition $\text{SNR} = \Delta_{\min}^2/\sigma^2$.
 \end{proof}
@@
 
-\begin{cor}\label{cor:L0} 
-%Suppose we have the initial assignment $z^{(0)}$ from Sub-Algorithm~\hyperref[alg:main]{1} and the identity permutation minimizes the misclustering error, i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} $ $ \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
-Suppose SNR $\gtrsim p^{-K/2} \log p$ and  $\sigma^2 = 1$. Then, the misclustering loss for the initialization is upper bounded as 
-\begin{equation}
-    L^{(0)} =  \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(0)}(i) = b } \onormSize{}{ [ \mS_{ \pi^{(0)}( z(i)):}  ]^s - [ \mS_{b:}  ]^s  }^2 \lesssim \frac{\Delta_{\min}^2}{r \log p},
-\end{equation}
-where $\pi^{(0)}$ minimizes the initial misclustering error, i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} $ $ \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$.
+\begin{cor}[Initial misclustering loss]\label{cor:L0} 
+Under the setup of Theorem~\ref{thm:initial}, assume SNR $\geq \tilde  C p^{-K/2} \log p$ for some positive universal constant $\tilde C$ and $\sigma^2 = 1$.
+With probability going to 1, the misclustering loss for the initialization is upper bounded as
+\begin{equation}
+    L^{(0)} =  \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(0)}(i) = b } \onormSize{}{ [ \mS_{ \pi^{(0)}( z(i)):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq \frac{\bar C \Delta_{\min}^2}{ \tilde C r \log p},
+\end{equation}
+where $\pi^{(0)}$ minimizes the initial misclustering error; i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} $ $ \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$, and $\bar C > 1$ is a positive universal constant.
 \end{cor}
 
-\begin{proof}
-
-{
-\color{blue}
+\begin{proof} [Proof of Corollary~\ref{cor:L0}]
 Without loss of generality, we assume $\pi^{(0)}$ is the identity mapping such that  $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
-Note that $\mX_{i:}^s$ only have $r$ different values.
+Note that $\mX_{i:}^s$ have only $r$ different values.
 We let $\mX_a^s = \mX_{i:}^s$ for all $i$ such that $z(i) = a, a \in [r]$.
@@
 \end{equation}
-Thus, when $p$ is large enough, we have 
+Therefore, when $p$ is large enough, we have 
 \begin{align}
@@
 \end{align}
-where the first inequality follows by the assumption $\min_{i \in [p]} \theta(i) \geq c$, the second inequality follows by the inequality~\eqref{eq:theta_p}, and the last inequality comes from the inequality~\eqref{eq:cor_1}.
+where the first inequality follows from the assumption $\min_{i \in [p]} \theta(i) \geq c$, the second inequality follows from the inequality~\eqref{eq:theta_p}, and the last inequality comes from the inequality~\eqref{eq:cor_1}.
 
@@
 \end{align}
-where the first inequality follows by the assumption of $\theta(i)$ and triangle inequality, the second inequality follows by Lemma~\ref{lem:norm_diff}, and the last inequality follows by~\eqref{eq:cor_2}.
-In addition, with Theorem~\ref{thm:initial} and the condition SNR $\gtrsim p^{-K/2} \log p$, for any $a \in [r]$
+where the first inequality follows from the assumption of $\theta(i)$ and triangle inequality, the second inequality follows from Lemma~\ref{lem:norm_diff}, and the last inequality follows from~\eqref{eq:cor_2}.
+In addition, with Theorem~\ref{thm:initial} and the condition SNR $\gtrsim p^{-K/2} \log p$, for all $a \in [r]$, we have
 \begin{equation}
@@
 \end{equation}
-with $p$ large enough. Then, for any $a \in [r]$, we have 
+when $p$ is large enough. Therefore, for all $a \in [r]$, we have 
 \begin{align}
@@
 \end{align}
-where the last inequality follows the inequality~\eqref{eq:cor_2}.
-
-Therefore, we finally obtain 
+where the last inequality follows from the inequality~\eqref{eq:cor_2}.
+
+Finally, we obtain 
 \begin{align}
@@
      & \lesssim \frac{1}{p}  \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i)  \onormSize{}{ \mX_{i:}^s -  \mX_{z^{(0)}(i)}^s  }^2\\
-     & \lesssim \frac{1}{p}  \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \of{ \onormSize{}{\mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 + \onormSize{}{ \hat \mx_{z^{(0)}(i)} - \mX_{z^{(0)}(i)}^s  }^2 }, \\
-     & \lesssim \frac{r^{K}}{p^{K}} \of{ p^{K/2}r + pr^2 + r^K}, \\
-     & \lesssim \frac{\Delta_{\min}^2}{r \log p}
-\end{align}
-where the first inequality follows by Lemma~\ref{lem:angle_gap_x}, and the third inequality follows inequalities~\eqref{eq:cor_3} and \eqref{eq:cor_4}.
-
-}
-
-% Note that 
-% \begin{align}
-%      \sum_{i \in [p]} \onorm{\mX_{i:}^s - \hat \mx_a}^2 &\leq  2\sum_{i \in [p]} \onorm{\mX_{i:}^s - \hat \mX_{i:}^s}^2 + 2\sum_{i \in [p]} \onorm{\hat \mX_{i:}^s - \hat \mx_a }^2 \\
-%     & \leq \frac{2}{\min_{i \in [p]} \onorm{\mX_{i:}}^2}  \off{ \sum_{i \in [p]} \onorm{\mX_{i:}}^2  \onorm{\mX_{i:}^s - \hat \mX_{i:}^s}^2 +  \sum_{i \in [p]}  \onorm{\mX_{i:}}^2  \onorm{\hat \mX_{i:}^s - \hat \mx_a }^2 }\\
-%     & \lesssim \frac{r^{K-1} (1 + \eta)}{p^{K-1}} \sum_{i \in [p]} \onorm{\mX_{i:}}^2  \onorm{\mX_{i:}^s - \hat \mX_{i:}^s}^2 \\
-%     & \lesssim  \frac{r^{K-1} (1 + \eta)}{p^{K-1}}  \onorm{\hat  \tX - \tX}_F^2 \\
-%     & \lesssim \frac{r^{K-1} (1 + \eta)}{p^{K-1}} p^{-K/2}.
-% \end{align}
-
-% Suppose $z^{(0)} = 1$.  Note that $\mX^s_{i:}$ only have $r$ different values. Let $\mX_{a}^s = \mX^s_{i:}$ with $z(i) = a, a \in [r]$. Note that 
-% \begin{equation}
-%     |z^{-1}(a) \cap (z^{(0)})^{-1} (a)| \geq |z^{-1}(a)| - p\ell(z^{(0)} ,  z) \gtrsim \frac{p}{r} - \frac{1}{\log p} \gtrsim \frac{p}{r},
-% \end{equation}
-% when $\Delta_{\min}^2 \geq p^{-K/2} \log p$. Then
-% \begin{align}
-%     \onorm{\mX_a^s - \hat \mx_a}^2 &= \frac{ \sum_{i \in z^{-1}(a) \cap (z^{(0)})^{-1} (a) } \onorm{ \mX_{i:}^s - \hat \mx_a}^2 }{|z^{-1}(a) \cap (z^{(0)})^{-1} (a) |} \\
-%     & \lesssim \frac{ r \sum_{i \in [p]} \onorm{\mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 }{ p}\\
-%     & \lesssim \frac{r^K(1 + \eta)}{p^{K}} p^{-K/2} 
-% \end{align}
-
-% Note that 
-% \begin{align}
-%     L^{(0)} &= \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \onorm{\mS_{z^{(0)}(i),:}^s - \mS_{z(i),:}^s}^2 \\
-%     & \lesssim \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \onorm{\mX^s_{z^{-1}(z^{(0)}):} - \mX^s_{i,:}}^2\\
-%     & \lesssim \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \off{ \onorm{ \mX^s_{z^{-1}(z^{(0)}):} - \hat \mx_{z^{(0)}(i)} }^2 + \onorm{ \hat \mx_{z^{(0)}(i)} - \mX^s_{i,:}}^2 }\\
-%     & \lesssim   \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \frac{ \onorm{\hat \mX_{i:} - \mX_{i:}}^2 }{ \onorm{ \mX_{i:}}^2 } +  \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i)  \frac{r^K(1 + \eta)}{p^{K}} p^{-K/2} \\
-%     & \lesssim \frac{r^K}{p^{K}} p^{-K/2} \\
-%     & \lesssim \frac{\Delta_{\min}^2}{r \log p}.
-%     %& \lesssim \frac{1}{p} \lambda^{-2(K-1)}_r(\mTheta \mM) \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \onorm{\mX^{(0)}_{i:}-\mX_{i:}}^2 \\
-%     %& \lesssim \frac{r^{K-1}}{p^K} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \off{ \onorm{\mX^{(0)}_{i:}-\hat \mX_{i:}}^2 +  \onorm{\hat \mX_{i:}-\mX_{i:}}^2 }
-% \end{align}
-% where $\frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \lesssim \frac{1}{p} \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i)^2$ due to the boundedness.
+     & \lesssim \frac{1}{p}  \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \of{ \onormSize{}{\mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 + \onormSize{}{ \hat \mx_{z^{(0)}(i)} - \mX_{z^{(0)}(i)}^s  }^2 } \\
+     & \leq \bar C \frac{r^{K}}{p^{K}} \of{ p^{K/2}r + pr^2 + r^K}, \\
+     & \leq \frac{\bar C\Delta_{\min}^2}{ \tilde C r \log p}
+\end{align}
+where the first inequality follows from Lemma~\ref{lem:angle_gap_x}, the third inequality follows from inequalities~\eqref{eq:cor_3} and \eqref{eq:cor_4}, and the last inequality follows from the assumption that SNR $\geq \tilde C p^{-K/2} \log p$.
 \end{proof}
@@
 \end{align}
-where the equations follows by the properties of trigonometric function and the inequality follows by the fact the $\cos \frac{\alpha}{2} \leq 1$ and $\sin \alpha = 2 \sin \frac{\alpha}{2} \cos \frac{\alpha}{2} > 0$ for $\alpha \in [0, \pi]$.
+where the equations follows from the properties of trigonometric function and the inequality follows from the fact the $\cos \frac{\alpha}{2} \leq 1$ and $\sin \alpha = 2 \sin \frac{\alpha}{2} \cos \frac{\alpha}{2} > 0$ for $\alpha \in [0, \pi]$.
 
@@
 
-\begin{defn}[Weighted padding vectors]\label{def:pad} For a vector $\ma = \entry{a_i} \in \bbR^d$, we define the padding vector of $\ma$ with the weight collection $\mw = \{\mw_i\}_{i = 1}^d$ and $\mw_i = \entry{w_{ik}} \in \bbR^{p_i}, i \in [d]$ as
-\begin{equation}
-    \pad_{\mw}(\ma) = [a_1 \circ \mw_1, \ldots, a_d \circ \mw_d]^T, \quad \text{ where} \quad  a_i \circ \mw_i = [a_i w_{i1},\ldots, a_i w_{i p_i}]^T, \text{ for all } i \in [d]
-\end{equation}
-and $\pad: \bbR^{d}\mapsto \bbR^{\sum_{i\in[d]}p_i}$ can be viewed as an operator.
-%, and by definition we have $\pad (\ma) \in \bbR^{\sum_{i \in[d]} p_i}$. 
+\begin{defn}[Weighted padding vectors]\label{def:pad} For a vector $\ma = \entry{a_i} \in \bbR^d$, we define the padding vector of $\ma$ with the weight collection $\mw = \{\mw_i\colon \mw_i = \entry{w_{ik}} \in \bbR^{p_i}\}_{i = 1}^d$ as
+\begin{equation}\label{eq:paddef}
+    \pad_{\mw}(\ma) = [a_1 \circ \mw_1, \ldots, a_d \circ \mw_d]^T, \quad \text{ where} \quad  a_i \circ \mw_i = [a_i w_{i1},\ldots, a_i w_{i p_i}]^T, \text{ for all } i \in [d].
+\end{equation}
+Here we also view $\pad_{\mw}(\cdot)\colon \bbR^{d}\mapsto \bbR^{\sum_{i\in[d]}p_i}$ as an operator. 
 We have the bounds of the weighted padding vector
@@
 \begin{lem}[Angle for weighted padding vectors]\label{lem:pad} Suppose we have two non-zero vectors $\ma, \mb \in \bbR^d$. Given the weight collection $\mw$, we have 
-\begin{equation}
-   \frac{\min_{i \in [d]} \onormSize{}{\mw_i}}{\max_{i \in [d]} \onormSize{}{\mw_i} } \sin (\ma, \mb) \leq \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb)) \leq \frac{ \max_{i \in [d]} \onormSize{}{\mw_i} }{\min_{i \in [d]} \onormSize{}{\mw_i}} \sin (\ma, \mb).
+\begin{equation}\label{eq:lempad}
+   \frac{\min_{i \in [d]} \onormSize{}{\mw_i}}{\max_{i \in [d]} \onormSize{}{\mw_i} } \sin (\ma, \mb) \stackrel{*}{\leq} \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb)) \stackrel{**}{\leq} \frac{ \max_{i \in [d]} \onormSize{}{\mw_i} }{\min_{i \in [d]} \onormSize{}{\mw_i}} \sin (\ma, \mb).
 \end{equation}
@@
 
-First, we prove the inequality $\sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb)) \leq \frac{ \max_{i \in [d]} \onormSize{}{\mw_i}^2 }{\min_{i \in [d]} \onormSize{}{\mw_i}^2} \sin (\ma, \mb)$.
-Decomposing $\mb$, we have
+First, we prove the inequality ** in~\eqref{eq:lempad}. Decomposing $\mb$ yields
 \begin{equation}
@@
 \end{equation}
-where $\ma^{\perp} \in \bbR^d$ is the orthogonal vector of $\ma$. By the Definition~\ref{def:pad}, we have 
+where $\ma^{\perp} \in \bbR^d$ is in the orthogonal complement space of $\ma$. By the Definition~\ref{def:pad}, we have 
 \begin{equation}\label{eq:pad_second_ineq}
@@
 \end{equation}
-Note that $\pad_{\mw}(\ma^{\perp})$ is not necessary equal to the orthogonal vector of $\pad(\ma)$, i.e., $\pad_{\mw}(\ma^{\perp}) \neq (\pad_{\mw}(\ma))^{\perp}$.
-Then by the geometry property of trigonometric functions, we obtain
+Note that $\pad_{\mw}(\ma^{\perp})$ is not necessary equal to the orthogonal vector of $\pad(\ma)$; i.e., $\pad_{\mw}(\ma^{\perp}) \neq (\pad_{\mw}(\ma))^{\perp}$.
+By the geometry property of trigonometric functions, we obtain
 \begin{align}
@@
 \end{align}
-where the second inequality follows by applying the property~\eqref{eq:pad_bound} to vectors $\mb, \ma^{\perp}$.
-
-Next, we prove $\frac{\min_{i \in [d]} \onormSize{}{\mw_i}^2}{\max_{i \in [d]} \onormSize{}{\mw_i}^2 } \sin (\ma, \mb) \leq \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb))$.
-With the decomposition of $\pad_{\mw}(\mb)$ and the inverse weighted padding operator, we have
+where the second inequality follows by applying the property~\eqref{eq:pad_bound} to vectors $\mb$ and $\ma^{\perp}$.
+
+Next, we prove inequality * in~\eqref{eq:lempad}. With the decomposition of $\pad_{\mw}(\mb)$ and the inverse weighted padding operator, we have 
 \begin{align}
@@
 \end{align}
-where the second inequality follows by applying the property~\eqref{eq:pad_bound} to vectors $\mb, \pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp})$.
+where the second inequality follows by applying the property~\eqref{eq:pad_bound} to vectors $\mb$ and $\pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp})$.
 \end{proof}
 
-\begin{lem}[Angle gap in $\tX$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor}.
+\begin{lem}[Angle gap in $\tX$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family}.
 Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying~\eqref{eq:degree}.
-Then the angle gap in $\tX$ is approximate to angle gap in $\tS$, i.e., for all $i,j$ such that $z(i) \neq z(j)$
-% \begin{equation}
-%     \Delta_{\min} \leq \min_{(i,j): z(i) \neq z(j)}  \onormSize{}Size{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s}  \lesssim \min_{(i,j): z(i) \neq z(j)} \onormSize{}Size{}{\mX_{i:}^s - \mX_{j:}^s}. 
-% \end{equation}
+Then the angle gap in $\tX$ is asymptotically lower bounded by the angle gap in $\tS$; i.e., for all $i,j$ such that $z(i) \neq z(j)$
 \begin{equation}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:angle_gap_x}] 
-Note that the vector $\mS_{z(i):}$ can be folded to a tensor $\tS' =\entry{\tS'_{a_2,\ldots,a_K}} \in \bbR^{r^{K-1}}$, i.e., $\text{vec}(\tS') = \mS_{z(i):}$.
-Let the weight vectors $\mw_{a_2, \cdots, a_K}$ correspond to the elements in $\tS'_{a_2,\ldots,a_K}$ as
+Note that the vector $\mS_{z(i):}$ can be folded to a tensor $\tS' =\entry{\tS'_{a_2,\ldots,a_K}} \in \bbR^{r^{K-1}}$; i.e., $\text{vec}(\tS') = \mS_{z(i):}$.
+Define weight vectors $\mw_{a_2, \cdots, a_K}$ correspond to the elements in $\tS'_{a_2,\ldots,a_K}$ by
 \begin{equation}
@@
 
-Consider the inner product of $\mX_{i:}$ and $\mX_{j:}$ for $z(i) \neq z(j)$. By the definition of weighted padding operator and the balanced assumption~\eqref{eq:pad_balance}, we have 
+Consider the inner product of $\mX_{i:}$ and $\mX_{j:}$ for $z(i) \neq z(j)$.
+By the definition of weighted padding operator~\eqref{eq:paddef} and the balanced assumption~\eqref{eq:pad_balance}, we have
 \begin{align}
@@
 \end{align}
-Thus, when $p$ large enough, the inner product $\ang{\mX_{i:},\mX_{j:}} $ as the same sign with $\ang{\mS_{z(i):}, \mS_{z(j):}}$. Next, we discuss the angle between $\mX_{i:}$ and $\mX_{j:}$ by cases.
-
-\begin{enumerate}
+Therefore, when $p$ large enough, the inner product $\ang{\mX_{i:},\mX_{j:}} $ has the same sign as $\ang{\mS_{z(i):}, \mS_{z(j):}}$.
+Next, we discuss the angle between $\mX_{i:}$ and $\mX_{j:}$ by two cases.
+
+\begin{enumerate}[wide]
     \item \textbf{Case 1:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} \leq 0$.
@@
     \end{align}
-    where the second inequality follows by Lemma~\ref{lem:pad}, and the last inequality follows by the balanced weight~\eqref{eq:pad_balance} and Lemma~\ref{lem:norm_diff}.
+    where the second inequality follows from Lemma~\ref{lem:pad}, and the last inequality follows from the balanced weight~\eqref{eq:pad_balance} and Lemma~\ref{lem:norm_diff}.
 \end{enumerate}
 
-Hence, we conclude that for all $i,j$ such that $z(i) \neq z(j)$
-% \begin{equation}
-%   \min_{(i,j): z(i) \neq z(j)} \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \gtrsim \min_{(i,j): z(i) \neq z(j)}  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
-% \end{equation}
+Hence, we conclude that for all $i,j$ such that $z(i) \neq z(j)$,
 \begin{equation}
@@
 
-% % generalized version
-% \begin{lem}[Angle gap in $\tX$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:general_dtbm}.
-Suppose Assumption~\ref{assmp:general_minimal_gap} holds and $\{\mtheta_k\}$ is balanced satisfying~\eqref{eq:general_balanced}.
-Then the angle gap in $\tX$ is approximate to angle gap in $\tS$, i.e.,
-% \begin{equation}
-%     \min\{1, \Delta_{\min}\} \lesssim \min_{k \in [K]} \min_{(i,j): z_k(i) \neq z_k(j)} \onorm{\mX_{k,i:}^s - \mX_{k,j:}^s}. % \lesssim \Delta_{\min}.
-% \end{equation}
-% \end{lem}
-
-% \begin{proof}[Proof of Lemma~\ref{lem:angle_gap_x}] 
-% For simplicity, we consider the case with $k = 1$ and omit the subscript $k$ in the matricizations $\mX,\mS$.
-
-% Note that the vector $\mS_{z_1(i):}$ can be folded to a tensor $\tS' =\entry{\tS'_{a_2,\ldots,a_K}} \in \bbR^{r_2 \times \cdots \times r_K}$, i.e., $\text{vec}(\tS') = \mS_{z_1(i):}$.
-Let the weight vector $\mw_{a_2 \cdots a_K}$ correspond to the elements $\tS'_{a_2,\ldots,a_K}$ as
-% \begin{equation}
-%     \mw_{a_2 \cdots a_K} = [ \mtheta_{2, z_2^{-1} (a_2)} \otimes \cdots \otimes \mtheta_{K, z_K^{-1} (a_K)}] \in \bbR^{|z_2^{-1} (a_2)| \times \cdots \times |z_K^{-1} (a_K)|},
-% \end{equation}
-% for all $a_k \in [r_k], k = 2,\ldots, K$, where $\otimes$ denotes the Kronecker product.
-Therefore, we have  $\mX_{i:} = \theta_1(i) \pad_{\mw}(\mS_{z_1(i):})$ where $\mw = \{ \mw_{a_2 \cdots a_K}\}$.
-Specifically, we have $\onorm{\mw_{a_2,\ldots,a_K}}^2 = \prod_{k = 2}^K \onorm{ \mtheta_{k,z_k^{-1}(a_k)}}^2$, and by the balanced assumption we have
-% \begin{equation}\label{eq:pad_balance}
-%     \max_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}}^2 = (1 + o(1))  \min_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}}^2.
-% \end{equation}
-
-% Consider the inner product of $\mX_{i:}$ and $\mX_{j:}$ for $z_1(i) \neq z_1(j)$. By the definition of weighted padding operator and the balanced assumption~\eqref{eq:pad_balance}, we have 
-% \begin{align}
-%     \ang{\mX_{i:},\mX_{j:}} &= \theta_1(i) \theta_1(j) \ang{ \pad_{\mw}(\mS_{z_1(i):}), \pad_{\mw}(\mS_{z_1(j):} )}\\
-%     &= \theta_1(i) \theta_1(j) \min_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}}^2 \ang{\mS_{z_1(i):}, \mS_{z_1(j):}}(1 + o(1)).
-% \end{align}
-% Thus, when $p$ large enough, the inner product $\ang{\mX_{i:},\mX_{j:}} $ as the same sign with $\ang{\mS_{z(i):}, \mS_{z(j):}}$.
-Next, we discuss the angle between $\mX_{k,i:}$ and $\mX_{k,j:}$ by cases.
-
-% \begin{enumerate}
-%     \item \textbf{Case 1:} Suppose $\ang{\mS_{z_1(i):}, \mS_{z_1(j):}} \leq 0$. Then, we also have $\ang{\mX_{i:},\mX_{j:}} \leq 0$, which implies $\onorm{\mX_{i:}^s -  \mX_{j:}^s} \geq \sqrt{2}$.
-%     \item \textbf{Case 2:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} > 0$.
-Then, we have $\cos(\mS_{z(i):}, \mS_{z(j):}) > 0$.
-Note that the fact $\sqrt{1 - \cos \alpha} = 2 \sin \frac{\alpha}{2} \lesssim \sin \alpha$ for the angle $\alpha \in [0, \frac{\pi}{2})$.
-Then, we have
-%     \begin{align}
-%         \onorm{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} &= \sqrt{ 1 - \cos(\mS_{z_1(i):}, \mS_{z_1(j):})}  \\
-%         & \lesssim \sin (\mS_{z_1(i):}, \mS_{z_1(j):})\\
-%         & \leq \frac{ \max_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}} }{  \min_{(a_2,\ldots, a_K)} \onorm{\mw_{a_2,\ldots, a_K}} }  \sin ( \pad_{\mw} (\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):})) \\
-%         & \leq (1 + o(1)) \onorm{\mX_{i:}^s -  \mX_{j:}^s},
-%     \end{align}
-%     where the second inequality follows by Lemma~\ref{lem:pad}, and the last inequality follows by the balanced assumption~\eqref{eq:pad_balance} and Lemma~\ref{lem:norm_diff}.
-% \end{enumerate}
-
-% Hence, we conclude that for all $i,j \in [p_1], z_1(i) \neq z_1(j)$ 
-% \begin{equation}
-%     \onorm{\mX_{i:}^s -  \mX_{j:}^s} \gtrsim \min \offf{ 1, \onorm{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s}}, 
-% \end{equation}
-% and the same conclusion for matricizations $\mX_k, \mS_k, k \in [K]$ can be obtained by the same procedures.
-Therefore, by the definition $\Delta_{\min} \coloneqq \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onorm{ \mS_{k,a:}^s - \mS_{k, b:}^s }$, we finally obtain that
-% \begin{equation}
-%     \min_{k \in [K]} \min_{(i,j): z_k(i) \neq z_k(j)} \onorm{\mX_{k,i:}^s - \mX_{k,j:}^s} \gtrsim \min\{1, \Delta_{\min}\}.
-% \end{equation}
-
-% % Define an order-$(K-1)$  $(r,...,r)$-dimensional weight tensor $\tW$ with entries $\tW(a_2,...,a_K) \coloneqq \prod_{k=2}^K \onorm{\mtheta_{z^{-1}(a_k)}}^2$ and let $ \mw = \Vec(\tW)$.
-By Lemma~\ref{lem:weighted},
-% % \begin{equation}
-% %     \ang{ \mX_{k,i:},  \mX_{k,j:} } = \theta(i) \theta(j) \ang{ \mS_{k,z(i):}, \mS_{k,z(j):} }_{\mw} = \theta(i) \theta(j) \ang{ \mS_{k,z(i):}, \mS_{k,z(j):} } + o(1),
-% % \end{equation}
-% % and 
-% % \begin{equation}
-% %     \onorm{ \mX_{k,i:} } = \theta(i) \onorm{ \mS_{k,z(i):} }_{\mw} = \theta(i) \onorm{ \mS_{k,z(i):} } + o(1).
-% % \end{equation}
-% % Then, we have 
-% % \begin{equation}
-% %     \cos ( \mX_{k,i:},  \mX_{k,j:}  ) = \frac{ \ang{ \mX_{k,i:},  \mX_{k,j:} } }{\onorm{ \mX_{k,i:} } \onorm{ \mX_{k,j:} }} = \frac{\ang{ \mS_{k,z(i):}, \mS_{k,z(j):} } }{\onorm{ \mS_{k,z(i):} } \onorm{ \mS_{k,z(j):} }} + o(1) = \cos (\mS_{k,z(i):},\mS_{k,z(j):} ) + o(1).
-% % \end{equation}
-% % Hence, we finally obtain that 
-% % \begin{align}
-% %     \min_{k \in [K]} \min_{(i,j): z_k(i) \neq z_k(j)} \onorm{\mX_{k,i:}^s - \mX_{k,j:}^s}  &=  \min_{k \in [K]} \min_{(i,j): z_k(i) \neq z_k(j)} \sqrt{ 1 - \cos ( \mX_{k,i:},  \mX_{k,j:}  )}\\
-% %     &= \min_{k \in [K]} \min_{(i,j): z_k(i) \neq z_k(j)} \sqrt{ 1 - \cos (\mS_{k,z(i):},\mS_{k,z(j):} ) }  + o(1) \\
-% %     &\asymp \Delta_{\min},
-% % \end{align}
-% % where the last inequality follows by the minimal gap assumption.
-
-% % \fixme{Jiaxin}{ Problem here: if $\Delta_{\min}^2 = O(p^{-K/2})$, then $o(1)$ is much larger than $\Delta_{\min}$...}
-% \end{proof}
-
-\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the assumption that $\min_{i \in [p]} \theta(i) \geq c$, we have the singular values of $\mTheta \mM$ are bounded as
+\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the parameter space~\eqref{eq:family} and assumption that $\min_{i \in [p]} \theta(i) \geq c$, the singular values of $\mTheta \mM$ are bounded as
 \begin{equation}\label{eq:membership2}
@@
 \end{equation}
-By the definition of singular value, we have 
+By the definition of singular values, we have 
 \begin{equation}
@@
 
-Given that $\min_{i \in [p]}\theta(i) \geq c$, we have  
+Since that $\min_{i \in [p]}\theta(i) \geq c$ by the constraints in parameter space, we have  
 \begin{equation}
@@
 \end{equation}
-where the last inequality follows by the constraint in parameter space~\eqref{eq:family}. Also, notice that 
+where the last inequality follows from the constraint in parameter space~\eqref{eq:family}. Finally, notice that 
 \begin{equation}
@@
 
-\begin{lem}[Singular-value-gap-free tensor estimation error bound]\label{lem:two-step_esterror} Let an order-$K$ tensor $\tA = \tX + \tZ \in \bbR^{p \times \cdots \times p}$, and $\tX$ has tucker rank $(r,...r)$ and $\tZ$ has independent sub-Gaussian entries with parameter $\sigma^2$.
-Let $\hat \tX$ denote the two-step estimated tensor in Step 2 of Sub-Algorithm~\hyperref[alg:main]{1} in the main paper.
+\begin{lem}[Singular-value gap-free tensor estimation error bound]\label{lem:two-step_esterror}Consider an order-$K$ tensor $\tA = \tX + \tZ \in \bbR^{p \times \cdots \times p}$, where $\tX$ has Tucker rank $(r,...r)$ and $\tZ$ has independent sub-Gaussian entries with parameter $\sigma^2$.
+Let $\hat \tX$ denote the double projection estimated tensor in Step 2 of Sub-algorithm~\hyperref[alg:main]{1} in the main paper.
 Then with probability at least $1 - C \exp\of{- cp }$, we have
@@
 
-\begin{proof}
+\begin{proof}[Proof of Lemma~\ref{lem:two-step_esterror}]
 See \citet[Proposition 1]{han2020exact}.
@@
 
-\begin{lem}[Upper bound of misclustering error]\label{lem:upper_mis} Let $z: [p] \mapsto [r]$ be a clustering assignment such that $|z^{-1}(a)| \asymp p/r$ for all $a \in [r]$.
-Let node $i$ corresponds to a vector $\mx_i  = \theta(i) \mv_{z(i)} \in \bbR^d$ where $\{\mv_a\}_{a = 1}^r$ are the cluster centers and $\mtheta = \entry{\theta(i)} \in \bbR^p_+$ is the positive degree heterogeneity.
+\begin{lem}[Upper bound of misclustering error]\label{lem:upper_mis} Let $z: [p] \mapsto [r]$ be a cluster assignment such that $|z^{-1}(a)| \asymp p/r$ for all $a \in [r]$.
+Let node $i$ correspond to a vector $\mx_i  = \theta(i) \mv_{z(i)} \in \bbR^d$, where $\{\mv_a\}_{a = 1}^r$ are the cluster centers and $\mtheta = \entry{\theta(i)} \in \bbR^p_+$ is the positive degree heterogeneity.
 Assume that  $\mtheta$ satisfies the balanced assumption~\eqref{eq:degree} such that ${\max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 \over \min_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 }= 1 + o(1)$.
@@
 Then, if
-\begin{equation}
+\begin{equation}\label{eq:upper_mis_cond}
     \min_{a \neq b \in [r]} \onormSize{}{\mv_a - \mv_b} \geq 2c,
@@
 \end{equation}
-where $S_0$ is defined in Step 3 of Sub-Algorithm~\hyperref[alg:main]{1} and
-\begin{equation}
-   S = \{i \in S_0^c: \onormSize{}{\hat \mx_i - \mx_i } \geq c \}.
+where $S_0$ is defined in Step 3 of Sub-algorithm~\hyperref[alg:main]{1} and
+\begin{equation}
+   S = \{i \in S_0^c: \onormSize{}{\hat \mx_i - \mv_{z(i)} } \geq c \}.
 \end{equation}
@@
 \begin{equation}
-    C_u = \{ i \in z^{-1}(u) \cap S_0^c: \onormSize{}{\hat \mx_i - \mx_i} < c \},
+    C_u = \{ i \in z^{-1}(u) \cap S_0^c: \onormSize{}{\hat \mx_i - \mv_{z(i)}} < c \},
 \end{equation}
@@
 \end{align}
-Following \citet[Lemma 6]{gao2018community}, we have 
-\begin{equation}
-    \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \leq \sum_{i \in S_0} \theta(i)^2 + \sum_{i \in S} \theta(i)^2 + \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2,
-\end{equation}
-and $|R_3| \leq |R_1|$. We only need to bound $\sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2$ to finish the proof. Note that 
+Note that $\cup_{u \in [r]}C_u = S_0^c/S^c$ and $C_u \cap C_v = \emptyset$ for any $u \neq v$.
+Suppose there exist $ i \in C_u$ and $j \in C_v$ with $u \neq v \in [r]$ and $\hat z(i) = \hat z(j)$.
+Then we have
+\begin{equation}
+    \onormSize{}{\mv_{z(i)} - \mv_{z(j)} } \leq  \onormSize{}{\mv_{z(i)} - \hat \mx_{i} } + \onormSize{}{\mv_{z(j)} - \hat \mx_j } < 2c,
+\end{equation}
+which contradicts to the assumption~\eqref{eq:upper_mis_cond}.
+Hence, the estimates $\hat z(i) \neq \hat z(j)$ for the nodes $ i \in C_u$ and $j \in C_v$ with $u \neq v$.
+By the definition of $R_2$, the nodes in $\cup_{u \in R_2} C_u$ have the same assignment with $z$ and $\hat z$.
+Then,
+%Following \citet[Lemma 6]{gao2018community} \fixme{Miaoyan}{Remove citation. fill in details in a self-contained way. }, 
+we have 
+\begin{equation}
+    \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \leq \sum_{i \in S_0} \theta(i)^2 + \sum_{i \in S} \theta(i)^2 + \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2.
+\end{equation}
+ We only need to bound $\sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2$ to finish the proof.
+Note that every $C_u$ with $u \in R_3$ contains at least two nodes assigned to different clusters by $\hat z$.
+Then, we have $|R_2| + 2 |R_3| \leq r$.
+Since $|R_1| + |R_2| + |R_3| = r$, we have $|R_3| \leq |R_1|$.
+Hence, we obtain
 \begin{align}
@@
 
-\newpage
-
 \section*{Proof of Theorem~\ref{thm:refinement}}
 
-% Extra condition: 
-% \begin{equation}\label{cond:extra}
-%   c_0 < \min_{i \in [p]} \theta(i), \quad \Delta_{\min}^2 \gtrsim p^{-K/2}\log p,
-% \end{equation}
-% the second inequality requires the initialization error tends to 0 as $p \rightarrow \infty$. Otherwise, we can not handle $\eta$ in $A_{11}$ very well.
-
-%\fixme{Miaoyan}{We claim the following %lower bound on $\alpha$
-%\[
-%\sin(\mp(\bar z), \mp_{\theta}(\bar %z))\lesssim \sin\alpha.
-%\]
-%Therefore, the condition on $\sin(\mp(\bar %z), \mp_{\theta}(\bar z))$ is necessary.
-%Construct $\bar z$ to be the ground-truth %clustering.
-Then $\tilde \mD=\mI_r$ and %$\mD=\text{diag}(\mp)^{-1}\text{diag}(\mp_%{\mtheta})$
-%\[
-%\sin \alpha=\sin(\mS\tilde \mD, \mS) \geq %\lambda_{\min}(\mD){\min_{a\in[r]}\FnormSi%ze{}{\mS_{a:}}\over %\max_{a\in[r]}\FnormSize{}{\mS_{a:}}}\gtrs%im ? \sin(\mp,\mp_{\mtheta})
-%\]
-%}
-
-%  \begin{defn}[Locally linear stability of degrees] Let $z$ be the ground-truth clustering, and define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\epsilon)=\{\bar z\colon \ell(\bar z, z)\leq \epsilon\}$.
-We define two cluster-size vectors for $\bar z$,
-%  \[
-%  \mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T,\quad \mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
-%  \]
-% We call the degree is $\varepsilon$-locally linearly stable if and only if 
-% \begin{equation}
-%     \sin(\mp(\bar z),\ \mp_{\mtheta}(\bar z))\lesssim \varepsilon \Delta_{\min} ,\quad \text{for all } \bar z\in \tN(z, \varepsilon).
-% \end{equation}
-
-% \end{defn}
-
-%  \begin{rmk}
-%  The local stability holds trivially for $\varepsilon=0$ based on the construction of parameter space~\eqref{eq:family}.
-%  \end{rmk}
-
-%  \begin{assumption}[Degree stability]\label{assmp:stability}
-%  Assume the local linear stability of degree holds in all neighborhoods $\tN(z,\epsilon)$ for any $\epsilon \leq \log^{-1}p$.
-%  \end{assumption}
-
-%\begin{rmk}[Comparison to~\cite{gao2018community}]
-%We consider the special case when $r=2$, $z^{-1}(1)=\{1,\ldots,p\}$, $z^{-1}(2)=\{p+1,\ldots,2p\}$, $\sigma=1$, and 
-%\[
-%\mS=
-%\begin{bmatrix}
-%1 & s_1\\
-%1 & s_2
-%\end{bmatrix}.
-%\]
-%Under our model, the signal level is
-%\[
-%\Delta^2_{\min}\asymp 1- {1+\sqrt{(s^2_1-1)(s^2_2-1))}\over s_1s_2} \asymp 
-%\]
-%The assumption on $\theta$ is
-%\begin{enumerate}
-  %  \item $\theta_1+\cdots+\theta_{p}=p$ and $\theta_{p+1}+\cdots+\theta_{2p}=p$.
- %   \item $\mp(\bar z)=(p(1\pm \varepsilon),\ p(1\pm \varepsilon))^T$ and $\mp_{\mtheta}(\bar z)=(p\pm \sum_{i\in[I], |I|\leq p\varepsilon}\theta_i,\ p\pm \sum_{j\in[J], |J|\leq p\varepsilon}\theta_{p+j})^T$,
- %   \[
- %   {\max_{|I|\leq p\varepsilon}\sum_{j\in[I]}\theta_{i} \over 
- %   \min_{|J|\leq p\varepsilon}\sum_{j\in[J]}\theta_{p+j}
- %   }\lesssim \varepsilon \sqrt{|1-s|}.
- %   \]
-%\end{enumerate}
-%This model can not be incorporated into~\cite{gao2018community}, because $\mS$ is not assortative.
-% \end{rmk}
-
 \begin{proof}[Proof of Theorem~\ref{thm:refinement}]
-We prove Theorem~\ref{thm:refinement} under the dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
+We prove Theorem~\ref{thm:refinement} under the symmetric dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
 We drop the subscript $k$ in the matricizations $\mM_k, \mS_k, \mX_k$.
-Without loss of generality, we assume that the variance $\sigma = 1$ and the identity permutation minimizes the initial misclustering error, i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
-
-%Here, we also prove Theorem~\ref{thm:refinement} under the symmetric case, i.e., assuming $p_k = p, r_k = r$, and $\tS \in \bbR^{r \times \cdots \times r}, \tX \in \bbR^{p \times \cdots \times p}$ are symmetric with $z_k = z, \mM_k = \mM, \mtheta_k = \mtheta$.
-Without loss of generality, we assume $\sigma = 1$ and $\pi^{(0)}$ is the identity mapping on $[r]$.
-
-{\bf Step 1 (Notations and conditions):} We first introduce additional notations and the necessary condition to complete the proof.
+Without loss of generality, we assume that the variance $\sigma = 1$, and that the identity permutation minimizes the initial misclustering error; i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
+
+{\bf Step 1 (Notation and conditions).} We first introduce additional notations and the necessary conditions used in the proof.
 We will verify that the conditions hold in our context under high probability in the last step of the proof.
 
-\textbf{Notations.}
-\begin{enumerate}
-
-%\item Norms. 
+\textbf{Notation.}
+\begin{enumerate}[wide]
+
 \item Projection.
+We use $\mI_d$ to denote the identity matrix of dimension $d$.
 For a vector $\mv \in \bbR^d$, let $\text{Proj}(\mv) \in \bbR^{d \times d}$ denote the projection matrix to $\mv$.
-Then $\mI_d - \text{Proj}(\mv)$ is the projection matrix to the orthogonal complement $\mv^{\perp}$.
-
-    \item Normalized membership matrices
+Then, $\mI_d - \text{Proj}(\mv)$ is the projection matrix to the orthogonal complement $\mv^{\perp}$.
+
+    \item We define normalized membership matrices
     \begin{align}
@@
     \end{align}
-    \item Estimator of $\tS$ in the $t$-th iteration, $\tS^{(t)}$, and the oracle estimator of $\tS$ given true assignment $z$, $\tilde \tS$,
+    \item We use $\tS^{(t)}$ to denote the estimator of $\tS$ in the $t$-th iteration, and we use $\tilde \tS$ to denote the oracle estimator of $\tS$ given true assignment $z$; i.e.,
     \begin{align}
@@
     \end{align}
-    \item Matricization of tensors
+    \item We define the matricizations of tensors
     \begin{align}
@@
     \end{align}
-    \item The angle-based misclustering loss in the $t$-th iteration %\fixme{Jiaxin}{Need $\theta(i)$ in the definition}
+    \item We define the angle-based misclustering loss in the $t$-th iteration 
     \begin{equation}
@@
 
-\begin{condition}\label{cond:origin} Let $\mathbb{O}_{p, r}$ denote the collection of all the $p$-by-$r$ matrices with orthonormal columns. For all $a \in [r]$, we have 
+\begin{condition}(Intermediate results) \label{cond:origin} Let $\mathbb{O}_{p, r}$ denote the collection of all the $p$-by-$r$ matrices with orthonormal columns. We have 
 \begin{equation}\label{eq:cond1}
-    \onormSize{}{\mE \mV}_\sigma \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}} \of{ p ^{1/2}+ r^{(K-1)/2}}, \quad \onormSize{}{\mE \mV}_F \lesssim \sqrt{\frac{ r^{2(K-1)}}{p^{K-2}}},\quad \onormSize{}{ \mW_{a:}^T \mE \mV} \lesssim \frac{r^K}{p^{K/2}},
-\end{equation}
-% \begin{equation}\label{eq:cond11}
-%     , \quad \onormSize{}{ \mW_{a:}^{(t),T} \mE \mV} \lesssim \frac{r^K}{p^{K/2}},
-% \end{equation}
+    \onormSize{}{\mE \mV}_\sigma \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}} \of{ p ^{1/2}+ r^{(K-1)/2}}, \quad \onormSize{}{\mE \mV}_F \lesssim \sqrt{\frac{ r^{2(K-1)}}{p^{K-2}}},\quad \onormSize{}{ \mW_{a:}^T \mE \mV} \lesssim \frac{r^K}{p^{K/2}} \text{ for all }a \in [r],
+\end{equation}
 \begin{equation}\label{eq:cond2}
@@
 \begin{equation}\label{eq:cond_oracle}
-    \xi \lesssim \exp\of{ - \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}},
-\end{equation}
-
-% \begin{equation}\label{eq:cond_intial}
-%   \sum_{i: z^{(t)}(i) \neq z(i)} \theta(i)^2 \lesssim p^{-K/2 + 1} \Delta_{\min}^{-2}, \quad  L^{(t)} \lesssim \Delta_{\min}^2 / r, \quad \text{for} \quad t = 0, 1, \ldots, T.
-% \end{equation}
+    \xi \leq \exp\of{ - M \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}},
+\end{equation}
 
 \begin{equation}\label{eq:cond_intial}
-    L^{(t)} \lesssim \frac{\Delta_{\min}^2}{r \log p}, \quad \text{for} \quad t = 0, 1, \ldots, T.
-\end{equation}
-% \fixme{Jiaxin}{See corollary 1}
-
+    L^{(t)} \leq \frac{\bar C}{\tilde C} \frac{\Delta_{\min}^2}{r \log p}, \quad \text{for} \quad t = 0, 1, \ldots, T,
+\end{equation}
+where $M$ is a positive universal constant in inequality~\eqref{eq:constant_M}, $\bar C, \tilde C$ are universal constants in Corollary~\ref{cor:L0} and assumption SNR $\geq \tilde C p^{-K/2} \log p$, respectively.
 Further, inequality~\eqref{eq:cond1} holds by replacing $\mV$ to $\mV^{(t)}$ and $\mW_{:a}$ to $\mW_{:a}^{(t),T}$ when initialization condition~\eqref{eq:cond_intial} holds.
@@
 
-{\bf Step 2 (Misclustering loss decomposition):} Next, we derive the upper bound of $L^{(t+1)}$ for $t = 0 ,1, \ldots, T-1$.
-By Sub-Algorithm~\hyperref[alg:main]{2}, we update the assignment in $t$-th iteration via
+{\bf Step 2 (Misclustering loss decomposition).} Next, we derive the upper bound of $L^{(t+1)}$ for $t = 0 ,1, \ldots, T-1$.
+By Sub-algorithm~\hyperref[alg:main]{2}, we update the assignment in $t$-th iteration via
     \begin{equation}
@@
     \end{equation}
-    following the facts that $\onormSize{}{\ma^s - \mb^s}^2 = 1 - \cos(\ma, \mb)$ for vectors $\ma,\mb$ of same dimension and $\mat(\tY^\text{d}) = \mY \mV^{(t)}$ where $\tY^\text{d}$ is the reduced tensor defined in Step 8 of Sub-Algorithm~\hyperref[alg:main]{2}.
+    following the facts that $\onormSize{}{\ma^s - \mb^s}^2 = 1 - \cos(\ma, \mb)$ for vectors $\ma,\mb$ of same dimension and $\mat(\tY^\text{d}) = \mY \mV^{(t)}$ where $\tY^\text{d}$ is the reduced tensor defined in Step 8 of Sub-algorithm~\hyperref[alg:main]{2}.
 Then the event $z^{(t+1)}(i) = b$ implies
@@
     \end{equation}
-    {\color{blue} Note that the event~\eqref{eq:event} also holds for the degenerate entity $i$ with $\onormSize{}{\mY_{i:} \mV^{(t)}} = 0$ due to the convention that $\ma^s = 0$ if $\ma = 0$. }
-    Arranging the terms in \eqref{eq:event}, we obtain the decomposition
+Note that the event~\eqref{eq:event} also holds for the degenerate entity $i$ with $\onormSize{}{\mY_{i:} \mV^{(t)}} = 0$ due to the convention that $\ma^s = {\bf 0}$ if $\ma = {\bf 0}$. 
+    Arranging the terms in \eqref{eq:event} yields the decomposition
     \begin{equation}\label{eq:decomp}
@@
     where
-    % \begin{equation}\label{eq:decomp}
-    %     2 \ang{ \mE_{j:} \mV, \off{  \tilde \mS_{z(j):} }^s - \off{  \tilde \mS_{b:} }^s } \leq 2 \ang{ \mX_{j:} \mV^{(t)}, \off{  \mS_{b:}^{(t)}  }^s -\off{  \mS_{z(j):}^{(t)}  }^s  } + F_{jb}^{(t)},
-    % \end{equation}
     \begin{align}
@@
 
-  %\fixme{Miaoyan}{Add extra condition $\max_{a\in[R]}\onorm{\theta_{z^{-1}(a)}}\lesssim \sqrt{p}$ in order to guarantee $\onorm{\mX_{j:}\mV^{(t)}}<c$.
-Specifically, $\onorm{\mX_{j:}\mV^{(t)}} \leq \onorm{\mX_{j:}}\onorm{\mV^{(t)}}_{\sigma} \lesssim p^{-(K-1)/2}\max_{a\in[R]}\onorm{\theta_{z^{-1}(a)}}^{K-1} \lesssim c<\infty$.
-The second inequality is from Lemma 4 of Han et al.
-Add Lemma 4.}
-    Then, we upper bound the event
+   Therefore, the event $ \ind \offf{ z^{(t+1)}(i) = b }$ can be upper bounded as
     \begin{align}
@@
      \onormSize{}{\mX_{i:} \mV^{(t)}}  = \theta(i) \onormSize{}{\mS_{i:} (\mTheta \mM)^{\otimes (K-1),T} \mW^{(t), \otimes^{K-1}}}
-       % &\geq \theta(i) \onormSize{}{\mS_{i:} (\mTheta \mM)^{\otimes (K-1),T} } \lambda^{K-1}_r(\mW^{(t)}) \\
          \geq \theta(i) \onormSize{}{\mS_{z(i):}} \lambda_r^{K-1}(\mTheta \mM) \lambda^{K-1}_r(\mW^{(t)})
@@
     \end{align}
-    where the first inequality follows by the property of eigenvalues, the last inequality follows by Lemmas~\ref{lem:singular_thetam}, \ref{lem:membership}, the assumption that $\min_{a \in [r]} \onormSize{}{\mS_{z(i):}} \geq c_3$, and $m >0$ is a positive constant related to $ c_3$.
-Plugging the lower bound of $ \onormSize{}{\mX_{i:} \mV^{(t)}}$ \eqref{eq:xv_lower} into the inequality \eqref{eq:ind_b}, we obtain that
+    where the first inequality follows from the property of eigenvalues; the last inequality follows from Lemma~\ref{lem:singular_thetam}, Lemma~\ref{lem:membership}, and assumption that $\min_{a \in [r]} \onormSize{}{\mS_{z(i):}} \geq c_3>0$; and $m >0$ is a positive constant related to $ c_3$.
+Plugging the lower bound of $ \onormSize{}{\mX_{i:} \mV^{(t)}}$ \eqref{eq:xv_lower} into the inequality \eqref{eq:ind_b} gives
     \begin{equation}\label{eq:decomp1}
@@
     \end{align}
-
-    {
-    \color{blue} Take the weighted summation of~\eqref{eq:decomp1} over $i \in [p]$. We have 
+ Taking the weighted summation of~\eqref{eq:decomp1} over $i \in [p]$ yields 
     \begin{equation}
@@
     \end{equation}
-    and we define
+    Similarly to $\xi$ in~\eqref{eq:decomp2}, we define
     \begin{equation}
@@
 
-    {\bf Step 3 (Derivation of contraction inequality):} In this step we derive the upper bound of $\zeta_{ib}$ and thereof obtain the contraction inequality. Note that 
+    {\bf Step 3 (Derivation of contraction inequality).} In this step we derive the upper bound of $\zeta_{ib}$ and obtain the contraction inequality~\eqref{eq:proof_5_ineq}. Note that 
     \begin{align}
@@
         & \leq \theta(i) \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \ind \offf{z^{(t+1)}(i) = b, \frac{1}{4}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} } \\
-        & \lesssim \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \theta(i)m^2\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    }\\
-        &\leq C \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \theta(i)m^2\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    },\label{eq:zeta}
-    \end{align}
-    where the first inequality follows by the inequality~\eqref{eq:hib} in Lemma~\ref{lem:upper_fgh}, and the last inequality follows by the assumption that $\min_{i \in [p]} \theta(i) \geq c$.
+        & \leq 64 \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ cm^2\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    }
+        %&\leq  \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    },\label{eq:zeta}
+    \end{align}
+    where the first inequality follows from the inequality~\eqref{eq:hib} in Lemma~\ref{lem:upper_fgh}, and the last inequality follows from the assumption that $\min_{i \in [p]} \theta(i) \geq c>0$.
+Choose the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ that satisfies inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
 Following \citet[Step 4, Proof of Theorem 2]{han2020exact} and Lemma~\ref{lem:upper_fgh}, we have
     \begin{align}\label{eq:f_sum}
-          \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{(F_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \leq C_1 L^{(t)},
-    \end{align}
-    and 
-    \begin{align}
-         \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}  &\lesssim \frac{1}{p} \sum_{i \in [p]}\theta(i) \sum_{b \in [r]/z(i)}   \ind \offf{ z^{(t+1)}(i) = b}  (\Delta_{\min}^2 + L^{(t)}) \\
-         & \leq C_2 (L^{(t+1)} + L^{(t)}), \label{eq:g_sum}
-    \end{align}
-    where the last inequality follows by the definition of $L^{(t)}$ and the constraint of $\mtheta$ in parameter space~\eqref{eq:family}.
-Choosing the constants $C, C_1, C_2$ in~\eqref{eq:zeta}~\eqref{eq:f_sum}~\eqref{eq:g_sum} carefully, we have
+          \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{(F_{ib}^{(t)})^2}{ cm^2 \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \leq \frac{C_0 \bar C}{c m^2 \tilde C^2} L^{(t)},
+    \end{align}
+   for a positive universal constant $C$ and 
+    \begin{align}
+         \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}  &\leq \frac{1}{512} \frac{1}{p} \sum_{i \in [p]}\theta(i) \sum_{b \in [r]/z(i)}   \ind \offf{ z^{(t+1)}(i) = b}  (\Delta_{\min}^2 + L^{(t)}) \\
+         & \leq \frac{1}{512} (L^{(t+1)} + L^{(t)}), \label{eq:g_sum}
+    \end{align}
+    where the last inequality follows from the definition of $L^{(t)}$ and the constraint of $\mtheta$ in parameter space~\eqref{eq:family}. 
+    %Choosing the constants $C, \bar C, C_2$ in~\eqref{eq:zeta}~\eqref{eq:f_sum}~\eqref{eq:g_sum} carefully \fixme{Miaoyan}{be specific. Add one sentence to explain}, 
+   For $\tilde C$ also satisfies 
+    \begin{equation}\label{eq:tilde_c3}
+         \frac{C_0 \bar C}{c m^2 \tilde C^2} \leq \frac{1}{512},
+    \end{equation}
+    we have 
     \begin{equation}\label{eq:zeta_upper}
-        \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}^{(t)} \leq  C_4 L^{(t+1)}  + C_5 L^{(t)},
-    \end{equation}
-    for some small positive constants $C_4, C_5$. Plugging the inequality~\eqref{eq:zeta_upper} into the decomposition~\eqref{eq:decomp2}, we obtain the contraction inequality 
+        \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}^{(t)} \leq  \frac{1}{8} L^{(t+1)}  + \frac{1}{4} L^{(t)}.
+    \end{equation}
+    %for some small \fixme{Miaoyan}{In what sense "small"? Add one sentence.} positive constants $C_4, C_5$. 
+    Plugging the inequality~\eqref{eq:zeta_upper} into the decomposition~\eqref{eq:decomp2}, we obtain the contraction inequality 
     \begin{equation}\label{eq:decomp_t}
-          L^{(t+1)} \leq M \xi  + \rho L^{(t)},
-    \end{equation}
-     where $M \geq 1$ and $\rho \in [0,1)$ is the contraction parameter.
-
-     Combining the inequality~\eqref{eq:cond_oracle} in Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}, we have proven Theorem~\ref{thm:refinement} with inequality~\eqref{eq:decomp_t}.
-    }
-
-    % On one hand, the event $A_{ib}$ is a critical component of $\xi$, where
-
-    % %We will investigate $\bbP(A_{ib} = 1)$ when verifying the Condition~\ref{cond:origin}.
-
-    % On the other hand, the terms $ F_{ib}^{(t)}, G_{ib}^{(t)}$, and $H_{ib}^{(t)}$ in $B_{ib}$ describe the misclustering error in that the $t$-th iteration, which intuitively can be represented by $L^{(t)}$.
-Specifically, under Condition~\ref{cond:origin}, Lemma~\ref{lem:upper_fgh} and \citet[Step 4, Proof of Theorem 2]{han2020exact} imply that
-    % %\fixme{Jiaxin}{Need to elaborate more. Step 4 is not completely the same as Han. Need to care about $\theta$}
-    % \begin{equation}
-    %     \frac{1}{p} \sum_{i \in [p] } \theta(i) \sum_{b \in [r]/z(i)} B_{ib} \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2  \leq c' L^{(t+1)} + \rho L^{(t)},
-    % \end{equation}
-    % where $c', \rho \in (0,1)$ are two small positive constants.
-Hence, by the upper bound~\eqref{eq:decomp1} and the defintion of $L^{(t+1)}$, we obtain the upper bound for the misclustering loss in the $t+1$-th iteration
-    % \begin{equation}\label{eq:decomp_t}
-    %     L^{(t+1)} \leq M \xi + \rho L^{(t)},
-    % \end{equation}
-    % where $M \geq 1$ and $\rho$ is called the contraction parameter.
-
-    {\bf Step 4 (Verification of Condition~\ref{cond:origin}):} Last, we verify the Condition~\ref{cond:origin} under high probability to finish the proof.
-Note that the inequalities~\eqref{eq:cond1}, \eqref{eq:cond2}, and \eqref{eq:cond3} describes the property of the sub-Gaussian noise tensor $\tE$, and the readers can find the proof directly in \citet[Step 5, Proof of Theorem 2]{han2020exact}.
-Here, we only include the verification of inequalities~\eqref{eq:cond_oracle} and \eqref{eq:cond_intial}.
-
-    Now, we verify the oracle loss condition~\eqref{eq:cond_oracle}. Recall the definition of $\xi$
+          L^{(t+1)} \leq \frac{3}{2} \xi  + \frac{1}{2} L^{(t)},
+    \end{equation}
+     where $\frac{1}{2}$ is the contraction parameter.
+
+     Therefore, with $\tilde C$ satisfying inequalities~\eqref{eq:tilde_c3}, \eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}, we obtain the conclusion in Theorem~\ref{thm:refinement} via inequality~\eqref{eq:decomp_t} combining the inequality~\eqref{eq:cond_oracle} in Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.
+
+    {\bf Step 4 (Verification of Condition~\ref{cond:origin}).} Last, we verify the Condition~\ref{cond:origin} under high probability to finish the proof.
+Note that the inequalities~\eqref{eq:cond1}, \eqref{eq:cond2}, and \eqref{eq:cond3} describe the property of the sub-Gaussian noise tensor $\tE$, and the readers can find the proof directly in \citet[Step 5, Proof of Theorem 2]{han2020exact}.
+Here, we include only the verification of inequalities~\eqref{eq:cond_oracle} and \eqref{eq:cond_intial}.
+
+    Now, we verify the oracle loss condition~\eqref{eq:cond_oracle}. Recall the definition of $\xi$,
     \begin{align}
@@
 
-    Let $e_i = \mE_{i:} \mV$ denote the aggregated noise vector for all $i \in [p]$, and $e_i$ are independent mean-zero sub-Gaussian vector in $\bbR^{r^{K-1}}$.
-Each coordinate in $e_i$ is a independent mean-zero sub-Gaussian variable with sub-Gaussian norm upper bounded by $C\sqrt{r^{K-1}/p^{K-1}}$ with some positive constant $C$.
-    %\fixme{Miaoyan}{\color{red} explain $e_j$. See my comments at page 34. Please correct all similar issues throughout the appendix.} 
-    Note that probability 
+    Let $e_i = \mE_{i:} \mV$ denote the aggregated noise vector for all $i \in [p]$, and $e_i$'s are independent zero-mean sub-Gaussian vector in $\bbR^{r^{K-1}}$.
+The entries in $e_i$ are independent zero-mean sub-Gaussian variables with sub-Gaussian norm upper bounded by $m_1\sqrt{r^{K-1}/p^{K-1}}$ with some positive constant $m_1$.
+We have the probability inequality
     \begin{align}
@@
 
-    For $P_1$, notice that the inner product $\ang{e_j,\mS_{z(j):}^s - \mS_{b:}^s} $ is a sub-Gaussian variable with sub-Gaussian norm bounded by $C \sqrt{ r^{K-1}/ p^{K-1} } \onormSize{}{ \mS_{z(i):}^s -  \mS_{b:}^s }$.
+    For $P_1$, notice that the inner product $\ang{e_j,\mS_{z(j):}^s - \mS_{b:}^s} $ is a sub-Gaussian variable with sub-Gaussian norm bounded by $m_2 \sqrt{ r^{K-1}/ p^{K-1} } \onormSize{}{ \mS_{z(i):}^s -  \mS_{b:}^s }$ with some positive constant $m_2$.
 Then, by Chernoff bound, we have
@@
     \end{equation}
-    where with some positive constant $t$
+    where with some positive constant $t>0$,
     \begin{align}
@@
 
-    For $P_{21}$, note that the term $\mW^T_{: z(i)} \mE \mV = \frac{ \sum_{ j \neq i, j \in [p]} \ind \{ z(j) = z(i) \} e_j }{\sum_{j \in [p]} \ind \{ z(j) = z(i) \} }$  is a sub-Gaussian vector with sub-Gaussian norm bounded by $C_1 \sqrt{r^K/ p^K}$.
-We have
+    For $P_{21}$, note that the term $\mW^T_{: z(i)} \mE \mV = \frac{ \sum_{ j \neq i, j \in [p]} \ind \{ z(j) = z(i) \} e_j }{\sum_{j \in [p]} \ind \{ z(j) = z(i) \} }$  is a sub-Gaussian vector with sub-Gaussian norm bounded by $m_3 \sqrt{r^K/ p^K}$ with some positive constant $m_3$.
+This implies
         \begin{align}
@@
     \end{align}
-    where the first inequality follows by the basic inequality in Lemma~\ref{lem:norm_diff}, the second inequality follows by the assumption that $ \min_{a \in [r]} \onormSize{}{\mS_{z(i):}} \geq c_3$ in~\eqref{eq:family}, and the last inequality follows by the Bernstein inequality.
-
-    For $P_{22}$, the inner product $ \ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s}$ is also a sub-Gaussian variable with sub-Gaussian norm $ C \sqrt{ r^{K-1}/ p^{K-1} } t$, conditioned on $\onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s } < t$.
+    where the first inequality follows from the basic inequality in Lemma~\ref{lem:norm_diff}, the second inequality follows from the assumption that $ \min_{a \in [r]} \onormSize{}{\mS_{z(i):}} \geq c_3>0$ in~\eqref{eq:family}, and the last inequality follows from the Bernstein inequality.
+
+    For $P_{22}$, the inner product $ \ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s}$ is also a sub-Gaussian variable with sub-Gaussian norm $ m_4 \sqrt{ r^{K-1}/ p^{K-1} } t$, conditioned on $\onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s } < t$ with some positive constant $m_4$.
 Then, by Chernoff bound, we have
@@
     \end{equation}
-    Take $t =  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }$ in $P_{21}$ and $P_{22}$, and plug the inequalities \eqref{eq:p21} and \eqref{eq:p22} into to the upper bound for $P_2$~\eqref{eq:p2}.
+    We take $t =  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }$ in $P_{21}$ and $P_{22}$, and plug the inequalities \eqref{eq:p21} and \eqref{eq:p22} into to the upper bound for $P_2$ in~\eqref{eq:p2}.
 We obtain that
@@
     \end{equation}
-    Therefore, combining the upper bounds~\eqref{eq:p1_2} and \eqref{eq:p2_2}, we have 
+    Combining the upper bounds~\eqref{eq:p1_2} and \eqref{eq:p2_2} gives 
        \begin{equation}\label{eq:xi_p}
@@
     \begin{align}
-        \bbE[\xi] & =  \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \bbP \offf{  \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2 \\
+        \bbE \xi & =  \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \bbP \offf{  \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2 \\
         &\lesssim \frac{1}{p} \sum_{i \in [p]}  \theta(i) \max_{i \in [p], b \in [r]}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2   \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \\
-        & \lesssim  \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 },
-    \end{align}
-    where the first inequality follows by the constraint that $\sum_{i \in [p]} \theta(i) = p$.
+        & \leq \exp \of{  - M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }, \label{eq:constant_M}
+    \end{align}
+    where $M$ is a positive constant, the first inequality follows from the constraint that $\sum_{i \in [p]} \theta(i) = p$, and the last inequality follows from~\eqref{eq:xi_p}.
 
@@
     \begin{equation}
-        \bbP\of{ \xi \lesssim \bbE[\xi] + \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } } \geq 1 -  C_2 \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 },
-    \end{equation}
-    and thus the condition~\eqref{eq:cond_oracle} holds with probability at least $1 -  C_2 \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }$ for some constant $C_2 > 0$.
+        \bbP\of{ \xi \lesssim \bbE \xi + \exp \of{  -M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } } \geq 1 -  C \exp \of{  - M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 },
+    \end{equation}
+    and thus the condition~\eqref{eq:cond_oracle} holds with probability at least $1 -  C \exp \of{  - M\frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }$ for some constant $C > 0$.
 
     Finally, we verify the bounded loss condition~\eqref{eq:cond_intial} by induction.
-With output $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{2} and the assumption SNR $\gtrsim p^{-K/2} \log p$, by Corollary~\ref{cor:L0}, we have
-    \begin{equation}
-        L^{(0)} \lesssim \frac{\Delta_{\min}^2}{r \log p}
-    \end{equation}
-    when $p$ is large enough, and thus condition~\eqref{eq:cond_intial} holds for $t = 0$.
-Assume the condition~\eqref{eq:cond_intial} also holds for all $t \leq t_0$.
-Then, by the decomposition~\eqref{eq:decomp_t}, we have
-    \begin{align}
-         L^{(t_0+1)} &\lesssim M \xi + \rho L^{(t_0)} \\
-         & \lesssim  \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } + \frac{\Delta_{\min}^2}{r \log p} \\
-         & \lesssim \frac{\Delta_{\min}^2}{r\log p} ,
-    \end{align}
-    where the second inequality follows by condition~\eqref{eq:cond_oracle} and the last inequality follows by the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2} \log p$.
+With output $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{2} and the assumption SNR $\geq \tilde C p^{-K/2} \log p$, by Corollary~\ref{cor:L0}, we have
+    \begin{equation}
+        L^{(0)} \leq \frac{\bar C \Delta_{\min}^2}{\tilde C r \log p},\quad \text{    when $p$ is large enough}.
+    \end{equation}
+Therefore, the condition~\eqref{eq:cond_intial} holds for $t = 0$. Assume the condition~\eqref{eq:cond_intial} also holds for all $t \leq t_0$. Then, by the decomposition~\eqref{eq:decomp_t}, we have 
+    \begin{align}
+         L^{(t_0+1)} &\leq \frac{3}{2} \xi + \frac{1}{2} L^{(t_0)} \\
+         & \leq \exp \of{  - M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } + \frac{\Delta_{\min}^2}{r \log p} \\
+         & \leq \frac{\bar C}{\tilde C}\frac{\Delta_{\min}^2}{r\log p} ,
+    \end{align}
+    where the second inequality follows from the condition~\eqref{eq:cond_oracle} and the last inequality follows from the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2} \log p$.
 Thus, the condition~\eqref{eq:cond_intial} holds for $t_0 + 1$, and the condition~\eqref{eq:cond_intial} is proved by induction.
@@
 
-% \begin{lem}[Bernstein's type bound]\label{lem:bernstein}
-%  Let $e_j = \mE_j \mV$. Let $\eta_j = \mW_{:z_j}^T \mE \mV$, i.e.,
-%  \begin{align}
-%      \eta_j = \frac{\sum_{l \in [p]} \mone\offf{ z_l = z_j} e_l }{\sum_{l \in [p]} \mone\offf{ z_l = z_j}}.
-%  \end{align}
-%  Then, we have
-%  \begin{align}
-%      \bbP\of{ \onorm{\eta_j}^2 \geq \frac{C r}{p} (r^{K-1} + 2 \sqrt{x r^{K-1}} + 2 x)  } \leq \exp \of{- \frac{x}{\mtheta_{\max}^K \alpha_2}}.
-%  \end{align}
-% \end{lem}
-
-% \begin{proof}[Proof of Lemma~\ref{lem:bernstein}]
-% Note that $\eta_j$ is still a mean-0 sub Gaussian random vector with independent entries where $\text{Var}( (\eta_j)_i ) \leq \frac{C r^K}{p^K} \mtheta_{\max}^K \alpha_2$.
-Then, we obtain the bound following the Bernstein inequality.
-% \end{proof}
-
 {\bf Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:refinement}}
@@
 \end{equation}
-%Define the misclustering error in the $t$-th iteration as $\ell^{(t)} = \ell(z^{(t)}, z)$, where $\ell(\cdot,\cdot)$ is the misclustering error in Section~\ref{sec:limits}.
-Then, we have $\ell^{(t)} \leq L^{(t)}/\Delta_{\min}^2$.
 \end{lem}
@@
     \end{equation}
-    where the last inequality follows by the assumption $\min_{i \in [p]} \theta(i) \geq c$.
+    where the last inequality follows from the assumption $\min_{i \in [p]} \theta(i) \geq c>0$.
 \end{proof}
 
-\begin{lem}[Singular property of membership matrices]\label{lem:membership} Suppose the condition~\eqref{eq:cond_intial} holds.
-Then, for any $a \in [r]$, we have $|\of{z^{(t)}}^{-1}(a)| \asymp p/r$.
+\begin{lem}[Singular-value property of membership matrices]\label{lem:membership} Under the setup of Theorem~\ref{thm:refinement}, suppose the condition~\eqref{eq:cond_intial} holds.
+Then, for all $a \in [r]$, we have $|\of{z^{(t)}}^{-1}(a)| \asymp p/r$.
 Moreover, we have
@@
 
-\begin{proof}[Proof of Lemma~\ref{lem:membership}] See \citet[Proof of Lemma 4]{han2020exact} for inequality~\eqref{eq:membership1}.
-
-For inequality~\eqref{eq:WWsingular}, note that for all $k \in [r]$
+\begin{proof}[Proof of Lemma~\ref{lem:membership}] The proof for the inequality~\eqref{eq:membership1} can be found in \citet[Proof of Lemma 4]{han2020exact}
+
+For inequality~\eqref{eq:WWsingular}, note that for all $k \in [r]$,
 \begin{align}
@@
 \end{align}
-where $\text{eigen}_k(\mA)$ denotes the $k$-th largest eigenvalue of the square matrix $\mA$, the first inequality follows the fact that $\mW^T\mW$ is a diagonal matrix with elements of order $r/p$, and the second equation follows by the definition of singular value.
+where $\text{eigen}_k(\mA)$ denotes the $k$-th largest eigenvalue of the square matrix $\mA$, the first inequality follows the fact that $\mW^T\mW$ is a diagonal matrix with elements of order $r/p$, and the second equation follows from the definition of singular value.
 \end{proof}
 
-\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin}, we have 
+\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement},  assume the the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ is large enough to satisfy the inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
+We have
     \begin{equation}\label{eq:fib}
@@
     \begin{equation}\label{eq:gib}
-        \max_{i \in [p]} \max_{b \neq z(i)} \frac{ \of{  G_{ib}^{(t)} }^2  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2}  \lesssim \of{ \Delta_{\min}^2 +L^{(t)}},
-    \end{equation}
-   % where $c$ is a small positive constant near to 0.
+        \max_{i \in [p]} \max_{b \neq z(i)} \frac{ \of{  G_{ib}^{(t)} }^2  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2}  \leq  \frac{1}{512} \of{ \Delta_{\min}^2 +L^{(t)}},
+    \end{equation}
     \begin{equation}\label{eq:hib}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:upper_fgh}]
-We prove the inequalities in Lemma~\ref{lem:upper_fgh} separately.
-\begin{enumerate}
-    \item Upper bound for $F_{ib}^{(t)}$, inequality \eqref{eq:fib}. Recall the definition of $F_{ib}^{(t)}$ 
+We prove the each of the inequalities in Lemma~\ref{lem:upper_fgh} separately.
+\begin{enumerate}[wide]
+    \item Upper bound for $F_{ib}^{(t)}$, i.e., inequality \eqref{eq:fib}. Recall the definition of $F_{ib}^{(t)}$,
     \begin{align}
@@
          & \quad \quad \quad \quad \quad + 8\of{ \ang{ \mE_{i:} (\mV - \mV^{(t)} ), [  \tilde \mS_{z(i):} ]^s - [ \tilde  \mS_{b:} ]^s  }}^2 \\
-         & \lesssim \of{ \onormSize{}{\mE_{i:} \mV}^2 + \onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2 } \max_{a \in [r]^s} \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  } \\
+         & \leq 8 \of{ \onormSize{}{\mE_{i:} \mV}^2 + \onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2 } \max_{a \in [r]^s} \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  } \\
           & \quad \quad \quad \quad \quad +\onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2  \onormSize{}{  [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }. \label{eq:fib_decomp}
     \end{align}
-    Note that for any $a \in [r]$ 
+    Note that for all $a \in [r]$,
     \begin{align}
@@
     \end{align}
-    where the second inequality follows the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4} in Lemma~\ref{lem:intermediate}, the third inequality follows the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}, and the last inequality follows the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.
-
-    Also note that 
+    where the second inequality follows from the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4} in Lemma~\ref{lem:intermediate}, the third inequality follows from the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}, and the last inequality follows from the assumption that $\Delta_{\min}^2 \geq \tilde C p^{-K/2}\log p$.
+
+    Note that 
     \begin{align}
@@
     \end{align}
-    where the second inequality follows by Lemma~\ref{lem:norm_diff}, and the last inequality follows by the assumptions on $\onormSize{}{\mS_{a:}}$ in the parameter space~\eqref{eq:family}, the inequality~\eqref{eq:cond1} Condition~\ref{cond:origin} and the assumption $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.
+    where the second inequality follows from Lemma~\ref{lem:norm_diff}, and the last inequality follows from the assumptions on $\onormSize{}{\mS_{a:}}$ in the parameter space~\eqref{eq:family}, the inequality~\eqref{eq:cond1} in Condition~\ref{cond:origin} and the assumption $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.
 
@@
 
-    \item Upper bound for $G_{ib}^{(t)}$, inequality~\eqref{eq:gib}. Recall the definition of $G_{ib}^{(t)}$ and rearrange the terms
+    \item Upper bound for $G_{ib}^{(t)}$, i.e., inequality~\eqref{eq:gib}. By definition of $G_{ib}^{(t)}$, we rearrange terms and obtain
     \begin{align}
@@
         &= 2 \ang{  [\mX_{i:} \mV^{(t)}]^s , \of{[  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s } - \of{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s }}\\
-        %& \leq 4 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [  \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }.
       &= G_1 + G_2 - G_3, \label{eq:gib_decomp}
@@
         & \leq \max_{a \in [r]} \onormSize{}{ [ \mW_{:a}^T \mY \mV^{(t)} ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^4\\
-        & \lesssim \frac{r^4}{\Delta_{\min}^4} (L^{(t)})^4 +  \frac{r^2 r^{4K} + p^2r^{2K+4}}{p^{2K}} \frac{(L^{(t)})^2}{\Delta_{\min}^4} \\
-        & \leq c\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}}, \label{eq:g1}
-    \end{align}
-    where the third inequality follows by the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:membership} and the last inequality follows by the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
-
-    For $G_2$, noted that $[\mX_{i:} \mV^{(t)}]^s = [ \mW_{z(i):}^T\mX \mV^{(t)}]^s$, we have 
+        & \leq C^4 \frac{r^4}{\Delta_{\min}^4} (L^{(t)})^4 +  \frac{r^2 r^{4K} + p^2r^{2K+4}}{p^{2K}} \frac{(L^{(t)})^2}{\Delta_{\min}^4} \\
+        & \leq C^4 \frac{\bar C}{\tilde C^3} \of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}}, \label{eq:g1}
+    \end{align}
+    where the third inequality follows from the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:membership} and the last inequality follows from the assumption that $\Delta_{\min}^2 \geq \tilde C p^{-K/2}\log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
+
+    For $G_2$, noticing that $[\mX_{i:} \mV^{(t)}]^s = [ \mW_{z(i):}^T\mX \mV^{(t)}]^s$, we have 
     \begin{align}
@@
         & \leq \frac{2}{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2} \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
-        & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} \of{ \frac{r^2}{\Delta_{\min}^2} (L^{(t)})^2 + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} }\\
-        & \leq c \Delta_{\min}^2 L^{(t)},\label{eq:g2}
-    \end{align}
-    where the second inequality follows by Lemma~\ref{lem:norm_diff}, the third inequality follows by the inequality~\eqref{eq:cond2} in Condition~\ref{cond:origin}, and inequalities~\eqref{eq:inter5}, \eqref{eq:j11_wxvt} in the proof of Lemma~\ref{lem:intermediate},  and the last inequality follows by the assumption $\Delta_{\min}^2 \geq p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
+        & \leq C' \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} \of{ \frac{r^2}{\Delta_{\min}^2} (L^{(t)})^2 + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} }\\
+        & \leq \frac{C'}{\tilde C^2} \Delta_{\min}^2 L^{(t)},\label{eq:g2}
+    \end{align}
+    where $C'$ is a positive universal constant, the second inequality follows from Lemma~\ref{lem:norm_diff}, the third inequality follows from the inequality~\eqref{eq:cond2} in Condition~\ref{cond:origin}, the inequalities~\eqref{eq:inter5} and \eqref{eq:j11_wxvt} in the proof of Lemma~\ref{lem:intermediate},  and the last inequality follows from the assumption $\Delta_{\min}^2 \geq \tilde C p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
 
@@
         \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2 &\leq \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + 2\max_{a \in [r]}\onormSize{}{[\mW_{:a}^T\mX \mV^{(t)}]^s - [\mW_{:a}^T\mX \mV]^s }^2\\
-        & \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2}, \label{eq:g3xv_wxv}
-    \end{align}
-    where the last inequality follows by the inequality~\eqref{eq:j1} in the proof of Lemma~\ref{lem:intermediate}.
-    % \begin{align}
-    %     %& \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2},
-    % \end{align}
-   % where the last inequality follows by the inequality~\eqref{eq:inter4} in Lemma~\ref{lem:intermediate}. 
+        & \leq \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + C \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2}, \label{eq:g3xv_wxv}
+    \end{align}
+    where the last inequality follows from the inequality~\eqref{eq:j1} in the proof of Lemma~\ref{lem:intermediate} and $C$ is a positive constant.
    Then we have 
@@
         & \leq 2 \of{ \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  +  \onormSize{}{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  } \\
-        & \quad \cdot \max_{a \in [r]} \onorm{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
-        & \lesssim \of{\onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} } \of{ \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2}}+ c \Delta_{\min}^2 L^{(t)} \\
-        & \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 (\Delta_{\min}^2 + L^{(t)}) + c\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}},\label{eq:g3}
-    \end{align}
-    where the third inequality follows by the procedures to derive~\eqref{eq:g1} and \eqref{eq:g2}
-    %, the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:intermediate},
-    , and the last inequality follows by the assumption $\Delta_{\min}^2 \geq p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
-
-    Plugging the inequalities~\eqref{eq:g1}, \eqref{eq:g2}, and \eqref{eq:g3} into the upper bound~\eqref{eq:gib_decomp}, we finish the proof of inequality~\eqref{eq:gib}.
-
-    % According to the inequality~\eqref{eq:inter_wa_wb} in the proof of Lemma~\ref{lem:intermediate}, we have
-    % \begin{equation}
-    %     \onorm{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  \lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2.
-    % \end{equation}
-    % Then, we have 
-    % \begin{align}
-    %     |G_3|^2 &\lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 \max_{a \in [r]} \onorm{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2 + |G_2|^2 \\
-    %     &\lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 \of{ \frac{r^2}{\Delta_{\min}^2} \of{L^{(t)}}^2 + K  \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} } + |G_2|^2\\
-    %     & \leq \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2  c \of{ \Delta_{\min}^2 + K L^{(t)} }+|G_2|^2, \label{eq:g3}
-    % \end{align}
-    % where the second inequality follows by Lemma~\ref{lem:intermediate}, and the last inequality follows by the assumption on $\Delta_{\min}^2$.
-
-    % Therefore, combining inequalities~\eqref{eq:g1}, \eqref{eq:g2}, \eqref{eq:g3} and apply the inequality of arithmetic and geometric means, .
-
-    \item Upper bound for $H_{ib}^{(t)}$, inequality~\eqref{eq:hib}. Recall the definition of $H_{ib}$ and rearrange the terms
+        & \quad \times \max_{a \in [r]} \onorm{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
+        & \leq C^2 \of{\onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + C \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} } \of{ \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2}}+\frac{C'}{\tilde C^2} \Delta_{\min}^2 L^{(t)} \\
+        & \leq \frac{C^2 \bar C^2}{\tilde C} \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 (\Delta_{\min}^2 + L^{(t)}) +  \frac{C^3 C' \bar C^2}{\tilde C^2 }\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}},\label{eq:g3}
+    \end{align}
+    where the third inequality follows from the same procedure to derive~\eqref{eq:g1} and \eqref{eq:g2}, and the last inequality follows from the assumption $\Delta_{\min}^2 \geq \tilde C p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
+
+    Choose the $\tilde C$ such that
+    \begin{equation}\label{eq:tilde_c1}
+        3 \of{  C^4 \frac{\bar C}{\tilde C^3}  + \frac{C'}{\tilde C^2} + \frac{C^2 \bar C^2}{\tilde C} + \frac{C^3 C' \bar C^2}{\tilde C^2 }} \leq \frac{1}{512}.
+    \end{equation}
+    Then,  we finish the proof of inequality~\eqref{eq:gib} by plugging the inequalities~\eqref{eq:g1}, \eqref{eq:g2}, and \eqref{eq:g3} into the upper bound~\eqref{eq:gib_decomp}.
+
+    \item Upper bound for $H_{ib}^{(t)}$, i.e., the inequality~\eqref{eq:hib}. By definition of $H_{ib}$, we rearrange terms and obtain
     \begin{align}
@@
          \leq  \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} 
-        \lesssim  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2\label{eq:h1},
-    \end{align}
-    following the derivation of $G_2$ in inequality~\eqref{eq:g2} and the assumption that $\Delta_{\min}^2 \geq p^{-K/2} \log p$.
-    % \begin{align}
-    %     |H_1| &\leq 2 \max_{a \in [r]}  \frac{1}{ \onormSize{}{  \mW_{:a}^T \mX \mV^{(t)}}^2} \onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2  \\
-    %     & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}}\\
-    %     & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K} \Delta_{\min}^2}  \onorm{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
-    %     & \lesssim  \onorm{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 ,
-    % \end{align}
-    % where the first two inequalities follow the proof of $G_2$~\eqref{eq:g2} and the last inequality follows by the assumption on $\Delta_{\min}^2$.
+        \leq \tilde C^{-2} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2\label{eq:h1},
+    \end{align}
+    following the derivation of $G_2$ in inequality~\eqref{eq:g2} and the assumption that $\Delta_{\min}^2 \geq \tilde C p^{-K/2} \log p$.
 
@@
     \begin{equation}\label{eq:h2}
-        |H_2| \lesssim 2 \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
-    \end{equation}
-    where the last inequality follows by the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
-
-    % For $H_2$,  sccording to the inequality~\eqref{eq:inter_wa_wb} in the proof of Lemma~\ref{lem:intermediate}, we have
-    % \begin{equation}
-    %     \onorm{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  \lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
-    % \end{equation}
-    % and thus 
-    % \begin{equation}\label{eq:h2}
-    %     |H_2| \lesssim  \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2.
-    % \end{equation}
+        |H_2| \lesssim 2\max_{a \in [r]}\onormSize{}{[\mW_{:a}^T\mX \mV^{(t)}]^s - [\mW_{:a}^T\mX \mV]^s }^2 \lesssim \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} \leq C \frac{\bar C^2}{ \tilde C^2} \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
+    \end{equation}
+    where the last inequality follows from the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
 
@@
     \begin{equation}\label{eq:h3}
-         |H_3| \lesssim \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
+         |H_3| \lesssim \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \leq 2 \tilde C^{-1} \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
     \end{equation}
@@
 
-    % For $H_3$, by Cauchy-Schwartz inequality, we have 
-    % \begin{align}
-    %     |H_3| &\lesssim    \onorm{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \\
-    %     &\lesssim  \frac{ r^{K-1/2} + K p^{1/2} r^{K/2+1/2} }{p^{K/2}} \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }\\
-    %     & \lesssim \Delta_{\min} \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }\\
-    %     & \lesssim \onorm{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2, \label{eq:h3}
-    % \end{align}
-    % where the third inequality follows by the assumption on $\Delta_{\min}^2$.
-
-    Therefore, combining inequalities~\eqref{eq:h1}, \eqref{eq:h2}, and \eqref{eq:h3}, we have finish the proof of inequality~\eqref{eq:hib}.
+    Choose $\tilde C$ such that 
+    \begin{equation}\label{eq:tilde_c2}
+        \tilde C^{-2} + C \frac{\bar C^2}{ \tilde C^2} + \tilde C^{-1} \leq \frac{1}{4}.
+    \end{equation}
+     Therefore, we finish the proof of inequality~\eqref{eq:hib} combining inequalities~\eqref{eq:h1}, \eqref{eq:h2}, and \eqref{eq:h3}.
 
@@
 
-     \begin{lem}[Relationship between misclustering loss and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin}, we have
+     \begin{lem}[Relationship between misclustering loss and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement}, we have
     \begin{equation}\label{eq:inter1}
@@
         \onormSize{}{\mE (\mV - \mV^{(t)})}_{\sigma} \lesssim \sqrt{ \frac{r^{K-1}(p r^{K-1} + p r)}{p^{K-1}}} 
- \frac{r}{\Delta_{\min}^2} L^{(t)}.    
+ \frac{r}{\Delta_{\min}^2} L^{(t)},
  \end{equation}
  \begin{equation}\label{eq:inter3}
-     \max_{b \in [r]} \onormSize{}{ [ \mW_{:b}^T \mY \mV ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV  ]^s } \lesssim \frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{r^{2K} + pr^{K+1}}{p^{K}} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}
+     \max_{b \in [r]} \onormSize{}{ [ \mW_{:b}^T \mY \mV ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV  ]^s } \leq C \of{\frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{r^{2K} + pr^{K+1}}{p^{K}} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}},
  \end{equation}
  \begin{equation}\label{eq:inter4}
-     \max_{b \in [r]} \onormSize{}{  [  \mW_{:b}^{(t), T}  \mY \mV]^s - [  \mW_{:b}^{(t), T}  \mY \mV^{(t)}]^s } \lesssim  \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} +  \frac{r L^{(t)}}{\Delta_{\min}}
+     \max_{b \in [r]} \onormSize{}{  [  \mW_{:b}^{(t), T}  \mY \mV]^s - [  \mW_{:b}^{(t), T}  \mY \mV^{(t)}]^s } \leq C  \of{ \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} +  \frac{r L^{(t)}}{\Delta_{\min}}},
      \end{equation}
     \begin{equation}\label{eq:inter5}
-         \max_{b \in [r]} \onormSize{}{ [\mW_{:b}^T \mY \mV^{(t)} ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s } \lesssim \frac{rL^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
-    \end{equation}
-
-    In addition, the inequality~\eqref{eq:inter4} also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$.
+         \max_{b \in [r]} \onormSize{}{ [\mW_{:b}^T \mY \mV^{(t)} ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s } \leq C \of{\frac{rL^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}},
+    \end{equation}
+    for some positive universal constant $C$. In addition, the inequality~\eqref{eq:inter4} also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$.
     \end{lem}
@@
     \begin{proof}[Proof of Lemma~\ref{lem:intermediate}] We follow and use several intermediate conclusions in \citet[Proof of Lemma 5]{han2020exact}. We prove each inequality separately.
-    \begin{enumerate}
+    \begin{enumerate}[wide]
     \item Inequality~\eqref{eq:inter1}. By \citet[Proof of Lemma 5]{han2020exact}, we have 
@@
 
-        %\item Inequalities~\eqref{eq:inter1} and \eqref{eq:inter2}.
-We follow the \citet[Proof of Lemma 5]{han2020exact} except upper bounding their misclassification error ``$h_k^{(t)}$" by $L^{(t)}/\Delta_{\min}^2$ rather than ``$ l_k^{(t)}/\Delta_k^2$".
-        %defined in their context. 
-        %The technique that replacing ``$ l_k^{(t)}/\Delta_k^2$" by $L^{(t)}/\Delta_{\min}^2$ is also used in the proofs for other inequalities. 
-    \item Inequality~\eqref{eq:inter3}. We upper bound the desired quantity by triangle inequality as
+    \item Inequality~\eqref{eq:inter3}. We upper bound the desired quantity by triangle inequality,
     \begin{equation}
@@
 
-    For $I_1$, we further bound $I_1$ by triangle inequality as
-    %and Lemma~\ref{lem:norm_diff} as
+    For $I_1$, we further bound $I_1$ by triangle inequality,
     \begin{equation}
@@
     where
-    % \begin{equation}
-    %     I_{11} = \onorm{ \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }}   } \quad \text{and} \quad I_{12} = \frac{1}{\onormSize{}{\mW_{:b}^T \mX \mV}}\onorm{ \of{\mW_{:b}^T - \mW_{:b}^{(t), T}} \mE \mV  }.
-    % \end{equation}
-     \begin{equation}
-        I_{11} = \onorm{ \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }}   } \quad \text{and} \quad I_{12} = \onorm{ \frac{\mW_{:b}^T \mE \mV}{ \onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mE \mV}{ \onormSize{}{  \mW_{:b}^{(t),T} \mX \mV }} }.
-    \end{equation}
-    Consider $I_{11}$. We first define the confusion matrix $\mD = \mM^T \mTheta^T \mW^{(t)} = \entry{D_{ab}} \in \bbR^{r \times r}$ where 
+ \begin{equation}
+        I_{11} = \onorm{ \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }}   }, \quad \text{and} \quad I_{12} = \onorm{ \frac{\mW_{:b}^T \mE \mV}{ \onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mE \mV}{ \onormSize{}{  \mW_{:b}^{(t),T} \mX \mV }} }.
+    \end{equation}
+    We first consider $I_{11}$. Define the confusion matrix $\mD = \mM^T \mTheta^T \mW^{(t)} = \entry{D_{ab}} \in \bbR^{r \times r}$ where 
     \begin{equation}
@@
     \end{equation}
-    and for all $b \in [r]$
+    and for all $b \in [r]$,
     \begin{equation}\label{eq:dbb_bound}
@@
     \begin{equation}
-        \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} = \off{\mS_{b:}}^s, \quad \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }} = [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}]^s,
-    \end{equation}
-    % \begin{equation}
-    %     D_a = \frac{\sum_{i \in [p]}  \theta(i) \ind \offf{ z(i) = a, z^{(t)}(i) = b } }{\sum_{i \in [p]} \theta(i)  \ind\offf{  z^{(t)}(i) = b }}, \text{  for all } a \neq b, \quad D = \sum_{a \neq b, a \in [r]} D_a.
-    % \end{equation}
+        \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} = \off{\mS_{b:}}^s, \quad \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }} = [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}]^s.
+    \end{equation}
 
@@
     \end{align}
-    where the first inequality follows by triangle inequality, and the last inequality follows by Lemma~\ref{lem:norm_diff}.
-Also, note that with bounds \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}, when $p$ is large enough we have
+    where the first inequality follows from the triangle inequality, and the last inequality follows from Lemma~\ref{lem:norm_diff}.
+Note that with bounds \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}, when $p$ is large enough, we have
     \begin{equation}\label{eq:i11_sin_dom}
@@
         & = \frac{  \onormSize{}{ [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}] \sin \alpha}}{  \onormSize{}{ D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}} } \\
-        & \leq \frac{ 1}{C_1}  \sum_{a \neq b, a \in [r]} D_{ab} \onorm{\mS_{a:}} \onorm{ \mS^s_{b:} - \mS^s_{a:}} \\
+        & \leq \frac{ 1}{ C_1}  \sum_{a \neq b, a \in [r]} D_{ab} \onorm{\mS_{a:}} \onorm{ \mS^s_{b:} - \mS^s_{a:}} \\
         & \lesssim \frac{r}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind\offf{z^{(t)}(i) = b} \onorm{ \mS^s_{b:} - \mS^s_{a:}}\\
@@
     \end{align}
-    where the second inequality follows by~\eqref{eq:i11_sin_num} and \eqref{eq:i11_sin_dom}, and the last two inequalities follow by the definition of $D_a, L^{(t)}$ and the constraint of $\onorm{\mS_{b:}}$ in parameter space~\eqref{eq:family}.
-
-   % \fixme{Jiaxin}{Need to control $\frac{1}{p}\sum_{i \in [p]} \theta(i) \ind\{ z(i) \neq z^{(t)}(i)\} \leq \log^{-1} p$.
-Actually can be implied from $L^{(t)}$.
-Note that $\frac{1}{p} \sum \theta(i) \leq L^{(t)}/\Delta_{\min}^2$.}
-    % Following the assumption in parameter space~\eqref{eq:family} that $c_3 \leq \onorm{\mS_{a:}} \leq c_4$ for all $a \in [r]$, we have 
-    % \begin{equation}\label{eq:a11_norm}
-    %     \onorm{(1 - \eta) \mS_{b:} + \sum_{a \neq b} \eta_a \mS_{a:}} \leq (1 - \eta) \onorm{\mS_{b:}} + \sum_{a \neq b, a\in [r]} \eta_a \onorm{\mS_{a:}} \leq \of{ 1 - \eta + \eta c_4 /c_3 } \onorm{\mS_{b:}},
-    % \end{equation}
-    % %\fixme{Jiaxin}{Definition of $\eta$ should include $\theta$. correct this in the following proof}
-    % where the first inequality uses triangle inequality, and 
-    % \begin{equation}\label{eq:a11_ang}
-    %     \ang{ \mS_{b:}, \mS_{a:} } \geq - \onorm{\mS_{b:}} \onorm{\mS_{a:}} \geq -c_4/c_3 \onorm{\mS_{b:}}^2,
-    % \end{equation}
-    % where the first inequality uses the fact that $ \cos \of{\mS_{b:}, \mS_{a:} } \geq -1$.
-Let $\alpha \in [0, \pi]$ denote the angle between $\mS_{b:}$ and $ (1 - \eta) \mS_{b:} + \sum_{a \neq b} \eta_a \mS_{a:}$, we have
-    % \begin{align}
-    %     \cos \alpha&= \frac{ \ang{ \mS_{b:},  (1 - \eta) \mS_{b:} + \sum_{a \neq b} \eta_a \mS_{a:}} }{ \onorm{ \mS_{b:} } \onorm{ (1 - \eta) \mS_{b:} + \sum_{a \neq b} \eta_a \mS_{a:} }}\\
-    %     &\geq \frac{ (1- \eta) \onorm{\mS_{b:}}^2 + \sum_{a \neq b} \eta_a \ang{\mS_{b:}, \mS_{a:}} }{ \of{ 1 - \eta + \eta c_4 /c_3  } \onorm{\mS_{b:}}^2 }\\
-    %     & \geq \frac{ (1- \eta - \eta c_4/c_3 )  \onorm{\mS_{b:}}^2  }{ \of{1 - \eta + \eta c_4 /c_3   } \onorm{\mS_{b:}}^2} \\
-    %     &\geq 1 - 2 \frac{c_4}{c_3} \eta.\label{eq:cos_sb_sbmix}
-    % \end{align}
-
-    % Note that  $\sum_{i \in [p]} \ind\offf{  z^{(t)}(i) = b } \asymp \frac{p}{r}$ by Lemma~\ref{lem:membership}. Then 
-    % \begin{equation}\label{eq:eta}
-    %     \eta = \frac{ \sum_{z^{(t)}(i) = b, z(i) \neq b }\theta(i)}{\sum_{z^{(t)}(i) = b } \theta(i)}  \lesssim \frac{r}{c_0 p} \sum_{i: z^{(t)}(i) \neq z(i) } \theta(i) \lesssim \frac{1}{\log p}
-    % \end{equation}
-    % where the first inequality follows by the assumption that $\min_{i \in [p]}\theta(i) \geq c_0$, the second and third inequality follows by the inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin} and the assumption on $\Delta_{\min}^2$.
-
-    % % \begin{align}
-    % %     \eta &= \frac{ \sum_{z^{(t)}(i) = b, z(i) \neq b }\theta(i)  \sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b , z(i) \neq b} }{ \sum_{z^{(t)}(i) = b } \theta(i) \sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } } \\
-    % %     & \leq \frac{ \sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b , z(i) \neq b} }{ \sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } } \\
-    % %     &\lesssim \frac{r}{p} \sum_{i \in [p]} \ind \offf{ z(i) \neq z^{(t)}(i)} \lesssim \ell^{(t)}.
-    % % \end{align}
-
-    % % Also, by the bounded loss condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}, we have 
-    % % \begin{equation}
-    % %     \eta \lesssim \frac{L^{(t)}}{ \Delta_{\min}^2} = \frac{1}{r \log p} \leq \frac{c_3\sqrt{2}}{4c_4},
-    % % \end{equation}
-    % % where the last inequality holds when $p$ is large enough and thereof $\alpha \leq \frac{\pi}{4}$.
-
-    % Then, when $p$ is large enough, we have $\eta \leq \frac{c_3\sqrt{2}}{4c_4}$ and $\alpha \leq \frac{\pi}{4}$. Note that by the property of trigonometric functions, we have
-    % \begin{equation}\label{eq:trig}
-    %      1- \cos \alpha = 2 \sin^2 \frac{\alpha}{2}, \quad \sin \frac{\alpha}{2} \leq \sin \alpha \leq \tan \alpha,
-    % \end{equation}
-    % and with triangle inequality
-    % \begin{equation}\label{eq:sin}
-    %     \onorm{ \off{(1 - \eta) \mS_{b:} + \sum_{a \neq b, a\in [r]} \eta_a \mS_{a:}} \sin \alpha } \leq \sum_{a \neq b, a\in [r]} \eta_a \onorm{\mS_{a:}} \sin \of{ \mS_{a:}, \mS_{b:} },  
-    % \end{equation}
-    % and 
-    % \begin{equation}\label{eq:cos}
-    %     \onorm{ \off{(1 - \eta) \mS_{b:} + \sum_{a \neq b, a\in [r]} \eta_a \mS_{a:}} \cos \alpha } \leq (1-\eta) \onorm{\mS_{b:}}.  
-    % \end{equation}
-
-    % Hence, we have the upper for $I_{11}$
-    % \begin{align}
-    %     I_{11} &\stackrel{\text{def }}{=}  2 \sin \frac{\alpha}{2}\\
-    %     &\stackrel{\text{\eqref{eq:trig}}}{\leq}  2 \tan \alpha \\
-    %     &\stackrel{\text{\eqref{eq:sin}, \eqref{eq:cos}}}{\leq} 2 \frac{ \sum_{a \neq b, a \in [r]} \eta_a \onorm{\mS_{a:}} \sin\of{ \mS_{a:}, \mS_{b:} } }{ (1- \eta) \onorm{\mS_{b:} } }\\
-    %   & \stackrel{\text{\eqref{eq:family}, Lemma~\ref{lem:norm_diff}}}{\lesssim} \sum_{a \neq b, a \in [r]}\eta_a \onorm{ \off{\mS_{b:}}^s - \off{ \mS_{a;} }^s }\\
-    %     &\stackrel{\text{ Lemma~\ref{lem:membership}}}{\lesssim} \frac{r}{p}  \sum_{i \in [p]} \theta(i) \sum_{a \neq b, a \in [r]} \ind \offf{ z(i) = b, z^{(i)} = a }  \onorm{ \off{\mS_{b:}}^s - \off{ \mS_{a:} }^s } \label{eq:i11_p} \\
-    %     & \leq \frac{r L^{(t)}}{\Delta_{\min}}. \label{eq:i11}
-    % \end{align}
-
-    % Thus, we have $\sin \frac{\alpha}{2} \leq \sin \alpha \leq \tan \alpha$ by the property of trigonometric functions. We have 
-    % \begin{align}
-    %     I_{11}  & \leq 2 \tan \alpha\\
-    %     & \leq 2 \frac{ \sum_{a \neq b, a \in [r]} \eta_a \onorm{\mS_{a:}} \sin\of{ \mS_{a:}, \mS_{b:} } }{ (1- \eta) \onorm{\mS_{b:} } }\\
-    %     & \lesssim \sum_{a \neq b} \eta_a \onorm{ \off{\mS_{b:}}^s - \off{ \mS_{a;} }^s }\\
-    %     & \lesssim \frac{r}{p} \theta(i) \sum_{i \in [p]} \sum_{a \neq b, a \in [r]} \ind \offf{ z(i) = b, z^{(i)} = a }  \onorm{ \off{\mS_{b:}}^s - \off{ \mS_{a:} }^s } \label{eq:i11_p} \\ 
-    %     & \leq L^{(t)}/\Delta_{\min}, 
-    % \end{align}
-    % where the second inequality follows by the definition of $\tan(\cdot)$ and triangle inequality, the third inequality follows by the Lemma~\ref{lem:norm_diff}, the third inequality follows by the fact hat $\eta_a \leq \frac{ \sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b , z(i) = a} }{ \sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } } $, the last two inequalities follows by the definition of $L^{(t)}$.
-
-    Consider $I_{12}$. By triangle inequality, we have 
+    where the second inequality follows from~\eqref{eq:i11_sin_num} and \eqref{eq:i11_sin_dom}, and the last two inequalities follow by the definition of $D_a$ and $L^{(t)}$, and the constraint of $\onorm{\mS_{b:}}$ in parameter space~\eqref{eq:family}.
+
+    We now consider $I_{12}$. By triangle inequality, we have 
     \begin{equation}
@@
     \end{equation}
-    where the second inequality follows by  \citet[Inequality (121), Proof of Lemma 5]{han2020exact} and the last inequality follows by Lemma~\ref{lem:singular_thetam} and \eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
-    % Also, by the definition of $\mW^{(t)}$ and Condition~\ref{cond:origin}, we have
-    % \begin{equation}
-    %     \onormSize{}{ \mW_{:b}^{(t),T} \mE \mV} \leq \onormSize{}{\mW_{:b}^{(t),T}} \onorm{\mE \mV}_{\sigma} \lesssim \sqrt{\frac{r^{K-1}}{p^K}} (\sqrt{p} + \sqrt{r^{K-1}}).
-    % \end{equation}
+    where the second inequality follows from~\citet[Inequality (121), Proof of Lemma 5]{han2020exact} and the last inequality follows from Lemma~\ref{lem:singular_thetam} and \eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
      Note that $\onorm{\mW_{:b}^T \mX \mV} = \onorm{\mS_{b:}} \geq c_3$ and $\onormSize{}{  \mW_{:b}^{(t),T} \mX \mV } \geq C_1$ by inequality~\eqref{eq:i11_sin_dom}. Therefore, we have 
@@
      \end{align}
-     where second inequality follows the inequalities~\eqref{eq:i12_wwe}, \eqref{eq:i12_wwx}, and \eqref{eq:cond1} in Condition~\ref{cond:origin}.
-
-    Hence, combining inequalities~\eqref{eq:i11} and~\eqref{eq:i12},  we have 
+     where second inequality follows from the inequalities~\eqref{eq:i12_wwe}, \eqref{eq:i12_wwx}, and \eqref{eq:cond1} in Condition~\ref{cond:origin}.
+
+    Hence, combining inequalities~\eqref{eq:i11} and~\eqref{eq:i12} yields  
     \begin{equation}\label{eq:i1}
@@
     \end{equation}
-    % Therefore, we complete the proof of inequality~\eqref{eq:inter3}.
-
-    % \begin{align}
-    %      \aabs{ \frac{1}{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{1}{\onormSize{}{  \mW_{:b}^{(t),T} \mX \mV }} } \leq \frac{\onormSize{}{ \mW_{:b}^T \mX \mV} + \onormSize{}{ \mW_{:b}^{(t),T} \mX \mV} }{\onormSize{}{  \mW_{:b}^T \mX \mV }\onormSize{}{  \mW_{:b}^{(t),T} \mX \mV }}
-    % \end{align}
-
-    % Note that $\onorm{\mS_{b:}} = \onorm{\mW_{:b}^T \mX \mV}$. We have
-    % \begin{equation}\label{eq:i12}
-    %      I_{12} \lesssim  \onorm{ \of{\mW_{:b}^T - \mW_{:b}^{(t), T}} \mE \mV} \lesssim \sqrt{\frac{r^{2K} +p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
-    % \end{equation}
-    % where the first inequality follows by the constraint~\eqref{eq:family}, and the second inequality follows by \citet[Equation (121), Proof of Lemma 5]{han2020exact}.
 
@@
      \end{equation}
-
-    % we have 
-    % \begin{equation}\label{eq:i2}
-    %     I_2 \leq \frac{ \onorm{ \mW_{:b}^T \mE \mV } }{ \onorm{\mW_{:b}^T \mX \mV}  } \lesssim  \onorm{ \mW_{:b}^T \mE \mV }  \lesssim \frac{r^K}{p^{K/2}},
-    % \end{equation}
-    % where the first inequality follows by triangle inequality, the second inequality follows~\eqref{eq:family}, and the last inequality follows by Condition~\ref{cond:origin}.
-
-    % For $I_3$, note that 
-    % \begin{align}
-    %     \onorm{\mW_{:b}^{(t),T} \mX \mV} = \onorm{ (1 - \eta) \mS_{b:} + \sum_{a \neq b, a \in [r]} \eta_a \mS_{a:}} \geq (1 - 2\eta) c_3 \gtrsim 1,
-    % \end{align}
-    % where the second inequality holds by~\eqref{eq:eta} when $p$ is large enough. Then, we have 
-    % \begin{equation}\label{eq:i3}
-    %     I_3 \leq \frac{ \onorm{ \mW_{:b}^{(t),T} \mE \mV } }{ \onorm{\mW_{:b}^{(t),T} \mX \mV}  }  \lesssim \onorm{ \mW_{:b}^{(t),T} \mE \mV } \lesssim \frac{r^K}{p^{K/2}},
-    % \end{equation}
-    % where the last inequality follows by Lemma~\ref{lem:membership} and Condition~\ref{cond:origin}.
-
        Therefore, combining the inequalities~\eqref{eq:i1}, \eqref{eq:i2}, and \eqref{eq:i3}, we finish the proof of inequality~\eqref{eq:inter3}.
@@
     \item Inequality~\eqref{eq:inter4}.
-Here we only show the detailed proof of inequality~\eqref{eq:inter4} with $\mW_{:b}^{(t)}$.
+Here we only show the proof of inequality~\eqref{eq:inter4} with $\mW_{:b}^{(t)}$.
 The proof also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$, and we omit the repeated procedures.
 
-    We upper bound the desired quantity by triangle inequality as
+    We upper bound the desired quantity by triangle inequality
     \begin{equation}
@@
 
-    For $J_1$, by triangle inequality 
-    %and Lemma~\ref{lem:norm_diff}
-    , we have 
+    For $J_1$, by triangle inequality,
+we have 
     \begin{equation}
@@
 
-    % \fixme{Jiaxin}{Start $J_{11}$}
-
-    Consider $J_{11}$.
+    We first consider $J_{11}$.
 Define the matrix $\mV^{k} \coloneqq \mW^{\otimes (k-1)} \otimes \mW^{(t), \otimes (K-k)}$ for $k = 2,\ldots, K-1$, and denote $\mV^1 = \mV^{(t)}, \mV^K = \mV$.
-Then, define the quantity
+Also, define the quantity
     \begin{equation}
@@
     Recall the definition of confusion matrix $\mD = \mM^T \mTheta^T \mW^{(t)} = \entry{D_{ab}} \in \bbR^{r \times r}$.
-    %, and let  $\mZ^k =  \mW^{\otimes (k-1)} \otimes \mW^{(t),\otimes (K-k-1)} \in \bbR^{ p^{K-2} \times r^{K-2} }$.  
  We have 
@@
          \ang{\mW_{:b}^{(t),T} \mX \mV^k , \mW_{:b}^{(t),T} \mX \mV^{k+1}} 
-         %&= \ang{ \mat_1(\tT_k), \mat_1(\tT_{k+1}) } \\
          &  =\ang{ \mat_{k+1}(\tT_k), \mat_{k+1}(\tT_{k+1}) } \label{eq:j11_tensormat} \\
-         %& = \ang{ \mW^{(t),T} \mX \mW_{:b}^{(t)} \otimes \mZ^k , \mW^T \mX \mW_{:b}^{(t)} \otimes \mZ^k} \\
-         %& = \ang{ \mD^T \mS (\mTheta \mM)^{\otimes (K-1), T}\mW_{:b}^{(t)} \otimes \mZ^k,  \mS (\mTheta \mM)^{\otimes (K-1), T}\mW_{:b}^{(t)} \otimes \mZ^k   }\\
   & =  \ang{ \mD^T \mS \mZ^k ,  \mS \mZ^k   }\\
@@
     \end{align}
-    where $\mZ^k = \mD_{:b} \otimes \mI_r^{\otimes (k-1)} \otimes  \mD^{\otimes (K-k-1)}$, the equations follow by the tensor algebra and the definitions, and the last inequality follows by the bounds of $\mD$ in \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}.
+    where $\mZ^k = \mD_{:b} \otimes \mI_r^{\otimes (k-1)} \otimes  \mD^{\otimes (K-k-1)}$, the equations follow by the tensor algebra and definitions, and the last inequality follows from the bounds of $\mD$ in \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}.
 
@@
     \end{equation}
-    where the second inequality follows by inequality~\eqref{eq:dab_bound}, and the fact that for all $b \in [r]$ 
+    where the second inequality follows from inequality~\eqref{eq:dab_bound}, and the fact that for all $b \in [r]$,
     \begin{equation}
@@
 
-    On one hand, recall Definition~\ref{def:stable} that for any cluster assignment $\bar z$
+    To bound $\sin \beta_{k1}$, recall Definition~\ref{def:stable} that for any cluster assignment $\bar z$ in the $\varepsilon$-neighborhood of true $z$,
     \begin{equation}
@@
     \end{equation}
-    Note that we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \lesssim \log^{-1}(p) $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}. With stability assumption, we have 
+    Note that we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \lesssim \log^{-1}(p) $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.
+With the locally linear stability assumption, we have
     \begin{equation}
@@
     \end{equation}
-    Note that $\text{diag}(\mp(z^{(t)})) \mD = \text{diag}(\mp_{\mtheta}(z^{(t)})) \tilde \mD$.
-Also, note that $\sin(\ma,\mb) =\min_{c \in \bbR} \frac{ \onormSize{}{\ma - c\mb} }{\onormSize{}{\ma}}$ for vectors $\ma,\mb$ of same dimension, and let $c_0 = \argmin_{c \in \bbR} \frac{\onormSize{}{\mp(z^{(t)}) - c\mp_{\mtheta}(z^{(t)})}}{\onormSize{}{\mp(z^{(t)})}}$.
+    Note that $\text{diag}(\mp(z^{(t)})) \mD = \text{diag}(\mp_{\mtheta}(z^{(t)})) \tilde \mD$, and $\sin(\ma,\mb) =\min_{c \in \bbR} \frac{ \onormSize{}{\ma - c\mb} }{\onormSize{}{\ma}}$ for vectors $\ma,\mb$ of same dimension.
+Let $c_0 = \argmin_{c \in \bbR} \frac{\onormSize{}{\mp(z^{(t)}) - c\mp_{\mtheta}(z^{(t)})}}{\onormSize{}{\mp(z^{(t)})}}$.
 Then, we have
@@
     \end{align}
-    where the last inequality follows Lemma~\ref{lem:membership} and constraint $\min_{i \in [p]}\theta(i) \geq c$ and thereof $\onormSize{}{ \mp(z^{(t)})} \lesssim p$ and $\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  \gtrsim p$.
+    where the last inequality follows from Lemma~\ref{lem:membership}, the constraint $\min_{i \in [p]}\theta(i) \geq c>0$, $\onormSize{}{ \mp(z^{(t)})} \lesssim p$ and $\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  \gtrsim p$.
 By the geometry property of trigonometric function, we have
-    \begin{align}
+    \begin{align}\label{eq:b1}
         \sin \beta_{k1} &= \min_{c \in \bbR} \frac{\onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes (\mD - c \tilde \mD) \otimes \mD^{\otimes K-k - 1} } }{ \onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k}} }\\
@@
     \end{align}
-    where the second inequality follows by the singular property of $\mD$ in \eqref{eq:singular_Du}, \eqref{eq:singular_Dl} and the constraint of $\mS$ in \eqref{eq:family}.
-
-    On the other hand, let $\mC = \text{diag}(\{\onormSize{}{\mS_{a:}}\}_{a \in [r]})$. We have 
-    \begin{align}
+    where the second inequality follows from the singular property of $\mD$ in \eqref{eq:singular_Du}, \eqref{eq:singular_Dl} and the constraint of $\mS$ in \eqref{eq:family}.
+
+    To bound $\sin \beta_{k2}$, let $\mC = \text{diag}(\{\onormSize{}{\mS_{a:}}\}_{a \in [r]})$. We have 
+    \begin{align}\label{eq:b2}
         \sin \beta_{k2} &\lesssim \frac{ \onorm{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes (\mI_r - \tilde \mD) \otimes \mD^{\otimes K-k-1} } }{\onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k} \otimes \mD^{\otimes K-k-1} }  }\\
@@
     \end{align}
-    where the third inequality follows by the singular property of $\mD$ and the boundedness of $\mS$, and the fourth inequality follows by the definition of $\tilde \mD$, boundedness of $\mS$, lower bound of $\mtheta$, and the singular property of $\mZ^k$ in inequality~\eqref{eq:singular_z}.
-
-    Therefore, we have shown the 
+    where the third inequality follows from the singular property of $\mD$ and the boundedness of $\mS$, and the fourth inequality follows from the definition of $\tilde \mD$, boundedness of $\mS$, the lower bound of $\mtheta$, and the singular property of $\mZ^k$ in inequality~\eqref{eq:singular_z}.
+
+  Combining~\eqref{eq:b1} and~\eqref{eq:b2} yields
     \begin{equation}
@@
     \end{equation}
-
-%     \fixme{Jiaxin}{Proof for $J_{11}$ starts here. Blue color is the latest proof.}
-
-%     {\color{blue}
-%     Consider $\mA \mx, \mB \mx$ where $\mB = \mD^T \mA$ and let $\alpha$ denote the angle between $\mA \mx, \mB \mx$.
-Specifically, $\mA = (\mV\mX)^T = \mI_r \mS^T$ and $\mB = \mD^T \mS^T$.
-Our goal is to upper bound $\sin \alpha$.
-Let $\tilde \mD$ is the normalized confusion matrix with $\tilde \mD_{ab} = \frac{\sum_{i \in [p]} \theta(i) \ind\{z^{(t)} = b, z(i) = a\}}{\sum_{i \in [p] }\theta(i) \ind\{z^{(t)} = b\}}$.
-Then we have
-%     \begin{equation}
-%         \sin \alpha \leq \sin (\mA \mx, \tilde \mB\mx) +  \sin (\tilde \mB \mx,  \mB\mx),
-%     \end{equation}
-%     where $\tilde \mB  = \tilde \mD^T \mA$. For the first term, we have 
-%     \begin{align}
-%         \sin (\mA \mx, \tilde \mB\mx) \leq  \frac{ \onorm{\mA \mx - \tilde \mB \mx } }{\onorm{\mA \mx}} \lesssim \onorm{(\mI - \tilde \mD^T) \mS^s}_F \lesssim \frac{1}{c_0 p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \onorm{\mS_{b:}^s -\mS_{z(i):}^s} \ind\{z^{(t)}(i) = b\} = \frac{L^{(t)}}{\Delta_{\min}}.
-%     \end{align}
-%     For the second term, we have 
-%     \begin{equation}
-%         \sin (\tilde \mB \mx,  \mB\mx) \lesssim \onorm{\tilde \mB - \mB}_F \lesssim \max_{b \in [r]} \aabs{ 1 -  \frac{\sum_{i \in [p]} \theta(i)\ind \offf{ z^{(t)}(i) = b } }{\sum_{i \in [p]} \ind \offf{ z^{(t)}(i) = b } }} \lesssim \frac{L^{(t)}}{\Delta_{\min}^2}.
-%     \end{equation}
-
-% }
-
-% \fixme{Miaoyan}{Merge the results of Lemma 14 into your proof. No need to write Lemma 14 as a stand-alone lemma.}
-%     \begin{lem}[Angle Stability]\label{lem:stability} Assume the local linear stability of degree holds in the neighborhood $\tN(z,\log^{-1}p)$. Let $\tilde \mB$ be.... Then, we have
-%     \[
-%     \sin(\tilde \mB \mx, \mB \mx)\lesssim {L^{(t)}\over \Delta_{\min}}.
-%     \]
-%     \end{lem}
-
-%     \begin{proof}[Proof of Lemma~\ref{lem:stability}]
-%     %By assumption, there exists a constant $\eta=\eta(\bar z)\in\bbR$ ($\bbR_{+}$?) such that $
-%   %\eta\mp_{\mtheta}(\bar z)= \mp(\bar z)$. Therefore, by the definition of angles, we have
-%   % \[
-%     %\sin(\tilde \mB \mx, \mB \mx)\leq \onormSize{}{\tilde \mB-\eta\mB}_F=0.
-%     %\]
-
-%     Note that 
-%     \begin{equation}
-%         \min_{c \in \bbR} \frac{ \onormSize{}{\mp(z^{(t)}) - c \mp_{\mtheta}(z^{(t)})}}{p} \asymp \sin (\mp(z^{(t)}) , \mp_{\mtheta}(z^{(t)})) \lesssim \varepsilon \Delta_{\min} \lesssim \frac{L^{(t)}}{\Delta_{\min}}.
-%     \end{equation}
-
-%     Note that $\text{diag}(\mp_{\mtheta}(\bar z)) \tilde \mD = \text{diag}(\mp(\bar z))  \mD$. Let $c_0 = \argmin_{c \in \bbR} \onormSize{}{\mp(z^{(t)}) - c \mp_{\mtheta}(z^{(t)})}$. We have
-%     \begin{align}
-%          \sin(\tilde \mB \mx, \mB \mx) & \leq  \frac{ \onormSize{}{\mD - c_0\tilde \mD}_F \onorm{\mA \mx} }{\lambda_r(\mD) \onorm{\mA \mx}}\\
-%          &\leq  \frac{\onorm{\mI - c_0\text{diag}(\mp(\bar z)) \text{diag}(\mp_{\mtheta}(\bar z))^{-1}}_F \onorm{\mD}_F \onorm{\mA \mx}}{ \lambda_r(\mD) \onorm{\mA \mx} }\\
-%          & \leq \onorm{ \text{diag} \of{ \frac{ \onormSize{}{\mtheta_{z^{(t), -1}(a)}}_1 - c_0 |z^{(t), -1}(a)| }{ \onormSize{}{\mtheta_{z^{(t), -1}(a)}}_1 }  }  }_F \\
-%          & \leq \frac{1}{p}  \onormSize{}{\mp(z^{(t)}) - c_0 \mp_{\mtheta}(z^{(t)})}\\
-%          & \leq \frac{1}{p} \sin (\mp(z^{(t)}) , \mp_{\mtheta}(z^{(t)}))\\
-%          & \lesssim \frac{L^{(t)}}{\Delta_{\min}},
-%     \end{align}
-%     where the third inequality follows by the singular property of $\mD$ in inequalities~\eqref{eq:singular_Dl} and \eqref{eq:singular_Du}.
-%     \end{proof}
-
-    % \fixme{Jiaxin}{concatenation may work}
-
-    % { \color{blue}
-
-    % \begin{lem}[Concatenation] \label{lem:concatenation}
-    % Consider two vectors $\mx = [\mx_1,\ldots, \mx_r]$ and $\mx' = [\mx'_1, \ldots, \mx'_r]$.
-Suppose $\onorm{\mx_a} \asymp \onorm{\mx'_a} \asymp 1$ for all $ a\in [r]$.
-Let $\alpha$ denote the angle between $\mx$ and $\mx'$, and $\alpha_a$ denote the angle between $\mx_a$ and $\mx'_a$.
-Then, we have
-    % \begin{equation}
-    %     %\min_{a \in [r]} (1 - \cos \alpha_a)  \lesssim 
-    %     \max_{a \in [r]} (1 - \cos \alpha_a) \lesssim  (1 - \cos\alpha ) \lesssim 1 - \min_{a \in [r]} \cos \alpha_a,
-    % \end{equation}
-
-    % \end{lem}
-
-    % \begin{proof}[Proof of Lemma~\ref{lem:concatenation}]
-    % For the first inequality, note that 
-    % \begin{align}
-    %     \cos \alpha = \frac{ \sum_{a \in [r]} \cos \alpha_a \onorm{\mx_a} \onorm{\mx'_a} }{ \sqrt{ \sum_{a \in [r]} \onorm{\mx_a}^2} \sqrt{\sum_{a \in [r]} \onorm{ \mx'_a}^2} } \leq \frac{ \sum_{a \in [r]} \cos \alpha_a \onorm{\mx_a} \onorm{\mx'_a}}{ \sum_{a \in [r]} \onorm{\mx_a}\onorm{ \mx'_a}  }.
-    % \end{align}
-    % Then, we have 
-    % \begin{equation}
-    %     1 - \cos \alpha \geq\sum_{a \in [r]}   \frac{ \onorm{\mx_a} \onorm{\mx'_a} }{\sum_{a \in [r]} \onorm{\mx_a}\onorm{ \mx'_a}   } (1 - \cos \alpha_b) \gtrsim \max_{a \in [r]} (1 - \cos \alpha_b),
-    % \end{equation}
-    % where the last inequality follows by the fact that $\onorm{\mx_a} \asymp \onorm{\mx'_a} \asymp 1$.
-
-    % For the second inequality, note that 
-    % \end{proof}
-
-    % Consider the quantity
-    % \begin{equation}
-    %     J = \onorm{ \frac{\mW^{(t),T} \mX \mV^k}{\onormSize{}{\mW^{(t),T} \mX \mV^k}_F } - \frac{\mW^{(t),T} \mX \mV^{k+1}}{\onormSize{}{\mW^{(t),T} \mX \mV^{k+1}}_F } }_F.
-    % \end{equation}
-    % Let $\mx = [ \mx_1, \ldots, \mx_r]$ where $\mx_a = \mW_{:a}^{(t),T} \mX \mV^k$, and let $\mx' = [ \mx'_1, \ldots, \mx'_r]$ where $\mx'_a = \mW_{:a}^{(t),T} \mX \mV^{k+1}$.
-Also, let $\alpha$ denote the angle between $\mx$ and $\mx'$, and $\alpha_a$ denote the angle between $\mx_a, \mx'_a$.
-Note that $\onorm{\mx_a} \asymp \onorm{\mx'_a} \asymp 1$.
-Then, by Lemma~\ref{lem:concatenation}, we have
-    % \begin{equation}
-    %     J_{11}^2 = (1 - \cos \alpha_b) \leq \max_{a \in [r]} (1 - \cos \alpha_a) \lesssim (1 - \cos \alpha) = J^2.
-    % \end{equation}
-    % Then, we only need to bound $J$. By matricization, we have 
-    % \begin{equation}
-    %     J = \onorm{ \frac{\mW^{T} \mX \mZ^k}{\onormSize{}{\mW^{T} \mX \mZ^k}_F } - \frac{\mW^{(t), T} \mX \mZ^k}{\onormSize{}{\mW^{(t), T} \mX \mZ^k}_F } }_F.
-    % \end{equation}
-    % Let $ \my = [\my_1,\ldots, \my_r]$ where $\my_a = \mW_{:a}^T \mX \mZ^k$ and $\my'= [\my'_1, \ldots, \my'_r]$ where $\my'_a = \mW_{:a}^{(t),T} \mX \mZ^k$.
-Let $\beta$ denote the angle between $\my, \my'$ and $\beta_a$ denote the angle between $\my_a, \my'_a$.
-Note that $\onorm{\my_a} \asymp \onorm{\my'_a} \asymp 1$.
-Also, note that for any $a \in [r]$
-    % \begin{equation}
-    %     1 - \cos \beta_a \lesssim (L^{(t)}/\Delta_{\min})^2,
-    % \end{equation}
-    % where the inequality may follow the procedures in $I_{11}$ and the padding lemma. Then, we have 
-    % \begin{equation}
-    %     (1 - \cos \alpha) = (1 - \cos \beta) \asymp \max_{b \in [r]} (1 - \cos \beta_a)  \lesssim (L^{(t)}/\Delta_{\min})^2.
-    % \end{equation}
-    % Thus we finish the proof.
-
-    % }
-
-%     \fixme{Jiaxin}{Trying to fix the above issue.}
-%     Next, we derive the upper bound of $\sin \beta_k$.
-Recall the conclusion in the proof of $I_{11}$.
-For arbitrary $b \in [r]$ and let $\alpha$ the angle between $\mS_{b:}$ and $D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}$, we have
-%     \begin{equation}
-%         \sin \alpha \lesssim \frac{r L^{(t)}}{\Delta_{\min}} \lesssim \frac{\Delta_{\min}}{\log p} \lesssim \frac{1}{\log p},
-%     \end{equation}
-%     under the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin} and the last inequality follows by the fact that $\Delta_{\min} = \tO(1)$.
-This indicates that when $p$ is large enough the cosine $\cos \alpha \geq \frac{1}{2}$.
-Then, by geometry, we have
-%     \begin{align}
-%         \onormSize{}{ \mS_{b:} - [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}]} = \frac{  \onormSize{}{ [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}] \sin \alpha} }{ \cos \alpha}
-%     \end{align}
-% \fixme{Jiaxin}{This inequality seems not correct in geometry...}
-
-%\fixme{Miaoyan}{The above bound has same form as (55). }
-
-%   Also, let $\mD = \mM^T \mTheta^T \mW^{(t)} = \entry{D_{ab}} \in \bbR^{r \times r}$ be the confusion matrix, where 
-%   \begin{equation}
-%       D_{ab} = \frac{\sum_{i \in [p]} \theta(i) \ind \offf{ z(i) = a ,z^{(t)}(i) = b } }{\sum_{i \in [p]} \theta(i) \ind\offf{  z^{(t)}(i) = b} }.
-%   \end{equation}
-%   Since $\sum_{i \in [p]} \theta(i) \ind\offf{  z^{(t)}(i) = b}  \gtrsim p/r$ by Lemma~\ref{lem:membership} and the constraint $\min_{i \in [p]} \theta(i) \geq c_0$, we have 
-%   \begin{equation}
-%         \sum_{a \neq b} D_{ab} \lesssim \frac{r}{p} \sum_{i\colon z^{(t)}(i) \neq z(i)} \theta(i) \lesssim\frac{L^{(t)}}{\Delta_{\min}^2}  \lesssim \frac{1}{\log p}, \quad 
-%   \end{equation}
-
-    % Let $\mD = \mM^T \mTheta^T \mW^{(t)} \in \bbR^{r \times r}$ be the confusion matrix and $\mD_{ab} = \frac{\sum_{i \in [p]} \theta(i) \ind \offf{ z(i) = a ,z^{(t)}(i) = b } }{\sum_{i \in [p]} \theta(i) \ind\offf{  z^{(t)}(i) = b} }$.
-By Lemma~\ref{lem:membership}, we have $\sum_{i \in [p]} \ind\offf{ z^{(t)}(i) = a}  \asymp \frac{p}{r}$ for all $ a \in [r]$.
-Then we have
-    % \begin{equation}\label{eq:d_property}
-    %   \sum_{a \neq b} \mD_{ab} \asymp \frac{r}{p}  \sum_{a \in [r]} \sum_{b \neq a}  \ind \offf{ z(i) = a ,z^{(t)}(i) = b }  = \ell^{(t)}, \quad \text{and} \quad \sum_{a \in [r]} \mD_{aa} \asymp 1 -   \ell^{(t)}
-    % \end{equation}  Let $\mZ^k =  \mW^{\otimes (k-1)} \otimes \mW^{(t),\otimes (K-k-1)} \in \bbR^{ p^{K-2} \times r^{K-2} }$, and by Lemma~\ref{lem:membership}
-    % \begin{align}
-    %     &\onorm{\mZ^k}_{\sigma} \lesssim  \onorm{\mW}^{(k-1)}_{\sigma} \onorm{\mW^{(t)}}_{\sigma}^{K-k-1} \lesssim \sqrt{\frac{r^{K-2}}{p^{K-2}}},\label{eq:z_singular1}\\
-    %     &\lambda_{r^{K-2}} ( \mZ^k) \gtrsim   \lambda_{r}^{k-1} (\mW) \lambda_{r}^{K-k-1}(\mW^{(t)}) \gtrsim \sqrt{\frac{r^{K-2}}{p^{K-2}}}.\label{eq:z_singular2}
-    % \end{align} We have
-    % \begin{align}
-    %     \ang{\mW_{:b}^{(t),T} \mX \mV^k , \mW_{:b}^{(t),T} \mX \mV^{k+1}} &= \ang{ \mW^{(t),T} \mX \mZ^k, \mW^T \mX \mZ^k} \\
-    %     & = \ang{ \mD^T \mS (\mM^{\otimes K-1})^T \mW_{:b}^{(t)} \otimes \mZ^k, \mS (\mM^{\otimes K-1})^T \mW_{:b}^{(t)} \otimes \mZ^k } \\
-    %     & = \sum_{a \in [r]}\of{ \mD_{aa} \onorm{\mS'_{a:}}^2 + \sum_{b \neq a}\mD_{ba} \ang{\mS_{a:}', \mS_{b:}'}} \\
-    %     & \gtrsim (1 -   \ell^{(t)}) \min_{a \in [r]}  \onorm{\mS'_{a:}}^2  - \ell^{(t)} \max_{a \in [r]} \onorm{\mS'_{a:}}^2,
-    % \end{align}
-    % where $\mS'_{a:}$ refers to the $a$-th column of $\mS (\mM^{\otimes K-1})^T\mW_{:b}^{(t)} \otimes \mZ^k$ for $a \in [r]$, the first equation follows by tensor matricization, the second and third equations follows by the definition of $\mX$ and $\mD$, and the last inequality follows by the property~\eqref{eq:d_property}.
-By Lemma~\ref{lem:membership}, inequalities~\eqref{eq:z_singular1}, \eqref{eq:z_singular2}, and the constraint $c_3 \leq \min_{a \in [r]}\onorm{\mS_{a:}} \leq \max_{a \in [r]}\onorm{\mS_{a:}} \leq c_4 $ we have
-    % \begin{align}
-    %     \onorm{\mS'_{a:}} \leq \max_{a \in [r]}\onorm{\mS_{a:}} \onorm{\mM}_{\sigma}^{K-1} \onorm{\mW_{:b}^{(t)} \otimes}_F \onorm{\mZ^k}_{\sigma} \lesssim C_4 \\
-    %     \onorm{\mS'_{a:}} \geq \min_{a \in [r]}\onorm{\mS_{a:}} \lambda_r^{K-1}(\mM) \onorm{\mW_{:b}^{(t)} \otimes}_F \lambda_{r^{K-2}} ( \mZ^k) \gtrsim C_3.
-    % \end{align}
-    % By the initialization condition that $\ell^{(t)} = o(\log^{-1} p)$, we have $\ang{\mW_{:b}^{(t),T} \mX \mV^k , \mW_{:b}^{(t),T} \mX \mV^{k+1}}  > 0$ when $p$ large enough.
-This implies the angle $\alpha_k \in [0,\pi/2]$, and the trigonometric functions satisfy $\sqrt{1 - \cos \alpha_k} = 2 \sin \frac{\alpha_k}{2} \lesssim \sin \alpha_k$ for the acute angle.
-    % \fixme{Jiaxin}{Below are previous proof.}
-    % To upper bound $J_{11}^k$, we derive the upper bound for $\onorm{\mX \mV^k - \mX \mV^{k+1}}_F$. Note that 
-    % \begin{align}\label{eq:a-b_f}
-    %     \onorm{\mX \mV^k - \mX \mV^{k+1}}_F &= \onorm{ \mW^{(t),T} \mX( \mI\otimes \mZ^k) - \mW^T \mX( \mI\otimes \mZ^k) }_F\\
-    %     & = \onorm{ \mS (\mM^{\otimes K-1})^T ( \mI\otimes \mZ^k)  - \mD \mS  (\mM^{\otimes K-1})^T ( \mI\otimes \mZ^k)  }_F \\
-    %     & \leq \onorm{ \mC (\mI - \mD^T) \mS^s }_F \onorm{(\mM^{\otimes K-1})^T ( \mI\otimes \mZ^k)}_{\sigma}\\
-    %     & \lesssim \onorm{ (\mI - \mD^T) \mS^s }_F \onorm{\mM}_{\sigma}^{K-1} \onorm{\mZ^k}^{K-1}\\
-    %     & \lesssim \frac{r}{p} \sum_{a \in [r]} \sum_{b \neq a} \ind\offf{z(i) = a, z^{(t)}(i) = b} \onorm{ \mS^s_{a:} - \mS_{b:}^s}_F \sqrt{\frac{p}{k}}, 
-    % \end{align}
-    % \fixme{Jiaxin}{This inequality is not correct due to the definition of $\mD$.}
-    % {\color{blue}
-    % \begin{align}
-    %     \onorm{(\mI -\mD^T)\mS^s}_F &\leq \sum_{b \in [r]}   \onormSize{}{ \mS_{b:}^s - [D_{bb} \mS_{b:}^s + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}^s]} \\
-    %     & \leq \sum_{b \in [r]}   \onormSize{}{ \mS_{b:}^s - \sum_{a \in [r]} D_{ab} \mS_{b:}^s + \sum_{a \neq b, a \in [r]} D_{ab} (\mS_{b:}^s- \mS_{a:}^s)}  \\
-    %     & \leq \sum_{b \in [r]} \of{ \onorm{(1 - \sum_{a \in [r]} D_{ab})\mS_{b:}^s }  + \sum_{a \neq b} D_{ab} \onorm{\mS_{b:}^s - \mS_{a:}^s}}\\
-    %     & \lesssim \frac{L^{(t)}}{\Delta_{\min}^2} + \frac{L^{(t)}}{\Delta_{\min}}
-    % \end{align}
-    % }
-
-   % \[
-    %\onormSize{}{\mC(\mI-\mD^T)\mS^s}_F\leq \sum_b\onormSize{}{\mS^s_{b\colon}-[(\mI-\mD^T)\mS^s]_{b\colon}}_F
-    %\]
-
-    %\fixme{Miaoyan}{how about using spectral norm? When $\mS^s$ is identity matrix.}
-   % \[
-   % \onormSize{}{\mD^T\mD}_\sigma\gtrsim \min_{b}\left(D_{bb}-\sum_{a\in[r],a\neq  b}D_{ab}\right)^2\gtrsim 1-4\max_{b}\sum_{a\in[r],a\neq b}D_{ab}.
-   % \]
-  %\[
-  %\onormSize{}{\mI-\mD^T}_\sigma \lesssim \max_{b}\sum_{a\in[r],a\neq b}D_{ab}.
-  %\]
-  %General $\mS^s$:
-  %\[
-  %\onormSize{}{(\mI-\mD^T)\mS^s}_\sigma
-  %\]
-
-    % where $\mC = \text{diag}(\onorm{\mS_{1:}}, \ldots, \onorm{\mS_{r:}})$ is a diagonal matrix with lower and upper bounded components, and $\mS^s \in \bbR^{r \times r^{K-1}}$ is the matrix with normalized rows $\mS_{a:}^s$, and the last inequality follows by the definition of $\mD$, triangle inequality and the singular property of $\mM$ and $\mZ^k$.
-
-    % Also, note that 
-    % \begin{align}\label{eq:wxv_bound}
-    %     \onorm{\mW_{:b}^{(t),T} \mX \mV^{k+1}} &= \onorm{\mW_{:b}^{(t),T} \mM \mS (\mM^{\otimes K-1})^T \mV^{k+1} } \\
-    %     &\geq \onorm{ \mW_{:b}^{(t),T} \mM \mS }_F \lambda_{r^{K-1}} \of{(\mM^{\otimes K-1})^T \mV^{k+1} } \\
-    %     & \gtrsim \lambda_r^{K-1}(\mM)^T \lambda_r^{k+1}(\mW) \lambda_r^{K-k-2}(\mW^{(t)}) \\
-    %     & \gtrsim c,
-    % \end{align}
-    % where $c$ is some positive constant, and the last inequality follows by the boundedness assumption on $\mS$ and the singular property in Lemma~\ref{lem:membership}.
-
-    % For two vectors $\mv_1, \mv_2 \in \bbR^d$, let $\text{Proj}(\mv_1) \in \bbR^{d \times d}$ denote the projection matrix from $\bbR^d$ to the direction of $\mv_1$. We have the fact that 
-    % \begin{align}\label{eq:proj_tri}
-    %     \onorm{(\mI - \text{Proj}(\mv_1)) \mv_2} \leq  \onorm{\mv_1 - \mv_2}.
-    % \end{align}
-
-    % Hence, we have 
-    % \begin{align}
-    %     J_{11}^k  \stackrel{\sin \frac{\alpha}{2} \lesssim \sin \alpha, \text{ for } \alpha \in [0,\pi/2]}{\lesssim} &  \sin \alpha_k \\
-    %     =  \quad \quad \quad \quad &\frac{ \onorm{(\mI - \text{Proj}(\mW_{:b}^{(t),T} \mX \mV^k )) \mW_{:b}^{(t),T} \mX \mV^{k+1}}  }{ \onorm{\mW_{:b}^{(t),T} \mX \mV^{k+1}} }\\
-    %     \stackrel{\eqref{eq:proj_tri},\eqref{eq:wxv_bound}  }{\lesssim}\quad \quad \quad  &  \onorm{\mW_{:b}^{(t),T} \mX \mV^k  - \mW_{:b}^{(t),T} \mX \mV^{k+1} }\\
-    %     \leq \quad \quad \quad \quad & \onorm{\mX \mV^k - \mX \mV^{k+1}}_F  \onorm{\mW_{:b}^{(t),T} }\\
-    %     \stackrel{\eqref{eq:a-b_f}  }{\lesssim} \quad \quad \quad \quad &  \frac{r}{p} \sum_{a \in [r]} \sum_{b \neq a} \ind\offf{z(i) = a, z^{(t)}(i) = b} \onorm{ \mS^s_{a:} - \mS_{b:}^s}_F \\
-    %   \leq  \quad \quad \quad \quad & \frac{r L^{(t)}}{\Delta_{\min}}.
-    % \end{align}
 
@@
 
-    % \fixme{Jiaxin}{End $J_{11}$}
-
-    Consider $J_{12}$. By triangle inequality, we have 
+    We now consider $J_{12}$. By triangle inequality, we have 
     \begin{align}
@@
     \end{equation}
-    where inequality follows by the bounds~\eqref{eq:singular_Dl} and \eqref{eq:singular_z}.
+    where the inequality follows from the bounds~\eqref{eq:singular_Dl} and \eqref{eq:singular_z}.
 
@@
     Notice that 
-    \begin{align}
-        \onormSize{}{\mX (\mV^k-\mV^{k+1})}_F &\leq \onormSize{}{(\mI - \mD^T) \mS (\mI_r^{\otimes (k-1)} \otimes  \mD^{\otimes (K-k-1)})}_F \\
-        &\leq \onormSize{}{(\mW^T - \mW^{(t),T}) \mTheta \mM}_F \onorm{\mS}_F \onorm{ \mD}_{\sigma}^{K-k-1}\\
-        & \lesssim \onormSize{}{ \mW^T - \mW^{(t),T} } \onorm{\mTheta \mM}_{\sigma} \\
+    \begin{align}\label{eq:step}
+        \onormSize{}{\mX (\mV^k-\mV^{k+1})}_F &\leq \onormSize{}{(\mI - \mD^T) \mS (\mI_r^{\otimes (k-1)} \otimes  \mD^{\otimes (K-k-1)})}_F\notag \\
+        &\leq \onormSize{}{(\mW^T - \mW^{(t),T}) \mTheta \mM}_F \onorm{\mS}_F \onorm{ \mD}_{\sigma}^{K-k-1}\notag \\
+        & \lesssim \onormSize{}{ \mW^T - \mW^{(t),T} } \onorm{\mTheta \mM}_{\sigma} \notag\\
         & \lesssim \frac{ \sqrt{r L^{(t)}}}{\Delta_{\min}},
     \end{align}
-    where the first inequality follows by the tensor algebra in 
-    inequality~\eqref{eq:j11inner}, and the second inequality follows by the fact  that $\mI = \mW^T \mTheta \mM$, and the last inequality follows by \citet[Proof of Lemma 5]{han2020exact}.
-Thus, we have
+    where the first inequality follows from the tensor algebra in 
+    inequality~\eqref{eq:j11inner}, the second inequality follows from the fact  that $\mI = \mW^T \mTheta \mM$, and the last inequality follows from \citet[Proof of Lemma 5]{han2020exact}.
+It follows from~\eqref{eq:step} and Lemma~\ref{lem:membership} that
     \begin{equation}
@@
     \end{align}
-    where the second inequality follows by inequalities~\eqref{eq:j12_wev}, \eqref{eq:j12_wxv}, and the inequality \eqref{eq:cond1} in Condition~\ref{cond:origin}.
-
-    % Consider $J_{12}$.
-Note that by Lemma~\ref{lem:membership}, $c_3 \leq \onorm{\mS_{b:}} = \onorm{\mW_{:b}^T \mX \mV}  $ and $\onorm{\mW_{:b}^{(t),T} \mX \mV}$ is bouneded.
-Also, by \citet[Inequality (123) (124) (125), Proof of Lemma 5]{han2020exact}, we have
-    % \begin{equation}\label{eq:j12}
-    %     J_{12} \lesssim \onorm{\mW_{:b}^{(t),T}  \mE (\mV - \mV^{(t)})  } \lesssim \sqrt{ \frac{rr^{2K} + pr^{K+2}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}.
-    % \end{equation}
+    where the second inequality follows from inequalities~\eqref{eq:j12_wev}, \eqref{eq:j12_wxv}, and the inequality \eqref{eq:cond1} in Condition~\ref{cond:origin}.
 
@@
     following the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4}. Therefore, we finish the proof of inequality~\eqref{eq:inter5}.
-    % The proof follows the proof of inequality~\eqref{eq:inter3} except (1) the term $I_{11}$ now is upper bounded as 
-    % \begin{equation}
-    %     I_{11} \lesssim  \frac{r}{p} \sum_{i \in [p]} \sum_{a \neq b, a \in [r]} \ind \offf{ z(i) = b, z^{(i)} = a }  \onorm{ \off{\mW_{:b}^T \mX \mV^{(t)}}^s - \off{ \mW_{:a}^T \mX \mV^{(t)} }^s },
-    % \end{equation}
-    % and (2) the term $I_{12}$ has the upper bound in \citet[Inequality (127), Proof of Lemma 5]{han2020exact}.
-For $I_{11}$, following the procedures to derive $J_{11}$ \eqref{eq:inter4_newcenter} in the proof of inequality~\eqref{eq:inter4}, we also have
-    % \begin{equation}\label{eq:inter_wa_wb}
-    %     \onorm{ \off{\mW_{:b}^T \mX \mV^{(t)}}^s - \off{ \mW_{:a}^T \mX \mV^{(t)} }^s } \lesssim  \onorm{ \off{\mS_{b:}}^s - \off{ \mS_{a:} }^s }.
-    % \end{equation}
-    % Therefore, we finish the proof of inequality~\eqref{eq:inter5}.
         \end{enumerate}

### 2022-01-18 tensor_clustering/tensor_clustering_ieee/ieee_v3.tex
v3

@@
 
-    {\bf Step 3 (Derivation of contraction inequality).} In this step we derive the upper bound of $\zeta_{ib}$ and obtain the contraction inequality~\eqref{eq:proof_5_ineq}. Note that 
+    {\bf Step 3 (Derivation of contraction inequality).} In this step we derive the upper bound of $\zeta_{ib}$ and obtain the contraction inequality~\eqref{eq:proof_5_ineq}.
+Choose the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ that satisfies the condition of Lemma~\ref{lem:upper_fgh}, inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
+Note that
     \begin{align}
@@
     where the first inequality follows from the inequality~\eqref{eq:hib} in Lemma~\ref{lem:upper_fgh}, and the last inequality follows from the assumption that $\min_{i \in [p]} \theta(i) \geq c>0$.
-Choose the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ that satisfies inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
 Following \citet[Step 4, Proof of Theorem 2]{han2020exact} and Lemma~\ref{lem:upper_fgh}, we have

### 2022-01-18 tensor_clustering/dTBM/R/tensor_clustering.R
update

@@ -1,638 +1,0 @@
-#' Weighted higher-order initialization
-#'
-#' Weighted higher-order initialization for multiway spherical clustering under degree-corrected tensor block model.
-#' This function takes the tensor/matrix observation, the number of clusters, and a logic variable indicating the symmetry
-#' as input. Output is the estimated clustering assignment.
-#'
-#'
-#' @param Y     array/matrix, order-3 tensor/matrix observation
-#' @param r     vector, the number of clusters on each mode; see "details"
-#' @param asymm logic variable, if "TRUE", assume the clustering assignment differs in different modes; if "FALSE", assume all the modes share the same clustering assignment
-#' @return a list containing the following:
-#'
-#' \code{z0} { a list of vectors recording the estimated clustering assignment }
-#'
-#' \code{s0} { a list of vectors recording the index of degenerate entities with random clustering assignment}
-#'
-#' @details   \code{r} should be a length 2 vector for matrix and length 3 vector for tensor observation;
-#'
-#' all the elements in \code{r} should be integer larger than 1;
-#'
-#' symmetric case only allow \code{r} with the same number of clusters on each mode;
-#'
-#' observations with non-identical dimension on each mode are only applicable with \code{asymm = T}.
-#'
-#'
-#'
-#' @export
-#' @examples
-#'
-#' test_data = sim_dTBM(seed = 1, imat = FALSE, asymm = FALSE, p = c(50,50,50), r = c(3,3,3),
-#'                     core_control = "control", s_min = 0.05, s_max = 1,
-#'                     dist = "normal", sigma = 0.5,
-#'                     theta_dist = "pareto", alpha = 4, beta = 3/4)
-#'
-#' initialization <- wkmeans(test_data$Y, r = c(3,3,3), asymm = FALSE)
-#'
-
-
-wkmeans <- function(Y, r, asymm) {
-  imat <- F
-
-  if (length(dim(Y)) == 2) {
-    cat("matrix case \n")
-    dim(Y) <- c(dim(Y), 1)
-    imat <- T
-  }
-
[…]

### 2022-01-18 tensor_clustering/ieee_submission/coverletter.tex
organize

\documentclass{article}[12pt]
\usepackage{enumitem}
\usepackage{lipsum}
\input{UWletterdefs}
\usepackage{ragged2e}
\justifying
\usepackage{amsmath,amssymb}

\begin{document}
Dear Editors,

I am writing to submit our manuscript ``Multiway Spherical Clustering via Degree-Corrected Tensor Block Models'' to the IEEE Transactions on Information Theory.

In this manuscript, we investigate the degree-corrected tensor block model (dTBM) for multiway clustering. The goal of multiway clustering is to identify the checkerboard structure in a noisy data tensor collected from social sciences, neuroscience, genetics, and computer vision applications. Compared with classical tensor block models, dTBM allows both block and individual effects and assumes a more flexible structure in multiway clustering. Studying the statistical and computational limits of dTBM and developing an optimal algorithm are of necessity and importance to the field of multiway clustering.

We present the phase transition of clustering performance under dTBM based on the notion of angle separability. We characterize three signal-to-noise regimes corresponding to different statistical-computational behaviors. In particular, we demonstrate that an intrinsic statistical-to-computational gap emerges only for tensors of order three or greater. Further, we develop an efficient polynomial-time algorithm that provably achieves exact clustering under mild signal conditions. The efficacy of our procedure is verified by two applications, one on human brain connectome project, and another on Peru Legislation network dataset.

Part of this work was presented at the {\bf 25th International Conference on Artificial Intelligence and Statistics (AISTATS)}. However, this IEEE submission covers much greater depth and extends the earlier publication in a substantial way. Newly added results are summarized below.
\begin{itemize}[topsep=0ex]
\item We have added four new motivating problems to highlight the applicability of our proposed models. The new applications include community detection in hypergraphes, clustering in multi-layer weighted networks, and multi-tissue multi-individual gene expression analysis. See Section II.B.

\item The earlier publication addresses only the special symmetric tensors. We have substantially extended our algorithm to more general models, including asymmetric tensors and non-Gaussian observations. New theoretical results are provided to justify the extension. We have also added two initialization schemes, one for independent and identically distributed sub-Gaussian noise, and another for independent but non-identically distributed Bernoulli noise. Theories are extended correspondingly. See Section V.A.

\item We have proposed a new procedure for hyperparameter selection. We have added new simulation studies to assess the efficacy in Section V.A and Table II.

\item We have added an entirely new real data application on Human Brain Connection Project (HCP). The extensive data analysis in Section VI.A and new Figures 9-11 demonstrate the potential of our method in neuroscience applications.
\end{itemize}

We believe our results will be of interest to a very broad readership – from those interested in new theoretical results to those interested in efficient algorithms in multiway clustering with social networks and neuroscience applications. Our method will help the practitioners efficiently analyze tensor datasets in various areas. Toward this end, we also submit the developed software package accompanying this paper.

We appreciate your consideration.

\end{document}

### 2022-01-18 tensor_clustering/others/ieee_v2.tex
organize

--- 2022-01-18 tensor_clustering/tensor_clustering_ieee/ieee_v3.tex
@@
 \end{equation}
-where $\mTheta' = \text{diag}(\mtheta') = \text{diag}(1,1, 1, 1)$, the core tensor $\tS' \in \bbR^{2 \times 2 \times 2}$ has mode-3 slices, and the membership matrix $\mM'\in\{0,1\}^{4\times 2}$ defines the clustering $z'\colon[4]\to[2]$,
+where $\mTheta' = \text{diag}(\mtheta') = \text{diag}(1,1, 1, 1)$, the core tensor $\tS' \in \bbR^{2 \times 2 \times 2}$ has mode-3 slices, and the membership matrix $\mM'\in\{0,1\}^{2\times 4}$ defines the clustering $z'\colon[4]\to[2]$,
 \begin{equation}
@@
     \end{bmatrix}, \quad \mM' = \begin{bmatrix}
-   1 & 0 \\
-   1 &0 \\
-    0 & 1 \\
-    0&1
+   1& 1 & 0 & 0\\
+    0 &0 &  1& 1 
     \end{bmatrix}.
@@
 \Algphase{Sub-algorithm 1: Weighted higher-order initialization}
-\INPUT Observation $\tY \in \bbR^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
+\INPUT Observation $\tY \in \bbR^{p\times \cdots \times p}$, number of clusters $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
 \State Compute factor matrix $ \mU_{\text{pre}} = \text{SVD}_{r} (\text{Mat}(\tY))$ and the $(K-1)$-mode projection $
@@
 \begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the setup as in Theorem~\ref{thm:initial}.
-Assume the local linear stability of degree holds in the neighborhood $\tN(z,\epsilon)$ for $\epsilon = \log^{-1}p$.
+Assume the local linear stability of degree holds in the neighborhood $\tN(z,\epsilon)$ for $\epsilon \geq \log^{-1}p$.
 Suppose
-$r=\tO(1)$ and $\text{SNR} \geq \tilde C p^{-K/2}\log p$ for some large constant $\tilde C$.
+$r=\tO(1)$ and $\text{SNR} \gtrsim  p^{-K/2}\log p$.
 Let $z^{(t)}$ denote the $t$-th iteration output in Sub-algorithm~\hyperref[alg:main]{2} with initialization $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{1}.
@@
 Specifically, in the Sub-algorithm~\hyperref[alg:main]{1}, we make the following changes: (1)
-Replace the matrcization $\mat(\tY)$ by $\mat_k(\tY)$; (2) Repeat the Steps 1-5 with mode-specified cluster number $r_k$; (3) Obtain the collection initialization $ \{z^{(0)}_k\}_{k=1}^K$.
+Replace the matrcization $\mat(\tY)$ by $\mat_k(\tY)$; (2) Repeat the Steps 1-5 with mode-specified number of clusters $r_k$; (3) Obtain the collection initialization $ \{z^{(0)}_k\}_{k=1}^K$.
 In the Sub-algorithm~\hyperref[alg:main]{2}, we make the following changes: (1) Take the collection $\{z^{(0)}_k\}_{k=1}^K$ as input, and update the block tensor $\tS^{(t)}$ with the collection $\{ z_k^{(t)}\}_{k=1}^K$, $
@@
 
-{\bf Hyper-parameter selection.} In our theoretical analysis, we have assumed the true cluster number $r$ is given to our algorithm.
-In practice, the cluster number $r$ is often unknown, and we now propose a method to choose $r$ from data.
+{\bf Hyper-parameter selection.} In our theoretical analysis, we have assumed the true number of clusters $r$ is given to our algorithm.
+In practice, the number of clusters $r$ is often unknown, and we now propose a method to choose $r$ from data.
 We impose the Bayesian information criterion (BIC) and choose the cluster number that minimizes BIC; i.e., under the symmetric Gaussian dTBM~\eqref{eq:model_tensor},
@@
     \cmidrule(lr){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9}
-         True cluster number $r$ & 2 & 4 & 2  & 4 & 2  & 4 & 2 & 4  \\
+         True number of clusters $r$ & 2 & 4 & 2  & 4 & 2  & 4 & 2 & 4  \\
          \midrule
-         Estimated cluster number $\hat r$ &  2(0)  & 3.9(0.25)& 2(0)    & 3.1(0.52) & 2(0)    & 4(0)   & 2(0)    & 3.9(0.31)   \\
+         Estimated number of clusters $\hat r$ &  2(0)  & 3.9(0.25)& 2(0)    & 3.1(0.52) & 2(0)    & 4(0)   & 2(0)    & 3.9(0.31)   \\
      \bottomrule
     \end{tabularx}
-    \caption{Estimated cluster number given by BIC criterion under the low noise level $(\sigma^2 = 0.25)$ and high noise level $(\sigma^2 = 0.5)$ settings.
+    \caption{Estimated number of clusters given by BIC criterion under the low noise $(\sigma^2 = 0.25)$ and high noise $(\sigma^2 = 0.5)$ settings.
 Numbers in parentheses are standard deviations of $\hat r$ over 30 replications.}
@@
 
-The third experiment evaluates the empirical performance of the BIC criterion to select unknown cluster number.
+The third experiment evaluates the empirical performance of the BIC criterion to select unknown number of clusters.
 We generate the data from an order-3 Gaussian model with $p = \{50,80\}$, $r = \{2,4\}$, and noise level $\sigma^2 \in \{ 0.25,1\}$.
 Table~\ref{tab:select} shows that our BIC criterion well chooses the true $r$ under most settings.
-Note that the BIC slightly underestimates the true cluster number $(r = 4)$ with smaller dimension and higher noise $(p = 50, \sigma=1)$, and the accuracy immediately increases with larger dimension $p = 80$.
+Note that the BIC slightly underestimates the true number of clusters $(r = 4)$ with smaller dimension and higher noise $(p = 50, \sigma=1)$, and the accuracy immediately increases with larger dimension $p = 80$.
 The improvement follows from the fact that a larger dimension $p$ indicates a larger sample size in the tensor block model.
-Therefore, we conclude that BIC criterion is a reasonable way to tune the cluster number.
+Therefore, we conclude that BIC criterion is a reasonable way to tune the number of clusters.
 
@@
 where the superscript $\cdot^{s}$ denotes the normalized vector; i.e., $\ma^s:=\ma/\onorm{\ma}$ if $\ma\neq 0$ and $\ma^{s}=0$ if $\ma=0$ for any vector $\ma$.
-We show that $L^{(t)}$ provides an upper bound for the misclassification error of interest via the inequality $\ell^{(t)}\leq {L^{(t)}\over \Delta^2_{\min}}$.
+We show that $L^{(t)}$ provides an upper bound for the misclassification error of interest via the inequality $\ell^{(t)}\lesssim {L^{(t)}\over \Delta^2_{\min}}$.
 Therefore, it suffices to control $L^{(t)}$.
@@
 \begin{equation}\label{eq:proof_5_ineq}
-    L^{(t+1)} \leq M \xi + \rho L^{(t)}, \quad \xi = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  A_{ib} \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
-\end{equation}
-where $M$ is a positive constant, $\rho \in (0,1)$ is the contraction parameter, and we call $\xi$ the oracle loss.
+    L^{(t+1)} \lesssim \xi + \rho L^{(t)}, \quad \xi = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  A_{ib} \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
+\end{equation}
+where $\rho \in (0,1)$ is the contraction parameter, and we call $\xi$ the oracle loss.
 Controlling the probability of event $B_{ib}$ and obtaining the $\rho L^{(t)}$ term in the right hand side of~\eqref{eq:proof_5_ineq} are the most challenging parts in the proof of Theorem~\ref{thm:refinement}.
@@
 Let $\{z'_k\} = \{z_k\}$, $\tS' = \tS$, and $\mtheta'_k = \mtheta_k$ for all $k = 2, \ldots, K$.
-Then the triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ is distinct from $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation.
+Then the triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ is distinct from \of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation.
 Similar conclusion holds when $\mS_{1,2:}$ is a zero vector.
@@
 
-\begin{cor}[Initial misclustering loss]\label{cor:L0} 
-Under the setup of Theorem~\ref{thm:initial}, assume SNR $\geq \tilde  C p^{-K/2} \log p$ for some positive universal constant $\tilde C$ and $\sigma^2 = 1$.
-With probability going to 1, the misclustering loss for the initialization is upper bounded as
-\begin{equation}
-    L^{(0)} =  \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(0)}(i) = b } \onormSize{}{ [ \mS_{ \pi^{(0)}( z(i)):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq \frac{\bar C \Delta_{\min}^2}{ \tilde C r \log p},
-\end{equation}
-where $\pi^{(0)}$ minimizes the initial misclustering error; i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} $ $ \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$, and $\bar C > 1$ is a positive universal constant.
+\begin{cor}[initial misclassification error]\label{cor:L0} 
+Suppose SNR $\gtrsim p^{-K/2} \log p$ and  $\sigma^2 = 1$. Then, the misclustering loss for the initialization is upper bounded as 
+\begin{equation}
+    L^{(0)} =  \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(0)}(i) = b } \onormSize{}{ [ \mS_{ \pi^{(0)}( z(i)):}  ]^s - [ \mS_{b:}  ]^s  }^2 \lesssim \frac{\Delta_{\min}^2}{r \log p},
+\end{equation}
+where $\pi^{(0)}$ minimizes the initial misclustering error; i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} $ $ \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$.
 \end{cor}
@@
      & \lesssim \frac{1}{p}  \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \of{ \onormSize{}{\mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 + \onormSize{}{ \hat \mx_{z^{(0)}(i)} - \mX_{z^{(0)}(i)}^s  }^2 } \\
-     & \leq \bar C \frac{r^{K}}{p^{K}} \of{ p^{K/2}r + pr^2 + r^K}, \\
-     & \leq \frac{\bar C\Delta_{\min}^2}{ \tilde C r \log p}
-\end{align}
-where the first inequality follows from Lemma~\ref{lem:angle_gap_x}, the third inequality follows from inequalities~\eqref{eq:cor_3} and \eqref{eq:cor_4}, and the last inequality follows from the assumption that SNR $\geq \tilde C p^{-K/2} \log p$.
+     & \lesssim \frac{r^{K}}{p^{K}} \of{ p^{K/2}r + pr^2 + r^K}, \\
+     & \lesssim \frac{\Delta_{\min}^2}{r \log p}
+\end{align}
+where the first inequality follows from Lemma~\ref{lem:angle_gap_x}, and the third inequality follows from inequalities~\eqref{eq:cor_3} and \eqref{eq:cor_4}.
 \end{proof}
@@
 
-\begin{lem}[Angle gap in $\tX$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family}.
+\begin{lem}[Angle gap in $\tX$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor}.
 Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying~\eqref{eq:degree}.
@@
 
-\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the parameter space~\eqref{eq:family} and assumption that $\min_{i \in [p]} \theta(i) \geq c$, the singular values of $\mTheta \mM$ are bounded as
+\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the assumption that $\min_{i \in [p]} \theta(i) \geq c$, the singular values of $\mTheta \mM$ are bounded as 
 \begin{equation}\label{eq:membership2}
@@
 Then, if
-\begin{equation}\label{eq:upper_mis_cond}
+\begin{equation}
     \min_{a \neq b \in [r]} \onormSize{}{\mv_a - \mv_b} \geq 2c,
@@
 \begin{equation}
-   S = \{i \in S_0^c: \onormSize{}{\hat \mx_i - \mv_{z(i)} } \geq c \}.
+   S = \{i \in S_0^c: \onormSize{}{\hat \mx_i - \mx_i } \geq c \}.
 \end{equation}
@@
 \begin{equation}
-    C_u = \{ i \in z^{-1}(u) \cap S_0^c: \onormSize{}{\hat \mx_i - \mv_{z(i)}} < c \},
+    C_u = \{ i \in z^{-1}(u) \cap S_0^c: \onormSize{}{\hat \mx_i - \mx_i} < c \},
 \end{equation}
@@
 \end{align}
-Note that $\cup_{u \in [r]}C_u = S_0^c/S^c$ and $C_u \cap C_v = \emptyset$ for any $u \neq v$.
-Suppose there exist $ i \in C_u$ and $j \in C_v$ with $u \neq v \in [r]$ and $\hat z(i) = \hat z(j)$.
-Then we have
-\begin{equation}
-    \onormSize{}{\mv_{z(i)} - \mv_{z(j)} } \leq  \onormSize{}{\mv_{z(i)} - \hat \mx_{i} } + \onormSize{}{\mv_{z(j)} - \hat \mx_j } < 2c,
-\end{equation}
-which contradicts to the assumption~\eqref{eq:upper_mis_cond}.
-Hence, the estimates $\hat z(i) \neq \hat z(j)$ for the nodes $ i \in C_u$ and $j \in C_v$ with $u \neq v$.
-By the definition of $R_2$, the nodes in $\cup_{u \in R_2} C_u$ have the same assignment with $z$ and $\hat z$.
-Then,
-%Following \citet[Lemma 6]{gao2018community} \fixme{Miaoyan}{Remove citation. fill in details in a self-contained way. }, 
-we have 
-\begin{equation}
-    \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \leq \sum_{i \in S_0} \theta(i)^2 + \sum_{i \in S} \theta(i)^2 + \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2.
-\end{equation}
- We only need to bound $\sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2$ to finish the proof.
-Note that every $C_u$ with $u \in R_3$ contains at least two nodes assigned to different clusters by $\hat z$.
-Then, we have $|R_2| + 2 |R_3| \leq r$.
-Since $|R_1| + |R_2| + |R_3| = r$, we have $|R_3| \leq |R_1|$.
-Hence, we obtain
+Following \citet[Lemma 6]{gao2018community} \fixme{Miaoyan}{Remove citation. fill in details in a self-contained way. }, we have 
+\begin{equation}
+    \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \leq \sum_{i \in S_0} \theta(i)^2 + \sum_{i \in S} \theta(i)^2 + \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2,
+\end{equation}
+and $|R_3| \leq |R_1|$. We only need to bound $\sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2$ to finish the proof. Note that 
 \begin{align}
@@
 \begin{equation}\label{eq:cond_oracle}
-    \xi \leq \exp\of{ - M \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}},
+    \xi \lesssim \exp\of{ - \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}},
 \end{equation}
@@
 \begin{equation}\label{eq:cond_intial}
-    L^{(t)} \leq \frac{\bar C}{\tilde C} \frac{\Delta_{\min}^2}{r \log p}, \quad \text{for} \quad t = 0, 1, \ldots, T,
-\end{equation}
-where $M$ is a positive universal constant in inequality~\eqref{eq:constant_M}, $\bar C, \tilde C$ are universal constants in Corollary~\ref{cor:L0} and assumption SNR $\geq \tilde C p^{-K/2} \log p$, respectively.
+    L^{(t)} \lesssim \frac{\Delta_{\min}^2}{r \log p}, \quad \text{for} \quad t = 0, 1, \ldots, T.
+\end{equation}
+
 Further, inequality~\eqref{eq:cond1} holds by replacing $\mV$ to $\mV^{(t)}$ and $\mW_{:a}$ to $\mW_{:a}^{(t),T}$ when initialization condition~\eqref{eq:cond_intial} holds.
@@
         & \leq \theta(i) \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \ind \offf{z^{(t+1)}(i) = b, \frac{1}{4}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} } \\
-        & \leq 64 \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ cm^2\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    }
-        %&\leq  \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    },\label{eq:zeta}
+        & \lesssim \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \theta(i)m^2\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    }\\
+        &\leq C \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    },\label{eq:zeta}
     \end{align}
     where the first inequality follows from the inequality~\eqref{eq:hib} in Lemma~\ref{lem:upper_fgh}, and the last inequality follows from the assumption that $\min_{i \in [p]} \theta(i) \geq c>0$.
-Choose the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ that satisfies inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
 Following \citet[Step 4, Proof of Theorem 2]{han2020exact} and Lemma~\ref{lem:upper_fgh}, we have
     \begin{align}\label{eq:f_sum}
-          \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{(F_{ib}^{(t)})^2}{ cm^2 \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \leq \frac{C_0 \bar C}{c m^2 \tilde C^2} L^{(t)},
-    \end{align}
-   for a positive universal constant $C$ and 
-    \begin{align}
-         \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}  &\leq \frac{1}{512} \frac{1}{p} \sum_{i \in [p]}\theta(i) \sum_{b \in [r]/z(i)}   \ind \offf{ z^{(t+1)}(i) = b}  (\Delta_{\min}^2 + L^{(t)}) \\
-         & \leq \frac{1}{512} (L^{(t+1)} + L^{(t)}), \label{eq:g_sum}
+          \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{(F_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \leq C_1 L^{(t)},
+    \end{align}
+    and 
+    \begin{align}
+         \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}  &\lesssim \frac{1}{p} \sum_{i \in [p]}\theta(i) \sum_{b \in [r]/z(i)}   \ind \offf{ z^{(t+1)}(i) = b}  (\Delta_{\min}^2 + L^{(t)}) \\
+         & \leq C_2 (L^{(t+1)} + L^{(t)}), \label{eq:g_sum}
     \end{align}
     where the last inequality follows from the definition of $L^{(t)}$ and the constraint of $\mtheta$ in parameter space~\eqref{eq:family}.
-    %Choosing the constants $C, \bar C, C_2$ in~\eqref{eq:zeta}~\eqref{eq:f_sum}~\eqref{eq:g_sum} carefully \fixme{Miaoyan}{be specific. Add one sentence to explain}, 
-   For $\tilde C$ also satisfies 
-    \begin{equation}\label{eq:tilde_c3}
-         \frac{C_0 \bar C}{c m^2 \tilde C^2} \leq \frac{1}{512},
-    \end{equation}
-    we have 
+Choosing the constants $C, C_1, C_2$ in~\eqref{eq:zeta}~\eqref{eq:f_sum}~\eqref{eq:g_sum} carefully \fixme{Miaoyan}{be specific.
+Add one sentence to explain}, we have
     \begin{equation}\label{eq:zeta_upper}
-        \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}^{(t)} \leq  \frac{1}{8} L^{(t+1)}  + \frac{1}{4} L^{(t)}.
-    \end{equation}
-    %for some small \fixme{Miaoyan}{In what sense "small"? Add one sentence.} positive constants $C_4, C_5$. 
+        \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}^{(t)} \leq  C_4 L^{(t+1)}  + C_5 L^{(t)},
+    \end{equation}
+    for some small \fixme{Miaoyan}{In what sense "small"?
+Add one sentence.} positive constants $C_4, C_5$.
 Plugging the inequality~\eqref{eq:zeta_upper} into the decomposition~\eqref{eq:decomp2}, we obtain the contraction inequality
     \begin{equation}\label{eq:decomp_t}
-          L^{(t+1)} \leq \frac{3}{2} \xi  + \frac{1}{2} L^{(t)},
-    \end{equation}
-     where $\frac{1}{2}$ is the contraction parameter.
-
-     Therefore, with $\tilde C$ satisfying inequalities~\eqref{eq:tilde_c3}, \eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}, we obtain the conclusion in Theorem~\ref{thm:refinement} via inequality~\eqref{eq:decomp_t} combining the inequality~\eqref{eq:cond_oracle} in Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.
+          L^{(t+1)} \leq M \xi  + \rho L^{(t)},
+    \end{equation}
+     where $M \geq 1$ and $\rho \in [0,1)$ is the contraction parameter.
+
+     Combining the inequality~\eqref{eq:cond_oracle} in Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}, we have proven Theorem~\ref{thm:refinement} with inequality~\eqref{eq:decomp_t}.
 
@@
     Let $e_i = \mE_{i:} \mV$ denote the aggregated noise vector for all $i \in [p]$, and $e_i$'s are independent zero-mean sub-Gaussian vector in $\bbR^{r^{K-1}}$.
-The entries in $e_i$ are independent zero-mean sub-Gaussian variables with sub-Gaussian norm upper bounded by $m_1\sqrt{r^{K-1}/p^{K-1}}$ with some positive constant $m_1$.
+The entries in $e_i$ are independent zero-mean sub-Gaussian variables with sub-Gaussian norm upper bounded by $C\sqrt{r^{K-1}/p^{K-1}}$ with some positive consta
 We have the probability inequality
@@
 
-    For $P_1$, notice that the inner product $\ang{e_j,\mS_{z(j):}^s - \mS_{b:}^s} $ is a sub-Gaussian variable with sub-Gaussian norm bounded by $m_2 \sqrt{ r^{K-1}/ p^{K-1} } \onormSize{}{ \mS_{z(i):}^s -  \mS_{b:}^s }$ with some positive constant $m_2$.
+    For $P_1$, notice that the inner product $\ang{e_j,\mS_{z(j):}^s - \mS_{b:}^s} $ is a sub-Gaussian variable with sub-Gaussian norm bounded by $C \sqrt{ r^{K-1}/ p^{K-1} } \onormSize{}{ \mS_{z(i):}^s -  \mS_{b:}^s }$.
 Then, by Chernoff bound, we have
@@
 
-    For $P_{21}$, note that the term $\mW^T_{: z(i)} \mE \mV = \frac{ \sum_{ j \neq i, j \in [p]} \ind \{ z(j) = z(i) \} e_j }{\sum_{j \in [p]} \ind \{ z(j) = z(i) \} }$  is a sub-Gaussian vector with sub-Gaussian norm bounded by $m_3 \sqrt{r^K/ p^K}$ with some positive constant $m_3$.
+    For $P_{21}$, note that the term $\mW^T_{: z(i)} \mE \mV = \frac{ \sum_{ j \neq i, j \in [p]} \ind \{ z(j) = z(i) \} e_j }{\sum_{j \in [p]} \ind \{ z(j) = z(i) \} }$  is a sub-Gaussian vector with sub-Gaussian norm bounded by $C_1 \sqrt{r^K/ p^K}$.
 This implies
@@
 
-    For $P_{22}$, the inner product $ \ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s}$ is also a sub-Gaussian variable with sub-Gaussian norm $ m_4 \sqrt{ r^{K-1}/ p^{K-1} } t$, conditioned on $\onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s } < t$ with some positive constant $m_4$.
+    For $P_{22}$, the inner product $ \ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s}$ is also a sub-Gaussian variable with sub-Gaussian norm $ C \sqrt{ r^{K-1}/ p^{K-1} } t$, conditioned on $\onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s } < t$.
 Then, by Chernoff bound, we have
@@
         &\lesssim \frac{1}{p} \sum_{i \in [p]}  \theta(i) \max_{i \in [p], b \in [r]}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2   \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \\
-        & \leq \exp \of{  - M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }, \label{eq:constant_M}
-    \end{align}
-    where $M$ is a positive constant, the first inequality follows from the constraint that $\sum_{i \in [p]} \theta(i) = p$, and the last inequality follows from~\eqref{eq:xi_p}.
+        & \lesssim  \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 },
+    \end{align}
+    where the first inequality follows from the constraint that $\sum_{i \in [p]} \theta(i) = p$, and the last inequality follows from~\eqref{eq:xi_p}.
 
@@
     \begin{equation}
-        \bbP\of{ \xi \lesssim \bbE \xi + \exp \of{  -M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } } \geq 1 -  C \exp \of{  - M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 },
-    \end{equation}
-    and thus the condition~\eqref{eq:cond_oracle} holds with probability at least $1 -  C \exp \of{  - M\frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }$ for some constant $C > 0$.
+        \bbP\of{ \xi \lesssim \bbE \xi + \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } } \geq 1 -  C_2 \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 },
+    \end{equation}
+    and thus the condition~\eqref{eq:cond_oracle} holds with probability at least $1 -  C_2 \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }$ for some constant $C_2 > 0$.
 
     Finally, we verify the bounded loss condition~\eqref{eq:cond_intial} by induction.
-With output $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{2} and the assumption SNR $\geq \tilde C p^{-K/2} \log p$, by Corollary~\ref{cor:L0}, we have
-    \begin{equation}
-        L^{(0)} \leq \frac{\bar C \Delta_{\min}^2}{\tilde C r \log p},\quad \text{    when $p$ is large enough}.
+With output $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{2} and the assumption SNR $\gtrsim p^{-K/2} \log p$, by Corollary~\ref{cor:L0}, we have
+    \begin{equation}
+        L^{(0)} \lesssim \frac{\Delta_{\min}^2}{r \log p},\quad \text{    when $p$ is large enough}.
     \end{equation}
@@
     \begin{align}
-         L^{(t_0+1)} &\leq \frac{3}{2} \xi + \frac{1}{2} L^{(t_0)} \\
-         & \leq \exp \of{  - M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } + \frac{\Delta_{\min}^2}{r \log p} \\
-         & \leq \frac{\bar C}{\tilde C}\frac{\Delta_{\min}^2}{r\log p} ,
+         L^{(t_0+1)} &\lesssim M \xi + \rho L^{(t_0)} \\
+         & \lesssim  \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } + \frac{\Delta_{\min}^2}{r \log p} \\
+         & \lesssim \frac{\Delta_{\min}^2}{r\log p} ,
     \end{align}
@@
 
-\begin{lem}[Singular-value property of membership matrices]\label{lem:membership} Under the setup of Theorem~\ref{thm:refinement}, suppose the condition~\eqref{eq:cond_intial} holds.
+\begin{lem}[Singular-value property of membership matrices]\label{lem:membership} Suppose the condition~\eqref{eq:cond_intial} holds.
 Then, for all $a \in [r]$, we have $|\of{z^{(t)}}^{-1}(a)| \asymp p/r$.
@@
 
-\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement},  assume the the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ is large enough to satisfy the inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
-We have
+\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin}, we have 
     \begin{equation}\label{eq:fib}
@@
     \begin{equation}\label{eq:gib}
-        \max_{i \in [p]} \max_{b \neq z(i)} \frac{ \of{  G_{ib}^{(t)} }^2  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2}  \leq  \frac{1}{512} \of{ \Delta_{\min}^2 +L^{(t)}},
+        \max_{i \in [p]} \max_{b \neq z(i)} \frac{ \of{  G_{ib}^{(t)} }^2  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2}  \lesssim \of{ \Delta_{\min}^2 +L^{(t)}},
     \end{equation}
@@
          & \quad \quad \quad \quad \quad + 8\of{ \ang{ \mE_{i:} (\mV - \mV^{(t)} ), [  \tilde \mS_{z(i):} ]^s - [ \tilde  \mS_{b:} ]^s  }}^2 \\
-         & \leq 8 \of{ \onormSize{}{\mE_{i:} \mV}^2 + \onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2 } \max_{a \in [r]^s} \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  } \\
+         & \lesssim \of{ \onormSize{}{\mE_{i:} \mV}^2 + \onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2 } \max_{a \in [r]^s} \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  } \\
           & \quad \quad \quad \quad \quad +\onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2  \onormSize{}{  [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }. \label{eq:fib_decomp}
@@
     \end{align}
-    where the second inequality follows from the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4} in Lemma~\ref{lem:intermediate}, the third inequality follows from the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}, and the last inequality follows from the assumption that $\Delta_{\min}^2 \geq \tilde C p^{-K/2}\log p$.
+    where the second inequality follows from the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4} in Lemma~\ref{lem:intermediate}, the third inequality follows from the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}, and the last inequality follows from the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.
 
@@
         & \leq \max_{a \in [r]} \onormSize{}{ [ \mW_{:a}^T \mY \mV^{(t)} ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^4\\
-        & \leq C^4 \frac{r^4}{\Delta_{\min}^4} (L^{(t)})^4 +  \frac{r^2 r^{4K} + p^2r^{2K+4}}{p^{2K}} \frac{(L^{(t)})^2}{\Delta_{\min}^4} \\
-        & \leq C^4 \frac{\bar C}{\tilde C^3} \of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}}, \label{eq:g1}
-    \end{align}
-    where the third inequality follows from the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:membership} and the last inequality follows from the assumption that $\Delta_{\min}^2 \geq \tilde C p^{-K/2}\log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
+        & \lesssim \frac{r^4}{\Delta_{\min}^4} (L^{(t)})^4 +  \frac{r^2 r^{4K} + p^2r^{2K+4}}{p^{2K}} \frac{(L^{(t)})^2}{\Delta_{\min}^4} \\
+        & \leq c\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}}, \label{eq:g1}
+    \end{align}
+    where the third inequality follows from the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:membership} and the last inequality follows from the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
 
@@
         & \leq \frac{2}{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2} \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
-        & \leq C' \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} \of{ \frac{r^2}{\Delta_{\min}^2} (L^{(t)})^2 + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} }\\
-        & \leq \frac{C'}{\tilde C^2} \Delta_{\min}^2 L^{(t)},\label{eq:g2}
-    \end{align}
-    where $C'$ is a positive universal constant, the second inequality follows from Lemma~\ref{lem:norm_diff}, the third inequality follows from the inequality~\eqref{eq:cond2} in Condition~\ref{cond:origin}, the inequalities~\eqref{eq:inter5} and \eqref{eq:j11_wxvt} in the proof of Lemma~\ref{lem:intermediate},  and the last inequality follows from the assumption $\Delta_{\min}^2 \geq \tilde C p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
+        & \lesssim \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} \of{ \frac{r^2}{\Delta_{\min}^2} (L^{(t)})^2 + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} }\\
+        & \leq c \Delta_{\min}^2 L^{(t)},\label{eq:g2}
+    \end{align}
+    where the second inequality follows from Lemma~\ref{lem:norm_diff}, the third inequality follows from the inequality~\eqref{eq:cond2} in Condition~\ref{cond:origin}, the inequalities~\eqref{eq:inter5} and \eqref{eq:j11_wxvt} in the proof of Lemma~\ref{lem:intermediate},  and the last inequality follows from the assumption $\Delta_{\min}^2 \geq p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
 
@@
         \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2 &\leq \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + 2\max_{a \in [r]}\onormSize{}{[\mW_{:a}^T\mX \mV^{(t)}]^s - [\mW_{:a}^T\mX \mV]^s }^2\\
-        & \leq \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + C \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2}, \label{eq:g3xv_wxv}
-    \end{align}
-    where the last inequality follows from the inequality~\eqref{eq:j1} in the proof of Lemma~\ref{lem:intermediate} and $C$ is a positive constant.
+        & \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2}, \label{eq:g3xv_wxv}
+    \end{align}
+    where the last inequality follows from the inequality~\eqref{eq:j1} in the proof of Lemma~\ref{lem:intermediate}.
    Then we have 
@@
         & \quad \times \max_{a \in [r]} \onorm{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
-        & \leq C^2 \of{\onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + C \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} } \of{ \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2}}+\frac{C'}{\tilde C^2} \Delta_{\min}^2 L^{(t)} \\
-        & \leq \frac{C^2 \bar C^2}{\tilde C} \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 (\Delta_{\min}^2 + L^{(t)}) +  \frac{C^3 C' \bar C^2}{\tilde C^2 }\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}},\label{eq:g3}
-    \end{align}
-    where the third inequality follows from the same procedure to derive~\eqref{eq:g1} and \eqref{eq:g2}, and the last inequality follows from the assumption $\Delta_{\min}^2 \geq \tilde C p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
-
-    Choose the $\tilde C$ such that
-    \begin{equation}\label{eq:tilde_c1}
-        3 \of{  C^4 \frac{\bar C}{\tilde C^3}  + \frac{C'}{\tilde C^2} + \frac{C^2 \bar C^2}{\tilde C} + \frac{C^3 C' \bar C^2}{\tilde C^2 }} \leq \frac{1}{512}.
-    \end{equation}
-    Then,  we finish the proof of inequality~\eqref{eq:gib} by plugging the inequalities~\eqref{eq:g1}, \eqref{eq:g2}, and \eqref{eq:g3} into the upper bound~\eqref{eq:gib_decomp}.
-
+        & \lesssim \of{\onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} } \of{ \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2}}+ c \Delta_{\min}^2 L^{(t)} \\
+        & \lesssim \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 (\Delta_{\min}^2 + L^{(t)}) + c\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}},\label{eq:g3}
+    \end{align}
+    where the third inequality follows from the same procedure to derive~\eqref{eq:g1} and \eqref{eq:g2}, and the last inequality follows from the assumption $\Delta_{\min}^2 \geq p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
+
+    Plugging the inequalities~\eqref{eq:g1}, \eqref{eq:g2}, and \eqref{eq:g3} into the upper bound~\eqref{eq:gib_decomp}, we finish the proof of inequality~\eqref{eq:gib}.    
     \item Upper bound for $H_{ib}^{(t)}$, i.e., the inequality~\eqref{eq:hib}. By definition of $H_{ib}$, we rearrange terms and obtain
@@
          \leq  \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} 
-        \leq \tilde C^{-2} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2\label{eq:h1},
-    \end{align}
-    following the derivation of $G_2$ in inequality~\eqref{eq:g2} and the assumption that $\Delta_{\min}^2 \geq \tilde C p^{-K/2} \log p$.
+        \lesssim  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2\label{eq:h1},
+    \end{align}
+    following the derivation of $G_2$ in inequality~\eqref{eq:g2} and the assumption that $\Delta_{\min}^2 \geq p^{-K/2} \log p$.
 
@@
     \begin{equation}\label{eq:h2}
-        |H_2| \lesssim 2\max_{a \in [r]}\onormSize{}{[\mW_{:a}^T\mX \mV^{(t)}]^s - [\mW_{:a}^T\mX \mV]^s }^2 \lesssim \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} \leq C \frac{\bar C^2}{ \tilde C^2} \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
+        |H_2| \lesssim 2 \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2 + \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
     \end{equation}
@@
     \begin{equation}\label{eq:h3}
-         |H_3| \lesssim \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \leq 2 \tilde C^{-1} \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
+         |H_3| \lesssim \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
     \end{equation}
@@
 
-    Choose $\tilde C$ such that 
-    \begin{equation}\label{eq:tilde_c2}
-        \tilde C^{-2} + C \frac{\bar C^2}{ \tilde C^2} + \tilde C^{-1} \leq \frac{1}{4}.
-    \end{equation}
-     Therefore, we finish the proof of inequality~\eqref{eq:hib} combining inequalities~\eqref{eq:h1}, \eqref{eq:h2}, and \eqref{eq:h3}.
+    Therefore, combining inequalities~\eqref{eq:h1}, \eqref{eq:h2}, and \eqref{eq:h3}, we have finish the proof of inequality~\eqref{eq:hib}.
 
@@
 
-     \begin{lem}[Relationship between misclustering loss and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement}, we have
+     \begin{lem}[Relationship between misclustering loss and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin}, we have
     \begin{equation}\label{eq:inter1}
@@
  \begin{equation}\label{eq:inter3}
-     \max_{b \in [r]} \onormSize{}{ [ \mW_{:b}^T \mY \mV ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV  ]^s } \leq C \of{\frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{r^{2K} + pr^{K+1}}{p^{K}} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}},
+     \max_{b \in [r]} \onormSize{}{ [ \mW_{:b}^T \mY \mV ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV  ]^s } \lesssim \frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{r^{2K} + pr^{K+1}}{p^{K}} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
  \end{equation}
  \begin{equation}\label{eq:inter4}
-     \max_{b \in [r]} \onormSize{}{  [  \mW_{:b}^{(t), T}  \mY \mV]^s - [  \mW_{:b}^{(t), T}  \mY \mV^{(t)}]^s } \leq C  \of{ \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} +  \frac{r L^{(t)}}{\Delta_{\min}}},
+     \max_{b \in [r]} \onormSize{}{  [  \mW_{:b}^{(t), T}  \mY \mV]^s - [  \mW_{:b}^{(t), T}  \mY \mV^{(t)}]^s } \lesssim  \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} +  \frac{r L^{(t)}}{\Delta_{\min}},
      \end{equation}
     \begin{equation}\label{eq:inter5}
-         \max_{b \in [r]} \onormSize{}{ [\mW_{:b}^T \mY \mV^{(t)} ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s } \leq C \of{\frac{rL^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}},
-    \end{equation}
-    for some positive universal constant $C$. In addition, the inequality~\eqref{eq:inter4} also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$.
+         \max_{b \in [r]} \onormSize{}{ [\mW_{:b}^T \mY \mV^{(t)} ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s } \lesssim \frac{rL^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}.
+    \end{equation}
+
+    In addition, the inequality~\eqref{eq:inter4} also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$.
     \end{lem}
@@
         & = \frac{  \onormSize{}{ [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}] \sin \alpha}}{  \onormSize{}{ D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}} } \\
-        & \leq \frac{ 1}{ C_1}  \sum_{a \neq b, a \in [r]} D_{ab} \onorm{\mS_{a:}} \onorm{ \mS^s_{b:} - \mS^s_{a:}} \\
+        & \leq \frac{ 1}{C_1}  \sum_{a \neq b, a \in [r]} D_{ab} \onorm{\mS_{a:}} \onorm{ \mS^s_{b:} - \mS^s_{a:}} \\
         & \lesssim \frac{r}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind\offf{z^{(t)}(i) = b} \onorm{ \mS^s_{b:} - \mS^s_{a:}}\\

### 2022-01-20 tensor_clustering/tensor_clustering_arxiv/tensor_clustering_arxiv.tex
package

--- 2022-01-18 tensor_clustering/tensor_clustering_ieee/ieee_v3.tex
@@
-\documentclass[lettersize,onecolumn,journal]{IEEEtran}
-\usepackage{amsmath,amsfonts}
-\usepackage{algorithm}
-\usepackage{array}
-\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-\usepackage{textcomp}
-\usepackage{stfloats}
-\usepackage{url}
-\usepackage{verbatim}
+\documentclass[11pt]{article}
 \usepackage{graphicx}
-\usepackage{cite}
-\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
-% updated with editorial comments 8/9/2021
+\usepackage{subfigure}
+\usepackage{caption}
+\usepackage[hidelinks]{hyperref}
+\usepackage[margin=1in]{geometry}
+% \geometry{
+%  left=30mm,
+%  right=30mm
+%  }
 
@@
 \usepackage{dsfont,multirow,hyperref,setspace,enumerate}
-\hypersetup{colorlinks,linkcolor=black,citecolor=black,urlcolor={black}}
+\hypersetup{colorlinks,linkcolor=blue,citecolor=blue,urlcolor={black}}
 \usepackage{lscape}
@@
 \usepackage{algpseudocode}
+
+\usepackage{textcomp}
+\usepackage{stfloats}
+\usepackage{verbatim}
+\usepackage{array}
 
@@
 
+\usepackage{appendix}
 \usepackage{fullpage} 
@@
 
+\newcommand{\distgap}{D_{\ell_2}}
 \newcommand{\Mat}{\text{Mat}}
@@
 \newcommand\Algphase[1]{%
-\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\columnwidth}{0.4pt}%
+\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\textwidth}{0.4pt}%
 \Statex\hspace*{-\algorithmicindent}\textbf{#1}%
-\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\columnwidth}{0.4pt}%
+\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\textwidth}{0.4pt}%
 }
@@
 
+\title{Multiway Spherical Clustering via Degree-Corrected\\ Tensor Block Models\footnote{Part of this paper was presented at 25th International Conference on Artificial Intelligence and Statistics (AISTATS).}}
+\date{}
+\author{%
+%\small
+Jiaxin Hu \\
+University of Wisconsin -- Madison\\
+\texttt{jhu267@wisc.edu} \\
+\and
+Miaoyan Wang \\
+University of Wisconsin -- Madison\\
+\texttt{miaoyan.wang@wisc.edu}\\
+}
+
 \begin{document}
 
-\title{Multiway Spherical Clustering via \\ Degree-Corrected Tensor Block Models}
-
-\author{ Jiaxin Hu and Miaoyan Wang\\
-University of Wisconsin - Madison
-%IEEE Publication Technology,~\IEEEmembership{Staff,~IEEE,}
-        % <-this % stops a space
-%\thanks{This paper was produced by the IEEE Publication Technology Group. They are in Piscataway, NJ.}% <-this % stops a space
-%\thanks{Manuscript received April 19, 2021; revised August 16, 2021.}
-}
-
-% The paper headers
-%\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2021}%
-%{Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}
-
-%\IEEEpubid{0000--0000/00\$00.00~\copyright~2021 IEEE}
-% Remember, if you use this you must call \IEEEpubidadjcol in the second
-% column for its text to clear the IEEEpubid mark.
-
 \maketitle
+%\doublespacing
 
@@
 
-\begin{IEEEkeywords}
-tensor clustering, degree correction, statistical-computational efficiency, human brain connectome networks
-\end{IEEEkeywords}
+\textbf{Keywords:} tensor clustering, degree correction, statistical-computational efficiency, human brain connectome networks
 
 \section{Introduction}\label{sec:intro}
-\IEEEPARstart{M}{ultiway} arrays have been widely collected in various fields including social networks \citep{anandkumar2014tensor}, neuroscience \citep{wang2017bayesian}, and computer science \citep{koniusz2016sparse}.
+Multiway arrays have been widely collected in various fields including social networks \citep{anandkumar2014tensor}, neuroscience \citep{wang2017bayesian}, and computer science \citep{koniusz2016sparse}.
 Tensors effectively represent the multiway data and serve as the foundation in higher-order data analysis.
@@
 Classical tensor block model suffers from drawbacks to model real world data in spite of the popularity.
-The key underlying assumption of block model is that all nodes in the same community are exchangeable; i.e., the nodes have no individual effects apart from the block effects.
+The key underlying assumption of block model is that all nodes in the same community are exchangeable; i.e., the nodes have no individual-specific parameters apart from the community-specific parameters.
 However, the exchangeability assumption is often non-realistic.
 Each node may contribute to the data variation by its own multiplicative effect.
+We call the unequal node-specific effects the \emph{degree heterogeneity}.
 Such degree heterogeneity appears commonly in social networks.
@@
 
-\begin{itemize}
+\begin{itemize}[wide]
     \item \textit{Block model for clustering.} Block models such as stochastic block model (SBM) and degree-corrected SBM have been widely used for matrix clustering problems.
@@
 
-    \item  \textit{Tensor block model.} The tensor block model (TBM, \cite{wang2019multiway, han2020exact, ghoshdastidar2017consistency}) is a higher-order extension of SBM, but it fails to allow degree heterogeneity.
-The Cartesian coordinates based analysis in \cite{han2020exact} is non-applicable to handle the extra complexity brought by degree heterogeneity.
-In contrast, our model addresses the degree heterogeneity, and we develop polar-coordinate based tools for the theoretical analysis.
-
-    \item \textit{Degree-corrected block model.} The hypergraph degree-corrected block model (hDCBM) proposed by \cite{ke2019community, yuan2018testing} accounts for degree heterogeneity.
-However, the hDCBM is designed only for binary observations, and the proposed spectral algorithm in \cite{ke2019community} achieves sub-optimal polynomial clustering rate in higher-order scenarios.
-In constrast, our model allows discrete and continuous entries, and achieves \emph{exponentially} fast rate in clustering tasks.
-More importantly, to our best knowledge, we are the first to provide the statistical and computational limits analyses for the degree-corrected block model in tensor clustering.
+    \item  \textit{Tensor block model.} The (non-degree) tensor block model (TBM) is a higher-order extension of SBM, and its statistical-computational properties are investigated in recent literatures~\citep{wang2019multiway, han2020exact, ghoshdastidar2017consistency}.
+Extending results from non-degree to degree-corrected model is highly challenging.
+Our dTBM parameter space is equipped with angle-based similarity and nuisance degree parameters.
+The extra complexity makes the Cartesian coordinates based analysis~\citep{han2020exact} non-applicable to our setting.
+Towards this goal, we have developed a new polar coordinates based analysis to control the model complexity.
+We also develop a new angle-based iteration algorithm to achieve optimal clustering rates \emph{without the need of estimating nuisance degree parameters}.
+
+    \item \textit{Degree-corrected block model.} The hypergraph degree-corrected block model (hDCBM) and its variant have been proposed in the literature~\citep{ke2019community, yuan2018testing}.
+For this popular model, however, the optimal statistical-computational rates remain an open problem.
+Our main contribution is to provide a sharp statistical and computational critical phase transition in dTBM literature.
+In addition, our algorithm results in a faster \emph{exponential} error rate, in contrast to the \emph{polynomial} rate in~\cite{ke2019community}.
+The original hDCBM~\citep{ke2019community} is designed for binary observations only, and we extend the model to both continuous and binary observations.
+We believe our results are novel and helpful to the community.
 See Fig~\ref{fig:phase_axis} for overview of our results.
@@
 
-Assume there exist $r$ disjoint communities among $p$ nodes, and the connection probabilities depend on the community assignments and node effects. 
+Assume there exist $r$ disjoint communities among $p$ nodes, and the connection probabilities depend on the community assignments and node-specific parameters. 
 Then, the equation~\eqref{eq:model_tensor} models $\bbE \tY$ with unknown degree heterogeneity $\mtheta$ and sub-Gaussianity parameter $\sigma^2 = 1/4$.
@@
 \State Update the block tensor $\tS^{(t)}$ via
-$\tS^{(t)} (i_1,...,i_K)= \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}(i_k) = j_k, k \in [K]\}.$
+$\tS^{(t)} (a_1,...,a_K)= \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}(i_k) = a_k, k \in [K]\}.$
 \State Calculate reduced tensor $\tY^{\text{d}} \in \bbR^{p \times r \times \cdots \times r}$ via
@@
 \begin{defn}[Locally linear stability] \label{def:stable}
-Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\epsilon)=\{\bar z\colon \ell(\bar z, z)\leq \epsilon\}$.
+Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\varepsilon)=\{\bar z\colon \ell(\bar z, z) \leq \varepsilon\}$.
 Let $\bar z\colon[p]\to [r]$ be a clustering function.
@@
 
-\begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the setup as in Theorem~\ref{thm:initial}.
-Assume the local linear stability of degree holds in the neighborhood $\tN(z,\epsilon)$ for $\epsilon = \log^{-1}p$.
-Suppose
-$r=\tO(1)$ and $\text{SNR} \geq \tilde C p^{-K/2}\log p$ for some large constant $\tilde C$.
+\begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the setup as in Theorem~\ref{thm:initial}. Suppose 
+$r=\tO(1)$ and $\text{SNR} \geq \tilde C p^{-K/2}\log p$ for some sufficiently large constant $\tilde C$.
+Assume the local linear stability of degree holds in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \geq \check{C}\log^{-1}p $ with some positive constant $\check{C}$.
 Let $z^{(t)}$ denote the $t$-th iteration output in Sub-algorithm~\hyperref[alg:main]{2} with initialization $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{1}.
@@
 In the Sub-algorithm~\hyperref[alg:main]{2}, we make the following changes: (1) Take the collection $\{z^{(0)}_k\}_{k=1}^K$ as input, and update the block tensor $\tS^{(t)}$ with the collection $\{ z_k^{(t)}\}_{k=1}^K$, $
-        \tS^{(t)}(i_1,\ldots, i_K) =  \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}_k(i_k) = j_k, k \in [K]\}$;
+        \tS^{(t)}(a_1,\ldots, a_K) =  \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}_k(i_k) = a_k, k \in [K]\}$;
 (2) Calculate reduced tensor $\tY^{\text{d}}_k$ for each mode via
@@
 \centering
-    \begin{tabularx}{\textwidth}{c *{8}{Y}}
+    \begin{tabularx}{\columnwidth}{c *{8}{Y}}
     \toprule
-    Settings & \multicolumn{2}{c}{$p = 50, \sigma^2 = 0.25$} & \multicolumn{2}{c}{$p = 50, \sigma^2 = 1$} & \multicolumn{2}{c}{$p = 80, \sigma^2 = 0.25$} & \multicolumn{2}{c}{$p = 80, \sigma^2 = 1$}\\
+    Settings & \multicolumn{2}{c}{\small  $p = 50, \sigma^2 = 0.25$} & \multicolumn{2}{c}{\small $p = 50, \sigma^2 = 1$} & \multicolumn{2}{c}{\small  $p = 80, \sigma^2 = 0.25$} & \multicolumn{2}{c}{\small  $p = 80, \sigma^2 = 1$}\\
     \cmidrule(lr){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9}
-         True cluster number $r$ & 2 & 4 & 2  & 4 & 2  & 4 & 2 & 4  \\
+         True cluster number $r$ &  \small 2 & \small 4 &  \small2  &  \small 4 &  \small2  & \small 4 &  \small 2 & \small 4  \\
          \midrule
-         Estimated cluster number $\hat r$ &  2(0)  & 3.9(0.25)& 2(0)    & 3.1(0.52) & 2(0)    & 4(0)   & 2(0)    & 3.9(0.31)   \\
+         Estimated cluster number $\hat r$ &   \small 2(0)  &  \small 3.9(0.2)&  \small 2(0)    &   \small 3.1(0.5) &  \small 2(0)    &  \small 4(0)   &  \small 2(0)    &  \small 3.9(0.3)   \\
      \bottomrule
@@
 \appendix
+\section*{Appendices}
 
@@
 
-\section*{Notation}
+\section{Notation}
 Before the proofs, we first introduce the notation used throughout the appendix and the generalized dTBM without symmetric assumptions.
@@
 \begin{equation}
-    \sum_{j \in z^{-1}(a)} \theta'_1(j) = c \sum_{j \in z^{-1}(a)} \theta_1(j) = 1,
+    \sum_{j \in z_1^{-1}(a)} \theta'_1(j) = c \sum_{j \in z_1^{-1}(a)} \theta_1(j) = 1,
 \end{equation}
@@
 
-\section*{Proof of Theorem~\ref{thm:stats}}
+\section{Proof of Theorem~\ref{thm:stats}}
 
@@
 
-\section*{Proof of Theorem~\ref{thm:comp}}
+\section{Proof of Theorem~\ref{thm:comp}}
 
@@
 
-\section*{Proof of Theorem~\ref{thm:initial}}
+\section{Proof of Theorem~\ref{thm:initial}}
 \begin{proof}[Proof of Theorem 4] We prove Theorem~\ref{thm:initial} under the symmetric dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
@@
 
-\section*{Proof of Theorem~\ref{thm:refinement}}
+\section{Proof of Theorem~\ref{thm:refinement}}
 
@@
     \end{equation}
-    Note that we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \lesssim \log^{-1}(p) $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.
-With the locally linear stability assumption, we have
+    Note that we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \leq \frac{\bar C}{\tilde C} r \log^{-1}(p) $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.
+Then,  with the locally linear stability assumption, the $\mtheta$ is $\ell^{(t)}$-locally linearly stable; i.e.,
     \begin{equation}

### 2022-01-20 tensor_clustering/package/dTBM/R/peru.R

@@ -1,0 +1,21 @@
+#' Peru Legislation data
+#' 
+#' The Peru Legislation data is obtained by preprocessing the original data in Lee et al., 2017. 
+#'
+#' The data frame "attr_data" is a 116 x 2 matrix consisting the name and party affiliation of 116 legislators
+#' in the top five parties. The legislators IDs are recorded in the row names of the matrix.
+#' 
+#' The data frame "laws_data" is a 5844 x 7 matrix recording the co-sponsorship of 116 legislators of 802 bills during the first half of 2006-2007 year.
+#' 
+#' The array "network_data" is a 116 x 116 x 116 binary tensor recording the presence of order-3 co-sponsorship among legislators based on "laws_data". 
+#' Specfically, the entry (i,j,k) is 1 if the legislators (i,j,k) have sponsored the same bill, and the entry (i,j,k) is 0  otherwise.
+#'
+#'@docType data
+#'@usage data(peru)
+#'
+#'@format  A list. Includes a 116-2 data frame named "attr_data", a 5844-7 data frame named "laws_data", and a 116-116-116 binary array named "networ_data".
+#'
+#'@keywords datasets
+#'
+#'
+"peru"

### 2022-02-18 tensor_clustering/tensor_clustering_aistats_camera/aistats_camera.tex
aistats

--- 2022-01-20 tensor_clustering/tensor_clustering_arxiv/tensor_clustering_arxiv.tex
@@
-\documentclass[11pt]{article}
-\usepackage{graphicx}
-\usepackage{subfigure}
-\usepackage{caption}
+\documentclass[twoside]{article}
+
+%\usepackage{aistats2022}
+% If your paper is accepted, change the options for the package
+% aistats2022 as follows:
+%
+\usepackage[accepted]{aistats2022}
+%
+% This option will print headings for the title of your paper and
+% headings for the authors names, plus a copyright note at the end of
+% the first column of the first page.
+
+% If you set papersize explicitly, activate the following three lines:
+\special{papersize = 8.5in, 11in}
+\setlength{\pdfpageheight}{11in}
+\setlength{\pdfpagewidth}{8.5in}
+% If you use natbib package, activate the following three lines:
+\usepackage[round,authoryear]{natbib}
+\renewcommand{\bibname}{References}
+\renewcommand{\bibsection}{\subsubsection*{\bibname}}
+
+% If you use BibTeX in apalike style, activate the following line:
+%\bibliographystyle{apalike}
+
+% page number margin
+
+%\setlength{\footskip}{10pt}
+
 \usepackage[hidelinks]{hyperref}
-\usepackage[margin=1in]{geometry}
-% \geometry{
-%  left=30mm,
-%  right=30mm
-%  }
 
@@
 \usepackage{enumerate}
-\usepackage[authoryear]{natbib}
 \usepackage{url}
@@
 \input macros.tex
+
 \newcommand*{\Scale}[2][4]{\scalebox{#1}{$#2$}}%
@@
 
-\title{Multiway Spherical Clustering via Degree-Corrected\\ Tensor Block Models\footnote{Part of this paper was presented at 25th International Conference on Artificial Intelligence and Statistics (AISTATS).}}
-\date{}
-\author{%
-%\small
-Jiaxin Hu \\
-University of Wisconsin -- Madison\\
-\texttt{jhu267@wisc.edu} \\
-\and
-Miaoyan Wang \\
-University of Wisconsin -- Madison\\
-\texttt{miaoyan.wang@wisc.edu}\\
-}
-
 \begin{document}
 
-\maketitle
-%\doublespacing
+% If your paper is accepted and the title of your paper is very long,
+% the style will print as headings an error message. Use the following
+% command to supply a shorter title of your paper so that it can be
+% used as headings.
+%
+%\runningtitle{I use this title instead because the last one was very long}
+
+% If your paper is accepted and the number of authors is large, the
+% style will print as headings an error message. Use the following
+% command to supply a shorter version of the authors names so that
+% they can be used as headings (for example, use only the surnames)
+%
+%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}
+
+\twocolumn[
+
+\aistatstitle{Multiway Spherical Clustering via Degree-Corrected\\ Tensor Block Models}
+
+\aistatsauthor{ Jiaxin Hu \And Miaoyan Wang }
+
+\aistatsaddress{ University of Wisconsin -- Madison \And  University of Wisconsin -- Madison} ]
 
@@
 Further, we develop an efficient polynomial-time algorithm that provably achieves exact clustering under mild signal conditions.
-The efficacy of our procedure is demonstrated through two data applications, one on human brain connectome project, and another on Peru Legislation network dataset.
+The efficacy of our procedure is demonstrated through both simulations and analyses of Peru Legislation dataset.
 \end{abstract}
 
-\textbf{Keywords:} tensor clustering, degree correction, statistical-computational efficiency, human brain connectome networks
-
-\section{Introduction}\label{sec:intro}
+\section{INTRODUCTION}\label{sec:intro}
 Multiway arrays have been widely collected in various fields including social networks \citep{anandkumar2014tensor}, neuroscience \citep{wang2017bayesian}, and computer science \citep{koniusz2016sparse}.
@@
 The most common model for higher-order clustering is called \emph{tensor block model} (TBM)~\citep{wang2019multiway}, which extends the usual matrix stochastic block model~\citep{abbe2017community} to tensors.
-The matrix analysis tools, however, are sub-optimal for higher-order clustering.
+Matrix analysis tools, however, are sub-optimal for higher-order clustering.
 Developing tensor tools for solving block models has received increased interest recently~\citep{ wang2019multiway,chi2020provable,han2020exact}.
@@
 Ignoring the degree heterogeneity may seriously mislead the clustering results.
-For example, regular block model fails to model the member affiliation in the Karate Club network~\citep{bickel2009nonparametric} without addressing degree heterogeneity.
+For example, regular block model fails to model the member affiliation in Karate Club network~\citep{bickel2009nonparametric} without addressing degree heterogeneity.
 
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
     \centering
-    \includegraphics[width = .9\textwidth]{intro2_arxiv.pdf}
-    \caption{Examples for order-3 tensor block model (TBM) with and without degree correction.
-Both TBM and dTBM have four communities on each mode, while dTBM allows a richer structure with degree heterogeneity.
+    \includegraphics[width = .9\columnwidth]{intro2_2.pdf}
+    \caption{ \small Examples for order-3 TBM with and without degree correction. Both TBM and dTBM have four communities on each mode, while dTBM allows a richer structure with degree heterogeneity.
     }
     \label{fig:intro}
-\end{figure*}
+\end{figure}
 
 {\bf Our contributions.} The primary goal of this paper is to provide both statistical and computational guarantees for dTBM. Our main contributions are summarized below.
-\begin{itemize}[leftmargin=*]
+\begin{itemize}[wide,topsep=-2pt,itemsep=2pt,parsep=4pt]
 
@@
     \centering
-    \includegraphics[width = 17cm]{phase.pdf}
-    \caption{SNR thresholds for statistical and computational limits in order-$K$ dTBM with dimension $(p,...,p)$ and $K \geq 2$.
+    \includegraphics[width = .9\textwidth]{phase.pdf}
+    \caption{ \small SNR thresholds for statistical and computational limits in order-$K$ dTBM with dimension $(p,...,p)$ and $K \geq 2$.
 The SNR gap between statistical possibility and computational efficiency  exists only for tensors with $K \geq 3$.
@@
 
+\thispagestyle{empty}
+
 {\bf Related work.} 
@@
 
-\begin{itemize}[wide]
-    \item \textit{Block model for clustering.} Block models such as stochastic block model (SBM) and degree-corrected SBM have been widely used for matrix clustering problems.
-The theoretical properties and algorithm performance for matrix block models have been well-studied~\citep{gao2018community}; see the review paper~\citep{abbe2017community} and the references therein.
-However, The tensor counterparts are relatively less understood.
-
-    \item  \textit{Tensor block model.} The (non-degree) tensor block model (TBM) is a higher-order extension of SBM, and its statistical-computational properties are investigated in recent literatures~\citep{wang2019multiway, han2020exact, ghoshdastidar2017consistency}.
+    \textit{Block model.} Block models such as stochastic block model (SBM) and degree-corrected SBM have been widely used for matrix clustering problems.
+See the review paper~\citep{abbe2017community} and the references therein.
+    %However, The tensor counterparts are relatively less understood.  
+    The (non-degree) tensor block model (TBM) is a higher-order extension of SBM, and its statistical-computational properties are investigated in recent literatures~\citep{wang2019multiway, han2020exact, ghoshdastidar2017consistency}.
 Extending results from non-degree to degree-corrected model is highly challenging.
@@
 
-    \item \textit{Degree-corrected block model.} The hypergraph degree-corrected block model (hDCBM) and its variant have been proposed in the literature~\citep{ke2019community, yuan2018testing}.
+     \textit{Degree-corrected block model.} The hypergraph degree-corrected block model (hDCBM) and its variant have been proposed in the literature~\citep{ke2019community, yuan2018testing}.
 For this popular model, however, the optimal statistical-computational rates remain an open problem.
@@
 We believe our results are novel and helpful to the community.
-See Fig~\ref{fig:phase_axis} for overview of our results.
-
-    \item \textit{Global-to-local algorithm strategy.} Our methods generalize the recent global-to-local strategy for matrix learning~\citep{gao2018community,chi2019nonconvex,yun2016optimal} to tensors~\citep{han2020exact,ahn2018hypergraph,kim2018stochastic}.
+See Figure~\ref{fig:phase_axis} for overview of our results.
+
+    \textit{Global-to-local algorithm strategy.} Our methods generalize the recent global-to-local strategy for matrix learning~\citep{gao2018community,chi2019nonconvex,yun2016optimal} to tensors~\citep{han2020exact,ahn2018hypergraph,kim2018stochastic}.
 Despite the conceptual similarity, we address several fundamental challenges associated with this non-convex, non-continuous problem.
@@
 
-\end{itemize}
-
 \begin{table*}[t]
 \resizebox{\textwidth}{!}{%
+\centering
     \begin{tabular}{c|ccccc}
@@
     }
-    \caption{Comparison between previous methods with our method. $^*$We list the result for order-K tensors with $K \geq 3$ and general number of communities $r = \tO(1)$. }\label{tab:comp}
+    \caption{\small Comparison between previous methods with our method. $^*$We list the result for order-K tensors with $K \geq 3$ and general number of communities $r = \tO(1)$. }\label{tab:comp}
+    \vspace{-.2cm}
 \end{table*}
+\thispagestyle{empty}
 
@@
 We use  $\onorm{\ma}=\sqrt{\sum_{i}a^2(i)}$ to denote the $\ell_2$-norm, $\onorm{\ma}_1=\sum_i |a_i|$ to denote the $\ell_1$ norm of $\ma$.
-For two vector $\ma, \mb$ of the same dimension, we denote the angle between $\ma, \mb$ by
-\begin{equation}
-    \cos \of{\ma, \mb} = \frac{\ang{ \ma, \mb}}{ \onorm{\ma} \onorm{\mb} },
-\end{equation}
-where $\ang{\ma,\mb}$ is the inner product of two vectors and $\cos \of{\ma, \mb} \in [-1,1]$. We make the convention that $\cos \of{\ma, \mb} = \cos \of{\ma^T, \mb^T}$.
-
+For two vector $\ma, \mb$ of the same dimension, we denote the angle between $\ma, \mb$ by $\cos \of{\ma, \mb} = \ang{ \ma, \mb} / \onorm{\ma} \onorm{\mb},$ where $\ang{\ma,\mb}$ is the inner product of two vectors and $\cos \of{\ma, \mb} \in [-1,1]$.
+We make the convention that $\cos \of{\ma, \mb} = \cos \of{\ma^T, \mb^T}$.
 Let $\tY  \in \bbR^{p_1 \times \cdots \times p_K}$ be an order-$K$ $(p_1,...,p_K)$-dimensional tensor.
 We use $\tY(i_1,\ldots,i_K)$ to denote the $(i_1,\ldots,i_K)$-th entry of $\tY$.
-The multilinear multiplication of a tensor $\tS\in \bbR^{r_1\times \cdots \times r_K}$ by matrices $\mM_k \in\mathbb{R}^{p_k\times r_k}$ results in an order-$d$ $(p_1,\ldots,p_K)$-dimensional tensor $\tX$, denoted
+The multilinear multiplication of a tensor $\tS\in \bbR^{r_1\times \cdots \times r_K}$ by matrices $\mM_k \in\mathbb{R}^{p_k\times r_k}$ results in an order-$d$ $(p_1,\ldots,p_K)$-dimensional tensor $\tX$
 \[
@@
 \]
-where the entries of $\tX$ are defined by
-\begin{align}
-    &\tX(i_1,\ldots, i_K)=\sum_{(j_1,\ldots,j_K)}\tS(j_1,\ldots,j_K)\mM_1(i_1,j_1)\cdots \mM_K(i_K,j_K).
-\end{align} 
+where the entries of $\tX$ are defined by $\tX(i_1,\ldots, i_K)
+    =\sum_{(j_1,\ldots,j_K)}\tS(j_1,\ldots,j_K)\mM_1(i_1,j_1)\cdots \mM_K(i_K,j_K).$
+% \small
+% \begin{align}
+%     &\tX(i_1,\ldots, i_K)\\
+%     &=\sum_{(j_1,\ldots,j_K)}\tS(j_1,\ldots,j_K)\mM_1(i_1,j_1)\cdots \mM_K(i_K,j_K).
+% \end{align} 
+% \normalsize
 For a matrix $\mY$, we use $\mY_{i:}$ (respectively, $\mY_{:i}$) to denote the $i$-th row (respectively, $i$-th column) of the matrix.
@@
 
-{\bf Organization.} The rest of this paper is organized as follows.
+% {\bf Organization.} The rest of this paper is organized as follows.
 Section~\ref{sec:model} introduces the degree-corrected tensor block model (dTBM) with three motivating examples and presents the identifiability of dTBM under the angle gap condition.
@@
 
-\section{Model formulation and motivations}\label{sec:model}
-
-\subsection{Degree-corrected tensor block model}
+\section{MODEL FORMULATION}\label{sec:model}
+
+\subsection{dTBM}
 
 Suppose we have an order-$K$ data tensor $\tY \in \bbR^{p \times \cdots \times p}$.
-For ease of notation, we focus on symmetric tensors in this section; the extension to general asymmetric tensors is provided in Section~\ref{subsec:exten}.
+For ease of notation, we focus on symmetric tensors in this section; see Supplement~\ref{sec:exten} for the extension to general asymmetric tensors.
 Assume there exist $r \geq 2$  disjoint communities among the $p$ nodes.
@@
 Let $\mtheta=(\theta(1),\ldots,\theta(p))^T$ denote the degree heterogeneity for $p$ nodes.
-We consider the order-$K$ dTBM~\citep{ghoshdastidar2017consistency,ke2019community},
+We consider the order-$K$ dTBM~\citep{ke2019community},
+\vspace{-0.2cm}
+\footnotesize
 \begin{equation}\label{eq:model_margin}
@@
 Note that the discrete nature of $\mM$ renders our model~\eqref{eq:model_tensor} more challenging than Tucker decomposition.
-We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$ if $\tY$ admits dTBM~\eqref{eq:model_tensor}.
-The goal of clustering is to estimate $z$ from a single noisy tensor $\tY$.
-We are particularly interested in the high-dimensional regime where $p$ grows whereas $r=\tO(1)$.
-
-\subsection{Motivating examples} Here, we provide four applications to illustrate the practical necessity of dTBM.
+We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$ if $\tY$ admits~\eqref{eq:model_tensor}.
+%The goal of clustering is to estimate $z$ from a single noisy tensor $\tY$. 
+We are particularly interested in high-dimensional regime where $p$ grows whereas $r=\tO(1)$. See Supplement~\ref{sec:examples} for motivated applications of dTBM.
+
+% \begin{rmk}[Comparison with non-degree models]
+% Our dTBM uses fewer block parameters than TBM.
+In particular, every non-degree $r_1$-block tensor can be represented by a \emph{degree-corrected} $r_2$-block tensor with $r_2\leq r_1$.
+In particular, there exist tensors with $r_1=p$ but $r_2=1$, so the reduction in model complexity can be dramatic from $p$ to 1.
+This fact highlights the benefits of introducing degree heterogeneity in higher-order clustering tasks.
+% \end{rmk}
+
+\subsection{Identifiability under Angle Gap}
+The goal of clustering is to estimate the partition function $z$ from model~\eqref{eq:model_tensor}.
+For ease of notation, we focus on symmetric tensors; the extension to non-symmetric tensors are similar.
+We use $\tP$ to denote the following parameter space for $(z,\tS,\mtheta)$,
+\footnotesize
+\begin{align}\label{eq:family}
+&\tP=\bigg\{(z,\tS,\mtheta)\colon  \mtheta\in\mathbb{R}^p_{+},\text{for }a\in[r], 
+{c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r},\notag \\
+&c_3\leq \onormSize{}{\text{Mat}(\tS)_{a:}}\leq c_4,
+\onorm{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)| \bigg\},
+\end{align}
+\normalsize 
+where $c_i>0$'s are universal constants. We briefly describe the rationale of the constraints in~\eqref{eq:family}. 
+First, the entrywise positivity constraint on  $\mtheta\in\mathbb{R}^p_{+}$ is imposed to avoid sign ambiguity between entries in $\mtheta_{z^{-1}(a)}$ and $\tS$.
+This constraint allows the trigonometric $\cos$ to describe the angle similarity in the Assumption~\ref{assmp:min_gap} below and Sub-algorithm~\hyperref[alg:main]{2} in Section~\ref{sec:alg}.
+Note that the positivity constraint can be achieved without sacrificing model flexibility, by using a slightly larger dimension of $\tS$ in the factorization~\eqref{eq:model_tensor}; see Supplement~\ref{subsec:positive}.
+Second, recall that the quantity $|z^{-1}(a)|$ denotes the number of nodes in $a$-th community.
+The constants $c_1, c_2$ in the $|z^{-1}(a)|$ bound assume the roughly balanced size across $r$ communities.
+Third, the constants $c_3, c_4 $ in the magnitude of $\text{Mat}(\tS)_{a:}$ requires no purely zero slide in $\tS$, so the core tensor $\tS$ is not trivially reduced to a lower rank.
+Lastly, the $\ell_1$ normalization $\onormSize{}{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|$ is imposed to avoid the scalar ambiguity between $\mtheta_{z^{-1}(a)}$ and $\tS$.
+This constraint, again, incurs no restriction to model flexibility but makes our presentation cleaner.
+See Supplement~\ref{subsec:comp_para} for the parameter space comparison with previous works.
+
+We now provide the identifiability conditions for our model before estimation procedures.
+When $r=1$, the decomposition~\eqref{eq:model_tensor} is always unique (up to cluster label permutation) in $\tP$, because dTBM is equivalent to the rank-1 tensor family under this case.
+When $r\geq 2$, the Tucker rank of signal tensor $\bbE\tY$ in~\eqref{eq:model_tensor} is bounded by, but not necessarily equal to, the number of blocks $r$~\citep{wang2019multiway}.
+Therefore, one can not apply the classical identifiability conditions for low-rank tensors to dTBM.
+Here, we introduce a key separation condition on the core tensor.
+
+\begin{assumption}[Angle gap] \label{assmp:min_gap}Let $\mS = \text{Mat}(\tS)$. Assume the minimal gap between normalized rows of $\mS$ is bounded away from zero; i.e., for $r \geq 2$,
+\begin{equation}\label{eq:minimal_gap}
+    \Delta_{\min} := \min_{a \neq b \in [r]} \onorm{\frac{\mS_{a:}}{\onormSize{}{ \mS_{a:}}} - \frac{\mS_{b:}}{\onormSize{}{ \mS_{b:}}} }>0.
+\end{equation}
+\end{assumption}
+We make the convention $\Delta_{\min} = 1$ for $r = 1$.
+Equivalently, \eqref{eq:minimal_gap} says that none of the two rows in $\mS$ are parallel; i.e., $\max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}}  = 1-\Delta^2_{\min}/2<1$.
+The quantity $\Delta_{\min}$ characterizes the non-redundancy among clusters measured by angle separation.
+Denominators involved in definition~\eqref{eq:minimal_gap} are well posed because of the lower bound on $\onorm{\mS_{a:}}$ in~\eqref{eq:family}.
+Following theorem shows that the angle separation is sufficient and necessary for the identifiability of dTBM.
+
+% Our first main result is the following theorem showing the sufficiency and necessarily of the angle gap separation condition for the parameter identifiability under dTBM.
+
+\begin{thm}[Model identifiability]\label{thm:unique} Consider the dTBM with $r\geq 2$.
+The parameterization~\eqref{eq:model_tensor} is unique in $\tP$ up to cluster label permutations, if and only if Assumption~\ref{assmp:min_gap} holds.
+\end{thm}
+
+The identifiability guarantee for the dTBM is more appealing than classical Tucker model.
+In the Tucker model, the factor matrix $\mM$ is identifiable only up to orthogonal rotations.
+In contrast, our model does not suffer from rotational invariance.
+%As we will show in Section~\ref{sec:alg}, each column of the membership matrix $\mM$ can be precisely recovered under our algorithm. 
+This property benefits the interpretation of dTBM in practice. 
+\thispagestyle{empty}
+\section{THEORETICAL LIMITS }\label{sec:limits}
+
+In this section, we study the statistical and computational limits of dTBM. We propose signal-to-noise ratio~(SNR),
+\begin{equation}\label{eq:gamma}
+     \text{SNR}:= \Delta^2_{\min}/\sigma^2 = p^{\gamma}, 
+\end{equation}
+with varying $\gamma \in \bbR$ that quantifies different regimes of interest.
+We call $\gamma$ the \emph{signal exponent}.
+Intuitively, a larger SNR, or equivalently a larger $\gamma$, benefits the clustering in the presence of noise.
+With quantification~\eqref{eq:gamma}, consider following parameter space,
+\small
+\begin{equation}\label{eq:gammafamily}
+    \tP(\gamma) = \tP\cap\{\tS \text{ satisfies SNR condition~\eqref{eq:gamma} with $\gamma$} \}.
+\end{equation}
+\normalsize
+The $1$-block dTBM does not belong to the space $\tP(\gamma)$ when $\gamma < 0$ by Assumption~\ref{assmp:min_gap}.
+Our goal is to characterize the clustering accuracy with respect to $\gamma$.
+Let $\hat z$ and $z$ be the estimated and true clustering functions in~\eqref{eq:family}.
+Define the misclustering error
+\[
+\ell(\hat z, z)={1\over p}\min_{\pi \in \Pi}\sum_{i\in[p]}\ind\{\hat z(i)\neq \pi \circ z(i)\},
+\]
+where $\pi: [r] \mapsto [r]$ is a permutation of cluster labels, $\circ$ denotes the composition operation, and $\Pi$ denotes the collection of all possible permutations.
+The infinitum over all permutations accounts for the ambiguity in cluster label permutation.
+
+In Sections~\ref{sec:statlimit} and \ref{sec:complimit}, we provide the lower bounds of $\ell (\hat z, z)$ for general Gaussian dTBMs~\eqref{eq:model_tensor} without symmetric assumptions.
+For general (asymmetric) Gaussian dTBMs, we assume Gaussian noise $\tE(i_1,\ldots,i_K)\stackrel{\text{i.i.d.}}{\sim} N(0,\sigma^2)$, and we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode.
+For notational simplicity, we still use $z$ and $\tP(\gamma)$ for this general (asymmetric) model.
+All lower bounds should be interpreted as the worst-case results across $K$ modes.
+
+\thispagestyle{empty}
+\subsection{Statistical Critical Values}\label{sec:statlimit}
+The statistical limit means the minimal SNR required for solving dTBMs with \emph{unlimited computational cost.} Our following result shows the minimax lower bound of SNR for exact recovery in dTBM.
+
+\begin{thm}[Statistical lower bound]\label{thm:stats} Consider general Gaussian dTBMs under the parameter space $\tP(\gamma)$ with $K\geq 1$.
+Assume $r \lesssim p^{1/3}$.
+If the signal exponent satisfies $\gamma < -(K-1)$, then, every estimator $\hat z_{\text{stat}}$ obeys
+\vspace{-.2cm}
+\begin{equation}
+    \sup_{(z, \tS, \mtheta) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
+\end{equation}
+\end{thm}
+\vspace{-.3cm}
+Theorem~\ref{thm:stats} demonstrates the impossibility of exact recovery of the assignment when $\gamma < -(K-1)$ in the high-dimensional regime $p\to \infty$ for fixed $r$.
+The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to, maximum likelihood estimation (MLE)~\citep{wang2019multiway} and trace maximization~\citep{ghoshdastidar2017uniform}.
+As we will show in Section~\ref{sec:alg}, the SNR threshold $-(K-1)$ is also a minimax upper bound, because MLE achieves exact recovery when $\gamma > -(K-1)$.
+Hence, the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM.
+
+\subsection{Computational Critical Values}\label{sec:complimit}
+%In this section, we derive the computational limits of dTBMs. 
+The computational limit means the minimal SNR required for exactly recovery with \emph{polynomial-time} computational cost.
+An important ingredient to establish the computational limits is the \emph{hypergraphic planted clique (HPC) conjecture} \citep{zhang2018tensor, brennan2020reducibility}.
+The HPC conjecture indicates the impossibility of fully recovering the planted cliques with polynomial-time algorithm when the clique size is less than the number of vertices in the hypergraph.
+The formal statement of HPC detection conjecture is provided in Supplement~\ref{subsec:thm_comp}.
+Under the HPC conjecture, we establish the SNR lower bound that is necessary for any \emph{polynomial-time} estimator to achieve exact clustering.
+
+\begin{thm}[Computational lower bound]\label{thm:comp} Consider general Gaussian dTBMs under the parameter space $\tP(\gamma)$ with $K\geq 2$.
+Assume HPC conjecture holds.
+If the signal exponent $\gamma < -K/2$, then, every \emph{polynomial-time estimator} $\hat z_{\text{comp}}$ obeys
+\begin{equation}
+     \liminf_{p\to \infty}\sup_{(z, \tS, \mtheta) \in \tP(\gamma)}  \bbE \left[ p\ell(\hat z_{\text{comp}}, z) \right]\geq 1.
+\end{equation}
+\vspace{-.3cm}
+\end{thm}
+Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$.
+Therefore, $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM.
+In Section~\ref{sec:alg}, we will show the condition $\gamma> -K/2$ suffices for our proposed polynomial-time estimator.
+Thus, $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM.
+
+\begin{rmk}[Statistical-computational gaps]
+Now, we have established the phase transition of exact clustering under order-$K$ dTBM by combing Theorems~\ref{thm:stats} and \ref{thm:comp}.
+Figure~\ref{fig:phase_axis} summarizes our results of critical SNRs when $K \geq 2$.
+In the weak SNR region $\gamma < -(K-1)$, no statistical estimator succeeds in degree-corrected higher-order clustering.
+In the strong SNR region $\gamma  > -K/2$, our proposed algorithm precisely recovers the clustering in polynomial time.
+In the moderate SNR regime, $-(K-1)\leq \gamma \leq -K/2$, the degree-corrected clustering problem is statistically easy but computationally hard.
+Particularly, dTBM reduces to matrix degree-corrected model when $K =2$, and the statistical and computational bounds show the same critical value.
+When $K =1$, dTBM reduces to the degree-corrected sub-Gaussian mixture model (GMM).
+% with model
+% \begin{equation}
+%     \mY = \mTheta \mM \mS + \mE,
+% \end{equation}
+% where $\mY \in \bbR^{p \times d}$ collects $n$ data points in $\bbR^d$, $\mS \in \bbR^{r \times d}$ collects the $d$-dimensional centroids for $r$ clusters, and $\mTheta \in \bbR^{p \times p}, \mM \in \{0,1\}^{p \times r}, \mE \in \bbR^{p \times d}$ have the same meaning as in dTBM.
+\cite{lu2016statistical} implies that polynomial-time algorithms are able to achieve the statistical minimax lower bound in GMM.
+Hence, the statistical-to-computational gap emerges only for higher-order tensors with $K \geq 3$, which reveals the intrinsic distinctions among (vector) one-dimensional clustering, (matrix) biclustering, and (tensor) higher-order clustering.
+%Besides, we find that the extra complexity from $\mtheta$ does not render the estimation of $z$ qualitatively harder; see comparison of our phase transition with non-degree TBM~\citep{han2020exact}. 
+\end{rmk}
+
+\begin{rmk}[Comparison with non-degree models]
+We compare our results to non-degree tensor models.
+The allowance of degree heterogeneity $\mtheta$ makes the model more flexible, but it incurs extra statistical and computational complexity.
+Fortunately, we find that the extra complexity does not render the estimation of $z$ qualitatively harder; see comparison of our phase transition with non-degree TBM~\citep{han2020exact}.
+\end{rmk}
+
+\section{ALGORITHM}\label{sec:alg}
+In this section, we present an efficient polynomial-time clustering algorithm under mild SNR.
+The procedure takes a global-to-local approach.
+See Figure~\ref{fig:demo} for illustration.
+The global step finds the basin of attraction with polynomial miclustering error, whereas the local iterations improve the initial clustering to exact recovery.
+Both steps are critical to obtain a satisfactory algorithm output.
+We use the symmetric tensor with i.i.d.\ sub-Gaussian noise to describe the algorithm and develop the theoretical guarantees.
+Extensions for asymmetric tensor and Bernoulli observation and practical issues are in Supplement~\ref{sec:exten}.
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=.9\columnwidth]{alg_demo.pdf}
+\caption{ \small Illustration of our global-to-local algorithm.}\label{fig:demo}
+\end{figure}
+
+\subsection{Initialization}
+
+We start with weighted higher-order clustering algorithm as initialization.
+We take an order-3 symmetric tensor as illustration for insight.
+Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
+By model~\eqref{eq:model_tensor}, for all $i \in [p]$, we have
+\begin{equation}\label{eq:kmeans}
+    \theta(i)^{-1} \mX_{i:} = \off{\text{Mat}( \tS \times_2 \mTheta \mM \times_3  \mTheta \mM )}_{z(i):}. 
+\end{equation}
+This implies that, all node $i$ belonging to $a$-th community (i.e., $z(i)=a$) share the same normalized mean vector $\theta(i)^{-1} \mX_{i:}$, and vice versa.
+Intuitively, one can apply $k$-means clustering to the vectors $\{ \theta(i)^{-1} \mX_{i:} \}_{i\in[p]}$, which leads to main idea of our Sub-algorithm~\hyperref[alg:main]{1}.
+Specifically, our initialization consists of denoising step and clustering step.
+The denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) estimates $\tX$ from $\tY$ by a double projection spectral method.
+% The first projection performs HOSVD~\citep{de2000multilinear} via $\mU_{\text{pre}} = \text{SVD}_{r} \of{ \text{Mat}(\tY) }$, where $\text{SVD}_r(\cdot)$ returns the top-$r$ left singular vectors.
+The second projection performs HOSVD on the projected $\tY$ onto the multilinear Kronecker space $\mU_{\text{pre}} \otimes  \mU_{\text{pre}}$; i.e.,
+% \begin{equation}\label{eq:two-step_factor}
+%     \hat \mU = \text{SVD}_{r} \of{\text{Mat}\of{ \tY\times_1  \mU_{\text{pre}} \mU_{\text{pre}}^T \times_2   \mU_{\text{pre}} \mU_{\text{pre}}^T }}.
+% \end{equation}
+% The final denoised tensor $\hat \tX$ is defined by
+% \begin{equation}\label{eq:two-step_est}
+%     \hat \tX = \tY \times_1 \hat \mU \hat 
+% \mU^T \times_2 \hat \mU \hat \mU^T \times_3 \hat \mU \hat \mU^T. 
+% \end{equation}
+The double projection improves usual matrix spectral methods in order to alleviate the noise effects for $K\geq 3$~\citep{han2020exact}.
+The clustering step (lines 3-5 in Sub-algorithm~\hyperref[alg:main]{1}) performs the weighted $k$-means clustering. 
+% We write $\hat \mX=\text{Mat}(\hat \tX)$, and normalize the rows into $\hat \mX^s_{i:}=\onormSize{}{\hat \mX_{i:}}^{-1}\hat \mX_{i:}$ as a surrogate of $\theta(i)^{-1} \mX_{i:}$.
+Then, a weighted $k$-means clustering is performed on the normalized rows with weights equal to $\onormSize{}{\hat \mX_{i:}}^2$.
+The choice of weights is to bound the $k$-means objective function by the Frobenius-norm accuracy of $\hat \tX$.
+Unlike existing clustering algorithm~\citep{ke2019community}, we apply the clustering on the unfolded tensor $\hat \mX$ rather than on the factors $\hat \mU$.
+This strategy relaxes the singular-value gap condition~\citep{gao2018community, han2020exact}.
+% We assign degenerate rows with purely zero entries to an arbitrarily random cluster; these nodes are negligible in high-dimensions because of the lower bound on $\onormSize{}{\text{Mat}(\tS)_{a:}}$ in~\eqref{eq:family}.
+The final result gives the initial cluster assignment $\hat z^{(0)}$.
+Full procedures are provided in Sub-algorithm~\hyperref[alg:main]{1}.
+
+\begin{figure*}[h]
+    \centering
+    \vspace{-.5cm}
+    \includegraphics[width=1\textwidth]{algorithm.pdf}
+  \vspace{-1cm}
+    \label{alg:main}
+\end{figure*}
+
+% \begin{algorithm}[h!]
+% \caption*{\bf Algorithm: Multiway spherical clustering for degree-corrected tensor block model }
+% \vspace{.15cm}
+% \begin{algorithmic}[1]
+% \Algphase{Sub-algorithm 1: Weighted higher-order initialization}
+% \INPUT Observation $\tY \in \bbR^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
+% \State Compute factor matrix $ \mU_{\text{pre}} = \text{SVD}_{r} (\text{Mat}(\tY))$ and the $(K-1)$-mode projection $
+% \tX_{\text{pre}} = \tY \times_1   \mU_{\text{pre}} \mU_{\text{pre}}^T \times_2 \cdots \times_{K-1}  \mU_{\text{pre}} \mU_{\text{pre}}^T.$
+% \State Compute factor matrix $\hat \mU = \text{SVD}_{r}(\text{Mat}(\tX_{\text{pre}}))$ and denoised tensor
+% $\hat \tX = \tY \times_1 \hat \mU \hat \mU^T \times_2 \cdots \times_K \hat \mU \hat \mU^T.$
+% \State {Let $\hat \mX = \text{Mat}(\hat \tX)$ and $S_0=\{i \in [p]: \onormSize{}{\hat \mX_{i:}} = 0\}$. Set $\hat z(i)$ randomly in $[r]$ for $i \in S_0$.}
+% \State{For all $i\in S_0^c$, compute normalized rows
+% $\hat \mX_{i:}^s :=\onormSize{}{\hat \mX_{i:}}^{-1} \hat \mX_{i:}.$
+% }
+% \State {Solve the clustering $\hat z\colon [p]\to[r]$ and centroids $ (\hat \mx_j)_{j\in[r_k]}$ using weighted $k$-means, such that}
+% \begin{align}
+%     &\sum_{i \in S_0^c }  \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{\hat z(i)} }^2 
+%     \leq 
+%     \eta \min_{\substack{\bar \mx_j, j\in[r], \bar z(i),i\in S_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\bar z(i)}}^2.
+% \end{align}
+% \OUTPUT Initial clustering $z^{(0)} \leftarrow \hat z$.
+
+% \Algphase{Sub-algorithm 2: Angle-based iteration}
+% \INPUT Observation $\tY \in \bbR^{p \times \cdots \times p}$, initialization $z^{(0)} \colon [p]\to[r]$ from Sub-algorithm 1, iteration number $T$.
+% \For {$t = 0$ to $T-1$}
+% \State Update the block tensor $\tS^{(t)}$ via
+% $\tS^{(t)} (a_1,...,a_K)= \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}(i_k) = a_k, k \in [K]\}.$
+% \State Calculate reduced tensor $\tY^{\text{d}} \in \bbR^{p \times r \times \cdots \times r}$ via
+% \begin{align}
+%     &\tY^{\text{d}}(i,a_2,\ldots,a_K) 
+%     = 
+% \text{Ave}\{\tY(i,i_2,\ldots,i_K): z^{(t)}(i_k) = a_k, k \neq 1 \}.
+% \end{align}
+
+% \State Let $\mY^{\text{d}} = \text{Mat}(\tY^{\text{d}})$ and $J_0 = \{ i\in[p]: \onorm{\mY^{\text{d}}_{i:}} = 0\}$. Set $z^{(t+1)}(i)$ randomly in $[r]$ for $i \in J_0$.
+
+% \State Let $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. For all $i \in J_0^c$ update the cluster assignment by
+% \begin{equation}
+%     z(i)^{(t+1)} = \argmax_{a \in [r]} \cos \left( \mY^{\text{d}}_{i:},\ \mS^{(t)}_{a:} \right).
+% \end{equation}
+
+% \EndFor
+
+% \OUTPUT Estimated clustering $z^{(T)}  \in [r]^{p}$.
+
+% \end{algorithmic}
+% \end{algorithm}
+
+We now establish the misclustering error rate of initialization. We call $\mtheta$ is balanced if the relative extent of heterogeneity is comparable across clusters
+\begin{equation}\label{eq:degree}
+{\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
+\end{equation}
+Note that, the assumption~\eqref{eq:degree} does not preclude degree heterogeneity.
+Indeed, within each of the clusters, the highest degree can be $\theta(i) = \Omega(p)$, whereas the lowest degree can be $\theta(i)=\tO(1)$.
+
+\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with i.i.d.\ noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
+Assume $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$.
+Let $ z^{(0)}$ denote the output of Sub-algorithm~\hyperref[alg:main]{1}.
+With probability going to 1, we have
+\begin{equation}\label{eq:ini}
+   \ell(z^{(0)}, z) \lesssim {r^K p^{-K/2} / \text{SNR}}. 
+\end{equation}
+\end{thm}
+
+\begin{rmk}[Comparison to previous results] For fixed SNR, our initialization error rate with $K=2$ agrees with the initialization error rate $\tO(p^{-1})$ in matrix models~\citep{gao2018community}.
+Furthermore, in the special case of non-degree TBMs with $\theta_1=\cdots=\theta_p=1$, we achieve the same initial misclustering error $\tO(p^{-K/2})$ as in non-degree models~\citep{han2020exact}.
+Theorem~\ref{thm:initial} implies the advantage of our algorithm in achieving both accuracy and model flexibility.
+\end{rmk}
+
+\thispagestyle{empty}
+
+\begin{rmk}[Failure of conventional tensor HOSVD] If we use conventional HOSVD for tensor denoising; that is, we use $\mU_{\text{pre}}$ in place of $\hat \mU$ in line 2, then the misclustering rate becomes $\tO(p^{-1})$ for all $K\geq 2$.
+This rate is substantially worse than our rate~\eqref{eq:ini}.
+\end{rmk}
+
+% \begin{rmk}[Singular-value gap-free clustering] Note that our clustering directly applies to the estimated mean tensor $\hat \tX$ rather than the leading tensor factors $\hat \mU$.
+Applying clustering to the tensor factors suffers from the non-identifiability issue due to the infinitely many orthogonal rotations when the number of blocks $r \geq 3$ in the absence of singular-value gaps.
+% Such ambiguity causes the trouble for effective clustering~\citep{abbe2020entrywise}.
+In contrast, our initialization algorithm applies the clustering to the overall mean tensor $\hat \tX$.
+This strategy avoids the non-identifiability issue regardless of the number of blocks and singular-value gaps.
+% \end{rmk}
+
+\subsection{Angle-based Iteration}
+Our Theorem~\ref{thm:initial} has shown the polynomially decaying error rate from our initialization.
+Now we improve the error rate to exponential decay using local iterations.
+We propose an angle-based local iteration to improve the outputs from Sub-algorithm~\hyperref[alg:main]{1}.
+To gain the intuition, consider an one-dimensional degree-corrected clustering problem with data vectors $\mx_i = \theta(i) \ms_{z(i)} + \mepsilon_i, i \in [p]$, where $\ms_i$'s are known cluster centroids, $\theta(i)$'s are unknown positive degrees, and $z\colon [p] \mapsto [r]$ is the cluster assignment of interest.
+The angle-based $k$-means algorithm estimates the assignment $z$ by minimizing the angle between data vectors and centroids; i.e.,
+ \begin{equation}\label{eq:angle_kmeans}
+     z(i) = \argmax_{a \in [r]} \cos ( \mx_i,\ \ms_{a} ),\ \text{ for all }i \in [p].
+ \end{equation}
+Classical Euclidean-distance based clustering~\citep{han2020exact} fails to recover $z$ in the presence of degree heterogeneity, even under noiseless case.
+In contrast, the angle-based $k$-means achieves accurate recovery without explicit estimation of $\mtheta$.
+Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as angle-based $k$-means, except that we use estimated centroids $\ms^{(t)}_a$ in place of $\ms_a$ based on estimated assignment in previous iterations.
+% Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as in angle-based $k$-means.
+We still take the order-3 tensor for illustration.
+Specifically, Sub-algorithm~\hyperref[alg:main]{2} updates estimated core tensor and cluster assignment in each iteration.
+We use superscript $\cdot^{(t)}$ to denote the estimate from $t$-th iteration, where $t=1,\ldots.$ For core tensor, we consider the following update strategy
+% \small
+% \begin{equation}
+%      \tS^{(t)}(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k\in[3]\}.
+% \end{equation}
+% \normalsize
+% Intuitively, $\tS^{(t)}$ becomes closer to the true core $\tS$ as $z^{(t)}$ is more precise.
+For cluster assignment, we first aggregate the slices of $\tY$ and obtain a reduced tensor $\tY^{\text{d}}\in \bbR^{p \times r \times r}$ with given $z^{(t)}$, where
+% \[
+% \tY^{\text{d}}(i,a_2,a_3)=\text{Ave}\{\tY(i,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k \neq 1\}.
+% \]
+% We use $\mY^d$ and $ \mS^{(t)}$ to denote the $\text{Mat}(\tY^{\text{d}})$ and $\text{Mat}(\tS^{(t)})$.
+The rows $\mY^d_{i:}$ and $\mS^{(t)}_{a:}$ correspond to the $\mx_i$ and $\ms_a$ in the one-dimensional clustering~\eqref{eq:angle_kmeans}.
+Then, we obtain the updated assignment by
+% \[
+% z(i)^{(t+1)} = \argmax_{a \in [r]} \cos \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} },\ \text{ for all } i \in [p],
+% \]
+% provided that $\mS^{(t)}_{a:}$ is a non-zero vector. Otherwise, if $\mS^{(t)}_{a:}$ is a zero vector, then we make the convention to assign $z^{(t+1)}(i)$ randomly in $[p]$. 
+See Sub-algorithm~\hyperref[alg:main]{2} for full procedures.
+
+We now establish the misclustering error rate of iterations under the stability assumption.
+
+\thispagestyle{empty}
+
+\begin{defn}[Locally linear stability] \label{def:stable}
+Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\varepsilon)=\{\bar z\colon \ell(\bar z, z) \leq \varepsilon\}$.
+Let $\bar z\colon[p]\to [r]$ be a clustering function.
+The degree is $\varepsilon$-locally linearly stable if and only if
+\begin{equation}\label{eq:local}
+    \sin(\mp(\bar z),\ \mp_{\mtheta}(\bar z))\lesssim \varepsilon \Delta_{\min} ,\text{ for all } \bar z\in \tN(z, \varepsilon),
+\end{equation}
+where $\mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T$ and $\mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T$.
+% We define two vectors associated with $\bar z$,
+% \small
+% \begin{align}
+%     &\mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T, \\
+%     &\mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
+% \end{align}
+% \normalsize
+\end{defn}
+Roughly speaking, the vector $\mp(\bar z)$ represents the raw cluster sizes, and $\mp_{\theta}(\bar z)$ represents the relative cluster sizes weighted by degrees. 
+The local stability holds trivially for $\varepsilon=0$ based on the construction of parameter space~\eqref{eq:family}.
+The condition~\eqref{eq:local} controls the impact of node degree to the $\mp_{\theta}(\cdot)$ with respect to the misclassification rate $\varepsilon$ and angle gap.
+
+\begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the setup as in Theorem~\ref{thm:initial}. Suppose 
+$r=\tO(1)$ and $\text{SNR} \geq \tilde C p^{-K/2}\log p$ for some sufficiently large constant $\tilde C$.
+Assume the local linear stability of degree holds in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \geq \check{C}\log^{-1}p $ with some positive constant $\check{C}$.
+Let $z^{(t)}$ denote the $t$-th iteration output in Sub-algorithm~\hyperref[alg:main]{2} with initialization $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{1}.
+With probability going to 1, there exists a contraction parameter $\rho \in (0,1)$ such that
+{\footnotesize
+\vspace{-.1cm}
+\begin{equation}\label{eq:final}
+     \ell(z, \hat z^{(t+1)}) \lesssim  \ \KeepStyleUnderBrace{
+   \text{SNR}^{-1}
+    \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}}_{\substack{\text{statistical error}} }+ \KeepStyleUnderBrace{ \rho^t \ell(z, z^{(0)}). }_{\substack{\text{computational}\\ \text{error}}}
+\end{equation}
+\vspace{-.4cm}
+}
+\end{thm}
+The iteration error is decomposed into two parts: statistical error and computational error.
+The statistical error is unavoidable with noisy data regardless $t$, whereas the computational error decays in an exponential rate as the number of iterations $t \rightarrow \infty$.
+Theorem~\ref{thm:refinement} implies that, with probability going to 1, our estimate $z^{(T)}$ achieves exact recovery within polynomial iterations; more precisely,
+\vspace{-.1cm}
+\begin{equation}
+     z^{(T)} = \pi \circ z, \quad \text{for all }T\gtrsim \log_{1/\rho} p,
+\end{equation}
+for some permutation $\pi \in \Pi$.
+Hence, our combined algorithm is \textit{computationally efficient} as long as SNR $\gtrsim p^{-K/2} \log p$.
+Note that, ignoring the logarithmic term, the minimal SNR requirement, $p^{-K/2}$, coincides with the computational lower bound in Theorem~\ref{thm:comp}.
+Therefore, our algorithm is optimal regarding the signal requirement and lies in the sharpest \emph{computationally efficient} regime in Figure~\ref{fig:phase_axis}.
+
+\thispagestyle{empty}
+%\subsection{Extensions \& Practical Issues}~\label{subsec:exten}
+
+\section{NUMERICAL STUDIES}\label{sec:simulation}
+
+ We evaluate the performance of our algorithm in this section.
+We report average errors and standard deviations across 30 replications in each experiment.
+Clustering accuracy is assessed by clustering error rate (CER, i.e., one minus rand index).
+Note that CER between $(\hat z, z)$ is equivalent to misclustering error $\ell(\hat z, z)$ up to constant multiplications~\citep{meilua2012local}, and lower CER indicates better performance.
+Extra simulation results and real data analysis for human brain connectome data are in Supplement~\ref{sec:extra_exp}.
+
+We generate order-3 tensors with \emph{assortative}~\citep{gao2018community} core tensors to control SNR; i.e., we set $\tS_{aaa} = s_1$ for $a \in [r]$ and others be $s_2$, where $s_1 > s_2 > 0$.
+Let $\alpha = s_1/s_2$.
+We set $\alpha$ close to 1 such that $1-\alpha=o(p)$.
+In particular, we have $\alpha = 1 + \Omega(p^{\gamma/2})$ with $\gamma<0$ by Assumption~\ref{assmp:min_gap} and definition~\eqref{eq:gamma}.
+Hence, we easily adjust SNR via varying $\alpha$.
+The assortative setting is proposed for simulations, and our algorithm is applicable for general tensors in practice.
+The cluster assignment $z$ is randomly generated with equal probability across  $r$ clusters for each mode.
+Without further explanation, we generate degree heterogeneity $\mtheta$ from absolute normal distribution by $\theta(i) = |X_i| + 1 - 1/\sqrt{2\pi}$ with $|X_i| \stackrel{\text{i.i.d.}}\sim N(0,1), i \in [p]$ and normalize $\mtheta$ to satisfy \eqref{eq:family}.
+We set $\sigma^2 = 1$ for Gaussian data.
+
+\subsection{Verification of Theoretical Results}\label{subsec:num_theory}
+
+The first experiment verifies  statistical-computational gap described in Section~\ref{sec:limits}.
+Consider the Gaussian model with $p = \{80, 100\}$, $r = 5$.
+We vary $\gamma $ in $ [-1.2, -0.4]$ and $[-2.1, -1.4]$ for matrix ($K=2$) and tensor $(K = 3)$ clustering, respectively.
+Note that finding MLE under dTBM is computationally intractable.
+We approximate MLE using an oracle estimator, i.e., the output of Sub-algorithm~\hyperref[alg:main]{2} initialized from true assignment.
+Figure~\ref{fig:phase}a shows that both our algorithm and oracle estimator start to decrease around the critical value $\gamma_{\text{stat}}  = \gamma_{\text{comp}}  = -1$ in matrix case.
+In contrast, Figure~\ref{fig:phase}b shows a significant gap in the phase transitions between the algorithm estimator and oracle estimator in tensor case.
+The oracle error rapidly decreases to 0 when $\gamma_{\text{stat}} = -2$, whereas the algorithm estimator tends to achieve exact clustering when $\gamma_{\text{comp}} = -1.5$.
+Figure~\ref{fig:phase} confirms the existence of the statistical-computational gap in our Theorems~\ref{thm:stats} and~\ref{thm:comp}.
+\vspace{-.2cm}
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=.95\columnwidth]{phase_anno.pdf}
+    \caption{ \footnotesize SNR phase transitions for clustering in dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.
+    }
+    \label{fig:phase}
+    \vspace{-.5cm}
+\end{figure}
+
+\vspace{-.3cm}
+\begin{figure}[htp!]
+    \centering
+     \includegraphics[width=.95\columnwidth]{ini_re_anno.pdf}
+    \caption{ \footnotesize CER versus signal exponent $(\gamma)$ for initialization only and for combined algorithm.
+We set $p = \{80, 100\}, r = 5, \gamma \in [-2.1, -1.4]$ under (a) Gaussian models and (b) Bernoulli models.
+}
+    \label{fig:ini_re}
+    \vspace{-.4cm}
+\end{figure}
+
+The second experiment verifies the performance guarantees of two algorithms: (i) weighted higher-order initialization; (ii) combined algorithm of weighted higher-order initialization and angle-based iteration.
+We consider both the Gaussian and Bernoulli models with $p = \{80, 100\}$, $r = 5$, $\gamma \in [-2.1, -1.4]$.
+Figure~\ref{fig:ini_re} shows the substantial improvement of combined algorithm over initialization, especially under weak and intermediate signals.
+This phenomenon agrees with the error rates in Theorems~\ref{thm:initial} and \ref{thm:refinement}
+and confirms the necessity of the local iterations.
+
+\subsection{Comparison with Other Methods}\label{subsec:comp}
+Compare our algorithm with methods below:
+\begin{itemize}[wide,topsep=-3pt,itemsep=0pt,parsep=1pt]
+    \item \textbf{\small HOSVD}: HOSVD on data tensor and $k$-means on the rows of the factor matrix;
+    \item \textbf{\small HOSVD+}: HOSVD on data tensor and $k$-means on the $\ell_2$-normalized rows of the factor matrix;
+    \item \textbf{\small HLloyd}~\citep{han2020exact}: High-order clustering algorithm developed for non-degree TBM;
+    \item \textbf{\small SCORE}~\citep{ke2019community}: Tensor-SCORE for clustering developed for binary tensors.
+\end{itemize}
+
+Among the four alternative algorithms, the \textbf{\small SCORE} is the closest method to ours.
+We set the tuning parameters of \textbf{\small SCORE} as in previous literature \citep{ke2019community}.
+The methods \textbf{\small SCORE} and \textbf{\small HOSVD+} are designed for dTBM~\eqref{eq:model_tensor}, whereas \textbf{\small HOSVD} and \textbf{\small HLloyd} are designed for non-degree models serving as benchmarks.
+We conduct two experiments to assess the impacts of (i) signal strength and (ii) degree heterogeneity under Gaussian and Bernoulli models with $ p = 100, r = 5$.
+We call our algorithm as \textbf{\small dTBM} in the comparison.
+
+We investigate the effects of signal to clustering performance by varying $\gamma \in [-1.5, -1.1]$.
+Figure~\ref{fig:comp_gamma} shows the consistent outperformance of our method \textbf{\small dTBM} among all algorithms.
+The sub-optimality of \textbf{\small SCORE} and \textbf{\small HOSVD+} indicates the necessity of local iterations on the clustering.
+Furthermore,  Figure~\ref{fig:comp_gamma} shows the inadequacy of non-degree algorithms in the presence of mild degree heterogeneity.
+The only exception is the slightly better performance of \textbf{\small HLloyd} over \textbf{\small HOSVD+} under Gaussian model.
+However, we find the advantage of \textbf{\small HLloyd} disappears with higher degree heterogeneity; see Supplement~\ref{sec:extrasim}.
+The experiment demonstrates the benefits of addressing heterogeneity in higher-order clustering tasks.   
+\begin{figure}[h!]
+ \vspace{-.3cm}
+    \centering
+    \includegraphics[width=.95\columnwidth]{comp_gamma_anno2.pdf}
+    \caption{ \footnotesize CER versus signal exponent (denoted $\gamma$) for different methods. We set $p = 100, r = 5, \gamma \in [-1.5, -1.1]$ under (a) Gaussian and (b) Bernoulli models.}
+    \label{fig:comp_gamma}
+    \vspace{-.4cm}
+\end{figure}
+% The only exception in Figure~\ref{fig:comp_gamma} is the slightly better performance of \textbf{\small HLloyd} over \textbf{\small HOSVD+} under Gaussian model.
+However, we find the advantage of \textbf{\small HLloyd} disappears with higher degree heterogeneity.
+
+The last experiment investigates the effects of degree heterogeneity to clustering performance.
+We use the same setting as in the first experiment in the Section~\ref{subsec:comp}, except that we fix the signal exponent $\gamma = -1.2$ and now generate the degree heterogeneity $\mtheta$ from Pareto distribution prior to normalization.
+The density function of Pareto distribution is $f(x|a,b) = a b^a x^{-(a+1)} \ind\{ x \geq b \}$, where $a$ is called \emph{shape} parameter.
+We vary the shape parameter $a \in [3,6]$ and choose $b$ such that $\bbE X = a(a - 1)^{-1}b = 1$ for $X$ following Pareto$(a,b)$.
+Note that a smaller $a$ leads to a larger variance in $\mtheta$ and hence a larger degree heterogeneity.
+Figure~\ref{fig:comp_theta} demonstrates the stability of degree-corrected algorithms (\textbf{\small dTBM}, \textbf{\small SCORE}, \textbf{\small HOSVD+}) over the entire range of degree heterogeneity under consideration.
+In contrast, non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) show poor performance with large heterogeneity, especially in Bernoulli cases.
+This experiment, again, highlights the benefit of addressing degree heterogeneity in higher-order clustering.
+\begin{figure}[h!]
+\vspace{-.2cm}
+    \centering
+    \includegraphics[width=.95\columnwidth]{comp_theta_anno2.pdf}
+    \caption{ \footnotesize CER versus shape parameter in degree (denoted $a\in[3,6]$) for different methods with $p = 100, r = 5, \gamma = -1.2$ under (a) Gaussian and (b) Bernoulli models.}
+    \label{fig:comp_theta}
+    \vspace{-.3cm}
+\end{figure}
+%\section{REAL DATA}\label{sec:real}
+
+\subsection{Peru Legislation Data Analysis}
+
+We apply our method to the legislation networks in the Congress of the Republic of Peru \citep{lee2017time}.
+Because of the frequent political power shifts in the Peruvian Congress during 2006-2011, we choose to focus on the data for the first half of 2006-2007 year.
+The dataset records the co-sponsorship of 116 legislators from top 5 parties and 802 bill proposals.
+We reconstruct legislation network as an order-3 binary tensor $\tY \in \{0,1\}^{116 \times 116 \times 116}$, where $\tY_{ijk} = 1$ if the legislators $(i,j,k)$ have sponsored the same bill, and $\tY_{ijk} = 0$ otherwise.
+True party affiliations of legislators are provided and serve as the ground truth.
+We apply various higher-order clustering methods to $\tY$ with $r = 5$.
+Table~\ref{tab:peru} shows that our \textbf{\small dTBM} achieves the best performance compared to others.
+The second best method is the two-stage algorithm \textbf{\small HLloyd}, followed by the spectral methods \textbf{\small SCORE} and \textbf{\small HOSVD+}, which consists with simulations under strong signal and moderate degree heterogeneity.
+%The comparison suggests that our method \textbf{\small dTBM} is more appealing in real-world applications.
+\vspace{-.5cm}
+\begin{table}[ht]
+    \centering
+    \resizebox{\columnwidth}{!}{%
+    \begin{tabular}{c | c cc c}
+    \hline
+        Method & \textbf{\small dTBM} 
+     %   &\textbf{\small HOSVD}
+        &\textbf{\small HOSVD+} & \textbf{\small HLloyd} &  \textbf{\small SCORE}\\
+         CER & \textbf{0.116}
+   %      &  0.22 
+         &0.213 & 0.149 &0.199\\
+         \hline
+    \end{tabular}
+    }
+    \vspace{-.2cm}
+    \caption{ \footnotesize Clustering errors (measured by CER) for various methods in the analysis of Peru Legislation dataset.}
+    \label{tab:peru}
+\end{table}
+
+\thispagestyle{empty}
+
+\subsection*{Acknowledgments}
+This research is supported in part by NSF grants DMS-1915978, DMS-2023239, EF-2133740, and funding from the Wisconsin Alumni Research foundation.
+We thank Zheng Tracy Ke, Rungang Han, Yuetian Luo for helpful discussions and for sharing software packages.
+
+\thispagestyle{empty}
+
+\bibliographystyle{apalike}
+\bibliography{tensor_wang}
+
+\thispagestyle{empty}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%% SUPPLEMENT (OPTIONAL) %%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\clearpage
+\appendix
+
+\thispagestyle{empty}
+
+% For one-column format, uncomment the following:
+\onecolumn \makesupplementtitle
+% For two-column format, uncomment the following:
+%\twocolumn[ \makesupplementtitle ]
+
+\setlength{\footskip}{10pt}
+
+\section{MOTIVATED EXAMPLES}\label{sec:examples}
+
+ Here, we provide four applications to illustrate the practical necessity of dTBM.
 
 \vspace{0.2cm}
-\paragraph{Tensor block model} Consider the model~\eqref{eq:model_tensor}.
+\paragraph{Tensor block model.} Consider the model~\eqref{eq:model_tensor}.
 Let $\theta(i)=1$ for all $ i \in [p]$.
@@
 
-\paragraph{Community detection in hypergraphs} Hypergraph network is a powerful tool to represent the complex entity relations with higher-order interactions~\citep{ke2019community}.
+\paragraph{Community detection in hypergraphs.} Hypergraph network is a powerful tool to represent the complex entity relations with higher-order interactions~\citep{ke2019community}.
 A typical undirected hypergraph is denoted as $H = (V,E)$, where $V = [p]$ is the set of nodes and $E$ is the set of undirected hyperedges.
@@
 \vspace{0.2cm}
-\paragraph{Multi-layer weighted network} 
+\paragraph{Multi-layer weighted network.} 
 Multi-layer weighted network data consists of multiple networks over the same set of nodes.
@@
 
-\paragraph{Gaussian higher-order clustering} Datasets in various fields such as medical image, genetics, and computer science are formulated as Gaussian tensors.
+\paragraph{Gaussian higher-order clustering.} Datasets in various fields such as medical image, genetics, and computer science are formulated as Gaussian tensors.
 One typical example is the multi-tissue gene expression dataset, which records the different gene expression in different individuals and different tissues.
@@
 
-\begin{rmk}[Comparison with non-degree models]
-Our dTBM uses fewer block parameters than TBM.
-In particular, every non-degree $r_1$-block tensor can be represented by a \emph{degree-corrected} $r_2$-block tensor with $r_2\leq r_1$.
-In particular, there exist tensors with $r_1=p$ but $r_2=1$, so the reduction in model complexity can be dramatic from $p$ to 1.
-This fact highlights the benefits of introducing degree heterogeneity in higher-order clustering tasks.
-\end{rmk}
-
-\subsection{Identifiability under angle gap condition}
-The goal of clustering is to estimate the partition function $z$ from model~\eqref{eq:model_tensor}.
-For ease of notation, we focus on symmetric tensors; the extension to non-symmetric tensors are similar.
-We use $\tP$ to denote the following parameter space for $(z,\tS,\mtheta)$,
-\begin{align}\label{eq:family}
-\tP=\bigg\{(z,\tS,\mtheta)\colon  \ \mtheta\in\mathbb{R}^p_{+},\ 
-{c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r}, c_3\leq \onormSize{}{\text{Mat}(\tS)_{a:}}\leq c_4, \onorm{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|, a\in[r]\bigg\},
-\end{align}
-\normalsize
-where $c_i>0$'s are universal constants. We briefly describe the rationale of the constraints in~\eqref{eq:family}. 
-First, the entrywise positivity constraint on  $\mtheta\in\mathbb{R}^p_{+}$ is imposed to avoid sign ambiguity between entries in $\mtheta_{z^{-1}(a)}$ and $\tS$.
-This constraint allows the trigonometric $\cos$ to describe the angle similarity in the Assumption~\ref{assmp:min_gap} below and Sub-algorithm~\hyperref[alg:main]{2} in Section~\ref{sec:alg}.
-Note that the positivity constraint can be achieved without sacrificing model flexibility, by using a slightly larger dimension of $\tS$ in the factorization~\eqref{eq:model_tensor}; see Example~\ref{ex:positive} below.
-Second, recall that the quantity $|z^{-1}(a)|$ denotes the number of nodes in $a$-th community.
-The constants $c_1, c_2$ in the $|z^{-1}(a)|$ bound assume the roughly balanced size across $r$ communities.
-Third, the constants $c_3, c_4 $ in the magnitude of $\text{Mat}(\tS)_{a:}$ requires no purely zero slide in $\tS$, so the core tensor $\tS$ is not trivially reduced to a lower rank.
-Lastly, the $\ell_1$ normalization $\onormSize{}{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|$ is imposed to avoid the scalar ambiguity between $\mtheta_{z^{-1}(a)}$ and $\tS$.
-This constraint, again, incurs no restriction to model flexibility but makes our presentation cleaner.
-
-\begin{example}[Positivity of degree parameters]~\label{ex:positive}
+\section{PARAMETER SPACE}
+
+Here, we provide the parameter space comparison with previous works and extra explanation for the positvity constraint of degree parameters.
+
+\subsection{Positivity of Degree Parameters}\label{subsec:positive}
+
 Here we provide an example to show the positivity constraints on $\mtheta$ incurs no loss on the model flexibility. 
@@
 In general, we can always reparameterize a block-$r$ tensor with mixed-signed degree using a block-$2r$ tensor with positive-signed degree.
-Since we assume $r=\tO(1)$ throughout the paper, the splitting does not affect the error rates of our interest.\end{example}
-
-We now provide the identifiability conditions for our model before estimation procedures.
-When $r=1$, the decomposition~\eqref{eq:model_tensor} is always unique (up to cluster label permutation) in $\tP$, because dTBM is equivalent to the rank-1 tensor family under this case.
-When $r\geq 2$, the Tucker rank of signal tensor $\bbE\tY$ in~\eqref{eq:model_tensor} is bounded by, but not necessarily equal to, the number of blocks $r$~\citep{wang2019multiway}.
-Therefore, one can not apply the classical identifiability conditions for low-rank tensors to dTBM.
-Here, we introduce a key separation condition on the core tensor.
-
-\begin{assumption}[Angle gap] \label{assmp:min_gap}Let $\mS = \text{Mat}(\tS)$. Assume the minimal gap between normalized rows of $\mS$ is bounded away from zero; i.e.,
-\begin{equation}\label{eq:minimal_gap}
-    \Delta_{\min} := \min_{a \neq b \in [r]} \onorm{\frac{\mS_{a:}}{\onormSize{}{ \mS_{a:}}} - \frac{\mS_{b:}}{\onormSize{}{ \mS_{b:}}} }>0, \quad \text{for} \quad r \geq 2.
-\end{equation}
-\end{assumption}
-We make the convention $\Delta_{\min} = 1$ for $r = 1$.
-Equivalently, \eqref{eq:minimal_gap} says that none of the two rows in $\mS$ are parallel; i.e., $\max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}}  = 1-\Delta^2_{\min}/2<1$.
-The quantity $\Delta_{\min}$ characterizes the non-redundancy among clusters measured by angle separation.
-The denominators involved in definition~\eqref{eq:minimal_gap} are well posed because of the lower bound on $\onorm{\mS_{a:}}$ in~\eqref{eq:family}.
-
-Our first main result is the following theorem showing the sufficiency and necessarily of the angle gap separation condition for the parameter identifiability under dTBM.
-
-\begin{thm}[Model identifiability]\label{thm:unique} Consider the dTBM with $r\geq 2$.
-The parameterization~\eqref{eq:model_tensor} is unique in $\tP$ up to cluster label permutations, if and only if Assumption~\ref{assmp:min_gap} holds.
-\end{thm}
-
-The identifiability guarantee for the dTBM is more appealing than classical Tucker model.
-In the Tucker model, the factor matrix $\mM$ is identifiable only up to orthogonal rotations.
-In contrast, our model does not suffer from rotational invariance.
-As we will show in Section~\ref{sec:alg}, each column of the membership matrix $\mM$ can be precisely recovered under our algorithm.
-This property benefits the interpretation of dTBM in practice.
-
-\section{Statistical-computational limits for higher-order tensors}\label{sec:limits}
-
-In this section, we study the statistical and computational limits of dTBM. We propose signal-to-noise ratio~(SNR),
-\begin{align}\label{eq:gamma}
-  \text{SNR}:= \Delta^2_{\min}/\sigma^2 = p^{\gamma}, 
-\end{align}
-with varying $\gamma \in \bbR$ that quantifies different regimes of interest.
-We call $\gamma$ the \emph{signal exponent}.
-Intuitively, a larger SNR, or equivalently a larger $\gamma$, benefits the clustering in the presence of noise.
-With quantification~\eqref{eq:gamma}, we consider the following parameter space,
-\begin{equation}\label{eq:gammafamily}
-    \tP(\gamma) = \tP\cap\{\tS \text{ satisfies SNR condition~\eqref{eq:gamma} with $\gamma$} \}.
-\end{equation}
-Note that $1$-block dTBM does not belong to the space $\tP(\gamma)$ when $\gamma < 0$ by Assumption~\ref{assmp:min_gap}.
-Our goal is to characterize the clustering accuracy with respect to $\gamma$.
-Let $\hat z$ and $z$ be the estimated and true clustering functions in the family~\eqref{eq:family}.
-Define the misclustering error by
-\[
-\ell(\hat z, z)={1\over p}\min_{\pi \in \Pi}\sum_{i\in[p]}\ind\{\hat z(i)\neq \pi \circ z(i)\},
-\]
-where $\pi: [r] \mapsto [r]$ is a permutation of cluster labels, $\circ$ denotes the composition operation, and $\Pi$ denotes the collection of all possible permutations.
-The infinitum over all permutations accounts for the ambiguity in cluster label permutation.
-
-In Sections~\ref{sec:statlimit} and \ref{sec:complimit}, we provide the lower bounds of $\ell (\hat z, z)$ for general Gaussian dTBMs~\eqref{eq:model_margin} without symmetric assumptions.
-For general (asymmetric) Gaussian dTBMs, we assume Gaussian noise $\tE(i_1,\ldots,i_K)\stackrel{\text{i.i.d.}}{\sim} N(0,\sigma^2)$, and we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode.
-For notational simplicity, we still use $z$ and $\tP(\gamma)$ for this general (asymmetric) model.
-All lower bounds should be interpreted as the worst-case results across $K$ modes.
-
-\subsection{Statistical critical values}\label{sec:statlimit}
-The statistical limit means the minimal SNR required for solving dTBMs with \emph{unlimited computational cost.} Our following result shows the minimax lower bound of SNR for exact recovery in dTBM.
-
-\begin{thm}[Statistical lower bound]\label{thm:stats} Consider general Gaussian dTBMs under the parameter space $\tP(\gamma)$ with $K\geq 1$.
-Assume $r \lesssim p^{1/3}$.
-If the signal exponent satisfies $\gamma < -(K-1)$, then, every estimator $\hat z_{\text{stat}}$ obeys
-\begin{equation}
-    \sup_{(z, \tS, \mtheta) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
-\end{equation}
-\end{thm}
-Theorem~\ref{thm:stats} demonstrates the impossibility of exact recovery of the assignment when $\gamma < -(K-1)$ in the high-dimensional regime $p\to \infty$ for fixed $r$.
-The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to, maximum likelihood estimation (MLE)~\citep{wang2019multiway} and trace maximization~\citep{ghoshdastidar2017uniform}.
-As we will show in Section~\ref{sec:alg}, the SNR threshold $-(K-1)$ is also a minimax upper bound, because MLE achieves exact recovery when $\gamma > -(K-1)$.
-Hence, the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM.
-
-\subsection{Computational critical values}\label{sec:complimit}
-In this section, we derive the computational limits of dTBMs.
-The computational limit means the minimal SNR required for exactly recovery with \emph{polynomial-time} computational cost.
-An important ingredient to establish the computational limits is the \emph{hypergraphic planted clique (HPC) conjecture} \citep{zhang2018tensor, brennan2020reducibility}.
-The HPC conjecture indicates the impossibility of fully recovering the planted cliques with polynomial-time algorithm when the clique size is less than the number of vertices in the hypergraph.
-The formal statement of HPC detection conjecture is provided in Definition~\ref{def:HPC} and Conjecture~\ref{hypo:HPC} as follows.
-
-\begin{defn}[Hypergraphic planted clique (HPC) detection]~\label{def:HPC} Consider an order-$K$ hypergraph $H = (V,E)$ where $V = [p]$ collects vertices and $E$ collects all the order-$K$ edges.
-Let $\tH_k(p, 1/2)$ denote the Erd\H{o}s-R\'{e}nyi $K$-hypergraph where the edge $(i_1,\ldots, i_K)$ belongs to $E$ with probability $1/2$.
-Further, we let $\tH_K(p, 1/2, \kappa)$ denote the hyhpergraph with planted cliques of size $\kappa$.
-Specifically, we generate a hypergraph from $\tH_k(p, 1/2)$, pick $\kappa$ vertices uniformly from $[p]$, denoted $K$, and then connect all the hyperedges with vertices in $K$.
-Note that the clique size $\kappa$ can be a function of $p$, denoted $\kappa_p$.
-The order-$K$ HPC detection aims to identify whether there exists a planted clique hidden in an Erd\H{o}s-R\'{e}nyi $K$-hypergraph.
-The HPC detection is formulated as the following hypothesis testing problem
-\begin{equation}
-    H_0:\ H \sim \tH_K(p,1/2) \quad \text{versus} \quad H_1: \ H \sim \tH_K(p,1/2, \kappa_p).
-\end{equation}
-\end{defn}
-
-\begin{conjecture}[HPC conjecture]\label{hypo:HPC} Consider the HPC detection problem in Definition~\ref{def:HPC}.
-Suppose the sequence $\{\kappa_p\}$ such that $\limsup_{p \rightarrow \infty} \log \kappa_p/ \log \sqrt{p} \leq (1 - \tau)$.
-Then, for every sequence of polynomial-time test $\{ \varphi_p\}: H \mapsto \{0,1\}$ we have
-\begin{equation}
-    \liminf_{p \rightarrow \infty} \bbP_{H_0} \of{ \varphi_p(H) =1 } +  \bbP_{H_1} \of{ \varphi_p(H) =0} \geq \frac{1}{2}.
-\end{equation}
-\end{conjecture}
-
-Under the HPC conjecture, we establish the SNR lower bound that is necessary for any \emph{polynomial-time} estimator to achieve exact clustering.
-
-\begin{thm}[Computational lower bound]\label{thm:comp} Consider general Gaussian dTBMs under the parameter space $\tP(\gamma)$ with $K\geq 2$.
-Assume HPC conjecture holds.
-If the signal exponent $\gamma < -K/2$, then, every \emph{polynomial-time estimator} $\hat z_{\text{comp}}$ obeys
-\begin{align}
-   \liminf_{p\to \infty}\sup_{(z, \tS, \mtheta) \in \tP(\gamma)}  \bbE \left[ p\ell(\hat z_{\text{comp}}, z) \right]\geq 1.
-\end{align}
-\end{thm}
-Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$.
-Therefore, $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM.
-In Section~\ref{sec:alg}, we will show the condition $\gamma> -K/2$ suffices for our proposed polynomial-time estimator.
-Thus, $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM.
-
-\begin{rmk}[Statistical-computational gaps]
-Now, we have established the phase transition of exact clustering under order-$K$ dTBM by combing Theorems~\ref{thm:stats} and \ref{thm:comp}.
-Figure~\ref{fig:phase_axis} summarizes our results of critical SNRs when $K \geq 2$.
-In the weak SNR region $\gamma < -(K-1)$, no statistical estimator succeeds in degree-corrected higher-order clustering.
-In the strong SNR region $\gamma  > -K/2$, our proposed algorithm precisely recovers the clustering in polynomial time.
-In the moderate SNR regime, $-(K-1)\leq \gamma \leq -K/2$, the degree-corrected clustering problem is statistically easy but computationally hard.
-Particularly, dTBM reduces to matrix degree-corrected model when $K =2$, and the statistical and computational bounds show the same critical value.
-When $K =1$, dTBM reduces to the degree-corrected sub-Gaussian mixture model (GMM) with model
-\begin{equation}
-    \mY = \mTheta \mM \mS + \mE,
-\end{equation}
-where $\mY \in \bbR^{p \times d}$ collects $n$ data points in $\bbR^d$, $\mS \in \bbR^{r \times d}$ collects the $d$-dimensional centroids for $r$ clusters, and $\mTheta \in \bbR^{p \times p}, \mM \in \{0,1\}^{p \times r}, \mE \in \bbR^{p \times d}$ have the same meaning as in dTBM.
-\cite{lu2016statistical} implies that polynomial-time algorithms are able to achieve the statistical minimax lower bound in GMM.
-Therefore, we conclude that the statistical-to-computational gap emerges only for higher-order tensors with $K \geq 3$.
-The result reveals the intrinsic distinctions among (vector) one-dimensional clustering, (matrix) biclustering, and (tensor) higher-order clustering.
-\end{rmk}
-
-\begin{rmk}[Comparison with non-degree models]
-We compare our results to non-degree tensor models.
-The allowance of degree heterogeneity $\mtheta$ makes the model more flexible, but it incurs extra statistical and computational complexity.
-Fortunately, we find that the extra complexity does not render the estimation of $z$ qualitatively harder; see the comparison of our phase transition with non-degree TBM~\citep{han2020exact}.
-\end{rmk}
-
-\section{Polynomial-time algorithm under mild SNR}\label{sec:alg}
-In this section, we present an efficient polynomial-time clustering algorithm under mild SNR.
-The procedure takes a global-to-local approach.
-See Figure~\ref{fig:demo} for illustration.
-The global step finds the basin of attraction with polynomial miclustering error, whereas the local iterations improve the initial clustering to exact recovery.
-Both steps are critical to obtain a satisfactory algorithm output.
-In what follows, we first use the symmetric tensor as a working example to describe the algorithm procedures to gain insight.
-Our theoretical analysis focuses on the noisy tensor with i.i.d.\ sub-Gaussian noise such as Gaussian and uniform observations.
-The extensions for asymmetric tensor and Bernoulli observation and other practical issues are in Section~\ref{subsec:exten}.
-
-\begin{figure}[ht!]
+Since we assume $r=\tO(1)$ throughout the paper, the splitting does not affect the error rates of our interest.
+
+ \subsection{Parameter Space Comparison with Previous Works}\label{subsec:comp_para}
+
+Table~\ref{tab:para} indicates that our parameter space is flexible compared to existing work.
+The balanced community assumption is mild in literature.
+In fact, our simulation requires only the positive community size (i.e., $\min_{a \in [r]}|z^{-1}(a)| \geq 1$) but no balanced community assumptions; see Section~\ref{sec:simulation}.
+The outperformance of dTBM shows the robustness of our algorithm to the imbalanced communities.
+
+\begin{table}[h]
     \centering
-\includegraphics[width=\columnwidth]{alg_demo1.pdf}
-\caption{Illustration of our global-to-local algorithm.}\label{fig:demo}
-\end{figure}
-
-\subsection{Weighted higher-order initialization}
-
-We start with weighted higher-order clustering algorithm as initialization.
-We take an order-3 symmetric tensor as illustration for insight.
-Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
-By model~\eqref{eq:model_tensor}, for all $i \in [p]$, we have
-\begin{equation}\label{eq:kmeans}
-    \theta(i)^{-1} \mX_{i:} = \off{\text{Mat}( \tS \times_2 \mTheta \mM \times_3  \mTheta \mM )}_{z(i):}. 
-\end{equation}
-This implies that, all node $i$ belonging to $a$-th community (i.e., $z(i)=a$) share the same normalized mean vector $\theta(i)^{-1} \mX_{i:}$, and vice versa.
-Intuitively, one can apply $k$-means clustering to the vectors $\{ \theta(i)^{-1} \mX_{i:} \}_{i\in[p]}$, which leads to main idea of our Sub-algorithm~\hyperref[alg:main]{1}.
-
-Specifically, our initialization consists of denoising step and clustering step.
-The denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) estimates $\tX$ from $\tY$ by a double projection spectral method.
-The first projection performs HOSVD~\citep{de2000multilinear} via $\mU_{\text{pre}} = \text{SVD}_{r} \of{ \text{Mat}(\tY) }$, where $\text{SVD}_r(\cdot)$ returns the top-$r$ left singular vectors.
-The second projection performs HOSVD on the projected $\tY$ onto the multilinear Kronecker space $\mU_{\text{pre}} \otimes  \mU_{\text{pre}}$; i.e.,
-\begin{equation}\label{eq:two-step_factor}
-    \hat \mU = \text{SVD}_{r} \of{\text{Mat}\of{ \tY\times_1  \mU_{\text{pre}} \mU_{\text{pre}}^T \times_2   \mU_{\text{pre}} \mU_{\text{pre}}^T }}.
-\end{equation}
-The final denoised tensor $\hat \tX$ is defined by
-\begin{equation}\label{eq:two-step_est}
-    \hat \tX = \tY \times_1 \hat \mU \hat 
-\mU^T \times_2 \hat \mU \hat \mU^T \times_3 \hat \mU \hat \mU^T. 
-\end{equation}
-The double projection improves usual matrix spectral methods in order to alleviate the noise effects for $K\geq 3$~\citep{han2020exact}.
-
-The clustering step (lines 3-5 in Sub-algorithm~\hyperref[alg:main]{1}) performs the weighted $k$-means clustering. 
-We write $\hat \mX=\text{Mat}(\hat \tX)$, and normalize the rows into $\hat \mX^s_{i:}=\onormSize{}{\hat \mX_{i:}}^{-1}\hat \mX_{i:}$ as a surrogate of $\theta(i)^{-1} \mX_{i:}$.
-Then, a weighted $k$-means clustering is performed on the normalized rows with weights equal to $\onormSize{}{\hat \mX_{i:}}^2$.
-The choice of weights is to bound the $k$-means objective function by the Frobenius-norm accuracy of $\hat \tX$.
-Unlike existing clustering algorithm~\citep{ke2019community}, we apply the clustering on the unfolded tensor $\hat \mX$ rather than on the factors $\hat \mU$.
-This strategy relaxes the singular-value gap condition~\citep{gao2018community, han2020exact}.
-We assign degenerate rows with purely zero entries to an arbitrarily random cluster; these nodes are negligible in high-dimensions because of the lower bound on $\onormSize{}{\text{Mat}(\tS)_{a:}}$ in~\eqref{eq:family}.
-The final result gives the initial cluster assignment $\hat z^{(0)}$.
-Full procedures are provided in Sub-algorithm~\hyperref[alg:main]{1}.
-
-\begin{algorithm}[h!]
-\caption*{\bf Algorithm: Multiway spherical clustering for degree-corrected tensor block model }
-\vspace{.15cm}
-\begin{algorithmic}[1]
-\Algphase{Sub-algorithm 1: Weighted higher-order initialization}
-\INPUT Observation $\tY \in \bbR^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
-\State Compute factor matrix $ \mU_{\text{pre}} = \text{SVD}_{r} (\text{Mat}(\tY))$ and the $(K-1)$-mode projection $
-\tX_{\text{pre}} = \tY \times_1   \mU_{\text{pre}} \mU_{\text{pre}}^T \times_2 \cdots \times_{K-1}  \mU_{\text{pre}} \mU_{\text{pre}}^T.$
-\State Compute factor matrix $\hat \mU = \text{SVD}_{r}(\text{Mat}(\tX_{\text{pre}}))$ and denoised tensor
-$\hat \tX = \tY \times_1 \hat \mU \hat \mU^T \times_2 \cdots \times_K \hat \mU \hat \mU^T.$
-\State {Let $\hat \mX = \text{Mat}(\hat \tX)$ and $S_0=\{i \in [p]: \onormSize{}{\hat \mX_{i:}} = 0\}$. Set $\hat z(i)$ randomly in $[r]$ for $i \in S_0$.}
-\State{For all $i\in S_0^c$, compute normalized rows
-$\hat \mX_{i:}^s :=\onormSize{}{\hat \mX_{i:}}^{-1} \hat \mX_{i:}.$
+    \resizebox{\textwidth}{!}{
+    \begin{tabular}{c|c ccc}
+    \hline
+    Assumptions in parameter space& \cite{gao2018community}& \cite{han2020exact}&\cite{ke2019community} &Ours\\
+    \hline
+         Balanced community size &$\surd$ &$\surd$&$\surd$ & $\surd$  \\
+         Balanced degree parameters & $\surd$&-&$\surd$ &$\surd$\\
+         Flexible in-group connections &$\times$ &$\surd$&$\times$& $\surd$\\
+         Gap among cluster centers & In-between cluster difference & Euclidean gap & Eigen gap & Angle gap\\
+         \hline
+    \end{tabular}
     }
-\State {Solve the clustering $\hat z\colon [p]\to[r]$ and centroids $ (\hat \mx_j)_{j\in[r_k]}$ using weighted $k$-means, such that}
-\begin{align}
-    &\sum_{i \in S_0^c }  \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{\hat z(i)} }^2 
-    \leq 
-    \eta \min_{\substack{\bar \mx_j, j\in[r], \bar z(i),i\in S_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\bar z(i)}}^2.
-\end{align}
-\OUTPUT Initial clustering $z^{(0)} \leftarrow \hat z$.
-
-\Algphase{Sub-algorithm 2: Angle-based iteration}
-\INPUT Observation $\tY \in \bbR^{p \times \cdots \times p}$, initialization $z^{(0)} \colon [p]\to[r]$ from Sub-algorithm 1, iteration number $T$.
-\For {$t = 0$ to $T-1$}
-\State Update the block tensor $\tS^{(t)}$ via
-$\tS^{(t)} (a_1,...,a_K)= \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}(i_k) = a_k, k \in [K]\}.$
-\State Calculate reduced tensor $\tY^{\text{d}} \in \bbR^{p \times r \times \cdots \times r}$ via
-\begin{align}
-    &\tY^{\text{d}}(i,a_2,\ldots,a_K) 
-    = 
-\text{Ave}\{\tY(i,i_2,\ldots,i_K): z^{(t)}(i_k) = a_k, k \neq 1 \}.
-\end{align}
-
-\State Let $\mY^{\text{d}} = \text{Mat}(\tY^{\text{d}})$ and $J_0 = \{ i\in[p]: \onorm{\mY^{\text{d}}_{i:}} = 0\}$. Set $z^{(t+1)}(i)$ randomly in $[r]$ for $i \in J_0$.
-
-\State Let $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. For all $i \in J_0^c$ update the cluster assignment by
-\begin{equation}
-    z(i)^{(t+1)} = \argmax_{a \in [r]} \cos \left( \mY^{\text{d}}_{i:},\ \mS^{(t)}_{a:} \right).
-\end{equation}
-
-\EndFor
-
-\OUTPUT Estimated clustering $z^{(T)}  \in [r]^{p}$.
-
-\end{algorithmic}
-\end{algorithm}
-
-We now establish the misclustering error rate of initialization. We call $\mtheta$ is balanced, if the relative extent of heterogeneity is comparable across clusters in that
-\begin{equation}\label{eq:degree}
-{\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
-\end{equation}
-Note that, the assumption~\eqref{eq:degree} does not preclude degree heterogeneity.
-Indeed, within each of the clusters, the highest degree can be $\theta(i) = \Omega(p)$, whereas the lowest degree can be $\theta(i)=\tO(1)$.
-
-\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with i.i.d.\ noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
-Assume $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$.
-Let $ z^{(0)}$ denote the output of Sub-algorithm~\hyperref[alg:main]{1}.
-With probability going to 1, we have
-\begin{equation}\label{eq:ini}
-   \ell(z^{(0)}, z) \lesssim {r^K p^{-K/2}\over \text{SNR}}. 
-\end{equation}
-\end{thm}
-
-\begin{rmk}[Comparison to previous results] For fixed SNR, our initialization error rate with $K=2$ agrees with the initialization error rate $\tO(p^{-1})$ in matrix models~\citep{gao2018community}.
-Furthermore, in the special case of non-degree TBMs with $\theta_1=\cdots=\theta_p=1$, we achieve the same initial misclustering error $\tO(p^{-K/2})$ as in non-degree models~\citep{han2020exact}.
-Theorem~\ref{thm:initial} implies the advantage of our algorithm in achieving both accuracy and model flexibility.
-\end{rmk}
-
-\begin{rmk}[Failure of conventional tensor HOSVD] If we use conventional HOSVD for tensor denoising; that is, we use $\mU_{\text{pre}}$ in place of $\hat \mU$ in line 2, then the misclustering rate becomes $\tO(p^{-1})$ for all $K\geq 2$.
-This rate is substantially worse than our current rate~\eqref{eq:ini}.
-\end{rmk}
-
-\begin{rmk}[Singular-value gap-free clustering] Note that our clustering directly applies to the estimated mean tensor $\hat \tX$ rather than the leading tensor factors $\hat \mU$.
-Applying clustering to the tensor factors suffers from the non-identifiability issue due to the infinitely many orthogonal rotations when the number of blocks $r \geq 3$ in the absence of singular-value gaps.
-Such ambiguity causes the trouble for effective clustering~\citep{abbe2020entrywise}.
-In contrast, our initialization algorithm applies the clustering to the overall mean tensor $\hat \tX$.
-This strategy avoids the non-identifiability issue regardless of the number of blocks and singular-value gaps.
-\end{rmk}
-
-\subsection{Angle-based iteration}
-Our Theorem~\ref{thm:initial} has shown the polynomially decaying error rate from our initialization.
-Now we improve the error rate to exponential decay using local iterations.
-We propose an angle-based local iteration to improve the outputs from Sub-algorithm~\hyperref[alg:main]{1}.
-To gain the intuition, consider an one-dimensional degree-corrected clustering problem with data vectors $\mx_i = \theta(i) \ms_{z(i)} + \mepsilon_i, i \in [p]$, where $\ms_i$'s are known cluster centroids, $\theta(i)$'s are unknown positive degrees, and $z\colon [p] \mapsto [r]$ is the cluster assignment of interest.
-The angle-based $k$-means algorithm estimates the assignment $z$ by minimizing the angle between data vectors and centroids; i.e.,
- \begin{equation}\label{eq:angle_kmeans}
-     z(i) = \argmax_{a \in [r]} \cos ( \mx_i,\ \ms_{a} ),\ \text{ for all }i \in [p].
- \end{equation}
-The classical Euclidean-distance based clustering~\citep{han2020exact} fails to recover $z$ in the presence of degree heterogeneity, even under noiseless case.
-In contrast, the proposed angle-based $k$-means achieves accurate recovery without explicit estimation of $\mtheta$.
-
-Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as in angle-based $k$-means.
-We still take the order-3 tensor for illustration.
-Specifically, Sub-algorithm~\hyperref[alg:main]{2} updates estimated core tensor and cluster assignment in each iteration.
-We use superscript $\cdot^{(t)}$ to denote the estimate from $t$-th iteration, where $t=1,\ldots.$ For core tensor, we consider the following update strategy
- \[
- \tS^{(t)}(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k\in[3]\}.
-\]
-Intuitively, $\tS^{(t)}$ becomes closer to the true core $\tS$ as $z^{(t)}$ is more precise.
-For cluster assignment, we first aggregate the slices of $\tY$ and obtain a reduced tensor $\tY^{\text{d}}\in \bbR^{p \times r \times r}$ with given $z^{(t)}$, where
-\[
-\tY^{\text{d}}(i,a_2,a_3)=\text{Ave}\{\tY(i,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k \neq 1\}.
-\]
-We use $\mY^d$ and $ \mS^{(t)}$ to denote the $\text{Mat}(\tY^{\text{d}})$ and $\text{Mat}(\tS^{(t)})$.
-The rows $\mY^d_{i:}$ and $\mS^{(t)}_{a:}$ correspond to the $\mx_i$ and $\ms_a$ in the one-dimensional clustering~\eqref{eq:angle_kmeans}.
-Then, we obtain the updated assignment by
-\[
-z(i)^{(t+1)} = \argmax_{a \in [r]} \cos \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} },\ \text{ for all } i \in [p],
-\]
-provided that $\mS^{(t)}_{a:}$ is a non-zero vector.
-Otherwise, if $\mS^{(t)}_{a:}$ is a zero vector, then we make the convention to assign $z^{(t+1)}(i)$ randomly in $[p]$.
-Full procedures for our angle-based iteration are described in Sub-algorithm~\hyperref[alg:main]{2}.
-
-We now establish the misclustering error rate of iterations under the stability assumption.
-
-\begin{defn}[Locally linear stability] \label{def:stable}
-Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\varepsilon)=\{\bar z\colon \ell(\bar z, z) \leq \varepsilon\}$.
-Let $\bar z\colon[p]\to [r]$ be a clustering function.
-We define two vectors associated with $\bar z$,
-\begin{align}
-    &\mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T, \quad \mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
-\end{align}
-
-We call the degree is $\varepsilon$-locally linearly stable if and only if 
-\begin{equation}\label{eq:local}
-    \sin(\mp(\bar z),\ \mp_{\mtheta}(\bar z))\lesssim \varepsilon \Delta_{\min} ,\quad \text{for all } \bar z\in \tN(z, \varepsilon).
-\end{equation}
-\end{defn}
-
-Roughly speaking, the vector $\mp(\bar z)$ represents the raw cluster sizes, and $\mp_{\theta}(\bar z)$ represents the relative cluster sizes weighted by degrees. 
-The local stability holds trivially for $\varepsilon=0$ based on the construction of parameter space~\eqref{eq:family}.
-The condition~\eqref{eq:local} controls the impact of node degree to the $\mp_{\theta}(\cdot)$ with respect to the misclassification rate $\varepsilon$ and angle gap.
-
-\begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the setup as in Theorem~\ref{thm:initial}. Suppose 
-$r=\tO(1)$ and $\text{SNR} \geq \tilde C p^{-K/2}\log p$ for some sufficiently large constant $\tilde C$.
-Assume the local linear stability of degree holds in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \geq \check{C}\log^{-1}p $ with some positive constant $\check{C}$.
-Let $z^{(t)}$ denote the $t$-th iteration output in Sub-algorithm~\hyperref[alg:main]{2} with initialization $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{1}.
-With probability going to 1, there exists a contraction parameter $\rho \in (0,1)$ such that
-\begin{align}\label{eq:final}
-    \ell(z, \hat z^{(t+1)}) \lesssim &\ \KeepStyleUnderBrace{
-   \text{SNR}^{-1}
-    \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}}_{\substack{\text{statistical error}} }+ \KeepStyleUnderBrace{ \rho^t \ell(z, z^{(0)}). }_{\substack{\text{computational error}}}
-\end{align}
-\end{thm}
-From the conclusion~\eqref{eq:final}, we find that the iteration error is decomposed into two parts: statistical error and computational error.
-The statistical error is unavoidable with noisy data regardless $t$, whereas the computational error decays in an exponential rate as the number of iterations $t \rightarrow \infty$.
-
-Theorem~\ref{thm:refinement} implies that, with probability going to 1, our estimate $z^{(T)}$ achieves exact recovery within polynomial iterations; more precisely,
-\begin{equation}
-     z^{(T)} = \pi \circ z, \quad \text{for all }T\gtrsim \log_{1/\rho} p,
-\end{equation}
-for some permutation $\pi \in \Pi$.
-Therefore, our combined algorithm is \textit{computationally efficient} as long as SNR $\gtrsim p^{-K/2} \log p$.
-Note that, ignoring the logarithmic term, the minimal SNR requirement, $p^{-K/2}$, coincides with the computational lower bound in Theorem~\ref{thm:comp}.
-Therefore, our algorithm is optimal regarding the signal requirement and lies in the sharpest \emph{computationally efficient} regime in Figure~\ref{fig:phase_axis}.
-
-\subsection{Extensions and practical issues}~\label{subsec:exten}
-\hspace{-.3cm} {\bf Extension for Bernoulli observations.} The main difficulty to establish the statistical guarantee for Bernoulli observations lies in the initialization Sub-algorithm~\hyperref[alg:main]{1}.
+    \caption{Comparison of parameter space of previous works with our work. }
+    \label{tab:para}
+\end{table}
+
+\section{EXTENSION AND PRACTICAL ISSUES}\label{sec:exten}
+
+Here, we provide the extension of our algorithm for Bernoulli observations and asymmetric tensors and other practical issues including computational complexity and hyper-parameter selection.
+
+ {\bf Extension for Bernoulli Observations.} The main difficulty to establish the statistical guarantee for Bernoulli observations lies in the initialization Sub-algorithm~\hyperref[alg:main]{1}.
 Theorem~\ref{thm:refinement} still holds for Bernoulli observations once the initialization accuracy satisfies the upper bound~\eqref{eq:ini} in Theorem~\ref{thm:initial}.
@@
 
-{\bf Extension for general dTBMs.} Our two-stage algorithm is able to be extended for the general (asymmetric) dTBMs.
+{\bf Extension for General dTBMs.} Our two-stage algorithm is able to be extended for the general (asymmetric) dTBMs.
 Specifically, in the Sub-algorithm~\hyperref[alg:main]{1}, we make the following changes: (1)
@@
 
-{\bf Computational complexity.} Our two-stage algorithm has a computational cost polynomial in tensor dimension $p$.
+{\bf Computational Complexity.} Our two-stage algorithm has a computational cost polynomial in tensor dimension $p$.
 Specifically, the complexity of Sub-algorithm~\hyperref[alg:main]{1} is $\tO(K p^{K+1} + Krp^K)$, where the first term is contributed by the double projection and the calculation of $\hat \tX$, and the second term comes from normalization and the $k$-means.
@@
 
-{\bf Hyper-parameter selection.} In our theoretical analysis, we have assumed the true cluster number $r$ is given to our algorithm.
+{\bf Hyper-parameter Selection.} In our theoretical analysis, we have assumed the true cluster number $r$ is given to our algorithm.
 In practice, the cluster number $r$ is often unknown, and we now propose a method to choose $r$ from data.
@@
 where $\hat \mY^{\text{d}}(r) = \mat(\hat \tY^{\text{d}}(r))$, $\hat \tY^{\text{d}}(r)(i,a_2,\ldots,a_K) =  \text{Ave}\{\tY(i,i_2,\ldots,i_K): \hat z(i_k, r) = a_k, k \neq 1 \}$, and $\hat z(i,r)$ denotes the community label for the $i$-th node with given cluster number $r$.
-We evaluate the performance of the BIC criterion in Section~\ref{subsec:num_theory}.
-
-\section{Numerical studies}\label{sec:simulation}
-
- We evaluate the performance of the weighted higher-order initialization and angle-based iteration in this section.
-We report average errors and standard deviations across 30 replications in each experiment.
-Clustering accuracy is assessed by clustering error rate (CER, i.e., one minus rand index).
-Note that CER between $(\hat z, z)$ is equivalent to misclustering error $\ell(\hat z, z)$ up to constant multiplications~\citep{meilua2012local}, and a lower CER indicates a better performance.
-
-We generate order-3 tensors with \emph{assortative}~\citep{gao2018community} core tensors to control SNR; i.e., we set $\tS_{aaa} = s_1$ for $a \in [r]$ and others be $s_2$, where $s_1 > s_2 > 0$.
-Let $\alpha = s_1/s_2$.
-We set $\alpha$ close to 1 such that $1-\alpha=o(p)$.
-In particular, we have $\alpha = 1 + \Omega(p^{\gamma/2})$ with $\gamma<0$ by Assumption~\ref{assmp:min_gap} and definition~\eqref{eq:gamma}.
-Hence, we easily adjust SNR via varying $\alpha$.
-Note that the assortative setting is proposed for simulations, and our algorithm is applicable for general tensors in practice.
-The cluster assignment $z$ is randomly generated with equal probability across  $r$ clusters for each mode.
-Without further explanation, we generate degree heterogeneity $\mtheta$ from absolute normal distribution by $\theta(i) = |X_i| + 1 - 1/\sqrt{2\pi}$ with $|X_i| \stackrel{\text{i.i.d.}}\sim N(0,1), i \in [p]$ and normalize $\mtheta$ to satisfy \eqref{eq:family}.
-Also, we set $\sigma^2 = 1$ for Gaussian data without further specification.
-
-\subsection{Verification of theoretical results}\label{subsec:num_theory}
-
-The first experiment verifies  statistical-computational gap described in Section~\ref{sec:limits}.
-Consider the Gaussian model with $p = \{80, 100\}$, $r = 5$.
-We vary $\gamma $ in $ [-1.2, -0.4]$ and $[-2.1, -1.4]$ for matrix ($K=2$) and tensor $(K = 3)$ clustering, respectively.
-Note that finding MLE under dTBM is computationally intractable.
-We approximate MLE using an oracle estimator, i.e., the output of Sub-algorithm~\hyperref[alg:main]{2} initialized from true assignment.
-Figure~\ref{fig:phase}a shows that both our algorithm and oracle estimator start to decrease around the critical value $\gamma_{\text{stat}}  = \gamma_{\text{comp}}  = -1$ in matrix case.
-In contrast, Figure~\ref{fig:phase}b shows a significant gap in the phase transitions between the algorithm estimator and oracle estimator in tensor case.
-The oracle error rapidly decreases to 0 when $\gamma_{\text{stat}} = -2$, whereas the algorithm estimator tends to achieve exact clustering when $\gamma_{\text{comp}} = -1.5$.
-Figure~\ref{fig:phase} confirms the existence of the statistical-computational gap in our Theorems~\ref{thm:stats} and~\ref{thm:comp}.
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=.8\columnwidth]{phase_anno3.pdf}
-    \caption{SNR phase transitions for clustering in dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.
-    }
-    \label{fig:phase}
-\end{figure}
-
-The second experiment verifies the performance guarantees of two algorithms: (i) weighted higher-order initialization; (ii) combined algorithm of weighted higher-order initialization and angle-based iteration.
-We consider both the Gaussian and Bernoulli models with $p = \{80, 100\}$, $r = 5$, $\gamma \in [-2.1, -1.4]$.
-Figure~\ref{fig:ini_re} shows the substantial improvement of combined algorithm over initialization, especially under weak and intermediate signals.
-This phenomenon agrees with the error rates in Theorems~\ref{thm:initial} and \ref{thm:refinement}
-and confirms the necessity of the local iterations.
-
-\begin{figure}[htp!]
-    \centering
-     \includegraphics[width=.8\columnwidth]{ini_re_anno3.pdf}
-    \caption{CER versus signal exponent $(\gamma)$ for initialization only and for combined algorithm.
-We set $p = \{80, 100\}, r = 5, \gamma \in [-2.1, -1.4]$ under (a) Gaussian models and (b) Bernoulli models.
-}
-    \label{fig:ini_re}
-
-\end{figure}
+We evaluate the performance of the BIC criterion in Section~\ref{subsec:sim_select}.
+
+\section{ADDITIONAL EXPERIMENTS}\label{sec:extra_exp}
+
+Here, we provide the additional experiments including the numerical verification of hyper-parameter selection criterion, extra simulation results for comparison, and real data analysis for human brain connectome data.
+
+\subsection{Verification of Hyper-parameter Selection}\label{subsec:sim_select}
+
+This experiment evaluates the empirical performance of the BIC criterion to select unknown cluster number in Section~\ref{sec:exten}.
+We generate the data from an order-3 Gaussian model with $p = \{50,80\}$, $r = \{2,4\}$, and noise level $\sigma^2 \in \{ 0.25,1\}$.
+Table~\ref{tab:select} shows that our BIC criterion well chooses the true $r$ under most settings.
+Note that the BIC slightly underestimates the true cluster number $(r = 4)$ with smaller dimension and higher noise $(p = 50, \sigma=1)$, and the accuracy immediately increases with larger dimension $p = 80$.
+The improvement follows from the fact that a larger dimension $p$ indicates a larger sample size in the tensor block model.
+Therefore, we conclude that BIC criterion is a reasonable way to tune the cluster number.
 
@@
     \end{tabularx}
-    \caption{Estimated cluster number given by BIC criterion under the low noise level $(\sigma^2 = 0.25)$ and high noise level $(\sigma^2 = 0.5)$ settings.
+    \caption{ \small Estimated cluster number given by BIC criterion under the low noise level $(\sigma^2 = 0.25)$ and high noise level $(\sigma^2 = 0.5)$ settings.
 Numbers in parentheses are standard deviations of $\hat r$ over 30 replications.}
@@
 
-The third experiment evaluates the empirical performance of the BIC criterion to select unknown cluster number.
-We generate the data from an order-3 Gaussian model with $p = \{50,80\}$, $r = \{2,4\}$, and noise level $\sigma^2 \in \{ 0.25,1\}$.
-Table~\ref{tab:select} shows that our BIC criterion well chooses the true $r$ under most settings.
-Note that the BIC slightly underestimates the true cluster number $(r = 4)$ with smaller dimension and higher noise $(p = 50, \sigma=1)$, and the accuracy immediately increases with larger dimension $p = 80$.
-The improvement follows from the fact that a larger dimension $p$ indicates a larger sample size in the tensor block model.
-Therefore, we conclude that BIC criterion is a reasonable way to tune the cluster number.
-
-\subsection{Comparison with other methods}\label{subsec:comp}
-
-We compare our algorithm with following higher-order clustering methods:
-\begin{itemize}[wide,topsep=-3pt,itemsep=0pt,parsep=1pt]
-    \item \textbf{\small HOSVD}: HOSVD on data tensor and $k$-means on the rows of the factor matrix;
-    \item \textbf{\small HOSVD+}: HOSVD on data tensor and $k$-means on the $\ell_2$-normalized rows of the factor matrix;
-    \item \textbf{\small HLloyd}~\citep{han2020exact}: High-order clustering algorithm developed for non-degree tensor block models;
-    \item \textbf{\small SCORE}~\citep{ke2019community}: Tensor-SCORE for clustering developed for binary tensors.
-\end{itemize}
-
-Among the four alternative algorithms, the \textbf{\small SCORE} is the closest method to ours.
-We set the tuning parameters of \textbf{\small SCORE} as in previous literature \citep{ke2019community}.
-The methods \textbf{\small SCORE} and \textbf{\small HOSVD+} are designed for degree models, whereas \textbf{\small HOSVD} and \textbf{\small HLloyd} are designed for non-degree models.
-We conduct two experiments to assess the impacts of (i) signal strength and (ii) degree heterogeneity, based on Gaussian and Bernoulli models with $ p = 100, r = 5$.
-We refer to our algorithm as \textbf{\small dTBM} in the comparison.
-
-We investigate the effects of signal to clustering performance by varying $\gamma \in [-1.5, -1.1]$.
-Figure~\ref{fig:comp_gamma} shows the consistent outperformance of our method \textbf{\small dTBM} among all algorithms.
-The sub-optimality of \textbf{\small SCORE} and \textbf{\small HOSVD+} indicates the necessity of local iterations on the clustering.
-Furthermore,  Figure~\ref{fig:comp_gamma} shows the inadequacy of non-degree algorithms in the presence of mild degree heterogeneity.
-The experiment demonstrates the benefits of addressing heterogeneity in higher-order clustering tasks.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=.8\columnwidth]{comp_gamma_anno3.pdf}
-    \caption{CER versus signal exponent (denoted $\gamma$) for different methods. We set $p = 100, r = 5, \gamma \in [-1.5, -1.1]$ under (a) Gaussian and (b) Bernoulli models.}
-    \label{fig:comp_gamma}
-\end{figure}
-
-The only exception in Figure~\ref{fig:comp_gamma} is the slightly better performance of \textbf{\small HLloyd} over \textbf{\small HOSVD+} under Gaussian model.
-However, we find the advantage of \textbf{\small HLloyd} disappears with higher degree heterogeneity.
-We perform extra simulations to verify the impact of degree effects.
-We use the same setting as in the first experiment in the Section~\ref{subsec:comp}, except that we now generate the degree heterogeneity $\mtheta$ from Pareto distribution prior to normalization.
-The density function of Pareto distribution is $f(x|a,b) = a b^a x^{-(a+1)} \ind\{ x \geq b \}$, where $a$ is called \emph{shape} parameter.
-We vary $a \in \{2,6\}$ and choose $b$ such that $\bbE X = a(a - 1)^{-1}b = 1$ for $X$ following Pareto$(a,b)$.
-Note that a smaller $a$ leads to a larger variance in $\mtheta$ and hence a larger degree heterogeneity.
+\subsection{Extra Simulation Results for Comparison}\label{sec:extrasim}
+
+We provide extra simulation results for the comparison with other higher-order clustering algorithms.
+We use the same setting as in the first experiment in the Section~\ref{subsec:comp}, except that we now generate the degree heterogeneity $\mtheta$ from Pareto distribution with shape parameter $a$ prior to normalization.
 We consider the Gaussian model under low $(a = 6)$ and high $(a = 2)$ degree heterogeneity.
-Figure~\ref{fig:comp_gamma_theta} shows that the errors for non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) increase with degree heterogeneity.
+Figure~\ref{fig:comp_gamma_theta} shows that the errors for non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) increases with degree heterogeneity.
 In addition, the advantage of \textbf{\small HLloyd} over \textbf{\small HOSVD+} disappears with higher degree heterogeneity.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=.8\columnwidth]{comp_theta_anno3.pdf}
-    \caption{CER versus shape parameter in degree (denoted $a\in[3,6]$) for different methods. We set $p = 100, r = 5, \gamma = -1.2$ under (a) Gaussian and (b) Bernoulli models.}
-    \label{fig:comp_theta}
-\end{figure}
+This experiment supports the conclusion we obtained in Section~\ref{subsec:comp}.
 
@@
     \includegraphics[width=.8\columnwidth]{comp_gamma_theta3.pdf}
-    \caption{CER comparison versus signal exponent (denoted $\gamma$) under (a) low (shape parameter $a = 6$)  (b) high (shape parameter $a = 2$) degree heterogeneity.
+    \caption{ \small CER comparison versus signal exponent (denoted $\gamma$) under (a) low (shape parameter $a = 6$)  (b) high (shape parameter $a = 2$) degree heterogeneity.
 We set $p = 100, r = 5, \gamma \in [-1.5, -1.1]$ under Gaussian model.}
@@
 
-The last experiment investigates the effects of degree heterogeneity to clustering performance.
-We fix the signal exponent $\gamma = -1.2$ and vary the extent of degree heterogeneity.
-In this experiment, we generate $\mtheta$ from Pareto distribution prior to normalization.
-We vary the shape parameter $a \in [3,6]$ in the Pareto distribution to investigate a range of degree heterogeneities.
-Figure~\ref{fig:comp_theta} demonstrates the stability of degree-corrected algorithms (\textbf{\small dTBM}, \textbf{\small SCORE}, \textbf{\small HOSVD+}) over the entire range of degree heterogeneity under consideration.
-In contrast, non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) show poor performance with large heterogeneity, especially in Bernoulli cases.
-This experiment, again, highlights the benefit of addressing degree heterogeneity in higher-order clustering.
-
-\section{Real data applications}\label{sec:real}
-\subsection{Human brain connectome data analysis}
+\subsection{Human Brain Connectome Data Analysis}
 
@@
     \centering
-    \includegraphics[width = .8\columnwidth]{brain_node_cluster.pdf}
-    \caption{Illustration of brain node clustering results for HCP data with (a) top and (b) side views. }
+    \includegraphics[width = 1\columnwidth]{brain_node_cluster.pdf}
+    \caption{ \small Illustration of brain node clustering results for HCP data with (a) top and (b) side views. }
     \label{fig:cluster_brain}
 \end{figure}
-
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width = 1\textwidth]{est_S_anno.pdf}
-    \caption{Mode 3 slices of estimated core tensor $\hat \tS$.
-(a) Average estimated slice weighted by the group size; (b)-(d) Group-specified enrichment, i.e., the difference between each slice of $\hat \tS$ and the averaged slice.
-}
-    \label{fig:ests}
-\end{figure*}
-
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width = 1\textwidth]{brain_connection.pdf}
-    \caption{Observed brain connections in the population and each group of individuals.
-(a) Average brain network; (b)-(d) Group-specified brain network enrichments in Groups 1-3.
-Red edges represent the positive enrichment and blue edges represent the negative enrichment.}
-    \label{fig:brain_conn}
-\end{figure*}
 
@@
 
-\subsection{Peru Legislation data analysis}
-
-We also apply our method to the legislation networks in the Congress of the Republic of Peru \citep{lee2017time}.
-Because of the frequent political power shifts in the Peruvian Congress during 2006-2011, we choose to focus on the data for the first half of 2006-2007 year.
-The dataset records the co-sponsorship of 116 legislators from top 5 parties and 802 bill proposals.
-We reconstruct legislation network as an order-3 binary tensor $\tY \in \{0,1\}^{116 \times 116 \times 116}$, where $\tY_{ijk} = 1$ if the legislators $(i,j,k)$ have sponsored the same bill, and $\tY_{ijk} = 0$ otherwise.
-The true party affiliations of legislators are provided and serve as the ground truth.
-We apply various higher-order clustering methods to $\tY$ with $r = 5$.
-Table~\ref{tab:peru} shows that our \textbf{\small dTBM} achieves the best performance compared to others.
-The second best method is the two-stage algorithm \textbf{\small HLloyd}, followed by the spectral methods \textbf{\small SCORE} and \textbf{\small HOSVD+}.
-This result is consistent with our simulations under strong signal and moderate degree heterogeneity.
-The comparison suggests that our method \textbf{\small dTBM} is more appealing in real-world applications.
-
-\begin{table}[ht]
+\begin{figure*}[htb]
     \centering
-    \begin{tabular}{c |c  c cc c}
-    \hline
-        Method & \textbf{\small dTBM} 
-        &\textbf{\small HOSVD}
-        &\textbf{\small HOSVD+} & \textbf{\small HLloyd} &  \textbf{\small SCORE}\\
-         CER & \textbf{0.116}
-         &  0.22 
-         &0.213 & 0.149 &0.199\\
-         \hline
-    \end{tabular}
-    \caption{Clustering errors (measured by CER) for various methods in the analysis of Peru Legislation dataset.}
-    \label{tab:peru}
-\end{table}
-
-\section{Proof Sketches}\label{sec:mainproof}
-
-In this section, we provide the proof sketches for the main Theorems~\ref{thm:initial}-\ref{thm:refinement}. Detail proofs and extra theoretical results are provided in Appendix.
-
-\subsection{Proof sketch of Theorem~\ref{thm:initial}}
-The proof of Theorem~\ref{thm:initial} is inspired by the proof idea of \citet[Lemma 1]{gao2018community}.
-The extra difficulties are the angle gap characterization and multilinear algebra property in tensors; we address both challenges in our proof.
-Specifically, we control the misclustering error by the estimation error of $\hat \tX$ calculated in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}.
-We prove the following inequality
-\begin{equation}\label{eq:proof_4}
-    \ell(z^{(0)},z) \lesssim \frac{1}{p}\min_{\pi \in \Pi} \sum_{i: z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2 \lesssim \frac{\sigma^2 r^{K-1}}{ \Delta_{\min}^2 p^K} \onormSize{}{\hat \tX - \tX}_F^2 \lesssim \frac{r^K p^{-K/2}}{\text{SNR}},
-\end{equation}
-where $\tX = \bbE \tY$ is the true mean.
-The first inequality in~\eqref{eq:proof_4} holds with the assumption $\min_{i \in [p]} \theta(i) \geq c>0$ in Theorem~\ref{thm:initial}.
-The second inequality relies on an important conclusion that the angle gap of mean tensor $\tX$ is lower bounded by that of core tensor $\tS$, i.e., the minimal angle gap $\Delta_{\min}$ defined in Assumption~\ref{assmp:min_gap}.
-Let $\ma^{s} \coloneqq \ma/\onorm{\ma}$ denote the normalized vector with the convention that $\ma^s = 0$ if $\ma = 0$.
-We want to show that
-\begin{equation}\label{eq:proof_4_gap}
-\min_{z(i) \neq z(j)} \onormSize{}{[\mX_{i:}]^s - [\mX_{j:}]^s} \gtrsim \Delta_{\min},
-\end{equation}
-where $\mX = \mat(\tX)$.
-The most challenging part in the proof of Theorem~\ref{thm:initial} lies in the derivation of inequality~\eqref{eq:proof_4_gap}, in which the proof of \cite{gao2018community} is no longer applicable due to different angle gap assumption in our dTBM.
-We develop the extra padding technique in Lemma~\ref{lem:pad} and balance assumption~\eqref{eq:degree} to derive~\eqref{eq:proof_4_gap}.
-Last, we finish the proof of Theorem~\ref{thm:initial} by showing the third inequality of~\eqref{eq:proof_4} using \citet[Proposition 1]{han2020exact}.
-
-\subsection{Proof sketch of Theorem 5}
-The proof of Theorem~\ref{thm:refinement} is inspired by the proof idea of \citet[Theorem 2]{han2020exact}.
-We develop extra polar-coordinate based techniques with angle gap characterization to address the nuisance degree heterogeneity.
-We introduce an intermediate quantity called misclustering loss
-\begin{equation}
-     L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
-\end{equation}
-where the superscript $\cdot^{s}$ denotes the normalized vector; i.e., $\ma^s:=\ma/\onorm{\ma}$ if $\ma\neq 0$ and $\ma^{s}=0$ if $\ma=0$ for any vector $\ma$.
-We show that $L^{(t)}$ provides an upper bound for the misclassification error of interest via the inequality $\ell^{(t)}\leq {L^{(t)}\over \Delta^2_{\min}}$.
-Therefore, it suffices to control $L^{(t)}$.
-Further, we introduce the oracle estimators for core tensor under the true cluster assignment via
-\begin{equation}
-    \tilde \tS = \tY \times_1 \mW^T \times_2 \cdots \times_K \mW^T, 
-\end{equation}
-where $\mW = \mM \of{ \text{diag}(\mone_{p}^T \mM) }^{-1}$ is the weighted true membership matrix.
-Let $ \mV = \mW^{\otimes (K-1)}$ denote the Kronecker product of $(K-1)$ copies of $\mW$ matrices, and we define the $t$-th iteration quantities $\mW^{(t)}, \mV^{(t)}$ corresponding to $\mM^{(t)}$ (or equivalently $z^{(t)}$).
-To evaluate $L^{(t+1)}$, we prove the bound
-\begin{equation}\label{eq:proof_5_event}
-    \ind \offf{ z^{(t+1)}(i) = b } = \ind \offf{       \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{b:}^{(t)}]^s }^2 \leq \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{z(i):}^{(t)}]^s }^2} \leq A_{ib} + B_{ib},
-\end{equation}
-where $\mY = \mat(\tY)$, $ \mS = \mat(\tS)$, $\mS^{(t)} = \mat(\tS^{(t)})$ and
-\begin{align}
-        A_{ib} &= \ind \offf{\ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s } \lesssim -  \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 },\\
-        B_{ib} &= \ind \offf{\onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 \lesssim F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }.
-\end{align}
-The terms $F_{ib}^{(t)}, G_{ib}^{(t)}, H_{ib}^{(t)}$ are controlled by $z^{(t)}, \tS^{(t)}$; see the detailed definitions in \eqref{eq:f}, \eqref{eq:g}, \eqref{eq:h}.
-Note that the event $A_{ib}$ only involves the oracle estimator independent of $t$, while all the terms related to the $t$-th iteration are in $B_{ib}$.
-Thus, the inequality~\eqref{eq:proof_5_event} decomposes the misclustering loss in the $(t+1)$-th iteration into the oracle loss and the loss in $t$-th iteration.
-This decomposition leads to the separation of statistical error and computational error in the final upper bound of Theorem~\ref{thm:refinement}.
-
-Specifically, we prove the contraction inequality
-\begin{equation}\label{eq:proof_5_ineq}
-    L^{(t+1)} \leq M \xi + \rho L^{(t)}, \quad \xi = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  A_{ib} \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
-\end{equation}
-where $M$ is a positive constant, $\rho \in (0,1)$ is the contraction parameter, and we call $\xi$ the oracle loss.
-Controlling the probability of event $B_{ib}$ and obtaining the $\rho L^{(t)}$ term in the right hand side of~\eqref{eq:proof_5_ineq} are the most challenging parts in the proof of Theorem~\ref{thm:refinement}.
-Note that the true and estimated core tensors are involved via their normalized rows such as $\mS_{a:}^s, \tilde \mS_{a:}^s, [\mS^{(t)}_{a:}]^s$.
-The Cartesian coordinate based analysis in \cite{han2020exact} is no longer applicable in our case.
-Instead, we use the polar-coordinate based analysis and the geometry property of trigonometric functions to derive the high probability upper bounds for $F_{ib}^{(t)}, G_{ib}^{(t)}, H_{ib}^{(t)}$.
-
-Further, by sub-Gaussian concentration, we prove the high probability upper bound for oracle loss
-\begin{equation}\label{eq:proof_5_xi}
-    \xi  \lesssim \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}.
-\end{equation}
-Combining the decomposition~\eqref{eq:proof_5_ineq} and the oracle bound~\eqref{eq:proof_5_xi}, we finish the proof of Theorem~\ref{thm:refinement}.
-
-\section*{Acknowledgments}
-This research is supported in part by NSF grants DMS-1915978, DMS-2023239, EF-2133740, and funding from the Wisconsin Alumni Research foundation.
-We thank Zheng Tracy Ke, Rungang Han, Yuetian Luo for helpful discussions and for sharing software packages.
-
-\bibliographystyle{apalike}
-\bibliography{tensor_wang}
-
-\newpage
-\appendix
-\section*{Appendices}
+    \includegraphics[width = 1\textwidth]{est_S_anno.pdf}
+    \caption{ \small Mode 3 slices of estimated core tensor $\hat \tS$.
+(a) Average estimated slice weighted by the group size; (b)-(d) Group-specified enrichment, i.e., the difference between each slice of $\hat \tS$ and the averaged slice.
+}
+    \label{fig:ests}
+\end{figure*}
+
+\begin{figure*}[htb]
+    \centering
+    \includegraphics[width = 1\textwidth]{brain_connection.pdf}
+    \caption{ \small Observed brain connections in the population and each group of individuals.
+(a) Average brain network; (b)-(d) Group-specified brain network enrichments in Groups 1-3.
+Red edges represent the positive enrichment and blue edges represent the negative enrichment.}
+    \label{fig:brain_conn}
+\end{figure*}
+
+\section{PROOFS}
 
@@
 
-\section{Notation}
+\subsection{Notation}
 Before the proofs, we first introduce the notation used throughout the appendix and the generalized dTBM without symmetric assumptions.
@@
 
-\section*{Proof of Theorem~\ref{thm:unique}}
+\subsection{Proof of Theorem~\ref{thm:unique}}
 
@@
 
-\section{Proof of Theorem~\ref{thm:stats}}
+\subsection{Proof of Theorem~\ref{thm:stats}}
 
@@
 \begin{equation}
-    \cos(\mx_1, \mx_2) = \frac{\and{\mx_1, \mx_2}}{\onorm{\mx_1} \onorm{\mx_2} } = \frac{(p/r)^{K-1} \ang{\mS_{1:}, \mS_{2:}}}{ \sqrt{(p/r)^{K-1} \onorm{\mS_{1:}}^2} \sqrt{(p/r)^{K-1} \onorm{\mS_{2:}}^2} } = \cos(\mS_{1:}, \mS_{2:}),
+    \cos(\mx_1, \mx_2) = \frac{\ang{\mx_1, \mx_2}}{\onorm{\mx_1} \onorm{\mx_2} } = \frac{(p/r)^{K-1} \ang{\mS_{1:}, \mS_{2:}}}{ \sqrt{(p/r)^{K-1} \onorm{\mS_{1:}}^2} \sqrt{(p/r)^{K-1} \onorm{\mS_{2:}}^2} } = \cos(\mS_{1:}, \mS_{2:}),
 \end{equation}
@@
 
-\section{Proof of Theorem~\ref{thm:comp}}
+\subsection{Proof of Theorem~\ref{thm:comp}} \label{subsec:thm_comp}
 
@@
 
-\section{Proof of Theorem~\ref{thm:initial}}
+{\bf Useful Definition and Conjecture for Theorem~\ref{thm:comp}}
+
+\begin{defn}[Hypergraphic planted clique (HPC) detection]~\label{def:HPC} Consider an order-$K$ hypergraph $H = (V,E)$ where $V = [p]$ collects vertices and $E$ collects all the order-$K$ edges.
+Let $\tH_k(p, 1/2)$ denote the Erd\H{o}s-R\'{e}nyi $K$-hypergraph where the edge $(i_1,\ldots, i_K)$ belongs to $E$ with probability $1/2$.
+Further, we let $\tH_K(p, 1/2, \kappa)$ denote the hyhpergraph with planted cliques of size $\kappa$.
+Specifically, we generate a hypergraph from $\tH_k(p, 1/2)$, pick $\kappa$ vertices uniformly from $[p]$, denoted $K$, and then connect all the hyperedges with vertices in $K$.
+Note that the clique size $\kappa$ can be a function of $p$, denoted $\kappa_p$.
+The order-$K$ HPC detection aims to identify whether there exists a planted clique hidden in an Erd\H{o}s-R\'{e}nyi $K$-hypergraph.
+The HPC detection is formulated as the following hypothesis testing problem
+\begin{equation}
+    H_0:\ H \sim \tH_K(p,1/2) \quad \text{versus} \quad H_1: \ H \sim \tH_K(p,1/2, \kappa_p).
+\end{equation}
+\end{defn}
+
+\begin{conjecture}[HPC conjecture]\label{hypo:HPC} Consider the HPC detection problem in Definition~\ref{def:HPC}.
+Suppose the sequence $\{\kappa_p\}$ such that $\limsup_{p \rightarrow \infty} \log \kappa_p/ \log \sqrt{p} \leq (1 - \tau)$.
+Then, for every sequence of polynomial-time test $\{ \varphi_p\}: H \mapsto \{0,1\}$ we have
+\begin{equation}
+    \liminf_{p \rightarrow \infty} \bbP_{H_0} \of{ \varphi_p(H) =1 } +  \bbP_{H_1} \of{ \varphi_p(H) =0} \geq \frac{1}{2}.
+\end{equation}
+\end{conjecture}
+
+\subsection{Proof of Theorem~\ref{thm:initial}}
 \begin{proof}[Proof of Theorem 4] We prove Theorem~\ref{thm:initial} under the symmetric dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
@@
 
-\section{Proof of Theorem~\ref{thm:refinement}}
+\subsection{Proof of Theorem~\ref{thm:refinement}}
 
@@
 
-    {\bf Step 3 (Derivation of contraction inequality).} In this step we derive the upper bound of $\zeta_{ib}$ and obtain the contraction inequality~\eqref{eq:proof_5_ineq}.
+    {\bf Step 3 (Derivation of contraction inequality).} In this step we derive the upper bound of $\zeta_{ib}$ and obtain the contraction inequality.
 Choose the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ that satisfies the condition of Lemma~\ref{lem:upper_fgh}, inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.

### 2022-02-18 tensor_clustering/tensor_clustering_aistats_camera/algorithm.tex
aistats

--- 2022-01-20 tensor_clustering/tensor_clustering_arxiv/tensor_clustering_arxiv.tex
@@
-\documentclass[11pt]{article}
-\usepackage{graphicx}
-\usepackage{subfigure}
-\usepackage{caption}
-\usepackage[hidelinks]{hyperref}
-\usepackage[margin=1in]{geometry}
-% \geometry{
-%  left=30mm,
-%  right=30mm
-%  }
+\documentclass[10pt]{article}
+\usepackage[paperheight=5.35in,paperwidth=7.1in,margin=0.05in]{geometry}
 
-\usepackage{algpseudocode}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{enumerate}
-\usepackage[authoryear]{natbib}
+\usepackage{setspace}
+\usepackage{amsmath,amssymb}
+\usepackage{amsthm}
+\usepackage{fancybox}
 \usepackage{url}
 \usepackage{enumitem}
+\usepackage{multirow}
+\usepackage{color}
+\usepackage{graphicx}
+\usepackage{setspace}
+\usepackage{comment}
+\usepackage{bm}
+\usepackage{mathtools}
+\mathtoolsset{showonlyrefs=true}
 
-\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
-\usepackage{algorithm}
-\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
-\hypersetup{colorlinks,linkcolor=blue,citecolor=blue,urlcolor={black}}
-\usepackage{lscape}
-\usepackage{afterpage}
-\usepackage{hyperref}
-
-\usepackage{microtype}
-\usepackage{wrapfig}
-\allowdisplaybreaks
 \usepackage{algorithm}
@@
 
-\usepackage{textcomp}
-\usepackage{stfloats}
-\usepackage{verbatim}
-\usepackage{array}
-
-\usepackage{tabularx,booktabs}
-\newcolumntype{Y}{>{\centering\arraybackslash}X}
-
-\theoremstyle{definition}
-\newtheorem{thm}{Theorem}
-\newtheorem{lem}{Lemma}
-\newtheorem{prop}{Proposition}[section]
-\newtheorem{pro}{Property}
-\newtheorem{cor}{Corollary}
-
-\theoremstyle{definition}
-\newtheorem{assumption}{Assumption}
-\newtheorem{defn}{Definition}
-\newtheorem{example}{Example}
-\newtheorem{rmk}{Remark}
-\newtheorem{condition}{Condition}
-\newtheorem{conjecture}{Conjecture}
-\newtheorem{property}{Property}
-\newtheorem{hypothesis}{Hypothesis}
-
-\newtheorem{innercustomgeneric}{\customgenericname}
-\providecommand{\customgenericname}{}
-\newcommand{\newcustomtheorem}[2]{%
-  \newenvironment{#1}[1]
-  {%
-   \renewcommand\customgenericname{#2}%
-   \renewcommand\theinnercustomgeneric{##1}%
-   \innercustomgeneric
-  }
-  {\endinnercustomgeneric}
-}
-
-\newcustomtheorem{customexample}{Example}
-
-\usepackage{appendix}
-\usepackage{fullpage} 
-\usepackage{microtype}
-\usepackage{wrapfig}
-\mathtoolsset{showonlyrefs}
-
-\newcommand{\of}[1]{\left(#1\right)}
-\newcommand{\off}[1]{\left[#1\right]}
-\newcommand{\offf}[1]{\left\{#1\right\}}
-\newcommand{\aabs}[1]{\left|#1\right|}
-\newcommand{\ang}[1]{\left\langle#1\right\rangle}
-
-\newcommand{\distgap}{D_{\ell_2}}
-\newcommand{\Mat}{\text{Mat}}
-
-\def\ci{\perp\!\!\!\perp}
-
-\usepackage{booktabs}
-\newcommand\doubleRule{\toprule\toprule}
-\allowdisplaybreaks
-
-\newcommand*{\KeepStyleUnderBrace}[1]{%f
-  \mathop{%
-    \mathchoice
-    {\underbrace{\displaystyle#1}}%
-    {\underbrace{\textstyle#1}}%
-    {\underbrace{\scriptstyle#1}}%
-    {\underbrace{\scriptscriptstyle#1}}%
-  }\limits
-}
+\usepackage{natbib}
+\usepackage{xr}
 
 \input macros.tex
-\newcommand*{\Scale}[2][4]{\scalebox{#1}{$#2$}}%
-\setlength\parindent{0pt}
-\usepackage[parfill]{parskip}
+\def\sign{\textup{sgn}}
+
 \usepackage{algorithm}
@@
 
-\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
-\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
+%\input macros.tex
+
+\def\@normalsize{\@setsize\normalsize{11pt}\xpt\@xpt}
 
@@
 \usepackage{caption}
-\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}
-
-\title{Multiway Spherical Clustering via Degree-Corrected\\ Tensor Block Models\footnote{Part of this paper was presented at 25th International Conference on Artificial Intelligence and Statistics (AISTATS).}}
-\date{}
-\author{%
-%\small
-Jiaxin Hu \\
-University of Wisconsin -- Madison\\
-\texttt{jhu267@wisc.edu} \\
-\and
-Miaoyan Wang \\
-University of Wisconsin -- Madison\\
-\texttt{miaoyan.wang@wisc.edu}\\
-}
 
@@
 
-\maketitle
-%\doublespacing
-
-\begin{abstract}
-We consider the problem of multiway clustering in the presence of unknown degree heterogeneity.
-Such data problems arise commonly in applications such as recommendation system, neuroimaging, community detection, and hypergraph partitions in social networks.
-The allowance of degree heterogeneity provides great flexibility in clustering models, but the extra complexity poses significant challenges in both statistics and computation.
-Here, we develop a degree-corrected tensor block model with estimation accuracy guarantees.
-We present the phase transition of clustering performance based on the notion of angle separability, and we characterize three signal-to-noise regimes corresponding to different statistical-computational behaviors.
-In particular, we demonstrate that an intrinsic statistical-to-computational gap emerges only for tensors of order three or greater.
-Further, we develop an efficient polynomial-time algorithm that provably achieves exact clustering under mild signal conditions.
-The efficacy of our procedure is demonstrated through two data applications, one on human brain connectome project, and another on Peru Legislation network dataset.
-\end{abstract}
-
-\textbf{Keywords:} tensor clustering, degree correction, statistical-computational efficiency, human brain connectome networks
-
-\section{Introduction}\label{sec:intro}
-Multiway arrays have been widely collected in various fields including social networks \citep{anandkumar2014tensor}, neuroscience \citep{wang2017bayesian}, and computer science \citep{koniusz2016sparse}.
-Tensors effectively represent the multiway data and serve as the foundation in higher-order data analysis.
-One data example is from multi-tissue multi-individual gene expression study~\citep{wang2019three,hore2016tensor}, where the data tensor consists of expression measurements indexed by (gene, individual, tissue) triplets.
-Another example is \emph{hypergraph} network~\citep{ghoshdastidar2017uniform,ghoshdastidar2017consistency,ahn2019community,ke2019community} in social science.
-A $K$-uniform hypergraph can be naturally represented as an order-$K$ tensor, where each entry indicates the presence of $K$-way hyperedge among nodes (a.k.a.\ entities).
-In both examples, identifying the similarity among tensor entities is important for scientific discovery.
-
-We study the problem of multiway clustering based on a data tensor.
-The goal of multiway clustering is to identify a checkerboard structure from a noisy data tensor.
-Figure~\ref{fig:intro} illustrates the noisy tensor and the underlying checkerboard structures discovered by multiway clustering methods.
-In the hypergraph example, the multiway clustering aims to identify the underlying block partition of nodes based on their higher-order connectivities; therefore, we also refer to the clustering as \emph{higher-order clustering}.
-The most common model for higher-order clustering is called \emph{tensor block model} (TBM)~\citep{wang2019multiway}, which extends the usual matrix stochastic block model~\citep{abbe2017community} to tensors.
-The matrix analysis tools, however, are sub-optimal for higher-order clustering.
-Developing tensor tools for solving block models has received increased interest recently~\citep{ wang2019multiway,chi2020provable,han2020exact}.
-
-Classical tensor block model suffers from drawbacks to model real world data in spite of the popularity.
-The key underlying assumption of block model is that all nodes in the same community are exchangeable; i.e., the nodes have no individual-specific parameters apart from the community-specific parameters.
-However, the exchangeability assumption is often non-realistic.
-Each node may contribute to the data variation by its own multiplicative effect.
-We call the unequal node-specific effects the \emph{degree heterogeneity}.
-Such degree heterogeneity appears commonly in social networks.
-Ignoring the degree heterogeneity may seriously mislead the clustering results.
-For example, regular block model fails to model the member affiliation in the Karate Club network~\citep{bickel2009nonparametric} without addressing degree heterogeneity.
-
-The \emph{degree-corrected tensor block model} (dTBM) has been proposed recently to account for the degree heterogeneity~\citep{ke2019community}.
-The dTBM combines a higher-order checkerboard structure with degree parameter $\mtheta=(\mtheta(1),\ldots,\mtheta(p))^T$ to allow heterogeneity among $p$ nodes.
-Figure~\ref{fig:intro} compares the underlying structures of TBM and dTBM with the same number of communities.
-The dTBM allows varying values within the same community, thereby allowing a richer structure.
-To solve dTBM, we project clustering objects to a unit sphere and perform iterative clustering based on angle similarity.
-We refer to the algorithm as the \textit{spherical} clustering; detailed procedures are in Section~\ref{sec:alg}.
-The spherical clustering avoids the estimation of nuisance degree heterogeneity.
-The usage of angle similarity brings new challenges to the theoretical results, and we develop new polar-coordinate based techniques in the proofs.
-
-\begin{figure*}[t]
-    \centering
-    \includegraphics[width = .9\textwidth]{intro2_arxiv.pdf}
-    \caption{Examples for order-3 tensor block model (TBM) with and without degree correction.
-Both TBM and dTBM have four communities on each mode, while dTBM allows a richer structure with degree heterogeneity.
-    }
-    \label{fig:intro}
-\end{figure*}
-
-{\bf Our contributions.} The primary goal of this paper is to provide both statistical and computational guarantees for dTBM. Our main contributions are summarized below.
-\begin{itemize}[leftmargin=*]
-
- \item We develop a general dTBM and establish the identifiability for the uniqueness of clustering using the notion of angle separability.
-
-\item  We present the phase transition of clustering performance with respect to three different statistical and computational behaviors.
-We characterize, for the first time, the critical signal-to-noise (SNR) thresholds in dTBMs, revealing the intrinsic distinctions among (vector) one-dimensional clustering, (matrix) biclustering, and (tensor) higher-order clustering.
-Specific SNR thresholds and algorithm behaviors are depicted in  Figure~\ref{fig:phase_axis}.
-
- \item We provide an angle-based algorithm that achieves exact clustering \emph{in polynomial time} under mild conditions.
-Simulation and data studies demonstrate the outperformance of our algorithm compared with existing higher-order clustering algorithms.
-\end{itemize}
-The last two contributions, to our best knowledge, are new to the literature of dTBMs.
-
-\begin{figure*}[t]
-    \centering
-    \includegraphics[width = 17cm]{phase.pdf}
-    \caption{SNR thresholds for statistical and computational limits in order-$K$ dTBM with dimension $(p,...,p)$ and $K \geq 2$.
-The SNR gap between statistical possibility and computational efficiency  exists only for tensors with $K \geq 3$.
-}
-    \label{fig:phase_axis}
-\end{figure*}
-
-{\bf Related work.} 
-Our work is closely related to but also distinct from several lines of existing research. Table~\ref{tab:comp} summarizes the most relevant models.
-
-\begin{itemize}[wide]
-    \item \textit{Block model for clustering.} Block models such as stochastic block model (SBM) and degree-corrected SBM have been widely used for matrix clustering problems.
-The theoretical properties and algorithm performance for matrix block models have been well-studied~\citep{gao2018community}; see the review paper~\citep{abbe2017community} and the references therein.
-However, The tensor counterparts are relatively less understood.
-
-    \item  \textit{Tensor block model.} The (non-degree) tensor block model (TBM) is a higher-order extension of SBM, and its statistical-computational properties are investigated in recent literatures~\citep{wang2019multiway, han2020exact, ghoshdastidar2017consistency}.
-Extending results from non-degree to degree-corrected model is highly challenging.
-Our dTBM parameter space is equipped with angle-based similarity and nuisance degree parameters.
-The extra complexity makes the Cartesian coordinates based analysis~\citep{han2020exact} non-applicable to our setting.
-Towards this goal, we have developed a new polar coordinates based analysis to control the model complexity.
-We also develop a new angle-based iteration algorithm to achieve optimal clustering rates \emph{without the need of estimating nuisance degree parameters}.
-
-    \item \textit{Degree-corrected block model.} The hypergraph degree-corrected block model (hDCBM) and its variant have been proposed in the literature~\citep{ke2019community, yuan2018testing}.
-For this popular model, however, the optimal statistical-computational rates remain an open problem.
-Our main contribution is to provide a sharp statistical and computational critical phase transition in dTBM literature.
-In addition, our algorithm results in a faster \emph{exponential} error rate, in contrast to the \emph{polynomial} rate in~\cite{ke2019community}.
-The original hDCBM~\citep{ke2019community} is designed for binary observations only, and we extend the model to both continuous and binary observations.
-We believe our results are novel and helpful to the community.
-See Fig~\ref{fig:phase_axis} for overview of our results.
-
-    \item \textit{Global-to-local algorithm strategy.} Our methods generalize the recent global-to-local strategy for matrix learning~\citep{gao2018community,chi2019nonconvex,yun2016optimal} to tensors~\citep{han2020exact,ahn2018hypergraph,kim2018stochastic}.
-Despite the conceptual similarity, we address several fundamental challenges associated with this non-convex, non-continuous problem.
-We show the insufficiency of the conventional tensor HOSVD~\citep{de2000multilinear}, and we develop a weighted higher-order initialization that relaxes the singular-value gap separation condition.
-Furthermore, our local iteration leverages the angle-based clustering in order to avoid explicit estimation of degree heterogeneity.
-Our bounds reveal the interesting interplay between the computational and statistical errors.
-We show that our final estimate \emph{provably} achieves the exact clustering within only polynomial-time complexity.
-
-\end{itemize}
-
-\begin{table*}[t]
-\resizebox{\textwidth}{!}{%
-    \begin{tabular}{c|ccccc}
-\hline
-    & \cite{gao2018community}&  \cite{han2020exact}&  \cite{ghoshdastidar2017consistency} &\cite{ke2019community} & \textbf{Ours}\\
-    \hline
-      Allow tensors of arbitrary order & $\times$ & $\surd$ & $\surd$ &$\surd$ & $\surd$  \\
-        Allow degree heterogeneity &  $\surd$ & $\times$ & $\surd$ &$\surd$ & $\surd$ \\
-        Singular-value gap-free clustering &  $\surd$ & $\surd$ & $\times$ & $\times$ &$\surd$ \\
-        Misclustering rate (for order $K^*$)& - & $\exp(-p^{K/2})$ & $p^{-1}$& $p^{-2}$ &  $\exp(-p^{K/2})$\\
-        \hline
-    \end{tabular}
-    }
-    \caption{Comparison between previous methods with our method. $^*$We list the result for order-K tensors with $K \geq 3$ and general number of communities $r = \tO(1)$. }\label{tab:comp}
-\end{table*}
-
-{\bf Notation.} We use lower-case letters (e.g., $a,b$) for scalars, lower-case boldface letters (e.g., $\ma,\mtheta$) for vectors, upper-case boldface letters (e.g., $\mX,\mY$) for matrices, and calligraphy letters (e.g., $\tX,\tY$) for tensors of order three or greater.
-We use $\mone_p$ to denote a vector of length $p$ with all entries to be 1.
-We use $|\cdot|$ for the cardinality of a set and $\ind\{\cdot\}$ for the indicator function.
-For an integer $p\in\mathbb{N}_{+}$, we use the shorthand $[p]= \offf{1,2,...,p}$.
-For a length-$p$ vector $\ma$, we use $a(i)\in\mathbb{R}$ to denote the $i$-th entry of $\ma$, and use $\ma_{I}$ to denote the sub-vector by restricting the indices in the set $I\subset [p]$.
-We use  $\onorm{\ma}=\sqrt{\sum_{i}a^2(i)}$ to denote the $\ell_2$-norm, $\onorm{\ma}_1=\sum_i |a_i|$ to denote the $\ell_1$ norm of $\ma$.
-For two vector $\ma, \mb$ of the same dimension, we denote the angle between $\ma, \mb$ by
-\begin{equation}
-    \cos \of{\ma, \mb} = \frac{\ang{ \ma, \mb}}{ \onorm{\ma} \onorm{\mb} },
-\end{equation}
-where $\ang{\ma,\mb}$ is the inner product of two vectors and $\cos \of{\ma, \mb} \in [-1,1]$. We make the convention that $\cos \of{\ma, \mb} = \cos \of{\ma^T, \mb^T}$.
-
-Let $\tY  \in \bbR^{p_1 \times \cdots \times p_K}$ be an order-$K$ $(p_1,...,p_K)$-dimensional tensor.
-We use $\tY(i_1,\ldots,i_K)$ to denote the $(i_1,\ldots,i_K)$-th entry of $\tY$.
-The multilinear multiplication of a tensor $\tS\in \bbR^{r_1\times \cdots \times r_K}$ by matrices $\mM_k \in\mathbb{R}^{p_k\times r_k}$ results in an order-$d$ $(p_1,\ldots,p_K)$-dimensional tensor $\tX$, denoted
-\[
-\tX=\tS \times_1 \mM_1 \times \cdots \times_K \mM_K,
-\]
-where the entries of $\tX$ are defined by
-\begin{align}
-    &\tX(i_1,\ldots, i_K)=\sum_{(j_1,\ldots,j_K)}\tS(j_1,\ldots,j_K)\mM_1(i_1,j_1)\cdots \mM_K(i_K,j_K).
-\end{align} 
-For a matrix $\mY$, we use $\mY_{i:}$ (respectively, $\mY_{:i}$) to denote the $i$-th row (respectively, $i$-th column) of the matrix.
-Similarly, for an order-3 tensor, we use $\tY_{::i}$ to denote the $i$-th matrix slide of the tensor.
-We use $\text{Ave}(\cdot)$ to denote the operation of taking averages across elements and $\text{Mat}_k(\cdot)$ to denote the unfolding operation that reshapes the tensor along mode $k$ into a matrix.
-For a symmetric tensor $\tY\in\mathbb{R}^{p\times \cdots \times p}$, we omit the subscript and use $\text{Mat}(\tY)\in\mathbb{R}^{p\times p^{K-1}}$ to denote the unfolding.
-For two sequences $\{a_p\}, \{b_p\}$, we denote $a_p\lesssim b_p$ or $a_p=\tO(b_p)$ if $\lim_{p\to\infty}a_p /b_p\leq c$ for some constant $c\geq 0$, $a_p=o(b_p)$ if $\lim_{p\to\infty}a_p/b_p =0$, and $a_p = \Omega(b_p)$ if both $b_p \lesssim a_p$ and $a_p\lesssim b_p$.
-Throughout the paper, we use the terms ``community'' and ``clusters'' exchangeably.
-
-{\bf Organization.} The rest of this paper is organized as follows.
-Section~\ref{sec:model} introduces the degree-corrected tensor block model (dTBM) with three motivating examples and presents the identifiability of dTBM under the angle gap condition.
-We show the phase transition and the existence of statistical-computational gaps for the higher-order dTBM in Section~\ref{sec:limits}.
-In Section~\ref{sec:alg}, we provide a polynomial-time two-stage algorithm with misclustering rate guarantees.
-Numerical studies including the simulation, comparison with other methods, and two real dataset analyses are in Sections~\ref{sec:simulation}-\ref{sec:real}.
-The main technical ideas we develop for addressing main theorems are provided in Section~\ref{sec:mainproof}.
-Detailed proofs and extra theoretical results are provided in Appendix.
-
-\section{Model formulation and motivations}\label{sec:model}
-
-\subsection{Degree-corrected tensor block model}
-
-Suppose we have an order-$K$ data tensor $\tY \in \bbR^{p \times \cdots \times p}$.
-For ease of notation, we focus on symmetric tensors in this section; the extension to general asymmetric tensors is provided in Section~\ref{subsec:exten}.
-Assume there exist $r \geq 2$  disjoint communities among the $p$ nodes.
-We represent the community assignment by a function $z \colon [p]\mapsto[r]$, where $z(i) = a$ for $i$-th node that belongs to the $a$-th community.
-Then, $z^{-1}(a)=\{i\in[p]\colon z(i)=a\}$ denotes the set of nodes that belong to the $a$-th community, and $|z^{-1}(a)|$ denotes the number of nodes in the $a$-th community.
-Let $\mtheta=(\theta(1),\ldots,\theta(p))^T$ denote the degree heterogeneity for $p$ nodes.
-We consider the order-$K$ dTBM~\citep{ghoshdastidar2017consistency,ke2019community},
-\begin{equation}\label{eq:model_margin}
-    \tY(i_1,\ldots,i_K) = \tS(z(i_1),\ldots,z(i_K)) \prod_{k = 1}^K\theta_{i_k} + \tE(i_1,\ldots,i_K), 
-\end{equation}
-\normalsize
-where $\tS \in \bbR^{r \times \cdots \times r}$ is an order-$K$ tensor collecting the block means among communities, and $\tE\in\mathbb{R}^{p\times \cdots \times p}$ is a noise tensor consisting of independent zero-mean sub-Gaussian entries with variance bounded by $\sigma^2$.
-The unknown parameters are $z$, $S$, and $\mtheta$.
-The dTBM can be equivalently written in a compact form of tensor-matrix product:
-\begin{equation}\label{eq:model_tensor}
-\bbE\tY = \tS \times_1 \mTheta \mM \times_2 \cdots \times_K  \mTheta \mM,
-\end{equation}
-where $\mTheta = \text{diag}(\theta(1),...,\theta(p)) \in \bbR^{p \times p}$ is a diagonal matrix, $\mM \in \offf{0,1}^{p \times r}$ is the membership matrix associated with community assignment $z$ such that $\mM(i,j)=\ind\{z(i)=j\}$.
-By definition, each row of $\mM$ has one copy of 1's and 0's elsewhere.
-Note that the discrete nature of $\mM$ renders our model~\eqref{eq:model_tensor} more challenging than Tucker decomposition.
-We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$ if $\tY$ admits dTBM~\eqref{eq:model_tensor}.
-The goal of clustering is to estimate $z$ from a single noisy tensor $\tY$.
-We are particularly interested in the high-dimensional regime where $p$ grows whereas $r=\tO(1)$.
-
-\subsection{Motivating examples} Here, we provide four applications to illustrate the practical necessity of dTBM.
-
-\vspace{0.2cm}
-\paragraph{Tensor block model} Consider the model~\eqref{eq:model_tensor}.
-Let $\theta(i)=1$ for all $ i \in [p]$.
-The model~\eqref{eq:model_tensor} reduces to the tensor block model, which is widely used in previous clustering algorithms~\citep{wang2019multiway,chi2020provable,han2020exact}.
-The theoretical results in TBM serve as benchmarks for dTBM.
-
-\vspace{0.2cm}
-
-\paragraph{Community detection in hypergraphs} Hypergraph network is a powerful tool to represent the complex entity relations with higher-order interactions~\citep{ke2019community}.
-A typical undirected hypergraph is denoted as $H = (V,E)$, where $V = [p]$ is the set of nodes and $E$ is the set of undirected hyperedges.
-Each hyperedge in $E$ is a subset of $V$, and we call the hyperedge an order-$K$ edge if the corresponding subset involves $K$ nodes.
-We call $H$ a $K$-uniform hypergraph if $E$ only contains order-$K$ edges.
-
-It is natural to represent the $K$-uniform hypergraph using a binary order-$K$ adjacency tensor.
-Let $\tY \in \{0,1\}^{p \times \cdots \times p}$ denote the adjacency tensor, where the entries encode the presence or absence of order-$K$ edges among $p$ nodes.
-Specifically, for all $(i_1,\ldots,i_K) \in [p]^K$, we have
-\begin{equation}
-    \tY(i_1,...,i_K) =  \begin{cases}
-    1  & \text{if }  (i_1,...,i_K) \in E,\\
-    0 & \text{if }  (i_1,...,i_K) \notin E.
-    \end{cases}
-\end{equation}
-
-Assume there exist $r$ disjoint communities among $p$ nodes, and the connection probabilities depend on the community assignments and node-specific parameters. 
-Then, the equation~\eqref{eq:model_tensor} models $\bbE \tY$ with unknown degree heterogeneity $\mtheta$ and sub-Gaussianity parameter $\sigma^2 = 1/4$.
-
-\vspace{0.2cm}
-\paragraph{Multi-layer weighted network} 
-Multi-layer weighted network data consists of multiple networks over the same set of nodes.
-One representative example is the brain connectome data~\citep{zhang2019tensor}.
-The multi-layer weighted network $\tY$ has dimension of $p \times p \times L$, where $p$ denotes the number of brain regions of interest, and $L$ denotes the number of layers (networks).
-Each of the $L$ networks describes one aspect of the brain connectivity, such as functional connectivity or structural connectivity.
-The resulting tensor $\tY$ consists of a mixture of slices with various data types.
-
-Assume there exist $r$ disjoint communities among $p$ nodes and $r_l$ disjoint communities among the $L$ layers.
-The multi-layer network community detection is modeled by the generalized asymmetric dTBM model~\eqref{eq:model_tensor}
-\begin{equation}
-    \bbE \tY = \tS \times_1 \mTheta \mM \times_2 \mTheta \mM \times_3 \mTheta_l \mM_l, 
-\end{equation}
-where $(\mtheta \in \bbR^p, \mM \in \{0,1\}^{p \times r})$ and $ (\mtheta_l \in \bbR^L, \mM_l \in \{0,1\}^{L \times r_l})$ are the degree heterogeneity and membership matrices corresponding to the community structure for $p$ nodes and $L$ layers, respectively.
-\vspace{0.2cm}
-
-\paragraph{Gaussian higher-order clustering} Datasets in various fields such as medical image, genetics, and computer science are formulated as Gaussian tensors.
-One typical example is the multi-tissue gene expression dataset, which records the different gene expression in different individuals and different tissues.
-The dataset, denoted as $\tY \in \bbR^{p \times n \times t}$, consists of the expression data for $p$ genes of $n$ individuals in $t$ tissues.
-
-Assume there exist $r_1, r_2, r_3$ disjoint clusters for $p$ genes, $n$ individuals, and $t$ tissues, respectively. We apply the generalized asymmetric dTBM model~\eqref{eq:model_tensor} 
-\begin{equation}
-    \bbE\tY = \tS \times_1 \mTheta_1 \mM_1 \times_2 \mTheta_2 \mM_2 \times_3 \mTheta_3 \mM_3, 
-\end{equation}
-where $\{(\mtheta_k, \mM_k)\}_{k=1}^3$ represents the heterogeneity and membership for genes, individuals, and tissues. \\
-
-\begin{rmk}[Comparison with non-degree models]
-Our dTBM uses fewer block parameters than TBM.
-In particular, every non-degree $r_1$-block tensor can be represented by a \emph{degree-corrected} $r_2$-block tensor with $r_2\leq r_1$.
-In particular, there exist tensors with $r_1=p$ but $r_2=1$, so the reduction in model complexity can be dramatic from $p$ to 1.
-This fact highlights the benefits of introducing degree heterogeneity in higher-order clustering tasks.
-\end{rmk}
-
-\subsection{Identifiability under angle gap condition}
-The goal of clustering is to estimate the partition function $z$ from model~\eqref{eq:model_tensor}.
-For ease of notation, we focus on symmetric tensors; the extension to non-symmetric tensors are similar.
-We use $\tP$ to denote the following parameter space for $(z,\tS,\mtheta)$,
-\begin{align}\label{eq:family}
-\tP=\bigg\{(z,\tS,\mtheta)\colon  \ \mtheta\in\mathbb{R}^p_{+},\ 
-{c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r}, c_3\leq \onormSize{}{\text{Mat}(\tS)_{a:}}\leq c_4, \onorm{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|, a\in[r]\bigg\},
-\end{align}
-\normalsize
-where $c_i>0$'s are universal constants. We briefly describe the rationale of the constraints in~\eqref{eq:family}. 
-First, the entrywise positivity constraint on  $\mtheta\in\mathbb{R}^p_{+}$ is imposed to avoid sign ambiguity between entries in $\mtheta_{z^{-1}(a)}$ and $\tS$.
-This constraint allows the trigonometric $\cos$ to describe the angle similarity in the Assumption~\ref{assmp:min_gap} below and Sub-algorithm~\hyperref[alg:main]{2} in Section~\ref{sec:alg}.
-Note that the positivity constraint can be achieved without sacrificing model flexibility, by using a slightly larger dimension of $\tS$ in the factorization~\eqref{eq:model_tensor}; see Example~\ref{ex:positive} below.
-Second, recall that the quantity $|z^{-1}(a)|$ denotes the number of nodes in $a$-th community.
-The constants $c_1, c_2$ in the $|z^{-1}(a)|$ bound assume the roughly balanced size across $r$ communities.
-Third, the constants $c_3, c_4 $ in the magnitude of $\text{Mat}(\tS)_{a:}$ requires no purely zero slide in $\tS$, so the core tensor $\tS$ is not trivially reduced to a lower rank.
-Lastly, the $\ell_1$ normalization $\onormSize{}{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|$ is imposed to avoid the scalar ambiguity between $\mtheta_{z^{-1}(a)}$ and $\tS$.
-This constraint, again, incurs no restriction to model flexibility but makes our presentation cleaner.
-
-\begin{example}[Positivity of degree parameters]~\label{ex:positive}
-Here we provide an example to show the positivity constraints on $\mtheta$ incurs no loss on the model flexibility. 
-Consider an order-3 dTBM with core tensor $\tS = 1$ and degree $\mtheta = (1,1,-1, -1)^T$. We have the mean tensor 
-\begin{equation}
-    \tX = \tS \times_1 \mTheta \mM \times_2 \mTheta \mM \times_3 \mTheta \mM,
-\end{equation}
-where $\mTheta = \text{diag}(\mtheta)$ and $\mM = (1,1,1,1)^T$.
-Note that $\tX \in \bbR^{4 \times 4 \times 4}$ is a 1-block tensor with \emph{mixed-signed} degree $\mtheta$, and the mode-3 slices of $\tX$ are
-\begin{equation}
-    \tX_{::1} = \tX_{::2} = - \tX_{::3} = -\tX_{::4}  = \begin{bmatrix}
-    1 & 1& -1& -1 \\
-    1 & 1& -1& -1 \\
-    -1& -1&1 & 1\\
-     -1& -1&1 & 1
-    \end{bmatrix}.
-\end{equation}
-Now, instead of original decomposition, we encode $\tX$ as a 2-block tensor with \emph{positive-signed} degree. Specifically, we write
-\begin{equation}
-     \tX = \tS' \times_1 \mTheta' \mM' \times_2 \mTheta' \mM'  \times_3 \mTheta' \mM' , 
-\end{equation}
-where $\mTheta' = \text{diag}(\mtheta') = \text{diag}(1,1, 1, 1)$, the core tensor $\tS' \in \bbR^{2 \times 2 \times 2}$ has mode-3 slices, and the membership matrix $\mM'\in\{0,1\}^{4\times 2}$ defines the clustering $z'\colon[4]\to[2]$,
-\begin{equation}
-    \tS'_{::1} = -\tS'_{::2} = \begin{bmatrix}
-    1 & -1 \\
-    -1 & 1
-    \end{bmatrix}, \quad \mM' = \begin{bmatrix}
-   1 & 0 \\
-   1 &0 \\
-    0 & 1 \\
-    0&1
-    \end{bmatrix}.
-\end{equation}
- The triplet $(z',\tS', \mtheta')$ lies in our parameter space~\eqref{eq:family}.
-In general, we can always reparameterize a block-$r$ tensor with mixed-signed degree using a block-$2r$ tensor with positive-signed degree.
-Since we assume $r=\tO(1)$ throughout the paper, the splitting does not affect the error rates of our interest.\end{example}
-
-We now provide the identifiability conditions for our model before estimation procedures.
-When $r=1$, the decomposition~\eqref{eq:model_tensor} is always unique (up to cluster label permutation) in $\tP$, because dTBM is equivalent to the rank-1 tensor family under this case.
-When $r\geq 2$, the Tucker rank of signal tensor $\bbE\tY$ in~\eqref{eq:model_tensor} is bounded by, but not necessarily equal to, the number of blocks $r$~\citep{wang2019multiway}.
-Therefore, one can not apply the classical identifiability conditions for low-rank tensors to dTBM.
-Here, we introduce a key separation condition on the core tensor.
-
-\begin{assumption}[Angle gap] \label{assmp:min_gap}Let $\mS = \text{Mat}(\tS)$. Assume the minimal gap between normalized rows of $\mS$ is bounded away from zero; i.e.,
-\begin{equation}\label{eq:minimal_gap}
-    \Delta_{\min} := \min_{a \neq b \in [r]} \onorm{\frac{\mS_{a:}}{\onormSize{}{ \mS_{a:}}} - \frac{\mS_{b:}}{\onormSize{}{ \mS_{b:}}} }>0, \quad \text{for} \quad r \geq 2.
-\end{equation}
-\end{assumption}
-We make the convention $\Delta_{\min} = 1$ for $r = 1$.
-Equivalently, \eqref{eq:minimal_gap} says that none of the two rows in $\mS$ are parallel; i.e., $\max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}}  = 1-\Delta^2_{\min}/2<1$.
-The quantity $\Delta_{\min}$ characterizes the non-redundancy among clusters measured by angle separation.
-The denominators involved in definition~\eqref{eq:minimal_gap} are well posed because of the lower bound on $\onorm{\mS_{a:}}$ in~\eqref{eq:family}.
-
-Our first main result is the following theorem showing the sufficiency and necessarily of the angle gap separation condition for the parameter identifiability under dTBM.
-
-\begin{thm}[Model identifiability]\label{thm:unique} Consider the dTBM with $r\geq 2$.
-The parameterization~\eqref{eq:model_tensor} is unique in $\tP$ up to cluster label permutations, if and only if Assumption~\ref{assmp:min_gap} holds.
-\end{thm}
-
-The identifiability guarantee for the dTBM is more appealing than classical Tucker model.
-In the Tucker model, the factor matrix $\mM$ is identifiable only up to orthogonal rotations.
-In contrast, our model does not suffer from rotational invariance.
-As we will show in Section~\ref{sec:alg}, each column of the membership matrix $\mM$ can be precisely recovered under our algorithm.
-This property benefits the interpretation of dTBM in practice.
-
-\section{Statistical-computational limits for higher-order tensors}\label{sec:limits}
-
-In this section, we study the statistical and computational limits of dTBM. We propose signal-to-noise ratio~(SNR),
-\begin{align}\label{eq:gamma}
-  \text{SNR}:= \Delta^2_{\min}/\sigma^2 = p^{\gamma}, 
-\end{align}
-with varying $\gamma \in \bbR$ that quantifies different regimes of interest.
-We call $\gamma$ the \emph{signal exponent}.
-Intuitively, a larger SNR, or equivalently a larger $\gamma$, benefits the clustering in the presence of noise.
-With quantification~\eqref{eq:gamma}, we consider the following parameter space,
-\begin{equation}\label{eq:gammafamily}
-    \tP(\gamma) = \tP\cap\{\tS \text{ satisfies SNR condition~\eqref{eq:gamma} with $\gamma$} \}.
-\end{equation}
-Note that $1$-block dTBM does not belong to the space $\tP(\gamma)$ when $\gamma < 0$ by Assumption~\ref{assmp:min_gap}.
-Our goal is to characterize the clustering accuracy with respect to $\gamma$.
-Let $\hat z$ and $z$ be the estimated and true clustering functions in the family~\eqref{eq:family}.
-Define the misclustering error by
-\[
-\ell(\hat z, z)={1\over p}\min_{\pi \in \Pi}\sum_{i\in[p]}\ind\{\hat z(i)\neq \pi \circ z(i)\},
-\]
-where $\pi: [r] \mapsto [r]$ is a permutation of cluster labels, $\circ$ denotes the composition operation, and $\Pi$ denotes the collection of all possible permutations.
-The infinitum over all permutations accounts for the ambiguity in cluster label permutation.
-
-In Sections~\ref{sec:statlimit} and \ref{sec:complimit}, we provide the lower bounds of $\ell (\hat z, z)$ for general Gaussian dTBMs~\eqref{eq:model_margin} without symmetric assumptions.
-For general (asymmetric) Gaussian dTBMs, we assume Gaussian noise $\tE(i_1,\ldots,i_K)\stackrel{\text{i.i.d.}}{\sim} N(0,\sigma^2)$, and we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode.
-For notational simplicity, we still use $z$ and $\tP(\gamma)$ for this general (asymmetric) model.
-All lower bounds should be interpreted as the worst-case results across $K$ modes.
-
-\subsection{Statistical critical values}\label{sec:statlimit}
-The statistical limit means the minimal SNR required for solving dTBMs with \emph{unlimited computational cost.} Our following result shows the minimax lower bound of SNR for exact recovery in dTBM.
-
-\begin{thm}[Statistical lower bound]\label{thm:stats} Consider general Gaussian dTBMs under the parameter space $\tP(\gamma)$ with $K\geq 1$.
-Assume $r \lesssim p^{1/3}$.
-If the signal exponent satisfies $\gamma < -(K-1)$, then, every estimator $\hat z_{\text{stat}}$ obeys
-\begin{equation}
-    \sup_{(z, \tS, \mtheta) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
-\end{equation}
-\end{thm}
-Theorem~\ref{thm:stats} demonstrates the impossibility of exact recovery of the assignment when $\gamma < -(K-1)$ in the high-dimensional regime $p\to \infty$ for fixed $r$.
-The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to, maximum likelihood estimation (MLE)~\citep{wang2019multiway} and trace maximization~\citep{ghoshdastidar2017uniform}.
-As we will show in Section~\ref{sec:alg}, the SNR threshold $-(K-1)$ is also a minimax upper bound, because MLE achieves exact recovery when $\gamma > -(K-1)$.
-Hence, the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM.
-
-\subsection{Computational critical values}\label{sec:complimit}
-In this section, we derive the computational limits of dTBMs.
-The computational limit means the minimal SNR required for exactly recovery with \emph{polynomial-time} computational cost.
-An important ingredient to establish the computational limits is the \emph{hypergraphic planted clique (HPC) conjecture} \citep{zhang2018tensor, brennan2020reducibility}.
-The HPC conjecture indicates the impossibility of fully recovering the planted cliques with polynomial-time algorithm when the clique size is less than the number of vertices in the hypergraph.
-The formal statement of HPC detection conjecture is provided in Definition~\ref{def:HPC} and Conjecture~\ref{hypo:HPC} as follows.
-
-\begin{defn}[Hypergraphic planted clique (HPC) detection]~\label{def:HPC} Consider an order-$K$ hypergraph $H = (V,E)$ where $V = [p]$ collects vertices and $E$ collects all the order-$K$ edges.
-Let $\tH_k(p, 1/2)$ denote the Erd\H{o}s-R\'{e}nyi $K$-hypergraph where the edge $(i_1,\ldots, i_K)$ belongs to $E$ with probability $1/2$.
-Further, we let $\tH_K(p, 1/2, \kappa)$ denote the hyhpergraph with planted cliques of size $\kappa$.
-Specifically, we generate a hypergraph from $\tH_k(p, 1/2)$, pick $\kappa$ vertices uniformly from $[p]$, denoted $K$, and then connect all the hyperedges with vertices in $K$.
-Note that the clique size $\kappa$ can be a function of $p$, denoted $\kappa_p$.
-The order-$K$ HPC detection aims to identify whether there exists a planted clique hidden in an Erd\H{o}s-R\'{e}nyi $K$-hypergraph.
-The HPC detection is formulated as the following hypothesis testing problem
-\begin{equation}
-    H_0:\ H \sim \tH_K(p,1/2) \quad \text{versus} \quad H_1: \ H \sim \tH_K(p,1/2, \kappa_p).
-\end{equation}
-\end{defn}
-
-\begin{conjecture}[HPC conjecture]\label{hypo:HPC} Consider the HPC detection problem in Definition~\ref{def:HPC}.
-Suppose the sequence $\{\kappa_p\}$ such that $\limsup_{p \rightarrow \infty} \log \kappa_p/ \log \sqrt{p} \leq (1 - \tau)$.
-Then, for every sequence of polynomial-time test $\{ \varphi_p\}: H \mapsto \{0,1\}$ we have
-\begin{equation}
-    \liminf_{p \rightarrow \infty} \bbP_{H_0} \of{ \varphi_p(H) =1 } +  \bbP_{H_1} \of{ \varphi_p(H) =0} \geq \frac{1}{2}.
-\end{equation}
-\end{conjecture}
-
-Under the HPC conjecture, we establish the SNR lower bound that is necessary for any \emph{polynomial-time} estimator to achieve exact clustering.
-
-\begin{thm}[Computational lower bound]\label{thm:comp} Consider general Gaussian dTBMs under the parameter space $\tP(\gamma)$ with $K\geq 2$.
-Assume HPC conjecture holds.
-If the signal exponent $\gamma < -K/2$, then, every \emph{polynomial-time estimator} $\hat z_{\text{comp}}$ obeys
-\begin{align}
-   \liminf_{p\to \infty}\sup_{(z, \tS, \mtheta) \in \tP(\gamma)}  \bbE \left[ p\ell(\hat z_{\text{comp}}, z) \right]\geq 1.
-\end{align}
-\end{thm}
-Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$.
-Therefore, $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM.
-In Section~\ref{sec:alg}, we will show the condition $\gamma> -K/2$ suffices for our proposed polynomial-time estimator.
-Thus, $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM.
-
-\begin{rmk}[Statistical-computational gaps]
-Now, we have established the phase transition of exact clustering under order-$K$ dTBM by combing Theorems~\ref{thm:stats} and \ref{thm:comp}.
-Figure~\ref{fig:phase_axis} summarizes our results of critical SNRs when $K \geq 2$.
-In the weak SNR region $\gamma < -(K-1)$, no statistical estimator succeeds in degree-corrected higher-order clustering.
-In the strong SNR region $\gamma  > -K/2$, our proposed algorithm precisely recovers the clustering in polynomial time.
-In the moderate SNR regime, $-(K-1)\leq \gamma \leq -K/2$, the degree-corrected clustering problem is statistically easy but computationally hard.
-Particularly, dTBM reduces to matrix degree-corrected model when $K =2$, and the statistical and computational bounds show the same critical value.
-When $K =1$, dTBM reduces to the degree-corrected sub-Gaussian mixture model (GMM) with model
-\begin{equation}
-    \mY = \mTheta \mM \mS + \mE,
-\end{equation}
-where $\mY \in \bbR^{p \times d}$ collects $n$ data points in $\bbR^d$, $\mS \in \bbR^{r \times d}$ collects the $d$-dimensional centroids for $r$ clusters, and $\mTheta \in \bbR^{p \times p}, \mM \in \{0,1\}^{p \times r}, \mE \in \bbR^{p \times d}$ have the same meaning as in dTBM.
-\cite{lu2016statistical} implies that polynomial-time algorithms are able to achieve the statistical minimax lower bound in GMM.
-Therefore, we conclude that the statistical-to-computational gap emerges only for higher-order tensors with $K \geq 3$.
-The result reveals the intrinsic distinctions among (vector) one-dimensional clustering, (matrix) biclustering, and (tensor) higher-order clustering.
-\end{rmk}
-
-\begin{rmk}[Comparison with non-degree models]
-We compare our results to non-degree tensor models.
-The allowance of degree heterogeneity $\mtheta$ makes the model more flexible, but it incurs extra statistical and computational complexity.
-Fortunately, we find that the extra complexity does not render the estimation of $z$ qualitatively harder; see the comparison of our phase transition with non-degree TBM~\citep{han2020exact}.
-\end{rmk}
-
-\section{Polynomial-time algorithm under mild SNR}\label{sec:alg}
-In this section, we present an efficient polynomial-time clustering algorithm under mild SNR.
-The procedure takes a global-to-local approach.
-See Figure~\ref{fig:demo} for illustration.
-The global step finds the basin of attraction with polynomial miclustering error, whereas the local iterations improve the initial clustering to exact recovery.
-Both steps are critical to obtain a satisfactory algorithm output.
-In what follows, we first use the symmetric tensor as a working example to describe the algorithm procedures to gain insight.
-Our theoretical analysis focuses on the noisy tensor with i.i.d.\ sub-Gaussian noise such as Gaussian and uniform observations.
-The extensions for asymmetric tensor and Bernoulli observation and other practical issues are in Section~\ref{subsec:exten}.
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=\columnwidth]{alg_demo1.pdf}
-\caption{Illustration of our global-to-local algorithm.}\label{fig:demo}
-\end{figure}
-
-\subsection{Weighted higher-order initialization}
-
-We start with weighted higher-order clustering algorithm as initialization.
-We take an order-3 symmetric tensor as illustration for insight.
-Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
-By model~\eqref{eq:model_tensor}, for all $i \in [p]$, we have
-\begin{equation}\label{eq:kmeans}
-    \theta(i)^{-1} \mX_{i:} = \off{\text{Mat}( \tS \times_2 \mTheta \mM \times_3  \mTheta \mM )}_{z(i):}. 
-\end{equation}
-This implies that, all node $i$ belonging to $a$-th community (i.e., $z(i)=a$) share the same normalized mean vector $\theta(i)^{-1} \mX_{i:}$, and vice versa.
-Intuitively, one can apply $k$-means clustering to the vectors $\{ \theta(i)^{-1} \mX_{i:} \}_{i\in[p]}$, which leads to main idea of our Sub-algorithm~\hyperref[alg:main]{1}.
-
-Specifically, our initialization consists of denoising step and clustering step.
-The denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) estimates $\tX$ from $\tY$ by a double projection spectral method.
-The first projection performs HOSVD~\citep{de2000multilinear} via $\mU_{\text{pre}} = \text{SVD}_{r} \of{ \text{Mat}(\tY) }$, where $\text{SVD}_r(\cdot)$ returns the top-$r$ left singular vectors.
-The second projection performs HOSVD on the projected $\tY$ onto the multilinear Kronecker space $\mU_{\text{pre}} \otimes  \mU_{\text{pre}}$; i.e.,
-\begin{equation}\label{eq:two-step_factor}
-    \hat \mU = \text{SVD}_{r} \of{\text{Mat}\of{ \tY\times_1  \mU_{\text{pre}} \mU_{\text{pre}}^T \times_2   \mU_{\text{pre}} \mU_{\text{pre}}^T }}.
-\end{equation}
-The final denoised tensor $\hat \tX$ is defined by
-\begin{equation}\label{eq:two-step_est}
-    \hat \tX = \tY \times_1 \hat \mU \hat 
-\mU^T \times_2 \hat \mU \hat \mU^T \times_3 \hat \mU \hat \mU^T. 
-\end{equation}
-The double projection improves usual matrix spectral methods in order to alleviate the noise effects for $K\geq 3$~\citep{han2020exact}.
-
-The clustering step (lines 3-5 in Sub-algorithm~\hyperref[alg:main]{1}) performs the weighted $k$-means clustering. 
-We write $\hat \mX=\text{Mat}(\hat \tX)$, and normalize the rows into $\hat \mX^s_{i:}=\onormSize{}{\hat \mX_{i:}}^{-1}\hat \mX_{i:}$ as a surrogate of $\theta(i)^{-1} \mX_{i:}$.
-Then, a weighted $k$-means clustering is performed on the normalized rows with weights equal to $\onormSize{}{\hat \mX_{i:}}^2$.
-The choice of weights is to bound the $k$-means objective function by the Frobenius-norm accuracy of $\hat \tX$.
-Unlike existing clustering algorithm~\citep{ke2019community}, we apply the clustering on the unfolded tensor $\hat \mX$ rather than on the factors $\hat \mU$.
-This strategy relaxes the singular-value gap condition~\citep{gao2018community, han2020exact}.
-We assign degenerate rows with purely zero entries to an arbitrarily random cluster; these nodes are negligible in high-dimensions because of the lower bound on $\onormSize{}{\text{Mat}(\tS)_{a:}}$ in~\eqref{eq:family}.
-The final result gives the initial cluster assignment $\hat z^{(0)}$.
-Full procedures are provided in Sub-algorithm~\hyperref[alg:main]{1}.
-
-\begin{algorithm}[h!]
+\begin{algorithm}[h]
 \caption*{\bf Algorithm: Multiway spherical clustering for degree-corrected tensor block model }
@@
 
-We now establish the misclustering error rate of initialization. We call $\mtheta$ is balanced, if the relative extent of heterogeneity is comparable across clusters in that
-\begin{equation}\label{eq:degree}
-{\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
-\end{equation}
-Note that, the assumption~\eqref{eq:degree} does not preclude degree heterogeneity.
-Indeed, within each of the clusters, the highest degree can be $\theta(i) = \Omega(p)$, whereas the lowest degree can be $\theta(i)=\tO(1)$.
+%We refer to the combination of Algorithms~\ref{alg:initial}+ \ref{alg:refinement} as provable dTBM algorithm.
 
-\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with i.i.d.\ noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
-Assume $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$.
-Let $ z^{(0)}$ denote the output of Sub-algorithm~\hyperref[alg:main]{1}.
-With probability going to 1, we have
-\begin{equation}\label{eq:ini}
-   \ell(z^{(0)}, z) \lesssim {r^K p^{-K/2}\over \text{SNR}}. 
-\end{equation}
-\end{thm}
-
-\begin{rmk}[Comparison to previous results] For fixed SNR, our initialization error rate with $K=2$ agrees with the initialization error rate $\tO(p^{-1})$ in matrix models~\citep{gao2018community}.
-Furthermore, in the special case of non-degree TBMs with $\theta_1=\cdots=\theta_p=1$, we achieve the same initial misclustering error $\tO(p^{-K/2})$ as in non-degree models~\citep{han2020exact}.
-Theorem~\ref{thm:initial} implies the advantage of our algorithm in achieving both accuracy and model flexibility.
-\end{rmk}
-
-\begin{rmk}[Failure of conventional tensor HOSVD] If we use conventional HOSVD for tensor denoising; that is, we use $\mU_{\text{pre}}$ in place of $\hat \mU$ in line 2, then the misclustering rate becomes $\tO(p^{-1})$ for all $K\geq 2$.
-This rate is substantially worse than our current rate~\eqref{eq:ini}.
-\end{rmk}
-
-\begin{rmk}[Singular-value gap-free clustering] Note that our clustering directly applies to the estimated mean tensor $\hat \tX$ rather than the leading tensor factors $\hat \mU$.
-Applying clustering to the tensor factors suffers from the non-identifiability issue due to the infinitely many orthogonal rotations when the number of blocks $r \geq 3$ in the absence of singular-value gaps.
-Such ambiguity causes the trouble for effective clustering~\citep{abbe2020entrywise}.
-In contrast, our initialization algorithm applies the clustering to the overall mean tensor $\hat \tX$.
-This strategy avoids the non-identifiability issue regardless of the number of blocks and singular-value gaps.
-\end{rmk}
-
-\subsection{Angle-based iteration}
-Our Theorem~\ref{thm:initial} has shown the polynomially decaying error rate from our initialization.
-Now we improve the error rate to exponential decay using local iterations.
-We propose an angle-based local iteration to improve the outputs from Sub-algorithm~\hyperref[alg:main]{1}.
-To gain the intuition, consider an one-dimensional degree-corrected clustering problem with data vectors $\mx_i = \theta(i) \ms_{z(i)} + \mepsilon_i, i \in [p]$, where $\ms_i$'s are known cluster centroids, $\theta(i)$'s are unknown positive degrees, and $z\colon [p] \mapsto [r]$ is the cluster assignment of interest.
-The angle-based $k$-means algorithm estimates the assignment $z$ by minimizing the angle between data vectors and centroids; i.e.,
- \begin{equation}\label{eq:angle_kmeans}
-     z(i) = \argmax_{a \in [r]} \cos ( \mx_i,\ \ms_{a} ),\ \text{ for all }i \in [p].
- \end{equation}
-The classical Euclidean-distance based clustering~\citep{han2020exact} fails to recover $z$ in the presence of degree heterogeneity, even under noiseless case.
-In contrast, the proposed angle-based $k$-means achieves accurate recovery without explicit estimation of $\mtheta$.
-
-Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as in angle-based $k$-means.
-We still take the order-3 tensor for illustration.
-Specifically, Sub-algorithm~\hyperref[alg:main]{2} updates estimated core tensor and cluster assignment in each iteration.
-We use superscript $\cdot^{(t)}$ to denote the estimate from $t$-th iteration, where $t=1,\ldots.$ For core tensor, we consider the following update strategy
- \[
- \tS^{(t)}(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k\in[3]\}.
-\]
-Intuitively, $\tS^{(t)}$ becomes closer to the true core $\tS$ as $z^{(t)}$ is more precise.
-For cluster assignment, we first aggregate the slices of $\tY$ and obtain a reduced tensor $\tY^{\text{d}}\in \bbR^{p \times r \times r}$ with given $z^{(t)}$, where
-\[
-\tY^{\text{d}}(i,a_2,a_3)=\text{Ave}\{\tY(i,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k \neq 1\}.
-\]
-We use $\mY^d$ and $ \mS^{(t)}$ to denote the $\text{Mat}(\tY^{\text{d}})$ and $\text{Mat}(\tS^{(t)})$.
-The rows $\mY^d_{i:}$ and $\mS^{(t)}_{a:}$ correspond to the $\mx_i$ and $\ms_a$ in the one-dimensional clustering~\eqref{eq:angle_kmeans}.
-Then, we obtain the updated assignment by
-\[
-z(i)^{(t+1)} = \argmax_{a \in [r]} \cos \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} },\ \text{ for all } i \in [p],
-\]
-provided that $\mS^{(t)}_{a:}$ is a non-zero vector.
-Otherwise, if $\mS^{(t)}_{a:}$ is a zero vector, then we make the convention to assign $z^{(t+1)}(i)$ randomly in $[p]$.
-Full procedures for our angle-based iteration are described in Sub-algorithm~\hyperref[alg:main]{2}.
-
-We now establish the misclustering error rate of iterations under the stability assumption.
-
-\begin{defn}[Locally linear stability] \label{def:stable}
-Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\varepsilon)=\{\bar z\colon \ell(\bar z, z) \leq \varepsilon\}$.
-Let $\bar z\colon[p]\to [r]$ be a clustering function.
-We define two vectors associated with $\bar z$,
-\begin{align}
-    &\mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T, \quad \mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
-\end{align}
-
-We call the degree is $\varepsilon$-locally linearly stable if and only if 
-\begin{equation}\label{eq:local}
-    \sin(\mp(\bar z),\ \mp_{\mtheta}(\bar z))\lesssim \varepsilon \Delta_{\min} ,\quad \text{for all } \bar z\in \tN(z, \varepsilon).
-\end{equation}
-\end{defn}
-
-Roughly speaking, the vector $\mp(\bar z)$ represents the raw cluster sizes, and $\mp_{\theta}(\bar z)$ represents the relative cluster sizes weighted by degrees. 
-The local stability holds trivially for $\varepsilon=0$ based on the construction of parameter space~\eqref{eq:family}.
-The condition~\eqref{eq:local} controls the impact of node degree to the $\mp_{\theta}(\cdot)$ with respect to the misclassification rate $\varepsilon$ and angle gap.
-
-\begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the setup as in Theorem~\ref{thm:initial}. Suppose 
-$r=\tO(1)$ and $\text{SNR} \geq \tilde C p^{-K/2}\log p$ for some sufficiently large constant $\tilde C$.
-Assume the local linear stability of degree holds in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \geq \check{C}\log^{-1}p $ with some positive constant $\check{C}$.
-Let $z^{(t)}$ denote the $t$-th iteration output in Sub-algorithm~\hyperref[alg:main]{2} with initialization $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{1}.
-With probability going to 1, there exists a contraction parameter $\rho \in (0,1)$ such that
-\begin{align}\label{eq:final}
-    \ell(z, \hat z^{(t+1)}) \lesssim &\ \KeepStyleUnderBrace{
-   \text{SNR}^{-1}
-    \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}}_{\substack{\text{statistical error}} }+ \KeepStyleUnderBrace{ \rho^t \ell(z, z^{(0)}). }_{\substack{\text{computational error}}}
-\end{align}
-\end{thm}
-From the conclusion~\eqref{eq:final}, we find that the iteration error is decomposed into two parts: statistical error and computational error.
-The statistical error is unavoidable with noisy data regardless $t$, whereas the computational error decays in an exponential rate as the number of iterations $t \rightarrow \infty$.
-
-Theorem~\ref{thm:refinement} implies that, with probability going to 1, our estimate $z^{(T)}$ achieves exact recovery within polynomial iterations; more precisely,
-\begin{equation}
-     z^{(T)} = \pi \circ z, \quad \text{for all }T\gtrsim \log_{1/\rho} p,
-\end{equation}
-for some permutation $\pi \in \Pi$.
-Therefore, our combined algorithm is \textit{computationally efficient} as long as SNR $\gtrsim p^{-K/2} \log p$.
-Note that, ignoring the logarithmic term, the minimal SNR requirement, $p^{-K/2}$, coincides with the computational lower bound in Theorem~\ref{thm:comp}.
-Therefore, our algorithm is optimal regarding the signal requirement and lies in the sharpest \emph{computationally efficient} regime in Figure~\ref{fig:phase_axis}.
-
-\subsection{Extensions and practical issues}~\label{subsec:exten}
-\hspace{-.3cm} {\bf Extension for Bernoulli observations.} The main difficulty to establish the statistical guarantee for Bernoulli observations lies in the initialization Sub-algorithm~\hyperref[alg:main]{1}.
-Theorem~\ref{thm:refinement} still holds for Bernoulli observations once the initialization accuracy satisfies the upper bound~\eqref{eq:ini} in Theorem~\ref{thm:initial}.
-
-We now provide a high-level explanation for the technical difficulty when applying Theorem~\ref{thm:initial} to Bernoulli observations.
-The derivation of Theorem~\ref{thm:initial} relies on the upper bound of the estimation error for the mean tensor in Lemma~\ref{lem:two-step_esterror}; i.e., with high probability
-\begin{equation}\label{eq:prop1}
-    \onormSize{}{\hat \tX - \tX}_F^2 \lesssim p^{K/2},
-\end{equation}
-where $\tX = \bbE\tY$ and $\hat \tX$ is defined in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}.
-Unfortunately, the inequality~\eqref{eq:prop1} holds only for i.i.d.\ sub-Gaussian observations while Bernoulli observations are generally not identically distributed.
-
-One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square} of Bernoulli tensor $\tY$.
-Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{\floor{p^{K/2}} \times \ceil{p^{K/2}} }$ denote the nearly square unfolded Bernoulli tensor.
-Define a new estimate
-\begin{equation}\label{eq:matrixsvd}
-    \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX) -  \mat_{sq}(\tY)}_F^2.
-\end{equation}
-The optimization~\eqref{eq:matrixsvd} is simply a matrix SVD problem. Following Lemma 7 in \cite{gao2018community}, with high probability, the new estimate satisfies
-\begin{equation}
-    \onormSize{}{\hat \tX' - \tX}_F^2 \lesssim p^{\ceil{K/2}}.
-\end{equation} 
-Replacing the estimate $\hat \tX$ by $\hat \tX'$ in Theorem~\ref{thm:initial}, the high probability upper bound for Bernoulli initialization is
-\begin{equation}\label{eq:ini_b}
-    \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}. 
-\end{equation}
-The Bernoulli bound~\eqref{eq:ini_b} is relatively looser than Gaussian bound~\eqref{eq:ini}, especially when $K$ is small.
-Nevertheless, our bound~\eqref{eq:ini_b} is already tighter than the pervious work~\citep{ke2019community}.
-The investigation of the gap between upper bound $p^{- \floor{K/2}}$ and the lower bound $p^{- K/2}$ for Bernoulli tensors will be left as future work.
-
-{\bf Extension for general dTBMs.} Our two-stage algorithm is able to be extended for the general (asymmetric) dTBMs.
-Specifically, in the Sub-algorithm~\hyperref[alg:main]{1}, we make the following changes: (1)
-Replace the matrcization $\mat(\tY)$ by $\mat_k(\tY)$; (2) Repeat the Steps 1-5 with mode-specified cluster number $r_k$; (3) Obtain the collection initialization $ \{z^{(0)}_k\}_{k=1}^K$.
-In the Sub-algorithm~\hyperref[alg:main]{2}, we make the following changes: (1) Take the collection $\{z^{(0)}_k\}_{k=1}^K$ as input, and update the block tensor $\tS^{(t)}$ with the collection $\{ z_k^{(t)}\}_{k=1}^K$, $
-        \tS^{(t)}(a_1,\ldots, a_K) =  \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}_k(i_k) = a_k, k \in [K]\}$;
-(2) Calculate reduced tensor $\tY^{\text{d}}_k$ for each mode via
-\begin{align}
-    &\tY^{\text{d}}_k(a_1,\ldots,a_{k-1}, i ,a_{k+1},\ldots,a_K) 
-    = \text{Ave}\{\tY(i_1,\ldots,i_{k-1},i,i_{k+1},\ldots,i_K): z^{(t)}(i_j) = a_j, j \neq k \}; 
-\end{align}
-
-(3) Repeat Step 8-10 with $\mat_k(\cdot), \tY_k^{\text{d}}$ for each $k \in [K]$ and obtain the collection $\{z^{(T)}_k\}_{k=1}^K$.
-
-Correspondingly, Theorems~\ref{thm:initial} and \ref{thm:refinement} still hold with $\ell(z^{(0)}_k, z_k)$ and $\ell(z^{(t+1)}_k, z_k)$  for all $k \in [K]$.
-The detailed model extension for general asymmetric dTBMS can be found in Appendix.
-
-{\bf Computational complexity.} Our two-stage algorithm has a computational cost polynomial in tensor dimension $p$.
-Specifically, the complexity of Sub-algorithm~\hyperref[alg:main]{1} is $\tO(K p^{K+1} + Krp^K)$, where the first term is contributed by the double projection and the calculation of $\hat \tX$, and the second term comes from normalization and the $k$-means.
-The cost of each update in Sub-algorithm~\hyperref[alg:main]{2} is $\tO(p^K + pr^K)$, where $p^K$ comes from the calculation of $\tS^{(t)}$ and $\tY^{\text{d}}$, and $pr^K$ comes from the normalization of $\tY^{\text{d}}$, the calculation of $\tS^{(t)}$, and the cluster assignment update in Step 10.
-
-{\bf Hyper-parameter selection.} In our theoretical analysis, we have assumed the true cluster number $r$ is given to our algorithm.
-In practice, the cluster number $r$ is often unknown, and we now propose a method to choose $r$ from data.
-We impose the Bayesian information criterion (BIC) and choose the cluster number that minimizes BIC; i.e., under the symmetric Gaussian dTBM~\eqref{eq:model_tensor},
-\begin{align}\label{eq:BIC}
- \hat r &= \argmin_{r \in \mathbb{Z}_+ } \of{p^K \log(\onormSize{}{\hat \tX - \tY}_F^2)  + p_e(r) K \log p},\\
- \text{where}  \quad \hat \tX &= \hat \tS(r) \times_1 \hat \mTheta(r) \hat \mM(r) \times_2 \cdots \times_K \hat \mTheta(r) \hat \mM(r),
-\end{align}
-where the triplet $(\hat z(r), \hat \mS(r),\hat \mtheta(r))$ are estimated parameters with cluster number $r$, and $p_e (r)= r^K + p(\log r + 1) - r$ is the effective number of parameters.
-Note that we have added the argument $(r)$ to related quantities as functions of $r$.
-In particular, the estimate $\hat \mtheta(r)$ in~\eqref{eq:BIC} is obtained by first calculating the reduced tensor $\hat \tY^{\text{d}}$ with $\hat z(r)$, and then normalizing the row norms $\onormSize{}{\hat \mY_{i:}^{\text{d}}}$ to 1 in each cluster; i.e.,
-\begin{equation}
- \hat \mtheta(r)=(\hat \theta(1,r),\ldots,\hat \theta(p,r))^T,\quad\text{where}\quad   \hat \theta(i,r) = \frac{\onormSize{}{\hat \mY^{\text{d}}(r)_{i:}}}{\sum_{j: \hat z(j,r) =  \hat z(i,r)} \onormSize{}{\hat \mY^{\text{d}}(r)_{j:}}},
-\end{equation}
-where $\hat \mY^{\text{d}}(r) = \mat(\hat \tY^{\text{d}}(r))$, $\hat \tY^{\text{d}}(r)(i,a_2,\ldots,a_K) =  \text{Ave}\{\tY(i,i_2,\ldots,i_K): \hat z(i_k, r) = a_k, k \neq 1 \}$, and $\hat z(i,r)$ denotes the community label for the $i$-th node with given cluster number $r$.
-We evaluate the performance of the BIC criterion in Section~\ref{subsec:num_theory}.
-
-\section{Numerical studies}\label{sec:simulation}
-
- We evaluate the performance of the weighted higher-order initialization and angle-based iteration in this section.
-We report average errors and standard deviations across 30 replications in each experiment.
-Clustering accuracy is assessed by clustering error rate (CER, i.e., one minus rand index).
-Note that CER between $(\hat z, z)$ is equivalent to misclustering error $\ell(\hat z, z)$ up to constant multiplications~\citep{meilua2012local}, and a lower CER indicates a better performance.
-
-We generate order-3 tensors with \emph{assortative}~\citep{gao2018community} core tensors to control SNR; i.e., we set $\tS_{aaa} = s_1$ for $a \in [r]$ and others be $s_2$, where $s_1 > s_2 > 0$.
-Let $\alpha = s_1/s_2$.
-We set $\alpha$ close to 1 such that $1-\alpha=o(p)$.
-In particular, we have $\alpha = 1 + \Omega(p^{\gamma/2})$ with $\gamma<0$ by Assumption~\ref{assmp:min_gap} and definition~\eqref{eq:gamma}.
-Hence, we easily adjust SNR via varying $\alpha$.
-Note that the assortative setting is proposed for simulations, and our algorithm is applicable for general tensors in practice.
-The cluster assignment $z$ is randomly generated with equal probability across  $r$ clusters for each mode.
-Without further explanation, we generate degree heterogeneity $\mtheta$ from absolute normal distribution by $\theta(i) = |X_i| + 1 - 1/\sqrt{2\pi}$ with $|X_i| \stackrel{\text{i.i.d.}}\sim N(0,1), i \in [p]$ and normalize $\mtheta$ to satisfy \eqref{eq:family}.
-Also, we set $\sigma^2 = 1$ for Gaussian data without further specification.
-
-\subsection{Verification of theoretical results}\label{subsec:num_theory}
-
-The first experiment verifies  statistical-computational gap described in Section~\ref{sec:limits}.
-Consider the Gaussian model with $p = \{80, 100\}$, $r = 5$.
-We vary $\gamma $ in $ [-1.2, -0.4]$ and $[-2.1, -1.4]$ for matrix ($K=2$) and tensor $(K = 3)$ clustering, respectively.
-Note that finding MLE under dTBM is computationally intractable.
-We approximate MLE using an oracle estimator, i.e., the output of Sub-algorithm~\hyperref[alg:main]{2} initialized from true assignment.
-Figure~\ref{fig:phase}a shows that both our algorithm and oracle estimator start to decrease around the critical value $\gamma_{\text{stat}}  = \gamma_{\text{comp}}  = -1$ in matrix case.
-In contrast, Figure~\ref{fig:phase}b shows a significant gap in the phase transitions between the algorithm estimator and oracle estimator in tensor case.
-The oracle error rapidly decreases to 0 when $\gamma_{\text{stat}} = -2$, whereas the algorithm estimator tends to achieve exact clustering when $\gamma_{\text{comp}} = -1.5$.
-Figure~\ref{fig:phase} confirms the existence of the statistical-computational gap in our Theorems~\ref{thm:stats} and~\ref{thm:comp}.
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=.8\columnwidth]{phase_anno3.pdf}
-    \caption{SNR phase transitions for clustering in dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.
-    }
-    \label{fig:phase}
-\end{figure}
-
-The second experiment verifies the performance guarantees of two algorithms: (i) weighted higher-order initialization; (ii) combined algorithm of weighted higher-order initialization and angle-based iteration.
-We consider both the Gaussian and Bernoulli models with $p = \{80, 100\}$, $r = 5$, $\gamma \in [-2.1, -1.4]$.
-Figure~\ref{fig:ini_re} shows the substantial improvement of combined algorithm over initialization, especially under weak and intermediate signals.
-This phenomenon agrees with the error rates in Theorems~\ref{thm:initial} and \ref{thm:refinement}
-and confirms the necessity of the local iterations.
-
-\begin{figure}[htp!]
-    \centering
-     \includegraphics[width=.8\columnwidth]{ini_re_anno3.pdf}
-    \caption{CER versus signal exponent $(\gamma)$ for initialization only and for combined algorithm.
-We set $p = \{80, 100\}, r = 5, \gamma \in [-2.1, -1.4]$ under (a) Gaussian models and (b) Bernoulli models.
-}
-    \label{fig:ini_re}
-
-\end{figure}
-
-\begin{table}[hbt]
-\centering
-    \begin{tabularx}{\columnwidth}{c *{8}{Y}}
-    \toprule
-    Settings & \multicolumn{2}{c}{\small  $p = 50, \sigma^2 = 0.25$} & \multicolumn{2}{c}{\small $p = 50, \sigma^2 = 1$} & \multicolumn{2}{c}{\small  $p = 80, \sigma^2 = 0.25$} & \multicolumn{2}{c}{\small  $p = 80, \sigma^2 = 1$}\\
-    \cmidrule(lr){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9}
-         True cluster number $r$ &  \small 2 & \small 4 &  \small2  &  \small 4 &  \small2  & \small 4 &  \small 2 & \small 4  \\
-         \midrule
-         Estimated cluster number $\hat r$ &   \small 2(0)  &  \small 3.9(0.2)&  \small 2(0)    &   \small 3.1(0.5) &  \small 2(0)    &  \small 4(0)   &  \small 2(0)    &  \small 3.9(0.3)   \\
-     \bottomrule
-    \end{tabularx}
-    \caption{Estimated cluster number given by BIC criterion under the low noise level $(\sigma^2 = 0.25)$ and high noise level $(\sigma^2 = 0.5)$ settings.
-Numbers in parentheses are standard deviations of $\hat r$ over 30 replications.}
-    \label{tab:select}
-\end{table}
-
-The third experiment evaluates the empirical performance of the BIC criterion to select unknown cluster number.
-We generate the data from an order-3 Gaussian model with $p = \{50,80\}$, $r = \{2,4\}$, and noise level $\sigma^2 \in \{ 0.25,1\}$.
-Table~\ref{tab:select} shows that our BIC criterion well chooses the true $r$ under most settings.
-Note that the BIC slightly underestimates the true cluster number $(r = 4)$ with smaller dimension and higher noise $(p = 50, \sigma=1)$, and the accuracy immediately increases with larger dimension $p = 80$.
-The improvement follows from the fact that a larger dimension $p$ indicates a larger sample size in the tensor block model.
-Therefore, we conclude that BIC criterion is a reasonable way to tune the cluster number.
-
-\subsection{Comparison with other methods}\label{subsec:comp}
-
-We compare our algorithm with following higher-order clustering methods:
-\begin{itemize}[wide,topsep=-3pt,itemsep=0pt,parsep=1pt]
-    \item \textbf{\small HOSVD}: HOSVD on data tensor and $k$-means on the rows of the factor matrix;
-    \item \textbf{\small HOSVD+}: HOSVD on data tensor and $k$-means on the $\ell_2$-normalized rows of the factor matrix;
-    \item \textbf{\small HLloyd}~\citep{han2020exact}: High-order clustering algorithm developed for non-degree tensor block models;
-    \item \textbf{\small SCORE}~\citep{ke2019community}: Tensor-SCORE for clustering developed for binary tensors.
-\end{itemize}
-
-Among the four alternative algorithms, the \textbf{\small SCORE} is the closest method to ours.
-We set the tuning parameters of \textbf{\small SCORE} as in previous literature \citep{ke2019community}.
-The methods \textbf{\small SCORE} and \textbf{\small HOSVD+} are designed for degree models, whereas \textbf{\small HOSVD} and \textbf{\small HLloyd} are designed for non-degree models.
-We conduct two experiments to assess the impacts of (i) signal strength and (ii) degree heterogeneity, based on Gaussian and Bernoulli models with $ p = 100, r = 5$.
-We refer to our algorithm as \textbf{\small dTBM} in the comparison.
-
-We investigate the effects of signal to clustering performance by varying $\gamma \in [-1.5, -1.1]$.
-Figure~\ref{fig:comp_gamma} shows the consistent outperformance of our method \textbf{\small dTBM} among all algorithms.
-The sub-optimality of \textbf{\small SCORE} and \textbf{\small HOSVD+} indicates the necessity of local iterations on the clustering.
-Furthermore,  Figure~\ref{fig:comp_gamma} shows the inadequacy of non-degree algorithms in the presence of mild degree heterogeneity.
-The experiment demonstrates the benefits of addressing heterogeneity in higher-order clustering tasks.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=.8\columnwidth]{comp_gamma_anno3.pdf}
-    \caption{CER versus signal exponent (denoted $\gamma$) for different methods. We set $p = 100, r = 5, \gamma \in [-1.5, -1.1]$ under (a) Gaussian and (b) Bernoulli models.}
-    \label{fig:comp_gamma}
-\end{figure}
-
-The only exception in Figure~\ref{fig:comp_gamma} is the slightly better performance of \textbf{\small HLloyd} over \textbf{\small HOSVD+} under Gaussian model.
-However, we find the advantage of \textbf{\small HLloyd} disappears with higher degree heterogeneity.
-We perform extra simulations to verify the impact of degree effects.
-We use the same setting as in the first experiment in the Section~\ref{subsec:comp}, except that we now generate the degree heterogeneity $\mtheta$ from Pareto distribution prior to normalization.
-The density function of Pareto distribution is $f(x|a,b) = a b^a x^{-(a+1)} \ind\{ x \geq b \}$, where $a$ is called \emph{shape} parameter.
-We vary $a \in \{2,6\}$ and choose $b$ such that $\bbE X = a(a - 1)^{-1}b = 1$ for $X$ following Pareto$(a,b)$.
-Note that a smaller $a$ leads to a larger variance in $\mtheta$ and hence a larger degree heterogeneity.
-We consider the Gaussian model under low $(a = 6)$ and high $(a = 2)$ degree heterogeneity.
-Figure~\ref{fig:comp_gamma_theta} shows that the errors for non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) increase with degree heterogeneity.
-In addition, the advantage of \textbf{\small HLloyd} over \textbf{\small HOSVD+} disappears with higher degree heterogeneity.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=.8\columnwidth]{comp_theta_anno3.pdf}
-    \caption{CER versus shape parameter in degree (denoted $a\in[3,6]$) for different methods. We set $p = 100, r = 5, \gamma = -1.2$ under (a) Gaussian and (b) Bernoulli models.}
-    \label{fig:comp_theta}
-\end{figure}
-
-\begin{figure}[htp!]
-    \centering
-    \includegraphics[width=.8\columnwidth]{comp_gamma_theta3.pdf}
-    \caption{CER comparison versus signal exponent (denoted $\gamma$) under (a) low (shape parameter $a = 6$)  (b) high (shape parameter $a = 2$) degree heterogeneity.
-We set $p = 100, r = 5, \gamma \in [-1.5, -1.1]$ under Gaussian model.}
-    \label{fig:comp_gamma_theta}
-\end{figure}
-
-The last experiment investigates the effects of degree heterogeneity to clustering performance.
-We fix the signal exponent $\gamma = -1.2$ and vary the extent of degree heterogeneity.
-In this experiment, we generate $\mtheta$ from Pareto distribution prior to normalization.
-We vary the shape parameter $a \in [3,6]$ in the Pareto distribution to investigate a range of degree heterogeneities.
-Figure~\ref{fig:comp_theta} demonstrates the stability of degree-corrected algorithms (\textbf{\small dTBM}, \textbf{\small SCORE}, \textbf{\small HOSVD+}) over the entire range of degree heterogeneity under consideration.
-In contrast, non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) show poor performance with large heterogeneity, especially in Bernoulli cases.
-This experiment, again, highlights the benefit of addressing degree heterogeneity in higher-order clustering.
-
-\section{Real data applications}\label{sec:real}
-\subsection{Human brain connectome data analysis}
-
-The Human Connectome Project (HCP) aims to construct the structural and functional neural connections in human brains~\citep{van2013wu}.
-We preprocess the original dataset following \cite{desikan2006automated} and partition the brain into 68 regions.
-The cleaned dataset includes brain networks for 136 individuals.
-Each brain network is represented by a 68-by-68 binary symmetric matrix, where the entry with value 1 indicates the presence of connection between node pairs, while the value 0 indicates the absence.
-We use $\tY \in \{0,1\}^{68 \times 68 \times 136}$ to denote the binary tensor.
-Individual attributes such as gender and sex are recorded.
-
-We apply our generalized algorithm to the HCP data with the numbers of clusters on three modes $r_1 = r_2 = 4$ and $r_3 = 3$.
-The selection of $r_1$ and $r_2$ follows the human brain anatomy and the symmetry in the brain network, and the $r_3$ is specified following previous analysis~\citep{hu2021generalized}.
-Because of the symmetry in the data, the estimated brain node clustering results are the same on the first and second modes.
-Figure~\ref{fig:cluster_brain} shows that brain connection exhibits a strong spatial separation structure.
-Specifically, the first cluster, named \emph{L.Hemis}, involves all the nodes in the left hemisphere.
-The nodes in the right hemisphere are further separated into three clusters led by the middle-part tissues in Temporal and Parietal lobes (\emph{R.Temporal}), the back-part tissues in Occipital lobe (\emph{R.Occipital}), and the front-part tissues in Frontal and Parietal lobes (\emph{R.Supra}).
-This clustering result is reasonable since the left and right hemispheres often play different roles in human brains.
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width = .8\columnwidth]{brain_node_cluster.pdf}
-    \caption{Illustration of brain node clustering results for HCP data with (a) top and (b) side views. }
-    \label{fig:cluster_brain}
-\end{figure}
-
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width = 1\textwidth]{est_S_anno.pdf}
-    \caption{Mode 3 slices of estimated core tensor $\hat \tS$.
-(a) Average estimated slice weighted by the group size; (b)-(d) Group-specified enrichment, i.e., the difference between each slice of $\hat \tS$ and the averaged slice.
-}
-    \label{fig:ests}
-\end{figure*}
-
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width = 1\textwidth]{brain_connection.pdf}
-    \caption{Observed brain connections in the population and each group of individuals.
-(a) Average brain network; (b)-(d) Group-specified brain network enrichments in Groups 1-3.
-Red edges represent the positive enrichment and blue edges represent the negative enrichment.}
-    \label{fig:brain_conn}
-\end{figure*}
-
-Figure~\ref{fig:ests} illustrates the estimated core tensor $\hat \tS$ with estimated clustering, and Figure~\ref{fig:brain_conn} visualizes the average brain connections and the connection enrichment in contrast to average networks in each group.
-In general, we find that the inner-hemisphere connection has stronger connection compared to inter-hemisphere connections (Figure~\ref{fig:ests}a).
-Also, the back and front parts (\emph{R.Occipital}, \emph{R.Supra}) are shown to have more interactions with temporal tissues than inner-cluster connections.
-In addition, the group 1 with 54\% females shows an enrichment on the inter-hemisphere connections (Figure~\ref{fig:ests}b), while group 4 with only 36\% females exhibits a reduction (Figure~\ref{fig:ests}d).
-This result agrees with previous findings in \cite{hu2021generalized}.
-The enrichment on the back-front connection is also recognized in group~3 (Figure~\ref{fig:ests}c).
-The interpretive patterns in our results demonstrate the usefulness of our clustering methods in the human brain connectome data application.
-
-\subsection{Peru Legislation data analysis}
-
-We also apply our method to the legislation networks in the Congress of the Republic of Peru \citep{lee2017time}.
-Because of the frequent political power shifts in the Peruvian Congress during 2006-2011, we choose to focus on the data for the first half of 2006-2007 year.
-The dataset records the co-sponsorship of 116 legislators from top 5 parties and 802 bill proposals.
-We reconstruct legislation network as an order-3 binary tensor $\tY \in \{0,1\}^{116 \times 116 \times 116}$, where $\tY_{ijk} = 1$ if the legislators $(i,j,k)$ have sponsored the same bill, and $\tY_{ijk} = 0$ otherwise.
-The true party affiliations of legislators are provided and serve as the ground truth.
-We apply various higher-order clustering methods to $\tY$ with $r = 5$.
-Table~\ref{tab:peru} shows that our \textbf{\small dTBM} achieves the best performance compared to others.
-The second best method is the two-stage algorithm \textbf{\small HLloyd}, followed by the spectral methods \textbf{\small SCORE} and \textbf{\small HOSVD+}.
-This result is consistent with our simulations under strong signal and moderate degree heterogeneity.
-The comparison suggests that our method \textbf{\small dTBM} is more appealing in real-world applications.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{c |c  c cc c}
-    \hline
-        Method & \textbf{\small dTBM} 
-        &\textbf{\small HOSVD}
-        &\textbf{\small HOSVD+} & \textbf{\small HLloyd} &  \textbf{\small SCORE}\\
-         CER & \textbf{0.116}
-         &  0.22 
-         &0.213 & 0.149 &0.199\\
-         \hline
-    \end{tabular}
-    \caption{Clustering errors (measured by CER) for various methods in the analysis of Peru Legislation dataset.}
-    \label{tab:peru}
-\end{table}
-
-\section{Proof Sketches}\label{sec:mainproof}
-
-In this section, we provide the proof sketches for the main Theorems~\ref{thm:initial}-\ref{thm:refinement}. Detail proofs and extra theoretical results are provided in Appendix.
-
-\subsection{Proof sketch of Theorem~\ref{thm:initial}}
-The proof of Theorem~\ref{thm:initial} is inspired by the proof idea of \citet[Lemma 1]{gao2018community}.
-The extra difficulties are the angle gap characterization and multilinear algebra property in tensors; we address both challenges in our proof.
-Specifically, we control the misclustering error by the estimation error of $\hat \tX$ calculated in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}.
-We prove the following inequality
-\begin{equation}\label{eq:proof_4}
-    \ell(z^{(0)},z) \lesssim \frac{1}{p}\min_{\pi \in \Pi} \sum_{i: z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2 \lesssim \frac{\sigma^2 r^{K-1}}{ \Delta_{\min}^2 p^K} \onormSize{}{\hat \tX - \tX}_F^2 \lesssim \frac{r^K p^{-K/2}}{\text{SNR}},
-\end{equation}
-where $\tX = \bbE \tY$ is the true mean.
-The first inequality in~\eqref{eq:proof_4} holds with the assumption $\min_{i \in [p]} \theta(i) \geq c>0$ in Theorem~\ref{thm:initial}.
-The second inequality relies on an important conclusion that the angle gap of mean tensor $\tX$ is lower bounded by that of core tensor $\tS$, i.e., the minimal angle gap $\Delta_{\min}$ defined in Assumption~\ref{assmp:min_gap}.
-Let $\ma^{s} \coloneqq \ma/\onorm{\ma}$ denote the normalized vector with the convention that $\ma^s = 0$ if $\ma = 0$.
-We want to show that
-\begin{equation}\label{eq:proof_4_gap}
-\min_{z(i) \neq z(j)} \onormSize{}{[\mX_{i:}]^s - [\mX_{j:}]^s} \gtrsim \Delta_{\min},
-\end{equation}
-where $\mX = \mat(\tX)$.
-The most challenging part in the proof of Theorem~\ref{thm:initial} lies in the derivation of inequality~\eqref{eq:proof_4_gap}, in which the proof of \cite{gao2018community} is no longer applicable due to different angle gap assumption in our dTBM.
-We develop the extra padding technique in Lemma~\ref{lem:pad} and balance assumption~\eqref{eq:degree} to derive~\eqref{eq:proof_4_gap}.
-Last, we finish the proof of Theorem~\ref{thm:initial} by showing the third inequality of~\eqref{eq:proof_4} using \citet[Proposition 1]{han2020exact}.
-
-\subsection{Proof sketch of Theorem 5}
-The proof of Theorem~\ref{thm:refinement} is inspired by the proof idea of \citet[Theorem 2]{han2020exact}.
-We develop extra polar-coordinate based techniques with angle gap characterization to address the nuisance degree heterogeneity.
-We introduce an intermediate quantity called misclustering loss
-\begin{equation}
-     L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
-\end{equation}
-where the superscript $\cdot^{s}$ denotes the normalized vector; i.e., $\ma^s:=\ma/\onorm{\ma}$ if $\ma\neq 0$ and $\ma^{s}=0$ if $\ma=0$ for any vector $\ma$.
-We show that $L^{(t)}$ provides an upper bound for the misclassification error of interest via the inequality $\ell^{(t)}\leq {L^{(t)}\over \Delta^2_{\min}}$.
-Therefore, it suffices to control $L^{(t)}$.
-Further, we introduce the oracle estimators for core tensor under the true cluster assignment via
-\begin{equation}
-    \tilde \tS = \tY \times_1 \mW^T \times_2 \cdots \times_K \mW^T, 
-\end{equation}
-where $\mW = \mM \of{ \text{diag}(\mone_{p}^T \mM) }^{-1}$ is the weighted true membership matrix.
-Let $ \mV = \mW^{\otimes (K-1)}$ denote the Kronecker product of $(K-1)$ copies of $\mW$ matrices, and we define the $t$-th iteration quantities $\mW^{(t)}, \mV^{(t)}$ corresponding to $\mM^{(t)}$ (or equivalently $z^{(t)}$).
-To evaluate $L^{(t+1)}$, we prove the bound
-\begin{equation}\label{eq:proof_5_event}
-    \ind \offf{ z^{(t+1)}(i) = b } = \ind \offf{       \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{b:}^{(t)}]^s }^2 \leq \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{z(i):}^{(t)}]^s }^2} \leq A_{ib} + B_{ib},
-\end{equation}
-where $\mY = \mat(\tY)$, $ \mS = \mat(\tS)$, $\mS^{(t)} = \mat(\tS^{(t)})$ and
-\begin{align}
-        A_{ib} &= \ind \offf{\ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s } \lesssim -  \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 },\\
-        B_{ib} &= \ind \offf{\onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 \lesssim F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }.
-\end{align}
-The terms $F_{ib}^{(t)}, G_{ib}^{(t)}, H_{ib}^{(t)}$ are controlled by $z^{(t)}, \tS^{(t)}$; see the detailed definitions in \eqref{eq:f}, \eqref{eq:g}, \eqref{eq:h}.
-Note that the event $A_{ib}$ only involves the oracle estimator independent of $t$, while all the terms related to the $t$-th iteration are in $B_{ib}$.
-Thus, the inequality~\eqref{eq:proof_5_event} decomposes the misclustering loss in the $(t+1)$-th iteration into the oracle loss and the loss in $t$-th iteration.
-This decomposition leads to the separation of statistical error and computational error in the final upper bound of Theorem~\ref{thm:refinement}.
-
-Specifically, we prove the contraction inequality
-\begin{equation}\label{eq:proof_5_ineq}
-    L^{(t+1)} \leq M \xi + \rho L^{(t)}, \quad \xi = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  A_{ib} \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
-\end{equation}
-where $M$ is a positive constant, $\rho \in (0,1)$ is the contraction parameter, and we call $\xi$ the oracle loss.
-Controlling the probability of event $B_{ib}$ and obtaining the $\rho L^{(t)}$ term in the right hand side of~\eqref{eq:proof_5_ineq} are the most challenging parts in the proof of Theorem~\ref{thm:refinement}.
-Note that the true and estimated core tensors are involved via their normalized rows such as $\mS_{a:}^s, \tilde \mS_{a:}^s, [\mS^{(t)}_{a:}]^s$.
-The Cartesian coordinate based analysis in \cite{han2020exact} is no longer applicable in our case.
-Instead, we use the polar-coordinate based analysis and the geometry property of trigonometric functions to derive the high probability upper bounds for $F_{ib}^{(t)}, G_{ib}^{(t)}, H_{ib}^{(t)}$.
-
-Further, by sub-Gaussian concentration, we prove the high probability upper bound for oracle loss
-\begin{equation}\label{eq:proof_5_xi}
-    \xi  \lesssim \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}.
-\end{equation}
-Combining the decomposition~\eqref{eq:proof_5_ineq} and the oracle bound~\eqref{eq:proof_5_xi}, we finish the proof of Theorem~\ref{thm:refinement}.
-
-\section*{Acknowledgments}
-This research is supported in part by NSF grants DMS-1915978, DMS-2023239, EF-2133740, and funding from the Wisconsin Alumni Research foundation.
-We thank Zheng Tracy Ke, Rungang Han, Yuetian Luo for helpful discussions and for sharing software packages.
-
-\bibliographystyle{apalike}
-\bibliography{tensor_wang}
-
-\newpage
-\appendix
-\section*{Appendices}
-
-We provide the proofs for all the theorems in our main paper. In each sub-section, we first show the proof of main theorem and then collect the useful lemmas in the end.
-
-\section{Notation}
-Before the proofs, we first introduce the notation used throughout the appendix and the generalized dTBM without symmetric assumptions.
-The parameter space and minimal gap assumption are also extended for the generalized dTBM.
-
-{\bf Preliminaries.}
-\begin{enumerate}[wide]
-    \item For mode $ k \in [K]$, denote the mode-$k$ tensor matricizations by
-    \begin{equation}
-        \mY_k = \mat_k \of{ \tY }, \quad \mS_k = \mat_k \of{\tS}, \quad \mE_k = \mat_k \of{ \tE}, \quad \mX_k = \mat_k \of{\tX}.
-    \end{equation}
-    \item For a vector $\ma$, let $\ma^{s} \coloneqq \ma/\onorm{\ma}$ denote the normalized vector. We make the convention that $\ma^s = {\bf 0}$ if $\ma = {\bf 0}$. 
-    \item For a matrix $\mA \in \bbR^{n \times m} $, let $\mA^{\otimes K}:=\mA\otimes \cdots \otimes \mA\in \bbR^{n^K \times m^K}$ denote the Kronecker product of $K$ copies of matrices $\mA $.
-    \item For a matrix $\mA$, let $\onormSize{}{\mA}_\sigma$ denote the spectral norm of matrix $\mA$, which is equal to the maximal singular value of $\mA$; let $\lambda_k(\mA)$ denote the $k$-th largest singular value of $\mA$; let $\onormSize{}{\mA}_F$ denote the Frobenius norm of matrix $\mA$.
-    \item For two sequence $a$ and $b$, let $a \asymp b$ if there exist two positive constants $c, C$ such that $cb \leq a\leq Cb$. 
-\end{enumerate}
-
-{\bf Model extension to generalized dTBM.}
-
- The general order-$K$ $(p_1, \ldots, p_K)$-dimensional dTBM model with $r_k$ communities and degree heterogeneity $\mtheta_k = \entry{\theta_k(i)} \in \bbR_+^{p_k}$ is represented by
-\begin{equation}\label{eq:general_dtbm}
-    \tY = \tX+ \tE,\quad \text{where}\quad \tX=\tS \times_1 \mTheta_1 \mM_1 \times_2 \cdots \times_K \mTheta_K \mM_K,
-\end{equation}
-where $\tY \in \bbR^{p_1 \times \cdots \times p_K}$ is the data tensor, $\tX\in \mathbb{R}^{p_1\times \cdots \times p_K}$ is the mean tensor, $\tS \in \bbR^{r_1 \times \cdots \times r_K}$ is the core tensor, $\tE \in \bbR^{p_1 \times \cdots \times p_K}$ is the noise tensor consisting of independent zero-mean sub-Gaussian entries with variance bounded by $\sigma^2$, $\mTheta_k = \text{diag}(\mtheta_k)$, and $\mM_k\in \{0,1\}^{p_k \times r_k}$ is the membership matrix corresponding to the assignment $z_k: [p_k] \mapsto [r_k]$, for all $k \in [K]$.
-
-For ease of notation, we use $\{z_k\}$ to denote the collection $\{z_k\}_{k=1}^K$, and $\{\mtheta_k\}$ to denote the collection $\{\mtheta_k\}_{k=1}^K$.
-Correspondingly, we consider the parameter space for the triplet $\of{\{z_k\}, \tS, \{\mtheta_k\}}$,
-
-\begin{align}
- \tP(\{r_k\}) = \Bigg\{ &\of{\{z_k\}, \tS, \{\mtheta_k\}}:\\
-   & \mtheta_k \in\mathbb{R}^p_{+}, {c_1 p_k\over r_k} |z_k^{-1}(a)| \leq {c_2 p_k\over r_k}, c_3 \leq  \onorm{\mS_{k,a:}} \leq c_4 ,\onormSize{}{\mtheta_{k,z_k^{-1}(a)}}_1=|z_k^{-1}(a)|, a \in [r_k], k\in[K]\Bigg\}.
-\label{eq:general_family}
-\end{align}
-
-We call the degree heterogeneity $\{\mtheta_k\}$ is balanced if for all $k \in [K]$,
-\begin{equation}\label{eq:general_balanced}
-    {\min_{a\in[r]} \onormSize{}{\mtheta_{k, z_k^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{k,z_k^{-1}(a)}}}.
-\end{equation}
-
-We also consider the generalized Assumption~\ref{assmp:min_gap} on angle gap.
-\begin{assumption}[Generalized angle gap]\label{assmp:general_minimal_gap} Recall $\mS_k=\mat_k(\tS)$.
-We assume the minimal gap between normalized rows of $\mS_k$ is bounded away from zero for all $k\in[K]$; i.e.,
-\begin{equation}
-     \Delta_{\min} \coloneqq \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onorm{ \mS_{k,a:}^s - \mS_{k, b:}^s } > 0.
-\end{equation}
-\end{assumption}
-Similarly, let $\text{SNR} = \Delta_{\min}^2/\sigma^2$ with the generalized minimal gap $\Delta_{\min}^2$ defined in Assumption~\ref{assmp:general_minimal_gap}. We define the regime
-\begin{equation}
-    \tP(\gamma) = \tP (\{r_k\}) \cap\{\tS \text{ satisfies $\text{SNR} = p^{\gamma}$ and $p_k \asymp p, \text{for all }k \in [K]$} \}.
-\end{equation}
-
-\section*{Proof of Theorem~\ref{thm:unique}}
-
-\begin{proof}[Proof of Theorem~\ref{thm:unique}]
-
-To study the identifiability, we consider the noiseless model with $\tE = 0$. Assume there exist two parameterizations satisfying
-\begin{equation}\label{eq:another}
-    \tX=\tS\times_1\Theta_1 \mM_1 \times_2 \cdots \times_K \Theta_K \mM'_K=\tS'\times_1\Theta'_1 \mM'_1 \times_2 \cdots \times_K \Theta'_K \mM'_K,
-\end{equation}
-where $\of{ \{z_k\}, \tS, \{ \mtheta_k \} } \in \tP(\{r_k\})$ and $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} } \in \tP(\{r_k'\})$ are two sets of parameters.
-We prove the sufficient and necessary conditions separately.
-
-\begin{enumerate}[wide]
-    \item[$(\Leftarrow)$] For the necessity, it suffices to construct two distinct parameters up to cluster label permutation, if the model~\eqref{eq:general_dtbm} violates Assumption~\ref{assmp:general_minimal_gap}.
-Without loss of generality, we assume $\onorm{ \mS_{1,1:}^s - \mS_{1,2:}^s } = 0$.
-
- If $\mS_{1,1:}$ is a zero vector, construct $\mtheta_1'$ such that $\mtheta'_{1,z_1^{-1}(1)} \neq \mtheta_{1,z_1^{-1}(1)}$.
-Let $\{z'_k\} = \{z_k\}$, $\tS' = \tS$, and $\mtheta'_k = \mtheta_k$ for all $k = 2, \ldots, K$.
-Then the triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ is distinct from $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation.
-Similar conclusion holds when $\mS_{1,2:}$ is a zero vector.
-
-If neither $\mS_{1,1:}$ nor $\mS_{1,2:}$ is a zero vector, there exists a positive constant $c$ such that $\mS_{1,1:} = c \mS_{1,2:}$.
-Thus, there exists a core tensor $\tS_0 \in \bbR^{r_1 -1 \times \cdots \times r_K}$ such that
-\begin{equation}
-    \tS = \tS_0 \times_1 \mC \mR, \quad \text{where} \quad \mC = \text{diag}(1, c, 1,...,1) \in \bbR^{r_1 \times r_1}, \quad \mR = \begin{pmatrix}
-    1& 0\\
-    1&0 \\
-    0 & \mone_{r_1-2}
-    \end{pmatrix} \in \bbR^{r_1 \times (r_1 -1)}.
-\end{equation}
-Let $\mD = \text{diag}(1+c, 1,...,1) \in \bbR^{r_1 -1 \times r_1 -1}$. Consider the parameterization
-\begin{equation}
-    \mM'_1 = \mM_1 \mR, \quad \tS' = \tS_0 \times_1 \mD, \quad \theta'_{1}(i) = \begin{cases}
-    \frac{1}{1+c} \theta_{1}(i) & i \in z_{1}^{-1}(1),\\
-     \frac{c}{1+c} \theta_{1}(i) & i \in z_{1}^{-1}(2),\\
-     \theta_{1}(i) & \text{ otherwise},
-    \end{cases}
-\end{equation}
-and $\mM'_k = \mM_k, \mtheta'_k = \mtheta_k$ for all $k = 2, \ldots, K$. Then we have constructed a
-triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ that is distinct from $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation.
-
-\item[$(\Rightarrow)$] For the sufficiency, it suffices to show that all possible triplets $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ are identical to $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation if the model~\eqref{eq:general_dtbm} satisfies Assumption~\eqref{assmp:general_minimal_gap}.
-We show the uniqueness of the three parameters, $\{\mM_k\}, \{\tS\}, \{\mtheta_k\}$ separately.
-
-First, we show the uniqueness of $\mM_k$ for all $k \in [K]$.
-Without loss of generality, we consider $k=1$ and show the first mode membership matrix; i.e., $\mM'_1 = \mM_1\mP_1$ where $\mP_1$ is a permutation matrix.
-The conclusion for $k\geq 2$ can be showed similarly and thus omitted.
-
-Consider an arbitrary node pair $(i,j)$.
-If $z_1(i) = z_1(j)$, then we have $\onormSize{}{\mX_{1, z_1(i):}^s - \mX_{1, z_1(j):}^s } = 0$ and thus $\onormSize{}{ (\mS')_{1, z'_1(i):}^{s} - (\mS')_{1, z'_1(j):}^{s} } = 0$ by Lemma~\ref{lem:angle}.
-Then, by Assumption~\eqref{assmp:general_minimal_gap}, we have $z'_1(i) = z'_1(j)$.
-Conversely, if $z_1(i) \neq z_1(j)$, then we have $ \onorm{\mX_{1,i:}^s - \mX_{1,j:}^s} \neq 0$ and thus $\onorm{ (\mS')_{1, z'_1(i):}^{s} - (\mS')_{1, z'_1(j):}^{s} } \neq 0$ by Lemma~\ref{lem:angle}.
-Hence, we have $z'_1(i) \neq z'_1(j)$.
-Therefore, we have proven that $z'_1$ is identical $z_i$ up to label permutation.
-
-Next, we show the uniqueness of $\mtheta_k$ for all $k \in [K]$ provided that $z_k = z_k'$. Similarly, consider $k=1$ only, and omit the procedure for $k\geq 2$.
-
-Consider an arbitrary $j \in [p_1]$ such that $z_1(j) = a$. Then for all the nodes $i \in  z_1^{-1}(a)$ in the same cluster of $j$, we have 
-\begin{equation}
-    \frac{\mX_{1,z_1(i):}}{\mX_{1,z_1(j):}} = \frac{\mX'_{1,z_1(i):}}{\mX'_{1,z_1(j):}}, \text{ which implies }  \frac{\theta_1(j)}{\theta_1(i)} = \frac{\theta'_1(j)}{\theta'_1(i)}.\label{eq:theta_uniq}
-\end{equation}
-Let $\theta'_1(j) = c\theta_1(j)$ for some positive constant $c$.
-By equation~\eqref{eq:theta_uniq}, we have $\theta'_1(i) = c \theta_1(i)$ for all $ i \in  z_1^{-1}(a)$.
-By the constraint $(\{z_k\}, \tS', \{\mtheta'_k\}) \in \tP(\{r_k\})$, we have
-\begin{equation}
-    \sum_{j \in z_1^{-1}(a)} \theta'_1(j) = c \sum_{j \in z_1^{-1}(a)} \theta_1(j) = 1,
-\end{equation}
-which implies $c = 1$. Hence, we have proven $\mtheta_1 = \mtheta'_1$ provided that $z_1 = z'_1$.
-
-Last, we show the uniqueness of $\tS$; i.e., $\tS'=\tS\times_1 \mP^{-1}_1\times_2\cdots \times_K \mP^{-1}_K$, where $\mP_k$'s are permutation matrices for all $k\in[K]$.
-Provided $z'_k = z_k, \mtheta'_k = \mtheta_k$, we have $\mM'_k = \mM_k \mP_k$ and $\mTheta'_k = \mTheta_k$ for all $k \in [K]$.
-
-Let $\mD_k = \off{ (\mTheta'_k \mM'_k)^T (\mTheta'_k \mM'_k) }^{-1} (\mTheta'_k \mM'_k)^T, k \in [K]$. By the parameterization~\eqref{eq:another}, we have 
-\begin{align}
-    \tS' &= \tX \times_1 \mD_1 \times_2 \cdots \times_K \mD_K \\
-    &= \tS \times_1 \mD_1 \mTheta_1 \mM_1 \times_1 \cdots \times_K \mD_K \mTheta_K \mM_K \\
-    &= \tS \times_1 \mP^{-1}_1 \times_2 \cdots \times_K \mP^{-1}_K.
-\end{align}
-
-\end{enumerate}
-
-Therefore, we finish the proof of Theorem~\ref{thm:unique}.
-\end{proof}
-
-{\bf Useful Lemma for the Proof of Theorem~\ref{thm:unique}}
-
-\begin{lem}[Motivation of angle-based clustering]\label{lem:angle} Consider the signal tensor $\tX$ in the generalized dTBM~\eqref{eq:general_dtbm} with $(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\{r_k\})$ and $r_k \geq 2$.
-Then, for any $k \in [K]$ and index pair $(i,j)\in[p_k]^2$, we have
-\begin{equation}
-    \onorm{ \mS_{k,z_k(i):}^s -  \mS_{k,z_k(j):}^s } = 0 \quad \text{if and only if} \quad \onorm{  \mX_{k, z_k(i):}^s -  \mX_{k,z_k(j):}^s } = 0.
-\end{equation}
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:angle}] Without loss of generality, we prove $k = 1$ only and drop the subscript $k$ in $\mX_k, \mS_k$ for notational convenience. 
-By tensor matricization, we have
-\begin{equation}
-    \mX_{j:} = \theta_1(j) \mS_{z_1(j):} \off{\mTheta_2 \mM_2 \otimes \cdots \otimes \mTheta_K \mM_K}^T.
-\end{equation}     
-Let $\tilde \mM = \mTheta_2 \mM_2 \otimes \cdots \otimes \mTheta_K \mM_K$. Notice that for two vectors $\ma, \mb$ and two positive constants $c_1, c_2 >0$, we have
-\begin{equation}
-\onorm{\ma^s - \mb^s} = \onorm{(c_1 \ma)^s - (c_2\mb)^s}.
-\end{equation}
-Thus it suffices to show the following statement holds for any index pair $(i,j)\in[p_1]^2$,
-\begin{equation}
-\onorm{ \mS_{z_1(i):}^s - \mS_{z_1(j):}^s} = 0 \quad \text{if and only if} \quad \onorm{ \off{\mS_{z_1(i):} \tilde \mM^T }^s - \off{\mS_{z_1(j):}\tilde \mM^T}^s} = 0.
-\end{equation}
-\begin{enumerate}[wide]
-    \item[$(\Leftarrow)$] Suppose $\onorm{ \off{\mS_{z_1(i):} \tilde \mM^T }^s - \off{\mS_{z_1(j):}\tilde \mM^T}^s} = 0$.
-There exists a positive constant $c$ such that $\mS_{z_1(i):} \tilde \mM^T= c \mS_{z_1(j):} \tilde \mM^T$.
-Note that
-\begin{equation}
-    \mS_{z_1(i):} = \mS_{z_1(i):} \tilde \mM^T \off{ \tilde \mM \of{ \tilde \mM^T  \tilde \mM}^{-1}},
-\end{equation}
-where $ \tilde \mM^T  \tilde \mM$ is an invertiable diagonal matrix with positive diagonal elements.
-Thus, we have $ \mS_{z_1(i):} = c  \mS_{z_1(j):}$, which implies $ \onorm{  \mS_{z_1(i):}^s -  \mS_{z_1(j):}^s } = 0 $.
-
-\item[$(\Rightarrow)$] Suppose $ \onorm{ \mS_{z_1(i):}^s - \mS_{z_1(j):}^s } = 0 $.
-There exists a positive constant $c$ such that $\mS_{z_1(i):} = c \mS_{z_1(j):}$, and thus $\mS_{z_1(i):} \tilde \mM^T = c \mS_{z_1(j):} \tilde \mM^T$, which implies $\onorm{\left[\mS_{z_1(i):} \tilde \mM^T\right]^s- \left[\mS_{z_1(j):} \tilde \mM^T\right]^s}=0$.
-\end{enumerate}
-Therefore, we finish the proof of Lemma~\ref{lem:angle}.
-\end{proof}
-
-\section{Proof of Theorem~\ref{thm:stats}}
-
-\begin{proof}[Proof of Theorem~\ref{thm:stats}]
-We will prove a more general conclusion than the main paper by allowing growing $r_k$'s.
-Consider the generalized dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$.
-Specifically, we will show that, under
-the assumptions $K\geq 1$, $r\lesssim p^{1/3}$ and SNR condition
-\[
-{\Delta^2_{\min}\over \sigma^2}\lesssim {r^{K-1}\over p^{K-1}},\quad \text{or equivalently}, \quad \gamma\leq -(K-1)(1+\log_p r),
-\]
- the desired conclusion in Theorem~\ref{thm:stats} holds; i.e, for all $k \in [K]$, every estimator $\hat z_{k,\text{stat}}$ obeys
-\begin{equation}\label{eq:max}
-    \sup_{(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{k,\text{stat}}, z_k) \right]\geq 1.
-\end{equation}
-
-Since the inequality~\eqref{eq:max} is a minimax lower bound, it suffices to show the inequality holds for a particular $(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)$.
-Specifically, we consider the estimation problem based on a particular parameter point $(\{z_k\}, \tS, \{\mtheta_k\})$ with the following three properties:
-\begin{equation}\label{eq:construction}
-    \text{(i)}\  \theta_k(i)=1 \text{ for all }i\in[p]; \quad  \text{(ii)}\ 
-    \Delta_{\min}\lesssim  \left({p\over r}\right)^{-\frac{K-1}{2}}\sigma; \quad  \text{(iii)}\ |z^{-1}_k(a)|={p \over r} \in \mathbb{Z}_+ \text{ for all }a\in[r],
-\end{equation}
-for all $k \in [K]$.
-Furthermore, we define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation.
-Based on \citet[Proof of Theorem 6]{han2020exact}, we consider the minimax rate over the restricted family of $\hat z_k$'s for which the following three conditions are satisfied:
-\begin{equation}
-    \text{(iv)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \quad \text{(v)} \ |T^c_k|\asymp {p\over r}; \quad \text{(vi)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
-\end{equation}
-\normalsize
-for all $k \in [K]$.
-The construction of $T$ is precisely the same as \citet[Proof of Theorem 6]{han2020exact}. 
-Then, following the proof of~\citet[Theorem 2]{gao2018community}, for all $k \in [K]$, we have
-\begin{equation}\label{eq:minimax}
-\inf_{\hat z_k}\sup_{z_k}\bbE \ell(\hat z_k, z_k) \gtrsim {1\over r^3 |T^c_k|}\sum_{i\in T^c_k}\inf_{\hat z_k}\left\{\bbP[\hat z_k(i)=2|z_k(i)=1] + \bbP[\hat z_k(i)=1|z_k(i)=2] \right\},
-\end{equation}
-where $\hat z_k$ and $z_k$ on the left hand side denote the generic clustering functions in $\tP(\gamma)$, $z_k$ on the right hand side denotes a particular parameter satisfying properties (i)-(vi), and the infimum on the right hand side is taken over the restricted family of $\hat z$ satisfying (iv)-(vi).
-Here, the factor $r^3=r\cdot r^2$ in~\eqref{eq:minimax} comes from two sources: $r^2\asymp {r\choose 2}$ comes from the multiple testing burden for all pairwise comparisons among $r$ clusters; and another $r$ comes from the number of elements $|T^c_k|\asymp p/r$ to be clustered.
-
-Next, we need to find the lower bound of the rightmost side in~\eqref{eq:minimax}.
-For simplicity, we show the bound for the mode-1 case $k = 1$ only.
-We drop the subscripts $1$ in $z_1, T_1, \mS_1, \mtheta_1$ and omit the repeated procedures for the cases of $k = 2,\ldots, K$.
-
-We consider the hypothesis test based on model~\eqref{eq:general_dtbm}. First, we reparameterize the model under the construction~\eqref{eq:construction}
-\begin{equation}
-    \mx_a = \off{\mat_1 \of{ \tS\times_2 \mM_2\times_3\cdots \times_K \mM_K  }}_{a:}, \quad \text{for all} \ a \in [r],
-\end{equation}
-where $\mx_a$'s are centroids in $\bbR^{p^{K-1}}$.
-Without loss of generality, we consider the lower bound for the summand in~\eqref{eq:minimax} for $i=1$.
-The analysis for other $i\in T^c$ are similar.
-For notational simplicity, we suppress the subscript $i$ and write $\my, \theta, z$ in place of $\my_1, \theta_1$ and $z(1)$, respectively.
-The equivalent vector problem for assessing the summand in~\eqref{eq:minimax} is
-\begin{equation}\label{eq:z}
-\my=\theta\mx_{z}+\me,
-\end{equation}
-where $\theta\in\bbR_{+}$ and $z\in \{1,2\}$ are unknown parameters, $\mx_1,\mx_2\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.
-Then, we consider the hypothesis testing under the model~\eqref{eq:z}:
-\begin{equation}\label{eq:test}
-H_0\colon z=1, \quad \text{v.s.}\quad H_\alpha\colon z=2.
-\end{equation}
-Note that the profile log-likelihood with respect to $z$ is
-\[
-\tL(z,  \theta(z);\my)\propto - \inf_{\theta>0}\onormSize{}{\my- \theta\mx_{z}}^2 \propto  \cos^2(\my, \mx_{z}) \ind\{ \langle \my,\mx_{z}\rangle  >0 \}, 
-\]
-and the MLE's of $\theta$ and $z$ are
-\begin{equation}\label{eq:MLE}
-    \hat \theta_{\text{MLE}}= \hat \theta(\hat z_{\text{MLE}})={\langle \my,\mx_{\hat z_{\text{MLE}}}\rangle \over \onorm{\mx_{\hat z_{\text{MLE}}}}^2}\vee 0 , \quad 
-\hat z_{\text{MLE}}= \argmax_{a\in\{1,2\}} \offf{\cos(\my,\ \mx_a)\vee 0}. 
-\end{equation}
-
-Then, the decision rule $\hat z_{\text{MLE}} \in\{1,2\}$ based on profile log-likelihood ratio is defined as
-\begin{equation}\label{eq:ratio}
-\hat z_{\text{MLE}}=
-\begin{cases}
-1 & \text{if }\cos(\my,\mx_1)\geq  \cos(\my,\mx_2) \text{ and }\langle \my,\mx_1\rangle > 0, \\
-2 & \text{if }\cos(\my,\mx_1)<\cos(\my,\mx_2) \text{ and }\langle \my,\mx_2\rangle > 0,\\
-\text{1 or 2 with equal probability} & \text{otherwise.}
-\end{cases}
-\end{equation}
-
-The Neyman-Pearson Lemma implies
-\begin{align}\label{eq:lower}
-&\inf_{\hat z}\left\{\bbP[\hat z=2|z=1] + \bbP[\hat z=1|z=2]\right\}= \bbP[\hat z_{\text{MLE}}=1|z=2]+\bbP[\hat z_{\text{MLE}}=2|z=1].
-\end{align}
-By symmetric, it suffices to bound $\bbP[\hat z_{\text{MLE}}=1|z=2]$. Using~\eqref{eq:ratio}, we obtain
-\begin{align}\label{eq:bound}
-\bbP[\hat z_{\text{MLE}}=1|z=2]&= \bbP\left[ \cos(\theta\mx_2+\me,\mx_1)\geq  \cos(\theta\mx_2+\me, \mx_2) \text{ and } \langle \theta\mx_2+\me,\mx_1\rangle > 0  \right] \notag \\
-&\stackrel{(*)}{\geq} \bbP \left[\left\langle \me, {\mx^s_1 -\mx^s_2\over\onorm{\mx^s_1- \mx^s_2}}  \right\rangle  \geq {\theta \over 2}\onormSize{}{\mx_2} \onormSize{}{\mx^s_1-\mx^s_2}\rangle\right]-\notag \\
-&\hspace{1cm} \bbP\left[\langle \me, \mx^s_1\rangle \leq -{\theta \over 2}\onorm{\mx_2}\left(2-\onormSize{}{\mx^s_1-\mx^s_2}^2\right)\right]\notag \\
-&\stackrel{(**)}{=}\Phi\left({\theta \over 2}\onorm{\mx_2}(2-\onormSize{}{\mx^s_1-\mx^s_2}^2) \right)-\Phi\left({\theta\over 2}\onorm{\mx_2}\onormSize{}{\mx^s_1-\mx^s_2} \right),
-\end{align}
-where $\Phi(\cdot)$ denotes the CDF for standard normal distribution.
-Here step $(*)$ is based on the inequality $\bbP(A\cup B)\geq \bbP(A)-\bbP(B^c)$ and the identity $1-\langle \mx^s_1,\mx^s_2 \rangle={1\over 2}\onorm{\mx^s_1-\mx^s_2}^2$; and step $(**)$ is based on isotropic property of i.i.d.\ Gaussian distribution
-\[
-\left\langle\me, {\mx^s_1-\mx^s_2\over \onorm{\mx^s_1-\mx^s_2}} \right\rangle \sim N(0, \sigma^2),\quad \langle \me, \mx^s_1 \rangle \sim N(0,\sigma^2).
-\]
-By construction~\eqref{eq:construction} of $(\{z_k\}, \tS, \{\mtheta_k\})$ with three properties and lower bound $\min_{a\in[r]}\onorm{\mS_{a:}}\geq c_3$ in the definition of $\tP(\gamma)$, we have $\theta^*=1$, $\onorm{\mx_2}\geq \onorm{\mS_{2:}}\min_{a\in[r]}\onorm{\mtheta_{\mz^{-1}(a)}}^{K-1} \gtrsim \left(p \over r\right)^{(K-1)/2}$.
-Also, note that under the construction~\eqref{eq:construction}
-\begin{equation}
-    \cos(\mx_1, \mx_2) = \frac{\and{\mx_1, \mx_2}}{\onorm{\mx_1} \onorm{\mx_2} } = \frac{(p/r)^{K-1} \ang{\mS_{1:}, \mS_{2:}}}{ \sqrt{(p/r)^{K-1} \onorm{\mS_{1:}}^2} \sqrt{(p/r)^{K-1} \onorm{\mS_{2:}}^2} } = \cos(\mS_{1:}, \mS_{2:}),
-\end{equation}
-which implies $\onormSize{}{\mx^s_1-\mx^s_2} = \onormSize{}{\mS_{1:}^s-\mS_{2:}^s} = \Delta_{\min} \leq 1$. Therefore, the equation~\eqref{eq:bound} is lower bounded by
-\begin{equation}\label{eq:normal}
-\bbP[\hat z_{\text{MLE}}=1|z=2] \geq \bbP\left[ \left({p\over r}\right)^{(K-1)/2}\Delta_{\min}\lesssim N(0,1)\lesssim  \left({p\over r}\right)^{(K-1)/2} \right]\geq C>0,
-\end{equation}
- where the existence of strictly positive constant $C$ is based on the SNR assumption~\eqref{eq:construction}. Combining~\eqref{eq:minimax}, \eqref{eq:lower} and~\eqref{eq:normal} yields
-\[
-\inf_{\hat z_1}\sup_{(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\gamma)}\bbE\ell(\hat z_1, z_1)\gtrsim C>0, 
-\]
-and henceforth for all $k \in [K]$
-\[
-\inf_{\hat z_k}\sup_{(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\gamma)}\bbE \left[p\ell(\hat z_k, z_k)\right]\geq 1.
-\]
-
-\end{proof}
-
-\section{Proof of Theorem~\ref{thm:comp}}
-
-\begin{proof}[Proof of Theorem~\ref{thm:comp}]
-The idea of proving computational hardness is to show the computational lower bound for a special class of degree-corrected tensor clustering model with $K\geq 2$.
-We construct the following special class of higher-order degree-corrected tensor clustering  model.
-For a given signal level $\gamma\in\bbR$ and noise variance $\sigma$, define a rank-2 symmetric tensor $\tS\in\bbR^{3\times \cdots\times 3}$ subject to
-\begin{equation}\label{eq:S}
-\tS=\tS(\gamma) = \begin{bmatrix}
-		1\\
-		1\\
-		1
-		\end{bmatrix}^{\otimes K}
-+\sigma p^{-\gamma/2} \begin{bmatrix}
-		1\\
-		-1\\
-		0
-	\end{bmatrix}^{\otimes K}.
-\end{equation}
-Then, we consider the signal tensor family
-\begin{align}
-\tP_{\text{shifted}}(\gamma)=\{\tX\colon & \tX=\tS\times_1\mM_1\times_2 \cdots \times_K\mM_K,\  \text{$\mM_k\in\{0,1\}^{p\times 3}$ is a membership matrix that}
-\\
-& \text{satisfies $|\mM_k(\colon,i)|\asymp p$ for all $i\in[3]$ and $k\in[K]$}\}.
-\end{align}
-We claim that the constructed family satisfies the following two properties:
-\begin{enumerate}[wide,label=(\roman*)]
-    \item For every $\gamma\in \mathbb{R}$, $\tP_{\text{shifted}}(\gamma)\subset \tP(\gamma)$, where $\tP(\gamma)$ is the degree-corrected cluster tensor family~\eqref{eq:gammafamily}.
-    \item For every $\gamma\in \mathbb{R}$, $\{\tX-1\colon \tX\in \tP_{\text{shifted}}(\gamma)\}\subset \tP_{\text{non-degree}}(\gamma)$, where $\tP_{\text{non-degree}}(\gamma)$ denotes the sub-family of rank-one tensor block model constructed in the proof of \citet[Theorem 7]{han2020exact}.
-\end{enumerate}
-The verification of the above two properties is provided in the end of this proof.
-
-Now, following the proof of~\citet[Theorem 7]{han2020exact}, when $\gamma<-K/2$, every polynomial-time algorithm estimator $(\hat \mM_k)_{k\in[K]}$ obeys
-\begin{align}\label{eq:p}
-\liminf_{p\to \infty} \sup_{\tX\in \tP_{\text{non-degree}}(\gamma)}\mathbb{P}(\exists k \in[K],\  \hat \mM_k \neq \mM_k)\geq 1/2,
-\end{align}
-under the HPC Conjecture~\ref{hypo:HPC}.
-The inequality~\eqref{eq:p} implies
-\[
-\liminf_{p\to \infty} \sup_{\tX\in \tP_{\text{non-degree}}(\gamma)}\max_{k\in[K]}\bbE[p\ell(z_k, \hat z_k)]\geq 1.
-\]
-Based on properties (i)-(ii), we conclude that
-\[
-\liminf_{p\to \infty} \sup_{\tX\in \tP(\gamma)}\max_{k\in[K]}\bbE[p\ell(z_k, \hat z_k)]\geq 1.
-\]
-We complete the proof by verifying the properties (i)-(ii).
-For (i), we verify that the angle gap for the core tensor $\tS$ in~\eqref{eq:S} is on the order of $\sigma p^{-\gamma/2}$.
-Specifically, write $\mone=(1,1,1)$ and $\me=(1,-1,0)$.
-We have
-\[
-\Mat(\tS)=
-\begin{bmatrix}
-\text{Vec}(\mone^{\otimes K-1})+\sigma p^{-\gamma/2}\text{Vec}\left(\me^{\otimes (K-1)}\right) \\
-\text{Vec}(\mone^{\otimes K-1})-\sigma p^{-\gamma/2}\text{Vec}\left(\me^{\otimes (K-1)}\right) \\
-\text{Vec}(\mone^{\otimes K-1})
-\end{bmatrix}.
-\]
-Based on the orthogonality $\langle \mone, \me\rangle=0$, the minimal angle gap among rows of $\Mat(\tS)$ is
-\begin{align}
-\Delta^2_{\min}(\tS)\asymp \tan^2(\Mat(\tS)_{1:}, \Mat(\tS)_{3:})
-=\left(\onorm{\me}_2 \over \onorm{\mone}_2\right)^{2(K-1)} \sigma^2 d^{-\gamma}\asymp \sigma^2 d^{-\gamma}.
-\end{align}
-Therefore, we have shown that $\tP_{\text{shifited}}(\gamma)=\tP(\gamma)$.
-Finally, the property (ii) follows directly by comparing the definition of $\tS$ in~\eqref{eq:S} with that in the proof of \citet[Theorem 7]{han2020exact}.
-\end{proof}
-
-\section{Proof of Theorem~\ref{thm:initial}}
-\begin{proof}[Proof of Theorem 4] We prove Theorem~\ref{thm:initial} under the symmetric dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
-We drop the subscript $k$ in the matricizations $\mM_k, \mX_k, \mS_k$.
-For simplicity, let $\hat z$ denote the output, $\hat z^{(0)}$, of Sub-algorithm~\hyperref[alg:main]{1}.
-
-First, by Lemma~\ref{lem:angle_gap_x}, there exists a positive constant such that $\min_{z(i) \neq z(j)} \onorm{ \mX^s_{i:} - \mX^s_{j:} } \geq c_0  \Delta_{\min}$.
-By the balance assumption on $\mtheta$ and Lemma~\ref{lem:upper_mis}, we have
- \begin{equation}\label{eq:theta_bound}
-          \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2  \leq  \sum_{i \in S_I} \theta(i)^2 + 4 \sum_{i \in S} \theta(i)^2 ,
-    \end{equation}
-    where 
-    \begin{equation}
-        S_0 = \{i: \onormSize{}{\hat \mX_{i:}} = 0\}, \quad S = \{i \in S_0^c: \onormSize{}{\hat \mx_{\hat z(i)} - \mX_{i:}^s } \geq c_0  \Delta_{\min}/2\}.
-    \end{equation}
-    On one hand, note that for any set $P \in [p]$,
-    \begin{align}
-        \sum_{i \in P} \onormSize{}{\mX_{i:}}^2 &= \sum_{i \in P} \onormSize{}{ \theta(i) \mS_{z(i):} (\mTheta \mM)^{T, \otimes (K-1)} }^2 \\
-        & \geq \sum_{i \in P} \theta(i)^2 \min_{a \in [r]} \onormSize{}{\mS_{a:}}^2 \lambda_{r}^{2(K-1)} (\mTheta \mM)\\
-        & \gtrsim \sum_{i \in P} \theta(i)^2 p^{K-1} r^{-(K-1)}, 
-    \end{align}
-    where the last inequality follows Lemma~\ref{lem:singular_thetam}, the assumption that $\min_{i \in [p]} \theta(i) \geq c$, and the constraint $\min_{a \in [r]} \onormSize{}{\mS_{a:}} \geq c_3$ in the parameter space~\eqref{eq:family}.
-Thus, we have
-    \begin{equation}\label{eq:theta_p}
-        \sum_{i \in P} \theta(i)^2 \lesssim \sum_{i \in P} \onormSize{}{\mX_{i:}}^2 p^{-(K-1)} r^{K-1}.
-    \end{equation}
-
-    On the other hand, note that 
-    \begin{align}
-         \sum_{i \in S} \onormSize{}{\mX_{i:}}^2 & \leq 2 \sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  +  2\sum_{i \in S} \onormSize{}{\hat \mX_{i:} - \mX_{i:}}^2 \label{eq:tr1}\\
-         & \leq \frac{8}{c_0^2 \Delta_{\min}^2  } \sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2   \onormSize{}{\hat \mx_{\hat z(i)} - \mX_{i:}^s }^2 +2 \onormSize{}{\hat \tX - \tX}_F^2 \label{eq:defs} \\
-         & \leq \frac{16}{c_0^2 \Delta_{\min}^2  }\sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  \off{ \onormSize{}{\hat \mx_{\hat z(i)} - \hat \mX_{i:}^s }^2 + \onormSize{}{\hat \mX_{i:}^s - \mX_{i:}^s }^2 } + 2\onormSize{}{\hat \tX - \tX}_F^2 \label{eq:tr2} \\
-         & \leq  \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 }\sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  \onormSize{}{\hat \mX_{i:}^s - \mX_{i:}^s }^2  + 2\onormSize{}{\hat \tX - \tX}_F^2 \label{eq:km} \\
-         & \leq \of{ \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 } + 2 }\onormSize{}{\hat \tX - \tX}_F^2 \label{eq:lem1}\\
-         & \lesssim \of{ \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 } + 2 } \of{ p^{K/2}r + pr^2 + r^K } \sigma^2 \label{eq:lem_err},
-    \end{align}
-    where inequalities \eqref{eq:tr1} and \eqref{eq:tr2} follow from the triangle inequality, \eqref{eq:defs} follows from the definition of $S$, \eqref{eq:km} follows from the update rule of $k$-means in Step 5 of Sub-algorithm~\hyperref[alg:main]{1}, \eqref{eq:lem1} follows from Lemma~\ref{lem:norm_diff}, and the last inequality~\eqref{eq:lem_err} follows from Lemma~\ref{lem:two-step_esterror}.
-Also, note that
-    \begin{equation}\label{eq:s0}
-         \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2 =  \sum_{i \in S_0} \onormSize{}{\hat \mX_{i:}- \mX_{i:}}^2 \leq \onormSize{}{\hat \tX - \tX}_F^2 \lesssim \of{ p^{K/2}r + pr^2 + r^K } \sigma^2,
-    \end{equation}
-    where the equation follows from the definition of $S_0$. Therefore, combining the inequalities~\eqref{eq:theta_bound}, \eqref{eq:theta_p}, \eqref{eq:lem_err}, and \eqref{eq:s0}, we have 
-    \begin{align}
-        \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2  &\lesssim \of{\sum_{i \in S} \onormSize{}{\mX_{i:}}^2 + \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2  } p^{-(K-1)} r^{K-1} \\
-        & \lesssim \frac{\sigma^2 r^{K-1}}{ \Delta_{\min}^2  p^{K-1} } \of{ p^{K/2}r + pr^2 + r^K }.
-    \end{align}
-    With the assumption that $\min_{i \in [p]} \theta(i) \geq c$, we finally obtain the result
-    \begin{equation}
-        \ell(z, z) \lesssim \frac{1}{p}\min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \lesssim \frac{r^K p^{-K/2} }{ \text{SNR} },
-    \end{equation}
-    where the last inequality follows from the definition $\text{SNR} = \Delta_{\min}^2/\sigma^2$.
-\end{proof}
-
-{\bf Useful Corollary of Theorem~\ref{thm:initial}}
-
-\begin{cor}[Initial misclustering loss]\label{cor:L0} 
-Under the setup of Theorem~\ref{thm:initial}, assume SNR $\geq \tilde  C p^{-K/2} \log p$ for some positive universal constant $\tilde C$ and $\sigma^2 = 1$.
-With probability going to 1, the misclustering loss for the initialization is upper bounded as
-\begin{equation}
-    L^{(0)} =  \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(0)}(i) = b } \onormSize{}{ [ \mS_{ \pi^{(0)}( z(i)):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq \frac{\bar C \Delta_{\min}^2}{ \tilde C r \log p},
-\end{equation}
-where $\pi^{(0)}$ minimizes the initial misclustering error; i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} $ $ \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$, and $\bar C > 1$ is a positive universal constant.
-\end{cor}
-
-\begin{proof} [Proof of Corollary~\ref{cor:L0}]
-Without loss of generality, we assume $\pi^{(0)}$ is the identity mapping such that  $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
-Note that $\mX_{i:}^s$ have only $r$ different values.
-We let $\mX_a^s = \mX_{i:}^s$ for all $i$ such that $z(i) = a, a \in [r]$.
-
-Notice that 
-\begin{equation}
-    \onormSize{}{\mX_{i:}}^2 \gtrsim p^{K-1} r^{-(K-1)} \quad \text{and} \quad \onormSize{}{\mX_{i:} - \hat \mX_{i:}}^2 \leq \onormSize{}{\hat \tX - \tX}_F^2 \lesssim  p^{K/2}r + pr^2 + r^K .
-\end{equation}
-Therefore, when $p$ is large enough, we have 
-\begin{align}
-    \sum_{i \in [p]} \onormSize{}{\mX_{i:}}^2 \onormSize{}{\hat \mX_i^s - \hat \mx_{z^{(0)}(i)}}^2 &\lesssim \sum_{i \in [p]} \of{\onormSize{}{\mX_{i:}}^2 - \onormSize{}{\mX_{i:} - \hat \mX_{i:}}^2} \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2\\
-   & \lesssim \sum_{i \in [p]} \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 \\
-   & \lesssim \eta  \sum_{i \in [p]} \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \mX_{i:}^s}^2 \\
-   & \lesssim \onormSize{}{\hat \tX - \tX}_F^2 \\
-   & \lesssim p^{K/2}r + pr^2 + r^K .\label{eq:cor_1}
-\end{align}
-
-Hence, we have 
-\begin{align}
-    \sum_{i \in [p]} \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 &\lesssim \sum_{i \in [p]} \theta(i)^2 \onormSize{}{\hat \mX_i^s - \hat \mx_{z^{(0)}(i)}}^2 \\
-    &\lesssim \frac{r^{K-1}}{p^{K-1}} \sum_{i \in [p]} \onormSize{}{\mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2\\
-    & \lesssim \frac{r^{K-1}}{p^{K-1}} \of{ p^{K/2}r + pr^2 + r^K },\label{eq:cor_2}
-\end{align}
-where the first inequality follows from the assumption $\min_{i \in [p]} \theta(i) \geq c$, the second inequality follows from the inequality~\eqref{eq:theta_p}, and the last inequality comes from the inequality~\eqref{eq:cor_1}.
-
-Next, we consider the following quantity,
-\begin{align}
-    \sum_{i \in [p]} \theta(i) \onormSize{}{ \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 &\lesssim \sum_{i \in [p]} \theta(i)^2 \onormSize{}{\mX_{i:}^s - \hat \mX_{i:}^s}^2 + \sum_{i \in [p]} \theta(i)^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)} }^2\\
-    & \lesssim \sum_{i \in [p]} \frac{ \theta(i)^2}{ \onormSize{}{\mX_{i:}}^2} \onormSize{}{\mX_{i:}  - \hat \mX_{i:}}^2 +  \sum_{i \in [p]} \theta(i)^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)} }^2\\
-    & \lesssim \frac{r^{K-1}}{p^{K-1}} \of{ p^{K/2}r + pr^2 + r^K} ,\label{eq:cor_3}
-\end{align}
-where the first inequality follows from the assumption of $\theta(i)$ and triangle inequality, the second inequality follows from Lemma~\ref{lem:norm_diff}, and the last inequality follows from~\eqref{eq:cor_2}.
-In addition, with Theorem~\ref{thm:initial} and the condition SNR $\gtrsim p^{-K/2} \log p$, for all $a \in [r]$, we have
-\begin{equation}
-    |z^{-1}(a) \cap (z^{(0)})^{-1} (a)| \geq |z^{-1}(a)| - p\ell(z^{(0)} ,  z) \gtrsim \frac{p}{r} - \frac{p}{\log p} \gtrsim \frac{p}{r},
-\end{equation}
-when $p$ is large enough. Therefore, for all $a \in [r]$, we have 
-\begin{align}
-       \onormSize{}{ \hat \mx_{a} - \mX_{a}^s  }^2  &= \frac{ \sum_{i \in z^{-1}(a) \cap (z^{(0)})^{-1} (a) } \onorm{ \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 }{|z^{-1}(a) \cap (z^{(0)})^{-1} (a) |}  \\
-       & \lesssim \frac{r}{p} \of{  \sum_{i \in [p]} \onormSize{}{ \mX_{i:}^s - \hat \mX_{i:}^s }^2 + \sum_{i \in [p]} \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 } \\
-       & \lesssim  \frac{r^{K}}{p^{K}}\of{ p^{K/2}r + pr^2 + r^K}, \label{eq:cor_4}
-\end{align}
-where the last inequality follows from the inequality~\eqref{eq:cor_2}.
-
-Finally, we obtain 
-\begin{align}
-     L^{(0)} &= \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(0)}(i) = b } \onormSize{}{ [ \mS_{  z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
-     & \lesssim \frac{1}{p}  \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i)  \onormSize{}{ \mX_{i:}^s -  \mX_{z^{(0)}(i)}^s  }^2\\
-     & \lesssim \frac{1}{p}  \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \of{ \onormSize{}{\mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 + \onormSize{}{ \hat \mx_{z^{(0)}(i)} - \mX_{z^{(0)}(i)}^s  }^2 } \\
-     & \leq \bar C \frac{r^{K}}{p^{K}} \of{ p^{K/2}r + pr^2 + r^K}, \\
-     & \leq \frac{\bar C\Delta_{\min}^2}{ \tilde C r \log p}
-\end{align}
-where the first inequality follows from Lemma~\ref{lem:angle_gap_x}, the third inequality follows from inequalities~\eqref{eq:cor_3} and \eqref{eq:cor_4}, and the last inequality follows from the assumption that SNR $\geq \tilde C p^{-K/2} \log p$.
-\end{proof}
-
-{\bf Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:initial}} 
-\begin{lem}[Basic inequality]\label{lem:norm_diff} For any two nonzero vectors $\mv_1,\mv_2$ of same dimension, we have 
-\[
-\sin(\mv_1 , \mv_2) \leq \onorm{{\mv_1^s}-{\mv_2^s}}\leq {2\onorm{\mv_1-\mv_2}\over \max\left(\onorm{\mv_1},\onorm{\mv_2}\right)}.
-\]
-\end{lem}
-\begin{proof}[Proof of Lemma~\ref{lem:norm_diff}]
-For the first inequality, let $\alpha \in [0,\pi]$ denote the angle between $\mv_1$ and $\mv_2$. We have 
-\begin{align}
-     \onorm{{\mv_1^s}-{\mv_2^s}} = \sqrt{2(1 - \cos \alpha)} = 2 \sin \frac{\alpha}{2} \geq \sin \alpha,
-\end{align}
-where the equations follows from the properties of trigonometric function and the inequality follows from the fact the $\cos \frac{\alpha}{2} \leq 1$ and $\sin \alpha = 2 \sin \frac{\alpha}{2} \cos \frac{\alpha}{2} > 0$ for $\alpha \in [0, \pi]$.
-
-For the second inequality, without loss of generality, we assume $\onorm{\mv_1}\geq \onorm{\mv_2}$. Then
-\begin{align}
-\onorm{{\mv_1^s}-{\mv_2^s}}&=\onorm{{\mv_1\over \onorm{\mv_1}}- {\mv_2\over \onorm{\mv_1}}+{\mv_2\over \onorm{\mv_1}}-{\mv_2\over \onorm{\mv_2}}}\\
-&\leq {\onorm{\mv_1-\mv_2}\over \onorm{\mv_1}}+{\onorm{\mv_2}\onorm{\mv_1}-\onorm{\mv_2}\over \onorm{\mv_1}\onorm{\mv_2}}\\
-&\leq {2\onorm{\mv_1-\mv_2}\over \onorm{\mv_2}}.
-\end{align}
-
-Therefore, Lemma~\ref{lem:norm_diff} is proved.
-\end{proof}
-
-\begin{defn}[Weighted padding vectors]\label{def:pad} For a vector $\ma = \entry{a_i} \in \bbR^d$, we define the padding vector of $\ma$ with the weight collection $\mw = \{\mw_i\colon \mw_i = \entry{w_{ik}} \in \bbR^{p_i}\}_{i = 1}^d$ as
-\begin{equation}\label{eq:paddef}
-    \pad_{\mw}(\ma) = [a_1 \circ \mw_1, \ldots, a_d \circ \mw_d]^T, \quad \text{ where} \quad  a_i \circ \mw_i = [a_i w_{i1},\ldots, a_i w_{i p_i}]^T, \text{ for all } i \in [d].
-\end{equation}
-Here we also view $\pad_{\mw}(\cdot)\colon \bbR^{d}\mapsto \bbR^{\sum_{i\in[d]}p_i}$ as an operator. 
-We have the bounds of the weighted padding vector
-\begin{equation}\label{eq:pad_bound}
-     \min_{i \in [d]} \onormSize{}{\mw_i}^2 \onormSize{}{\ma}^2 \leq \onormSize{}{\pad_{\mw}(\ma)}^2 \leq \max_{i \in [d]} \onormSize{}{\mw_i}^2 \onormSize{}{\ma}^2.
-\end{equation}
-Further, we define the inverse weighted padding operator $\pad^{-1}: \bbR^{\sum_{i\in[d]}p_i}\mapsto \bbR^d$ which satisfies 
-\begin{equation}
-    \pad^{-1}_{\mw}(\pad_{\mw}(\ma)) = \ma.
-\end{equation}
-\end{defn}
-
-\begin{lem}[Angle for weighted padding vectors]\label{lem:pad} Suppose we have two non-zero vectors $\ma, \mb \in \bbR^d$. Given the weight collection $\mw$, we have 
-\begin{equation}\label{eq:lempad}
-   \frac{\min_{i \in [d]} \onormSize{}{\mw_i}}{\max_{i \in [d]} \onormSize{}{\mw_i} } \sin (\ma, \mb) \stackrel{*}{\leq} \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb)) \stackrel{**}{\leq} \frac{ \max_{i \in [d]} \onormSize{}{\mw_i} }{\min_{i \in [d]} \onormSize{}{\mw_i}} \sin (\ma, \mb).
-\end{equation}
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:pad}] We prove the two inequalities separately with similar ideas.
-
-First, we prove the inequality ** in~\eqref{eq:lempad}. Decomposing $\mb$ yields
-\begin{equation}
-    \mb = \cos (\ma, \mb) \frac{\onormSize{}{\mb}}{\onormSize{}{\ma}} \ma + \sin (\ma, \mb) \frac{\onormSize{}{\mb}}{\onormSize{}{\ma^{\perp}}}\ma^{\perp},
-\end{equation}
-where $\ma^{\perp} \in \bbR^d$ is in the orthogonal complement space of $\ma$. By the Definition~\ref{def:pad}, we have 
-\begin{equation}\label{eq:pad_second_ineq}
-    \pad_{\mw}(\mb) = \cos (\ma, \mb) \frac{\onormSize{}{b}}{\onormSize{}{a}} \pad_{\mw}(\ma) + \sin (\ma, \mb) \frac{\onormSize{}{b}}{\onormSize{}{a^{\perp}}} \pad_{\mw}(\ma^{\perp}).
-\end{equation}
-Note that $\pad_{\mw}(\ma^{\perp})$ is not necessary equal to the orthogonal vector of $\pad(\ma)$; i.e., $\pad_{\mw}(\ma^{\perp}) \neq (\pad_{\mw}(\ma))^{\perp}$.
-By the geometry property of trigonometric functions, we obtain
-\begin{align}
-    \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb))  &\leq \frac{  \onormSize{}{\mb} \onormSize{}{\pad_{\mw}(\ma^{\perp})} }{ \onormSize{}{\ma^{\perp}} \onormSize{}{ \pad_{\mw}(\mb)}} \sin (\ma, \mb)  \\
-    & \leq  \frac{ \max_{i \in [d]} \onormSize{}{\mw_i} }{\min_{i \in [d]} \onormSize{}{\mw_i}} \sin (\ma, \mb),
-\end{align}
-where the second inequality follows by applying the property~\eqref{eq:pad_bound} to vectors $\mb$ and $\ma^{\perp}$.
-
-Next, we prove inequality * in~\eqref{eq:lempad}. With the decomposition of $\pad_{\mw}(\mb)$ and the inverse weighted padding operator, we have 
-\begin{align}
-    \mb &=  \cos(\pad_{\mw}(\ma), \pad_{\mw}(\mb)) \frac{\onormSize{}{\pad_{\mw}(\mb)}}{\onormSize{}{\pad_{\mw}(\ma)}} \ma +  \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb))\frac{\onormSize{}{\pad_{\mw}(\mb)}}{\onormSize{}{(\pad_{\mw}(\ma))^{\perp}}} \pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp}).
-\end{align}
-
-Therefore, we obtain 
-\begin{align}
-    \sin(\ma, \mb) &\leq \frac{ \onormSize{}{\pad_{\mw}(\mb) } \onormSize{}{\pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp})  } }{ \onormSize{}{(\pad_{\mw}(\ma))^{\perp} } \onormSize{}{\mb} }  \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb))\\
-    & \leq \frac{ \max_{i \in [d]} \onormSize{}{\mw_i} }{\min_{i \in [d]} \onormSize{}{\mw_i}} \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb)),
-\end{align}
-where the second inequality follows by applying the property~\eqref{eq:pad_bound} to vectors $\mb$ and $\pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp})$.
-\end{proof}
-
-\begin{lem}[Angle gap in $\tX$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family}.
-Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying~\eqref{eq:degree}.
-Then the angle gap in $\tX$ is asymptotically lower bounded by the angle gap in $\tS$; i.e., for all $i,j$ such that $z(i) \neq z(j)$
-\begin{equation}
-    \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \gtrsim  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
-\end{equation}
-
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:angle_gap_x}] 
-Note that the vector $\mS_{z(i):}$ can be folded to a tensor $\tS' =\entry{\tS'_{a_2,\ldots,a_K}} \in \bbR^{r^{K-1}}$; i.e., $\text{vec}(\tS') = \mS_{z(i):}$.
-Define weight vectors $\mw_{a_2, \cdots, a_K}$ correspond to the elements in $\tS'_{a_2,\ldots,a_K}$ by
-\begin{equation}
-    \mw_{a_2 \cdots a_K} = [ \mtheta_{z^{-1} (a_2)}^T \otimes \cdots \otimes \mtheta_{z^{-1} (a_K)}^T] \in \bbR^{|z^{-1} (a_2)| \times \cdots \times |z^{-1} (a_K)|},
-\end{equation}
-for all $a_k \in [r], k = 2,\ldots, K$, where $\otimes$ denotes the Kronecker product.
-Therefore, we have  $\mX_{i:} = \theta(i) \pad_{\mw}(\mS_{z(i):})$ where $\mw = \{ \mw_{a_2, \cdots, a_K}\}_{a_k \in [r], k\in [K]/\{1\}}$.
-Specifically, we have $\onormSize{}{\mw_{a_2, \ldots, a_K}}^2 = \prod_{k = 2}^K \onormSize{}{ \mtheta_{z^{-1}(a_k)}}^2$, and by the balanced assumption~\eqref{eq:degree} we have
-\begin{equation}\label{eq:pad_balance}
-    \max_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2 = (1 + o(1))  \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2.
-\end{equation}
-
-Consider the inner product of $\mX_{i:}$ and $\mX_{j:}$ for $z(i) \neq z(j)$.
-By the definition of weighted padding operator~\eqref{eq:paddef} and the balanced assumption~\eqref{eq:pad_balance}, we have
-\begin{align}
-    \ang{\mX_{i:},\mX_{j:}} &= \theta(i) \theta(j) \ang{ \pad_{\mw}(\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):} )}\\
-    &= \theta(i) \theta(j) \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2 \ang{\mS_{z(i):}, \mS_{z(j):}}(1 + o(1)).
-\end{align}
-Therefore, when $p$ large enough, the inner product $\ang{\mX_{i:},\mX_{j:}} $ has the same sign as $\ang{\mS_{z(i):}, \mS_{z(j):}}$.
-Next, we discuss the angle between $\mX_{i:}$ and $\mX_{j:}$ by two cases.
-
-\begin{enumerate}[wide]
-    \item \textbf{Case 1:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} \leq 0$.
-Then, we also have $\ang{\mX_{i:},\mX_{j:}} \leq 0$, which implies $\onormSize{}{\mX_{i:}^s -  \mX_{j:}^s} \geq \sqrt{2}$.
-Note that $\onormSize{}{\mS_{z(i):}^s - \mS_{z(j):}^s} \leq 2$ by the definition of angle gap.
-We have $\onormSize{}{\mX_{i:}^s -  \mX_{j:}^s} \gtrsim \onormSize{}{\mS_{z(i):}^s - \mS_{z(j):}^s}$.
-    \item \textbf{Case 2:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} > 0$.
-Then, we have $\cos(\mS_{z(i):}, \mS_{z(j):}) > 0$.
-Note that the fact $\sqrt{1 - \cos \alpha} = 2 \sin \frac{\alpha}{2} \lesssim \sin \alpha$ for the angle $\alpha \in [0, \frac{\pi}{2})$.
-Then, we have
-    \begin{align}
-        \onormSize{}{\mS_{z(i):}^s - \mS_{z(j):}^s} &= \sqrt{ 1 - \cos(\mS_{z(i):}, \mS_{z(j):})}  \\
-        & \lesssim \sin (\mS_{z(i):}, \mS_{z(j):})\\
-        & \leq \frac{ \max_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}} }{  \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}} }  \sin ( \pad_{\mw} (\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):})) \\
-        & \leq (1 + o(1)) \onormSize{}{\mX_{i:}^s -  \mX_{j:}^s},
-    \end{align}
-    where the second inequality follows from Lemma~\ref{lem:pad}, and the last inequality follows from the balanced weight~\eqref{eq:pad_balance} and Lemma~\ref{lem:norm_diff}.
-\end{enumerate}
-
-Hence, we conclude that for all $i,j$ such that $z(i) \neq z(j)$,
-\begin{equation}
-    \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \gtrsim  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
-\end{equation}
-\end{proof}
-
-\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the parameter space~\eqref{eq:family} and assumption that $\min_{i \in [p]} \theta(i) \geq c$, the singular values of $\mTheta \mM$ are bounded as
-\begin{equation}\label{eq:membership2}
-     \sqrt{p/r} \lesssim \sqrt{ \min_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 }  \leq \lambda_{r}(\mTheta \mM) \leq \onormSize{}{\mTheta \mM}_{\sigma} \leq\sqrt{ \max_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 } \lesssim  p /r .
-\end{equation}
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:singular_thetam}] Note that 
-\begin{equation}
-    (\mTheta\mM)^T \mTheta \mM = \mD, \quad \text{with} \quad \mD = \text{diag}(D_1,\ldots, D_r), \ D_a = \onormSize{}{\mtheta_{z^{-1}(a)}}^2, a \in [r].
-\end{equation}
-By the definition of singular values, we have 
-\begin{equation}
-     \sqrt{\min_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 } \leq \lambda_r(\mTheta \mM) \leq \onormSize{}{\mTheta\mM}_{\sigma} \leq \sqrt{ \max_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2}.
-\end{equation}
-
-Since that $\min_{i \in [p]}\theta(i) \geq c$ by the constraints in parameter space, we have  
-\begin{equation}
-   \min_{a \in [r]} \onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 \geq c^2 \min_{a \in [r]} |z^{-1}(a)| \gtrsim \frac{p}{r},
-\end{equation}
-where the last inequality follows from the constraint in parameter space~\eqref{eq:family}. Finally, notice that 
-\begin{equation}
-    \sqrt{\max_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 } \leq \max_{a \in [r]} \sqrt{ \onormSize{}{ \mtheta_{ z^{-1}(a)} }^2_1} \lesssim \frac{p}{r}.
-\end{equation}
-
- Therefore, we complete the proof of Lemma~\ref{lem:singular_thetam}.
-\end{proof}
-
-\begin{lem}[Singular-value gap-free tensor estimation error bound]\label{lem:two-step_esterror}Consider an order-$K$ tensor $\tA = \tX + \tZ \in \bbR^{p \times \cdots \times p}$, where $\tX$ has Tucker rank $(r,...r)$ and $\tZ$ has independent sub-Gaussian entries with parameter $\sigma^2$.
-Let $\hat \tX$ denote the double projection estimated tensor in Step 2 of Sub-algorithm~\hyperref[alg:main]{1} in the main paper.
-Then with probability at least $1 - C \exp\of{- cp }$, we have
-\begin{align}
-    \onormSize{}{\hat \tX - \tX}_F^2 \leq C \sigma^2 \of{ p^{K/2}r + pr^2 + r^K },
-\end{align}
-where $C, c$ are some positive constants.
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:two-step_esterror}]
-See \citet[Proposition 1]{han2020exact}.
-\end{proof}
-
-\begin{lem}[Upper bound of misclustering error]\label{lem:upper_mis} Let $z: [p] \mapsto [r]$ be a cluster assignment such that $|z^{-1}(a)| \asymp p/r$ for all $a \in [r]$.
-Let node $i$ correspond to a vector $\mx_i  = \theta(i) \mv_{z(i)} \in \bbR^d$, where $\{\mv_a\}_{a = 1}^r$ are the cluster centers and $\mtheta = \entry{\theta(i)} \in \bbR^p_+$ is the positive degree heterogeneity.
-Assume that  $\mtheta$ satisfies the balanced assumption~\eqref{eq:degree} such that ${\max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 \over \min_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 }= 1 + o(1)$.
-Consider an arbitrary estimate $\hat z$ with $\hat \mx_i = \hat \mv_{\hat z(i)}$ for all $ i \in S$.
-Then, if
-\begin{equation}\label{eq:upper_mis_cond}
-    \min_{a \neq b \in [r]} \onormSize{}{\mv_a - \mv_b} \geq 2c,
-\end{equation}
- for some constant $c >0$, we have 
-\begin{equation}
-    \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \leq \sum_{i \in S_0} \theta(i)^2 + 4 \sum_{i \in S} \theta(i)^2,
-\end{equation}
-where $S_0$ is defined in Step 3 of Sub-algorithm~\hyperref[alg:main]{1} and
-\begin{equation}
-   S = \{i \in S_0^c: \onormSize{}{\hat \mx_i - \mv_{z(i)} } \geq c \}.
-\end{equation}
-
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:upper_mis}]
-
-For each cluster $u\in[r]$, we use $C_u$ to collect the subset of points for which the estimated and true positions $\hat \mx_i, \mx_i$ are within distance $c$. Specifically, define
-\begin{equation}
-    C_u = \{ i \in z^{-1}(u) \cap S_0^c: \onormSize{}{\hat \mx_i - \mv_{z(i)}} < c \},
-\end{equation}
-and divide $[r]$ into three groups based on $C_u$ as 
-\begin{align}
-    R_1 &= \{ u \in [r]: C_u = \emptyset \},\\
-    R_2 &= \{ u \in [r]: C_u \neq \emptyset, \text{ for all } i, j \in C_u, \hat z(i) = \hat z(j) \},\\
-    R_3 &= \{ u \in [r]: C_u \neq \emptyset, \text{ there exist } i, j \in C_u, \hat z(i) \neq \hat z(j) \}.
-\end{align}
-Note that $\cup_{u \in [r]}C_u = S_0^c/S^c$ and $C_u \cap C_v = \emptyset$ for any $u \neq v$.
-Suppose there exist $ i \in C_u$ and $j \in C_v$ with $u \neq v \in [r]$ and $\hat z(i) = \hat z(j)$.
-Then we have
-\begin{equation}
-    \onormSize{}{\mv_{z(i)} - \mv_{z(j)} } \leq  \onormSize{}{\mv_{z(i)} - \hat \mx_{i} } + \onormSize{}{\mv_{z(j)} - \hat \mx_j } < 2c,
-\end{equation}
-which contradicts to the assumption~\eqref{eq:upper_mis_cond}.
-Hence, the estimates $\hat z(i) \neq \hat z(j)$ for the nodes $ i \in C_u$ and $j \in C_v$ with $u \neq v$.
-By the definition of $R_2$, the nodes in $\cup_{u \in R_2} C_u$ have the same assignment with $z$ and $\hat z$.
-Then,
-%Following \citet[Lemma 6]{gao2018community} \fixme{Miaoyan}{Remove citation. fill in details in a self-contained way. }, 
-we have 
-\begin{equation}
-    \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \leq \sum_{i \in S_0} \theta(i)^2 + \sum_{i \in S} \theta(i)^2 + \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2.
-\end{equation}
- We only need to bound $\sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2$ to finish the proof.
-Note that every $C_u$ with $u \in R_3$ contains at least two nodes assigned to different clusters by $\hat z$.
-Then, we have $|R_2| + 2 |R_3| \leq r$.
-Since $|R_1| + |R_2| + |R_3| = r$, we have $|R_3| \leq |R_1|$.
-Hence, we obtain
-\begin{align}
-    \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2 &\leq |R_3| \max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 \\
-    & \leq |R_1| \max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2  \\
-    & \leq {\max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 \over \min_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 } \sum_{i \in \cup_{u \in R_1} z^{-1}(u)} \theta(i)^2\\
-    & \leq 2 \sum_{i \in S} \theta(i)^2,
-\end{align}
-where the last inequality holds by the balanced assumption on $\mtheta$ when $p$ is large enough, and the fact that $ \cup_{u \in R_1} z^{-1}(u) \subset S$.
-\end{proof}
-
-\section{Proof of Theorem~\ref{thm:refinement}}
-
-\begin{proof}[Proof of Theorem~\ref{thm:refinement}]
-We prove Theorem~\ref{thm:refinement} under the symmetric dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
-We drop the subscript $k$ in the matricizations $\mM_k, \mS_k, \mX_k$.
-Without loss of generality, we assume that the variance $\sigma = 1$, and that the identity permutation minimizes the initial misclustering error; i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
-
-{\bf Step 1 (Notation and conditions).} We first introduce additional notations and the necessary conditions used in the proof.
-We will verify that the conditions hold in our context under high probability in the last step of the proof.
-
-\textbf{Notation.}
-\begin{enumerate}[wide]
-
-\item Projection.
-We use $\mI_d$ to denote the identity matrix of dimension $d$.
-For a vector $\mv \in \bbR^d$, let $\text{Proj}(\mv) \in \bbR^{d \times d}$ denote the projection matrix to $\mv$.
-Then, $\mI_d - \text{Proj}(\mv)$ is the projection matrix to the orthogonal complement $\mv^{\perp}$.
-
-    \item We define normalized membership matrices
-    \begin{align}
-        \mW = \mM \of{ \text{diag}(\mone_{p}^T \mM) }^{-1}, \quad \mW^{(t)} = \mM^{(t)} \of{ \text{diag}(\mone_{p}^T \mM^{(t)}) }^{-1},
-    \end{align}
-    and the dual normalized membership matrices
-    \begin{align}
-         \mV &= \mW^{\otimes (K-1)}, \quad \mV^{(t)} =\of{ \mW^{(t)}}^{\otimes (K-1)}.
-    \end{align}
-    \item We use $\tS^{(t)}$ to denote the estimator of $\tS$ in the $t$-th iteration, and we use $\tilde \tS$ to denote the oracle estimator of $\tS$ given true assignment $z$; i.e.,
-    \begin{align}
-        \tS^{(t)} = \tY \times_1 \of{\mW^{(t)}}^T \times_2 \cdots \times_K \of{\mW^{(t)}}^T, \quad \tilde \tS &= \tY \times_1 \mW^T \times_2 \cdots \times_K \mW^T.
-    \end{align}
-    \item We define the matricizations of tensors
-    \begin{align}
-        &\mS = \mat(\tS), \quad \mS^{(t)} = \mat(\tS^{(t)}), \quad \tilde \mS = \mat(\tilde \tS)\\
-        &\mY = \mat(\tY), \quad \mX = \mat(\tX), \quad \mE = \mat(\tE).
-    \end{align}
-    \item We define the angle-based misclustering loss in the $t$-th iteration 
-    \begin{equation}
-         L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2,
-    \end{equation}
-     and the oracle loss
-    \begin{align}
-         \xi & = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \Big\{\ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s}  \leq - \frac{ \theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \Big\} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2,
-    \end{align}
-    where $m$ is a positive universal constant defined in~\eqref{eq:xv_lower}.
-\end{enumerate}
-
-\begin{condition}(Intermediate results) \label{cond:origin} Let $\mathbb{O}_{p, r}$ denote the collection of all the $p$-by-$r$ matrices with orthonormal columns. We have 
-\begin{equation}\label{eq:cond1}
-    \onormSize{}{\mE \mV}_\sigma \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}} \of{ p ^{1/2}+ r^{(K-1)/2}}, \quad \onormSize{}{\mE \mV}_F \lesssim \sqrt{\frac{ r^{2(K-1)}}{p^{K-2}}},\quad \onormSize{}{ \mW_{a:}^T \mE \mV} \lesssim \frac{r^K}{p^{K/2}} \text{ for all }a \in [r],
-\end{equation}
-\begin{equation}\label{eq:cond2}
-    \sup_{ \mU_k \in \mathbb{O}_{p, r}, k = 2,\ldots, K } \onormSize{}{ \mE (\mU_{2} \otimes \cdots \otimes \mU_K)}_\sigma \lesssim \of{ \sqrt{r^{K-1}} + K\sqrt{pr}},
-\end{equation}
-
-\begin{equation}\label{eq:cond3}
-    \sup_{ \mU_k \in \mathbb{O}_{p, r}, k = 2,\ldots, K } \onormSize{}{ \mE (\mU_{2} \otimes \cdots \otimes \mU_K)}_F \lesssim \of{ \sqrt{pr^{K-1}} + K\sqrt{pr}},
-\end{equation}
-
-\begin{equation}\label{eq:cond_oracle}
-    \xi \leq \exp\of{ - M \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}},
-\end{equation}
-
-\begin{equation}\label{eq:cond_intial}
-    L^{(t)} \leq \frac{\bar C}{\tilde C} \frac{\Delta_{\min}^2}{r \log p}, \quad \text{for} \quad t = 0, 1, \ldots, T,
-\end{equation}
-where $M$ is a positive universal constant in inequality~\eqref{eq:constant_M}, $\bar C, \tilde C$ are universal constants in Corollary~\ref{cor:L0} and assumption SNR $\geq \tilde C p^{-K/2} \log p$, respectively.
-Further, inequality~\eqref{eq:cond1} holds by replacing $\mV$ to $\mV^{(t)}$ and $\mW_{:a}$ to $\mW_{:a}^{(t),T}$ when initialization condition~\eqref{eq:cond_intial} holds.
-\end{condition}
-
-{\bf Step 2 (Misclustering loss decomposition).} Next, we derive the upper bound of $L^{(t+1)}$ for $t = 0 ,1, \ldots, T-1$.
-By Sub-algorithm~\hyperref[alg:main]{2}, we update the assignment in $t$-th iteration via
-    \begin{equation}
-        z^{(t+1)}(i) = \argmin_{a \in [r]} \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{a:}^{(t)}]^s }^2,
-    \end{equation}
-    following the facts that $\onormSize{}{\ma^s - \mb^s}^2 = 1 - \cos(\ma, \mb)$ for vectors $\ma,\mb$ of same dimension and $\mat(\tY^\text{d}) = \mY \mV^{(t)}$ where $\tY^\text{d}$ is the reduced tensor defined in Step 8 of Sub-algorithm~\hyperref[alg:main]{2}.
-Then the event $z^{(t+1)}(i) = b$ implies
-    \begin{equation}\label{eq:event}
-        \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{b:}^{(t)}]^s }^2 \leq \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{z(i):}^{(t)}]^s }^2.
-    \end{equation}
-Note that the event~\eqref{eq:event} also holds for the degenerate entity $i$ with $\onormSize{}{\mY_{i:} \mV^{(t)}} = 0$ due to the convention that $\ma^s = {\bf 0}$ if $\ma = {\bf 0}$. 
-    Arranging the terms in \eqref{eq:event} yields the decomposition
-    \begin{equation}\label{eq:decomp}
-        2 \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s } \leq \onormSize{}{\mX_{i:} \mV^{(t)}} \of{ - \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 + G_{ib}^{(t)} + H_{ib}^{(t)} } +  F_{ib}^{(t)},
-    \end{equation}
-    where
-    \begin{align}
-        F_{ib}^{(t)} &= 2 \ang{\mE_{i:} \mV^{(t)}, \of{ [  \tilde \mS_{z(i):} ]^s  -  [  \mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  -  [  \mS_{b:}^{(t)} ]^s  }  }  + 2 \ang{ \mE_{i:} \of{ \mV - \mV^{(t)} }, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }, \label{eq:f}\\
-        G_{ib}^{(t)} &=   \of{ \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2}   \\
-        & \quad \quad \quad \quad \quad -   \of{ \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{b:}^{(t)}  ]^s}^2 -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 }, \label{eq:g} \\
-        H_{ib}^{(t)} &=   \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2  + \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 .
-\label{eq:h}
-    \end{align}
-
-   Therefore, the event $ \ind \offf{ z^{(t+1)}(i) = b }$ can be upper bounded as
-    \begin{align}
-         \ind \offf{ z^{(t+1)}(i) = b  }& \leq \ind \offf{ z^{(t+1)}(i) = b ,   \ang{ \mE_{j:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{1}{4} \onormSize{}{\mX_{i:} \mV^{(t)}} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \\
-         & + \ind \offf{z^{(t+1)}(i) = b, \frac{1}{2}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq \onormSize{}{\mX_{i:} \mV^{(t)}}^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }.
-\label{eq:ind_b}
-    \end{align}
-   Note that 
-    \begin{align}
-     \onormSize{}{\mX_{i:} \mV^{(t)}}  = \theta(i) \onormSize{}{\mS_{i:} (\mTheta \mM)^{\otimes (K-1),T} \mW^{(t), \otimes^{K-1}}}
-         \geq \theta(i) \onormSize{}{\mS_{z(i):}} \lambda_r^{K-1}(\mTheta \mM) \lambda^{K-1}_r(\mW^{(t)})
-         \geq\theta(i) m , \label{eq:xv_lower}
-    \end{align}
-    where the first inequality follows from the property of eigenvalues; the last inequality follows from Lemma~\ref{lem:singular_thetam}, Lemma~\ref{lem:membership}, and assumption that $\min_{a \in [r]} \onormSize{}{\mS_{z(i):}} \geq c_3>0$; and $m >0$ is a positive constant related to $ c_3$.
-Plugging the lower bound of $ \onormSize{}{\mX_{i:} \mV^{(t)}}$ \eqref{eq:xv_lower} into the inequality \eqref{eq:ind_b} gives
-    \begin{equation}\label{eq:decomp1}
-        \ind \offf{ z^{(t+1)}(i) = b  } \leq A_{ib} + B_{ib},
-    \end{equation}
-    where 
-    \begin{align}
-        A_{ib} &= \ind \offf{z^{(t+1)}(i) = b, \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 },\\
-        B_{ib} &= \ind \offf{z^{(t+1)}(i) = b, \frac{1}{2}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }.
-    \end{align}
- Taking the weighted summation of~\eqref{eq:decomp1} over $i \in [p]$ yields 
-    \begin{equation}
-        L^{(t+1)} \leq \xi + \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}^{(t)}, 
-    \end{equation}
-    where $\xi$ is the oracle loss such that 
-    \begin{equation}\label{eq:decomp2}
-        \xi =   \frac{1}{p} \sum_{i \in [p] } \theta(i) \sum_{b \in [r]/z(i)} A_{ib} \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2.
-    \end{equation}
-    Similarly to $\xi$ in~\eqref{eq:decomp2}, we define
-    \begin{equation}
-        \zeta_{ib}^{(t)} =   \theta(i) B_{ib} \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2.
-    \end{equation}
-
-    {\bf Step 3 (Derivation of contraction inequality).} In this step we derive the upper bound of $\zeta_{ib}$ and obtain the contraction inequality~\eqref{eq:proof_5_ineq}.
-Choose the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ that satisfies the condition of Lemma~\ref{lem:upper_fgh}, inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
-Note that
-    \begin{align}
-        \zeta_{ib}^{(t)} &= \theta(i) \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \ind \offf{z^{(t+1)}(i) = b, \frac{1}{2}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }\\
-        & \leq \theta(i) \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \ind \offf{z^{(t+1)}(i) = b, \frac{1}{4}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} } \\
-        & \leq 64 \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ cm^2\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    }
-        %&\leq  \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    },\label{eq:zeta}
-    \end{align}
-    where the first inequality follows from the inequality~\eqref{eq:hib} in Lemma~\ref{lem:upper_fgh}, and the last inequality follows from the assumption that $\min_{i \in [p]} \theta(i) \geq c>0$.
-Following \citet[Step 4, Proof of Theorem 2]{han2020exact} and Lemma~\ref{lem:upper_fgh}, we have
-    \begin{align}\label{eq:f_sum}
-          \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{(F_{ib}^{(t)})^2}{ cm^2 \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \leq \frac{C_0 \bar C}{c m^2 \tilde C^2} L^{(t)},
-    \end{align}
-   for a positive universal constant $C$ and 
-    \begin{align}
-         \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}  &\leq \frac{1}{512} \frac{1}{p} \sum_{i \in [p]}\theta(i) \sum_{b \in [r]/z(i)}   \ind \offf{ z^{(t+1)}(i) = b}  (\Delta_{\min}^2 + L^{(t)}) \\
-         & \leq \frac{1}{512} (L^{(t+1)} + L^{(t)}), \label{eq:g_sum}
-    \end{align}
-    where the last inequality follows from the definition of $L^{(t)}$ and the constraint of $\mtheta$ in parameter space~\eqref{eq:family}. 
-    %Choosing the constants $C, \bar C, C_2$ in~\eqref{eq:zeta}~\eqref{eq:f_sum}~\eqref{eq:g_sum} carefully \fixme{Miaoyan}{be specific. Add one sentence to explain}, 
-   For $\tilde C$ also satisfies 
-    \begin{equation}\label{eq:tilde_c3}
-         \frac{C_0 \bar C}{c m^2 \tilde C^2} \leq \frac{1}{512},
-    \end{equation}
-    we have 
-    \begin{equation}\label{eq:zeta_upper}
-        \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}^{(t)} \leq  \frac{1}{8} L^{(t+1)}  + \frac{1}{4} L^{(t)}.
-    \end{equation}
-    %for some small \fixme{Miaoyan}{In what sense "small"? Add one sentence.} positive constants $C_4, C_5$. 
-    Plugging the inequality~\eqref{eq:zeta_upper} into the decomposition~\eqref{eq:decomp2}, we obtain the contraction inequality 
-    \begin{equation}\label{eq:decomp_t}
-          L^{(t+1)} \leq \frac{3}{2} \xi  + \frac{1}{2} L^{(t)},
-    \end{equation}
-     where $\frac{1}{2}$ is the contraction parameter.
-
-     Therefore, with $\tilde C$ satisfying inequalities~\eqref{eq:tilde_c3}, \eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}, we obtain the conclusion in Theorem~\ref{thm:refinement} via inequality~\eqref{eq:decomp_t} combining the inequality~\eqref{eq:cond_oracle} in Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.
-
-    {\bf Step 4 (Verification of Condition~\ref{cond:origin}).} Last, we verify the Condition~\ref{cond:origin} under high probability to finish the proof.
-Note that the inequalities~\eqref{eq:cond1}, \eqref{eq:cond2}, and \eqref{eq:cond3} describe the property of the sub-Gaussian noise tensor $\tE$, and the readers can find the proof directly in \citet[Step 5, Proof of Theorem 2]{han2020exact}.
-Here, we include only the verification of inequalities~\eqref{eq:cond_oracle} and \eqref{eq:cond_intial}.
-
-    Now, we verify the oracle loss condition~\eqref{eq:cond_oracle}. Recall the definition of $\xi$,
-    \begin{align}
-        \xi = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \offf{  \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2,
-    \end{align}
-
-    Let $e_i = \mE_{i:} \mV$ denote the aggregated noise vector for all $i \in [p]$, and $e_i$'s are independent zero-mean sub-Gaussian vector in $\bbR^{r^{K-1}}$.
-The entries in $e_i$ are independent zero-mean sub-Gaussian variables with sub-Gaussian norm upper bounded by $m_1\sqrt{r^{K-1}/p^{K-1}}$ with some positive constant $m_1$.
-We have the probability inequality
-    \begin{align}
-        \bbP \of{ \ang{ e_i, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{ \theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \leq P_1 + P_2 + P_3,
-    \end{align}
-    where 
-    \begin{align}
-        P_1 &= \bbP \of{ \ang{e_i, [  \mS_{z(i):} ]^s - [ \mS_{b:} ]^s} \leq -\frac{\theta(i)m}{8}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 }, \\
-        P_2 &= \bbP \of{ \ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s} \leq -\frac{\theta(i)m}{16}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 },\\
-        P_3 &= \bbP \of{ \ang{e_i, [  \mS_{b:} ]^s - [\tilde  \mS_{b:} ]^s} \leq -\frac{\theta(i)m}{16}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 }.
-    \end{align}
-
-    For $P_1$, notice that the inner product $\ang{e_j,\mS_{z(j):}^s - \mS_{b:}^s} $ is a sub-Gaussian variable with sub-Gaussian norm bounded by $m_2 \sqrt{ r^{K-1}/ p^{K-1} } \onormSize{}{ \mS_{z(i):}^s -  \mS_{b:}^s }$ with some positive constant $m_2$.
-Then, by Chernoff bound, we have
-    \begin{equation}\label{eq:p1_2}
-         P_1 \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(j):}  ]^s - [ \mS_{b:}  ]^s  }^2 }.
-    \end{equation}
-
-    For $P_2$ and $P_3$, we only need to derive the upper bound of $P_2$ due to the symmetry. By the law of total probability, we have 
-    \begin{equation}\label{eq:p2}
-        P_2 \leq P_{21} + P_{22},
-    \end{equation}
-    where with some positive constant $t>0$,
-    \begin{align}
-        P_{21}& =  \bbP \of{ t \leq   \onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s }},\\
-        P_{22} &= \bbP \of{ \left.
-\ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s} \leq -\frac{\theta(i)m}{16}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \right| \onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s } < t  }.
-    \end{align}
-
-    For $P_{21}$, note that the term $\mW^T_{: z(i)} \mE \mV = \frac{ \sum_{ j \neq i, j \in [p]} \ind \{ z(j) = z(i) \} e_j }{\sum_{j \in [p]} \ind \{ z(j) = z(i) \} }$  is a sub-Gaussian vector with sub-Gaussian norm bounded by $m_3 \sqrt{r^K/ p^K}$ with some positive constant $m_3$.
-This implies
-        \begin{align}
-        P_{21} \leq \bbP\of{ t \onormSize{}{\mS_{z(i):} } \leq \onormSize{}{ \tilde  \mS_{z(i):}  -  \mS_{z(i):} }  } \leq \bbP\of{c_3 t \leq  \onormSize{}{ \mW^T_{: z(i)} \mE \mV  } } 
-        \lesssim \exp\of{ - \frac{p^K t^2}{r^K} }, \label{eq:p21}
-    \end{align}
-    where the first inequality follows from the basic inequality in Lemma~\ref{lem:norm_diff}, the second inequality follows from the assumption that $ \min_{a \in [r]} \onormSize{}{\mS_{z(i):}} \geq c_3>0$ in~\eqref{eq:family}, and the last inequality follows from the Bernstein inequality.
-
-    For $P_{22}$, the inner product $ \ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s}$ is also a sub-Gaussian variable with sub-Gaussian norm $ m_4 \sqrt{ r^{K-1}/ p^{K-1} } t$, conditioned on $\onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s } < t$ with some positive constant $m_4$.
-Then, by Chernoff bound, we have
-    \begin{equation}\label{eq:p22}
-        P_{22} \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1} t^2}   \onormSize{}{ [ \mS_{z(j):}  ]^s - [ \mS_{b:}  ]^s  }^4 }. 
-    \end{equation}
-    We take $t =  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }$ in $P_{21}$ and $P_{22}$, and plug the inequalities \eqref{eq:p21} and \eqref{eq:p22} into to the upper bound for $P_2$ in~\eqref{eq:p2}.
-We obtain that
-    \begin{equation}\label{eq:p2_2}
-        P_2 \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 }.
-    \end{equation}
-    Combining the upper bounds~\eqref{eq:p1_2} and \eqref{eq:p2_2} gives 
-       \begin{equation}\label{eq:xi_p}
-          \bbP \of{ \ang{ e_i, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \lesssim  \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 }.
-    \end{equation}
-
-    Hence, we have 
-    \begin{align}
-        \bbE \xi & =  \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \bbP \offf{  \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2 \\
-        &\lesssim \frac{1}{p} \sum_{i \in [p]}  \theta(i) \max_{i \in [p], b \in [r]}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2   \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \\
-        & \leq \exp \of{  - M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }, \label{eq:constant_M}
-    \end{align}
-    where $M$ is a positive constant, the first inequality follows from the constraint that $\sum_{i \in [p]} \theta(i) = p$, and the last inequality follows from~\eqref{eq:xi_p}.
-
-    By Markov's inequality, we have 
-    \begin{equation}
-        \bbP\of{ \xi \lesssim \bbE \xi + \exp \of{  -M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } } \geq 1 -  C \exp \of{  - M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 },
-    \end{equation}
-    and thus the condition~\eqref{eq:cond_oracle} holds with probability at least $1 -  C \exp \of{  - M\frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }$ for some constant $C > 0$.
-
-    Finally, we verify the bounded loss condition~\eqref{eq:cond_intial} by induction.
-With output $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{2} and the assumption SNR $\geq \tilde C p^{-K/2} \log p$, by Corollary~\ref{cor:L0}, we have
-    \begin{equation}
-        L^{(0)} \leq \frac{\bar C \Delta_{\min}^2}{\tilde C r \log p},\quad \text{    when $p$ is large enough}.
-    \end{equation}
-Therefore, the condition~\eqref{eq:cond_intial} holds for $t = 0$. Assume the condition~\eqref{eq:cond_intial} also holds for all $t \leq t_0$. Then, by the decomposition~\eqref{eq:decomp_t}, we have 
-    \begin{align}
-         L^{(t_0+1)} &\leq \frac{3}{2} \xi + \frac{1}{2} L^{(t_0)} \\
-         & \leq \exp \of{  - M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } + \frac{\Delta_{\min}^2}{r \log p} \\
-         & \leq \frac{\bar C}{\tilde C}\frac{\Delta_{\min}^2}{r\log p} ,
-    \end{align}
-    where the second inequality follows from the condition~\eqref{eq:cond_oracle} and the last inequality follows from the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2} \log p$.
-Thus, the condition~\eqref{eq:cond_intial} holds for $t_0 + 1$, and the condition~\eqref{eq:cond_intial} is proved by induction.
-\end{proof}
-
-{\bf Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:refinement}}
-
-\begin{lem}[Misclustering error and loss]\label{lem:mis} Define the misclustering error in the $t$-th iteration as $\ell^{(t)} = \ell(z^{(t)}, z)$. We have 
-\begin{equation}
-    \ell^{(t)} \lesssim \frac{1}{p}\sum_{i \in [p]} \theta(i) \ind\offf{ z^{(t)}(i) \neq z(i) } \leq \frac{L^{(t)}}{\Delta_{\min}^2}.
-\end{equation}
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:mis}]
-By the definition of minimal gap in Assumption~\ref{assmp:min_gap}, we have 
-\begin{equation}
-         L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \geq \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \Delta_{\min}^2 \geq c \ell^{(t)} \Delta_{\min}^2,
-    \end{equation}
-    where the last inequality follows from the assumption $\min_{i \in [p]} \theta(i) \geq c>0$.
-\end{proof}
-
-\begin{lem}[Singular-value property of membership matrices]\label{lem:membership} Under the setup of Theorem~\ref{thm:refinement}, suppose the condition~\eqref{eq:cond_intial} holds.
-Then, for all $a \in [r]$, we have $|\of{z^{(t)}}^{-1}(a)| \asymp p/r$.
-Moreover, we have
-\begin{equation}\label{eq:membership1}
-    \lambda_{r}(\mM) \asymp \onorm{\mM}_{\sigma} \asymp \sqrt{p/r}, \quad  \lambda_{r}(\mW) \asymp \onorm{\mW}_{\sigma} \asymp \sqrt{r/p}.
-\end{equation}
-The inequalities~\eqref{eq:membership1} also hold by replacing $\mM$ and $\mW$ to $\mM^{(t)}$ and $\mW^{(t)}$ respectively. 
-Further, we have 
-\begin{equation}\label{eq:WWsingular}
-   \lambda_{r}(\mW\mW^T) \asymp \onorm{\mW\mW^T}_{\sigma} \asymp r/p,
-\end{equation}
-which is also true for $\mW^{(t)}\mW^{(t),T}$.
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:membership}] The proof for the inequality~\eqref{eq:membership1} can be found in \citet[Proof of Lemma 4]{han2020exact}
-
-For inequality~\eqref{eq:WWsingular}, note that for all $k \in [r]$,
-\begin{align}
-    \lambda_k(\mW\mW^T) = \sqrt{\text{eigen}_k (\mW \mW^T\mW\mW^T)} \asymp \sqrt{\frac{r}{p} \text{eigen}_k (\mW \mW^T)} = \sqrt{\frac{r}{p} \lambda^2_k(\mW)} \asymp \frac{r}{p},
-\end{align}
-where $\text{eigen}_k(\mA)$ denotes the $k$-th largest eigenvalue of the square matrix $\mA$, the first inequality follows the fact that $\mW^T\mW$ is a diagonal matrix with elements of order $r/p$, and the second equation follows from the definition of singular value.
-\end{proof}
-
-\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement},  assume the the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ is large enough to satisfy the inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
-We have
-    \begin{equation}\label{eq:fib}
-        \max_{i \in [p]} \max_{b \neq z(i)} \frac{\of{F_{ib}^{(t)}}^2}{ \onormSize{}{ [  \mS_{z(i):} ]^s  -  [ \mS_{b:} ]^s }^2 } \lesssim  \frac{r L^{(t)}}{ \Delta_{\min}^2} \onormSize{}{\mE_{i:} \mV}^2 + \of{1 +  \frac{ r L^{(t)}}{ \Delta_{\min}^2}} \onormSize{}{\mE_{i:} (\mV - \mV^{(t)})}^2 ,
-    \end{equation}
-    \begin{equation}\label{eq:gib}
-        \max_{i \in [p]} \max_{b \neq z(i)} \frac{ \of{  G_{ib}^{(t)} }^2  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2}  \leq  \frac{1}{512} \of{ \Delta_{\min}^2 +L^{(t)}},
-    \end{equation}
-    \begin{equation}\label{eq:hib}
-        \max_{i \in [p]} \max_{b \neq z(i)}  \frac{ \aabs{  H_{ib}^{(t)} }  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2} \leq \frac{1}{4}.
-    \end{equation}
-    \end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:upper_fgh}]
-We prove the each of the inequalities in Lemma~\ref{lem:upper_fgh} separately.
-\begin{enumerate}[wide]
-    \item Upper bound for $F_{ib}^{(t)}$, i.e., inequality \eqref{eq:fib}. Recall the definition of $F_{ib}^{(t)}$,
-    \begin{align}
-         F_{ib}^{(t)} = 2 \ang{\mE_{i:} \mV^{(t)}, \of{ [ \tilde \mS_{z(i):} ]^s  - [\mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  - [ \mS_{b:}^{(t)} ]^s  }  }  + 2 \ang{ \mE_{i:}(\mV - \mV^{(t)} ), [ \tilde \mS_{z(i):} ]^s - [ \tilde \mS_{b:} ]^s  }.
-    \end{align}
-    By Cauchy-Schwartz inequality, we have 
-    \begin{align}
-        \of{F_{ib}^{(t)}}^2 & \leq 8 \of{\ang{\mE_{i:} \mV^{(t)}, \of{ [  \tilde \mS_{z(i):} ]^s  -  [  \mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  -  [  \mS_{b:}^{(t)} ]^s  }  }  }^2 \\
-         & \quad \quad \quad \quad \quad + 8\of{ \ang{ \mE_{i:} (\mV - \mV^{(t)} ), [  \tilde \mS_{z(i):} ]^s - [ \tilde  \mS_{b:} ]^s  }}^2 \\
-         & \leq 8 \of{ \onormSize{}{\mE_{i:} \mV}^2 + \onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2 } \max_{a \in [r]^s} \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  } \\
-          & \quad \quad \quad \quad \quad +\onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2  \onormSize{}{  [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }. \label{eq:fib_decomp}
-    \end{align}
-    Note that for all $a \in [r]$,
-    \begin{align}
-        \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  }^2 &= \onormSize{}{ [ \mW_{:a}^T \mY \mV ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }^2 \\
-        &\leq 2 \onormSize{}{ [ \mW_{:a}^T \mY \mV ]^s  - [ \mW_{:a}^{(t),T} \mY \mV ]^s  }^2 + 2 \onormSize{}{[ \mW_{:a}^{(t),T} \mY \mV ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }^2\\
-        & \lesssim \frac{ r^2 (L^{(t)})^2}{ \Delta_{\min}^2} + \frac{ r r^{2K} + p r^{K+2}}{p^K} \frac{L^{(t)}}{\Delta_{\min}^2}\\
-        & \lesssim  r L^{(t)}+  \frac{ r r^{2K} + p r^{K+2}}{p^K} \frac{L^{(t)}}{\Delta_{\min}^2}\\
-        & \lesssim  r L^{(t)},\label{eq:fib_1}
-    \end{align}
-    where the second inequality follows from the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4} in Lemma~\ref{lem:intermediate}, the third inequality follows from the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}, and the last inequality follows from the assumption that $\Delta_{\min}^2 \geq \tilde C p^{-K/2}\log p$.
-
-    Note that 
-    \begin{align}
-        \onormSize{}{  [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }^2 &=  \onormSize{}{  [  \tilde \mS_{z(i):} ]^s -[ \mS_{z(i):} ]^s +  [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s + [ \mS_{b:} ]^s -[  \tilde \mS_{b:} ]^s  }^2 \\
-        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2 + \max_{a \in [r]} \onormSize{}{ [ \mS_{a:} ]^s -[  \tilde \mS_{a:} ]^s }^2 \\
-        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2  + \max_{a \in [r]} \frac{1}{\onormSize{}{\mS_{a:} }^2} \onormSize{}{ \mW_{:a}^T \mE \mV}^2\\
-        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2, \label{eq:fib_2}
-    \end{align}
-    where the second inequality follows from Lemma~\ref{lem:norm_diff}, and the last inequality follows from the assumptions on $\onormSize{}{\mS_{a:}}$ in the parameter space~\eqref{eq:family}, the inequality~\eqref{eq:cond1} in Condition~\ref{cond:origin} and the assumption $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.
-
-    Therefore, we finish the proof of inequality~\eqref{eq:fib} by plugging the inequalities~\eqref{eq:fib_1} and \eqref{eq:fib_2} into the upper bound~\eqref{eq:fib_decomp}.
-
-    \item Upper bound for $G_{ib}^{(t)}$, i.e., inequality~\eqref{eq:gib}. By definition of $G_{ib}^{(t)}$, we rearrange terms and obtain
-    \begin{align}
-        G_{ib}^{(t)} &=   \of{ \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2}   \\
-        & \quad \quad \quad \quad \quad -   \of{ \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{b:}^{(t)}  ]^s}^2 -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 } \\
-        &= 2 \ang{  [\mX_{i:} \mV^{(t)}]^s , \of{[  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s } - \of{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s }}\\
-      &= G_1 + G_2 - G_3, \label{eq:gib_decomp}
-    \end{align}
-    where 
-    \begin{align}
-        G_1 &= \onormSize{}{ [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 - \onormSize{}{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s }^2,\\
-        G_2 &= 2 \ang{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s,   [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s},\\
-        G_3 & = 2 \ang{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s,  [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s}.
-    \end{align}
-
-    For $G_1$, we have 
-    \begin{align}
-        |G_1|^2 &\leq \aabs{ \onormSize{}{ [ \mW_{:z(i)}^T \mY \mV^{(t)} ]^s - [ \mS_{z(i):}^{(t)}  ]^s}^2 - \onormSize{}{ [ \mW_{:b}^T \mY \mV^{(t)} ]^s -  [ \mS_{b:}^{(t)}  ]^s }^2 }^2 \\
-        & \leq \max_{a \in [r]} \onormSize{}{ [ \mW_{:a}^T \mY \mV^{(t)} ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^4\\
-        & \leq C^4 \frac{r^4}{\Delta_{\min}^4} (L^{(t)})^4 +  \frac{r^2 r^{4K} + p^2r^{2K+4}}{p^{2K}} \frac{(L^{(t)})^2}{\Delta_{\min}^4} \\
-        & \leq C^4 \frac{\bar C}{\tilde C^3} \of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}}, \label{eq:g1}
-    \end{align}
-    where the third inequality follows from the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:membership} and the last inequality follows from the assumption that $\Delta_{\min}^2 \geq \tilde C p^{-K/2}\log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
-
-    For $G_2$, noticing that $[\mX_{i:} \mV^{(t)}]^s = [ \mW_{z(i):}^T\mX \mV^{(t)}]^s$, we have 
-    \begin{align}
-        |G_2|^2 &\leq 2\onormSize{}{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 \onormSize{}{ [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s }^2 \\
-        & \leq \frac{2}{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2} \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
-        & \leq C' \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} \of{ \frac{r^2}{\Delta_{\min}^2} (L^{(t)})^2 + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} }\\
-        & \leq \frac{C'}{\tilde C^2} \Delta_{\min}^2 L^{(t)},\label{eq:g2}
-    \end{align}
-    where $C'$ is a positive universal constant, the second inequality follows from Lemma~\ref{lem:norm_diff}, the third inequality follows from the inequality~\eqref{eq:cond2} in Condition~\ref{cond:origin}, the inequalities~\eqref{eq:inter5} and \eqref{eq:j11_wxvt} in the proof of Lemma~\ref{lem:intermediate},  and the last inequality follows from the assumption $\Delta_{\min}^2 \geq \tilde C p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
-
-    For $G_3$, note that by triangle inequality
-    \begin{align}
-        \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2 &\leq \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + 2\max_{a \in [r]}\onormSize{}{[\mW_{:a}^T\mX \mV^{(t)}]^s - [\mW_{:a}^T\mX \mV]^s }^2\\
-        & \leq \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + C \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2}, \label{eq:g3xv_wxv}
-    \end{align}
-    where the last inequality follows from the inequality~\eqref{eq:j1} in the proof of Lemma~\ref{lem:intermediate} and $C$ is a positive constant.
-   Then we have 
-    \begin{align}
-        |G_3|^2 & \leq 2 \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2 \\
-        & \leq 2 \of{ \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  +  \onormSize{}{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  } \\
-        & \quad \times \max_{a \in [r]} \onorm{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
-        & \leq C^2 \of{\onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + C \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} } \of{ \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2}}+\frac{C'}{\tilde C^2} \Delta_{\min}^2 L^{(t)} \\
-        & \leq \frac{C^2 \bar C^2}{\tilde C} \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 (\Delta_{\min}^2 + L^{(t)}) +  \frac{C^3 C' \bar C^2}{\tilde C^2 }\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}},\label{eq:g3}
-    \end{align}
-    where the third inequality follows from the same procedure to derive~\eqref{eq:g1} and \eqref{eq:g2}, and the last inequality follows from the assumption $\Delta_{\min}^2 \geq \tilde C p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
-
-    Choose the $\tilde C$ such that
-    \begin{equation}\label{eq:tilde_c1}
-        3 \of{  C^4 \frac{\bar C}{\tilde C^3}  + \frac{C'}{\tilde C^2} + \frac{C^2 \bar C^2}{\tilde C} + \frac{C^3 C' \bar C^2}{\tilde C^2 }} \leq \frac{1}{512}.
-    \end{equation}
-    Then,  we finish the proof of inequality~\eqref{eq:gib} by plugging the inequalities~\eqref{eq:g1}, \eqref{eq:g2}, and \eqref{eq:g3} into the upper bound~\eqref{eq:gib_decomp}.
-
-    \item Upper bound for $H_{ib}^{(t)}$, i.e., the inequality~\eqref{eq:hib}. By definition of $H_{ib}$, we rearrange terms and obtain
-    \begin{align}
-        H_{ib} &= \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 + \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
-        & =  \onormSize{}{ [ \mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 \\
-        & \quad \quad \quad \quad \quad + \of{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }  } \\
-        & \quad \quad \quad \quad \quad  - \of{\onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mY \mV^{(t)} ]^s } -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }   } \\
-        & = H_1 + H_2 + H_3,
-    \end{align}
-    where 
-    \begin{align}
-        H_1 &= \onormSize{}{ [ \mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 - \onormSize{}{ [ \mW_{:b}^T \mX \mV^{(t)}]^s -  [ \mW_{:b}^T \mY \mV^{(t)}]^s}^2, \\
-        H_2 &=  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }^2  , \\
-        H_3 & = 2 \ang{  [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s,  [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }.
-    \end{align}
-    For $H_1$, we have 
-    \begin{align}
-        |H_1| \leq  \frac{4 \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 }{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2}
-         \leq  \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} 
-        \leq \tilde C^{-2} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2\label{eq:h1},
-    \end{align}
-    following the derivation of $G_2$ in inequality~\eqref{eq:g2} and the assumption that $\Delta_{\min}^2 \geq \tilde C p^{-K/2} \log p$.
-
-    For $H_2$, by the inequality~\eqref{eq:g3xv_wxv}, we have 
-    \begin{equation}\label{eq:h2}
-        |H_2| \lesssim 2\max_{a \in [r]}\onormSize{}{[\mW_{:a}^T\mX \mV^{(t)}]^s - [\mW_{:a}^T\mX \mV]^s }^2 \lesssim \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} \leq C \frac{\bar C^2}{ \tilde C^2} \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
-    \end{equation}
-    where the last inequality follows from the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
-
-      For $H_3$,  by Cauchy-Schwartz inequality, we have 
-    \begin{equation}\label{eq:h3}
-         |H_3| \lesssim \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \leq 2 \tilde C^{-1} \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,
-    \end{equation}
-    following the inequalities~\eqref{eq:g3xv_wxv} and \eqref{eq:h1}.
-
-    Choose $\tilde C$ such that 
-    \begin{equation}\label{eq:tilde_c2}
-        \tilde C^{-2} + C \frac{\bar C^2}{ \tilde C^2} + \tilde C^{-1} \leq \frac{1}{4}.
-    \end{equation}
-     Therefore, we finish the proof of inequality~\eqref{eq:hib} combining inequalities~\eqref{eq:h1}, \eqref{eq:h2}, and \eqref{eq:h3}.
-
-\end{enumerate}
-
-\end{proof}
-
-     \begin{lem}[Relationship between misclustering loss and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement}, we have
-    \begin{equation}\label{eq:inter1}
-        \onormSize{}{\mV - \mV^{(t)}}_{\sigma} \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}}  \frac{r}{\Delta_{\min}^2} L^{(t)},
-    \end{equation}
-    \begin{equation}\label{eq:inter2}
-        \onormSize{}{\mE (\mV - \mV^{(t)})}_{\sigma} \lesssim \sqrt{ \frac{r^{K-1}(p r^{K-1} + p r)}{p^{K-1}}} 
- \frac{r}{\Delta_{\min}^2} L^{(t)},
- \end{equation}
- \begin{equation}\label{eq:inter3}
-     \max_{b \in [r]} \onormSize{}{ [ \mW_{:b}^T \mY \mV ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV  ]^s } \leq C \of{\frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{r^{2K} + pr^{K+1}}{p^{K}} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}},
- \end{equation}
- \begin{equation}\label{eq:inter4}
-     \max_{b \in [r]} \onormSize{}{  [  \mW_{:b}^{(t), T}  \mY \mV]^s - [  \mW_{:b}^{(t), T}  \mY \mV^{(t)}]^s } \leq C  \of{ \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} +  \frac{r L^{(t)}}{\Delta_{\min}}},
-     \end{equation}
-    \begin{equation}\label{eq:inter5}
-         \max_{b \in [r]} \onormSize{}{ [\mW_{:b}^T \mY \mV^{(t)} ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s } \leq C \of{\frac{rL^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}},
-    \end{equation}
-    for some positive universal constant $C$. In addition, the inequality~\eqref{eq:inter4} also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$.
-    \end{lem}
-
-    \begin{proof}[Proof of Lemma~\ref{lem:intermediate}] We follow and use several intermediate conclusions in \citet[Proof of Lemma 5]{han2020exact}. We prove each inequality separately.
-    \begin{enumerate}[wide]
-    \item Inequality~\eqref{eq:inter1}. By \citet[Proof of Lemma 5]{han2020exact}, we have 
-    \begin{equation}
-         \onormSize{}{\mV - \mV^{(t)}}_{\sigma} \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}} r \ell^{(t)}.
-    \end{equation}
-    Then, we complete the proof of inequality~\eqref{eq:inter1} by applying Lemma~\ref{lem:mis} to the above inequality.
-    \item Inequality~\eqref{eq:inter2}. By \citet[Proof of Lemma 5]{han2020exact}, we have 
-    \begin{equation}
-           \onormSize{}{\mE (\mV - \mV^{(t)})}_{\sigma} \lesssim \sqrt{ \frac{r^{K-1}(p r^{K-1} + p r)}{p^{K-1}}} r \ell^{(t)}.
-    \end{equation}
-    Also, we complete the proof of inequality~\eqref{eq:inter1} by applying Lemma~\ref{lem:mis} to the above inequality.
-
-    \item Inequality~\eqref{eq:inter3}. We upper bound the desired quantity by triangle inequality,
-    \begin{equation}
-        \onormSize{}{ [ \mW_{:b}^T \mY \mV ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV  ]^s } \leq I_1 + I_2 + I_3,
-    \end{equation}
-    where 
-    \begin{align}
-        I_1 &=  \onorm{ \frac{ \mW_{:b}^T \mY \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{ \mW_{:b}^{(t), T} \mY \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }}   },\\
-        I_2 & = \onorm{\of{ \frac{1}{\onormSize{}{\mW_{:b}^T \mY \mV}} -  \frac{1}{\onormSize{}{\mW_{:b}^T \mX \mV}}  } \mW_{:b}^T \mY \mV },\\
-        I_3 &= \onorm{\of{ \frac{1}{\onormSize{}{\mW_{:b}^{(t), T} \mY \mV  }} -  \frac{1}{\onormSize{}{\mW_{:b}^{(t), T} \mX \mV  }}  } \mW_{:b}^{(t), T} \mY \mV }.
-    \end{align}
-    Next, we upper bound the quantities $I_1, I_2, I_3$ separately.
-
-    For $I_1$, we further bound $I_1$ by triangle inequality,
-    \begin{equation}
-        I_1 \leq I_{11} + I_{12},
-    \end{equation}
-    where
- \begin{equation}
-        I_{11} = \onorm{ \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }}   }, \quad \text{and} \quad I_{12} = \onorm{ \frac{\mW_{:b}^T \mE \mV}{ \onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mE \mV}{ \onormSize{}{  \mW_{:b}^{(t),T} \mX \mV }} }.
-    \end{equation}
-    We first consider $I_{11}$. Define the confusion matrix $\mD = \mM^T \mTheta^T \mW^{(t)} = \entry{D_{ab}} \in \bbR^{r \times r}$ where 
-    \begin{equation}
-        D_{ab} = \frac{\sum_{i \in [p]}  \theta(i) \ind \offf{ z(i) = a, z^{(t)}(i) = b } }{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }}, \text{  for all } a, b \in [r].
-    \end{equation}
-    By Lemma~\ref{lem:membership},  we have $\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } \gtrsim p/r$. Then, we have
-    \begin{equation}\label{eq:dab_bound}
-        \sum_{a \neq b, a,b \in [r]} D_{ab} \lesssim \frac{r}{p} \sum_{i \colon z^{(t)}(i) \neq z(i)} \theta(i) \lesssim\frac{L^{(t)}}{\Delta_{\min}^2}  \lesssim \frac{1}{\log p}, 
-    \end{equation}
-    and for all $b \in [r]$,
-    \begin{equation}\label{eq:dbb_bound}
-        D_{bb} = \frac{\sum_{i \in [p]}  \theta(i) \ind \offf{ z(i) = z^{(t)}(i) = b } }{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }} \geq \frac{c(\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } - p\ell^{(t)})}{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }} \gtrsim 1 - \frac{1}{\log p},
-    \end{equation}
-    under the inequality \eqref{eq:cond_intial} in Condition~\ref{cond:origin}. By the definition of $\mW,\mW^{(t)},\mV$, we have 
-    \begin{equation}
-        \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} = \off{\mS_{b:}}^s, \quad \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }} = [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}]^s.
-    \end{equation}
-
-    Let $\alpha$ denote the angle between $\mS_{b:}$ and $D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}$. To roughly estimate the range of $\alpha$, we consider the inner product 
-    \begin{align}
-        \ang{\mS_{b:},  D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}}  &= D_{bb} \onorm{\mS_{b:}}^2 + \sum_{a \neq b} D_{ab}\ang{\mS_{b:}, \mS_{a:}}\\
-        &\geq D_{bb} \onorm{\mS_{b:}}^2 -  \sum_{a \neq b, a \in [r]} D_{ab}   \onorm{\mS_{b:}} \max_{a \in [r]}  \onorm{\mS_{a:}} \\
-        & \geq C,
-    \end{align}
-    where $C$ is a positive constant, and the last inequality holds when $p$ is large enough following the constraint of $\onorm{\mS_{b:}}$ in parameter space~\eqref{eq:family} and the bounds of $\mD$ in \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}.
-
-    The positive inner product between $\mS_{b:}$ and $D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}$ indicates $\alpha \in [0,  \pi/2)$, and thus $2\sin \frac{\alpha}{2} \leq \sqrt{2} \sin \alpha$.
-Then, by the geometry property of trigonometric function, we have
-    \begin{align}
-       \onormSize{}{ [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}] \sin \alpha} &= \onormSize{}{(\mI_d - \text{Proj}(\mS_{b:})) \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}}\\
-       &\leq \sum_{a \neq b, a \in [r]} D_{ab} \onorm{(\mI_d - \text{Proj}(\mS_{b:}))\mS_{a:}  }\\
-       &= \sum_{a \neq b, a \in [r]} D_{ab} \onorm{ \mS_{a:} \sin (\mS_{b:}, \mS_{a:}) }\\
-       & \leq \sum_{a \neq b, a \in [r]} D_{ab} \onorm{\mS_{a:}} \onorm{ \mS^s_{b:} - \mS^s_{a:}}, \label{eq:i11_sin_num}
-    \end{align}
-    where the first inequality follows from the triangle inequality, and the last inequality follows from Lemma~\ref{lem:norm_diff}.
-Note that with bounds \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}, when $p$ is large enough, we have
-    \begin{equation}\label{eq:i11_sin_dom}
-           \onormSize{}{\mW_{:b}^{(t), T} \mX \mV } = \onormSize{}{ D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}}  \geq D_{bb} \onorm{\mS_{b:}} - \sum_{a \neq b, a \in [r]} D_{ab} \onorm{ \mS_{a:}} \geq  C_1,
-    \end{equation}
-    for some positive constant $C_1$. Notice that $I_{11} = \sqrt{1 - \cos \alpha} = 2 \sin \frac{\alpha}{2}$. Therefore, we obtain
-    \begin{align}
-        I_{11} &\leq \sqrt{2} \sin \alpha \\
-        & = \frac{  \onormSize{}{ [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}] \sin \alpha}}{  \onormSize{}{ D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}} } \\
-        & \leq \frac{ 1}{ C_1}  \sum_{a \neq b, a \in [r]} D_{ab} \onorm{\mS_{a:}} \onorm{ \mS^s_{b:} - \mS^s_{a:}} \\
-        & \lesssim \frac{r}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind\offf{z^{(t)}(i) = b} \onorm{ \mS^s_{b:} - \mS^s_{a:}}\\
-        & \leq \frac{r L^{(t)}}{\Delta_{\min}},\label{eq:i11}
-    \end{align}
-    where the second inequality follows from~\eqref{eq:i11_sin_num} and \eqref{eq:i11_sin_dom}, and the last two inequalities follow by the definition of $D_a$ and $L^{(t)}$, and the constraint of $\onorm{\mS_{b:}}$ in parameter space~\eqref{eq:family}.
-
-    We now consider $I_{12}$. By triangle inequality, we have 
-    \begin{equation}
-        I_{12} \leq \frac{1}{\onormSize{}{  \mW_{:b}^T \mX \mV }}\onormSize{}{ (\mW_{:b}^T - \mW_{:b}^{(t),T}) \mE \mV } + \frac{ \onormSize{}{(\mW_{:b}^T - \mW_{:b}^{(t),T}) \mX \mV} }{\onormSize{}{  \mW_{:b}^T \mX \mV } \onormSize{}{  \mW_{:b}^{(t),T} \mX \mV } }\onormSize{}{ \mW_{:b}^{(t),T} \mE \mV}.
-    \end{equation}
-    By \citet[Proof of Lemma 5]{han2020exact}, we have 
-    \begin{equation}\label{eq:i12_wwe}
-        \onormSize{}{ (\mW_{:b}^T - \mW_{:b}^{(t),T}) \mE \mV }  \lesssim \sqrt{\frac{ r^{2K} + p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}.
-    \end{equation}
-    Notice that 
-    \begin{equation}\label{eq:i12_wwx}
-        \onormSize{}{(\mW_{:b}^T - \mW_{:b}^{(t),T}) \mX \mV} \leq \onormSize{}{\mW_{:b}^T - \mW_{:b}^{(t),T}} \onorm{\mX\mV}_F \lesssim \frac{r^{3/2}L^{(t)}}{\sqrt{p}\Delta_{\min}^2} \onormSize{}{\mS} \onormSize{}{\mTheta \mM}_{\sigma} \lesssim \frac{ \sqrt{r L^{(t)}} }{\Delta_{\min}},
-    \end{equation}
-    where the second inequality follows from~\citet[Inequality (121), Proof of Lemma 5]{han2020exact} and the last inequality follows from Lemma~\ref{lem:singular_thetam} and \eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
-     Note that $\onorm{\mW_{:b}^T \mX \mV} = \onorm{\mS_{b:}} \geq c_3$ and $\onormSize{}{  \mW_{:b}^{(t),T} \mX \mV } \geq C_1$ by inequality~\eqref{eq:i11_sin_dom}. Therefore, we have 
-     \begin{align}
-         I_{12} &\lesssim  \onormSize{}{ (\mW_{:b}^T - \mW_{:b}^{(t),T}) \mE \mV }  +   \onormSize{}{(\mW_{:b}^T - \mW_{:b}^{(t),T}) \mX \mV}   \onormSize{}{ \mW_{:b}^{(t),T} \mE \mV} \\
-         &\lesssim \sqrt{\frac{ r^{2K} + p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} + \frac{ \sqrt{r L^{(t)}} }{\Delta_{\min}}\sqrt{\frac{r^{2K}}{p^K}} \\
-         & \lesssim \sqrt{\frac{ r^{2K} + p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},\label{eq:i12}
-     \end{align}
-     where second inequality follows from the inequalities~\eqref{eq:i12_wwe}, \eqref{eq:i12_wwx}, and \eqref{eq:cond1} in Condition~\ref{cond:origin}.
-
-    Hence, combining inequalities~\eqref{eq:i11} and~\eqref{eq:i12} yields  
-    \begin{equation}\label{eq:i1}
-        I_1 \lesssim \frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{\frac{ r^{2K} + p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}.
-    \end{equation}
-
-        For $I_2$ and $I_3$, recall that $\onorm{\mW_{:b}^T \mX \mV} = \onorm{\mS_{b:}} \geq c_3$ and $\onormSize{}{  \mW_{:b}^{(t),T} \mX \mV } \geq C_1$ by inequality~\eqref{eq:i11_sin_dom}.
-By triangle inequality and \eqref{eq:cond1} in Condition~\ref{cond:origin}, we have
-    \begin{equation}\label{eq:i2}
-         I_2 \leq \frac{ \onormSize{}{ \mW_{:b}^T \mE \mV } }{ \onormSize{}{\mW_{:b}^T \mX \mV}  } \lesssim  \onormSize{}{ \mW_{:b}^T \mE \mV }  \lesssim \frac{r^K}{p^{K/2}},
-    \end{equation}
-    and 
-    \begin{equation}\label{eq:i3}
-        I_3 \leq \frac{ \onormSize{}{ \mW_{:b}^{(t),T} \mE \mV } }{ \onormSize{}{\mW_{:b}^{(t),T} \mX \mV}  }  \lesssim \onormSize{}{ \mW_{:b}^{(t),T} \mE \mV } \lesssim \frac{r^K}{p^{K/2}}.
-     \end{equation}
-       Therefore, combining the inequalities~\eqref{eq:i1}, \eqref{eq:i2}, and \eqref{eq:i3}, we finish the proof of inequality~\eqref{eq:inter3}.
-
-    \item Inequality~\eqref{eq:inter4}.
-Here we only show the proof of inequality~\eqref{eq:inter4} with $\mW_{:b}^{(t)}$.
-The proof also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$, and we omit the repeated procedures.
-
-    We upper bound the desired quantity by triangle inequality
-    \begin{equation}
-         \onormSize{}{  [   \mW_{:b}^{(t), T}  \mY \mV]^s -  [  \mW_{:b}^{(t), T}  \mY \mV^{(t)}]^s }  \leq J_1 + J_2 + J_3,
-    \end{equation}
-    where 
-    \begin{align}
-        J_1 &= \onorm{ \frac{\mW_{:b}^{(t),T} \mY \mV }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mY \mV^{(t)} }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV^{(t)} }}  },\\
-        J_2 & = \onorm{\of{ \frac{1}{\onormSize{}{\mW_{:b}^{(t),T} \mY \mV}} -  \frac{1}{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV}}  } \mW_{:b}^{(t),T} \mY \mV },\\
-        J_3 &= \onorm{\of{ \frac{1}{\onormSize{}{\mW_{:b}^{(t),T} \mY \mV^{(t)}}} -  \frac{1}{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)}} }  } \mW_{:b}^{(t), T} \mY \mV^{(t)} }.
-    \end{align}
-    Next, we upper bound the quantities $J_1, J_2, J_3$ separately.
-
-    For $J_1$, by triangle inequality,
-we have 
-    \begin{equation}
-        J_1 \leq J_{11} + J_{12},
-    \end{equation}
-    where 
-    \begin{equation}
-        J_{11} =  \onorm{ \frac{\mW_{:b}^{(t),T} \mX \mV }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mX \mV^{(t)} }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV^{(t)} }}  } \quad \text{and} \quad \quad J_{12} =  \onorm{ \frac{\mW_{:b}^{(t),T}  \mE \mV}{ \onormSize{}{\mW_{:b}^{(t),T}  \mX \mV} } - \frac{\mW_{:b}^{(t),T}  \mE \mV^{(t)}}{ \onormSize{}{\mW_{:b}^{(t),T}  \mX \mV^{(t)}} }   }.
-    \end{equation}
-
-    We first consider $J_{11}$.
-Define the matrix $\mV^{k} \coloneqq \mW^{\otimes (k-1)} \otimes \mW^{(t), \otimes (K-k)}$ for $k = 2,\ldots, K-1$, and denote $\mV^1 = \mV^{(t)}, \mV^K = \mV$.
-Also, define the quantity
-    \begin{equation}
-        J_{11}^k = \onormSize{}{ [\mW_{:b}^{(t),T} \mX \mV^k ]^s - [\mW_{:b}^{(t),T} \mX \mV^{k+1} ]^s },
-     \end{equation}
-    for $k = 1,\ldots, K-1$.
-Let $\beta_k$ denote the angle between $\mW_{:b}^{(t),T} \mX \mV^k$ and $\mW_{:b}^{(t),T} \mX \mV^{k+1}$.
-With the same idea to prove $I_{11}$ in inequality~\eqref{eq:i11}, we bound $J_{11}^k$ by the trigonometric function of $\beta_k$.
-
-    To roughly estimate the range of $\beta_k$, we consider the inner product between $\mW_{:b}^{(t),T} \mX \mV^k$ and $\mW_{:b}^{(t),T} \mX \mV^{k+1}$.
-Before the specific derivation of the inner product, note that
-    \begin{equation}
-        \mW_{:b}^{(t),T} \mX \mV^k = \mat_1(\tT_k), \quad  \mW_{:b}^{(t),T} \mX \mV^{k+1} = \mat_1(\tT_{k+1}),
-    \end{equation}
-    where 
-    \begin{align}
-        \tT_k &= \tX \times_1  \mW_{:b}^{(t),T} \times_2 \mW^T \times_3 \cdots \times_k \mW^T \times_{k+1} \mW^{(t),T} \times_{k+1} \cdots \times_K \mW^{(t),T}\\
-        \tT_{k+1} &= \tX \times_1  \mW_{:b}^{(t),T} \times_2 \mW^T \times_3 \cdots \times_k \mW^T \times_{k+1} \mW^{T} \times_{k+1} \cdots \times_K \mW^{(t),T}. 
-    \end{align}
-    Recall the definition of confusion matrix $\mD = \mM^T \mTheta^T \mW^{(t)} = \entry{D_{ab}} \in \bbR^{r \times r}$.
- We have 
-    \begin{align}
-         \ang{\mW_{:b}^{(t),T} \mX \mV^k , \mW_{:b}^{(t),T} \mX \mV^{k+1}} 
-         &  =\ang{ \mat_{k+1}(\tT_k), \mat_{k+1}(\tT_{k+1}) } \label{eq:j11_tensormat} \\
-  & =  \ang{ \mD^T \mS \mZ^k ,  \mS \mZ^k   }\\
-         & = \sum_{b \in [r]} \of{ D_{bb} \onormSize{}{\mS_{b:}\mZ^k}^2 + \sum_{a \neq b, a \in [r]} D_{ab} \ang{\mS_{a:} \mZ^k, \mS_{b:} \mZ^k}} \\
-         & \gtrsim (1 - \log p^{-1}) \min_{a \in [r]}\onormSize{}{\mS_{a:}\mZ^k}^2 - \log p^{-1} \max_{a \in [r]} \onormSize{}{\mS_{a:}\mZ^k}^2, \label{eq:j11inner}
-    \end{align}
-    where $\mZ^k = \mD_{:b} \otimes \mI_r^{\otimes (k-1)} \otimes  \mD^{\otimes (K-k-1)}$, the equations follow by the tensor algebra and definitions, and the last inequality follows from the bounds of $\mD$ in \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}.
-
-    Note that 
-    \begin{equation}\label{eq:singular_Du}
-        \onorm{\mD}_{\sigma} \leq \onorm{\mD}_{F} \leq  \sqrt{\sum_{b \in [r]} D_{bb}^2 + (\sum_{a \neq b, a ,b \in [r]} D_{ab} )^2} \lesssim \sqrt{r  + \log^2 p^{-1}} \lesssim 1,
-    \end{equation}
-    where the second inequality follows from inequality~\eqref{eq:dab_bound}, and the fact that for all $b \in [r]$,
-    \begin{equation}
-        D_{bb} \lesssim \frac{r}{p} \sum_{i\colon z(i) = b} \theta(i) \lesssim 1.
-    \end{equation}
-    Also, we have 
-    \begin{equation}\label{eq:singular_Dl}
-        \lambda_r(\mD) \geq \lambda_r(\mW^{(t)}) \lambda_r(\mTheta \mM) \gtrsim 1,
-    \end{equation}
-    following the Lemma~\ref{lem:singular_thetam} and Lemma~\ref{lem:membership}. Then, for all $k \in [K]$, we have 
-    \begin{equation}\label{eq:singular_z}
-       1 \lesssim \onorm{\mD_{:b}} \lambda_r(\mD)^{K-k-1}  \leq \lambda_{r^{K-2}}(\mZ^{k}) \leq \onormSize{}{\mZ^k}_{\sigma} \leq \onorm{\mD_{:b}} \onorm{\mD}_{\sigma}^{K-k-1} \lesssim 1.
-    \end{equation}
-    Thus, we have bounds 
-    \begin{equation}
-         \max_{a \in [r]} \onormSize{}{\mS_{a:} \mZ^k} \leq \max_{a \in [r]} \onorm{\mS_{a:}} \onormSize{}{\mZ^k}_{\sigma} \lesssim 1, \quad \min_{a \in [r]} \onormSize{}{\mS_{a:}\mZ^k}  \geq \min_{a \in [r]} \onorm{\mS_{a:}} \lambda_{r^{K-2}}(\mZ^{k}) \gtrsim 1.
-    \end{equation}
-    Hence, when $p$ is large enough, the inner product~\eqref{eq:j11inner} is positive, which implies $\beta_k \in [0, \pi/2)$ and thus $2 \sin \frac{\beta_k}{2} \leq \sqrt{2} \sin \beta_k$.
-
-    Next, we upper bound the trigonometric function $\sin \beta_k$. Note that
-    \begin{align}
-         \sin \beta_k &= \sin ({\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k} }, {\mD_{:b}^{T} \mS \mI_r^{\otimes k} \otimes \mD^{\otimes K-k-1} } )\\
-         &\leq \sin \beta_{k1} + \sin \beta_{k2},
-    \end{align}
-    where
-    \begin{align}
-         \sin \beta_{k1} &= \sin ({\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k} }, {\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \tilde \mD \otimes \mD^{\otimes K-k-1} } ),\\
-         \sin \beta_{k2} & = \sin ({\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \tilde \mD \otimes \mD^{\otimes K-k-1} } , {\mD_{:b}^{T} \mS \mI_r^{\otimes k} \otimes \mD^{\otimes K-k-1} }   ),
-    \end{align}
-    and $\tilde \mD$ is the normalized confusion matrix with entries $\tilde \mD_{ab} = \frac{\sum_{i \in [p]} \theta(i) \ind\{z^{(t)} = b, z(i) = a\}}{\sum_{i \in [p] }\theta(i) \ind\{z^{(t)} = b\}}$.
-
-    To bound $\sin \beta_{k1}$, recall Definition~\ref{def:stable} that for any cluster assignment $\bar z$ in the $\varepsilon$-neighborhood of true $z$,
-    \begin{equation}
-        \mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T,\quad \mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
-    \end{equation}
-    Note that we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \leq \frac{\bar C}{\tilde C} r \log^{-1}(p) $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.
-Then,  with the locally linear stability assumption, the $\mtheta$ is $\ell^{(t)}$-locally linearly stable; i.e.,
-    \begin{equation}
-         \sin (\mp(z^{(t)}), \mp_{\mtheta}(z^{(t)})) \lesssim \frac{L^{(t)}}{\Delta_{\min}}. 
-    \end{equation}
-    Note that $\text{diag}(\mp(z^{(t)})) \mD = \text{diag}(\mp_{\mtheta}(z^{(t)})) \tilde \mD$, and $\sin(\ma,\mb) =\min_{c \in \bbR} \frac{ \onormSize{}{\ma - c\mb} }{\onormSize{}{\ma}}$ for vectors $\ma,\mb$ of same dimension.
-Let $c_0 = \argmin_{c \in \bbR} \frac{\onormSize{}{\mp(z^{(t)}) - c\mp_{\mtheta}(z^{(t)})}}{\onormSize{}{\mp(z^{(t)})}}$.
-Then, we have
-    \begin{align}
-       \min_{c \in \bbR} \onormSize{}{ \mD - c \tilde \mD}_F &\leq  \onormSize{}{ \mI_r - c_0 \text{diag}(\mp(z^{(t)})) \text{diag}^{-1}(\mp_{\mtheta}(z^{(t)}))}_F \onormSize{}{\mD}_F\\
-       &\lesssim \frac{ \onormSize{}{ \mp(z^{(t)}) - c_0\mp_{\mtheta}(z^{(t)})} }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  } \\
-       & = \frac{\onormSize{}{ \mp(z^{(t)})} }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1 } \sin (\mp(z^{(t)}), \mp_{\mtheta}(z^{(t)}))\\
-       & \lesssim \frac{L^{(t)}}{\Delta_{\min}},
-    \end{align}
-    where the last inequality follows from Lemma~\ref{lem:membership}, the constraint $\min_{i \in [p]}\theta(i) \geq c>0$, $\onormSize{}{ \mp(z^{(t)})} \lesssim p$ and $\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  \gtrsim p$.
-By the geometry property of trigonometric function, we have
-    \begin{align}\label{eq:b1}
-        \sin \beta_{k1} &= \min_{c \in \bbR} \frac{\onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes (\mD - c \tilde \mD) \otimes \mD^{\otimes K-k - 1} } }{ \onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k}} }\\
-        &\leq \frac{ \onormSize{}{\mD_{:b}^{T} \mS} \onormSize{}{\mD - c_0 \tilde \mD}_{\sigma} \onormSize{}{\mD}_{\sigma}^{K-k-1} }{\onormSize{}{\mD_{:b}^{T} \mS} \lambda_r^{K-k}(\mD) }\\
-        & \lesssim  \onormSize{}{\mD - c_0 \tilde \mD}_{F}\\
-        & \lesssim \frac{L^{(t)}}{\Delta_{\min}},
-    \end{align}
-    where the second inequality follows from the singular property of $\mD$ in \eqref{eq:singular_Du}, \eqref{eq:singular_Dl} and the constraint of $\mS$ in \eqref{eq:family}.
-
-    To bound $\sin \beta_{k2}$, let $\mC = \text{diag}(\{\onormSize{}{\mS_{a:}}\}_{a \in [r]})$. We have 
-    \begin{align}\label{eq:b2}
-        \sin \beta_{k2} &\lesssim \frac{ \onorm{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes (\mI_r - \tilde \mD) \otimes \mD^{\otimes K-k-1} } }{\onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k} \otimes \mD^{\otimes K-k-1} }  }\\
-        & \lesssim \frac{ \onormSize{}{(\mI_r - \tilde \mD^T) \mS \mZ^k}_F }{\onormSize{}{\mD_{:b}^{T} \mS} \lambda_r^{K-k-1}(\mD)} \\
-        & \lesssim \onormSize{ }{(\mI_r - \tilde \mD^T) \mS \mC^{-1}}_F \onormSize{}{\mC \mZ^k}_{\sigma}\\
-        & \lesssim \frac{r}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \onormSize{}{\mS_{b:}^s - \mS_{z(i):}^s} \\
-        & \lesssim \frac{L^{(t)}}{\Delta_{\min}}, 
-    \end{align}
-    where the third inequality follows from the singular property of $\mD$ and the boundedness of $\mS$, and the fourth inequality follows from the definition of $\tilde \mD$, boundedness of $\mS$, the lower bound of $\mtheta$, and the singular property of $\mZ^k$ in inequality~\eqref{eq:singular_z}.
-
-  Combining~\eqref{eq:b1} and~\eqref{eq:b2} yields
-    \begin{equation}
-        \sin \beta_k \leq \sin \beta_{k1} + \sin \beta_{k2} \lesssim \frac{L^{(t)}}{\Delta_{\min}}.
-    \end{equation}
-
-       Finally, by triangle inequality, we obtain
-    \begin{equation}\label{eq:j1}
-        J_{11} \leq \sum_{k = 1}^{K-1} J_{11}^k \lesssim  \sum_{k = 1}^{K-1} \sin \beta_k  \lesssim (K-1)\frac{r L^{(t)}}{\Delta_{\min}}.
-    \end{equation}
-
-    We now consider $J_{12}$. By triangle inequality, we have 
-    \begin{align}
-        J_{12} &\leq \frac{ 1}{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }} \onormSize{}{\mW_{:b}^{(t),T} \mE (\mV-\mV^{(t)})}   +
-        \frac{ \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})} }{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)} }} \onormSize{}{\mW_{:b}^{(t),T} \mE \mV^{(t)} }.
-    \end{align}
-
-     Note that 
-    \begin{equation}\label{eq:j11_wxvt}
-       \onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)}} = \onormSize{}{\mD^T \mS \mZ^1} \geq \lambda_r(\mD) \onorm{\mS} \lambda_{r^{K-2}}(\mZ^1) \gtrsim 1,
-    \end{equation}
-    where the inequality follows from the bounds~\eqref{eq:singular_Dl} and \eqref{eq:singular_z}.
-
-    By \citet[Proof of Lemma 5]{han2020exact}, we have 
-    \begin{equation}\label{eq:j12_wev}
-        \onormSize{}{\mW_{:b}^{(t),T} \mE (\mV-\mV^{(t)})} \lesssim \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{(K-1)\sqrt{L^{(t)}}}{\Delta_{\min}}.
-    \end{equation}
-    Notice that 
-    \begin{align}\label{eq:step}
-        \onormSize{}{\mX (\mV^k-\mV^{k+1})}_F &\leq \onormSize{}{(\mI - \mD^T) \mS (\mI_r^{\otimes (k-1)} \otimes  \mD^{\otimes (K-k-1)})}_F\notag \\
-        &\leq \onormSize{}{(\mW^T - \mW^{(t),T}) \mTheta \mM}_F \onorm{\mS}_F \onorm{ \mD}_{\sigma}^{K-k-1}\notag \\
-        & \lesssim \onormSize{}{ \mW^T - \mW^{(t),T} } \onorm{\mTheta \mM}_{\sigma} \notag\\
-        & \lesssim \frac{ \sqrt{r L^{(t)}}}{\Delta_{\min}},
-    \end{align}
-    where the first inequality follows from the tensor algebra in 
-    inequality~\eqref{eq:j11inner}, the second inequality follows from the fact  that $\mI = \mW^T \mTheta \mM$, and the last inequality follows from \citet[Proof of Lemma 5]{han2020exact}.
-It follows from~\eqref{eq:step} and Lemma~\ref{lem:membership} that
-    \begin{equation}
-           \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})}\leq \onormSize{}{\mW_{:b}^{(t),T} }  \sum_{k = 1}^{K-1} \onormSize{}{\mX (\mV^k-\mV^{k+1})}_F \lesssim \frac{ \sqrt{r L^{(t)}}}{\sqrt{p} \Delta_{\min}}.\label{eq:j12_wxv}
-    \end{equation}
-    Note that $\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }$ and $\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)} }$ are lower bounded by inequalities~\eqref{eq:i11_sin_dom} and \eqref{eq:j11_wxvt}, respectively.
-We have
-    \begin{align}
-        J_{12} &\lesssim \onormSize{}{\mW_{:b}^{(t),T} \mE (\mV-\mV^{(t)})} + \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})}\onormSize{}{\mW_{:b}^{(t),T} \mE \mV^{(t)} }\\
-        &\lesssim \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} + \frac{ \sqrt{r L^{(t)}}}{\sqrt{p} \Delta_{\min}} \sqrt{\frac{r^{2K}}{p^K}} \\
-        & \lesssim  \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
-    \end{align}
-    where the second inequality follows from inequalities~\eqref{eq:j12_wev}, \eqref{eq:j12_wxv}, and the inequality \eqref{eq:cond1} in Condition~\ref{cond:origin}.
-
-    For $J_2$ and $J_3$, recall that $\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }$ and $\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)} }$ are lower bounded by inequalities~\eqref{eq:i11_sin_dom} and \eqref{eq:j11_wxvt}, respectively.
-By triangle inequality and inequality \eqref{eq:cond1} in Condition~\ref{cond:origin}, we have
-    \begin{equation}\label{eq:j2}
-        J_2 \leq \frac{ \onormSize{}{\mW_{:b}^{(t),T} \mE \mV}  }{  \onormSize{}{\mW_{:b}^{(t),T} \mX \mV} } \lesssim \onormSize{}{\mW_{:b}^{(t),T} \mE \mV} \lesssim \frac{r^K}{p^{K/2}},
-    \end{equation}
-    and 
-    \begin{equation}\label{eq:j3}
-        J_3 \leq \frac{ \onormSize{}{\mW_{:b}^{(t),T} \mE \mV^{(t)}}  }{  \onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)}} } \lesssim \onormSize{}{\mW_{:b}^{(t),T} \mE \mV} \lesssim \frac{r^K}{p^{K/2}}.
-    \end{equation}
-
-    Therefore, combining the inequalities~\eqref{eq:j1}, \eqref{eq:j2}, and \eqref{eq:j3}, we finish the proof of inequality~\eqref{eq:inter4}.
-
-    \item Inequality~\eqref{eq:inter5}. By triangle inequality, we upper bound the desired quantity 
-    \begin{align}
-        \onormSize{}{ [ \mW_{:b}^T \mY \mV^{(t)} ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s } &\leq \onormSize{}{[ \mW_{:b}^T \mY \mV^{(t)} ]^s - [ \mW_{:b}^T \mY \mV ]^s }  + \onormSize{}{[\mW_{:b}^T \mY \mV ]^s  - [ \mW_{:b}^{(t),T} \mY \mV ]^s } \\
-        & \quad \quad \quad\quad\quad+ \onormSize{}{[\mW_{:b}^{(t),T} \mY \mV ]^s  -  [\mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s }\\
-        &\lesssim \frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
-    \end{align}
-    following the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4}. Therefore, we finish the proof of inequality~\eqref{eq:inter5}.
-        \end{enumerate}
-
-    \end{proof}
+%\fixme{Miaoyan}{add discussion as in my JMLR binary tensor paper (stat-comp dominating terms)}
 

### 2022-02-18 tensor_clustering/tensor_clustering_aistats_camera/aistats_camera.tex
aistats

@@
 Suppose we have an order-$K$ data tensor $\tY \in \bbR^{p \times \cdots \times p}$.
-For ease of notation, we focus on symmetric tensors in this section; see Supplement~\ref{sec:exten} for the extension to general asymmetric tensors.
+For ease of notation, we focus on symmetric tensors in this section.
 Assume there exist $r \geq 2$  disjoint communities among the $p$ nodes.
@@
 %The goal of clustering is to estimate $z$ from a single noisy tensor $\tY$. 
-We are particularly interested in high-dimensional regime where $p$ grows whereas $r=\tO(1)$. See Supplement~\ref{sec:examples} for motivated applications of dTBM.
+We are particularly interested in high-dimensional regime where $p$ grows whereas $r=\tO(1)$.
+The extension to general asymmetrical dTBM is obtained via replacing $(\mM, \mTheta)$ in~\eqref{eq:model_tensor} by mode-specific parameters $(\mM_k, \mTheta_k)$ for every mode $k\in[K]$.
+Here, we give two special cases of dTBM.
+
+\begin{example}[Gaussian TBM]
+Let $\theta(i)= 1$ for all $i \in [p]$ and $\tE$ be a noise tensor with i.i.d.\ $N(0,\sigma^2)$ entries.
+Our dTBM reduces to a non-degree Gaussian TBM~\citep{wang2019multiway, han2020exact}, which is widely used in previous clustering algorithms~\citep{wang2019multiway,chi2020provable}.
+%The theoretical results in TBM serve as benchmarks for dTBM. 
+\end{example}
+
+\begin{example}[Binary dTBM]
+Consider a $K$-uniform hypergraph $H = (V,E)$, where $V = [p]$ collects the nodes with $r$ disjoint communities and $E$ collects all the $K$-way hyperedges.
+Let $\tY \in \{0,1\}^{p \times \cdots \times p}$ denote the adjacency tensor, where the entries encode the presence or absence of hyperdeges among $p$ nodes.
+Specifically, let $\tY(i_1,...,i_K) = 1$ if  $(i_1,...,i_K) \in E$, otherwise, $\tY(i_1,...,i_K) = 0$, for all $(i_1,\ldots,i_K) \in [p]^K$.
+The equation~\eqref{eq:model_tensor} models $\mathbb{E}\tY$ with degree heterogeneity and subgaussianity parameter $\sigma^2 = 1/4$.
+\end{example}
 
@@
 In this section, we study the statistical and computational limits of dTBM. We propose signal-to-noise ratio~(SNR),
+\vspace{-.2cm}
 \begin{equation}\label{eq:gamma}
@@
 Figure~\ref{fig:phase_axis} summarizes our results of critical SNRs when $K \geq 2$.
-In the weak SNR region $\gamma < -(K-1)$, no statistical estimator succeeds in degree-corrected higher-order clustering.
+% In the weak SNR region $\gamma < -(K-1)$, no statistical estimator succeeds in degree-corrected higher-order clustering.
 In the strong SNR region $\gamma  > -K/2$, our proposed algorithm precisely recovers the clustering in polynomial time.
@@
 Hence, the statistical-to-computational gap emerges only for higher-order tensors with $K \geq 3$, which reveals the intrinsic distinctions among (vector) one-dimensional clustering, (matrix) biclustering, and (tensor) higher-order clustering.
-%Besides, we find that the extra complexity from $\mtheta$ does not render the estimation of $z$ qualitatively harder; see comparison of our phase transition with non-degree TBM~\citep{han2020exact}. 
+Besides, we find that the extra complexity from $\mtheta$ does not render the estimation of $z$ qualitatively harder; see comparison of our phase transition with non-degree TBM~\citep{han2020exact}. 
 \end{rmk}
 
-\begin{rmk}[Comparison with non-degree models]
-We compare our results to non-degree tensor models.
+% \begin{rmk}[Comparison with non-degree models]
+% We compare our results to non-degree tensor models.
 The allowance of degree heterogeneity $\mtheta$ makes the model more flexible, but it incurs extra statistical and computational complexity.
 Fortunately, we find that the extra complexity does not render the estimation of $z$ qualitatively harder; see comparison of our phase transition with non-degree TBM~\citep{han2020exact}.
-\end{rmk}
+% \end{rmk}
 
@@
 Both steps are critical to obtain a satisfactory algorithm output.
-We use the symmetric tensor with i.i.d.\ sub-Gaussian noise to describe the algorithm and develop the theoretical guarantees.
-Extensions for asymmetric tensor and Bernoulli observation and practical issues are in Supplement~\ref{sec:exten}.
+%We use the symmetric tensor with i.i.d.\ sub-Gaussian noise to describe the algorithm and develop the theoretical guarantees.
 
@@
 \centering
-\includegraphics[width=.9\columnwidth]{alg_demo.pdf}
+\includegraphics[width=.8\columnwidth]{alg_demo.pdf}
 \caption{ \small Illustration of our global-to-local algorithm.}\label{fig:demo}
 \end{figure}
+
+\vspace{-.3cm}
 
@@
 
+\vspace{-.3cm}
 \begin{rmk}[Comparison to previous results] For fixed SNR, our initialization error rate with $K=2$ agrees with the initialization error rate $\tO(p^{-1})$ in matrix models~\citep{gao2018community}.
@@
 Note that CER between $(\hat z, z)$ is equivalent to misclustering error $\ell(\hat z, z)$ up to constant multiplications~\citep{meilua2012local}, and lower CER indicates better performance.
-Extra simulation results and real data analysis for human brain connectome data are in Supplement~\ref{sec:extra_exp}.
 
@@
 
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=.95\columnwidth]{phase_anno.pdf}
+    \caption{ \footnotesize SNR phase transitions for clustering in dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.
+    }
+    \label{fig:phase}
+    \vspace{-.5cm}
+\end{figure}
+
+\begin{figure}[htb]
+    \centering
+     \includegraphics[width=.95\columnwidth]{ini_re_anno.pdf}
+    \caption{ \footnotesize CER versus signal exponent $(\gamma)$ for initialization only and for combined algorithm.
+We set $p = \{80, 100\}, r = 5, \gamma \in [-2.1, -1.4]$ under (a) Gaussian models and (b) Bernoulli models.
+}
+    \label{fig:ini_re}
+    \vspace{-.5cm}
+\end{figure}
+
 The first experiment verifies  statistical-computational gap described in Section~\ref{sec:limits}.
@@
 Figure~\ref{fig:phase} confirms the existence of the statistical-computational gap in our Theorems~\ref{thm:stats} and~\ref{thm:comp}.
-\vspace{-.2cm}
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=.95\columnwidth]{phase_anno.pdf}
-    \caption{ \footnotesize SNR phase transitions for clustering in dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.
-    }
-    \label{fig:phase}
-    \vspace{-.5cm}
-\end{figure}
-
-\vspace{-.3cm}
-\begin{figure}[htp!]
-    \centering
-     \includegraphics[width=.95\columnwidth]{ini_re_anno.pdf}
-    \caption{ \footnotesize CER versus signal exponent $(\gamma)$ for initialization only and for combined algorithm.
-We set $p = \{80, 100\}, r = 5, \gamma \in [-2.1, -1.4]$ under (a) Gaussian models and (b) Bernoulli models.
-}
-    \label{fig:ini_re}
-    \vspace{-.4cm}
-\end{figure}
 
@@
 The only exception is the slightly better performance of \textbf{\small HLloyd} over \textbf{\small HOSVD+} under Gaussian model.
-However, we find the advantage of \textbf{\small HLloyd} disappears with higher degree heterogeneity; see Supplement~\ref{sec:extrasim}.
+However, we find the advantage of \textbf{\small HLloyd} disappears with higher degree heterogeneity; see Supplement~\ref{sec:extra_exp}.
 The experiment demonstrates the benefits of addressing heterogeneity in higher-order clustering tasks.
-\begin{figure}[h!]
+
+% The only exception in Figure~\ref{fig:comp_gamma} is the slightly better performance of \textbf{\small HLloyd} over \textbf{\small HOSVD+} under Gaussian model.
+However, we find the advantage of \textbf{\small HLloyd} disappears with higher degree heterogeneity.
+
+The last experiment investigates the effects of degree heterogeneity to clustering performance.
+We use the same setting as in the first experiment in the Section~\ref{subsec:comp}, except that we fix the signal exponent $\gamma = -1.2$ and generate the degree heterogeneity $\mtheta$ from Pareto distribution prior to normalization.
+The density function of Pareto distribution is $f(x|a,b) = a b^a x^{-(a+1)} \ind\{ x \geq b \}$, where $a$ is called \emph{shape} parameter.
+We vary the shape parameter $a \in [3,6]$ and choose $b$ such that $\bbE X = a(a - 1)^{-1}b = 1$ for $X$ following Pareto$(a,b)$.
+Note that a smaller $a$ leads to a larger variance in $\mtheta$ and hence a larger degree heterogeneity.
+Figure~\ref{fig:comp_theta} demonstrates the stability of degree-corrected algorithms (\textbf{\small dTBM}, \textbf{\small SCORE}, \textbf{\small HOSVD+}) over the entire range of degree heterogeneity under consideration.
+In contrast, non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) show poor performance with large heterogeneity, especially in Bernoulli cases.
+This experiment, again, highlights the benefit of addressing degree heterogeneity in clustering.
+
+\begin{figure}[htb]
  \vspace{-.3cm}
@@
 \end{figure}
-% The only exception in Figure~\ref{fig:comp_gamma} is the slightly better performance of \textbf{\small HLloyd} over \textbf{\small HOSVD+} under Gaussian model.
-However, we find the advantage of \textbf{\small HLloyd} disappears with higher degree heterogeneity.
-
-The last experiment investigates the effects of degree heterogeneity to clustering performance.
-We use the same setting as in the first experiment in the Section~\ref{subsec:comp}, except that we fix the signal exponent $\gamma = -1.2$ and now generate the degree heterogeneity $\mtheta$ from Pareto distribution prior to normalization.
-The density function of Pareto distribution is $f(x|a,b) = a b^a x^{-(a+1)} \ind\{ x \geq b \}$, where $a$ is called \emph{shape} parameter.
-We vary the shape parameter $a \in [3,6]$ and choose $b$ such that $\bbE X = a(a - 1)^{-1}b = 1$ for $X$ following Pareto$(a,b)$.
-Note that a smaller $a$ leads to a larger variance in $\mtheta$ and hence a larger degree heterogeneity.
-Figure~\ref{fig:comp_theta} demonstrates the stability of degree-corrected algorithms (\textbf{\small dTBM}, \textbf{\small SCORE}, \textbf{\small HOSVD+}) over the entire range of degree heterogeneity under consideration.
-In contrast, non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) show poor performance with large heterogeneity, especially in Bernoulli cases.
-This experiment, again, highlights the benefit of addressing degree heterogeneity in higher-order clustering.
-\begin{figure}[h!]
+
+\begin{figure}[htb]
 \vspace{-.2cm}
@@
 \end{figure}
+
 %\section{REAL DATA}\label{sec:real}
@@
     }
-    \vspace{-.2cm}
+    \vspace{-.3cm}
     \caption{ \footnotesize Clustering errors (measured by CER) for various methods in the analysis of Peru Legislation dataset.}
@@
 \setlength{\footskip}{10pt}
-
-\section{MOTIVATED EXAMPLES}\label{sec:examples}
-
- Here, we provide four applications to illustrate the practical necessity of dTBM.
-
-\vspace{0.2cm}
-\paragraph{Tensor block model.} Consider the model~\eqref{eq:model_tensor}.
-Let $\theta(i)=1$ for all $ i \in [p]$.
-The model~\eqref{eq:model_tensor} reduces to the tensor block model, which is widely used in previous clustering algorithms~\citep{wang2019multiway,chi2020provable,han2020exact}.
-The theoretical results in TBM serve as benchmarks for dTBM.
-
-\vspace{0.2cm}
-
-\paragraph{Community detection in hypergraphs.} Hypergraph network is a powerful tool to represent the complex entity relations with higher-order interactions~\citep{ke2019community}.
-A typical undirected hypergraph is denoted as $H = (V,E)$, where $V = [p]$ is the set of nodes and $E$ is the set of undirected hyperedges.
-Each hyperedge in $E$ is a subset of $V$, and we call the hyperedge an order-$K$ edge if the corresponding subset involves $K$ nodes.
-We call $H$ a $K$-uniform hypergraph if $E$ only contains order-$K$ edges.
-
-It is natural to represent the $K$-uniform hypergraph using a binary order-$K$ adjacency tensor.
-Let $\tY \in \{0,1\}^{p \times \cdots \times p}$ denote the adjacency tensor, where the entries encode the presence or absence of order-$K$ edges among $p$ nodes.
-Specifically, for all $(i_1,\ldots,i_K) \in [p]^K$, we have
-\begin{equation}
-    \tY(i_1,...,i_K) =  \begin{cases}
-    1  & \text{if }  (i_1,...,i_K) \in E,\\
-    0 & \text{if }  (i_1,...,i_K) \notin E.
-    \end{cases}
-\end{equation}
-
-Assume there exist $r$ disjoint communities among $p$ nodes, and the connection probabilities depend on the community assignments and node-specific parameters. 
-Then, the equation~\eqref{eq:model_tensor} models $\bbE \tY$ with unknown degree heterogeneity $\mtheta$ and sub-Gaussianity parameter $\sigma^2 = 1/4$.
-
-\vspace{0.2cm}
-\paragraph{Multi-layer weighted network.} 
-Multi-layer weighted network data consists of multiple networks over the same set of nodes.
-One representative example is the brain connectome data~\citep{zhang2019tensor}.
-The multi-layer weighted network $\tY$ has dimension of $p \times p \times L$, where $p$ denotes the number of brain regions of interest, and $L$ denotes the number of layers (networks).
-Each of the $L$ networks describes one aspect of the brain connectivity, such as functional connectivity or structural connectivity.
-The resulting tensor $\tY$ consists of a mixture of slices with various data types.
-
-Assume there exist $r$ disjoint communities among $p$ nodes and $r_l$ disjoint communities among the $L$ layers.
-The multi-layer network community detection is modeled by the generalized asymmetric dTBM model~\eqref{eq:model_tensor}
-\begin{equation}
-    \bbE \tY = \tS \times_1 \mTheta \mM \times_2 \mTheta \mM \times_3 \mTheta_l \mM_l, 
-\end{equation}
-where $(\mtheta \in \bbR^p, \mM \in \{0,1\}^{p \times r})$ and $ (\mtheta_l \in \bbR^L, \mM_l \in \{0,1\}^{L \times r_l})$ are the degree heterogeneity and membership matrices corresponding to the community structure for $p$ nodes and $L$ layers, respectively.
-\vspace{0.2cm}
-
-\paragraph{Gaussian higher-order clustering.} Datasets in various fields such as medical image, genetics, and computer science are formulated as Gaussian tensors.
-One typical example is the multi-tissue gene expression dataset, which records the different gene expression in different individuals and different tissues.
-The dataset, denoted as $\tY \in \bbR^{p \times n \times t}$, consists of the expression data for $p$ genes of $n$ individuals in $t$ tissues.
-
-Assume there exist $r_1, r_2, r_3$ disjoint clusters for $p$ genes, $n$ individuals, and $t$ tissues, respectively. We apply the generalized asymmetric dTBM model~\eqref{eq:model_tensor} 
-\begin{equation}
-    \bbE\tY = \tS \times_1 \mTheta_1 \mM_1 \times_2 \mTheta_2 \mM_2 \times_3 \mTheta_3 \mM_3, 
-\end{equation}
-where $\{(\mtheta_k, \mM_k)\}_{k=1}^3$ represents the heterogeneity and membership for genes, individuals, and tissues. \\
 
@@
 
-\begin{table}[h]
+\begin{table}[hbt]
     \centering
@@
 
-\section{EXTENSION AND PRACTICAL ISSUES}\label{sec:exten}
-
-Here, we provide the extension of our algorithm for Bernoulli observations and asymmetric tensors and other practical issues including computational complexity and hyper-parameter selection.
-
- {\bf Extension for Bernoulli Observations.} The main difficulty to establish the statistical guarantee for Bernoulli observations lies in the initialization Sub-algorithm~\hyperref[alg:main]{1}.
-Theorem~\ref{thm:refinement} still holds for Bernoulli observations once the initialization accuracy satisfies the upper bound~\eqref{eq:ini} in Theorem~\ref{thm:initial}.
-
-We now provide a high-level explanation for the technical difficulty when applying Theorem~\ref{thm:initial} to Bernoulli observations.
-The derivation of Theorem~\ref{thm:initial} relies on the upper bound of the estimation error for the mean tensor in Lemma~\ref{lem:two-step_esterror}; i.e., with high probability
-\begin{equation}\label{eq:prop1}
-    \onormSize{}{\hat \tX - \tX}_F^2 \lesssim p^{K/2},
-\end{equation}
-where $\tX = \bbE\tY$ and $\hat \tX$ is defined in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}.
-Unfortunately, the inequality~\eqref{eq:prop1} holds only for i.i.d.\ sub-Gaussian observations while Bernoulli observations are generally not identically distributed.
-
-One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square} of Bernoulli tensor $\tY$.
-Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{\floor{p^{K/2}} \times \ceil{p^{K/2}} }$ denote the nearly square unfolded Bernoulli tensor.
-Define a new estimate
-\begin{equation}\label{eq:matrixsvd}
-    \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX) -  \mat_{sq}(\tY)}_F^2.
-\end{equation}
-The optimization~\eqref{eq:matrixsvd} is simply a matrix SVD problem. Following Lemma 7 in \cite{gao2018community}, with high probability, the new estimate satisfies
-\begin{equation}
-    \onormSize{}{\hat \tX' - \tX}_F^2 \lesssim p^{\ceil{K/2}}.
-\end{equation} 
-Replacing the estimate $\hat \tX$ by $\hat \tX'$ in Theorem~\ref{thm:initial}, the high probability upper bound for Bernoulli initialization is
-\begin{equation}\label{eq:ini_b}
-    \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}. 
-\end{equation}
-The Bernoulli bound~\eqref{eq:ini_b} is relatively looser than Gaussian bound~\eqref{eq:ini}, especially when $K$ is small.
-Nevertheless, our bound~\eqref{eq:ini_b} is already tighter than the pervious work~\citep{ke2019community}.
-The investigation of the gap between upper bound $p^{- \floor{K/2}}$ and the lower bound $p^{- K/2}$ for Bernoulli tensors will be left as future work.
-
-{\bf Extension for General dTBMs.} Our two-stage algorithm is able to be extended for the general (asymmetric) dTBMs.
-Specifically, in the Sub-algorithm~\hyperref[alg:main]{1}, we make the following changes: (1)
-Replace the matrcization $\mat(\tY)$ by $\mat_k(\tY)$; (2) Repeat the Steps 1-5 with mode-specified cluster number $r_k$; (3) Obtain the collection initialization $ \{z^{(0)}_k\}_{k=1}^K$.
-In the Sub-algorithm~\hyperref[alg:main]{2}, we make the following changes: (1) Take the collection $\{z^{(0)}_k\}_{k=1}^K$ as input, and update the block tensor $\tS^{(t)}$ with the collection $\{ z_k^{(t)}\}_{k=1}^K$, $
-        \tS^{(t)}(a_1,\ldots, a_K) =  \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}_k(i_k) = a_k, k \in [K]\}$;
-(2) Calculate reduced tensor $\tY^{\text{d}}_k$ for each mode via
-\begin{align}
-    &\tY^{\text{d}}_k(a_1,\ldots,a_{k-1}, i ,a_{k+1},\ldots,a_K) 
-    = \text{Ave}\{\tY(i_1,\ldots,i_{k-1},i,i_{k+1},\ldots,i_K): z^{(t)}(i_j) = a_j, j \neq k \}; 
-\end{align}
-
-(3) Repeat Step 8-10 with $\mat_k(\cdot), \tY_k^{\text{d}}$ for each $k \in [K]$ and obtain the collection $\{z^{(T)}_k\}_{k=1}^K$.
-
-Correspondingly, Theorems~\ref{thm:initial} and \ref{thm:refinement} still hold with $\ell(z^{(0)}_k, z_k)$ and $\ell(z^{(t+1)}_k, z_k)$  for all $k \in [K]$.
-The detailed model extension for general asymmetric dTBMS can be found in Appendix.
-
-{\bf Computational Complexity.} Our two-stage algorithm has a computational cost polynomial in tensor dimension $p$.
-Specifically, the complexity of Sub-algorithm~\hyperref[alg:main]{1} is $\tO(K p^{K+1} + Krp^K)$, where the first term is contributed by the double projection and the calculation of $\hat \tX$, and the second term comes from normalization and the $k$-means.
-The cost of each update in Sub-algorithm~\hyperref[alg:main]{2} is $\tO(p^K + pr^K)$, where $p^K$ comes from the calculation of $\tS^{(t)}$ and $\tY^{\text{d}}$, and $pr^K$ comes from the normalization of $\tY^{\text{d}}$, the calculation of $\tS^{(t)}$, and the cluster assignment update in Step 10.
-
-{\bf Hyper-parameter Selection.} In our theoretical analysis, we have assumed the true cluster number $r$ is given to our algorithm.
-In practice, the cluster number $r$ is often unknown, and we now propose a method to choose $r$ from data.
-We impose the Bayesian information criterion (BIC) and choose the cluster number that minimizes BIC; i.e., under the symmetric Gaussian dTBM~\eqref{eq:model_tensor},
-\begin{align}\label{eq:BIC}
- \hat r &= \argmin_{r \in \mathbb{Z}_+ } \of{p^K \log(\onormSize{}{\hat \tX - \tY}_F^2)  + p_e(r) K \log p},\\
- \text{where}  \quad \hat \tX &= \hat \tS(r) \times_1 \hat \mTheta(r) \hat \mM(r) \times_2 \cdots \times_K \hat \mTheta(r) \hat \mM(r),
-\end{align}
-where the triplet $(\hat z(r), \hat \mS(r),\hat \mtheta(r))$ are estimated parameters with cluster number $r$, and $p_e (r)= r^K + p(\log r + 1) - r$ is the effective number of parameters.
-Note that we have added the argument $(r)$ to related quantities as functions of $r$.
-In particular, the estimate $\hat \mtheta(r)$ in~\eqref{eq:BIC} is obtained by first calculating the reduced tensor $\hat \tY^{\text{d}}$ with $\hat z(r)$, and then normalizing the row norms $\onormSize{}{\hat \mY_{i:}^{\text{d}}}$ to 1 in each cluster; i.e.,
-\begin{equation}
- \hat \mtheta(r)=(\hat \theta(1,r),\ldots,\hat \theta(p,r))^T,\quad\text{where}\quad   \hat \theta(i,r) = \frac{\onormSize{}{\hat \mY^{\text{d}}(r)_{i:}}}{\sum_{j: \hat z(j,r) =  \hat z(i,r)} \onormSize{}{\hat \mY^{\text{d}}(r)_{j:}}},
-\end{equation}
-where $\hat \mY^{\text{d}}(r) = \mat(\hat \tY^{\text{d}}(r))$, $\hat \tY^{\text{d}}(r)(i,a_2,\ldots,a_K) =  \text{Ave}\{\tY(i,i_2,\ldots,i_K): \hat z(i_k, r) = a_k, k \neq 1 \}$, and $\hat z(i,r)$ denotes the community label for the $i$-th node with given cluster number $r$.
-We evaluate the performance of the BIC criterion in Section~\ref{subsec:sim_select}.
-
 \section{ADDITIONAL EXPERIMENTS}\label{sec:extra_exp}
-
-Here, we provide the additional experiments including the numerical verification of hyper-parameter selection criterion, extra simulation results for comparison, and real data analysis for human brain connectome data.
-
-\subsection{Verification of Hyper-parameter Selection}\label{subsec:sim_select}
-
-This experiment evaluates the empirical performance of the BIC criterion to select unknown cluster number in Section~\ref{sec:exten}.
-We generate the data from an order-3 Gaussian model with $p = \{50,80\}$, $r = \{2,4\}$, and noise level $\sigma^2 \in \{ 0.25,1\}$.
-Table~\ref{tab:select} shows that our BIC criterion well chooses the true $r$ under most settings.
-Note that the BIC slightly underestimates the true cluster number $(r = 4)$ with smaller dimension and higher noise $(p = 50, \sigma=1)$, and the accuracy immediately increases with larger dimension $p = 80$.
-The improvement follows from the fact that a larger dimension $p$ indicates a larger sample size in the tensor block model.
-Therefore, we conclude that BIC criterion is a reasonable way to tune the cluster number.
-
-\begin{table}[hbt]
-\centering
-    \begin{tabularx}{\columnwidth}{c *{8}{Y}}
-    \toprule
-    Settings & \multicolumn{2}{c}{\small  $p = 50, \sigma^2 = 0.25$} & \multicolumn{2}{c}{\small $p = 50, \sigma^2 = 1$} & \multicolumn{2}{c}{\small  $p = 80, \sigma^2 = 0.25$} & \multicolumn{2}{c}{\small  $p = 80, \sigma^2 = 1$}\\
-    \cmidrule(lr){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9}
-         True cluster number $r$ &  \small 2 & \small 4 &  \small2  &  \small 4 &  \small2  & \small 4 &  \small 2 & \small 4  \\
-         \midrule
-         Estimated cluster number $\hat r$ &   \small 2(0)  &  \small 3.9(0.2)&  \small 2(0)    &   \small 3.1(0.5) &  \small 2(0)    &  \small 4(0)   &  \small 2(0)    &  \small 3.9(0.3)   \\
-     \bottomrule
-    \end{tabularx}
-    \caption{ \small Estimated cluster number given by BIC criterion under the low noise level $(\sigma^2 = 0.25)$ and high noise level $(\sigma^2 = 0.5)$ settings.
-Numbers in parentheses are standard deviations of $\hat r$ over 30 replications.}
-    \label{tab:select}
-\end{table}
-
-\subsection{Extra Simulation Results for Comparison}\label{sec:extrasim}
 
@@
 \end{figure}
-
-\subsection{Human Brain Connectome Data Analysis}
-
-The Human Connectome Project (HCP) aims to construct the structural and functional neural connections in human brains~\citep{van2013wu}.
-We preprocess the original dataset following \cite{desikan2006automated} and partition the brain into 68 regions.
-The cleaned dataset includes brain networks for 136 individuals.
-Each brain network is represented by a 68-by-68 binary symmetric matrix, where the entry with value 1 indicates the presence of connection between node pairs, while the value 0 indicates the absence.
-We use $\tY \in \{0,1\}^{68 \times 68 \times 136}$ to denote the binary tensor.
-Individual attributes such as gender and sex are recorded.
-
-We apply our generalized algorithm to the HCP data with the numbers of clusters on three modes $r_1 = r_2 = 4$ and $r_3 = 3$.
-The selection of $r_1$ and $r_2$ follows the human brain anatomy and the symmetry in the brain network, and the $r_3$ is specified following previous analysis~\citep{hu2021generalized}.
-Because of the symmetry in the data, the estimated brain node clustering results are the same on the first and second modes.
-Figure~\ref{fig:cluster_brain} shows that brain connection exhibits a strong spatial separation structure.
-Specifically, the first cluster, named \emph{L.Hemis}, involves all the nodes in the left hemisphere.
-The nodes in the right hemisphere are further separated into three clusters led by the middle-part tissues in Temporal and Parietal lobes (\emph{R.Temporal}), the back-part tissues in Occipital lobe (\emph{R.Occipital}), and the front-part tissues in Frontal and Parietal lobes (\emph{R.Supra}).
-This clustering result is reasonable since the left and right hemispheres often play different roles in human brains.
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width = 1\columnwidth]{brain_node_cluster.pdf}
-    \caption{ \small Illustration of brain node clustering results for HCP data with (a) top and (b) side views. }
-    \label{fig:cluster_brain}
-\end{figure}
-
-Figure~\ref{fig:ests} illustrates the estimated core tensor $\hat \tS$ with estimated clustering, and Figure~\ref{fig:brain_conn} visualizes the average brain connections and the connection enrichment in contrast to average networks in each group.
-In general, we find that the inner-hemisphere connection has stronger connection compared to inter-hemisphere connections (Figure~\ref{fig:ests}a).
-Also, the back and front parts (\emph{R.Occipital}, \emph{R.Supra}) are shown to have more interactions with temporal tissues than inner-cluster connections.
-In addition, the group 1 with 54\% females shows an enrichment on the inter-hemisphere connections (Figure~\ref{fig:ests}b), while group 4 with only 36\% females exhibits a reduction (Figure~\ref{fig:ests}d).
-This result agrees with previous findings in \cite{hu2021generalized}.
-The enrichment on the back-front connection is also recognized in group~3 (Figure~\ref{fig:ests}c).
-The interpretive patterns in our results demonstrate the usefulness of our clustering methods in the human brain connectome data application.
-
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width = 1\textwidth]{est_S_anno.pdf}
-    \caption{ \small Mode 3 slices of estimated core tensor $\hat \tS$.
-(a) Average estimated slice weighted by the group size; (b)-(d) Group-specified enrichment, i.e., the difference between each slice of $\hat \tS$ and the averaged slice.
-}
-    \label{fig:ests}
-\end{figure*}
-
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width = 1\textwidth]{brain_connection.pdf}
-    \caption{ \small Observed brain connections in the population and each group of individuals.
-(a) Average brain network; (b)-(d) Group-specified brain network enrichments in Groups 1-3.
-Red edges represent the positive enrichment and blue edges represent the negative enrichment.}
-    \label{fig:brain_conn}
-\end{figure*}
 

### 2022-02-20 tensor_clustering/tensor_clustering_aistats_camera/aistats_camera.tex
last version

@@
 The most common model for higher-order clustering is called \emph{tensor block model} (TBM)~\citep{wang2019multiway}, which extends the usual matrix stochastic block model~\citep{abbe2017community} to tensors.
-Matrix analysis tools, however, are sub-optimal for higher-order clustering.
-Developing tensor tools for solving block models has received increased interest recently~\citep{ wang2019multiway,chi2020provable,han2020exact}.
+{Matrix analysis tools, however, are sub-optimal for higher-order clustering.} Developing tensor tools for solving block models has received increased interest recently~\citep{ wang2019multiway,chi2020provable,han2020exact}.
 
@@
 
-{\bf Our contributions.} The primary goal of this paper is to provide both statistical and computational guarantees for dTBM. Our main contributions are summarized below.
+{\bf Our Contributions.} The primary goal of this paper is to provide both statistical and computational guarantees for dTBM. Our main contributions are summarized below.
 \begin{itemize}[wide,topsep=-2pt,itemsep=2pt,parsep=4pt]
@@
 
-{\bf Related work.} 
+{\bf Related Work.} 
 Our work is closely related to but also distinct from several lines of existing research. Table~\ref{tab:comp} summarizes the most relevant models.
@@
 
-\subsection{dTBM}
+\subsection{Degree-corrected Tensor Block Model}
 
@@
 Note that the discrete nature of $\mM$ renders our model~\eqref{eq:model_tensor} more challenging than Tucker decomposition.
-We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$ if $\tY$ admits~\eqref{eq:model_tensor}.
+{We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$ if $\tY$ admits~\eqref{eq:model_tensor}.
+}
 %The goal of clustering is to estimate $z$ from a single noisy tensor $\tY$. 
@@
 Specifically, let $\tY(i_1,...,i_K) = 1$ if  $(i_1,...,i_K) \in E$, otherwise, $\tY(i_1,...,i_K) = 0$, for all $(i_1,\ldots,i_K) \in [p]^K$.
-The equation~\eqref{eq:model_tensor} models $\mathbb{E}\tY$ with degree heterogeneity and subgaussianity parameter $\sigma^2 = 1/4$.
+{The equation~\eqref{eq:model_tensor} models $\mathbb{E}\tY$ with degree heterogeneity and subgaussianity parameter $\sigma^2 = 1/4$.
+}
 \end{example}
@@
 \footnotesize
+{
+\color{red}
 \begin{align}\label{eq:family}
@@
 \end{align}
+}
 \normalsize 
@@
 This constraint allows the trigonometric $\cos$ to describe the angle similarity in the Assumption~\ref{assmp:min_gap} below and Sub-algorithm~\hyperref[alg:main]{2} in Section~\ref{sec:alg}.
-Note that the positivity constraint can be achieved without sacrificing model flexibility, by using a slightly larger dimension of $\tS$ in the factorization~\eqref{eq:model_tensor}; see Supplement~\ref{subsec:positive}.
+Note that the positivity constraint can be achieved without sacrificing model flexibility, by using a slightly larger dimension of $\tS$ in the factorization~\eqref{eq:model_tensor}; {see Supplement~\ref{subsec:positive}}.
 Second, recall that the quantity $|z^{-1}(a)|$ denotes the number of nodes in $a$-th community.
@@
 This constraint, again, incurs no restriction to model flexibility but makes our presentation cleaner.
-See Supplement~\ref{subsec:comp_para} for the parameter space comparison with previous works.
+See Supplement~\ref{subsec:comp_para} for the parameter space comparison with previous work \citep{gao2018community,han2020exact, ke2019community}.
 
@@
 Denominators involved in definition~\eqref{eq:minimal_gap} are well posed because of the lower bound on $\onorm{\mS_{a:}}$ in~\eqref{eq:family}.
-Following theorem shows that the angle separation is sufficient and necessary for the identifiability of dTBM.
+{The following theorem shows that the angle separation is sufficient and necessary for the identifiability of dTBM.
+}
 
@@
 Our goal is to characterize the clustering accuracy with respect to $\gamma$.
-Let $\hat z$ and $z$ be the estimated and true clustering functions in~\eqref{eq:family}.
-Define the misclustering error
+{Let $\hat z$ and $z$ be the estimated and true clustering functions in family~\eqref{eq:family}.
+Define the misclustering error by}
 \[
@@
 The HPC conjecture indicates the impossibility of fully recovering the planted cliques with polynomial-time algorithm when the clique size is less than the number of vertices in the hypergraph.
-The formal statement of HPC detection conjecture is provided in Supplement~\ref{subsec:thm_comp}.
-Under the HPC conjecture, we establish the SNR lower bound that is necessary for any \emph{polynomial-time} estimator to achieve exact clustering.
+{The formal statement of HPC detection conjecture is provided in Supplement~\ref{subsec:thm_comp}.} Under the HPC conjecture, we establish the SNR lower bound that is necessary for any \emph{polynomial-time} estimator to achieve exact clustering.
 
@@
 Particularly, dTBM reduces to matrix degree-corrected model when $K =2$, and the statistical and computational bounds show the same critical value.
-When $K =1$, dTBM reduces to the degree-corrected sub-Gaussian mixture model (GMM).
+{When $K =1$, dTBM reduces to the degree-corrected sub-Gaussian mixture model (GMM).
+Earlier work~\citep{lu2016statistical} implies that polynomial-time algorithms are able to achieve the statistical minimax lower bound in GMM.}
 % with model
@@
 % where $\mY \in \bbR^{p \times d}$ collects $n$ data points in $\bbR^d$, $\mS \in \bbR^{r \times d}$ collects the $d$-dimensional centroids for $r$ clusters, and $\mTheta \in \bbR^{p \times p}, \mM \in \{0,1\}^{p \times r}, \mE \in \bbR^{p \times d}$ have the same meaning as in dTBM.
-\cite{lu2016statistical} implies that polynomial-time algorithms are able to achieve the statistical minimax lower bound in GMM.
  Hence, the statistical-to-computational gap emerges only for higher-order tensors with $K \geq 3$, which reveals the intrinsic distinctions among (vector) one-dimensional clustering, (matrix) biclustering, and (tensor) higher-order clustering.
-Besides, we find that the extra complexity from $\mtheta$ does not render the estimation of $z$ qualitatively harder; see comparison of our phase transition with non-degree TBM~\citep{han2020exact}. 
+ {We also find that the extra complexity from $\mtheta$ does not render the estimation of $z$ qualitatively harder; see comparison of our phase transition with non-degree TBM~\citep{han2020exact}. }
 \end{rmk}
@@
 We start with weighted higher-order clustering algorithm as initialization.
-We take an order-3 symmetric tensor as illustration for insight.
-Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
+{We take an order-3 symmetric tensor as illustration for insight.} Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
 By model~\eqref{eq:model_tensor}, for all $i \in [p]$, we have
@@
 
-We now establish the misclustering error rate of initialization. We call $\mtheta$ is balanced if the relative extent of heterogeneity is comparable across clusters
+We now establish the misclustering error rate of initialization. {We call $\mtheta$ is balanced if the relative extent of heterogeneity is comparable across clusters in that}
 \begin{equation}\label{eq:degree}
@@
  \end{equation}
-Classical Euclidean-distance based clustering~\citep{han2020exact} fails to recover $z$ in the presence of degree heterogeneity, even under noiseless case.
-In contrast, the angle-based $k$-means achieves accurate recovery without explicit estimation of $\mtheta$.
-Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as angle-based $k$-means, except that we use estimated centroids $\ms^{(t)}_a$ in place of $\ms_a$ based on estimated assignment in previous iterations.
+{The classical Euclidean-distance based clustering~\citep{han2020exact} fails to recover $z$ in the presence of degree heterogeneity, even under noiseless case.
+In contrast, the angle-based $k$-means clustering achieves accurate recovery without explicit estimation of $\mtheta$.
+}Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as angle-based $k$-means, except that we use estimated centroids $\ms^{(t)}_a$ in place of $\ms_a$ based on estimated assignment in previous iterations.
 % Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as in angle-based $k$-means.
@@
 % provided that $\mS^{(t)}_{a:}$ is a non-zero vector. Otherwise, if $\mS^{(t)}_{a:}$ is a zero vector, then we make the convention to assign $z^{(t+1)}(i)$ randomly in $[p]$. 
-See Sub-algorithm~\hyperref[alg:main]{2} for full procedures.
+{See Sub-algorithm~\hyperref[alg:main]{2} for full procedures.
+}
 
@@
 \end{equation}
-\vspace{-.4cm}
+\vspace{-.5cm}
 }
@@
 \end{equation}
+\vspace{-.1cm}
 for some permutation $\pi \in \Pi$.
@@
 
- We evaluate the performance of our algorithm in this section.
+ We evaluate the performance of our algorithm\footnote{The R package and data used are available at \texttt{https://cran.r-project.org/package=dTBM}.
+} in this section.
 We report average errors and standard deviations across 30 replications in each experiment.
 Clustering accuracy is assessed by clustering error rate (CER, i.e., one minus rand index).
-Note that CER between $(\hat z, z)$ is equivalent to misclustering error $\ell(\hat z, z)$ up to constant multiplications~\citep{meilua2012local}, and lower CER indicates better performance.
+Note that CER between $(\hat z, z)$ is equivalent to misclustering error $\ell(\hat z, z)$ up to constant multiplications~\citep{meilua2012local}, {and lower CERs indicate better performances.}
 
@@
 Hence, we easily adjust SNR via varying $\alpha$.
-The assortative setting is proposed for simulations, and our algorithm is applicable for general tensors in practice.
-The cluster assignment $z$ is randomly generated with equal probability across  $r$ clusters for each mode.
+{The assortative setting is proposed for simulations, and our algorithm is applicable for general tensors in practice.} The cluster assignment $z$ is randomly generated with equal probability across  $r$ clusters for each mode.
 Without further explanation, we generate degree heterogeneity $\mtheta$ from absolute normal distribution by $\theta(i) = |X_i| + 1 - 1/\sqrt{2\pi}$ with $|X_i| \stackrel{\text{i.i.d.}}\sim N(0,1), i \in [p]$ and normalize $\mtheta$ to satisfy \eqref{eq:family}.
@@
     \includegraphics[width=.95\columnwidth]{phase_anno.pdf}
+    \vspace{-.2cm}
     \caption{ \footnotesize SNR phase transitions for clustering in dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.
@@
      \includegraphics[width=.95\columnwidth]{ini_re_anno.pdf}
+     \vspace{-.2cm}
     \caption{ \footnotesize CER versus signal exponent $(\gamma)$ for initialization only and for combined algorithm.
@@
 \subsection{Comparison with Other Methods}\label{subsec:comp}
-Compare our algorithm with methods below:
+{We compare our algorithm with higher-order clustering methods below:}
 \begin{itemize}[wide,topsep=-3pt,itemsep=0pt,parsep=1pt]
@@
 We set the tuning parameters of \textbf{\small SCORE} as in previous literature \citep{ke2019community}.
-The methods \textbf{\small SCORE} and \textbf{\small HOSVD+} are designed for dTBM~\eqref{eq:model_tensor}, whereas \textbf{\small HOSVD} and \textbf{\small HLloyd} are designed for non-degree models serving as benchmarks.
-We conduct two experiments to assess the impacts of (i) signal strength and (ii) degree heterogeneity under Gaussian and Bernoulli models with $ p = 100, r = 5$.
-We call our algorithm as \textbf{\small dTBM} in the comparison.
+{The methods \textbf{\small SCORE} and \textbf{\small HOSVD+} are designed for dTBM~\eqref{eq:model_tensor}, whereas \textbf{\small HOSVD} and \textbf{\small HLloyd} are designed for non-degree models.
+We conduct two experiments to assess the impacts of (i) signal strength and (ii) degree heterogeneity under Gaussian and Bernoulli models with $ p = 100, r = 5$.} {We call our algorithm as \textbf{\small dTBM} in comparison.}
 
@@
 The only exception is the slightly better performance of \textbf{\small HLloyd} over \textbf{\small HOSVD+} under Gaussian model.
-However, we find the advantage of \textbf{\small HLloyd} disappears with higher degree heterogeneity; see Supplement~\ref{sec:extra_exp}.
+{However, we find the advantage of \textbf{\small HLloyd} disappears with higher degree heterogeneity; see Supplement~\ref{sec:extra_exp}.}
 The experiment demonstrates the benefits of addressing heterogeneity in higher-order clustering tasks.
@@
 In contrast, non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) show poor performance with large heterogeneity, especially in Bernoulli cases.
-This experiment, again, highlights the benefit of addressing degree heterogeneity in clustering.
+{This experiment, again, highlights the benefit of addressing degree heterogeneity in clustering.
+}
 
@@
 We reconstruct legislation network as an order-3 binary tensor $\tY \in \{0,1\}^{116 \times 116 \times 116}$, where $\tY_{ijk} = 1$ if the legislators $(i,j,k)$ have sponsored the same bill, and $\tY_{ijk} = 0$ otherwise.
-True party affiliations of legislators are provided and serve as the ground truth.
-We apply various higher-order clustering methods to $\tY$ with $r = 5$.
+{True party affiliations of legislators are provided and serve as the ground truth.} We apply various higher-order clustering methods to $\tY$ with $r = 5$.
 Table~\ref{tab:peru} shows that our \textbf{\small dTBM} achieves the best performance compared to others.
-The second best method is the two-stage algorithm \textbf{\small HLloyd}, followed by the spectral methods \textbf{\small SCORE} and \textbf{\small HOSVD+}, which consists with simulations under strong signal and moderate degree heterogeneity.
+{The second best method is the two-stage algorithm \textbf{\small HLloyd}, followed by the spectral methods \textbf{\small SCORE} and \textbf{\small HOSVD+}.
+The result is consistent with simulations under moderate heterogeneity.
+}
 %The comparison suggests that our method \textbf{\small dTBM} is more appealing in real-world applications.
-\vspace{-.5cm}
+\vspace{-.1cm}
 \begin{table}[ht]
@@
 
- \subsection{Parameter Space Comparison with Previous Works}\label{subsec:comp_para}
+ \subsection{Parameter Space Comparison with Previous Work}\label{subsec:comp_para}
 

### 2023-01-15 tensor_clustering/ieee_revision/response_s2.tex
update

\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{caption}

\usepackage{xr-hyper}
% this is for overleaf cross reference
\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}

\usepackage[hidelinks]{hyperref}
\usepackage[margin=1in]{geometry}
% \geometry{
%  left=30mm,
%  right=30mm
%  }

\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage[authoryear]{natbib}
\usepackage{url}
\usepackage{enumitem}

\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{algorithm}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor=blue,citecolor=blue,urlcolor={black}}
\usepackage{lscape}
\usepackage{afterpage}
\usepackage{hyperref}

\usepackage{microtype}
\usepackage{wrapfig}
\allowdisplaybreaks
\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{verbatim}
\usepackage{array}

\usepackage{tabularx,booktabs}
\newcolumntype{Y}{>{\centering\arraybackslash}X}

\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{cor}{Corollary}

\theoremstyle{definition}
\newtheorem{assumption}{Assumption}
\newtheorem{defn}{Definition}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}
\newtheorem{condition}{Condition}
\newtheorem{conjecture}{Conjecture}
\newtheorem{property}{Property}
\newtheorem{hypothesis}{Hypothesis}

\newtheorem{innercustomgeneric}{\customgenericname}
\providecommand{\customgenericname}{}
\newcommand{\newcustomtheorem}[2]{%
  \newenvironment{#1}[1]
  {%
   \renewcommand\customgenericname{#2}%
   \renewcommand\theinnercustomgeneric{##1}%
   \innercustomgeneric
  }
  {\endinnercustomgeneric}
}

\newcustomtheorem{customexample}{Example}

\usepackage{appendix}
\usepackage{fullpage} 
\usepackage{microtype}
\usepackage{wrapfig}
\mathtoolsset{showonlyrefs}

\newcommand{\of}[1]{\left(#1\right)}
\newcommand{\off}[1]{\left[#1\right]}
\newcommand{\offf}[1]{\left\{#1\right\}}
\newcommand{\aabs}[1]{\left|#1\right|}
\newcommand{\ang}[1]{\left\langle#1\right\rangle}

\newcommand{\distgap}{D_{\ell_2}}
\newcommand{\Mat}{\text{Mat}}

\def\ci{\perp\!\!\!\perp}

\newtheorem{innercustomgeneric}{\customgenericname}
\providecommand{\customgenericname}{}
\newcommand{\newcustomtheorem}[2]{%
  \newenvironment{#1}[1]
  {%
   \renewcommand\customgenericname{#2}%
   \renewcommand\theinnercustomgeneric{##1}%
   \innercustomgeneric
  }
  {\endinnercustomgeneric}
}

\newcustomtheorem{customthm}{Theorem}
\newcustomtheorem{customlemma}{Lemma}
\newcustomtheorem{customrmk}{Remark}
\newcustomtheorem{customprop}{Proposition}

\usepackage{booktabs}
\newcommand\doubleRule{\toprule\toprule}
\allowdisplaybreaks

\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}

\input macros.tex
\newcommand*{\Scale}[2][4]{\scalebox{#1}{$#2$}}%
\setlength\parindent{0pt}
\usepackage[parfill]{parskip}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsmath}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\newcommand\Algphase[1]{%
\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\textwidth}{0.4pt}%
\Statex\hspace*{-\algorithmicindent}\textbf{#1}%
\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\textwidth}{0.4pt}%
}
\usepackage{caption}
\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}

\myexternaldocument{ieee_revise_second}

\title{\textbf{Response Letter}}
\date{}

\begin{document}

\maketitle
%\doublespacing

\begin{center}
    \textbf{Response to Major Comments from Editor}
\end{center}

\textit{Reviewer 1 has identiﬁed a number of technical concerns in the proofs of some results. Please make sure that these are fully addressed in your revised manuscript.}

\textbf{Response:} We thank editor and reviewers for the comments and suggestions. Reviewer 1 has mainly commented on the Bernoulli initialization in Section~\ref{subsec:ber} and the lower bound statement in Section~\ref{sec:limits}. In comment \# 1, Reviewer 1 has identified the technical issue in the proof of Proposition~\ref{prop:ber}.  In comment \# 3, Reviewer 1 has concerned the proving hardness of our impossibility statements compared with tensor block model (TBM) and pointed out the technical issues in the proof of Theorem~\ref{thm:stats}.

 For comment \# 1, we have developed a new proof of Proposition~\ref{prop:ber} with a new lemma applicable for binary low-rank matrix estimation. The conclusion in Proposition~\ref{prop:ber} still holds. We have put the new proof in Appendix~\ref{sec:initial_prove} with a more rigorous presentation.

 For comment \# 3, we have proved a new stronger statistical impossibility result in Theorem~\ref{thm:stats}, inspired by the reviewer. With the new statement, we have added proof sketches in Section~\ref{sec:mainproof} to highlight the usage of dTBM-specific techniques in our proofs and revised the discussion in Section~\ref{sec:tbm} to emphasize the comparison with TBM. We have provided the proof of the new Theorem~\ref{thm:stats} in Appendix~\ref{sec:statprove1}. We have addressed all reviewer's technical concerns in the new proof.

 In addition to the aforementioned main points, we have addressed all comments from Reviewers 1 and 2. See our response in \textbf{Point-by-point response to Reviewer 1} and \textbf{Point-by-point response to Reviewer 2} for details.

\newpage
\begin{center}
    \textbf{Point-by-point response to Reviewer 1}
\end{center}

\textit{The manuscript has been revised substantially and I appreciate the authors’ efforts on providing additional theoretical and numeric results. Nevertheless, I think the response does not fully address my concerns and the added stuff may not be even technically sound, which I will elaborate as follow.}

\begin{enumerate}
    \item \textit{My first comment in the previous review concerns with the applicability of the proposed framework on Bernoulli setting (hypergraph). The authors introduce the Section IV-C to discuss the initialization strategy for dense Bernoulli data (sub-algorithm 3) with guarantee (Proposition 1).}

    \textbf{Response:} There seem to be two sub-questions in this comment; we answer them one-by-one.

    \begin{enumerate}[wide]
        \item \textit{The proposed strategy relies on the square-unfolding of tensors (i.e., $\Mat_{sq}$). Since this is not a common tensor operation, the authors should provide the mathematical definition. My major concern is the soundness of Proposition 1, which is proved using one line by applying the result by Gao et al. (2018). The referenced paper is a benchmark work on matrix dSBM (so that the input matrix Y in their context is an exact adjacency matrix generated by dSBM). I doubt that, under tensor dSBM, is $\Mat_{sq}(\tY)$ a valid adjacency matrix? From my perspective, the $\Mat_{sq}(\tY)$ may not be a squared matrix and the independent noises may also derive from the setting studied by Gao et al. (2018). In any event, the authors should be more rigours on proving their mathematical statement instead of just saying “Following ...,”.}

        \textbf{Response:} We thank reviewer for the suggestions and pointing out the issues in the proof of Proposition~\ref{prop:ber}.

        First, we have added the definition of operation $\Mat_{sq}$ and the interpretation of the matricization $\Mat_{sq}(\tY)$. The matrix  $\Mat_{sq}(\tY)$ is not square and symmetric, but we are able to interpret $\Mat_{sq}(\tY)$ as a valid adjacency matrix for a bipartite network with connections between two groups of nodes. We quote the corresponding revision in Section~\ref{subsec:ber}:

        \begin{quote}
            ``...
            One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square}, $\Mat_{sq}(\cdot)$, of Bernoulli tensor $\tY \in \{0,1\}^{p_1 \times \cdots \times p_K}$. Specifically, the square matricization $\mat_{sq} (\tY) \in \{0,1\}^{p^{\floor{K/2}} \times p^{\ceil{K/2}} }$ has entries $[\Mat_{sq}(\tY)](j_1, j_2) = \tY(i_1, \ldots, i_K)$, where
\begin{align}
    j_1 &= i_1 + p_1(i_2 - 1) + \cdots + p_1 \cdots p_{\floor{K/2}-1} (i_{\floor{K/2}} - 1),\\
    j_2 &= i_{\ceil{K/2}} + p_{ \ceil{K/2}}(i_{\ceil{K/2} + 1} - 1) + \cdots + p_{ \ceil{K/2}} \cdot p_{K-1} (i_K-1).
\end{align}
The matrix $\Mat_{sq}(\tY)$ is asymmetric. We are able to interpret $\Mat_{sq}(\tY)$ as the adjacency matrix for a bipartite network with connections between two groups of node. The two groups of nodes in the bipartite network have $p_1 \cdots p_{\floor{K/2}}$ and $ p_{\ceil{K/2}} \cdots p_K$ nodes, respectively. The entry $[\Mat_{sq}(\tY)](j_1, j_2)$ refers to the presence of connection between the nodes indexed by combinations $(i_1, \ldots, i_{\floor{K/2}})$ and $(i_{\ceil{K/2}}, \ldots, i_K)$. We summarize the procedure in Sub-algorithm 3.
 ... "
        \end{quote}

        Second, we agree that \citet[Lemma 7]{gao2018community} is non-applicable in our case due to the asymmetry and the different estimation procedure to obtain $\hat \tX'$. Instead, we have applied a new lemma for low-rank matrix estimation to upper bound the estimation error via square unfolding. The conclusion in Proposition~\ref{prop:ber} still holds with the new lemma. We have revised the proof more rigorously without ``following..." terms and put the proof to Appendix~\ref{sec:initial_prove} for a more detailed presentation.

        For self-consistency, we attach the Proposition~\ref{prop:ber} with proof and the new lemma here:
        \begin{quote}
            `` ...
            \begin{customprop}{1}[Error for Bernoulli initialization] Consider the Bernoulli dTBM in the parameter space $\tP$ and Assumption~\ref{assmp:min_gap} holds. Assume that $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$. Let $ z^{(0)}$ denote the output of Sub-algorithm~3. With probability going to 1, we have
\begin{equation}\label{eq:ini_b}
 \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}, \quad \text{and} \quad L(z^{(0)}, z) \lesssim  {\sigma^2 r^K p^{-\floor{K/2}}}.
\end{equation}
            \end{customprop}
... "
        \end{quote}

        \begin{quote}

``... \begin{proof}[Proof of Proposition~\ref{prop:ber}] Sub-algorithm~3 shares the same algorithm strategy as Sub-algorithm~\hyperref[alg:main]{1} but with a different estimation of the mean tensor, $\hat \tX'$. Hence, the proof of Proposition~\ref{prop:ber} follows the same proof idea with the proof of Theorem~\ref{thm:initial}. Replacing the estimation $\hat \tX$ by $\hat \tX'$ in the proof of Theorem~\ref{thm:initial}, we have 
\begin{align}
        \min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2  &\lesssim \of{\sum_{i \in S} \onormSize{}{\mX_{i:}}^2 + \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2  } p^{-(K-1)} r^{K-1}. \label{eq:binary_theta}
    \end{align}
By inequalities~\eqref{eq:lem1} and \eqref{eq:s0}, we have 
\begin{align}
   \sum_{i \in S} \onormSize{}{\mX_{i:}}^2 &\leq  \of{ \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 } + 2 }\onormSize{}{\hat \tX' - \tX}_F^2, \label{eq:binary_theta1}  \\
   \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2 &\leq  \onormSize{}{\hat \tX' - \tX}_F^2. \label{eq:binary_theta2}
\end{align}
Hence, it suffices to find the upper bound of the estimation error $\onormSize{}{\hat \tX' - \tX}_F^2$ to complete our proof. Note that the matricization $\Mat_{sq}(\tX) \in \bbR^{p^{\floor{K/2}} \times p^{\ceil{K/2}}}$ has $\text{rank}(\Mat_{sq}(\tX)) \leq r^{\ceil{K/2}}$, and Bernoulli random variables follow the sub-Gaussian distribution with bounded variance $\sigma^2 = 1/4$. Apply Lemma~\ref{lem:lowrank} to $\mY = \Mat_{sq}(\tY), \mX = \Mat_{sq}(\tX)$, and $\hat \mX = \Mat_{sq}(\hat \tX')$. Then, with probability tending to 1 as $ p \rightarrow \infty$, we have 
        \begin{equation}\label{eq:binary_est}
            \onormSize{}{\hat \tX' - \tX}_F^2 = \onormSize{}{\Mat_{sq}(\hat \tX') - \Mat_{sq}(\tX)}_F^2 \lesssim p^{\ceil{K/2}}.
        \end{equation}
Combining the estimation error~\eqref{eq:binary_est} with inequalities~\eqref{eq:binary_theta1}, \eqref{eq:binary_theta2}, and \eqref{eq:binary_theta}, we obtain 
\begin{equation}\label{eq:binary_theta_sum}
    \min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2 \lesssim  \frac{\sigma^2 r^{K-1}}{ \Delta_{\min}^2  p^{K-1} } p^{\ceil{K/2}}.
\end{equation}
Replace the inequality~\eqref{eq:theta_sum} in the proof of Theorem~\ref{thm:initial} by inequality~\eqref{eq:binary_theta_sum}. With the the same procedures to obtain $\ell(\hat z^{(0)}, z)$ and $L(\hat z^{(0)}, z)$ for Theorem~\ref{thm:initial} , we finish the proof of Proposition~\ref{prop:ber}.
\end{proof}

\begin{customlemma}{9}[Low-rank matrix estimation]  Let $\mY = \mX + \mE \in \bbR^{m \times n}$, where $\mE$ contains independent mean-zero sub-Gaussian entries with bounded variance $\sigma^2$. Suppose $\text{rank}(\mX) = r$. Consider the least square estimator 
        \begin{equation}
            \hat \mX = \argmin_{\mX' \in \bbR^{m \times n}, \text{rank}(\mX') \leq r} \onormSize{}{\mX' - \mY}_F^2.
        \end{equation}
        There exist positive constant $C_1, C_2$ such that
        \begin{equation}
            \onormSize{}{\hat \mX - \mX}_F^2 \leq C_1 \sigma^2 nr,
        \end{equation}
        with probability at least $1 - \exp(-C_2 nr)$.
\end{customlemma}

\begin{proof}[Proof of Lemma~\ref{lem:lowrank}] Note that $\onormSize{}{\hat \mX - \mY}_F^2 \leq \onormSize{}{\mX - \mY}_F^2$ by the definition of least square estimator. We have 
        \begin{align}
            \onormSize{}{\hat \mX - \mX}_F^2 & \leq 2 \ang{ \hat \mX - \mX, \mY - \mX} \\ & \leq 2 \onormSize{}{\hat \mX - \mX}_F \sup_{\mT \in \bbR^{m \times n}, \text{rank}(\mT) \leq 2r, \onormSize{}{\mT}_F = 1} \ang{\mT, \mY - \mX} \\
            & \leq C \sigma  \onormSize{}{\hat \mX - \mX}_F \sqrt{nr},
        \end{align}
        with probability at least $1 - \exp(-C_2 nr)$, where the second inequality follows by re-arrangement, and the last inequality follows from \citet[Lemma E5]{han2022optimal} under the matrix case. "
\end{proof}
        \end{quote}

%         We attach our revision in Section IV-C here:
%         \begin{quote}
%         ``...

%         One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square}, $\Mat_{sq}(\cdot)$, of Bernoulli tensor $\tY \in \{0,1\}^{p_1 \times \cdots \times p_K}$. Specifically, we have $\mat_{sq} (\tY) \in \{0,1\}^{p^{\floor{K/2}} \times p^{\ceil{K/2}} }$ with entries $[\Mat_{sq}(\tY)](j_1, j_2) = \tY(i_1, \ldots, i_K)$, where
% \begin{align}
%     j_1 &= i_1 + p_1(i_2 - 1) + \cdots + p_1 \cdots p_{\floor{K/2}-1} (i_{\floor{K/2}} - 1),\\
%     j_2 &= i_{\ceil{K/2}} + p_{ \ceil{K/2}}(i_{\ceil{K/2}} - 1) + \cdots + p_{ \ceil{K/2}} \cdot p_{K-1} (i_K-1).
% \end{align}
% The square matricization $\Mat_{sq}(\tY)$ is an asymmetric matrix. We can interpret $\Mat_{sq}(\tY)$ as the adjacency matrix for a bipartite network with connections between two groups of node. The first and second groups have $p_1 \cdots p_{\floor{K/2}}$ and $ p_{\floor{K/2}} \cdots p_K$ nodes, respectively. The entry $[\Mat_{sq}(\tY)](j_1, j_2)$ with $j_1, j_2$ in above equations refer to the presence of connections between the nodes corresponding to combinations $(i_1, \ldots, i_{\floor{K/2}})$ and $(i_{\ceil{K/2}}, \ldots, i_K)$. We summarize the procedure in Sub-algorithm 3.

% \begin{algorithm}[h!]
% \caption*{\bf Sub-algorithm 3: Weighted higher-order initialization for Bernoulli observation}
% \vspace{.15cm}
% \begin{algorithmic}[1] 
% \INPUT Bernoulli tensor $\tY \in \{0,1\}^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
% \State  Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{p^{\floor{K/2}} \times p^{\ceil{K/2}} }$ denote the nearly square unfolded tensor. Compute the estimate $\tX'$, where
% \begin{equation}\label{eq:matrixsvd}
%     \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX) -  \mat_{sq}(\tY)}_F^2.
% \end{equation}
% \State Implement lines 3-5 of Sub-algorithm~\hyperref[alg:main]{1} with $\hat \tX$ replaced by $\hat \tX'$ in~\eqref{eq:matrixsvd}.
% \OUTPUT Initial clustering $z^{(0)} \leftarrow \hat z$.
% \end{algorithmic}
% \end{algorithm}

% \begin{prop}[Error for Bernoulli initialization]\label{cor:ber} Consider the Bernoulli dTBM in the parameter space $\tP$ and Assumption~\ref{assmp:min_gap} holds. Assume that $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$. Let $ z^{(0)}$ denote the output of Sub-algorithm~3. With probability going to 1, we have
% \begin{equation}\label{eq:ini_b}
%  \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}, \quad \text{and} \quad L(z^{(0)}, z) \lesssim  {\sigma^2 r^K p^{-\floor{K/2}}}.
% \end{equation}
% \end{prop}

%  \begin{proof}[Proof of Proposition~\ref{cor:ber}] Note that $\Mat_{sq}(\tX) \in \bbR^{p^{\floor{K/2}} \times p^{\ceil{K/2}}}$ and $\text{rank}(\Mat_{sq}(\tX)) \leq r^{\ceil{K/2}}$. Apply Lemma~\ref{lem:lowrank_binary} with $\mY = \Mat_{sq}(\tY), \mX = \Mat_{sq}(\tX)$, and $\hat \mX = \Mat_{sq}(\hat \tX')$. With probability tending to 1 as $ p \rightarrow \infty$, we have 
%         \begin{equation}
%             \onormSize{}{\hat \tX' - \tX}_F^2 = \onormSize{}{\Mat_{sq}(\hat \tX') - \Mat_{sq}(\tX)}_F^2 \lesssim p^{\ceil{K/2}}.
%         \end{equation}
%         We then obtain the upper bound by using $\hat \tX'$ in place of $\hat \tX$ in Theorem~\ref{thm:initial}.\fixme{Miaoyan}{you did not mention anywhere in the proof where the assumptions of $\theta$ is used...Change ``we then obtain the upper bound....'' by the full proof.
%         }
%         \end{proof}

% \begin{lem}[Low-rank binary matrix estimation] \label{lem:lowrank_binary} Let $\mY \in \{0, 1\}^{m \times n}$ denote a binary matrix with independent Bernoulli entries. \fixme{Miaoyan}{remove binary. The proof does not use binary at all. Every mathematical Lemma should mention only *minimal* necessary conditions.  Never, never put redundant and useless conditions.} Suppose that $n \geq m$ and $\bbE[\mY] = \mX \in [0,1]^{m \times n}$ with fixed rank $\text{rank}(\mX) \leq r$. Consider the least square estimator 
%         \begin{equation}
%             \hat \mX = \argmin_{\mX' \in \bbR^{m \times n}, \text{rank}(\mX') \leq r} \onormSize{}{\mX' - \mY}_F^2.
%         \end{equation}
%         There exist positive constant $C_1, C_2$ such that
%         \begin{equation}
%             \onormSize{}{\hat \mX - \mX}_F^2 \leq C_1 nr,
%         \end{equation}
%         with probability at least $1 - \exp(-C_2 nr)$.
%         \end{lem}
%  "
%         \end{quote}

%         For self-consistency, we also attach the proof for Lemma~\ref{lem:lowrank_binary} in Appendix C here:
%         \begin{quote}
%         ``
%             \begin{proof}[Proof of Lemma~\ref{lem:lowrank_binary}] Note that $\onormSize{}{\hat \mX - \mY}_F^2 \leq \onormSize{}{\mX - \mY}_F^2$ by the definition of least square estimator. Re-arranging the terms, we have 
%         \begin{align}
%             \onormSize{}{\hat \mX - \mX}_F^2 & \leq 2 \ang{ \hat \mX - \mX, \mY - \mX} \\ & \leq 2 \onormSize{}{\hat \mX - \mX}_F \sup_{\mT \in \bbR^{m \times n}, \text{rank}(\mT) \leq 2r, \onormSize{}{\mT}_F = 1} \ang{\mT, \mY - \mX} \\
%             & \leq C  \onormSize{}{\hat \mX - \mX}_F \sqrt{nr}
%         \end{align}
%         with probability at least $1 - \exp(-C_2 nr)$ for some positive constant $C_2$, where the second inequality follows from \citet[Lemma 1]{wang2019multiway}\fixme{Miaoyan}{provide a full proof for self-containness.}, and the last inequality follows from \citet[Lemma E5]{han2022optimal} \fixme{Miaoyan}{Not reader friendly! This is a simple results in matrix theory. Do not over-complicate things by involving a harder lemma.
%         Find a textbook for easy formulation for reference. }under the matrix case and the fact that the Orlicz norm of Bernoulli noise $\onormSize{}{\mY_{ij} - \mX_{ij}}_{\varphi_2} = \sup_{q \geq 1} \bbE[|\mY_{ij} - \mX_{ij}|^q]^{1/q}/q^{1/2} \leq 1$ since $|Y_{ij} - X_{ij}| \leq 1$ for all $i \in [m], j \in [n]$.
%         \end{proof} "
%         \end{quote}

        \item \textit{Also, Comparison in Remark 3 seems unfair. While the final exponential error rate developed in this paper is tight, it requires quite stronger SNR for a warm initialization. So, it does not make sense to compare the final rate without talking about under what SNR conditions these rates are achievable. Could the authors compare the SNR conditions under which both methods (the proposed one and Ke et al., 2019) achieve exact recovery?}

        \textbf{Response:} It seems that reviewer intends to comment the Remark 8, which compares the error rates between our algorithm and \cite{ke2019community}, rather than the Remark 3 for degree assumption, which is irrelevant to algorithm performance. We answer this question with focus on the Remark 8.

        The method in \cite{ke2019community} adopts a signal notion based on the singular gap of the core tensor, denoted as $\Delta_{\rm singular}$. However, we are not able to infer the exact recovery of the spectral method by our angle-base SNR condition. There exists a case in which our algorithm achieves exact recovery but the spectral method in \cite{ke2019community} fails. Hence, for fair comparison, we compare the best performance of our algorithm and \cite{ke2019community} under the strongest signal setting of each model. We have added a new remark in Section~\ref{subsec:ber} and elaborated the comparison with \cite{ke2019community} with SNR consideration:
        \begin{quote}
        ``...
            \begin{customrmk}{10}[Comparison with previous methods]   Previous work \citep{ke2019community} develops a spectral clustering method for Bernoulli dTBM. \cite{ke2019community} adopts a different signal notion based on the singular gap in the core tensor, denoted as $\Delta_{\rm singular}$. By \citet[Theorem 1]{ke2019community}, the spectral method achieves exact recovery with $\Delta_{\rm singular} \gtrsim p^{-1/2}$. However, we are not able to infer the exact recovery of spectral method by our angle-base SNR condition. Consider an order-2 dTBM with $p > 2, \sigma^2 = 1$, $\mtheta = \mathbf{1}$, equal size assignment $|z^{-1}(a)| = p/r$ for all $ a \in [r]$, and core matrix $\mS = \mI_2$. The singular gap under this setting is $\Delta_{\rm singular} = \min\{ \lambda_1 - \lambda_2, \lambda_2 \} = 0$, where $\lambda_1 \geq \lambda_2$ are singular values of $\mS$. In contrast, our angle gap $\Delta_{\min}^2 = 2$ satisfies the SNR condition in Theorem~\ref{thm:refinement}. Then, our algorithm achieves the exact recovery, but the spectral method in \cite{ke2019community} fails.

Hence, for fair comparison, we compare the best performance of our algorithm and \cite{ke2019community} under the strongest signal setting of each model. Since both methods contain an iteration procedure, we set the iteration number to infinity to avoid the computational error. Considering the largest angle-based SNR $\asymp 1$ in Theorem~\ref{thm:refinement}, our Bernoulli clustering achieves exponential error rate of order $\exp(-p^{(K-1)})$; considering the largest singular gap $\Delta_{\rm singular} \asymp 1$ in Theorem 1 of \cite{ke2019community}, the spectral clustering has a polynomial error rate of order $p^{-2}$. Our algorithm still shows a better theoretical accuracy than the competitive work for Bernoulli observations.   "
            \end{customrmk}

        \end{quote}

        % In Remark 8, we compare the best performances that our algorithm and method in \cite{ke2019community} can achieve. Both methods contain an iteration procedure, and \cite{ke2019community} adopts a singular gap signal notion different with our angle gap notion. To obtain the best performance, we set the number of iterations to infinity for both methods to avoid computational error and consider the error rates under the strongest angle or singular based SNR, respectively.

        % We consider the algorithm error rates of both methods with infinity number of iterations and with the strongest angle and singular SNR

        % Since both methods contain an iteration procedure, we set both iteration numbers to infinity to avoid the computational error; we

    \end{enumerate}

    \item \textit{My second concern was on the imposed identifiability conditions and I think the authors have fully addressed it.}

    \item \textit{My last comment surrounding the two lower bounds is the major one, but I don’t think the authors have addressed it appropriately. }

    \textbf{Response:} This comment concerns the proving hardness of the lower bound statements in dTBM and the flaws in the proof of Theorem~\ref{thm:stats}. We separate the comment by these two aspects and address each aspect one-by-one.

    \begin{enumerate}[wide]
        \item \textit{The authors argued that ``Our results show the similar conclusion but under different conditions.”. I agree that $\Delta_{\rm ang}$ and $\Delta_{\rm Euc}$ (using the notation is the response) are different quantities and $\Delta_{\rm ang}$ can be much smaller than $\Delta_{\rm Euc}$ in the proposed setting. However, that does NOT affect the hardness in proving the lower bound statement. For simplicity, I will just show that for statistical lower bound, which claims that}
    \begin{equation}
        \inf_{\hat z} \sup_{(z,S, \theta) \in P(\gamma)} \bbE[p\ell(\hat z, z )] \geq 1.
    \end{equation}
    \textit{Here $P(\gamma)$ refers to the parameter space where the core tensor S satisfies $\Delta_{\rm ang}^2 /\sigma^2 \leq p^{\gamma}$  and $z, \theta$ satisfy some regularity conditions (e.g., balance). To prove the lower bound, it suffices to fix particular $\theta$ and $S$ and take supreme over the community variable $z$. Then, a trivial selection is to set $\theta_i = 1$ (homogeneous degree) and some S such that $\Delta_{\rm ang}/\Delta_{\rm Euc} = 1 + o(1)$. Such $S$ can be constructed randomly: $S_{ijk} = 1 + \epsilon_{ijk}$ where $\epsilon_{ijk}$  are independent mean-zero Gaussian with variance tends to zero. After constructing such $\theta_i$ and $S$, the heterogeneity issue vanishes and $\Delta_{\rm Euc}$ and $\Delta_{\rm ang}$ becomes equivalent.}

    \textit{The lower bounds arguments for TBM and dTBM become really different if the author considers the following stronger statement:}
    \begin{equation}
        \inf_{S \in P(\gamma)} \inf_{\hat z} \sup_{(z, \theta)} \bbE[p \ell(\hat z, z)] \geq 1
    \end{equation}
    \textit{This statement is stronger than the previous one, as it suggests statistical impossibility whenever the SNR condition is not met, while the current one actually suggests statistical impossibility for at least a specific $S$ that does not meet SNR requirement.}

    \textbf{Response:} 
    %The former part of the comment concerns the proving hardness of lower bound statements in dTBM compared with TBM. 
    We thank reviewer for the suggestions. In short, we have provided a stronger statistical impossibility statement for dTBM, inspired by the reviewer. With the new statement, we argue that dTBM-specific techniques are required to obtain our statistical and computational impossibilities, though the proof idea shares the same spirit with reviewer's proof in comment. We have made several revisions in Section~\ref{sec:limits}, Section~\ref{sec:tbm}, Section~\ref{sec:mainproof}, and Appendix~\ref{sec:statprove1} to address the comment. We attach each revision with explanations below.

    First, we have provided a new impossibility result in Theorem~\ref{thm:stats}. The new Theorem~\ref{thm:stats} suggests the statistical impossibility of dTBM whenever the core tensor $\tS$ leads to a SNR with signal exponent $\gamma < -(K-1)$. This new impossibility statement is stronger than previous one which suggests the worst case impossibility for a particular $\tS$. We quote the revised theorem and discussion in Section~\ref{sec:limits} here:

    \begin{quote}
    ``
    \begin{customthm}{2}[Statistical critical value]  Consider general Gaussian dTBMs with parameter space $\tP(\gamma)$ with $K\geq 1$. Then, we have the following statistical phase transition. 
    \begin{itemize}[wide]
    \item 
    \textbf{Impossibility.} Assume $r\lesssim p^{1/3}$. Let $\tP_{\tS}(\gamma) \coloneqq \{ \tS: c_3\leq \onormSize{}{\text{Mat}(\tS)_{a:}}\leq c_4, a\in[r]\} \cap \{ \tS: \Delta_{\min}^2 = p^{\gamma}\}$ denote the space for valid $\tS$ satisfying SNR condition~\eqref{eq:gamma}, and $\tP_{z ,\mtheta} \coloneqq \{ \mtheta\in\mathbb{R}^p_{+},\ 
{c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r}, \onorm{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|, a \in [r] \}$ denote the space for valid $(z, \mtheta)$,  where $c_1, c_2, c_3, c_4$ are the constants in parameter space \eqref{eq:family}. If the signal exponent satisfies $\gamma < -(K-1)$, then, for any true core tensor $\tS \in \tP_{\tS}(\gamma)$, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation; that is,
    \begin{equation}
  \gamma< -(K-1) \quad \Rightarrow  \quad \liminf_{p \rightarrow \infty}\inf_{\tS \in  \tP_{\tS}(\gamma)}  \inf_{\hat z_{\text{stat}}}\sup_{ (z, \mtheta) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
\end{equation}
\end{itemize}
    \end{customthm} ... "
    \end{quote}

    \begin{quote}
        `` ...
        The proofs for the two parts in Theorem~\ref{thm:stats} are in the Appendix II, Section~\ref{sec:statprove1} and Section~\ref{sec:statprove2}, respectively. The first part of Theorem~\ref{thm:stats} demonstrates impossibility of exact recovery whenever the core tensor $\tS$ satisfies SNR condition~\eqref{eq:gamma} with exponent $\gamma < -(K-1)$. The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to MLE and trace maximization~\citep{ghoshdastidar2017uniform}. The minimax bound~\eqref{eq:minminmax} indicates the worst case impossibility for a particular core tensor $\tS$ with signal exponent $\gamma < -(K-1)$; i.e., under the assumptions of Theorem~\ref{thm:stats}, 
\begin{equation}
    \gamma < -(K-1) \quad \Rightarrow \quad \liminf_{p \rightarrow \infty}  \inf_{\hat z_{\text{stat}}}\sup_{ (z,\tS, \mtheta) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
\end{equation}
Such worst case impossibility is frequently studied in related works \citep{han2020exact, gao2018community} while our lower bound~\eqref{eq:minminmax} provides a stronger impossibility statement for arbitrary core tensor with weak signal.   ..."
    \end{quote}

    Second, we have added proof sketches for the impossibility results to better highlight the dTBM-specific techniques we used. In fact, our proof idea is common in minimax analysis and shares the same spirit with reviewer's proof idea in the comment: find the lower bound of minimax error by constructing a particular set of true parameters. However, angle-based signal notion brings extra difficulties in parameter constructions, and the previous TBM construction is no longer applicable to address the arbitrariness of core tensor. We quote the new proof sketches in new Section~\ref{subsec:impossible} here:

    \begin{quote}
    `` ...

    The proofs of impossibility in Theorems~\ref{thm:stats} and \ref{thm:comp} share the same proof idea with \citet[Theorems 6 and 7]{han2020exact} and \citet[Theorem 2]{gao2018community}. In both proofs of statistical and computational impossibilities, the key idea is to construct a particular set of parameters to lower bound the minimax rate. Specifically, for statistical impossibility in Theorem~\ref{thm:stats}, we construct a particular $(z^*_{\rm stats}, \mtheta^*_{\rm stats})  \in \tP_{z, \mtheta} $ such that for all $\tS^* \in \tP_{\tS}(\gamma)$
\begin{equation}
    \inf_{\hat z_{\rm stats}} \sup_{(z, \mtheta) \in \tP_{z, \mtheta}} \bbE[p \ell(\hat z_{\rm stat}, z)] \geq \inf_{\hat z_{\rm stats}} \bbE[p \ell(\hat z_{\rm stat}, z_{\rm stats}^*)| (z^*_{\rm stats}, \tS^*,\mtheta^*_{\rm stats})] \geq 1;
\end{equation}
for computational impossibility in Theorem~\ref{thm:comp}, we construct a particular $(z^*_{\rm comp}, \tS^*_{\rm comp}, \mtheta^*_{\rm comp})$ $\in \tP(\gamma)$ such that 
\begin{equation}
    \inf_{\hat z_{\rm comp}} \sup_{(z, \tS, \mtheta) \in \tP(\gamma)} \bbE[p \ell(\hat z_{\rm comp}, z)] \geq \inf_{\hat z_{\rm comp}} \bbE[p \ell(\hat z_{\rm comp}, z_{\rm comp}^*)| (z^*_{\rm comp}, \tS^*_{\rm comp},\mtheta^*_{\rm comp})] \geq 1.
\end{equation}

The constructions of $(z^*_{\rm stats}, \mtheta^*_{\rm stats})$ and $(z^*_{\rm comp}, \tS^*_{\rm comp}, \mtheta^*_{\rm comp})$ are the most critical steps. With good constructions, the lower bound ``$\geq 1$" can be verified by classical statistical conclusions (e.g.\ Neyman-Pearson Lemma) or prior work (e.g.\ HPC Conjecture).

A notable detail in the proof of statistical impossibility is the arbitrariness of $\tS^*$. The first infimum over $ \tP_{\tS}(\gamma)$ in the minimax rate~\eqref{eq:minminmax} requires that the lower bound~\eqref{eq:stats_impo_sketch} holds for any $\tS^* \in \tP_{\tS}(\gamma)$. The arbitrary choice of $\tS^*$ brings extra difficulties in the parameter construction, and consequently a non-trivial $\mtheta^*_{\rm stats} \neq \mathbf{1}$ is chosen to address the arbitrariness. Previous TBM construction in the proof of \citet[Theorem 6]{han2020exact} with $\mtheta^*_{\rm stats} = \mathbf{1}$ is no longer applicable in our case. Meanwhile, our construction $(z^*_{\rm comp}, \tS^*_{\rm comp}, \mtheta^*_{\rm comp})$ leads to a rank-2 mean tensor to relate the HPC Conjecture while TBM \citet[Theorem 7]{han2020exact} constructs a rank-1 mean tensor. Hence, we emphasize that dTBM-specific techniques are required to obtain our impossibility results, though the proof idea is common for minimax lower bound analysis. ... "
    \end{quote}

    Here are two details differ with reviewer's comment. First, reviewer's construction with $\mtheta^* = \mathbf{1}$ is non-applicable for our new statistical impossibility. The TBM construction with $\mtheta^*_{\rm stats} = \mathbf{1}$ is not able to handle the arbitrary choice of $\tS^*$, because the Euclidean gap condition $\Delta_{\rm Euc}^2 \leq p^{\gamma}$ is not guaranteed with arbitrary $\tS^*$ satisfying $\Delta_{\rm ang}^2 \leq p^{\gamma}$; see Example \ref{example:euc_alg} in our main text. In consequence, we design a special $\mtheta^*_{\rm stats} \neq \mathbf{1}$ in the proof.
    %Because the core tensor $\tS$ with angle gap $\Delta_{\rm ang}^2 \leq p^{\gamma}$ is not guaranteed to have Euclidean gap $\Delta_{\rm Euc}^2 \leq p^{\gamma}$; see Example \ref{example:euc_alg} in our main text. In consequence, we construct a special $\mtheta^*_{\rm stats} \neq \mathbf{1}$ to address the arbitrariness of $\tS^*$ with angle gap requirement. 
    Second, we consider one particular core tensor $\tS^*$ at a time, rather than the random tensor with entries $\tS_{ijk}^* = 1 + \epsilon_{ijk}, \epsilon_{ijk} \sim N(0, \delta^2)$ for some $\delta \rightarrow 0$ as $p \rightarrow \infty$. The expectation in minimax error is taken with respect to the noise in observation, rather than the randomness in the true parameters.

    We have also revised the discussion in Section~\ref{sec:tbm} based on the new theorem and proof sketches:

    \begin{quote}
    ``...

    Third, from the perspective of proofs, we develop new dTBM-specific techniques to handle the extra degree heterogeneity. In our Theorem~\ref{thm:stats}, we construct a special non-trivial degree heterogeneity to establish the lower bound for arbitrary core tensor with small angle gap, while, TBM~\citep{han2020exact} considers the constructions without degree parameter.
%we consider the profile MLE by maximizing out the nuisance degree parameter, while TBM~\citep{han2020exact} considers the usual MLE without degree parameter.
In our Theorem~\ref{thm:comp}, we construct a rank-2 tensor to relate HPC conjecture to $\Delta^2_{\text{ang}}$, while TBM~\citep{han2020exact} constructs a rank-1 tensor to relate HPC conjecture to $\Delta^2_{\text{Euc}}$. The asymptotic non-equivalence between $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$ renders our proof technically more involved. ..."
    \end{quote}

    Last, we have added the comparison with TBM in terms of the stronger statistical impossibility. Though the presentation in the Theorem 6 of \cite{han2020exact} suggests a weaker worst case impossibility for TBM, the proof of Theorem 6 indeed implies a stronger statistical impossibility for arbitrary core tensor with small signal. Both models show similar conclusions under different signal notions again, but we emphasize that different techniques are required to address the different signal notions. We quote the new discussion in Section~\ref{sec:tbm} here:

    \begin{quote}
        `` ...

         Last, we discuss the statistical impossibility statements. Our Theorem~\ref{thm:stats} implies the statistical impossibility whenever the core tensor $\tS$ leads to an angle-based SNR below the critical threshold. The Theorem 6 in \cite{han2020exact} implies the worst case statistical impossibility for a particular core tensor $\tS$ with Euclidean-based SNR below the statistical limit. Hence, our Theorem~\ref{thm:stats} shows a stronger statistical impossibility for dTBM than that presented in TBM \citet[Theorem 6]{han2020exact}. However, inspecting the proof of \cite{han2020exact}, the proof of Theorem 6 indeed implies a stronger TBM impossibility statement for arbitrary core tensor; i.e.,
 \begin{equation}
     \gamma < -(K-1) \quad \Rightarrow \quad \liminf_{p \rightarrow \infty} \inf_{ \tS \in \tP_{\tS, {\rm TBM}} \cap \{\Delta_{\rm Euc}^2 = p^{\gamma} \} }\inf_{\hat z_{\rm stats}}  \sup_{z \in \tP_{z, {\rm TBM}}}  \bbE[p \ell(\hat z_{\rm stats}, z)] \geq 1,
 \end{equation}
 where $\tP_{\tS, {\rm TBM}}$ and $\tP_{z, {\rm TBM}}$ refer to the space for core tensor $\tS$ and assignment $z$ under TBM, respectively. Again, in terms of the strong statistical impossibility, both models show similar conclusions but under different conditions. Since two impossibilities consider different core tensor regimes with non-equivalent $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$, we emphasize that different proof techniques are required to obtain these similar conclusions.  See our proof sketch in Section~\ref{subsec:impossible}, Appendices~\ref{sec:statprove1} and \ref{sec:compprove1} for detail technical differences. "
    \end{quote}

    For self-consistency, we also attach the proof of the new impossibility theorem in Appendix~\ref{sec:statprove1}:

    \begin{quote}
    ``...
        \begin{proof}[Proof of Theorem~\ref{thm:stats} (Impossibility)]Consider the general asymetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$. For simplicity, we show the minimax rate for the estimation on the first mode $\hat z_1$; the proof for other modes are essentially the same.

   To prove the minimax rate~\eqref{eq:minminmax}, it suffices to take an arbitrary $\tS^* \in  \tP_{\tS}(\gamma)$ wih $\gamma < -(K-1)$ and construct $(z^*_k, \mtheta^*_k)$ such that 
   \begin{equation}
       \inf_{\hat z_1} \bbE \left[ p\ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right]\geq 1.
   \end{equation}

   We first define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation. Based on \citet[Proof of Theorem 6]{han2020exact}, we consider the restricted family of $\hat z_k$'s for which the following three conditions are satisfied:
   \begin{align}
       &\text{(a)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \quad \text{(b)} \ |T^c_k|\asymp {p\over r}; \\
       &\text{(c)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
   \end{align}
for all $k \in [K]$.
   Now, we consider the construction:
   \begin{enumerate}
       \item[(i)] $\{z_k^*\}$ satisfies properties (a)-(c) with misclassification sets $T_k^c$ for all $k \in [K]$;
       \item [(ii)] $\{\mtheta_k^*\}$ such that $\mtheta_k^*(i) \leq \sigma r^{(K-1)/2} p^{-(K-1)/2}$ for all $i \in T_k^c, k \in [K]$ and $\max_{k \in [K], a \in [r]} $ $\onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}^2_2 \asymp p/r$.
   \end{enumerate}
   % There exists a $(z^*, \mtheta^*) \in \tP_{z, \mtheta}$ satisfying the property (i) and (ii) when $r < p$; otherwise, when $r = p$, we must have $\mtheta^*(i) = 1$ due to the $\ell_1$ norm constraint and we can not satisfy the property (ii).

   Combining the inequalities (39) and (40) in the proof of Theorem 2 in \cite{gao2018community}, we have 
   \begin{align}
        & \inf_{\hat z_1 } \bbE \left[ \ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right] \geq  \\
        & \quad  \frac{C}{r^3 |T_1^c|} \sum_{i \in T_1^c} \inf_{\hat z_1  (i)} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, z_k^*, \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, z_k^*, \mtheta_k^*, \tS^*] \},
   \end{align}
   where $C$ is some positive constant,  $\hat z_1$ on the left hand side denote the generic assignment functions in $\tP(\gamma)$, and the infimum on the right hand side is taken over the generic assignment function family of $\hat z_1(i)$ for all nodes $i \in T_1^c$. Here, the factor $r^3=r\cdot r^2$ in~\eqref{eq:inf_lower} comes from two sources: $r^2\asymp {r\choose 2}$ comes from the multiple testing burden for all pairwise comparisons among $r$ clusters; and another $r$ comes from the number of elements $|T^c_k|\asymp p/r$ to be clustered.

   Next, we need to find the lower bound of the rightmost side in~\eqref{eq:inf_lower}. 
   %For simplicity, we show the bound for the mode-1 case $k = 1$ only. We drop the subscripts $1$ in $z_1, T_1, \mS_1, \mtheta_1$ and omit the repeated procedures for the cases of $k = 2,\ldots, K$. 
We consider the hypothesis test based on model~\eqref{eq:general_dtbm}. First, we reparameterize the model under the construction (i)-(ii).
\begin{equation}
    \mx_a^* = \off{\mat_1 \of{ \tS^*\times_2 \mTheta_2^* \mM_2^*\times_3\cdots \times_K \mTheta^*_K \mM_K^* }}_{a:}, \quad \text{for all} \ a \in [r],
\end{equation}
where $\mx_a^*$'s are centroids in $\bbR^{p^{K-1}}$. Without loss of generality, we consider the lower bound for the summand in~\eqref{eq:inf_lower} for $i=1$. The analysis for other $i\in T^c_1$ are similar. For notational simplicity, we suppress the subscript $i$ and write $\my, \theta^*, z$ in place of $\my_1, \mtheta_1^*(1)$ and $z_1(1)$, respectively. The equivalent vector problem for assessing the summand in~\eqref{eq:inf_lower} is
\begin{equation}
\my=\theta^* \mx_{z}^*+\me,
\end{equation}
where $z\in \{1,2\}$ is an unknown parameter, $\theta^* \in \bbR_+$ is the given heterogeneity degree, $\mx_1^*,\mx_2^*\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.  Then, we consider the hypothesis testing under the model~\eqref{eq:z}:
\begin{equation}
 H_0: z = 1 \text{ and } \my = \theta^* \mx_1^* + \me \quad \leftrightarrow \quad H_1: z = 2 \text{ and } \my = \theta^* \mx_2^* + \me,
\end{equation}

%   For an arbitrary $i \in T_1^c$, the right hand side in \eqref{eq:inf_lower} corresponds to the following hypothesis test:
%   \begin{equation}\label{eq:hypo}
%       H_0: \my = \mtheta_1^*(i) \mx_1^* + \me \quad \leftrightarrow \quad H_1: \my = \mtheta_1^*(i) \mx_2^* + \me,
%   \end{equation}
%   where $\my = [\mat_1(\tY)]_{i:}$, $\mx_a^* = [\mat_1( \tS^* \times_2 \mTheta^* \mM_2^* \times_3 \cdots \times_K \mTheta^* \mM_K^*)]_{a:}$, and $\me$ consists of i.i.d.\ $N(0,\sigma^2)$ entries. 
   The hypothesis testing~\eqref{eq:test} is a simple versus simple testing, since the assignment $z$ is the only unknown parameter in the test. 
   By Neyman-Pearson lemma, the likelihood ratio test is optimal with minimal Type I + II error. Under Gaussian model, the likelihood ratio test of \eqref{eq:test} is equivalent to the least square estimator $\hat z_{LS} = \argmin_{a = \{1,2\}} \onormSize{}{\my - \theta^* \mx_a^*}_F^2$.

   Let $\mS = \mat_1(\tS)$. Note that 
   \begin{align}
       \onormSize{}{ \theta^* \mx_1^*  - \theta^* \mx_2^*}_F &\leq  \theta^* \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*)  \\
       & \leq \theta^* \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F   \max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}_2^{K-1} \\
       & \leq  \sigma r^{(K-1)/2} p^{-(K-1)/2} 2 c_4 p^{(K-1)/2} r^{-(K-1)/2}\\
       & \leq 2 c_4 \sigma, 
   \end{align}
%   \begin{align}
%       \onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F &\leq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*)  \\
%       & \leq \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F   \max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}_2^{K-1} \\
%       & \leq  \sigma r^{(K-1)/2} p^{-(K-1)/2} 2 c_4 p^{(K-1)/2} r^{-(K-1)/2}\\
%       & \leq 2 c_4 \sigma, \label{eq:bound}
%   \end{align}
   where $\lambda_{\max}$ denote the maximal singular value, the second inequality follows from Lemma~\ref{lem:singular_thetam}, and the third inequality follows from property (ii) and the boundedness constraint in $\tP_{\tS}(\gamma)$ such that $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F  \leq \onormSize{}{\mS_{1:}^*}_F + \onormSize{}{\mS_{2:}^*}_F \leq 2c_4$.

   Hence, we have 
   \begin{align}
       & \inf_{\hat z_1(1)} \{ \bbP[\hat z_1(1) = 1| z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(1) = 2| z_1^*(1) = 1, z^*_k,  \mtheta_k^*, \tS^*] \} \\
       & \quad = 2 \bbP[ \hat z_{LS} = 1 |  z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^* ] \\
      & \quad  = 2 \bbP[ \onormSize{}{\my - \theta^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \theta^* \mx_2^*}_F^2 | z_1^*(1) = 2, z^*_k , \mtheta_k^*, \tS^*   ] \\
      & \quad = 2 \bbP \off{ 2 \ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \geq  \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F^2 }\\
       & \quad =  2 \bbP[ N(0,1) \geq \theta^* \onormSize{}{\mx_1^* - \mx_2^*}_F /(2\sigma) ] \\
       & \quad \geq 2 \bbP[ N(0,1) \geq c_4 ]   \geq c,
   \end{align}

where the first equation holds by symmetry, the third equation holds by rearrangement, the fourth equation holds by the fact that $\ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \sim N(0, \sigma \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F)$, and $c$ is some positive constant in the last inequality.
   %for some positive constant $c$, where the calculation for the last equation follows from the calculation in equation (102) in \cite{han2020exact}.

   Plugging the inequality~\eqref{eq:inf} into the inequality~\eqref{eq:inf_lower} for all $i \in T_1^c$, then, we have 
   \begin{equation}
       \liminf_{p \rightarrow \infty}  \inf_{\hat z_1 } \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right]  \geq \liminf_{p \rightarrow \infty} \frac{ C cp }{r^3} \geq C c,
   \end{equation}
   where the last inequality follows by the condition $r = o(p^{1/3})$. By the discrete nature of the misclassification error, we obtain our conclusion
   \begin{equation}
       \liminf_{p \rightarrow \infty} \inf_{\tS^* \in  \tP_{\tS}(\gamma)}  \inf_{\hat z_{\text{stat}} }\sup_{ (z^*, \mtheta^*) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]  \geq 1. 
   \end{equation}
"
   \end{proof}
    \end{quote}

    % In fact, our proof idea of minimax lower bound statements (impossibility in Theorems~\ref{thm:stats} and \ref{thm:comp}) shares the same spirit with reviewer's comment. We briefly describe the proof idea of impossibility below:

    % Let $\tP_z$ denote the space for all possible (polynomial-time) community assignment estimation $\hat z$ and $\tP(\gamma) = \tP \cap \{ \Delta_{\rm ang}^2/\sigma^2 = p^{\gamma}\}$ denote the dTBM parameter space with SNR condition. We prove the minimax lower bound by constructing fixed true parameters $(z^*, \mtheta^*, \tS^*) \in \tP(\gamma)$ such that 
    % \begin{equation}\label{eq:impossible_prove}
    %     \inf_{\hat z \in \tP_z} \sup_{(z, \mtheta, \tS) \in \tP(\gamma)} \bbE[p \ell(\hat z, z)] \geq  \inf_{\hat z \in \tP_z} \bbE[p \ell(\hat z, z^*)| (z^*, \mtheta^*, \tS^*)] \geq 1.
    % \end{equation}
    % Here are two details differ with reviewer's comment. First, the SNR condition in $\tP(\gamma)$ is $\{ \Delta_{\rm ang}^2/\sigma^2 = p^{\gamma}\}$ rather than $\{ \Delta_{\rm ang}^2/\sigma^2 \leq p^{\gamma}\}$; otherwise, for any $\gamma \leq 0$, we can always take $\tS^* \in \{\Delta_{\rm ang}^2 = 0\}$ to make dTBM non-identifiable and thereof theoretically impossible. Second, we consider one particular core tensor $\tS^*$ rather than the random tensor with entries $\tS_{ijk} = 1 + \epsilon_{ijk}, \epsilon_{ijk} \sim N(0, \delta^2)$ for some $\delta \rightarrow 0$ as $p \rightarrow \infty$. The expectation in minimax error~\eqref{eq:impossible_prove} is taken with respect to the noise in observation, rather than the randomness in the true parameters.

    % The construction of $(z^*, \mtheta^*, \tS^*) \in \tP(\gamma)$ is the key step in the proof. With a good construction, the second inequality in~\eqref{eq:impossible_prove} can be easily verified with prior works. For statistical lower bound, we take one $(z^*, \mtheta^*, \tS^*)$ with $\mtheta^* = \mathbf{1}$ and $\Delta_{\rm Euc}^2(\tS^*) = p^{\gamma}$; for computational lower bound, we construct a rank-2 $\tS^*$ to relate the HPC conjecture. The inequality~\eqref{eq:impossible_prove} then holds with \citet[Proof of Theorems 6 and 7]{han2020exact}.

    % From the perspective of proof, we agree that the minimax lower bound statements can be obtained by exploiting prior results via proof idea~\eqref{eq:impossible_prove}. However, we argue that \textbf{none of our lower bound results are corollaries of \cite{han2020exact} and dTBM-specific techniques are required.} More importantly, we want to emphasize that the proof of dTBM achievability is more difficult than that for TBM due to the extra heterogeneity.

    % In the revision, we have proved a stronger lower bound statement for statistical impossibility as following comment suggested and added more discussion in Section~\ref{sec:tbm}.

    % Define $\tP_z \coloneqq \offf{ z:  c_1 p/r \leq |z^{-1}(a)| \leq c_2 p/r, a \in [r]}$, where $c_1, c_2$ are the positive constants in dTBM parameter space~\eqref{eq:family}. The space $\tP_z$ collects all possible community assignment $z$ under the dTBM or TBM. Let $\tP_{\rm dTBM}(\gamma) \coloneqq \tP_{\rm dTBM} \cap \{\Delta_{\rm ang}^2 = p^{\gamma}\}$ refer to the space $\tP(\gamma)$ in our main text. We prove the minimax lower bound by constructing fixed true parameters $(z^*, \mtheta^*, \tS^*) \in \tP_{\rm dTBM}(\gamma)$ such that 
    % \begin{equation}
    %     \inf_{\hat z \in \tP_z} \sup_{(z, \mtheta, \tS) \in \tP_{\rm dTBM}(\gamma)} \bbE[p \ell(\hat z, z)] \geq  \inf_{\hat z \in \tP_z} \bbE[p \ell(\hat z, z^*)| (z^*, \mtheta^*, \tS^*)] \geq 1.
    % \end{equation}

    % Let $\tP_{\rm TBM}(\gamma) \coloneqq \tP_{\rm TBM} \cap \{\Delta_{\rm Euc}^2 = p^{\gamma}\}$ and $\tP_{\rm dTBM}(\gamma) \coloneqq \tP_{\rm dTBM} \cap \{\Delta_{\rm ang}^2 = p^{\gamma}\}$ denote the parameter space for TBM and dTBM with SNR requirement, where $\tP_{\rm TBM}(\gamma)$ refers to the equation (26) in \cite{han2020exact} and $\tP_{\rm dTBM}(\gamma)$ refers to the space $\tP(\gamma)$ in our main text.

%     \textbf{Response:} We thank reviewer for the suggestions. Inspired by the reviewer, we have proved a stronger statement for statistical impossibility. The new statement suggests the dTBM statistical impossibility whenever the core tensor $\tS$ leads to a week SNR with exponent $\gamma < -(K-1)$. The proof of our new statement relies on special construction of heterogeneity $\mtheta^*$ in the worst case, and previous conclusions for TBM is non-applicable under the construction. We attach our new statement in Theorem~\ref{thm:stats}, Section~\ref{sec:limits}:

%     For self-consistency, we attach the proof of the new impossibility result in Section~\ref{sec:statprove1}:

%     \begin{quote}
%       ``
%       \begin{proof}[Proof of Theorem~\ref{thm:stats} (Impossibility)]Consider the general asymetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$. For simplicity, we show the minimax rate for the estimation on the first mode $\hat z_1$; the proof for other modes are essentially the same.

%   To prove the minimax rate~\eqref{eq:minminmax}, it suffices to take an arbitrary $\tS^* \in  \tP_{\tS}(\gamma)$ wih $\gamma < -(K-1)$ and construct $(z^*_k, \mtheta^*_k)$ such that 
%   \begin{equation}
%       \inf_{\hat z_1} \bbE \left[ p\ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right]\geq 1.
%   \end{equation}

%   We first define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation. Based on \citet[Proof of Theorem 6]{han2020exact}, we consider the restricted family of $\hat z_k$'s for which the following three conditions are satisfied:
%   \begin{align}
%         &\text{(a)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \quad \text{(b)} \ |T^c_k|\asymp {p\over r}; \\
%         &\text{(c)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
%   \end{align}
% for all $k \in [K]$.
%   Now, we consider the construction:
%   \begin{enumerate}
%       \item[(i)] $\{z_k^*\}$ satisfies properties (a)-(c) with misclassification sets $T_k^c$ for all $k \in [K]$;
%       \item [(ii)] $\{\mtheta_k^*\}$ such that $\mtheta_k^*(i) \leq \sigma r^{(K-1)/2} p^{-(K-1)/2}$ for all $i \in T_k^c, k \in [K]$ and $\max_{k \in [K], a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}^2_2 \asymp p/r$.
%   \end{enumerate}
%   % There exists a $(z^*, \mtheta^*) \in \tP_{z, \mtheta}$ satisfying the property (i) and (ii) when $r < p$; otherwise, when $r = p$, we must have $\mtheta^*(i) = 1$ due to the $\ell_1$ norm constraint and we can not satisfy the property (ii).

%   Then, following the proof of \cite{gao2018community}, we have 
%   \begin{align}
%         & \inf_{\hat z_1 } \bbE \left[ \ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right] \geq  \\
%         &\frac{C}{r^3 |T_1^c|} \sum_{i \in T_1^c} \inf_{\hat z_1  (i)} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, z_k^*, \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, z_k^*, \mtheta_k^*, \tS^*] \},
%   \end{align}
%   where $C$ is some positive constant,  $\hat z_1$ on the left hand side denote the generic assignment functions in $\tP(\gamma)$, and the infimum on the right hand side is taken over the generic assignment function family of $\hat z_1(i)$ for all node $i \in T_1^c$. Here, the factor $r^3=r\cdot r^2$ in~\eqref{eq:inf_lower} comes from two sources: $r^2\asymp {r\choose 2}$ comes from the multiple testing burden for all pairwise comparisons among $r$ clusters; and another $r$ comes from the number of elements $|T^c_k|\asymp p/r$ to be clustered.

%   Next, we need to find the lower bound of the rightmost side in~\eqref{eq:inf_lower}. 
%   %For simplicity, we show the bound for the mode-1 case $k = 1$ only. We drop the subscripts $1$ in $z_1, T_1, \mS_1, \mtheta_1$ and omit the repeated procedures for the cases of $k = 2,\ldots, K$. 
% We consider the hypothesis test based on model~\eqref{eq:general_dtbm}. First, we reparameterize the model under the construction (i)-(ii).
% \begin{equation}
%     \mx_a^* = \off{\mat_1 \of{ \tS^*\times_2 \mTheta_2^* \mM_2^*\times_3\cdots \times_K \mTheta^* \mM_K^* }}_{a:}, \quad \text{for all} \ a \in [r],
% \end{equation}
% where $\mx_a^*$'s are centroids in $\bbR^{p^{K-1}}$. Without loss of generality, we consider the lower bound for the summand in~\eqref{eq:inf_lower} for $i=1$. The analysis for other $i\in T^c_1$ are similar. For notational simplicity, we suppress the subscript $i$ and write $\my, \theta^*, z$ in place of $\my_1, \mtheta_1^*(1)$ and $z_1(1)$, respectively. The equivalent vector problem for assessing the summand in~\eqref{eq:inf_lower} is
% \begin{equation}
% \my=\theta^* \mx_{z}^*+\me,
% \end{equation}
% where $z\in \{1,2\}$ is an unknown parameter, $\theta^* \in \bbR_+$ is the given heterogeneity degree, $\mx_1^*,\mx_2^*\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.  Then, we consider the hypothesis testing under the model~\eqref{eq:z}:
% \begin{equation}
%  H_0: z = 1 \text{ and } \my = \theta^* \mx_1^* + \me \quad \leftrightarrow \quad H_1: z = 2 \text{ and } \my = \theta^* \mx_2^* + \me,
% \end{equation}

% %   For an arbitrary $i \in T_1^c$, the right hand side in \eqref{eq:inf_lower} corresponds to the following hypothesis test:
% %   \begin{equation}\label{eq:hypo}
% %       H_0: \my = \mtheta_1^*(i) \mx_1^* + \me \quad \leftrightarrow \quad H_1: \my = \mtheta_1^*(i) \mx_2^* + \me,
% %   \end{equation}
% %   where $\my = [\mat_1(\tY)]_{i:}$, $\mx_a^* = [\mat_1( \tS^* \times_2 \mTheta^* \mM_2^* \times_3 \cdots \times_K \mTheta^* \mM_K^*)]_{a:}$, and $\me$ consists of i.i.d.\ $N(0,\sigma^2)$ entries. 
%   The hypothesis test~\eqref{eq:test} is a simple versus simple testing, since the assignment $z$ is the only unknown parameter in the test. 
%   By Neyman-Pearson lemma, the likelihood ratio test is optimal with minimal Type I + II error. Under Gaussian model, the likelihood ratio test is equivalent to the least square estimator $\hat z_{LS} = \argmin_{a = \{1,2\}} \onormSize{}{\my - \theta^* \mx_a^*}_F^2$.

%   Let $\mS = \mat_1(\tS)$. Note that 
%   \begin{align}
%       \onormSize{}{ \theta^* \mx_1^*  - \theta^* \mx_2^*}_F &\leq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*)  \\
%       & \leq \theta^* \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F   \max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}_2^{K-1} \\
%       & \leq  \sigma r^{(K-1)/2} p^{-(K-1)/2} 2 c_4 p^{(K-1)/2} r^{-(K-1)/2}\\
%       & \leq 2 c_4 \sigma, 
%   \end{align}
% %   \begin{align}
% %       \onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F &\leq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*)  \\
% %       & \leq \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F   \max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}_2^{K-1} \\
% %       & \leq  \sigma r^{(K-1)/2} p^{-(K-1)/2} 2 c_4 p^{(K-1)/2} r^{-(K-1)/2}\\
% %       & \leq 2 c_4 \sigma, \label{eq:bound}
% %   \end{align}
%   where the second inequality follows Lemma~\ref{lem:singular_thetam} and the third inequality follows from property (ii) and the boundedness constraint in $\tP_{\tS}(\gamma)$ such that $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F  \leq \onormSize{}{\mS_{1:}^*}_F + \onormSize{}{\mS_{2:}^*}_F \leq 2c_4$.

%   Hence, we have 
%   \begin{align}
%       & \inf_{\hat z_1(1)} \{ \bbP[\hat z_1(1) = 1| z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(1) = 2| z_1^*(1) = 1, z^*_k,  \mtheta_k^*, \tS^*] \} \\
%       & \quad = 2 \bbP[ \hat z_{LS} = 1 |  z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^* ] \\
%       & \quad  = 2 \bbP[ \onormSize{}{\my - \theta^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \theta^* \mx_2^*}_F^2 | z_1^*(1) = 2, z^*_k , \mtheta_k^*, \tS^*   ] \\
%       & \quad = 2 \bbP \off{ 2 \ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \geq  \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F^2 }\\
%       & \quad =  2 \bbP[ N(0,1) \geq \theta^* \onormSize{}{\mx_1^* - \mx_2^*}_F /(2\sigma) ] \\
%       & \quad \geq 2 \bbP[ N(0,1) \geq c_4 ]   \geq c,
%   \end{align}
% %   \begin{align}
% %       & \inf_{\hat z_1(1)} \{ \bbP[\hat z_1(1) = 1| z_1^*(i) = 2, z^*_k,  \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(1) = 2| z_1^*(i) = 1, z^*_k,  \mtheta_k^*, \tS^*] \} \\
% %       & \quad = 2 \bbP[ \hat z_{LS}(i) = 1 |  z_1^*(i) = 2, z^*_k,  \mtheta_k^*, \tS^* ] \\
% %       & \quad  = 2 \bbP[ \onormSize{}{\my - \mtheta_1(i)^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \mtheta_1(i)^* \mx_2^*}_F^2 | z_1^*(i) = 2, z^*_k , \mtheta_k^*, \tS^*   ] \\
% %       & \quad =  2 \bbP[ N(0,1) \geq \mtheta_1(i)^* \onormSize{}{\mx_1^* - \mx_2^*}_F /\sigma ] \\
% %       & \quad \geq 2 \bbP[ N(0,1) \geq 2 c_4 ]   \geq c, \label{eq:inf}
% %   \end{align}
% where the first equation holds by symmetry, the third equation holds by rearrangement, the fourth equation holds by the fact that $\ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \sim N(0, \sigma \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F)$, and $c$ is some positive constant in the last inequality.
%   %for some positive constant $c$, where the calculation for the last equation follows from the calculation in equation (102) in \cite{han2020exact}.

%   Plugging the inequality~\eqref{eq:inf} to inequality~\eqref{eq:inf_lower}, then, we have 
%   \begin{equation}
%       \liminf_{p \rightarrow \infty}  \inf_{\hat z_1 } \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right]  \geq \liminf_{p \rightarrow \infty} \frac{ C cp }{r^3} \geq C c,
%   \end{equation}
%   where the last inequality follows by the condition $r = o(p^{1/3})$. By the discrete nature of the misclassification error, we have 
%   \begin{equation}
%       \liminf_{p \rightarrow \infty} \inf_{\tS^* \in  \tP_{\tS}(\gamma)}  \inf_{\hat z_{\text{stat}} }\sup_{ (z^*, \mtheta^*) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]  \geq 1. 
%   \end{equation}
% %   for any $r < p$ and $p$,  we have 
% %   \begin{equation}
% %         \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq \frac{ cp }{r^3} > 0,
% %   \end{equation}
% %   which indicates 
% %   \begin{equation}
% %         \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq 1, \quad \text{for every } p
% %   \end{equation}
% %   by the discrete nature of $p\ell(\hat z, z^*)$.
%   \end{proof} "
%     \end{quote}

    \item \textit{Finally, I checked the author’s proof towards the statement. I think the proofs need to be revised substantially as I found several flaws:}

    \textbf{Response:} We thank reviewer for pointing out these issues. We have provided a new proof of Theorem~\ref{thm:stats} for the new statistical impossibility statement. We will reply to each issue based on the related parts in the new proof.

    %Since we have proved a new stronger statement for statistical impossibility, most of following concerns no longer exist in the new proof of Theorem~\ref{thm:stats}. We still thank reviewer for pointing out these concerns and answer each sub-question for clarification.

    \begin{itemize}
        \item \textit{Page 32, Eqn (31) suggests that $\theta$ is fixed and the decision problems are reduced to $z$ from that point. But in latter argument $\theta$ is taken as unknown parameter.}

        \textbf{Response:} We agree that the inconsistent status of $\mtheta$ is a flaw in previous proof of Theorem~\ref{thm:stats}. Once we fix the true $\mtheta$ at the beginning (Eqn (31) of previous proof), we should take true $\mtheta$ as a known parameter in the rest of the proof. Specifically, in previous Eqn (33), the model of $\my$
        \begin{equation}
            \my = \theta \mx_z + \me
        \end{equation}
        should take $\theta$ as a known parameter. Because the distribution of observation $\my$ corresponds to the conditional probability $\bbP[\cdot | z_k(i)]$ in previous Eqn (31). When the true assignment $z_k(i)$ is known, the distribution of $\my$ should be known with true heterogeneity $\theta$ and true centroid $\mx_{z_k(i)}$. To avoid this flaw in the new proof, we use new notations $z^*_k, \mtheta^*_k, \tS^*_k$ for true parameters to better distinguish the known and unknown parameters. We also re-write the conditional probability as $\bbP[\cdot | z^*_k(i), z_k^*, \mtheta^*_k, \tS^*]$ for better expression.  We attach the related parts in the new proof in Appendix~\ref{sec:statprove1} here:

        \begin{quote}
            ``...
            To prove the minimax rate~\eqref{eq:minminmax}, it suffices to take an arbitrary $\tS^* \in  \tP_{\tS}(\gamma)$ wih $\gamma < -(K-1)$ and construct $(z^*_k, \mtheta^*_k)$ such that 
   \begin{equation}
       \inf_{\hat z_1} \bbE \left[ p\ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right]\geq 1.
   \end{equation}
   ..."
        \end{quote}

        \begin{quote}
            ``... we have 
   \begin{align}
        & \inf_{\hat z_1 } \bbE \left[ \ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right] \geq  \\
        & \frac{C}{r^3 |T_1^c|} \sum_{i \in T_1^c} \inf_{\hat z_1  (i)} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, z_k^*, \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, z_k^*, \mtheta_k^*, \tS^*] \},
   \end{align}
    where $C$ is some positive constant, ..."
        \end{quote}

        \begin{quote}
            `` ...  The equivalent vector problem for assessing the summand in~\eqref{eq:inf_lower} is
\begin{equation}
\my=\theta^* \mx_{z}^*+\me,
\end{equation}
where $z\in \{1,2\}$ is an unknown parameter, $\theta^* \in \bbR_+$ is the given heterogeneity degree, $\mx_1^*,\mx_2^*\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.  ..."
        \end{quote}

        \item \textit{Page 33, Line 9: let’s say the authors may want to make $\theta$ unknown and involve it in the decision problem. The authors formulate a hypothesis testing for model (33). My question is, is this really a hypotheses testing? Eqn. (30) involves both $z$ and $\theta$ as unknown parameters. Under null ($z = 1$), the distribution of $y$ is still unknown.}

        \textbf{Response:} This is a follow-up question of last comment. We agree that involving $\mtheta$ as an unknown parameter in the model of $\my$ and hypothesis testing is in appropriate. Following our last response, we consider the true heterogeneity $\mtheta^*$ as known in the model of $\my$. Hence, we have a valid simple hypothesis testing problem, in which the assignment function $z$ is the only unknown parameter in the test. Under the null hypothesis $(z = 1)$, the observation $\my = \theta^* \mx_1^* + \me$ follows multivariate Gaussian distribution with known mean vector $\theta^* \mx_1^*$ and covariance matrix $\sigma^2 \mI$. For clarification, we re-write the hypothesis testing with the model of $\my$. We attach the related parts in the new proof in Appendix~\ref{sec:statprove1} here:

        \begin{quote}
            ``...
            The equivalent vector problem for assessing the summand in~\eqref{eq:inf_lower} is
\begin{equation}
\my=\theta^* \mx_{z}^*+\me,
\end{equation}
where $z\in \{1,2\}$ is an unknown parameter, $\theta^* \in \bbR_+$ is the given heterogeneity degree, $\mx_1^*,\mx_2^*\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.  Then, we consider the hypothesis testing under the model~\eqref{eq:z}:
\begin{equation}
 H_0: z = 1 \text{ and } \my = \theta^* \mx_1^* + \me \quad \leftrightarrow \quad H_1: z = 2 \text{ and } \my = \theta^* \mx_2^* + \me,
\end{equation}
...
            "
        \end{quote}

        % Based on last sub-question, we agree that the $\theta$ should be considered as a known parameter in model (33). The hypothesis testing from model (33) is equivalent to
        % \begin{equation}\label{eq:hypo_revise}
        %     H_0: \my = \theta \mx_1 + \me \quad \text{v.s.} \quad H_1: \my = \theta \mx_2 + \me,
        % \end{equation}
        % where $\theta, \mx_1, \mx_2$ are known parameters, $\me$ consists of i.i.d.\ $N(0, \sigma^2)$ entries. Now, the null and alternative distri-0bution of $\my$ is known, and the test~\eqref{eq:hypo_revise} is a valid hypothesis testing problem.

        \item \textit{A follow-up question is why could you apply Neyman-Pearson Lemma to lower bound the decision risks? Apparently this is not a simple testing, and the authors should provide the justification on why the presented “profiled MLE” works (technically) in determining the lower bound.}

        \textbf{Response:} The revised hypothesis testing in new Eqn~\eqref{eq:test} is a simple $(z = 1)$ versus simple $(z = 2)$ testing problem. Because the assignment function $z$ is the only unknown parameter in test. Therefore, we are able to apply Neyman-Pearson Lemma to our testing problem. By Neyman-Pearson Lemma, the likelihood ratio test, which coincides to the least square estimator of $z$ under our Gaussian model, leads to the minimal Type I + II error. As $\theta^*$ is a known parameter in the new proof, we no longer use the ``profile MLE" to determine the lower bound. We attach the related parts in the new proof in Appendix~\ref{sec:statprove1} here:

        \begin{quote}
            ``
            ... The hypothesis testing~\eqref{eq:test} is a simple versus simple testing, since the assignment $z$ is the only unknown parameter in the test. By Neyman-Pearson lemma, the likelihood ratio test is optimal with minimal Type I + II error. Under Gaussian model, the likelihood ratio test of \eqref{eq:test} is equivalent to the least square estimator $\hat z_{LS} = \argmin_{a = \{1,2\}} \onormSize{}{\my - \theta^* \mx_a^*}_F^2$.  ...
            "
        \end{quote}

        \item \textit{Although I don’t understand why $\hat \theta_{\rm MLE}$ is introduced here, I feel its calculation is incorrect. Does the authors consider the fact that there is a pre-specified parameter space for $\theta$ and it cannot take all values on $\bbR^+$ ?}

        \textbf{Response:} In the new proof, we no longer introduce the estimator $\hat \theta_{\rm MLE}$ and $\hat z_{\rm MLE}$ defined in the previous proof. Instead, we consider the least square estimator by Neyman-Pearson Lemma. In general, we agree that the MLE of $(z, \mtheta, \tS)$ should lie in the pre-specified parameter space $\tP$ to guarantee the identifiability. We emphasize this point in the formal definition of MLE~\eqref{eq:mle} in the main text. We attach the definition in Section~\ref{sec:limits} here:
        \begin{quote}
            ``
            We consider the Gaussian MLE, denoted as $(\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}})$, over the estimation space $\tP$, where 
\begin{equation}
    (\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}}) = \argmin_{ (z, \mtheta, \tS) \in \tP} \onormSize{}{\tY - \tX(z,\mtheta, \tS)}_F^2.
\end{equation}
...
            "
        \end{quote}

    \end{itemize}
    \end{enumerate}

\end{enumerate}

\newpage
\begin{center}
    \textbf{Point-by-point response to Reviewer 2}
\end{center}

\textit{I thank the authors for the signiﬁcant revision that clariﬁes the issues that I raised. I still suggest the authors to emphasise early that this paper focuses on dense regime. For instance, they could include sparsity in Table 1 and mention that Ahn et al (2018) proves exact recovery in the sparse regime (which this paper and few others in the list don't). This would provide a clear (unbiased) comparison of the results.}

\textbf{Response:} We thank reviewer for the reference. We have included the reference \cite{ahn2018hypergraph} in Table~\ref{tab:comp}:

\begin{quote}
    \begin{table*}[ht]
\resizebox{\textwidth}{!}{%
    \begin{tabular}{c|cccccc}
\hline
    & \cite{gao2018community}&  \cite{ahn2018hypergraph} &\cite{han2020exact}&  \cite{ghoshdastidar2017consistency} &\cite{ke2019community} & \textbf{Ours}\\
    \hline
      Allow tensors of arbitrary order & $\times$ & $\surd$&  $\surd$ & $\surd$ &$\surd$ & $\surd$  \\
        Allow degree heterogeneity &  $\surd$ & $\times$ & $\times$ & $\surd$ &$\surd$ & $\surd$ \\
        Singular-value gap-free clustering &  $\surd$ & $\surd$ & $\surd$ & $\times$ & $\times$ &$\surd$ \\
        Misclustering rate (for order $K^*$)& - & ${p^{-(K-1)} \alpha^{-1}_p}^{**}$ & $\exp(-p^{K/2})$ & $p^{-1}$& $p^{-2}$ &  $\exp(-p^{K/2})$\\
        Consider sparse observation & $\times$ & $\surd$ & $\times$ & $\times$ & $\times$ &$\times$\\
        \hline
    \end{tabular}
    }
    \caption{ Comparison between previous methods with our method. $^*$We list the result for order-K tensors with $K \geq 3$ and general number of communities $r = \tO(1)$. $^{**}$The parameter $\alpha = f(p) > 0$ denotes the sparsity level which is some function of dimension $p$. }
\end{table*}
\end{quote}

We have added a brief clarification in the related work, Section~\ref{sec:intro}:
\begin{quote}
    ``... Some works~\citep{ahn2018hypergraph} study the TBM with sparse observations, while, others~\citep{wang2019multiway, han2020exact} and our work focus on the dense regime. ..."
\end{quote}

We also added the reference in the extension to Bernoulli clustering, Section~\ref{subsec:ber}:
\begin{quote}
    ``The sparsity is often a popular feature in hypergraphs~\citep{florescu2016spectral,ke2019community, ahn2018hypergraph}.
    "
\end{quote}

\newpage
% \section{Notes}

% \textbf{Claim:} The statistical impossibility in dTBM can be easily obtained from the TBM paper.

%     \textcolor{blue}{ Note that ``easily obtained" is not ``directly obtained". See following arguments.}

%     Let $\tP_{\rm dTBM}$ and $\tP_{\rm TBM}$ denote the true parameter spaces for dTBM and TBM models. There is no belonging relation between $\tP_{\rm dTBM}$ and $\tP_{\rm TBM}$. Examples: if $\mtheta \neq \mathbf{1}$, then we have $(z, \mtheta, \tS) \in \tP_{\rm dTBM}$ but $(z, \mtheta, \tS) \notin \tP_{\rm TBM}$; if $\tS$ has $\onormSize{}{\mS_{a:}}_F = 0$ and $\mtheta = \mathbf{1}$, then we have $(z, \mtheta, \tS) \in \tP_{\rm TBM}$ but $(z, \mtheta, \tS) \notin \tP_{\rm dTBM}$. Also, there is no belonging relation between $\tP_{\rm dTBM}(\gamma) \coloneqq \tP_{\rm dTBM} \cap \{ \Delta_{\rm ang}^2 \lesssim p^{\gamma} \}$ and $\tP_{\rm TBM}(\gamma)  \coloneqq \tP_{\rm TBM} \cap \{ \Delta_{\rm Euc}^2 \lesssim p^{\gamma} \}$.

%     Hence, for any estimator $\hat z$, we can not infer the worst case error from one case to another case; i.e.,
%     \begin{equation}
%         \sup_{(z, \mtheta, \tS) \in \tP_{\rm dTBM}(\gamma)} p\ell(\hat z, z)  >  1 \quad \nRightarrow \text{and} \nLeftarrow \quad  \sup_{(z, \mtheta, \tS) \in \tP_{\rm TBM}(\gamma)} p\ell(\hat z, z) > 1.
%     \end{equation}

%     That's why we can not ``\emph{directly} obtain" the dTBM impossibility from TBM impossibility, as previous Remark 1 said. But we can ``\emph{easily}" obtain dTBM result by using the proof of TBM impossibility.

%     Note that there exists a set of parameters $(z^*, \mtheta^*, \tS^*)$ with $\mtheta^* = \mathbf{1}$ such that $(z^*, \mtheta^*, \tS^*) \in \tP_{\rm dTBM}(\gamma) \cap \tP_{\rm TBM}(\gamma)$. In the proof of TBM impossibility in \cite{han2020exact}, we eliminate the supreme by considering the special case with $(z^*, \mtheta^*, \tS^*)$ 
%     \begin{equation}
%         \inf_{\hat z: (\hat z, \hat \mtheta, \hat \tS) \in \tP_{\rm TBM}}  \sup_{(z, \mtheta, \tS) \in \tP_{\rm TBM}(\gamma)} p\ell(\hat z, z) \geq \inf_{\hat z: (\hat z, \hat \mtheta, \hat \tS) \in \tP_{\rm TBM}}    p\ell(\hat z, z^*)  > 1.
%     \end{equation}
%     Then, in dTBM, we also can eliminate the supreme by considering the same special case with $(z^*, \mtheta^*, \tS^*)$ 
%     \begin{equation}
%         \inf_{\hat z: (\hat z, \hat \mtheta, \hat \tS) \in \tP_{\rm dTBM}}  \sup_{(z, \mtheta, \tS) \in \tP_{\rm dTBM}(\gamma)} p\ell(\hat z, z) \geq \inf_{\hat z: (\hat z, \hat \mtheta, \hat \tS) \in \tP_{\rm dTBM}}    p\ell(\hat z, z^*)
%     \end{equation}
%     Notice that the infimum are taken under different regimes. We only need to show that 
%     \begin{equation}\label{eq:inf}
%         \inf_{\hat z: (\hat z, \hat \mtheta, \hat \tS) \in \tP_{\rm dTBM}}    p\ell(\hat z, z^*) = \inf_{\hat z: (\hat z, \hat \mtheta, \hat \tS) \in \tP_{\rm TBM}} p\ell(\hat z, z^*). 
%     \end{equation}
%     The equation~\eqref{eq:inf} is true. Because under $(z^*, \mtheta^*, \tS^*)$ with $\mtheta^* = \mathbf{1}$, it is equivalent to consider the UMP of the vector hypothesis test problem
%     \begin{equation}
%         H_0: \my = \mx_1 + \me \quad vs \quad H_1: \my = \mx_2 + \me,
%     \end{equation}
%     where $\mx_a$ are known centroids and $\me$ consists of i.i.d.\ $N(0,\sigma^2)$. By Neyman-Pearson Lemma, the UMP is the non-degree least square estimator $\hat z_{\rm UMP} = \argmin_{a \in \{1,2\}} \onormSize{}{\my - \mx_a}^2_F$, which is independent with the estimation $\hat \mtheta$. Then, we finish the proof
%     \begin{equation}
%         \inf_{\hat z: (\hat z, \hat \mtheta, \tS) \in \tP_{\rm dTBM}}    p\ell(\hat z, z^*) = p\ell(\hat z_{\rm UMP}, z^*) = \inf_{\hat z: (\hat z, \hat \mtheta, \tS) \in \tP_{\rm TBM}} p\ell(\hat z, z^*)  > 1
%     \end{equation}

%  \textbf{Claim:} The minimax errors over $\hat z$ and over $(\hat z, \hat \mtheta)$ are the same in our case; i.e.,
%     \begin{equation}
%         \inf_{\hat z: (\hat z, \hat \mtheta, \hat \tS) \in \tP_{\rm dTBM}} \sup_{(z^*, \mtheta^*, \tS^*) \in \tP_{\rm dTBM}(\gamma)} \ell(\hat z, z^*) =  \inf_{(\hat z, \hat \mtheta): (\hat z, \hat \mtheta, \hat \tS) \in \tP_{\rm dTBM}} \sup_{(z^*, \mtheta^*, \tS^*) \in \tP_{\rm dTBM}(\gamma)} \ell(\hat z, z^*).
%     \end{equation}

%     \textbf{Proof:} Note that the parts after the infimum are the same. Let $f(\hat z) = \sup_{(z^*, \mtheta^*, \tS^*) \in \tP_{\rm dTBM}(\gamma)} \ell(\hat z, z^*)$. Also note that for some parameters $x,y$, function $f$, and area $\mathbb{A}$ we use the notation
%     \begin{equation}
%         \{ x: f(x,y) \in \mathbb{A}\} \coloneqq \{ x: \text{there exists a } y \text{ such that } f(x,y) \in \mathbb{A}\} = \{x: \exists y \text{ s.t.} f(x,y) \in \mathbb{A} \}.
%     \end{equation}
%     It is equivalent to consider
%     \begin{equation}
%         \inf_{\hat z: \exists (\hat \mtheta, \hat \tS) \text{ s.t.} (\hat z, \hat \mtheta, \hat \tS) \in \tP_{\rm dTBM}}  f(\hat z) \quad \text{and} \quad \inf_{(\hat z, \hat \mtheta): \exists \hat \tS \text{ s.t.}  (\hat z, \hat \mtheta, \hat \tS) \in \tP_{\rm dTBM}} f(\hat z).
%     \end{equation}
%     Then, we have
%     \begin{align}
%          \inf_{(\hat z, \hat \mtheta): \exists \hat \tS \text{ s.t.}  (\hat z, \hat \mtheta, \hat \tS) \in \tP_{\rm dTBM}} f(\hat z) &=  \inf_{\hat z': \exists (\hat \mtheta, \hat \tS) \text{ s.t.} (\hat z', \hat \mtheta, \hat \tS) \in \tP_{\rm dTBM}}  \inf_{  \hat \mtheta': \exists \hat \tS \text{ s.t.} (\hat z', \hat \mtheta', \hat \tS) \in  \tP_{\rm dTBM} } f(\hat z')  \\
%          & =  \inf_{\hat z': \exists (\hat \mtheta, \hat \tS) \text{ s.t.} (\hat z', \hat \mtheta, \hat \tS) \in \tP_{\rm dTBM}} f(\hat z'),
%     \end{align}
%     where the last equation follows from the fact that $f(\hat z)$ is independent with $\hat \mtheta$. Therefore, we have proven the claim.

% \textbf{Claim:} Let $(z^*, \mtheta^*, \tS^*)$ denote the true parameters in Eqn (31). The hypothesis test corresponds to Eqn (32) should be 
%     \begin{equation}
%         H_0: \my = \theta^* \mx_1^* + \me, \quad vs \quad H_1: \my = \theta^* \mx_2^* + \me,
%     \end{equation}
%     where $\mx_a^* = [\Mat_1(\tS^* \times_2 \mM_2^* \times_2 \cdots \times_K \mM_k^*)]_{a:}$.

%     \textbf{Proof:} In all impossibility proofs (Eqn(32) in our proof, \cite{han2020exact}, and \cite{gao2018community}), we want to lower bound the following probability
%     \begin{equation}\label{eq:hypo}
%         \inf_{\hat z \in \tP} \{ \bbP(\hat z = 1| z^* = 2) + \bbP(\hat z = 2|z^* = 1) \},
%     \end{equation}
%     where the $\tP$ is the estimation space of $\hat z$ and $z^*$ is the true assignment. In our dTBM, we have $\tP = \{\hat z: (\hat z, \hat \mtheta, \hat \tS) \in \tP_{\rm dTBM}\}$, and the $\bbP(\cdot | z^*)$ should refer to the distribution of observation $\my$ with true parameter $z^*, \mtheta^*, \tS^*$, denoted as $\bbP(\cdot | z^*, \mtheta^*, \tS^*)$. Then, the hypothesis test corresponds to the probability~\eqref{eq:hypo} is 
%     \begin{equation}
%         H_0: \my = \theta^* \mx^*_1 + \me \quad vs \quad H_1: \my = \theta^* \mx_2^* + \me.
%     \end{equation}
%     The hypothesis test should not be affected by the estimation procedure of $\hat z$.

%     If we write the hypothesis test with unknown $\theta$
%     \begin{equation}
%         H_0: \my = \theta \mx^*_1 + \me \quad vs \quad H_1: \my = \theta \mx_2^* + \me,
%     \end{equation}
%     then we do not have a valid null/alternative distribution. In addition, from the estimation perspective, parameter $\mx_a$ is also unknown in dTBM problem.

%     \textbf{Claim:} Both \cite{han2020exact} and \cite{gao2018community} proves minimax bound
%     \begin{equation}\label{eq:minimax}
%         \inf_{\hat z} \sup_{(z, \tS, \mtheta) \in \tP(\gamma)} \bbE[p \ell(\hat z, z)] \geq 1,
%     \end{equation}
%     rather than the minimin bound
%     \begin{equation}\label{eq:minimin}
%          \inf_{\tS \in \tP(\gamma)} \inf_{\hat z} \sup_{(z, \mtheta)} \bbE[p \ell(\hat z, z)] \geq 1.
%     \end{equation}

%     \textbf{Proof:}

%     - For TBM \cite{han2020exact}, the parameter space is defined as 
%     \begin{equation}\label{eq:space}
%         \tP_{\rm TBM}(\gamma) = \tP_{\rm TBM} \cap \{\Delta_{\rm Euc}^2 \gtrsim p^{\gamma}\}.
%     \end{equation}
%     To prove the minimax rate~\eqref{eq:minimax}, we consider the boundary case $\tS^*$ such that $\Delta_{\rm Euc}^2(\tS^*) = p^{\gamma}$.

%     For minimin bound~\eqref{eq:minimin}, with any $\gamma$, we can always have 
%     \begin{equation}
%          \inf_{\tS \in \tP_{\rm TBM}(\gamma)} \inf_{\hat z} \sup_{z} \bbE[p \ell(\hat z, z)] \leq \inf_{\hat z} \sup_{z^*} \bbE[p \ell(\hat z, z^*)] \leq 0,
%     \end{equation}
%     where the inequality holds with true parameter $(z^*, \tS^*)$ which lies in the strong signal regime to exact recovery.

%   - For DCBM \cite{gao2018community}, we have parameter space 
%   \begin{equation}\label{eq:space_dcbm}
%       \tP_{\rm DCBM}(\gamma) = \tP_{\rm DCBM} \cap \{ \min_u B_{uu} - \max_{u \neq v} B_{uv} \geq p-q = \gamma \},
%   \end{equation}
%   where $\min_u B_{uu} - \max_{u \neq v} B_{uv}$ denotes the minimal difference between the diagonal and off-diagonal elements in the mean matrix, and $p-q$ can be considered as the SNR in this case.

%   To prove the minimax rate~\eqref{eq:minimax}, \cite{gao2018community} considers the boundary case where $\min_u B^*_{uu} - \max_{u \neq v} B^*_{uv} = p^* - q^*$.

%   For minimin bound~\eqref{eq:minimin}, with any $\gamma \in [0,1]$, ww can always have 
%   \begin{equation}
%       \inf_{B \in \tP_{\rm DCBM}(\gamma)} \inf_{\hat z} \sup_{z, \mtheta} \bbE[p \ell(\hat z, z)]  \leq \inf_{\hat z} \sup_{z^*, \mtheta^*} \bbE[p \ell(\hat z, z^*)] \leq 0,
%   \end{equation}
%   where the inequality holds with true parameters $B^*$ such that $p^* - q^*$ lies in the strong signal regime.

%   \textit{Therefore, the minimin bound~\eqref{eq:minimin} is trivial or meaningless under the TBM or DCBM case with parameter spaces~\eqref{eq:space} or \eqref{eq:space_dcbm}.}

%   - In our case, we have parameter space 
%   \begin{equation}
%       \tP_{\rm dTBM}(\gamma) = \tP_{\rm dTBM} \cap \{ \Delta_{\rm ang }^2 = p^{\gamma}\},
%   \end{equation}
%   For the minimax rate~\eqref{eq:minimax}, we consider the boundary case $(z^*, S^*, \mtheta^*) \in \tP_{\rm dTBM}(\gamma) \cap \tP_{\rm TBM}(\gamma)$.

%   However, the minimin rate~\eqref{eq:minimin} is not trivial for dTBM since $\tP_{\rm dTBM}(\gamma)$ does not include the $\tS^*$ that leads to arbitrarily large signal. If we revise the space as $ \tP_{\rm dTBM}'(\gamma) = \tP_{\rm dTBM} \cap \{ \Delta_{\rm ang }^2 \geq p^{\gamma}\}$, then the minimin rate is trivial as TBM and DCBM. This definition revision from $\tP_{\rm dTBM}$ to $\tP'_{\rm dTBM}$ won't affect results in old submissions but lead to an easier defense against the minimin rate comment.

%   \newpage

%   \textbf{Minimax vs Minimin rate.} The reviewer mention the minimax and minimin rate under the dTBM:
%   \begin{align}
%       {\rm Minimax:} \quad  \inf_{\hat z} \sup_{(z, \theta, S) \in \tP(\gamma)} \bbE[p l(\hat z,  z)] \geq 1\quad \quad 
%       {\rm Minimin:} \quad \inf_{S \in P(\gamma)} \inf_{\hat z} \sup_{(z, \theta) \in \tP(\gamma)} \bbE[p l(\hat z,  z)]\geq 1.
%   \end{align}
%   To analyze the minimax and minimin rate rigorously, we firstly recall and define several spaces for the estimator and true parameters.
%   \begin{enumerate}[wide]
%       \item $\tP_{\rm dTBM}, \tP_{\rm TBM}, \tP_{\rm DCBM}$: general parameter spaces for our dTBM, TBM \cite{han2020exact}, and matrix DCBM \cite{gao2018community};
%       \item $\tP_{\rm dTBM} (\gamma) = \tP_{\rm dTBM} \cap \{ \Delta_{\rm ang}^2 = p^{\gamma} \}$: dTBM parameter space with angle gap equal to $p^{\gamma}$;
%       \item $\tP_{\rm TBM} (\gamma) = \tP_{\rm TBM} \cap \{ \Delta_{\rm Euc}^2 = p^{\gamma} \}$: TBM parameter space with Euclidean gap equal to $p^{\gamma}$;
%       \item $\tP_{\rm DCBM} (\gamma) = \tP_{\rm TBM} \cap \{ \min_u B_{uu} - \max_{u \neq v} B_{uv} = \gamma \}$: DCBM parameter space with diagonal-offdiagonal gap equal to $\gamma$;
%       \item $\tP'_{\rm dTBM}(\gamma) = \tP_{\rm dTBM} \cap \{ \Delta_{\rm ang}^2 \geq  p^{\gamma} \}$:  dTBM parameter space with angle gap equal or \textbf{larger} than $p^{\gamma}$; similar for $\tP'_{\rm TBM}(\gamma)$ and $\tP'_{\rm DCBM}(\gamma)$;
%       \item $\tP_{z} = \{ z: c_1 p/r |z^{-1}(a)| \leq c_2 p/r \}$: space for all possible assignment satisfying the group size balance assumption; we omit the subscript $k$ for simplicity; $\tP_z$ should be the estimation space for $\hat z$ in dTBM, TBM, and DCBM, because all three models share the same balance constraint on $z$. 
%       \item $\tP_{\mtheta, {\rm dTBM}} = \{ \mtheta: \text{there exists } (z, \tS) \text{ such that } (z, \mtheta, \tS) \in \tP_{\rm dTBM}\}$: the space for all $\mtheta$ satisfying the dTBM requirement; similar to replace the subscript by $\tS$ or $(z,\mtheta)$ and similar to define $\tP_{\mtheta, {\rm dTBM}}(\gamma)$ and with other models.
%     \end{enumerate}
%   Equipped with these spaces, we re-state the minimax and minimin rates rigorously:
%   \begin{align}
%       {\rm Minimax:}& \quad  \inf_{\hat z \in \tP_z} \sup_{(z, \mtheta, \tS) \in \tP_{\rm dTBM}(\gamma)} \bbE[p \ell(\hat z,  z)] \geq 1 \label{eq:minimax} \\ 
%       {\rm Minimin:}& \quad \inf_{\tS \in \tP_{\tS, {\rm dTBM}}(\gamma)} \inf_{\hat z \in \tP_z} \sup_{(z, \mtheta) \in \tP_{(z,\mtheta), \rm dTBM}(\gamma)} \bbE[p \ell(\hat z,  z)]\geq 1. \label{eq:minimin}
%   \end{align}
%   Moreover, the space $\tP(\gamma)$ under the supreme can be replaced by the space $\tP'(\gamma)$, and these rates also holds for TBM and DCBM.

%   The comparison between minimax and minimin rates depends on spaces related to the signal; the minimax rates are invariant to the choice of $\tP(\gamma)$ and $\tP'(\gamma)$ while the minimin results are quite different with different spaces. The minimax rates under all models have been studied, so we focus on the minimin results here.

%   \begin{itemize}
%   \item \textbf{Choose $\tP'(\gamma).$} In \cite{han2020exact} and \cite{gao2018community}, we are actually using $\tP'(\gamma)$ rather than $\tP(\gamma)$. The minimin lower bound becomes meaningless under this case for all dTBM, TBM, and DCBM. We take TBM for detailed illustration.

%   \textbf{Claim:} For any $\gamma \leq 0$, 
%   \begin{equation}
%       \inf_{\tS \in \tP'_{\tS, {\rm TBM}}(\gamma)} \inf_{\hat z \in \tP_z} \sup_{z \in \tP'_{z, \rm TBM}(\gamma)} \bbE[p \ell(\hat z,  z)] = 0.
%   \end{equation}

%   The intuition is that we need to consider all $\tS^*$ with large signal in $\tP'(\gamma)$ due to the first infimum.

%   \textbf{Proof:} For any $\gamma \leq 0$, we always have $\tS^* \in \tP'_{\tS, {\rm TBM}}(\gamma)$ such that $\Delta_{\rm Euc}^2 (\tS^*) = c$ for some positive constant $c$. Then, we have 
%   \begin{align}
%       \inf_{\tS \in \tP'_{\tS, {\rm TBM}}(\gamma)} \inf_{\hat z \in \tP_z} \sup_{z \in \tP'_{z, \rm TBM}(\gamma)} \bbE[p \ell(\hat z,  z)] & \leq  \inf_{\hat z \in \tP_z} \sup_{z \in \tP'_{z, \rm TBM}(\gamma)} \bbE[p \ell(\hat z, z)| \Delta_{\rm Euc}^2(\tS^*) = c ]  \\
%       & \leq \sup_{z^* \in \tP'_{z, \rm TBM}(\gamma)} \bbE[p \ell(\hat z_{\rm alg}, z^*)| \Delta_{\rm Euc}^2(\tS^*) = c ]\\
%       & \leq 0,
%   \end{align}
%   where $\hat z_{\rm alg}$ refers to the algorithm output in \cite{han2020exact}, and the last inequality holds since the algorithm guarantee Theorem 3 or 4 hold with arbitrary $z^* \in \tP_z$ and $\tP'_{z, \rm TBM}(\gamma) = \tP_z$.

%       \item \textbf{Choose $\tP(\gamma).$} Our dTBM shows different conclusion with TBM and DCBM in this case.

%     %   \textbf{Claim:} For TBM and DCBM, the minimax is equal to the minimin rate; i.e., for example, 
%     %   \begin{equation}
%     %       \inf_{\hat z \in \tP_z} \sup_{(z, \tS) \in \tP_{z, \rm TBM}(\gamma)} \bbE[p \ell(\hat z,  z)]  =  \inf_{\tS \in \tP_{\tS, {\rm TBM}}(\gamma)} \inf_{\hat z \in \tP_z} \sup_{z \in \tP_{z, \rm TBM}(\gamma)} \bbE[p \ell(\hat z,  z)]. 
%     %   \end{equation}

%     %   \textbf{Proof:}

%   \end{itemize}

%   \textbf{Minimax rate of dTBM.} We consider the equivalent statement of the minimin/minimax rate under dTBM. We omit the subscript dTBM in $\tP$ here.

%   \begin{align}
%       \inf_{\tS \in \tP_{\tS}(\gamma)} \inf_{\hat z \in \tP_z} \sup_{(z, \mtheta) \in \tP_{z, \mtheta}(\gamma)} \bbE[p \ell(\hat z, z)] \geq 1  \Leftrightarrow  \text{for all } \tS^* \text{ s.t. } \Delta_{\rm ang}(\tS^*)^2 = p^{\gamma}, \inf_{\hat z \in \tP_z} \sup_{(z, \mtheta) \in \tP_{z, \mtheta}(\gamma)} \bbE[p \ell(\hat z, z)| \tS^*] \geq 1,
%   \end{align}
%   and similarly, 
%     \begin{align}
%       \inf_{(\mtheta, \tS) \in \tP_{\mtheta, \tS}(\gamma)} \inf_{\hat z \in \tP_z} \sup_{z \in \tP_{z}(\gamma)} \bbE[p \ell(\hat z, z)] \geq 1  \Leftrightarrow  \text{for all } (\mtheta^*, \tS^*) \in \tP_{\mtheta, \tS}(\gamma), \inf_{\hat z \in \tP_z} \sup_{z \in \tP_{z}(\gamma)} \bbE[p \ell(\hat z, z)| \mtheta^*, \tS^*] \geq 1.
%   \end{align}

%   The key step to prove the minimin or minimax rate is to lower bound error with given true parameters $(z^*, \mtheta^*, \tS^*)$ 
%   \begin{equation}
%       \inf_{\hat z \in \tP_z} \bbE[p \ell(\hat z, z)| z^*, \mtheta^*, \tS^*] \gtrsim \frac{1}{|T^c|} \sum_{i \in T^c}\inf_{\hat z \in \tP_Z} \bbP(\hat z(i) = 1| z^*(i) = 2, \mtheta^*, \tS^*) +  \bbP(\hat z(i) = 2| z^*(i) = 1, \mtheta^*, \tS^*),
%   \end{equation}
%   where $T^c$ is a set defined in \cite{han2020exact} and $|T^c| \asymp c_2$ for some constant $c_2$.

%   For any $i \in T^c$, the probability on the right hand side corresponding to the hypothesis test
%   \begin{equation}
%       H_0: \my = \mtheta^*(i) \mx^*_1 + \me, \quad \leftrightarrow \quad H_1: \my = \mtheta^*(i) \mx^*_2 + \me,
%   \end{equation}
%   and the best estimator is $\hat z_{LS}(i) = \argmin_{a = 1,2} \onormSize{}{\my - \mtheta^*(i) \mx^*_a} _F^2$ by Neyman-Pearson Lemma, where $\mx^*_a = \mat_1(\tS^* \times_2 \mTheta^* \mM^* \times \cdots \times_K \mTheta^* \mM^*)_{a:}$.

%   By the calculation in equation (102) in \cite{han2020exact}, we have 
%   \begin{equation}
%       \bbP(\hat z(i) = 1| z^*(i) = 2, \mtheta^*, \tS^*) +  \bbP(\hat z(i) = 2| z^*(i) = 1, \mtheta^*, \tS^*) \geq 2 \bbP \of{ N(0,1) \geq \frac{\onormSize{}{\mtheta^*(i) \mx^*_1 - \mtheta^*(i) \mx^*_2}_F}{\sigma}}.
%   \end{equation}
%   Due to the summation over $T^c$ and the number $(1,2)$ can be replace by any $(a,b), a \neq b$, we define $f(z, \mtheta, \tS) = \min_{i \in T^c, a \neq b}\onormSize{}{\mtheta(i) \mx_a - \mtheta(i) \mx_b}_F$. Then, we can lower bounded the rate $\inf_{\hat z \in \tP_z} \bbE[p \ell(\hat z, z)| z^*, \mtheta^*, \tS^*]$ by a constant when $f(z^*, \mtheta^*, \tS^*) \lesssim \sigma$. Hence, we re-state the minimax rates with $f.$

%   \begin{align}
%       &(1) \inf_{\hat \in \tP_z} \sup_{(z, \mtheta, \tS) \in \tP(\gamma)} \bbE[p \ell(\hat z, z)| \mtheta^*, \tS^*] \geq 1  \Leftrightarrow  \text{there exists a }(z^*, \mtheta^*, \tS^*)  \in \tP(\gamma) \text{ s.t. } f(z^*, \mtheta^*, \tS^*) \lesssim \sigma \\
%         &(2) \inf_{\tS \in \tP_{\tS}(\gamma)} \inf_{\hat z \in \tP_z} \sup_{(z, \mtheta) \in \tP_{z, \mtheta}(\gamma)} \bbE[p \ell(\hat z, z)| \mtheta^*, \tS^*] \geq 1   \\
%         & \quad \quad \Leftrightarrow  \text{for all } \tS^* \in \tP_{\tS}(\gamma), \text{ there exists a }(z^*, \mtheta^*)  \in \tP_{z, \mtheta}(\gamma) \text{ s.t. } f(z^*, \mtheta^*, \tS^*) \lesssim \sigma \\
%         &(3) \inf_{(\mtheta, \tS) \in \tP_{\mtheta, \tS}(\gamma)} \inf_{\hat z \in \tP_z} \sup_{z \in \tP_{z}(\gamma)}  \bbE[p \ell(\hat z, z)| \mtheta^*, \tS^*] \geq 1   \\
%         & \quad \quad \Leftrightarrow  \text{for all } (\mtheta^*, \tS^*) \in \tP_{\mtheta, \tS} (\gamma), \text{ there exists a }z^*  \in \tP_{z}(\gamma) \text{ s.t. } f(z^*, \mtheta^*, \tS^*) \lesssim \sigma.
%   \end{align}

%   (1) is easy to prove under both TBM and dTBM, since we only need to design a specific $(z^*, \mtheta^*, \tS^*)$ such that $   f(z^*, \mtheta^*, \tS^*) \leq \sigma$.

%   (2) is easy in TBM because $f(z, \mtheta, \tS) = \min \onormSize{}{\mx_a - \mx_b}_F \leq p^{(K-1)/2} \Delta_{Euc}$, and thus $f(z, \mtheta, \tS) \lesssim \sigma $ for any $\tS$ such that $\Delta_{Euc}^2(\tS) \leq p^{-(K-1)} \sigma^2$.

%   For (2) under dTBM, we consider the upper bound
%   \begin{align}
%       f(z^*, \mtheta^*, \tS^*) & = \min_{i \in T^c} \mtheta^*(i) \min_{a \neq b }\onormSize{}{\mx_a^* - \mx_b^*}_F \\
%       & \leq  \min_{i \in T^c} \mtheta^*(i) \min_{a \neq b } \onormSize{}{\mS^*_{a:} - \mS^*_{b:}}_F \lambda_{\max}^{K-1}(\mTheta^* \mM^*)   \\
%       & \leq \min_{i \in T^c} \mtheta^*(i) \min_{a \neq b } \onormSize{}{\mS^*_{a:} - \mS^*_{b:}}_F (\sqrt{\max_a \onormSize{}{\mtheta_{z^{-1}(a)}^* }^2})^{(K-1)}
%   \end{align}
%   where the last inequality follows from Lemma 6.

%   \newpage

%   \section{New minimax results}

%   Let $\tP_{\tS} \coloneqq \{ \tS: c_3\leq \onormSize{}{\text{Mat}(\tS)_{a:}}\leq c_4, a\in[r]\}$ denote space for core tensor $\tS$ under dTBM;  let $\tP_{z} \coloneqq \{ z: {c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r}, a \in [r]\}$ denote the space for assignment $z$ under dTBM; let $\tP_{z ,\mtheta} \coloneqq \{ \mtheta\in\mathbb{R}^p_{+},\ 
% {c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r}, \onorm{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|, a \in [r] \}$ denote the space for $(z, \mtheta)$ under dTBM,  where $c_1, c_2, c_3, c_4$ are constants in $\tP$.

%   \begin{lem}[Trivial statistical impossibility]\label{lem:trivial}  Consider general Gaussian dTBMs with parameter space $\tP(\gamma)$ with $K\geq 1$. Assume $ r =  o(p^{1/3})$. For any true core tensors $\tS^* \in \tP_{\tS} $ such that $\Delta_{\min}^2 (\tS^*)\leq 2$ (or equivalently $\gamma = 0$), no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation under the worst case; i.e.,
%   \begin{equation}\label{eq:minminmax}
%   \liminf_{p \rightarrow \infty}\inf_{\tS^* \in  \tP_{\tS} \cap \{\Delta_{\min}^2(\tS^*) \leq 2\}}  \inf_{\hat z_{\text{stat}} \in \tP_z}\sup_{ (z^*, \mtheta^*) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
% \end{equation}
%   \end{lem}

%   \begin{rmk}The Lemma~\ref{lem:trivial} indicates that for any possible $\tS^*$ with   $\Delta_{\min}^2 (\tS^*)\leq 2$ there exist parameters $(z^*, \mtheta^*)$ such that no estimator is able to exactly recover the assignment under the dTBM with true parameter $(z^*, \tS^*,\mtheta^*)$. Note that $\onormSize{}{\mv_1^s - \mv_2^s}^2_F \leq 2$ for arbitrary vectors $\mv_1, \mv_2$. Then, the regime $ \tP_{\tS} \cap \{\Delta_{\min}^2 (\tS^*)\leq 2\} = \tP_{\tS}$ includes all possible core tensor $\tS^*$ under dTBM. Hence, we say this minimax lower bound~\eqref{eq:minminmax} is trivial.
%   \end{rmk}

%   \begin{cor}[Statistical impossibility]\label{cor:impossible} Consider general Gaussian dTBMs with parameter space $\tP(\gamma)$ with $K\geq 1$.  Assume $ r < c p^{1/3}$ for some positive constant $c$. If the signal exponent satisfies $\gamma < -(K-1)$, then, for any true core tensor $\tS^* \in \tP_{\tS}$ such that $\Delta_{\min}^2 (\tS^*) = p^{\gamma}$, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation under the worst case; i.e.,
%   \begin{equation}
%   \inf_{\tS^* \in \tP_{\tS} \cap \{\Delta_{\min}^2(\tS^*) = p^{\gamma} \} }  \inf_{\hat z_{\text{stat}} \in \tP_z }\sup_{ (z^*, \mtheta^*) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
% \end{equation}
%   \end{cor}

%   \begin{proof}[Proof of Corollary~\ref{cor:impossible}] Note that for any $\tS^*$
%   \begin{equation}
%       \{ (z, \mtheta): (z, \tS^*, \mtheta) \in \tP(\gamma) \} = \{ (z, \mtheta):  \mtheta\in\mathbb{R}^p_{+},\ 
% {c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r}, \onorm{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|, a\in[r] \},
%   \end{equation}
%           which is independent with $\gamma$. Then, we have $\{ (z, \mtheta): (z, \tS^*, \mtheta) \in \tP(\gamma) \} = \{ (z, \mtheta): (z, \tS^*, \mtheta) \in \tP(0) \}$. 
%           Also note that $ \tP_{\tS} \cap \{\Delta_{\min}^2(\tS^*) = p^{\gamma} \} \subset  \tP_{\tS} \cap \{\Delta_{\min}^2(\tS^*) \leq 2\}$. Hence, we have 
%           \begin{align}
%                 & \inf_{\tS^* \in \tP_{\tS} \cap \{\Delta_{\min}^2(\tS^*) = p^{\gamma} \} }  \inf_{\hat z_{\text{stat}} \in \tP_z }\sup_{ (z^*, \mtheta^*): (z^*, \tS^*, \mtheta^*) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right] \\
%                 & \quad \quad \geq \inf_{\tS^* \in \tP_{\tS} \cap \{\Delta_{\min}^2(\tS^*) \leq 2\}}  \inf_{\hat z_{\text{stat}} \in \tP_z}\sup_{ (z^*, \mtheta^*): (z^*, \tS^*, \mtheta^*) \in \tP(0)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
%           \end{align}
%   \end{proof}

%   \begin{proof}[Proof of Lemma~\ref{lem:trivial}] Consider the general asymetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$. For simplicity, we show the minimax rate for the estimation on the first mode $\hat z_1$; the proof for other modes are essentially the same.

%   To prove the minimax rate~\eqref{eq:minminmax}, it suffices to take an arbitrary $\tS^* \in  \tP_{\tS} \cap \{\Delta_{\min}^2(\tS^*) \leq 2\}$ and construct $(z^*, \mtheta^*)$ such that 
%   \begin{equation}
%       \inf_{\hat z_1 \in \tP_z} \bbE \left[ p\ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right]\geq 1.
%   \end{equation}

%   We consider the following construction:
%   \begin{enumerate}
%       \item[(i)] $\{z_k^*\}$ the same in \citet[Proof of Theorem 6]{han2020exact} with misclassification set $T^c_k \subset [p]$ and $|T^c_k| \asymp p/r$ for all $k \in [K]$.
%       \item [(ii)] $\{\mtheta_k^*\}$ such that $\mtheta_1^*(i) \leq \sigma r^{(K-1)/2} p^{-(K-1)/2}$ for all $i \in T_1^c$ and $\max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}^2_2 \asymp p/r$.
%   \end{enumerate}
%   % There exists a $(z^*, \mtheta^*) \in \tP_{z, \mtheta}$ satisfying the property (i) and (ii) when $r < p$; otherwise, when $r = p$, we must have $\mtheta^*(i) = 1$ due to the $\ell_1$ norm constraint and we can not satisfy the property (ii).

%   Then, following the proof of \cite{gao2018community}, we have 
%   \begin{equation}\label{eq:inf_lower}
%         \inf_{\hat z_1 \in \tP_z} \bbE \left[ \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq \frac{1}{r^3 |T_1^c|} \sum_{i \in T_1^c} \inf_{\hat z_1 \in \tP_1} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, \mtheta_k^*, \tS^*] \}. 
%   \end{equation}

%   The right hand side corresponds to the following hypothesis test for vectors: for an arbitrary $i \in T_1^c$, 
%   \begin{equation}\label{eq:hypo}
%       H_0: \my = \mtheta_1^*(i) \mx_1^* + \me \quad \leftrightarrow \quad H_1: \my = \mtheta_1^*(i) \mx_2^* + \me,
%   \end{equation}
%   where $\my = [\mat_1(\tY)]_{i:}$, $\mx_a^* = [\mat_1( \tS^* \times_2 \mTheta^* \mM_2^* \times_3 \cdots \times_K \mTheta^* \mM_K^*)]_{a:}$, and $\me$ consists of i.i.d.\ $N(0,\sigma^2)$ entries. By Neyman-Pearson lemma, the least square estimator is the best test for above hypothesis testing, where $\hat z_{LS}(i) = \argmin_{a = 1,2} \onormSize{}{\my - \mtheta_1(i)^* \mx_a^*}_F^2$.

%   Note that 
%   \begin{align}
%       \onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F &\leq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*)  \\
%       & \leq \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F   \max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}_2^{K-1} \\
%       & \leq  \sigma r^{(K-1)/2} p^{-(K-1)/2} 2 c_4 p^{(K-1)/2} r^{-(K-1)/2}\\
%       & \leq 2 c_4 \sigma, \label{eq:bound}
%   \end{align}
%   where the second inequality follows Lemma 6, the third inequality follows from property (ii) and the boundedness constraint in $\tP_{\tS}$ such that $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F  \leq \onormSize{}{\mS_{1:}^*}_F + \onormSize{}{\mS_{2:}^*}_F \leq 2c_4$.

%   Hence, we have 
%   \begin{align}
%       & \inf_{\hat z_1 \in \tP_1} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, \mtheta_k^*, \tS^*] \} \\
%       & \quad = 2 \bbP[ \hat z_{LS}(i) = 1 |  z_1^*(i) = 2, \mtheta_k^*, \tS^* ] \\
%       & \quad  = 2 \bbP[ \onormSize{}{\my - \mtheta_1(i)^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \mtheta_1(i)^* \mx_2^*}_F^2 | z_1^*(i) = 2, \mtheta_k^*, \tS^*   ] \\
%       & \quad =  2 \bbP[ N(0,1) \geq \mtheta_1(i)^* \onormSize{}{\mx_1^* - \mx_2^*}_F /\sigma ] \\
%       & \quad \geq 2 \bbP[ N(0,1) \geq 2 c_4 ]   \geq c, \label{eq:inf}
%   \end{align}
%   for some positive constant $c$, where the calculation for the last equation follows from the calculation in equation (102) in \cite{han2020exact}.

%   Plugging the inequality~\eqref{eq:inf} to inequality~\eqref{eq:inf_lower}, then, we have 
%   \begin{equation}
%       \liminf_{p \rightarrow \infty}  \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right]  \geq \liminf_{p \rightarrow \infty} \frac{ cp }{r^3} \geq c,
%   \end{equation}
%   where the last inequality follows by the condition $r = o(p^{1/3})$. By the discrete nature of the misclassification error, we have 
%   \begin{equation}
%       \liminf_{p \rightarrow \infty} \inf_{\tS^* \in  \tP_{\tS} \cap \{\Delta_{\min}^2(\tS^*) \leq 2\}}  \inf_{\hat z_{\text{stat}} \in \tP_z}\sup_{ (z^*, \mtheta^*) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]  \geq 1. 
%   \end{equation}
% %   for any $r < p$ and $p$,  we have 
% %   \begin{equation}
% %         \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq \frac{ cp }{r^3} > 0,
% %   \end{equation}
% %   which indicates 
% %   \begin{equation}
% %         \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq 1, \quad \text{for every } p
% %   \end{equation}
% %   by the discrete nature of $p\ell(\hat z, z^*)$.
%   \end{proof}

%   \begin{rmk}[Where the triviality comes from]
%   The proof of Lemma~\ref{lem:trivial} does not use the angle gap of $\tS^*$ but use the Euclidean gap of $\tS^*$ when bound the key quantity $ \onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F$ in \eqref{eq:bound}. By the boundedness constraint in $\tP_{\tS}$, we have $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F  \leq 2 c_4$. Then we can construct a small $\mtheta^*(i)$ to eliminate the constant Euclidean gap and make the classification problem very hard.  That's why the minimax bound~\eqref{eq:minminmax} is trivial with respect to the angle gap.
%   \end{rmk}

%   \begin{rmk}[Strongest position of $\mtheta$]
%   The proof of Lemma~\ref{lem:trivial} also indicates the strongest position of $\mtheta^*$ should be under the sup. Note that the key quantity in \eqref{eq:bound} relies on the Euclidean gap of $\tS^*$; i.e., 
%   \begin{equation}
%      \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\min}(\mTheta_k^*\mM_k^*)   \leq  \onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F \leq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*).
%   \end{equation}
%   We can not infer the Euclidean gap of $\tS^*$ from the angle gap of $\tS^*$. Therefore, we need to design a specific $\mtheta^*$ to eliminate the gap $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F$ to make a hard case to solve the hypothesis test~\eqref{eq:hypo}.

%   If we take $\mtheta$ under the inf, then we need to consider arbitrary combination of $\mtheta^*$ and $\tS^*$. Take $(\mtheta^*, \tS^*)$ such that $\onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F  \geq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\min}(\mTheta_k^*\mM_k^*)   \geq c \sigma p^{(K-1)/2} r^{-(K-1)/2}$. Plugging into the inequality~\eqref{eq:inf}, we have 
%   $ \inf_{\hat z_1 \in \tP_1} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, \mtheta_k^*, \tS^*] \} =  0$ and thus $\inf_{\hat z_1 \in \tP_z} \bbE \left[ p\ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right]  \geq 0$, which is meaningless because the misclassification error always $\geq 0$. 
%   \end{rmk}

%   \begin{rmk}[Comparison with TBM] In TBM, the statistical impossibility is not trivial. We will have following results:

%   Assume $ r \lesssim p^{1/3}$. If the signal exponent satisfies $\gamma < -(K-1)$, then, for any true core tensor $\tS^* \in \tP_{\tS}$ such that $\Delta_{\rm Euc}^2 (\tS^*) = p^{\gamma}$, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation under the worst case; i.e.,

%   \begin{equation}
%   \inf_{\tS^* \in \tP_{TBM, \tS} \cap \{\Delta_{\rm Euc}^2(\tS^*) = p^{\gamma} \} }  \inf_{\hat z_{\text{stat}} \in \tP_z }\sup_{ (z^*, \mtheta^*): (z^*, \tS^*, \mtheta^*) \in \tP_{TBM}(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
% \end{equation}
% The prove of above result is exactly the Proof of Lemma~\ref{lem:trivial} after replacing $\mtheta^* = \mathbf{1}$. Particularly, the inequality~\eqref{eq:bound} becomes 
% \begin{equation}
%      \onormSize{}{  \mx_1^*  -  \mx_2^*}_F  \leq \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F p^{(K-1)/2} r^{-(K-1)/2},
% \end{equation}
% and thus we need the Euclidean gap $\Delta_{\rm Euc}^2 (\tS^*) \leq \sigma^2  p^{-(K-1)} r^{K-1}$ to have $ \onormSize{}{  \mx_1^*  -  \mx_2^*}_F  \leq \sigma$. That's the why the TBM minimax rate is not trivial.
%   \end{rmk}

\bibliography{tensor_wang}
\bibliographystyle{apalike}

\end{document}

### 2023-01-20 tensor_clustering/ieee_final/ieee_final_source/ieee_final.tex
final

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%%%%%%%%%%%%%%%%%%%%%%%% my packages

\usepackage{amsmath,amsfonts,amssymb,amsthm,bm,mathtools}
\usepackage{algorithm,algpseudocode}
\usepackage{array}
%\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
% updated with editorial comments 8/9/2021

\usepackage{enumerate}
\usepackage[numbers]{natbib}
\usepackage{enumitem}

%\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\usepackage{dsfont,multirow,hyperref,enumerate}
\hypersetup{colorlinks,linkcolor=black,citecolor=black,urlcolor={black}}
\usepackage{lscape}
%\usepackage{afterpage}

\usepackage{microtype}
\usepackage{wrapfig}
\allowdisplaybreaks

\usepackage{tabularx,booktabs}
\newcolumntype{Y}{>{\centering\arraybackslash}X}

\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{cor}{Corollary}

\theoremstyle{definition}
\newtheorem{assumption}{Assumption}
\newtheorem{defn}{Definition}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}
\newtheorem{condition}{Condition}
\newtheorem{conjecture}{Conjecture}
\newtheorem{property}{Property}
\newtheorem{hypothesis}{Hypothesis}

\newtheorem{innercustomgeneric}{\customgenericname}
\providecommand{\customgenericname}{}
\newcommand{\newcustomtheorem}[2]{%
  \newenvironment{#1}[1]
  {%
   \renewcommand\customgenericname{#2}%
   \renewcommand\theinnercustomgeneric{##1}%
   \innercustomgeneric
  }
  {\endinnercustomgeneric}
}

\newcustomtheorem{customexample}{Example}

%\usepackage{fullpage}

\mathtoolsset{showonlyrefs}

\newcommand{\of}[1]{\left(#1\right)}
\newcommand{\off}[1]{\left[#1\right]}
\newcommand{\offf}[1]{\left\{#1\right\}}
\newcommand{\aabs}[1]{\left|#1\right|}
\newcommand{\ang}[1]{\left\langle#1\right\rangle}

\newcommand{\Mat}{\text{Mat}}

\def\ci{\perp\!\!\!\perp}

\usepackage{booktabs}
\newcommand\doubleRule{\toprule\toprule}
\allowdisplaybreaks

\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}

\input macros.tex
\newcommand*{\Scale}[2][4]{\scalebox{#1}{$#2$}}%
\setlength\parindent{0pt}
\usepackage[parfill]{parskip}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\newcommand\Algphase[1]{%
\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\textwidth}{0.4pt}%
\Statex\hspace*{-\algorithmicindent}\textbf{#1}%
\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\textwidth}{0.4pt}%
}
\usepackage{caption}
\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}

\newcounter{MYtempeqncnt}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Multiway Spherical Clustering via \\ Degree-Corrected Tensor Block Models}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Jiaxin~Hu
        and~Miaoyan~Wang% <-this % stops a space 
\thanks{The work of Miaoyan Wang was supported in part by NSF CAREER DMS-2141865, DMS-1915978, DMS-2023239, EF-2133740, and funding from the Wisconsin Alumni Research foundation. (\textit{Corresponding author: Miaoyan Wang.})}
\thanks{Jiaxin Hu and Miaoyan Wang are with the Department of Statistics, University of Wisconsin-Madison, Madison, WI, 53706 USA. (e-mail: jhu267@wisc.edu, miaoyan.wang@wisc.edu)}% <-this % stops a space
\thanks{Part of this paper was presented at 25th International Conference on Artificial Intelligence and Statistics (AISTATS).}% <-this % stops a space
}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~00, No.~0, 000~0000}{Hu \MakeLowercase{\textit{et al.}}: Multiway Spherical Clustering}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
We consider the problem of multiway clustering in the presence of unknown degree heterogeneity. Such data problems arise commonly in applications such as recommendation system, neuroimaging, community detection, and hypergraph partitions in social networks. The allowance of degree heterogeneity provides great flexibility in clustering models, but the extra complexity poses significant challenges in both statistics and computation. Here, we develop a degree-corrected tensor block model with estimation accuracy guarantees. We present the phase transition of clustering performance based on the notion of angle separability, and we characterize three signal-to-noise regimes corresponding to different statistical-computational behaviors. In particular, we demonstrate that an intrinsic statistical-to-computational gap emerges only for tensors of order three or greater. Further, we develop an efficient polynomial-time algorithm that provably achieves exact clustering under mild signal conditions. The efficacy of our procedure is demonstrated through two data applications, one on human brain connectome project, and another on Peru Legislation network dataset. 
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
tensor clustering, degree correction, statistical-computational efficiency, human brain connectome networks
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
% \IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
% for IEEE journal papers produced under \LaTeX\ using
% IEEEtran.cls version 1.8b and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)

\IEEEPARstart{M}{ultiway} arrays have been widely collected in various fields including social networks \citep{anandkumar2014tensor}, neuroscience \citep{wang2017bayesian}, and computer science \citep{koniusz2016sparse}. Tensors effectively represent the multiway data and serve as the foundation in higher-order data analysis. One data example is from multi-tissue multi-individual gene expression study~\citep{wang2019three,hore2016tensor}, where the data tensor consists of expression measurements indexed by (gene, individual, tissue) triplets. Another example is \emph{hypergraph} network~\citep{ghoshdastidar2017uniform,ghoshdastidar2017consistency,ahn2019community,ke2019community} in social science. A $K$-uniform hypergraph can be naturally represented as an order-$K$ tensor, where each entry indicates the presence of $K$-way hyperedge among nodes (a.k.a.\ entities). In both examples, identifying the similarity among tensor entities is important for scientific discovery.

We study the problem of multiway clustering based on a data tensor. The goal of multiway clustering is to identify a checkerboard structure from a noisy data tensor. Figure~\ref{fig:intro} illustrates the noisy tensor and the underlying checkerboard structures discovered by multiway clustering methods. In the hypergraph example, the multiway clustering aims to identify the underlying block partition of nodes based on their higher-order connectivities; therefore, we also refer to the clustering as \emph{higher-order clustering}. The most common model for higher-order clustering is called \emph{tensor block model} (TBM)~\citep{wang2019multiway}, which extends the usual matrix stochastic block model~\citep{abbe2017community} to tensors. The matrix analysis tools, however, are sub-optimal for higher-order clustering. Developing tensor tools for solving block models has received increased interest recently~\citep{ wang2019multiway,chi2020provable,han2020exact}.

The classical tensor block model suffers from drawbacks to model real world data in spite of the popularity. The key underlying assumption of block model is that all nodes in the same community are exchangeable; i.e., the nodes have no individual-specific parameters apart from the community-specific parameters. However, the exchangeability assumption is often non-realistic. Each node may contribute to the data variation by its own multiplicative effect. We call the unequal node-specific effects the \emph{degree heterogeneity}. Such degree heterogeneity appears commonly in social networks. Ignoring the degree heterogeneity may seriously mislead the clustering results. For example, the regular block model fails to model the member affiliation in the Karate Club network~\citep{bickel2009nonparametric} without addressing degree heterogeneity.

The \emph{degree-corrected tensor block model} (dTBM) has been proposed recently to account for the degree heterogeneity~\citep{ke2019community}. The dTBM combines a higher-order checkerboard structure with degree parameter $\mtheta=(\mtheta(1),\ldots,\mtheta(p))^T$ to allow heterogeneity among $p$ nodes.  Figure~\ref{fig:intro} compares the underlying structures of TBM and dTBM with the same number of communities. The dTBM allows varying values within the same community, thereby allowing a richer structure. To solve dTBM, we project clustering objects to a unit sphere and perform iterative clustering based on angle similarity. We refer to the algorithm as the \textit{spherical} clustering; detailed procedures are in Section~\ref{sec:alg}. The spherical clustering avoids the estimation of nuisance degree heterogeneity. The usage of angle similarity brings new challenges to the theoretical results, and we develop new polar-coordinate based techniques in the proofs.

\begin{figure*}[t]
    \centering
    \includegraphics[width = .9\textwidth]{intro2_arxiv.pdf}
    \caption{Examples for order-3 tensor block model (TBM) with and without degree correction. Both TBM and dTBM have four communities on each mode, while dTBM allows a richer structure with degree heterogeneity.
    }
    \label{fig:intro}
\end{figure*}

{\bf Our contributions.} The primary goal of this paper is to provide both statistical and computational guarantees for dTBM. Our main contributions are summarized below.
\begin{itemize}[leftmargin=*]

 \item We develop a general dTBM and establish the identifiability for the uniqueness of clustering using the notion of angle separability.

\item  We present the phase transition of clustering performance with respect to three different statistical and computational behaviors.  We characterize, for the first time, the critical signal-to-noise (SNR) thresholds in dTBMs, revealing the intrinsic distinctions among (vector) one-dimensional clustering, (matrix) biclustering, and (tensor) higher-order clustering. Specific SNR thresholds and algorithm behaviors are depicted in  Figure~\ref{fig:phase_axis}.

 \item We provide an angle-based algorithm that achieves exact clustering \emph{in polynomial time} under mild conditions. Simulation and data studies demonstrate that our algorithm outperforms existing higher-order clustering algorithms. 
\end{itemize}
The last two contributions, to our best knowledge, are new to the literature of dTBMs.

\begin{figure*}[t]
    \centering
    \includegraphics[width = 17cm]{phase.pdf}
    \caption{SNR thresholds for statistical and computational limits in order-$K$ dTBM with dimension $(p,...,p)$ and $K \geq 2$. The SNR gap between statistical possibility and computational efficiency  exists only for tensors with $K \geq 3$. }
    \label{fig:phase_axis}
\end{figure*}

{\bf Related work.} 
Our work is closely related to but also distinct from several lines of existing research. Table~\ref{tab:comp} summarizes the most relevant models.

\begin{itemize}[wide]
    \item \textit{Block model for clustering.} The block model such as stochastic block model (SBM) and degree-corrected SBM has been widely used for matrix clustering problems. The theoretical properties and algorithm performance for matrix block models have been well-studied~\citep{gao2018community}; see the review paper~\citep{abbe2017community} and the references therein. However, The tensor counterparts are relatively less understood.

    \item  \textit{Tensor block model.} The (non-degree) tensor block model (TBM) is a higher-order extension of SBM, and its statistical-computational properties are investigated in recent literatures~\citep{wang2019multiway, han2020exact, ghoshdastidar2017consistency}. Some works~\citep{ahn2018hypergraph} study the TBM with sparse observations, while, others~\citep{wang2019multiway, han2020exact} and our work focus on the dense regime. Extending results from non-degree to degree-corrected model is highly challenging. Our dTBM parameter space is equipped with angle-based similarity and nuisance degree parameters. The extra complexity makes the Cartesian coordinates based analysis~\citep{han2020exact} non-applicable to our setting. Towards this goal, we have developed a new polar coordinates based analysis to control the model complexity. We also develop a new angle-based iteration algorithm to achieve optimal clustering rates \emph{without the need of estimating nuisance degree parameters}.

    \item \textit{Degree-corrected block model.} The hypergraph degree-corrected block model (hDCBM) and its variant have been proposed in the literature~\citep{ke2019community, yuan2022testing}. For this popular model, however, the optimal statistical-computational rates remain an open problem. Our main contribution is to provide a sharp statistical and computational critical phase transition in dTBM literature. In addition, our algorithm results in a faster \emph{exponential} error rate, in contrast to the \emph{polynomial} rate in~\cite{ke2019community}. The original hDCBM~\citep{ke2019community} is designed for binary observations only, and we extend the model to both continuous and binary observations. We believe our results are novel and helpful to the community. See Fig~\ref{fig:phase_axis} for overview of our results.

    \item \textit{Global-to-local algorithm strategy.} Our methods generalize the recent global-to-local strategy for matrix learning~\citep{gao2018community,chi2019nonconvex,yun2016optimal} to tensors~\citep{han2020exact,ahn2018hypergraph,kim2018stochastic}. Despite the conceptual similarity, we address several fundamental challenges associated with this non-convex, non-continuous problem. We show the insufficiency of the conventional tensor HOSVD~\citep{de2000multilinear}, and we develop a weighted higher-order initialization that relaxes the singular-value gap separation condition. Furthermore, our local iteration leverages the angle-based clustering in order to avoid explicit estimation of degree heterogeneity. Our bounds reveal the interesting interplay between the computational and statistical errors. We show that our final estimate \emph{provably} achieves the exact clustering within only polynomial-time complexity.

\end{itemize}

\begin{table*}[t]
\resizebox{\textwidth}{!}{%
    \begin{tabular}{c|cccccc}
\hline
    & Gao et al. (2018)\cite{gao2018community}&  Ahn et al. (2018)\cite{ahn2018hypergraph} &Han et al. (2020)\cite{han2020exact}& Ghoshdastidar et al. (2019)\cite{ghoshdastidar2017consistency} & Ke et al. (2019)\cite{ke2019community} & \textbf{Ours}\\
    \hline
      Allow tensors of arbitrary order & $\times$ & $\surd$&  $\surd$ & $\surd$ &$\surd$ & $\surd$  \\
        Allow degree heterogeneity &  $\surd$ & $\times$ & $\times$ & $\surd$ &$\surd$ & $\surd$ \\
        Singular-value gap-free clustering &  $\surd$ & $\surd$ & $\surd$ & $\times$ & $\times$ &$\surd$ \\
        Misclustering rate (for order $K^*$)& - & ${p^{-(K-1)} \alpha^{-1}}^{**}$ & $\exp(-p^{K/2})$ & $p^{-1}$& $p^{-2}$ &  $\exp(-p^{K/2})$\\
        Consider sparse observation & $\times$ & $\surd$ & $\times$ & $\times$ & $\times$ &$\times$\\
        \hline
    \end{tabular}
    }
    \caption{ Comparison between previous methods with our method. $^*$We list the result for order-K tensors with $K \geq 3$ and general number of communities $r = \tO(1)$. $^{**}$The parameter $\alpha = f(p) > 0$ denotes the sparsity level which is some function of dimension $p$. }\label{tab:comp}
\end{table*}

{\bf Notation.} We use lower-case letters (e.g., $a,b$) for scalars, lower-case boldface letters (e.g., $\ma,\mtheta$) for vectors, upper-case boldface letters (e.g., $\mX,\mY$) for matrices, and calligraphy letters (e.g., $\tX,\tY$) for tensors of order three or greater. We use $\mone_p$ to denote a vector of length $p$ with all entries to be 1. We use $|\cdot|$ for the cardinality of a set and $\ind\{\cdot\}$ for the indicator function. For an integer $p\in\mathbb{N}_{+}$, we use the shorthand $[p]= \offf{1,2,...,p}$. For a length-$p$ vector $\ma$, we use $a(i)\in\mathbb{R}$ to denote the $i$-th entry of $\ma$, and use $\ma_{I}$ to denote the sub-vector by restricting the indices in the set $I\subset [p]$.  We use  $\onorm{\ma}=\sqrt{\sum_{i}a^2(i)}$ to denote the $\ell_2$-norm, $\onorm{\ma}_1=\sum_i |a_i|$ to denote the $\ell_1$ norm of $\ma$. For two vector $\ma, \mb$ of the same dimension, we denote the angle between $\ma, \mb$ by 
\begin{equation}
    \cos \of{\ma, \mb} = \frac{\ang{ \ma, \mb}}{ \onorm{\ma} \onorm{\mb} },
\end{equation}
where $\ang{\ma,\mb}$ is the inner product of two vectors and $\cos \of{\ma, \mb} \in [-1,1]$. We make the convention that $\cos \of{\ma, \mb} = \cos \of{\ma^T, \mb^T}$.

Let $\tY  \in \bbR^{p_1 \times \cdots \times p_K}$ be an order-$K$ $(p_1,...,p_K)$-dimensional tensor. We use $\tY(i_1,\ldots,i_K)$ to denote the $(i_1,\ldots,i_K)$-th entry of $\tY$. The multilinear multiplication of a tensor $\tS\in \bbR^{r_1\times \cdots \times r_K}$ by matrices $\mM_k \in\mathbb{R}^{p_k\times r_k}$ results in an order-$d$ $(p_1,\ldots,p_K)$-dimensional tensor $\tX$, denoted
\[
\tX=\tS \times_1 \mM_1 \times \cdots \times_K \mM_K,
\]
where the entries of $\tX$ are defined by
\small
\begin{align}
    &\tX(i_1,\ldots, i_K) \\
    & \quad =\sum_{(j_1,\ldots,j_K)}\tS(j_1,\ldots,j_K)\mM_1(i_1,j_1)\cdots \mM_K(i_K,j_K).
\end{align} 
\normalsize
For a matrix $\mY$, we use $\mY_{i:}$ (respectively, $\mY_{:i}$) to denote the $i$-th row (respectively, $i$-th column) of the matrix. Similarly, for an order-3 tensor, we use $\tY_{::i}$ to denote the $i$-th matrix slide of the tensor. We use $\text{Ave}(\cdot)$ to denote the operation of taking averages across elements and $\text{Mat}_k(\cdot)$ to denote the unfolding operation that reshapes the tensor along mode $k$ into a matrix.   {For a symmetric tensor $\tX \in\mathbb{R}^{p\times \cdots \times p}$, we omit the subscript and use $\text{Mat}(\tX)\in\mathbb{R}^{p\times p^{K-1}}$ to denote the unfolding.}  For two sequences $\{a_p\}, \{b_p\}$, we denote $a_p\lesssim b_p$ or $a_p=\tO(b_p)$ if $\lim_{p\to\infty}a_p /b_p\leq c$ for some constant $c\geq 0$, $a_p=o(b_p)$ if $\lim_{p\to\infty}a_p/b_p =0$, and $a_p = \Omega(b_p)$ if both $b_p \lesssim a_p$ and $a_p\lesssim b_p$. Throughout the paper, we use the terms ``community'' and ``clusters'' exchangeably.

{\bf Organization.} The rest of this paper is organized as follows. Section~\ref{sec:model} introduces the degree-corrected tensor block model (dTBM) with three motivating examples and presents the identifiability of dTBM under the angle gap condition. We show the phase transition and the existence of statistical-computational gaps for the higher-order dTBM in Section~\ref{sec:limits}. In Section~\ref{sec:alg}, we provide a polynomial-time two-stage algorithm with misclustering rate guarantees. { Extension to Bernoulli models is also presented. In Section~\ref{sec:tbm}, we compare our work with non-degree tensor block models.} Numerical studies including the simulation, comparison with other methods, and two real dataset analyses are in Sections~\ref{sec:simulation}-\ref{sec:real}. The main technical ideas we develop for addressing main theorems are provided in Section~\ref{sec:mainproof}. Detailed proofs and extra theoretical results are provided in Appendix.

\section{Model formulation and motivations}\label{sec:model}

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

\subsection{Degree-corrected tensor block model}

Suppose we have an order-$K$ data tensor $\tY \in \bbR^{p \times \cdots \times p}$. %; the extension to general case with asymmetric mean tensor is provided in Section~\ref{subsec:exten}.} 
%we focus on symmetric tensors in this section; the extension to general asymmetric tensors is provided in Section~\ref{subsec:exten}. 
Assume there exist $r \geq 2$  disjoint communities among the $p$ nodes. We represent the community assignment by a function $z \colon [p]\mapsto[r]$, where $z(i) = a$ for $i$-th node that belongs to the $a$-th community. Then, $z^{-1}(a)=\{i\in[p]\colon z(i)=a\}$ denotes the set of nodes that belong to the $a$-th community, and $|z^{-1}(a)|$ denotes the number of nodes in the $a$-th community. Let $\mtheta=(\theta(1),\ldots,\theta(p))^T$ denote the degree heterogeneity for $p$ nodes. We consider the order-$K$ dTBM~\citep{ghoshdastidar2017consistency,ke2019community},
\small
\begin{equation}\label{eq:model_margin}
    \tY(i_1,\ldots,i_K) = \tS(z(i_1),\ldots,z(i_K)) \prod_{k = 1}^K\theta_{i_k} + \tE(i_1,\ldots,i_K), 
\end{equation}
\normalsize
where $\tS \in \bbR^{r \times \cdots \times r}$ is an order-$K$ tensor collecting the block means among communities, and %\textcolor{blue}{$\tE\in\mathbb{R}^{p\times \cdots \times p}$ is a symmetric noise tensor consisting of independent zero-mean sub-Gaussian entries $(i_1,\ldots, i_K)$ for indices $i_1 \leq \cdots \leq i_K$ with variance bounded by $\sigma^2$. The entries in $\tE$ satisfy  $\tE(i_1,\ldots,i_K) = \tE(\pi(i_1),\ldots,\pi(i_K))$ for any permutation $\pi$ on $\{i_1, \ldots, i_K\}$. The symmetric constraint in $\tE$ is relaxed for general dTBM.} 
$\tE\in\mathbb{R}^{p\times \cdots \times p}$ is a noise tensor consisting of independent zero-mean sub-Gaussian entries with variance bounded by $\sigma^2$. 
The unknown parameters are $z$, $S$, and $\mtheta$. The dTBM can be equivalently written in a compact form of tensor-matrix product:
\begin{equation}\label{eq:model_tensor}
\bbE\tY = \tS \times_1 \mTheta \mM \times_2 \cdots \times_K  \mTheta \mM,
\end{equation}
where $\mTheta = \text{diag}(\theta(1),...,\theta(p)) \in \bbR^{p \times p}$ is a diagonal matrix, $\mM \in \offf{0,1}^{p \times r}$ is the membership matrix associated with community assignment $z$ such that $\mM(i,j)=\ind\{z(i)=j\}$. By definition, each row of $\mM$ has one copy of 1's and 0's elsewhere. Note that the discrete nature of $\mM$ renders our model~\eqref{eq:model_tensor} more challenging than Tucker decomposition. We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$ if $\tY$ admits dTBM~\eqref{eq:model_tensor}   {and let $\tX = \bbE \tY$ denote the mean tensor.} The goal of clustering is to estimate $z$ from a single noisy tensor $\tY$. We are particularly interested in the high-dimensional regime where $p$ grows whereas $r=\tO(1)$.

For ease of notation, we have focused on the case with symmetric mean tensor $\bbE \tY$. This assumption simplifies the notation because all modes have the same $(\mTheta, \mM, z)$; the noise tensor $\tE$ and the data tensor $\tY$ are still possibly asymmetric. In general, we allow asymmetric mean tensors with $\{(\mTheta_k, \mM_k, z_k)\}_{k=1}^K$, one for each mode. The extension can be found in Appendix II.

\subsection{Motivating examples}\label{subsec:motiv} Here, we provide four applications to illustrate the practical necessity of dTBM.

\paragraph{Tensor block model} Consider the model~\eqref{eq:model_tensor}. Let $\theta(i)=1$ for all $ i \in [p]$. The model~\eqref{eq:model_tensor} reduces to the tensor block model, which is widely used in previous clustering algorithms~\citep{wang2019multiway,chi2020provable,han2020exact}. The theoretical results in TBM serve as benchmarks for dTBM.

\paragraph{Community detection in hypergraphs} The hypergraph network is a powerful tool to represent the complex entity relations with higher-order interactions~\citep{ke2019community}. A typical undirected hypergraph is denoted as $H = (V,E)$, where $V = [p]$ is the set of nodes and $E$ is the set of undirected hyperedges. Each hyperedge in $E$ is a subset of $V$, and we call the hyperedge an order-$K$ edge if the corresponding subset involves $K$ nodes. We call $H$ a $K$-uniform hypergraph if $E$ only contains order-$K$ edges.

It is natural to represent the $K$-uniform hypergraph using a binary order-$K$ adjacency tensor. Let $\tY \in \{0,1\}^{p \times \cdots \times p}$ denote the adjacency tensor, where the entries encode the presence or absence of order-$K$ edges among $p$ nodes. Specifically, for all $(i_1,\ldots,i_K) \in [p]^K$, we have
\begin{equation}
    \tY(i_1,...,i_K) =  \begin{cases}
    1  & \text{if }  (i_1,...,i_K) \in E,\\
    0 & \text{if }  (i_1,...,i_K) \notin E.
    \end{cases}
\end{equation}

Assume there exist $r$ disjoint communities among $p$ nodes, and the connection probabilities depend on the community assignments and node-specific parameters. 
Then, the equation~\eqref{eq:model_tensor} models $\bbE \tY$ with unknown degree heterogeneity $\mtheta$ and sub-Gaussianity parameter $\sigma^2 = 1/4$.

\paragraph{Multi-layer weighted network} 
Multi-layer weighted network data consists of multiple networks over the same set of nodes. One representative example is the brain connectome data~\citep{zhang2019tensor}. The multi-layer weighted network $\tY$ has dimension of $p \times p \times L$, where $p$ denotes the number of brain regions of interest, and $L$ denotes the number of layers (networks). Each of the $L$ networks describes one aspect of the brain connectivity, such as functional connectivity or structural connectivity. The resulting tensor $\tY$ consists of a mixture of slices with various data types.

Assume there exist $r$ disjoint communities among $p$ nodes and $r_l$ disjoint communities among the $L$ layers. The multi-layer network community detection is modeled by the general asymmetric asymmetric dTBM model~\eqref{eq:model_tensor}
\begin{equation}
    \bbE \tY = \tS \times_1 \mTheta \mM \times_2 \mTheta \mM \times_3 \mTheta_l \mM_l, 
\end{equation}
where $(\mtheta \in \bbR^p, \mM \in \{0,1\}^{p \times r})$ and $ (\mtheta_l \in \bbR^L, \mM_l \in \{0,1\}^{L \times r_l})$ are the degree heterogeneity and membership matrices corresponding to the community structure for $p$ nodes and $L$ layers, respectively.

\paragraph{Gaussian higher-order clustering} Datasets in various fields such as medical image, genetics, and computer science are formulated as Gaussian tensors. One typical example is the multi-tissue gene expression dataset, which records the different gene expression in different individuals and different tissues. The dataset, denoted as $\tY \in \bbR^{p \times n \times t}$, consists of the expression data for $p$ genes of $n$ individuals in $t$ tissues.

Assume there exist $r_1, r_2, r_3$ disjoint clusters for $p$ genes, $n$ individuals, and $t$ tissues, respectively. We apply the general asymmetric dTBM model~\eqref{eq:model_tensor} 
\begin{equation}
    \bbE\tY = \tS \times_1 \mTheta_1 \mM_1 \times_2 \mTheta_2 \mM_2 \times_3 \mTheta_3 \mM_3, 
\end{equation}
where $\{(\mtheta_k, \mM_k)\}_{k=1}^3$ represents the heterogeneity and membership for genes, individuals, and tissues.

\begin{rmk}[Comparison with non-degree models]
Our dTBM uses fewer block parameters than TBM. In particular, every non-degree $r_1$-block tensor can be represented by a \emph{degree-corrected} $r_2$-block tensor with $r_2\leq r_1$. In particular, there exist tensors with $r_1=p$ but $r_2=1$, so the reduction in model complexity can be dramatic from $p$ to 1. This fact highlights the benefits of introducing degree heterogeneity in higher-order clustering tasks.
\end{rmk}

\subsection{Identifiability under angle gap condition}\label{subsec:identify}
The goal of clustering is to estimate the partition function $z$ from model~\eqref{eq:model_tensor}. For ease of notation, we focus on symmetric tensors; the extension to non-symmetric tensors are similar. We use $\tP$ to denote the following parameter space for $(z,\tS,\mtheta)$,
\small
\begin{align}
\tP=  &\bigg\{  (z,\tS,\mtheta)\colon  \ \mtheta\in\mathbb{R}^p_{+},\ 
{c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r}, \\
& c_3\leq \onormSize{}{\text{Mat}(\tS)_{a:}}\leq c_4, \onorm{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|, a\in[r]\bigg\}\label{eq:family}
\end{align}
\normalsize
where $c_i>0$'s are universal constants. We briefly describe the rationale of the constraints in~\eqref{eq:family}. 
First, the entrywise positivity constraint on  $\mtheta\in\mathbb{R}^p_{+}$ is imposed to avoid sign ambiguity between entries in $\mtheta_{z^{-1}(a)}$ and $\tS$. This constraint allows the trigonometric $\cos$ to describe the angle similarity in the Assumption~\ref{assmp:min_gap} below and Sub-algorithm~\hyperref[alg:main]{2} in Section~\ref{sec:alg}. Note that the positivity constraint can be achieved without sacrificing model flexibility, by using a slightly larger dimension of $\tS$ in the factorization~\eqref{eq:model_tensor}; see Example~\ref{ex:positive} below. Second, recall that the quantity $|z^{-1}(a)|$ denotes the number of nodes in $a$-th community. The constants $c_1, c_2$ in the $|z^{-1}(a)|$ bound assume the roughly balanced size across $r$ communities.  
{Third, the constant $c_3$ requires that all slides in $\tS$ have non-degenerate norm. Particularly, the lower bound $c_3$ excludes the no purely zero slide case to avoid trivial non-identifiability of model \eqref{eq:model_tensor}; see Example~\ref{example:c3} below. The upper bound $c_4$ is a technical constraint to avoid the slides with unbounded norm as dimension grows; in practice, the constraint $\max_{a:} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large $c_4 \geq \onormSize{}{\tY}_F$.} 
%{ Third, the constant $c_3$ requires no purely zero slide in $\tS$, and thus the model~\eqref{eq:model_tensor} is not trivially non-identifiable; see Example~\ref{example:c3} below. The upper bound $c_4$ is a technical constraint to avoid the unbounded entries in the core tensor; in practice, the constraint $\max_{a:} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large $c_4 \geq \onormSize{}{\tY}_F$. }  
%Third, the constants $c_3, c_4 $ in the magnitude of $\text{Mat}(\tS)_{a:}$ requires no purely zero slide in $\tS$, so the core tensor $\tS$ is not trivially reduced to a lower rank. 
Lastly, the $\ell_1$ normalization $\onormSize{}{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|$ is imposed to avoid the scalar ambiguity between $\mtheta_{z^{-1}(a)}$ and $\tS$. This constraint, again, incurs no restriction to model flexibility but makes our presentation cleaner.   {Our constraints in $\tP$ are mild compared with previous literature; see Table~\ref{tab:para} for comparison.}

\begin{example}[Positivity of degree parameters]~\label{ex:positive}
Here we provide an example to show the positivity constraints on $\mtheta$ incurs no loss on the model flexibility. 
Consider an order-3 dTBM with core tensor $\tS = 1$ and degree $\mtheta = (1,1,-1, -1)^T$. We have the mean tensor 
\begin{equation}
    \tX = \tS \times_1 \mTheta \mM \times_2 \mTheta \mM \times_3 \mTheta \mM,
\end{equation}
where $\mTheta = \text{diag}(\mtheta)$ and $\mM = (1,1,1,1)^T$. Note that $\tX \in \bbR^{4 \times 4 \times 4}$ is a 1-block tensor with \emph{mixed-signed} degree $\mtheta$, and the mode-3 slices of $\tX$ are
\begin{equation}
    \tX_{::1} = \tX_{::2} = - \tX_{::3} = -\tX_{::4}  = \begin{bmatrix}
    1 & 1& -1& -1 \\
    1 & 1& -1& -1 \\
    -1& -1&1 & 1\\
     -1& -1&1 & 1
    \end{bmatrix}.
\end{equation}
Now, instead of original decomposition, we encode $\tX$ as a 2-block tensor with \emph{positive-signed} degree. Specifically, we write
\begin{equation}
     \tX = \tS' \times_1 \mTheta' \mM' \times_2 \mTheta' \mM'  \times_3 \mTheta' \mM' , 
\end{equation}
where $\mTheta' = \text{diag}(\mtheta') = \text{diag}(1,1, 1, 1)$, the core tensor $\tS' \in \bbR^{2 \times 2 \times 2}$ has mode-3 slices, and the membership matrix $\mM'\in\{0,1\}^{4\times 2}$ defines the clustering $z'\colon[4]\to[2]$,
\begin{equation}
    \tS'_{::1} = -\tS'_{::2} = \begin{bmatrix}
    1 & -1 \\
    -1 & 1
    \end{bmatrix}, \quad \mM' = \begin{bmatrix}
   1 & 0 \\
   1 &0 \\
    0 & 1 \\
    0&1
    \end{bmatrix}.
\end{equation}
 The triplet $(z',\tS', \mtheta')$ lies in our parameter space~\eqref{eq:family}. In general, we can always reparameterize a block-$r$ tensor with mixed-signed degree using a block-$2r$ tensor with positive-signed degree. Since we assume $r=\tO(1)$ throughout the paper, the splitting does not affect the error rates of our interest.
 \end{example}

  \begin{example}[Non-identifiability with purely zero core slice]\label{example:c3} 
 Consider an order-2 dTBM with core tensor $\mS = \begin{pmatrix} 0 & 0\\
    1 & -1
    \end{pmatrix}$ and mean tensor 
    \begin{equation}
        \tX =  \mTheta_1 \mM \mS   \mM^T \mTheta_2, \quad \text{with } \mM = \begin{bmatrix} 1 & 0 \\
        1 & 0 \\
        0 & 1 \\
        0 & 1 \\
        \end{bmatrix}
    \end{equation}
    and $ \mTheta_1 =\mTheta_2= \text{diag}(1,1,1,1)$,
Replacing $\mTheta_1$ by $\mTheta'_1 = (3/2, 1/2, 1,1)$ leads to the same mean tensor $\tX$. 
\end{example}

 \begin{table*}[th]
    \centering
    \resizebox{.85\textwidth}{!}{
    \begin{tabular}{c|c ccc}
    \hline
    Assumptions in parameter space& Gao et al. (2018)\cite{gao2018community}& Han et al. (2020)\cite{han2020exact}& Ke et al. (2019)\cite{ke2019community} & {\bf Ours}\\
    \hline
         Balanced community sizes &$\surd$ &$\surd$&$\surd$ & $\surd$  \\
          Bounded core tensors & $\surd$ & $\times$ & $\surd$ &$\surd$\\
         Balanced degrees& $\surd$&-&$\surd$ &$\surd$\\
         Flexible in-group connections &$\times$ &$\surd$&$\surd$& $\surd$\\
         Gaps among cluster centers & In-between cluster difference & Euclidean gap & Eigen gap & Angle gap\\
         \hline
    \end{tabular}
    }
    \caption{Parameter space comparison between previous work with our assumption.}
    \label{tab:para}
\end{table*}

We now provide the identifiability conditions for our model before estimation procedures. When $r=1$, the decomposition~\eqref{eq:model_tensor} is always unique (up to cluster label permutation) in $\tP$, because dTBM is equivalent to the rank-1 tensor family under this case. When $r\geq 2$, the Tucker rank of signal tensor $\bbE\tY$ in~\eqref{eq:model_tensor} is bounded by, but not necessarily equal to, the number of blocks $r$~\citep{wang2019multiway}. Therefore, one can not apply the classical identifiability conditions for low-rank tensors to dTBM. Here, we introduce a key separation condition on the core tensor.

\begin{assumption}[Angle gap] \label{assmp:min_gap}Let $\mS = \text{Mat}(\tS)$. Assume the minimal gap between normalized rows of $\mS$ is bounded away from zero; i.e.,
\begin{equation}\label{eq:minimal_gap}
    \Delta_{\min} := \min_{a \neq b \in [r]} \onorm{\frac{\mS_{a:}}{\onormSize{}{ \mS_{a:}}} - \frac{\mS_{b:}}{\onormSize{}{ \mS_{b:}}} }>0, \quad \text{for} \quad r \geq 2.
\end{equation}
\end{assumption}
We make the convention $\Delta_{\min} = 1$ for $r = 1$. Equivalently, \eqref{eq:minimal_gap} says that none of the two rows in $\mS$ are parallel; i.e., $\max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}}  = 1-\Delta^2_{\min}/2<1$. The quantity $\Delta_{\min}$ characterizes the non-redundancy among clusters measured by angle separation. The denominators involved in definition~\eqref{eq:minimal_gap} are well posed because of the lower bound on $\onorm{\mS_{a:}}$ in~\eqref{eq:family}.

Our first main result is the following theorem showing the sufficiency and necessarily of the angle gap separation condition for the parameter identifiability under dTBM.

\begin{thm}[Model identifiability]\label{thm:unique} Consider the dTBM with $r\geq 2$ {and $K \geq 2$}. The parameterization~\eqref{eq:model_tensor} is unique in $\tP$ up to cluster label permutations, if and only if Assumption~\ref{assmp:min_gap} holds.
\end{thm}

The identifiability guarantee for the dTBM is stronger than classical Tucker model. In the Tucker model, the factor matrix $\mM$ is identifiable only up to orthogonal rotations. In contrast, our model does not suffer from rotational invariance. As we will show in Section~\ref{sec:alg}, each column of the membership matrix $\mM$ can be precisely recovered under our algorithm. This property benefits the interpretation of dTBM in practice.

\section{Statistical-computational critical values for higher-order tensors}\label{sec:limits}

\subsection{Assumptions} \label{sec:prelim}

We propose the signal-to-noise ratio~(SNR),
\begin{align}\label{eq:gamma}
  \text{SNR}:= \Delta^2_{\min}/\sigma^2 = p^{\gamma}, 
\end{align}
with varying $\gamma \in \bbR$ that quantifies different regimes of interest. We call $\gamma$ the \emph{signal exponent}. Intuitively, a larger SNR, or equivalently a larger $\gamma$, benefits the clustering in the presence of noise.

With quantification~\eqref{eq:gamma}, we consider the following parameter space,
\begin{equation}\label{eq:gammafamily}
    \tP(\gamma) = \tP\cap\{\tS \text{ satisfies SNR condition~\eqref{eq:gamma} with $\gamma$} \}.
\end{equation}
The $1$-block dTBM does not belong to the space $\tP(\gamma)$ when $\gamma < 0$ by Assumption~\ref{assmp:min_gap}. Our goal is to characterize the clustering accuracy with respect to $\gamma$ under the space $\tP(\gamma)$.

In our algorithmic development, we often refer to the regime of balanced degree heterogeneity. We call the degree $\mtheta$ \emph{balanced} if
\begin{equation}\label{eq:degree}
{\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
\end{equation}

The following lemma provides the rational of balanced degree assumption. We show the close relation between angle gaps in the mean tensor $\tX$ and the core tensor $\tS$ under balanced degree heterogeneity.

\begin{lem}[Angle gaps in $\tX$ and $\tS$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family} {with $r \geq 2$. Suppose $\mtheta$ is balanced satisfying~\eqref{eq:degree} {and $\min_{i\in[p]}\theta(i)\geq c$ from some constant $c>0$}.} Then, {as $p \rightarrow \infty$}, for all $i,j$ such that $z(i) \neq z(j)$, we have
%. Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying~\eqref{eq:degree}. Then, for all $i,j$ such that $z(i) \neq z(j)$, we have
\begin{equation}
 % \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  } \asymp  \onorm{ \frac{\mS_{z(i):}}{\onormSize{}{\mS_{z(i):}}}  -  \frac{\mS_{z(j):}}{\onormSize{}{\mS_{z(j):}}}  },
 \cos(\mX_{i:}, \mX_{j:})\asymp  \cos(\mS_{z(i):}, \mS_{z(j):}),
\end{equation}

where $\mX =\mat(\tX)$ and $\mS = \mat (\tS)$.
\end{lem}
In practice, an estimation algorithm has access to a noisy version of $\tX$ but not $\tS$. %Intuitively, the performance of dTBM is determined by both the core tensor and degree heterogeneity. 
%Since we view degrees as nuisance parameters, 
Our goal is to establish the algorithm performance with respect to the signal $\Delta^2_{\min}$ in the core tensor. By Lemma~\ref{lem:angle_gap_x}, the mapping from the core tensor $\mS_{z(i):}$ to the mean tensor $\mX_{z(i):}$ preserves the angle information $\Delta_{\min}^2$ under balanced degree heterogeneity~\eqref{eq:degree}. Therefore, the balanced degree assumption helps to exclude the cases in which the degree heterogeneity distorts the algorithm guarantees.

Here, we provide an example to illustrate the insufficiency of $\Delta_{\min}^2$ in the absence of balanced degrees.

\begin{example}[Insufficiency of $\Delta_{\min}^2$ in the absence of balanced degrees] Consider an order-2 $(p,p)$-dimensional dTBM with core matrix
\begin{equation}\label{eq:modelexample}
    \mS = \begin{pmatrix} 1 &a\\
    1 & -a
    \end{pmatrix}, 
\end{equation}
and $ \mtheta \text{ such that } \onormSize{}{\mtheta_{z^{-1}(1)}}^2 = p^m \onormSize{}{\mtheta_{z^{-1}(2)}}^2$,
where {$m \in [-1,1]$} is a scalar parameter controlling the skewness of degrees. Let $\Delta_{\mX}^2$ denote the minimal angle gap of the mean tensor, defined by
\begin{equation}\label{eq:delta_x}
    \Delta_{\mX}^2 \coloneqq \min_{i,j \in [p], z(i) \neq z(j)} \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  }, \quad %\quad  \tX = \tS \times_1 \mTheta \mM \times_2 \cdots \times_K \mTheta \mM.
\end{equation}
where $ \mX = \mat_1(\tX)$.
Take $ a = p^{-1/4}$ in the model setup~\eqref{eq:modelexample}. We have 
\begin{align}
    \Delta_{\min}^2 &= \frac{2 a^2}{1 + a^2} \asymp p^{-1/2}, \quad  \\
    \Delta_{\mX}^2 & = \frac{2\onormSize{}{\mtheta_{z^{-1}(2)}}^{2} a^2}{\onormSize{}{\mtheta_{z^{-1}(1)}}^{2} +  \onormSize{}{\mtheta_{z^{-1}(2)}}^{2}a^2} \asymp p^{-1/2-m}.
\end{align}

{Based on our theory in later Sections, the dTBM is impossible to solve when $\Delta^2_{\mX} \lesssim p^{-1}$
%but can be effectively solved when $\Delta^2_{\min} \gtrsim p^{-1/2}$; 
}; that is, the dTBM estimation depends on the relative magnitude of $m$ vs. $1/2$. In such a setting, the proposed signal notion $\Delta^2_{\min}$ alone fails to fully characterize dTBM. 
\end{example}

\begin{rmk}[Flexibility in balanced degree assumption] One important note is that our balance assumption~\eqref{eq:degree} does not preclude the mild degree heterogeneity. In fact, within each of the clusters, we allow the highest degree at the order $\tO(p)$, whereas the lowest degree at the order $\Omega(1)$. This range is more relaxed than previous work \citep{gao2018community} that restricts the highest degree in the sub-linear regime $o(p)$ and the lowest degree at the order $\Omega(1)$. 
\end{rmk}

\begin{rmk}[Similar assumptions in literature]
%We propose he balance assumption~\eqref{eq:degree} to avoid the extreme inflation or elimination of the core tensor gap $\Delta_{\min}^2$ after the weighted mapping by degree heterogeneity. 
Similar degree regulations are not rare in literature. In higher-order tensor model \citep{ke2019community}, the degree assumption $\max_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}} \leq $ $ C \min_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}$ is made to ensure balancedness across communities. %Since they have no $\ell_1$ constraint on $\mtheta$, their balance assumption is equivalent to ours~\eqref{eq:degree} multiplying some constants to the degrees. " 
In \cite{gao2018community}, the degree distribution is restricted to ${1\over |z^{-1}(a)|}\sum_{i\in z^{-1}(a)}\theta_i=1+o(1)$ for all communities. 
\end{rmk}

Last, let $\hat z$ and $z$ be the estimated and true clustering functions in the family~\eqref{eq:family}. Define the misclustering error by
\[
\ell(\hat z, z)={1\over p}\min_{\pi \in \Pi}\sum_{i\in[p]}\ind\{\hat z(i)\neq \pi \circ z(i)\},
\]
where $\pi: [r] \mapsto [r]$ is a permutation of cluster labels, $\circ$ denotes the composition operation, and $\Pi$ denotes the collection of all possible permutations. The infinitum over all permutations accounts for the ambiguity in cluster label permutation.

In Sections~\ref{sec:statlimit} and \ref{sec:complimit}, we provide the phase transition of $\ell (\hat z, z)$ for general Gaussian dTBMs~\eqref{eq:model_margin} without symmetric assumptions. For general (asymmetric) Gaussian dTBMs, we assume Gaussian noise $\tE(i_1,\ldots,i_K)\stackrel{\text{i.i.d.}}{\sim} N(0,\sigma^2)$, and we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode. For notational simplicity, we still use $z$ and $\tP(\gamma)$ for this general (asymmetric) model. All results should be interpreted as the worst-case results across $K$ modes.

\subsection{Statistical critical value}\label{sec:statlimit}
The statistical critical value means the SNR required for solving dTBMs with \emph{unlimited computational cost.}   Our following result shows the minimax lower bound for exact recovery and the matching upper bound for maximum likelihood estimator (MLE).  We consider the Gaussian MLE, denoted as $(\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}})$, over the estimation space $\tP$, where 
\begin{equation}\label{eq:mle}
    (\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}}) = \argmin_{ (z, \mtheta, \tS) \in \tP} \onormSize{}{\tY - \tX(z,\mtheta, \tS)}_F^2.
\end{equation}

\begin{thm}[Statistical critical value]\label{thm:stats} Consider general Gaussian dTBMs with parameter space $\tP(\gamma)$ with $K\geq 1$. Then, we have the following statistical phase transition.

\begin{itemize}[wide]
    \item 
    \textbf{Impossibility.}  {Assume $p \rightarrow \infty$ and $2 \leq r\lesssim p^{1/3}$.} Let $\tP_{\tS}(\gamma) \coloneqq \{ \tS: c_3\leq \onormSize{}{\text{Mat}(\tS)_{a:}}\leq c_4, a\in[r]\} \cap \{ \tS: \Delta_{\min}^2 = p^{\gamma}\}$ denote the space for valid $\tS$ satisfying SNR condition~\eqref{eq:gamma}, and $\tP_{z ,\mtheta} \coloneqq \{ \mtheta\in\mathbb{R}^p_{+},\ 
{c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r}, \onorm{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|, a \in [r] \}$ denote the space for valid $(z, \mtheta)$,  where $c_1, c_2, c_3, c_4$ are the constants in parameter space \eqref{eq:family}. If the signal exponent satisfies $\gamma < -(K-1)$, then, for any true core tensor $\tS \in \tP_{\tS}(\gamma)$, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation; that is, when $\gamma< -(K-1)$, we have 
    \begin{equation}\label{eq:minminmax}
   \liminf_{p \rightarrow \infty}\inf_{\tS \in  \tP_{\tS}(\gamma)}  \inf_{\hat z_{\text{stat}}}\sup_{ (z, \mtheta) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
\end{equation}
{ 
    Further, we define the parameter space $\tP'(\gamma') \coloneqq \tP\cap \{ \Delta_{\mX}^2 = p^{\gamma'} \}$, where $\Delta_{\mX}^2$ is the mean tensor minimal gap in \eqref{eq:delta_x}. When $\gamma' < -(K-1)$, we have 
    %the impossibility can also be described by the minimal angle in mean tensor $\Delta_{\mX}^2$ defined in \eqref{eq:delta_x}. 
    %Let $\tP'(\gamma') \coloneqq \tP\cap \{ \Delta_{\mX}^2 = p^{\gamma'} \}$. We have 
    \begin{equation}
        \liminf_{p \rightarrow \infty} \inf_{\hat z_{\rm stat}} \sup_{(z, \mtheta, \tS) \in \tP'(\gamma')}  \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
    \end{equation}
    }

%     \textbf{Impossibility.} 
% Assume $r\lesssim p^{1/3}$. If the signal exponent satisfies $\gamma < -(K-1)$, then, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation; that is,
% \begin{equation}
%   \gamma< -(K-1) \quad \Rightarrow  \inf_{\hat z_{\text{stat}}}\sup_{(z, \tS, \mtheta)\in\tP(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
% \end{equation}
\item \textbf{MLE achievability.} 
Suppose the signal exponent satisfies $\gamma >-(K-1)+c_0$ for an arbitrary constant {$c_0>0$}. Furthermore, assume {$\mtheta$} is balanced and $\min_{i\in[p]}\theta(i)\geq c$ from some constant $c>0$. Then, {when $p \rightarrow \infty$, fixed $r \geq 1$, }the MLE in~\eqref{eq:mle} achieves exact recovery in high probability; that is,
%Suppose the signal exponent satisfies $\gamma >-(K-1)+c$ for an arbitrary constant $c>0$. Furthermore, assume $\theta$ is balanced and $\min_{i\in[p]}\theta(i)\geq c$ from some constant $c>0$. Then, the MLE in~\eqref{eq:mle} achieves exact recovery in high probability; that is, when $\gamma> -(K-1)+c$, we have 
\begin{equation}
\ell(\hat z_{\text{MLE}}, z) \lesssim \text{SNR}^{-1}\exp\left(-{p^{K-1}\text{SNR}\over r^{K-1}}\right) \to 0,
\end{equation}
with probability going to 1. 
\end{itemize}
\end{thm}

The proofs for the two parts in Theorem~\ref{thm:stats} are in the Appendix II, Section~\ref{sec:statprove1} and Section~\ref{sec:statprove2}, respectively. The first part of Theorem~\ref{thm:stats} demonstrates impossibility of exact recovery whenever the core tensor $\tS$ satisfies SNR condition~\eqref{eq:gamma} with exponent $\gamma < -(K-1)$. The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to MLE and trace maximization~\citep{ghoshdastidar2017uniform}. The minimax bound~\eqref{eq:minminmax} indicates the worst case impossibility for a particular core tensor $\tS$ with signal exponent $\gamma < -(K-1)$; i.e., under the assumptions of Theorem~\ref{thm:stats}, when $\gamma < -(K-1)$, we have 
\begin{equation}
    \liminf_{p \rightarrow \infty}  \inf_{\hat z_{\text{stat}}}\sup_{ (z,\tS, \mtheta) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
\end{equation}
Such worst case impossibility is frequently studied in related works \citep{han2020exact, gao2018community} while our lower bound~\eqref{eq:minminmax} provides a stronger impossibility statement for arbitrary core tensor with weak signal.

The second part of Theorem~\ref{thm:stats} shows the exact recovery of MLE when the $\gamma>-(K-1)+c$ for an arbitrary constant $c>0$. Combining the impossibility and achievability results, we conclude that the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM with respect to our SNR.

\subsection{Computational critical value}\label{sec:complimit}
In this section, we derive the computational critical value of dTBMs. The computational critical value means the minimal SNR required for exactly recovery with \emph{polynomial-time} computational cost. An important ingredient to establish the computational limits is the \emph{hypergraphic planted clique (HPC) conjecture} \citep{zhang2018tensor, brennan2020reducibility}. The HPC conjecture indicates the impossibility of fully recovering the planted cliques with polynomial-time algorithm when the clique size is less than the number of vertices in the hypergraph. The formal statement of HPC detection conjecture is provided in Definition~\ref{def:HPC} and Conjecture~\ref{hypo:HPC} as follows.

\begin{defn}[Hypergraphic planted clique (HPC) detection]~\label{def:HPC} Consider an order-$K$ hypergraph $H = (V,E)$ where $V = [p]$ collects vertices and $E$ collects all the order-$K$ edges. Let $\tH_k(p, 1/2)$ denote the Erd\H{o}s-R\'{e}nyi $K$-hypergraph where the edge $(i_1,\ldots, i_K)$ belongs to $E$ with probability $1/2$. Further, we let $\tH_K(p, 1/2, \kappa)$ denote the hyhpergraph with planted cliques of size $\kappa$. Specifically, we generate a hypergraph from $\tH_k(p, 1/2)$, pick $\kappa$ vertices uniformly from $[p]$, denoted $K$, and then connect all the hyperedges with vertices in $K$. Note that the clique size $\kappa$ can be a function of $p$, denoted $\kappa_p$. The order-$K$ HPC detection aims to identify whether there exists a planted clique hidden in an Erd\H{o}s-R\'{e}nyi $K$-hypergraph. The HPC detection is formulated as the following hypothesis testing problem
\begin{equation}
    H_0:\ H \sim \tH_K(p,1/2) \quad \text{versus} \quad H_1: \ H \sim \tH_K(p,1/2, \kappa_p).
\end{equation}
\end{defn}

\begin{conjecture}[HPC conjecture]\label{hypo:HPC} Consider the HPC detection problem in Definition~\ref{def:HPC}. Suppose the sequence $\{\kappa_p\}$ such that $\limsup_{p \rightarrow \infty} \log \kappa_p/ \log \sqrt{p} \leq (1 - \tau)$. Then, for every sequence of polynomial-time test $\{ \varphi_p\}: H \mapsto \{0,1\}$ we have 
\begin{equation}
    \liminf_{p \rightarrow \infty} \bbP_{H_0} \of{ \varphi_p(H) =1 } +  \bbP_{H_1} \of{ \varphi_p(H) =0} \geq \frac{1}{2}.
\end{equation}
\end{conjecture}

Under the HPC conjecture, we establish the SNR lower bound that is necessary for any \emph{polynomial-time} estimator to achieve exact clustering.

\begin{thm}[Computational critical value]\label{thm:comp} 
Consider general Gaussian dTBMs under the parameter space $\tP$ with $K\geq 2$. Then, we have the following computational phase transition.

\begin{itemize}[wide]
    \item \textbf{Impossibility.} Assume HPC conjecture holds {and $r \geq 2$}. If the signal exponent satisfies $\gamma < -K/2$, then, no \emph{polynomial-time estimator} $\hat z_{\text{comp}}$ achieves exact recovery in expectation as $p\to \infty$; that is, when $\gamma<-K/2$, we have 
\begin{align}
   \liminf_{p\to \infty}\sup_{(z, \tS, \mtheta) \in \tP(\gamma)}  \bbE \left[ p\ell(\hat z_{\text{comp}}, z) \right]\geq 1.
\end{align}
\item \textbf{Polynomial-time algorithm achievability.} Suppose the parameter space satisfies $\gamma  >  -K/2+c_0$ for an arbitrary constant {$c_0>0$}. Furthermore, assume {fixed $r \geq 1, K \geq 2$,} the degree is balanced, lower bounded in that $\min_{i\in[p]}\theta_i\geq c$ for some constant $c>0$, and satisfies the local linear stability in Definition~\ref{def:stable} {in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \geq \check{C}\log^{-1}p $ with some positive constant $\check{C}$.}
%in the neighborhood $\tN(z,\varepsilon)$ for $\varepsilon \geq \log^{-1}(p)$. 
Then, {as $p \rightarrow \infty$}, there exists a polynomial-time algorithm $\hat z_{\text{ploy}}$ that achieves exact recovery in high probability; that is,
%Suppose the parameter space satisfies $\gamma  >  -K/2+c$ for an arbitrary constant $c>0$. Furthermore, assume the degree is balanced, lower bounded in that $\min_{i\in[p]}\theta_i\geq c$ for some constant $c>0$, and satisfies the local linear stability in Definition~\ref{def:stable} in the neighborhood $\tN(z,\varepsilon)$ for $\varepsilon \geq \log^{-1}(p)$. Then, there exists a polynomial-time algorithm $\hat z_{\text{ploy}}$ that achieves exact recovery in high probability; that is, when $\gamma>-K/2+c$, we have 
\begin{equation}
     % \ell(\hat z_{\text{poly}}, z) \lesssim \exp \of{ - \frac{p^{K-1}}{r^{K-1}} \Delta_{\min}^2 }.
 \ell(\hat z_{\text{poly}}, z) \lesssim \text{SNR}^{-1}\exp \of{ - \frac{p^{K-1}\text{SNR}}{r^{K-1}} } \to 0,
\end{equation}
with probability going to 1. 
\end{itemize}
\end{thm}

The proofs for the two parts in Theorem~\ref{thm:comp} are in the Appendix II, Section~\ref{sec:compprove1} and Section~\ref{sec:statprove2}, respectively. The first part of Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$, and the second part shows the existence of such algorithm when {$\gamma > -K/2+c_0$  for an arbitrary constant $c_0$ and} under extra technical assumptions. In Section~\ref{sec:alg}, we will present an efficient polynomial-time algorithm in this setting. Therefore, we conclude that $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM with respect to our SNR.

\begin{rmk}[Statistical-computational gaps]
Now, we have established the phase transition of exact clustering under order-$K$ dTBM by combing Theorems~\ref{thm:stats} and \ref{thm:comp}. Figure~\ref{fig:phase_axis} summarizes our results of critical SNRs when $K \geq 2$. In the weak SNR region $\gamma < -(K-1)$, no statistical estimator succeeds in degree-corrected higher-order clustering. In the strong SNR region $\gamma  > -K/2$, our proposed algorithm precisely recovers the clustering in polynomial time. In the moderate SNR regime, $-(K-1)\leq \gamma \leq -K/2$, the degree-corrected clustering problem is statistically easy but computationally hard. Particularly, dTBM reduces to matrix degree-corrected model when $K =2$, and the statistical and computational bounds show the same critical value. When $K =1$, dTBM reduces to the degree-corrected sub-Gaussian mixture model (GMM) with model
\begin{equation}
    \mY = \mTheta \mM \mS + \mE,
\end{equation}
where $\mY \in \bbR^{p \times d}$ collects $n$ data points in $\bbR^d$, $\mS \in \bbR^{r \times d}$ collects the $d$-dimensional centroids for $r$ clusters, and $\mTheta \in \bbR^{p \times p}, \mM \in \{0,1\}^{p \times r}, \mE \in \bbR^{p \times d}$ have the same meaning as in dTBM. \cite{lu2016statistical} implies that polynomial-time algorithms are able to achieve the statistical minimax lower bound in GMM. Therefore, we conclude that the statistical-to-computational gap emerges only for higher-order tensors with $K \geq 3$. The result reveals the intrinsic distinctions among (vector) one-dimensional clustering, (matrix) biclustering, and (tensor) higher-order clustering. 
\end{rmk}

\section{Polynomial-time algorithm under mild SNR}\label{sec:alg}
In this section, we present an efficient polynomial-time clustering algorithm under mild SNR. The procedure takes a global-to-local approach. See Figure~\ref{fig:demo} for illustration. The global step finds the basin of attraction with polynomial miclustering error, whereas the local iterations improve the initial clustering to exact recovery. Both steps are critical to obtain a satisfactory algorithm output. In what follows, we first use the symmetric tensor as a working example to describe the algorithm procedures to gain insight.    {Our theoretical analysis focuses on the symmetric mean tensor with independent sub-Gaussian noises} 
%i.i.d.\ sub-Gaussian noise
such as Gaussian and uniform observations.   {The extensions for Bernoulli observations, asymmetric mean tensors, and other practical issues are in Sections~\ref{subsec:ber} and \ref{subsec:exten}.}

\begin{figure}[ht!]
\centering
\includegraphics[width=\columnwidth]{alg_demo.pdf}
\caption{Illustration of our global-to-local algorithm.}\label{fig:demo}
\end{figure}

To construct algorithm guarantees,  we introduce the misclustering loss between an estimator $\hat z$ and the true $z$:
\begin{align}
     L(\hat z, z) = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind &\offf{ \hat z(i) = b } \\
     & \cdot \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2, \label{eq:defnofL}
\end{align}
where the superscript $\cdot^{s}$ denotes the normalized vector; i.e., $\ma^s:=\ma/\onorm{\ma}$ if $\ma\neq 0$ and $\ma^{s}=0$ if $\ma=0$ for any vector $\ma$. The following lemma indicates the close relationship between the loss $L(\hat z, z)$ and error $\ell(\hat z, z)$. The loss $L(\hat z, z)$ serves as an important intermediate quantity to control the misclustering error.

\begin{lem}[Relationship between misclustering error and loss]\label{lem:mis} {Consider the dTBM under the parameter space $\tP$.} Suppose $\min_{i \in [p]}\theta(i) > c$ for some constant $c > 0$. We have $\ell(\hat z, z) \Delta_{\min}^2 \leq L(\hat z, z)$.
%\fixme{miaoyan}{under bounded degree assumptions? please also check all lemmas in supp to make sure conditions are well specified.}
\end{lem}

\subsection{Weighted higher-order initialization}

We start with weighted higher-order clustering algorithm as initialization.   {We take an order-3 tensor and the clustering on the first mode as illustration for insight.} Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$. 
By model~\eqref{eq:model_tensor}, for all $i \in [p]$, we have
\begin{equation}\label{eq:kmeans}
    \theta(i)^{-1} \mX_{i:} = \off{\text{Mat}( \tS \times_2 \mTheta \mM \times_3  \mTheta \mM )}_{z(i):}. 
\end{equation}
This implies that, all node $i$ belonging to $a$-th community (i.e., $z(i)=a$) share the same normalized mean vector $\theta(i)^{-1} \mX_{i:}$, and vice versa. Intuitively, one can apply $k$-means clustering to the vectors $\{ \theta(i)^{-1} \mX_{i:} \}_{i\in[p]}$, which leads to main idea of our Sub-algorithm~\hyperref[alg:main]{1}.

Specifically, our initialization consists of denoising step and clustering step. The denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) estimates $\tX$ from $\tY$ by a double projection spectral method.  
The first projection performs HOSVD~\citep{de2000multilinear} via $\mU_{\text{pre},k} = \text{SVD}_{r} \of{ \text{Mat}_k(\tY) }, k \in [3]$, where $\text{SVD}_r(\cdot)$ returns the top-$r$ left singular vectors. The second projection performs HOSVD on the projected $\tY$ onto the multilinear Kronecker space $\mU_{\text{pre},k} \otimes  \mU_{\text{pre},k}$; i.e.,
\begin{equation}\label{eq:two-step_factor}
    \hat \mU_1 = \text{SVD}_{r} \of{\text{Mat}_1\of{ \tY\times_2  \mU_{\text{pre},2} \mU_{\text{pre},2}^T \times_3   \mU_{\text{pre},3} \mU_{\text{pre},3}^T }}.
\end{equation}
and similar for $ \hat \mU_2 ,  \hat \mU_3 $.
The final denoised tensor $\hat \tX$ is defined by
\begin{equation}\label{eq:two-step_est}
    \hat \tX = \tY \times_1 \hat \mU_1 \hat 
\mU^T_1 \times_2 \hat \mU_2 \hat \mU^T_2 \times_3 \hat \mU_3 \hat \mU^T_3. 
\end{equation}
The double projection improves usual matrix spectral methods in order to alleviate the noise effects for $K\geq 3$~\citep{han2020exact}.

The clustering step (lines 3-5 in Sub-algorithm~\hyperref[alg:main]{1}) performs the weighted $k$-means clustering. 
We write $\hat \mX=\text{Mat}_1(\hat \tX)$, and normalize the rows into $\hat \mX^s_{i:}=\onormSize{}{\hat \mX_{i:}}^{-1}\hat \mX_{i:}$ as a surrogate of $\theta(i)^{-1} \mX_{i:}$. Then, a weighted $k$-means clustering is performed on the normalized rows with weights equal to $\onormSize{}{\hat \mX_{i:}}^2$. The choice of weights is to bound the $k$-means objective function by the Frobenius-norm accuracy of $\hat \tX$. Unlike existing clustering algorithm~\citep{ke2019community}, we apply the clustering on the unfolded tensor $\hat \mX$ rather than on the factors $\hat \mU_1$. This strategy relaxes the singular-value gap condition~\citep{gao2018community, han2020exact}.
We assign degenerate rows with purely zero entries to an arbitrarily random cluster; these nodes are negligible in high-dimensions because of the lower bound on $\onormSize{}{\text{Mat}(\tS)_{a:}}$ in~\eqref{eq:family}. The final result gives the initial cluster assignment $z^{(0)}$. Full procedures for clustering are provided in Sub-algorithm~\hyperref[alg:main]{1}.

\begin{algorithm*}[th!]
\caption*{\bf Algorithm: Multiway spherical clustering for degree-corrected tensor block model }
\vspace{.15cm}
\begin{algorithmic}[1]
\Algphase{Sub-algorithm 1: Weighted higher-order initialization}
\INPUT Observation $\tY \in \bbR^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.

\State {
  Compute factor matrix $ \mU_{\text{pre},k} = \text{SVD}_{r} (\text{Mat}_k(\tY)), k \in [K]$ and the $(K-1)$-mode projection 
\begin{equation}
    \tX_{\text{pre},k} = \tY \times_1   \mU_{\text{pre},1} \mU_{\text{pre},1}^T \times_2 \cdots  \times_{k-1} \mU_{\text{pre},k-1} \mU_{\text{pre},k-1}^T \times_{k+1} \mU_{\text{pre},k-1} \mU_{\text{pre},k+1}^T\times_{K+1}  \times_K \mU_{\text{pre},K} \mU_{\text{pre}, K}^T.
\end{equation}
}
\State {

Compute factor matrix $\hat \mU_k = \text{SVD}_{r}(\text{Mat}_k(\tX_{\text{pre},k})), k \in [K]$ and denoised tensor
\begin{equation}
    \hat \tX = \tY \times_1 \hat \mU_1 \hat \mU^T_1 \times_2 \cdots \times_K \hat \mU_K \hat \mU^T_K.
\end{equation}
} 
\For{  {$k \in [K]$}}
\State {    Let $\hat \mX = \text{Mat}_k(\hat \tX)$ and $S_0=\{i \in [p]: \onormSize{}{\hat \mX_{i:}} = 0\}$. Set $\hat z(i)$ randomly in $[r]$ for $i \in S_0$.}
\State{  For all $i\in S_0^c$, compute normalized rows
$\hat \mX_{i:}^s :=\onormSize{}{\hat \mX_{i:}}^{-1} \hat \mX_{i:}.$
}
\State { Solve the clustering $\hat z_k \colon [p]\to[r]$ and centroids $ (\hat \mx_j)_{j\in[r_k]}$ using weighted $k$-means, such that}
\begin{align}
    &\sum_{i \in S_0^c }  \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{\hat z_k(i)} }^2 
    \leq 
    \eta \min_{\substack{\bar \mx_j, j\in[r], \bar z_k(i),i\in S_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\hat z_k(i)}}^2.
\end{align}
\EndFor

\OUTPUT {   Initial clustering $z^{(0)}_k \leftarrow \hat z_k, k \in [K]$.}

\Algphase{Sub-algorithm 2: Angle-based iteration}\label{alg:2}
\INPUT Observation $\tY \in \bbR^{p \times \cdots \times p}$, initialization $z^{(0)}_k \colon [p]\to[r], k \in [K]$ from Sub-algorithm 1, iteration number $T$.
\For {$t = 0$ to $T-1$}
\State Update the block tensor $\tS^{(t)}$ via
$\tS^{(t)} (a_1,...,a_K)= \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}_k(i_k) = a_k, k \in [K]\}.$
\For{ $k \in [K]$}
\State {   Calculate reduced tensor $\tY^{\text{d}}_k \in \bbR^{r \times r \times p \times r \times \cdots \times r}$ via
\begin{equation}
    \tY^{\text{d}}_k(a_1,\ldots,a_{k-1}, i ,a_{k+1},\ldots,a_K) 
    = \text{Ave}\{\tY(i_1,\ldots,i_{k-1},i,i_{k+1},\ldots,i_K): z^{(t)}(i_j) = a_j, j \neq k \}
\end{equation}}
% \begin{align}
%     &\tY^{\text{d}}(i,a_2,\ldots,a_K) 
%     = 
% \text{Ave}\{\tY(i,i_2,\ldots,i_K): z^{(t)}(i_k) = a_k, k \neq 1 \}.
% \end{align}

\State {   Let $\mY_k^{\text{d}} = \text{Mat}_k(\tY^{\text{d}})$ and $J_0 = \{ i\in[p]: \onorm{\mY^{\text{d}}_{i:}} = 0\}$. Set $z_k^{(t+1)}(i)$ randomly in $[r]$ for $i \in J_0$.}

\State {   Let $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. For all $i \in J_0^c$ update the cluster assignment by
\begin{equation}
    z(i)^{(t+1)}_k = \argmax_{a \in [r]} \cos \left( \mY^{\text{d}}_{k,i:},\ \mS^{(t)}_{a:} \right).
\end{equation}}
\EndFor
\EndFor

\OUTPUT {    Estimated clustering $z^{(T)}_k  \in [r]^{p}, k \in [K]$.}

\end{algorithmic}
\end{algorithm*}

We now establish the misclustering error rate of initialization.

\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with {fixed $r \geq 1$, $K \geq 2$}, i.i.d.\ noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}. Assume $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$ and let $\Delta_{\mX}$ denote the minimal gap in mean tensor defined in~\eqref{eq:delta_x}.  Let $ z^{(0)}_k$ denote the output of Sub-algorithm~\hyperref[alg:main]{1}. With probability going to 1, {as $p \rightarrow \infty$}, we have
\begin{equation}
    \ell(z^{(0)}_k, z) \lesssim {\sigma^2 r^K p^{-K/2}\over \Delta_{\mX}^2}.
\end{equation}
Further, assume $\mtheta$ is balanced as~\eqref{eq:degree}. We have
\begin{equation}\label{eq:ini}
 \ell(z_k^{(0)}, z) \lesssim {r^K p^{-K/2}\over \text{SNR}} \  \text{and} \ L(z^{(0)}_k, z) \lesssim  {\sigma^2 r^K p^{-K/2}},
\end{equation}
{with probability going to 1 as $p \rightarrow \infty$.}
\end{thm}

\begin{rmk}[Comparison to previous results] For fixed SNR, our initialization error rate with $K=2$ agrees with the initialization error rate $\tO(p^{-1})$ in matrix models~\citep{gao2018community}. Furthermore, in the special case of non-degree TBMs with $\theta_1=\cdots=\theta_p=1$, we achieve the same initial misclustering error $\tO(p^{-K/2})$ as in non-degree models~\citep{han2020exact}. Theorem~\ref{thm:initial} implies the advantage of our algorithm in achieving both accuracy and model flexibility. 
\end{rmk}

\begin{rmk}[Failure of conventional tensor HOSVD] If we use conventional HOSVD for tensor denoising; that is, we use $\mU_{\text{pre},k}$ in place of $\hat \mU,k$ in line 2, then the misclustering rate becomes $\tO(p^{-1})$ for all $K\geq 2$. This rate is substantially worse than our current rate~\eqref{eq:ini}.
\end{rmk}

\begin{rmk}[Singular-value gap-free clustering] Note that our clustering directly applies to the estimated mean tensor $\hat \tX$ rather than the leading tensor factors $\hat \mU_k$. Applying clustering to the tensor factors suffers from the non-identifiability issue due to the infinitely many orthogonal rotations when the number of blocks $r \geq 3$ in the absence of singular-value gaps. 
Such ambiguity causes the trouble for effective clustering~\citep{abbe2020entrywise}. In contrast, our initialization algorithm applies the clustering to the overall mean tensor $\hat \tX$. This strategy avoids the non-identifiability issue regardless of the number of blocks and singular-value gaps.  
\end{rmk}

\subsection{Angle-based iteration}\label{subsec:angle}
Our Theorem~\ref{thm:initial} has shown the polynomially decaying error rate from our initialization. Now we improve the error rate to exponential decay using local iterations. %\textcolor{blue}{We still focus on the first mode clustering for illustration.} 
We propose an angle-based local iteration to improve the outputs from Sub-algorithm~\hyperref[alg:main]{1}. 
To gain the intuition, consider an one-dimensional degree-corrected clustering problem with data vectors $\mx_i = \theta(i) \ms_{z(i)} + \mepsilon_i, i \in [p]$, where $\ms_i$'s are known cluster centroids, $\theta(i)$'s are unknown positive degrees, and $z\colon [p] \mapsto [r]$ is the cluster assignment of interest. The angle-based $k$-means algorithm estimates the assignment $z$ by minimizing the angle between data vectors and centroids; i.e., 
 \begin{equation}\label{eq:angle_kmeans}
     z(i) = \argmax_{a \in [r]} \cos ( \mx_i,\ \ms_{a} ),\ \text{ for all }i \in [p].
 \end{equation}
The classical Euclidean-distance based clustering~\citep{han2020exact} fails to recover $z$ in the presence of degree heterogeneity, even under noiseless case. In contrast, the proposed angle-based $k$-means achieves accurate recovery without explicit estimation of $\mtheta$.

Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as in angle-based $k$-means. We still take the order-3 tensor for illustration. Specifically, Sub-algorithm~\hyperref[alg:main]{2} updates estimated core tensor and cluster assignment in each iteration. We use superscript $\cdot^{(t)}$ to denote the estimate from $t$-th iteration, where $t=1,\ldots.$ For core tensor, we consider the following update strategy  
 \[
 \tS^{(t)}(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z^{(t)}_k(i_k)=a_k, k\in[3]\}.
\]
Intuitively, $\tS^{(t)}$ becomes closer to the true core $\tS$ as $z^{(t)}$ is more precise. For cluster assignment, we first aggregate the slices of $\tY$ and obtain a reduced tensor $\tY^{\text{d}}\in \bbR^{p \times r \times r}$ with given $z^{(t)}$, where
\[
\tY^{\text{d}}_1(i,a_2,a_3)=\text{Ave}\{\tY(i,i_2,i_3)\colon z^{(t)}_k(i_k)=a_k, k \neq 1\}.
\]
We use $\mY_k^d$ and $ \mS^{(t)}_k$ to denote the $\text{Mat}_k(\tY^{\text{d}})$ and $\text{Mat}_k(\tS^{(t)})$. The rows $\mY^d_{k,i:}$ and $\mS^{(t)}_{k,a:}$ correspond to the $\mx_i$ and $\ms_a$ in the one-dimensional clustering~\eqref{eq:angle_kmeans}. Then, we obtain the updated assignment by
\[
z_k(i)^{(t+1)} = \argmax_{a \in [r]} \cos \of{ \mY^{\text{d}}_{k,i:}, \mS^{(t)}_{k,a:} },\ \text{ for all } i \in [p],
\]
provided that $\mS^{(t)}_{k,a:}$ is a non-zero vector. Otherwise, if $\mS^{(t)}_{k,a:}$ is a zero vector, then we make the convention to assign $z^{(t+1)}_k(i)$ randomly in $[p]$. Full procedures for our angle-based iteration are described in Sub-algorithm~\hyperref[alg:main]{2}.

We now establish the misclustering error rate of iterations under the stability assumption. 
\begin{defn}[Locally linear stability] \label{def:stable}
Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\epsilon)=\{\bar z\colon \ell(\bar z, z)\leq \epsilon\}$. Let $\bar z\colon[p]\to [r]$ be a clustering function. We define two vectors associated with $\bar z$,
\begin{align}
    \mp(\bar z) & =(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T, \\
    \mp_{\mtheta}(\bar z)&=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
\end{align}

We call the degree is $\varepsilon$-locally linearly stable if and only if 
\begin{equation}\label{eq:local}
    \sin(\mp(\bar z),\ \mp_{\mtheta}(\bar z))\lesssim \varepsilon \Delta_{\min} ,\quad \text{for all } \bar z\in \tN(z, \varepsilon).
\end{equation}
\end{defn}

Roughly speaking, the vector $\mp(\bar z)$ represents the raw cluster sizes, and $\mp_{\theta}(\bar z)$ represents the relative cluster sizes weighted by degrees. 
The local stability holds trivially for $\varepsilon=0$ based on the construction of parameter space~\eqref{eq:family}. The condition~\eqref{eq:local} controls the impact of node degree to the $\mp_{\theta}(\cdot)$ with respect to the misclassification rate $\varepsilon$ and angle gap.  Intuitively, the condition~\eqref{eq:local} controls the skewness of degree so that the angle between raw cluster size and degree-weighted cluster size is well controlled. The stability assumption is proposed for technical convenience, and we relax this condition in numerical studies; see Section~\ref{sec:simulation}.

\begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the general sub-Gaussian dTBM with {fixed $r \geq 1$, $K \geq 2$}, independent noise under the parameter space $\tP$, and Assumption~\ref{assmp:min_gap}. {Assume the local linear stability of degree holds in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \geq \check{C}\log^{-1}p $ with some positive constant $\check{C}$.}
%Assume the local linear stability of degree holds in the neighborhood $\tN(z, \epsilon)$ for $\epsilon \geq \log^{-1}(p)$.
Let $\{z^{(0)}_k\}_{k=1}^K$ be the initialization for Sub-Algorithm~\hyperref[alg:main]{2} and $z^{(t)}_k$ be the $t$-th iteration output on $k$-th mode. Suppose $\min_{i \in [p]}\theta(i) \geq c $ for some {constant} $c > 0$, the $\text{SNR} \geq \tilde C p^{-(K-1)}\log p$ for some sufficiently large constant $\tilde C$, and the initialization satisfies 
%Consider the general sub-Gaussian dTBM with independent noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}. Assume the local linear stability of degree holds in the neighborhood $\tN(z, \epsilon)$ for $\epsilon \geq \log^{-1}(p)$. Let $\{z^{(0)}_k\}_{k=1}^K$ be the initialization for Sub-Algorithm~\hyperref[alg:main]{2} and $z^{(t)}_k$ be the $t$-th iteration output on $k$-th mode. Suppose $\min_{i \in [p]}\theta(i) \geq c $ for some $c > 0$, the $\text{SNR} \geq \tilde C p^{-(K-1)}\log p$ for some sufficiently large constant $\tilde C$, and the initialization satisfies 
\begin{equation}
    L(z^{(0)}_k, z) \lesssim \frac{\Delta_{\min}^2}{r \log p}, \quad k \in [K].
\end{equation}
 With probability going to 1 {as $p \rightarrow \infty$}, there exists a contraction parameter $\rho \in (0,1)$ such that 
 \small
\begin{align}\label{eq:final}
    \ell(z, \hat z^{(t+1)}_k) \lesssim &\ \KeepStyleUnderBrace{
   \text{SNR}^{-1}
    \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}}_{\substack{\text{statistical error}} }+ \KeepStyleUnderBrace{ \rho^t \ell(z, z^{(0)}_k). }_{\substack{\text{computational error}}}
\end{align}
\normalsize
\end{thm}
From the conclusion~\eqref{eq:final}, we find that the iteration error is decomposed into two parts: statistical error and computational error. The statistical error is unavoidable with noisy data regardless $t$, whereas the computational error decays in an exponential rate as the number of iterations $t \rightarrow \infty$.

\begin{cor}[Exact recovery of dTBM with weighted higher-order initialization]  Let the initialization $\{z^{(0)}_k\}_{k = 1}^K$ be the output from Sub-algorithm~\hyperref[alg:main]{1}. Assume $\text{SNR} \gtrsim p^{-K/2} \log p$. Combining  Combining {all parameter assumptions and the results in} Theorems~\ref{thm:initial} and \ref{thm:refinement}, with probability going to 1 {as $p \rightarrow \infty$}, our estimate $z^{(T)}_k$ achieves exact recovery within polynomial iterations; more precisely,
\begin{equation}
     z^{(T)}_k = \pi_k \circ z, \quad \text{for all }T\gtrsim \log_{1/\rho} p\  \text{and}\ k \in [K ].
\end{equation}
for some permutation $\pi_k \in \Pi$. 
\end{cor}
Therefore, our combined algorithm is \textit{computationally efficient} as long as SNR $\gtrsim p^{-K/2} \log p$. Note that, ignoring the logarithmic term, the minimal SNR requirement, $p^{-K/2}$, coincides with the computational lower bound in Theorem~\ref{thm:comp}. Therefore, our algorithm is optimal regarding the signal requirement and lies in the sharpest \emph{computationally efficient} regime in Figure~\ref{fig:phase_axis}.

\subsection{Extension to Bernoulli observations}\label{subsec:ber}

Bernoulli or network observations are common in multiple fields. Our iteration Theorem~\ref{thm:refinement} holds for Bernoulli models, but our initialization Theorem~\ref{thm:initial} does not. Moreover, our current dTBM is insufficient to address sparsity with decaying mean tensor. Here, we provide extra discussions for Bernoulli initialization and strategies under sparse settings.

\begin{itemize}[wide]
    \item \textit{Extension to dense binary dTBMs.} The main difficulty to establish initialization guarantees for Bernoulli observations lies in the denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}). We now provide a high-level explanation for the technical difficulty when applying Theorem~\ref{thm:initial} to Bernoulli observations.

    The derivation of Theorem~\ref{thm:initial} relies on the upper bound of the estimation error for the mean tensor in Lemma~\ref{lem:two-step_esterror}; i.e., with high probability
\begin{equation}\label{eq:prop1}
    \onormSize{}{\hat \tX - \tX}_F^2 \lesssim p^{K/2},
\end{equation}
where $\tX = \bbE\tY$ and $\hat \tX$ is defined in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}. Unfortunately, the inequality~\eqref{eq:prop1} holds only for i.i.d.\ sub-Gaussian observations, while Bernoulli observations are generally not identically distributed.

One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square}, $\Mat_{sq}(\cdot)$, of Bernoulli tensor $\tY \in \{0,1\}^{p_1 \times \cdots \times p_K}$. Specifically, the square matricization $\mat_{sq} (\tY) \in \{0,1\}^{p^{\floor{K/2}} \times p^{\ceil{K/2}} }$ has entries $[\Mat_{sq}(\tY)](j_1, j_2) = \tY(i_1, \ldots, i_K)$, where
\begin{align}
    j_1 &= i_1 + p_1(i_2 - 1) + \cdots + p_1 \cdots p_{\floor{K/2}-1} (i_{\floor{K/2}} - 1),\\
    j_2 &= i_{\ceil{K/2}} + p_{ \ceil{K/2}}(i_{\ceil{K/2} + 1} - 1) + \cdots \\
    & \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + p_{ \ceil{K/2}} \cdot p_{K-1} (i_K-1).
\end{align}
The matrix $\Mat_{sq}(\tY)$ is asymmetric. We are able to interpret $\Mat_{sq}(\tY)$ as the adjacency matrix for a bipartite network with connections between two groups of node. The two groups of nodes in the bipartite network have $p_1 \cdots p_{\floor{K/2}}$ and $ p_{\ceil{K/2}} \cdots p_K$ nodes, respectively. The entry $[\Mat_{sq}(\tY)](j_1, j_2)$ refers to the presence of connection between the nodes indexed by combinations $(i_1, \ldots, i_{\floor{K/2}})$ and $(i_{\ceil{K/2}}, \ldots, i_K)$. We summarize the procedure in Sub-algorithm 3.

\begin{algorithm}[h!]
\caption*{\bf Sub-algorithm 3: Weighted higher-order initialization for Bernoulli observation}
\vspace{.15cm}
\begin{algorithmic}[1] 
\INPUT Bernoulli tensor $\tY \in \{0,1\}^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
\State  Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{p^{\floor{K/2}} \times p^{\ceil{K/2}} }$ denote the nearly square unfolded tensor. Compute the estimate $\tX'$, where
\small
\begin{equation}\label{eq:matrixsvd}
    \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX) -  \mat_{sq}(\tY)}_F^2.
\end{equation}
\normalsize
\State Implement lines 3-5 of Sub-algorithm~\hyperref[alg:main]{1} with $\hat \tX$ replaced by $\hat \tX'$ in~\eqref{eq:matrixsvd}.
\OUTPUT Initial clustering $z^{(0)} \leftarrow \hat z$.
\end{algorithmic}
\end{algorithm}

\begin{prop}[Error for Bernoulli initialization]\label{prop:ber} Consider the Bernoulli dTBM in the parameter space $\tP$ {with fixed $r \geq 1, K \geq 2$} and Assumption~\ref{assmp:min_gap} holds. Assume that $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$. Let $ z^{(0)}$ denote the output of Sub-algorithm~3. With probability going to 1 {as $p \rightarrow \infty$}, we have
\begin{equation}\label{eq:ini_b}
 \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}, \quad \text{and} \quad L(z^{(0)}, z) \lesssim  {\sigma^2 r^K p^{-\floor{K/2}}}.
\end{equation}
\end{prop}

\begin{rmk}[Comparison with Gaussian model] The Bernoulli bound $\tO(p^{- \floor{K/2} })$ in Proposition~\ref{prop:ber} is relatively looser than the Gaussian bound $\tO(p^{- K/2 })$ in Theorem~\ref{thm:initial}. The gap between Bernoulli and Gaussian error decreases as the order $K$ increases. Nevertheless, combining with angle iteration Sub-algorithm~\hyperref[alg:main]{2}, Bernoulli clustering still achieves exponential error rate $\exp \of{ - p^{(K-1)}}$ at a price of a larger SNR.
%Nevertheless, the Bernoulli clustering still achieves exponential error rate $\exp \of{ - p^ \floor{K/2}}$ with large SNR combining with our angle iteration Sub-algorithm~\hyperref[alg:main]{2}. 
%The exponential rate is tighter than the rate error $\tO(p^{-2})$ in previous work~\citep{ke2019community}. 
The investigation of the gap between upper bound $p^{- \floor{K/2}}$ and the lower bound $p^{- K/2}$ for Bernoulli tensors will be left as future work. In numerical experiments, we will use our original initialization, Sub-algorithm 1, to verify the robustness to Bernoulli observations.
\end{rmk}

\begin{rmk}[Comparison with previous methods]  Previous work \citep{ke2019community} develops a spectral clustering method for Bernoulli dTBM. \cite{ke2019community} adopts a different signal notion based on the singular gap in the core tensor, denoted as $\Delta_{\rm singular}$. By \cite[Theorem 1]{ke2019community}, the spectral method achieves exact recovery with $\Delta_{\rm singular} \gtrsim p^{-1/2}$. However, we are not able to infer the exact recovery of spectral method by our angle-base SNR condition. Consider an order-2 dTBM with $p > 2, \sigma^2 = 1$, $\mtheta = \mathbf{1}$, equal size assignment $|z^{-1}(a)| = p/r$ for all $ a \in [r]$, and core matrix $\mS = \mI_2$. The singular gap under this setting is $\Delta_{\rm singular} = \min\{ \lambda_1 - \lambda_2, \lambda_2 \} = 0$, where $\lambda_1 \geq \lambda_2$ are singular values of $\mS$. In contrast, our angle gap $\Delta_{\min}^2 = 2$ satisfies the SNR condition in Theorem~\ref{thm:refinement}. Then, our algorithm achieves the exact recovery, but the spectral method in \cite{ke2019community} fails.
% \begin{equation}
%     \mS = \begin{pmatrix} 1 &  0 \\
%     0 & 1
%     \end{pmatrix}, 
% \end{equation}
%for $\gamma \leq 0$. 
% The singular gap under this setting is defined as $\Delta_{\rm singular} = \min\{ \lambda_1 - \lambda_2, \lambda_2 \}$, where $\lambda_1 \geq \lambda_2$ are singular values of $\mS$. When $\gamma_1 = \gamma_2 = 0$, we have angle gap $\Delta_{\min}^2 = 2$ larger than the singular gap $\Delta_{\rm singular}^2 = 0$. Our algorithm achieves the exact recovery, while, the spectral method in \cite{ke2019community} fails.
%; when $\gamma_1 > \gamma_2 = -1$, we have angle gap $\Delta_{\min}^2 < p^{\gamma_1}$ smaller than the singular gap $\Delta_{\rm singular}^2 > p^{\gamma_1}$.

Hence, for fair comparison, we compare the best performance of our algorithm and \cite{ke2019community} under the strongest signal setting of each model. Since both methods contain an iteration procedure, we set the iteration number to infinity to avoid the computational error. Considering the largest angle-based SNR $\asymp 1$ in Theorem~\ref{thm:refinement}, our Bernoulli clustering achieves exponential error rate of order $\exp(-p^{(K-1)})$; considering the largest singular gap $\Delta_{\rm singular} \asymp 1$ in Theorem 1 of \cite{ke2019community}, the spectral clustering has a polynomial error rate of order $p^{-2}$. Our algorithm still shows a better theoretical accuracy than the competitive work for Bernoulli observations. 
\end{rmk}

\item \textit{Extension to sparse binary dTBMs.} The sparsity is often a popular feature in hypergraphs~\citep{florescu2016spectral,ke2019community, ahn2018hypergraph}. Specifically, the sparse binary dTBM assumes that, the entries of $\tY$ follow independent Bernoulli distributions with the mean
     \begin{equation}\label{eq:sparse_dtbm}
    \bbE\tY = \alpha_p \tS \times_1 \mTheta \mM \times_2  \cdots \times_K \mTheta \mM,
\end{equation}
where the extra scalar parameter $\alpha_p \in (0,1]$ is function of $p$ that controls the sparsity. A smaller $\alpha_p$ indicates a higher level of sparsity. Our current work focuses on dense dTBM with $\alpha_p=1$.  While sparse dTBM is an interesting application, the algorithm and its analysis require different techniques. Below, we discuss possible modifications of the algorithm.

The sparsity affects our initialization guarantee in our Theorem~\ref{thm:initial}. In our initialization, the spectral denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) implements matrix SVD to unfolded tensors. %, and Theorem~\ref{thm:initial} indicates the good performance under the dense case. 
However, SVD-based methods are believed to fail in extremely sparse SBM due to the localization phenomenon in the singular vectors \citep{florescu2016spectral}. Inspired by \cite{florescu2016spectral}, we adapt the diagonal-deleted HOSVD (D-HOSVD) \citep{ke2019community} as the initialization in our higher-order clustering.

The sparsity also affects the iteration guarantee in our Theorem~\ref{thm:refinement}. The decaying mean tensor leads to a worse statistical error of order $\tO( -\alpha_p p^{K-1})$ on $\hat \tX$. The theoretical analyses for sparse binary dTBM and algorithms are left as future direction. Instead, we add numerical experiments to evaluate the robustness of our algorithm and the improvement of D-HOSVD initialization in the sparse dTBM; see Appendix I.

\end{itemize}

\subsection{Practical issues}~\label{subsec:exten}

{\bf Computational complexity.} Our two-stage algorithm has a computational cost polynomial in tensor dimension $p$. Specifically, the complexity of Sub-algorithm~\hyperref[alg:main]{1} is $\tO(K p^{K+1} + Krp^K)$, where the first term is contributed by the double projection and the calculation of $\hat \tX$, and the second term comes from normalization and the $k$-means. The cost of each update in Sub-algorithm~\hyperref[alg:main]{2} is $\tO(p^K + pr^K)$, where $p^K$ comes from the calculation of $\tS^{(t)}$ and $\tY^{\text{d}}$, and $pr^K$ comes from the normalization of $\tY^{\text{d}}$, the calculation of $\tS^{(t)}$, and the cluster assignment update in Step~10.

{\bf Hyper-parameter selection.} In our theoretical analysis, we have assumed the true cluster number $r$ is given to our algorithm. In practice, the cluster number $r$ is often unknown, and we now propose a method to choose $r$ from data. We impose the Bayesian information criterion (BIC) and choose the cluster number that minimizes BIC; i.e., under the symmetric Gaussian dTBM~\eqref{eq:model_tensor},
\begin{align}\label{eq:BIC}
 \hat r = \argmin_{r \in \mathbb{Z}_+ } \of{p^K \log(\onormSize{}{\hat \tX - \tY}_F^2)  + p_e(r) K \log p},
\end{align}
with $\hat \tX = \hat \tS(r) \times_1 \hat \mTheta(r) \hat \mM(r) \times_2 \cdots \times_K \hat \mTheta(r) \hat \mM(r),$ where the triplet $(\hat z(r), \hat \mS(r),\hat \mtheta(r))$ are estimated parameters with cluster number $r$, and $p_e (r)= r^K + p(\log r + 1) - r$ is the effective number of parameters. Note that we have added the argument $(r)$ to related quantities as functions of $r$. In particular, the estimate $\hat \mtheta(r)$ in~\eqref{eq:BIC} is obtained by first calculating the reduced tensor $\hat \tY^{\text{d}}$ with $\hat z(r)$, and then normalizing the row norms $\onormSize{}{\hat \mY_{i:}^{\text{d}}}$ to 1 in each cluster; i.e., 
\begin{equation}
 \hat \mtheta(r)=(\hat \theta(1,r),\ldots,\hat \theta(p,r))^T,
\end{equation}
with $\hat \theta(i,r) = {\onormSize{}{\hat \mY^{\text{d}}(r)_{i:}}}/{\sum_{j: \hat z(j,r) =  \hat z(i,r)} \onormSize{}{\hat \mY^{\text{d}}(r)_{j:}}}$, $\hat \mY^{\text{d}}(r) = \mat(\hat \tY^{\text{d}}(r))$, $\hat \tY^{\text{d}}(r)(i,a_2,\ldots,a_K) =  \text{Ave}\{\tY(i,i_2,\ldots,i_K): \hat z(i_k, r) = a_k, k \neq 1 \}$, and $\hat z(i,r)$ denotes the community label for the $i$-th node with given cluster number $r$. We evaluate the performance of the BIC criterion in Section~\ref{subsec:num_theory}.

\section{Comparison with Non-degree Tensor Block Model} \label{sec:tbm}

We discuss the connections and difference between dTBM and TBM~\citep{han2020exact} from three aspects: signal notions, theoretical results, and algorithms. Without loss of generality, let $\sigma^2=1$.

\begin{itemize}[wide]
    \item \textit{Signal notion.} The signal levels in both TBM~\citep{han2020exact} and our dTBM are functions of the core tensor $\tS$. We emphasize that the signal notions are different between the two models. In particular, the Euclidean-based signal notion in TBM~\cite{han2020exact} fails to accurately describe the phase transition in our dTBM due to the possible heterogeneity in degree $\mtheta$. To compare, we denote our angle-based signal notion in~\eqref{eq:gamma} and the Euclidean-based SNR in \cite{han2020exact} as $\Delta_{\text{ang}}^2$ and $\Delta_{\text{Euc}}^2$, respectively:
    \begin{align}
        \Delta_{\text{ang}}^2 &=  2(1 - \max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}} ), \\ 
 \Delta_{\text{Euc}}^2 &= \min_{a \neq b \in [r]} \onormSize{}{\mS_{a:} - \mS_{b:}}^2.
    \end{align}
By Lemma~\ref{lem:norm_diff} in the Appendix II, we have 
\begin{equation}\label{eq:signalcompare}
     \Delta_{\text{ang}}^2  \max_{a \in [r]}\onormSize{}{\mS_{a:}}^2 \leq \Delta_{\text{Euc}}^2.
\end{equation}
The above inequality indicates the condition $\Delta_{\text{Euc}}^2 \leq p^{\gamma}$ is sufficient but not necessary for $\Delta_{\text{ang}}^2 \leq p^{\gamma}$. In fact, if we were to use $\Delta_{\text{Euc}}^2$ for both models, then the phase transition of dTBM can be arbitrarily worse than that for TBM.

Here, we provide an example to illustrate the dramatical difference between TBM and dTBM with the same core tensor.

\begin{example}[Comparison with Euclidean-based signal notion] \label{example:euc_alg} Consider a biclustering model with $\mtheta=1$ and an order-2 core matrix 
\begin{equation}
    \mS = \begin{pmatrix} p^{(\gamma+1)/2 } + 2  & 2 p^{(\gamma+1)/2} + 4\\
    2 & 4
    \end{pmatrix},\quad \text{with}\ \gamma \leq -1.
\end{equation}
The core matrix $\mS$ lies in the parameter spaces of TBM and our dTBM. Here, the constraint $\gamma \leq -1$ is added to ensure the bounded condition of $\mS$ in our parameter space in \eqref{eq:family}. The angle-based and Euclidean-based signal levels of $\mS$ are 
\begin{equation}
    \Delta_{\text{ang }}^2(\mS) = 0 \ \left(\leq p^{\gamma}\right), \quad \Delta_{\text{Euc}}^2(\mS) = 5 p^{\gamma + 1} \ \left(\geq p^{\gamma}\right).
\end{equation}
We conclude that TBM with $\mS$ achieves exact recovery with a polynomial-time algorithm; see \cite[Theorem 4]{han2020exact}. By contrast, the dTBM with the same $\mS$ and input $r=2$ violets the identifiability condition, and thus fails to be solved by all estimators; see our Theorem~\ref{thm:unique}. 
\end{example}

    \item \textit{Theoretical results.} In both works, we study the phase transition of TBM and dTBM with respect to the Euclidean and angle-based SNRs. We briefly summarize the results in \cite{han2020exact} and compare with ours.

    \textit{Statistical critical value:}
    \begin{align}
        \text{Ours:}& \ \Delta_{\text{ang}}^2 \lesssim p^{-(K-1)} \Rightarrow \text{statistically impossible;} \\
        & \ \Delta_{\text{ang}}^2 \gtrsim   p^{-(K-1)} \Rightarrow \text{MLE achieves exact recovery;} \\
        \text{Han's:}& \ \Delta_{\text{Euc}}^2 \lesssim p^{-(K-1)} \Rightarrow \text{statistically impossible;} \\
        &\ \Delta_{\text{Euc}}^2 \gtrsim   p^{-(K-1)} \Rightarrow \text{MLE achieves exact recovery}.
    \end{align}

     \textit{Computational critical value:}
    \begin{align}
        \text{Ours:}& \ \Delta_{\text{ang}}^2 \lesssim p^{-K/2} \Rightarrow \text{computationally impossible;} \\
        &\ \Delta_{\text{ang}}^2 \gtrsim   p^{-K/2} \Rightarrow \text{ computationally efficient;} \\
        \text{Han's:}& \ \Delta_{\text{Euc}}^2 \lesssim p^{-K/2} \Rightarrow \text{computationally impossible;} \\
        &\ \Delta_{\text{Euc}}^2 \gtrsim   p^{-K/2} \Rightarrow \text{computationally efficient}.
    \end{align}

    The above comparison reveals four major differences.

    First, none of our results in Section~\ref{sec:limits} are corollaries of \cite{han2020exact}. Both models show the similar conclusion but under different conditions. While the TBM impossibility~\citep{han2020exact} provides a necessary condition for our dTBM impossibility, we find that such a condition is often loose. There exists a regime of $\tS$ in which TBM problems are computationally efficient but dTBM problems are statistically impossible; see Example~\ref{example:euc_alg}. This observation has motivated us to develop the new signal notion $\Delta^2_{\text{ang}}$ for sharp dTBM phase transition conditions.

 Second, to find the phase transition, we need to show both the impossibility and achievability when SNR is below and above the critical value, respectively. While the TBM impossibility can serve as a loose condition of our dTBM impossibility, more efforts are required to show the achievability. In particular, since TBM is a more restrictive model than dTBM, the achievability in \cite{han2020exact} does not imply the achievability of dTBM in a larger parameter space. The latter requires us to develop new MLE and polynomial algorithms for dTBM achievability.

 Third, from the perspective of proofs, we develop new dTBM-specific techniques to handle the extra degree heterogeneity. In our Theorem~\ref{thm:stats}, we construct a special non-trivial degree heterogeneity to establish the lower bound for arbitrary core tensor with small angle gap, while, TBM~\citep{han2020exact} considers the constructions without degree parameter.
%we consider the profile MLE by maximizing out the nuisance degree parameter, while TBM~\citep{han2020exact} considers the usual MLE without degree parameter.
In our Theorem~\ref{thm:comp}, we construct a rank-2 tensor to relate HPC conjecture to $\Delta^2_{\text{ang}}$, while TBM~\citep{han2020exact} constructs a rank-1 tensor to relate HPC conjecture to $\Delta^2_{\text{Euc}}$. The asymptotic non-equivalence between $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$ renders our proof technically more involved.

 Last, we discuss the statistical impossibility statements. Our Theorem~\ref{thm:stats} implies the statistical impossibility whenever the core tensor $\tS$ leads to an angle-based SNR below the critical threshold. The Theorem 6 in \cite{han2020exact} implies the worst case statistical impossibility for a particular core tensor $\tS$ with Euclidean-based SNR below the statistical limit. Hence, our Theorem~\ref{thm:stats} shows a stronger statistical impossibility for dTBM than that presented in TBM \cite[Theorem 6]{han2020exact}. However, inspecting the proof of \cite{han2020exact}, the proof of Theorem 6 indeed implies a stronger TBM impossibility statement for arbitrary core tensor; i.e., when $\gamma < -(K-1)$
 \small
 \begin{equation}
     \liminf_{p \rightarrow \infty} \inf_{ \tS \in \tP_{\tS, {\rm TBM}} \cap \{\Delta_{\rm Euc}^2 = p^{\gamma} \} }\inf_{\hat z_{\rm stats}}  \sup_{z \in \tP_{z, {\rm TBM}}}  \bbE[p \ell(\hat z_{\rm stats}, z)] \geq 1,
 \end{equation}
 \normalsize
 where $\tP_{\tS, {\rm TBM}}$ and $\tP_{z, {\rm TBM}}$ refer to the space for core tensor $\tS$ and assignment $z$ under TBM, respectively. Again, in terms of the strong statistical impossibility, both models show similar conclusions but under different conditions. Since two impossibilities consider different core tensor regimes with non-equivalent $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$, we emphasize that different proof techniques are required to obtain these similar conclusions. See our proof sketch in Section~\ref{subsec:impossible}, Appendices~\ref{sec:statprove1} and \ref{sec:compprove1} for detail technical differences.

 \item \textit{Algorithms.} Both \cite{han2020exact} and our work propose the two-step algorithm, which combines warm initialization and iterative refinement to achieve exact recovery. This local-to-global strategy is not new in clustering literature~\citep{gao2022iterative, chien2019minimax}. The highlight of our algorithm is the angle-based update in lines 10-14, Sub-algorithm~\hyperref[alg:main]{2}, which is specifically designed for dTBM to avoid the estimation of $\mtheta$. This angle-based update brings new proof challenges. We develop polar-coordinate based techniques to establish the error rate for the proposed algorithm. "
\end{itemize}

\section{Numerical studies}\label{sec:simulation}

 We evaluate the performance of the weighted higher-order initialization and angle-based iteration in this section. We report average errors and standard deviations across 30 replications in each experiment. Clustering accuracy is assessed by clustering error rate (CER, i.e., one minus rand index). Note that CER between $(\hat z, z)$ is equivalent to misclustering error $\ell(\hat z, z)$ up to constant multiplications~\citep{meilua2012local}, and a lower CER indicates a better performance.

We generate order-3 tensors with \emph{assortative}~\citep{gao2018community} core tensors to control SNR; i.e., we set $\tS_{aaa} = s_1$ for $a \in [r]$ and others be $s_2$, where $s_1 > s_2 > 0$. Let $\alpha = s_1/s_2$. We set $\alpha$ close to 1 such that $1-\alpha=o(p)$. In particular, we have $\alpha = 1 + \Omega(p^{\gamma/2})$ with $\gamma<0$ by Assumption~\ref{assmp:min_gap} and definition~\eqref{eq:gamma}. Hence, we easily adjust SNR via varying $\alpha$. Note that the assortative setting is proposed for simulations, and our algorithm is applicable for general tensors in practice. The cluster assignment $z$ is randomly generated with equal probability across  $r$ clusters for each mode. Without further explanation, we generate degree heterogeneity $\mtheta$ from absolute normal distribution by $\theta(i) = |X_i| + 1 - 1/\sqrt{2\pi}$ with $|X_i| \stackrel{\text{i.i.d.}}\sim N(0,1), i \in [p]$ and normalize $\mtheta$ to satisfy \eqref{eq:family}. Also, we set $\sigma^2 = 1$ for Gaussian data without further specification.

\subsection{Verification of theoretical results}\label{subsec:num_theory}

The first experiment verifies  statistical-computational gap described in Section~\ref{sec:limits}. Consider the Gaussian model with $p = \{80, 100\}$, $r = 5$. We vary $\gamma $ in $ [-1.2, -0.4]$ and $[-2.1, -1.4]$ for matrix ($K=2$) and tensor $(K = 3)$ clustering, respectively. Note that finding MLE under dTBM is computationally intractable. We approximate MLE using an oracle estimator, i.e., the output of Sub-algorithm~\hyperref[alg:main]{2} initialized from true assignment. Figure~\ref{fig:phase}a shows that both our algorithm and oracle estimator start to decrease around the critical value $\gamma_{\text{stat}}  = \gamma_{\text{comp}}  = -1$ in matrix case. In contrast, Figure~\ref{fig:phase}b shows a significant gap in the phase transitions between the algorithm estimator and oracle estimator in tensor case. The oracle error rapidly decreases to 0 when $\gamma_{\text{stat}} = -2$, whereas the algorithm estimator tends to achieve exact clustering when $\gamma_{\text{comp}} = -1.5$. Figure~\ref{fig:phase} confirms the existence of the statistical-computational gap in our Theorems~\ref{thm:stats} and~\ref{thm:comp}.

\begin{figure}[htb]
    \centering
    \includegraphics[width=\columnwidth]{phase_anno3.pdf}
    \caption{SNR phase transitions for clustering in dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.
    }
    \label{fig:phase}
\end{figure}

The second experiment verifies the performance guarantees of two algorithms: (i) weighted higher-order initialization; (ii) combined algorithm of weighted higher-order initialization and angle-based iteration. We consider both the Gaussian and Bernoulli models with $p = \{80, 100\}$, $r = 5$, $\gamma \in [-2.1, -1.4]$. Figure~\ref{fig:ini_re} shows the substantial improvement of combined algorithm over initialization, especially under weak and intermediate signals. This phenomenon agrees with the error rates in Theorems~\ref{thm:initial} and \ref{thm:refinement} 
and confirms the necessity of the local iterations.

\begin{figure}[htp!]
    \centering
     \includegraphics[width=\columnwidth]{ini_re_anno3.pdf}
    \caption{CER versus signal exponent $(\gamma)$ for initialization only and for combined algorithm. We set $p = \{80, 100\}, r = 5, \gamma \in [-2.1, -1.4]$ under (a) Gaussian models and (b) Bernoulli models. }
    \label{fig:ini_re}
\end{figure}

\begin{table*}[hbt]
\centering
    \begin{tabularx}{\textwidth}{c *{8}{Y}}
    \toprule
    Settings & \multicolumn{2}{c}{$p = 50, \sigma^2 = 0.25$} & \multicolumn{2}{c}{$p = 50, \sigma^2 = 1$} & \multicolumn{2}{c}{$p = 80, \sigma^2 = 0.25$} & \multicolumn{2}{c}{$p = 80, \sigma^2 = 1$}\\
    \cmidrule(lr){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9}
         True cluster number $r$ & 2 & 4 & 2  & 4 & 2  & 4 & 2 & 4  \\
         \midrule
         Estimated cluster number $\hat r$ &  2(0)  & 3.9(0.25)& 2(0)    & 3.1(0.52) & 2(0)    & 4(0)   & 2(0)    & 3.9(0.31)   \\
     \bottomrule
    \end{tabularx}
    \caption{Estimated cluster number given by BIC criterion under the low noise level $(\sigma^2 = 0.25)$ and high noise level $(\sigma^2 = 0.5)$ settings. Numbers in parentheses are standard deviations of $\hat r$ over 30 replications.}
    \label{tab:select}
\end{table*}

The third experiment evaluates the empirical performance of the BIC criterion to select unknown cluster number. We generate the data from an order-3 Gaussian model with $p = \{50,80\}$, $r = \{2,4\}$, and noise level $\sigma^2 \in \{ 0.25,1\}$. Table~\ref{tab:select} shows that our BIC criterion well chooses the true $r$ under most settings.  Note that the BIC slightly underestimates the true cluster number $(r = 4)$ with smaller dimension and higher noise $(p = 50, \sigma=1)$, and the accuracy immediately increases with larger dimension $p = 80$. The improvement follows from the fact that a larger dimension $p$ indicates a larger sample size in the tensor block model. Therefore, we conclude that BIC criterion is a reasonable way to tune the cluster number.

\subsection{Comparison with other methods}\label{subsec:comp}

We compare our algorithm with following higher-order clustering methods:
\begin{itemize}[wide,topsep=-3pt,itemsep=0pt,parsep=1pt]
    \item \textbf{\small HOSVD}: HOSVD on data tensor and $k$-means on the rows of the factor matrix;
    \item \textbf{\small HOSVD+}: HOSVD on data tensor and $k$-means on the $\ell_2$-normalized rows of the factor matrix;
    \item \textbf{\small HLloyd}~\citep{han2020exact}: High-order clustering algorithm developed for non-degree tensor block models;
    \item \textbf{\small SCORE}~\citep{ke2019community}: Tensor-SCORE for clustering developed for sparse binary tensors.
\end{itemize}

Among the four alternative algorithms, the \textbf{\small SCORE} is the closest method to ours. We set the tuning parameters of \textbf{\small SCORE} as in previous literature \citep{ke2019community}. The methods \textbf{\small SCORE} and \textbf{\small HOSVD+} are designed for degree models, whereas \textbf{\small HOSVD} and \textbf{\small HLloyd} are designed for non-degree models. We conduct two experiments to assess the impacts of (i) signal strength and (ii) degree heterogeneity, based on Gaussian and Bernoulli models with $ p = 100, r = 5$. We refer to our algorithm as \textbf{\small dTBM} in the comparison.

We investigate the effects of signal to clustering performance by varying $\gamma \in [-1.5, -1.1]$. Figure~\ref{fig:comp_gamma} shows that our method \textbf{\small dTBM} outperforms all other algorithms. The sub-optimality of \textbf{\small SCORE} and \textbf{\small HOSVD+} indicates the necessity of local iterations on the clustering. Furthermore,  Figure~\ref{fig:comp_gamma} shows the inadequacy of non-degree algorithms in the presence of mild degree heterogeneity. 
The experiment demonstrates the benefits of addressing heterogeneity in higher-order clustering tasks.

\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{comp_gamma_anno3.pdf}
    \caption{CER versus signal exponent (denoted $\gamma$) for different methods. We set $p = 100, r = 5, \gamma \in [-1.5, -1.1]$ under (a) Gaussian and (b) Bernoulli models.}
    \label{fig:comp_gamma}
\end{figure}

The only exception in Figure~\ref{fig:comp_gamma} is the slightly better performance of \textbf{\small HLloyd} over \textbf{\small HOSVD+} under Gaussian model. However, we find the advantage of \textbf{\small HLloyd} disappears with higher degree heterogeneity. We perform extra simulations to verify the impact of degree effects. We use the same setting as in the first experiment in the Section~\ref{subsec:comp}, except that we now generate the degree heterogeneity $\mtheta$ from Pareto distribution prior to normalization. The density function of Pareto distribution is $f(x|a,b) = a b^a x^{-(a+1)} \ind\{ x \geq b \}$, where $a$ is called \emph{shape} parameter. We vary $a \in \{2,6\}$ and choose $b$ such that $\bbE X = a(a - 1)^{-1}b = 1$ for $X$ following Pareto$(a,b)$. Note that a smaller $a$ leads to a larger variance in $\mtheta$ and hence a larger degree heterogeneity. We consider the Gaussian model under low $(a = 6)$ and high $(a = 2)$ degree heterogeneity. Figure~\ref{fig:comp_gamma_theta} shows that the errors for non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) increase with degree heterogeneity. In addition, the advantage of \textbf{\small HLloyd} over \textbf{\small HOSVD+} disappears with higher degree heterogeneity.

\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{comp_theta_anno3.pdf}
    \caption{CER versus shape parameter in degree (denoted $a\in[3,6]$) for different methods. We set $p = 100, r = 5, \gamma = -1.2$ under (a) Gaussian and (b) Bernoulli models.}
    \label{fig:comp_theta}
\end{figure}

\begin{figure}[htp!]
    \centering
    \includegraphics[width=\columnwidth]{comp_gamma_theta3.pdf}
    \caption{CER comparison versus signal exponent (denoted $\gamma$) under (a) low (shape parameter $a = 6$)  (b) high (shape parameter $a = 2$) degree heterogeneity. We set $p = 100, r = 5, \gamma \in [-1.5, -1.1]$ under Gaussian model.}
    \label{fig:comp_gamma_theta}
\end{figure}

The last experiment investigates the effects of degree heterogeneity to clustering performance. We fix the signal exponent $\gamma = -1.2$ and vary the extent of degree heterogeneity. In this experiment, we generate $\mtheta$ from Pareto distribution prior to normalization. We vary the shape parameter $a \in [3,6]$ in the Pareto distribution to investigate a range of degree heterogeneities. Figure~\ref{fig:comp_theta} demonstrates the stability of degree-corrected algorithms (\textbf{\small dTBM}, \textbf{\small SCORE}, \textbf{\small HOSVD+}) over the entire range of degree heterogeneity under consideration. In contrast, non-degree algorithms (\textbf{\small HLloyd}, \textbf{\small HOSVD}) show poor performance with large heterogeneity, especially in Bernoulli cases. This experiment, again, highlights the benefit of addressing degree heterogeneity in higher-order clustering.

\section{Real data applications}\label{sec:real}
\subsection{Human brain connectome data analysis}

The Human Connectome Project (HCP) aims to construct the structural and functional neural connections in human brains~\citep{van2013wu}. We preprocess the original dataset following \cite{desikan2006automated} and partition the brain into 68 regions. The cleaned dataset includes brain networks for 136 individuals. Each brain network is represented by a 68-by-68 binary symmetric matrix, where the entry with value 1 indicates the presence of connection between node pairs, while the value 0 indicates the absence. We use $\tY \in \{0,1\}^{68 \times 68 \times 136}$ to denote the binary tensor. Individual attributes such as gender and sex are recorded.

We apply our general asymmetric algorithm to the HCP data with the numbers of clusters on three modes $r_1 = r_2 = 4$ and $r_3 = 3$. The selection of $r_1$ and $r_2$ follows the human brain anatomy and the symmetry in the brain network, and the $r_3$ is specified following previous analysis~\citep{hu2021generalized}. Because of the symmetry in the data, the estimated brain node clustering results are the same on the first and second modes. Figure~\ref{fig:cluster_brain} shows that brain connection exhibits a strong spatial separation structure. Specifically, the first cluster, named \emph{L.Hemis}, involves all the nodes in the left hemisphere. The nodes in the right hemisphere are further separated into three clusters led by the middle-part tissues in Temporal and Parietal lobes (\emph{R.Temporal}), the back-part tissues in Occipital lobe (\emph{R.Occipital}), and the front-part tissues in Frontal and Parietal lobes (\emph{R.Supra}). This clustering result is reasonable since the left and right hemispheres often play different roles in human brains.

\begin{figure*}[htb]
    \centering
    \includegraphics[width = .7\textwidth]{brain_node_cluster.pdf}
    \caption{Illustration of brain node clustering results for HCP data with (a) top and (b) side views. }
    \label{fig:cluster_brain}
\end{figure*}

\begin{figure*}[htb]
    \centering
    \includegraphics[width = 1\textwidth]{est_S_anno.pdf}
    \caption{Mode 3 slices of estimated core tensor $\hat \tS$. (a) Average estimated slice weighted by the group size; (b)-(d) Group-specified enrichment, i.e., the difference between each slice of $\hat \tS$ and the averaged slice. }
    \label{fig:ests}
\end{figure*}

\begin{figure*}[htb]
    \centering
    \includegraphics[width = 1\textwidth]{brain_connection.pdf}
    \caption{Observed brain connections in the population and each group of individuals. (a) Average brain network; (b)-(d) Group-specified brain network enrichments in Groups 1-3. Red edges represent the positive enrichment and blue edges represent the negative enrichment.}
    \label{fig:brain_conn}
\end{figure*}

Figure~\ref{fig:ests} illustrates the estimated core tensor $\hat \tS$ with estimated clustering, and Figure~\ref{fig:brain_conn} visualizes the average brain connections and the connection enrichment in contrast to average networks in each group. In general, we find that the inner-hemisphere connection has stronger connection compared to inter-hemisphere connections (Figure~\ref{fig:ests}a). Also, the back and front parts (\emph{R.Occipital}, \emph{R.Supra}) are shown to have more interactions with temporal tissues than inner-cluster connections. In addition, the group 1 with 54\% females shows an enrichment on the inter-hemisphere connections (Figure~\ref{fig:ests}b), while group 4 with only 36\% females exhibits a reduction (Figure~\ref{fig:ests}d). This result agrees with previous findings in \cite{hu2021generalized}. The enrichment on the back-front connection is also recognized in group~3 (Figure~\ref{fig:ests}c). The interpretive patterns in our results demonstrate the usefulness of our clustering methods in the human brain connectome data application.

\subsection{Peru Legislation data analysis}

We also apply our method to the legislation networks in the Congress of the Republic of Peru \citep{lee2017time}. Because of the frequent political power shifts in the Peruvian Congress during 2006-2011, we choose to focus on the data for the first half of 2006-2007 year. The dataset records the co-sponsorship of 116 legislators from top 5 parties and 802 bill proposals. We reconstruct legislation network as an order-3 binary tensor $\tY \in \{0,1\}^{116 \times 116 \times 116}$, where $\tY_{ijk} = 1$ if the legislators $(i,j,k)$ have sponsored the same bill, and $\tY_{ijk} = 0$ otherwise. The true party affiliations of legislators are provided and serve as the ground truth. We apply various higher-order clustering methods to $\tY$ with $r = 5$. Table~\ref{tab:peru} shows that our \textbf{\small dTBM} achieves the best performance compared to others. The second best method is the two-stage algorithm \textbf{\small HLloyd}, followed by the spectral methods \textbf{\small SCORE} and \textbf{\small HOSVD+}. This result is consistent with our simulations under strong signal and moderate degree heterogeneity. The comparison suggests that our method \textbf{\small dTBM} is more appealing in real-world applications.

\begin{table}[ht]
\renewcommand{\arraystretch}{1.3}
    \centering
    \begin{tabular}{c |c  c cc c} 
    \hline 
        Method & \textbf{\small dTBM} 
        &\textbf{\small HOSVD}
        &\textbf{\small HOSVD+} & \textbf{\small HLloyd} &  \textbf{\small SCORE}\\
         CER & \textbf{0.116}
         &  0.22 
         &0.213 & 0.149 &0.199\\
         \hline
    \end{tabular}
    \caption{Clustering errors (measured by CER) for various methods in the analysis of Peru Legislation dataset.}
    \label{tab:peru}
\end{table}

\section{Proof Sketches}\label{sec:mainproof}

In this section, we provide the proof sketches for the main Theorem~\ref{thm:stats} (Impossibility), Theorem~\ref{thm:comp} (Impossibility), and Theorems~\ref{thm:initial}-\ref{thm:refinement}. Detail proofs and extra theoretical results are provided in Appendix II.

\subsection{Proof sketch of Theorem~\ref{thm:stats} (Impossibility) and Theorem~\ref{thm:comp} (Impossibility)} \label{subsec:impossible}

The proofs of impossibility in Theorems~\ref{thm:stats} and \ref{thm:comp} share the same proof idea with \cite[Theorems 6 and 7]{han2020exact} and \cite[Theorem 2]{gao2018community}. In both proofs of statistical and computational impossibilities, the key idea is to construct a particular set of parameters to lower bound the minimax rate. Specifically, for statistical impossibility in Theorem~\ref{thm:stats}, we construct a particular $(z^*_{\rm stats}, \mtheta^*_{\rm stats})  \in \tP_{z, \mtheta} $ such that for all $\tS^* \in \tP_{\tS}(\gamma)$
\begin{align}
    & \inf_{\hat z_{\rm stats}} \sup_{(z, \mtheta) \in \tP_{z, \mtheta}} \bbE[p \ell(\hat z_{\rm stat}, z)] \\
     & \ \geq \inf_{\hat z_{\rm stats}} \bbE[p \ell(\hat z_{\rm stat}, z_{\rm stats}^*)| (z^*_{\rm stats}, \tS^*,\mtheta^*_{\rm stats})] \geq 1;\label{eq:stats_impo_sketch}  
\end{align}
for computational impossibility in Theorem~\ref{thm:comp}, we construct a particular $(z^*_{\rm comp}, \tS^*_{\rm comp}, \mtheta^*_{\rm comp}) \in \tP(\gamma)$ such that
\begin{align}
    &\inf_{\hat z_{\rm comp}} \sup_{(z, \tS, \mtheta) \in \tP(\gamma)} \bbE[p \ell(\hat z_{\rm comp}, z)] \\
    &\ \geq \inf_{\hat z_{\rm comp}} \bbE[p \ell(\hat z_{\rm comp}, z_{\rm comp}^*)| (z^*_{\rm comp}, \tS^*_{\rm comp},\mtheta^*_{\rm comp})] \geq 1.
\end{align}

The constructions of $(z^*_{\rm stats}, \mtheta^*_{\rm stats})$ and $(z^*_{\rm comp}, \tS^*_{\rm comp}, \mtheta^*_{\rm comp})$ are the most critical steps. With good constructions, the lower bound ``$\geq 1$" can be verified by classical statistical conclusions (e.g.\ Neyman-Pearson Lemma) or prior work (e.g.\ HPC Conjecture).

A notable detail in the proof of statistical impossibility is the arbitrariness of $\tS^*$. The first infimum over $ \tP_{\tS}(\gamma)$ in the minimax rate~\eqref{eq:minminmax} requires that the lower bound~\eqref{eq:stats_impo_sketch} holds for any $\tS^* \in \tP_{\tS}(\gamma)$. The arbitrary choice of $\tS^*$ brings extra difficulties in the parameter construction, and consequently a non-trivial $\mtheta^*_{\rm stats} \neq \mathbf{1}$ is chosen to address the arbitrariness. Previous TBM construction in the proof of \cite[Theorem 6]{han2020exact} with $\mtheta^*_{\rm stats} = \mathbf{1}$ is no longer applicable in our case. Meanwhile, our construction $(z^*_{\rm comp}, \tS^*_{\rm comp}, \mtheta^*_{\rm comp})$ leads to a rank-2 mean tensor to relate the HPC Conjecture while TBM \cite[Theorem 7]{han2020exact} constructs a rank-1 mean tensor. Hence, we emphasize that dTBM-specific techniques are required to obtain our impossibility results, though the proof idea is common for minimax lower bound analysis.

\subsection{Proof sketch of Theorem~\ref{thm:initial}}
The proof of Theorem~\ref{thm:initial} is inspired by the proof idea of \cite[Lemma 1]{gao2018community}. The extra difficulties are the angle gap characterization and multilinear algebra property in tensors; we address both challenges in our proof. Specifically, we control the misclustering error by the estimation error of $\hat \tX$ calculated in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}.  We prove the following inequality
\begin{align}
    \ell(z^{(0)},z) &\lesssim \frac{1}{p}\min_{\pi \in \Pi} \sum_{i: z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2 \\
    &\lesssim \frac{\sigma^2 r^{K-1}}{ \Delta_{\min}^2 p^K} \onormSize{}{\hat \tX - \tX}_F^2 \\
    &\lesssim \frac{r^K p^{-K/2}}{\text{SNR}}, \label{eq:proof_4}
\end{align}
where $\tX = \bbE \tY$ is the true mean. The first inequality in~\eqref{eq:proof_4} holds with the assumption $\min_{i \in [p]} \theta(i) \geq c>0$ in Theorem~\ref{thm:initial}. The second inequality relies on an important conclusion that the angle gap of mean tensor $\tX$ is lower bounded by that of core tensor $\tS$, i.e., the minimal angle gap $\Delta_{\min}$ defined in Assumption~\ref{assmp:min_gap}. Let $\ma^{s} \coloneqq \ma/\onorm{\ma}$ denote the normalized vector with the convention that $\ma^s = 0$ if $\ma = 0$. We want to show that
\begin{equation}\label{eq:proof_4_gap}
\min_{z(i) \neq z(j)} \onormSize{}{[\mX_{i:}]^s - [\mX_{j:}]^s} \gtrsim \Delta_{\min},
\end{equation}
where $\mX = \mat(\tX)$. The most challenging part in the proof of Theorem~\ref{thm:initial} lies in the derivation of inequality~\eqref{eq:proof_4_gap}, in which the proof of \cite{gao2018community} is no longer applicable due to different angle gap assumption in our dTBM. We develop the extra padding technique in Lemma~\ref{lem:pad} and balance assumption~\eqref{eq:degree} to derive~\eqref{eq:proof_4_gap}. Last, we finish the proof of Theorem~\ref{thm:initial} by showing the third inequality of~\eqref{eq:proof_4} using \cite[Proposition 1]{han2020exact}.

\subsection{Proof sketch of Theorem~\ref{thm:refinement}}\label{sec:thm5}
The proof of Theorem~\ref{thm:refinement} is inspired by the proof idea of \cite[Theorem 2]{han2020exact}. We develop extra polar-coordinate based techniques with angle gap characterization to address the nuisance degree heterogeneity. Recall the intermediate quantity, misclustering loss, defined in~\eqref{eq:defnofL}
\small
\begin{align}
    L^{(t)} : &=L(z,z^{(t)})\\
    &= \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
\end{align}
\normalsize
where the superscript $\cdot^{s}$ denotes the normalized vector; i.e., $\ma^s:=\ma/\onorm{\ma}$ if $\ma\neq 0$ and $\ma^{s}=0$ if $\ma=0$ for any vector $\ma$. We show that $L^{(t)}$ provides an upper bound for the misclassification error of interest via the inequality $\ell^{(t)}\leq {L^{(t)}\over \Delta^2_{\min}}$. Therefore, it suffices to control $L^{(t)}$. Further, we introduce the oracle estimators for core tensor under the true cluster assignment via 
\begin{equation}
    \tilde \tS = \tY \times_1 \mW^T \times_2 \cdots \times_K \mW^T, 
\end{equation}
where $\mW = \mM \of{ \text{diag}(\mone_{p}^T \mM) }^{-1}$ is the weighted true membership matrix. Let $ \mV = \mW^{\otimes (K-1)}$ denote the Kronecker product of $(K-1)$ copies of $\mW$ matrices, and we define the $t$-th iteration quantities $\mW^{(t)}, \mV^{(t)}$ corresponding to $\mM^{(t)}$ (or equivalently $z^{(t)}$). To evaluate $L^{(t+1)}$, we prove the bound
\begin{align}
    &\ind \offf{ z^{(t+1)}(i) = b } \\
    &\ = \ind \offf{       \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{b:}^{(t)}]^s }^2 \leq \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{z(i):}^{(t)}]^s }^2}\\
    &\ \leq A_{ib} + B_{ib}, \label{eq:proof_5_event}
\end{align}
where $\mY = \mat(\tY)$, $ \mS = \mat(\tS)$, $\mS^{(t)} = \mat(\tS^{(t)})$ and
\small
\begin{align}
        A_{ib} &= \ind \offf{\ang{ \mE_{i:} \mV, \off{  \tilde \mS_{z(i):} }^s - \off{  \tilde \mS_{b:} }^s } \lesssim -  \onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 },\\
        B_{ib} &= \ind \offf{\onorm{ \off{ \mS_{z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2 \lesssim F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }.
\end{align}
\normalsize
The terms $F_{ib}^{(t)}, G_{ib}^{(t)}, H_{ib}^{(t)}$ are controlled by $z^{(t)}, \tS^{(t)}$; see the detailed definitions in \eqref{eq:f}, \eqref{eq:g}, \eqref{eq:h}. Note that the event $A_{ib}$ only involves the oracle estimator independent of $t$, while all the terms related to the $t$-th iteration are in $B_{ib}$. Thus, the inequality~\eqref{eq:proof_5_event} decomposes the misclustering loss in the $(t+1)$-th iteration into the oracle loss and the loss in $t$-th iteration. This decomposition leads to the separation of statistical error and computational error in the final upper bound of Theorem~\ref{thm:refinement}.

Specifically, we prove the contraction inequality
\begin{align}
     &L^{(t+1)} \leq M \xi + \rho L^{(t)}, \\
     \text{with }&\xi = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  A_{ib} \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2, \label{eq:proof_5_ineq}
\end{align}
where $M$ is a positive constant, $\rho \in (0,1)$ is the contraction parameter, and we call $\xi$ the oracle loss. Controlling the probability of event $B_{ib}$ and obtaining the $\rho L^{(t)}$ term in the right hand side of~\eqref{eq:proof_5_ineq} are the most challenging parts in the proof of Theorem~\ref{thm:refinement}. Note that the true and estimated core tensors are involved via their normalized rows such as $\mS_{a:}^s, \tilde \mS_{a:}^s, [\mS^{(t)}_{a:}]^s$. The Cartesian coordinate based analysis in \cite{han2020exact} is no longer applicable in our case. Instead, we use the polar-coordinate based analysis and the geometry property of trigonometric functions to derive the high probability upper bounds for $F_{ib}^{(t)}, G_{ib}^{(t)}, H_{ib}^{(t)}$.

Further, by sub-Gaussian concentration, we prove the high probability upper bound for oracle loss
\begin{equation}\label{eq:proof_5_xi}
    \xi  \lesssim {\text{SNR}^{-1}}\exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}.
\end{equation}
Combining the decomposition~\eqref{eq:proof_5_ineq} and the oracle bound~\eqref{eq:proof_5_xi}, we finish the proof of Theorem~\ref{thm:refinement}.

The proof of MLE error shares the similar idea as Theorems~\ref{thm:initial}-\ref{thm:refinement}. we first show a weaker polynomial rate for MLE, and then improve the rate from polynomial to exponential throughout the iterations. The only difference is that the MLE remains the same over iterations due to its global optimality. See Appendix II, Section~\ref{sec:statprove2} for the detailed proof.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
\section{Additional numerical experiments}
\textbf{Bernoulli phase transition.} The first additional experiment verifies the statistical-computational gap in Section~\ref{sec:limits} under the Bernoulli model. Consider the Bernoulli model with $p = \{80, 100\}$, $r = 5$. We vary $\gamma $ in $ [-1.2, -0.4]$ and $[-2.1, -1.4]$ for matrix ($K=2$) and tensor $(K = 3)$ clustering, respectively. We  approximate MLE using an oracle estimator, i.e., the output of Sub-algorithm~\hyperref[alg:main]{2} initialized from the true assignment. Figure~\ref{fig:phase_binary} shows a similar pattern as Figure~\ref{fig:phase}. The algorithm and oracle estimators have no gap in the matrix case, while an error gap emerges between the critical values $\gamma_{\text{stat}} = -2$ and $\gamma_{\text{comp}} = -1.5$ in the tensor case. Figure~\ref{fig:phase} suggests the statistical-computational gap in Bernoulli models.

\begin{figure}[htb]
    \centering
    \includegraphics[width = \columnwidth]{binary_phase.pdf}
    \caption{SNR phase transitions for Bernoulli dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.}
    \label{fig:phase_binary}
\end{figure}

{\bf Sparsity.} The second additional experiment evaluates the algorithm performances under the sparse binary dTBM~\eqref{eq:sparse_dtbm}. We fix the signal exponent $\gamma = -1.2$ and vary the sparsity parameter $\alpha_p \in [0.05, 0.9]$. A smaller $\alpha_p$ leads to a higher probability of zero entries in the observation. In addition to the three algorithms mentioned in Section~\ref{subsec:comp} (denoted {\bf \small Initialization}, {\bf \small dTBM}, and {\bf \small SCORE}), we consider other three algorithms based on the discussion in Section~\ref{subsec:ber}: 
  \begin{itemize} 
  \item \textbf{\small D-HOSVD}, the diagonal-deleted HOSVD in \cite{ke2019community}; 
  \item \textbf{\small D-HOSVD + Angle}, the combined algorithm of our angle-based iteration with initialization from \textbf{\small D-HOSVD};
  \item \textbf{\small SCORE + Angle}, the combined algorithms of our angle-based iteration with initialization from \textbf{\small SCORE}.
  \end{itemize}

\begin{figure}[htp!]
    \centering
    \includegraphics[width=1\columnwidth]{sparsity.pdf}
    \caption{CER comparison versus sparsity parameter $\alpha_p$ in $[0.05, 0.9]$. We set $p = 100, r = 5$ and $\gamma = -1.2$ under sparse binary dTBM.}
    \label{fig:sparse}
\end{figure}

Figure~\ref{fig:sparse} shows a slightly larger error in \textbf{\small dTBM} than that in \textbf{\small SCORE}, \textbf{\small D-HOSVD + Angle}, and \textbf{\small SCORE + Angle} under the sparse setting with $\alpha_p < 0.3$. 
The small gap between  \textbf{\small dTBM} and other sparse-specific methods implies the robustness of our algorithm. In addition, comparing \textbf{\small SCORE} versus \textbf{\small SCORE + Angle} (or \textbf{\small D-HOSVD} versus \textbf{\small D-HOSVD + Angle}) indicates the benefit of our angle iterations under the sparse dTBM. In the intermediate and dense cases with $\alpha_p \geq 0.3$, our proposed \textbf{\small dTBM} has a clear improvement over others, which again verifies the success of our algorithm in dense settings.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{Proofs}

We provide the proofs for all the theorems in our main paper. In each sub-section, we first show the proof of main theorem and then collect the useful lemmas in the end.   {We combine the proofs of MLE achievement in Theorem~\ref{thm:stats} and polynomial-time achievement in Theorem~\ref{thm:refinement} in the last section due to the similar idea. }

\subsection{Notation}
Before the proofs, we first introduce the notation used throughout the appendix and the general dTBM without symmetric assumptions. The parameter space and minimal gap assumption are also extended for the general asymetric dTBM.

{\bf Preliminaries.}
\begin{enumerate}[wide]
    \item For mode $ k \in [K]$, denote mode-$k$ tensor matricizations by
    \begin{align}
        &\mY_k = \mat_k \of{ \tY }, \quad \mS_k = \mat_k \of{\tS}, \\ &\mE_k = \mat_k \of{ \tE}, \quad \mX_k = \mat_k \of{\tX}.
    \end{align}
    \item For a vector $\ma$, let $\ma^{s} \coloneqq \ma/\onorm{\ma}$ denote the normalized vector. We make the convention that $\ma^s = {\bf 0}$ if $\ma = {\bf 0}$. 
    \item For a matrix $\mA \in \bbR^{n \times m} $, let $\mA^{\otimes K}:=\mA\otimes \cdots \otimes \mA\in \bbR^{n^K \times m^K}$ denote the Kronecker product of $K$ copies of matrices $\mA $.
    \item For a matrix $\mA$, let $\onormSize{}{\mA}_\sigma$ denote the spectral norm of matrix $\mA$, which is equal to the maximal singular value of $\mA$; let $\lambda_k(\mA)$ denote the $k$-th largest singular value of $\mA$; let $\onormSize{}{\mA}_F$ denote the Frobenius norm of matrix $\mA$.
    \item For two sequence $a$ and $b$, let $a \asymp b$ if there exist two positive constants $c, C$ such that $cb \leq a\leq Cb$. 
\end{enumerate}

{\bf Extension to general asymetric dTBM.}

 The general order-$K$ $(p_1, \ldots, p_K)$-dimensional dTBM with $r_k$ communities and degree heterogeneity $\mtheta_k = \entry{\theta_k(i)} \in \bbR_+^{p_k}$ is represented by
\begin{equation}\label{eq:general_dtbm}
    \tY = \tX+ \tE,\quad \text{where}\quad \tX=\tS \times_1 \mTheta_1 \mM_1 \times_2 \cdots \times_K \mTheta_K \mM_K,
\end{equation}
where $\tY \in \bbR^{p_1 \times \cdots \times p_K}$ is the data tensor, $\tX\in \mathbb{R}^{p_1\times \cdots \times p_K}$ is the mean tensor, $\tS \in \bbR^{r_1 \times \cdots \times r_K}$ is the core tensor, $\tE \in \bbR^{p_1 \times \cdots \times p_K}$ is the noise tensor consisting of independent zero-mean sub-Gaussian entries with variance bounded by $\sigma^2$, $\mTheta_k = \text{diag}(\mtheta_k)$, and $\mM_k\in \{0,1\}^{p_k \times r_k}$ is the membership matrix corresponding to the assignment $z_k: [p_k] \mapsto [r_k]$, for all $k \in [K]$.

For ease of notation, we use $\{z_k\}$ to denote the collection $\{z_k\}_{k=1}^K$, and $\{\mtheta_k\}$ to denote the collection $\{\mtheta_k\}_{k=1}^K$. Correspondingly, we consider the parameter space for the triplet $\of{\{z_k\}, \tS, \{\mtheta_k\}}$,
\small
\begin{align}
 & \tP(\{r_k\}) = \Big\{ \of{\{z_k\}, \tS, \{\mtheta_k\}}: \mtheta_k \in\mathbb{R}^p_{+}, {c_1 p_k\over r_k} |z_k^{-1}(a)| \leq {c_2 p_k\over r_k}, \\ & \ c_3 \leq  \onorm{\mS_{k,a:}} \leq c_4 ,\onormSize{}{\mtheta_{k,z_k^{-1}(a)}}_1=|z_k^{-1}(a)|, a \in [r_k], k\in[K]\Big\}. \label{eq:general_family}
\end{align}
\normalsize

We call the degree heterogeneity $\{\mtheta_k\}$ is balanced if for all $k \in [K]$,
\begin{equation}\label{eq:general_balanced}
    {\min_{a\in[r]} \onormSize{}{\mtheta_{k, z_k^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{k,z_k^{-1}(a)}}}.
\end{equation}

We also consider the generalized Assumption~\ref{assmp:min_gap} on angle gap.
\begin{assumption}[Generalized angle gap]\label{assmp:general_minimal_gap} Recall $\mS_k=\mat_k(\tS)$. We assume the minimal gap between normalized rows of $\mS_k$ is bounded away from zero for all $k\in[K]$; i.e.,
\begin{equation}
     \Delta_{\min} \coloneqq \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onorm{ \mS_{k,a:}^s - \mS_{k, b:}^s } > 0.
\end{equation}
\end{assumption}
Similarly, let $\text{SNR} = \Delta_{\min}^2/\sigma^2$ with the generalized minimal gap $\Delta_{\min}^2$ defined in Assumption~\ref{assmp:general_minimal_gap}. We define the regime
\small
\begin{align}
    \tP(\gamma) = \tP (\{r_k\}) \cap\{\tS \text{ satisfies $\text{SNR} = p^{\gamma}$ and $p_k \asymp p, k \in [K]$} \}.
\end{align}
\normalsize

\subsection{Proof of Theorem~\ref{thm:unique}}

\begin{proof}[Proof of Theorem~\ref{thm:unique}]

To study the identifiability, we consider the noiseless model with $\tE = 0$. Assume there exist two parameterizations satisfying
\begin{align}
    \tX&=\tS\times_1\Theta_1 \mM_1 \times_2 \cdots \times_K \Theta_K \mM'_K\\
    &=\tS'\times_1\Theta'_1 \mM'_1 \times_2 \cdots \times_K \Theta'_K \mM'_K,\label{eq:another}
\end{align}
where $\of{ \{z_k\}, \tS, \{ \mtheta_k \} } \in \tP(\{r_k\})$ and $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} } \in \tP(\{r_k'\})$ are two sets of parameters. We prove the sufficient and necessary conditions separately.

\begin{enumerate}[wide]
    \item[$(\Leftarrow)$] For the necessity, it suffices to construct two distinct parameters up to cluster label permutation, if the model~\eqref{eq:general_dtbm} violates Assumption~\ref{assmp:general_minimal_gap}. {Note that $\Delta_{\min}^2 = 1$ when there exists a $k \in [K]$ such that $r_k = 1$. Hence, we consider the case that  $r_k \geq 2$ for all $k \in [K]$.}  Without loss of generality, we assume $\onorm{ \mS_{1,1:}^s - \mS_{1,2:}^s } = 0$.

% If $\mS_{1,1:}$ is a zero vector, construct $\mtheta_1'$ such that $\mtheta'_{1,z_1^{-1}(1)} \neq \mtheta_{1,z_1^{-1}(1)}$. Let $\{z'_k\} = \{z_k\}$, $\tS' = \tS$, and $\mtheta'_k = \mtheta_k$ for all $k = 2, \ldots, K$. Then the triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ is distinct from $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation. Similar conclusion holds when $\mS_{1,2:}$ is a zero vector. 
{

By constraints in parameter space~\eqref{eq:general_family}, neither $\mS_{1,1:}$ nor $\mS_{1,2:}$ is a zero vector.
}
There exists a positive constant $c$ such that $\mS_{1,1:} = c \mS_{1,2:}$. Thus, there exists a core tensor $\tS_0 \in \bbR^{r_1 -1 \times \cdots \times r_K}$ such that 
\begin{equation}
    \tS = \tS_0 \times_1 \mC \mR,
\end{equation}
where $\mC = \text{diag}(1, c, 1,...,1) \in \bbR^{r_1 \times r_1}$ and 
\begin{equation}
    \mR = \begin{pmatrix}
    1& 0\\
    1&0 \\
    0 & \mone_{r_1-2}
    \end{pmatrix} \in \bbR^{r_1 \times (r_1 -1)}.
\end{equation}
Let $\mD = \text{diag}(1+c, 1,...,1) \in \bbR^{r_1 -1 \times r_1 -1}$. Consider the parameterization $\mM'_1 = \mM_1 \mR,\tS' = \tS_0 \times_1 \mD,$ and 
\begin{equation}
     \theta'_{1}(i) = \begin{cases}
    \frac{1}{1+c} \theta_{1}(i) & i \in z_{1}^{-1}(1),\\
     \frac{c}{1+c} \theta_{1}(i) & i \in z_{1}^{-1}(2),\\
     \theta_{1}(i) & \text{ otherwise},
    \end{cases}
\end{equation}
and $\mM'_k = \mM_k, \mtheta'_k = \mtheta_k$ for all $k = 2, \ldots, K$. Then we have constructed a
triplet $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ that is distinct from $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation.

\item[$(\Rightarrow)$] For the sufficiency, it suffices to show that all possible triplets $\of{ \{z'_k\}, \tS', \{ \mtheta'_k \} }$ are identical to $\of{ \{z_k\}, \tS, \{ \mtheta_k \} }$ up to label permutation if the model~\eqref{eq:general_dtbm} satisfies Assumption~\eqref{assmp:general_minimal_gap}. We show the uniqueness of the three parameters, $\{\mM_k\}, \{\tS\}, \{\mtheta_k\}$ separately.

First, we show the uniqueness of $\mM_k$ for all $k \in [K]$. 
{When $r_k = 1$, all possible $\mM_k$ is equal to the vector ${\bf 1}_{p_k}$, and the uniqueness holds trivially. Hence, we consider the case that $r_k \geq 2$. Without loss of generality, we consider $k=1$ with $r_1 \geq 2$ and show the uniqueness of the first mode membership matrix;}  
%Without loss of generality, we consider $k=1$ and show the first mode membership matrix; 
i.e., $\mM'_1 = \mM_1\mP_1$ where $\mP_1$ is a permutation matrix. The conclusion for $k\geq 2$ can be showed similarly and thus omitted.

Consider an arbitrary node pair $(i,j)$. If $z_1(i) = z_1(j)$, then we have $\onormSize{}{\mX_{1, z_1(i):}^s - \mX_{1, z_1(j):}^s } = 0$ and thus $\onormSize{}{ (\mS')_{1, z'_1(i):}^{s} - (\mS')_{1, z'_1(j):}^{s} } = 0$ by Lemma~\ref{lem:angle}. Then, by Assumption~\eqref{assmp:general_minimal_gap}, we have $z'_1(i) = z'_1(j)$. Conversely, if $z_1(i) \neq z_1(j)$, then we have $ \onorm{\mX_{1,i:}^s - \mX_{1,j:}^s} \neq 0$ and thus $\onorm{ (\mS')_{1, z'_1(i):}^{s} - (\mS')_{1, z'_1(j):}^{s} } \neq 0$ by Lemma~\ref{lem:angle}. Hence, we have $z'_1(i) \neq z'_1(j)$. Therefore, we have proven that $z'_1$ is identical $z_i$ up to label permutation.

Next, we show the uniqueness of $\mtheta_k$ for all $k \in [K]$ provided that $z_k = z_k'$. Similarly, consider $k=1$ only, and omit the procedure for $k\geq 2$.

Consider an arbitrary $j \in [p_1]$ such that $z_1(j) = a$. Then for all the nodes $i \in  z_1^{-1}(a)$ in the same cluster of $j$, we have 
\begin{equation}
    \frac{\mX_{1,z_1(i):}}{\mX_{1,z_1(j):}} = \frac{\mX'_{1,z_1(i):}}{\mX'_{1,z_1(j):}}, \text{ which implies }  \frac{\theta_1(j)}{\theta_1(i)} = \frac{\theta'_1(j)}{\theta'_1(i)}.\label{eq:theta_uniq}
\end{equation}
Let $\theta'_1(j) = c\theta_1(j)$ for some positive constant $c$. By equation~\eqref{eq:theta_uniq}, we have $\theta'_1(i) = c \theta_1(i)$ for all $ i \in  z_1^{-1}(a)$. By the constraint $(\{z_k\}, \tS', \{\mtheta'_k\}) \in \tP(\{r_k\})$, we have 
\begin{equation}
    \sum_{j \in z_1^{-1}(a)} \theta'_1(j) = c \sum_{j \in z_1^{-1}(a)} \theta_1(j) = 1,
\end{equation}
which implies $c = 1$. Hence, we have proven $\mtheta_1 = \mtheta'_1$ provided that $z_1 = z'_1$.

Last, we show the uniqueness of $\tS$; i.e., $\tS'=\tS\times_1 \mP^{-1}_1\times_2\cdots \times_K \mP^{-1}_K$, where $\mP_k$'s are permutation matrices for all $k\in[K]$.  Provided $z'_k = z_k, \mtheta'_k = \mtheta_k$, we have $\mM'_k = \mM_k \mP_k$ and $\mTheta'_k = \mTheta_k$ for all $k \in [K]$.

Let $\mD_k = \off{ (\mTheta'_k \mM'_k)^T (\mTheta'_k \mM'_k) }^{-1} (\mTheta'_k \mM'_k)^T, k \in [K]$. By the parameterization~\eqref{eq:another}, we have 
\begin{align}
    \tS' &= \tX \times_1 \mD_1 \times_2 \cdots \times_K \mD_K \\
    &= \tS \times_1 \mD_1 \mTheta_1 \mM_1 \times_1 \cdots \times_K \mD_K \mTheta_K \mM_K \\
    &= \tS \times_1 \mP^{-1}_1 \times_2 \cdots \times_K \mP^{-1}_K.
\end{align}

\end{enumerate}

Therefore, we finish the proof of Theorem~\ref{thm:unique}.
\end{proof}

{\bf Useful Lemma for the Proof of Theorem~\ref{thm:unique}}

\begin{lem}[Motivation of angle-based clustering]\label{lem:angle} Consider the signal tensor $\tX$ in the general asymmetric dTBM~\eqref{eq:general_dtbm} with $(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\{r_k\})$ and $r_k \geq 2$. Then, for any $k \in [K]$ and index pair $(i,j)\in[p_k]^2$, we have 
\begin{align}
     & \onorm{ \mS_{k,z_k(i):}^s -  \mS_{k,z_k(j):}^s } = 0 \quad  \text{if and only if} \\
    & \quad \quad \quad \quad \quad  \onorm{  \mX_{k, z_k(i):}^s -  \mX_{k,z_k(j):}^s } = 0.
\end{align}
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:angle}] Without loss of generality, we prove $k = 1$ only and drop the subscript $k$ in $\mX_k, \mS_k$ for notational convenience. 
By tensor matricization, we have
\begin{equation}
    \mX_{j:} = \theta_1(j) \mS_{z_1(j):} \off{\mTheta_2 \mM_2 \otimes \cdots \otimes \mTheta_K \mM_K}^T.
\end{equation}     
Let $\tilde \mM = \mTheta_2 \mM_2 \otimes \cdots \otimes \mTheta_K \mM_K$. Notice that for two vectors $\ma, \mb$ and two positive constants $c_1, c_2 >0$, we have
\begin{equation}
\onorm{\ma^s - \mb^s} = \onorm{(c_1 \ma)^s - (c_2\mb)^s}.
\end{equation}
Thus it suffices to show the following statement holds for any index pair $(i,j)\in[p_1]^2$,
\begin{align}
    &\onorm{ \mS_{z_1(i):}^s - \mS_{z_1(j):}^s} = 0 \quad \text{if and only if} \\
    & \quad \quad \quad \quad \quad \onorm{ \off{\mS_{z_1(i):} \tilde \mM^T }^s - \off{\mS_{z_1(j):}\tilde \mM^T}^s} = 0.
\end{align}
\begin{enumerate}[wide]
    \item[$(\Leftarrow)$] Suppose $\onorm{ \off{\mS_{z_1(i):} \tilde \mM^T }^s - \off{\mS_{z_1(j):}\tilde \mM^T}^s} = 0$. There exists a positive constant $c$ such that $\mS_{z_1(i):} \tilde \mM^T= c \mS_{z_1(j):} \tilde \mM^T$. Note that
\begin{equation}
    \mS_{z_1(i):} = \mS_{z_1(i):} \tilde \mM^T \off{ \tilde \mM \of{ \tilde \mM^T  \tilde \mM}^{-1}},
\end{equation}
where $ \tilde \mM^T  \tilde \mM$ is an invertiable diagonal matrix with positive diagonal elements. Thus, we have $ \mS_{z_1(i):} = c  \mS_{z_1(j):}$, which implies $ \onorm{  \mS_{z_1(i):}^s -  \mS_{z_1(j):}^s } = 0 $.

\item[$(\Rightarrow)$] Suppose $ \onorm{ \mS_{z_1(i):}^s - \mS_{z_1(j):}^s } = 0 $. There exists a positive constant $c$ such that $\mS_{z_1(i):} = c \mS_{z_1(j):}$, and thus $\mS_{z_1(i):} \tilde \mM^T = c \mS_{z_1(j):} \tilde \mM^T$, which implies $\onorm{\left[\mS_{z_1(i):} \tilde \mM^T\right]^s- \left[\mS_{z_1(j):} \tilde \mM^T\right]^s}=0$.
\end{enumerate}
Therefore, we finish the proof of Lemma~\ref{lem:angle}.
\end{proof}

\subsection{Proof of Lemma~\ref{lem:angle_gap_x} and Lemma~\ref{lem:mis}}

\begin{proof}[Proof of Lemma~\ref{lem:angle_gap_x}]

Note that the vector $\mS_{z(i):}$ can be folded to a tensor $\tS' =\entry{\tS'_{a_2,\ldots,a_K}} \in \bbR^{r^{K-1}}$; i.e., $\text{vec}(\tS') = \mS_{z(i):}$. Define weight vectors $\mw_{a_2, \cdots, a_K}$ correspond to the elements in $\tS'_{a_2,\ldots,a_K}$ by
\begin{equation}
    \mw_{a_2 \cdots a_K} = [ \mtheta_{z^{-1} (a_2)}^T \otimes \cdots \otimes \mtheta_{z^{-1} (a_K)}^T] \in \bbR^{|z^{-1} (a_2)| \times \cdots \times |z^{-1} (a_K)|},
\end{equation}
for all $a_k \in [r], k = 2,\ldots, K$, where $\otimes$ denotes the Kronecker product. Therefore, we have  $\mX_{i:} = \theta(i) \pad_{\mw}(\mS_{z(i):})$ where $\mw = \{ \mw_{a_2, \cdots, a_K}\}_{a_k \in [r], k\in [K]/\{1\}}$. Specifically, we have $\onormSize{}{\mw_{a_2, \ldots, a_K}}^2 = \prod_{k = 2}^K \onormSize{}{ \mtheta_{z^{-1}(a_k)}}^2$, and by the balanced assumption~\eqref{eq:degree} we have 
\small
\begin{equation}\label{eq:pad_balance}
    \max_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2 = (1 + o(1))  \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2.
\end{equation}
\normalsize

Consider the inner product of $\mX_{i:}$ and $\mX_{j:}$ for $z(i) \neq z(j)$. By the definition of weighted padding operator~\eqref{eq:paddef} and the balanced assumption~\eqref{eq:pad_balance}, we have 
\begin{align}
    &\ang{\mX_{i:},\mX_{j:}} \\
    &\ = \theta(i) \theta(j) \ang{ \pad_{\mw}(\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):} )}\\
    &\ = \theta(i) \theta(j) \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2 \ang{\mS_{z(i):}, \mS_{z(j):}}(1 + o(1)).
\end{align}
Therefore, when $p$ large enough, the inner product $\ang{\mX_{i:},\mX_{j:}} $ has the same sign as $\ang{\mS_{z(i):}, \mS_{z(j):}}$.

{
 Then, we have 
\begin{align}
    \cos (\mS_{z_1(i):}, \mS_{z_1(j):}) &= \frac{ \ang{  \mS_{z_1(i):}, \mS_{z_1(j):} }}{ \onormSize{}{\mS_{z_1(i):}} \onormSize{}{\mS_{z_1(j):}} } \\
    &= (1+o(1))\frac{ \ang{  \mX_{i:}, \mX_{j:} }}{ \onormSize{}{ \mX_{i:}} \onormSize{}{ \mX_{j:}} }\\
    &= (1 + o(1)) \cos(\mX_{i:}, \mX_{j:}),
\end{align}
where the second inequality follows by the balance assumption on $\mtheta$.
%and the last inequality holds when $p$ is large enough.

Further, notice that $\onormSize{}{\mv_1^s - \mv_2^s}^2 =  2(1 - \cos(\mv_1, \mv_2))$. For all $i,j$ such that $z(i) \neq z(j)$, when $p \rightarrow \infty$, we have
\begin{equation}
    \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \asymp \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
\end{equation}

}

% {
% Notice that $\onormSize{}{\mv_1^s - \mv_2^s}^2 =  2(1 - \cos(\mv_1, \mv_2))$ for two vectors $\mv_1, \mv_2$ of the same dimension. We have 
% \begin{align}
%     \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s}  &= 2 - 2 \frac{ \ang{  \mS_{z_1(i):}, \mS_{z_1(j):} }}{ \onormSize{}{\mS_{z_1(i):}} \onormSize{}{\mS_{z_1(j):}} } \\
%     & = 2 - 2 (1+o(1))\frac{ \ang{  \mX_{i:}, \mX_{j:} }}{ \onormSize{}{ \mX_{i:}} \onormSize{}{ \mX_{j:}} } \\
%     & = (1 + o(1)) \onormSize{}{\mX_{i:}^s - \mX_{j:}^s},
% \end{align}
% where the second inequality follows by the balance assumption on $\mtheta$, and the last inequality holds when $p$ is large enough.

% Hence, we conclude that for all $i,j$ such that $z(i) \neq z(j)$,
% \begin{equation}
%     \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \asymp \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
% \end{equation}
% }

\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:mis}]
By the definition of minimal gap in Assumption~\ref{assmp:min_gap}, we have 
\begin{align}
     L^{(t)} &= \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
     &\geq \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \Delta_{\min}^2 \\
     & \geq c \ell^{(t)} \Delta_{\min}^2,
\end{align}
    where the last inequality follows from the assumption $\min_{i \in [p]} \theta(i) \geq c>0$.
\end{proof}

\subsection{Proof of Theorem~\ref{thm:stats} (Impossibility)}\label{sec:statprove1}

\begin{proof}[Proof of Theorem~\ref{thm:stats} (Impossibility)]Consider the general asymetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$ {with $K\geq 2$, $2 \leq r\lesssim p^{1/3}$ as $p \rightarrow \infty$}. For simplicity, we show the minimax rate for the estimation on the first mode $\hat z_1$; the proof for other modes are essentially the same.

   To prove the minimax rate~\eqref{eq:minminmax}, it suffices to take an arbitrary $\tS^* \in  \tP_{\tS}(\gamma)$ wih $\gamma < -(K-1)$ and construct $(z^*_k, \mtheta^*_k)$ such that 
   \begin{equation}
       \inf_{\hat z_1} \bbE \left[ p\ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right]\geq 1.
   \end{equation}

   We first define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation. Based on \cite[Proof of Theorem 6]{han2020exact}, we consider the restricted family of $\hat z_k$'s for which the following three conditions are satisfied:
   \begin{align}
        &\text{(a)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \quad \text{(b)} \ |T^c_k|\asymp {p\over r}; \\
        &\text{(c)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
   \end{align}
for all $k \in [K]$.
   Now, we consider the construction:
   \begin{enumerate}
       \item[(i)] $\{z_k^*\}$ satisfies properties (a)-(c) with misclassification sets $T_k^c$ for all $k \in [K]$;
       \item [(ii)] $\{\mtheta_k^*\}$ such that $\mtheta_k^*(i) \leq \sigma r^{(K-1)/2} p^{-(K-1)/2}$ for all $i \in T_k^c, k \in [K]$ and $\max_{k \in [K], a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}^2_2 \asymp p/r$.
   \end{enumerate}
   % There exists a $(z^*, \mtheta^*) \in \tP_{z, \mtheta}$ satisfying the property (i) and (ii) when $r < p$; otherwise, when $r = p$, we must have $\mtheta^*(i) = 1$ due to the $\ell_1$ norm constraint and we can not satisfy the property (ii).

   Combining the inequalities (39) and (40) in the proof of Theorem 2 in \cite{gao2018community}, we have 
   \begin{align}
        & \inf_{\hat z_1 } \bbE \left[ \ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right] \geq  \\
        & \  \frac{C}{r^3 |T_1^c|} \sum_{i \in T_1^c} \inf_{\hat z_1  (i)} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, z_k^*, \mtheta_k^*, \tS^*] \\
        & \quad \quad \quad  \quad  \quad \quad +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, z_k^*, \mtheta_k^*, \tS^*] \},\label{eq:inf_lower}
   \end{align}
   where $C$ is some positive constant,  $\hat z_1$ on the left hand side denote the generic assignment functions in $\tP(\gamma)$, and the infimum on the right hand side is taken over the generic assignment function family of $\hat z_1(i)$ for all nodes $i \in T_1^c$. Here, the factor $r^3=r\cdot r^2$ in~\eqref{eq:inf_lower} comes from two sources: $r^2\asymp {r\choose 2}$ comes from the multiple testing burden for all pairwise comparisons among $r$ clusters; and another $r$ comes from the number of elements $|T^c_k|\asymp p/r$ to be clustered.

   Next, we need to find the lower bound of the rightmost side in~\eqref{eq:inf_lower}. 
   %For simplicity, we show the bound for the mode-1 case $k = 1$ only. We drop the subscripts $1$ in $z_1, T_1, \mS_1, \mtheta_1$ and omit the repeated procedures for the cases of $k = 2,\ldots, K$. 
We consider the hypothesis test based on model~\eqref{eq:general_dtbm}. First, we reparameterize the model under the construction (i)-(ii).
\begin{equation}
    \mx_a^* = \off{\mat_1 \of{ \tS^*\times_2 \mTheta_2^* \mM_2^*\times_3\cdots \times_K \mTheta^*_K \mM_K^* }}_{a:},
\end{equation}
for all $a \in [r]$, where $\mx_a^*$'s are centroids in $\bbR^{p^{K-1}}$. Without loss of generality, we consider the lower bound for the summand in~\eqref{eq:inf_lower} for $i=1$. The analysis for other $i\in T^c_1$ are similar. For notational simplicity, we suppress the subscript $i$ and write $\my, \theta^*, z$ in place of $\my_1, \mtheta_1^*(1)$ and $z_1(1)$, respectively. The equivalent vector problem for assessing the summand in~\eqref{eq:inf_lower} is
\begin{equation}\label{eq:z}
\my=\theta^* \mx_{z}^*+\me,
\end{equation}
where $z\in \{1,2\}$ is an unknown parameter, $\theta^* \in \bbR_+$ is the given heterogeneity degree, $\mx_1^*,\mx_2^*\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.  Then, we consider the hypothesis testing under the model~\eqref{eq:z}:
\small
\begin{equation}\label{eq:test}
 H_0: z = 1, \my = \theta^* \mx_1^* + \me \  \leftrightarrow \  H_1: z = 2, \my = \theta^* \mx_2^* + \me,
\end{equation}
\normalsize

%   For an arbitrary $i \in T_1^c$, the right hand side in \eqref{eq:inf_lower} corresponds to the following hypothesis test:
%   \begin{equation}\label{eq:hypo}
%       H_0: \my = \mtheta_1^*(i) \mx_1^* + \me \quad \leftrightarrow \quad H_1: \my = \mtheta_1^*(i) \mx_2^* + \me,
%   \end{equation}
%   where $\my = [\mat_1(\tY)]_{i:}$, $\mx_a^* = [\mat_1( \tS^* \times_2 \mTheta^* \mM_2^* \times_3 \cdots \times_K \mTheta^* \mM_K^*)]_{a:}$, and $\me$ consists of i.i.d.\ $N(0,\sigma^2)$ entries. 
   The hypothesis testing~\eqref{eq:test} is a simple versus simple testing, since the assignment $z$ is the only unknown parameter in the test. 
   By Neyman-Pearson lemma, the likelihood ratio test is optimal with minimal Type I + II error. Under Gaussian model, the likelihood ratio test of \eqref{eq:test} is equivalent to the least square estimator $\hat z_{LS} = \argmin_{a = \{1,2\}} \onormSize{}{\my - \theta^* \mx_a^*}_F^2$.

   Let $\mS = \mat_1(\tS)$. Note that 
   \begin{align}
       &\onormSize{}{ \theta^* \mx_1^*  - \theta^* \mx_2^*}_F\\
       &\leq  \theta^* \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*)  \\
       & \leq \theta^* \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F   \max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}_2^{K-1} \\
       & \leq  \sigma r^{(K-1)/2} p^{-(K-1)/2} 2 c_4 p^{(K-1)/2} r^{-(K-1)/2}\\
       & \leq 2 c_4 \sigma, \label{eq:bound}
   \end{align}
%   \begin{align}
%       \onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F &\leq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*)  \\
%       & \leq \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F   \max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}_2^{K-1} \\
%       & \leq  \sigma r^{(K-1)/2} p^{-(K-1)/2} 2 c_4 p^{(K-1)/2} r^{-(K-1)/2}\\
%       & \leq 2 c_4 \sigma, \label{eq:bound}
%   \end{align}
   where $\lambda_{\max}$ denote the maximal singular value, the second inequality follows from Lemma~\ref{lem:singular_thetam}, and the third inequality follows from property (ii) and the boundedness constraint in $\tP_{\tS}(\gamma)$ such that $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F  \leq \onormSize{}{\mS_{1:}^*}_F + \onormSize{}{\mS_{2:}^*}_F \leq 2c_4$.

   Hence, we have 
   \begin{align}
       & \inf_{\hat z_1(1)} \{ \bbP[\hat z_1(1) = 1| z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^*] \\
       & \quad \quad \quad  \quad \quad  +   \bbP[\hat z_1(1) = 2| z_1^*(1) = 1, z^*_k,  \mtheta_k^*, \tS^*] \} \\
       & \quad = 2 \bbP[ \hat z_{LS} = 1 |  z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^* ] \\
      & \quad  = 2 \bbP[ \onormSize{}{\my - \theta^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \theta^* \mx_2^*}_F^2 | z_1^*(1) = 2, z^*_k , \mtheta_k^*, \tS^*   ] \\
      & \quad = 2 \bbP \off{ 2 \ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \geq  \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F^2 }\\
       & \quad =  2 \bbP[ N(0,1) \geq \theta^* \onormSize{}{\mx_1^* - \mx_2^*}_F /(2\sigma) ] \\
       & \quad \geq 2 \bbP[ N(0,1) \geq c_4 ]   \geq c, \label{eq:inf}
   \end{align}
%   \begin{align}
%       & \inf_{\hat z_1(1)} \{ \bbP[\hat z_1(1) = 1| z_1^*(i) = 2, z^*_k,  \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(1) = 2| z_1^*(i) = 1, z^*_k,  \mtheta_k^*, \tS^*] \} \\
%       & \quad = 2 \bbP[ \hat z_{LS}(i) = 1 |  z_1^*(i) = 2, z^*_k,  \mtheta_k^*, \tS^* ] \\
%       & \quad  = 2 \bbP[ \onormSize{}{\my - \mtheta_1(i)^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \mtheta_1(i)^* \mx_2^*}_F^2 | z_1^*(i) = 2, z^*_k , \mtheta_k^*, \tS^*   ] \\
%       & \quad =  2 \bbP[ N(0,1) \geq \mtheta_1(i)^* \onormSize{}{\mx_1^* - \mx_2^*}_F /\sigma ] \\
%       & \quad \geq 2 \bbP[ N(0,1) \geq 2 c_4 ]   \geq c, \label{eq:inf}
%   \end{align}
where the first equation holds by symmetry, the third equation holds by rearrangement, the fourth equation holds by the fact that $\ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \sim N(0, \sigma \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F)$, and $c$ is some positive constant in the last inequality.
   %for some positive constant $c$, where the calculation for the last equation follows from the calculation in equation (102) in \cite{han2020exact}.

   Plugging the inequality~\eqref{eq:inf} into the inequality~\eqref{eq:inf_lower} for all $i \in T_1^c$, then, we have 
   \begin{equation}
       \liminf_{p \rightarrow \infty}  \inf_{\hat z_1 } \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right]  \geq \liminf_{p \rightarrow \infty} \frac{ C cp }{r^3} \geq C c,
   \end{equation}
   where the last inequality follows by the condition $r = o(p^{1/3})$. By the discrete nature of the misclassification error, we obtain our conclusion
   \begin{equation}
       \liminf_{p \rightarrow \infty} \inf_{\tS^* \in  \tP_{\tS}(\gamma)}  \inf_{\hat z_{\text{stat}} }\sup_{ (z^*, \mtheta^*) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]  \geq 1. 
   \end{equation}
%   for any $r < p$ and $p$,  we have 
%   \begin{equation}
%         \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq \frac{ cp }{r^3} > 0,
%   \end{equation}
%   which indicates 
%   \begin{equation}
%         \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq 1, \quad \text{for every } p
%   \end{equation}
%   by the discrete nature of $p\ell(\hat z, z^*)$.

{

Last, with constructed $z^*_k, \mtheta^*_k$ satisfying properties (i) and (ii) and $\gamma' < -(K-1)$, we construct a core tensor $\tS^*$ such that $\Delta_{\mX^*}^2 \leq p^{-(K-1)}$. Based on the property (ii) and the boundedness constraint of $\tS^*$ in $\tP$, we still have $\onormSize{}{ \theta^* \mx_1^*  - \theta^* \mx_2^*}_F \leq 2c_4 \sigma$. Hence, we obtain the desired result 
\begin{align}
    &\liminf_{p \rightarrow \infty} \inf_{\hat z_1} \sup_{(z, \mtheta, \tS) \in \tP'(\gamma')} \bbE \left[ p\ell(\hat z_1, z_1) \right]  \\
    &\quad \geq   \liminf_{p \rightarrow \infty} \inf_{\hat z_{\rm stat}}  \bbE \left[ p\ell(\hat z_1, z^*_1) |  z^*_k, \mtheta^*_k, \tS^*  \right]  \geq 1.
\end{align}

%for any parameter $(z^*, \mtheta^*, \tS^*) \in \tP$ such that $\Delta_{\mX^*}^2 \lesssim p^{-(K-1)}$, we still have $ \onormSize{}{ \theta^* \mx_1^*  - \theta^* \mx_2^*}_F \leq 2c_4 \sigma$ based on the property (ii) and the boundedness constraint of $\tS^*$ in $\tP$. Hence, the same impossibility conclusion holds with $\Delta_{\mX}^2 \lesssim p^{-(K-1)}$.

}

   \end{proof}

\subsection{Proof of Theorem~\ref{thm:comp} (Impossibility)}\label{sec:compprove1}

\begin{proof}[Proof of Theorem~\ref{thm:comp} (Impossibility)]
The idea of proving computational hardness is to show the computational lower bound for a special class of degree-corrected tensor clustering model with $K\geq 2$ {and $ r \geq 2$}. We construct the following special class of higher-order degree-corrected tensor clustering  model. For a given signal level $\gamma\in\bbR$ and noise variance $\sigma$, define a rank-2 symmetric tensor $\tS\in\bbR^{3\times \cdots\times 3}$ subject to
\begin{equation}\label{eq:S}
\tS=\tS(\gamma) = \begin{bmatrix}
		1\\
		1\\
		1
		\end{bmatrix}^{\otimes K}
+\sigma p^{-\gamma/2} \begin{bmatrix}
		1\\
		-1\\
		0
	\end{bmatrix}^{\otimes K}.
\end{equation}
Then, we consider the signal tensor family
\small
\begin{align}
\tP_{\text{shifted}}(\gamma)=\{\tX\colon  &\tX=\tS\times_1\mM_1\times_2 \cdots \times_K\mM_K,\ \text{where}  \\
& \text{membership matrix $\mM_k\in\{0,1\}^{p\times 3}$ satisfies} \\
 &\text{$|\mM_k(\colon,i)|\asymp p$ for all $i\in[3]$ and $k\in[K]$}\}.\\
\end{align}
\normalsize
We claim that the constructed family satisfies the following two properties:
\begin{enumerate}[wide,label=(\roman*)]
    \item For every $\gamma\in \mathbb{R}$, $\tP_{\text{shifted}}(\gamma)\subset \tP(\gamma)$, where $\tP(\gamma)$ is the degree-corrected cluster tensor family~\eqref{eq:gammafamily}.
    \item For every $\gamma\in \mathbb{R}$, $\{\tX-1\colon \tX\in \tP_{\text{shifted}}(\gamma)\}\subset \tP_{\text{non-degree}}(\gamma)$, where $\tP_{\text{non-degree}}(\gamma)$ denotes the sub-family of rank-one tensor block model constructed in the proof of \cite[Theorem 7]{han2020exact}. 
\end{enumerate}
The verification of the above two properties is provided in the end of this proof.

Now, following the proof of~\cite[Theorem 7]{han2020exact}, when $\gamma<-K/2$, every polynomial-time algorithm estimator $(\hat \mM_k)_{k\in[K]}$ obeys
\begin{align}\label{eq:p}
\liminf_{p\to \infty} \sup_{\tX\in \tP_{\text{non-degree}}(\gamma)}\mathbb{P}(\exists k \in[K],\  \hat \mM_k \neq \mM_k)\geq 1/2,
\end{align}
under the HPC Conjecture~\ref{hypo:HPC}.
The inequality~\eqref{eq:p} implies
\[
\liminf_{p\to \infty} \sup_{\tX\in \tP_{\text{non-degree}}(\gamma)}\max_{k\in[K]}\bbE[p\ell(z_k, \hat z_k)]\geq 1.
\]
Based on properties (i)-(ii), we conclude that
\[
\liminf_{p\to \infty} \sup_{\tX\in \tP(\gamma)}\max_{k\in[K]}\bbE[p\ell(z_k, \hat z_k)]\geq 1.
\]
We complete the proof by verifying the properties (i)-(ii). For (i), we verify that the angle gap for the core tensor $\tS$ in~\eqref{eq:S} is on the order of $\sigma p^{-\gamma/2}$. Specifically, write $\mone=(1,1,1)$ and $\me=(1,-1,0)$. We have
\[
\Mat(\tS)=
\begin{bmatrix}
\text{Vec}(\mone^{\otimes K-1})+\sigma p^{-\gamma/2}\text{Vec}\left(\me^{\otimes (K-1)}\right) \\
\text{Vec}(\mone^{\otimes K-1})-\sigma p^{-\gamma/2}\text{Vec}\left(\me^{\otimes (K-1)}\right) \\
\text{Vec}(\mone^{\otimes K-1})
\end{bmatrix}.
\]
Based on the orthogonality $\langle \mone, \me\rangle=0$, the minimal angle gap among rows of $\Mat(\tS)$ is
\begin{align}
\Delta^2_{\min}(\tS)&\asymp \tan^2(\Mat(\tS)_{1:}, \Mat(\tS)_{3:})
\\
&=\left(\onorm{\me}_2 \over \onorm{\mone}_2\right)^{2(K-1)} \sigma^2 d^{-\gamma}\\
&\asymp \sigma^2 d^{-\gamma}.
\end{align}
Therefore, we have shown that $\tP_{\text{shifited}}(\gamma)=\tP(\gamma)$. Finally, the property (ii) follows directly by comparing the definition of $\tS$ in~\eqref{eq:S} with that in the proof of \cite[Theorem 7]{han2020exact}. 
\end{proof}

\subsection{Proof of Theorem~\ref{thm:initial} and Proposition~\ref{prop:ber}} \label{sec:initial_prove}

\begin{proof}[Proof of Theorem~\ref{thm:initial}] We prove Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with symmetric mean tensor and parameters $(z, \tS, \mtheta)$ {fixed $r\geq 1, K \geq 2$}, and i.i.d.\ noise. {For the case $r = 1$, we have $L(z^{(0)},z) = 0, \ell(z^{(0)}, z) = 0$ trivially. Hence, we focus on the proof of the first mode clustering $z^{(0)}_1$ with $r \geq 2$}; 
%and i.i.d.\ noise. We focus on the proof of the first mode clustering $z^{(0)}_1$; 
the proofs for the other modes can be extended similarly. We drop the subscript $k$ in the matricizations $\mM_k, \mX_k, \mS_k$ and in the estimate $z^{(0)}_1$. We firstly show the proof with balanced $\mtheta$.
%For simplicity, let $\hat z$ denote the output, $z^{(0)}$, of Sub-algorithm~\hyperref[alg:main]{1}.

\textbf{We firstly show the upper bound for misclustering error $\ell(z^{(0)}, z)$.}
%and then prove the upper bound for the loss $L^{z^{(0)}, z}$

First, by Lemma~\ref{lem:angle_gap_x}, there exists a positive constant such that $\min_{z(i) \neq z(j)} \onorm{ \mX^s_{i:} - \mX^s_{j:} } \geq c_0  \Delta_{\min}$. By the balance assumption on $\mtheta$ and Lemma~\ref{lem:upper_mis}, we have 
 \begin{equation}\label{eq:theta_bound}
          \min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2  \leq  \sum_{i \in S_I} \theta(i)^2 + 4 \sum_{i \in S} \theta(i)^2 ,
    \end{equation}
    where 
    \small
    \begin{equation}
        S_0 = \{i: \onormSize{}{\hat \mX_{i:}} = 0\}, S = \{i \in S_0^c: \onormSize{}{\hat \mx_{z^{(0)}(i)} - \mX_{i:}^s } \geq c_0  \Delta_{\min}/2\}.
    \end{equation}
    \normalsize
    On one hand, note that for any set $P \in [p]$,
    \begin{align}
        \sum_{i \in P} \onormSize{}{\mX_{i:}}^2 &= \sum_{i \in P} \onormSize{}{ \theta(i) \mS_{z(i):} (\mTheta \mM)^{T, \otimes (K-1)} }^2 \\
        & \geq \sum_{i \in P} \theta(i)^2 \min_{a \in [r]} \onormSize{}{\mS_{a:}}^2 \lambda_{r}^{2(K-1)} (\mTheta \mM)\\
        & \gtrsim \sum_{i \in P} \theta(i)^2 p^{K-1} r^{-(K-1)}, 
    \end{align}
    where the last inequality follows Lemma~\ref{lem:singular_thetam}, the assumption that $\min_{i \in [p]} \theta(i) \geq c$, and the constraint $\min_{a \in [r]} \onormSize{}{\mS_{a:}} \geq c_3$ in the parameter space~\eqref{eq:family}. Thus, we have 
    \begin{equation}\label{eq:theta_p}
        \sum_{i \in P} \theta(i)^2 \lesssim \sum_{i \in P} \onormSize{}{\mX_{i:}}^2 p^{-(K-1)} r^{K-1}.
    \end{equation}

    On the other hand, note that 
    \begin{align}
         &\sum_{i \in S} \onormSize{}{\mX_{i:}}^2\\
         & \leq 2 \sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  +  2\sum_{i \in S} \onormSize{}{\hat \mX_{i:} - \mX_{i:}}^2 \label{eq:tr1}\\
         & \leq \frac{8}{c_0^2 \Delta_{\min}^2  } \sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2   \onormSize{}{\hat \mx_{z^{(0)}(i)} - \mX_{i:}^s }^2 
          +2 \onormSize{}{\hat \tX - \tX}_F^2 \label{eq:defs} \\
         & \leq \frac{16}{c_0^2 \Delta_{\min}^2  }\sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  \off{ \onormSize{}{\hat \mx_{z^{(0)}(i)} - \hat \mX_{i:}^s }^2 + \onormSize{}{\hat \mX_{i:}^s - \mX_{i:}^s }^2 } \\
         & \quad \quad \quad \quad \quad \quad + 2\onormSize{}{\hat \tX - \tX}_F^2 \label{eq:tr2} \\
         & \leq  \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 }\sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  \onormSize{}{\hat \mX_{i:}^s - \mX_{i:}^s }^2  + 2\onormSize{}{\hat \tX - \tX}_F^2 \label{eq:km} \\
         & \leq \of{ \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 } + 2 }\onormSize{}{\hat \tX - \tX}_F^2 \label{eq:lem1}\\
         & \lesssim \of{ \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 } + 2 } \of{ p^{K/2}r + pr^2 + r^K } \sigma^2 \label{eq:lem_err},
    \end{align}
    where inequalities \eqref{eq:tr1} and \eqref{eq:tr2} follow from the triangle inequality, \eqref{eq:defs} follows from the definition of $S$, \eqref{eq:km} follows from the update rule of $k$-means in Step 5 of Sub-algorithm~\hyperref[alg:main]{1}, \eqref{eq:lem1} follows from Lemma~\ref{lem:norm_diff}, and the last inequality~\eqref{eq:lem_err} follows from Lemma~\ref{lem:two-step_esterror}. Also, note that 
    \begin{align}
        \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2 &=  \sum_{i \in S_0} \onormSize{}{\hat \mX_{i:}- \mX_{i:}}^2 \\
        &\leq \onormSize{}{\hat \tX - \tX}_F^2 \\
        &\lesssim \of{ p^{K/2}r + pr^2 + r^K } \sigma^2,\label{eq:s0}
    \end{align}
    where the equation follows from the definition of $S_0$. Therefore, combining the inequalities~\eqref{eq:theta_bound}, \eqref{eq:theta_p}, \eqref{eq:lem_err}, and \eqref{eq:s0}, we have 
    \begin{align}
        &\min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2  \\
        &\lesssim \of{\sum_{i \in S} \onormSize{}{\mX_{i:}}^2 + \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2  } p^{-(K-1)} r^{K-1} \\
        & \lesssim \frac{\sigma^2 r^{K-1}}{ \Delta_{\min}^2  p^{K-1} } \of{ p^{K/2}r + pr^2 + r^K }.\label{eq:theta_sum}
    \end{align}
    With the assumption that $\min_{i \in [p]} \theta(i) \geq c$, we finally obtain the result
    \begin{equation}
        \ell(z^{(0)}, z) \lesssim \frac{1}{p}\min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2 \lesssim \frac{r^K p^{-K/2} }{ \text{SNR} },
    \end{equation}
    where the last inequality follows from the definition $\text{SNR} = \Delta_{\min}^2/\sigma^2$.

      {Without the balanced $\mtheta$, we have $\min_{z(i) \neq z(j)} \onorm{ \mX^s_{i:} - \mX^s_{j:} } \geq c_0  \Delta_{\mX}$. Replacing the definition of $S$ with $\Delta_{\mX}$, we obtain the desired result.}

    \textbf{Next, we show the bound for $L(z^{(0)}, z).$}

    Note that $\mX_{i:}^s$ have only $r$ different values. We let $\mX_a^s = \mX_{i:}^s$ for all $i$ such that $z(i) = a, a \in [r]$. 
    %We firstly show the results for initialization and then prove the MLE results.
    Notice that 
\begin{equation}
    \onormSize{}{\mX_{i:}}^2 \gtrsim p^{K-1} r^{-(K-1)} 
\end{equation}
and
\begin{equation}
    \onormSize{}{\mX_{i:} - \hat \mX_{i:}}^2 \leq \onormSize{}{\hat \tX - \tX}_F^2 \lesssim  p^{K/2}r + pr^2 + r^K .
\end{equation}
Therefore, when $p$ is large enough, we have 
\begin{align}
    &\sum_{i \in [p]} \onormSize{}{\mX_{i:}}^2 \onormSize{}{\hat \mX_i^s - \hat \mx_{z^{(0)}(i)}}^2\\
    &\lesssim \sum_{i \in [p]} \of{\onormSize{}{\mX_{i:}}^2 - \onormSize{}{\mX_{i:} - \hat \mX_{i:}}^2} \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2\\
   & \lesssim \sum_{i \in [p]} \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 \\
   & \lesssim \eta  \sum_{i \in [p]} \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \mX_{i:}^s}^2 \\
   & \lesssim \onormSize{}{\hat \tX - \tX}_F^2 \\
   & \lesssim p^{K/2}r + pr^2 + r^K .\label{eq:cor_1}
\end{align}

Hence, we have 
\begin{align}
    \sum_{i \in [p]} \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 &\lesssim \sum_{i \in [p]} \theta(i)^2 \onormSize{}{\hat \mX_i^s - \hat \mx_{z^{(0)}(i)}}^2 \\
    &\lesssim \frac{r^{K-1}}{p^{K-1}} \sum_{i \in [p]} \onormSize{}{\mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2\\
    & \lesssim \frac{r^{K-1}}{p^{K-1}} \of{ p^{K/2}r + pr^2 + r^K },\label{eq:cor_2}
\end{align}
where the first inequality follows from the assumption $\min_{i \in [p]} \theta(i) \geq c$, the second inequality follows from the inequality~\eqref{eq:theta_p}, and the last inequality comes from the inequality~\eqref{eq:cor_1}.

Next, we consider the following quantity,
\begin{align}
    &\sum_{i \in [p]} \theta(i) \onormSize{}{ \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2\\
    &\lesssim \sum_{i \in [p]} \theta(i)^2 \onormSize{}{\mX_{i:}^s - \hat \mX_{i:}^s}^2 + \sum_{i \in [p]} \theta(i)^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)} }^2\\
    & \lesssim \sum_{i \in [p]} \frac{ \theta(i)^2}{ \onormSize{}{\mX_{i:}}^2} \onormSize{}{\mX_{i:}  - \hat \mX_{i:}}^2 +  \sum_{i \in [p]} \theta(i)^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)} }^2\\
    & \lesssim \frac{r^{K-1}}{p^{K-1}} \of{ p^{K/2}r + pr^2 + r^K} ,\label{eq:cor_3}
\end{align}
where the first inequality follows from the assumption of $\theta(i)$ and triangle inequality, the second inequality follows from Lemma~\ref{lem:norm_diff}, and the last inequality follows from~\eqref{eq:cor_2}. In addition, with Theorem~\ref{thm:initial} and the condition SNR $\gtrsim p^{-K/2} \log p$, for all $a \in [r]$, we have
\small
\begin{equation}
    |z^{-1}(a) \cap (z^{(0)})^{-1} (a)| \geq |z^{-1}(a)| - p\ell(z^{(0)} ,  z) \gtrsim \frac{p}{r} - \frac{p}{\log p} \gtrsim \frac{p}{r},
\end{equation}
\normalsize
when $p$ is large enough. Therefore, for all $a \in [r]$, we have
\small
\begin{align}
       \onormSize{}{ \hat \mx_{a} - \mX_{a}^s  }^2  &= \frac{ \sum_{i \in z^{-1}(a) \cap (z^{(0)})^{-1} (a) } \onorm{ \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 }{|z^{-1}(a) \cap (z^{(0)})^{-1} (a) |}  \\
       & \lesssim \frac{r}{p} \of{  \sum_{i \in [p]} \onormSize{}{ \mX_{i:}^s - \hat \mX_{i:}^s }^2 + \sum_{i \in [p]} \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 } \\
       & \lesssim  \frac{r^{K}}{p^{K}}\of{ p^{K/2}r + pr^2 + r^K}, \label{eq:cor_4}
\end{align}
\normalsize
where the last inequality follows from the inequality~\eqref{eq:cor_2}.

Finally, we obtain 
\begin{align}
     L^{(0)} &= \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(0)}(i) = b } \onormSize{}{ [ \mS_{  z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
     & \lesssim \frac{1}{p}  \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i)  \onormSize{}{ \mX_{i:}^s -  \mX_{z^{(0)}(i)}^s  }^2\\
     & \lesssim \frac{1}{p}  \sum_{i \in [p], z^{(0)}(i) \neq z(i)} \theta(i) \Big( \onormSize{}{\mX_{i:}^s - \hat \mx_{z^{(0)}(i)}}^2 \\
     &  \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad\quad \quad + \onormSize{}{ \hat \mx_{z^{(0)}(i)} - \mX_{z^{(0)}(i)}^s  }^2 \Big) \\
     & \leq \bar C \frac{r^{K}}{p^{K}} \of{ p^{K/2}r + pr^2 + r^K}, \\
     & \leq \frac{\bar C\Delta_{\min}^2}{ \tilde C r \log p}
\end{align}
where the first inequality follows from Lemma~\ref{lem:angle_gap_x}, the third inequality follows from inequalities~\eqref{eq:cor_3} and \eqref{eq:cor_4}, and the last inequality follows from the assumption that SNR $\geq \tilde C p^{-K/2} \log p$.
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:ber}]  Sub-algorithm~3 shares the same algorithm strategy as Sub-algorithm~\hyperref[alg:main]{1} but with a different estimation of the mean tensor, $\hat \tX'$. Hence, the proof of Proposition~\ref{prop:ber} follows the same proof idea with the proof of Theorem~\ref{thm:initial}. Replacing the estimation $\hat \tX$ by $\hat \tX'$ in the proof of Theorem~\ref{thm:initial}, we have 
\begin{align}
        & \min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2  \\
        &\quad \lesssim \of{\sum_{i \in S} \onormSize{}{\mX_{i:}}^2 + \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2  } p^{-(K-1)} r^{K-1}. \label{eq:binary_theta}
    \end{align}
    \normalsize
By inequalities~\eqref{eq:lem1} and \eqref{eq:s0}, we have 
\begin{align}
   \sum_{i \in S} \onormSize{}{\mX_{i:}}^2 &\leq  \of{ \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 } + 2 }\onormSize{}{\hat \tX' - \tX}_F^2, \label{eq:binary_theta1}  \\
   \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2 &\leq  \onormSize{}{\hat \tX' - \tX}_F^2. \label{eq:binary_theta2}
\end{align}
Hence, it suffices to find the upper bound of the estimation error $\onormSize{}{\hat \tX' - \tX}_F^2$ to complete our proof. Note that the matricization $\Mat_{sq}(\tX) \in \bbR^{p^{\floor{K/2}} \times p^{\ceil{K/2}}}$ has $\text{rank}(\Mat_{sq}(\tX)) \leq r^{\ceil{K/2}}$, and Bernoulli random variables follow the sub-Gaussian distribution with bounded variance $\sigma^2 = 1/4$. Apply Lemma~\ref{lem:lowrank} to $\mY = \Mat_{sq}(\tY), \mX = \Mat_{sq}(\tX)$, and $\hat \mX = \Mat_{sq}(\hat \tX')$. Then, with probability tending to 1 as $ p \rightarrow \infty$, we have 
        \begin{equation}\label{eq:binary_est}
            \onormSize{}{\hat \tX' - \tX}_F^2 = \onormSize{}{\Mat_{sq}(\hat \tX') - \Mat_{sq}(\tX)}_F^2 \lesssim p^{\ceil{K/2}}.
        \end{equation}
Combining the estimation error~\eqref{eq:binary_est} with inequalities~\eqref{eq:binary_theta1}, \eqref{eq:binary_theta2}, and \eqref{eq:binary_theta}, we obtain 
\begin{equation}\label{eq:binary_theta_sum}
    \min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2 \lesssim  \frac{\sigma^2 r^{K-1}}{ \Delta_{\min}^2  p^{K-1} } p^{\ceil{K/2}}.
\end{equation}
Replace the inequality~\eqref{eq:theta_sum} in the proof of Theorem~\ref{thm:initial} by inequality~\eqref{eq:binary_theta_sum}. With the the same procedures to obtain $\ell(\hat z^{(0)}, z)$ and $L(\hat z^{(0)}, z)$ for Theorem~\ref{thm:initial} , we finish the proof of Proposition~\ref{prop:ber}.
\end{proof}

{\bf Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:initial}}

\begin{lem}[Basic inequality]\label{lem:norm_diff} For any two nonzero vectors $\mv_1,\mv_2$ of same dimension, we have 
\[
\sin(\mv_1 , \mv_2) \leq \onorm{{\mv_1^s}-{\mv_2^s}}\leq {2\onorm{\mv_1-\mv_2}\over \max\left(\onorm{\mv_1},\onorm{\mv_2}\right)}.
\]
\end{lem}
\begin{proof}[Proof of Lemma~\ref{lem:norm_diff}]
For the first inequality, let $\alpha \in [0,\pi]$ denote the angle between $\mv_1$ and $\mv_2$. We have 
\begin{align}
     \onorm{{\mv_1^s}-{\mv_2^s}} = \sqrt{2(1 - \cos \alpha)} = 2 \sin \frac{\alpha}{2} \geq \sin \alpha,
\end{align}
where the equations follows from the properties of trigonometric function and the inequality follows from the fact the $\cos \frac{\alpha}{2} \leq 1$ and $\sin \alpha = 2 \sin \frac{\alpha}{2} \cos \frac{\alpha}{2} > 0$ for $\alpha \in [0, \pi]$.

For the second inequality, without loss of generality, we assume $\onorm{\mv_1}\geq \onorm{\mv_2}$. Then
\begin{align}
\onorm{{\mv_1^s}-{\mv_2^s}}&=\onorm{{\mv_1\over \onorm{\mv_1}}- {\mv_2\over \onorm{\mv_1}}+{\mv_2\over \onorm{\mv_1}}-{\mv_2\over \onorm{\mv_2}}}\\
&\leq {\onorm{\mv_1-\mv_2}\over \onorm{\mv_1}}+{\onorm{\mv_2}\onorm{\mv_1}-\onorm{\mv_2}\over \onorm{\mv_1}\onorm{\mv_2}}\\
&\leq {2\onorm{\mv_1-\mv_2}\over \onorm{\mv_2}}.
\end{align}

Therefore, Lemma~\ref{lem:norm_diff} is proved.
\end{proof}

\begin{defn}[Weighted padding vectors]\label{def:pad} For a vector $\ma = \entry{a_i} \in \bbR^d$, we define the padding vector of $\ma$ with the weight collection $\mw = \{\mw_i\colon \mw_i = \entry{w_{ik}} \in \bbR^{p_i}\}_{i = 1}^d$ as
\begin{equation}\label{eq:paddef}
    \pad_{\mw}(\ma) = [a_1 \circ \mw_1, \ldots, a_d \circ \mw_d]^T, 
\end{equation}
where $ a_i \circ \mw_i = [a_i w_{i1},\ldots, a_i w_{i p_i}]^T, \text{ for all } i \in [d].$
Here we also view $\pad_{\mw}(\cdot)\colon \bbR^{d}\mapsto \bbR^{\sum_{i\in[d]}p_i}$ as an operator. 
We have the bounds of the weighted padding vector
\begin{equation}\label{eq:pad_bound}
     \min_{i \in [d]} \onormSize{}{\mw_i}^2 \onormSize{}{\ma}^2 \leq \onormSize{}{\pad_{\mw}(\ma)}^2 \leq \max_{i \in [d]} \onormSize{}{\mw_i}^2 \onormSize{}{\ma}^2.
\end{equation}
Further, we define the inverse weighted padding operator $\pad^{-1}: \bbR^{\sum_{i\in[d]}p_i}\mapsto \bbR^d$ which satisfies 
\begin{equation}
    \pad^{-1}_{\mw}(\pad_{\mw}(\ma)) = \ma.
\end{equation}
\end{defn}

\begin{lem}[Angle for weighted padding vectors]\label{lem:pad} Suppose we have two non-zero vectors $\ma, \mb \in \bbR^d$. Given the weight collection $\mw$, we have 
\begin{align}
    \frac{\min_{i \in [d]} \onormSize{}{\mw_i}}{\max_{i \in [d]} \onormSize{}{\mw_i} } \sin (\ma, \mb) & \stackrel{*}{\leq} \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb)) \\
    & \stackrel{**}{\leq} \frac{ \max_{i \in [d]} \onormSize{}{\mw_i} }{\min_{i \in [d]} \onormSize{}{\mw_i}} \sin (\ma, \mb). \label{eq:lempad}
\end{align}
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:pad}] We prove the two inequalities separately with similar ideas.

First, we prove the inequality ** in~\eqref{eq:lempad}. Decomposing $\mb$ yields
\begin{equation}
    \mb = \cos (\ma, \mb) \frac{\onormSize{}{\mb}}{\onormSize{}{\ma}} \ma + \sin (\ma, \mb) \frac{\onormSize{}{\mb}}{\onormSize{}{\ma^{\perp}}}\ma^{\perp},
\end{equation}
where $\ma^{\perp} \in \bbR^d$ is in the orthogonal complement space of $\ma$. By the Definition~\ref{def:pad}, we have 
\begin{equation}\label{eq:pad_second_ineq}
    \pad_{\mw}(\mb) = \cos (\ma, \mb) \frac{\onormSize{}{b}}{\onormSize{}{a}} \pad_{\mw}(\ma) + \sin (\ma, \mb) \frac{\onormSize{}{b}}{\onormSize{}{a^{\perp}}} \pad_{\mw}(\ma^{\perp}).
\end{equation}
Note that $\pad_{\mw}(\ma^{\perp})$ is not necessary equal to the orthogonal vector of $\pad(\ma)$; i.e., $\pad_{\mw}(\ma^{\perp}) \neq (\pad_{\mw}(\ma))^{\perp}$. By the geometry property of trigonometric functions, we obtain
\begin{align}
    \sin(\pad_{\mw}(\ma),  \pad_{\mw}(\mb))  &\leq \frac{  \onormSize{}{\mb} \onormSize{}{\pad_{\mw}(\ma^{\perp})} }{ \onormSize{}{\ma^{\perp}} \onormSize{}{ \pad_{\mw}(\mb)}} \sin (\ma, \mb)  \\
    & \leq  \frac{ \max_{i \in [d]} \onormSize{}{\mw_i} }{\min_{i \in [d]} \onormSize{}{\mw_i}} \sin (\ma, \mb),
\end{align}
where the second inequality follows by applying the property~\eqref{eq:pad_bound} to vectors $\mb$ and $\ma^{\perp}$.

Next, we prove inequality * in~\eqref{eq:lempad}. With the decomposition of $\pad_{\mw}(\mb)$ and the inverse weighted padding operator, we have 
\small
\begin{align}
    \mb &=  \cos(\pad_{\mw}(\ma), \pad_{\mw}(\mb)) \frac{\onormSize{}{\pad_{\mw}(\mb)}}{\onormSize{}{\pad_{\mw}(\ma)}} \ma \\
    & +  \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb))\frac{\onormSize{}{\pad_{\mw}(\mb)}}{\onormSize{}{(\pad_{\mw}(\ma))^{\perp}}} \pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp}).
\end{align}
\normalsize

Therefore, we obtain 
\begin{align}
    &\sin(\ma, \mb) \\
    &\leq \frac{ \onormSize{}{\pad_{\mw}(\mb) } \onormSize{}{\pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp})  } }{ \onormSize{}{(\pad_{\mw}(\ma))^{\perp} } \onormSize{}{\mb} }  \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb))\\
    & \leq \frac{ \max_{i \in [d]} \onormSize{}{\mw_i} }{\min_{i \in [d]} \onormSize{}{\mw_i}} \sin(\pad_{\mw}(\ma), \pad_{\mw}(\mb)),
\end{align}
where the second inequality follows by applying the property~\eqref{eq:pad_bound} to vectors $\mb$ and $\pad_{\mw}^{-1}((\pad_{\mw}(\ma))^{\perp})$.
\end{proof}

\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the parameter space~\eqref{eq:family} and assumption that $\min_{i \in [p]} \theta(i) \geq c$ {for some constant $c >0$}, the singular values of $\mTheta \mM$ are bounded as 
\begin{align}
    \sqrt{p/r} &\lesssim \sqrt{ \min_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 }  \leq \lambda_{r}(\mTheta \mM) \\
    &\ \leq \onormSize{}{\mTheta \mM}_{\sigma} \leq\sqrt{ \max_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 } \lesssim  p /r . \label{eq:membership2}
\end{align}
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:singular_thetam}] Note that 
\begin{equation}
    (\mTheta\mM)^T \mTheta \mM = \mD,
\end{equation}
with $\mD = \text{diag}(D_1,\ldots, D_r)$ where $D_a = \onormSize{}{\mtheta_{z^{-1}(a)}}^2, a \in [r]$.
By the definition of singular values, we have 
\begin{equation}
     \sqrt{\min_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 } \leq \lambda_r(\mTheta \mM) \leq \onormSize{}{\mTheta\mM}_{\sigma} \leq \sqrt{ \max_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2}.
\end{equation}

Since that $\min_{i \in [p]}\theta(i) \geq c$ by the constraints in parameter space, we have  
\begin{equation}
   \min_{a \in [r]} \onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 \geq c^2 \min_{a \in [r]} |z^{-1}(a)| \gtrsim \frac{p}{r},
\end{equation}
where the last inequality follows from the constraint in parameter space~\eqref{eq:family}. Finally, notice that 
\begin{equation}
    \sqrt{\max_{a \in [r]}\onormSize{}{ \mtheta_{ z^{-1}(a)} }^2 } \leq \max_{a \in [r]} \sqrt{ \onormSize{}{ \mtheta_{ z^{-1}(a)} }^2_1} \lesssim \frac{p}{r}.
\end{equation}

 Therefore, we complete the proof of Lemma~\ref{lem:singular_thetam}.
\end{proof}

\begin{lem}[Singular-value gap-free tensor estimation error bound]\label{lem:two-step_esterror}Consider an order-$K$ tensor $\tA = \tX + \tZ \in \bbR^{p \times \cdots \times p}$, where $\tX$ has Tucker rank $(r,...r)$ and $\tZ$ has independent sub-Gaussian entries with parameter $\sigma^2$. Let $\hat \tX$ denote the double projection estimated tensor in Step 2 of Sub-algorithm~\hyperref[alg:main]{1} in the main paper. Then with probability at least $1 - C \exp\of{- cp }$, we have
\begin{align}
    \onormSize{}{\hat \tX - \tX}_F^2 \leq C \sigma^2 \of{ p^{K/2}r + pr^2 + r^K },
\end{align}
where $C, c$ are some positive constants.
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:two-step_esterror}]
See \cite[Proposition 1]{han2020exact}.
\end{proof}

\begin{lem}[Upper bound of misclustering error]\label{lem:upper_mis} Let $z: [p] \mapsto [r]$ be a cluster assignment such that $|z^{-1}(a)| \asymp p/r$ for all $a \in [r]$ {with $r \geq 2$}. Let node $i$ correspond to a vector $\mx_i  = \theta(i) \mv_{z(i)} \in \bbR^d$, where $\{\mv_a\}_{a = 1}^r$ are the cluster centers and $\mtheta = \entry{\theta(i)} \in \bbR^p_+$ is the positive degree heterogeneity.  Assume that  $\mtheta$ satisfies the balanced assumption~\eqref{eq:degree} such that ${\max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 \over \min_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 }= 1 + o(1)$. Consider an arbitrary estimate $\hat z$ with $\hat \mx_i = \hat \mv_{\hat z(i)}$ for all $ i \in S$. Then, if
\begin{equation}\label{eq:upper_mis_cond}
    \min_{a \neq b \in [r]} \onormSize{}{\mv_a - \mv_b} \geq 2c,
\end{equation}
 for some constant $c >0$, we have 
\begin{equation}
    \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \leq \sum_{i \in S_0} \theta(i)^2 + 4 \sum_{i \in S} \theta(i)^2,
\end{equation}
where $S_0$ is defined in {Step 4} of Sub-algorithm~\hyperref[alg:main]{1} and
\begin{equation}
   S = \{i \in S_0^c: \onormSize{}{\hat \mx_i - \mv_{z(i)} } \geq c \}.
\end{equation}

\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:upper_mis}]

For each cluster $u\in[r]$, we use $C_u$ to collect the subset of points for which the estimated and true positions $\hat \mx_i, \mx_i$ are within distance $c$. Specifically, define
\begin{equation}
    C_u = \{ i \in z^{-1}(u) \cap S_0^c: \onormSize{}{\hat \mx_i - \mv_{z(i)}} < c \},
\end{equation}
and divide $[r]$ into three groups based on $C_u$ as 
\begin{align}
    R_1 &= \{ u \in [r]: C_u = \emptyset \},\\
    R_2 &= \{ u \in [r]: C_u \neq \emptyset, \text{ for all } i, j \in C_u, \hat z(i) = \hat z(j) \},\\
    R_3 &= \{ u \in [r]: C_u \neq \emptyset, \text{ there exist } i, j \in C_u, \hat z(i) \neq \hat z(j) \}.
\end{align}
Note that $\cup_{u \in [r]}C_u = S_0^c/S^c$ and $C_u \cap C_v = \emptyset$ for any $u \neq v$. Suppose there exist $ i \in C_u$ and $j \in C_v$ with $u \neq v \in [r]$ and $\hat z(i) = \hat z(j)$. Then we have 
\begin{equation}
    \onormSize{}{\mv_{z(i)} - \mv_{z(j)} } \leq  \onormSize{}{\mv_{z(i)} - \hat \mx_{i} } + \onormSize{}{\mv_{z(j)} - \hat \mx_j } < 2c,
\end{equation}
which contradicts to the assumption~\eqref{eq:upper_mis_cond}. Hence, the estimates $\hat z(i) \neq \hat z(j)$ for the nodes $ i \in C_u$ and $j \in C_v$ with $u \neq v$. By the definition of $R_2$, the nodes in $\cup_{u \in R_2} C_u$ have the same assignment with $z$ and $\hat z$. Then, 
%Following \cite[Lemma 6]{gao2018community} \fixme{Miaoyan}{Remove citation. fill in details in a self-contained way. }, 
we have 
\small
\begin{equation}
    \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \leq \sum_{i \in S_0} \theta(i)^2 + \sum_{i \in S} \theta(i)^2 + \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2.
\end{equation}
\normalsize
 We only need to bound $\sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2$ to finish the proof. Note that every $C_u$ with $u \in R_3$ contains at least two nodes assigned to different clusters by $\hat z$. Then, we have $|R_2| + 2 |R_3| \leq r$. Since $|R_1| + |R_2| + |R_3| = r$, we have $|R_3| \leq |R_1|$. Hence, we obtain
\begin{align}
    \sum_{ i \in \cup_{u \in R_3} C_u} \theta(i)^2 &\leq |R_3| \max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 \\
    & \leq |R_1| \max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2  \\
    & \leq {\max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 \over \min_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 } \sum_{i \in \cup_{u \in R_1} z^{-1}(u)} \theta(i)^2\\
    & \leq 2 \sum_{i \in S} \theta(i)^2,
\end{align}
where the last inequality holds by the balanced assumption on $\mtheta$ when $p$ is large enough, and the fact that $ \cup_{u \in R_1} z^{-1}(u) \subset S$.
\end{proof}

\begin{lem}[Low-rank matrix estimation] \label{lem:lowrank} Let $\mY = \mX + \mE \in \bbR^{m \times n}$, where {$n > m$ and} $\mE$ contains independent mean-zero sub-Gaussian entries with bounded variance $\sigma^2$. Suppose $\text{rank}(\mX) = r$. Consider the least square estimator 
        \begin{equation}
            \hat \mX = \argmin_{\mX' \in \bbR^{m \times n}, \text{rank}(\mX') \leq r} \onormSize{}{\mX' - \mY}_F^2.
        \end{equation}
        There exist positive constant $C_1, C_2$ such that
        \begin{equation}
            \onormSize{}{\hat \mX - \mX}_F^2 \leq C_1 \sigma^2 nr,
        \end{equation}
        with probability at least $1 - \exp(-C_2 nr)$.

% denote a binary matrix with independent Bernoulli entries. Suppose that $n \geq m$ and $\bbE[\mY] = \mX \in [0,1]^{m \times n}$ with fixed rank $\text{rank}(\mX) \leq r$. Consider the least square estimator 
%         \begin{equation}
%             \hat \mX = \argmin_{\mX' \in \bbR^{m \times n}, \text{rank}(\mX') \leq r} \onormSize{}{\mX' - \mY}_F^2.
%         \end{equation}
%         There exist positive constant $C_1, C_2$ such that
%         \begin{equation}
%             \onormSize{}{\hat \mX - \mX}_F^2 \leq C_1 nr,
%         \end{equation}
%         with probability at least $1 - \exp(-C_2 nr)$.
        \end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:lowrank}] Note that $\onormSize{}{\hat \mX - \mY}_F^2 \leq \onormSize{}{\mX - \mY}_F^2$ by the definition of least square estimator. 
{

We have 
\small
        \begin{align}
            &\onormSize{}{\hat \mX - \mX}_F^2\\
            & \leq 2 \ang{ \hat \mX - \mX, \mY - \mX} \\ & \leq 2 \onormSize{}{\hat \mX - \mX}_F \sup_{\mT \in \bbR^{m \times n}, \text{rank}(\mT) \leq 2r, \onormSize{}{\mT}_F = 1} \ang{\mT, \mY - \mX} \label{eq:lem9_1}
            %& \leq C \sigma  \onormSize{}{\hat \mX - \mX}_F \sqrt{nr},
        \end{align}
        \normalsize
        with probability at least $1 - \exp(-C_2 nr)$, where the second inequality follows by re-arrangement.

        Consider the SVD for matrix $\mT = \mU \Sigma \mV^T$ with orthogonal matrices $\mU \in \bbR^{m \times 2r}, \mV \in \bbR^{n \times 2r}$ and diagonal matrix $\Sigma \in \bbR^{2r \times 2r}$. We have 
        \begin{align}
            &\sup_{\mT \in \bbR^{m \times n}, \text{rank}(\mT) \leq 2r, \onormSize{}{\mT}_F = 1} \ang{\mT, \mY - \mX} \\
            =& \sup_{\mT \in \bbR^{m \times n}, \text{rank}(\mT) \leq 2r, \onormSize{}{\mT}_F = 1} \ang{\mU \Sigma, \mE \mV} \\
            =&  \sup_{\mv \in \bbR^{2nr}} \mv^T \me \leq C \sigma \sqrt{nr}, \label{eq:lem9_2}
        \end{align}
        with probability $1 - \exp(-C_2 nr)$, where $C, C_2$ are two positive constants, the vectorization $\me = \text{Vec}(\mE \mV) \in \bbR^{2nr}$ has independent mean-zero sub-Gaussian entries with bounded variance $\sigma^2$ due to the orthogonality of $\mV$, and the last inequality follows from \cite[Theorem 1.19]{rigollet2015high}.

        Combining inequalities~\eqref{eq:lem9_1} and \eqref{eq:lem9_2}, we obtain the desired conclusion.
        }

% We have 
%         \begin{align}
%             &\onormSize{}{\hat \mX - \mX}_F^2 \\
%             & \leq 2 \ang{ \hat \mX - \mX, \mY - \mX} \\ & \leq 2 \onormSize{}{\hat \mX - \mX}_F \sup_{\mT \in \bbR^{m \times n}, \text{rank}(\mT) \leq 2r, \onormSize{}{\mT}_F = 1} \ang{\mT, \mY - \mX} \\
%             & \leq C \sigma  \onormSize{}{\hat \mX - \mX}_F \sqrt{nr},
%         \end{align}
%         with probability at least $1 - \exp(-C_2 nr)$, where the second inequality follows by re-arrangement, and the last inequality follows from \cite[Lemma E5]{han2022optimal} under the matrix case.
\end{proof}

\subsection{Proofs of Theorem~\ref{thm:stats} (Achievability) and Theorem~\ref{thm:refinement}}\label{sec:statprove2}

\begin{proof}[Proof of Theorem~\ref{thm:stats} (Achievability) and Theorem~\ref{thm:refinement}]

The proofs of Theorem~\ref{thm:stats} (Achievability) and Theorem~\ref{thm:refinement} share the same idea. We prove the contraction step by step. In each step, we show the specific procedures for the algorithm loss and address the MLE loss by stating the difference.

We consider dTBM~\eqref{eq:model_tensor} with symmetric mean tensor and parameters $(z, \tS, \mtheta)$ {fixed $r\geq 1, K \geq 2$}, and i.i.d.\ noise.  Let $(\hat z, \hat \mtheta, \hat \tS)$ denote the MLE in \eqref{eq:mle}, and $(z^{(0)}_k, \mtheta^{(0)}, \tS^{(0)})$ denote parameters related to the initialization.  {For the case $r = 1$, we have $\ell(z^{(t)}, z) = 0$ trivially for all $t \geq 0$. Hence, we focus on the proof of the first mode clustering $z^{(t+1)}_1$ with $r \geq 2$}; 
%Let $(\hat z, \hat \mtheta, \hat \tS)$ denote the MLE in \eqref{eq:mle}, and $(z^{(0)}_k, \mtheta^{(0)}, \tS^{(0)})$ denote parameters related to the initialization. We focus on the proof of $z^{(t+1)}_1$; 
the extension for other modes can be obtained similarly. We drop the subscript $k$ in the matricizations $\mTheta, \mM_k, \mS_k, \mX_k$ and in estimates $z^{(0)}, z^{(t+1)}_1, z^{(t)}_k$ for ease of the notation.  Without loss of generality, we assume that the variance $\sigma = 1$, and that the identity permutation minimizes the initial misclustering error; i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$, and so for $\hat z$.

{\bf Step 1 (Notation and conditions).} We first introduce additional notations and the necessary conditions used in the proof. We will verify that the conditions hold in our context under high probability in the last step of the proof.

{

\textbf{Notation.}
\begin{enumerate}[wide]

\item Projection. We use $\mI_d$ to denote the identity matrix of dimension $d$. For a vector $\mv \in \bbR^d$, let $\text{Proj}(\mv) \in \bbR^{d \times d}$ denote the projection matrix to $\mv$. Then, $\mI_d - \text{Proj}(\mv)$ is the projection matrix to the orthogonal complement $\mv^{\perp}$.

    \item We define normalized membership matrices
    \small
    \begin{equation}
        \mW = \mM \of{ \text{diag}(\mone_{p}^T \mM) }^{-1}, \mW^{(t)} = \mM^{(t)} \of{ \text{diag}(\mone_{p}^T \mM^{(t)}) }^{-1},
    \end{equation}
    \normalsize
    weighted normalized membership matrices
    \begin{align}
        &\mP = \mTheta \mM ( \text{diag}( \onormSize{}{\mtheta_{z^{-1}(1)}}^2 , \ldots, \onormSize{}{\mtheta_{z^{-1}(r)}}^2 ) )^{-1}, \\
        &\hat \mP = \hat \mTheta \hat \mM ( \text{diag}( \onormSize{}{\hat \mtheta_{z^{-1}(1)}}^2 , \ldots, \onormSize{}{\hat \mtheta_{z^{-1}(r)}}^2 ) )^{-1},
    \end{align}
    and the dual normalized and dual weighted normalized membership matrices
    \begin{align}
         &\mV = \mW^{\otimes (K-1)}, \quad \mV^{(t)} =\of{ \mW^{(t)}}^{\otimes (K-1)}\\
         &\mQ = \mP^{\otimes K-1}, \quad  \hat \mQ = \hat \mP^{\otimes K-1}.
    \end{align}
    Also, let $\mB = (\mTheta \mM)^{\otimes (K-1)}, \hat \mB = (\hat \mTheta \hat \mM)^{\otimes (K-1)}$. By the definition, we have $\mB^T \mQ =\hat \mB^T \hat \mQ =\mI_{r^{K-1}}$.
    \item We use $\tS^{(t)}$ to denote the estimator of $\tS$ in the $t$-th iteration, $\hat \tS$ for MLE, $\tilde \tS$ to denote the oracle estimator of $\tS$ given true assignment $z$, and $\bar \tS$ for weighted oracle estimator; i.e.,
    \begin{align}
        &\tS^{(t)} = \tY \times_1 \of{\mW^{(t)}}^T \times_2 \cdots \times_K \of{\mW^{(t)}}^T,\\ &\tilde \tS = \tY \times_1 \mW^T \times_2 \cdots \times_K \mW^T \\
        &\hat \tS =  \tY \times_1 \hat \mP^T \times_2 \cdots \times_K \hat \mP^T \\
        &\bar \tS  = \tY \times_1 \mP^T \times_2 \cdots \times_K \mP^T.
    \end{align}
    \item We define the matricizations of tensors
    \begin{align}
        \mS = \mat(\tS), \  \mY = \mat(\tY), \  \mX = \mat(\tX), \  \mE = \mat(\tE)
    \end{align}
    \begin{equation}
        \mS^{(t)} = \mat(\tS^{(t)}), \   \hat \mS = \mat(\hat \tS), \  \tilde \mS = \mat(\tilde \tS), \  \bar \mS = \mat( \bar \tS )
    \end{equation}
    \item We define the extended core tensor on $K-1$ modes
    \begin{equation}
        \mA = \mS \mB^T , \quad \bar \mA = \bar \mS \mB^T, \quad \hat \mA = \hat \mS \hat \mB^T.
    \end{equation}
    By the assumption in parameter space~\eqref{eq:family}, we have $\mA = \mP \mX = \mW \mX, \quad \hat \mA = \hat \mP \hat \mX = \hat \mW \hat \mX.$

    \item We define the angle-based misclustering loss in the $t$-th iteration and loss for MLE
    \begin{align}
        L^{(t)} &= \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \{ z^{(t)}(i) = b \} \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2, \\
         L(\hat z)& = \frac{1}{p} \sum_{i \in [p]} \theta(i)^2 \sum_{b \in [r]} \ind\{ \hat z(i) = b\} \onormSize{}{[\mA_{z(i):}]^s - [\mA_{b:}]^s}^2.
    \end{align}
    We also define the loss for oracle and weighted oracle estimators
    \begin{align}
         \xi & = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \Big\{\ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s} \\
         & \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \leq - \frac{ \theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \Big\} \\
         & \quad \quad \quad \quad \quad \quad \quad \quad \cdot \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2,\\
         \xi' & = \frac{1}{p} \sum_{i \in [p]} \theta(i)^2 \sum_{b \in [r]} \ind \Big\{\ang{\mE_{i:}, [\bar \mA_{z(i):}]^s - [\bar \mA_{b:}]^s } \\
         & \quad \quad \quad \quad \quad \quad \quad \quad  \leq - \frac{m'}{4}  \sqrt{\frac{p^{K-1}}{r^{K-1}}} \onormSize{}{ [\mA_{z(i):}]^s -  [\mA_{b:}]^s  }_F^2 \Big\} \\
         & \quad \quad \quad \quad \quad \quad \quad \quad \cdot \onormSize{}{[\mA_{z(i):}]^s - [\mA_{b:}]^s}^2.
    \end{align}
    where $m$ and $m'$ are some positive universal constants.
    %defined in~\eqref{eq:xv_lower} and .
\end{enumerate}
Then we introduce the necessary conditions in Condition~\ref{cond:origin}.

\begin{figure*}[t] 
% ensure that we have normalsize text \normalsize 
% Store the current equation number. 
\begin{condition}(Intermediate results) \label{cond:origin} Let $\mathbb{O}_{p, r}$ denote the collection of all the $p$-by-$r$ matrices with orthonormal columns. We have 
\setcounter{MYtempeqncnt}{\value{equation}} 
% Set the equation number to one less than the one % desired for the first equation here.

% The value here will have to changed if equations 
% are added or removed prior to the place these 
% equations are referenced in the main text. 
\setcounter{equation}{62} 
\begin{equation}\label{eq:cond1}
    \onormSize{}{\mE \mV}_\sigma \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}} \of{ p ^{1/2}+ r^{(K-1)/2}}, \quad \onormSize{}{\mE \mV}_F \lesssim \sqrt{\frac{ r^{2(K-1)}}{p^{K-2}}},\quad \onormSize{}{ \mW_{a:}^T \mE \mV} \lesssim \frac{r^K}{p^{K/2}}, \text{  for all }a \in [r], 
\end{equation}
\begin{equation}\label{eq:cond2}
    \sup_{ \mU_k \in \mathbb{O}_{p, r}, k = 2,\ldots, K } \onormSize{}{ \mE (\mU_{2} \otimes \cdots \otimes \mU_K)}_\sigma \lesssim \of{ \sqrt{r^{K-1}} + K\sqrt{pr}},
\end{equation}
\begin{equation}\label{eq:cond3}
    \sup_{ \mU_k \in \mathbb{O}_{p, r}, k = 2,\ldots, K } \onormSize{}{ \mE (\mU_{2} \otimes \cdots \otimes \mU_K)}_F \lesssim \of{ \sqrt{pr^{K-1}} + K\sqrt{pr}},
\end{equation}
\begin{equation}\label{eq:cond_oracle}
    \xi \leq \exp\of{ - M \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}}, \quad  \xi' \lesssim \exp\of{ -  \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}},
\end{equation}

\begin{equation}\label{eq:cond_intial}
    L^{(t)} \leq \frac{\bar C}{\tilde C} \frac{\Delta_{\min}^2}{r \log p}, \quad \text{for} \quad t = 0, 1, \ldots, T, \quad  L(\hat z) \leq \frac{\bar C}{\tilde C} \frac{\Delta_{\min}^2}{r \log p},
\end{equation}
% Restore the current equation number. 
\setcounter{equation}{\value{MYtempeqncnt}} 
where $M$ is a positive universal constant in inequality~\eqref{eq:constant_M}, $\bar C, \tilde C$ are universal constants in {the proof of Theorem~\ref{thm:initial}} and assumption SNR $\geq \tilde C p^{-K/2} \log p$, respectively. Further, inequality~\eqref{eq:cond1} holds by replacing $\mV$ to $\mV^{(t)}, \mQ, \hat \mQ$ and $\mW_{:a}$ to $\mW_{:a}^{(t),T}, \mP_{:a}^T, \hat \mP_{:a}^T$ when initialization condition~\eqref{eq:cond_intial} holds.
\end{condition}
% The IEEE uses as a separator 
\hrulefill 
% The spacer can be tweaked to stop underfull vboxes. 
\vspace*{4pt} 
\end{figure*}
\setcounter{equation}{67} 
% \begin{equation}\label{eq:cond1}
%     \onormSize{}{\mE \mV}_\sigma \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}} \of{ p ^{1/2}+ r^{(K-1)/2}}, \quad \onormSize{}{\mE \mV}_F \lesssim \sqrt{\frac{ r^{2(K-1)}}{p^{K-2}}},\quad \onormSize{}{ \mW_{a:}^T \mE \mV} \lesssim \frac{r^K}{p^{K/2}} \text{ for all }a \in [r], 
% \end{equation}
% \begin{equation}\label{eq:cond2}
%     \sup_{ \mU_k \in \mathbb{O}_{p, r}, k = 2,\ldots, K } \onormSize{}{ \mE (\mU_{2} \otimes \cdots \otimes \mU_K)}_\sigma \lesssim \of{ \sqrt{r^{K-1}} + K\sqrt{pr}},
% \end{equation}

% \begin{equation}\label{eq:cond3}
%     \sup_{ \mU_k \in \mathbb{O}_{p, r}, k = 2,\ldots, K } \onormSize{}{ \mE (\mU_{2} \otimes \cdots \otimes \mU_K)}_F \lesssim \of{ \sqrt{pr^{K-1}} + K\sqrt{pr}},
% \end{equation}

% \begin{equation}\label{eq:cond_oracle}
%     \xi \leq \exp\of{ - M \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}}, \quad  \xi' \lesssim \exp\of{ -  \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}},
% \end{equation}

% \begin{equation}\label{eq:cond_intial}
%     L^{(t)} \leq \frac{\bar C}{\tilde C} \frac{\Delta_{\min}^2}{r \log p}, \quad \text{for} \quad t = 0, 1, \ldots, T, \quad  L(\hat z) \leq \frac{\bar C}{\tilde C} \frac{\Delta_{\min}^2}{r \log p}
% \end{equation}

}

{\bf Step 2 (Misclustering loss decomposition).} Next, we derive the upper bound of $L^{(t+1)}$ for $t = 0 ,1, \ldots, T-1$. By Sub-algorithm~\hyperref[alg:main]{2}, we update the assignment in $t$-th iteration via
    \begin{equation}
        z^{(t+1)}(i) = \argmin_{a \in [r]} \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{a:}^{(t)}]^s }^2,
    \end{equation}
    following the facts that $\onormSize{}{\ma^s - \mb^s}^2 = 1 - \cos(\ma, \mb)$ for vectors $\ma,\mb$ of same dimension and $\mat(\tY^\text{d}) = \mY \mV^{(t)}$ where $\tY^\text{d}$ is the reduced tensor defined in Step 8 of Sub-algorithm~\hyperref[alg:main]{2}. Then the event $z^{(t+1)}(i) = b$ implies
    \begin{equation}\label{eq:event}
        \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{b:}^{(t)}]^s }^2 \leq \onormSize{}{ [ \mY_{i:} \mV^{(t)}  ]^s - [\mS_{z(i):}^{(t)}]^s }^2.
    \end{equation}
Note that the event~\eqref{eq:event} also holds for the degenerate entity $i$ with $\onormSize{}{\mY_{i:} \mV^{(t)}} = 0$ due to the convention that $\ma^s = {\bf 0}$ if $\ma = {\bf 0}$. 
    Arranging the terms in \eqref{eq:event} yields the decomposition
    \small
    \begin{align}
        &2 \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s } \\
        & \quad \leq \onormSize{}{\mX_{i:} \mV^{(t)}} \of{ - \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 + G_{ib}^{(t)} + H_{ib}^{(t)} } +  F_{ib}^{(t)},\label{eq:decomp}
    \end{align}
    \normalsize
    where
    \small
    \begin{align}
        F_{ib}^{(t)} &= 2 \ang{\mE_{i:} \mV^{(t)}, \of{ [  \tilde \mS_{z(i):} ]^s  -  [  \mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  -  [  \mS_{b:}^{(t)} ]^s  }  }  \\
        & + 2 \ang{ \mE_{i:} \of{ \mV - \mV^{(t)} }, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }, \label{eq:f}\\
        G_{ib}^{(t)} &=   \Big( \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 \\
        & \quad \quad \quad \quad \quad \quad -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2\Big)   \\
        & -   \Big( \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{b:}^{(t)}  ]^s}^2 \\
        & \quad \quad \quad \quad \quad \quad   -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 \Big), \label{eq:g} \\
        H_{ib}^{(t)} &=   \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s \\
        & -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2  + \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 . \label{eq:h}
    \end{align}
    \normalsize

   Therefore, the event $ \ind \offf{ z^{(t+1)}(i) = b }$ can be upper bounded as
   \small
    \begin{align}
         & \ind \offf{ z^{(t+1)}(i) = b  }\\
         & \leq \ind \Bigg\{ z^{(t+1)}(i) = b ,   \ang{ \mE_{j:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \\
         & \quad \quad \quad \quad \quad \quad \quad  \leq - \frac{1}{4} \onormSize{}{\mX_{i:} \mV^{(t)}} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \Bigg\} \\
         & + \ind \Bigg\{z^{(t+1)}(i) = b, \frac{1}{2}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
         & \quad \quad \quad \quad \quad \quad \quad \leq \onormSize{}{\mX_{i:} \mV^{(t)}}^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} \Bigg\}. \label{eq:ind_b}
    \end{align}
    \normalsize
   Note that 
    \begin{align}
     \onormSize{}{\mX_{i:} \mV^{(t)}}  &= \theta(i) \onormSize{}{\mS_{i:} (\mTheta \mM)^{\otimes (K-1),T} \mW^{(t), \otimes^{K-1}}}\\
         &\geq \theta(i) \onormSize{}{\mS_{z(i):}} \lambda_r^{K-1}(\mTheta \mM) \lambda^{K-1}_r(\mW^{(t)})\\
         &\geq\theta(i) m , \label{eq:xv_lower}
    \end{align}
    where the first inequality follows from the property of eigenvalues; the last inequality follows from Lemma~\ref{lem:singular_thetam}, Lemma~\ref{lem:membership}, and assumption that $\min_{a \in [r]} \onormSize{}{\mS_{z(i):}} \geq c_3>0$; and $m >0$ is a positive constant related to $ c_3$. Plugging the lower bound of $ \onormSize{}{\mX_{i:} \mV^{(t)}}$ \eqref{eq:xv_lower} into the inequality \eqref{eq:ind_b} gives
    \begin{equation}\label{eq:decomp1}
        \ind \offf{ z^{(t+1)}(i) = b  } \leq A_{ib} + B_{ib},
    \end{equation}
    where 
    \small
    \begin{align}
        A_{ib} &= \ind \Bigg\{z^{(t+1)}(i) = b, \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \\
        & \quad \quad \quad \quad \quad \quad \quad \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \Bigg\},\\
        B_{ib} &= \ind \Bigg\{z^{(t+1)}(i) = b, \frac{1}{2}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
        & \quad \quad \quad \quad \quad \quad \quad  \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} \Bigg\}.
    \end{align}
    \normalsize
 Taking the weighted summation of~\eqref{eq:decomp1} over $i \in [p]$ yields 
    \begin{equation}
        L^{(t+1)} \leq \xi + \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}^{(t)}, 
    \end{equation}
    where $\xi$ is the oracle loss such that 
    \begin{equation}\label{eq:decomp2}
        \xi =   \frac{1}{p} \sum_{i \in [p] } \theta(i) \sum_{b \in [r]/z(i)} A_{ib} \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2.
    \end{equation}
    Similarly to $\xi$ in~\eqref{eq:decomp2}, we define
    \begin{equation}
        \zeta_{ib}^{(t)} =   \theta(i) B_{ib} \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2.
    \end{equation}

    {

    \textbf{Now, we show the decomposition for MLE loss.}

    By the definition of Gaussian MLE, the estimator $\hat \mtheta$ satisfies $\hat \theta(i)  = \ang{ \mY_{i:} , \hat \mA_{\hat z(i):} }/{\onormSize{}{ \hat \mA_{\hat z(i):}}_F^2}$ for all $i \in [p]$.
Hence, we have 
\begin{equation}
    \hat z(i) = \argmin_{a \in [r_1]} \onormSize{}{ [\mY_{ i:}]^s - [\hat \mA_{a:}]^s}_F^2,
\end{equation}
and the decomposition 
 \begin{equation}
        L(\hat z) \leq \xi' + \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}', 
    \end{equation}
    where $ \zeta_{ib}' =   \theta(i)^2 B_{ib}' \onormSize{}{ [ \mA_{ z(i):}  ]^s - [ \mA_{b:}  ]^s  }^2$ and 
    \small
    \begin{align}
        A'_{ib} &= \ind \Bigg\{  \hat z(i)  = b,   \ang{\mE_{i:}, [\bar \mA_{z(i):}]^s - [\bar \mA_{b:}]^s }  \\
        & \quad \quad \quad \quad \quad \quad \quad  \leq - \frac{m'}{4}  \sqrt{\frac{p^{K-1}}{r^{K-1}}} \onormSize{}{ [\mA_{z(i):}]^s -  [\mA_{b:}]^s  }_F^2 \Bigg\}, \\
    B'_{ib} &= \ind \Bigg\{ \hat z(i)  = b,     - \frac{1}{2}  \onormSize{}{ [\mA_{z(i):}]^s -  [\mA_{b:}]^s  }_F^2  \\
    & \quad \quad \quad \quad \quad \quad \quad  \leq \sqrt{\frac{r^{K-1}}{(m')^2 p^{K-1}}}  \hat F_{ib} + \hat G_{ib} + \hat H_{ib} \Bigg\}
    \end{align}
    \normalsize
    with terms 
    \begin{align}
    \hat F_{ib} &= 2 \ang{ \mE_{i:}, ([\bar  \mA_{z(i):}]^s - [\hat \mA_{a:}]^s) - ([\bar  \mA_{b:}]^s - [\hat \mA_{b:}]^s) }, \\
    \hat G_{ib}& = \of{ \onormSize{}{\mX_{i:}^s - [\hat \mA_{z(i):}]^s}_F^2 - \onormSize{}{ \mX^s_{i:} - [\mP_{:z(i)}^T \mY \hat \mQ \hat \mB^T ]^s }_F^2  } \\
    & \quad \quad - \of{ \onormSize{}{\mX^s_{i:} - [\hat \mA_{b:}]^s}_F^2 - \onormSize{}{ \mX^s_{i:} - [\mP_{:b}^T \mY \hat \mQ \hat \mB^T]^s }_F^2   }, \\
   \hat  H_{ib}& = \onormSize{}{ \mX^s_{i:} - [\mP_{:z(i)}^T \mY \hat \mQ \hat \mB^T]^s }_F^2 - \onormSize{}{ \mX^s_{i:} - [\mP_{:b}^T \mY \hat \mQ \hat \mB^T]^s }_F^2 \\
   & \quad \quad + \onormSize{}{ \mA^s_{z(i):} -  \mA^s_{b:}  }_F^2.
\end{align}

    }

    {\bf Step 3 (Derivation of contraction inequality).} In this step we derive the upper bound of $\zeta_{ib}$ and obtain the contraction inequality~\eqref{eq:proof_5_ineq}. We show the analysis in the following one-column box for a better presentation.

\begin{figure*}[t] 
% ensure that we have normalsize text \normalsize 
% Store the current equation number. 
{\bf Step 3.}  Choose the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ that satisfies the condition of Lemma~\ref{lem:upper_fgh}, inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}. Note that  \begin{align}
        \zeta_{ib}^{(t)} &= \theta(i) \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \ind \offf{z^{(t+1)}(i) = b, \frac{1}{2}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }\\
        & \leq \theta(i) \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \ind \offf{z^{(t+1)}(i) = b, \frac{1}{4}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} } \\
        & \leq 64 \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ cm^2\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    }
        %&\leq  \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    },\label{eq:zeta}
    \end{align}
    where the first inequality follows from the inequality~\eqref{eq:hib} in Lemma~\ref{lem:upper_fgh}, and the last inequality follows from the assumption that $\min_{i \in [p]} \theta(i) \geq c>0$. Following \cite[Step 4, Proof of Theorem 2]{han2020exact} and Lemma~\ref{lem:upper_fgh}, we have 
\setcounter{MYtempeqncnt}{\value{equation}} 
% Set the equation number to one less than the one % desired for the first equation here.

% The value here will have to changed if equations 
% are added or removed prior to the place these 
% equations are referenced in the main text. 
\setcounter{equation}{75} 
\begin{equation}\label{eq:f_sum}
          \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{(F_{ib}^{(t)})^2}{ cm^2 \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \leq \frac{C_0 \bar C}{c m^2 \tilde C^2} L^{(t)},
    \end{equation}
    for a positive universal constant $C$ and 
    \begin{align}
         \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}  &\leq \frac{1}{512} \frac{1}{p} \sum_{i \in [p]}\theta(i) \sum_{b \in [r]/z(i)}   \ind \offf{ z^{(t+1)}(i) = b}  (\Delta_{\min}^2 + L^{(t)}) \\
         & \leq \frac{1}{512} (L^{(t+1)} + L^{(t)}), \label{eq:g_sum}
    \end{align}
 where the last inequality follows from the definition of $L^{(t)}$ and the constraint of $\mtheta$ in parameter space~\eqref{eq:family}. For $\tilde C$ also satisfies 
    \begin{equation}\label{eq:tilde_c3}
         \frac{C_0 \bar C}{c m^2 \tilde C^2} \leq \frac{1}{512},
    \end{equation}
    we have 
    \begin{equation}\label{eq:zeta_upper}
        \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}^{(t)} \leq  \frac{1}{8} L^{(t+1)}  + \frac{1}{4} L^{(t)}.
    \end{equation}
    %for some small \fixme{Miaoyan}{In what sense "small"? Add one sentence.} positive constants $C_4, C_5$. 
    Plugging the inequality~\eqref{eq:zeta_upper} into the decomposition~\eqref{eq:decomp2}, we obtain the contraction inequality 
    \begin{equation}\label{eq:decomp_t}
          L^{(t+1)} \leq \frac{3}{2} \xi  + \frac{1}{2} L^{(t)},
    \end{equation}
     where $\frac{1}{2}$ is the contraction parameter.

     Therefore, with $\tilde C$ satisfying inequalities~\eqref{eq:tilde_c3}, \eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}, we obtain the conclusion in Theorem~\ref{thm:refinement} via inequality~\eqref{eq:decomp_t} combining the inequality~\eqref{eq:cond_oracle} in Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.

     \textbf{We also have the contraction inequality for MLE.}

     Following the same derivation of~\eqref{eq:decomp_t} with the upper bound of $\hat F_{ib}, \hat G_{ib}, \hat H_{ib}$ in Lemma~\ref{lem:intermediate}, we also have
    \begin{equation}\label{eq:decomp_t_mle}
          L(\hat z) \leq \frac{3}{2} \xi'  + \frac{1}{2} L(\hat z),
    \end{equation}
    which indicates the conclusion $\ell(\hat z,z) \lesssim \Delta_{\min}^2  \exp \of{ - \frac{p^{K-1}}{r^{K-1}} \Delta_{\min}^2  }$.
% Restore the current equation number. 
\setcounter{equation}{\value{MYtempeqncnt}}

% The IEEE uses as a separator 
\hrulefill 
% The spacer can be tweaked to stop underfull vboxes. 
\vspace*{4pt} 
\end{figure*}
\setcounter{equation}{78}

   %  Choose the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ that satisfies the condition of Lemma~\ref{lem:upper_fgh}, inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}. Note that 
   %  \begin{align}
   %      \zeta_{ib}^{(t)} &= \theta(i) \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \ind \offf{z^{(t+1)}(i) = b, \frac{1}{2}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} + H_{ib}^{(t)} }\\
   %      & \leq \theta(i) \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \ind \offf{z^{(t+1)}(i) = b, \frac{1}{4}\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq (\theta(i) m)^{-1} F_{ib}^{(t)} + G_{ib}^{(t)} } \\
   %      & \leq 64 \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ cm^2\onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    }
   %      %&\leq  \ind \offf{ z^{(t+1)}(i) = b} \of{  \frac{(F_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} + \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}    },\label{eq:zeta}
   %  \end{align}
   %  where the first inequality follows from the inequality~\eqref{eq:hib} in Lemma~\ref{lem:upper_fgh}, and the last inequality follows from the assumption that $\min_{i \in [p]} \theta(i) \geq c>0$. Following \cite[Step 4, Proof of Theorem 2]{han2020exact} and Lemma~\ref{lem:upper_fgh}, we have 
   %  \begin{align}\label{eq:f_sum}
   %        \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{(F_{ib}^{(t)})^2}{ cm^2 \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \leq \frac{C_0 \bar C}{c m^2 \tilde C^2} L^{(t)},
   %  \end{align}
   % for a positive universal constant $C$ and 
   %  \begin{align}
   %       \frac{1}{p} \sum_{i \in [p]}\sum_{b \in [r]/z(i)} \ind \offf{ z^{(t+1)}(i) = b} \frac{\theta(i)(G_{ib}^{(t)})^2}{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}  &\leq \frac{1}{512} \frac{1}{p} \sum_{i \in [p]}\theta(i) \sum_{b \in [r]/z(i)}   \ind \offf{ z^{(t+1)}(i) = b}  (\Delta_{\min}^2 + L^{(t)}) \\
   %       & \leq \frac{1}{512} (L^{(t+1)} + L^{(t)}), \label{eq:g_sum}
   %  \end{align}
   %  where the last inequality follows from the definition of $L^{(t)}$ and the constraint of $\mtheta$ in parameter space~\eqref{eq:family}. 
   %  %Choosing the constants $C, \bar C, C_2$ in~\eqref{eq:zeta}~\eqref{eq:f_sum}~\eqref{eq:g_sum} carefully \fixme{Miaoyan}{be specific. Add one sentence to explain}, 
   % For $\tilde C$ also satisfies 
   %  \begin{equation}\label{eq:tilde_c3}
   %       \frac{C_0 \bar C}{c m^2 \tilde C^2} \leq \frac{1}{512},
   %  \end{equation}
   %  we have 
   %  \begin{equation}\label{eq:zeta_upper}
   %      \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}^{(t)} \leq  \frac{1}{8} L^{(t+1)}  + \frac{1}{4} L^{(t)}.
   %  \end{equation}
   %  %for some small \fixme{Miaoyan}{In what sense "small"? Add one sentence.} positive constants $C_4, C_5$. 
   %  Plugging the inequality~\eqref{eq:zeta_upper} into the decomposition~\eqref{eq:decomp2}, we obtain the contraction inequality 
   %  \begin{equation}\label{eq:decomp_t}
   %        L^{(t+1)} \leq \frac{3}{2} \xi  + \frac{1}{2} L^{(t)},
   %  \end{equation}
   %   where $\frac{1}{2}$ is the contraction parameter.

   %   Therefore, with $\tilde C$ satisfying inequalities~\eqref{eq:tilde_c3}, \eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}, we obtain the conclusion in Theorem~\ref{thm:refinement} via inequality~\eqref{eq:decomp_t} combining the inequality~\eqref{eq:cond_oracle} in Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.

   %   {

   %   \textbf{We also have the contraction inequality for MLE.}

   %   Following the same derivation of~\eqref{eq:decomp_t} with the upper bound of $\hat F_{ib}, \hat G_{ib}, \hat H_{ib}$ in Lemma~\ref{lem:intermediate}, we also have
   %  \begin{equation}\label{eq:decomp_t_mle}
   %        L(\hat z) \leq \frac{3}{2} \xi'  + \frac{1}{2} L(\hat z),
   %  \end{equation}
   %  which indicates the conclusion $\ell(\hat z,z) \lesssim \Delta_{\min}^2  \exp \of{ - \frac{p^{K-1}}{r^{K-1}} \Delta_{\min}^2  }$.
   %   }

    {\bf Step 4 (Verification of Condition~\ref{cond:origin}).} Last, we verify the Condition~\ref{cond:origin} under high probability to finish the proof. Note that the inequalities~\eqref{eq:cond1}, \eqref{eq:cond2}, and \eqref{eq:cond3} describe the property of the sub-Gaussian noise tensor $\tE$, and the readers can find the proof directly in \cite[Step 5, Proof of Theorem 2]{han2020exact}. The initial condition~\eqref{eq:cond_intial} for MLE is satisfied by Lemma~\ref{lem:poly_mle_degree}. Here, we include only the verification of inequalities~\eqref{eq:cond_oracle} and \eqref{eq:cond_intial} for algorithm estimators.

    Now, we verify the oracle loss condition~\eqref{eq:cond_oracle}. Recall the definition of $\xi$,
    \small
    \begin{align}
        \xi & = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \Big\{\ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s} \\
         & \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \leq - \frac{ \theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \Big\} \\
         & \quad \quad \quad \quad \quad \quad \quad \quad \cdot \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2.
    \end{align}
    \normalsize
    % \begin{align}
    %     \xi = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \offf{  \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2,
    % \end{align}

    Let $e_i = \mE_{i:} \mV$ denote the aggregated noise vector for all $i \in [p]$, and $e_i$'s are independent zero-mean sub-Gaussian vector in $\bbR^{r^{K-1}}$. The entries in $e_i$ are independent zero-mean sub-Gaussian variables with sub-Gaussian norm upper bounded by $m_1\sqrt{r^{K-1}/p^{K-1}}$ with some positive constant $m_1$. We have the probability inequality
    \begin{align}
        &\bbP \of{ \ang{ e_i, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s } \leq - \frac{ \theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2}\\
        &\quad \quad \leq P_1 + P_2 + P_3,
    \end{align}
    where 
    \small
    \begin{align}
        P_1 &= \bbP \of{ \ang{e_i, [  \mS_{z(i):} ]^s - [ \mS_{b:} ]^s} \leq -\frac{\theta(i)m}{8}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 }, \\
        P_2 &= \bbP \of{ \ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s} \leq -\frac{\theta(i)m}{16}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 },\\
        P_3 &= \bbP \of{ \ang{e_i, [  \mS_{b:} ]^s - [\tilde  \mS_{b:} ]^s} \leq -\frac{\theta(i)m}{16}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 }.
    \end{align}
    \normalsize
    For $P_1$, notice that the inner product $\ang{e_j,\mS_{z(j):}^s - \mS_{b:}^s} $ is a sub-Gaussian variable with sub-Gaussian norm bounded by $m_2 \sqrt{ r^{K-1}/ p^{K-1} } \onormSize{}{ \mS_{z(i):}^s -  \mS_{b:}^s }$ with some positive constant $m_2$. Then, by Chernoff bound, we have  
    \begin{equation}\label{eq:p1_2}
         P_1 \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(j):}  ]^s - [ \mS_{b:}  ]^s  }^2 }.
    \end{equation}

    For $P_2$ and $P_3$, we only need to derive the upper bound of $P_2$ due to the symmetry. By the law of total probability, we have 
    \begin{equation}\label{eq:p2}
        P_2 \leq P_{21} + P_{22},
    \end{equation}
    where with some positive constant $t>0$,
    \begin{align}
        P_{21}& =  \bbP \of{ t \leq   \onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s }},\\
        P_{22} &= \bbP \Bigg(  \ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s} \leq -\frac{\theta(i)m}{16} \\
        &\quad \cdot \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2   \bigg| \onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s } < t  \Bigg).
    \end{align}

    For $P_{21}$, note that the term $\mW^T_{: z(i)} \mE \mV = \frac{ \sum_{ j \neq i, j \in [p]} \ind \{ z(j) = z(i) \} e_j }{\sum_{j \in [p]} \ind \{ z(j) = z(i) \} }$  is a sub-Gaussian vector with sub-Gaussian norm bounded by $m_3 \sqrt{r^K/ p^K}$ with some positive constant $m_3$. This implies
        \begin{align}
        P_{21} & \leq \bbP\of{ t \onormSize{}{\mS_{z(i):} } \leq \onormSize{}{ \tilde  \mS_{z(i):}  -  \mS_{z(i):} }  }\\
        & \leq \bbP\of{c_3 t \leq  \onormSize{}{ \mW^T_{: z(i)} \mE \mV  } } \\
        &\lesssim \exp\of{ - \frac{p^K t^2}{r^K} }, \label{eq:p21}
    \end{align}
    where the first inequality follows from the basic inequality in Lemma~\ref{lem:norm_diff}, the second inequality follows from the assumption that $ \min_{a \in [r]} \onormSize{}{\mS_{z(i):}} \geq c_3>0$ in~\eqref{eq:family}, and the last inequality follows from the Bernstein inequality.

    For $P_{22}$, the inner product $ \ang{e_i, [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s}$ is also a sub-Gaussian variable with sub-Gaussian norm $ m_4 \sqrt{ r^{K-1}/ p^{K-1} } t$, conditioned on $\onormSize{}{ [ \tilde  \mS_{z(i):} ]^s - [ \mS_{z(i):} ]^s } < t$ with some positive constant $m_4$. Then, by Chernoff bound, we have 
    \begin{equation}\label{eq:p22}
        P_{22} \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1} t^2}   \onormSize{}{ [ \mS_{z(j):}  ]^s - [ \mS_{b:}  ]^s  }^4 }. 
    \end{equation}
    We take $t =  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }$ in $P_{21}$ and $P_{22}$, and plug the inequalities \eqref{eq:p21} and \eqref{eq:p22} into to the upper bound for $P_2$ in~\eqref{eq:p2}. We obtain that 
    \begin{equation}\label{eq:p2_2}
        P_2 \lesssim \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 }.
    \end{equation}
    Combining the upper bounds~\eqref{eq:p1_2} and \eqref{eq:p2_2} gives 
    \begin{align}
        &\bbP \of{ \ang{ e_i, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2} \\
        & \lesssim  \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 }.\label{eq:xi_p}
    \end{align}

    Hence, we have 
    \begin{align}
        \bbE \xi & =  \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \bbP \Bigg\{  \ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s }  \\
        & \quad \quad  \leq - \frac{\theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \Bigg\} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2 \\
        &\lesssim \frac{1}{p} \sum_{i \in [p]}  \theta(i) \max_{i \in [p], b \in [r]}  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
        & \quad \quad \quad \quad  \quad \quad \quad \quad \cdot \exp \of{  - \frac{p^{K-1}}{r^{K-1}}   \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 } \\
        & \leq \exp \of{  - M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }, \label{eq:constant_M}
    \end{align}
    where $M$ is a positive constant, the first inequality follows from the constraint that $\sum_{i \in [p]} \theta(i) = p$, and the last inequality follows from~\eqref{eq:xi_p}.

    By Markov's inequality, we have 
    \begin{align}
        & \bbP\of{ \xi \lesssim \bbE \xi + \exp \of{  - \frac{M p^{K-1}}{2 r^{K-1}}   \Delta_{\min}^2 } } \\
        & \quad \quad  \quad \quad \quad \quad \geq 1 -  C \exp \of{  - \frac{M p^{K-1}}{2 r^{K-1}}   \Delta_{\min}^2 },
    \end{align}
    and thus the condition~\eqref{eq:cond_oracle} holds with probability at least $1 -  C \exp \of{  - \frac{M p^{K-1}}{2 r^{K-1}}   \Delta_{\min}^2 }$ for some constant $C > 0$.

    {

    \textbf{The initialization condition for MLE also holds.}

    For $\xi'$, notice that $\ang{\mE_i, \mA^s_{a:} - \mA^s_{b:} }$ is a sub-Gaussian vector with variance bounded by $\onormSize{}{\mA^s_{a:} - \mA^s_{b:}}^2$ and 
    \begin{align}
        \bbP \of{ t \leq \onormSize{}{ [\bar \mA_{a:}]^s - \mA^s_{a:} }} & \leq \of{ t \leq \onormSize{}{ [\mP_{:a}^T \mY \mQ]^s - [\mP_{:a}^T \mX \mQ]^s }} \\
        &\leq \bbP( t \min_{a \in [r]}\onormSize{}{\mS_{a:}} \leq \onormSize{}{\mP_{:a}^T \mE \mQ} ) \\
        &\lesssim \exp \of{ - \frac{p^{K}t^2}{r^{K}}},
    \end{align}
   where the first inequality follows from the property in later inequality \eqref{eq:angle_aq}. We also have 
   \begin{equation}
        \xi' \lesssim  \of{  -  \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }.
   \end{equation}

    }

    Finally, we verify the bounded loss condition~\eqref{eq:cond_intial} for algorithm estimator by induction.  With output $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{2} and the assumption SNR $\geq \tilde C p^{-K/2} \log p$, by Theorem~\ref{thm:initial}, we have 
    \begin{equation}
        L^{(0)} \leq \frac{\bar C \Delta_{\min}^2}{\tilde C r \log p},\quad \text{    when $p$ is large enough}.
    \end{equation}
Therefore, the condition~\eqref{eq:cond_intial} holds for $t = 0$. Assume the condition~\eqref{eq:cond_intial} also holds for all $t \leq t_0$. Then, by the decomposition~\eqref{eq:decomp_t}, we have 
    \begin{align}
         L^{(t_0+1)} &\leq \frac{3}{2} \xi + \frac{1}{2} L^{(t_0)} \\
         & \leq \exp \of{  - M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } + \frac{\Delta_{\min}^2}{r \log p} \\
         & \leq \frac{\bar C}{\tilde C}\frac{\Delta_{\min}^2}{r\log p} ,
    \end{align}
    where the second inequality follows from the condition~\eqref{eq:cond_oracle} and the last inequality follows from the assumption that $\Delta_{\min}^2 \gtrsim p^{-K/2} \log p$. Thus, the condition~\eqref{eq:cond_intial} holds for $t_0 + 1$, and the condition~\eqref{eq:cond_intial} is proved by induction.
\end{proof}

{\bf Useful Lemmas for the Proof of Theorem~\ref{thm:refinement}}

\begin{lem}[Singular-value property of membership matrices]\label{lem:membership} Under the setup of Theorem~\ref{thm:refinement}, suppose the condition~\eqref{eq:cond_intial} holds. Then, for all $a \in [r]$, we have $|\of{z^{(t)}}^{-1}(a)| \asymp p/r$. Moreover, we have 
\begin{align}
     &\lambda_{r}(\mM) \asymp \onorm{\mM}_{\sigma} \asymp \sqrt{p/r}, \  \lambda_{r}(\mW) \asymp \onorm{\mW}_{\sigma} \asymp \sqrt{r/p},\\
     &\lambda_{r}(\mP) \asymp \onorm{\mP}_{\sigma}    \asymp  \min_{a \in [r] } \onormSize{}{\mtheta_{z^{-1}(a)}}^{-1}  \lesssim  \sqrt{r/p}. \label{eq:membership1}
\end{align}
The inequalities~\eqref{eq:membership1} also hold by replacing $\mM$ and $\mW$ to $\mM^{(t)}$ and $\mW^{(t)}$ respectively. 
Further, we have 
\begin{equation}\label{eq:WWsingular}
   \lambda_{r}(\mW\mW^T) \asymp \onorm{\mW\mW^T}_{\sigma} \asymp r/p,
\end{equation}
which is also true for $\mW^{(t)}\mW^{(t),T}$.
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:membership}] The proof for the inequality~\eqref{eq:membership1} for $\mM, \mW$ can be found in \cite[Proof of Lemma 4]{han2020exact}. The inequalities for $\mP$ follows the same derivation with balance assumption on $\mtheta$ and $\min_{i \in [p]}\theta(i) \geq c$.

For inequality~\eqref{eq:WWsingular}, note that for all $k \in [r]$,
\begin{align}
    \lambda_k(\mW\mW^T) &= \sqrt{\text{eigen}_k (\mW \mW^T\mW\mW^T)} \\
    &\asymp \sqrt{\frac{r}{p} \text{eigen}_k (\mW \mW^T)} \\
    &= \sqrt{\frac{r}{p} \lambda^2_k(\mW)} \asymp \frac{r}{p},
\end{align}
where $\text{eigen}_k(\mA)$ denotes the $k$-th largest eigenvalue of the square matrix $\mA$, the first inequality follows the fact that $\mW^T\mW$ is a diagonal matrix with elements of order $r/p$, and the second equation follows from the definition of singular value.
\end{proof}

\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement}  {with fixed $r \geq 2$},  assume the the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ is large enough to satisfy the inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}. {As $p \rightarrow \infty$}, we have 
\small
\begin{align}
    &\max_{i \in [p]} \max_{b \neq z(i)} \frac{\of{F_{ib}^{(t)}}^2}{ \onormSize{}{ [  \mS_{z(i):} ]^s  -  [ \mS_{b:} ]^s }^2 } \\
    &\lesssim  \frac{r L^{(t)}}{ \Delta_{\min}^2} \onormSize{}{\mE_{i:} \mV}^2 + \of{1 +  \frac{ r L^{(t)}}{ \Delta_{\min}^2}} \onormSize{}{\mE_{i:} (\mV - \mV^{(t)})}^2 ,\label{eq:fib}
\end{align}
    \begin{equation}\label{eq:gib}
        \max_{i \in [p]} \max_{b \neq z(i)} \frac{ \of{  G_{ib}^{(t)} }^2  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2}  \leq  \frac{1}{512} \of{ \Delta_{\min}^2 +L^{(t)}},
    \end{equation}
    \begin{equation}\label{eq:hib}
        \max_{i \in [p]} \max_{b \neq z(i)}  \frac{ \aabs{  H_{ib}^{(t)} }  }{ \onormSize{}{ [\mS_{z(i):} ]^s - [\mS_{b:}]^s }^2} \leq \frac{1}{4}.
    \end{equation}
    \normalsize

    Similarly, when the SNR $\geq \tilde C p^{-(K-1)}\log p$ with large constant $\tilde C$, we have 
    \small
     \begin{equation}\label{eq:fib_cor}
        \max_{i \in [p]} \max_{b \neq z(i)} \frac{\of{\hat F_{ib}}^2}{ \onormSize{}{ [  \mA_{z(i):} ]^s  -  [ \mA_{b:} ]^s }^2 } \lesssim p^{K-1} \frac{r L(\hat z)}{ \Delta_{\min}^2}
        %+ \of{1 +  \frac{ r L(\hat z)}{ \Delta_{\min}^2}} \onormSize{}{\mE_{i:} (\mV - \mV^{(t)})}^2 ,
    \end{equation}
    \begin{equation}\label{eq:gib_cor}
        \max_{i \in [p]} \max_{b \neq z(i)} \frac{ \of{  \hat G_{ib} }^2  }{ \onormSize{}{ [\mA_{z(i):} ]^s - [\mA_{b:}]^s }^2}  \leq  \frac{1}{512} \of{ \Delta_{\min}^2 +L(\hat z)},
    \end{equation}
    \begin{equation}\label{eq:hib_cor}
        \max_{i \in [p]} \max_{b \neq z(i)}  \frac{ \aabs{  \hat H_{ib} }  }{ \onormSize{}{ [\mA_{z(i):} ]^s - [\mA_{b:}]^s }^2} \leq \frac{1}{4}.
    \end{equation}
    \end{lem}
    \normalsize

\begin{proof}[Proof of Lemma~\ref{lem:upper_fgh}]
We prove the the first three inequalities in Lemma~\ref{lem:upper_fgh} separately.
\begin{enumerate}[wide]
    \item Upper bound for $F_{ib}^{(t)}$, i.e., inequality \eqref{eq:fib}. Recall the definition of $F_{ib}^{(t)}$,
    \small
    \begin{align}
         F_{ib}^{(t)} &= 2 \ang{\mE_{i:} \mV^{(t)}, \of{ [ \tilde \mS_{z(i):} ]^s  - [\mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  - [ \mS_{b:}^{(t)} ]^s  }  }  \\
         &+ 2 \ang{ \mE_{i:}(\mV - \mV^{(t)} ), [ \tilde \mS_{z(i):} ]^s - [ \tilde \mS_{b:} ]^s  }.
    \end{align}
    \normalsize
    By Cauchy-Schwartz inequality, we have 
    \small
    \begin{align}
        &\of{F_{ib}^{(t)}}^2 \\
        & \leq 8 \of{\ang{\mE_{i:} \mV^{(t)}, \of{ [  \tilde \mS_{z(i):} ]^s  -  [  \mS_{z(i):}^{(t)} ]^s  }  - \of{  [  \tilde \mS_{b:} ]^s  -  [  \mS_{b:}^{(t)} ]^s  }  }  }^2 \\
         & \quad \quad \quad \quad \quad + 8\of{ \ang{ \mE_{i:} (\mV - \mV^{(t)} ), [  \tilde \mS_{z(i):} ]^s - [ \tilde  \mS_{b:} ]^s  }}^2 \\
         & \leq 8 \of{ \onormSize{}{\mE_{i:} \mV}^2 + \onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2 } \max_{a \in [r]^s} \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  } \\
          & \quad \quad \quad \quad \quad +\onormSize{}{\mE_{i:} ( \mV - \mV^{(t)} )}^2  \onormSize{}{  [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }. \label{eq:fib_decomp}
    \end{align}
    \normalsize
    Note that for all $a \in [r]$,
    \begin{align}
        \onormSize{}{  [  \tilde \mS_{a:} ]^s - [   \mS_{a:}^{(t)} ]^s  }^2 &= \onormSize{}{ [ \mW_{:a}^T \mY \mV ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }^2 \\
        &\leq 2 \onormSize{}{ [ \mW_{:a}^T \mY \mV ]^s  - [ \mW_{:a}^{(t),T} \mY \mV ]^s  }^2 \\
        & \quad + 2 \onormSize{}{[ \mW_{:a}^{(t),T} \mY \mV ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)} ]^s }^2\\
        & \lesssim \frac{ r^2 (L^{(t)})^2}{ \Delta_{\min}^2} + \frac{ r r^{2K} + p r^{K+2}}{p^K} \frac{L^{(t)}}{\Delta_{\min}^2}\\
        & \lesssim  r L^{(t)}+  \frac{ r r^{2K} + p r^{K+2}}{p^K} \frac{L^{(t)}}{\Delta_{\min}^2}\\
        & \lesssim  r L^{(t)},\label{eq:fib_1}
    \end{align}
    where the second inequality follows from the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4} in Lemma~\ref{lem:intermediate}, the third inequality follows from the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}, and the last inequality follows from the assumption that $\Delta_{\min}^2 \geq \tilde C p^{-K/2}\log p$.

    Note that 
    \small
    \begin{align}
        &\onormSize{}{  [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s  }^2 \\
        &=  \onormSize{}{  [  \tilde \mS_{z(i):} ]^s -[ \mS_{z(i):} ]^s +  [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s + [ \mS_{b:} ]^s -[  \tilde \mS_{b:} ]^s  }^2 \\
        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2 + \max_{a \in [r]} \onormSize{}{ [ \mS_{a:} ]^s -[  \tilde \mS_{a:} ]^s }^2 \\
        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2  + \max_{a \in [r]} \frac{1}{\onormSize{}{\mS_{a:} }^2} \onormSize{}{ \mW_{:a}^T \mE \mV}^2\\
        & \lesssim \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{b:} ]^s }^2, \label{eq:fib_2}
    \end{align}
    \normalsize
    where the second inequality follows from Lemma~\ref{lem:norm_diff}, and the last inequality follows from the assumptions on $\onormSize{}{\mS_{a:}}$ in the parameter space~\eqref{eq:family}, the inequality~\eqref{eq:cond1} in Condition~\ref{cond:origin} and the assumption $\Delta_{\min}^2 \gtrsim p^{-K/2}\log p$.

    Therefore, we finish the proof of inequality~\eqref{eq:fib} by plugging the inequalities~\eqref{eq:fib_1} and \eqref{eq:fib_2} into the upper bound~\eqref{eq:fib_decomp}.

    \item Upper bound for $G_{ib}^{(t)}$, i.e., inequality~\eqref{eq:gib}. By definition of $G_{ib}^{(t)}$, we rearrange terms and obtain
    \begin{align}
        G_{ib}^{(t)} &=   \Big( \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 \\
        &\quad \quad \quad \quad \quad-  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2\Big)   \\
        & \quad -   \Big( \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mS_{b:}^{(t)}  ]^s}^2 \\
        & \quad \quad \quad \quad \quad -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 \Big) \\
        &= 2 \Big\langle  [\mX_{i:} \mV^{(t)}]^s , \of{[  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s } \\
        &\quad \quad \quad \quad \quad  \quad \quad \quad \quad  - \of{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s } \Big\rangle \\
      &= G_1 + G_2 - G_3, \label{eq:gib_decomp}
    \end{align}
    where 
    \small
    \begin{align}
        G_1 &= \onormSize{}{ [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s}^2 - \onormSize{}{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s }^2,\\
        G_2 &= 2 \ang{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s,   [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s},\\
        G_3 & = 2 \ang{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s,  [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mS_{b:}^{(t)}  ]^s}.
    \end{align}
    \normalsize

    For $G_1$, we have 
    \begin{align}
        |G_1|^2 &\leq \Big| \onormSize{}{ [ \mW_{:z(i)}^T \mY \mV^{(t)} ]^s - [ \mS_{z(i):}^{(t)}  ]^s}^2 \\
        & \quad \quad \quad \quad \quad  \quad \quad \quad - \onormSize{}{ [ \mW_{:b}^T \mY \mV^{(t)} ]^s -  [ \mS_{b:}^{(t)}  ]^s }^2 \Big|^2 \\
        & \leq \max_{a \in [r]} \onormSize{}{ [ \mW_{:a}^T \mY \mV^{(t)} ]^s - [ \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^4\\
        & \leq C^4 \frac{r^4}{\Delta_{\min}^4} (L^{(t)})^4 +  \frac{r^2 r^{4K} + p^2r^{2K+4}}{p^{2K}} \frac{(L^{(t)})^2}{\Delta_{\min}^4} \\
        & \leq C^4 \frac{\bar C}{\tilde C^3} \of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}}, \label{eq:g1}
    \end{align}
    where the third inequality follows from the inequality~\eqref{eq:inter5} in Lemma~\ref{lem:intermediate} and the last inequality follows from the assumption that $\Delta_{\min}^2 \geq \tilde C p^{-K/2}\log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.

    For $G_2$, noticing that $[\mX_{i:} \mV^{(t)}]^s = [ \mW_{z(i):}^T\mX \mV^{(t)}]^s$, we have 
    \begin{align}
        |G_2|^2 &\leq 2\onormSize{}{ [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 \\
        & \quad \quad \quad \quad \quad  \quad \quad \quad \cdot \onormSize{}{ [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s -  [  \mS_{z(i):}^{(t)}  ]^s }^2 \\
        & \leq \frac{2}{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2} \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 \\
        & \quad \quad \quad \quad \quad  \quad \cdot \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
        & \leq C' \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} \\
        &  \quad \quad \quad  \quad  \cdot \of{ \frac{r^2}{\Delta_{\min}^2} (L^{(t)})^2 + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2} }\\
        & \leq \frac{C'}{\tilde C^2} \Delta_{\min}^2 L^{(t)},\label{eq:g2}
    \end{align}
    where $C'$ is a positive universal constant, the second inequality follows from Lemma~\ref{lem:norm_diff}, the third inequality follows from the inequality~\eqref{eq:cond2} in Condition~\ref{cond:origin}, the inequalities~\eqref{eq:inter5} and \eqref{eq:j11_wxvt} in the proof of Lemma~\ref{lem:intermediate},  and the last inequality follows from the assumption $\Delta_{\min}^2 \geq \tilde C p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.

    For $G_3$, note that by triangle inequality
    \begin{align}
        &\onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2 \\& \leq \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + 2\max_{a \in [r]}\onormSize{}{[\mW_{:a}^T\mX \mV^{(t)}]^s - [\mW_{:a}^T\mX \mV]^s }^2\\
        & \leq \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + C \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2}, \label{eq:g3xv_wxv}
    \end{align}
    where the last inequality follows from the inequality~\eqref{eq:j1} in the proof of Lemma~\ref{lem:intermediate} and $C$ is a positive constant.
   Then we have 
   \small
    \begin{align}
        |G_3|^2 & \leq 2 \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2  \\
        &  \quad \quad \quad \quad \quad  \quad \cdot \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2 \\
        & \leq 2 \Big( \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  \\
        & \quad \quad \quad \quad \quad  \quad  +  \onormSize{}{ [  \mW_{:b}^T \mY \mV^{(t)} ]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s}^2  \Big) \\
        & \quad \cdot \max_{a \in [r]} \onorm{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
        & \leq C^2 \of{\onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 + C \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} } \\
        &  \quad \cdot \of{ \frac{r^2(L^{(t)})^2}{ \Delta_{\min}^2} + \frac{r r^{2K} + pr^{K+2}}{p^{K}} \frac{L^{(t)}}{\Delta_{\min}^2}}+\frac{C'}{\tilde C^2} \Delta_{\min}^2 L^{(t)} \\
        & \leq \frac{C^2 \bar C^2}{\tilde C} \onormSize{}{\mS_{z(i):}^s - \mS_{b:}^s}^2 (\Delta_{\min}^2 + L^{(t)}) \\
        & \quad +  \frac{C^3 C' \bar C^2}{\tilde C^2 }\of{\Delta_{\min}^4 +  \Delta_{\min}^2 L^{(t)}},\label{eq:g3}
    \end{align}
    \normalsize
    where the third inequality follows from the same procedure to derive~\eqref{eq:g1} and \eqref{eq:g2}, and the last inequality follows from the assumption $\Delta_{\min}^2 \geq \tilde C p^{-K/2} \log p$ and inequality~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.

    Choose the $\tilde C$ such that
    \begin{equation}\label{eq:tilde_c1}
        3 \of{  C^4 \frac{\bar C}{\tilde C^3}  + \frac{C'}{\tilde C^2} + \frac{C^2 \bar C^2}{\tilde C} + \frac{C^3 C' \bar C^2}{\tilde C^2 }} \leq \frac{1}{512}.
    \end{equation}
    Then,  we finish the proof of inequality~\eqref{eq:gib} by plugging the inequalities~\eqref{eq:g1}, \eqref{eq:g2}, and \eqref{eq:g3} into the upper bound~\eqref{eq:gib_decomp}.

    \item Upper bound for $H_{ib}^{(t)}$, i.e., the inequality~\eqref{eq:hib}. By definition of $H_{ib}$, we rearrange terms and obtain
    \small
    \begin{align}
        H_{ib} &= \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 \\
        & - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s -  [  \mW_{:b}^T \mY \mV^{(t)} ]^s}^2 + \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \\
        & =  \onormSize{}{ [ \mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 \\
        &  + \of{ \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }  } \\
        &  - \Big(\onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mY \mV^{(t)} ]^s } \\
        & \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad -  \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }   \Big) \\
        & = H_1 + H_2 + H_3,
    \end{align}
    \normalsize
    where 
    \small
    \begin{align}
        H_1 &= \onormSize{}{ [ \mX_{i:} \mV^{(t)}]^s -  [  \mW_{:z(i)}^T \mY \mV^{(t)} ]^s}^2 \\
        & \quad \quad \quad \quad \quad  \quad \quad \quad - \onormSize{}{ [ \mW_{:b}^T \mX \mV^{(t)}]^s -  [ \mW_{:b}^T \mY \mV^{(t)}]^s}^2, \\
        H_2 &=  \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 - \onormSize{}{ [\mX_{i:} \mV^{(t)}]^s-  [  \mW_{:b}^T \mX \mV^{(t)} ]^s }^2  , \\
        H_3 & = 2 \Big\langle [\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s, \\
        &  \quad \quad \quad \quad \quad  \quad \quad \quad [  \mW_{:b}^T \mY \mV^{(t)} ]^s -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s \Big\rangle.
    \end{align}
    \normalsize
    For $H_1$, we have 
    \begin{align}
        |H_1| &\leq  \frac{4 \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 }{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2}\\
        &\leq  \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} \\
        &\leq \tilde C^{-2} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2\label{eq:h1},
    \end{align}
    following the derivation of $G_2$ in inequality~\eqref{eq:g2} and the assumption that $\Delta_{\min}^2 \geq \tilde C p^{-K/2} \log p$.

    For $H_2$, by the inequality~\eqref{eq:g3xv_wxv}, we have 
    \begin{align}
        |H_2| &\lesssim 2\max_{a \in [r]}\onormSize{}{[\mW_{:a}^T\mX \mV^{(t)}]^s - [\mW_{:a}^T\mX \mV]^s }^2 \\
        &\lesssim \frac{r^2 (L^{(t)})^2}{\Delta_{\min}^2} \\
        &\leq C \frac{\bar C^2}{ \tilde C^2} \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2, \label{eq:h2}
    \end{align}
    where the last inequality follows from the condition~\eqref{eq:cond_intial} in Condition~\ref{cond:origin}.

      For $H_3$,  by Cauchy-Schwartz inequality, we have 
      \begin{align}
           |H_3| & \lesssim \onormSize{}{[\mX_{i:} \mV^{(t)}]^s  -  [  \mW_{:b}^T \mX \mV^{(t)} ]^s} |H_1|^{1/2} \\
           &\leq 2 \tilde C^{-1} \onormSize{}{ [ \mS_{z(i):} ]^s - [ \mS_{a:}]^s }^2,\label{eq:h3}
      \end{align}
    following the inequalities~\eqref{eq:g3xv_wxv} and \eqref{eq:h1}.

    Choose $\tilde C$ such that 
    \begin{equation}\label{eq:tilde_c2}
        \tilde C^{-2} + C \frac{\bar C^2}{ \tilde C^2} + \tilde C^{-1} \leq \frac{1}{4}.
    \end{equation}
     Therefore, we finish the proof of inequality~\eqref{eq:hib} combining inequalities~\eqref{eq:h1}, \eqref{eq:h2}, and \eqref{eq:h3}.

\end{enumerate}
{

\textbf{Next, we show the upper bounds for $\hat F_{ib}, \hat G_{ib}$ and $\hat H_{ib}$.}

By Lemma~\ref{lem:angle_gap_x}, we have 
\begin{equation}
    \onormSize{}{\mS_{a:}^s - \mS_{b:}^s}  = (1 + o(1)) \onormSize{}{\mA_{a:}^s - \mA_{b:}^s}.
\end{equation}
Also, notice that the matrix product of $ \mB^T$ corresponding to the padding operation in Lemma~\ref{lem:pad}, and the padding weights are balanced such that $\onormSize{}{\mv \mB}  = (1 + o(1)) \max_{a} \onormSize{}{\mtheta_{z^{-1}(a)}}^{(K-1)/2} \onormSize{}{\mv} $ for all $\mv \in \bbR^{r(K-1)}$. For two vectors $\mv_1, \mv_2 \in \bbR^{r^{K-1}}$, we have 
\begin{equation}\label{eq:b_angle}
    \onormSize{}{\mv_1^s - \mv_2^s} = (1 + o(1)) \onormSize{}{[\mv_1 \mB^T]^s -  [\mv_2 \mB^T]^s  }.
\end{equation}
The equation~\eqref{eq:b_angle} also holds for $\hat \mB^T$.

Note that for all $i \in [p]$ we have 
\small
\begin{align}
    \onormSize{}{\mA_{i:} \hat \mQ} &= \onormSize{}{\mS_{z(i:)} \mB^T \hat \mQ }\\
    &=  \onormSize{}{\mS_{z(i:)} \hat \mD^{\otimes (K-1)} }\\
    &=  (1 + o(1)) \onormSize{}{\mS_{z(i:)}} \\
    &=  (1 + o(1)) \max_{a} \onormSize{}{\mtheta_{z^{-1}(a)}}^{-(K-1)/2} \onormSize{}{\mA_{i:}},  \label{eq:aq}
\end{align}
\normalsize
where the third inequality follows from the singular property of MLE confusion matrix~\eqref{eq:mle_confusion} and the last inequality follows from the fact that $\mA_i = \mS_{z(i:)} \mB^T$ and Lemma~\ref{lem:membership}. Above equation indicates that $\mA_{i:}$ is the span space of the singular values as $p \rightarrow \infty$. Also, notice that the row space of $\mP_{:a}^T \mY \hat \mQ  \hat \mB^T$ is equal to the column space of $\hat \mQ$, and $\mA_{i:} \neq \mP_{:a}^T \mY \hat \mQ  \hat \mB^T$ in noisy case.

Hence, for all $a \in [r]$, we have 
\begin{align}
    &\onormSize{}{ [\mX_i \hat \mQ]^s - [\mP_{:a}^T \mY \hat \mQ ]^s } \\
    &= \onorm{ \frac{\mA_{z(i:)} \hat \mQ  }{\onormSize{}{\mA_{z(i:)} \hat \mQ }} - \frac{\mP_{:a}^T \mY \hat \mQ  }{ \onormSize{}{\mP_{:a}^T \mY \hat \mQ} } } \\
    &= (1+o(1))  \onorm{ \frac{\mA_{z(i:)} }{\onormSize{}{\mA_{z(i:)} }} - \frac{\mP_{:a}^T \mY \hat \mQ \hat \mB^T  }{ \onormSize{}{\mP_{:a}^T \mY \hat \mQ \hat \mB^T} } } \\
    &= (1+o(1)) \onormSize{}{ [\mX_i]^s - [\mP_{:a}^T \mY \hat \mQ \hat \mB^T ]^s } \label{eq:angle_aq}
\end{align}
where the second equation follows from~\eqref{eq:aq}, $\onormSize{}{\mP_{:a}^T \mY \hat \mQ \hat \mB^T} = (1 + o(1)) \max_{a} \onormSize{}{\mtheta_{z^{-1}(a)}}^{(K-1)/2} \onormSize{}{\mP_{:a}^T \mY \hat \mQ } $, and singular property of $\hat \mB^T$. Similar result holds after replacing $\mP_{:a}^T \mY \hat \mQ$ by  $\mP_{:a}^T \mY\mQ$ or $\mP_{:a}^T \mY \hat \mQ$.

We are now ready to show the upper bounds for $\hat F_{ib},\hat G_{ib}$ and $\hat H_{ib}$.

 For $\hat F_{ib}$, we have 
\begin{align}
    (\hat F_{ib})^2 &\leq \onormSize{}{\mE_{i:}}^2 \onormSize{}{ [\bar  \mA_{a:}]^s - [\hat \mA_{a:}]^s }^2\\
    &\leq \onormSize{}{\mE_{i:}}^2  \Big[ \onormSize{}{ [ \bar \mS_{a:} \mB^T]^s - [\bar \mS_{a:} \hat \mB^T]^s } \\
    & \quad \quad \quad \quad \quad  \quad  \quad \quad \quad \quad +  \onormSize{}{ [ \bar \mS_{a:} \hat \mB^T]^s -  [\hat \mS_{a:} \hat \mB^T]^s}  \Big]^2 \\
    &\lesssim \onormSize{}{\mE_{i:}}^2  \off{ \onormSize{}{ [ \bar \mS_{a:} \mB^T \hat \mQ]^s - [\bar \mS_{a:}]^s } +  \onormSize{}{ [ \bar \mS_{a:} ]^s -  [ \hat \mS_{a:} ]^s}  }^2.
\end{align}
Following similar derivations in inequalities~\eqref{eq:fib_1}, \eqref{eq:fib_2}, and the upper bound for $J_{1}$, respectively, we have 
\begin{equation}
    \onormSize{}{ [ \bar \mS_{a:} ]^s -  [ \hat \mS_{a:} ]^s}  \lesssim r L(\hat z), \quad  \onormSize{}{ [ \bar \mS_{a:} ]^s -  [ \bar \mS_{b:} ]^s} \lesssim \onormSize{}{\mS_{a:}^s - \mS_{b:}^s}^2, 
\end{equation}
and 
\begin{align}
    \onormSize{}{ [ \bar \mS_{a:} \mB^T \hat \mQ]^s - [\bar \mS_{a:}]^s } \lesssim L(\hat z).
\end{align}
We then obtain the upper bound for $\hat F_{ib}$ by noticing that $\onormSize{}{\mE_i}^2 \lesssim p^{K-1}$.

For $\hat G_{ib}$ and $\hat H_{ib}$, by the property~\eqref{eq:angle_aq}, we have 
    \begin{align}
        &(1+o(1))\hat G_{ib}\\
        & = \of{ \onormSize{}{[\mX_{i:}\hat \mQ]^s - [\hat \mS_{a:}]^s}_F^2 - \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:a}^T \mY \hat \mQ ]^s }_F^2  } \\
        & \quad \quad- \of{ \onormSize{}{[\mX_{i:} \hat \mQ]^s - [\hat \mS_{b:}]^s}_F^2 - \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:b}^T \mY \hat \mQ ]^s }_F^2   }. \\
        &(1+o(1))\hat H_{ib}  \\
        & = \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:a}^T \mY \hat \mQ ]^s }_F^2 - \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:b}^T \mY \hat \mQ]^s }_F^2 \\
        & \quad \quad + \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^2.
    \end{align}
    We obtain the upper bounds following the proof for inequalities~\eqref{eq:gib} and \eqref{eq:hib}.
}

\end{proof}

     \begin{lem}[Relationship between misclustering loss and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement} {with fixed $r \geq 2$, as $p \rightarrow \infty$}, we have
     \small
    \begin{equation}\label{eq:inter1}
        \onormSize{}{\mV - \mV^{(t)}}_{\sigma} \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}}  \frac{r}{\Delta_{\min}^2} L^{(t)},
    \end{equation}
    \begin{equation}\label{eq:inter2}
        \onormSize{}{\mE (\mV - \mV^{(t)})}_{\sigma} \lesssim \sqrt{ \frac{r^{K-1}(p r^{K-1} + p r)}{p^{K-1}}} 
 \frac{r}{\Delta_{\min}^2} L^{(t)},
 \end{equation}
 \begin{align}
     &\max_{b \in [r]} \onormSize{}{ [ \mW_{:b}^T \mY \mV ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV  ]^s } \\
     & \quad \quad \quad \leq C \of{\frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{r^{2K} + pr^{K+1}}{p^{K}} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}}, \label{eq:inter3}
 \end{align}
 \begin{align}
     &\max_{b \in [r]} \onormSize{}{  [  \mW_{:b}^{(t), T}  \mY \mV]^s - [  \mW_{:b}^{(t), T}  \mY \mV^{(t)}]^s } \\
      &\quad \quad \quad  \leq C  \of{ \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} +  \frac{r L^{(t)}}{\Delta_{\min}}}, \label{eq:inter4}
 \end{align}
 \begin{align}
     & \max_{b \in [r]} \onormSize{}{ [\mW_{:b}^T \mY \mV^{(t)} ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s }\\
     &\quad \quad \quad  \leq C \of{\frac{rL^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}}, \label{eq:inter5}
 \end{align}
 \normalsize
    for some positive universal constant $C$. In addition, the inequality~\eqref{eq:inter4} also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$. Further, the above inequalities holds after replacing $\mW$ to $\mP$, $\mV$ to $\mQ$, and $L^{(t)}$ to $L(\hat z)$.
    \end{lem}

    \begin{proof}[Proof of Lemma~\ref{lem:intermediate}] We follow and use several intermediate conclusions in \cite[Proof of Lemma 5]{han2020exact}. We prove each inequality separately.
    \begin{enumerate}[wide]
    \item Inequality~\eqref{eq:inter1}. By \cite[Proof of Lemma 5]{han2020exact}, we have 
    \begin{equation}
         \onormSize{}{\mV - \mV^{(t)}}_{\sigma} \lesssim \sqrt{\frac{r^{K-1}}{p^{K-1}}} r \ell^{(t)}.
    \end{equation}
    Then, we complete the proof of inequality~\eqref{eq:inter1} by applying Lemma~\ref{lem:mis} to the above inequality.
    \item Inequality~\eqref{eq:inter2}. By \cite[Proof of Lemma 5]{han2020exact}, we have 
    \begin{equation}
           \onormSize{}{\mE (\mV - \mV^{(t)})}_{\sigma} \lesssim \sqrt{ \frac{r^{K-1}(p r^{K-1} + p r)}{p^{K-1}}} r \ell^{(t)}.
    \end{equation}
    Also, we complete the proof of inequality~\eqref{eq:inter1} by applying Lemma~\ref{lem:mis} to the above inequality.

    \item Inequality~\eqref{eq:inter3}. We upper bound the desired quantity by triangle inequality,
    \begin{equation}
        \onormSize{}{ [ \mW_{:b}^T \mY \mV ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV  ]^s } \leq I_1 + I_2 + I_3,
    \end{equation}
    where 
    \begin{align}
        I_1 &=  \onorm{ \frac{ \mW_{:b}^T \mY \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{ \mW_{:b}^{(t), T} \mY \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }}   },\\
        I_2 & = \onorm{\of{ \frac{1}{\onormSize{}{\mW_{:b}^T \mY \mV}} -  \frac{1}{\onormSize{}{\mW_{:b}^T \mX \mV}}  } \mW_{:b}^T \mY \mV },\\
        I_3 &= \onorm{\of{ \frac{1}{\onormSize{}{\mW_{:b}^{(t), T} \mY \mV  }} -  \frac{1}{\onormSize{}{\mW_{:b}^{(t), T} \mX \mV  }}  } \mW_{:b}^{(t), T} \mY \mV }.
    \end{align}
    Next, we upper bound the quantities $I_1, I_2, I_3$ separately.

    For $I_1$, we further bound $I_1$ by triangle inequality,
    \begin{equation}
        I_1 \leq I_{11} + I_{12},
    \end{equation}
    where
 \begin{equation}
        I_{11} = \onorm{ \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }}   },
    \end{equation}
    and 
    \begin{equation}
        I_{12} = \onorm{ \frac{\mW_{:b}^T \mE \mV}{ \onormSize{}{  \mW_{:b}^T \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mE \mV}{ \onormSize{}{  \mW_{:b}^{(t),T} \mX \mV }} }.
    \end{equation}
    We first consider $I_{11}$. Define the confusion matrix $\mD = \mM^T \mTheta^T \mW^{(t)} = \entry{D_{ab}} \in \bbR^{r \times r}$ where 
    \begin{equation}
        D_{ab} = \frac{\sum_{i \in [p]}  \theta(i) \ind \offf{ z(i) = a, z^{(t)}(i) = b } }{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }}, \text{  for all } a, b \in [r].
    \end{equation}
    By Lemma~\ref{lem:membership},  we have $\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } \gtrsim p/r$. Then, we have
    \small
    \begin{equation}\label{eq:dab_bound}
        \sum_{a \neq b, a,b \in [r]} D_{ab} \lesssim \frac{r}{p} \sum_{i \colon z^{(t)}(i) \neq z(i)} \theta(i) \lesssim\frac{L^{(t)}}{\Delta_{\min}^2}  \lesssim \frac{1}{\log p}, 
    \end{equation}
    \normalsize
    and for all $b \in [r]$,
    \begin{align}
        D_{bb} &= \frac{\sum_{i \in [p]}  \theta(i) \ind \offf{ z(i) = z^{(t)}(i) = b } }{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }} \\
        &\geq \frac{c(\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } - p\ell^{(t)})}{\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b }} \\
        &\gtrsim 1 - \frac{1}{\log p},\label{eq:dbb_bound}
    \end{align}
    under the inequality \eqref{eq:cond_intial} in Condition~\ref{cond:origin}. By the definition of $\mW,\mW^{(t)},\mV$, we have 
    \begin{align}
        \frac{ \mW_{:b}^T \mX \mV  }{\onormSize{}{  \mW_{:b}^T \mX \mV }} = \off{\mS_{b:}}^s,
    \end{align}
    and 
    \begin{equation}
        \frac{ \mW_{:b}^{(t), T} \mX \mV  }{ \onormSize{}{\mW_{:b}^{(t), T} \mX \mV }} = [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}]^s.
    \end{equation}

    Let $\alpha$ denote the angle between $\mS_{b:}$ and $D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}$. To roughly estimate the range of $\alpha$, we consider the inner product 
    \begin{align}
        &\ang{\mS_{b:},  D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}} \\  &= D_{bb} \onorm{\mS_{b:}}^2 + \sum_{a \neq b} D_{ab}\ang{\mS_{b:}, \mS_{a:}}\\
        &\geq D_{bb} \onorm{\mS_{b:}}^2 -  \sum_{a \neq b, a \in [r]} D_{ab}   \onorm{\mS_{b:}} \max_{a \in [r]}  \onorm{\mS_{a:}} \\
        & \geq C,
    \end{align}
    where $C$ is a positive constant, and the last inequality holds when $p$ is large enough following the constraint of $\onorm{\mS_{b:}}$ in parameter space~\eqref{eq:family} and the bounds of $\mD$ in \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}.

    The positive inner product between $\mS_{b:}$ and $D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}$ indicates $\alpha \in [0,  \pi/2)$, and thus $2\sin \frac{\alpha}{2} \leq \sqrt{2} \sin \alpha$. Then, by the geometry property of trigonometric function, we have 
    \begin{align}
       &\onormSize{}{ [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}] \sin \alpha} \\
       &= \onormSize{}{(\mI_d - \text{Proj}(\mS_{b:})) \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}}\\
       &\leq \sum_{a \neq b, a \in [r]} D_{ab} \onorm{(\mI_d - \text{Proj}(\mS_{b:}))\mS_{a:}  }\\
       &= \sum_{a \neq b, a \in [r]} D_{ab} \onorm{ \mS_{a:} \sin (\mS_{b:}, \mS_{a:}) }\\
       & \leq \sum_{a \neq b, a \in [r]} D_{ab} \onorm{\mS_{a:}} \onorm{ \mS^s_{b:} - \mS^s_{a:}}, \label{eq:i11_sin_num}
    \end{align}
    where the first inequality follows from the triangle inequality, and the last inequality follows from Lemma~\ref{lem:norm_diff}. Note that with bounds \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}, when $p$ is large enough, we have
    \begin{align}
          \onormSize{}{\mW_{:b}^{(t), T} \mX \mV } &= \onormSize{}{ D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}} \\
          &\geq D_{bb} \onorm{\mS_{b:}} - \sum_{a \neq b, a \in [r]} D_{ab} \onorm{ \mS_{a:}} \\
          &\geq  C_1,\label{eq:i11_sin_dom}
    \end{align}
    for some positive constant $C_1$. Notice that $I_{11} = \sqrt{1 - \cos \alpha} = 2 \sin \frac{\alpha}{2}$. Therefore, we obtain
    \begin{align}
        I_{11} &\leq \sqrt{2} \sin \alpha \\
        & = \frac{  \onormSize{}{ [D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}] \sin \alpha}}{  \onormSize{}{ D_{bb} \mS_{b:} + \sum_{a \neq b, a \in [r]} D_{ab} \mS_{a:}} } \\
        & \leq \frac{ 1}{ C_1}  \sum_{a \neq b, a \in [r]} D_{ab} \onorm{\mS_{a:}} \onorm{ \mS^s_{b:} - \mS^s_{a:}} \\
        & \lesssim \frac{r}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind\offf{z^{(t)}(i) = b} \onorm{ \mS^s_{b:} - \mS^s_{a:}}\\
        & \leq \frac{r L^{(t)}}{\Delta_{\min}},\label{eq:i11}
    \end{align}
    where the second inequality follows from~\eqref{eq:i11_sin_num} and \eqref{eq:i11_sin_dom}, and the last two inequalities follow by the definition of $D_a$ and $L^{(t)}$, and the constraint of $\onorm{\mS_{b:}}$ in parameter space~\eqref{eq:family}.

    We now consider $I_{12}$. By triangle inequality, we have 
    \begin{align}
        I_{12} &\leq \frac{1}{\onormSize{}{  \mW_{:b}^T \mX \mV }}\onormSize{}{ (\mW_{:b}^T - \mW_{:b}^{(t),T}) \mE \mV }\\
        &\quad \quad + \frac{ \onormSize{}{(\mW_{:b}^T - \mW_{:b}^{(t),T}) \mX \mV} }{\onormSize{}{  \mW_{:b}^T \mX \mV } \onormSize{}{  \mW_{:b}^{(t),T} \mX \mV } }\onormSize{}{ \mW_{:b}^{(t),T} \mE \mV}.
    \end{align}
    By \cite[Proof of Lemma 5]{han2020exact}, we have 
    \begin{equation}\label{eq:i12_wwe}
        \onormSize{}{ (\mW_{:b}^T - \mW_{:b}^{(t),T}) \mE \mV }  \lesssim \sqrt{\frac{ r^{2K} + p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}.
    \end{equation}
    Notice that 
    \begin{align}
        \onormSize{}{(\mW_{:b}^T - \mW_{:b}^{(t),T}) \mX \mV} &\leq \onormSize{}{\mW_{:b}^T - \mW_{:b}^{(t),T}} \onorm{\mX\mV}_F \\
        &\lesssim \frac{r^{3/2}L^{(t)}}{\sqrt{p}\Delta_{\min}^2} \onormSize{}{\mS} \onormSize{}{\mTheta \mM}_{\sigma} \\
        &\lesssim \frac{ \sqrt{r L^{(t)}} }{\Delta_{\min}}, \label{eq:i12_wwx}
    \end{align}
    where the second inequality follows from~\cite[Inequality (121), Proof of Lemma 5]{han2020exact} and the last inequality follows from Lemma~\ref{lem:singular_thetam} and \eqref{eq:cond_intial} in Condition~\ref{cond:origin}.
     Note that $\onorm{\mW_{:b}^T \mX \mV} = \onorm{\mS_{b:}} \geq c_3$ and $\onormSize{}{  \mW_{:b}^{(t),T} \mX \mV } \geq C_1$ by inequality~\eqref{eq:i11_sin_dom}. Therefore, we have 
     \begin{align}
         I_{12} &\lesssim  \onormSize{}{ (\mW_{:b}^T - \mW_{:b}^{(t),T}) \mE \mV } \\
         & \quad \quad  \quad \quad  \quad \quad +   \onormSize{}{(\mW_{:b}^T - \mW_{:b}^{(t),T}) \mX \mV}   \onormSize{}{ \mW_{:b}^{(t),T} \mE \mV} \\
         &\lesssim \sqrt{\frac{ r^{2K} + p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} + \frac{ \sqrt{r L^{(t)}} }{\Delta_{\min}}\sqrt{\frac{r^{2K}}{p^K}} \\
         & \lesssim \sqrt{\frac{ r^{2K} + p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},\label{eq:i12}
     \end{align}
     where second inequality follows from the inequalities~\eqref{eq:i12_wwe}, \eqref{eq:i12_wwx}, and \eqref{eq:cond1} in Condition~\ref{cond:origin}.

    Hence, combining inequalities~\eqref{eq:i11} and~\eqref{eq:i12} yields  
    \begin{equation}\label{eq:i1}
        I_1 \lesssim \frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{\frac{ r^{2K} + p r^{K+1}}{p^K}} \frac{\sqrt{L^{(t)}}}{\Delta_{\min}}.
    \end{equation}

        For $I_2$ and $I_3$, recall that $\onorm{\mW_{:b}^T \mX \mV} = \onorm{\mS_{b:}} \geq c_3$ and $\onormSize{}{  \mW_{:b}^{(t),T} \mX \mV } \geq C_1$ by inequality~\eqref{eq:i11_sin_dom}. By triangle inequality and \eqref{eq:cond1} in Condition~\ref{cond:origin}, we have 
    \begin{equation}\label{eq:i2}
         I_2 \leq \frac{ \onormSize{}{ \mW_{:b}^T \mE \mV } }{ \onormSize{}{\mW_{:b}^T \mX \mV}  } \lesssim  \onormSize{}{ \mW_{:b}^T \mE \mV }  \lesssim \frac{r^K}{p^{K/2}},
    \end{equation}
    and 
    \begin{equation}\label{eq:i3}
        I_3 \leq \frac{ \onormSize{}{ \mW_{:b}^{(t),T} \mE \mV } }{ \onormSize{}{\mW_{:b}^{(t),T} \mX \mV}  }  \lesssim \onormSize{}{ \mW_{:b}^{(t),T} \mE \mV } \lesssim \frac{r^K}{p^{K/2}}.
     \end{equation}
       Therefore, combining the inequalities~\eqref{eq:i1}, \eqref{eq:i2}, and \eqref{eq:i3}, we finish the proof of inequality~\eqref{eq:inter3}.

    \item Inequality~\eqref{eq:inter4}. Here we only show the proof of inequality~\eqref{eq:inter4} with $\mW_{:b}^{(t)}$. The proof also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$, and we omit the repeated procedures.

    We upper bound the desired quantity by triangle inequality
    \begin{equation}
         \onormSize{}{  [   \mW_{:b}^{(t), T}  \mY \mV]^s -  [  \mW_{:b}^{(t), T}  \mY \mV^{(t)}]^s }  \leq J_1 + J_2 + J_3,
    \end{equation}
    where 
    \small
    \begin{align}
        J_1 &= \onorm{ \frac{\mW_{:b}^{(t),T} \mY \mV }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mY \mV^{(t)} }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV^{(t)} }}  },\\
        J_2 & = \onorm{\of{ \frac{1}{\onormSize{}{\mW_{:b}^{(t),T} \mY \mV}} -  \frac{1}{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV}}  } \mW_{:b}^{(t),T} \mY \mV },\\
        J_3 &= \onorm{\of{ \frac{1}{\onormSize{}{\mW_{:b}^{(t),T} \mY \mV^{(t)}}} -  \frac{1}{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)}} }  } \mW_{:b}^{(t), T} \mY \mV^{(t)} }.
    \end{align}
    \normalsize
    Next, we upper bound the quantities $J_1, J_2, J_3$ separately.

    For $J_1$, by triangle inequality,
we have 
    \begin{equation}
        J_1 \leq J_{11} + J_{12},
    \end{equation}
    where 
    \begin{equation}
        J_{11} =  \onorm{ \frac{\mW_{:b}^{(t),T} \mX \mV }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV }} - \frac{\mW_{:b}^{(t),T} \mX \mV^{(t)} }{\onormSize{}{ \mW_{:b}^{(t),T} \mX \mV^{(t)} }}  }
    \end{equation}
    and
    \begin{equation}
         J_{12} =  \onorm{ \frac{\mW_{:b}^{(t),T}  \mE \mV}{ \onormSize{}{\mW_{:b}^{(t),T}  \mX \mV} } - \frac{\mW_{:b}^{(t),T}  \mE \mV^{(t)}}{ \onormSize{}{\mW_{:b}^{(t),T}  \mX \mV^{(t)}} }   }.
    \end{equation}

    We first consider $J_{11}$.  Define the matrix $\mV^{k} \coloneqq \mW^{\otimes (k-1)} \otimes \mW^{(t), \otimes (K-k)}$ for $k = 2,\ldots, K-1$, and denote $\mV^1 = \mV^{(t)}, \mV^K = \mV$. Also, define the quantity
    \begin{equation}
        J_{11}^k = \onormSize{}{ [\mW_{:b}^{(t),T} \mX \mV^k ]^s - [\mW_{:b}^{(t),T} \mX \mV^{k+1} ]^s },
     \end{equation}
    for $k = 1,\ldots, K-1$. Let $\beta_k$ denote the angle between $\mW_{:b}^{(t),T} \mX \mV^k$ and $\mW_{:b}^{(t),T} \mX \mV^{k+1}$. With the same idea to prove $I_{11}$ in inequality~\eqref{eq:i11}, we bound $J_{11}^k$ by the trigonometric function of $\beta_k$.

    To roughly estimate the range of $\beta_k$, we consider the inner product between $\mW_{:b}^{(t),T} \mX \mV^k$ and $\mW_{:b}^{(t),T} \mX \mV^{k+1}$. Before the specific derivation of the inner product, note that 
    \begin{equation}
        \mW_{:b}^{(t),T} \mX \mV^k = \mat_1(\tT_k), \quad  \mW_{:b}^{(t),T} \mX \mV^{k+1} = \mat_1(\tT_{k+1}),
    \end{equation}
    where 
    \begin{align}
        \tT_k &= \tX \times_1  \mW_{:b}^{(t),T} \times_2 \mW^T \times_3 \cdots \times_k \mW^T  \\
        & \quad \quad  \quad \quad \quad \quad \quad \quad \times_{k+1} \mW^{(t),T} \times_{k+2} \cdots \times_K \mW^{(t),T}\\
        \tT_{k+1} &= \tX \times_1  \mW_{:b}^{(t),T} \times_2 \mW^T \times_3 \cdots \times_k \mW^T \\
        & \quad \quad  \quad \quad \quad \quad \quad \quad \times_{k+1} \mW^{T} \times_{k+2} \cdots \times_K \mW^{(t),T}. 
    \end{align}
    Recall the definition of confusion matrix $\mD = \mM^T \mTheta^T \mW^{(t)} = \entry{D_{ab}} \in \bbR^{r \times r}$.
 We have 
 \small
    \begin{align}
         &\ang{\mW_{:b}^{(t),T} \mX \mV^k , \mW_{:b}^{(t),T} \mX \mV^{k+1}} \\
         &  =\ang{ \mat_{k+1}(\tT_k), \mat_{k+1}(\tT_{k+1}) } \label{eq:j11_tensormat} \\
  & =  \ang{ \mD^T \mS \mZ^k ,  \mS \mZ^k   }\\
         & = \sum_{b \in [r]} \of{ D_{bb} \onormSize{}{\mS_{b:}\mZ^k}^2 + \sum_{a \neq b, a \in [r]} D_{ab} \ang{\mS_{a:} \mZ^k, \mS_{b:} \mZ^k}} \\
         & \gtrsim (1 - \log p^{-1}) \min_{a \in [r]}\onormSize{}{\mS_{a:}\mZ^k}^2 - \log p^{-1} \max_{a \in [r]} \onormSize{}{\mS_{a:}\mZ^k}^2, \label{eq:j11inner}
    \end{align}
    \normalsize
    where $\mZ^k = \mD_{:b} \otimes \mI_r^{\otimes (k-1)} \otimes  \mD^{\otimes (K-k-1)}$, the equations follow by the tensor algebra and definitions, and the last inequality follows from the bounds of $\mD$ in \eqref{eq:dab_bound} and \eqref{eq:dbb_bound}.

    Note that 
    \begin{align}
        \onorm{\mD}_{\sigma} &\leq \onorm{\mD}_{F} \\
        &\leq  \sqrt{\sum_{b \in [r]} D_{bb}^2 + (\sum_{a \neq b, a ,b \in [r]} D_{ab} )^2} \\
        &\lesssim \sqrt{r  + \log^2 p^{-1}} \lesssim 1,\label{eq:singular_Du}
    \end{align}
    where the second inequality follows from inequality~\eqref{eq:dab_bound}, and the fact that for all $b \in [r]$,
    \begin{equation}
        D_{bb} \lesssim \frac{r}{p} \sum_{i\colon z(i) = b} \theta(i) \lesssim 1.
    \end{equation}
    Also, we have 
    \begin{equation}\label{eq:singular_Dl}
        \lambda_r(\mD) \geq \lambda_r(\mW^{(t)}) \lambda_r(\mTheta \mM) \gtrsim 1,
    \end{equation}
    following the Lemma~\ref{lem:singular_thetam} and Lemma~\ref{lem:membership}. Then, for all $k \in [K]$, we have 
    \begin{align}
        1 &\lesssim \onorm{\mD_{:b}} \lambda_r(\mD)^{K-k-1}  \leq \lambda_{r^{K-2}}(\mZ^{k}) \\
        &\leq \onormSize{}{\mZ^k}_{\sigma} \leq \onorm{\mD_{:b}} \onorm{\mD}_{\sigma}^{K-k-1} \lesssim 1.\label{eq:singular_z}
    \end{align}
    Thus, we have bounds 
    \begin{equation}
         \max_{a \in [r]} \onormSize{}{\mS_{a:} \mZ^k} \leq \max_{a \in [r]} \onorm{\mS_{a:}} \onormSize{}{\mZ^k}_{\sigma} \lesssim 1, 
    \end{equation}
    \begin{equation}
        \min_{a \in [r]} \onormSize{}{\mS_{a:}\mZ^k}  \geq \min_{a \in [r]} \onorm{\mS_{a:}} \lambda_{r^{K-2}}(\mZ^{k}) \gtrsim 1.
    \end{equation}
    Hence, when $p$ is large enough, the inner product~\eqref{eq:j11inner} is positive, which implies $\beta_k \in [0, \pi/2)$ and thus $2 \sin \frac{\beta_k}{2} \leq \sqrt{2} \sin \beta_k$.

    Next, we upper bound the trigonometric function $\sin \beta_k$. Note that
    \begin{align}
         \sin \beta_k &= \sin ({\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k} }, {\mD_{:b}^{T} \mS \mI_r^{\otimes k} \otimes \mD^{\otimes K-k-1} } )\\
         &\leq \sin \beta_{k1} + \sin \beta_{k2},
    \end{align}
    where
    \begin{align}
         \sin \beta_{k1} &= \sin ({\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k} }, \\
         & \quad \quad \quad \quad \quad \quad\quad \quad\quad \quad {\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \tilde \mD \otimes \mD^{\otimes K-k-1} } ),\\
         \sin \beta_{k2} & = \sin ({\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \tilde \mD \otimes \mD^{\otimes K-k-1} } ,\\
         &\quad \quad \quad \quad \quad \quad\quad \quad\quad \quad  {\mD_{:b}^{T} \mS \mI_r^{\otimes k} \otimes \mD^{\otimes K-k-1} }   ),
    \end{align}
    and $\tilde \mD$ is the normalized confusion matrix with entries $\tilde \mD_{ab} = \frac{\sum_{i \in [p]} \theta(i) \ind\{z^{(t)} = b, z(i) = a\}}{\sum_{i \in [p] }\theta(i) \ind\{z^{(t)} = b\}}$.

    % {
    % \color{blue}

    % Consider two vectors 
    % \begin{equation}
    %     \mp( z^{(t)})=(| z^{(t),-1}(1)|, \ldots,| z^{(t),-1}(r)|)^T,\quad \mp_{\mtheta}( z^{(t)})=(\onormSize{}{\mtheta_{ z^{(t),-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{(t), -1}(r)}}_1)^T.
    % \end{equation}
    %  Note that we have $\ell^{(t)} \lesssim p^{-1} \sum_{i \in [p]} \theta(i) \ind\{ z^{(t)} = b \}  \leq \frac{L^{(t)}}{\Delta_{\min}^2} $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}. For all $b \in [r]$, We have 
    % \begin{equation}
    %     | z^{(t),-1}(b)| =  (1 + \tO(L^{(t)}/\Delta_{\min}^2 ))  | z^{-1}(b)|,
    % \end{equation}
    % and 
    % \begin{align}
    %     \onormSize{}{\mtheta_{ z^{(t),-1}(b)}}_1 &= \sum_{ i \in [p]} \theta(i) \ind\{ z^{(t)}(i) = z(i) = b \} + \sum_{ i \in [p]} \theta(i) \ind\{ z^{(t)}(i) = b,  z(i) \neq b \} \\
    %     &\ +  \sum_{ i \in [p]} \theta(i) \ind\{ z^{(t)}(i) \neq  b,  z(i) = b \} - \sum_{ i \in [p]} \theta(i) \ind\{ z^{(t)}(i) \neq  b,  z(i) = b \}  \\
    %     & =  | z^{-1}(b)| + \tO(  \sum_{i \in [p]} \theta(i) \ind\{ z^{(t)} \neq b \} ) = (1 + \tO(L^{(t)}/\Delta_{\min}^2 ))  | z^{-1}(b)|.
    % \end{align}
    % Hence, we have 
    % \begin{equation}
    %     \onormSize{}{  \mp( z^{(t)}) -  \mp_{\mtheta}( z^{(t)}) } \leq r \tO(L^{(t)}/\Delta_{\min}^2 )  \max_{b \in [r]}| z^{-1}(b)| \lesssim  \tO(L^{(t)}/\Delta_{\min}^2 ) p
    % \end{equation}
    % with assumption $\onormSize{}{ z^{-1}(b)} \lesssim p$ in parameter space~\eqref{eq:family}.
    %  Note that $\text{diag}(\mp(z^{(t)})) \mD = \text{diag}(\mp_{\mtheta}(z^{(t)})) \tilde \mD$. Then, we have
    % \begin{align}
    %     \onormSize{}{ \mD -  \tilde \mD}_F &\leq  \onormSize{}{ \mI_r -  \text{diag}(\mp(z^{(t)})) \text{diag}^{-1}(\mp_{\mtheta}(z^{(t)}))}_F \onormSize{}{\mD}_F\\
    %   &\lesssim \frac{ \onormSize{}{ \mp(z^{(t)}) - \mp_{\mtheta}(z^{(t)})} }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  } \\
    %   & = \frac{\tO(L^{(t)}/\Delta_{\min}^2 ) p }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1 } \\
    %   & \lesssim \frac{L^{(t)}}{\Delta_{\min}},\fixme{miaoyan}{$\Delta^2_{min}$?}
    %   \end{align}
    % where the last inequality follows from the constraint $\min_{i \in [p]}\theta(i) \geq c>0$  and $\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  \gtrsim p$.  By the geometry property of trigonometric function, we have
    % \begin{align}\label{eq:b1}
    %     \sin \beta_{k1} &= \min_{c \in \bbR} \frac{\onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes (\mD -  \tilde \mD) \otimes \mD^{\otimes K-k - 1} } }{ \onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k}} }\\
    %     &\leq \frac{ \onormSize{}{\mD_{:b}^{T} \mS} \onormSize{}{\mD - \tilde \mD}_{\sigma} \onormSize{}{\mD}_{\sigma}^{K-k-1} }{\onormSize{}{\mD_{:b}^{T} \mS} \lambda_r^{K-k}(\mD) }\\
    %     & \lesssim  \onormSize{}{\mD - \tilde \mD}_{F}\\
    %     & \lesssim \frac{L^{(t)}}{\Delta_{\min}},
    % \end{align}
    % where the second inequality follows from the singular property of $\mD$ in \eqref{eq:singular_Du}, \eqref{eq:singular_Dl} and the constraint of $\mS$ in \eqref{eq:family}.

    % }

    To bound $\sin \beta_{k1}$, recall Definition~\ref{def:stable} that for any cluster assignment $\bar z$ in the $\varepsilon$-neighborhood of true $z$,
    \begin{align}
        &\mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T,\quad\\
        &\mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
    \end{align}
    {Note that we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \leq \frac{\bar C}{\tilde C} r \log^{-1}(p) $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}. Then,  with the locally linear stability assumption, the $\mtheta$ is $\ell^{(t)}$-locally linearly stable; i.e.,}
    %With the locally linear stability assumption, we have 
    \begin{equation}
         \sin (\mp(z^{(t)}), \mp_{\mtheta}(z^{(t)})) \lesssim \frac{L^{(t)}}{\Delta_{\min}}. 
    \end{equation}
    Note that $\text{diag}(\mp(z^{(t)})) \mD = \text{diag}(\mp_{\mtheta}(z^{(t)})) \tilde \mD$, and $\sin(\ma,\mb) =\min_{c \in \bbR} \frac{ \onormSize{}{\ma - c\mb} }{\onormSize{}{\ma}}$ for vectors $\ma,\mb$ of same dimension. Let $c_0 = \argmin_{c \in \bbR} \frac{\onormSize{}{\mp(z^{(t)}) - c\mp_{\mtheta}(z^{(t)})}}{\onormSize{}{\mp(z^{(t)})}}$. Then, we have
    \begin{align}
      &\min_{c \in \bbR} \onormSize{}{ \mD - c \tilde \mD}_F \\
      &\leq  \onormSize{}{ \mI_r - c_0 \text{diag}(\mp(z^{(t)})) \text{diag}^{-1}(\mp_{\mtheta}(z^{(t)}))}_F \onormSize{}{\mD}_F\\
      &\lesssim \frac{ \onormSize{}{ \mp(z^{(t)}) - c_0\mp_{\mtheta}(z^{(t)})} }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  } \\
      & = \frac{\onormSize{}{ \mp(z^{(t)})} }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1 } \sin (\mp(z^{(t)}), \mp_{\mtheta}(z^{(t)}))\\
      & \lesssim \frac{L^{(t)}}{\Delta_{\min}},
    \end{align}
    where the last inequality follows from Lemma~\ref{lem:membership}, the constraint $\min_{i \in [p]}\theta(i) \geq c>0$, $\onormSize{}{ \mp(z^{(t)})} \lesssim p$ and $\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  \gtrsim p$.

    By the geometry property of trigonometric function, we have
    \begin{align}
        \sin \beta_{k1} &= \min_{c \in \bbR} \frac{\onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes (\mD - c \tilde \mD) \otimes \mD^{\otimes K-k - 1} } }{ \onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k}} }\\
        &\leq \frac{ \onormSize{}{\mD_{:b}^{T} \mS} \onormSize{}{\mD - c_0 \tilde \mD}_{\sigma} \onormSize{}{\mD}_{\sigma}^{K-k-1} }{\onormSize{}{\mD_{:b}^{T} \mS} \lambda_r^{K-k}(\mD) }\\
        & \lesssim  \onormSize{}{\mD - c_0 \tilde \mD}_{F}\\
        & \lesssim \frac{L^{(t)}}{\Delta_{\min}}, \label{eq:b1}
    \end{align}
    where the second inequality follows from the singular property of $\mD$ in \eqref{eq:singular_Du}, \eqref{eq:singular_Dl} and the constraint of $\mS$ in \eqref{eq:family}.

    To bound $\sin \beta_{k2}$, let $\mC = \text{diag}(\{\onormSize{}{\mS_{a:}}\}_{a \in [r]})$. We have 
    \begin{align}
        \sin \beta_{k2} &\lesssim \frac{ \onorm{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes (\mI_r - \tilde \mD) \otimes \mD^{\otimes K-k-1} } }{\onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k} \otimes \mD^{\otimes K-k-1} }  }\\
        & \lesssim \frac{ \onormSize{}{(\mI_r - \tilde \mD^T) \mS \mZ^k}_F }{\onormSize{}{\mD_{:b}^{T} \mS} \lambda_r^{K-k-1}(\mD)} \\
        & \lesssim \onormSize{ }{(\mI_r - \tilde \mD^T) \mS \mC^{-1}}_F \onormSize{}{\mC \mZ^k}_{\sigma}\\
        & \lesssim \frac{r}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind\{z^{(t)}(i) = b\} \onormSize{}{\mS_{b:}^s - \mS_{z(i):}^s} \\
        & \lesssim \frac{L^{(t)}}{\Delta_{\min}}, \label{eq:b2} %\fixme{miaoyan}{reason for last line?} 
    \end{align}
    where the third inequality follows from the singular property of $\mD$ and the boundedness of $\mS$, and the fourth inequality follows from the definition of $\tilde \mD$, boundedness of $\mS$, the lower bound of $\mtheta$, and the singular property of $\mZ^k$ in inequality~\eqref{eq:singular_z}, and the last line follows from the definition of $L^{(t)}$.

  Combining~\eqref{eq:b1} and~\eqref{eq:b2} yields
    \begin{equation}
        \sin \beta_k \leq \sin \beta_{k1} + \sin \beta_{k2} \lesssim \frac{L^{(t)}}{\Delta_{\min}}.
    \end{equation}

       Finally, by triangle inequality, we obtain
    \begin{equation}\label{eq:j1}
        J_{11} \leq \sum_{k = 1}^{K-1} J_{11}^k \lesssim  \sum_{k = 1}^{K-1} \sin \beta_k  \lesssim (K-1)\frac{r L^{(t)}}{\Delta_{\min}}.
    \end{equation}

    We now consider $J_{12}$. By triangle inequality, we have 
    \begin{align}
        J_{12} &\leq \frac{ 1}{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }} \onormSize{}{\mW_{:b}^{(t),T} \mE (\mV-\mV^{(t)})}  \\
        & \quad \quad +
        \frac{ \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})} }{\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)} }} \onormSize{}{\mW_{:b}^{(t),T} \mE \mV^{(t)} }.
    \end{align}
     Note that 
     \begin{align}
          \onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)}} &= \onormSize{}{\mD^T \mS \mZ^1} 
          \\ &\geq \lambda_r(\mD) \onorm{\mS} \lambda_{r^{K-2}}(\mZ^1) \gtrsim 1, \label{eq:j11_wxvt}
     \end{align}
    where the inequality follows from the bounds~\eqref{eq:singular_Dl} and \eqref{eq:singular_z}.

    By \cite[Proof of Lemma 5]{han2020exact}, we have 
    \begin{align}
       & \onormSize{}{\mW_{:b}^{(t),T} \mE (\mV-\mV^{(t)})} \\
        & \quad \quad \quad \quad \quad \lesssim \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{(K-1)\sqrt{L^{(t)}}}{\Delta_{\min}}. \label{eq:j12_wev}
    \end{align}
    Notice that 
    \begin{align}
        & \onormSize{}{\mX (\mV^k-\mV^{k+1})}_F \\ &\leq \onormSize{}{(\mI - \mD^T) \mS (\mI_r^{\otimes (k-1)} \otimes  \mD^{\otimes (K-k-1)})}_F\notag \\
        &\leq \onormSize{}{(\mW^T - \mW^{(t),T}) \mTheta \mM}_F \onorm{\mS}_F \onorm{ \mD}_{\sigma}^{K-k-1}\notag \\
        & \lesssim \onormSize{}{ \mW^T - \mW^{(t),T} } \onorm{\mTheta \mM}_{\sigma} \notag\\
        & \lesssim \frac{ \sqrt{r L^{(t)}}}{\Delta_{\min}}, \label{eq:step}
    \end{align}
    where the first inequality follows from the tensor algebra in 
    inequality~\eqref{eq:j11inner}, the second inequality follows from the fact  that $\mI = \mW^T \mTheta \mM$, and the last inequality follows from \cite[Proof of Lemma 5]{han2020exact}. It follows from~\eqref{eq:step} and Lemma~\ref{lem:membership} that
    \small
    \begin{align}
        \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})}&\leq \onormSize{}{\mW_{:b}^{(t),T} }  \sum_{k = 1}^{K-1} \onormSize{}{\mX (\mV^k-\mV^{k+1})}_F \\ &\lesssim \frac{ \sqrt{r L^{(t)}}}{\sqrt{p} \Delta_{\min}}.\label{eq:j12_wxv}
    \end{align}
    \normalsize
    Note that $\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }$ and $\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)} }$ are lower bounded by inequalities~\eqref{eq:i11_sin_dom} and \eqref{eq:j11_wxvt}, respectively. We have 
    \begin{align}
        J_{12} &\lesssim \onormSize{}{\mW_{:b}^{(t),T} \mE (\mV-\mV^{(t)})} \\
        & \quad \quad \quad \quad + \onormSize{}{\mW_{:b}^{(t),T} \mX (\mV-\mV^{(t)})}\onormSize{}{\mW_{:b}^{(t),T} \mE \mV^{(t)} }\\
        &\lesssim \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}} + \frac{ \sqrt{r L^{(t)}}}{\sqrt{p} \Delta_{\min}} \sqrt{\frac{r^{2K}}{p^K}} \\
        & \lesssim  \sqrt{ \frac{r^{2K+1} + p r^{2 + K}}{p^K} } \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
    \end{align}
    where the second inequality follows from inequalities~\eqref{eq:j12_wev}, \eqref{eq:j12_wxv}, and the inequality \eqref{eq:cond1} in Condition~\ref{cond:origin}.

    For $J_2$ and $J_3$, recall that $\onormSize{}{\mW_{:b}^{(t),T} \mX \mV }$ and $\onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)} }$ are lower bounded by inequalities~\eqref{eq:i11_sin_dom} and \eqref{eq:j11_wxvt}, respectively. By triangle inequality and inequality \eqref{eq:cond1} in Condition~\ref{cond:origin}, we have 
    \begin{equation}\label{eq:j2}
        J_2 \leq \frac{ \onormSize{}{\mW_{:b}^{(t),T} \mE \mV}  }{  \onormSize{}{\mW_{:b}^{(t),T} \mX \mV} } \lesssim \onormSize{}{\mW_{:b}^{(t),T} \mE \mV} \lesssim \frac{r^K}{p^{K/2}},
    \end{equation}
    and 
    \begin{equation}\label{eq:j3}
        J_3 \leq \frac{ \onormSize{}{\mW_{:b}^{(t),T} \mE \mV^{(t)}}  }{  \onormSize{}{\mW_{:b}^{(t),T} \mX \mV^{(t)}} } \lesssim \onormSize{}{\mW_{:b}^{(t),T} \mE \mV} \lesssim \frac{r^K}{p^{K/2}}.
    \end{equation}

    Therefore, combining the inequalities~\eqref{eq:j1}, \eqref{eq:j2}, and \eqref{eq:j3}, we finish the proof of inequality~\eqref{eq:inter4}.

    \item Inequality~\eqref{eq:inter5}. By triangle inequality, we upper bound the desired quantity 
    \begin{align}
        &\onormSize{}{ [ \mW_{:b}^T \mY \mV^{(t)} ]^s   -  [ \mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s } \\
        &\leq \onormSize{}{[ \mW_{:b}^T \mY \mV^{(t)} ]^s - [ \mW_{:b}^T \mY \mV ]^s } \\ & \quad \quad \quad\quad\quad + \onormSize{}{[\mW_{:b}^T \mY \mV ]^s  - [ \mW_{:b}^{(t),T} \mY \mV ]^s } \\
        & \quad \quad \quad\quad\quad+ \onormSize{}{[\mW_{:b}^{(t),T} \mY \mV ]^s  -  [\mW_{:b}^{(t), T} \mY \mV^{(t)} ]^s }\\
        &\lesssim \frac{r L^{(t)}}{\Delta_{\min}} + \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L^{(t)}}}{\Delta_{\min}},
    \end{align}
    \normalsize
    following the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4}. Therefore, we finish the proof of inequality~\eqref{eq:inter5}.
        \end{enumerate}
  {

  \textbf{Next, we show the intermediate inequalities holds with $\mP, \mQ$ and $L(\hat z)$.}

        Consider the MLE confusion matrix $\hat \mD = \mM^T \mTheta^T \hat \mP = \entry{\hat D_{ab}} \in \bbR^{r \times r}$ with entries
        \small
\begin{align}
    &\hat D_{ab}
    = \frac{ \sum_{i \in [p]} \theta(i) \hat \theta(i) \ind\{ z(i) = a, \hat z(i) = b  \} }{\onormSize{}{\hat \mtheta_{\hat z^{-1}(b)}}^2}\\
    &\ = \frac{ \sum_{i \in [p]} (1 + o(p^{K-2}))(\hat \theta(i))^2 \ind\{ z(i) = a, \hat z(i) = b  \} }{\onormSize{}{\hat \mtheta_{\hat z^{-1}(b)}}^2}, \label{eq:mle_confusion}
\end{align}
\normalsize
where the second equation follows from Lemma~\ref{lem:poly_mle_degree}, and thus $\sum_{a \in [r]}\hat \mD_{ab} = 1 + o(1).$ By the derivation of~\eqref{eq:dab_bound}, \eqref{eq:dbb_bound}, \eqref{eq:singular_Dl}, and \eqref{eq:singular_Du}, we have 
\begin{equation}
     \sum_{a \neq b \in [r]} \hat D_{ab} \lesssim \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\hat \theta(i))^2 \lesssim \frac{1}{\log p},
\end{equation}
\begin{equation}
     \hat D_{bb} \gtrsim 1- \frac{1}{\log p}, \quad \lambda_{\min}(\hat \mD) \asymp \onormSize{}{\hat \mD}_{\sigma} = (1 + o(1)).
\end{equation}
for all $a \neq b \in [r]$.

Now, we are ready to show the intermediate inequalities. First, by Lemma~\ref{lem:angle_gap_x} and $\min_{ i \in [p]} \theta(i) \geq c$, we have 
\begin{equation}
    \onormSize{}{\mS_{a:}^s - \mS_{b:}^s} \asymp \onormSize{}{\mA_{a:}^s - \mA_{b:}^s}.
\end{equation}
Then we can replace the $L^{(t)}$ by $L(\hat z)$ in the proof of Lemma~\ref{lem:intermediate}. The analogies of inequalities~\eqref{eq:inter1}, \eqref{eq:inter2}, \eqref{eq:inter3},  \eqref{eq:inter4}, and \eqref{eq:inter5}  hold by using the MLE confusion matrix and the definition of $L(\hat z)$. 
%It suffices to two show the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4} holds with $\mP, \mQ$.

Particularly, for the analogy of \eqref{eq:inter4}, the usage of MLE confusion matrix avoids the stability condition on $\mtheta$. Let $\bar \mD$ be the normalized version of $\hat \mD$. The angle in inequality~\eqref{eq:b1} decays to 0 at speed $p^{-(K-2)} \lesssim \Delta_{\min}$ when $K \geq 3$, and the inequality~\eqref{eq:b2} holds by the fact that 
\begin{align}
    \onormSize{}{(\mI_r - \bar \mD) \mS \mC^{-1}}_F 
    &\lesssim \frac{r}{p} \sum_{i \in [p]} (\theta(i))^2 \sum_{b \in [r]} \onormSize{}{\mS_{b:}^s - \mS_{z(i):}^s} \\
    &\lesssim \frac{r}{p} \sum_{i \in [p]} (\theta(i))^2 \sum_{b \in [r]} \onormSize{}{\mA_{b:}^s - \mA_{z(i):}^s}.
\end{align} 
  }

    \end{proof}

\begin{lem}[Polynomial estimation error of MLE]\label{lem:poly_mle_degree}  Let $(\hat z, \hat \mtheta, \hat \tS)$ denote the MLE in \eqref{eq:mle} {with fixed $K \geq 2$}, and $\hat \tX$ denote the mean tensor consisting of parameter $(\hat z, \hat \mtheta, \hat \tS)$. {With high probability going to 1 as $p \rightarrow \infty$, we} have 
\begin{equation}
    \onormSize{}{\tX - \hat \tX}_F^2 \lesssim \sigma^2 \of{ r^{K} + K p r},
\end{equation}
with probability going to 1. 
When {$\text{SNR} \gtrsim p^{-(K-1)} \log p$,  $\mtheta$ is balanced, and $\min_{i \in [p]} \theta(i) \geq c$ for some constant $c$}, the MLE satisfies
\begin{align}
    &\frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\theta(i))^2 \lesssim \frac{1}{r \log p}, \quad\\
    &\frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\hat \theta(i))^2 \lesssim \frac{1}{r \log p} ,\quad   \\
    & \text{and } L(\hat z) \lesssim \frac{\Delta_{\min}^2}{r \log p},
\end{align}
% The MLE loss is upper bounded as 
% \begin{equation}
%      L(\hat z) = \frac{1}{p} \sum_{i \in [p]} \theta(i)^2 \sum_{b \in [r]} \ind\{ \hat z(i) = b\} \onormSize{}{[\mA_{\hat \pi(z(i)):}]^s - [\mA_{b:}]^s}^2 \lesssim 
% \end{equation}
% where $\mA \coloneqq \mS (\mM^T \mTheta^T)^{\otimes (K-1)}$, and $\hat \pi$ is the permutation minimizes the MLE error.  
Further, we have 
\begin{equation}
     \theta(i)^2 = (1 + o(p^{-(K-2)})) \hat \theta(i)^2.
 \end{equation}
\end{lem}

\begin{proof}[Proof of Lemma~\ref{lem:poly_mle_degree}] Without loss of generality, we assume $\sigma^2 = 1$ and identity mapping minimizes the misclustering error for MLE.
For arbitrary two sets of parameters $(z, \mtheta, \tS), (z_k', \mtheta_k', \tS') \in \tP_b(\gamma)$ and corresponding mean tensors $\tX, \tX'$, we have 
\begin{align}
    \text{rank}(\mat_k(\tX) - \mat_k(\tX')) &\leq \text{rank}(\mat_k(\tX)) + \text{rank}(\mat_k(\tX)) \\
    &\leq 2 r_k, \quad k \in [K].
\end{align}

Hence, we have
\begin{equation}\label{eq:dtbm_tucker}
    \tX - \tX' \in \tQ(2r_1, \ldots, 2 r_K),
\end{equation}
where $\tQ(r_1, \ldots, r_K) \coloneqq \{ \text{Tucker tensor with rank }(r_1, \ldots,  r_K) \} $.

Then, we obtain that 
\begin{align}
     & \bbP(\FnormSize{}{\tX - \hat \tX_{ML}} \geq t) \\
     &\leq 2  \bbP\of{ \sup_{ \tX, \tX' \in \tP(r_1, 
\ldots, r_K)} \ang{ \frac{\tX - \tX'}{\FnormSize{}{\tX - \tX'}}, \tE  } \geq t } \\
     &\leq 2  \bbP\of{ \sup_{ \tT \in \tQ(2r_1, \ldots, 2r_K) \cap \{ \onormSize{}{\tT}_F = 1\} } \ang{ \tT, \tE  } \geq t } \\
     &\lesssim  \exp(-Kpr),
\end{align}
with the choice $t\asymp \sigma \sqrt{(Kpr + r^K)}$. Here the first inequality follows from \cite[Lemma 1]{wang2019multiway}, the second inequality follows from~\eqref{eq:dtbm_tucker}, and the last inequality follows from \cite[Lemma E5]{han2022optimal}.

When $\Delta_{\min}^2 \gtrsim p^{-(K-1)} \log p$, we replace the vector $\hat x_{\hat z(i)}$ and $\hat \mX$ by our MLE estimator in the proof of Theorem~\ref{thm:initial}. With estimation error $\onormSize{}{\tX - \hat \tX}_F^2 \lesssim \of{ r^{K} + K p r}$ and $\Delta_{\min}^2 \gtrsim p^{-(K-1)} \log p$, we have 
\begin{align}
      \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\theta(i))^2 &\lesssim \frac{r^{K-1}}{\Delta_{\min}^2 p^{K}} \onormSize{}{\tX - \hat \tX}_F^2 \\
      &\lesssim \frac{r^{K-2}}{p^{K-1} \Delta_{\min}^2} \\
      & \lesssim \frac{1}{r \log p},
\end{align}
and 
\begin{equation}
     L(\hat z) \lesssim \frac{\Delta_{\min}^2}{r \log p}.
\end{equation}
Above result holds for $\hat \theta(i)$ after switching the parameters $\mX$ with $\hat \mX$ and switch $\mtheta$ with $\hat \mtheta$ in the proof.

Last, notice that for all $a \in [r]$
% following the proof of inequality~\eqref{eq:cor_4} in Corollary~\ref{cor:L0} with $\onormSize{}{\tX - \hat \tX}^2_F \lesssim pr$,  we have 
% \begin{equation}
%     \onormSize{}{ \mW_{:a}^T \mX -  \hat \mW_{:a}^T \hat \mX }
% \end{equation}
%  for all $a \in [r]$,
\begin{align}
    &(1 - O(1)) \frac{p^2}{r^2}\onormSize{}{\mW_{:a}^T \mX - \hat \mW_{:a}^T \hat \mX}_F^2 \\
    &\leq  \onormSize{}{ \sum_{\hat z(i) = z(i) = a} (\theta(i) \mW_{:a}^T \mX  - \hat \theta(i) \hat \mW_{:a}^T \hat \mX)}_F^2 \\
    &\leq \onormSize{}{\tX - \hat \tX}_F^2 \leq p r,
\end{align}
where the first inequality follows from the facts that  $\ell(\hat z, z) \lesssim \frac{1}{\log p}, |z^{-1}(a)| \asymp p /r$, 
\begin{align}
     | z^{-1}(a)| - C \frac{p}{r} \ell(\hat z, z)  \leq |\hat z^{-1}(a)| \leq | z^{-1}(a)| + C \frac{p}{r} \ell(\hat z, z), 
\end{align}
\begin{equation}
  | z^{-1}(a)|  - C \frac{p}{r} \ell( \hat z,z)  \leq \sum_{ z(i) = z(i) = a} \theta(i) \leq | z^{-1}(a)|, 
\end{equation}
and 
\begin{equation}
    |\hat z^{-1}(a)|  - C \frac{p}{r} \ell(\hat z, z)  \leq \sum_{\hat z(i) = z(i) = a} \hat \theta(i) \leq |\hat z^{-1}(a)|.
\end{equation}

Hence, for all $i \in [p]$
\begin{align}
  &(\theta(i) - \hat \theta(i))^2 \onormSize{}{ \mW_{:a}^T \mX }_F^2 - O(p) \\
  &\leq  \onormSize{}{ (\theta(i) - \hat \theta(i))\mW_{:a}^T \mX }_F^2 - \onormSize{}{ \hat \theta(i) (\mW_{:a}^T \mX - \hat \mW_{:a}^T \hat \mX) }_F^2 \\
  &\leq \onormSize{}{\tX - \hat \tX}_F^2 \leq p r,\label{eq:theta_hat}
\end{align}
where the first inequality follows from $\onormSize{}{\mW_{:a}^T \mX - \hat \mW_{:a}^T \hat \mX}_F^2 \lesssim 1/p$ and $\hat \theta(i) \lesssim \frac{p}{r}$. Notice that for all $a \in [r]$
\begin{equation}
    \onormSize{}{\mW_{:a}^T \mX}_F^2 \geq \onormSize{}{\mS_{a:}}_F^2 \lambda_{\min}^{2(K-1)}(\mTheta \mM) \gtrsim p^{K-1}.
\end{equation}
The inequality indicates that $\theta(i)^2 = (1 + o(p^{-(K-2)})) \hat \theta(i)^2$.

\end{proof}

% use section* for acknowledgment
\section*{Acknowledgment}

This research is supported in part by NSF CAREER DMS-2141865, DMS-1915978, DMS-2023239, EF-2133740, and funding from the Wisconsin Alumni Research foundation. We thank Zheng Tracy Ke, Rungang Han, Yuetian Luo for helpful discussions and for sharing software packages.

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
% \begin{thebibliography}{1}

% \bibitem{IEEEhowto:kopka}
% H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

% \end{thebibliography}

\bibliographystyle{IEEEtran}
\bibliography{tensor_wang}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

% \begin{IEEEbiography}{Michael Shell}
% Biography text here.
% \end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{Jiaxin Hu} received the B.S.\ degree from Wuhan University in 2019 and Master's degree in statistics from the University of Wisconsin-Madison in 2020. She is currently a Ph.D.\ student with the Department of Statistics, University of Wisconsin-Madison. Her research interests mainly include tensor data and network analysis with applications in genetics and ecology.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Miaoyan Wang} received the B.S.\ degree from Fudan University and the Ph.D.\ degree from the University of Chicago. She is currently the Assistant Professor with the Department of Statistics, University of Wisconsin-Madison. She received the NSF CAREER Award in 2022. Her work focuses on tensor/matrix decomposition, high-dimensional statistics, multilinear optimization, and applications in genetics and neuroimaging. 
\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2023-01-21 tensor_clustering/ieee_final/ieee_final_source/ieee_final.tex
final_version

@@
 (e-mail: jhu267@wisc.edu, miaoyan.wang@wisc.edu)}% <-this % stops a space
-\thanks{Part of this paper was presented at 25th International Conference on Artificial Intelligence and Statistics (AISTATS).}% <-this % stops a space
+\thanks{This paper was presented in part at 25th International Conference on Artificial Intelligence and Statistics (AISTATS).}% <-this % stops a space
 }
@@
 % The paper headers
-\markboth{Journal of \LaTeX\ Class Files,~Vol.~00, No.~0, 000~0000}{Hu \MakeLowercase{\textit{et al.}}: Multiway Spherical Clustering}
+\markboth{Journal of \LaTeX\ Class Files,~Vol.~00, No.~0, 000~0000}{Hu \MakeLowercase{\textit{et al.}}: Multiway Spherical Clustering via Degree-Corrected Tensor Block Models}
 % The only time the second header will appear is for the odd numbered pages
@@
 Towards this goal, we have developed a new polar coordinates based analysis to control the model complexity.
-We also develop a new angle-based iteration algorithm to achieve optimal clustering rates \emph{without the need of estimating nuisance degree parameters}.
+We have also developed a new angle-based iteration algorithm to achieve optimal clustering rates \emph{without the need of estimating nuisance degree parameters}.
 
@@
 We believe our results are novel and helpful to the community.
-See Fig~\ref{fig:phase_axis} for overview of our results.
+See Figure~\ref{fig:phase_axis} for overview of our results.
 
@@
 We use $\tY(i_1,\ldots,i_K)$ to denote the $(i_1,\ldots,i_K)$-th entry of $\tY$.
-The multilinear multiplication of a tensor $\tS\in \bbR^{r_1\times \cdots \times r_K}$ by matrices $\mM_k \in\mathbb{R}^{p_k\times r_k}$ results in an order-$d$ $(p_1,\ldots,p_K)$-dimensional tensor $\tX$, denoted
+The multilinear multiplication of a tensor $\tS\in \bbR^{r_1\times \cdots \times r_K}$ by matrices $\mM_k \in\mathbb{R}^{p_k\times r_k}$ results in an order-$K$ $(p_1,\ldots,p_K)$-dimensional tensor $\tX$, denoted
 \[
@@
 We use $\text{Ave}(\cdot)$ to denote the operation of taking averages across elements and $\text{Mat}_k(\cdot)$ to denote the unfolding operation that reshapes the tensor along mode $k$ into a matrix.
-{For a symmetric tensor $\tX \in\mathbb{R}^{p\times \cdots \times p}$, we omit the subscript and use $\text{Mat}(\tX)\in\mathbb{R}^{p\times p^{K-1}}$ to denote the unfolding.}  For two sequences $\{a_p\}, \{b_p\}$, we denote $a_p\lesssim b_p$ or $a_p=\tO(b_p)$ if $\lim_{p\to\infty}a_p /b_p\leq c$ for some constant $c\geq 0$, $a_p=o(b_p)$ if $\lim_{p\to\infty}a_p/b_p =0$, and $a_p = \Omega(b_p)$ if both $b_p \lesssim a_p$ and $a_p\lesssim b_p$.
+{For a symmetric tensor $\tX \in\mathbb{R}^{p\times \cdots \times p}$, we omit the subscript and use $\text{Mat}(\tX)\in\mathbb{R}^{p\times p^{K-1}}$ to denote the unfolding.}  For two sequences $\{a_p\}, \{b_p\}$, we denote $a_p\lesssim b_p$ or $a_p=\tO(b_p)$ if $\lim_{p\to\infty}a_p /b_p\leq c${, $a_p \gtrsim b_p$ or $a_p = \Omega(b_p)$ if $\lim_{p\to\infty}a_p /b_p\geq c$, for some constant $c > 0$,}  $a_p=o(b_p)$ if $\lim_{p\to\infty}a_p/b_p =0$, and $a_p \asymp b_p$ if both $b_p \lesssim a_p$ and $a_p\lesssim b_p$.
 Throughout the paper, we use the terms ``community'' and ``clusters'' exchangeably.
@@
 
-Suppose we have an order-$K$ data tensor $\tY \in \bbR^{p \times \cdots \times p}$. %; the extension to general case with asymmetric mean tensor is provided in Section~\ref{subsec:exten}.} 
+Suppose that we have an order-$K$ data tensor $\tY \in \bbR^{p \times \cdots \times p}$. %; the extension to general case with asymmetric mean tensor is provided in Section~\ref{subsec:exten}.} 
 %we focus on symmetric tensors in this section; the extension to general asymmetric tensors is provided in Section~\ref{subsec:exten}. 
-Assume there exist $r \geq 2$  disjoint communities among the $p$ nodes.
+Assume that there exist $r \geq 1$  disjoint communities among the $p$ nodes.
 We represent the community assignment by a function $z \colon [p]\mapsto[r]$, where $z(i) = a$ for $i$-th node that belongs to the $a$-th community.
@@
 We consider the order-$K$ dTBM~\citep{ghoshdastidar2017consistency,ke2019community},
-\small
 \begin{equation}\label{eq:model_margin}
@@
 In general, we allow asymmetric mean tensors with $\{(\mTheta_k, \mM_k, z_k)\}_{k=1}^K$, one for each mode.
-The extension can be found in Appendix II.
+The extension can be found in Appendix B.
 
@@
 
-Assume there exist $r$ disjoint communities among $p$ nodes, and the connection probabilities depend on the community assignments and node-specific parameters. 
+Assume that there exist $r$ disjoint communities among $p$ nodes, and the connection probabilities depend on the community assignments and node-specific parameters. 
 Then, the equation~\eqref{eq:model_tensor} models $\bbE \tY$ with unknown degree heterogeneity $\mtheta$ and sub-Gaussianity parameter $\sigma^2 = 1/4$.
@@
 
-Assume there exist $r$ disjoint communities among $p$ nodes and $r_l$ disjoint communities among the $L$ layers.
-The multi-layer network community detection is modeled by the general asymmetric asymmetric dTBM model~\eqref{eq:model_tensor}
+Assume that there exist $r$ disjoint communities among $p$ nodes and $r_l$ disjoint communities among the $L$ layers.
+The multi-layer network community detection is modeled by the general asymmetric dTBM model~\eqref{eq:model_tensor}
 \begin{equation}
@@
 \paragraph{Gaussian higher-order clustering} Datasets in various fields such as medical image, genetics, and computer science are formulated as Gaussian tensors.
-One typical example is the multi-tissue gene expression dataset, which records the different gene expression in different individuals and different tissues.
+One typical example is the multi-tissue gene expression dataset, which records different gene expressions in different individuals and different tissues.
 The dataset, denoted as $\tY \in \bbR^{p \times n \times t}$, consists of the expression data for $p$ genes of $n$ individuals in $t$ tissues.
 
-Assume there exist $r_1, r_2, r_3$ disjoint clusters for $p$ genes, $n$ individuals, and $t$ tissues, respectively. We apply the general asymmetric dTBM model~\eqref{eq:model_tensor} 
+Assume that there exist $r_1, r_2, r_3$ disjoint clusters for $p$ genes, $n$ individuals, and $t$ tissues, respectively. We apply the general asymmetric dTBM model~\eqref{eq:model_tensor} 
 \begin{equation}
@@
 \end{equation}
-where $\{(\mtheta_k, \mM_k)\}_{k=1}^3$ represents the heterogeneity and membership for genes, individuals, and tissues.
+where $\{(\mtheta_k, \mM_k)\}_{k=1}^3$ represents the degree heterogeneity and membership for genes, individuals, and tissues.
 
@@
 We use $\tP$ to denote the following parameter space for $(z,\tS,\mtheta)$,
+
+\vspace{-0.5cm}
 \small
@@
 Note that the positivity constraint can be achieved without sacrificing model flexibility, by using a slightly larger dimension of $\tS$ in the factorization~\eqref{eq:model_tensor}; see Example~\ref{ex:positive} below.
-Second, recall that the quantity $|z^{-1}(a)|$ denotes the number of nodes in $a$-th community.
-The constants $c_1, c_2$ in the $|z^{-1}(a)|$ bound assume the roughly balanced size across $r$ communities.
+Second, recall that the quantity $|z^{-1}(a)|$ denotes the number of nodes in the $a$-th community.
+The constants $c_1, c_2$ in the $|z^{-1}(a)|$ bounds assume the roughly balanced size across $r$ communities.
 {Third, the constant $c_3$ requires that all slides in $\tS$ have non-degenerate norm.
-Particularly, the lower bound $c_3$ excludes the no purely zero slide case to avoid trivial non-identifiability of model \eqref{eq:model_tensor}; see Example~\ref{example:c3} below.
-The upper bound $c_4$ is a technical constraint to avoid the slides with unbounded norm as dimension grows; in practice, the constraint $\max_{a:} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large $c_4 \geq \onormSize{}{\tY}_F$.}
+Particularly, the lower bound $c_3$ excludes the purely zero slide to avoid trivial non-identifiability of model \eqref{eq:model_tensor}; see Example~\ref{example:c3} below.
+The upper bound $c_4$ is a technical constraint to avoid the slides with diverging norm as dimension grows.}
+%; in practice, the constraint $\max_{a:} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large $c_4 \geq \onormSize{}{\tY}_F$.
 %{ Third, the constant $c_3$ requires no purely zero slide in $\tS$, and thus the model~\eqref{eq:model_tensor} is not trivially non-identifiable; see Example~\ref{example:c3} below.
@@
 \begin{example}[Positivity of degree parameters]~\label{ex:positive}
-Here we provide an example to show the positivity constraints on $\mtheta$ incurs no loss on the model flexibility. 
+Here we provide an example to show the positivity constraint on $\mtheta$ incurs no loss on the model flexibility. 
 Consider an order-3 dTBM with core tensor $\tS = 1$ and degree $\mtheta = (1,1,-1, -1)^T$. We have the mean tensor 
@@
 \end{equation}
-where $\mTheta' = \text{diag}(\mtheta') = \text{diag}(1,1, 1, 1)$, the core tensor $\tS' \in \bbR^{2 \times 2 \times 2}$ has mode-3 slices, and the membership matrix $\mM'\in\{0,1\}^{4\times 2}$ defines the clustering $z'\colon[4]\to[2]$,
+where $\mTheta' = \text{diag}(\mtheta') = \text{diag}(1,1, 1, 1)$, the core tensor $\tS' \in \bbR^{2 \times 2 \times 2}$ has following mode-3 slices, and the membership matrix $\mM'\in\{0,1\}^{4\times 2}$ defines the clustering $z'\colon[4]\to[2]$; i.e.,
 \begin{equation}
@@
  The triplet $(z',\tS', \mtheta')$ lies in our parameter space~\eqref{eq:family}.
-In general, we can always reparameterize a block-$r$ tensor with mixed-signed degree using a block-$2r$ tensor with positive-signed degree.
+In general, we can always reparameterize an $r$-block tensor with mixed-signed degree using a $2r$-block tensor with positive-signed degree.
 Since we assume $r=\tO(1)$ throughout the paper, the splitting does not affect the error rates of our interest.
@@
     1 & -1
-    \end{pmatrix}$ and mean tensor 
+    \end{pmatrix}$ degree matrices $ \mTheta_1 =\mTheta_2= \text{diag}(1,1,1,1)$, and mean tensor 
     \begin{equation}
@@
         0 & 1 \\
-        \end{bmatrix}
-    \end{equation}
-    and $ \mTheta_1 =\mTheta_2= \text{diag}(1,1,1,1)$,
+        \end{bmatrix}.
+    \end{equation}
 Replacing $\mTheta_1$ by $\mTheta'_1 = (3/2, 1/2, 1,1)$ leads to the same mean tensor $\tX$. 
@@
 
-\begin{assumption}[Angle gap] \label{assmp:min_gap}Let $\mS = \text{Mat}(\tS)$. Assume the minimal gap between normalized rows of $\mS$ is bounded away from zero; i.e.,
+\begin{assumption}[Angle gap] \label{assmp:min_gap}Let $\mS = \text{Mat}(\tS)$. Assume that the minimal gap between normalized rows of $\mS$ is bounded away from zero; i.e.,
 \begin{equation}\label{eq:minimal_gap}
@@
 
-Our first main result is the following theorem showing the sufficiency and necessarily of the angle gap separation condition for the parameter identifiability under dTBM.
+Our first main result is the following theorem showing the sufficiency and necessity of the angle gap separation condition for the parameter identifiability under dTBM.
 
@@
 Intuitively, a larger SNR, or equivalently a larger $\gamma$, benefits the clustering in the presence of noise.
-
 With quantification~\eqref{eq:gamma}, we consider the following parameter space,
@@
 \end{equation}
-The $1$-block dTBM does not belong to the space $\tP(\gamma)$ when $\gamma < 0$ by Assumption~\ref{assmp:min_gap}.
+The $1$-block dTBM does not belong to the space $\tP(\gamma)$ when $\gamma < 0$, due to the convention in Assumption~\ref{assmp:min_gap}.
 Our goal is to characterize the clustering accuracy with respect to $\gamma$ under the space $\tP(\gamma)$.
@@
 
-The following lemma provides the rational of balanced degree assumption.
+The following lemma provides the rationale of balanced degree assumption.
 We show the close relation between angle gaps in the mean tensor $\tX$ and the core tensor $\tS$ under balanced degree heterogeneity.
@@
 \end{equation}
-where $ \mX = \mat_1(\tX)$.
+where $ \mX = \mat(\tX)$.
 Take $ a = p^{-1/4}$ in the model setup~\eqref{eq:modelexample}. We have 
@@
 
-{Based on our theory in later Sections, the dTBM is impossible to solve when $\Delta^2_{\mX} \lesssim p^{-1}$
+{Based on the Theorem~\ref{thm:stats} in Section~\ref{sec:limits}, the dTBM is impossible to solve when $\Delta^2_{\mX} \lesssim p^{-1}$ even though $\Delta_{\min}^2 \asymp p^{-1/2}$
 %but can be effectively solved when $\Delta^2_{\min} \gtrsim p^{-1/2}$; 
@@
 Similar degree regulations are not rare in literature.
-In higher-order tensor model \citep{ke2019community}, the degree assumption $\max_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}} \leq $ $ C \min_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}$ is made to ensure balancedness across communities.
+In higher-order tensor model \citep{ke2019community}, the degree assumption $\max_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}} \leq $ $ C \min_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}$ is made to ensure degree balance across communities.
 %Since they have no $\ell_1$ constraint on $\mtheta$, their balance assumption is equivalent to ours~\eqref{eq:degree} multiplying some constants to the degrees.
@@
 where $\pi: [r] \mapsto [r]$ is a permutation of cluster labels, $\circ$ denotes the composition operation, and $\Pi$ denotes the collection of all possible permutations.
-The infinitum over all permutations accounts for the ambiguity in cluster label permutation.
-
-In Sections~\ref{sec:statlimit} and \ref{sec:complimit}, we provide the phase transition of $\ell (\hat z, z)$ for general Gaussian dTBMs~\eqref{eq:model_margin} without symmetric assumptions.
-For general (asymmetric) Gaussian dTBMs, we assume Gaussian noise $\tE(i_1,\ldots,i_K)\stackrel{\text{i.i.d.}}{\sim} N(0,\sigma^2)$, and we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode.
+The infimum over all permutations accounts for the ambiguity in cluster label permutation.
+
+In Sections~\ref{sec:statlimit} and \ref{sec:complimit}, we provide the phase transition of $\ell (\hat z, z)$ for general Gaussian dTBMs~\eqref{eq:model_tensor} without symmetric assumptions.
+For general (asymmetric) Gaussian dTBMs, we assume Gaussian noise $\tE(i_1,\ldots,i_K)\stackrel{\text{i.i.d.}}{\sim} N(0,\sigma^2)$, and we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $\{z_k\}_{k\in[K]}$, one for each mode.
 For notational simplicity, we still use $z$ and $\tP(\gamma)$ for this general (asymmetric) model.
@@
 The statistical critical value means the SNR required for solving dTBMs with \emph{unlimited computational cost.}   Our following result shows the minimax lower bound for exact recovery and the matching upper bound for maximum likelihood estimator (MLE).
-We consider the Gaussian MLE, denoted as $(\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}})$, over the estimation space $\tP$, where
+We consider the Gaussian MLE, denoted as $(\hat z_{\text{MLE}},\hat \tS_{\text{MLE}},  \hat \mtheta_{\text{MLE}})$, over the estimation space $\tP$, where
 \begin{equation}\label{eq:mle}
-    (\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}}) = \argmin_{ (z, \mtheta, \tS) \in \tP} \onormSize{}{\tY - \tX(z,\mtheta, \tS)}_F^2.
-\end{equation}
-
-\begin{thm}[Statistical critical value]\label{thm:stats} Consider general Gaussian dTBMs with parameter space $\tP(\gamma)$ with $K\geq 1$. Then, we have the following statistical phase transition.
+    (\hat z_{\text{MLE}}, \hat \tS_{\text{MLE}}, \hat \mtheta_{\text{MLE}}) = \argmin_{ (z,\tS,\mtheta) \in \tP} \onormSize{}{\tY - \tX(z,\tS,\mtheta)}_F^2.
+\end{equation}
+
+\begin{thm}[Statistical critical value]\label{thm:stats} Consider general Gaussian dTBMs with parameter space $\tP(\gamma)$ and $K\geq 2$. Then, we have the following statistical phase transition.
 
@@
     \begin{equation}
-        \liminf_{p \rightarrow \infty} \inf_{\hat z_{\rm stat}} \sup_{(z, \mtheta, \tS) \in \tP'(\gamma')}  \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
+        \liminf_{p \rightarrow \infty} \inf_{\hat z_{\rm stat}} \sup_{(z, \tS, \mtheta) \in \tP'(\gamma')}  \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
     \end{equation}
@@
 \item \textbf{MLE achievability.} 
-Suppose the signal exponent satisfies $\gamma >-(K-1)+c_0$ for an arbitrary constant {$c_0>0$}.
-Furthermore, assume {$\mtheta$} is balanced and $\min_{i\in[p]}\theta(i)\geq c$ from some constant $c>0$.
-Then, {when $p \rightarrow \infty$, fixed $r \geq 1$, }the MLE in~\eqref{eq:mle} achieves exact recovery in high probability; that is,
+Suppose that the signal exponent satisfies $\gamma >-(K-1)+c_0$ for an arbitrary constant {$c_0>0$}.
+Furthermore, assume that {$\mtheta$} is balanced and $\min_{i\in[p]}\theta(i)\geq c$ from some constant $c>0$.
+Then, {when $p \rightarrow \infty$, for fixed $r \geq 1$, }the MLE in~\eqref{eq:mle} achieves exact recovery in high probability; that is,
 %Suppose the signal exponent satisfies $\gamma >-(K-1)+c$ for an arbitrary constant $c>0$.
@@
 
-The proofs for the two parts in Theorem~\ref{thm:stats} are in the Appendix II, Section~\ref{sec:statprove1} and Section~\ref{sec:statprove2}, respectively.
+The proofs for the two parts in Theorem~\ref{thm:stats} are in the Appendix B, Section~\ref{sec:statprove1} and Section~\ref{sec:statprove2}, respectively.
 The first part of Theorem~\ref{thm:stats} demonstrates impossibility of exact recovery whenever the core tensor $\tS$ satisfies SNR condition~\eqref{eq:gamma} with exponent $\gamma < -(K-1)$.
@@
 \end{equation}
-Such worst case impossibility is frequently studied in related works \citep{han2020exact, gao2018community} while our lower bound~\eqref{eq:minminmax} provides a stronger impossibility statement for arbitrary core tensor with weak signal.
-
-The second part of Theorem~\ref{thm:stats} shows the exact recovery of MLE when the $\gamma>-(K-1)+c$ for an arbitrary constant $c>0$.
+Such worst case impossibility is studied in related works \citep{han2020exact, gao2018community} while our lower bound~\eqref{eq:minminmax} provides a stronger impossibility statement for arbitrary core tensors with weak signals.
+The second part of Theorem~\ref{thm:stats} shows the exact recovery of MLE when $\gamma>-(K-1)+c_0$ for an arbitrary constant $c_0>0$.
 Combining the impossibility and achievability results, we conclude that the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM with respect to our SNR.
@@
 \subsection{Computational critical value}\label{sec:complimit}
-In this section, we derive the computational critical value of dTBMs.
-The computational critical value means the minimal SNR required for exactly recovery with \emph{polynomial-time} computational cost.
+%In this section, we derive the computational critical value of dTBMs. 
+The computational critical value means the minimal SNR required for exact recovery with \emph{polynomial-time} computational cost.
 An important ingredient to establish the computational limits is the \emph{hypergraphic planted clique (HPC) conjecture} \citep{zhang2018tensor, brennan2020reducibility}.
@@
 
-\begin{defn}[Hypergraphic planted clique (HPC) detection]~\label{def:HPC} Consider an order-$K$ hypergraph $H = (V,E)$ where $V = [p]$ collects vertices and $E$ collects all the order-$K$ edges.
+\begin{defn}[Hypergraphic planted clique (HPC) detection]~\label{def:HPC}Consider an order-$K$ hypergraph $H = (V,E)$ where $V = [p]$ collects vertices and $E$ collects all the order-$K$ edges.
 Let $\tH_k(p, 1/2)$ denote the Erd\H{o}s-R\'{e}nyi $K$-hypergraph where the edge $(i_1,\ldots, i_K)$ belongs to $E$ with probability $1/2$.
@@
 
-\begin{conjecture}[HPC conjecture]\label{hypo:HPC} Consider the HPC detection problem in Definition~\ref{def:HPC}.
-Suppose the sequence $\{\kappa_p\}$ such that $\limsup_{p \rightarrow \infty} \log \kappa_p/ \log \sqrt{p} \leq (1 - \tau)$.
+\begin{conjecture}[HPC conjecture]\label{hypo:HPC} Consider the HPC detection problem in Definition~\ref{def:HPC} with $K \geq 2$.
+Suppose the sequence $\{\kappa_p\}$ such that $\limsup_{p \rightarrow \infty} \log \kappa_p/ \log \sqrt{p} \leq (1 - \tau)$ for any $\tau > 0$.
 Then, for every sequence of polynomial-time test $\{ \varphi_p\}: H \mapsto \{0,1\}$ we have
 \begin{equation}
-    \liminf_{p \rightarrow \infty} \bbP_{H_0} \of{ \varphi_p(H) =1 } +  \bbP_{H_1} \of{ \varphi_p(H) =0} \geq \frac{1}{2}.
+    \liminf_{p \rightarrow \infty} \bbP_{H_0} \of{ \varphi_p(H) =1 } +  \bbP_{H_1} \of{ \varphi_p(H) =0} > \frac{1}{2}.
 \end{equation}
@@
 \end{align}
-\item \textbf{Polynomial-time algorithm achievability.} Suppose the parameter space satisfies $\gamma  >  -K/2+c_0$ for an arbitrary constant {$c_0>0$}.
-Furthermore, assume {fixed $r \geq 1, K \geq 2$,} the degree is balanced, lower bounded in that $\min_{i\in[p]}\theta_i\geq c$ for some constant $c>0$, and satisfies the local linear stability in Definition~\ref{def:stable} {in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \geq \check{C}\log^{-1}p $ with some positive constant $\check{C}$.}
+\item \textbf{Polynomial-time algorithm achievability.} Suppose that we have {fixed $r \geq 1, K \geq 2$}, and the signal exponent satisfies $\gamma  >  -K/2+c_0$ for an arbitrary constant {$c_0>0$}.
+Furthermore, assume that the degree $\mtheta$ is balanced, lower bounded in that $\min_{i\in[p]}\theta_i\geq c$ for some constant $c>0$, and satisfies the locally linear stability in Definition~\ref{def:stable} {in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \gtrsim \log^{-1}p $.}
+%with some positive constant $\check{C}$.
 %in the neighborhood $\tN(z,\varepsilon)$ for $\varepsilon \geq \log^{-1}(p)$. 
@@
 
-The proofs for the two parts in Theorem~\ref{thm:comp} are in the Appendix II, Section~\ref{sec:compprove1} and Section~\ref{sec:statprove2}, respectively.
-The first part of Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$, and the second part shows the existence of such algorithm when {$\gamma > -K/2+c_0$  for an arbitrary constant $c_0$ and} under extra technical assumptions.
+The proofs for the two parts in Theorem~\ref{thm:comp} are in the Appendix B, Section~\ref{sec:compprove1} and Section~\ref{sec:statprove2}, respectively.
+The first part of Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$, and the second part shows the existence of such algorithm when {$\gamma > -K/2+c_0$  for an arbitrary constant $c_0 > 0$} under extra technical assumptions.
 In Section~\ref{sec:alg}, we will present an efficient polynomial-time algorithm in this setting.
@@
 \begin{rmk}[Statistical-computational gaps]
-Now, we have established the phase transition of exact clustering under order-$K$ dTBM by combing Theorems~\ref{thm:stats} and \ref{thm:comp}.
+Now, we have established the phase transition of exact clustering under order-$K$ dTBM by combining Theorems~\ref{thm:stats} and \ref{thm:comp}.
 Figure~\ref{fig:phase_axis} summarizes our results of critical SNRs when $K \geq 2$.
@@
 \cite{lu2016statistical} implies that polynomial-time algorithms are able to achieve the statistical minimax lower bound in GMM.
-Therefore, we conclude that the statistical-to-computational gap emerges only for higher-order tensors with $K \geq 3$.
+Therefore, we conclude that the statistical-computational gap emerges only for higher-order tensors with $K \geq 3$.
 The result reveals the intrinsic distinctions among (vector) one-dimensional clustering, (matrix) biclustering, and (tensor) higher-order clustering.
@@
 See Figure~\ref{fig:demo} for illustration.
-The global step finds the basin of attraction with polynomial miclustering error, whereas the local iterations improve the initial clustering to exact recovery.
+The global step finds the basin of attraction with polynomial misclustering error, whereas the local iterations improve the initial clustering to exact recovery.
 Both steps are critical to obtain a satisfactory algorithm output.
 In what follows, we first use the symmetric tensor as a working example to describe the algorithm procedures to gain insight.
-{Our theoretical analysis focuses on the symmetric mean tensor with independent sub-Gaussian noises}
+{Our theoretical analysis focuses on dTBMs with symmetric mean tensor and independent sub-Gaussian noises}
 %i.i.d.\ sub-Gaussian noise
-such as Gaussian and uniform observations.   {The extensions for Bernoulli observations, asymmetric mean tensors, and other practical issues are in Sections~\ref{subsec:ber} and \ref{subsec:exten}.}
+such as Gaussian and uniform observations.   {The extensions for Bernoulli observations and other practical issues are in Sections~\ref{subsec:ber} and \ref{subsec:exten}.}
 
@@
 \centering
-\includegraphics[width=\columnwidth]{alg_demo.pdf}
+\includegraphics[width=.9\columnwidth]{alg_demo.pdf}
 \caption{Illustration of our global-to-local algorithm.}\label{fig:demo}
@@
 \end{equation}
-This implies that, all node $i$ belonging to $a$-th community (i.e., $z(i)=a$) share the same normalized mean vector $\theta(i)^{-1} \mX_{i:}$, and vice versa.
+This implies that, all node $i$ belonging to the $a$-th community (i.e., $z(i)=a$) share the same normalized mean vector $\theta(i)^{-1} \mX_{i:}$, and vice versa.
 Intuitively, one can apply $k$-means clustering to the vectors $\{ \theta(i)^{-1} \mX_{i:} \}_{i\in[p]}$, which leads to main idea of our Sub-algorithm~\hyperref[alg:main]{1}.
 
-Specifically, our initialization consists of denoising step and clustering step.
+Specifically, our initialization consists of the denoising step and the clustering step.
 The denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) estimates $\tX$ from $\tY$ by a double projection spectral method.
@@
 The choice of weights is to bound the $k$-means objective function by the Frobenius-norm accuracy of $\hat \tX$.
-Unlike existing clustering algorithm~\citep{ke2019community}, we apply the clustering on the unfolded tensor $\hat \mX$ rather than on the factors $\hat \mU_1$.
+Unlike existing clustering algorithm~\citep{ke2019community}, we apply the clustering on the unfolded tensor $\hat \mX$ rather than on the factors $\hat \mU_k$.
 This strategy relaxes the singular-value gap condition~\citep{gao2018community, han2020exact}.
@@
 \State {
-  Compute factor matrix $ \mU_{\text{pre},k} = \text{SVD}_{r} (\text{Mat}_k(\tY)), k \in [K]$ and the $(K-1)$-mode projection 
-\begin{equation}
-    \tX_{\text{pre},k} = \tY \times_1   \mU_{\text{pre},1} \mU_{\text{pre},1}^T \times_2 \cdots  \times_{k-1} \mU_{\text{pre},k-1} \mU_{\text{pre},k-1}^T \times_{k+1} \mU_{\text{pre},k-1} \mU_{\text{pre},k+1}^T\times_{K+1}  \times_K \mU_{\text{pre},K} \mU_{\text{pre}, K}^T.
+  Compute factor matrices $ \mU_{\text{pre},k} = \text{SVD}_{r} (\text{Mat}_k(\tY)), k \in [K]$ and the $(K-1)$-mode projections 
+\begin{equation}
+    \tX_{\text{pre},k} = \tY \times_1   \mU_{\text{pre},1} \mU_{\text{pre},1}^T \times_2 \cdots  \times_{k-1} \mU_{\text{pre},k-1} \mU_{\text{pre},k-1}^T \times_{k+1} \mU_{\text{pre},k+1} \mU_{\text{pre},k+1}^T\times_{k+2} \cdots  \times_K \mU_{\text{pre},K} \mU_{\text{pre}, K}^T.
 \end{equation}
@@
 \State {
-
-Compute factor matrix $\hat \mU_k = \text{SVD}_{r}(\text{Mat}_k(\tX_{\text{pre},k})), k \in [K]$ and denoised tensor
+Compute factor matrices $\hat \mU_k = \text{SVD}_{r}(\text{Mat}_k(\tX_{\text{pre},k})), k \in [K]$ and the denoised tensor
 \begin{equation}
@@
 }
-\State { Solve the clustering $\hat z_k \colon [p]\to[r]$ and centroids $ (\hat \mx_j)_{j\in[r_k]}$ using weighted $k$-means, such that}
+\State { Solve the clustering $\hat z_k \colon [p]\to[r]$ and centroids $\{\hat \mx_j\}_{j\in[r]}$ using weighted $k$-means, such that}
 \begin{align}
@@
     \leq 
-    \eta \min_{\substack{\bar \mx_j, j\in[r], \bar z_k(i),i\in S_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\hat z_k(i)}}^2.
+    \eta \min_{\substack{\bar \mx_j, j\in[r], \bar z_k(i),i\in S_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\bar z_k(i)}}^2.
 \end{align}
@@
 \For{ $k \in [K]$}
-\State {   Calculate reduced tensor $\tY^{\text{d}}_k \in \bbR^{r \times r \times p \times r \times \cdots \times r}$ via
+\State {   Calculate the reduced tensor $\tY^{\text{d}}_k \in \bbR^{r \times \cdots \times r \times p \times r \times \cdots \times r}$ via
 \begin{equation}
@@
 
-\State {   Let $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. For all $i \in J_0^c$ update the cluster assignment by
-\begin{equation}
-    z(i)^{(t+1)}_k = \argmax_{a \in [r]} \cos \left( \mY^{\text{d}}_{k,i:},\ \mS^{(t)}_{a:} \right).
+\State {   Let $\mS^{(t)}_k = \text{Mat}_k(\tS^{(t)})$. For all $i \in J_0^c$, update the cluster assignment by
+\begin{equation}
+    z(i)^{(t+1)}_k = \argmax_{a \in [r]} \cos \left( \mY^{\text{d}}_{k,i:},\ \mS^{(t)}_{k, a:} \right).
 \end{equation}}
@@
 
-\OUTPUT {    Estimated clustering $z^{(T)}_k  \in [r]^{p}, k \in [K]$.}
+\OUTPUT {    Estimated clustering $z^{(T)}_k: [p] \mapsto [r], k \in [K]$.}
 
@@
 
-\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with {fixed $r \geq 1$, $K \geq 2$}, i.i.d.\ noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
-Assume $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$ and let $\Delta_{\mX}$ denote the minimal gap in mean tensor defined in~\eqref{eq:delta_x}.
-Let $ z^{(0)}_k$ denote the output of Sub-algorithm~\hyperref[alg:main]{1}.
+\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with {fixed $r \geq 1$, $K \geq 2$}, i.i.d.\ noise under the parameter space $\tP$, and Assumption~\ref{assmp:min_gap}.
+Assume $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$.
+Let $\Delta_{\mX}$ denote the minimal gap in mean tensor defined in~\eqref{eq:delta_x}, and let $ z^{(0)}_k$ denote the output of Sub-algorithm~\hyperref[alg:main]{1}.
 With probability going to 1, {as $p \rightarrow \infty$}, we have
@@
 \end{equation}
-Further, assume $\mtheta$ is balanced as~\eqref{eq:degree}. We have
+Further, assume that $\mtheta$ is balanced as~\eqref{eq:degree}. We have
 \begin{equation}\label{eq:ini}
@@
 \begin{rmk}[Comparison to previous results] For fixed SNR, our initialization error rate with $K=2$ agrees with the initialization error rate $\tO(p^{-1})$ in matrix models~\citep{gao2018community}.
-Furthermore, in the special case of non-degree TBMs with $\theta_1=\cdots=\theta_p=1$, we achieve the same initial misclustering error $\tO(p^{-K/2})$ as in non-degree models~\citep{han2020exact}.
+Furthermore, in the special case of non-degree TBMs with $\mtheta = \mathbf{1}_p$, we achieve the same initial misclustering error $\tO(p^{-K/2})$ as in non-degree models~\citep{han2020exact}.
 Theorem~\ref{thm:initial} implies the advantage of our algorithm in achieving both accuracy and model flexibility.
@@
 
-\begin{rmk}[Failure of conventional tensor HOSVD] If we use conventional HOSVD for tensor denoising; that is, we use $\mU_{\text{pre},k}$ in place of $\hat \mU,k$ in line 2, then the misclustering rate becomes $\tO(p^{-1})$ for all $K\geq 2$.
+\begin{rmk}[Failure of conventional tensor HOSVD] If we use conventional HOSVD for tensor denoising; that is, we use $\mU_{\text{pre},k}$ in place of $\hat \mU_k$ in line 2, then the misclustering rate becomes $\tO(p^{-1})$ for all $K\geq 2$.
 This rate is substantially worse than our current rate~\eqref{eq:ini}.
@@
 The classical Euclidean-distance based clustering~\citep{han2020exact} fails to recover $z$ in the presence of degree heterogeneity, even under noiseless case.
-In contrast, the proposed angle-based $k$-means achieves accurate recovery without explicit estimation of $\mtheta$.
-
-Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as in angle-based $k$-means.
+In contrast, the proposed angle-based $k$-means algorithm
+achieves accurate recovery without the explicit estimation of $\mtheta$.
+
+Our Sub-algorithm~\hyperref[alg:main]{2} shares the same spirit as in the angle-based $k$-means.
 We still take the order-3 tensor for illustration.
 Specifically, Sub-algorithm~\hyperref[alg:main]{2} updates estimated core tensor and cluster assignment in each iteration.
-We use superscript $\cdot^{(t)}$ to denote the estimate from $t$-th iteration, where $t=1,\ldots.$ For core tensor, we consider the following update strategy
+We use superscript $\cdot^{(t)}$ to denote the estimate from the $t$-th iteration, where $t=1, 2, \ldots.$ For core tensor, we consider the following update strategy
  \[
@@
 \]
-Intuitively, $\tS^{(t)}$ becomes closer to the true core $\tS$ as $z^{(t)}$ is more precise.
-For cluster assignment, we first aggregate the slices of $\tY$ and obtain a reduced tensor $\tY^{\text{d}}\in \bbR^{p \times r \times r}$ with given $z^{(t)}$, where
+Intuitively, $\tS^{(t)}$ becomes closer to the true core $\tS$ as $z^{(t)}_k$ is more precise.
+For cluster assignment, we first aggregate the slices of $\tY$ and obtain the reduced tensor $\tY^{\text{d}}_1 \in \bbR^{p \times r \times r}$ on the first mode with given $z^{(t)}_k$, where
 \[
@@
 \]
+Similarly, we also obtain $\tY^{\text{d}}_2, \tY^{\text{d}}_3$.
 We use $\mY_k^d$ and $ \mS^{(t)}_k$ to denote the $\text{Mat}_k(\tY^{\text{d}})$ and $\text{Mat}_k(\tS^{(t)})$.
@@
 provided that $\mS^{(t)}_{k,a:}$ is a non-zero vector.
-Otherwise, if $\mS^{(t)}_{k,a:}$ is a zero vector, then we make the convention to assign $z^{(t+1)}_k(i)$ randomly in $[p]$.
+Otherwise, if $\mS^{(t)}_{k,a:}$ is a zero vector, then we make the convention to assign $z^{(t+1)}_k(i)$ randomly in $[r]$.
 Full procedures for our angle-based iteration are described in Sub-algorithm~\hyperref[alg:main]{2}.
@@
 The local stability holds trivially for $\varepsilon=0$ based on the construction of parameter space~\eqref{eq:family}.
-The condition~\eqref{eq:local} controls the impact of node degree to the $\mp_{\theta}(\cdot)$ with respect to the misclassification rate $\varepsilon$ and angle gap.
+The condition~\eqref{eq:local} controls the impact of node degree to the $\mp_{\theta}(\cdot)$ with respect to the misclustering rate $\varepsilon$ and angle gap.
 Intuitively, the condition~\eqref{eq:local} controls the skewness of degree so that the angle between raw cluster size and degree-weighted cluster size is well controlled.
@@
 \begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the general sub-Gaussian dTBM with {fixed $r \geq 1$, $K \geq 2$}, independent noise under the parameter space $\tP$, and Assumption~\ref{assmp:min_gap}.
-{Assume the local linear stability of degree holds in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \geq \check{C}\log^{-1}p $ with some positive constant $\check{C}$.}
+{Assume that the locally linear stability of degree holds in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \gtrsim \log^{-1}p $.}
 %Assume the local linear stability of degree holds in the neighborhood $\tN(z, \epsilon)$ for $\epsilon \geq \log^{-1}(p)$.
-Let $\{z^{(0)}_k\}_{k=1}^K$ be the initialization for Sub-Algorithm~\hyperref[alg:main]{2} and $z^{(t)}_k$ be the $t$-th iteration output on $k$-th mode.
-Suppose $\min_{i \in [p]}\theta(i) \geq c $ for some {constant} $c > 0$, the $\text{SNR} \geq \tilde C p^{-(K-1)}\log p$ for some sufficiently large constant $\tilde C$, and the initialization satisfies
+Let $\{z^{(0)}_k\}_{k=1}^K$ be the initialization for Sub-algorithm~\hyperref[alg:main]{2} and $z^{(t)}_k$ be the $t$-th iteration output on the $k$-th mode.
+Suppose $\min_{i \in [p]}\theta(i) \geq c $ for some {constant} $c > 0$, the $\text{SNR} \geq \tilde C p^{-(K-1)}\log p$ for some sufficiently large positive constant $\tilde C$, and the initialization satisfies
 %Consider the general sub-Gaussian dTBM with independent noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
@@
  With probability going to 1 {as $p \rightarrow \infty$}, there exists a contraction parameter $\rho \in (0,1)$ such that
+
+ \vspace{-0.5cm}
  \small
@@
 Assume $\text{SNR} \gtrsim p^{-K/2} \log p$.
-Combining  Combining {all parameter assumptions and the results in} Theorems~\ref{thm:initial} and \ref{thm:refinement}, with probability going to 1 {as $p \rightarrow \infty$}, our estimate $z^{(T)}_k$ achieves exact recovery within polynomial iterations; more precisely,
+Combining {all parameter assumptions and the results in} Theorems~\ref{thm:initial} and \ref{thm:refinement}, with probability going to 1 {as $p \rightarrow \infty$}, our estimate $z^{(T)}_k$ achieves exact recovery within polynomial iterations; more precisely,
 \begin{equation}
@@
 Therefore, our combined algorithm is \textit{computationally efficient} as long as SNR $\gtrsim p^{-K/2} \log p$.
-Note that, ignoring the logarithmic term, the minimal SNR requirement, $p^{-K/2}$, coincides with the computational lower bound in Theorem~\ref{thm:comp}.
+Note that, ignoring the logarithmic term, the minimal SNR requirement, $p^{-K/2}$, coincides with the computational critical value in Theorem~\ref{thm:comp}.
 Therefore, our algorithm is optimal regarding the signal requirement and lies in the sharpest \emph{computationally efficient} regime in Figure~\ref{fig:phase_axis}.
@@
 The matrix $\Mat_{sq}(\tY)$ is asymmetric.
-We are able to interpret $\Mat_{sq}(\tY)$ as the adjacency matrix for a bipartite network with connections between two groups of node.
+We interpret $\Mat_{sq}(\tY)$ as the adjacency matrix for a bipartite network with connections between two groups of nodes.
 The two groups of nodes in the bipartite network have $p_1 \cdots p_{\floor{K/2}}$ and $ p_{\ceil{K/2}} \cdots p_K$ nodes, respectively.
@@
 \State Implement lines 3-5 of Sub-algorithm~\hyperref[alg:main]{1} with $\hat \tX$ replaced by $\hat \tX'$ in~\eqref{eq:matrixsvd}.
-\OUTPUT Initial clustering $z^{(0)} \leftarrow \hat z$.
+\OUTPUT Initial clustering $z^{(0)}_k \leftarrow \hat z_k, k \in [K]$.
 \end{algorithmic}
@@
 
-\begin{prop}[Error for Bernoulli initialization]\label{prop:ber} Consider the Bernoulli dTBM in the parameter space $\tP$ {with fixed $r \geq 1, K \geq 2$} and Assumption~\ref{assmp:min_gap} holds.
-Assume that $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$.
-Let $ z^{(0)}$ denote the output of Sub-algorithm~3.
+\begin{prop}[Error for Bernoulli initialization]\label{prop:ber} Consider the Bernoulli dTBM in the parameter space $\tP$ {with fixed $r \geq 1, K \geq 2$}.
+Assume that Assumption~\ref{assmp:min_gap} holds, $\mtheta$ is balanced, and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$.
+Let $ z^{(0)}_k$ denote the output of Sub-algorithm~3.
 With probability going to 1 {as $p \rightarrow \infty$}, we have
 \begin{equation}\label{eq:ini_b}
- \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}, \quad \text{and} \quad L(z^{(0)}, z) \lesssim  {\sigma^2 r^K p^{-\floor{K/2}}}.
+ \ell(z^{(0)}_k, z_k) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}, \quad \text{and} \quad L(z^{(0)}_k, z_k) \lesssim  {\sigma^2 r^K p^{-\floor{K/2}}}.
 \end{equation}
@@
 However, we are not able to infer the exact recovery of spectral method by our angle-base SNR condition.
-Consider an order-2 dTBM with $p > 2, \sigma^2 = 1$, $\mtheta = \mathbf{1}$, equal size assignment $|z^{-1}(a)| = p/r$ for all $ a \in [r]$, and core matrix $\mS = \mI_2$.
+Consider an order-2 dTBM with $p > 2, \sigma^2 = 1$, $\mtheta = \mathbf{1}_p$, equal size assignment $|z^{-1}(a)| = p/r$ for all $ a \in [r]$, and core matrix equal to the 2-dimensional identity matrix $\mS = \mI_2$.
 The singular gap under this setting is $\Delta_{\rm singular} = \min\{ \lambda_1 - \lambda_2, \lambda_2 \} = 0$, where $\lambda_1 \geq \lambda_2$ are singular values of $\mS$.
@@
 However, SVD-based methods are believed to fail in extremely sparse SBM due to the localization phenomenon in the singular vectors \citep{florescu2016spectral}.
-Inspired by \cite{florescu2016spectral}, we adapt the diagonal-deleted HOSVD (D-HOSVD) \citep{ke2019community} as the initialization in our higher-order clustering.
+Inspired by \cite{florescu2016spectral}, we adopt the diagonal-deleted HOSVD (D-HOSVD) \citep{ke2019community} as the initialization in our higher-order clustering.
 
@@
 The decaying mean tensor leads to a worse statistical error of order $\tO( -\alpha_p p^{K-1})$ on $\hat \tX$.
-The theoretical analyses for sparse binary dTBM and algorithms are left as future direction.
-Instead, we add numerical experiments to evaluate the robustness of our algorithm and the improvement of D-HOSVD initialization in the sparse dTBM; see Appendix I.
+The theoretical analyses for sparse binary dTBM and algorithms are left as future directions.
+Instead, we add numerical experiments to evaluate the robustness of our algorithm and the improvement of D-HOSVD initialization in the sparse dTBM; see Appendix A.
 
@@
 Specifically, the complexity of Sub-algorithm~\hyperref[alg:main]{1} is $\tO(K p^{K+1} + Krp^K)$, where the first term is contributed by the double projection and the calculation of $\hat \tX$, and the second term comes from normalization and the $k$-means.
-The cost of each update in Sub-algorithm~\hyperref[alg:main]{2} is $\tO(p^K + pr^K)$, where $p^K$ comes from the calculation of $\tS^{(t)}$ and $\tY^{\text{d}}$, and $pr^K$ comes from the normalization of $\tY^{\text{d}}$, the calculation of $\tS^{(t)}$, and the cluster assignment update in Step~10.
+The cost of each update in Sub-algorithm~\hyperref[alg:main]{2} is $\tO(p^K + pr^K)$, where $p^K$ comes from the calculation of $\tS^{(t)}$ and $\tY^{\text{d}}_k$, and $pr^K$ comes from the normalization of $\tY^{\text{d}}_k$, the calculation of $\tS^{(t)}$, and the cluster assignment update in Step~13.
 
@@
 \end{align}
-with $\hat \tX = \hat \tS(r) \times_1 \hat \mTheta(r) \hat \mM(r) \times_2 \cdots \times_K \hat \mTheta(r) \hat \mM(r),$ where the triplet $(\hat z(r), \hat \mS(r),\hat \mtheta(r))$ are estimated parameters with cluster number $r$, and $p_e (r)= r^K + p(\log r + 1) - r$ is the effective number of parameters.
+with $\hat \tX = \hat \tS(r) \times_1 \hat \mTheta(r) \hat \mM(r) \times_2 \cdots \times_K \hat \mTheta(r) \hat \mM(r),$ where the triplet $(\hat z(r), \hat \tS(r),\hat \mtheta(r))$ are estimated parameters with cluster number $r$, and $p_e (r)= r^K + p(\log r + 1) - r$ is the effective number of parameters.
 Note that we have added the argument $(r)$ to related quantities as functions of $r$.
@@
 
-We discuss the connections and difference between dTBM and TBM~\citep{han2020exact} from three aspects: signal notions, theoretical results, and algorithms.
+We discuss the connections and differences between dTBM and TBM~\citep{han2020exact} from three aspects: signal notions, theoretical results, and algorithms.
 Without loss of generality, let $\sigma^2=1$.
@@
     \end{align}
-By Lemma~\ref{lem:norm_diff} in the Appendix II, we have 
+By Lemma~\ref{lem:norm_diff} in the Appendix B, we have 
 \begin{equation}\label{eq:signalcompare}
@@
 \end{equation}
-The above inequality indicates the condition $\Delta_{\text{Euc}}^2 \leq p^{\gamma}$ is sufficient but not necessary for $\Delta_{\text{ang}}^2 \leq p^{\gamma}$.
+The above inequality indicates that the condition $\Delta_{\text{Euc}}^2 \leq p^{\gamma}$ is sufficient but not necessary for $\Delta_{\text{ang}}^2 \leq p^{\gamma}$.
 In fact, if we were to use $\Delta_{\text{Euc}}^2$ for both models, then the phase transition of dTBM can be arbitrarily worse than that for TBM.
@@
 
-\begin{example}[Comparison with Euclidean-based signal notion] \label{example:euc_alg} Consider a biclustering model with $\mtheta=1$ and an order-2 core matrix 
+\begin{example}[Comparison with Euclidean-based signal notion] \label{example:euc_alg} Consider a biclustering model with $\mtheta=\mathbf{1}$ and an order-2 core matrix 
 \begin{equation}
@@
  Last, we discuss the statistical impossibility statements.
-Our Theorem~\ref{thm:stats} implies the statistical impossibility whenever the core tensor $\tS$ leads to an angle-based SNR below the critical threshold.
-The Theorem 6 in \cite{han2020exact} implies the worst case statistical impossibility for a particular core tensor $\tS$ with Euclidean-based SNR below the statistical limit.
+Our Theorem~\ref{thm:stats} implies the statistical impossibility whenever the core tensor $\tS$ leads to an angle-based SNR below the critical value, while, Theorem 6 in \cite{han2020exact} implies the worst case statistical impossibility for a particular core tensor $\tS$ with Euclidean-based SNR below the statistical limit.
 Hence, our Theorem~\ref{thm:stats} shows a stronger statistical impossibility for dTBM than that presented in TBM \cite[Theorem 6]{han2020exact}.
 However, inspecting the proof of \cite{han2020exact}, the proof of Theorem 6 indeed implies a stronger TBM impossibility statement for arbitrary core tensor; i.e., when $\gamma < -(K-1)$
+
+ \vspace{-0.5cm}
  \small
@@
  where $\tP_{\tS, {\rm TBM}}$ and $\tP_{z, {\rm TBM}}$ refer to the space for core tensor $\tS$ and assignment $z$ under TBM, respectively.
-Again, in terms of the strong statistical impossibility, both models show similar conclusions but under different conditions.
+Again, in terms of the strong statistical impossibility, both models show the similar conclusion but under different conditions.
 Since two impossibilities consider different core tensor regimes with non-equivalent $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$, we emphasize that different proof techniques are required to obtain these similar conclusions.
@@
 We develop polar-coordinate based techniques to establish the error rate for the proposed algorithm.
-"
 \end{itemize}
@@
 Clustering accuracy is assessed by clustering error rate (CER, i.e., one minus rand index).
-Note that CER between $(\hat z, z)$ is equivalent to misclustering error $\ell(\hat z, z)$ up to constant multiplications~\citep{meilua2012local}, and a lower CER indicates a better performance.
+The CER between $(\hat z, z)$ is equivalent to misclustering error $\ell(\hat z, z)$ up to constant multiplications~\citep{meilua2012local}, and a lower CER indicates a better performance.
 
@@
 Hence, we easily adjust SNR via varying $\alpha$.
-Note that the assortative setting is proposed for simulations, and our algorithm is applicable for general tensors in practice.
+The assortative setting is proposed for simulations, and our algorithm is applicable for general tensors in practice.
 The cluster assignment $z$ is randomly generated with equal probability across  $r$ clusters for each mode.
@@
 Table~\ref{tab:select} shows that our BIC criterion well chooses the true $r$ under most settings.
-Note that the BIC slightly underestimates the true cluster number $(r = 4)$ with smaller dimension and higher noise $(p = 50, \sigma=1)$, and the accuracy immediately increases with larger dimension $p = 80$.
+Note that the BIC slightly underestimates the true cluster number $(r = 4)$ with smaller dimension and higher noise $(p = 50, \sigma^2=1)$, and the accuracy immediately increases with larger dimension $p = 80$.
 The improvement follows from the fact that a larger dimension $p$ indicates a larger sample size in the tensor block model.
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\columnwidth]{comp_theta_anno3.pdf}
-    \caption{CER versus shape parameter in degree (denoted $a\in[3,6]$) for different methods. We set $p = 100, r = 5, \gamma = -1.2$ under (a) Gaussian and (b) Bernoulli models.}
-    \label{fig:comp_theta}
-\end{figure}
-
 \begin{figure}[htp!]
@@
     \label{fig:comp_gamma_theta}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\columnwidth]{comp_theta_anno3.pdf}
+    \caption{CER versus shape parameter in degree (denoted $a\in[3,6]$) for different methods. We set $p = 100, r = 5, \gamma = -1.2$ under (a) Gaussian and (b) Bernoulli models.}
+    \label{fig:comp_theta}
 \end{figure}
@@
 In this section, we provide the proof sketches for the main Theorem~\ref{thm:stats} (Impossibility), Theorem~\ref{thm:comp} (Impossibility), and Theorems~\ref{thm:initial}-\ref{thm:refinement}.
-Detail proofs and extra theoretical results are provided in Appendix II.
+Detail proofs and extra theoretical results are provided in Appendix B.
 
@@
 \end{align}
-where $\tX = \bbE \tY$ is the true mean.
-The first inequality in~\eqref{eq:proof_4} holds with the assumption $\min_{i \in [p]} \theta(i) \geq c>0$ in Theorem~\ref{thm:initial}.
-The second inequality relies on an important conclusion that the angle gap of mean tensor $\tX$ is lower bounded by that of core tensor $\tS$, i.e., the minimal angle gap $\Delta_{\min}$ defined in Assumption~\ref{assmp:min_gap}.
+where $\tX = \bbE \tY$ is the true mean. The first inequality in~\eqref{eq:proof_4} holds with the assumption $\min_{i \in [p]} \theta(i) \geq c>0$ in Theorem~\ref{thm:initial}. 
+{The second inequality relies on the key Lemma~\ref{lem:angle_gap_x}, which indicates}
+%The second inequality relies on an important conclusion that the angle gap of mean tensor $\tX$ is lower bounded by that of core tensor $\tS$, i.e., the minimal angle gap $\Delta_{\min}$ defined in Assumption~\ref{assmp:min_gap}.
 Let $\ma^{s} \coloneqq \ma/\onorm{\ma}$ denote the normalized vector with the convention that $\ma^s = 0$ if $\ma = 0$.
@@
 where $\mX = \mat(\tX)$.
-The most challenging part in the proof of Theorem~\ref{thm:initial} lies in the derivation of inequality~\eqref{eq:proof_4_gap}, in which the proof of \cite{gao2018community} is no longer applicable due to different angle gap assumption in our dTBM.
-We develop the extra padding technique in Lemma~\ref{lem:pad} and balance assumption~\eqref{eq:degree} to derive~\eqref{eq:proof_4_gap}.
+The most challenging part in the proof of Theorem~\ref{thm:initial} lies in the derivation of inequality~\eqref{eq:proof_4_gap} (or the proof of Lemma~\ref{lem:angle_gap_x}), in which the proof of \cite{gao2018community} is no longer applicable due to different angle gap assumption in our dTBM.
+To address the angle gap notion, we develop the extra padding technique in Lemma~\ref{lem:pad} and balance assumption~\eqref{eq:degree}.
 Last, we finish the proof of Theorem~\ref{thm:initial} by showing the third inequality of~\eqref{eq:proof_4} using \cite[Proposition 1]{han2020exact}.
@@
 Recall the intermediate quantity, misclustering loss, defined in~\eqref{eq:defnofL}
+
+\vspace{-0.5cm}
 \small
@@
     L^{(t)} : &=L(z,z^{(t)})\\
-    &= \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
+    &= \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2.
 \end{align}
 \normalsize
-where the superscript $\cdot^{s}$ denotes the normalized vector; i.e., $\ma^s:=\ma/\onorm{\ma}$ if $\ma\neq 0$ and $\ma^{s}=0$ if $\ma=0$ for any vector $\ma$.
-We show that $L^{(t)}$ provides an upper bound for the misclassification error of interest via the inequality $\ell^{(t)}\leq {L^{(t)}\over \Delta^2_{\min}}$.
+%where the superscript $\cdot^{s}$ denotes the normalized vector; i.e., $\ma^s:=\ma/\onorm{\ma}$ if $\ma\neq 0$ and $\ma^{s}=0$ if $\ma=0$ for any vector $\ma$. 
+We show that $L^{(t)}$ provides an upper bound for the misclustering error of interest via the inequality $\ell^{(t)}\leq {L^{(t)}\over \Delta^2_{\min}}$ in Lemma~\ref{lem:mis}.
 Therefore, it suffices to control $L^{(t)}$.
@@
 where $\mY = \mat(\tY)$, $ \mS = \mat(\tS)$, $\mS^{(t)} = \mat(\tS^{(t)})$ and
+
+\vspace{-0.5cm}
 \small
@@
 The proof of MLE error shares the similar idea as Theorems~\ref{thm:initial}-\ref{thm:refinement}.
-we first show a weaker polynomial rate for MLE, and then improve the rate from polynomial to exponential throughout the iterations.
+We first show a weaker polynomial rate for MLE and then improve the rate from polynomial to exponential through the iterations.
 The only difference is that the MLE remains the same over iterations due to its global optimality.
-See Appendix II, Section~\ref{sec:statprove2} for the detailed proof.
+See Appendix B, Section~\ref{sec:statprove2} for the detailed proof.
 
@@
 Before the proofs, we first introduce the notation used throughout the appendix and the general dTBM without symmetric assumptions.
-The parameter space and minimal gap assumption are also extended for the general asymetric dTBM.
+The parameter space and minimal gap assumption are also extended for the general asymmetric dTBM.
 
@@
     \item For a matrix $\mA$, let $\onormSize{}{\mA}_\sigma$ denote the spectral norm of matrix $\mA$, which is equal to the maximal singular value of $\mA$; let $\lambda_k(\mA)$ denote the $k$-th largest singular value of $\mA$; let $\onormSize{}{\mA}_F$ denote the Frobenius norm of matrix $\mA$.
-    \item For two sequence $a$ and $b$, let $a \asymp b$ if there exist two positive constants $c, C$ such that $cb \leq a\leq Cb$. 
+    %\item For two sequences $a$ and $b$, let $a \asymp b$ if there exist two positive constants $c, C$ such that $cb \leq a\leq Cb$. 
 \end{enumerate}
 
-{\bf Extension to general asymetric dTBM.}
+{\bf Extension to general asymmetric dTBM.}
 
@@
 \begin{equation}\label{eq:general_dtbm}
-    \tY = \tX+ \tE,\quad \text{where}\quad \tX=\tS \times_1 \mTheta_1 \mM_1 \times_2 \cdots \times_K \mTheta_K \mM_K,
+    \tY = \tX+ \tE, \ \text{where}\ \tX=\tS \times_1 \mTheta_1 \mM_1 \times_2 \cdots \times_K \mTheta_K \mM_K,
 \end{equation}
@@
 Correspondingly, we consider the parameter space for the triplet $\of{\{z_k\}, \tS, \{\mtheta_k\}}$,
+
+\vspace{-1cm}
 \small
 \begin{align}
- & \tP(\{r_k\}) = \Big\{ \of{\{z_k\}, \tS, \{\mtheta_k\}}: \mtheta_k \in\mathbb{R}^p_{+}, {c_1 p_k\over r_k} |z_k^{-1}(a)| \leq {c_2 p_k\over r_k}, \\ & \ c_3 \leq  \onorm{\mS_{k,a:}} \leq c_4 ,\onormSize{}{\mtheta_{k,z_k^{-1}(a)}}_1=|z_k^{-1}(a)|, a \in [r_k], k\in[K]\Big\}.
-\label{eq:general_family}
+  \tP(\{r_k\}) = \Big\{ &\of{\{z_k\}, \tS, \{\mtheta_k\}}: \mtheta_k \in\mathbb{R}^p_{+}, {c_1 p_k\over r_k} |z_k^{-1}(a)| \leq {c_2 p_k\over r_k}, \\ & c_3 \leq  \onorm{\mS_{k,a:}} \leq c_4 ,\onormSize{}{\mtheta_{k,z_k^{-1}(a)}}_1=|z_k^{-1}(a)|,\\
+  & \text{for all }a \in [r_k], k\in[K]\Big\}. \label{eq:general_family}
 \end{align}
@@
 Similarly, let $\text{SNR} = \Delta_{\min}^2/\sigma^2$ with the generalized minimal gap $\Delta_{\min}^2$ defined in Assumption~\ref{assmp:general_minimal_gap}. We define the regime
+
+\vspace{-0.5cm}
 \small
@@
 
-To study the identifiability, we consider the noiseless model with $\tE = 0$. Assume there exist two parameterizations satisfying
+To study the identifiability, we consider the noiseless model with $\tE = 0$. Assume that there exist two parameterizations satisfying
 \begin{align}
@@
     \item[$(\Leftarrow)$] For the necessity, it suffices to construct two distinct parameters up to cluster label permutation, if the model~\eqref{eq:general_dtbm} violates Assumption~\ref{assmp:general_minimal_gap}.
-{Note that $\Delta_{\min}^2 = 1$ when there exists a $k \in [K]$ such that $r_k = 1$.
+{Note that $\Delta_{\min}^2 = 1$ when there exists $k \in [K]$ such that $r_k = 1$.
 Hence, we consider the case that  $r_k \geq 2$ for all $k \in [K]$.}  Without loss of generality, we assume $\onorm{ \mS_{1,1:}^s - \mS_{1,2:}^s } = 0$.
@@
 First, we show the uniqueness of $\mM_k$ for all $k \in [K]$. 
-{When $r_k = 1$, all possible $\mM_k$ is equal to the vector ${\bf 1}_{p_k}$, and the uniqueness holds trivially.
+{When $r_k = 1$, all possible $\mM_k$'s are equal to the vector ${\bf 1}_{p_k}$, and the uniqueness holds trivially.
 Hence, we consider the case that $r_k \geq 2$.
@@
 
-\begin{lem}[Motivation of angle-based clustering]\label{lem:angle} Consider the signal tensor $\tX$ in the general asymmetric dTBM~\eqref{eq:general_dtbm} with $(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\{r_k\})$ and $r_k \geq 2$.
+\begin{lem}[Motivation of angle-based clustering]\label{lem:angle} Consider the signal tensor $\tX$ in the general asymmetric dTBM~\eqref{eq:general_dtbm} with $(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\{r_k\})$ and $r_k \geq 2, k \in [K]$.
 Then, for any $k \in [K]$ and index pair $(i,j)\in[p_k]^2$, we have
@@
 Note that the vector $\mS_{z(i):}$ can be folded to a tensor $\tS' =\entry{\tS'_{a_2,\ldots,a_K}} \in \bbR^{r^{K-1}}$; i.e., $\text{vec}(\tS') = \mS_{z(i):}$.
-Define weight vectors $\mw_{a_2, \cdots, a_K}$ correspond to the elements in $\tS'_{a_2,\ldots,a_K}$ by
+Define weight vectors $\mw_{a_2, \cdots, a_K}$ corresponding to the elements in $\tS'_{a_2,\ldots,a_K}$ by
 \begin{equation}
@@
 Specifically, we have $\onormSize{}{\mw_{a_2, \ldots, a_K}}^2 = \prod_{k = 2}^K \onormSize{}{ \mtheta_{z^{-1}(a_k)}}^2$, and by the balanced assumption~\eqref{eq:degree} we have
+
+\vspace{-0.5cm}
 \small
@@
 
-\begin{proof}[Proof of Theorem~\ref{thm:stats} (Impossibility)]Consider the general asymetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$ {with $K\geq 2$, $2 \leq r\lesssim p^{1/3}$ as $p \rightarrow \infty$}.
+\begin{proof}[Proof of Theorem~\ref{thm:stats} (Impossibility)]Consider the general asymmetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$ {with $K\geq 2$, $2 \leq r\lesssim p^{1/3}$ as $p \rightarrow \infty$}.
 For simplicity, we show the minimax rate for the estimation on the first mode $\hat z_1$; the proof for other modes are essentially the same.
@@
    \begin{equation}
-       \inf_{\hat z_1} \bbE \left[ p\ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right]\geq 1.
+       \inf_{\hat z_1} \bbE \left[ p\ell(\hat z_1, z_1^*) | (z^*_k, \tS^*,  \mtheta^*_k)  \right]\geq 1.
    \end{equation}
@@
    \begin{align}
-        & \inf_{\hat z_1 } \bbE \left[ \ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right] \geq  \\
-        & \  \frac{C}{r^3 |T_1^c|} \sum_{i \in T_1^c} \inf_{\hat z_1  (i)} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, z_k^*, \mtheta_k^*, \tS^*] \\
-        & \quad \quad \quad  \quad  \quad \quad +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, z_k^*, \mtheta_k^*, \tS^*] \},\label{eq:inf_lower}
+        & \inf_{\hat z_1 } \bbE \left[ \ell(\hat z_1, z_1^*) | (z^*_k, \tS^*, \mtheta^*_k)  \right] \geq  \\
+        & \  \frac{C}{r^3 |T_1^c|} \sum_{i \in T_1^c} \inf_{\hat z_1  (i)} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2,z^*_k, \tS^*, \mtheta^*_k] \\
+        & \quad \quad \quad  \quad  \quad \quad +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, z^*_k, \tS^*, \mtheta^*_k] \},\label{eq:inf_lower}
    \end{align}
@@
 Then, we consider the hypothesis testing under the model~\eqref{eq:z}:
+
+\vspace{-0.5cm}
 \small
@@
 %   \end{align}
-   where $\lambda_{\max}$ denote the maximal singular value, the second inequality follows from Lemma~\ref{lem:singular_thetam}, and the third inequality follows from property (ii) and the boundedness constraint in $\tP_{\tS}(\gamma)$ such that $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F  \leq \onormSize{}{\mS_{1:}^*}_F + \onormSize{}{\mS_{2:}^*}_F \leq 2c_4$.
+   where $\lambda_{\max}(\cdot)$ denotes the maximal singular value, the second inequality follows from Lemma~\ref{lem:singular_thetam}, and the third inequality follows from property (ii) and the boundedness constraint in $\tP_{\tS}(\gamma)$ such that $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F  \leq \onormSize{}{\mS_{1:}^*}_F + \onormSize{}{\mS_{2:}^*}_F \leq 2c_4$.
 
@@
    \begin{align}
-       & \inf_{\hat z_1(1)} \{ \bbP[\hat z_1(1) = 1| z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^*] \\
-       & \quad \quad \quad  \quad \quad  +   \bbP[\hat z_1(1) = 2| z_1^*(1) = 1, z^*_k,  \mtheta_k^*, \tS^*] \} \\
-       & \quad = 2 \bbP[ \hat z_{LS} = 1 |  z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^* ] \\
-      & \quad  = 2 \bbP[ \onormSize{}{\my - \theta^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \theta^* \mx_2^*}_F^2 | z_1^*(1) = 2, z^*_k , \mtheta_k^*, \tS^*   ] \\
-      & \quad = 2 \bbP \off{ 2 \ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \geq  \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F^2 }\\
+       & \inf_{\hat z_1(1)} \{ \bbP[\hat z_1(1) = 1| z_1^*(1) = 2, z^*_k, \tS^*, \mtheta^*_k] \\
+       & \quad \quad \quad  \quad \quad  +   \bbP[\hat z_1(1) = 2| z_1^*(1) = 1, z^*_k, \tS^*, \mtheta^*_k] \} \\
+       & \quad = 2 \bbP[ \hat z_{LS} = 1 |  z_1^*(1) = 2, z^*_k, \tS^*, \mtheta^*_k ] \\
+      & \quad  = 2 \bbP[ \onormSize{}{\my - \theta^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \theta^* \mx_2^*}_F^2 | z_1^*(1) = 2, z^*_k, \tS^*, \mtheta^*_k   ]\\
+      & \quad = 2 \bbP [ 2 \langle \me, \theta^* \mx_1^* - \theta^*\mx_2^* \rangle \geq  \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F^2 ]\\
        & \quad =  2 \bbP[ N(0,1) \geq \theta^* \onormSize{}{\mx_1^* - \mx_2^*}_F /(2\sigma) ] \\
@@
 %   \end{align}
-where the first equation holds by symmetry, the third equation holds by rearrangement, the fourth equation holds by the fact that $\ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \sim N(0, \sigma \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F)$, and $c$ is some positive constant in the last inequality.
+where the first equation holds by symmetry, the third equation holds by rearrangement, the fourth equation holds from the fact that $\ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \sim N(0, \sigma \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F)$, and $c$ is some positive constant in the last inequality.
    %for some positive constant $c$, where the calculation for the last equation follows from the calculation in equation (102) in \cite{han2020exact}.
@@
    \end{equation}
-   where the last inequality follows by the condition $r = o(p^{1/3})$. By the discrete nature of the misclassification error, we obtain our conclusion
+   where the last inequality follows by the condition $r = o(p^{1/3})$. By the discrete nature of the misclustering error, we obtain our conclusion
    \begin{equation}
@@
 \begin{align}
-    &\liminf_{p \rightarrow \infty} \inf_{\hat z_1} \sup_{(z, \mtheta, \tS) \in \tP'(\gamma')} \bbE \left[ p\ell(\hat z_1, z_1) \right]  \\
-    &\quad \geq   \liminf_{p \rightarrow \infty} \inf_{\hat z_{\rm stat}}  \bbE \left[ p\ell(\hat z_1, z^*_1) |  z^*_k, \mtheta^*_k, \tS^*  \right]  \geq 1.
+    &\liminf_{p \rightarrow \infty} \inf_{\hat z_1} \sup_{(z, \tS, \mtheta) \in \tP'(\gamma')} \bbE \left[ p\ell(\hat z_1, z_1) \right]  \\
+    &\quad \geq   \liminf_{p \rightarrow \infty} \inf_{\hat z_{\rm stat}}  \bbE \left[ p\ell(\hat z_1, z^*_1) |  z^*_k, \tS^*, \mtheta^*_k  \right]  \geq 1.
 \end{align}
@@
 Then, we consider the signal tensor family
+
+\vspace{-0.5cm}
 \small
@@
 
-\begin{proof}[Proof of Theorem~\ref{thm:initial}] We prove Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with symmetric mean tensor and parameters $(z, \tS, \mtheta)$ {fixed $r\geq 1, K \geq 2$}, and i.i.d.\ noise.
+\begin{proof}[Proof of Theorem~\ref{thm:initial}] We prove Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with symmetric mean tensor, parameters $(z, \tS, \mtheta)$, {fixed $r\geq 1, K \geq 2$}, and i.i.d.\ noise.
 {For the case $r = 1$, we have $L(z^{(0)},z) = 0, \ell(z^{(0)}, z) = 0$ trivially.
@@
     where
+
+\vspace{-0.5cm}
     \small
@@
     \end{align}
-    where inequalities \eqref{eq:tr1} and \eqref{eq:tr2} follow from the triangle inequality, \eqref{eq:defs} follows from the definition of $S$, \eqref{eq:km} follows from the update rule of $k$-means in Step 5 of Sub-algorithm~\hyperref[alg:main]{1}, \eqref{eq:lem1} follows from Lemma~\ref{lem:norm_diff}, and the last inequality~\eqref{eq:lem_err} follows from Lemma~\ref{lem:two-step_esterror}.
+    where inequalities \eqref{eq:tr1} and \eqref{eq:tr2} follow from the triangle inequality, \eqref{eq:defs} follows from the definition of $S$, \eqref{eq:km} follows from the update rule of $k$-means in Step 6 of Sub-algorithm~\hyperref[alg:main]{1}, \eqref{eq:lem1} follows from Lemma~\ref{lem:norm_diff}, and the last inequality~\eqref{eq:lem_err} follows from Lemma~\ref{lem:two-step_esterror}.
 Also, note that
@@
 \end{align}
-where the first inequality follows from the assumption $\min_{i \in [p]} \theta(i) \geq c$, the second inequality follows from the inequality~\eqref{eq:theta_p}, and the last inequality comes from the inequality~\eqref{eq:cor_1}.
+where the first inequality follows from the assumption $\min_{i \in [p]} \theta(i) \geq c > 0$, the second inequality follows from the inequality~\eqref{eq:theta_p}, and the last inequality comes from the inequality~\eqref{eq:cor_1}.
 
@@
 In addition, with Theorem~\ref{thm:initial} and the condition SNR $\gtrsim p^{-K/2} \log p$, for all $a \in [r]$, we have
+
+\vspace{-0.5cm}
 \small
@@
 when $p$ is large enough. Therefore, for all $a \in [r]$, we have
+
+\vspace{-0.5cm}
 \small
@@
 \end{align}
-where the equations follows from the properties of trigonometric function and the inequality follows from the fact the $\cos \frac{\alpha}{2} \leq 1$ and $\sin \alpha = 2 \sin \frac{\alpha}{2} \cos \frac{\alpha}{2} > 0$ for $\alpha \in [0, \pi]$.
+where the equations follow from the properties of trigonometric function and the inequality follows from the fact the $\cos \frac{\alpha}{2} \leq 1$ and $\sin \alpha = 2 \sin \frac{\alpha}{2} \cos \frac{\alpha}{2} > 0$ for $\alpha \in [0, \pi]$.
 
@@
 
-\begin{lem}[Angle for weighted padding vectors]\label{lem:pad} Suppose we have two non-zero vectors $\ma, \mb \in \bbR^d$. Given the weight collection $\mw$, we have 
+\begin{lem}[Angle for weighted padding vectors]\label{lem:pad} Suppose that we have two non-zero vectors $\ma, \mb \in \bbR^d$. Given the weight collection $\mw$, we have 
 \begin{align}
@@
 Next, we prove inequality * in~\eqref{eq:lempad}. With the decomposition of $\pad_{\mw}(\mb)$ and the inverse weighted padding operator, we have
+
+\vspace{-0.5cm} 
 \small
@@
 we have
+
+\vspace{-0.5cm}
 \small
@@
         \end{equation}
-        There exist positive constant $C_1, C_2$ such that
+        There exist positive constants $C_1, C_2$ such that
         \begin{equation}
@@
 
-We consider dTBM~\eqref{eq:model_tensor} with symmetric mean tensor and parameters $(z, \tS, \mtheta)$ {fixed $r\geq 1, K \geq 2$}, and i.i.d.\ noise.
-Let $(\hat z, \hat \mtheta, \hat \tS)$ denote the MLE in \eqref{eq:mle}, and $(z^{(0)}_k, \mtheta^{(0)}, \tS^{(0)})$ denote parameters related to the initialization.
-{For the case $r = 1$, we have $\ell(z^{(t)}, z) = 0$ trivially for all $t \geq 0$.
+We consider dTBM~\eqref{eq:model_tensor} with symmetric mean tensor, parameters $(z, \tS, \mtheta)$, {fixed $r\geq 1, K \geq 2$}, and i.i.d.\ noise.
+Let $(\hat z,\hat \tS,  \hat \mtheta)$ denote the MLE in \eqref{eq:mle}, and $(z^{(0)}_k, \tS^{(0)}, \mtheta^{(0)}_k)$ denote parameters related to the initialization.
+{For the case $r = 1$, $\ell(z^{(t)}_k, z) = 0$ trivially for all $t \geq 0, k \in [k]$.
 Hence, we focus on the proof of the first mode clustering $z^{(t+1)}_1$ with $r \geq 2$};
@@
 the extension for other modes can be obtained similarly.
-We drop the subscript $k$ in the matricizations $\mTheta, \mM_k, \mS_k, \mX_k$ and in estimates $z^{(0)}, z^{(t+1)}_1, z^{(t)}_k$ for ease of the notation.
+We drop the subscript $k$ in the matricizations $\mTheta, \mM_k, \mS_k, \mX_k$ and in estimates $z^{(0)}_k, z^{(t+1)}_k, z^{(t)}_k$ for ease of the notation.
 Without loss of generality, we assume that the variance $\sigma = 1$, and that the identity permutation minimizes the initial misclustering error; i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$, and so for $\hat z$.
@@
     \begin{align}
-         &\mV = \mW^{\otimes (K-1)}, \quad \mV^{(t)} =\of{ \mW^{(t)}}^{\otimes (K-1)}\\
+         &\mV = \mW^{\otimes (K-1)}, \quad \mV^{(t)} =\of{ \mW^{(t)}}^{\otimes (K-1)},\\
          &\mQ = \mP^{\otimes K-1}, \quad  \hat \mQ = \hat \mP^{\otimes K-1}.
@@
     \begin{align}
-        &\tS^{(t)} = \tY \times_1 \of{\mW^{(t)}}^T \times_2 \cdots \times_K \of{\mW^{(t)}}^T,\\ &\tilde \tS = \tY \times_1 \mW^T \times_2 \cdots \times_K \mW^T \\
-        &\hat \tS =  \tY \times_1 \hat \mP^T \times_2 \cdots \times_K \hat \mP^T \\
+        &\tS^{(t)} = \tY \times_1 \of{\mW^{(t)}}^T \times_2 \cdots \times_K \of{\mW^{(t)}}^T,\\ &\tilde \tS = \tY \times_1 \mW^T \times_2 \cdots \times_K \mW^T, \\
+        &\hat \tS =  \tY \times_1 \hat \mP^T \times_2 \cdots \times_K \hat \mP^T, \\
         &\bar \tS  = \tY \times_1 \mP^T \times_2 \cdots \times_K \mP^T.
@@
     \begin{align}
-        \mS = \mat(\tS), \  \mY = \mat(\tY), \  \mX = \mat(\tX), \  \mE = \mat(\tE)
-    \end{align}
-    \begin{equation}
-        \mS^{(t)} = \mat(\tS^{(t)}), \   \hat \mS = \mat(\hat \tS), \  \tilde \mS = \mat(\tilde \tS), \  \bar \mS = \mat( \bar \tS )
+        \mS = \mat(\tS), \  \mY = \mat(\tY), \  \mX = \mat(\tX), \  \mE = \mat(\tE),
+    \end{align}
+    \begin{equation}
+        \mS^{(t)} = \mat(\tS^{(t)}), \   \hat \mS = \mat(\hat \tS), \  \tilde \mS = \mat(\tilde \tS), \  \bar \mS = \mat( \bar \tS ).
     \end{equation}
@@
 % equations are referenced in the main text. 
-\setcounter{equation}{62} 
+\setcounter{equation}{61} 
 \begin{equation}\label{eq:cond1}
@@
 % Restore the current equation number. 
-\setcounter{equation}{\value{MYtempeqncnt}} 
-where $M$ is a positive universal constant in inequality~\eqref{eq:constant_M}, $\bar C, \tilde C$ are universal constants in {the proof of Theorem~\ref{thm:initial}} and assumption SNR $\geq \tilde C p^{-K/2} \log p$, respectively.
+where $M$ is a positive universal constant in inequality~\eqref{eq:constant_M}, $\bar C, \tilde C$ are positive universal constants in {the proof of Theorem~\ref{thm:initial}} and assumption SNR $\geq \tilde C p^{-K/2} \log p$, respectively.
 Further, inequality~\eqref{eq:cond1} holds by replacing $\mV$ to $\mV^{(t)}, \mQ, \hat \mQ$ and $\mW_{:a}$ to $\mW_{:a}^{(t),T}, \mP_{:a}^T, \hat \mP_{:a}^T$ when initialization condition~\eqref{eq:cond_intial} holds.
 \end{condition}
+\setcounter{equation}{\value{MYtempeqncnt}} 
 % The IEEE uses as a separator 
@@
 \end{figure*}
-\setcounter{equation}{67} 
+\setcounter{equation}{66} 
 % \begin{equation}\label{eq:cond1}
@@
     Arranging the terms in \eqref{eq:event} yields the decomposition
+
+    \vspace{-0.5cm}
     \small
@@
     where
+
+    \vspace{-0.5cm}
     \small
@@
    Therefore, the event $ \ind \offf{ z^{(t+1)}(i) = b }$ can be upper bounded as
+
+   \vspace{-0.5cm}
    \small
@@
     where $ \zeta_{ib}' =   \theta(i)^2 B_{ib}' \onormSize{}{ [ \mA_{ z(i):}  ]^s - [ \mA_{b:}  ]^s  }^2$ and
+
+    \vspace{-0.5cm}
     \small
@@
 % Store the current equation number. 
-{\bf Step 3.}  Choose the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ that satisfies the condition of Lemma~\ref{lem:upper_fgh}, inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
+{\bf Step 3.}  Choose the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ that satisfies the condition of Lemma~\ref{lem:upper_fgh}, inequalities~\eqref{eq:tilde_c1}, and \eqref{eq:tilde_c2}.
 Note that  \begin{align}
@@
 % equations are referenced in the main text. 
-\setcounter{equation}{75} 
+\setcounter{equation}{74} 
 \begin{equation}\label{eq:f_sum}
@@
 \end{figure*}
-\setcounter{equation}{78}
+\setcounter{equation}{77}
 
@@
     Now, we verify the oracle loss condition~\eqref{eq:cond_oracle}. Recall the definition of $\xi$,
+
+    \vspace{-0.5cm}
     \small
@@
     \end{equation}
-Therefore, the condition~\eqref{eq:cond_intial} holds for $t = 0$. Assume the condition~\eqref{eq:cond_intial} also holds for all $t \leq t_0$. Then, by the decomposition~\eqref{eq:decomp_t}, we have 
+Therefore, the condition~\eqref{eq:cond_intial} holds for $t = 0$.
+Assume that the condition~\eqref{eq:cond_intial} also holds for all $t \leq t_0$.
+Then, by the decomposition~\eqref{eq:decomp_t}, we have
     \begin{align}
@@
 
-\begin{lem}[Singular-value property of membership matrices]\label{lem:membership} Under the setup of Theorem~\ref{thm:refinement}, suppose the condition~\eqref{eq:cond_intial} holds.
+\begin{lem}[Singular-value property of membership matrices]\label{lem:membership} Under the setup of Theorem~\ref{thm:refinement}, suppose that the condition~\eqref{eq:cond_intial} holds.
 Then, for all $a \in [r]$, we have $|\of{z^{(t)}}^{-1}(a)| \asymp p/r$.
@@
 
-\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement}  {with fixed $r \geq 2$},  assume the the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ is large enough to satisfy the inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
+\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement}  {with fixed $r \geq 2$},  assume the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ is large enough to satisfy the inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
 {As $p \rightarrow \infty$}, we have
+
+\vspace{-0.5cm}
 \small
@@
 
-    Similarly, when the SNR $\geq \tilde C p^{-(K-1)}\log p$ with large constant $\tilde C$, we have 
+    Similarly, when the SNR $\geq \tilde C p^{-(K-1)}\log p$ with a large constant $\tilde C$, we have
+
+    \vspace{-0.5cm}
     \small
@@
     \item Upper bound for $F_{ib}^{(t)}$, i.e., inequality \eqref{eq:fib}. Recall the definition of $F_{ib}^{(t)}$,
+
+    \vspace{-0.5cm}
     \small
@@
     By Cauchy-Schwartz inequality, we have
+
+    \vspace{-0.5cm}
     \small
@@
     Note that
+
+    \vspace{-0.5cm}
     \small
@@
         & \leq \frac{2}{ \onormSize{}{ \mW_{z(i):}^T\mX \mV^{(t)}}^2} \max_{a \in [r]}\onormSize{}{ \mW_{:a}^T \mE \mV^{(t)} }^2 \\
-        & \quad \quad \quad \quad \quad  \quad \cdot \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
+        & \quad \quad \quad \quad  \quad \cdot \max_{a \in [r]} \onormSize{}{ [  \mW_{:a}^T \mY \mV^{(t)} ]^s -  [   \mW_{:a}^{(t),T} \mY \mV^{(t)}  ]^s}^2\\
         & \leq C' \frac{ r^{2K-1} + K p r^{K+1} }{p^{K}} \\
@@
    Then we have
+
+    \vspace{-0.5cm}
    \small
@@
     \item Upper bound for $H_{ib}^{(t)}$, i.e., the inequality~\eqref{eq:hib}. By definition of $H_{ib}$, we rearrange terms and obtain
+
+    \vspace{-0.5cm}
     \small
@@
 \end{equation}
-Also, notice that the matrix product of $ \mB^T$ corresponding to the padding operation in Lemma~\ref{lem:pad}, and the padding weights are balanced such that $\onormSize{}{\mv \mB}  = (1 + o(1)) \max_{a} \onormSize{}{\mtheta_{z^{-1}(a)}}^{(K-1)/2} \onormSize{}{\mv} $ for all $\mv \in \bbR^{r(K-1)}$.
+Also, notice that the matrix product of $ \mB^T$ corresponds to the padding operation in Lemma~\ref{lem:pad}, and the padding weights are balanced such that $\onormSize{}{\mv \mB}  = (1 + o(1)) \max_{a} \onormSize{}{\mtheta_{z^{-1}(a)}}^{(K-1)/2} \onormSize{}{\mv} $ for all $\mv \in \bbR^{r(K-1)}$.
 For two vectors $\mv_1, \mv_2 \in \bbR^{r^{K-1}}$, we have
@@
 Note that for all $i \in [p]$ we have
+
+    \vspace{-0.5cm}
 \small
@@
 \end{align}
-Following similar derivations in inequalities~\eqref{eq:fib_1}, \eqref{eq:fib_2}, and the upper bound for $J_{1}$, respectively, we have 
+Following similar derivations in inequalities~\eqref{eq:fib_1}, \eqref{eq:fib_2}, and the upper bound for $J_{1}$ in the proof of Lemma~\ref{lem:intermediate}, respectively, we have 
 \begin{equation}
@@
         & = \of{ \onormSize{}{[\mX_{i:}\hat \mQ]^s - [\hat \mS_{a:}]^s}_F^2 - \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:a}^T \mY \hat \mQ ]^s }_F^2  } \\
-        & \quad \quad- \of{ \onormSize{}{[\mX_{i:} \hat \mQ]^s - [\hat \mS_{b:}]^s}_F^2 - \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:b}^T \mY \hat \mQ ]^s }_F^2   }. \\
+        & \quad \quad- \of{ \onormSize{}{[\mX_{i:} \hat \mQ]^s - [\hat \mS_{b:}]^s}_F^2 - \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:b}^T \mY \hat \mQ ]^s }_F^2   }, \\
         &(1+o(1))\hat H_{ib}  \\
@@
      \begin{lem}[Relationship between misclustering loss and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement} {with fixed $r \geq 2$, as $p \rightarrow \infty$}, we have
+
+    \vspace{-0.4cm}
      \small
@@
     By Lemma~\ref{lem:membership},  we have $\sum_{i \in [p]}  \ind\offf{  z^{(t)}(i) = b } \gtrsim p/r$. Then, we have
+
+    \vspace{-0.5cm}
     \small
@@
  We have
+
+    \vspace{-0.5cm}
  \small
@@
          \sin \beta_{k1} &= \sin ({\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k} }, \\
-         & \quad \quad \quad \quad \quad \quad\quad \quad\quad \quad {\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \tilde \mD \otimes \mD^{\otimes K-k-1} } ),\\
+         & \quad   \quad\quad \quad\quad \quad {\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \tilde \mD \otimes \mD^{\otimes K-k-1} } ),\\
          \sin \beta_{k2} & = \sin ({\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \tilde \mD \otimes \mD^{\otimes K-k-1} } ,\\
@@
     \end{align}
+
     {Note that we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \leq \frac{\bar C}{\tilde C} r \log^{-1}(p) $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.
@@
 It follows from~\eqref{eq:step} and Lemma~\ref{lem:membership} that
+
+    \vspace{-0.5cm}
     \small
@@
         Consider the MLE confusion matrix $\hat \mD = \mM^T \mTheta^T \hat \mP = \entry{\hat D_{ab}} \in \bbR^{r \times r}$ with entries
+
+    \vspace{-0.5cm}
         \small
@@
 
-\begin{lem}[Polynomial estimation error of MLE]\label{lem:poly_mle_degree}  Let $(\hat z, \hat \mtheta, \hat \tS)$ denote the MLE in \eqref{eq:mle} {with fixed $K \geq 2$}, and $\hat \tX$ denote the mean tensor consisting of parameter $(\hat z, \hat \mtheta, \hat \tS)$.
+\begin{lem}[Polynomial estimation error of MLE]\label{lem:poly_mle_degree}  Let $(\hat z, \hat \tS, \hat \mtheta)$ denote the MLE in \eqref{eq:mle} {with fixed $K \geq 2$ and symmetric mean tensor}, and $\hat \tX$ denote the mean tensor consisting of parameter $(\hat z, \hat \tS, \hat \mtheta)$.
 {With high probability going to 1 as $p \rightarrow \infty$, we} have
@@
 with probability going to 1. 
-When {$\text{SNR} \gtrsim p^{-(K-1)} \log p$,  $\mtheta$ is balanced, and $\min_{i \in [p]} \theta(i) \geq c$ for some constant $c$}, the MLE satisfies
+When {$\text{SNR} \gtrsim p^{-(K-1)} \log p$,  $\mtheta$ is balanced, and $\min_{i \in [p]} \theta(i) \geq c$ for some positive constant $c$}, the MLE satisfies
 \begin{align}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:poly_mle_degree}] Without loss of generality, we assume $\sigma^2 = 1$ and identity mapping minimizes the misclustering error for MLE.
-For arbitrary two sets of parameters $(z, \mtheta, \tS), (z_k', \mtheta_k', \tS') \in \tP_b(\gamma)$ and corresponding mean tensors $\tX, \tX'$, we have 
-\begin{align}
-    \text{rank}(\mat_k(\tX) - \mat_k(\tX')) &\leq \text{rank}(\mat_k(\tX)) + \text{rank}(\mat_k(\tX)) \\
-    &\leq 2 r_k, \quad k \in [K].
+For arbitrary two sets of parameters $(z, \tS, \mtheta), (z', \tS', \mtheta') \in \tP(\gamma)$ and corresponding mean tensors $\tX, \tX'$, we have 
+\begin{align}
+    &\text{rank}(\mat_k(\tX) - \mat_k(\tX')) \\
+    & \quad \quad \leq \text{rank}(\mat_k(\tX)) + \text{rank}(\mat_k(\tX)) \\
+    &\quad \quad \leq 2 r, \quad k \in [K].
 \end{align}
@@
 \begin{equation}\label{eq:dtbm_tucker}
-    \tX - \tX' \in \tQ(2r_1, \ldots, 2 r_K),
-\end{equation}
-where $\tQ(r_1, \ldots, r_K) \coloneqq \{ \text{Tucker tensor with rank }(r_1, \ldots,  r_K) \} $.
+    \tX - \tX' \in \tQ(2r, \ldots, 2 r),
+\end{equation}
+where $\tQ(r, \ldots, r) \coloneqq \{ \text{Tucker tensor with rank }(r, \ldots,  r) \} $.
 
@@
      & \bbP(\FnormSize{}{\tX - \hat \tX_{ML}} \geq t) \\
-     &\leq 2  \bbP\of{ \sup_{ \tX, \tX' \in \tP(r_1, 
-\ldots, r_K)} \ang{ \frac{\tX - \tX'}{\FnormSize{}{\tX - \tX'}}, \tE  } \geq t } \\
-     &\leq 2  \bbP\of{ \sup_{ \tT \in \tQ(2r_1, \ldots, 2r_K) \cap \{ \onormSize{}{\tT}_F = 1\} } \ang{ \tT, \tE  } \geq t } \\
+     &\leq 2  \bbP\of{ \sup_{ \tX, \tX' \in \tP(r, 
+\ldots, r)} \ang{ \frac{\tX - \tX'}{\FnormSize{}{\tX - \tX'}}, \tE  } \geq t } \\
+     &\leq 2  \bbP\of{ \sup_{ \tT \in \tQ(2r, \ldots, 2r) \cap \{ \onormSize{}{\tT}_F = 1\} } \ang{ \tT, \tE  } \geq t } \\
      &\lesssim  \exp(-Kpr),
@@
 % if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{Jiaxin Hu} received the B.S.\ degree from Wuhan University in 2019 and Master's degree in statistics from the University of Wisconsin-Madison in 2020.
+\begin{IEEEbiographynophoto}{Jiaxin Hu} received the B.S.\ degree from Wuhan University in 2019 and Master's degree in statistics from the University of Wisconsin - Madison in 2020.
 She is currently a Ph.D.\ student with the Department of Statistics, University of Wisconsin-Madison.
-Her research interests mainly include tensor data and network analysis with applications in genetics and ecology.
+Her research interests mainly include tensor methods, network analysis, and applications in genetics and social science.
 \end{IEEEbiographynophoto}
@@
 
-\begin{IEEEbiographynophoto}{Miaoyan Wang} received the B.S.\ degree from Fudan University and the Ph.D.\ degree from the University of Chicago.
-She is currently the Assistant Professor with the Department of Statistics, University of Wisconsin-Madison.
-She received the NSF CAREER Award in 2022.
-Her work focuses on tensor/matrix decomposition, high-dimensional statistics, multilinear optimization, and applications in genetics and neuroimaging.
+\begin{IEEEbiographynophoto}{Miaoyan Wang} received B.S.
+in Mathematics from Fudan University in 2010 and Ph.D.
+in Statistics from the University of Chicago in 2015.
+She is currently an Assistant Professor of Statistics at the University of Wisconsin - Madison.
+She is also a faculty affiliate in Mathematical Foundations of Machine Learning and Institute for Foundations of Data Science.
+Her research lies in the intersection of statistics, machine learning, and genetics.
+She is the recipient of NSF CAREER Award (2022), Best Student Paper Awards (with her as advisor) from American Statistical Association (three times in 2021-2023) and New New England Statistical Society (2022).
 \end{IEEEbiographynophoto}

### 2023-01-24 tensor_clustering/IEEE/ieee_revision/response_s1.tex
organize

\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{caption}

\usepackage{xr-hyper}
% this is for overleaf cross reference
\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}

\usepackage[hidelinks]{hyperref}
\usepackage[margin=1in]{geometry}
% \geometry{
%  left=30mm,
%  right=30mm
%  }

\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage[authoryear]{natbib}
\usepackage{url}
\usepackage{enumitem}

\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{algorithm}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor=blue,citecolor=blue,urlcolor={black}}
\usepackage{lscape}
\usepackage{afterpage}
\usepackage{hyperref}

\usepackage{microtype}
\usepackage{wrapfig}
\allowdisplaybreaks
\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{verbatim}
\usepackage{array}

\usepackage{tabularx,booktabs}
\newcolumntype{Y}{>{\centering\arraybackslash}X}

\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{cor}{Corollary}

\theoremstyle{definition}
\newtheorem{assumption}{Assumption}
\newtheorem{defn}{Definition}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}
\newtheorem{condition}{Condition}
\newtheorem{conjecture}{Conjecture}
\newtheorem{property}{Property}
\newtheorem{hypothesis}{Hypothesis}

\newtheorem{innercustomgeneric}{\customgenericname}
\providecommand{\customgenericname}{}
\newcommand{\newcustomtheorem}[2]{%
  \newenvironment{#1}[1]
  {%
   \renewcommand\customgenericname{#2}%
   \renewcommand\theinnercustomgeneric{##1}%
   \innercustomgeneric
  }
  {\endinnercustomgeneric}
}

\newcustomtheorem{customexample}{Example}

\usepackage{appendix}
\usepackage{fullpage} 
\usepackage{microtype}
\usepackage{wrapfig}
\mathtoolsset{showonlyrefs}

\newcommand{\of}[1]{\left(#1\right)}
\newcommand{\off}[1]{\left[#1\right]}
\newcommand{\offf}[1]{\left\{#1\right\}}
\newcommand{\aabs}[1]{\left|#1\right|}
\newcommand{\ang}[1]{\left\langle#1\right\rangle}

\newcommand{\distgap}{D_{\ell_2}}
\newcommand{\Mat}{\text{Mat}}

\def\ci{\perp\!\!\!\perp}

\newtheorem{innercustomgeneric}{\customgenericname}
\providecommand{\customgenericname}{}
\newcommand{\newcustomtheorem}[2]{%
  \newenvironment{#1}[1]
  {%
   \renewcommand\customgenericname{#2}%
   \renewcommand\theinnercustomgeneric{##1}%
   \innercustomgeneric
  }
  {\endinnercustomgeneric}
}

\newcustomtheorem{customthm}{Theorem}
\newcustomtheorem{customlemma}{Lemma}

\usepackage{booktabs}
\newcommand\doubleRule{\toprule\toprule}
\allowdisplaybreaks

\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}

\input macros.tex
\newcommand*{\Scale}[2][4]{\scalebox{#1}{$#2$}}%
\setlength\parindent{0pt}
\usepackage[parfill]{parskip}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsmath}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\newcommand\Algphase[1]{%
\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\textwidth}{0.4pt}%
\Statex\hspace*{-\algorithmicindent}\textbf{#1}%
\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\textwidth}{0.4pt}%
}
\usepackage{caption}
\def\fixme#1#2{\textbf{\color{red}[FIXME (#1): #2]}}

\myexternaldocument{ieee_revise}

\title{\textbf{Response Letter}}
\date{}

\begin{document}

\maketitle
\vspace{-2cm}
\begin{center}
    \textbf{Response to Major Comments from Editor}
\end{center}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item \textit{The similarity of the manuscript to the previous paper ``Exact clustering in tensor block model: Statistical optimality and computational limit". Please clearly highlight the connections with \cite{han2020exact} and explain the differences.  }

\textbf{Response:} We have added a new Section~\ref{sec:tbm}, ``Comparison with Non-degree Tensor Block Model", to discuss the connections and difference between our work and \cite{han2020exact} from three aspects: signal notions, theoretical results, and algorithms. We quote the new section below:

\begin{quote}

``...We discuss the connections and difference between our dTBM and TBM~\citep{han2020exact} from three aspects: signal notions, theoretical results, and algorithms. Without loss of generality, let $\sigma^2=1$.

\begin{itemize}[wide]
    \item \textit{Signal notion.} The signal levels in both TBM~\citep{han2020exact} and our dTBM are functions of the core tensor $\tS$. We emphasize that the signal notions are different between the two models. In particular, the Euclidean-based signal notion in TBM~\cite{han2020exact} fails to accurately describe the phase transition in our dTBM due to the possible heterogeneity in degree $\mtheta$. To compare, we denote our angle-based signal notion in~\eqref{eq:gamma} and the Euclidean-based SNR in \cite{han2020exact} as $\Delta_{\text{ang}}^2$ and $\Delta_{\text{Euc}}^2$, respectively:
\begin{equation}
     \Delta_{\text{ang}}^2 =  2(1 - \max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}} ), \quad \Delta_{\text{Euc}}^2 = \min_{a \neq b \in [r]} \onormSize{}{\mS_{a:} - \mS_{b:}}^2.
\end{equation}
By Lemma~\ref{lem:norm_diff} in the Appendix II, we have 
\begin{equation}\label{eq:signalcompare}
     \Delta_{\text{ang}}^2  \max_{a \in [r]}\onormSize{}{\mS_{a:}}^2 \leq \Delta_{\text{Euc}}^2.
\end{equation}
The above inequality indicates the condition $\Delta_{\text{Euc}}^2 \leq p^{\gamma}$ is {\bf sufficient but not necessary} for $\Delta_{\text{ang}}^2 \leq p^{\gamma}$. In fact, if we were to use $\Delta_{\text{Euc}}^2$ for both models, then the phase transition of dTBM can be arbitrarily worse than that for TBM.

Here, we provide an example to illustrate the dramatical difference between TBM and dTBM with the same core tensor.

\begin{example}[Comparison with Euclidean-based signal notion] \label{example:euc_alg} Consider a biclustering model with $\mtheta=1$ and an order-2 core matrix 
\begin{equation}
    \mS = \begin{pmatrix} p^{(\gamma+1)/2 } + 2  & 2 p^{(\gamma+1)/2} + 4\\
    2 & 4
    \end{pmatrix},\quad \text{with}\ \gamma \leq -1.
\end{equation}
The core matrix $\mS$ lies in the parameter spaces of TBM and our dTBM. Here, the constraint $\gamma \leq -1$ is added to ensure the bounded condition of $\mS$ in our parameter space in \eqref{eq:family}. The angle-based and Euclidean-based signal levels of $\mS$ are 
\begin{equation}
    \Delta_{\text{ang }}^2(\mS) = 0 \ \left(\leq p^{\gamma}\right), \quad \Delta_{\text{Euc}}^2(\mS) = 5 p^{\gamma + 1} \ \left(\geq p^{\gamma}\right).
\end{equation}
We conclude that TBM with $\mS$ achieves exact recovery with a polynomial-time algorithm; see \citet[Theorem 4]{han2020exact}. By contrast, the dTBM with the same $\mS$ and input $r=2$ violets the identifiability condition, and thus fails to be solved by all estimators; see our Theorem~\ref{thm:unique}. 
\end{example}

    \item \textit{Theoretical results.} In both works, we study the phase transition of TBM and dTBM with respect to the Euclidean and angle-based SNRs. We briefly summarize the results in \cite{han2020exact} and compare with ours.

    \textit{Statistical critical value:}
    \begin{align}
        \text{Ours:}& \ \Delta_{\text{ang}}^2 \lesssim p^{-(K-1)} \Rightarrow \text{statistically impossible;} \quad \Delta_{\text{ang}}^2 \gtrsim   p^{-(K-1)} \Rightarrow \text{MLE achieves exact recovery;} \\
        \text{Han's:}& \ \Delta_{\text{Euc}}^2 \lesssim p^{-(K-1)} \Rightarrow \text{statistically impossible;} \quad \Delta_{\text{Euc}}^2 \gtrsim   p^{-(K-1)} \Rightarrow \text{MLE achieves exact recovery}.
    \end{align}

     \textit{Computational critical value:}
    \begin{align}
        \text{Ours:}& \ \Delta_{\text{ang}}^2 \lesssim p^{-K/2} \Rightarrow \text{computationally impossible;} \quad \Delta_{\text{ang}}^2 \gtrsim   p^{-K/2} \Rightarrow \text{ computationally efficient;} \\
        \text{Han's:}& \ \Delta_{\text{Euc}}^2 \lesssim p^{-K/2} \Rightarrow \text{computationally impossible;} \quad \Delta_{\text{Euc}}^2 \gtrsim   p^{-K/2} \Rightarrow \text{computationally efficient}.
    \end{align}

The above comparison reveals three major differences.

 %Our dTBM phase transition re-discovers the statistical-computational gaps in TBM. %This finding indicates the intrinsic behavior distinctions among higher-order, matrix, and vector problem may be generally exists in many higher-order problems.
First, none of our results in Section~\ref{sec:limits} are corollaries of \cite{han2020exact}. Both models show the similar conclusion {\bf but under different conditions}. While the TBM impossibility~\citep{han2020exact} provides a necessary condition for our dTBM impossibility, we find that such a condition is often loose. There exists a regime of $\tS$ in which TBM problems are computationally efficient but dTBM problems are statistically impossible; see Example~\ref{example:euc_alg}. This observation has motivated us to develop the new signal notion $\Delta^2_{\text{ang}}$ for sharp dTBM phase transition conditions.

 Second, to find the phase transition, we need to show both the impossibility {\bf and} achievability when SNR is below {\bf and} above the critical value, respectively. While the TBM impossibility can serve as a loose condition of our dTBM impossibility, more efforts are required to show the achievability. In particular, since TBM is a more restrictive model than dTBM, the achievability in \cite{han2020exact} does not imply the achievability of dTBM in a larger parameter space. The latter requires us to develop new MLE and polynomial algorithms for dTBM achievability.  %For impossibility, the minimax bounds in dTBM search over a larger range of estimators than TBM. Though the thresholds for TBM impossibility can serve as the loose lower bounds of the dTBM impossibility, more efforts are required to avoid the gap between impossibility and achievablity.

Third, from the perspective of proofs, we develop new tools to handle the extra degree heterogeneity. In our Theorem~\ref{thm:stats}, we consider the profile MLE by maximizing out the nuisance degree parameter, while TBM~\citep{han2020exact} considers the usual MLE without degree parameter. In our Theorem~\ref{thm:comp}, we construct a rank-2 tensor to relate HPC conjecture to $\Delta^2_{\text{ang}}$, while TBM~\citep{han2020exact} constructs a rank-1 tensor to relate HPC conjecture to $\Delta^2_{\text{Euc}}$. The asymptotic non-equivalence between $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$ renders our proof technically more involved.

    \item \textit{Algorithms.} Both \cite{han2020exact} and our work propose the two-step algorithm, which combines warm initialization and iterative refinement to achieve exact recovery. This local-to-global strategy is not new in clustering literature~\citep{gao2022iterative, chien2019minimax}. The highlight of our algorithm is the angle-based update in lines 10-14, Sub-algorithm~\hyperref[alg:main]{2}, which is specifically designed for dTBM to avoid the estimation of $\mtheta$. This angle-based update brings new proof challenges. We develop polar-coordinate based techniques to establish the error rate for the proposed algorithm. "
\end{itemize}

\end{quote}

\item \textit{Guarantees for the MLE and the claim that it matches the statistical lower bound. Please include an analysis of the MLE and clarify the claim.}

\textbf{Response:} We have added the MLE guarantee in the revised Theorem~\ref{thm:stats} in Section~\ref{sec:statlimit}.  The MLE error rate matches the statistical lower bound. Detailed proofs have been added to Appendix II, Section~\ref{sec:statprove2}. Inspired by this comment, we have also revised the exposition of Theorems~\ref{thm:stats}-\ref{thm:comp} to better present the stat-computational lower and upper bounds. 
%We also added a new Section~\ref{sec:limits}-A to discuss the preliminary settings in our work. 
We quote the revision in Section~\ref{sec:limits} here:

%New Section~\ref{sec:prelim}:
%\begin{quote}
%`` ...
%In our work, we give a special attention to the regime of balanced degree heterogeneity. We call the degree $\mtheta$ \emph{balanced} if
%\begin{equation}\label{eq:degree}
%{\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
%\end{equation}
%Following lemma shows the close relation between the angle gaps in mean tensor $\tX$ and core tensor $\tS$ under the balanced degree.

%\begin{lem}[Angle gap in $\tX$ and $\tS$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family}. Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying~\eqref{eq:degree}. Then, for all $i,j$ such that $z(i) \neq z(j)$

%\begin{equation}
  %\onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  } \asymp  \onorm{ \frac{\mS_{z(i):}}{\onormSize{}{\mS_{z(i):}}}  -  \frac{\mS_{z(j):}}{\onormSize{}{\mS_{z(j):}}}  },
%\end{equation}

%where $\mX =\mat(\tX)$ and $\mS = \mat (\tS)$.
%\end{lem}

%Intuitively, the performance of dTBM is determined by the signal in core tensor and the distribution of degree heterogeneity. Since we view the individual degree as a nuisance variable, our main interest focuses on the dTBM performance with varying core tensor signal. By Lemma~\ref{lem:angle_gap_x}, the weighted mapping from core tensor $\mS_{z(i):}$ to mean tensor $\mX_{z(i):}$ with balanced degree does not render the angular signal $\Delta_{\min}^2$ in core tensor. Therefore, the balance assumption helps to exclude the cases in which the degree heterogeneity brings dramatic effects to dTBM performance. Indeed, as we will see in following subsections, under the balance assumption~\eqref{eq:degree}, the core signal $\Delta_{\min}^2$ is sufficient to establish the dTBM phase transition.

%In contrast, the unbalanced degree may lead to an arbitrarily worse dTBM phase transition with respect to the core signal $\Delta_{\min}^2$. Here, we provide an example to illustrate the failure of $\Delta_{\min}^2$ with the unbalanced degree.

% The balance assumption \eqref{eq:degree} ensures the sufficiency of core tensor signal $\Delta_{\min}^2$ to describe the phase transition of dTBM.
% Without balanced degree,

%The dTBM phase transition without balanced degree may be arbitrary worse than that with balanced degree.

%\begin{example}[Failure of  $\Delta_{\min}^2$ with unbalanced degree] Consider the order-2 dTBM of 2 communities with core matrix
%\begin{equation}
%    \mS = \begin{pmatrix} 1 & a\\
%    1 & -a
%    \end{pmatrix}, \quad \text{and} \quad \mtheta \text{ such that } \onormSize{}{\mtheta_{z^{-1}(1)}}^2 = M \onormSize{}{\mtheta_{z^{-1}(2)}}^2,
%\end{equation}
%where $\tO(p^{-1}) \leq M \leq \tO(p)$ by the parameter space~\eqref{eq:family}. Let $\Delta_{\mX}^2$ denote the angle between the mean tensor, where 
%\begin{equation}\label{eq:delta_x}
%    \Delta_{\mX}^2 \coloneqq \min_{i,j \in [p], z(i) \neq z(j)} \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  }, \quad \mX = \mat(\tX). %\quad  \tX = \tS \times_1 \mTheta \mM \times_2 \cdots \times_K \mTheta \mM.
%\end{equation}
%With core matrix $\mS$, we have 
%\begin{equation}
%    \Delta_{\min}^2 = \frac{2 a^2}{1 + a^2}, \quad  \Delta_{\mX}^2 = \frac{2\onormSize{}{\mtheta_{z^{-1}(2)}}^{2} a^2}{ \onormSize{}{\mtheta_{z^{-1}(1)}}^{2} +  \onormSize{}{\mtheta_{z^{-1}(2)}}^{2}a^2}.
%\end{equation}
%Take $ a = p^{-1/4}, M = p^{1/2 + \epsilon}$ for some $\epsilon > 0$. We have \begin{equation}
%    \Delta_{\min}^2 = p^{-1/2} \quad \text{and} \quad    \Delta_{\mX}^2 = p^{-1-\epsilon}.
%\end{equation}
% \begin{align}
%     &\text{if } a = p^{-1/2}, M = 1& \quad &\Rightarrow \quad \Delta_{\min}^2 =  \Delta_{\mX}^2 = p^{-1}; \\
%     &\text{if } a = p^{-1/4}, M = p^{1/2} & \quad &\Rightarrow \quad \Delta_{\min}^2 = p^{-1/2}, \   \Delta_{\mX}^2 = p^{-1}.
% \end{align}
%By following Theorem~\ref{thm:stats}, no estimator exact recovers dTBM under $\Delta_{\mX}^2 < p^{-1}$. However, the condition $ \Delta_{\min}^2 = p^{-1/2}$ with balanced degree leads to a computationally efficient dTBM by Theorem~\ref{thm:refinement}.
%\end{example}

%\begin{rmk}[Flexibility in balanced degree] One important note is that the balance assumption~\eqref{eq:degree} does not preclude the dramatic degree heterogeneity. In fact, within each of the clusters, the highest degree can be $\theta(i) = \Omega(p)$, whereas the lowest degree can be $\theta(i)=\tO(1)$. This range constraints on $\mtheta$ is looser than previous work \citep{gao2018community} that restricts the highest degree in a sub-linear rate $o(p)$. 
%\end{rmk}

%\begin{rmk}[Similar assumption in literature]
%We propose he balance assumption~\eqref{eq:degree} to avoid the extreme inflation or elimination of the core tensor gap $\Delta_{\min}^2$ after the weighted mapping by degree heterogeneity. 
%Similar regulations are not rare in literature. In higher-order model \cite{ke2019community}, similar assumption ${\max_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}\leq C \min_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}$ is also made to ensure the degree distributions are balanced on each community. Since they have no $\ell_1$ constraint on $\mtheta$, their balance assumption is equivalent to ours~\eqref{eq:degree} multiplying some constants to the degrees. "
%\end{rmk}

%\end{quote}

New Section~\ref{sec:statlimit}:

\begin{quote}
``... We consider the Gaussian MLE, denoted as $(\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}})$, over the estimation space $\tP(\gamma)$, where 
\begin{equation}\label{eq:mle}
    (\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}}) = \argmin_{ (z, \mtheta, \tS) \in \tP (\gamma)} \onormSize{}{\tY - \tX(z,\mtheta, \tS)}_F^2.
\end{equation}

\begin{customthm}{2}[Statistical critical value] Consider general Gaussian dTBMs with parameter space $\tP(\gamma)$ with $K\geq 1$. Then, we have the following statistical phase transition.

\begin{itemize}[wide]
    \item \textbf{Impossibility.} 
%     Considerthe minimal gap between the mean tensor as 
% \begin{equation}
%     \Delta_{\mX}^2 \coloneqq \min_{i,j \in [p], z(i) \neq z(j)} \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  }, \quad \text{where} \quad \mX = \mat(\tX).
% \end{equation}
Assume $r\lesssim p^{1/3}$. If the signal exponent satisfies $\gamma < -(K-1)$, then, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation; that is,
\begin{equation}
  \gamma< -(K-1) \quad \Rightarrow  \inf_{\hat z_{\text{stat}}}\sup_{(z, \tS, \mtheta)\in\tP(\lambda)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
\end{equation}
\item \textbf{MLE achievability.} If the signal exponent satisfies $\gamma >-(K-1)+c$ for an arbitrary constant $c>0$, then, the MLE in~\eqref{eq:mle} achieves exact recovery in high probability; that is,
\begin{equation}
\gamma> -(K-1)+c \quad  \Rightarrow  \quad \ell(\hat z_{\text{MLE}}, z) \lesssim \text{SNR}^{-1}\exp\left(-{p^{K-1}\text{SNR}\over r^{K-1}}\right) \to 0,
\end{equation}
with probability going to 1. 
\end{itemize}
\end{customthm}

The proofs for the two parts in Theorem~\ref{thm:stats} are in the Appendix II, Section~\ref{sec:statprove1} and Section~\ref{sec:statprove2}, respectively. The first part of Theorem~\ref{thm:stats} demonstrates the impossibility of exact recovery when $\gamma < -(K-1)$. Inspection of the proof shows that the impossibility holds with or without the balanced degree assumption. %Moreover, with general degree, Theorem~\ref{thm:stats} also indicates the impossibility when the mean tensor gap is below the threshold $p^{-(K-1)}$. 
The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to MLE and trace maximization~\citep{ghoshdastidar2017uniform}.

The second part of Theorem~\ref{thm:stats} shows the exact recovery of MLE when the $\gamma>-(K-1)+c$ for an arbitrary constant $c>0$. Combining the impossibility and achievability results, we conclude that the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM with respect to our SNR. "
\end{quote}

New Section~\ref{sec:thm5}:
\begin{quote}
``...The proof of MLE error shares the similar idea as Theorems~\ref{thm:initial}-\ref{thm:refinement}. we first show a weaker polynomial rate for MLE, and then improve the rate from polynomial to exponential throughout the iterations. The only difference is that the MLE remains the same over iterations due to its global optimality. See Appendix II, Section~\ref{sec:statprove2} for the detailed proof.''
\end{quote}

New Section~\ref{sec:complimit}:

\begin{quote}
``...
\begin{customthm}{3}[Computational critical value] Consider general Gaussian dTBMs under the parameter space $\tP$ with $K\geq 2$. Then, we have the following computational phase transition.

\begin{itemize}[wide]
    \item \textbf{Impossibility.} Assume HPC conjecture holds. If the signal exponent satisfies $\gamma < -K/2$, then, no \emph{polynomial-time estimator} $\hat z_{\text{comp}}$ achieves exact recovery in expectation as $p\to \infty$; that is, 
\begin{align}
  \gamma<-K/2 \ \Rightarrow \ \liminf_{p\to \infty}\sup_{(z, \tS, \mtheta) \in \tP(\gamma)}  \bbE \left[ p\ell(\hat z_{\text{comp}}, z) \right]\geq 1.
\end{align}
\item \textbf{Polynomial-time algorithm achievability.} Suppose the parameter space satisfies $\gamma  >  -K/2+c$ for an arbitrary constant $c>0$. Furthermore, assume the degree is lower bounded in that $\min_{i\in[p]}\theta_i\geq c$ for some constant $c>0$, and $\mtheta$ satisfies the local linear stability in Definition~\ref{def:stable} in the neighborhood $\tN(z,\varepsilon)$ for $\varepsilon \geq \log^{-1}(p)$. Then, there exists a polynomial-time algorithm $\hat z_{\text{ploy}}$ that achieves exact recovery in high probability; that is,
\begin{equation}
     % \ell(\hat z_{\text{poly}}, z) \lesssim \exp \of{ - \frac{p^{K-1}}{r^{K-1}} \Delta_{\min}^2 }.
  \gamma>-K/2+c\ \Rightarrow   \  \ell(\hat z_{\text{poly}}, z) \lesssim \text{SNR}^{-1}\exp \of{ - \frac{p^{K-1}\text{SNR}}{r^{K-1}} } \to 0,
\end{equation}
with probability going to 1. 
\end{itemize}
\end{customthm}

The proofs for the two parts in Theorem~\ref{thm:comp} are in the Appendix II, Section~\ref{sec:compprove1} and Section~\ref{sec:statprove2}, respectively. The first part of Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$, and the second part shows the existence of such algorithm when $\gamma > -K/2+c$ under extra technical assumptions. In Section~\ref{sec:alg}, we will present an efficient polynomial-time algorithm in this setting. Therefore, we conclude that $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM with respect to our SNR. ''
\end{quote}

\end{enumerate}

\newpage

\begin{center}
    \textbf{Point-by-point Response to Reviewer 1}
\end{center}

\textit{The paper studies the problem of finding the planted block structure in a tensor, generated from the degree-corrected tensor block model (dTBM). This problem arises in different contexts, such as community detection in hypergraph, higher-order clustering etc. A simpler version of the problem (non-degree corrected TBM) has been studied in previous works. The main contribution of the paper is to extend the guarantees for recovering planted blocks in TBM to the case of dTBM. In particular, a suitable SNR is proposed and statistical lower bound (Thm 2) and computational lower bound under HPC (Thm 3) are derived. Thm 4-5 provide error rate for a poly-time algorithm.}

\textit{Overall, the paper is well-written and comes with additional numerical studies, also on real data sets. The contribution of the paper is quite good. There are few aspects that are not addressed in the paper (listed below). It is to be noted that all the mentioned may not be easily resolved, but brief discussions on these aspects would improve the paper.}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]

\item \textit{The statistical upper bound (guarantees for MLE) seem to be missing since Thm 4-5 focuses on a poly-time approach. Please explicitly describe if (and how) one can derived error rate for MLE from Thm 5.}

\textbf{Response:} Thank you for the suggestion. We have added the MLE guarantee in the revised Theorem~\ref{thm:stats}, Section~\ref{sec:statlimit}. Detailed proof has been added in the Appendix II, Section~\ref{sec:statprove2}. The proof shares the similar idea as Theorems~\ref{thm:initial}-\ref{thm:refinement}; the only difference is that the MLE remains the same over iterations due to its global optimality. See our response to \textbf{\# 2 Major Comments from Editor} for details.

%The MLE guarantee matches with the statistical lower bound, and thus rigorously supports our claim that the statistical critical value is $\gamma_{\text{stat}} = -(K-1)$.

\item \textit{While the dTBM model is general, it ignores the effect of sparsity in the context of hypergraphs. In case of community detection in sparse hypergraphs, $\bbE[\tY]$ includes a scalar that vanishes with number of nodes/tensor dimension. As in the case of graph SBM, the analysis gets a bit tricky for sparse hypergraphs (which is not covered in this paper). The authors could look at \cite{florescu2016spectral} which studies this setting and the proposed method also has similarities with the algorithm presented here.}

\textbf{Response:} We thank reviewer for the suggestions. Our current work focuses on \emph{dense} dTBM. While sparse dTBM is an interesting application, the theory and algorithm require different techniques. 
Inspired by the comment, we have extended our algorithm and experiment to sparse dTBM with a decaying sparse parameter. We adapt diagonal-deleted HOSVD as a new initialization strategy under extremely sparse settings.

We have added the following discussion in Section \ref{subsec:ber}:

\begin{quote}
    ``... \textit{Extension to sparse binary dTBMs.} The sparsity is often a popular feature in hypergraphs~\citep{florescu2016spectral,ke2019community}. Specifically, the sparse binary dTBM assumes that, the entries of $\tY$ follow independent Bernoulli distributions with the mean
     \begin{equation}\label{eq:sparse_dtbm}
    \bbE\tY = \alpha_p \tS \times_1 \mTheta \mM \times_2  \cdots \times_K \mTheta \mM,
\end{equation}
where the extra scalar parameter $\alpha_p \in (0,1]$ is function of $p$ that controls the sparsity. A smaller $\alpha_p$ indicates a higher level of sparsity. Our current work focuses on dense dTBM with $\alpha_p=1$.  While sparse dTBM is an interesting application, the algorithm and its analysis require different techniques. Below, we discuss possible modifications of the algorithm.

%The parameters $(z, \mtheta, \alpha_p \tS)$ are still identifiable up to permutation under our Assumption~\ref{assmp:min_gap}. However, the constant lower bound assumption on the core tensor is no longer suitable to describe the decaying sparse parameter. Therefore, our phase transition will not directly apply; we expect the extra sparsity parameter $\alpha_p$ should also be included in the signal notion, in order to characterize the phase transition.

The sparsity affects our initialization guarantee in our Theorem~\ref{thm:initial}. In our initialization, the spectral denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) implements matrix SVD to unfolded tensors. %, and Theorem~\ref{thm:initial} indicates the good performance under the dense case. 
However, SVD-based methods are believed to fail in extremely sparse SBM due to the localization phenomenon in the singular vectors \citep{florescu2016spectral}. Inspired by \cite{florescu2016spectral}, we adapt the diagonal-deleted HOSVD (D-HOSVD) \citep{ke2019community} as the initialization in our higher-order clustering.

The sparsity also affects the iteration guarantee in our Theorem~\ref{thm:refinement}. The decaying mean tensor leads to a worse statistical error of order $\tO( -\alpha_p p^{K-1})$ on $\hat \tX$. The theoretical analyses for sparse binary dTBM and algorithms are left as future direction. Instead, we add numerical experiments to evaluate the robustness of our algorithm and the improvement of D-HOSVD initialization in the sparse dTBM; see Appendix I."
\end{quote}

We have added a numerical experiment and Figure~\ref{fig:sparse} under the sparse binary dTBM in Appendix I:

\begin{quote}

\begin{figure}[h!]
    \centering
    \includegraphics[width=.8\columnwidth]{sparsity.pdf}
    \caption{CER comparison versus sparsity parameter $\alpha_p$ in $[0.05, 0.9]$. We set $p = 100, r = 5$, and $\gamma = -1.2$ under sparse binary dTBM.}
    \label{fig:sparse}
\end{figure}

`` ...The second additional experiment evaluates the algorithm performances under the sparse binary dTBM~\eqref{eq:sparse_dtbm}. We fix the signal exponent $\gamma = -1.2$ and vary the sparsity parameter $\alpha_p \in [0.05, 0.9]$. A smaller $\alpha_p$ leads to a higher probability of zero entries in the observation. In addition to the three algorithms mentioned in Section~\ref{subsec:comp}  (denoted {\bf \small Initialization}, {\bf \small dTBM}, and {\bf \small SCORE}), we consider other three algorithms based on the discussion in Section~\ref{subsec:ber}: 
  \begin{itemize} 
  \item \textbf{\small D-HOSVD}, the diagonal-deleted HOSVD in \cite{ke2019community}; 
  \item \textbf{\small D-HOSVD + Angle}, the combined algorithm of our angle-based iteration with initialization from \textbf{\small D-HOSVD};
  \item \textbf{\small SCORE + Angle}, the combined algorithms of our angle-based iteration with initialization from \textbf{\small SCORE}.
  \end{itemize}

Figure~\ref{fig:sparse} shows a slightly larger error in \textbf{\small dTBM} than that in \textbf{\small SCORE}, \textbf{\small D-HOSVD + Angle}, and \textbf{\small SCORE + Angle} under the sparse setting with $\alpha_p < 0.3$. 
%The outperformance of \textbf{\small D-HOSVD} based iterative methods confirms the benefits of diagonal deletion in the sparse case. Meanwhile, 
The small gap between  \textbf{\small dTBM} and other sparse-specific methods implies the robustness of our algorithm. In addition, comparing \textbf{\small SCORE} versus \textbf{\small SCORE + Angle} (or \textbf{\small D-HOSVD} versus \textbf{\small D-HOSVD + Angle}) indicates the benefit of our angle iterations under the sparse dTBM. In the intermediate and dense cases with $\alpha_p \geq 0.3$, our proposed \textbf{\small dTBM} has a clear improvement over others, which again verifies the success of our algorithm in dense settings."
\end{quote}

\item \textit{ (Critical) The statistical lower bound is derived under non-degree TBM (since all $\theta(i) = 1$). From this, it is still not clear whether the lower bound would signiﬁcantly increase when there is a skewed degree correction. The authors’ claim that the statistical lower bound matches MLE is diﬃcult to verify (since I could not ﬁnd an explicit result for MLE). Moreover, the stated upper bound (Thm 4-5) assume some uniformity in $\theta$ (see top of page-11), and so the (matched) upper/lower bounds could only be for the case of relatively homogeneous degrees. A more clear discussion on this would be useful. The present results leaves open the possibility that phase-transition SNR for dTBM could be arbitrarily worse than that of TBM, in which case making some assumption on balance is reasonable.}

\textbf{Response:}

% \fixme{Miaoyan}{Say ``There seem to be three questions in this comments. We address them one by one. 
% \begin{enumerate}
% \item Regarding ``whether the lower bound would significatly increase when there is a skewed degree correction''. (move your response here)
% \item Regarding ``explicit result for MLE'' (move your response here)
% \item Regarding ``The stated upper bound assume some...''. 
% \end{enumerate}}

We thank reviewer for the suggestions. There seem to be three questions in this comment. We address them one by one.

\begin{enumerate}
    \item Regarding ``whether the lower bound would significantly increase when there is a skewed degree correction". We have revised the proof of statistical minimax bound with relaxed degree assumptions. Briefly, we relax the previous non-degree construction $\theta(i)\equiv 1$ to the skewed-degree construction under assumption~\eqref{eq:degree}. The latter allows moderate degree heterogeneity; for example, within each of the cluster, we allow the highest degree to be $\theta(i) = \Omega(p)$, whereas the lowest degree to be $\theta(i)=\tO(1)$. The revised minimax bound does not increase in the presence of moderately skewed degree. See our new Theorem~\ref{thm:stats} in Section~\ref{sec:limits} and new proof in Appendix II, Section~\ref{sec:statprove2}.

    \item  Regarding ``explicit result for MLE". We have added the MLE guarantee in new Theorem~\ref{thm:stats}. The MLE error matches with statistical lower bound and ensures the dTBM achievability above the statistical critical value. We have also revised the exposure of Theorems~\ref{thm:stats}-\ref{thm:comp} to include both impossibility and achievability results. See our new Section~\ref{sec:statlimit} and Section~\ref{sec:complimit}. See our response to {\bf \# 2 Major Comments from Editor} for details.

    \item Regarding ``The stated upper bound assume some uniformity in $\theta$". We have added a new Section~\ref{sec:prelim} to discuss the rational for the balanced degree assumption~\eqref{eq:degree}. Briefly, we focus on the regime with balanced heterogeneity due to the relationship between dTBM performance and core tensor signal $\Delta_{\min}^2$. Indeed, without the balance assumption, the phase transition of dTBM can be arbitrarily worse than that of TBM. We quote our new discussion below:

New Section~\ref{sec:prelim}:
    \begin{quote}
    ``...In our work, we focus on the regime of balanced degree heterogeneity. We call the degree $\mtheta$ \emph{balanced} if
\begin{equation}\label{eq:degree}
{\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
\end{equation}
The following lemma provides the rational of balanced degree assumption. We show the close relation between angle gaps in the mean tensor $\tX$ and the core tensor $\tS$ under balanced degree heterogeneity.

\begin{lem}[Angle gaps in $\tX$ and $\tS$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family}. Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying~\eqref{eq:degree}. Then, for all $i,j$ such that $z(i) \neq z(j)$, we have
\begin{equation}
 % \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  } \asymp  \onorm{ \frac{\mS_{z(i):}}{\onormSize{}{\mS_{z(i):}}}  -  \frac{\mS_{z(j):}}{\onormSize{}{\mS_{z(j):}}}  },
 \cos(\mX_{i:}, \mX_{j:})\asymp  \cos(\mS_{z(i):}, \mS_{z(j):}),
\end{equation}

where $\mX =\mat(\tX)$ and $\mS = \mat (\tS)$.
\end{lem}

%Intuitively, the performance of dTBM is determined by both the core tensor and degree heterogeneity. Since we view degrees as nuisance parameters, our main interest is to characterize dTBM performance with respect to the signal $\Delta^2_{\min}$ in the core tensor. By Lemma~\ref{lem:angle_gap_x}, the mapping from the core tensor $\mS_{z(i):}$ to the mean tensor $\mX_{z(i):}$ preserves the angle information $\Delta_{\min}^2$ under balanced degree heterogeneity~\eqref{eq:degree}. Therefore, the balance assumption helps to exclude the cases in which the degree heterogeneity distorts the dTBM performance.

In practice, an estimation algorithm has access to a noisy version of $\tX$ but not $\tS$. Our goal is to establish the algorithm performance with respect to the signal $\Delta^2_{\min}$ in the core tensor. By Lemma~\ref{lem:angle_gap_x}, the mapping from the core tensor $\mS_{z(i):}$ to the mean tensor $\mX_{z(i):}$ preserves the angle information $\Delta_{\min}^2$ under balanced degree heterogeneity~\eqref{eq:degree}. Therefore, the balanced degree assumption helps to exclude the cases in which the degree heterogeneity distorts the algorithm garantees.

%Indeed, as we will see in Sections~\ref{sec:statlimit}-\ref{sec:complimit}, under the balance assumption~\eqref{eq:degree}, the $\Delta_{\min}^2$ is sufficient to establish the dTBM phase transition.

%In contrast, a highly-skewed degree violating~\eqref{eq:degree} may lead to a worse performance of dTBM than that of TBM. 
Here, we provide an example to illustrate the insufficiency of $\Delta_{\min}^2$ in the absence of balanced degrees.

% The balance assumption \eqref{eq:degree} ensures the sufficiency of core tensor signal $\Delta_{\min}^2$ to describe the phase transition of dTBM.
% Without balanced degree,

%The dTBM phase transition without balanced degree may be arbitrary worse than that with balanced degree.

\begin{example}[Insufficiency of $\Delta_{\min}^2$ in the absence of balanced degrees] Consider an order-2 $(p,p)$-dimensional dTBM with core matrix
\begin{equation}\label{eq:modelexample}
    \mS = \begin{pmatrix} 1 &a\\
    1 & -a
    \end{pmatrix}, \quad \text{and} \ \mtheta \text{ such that } \onormSize{}{\mtheta_{z^{-1}(1)}}^2 = p^m \onormSize{}{\mtheta_{z^{-1}(2)}}^2,
\end{equation}
where $m\geq 0$ is a scalar parameter controlling the skewness of degrees. Let $\Delta_{\mX}^2$ denote the angle gap of the mean tensor, defined by
\begin{equation}\label{eq:delta_x}
    \Delta_{\mX}^2 \coloneqq \min_{i,j \in [p], z(i) \neq z(j)} \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  }, \quad \mX = \mat(\tX). %\quad  \tX = \tS \times_1 \mTheta \mM \times_2 \cdots \times_K \mTheta \mM.
\end{equation}
Take $ a = p^{-1/4}$ in the model setup~\eqref{eq:modelexample}. We have 
\begin{equation}
    \Delta_{\min}^2 = \frac{2 a^2}{1 + a^2} \asymp p^{-1/2}, \quad  \Delta_{\mX}^2 = \frac{2\onormSize{}{\mtheta_{z^{-1}(2)}}^{2} a^2}{\onormSize{}{\mtheta_{z^{-1}(1)}}^{2} +  \onormSize{}{\mtheta_{z^{-1}(2)}}^{2}a^2} \asymp p^{-1/2-m}.
\end{equation}
%Take $ a = p^{-1/4}. We have 
%\begin{equation}
   % \Delta_{\min}^2 = p^{-1/2} \quad \text{and} \quad    \Delta_{\mX}^2 = {1\over p^{m-1/2}+1} = \min(p^{1/2-m},1). 
    %{2 a^2 \over M+a^2} = p^{-1-\epsilon}.
%\end{equation}
% \begin{align}
%     &\text{if } a = p^{-1/2}, M = 1& \quad &\Rightarrow \quad \Delta_{\min}^2 =  \Delta_{\mX}^2 = p^{-1}; \\
%     &\text{if } a = p^{-1/4}, M = p^{1/2} & \quad &\Rightarrow \quad \Delta_{\min}^2 = p^{-1/2}, \   \Delta_{\mX}^2 = p^{-1}.
% \end{align}
%By the following Theorem~\ref{thm:stats}, no estimator achieves exact recovery of dTBM under $\Delta_{\mX}^2 < p^{-1}$. 
%Under the balanced degree assumption, the dTBM is computationally solvable based on our Theorem~\ref{thm:comp} in Section~\ref{sec:complimit}.
Based on our theory in later Sections, the stat-computational achievability of dTBM occurs when $\Delta^2_{\mX}>p^{-1}$; that is, the dTBM estimation depends on the relative magnitude of $m$ vs. $1/2$. In such a setting, the proposed signal notion $\Delta^2_{\min}$ alone fails to fully characterize dTBM. %However, the condition $ \Delta_{\min}^2 = p^{-1/2}$ with balanced degree leads to a computationally efficient dTBM by Theorem~\ref{thm:refinement}.
%$-1\leq \varepsilon \leq 1/2$.
 %($\tO(p^{-1}) \leq M \leq \tO(p)$ by the parameter space~\eqref{eq:family}). 
\end{example}

\begin{rmk}[Flexibility in balanced degree assumption] One important note is that our balance assumption~\eqref{eq:degree} does not preclude the mild degree heterogeneity. In fact, within each of the clusters, we allow the highest degree at the order $\tO(p)$, whereas the lowest degree at the order $\Omega(1)$. This range is more relaxed than previous work \citep{gao2018community} that restricts the highest degree in the sub-linear regime $o(p)$ and the lowest degree at the order $\Omega(1)$. 
\end{rmk}

\begin{rmk}[Similar assumptions in literature]
%We propose he balance assumption~\eqref{eq:degree} to avoid the extreme inflation or elimination of the core tensor gap $\Delta_{\min}^2$ after the weighted mapping by degree heterogeneity. 
Similar degree regulations are not rare in literature. In higher-order tensor model \citep{ke2019community}, the degree assumption $\max_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}} \leq $ $ C \min_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}$ is made to ensure balancedness across communities. %Since they have no $\ell_1$ constraint on $\mtheta$, their balance assumption is equivalent to ours~\eqref{eq:degree} multiplying some constants to the degrees. " 
In \cite{gao2018community}, the degree distribution is restricted to ${1\over |z^{-1}(a)|}\sum_{i\in z^{-1}(a)}\theta_i=1+o(1)$ for all communities. 
\end{rmk}

    \end{quote}

\end{enumerate}

%Detailed revisions in Section~\ref{sec:limits} are in our response to \textbf{\# 2 Major Comments from Editor}.

% For the statistical lower bound, we have revised the proof of minimax bound with general degree heterogeneity and the lower bound would not significantly increases. We have added the MLE guarantee to ensure the achievability above the statistical critical value. We have also revised the exposure of Theorem~\ref{thm:stats} and Theorem~\ref{thm:comp} to include both impossibility and achievability results.

% We have changed the exposure of Theorems~\ref{thm:stats} and \ref{thm:comp} to include both the impossibility and achievement of dTBM problem when signal is below and above the critical values. We quote the revised paragraphs in Sections~\ref{sec:statlimit} and \ref{sec:complimit}.

%We focus on the regime with balanced $\mtheta$ due to our main interest on the relationship between dTBM performance and core tensor signal $\Delta_{\min}^2$. Indeed, without balance assumption, the phase transition of dTBM can be arbitrarily worse than that with balanced assumption or that of TBM. We have added a new Section~\ref{sec:prelim} to discuss the settings. \fixme{Miaoyan}{Quote your discussion here.}
% For the balance assumption on degree heterogeneity~\eqref{eq:degree}, we have added a preliminary section~\ref{sec:prelim} in Section~\ref{sec:limits} to discuss our signal notion and the balance regulation on $\mtheta$. The balance assumption is proposed to describe the performance of dTBM by the signal only in core tensor. We show an example to illustrate that the phase transition of dTBM would be arbitrary worse than TBM with unbalanced heterogeneity. Our statistical and computational limits are sharp under the regime of balance degree. We quote the discussion in the preliminary Section~\ref{sec:prelim} here:

% Our statistical lower bound is derived under the dTBM since we consider the estimator space without the equal degree assumption. The condition (i), $\theta(i) = 1$ for all $i \in [p]$, was assumed on true heterogeneity for technical convenience. We have now relaxed the equal degree condition (i) in the proof of Theorem 2 to avoid the ambiguity. We quote the revised proof below:

% \begin{quote}
%     ``
%     ... Specifically, we consider the estimation problem based on a particular parameter point $(\{z_k\}, \tS, \{\mtheta_k\})$ with the following two properties:
% \begin{equation}\label{eq:construction}
% \text{(i)}\ 
%     \Delta_{\min}\lesssim  \left({p\over r}\right)^{-\frac{K-1}{2}}\sigma; \quad  \text{(ii)}\ |z^{-1}_k(a)|={p \over r} \in \mathbb{Z}_+ \text{ for all }a\in[r],
% \end{equation}
% for all $k \in [K]$.

% Furthermore, we define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation. The construction of $T$ is precisely the same as \citet[Proof of Theorem 2]{gao2018community}. By the constraint $\onormSize{}{\mtheta_{z^{-1}(a)}}_1 = |z^{-1}(a)|, a \in [r]$ in parameter space (3), the $\max_{i \in T^c} \theta(i) \leq C$ for some positive constant $C$. Then, following the proof of~\citet[Theorem 2]{gao2018community}, ...."
% \end{quote}

% We have showed that MLE matches the statistical limit under the balance assumption in a new theorem. See our response to \textbf{\# 2 Major Comments from Editor.}

% \fixme{Jiaxin}{Need to rethink this question about the signal notion and phase transition.}
% {
% \color{red}

% The phase transition for dTBM is arbitrary worse than that of TBM, no matter with angle-based or Euclidean-based SNR. See Example 1. That's why we need to propose angle based SNR. 
% %To see the necessity of balance, we need to show the SNR phase transition will arbitrary worse than that with balance assumption.

% Notice that 
% \begin{equation}
%     \{\text{$\mtheta$ is balanced} ,  \Delta_{\mS, \min}^2 \coloneqq \min \onormSize{}{\mS_{a:}^s - \mS_{b:}^s}^2 \leq p^{\gamma} \} \quad \subset \quad \{ \Delta_{\mA, \min}^2 \coloneqq \min \onormSize{}{\mA_{a:}^s - \mA_{b:}^s}^2 \leq  p^{\gamma}\},
% \end{equation}
% where $\mA = \mS (\mM^T \mTheta^T)^{\otimes K-1}$.

% Our current statistical and computational limits are sharp under the former notion, $\Delta_{\mS, \min}^2$ with balance. The statistical lower bound and initialization guarantee still hold with latter notion $\Delta_{\mA, \min}^2$; but the computational limits, algorithm guarantees, MLE are not easy to extend with $\Delta_{\mA, \min}^2$.

% Specifically, in the computational lower bound,  we pick up a special example with $\theta = 1$. This special case lies in the parameter spaces generated by both signal notions. We apply Han's result to the special case, and this indicates computational lower bound can be much higher than current one; i.e., when $\Delta_{\mS, \min}^2$ with balance or $\Delta_{\mA, \min}^2 $ is larger than $p^{-K/2}$, it still may be computationally inefficient. The good news is: when $\Delta_{\mS, \min}^2 = p^{-K/2} \log p$ with balance, we have a polynomial time algorithm achieves exact recovery. Therefore, the threshold  $p^{-K/2}$ is sharp if we use the signal notion $\Delta_{\mS, \min}^2$ with balance.

%  In the iteration, we do not know the $\mtheta$ in practice, do not estimate $\tX$ in algorithm, and need the gap between folded vectors $\mS_{a:}$'s. Hence, it may not be easy to show the algorithm guarantee with later notion $\Delta_{\mA, \min}^2$; the MLE upper bound relies on iteration results  may also not be easy to extend.

% }

\item \textit{Please provide some intuition on Deﬁnition 2, particularly Eqn (10).}

\textbf{Response:} We have added more explanations for Definition~\ref{def:stable} and equation~\eqref{eq:local} in Section~\ref{subsec:angle}:
\begin{quote}
    ``... Roughly speaking, the vector $\mp(\bar z)$ represents the raw cluster sizes, and $\mp_\theta(\bar z)$ represents the relative cluster sizes weighted by degrees. ...The condition~\eqref{eq:local} controls the impact of node degree to the $\mp_{\theta}(\cdot)$ with respect to the misclassification rate $\varepsilon$ and angle gap $\Delta_{\min}$. Intuitively, the condition~\eqref{eq:local} controls the skewness of degree so that the angle between raw cluster size and degree-weighted cluster size is well controlled. The stability assumption is proposed for technical convenience, and we relax this condition in numerical studies; see Section~\ref{sec:simulation}."
\end{quote}

%We thank reviewer for the comment. Previous Definition 2 and equation (10) ensures the small gap between raw cluster size $|\hat z^{-1}(b)|$ and weighted cluster size $|\sum_{i \in \hat z^{-1}(b)} \theta(i)|$. With a better technique, we have shown the Definition 2 and the local linear stability condition on the heterogeneity $\mtheta$ are unnecessary for the algorithm guarantee. We have removed related parts from our manuscript.
\end{enumerate}

\newpage 
\begin{center}
    \textbf{Point-by-point Response to Reviewer 2}
\end{center}

\textit{Comments to the Author: This manuscript studies clustering under degree-corrected tensor block models (dTBMs). The dTBM is a generalization of the degree-corrected stochastic block model (DCSBM) for matrices/networks. It is also a generalization of the (non-degree) tensor block model (TBM) which is a higher-order extension of the stochastic block model (SBM).}

\textit{Contributions: The identifiability for the uniqueness of clustering under dTBMs is studied.
Statistical and computational lower bounds for clustering under dTBMs are established.
A two-stage, angle-based algorithm is proposed to achieve exact clustering in polynomial time under mild conditions. 
The last contribution is the most important one. The algorithm proposed uses the ``global-to-local strategy" which is commonly seen in community detection literature. An angle-based clustering is developed to handle the degree heterogeneity of dTBMs. Then theoretical analysis shows that it provably achieves the exact clustering. I feel this is a solid contribution, making the manuscript worth publication.}

\textit{Nevertheless, I have a few concerns and comments.}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]

\item \textit{One concern I have is that this manuscript is very similar to one of the authors' previous paper \cite{han2020exact} on clustering under the non-degree tensor block model. They share overall the same spirit, taste, and structure. For example:
The sessions for the computational and statistical lower bounds in this manuscript are very similar to the corresponding sessions in \cite{han2020exact} and they have the same results.
\cite{han2020exact} also gives a two-stage algorithm using the ``global-to-local strategy" and also achieves the exact clustering. \cite{han2020exact} gives the same threshold for the exact clustering as in this manuscript.
I think this manuscript should point out its connection with \cite{han2020exact}. The current version mentions that the method of \cite{han2020exact} fails under dTBMs but I think \cite{han2020exact} deserves more discussion given the similarity.}

\textbf{Response:} Thank you for your suggestion. We intentionally make the paper structure similar to~\cite{han2020exact} in order for easier comparison with non-degree models. Our results (e.g. the stat-comp lower bounds, the threshold for exact clustering) are all different from~\cite{han2020exact}. The confusion may come from the choice of our exposition: we chose the same notations as in~\cite{han2020exact} but actually they represent different quantities.

We have added a new Section~\ref{sec:tbm}, ``Comparison with Non-degree Tensor Block Model'', to discuss the connection and difference between our dTBM work and TBM in \cite{han2020exact} from three aspects: signal notions, theoretical results, and algorithms. Please see our response to {\bf \# 1, Major Comments from Editor} for details.

Briefly, we separate the discussion from three aspects: signal notion, theoretical results, and algorithm. We develop an angle-based signal notion to address the degree heterogeneity in dTBM. Our theoretical results are different with, and not easy extension of, the result in \cite{han2020exact}. Our algorithm share the classical local-to-global idea, but our angle-based iteration is specifically designed for dTBM and requires new proof techniques.  
%Particularly, our results in Section~\ref{sec:limits} are different with the result in \cite{han2020exact}. Though the critical values look the same, different signal notions lead to different interpretations of the results. The different parameter spaces and identifiability conditions also require different procedures to obtain the critical values. Also, our two-stage algorithm is different than that in \cite{han2020exact}. The angle-based update is specially designed for our degree-corrected model and brings a big challenge to establish the algorithm guarantee with polar coordinate based analysis.

\item \textit{My next comment is about whether Session III B (computational lower bound) can be much simplified. My understanding is that the non-degree tensor block model is a special case of dTBMs, and hence lower bounds givens in \cite{han2020exact} also hold under dTBMs. As a result, I also feel the 2nd contribution this manuscript claim (ie, statistical and computational lower bounds) are relatively minor.}

\textbf{Response:} Thank you for your suggestion. The Session III B (computational lower bound) cannot be derived from earlier non-degree work~\citep{han2020exact}. The short answer is that, our results show the similar conclusion but {\bf under different conditions.} While the TBM impossibility~\citep{han2020exact} provides a necessary condition for our dTBM impossibility, we find that {\bf such a condition is loose} in our contexts. This has motivated to develop new techniques to derive sharp lower bounds.

Inspired by your comment, we have revised the exposition of our theory. See our response to \textbf{\# 1 Major Comments from Editor} for details.

We quote the newly-added texts here:
\begin{quote}
    ``...\emph{Theoretical results.} In both works, we study the phase transition of TBM and dTBM with respect to the Euclidean and angle-based SNRs. We briefly summarize the results in \cite{han2020exact} and compare with ours.

    \textit{Statistical critical value:}
    \begin{align}
        \text{Ours:}& \ \Delta_{\text{ang}}^2 \lesssim p^{-(K-1)} \Rightarrow \text{statistically impossible;} \quad \Delta_{\text{ang}}^2 \gtrsim   p^{-(K-1)} \Rightarrow \text{MLE achieves exact recovery;} \\
        \text{Han's:}& \ \Delta_{\text{Euc}}^2 \lesssim p^{-(K-1)} \Rightarrow \text{statistically impossible;} \quad \Delta_{\text{Euc}}^2 \gtrsim   p^{-(K-1)} \Rightarrow \text{MLE achieves exact recovery}.
    \end{align}

     \textit{Computational critical value:}
    \begin{align}
        \text{Ours:}& \ \Delta_{\text{ang}}^2 \lesssim p^{-K/2} \Rightarrow \text{computationally impossible;} \quad \Delta_{\text{ang}}^2 \gtrsim   p^{-K/2} \Rightarrow \text{ computationally efficient;} \\
        \text{Han's:}& \ \Delta_{\text{Euc}}^2 \lesssim p^{-K/2} \Rightarrow \text{computationally impossible;} \quad \Delta_{\text{Euc}}^2 \gtrsim   p^{-K/2} \Rightarrow \text{computationally efficient}.
    \end{align}

The above comparison reveals three major differences.

 %Our dTBM phase transition re-discovers the statistical-computational gaps in TBM. %This finding indicates the intrinsic behavior distinctions among higher-order, matrix, and vector problem may be generally exists in many higher-order problems.
First, none of our results in Section~\ref{sec:limits} are corollaries of \cite{han2020exact}. Both models show the similar conclusion {\bf but under different conditions}. While the TBM impossibility~\citep{han2020exact} provides a necessary condition for our dTBM impossibility, we find that such a condition is often loose. There exists a regime of $\tS$ in which TBM problems are computationally efficient but dTBM problems are statistically impossible; see Example~\ref{example:euc_alg}. This observation has motivated us to develop the new signal notion $\Delta^2_{\text{ang}}$ for sharp dTBM phase transition conditions.

 Second, to find the phase transition, we need to show both the impossibility {\bf and} achievability when SNR is below {\bf and} above the critical value, respectively. While the TBM impossibility can serve as a loose condition of our dTBM impossibility, more efforts are required to show the achievability. In particular, since TBM is a more restrictive model than dTBM, the achievability in \cite{han2020exact} does not imply the achievability of dTBM in a larger parameter space. The latter requires us to develop new MLE and polynomial algorithms for dTBM achievability.  %For impossibility, the minimax bounds in dTBM search over a larger range of estimators than TBM. Though the thresholds for TBM impossibility can serve as the loose lower bounds of the dTBM impossibility, more efforts are required to avoid the gap between impossibility and achievablity.

Third, from the perspective of proofs, we develop new tools to handle the extra degree heterogeneity. In our Theorem~\ref{thm:stats}, we consider the profile MLE by maximizing out the nuisance degree parameter, while TBM~\citep{han2020exact} considers the usual MLE without degree parameter. In our Theorem~\ref{thm:comp}, we construct a rank-2 tensor to relate HPC conjecture to $\Delta^2_{\text{ang}}$, while TBM~\citep{han2020exact} constructs a rank-1 tensor to relate HPC conjecture to $\Delta^2_{\text{Euc}}$. The asymptotic non-equivalence between $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$ renders our proof technically more involved...''
   \end{quote}

   \begin{quote}
   ``... \emph{Signal Notion.} We emphasize that the signal notions are different between the two models. In particular, the Euclidean-based signal notion in TBM~\cite{han2020exact} fails to accurately describe the phase transition in our dTBM due to the possible heterogeneity in degree $\mtheta$. To compare, we denote our angle-based signal notion in~\eqref{eq:gamma} and the Euclidean-based SNR in \cite{han2020exact} as $\Delta_{\text{ang}}^2$ and $\Delta_{\text{Euc}}^2$, respectively:
\begin{equation}
     \Delta_{\text{ang}}^2 =  2(1 - \max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}} ), \quad \Delta_{\text{Euc}}^2 = \min_{a \neq b \in [r]} \onormSize{}{\mS_{a:} - \mS_{b:}}^2.
\end{equation}
By Lemma~\ref{lem:norm_diff} in the Appendix II, we have 
\begin{equation}\label{eq:signalcompare}
     \Delta_{\text{ang}}^2  \max_{a \in [r]}\onormSize{}{\mS_{a:}}^2 \leq \Delta_{\text{Euc}}^2.
\end{equation}
The above inequality indicates the condition $\Delta_{\text{Euc}}^2 \leq p^{\gamma}$ is {\bf sufficient but not necessary} for $\Delta_{\text{ang}}^2 \leq p^{\gamma}$. In fact, if we were to use $\Delta_{\text{Euc}}^2$ for both models, then the phase transition of dTBM can be arbitrarily worse than that for TBM.

Here, we provide an example to illustrate the dramatical difference between TBM and dTBM with the same core tensor.

\begin{example}[Comparison with Euclidean-based signal notion] \label{example:euc_alg} Consider a biclustering model with $\mtheta=1$ and an order-2 core matrix 
\begin{equation}
    \mS = \begin{pmatrix} p^{(\gamma+1)/2 } + 2  & 2 p^{(\gamma+1)/2} + 4\\
    2 & 4
    \end{pmatrix},\quad \text{with}\ \gamma \leq -1.
\end{equation}
The core matrix $\mS$ lies in the parameter spaces of TBM and our dTBM. Here, the constraint $\gamma \leq -1$ is added to ensure the bounded condition of $\mS$ in our parameter space in \eqref{eq:family}. The angle-based and Euclidean-based signal levels of $\mS$ are 
\begin{equation}
    \Delta_{\text{ang }}^2(\mS) = 0 \ \left(\leq p^{\gamma}\right), \quad \Delta_{\text{Euc}}^2(\mS) = 5 p^{\gamma + 1} \ \left(\geq p^{\gamma}\right).
\end{equation}
We conclude that TBM with $\mS$ achieves exact recovery with a polynomial-time algorithm; see \citet[Theorem 4]{han2020exact}. By contrast, the dTBM with the same $\mS$ and input $r=2$ violets the identifiability condition, and thus fails to be solved by all estimators; see our Theorem~\ref{thm:unique}.''
\end{example}

   \end{quote}
   We have revised our Section~\ref{sec:limits} to better present the lower bounds. See our response to \textbf{\# 1 Major Comments from Editor}.

\item \textit{My third comment is on Condition (7), i.e., ``balance" of $\theta$. This is quite artificial and impractical and is needed only for technical reasons in the theoretical analysis of the proposed algorithm. In fact, as far as I know, this (or similar condition) is not needed in DCSBM literature for community detection. I suspect (7) can be removed by some better proof technique or modification of the proposed method.}

\textbf{Response:} Thank you for your suggestion. See our response to \textbf{\#3(a) and 3(c), Reviewer 1} (page 7 of this response letter) for the detailed response.

While we agree the assumption is more for technical reasons, we suspect some ``balanced'' assumption or its variation seems necessary (and common in literature) for the achievability of computational critical values. 
Our constraints in $\tP$ are mild compared with previous literature. We have also added following comparison in Section~\ref{subsec:identify}:
\begin{quote}
    \begin{table}[h]
    \centering
    \resizebox{\textwidth}{!}{
    \begin{tabular}{c|c ccc}
    \hline
    Assumptions in parameter space& \cite{gao2018community}& \cite{han2020exact}& \cite{ke2019community} &Ours\\
    \hline
         Balanced community size &$\surd$ &$\surd$&$\surd$ & $\surd$  \\
        Balanced degree assumption & $\surd$&-&$\surd$ &$\surd$\\
         Bounded core tensor assumption & $\surd$ & $\times$ & $\surd$ &$\surd$\\
         Flexible in-group connections &$\times$ &$\surd$&$\surd$& $\surd$\\
         Gap among cluster centers & In-between cluster difference & Euclidean gap & Eigen gap & Angle gap\\
         \hline
    \end{tabular}
    }
    \caption{Parameter space comparison between previous work with our assumption.}
\end{table}

%We have added a new preliminary Section~\ref{sec:prelim} to discuss the regime with balanced $\mtheta$. The balance assumption is proposed to avoid the arbitrary bad phase transition of dTBM with respect to our signal notion. Similar regulations on degree also implicitly occur in DCSBM literature \cite{gao2018community, ke2019community}, in which more stringent constraints are applied on core tensor than ours. See our response to \textbf{ \# 3, Reviewer 1} for detailed discussion on balance assumption.

\end{enumerate}
\end{enumerate}

\newpage
\begin{center}
    \textbf{Point-by-point Response to Reviewer 3}
\end{center}

\textit{In general, this paper is well written and organized, numeric studies including simulation and real data application illustrate that the proposed model and method do have some merits. But I have several concerns regarding the theory and technical contribution.}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item \textit{Firstly, the mathematical formulation lacks motivation. From my knowledge, most existing degree-corrected block models are proposed for network data and Bernoulli model. While the proposed paper can be applied to Bernoulli as it could be taken as a special sub-Gaussian but I believe such treatment would not be suitable and/or optimal if one only cares about Bernoulli setting. Also, I notice several formulation inconsistencies. {For example, the author said they focus on symmetric tensor, but they assume the noise tensor $\mathcal{E}$ consists of independent entries, which does not make sense.}}

\textbf{Response:} Thank you for your suggestions. There seem to be three questions in this comment; we address them one-by-one.

\begin{enumerate}
\item Regarding ``motivation of general sub-Gaussian model''.
Our proposal of sub-Gaussian dTBM is intended to cover multiple types of data in applications. Continuous tensors arise commonly in fields such as genetics and medical image studies. From theoretical perspective, our sub-Gaussian framework allows both binary and continuous observations. Conversely, binary-specific methods may not be suitable for general (continuous) sub-Gaussian data. %We have added a motivating example in Section~\ref{subsec:motiv} to emphasize the popularity of continuous data. 
%\begin{quote}
   % `` ...
    %\textit{Gaussian tensor mixture model:} The Gaussian mixture model is a common model in pattern recognition and image processing. We say an order-$K$ sample tensor $\tY \in \bbR^{p_1 \times \cdots \times p_K}$ follows a tensor Gaussian distribution with mean tensor $\tS \in \bbR^{p_1 \times \cdots \times p_K}$ if $\text{vec}(\tY)$ is a Gaussian vector with mean $\text{vec}(\tS)$. Tensor Gaussian mixture model assumes the tensor data $\tY_i$ for all $i \in [N]$ follow the mixture tensor Gaussian distribution with density $f(\tY_i) = \sum_{a = 1}^r \pi_a \phi_a(\tS_a), \quad i \in [N]$, where $r$ is the number of clusters, $\pi_a$ is the mixture weight, $\tS_a$ is the mean tensor, and $\phi_a$ denotes the tensor Gaussian density of the $a$-th cluster with mean $\tS_a$ for all $a \in [r]$.

%Let $\tY \in \bbR^{N \times p_1 \times \cdots \times p_K }$ denote the combined data with slices $\tY(i,:) = \tY_i$ and $\tS \in \bbR^{r \times p_1 \times \cdots \times p_K}$ denote the combined mean tensor with slices $\tS(a,:) = \tS_a$. The tensor Gaussian mixture model with prior probability $\pi_a$ is equal to a special dTBM model
%\begin{equation}
   % \bbE[\tY] = \tS \times_1 \mM \times_2 \mI_{p_1} \times_3 \cdots \times_{K+1} \mI_{p_K},
%\end{equation}
%where $\mI_p$ is the identity matrix of dimension $p$, $\mM \in \{0,1\}^{N\times r}$ is the membership matrix for $N$ samples such that $\pi_a = |\{ i \in [N]: \mM_{i,a} = 1\} |/N$ for all $a \in [r]$. Assume that $\tS$ has a community structure. The tensor Gaussian mixture further becomes 
%\begin{equation}
   % \bbE[\tY] = \tC \times_1 \mM \times_2 \mTheta_1 \mM_1 \times_3 \cdots \times_{K+1} \mTheta_K \mM_K,
%\end{equation}
%where $\{ \mTheta_k, \mM_k\}_{k=1}^K$ refers to the heterogeneity and membership on the mean of Gaussian tensors. "
%\end{quote}

% First, our sub-Gaussian dTBM framework is well-motivated from the real life applications with multiple types of data. The general framework unifies a wide range of higher-order clustering models and facilities a comprehensive analysis for distinct observations. In contrast, the performances of existing Bernoulli methods are not guaranteed with general sub-Gaussian data. We have added more motivating examples to show and emphasize the importance of model flexibility. 
% {
% \color{blue}
% \begin{example}[Higher-order degree-corrected Gaussian mixture model] Gaussian mixture model is widely applied in applications including pattern recognition, image processing, and machine learning. We say an order-$K$ sample tensor $\tX \in \bbR^{p_1 \times \cdots \times p_K}$ follows a tensor Gaussian distribution with mean tensor $\tS \in \bbR^{p_1 \times \cdots \times p_K}$, denoted as $\tX \sim \tN(\tS; \mI_{p_1}, \ldots, \mI_{p_K})$, if $\text{vec}(\tY) \sim \tN( \text{vec}(\tS), \mI_{p_1 \cdots p_K})$, where $\mI_p$ is the identity matrix of dimension $p$.

% Higher-order degree-corrected Gaussian mixture model assumes the tensor data $\tY_i \in \bbR^{p_1 \times \cdots \times p_K}$ for all $i \in [N]$ follow the mixture tensor Gaussian distribution with density 
% \begin{equation} \label{eq:gaussian_mix}
%     f(\tY_i) = \sum_{a = 1}^r \pi_a \phi_a( \theta_i \tS_a; \mI_{p_1}, \ldots, \mI_{p_K}), \quad i \in [N],
% \end{equation}
% where $\theta_i$ is the heterogeneity of the $i$-th observation, $r$ is the number of clusters, $\pi_a$ is the mixture weight, $\tS_a$ is the mean tensor  and $\phi_a$ denotes the tensor Gaussian density of the $a$-th cluster, respectively, for all $r \in [a]$.

% Let $\tY \in \bbR^{N \times p_1 \times \cdots \times p_K }$ denote the combined data with slices $\tY(i,:) = \tY_i$. Let $\tS \in \bbR^{r \times p_1 \times \cdots \times p_K}$ denote the combined mean tensor with slices $\tS(a,:) = \tS_a$. Further, we equip the mean tensor with community structure 
% \begin{equation}
%     \tS = \tC \times_2 \mM_1 \times_3 \cdots \times_{K+1} \mM_K,
% \end{equation}
% where $\tC \in \bbR^{r \times r_1 \times \cdots \times r_K}$ is the core tensor, and $\mM_k \in \{0,1\}^{p_k \times r_k}$ for all $k \in [K]$ are membership matrices. Consider the dTBM model
% \begin{equation}\label{eq:gaussian_mix_tbm}
%     \bbE[\tY] = \tC \times_1  \mTheta \mM \times_2 \mM_1 \times_3 \cdots \times_{K+1} \mM_K,
% \end{equation}
% where $\mTheta = \text{diag}(\theta_1, \cdots ,\theta_N)$ is the heterogeneity vector, $\mM \in \{0,1\}^{N \times r}$ is the membership matrix for $N$ samples. The higher-order degree-corrected Gaussian mixture model~\eqref{eq:gaussian_mix} is equivalent to our dTBM model~\eqref{eq:gaussian_mix_tbm} with prior probability $\pi_a = |\{ i \in [N]: \mM_{i,a} = 1\} |/N$ for all $a \in [r]$.
% \end{example}
% }
\item Regarding ``sub-optimality of Bernoulli as a special sub-Gaussian model''. We discuss below which of our results are still optimal for Bernoulli data, and which are not.

\begin{itemize}
\item {\bf sub-Algorithm I.} Our initialization gaurantee is not suitable for Bernoulli data. To address this, we have proposed a remedy for Bernoulli initialization with square unfolding; see Section~\ref{subsec:ber}. The Bernoulli initialization error rate $\tO(p^{-\floor{K/2}})$ is slightly weaker than Gaussian initialization error rate $\tO(p^{-K/2})$. Combining the Bernoulli initialization with sub-Algorithm II, our method achieves an exponential error rate for Bernoulli data.
%and beats the state-of-art~\citep{ke2019community}.

\item  {\bf sub-Algorithm II.} Our angle-based iteration guarantee is suitable for Bernoulli data. With Bernoulli data and a proper initialization, our iteration reduces the misclustering error from a polynomial rate to an exponential rate $\tO(\exp(-p^{K-1}))$, while the state-of-art algorithm for binary hypergraph \citep{ke2019community} leads to a polynomial error rate $\tO(p^{-2})$.

\item {\bf Theorems~\ref{thm:stats}-\ref{thm:comp}.} The phase transitions are not applicable for Bernoulli data. The analyses of statistical phase transition and computational impossibility are based on Gaussian noise, and the polynomial-time algorithm achievability requires a larger SNR of order $\tO(p^{-\floor{K/2}})$ for Bernoulli data. We have added a numerical experiment in Appendix I to verify the existence of statistical-computational gap in Bernoulli tensor clustering. 
\end{itemize}

New Section~\ref{subsec:ber}: 
\begin{quote}
    ``...

    \textit{Initialization guarantee for Bernoulli data.} The main difficulty to establish the initialization guarantee for Bernoulli observations lies in the denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}). We now provide a high-level explanation for the technical difficulty when applying Theorem~\ref{thm:initial} to Bernoulli observations.

    The derivation of Theorem~\ref{thm:initial} relies on the upper bound of the estimation error for the mean tensor in Lemma~\ref{lem:two-step_esterror}; i.e., with high probability
\begin{equation}\label{eq:prop1}
    \onormSize{}{\hat \tX - \tX}_F^2 \lesssim p^{K/2},
\end{equation}
where $\tX = \bbE\tY$ and $\hat \tX$ is defined in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}. Unfortunately, the inequality~\eqref{eq:prop1} holds only for i.i.d.\ sub-Gaussian observations, while Bernoulli observations are generally not identically distributed.

One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square} of Bernoulli tensor $\tY$. We summarize the procedure in sub-algorithm 3.

\begin{algorithm}[h!]
\caption*{\bf Sub-algorithm 3: Weighted higher-order initialization for Bernoulli observation}
\vspace{.15cm}
\begin{algorithmic}[1] 
\INPUT Bernoulli tensor $\tY \in \{0,1\}^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
\State  Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{\floor{p^{K/2}} \times \ceil{p^{K/2}} }$ denote the nearly square unfolded tensor. Compute the estimate $\tX'$, where
\begin{equation}\label{eq:matrixsvd}
    \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX) -  \mat_{sq}(\tY)}_F^2.
\end{equation}
\State Implement lines 3-5 of Sub-algorithm~\hyperref[alg:main]{1} with $\hat \tX$ replaced by $\hat \tX'$ in~\eqref{eq:matrixsvd}.
\OUTPUT Initial clustering $z^{(0)} \leftarrow \hat z$.
\end{algorithmic}
\end{algorithm}

\begin{prop}[Error for Bernoulli initialization]\label{cor:ber} Consider the Bernoulli dTBM in the parameter space $\tP$ and Assumption~\ref{assmp:min_gap} holds. Assume that $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$. Let $ z^{(0)}$ denote the output of Sub-algorithm~3. With probability going to 1, we have
\begin{equation}\label{eq:ini_b}
 \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}, \quad \text{and} \quad L(z^{(0)}, z) \lesssim  {\sigma^2 r^K p^{-\floor{K/2}}}.
\end{equation}
\end{prop}

\begin{proof}[Proof of Proposition~\ref{cor:ber}] Following Lemma 7 in \cite{gao2018community}, with high probability, the new estimate satisfies
\begin{equation}
    \onormSize{}{\hat \tX' - \tX}_F^2 \lesssim p^{\ceil{K/2}}.
\end{equation} 
We obtain the upper bound by using $\hat \tX'$ in place of $\hat \tX$  in Theorem~\ref{thm:initial}.
\end{proof}

\begin{rmk}[Comparison with previous methods] The Bernoulli bound $\tO(p^{- \floor{K/2} })$ in Proposition~\ref{cor:ber} is relatively looser than the Gaussian bound \tO(p^{- K/2 })$ in Theorem~\ref{thm:initial}. The gap between Bernoulli and Gaussian error decreases as the order $K$ increases. Nevertheless, combining with angle iteration Sub-algorithm~\hyperref[alg:main]{2}, Bernoulli clustering still achieves exponential error rate $\exp \of{ - p^{(K-1)}}$ at a price of a larger SNR.
%Nevertheless, the Bernoulli clustering still achieves exponential error rate $\exp \of{ - p^ \floor{K/2}}$ with large SNR combining with our angle iteration Sub-algorithm~\hyperref[alg:main]{2}. 
The exponential rate is tighter than the rate error $\tO(p^{-2})$ in previous work~\citep{ke2019community}. The investigation of the gap between upper bound $p^{- \floor{K/2}}$ and the lower bound $p^{- K/2}$ for Bernoulli tensors will be left as future work. In numerical experiments, we will use our original initialization, Sub-algorithm 1, to verify the robustness to Bernoulli observations."
\end{rmk}

\end{quote}

New Appendix I: 
\begin{quote}
   `` \textbf{Bernoulli phase transition.} The first additional experiment verifies the statistical-computational gap in Section~\ref{sec:limits} under the Bernoulli model. Consider the Bernoulli model with $p = \{80, 100\}$, $r = 5$. We vary $\gamma $ in $ [-1.2, -0.4]$ and $[-2.1, -1.4]$ for matrix ($K=2$) and tensor $(K = 3)$ clustering, respectively. We  approximate MLE using an oracle estimator, i.e., the output of Sub-algorithm~\hyperref[alg:main]{2} initialized from the true assignment. Figure~\ref{fig:phase_binary} shows a similar pattern as Figure~\ref{fig:phase}. The algorithm and oracle estimators have no gap in the matrix case, while an error gap emerges between the critical values $\gamma_{\text{stat}} = -2$ and $\gamma_{\text{comp}} = -1.5$ in the tensor case. Figure~\ref{fig:phase} suggests the statistical-computational gap in Bernoulli models."

\begin{figure}[htb]
    \centering
    \includegraphics[width = .8\columnwidth]{binary_phase.pdf}
    \caption{SNR phase transitions for Bernoulli dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.}
    \label{fig:phase_binary}
\end{figure}
\end{quote}

\item Regarding ``Formulation inconsistency". We agree that the original exposition is confusing and also unnecessary. We were attempting to simplify the notation by focusing on symmetric signal tensors (not necessarily symmetric data). Based on your comment, the revised paragraph in Section~\ref{sec:model} now reads

\begin{quote}
    ``... For ease of notation, we focus on the case with symmetric mean tensor $\bbE \tY$. This assumption simplifies the notation because all modes have the same $(\mTheta, \mM, z)$; the noise tensor $\tE$ and the data tensor $\tY$ are still possibly asymmetric. In general, we allow asymmetric mean tensor with $\{(\mTheta_k, \mM_k, z_k)\}_{k=1}^K$, one for each mode. The extension can be found in Appendix II...."
\end{quote}

In Section~\ref{sec:prelim}:
\begin{quote}
``...we provide...for general Gaussian dTBMs~\eqref{eq:model_margin} without symmetric assumptions. For general (asymmetric) Gaussian dTBMs, ....we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode. For notational simplicity, we still use $z$ and $\tP(\gamma)$ for this general (asymmetric) model. All results should be interpreted as the worst-case results across $K$ modes....''
\end{quote}

Consequently, we have revised our algorithm to reflect the asymmetry of data. We quote the revised algorithm in the last page of the response letter. We have also revised the notation in Theorems~\ref{thm:initial}-\ref{thm:refinement} by adding subscripts; that is, we now use $z_k^{(0)}$, $z_k^{(t+1)},\ldots$ in place of $z^{(0)}$, $z^{(t+1)}, \ldots$, throughout the paper.

%\begin{quote}
   % \begin{thm}[Error for weighted higher-order initialization] Consider the general sub-Gaussian dTBM with i.i.d.\ noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}. Assume $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$ and let $\Delta_{\mX}$ denote the minimal gap in mean tensor defined in~\eqref{eq:delta_x}.  Let $ z^{(0)}_k$ denote the output of Sub-algorithm~\hyperref[alg:main]{1}. With probability going to 1, we have
%\begin{equation}
%    \ell(z^{(0)}_k, z) \lesssim {\sigma^2 r^K p^{-K/2}\over \Delta_{\mX}^2}.
%\end{equation}
%Further, assume $\mtheta$ is balanced as~\eqref{eq:degree}. We have
%\begin{equation}\label{eq:ini}
% \ell(z_k^{(0)}, z) \lesssim {r^K p^{-K/2}\over \text{SNR}} \quad \text{and} \quad L(z^{(0)}_k, z) \lesssim  {\sigma^2 r^K p^{-K/2}}
%\end{equation}
%\end{thm}

%\begin{thm}[Error for angle-based iteration] Consider the general sub-Gaussian dTBM with independent noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}. Let $\{z^{(0)}_k\}_{k=1}^K$ be the initialization for Sub-Algorithm~\hyperref[alg:main]{2} and $z^{(t)}_k$ be the $t$-th iteration output on $k$-th mode. Assume $\min_{i \in [p]}\theta(i) \geq c $ for some $c > 0$, the $\text{SNR} \geq \tilde C p^{-(K-1)}\log p$ for some sufficiently large constant $\tilde C$, and the initialization satisfies 
%\begin{equation}
%    L(z^{(0)}_k, z) \lesssim \frac{\Delta_{\min}^2}{r \log p}, \quad k \in [K].
%\end{equation}
%% With probability going to 1, there exists a contraction parameter $\rho \in (0,1)$ such that 
%\begin{align}\label{eq:final}
%    \ell(z, \hat z^{(t+1)}_k) \lesssim &\ \KeepStyleUnderBrace{
%   \text{SNR}^{-1}
%    \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}}_{\substack{\text{statistical error}} }+ \KeepStyleUnderBrace{ \rho^t \ell(z, z^{(0)}_k). }_{\substack{\text{computational error}}}
%\end{align}
%\end{thm}

%\end{quote}

%Last, we have clarified the distribution of $\tE$ in equation~\eqref{eq:model_margin}. Throughout the paper, we consider the model with symmetric noise for simpler presentation; our theoretical results are based on the noise tensors with independent noise. We summarize the clarification here:
% \begin{quote}
%     ``..., and $\tE\in\mathbb{R}^{p\times \cdots \times p}$ is a symmetric noise tensor consisting of independent zero-mean sub-Gaussian entries $(i_1,\ldots, i_K)$ for indices $i_1 \leq \cdots \leq i_K$ with variance bounded by $\sigma^2$. The entries in $\tE$ satisfy  $\tE(i_1,\ldots,i_K) = \tE(\pi(i_1),\ldots,\pi(i_K))$ for any permutation $\pi$ on $\{i_1, \ldots, i_K\}$. The symmetric constraint in $\tE$ is relaxed for general dTBM. ..."
% \end{quote}
\end{enumerate}

\item \textit{I also found some technical assumptions strong.  For example, in Eq. (3), the authors restricted the matricization of core tensor has norm bounded from below and above. I don't see why that is necessary. The author argues ``...so the core tensor $\tS$ is not trivially reduced to a lower rank", but the tensor block model should be identifiable even when $\tS$ has degenerated rank. This leads to another question on the defined SNR in Eq. (4) (which requires minimal $\tS$ sliced norm). It is doubtful that this is a true quantity to work on.}

\textbf{Response:} This comment concerns the lower and upper bounds for the core tensor. We address them separately.

\begin{enumerate}
\item Lower bound. Regarding ``...but tensor block model should be identifiable even when $\tS$ has degenerated rank.'' This is in fact what we intended to convey. Indeed, our remark of Theorem~\ref{thm:unique} reads: ``...the identifiability guarantee for the dTBM is stronger than classical Tucker model. .'' To avoid the confusion, we have revised the sentence into

\begin{quote}
``...Third, the lower bound constant $c_3>0$ requires no purely zero slide in $\tS$, and thus the model~\eqref{eq:model_tensor} is not trivially non-identifiable; see Example~\ref{example:c3}. 
...
\end{quote}

The lower boundness of core tensor slices is {\bf necessary} for the model identifiability. We have added following example in Section~\ref{subsec:identify} to illustrate the necessity of the lower bound $c_3$.

\begin{quote}
    `` ...

    \begin{example}[Non-identifiability with purely zero core slice] Consider an order-2 dTBM with core tensor $\mS = \begin{pmatrix} 0 & 0\\
    1 & -1
    \end{pmatrix}$ and mean tensor 
    \begin{equation}
        \tX =  \mTheta_1 \mM \mS   \mM^T \mTheta_2, \quad \text{with } \mM = \begin{bmatrix} 1 & 0 \\
        1 & 0 \\
        0 & 1 \\
        0 & 1 \\
        \end{bmatrix}, \quad \mTheta_1 =\mTheta_2= \text{diag}(1,1,1,1).
    \end{equation}
Replacing $\mTheta_1$ by $\mTheta'_1 = (3/2, 1/2, 1,1)$ leads to the same mean tensor $\tX$. "
\end{example}
\end{quote}
We note that the purely zero slice in $\tS$ always makes the degree heterogeneity indistinguishable. Therefore, we exclude such case for model identifiability.

\item Upper bound. The upper bound $c_4>0$ is a technical constraint to avoid unbounded entries in the core tensor; in practices, the constraint $\max_{a\in[r]} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large choice of $c_4 \geq \onorm{\tY}$. We have changed the wording in Section~\ref{subsec:identify} into

\begin{quote}
    ``... The upper bound $c_4>0$ is a technical constraint to avoid unbounded entries in the core tensor. We introduce this assumption for proof convenience; in practical algorithms, the constraint $\max_{a\in[r]} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large choice of $c_4 \geq \onormSize{}{\tY}_F$. "
\end{quote}

\item Comparison of assumptions with previous work. Our constraints in $\tP$ are mild compared with previous literature. We have also added following comparison in Section~\ref{subsec:identify}:
\begin{quote}
    \begin{table}[h]
    \centering
    \resizebox{\textwidth}{!}{
    \begin{tabular}{c|c ccc}
    \hline
    Assumptions in parameter space& \cite{gao2018community}& \cite{han2020exact}& \cite{ke2019community} &Ours\\
    \hline
         Balanced community sizes &$\surd$ &$\surd$&$\surd$ & $\surd$  \\
         Bounded core tensors& $\surd$ & $\times$ & $\surd$ &$\surd$\\
              Balanced degrees & $\surd$&-&$\surd$ &$\surd$\\
         Flexible in-group connections &$\times$ &$\surd$&$\surd$& $\surd$\\
         Gaps among cluster centers & In-between cluster difference & Euclidean gap & Eigen gap & Angle gap\\
         \hline
    \end{tabular}
    }
    \caption{Parameter space comparison between previous work with our assumption.}
\end{table}
\end{quote}

% We agree that our dTBM is identifiable with some degenerated rank, and we thank the reviewer for pointing out the incorrect expression. The upper bound is technical constraint for theoretical analysis; in practice, the constraint $\max_{a:} \onorm{\mat(\mS)_{a:}} \leq c_4$ would likely never be active with a large $c_4 \geq \onorm{\tY}$.  Nevertheless, the lower bound of core tensor $\tS$ is still necessary for degree-corrected models. The lower bound is critical for identifiability; a purely zero slice in $\tS$ makes the heterogeneity indistinguishable. We have added a short counter-example to illustrate the necessity of non-zero core tensor slice.

% {\color{blue}

% \begin{example}[Non-identifiability with pure zero core slice] Consider an order-2 dTBM with core tensor $\mS = \begin{pmatrix} 0 & 0\\
%     1 & -1
%     \end{pmatrix}$ and mean tensor 
%     \begin{equation}
%         \tX =  \mTheta_1 \mM \mS   \mM^T \mTheta_2, \quad \text{with } \mM = \begin{bmatrix} 1 & 0 \\
%         1 & 0 \\
%         0 & 1 \\
%         0 & 1 \\
%         \end{bmatrix}, \quad \mTheta_1 =\mTheta_2= \text{diag}(1,1,1,1).
%     \end{equation}
% Replacing $\mTheta_1$ by $\mTheta_1 = (3/2, 1/2, 1,1)$ leads to the same mean tensor $\tX$.
% \end{example}
% }

% In addition, the constraints in our parameter space are mild compared with previous degree-corrected block models \citep{gao2018community, ke2019community}. We have added a parameter space comparison table.

% {
% \color{blue}

% \begin{table}[h]
%     \centering
%     \resizebox{\textwidth}{!}{
%     \begin{tabular}{c|c ccc}
%     \hline
%     Assumptions in parameter space& \cite{gao2018community}& \cite{han2020exact}& \cite{ke2019community} &Ours\\
%     \hline
%          Balanced community size &$\surd$ &$\surd$&$\surd$ & $\surd$  \\
%          Balanced degree parameters & $\surd$&-&$\surd$ &$\surd$\\
%          Bounded connection slices & $\surd$ & $\times$ & $\surd$ &$\surd$\\
%          Flexible in-group connections &$\times$ &$\surd$&$\times$& $\surd$\\
%          Gap among cluster centers & In-between cluster difference & Euclidean gap & Eigen gap & Angle gap\\
%          \hline
%     \end{tabular}
%     }
%     \caption{Parameter space comparison between previous work with our assumption.}
% \end{table}
% }

\end{enumerate}
\item \textit{Another big concern I have is the whole Section 3. The statistical and computational lower bounds are exactly the same as the tensor block model in \cite{han2020exact}. Therefore, the whole results in Section 3 can be simply obtained as corollary from \cite{han2020exact}. since the authors are working on a more restrictive model and the lower bounds must be larger than simpler model.}

\textbf{Response:}

In fact, none of our results in Section 3 can be derived from earlier non-degree work~\citep{han2020exact}. The short answer is that, our results show the similar conclusion but {\bf under different conditions.} While the TBM impossibility~\citep{han2020exact} provides a necessary condition for our dTBM impossibility, we find that {\bf such a condition is loose} in our contexts. This has motivated to develop new techniques to derive sharp lower bounds. 
Inspired by your comment, we have revised the exposition of our theory. See our response to \textbf{\# 1 Major Comments from Editor} for details.

We quote the newly-added texts here:
\begin{quote}
    ``...\emph{Theoretical results.} In both works, we study the phase transition of TBM and dTBM with respect to the Euclidean and angle-based SNRs. We briefly summarize the results in \cite{han2020exact} and compare with ours.

    \textit{Statistical critical value:}
    \begin{align}
        \text{Ours:}& \ \Delta_{\text{ang}}^2 \lesssim p^{-(K-1)} \Rightarrow \text{statistically impossible;} \quad \Delta_{\text{ang}}^2 \gtrsim   p^{-(K-1)} \Rightarrow \text{MLE achieves exact recovery;} \\
        \text{Han's:}& \ \Delta_{\text{Euc}}^2 \lesssim p^{-(K-1)} \Rightarrow \text{statistically impossible;} \quad \Delta_{\text{Euc}}^2 \gtrsim   p^{-(K-1)} \Rightarrow \text{MLE achieves exact recovery}.
    \end{align}

     \textit{Computational critical value:}
    \begin{align}
        \text{Ours:}& \ \Delta_{\text{ang}}^2 \lesssim p^{-K/2} \Rightarrow \text{computationally impossible;} \quad \Delta_{\text{ang}}^2 \gtrsim   p^{-K/2} \Rightarrow \text{ computationally efficient;} \\
        \text{Han's:}& \ \Delta_{\text{Euc}}^2 \lesssim p^{-K/2} \Rightarrow \text{computationally impossible;} \quad \Delta_{\text{Euc}}^2 \gtrsim   p^{-K/2} \Rightarrow \text{computationally efficient}.
    \end{align}

The above comparison reveals three major differences.

 %Our dTBM phase transition re-discovers the statistical-computational gaps in TBM. %This finding indicates the intrinsic behavior distinctions among higher-order, matrix, and vector problem may be generally exists in many higher-order problems.
First, none of our results in Section~\ref{sec:limits} are corollaries of \cite{han2020exact}. Both models show the similar conclusion {\bf but under different conditions}. While the TBM impossibility~\citep{han2020exact} provides a necessary condition for our dTBM impossibility, we find that such a condition is often loose. There exists a regime of $\tS$ in which TBM problems are computationally efficient but dTBM problems are statistically impossible; see Example~\ref{example:euc_alg}. This observation has motivated us to develop the new signal notion $\Delta^2_{\text{ang}}$ for sharp dTBM phase transition conditions.

 Second, to find the phase transition, we need to show both the impossibility {\bf and} achievability when SNR is below {\bf and} above the critical value, respectively. While the TBM impossibility can serve as a loose condition of our dTBM impossibility, more efforts are required to show the achievability. In particular, since TBM is a more restrictive model than dTBM, the achievability in \cite{han2020exact} does not imply the achievability of dTBM in a larger parameter space. The latter requires us to develop new MLE and polynomial algorithms for dTBM achievability.  %For impossibility, the minimax bounds in dTBM search over a larger range of estimators than TBM. Though the thresholds for TBM impossibility can serve as the loose lower bounds of the dTBM impossibility, more efforts are required to avoid the gap between impossibility and achievablity.

Third, from the perspective of proofs, we develop new tools to handle the extra degree heterogeneity. In our Theorem~\ref{thm:stats}, we consider the profile MLE by maximizing out the nuisance degree parameter, while TBM~\citep{han2020exact} considers the usual MLE without degree parameter. In our Theorem~\ref{thm:comp}, we construct a rank-2 tensor to relate HPC conjecture to $\Delta^2_{\text{ang}}$, while TBM~\citep{han2020exact} constructs a rank-1 tensor to relate HPC conjecture to $\Delta^2_{\text{Euc}}$. The asymptotic non-equivalence between $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$ renders our proof technically more involved...''
   \end{quote}

   \begin{quote}
   ``... \emph{Signal notion.} We emphasize that the signal notions are different between the two models. In particular, the Euclidean-based signal notion in TBM~\cite{han2020exact} fails to accurately describe the phase transition in our dTBM due to the possible heterogeneity in degree $\mtheta$. To compare, we denote our angle-based signal notion in~\eqref{eq:gamma} and the Euclidean-based SNR in \cite{han2020exact} as $\Delta_{\text{ang}}^2$ and $\Delta_{\text{Euc}}^2$, respectively:
\begin{equation}
     \Delta_{\text{ang}}^2 =  2(1 - \max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}} ), \quad \Delta_{\text{Euc}}^2 = \min_{a \neq b \in [r]} \onormSize{}{\mS_{a:} - \mS_{b:}}^2.
\end{equation}
By Lemma~\ref{lem:norm_diff} in the Appendix II, we have 
\begin{equation}\label{eq:signalcompare}
     \Delta_{\text{ang}}^2  \max_{a \in [r]}\onormSize{}{\mS_{a:}}^2 \leq \Delta_{\text{Euc}}^2.
\end{equation}
The above inequality indicates the condition $\Delta_{\text{Euc}}^2 \leq p^{\gamma}$ is {\bf sufficient but not necessary} for $\Delta_{\text{ang}}^2 \leq p^{\gamma}$. In fact, if we were to use $\Delta_{\text{Euc}}^2$ for both models, then the phase transition of dTBM can be arbitrarily worse than that for TBM.

Here, we provide an example to illustrate the dramatical difference between TBM and dTBM with the same core tensor.

\begin{example}[Comparison with Euclidean-based signal notion] \label{example:euc_alg} Consider a biclustering model with $\mtheta=1$ and an order-2 core matrix 
\begin{equation}
    \mS = \begin{pmatrix} p^{(\gamma+1)/2 } + 2  & 2 p^{(\gamma+1)/2} + 4\\
    2 & 4
    \end{pmatrix},\quad \text{with}\ \gamma \leq -1.
\end{equation}
The core matrix $\mS$ lies in the parameter spaces of TBM and our dTBM. Here, the constraint $\gamma \leq -1$ is added to ensure the bounded condition of $\mS$ in our parameter space in \eqref{eq:family}. The angle-based and Euclidean-based signal levels of $\mS$ are 
\begin{equation}
    \Delta_{\text{ang }}^2(\mS) = 0 \ \left(\leq p^{\gamma}\right), \quad \Delta_{\text{Euc}}^2(\mS) = 5 p^{\gamma + 1} \ \left(\geq p^{\gamma}\right).
\end{equation}
We conclude that TBM with $\mS$ achieves exact recovery with a polynomial-time algorithm; see \citet[Theorem 4]{han2020exact}. By contrast, the dTBM with the same $\mS$ and input $r=2$ violets the identifiability condition, and thus fails to be solved by all estimators; see our Theorem~\ref{thm:unique}.'' 
\end{example}

   \end{quote}
   We have revised our Section~\ref{sec:limits} to better present the lower bounds. See our response to \textbf{\# 1 Major Comments from Editor}.

\newpage

\begin{algorithm}[h!]
\caption*{\bf Algorithm: Multiway spherical clustering for degree-corrected tensor block model }
\vspace{.15cm}
\begin{algorithmic}[1]
\Algphase{Sub-algorithm 1: Weighted higher-order initialization}
\INPUT Observation $\tY \in \bbR^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.

\State {
\color{blue}Compute factor matrix $ \mU_{\text{pre},k} = \text{SVD}_{r} (\text{Mat}_k(\tY)), k \in [K]$ and the $(K-1)$-mode projection 
\begin{equation}
    \tX_{\text{pre},k} = \tY \times_1   \mU_{\text{pre},1} \mU_{\text{pre},1}^T \times_2 \cdots  \times_{k-1} \mU_{\text{pre},k-1} \mU_{\text{pre},k-1}^T \times_{k+1} \mU_{\text{pre},k-1} \mU_{\text{pre},k+1}^T\times_{K+1}  \times_K \mU_{\text{pre},K} \mU_{\text{pre}, K}^T.
\end{equation}
}
\State {
\color{blue}
Compute factor matrix $\hat \mU_k = \text{SVD}_{r}(\text{Mat}_k(\tX_{\text{pre},k})), k \in [K]$ and denoised tensor
\begin{equation}
    \hat \tX = \tY \times_1 \hat \mU_1 \hat \mU^T_1 \times_2 \cdots \times_K \hat \mU_K \hat \mU^T_K.
\end{equation}
} 
\For{\textcolor{blue}{$k \in [K]$}}
\State { \color{blue} Let $\hat \mX = \text{Mat}_k(\hat \tX)$ and $S_0=\{i \in [p]: \onormSize{}{\hat \mX_{i:}} = 0\}$. Set $\hat z(i)$ randomly in $[r]$ for $i \in S_0$.}
\State{  For all $i\in S_0^c$, compute normalized rows
$\hat \mX_{i:}^s :=\onormSize{}{\hat \mX_{i:}}^{-1} \hat \mX_{i:}.$
}
\State { Solve the clustering $\hat z_k \colon [p]\to[r]$ and centroids $ (\hat \mx_j)_{j\in[r_k]}$ using weighted $k$-means, such that}
\begin{align}
    &\sum_{i \in S_0^c }  \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{\hat z_k(i)} }^2 
    \leq 
    \eta \min_{\substack{\bar \mx_j, j\in[r], \bar z_k(i),i\in S_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\hat z_k(i)}}^2.
\end{align}
\EndFor

\OUTPUT {\color{blue} Initial clustering $z^{(0)}_k \leftarrow \hat z_k, k \in [K]$.}

\Algphase{Sub-algorithm 2: Angle-based iteration}
\INPUT Observation $\tY \in \bbR^{p \times \cdots \times p}$, initialization $z^{(0)}_k \colon [p]\to[r], k \in [K]$ from Sub-algorithm 1, iteration number $T$.
\For {$t = 0$ to $T-1$}
\State Update the block tensor $\tS^{(t)}$ via
$\tS^{(t)} (a_1,...,a_K)= \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}_k(i_k) = a_k, k \in [K]\}.$
\For{ $k \in [K]$}
\State {\color{blue} Calculate reduced tensor $\tY^{\text{d}}_k \in \bbR^{r \times r \times p \times r \times \cdots \times r}$ via
\begin{equation}
    \tY^{\text{d}}_k(a_1,\ldots,a_{k-1}, i ,a_{k+1},\ldots,a_K) 
    = \text{Ave}\{\tY(i_1,\ldots,i_{k-1},i,i_{k+1},\ldots,i_K): z^{(t)}(i_j) = a_j, j \neq k \}
\end{equation}}
% \begin{align}
%     &\tY^{\text{d}}(i,a_2,\ldots,a_K) 
%     = 
% \text{Ave}\{\tY(i,i_2,\ldots,i_K): z^{(t)}(i_k) = a_k, k \neq 1 \}.
% \end{align}

\State { \color{blue} Let $\mY_k^{\text{d}} = \text{Mat}_k(\tY^{\text{d}})$ and $J_0 = \{ i\in[p]: \onorm{\mY^{\text{d}}_{i:}} = 0\}$. Set $z_k^{(t+1)}(i)$ randomly in $[r]$ for $i \in J_0$.}

\State {\color{blue} Let $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. For all $i \in J_0^c$ update the cluster assignment by
\begin{equation}
    z(i)^{(t+1)}_k = \argmax_{a \in [r]} \cos \left( \mY^{\text{d}}_{k,i:},\ \mS^{(t)}_{a:} \right).
\end{equation}}
\EndFor
\EndFor

\OUTPUT { \color{blue} Estimated clustering $z^{(T)}_k  \in [r]^{p}, k \in [K]$.}

\end{algorithmic}
\end{algorithm}

\end{enumerate}

% \begin{lem}[perturbation for degree-corrected clustering]. Let $z^*$ be the true clustering. Fix an arbitrary clustering $z$ with $\text{MisClust}(z,z^*):=\varepsilon \leq ...$ (a small constant). 
% Under assumptions....., we have
% \begin{equation}\label{eq:mode1}
% \min_{(\tS, \Theta)}\Fnorm{\tX(z^*, \tS^*, \Theta^*)-\tX(z; \tS, \Theta)}^2\gtrsim (...\text{factors involving $p^K$... }) \varepsilon \text{Angle-gap}^2 + o(\varepsilon).
% \end{equation}
% \end{lem}

% \begin{rmk}
% Everything is deterministic. Here, $\tX(z^*,...)$ is the degree-corrected signal tensor. Two ways to prove: 1) Find (deterministic)
% \[
% (\tS_{\min}, \Theta_{\min})=\argmin_{(\tS, \Theta)}\Fnorm{\tX(z^*)-\tX(z, \tS, \Theta)}^2,
% \]
% and then plug into~\eqref{eq:mode1} to evaluate the perturbation; 2) Prove a stronger version of identifiability result
% \[
% \KeepStyleUnderBrace{\min_{(\tS, \Theta, \tS^*, \Theta^*)}\Fnorm{\tX(z^*, \tS^*, \Theta^*)-\tX(z; \tS, \Theta)}^2}_{\text{distance between two points in the parameter space}}\gtrsim (...\text{factors involving $p^K$... })  \text{Angle-gap}^2 \KeepStyleUnderBrace{\varepsilon}_{\text{distance between $z$ and $z^*$}}.
% \]
% Our identifiability theorem is a special case of the above inequality by right hand side = 0.

% Lemma 1 is simpler than approaches in my NIPS paper, because no $\tY$ or expectation is needed. 
% In addition, I omitted the subscript $k$ in~\eqref{eq:mode1} for intuition. Precisely, ~\eqref{eq:mode1} holds only for $z_1$ while fixing $\{z_k\}_{k\geq 2}$. 
% \end{rmk}
% \begin{lem}[Union bound]
% \begin{align}
% &\mathbb{P}(\text{MisClust}(\hat z_{\text{MLE}}, z^*)\geq \varepsilon)\\
% \stackrel{\text{\color{red}You should write this step as instinct}}{=}&\mathbb{P}(\exists (z, \tS, \Theta) \ \text{s.t.}\  \text{MisClust}(z,z^*)\geq \varepsilon \ \text{and}\ \text{Log-lik}(z, \tS, \Theta) \geq \text{Log-lik}(z^*))\\
% %&\leq \sum_{z \text{ for which MisClust}(z,z^*)\geq \varepsilon}\mathbb{P}(\text{Log-lik}(z)\geq \text{Log-lik}(z^*))\\
% \leq &\sum_{z \text{ for which MisClust}(z,z^*)\geq \varepsilon}\sum_{(\tS, \Theta)}\mathbb{P}(\Fnorm{\tY-\tX(z, \tS, \Theta)}^2\leq \Fnorm{\tY-\tX(z^*)}^2)\\
% \leq & \sum_{z:\ \text{MisClust}(z,z^*)\geq \varepsilon}\sum_{(\tS, \Theta)} \mathbb{P}\left({\langle \tE, \tX(z^*)-\tX(z, \tS, \Theta)\rangle  \over\Fnorm{\tX(z^*)-\tX(z, \tS, \Theta)}}\gtrsim \Fnorm{\tX(z^*)-\tX(z, \tS, \Theta)} \right)\\
% \leq& \text{(desired conclusion)}
% \end{align}

% In the last step, only randomness is on $\tE$. The term $\Fnorm{\tX(z^*)-\tX(z,...)}$ can be replaced by $...\sqrt{\varepsilon}$ because of condition in the sum ``$z$ for which MisClust$(z,z^*)\geq \varepsilon$ '' and Lemma 1.

% \fixme{Miaoyan}{Please complete full steps using ideas from Lemma 1 and Thm 1 in my NIPS paper. Lemma 2 is simple. I sketched out the proof for Lemma 1 for $R=2$. For general $R$ (number of blocks), it should be doable (I did not do) with just more notations}

% \end{lem}

\bibliography{tensor_wang}
\bibliographystyle{apalike}

\end{document}

### 2023-01-24 tensor_clustering/IEEE/ieee_revision/ieee_revise_s1.5.tex
organize

--- 2022-01-18 tensor_clustering/tensor_clustering_ieee/ieee_v3.tex
@@
 
+% \usepackage{xr-hyper}
+% % this is for overleaf cross reference
+% \makeatletter
+% \newcommand*{\addFileDependency}[1]{% argument=file name and extension
+%   \typeout{(#1)}
+%   \@addtofilelist{#1}
+%   \IfFileExists{#1}{}{\typeout{No file #1.}}
+% }
+% \makeatother
+
+% \newcommand*{\myexternaldocument}[1]{%
+%     \externaldocument{#1}%
+%     \addFileDependency{#1.tex}%
+%     \addFileDependency{#1.aux}%
+% }
+% \myexternaldocument{response}
+
 \begin{document}
@@
 
-\author{ Jiaxin Hu and Miaoyan Wang\\
-University of Wisconsin - Madison
+\author{ Jiaxin Hu and Miaoyan Wang$^*$\thanks{$^*$Part of this paper was presented at 25th International Conference on Artificial Intelligence and Statistics (AISTATS).}\\
+Department of Statistics, University of Wisconsin - Madison
 %IEEE Publication Technology,~\IEEEmembership{Staff,~IEEE,}
@@
 
-Classical tensor block model suffers from drawbacks to model real world data in spite of the popularity.
-The key underlying assumption of block model is that all nodes in the same community are exchangeable; i.e., the nodes have no individual effects apart from the block effects.
+The classical tensor block model suffers from drawbacks to model real world data in spite of the popularity.
+The key underlying assumption of block model is that all nodes in the same community are exchangeable; i.e., the nodes have no individual-specific parameters apart from the community-specific parameters.
 However, the exchangeability assumption is often non-realistic.
 Each node may contribute to the data variation by its own multiplicative effect.
+We call the unequal node-specific effects the \emph{degree heterogeneity}.
 Such degree heterogeneity appears commonly in social networks.
 Ignoring the degree heterogeneity may seriously mislead the clustering results.
-For example, regular block model fails to model the member affiliation in the Karate Club network~\citep{bickel2009nonparametric} without addressing degree heterogeneity.
+For example, the regular block model fails to model the member affiliation in the Karate Club network~\citep{bickel2009nonparametric} without addressing degree heterogeneity.
 
@@
  \item We provide an angle-based algorithm that achieves exact clustering \emph{in polynomial time} under mild conditions.
-Simulation and data studies demonstrate the outperformance of our algorithm compared with existing higher-order clustering algorithms.
+Simulation and data studies demonstrate that our algorithm outperforms existing higher-order clustering algorithms.
 \end{itemize}
@@
 
-\begin{itemize}
-    \item \textit{Block model for clustering.} Block models such as stochastic block model (SBM) and degree-corrected SBM have been widely used for matrix clustering problems.
+\begin{itemize}[wide]
+    \item \textit{Block model for clustering.} The block model such as stochastic block model (SBM) and degree-corrected SBM has been widely used for matrix clustering problems.
 The theoretical properties and algorithm performance for matrix block models have been well-studied~\citep{gao2018community}; see the review paper~\citep{abbe2017community} and the references therein.
@@
 We use $\text{Ave}(\cdot)$ to denote the operation of taking averages across elements and $\text{Mat}_k(\cdot)$ to denote the unfolding operation that reshapes the tensor along mode $k$ into a matrix.
-For a symmetric tensor $\tY\in\mathbb{R}^{p\times \cdots \times p}$, we omit the subscript and use $\text{Mat}(\tY)\in\mathbb{R}^{p\times p^{K-1}}$ to denote the unfolding.
-For two sequences $\{a_p\}, \{b_p\}$, we denote $a_p\lesssim b_p$ or $a_p=\tO(b_p)$ if $\lim_{p\to\infty}a_p /b_p\leq c$ for some constant $c\geq 0$, $a_p=o(b_p)$ if $\lim_{p\to\infty}a_p/b_p =0$, and $a_p = \Omega(b_p)$ if both $b_p \lesssim a_p$ and $a_p\lesssim b_p$.
+\textcolor{blue}{For a symmetric tensor $\tX \in\mathbb{R}^{p\times \cdots \times p}$, we omit the subscript and use $\text{Mat}(\tX)\in\mathbb{R}^{p\times p^{K-1}}$ to denote the unfolding.}  For two sequences $\{a_p\}, \{b_p\}$, we denote $a_p\lesssim b_p$ or $a_p=\tO(b_p)$ if $\lim_{p\to\infty}a_p /b_p\leq c$ for some constant $c\geq 0$, $a_p=o(b_p)$ if $\lim_{p\to\infty}a_p/b_p =0$, and $a_p = \Omega(b_p)$ if both $b_p \lesssim a_p$ and $a_p\lesssim b_p$.
 Throughout the paper, we use the terms ``community'' and ``clusters'' exchangeably.
@@
 In Section~\ref{sec:alg}, we provide a polynomial-time two-stage algorithm with misclustering rate guarantees.
-Numerical studies including the simulation, comparison with other methods, and two real dataset analyses are in Sections~\ref{sec:simulation}-\ref{sec:real}.
+{\color{blue}Extension to Bernoulli models is also presented.
+In Section~\ref{sec:tbm}, we compare our work with non-degree tensor block models.} Numerical studies including the simulation, comparison with other methods, and two real dataset analyses are in Sections~\ref{sec:simulation}-\ref{sec:real}.
 The main technical ideas we develop for addressing main theorems are provided in Section~\ref{sec:mainproof}.
@@
 
-Suppose we have an order-$K$ data tensor $\tY \in \bbR^{p \times \cdots \times p}$.
-For ease of notation, we focus on symmetric tensors in this section; the extension to general asymmetric tensors is provided in Section~\ref{subsec:exten}.
+Suppose we have an order-$K$ data tensor $\tY \in \bbR^{p \times \cdots \times p}$. %; the extension to general case with asymmetric mean tensor is provided in Section~\ref{subsec:exten}.} 
+%we focus on symmetric tensors in this section; the extension to general asymmetric tensors is provided in Section~\ref{subsec:exten}. 
 Assume there exist $r \geq 2$  disjoint communities among the $p$ nodes.
@@
 \normalsize
-where $\tS \in \bbR^{r \times \cdots \times r}$ is an order-$K$ tensor collecting the block means among communities, and $\tE\in\mathbb{R}^{p\times \cdots \times p}$ is a noise tensor consisting of independent zero-mean sub-Gaussian entries with variance bounded by $\sigma^2$.
-The unknown parameters are $z$, $S$, and $\mtheta$.
-The dTBM can be equivalently written in a compact form of tensor-matrix product:
+where $\tS \in \bbR^{r \times \cdots \times r}$ is an order-$K$ tensor collecting the block means among communities, and %\textcolor{blue}{$\tE\in\mathbb{R}^{p\times \cdots \times p}$ is a symmetric noise tensor consisting of independent zero-mean sub-Gaussian entries $(i_1,\ldots, i_K)$ for indices $i_1 \leq \cdots \leq i_K$ with variance bounded by $\sigma^2$.
+The entries in $\tE$ satisfy  $\tE(i_1,\ldots,i_K) = \tE(\pi(i_1),\ldots,\pi(i_K))$ for any permutation $\pi$ on $\{i_1, \ldots, i_K\}$.
+The symmetric constraint in $\tE$ is relaxed for general dTBM.}
+$\tE\in\mathbb{R}^{p\times \cdots \times p}$ is a noise tensor consisting of independent zero-mean sub-Gaussian entries with variance bounded by $\sigma^2$. 
+The unknown parameters are $z$, $S$, and $\mtheta$. The dTBM can be equivalently written in a compact form of tensor-matrix product:
 \begin{equation}\label{eq:model_tensor}
@@
 Note that the discrete nature of $\mM$ renders our model~\eqref{eq:model_tensor} more challenging than Tucker decomposition.
-We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$ if $\tY$ admits dTBM~\eqref{eq:model_tensor}.
-The goal of clustering is to estimate $z$ from a single noisy tensor $\tY$.
+We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$ if $\tY$ admits dTBM~\eqref{eq:model_tensor} \textcolor{blue}{and let $\tX = \bbE \tY$ denote the mean tensor.} The goal of clustering is to estimate $z$ from a single noisy tensor $\tY$.
 We are particularly interested in the high-dimensional regime where $p$ grows whereas $r=\tO(1)$.
 
-\subsection{Motivating examples} Here, we provide four applications to illustrate the practical necessity of dTBM.
+\textcolor{blue}{For ease of notation, we have focused on the case with symmetric mean tensor $\bbE \tY$.
+This assumption simplifies the notation because all modes have the same $(\mTheta, \mM, z)$; the noise tensor $\tE$ and the data tensor $\tY$ are still possibly asymmetric.
+In general, we allow asymmetric mean tensors with $\{(\mTheta_k, \mM_k, z_k)\}_{k=1}^K$, one for each mode.
+The extension can be found in Appendix II.}
+
+\subsection{Motivating examples}\label{subsec:motiv} Here, we provide four applications to illustrate the practical necessity of dTBM.
 
@@
 
-\paragraph{Community detection in hypergraphs} Hypergraph network is a powerful tool to represent the complex entity relations with higher-order interactions~\citep{ke2019community}.
+\paragraph{Community detection in hypergraphs} The hypergraph network is a powerful tool to represent the complex entity relations with higher-order interactions~\citep{ke2019community}.
 A typical undirected hypergraph is denoted as $H = (V,E)$, where $V = [p]$ is the set of nodes and $E$ is the set of undirected hyperedges.
@@
 Assume there exist $r$ disjoint communities among $p$ nodes and $r_l$ disjoint communities among the $L$ layers.
-The multi-layer network community detection is modeled by the generalized asymmetric dTBM model~\eqref{eq:model_tensor}
+The multi-layer network community detection is modeled by the general asymmetric asymmetric dTBM model~\eqref{eq:model_tensor}
 \begin{equation}
@@
 
-Assume there exist $r_1, r_2, r_3$ disjoint clusters for $p$ genes, $n$ individuals, and $t$ tissues, respectively. We apply the generalized asymmetric dTBM model~\eqref{eq:model_tensor} 
+Assume there exist $r_1, r_2, r_3$ disjoint clusters for $p$ genes, $n$ individuals, and $t$ tissues, respectively. We apply the general asymmetric dTBM model~\eqref{eq:model_tensor} 
 \begin{equation}
@@
 \end{equation}
-where $\{(\mtheta_k, \mM_k)\}_{k=1}^3$ represents the heterogeneity and membership for genes, individuals, and tissues. \\
+where $\{(\mtheta_k, \mM_k)\}_{k=1}^3$ represents the heterogeneity and membership for genes, individuals, and tissues. 
+\vspace{0.2cm}
+
+%{
+%\color{blue}
+%\paragraph{Tensor Gaussian mixture model} Gaussian mixture model is widely applied in applications including pattern recognition, image processing, and machine learning.
+We say an order-$K$ sample tensor $\tY \in \bbR^{p_1 \times \cdots \times p_K}$ follows a tensor Gaussian distribution with mean tensor $\tS \in \bbR^{p_1 \times \cdots \times p_K}$ if $\text{vec}(\tY)$ is a Gaussian vector with mean $\text{vec}(\tS)$.
+Tensor Gaussian mixture model assumes the tensor data $\tY_i$ for all $i \in [N]$ follow the mixture tensor Gaussian distribution with density $f(\tY_i) = \sum_{a = 1}^r \pi_a \phi_a(\tS_a), i \in [N]$, where $r$ is the number of clusters, $\pi_a$ is the mixture weight, $\tS_a$ is the mean tensor, and $\phi_a$ denotes the tensor Gaussian density of the $a$-th cluster with mean $\tS_a$ for all $a \in [r]$.
+
+%Let $\tY \in \bbR^{N \times p_1 \times \cdots \times p_K }$ denote the combined data with slices $\tY(i,:) = \tY_i$ and $\tS \in \bbR^{r \times p_1 \times \cdots \times p_K}$ denote the combined mean tensor with slices $\tS(a,:) = \tS_a$.
+The tensor Gaussian mixture model with prior probability $\pi_a$ is equal to a special dTBM model
+%\begin{equation}
+   % \bbE[\tY] = \tS \times_1 \mM \times_2 \mI_{p_1} \times_3 \cdots \times_{K+1} \mI_{p_K},
+%\end{equation}
+%where $\mI_p$ is the identity matrix of dimension $p$, $\mM \in \{0,1\}^{N\times r}$ is the membership matrix for $N$ samples such that $\pi_a = |\{ i \in [N]: \mM_{i,a} = 1\} |/N$ for all $a \in [r]$.
+Assume that $\tS$ has a community structure.
+The tensor Gaussian mixture further becomes
+%\begin{equation}
+   % \bbE[\tY] = \tC \times_1 \mM \times_2 \mTheta_1 \mM_1 \times_3 \cdots \times_{K+1} \mTheta_K \mM_K,
+%\end{equation}
+%where $\{ \mTheta_k, \mM_k\}_{k=1}^K$ refers to the heterogeneity and membership on the mean of Gaussian tensors. 
+%}
 
@@
 
-\subsection{Identifiability under angle gap condition}
+\subsection{Identifiability under angle gap condition}\label{subsec:identify}
 The goal of clustering is to estimate the partition function $z$ from model~\eqref{eq:model_tensor}.
@@
 The constants $c_1, c_2$ in the $|z^{-1}(a)|$ bound assume the roughly balanced size across $r$ communities.
-Third, the constants $c_3, c_4 $ in the magnitude of $\text{Mat}(\tS)_{a:}$ requires no purely zero slide in $\tS$, so the core tensor $\tS$ is not trivially reduced to a lower rank.
+\textcolor{red}{Third, the constant $c_3$ requires that all slides in $\tS$ have non-degenerate norm.
+Particularly, the lower bound $c_3$ excludes the no purely zero slide case to avoid trivial non-identifiability of model \eqref{eq:model_tensor}; see Example~\ref{example:c3} below.
+The upper bound $c_4$ is a technical constraint to avoid the slides with unbounded norm as dimension grows; in practice, the constraint $\max_{a:} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large $c_4 \geq \onormSize{}{\tY}_F$.
+} %\textcolor{blue}{ Third, the constant $c_3$ requires no purely zero slide in $\tS$, and thus the model~\eqref{eq:model_tensor} is not trivially non-identifiable; see Example~\ref{example:c3} below.
+The upper bound $c_4$ is a technical constraint to avoid the unbounded entries in the core tensor; in practice, the constraint $\max_{a:} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large $c_4 \geq \onormSize{}{\tY}_F$.
+}
+%Third, the constants $c_3, c_4 $ in the magnitude of $\text{Mat}(\tS)_{a:}$ requires no purely zero slide in $\tS$, so the core tensor $\tS$ is not trivially reduced to a lower rank. 
 Lastly, the $\ell_1$ normalization $\onormSize{}{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|$ is imposed to avoid the scalar ambiguity between $\mtheta_{z^{-1}(a)}$ and $\tS$.
 This constraint, again, incurs no restriction to model flexibility but makes our presentation cleaner.
+\textcolor{blue}{Our constraints in $\tP$ are mild compared with previous literature; see Table~\ref{tab:para} for comparison.}
 
@@
 
+ {
+ \color{blue}
+    \begin{example}[Non-identifiability with purely zero core slice]\label{example:c3} 
+ Consider an order-2 dTBM with core tensor $\mS = \begin{pmatrix} 0 & 0\\
+    1 & -1
+    \end{pmatrix}$ and mean tensor 
+    \begin{equation}
+        \tX =  \mTheta_1 \mM \mS   \mM^T \mTheta_2, \quad \text{with } \mM = \begin{bmatrix} 1 & 0 \\
+        1 & 0 \\
+        0 & 1 \\
+        0 & 1 \\
+        \end{bmatrix}, \quad \mTheta_1 =\mTheta_2= \text{diag}(1,1,1,1).
+    \end{equation}
+Replacing $\mTheta_1$ by $\mTheta'_1 = (3/2, 1/2, 1,1)$ leads to the same mean tensor $\tX$. 
+\end{example}
+
+     \begin{table}[h]
+    \centering
+    \resizebox{\textwidth}{!}{
+    \begin{tabular}{c|c ccc}
+    \hline
+    Assumptions in parameter space& \cite{gao2018community}& \cite{han2020exact}& \cite{ke2019community} &Ours\\
+    \hline
+         Balanced community sizes &$\surd$ &$\surd$&$\surd$ & $\surd$  \\
+          Bounded core tensors & $\surd$ & $\times$ & $\surd$ &$\surd$\\
+         Balanced degrees& $\surd$&-&$\surd$ &$\surd$\\
+         Flexible in-group connections &$\times$ &$\surd$&$\surd$& $\surd$\\
+         Gaps among cluster centers & In-between cluster difference & Euclidean gap & Eigen gap & Angle gap\\
+         \hline
+    \end{tabular}
+    }
+    \caption{Parameter space comparison between previous work with our assumption.}
+    \label{tab:para}
+\end{table}
+ }
+
 We now provide the identifiability conditions for our model before estimation procedures.
@@
 
-\begin{thm}[Model identifiability]\label{thm:unique} Consider the dTBM with $r\geq 2$.
+\begin{thm}[Model identifiability]\label{thm:unique} Consider the dTBM with $r\geq 1$ and \textcolor{red}{ $K \geq 2$}.
 The parameterization~\eqref{eq:model_tensor} is unique in $\tP$ up to cluster label permutations, if and only if Assumption~\ref{assmp:min_gap} holds.
@@
 
-The identifiability guarantee for the dTBM is more appealing than classical Tucker model.
+The identifiability guarantee for the dTBM is stronger than classical Tucker model.
 In the Tucker model, the factor matrix $\mM$ is identifiable only up to orthogonal rotations.
@@
 
-\section{Statistical-computational limits for higher-order tensors}\label{sec:limits}
-
-In this section, we study the statistical and computational limits of dTBM. We propose signal-to-noise ratio~(SNR),
+\section{Statistical-computational critical values for higher-order tensors}\label{sec:limits}
+
+%In this section, we study the statistical and computational limits of dTBM.
+
+% {
+% \color{blue}
+% \subsection{Preliminary (Version 1)}
+
+% We consider the angle gap among the slices of the mean tensor
+% \begin{equation}
+%     \bar \Delta_{\min}^2 = \min_{i,j \in [p], z(i) \neq z(j)} \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  },
+% \end{equation}
+% where $\mX = \mat(\tX), \tX = \tS \times_1 \mTheta \mM \times_2 \cdots \times_K \mTheta \mM$ matricization of the mean tensor.
+Unlike Assumption~\ref{assmp:min_gap}, the gap $ \bar \Delta_{\min}^2$ is a function of all parameters $(z,\mtheta, \tS)$.
+We propose the signal-to-noise ratio SNR
+% \begin{equation}\eqref{eq:gamma}
+%      \text{SNR}:= \bar \Delta^2_{\min}/\sigma^2 = p^{\gamma}.
+% \end{equation}
+% with varying $\gamma \in \bbR$ that quantifies different regimes of interest.
+We call $\gamma$ the \emph{signal exponent}.
+Intuitively, a larger SNR, or equivalently a larger $\gamma$, benefits the clustering in the presence of noise.
+With quantification~\eqref{eq:gamma}, we consider the following parameter space,
+% \begin{equation}\label{eq:gammafamily}
+%     \tP(\gamma) = \tP\cap\{\tS \text{ satisfies SNR condition~\eqref{eq:gamma} with $\gamma$}\}.
+% \end{equation}
+% The $1$-block dTBM does not belong to the space $\tP(\gamma)$ when $\gamma < 0$ by Assumption~\ref{assmp:min_gap}.
+Our goal is to characterize the clustering accuracy with respect to $\gamma$ under the space $\tP(\gamma)$.
+
+% \begin{rmk}[Comparison with Euclidean-based SNR] The Euclidean-based SNR is proposed to describe the phase transition in TBM.
+To compare, recall our angle-based SNR in~\eqref{eq:gamma}, denoted $\Delta_{\text{ang}}^2$ and Euclidean-based SNR in \cite{han2020exact}, denoted $\Delta_{\text{Euc}}^2$, with noise level $\sigma^2 = 1$:
+% \begin{equation}
+%      \Delta_{\text{alg}}^2 =  2(1 - \max_{i, j, z(i)\neq z(j)\in [r]}\cos \of{\mX_{i:},\  \mX_{j:}} ), \quad \Delta_{\text{Euc}}^2 = \min_{a \neq b \in [r]} \onormSize{}{\mS_{a:} - \mS_{b:}}^2.
+% \end{equation}
+
+% \begin{example}[Comparison with Euclidean-based SNR] Here, we provide an example to see the dramatic difference between $ \Delta_{\text{alg}}^2$ and  $\Delta_{\text{Euc}}^2$.
+Consider the matrix observations and the order-2 core matrix
+% \begin{equation}
+%     \mS = \begin{pmatrix} p^{(\gamma+1)/2 } + 2  & p^{(\gamma+1)/2} + 2\\
+%     2 & 2
+%     \end{pmatrix},
+% \end{equation}
+% for $\gamma \leq -1$. The angle-based with $\mtheta = 1$ and Euclidean-based SNRs are 
+% \begin{equation}
+%     \Delta_{\text{alg}}^2(\mS) = 0 \lesssim p^{\gamma}, \quad \Delta_{\text{Euc}}^2(\mS) = 2 p^{\gamma + 1} \gtrsim p^{\gamma}.
+% \end{equation}
+% The TBMs with $\mS$ and proper membership $\mM$ is computationally efficient by \cite{han2020exact}.
+However, the dTBMs in $\tP$ with $\mS$ violets the identifiable condition, Assumption~\ref{assmp:min_gap}, and thus fails to be solved by any estimator.
+% \end{example}
+% Above example indicates that the phase transition of dTBM with respect to $\Delta_{\text{Euc}}^2$ is arbitrary worse than that of TBM, and thus $\Delta_{\text{Euc}}^2$ is not suitable to describe the transition of dTBM.
+% \end{rmk}
+
+% In addition, we consider the \emph{balanced} heterogeneity where 
+% \begin{equation}\label{eq:degree}
+% {\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
+% \end{equation}
+% Following lemma shows the close relation between the angle gaps in mean tensor $\tX$ and core tensor $\tS$ under the balanced heterogeneity.
+
+% \begin{lem}[Angle gap in $\tX$ and $\tS$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family}.
+Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying~\eqref{eq:degree}.
+Then, for all $i,j$ such that $z(i) \neq z(j)$
+% {
+% \color{blue}
+% \begin{equation}
+%   \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  } \asymp  \onorm{ \frac{\mS_{z(i):}}{\onormSize{}{\mS_{z(i):}}}  -  \frac{\mS_{z(j):}}{\onormSize{}{\mS_{z(j):}}}  }.
+% \end{equation}
+% }
+% \end{lem}
+
+% \begin{rmk}[Core tensor signal as SNR.]In our work, we pay a special attention under the regime with balanced $\mtheta$ in estimation.
+With balanced heterogeneity, the $\text{SNR} \asymp \Delta_{\min}^2/ \sigma^2$ and the transition of dTBM is completely described by the signal in core tensor.
+The core tensor signal is more interpretable and easily verified by scientific knowledge in practice, which makes the phase transition with respect to $\Delta_{\min}^2$ be of our main interest.
+Without balanced heterogeneity, the gap in core tensor fails to represent our SNR and describes dTBM transition accurately.
+Here, we provide an example to illustrate the failure of $\Delta_{\min}^2$ with unbalanced heterogeneity.
+
+% \begin{example}[Unbalanced heterogeneity] Consider the order-2 dTBM with core matrix
+% \begin{equation}
+%     \mS = \begin{pmatrix} 1 & a\\
+%     1 & -a
+%     \end{pmatrix}, \quad \text{and} \quad \mtheta \text{ such that } \onormSize{}{\mtheta_{z^{-1}(1)}}^2 = M \onormSize{}{\mtheta_{z^{-1}(2)}}^2,
+% \end{equation}
+% where $\tO(p^{-1}) \leq M \leq \tO(p)$ by the parameter space~\eqref{eq:family}.
+% We have 
+% \begin{equation}
+%     \Delta_{\min}^2 = \frac{2 a^2}{1 + a^2}, \quad \bar \Delta_{\min}^2 = \frac{2\onormSize{}{\mtheta_{z^{-1}(2)}}^{2} a^2}{ \onormSize{}{\mtheta_{z^{-1}(1)}}^{2} +  \onormSize{}{\mtheta_{z^{-1}(2)}}^{2}a^2}.
+% \end{equation}
+% Then,
+% \begin{align}
+%     &\text{if } a = p^{-1/2}, M = 1& \quad &\Rightarrow \quad \Delta_{\min}^2 = \bar \Delta_{\min}^2 = p^{-1}; \\
+%     &\text{if } a = p^{-1/4}, M = p^{1/2} & \quad &\Rightarrow \quad \Delta_{\min}^2 = p^{-1/2}, \   \bar \Delta_{\min}^2 = p^{-1}.
+% \end{align}
+% Above example indicates that different core tensor gaps $\Delta_{\min}^2$ may lead to the same mean tensor gap $\bar \Delta_{\min}^2$ under the unbalanced heterogeneity. 
+% \end{example}
+% \end{rmk}
+
+% \begin{rmk}[Balanced heterogeneity] The balance assumption~\eqref{eq:degree} does not preclude the dramatic degree heterogeneity.
+Indeed, within each of the clusters, the highest degree can be $\theta(i) = \Omega(p)$, whereas the lowest degree can be $\theta(i)=\tO(1)$.
+This range constraints on $\mtheta$ is looser than previous work \citep{gao2018community} that restricts the highest degree in a sub-linear rate $o(p)$.
+
+% We propose he balance assumption~\eqref{eq:degree} to avoid the extreme inflation or elimination of the core tensor gap $\Delta_{\min}^2$ after the weighted mapping by degree heterogeneity.
+Similar regulations are common in literature but with different exposure.
+In higher-order model \cite{ke2019community}, similar assumption ${\max_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}\leq C \min_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}$ is made to ensure the degree distributions are balanced on each community; our balance assumption~\eqref{eq:degree} is equivalent to their assumption after relaxing the $\ell_1$ constraint on $\onormSize{}{\mtheta_{z^{-1}(a)}}_1$ as \cite{ke2019community}.
+In matrix case \cite{gao2018community},  assortative connection, i.e., $\mS(a,b) \leq p_2 < p_1 \leq \mS(a,a)$ for all $a \neq b$, constraint $\onormSize{}{\mtheta_{z^{-1}(a)}}_1 = |z^{-1}(a)| + o(1)$, and the sub-linearity $\theta(i) = o(\sqrt{p})$ are assumed; the $\ell_1$ constraint and sub-linearity indicate the balanced assumption ${\max_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}\leq C \min_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}$; the constant 1 in our assumption~\eqref{eq:degree} can be relaxed to $C$ if we consider the assortative dTBM with diagonal dominant core tensor.
+
+% \end{rmk}
+
+% }
+
+% Last, let $\hat z$ and $z$ be the estimated and true clustering functions in the family~\eqref{eq:family}. Define the misclustering error by
+% \[
+% \ell(\hat z, z)={1\over p}\min_{\pi \in \Pi}\sum_{i\in[p]}\ind\{\hat z(i)\neq \pi \circ z(i)\},
+% \]
+% where $\pi: [r] \mapsto [r]$ is a permutation of cluster labels, $\circ$ denotes the composition operation, and $\Pi$ denotes the collection of all possible permutations.
+The infinitum over all permutations accounts for the ambiguity in cluster label permutation.
+
+% In Sections~\ref{sec:statlimit} and \ref{sec:complimit}, we provide the lower bounds of $\ell (\hat z, z)$ for general Gaussian dTBMs~\eqref{eq:model_margin} without symmetric assumptions.
+For general (asymmetric) Gaussian dTBMs, we assume Gaussian noise $\tE(i_1,\ldots,i_K)\stackrel{\text{i.i.d.}}{\sim} N(0,\sigma^2)$, and we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode.
+For notational simplicity, we still use $z$ and $\tP(\gamma)$ for this general (asymmetric) model.
+All lower bounds should be interpreted as the worst-case results across $K$ modes.
+
+\subsection{Assumptions} \label{sec:prelim}
+
+We propose the signal-to-noise ratio~(SNR),
 \begin{align}\label{eq:gamma}
@@
 Intuitively, a larger SNR, or equivalently a larger $\gamma$, benefits the clustering in the presence of noise.
+
 With quantification~\eqref{eq:gamma}, we consider the following parameter space,
@@
 \end{equation}
-Note that $1$-block dTBM does not belong to the space $\tP(\gamma)$ when $\gamma < 0$ by Assumption~\ref{assmp:min_gap}.
-Our goal is to characterize the clustering accuracy with respect to $\gamma$.
-Let $\hat z$ and $z$ be the estimated and true clustering functions in the family~\eqref{eq:family}.
-Define the misclustering error by
+The $1$-block dTBM does not belong to the space $\tP(\gamma)$ when $\gamma < 0$ by Assumption~\ref{assmp:min_gap}.
+Our goal is to characterize the clustering accuracy with respect to $\gamma$ under the space $\tP(\gamma)$.
+
+{
+\color{blue}In our algorithmic development, we often refer to the regime of balanced degree heterogeneity. We call the degree $\mtheta$ \emph{balanced} if
+\begin{equation}\label{eq:degree}
+{\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
+\end{equation}
+
+The following lemma provides the rational of balanced degree assumption.
+We show the close relation between angle gaps in the mean tensor $\tX$ and the core tensor $\tS$ under balanced degree heterogeneity.
+
+\begin{lem}[Angle gaps in $\tX$ and $\tS$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family} \textcolor{red}{with $r \geq 2$.
+Suppose $\mtheta$ is balanced satisfying~\eqref{eq:degree} \textcolor{red}{and $\min_{i\in[p]}\theta(i)\geq c$ from some constant $c>0$}.
+} Then, \textcolor{red}{as $p \rightarrow \infty$}, for all $i,j$ such that $z(i) \neq z(j)$, we have
+\begin{equation}
+ % \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  } \asymp  \onorm{ \frac{\mS_{z(i):}}{\onormSize{}{\mS_{z(i):}}}  -  \frac{\mS_{z(j):}}{\onormSize{}{\mS_{z(j):}}}  },
+ \cos(\mX_{i:}, \mX_{j:})\asymp  \cos(\mS_{z(i):}, \mS_{z(j):}),
+\end{equation}
+
+where $\mX =\mat(\tX)$ and $\mS = \mat (\tS)$.
+\end{lem}
+In practice, an estimation algorithm has access to a noisy version of $\tX$ but not $\tS$. %Intuitively, the performance of dTBM is determined by both the core tensor and degree heterogeneity. 
+%Since we view degrees as nuisance parameters, 
+Our goal is to establish the algorithm performance with respect to the signal $\Delta^2_{\min}$ in the core tensor.
+By Lemma~\ref{lem:angle_gap_x}, the mapping from the core tensor $\mS_{z(i):}$ to the mean tensor $\mX_{z(i):}$ preserves the angle information $\Delta_{\min}^2$ under balanced degree heterogeneity~\eqref{eq:degree}.
+Therefore, the balanced degree assumption helps to exclude the cases in which the degree heterogeneity distorts the algorithm guarantees.
+%Therefore, the balance assumption helps to exclude the cases in which the degree heterogeneity distorts the dTBM performance.
+
+%Indeed, as we will see in Sections~\ref{sec:statlimit}-\ref{sec:complimit}, under the balance assumption~\eqref{eq:degree}, the $\Delta_{\min}^2$ is sufficient to establish the dTBM phase transition.
+
+%In contrast, a highly-skewed degree violating~\eqref{eq:degree} may lead to a worse performance of dTBM than that of TBM. 
+Here, we provide an example to illustrate the insufficiency of $\Delta_{\min}^2$ in the absence of balanced degrees.
+
+% The balance assumption \eqref{eq:degree} ensures the sufficiency of core tensor signal $\Delta_{\min}^2$ to describe the phase transition of dTBM.
+% Without balanced degree,
+
+%The dTBM phase transition without balanced degree may be arbitrary worse than that with balanced degree.
+
+\begin{example}[Insufficiency of $\Delta_{\min}^2$ in the absence of balanced degrees] Consider an order-2 $(p,p)$-dimensional dTBM with core matrix
+\begin{equation}\label{eq:modelexample}
+    \mS = \begin{pmatrix} 1 &a\\
+    1 & -a
+    \end{pmatrix}, \quad \text{and} \ \mtheta \text{ such that } \onormSize{}{\mtheta_{z^{-1}(1)}}^2 = p^m \onormSize{}{\mtheta_{z^{-1}(2)}}^2,
+\end{equation}
+where \textcolor{red}{$m \in [-1,1]$} is a scalar parameter controlling the skewness of degrees. Let $\Delta_{\mX}^2$ denote the minimal angle gap of the mean tensor, defined by
+\begin{equation}\label{eq:delta_x}
+    \Delta_{\mX}^2 \coloneqq \min_{i,j \in [p], z(i) \neq z(j)} \onorm{ \frac{\mX_{i:}}{\onormSize{}{\mX_{i:}}}  -  \frac{\mX_{j:}}{\onormSize{}{\mX_{j:}}}  }, \quad \mX = \mat_1(\tX).
+%\quad  \tX = \tS \times_1 \mTheta \mM \times_2 \cdots \times_K \mTheta \mM.
+\end{equation}
+% {
+% \color{red}
+% Take $ a = c$ for some positive constant $c$ in the model setup~\eqref{eq:modelexample}. We have 
+% \begin{equation}
+%     \Delta_{\min}^2 = \frac{2 a^2}{1 + a^2} \asymp 1, \quad  \Delta_{\mX}^2 = \frac{2\onormSize{}{\mtheta_{z^{-1}(2)}}^{2} a^2}{\onormSize{}{\mtheta_{z^{-1}(1)}}^{2} +  \onormSize{}{\mtheta_{z^{-1}(2)}}^{2}a^2} \asymp p^{-m}.
+% \end{equation}
+
+% The notion $\Delta_{\min}^2$ achieves the maximum $\tO(1)$ while the the dTBM is still impossible when $m = -1$, i.e., $\Delta^2_{\mX} \lesssim p^{-1}$, based on our theory in later Sections.
+That is, the dTBM estimation depends on the magnitude of $m$.
+In such a setting, the proposed signal notion $\Delta^2_{\min}$ alone fails to fully characterize dTBM, and the phase transition of $\Delta_{\min}^2$ can be arbitrarily worse than that with balance assumption.
+% }
+ Take $ a = p^{-1/4}$ in the model setup~\eqref{eq:modelexample}. We have 
+\begin{equation}
+    \Delta_{\min}^2 = \frac{2 a^2}{1 + a^2} \asymp p^{-1/2}, \quad  \Delta_{\mX}^2 = \frac{2\onormSize{}{\mtheta_{z^{-1}(2)}}^{2} a^2}{\onormSize{}{\mtheta_{z^{-1}(1)}}^{2} +  \onormSize{}{\mtheta_{z^{-1}(2)}}^{2}a^2} \asymp p^{-1/2-m}.
+\end{equation}
+%Take $ a = p^{-1/4}. We have 
+%\begin{equation}
+   % \Delta_{\min}^2 = p^{-1/2} \quad \text{and} \quad    \Delta_{\mX}^2 = {1\over p^{m-1/2}+1} = \min(p^{1/2-m},1). 
+    %{2 a^2 \over M+a^2} = p^{-1-\epsilon}.
+%\end{equation}
+% \begin{align}
+%     &\text{if } a = p^{-1/2}, M = 1& \quad &\Rightarrow \quad \Delta_{\min}^2 =  \Delta_{\mX}^2 = p^{-1}; \\
+%     &\text{if } a = p^{-1/4}, M = p^{1/2} & \quad &\Rightarrow \quad \Delta_{\min}^2 = p^{-1/2}, \   \Delta_{\mX}^2 = p^{-1}.
+% \end{align}
+%By the following Theorem~\ref{thm:stats}, no estimator achieves exact recovery of dTBM under $\Delta_{\mX}^2 < p^{-1}$. 
+%Under the balanced degree assumption, the dTBM is computationally solvable based on our Theorem~\ref{thm:comp} in Section~\ref{sec:complimit}.
+{\color{red}Based on our theory in later Sections, the dTBM is impossible to solve when $\Delta^2_{\mX} \lesssim p^{-1}$; 
+}
+%Based on our theory in later Sections, the stat-computational achievability of dTBM occurs when $\Delta^2_{\mX}>p^{-1}$; 
+that is, the dTBM estimation depends on the relative magnitude of $m$ vs. $1/2$. In such a setting, the proposed signal notion $\Delta^2_{\min}$ alone fails to fully characterize dTBM. 
+\textcolor{red}{}
+
+%However, the condition $ \Delta_{\min}^2 = p^{-1/2}$ with balanced degree leads to a computationally efficient dTBM by Theorem~\ref{thm:refinement}.
+%$-1\leq \varepsilon \leq 1/2$.
+ %($\tO(p^{-1}) \leq M \leq \tO(p)$ by the parameter space~\eqref{eq:family}). 
+\end{example}
+
+\begin{rmk}[Flexibility in balanced degree assumption] One important note is that our balance assumption~\eqref{eq:degree} does not preclude the mild degree heterogeneity.
+In fact, within each of the clusters, we allow the highest degree at the order $\tO(p)$, whereas the lowest degree at the order $\Omega(1)$.
+This range is more relaxed than previous work \citep{gao2018community} that restricts the highest degree in the sub-linear regime $o(p)$ and the lowest degree at the order $\Omega(1)$.
+\end{rmk}
+
+\begin{rmk}[Similar assumptions in literature]
+%We propose he balance assumption~\eqref{eq:degree} to avoid the extreme inflation or elimination of the core tensor gap $\Delta_{\min}^2$ after the weighted mapping by degree heterogeneity. 
+Similar degree regulations are not rare in literature.
+In higher-order tensor model \citep{ke2019community}, the degree assumption $\max_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}} \leq $ $ C \min_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}$ is made to ensure balancedness across communities.
+%Since they have no $\ell_1$ constraint on $\mtheta$, their balance assumption is equivalent to ours~\eqref{eq:degree} multiplying some constants to the degrees.
+"
+In \cite{gao2018community}, the degree distribution is restricted to ${1\over |z^{-1}(a)|}\sum_{i\in z^{-1}(a)}\theta_i=1+o(1)$ for all communities. 
+\end{rmk}
+}
+
+Last, let $\hat z$ and $z$ be the estimated and true clustering functions in the family~\eqref{eq:family}. Define the misclustering error by
 \[
@@
 
-In Sections~\ref{sec:statlimit} and \ref{sec:complimit}, we provide the lower bounds of $\ell (\hat z, z)$ for general Gaussian dTBMs~\eqref{eq:model_margin} without symmetric assumptions.
+In Sections~\ref{sec:statlimit} and \ref{sec:complimit}, we provide the phase transition of $\ell (\hat z, z)$ for general Gaussian dTBMs~\eqref{eq:model_margin} without symmetric assumptions.
 For general (asymmetric) Gaussian dTBMs, we assume Gaussian noise $\tE(i_1,\ldots,i_K)\stackrel{\text{i.i.d.}}{\sim} N(0,\sigma^2)$, and we extend the parameter space~\eqref{eq:family} to allow $K$ clustering functions $(z_k)_{k\in[K]}$, one for each mode.
 For notational simplicity, we still use $z$ and $\tP(\gamma)$ for this general (asymmetric) model.
-All lower bounds should be interpreted as the worst-case results across $K$ modes.
-
-\subsection{Statistical critical values}\label{sec:statlimit}
-The statistical limit means the minimal SNR required for solving dTBMs with \emph{unlimited computational cost.} Our following result shows the minimax lower bound of SNR for exact recovery in dTBM.
-
-\begin{thm}[Statistical lower bound]\label{thm:stats} Consider general Gaussian dTBMs under the parameter space $\tP(\gamma)$ with $K\geq 1$.
+All results should be interpreted as the worst-case results across $K$ modes.
+
+\subsection{Statistical critical value}\label{sec:statlimit}
+The statistical critical value means the SNR required for solving dTBMs with \emph{unlimited computational cost.} \textcolor{blue}{Our following result shows the minimax lower bound for exact recovery and the matching upper bound for maximum likelihood estimator (MLE).
+}
+{
+\color{blue} We consider the Gaussian MLE, denoted as $(\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}})$, over the estimation space $\tP(\gamma)$, where 
+\begin{equation}\label{eq:mle}
+    (\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}}) = \argmin_{ (z, \mtheta, \tS) \in \tP (\gamma)} \onormSize{}{\tY - \tX(z,\mtheta, \tS)}_F^2.
+\end{equation}
+
+\begin{thm}[Statistical critical value]\label{thm:stats} Consider general Gaussian dTBMs with parameter space $\tP(\gamma)$ with $K\geq 1$. Then, we have the following statistical phase transition.
+
+\begin{itemize}[wide]
+    \item \textbf{Impossibility.} \textcolor{red}{Assume $p \rightarrow \infty$ and $2 \leq r\lesssim p^{1/3}$.} If the signal exponent satisfies $\gamma < -(K-1)$, then, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation; that is,
+\begin{equation}
+  \gamma< -(K-1) \quad \Rightarrow  \inf_{\hat z_{\text{stat}}}\sup_{(z, \tS, \mtheta)\in \textcolor{red}{\tP(\gamma)}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
+\end{equation}
+\textcolor{red}{The impossibility of MLE also holds for the general Gaussian dTBMs with parameters $(z, \mtheta, \tS) \in \tP$ such that $\Delta_{\mX}^2 \lesssim p^{-(K-1)}$, where $\Delta_{\mX}^2$ is the minimal angle in mean tensor defined in \eqref{eq:delta_x}.}
+
+\item \textbf{MLE achievability.} Suppose the signal exponent satisfies $\gamma >-(K-1)+c_0$ for an arbitrary constant \textcolor{red}{$c_0>0$}.
+Furthermore, assume \textcolor{red}{$\mtheta$} is balanced and $\min_{i\in[p]}\theta(i)\geq c$ from some constant $c>0$.
+Then, \textcolor{red}{when $p \rightarrow \infty$, fixed $r \geq 1$, and $K \geq 3$, }the MLE in~\eqref{eq:mle} achieves exact recovery in high probability; that is,
+\begin{equation}
+\gamma> -(K-1)+c \quad  \Rightarrow  \quad \ell(\hat z_{\text{MLE}}, z) \lesssim \text{SNR}^{-1}\exp\left(-{p^{K-1}\text{SNR}\over r^{K-1}}\right) \to 0,
+\end{equation}
+with probability going to 1. 
+\end{itemize}
+\end{thm}
+
+The proofs for the two parts in Theorem~\ref{thm:stats} are in the Appendix II, Section~\ref{sec:statprove1} and Section~\ref{sec:statprove2}, respectively.
+The first part of Theorem~\ref{thm:stats} demonstrates the impossibility of exact recovery when $\gamma < -(K-1)$.
+Inspection of the proof shows that the impossibility holds with or without the balanced degree assumption.
+%Moreover, with general degree, Theorem~\ref{thm:stats} also indicates the impossibility when the mean tensor gap is below the threshold $p^{-(K-1)}$.
+The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to MLE and trace maximization~\citep{ghoshdastidar2017uniform}.
+
+The second part of Theorem~\ref{thm:stats} shows the exact recovery of MLE when the $\gamma>-(K-1)+c_0$ for an arbitrary constant \textcolor{red}{$c_0>0$}.
+Combining the impossibility and achievability results, we conclude that the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM with respect to our SNR.
+}
+
+% \begin{thm}[Statistical lower bound]\label{thm:stats} Consider general Gaussian dTBMs under the parameter space $\tP(\gamma)$ with $K\geq 1$.
 Assume $r \lesssim p^{1/3}$.
 If the signal exponent satisfies $\gamma < -(K-1)$, then, every estimator $\hat z_{\text{stat}}$ obeys
-\begin{equation}
-    \sup_{(z, \tS, \mtheta) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
-\end{equation}
-\end{thm}
-Theorem~\ref{thm:stats} demonstrates the impossibility of exact recovery of the assignment when $\gamma < -(K-1)$ in the high-dimensional regime $p\to \infty$ for fixed $r$.
-The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to, maximum likelihood estimation (MLE)~\citep{wang2019multiway} and trace maximization~\citep{ghoshdastidar2017uniform}.
-As we will show in Section~\ref{sec:alg}, the SNR threshold $-(K-1)$ is also a minimax upper bound, because MLE achieves exact recovery when $\gamma > -(K-1)$.
+% \begin{equation}
+%     \sup_{(z, \tS, \mtheta) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
+% \end{equation}
+% \end{thm}
+% Theorem~\ref{thm:stats} demonstrates the impossibility of exact recovery of the assignment when $\gamma < -(K-1)$ in the high-dimensional regime $p\to \infty$ for fixed $r$ \textcolor{blue}{ and balanced $\mtheta$.
+Moreover, with general degree,  Theorem~\ref{thm:stats} also indicates the impossibility when the mean tensor gap is below the threshold $p^{-(K-1)}$.} The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to MLE and trace maximization~\citep{ghoshdastidar2017uniform}.
+
+% As we will show in Section~\ref{sec:alg}, the SNR threshold $-(K-1)$ is also a minimax upper bound, because MLE achieves exact recovery when $\gamma > -(K-1)$.
 Hence, the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM.
 
-\subsection{Computational critical values}\label{sec:complimit}
-In this section, we derive the computational limits of dTBMs.
-The computational limit means the minimal SNR required for exactly recovery with \emph{polynomial-time} computational cost.
+{
+\color{blue}
+% We consider the MLE of Gaussian dTBM under mild assumptions on the heterogeneity.
+
+% We call $\mtheta$ is balanced, if the relative extent of heterogeneity is comparable across clusters in that
+% \begin{equation}\label{eq:degree}
+% {\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
+% \end{equation}
+
+% The balance assumption~\eqref{eq:degree} does not preclude degree heterogeneity.
+Indeed, within each of the clusters, the highest degree can be $\theta(i) = \Omega(p)$, whereas the lowest degree can be $\theta(i)=\tO(1)$.
+The assumptions~\eqref{eq:degree} is proposed for technical reason.
+In practice, we implement numerical verification for the limit results and the polynomial algorithm simulations without the balance constraint on $\mtheta$; see Section~\ref{sec:simulation}.
+
+% Define the parameter space 
+% \begin{equation}
+%     \tP_{b}(\gamma) \coloneqq \tP(\gamma)  \cap \{ \mtheta: \mtheta \text{ is balanced, } \min_{i \in [p]}\theta(i) > c \},
+% \end{equation}
+% $c > 0$ is an universal positive lower bound of the minimal heterogeneity.
+
+% Next, we establish the MLE upper bound to show the achievement above the impossibility threshold $ \gamma = -(K-1)$.
+
+% \begin{thm}[Statistical lower bound: MLE achievement]\label{thm:mle} Consider general Gaussian dTBM over $ \tP(\gamma)$ with $K \geq 1$ and balanced $\mtheta$.
+If the signal exponent satisfies $\gamma \geq -(K-1) + \epsilon$ for arbitrary $\epsilon > 0$, the MLE in~\eqref{eq:mle} obeys
+% \begin{equation}
+%     \ell(\hat z_{\text{MLE}}, z) \lesssim \exp \of{ - \frac{p^{K-1}}{r^{K-1}} \Delta_{\min}^2 }.
+% \end{equation}
+% \end{thm}
+
+% Theorem~\ref{thm:mle} shows the exact recovery of MLE with signal exponent $\gamma > -(K-1)$ and mild regulation on heterogeneity $\mtheta$.
+Therefore, MLE matches the statistical lower bound in Theorem~\ref{thm:stats}, and the SNR threshold $-(K-1)$ is also a minimax upper bound.
+Hence, the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM.
+
+}
+
+\subsection{Computational critical value}\label{sec:complimit}
+In this section, we derive the computational critical value of dTBMs.
+The computational critical value means the minimal SNR required for exactly recovery with \emph{polynomial-time} computational cost.
 An important ingredient to establish the computational limits is the \emph{hypergraphic planted clique (HPC) conjecture} \citep{zhang2018tensor, brennan2020reducibility}.
@@
 
-\begin{thm}[Computational lower bound]\label{thm:comp} Consider general Gaussian dTBMs under the parameter space $\tP(\gamma)$ with $K\geq 2$.
-Assume HPC conjecture holds.
-If the signal exponent $\gamma < -K/2$, then, every \emph{polynomial-time estimator} $\hat z_{\text{comp}}$ obeys
-\begin{align}
-   \liminf_{p\to \infty}\sup_{(z, \tS, \mtheta) \in \tP(\gamma)}  \bbE \left[ p\ell(\hat z_{\text{comp}}, z) \right]\geq 1.
-\end{align}
+{
+\color{blue}
+\begin{thm}[Computational critical value]\label{thm:comp} 
+Consider general Gaussian dTBMs under the parameter space $\tP$ with $K\geq 2$. Then, we have the following computational phase transition.
+
+\begin{itemize}[wide]
+    \item \textbf{Impossibility.} Assume HPC conjecture holds \textcolor{red}{and $r \geq 2$}.
+If the signal exponent satisfies $\gamma < -K/2$, then, no \emph{polynomial-time estimator} $\hat z_{\text{comp}}$ achieves exact recovery in expectation as $p\to \infty$; that is,
+\begin{align}
+  \gamma<-K/2 \ \Rightarrow \ \liminf_{p\to \infty}\sup_{(z, \tS, \mtheta) \in \tP(\gamma)}  \bbE \left[ p\ell(\hat z_{\text{comp}}, z) \right]\geq 1.
+\end{align}
+\item \textbf{Polynomial-time algorithm achievability.} Suppose the parameter space satisfies $\gamma  >  -K/2+c_0$ for an arbitrary constant \textcolor{red}{$c_0>0$}.
+Furthermore, assume \textcolor{red}{fixed $r \geq 1, K \geq 2$,} the degree is balanced, lower bounded in that $\min_{i\in[p]}\theta_i\geq c$ for some constant $c>0$, and satisfies the local linear stability in Definition~\ref{def:stable} \textcolor{red}{in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \geq \check{C}\log^{-1}p $ with some positive constant $\check{C}$.}
+%in the neighborhood $\tN(z,\varepsilon)$ for $\varepsilon \geq \log^{-1}(p)$. 
+Then, \textcolor{red}{as $p \rightarrow \infty$}, there exists a polynomial-time algorithm $\hat z_{\text{ploy}}$ that achieves exact recovery in high probability; that is,
+\begin{equation}
+     % \ell(\hat z_{\text{poly}}, z) \lesssim \exp \of{ - \frac{p^{K-1}}{r^{K-1}} \Delta_{\min}^2 }.
+  \gamma>-K/2+c\ \Rightarrow   \  \ell(\hat z_{\text{poly}}, z) \lesssim \text{SNR}^{-1}\exp \of{ - \frac{p^{K-1}\text{SNR}}{r^{K-1}} } \to 0,
+\end{equation}
+with probability going to 1. 
+\end{itemize}
+
 \end{thm}
-Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$.
-Therefore, $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM.
+}
+
+% under the parameter space $\tP(\gamma)$ with $K\geq 2$. Assume HPC conjecture holds. If the signal exponent $\gamma < -K/2$, then, every \emph{polynomial-time estimator} $\hat z_{\text{comp}}$ obeys
+% \begin{align}
+%   \liminf_{p\to \infty}\sup_{(z, \tS, \mtheta) \in \tP(\gamma)}  \bbE \left[ p\ell(\hat z_{\text{comp}}, z) \right]\geq 1.
+% \end{align}
+{
+\color{blue}
+The proofs for the two parts in Theorem~\ref{thm:comp} are in the Appendix II, Section~\ref{sec:compprove1} and Section~\ref{sec:statprove2}, respectively.
+The first part of Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$, and the second part shows the existence of such algorithm when $\gamma > -K/2+c_0$ \textcolor{red}{for an arbitrary constant $c_0 > 0$} under extra technical assumptions.
+In Section~\ref{sec:alg}, we will present an efficient polynomial-time algorithm in this setting.
+Therefore, we conclude that $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM with respect to our SNR.
+
+% Therefore, $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM.
 In Section~\ref{sec:alg}, we will show the condition $\gamma> -K/2$ suffices for our proposed polynomial-time estimator.
-Thus, $\gamma_{\text{comp}}:=-K/2$ is the critical value for computational performance of dTBM.
+Thus, $\gamma_{\text{comp}}:=-K/2$
+}
 
@@
 
-\begin{rmk}[Comparison with non-degree models]
-We compare our results to non-degree tensor models.
+% \begin{rmk}[Comparison with non-degree models]
+% We compare our results to non-degree tensor models.
 The allowance of degree heterogeneity $\mtheta$ makes the model more flexible, but it incurs extra statistical and computational complexity.
 Fortunately, we find that the extra complexity does not render the estimation of $z$ qualitatively harder; see the comparison of our phase transition with non-degree TBM~\citep{han2020exact}.
-\end{rmk}
+% \end{rmk}
 
@@
 In what follows, we first use the symmetric tensor as a working example to describe the algorithm procedures to gain insight.
-Our theoretical analysis focuses on the noisy tensor with i.i.d.\ sub-Gaussian noise such as Gaussian and uniform observations.
-The extensions for asymmetric tensor and Bernoulli observation and other practical issues are in Section~\ref{subsec:exten}.
+\textcolor{blue}{Our theoretical analysis focuses on the symmetric mean tensor with independent sub-Gaussian noises}
+%i.i.d.\ sub-Gaussian noise
+such as Gaussian and uniform observations.
+\textcolor{blue}{The extensions for Bernoulli observations, asymmetric mean tensors, and other practical issues are in Sections~\ref{subsec:ber} and \ref{subsec:exten}.}
 
@@
 
+{
+\color{blue}
+
+To construct algorithm guarantees,  we introduce the misclustering loss between an estimator $\hat z$ and the true $z$:
+\begin{equation}\label{eq:defnofL}
+    L(\hat z, z) = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ \hat z(i) = b } \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
+\end{equation}
+where the superscript $\cdot^{s}$ denotes the normalized vector; i.e., $\ma^s:=\ma/\onorm{\ma}$ if $\ma\neq 0$ and $\ma^{s}=0$ if $\ma=0$ for any vector $\ma$.
+The following lemma indicates the close relationship between the loss $L(\hat z, z)$ and error $\ell(\hat z, z)$.
+The loss $L(\hat z, z)$ serves as an important intermediate quantity to control the misclustering error.
+
+\begin{lem}[Misclustering error and loss]\label{lem:mis}  \textcolor{red}{Consider the dTBM under the parameter space $\tP$.} Suppose $\min_{i \in [p]}\theta(i) > c$ for some constant $c > 0$.
+We have $\ell(\hat z, z) \Delta_{\min}^2 \leq L(\hat z, z)$.
+%\fixme{miaoyan}{under bounded degree assumptions? please also check all lemmas in supp to make sure conditions are well specified.}
+\end{lem}
+
+}
+
 \subsection{Weighted higher-order initialization}
@@
 We start with weighted higher-order clustering algorithm as initialization.
-We take an order-3 symmetric tensor as illustration for insight.
-Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
+\textcolor{blue}{We take an order-3 tensor and the clustering on the first mode as illustration for insight.} Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
 By model~\eqref{eq:model_tensor}, for all $i \in [p]$, we have
@@
 The denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) estimates $\tX$ from $\tY$ by a double projection spectral method.
-The first projection performs HOSVD~\citep{de2000multilinear} via $\mU_{\text{pre}} = \text{SVD}_{r} \of{ \text{Mat}(\tY) }$, where $\text{SVD}_r(\cdot)$ returns the top-$r$ left singular vectors.
-The second projection performs HOSVD on the projected $\tY$ onto the multilinear Kronecker space $\mU_{\text{pre}} \otimes  \mU_{\text{pre}}$; i.e.,
+The first projection performs HOSVD~\citep{de2000multilinear} via $\mU_{\text{pre},k} = \text{SVD}_{r} \of{ \text{Mat}_k(\tY) }, k \in [3]$, where $\text{SVD}_r(\cdot)$ returns the top-$r$ left singular vectors.
+The second projection performs HOSVD on the projected $\tY$ onto the multilinear Kronecker space $\mU_{\text{pre},k} \otimes  \mU_{\text{pre},k}$; i.e.,
 \begin{equation}\label{eq:two-step_factor}
-    \hat \mU = \text{SVD}_{r} \of{\text{Mat}\of{ \tY\times_1  \mU_{\text{pre}} \mU_{\text{pre}}^T \times_2   \mU_{\text{pre}} \mU_{\text{pre}}^T }}.
-\end{equation}
+    \hat \mU_1 = \text{SVD}_{r} \of{\text{Mat}_1\of{ \tY\times_2  \mU_{\text{pre},2} \mU_{\text{pre},2}^T \times_3   \mU_{\text{pre},3} \mU_{\text{pre},3}^T }}.
+\end{equation}
+and similar for $ \hat \mU_2 ,  \hat \mU_3 $.
 The final denoised tensor $\hat \tX$ is defined by
 \begin{equation}\label{eq:two-step_est}
-    \hat \tX = \tY \times_1 \hat \mU \hat 
-\mU^T \times_2 \hat \mU \hat \mU^T \times_3 \hat \mU \hat \mU^T. 
+    \hat \tX = \tY \times_1 \hat \mU_1 \hat 
+\mU^T_1 \times_2 \hat \mU_2 \hat \mU^T_2 \times_3 \hat \mU_3 \hat \mU^T_3. 
 \end{equation}
@@
 The clustering step (lines 3-5 in Sub-algorithm~\hyperref[alg:main]{1}) performs the weighted $k$-means clustering. 
-We write $\hat \mX=\text{Mat}(\hat \tX)$, and normalize the rows into $\hat \mX^s_{i:}=\onormSize{}{\hat \mX_{i:}}^{-1}\hat \mX_{i:}$ as a surrogate of $\theta(i)^{-1} \mX_{i:}$.
+We write $\hat \mX=\text{Mat}_1(\hat \tX)$, and normalize the rows into $\hat \mX^s_{i:}=\onormSize{}{\hat \mX_{i:}}^{-1}\hat \mX_{i:}$ as a surrogate of $\theta(i)^{-1} \mX_{i:}$.
 Then, a weighted $k$-means clustering is performed on the normalized rows with weights equal to $\onormSize{}{\hat \mX_{i:}}^2$.
 The choice of weights is to bound the $k$-means objective function by the Frobenius-norm accuracy of $\hat \tX$.
-Unlike existing clustering algorithm~\citep{ke2019community}, we apply the clustering on the unfolded tensor $\hat \mX$ rather than on the factors $\hat \mU$.
+Unlike existing clustering algorithm~\citep{ke2019community}, we apply the clustering on the unfolded tensor $\hat \mX$ rather than on the factors $\hat \mU_1$.
 This strategy relaxes the singular-value gap condition~\citep{gao2018community, han2020exact}.
 We assign degenerate rows with purely zero entries to an arbitrarily random cluster; these nodes are negligible in high-dimensions because of the lower bound on $\onormSize{}{\text{Mat}(\tS)_{a:}}$ in~\eqref{eq:family}.
-The final result gives the initial cluster assignment $\hat z^{(0)}$.
-Full procedures are provided in Sub-algorithm~\hyperref[alg:main]{1}.
+The final result gives the initial cluster assignment $z^{(0)}$.
+Full procedures for clustering are provided in Sub-algorithm~\hyperref[alg:main]{1}.
 
@@
 \INPUT Observation $\tY \in \bbR^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
-\State Compute factor matrix $ \mU_{\text{pre}} = \text{SVD}_{r} (\text{Mat}(\tY))$ and the $(K-1)$-mode projection $
-\tX_{\text{pre}} = \tY \times_1   \mU_{\text{pre}} \mU_{\text{pre}}^T \times_2 \cdots \times_{K-1}  \mU_{\text{pre}} \mU_{\text{pre}}^T.$
-\State Compute factor matrix $\hat \mU = \text{SVD}_{r}(\text{Mat}(\tX_{\text{pre}}))$ and denoised tensor
-$\hat \tX = \tY \times_1 \hat \mU \hat \mU^T \times_2 \cdots \times_K \hat \mU \hat \mU^T.$
-\State {Let $\hat \mX = \text{Mat}(\hat \tX)$ and $S_0=\{i \in [p]: \onormSize{}{\hat \mX_{i:}} = 0\}$. Set $\hat z(i)$ randomly in $[r]$ for $i \in S_0$.}
-\State{For all $i\in S_0^c$, compute normalized rows
+
+\State {
+\color{blue}Compute factor matrix $ \mU_{\text{pre},k} = \text{SVD}_{r} (\text{Mat}_k(\tY)), k \in [K]$ and the $(K-1)$-mode projection 
+\begin{equation}
+    \tX_{\text{pre},k} = \tY \times_1   \mU_{\text{pre},1} \mU_{\text{pre},1}^T \times_2 \cdots  \times_{k-1} \mU_{\text{pre},k-1} \mU_{\text{pre},k-1}^T \times_{k+1} \mU_{\text{pre},k-1} \mU_{\text{pre},k+1}^T\times_{K+1}  \times_K \mU_{\text{pre},K} \mU_{\text{pre}, K}^T.
+\end{equation}
+}
+\State {
+\color{blue}
+Compute factor matrix $\hat \mU_k = \text{SVD}_{r}(\text{Mat}_k(\tX_{\text{pre},k})), k \in [K]$ and denoised tensor
+\begin{equation}
+    \hat \tX = \tY \times_1 \hat \mU_1 \hat \mU^T_1 \times_2 \cdots \times_K \hat \mU_K \hat \mU^T_K.
+\end{equation}
+} 
+\For{\textcolor{blue}{$k \in [K]$}}
+\State { \color{blue} Let $\hat \mX = \text{Mat}_k(\hat \tX)$ and $S_0=\{i \in [p]: \onormSize{}{\hat \mX_{i:}} = 0\}$. Set $\hat z(i)$ randomly in $[r]$ for $i \in S_0$.}
+\State{  For all $i\in S_0^c$, compute normalized rows
 $\hat \mX_{i:}^s :=\onormSize{}{\hat \mX_{i:}}^{-1} \hat \mX_{i:}.$
 }
-\State {Solve the clustering $\hat z\colon [p]\to[r]$ and centroids $ (\hat \mx_j)_{j\in[r_k]}$ using weighted $k$-means, such that}
-\begin{align}
-    &\sum_{i \in S_0^c }  \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{\hat z(i)} }^2 
+\State { Solve the clustering $\hat z_k \colon [p]\to[r]$ and centroids $ (\hat \mx_j)_{j\in[r_k]}$ using weighted $k$-means, such that}
+\begin{align}
+    &\sum_{i \in S_0^c }  \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{\hat \mX_{i:}^s - \hat \mx_{\hat z_k(i)} }^2 
     \leq 
-    \eta \min_{\substack{\bar \mx_j, j\in[r], \bar z(i),i\in S_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\bar z(i)}}^2.
-\end{align}
-\OUTPUT Initial clustering $z^{(0)} \leftarrow \hat z$.
-
-\Algphase{Sub-algorithm 2: Angle-based iteration}
-\INPUT Observation $\tY \in \bbR^{p \times \cdots \times p}$, initialization $z^{(0)} \colon [p]\to[r]$ from Sub-algorithm 1, iteration number $T$.
+    \eta \min_{\substack{\bar \mx_j, j\in[r], \bar z_k(i),i\in S_0^{c}}} \sum_{i \in S^c } \onormSize{}{\hat \mX_{i:}}^2 \onormSize{}{ \hat \mX_{i:}^s -   \bar \mx_{\hat z_k(i)}}^2.
+\end{align}
+\EndFor
+
+\OUTPUT {\color{blue} Initial clustering $z^{(0)}_k \leftarrow \hat z_k, k \in [K]$.}
+
+\Algphase{Sub-algorithm 2: Angle-based iteration}\label{alg:2}
+\INPUT Observation $\tY \in \bbR^{p \times \cdots \times p}$, initialization $z^{(0)}_k \colon [p]\to[r], k \in [K]$ from Sub-algorithm 1, iteration number $T$.
 \For {$t = 0$ to $T-1$}
 \State Update the block tensor $\tS^{(t)}$ via
-$\tS^{(t)} (i_1,...,i_K)= \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}(i_k) = j_k, k \in [K]\}.$
-\State Calculate reduced tensor $\tY^{\text{d}} \in \bbR^{p \times r \times \cdots \times r}$ via
-\begin{align}
-    &\tY^{\text{d}}(i,a_2,\ldots,a_K) 
-    = 
-\text{Ave}\{\tY(i,i_2,\ldots,i_K): z^{(t)}(i_k) = a_k, k \neq 1 \}.
-\end{align}
-
-\State Let $\mY^{\text{d}} = \text{Mat}(\tY^{\text{d}})$ and $J_0 = \{ i\in[p]: \onorm{\mY^{\text{d}}_{i:}} = 0\}$. Set $z^{(t+1)}(i)$ randomly in $[r]$ for $i \in J_0$.
-
-\State Let $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. For all $i \in J_0^c$ update the cluster assignment by
-\begin{equation}
-    z(i)^{(t+1)} = \argmax_{a \in [r]} \cos \left( \mY^{\text{d}}_{i:},\ \mS^{(t)}_{a:} \right).
-\end{equation}
-
+$\tS^{(t)} (a_1,...,a_K)= \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}_k(i_k) = a_k, k \in [K]\}.$
+\For{ $k \in [K]$}
+\State {\color{blue} Calculate reduced tensor $\tY^{\text{d}}_k \in \bbR^{r \times r \times p \times r \times \cdots \times r}$ via
+\begin{equation}
+    \tY^{\text{d}}_k(a_1,\ldots,a_{k-1}, i ,a_{k+1},\ldots,a_K) 
+    = \text{Ave}\{\tY(i_1,\ldots,i_{k-1},i,i_{k+1},\ldots,i_K): z^{(t)}(i_j) = a_j, j \neq k \}
+\end{equation}}
+% \begin{align}
+%     &\tY^{\text{d}}(i,a_2,\ldots,a_K) 
+%     = 
+% \text{Ave}\{\tY(i,i_2,\ldots,i_K): z^{(t)}(i_k) = a_k, k \neq 1 \}.
+% \end{align}
+
+\State {\color{blue} Let $\mY_k^{\text{d}} = \text{Mat}_k(\tY^{\text{d}})$ and $J_0 = \{ i\in[p]: \onorm{\mY^{\text{d}}_{i:}} = 0\}$. Set $z_k^{(t+1)}(i)$ randomly in $[r]$ for $i \in J_0$.}
+
+\State {\color{blue} Let $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. For all $i \in J_0^c$ update the cluster assignment by
+\begin{equation}
+    z(i)^{(t+1)}_k = \argmax_{a \in [r]} \cos \left( \mY^{\text{d}}_{k,i:},\ \mS^{(t)}_{a:} \right).
+\end{equation}}
 \EndFor
-
-\OUTPUT Estimated clustering $z^{(T)}  \in [r]^{p}$.
+\EndFor
+
+\OUTPUT { \color{blue} Estimated clustering $z^{(T)}_k  \in [r]^{p}, k \in [K]$.}
 
@@
 
-We now establish the misclustering error rate of initialization. We call $\mtheta$ is balanced, if the relative extent of heterogeneity is comparable across clusters in that
-\begin{equation}\label{eq:degree}
-{\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}.
-\end{equation}
-Note that, the assumption~\eqref{eq:degree} does not preclude degree heterogeneity.
-Indeed, within each of the clusters, the highest degree can be $\theta(i) = \Omega(p)$, whereas the lowest degree can be $\theta(i)=\tO(1)$.
-
-\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with i.i.d.\ noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
+We now establish the misclustering error rate of initialization.
+{
+\color{blue}
+\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with \textcolor{red}{fixed $r \geq 1$, $K \geq 2$}, i.i.d.\ noise under the parameter space $\tP$, and Assumption~\ref{assmp:min_gap}.
+Assume $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$ and let $\Delta_{\mX}$ denote the minimal gap in mean tensor defined in~\eqref{eq:delta_x}.
+Let $ z^{(0)}_k$ denote the output of Sub-algorithm~\hyperref[alg:main]{1}.
+With probability going to 1 \textcolor{red}{as $p \rightarrow \infty$}, we have
+\begin{equation}
+    \ell(z^{(0)}_k, z) \lesssim {\sigma^2 r^K p^{-K/2}\over \Delta_{\mX}^2}.
+\end{equation}
+Further, assume $\mtheta$ is balanced as~\eqref{eq:degree}. We have
+\begin{equation}\label{eq:ini}
+ \ell(z_k^{(0)}, z) \lesssim {r^K p^{-K/2}\over \text{SNR}} \quad \text{and} \quad L(z^{(0)}_k, z) \lesssim  {\sigma^2 r^K p^{-K/2}},
+\end{equation}
+\textcolor{red}{with probability going to 1 as $p \rightarrow \infty$.}
+\end{thm}
+}
+
+% \begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with i.i.d.\ noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
 Assume $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$.
@@
 With probability going to 1, we have
-\begin{equation}\label{eq:ini}
-   \ell(z^{(0)}, z) \lesssim {r^K p^{-K/2}\over \text{SNR}}. 
-\end{equation}
-\end{thm}
+% \begin{equation}\label{eq:ini}
+%     \ell(z^{(0)}, z) \lesssim {r^K p^{-K/2}\over \text{SNR}}
+% \end{equation}
+% \end{thm}
 
@@
 
-\begin{rmk}[Failure of conventional tensor HOSVD] If we use conventional HOSVD for tensor denoising; that is, we use $\mU_{\text{pre}}$ in place of $\hat \mU$ in line 2, then the misclustering rate becomes $\tO(p^{-1})$ for all $K\geq 2$.
+\begin{rmk}[Failure of conventional tensor HOSVD] If we use conventional HOSVD for tensor denoising; that is, we use $\mU_{\text{pre},k}$ in place of $\hat \mU,k$ in line 2, then the misclustering rate becomes $\tO(p^{-1})$ for all $K\geq 2$.
 This rate is substantially worse than our current rate~\eqref{eq:ini}.
@@
 
-\begin{rmk}[Singular-value gap-free clustering] Note that our clustering directly applies to the estimated mean tensor $\hat \tX$ rather than the leading tensor factors $\hat \mU$.
+\begin{rmk}[Singular-value gap-free clustering] Note that our clustering directly applies to the estimated mean tensor $\hat \tX$ rather than the leading tensor factors $\hat \mU_k$.
 Applying clustering to the tensor factors suffers from the non-identifiability issue due to the infinitely many orthogonal rotations when the number of blocks $r \geq 3$ in the absence of singular-value gaps.
@@
 
-\subsection{Angle-based iteration}
+\subsection{Angle-based iteration}\label{subsec:angle}
 Our Theorem~\ref{thm:initial} has shown the polynomially decaying error rate from our initialization.
 Now we improve the error rate to exponential decay using local iterations.
+%\textcolor{blue}{We still focus on the first mode clustering for illustration.}
 We propose an angle-based local iteration to improve the outputs from Sub-algorithm~\hyperref[alg:main]{1}. 
@@
  \[
- \tS^{(t)}(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k\in[3]\}.
+ \tS^{(t)}(a_1,a_2,a_3)=\text{Ave}\{\tY(i_1,i_2,i_3)\colon z^{(t)}_k(i_k)=a_k, k\in[3]\}.
 \]
@@
 \[
-\tY^{\text{d}}(i,a_2,a_3)=\text{Ave}\{\tY(i,i_2,i_3)\colon z^{(t)}(i_k)=a_k, k \neq 1\}.
+\tY^{\text{d}}_1(i,a_2,a_3)=\text{Ave}\{\tY(i,i_2,i_3)\colon z^{(t)}_k(i_k)=a_k, k \neq 1\}.
 \]
-We use $\mY^d$ and $ \mS^{(t)}$ to denote the $\text{Mat}(\tY^{\text{d}})$ and $\text{Mat}(\tS^{(t)})$.
-The rows $\mY^d_{i:}$ and $\mS^{(t)}_{a:}$ correspond to the $\mx_i$ and $\ms_a$ in the one-dimensional clustering~\eqref{eq:angle_kmeans}.
+We use $\mY_k^d$ and $ \mS^{(t)}_k$ to denote the $\text{Mat}_k(\tY^{\text{d}})$ and $\text{Mat}_k(\tS^{(t)})$.
+The rows $\mY^d_{k,i:}$ and $\mS^{(t)}_{k,a:}$ correspond to the $\mx_i$ and $\ms_a$ in the one-dimensional clustering~\eqref{eq:angle_kmeans}.
 Then, we obtain the updated assignment by
 \[
-z(i)^{(t+1)} = \argmax_{a \in [r]} \cos \of{ \mY^{\text{d}}_{i:}, \mS^{(t)}_{a:} },\ \text{ for all } i \in [p],
+z_k(i)^{(t+1)} = \argmax_{a \in [r]} \cos \of{ \mY^{\text{d}}_{k,i:}, \mS^{(t)}_{k,a:} },\ \text{ for all } i \in [p],
 \]
-provided that $\mS^{(t)}_{a:}$ is a non-zero vector.
-Otherwise, if $\mS^{(t)}_{a:}$ is a zero vector, then we make the convention to assign $z^{(t+1)}(i)$ randomly in $[p]$.
+provided that $\mS^{(t)}_{k,a:}$ is a non-zero vector.
+Otherwise, if $\mS^{(t)}_{k,a:}$ is a zero vector, then we make the convention to assign $z^{(t+1)}_k(i)$ randomly in $[p]$.
 Full procedures for our angle-based iteration are described in Sub-algorithm~\hyperref[alg:main]{2}.
@@
 We now establish the misclustering error rate of iterations under the stability assumption. 
-
 \begin{defn}[Locally linear stability] \label{def:stable}
@@
 The condition~\eqref{eq:local} controls the impact of node degree to the $\mp_{\theta}(\cdot)$ with respect to the misclassification rate $\varepsilon$ and angle gap.
-
-\begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the setup as in Theorem~\ref{thm:initial}.
-Assume the local linear stability of degree holds in the neighborhood $\tN(z,\epsilon)$ for $\epsilon = \log^{-1}p$.
-Suppose
-$r=\tO(1)$ and $\text{SNR} \geq \tilde C p^{-K/2}\log p$ for some large constant $\tilde C$.
-Let $z^{(t)}$ denote the $t$-th iteration output in Sub-algorithm~\hyperref[alg:main]{2} with initialization $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{1}.
-With probability going to 1, there exists a contraction parameter $\rho \in (0,1)$ such that
+\textcolor{blue}{Intuitively, the condition~\eqref{eq:local} controls the skewness of degree so that the angle between raw cluster size and degree-weighted cluster size is well controlled.
+The stability assumption is proposed for technical convenience, and we relax this condition in numerical studies; see Section~\ref{sec:simulation}.}
+
+{
+\color{blue}
+
+%\fixme{Jiaxin}{We do not need the stability condition.
+The difference between $|\bar z^{-1}(b)|$ and $\onormSize{}{\mtheta_{\bar z^{-1}(b)}}$ can be bounded by $L(\bar z)$.
+See the new proof on page 52-53.}
+
+\begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the general sub-Gaussian dTBM with \textcolor{red}{fixed $r \geq 1$, $K \geq 2$}, independent noise under the parameter space $\tP$, and Assumption~\ref{assmp:min_gap}.
+\textcolor{red}{Assume the local linear stability of degree holds in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \geq \check{C}\log^{-1}p $ with some positive constant $\check{C}$.}
+%Assume the local linear stability of degree holds in the neighborhood $\tN(z, \epsilon)$ for $\epsilon \geq \log^{-1}(p)$.
+Let $\{z^{(0)}_k\}_{k=1}^K$ be the initialization for Sub-Algorithm~\hyperref[alg:main]{2} and $z^{(t)}_k$ be the $t$-th iteration output on $k$-th mode.
+Suppose $\min_{i \in [p]}\theta(i) \geq c $ for some \textcolor{red}{constant} $c > 0$, the $\text{SNR} \geq \tilde C p^{-(K-1)}\log p$ for some sufficiently large constant $\tilde C$, and the initialization satisfies
+\begin{equation}
+    L(z^{(0)}_k, z) \lesssim \frac{\Delta_{\min}^2}{r \log p}, \quad k \in [K].
+\end{equation}
+ With probability going to 1 \textcolor{red}{as $p \rightarrow \infty$}, there exists a contraction parameter $\rho \in (0,1)$ such that 
 \begin{align}\label{eq:final}
-    \ell(z, \hat z^{(t+1)}) \lesssim &\ \KeepStyleUnderBrace{
+    \ell(z, \hat z^{(t+1)}_k) \lesssim &\ \KeepStyleUnderBrace{
    \text{SNR}^{-1}
-    \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}}_{\substack{\text{statistical error}} }+ \KeepStyleUnderBrace{ \rho^t \ell(z, z^{(0)}). }_{\substack{\text{computational error}}}
+    \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}}_{\substack{\text{statistical error}} }+ \KeepStyleUnderBrace{ \rho^t \ell(z, z^{(0)}_k). }_{\substack{\text{computational error}}}
 \end{align}
@@
 
-Theorem~\ref{thm:refinement} implies that, with probability going to 1, our estimate $z^{(T)}$ achieves exact recovery within polynomial iterations; more precisely,
-\begin{equation}
-     z^{(T)} = \pi \circ z, \quad \text{for all }T\gtrsim \log_{1/\rho} p,
-\end{equation}
-for some permutation $\pi \in \Pi$.
+\begin{cor}[Exact recovery of dTBM with weighted higher-order initialization]  Let the initialization $\{z^{(0)}_k\}_{k = 1}^K$ be the output from Sub-algorithm~\hyperref[alg:main]{1}.
+Assume $\text{SNR} \gtrsim p^{-K/2} \log p$.
+Combining \textcolor{red}{all parameter assumptions and the results in} Theorems~\ref{thm:initial} and \ref{thm:refinement}, with probability going to 1 \textcolor{red}{as $p \rightarrow \infty$}, our estimate $z^{(T)}_k$ achieves exact recovery within polynomial iterations; more precisely,
+\begin{equation}
+     z^{(T)}_k = \pi_k \circ z, \quad \text{for all }T\gtrsim \log_{1/\rho} p\  \text{and}\ k \in [K ].
+\end{equation}
+for some permutation $\pi_k \in \Pi$. 
+\end{cor}
 Therefore, our combined algorithm is \textit{computationally efficient} as long as SNR $\gtrsim p^{-K/2} \log p$.
@@
 
-\subsection{Extensions and practical issues}~\label{subsec:exten}
-\hspace{-.3cm} {\bf Extension for Bernoulli observations.} The main difficulty to establish the statistical guarantee for Bernoulli observations lies in the initialization Sub-algorithm~\hyperref[alg:main]{1}.
-Theorem~\ref{thm:refinement} still holds for Bernoulli observations once the initialization accuracy satisfies the upper bound~\eqref{eq:ini} in Theorem~\ref{thm:initial}.
-
+% Theorem~\ref{thm:refinement} implies that, with probability going to 1, our estimate $z^{(T)}$ achieves exact recovery within polynomial iterations; more precisely,
+% \begin{equation}
+%      z^{(T)} = \pi \circ z, \quad \text{for all }T\gtrsim \log_{1/\rho} p,
+% \end{equation}
+% for some permutation $\pi \in \Pi$.
+Therefore, our combined algorithm is \textit{computationally efficient} as long as SNR $\gtrsim p^{-K/2} \log p$.
+Note that, ignoring the logarithmic term, the minimal SNR requirement, $p^{-K/2}$, coincides with the computational lower bound in Theorem~\ref{thm:comp}.
+Therefore, our algorithm is optimal regarding the signal requirement and lies in the sharpest \emph{computationally efficient} regime in Figure~\ref{fig:phase_axis}.
+
+}
+
+% \begin{defn}[Locally linear stability] \label{def:stable}
+% Define the $\varepsilon$-neighborhood of $z$ by $\tN(z,\epsilon)=\{\bar z\colon \ell(\bar z, z)\leq \epsilon\}$.
+Let $\bar z\colon[p]\to [r]$ be a clustering function.
+We define two vectors associated with $\bar z$,
+% \begin{align}
+%     &\mp(\bar z)=(|\bar z^{-1}(1)|, \ldots,|\bar z^{-1}(r)|)^T, \quad \mp_{\mtheta}(\bar z)=(\onormSize{}{\mtheta_{\bar z^{-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{-1}(r)}}_1)^T.
+% \end{align}
+
+% We call the degree is $\varepsilon$-locally linearly stable if and only if 
+% \begin{equation}\label{eq:local}
+%     \sin(\mp(\bar z),\ \mp_{\mtheta}(\bar z))\lesssim \varepsilon \Delta_{\min} ,\quad \text{for all } \bar z\in \tN(z, \varepsilon).
+% \end{equation}
+% \end{defn}
+
+% Roughly speaking, the vector $\mp(\bar z)$ represents the raw cluster sizes, and $\mp_{\theta}(\bar z)$ represents the relative cluster sizes weighted by degrees. 
+% The local stability holds trivially for $\varepsilon=0$ based on the construction of parameter space~\eqref{eq:family}.
+The condition~\eqref{eq:local} controls the impact of node degree to the $\mp_{\theta}(\cdot)$ with respect to the misclassification rate $\varepsilon$ and angle gap.
+
+% \begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the setup as in Theorem~\ref{thm:initial}.
+Assume the local linear stability of degree holds in the neighborhood $\tN(z,\epsilon)$ for $\epsilon \geq \log^{-1}p$.
+Suppose
+% $r=\tO(1)$ and $\text{SNR} \geq \tilde C p^{-K/2}\log p$ for some sufficiently large constant $\tilde C$.
+Let $z^{(t)}$ denote the $t$-th iteration output in Sub-algorithm~\hyperref[alg:main]{2} with initialization $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{1}.
+With probability going to 1, there exists a contraction parameter $\rho \in (0,1)$ such that
+% \begin{align}\label{eq:final}
+%     \ell(z, \hat z^{(t+1)}) \lesssim &\ \KeepStyleUnderBrace{
+%   \text{SNR}^{-1}
+%     \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}}_{\substack{\text{statistical error}} }+ \KeepStyleUnderBrace{ \rho^t \ell(z, z^{(0)}). }_{\substack{\text{computational error}}}
+% \end{align}
+% \end{thm}
+% From the conclusion~\eqref{eq:final}, we find that the iteration error is decomposed into two parts: statistical error and computational error.
+The statistical error is unavoidable with noisy data regardless $t$, whereas the computational error decays in an exponential rate as the number of iterations $t \rightarrow \infty$.
+
+% Theorem~\ref{thm:refinement} implies that, with probability going to 1, our estimate $z^{(T)}$ achieves exact recovery within polynomial iterations; more precisely,
+% \begin{equation}
+%      z^{(T)} = \pi \circ z, \quad \text{for all }T\gtrsim \log_{1/\rho} p,
+% \end{equation}
+% for some permutation $\pi \in \Pi$.
+Therefore, our combined algorithm is \textit{computationally efficient} as long as SNR $\gtrsim p^{-K/2} \log p$.
+Note that, ignoring the logarithmic term, the minimal SNR requirement, $p^{-K/2}$, coincides with the computational lower bound in Theorem~\ref{thm:comp}.
+Therefore, our algorithm is optimal regarding the signal requirement and lies in the sharpest \emph{computationally efficient} regime in Figure~\ref{fig:phase_axis}.
+
+{
+\color{blue}
+
+\subsection{Extension to Bernoulli observations}\label{subsec:ber}
+
+Bernoulli or network observations are common in multiple fields.
+Our iteration Theorem~\ref{thm:refinement} holds for Bernoulli models, but our initialization Theorem~\ref{thm:initial} does not.
+Moreover, our current dTBM is insufficient to address sparsity with decaying mean tensor.
+Here, we provide extra discussions for Bernoulli initialization and strategies under sparse settings.
+
+\begin{itemize}[wide]
+    \item \textit{Extension to dense binary dTBMs.} The main difficulty to establish initialization guarantees for Bernoulli observations lies in the denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}).
 We now provide a high-level explanation for the technical difficulty when applying Theorem~\ref{thm:initial} to Bernoulli observations.
+
     The derivation of Theorem~\ref{thm:initial} relies on the upper bound of the estimation error for the mean tensor in Lemma~\ref{lem:two-step_esterror}; i.e., with high probability
@@
 where $\tX = \bbE\tY$ and $\hat \tX$ is defined in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}.
+Unfortunately, the inequality~\eqref{eq:prop1} holds only for i.i.d.\ sub-Gaussian observations, while Bernoulli observations are generally not identically distributed.
+
+One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square} of Bernoulli tensor $\tY$. We summarize the procedure in Sub-algorithm 3.
+
+\begin{algorithm}[h!]
+\caption*{\bf Sub-algorithm 3: Weighted higher-order initialization for Bernoulli observation}
+\vspace{.15cm}
+\begin{algorithmic}[1] 
+\INPUT Bernoulli tensor $\tY \in \{0,1\}^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
+\State  Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{\floor{p^{K/2}} \times \ceil{p^{K/2}} }$ denote the nearly square unfolded tensor. Compute the estimate $\tX'$, where
+\begin{equation}\label{eq:matrixsvd}
+    \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX) -  \mat_{sq}(\tY)}_F^2.
+\end{equation}
+\State Implement lines 3-5 of Sub-algorithm~\hyperref[alg:main]{1} with $\hat \tX$ replaced by $\hat \tX'$ in~\eqref{eq:matrixsvd}.
+\OUTPUT Initial clustering $z^{(0)} \leftarrow \hat z$.
+\end{algorithmic}
+\end{algorithm}
+
+\begin{prop}[Error for Bernoulli initialization]\label{cor:ber} Consider the Bernoulli dTBM in the parameter space $\tP$ \textcolor{red}{with fixed $r \geq 1, K \geq 2$} and Assumption~\ref{assmp:min_gap} holds.
+Assume that $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$.
+Let $ z^{(0)}$ denote the output of Sub-algorithm~3.
+With probability going to 1 \textcolor{red}{as $p \rightarrow \infty$}, we have
+\begin{equation}\label{eq:ini_b}
+ \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}, \quad \text{and} \quad L(z^{(0)}, z) \lesssim  {\sigma^2 r^K p^{-\floor{K/2}}}.
+\end{equation}
+\end{prop}
+
+\begin{proof}[Proof of Proposition~\ref{cor:ber}] Following Lemma 7 in \cite{gao2018community}, with high probability, the new estimate satisfies
+\begin{equation}
+    \onormSize{}{\hat \tX' - \tX}_F^2 \lesssim p^{\ceil{K/2}}.
+\end{equation} 
+We obtain the upper bound by using $\hat \tX'$ in place of $\hat \tX$ in Theorem~\ref{thm:initial}.
+\end{proof}
+
+\begin{rmk}[Comparison with previous methods] The Bernoulli bound $\tO(p^{- \floor{K/2} })$ in Proposition~\ref{cor:ber} is relatively looser than the Gaussian bound $\tO(p^{- K/2 })$ in Theorem~\ref{thm:initial}.
+The gap between Bernoulli and Gaussian error decreases as the order $K$ increases.
+Nevertheless, combining with angle iteration Sub-algorithm~\hyperref[alg:main]{2}, Bernoulli clustering still achieves exponential error rate $\exp \of{ - p^{(K-1)}}$ at a price of a larger SNR.
+%Nevertheless, the Bernoulli clustering still achieves exponential error rate $\exp \of{ - p^ \floor{K/2}}$ with large SNR combining with our angle iteration Sub-algorithm~\hyperref[alg:main]{2}. 
+The exponential rate is tighter than the rate error $\tO(p^{-2})$ in previous work~\citep{ke2019community}.
+The investigation of the gap between upper bound $p^{- \floor{K/2}}$ and the lower bound $p^{- K/2}$ for Bernoulli tensors will be left as future work.
+In numerical experiments, we will use our original initialization, Sub-algorithm 1, to verify the robustness to Bernoulli observations.
+\end{rmk}
+
+  \item \textit{Extension to sparse binary dTBMs.} The sparsity is often a popular feature in hypergraphs~\citep{florescu2016spectral,ke2019community}.
+Specifically, the sparse binary dTBM assumes that, the entries of $\tY$ follow independent Bernoulli distributions with the mean
+     \begin{equation}\label{eq:sparse_dtbm}
+    \bbE\tY = \alpha_p \tS \times_1 \mTheta \mM \times_2  \cdots \times_K \mTheta \mM,
+\end{equation}
+where the extra scalar parameter $\alpha_p \in (0,1]$ is function of $p$ that controls the sparsity.
+A smaller $\alpha_p$ indicates a higher level of sparsity.
+Our current work focuses on dense dTBM with $\alpha_p=1$.
+While sparse dTBM is an interesting application, the algorithm and its analysis require different techniques.
+Below, we discuss possible modifications of the algorithm.
+
+%The parameters $(z, \mtheta, \alpha_p \tS)$ are still identifiable up to permutation under our Assumption~\ref{assmp:min_gap}.
+However, the constant lower bound assumption on the core tensor is no longer suitable to describe the decaying sparse parameter.
+Therefore, our phase transition will not directly apply; we expect the extra sparsity parameter $\alpha_p$ should also be included in the signal notion, in order to characterize the phase transition.
+
+The sparsity affects our initialization guarantee in our Theorem~\ref{thm:initial}.
+In our initialization, the spectral denoising step (lines 1-2 in Sub-algorithm~\hyperref[alg:main]{1}) implements matrix SVD to unfolded tensors.
+%, and Theorem~\ref{thm:initial} indicates the good performance under the dense case.
+However, SVD-based methods are believed to fail in extremely sparse SBM due to the localization phenomenon in the singular vectors \citep{florescu2016spectral}.
+Inspired by \cite{florescu2016spectral}, we adapt the diagonal-deleted HOSVD (D-HOSVD) \citep{ke2019community} as the initialization in our higher-order clustering.
+
+The sparsity also affects the iteration guarantee in our Theorem~\ref{thm:refinement}.
+The decaying mean tensor leads to a worse statistical error of order $\tO( -\alpha_p p^{K-1})$ on $\hat \tX$.
+The theoretical analyses for sparse binary dTBM and algorithms are left as future direction.
+Instead, we add numerical experiments to evaluate the robustness of our algorithm and the improvement of D-HOSVD initialization in the sparse dTBM; see Appendix I.
+
+% Inspired by matrix D-SVD, it is beneficial to propose an initialization strategy based on Diagonal Deletion HOSVD (D-HOSVD).
+
+% \begin{algorithm}[h!]
+% \caption*{\bf Algorithm 3: Diagonal Deletion HOSVD initialization for sparse Bernoulli observation}
+% \vspace{.15cm}
+% \begin{algorithmic}[1] 
+% \INPUT Observation $\tY \in \bbR^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
+% \State Calculate the diagonal deleted matrix $\mY_{\text{del}} = \text{Mat}(\tY) \text{Mat}(\tY)^T - \text{diag} ( \text{Mat}(\tY) \text{Mat}(\tY)^T) \in \bbR^{r \times r}$.
+Compute the factor matrix from D-HOSVD
+% \begin{equation}
+%   \hat  \mU_{\text{del}} =  \text{SVD}_{r} (\mY_{\text{del}}), 
+% \end{equation}
+% and the denoised tensor $\tX'' = \tY \times_1 \hat \mU_{\text{del}} \hat \mU^T_{\text{del}}  \times_2 \cdots \times_K \hat \mU_{\text{del}}  \hat \mU^T_{\text{del}}$.
+% \State Implement lines 3-5 of Sub-algorithm~\hyperref[alg:main]{1} with $\tX''$.
+% \OUTPUT Initial clustering $z^{(0)} \leftarrow \hat z$.
+% \end{algorithmic}
+% \end{algorithm}
+
+% Previous work \cite{ke2019community} also adapts D-HOSVD as a warm initialization for singular vector clustering while our Algorithm 3 still clusters over the whole denoised tensor.
+Instead of bounding the singular space $\hat \mU_{\text{del}}^T \mU_{\text{true}}$ in \cite{ke2019community}, the key step is to upper bound the estimation error $\onormSize{}{\tX'' - \tX}_F$ without the singular gap condition.
+New techniques for singular-free tensor estimation with Bernoulli data and D-HOSVD should be developed.
+For iteration, the statistical error increases to $\tO( \exp \of{ - \alpha_p p^{K-1}})$ by the extra sparsity parameter $\alpha_p$.
+The contraction inequality~\eqref{eq:final} is conjectured to be valid due to the angle-based update.
+Numerical simulations in Section~\ref{sec:simulation} verifies the improvement of D-HOSVD initialization and the robustness of angle-based iteration under the sparse setting.
+We leave the specific theoretical analyses for sparse binary dTBM in the future.
+
+\end{itemize}
+
+%   The main difficulty to establish the statistical guarantee for Bernoulli observations lies in the initialization Sub-algorithm~\hyperref[alg:main]{1}.
+Theorem~\ref{thm:refinement} still holds for Bernoulli observations once the initialization accuracy satisfies the upper bound~\eqref{eq:ini} in Theorem~\ref{thm:initial}.
+
+% We now provide a high-level explanation for the technical difficulty when applying Theorem~\ref{thm:initial} to Bernoulli observations.
+The derivation of Theorem~\ref{thm:initial} relies on the upper bound of the estimation error for the mean tensor in Lemma~\ref{lem:two-step_esterror}; i.e., with high probability
+% \begin{equation}\label{eq:prop1}
+%     \onormSize{}{\hat \tX - \tX}_F^2 \lesssim p^{K/2},
+% \end{equation}
+% where $\tX = \bbE\tY$ and $\hat \tX$ is defined in Step 2 of Sub-algorithm~\hyperref[alg:main]{1}.
 Unfortunately, the inequality~\eqref{eq:prop1} holds only for i.i.d.\ sub-Gaussian observations while Bernoulli observations are generally not identically distributed.
 
-One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square} of Bernoulli tensor $\tY$.
+% One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square} of Bernoulli tensor $\tY$.
 Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{\floor{p^{K/2}} \times \ceil{p^{K/2}} }$ denote the nearly square unfolded Bernoulli tensor.
 Define a new estimate
-\begin{equation}\label{eq:matrixsvd}
-    \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX) -  \mat_{sq}(\tY)}_F^2.
-\end{equation}
-The optimization~\eqref{eq:matrixsvd} is simply a matrix SVD problem. Following Lemma 7 in \cite{gao2018community}, with high probability, the new estimate satisfies
-\begin{equation}
-    \onormSize{}{\hat \tX' - \tX}_F^2 \lesssim p^{\ceil{K/2}}.
-\end{equation} 
-Replacing the estimate $\hat \tX$ by $\hat \tX'$ in Theorem~\ref{thm:initial}, the high probability upper bound for Bernoulli initialization is
-\begin{equation}\label{eq:ini_b}
-    \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}. 
-\end{equation}
-The Bernoulli bound~\eqref{eq:ini_b} is relatively looser than Gaussian bound~\eqref{eq:ini}, especially when $K$ is small.
+% \begin{equation}\label{eq:matrixsvd}
+%     \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX) -  \mat_{sq}(\tY)}_F^2.
+% \end{equation}
+% The optimization~\eqref{eq:matrixsvd} is simply a matrix SVD problem. Following Lemma 7 in \cite{gao2018community}, with high probability, the new estimate satisfies
+% \begin{equation}
+%     \onormSize{}{\hat \tX' - \tX}_F^2 \lesssim p^{\ceil{K/2}}.
+% \end{equation} 
+% Replacing the estimate $\hat \tX$ by $\hat \tX'$ in Theorem~\ref{thm:initial}, the high probability upper bound for Bernoulli initialization is
+% \begin{equation}\label{eq:ini_b}
+%     \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}. 
+% \end{equation}
+% The Bernoulli bound~\eqref{eq:ini_b} is relatively looser than Gaussian bound~\eqref{eq:ini}, especially when $K$ is small.
 Nevertheless, our bound~\eqref{eq:ini_b} is already tighter than the pervious work~\citep{ke2019community}.
@@
 
-{\bf Extension for general dTBMs.} Our two-stage algorithm is able to be extended for the general (asymmetric) dTBMs.
+\subsection{Practical issues}~\label{subsec:exten}
+}
+
+% {\bf Extension for general dTBMs.} Our two-stage algorithm is able to be extended for the general (asymmetric) dTBMs.
 Specifically, in the Sub-algorithm~\hyperref[alg:main]{1}, we make the following changes: (1)
-Replace the matrcization $\mat(\tY)$ by $\mat_k(\tY)$; (2) Repeat the Steps 1-5 with mode-specified cluster number $r_k$; (3) Obtain the collection initialization $ \{z^{(0)}_k\}_{k=1}^K$.
+% Replace the matrcization $\mat(\tY)$ by $\mat_k(\tY)$; (2) Repeat the Steps 1-5 with mode-specified cluster number $r_k$; (3) Obtain the collection initialization $ \{z^{(0)}_k\}_{k=1}^K$.
 In the Sub-algorithm~\hyperref[alg:main]{2}, we make the following changes: (1) Take the collection $\{z^{(0)}_k\}_{k=1}^K$ as input, and update the block tensor $\tS^{(t)}$ with the collection $\{ z_k^{(t)}\}_{k=1}^K$, $
-        \tS^{(t)}(i_1,\ldots, i_K) =  \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}_k(i_k) = j_k, k \in [K]\}$;
-(2) Calculate reduced tensor $\tY^{\text{d}}_k$ for each mode via
-\begin{align}
-    &\tY^{\text{d}}_k(a_1,\ldots,a_{k-1}, i ,a_{k+1},\ldots,a_K) 
-    = \text{Ave}\{\tY(i_1,\ldots,i_{k-1},i,i_{k+1},\ldots,i_K): z^{(t)}(i_j) = a_j, j \neq k \}; 
-\end{align}
-
-(3) Repeat Step 8-10 with $\mat_k(\cdot), \tY_k^{\text{d}}$ for each $k \in [K]$ and obtain the collection $\{z^{(T)}_k\}_{k=1}^K$.
-
-Correspondingly, Theorems~\ref{thm:initial} and \ref{thm:refinement} still hold with $\ell(z^{(0)}_k, z_k)$ and $\ell(z^{(t+1)}_k, z_k)$  for all $k \in [K]$.
+%         \tS^{(t)}(a_1,\ldots, a_K) =  \text{Ave} \{\tY(i_1,\ldots,i_K): z^{(t)}_k(i_k) = a_k, k \in [K]\}$;
+% (2) Calculate reduced tensor $\tY^{\text{d}}_k$ for each mode via
+% \begin{align}
+%     &\tY^{\text{d}}_k(a_1,\ldots,a_{k-1}, i ,a_{k+1},\ldots,a_K) 
+%     = \text{Ave}\{\tY(i_1,\ldots,i_{k-1},i,i_{k+1},\ldots,i_K): z^{(t)}(i_j) = a_j, j \neq k \}; 
+% \end{align}
+
+% (3) Repeat Step 8-10 with $\mat_k(\cdot), \tY_k^{\text{d}}$ for each $k \in [K]$ and obtain the collection $\{z^{(T)}_k\}_{k=1}^K$.
+
+% Correspondingly, Theorems~\ref{thm:initial} and \ref{thm:refinement} still hold with $\ell(z^{(0)}_k, z_k)$ and $\ell(z^{(t+1)}_k, z_k)$  for all $k \in [K]$.
 The detailed model extension for general asymmetric dTBMS can be found in Appendix.
@@
 Specifically, the complexity of Sub-algorithm~\hyperref[alg:main]{1} is $\tO(K p^{K+1} + Krp^K)$, where the first term is contributed by the double projection and the calculation of $\hat \tX$, and the second term comes from normalization and the $k$-means.
-The cost of each update in Sub-algorithm~\hyperref[alg:main]{2} is $\tO(p^K + pr^K)$, where $p^K$ comes from the calculation of $\tS^{(t)}$ and $\tY^{\text{d}}$, and $pr^K$ comes from the normalization of $\tY^{\text{d}}$, the calculation of $\tS^{(t)}$, and the cluster assignment update in Step 10.
+The cost of each update in Sub-algorithm~\hyperref[alg:main]{2} is $\tO(p^K + pr^K)$, where $p^K$ comes from the calculation of $\tS^{(t)}$ and $\tY^{\text{d}}$, and $pr^K$ comes from the normalization of $\tY^{\text{d}}$, the calculation of $\tS^{(t)}$, and the cluster assignment update in Step~10.
 
@@
 We evaluate the performance of the BIC criterion in Section~\ref{subsec:num_theory}.
+
+{
+\color{blue}
+
+\section{Comparison with Non-degree Tensor Block Model} \label{sec:tbm}
+We discuss the connections and difference between our dTBM and TBM~\citep{han2020exact} from three aspects: signal notions, theoretical results, and algorithms.
+Without loss of generality, let $\sigma^2=1$.
+
+\begin{itemize}[wide]
+    \item \textit{Signal notion.} The signal levels in both TBM~\citep{han2020exact} and our dTBM are functions of the core tensor $\tS$.
+We emphasize that the signal notions are different between the two models.
+In particular, the Euclidean-based signal notion in TBM~\cite{han2020exact} fails to accurately describe the phase transition in our dTBM due to the possible heterogeneity in degree $\mtheta$.
+To compare, we denote our angle-based signal notion in~\eqref{eq:gamma} and the Euclidean-based SNR in \cite{han2020exact} as $\Delta_{\text{ang}}^2$ and $\Delta_{\text{Euc}}^2$, respectively:
+\begin{equation}
+     \Delta_{\text{ang}}^2 =  2(1 - \max_{a \neq b\in [r]}\cos \of{\mS_{a:},\  \mS_{b:}} ), \quad \Delta_{\text{Euc}}^2 = \min_{a \neq b \in [r]} \onormSize{}{\mS_{a:} - \mS_{b:}}^2.
+\end{equation}
+By Lemma~\ref{lem:norm_diff} in the Appendix II, we have 
+\begin{equation}\label{eq:signalcompare}
+     \Delta_{\text{ang}}^2  \max_{a \in [r]}\onormSize{}{\mS_{a:}}^2 \leq \Delta_{\text{Euc}}^2.
+\end{equation}
+The above inequality indicates the condition $\Delta_{\text{Euc}}^2 \leq p^{\gamma}$ is sufficient but not necessary for $\Delta_{\text{ang}}^2 \leq p^{\gamma}$.
+In fact, if we were to use $\Delta_{\text{Euc}}^2$ for both models, then the phase transition of dTBM can be arbitrarily worse than that for TBM.
+
+Here, we provide an example to illustrate the dramatical difference between TBM and dTBM with the same core tensor.
+
+\begin{example}[Comparison with Euclidean-based signal notion] \label{example:euc_alg} Consider a biclustering model with $\mtheta=1$ and an order-2 core matrix 
+\begin{equation}
+    \mS = \begin{pmatrix} p^{(\gamma+1)/2 } + 2  & 2 p^{(\gamma+1)/2} + 4\\
+    2 & 4
+    \end{pmatrix},\quad \text{with}\ \gamma \leq -1.
+\end{equation}
+The core matrix $\mS$ lies in the parameter spaces of TBM and our dTBM.
+Here, the constraint $\gamma \leq -1$ is added to ensure the bounded condition of $\mS$ in our parameter space in \eqref{eq:family}.
+The angle-based and Euclidean-based signal levels of $\mS$ are
+\begin{equation}
+    \Delta_{\text{ang }}^2(\mS) = 0 \ \left(\leq p^{\gamma}\right), \quad \Delta_{\text{Euc}}^2(\mS) = 5 p^{\gamma + 1} \ \left(\geq p^{\gamma}\right).
+\end{equation}
+We conclude that TBM with $\mS$ achieves exact recovery with a polynomial-time algorithm; see \citet[Theorem 4]{han2020exact}.
+By contrast, the dTBM with the same $\mS$ and input $r=2$ violets the identifiability condition, and thus fails to be solved by all estimators; see our Theorem~\ref{thm:unique}.
+\end{example}
+
+    \item \textit{Theoretical results.} In both works, we study the phase transition of TBM and dTBM with respect to the Euclidean and angle-based SNRs.
+We briefly summarize the results in \cite{han2020exact} and compare with ours.
+
+    \textit{Statistical critical value:}
+    \begin{align}
+        \text{Ours:}& \ \Delta_{\text{ang}}^2 \lesssim p^{-(K-1)} \Rightarrow \text{statistically impossible;} \quad \Delta_{\text{ang}}^2 \gtrsim   p^{-(K-1)} \Rightarrow \text{MLE achieves exact recovery;} \\
+        \text{Han's:}& \ \Delta_{\text{Euc}}^2 \lesssim p^{-(K-1)} \Rightarrow \text{statistically impossible;} \quad \Delta_{\text{Euc}}^2 \gtrsim   p^{-(K-1)} \Rightarrow \text{MLE achieves exact recovery}.
+    \end{align}
+
+     \textit{Computational critical value:}
+    \begin{align}
+        \text{Ours:}& \ \Delta_{\text{ang}}^2 \lesssim p^{-K/2} \Rightarrow \text{computationally impossible;} \quad \Delta_{\text{ang}}^2 \gtrsim   p^{-K/2} \Rightarrow \text{ computationally efficient;} \\
+        \text{Han's:}& \ \Delta_{\text{Euc}}^2 \lesssim p^{-K/2} \Rightarrow \text{computationally impossible;} \quad \Delta_{\text{Euc}}^2 \gtrsim   p^{-K/2} \Rightarrow \text{computationally efficient}.
+    \end{align}
+
+The above comparison reveals three major differences.
+
+ %Our dTBM phase transition re-discovers the statistical-computational gaps in TBM.
+%This finding indicates the intrinsic behavior distinctions among higher-order, matrix, and vector problem may be generally exists in many higher-order problems.
+First, none of our results in Section~\ref{sec:limits} are corollaries of \cite{han2020exact}.
+Both models show the similar conclusion but under different conditions.
+While the TBM impossibility~\citep{han2020exact} provides a necessary condition for our dTBM impossibility, we find that such a condition is often loose.
+There exists a regime of $\tS$ in which TBM problems are computationally efficient but dTBM problems are statistically impossible; see Example~\ref{example:euc_alg}.
+This observation has motivated us to develop the new signal notion $\Delta^2_{\text{ang}}$ for sharp dTBM phase transition conditions.
+
+ Second, to find the phase transition, we need to show both the impossibility and achievability when SNR is below and above the critical value, respectively.
+While the TBM impossibility can serve as a loose condition of our dTBM impossibility, more efforts are required to show the achievability.
+In particular, since TBM is a more restrictive model than dTBM, the achievability in \cite{han2020exact} does not imply the achievability of dTBM in a larger parameter space.
+The latter requires us to develop new MLE and polynomial algorithms for dTBM achievability.
+%For impossibility, the minimax bounds in dTBM search over a larger range of estimators than TBM.
+Though the thresholds for TBM impossibility can serve as the loose lower bounds of the dTBM impossibility, more efforts are required to avoid the gap between impossibility and achievablity.
+
+Third, from the perspective of proofs, we develop new tools to handle the extra degree heterogeneity.
+In our Theorem~\ref{thm:stats}, we consider the profile MLE by maximizing out the nuisance degree parameter, while TBM~\citep{han2020exact} considers the usual MLE without degree parameter.
+In our Theorem~\ref{thm:comp}, we construct a rank-2 tensor to relate HPC conjecture to $\Delta^2_{\text{ang}}$, while TBM~\citep{han2020exact} constructs a rank-1 tensor to relate HPC conjecture to $\Delta^2_{\text{Euc}}$.
+The asymptotic non-equivalence between $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$ renders our proof technically more involved.
+
+    \item \textit{Algorithms.} Both \cite{han2020exact} and our work propose the two-step algorithm, which combines warm initialization and iterative refinement to achieve exact recovery.
+This local-to-global strategy is not new in clustering literature~\citep{gao2022iterative, chien2019minimax}.
+The highlight of our algorithm is the angle-based update in lines 10-14, Sub-algorithm~\hyperref[alg:main]{2}, which is specifically designed for dTBM to avoid the estimation of $\mtheta$.
+This angle-based update brings new proof challenges.
+We develop polar-coordinate based techniques to establish the error rate for the proposed algorithm.
+\end{itemize}
+}
 
@@
     \item \textbf{\small HLloyd}~\citep{han2020exact}: High-order clustering algorithm developed for non-degree tensor block models;
-    \item \textbf{\small SCORE}~\citep{ke2019community}: Tensor-SCORE for clustering developed for binary tensors.
+    \item \textbf{\small SCORE}~\citep{ke2019community}: Tensor-SCORE for clustering developed for sparse binary tensors.
 \end{itemize}
@@
 We investigate the effects of signal to clustering performance by varying $\gamma \in [-1.5, -1.1]$.
-Figure~\ref{fig:comp_gamma} shows the consistent outperformance of our method \textbf{\small dTBM} among all algorithms.
+Figure~\ref{fig:comp_gamma} shows that our method \textbf{\small dTBM} outperforms all other algorithms.
 The sub-optimality of \textbf{\small SCORE} and \textbf{\small HOSVD+} indicates the necessity of local iterations on the clustering.
@@
 
-We apply our generalized algorithm to the HCP data with the numbers of clusters on three modes $r_1 = r_2 = 4$ and $r_3 = 3$.
+We apply our general asymmetric algorithm to the HCP data with the numbers of clusters on three modes $r_1 = r_2 = 4$ and $r_3 = 3$.
 The selection of $r_1$ and $r_2$ follows the human brain anatomy and the symmetry in the brain network, and the $r_3$ is specified following previous analysis~\citep{hu2021generalized}.
@@
 
-In this section, we provide the proof sketches for the main Theorems~\ref{thm:initial}-\ref{thm:refinement}. Detail proofs and extra theoretical results are provided in Appendix.
+In this section, we provide the proof sketches for the main Theorems~\ref{thm:initial}-\ref{thm:refinement}. Detail proofs and extra theoretical results are provided in Appendix II.
 
@@
 
-\subsection{Proof sketch of Theorem 5}
+\subsection{Proof sketch of Theorem~\ref{thm:refinement}}\label{sec:thm5}
 The proof of Theorem~\ref{thm:refinement} is inspired by the proof idea of \citet[Theorem 2]{han2020exact}.
 We develop extra polar-coordinate based techniques with angle gap characterization to address the nuisance degree heterogeneity.
-We introduce an intermediate quantity called misclustering loss
-\begin{equation}
-     L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
+Recall the intermediate quantity, misclustering loss, defined in~\eqref{eq:defnofL}
+\begin{equation}
+     L^{(t)}:=L(z,z^{(t)})= \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onorm{ \off{ \mS_{ z(i):}  }^s - \off{ \mS_{b:}  }^s  }^2,
 \end{equation}
@@
 \begin{equation}\label{eq:proof_5_xi}
-    \xi  \lesssim \exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}.
+    \xi  \lesssim {\text{SNR}^{-1}}\exp\of{- \frac{p^{K-1}\text{SNR}}{r^{K-1}}}.
 \end{equation}
@@
 
+{\color{blue}
+The proof of MLE error shares the similar idea as Theorems~\ref{thm:initial}-\ref{thm:refinement}.
+we first show a weaker polynomial rate for MLE, and then improve the rate from polynomial to exponential throughout the iterations.
+The only difference is that the MLE remains the same over iterations due to its global optimality.
+See Appendix II, Section~\ref{sec:statprove2} for the detailed proof.}
+
 \section*{Acknowledgments}
-This research is supported in part by NSF grants DMS-1915978, DMS-2023239, EF-2133740, and funding from the Wisconsin Alumni Research foundation.
+This research is supported in part by NSF CAREER DMS-2141865, DMS-1915978, DMS-2023239, EF-2133740, and funding from the Wisconsin Alumni Research foundation.
 We thank Zheng Tracy Ke, Rungang Han, Yuetian Luo for helpful discussions and for sharing software packages.
@@
 
-We provide the proofs for all the theorems in our main paper. In each sub-section, we first show the proof of main theorem and then collect the useful lemmas in the end.
-
-\section*{Notation}
-Before the proofs, we first introduce the notation used throughout the appendix and the generalized dTBM without symmetric assumptions.
-The parameter space and minimal gap assumption are also extended for the generalized dTBM.
+{\color{blue}
+\section*{Appendix I. Additional numerical experiments}
+\textbf{Bernoulli phase transition.} The first additional experiment verifies the statistical-computational gap in Section~\ref{sec:limits} under the Bernoulli model.
+Consider the Bernoulli model with $p = \{80, 100\}$, $r = 5$.
+We vary $\gamma $ in $ [-1.2, -0.4]$ and $[-2.1, -1.4]$ for matrix ($K=2$) and tensor $(K = 3)$ clustering, respectively.
+We  approximate MLE using an oracle estimator, i.e., the output of Sub-algorithm~\hyperref[alg:main]{2} initialized from the true assignment.
+Figure~\ref{fig:phase_binary} shows a similar pattern as Figure~\ref{fig:phase}.
+The algorithm and oracle estimators have no gap in the matrix case, while an error gap emerges between the critical values $\gamma_{\text{stat}} = -2$ and $\gamma_{\text{comp}} = -1.5$ in the tensor case.
+Figure~\ref{fig:phase_binary} suggests the statistical-computational gap in Bernoulli models.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width = .8\columnwidth]{binary_phase.pdf}
+    \caption{SNR phase transitions for Bernoulli dTBM with $p = \{80, 100\}, r = 5$ under (a) matrix case with $\gamma \in [-1.2, -0.4]$ and (b) tensor case with $ \gamma \in [-2.1, -1.4]$.}
+    \label{fig:phase_binary}
+\end{figure}
+
+{\bf Sparsity.} The second additional experiment evaluates the algorithm performances under the sparse binary dTBM~\eqref{eq:sparse_dtbm}.
+We fix the signal exponent $\gamma = -1.2$ and vary the sparsity parameter $\alpha_p \in [0.05, 0.9]$.
+A smaller $\alpha_p$ leads to a higher probability of zero entries in the observation.
+In addition to the three algorithms mentioned in Section~\ref{subsec:comp} (denoted {\bf \small Initialization}, {\bf \small dTBM}, and {\bf \small SCORE}), we consider other three algorithms based on the discussion in Section~\ref{subsec:ber}:
+  \begin{itemize} 
+  \item \textbf{\small D-HOSVD}, the diagonal-deleted HOSVD in \cite{ke2019community}; 
+  \item \textbf{\small D-HOSVD + Angle}, the combined algorithm of our angle-based iteration with initialization from \textbf{\small D-HOSVD};
+  \item \textbf{\small SCORE + Angle}, the combined algorithms of our angle-based iteration with initialization from \textbf{\small SCORE}.
+  \end{itemize}
+
+\begin{figure}[htp!]
+    \centering
+    \includegraphics[width=.8\columnwidth]{sparsity.pdf}
+    \caption{CER comparison versus sparsity parameter $\alpha_p$ in $[0.05, 0.9]$. We set $p = 100, r = 5$ and $\gamma = -1.2$ under sparse binary dTBM.}
+    \label{fig:sparse}
+\end{figure}
+
+Figure~\ref{fig:sparse} shows a slightly larger error in \textbf{\small dTBM} than that in \textbf{\small SCORE}, \textbf{\small D-HOSVD + Angle}, and \textbf{\small SCORE + Angle} under the sparse setting with $\alpha_p < 0.3$.
+The small gap between  \textbf{\small dTBM} and other sparse-specific methods implies the robustness of our algorithm.
+In addition, comparing \textbf{\small SCORE} versus \textbf{\small SCORE + Angle} (or \textbf{\small D-HOSVD} versus \textbf{\small D-HOSVD + Angle}) indicates the benefit of our angle iterations under the sparse dTBM.
+In the intermediate and dense cases with $\alpha_p \geq 0.3$, our proposed \textbf{\small dTBM} has a clear improvement over others, which again verifies the success of our algorithm in dense settings.
+
+}
+
+\section*{Appendix II. Proofs}
+
+We provide the proofs for all the theorems in our main paper.
+In each sub-section, we first show the proof of main theorem and then collect the useful lemmas in the end.
+\textcolor{blue}{We combine the proofs of MLE achievement in Theorem~\ref{thm:stats} and polynomial-time achievement in Theorem~\ref{thm:refinement} in the last section due to the similar idea.
+}
+
+\subsection{Notation}
+Before the proofs, we first introduce the notation used throughout the appendix and the general dTBM without symmetric assumptions.
+The parameter space and minimal gap assumption are also extended for the general asymetric dTBM.
 
@@
 
-{\bf Model extension to generalized dTBM.}
-
- The general order-$K$ $(p_1, \ldots, p_K)$-dimensional dTBM model with $r_k$ communities and degree heterogeneity $\mtheta_k = \entry{\theta_k(i)} \in \bbR_+^{p_k}$ is represented by
+{\bf Extension to general asymetric dTBM.}
+
+ The general order-$K$ $(p_1, \ldots, p_K)$-dimensional dTBM with $r_k$ communities and degree heterogeneity $\mtheta_k = \entry{\theta_k(i)} \in \bbR_+^{p_k}$ is represented by
 \begin{equation}\label{eq:general_dtbm}
@@
     \item[$(\Leftarrow)$] For the necessity, it suffices to construct two distinct parameters up to cluster label permutation, if the model~\eqref{eq:general_dtbm} violates Assumption~\ref{assmp:general_minimal_gap}.
-Without loss of generality, we assume $\onorm{ \mS_{1,1:}^s - \mS_{1,2:}^s } = 0$.
-
- If $\mS_{1,1:}$ is a zero vector, construct $\mtheta_1'$ such that $\mtheta'_{1,z_1^{-1}(1)} \neq \mtheta_{1,z_1^{-1}(1)}$.
+\textcolor{red}{Note that $\Delta_{\min}^2 = 1$ when there exists a $k \in [K]$ such that $r_k = 1$.
+Hence, we consider the case that  $r_k \geq 2$ for all $k \in [K]$.} Without loss of generality, we assume $\onorm{ \mS_{1,1:}^s - \mS_{1,2:}^s } = 0$.
+
+% If $\mS_{1,1:}$ is a zero vector, construct $\mtheta_1'$ such that $\mtheta'_{1,z_1^{-1}(1)} \neq \mtheta_{1,z_1^{-1}(1)}$.
 Let $\{z'_k\} = \{z_k\}$, $\tS' = \tS$, and $\mtheta'_k = \mtheta_k$ for all $k = 2, \ldots, K$.
@@
 Similar conclusion holds when $\mS_{1,2:}$ is a zero vector.
-
-If neither $\mS_{1,1:}$ nor $\mS_{1,2:}$ is a zero vector, there exists a positive constant $c$ such that $\mS_{1,1:} = c \mS_{1,2:}$.
-Thus, there exists a core tensor $\tS_0 \in \bbR^{r_1 -1 \times \cdots \times r_K}$ such that
+{
+\color{blue}
+By constraints in parameter space~\eqref{eq:general_family}, neither $\mS_{1,1:}$ nor $\mS_{1,2:}$ is a zero vector.
+}
+There exists a positive constant $c$ such that $\mS_{1,1:} = c \mS_{1,2:}$. Thus, there exists a core tensor $\tS_0 \in \bbR^{r_1 -1 \times \cdots \times r_K}$ such that 
 \begin{equation}
@@
 First, we show the uniqueness of $\mM_k$ for all $k \in [K]$.
-Without loss of generality, we consider $k=1$ and show the first mode membership matrix; i.e., $\mM'_1 = \mM_1\mP_1$ where $\mP_1$ is a permutation matrix.
+\textcolor{red}{When $r_k = 1$, all possible $\mM_k$ is equal to the vector ${\bf 1}_{p_k}$, and the uniqueness holds trivially.
+Hence, we consider the case that $r_k \geq 2$.
+Without loss of generality, we consider $k=1$ with $r_1 \geq 2$ and show the uniqueness of the first mode membership matrix;}  i.e., $\mM'_1 = \mM_1\mP_1$ where $\mP_1$ is a permutation matrix.
 The conclusion for $k\geq 2$ can be showed similarly and thus omitted.
@@
 
-\begin{lem}[Motivation of angle-based clustering]\label{lem:angle} Consider the signal tensor $\tX$ in the generalized dTBM~\eqref{eq:general_dtbm} with $(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\{r_k\})$ and $r_k \geq 2$.
+\begin{lem}[Motivation of angle-based clustering]\label{lem:angle} Consider the signal tensor $\tX$ in the general asymmetric dTBM~\eqref{eq:general_dtbm} with $(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\{r_k\})$, $r_k \geq 2$ \textcolor{red}{ and $K \geq 2$}.
 Then, for any $k \in [K]$ and index pair $(i,j)\in[p_k]^2$, we have
@@
 
-\section*{Proof of Theorem~\ref{thm:stats}}
-
-\begin{proof}[Proof of Theorem~\ref{thm:stats}]
+{
+\color{blue}
+
+\subsection{Proof of Lemma~\ref{lem:angle_gap_x} and Lemma~\ref{lem:mis}}
+
+\begin{proof}[Proof of Lemma~\ref{lem:angle_gap_x}]
+
+Note that the vector $\mS_{z(i):}$ can be folded to a tensor $\tS' =\entry{\tS'_{a_2,\ldots,a_K}} \in \bbR^{r^{K-1}}$; i.e., $\text{vec}(\tS') = \mS_{z(i):}$.
+Define weight vectors $\mw_{a_2, \cdots, a_K}$ correspond to the elements in $\tS'_{a_2,\ldots,a_K}$ by
+\begin{equation}
+    \mw_{a_2 \cdots a_K} = [ \mtheta_{z^{-1} (a_2)}^T \otimes \cdots \otimes \mtheta_{z^{-1} (a_K)}^T] \in \bbR^{|z^{-1} (a_2)| \times \cdots \times |z^{-1} (a_K)|},
+\end{equation}
+for all $a_k \in [r], k = 2,\ldots, K$, where $\otimes$ denotes the Kronecker product.
+Therefore, we have  $\mX_{i:} = \theta(i) \pad_{\mw}(\mS_{z(i):})$ where $\mw = \{ \mw_{a_2, \cdots, a_K}\}_{a_k \in [r], k\in [K]/\{1\}}$.
+Specifically, we have $\onormSize{}{\mw_{a_2, \ldots, a_K}}^2 = \prod_{k = 2}^K \onormSize{}{ \mtheta_{z^{-1}(a_k)}}^2$, and by the balanced assumption~\eqref{eq:degree} we have
+\begin{equation}\label{eq:pad_balance}
+    \max_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2 = (1 + o(1))  \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2.
+\end{equation}
+
+Consider the inner product of $\mX_{i:}$ and $\mX_{j:}$ for $z(i) \neq z(j)$.
+By the definition of weighted padding operator~\eqref{eq:paddef} and the balanced assumption~\eqref{eq:pad_balance}, we have
+\begin{align}
+    \ang{\mX_{i:},\mX_{j:}} &= \theta(i) \theta(j) \ang{ \pad_{\mw}(\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):} )}\\
+    &= \theta(i) \theta(j) \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2 \ang{\mS_{z(i):}, \mS_{z(j):}}(1 + o(1)).
+\end{align}
+Therefore, when $p$ large enough, the inner product $\ang{\mX_{i:},\mX_{j:}} $ has the same sign as $\ang{\mS_{z(i):}, \mS_{z(j):}}$.
+
+{
+\color{blue}
+
+{
+\color{red} Then, we have 
+\begin{equation}
+    \cos (\mS_{z_1(i):}, \mS_{z_1(j):}) = \frac{ \ang{  \mS_{z_1(i):}, \mS_{z_1(j):} }}{ \onormSize{}{\mS_{z_1(i):}} \onormSize{}{\mS_{z_1(j):}} }  = (1+o(1))\frac{ \ang{  \mX_{i:}, \mX_{j:} }}{ \onormSize{}{ \mX_{i:}} \onormSize{}{ \mX_{j:}} } = (1 + o(1)) \cos(\mX_{i:}, \mX_{j:}),
+\end{equation}
+where the second inequality follows by the balance assumption on $\mtheta$, and the last inequality holds when $p$ is large enough.
+
+Further, notice that $\onormSize{}{\mv_1^s - \mv_2^s}^2 =  2(1 - \cos(\mv_1, \mv_2))$. For all $i,j$ such that $z(i) \neq z(j)$, when $p \rightarrow \infty$, we have
+\begin{equation}
+    \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \asymp \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
+\end{equation}
+
+}
+
+% Notice that $\onormSize{}{\mv_1^s - \mv_2^s}^2 =  2(1 - \cos(\mv_1, \mv_2))$ for two vectors $\mv_1, \mv_2$ of the same dimension. We have 
+% \begin{align}
+%     \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s}  &= 2 - 2 \frac{ \ang{  \mS_{z_1(i):}, \mS_{z_1(j):} }}{ \onormSize{}{\mS_{z_1(i):}} \onormSize{}{\mS_{z_1(j):}} } \\
+%     & = 2 - 2 (1+o(1))\frac{ \ang{  \mX_{i:}, \mX_{j:} }}{ \onormSize{}{ \mX_{i:}} \onormSize{}{ \mX_{j:}} } \\
+%     & = (1 + o(1)) \onormSize{}{\mX_{i:}^s - \mX_{j:}^s},
+% \end{align}
+% where the second inequality follows by the balance assumption on $\mtheta$, and the last inequality holds when $p$ is large enough.
+
+% Hence, we conclude that for all $i,j$ such that $z(i) \neq z(j)$,
+% \begin{equation}
+%     \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \asymp \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
+% \end{equation}
+
+}
+
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:mis}]
+By the definition of minimal gap in Assumption~\ref{assmp:min_gap}, we have 
+\begin{equation}
+         L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \geq \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \Delta_{\min}^2 \geq c \ell^{(t)} \Delta_{\min}^2,
+    \end{equation}
+    where the last inequality follows from the assumption $\min_{i \in [p]} \theta(i) \geq c>0$.
+\end{proof}
+
+}
+
+\subsection{Proof of Theorem~\ref{thm:stats} (Impossibility)}\label{sec:statprove1}
+
+\begin{proof}[Proof of Theorem~\ref{thm:stats} (Impossibility)]
 We will prove a more general conclusion than the main paper by allowing growing $r_k$'s. 
-Consider the generalized dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$.
-Specifically, we will show that, under
-the assumptions $K\geq 1$, $r\lesssim p^{1/3}$ and SNR condition
+%We firstly show the proof when the parameter satisfies condition (b). 
+Consider the general asymetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$. Specifically, we will show that, under 
+the assumptions $K\geq 1$, \textcolor{red}{$2 \leq r\lesssim p^{1/3}$ as 
+$p \rightarrow \infty$}, and SNR condition
 \[
@@
 Since the inequality~\eqref{eq:max} is a minimax lower bound, it suffices to show the inequality holds for a particular $(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)$. 
-Specifically, we consider the estimation problem based on a particular parameter point $(\{z_k\}, \tS, \{\mtheta_k\})$ with the following three properties:
+% Specifically, we consider the estimation problem based on a particular parameter point $(\{z_k\}, \tS, \{\mtheta_k\})$ with the following three properties:
+% \begin{equation}\label{eq:construction}
+%   \text{(i)}\  \theta_k(i)=1 \text{ for all }i\in[p]; \quad  \text{(ii)}\ 
+%     \Delta_{\min}\lesssim  \left({p\over r}\right)^{-\frac{K-1}{2}}\sigma; \quad  \text{(iii)}\ |z^{-1}_k(a)|={p \over r} \in \mathbb{Z}_+ \text{ for all }a\in[r],
+% \end{equation}
+% for all $k \in [K]$.
+
+{
+\color{blue} 
+%\fixme{Jiaxin}{The condition (i) can be replaced by the balance assumption after using the construction of $T$ in \cite{gao2018community}.
+The only difference between $T_{gao}$ and $T_{han}$ is: $T_{gao}$ collects the indices with large degree and ensures the boundedness of $\theta(i)$ for $i \in T^c$ while $T_{han}$ chooses the indices randomly.
+Modification is below.}
+
+Specifically, we consider the estimation problem based on a particular parameter point $(\{z_k\}, \tS, \{\mtheta_k\})$ with the following two properties:
 \begin{equation}\label{eq:construction}
-    \text{(i)}\  \theta_k(i)=1 \text{ for all }i\in[p]; \quad  \text{(ii)}\ 
+\text{(i)}\  {\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}; \quad  \text{(ii)}\ 
     \Delta_{\min}\lesssim  \left({p\over r}\right)^{-\frac{K-1}{2}}\sigma; \quad  \text{(iii)}\ |z^{-1}_k(a)|={p \over r} \in \mathbb{Z}_+ \text{ for all }a\in[r],
@@
 for all $k \in [K]$.
+
+%Furthermore, we define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation.....
+
+}
+
 Furthermore, we define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation.
@@
 \end{equation}
+% {
+% \color{blue} 
+% \begin{equation}
+%     \text{(iv)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \quad \text{(v)} \ |T^c_k|\asymp {p\over r}; \quad \text{(vi)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
+% \end{equation}
+
+% }
+
 \normalsize
 for all $k \in [K]$.
-The construction of $T$ is precisely the same as \citet[Proof of Theorem 6]{han2020exact}. 
-Then, following the proof of~\citet[Theorem 2]{gao2018community}, for all $k \in [K]$, we have
+
+{
+\color{blue}
+ The construction of $T$ is precisely the same as \citet[Proof of Theorem 2]{gao2018community}.
+By the constraint $\onormSize{}{\mtheta_{z^{-1}(a)}}_1 = |z^{-1}(a)|, a \in [r]$  in parameter space~\eqref{eq:family}, the $\max_{i \in T^c} \theta(i) \leq C$ for some positive constant $C$.
+Then, following the proof of~\citet[Theorem 2]{gao2018community},
+}
+% The construction of $T$ is precisely the same as \citet[Proof of Theorem 6]{han2020exact}. 
+% Then, following the proof of~\citet[Theorem 2]{gao2018community},
+for all $k \in [K]$, we have
 \begin{equation}\label{eq:minimax}
@@
 By construction~\eqref{eq:construction} of $(\{z_k\}, \tS, \{\mtheta_k\})$ with three properties and lower bound $\min_{a\in[r]}\onorm{\mS_{a:}}\geq c_3$ in the definition of $\tP(\gamma)$, we have $\theta^*=1$, $\onorm{\mx_2}\geq \onorm{\mS_{2:}}\min_{a\in[r]}\onorm{\mtheta_{\mz^{-1}(a)}}^{K-1} \gtrsim \left(p \over r\right)^{(K-1)/2}$.
-Also, note that under the construction~\eqref{eq:construction}
-\begin{equation}
-    \cos(\mx_1, \mx_2) = \frac{\and{\mx_1, \mx_2}}{\onorm{\mx_1} \onorm{\mx_2} } = \frac{(p/r)^{K-1} \ang{\mS_{1:}, \mS_{2:}}}{ \sqrt{(p/r)^{K-1} \onorm{\mS_{1:}}^2} \sqrt{(p/r)^{K-1} \onorm{\mS_{2:}}^2} } = \cos(\mS_{1:}, \mS_{2:}),
-\end{equation}
-which implies $\onormSize{}{\mx^s_1-\mx^s_2} = \onormSize{}{\mS_{1:}^s-\mS_{2:}^s} = \Delta_{\min} \leq 1$. Therefore, the equation~\eqref{eq:bound} is lower bounded by
+% Also, note that under the construction~\eqref{eq:construction}
+% \begin{equation}
+%     \cos(\mx_1, \mx_2) = \frac{\and{\mx_1, \mx_2}}{\onorm{\mx_1} \onorm{\mx_2} } = \frac{(p/r)^{K-1} \ang{\mS_{1:}, \mS_{2:}}}{ \sqrt{(p/r)^{K-1} \onorm{\mS_{1:}}^2} \sqrt{(p/r)^{K-1} \onorm{\mS_{2:}}^2} } = \cos(\mS_{1:}, \mS_{2:}),
+% \end{equation}
+% which implies $\onormSize{}{\mx^s_1-\mx^s_2} = \onormSize{}{\mS_{1:}^s-\mS_{2:}^s} = \Delta_{\min} \leq 1$. 
+\textcolor{blue}{Also, by the balance assumption (i), we have $\onormSize{}{\mx^s_1-\mx^s_2} = (1+o(1)) \onormSize{}{\mS_{1:}^s-\mS_{2:}^s} = \Delta_{\min} \leq 1$}.
+
+Therefore, the equation~\eqref{eq:bound} is lower bounded by
 \begin{equation}\label{eq:normal}
@@
 
+{
+\color{red}
+
+The same conclusion holds after replacing the condition (i) and (ii) by  $\Delta_{\mX}^2 \lesssim p^{-(K-1)}$ and constructing the sets $T$ as \citet[Proof of Theorem 6]{han2020exact}.
+
+}
 \end{proof}
 
-\section*{Proof of Theorem~\ref{thm:comp}}
-
-\begin{proof}[Proof of Theorem~\ref{thm:comp}]
-The idea of proving computational hardness is to show the computational lower bound for a special class of degree-corrected tensor clustering model with $K\geq 2$.
+\subsection{Proof of Theorem~\ref{thm:comp} (Impossibility)}\label{sec:compprove1}
+
+\begin{proof}[Proof of Theorem~\ref{thm:comp} (Impossibility)]
+The idea of proving computational hardness is to show the computational lower bound for a special class of degree-corrected tensor clustering model with $K\geq 2$ \textcolor{red}{and $ r \geq 2$}.
 We construct the following special class of higher-order degree-corrected tensor clustering  model.
@@
 
-\section*{Proof of Theorem~\ref{thm:initial}}
-\begin{proof}[Proof of Theorem 4] We prove Theorem~\ref{thm:initial} under the symmetric dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
-We drop the subscript $k$ in the matricizations $\mM_k, \mX_k, \mS_k$.
-For simplicity, let $\hat z$ denote the output, $\hat z^{(0)}$, of Sub-algorithm~\hyperref[alg:main]{1}.
+\subsection{Proof of Theorem~\ref{thm:initial}}
+\begin{proof}[Proof of Theorem~\ref{thm:initial}] We prove Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with symmetric mean tensor, parameters $(z, \tS, \mtheta)$, \textcolor{red}{fixed $r\geq 1, K \geq 2$}, and i.i.d.\ noise.
+\textcolor{red}{For the case $r = 1$, we have $L(z^{(0)},z) = 0, \ell(z^{(0)}, z) = 0$ trivially.
+Hence, we focus on the proof of the first mode clustering $z^{(0)}_1$ with $r \geq 2$}; the proofs for the other modes can be extended similarly.
+We drop the subscript $k$ in the matricizations $\mM_k, \mX_k, \mS_k$ and in the estimate $z^{(0)}_1$.
+We firstly show the proof under the balance $\mtheta$.
+%For simplicity, let $\hat z$ denote the output, $z^{(0)}$, of Sub-algorithm~\hyperref[alg:main]{1}.
+
+\textbf{We firstly show the upper bound for misclustering error $\ell(z^{(0)}, z)$.}
+%and then prove the upper bound for the loss $L^{z^{(0)}, z}$
 
@@
  \begin{equation}\label{eq:theta_bound}
-          \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2  \leq  \sum_{i \in S_I} \theta(i)^2 + 4 \sum_{i \in S} \theta(i)^2 ,
+          \min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2  \leq  \sum_{i \in S_I} \theta(i)^2 + 4 \sum_{i \in S} \theta(i)^2 ,
     \end{equation}
@@
     \begin{equation}
-        S_0 = \{i: \onormSize{}{\hat \mX_{i:}} = 0\}, \quad S = \{i \in S_0^c: \onormSize{}{\hat \mx_{\hat z(i)} - \mX_{i:}^s } \geq c_0  \Delta_{\min}/2\}.
+        S_0 = \{i: \onormSize{}{\hat \mX_{i:}} = 0\}, \quad S = \{i \in S_0^c: \onormSize{}{\hat \mx_{z^{(0)}(i)} - \mX_{i:}^s } \geq c_0  \Delta_{\min}/2\}.
     \end{equation}
@@
          \sum_{i \in S} \onormSize{}{\mX_{i:}}^2 & \leq 2 \sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  +  2\sum_{i \in S} \onormSize{}{\hat \mX_{i:} - \mX_{i:}}^2 \label{eq:tr1}\\
-         & \leq \frac{8}{c_0^2 \Delta_{\min}^2  } \sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2   \onormSize{}{\hat \mx_{\hat z(i)} - \mX_{i:}^s }^2 +2 \onormSize{}{\hat \tX - \tX}_F^2 \label{eq:defs} \\
-         & \leq \frac{16}{c_0^2 \Delta_{\min}^2  }\sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  \off{ \onormSize{}{\hat \mx_{\hat z(i)} - \hat \mX_{i:}^s }^2 + \onormSize{}{\hat \mX_{i:}^s - \mX_{i:}^s }^2 } + 2\onormSize{}{\hat \tX - \tX}_F^2 \label{eq:tr2} \\
+         & \leq \frac{8}{c_0^2 \Delta_{\min}^2  } \sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2   \onormSize{}{\hat \mx_{z^{(0)}(i)} - \mX_{i:}^s }^2 +2 \onormSize{}{\hat \tX - \tX}_F^2 \label{eq:defs} \\
+         & \leq \frac{16}{c_0^2 \Delta_{\min}^2  }\sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  \off{ \onormSize{}{\hat \mx_{z^{(0)}(i)} - \hat \mX_{i:}^s }^2 + \onormSize{}{\hat \mX_{i:}^s - \mX_{i:}^s }^2 } + 2\onormSize{}{\hat \tX - \tX}_F^2 \label{eq:tr2} \\
          & \leq  \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 }\sum_{i \in S} \onormSize{}{\hat \mX_{i:}}^2  \onormSize{}{\hat \mX_{i:}^s - \mX_{i:}^s }^2  + 2\onormSize{}{\hat \tX - \tX}_F^2 \label{eq:km} \\
@@
     \begin{align}
-        \min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2  &\lesssim \of{\sum_{i \in S} \onormSize{}{\mX_{i:}}^2 + \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2  } p^{-(K-1)} r^{K-1} \\
+        \min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2  &\lesssim \of{\sum_{i \in S} \onormSize{}{\mX_{i:}}^2 + \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2  } p^{-(K-1)} r^{K-1} \\
         & \lesssim \frac{\sigma^2 r^{K-1}}{ \Delta_{\min}^2  p^{K-1} } \of{ p^{K/2}r + pr^2 + r^K }.
@@
     \begin{equation}
-        \ell(z, z) \lesssim \frac{1}{p}\min_{\pi \in \Pi} \sum_{i : \hat z(i) \neq \pi(z(i))} \theta(i)^2 \lesssim \frac{r^K p^{-K/2} }{ \text{SNR} },
+        \ell(z^{(0)}, z) \lesssim \frac{1}{p}\min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2 \lesssim \frac{r^K p^{-K/2} }{ \text{SNR} },
     \end{equation}
     where the last inequality follows from the definition $\text{SNR} = \Delta_{\min}^2/\sigma^2$.
-\end{proof}
-
-{\bf Useful Corollary of Theorem~\ref{thm:initial}}
-
-\begin{cor}[Initial misclustering loss]\label{cor:L0} 
-Under the setup of Theorem~\ref{thm:initial}, assume SNR $\geq \tilde  C p^{-K/2} \log p$ for some positive universal constant $\tilde C$ and $\sigma^2 = 1$.
-With probability going to 1, the misclustering loss for the initialization is upper bounded as
-\begin{equation}
-    L^{(0)} =  \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(0)}(i) = b } \onormSize{}{ [ \mS_{ \pi^{(0)}( z(i)):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq \frac{\bar C \Delta_{\min}^2}{ \tilde C r \log p},
-\end{equation}
-where $\pi^{(0)}$ minimizes the initial misclustering error; i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} $ $ \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$, and $\bar C > 1$ is a positive universal constant.
-\end{cor}
-
-\begin{proof} [Proof of Corollary~\ref{cor:L0}]
-Without loss of generality, we assume $\pi^{(0)}$ is the identity mapping such that  $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
-Note that $\mX_{i:}^s$ have only $r$ different values.
-We let $\mX_a^s = \mX_{i:}^s$ for all $i$ such that $z(i) = a, a \in [r]$.
-
+
+    \textcolor{blue}{Without the balanced $\mtheta$, we have $\min_{z(i) \neq z(j)} \onorm{ \mX^s_{i:} - \mX^s_{j:} } \geq c_0  \Delta_{\mX}$.
+Replacing the definition of $S$ with $\Delta_{\mX}$, we obtain the desired result.}
+
+    \textbf{Next, we show the bound for $L(z^{(0)}, z).$}
+
+    Note that $\mX_{i:}^s$ have only $r$ different values. We let $\mX_a^s = \mX_{i:}^s$ for all $i$ such that $z(i) = a, a \in [r]$. 
+    %We firstly show the results for initialization and then prove the MLE results.
     Notice that 
@@
 where the first inequality follows from Lemma~\ref{lem:angle_gap_x}, the third inequality follows from inequalities~\eqref{eq:cor_3} and \eqref{eq:cor_4}, and the last inequality follows from the assumption that SNR $\geq \tilde C p^{-K/2} \log p$.
+
 \end{proof}
+
+% \begin{cor}[Initial misclustering loss]\label{cor:L0} 
+% Under the setup of Theorem~\ref{thm:initial}, assume SNR $\geq \tilde  C p^{-K/2} \log p$ for some positive universal constant $\tilde C$ and $\sigma^2 = 1$.
+With probability going to 1, the misclustering loss for the initialization is upper bounded as
+% \begin{equation}
+%     L^{(0)} =  \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(0)}(i) = b } \onormSize{}{ [ \mS_{ \pi^{(0)}( z(i)):}  ]^s - [ \mS_{b:}  ]^s  }^2 \leq \frac{\bar C \Delta_{\min}^2}{ \tilde C r \log p},
+% \end{equation}
+% where $\pi^{(0)}$ minimizes the initial misclustering error; i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} $ $ \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$, and $\bar C > 1$ is a positive universal constant.
+
+% {
+% \color{blue}
+
+% }
+
+% \end{cor}
+
+% \begin{proof} [Proof of Corollary~\ref{cor:L0}]
+% Without loss of generality, we assume $\pi^{(0)}$ is the identity mapping such that  $\pi^{(0)}(a) = a$ for all $ a \in [r]$, and so for $\hat z$.
+
+% {
+% \color{blue}
+
+% \textbf{Now, we show the results for MLE.}
+
+% We replace the vector $\hat x_{\hat z(i)}$ by our MLE $\hat \mW_{:\hat z(i) }\hat \mX$ in the proof of Theorem 4. Then, we have 
+% \begin{align}
+%     \frac{1}{p}\sum_{i \in [p], \hat z(i) \neq z(i)} (\theta^{(i)})^2 \lesssim \frac{r^{K-1}}{\Delta_{\min}^2 p^{K}} \onormSize{}{\tX - \hat \tX}_F^2  \lesssim \frac{r^{K-2}}{p^{K-1} \Delta_{\min}^2}.
+% \end{align}
+% By the assumption that $\Delta_{\min}^2 \gtrsim \frac{p^{(K-1)}}{r^{K-1}} \log p$ and  $\min_{i \in [p]}\theta(i) \geq c$, we have 
+% \begin{equation}
+%   \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\theta(i))^2 \lesssim \frac{1}{r \log p}, \quad  \ell(\hat z, z)  \lesssim \frac{1}{r \log p}.
+% \end{equation}
+% Note that $(\hat z, \hat \mtheta, \hat \tS) \in \tP(\gamma)$ and the minimal angle gap in $\hat \tS$ also larger than $\Delta_{\min}^2$.
+By Lemma~\ref{lem:angle_gap_x}, we have $\min_{a \neq b \in [r]} \onormSize{}{ [\hat \mW_{:a}^T \hat \mX]^s -  [\hat \mW_{:b}^T \hat \mX]^s } \gtrsim \Delta_{\min}$.
+Then, we switch the parameters $\mX$ with $\hat \mX$ and switch $\mtheta$ with $\hat \mtheta$ in the proof of Theorem~\ref{thm:initial}.
+We obtain
+% \begin{equation}
+%     \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\hat \theta(i))^2 \lesssim \frac{1}{r \log p}.
+% \end{equation}
+
+% Next, we plug the estimator $\hat z$ into the proof of Corollary~\ref{cor:L0} with $\onormSize{}{\tX - \hat \tX}^2_F \lesssim pr$ and $\ell(\hat z, z) \lesssim \frac{1}{\log p}$. We obtain that 
+% \begin{align}
+%     L(\hat z) \lesssim \frac{r^K}{p^{K}} \onormSize{}{\tX - \hat \tX}^2_F  \lesssim \frac{r \Delta_{\min}^2}{\log p}.
+% \end{align}
+
+% }
+
+% \end{proof}
 
@@
 
-\begin{lem}[Angle gap in $\tX$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family}.
+% \begin{cor}[Angle for vector-matrix production]\label{cor:angle_matrix} Suppose we have two non-zero vectors $\a,\mb \in \bbR^d$ and a matrix $\mM \in \bbR^{m \times d}$. We have 
+% \begin{equation}
+%   \frac{|\egien_{\min}( \mM )| }{|\egien_{\max}( \mM ) | } \sin(\ma, \mb) \leq \sin( \mM \ma, \mM \mb ) \leq \frac{|\egien_{\max}( \mM )| }{ \onormSize{}{\mM }  } \sin(\ma, \mb).
+% \end{equation}
+% \end{cor}
+
+% {
+% \color{blue}
+
+% \begin{lem}[Angle gap in $\tX$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family}.
+Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying $\max_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2 \leq C \min_{a \in [r]} \onormSize{}{\mtheta_{z^{-1}(a)}}^2$ for some positive constant $C$.
+Then the angle gap in $\tX$ is asymptotically lower bounded by the angle gap in $\tS$; i.e., for all $i,j$ such that $z(i) \neq z(j)$
+
+% \begin{equation}
+%     \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \asymp  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
+% \end{equation}
+
+% \end{lem}
+
+% \begin{proof}[Proof of Lemma~\ref{lem:angle_gap_x}] 
+% Note that the vector $\mS_{z(i):}$ can be folded to a tensor $\tS' =\entry{\tS'_{a_2,\ldots,a_K}} \in \bbR^{r^{K-1}}$; i.e., $\text{vec}(\tS') = \mS_{z(i):}$.
+Define weight vectors $\mw_{a_2, \cdots, a_K}$ correspond to the elements in $\tS'_{a_2,\ldots,a_K}$ by
+% \begin{equation}
+%     \mw_{a_2 \cdots a_K} = [ \mtheta_{z^{-1} (a_2)}^T \otimes \cdots \otimes \mtheta_{z^{-1} (a_K)}^T] \in \bbR^{|z^{-1} (a_2)| \times \cdots \times |z^{-1} (a_K)|},
+% \end{equation}
+% for all $a_k \in [r], k = 2,\ldots, K$, where $\otimes$ denotes the Kronecker product.
+Therefore, we have  $\bar \mX_{i:} \coloneqq \theta(i)^{-1} \mX_{i:} =  \pad_{\mw}(\mS_{z(i):})$ where $\mw = \{ \mw_{a_2, \cdots, a_K}\}_{a_k \in [r], k\in [K]/\{1\}}$ and $\bar \mX_{i:} - \bar \mX_{i:} = \pad_{\mw}(\mS_{z(i):} - \mS_{z(j):})$.
+Specifically, we have $\onormSize{}{\mw_{a_2, \ldots, a_K}}^2 = \prod_{k = 2}^K \onormSize{}{ \mtheta_{z^{-1}(a_k)}}^2$, and .
+
+% By Lemma~\ref{lem:norm_diff}, we have
+% \begin{align}
+%     \onormSize{}{\bar \mX_{i:}^s - \bar \mX_{j:}^s} \leq 2\frac{ \onormSize{}{\bar \mX_{i:} - \bar \mX_{j:}}}{ \onormSize{}{\bar \mX_{i:}} } \leq 2\frac{\max_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}} }{ \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}  } \frac{\onormSize{}{ \mS_{z(i):} - \mS_{z(j):} }}{ \onormSize{}{\mS_{z(i):}}} \leq \frac{2}{C}\frac{\onormSize{}{ \mS_{z(i):} - \mS_{z(j):} }}{ \onormSize{}{\mS_{z(i):}}}
+% \end{align}
+% and 
+% \begin{align}
+%       \onormSize{}{\bar \mX_{i:}^s - \bar \mX_{j:}^s}  \geq \sin (\bar \mX_{i:},  \bar \mX_{j:} ) \geq \frac{\min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}} }{ \max_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}  } \sin ( \mS_{z(i):}, \mS_{z(j):} ) \geq C  \sin ( \mS_{z(i):}, \mS_{z(j):} ).
+% \end{align}
+
+% Hence, we conclude that for all $i,j$ such that $z(i) \neq z(j)$,
+% \begin{equation}
+%     \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \gtrsim  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
+% \end{equation}
+% \end{proof}
+
+% }
+
+% \begin{lem}[Angle gap in $\tX$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family}.
 Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying~\eqref{eq:degree}.
 Then the angle gap in $\tX$ is asymptotically lower bounded by the angle gap in $\tS$; i.e., for all $i,j$ such that $z(i) \neq z(j)$
-\begin{equation}
-    \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \gtrsim  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
-\end{equation}
-
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:angle_gap_x}] 
-Note that the vector $\mS_{z(i):}$ can be folded to a tensor $\tS' =\entry{\tS'_{a_2,\ldots,a_K}} \in \bbR^{r^{K-1}}$; i.e., $\text{vec}(\tS') = \mS_{z(i):}$.
-Define weight vectors $\mw_{a_2, \cdots, a_K}$ correspond to the elements in $\tS'_{a_2,\ldots,a_K}$ by
-\begin{equation}
-    \mw_{a_2 \cdots a_K} = [ \mtheta_{z^{-1} (a_2)}^T \otimes \cdots \otimes \mtheta_{z^{-1} (a_K)}^T] \in \bbR^{|z^{-1} (a_2)| \times \cdots \times |z^{-1} (a_K)|},
-\end{equation}
-for all $a_k \in [r], k = 2,\ldots, K$, where $\otimes$ denotes the Kronecker product.
-Therefore, we have  $\mX_{i:} = \theta(i) \pad_{\mw}(\mS_{z(i):})$ where $\mw = \{ \mw_{a_2, \cdots, a_K}\}_{a_k \in [r], k\in [K]/\{1\}}$.
-Specifically, we have $\onormSize{}{\mw_{a_2, \ldots, a_K}}^2 = \prod_{k = 2}^K \onormSize{}{ \mtheta_{z^{-1}(a_k)}}^2$, and by the balanced assumption~\eqref{eq:degree} we have
-\begin{equation}\label{eq:pad_balance}
-    \max_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2 = (1 + o(1))  \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2.
-\end{equation}
-
-Consider the inner product of $\mX_{i:}$ and $\mX_{j:}$ for $z(i) \neq z(j)$.
-By the definition of weighted padding operator~\eqref{eq:paddef} and the balanced assumption~\eqref{eq:pad_balance}, we have
-\begin{align}
-    \ang{\mX_{i:},\mX_{j:}} &= \theta(i) \theta(j) \ang{ \pad_{\mw}(\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):} )}\\
-    &= \theta(i) \theta(j) \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}}^2 \ang{\mS_{z(i):}, \mS_{z(j):}}(1 + o(1)).
-\end{align}
-Therefore, when $p$ large enough, the inner product $\ang{\mX_{i:},\mX_{j:}} $ has the same sign as $\ang{\mS_{z(i):}, \mS_{z(j):}}$.
-Next, we discuss the angle between $\mX_{i:}$ and $\mX_{j:}$ by two cases.
-
-\begin{enumerate}[wide]
-    \item \textbf{Case 1:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} \leq 0$.
+% % \begin{equation}
+% %     \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \gtrsim  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
+% % \end{equation}
+% {
+% \color{blue}
+% \begin{equation}
+%     \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} = (1 + o(1))  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
+% \end{equation}
+% }
+% \end{lem}
+
+% Next, we discuss the angle between $\mX_{i:}$ and $\mX_{j:}$ by two cases.
+
+% \begin{enumerate}[wide]
+%     \item \textbf{Case 1:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} \leq 0$.
 Then, we also have $\ang{\mX_{i:},\mX_{j:}} \leq 0$, which implies $\onormSize{}{\mX_{i:}^s -  \mX_{j:}^s} \geq \sqrt{2}$.
@@
 We have $\onormSize{}{\mX_{i:}^s -  \mX_{j:}^s} \gtrsim \onormSize{}{\mS_{z(i):}^s - \mS_{z(j):}^s}$.
-    \item \textbf{Case 2:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} > 0$.
+%     \item \textbf{Case 2:} Suppose $\ang{\mS_{z(i):}, \mS_{z(j):}} > 0$.
 Then, we have $\cos(\mS_{z(i):}, \mS_{z(j):}) > 0$.
@@
 Then, we have
-    \begin{align}
-        \onormSize{}{\mS_{z(i):}^s - \mS_{z(j):}^s} &= \sqrt{ 1 - \cos(\mS_{z(i):}, \mS_{z(j):})}  \\
-        & \lesssim \sin (\mS_{z(i):}, \mS_{z(j):})\\
-        & \leq \frac{ \max_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}} }{  \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}} }  \sin ( \pad_{\mw} (\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):})) \\
-        & \leq (1 + o(1)) \onormSize{}{\mX_{i:}^s -  \mX_{j:}^s},
-    \end{align}
-    where the second inequality follows from Lemma~\ref{lem:pad}, and the last inequality follows from the balanced weight~\eqref{eq:pad_balance} and Lemma~\ref{lem:norm_diff}.
-\end{enumerate}
-
-Hence, we conclude that for all $i,j$ such that $z(i) \neq z(j)$,
-\begin{equation}
-    \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \gtrsim  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
-\end{equation}
-\end{proof}
-
-\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the parameter space~\eqref{eq:family} and assumption that $\min_{i \in [p]} \theta(i) \geq c$, the singular values of $\mTheta \mM$ are bounded as
+%     \begin{align}
+%         \onormSize{}{\mS_{z(i):}^s - \mS_{z(j):}^s} &= \sqrt{ 1 - \cos(\mS_{z(i):}, \mS_{z(j):})}  \\
+%         & \lesssim \sin (\mS_{z(i):}, \mS_{z(j):})\\
+%         & \leq \frac{ \max_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}} }{  \min_{(a_2,\ldots, a_K)} \onormSize{}{\mw_{a_2,\ldots, a_K}} }  \sin ( \pad_{\mw} (\mS_{z(i):}), \pad_{\mw}(\mS_{z(j):})) \\
+%         & \leq (1 + o(1)) \onormSize{}{\mX_{i:}^s -  \mX_{j:}^s},
+%     \end{align}
+%     where the second inequality follows from Lemma~\ref{lem:pad}, and the last inequality follows from the balanced weight~\eqref{eq:pad_balance} and Lemma~\ref{lem:norm_diff}.
+% \end{enumerate}
+
+% Hence, we conclude that for all $i,j$ such that $z(i) \neq z(j)$,
+% \begin{equation}
+%     \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \gtrsim  \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
+% \end{equation}
+
+\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the parameter space~\eqref{eq:family} and assumption that $\min_{i \in [p]} \theta(i) \geq c$ \textcolor{red}{for some constant $c >0$}, the singular values of $\mTheta \mM$ are bounded as
 \begin{equation}\label{eq:membership2}
@@
 
-\begin{lem}[Upper bound of misclustering error]\label{lem:upper_mis} Let $z: [p] \mapsto [r]$ be a cluster assignment such that $|z^{-1}(a)| \asymp p/r$ for all $a \in [r]$.
+\begin{lem}[Upper bound of misclustering error]\label{lem:upper_mis} Let $z: [p] \mapsto [r]$ be a cluster assignment such that $|z^{-1}(a)| \asymp p/r$ for all $a \in [r]$ \textcolor{red}{with $r \geq 2$}.
 Let node $i$ correspond to a vector $\mx_i  = \theta(i) \mv_{z(i)} \in \bbR^d$, where $\{\mv_a\}_{a = 1}^r$ are the cluster centers and $\mtheta = \entry{\theta(i)} \in \bbR^p_+$ is the positive degree heterogeneity.
@@
 \end{equation}
-where $S_0$ is defined in Step 3 of Sub-algorithm~\hyperref[alg:main]{1} and
+where $S_0$ is defined in \textcolor{red}{Step 4} of Sub-algorithm~\hyperref[alg:main]{1} and
 \begin{equation}
@@
 
-\section*{Proof of Theorem~\ref{thm:refinement}}
-
-\begin{proof}[Proof of Theorem~\ref{thm:refinement}]
-We prove Theorem~\ref{thm:refinement} under the symmetric dTBM~\eqref{eq:model_tensor} with parameters $(z, \tS, \mtheta)$.
-We drop the subscript $k$ in the matricizations $\mM_k, \mS_k, \mX_k$.
-Without loss of generality, we assume that the variance $\sigma = 1$, and that the identity permutation minimizes the initial misclustering error; i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$.
+\subsection{Proofs of Theorem~\ref{thm:stats} (Achievability) and Theorem~\ref{thm:refinement}}\label{sec:statprove2}
+
+\begin{proof}[Proof of Theorem~\ref{thm:stats} (Achievability) and Theorem~\ref{thm:refinement}]
+
+{
+\color{blue}
+
+The proofs of Theorem~\ref{thm:stats} (Achievability) and Theorem~\ref{thm:refinement} share the same idea.
+We prove the contraction step by step.
+In each step, we show the specific procedures for the algorithm loss and address the MLE loss by stating the difference.
+
+%We prove Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with symmetric mean tensor, parameters $(z, \tS, \mtheta)$, \textcolor{red}{fixed $r\geq 1, K \geq 2$}, and i.i.d.\ noise.
+\textcolor{red}{For the case $r = 1$, we have $L(z^{(0)},z) = 0, \ell(z^{(0)}, z) = 0$ trivially.
+Hence, we focus on the proof of the first mode clustering $z^{(0)}_1$ with $r \geq 2$};
+
+We consider dTBM~\eqref{eq:model_tensor} with symmetric mean tensor, parameters $(z, \tS, \mtheta)$, \textcolor{red}{fixed $r\geq 1, K \geq 2$}, and i.i.d.\ noise.
+Let $(\hat z, \hat \mtheta, \hat \tS)$ denote the MLE in \eqref{eq:mle}, and $(z^{(0)}_k, \mtheta^{(0)}, \tS^{(0)})$ denote parameters related to the initialization.
+\textcolor{red}{For the case $r = 1$, we have $\ell(z^{(t)}, z) = 0$ trivially for all $t \geq 0$.
+Hence, we focus on the proof of the first mode clustering $z^{(t+1)}_1$ with $r \geq 2$}; the extension for other modes can be obtained similarly.
+We drop the subscript $k$ in the matricizations $\mTheta, \mM_k, \mS_k, \mX_k$ and in estimates $z^{(0)}, z^{(t+1)}_1, z^{(t)}_k$ for ease of the notation.
+Without loss of generality, we assume that the variance $\sigma = 1$, and that the identity permutation minimizes the initial misclustering error; i.e., $\pi^{(0)} = \argmin_{\pi \in \Pi} \sum_{i \in [p]}\ind \offf{z^{(0)}(i) \neq \pi \circ z(i) }$ and $\pi^{(0)}(a) = a$ for all $ a \in [r]$, and so for $\hat z$.
+}
 
@@
 We will verify that the conditions hold in our context under high probability in the last step of the proof.
+
+{
+\color{blue}
 
@@
 
-    \item We define normalized membership matrices
-    \begin{align}
-        \mW = \mM \of{ \text{diag}(\mone_{p}^T \mM) }^{-1}, \quad \mW^{(t)} = \mM^{(t)} \of{ \text{diag}(\mone_{p}^T \mM^{(t)}) }^{-1},
-    \end{align}
-    and the dual normalized membership matrices
-    \begin{align}
-         \mV &= \mW^{\otimes (K-1)}, \quad \mV^{(t)} =\of{ \mW^{(t)}}^{\otimes (K-1)}.
-    \end{align}
-    \item We use $\tS^{(t)}$ to denote the estimator of $\tS$ in the $t$-th iteration, and we use $\tilde \tS$ to denote the oracle estimator of $\tS$ given true assignment $z$; i.e.,
-    \begin{align}
-        \tS^{(t)} = \tY \times_1 \of{\mW^{(t)}}^T \times_2 \cdots \times_K \of{\mW^{(t)}}^T, \quad \tilde \tS &= \tY \times_1 \mW^T \times_2 \cdots \times_K \mW^T.
+    \item We define normalized and weighted normalized membership matrices
+    \begin{align}
+        &\mW = \mM \of{ \text{diag}(\mone_{p}^T \mM) }^{-1}, \quad& &\mW^{(t)} = \mM^{(t)} \of{ \text{diag}(\mone_{p}^T \mM^{(t)}) }^{-1}, \\
+        &\mP = \mTheta \mM ( \text{diag}( \onormSize{}{\mtheta_{z^{-1}(1)}}^2 , \ldots, \onormSize{}{\mtheta_{z^{-1}(r)}}^2 ) )^{-1}, \quad& & \hat \mP = \hat \mTheta \hat \mM ( \text{diag}( \onormSize{}{\hat \mtheta_{z^{-1}(1)}}^2 , \ldots, \onormSize{}{\hat \mtheta_{z^{-1}(r)}}^2 ) )^{-1},
+    \end{align}
+    and the dual normalized and dual weighted normalized membership matrices
+    \begin{align}
+         \mV = \mW^{\otimes (K-1)}, \quad \mV^{(t)} =\of{ \mW^{(t)}}^{\otimes (K-1)} , \quad \mQ = \mP^{\otimes K-1}, \quad  \hat \mQ = \hat \mP^{\otimes K-1}.
+    \end{align}
+    Also, let $\mB = (\mTheta \mM)^{\otimes (K-1)}, \hat \mB = (\hat \mTheta \hat \mM)^{\otimes (K-1)}$. By the definition, we have $\mB^T \mQ =\hat \mB^T \hat \mQ =\mI_{r^{K-1}}$.
+    \item We use $\tS^{(t)}$ to denote the estimator of $\tS$ in the $t$-th iteration, $\hat \tS$ for MLE, $\tilde \tS$ to denote the oracle estimator of $\tS$ given true assignment $z$, and $\bar \tS$ for weighted oracle estimator; i.e.,
+    \begin{align}
+        &\tS^{(t)} = \tY \times_1 \of{\mW^{(t)}}^T \times_2 \cdots \times_K \of{\mW^{(t)}}^T,& \quad &\tilde \tS = \tY \times_1 \mW^T \times_2 \cdots \times_K \mW^T \\
+        &\hat \tS =  \tY \times_1 \hat \mP^T \times_2 \cdots \times_K \hat \mP^T &  \quad  & \bar \tS  = \tY \times_1 \mP^T \times_2 \cdots \times_K \mP^T.
     \end{align}
@@
     \begin{align}
-        &\mS = \mat(\tS), \quad \mS^{(t)} = \mat(\tS^{(t)}), \quad \tilde \mS = \mat(\tilde \tS)\\
-        &\mY = \mat(\tY), \quad \mX = \mat(\tX), \quad \mE = \mat(\tE).
-    \end{align}
-    \item We define the angle-based misclustering loss in the $t$-th iteration 
-    \begin{equation}
-         L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2,
-    \end{equation}
-     and the oracle loss
-    \begin{align}
-         \xi & = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \Big\{\ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s}  \leq - \frac{ \theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \Big\} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2,
-    \end{align}
-    where $m$ is a positive universal constant defined in~\eqref{eq:xv_lower}.
+        &\mS = \mat(\tS), \quad \mY = \mat(\tY), \quad \mX = \mat(\tX), \quad \mE = \mat(\tE)\\
+        &\mS^{(t)} = \mat(\tS^{(t)}), \quad  \hat \mS = \mat(\hat \tS), \quad \tilde \mS = \mat(\tilde \tS), \quad \bar \mS = \mat( \bar \tS ).
+    \end{align}
+    \item We define the extended core tensor on $K-1$ modes
+    \begin{equation}
+        \mA = \mS \mB^T , \quad \bar \mA = \bar \mS \mB^T, \quad \hat \mA = \hat \mS \hat \mB^T.
+    \end{equation}
+    By the assumption in parameter space~\eqref{eq:family}, we have $\mA = \mP \mX = \mW \mX, \quad \hat \mA = \hat \mP \hat \mX = \hat \mW \hat \mX.$
+
+    \item We define the angle-based misclustering loss in the $t$-th iteration and loss for MLE
+    \begin{align}
+        L^{(t)} &= \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \{ z^{(t)}(i) = b \} \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2, \\
+         L(\hat z)& = \frac{1}{p} \sum_{i \in [p]} \theta(i)^2 \sum_{b \in [r]} \ind\{ \hat z(i) = b\} \onormSize{}{[\mA_{z(i):}]^s - [\mA_{b:}]^s}^2.
+    \end{align}
+    We also define the loss for oracle and weighted oracle estimators
+    \begin{align}
+         \xi & = \frac{1}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind \Big\{\ang{ \mE_{i:} \mV, [  \tilde \mS_{z(i):} ]^s - [  \tilde \mS_{b:} ]^s}  \leq - \frac{ \theta(i) m}{4} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \Big\} \onormSize{}{ [ \mS_{z(i):}  ]^s - [ \mS_{ b:}  ]^s  }^2,\\
+         \xi' & = \frac{1}{p} \sum_{i \in [p]} \theta(i)^2 \sum_{b \in [r]} \ind \Big\{\ang{\mE_{i:}, [\bar \mA_{z(i):}]^s - [\bar \mA_{b:}]^s }  \leq - \frac{m'}{4}  \sqrt{\frac{p^{K-1}}{r^{K-1}}} \onormSize{}{ [\mA_{z(i):}]^s -  [\mA_{b:}]^s  }_F^2 \Big\}  \onormSize{}{[\mA_{z(i):}]^s - [\mA_{b:}]^s}^2.
+    \end{align}
+    where $m$ and $m'$ are some positive universal constants.
+    %defined in~\eqref{eq:xv_lower} and .
 \end{enumerate}
@@
 \begin{equation}\label{eq:cond_oracle}
-    \xi \leq \exp\of{ - M \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}},
+    \xi \leq \exp\of{ - M \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}}, \quad  \xi' \lesssim \exp\of{ -  \frac{\Delta_{\min}^2 p^{K-1}}{r^{K-1}}},
 \end{equation}
@@
 \begin{equation}\label{eq:cond_intial}
-    L^{(t)} \leq \frac{\bar C}{\tilde C} \frac{\Delta_{\min}^2}{r \log p}, \quad \text{for} \quad t = 0, 1, \ldots, T,
-\end{equation}
-where $M$ is a positive universal constant in inequality~\eqref{eq:constant_M}, $\bar C, \tilde C$ are universal constants in Corollary~\ref{cor:L0} and assumption SNR $\geq \tilde C p^{-K/2} \log p$, respectively.
-Further, inequality~\eqref{eq:cond1} holds by replacing $\mV$ to $\mV^{(t)}$ and $\mW_{:a}$ to $\mW_{:a}^{(t),T}$ when initialization condition~\eqref{eq:cond_intial} holds.
+    L^{(t)} \leq \frac{\bar C}{\tilde C} \frac{\Delta_{\min}^2}{r \log p}, \quad \text{for} \quad t = 0, 1, \ldots, T, \quad  L(\hat z) \leq \frac{\bar C}{\tilde C} \frac{\Delta_{\min}^2}{r \log p}
+\end{equation}
+where $M$ is a positive universal constant in inequality~\eqref{eq:constant_M}, $\bar C, \tilde C$ are universal constants in \textcolor{red}{the proof of Theorem~\ref{thm:initial}} and assumption SNR $\geq \tilde C p^{-K/2} \log p$, respectively.
+Further, inequality~\eqref{eq:cond1} holds by replacing $\mV$ to $\mV^{(t)}, \mQ, \hat \mQ$ and $\mW_{:a}$ to $\mW_{:a}^{(t),T}, \mP_{:a}^T, \hat \mP_{:a}^T$ when initialization condition~\eqref{eq:cond_intial} holds.
 \end{condition}
+
+}
 
@@
 
+    {
+    \color{blue}
+
+    \textbf{Now, we show the decomposition for MLE loss.}
+
+    By the definition of Gaussian MLE, the estimator $\hat \mtheta$ satisfies $\hat \theta(i)  = \ang{ \mY_{i:} , \hat \mA_{\hat z(i):} }/{\onormSize{}{ \hat \mA_{\hat z(i):}}_F^2}$ for all $i \in [p]$.
+Hence, we have 
+\begin{equation}
+    \hat z(i) = \argmin_{a \in [r_1]} \onormSize{}{ [\mY_{ i:}]^s - [\hat \mA_{a:}]^s}_F^2,
+\end{equation}
+and the decomposition 
+ \begin{equation}
+        L(\hat z) \leq \xi' + \frac{1}{p}\sum_{i \in [p] }  \sum_{b \in [r]/z(i)}  \zeta_{ib}', 
+    \end{equation}
+    where $ \zeta_{ib}' =   \theta(i)^2 B_{ib}' \onormSize{}{ [ \mA_{ z(i):}  ]^s - [ \mA_{b:}  ]^s  }^2$ and 
+    \begin{align}
+        A'_{ib} &= \ind \offf{  \hat z(i)  = b,   \ang{\mE_{i:}, [\bar \mA_{z(i):}]^s - [\bar \mA_{b:}]^s }  \leq - \frac{m'}{4}  \sqrt{\frac{p^{K-1}}{r^{K-1}}} \onormSize{}{ [\mA_{z(i):}]^s -  [\mA_{b:}]^s  }_F^2 }, \\
+    B'_{ib} &= \ind \offf{  \hat z(i)  = b,     - \frac{1}{2}  \onormSize{}{ [\mA_{z(i):}]^s -  [\mA_{b:}]^s  }_F^2  \leq \sqrt{\frac{r^{K-1}}{(m')^2 p^{K-1}}}  \hat F_{ib} + \hat G_{ib} + \hat H_{ib} }
+    \end{align}
+    with terms 
+    \begin{align}
+    \hat F_{ib} &= 2 \ang{ \mE_{i:}, ([\bar  \mA_{z(i):}]^s - [\hat \mA_{a:}]^s) - ([\bar  \mA_{b:}]^s - [\hat \mA_{b:}]^s) }, \\
+    \hat G_{ib}& = \of{ \onormSize{}{\mX_{i:}^s - [\hat \mA_{z(i):}]^s}_F^2 - \onormSize{}{ \mX^s_{i:} - [\mP_{:z(i)}^T \mY \hat \mQ \hat \mB^T ]^s }_F^2  } - \of{ \onormSize{}{\mX^s_{i:} - [\hat \mA_{b:}]^s}_F^2 - \onormSize{}{ \mX^s_{i:} - [\mP_{:b}^T \mY \hat \mQ \hat \mB^T]^s }_F^2   }, \\
+   \hat  H_{ib}& = \onormSize{}{ \mX^s_{i:} - [\mP_{:z(i)}^T \mY \hat \mQ \hat \mB^T]^s }_F^2 - \onormSize{}{ \mX^s_{i:} - [\mP_{:b}^T \mY \hat \mQ \hat \mB^T]^s }_F^2 + \onormSize{}{ \mA^s_{z(i):} -  \mA^s_{b:}  }_F^2.
+\end{align}
+
+    }
+
     {\bf Step 3 (Derivation of contraction inequality).} In this step we derive the upper bound of $\zeta_{ib}$ and obtain the contraction inequality~\eqref{eq:proof_5_ineq}.
@@
 
+     {
+     \color{blue}
+     \textbf{We also have the contraction inequality for MLE.}
+
+     Following the same derivation of~\eqref{eq:decomp_t} with the upper bound of $\hat F_{ib}, \hat G_{ib}, \hat H_{ib}$ in Lemma~\ref{lem:intermediate}, we also have
+    \begin{equation}\label{eq:decomp_t}
+          L(\hat z) \leq \frac{3}{2} \xi'  + \frac{1}{2} L(\hat z),
+    \end{equation}
+    which indicates the conclusion $\ell(\hat z,z) \lesssim \Delta_{\min}^2  \exp \of{ - \frac{p^{K-1}}{r^{K-1}} \Delta_{\min}^2  }$.
+     }
+
     {\bf Step 4 (Verification of Condition~\ref{cond:origin}).} Last, we verify the Condition~\ref{cond:origin} under high probability to finish the proof.
 Note that the inequalities~\eqref{eq:cond1}, \eqref{eq:cond2}, and \eqref{eq:cond3} describe the property of the sub-Gaussian noise tensor $\tE$, and the readers can find the proof directly in \citet[Step 5, Proof of Theorem 2]{han2020exact}.
-Here, we include only the verification of inequalities~\eqref{eq:cond_oracle} and \eqref{eq:cond_intial}.
+The initial condition~\eqref{eq:cond_intial} for MLE is satisfied by Lemma~\ref{lem:poly_mle_degree}.
+Here, we include only the verification of inequalities~\eqref{eq:cond_oracle} and \eqref{eq:cond_intial} for algorithm estimators.
 
@@
     \begin{equation}
-        \bbP\of{ \xi \lesssim \bbE \xi + \exp \of{  -M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 } } \geq 1 -  C \exp \of{  - M \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 },
-    \end{equation}
-    and thus the condition~\eqref{eq:cond_oracle} holds with probability at least $1 -  C \exp \of{  - M\frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }$ for some constant $C > 0$.
-
-    Finally, we verify the bounded loss condition~\eqref{eq:cond_intial} by induction.
-With output $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{2} and the assumption SNR $\geq \tilde C p^{-K/2} \log p$, by Corollary~\ref{cor:L0}, we have
+        \bbP\of{ \xi \lesssim \bbE \xi + \exp \of{  - \frac{M p^{K-1}}{2 r^{K-1}}   \Delta_{\min}^2 } } \geq 1 -  C \exp \of{  - \frac{M p^{K-1}}{2 r^{K-1}}   \Delta_{\min}^2 },
+    \end{equation}
+    and thus the condition~\eqref{eq:cond_oracle} holds with probability at least $1 -  C \exp \of{  - \frac{M p^{K-1}}{2 r^{K-1}}   \Delta_{\min}^2 }$ for some constant $C > 0$.
+
+    {
+    \color{blue}
+
+    \textbf{The initialization condition for MLE also holds.}
+
+    For $\xi'$, notice that $\ang{\mE_i, \mA^s_{a:} - \mA^s_{b:} }$ is a sub-Gaussian vector with variance bounded by $\onormSize{}{\mA^s_{a:} - \mA^s_{b:}}^2$ and 
+    \begin{align}
+        \bbP \of{ t \leq \onormSize{}{ [\bar \mA_{a:}]^s - \mA^s_{a:} }}  \leq \of{ t \leq \onormSize{}{ [\mP_{:a}^T \mY \mQ]^s - [\mP_{:a}^T \mX \mQ]^s }} \leq \bbP( t \min_{a \in [r]}\onormSize{}{\mS_{a:}} \leq \onormSize{}{\mP_{:a}^T \mE \mQ} ) \lesssim \exp \of{ - \frac{p^{K}t^2}{r^{K}}},
+    \end{align}
+   where the first inequality follows from the property in later inequality \eqref{eq:angle_aq}. We also have 
+   \begin{equation}
+        \xi' \lesssim  \exp \of{  -  \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }.
+   \end{equation}
+
+    }
+
+    Finally, we verify the bounded loss condition~\eqref{eq:cond_intial} for algorithm estimator by induction.
+With output $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{2} and the assumption SNR $\geq \tilde C p^{-K/2} \log p$, by Theorem~\ref{thm:initial}, we have
     \begin{equation}
@@
 {\bf Useful Definitions and Lemmas for the Proof of Theorem~\ref{thm:refinement}}
-
-\begin{lem}[Misclustering error and loss]\label{lem:mis} Define the misclustering error in the $t$-th iteration as $\ell^{(t)} = \ell(z^{(t)}, z)$. We have 
-\begin{equation}
-    \ell^{(t)} \lesssim \frac{1}{p}\sum_{i \in [p]} \theta(i) \ind\offf{ z^{(t)}(i) \neq z(i) } \leq \frac{L^{(t)}}{\Delta_{\min}^2}.
-\end{equation}
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:mis}]
-By the definition of minimal gap in Assumption~\ref{assmp:min_gap}, we have 
-\begin{equation}
-         L^{(t)} = \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \onormSize{}{ [ \mS_{ z(i):}  ]^s - [ \mS_{b:}  ]^s  }^2 \geq \frac{1}{p}  \sum_{i \in [p]} \theta(i) \sum_{b \in [r]}  \ind \offf{ z^{(t)}(i) = b } \Delta_{\min}^2 \geq c \ell^{(t)} \Delta_{\min}^2,
-    \end{equation}
-    where the last inequality follows from the assumption $\min_{i \in [p]} \theta(i) \geq c>0$.
-\end{proof}
 
@@
 \begin{equation}\label{eq:membership1}
-    \lambda_{r}(\mM) \asymp \onorm{\mM}_{\sigma} \asymp \sqrt{p/r}, \quad  \lambda_{r}(\mW) \asymp \onorm{\mW}_{\sigma} \asymp \sqrt{r/p}.
+    \lambda_{r}(\mM) \asymp \onorm{\mM}_{\sigma} \asymp \sqrt{p/r}, \   \lambda_{r}(\mW) \asymp \onorm{\mW}_{\sigma} \asymp \sqrt{r/p},\   \lambda_{r}(\mP) \asymp \onorm{\mP}_{\sigma}    \asymp  \min_{a \in [r] } \onormSize{}{\mtheta_{z^{-1}(a)}}^{-1}  \lesssim  \sqrt{r/p} .
 \end{equation}
@@
 
-\begin{proof}[Proof of Lemma~\ref{lem:membership}] The proof for the inequality~\eqref{eq:membership1} can be found in \citet[Proof of Lemma 4]{han2020exact}
+\begin{proof}[Proof of Lemma~\ref{lem:membership}] The proof for the inequality~\eqref{eq:membership1} for $\mM, \mW$ can be found in \citet[Proof of Lemma 4]{han2020exact}.
+The inequalities for $\mP$ follows the same derivation with balance assumption on $\mtheta$ and $\min_{i \in [p]}\theta(i) \geq c$.
 
@@
 
-\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement},  assume the the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ is large enough to satisfy the inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
-We have
+\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement} \textcolor{red}{with fixed $r \geq 2$},  assume the the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ is large enough to satisfy the inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
+\textcolor{red}{As $p \rightarrow \infty$}, we have
     \begin{equation}\label{eq:fib}
@@
     \end{equation}
+
+    Similarly, when the SNR $\geq \tilde C p^{-(K-1)}\log p$ with large constant $\tilde C$, we have 
+     \begin{equation}\label{eq:fib_cor}
+        \max_{i \in [p]} \max_{b \neq z(i)} \frac{\of{\hat F_{ib}}^2}{ \onormSize{}{ [  \mA_{z(i):} ]^s  -  [ \mA_{b:} ]^s }^2 } \lesssim p^{K-1} \frac{r L(\hat z)}{ \Delta_{\min}^2}
+        %+ \of{1 +  \frac{ r L(\hat z)}{ \Delta_{\min}^2}} \onormSize{}{\mE_{i:} (\mV - \mV^{(t)})}^2 ,
+    \end{equation}
+    \begin{equation}\label{eq:gib_cor}
+        \max_{i \in [p]} \max_{b \neq z(i)} \frac{ \of{  \hat G_{ib} }^2  }{ \onormSize{}{ [\mA_{z(i):} ]^s - [\mA_{b:}]^s }^2}  \leq  \frac{1}{512} \of{ \Delta_{\min}^2 +L(\hat z)},
+    \end{equation}
+    \begin{equation}\label{eq:hib_cor}
+        \max_{i \in [p]} \max_{b \neq z(i)}  \frac{ \aabs{  \hat H_{ib} }  }{ \onormSize{}{ [\mA_{z(i):} ]^s - [\mA_{b:}]^s }^2} \leq \frac{1}{4}.
+    \end{equation}
     \end{lem}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:upper_fgh}]
-We prove the each of the inequalities in Lemma~\ref{lem:upper_fgh} separately.
+We prove the the first three inequalities in Lemma~\ref{lem:upper_fgh} separately.
 \begin{enumerate}[wide]
@@
 \end{enumerate}
+{
+\color{blue}
+\textbf{Next, we show the upper bounds for $\hat F_{ib}, \hat G_{ib}$ and $\hat H_{ib}$.}
+
+By Lemma~\ref{lem:angle_gap_x}, we have 
+\begin{equation}
+    \onormSize{}{\mS_{a:}^s - \mS_{b:}^s}  = (1 + o(1)) \onormSize{}{\mA_{a:}^s - \mA_{b:}^s}.
+\end{equation}
+Also, notice that the matrix product of $ \mB^T$ corresponding to the padding operation in Lemma~\ref{lem:pad}, and the padding weights are balanced such that $\onormSize{}{\mv \mB}  = (1 + o(1)) \max_{a} \onormSize{}{\mtheta_{z^{-1}(a)}}^{(K-1)/2} \onormSize{}{\mv} $ for all $\mv \in \bbR^{r(K-1)}$.
+For two vectors $\mv_1, \mv_2 \in \bbR^{r^{K-1}}$, we have
+\begin{equation}\label{eq:b_angle}
+    \onormSize{}{\mv_1^s - \mv_2^s} = (1 + o(1)) \onormSize{}{[\mv_1 \mB^T]^s -  [\mv_2 \mB^T]^s  }.
+\end{equation}
+The equation~\eqref{eq:b_angle} also holds for $\hat \mB^T$.
+
+Note that for all $i \in [p]$ we have 
+\begin{equation}\label{eq:aq}
+    \onormSize{}{\mA_{i:} \hat \mQ} = \onormSize{}{\mS_{z(i:)} \mB^T \hat \mQ } =  \onormSize{}{\mS_{z(i:)} \hat \mD^{\otimes (K-1)} } =  (1 + o(1)) \onormSize{}{\mS_{z(i:)}} =  (1 + o(1)) \max_{a} \onormSize{}{\mtheta_{z^{-1}(a)}}^{-(K-1)/2} \onormSize{}{\mA_{i:}},
+\end{equation}
+where the third inequality follows from the singular property of MLE confusion matrix~\eqref{eq:mle_confusion} and the last inequality follows from the fact that $\mA_i = \mS_{z(i:)} \mB^T$ and Lemma~\ref{lem:membership}.
+Above equation indicates that $\mA_{i:}$ is the span space of the singular values as $p \rightarrow \infty$.
+Also, notice that the row space of $\mP_{:a}^T \mY \hat \mQ  \hat \mB^T$ is equal to the column space of $\hat \mQ$, and $\mA_{i:} \neq \mP_{:a}^T \mY \hat \mQ  \hat \mB^T$ in noisy case.
+
+Hence, for all $a \in [r]$, we have 
+\begin{align}
+    \onormSize{}{ [\mX_i \hat \mQ]^s - [\mP_{:a}^T \mY \hat \mQ ]^s } &= \onorm{ \frac{\mA_{z(i:)} \hat \mQ  }{\onormSize{}{\mA_{z(i:)} \hat \mQ }} - \frac{\mP_{:a}^T \mY \hat \mQ  }{ \onormSize{}{\mP_{:a}^T \mY \hat \mQ} } } \\
+    &= (1+o(1))  \onorm{ \frac{\mA_{z(i:)} }{\onormSize{}{\mA_{z(i:)} }} - \frac{\mP_{:a}^T \mY \hat \mQ \hat \mB^T  }{ \onormSize{}{\mP_{:a}^T \mY \hat \mQ \hat \mB^T} } } \\
+    &= (1+o(1)) \onormSize{}{ [\mX_i]^s - [\mP_{:a}^T \mY \hat \mQ \hat \mB^T ]^s } \label{eq:angle_aq}
+\end{align}
+where the second equation follows from~\eqref{eq:aq}, $\onormSize{}{\mP_{:a}^T \mY \hat \mQ \hat \mB^T} = (1 + o(1)) \max_{a} \onormSize{}{\mtheta_{z^{-1}(a)}}^{(K-1)/2} \onormSize{}{\mP_{:a}^T \mY \hat \mQ } $, and singular property of $\hat \mB^T$.
+Similar result holds after replacing $\mP_{:a}^T \mY \hat \mQ$ by  $\mP_{:a}^T \mY\mQ$ or $\mP_{:a}^T \mY \hat \mQ$.
+
+We are now ready to show the upper bounds for $\hat F_{ib},\hat G_{ib}$ and $\hat H_{ib}$.
+
+ For $\hat F_{ib}$, we have 
+\begin{align}
+    (\hat F_{ib})^2 &\leq \onormSize{}{\mE_{i:}}^2 \onormSize{}{ [\bar  \mA_{a:}]^s - [\hat \mA_{a:}]^s }^2\\
+    &\leq \onormSize{}{\mE_{i:}}^2  \off{ \onormSize{}{ [ \bar \mS_{a:} \mB^T]^s - [\bar \mS_{a:} \hat \mB^T]^s } +  \onormSize{}{ [ \bar \mS_{a:} \hat \mB^T]^s -  [\hat \mS_{a:} \hat \mB^T]^s}  }^2 \\
+    &\lesssim \onormSize{}{\mE_{i:}}^2  \off{ \onormSize{}{ [ \bar \mS_{a:} \mB^T \hat \mQ]^s - [\bar \mS_{a:}]^s } +  \onormSize{}{ [ \bar \mS_{a:} ]^s -  [ \hat \mS_{a:} ]^s}  }^2.
+\end{align}
+Following similar derivations in inequalities~\eqref{eq:fib_1}, \eqref{eq:fib_2}, and the upper bound for $J_{1}$, respectively, we have 
+\begin{equation}
+    \onormSize{}{ [ \bar \mS_{a:} ]^s -  [ \hat \mS_{a:} ]^s}  \lesssim r L(\hat z), \quad  \onormSize{}{ [ \bar \mS_{a:} ]^s -  [ \bar \mS_{b:} ]^s} \lesssim \onormSize{}{\mS_{a:}^s - \mS_{b:}^s}^2, \quad  \onormSize{}{ [ \bar \mS_{a:} \mB^T \hat \mQ]^s - [\bar \mS_{a:}]^s } \lesssim L(\hat z).
+\end{equation}
+We then obtain the upper bound for $\hat F_{ib}$ by noticing that $\onormSize{}{\mE_i}^2 \lesssim p^{K-1}$.
+
+For $\hat G_{ib}$ and $\hat H_{ib}$, by the property~\eqref{eq:angle_aq}, we have 
+    \begin{align}
+        (1+o(1))\hat G_{ib}& = \of{ \onormSize{}{[\mX_{i:}\hat \mQ]^s - [\hat \mS_{a:}]^s}_F^2 - \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:a}^T \mY \hat \mQ ]^s }_F^2  } \\
+        & \quad - \of{ \onormSize{}{[\mX_{i:} \hat \mQ]^s - [\hat \mS_{b:}]^s}_F^2 - \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:b}^T \mY \hat \mQ ]^s }_F^2   }. \\
+        (1+o(1))\hat H_{ib} & = \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:a}^T \mY \hat \mQ ]^s }_F^2 - \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:b}^T \mY \hat \mQ]^s }_F^2 + \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^2.
+    \end{align}
+    We obtain the upper bounds following the proof for inequalities~\eqref{eq:gib} and \eqref{eq:hib}.
+}
 
@@
 
-     \begin{lem}[Relationship between misclustering loss and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement}, we have
+     \begin{lem}[Relationship between misclustering loss and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement} \textcolor{red}{with fixed $r \geq 2$, as $p \rightarrow \infty$}, we have
     \begin{equation}\label{eq:inter1}
@@
     \end{equation}
-    for some positive universal constant $C$. In addition, the inequality~\eqref{eq:inter4} also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$.
+    for some positive universal constant $C$.
+In addition, the inequality~\eqref{eq:inter4} also holds by replacing $\mW_{:b}^{(t)}$ to $\mW_{:b}$.
+Further, the above inequalities holds after replacing $\mW$ to $\mP$, $\mV$ to $\mQ$, and $L^{(t)}$ to $L(\hat z)$.
     \end{lem}
@@
 
+    % {
+    % \color{blue}
+
+    % Consider two vectors 
+    % \begin{equation}
+    %     \mp( z^{(t)})=(| z^{(t),-1}(1)|, \ldots,| z^{(t),-1}(r)|)^T,\quad \mp_{\mtheta}( z^{(t)})=(\onormSize{}{\mtheta_{ z^{(t),-1}(1)}}_1,\ldots,\onormSize{}{\mtheta_{\bar z^{(t), -1}(r)}}_1)^T.
+    % \end{equation}
+    %  Note that we have $\ell^{(t)} \lesssim p^{-1} \sum_{i \in [p]} \theta(i) \ind\{ z^{(t)} = b \}  \leq \frac{L^{(t)}}{\Delta_{\min}^2} $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.
+For all $b \in [r]$, We have
+    % \begin{equation}
+    %     | z^{(t),-1}(b)| =  (1 + \tO(L^{(t)}/\Delta_{\min}^2 ))  | z^{-1}(b)|,
+    % \end{equation}
+    % and 
+    % \begin{align}
+    %     \onormSize{}{\mtheta_{ z^{(t),-1}(b)}}_1 &= \sum_{ i \in [p]} \theta(i) \ind\{ z^{(t)}(i) = z(i) = b \} + \sum_{ i \in [p]} \theta(i) \ind\{ z^{(t)}(i) = b,  z(i) \neq b \} \\
+    %     &\ +  \sum_{ i \in [p]} \theta(i) \ind\{ z^{(t)}(i) \neq  b,  z(i) = b \} - \sum_{ i \in [p]} \theta(i) \ind\{ z^{(t)}(i) \neq  b,  z(i) = b \}  \\
+    %     & =  | z^{-1}(b)| + \tO(  \sum_{i \in [p]} \theta(i) \ind\{ z^{(t)} \neq b \} ) = (1 + \tO(L^{(t)}/\Delta_{\min}^2 ))  | z^{-1}(b)|.
+    % \end{align}
+    % Hence, we have 
+    % \begin{equation}
+    %     \onormSize{}{  \mp( z^{(t)}) -  \mp_{\mtheta}( z^{(t)}) } \leq r \tO(L^{(t)}/\Delta_{\min}^2 )  \max_{b \in [r]}| z^{-1}(b)| \lesssim  \tO(L^{(t)}/\Delta_{\min}^2 ) p
+    % \end{equation}
+    % with assumption $\onormSize{}{ z^{-1}(b)} \lesssim p$ in parameter space~\eqref{eq:family}.
+    %  Note that $\text{diag}(\mp(z^{(t)})) \mD = \text{diag}(\mp_{\mtheta}(z^{(t)})) \tilde \mD$. Then, we have
+    % \begin{align}
+    %     \onormSize{}{ \mD -  \tilde \mD}_F &\leq  \onormSize{}{ \mI_r -  \text{diag}(\mp(z^{(t)})) \text{diag}^{-1}(\mp_{\mtheta}(z^{(t)}))}_F \onormSize{}{\mD}_F\\
+    %   &\lesssim \frac{ \onormSize{}{ \mp(z^{(t)}) - \mp_{\mtheta}(z^{(t)})} }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  } \\
+    %   & = \frac{\tO(L^{(t)}/\Delta_{\min}^2 ) p }{\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1 } \\
+    %   & \lesssim \frac{L^{(t)}}{\Delta_{\min}},\fixme{miaoyan}{$\Delta^2_{min}$?}
+    %   \end{align}
+    % where the last inequality follows from the constraint $\min_{i \in [p]}\theta(i) \geq c>0$  and $\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  \gtrsim p$.
+By the geometry property of trigonometric function, we have
+    % \begin{align}\label{eq:b1}
+    %     \sin \beta_{k1} &= \min_{c \in \bbR} \frac{\onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes (\mD -  \tilde \mD) \otimes \mD^{\otimes K-k - 1} } }{ \onormSize{}{\mD_{:b}^{T} \mS \mI_r^{\otimes k-1} \otimes \mD^{\otimes K-k}} }\\
+    %     &\leq \frac{ \onormSize{}{\mD_{:b}^{T} \mS} \onormSize{}{\mD - \tilde \mD}_{\sigma} \onormSize{}{\mD}_{\sigma}^{K-k-1} }{\onormSize{}{\mD_{:b}^{T} \mS} \lambda_r^{K-k}(\mD) }\\
+    %     & \lesssim  \onormSize{}{\mD - \tilde \mD}_{F}\\
+    %     & \lesssim \frac{L^{(t)}}{\Delta_{\min}},
+    % \end{align}
+    % where the second inequality follows from the singular property of $\mD$ in \eqref{eq:singular_Du}, \eqref{eq:singular_Dl} and the constraint of $\mS$ in \eqref{eq:family}.
+
+    % }
+
     To bound $\sin \beta_{k1}$, recall Definition~\ref{def:stable} that for any cluster assignment $\bar z$ in the $\varepsilon$-neighborhood of true $z$,
@@
     \end{equation}
-    Note that we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \lesssim \log^{-1}(p) $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.
-With the locally linear stability assumption, we have
+     \textcolor{red}{Note that we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \leq \frac{\bar C}{\tilde C} r \log^{-1}(p) $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.
+Then,  with the locally linear stability assumption, the $\mtheta$ is $\ell^{(t)}$-locally linearly stable; i.e.,}
+     %With the locally linear stability assumption, we have 
     \begin{equation}
@@
     where the last inequality follows from Lemma~\ref{lem:membership}, the constraint $\min_{i \in [p]}\theta(i) \geq c>0$, $\onormSize{}{ \mp(z^{(t)})} \lesssim p$ and $\min_{a \in [r]} \onormSize{}{\mtheta_{z^{(t),-1}(a)} }_1  \gtrsim p$.
+
     By the geometry property of trigonometric function, we have
@@
         & \lesssim \onormSize{ }{(\mI_r - \tilde \mD^T) \mS \mC^{-1}}_F \onormSize{}{\mC \mZ^k}_{\sigma}\\
-        & \lesssim \frac{r}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \onormSize{}{\mS_{b:}^s - \mS_{z(i):}^s} \\
-        & \lesssim \frac{L^{(t)}}{\Delta_{\min}}, 
-    \end{align}
-    where the third inequality follows from the singular property of $\mD$ and the boundedness of $\mS$, and the fourth inequality follows from the definition of $\tilde \mD$, boundedness of $\mS$, the lower bound of $\mtheta$, and the singular property of $\mZ^k$ in inequality~\eqref{eq:singular_z}.
+        & \lesssim \frac{r}{p} \sum_{i \in [p]} \theta(i) \sum_{b \in [r]} \ind\{z^{(t)}(i) = b\} \onormSize{}{\mS_{b:}^s - \mS_{z(i):}^s} \\
+        & \lesssim \frac{L^{(t)}}{\Delta_{\min}}, %\fixme{miaoyan}{reason for last line?}
+    \end{align}
+    where the third inequality follows from the singular property of $\mD$ and the boundedness of $\mS$, and the fourth inequality follows from the definition of $\tilde \mD$, boundedness of $\mS$, the lower bound of $\mtheta$, and the singular property of $\mZ^k$ in inequality~\eqref{eq:singular_z}, and the last line follows from the definition of $L^{(t)}$.
 
@@
         \end{enumerate}
+  {
+  \color{blue}
+  \textbf{Next, we show the intermediate inequalities holds with $\mP, \mQ$ and $L(\hat z)$.}
+
+        Consider the MLE confusion matrix $\hat \mD = \mM^T \mTheta^T \hat \mP = \entry{\hat D_{ab}} \in \bbR^{r \times r}$ with entries
+\begin{align}\label{eq:mle_confusion}
+    \hat D_{ab} = \frac{ \sum_{i \in [p]} \theta(i) \hat \theta(i) \ind\{ z(i) = a, \hat z(i) = b  \} }{\onormSize{}{\hat \mtheta_{\hat z^{-1}(b)}}^2} = \frac{ \sum_{i \in [p]} (1 + o(p^{K-2}))(\hat \theta(i))^2 \ind\{ z(i) = a, \hat z(i) = b  \} }{\onormSize{}{\hat \mtheta_{\hat z^{-1}(b)}}^2},
+\end{align}
+where the second equation follows from Lemma~\ref{lem:poly_mle_degree}, and thus $\sum_{a \in [r]}\hat \mD_{ab} = 1 + o(1).$ By the derivation of~\eqref{eq:dab_bound}, \eqref{eq:dbb_bound}, \eqref{eq:singular_Dl}, and \eqref{eq:singular_Du}, we have
+\begin{equation}
+     \sum_{a \neq b \in [r]} \hat D_{ab} \lesssim \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\hat \theta(i))^2 \lesssim \frac{1}{\log p}, \quad \hat D_{bb} \gtrsim 1- \frac{1}{\log p}, \quad \lambda_{\min}(\hat \mD) \asymp \onormSize{}{\hat \mD}_{\sigma} = (1 + o(1)).
+\end{equation}
+for all $a \neq b \in [r]$.
+
+Now, we are ready to show the intermediate inequalities. First, by Lemma~\ref{lem:angle_gap_x} and $\min_{ i \in [p]} \theta(i) \geq c$, we have 
+\begin{equation}
+    \onormSize{}{\mS_{a:}^s - \mS_{b:}^s} \asymp \onormSize{}{\mA_{a:}^s - \mA_{b:}^s}.
+\end{equation}
+Then we can replace the $L^{(t)}$ by $L(\hat z)$ in the proof of Lemma~\ref{lem:intermediate}.
+The analogies of inequalities~\eqref{eq:inter1}, \eqref{eq:inter2}, \eqref{eq:inter3},  \eqref{eq:inter4}, and \eqref{eq:inter5}  hold by using the MLE confusion matrix and the definition of $L(\hat z)$.
+%It suffices to two show the inequalities~\eqref{eq:inter3} and \eqref{eq:inter4} holds with $\mP, \mQ$.
+
+Particularly, for the analogy of \eqref{eq:inter4}, the usage of MLE confusion matrix avoids the stability condition on $\mtheta$.
+Let $\bar \mD$ be the normalized version of $\hat \mD$.
+The angle in inequality~\eqref{eq:b1} decays to 0 at speed $p^{-(K-2)} \lesssim \Delta_{\min}$ when $K \geq 3$, and the inequality~\eqref{eq:b2} holds by the fact that
+\begin{equation}
+    \onormSize{}{(\mI_r - \bar \mD) \mS \mC^{-1}}_F \lesssim \frac{r}{p} \sum_{i \in [p]} (\theta(i))^2 \sum_{b \in [r]} \onormSize{}{\mS_{b:}^s - \mS_{z(i):}^s} \lesssim \frac{r}{p} \sum_{i \in [p]} (\theta(i))^2 \sum_{b \in [r]} \onormSize{}{\mA_{b:}^s - \mA_{z(i):}^s}.
+\end{equation}
+  }
 
@@
 
+{
+\color{blue}
+
+\begin{lem}[Polynomial estimation error of MLE]\label{lem:poly_mle_degree}  Let $(\hat z, \hat \mtheta, \hat \tS)$ denote the MLE in \eqref{eq:mle} \textcolor{red}{with fixed $K \geq 2$}, and $\hat \tX$ denote the mean tensor consisting of parameter $(\hat z, \hat \mtheta, \hat \tS)$.
+\textcolor{red}{With high probability going to 1 as $p \rightarrow \infty$, we} have
+\begin{equation}
+    \onormSize{}{\tX - \hat \tX}_F^2 \lesssim \sigma^2 \of{ r^{K} + K p r},
+\end{equation}
+with probability going to 1. 
+When \textcolor{red}{  $\text{SNR} \gtrsim p^{-(K-1)} \log p$,  $\mtheta$ is balanced and $\min_{i \in [p]} \theta(i) \geq c$ for some constant $c$}, the MLE satisfies
+\begin{equation}
+    \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\theta(i))^2 \lesssim \frac{1}{r \log p}, \quad \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\hat \theta(i))^2 \lesssim \frac{1}{r \log p} ,\quad   L(\hat z) \lesssim \frac{\Delta_{\min}^2}{r \log p},
+\end{equation}
+% The MLE loss is upper bounded as 
+% \begin{equation}
+%      L(\hat z) = \frac{1}{p} \sum_{i \in [p]} \theta(i)^2 \sum_{b \in [r]} \ind\{ \hat z(i) = b\} \onormSize{}{[\mA_{\hat \pi(z(i)):}]^s - [\mA_{b:}]^s}^2 \lesssim 
+% \end{equation}
+% where $\mA \coloneqq \mS (\mM^T \mTheta^T)^{\otimes (K-1)}$, and $\hat \pi$ is the permutation minimizes the MLE error.  
+Further, we have 
+\begin{equation}
+     \theta(i)^2 = (1 + o(p^{-(K-2)})) \hat \theta(i)^2.
+ \end{equation}
+\end{lem}
+
+\begin{proof}[Proof of Lemma~\ref{lem:poly_mle_degree}] Without loss of generality, we assume $\sigma^2 = 1$ and identity mapping minimizes the misclustering error for MLE.
+For arbitrary two sets of parameters $(z, \mtheta, \tS), (z_k', \mtheta_k', \tS') \in \tP_b(\gamma)$ and corresponding mean tensors $\tX, \tX'$, we have 
+\begin{equation}
+    \text{rank}(\mat_k(\tX) - \mat_k(\tX')) \leq \text{rank}(\mat_k(\tX)) + \text{rank}(\mat_k(\tX)) \leq 2 r_k, \quad k \in [K].
+\end{equation}
+Hence, we have
+\begin{equation}\label{eq:dtbm_tucker}
+    \tX - \tX' \in \tQ(2r_1, \ldots, 2 r_K) \coloneqq \{ \text{Tucker tensor with rank }(2r_1, \ldots, 2 r_K) \}.
+\end{equation}
+
+Then, we obtain that 
+\begin{align}
+     \bbP(\FnormSize{}{\tX - \hat \tX_{ML}} \geq t) &\leq 2  \bbP\of{ \sup_{ \tX, \tX' \in \tP(r_1, 
+\ldots, r_K)} \ang{ \frac{\tX - \tX'}{\FnormSize{}{\tX - \tX'}}, \tE  } \geq t } \\
+     &\leq 2  \bbP\of{ \sup_{ \tT \in \tQ(2r_1, \ldots, 2r_K) \cap \{ \onormSize{}{\tT}_F = 1\} } \ang{ \tT, \tE  } \geq t } \\
+     &\lesssim  \exp(-Kpr),
+\end{align}
+with the choice $t\asymp \sigma \sqrt{(Kpr + r^K)}$.
+Here the first inequality follows from \citet[Lemma 1]{wang2019multiway}, the second inequality follows from~\eqref{eq:dtbm_tucker}, and the last inequality follows from \citet[Lemma E5]{han2022optimal}.
+\end{proof}
+
+When $\Delta_{\min}^2 \gtrsim p^{-(K-1)} \log p$, we replace the vector $\hat x_{\hat z(i)}$ and $\hat \mX$ by our MLE estimator in the proof of Theorem~\ref{thm:initial}.
+With estimation error $\onormSize{}{\tX - \hat \tX}_F^2 \lesssim \of{ r^{K} + K p r}$ and $\Delta_{\min}^2 \gtrsim p^{-(K-1)} \log p$, we have
+\begin{align}
+      \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\theta(i))^2 \lesssim \frac{r^{K-1}}{\Delta_{\min}^2 p^{K}} \onormSize{}{\tX - \hat \tX}_F^2  \lesssim \frac{r^{K-2}}{p^{K-1} \Delta_{\min}^2} \lesssim \frac{1}{r \log p}, \quad \text{and} \quad  L(\hat z) \lesssim \frac{\Delta_{\min}^2}{r \log p}.
+\end{align}
+Above result holds for $\hat \theta(i)$ after switching the parameters $\mX$ with $\hat \mX$ and switch $\mtheta$ with $\hat \mtheta$ in the proof.
+
+Last, notice that for all $a \in [r]$
+% following the proof of inequality~\eqref{eq:cor_4} in Corollary~\ref{cor:L0} with $\onormSize{}{\tX - \hat \tX}^2_F \lesssim pr$,  we have 
+% \begin{equation}
+%     \onormSize{}{ \mW_{:a}^T \mX -  \hat \mW_{:a}^T \hat \mX }
+% \end{equation}
+%  for all $a \in [r]$,
+\begin{align}
+    (1 - O(1)) \frac{p^2}{r^2}\onormSize{}{\mW_{:a}^T \mX - \hat \mW_{:a}^T \hat \mX}_F^2 \leq  \onormSize{}{ \sum_{\hat z(i) = z(i) = a} (\theta(i) \mW_{:a}^T \mX  - \hat \theta(i) \hat \mW_{:a}^T \hat \mX)}_F^2 \leq \onormSize{}{\tX - \hat \tX}_F^2 \leq p r,
+\end{align}
+where the first inequality follows from the facts that  $\ell(\hat z, z) \lesssim \frac{1}{\log p}, |z^{-1}(a)| \asymp p /r$, 
+\begin{align}
+     | z^{-1}(a)| - C \frac{p}{r} \ell(\hat z, z)  \leq |\hat z^{-1}(a)| \leq | z^{-1}(a)| + C \frac{p}{r} \ell(\hat z, z), 
+\end{align}
+\begin{equation}
+  | z^{-1}(a)|  - C \frac{p}{r} \ell( \hat z,z)  \leq \sum_{ z(i) = z(i) = a} \theta(i) \leq | z^{-1}(a)|, \quad \text{and} \quad |\hat z^{-1}(a)|  - C \frac{p}{r} \ell(\hat z, z)  \leq \sum_{\hat z(i) = z(i) = a} \hat \theta(i) \leq |\hat z^{-1}(a)|.
+\end{equation}
+
+Hence, for all $i \in [p]$
+\begin{align}\label{eq:theta_hat}
+  (\theta(i) - \hat \theta(i))^2 \onormSize{}{ \mW_{:a}^T \mX }_F^2 - O(p) \leq  \onormSize{}{ (\theta(i) - \hat \theta(i))\mW_{:a}^T \mX }_F^2 - \onormSize{}{ \hat \theta(i) (\mW_{:a}^T \mX - \hat \mW_{:a}^T \hat \mX) }_F^2 \leq \onormSize{}{\tX - \hat \tX}_F^2 \leq p r,
+\end{align}
+where the first inequality follows from $\onormSize{}{\mW_{:a}^T \mX - \hat \mW_{:a}^T \hat \mX}_F^2 \lesssim 1/p$ and $\hat \theta(i) \lesssim \frac{p}{r}$. Notice that for all $a \in [r]$
+\begin{equation}
+    \onormSize{}{\mW_{:a}^T \mX}_F^2 \geq \onormSize{}{\mS_{a:}}_F^2 \lambda_{\min}^{2(K-1)}(\mTheta \mM) \gtrsim p^{K-1}.
+\end{equation}
+The inequality indicates that $\theta(i)^2 = (1 + o(p^{-(K-2)})) \hat \theta(i)^2$.
+
+}
+
+% \begin{cor}[Initial errors of MLE]\label{cor:mle_initial} When $\Delta_{\min}^2 \gtrsim p^{-(K-1)} \log p$, we have 
+% \begin{equation}
+%     \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\theta(i))^2 \lesssim \frac{1}{r \log p}, \quad \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\hat \theta(i))^2 \lesssim \frac{1}{r \log p}.
+% \end{equation}and 
+% \begin{equation}
+%     L(\hat z) \lesssim \frac{\Delta_{\min}^2}{r \log p},
+% \end{equation}
+% Further, we have 
+% \begin{equation}
+%     \theta(i) = (1 + o(1)) \hat \theta(i).
+% \end{equation}
+% \end{cor}
+
+% \begin{proof}[Proof of Corollary~\ref{cor:mle_initial}] Recall the convention that the variance $\sigma = 1$, and that the identity permutation minimizes the initial misclustering error; i.e., $\hat \pi = \argmin_{\pi \in \Pi} \sum_{i \in [p]}\ind \offf{\hat z(i) \neq \pi \circ z(i) }$ and $\hat \pi(a) = a$ for all $ a \in [r]$.
+
+% We replace the vector $\hat x_{\hat z(i)}$ by our MLE $\hat \mW_{:\hat z(i) }\hat \mX$ in the proof of Theorem 4. Then, we have 
+% \begin{align}
+%     \frac{1}{p}\sum_{i \in [p], \hat z(i) \neq z(i)} (\theta^{(i)})^2 \lesssim \frac{r^{K-1}}{\Delta_{\min}^2 p^{K}} \onormSize{}{\tX - \hat \tX}_F^2  \lesssim \frac{r^{K-2}}{p^{K-1} \Delta_{\min}^2}.
+% \end{align}
+% By the assumption that $\Delta_{\min}^2 \gtrsim \frac{p^{(K-1)}}{r^{K-1}} \log p$ and  $\min_{i \in [p]}\theta(i) \geq c$, we have 
+% \begin{equation}
+%   \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\theta(i))^2 \lesssim \frac{1}{r \log p}, \quad  \ell(\hat z, z)  \lesssim \frac{1}{r \log p}.
+% \end{equation}
+% Note that $(\hat z, \hat \mtheta, \hat \tS) \in \tP(\gamma)$ and the minimal angle gap in $\hat \tS$ also larger than $\Delta_{\min}^2$.
+By Lemma~\ref{lem:angle_gap_x}, we have $\min_{a \neq b \in [r]} \onormSize{}{ [\hat \mW_{:a}^T \hat \mX]^s -  [\hat \mW_{:b}^T \hat \mX]^s } \gtrsim \Delta_{\min}$.
+Then, we switch the parameters $\mX$ with $\hat \mX$ and switch $\mtheta$ with $\hat \mtheta$ in the proof of Theorem~\ref{thm:initial}.
+We obtain
+% \begin{equation}
+%     \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\hat \theta(i))^2 \lesssim \frac{1}{r \log p}.
+% \end{equation}
+
+% Next, we plug the estimator $\hat z$ into the proof of Corollary~\ref{cor:L0} with $\onormSize{}{\tX - \hat \tX}^2_F \lesssim pr$ and $\ell(\hat z, z) \lesssim \frac{1}{\log p}$. We obtain that 
+% \begin{align}
+%     L(\hat z) \lesssim \frac{r^K}{p^{K}} \onormSize{}{\tX - \hat \tX}^2_F  \lesssim \frac{r \Delta_{\min}^2}{\log p}.
+% \end{align}
+
+% Last, notice that for all $a \in [r]$
+% % following the proof of inequality~\eqref{eq:cor_4} in Corollary~\ref{cor:L0} with $\onormSize{}{\tX - \hat \tX}^2_F \lesssim pr$,  we have 
+% % \begin{equation}
+% %     \onormSize{}{ \mW_{:a}^T \mX -  \hat \mW_{:a}^T \hat \mX }
+% % \end{equation}
+% %  for all $a \in [r]$,
+% \begin{align}
+%     (1 - O(1)) \frac{p^2}{r^2}\onormSize{}{\mW_{:a}^T \mX - \hat \mW_{:a}^T \hat \mX}_F^2 \leq  \onormSize{}{ \sum_{\hat z(i) = z(i) = a} (\theta(i) \mW_{:a}^T \mX  - \hat \theta(i) \hat \mW_{:a}^T \hat \mX)}_F^2 \leq \onormSize{}{\tX - \hat \tX}_F^2 \leq p r,
+% \end{align}
+% where the first inequality follows from the facts that  $\ell(\hat z, z) \lesssim \frac{1}{\log p}, |z^{-1}(a)| \asymp p /r$, 
+% \begin{align}
+%      | z^{-1}(a)| - C \frac{p}{r} \ell(\hat z, z)  \leq |\hat z^{-1}(a)| \leq | z^{-1}(a)| + C \frac{p}{r} \ell(\hat z, z), 
+% \end{align}
+% \begin{equation}
+%   | z^{-1}(a)|  - C \frac{p}{r} \ell( \hat z,z)  \leq \sum_{ z(i) = z(i) = a} \theta(i) \leq | z^{-1}(a)|, \quad \text{and} \quad |\hat z^{-1}(a)|  - C \frac{p}{r} \ell(\hat z, z)  \leq \sum_{\hat z(i) = z(i) = a} \hat \theta(i) \leq |\hat z^{-1}(a)|.
+% \end{equation}
+
+% Hence, for all $i \in [p]$
+% \begin{align}\label{eq:theta_hat}
+%   (\theta(i) - \hat \theta(i))^2 \onormSize{}{ \mW_{:a}^T \mX }_F^2 - O(p) \leq  \onormSize{}{ (\theta(i) - \hat \theta(i))\mW_{:a}^T \mX }_F^2 - \onormSize{}{ \hat \theta(i) (\mW_{:a}^T \mX - \hat \mW_{:a}^T \hat \mX) }_F^2 \leq \onormSize{}{\tX - \hat \tX}_F^2 \leq p r,
+% \end{align}
+% where the first inequality follows from $\onormSize{}{\mW_{:a}^T \mX - \hat \mW_{:a}^T \hat \mX}_F^2 \lesssim 1/p$ and $\hat \theta(i) \lesssim \frac{p}{r}$. Notice that for all $a \in [r]$
+% \begin{equation}
+%     \onormSize{}{\mW_{:a}^T \mX}^2 \geq \onormSize{}{\mS_{a:}}^2 \lambda_{\min}^{2(K-1)}(\mTheta \mM) \gtrsim p^{K-1}.
+% \end{equation}
+% The inequality indicates that $\theta(i) = (1 + o(1)) \hat \theta(i)$.
+
+% Note that for all $a \in [r]$, we have at least one $i \in [p]$ such that $\hat z(i) = z(i) = a$, otherwise, we have degenerate number of communities.
+Then, by Lemma~\ref{lem:poly_mle_degree}, we have
+% \begin{equation}
+%     \max_{i \in [p]} \onormSize{}{\mX_{i:} - \hat \mX_{i:} }^2 \leq \onormSize{}{\tX - \hat \tX} \lesssim pr,
+% \end{equation}
+% and thus for all $a \in [r]$
+% \begin{equation}
+%     \onormSize{}{ \theta(i) \mW_{:a}^T \mX}^2 \onormSize{}{[\mW_{:a}^T \mX]^s - [\hat \mW_{:a}^T \hat \mX]^s }^2 \lesssim \onormSize{}{\theta(i)\mW_{:a}^T \mX - \hat \theta(i)\hat \mW_{:a}^T \hat \mX }^2 = \max_{i \in [p]}\onormSize{}{\mX_{i:} - \hat \mX_{i:} }^2 \lesssim pr,
+% \end{equation}
+% where the first inequality follows from Lemma~\ref{lem:norm_diff} and the fact that $[c\mv]^s = \mv^s$ for positive constant $c$.
+
+% Following
+
+% Note that 
+% \begin{equation}
+%     \onormSize{}{\mW_{:a}^T \mX}^2 \geq \onormSize{}{\mS_{a:}}^2 \lambda_{\min}^{2(K-1)}(\mTheta \mM) \gtrsim p^{K-1},
+% \end{equation}
+% where the last inequality follows from Lemma~\ref{lem:singular_thetam} and the assumption $\min_{a \in [r]}\onormSize{}{\mS_{a:}}^2 \geq c_3$.
+
+% By the last inequality in the proof of Corollary~\ref{cor:L0}, we have 
+% \begin{align}
+%     L(\hat z) &\lesssim \frac{1}{p} \sum_{i \in [p], \hat z(i)  \neq z(i)} \theta(i) \onormSize{}{\mX_{i:}^s - [\mW_{:\hat z(i)}^T \mX]^s }^2 \\
+%     & \lesssim  \frac{1}{p} \sum_{i \in [p], \hat z(i)  \neq z(i)} \theta(i)  \of{ \onormSize{}{\mX_{i:}^s - [\hat \mW_{:\hat z(i)}^T \hat \mX]^s }^2 +  \onormSize{}{ [\hat \mW_{:\hat z(i)}^T \hat \mX]^s - [\mW_{:\hat z(i)}^T \mX]^s }^2 } \\
+%     & \lesssim   \frac{1}{p} \sum_{i \in [p], \hat z(i)  \neq z(i)} \frac{r^{K-1}}{p^{K-1}} \onormSize{}{\mX_{i:}}^2  \onormSize{}{\mX_{i:}^s - [\hat \mW_{:\hat z(i)}^T \hat \mX]^s }^2 + \theta(i) \frac{r^{K-1}}{p^{K-1}} pr
+% \end{align}
+
+%\end{proof}
+
+% {
+% \color{blue}
+
+% \section*{Proof of Theorem~\ref{thm:mle}}
+
+% \begin{proof}[Proof of Theorem~\ref{thm:mle}] The Lemma~\ref{lem:poly_mle_degree} and its Corollary~\ref{cor:mle_initial} indicate that the MLE satisfies similar initialization condition in Condition~\ref{cond:origin} of the proof for Theorem~\ref{thm:comp}.
+The Lemma~\ref{lem:contract_mle} indicates the likelihood loss also has a contraction property around the global optimum MLE.
+Note that the contraction inequality in Lemma~\ref{lem:contract_mle} is different with the contraction in Theorem~\ref{thm:comp} due to the different objective function for loss.
+Though, the proof idea for Lemma~\ref{lem:contract_mle} and Theorem~\ref{thm:comp} are similar.
+
+% Combining the results in Lemma~\ref{lem:poly_mle_degree}, Corollary~\ref{cor:mle_initial}, and Lemma~\ref{lem:contract_mle}, we obtain that 
+% \begin{equation}
+%     \ell(\hat z, z) \leq \Delta_{\min}^2 \exp \of{ - \frac{p^{K-1}}{r^{K-1}} \Delta_{\min}^2 },
+% \end{equation}
+% when $\Delta_{\min}^2 \geq C p^{-(K-1)}\log p$ for large positive constant $C$.
+% \end{proof}
+
+% {\bf Useful Lemmas and Corollaries for the proof of Theorem~\ref{thm:mle}}
+
+% \begin{lem}[Loss decomposition of MLE]\label{lem:contract_mle} Consider the angle-based loss function for MLE
+%   \begin{equation}\label{eq:mle_loss}
+%     L(\hat z) = \frac{1}{p} \sum_{i \in [p]} (\theta(i))^2 \sum_{b \in [r]} \ind\{ \hat z(i) = b\} \onormSize{}{[\mA_{a:}]^s - [\mA_{b:}]^s}^2.
+% \end{equation}
+% When $\Delta_{\min}^2 \log p$, the MLE of the assignment $z$ satisfies 
+% \begin{equation}
+%     L(\hat z) \leq C \exp\of{ - \frac{p^{K-1}}{r^{K-1}} \Delta_{\min}^2}   + \rho' L(\hat z),
+% \end{equation}
+% where $C$ is a positive constant and $\rho' \in (0,1)$ is a contraction constant.
+% \end{lem}
+
+% \begin{proof}[Proof of Lemma~\ref{lem:contract_mle}]  
+%  We keep using the conventions and notations in the proof of Theorem~\ref{thm:comp}.
+Let $(\hat z, \hat \mtheta, \hat \tS)$ denote the MLE in \eqref{eq:mle}.
+We start by claiming more specific notations related to MLE.
+
+% \textbf{Notation.} 
+% \begin{enumerate}[wide]
+%     \item Let the parameters with super-script $\hat \cdot$ (e.g.\ $\hat \mW, \hat \mV, \hat \tS$) denote the analogy of the parameters with super-script $(t)$ (e.g.\ $\mW^{(t)}, \mV^{(t)}, \tS^{(t)}$) in the proof of Theorem~\ref{thm:comp}.
+%     \item Dual dual weighted membership matrices 
+%     \begin{equation}
+%         \mB = (\mTheta \mM)^{\otimes (K-1)}, \quad  \hat \mB = (\hat \mTheta \hat \mM)^{\otimes (K-1)}
+%     \end{equation}
+%     \item Consider the weighted normalized membership matrices
+%     \begin{equation}
+%         \mP = \mTheta \mM ( \text{diag}( \onormSize{}{\mtheta_{z^{-1}(1)}}^2 , \ldots, \onormSize{}{\mtheta_{z^{-1}(r)}}^2 ) )^{-1}, \quad  \hat \mP = \hat \mTheta \hat \mM ( \text{diag}( \onormSize{}{\hat \mtheta_{z^{-1}(1)}}^2 , \ldots, \onormSize{}{\hat \mtheta_{z^{-1}(r)}}^2 ) )^{-1},
+%     \end{equation}
+%     and the dual weighted normalized membership matrices 
+%     \begin{equation}
+%         \mQ = \mP^{\otimes K-1}, \quad \hat \mQ = \hat \mP^{\otimes K-1}.
+%     \end{equation}
+%     By the definition, we have $\mB^T \mQ =\hat \mB^T \hat \mQ =\mI$.
+%     \item Consider the weighted oracle estimator and the MLE 
+%     \begin{equation}
+%     \bar \tS  = \tY \times_1 \mP^T \times_2 \cdots \times_K \mP^T,  \quad  \hat \tS =  \tY \times_1 \hat \mP^T \times_2 \cdots \times_K \hat \mP^T.
+% \end{equation} Similar as Lemma~\ref{lem:membership} and with the balance assumption, we have 
+%     \begin{equation}\label{eq:singular_P}
+%   \lambda_{r}(\mP) \asymp \onorm{\mP}_{\sigma}    \asymp   \sqrt{\frac{1}{\min_{a \in [r] } \onormSize{}{\mtheta_{z^{-1}(a)}}^2  }  } \lesssim   \sqrt{ \frac{r}{p}} .
+%     \end{equation}
+
+% \item Consider the matricizations for core tensors
+% \begin{align}
+%   \mS = \mP^T \mX \mQ, \quad  \bar \mS =  \mP^T \mY \mQ, \quad \hat \mS  =  \hat \mP^T \mY \hat \mQ.
+% \end{align}
+% Also, consider the matrices
+% \begin{equation}
+%     \mA = \mS \mB^T , \quad \bar \mA = \bar \mS \mB^T, \quad \hat \mA = \hat \mS \hat \mB^T.
+% \end{equation}
+% By the assumption that $\onormSize{}{\mtheta_{z^{-1}(a)}}_1 = |z^{-1}(a)|$ for all $a \in [r]$, we have 
+%     \begin{equation}
+%         \mA = \mP \mX = \mW \mX, \quad \hat \mA = \hat \mP \hat \mX = \hat \mW \hat \mX.
+%     \end{equation}
+%   %\item  The MLE loss function~\eqref{eq:mle_loss} is different with the loss function $L^{(t)}$ by replacing the angle gaps between $\mS_{a:}$'s to the angle gaps between the rows of the mean tensor $\mP_{:a}^T \mX$.
+
+% \end{enumerate}
+
+% \textbf{Misclustering loss decomposition.}
+
+% By the MLE property, we have 
+% \begin{equation}
+%     \hat z(i) = \argmin_{a \in [r_1]} \onormSize{}{ [\mY_{ i:}]^s - [\hat \mA_{a:}]^s}_F^2.
+% \end{equation}
+
+% Consider the node $i \in [p]$ such that $z(i) = a, \hat z(i) = b$ and $a \neq b \in [r]$. Then, we have 
+% \begin{equation}\label{eq:mle_event}
+%      \onormSize{}{[\mY_{i:}]^s - [\hat \mA_{b:}]^s}_F^2 \leq  \onormSize{}{[\mY_{i:}]^s - [\hat \mA_{a:}]^s}_F^2.
+% \end{equation}
+% Let $\bar \tS$ denote the weighted oracle estimator, where
+% \begin{equation}
+%     \bar \tS  = \tY \times_1 \mP^T \times_2 \cdots \times_K \mP^T, \quad \bar \mS =  \mP^T \mY \mQ
+% \end{equation}
+% and $\bar \mA = \bar \mS \mB^T$. Above event~\eqref{eq:mle_event} implies the decomposition
+% \begin{equation}
+%     2 \ang{\mE_{i:}, [\bar \mA_{a:}]^s - [\bar \mA_{b:}]^s } \leq \onormSize{}{\mX_{i:}}(- \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^2 +  \hat G_{ib} + \hat H_{ib}) + \hat F_{ib} ,
+% \end{equation}
+% where 
+% \begin{align}
+%     \hat F_{ib} &= 2 \ang{ \mE_{i:}, ([\bar  \mA_{a:}]^s - [\hat \mA_{a:}]^s) - ([\bar  \mA_{b:}]^s - [\hat \mA_{b:}]^s) }, \\
+%     \hat G_{ib}& = \of{ \onormSize{}{\mX_{i:}^s - [\hat \mA_{a:}]^s}_F^2 - \onormSize{}{ \mX^s_{i:} - [\mP_{:a}^T \mY \hat \mQ \hat \mB^T ]^s }_F^2  } - \of{ \onormSize{}{\mX^s_{i:} - [\hat \mA_{b:}]^s}_F^2 - \onormSize{}{ \mX^s_{i:} - [\mP_{:b}^T \mY \hat \mQ \hat \mB^T]^s }_F^2   }, \\
+%   \hat  H_{ib}& = \onormSize{}{ \mX^s_{i:} - [\mP_{:a}^T \mY \hat \mQ \hat \mB^T]^s }_F^2 - \onormSize{}{ \mX^s_{i:} - [\mP_{:b}^T \mY \hat \mQ \hat \mB^T]^s }_F^2 + \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^2.
+% \end{align}
+% Note that 
+% \begin{equation}
+%     \onormSize{}{\mX_{i:}} \geq \min_{i \in [p_1]} \theta(i) \onormSize{}{\mS_{a:}} \lambda_{\min}(\mQ)\geq m \sqrt{\frac{p^{K-1}}{r^{K-1}}},
+% \end{equation}
+% for some positive constant $m$.
+
+% Then, we have 
+% \begin{equation}
+%     \ind\{   \hat z(i)  = b  \} \leq A'_{ib} + B'_{ib},  
+% \end{equation}
+% where 
+% \begin{align}
+%     A'_{ib} &= \ind \offf{  \hat z(i)  = b,   \ang{\mE_{i:}, [\bar \mA_{a:}]^s - [\bar \mA_{b:}]^s }  \leq - \frac{m}{4}  \sqrt{\frac{p^{K-1}}{r^{K-1}}} \onormSize{}{ [\mA_{a:}]^s -  [\mA_{b:}]^s  }_F^2 }, \\
+%     B'_{ib} &= \ind \offf{  \hat z(i)  = b,     - \frac{1}{2}  \onormSize{}{ [\mA_{a:}]^s -  [\mA_{b:}]^s  }_F^2  \leq \sqrt{\frac{r^{K-1}}{m^2 p^{K-1}}}  F_{ib} + G_{ib} + H_{ib} }.
+% \end{align}
+
+% \textbf{Derivation of contraction inequality.} Following the proof for Theorem~\ref{thm:comp} with Corollary~\ref{cor:upper_fgh_cor}, we have 
+% \begin{equation}
+%     L(\hat z) \leq \xi' + \frac{1}{2} L(\hat z),
+% \end{equation}
+% where 
+% \begin{equation}
+%     \xi' =  \frac{1}{p} \sum_{i \in [p]} (\theta(i))^2 \sum_{b \in [r]} A_{ib'} \onormSize{}{[\mA_{a:}]^s - [\mA_{b:}]^s}^2. 
+% \end{equation}
+% Now, it suffices to show $\bbE[\xi'] \lesssim \exp \of{ - \frac{p^{K-1}}{r^{K-1}}\Delta_{\min}^2 }$.
+
+% Note that 
+% \begin{align}
+%     \bbE[A_{ib}] \leq &  \bbP \of{\ang{\mE_i, \mA^s_{a:} - \mA^s_{b:} } \leq - \frac{m}{8} \sqrt{\frac{p^{K-1}}{r^{K-1}}}  \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^2 }  \\
+%     & + 2 \bbP \of{\ang{\mE_i, [\bar \mA_{a:}]^s - \mA^s_{a:}} \leq -\frac{m}{16} \sqrt{\frac{p^{K-1}}{r^{K-1}}} \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^2 }.
+% \end{align}
+
+% For the first probability, by Gaussian concentration, we have 
+% \begin{equation}
+%     \bbP \of{\ang{\mE_i, \mA^s_{a:} - \mA^s_{b:} } \leq - \frac{m}{8} \sqrt{\frac{p^{K-1}}{r^{K-1}}}  \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^2 }  \lesssim \exp \of{ - \frac{p^{K-1}}{r^{K-1}}  \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^2 }.
+% \end{equation}
+
+% For the second probability, we decompose the probability 
+% \begin{equation}
+%     \bbP \of{\ang{\mE_i, [\bar \mA_{a:}]^s - \mA^s_{a:}} \leq -\frac{m}{16} \sqrt{\frac{p^{K-1}}{r^{K-1}}} \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^2 } \leq P_{21} + P_{22},
+% \end{equation}
+% where 
+% \begin{align}
+%     P_{21} &= \bbP \of{ t \leq \onormSize{}{ [\bar \mA_{a:}]^s - \mA^s_{a:} }}  \\
+%     P_{22}& = \bbP \of{\ang{\mE_i, [\bar \mA_{a:}]^s - \mA^s_{a:}} \leq -\frac{m}{16} \sqrt{\frac{p^{K-1}}{r^{K-1}}} \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^2 \Big|  \onormSize{}{ [\tilde \mA_{a:}]^s - \mA^s_{a:} }  < t}.
+% \end{align}
+% For $P_{21}$, we have 
+% \begin{equation}
+%      P_{21} \leq \bbP \of{ t \leq \onormSize{}{ [\mP_{:a}^T \mY \mQ]^s - [\mP_{:a}^T \mX \mQ]^s }} \leq \bbP( t \min_{a \in [r]}\onormSize{}{\mS_{a:}} \leq \onormSize{}{\mP_{:a}^T \mE \mQ} ) \lesssim \exp \of{ - \frac{p^{K}t^2}{r^{K}}},
+% \end{equation}
+% where the first inequality follows from the property in later inequality \eqref{eq:angle_aq}, and the last inequality follows from the Bernstein inequality and the fact that $\min_{a \in [r]}\onormSize{}{\mtheta_{z^{-1}(a)}^2} \gtrsim \frac{r}{p}$.
+
+% For $P_{22}$, note that $\ang{\mE_i, [\bar \mA_{a:}]^s - \mA^s_{a:}}$ is a sub-Gaussian variable with bounded variance $t^2$ conditional $\onormSize{}{ [\bar \mA_{a:}]^s - \mA^s_{a:} }  < t$.
+By sub-Gaussian concentration, we have
+% \begin{equation}
+%      P_{22} \lesssim \exp\of{ - \frac{p^{K-1}}{r^{K-1} t^2 }  \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^4}.
+% \end{equation}
+% Take $t = \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F$. Combining the upper bounds for $P_{21}$ and $P_{22}$, we have 
+% \begin{equation}
+%     \bbE[A_{ib}]  \lesssim \exp \of{ - \frac{p^{K-1}}{r^{K-1}} \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^2 } \lesssim \exp \of{ - \frac{p^{K-1}}{r^{K-1}} \onormSize{}{ \mS^s_{a:} -  \mS^s_{b:}  }_F^2 }  \lesssim \exp \of{ - \frac{p^{K-1}}{r^{K-1}} \Delta_{\min}^2},
+% \end{equation}
+% where the second inequality follows from Lemma~\ref{lem:angle_gap_x}. 
+% \end{proof}
+
+% \begin{lem}[Confusion matrix of MLE]\label{lem:new_confusion} Assume $K \geq 3$.
+Consider the confusion matrix $\hat \mD = \mM^T \mTheta^T \hat \mP = \entry{D_{ab}} \in \bbR^{r \times r}$, where for all $a , b \in [r]$
+% \begin{align}
+%     \mD_{ab} = \frac{ \sum_{i \in [p]} \theta(i) \hat \theta(i) \ind\{ z(i) = a, \hat z(i) = b  \} }{\onormSize{}{\hat \mtheta_{\hat z^{-1}(b)}}^2} 
+% \end{align}
+% We have 
+% \begin{equation}
+%     \sum_{a \in [r]}\hat \mD_{ab} = 1 + o(1).
+% \end{equation}
+% Thus, we have 
+% \begin{equation}
+%      \sum_{a \neq b \in [r]} \hat \mD_{ab} \lesssim \frac{1}{\log p}, \quad \hat \mD_{bb} \gtrsim 1- \frac{1}{\log p}, \text{ for all } b \in [r].
+% \end{equation}
+% Further, we have 
+% \begin{equation}
+%     \lambda_{\min}(\hat \mD) \asymp \onormSize{}{\mD}_{\sigma} = (1 + o(1)).
+% \end{equation}
+% \end{lem}
+
+% \begin{proof}[Proof of Lemma~\ref{lem:new_confusion}] By Corollary~\ref{cor:mle_initial}, we have 
+% \begin{equation}
+%     \mD_{ab}  =  \frac{ \sum_{i \in [p]} (1 + o(1))(\hat \theta(i))^2 \ind\{ z(i) = a, \hat z(i) = b  \} }{\onormSize{}{\hat \mtheta_{\hat z^{-1}(b)}}^2}.
+% \end{equation}
+% Then, by the definition of $\hat \mD$ and the assumption $\min_{i \in [p]}\theta(i) \geq c$, we have 
+% \begin{equation}
+%       \sum_{a \neq b \in [r]} \hat \mD_{ab} \lesssim \frac{1}{p} \sum_{i \in [p]} \ind\{ \hat z(i) \neq z(i) \} (\hat \theta(i))^2 \lesssim \frac{1}{\log p},
+% \end{equation}
+% where the last inequality follows from Corollary~\ref{cor:mle_initial}.
+The lower bound of $\hat \mD_{bb}$ and the bounds for singular values follows the proof of \eqref{eq:dbb_bound}, \eqref{eq:singular_Du}, and \eqref{eq:singular_Dl}.
+% \end{proof}
+
+% \begin{cor}[Upper bound for $\hat F_{ib},\hat G_{ib}$ and $\hat H_{ib}$] \label{cor:upper_fgh_cor} We have 
+%     \begin{equation}\label{eq:fib_cor}
+%         \max_{i \in [p]} \max_{b \neq z(i)} \frac{\of{\hat F_{ib}}^2}{ \onormSize{}{ [  \mA_{z(i):} ]^s  -  [ \mA_{b:} ]^s }^2 } \lesssim p^{K-1} \frac{r L(\hat z)}{ \Delta_{\min}^2}
+%         %+ \of{1 +  \frac{ r L(\hat z)}{ \Delta_{\min}^2}} \onormSize{}{\mE_{i:} (\mV - \mV^{(t)})}^2 ,
+%     \end{equation}
+%     \begin{equation}\label{eq:gib_cor}
+%         \max_{i \in [p]} \max_{b \neq z(i)} \frac{ \of{  \hat G_{ib} }^2  }{ \onormSize{}{ [\mA_{z(i):} ]^s - [\mA_{b:}]^s }^2}  \leq  \frac{1}{512} \of{ \Delta_{\min}^2 +L(\hat z)},
+%     \end{equation}
+%     \begin{equation}\label{eq:hib_cor}
+%         \max_{i \in [p]} \max_{b \neq z(i)}  \frac{ \aabs{  \hat H_{ib} }  }{ \onormSize{}{ [\mA_{z(i):} ]^s - [\mA_{b:}]^s }^2} \leq \frac{1}{4}.
+%     \end{equation}
+% \end{cor}
+
+% \begin{proof}[Proof of Corollary~\ref{cor:upper_fgh_cor}] By Lemma~\ref{lem:angle_gap_x}, we have 
+% \begin{equation}
+%     \onormSize{}{\mS_{a:}^s - \mS_{b:}^s}  = (1 + o(1)) \onormSize{}{\mA_{a:}^s - \mA_{b:}^s}.
+% \end{equation}
+% Also, notice that the matrix product of $ \mB^T$ corresponding to the padding operation in Lemma~\ref{lem:pad}, and the padding weights are balanced such that $\onormSize{}{\mv \mB}  = (1 + o(1)) \max_{a} \onormSize{}{\mtheta_{z^{-1}(a)}}^{(K-1)/2} \onormSize{}{\mv} $ for all $\mv \in \bbR^{r(K-1)}$.
+For two vectors $\mv_1, \mv_2 \in \bbR^{r^{K-1}}$, we have
+% \begin{equation}\label{eq:b_angle}
+%     \onormSize{}{\mv_1^s - \mv_2^s} = (1 + o(1)) \onormSize{}{[\mv_1 \mB^T]^s -  [\mv_2 \mB^T]^s  }.
+% \end{equation}
+% The equation~\eqref{eq:b_angle} also holds for $\hat \mB^T$.
+
+% Note that for all $i \in [p]$ we have 
+% \begin{equation}\label{eq:aq}
+%     \onormSize{}{\mA_{i:} \hat \mQ} = \onormSize{}{\mS_{z(i:)} \mB^T \hat \mQ } =  \onormSize{}{\mS_{z(i:)} \hat \mD^{\otimes (K-1)} } =  (1 + o(1)) \onormSize{}{\mS_{z(i:)}} =  (1 + o(1)) \max_{a} \onormSize{}{\mtheta_{z^{-1}(a)}}^{-(K-1)/2} \onormSize{}{\mA_{i:}},
+% \end{equation}
+% where the third inequality follows from the singular property of confusion matrix in Lemma~\ref{lem:new_confusion} and the last inequality follows from the fact that $\mA_i = \mS_{z(i:)} \mB^T$.
+Above equation indicates that $\mA_{i:}$ is the span space of the singular values as $p \rightarrow \infty$.
+Also, notice that the row space of $\mP_{:a}^T \mY \hat \mQ  \hat \mB^T$ is equal to the column space of $\hat \mQ$, and $\mA_{i:} \neq \mP_{:a}^T \mY \hat \mQ  \hat \mB^T$ in noisy case.
+
+% Hence, for all $a \in [r]$, we have 
+% \begin{align}
+%     \onormSize{}{ [\mX_i \hat \mQ]^s - [\mP_{:a}^T \mY \hat \mQ ]^s } &= \onorm{ \frac{\mA_{z(i:)} \hat \mQ  }{\onormSize{}{\mA_{z(i:)} \hat \mQ }} - \frac{\mP_{:a}^T \mY \hat \mQ  }{ \onormSize{}{\mP_{:a}^T \mY \hat \mQ} } } \\
+%     &= (1+o(1))  \onorm{ \frac{\mA_{z(i:)} }{\onormSize{}{\mA_{z(i:)} }} - \frac{\mP_{:a}^T \mY \hat \mQ \hat \mB^T  }{ \onormSize{}{\mP_{:a}^T \mY \hat \mQ \hat \mB^T} } } \\
+%     &= (1+o(1)) \onormSize{}{ [\mX_i]^s - [\mP_{:a}^T \mY \hat \mQ \hat \mB^T ]^s } \label{eq:angle_aq}
+% \end{align}
+% where the second equation follows from~\eqref{eq:aq}, $\onormSize{}{\mP_{:a}^T \mY \hat \mQ \hat \mB^T} = (1 + o(1)) \max_{a} \onormSize{}{\mtheta_{z^{-1}(a)}}^{(K-1)/2} \onormSize{}{\mP_{:a}^T \mY \hat \mQ } $, and singular property of $\hat \mB^T$.
+Similar result holds after replacing $\mP_{:a}^T \mY \hat \mQ$ by  $\mP_{:a}^T \mY\mQ$ or $\mP_{:a}^T \mY \hat \mQ$.
+
+% We are now ready to show the upper bounds for $\hat F_{ib},\hat G_{ib}$ and $\hat H_{ib}$.
+
+% \begin{enumerate}
+% \item For $\hat F_{ib}$, we have 
+% \begin{align}
+%     (\hat F_{ib})^2 &\leq \onormSize{}{\mE_{i:}}^2 \onormSize{}{ [\bar  \mA_{a:}]^s - [\hat \mA_{a:}]^s }^2\\
+%     &\leq \onormSize{}{\mE_{i:}}^2  \off{ \onormSize{}{ [ \bar \mS_{a:} \mB^T]^s - [\bar \mS_{a:} \hat \mB^T]^s } +  \onormSize{}{ [ \bar \mS_{a:} \hat \mB^T]^s -  [\hat \mS_{a:} \hat \mB^T]^s}  }^2 \\
+%     &\lesssim \onormSize{}{\mE_{i:}}^2  \off{ \onormSize{}{ [ \bar \mS_{a:} \mB^T \hat \mQ]^s - [\bar \mS_{a:}]^s } +  \onormSize{}{ [ \bar \mS_{a:} ]^s -  [ \hat \mS_{a:} ]^s}  }^2.
+% \end{align}
+% By similar proof for inequality~\eqref{eq:fib_1} and \eqref{eq:fib_2} and Lemma~\ref{cor:inter_weight}, we have 
+% \begin{equation}
+%     \onormSize{}{ [ \bar \mS_{a:} ]^s -  [ \hat \mS_{a:} ]^s}  \lesssim r L(\hat z), \quad  \onormSize{}{ [ \bar \mS_{a:} ]^s -  [ \bar \mS_{b:} ]^s} \lesssim \onormSize{}{\mS_{a:}^s - \mS_{b:}^s}^2.
+% \end{equation}
+% Following the proof to derive $J_{1}$ in the proof for inequality~\eqref{eq:inter4}, we have 
+% \begin{equation}
+%      \onormSize{}{ [ \bar \mS_{a:} \mB^T \hat \mQ]^s - [\bar \mS_{a:}]^s } \lesssim L(\hat z).
+% \end{equation}
+% We then obtain the inequality~\eqref{eq:fib_cor} by noticing that $\onormSize{}{\mE_i}^2 \lesssim p^{K-1}$.
+
+%     \item For $\hat G_{ib} $, by the property~\eqref{eq:angle_aq}, we have 
+%     \begin{align}
+%         (1+o(1))\hat G_{ib}& = \of{ \onormSize{}{[\mX_{i:}\hat \mQ]^s - [\hat \mS_{a:}]^s}_F^2 - \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:a}^T \mY \hat \mQ ]^s }_F^2  } \\
+%         & \quad - \of{ \onormSize{}{[\mX_{i:} \hat \mQ]^s - [\hat \mS_{b:}]^s}_F^2 - \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:b}^T \mY \hat \mQ ]^s }_F^2   }.
+%     \end{align}
+%     Following the proof for inequality~\eqref{eq:gib} with Lemma~\ref{cor:inter_weight}, we obtain the inequality~\eqref{eq:gib_cor}.
+% \end{enumerate}
+
+% \item For $\hat H_{ib}$, by the property~\eqref{eq:angle_aq}, we have 
+% \begin{align}
+%     (1+o(1))\hat H_{ib} \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:a}^T \mY \hat \mQ ]^s }_F^2 - \onormSize{}{ [\mX_{i:} \hat \mQ]^s - [\mP_{:b}^T \mY \hat \mQ]^s }_F^2 + \onormSize{}{ \mA^s_{a:} -  \mA^s_{b:}  }_F^2.
+% \end{align}
+% Following the proof for inequality~\eqref{eq:hib} with Lemma~\ref{cor:inter_weight}, we obtain the inequality~\eqref{eq:hib_cor}.
+
+% \end{proof}
+
+% \begin{cor}[Intermediate inequalities with weighted normalized matrix]\label{cor:inter_weight} The inequalities~\eqref{eq:inter3}, \eqref{eq:inter4}, and \eqref{eq:inter5} holds after replacing $\mW, \mW^{(t)}$ by $\mP, \hat \mP$, replacing $\mV, \mV^{(t)}$ by $\mQ, \hat \mQ$, and replacing $L^{(t)}$ by $L(\hat z)$.
+That is
+% \begin{equation}\label{eq:inter3_cor}
+%      \max_{b \in [r]} \onormSize{}{ [ \mP_{:b}^T \mY \mQ ]^s   -  [ \hat \mP_{:b}^{T} \mY \mQ  ]^s } \leq C \of{\frac{r L(\hat z)}{\Delta_{\min}} + \sqrt{ \frac{r^{2K} + pr^{K+1}}{p^{K}} } \frac{\sqrt{L(\hat z)}}{\Delta_{\min}}},
+%  \end{equation}
+%  \begin{equation}\label{eq:inter4_cor}
+%      \max_{b \in [r]} \onormSize{}{  [  \hat \mP_{:b}^{ T}  \mY \mQ]^s - [  \hat \mP_{:b}^{T}  \mY \hat \mQ]^s } \leq C  \of{ \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L(\hat z)}}{\Delta_{\min}} +  \frac{r L(\hat z)}{\Delta_{\min}}},
+%      \end{equation}
+%     \begin{equation}\label{eq:inter5_cor}
+%          \max_{b \in [r]} \onormSize{}{ [\mP_{:b}^T \mY \hat \mQ ]^s   -  [ \mP_{:b}^{(t), T} \mY \hat \mQ ]^s } \leq C \of{\frac{rL(\hat z)}{\Delta_{\min}} + \sqrt{ \frac{rr^{2K} + p r^{K+2}}{p^{K}}}  \frac{\sqrt{L(\hat z)}}{\Delta_{\min}}},
+%     \end{equation}
+%     for some positive universal constant $C$. In addition, the inequality~\eqref{eq:inter4} also holds by replacing $\hat \mP_{:b}$ to $\mP_{:b}$.
+% \end{cor}
+
+% \begin{proof}[Proof of Corollary~\ref{cor:inter_weight}] By Lemma~\ref{lem:angle_gap_x}, we have 
+% \begin{equation}
+%     \onormSize{}{\mS_{a:}^s - \mS_{b:}^s} \asymp \onormSize{}{\mA_{a:}^s - \mA_{b:}^s}.
+% \end{equation}
+% Then we can replace the $L^{(t)}$ by $L(\hat z)$ in the proof of Lemma~\ref{lem:intermediate}.
+Also, replace the confusion matrix $\mD$ by $\hat \mD$ in Lemma~\ref{lem:new_confusion} and use the singular property of $\mTheta \mM, \mP$ in Lemma~\ref{lem:singular_thetam} and~\eqref{eq:singular_P}.
+We obtain the inequalities~\eqref{eq:inter3_cor} and~\eqref{eq:inter4_cor}.
+
+% Particularly, in the proof of \eqref{eq:inter4_cor}, we replace the normalized confusion matrix $\tilde \mD$ by $\hat \mD$, and thus avoid the stability condition on $\mtheta$.
+The angle in inequality~\eqref{eq:b1} decays to 0, and the inequality~\eqref{eq:b2} holds by the fact that
+% \begin{equation}
+%     \onormSize{}{(\mI_r - \hat \mD) \mS \mC^{-1}}_F \lesssim \frac{r}{p} \sum_{i \in [p]} (\theta(i))^2 \sum_{b \in [r]} \onormSize{}{\mS_{b:}^s - \mS_{z(i):}^s} \lesssim \frac{r}{p} \sum_{i \in [p]} (\theta(i))^2 \sum_{b \in [r]} \onormSize{}{\mA_{b:}^s - \mA_{z(i):}^s},
+% \end{equation}
+% where the first inequality follows by the Lemma~\ref{lem:new_confusion} and $\theta(i) = (1 + o(1)) \hat \theta(i)$ in Corollary~\ref{cor:mle_initial}.
+
+% Similar as the proof of Lemma~\ref{lem:intermediate}, the inequality~\eqref{eq:inter5_cor} is a direct consequence of inequalities~\eqref{eq:inter3_cor} and \eqref{eq:inter4_cor}.
+%Therefore, we only verify the inequalities \eqref{eq:inter3_cor} and \eqref{eq:inter4_cor} by comparing the proof of Lemma~\ref{lem:intermediate}.
+
+% \end{proof}
+
+% }
+
 \end{document}

### 2023-01-24 tensor_clustering/IEEE/ieee_revision/ieee_revise_s1.tex
organize

--- 2023-01-24 tensor_clustering/IEEE/ieee_revision/ieee_revise_s1.5.tex
@@
 The constants $c_1, c_2$ in the $|z^{-1}(a)|$ bound assume the roughly balanced size across $r$ communities.
-\textcolor{red}{Third, the constant $c_3$ requires that all slides in $\tS$ have non-degenerate norm.
-Particularly, the lower bound $c_3$ excludes the no purely zero slide case to avoid trivial non-identifiability of model \eqref{eq:model_tensor}; see Example~\ref{example:c3} below.
-The upper bound $c_4$ is a technical constraint to avoid the slides with unbounded norm as dimension grows; in practice, the constraint $\max_{a:} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large $c_4 \geq \onormSize{}{\tY}_F$.
-} %\textcolor{blue}{ Third, the constant $c_3$ requires no purely zero slide in $\tS$, and thus the model~\eqref{eq:model_tensor} is not trivially non-identifiable; see Example~\ref{example:c3} below.
+\textcolor{blue}{ Third, the constant $c_3$ requires no purely zero slide in $\tS$, and thus the model~\eqref{eq:model_tensor} is not trivially non-identifiable; see Example~\ref{example:c3} below.
 The upper bound $c_4$ is a technical constraint to avoid the unbounded entries in the core tensor; in practice, the constraint $\max_{a:} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large $c_4 \geq \onormSize{}{\tY}_F$.
@@
 
-\begin{thm}[Model identifiability]\label{thm:unique} Consider the dTBM with $r\geq 1$ and \textcolor{red}{ $K \geq 2$}.
+\begin{thm}[Model identifiability]\label{thm:unique} Consider the dTBM with $r\geq 2$.
 The parameterization~\eqref{eq:model_tensor} is unique in $\tP$ up to cluster label permutations, if and only if Assumption~\ref{assmp:min_gap} holds.
@@
 
-\begin{lem}[Angle gaps in $\tX$ and $\tS$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family} \textcolor{red}{with $r \geq 2$.
-Suppose $\mtheta$ is balanced satisfying~\eqref{eq:degree} \textcolor{red}{and $\min_{i\in[p]}\theta(i)\geq c$ from some constant $c>0$}.
-} Then, \textcolor{red}{as $p \rightarrow \infty$}, for all $i,j$ such that $z(i) \neq z(j)$, we have
+\begin{lem}[Angle gaps in $\tX$ and $\tS$]\label{lem:angle_gap_x} Consider the dTBM model~\eqref{eq:model_tensor} under the parameter space $\tP$ in \eqref{eq:family}.
+Suppose Assumption~\ref{assmp:min_gap} holds and $\mtheta$ is balanced satisfying~\eqref{eq:degree}.
+Then, for all $i,j$ such that $z(i) \neq z(j)$, we have
 \begin{equation}
@@
 \end{equation}
-where \textcolor{red}{$m \in [-1,1]$} is a scalar parameter controlling the skewness of degrees. Let $\Delta_{\mX}^2$ denote the minimal angle gap of the mean tensor, defined by
+where $m\geq 0$ is a scalar parameter controlling the skewness of degrees. Let $\Delta_{\mX}^2$ denote the minimal angle gap of the mean tensor, defined by
 \begin{equation}\label{eq:delta_x}
@@
 \end{equation}
-% {
-% \color{red}
-% Take $ a = c$ for some positive constant $c$ in the model setup~\eqref{eq:modelexample}. We have 
-% \begin{equation}
-%     \Delta_{\min}^2 = \frac{2 a^2}{1 + a^2} \asymp 1, \quad  \Delta_{\mX}^2 = \frac{2\onormSize{}{\mtheta_{z^{-1}(2)}}^{2} a^2}{\onormSize{}{\mtheta_{z^{-1}(1)}}^{2} +  \onormSize{}{\mtheta_{z^{-1}(2)}}^{2}a^2} \asymp p^{-m}.
-% \end{equation}
-
-% The notion $\Delta_{\min}^2$ achieves the maximum $\tO(1)$ while the the dTBM is still impossible when $m = -1$, i.e., $\Delta^2_{\mX} \lesssim p^{-1}$, based on our theory in later Sections.
-That is, the dTBM estimation depends on the magnitude of $m$.
-In such a setting, the proposed signal notion $\Delta^2_{\min}$ alone fails to fully characterize dTBM, and the phase transition of $\Delta_{\min}^2$ can be arbitrarily worse than that with balance assumption.
-% }
 Take $ a = p^{-1/4}$ in the model setup~\eqref{eq:modelexample}. We have 
@@
 %Under the balanced degree assumption, the dTBM is computationally solvable based on our Theorem~\ref{thm:comp} in Section~\ref{sec:complimit}.
-{\color{red}Based on our theory in later Sections, the dTBM is impossible to solve when $\Delta^2_{\mX} \lesssim p^{-1}$; 
-}
-%Based on our theory in later Sections, the stat-computational achievability of dTBM occurs when $\Delta^2_{\mX}>p^{-1}$; 
-that is, the dTBM estimation depends on the relative magnitude of $m$ vs. $1/2$. In such a setting, the proposed signal notion $\Delta^2_{\min}$ alone fails to fully characterize dTBM. 
-\textcolor{red}{}
-
+Based on our theory in later Sections, the stat-computational achievability of dTBM occurs when $\Delta^2_{\mX}>p^{-1}$; that is, the dTBM estimation depends on the relative magnitude of $m$ vs.
+$1/2$.
+In such a setting, the proposed signal notion $\Delta^2_{\min}$ alone fails to fully characterize dTBM.
 %However, the condition $ \Delta_{\min}^2 = p^{-1/2}$ with balanced degree leads to a computationally efficient dTBM by Theorem~\ref{thm:refinement}.
@@
 "
-In \cite{gao2018community}, the degree distribution is restricted to ${1\over |z^{-1}(a)|}\sum_{i\in z^{-1}(a)}\theta_i=1+o(1)$ for all communities. 
+In \cite{gao2018community}, the degree distribution is restricted to ${1\over |z^{-1}(a)|}\sum_{i\in z^{-1}(a)}\theta_i=1+o(1)$ for all communities. ''
 \end{rmk}
@@
 \begin{itemize}[wide]
-    \item \textbf{Impossibility.} \textcolor{red}{Assume $p \rightarrow \infty$ and $2 \leq r\lesssim p^{1/3}$.} If the signal exponent satisfies $\gamma < -(K-1)$, then, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation; that is,
-\begin{equation}
-  \gamma< -(K-1) \quad \Rightarrow  \inf_{\hat z_{\text{stat}}}\sup_{(z, \tS, \mtheta)\in \textcolor{red}{\tP(\gamma)}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
-\end{equation}
-\textcolor{red}{The impossibility of MLE also holds for the general Gaussian dTBMs with parameters $(z, \mtheta, \tS) \in \tP$ such that $\Delta_{\mX}^2 \lesssim p^{-(K-1)}$, where $\Delta_{\mX}^2$ is the minimal angle in mean tensor defined in \eqref{eq:delta_x}.}
-
-\item \textbf{MLE achievability.} Suppose the signal exponent satisfies $\gamma >-(K-1)+c_0$ for an arbitrary constant \textcolor{red}{$c_0>0$}.
-Furthermore, assume \textcolor{red}{$\mtheta$} is balanced and $\min_{i\in[p]}\theta(i)\geq c$ from some constant $c>0$.
-Then, \textcolor{red}{when $p \rightarrow \infty$, fixed $r \geq 1$, and $K \geq 3$, }the MLE in~\eqref{eq:mle} achieves exact recovery in high probability; that is,
+    \item \textbf{Impossibility.} 
+Assume $r\lesssim p^{1/3}$. If the signal exponent satisfies $\gamma < -(K-1)$, then, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation; that is,
+\begin{equation}
+  \gamma< -(K-1) \quad \Rightarrow  \inf_{\hat z_{\text{stat}}}\sup_{(z, \tS, \mtheta)\in\tP(\lambda)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
+\end{equation}
+\item \textbf{MLE achievability.} Suppose the signal exponent satisfies $\gamma >-(K-1)+c$ for an arbitrary constant $c>0$.
+Furthermore, assume $\theta$ is balanced and $\min_{i\in[p]}\theta(i)\geq c$ from some constant $c>0$.
+Then, the MLE in~\eqref{eq:mle} achieves exact recovery in high probability; that is,
 \begin{equation}
@@
 
-The second part of Theorem~\ref{thm:stats} shows the exact recovery of MLE when the $\gamma>-(K-1)+c_0$ for an arbitrary constant \textcolor{red}{$c_0>0$}.
+The second part of Theorem~\ref{thm:stats} shows the exact recovery of MLE when the $\gamma>-(K-1)+c$ for an arbitrary constant $c>0$.
 Combining the impossibility and achievability results, we conclude that the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM with respect to our SNR.
@@
 \begin{itemize}[wide]
-    \item \textbf{Impossibility.} Assume HPC conjecture holds \textcolor{red}{and $r \geq 2$}.
+    \item \textbf{Impossibility.} Assume HPC conjecture holds.
 If the signal exponent satisfies $\gamma < -K/2$, then, no \emph{polynomial-time estimator} $\hat z_{\text{comp}}$ achieves exact recovery in expectation as $p\to \infty$; that is,
@@
 \end{align}
-\item \textbf{Polynomial-time algorithm achievability.} Suppose the parameter space satisfies $\gamma  >  -K/2+c_0$ for an arbitrary constant \textcolor{red}{$c_0>0$}.
-Furthermore, assume \textcolor{red}{fixed $r \geq 1, K \geq 2$,} the degree is balanced, lower bounded in that $\min_{i\in[p]}\theta_i\geq c$ for some constant $c>0$, and satisfies the local linear stability in Definition~\ref{def:stable} \textcolor{red}{in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \geq \check{C}\log^{-1}p $ with some positive constant $\check{C}$.}
-%in the neighborhood $\tN(z,\varepsilon)$ for $\varepsilon \geq \log^{-1}(p)$. 
-Then, \textcolor{red}{as $p \rightarrow \infty$}, there exists a polynomial-time algorithm $\hat z_{\text{ploy}}$ that achieves exact recovery in high probability; that is,
+\item \textbf{Polynomial-time algorithm achievability.} Suppose the parameter space satisfies $\gamma  >  -K/2+c$ for an arbitrary constant $c>0$.
+Furthermore, assume the degree is balanced, lower bounded in that $\min_{i\in[p]}\theta_i\geq c$ for some constant $c>0$, and satisfies the local linear stability in Definition~\ref{def:stable} in the neighborhood $\tN(z,\varepsilon)$ for $\varepsilon \geq \log^{-1}(p)$.
+Then, there exists a polynomial-time algorithm $\hat z_{\text{ploy}}$ that achieves exact recovery in high probability; that is,
 \begin{equation}
@@
 The proofs for the two parts in Theorem~\ref{thm:comp} are in the Appendix II, Section~\ref{sec:compprove1} and Section~\ref{sec:statprove2}, respectively.
-The first part of Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$, and the second part shows the existence of such algorithm when $\gamma > -K/2+c_0$ \textcolor{red}{for an arbitrary constant $c_0 > 0$} under extra technical assumptions.
+The first part of Theorem~\ref{thm:comp} indicates the impossibility of exact recovery by polynomial-time algorithms when $\gamma < -K/2$, and the second part shows the existence of such algorithm when $\gamma > -K/2+c$ under extra technical assumptions.
 In Section~\ref{sec:alg}, we will present an efficient polynomial-time algorithm in this setting.
@@
 
-\begin{lem}[Misclustering error and loss]\label{lem:mis}  \textcolor{red}{Consider the dTBM under the parameter space $\tP$.} Suppose $\min_{i \in [p]}\theta(i) > c$ for some constant $c > 0$.
-We have $\ell(\hat z, z) \Delta_{\min}^2 \leq L(\hat z, z)$.
+\begin{lem}[Misclustering error and loss]\label{lem:mis}Suppose $\min_{i \in [p]}\theta(i) > c$ for some $c > 0$. We have $\ell(\hat z, z) \Delta_{\min}^2 \leq L(\hat z, z)$.
 %\fixme{miaoyan}{under bounded degree assumptions? please also check all lemmas in supp to make sure conditions are well specified.}
@@
 \color{blue}
-\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with \textcolor{red}{fixed $r \geq 1$, $K \geq 2$}, i.i.d.\ noise under the parameter space $\tP$, and Assumption~\ref{assmp:min_gap}.
+\begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with i.i.d.\ noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
 Assume $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$ and let $\Delta_{\mX}$ denote the minimal gap in mean tensor defined in~\eqref{eq:delta_x}.
 Let $ z^{(0)}_k$ denote the output of Sub-algorithm~\hyperref[alg:main]{1}.
-With probability going to 1 \textcolor{red}{as $p \rightarrow \infty$}, we have
+With probability going to 1, we have
 \begin{equation}
@@
 \begin{equation}\label{eq:ini}
- \ell(z_k^{(0)}, z) \lesssim {r^K p^{-K/2}\over \text{SNR}} \quad \text{and} \quad L(z^{(0)}_k, z) \lesssim  {\sigma^2 r^K p^{-K/2}},
-\end{equation}
-\textcolor{red}{with probability going to 1 as $p \rightarrow \infty$.}
+ \ell(z_k^{(0)}, z) \lesssim {r^K p^{-K/2}\over \text{SNR}} \quad \text{and} \quad L(z^{(0)}_k, z) \lesssim  {\sigma^2 r^K p^{-K/2}}.
+\end{equation}
 \end{thm}
@@
 
-\begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the general sub-Gaussian dTBM with \textcolor{red}{fixed $r \geq 1$, $K \geq 2$}, independent noise under the parameter space $\tP$, and Assumption~\ref{assmp:min_gap}.
-\textcolor{red}{Assume the local linear stability of degree holds in the neighborhood $\tN(z,\varepsilon)$ for all $\varepsilon \leq E_0$ and some $E_0 \geq \check{C}\log^{-1}p $ with some positive constant $\check{C}$.}
-%Assume the local linear stability of degree holds in the neighborhood $\tN(z, \epsilon)$ for $\epsilon \geq \log^{-1}(p)$.
+\begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the general sub-Gaussian dTBM with independent noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
+Assume the local linear stability of degree holds in the neighborhood $\tN(z, \epsilon)$ for $\epsilon \geq \log^{-1}(p)$.
 Let $\{z^{(0)}_k\}_{k=1}^K$ be the initialization for Sub-Algorithm~\hyperref[alg:main]{2} and $z^{(t)}_k$ be the $t$-th iteration output on $k$-th mode.
-Suppose $\min_{i \in [p]}\theta(i) \geq c $ for some \textcolor{red}{constant} $c > 0$, the $\text{SNR} \geq \tilde C p^{-(K-1)}\log p$ for some sufficiently large constant $\tilde C$, and the initialization satisfies
+Suppose $\min_{i \in [p]}\theta(i) \geq c $ for some $c > 0$, the $\text{SNR} \geq \tilde C p^{-(K-1)}\log p$ for some sufficiently large constant $\tilde C$, and the initialization satisfies
 \begin{equation}
@@
 \end{equation}
- With probability going to 1 \textcolor{red}{as $p \rightarrow \infty$}, there exists a contraction parameter $\rho \in (0,1)$ such that 
+ With probability going to 1, there exists a contraction parameter $\rho \in (0,1)$ such that 
 \begin{align}\label{eq:final}
@@
 Assume $\text{SNR} \gtrsim p^{-K/2} \log p$.
-Combining \textcolor{red}{all parameter assumptions and the results in} Theorems~\ref{thm:initial} and \ref{thm:refinement}, with probability going to 1 \textcolor{red}{as $p \rightarrow \infty$}, our estimate $z^{(T)}_k$ achieves exact recovery within polynomial iterations; more precisely,
+Combining Theorems~\ref{thm:initial} and \ref{thm:refinement}, with probability going to 1, our estimate $z^{(T)}_k$ achieves exact recovery within polynomial iterations; more precisely,
 \begin{equation}
@@
 
-\begin{prop}[Error for Bernoulli initialization]\label{cor:ber} Consider the Bernoulli dTBM in the parameter space $\tP$ \textcolor{red}{with fixed $r \geq 1, K \geq 2$} and Assumption~\ref{assmp:min_gap} holds.
+\begin{prop}[Error for Bernoulli initialization]\label{cor:ber} Consider the Bernoulli dTBM in the parameter space $\tP$ and Assumption~\ref{assmp:min_gap} holds.
 Assume that $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$.
 Let $ z^{(0)}$ denote the output of Sub-algorithm~3.
-With probability going to 1 \textcolor{red}{as $p \rightarrow \infty$}, we have
+With probability going to 1, we have
 \begin{equation}\label{eq:ini_b}
@@
 
-\begin{rmk}[Comparison with previous methods] The Bernoulli bound $\tO(p^{- \floor{K/2} })$ in Proposition~\ref{cor:ber} is relatively looser than the Gaussian bound $\tO(p^{- K/2 })$ in Theorem~\ref{thm:initial}.
+\begin{rmk}[Comparison with previous methods] The Bernoulli bound $\tO(p^{- \floor{K/2} })$ in Proposition~\ref{cor:ber} is relatively looser than the Gaussian bound \tO(p^{- K/2 })$ in Theorem~\ref{thm:initial}.
 The gap between Bernoulli and Gaussian error decreases as the order $K$ increases.
@@
 We develop polar-coordinate based techniques to establish the error rate for the proposed algorithm.
+"
 \end{itemize}
@@
 The algorithm and oracle estimators have no gap in the matrix case, while an error gap emerges between the critical values $\gamma_{\text{stat}} = -2$ and $\gamma_{\text{comp}} = -1.5$ in the tensor case.
-Figure~\ref{fig:phase_binary} suggests the statistical-computational gap in Bernoulli models.
+Figure~\ref{fig:phase} suggests the statistical-computational gap in Bernoulli models.
 
@@
     \item[$(\Leftarrow)$] For the necessity, it suffices to construct two distinct parameters up to cluster label permutation, if the model~\eqref{eq:general_dtbm} violates Assumption~\ref{assmp:general_minimal_gap}.
-\textcolor{red}{Note that $\Delta_{\min}^2 = 1$ when there exists a $k \in [K]$ such that $r_k = 1$.
-Hence, we consider the case that  $r_k \geq 2$ for all $k \in [K]$.} Without loss of generality, we assume $\onorm{ \mS_{1,1:}^s - \mS_{1,2:}^s } = 0$.
+Without loss of generality, we assume $\onorm{ \mS_{1,1:}^s - \mS_{1,2:}^s } = 0$.
 
@@
 First, we show the uniqueness of $\mM_k$ for all $k \in [K]$.
-\textcolor{red}{When $r_k = 1$, all possible $\mM_k$ is equal to the vector ${\bf 1}_{p_k}$, and the uniqueness holds trivially.
-Hence, we consider the case that $r_k \geq 2$.
-Without loss of generality, we consider $k=1$ with $r_1 \geq 2$ and show the uniqueness of the first mode membership matrix;}  i.e., $\mM'_1 = \mM_1\mP_1$ where $\mP_1$ is a permutation matrix.
+Without loss of generality, we consider $k=1$ and show the first mode membership matrix; i.e., $\mM'_1 = \mM_1\mP_1$ where $\mP_1$ is a permutation matrix.
 The conclusion for $k\geq 2$ can be showed similarly and thus omitted.
@@
 
-\begin{lem}[Motivation of angle-based clustering]\label{lem:angle} Consider the signal tensor $\tX$ in the general asymmetric dTBM~\eqref{eq:general_dtbm} with $(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\{r_k\})$, $r_k \geq 2$ \textcolor{red}{ and $K \geq 2$}.
+\begin{lem}[Motivation of angle-based clustering]\label{lem:angle} Consider the signal tensor $\tX$ in the general asymmetric dTBM~\eqref{eq:general_dtbm} with $(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\{r_k\})$ and $r_k \geq 2$.
 Then, for any $k \in [K]$ and index pair $(i,j)\in[p_k]^2$, we have
@@
 
-{
-\color{red} Then, we have 
-\begin{equation}
-    \cos (\mS_{z_1(i):}, \mS_{z_1(j):}) = \frac{ \ang{  \mS_{z_1(i):}, \mS_{z_1(j):} }}{ \onormSize{}{\mS_{z_1(i):}} \onormSize{}{\mS_{z_1(j):}} }  = (1+o(1))\frac{ \ang{  \mX_{i:}, \mX_{j:} }}{ \onormSize{}{ \mX_{i:}} \onormSize{}{ \mX_{j:}} } = (1 + o(1)) \cos(\mX_{i:}, \mX_{j:}),
-\end{equation}
+Notice that $\onormSize{}{\mv_1^s - \mv_2^s}^2 =  2(1 - \cos(\mv_1, \mv_2))$ for two vectors $\mv_1, \mv_2$ of the same dimension. We have 
+\begin{align}
+    \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s}  &= 2 - 2 \frac{ \ang{  \mS_{z_1(i):}, \mS_{z_1(j):} }}{ \onormSize{}{\mS_{z_1(i):}} \onormSize{}{\mS_{z_1(j):}} } \\
+    & = 2 - 2 (1+o(1))\frac{ \ang{  \mX_{i:}, \mX_{j:} }}{ \onormSize{}{ \mX_{i:}} \onormSize{}{ \mX_{j:}} } \\
+    & = (1 + o(1)) \onormSize{}{\mX_{i:}^s - \mX_{j:}^s},
+\end{align}
 where the second inequality follows by the balance assumption on $\mtheta$, and the last inequality holds when $p$ is large enough.
 
-Further, notice that $\onormSize{}{\mv_1^s - \mv_2^s}^2 =  2(1 - \cos(\mv_1, \mv_2))$. For all $i,j$ such that $z(i) \neq z(j)$, when $p \rightarrow \infty$, we have
+Hence, we conclude that for all $i,j$ such that $z(i) \neq z(j)$,
 \begin{equation}
@@
 \end{equation}
-
-}
-
-% Notice that $\onormSize{}{\mv_1^s - \mv_2^s}^2 =  2(1 - \cos(\mv_1, \mv_2))$ for two vectors $\mv_1, \mv_2$ of the same dimension. We have 
-% \begin{align}
-%     \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s}  &= 2 - 2 \frac{ \ang{  \mS_{z_1(i):}, \mS_{z_1(j):} }}{ \onormSize{}{\mS_{z_1(i):}} \onormSize{}{\mS_{z_1(j):}} } \\
-%     & = 2 - 2 (1+o(1))\frac{ \ang{  \mX_{i:}, \mX_{j:} }}{ \onormSize{}{ \mX_{i:}} \onormSize{}{ \mX_{j:}} } \\
-%     & = (1 + o(1)) \onormSize{}{\mX_{i:}^s - \mX_{j:}^s},
-% \end{align}
-% where the second inequality follows by the balance assumption on $\mtheta$, and the last inequality holds when $p$ is large enough.
-
-% Hence, we conclude that for all $i,j$ such that $z(i) \neq z(j)$,
-% \begin{equation}
-%     \onormSize{}{\mX_{i:}^s - \mX_{j:}^s} \asymp \onormSize{}{\mS_{z_1(i):}^s - \mS_{z_1(j):}^s} \gtrsim  \Delta_{\min}.
-% \end{equation}
 
@@
 Consider the general asymetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$. Specifically, we will show that, under 
-the assumptions $K\geq 1$, \textcolor{red}{$2 \leq r\lesssim p^{1/3}$ as 
-$p \rightarrow \infty$}, and SNR condition
+the assumptions $K\geq 1$, $r\lesssim p^{1/3}$ and SNR condition
 \[
@@
 {
-\color{red}
-
-The same conclusion holds after replacing the condition (i) and (ii) by  $\Delta_{\mX}^2 \lesssim p^{-(K-1)}$ and constructing the sets $T$ as \citet[Proof of Theorem 6]{han2020exact}.
+\color{blue}
+
+%When the parameter satisfies condition (a), we obtain the same conclusion by replacing the condition (i) and (ii) by  $\Delta_{\mX}^2 \leq p^{-(K-1)}$ and constructing the sets $T$ as \citet[Proof of Theorem 6]{han2020exact}.
 
@@
 \begin{proof}[Proof of Theorem~\ref{thm:comp} (Impossibility)]
-The idea of proving computational hardness is to show the computational lower bound for a special class of degree-corrected tensor clustering model with $K\geq 2$ \textcolor{red}{and $ r \geq 2$}.
+The idea of proving computational hardness is to show the computational lower bound for a special class of degree-corrected tensor clustering model with $K\geq 2$.
 We construct the following special class of higher-order degree-corrected tensor clustering  model.
@@
 \subsection{Proof of Theorem~\ref{thm:initial}}
-\begin{proof}[Proof of Theorem~\ref{thm:initial}] We prove Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with symmetric mean tensor, parameters $(z, \tS, \mtheta)$, \textcolor{red}{fixed $r\geq 1, K \geq 2$}, and i.i.d.\ noise.
-\textcolor{red}{For the case $r = 1$, we have $L(z^{(0)},z) = 0, \ell(z^{(0)}, z) = 0$ trivially.
-Hence, we focus on the proof of the first mode clustering $z^{(0)}_1$ with $r \geq 2$}; the proofs for the other modes can be extended similarly.
+\begin{proof}[Proof of Theorem~\ref{thm:initial}] We prove Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with symmetric mean tensor and parameters $(z, \tS, \mtheta)$ and i.i.d.\ noise.
+We focus on the proof of the first mode clustering $z^{(0)}_1$; the proofs for the other modes can be extended similarly.
 We drop the subscript $k$ in the matricizations $\mM_k, \mX_k, \mS_k$ and in the estimate $z^{(0)}_1$.
@@
 
-\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the parameter space~\eqref{eq:family} and assumption that $\min_{i \in [p]} \theta(i) \geq c$ \textcolor{red}{for some constant $c >0$}, the singular values of $\mTheta \mM$ are bounded as
+\begin{lem}[Singular value of weighted membership matrix]\label{lem:singular_thetam} Under the parameter space~\eqref{eq:family} and assumption that $\min_{i \in [p]} \theta(i) \geq c$, the singular values of $\mTheta \mM$ are bounded as
 \begin{equation}\label{eq:membership2}
@@
 
-\begin{lem}[Upper bound of misclustering error]\label{lem:upper_mis} Let $z: [p] \mapsto [r]$ be a cluster assignment such that $|z^{-1}(a)| \asymp p/r$ for all $a \in [r]$ \textcolor{red}{with $r \geq 2$}.
+\begin{lem}[Upper bound of misclustering error]\label{lem:upper_mis} Let $z: [p] \mapsto [r]$ be a cluster assignment such that $|z^{-1}(a)| \asymp p/r$ for all $a \in [r]$.
 Let node $i$ correspond to a vector $\mx_i  = \theta(i) \mv_{z(i)} \in \bbR^d$, where $\{\mv_a\}_{a = 1}^r$ are the cluster centers and $\mtheta = \entry{\theta(i)} \in \bbR^p_+$ is the positive degree heterogeneity.
@@
 \end{equation}
-where $S_0$ is defined in \textcolor{red}{Step 4} of Sub-algorithm~\hyperref[alg:main]{1} and
+where $S_0$ is defined in Step 3 of Sub-algorithm~\hyperref[alg:main]{1} and
 \begin{equation}
@@
 
-%We prove Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with symmetric mean tensor, parameters $(z, \tS, \mtheta)$, \textcolor{red}{fixed $r\geq 1, K \geq 2$}, and i.i.d.\ noise.
-\textcolor{red}{For the case $r = 1$, we have $L(z^{(0)},z) = 0, \ell(z^{(0)}, z) = 0$ trivially.
-Hence, we focus on the proof of the first mode clustering $z^{(0)}_1$ with $r \geq 2$};
-
-We consider dTBM~\eqref{eq:model_tensor} with symmetric mean tensor, parameters $(z, \tS, \mtheta)$, \textcolor{red}{fixed $r\geq 1, K \geq 2$}, and i.i.d.\ noise.
+We consider dTBM~\eqref{eq:model_tensor} with symmetric mean tensor and parameters $(z, \tS, \mtheta)$.
 Let $(\hat z, \hat \mtheta, \hat \tS)$ denote the MLE in \eqref{eq:mle}, and $(z^{(0)}_k, \mtheta^{(0)}, \tS^{(0)})$ denote parameters related to the initialization.
-\textcolor{red}{For the case $r = 1$, we have $\ell(z^{(t)}, z) = 0$ trivially for all $t \geq 0$.
-Hence, we focus on the proof of the first mode clustering $z^{(t+1)}_1$ with $r \geq 2$}; the extension for other modes can be obtained similarly.
+We focus on the proof of $z^{(t+1)}_1$; the extension for other modes can be obtained similarly.
 We drop the subscript $k$ in the matricizations $\mTheta, \mM_k, \mS_k, \mX_k$ and in estimates $z^{(0)}, z^{(t+1)}_1, z^{(t)}_k$ for ease of the notation.
@@
 \end{equation}
-where $M$ is a positive universal constant in inequality~\eqref{eq:constant_M}, $\bar C, \tilde C$ are universal constants in \textcolor{red}{the proof of Theorem~\ref{thm:initial}} and assumption SNR $\geq \tilde C p^{-K/2} \log p$, respectively.
+where $M$ is a positive universal constant in inequality~\eqref{eq:constant_M}, $\bar C, \tilde C$ are universal constants in Corollary~\ref{cor:L0} and assumption SNR $\geq \tilde C p^{-K/2} \log p$, respectively.
 Further, inequality~\eqref{eq:cond1} holds by replacing $\mV$ to $\mV^{(t)}, \mQ, \hat \mQ$ and $\mW_{:a}$ to $\mW_{:a}^{(t),T}, \mP_{:a}^T, \hat \mP_{:a}^T$ when initialization condition~\eqref{eq:cond_intial} holds.
@@
    \begin{equation}
-        \xi' \lesssim  \exp \of{  -  \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }.
+        \xi' \lesssim  \of{  -  \frac{p^{K-1}}{r^{K-1}}   \Delta_{\min}^2 }.
    \end{equation}
@@
 
-\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement} \textcolor{red}{with fixed $r \geq 2$},  assume the the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ is large enough to satisfy the inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
-\textcolor{red}{As $p \rightarrow \infty$}, we have
+\begin{lem}[Upper bound for $F_{ib}^{(t)}, G_{ib}^{(t)}$ and $H_{ib}^{(t)}$]\label{lem:upper_fgh} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement},  assume the the constant $\tilde C$ in the condition SNR $\geq \tilde C p^{-K/2} \log p$ is large enough to satisfy the inequalities~\eqref{eq:tilde_c1} and \eqref{eq:tilde_c2}.
+We have
     \begin{equation}\label{eq:fib}
@@
 
-     \begin{lem}[Relationship between misclustering loss and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement} \textcolor{red}{with fixed $r \geq 2$, as $p \rightarrow \infty$}, we have
+     \begin{lem}[Relationship between misclustering loss and intermediate parameters]\label{lem:intermediate} Under the Condition~\ref{cond:origin} and the setup of Theorem~\ref{thm:refinement}, we have
     \begin{equation}\label{eq:inter1}
@@
     \end{equation}
-     \textcolor{red}{Note that we have $\ell^{(t)} \leq \frac{L^{(t)}}{\Delta_{\min}^2} \leq \frac{\bar C}{\tilde C} r \log^{-1}(p) $ by Condition~\ref{cond:origin} and Lemma~\ref{lem:mis}.
-Then,  with the locally linear stability assumption, the $\mtheta$ is $\ell^{(t)}$-locally linearly stable; i.e.,}
-     %With the locally linear stability assumption, we have 
+    With the locally linear stability assumption, we have 
     \begin{equation}
@@
 
-\begin{lem}[Polynomial estimation error of MLE]\label{lem:poly_mle_degree}  Let $(\hat z, \hat \mtheta, \hat \tS)$ denote the MLE in \eqref{eq:mle} \textcolor{red}{with fixed $K \geq 2$}, and $\hat \tX$ denote the mean tensor consisting of parameter $(\hat z, \hat \mtheta, \hat \tS)$.
-\textcolor{red}{With high probability going to 1 as $p \rightarrow \infty$, we} have
+\begin{lem}[Polynomial estimation error of MLE]\label{lem:poly_mle_degree}  Let $(\hat z, \hat \mtheta, \hat \tS)$ denote the MLE in \eqref{eq:mle}, and $\hat \tX$ denote the mean tensor consisting of parameter $(\hat z, \hat \mtheta, \hat \tS)$.
+We have
 \begin{equation}
@@
 with probability going to 1. 
-When \textcolor{red}{  $\text{SNR} \gtrsim p^{-(K-1)} \log p$,  $\mtheta$ is balanced and $\min_{i \in [p]} \theta(i) \geq c$ for some constant $c$}, the MLE satisfies
+When $\Delta_{\min}^2 \gtrsim p^{-(K-1)} \log p$, the MLE satisfies
 \begin{equation}

### 2023-01-24 tensor_clustering/IEEE/ieee_revision/ieee_revise_s2.tex
organize

--- 2023-01-24 tensor_clustering/IEEE/ieee_revision/ieee_revise_s1.tex
@@
     \item  \textit{Tensor block model.} The (non-degree) tensor block model (TBM) is a higher-order extension of SBM, and its statistical-computational properties are investigated in recent literatures~\citep{wang2019multiway, han2020exact, ghoshdastidar2017consistency}.
-Extending results from non-degree to degree-corrected model is highly challenging.
+\textcolor{blue}{Some works~\citep{ahn2018hypergraph} study the TBM with sparse observations, while, others~\citep{wang2019multiway, han2020exact} and our work focus on the dense regime.} Extending results from non-degree to degree-corrected model is highly challenging.
 Our dTBM parameter space is equipped with angle-based similarity and nuisance degree parameters.
@@
 
+{ 
+\color{blue}
+
 \begin{table*}[t]
 \resizebox{\textwidth}{!}{%
-    \begin{tabular}{c|ccccc}
+    \begin{tabular}{c|cccccc}
 \hline
-    & \cite{gao2018community}&  \cite{han2020exact}&  \cite{ghoshdastidar2017consistency} &\cite{ke2019community} & \textbf{Ours}\\
+    & \cite{gao2018community}&  \cite{ahn2018hypergraph} &\cite{han2020exact}&  \cite{ghoshdastidar2017consistency} &\cite{ke2019community} & \textbf{Ours}\\
     \hline
-      Allow tensors of arbitrary order & $\times$ & $\surd$ & $\surd$ &$\surd$ & $\surd$  \\
-        Allow degree heterogeneity &  $\surd$ & $\times$ & $\surd$ &$\surd$ & $\surd$ \\
-        Singular-value gap-free clustering &  $\surd$ & $\surd$ & $\times$ & $\times$ &$\surd$ \\
-        Misclustering rate (for order $K^*$)& - & $\exp(-p^{K/2})$ & $p^{-1}$& $p^{-2}$ &  $\exp(-p^{K/2})$\\
+      Allow tensors of arbitrary order & $\times$ & $\surd$&  $\surd$ & $\surd$ &$\surd$ & $\surd$  \\
+        Allow degree heterogeneity &  $\surd$ & $\times$ & $\times$ & $\surd$ &$\surd$ & $\surd$ \\
+        Singular-value gap-free clustering &  $\surd$ & $\surd$ & $\surd$ & $\times$ & $\times$ &$\surd$ \\
+        Misclustering rate (for order $K^*$)& - & ${p^{-(K-1)} \alpha^{-1}}^{**}$ & $\exp(-p^{K/2})$ & $p^{-1}$& $p^{-2}$ &  $\exp(-p^{K/2})$\\
+        Consider sparse observation & $\times$ & $\surd$ & $\times$ & $\times$ & $\times$ &$\times$\\
         \hline
@@
     }
-    \caption{Comparison between previous methods with our method. $^*$We list the result for order-K tensors with $K \geq 3$ and general number of communities $r = \tO(1)$. }\label{tab:comp}
+    \caption{ \color{blue} Comparison between previous methods with our method.
+$^*$We list the result for order-K tensors with $K \geq 3$ and general number of communities $r = \tO(1)$.
+$^{**}$The parameter $\alpha = f(p) > 0$ denotes the sparsity level which is some function of dimension $p$.
+}\label{tab:comp}
 \end{table*}
+}
 
@@
 We use $\text{Ave}(\cdot)$ to denote the operation of taking averages across elements and $\text{Mat}_k(\cdot)$ to denote the unfolding operation that reshapes the tensor along mode $k$ into a matrix.
-\textcolor{blue}{For a symmetric tensor $\tX \in\mathbb{R}^{p\times \cdots \times p}$, we omit the subscript and use $\text{Mat}(\tX)\in\mathbb{R}^{p\times p^{K-1}}$ to denote the unfolding.}  For two sequences $\{a_p\}, \{b_p\}$, we denote $a_p\lesssim b_p$ or $a_p=\tO(b_p)$ if $\lim_{p\to\infty}a_p /b_p\leq c$ for some constant $c\geq 0$, $a_p=o(b_p)$ if $\lim_{p\to\infty}a_p/b_p =0$, and $a_p = \Omega(b_p)$ if both $b_p \lesssim a_p$ and $a_p\lesssim b_p$.
+{For a symmetric tensor $\tX \in\mathbb{R}^{p\times \cdots \times p}$, we omit the subscript and use $\text{Mat}(\tX)\in\mathbb{R}^{p\times p^{K-1}}$ to denote the unfolding.}  For two sequences $\{a_p\}, \{b_p\}$, we denote $a_p\lesssim b_p$ or $a_p=\tO(b_p)$ if $\lim_{p\to\infty}a_p /b_p\leq c$ for some constant $c\geq 0$, $a_p=o(b_p)$ if $\lim_{p\to\infty}a_p/b_p =0$, and $a_p = \Omega(b_p)$ if both $b_p \lesssim a_p$ and $a_p\lesssim b_p$.
 Throughout the paper, we use the terms ``community'' and ``clusters'' exchangeably.
@@
 In Section~\ref{sec:alg}, we provide a polynomial-time two-stage algorithm with misclustering rate guarantees.
-{\color{blue}Extension to Bernoulli models is also presented.
+{ Extension to Bernoulli models is also presented.
 In Section~\ref{sec:tbm}, we compare our work with non-degree tensor block models.} Numerical studies including the simulation, comparison with other methods, and two real dataset analyses are in Sections~\ref{sec:simulation}-\ref{sec:real}.
@@
 Note that the discrete nature of $\mM$ renders our model~\eqref{eq:model_tensor} more challenging than Tucker decomposition.
-We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$ if $\tY$ admits dTBM~\eqref{eq:model_tensor} \textcolor{blue}{and let $\tX = \bbE \tY$ denote the mean tensor.} The goal of clustering is to estimate $z$ from a single noisy tensor $\tY$.
+We call a tensor $\tY$ an $r$-block tensor with degree $\mtheta$ if $\tY$ admits dTBM~\eqref{eq:model_tensor}   {and let $\tX = \bbE \tY$ denote the mean tensor.} The goal of clustering is to estimate $z$ from a single noisy tensor $\tY$.
 We are particularly interested in the high-dimensional regime where $p$ grows whereas $r=\tO(1)$.
 
-\textcolor{blue}{For ease of notation, we have focused on the case with symmetric mean tensor $\bbE \tY$.
+  {For ease of notation, we have focused on the case with symmetric mean tensor $\bbE \tY$.
 This assumption simplifies the notation because all modes have the same $(\mTheta, \mM, z)$; the noise tensor $\tE$ and the data tensor $\tY$ are still possibly asymmetric.
@@
 The constants $c_1, c_2$ in the $|z^{-1}(a)|$ bound assume the roughly balanced size across $r$ communities.
-\textcolor{blue}{ Third, the constant $c_3$ requires no purely zero slide in $\tS$, and thus the model~\eqref{eq:model_tensor} is not trivially non-identifiable; see Example~\ref{example:c3} below.
+{ Third, the constant $c_3$ requires no purely zero slide in $\tS$, and thus the model~\eqref{eq:model_tensor} is not trivially non-identifiable; see Example~\ref{example:c3} below.
 The upper bound $c_4$ is a technical constraint to avoid the unbounded entries in the core tensor; in practice, the constraint $\max_{a:} \onorm{\mat(\tS)_{a:}} \leq c_4$ would likely never be active with a large $c_4 \geq \onormSize{}{\tY}_F$.
@@
 This constraint, again, incurs no restriction to model flexibility but makes our presentation cleaner.
-\textcolor{blue}{Our constraints in $\tP$ are mild compared with previous literature; see Table~\ref{tab:para} for comparison.}
+{Our constraints in $\tP$ are mild compared with previous literature; see Table~\ref{tab:para} for comparison.}
 
@@
  {
- \color{blue}
+
     \begin{example}[Non-identifiability with purely zero core slice]\label{example:c3} 
@@
 {
-\color{blue}In our algorithmic development, we often refer to the regime of balanced degree heterogeneity. We call the degree $\mtheta$ \emph{balanced} if
+ In our algorithmic development, we often refer to the regime of balanced degree heterogeneity. We call the degree $\mtheta$ \emph{balanced} if
 \begin{equation}\label{eq:degree}
@@
 \subsection{Statistical critical value}\label{sec:statlimit}
-The statistical critical value means the SNR required for solving dTBMs with \emph{unlimited computational cost.} \textcolor{blue}{Our following result shows the minimax lower bound for exact recovery and the matching upper bound for maximum likelihood estimator (MLE).
+The statistical critical value means the SNR required for solving dTBMs with \emph{unlimited computational cost.}   {Our following result shows the minimax lower bound for exact recovery and the matching upper bound for maximum likelihood estimator (MLE).
 }
 {
-\color{blue} We consider the Gaussian MLE, denoted as $(\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}})$, over the estimation space $\tP(\gamma)$, where 
+\color{blue}
+  We consider the Gaussian MLE, denoted as $(\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}})$, over the estimation space $\tP$, where 
 \begin{equation}\label{eq:mle}
-    (\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}}) = \argmin_{ (z, \mtheta, \tS) \in \tP (\gamma)} \onormSize{}{\tY - \tX(z,\mtheta, \tS)}_F^2.
-\end{equation}
+    (\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}}) = \argmin_{ (z, \mtheta, \tS) \in \tP} \onormSize{}{\tY - \tX(z,\mtheta, \tS)}_F^2.
+\end{equation}
+}
 
@@
 \begin{itemize}[wide]
-    \item \textbf{Impossibility.} 
-Assume $r\lesssim p^{1/3}$. If the signal exponent satisfies $\gamma < -(K-1)$, then, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation; that is,
-\begin{equation}
-  \gamma< -(K-1) \quad \Rightarrow  \inf_{\hat z_{\text{stat}}}\sup_{(z, \tS, \mtheta)\in\tP(\lambda)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
-\end{equation}
+    \item { \color{blue}
+    \textbf{Impossibility.}  Assume $r\lesssim p^{1/3}$.
+Let $\tP_{\tS}(\gamma) \coloneqq \{ \tS: c_3\leq \onormSize{}{\text{Mat}(\tS)_{a:}}\leq c_4, a\in[r]\} \cap \{ \tS: \Delta_{\min}^2 = p^{\gamma}\}$ denote the space for valid $\tS$ satisfying SNR condition~\eqref{eq:gamma}, and $\tP_{z ,\mtheta} \coloneqq \{ \mtheta\in\mathbb{R}^p_{+},\
+{c_1p\over r}\leq |z^{-1}(a)|\leq {c_2 p\over r}, \onorm{\mtheta_{z^{-1}(a)}}_1=|z^{-1}(a)|, a \in [r] \}$ denote the space for valid $(z, \mtheta)$,  where $c_1, c_2, c_3, c_4$ are the constants in parameter space \eqref{eq:family}.
+If the signal exponent satisfies $\gamma < -(K-1)$, then, for any true core tensor $\tS \in \tP_{\tS}(\gamma)$, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation; that is,
+    \begin{equation}\label{eq:minminmax}
+  \gamma< -(K-1) \quad \Rightarrow  \quad \liminf_{p \rightarrow \infty}\inf_{\tS \in  \tP_{\tS}(\gamma)}  \inf_{\hat z_{\text{stat}}}\sup_{ (z, \mtheta) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
+\end{equation}
+    }
+
+%     \textbf{Impossibility.} 
+% Assume $r\lesssim p^{1/3}$. If the signal exponent satisfies $\gamma < -(K-1)$, then, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation; that is,
+% \begin{equation}
+%   \gamma< -(K-1) \quad \Rightarrow  \inf_{\hat z_{\text{stat}}}\sup_{(z, \tS, \mtheta)\in\tP(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
+% \end{equation}
 \item \textbf{MLE achievability.} Suppose the signal exponent satisfies $\gamma >-(K-1)+c$ for an arbitrary constant $c>0$.
@@
 
+{
+\color{blue} 
 The proofs for the two parts in Theorem~\ref{thm:stats} are in the Appendix II, Section~\ref{sec:statprove1} and Section~\ref{sec:statprove2}, respectively.
+The first part of Theorem~\ref{thm:stats} demonstrates impossibility of exact recovery whenever the core tensor $\tS$ satisfies SNR condition~\eqref{eq:gamma} with exponent $\gamma < -(K-1)$.
+The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to MLE and trace maximization~\citep{ghoshdastidar2017uniform}.
+The minimax bound~\eqref{eq:minminmax} indicates the worst case impossibility for a particular core tensor $\tS$ with signal exponent $\gamma < -(K-1)$; i.e., under the assumptions of Theorem~\ref{thm:stats},
+\begin{equation}
+    \gamma < -(K-1) \quad \Rightarrow \quad \liminf_{p \rightarrow \infty}  \inf_{\hat z_{\text{stat}}}\sup_{ (z,\tS, \mtheta) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
+\end{equation}
+Such worst case impossibility is frequently studied in related works \citep{han2020exact, gao2018community} while our lower bound~\eqref{eq:minminmax} provides a stronger impossibility statement for arbitrary core tensor with weak signal.
+
+%The worst case impossibility is weaker than our impossibility~\eqref{eq:minminmax} for arbitrary core tensor with weak signal.  
+%Inspection of the proof shows that the impossibility holds with or without the balanced degree assumption.
+%Moreover, with general degree, Theorem~\ref{thm:stats} also indicates the impossibility when the mean tensor gap is below the threshold $p^{-(K-1)}$.
+
+}
+
+% The proofs for the two parts in Theorem~\ref{thm:stats} are in the Appendix II, Section~\ref{sec:statprove1} and Section~\ref{sec:statprove2}, respectively.
 The first part of Theorem~\ref{thm:stats} demonstrates the impossibility of exact recovery when $\gamma < -(K-1)$.
@@
 %Moreover, with general degree, Theorem~\ref{thm:stats} also indicates the impossibility when the mean tensor gap is below the threshold $p^{-(K-1)}$.
-The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to MLE and trace maximization~\citep{ghoshdastidar2017uniform}.
+% The proof is information-theoretical, and therefore the results apply to all statistical estimators, including but not limited to MLE and trace maximization~\citep{ghoshdastidar2017uniform}.
 
@@
 Combining the impossibility and achievability results, we conclude that the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM with respect to our SNR.
-}
 
@@
 
-{
-\color{blue}
 % We consider the MLE of Gaussian dTBM under mild assumptions on the heterogeneity.
@@
 Hence, the boundary $\gamma_{\text{stat}} \coloneqq -(K-1)$ is the critical value for statistical performance of dTBM.
-
-}
 
@@
 {
-\color{blue}
+
 \begin{thm}[Computational critical value]\label{thm:comp} 
@@
 {
-\color{blue}
+
 The proofs for the two parts in Theorem~\ref{thm:comp} are in the Appendix II, Section~\ref{sec:compprove1} and Section~\ref{sec:statprove2}, respectively.
@@
 In what follows, we first use the symmetric tensor as a working example to describe the algorithm procedures to gain insight.
-\textcolor{blue}{Our theoretical analysis focuses on the symmetric mean tensor with independent sub-Gaussian noises}
+{Our theoretical analysis focuses on the symmetric mean tensor with independent sub-Gaussian noises}
 %i.i.d.\ sub-Gaussian noise
-such as Gaussian and uniform observations.
-\textcolor{blue}{The extensions for Bernoulli observations, asymmetric mean tensors, and other practical issues are in Sections~\ref{subsec:ber} and \ref{subsec:exten}.}
+such as Gaussian and uniform observations.   {The extensions for Bernoulli observations, asymmetric mean tensors, and other practical issues are in Sections~\ref{subsec:ber} and \ref{subsec:exten}.}
 
@@
 {
-\color{blue}
 
@@
 We start with weighted higher-order clustering algorithm as initialization.
-\textcolor{blue}{We take an order-3 tensor and the clustering on the first mode as illustration for insight.} Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
+{We take an order-3 tensor and the clustering on the first mode as illustration for insight.} Consider noiseless case with $\tX = \bbE\tY$ and $\mX = \text{Mat}(\tX)$.
 By model~\eqref{eq:model_tensor}, for all $i \in [p]$, we have
@@
 \State {
-\color{blue}Compute factor matrix $ \mU_{\text{pre},k} = \text{SVD}_{r} (\text{Mat}_k(\tY)), k \in [K]$ and the $(K-1)$-mode projection 
+  Compute factor matrix $ \mU_{\text{pre},k} = \text{SVD}_{r} (\text{Mat}_k(\tY)), k \in [K]$ and the $(K-1)$-mode projection 
 \begin{equation}
@@
 \State {
-\color{blue}
+
 Compute factor matrix $\hat \mU_k = \text{SVD}_{r}(\text{Mat}_k(\tX_{\text{pre},k})), k \in [K]$ and denoised tensor
@@
 } 
-\For{\textcolor{blue}{$k \in [K]$}}
-\State { \color{blue} Let $\hat \mX = \text{Mat}_k(\hat \tX)$ and $S_0=\{i \in [p]: \onormSize{}{\hat \mX_{i:}} = 0\}$. Set $\hat z(i)$ randomly in $[r]$ for $i \in S_0$.}
+\For{  {$k \in [K]$}}
+\State {    Let $\hat \mX = \text{Mat}_k(\hat \tX)$ and $S_0=\{i \in [p]: \onormSize{}{\hat \mX_{i:}} = 0\}$. Set $\hat z(i)$ randomly in $[r]$ for $i \in S_0$.}
 \State{  For all $i\in S_0^c$, compute normalized rows
@@
 
-\OUTPUT {\color{blue} Initial clustering $z^{(0)}_k \leftarrow \hat z_k, k \in [K]$.}
+\OUTPUT {   Initial clustering $z^{(0)}_k \leftarrow \hat z_k, k \in [K]$.}
 
@@
 \For{ $k \in [K]$}
-\State {\color{blue} Calculate reduced tensor $\tY^{\text{d}}_k \in \bbR^{r \times r \times p \times r \times \cdots \times r}$ via
+\State {   Calculate reduced tensor $\tY^{\text{d}}_k \in \bbR^{r \times r \times p \times r \times \cdots \times r}$ via
 \begin{equation}
@@
 
-\State {\color{blue} Let $\mY_k^{\text{d}} = \text{Mat}_k(\tY^{\text{d}})$ and $J_0 = \{ i\in[p]: \onorm{\mY^{\text{d}}_{i:}} = 0\}$. Set $z_k^{(t+1)}(i)$ randomly in $[r]$ for $i \in J_0$.}
-
-\State {\color{blue} Let $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. For all $i \in J_0^c$ update the cluster assignment by
+\State {   Let $\mY_k^{\text{d}} = \text{Mat}_k(\tY^{\text{d}})$ and $J_0 = \{ i\in[p]: \onorm{\mY^{\text{d}}_{i:}} = 0\}$. Set $z_k^{(t+1)}(i)$ randomly in $[r]$ for $i \in J_0$.}
+
+\State {   Let $\mS^{(t)} = \text{Mat}(\tS^{(t)})$. For all $i \in J_0^c$ update the cluster assignment by
 \begin{equation}
@@
 
-\OUTPUT { \color{blue} Estimated clustering $z^{(T)}_k  \in [r]^{p}, k \in [K]$.}
+\OUTPUT {    Estimated clustering $z^{(T)}_k  \in [r]^{p}, k \in [K]$.}
 
@@
 {
-\color{blue}
+
 \begin{thm}[Error for weighted higher-order initialization]\label{thm:initial} Consider the general sub-Gaussian dTBM with i.i.d.\ noise under the parameter space $\tP$ and Assumption~\ref{assmp:min_gap}.
@@
 The condition~\eqref{eq:local} controls the impact of node degree to the $\mp_{\theta}(\cdot)$ with respect to the misclassification rate $\varepsilon$ and angle gap.
-\textcolor{blue}{Intuitively, the condition~\eqref{eq:local} controls the skewness of degree so that the angle between raw cluster size and degree-weighted cluster size is well controlled.
+{Intuitively, the condition~\eqref{eq:local} controls the skewness of degree so that the angle between raw cluster size and degree-weighted cluster size is well controlled.
 The stability assumption is proposed for technical convenience, and we relax this condition in numerical studies; see Section~\ref{sec:simulation}.}
@@
 
-One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square} of Bernoulli tensor $\tY$. We summarize the procedure in Sub-algorithm 3.
+{
+\color{blue}
+
+One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square}, $\Mat_{sq}(\cdot)$, of Bernoulli tensor $\tY \in \{0,1\}^{p_1 \times \cdots \times p_K}$.
+Specifically, the square matricization $\mat_{sq} (\tY) \in \{0,1\}^{p^{\floor{K/2}} \times p^{\ceil{K/2}} }$ has entries $[\Mat_{sq}(\tY)](j_1, j_2) = \tY(i_1, \ldots, i_K)$, where
+\begin{align}
+    j_1 &= i_1 + p_1(i_2 - 1) + \cdots + p_1 \cdots p_{\floor{K/2}-1} (i_{\floor{K/2}} - 1),\\
+    j_2 &= i_{\ceil{K/2}} + p_{ \ceil{K/2}}(i_{\ceil{K/2} + 1} - 1) + \cdots + p_{ \ceil{K/2}} \cdot p_{K-1} (i_K-1).
+\end{align}
+The matrix $\Mat_{sq}(\tY)$ is asymmetric.
+We are able to interpret $\Mat_{sq}(\tY)$ as the adjacency matrix for a bipartite network with connections between two groups of node.
+The two groups of nodes in the bipartite network have $p_1 \cdots p_{\floor{K/2}}$ and $ p_{\ceil{K/2}} \cdots p_K$ nodes, respectively.
+The entry $[\Mat_{sq}(\tY)](j_1, j_2)$ refers to the presence of connection between the nodes indexed by combinations $(i_1, \ldots, i_{\floor{K/2}})$ and $(i_{\ceil{K/2}}, \ldots, i_K)$.
+We summarize the procedure in Sub-algorithm 3.
 
@@
 \INPUT Bernoulli tensor $\tY \in \{0,1\}^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
-\State  Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{\floor{p^{K/2}} \times \ceil{p^{K/2}} }$ denote the nearly square unfolded tensor. Compute the estimate $\tX'$, where
+\State  Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{p^{\floor{K/2}} \times p^{\ceil{K/2}} }$ denote the nearly square unfolded tensor. Compute the estimate $\tX'$, where
 \begin{equation}\label{eq:matrixsvd}
@@
 
-\begin{prop}[Error for Bernoulli initialization]\label{cor:ber} Consider the Bernoulli dTBM in the parameter space $\tP$ and Assumption~\ref{assmp:min_gap} holds.
+\begin{prop}[Error for Bernoulli initialization]\label{prop:ber} Consider the Bernoulli dTBM in the parameter space $\tP$ and Assumption~\ref{assmp:min_gap} holds.
 Assume that $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$.
@@
 
-\begin{proof}[Proof of Proposition~\ref{cor:ber}] Following Lemma 7 in \cite{gao2018community}, with high probability, the new estimate satisfies
-\begin{equation}
-    \onormSize{}{\hat \tX' - \tX}_F^2 \lesssim p^{\ceil{K/2}}.
-\end{equation} 
-We obtain the upper bound by using $\hat \tX'$ in place of $\hat \tX$ in Theorem~\ref{thm:initial}.
-\end{proof}
-
-\begin{rmk}[Comparison with previous methods] The Bernoulli bound $\tO(p^{- \floor{K/2} })$ in Proposition~\ref{cor:ber} is relatively looser than the Gaussian bound \tO(p^{- K/2 })$ in Theorem~\ref{thm:initial}.
+%  \begin{proof}[Proof of Proposition~\ref{cor:ber}] Note that $\Mat_{sq}(\tX) \in \bbR^{p^{\floor{K/2}} \times p^{\ceil{K/2}}}$ and $\text{rank}(\Mat_{sq}(\tX)) \leq r^{\ceil{K/2}}$.
+Apply Lemma~\ref{lem:lowrank_binary} with $\mY = \Mat_{sq}(\tY), \mX = \Mat_{sq}(\tX)$, and $\hat \mX = \Mat_{sq}(\hat \tX')$.
+With probability tending to 1 as $ p \rightarrow \infty$, we have
+%         \begin{equation}
+%             \onormSize{}{\hat \tX' - \tX}_F^2 = \onormSize{}{\Mat_{sq}(\hat \tX') - \Mat_{sq}(\tX)}_F^2 \lesssim p^{\ceil{K/2}}.
+%         \end{equation}
+%         We then obtain the upper bound by using $\hat \tX'$ in place of $\hat \tX$ in Theorem~\ref{thm:initial}.
+%         \end{proof}
+
+}
+
+% ------------- original
+% One possible remedy is to apply singular value decomposition to the \emph{square unfolding}~\citep{mu2014square} of Bernoulli tensor $\tY$. We summarize the procedure in Sub-algorithm 3.
+
+% \begin{algorithm}[h!]
+% \caption*{\bf Sub-algorithm 3: Weighted higher-order initialization for Bernoulli observation}
+% \vspace{.15cm}
+% \begin{algorithmic}[1] 
+% \INPUT Bernoulli tensor $\tY \in \{0,1\}^{p\times \cdots \times p}$, cluster number $r$, relaxation factor $\eta > 1$ in $k$-means clustering.
+% \State  Let the matrix $\mat_{sq} (\tY) \in \{0,1\}^{\floor{p^{K/2}} \times \ceil{p^{K/2}} }$ denote the nearly square unfolded tensor. Compute the estimate $\tX'$, where
+% \begin{equation}\label{eq:matrixsvd}
+%     \hat \tX' = \argmin_{\text{rank}(\mat_{sq}(\tX)) \leq r^{\ceil{K/2}}} \onormSize{}{ \mat_{sq}(\tX) -  \mat_{sq}(\tY)}_F^2.
+% \end{equation}
+% \State Implement lines 3-5 of Sub-algorithm~\hyperref[alg:main]{1} with $\hat \tX$ replaced by $\hat \tX'$ in~\eqref{eq:matrixsvd}.
+% \OUTPUT Initial clustering $z^{(0)} \leftarrow \hat z$.
+% \end{algorithmic}
+% \end{algorithm}
+
+% \begin{prop}[Error for Bernoulli initialization]\label{cor:ber} Consider the Bernoulli dTBM in the parameter space $\tP$ and Assumption~\ref{assmp:min_gap} holds.
+Assume that $\mtheta$ is balanced and $\min_{i\in[p]}\theta(i) \geq c$ for some constant $c>0$.
+Let $ z^{(0)}$ denote the output of Sub-algorithm~3.
+With probability going to 1, we have
+% \begin{equation}\label{eq:ini_b}
+%  \ell(z^{(0)}, z) \lesssim \frac{r^K p^{- \floor{K/2} }}{\text{SNR}}, \quad \text{and} \quad L(z^{(0)}, z) \lesssim  {\sigma^2 r^K p^{-\floor{K/2}}}.
+% \end{equation}
+% \end{prop}
+
+% \begin{proof}[Proof of Proposition~\ref{cor:ber}] Following Lemma 7 in \cite{gao2018community}, with high probability, the new estimate satisfies
+% \begin{equation}
+%     \onormSize{}{\hat \tX' - \tX}_F^2 \lesssim p^{\ceil{K/2}}.
+% \end{equation} 
+% We obtain the upper bound by using $\hat \tX'$ in place of $\hat \tX$ in Theorem~\ref{thm:initial}.
+% \end{proof}
+
+% \begin{rmk}[Comparison with previous methods] The Bernoulli bound $\tO(p^{- \floor{K/2} })$ in Proposition~\ref{cor:ber} is relatively looser than the Gaussian bound $\tO(p^{- K/2 })$ in Theorem~\ref{thm:initial}.
+The gap between Bernoulli and Gaussian error decreases as the order $K$ increases.
+Nevertheless, combining with angle iteration Sub-algorithm~\hyperref[alg:main]{2}, Bernoulli clustering still achieves exponential error rate $\exp \of{ - p^{(K-1)}}$ at a price of a larger SNR.
+% %Nevertheless, the Bernoulli clustering still achieves exponential error rate $\exp \of{ - p^ \floor{K/2}}$ with large SNR combining with our angle iteration Sub-algorithm~\hyperref[alg:main]{2}. 
+% The exponential rate is tighter than the rate error $\tO(p^{-2})$ in previous work~\citep{ke2019community}.
+The investigation of the gap between upper bound $p^{- \floor{K/2}}$ and the lower bound $p^{- K/2}$ for Bernoulli tensors will be left as future work.
+In numerical experiments, we will use our original initialization, Sub-algorithm 1, to verify the robustness to Bernoulli observations.
+% \end{rmk}
+
+{
+\color{blue}
+
+\begin{rmk}[Comparison with Gaussian model] The Bernoulli bound $\tO(p^{- \floor{K/2} })$ in Proposition~\ref{prop:ber} is relatively looser than the Gaussian bound $\tO(p^{- K/2 })$ in Theorem~\ref{thm:initial}.
 The gap between Bernoulli and Gaussian error decreases as the order $K$ increases.
@@
 %Nevertheless, the Bernoulli clustering still achieves exponential error rate $\exp \of{ - p^ \floor{K/2}}$ with large SNR combining with our angle iteration Sub-algorithm~\hyperref[alg:main]{2}. 
-The exponential rate is tighter than the rate error $\tO(p^{-2})$ in previous work~\citep{ke2019community}.
+%The exponential rate is tighter than the rate error $\tO(p^{-2})$ in previous work~\citep{ke2019community}. 
 The investigation of the gap between upper bound $p^{- \floor{K/2}}$ and the lower bound $p^{- K/2}$ for Bernoulli tensors will be left as future work.
@@
 
-  \item \textit{Extension to sparse binary dTBMs.} The sparsity is often a popular feature in hypergraphs~\citep{florescu2016spectral,ke2019community}.
+\begin{rmk}[Comparison with previous methods]  Previous work \citep{ke2019community} develops a spectral clustering method for Bernoulli dTBM.
+\cite{ke2019community} adopts a different signal notion based on the singular gap in the core tensor, denoted as $\Delta_{\rm singular}$.
+By \citet[Theorem 1]{ke2019community}, the spectral method achieves exact recovery with $\Delta_{\rm singular} \gtrsim p^{-1/2}$.
+However, we are not able to infer the exact recovery of spectral method by our angle-base SNR condition.
+Consider an order-2 dTBM with $p > 2, \sigma^2 = 1$, $\mtheta = \mathbf{1}$, equal size assignment $|z^{-1}(a)| = p/r$ for all $ a \in [r]$, and core matrix $\mS = \mI_2$.
+The singular gap under this setting is $\Delta_{\rm singular} = \min\{ \lambda_1 - \lambda_2, \lambda_2 \} = 0$, where $\lambda_1 \geq \lambda_2$ are singular values of $\mS$.
+In contrast, our angle gap $\Delta_{\min}^2 = 2$ satisfies the SNR condition in Theorem~\ref{thm:refinement}.
+Then, our algorithm achieves the exact recovery, but the spectral method in \cite{ke2019community} fails.
+% \begin{equation}
+%     \mS = \begin{pmatrix} 1 &  0 \\
+%     0 & 1
+%     \end{pmatrix}, 
+% \end{equation}
+%for $\gamma \leq 0$. 
+% The singular gap under this setting is defined as $\Delta_{\rm singular} = \min\{ \lambda_1 - \lambda_2, \lambda_2 \}$, where $\lambda_1 \geq \lambda_2$ are singular values of $\mS$.
+When $\gamma_1 = \gamma_2 = 0$, we have angle gap $\Delta_{\min}^2 = 2$ larger than the singular gap $\Delta_{\rm singular}^2 = 0$.
+Our algorithm achieves the exact recovery, while, the spectral method in \cite{ke2019community} fails.
+%; when $\gamma_1 > \gamma_2 = -1$, we have angle gap $\Delta_{\min}^2 < p^{\gamma_1}$ smaller than the singular gap $\Delta_{\rm singular}^2 > p^{\gamma_1}$.
+
+Hence, for fair comparison, we compare the best performance of our algorithm and \cite{ke2019community} under the strongest signal setting of each model.
+Since both methods contain an iteration procedure, we set the iteration number to infinity to avoid the computational error.
+Considering the largest angle-based SNR $\asymp 1$ in Theorem~\ref{thm:refinement}, our Bernoulli clustering achieves exponential error rate of order $\exp(-p^{(K-1)})$; considering the largest singular gap $\Delta_{\rm singular} \asymp 1$ in Theorem 1 of \cite{ke2019community}, the spectral clustering has a polynomial error rate of order $p^{-2}$.
+Our algorithm still shows a better theoretical accuracy than the competitive work for Bernoulli observations.
+\end{rmk}
+
+}
+
+  \item \textit{Extension to sparse binary dTBMs.} \textcolor{blue}{The sparsity is often a popular feature in hypergraphs~\citep{florescu2016spectral,ke2019community, ahn2018hypergraph}}.
 Specifically, the sparse binary dTBM assumes that, the entries of $\tY$ follow independent Bernoulli distributions with the mean
@@
 
-The above comparison reveals three major differences.
+{
+\color{blue} The above comparison reveals four major differences. 
+}
 
@@
 
-Third, from the perspective of proofs, we develop new tools to handle the extra degree heterogeneity.
+%Third, from the perspective of proofs, we develop new tools to handle the extra degree heterogeneity.
 In our Theorem~\ref{thm:stats}, we consider the profile MLE by maximizing out the nuisance degree parameter, while TBM~\citep{han2020exact} considers the usual MLE without degree parameter.
@@
 The asymptotic non-equivalence between $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$ renders our proof technically more involved.
+
+ {
+ \color{blue}
+
+Third, from the perspective of proofs, we develop new dTBM-specific techniques to handle the extra degree heterogeneity.
+In our Theorem~\ref{thm:stats}, we construct a special non-trivial degree heterogeneity to establish the lower bound for arbitrary core tensor with small angle gap, while, TBM~\citep{han2020exact} considers the constructions without degree parameter.
+%we consider the profile MLE by maximizing out the nuisance degree parameter, while TBM~\citep{han2020exact} considers the usual MLE without degree parameter.
+In our Theorem~\ref{thm:comp}, we construct a rank-2 tensor to relate HPC conjecture to $\Delta^2_{\text{ang}}$, while TBM~\citep{han2020exact} constructs a rank-1 tensor to relate HPC conjecture to $\Delta^2_{\text{Euc}}$.
+The asymptotic non-equivalence between $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$ renders our proof technically more involved.
+
+ Last, we discuss the statistical impossibility statements.
+Our Theorem~\ref{thm:stats} implies the statistical impossibility whenever the core tensor $\tS$ leads to an angle-based SNR below the critical threshold.
+The Theorem 6 in \cite{han2020exact} implies the worst case statistical impossibility for a particular core tensor $\tS$ with Euclidean-based SNR below the statistical limit.
+Hence, our Theorem~\ref{thm:stats} shows a stronger statistical impossibility for dTBM than that presented in TBM \citet[Theorem 6]{han2020exact}.
+However, inspecting the proof of \cite{han2020exact}, the proof of Theorem 6 indeed implies a stronger TBM impossibility statement for arbitrary core tensor; i.e.,
+ \begin{equation}
+     \gamma < -(K-1) \quad \Rightarrow \quad \liminf_{p \rightarrow \infty} \inf_{ \tS \in \tP_{\tS, {\rm TBM}} \cap \{\Delta_{\rm Euc}^2 = p^{\gamma} \} }\inf_{\hat z_{\rm stats}}  \sup_{z \in \tP_{z, {\rm TBM}}}  \bbE[p \ell(\hat z_{\rm stats}, z)] \geq 1,
+ \end{equation}
+ where $\tP_{\tS, {\rm TBM}}$ and $\tP_{z, {\rm TBM}}$ refer to the space for core tensor $\tS$ and assignment $z$ under TBM, respectively.
+Again, in terms of the strong statistical impossibility, both models show similar conclusions but under different conditions.
+Since two impossibilities consider different core tensor regimes with non-equivalent $\Delta^2_{\text{ang}}$ and $\Delta^2_{\text{Euc}}$, we emphasize that different proof techniques are required to obtain these similar conclusions.
+See our proof sketch in Section~\ref{subsec:impossible}, Appendices~\ref{sec:statprove1} and \ref{sec:compprove1} for detail technical differences.
+
+ }
 
@@
 
-In this section, we provide the proof sketches for the main Theorems~\ref{thm:initial}-\ref{thm:refinement}. Detail proofs and extra theoretical results are provided in Appendix II.
+{
+\color{blue}
+
+In this section, we provide the proof sketches for the main Theorem~\ref{thm:stats} (Impossibility), Theorem~\ref{thm:comp} (Impossibility), and Theorems~\ref{thm:initial}-\ref{thm:refinement}.
+Detail proofs and extra theoretical results are provided in Appendix II.
+
+\subsection{Proof sketch of Theorem~\ref{thm:stats} (Impossibility) and Theorem~\ref{thm:comp} (Impossibility)} \label{subsec:impossible}
+
+The proofs of impossibility in Theorems~\ref{thm:stats} and \ref{thm:comp} share the same proof idea with \citet[Theorems 6 and 7]{han2020exact} and \citet[Theorem 2]{gao2018community}.
+In both proofs of statistical and computational impossibilities, the key idea is to construct a particular set of parameters to lower bound the minimax rate.
+Specifically, for statistical impossibility in Theorem~\ref{thm:stats}, we construct a particular $(z^*_{\rm stats}, \mtheta^*_{\rm stats})  \in \tP_{z, \mtheta} $ such that
+\begin{equation}\label{eq:stats_impo_sketch}
+    \inf_{\hat z_{\rm stats}} \sup_{(z, \mtheta) \in \tP_{z, \mtheta}} \bbE[p \ell(\hat z_{\rm stat}, z)] \geq \inf_{\hat z_{\rm stats}} \bbE[p \ell(\hat z_{\rm stat}, z_{\rm stats}^*)| (z^*_{\rm stats}, \tS^*,\mtheta^*_{\rm stats})] \geq 1, \quad \text{ for all $\tS^* \in \tP_{\tS}(\gamma)$};
+\end{equation}
+for computational impossibility in Theorem~\ref{thm:comp}, we construct a particular $(z^*_{\rm comp}, \tS^*_{\rm comp}, \mtheta^*_{\rm comp}) \in \tP(\gamma)$ such that 
+\begin{equation}
+    \inf_{\hat z_{\rm comp}} \sup_{(z, \tS, \mtheta) \in \tP(\gamma)} \bbE[p \ell(\hat z_{\rm comp}, z)] \geq \inf_{\hat z_{\rm comp}} \bbE[p \ell(\hat z_{\rm comp}, z_{\rm comp}^*)| (z^*_{\rm comp}, \tS^*_{\rm comp},\mtheta^*_{\rm comp})] \geq 1.
+\end{equation}
+
+The constructions of $(z^*_{\rm stats}, \mtheta^*_{\rm stats})$ and $(z^*_{\rm comp}, \tS^*_{\rm comp}, \mtheta^*_{\rm comp})$ are the most critical steps.
+With good constructions, the lower bound ``$\geq 1$" can be verified by classical statistical conclusions (e.g.\ Neyman-Pearson Lemma) or prior work (e.g.\ HPC Conjecture).
+
+A notable detail in the proof of statistical impossibility is the arbitrariness of $\tS^*$.
+The first infimum over $ \tP_{\tS}(\gamma)$ in the minimax rate~\eqref{eq:minminmax} requires that the lower bound~\eqref{eq:stats_impo_sketch} holds for any $\tS^* \in \tP_{\tS}(\gamma)$.
+The arbitrary choice of $\tS^*$ brings extra difficulties in the parameter construction, and consequently a non-trivial $\mtheta^*_{\rm stats} \neq \mathbf{1}$ is chosen to address the arbitrariness.
+Previous TBM construction in the proof of \citet[Theorem 6]{han2020exact} with $\mtheta^*_{\rm stats} = \mathbf{1}$ is no longer applicable in our case.
+Meanwhile, our construction $(z^*_{\rm comp}, \tS^*_{\rm comp}, \mtheta^*_{\rm comp})$ leads to a rank-2 mean tensor to relate the HPC Conjecture while TBM \citet[Theorem 7]{han2020exact} constructs a rank-1 mean tensor.
+Hence, we emphasize that dTBM-specific techniques are required to obtain our impossibility results, though the proof idea is common for minimax lower bound analysis.
+
+}
 
@@
 
-{\color{blue}
+{  
 The proof of MLE error shares the similar idea as Theorems~\ref{thm:initial}-\ref{thm:refinement}.
@@
 
-{\color{blue}
+{  
 \section*{Appendix I. Additional numerical experiments}
@@
 In each sub-section, we first show the proof of main theorem and then collect the useful lemmas in the end.
-\textcolor{blue}{We combine the proofs of MLE achievement in Theorem~\ref{thm:stats} and polynomial-time achievement in Theorem~\ref{thm:refinement} in the last section due to the similar idea.
+{We combine the proofs of MLE achievement in Theorem~\ref{thm:stats} and polynomial-time achievement in Theorem~\ref{thm:refinement} in the last section due to the similar idea.
 }
@@
 {
-\color{blue}
+
 By constraints in parameter space~\eqref{eq:general_family}, neither $\mS_{1,1:}$ nor $\mS_{1,2:}$ is a zero vector.
@@
 
+% {
+% \color{blue}
+% \begin{proof}[Proof of Lemma~\ref{lem:lowrank_binary}] Note that $\onormSize{}{\hat \mX - \mY}_F^2 \leq \onormSize{}{\mX - \mY}_F^2$ by the definition of least square estimator.
+Re-arranging the terms, we have
+%         \begin{align}
+%             \onormSize{}{\hat \mX - \mX}_F^2 & \leq 2 \ang{ \hat \mX - \mX, \mY - \mX} \\ & \leq 2 \onormSize{}{\hat \mX - \mX}_F \sup_{\mT \in \bbR^{m \times n}, \text{rank}(\mT) \leq 2r, \onormSize{}{\mT}_F = 1} \ang{\mT, \mY - \mX} \\
+%             & \leq C  \onormSize{}{\hat \mX - \mX}_F \sqrt{nr}
+%         \end{align}
+%         with probability at least $1 - \exp(-C_2 nr)$ for some positive constant $C_2$, where the second inequality follows from \citet[Lemma 1]{wang2019multiway}, and the last inequality follows from \citet[Lemma E5]{han2022optimal} under the matrix case and the fact that the Orlicz norm of Bernoulli noise $\onormSize{}{\mY_{ij} - \mX_{ij}}_{\varphi_2} = \sup_{q \geq 1} \bbE[|\mY_{ij} - \mX_{ij}|^q]^{1/q}/q^{1/2} \leq 1$  since $|Y_{ij} - X_{ij}| \leq 1$ for all $i \in [m], j \in [n]$.
+%         \end{proof}
+
+% }
+
 \subsection{Proof of Theorem~\ref{thm:stats} (Impossibility)}\label{sec:statprove1}
-
-\begin{proof}[Proof of Theorem~\ref{thm:stats} (Impossibility)]
-We will prove a more general conclusion than the main paper by allowing growing $r_k$'s. 
-%We firstly show the proof when the parameter satisfies condition (b). 
-Consider the general asymetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$. Specifically, we will show that, under 
-the assumptions $K\geq 1$, $r\lesssim p^{1/3}$ and SNR condition
-\[
-{\Delta^2_{\min}\over \sigma^2}\lesssim {r^{K-1}\over p^{K-1}},\quad \text{or equivalently}, \quad \gamma\leq -(K-1)(1+\log_p r),
-\]
- the desired conclusion in Theorem~\ref{thm:stats} holds; i.e, for all $k \in [K]$, every estimator $\hat z_{k,\text{stat}}$ obeys
-\begin{equation}\label{eq:max}
-    \sup_{(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{k,\text{stat}}, z_k) \right]\geq 1.
-\end{equation}
-
-Since the inequality~\eqref{eq:max} is a minimax lower bound, it suffices to show the inequality holds for a particular $(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)$. 
-% Specifically, we consider the estimation problem based on a particular parameter point $(\{z_k\}, \tS, \{\mtheta_k\})$ with the following three properties:
-% \begin{equation}\label{eq:construction}
-%   \text{(i)}\  \theta_k(i)=1 \text{ for all }i\in[p]; \quad  \text{(ii)}\ 
-%     \Delta_{\min}\lesssim  \left({p\over r}\right)^{-\frac{K-1}{2}}\sigma; \quad  \text{(iii)}\ |z^{-1}_k(a)|={p \over r} \in \mathbb{Z}_+ \text{ for all }a\in[r],
-% \end{equation}
-% for all $k \in [K]$.
-
 {
 \color{blue}
-%\fixme{Jiaxin}{The condition (i) can be replaced by the balance assumption after using the construction of $T$ in \cite{gao2018community}.
+
+\begin{proof}[Proof of Theorem~\ref{thm:stats} (Impossibility)]Consider the general asymetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$.
+For simplicity, we show the minimax rate for the estimation on the first mode $\hat z_1$; the proof for other modes are essentially the same.
+
+   To prove the minimax rate~\eqref{eq:minminmax}, it suffices to take an arbitrary $\tS^* \in  \tP_{\tS}(\gamma)$ wih $\gamma < -(K-1)$ and construct $(z^*_k, \mtheta^*_k)$ such that 
+   \begin{equation}
+       \inf_{\hat z_1} \bbE \left[ p\ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right]\geq 1.
+   \end{equation}
+
+   We first define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation.
+Based on \citet[Proof of Theorem 6]{han2020exact}, we consider the restricted family of $\hat z_k$'s for which the following three conditions are satisfied:
+\begin{equation}
+    \text{(a)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \quad \text{(b)} \ |T^c_k|\asymp {p\over r}; \quad \text{(c)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
+\end{equation} 
+for all $k \in [K]$.
+   Now, we consider the construction:
+   \begin{enumerate}
+       \item[(i)] $\{z_k^*\}$ satisfies properties (a)-(c) with misclassification sets $T_k^c$ for all $k \in [K]$;
+       \item [(ii)] $\{\mtheta_k^*\}$ such that $\mtheta_k^*(i) \leq \sigma r^{(K-1)/2} p^{-(K-1)/2}$ for all $i \in T_k^c, k \in [K]$ and $\max_{k \in [K], a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}^2_2 \asymp p/r$.
+   \end{enumerate}
+   % There exists a $(z^*, \mtheta^*) \in \tP_{z, \mtheta}$ satisfying the property (i) and (ii) when $r < p$; otherwise, when $r = p$, we must have $\mtheta^*(i) = 1$ due to the $\ell_1$ norm constraint and we can not satisfy the property (ii).
+
+   Combining the inequalities (39) and (40) in the proof of Theorem 2 in \cite{gao2018community}, we have 
+   \begin{align}
+        & \inf_{\hat z_1 } \bbE \left[ \ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right] \geq  \\
+        & \quad \quad \quad \frac{C}{r^3 |T_1^c|} \sum_{i \in T_1^c} \inf_{\hat z_1  (i)} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, z_k^*, \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, z_k^*, \mtheta_k^*, \tS^*] \},\label{eq:inf_lower}
+   \end{align}
+   where $C$ is some positive constant,  $\hat z_1$ on the left hand side denote the generic assignment functions in $\tP(\gamma)$, and the infimum on the right hand side is taken over the generic assignment function family of $\hat z_1(i)$ for all nodes $i \in T_1^c$.
+Here, the factor $r^3=r\cdot r^2$ in~\eqref{eq:inf_lower} comes from two sources: $r^2\asymp {r\choose 2}$ comes from the multiple testing burden for all pairwise comparisons among $r$ clusters; and another $r$ comes from the number of elements $|T^c_k|\asymp p/r$ to be clustered.
+
+   Next, we need to find the lower bound of the rightmost side in~\eqref{eq:inf_lower}. 
+   %For simplicity, we show the bound for the mode-1 case $k = 1$ only. We drop the subscripts $1$ in $z_1, T_1, \mS_1, \mtheta_1$ and omit the repeated procedures for the cases of $k = 2,\ldots, K$. 
+We consider the hypothesis test based on model~\eqref{eq:general_dtbm}. First, we reparameterize the model under the construction (i)-(ii).
+\begin{equation}
+    \mx_a^* = \off{\mat_1 \of{ \tS^*\times_2 \mTheta_2^* \mM_2^*\times_3\cdots \times_K \mTheta^*_K \mM_K^* }}_{a:}, \quad \text{for all} \ a \in [r],
+\end{equation}
+where $\mx_a^*$'s are centroids in $\bbR^{p^{K-1}}$.
+Without loss of generality, we consider the lower bound for the summand in~\eqref{eq:inf_lower} for $i=1$.
+The analysis for other $i\in T^c_1$ are similar.
+For notational simplicity, we suppress the subscript $i$ and write $\my, \theta^*, z$ in place of $\my_1, \mtheta_1^*(1)$ and $z_1(1)$, respectively.
+The equivalent vector problem for assessing the summand in~\eqref{eq:inf_lower} is
+\begin{equation}\label{eq:z}
+\my=\theta^* \mx_{z}^*+\me,
+\end{equation}
+where $z\in \{1,2\}$ is an unknown parameter, $\theta^* \in \bbR_+$ is the given heterogeneity degree, $\mx_1^*,\mx_2^*\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.
+Then, we consider the hypothesis testing under the model~\eqref{eq:z}:
+\begin{equation}\label{eq:test}
+ H_0: z = 1 \text{ and } \my = \theta^* \mx_1^* + \me \quad \leftrightarrow \quad H_1: z = 2 \text{ and } \my = \theta^* \mx_2^* + \me,
+\end{equation}
+
+%   For an arbitrary $i \in T_1^c$, the right hand side in \eqref{eq:inf_lower} corresponds to the following hypothesis test:
+%   \begin{equation}\label{eq:hypo}
+%       H_0: \my = \mtheta_1^*(i) \mx_1^* + \me \quad \leftrightarrow \quad H_1: \my = \mtheta_1^*(i) \mx_2^* + \me,
+%   \end{equation}
+%   where $\my = [\mat_1(\tY)]_{i:}$, $\mx_a^* = [\mat_1( \tS^* \times_2 \mTheta^* \mM_2^* \times_3 \cdots \times_K \mTheta^* \mM_K^*)]_{a:}$, and $\me$ consists of i.i.d.\ $N(0,\sigma^2)$ entries. 
+   The hypothesis testing~\eqref{eq:test} is a simple versus simple testing, since the assignment $z$ is the only unknown parameter in the test. 
+   By Neyman-Pearson lemma, the likelihood ratio test is optimal with minimal Type I + II error.
+Under Gaussian model, the likelihood ratio test of \eqref{eq:test} is equivalent to the least square estimator $\hat z_{LS} = \argmin_{a = \{1,2\}} \onormSize{}{\my - \theta^* \mx_a^*}_F^2$.
+
+   Let $\mS = \mat_1(\tS)$. Note that 
+   \begin{align}
+       \onormSize{}{ \theta^* \mx_1^*  - \theta^* \mx_2^*}_F &\leq  \theta^* \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*)  \\
+       & \leq \theta^* \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F   \max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}_2^{K-1} \\
+       & \leq  \sigma r^{(K-1)/2} p^{-(K-1)/2} 2 c_4 p^{(K-1)/2} r^{-(K-1)/2}\\
+       & \leq 2 c_4 \sigma, \label{eq:bound}
+   \end{align}
+%   \begin{align}
+%       \onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F &\leq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*)  \\
+%       & \leq \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F   \max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}_2^{K-1} \\
+%       & \leq  \sigma r^{(K-1)/2} p^{-(K-1)/2} 2 c_4 p^{(K-1)/2} r^{-(K-1)/2}\\
+%       & \leq 2 c_4 \sigma, \label{eq:bound}
+%   \end{align}
+   where $\lambda_{\max}$ denote the maximal singular value, the second inequality follows from Lemma~\ref{lem:singular_thetam}, and the third inequality follows from property (ii) and the boundedness constraint in $\tP_{\tS}(\gamma)$ such that $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F  \leq \onormSize{}{\mS_{1:}^*}_F + \onormSize{}{\mS_{2:}^*}_F \leq 2c_4$.
+
+   Hence, we have 
+   \begin{align}
+       & \inf_{\hat z_1(1)} \{ \bbP[\hat z_1(1) = 1| z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(1) = 2| z_1^*(1) = 1, z^*_k,  \mtheta_k^*, \tS^*] \} \\
+       & \quad = 2 \bbP[ \hat z_{LS} = 1 |  z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^* ] \\
+      & \quad  = 2 \bbP[ \onormSize{}{\my - \theta^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \theta^* \mx_2^*}_F^2 | z_1^*(1) = 2, z^*_k , \mtheta_k^*, \tS^*   ] \\
+      & \quad = 2 \bbP \off{ 2 \ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \geq  \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F^2 }\\
+       & \quad =  2 \bbP[ N(0,1) \geq \theta^* \onormSize{}{\mx_1^* - \mx_2^*}_F /(2\sigma) ] \\
+       & \quad \geq 2 \bbP[ N(0,1) \geq c_4 ]   \geq c, \label{eq:inf}
+   \end{align}
+%   \begin{align}
+%       & \inf_{\hat z_1(1)} \{ \bbP[\hat z_1(1) = 1| z_1^*(i) = 2, z^*_k,  \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(1) = 2| z_1^*(i) = 1, z^*_k,  \mtheta_k^*, \tS^*] \} \\
+%       & \quad = 2 \bbP[ \hat z_{LS}(i) = 1 |  z_1^*(i) = 2, z^*_k,  \mtheta_k^*, \tS^* ] \\
+%       & \quad  = 2 \bbP[ \onormSize{}{\my - \mtheta_1(i)^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \mtheta_1(i)^* \mx_2^*}_F^2 | z_1^*(i) = 2, z^*_k , \mtheta_k^*, \tS^*   ] \\
+%       & \quad =  2 \bbP[ N(0,1) \geq \mtheta_1(i)^* \onormSize{}{\mx_1^* - \mx_2^*}_F /\sigma ] \\
+%       & \quad \geq 2 \bbP[ N(0,1) \geq 2 c_4 ]   \geq c, \label{eq:inf}
+%   \end{align}
+where the first equation holds by symmetry, the third equation holds by rearrangement, the fourth equation holds by the fact that $\ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \sim N(0, \sigma \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F)$, and $c$ is some positive constant in the last inequality.
+   %for some positive constant $c$, where the calculation for the last equation follows from the calculation in equation (102) in \cite{han2020exact}.
+
+   Plugging the inequality~\eqref{eq:inf} into the inequality~\eqref{eq:inf_lower} for all $i \in T_1^c$, then, we have 
+   \begin{equation}
+       \liminf_{p \rightarrow \infty}  \inf_{\hat z_1 } \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right]  \geq \liminf_{p \rightarrow \infty} \frac{ C cp }{r^3} \geq C c,
+   \end{equation}
+   where the last inequality follows by the condition $r = o(p^{1/3})$. By the discrete nature of the misclassification error, we obtain our conclusion
+   \begin{equation}
+       \liminf_{p \rightarrow \infty} \inf_{\tS^* \in  \tP_{\tS}(\gamma)}  \inf_{\hat z_{\text{stat}} }\sup_{ (z^*, \mtheta^*) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]  \geq 1.
+   \end{equation}
+%   for any $r < p$ and $p$,  we have 
+%   \begin{equation}
+%         \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq \frac{ cp }{r^3} > 0,
+%   \end{equation}
+%   which indicates 
+%   \begin{equation}
+%         \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq 1, \quad \text{for every } p
+%   \end{equation}
+%   by the discrete nature of $p\ell(\hat z, z^*)$.
+   \end{proof}
+}
+
+% \subsection{Proof of Theorem~\ref{thm:stats} (Impossibility)}\label{sec:statprove1}
+
+% {
+% \color{blue}
+% \begin{proof}[Proof of Theorem~\ref{thm:stats} (Impossibility)] We will prove a more general conclusion than the main paper by allowing growing $r_k$'s. 
+% Consider the general asymetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$.
+Let $\tP_{z} = \{ \{z_k\}: \text{there exist } (\tS, \{\mtheta_k\}) \text{ such that } (\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\{r_k\}) \}$ denote the space for all possible assignment $\{z_k\}$ under dTBM.
+Specifically, we will show that, under
+% the assumptions $K\geq 1$, $r\lesssim p^{1/3}$ and SNR condition
+% \[
+% {\Delta^2_{\min}\over \sigma^2}\lesssim {r^{K-1}\over p^{K-1}},\quad \text{or equivalently}, \quad \gamma\leq -(K-1)(1+\log_p r),
+% \]
+%  the desired conclusion in Theorem~\ref{thm:stats} holds; i.e, for all $ k \in [K]$, we have 
+% \begin{equation}\label{eq:max}
+%     \inf_{\{\hat z_{k,\text{stat}}\} \in \tP_z} \sup_{(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{k,\text{stat}}, z_k) \right]\geq 1,
+% \end{equation}
+% where $\tP_{z} = \{ \{z_k\}: c_1 p_k/r_k \leq |z_k^{-1}(a)| \leq c_2p_k/r_k, \text{for all }z \in [r_k], k \in [K] \}$ with positive constants $c_1, c_2$ in~\eqref{eq:general_family} denotes the space for all possible assignment under dTBM.
+% %$\tP_{z} = \{ \{z_k\}: \text{there exist } (\tS, \{\mtheta_k\}) \text{ such that } (\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\{r_k\}) \}$
+
+% Since the inequality~\eqref{eq:max} is a minimax lower bound, it suffices to show the inequality holds for a particular $(\{z_k^*\}, \tS^*, \{\mtheta_k^*\}) \in \tP(\gamma)$.
+We specify the parameters with following properties:
+% \begin{align}
+%     (\text{i}) \quad  \{z_k^*\} \text{ the same in \citet[Proof of Theorem 6]{han2020exact}}; \quad (\text{ii}) \quad \mtheta^*_k = \mathbf{1}, k \in [K]; 
+% \end{align}
+% and 
+% \begin{equation}
+%     (\text{iii}) \quad  \Delta_{\min}^2(\tS^*) \lesssim \of{\frac{p}{r}}^{-(K-1)}; \quad (\text{iv}) \quad  \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onormSize{}{\mat_k(\tS^*)_{a:} - \mat_k(\tS^*)_{b:}}_F^2 \lesssim  \of{\frac{p}{r}}^{-(K-1)}.
+% \end{equation}
+% The properties (iii) and (iv) do not lead to an empty set. Because by Lemma~\ref{lem:norm_diff} for all $k \in [K]$ and $a \neq b \in [r_k]$
+% \begin{equation}
+%      \onormSize{}{\mat_k(\tS^*)_{a:}^s - \mat_k(\tS^*)_{b:}^s}_F^2 c_3  \leq 2  \onormSize{}{\mat_k(\tS^*)_{a:} - \mat_k(\tS^*)_{b:}}_F^2,
+% \end{equation}
+% where $c_3$ is the positive constant in parameter space~\eqref{eq:general_family} and thus
+% \begin{equation}
+%   \{\tS: \min_{k \in [K]} \min_{a \neq b \in [r_k]} \onormSize{}{\mat_k(\tS)_{a:} - \mat_k(\tS)_{b:}}_F^2 \lesssim  \of{\frac{p}{r}}^{-(K-1)} \}  \subset \{\tS: \Delta_{\min}^2(\tS) \lesssim \of{\frac{p}{r}}^{-(K-1)}\}.
+% \end{equation}
+% Hence, the true parameter $(\{z_k^*\}, \tS^*, \{\mtheta_k^*\})$ satisfying properties (i)-(iv) coincides with the specified TBM parameters in \citet[Proof of Theorem 6]{han2020exact}.
+% Also, $\tP_z$ is equal to the collection of all possible $\{z_k\}$ under the TBM~\cite{han2020exact} due to the same group size balance constraint for $\{z_k\}$.
+
+% Then, we have the minimax lower bound~\eqref{eq:max}
+% \begin{align}
+%      \inf_{\{\hat z_{k,\text{stat}}\} \in \tP_z} \sup_{(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{k,\text{stat}}, z_k) \right] \geq  \inf_{\{\hat z_{k,\text{stat}}\} \in \tP_z} \bbE \left[ p\ell(\hat z_{k,\text{stat}}, z_k^*) |(\{z_k^*\}, \{\mtheta_k^*\}, \tS^*)\right] \geq 1,
+% \end{align}
+% where the last inequality follows from the inequalities (100) and (102) in \citet[Proof of Theorem 6]{han2020exact}.
+
+% \end{proof}
+
+% }
+
+% \begin{proof}[Proof of Theorem~\ref{thm:stats} (Impossibility)]
+% We will prove a more general conclusion than the main paper by allowing growing $r_k$'s. 
+% %We firstly show the proof when the parameter satisfies condition (b). 
+% Consider the general asymetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$. Specifically, we will show that, under 
+% the assumptions $K\geq 1$, $r\lesssim p^{1/3}$ and SNR condition
+% \[
+% {\Delta^2_{\min}\over \sigma^2}\lesssim {r^{K-1}\over p^{K-1}},\quad \text{or equivalently}, \quad \gamma\leq -(K-1)(1+\log_p r),
+% \]
+%  the desired conclusion in Theorem~\ref{thm:stats} holds; i.e, for all $k \in [K]$, every estimator $\hat z_{k,\text{stat}}$ obeys
+% \begin{equation}\label{eq:max}
+%     \sup_{(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)} \bbE \left[ p\ell(\hat z_{k,\text{stat}}, z_k) \right]\geq 1.
+% \end{equation}
+
+% Since the inequality~\eqref{eq:max} is a minimax lower bound, it suffices to show the inequality holds for a particular $(\{z_k\}, \tS, \{\mtheta_k\}) \in \tP(\gamma)$. 
+% % Specifically, we consider the estimation problem based on a particular parameter point $(\{z_k\}, \tS, \{\mtheta_k\})$ with the following three properties:
+% % \begin{equation}\label{eq:construction}
+% %   \text{(i)}\  \theta_k(i)=1 \text{ for all }i\in[p]; \quad  \text{(ii)}\ 
+% %     \Delta_{\min}\lesssim  \left({p\over r}\right)^{-\frac{K-1}{2}}\sigma; \quad  \text{(iii)}\ |z^{-1}_k(a)|={p \over r} \in \mathbb{Z}_+ \text{ for all }a\in[r],
+% % \end{equation}
+% % for all $k \in [K]$.
+
+% {
+
+% %\fixme{Jiaxin}{The condition (i) can be replaced by the balance assumption after using the construction of $T$ in \cite{gao2018community}.
 The only difference between $T_{gao}$ and $T_{han}$ is: $T_{gao}$ collects the indices with large degree and ensures the boundedness of $\theta(i)$ for $i \in T^c$ while $T_{han}$ chooses the indices randomly.
@@
 
-Specifically, we consider the estimation problem based on a particular parameter point $(\{z_k\}, \tS, \{\mtheta_k\})$ with the following two properties:
-\begin{equation}\label{eq:construction}
-\text{(i)}\  {\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}; \quad  \text{(ii)}\ 
-    \Delta_{\min}\lesssim  \left({p\over r}\right)^{-\frac{K-1}{2}}\sigma; \quad  \text{(iii)}\ |z^{-1}_k(a)|={p \over r} \in \mathbb{Z}_+ \text{ for all }a\in[r],
-\end{equation}
-for all $k \in [K]$.
-
-%Furthermore, we define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation.....
-
-}
-
-Furthermore, we define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation.
+% Specifically, we consider the estimation problem based on a particular parameter point $(\{z_k\}, \tS, \{\mtheta_k\})$ with the following two properties:
+% \begin{equation}\label{eq:construction}
+% \text{(i)}\  {\min_{a\in[r]} \onormSize{}{\mtheta_{z^{-1}(a)}}=\left(1+o(1)\right)\max_{a\in[r]}\onormSize{}{\mtheta_{z^{-1}(a)}}}; \quad  \text{(ii)}\ 
+%     \Delta_{\min}\lesssim  \left({p\over r}\right)^{-\frac{K-1}{2}}\sigma; \quad  \text{(iii)}\ |z^{-1}_k(a)|={p \over r} \in \mathbb{Z}_+ \text{ for all }a\in[r],
+% \end{equation}
+% for all $k \in [K]$.
+
+% %Furthermore, we define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation.....
+
+% }
+
+% Furthermore, we define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation.
 Based on \citet[Proof of Theorem 6]{han2020exact}, we consider the minimax rate over the restricted family of $\hat z_k$'s for which the following three conditions are satisfied:
-\begin{equation}
-    \text{(iv)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \quad \text{(v)} \ |T^c_k|\asymp {p\over r}; \quad \text{(vi)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
-\end{equation}
+% \begin{equation}
+%     \text{(iv)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \quad \text{(v)} \ |T^c_k|\asymp {p\over r}; \quad \text{(vi)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
+% \end{equation}
+% % {
+% %    
+% % \begin{equation}
+% %     \text{(iv)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \quad \text{(v)} \ |T^c_k|\asymp {p\over r}; \quad \text{(vi)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
+% % \end{equation}
+
+% % }
+
+% \normalsize
+% for all $k \in [K]$.
+
 % {
-% \color{blue} 
-% \begin{equation}
-%     \text{(iv)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \quad \text{(v)} \ |T^c_k|\asymp {p\over r}; \quad \text{(vi)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
-% \end{equation}
-
-% }
-
-\normalsize
-for all $k \in [K]$.
-
-{
-\color{blue}
- The construction of $T$ is precisely the same as \citet[Proof of Theorem 2]{gao2018community}.
+
+%  The construction of $T$ is precisely the same as \citet[Proof of Theorem 2]{gao2018community}.
 By the constraint $\onormSize{}{\mtheta_{z^{-1}(a)}}_1 = |z^{-1}(a)|, a \in [r]$  in parameter space~\eqref{eq:family}, the $\max_{i \in T^c} \theta(i) \leq C$ for some positive constant $C$.
 Then, following the proof of~\citet[Theorem 2]{gao2018community},
-}
-% The construction of $T$ is precisely the same as \citet[Proof of Theorem 6]{han2020exact}. 
-% Then, following the proof of~\citet[Theorem 2]{gao2018community},
-for all $k \in [K]$, we have
-\begin{equation}\label{eq:minimax}
-\inf_{\hat z_k}\sup_{z_k}\bbE \ell(\hat z_k, z_k) \gtrsim {1\over r^3 |T^c_k|}\sum_{i\in T^c_k}\inf_{\hat z_k}\left\{\bbP[\hat z_k(i)=2|z_k(i)=1] + \bbP[\hat z_k(i)=1|z_k(i)=2] \right\},
-\end{equation}
-where $\hat z_k$ and $z_k$ on the left hand side denote the generic clustering functions in $\tP(\gamma)$, $z_k$ on the right hand side denotes a particular parameter satisfying properties (i)-(vi), and the infimum on the right hand side is taken over the restricted family of $\hat z$ satisfying (iv)-(vi).
+% }
+% % The construction of $T$ is precisely the same as \citet[Proof of Theorem 6]{han2020exact}. 
+% % Then, following the proof of~\citet[Theorem 2]{gao2018community},
+% for all $k \in [K]$, we have
+% \begin{equation}\label{eq:minimax}
+% \inf_{\hat z_k}\sup_{z_k}\bbE \ell(\hat z_k, z_k) \gtrsim {1\over r^3 |T^c_k|}\sum_{i\in T^c_k}\inf_{\hat z_k}\left\{\bbP[\hat z_k(i)=2|z_k(i)=1] + \bbP[\hat z_k(i)=1|z_k(i)=2] \right\},
+% \end{equation}
+% where $\hat z_k$ and $z_k$ on the left hand side denote the generic clustering functions in $\tP(\gamma)$, $z_k$ on the right hand side denotes a particular parameter satisfying properties (i)-(vi), and the infimum on the right hand side is taken over the restricted family of $\hat z$ satisfying (iv)-(vi).
 Here, the factor $r^3=r\cdot r^2$ in~\eqref{eq:minimax} comes from two sources: $r^2\asymp {r\choose 2}$ comes from the multiple testing burden for all pairwise comparisons among $r$ clusters; and another $r$ comes from the number of elements $|T^c_k|\asymp p/r$ to be clustered.
 
-Next, we need to find the lower bound of the rightmost side in~\eqref{eq:minimax}.
+% Next, we need to find the lower bound of the rightmost side in~\eqref{eq:minimax}.
 For simplicity, we show the bound for the mode-1 case $k = 1$ only.
@@
 
-We consider the hypothesis test based on model~\eqref{eq:general_dtbm}. First, we reparameterize the model under the construction~\eqref{eq:construction}
-\begin{equation}
-    \mx_a = \off{\mat_1 \of{ \tS\times_2 \mM_2\times_3\cdots \times_K \mM_K }}_{a:}, \quad \text{for all} \ a \in [r],
-\end{equation}
-where $\mx_a$'s are centroids in $\bbR^{p^{K-1}}$.
+% We consider the hypothesis test based on model~\eqref{eq:general_dtbm}. First, we reparameterize the model under the construction~\eqref{eq:construction}
+% \begin{equation}
+%     \mx_a = \off{\mat_1 \of{ \tS\times_2 \mM_2\times_3\cdots \times_K \mM_K }}_{a:}, \quad \text{for all} \ a \in [r],
+% \end{equation}
+% where $\mx_a$'s are centroids in $\bbR^{p^{K-1}}$.
 Without loss of generality, we consider the lower bound for the summand in~\eqref{eq:minimax} for $i=1$.
@@
 The equivalent vector problem for assessing the summand in~\eqref{eq:minimax} is
-\begin{equation}\label{eq:z}
-\my=\theta\mx_{z}+\me,
-\end{equation}
-where $\theta\in\bbR_{+}$ and $z\in \{1,2\}$ are unknown parameters, $\mx_1,\mx_2\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.
-Then, we consider the hypothesis testing under the model~\eqref{eq:z}:
-\begin{equation}\label{eq:test}
-H_0\colon z=1, \quad \text{v.s.}\quad H_\alpha\colon z=2.
-\end{equation}
-Note that the profile log-likelihood with respect to $z$ is
-\[
-\tL(z,  \theta(z);\my)\propto - \inf_{\theta>0}\onormSize{}{\my- \theta\mx_{z}}^2 \propto  \cos^2(\my, \mx_{z}) \ind\{ \langle \my,\mx_{z}\rangle  >0 \}, 
-\]
-and the MLE's of $\theta$ and $z$ are
-\begin{equation}\label{eq:MLE}
-    \hat \theta_{\text{MLE}}= \hat \theta(\hat z_{\text{MLE}})={\langle \my,\mx_{\hat z_{\text{MLE}}}\rangle \over \onorm{\mx_{\hat z_{\text{MLE}}}}^2}\vee 0 , \quad 
-\hat z_{\text{MLE}}= \argmax_{a\in\{1,2\}} \offf{\cos(\my,\ \mx_a)\vee 0}. 
-\end{equation}
-
-Then, the decision rule $\hat z_{\text{MLE}} \in\{1,2\}$ based on profile log-likelihood ratio is defined as
-\begin{equation}\label{eq:ratio}
-\hat z_{\text{MLE}}=
-\begin{cases}
-1 & \text{if }\cos(\my,\mx_1)\geq  \cos(\my,\mx_2) \text{ and }\langle \my,\mx_1\rangle > 0, \\
-2 & \text{if }\cos(\my,\mx_1)<\cos(\my,\mx_2) \text{ and }\langle \my,\mx_2\rangle > 0,\\
-\text{1 or 2 with equal probability} & \text{otherwise.}
-\end{cases}
-\end{equation}
-
-The Neyman-Pearson Lemma implies
-\begin{align}\label{eq:lower}
-&\inf_{\hat z}\left\{\bbP[\hat z=2|z=1] + \bbP[\hat z=1|z=2]\right\}= \bbP[\hat z_{\text{MLE}}=1|z=2]+\bbP[\hat z_{\text{MLE}}=2|z=1].
-\end{align}
-By symmetric, it suffices to bound $\bbP[\hat z_{\text{MLE}}=1|z=2]$. Using~\eqref{eq:ratio}, we obtain
-\begin{align}\label{eq:bound}
-\bbP[\hat z_{\text{MLE}}=1|z=2]&= \bbP\left[ \cos(\theta\mx_2+\me,\mx_1)\geq  \cos(\theta\mx_2+\me, \mx_2) \text{ and } \langle \theta\mx_2+\me,\mx_1\rangle > 0  \right] \notag \\
-&\stackrel{(*)}{\geq} \bbP \left[\left\langle \me, {\mx^s_1 -\mx^s_2\over\onorm{\mx^s_1- \mx^s_2}}  \right\rangle  \geq {\theta \over 2}\onormSize{}{\mx_2} \onormSize{}{\mx^s_1-\mx^s_2}\rangle\right]-\notag \\
-&\hspace{1cm} \bbP\left[\langle \me, \mx^s_1\rangle \leq -{\theta \over 2}\onorm{\mx_2}\left(2-\onormSize{}{\mx^s_1-\mx^s_2}^2\right)\right]\notag \\
-&\stackrel{(**)}{=}\Phi\left({\theta \over 2}\onorm{\mx_2}(2-\onormSize{}{\mx^s_1-\mx^s_2}^2) \right)-\Phi\left({\theta\over 2}\onorm{\mx_2}\onormSize{}{\mx^s_1-\mx^s_2} \right),
-\end{align}
-where $\Phi(\cdot)$ denotes the CDF for standard normal distribution.
+% \begin{equation}\label{eq:z}
+% \my=\theta\mx_{z}+\me,
+% \end{equation}
+% where $\theta\in\bbR_{+}$ and $z\in \{1,2\}$ are unknown parameters, {\color{red} \fixme{Jiaxin}{ Why $\theta$ is an unknown parameter?
+Equation~\eqref{eq:z} should represent the distribution of $y$ when all true parameters in equation~\eqref{eq:construction} are given, where true parameters include $\mx_z$ and $\theta$.
+Given $\theta_i = 1$, the full hypothesis test corresponding to the right hand side of equation~\eqref{eq:minimax} should be $H_0: y = \mx_1 + \me$ versus $H_1: y = \mx_2 + \me$.
+By Neyman-Pearson Lemma, the best estimator becomes the non-degree least square, \textit{no matter the estimation space includes known (TBM) $\theta$ or unknown (dTBM) $\theta$.} Then, we can directly use TBM results.
+The key step for the new proof should be: }}
+% {
+% \color{red} 
+% Given true parameter $\theta(i) = 1$, 
+% \begin{align}
+%     &\inf_{\hat z_k, (\hat z, \hat \theta, \tS ) \in \tP_{\rm dTBM}} \left\{\bbP[\hat z_k(i)=2|z_k(i)=1] + \bbP[\hat z_k(i)=1|z_k(i)=2] \right\} \\
+%     = & \inf_{\hat z_k, (\hat z, \tS ) \in \tP_{\rm TBM}} \left\{\bbP[\hat z_k(i)=2|z_k(i)=1] + \bbP[\hat z_k(i)=1|z_k(i)=2] \right\}.
+% \end{align}
+% The best estimator is $\hat z(i) = \argmin_{a \in \{1,2\}} \onormSize{}{\my - \mx_a}_2^2$.
+
+% }
+
+% $\mx_1,\mx_2\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.  Then, we consider the hypothesis testing under the model~\eqref{eq:z}:
+% \begin{equation}\label{eq:test}
+% H_0\colon z=1, \quad \text{v.s.}\quad H_\alpha\colon z=2.
+% \end{equation}
+% Note that the profile log-likelihood with respect to $z$ is
+% \[
+% \tL(z,  \theta(z);\my)\propto - \inf_{\theta>0}\onormSize{}{\my- \theta\mx_{z}}^2 \propto  \cos^2(\my, \mx_{z}) \ind\{ \langle \my,\mx_{z}\rangle  >0 \}, 
+% \]
+% and the MLE's of $\theta$ and $z$ are
+% \begin{equation}\label{eq:MLE}
+%     \hat \theta_{\text{MLE}}= \hat \theta(\hat z_{\text{MLE}})={\langle \my,\mx_{\hat z_{\text{MLE}}}\rangle \over \onorm{\mx_{\hat z_{\text{MLE}}}}^2}\vee 0 , \quad 
+% \hat z_{\text{MLE}}= \argmax_{a\in\{1,2\}} \offf{\cos(\my,\ \mx_a)\vee 0}. 
+% \end{equation}
+
+% Then, the decision rule $\hat z_{\text{MLE}} \in\{1,2\}$ based on profile log-likelihood ratio is defined as
+% \begin{equation}\label{eq:ratio}
+% \hat z_{\text{MLE}}=
+% \begin{cases}
+% 1 & \text{if }\cos(\my,\mx_1)\geq  \cos(\my,\mx_2) \text{ and }\langle \my,\mx_1\rangle > 0, \\
+% 2 & \text{if }\cos(\my,\mx_1)<\cos(\my,\mx_2) \text{ and }\langle \my,\mx_2\rangle > 0,\\
+% \text{1 or 2 with equal probability} & \text{otherwise.}
+% \end{cases}
+% \end{equation}
+
+% The Neyman-Pearson Lemma implies
+% \begin{align}\label{eq:lower}
+% &\inf_{\hat z}\left\{\bbP[\hat z=2|z=1] + \bbP[\hat z=1|z=2]\right\}= \bbP[\hat z_{\text{MLE}}=1|z=2]+\bbP[\hat z_{\text{MLE}}=2|z=1].
+% \end{align}
+% By symmetric, it suffices to bound $\bbP[\hat z_{\text{MLE}}=1|z=2]$. Using~\eqref{eq:ratio}, we obtain
+% \begin{align}\label{eq:bound}
+% \bbP[\hat z_{\text{MLE}}=1|z=2]&= \bbP\left[ \cos(\theta\mx_2+\me,\mx_1)\geq  \cos(\theta\mx_2+\me, \mx_2) \text{ and } \langle \theta\mx_2+\me,\mx_1\rangle > 0  \right] \notag \\
+% &\stackrel{(*)}{\geq} \bbP \left[\left\langle \me, {\mx^s_1 -\mx^s_2\over\onorm{\mx^s_1- \mx^s_2}}  \right\rangle  \geq {\theta \over 2}\onormSize{}{\mx_2} \onormSize{}{\mx^s_1-\mx^s_2}\rangle\right]-\notag \\
+% &\hspace{1cm} \bbP\left[\langle \me, \mx^s_1\rangle \leq -{\theta \over 2}\onorm{\mx_2}\left(2-\onormSize{}{\mx^s_1-\mx^s_2}^2\right)\right]\notag \\
+% &\stackrel{(**)}{=}\Phi\left({\theta \over 2}\onorm{\mx_2}(2-\onormSize{}{\mx^s_1-\mx^s_2}^2) \right)-\Phi\left({\theta\over 2}\onorm{\mx_2}\onormSize{}{\mx^s_1-\mx^s_2} \right),
+% \end{align}
+% where $\Phi(\cdot)$ denotes the CDF for standard normal distribution.
 Here step $(*)$ is based on the inequality $\bbP(A\cup B)\geq \bbP(A)-\bbP(B^c)$ and the identity $1-\langle \mx^s_1,\mx^s_2 \rangle={1\over 2}\onorm{\mx^s_1-\mx^s_2}^2$; and step $(**)$ is based on isotropic property of i.i.d.\ Gaussian distribution
-\[
-\left\langle\me, {\mx^s_1-\mx^s_2\over \onorm{\mx^s_1-\mx^s_2}} \right\rangle \sim N(0, \sigma^2),\quad \langle \me, \mx^s_1 \rangle \sim N(0,\sigma^2).
-\]
-By construction~\eqref{eq:construction} of $(\{z_k\}, \tS, \{\mtheta_k\})$ with three properties and lower bound $\min_{a\in[r]}\onorm{\mS_{a:}}\geq c_3$ in the definition of $\tP(\gamma)$, we have $\theta^*=1$, $\onorm{\mx_2}\geq \onorm{\mS_{2:}}\min_{a\in[r]}\onorm{\mtheta_{\mz^{-1}(a)}}^{K-1} \gtrsim \left(p \over r\right)^{(K-1)/2}$.
-% Also, note that under the construction~\eqref{eq:construction}
-% \begin{equation}
-%     \cos(\mx_1, \mx_2) = \frac{\and{\mx_1, \mx_2}}{\onorm{\mx_1} \onorm{\mx_2} } = \frac{(p/r)^{K-1} \ang{\mS_{1:}, \mS_{2:}}}{ \sqrt{(p/r)^{K-1} \onorm{\mS_{1:}}^2} \sqrt{(p/r)^{K-1} \onorm{\mS_{2:}}^2} } = \cos(\mS_{1:}, \mS_{2:}),
-% \end{equation}
-% which implies $\onormSize{}{\mx^s_1-\mx^s_2} = \onormSize{}{\mS_{1:}^s-\mS_{2:}^s} = \Delta_{\min} \leq 1$. 
-\textcolor{blue}{Also, by the balance assumption (i), we have $\onormSize{}{\mx^s_1-\mx^s_2} = (1+o(1)) \onormSize{}{\mS_{1:}^s-\mS_{2:}^s} = \Delta_{\min} \leq 1$}.
-
-Therefore, the equation~\eqref{eq:bound} is lower bounded by
-\begin{equation}\label{eq:normal}
-\bbP[\hat z_{\text{MLE}}=1|z=2] \geq \bbP\left[ \left({p\over r}\right)^{(K-1)/2}\Delta_{\min}\lesssim N(0,1)\lesssim  \left({p\over r}\right)^{(K-1)/2} \right]\geq C>0,
-\end{equation}
- where the existence of strictly positive constant $C$ is based on the SNR assumption~\eqref{eq:construction}. Combining~\eqref{eq:minimax}, \eqref{eq:lower} and~\eqref{eq:normal} yields
-\[
-\inf_{\hat z_1}\sup_{(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\gamma)}\bbE\ell(\hat z_1, z_1)\gtrsim C>0, 
-\]
-and henceforth for all $k \in [K]$
-\[
-\inf_{\hat z_k}\sup_{(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\gamma)}\bbE \left[p\ell(\hat z_k, z_k)\right]\geq 1.
-\]
-
-{
-\color{blue}
-
-%When the parameter satisfies condition (a), we obtain the same conclusion by replacing the condition (i) and (ii) by  $\Delta_{\mX}^2 \leq p^{-(K-1)}$ and constructing the sets $T$ as \citet[Proof of Theorem 6]{han2020exact}.
-
-}
-\end{proof}
+% \[
+% \left\langle\me, {\mx^s_1-\mx^s_2\over \onorm{\mx^s_1-\mx^s_2}} \right\rangle \sim N(0, \sigma^2),\quad \langle \me, \mx^s_1 \rangle \sim N(0,\sigma^2).
+% \]
+% By construction~\eqref{eq:construction} of $(\{z_k\}, \tS, \{\mtheta_k\})$ with three properties and lower bound $\min_{a\in[r]}\onorm{\mS_{a:}}\geq c_3$ in the definition of $\tP(\gamma)$, we have $\theta^*=1$, $\onorm{\mx_2}\geq \onorm{\mS_{2:}}\min_{a\in[r]}\onorm{\mtheta_{\mz^{-1}(a)}}^{K-1} \gtrsim \left(p \over r\right)^{(K-1)/2}$.
+% % Also, note that under the construction~\eqref{eq:construction}
+% % \begin{equation}
+% %     \cos(\mx_1, \mx_2) = \frac{\and{\mx_1, \mx_2}}{\onorm{\mx_1} \onorm{\mx_2} } = \frac{(p/r)^{K-1} \ang{\mS_{1:}, \mS_{2:}}}{ \sqrt{(p/r)^{K-1} \onorm{\mS_{1:}}^2} \sqrt{(p/r)^{K-1} \onorm{\mS_{2:}}^2} } = \cos(\mS_{1:}, \mS_{2:}),
+% % \end{equation}
+% % which implies $\onormSize{}{\mx^s_1-\mx^s_2} = \onormSize{}{\mS_{1:}^s-\mS_{2:}^s} = \Delta_{\min} \leq 1$. 
+%   {Also, by the balance assumption (i), we have $\onormSize{}{\mx^s_1-\mx^s_2} = (1+o(1)) \onormSize{}{\mS_{1:}^s-\mS_{2:}^s} = \Delta_{\min} \leq 1$}.
+
+% Therefore, the equation~\eqref{eq:bound} is lower bounded by
+% \begin{equation}\label{eq:normal}
+% \bbP[\hat z_{\text{MLE}}=1|z=2] \geq \bbP\left[ \left({p\over r}\right)^{(K-1)/2}\Delta_{\min}\lesssim N(0,1)\lesssim  \left({p\over r}\right)^{(K-1)/2} \right]\geq C>0,
+% \end{equation}
+%  where the existence of strictly positive constant $C$ is based on the SNR assumption~\eqref{eq:construction}. Combining~\eqref{eq:minimax}, \eqref{eq:lower} and~\eqref{eq:normal} yields
+% \[
+% \inf_{\hat z_1}\sup_{(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\gamma)}\bbE\ell(\hat z_1, z_1)\gtrsim C>0, 
+% \]
+% and henceforth for all $k \in [K]$
+% \[
+% \inf_{\hat z_k}\sup_{(\{z_k\},\tS,\{\mtheta_k\})\in \tP(\gamma)}\bbE \left[p\ell(\hat z_k, z_k)\right]\geq 1.
+% \]
+
+% {
+
+% %When the parameter satisfies condition (a), we obtain the same conclusion by replacing the condition (i) and (ii) by  $\Delta_{\mX}^2 \leq p^{-(K-1)}$ and constructing the sets $T$ as \citet[Proof of Theorem 6]{han2020exact}.
+
+% }
+% \end{proof}
 
@@
 
-\subsection{Proof of Theorem~\ref{thm:initial}}
+{
+\color{blue}
+\subsection{Proof of Theorem~\ref{thm:initial} and Proposition~\ref{prop:ber}} \label{sec:initial_prove}
+}
+
 \begin{proof}[Proof of Theorem~\ref{thm:initial}] We prove Theorem~\ref{thm:initial} under the dTBM~\eqref{eq:model_tensor} with symmetric mean tensor and parameters $(z, \tS, \mtheta)$ and i.i.d.\ noise.
@@
     where the equation follows from the definition of $S_0$. Therefore, combining the inequalities~\eqref{eq:theta_bound}, \eqref{eq:theta_p}, \eqref{eq:lem_err}, and \eqref{eq:s0}, we have 
-    \begin{align}
+    \begin{align}\label{eq:theta_sum}
         \min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2  &\lesssim \of{\sum_{i \in S} \onormSize{}{\mX_{i:}}^2 + \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2  } p^{-(K-1)} r^{K-1} \\
@@
 
-    \textcolor{blue}{Without the balanced $\mtheta$, we have $\min_{z(i) \neq z(j)} \onorm{ \mX^s_{i:} - \mX^s_{j:} } \geq c_0  \Delta_{\mX}$.
+      {Without the balanced $\mtheta$, we have $\min_{z(i) \neq z(j)} \onorm{ \mX^s_{i:} - \mX^s_{j:} } \geq c_0  \Delta_{\mX}$.
 Replacing the definition of $S$ with $\Delta_{\mX}$, we obtain the desired result.}
@@
 \end{proof}
+
+{
+\color{blue}
+
+\begin{proof}[Proof of Proposition~\ref{prop:ber}]  Sub-algorithm~3 shares the same algorithm strategy as Sub-algorithm~\hyperref[alg:main]{1} but with a different estimation of the mean tensor, $\hat \tX'$.
+Hence, the proof of Proposition~\ref{prop:ber} follows the same proof idea with the proof of Theorem~\ref{thm:initial}.
+Replacing the estimation $\hat \tX$ by $\hat \tX'$ in the proof of Theorem~\ref{thm:initial}, we have
+\begin{align}
+        \min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2  &\lesssim \of{\sum_{i \in S} \onormSize{}{\mX_{i:}}^2 + \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2  } p^{-(K-1)} r^{K-1}.
+\label{eq:binary_theta}
+    \end{align}
+By inequalities~\eqref{eq:lem1} and \eqref{eq:s0}, we have 
+\begin{align}
+   \sum_{i \in S} \onormSize{}{\mX_{i:}}^2 &\leq  \of{ \frac{16(1 + \eta)}{c_0^2 \Delta_{\min}^2 } + 2 }\onormSize{}{\hat \tX' - \tX}_F^2, \label{eq:binary_theta1}  \\
+   \sum_{i \in S_0} \onormSize{}{\mX_{i:}}^2 &\leq  \onormSize{}{\hat \tX' - \tX}_F^2. \label{eq:binary_theta2}
+\end{align}
+Hence, it suffices to find the upper bound of the estimation error $\onormSize{}{\hat \tX' - \tX}_F^2$ to complete our proof.
+Note that the matricization $\Mat_{sq}(\tX) \in \bbR^{p^{\floor{K/2}} \times p^{\ceil{K/2}}}$ has $\text{rank}(\Mat_{sq}(\tX)) \leq r^{\ceil{K/2}}$, and Bernoulli random variables follow the sub-Gaussian distribution with bounded variance $\sigma^2 = 1/4$.
+Apply Lemma~\ref{lem:lowrank} to $\mY = \Mat_{sq}(\tY), \mX = \Mat_{sq}(\tX)$, and $\hat \mX = \Mat_{sq}(\hat \tX')$.
+Then, with probability tending to 1 as $ p \rightarrow \infty$, we have
+        \begin{equation}\label{eq:binary_est}
+            \onormSize{}{\hat \tX' - \tX}_F^2 = \onormSize{}{\Mat_{sq}(\hat \tX') - \Mat_{sq}(\tX)}_F^2 \lesssim p^{\ceil{K/2}}.
+        \end{equation}
+Combining the estimation error~\eqref{eq:binary_est} with inequalities~\eqref{eq:binary_theta1}, \eqref{eq:binary_theta2}, and \eqref{eq:binary_theta}, we obtain 
+\begin{equation}\label{eq:binary_theta_sum}
+    \min_{\pi \in \Pi} \sum_{i : z^{(0)}(i) \neq \pi(z(i))} \theta(i)^2 \lesssim  \frac{\sigma^2 r^{K-1}}{ \Delta_{\min}^2  p^{K-1} } p^{\ceil{K/2}}.
+\end{equation}
+Replace the inequality~\eqref{eq:theta_sum} in the proof of Theorem~\ref{thm:initial} by inequality~\eqref{eq:binary_theta_sum}.
+With the the same procedures to obtain $\ell(\hat z^{(0)}, z)$ and $L(\hat z^{(0)}, z)$ for Theorem~\ref{thm:initial} , we finish the proof of Proposition~\ref{prop:ber}.
+\end{proof}
+
+}
 
@@
 \end{proof}
-
-\subsection{Proofs of Theorem~\ref{thm:stats} (Achievability) and Theorem~\ref{thm:refinement}}\label{sec:statprove2}
-
-\begin{proof}[Proof of Theorem~\ref{thm:stats} (Achievability) and Theorem~\ref{thm:refinement}]
-
 {
 \color{blue}
+\begin{lem}[Low-rank matrix estimation] \label{lem:lowrank} Let $\mY = \mX + \mE \in \bbR^{m \times n}$, where \textcolor{red}{$n > m$ and}  $\mE$ contains independent mean-zero sub-Gaussian entries with bounded variance $\sigma^2$.
+Suppose $\text{rank}(\mX) = r$.
+Consider the least square estimator
+        \begin{equation}
+            \hat \mX = \argmin_{\mX' \in \bbR^{m \times n}, \text{rank}(\mX') \leq r} \onormSize{}{\mX' - \mY}_F^2.
+        \end{equation}
+        There exist positive constant $C_1, C_2$ such that
+        \begin{equation}
+            \onormSize{}{\hat \mX - \mX}_F^2 \leq C_1 \sigma^2 nr,
+        \end{equation}
+        with probability at least $1 - \exp(-C_2 nr)$.
+
+% denote a binary matrix with independent Bernoulli entries.
+Suppose that $n \geq m$ and $\bbE[\mY] = \mX \in [0,1]^{m \times n}$ with fixed rank $\text{rank}(\mX) \leq r$.
+Consider the least square estimator
+%         \begin{equation}
+%             \hat \mX = \argmin_{\mX' \in \bbR^{m \times n}, \text{rank}(\mX') \leq r} \onormSize{}{\mX' - \mY}_F^2.
+%         \end{equation}
+%         There exist positive constant $C_1, C_2$ such that
+%         \begin{equation}
+%             \onormSize{}{\hat \mX - \mX}_F^2 \leq C_1 nr,
+%         \end{equation}
+%         with probability at least $1 - \exp(-C_2 nr)$.
+        \end{lem}
+
+\begin{proof}[Proof of Lemma~\ref{lem:lowrank}] Note that $\onormSize{}{\hat \mX - \mY}_F^2 \leq \onormSize{}{\mX - \mY}_F^2$ by the definition of least square estimator. We have 
+        \begin{align}
+            \onormSize{}{\hat \mX - \mX}_F^2 & \leq 2 \ang{ \hat \mX - \mX, \mY - \mX} \\ & \leq 2 \onormSize{}{\hat \mX - \mX}_F \sup_{\mT \in \bbR^{m \times n}, \text{rank}(\mT) \leq 2r, \onormSize{}{\mT}_F = 1} \ang{\mT, \mY - \mX} \\
+            & \leq C \sigma  \onormSize{}{\hat \mX - \mX}_F \sqrt{nr},
+        \end{align}
+        with probability at least $1 - \exp(-C_2 nr)$, where the second inequality follows by re-arrangement, and the last inequality follows from \citet[Lemma E5]{han2022optimal} under the matrix case.
+
+        {\color{red} \textbf{Alternative proof for the last line.}
+        Consider the SVD for matrix $\mT = \mU \Sigma \mV^T$ with orthogonal matrices $\mU \in \bbR^{m \times 2r}, \mV \in \bbR^{n \times 2r}$ and diagonal matrix $\Sigma \in \bbR^{2r \times 2r}$.
+We have
+        \begin{align}
+            \sup_{\mT \in \bbR^{m \times n}, \text{rank}(\mT) \leq 2r, \onormSize{}{\mT}_F = 1} \ang{\mT, \mY - \mX} = \sup_{\mT \in \bbR^{m \times n}, \text{rank}(\mT) \leq 2r, \onormSize{}{\mT}_F = 1} \ang{\mU \Sigma, \mE \mV}
+            =  \sup_{\mv \in \bbR^{2nr}} \mv^T \me \leq C \sigma \sqrt{nr},
+        \end{align}
+        with probability $1 - \exp(-C_2 nr)$, where $C, C_2$ are two positive constants, the vectorization $\me = \text{Vec}(\mE \mV) \in \bbR^{2nr}$ has independent mean-zero sub-Gaussian entries with bounded variance $\sigma^2$ due to the orthogonality of $\mV$, and the last inequality follows from \citet[Theorem 1.19]{rigollet2015high}.
+        }
+\end{proof}
+}
+
+\subsection{Proofs of Theorem~\ref{thm:stats} (Achievability) and Theorem~\ref{thm:refinement}}\label{sec:statprove2}
+
+\begin{proof}[Proof of Theorem~\ref{thm:stats} (Achievability) and Theorem~\ref{thm:refinement}]
+
+{
 
@@
     {
-    \color{blue}
 
@@
      {
-     \color{blue}
+
      \textbf{We also have the contraction inequality for MLE.}
@@
 {
-\color{blue}
+
 \textbf{Next, we show the upper bounds for $\hat F_{ib}, \hat G_{ib}$ and $\hat H_{ib}$.}
@@
   {
-  \color{blue}
+
   \textbf{Next, we show the intermediate inequalities holds with $\mP, \mQ$ and $L(\hat z)$.}

### 2023-01-24 tensor_clustering/IEEE/ieee_submission/ieee_v3.tex
organize

--- 2022-01-18 tensor_clustering/tensor_clustering_ieee/ieee_v3.tex
@@
 \begin{thm}[Error for angle-based iteration]\label{thm:refinement} Consider the setup as in Theorem~\ref{thm:initial}.
-Assume the local linear stability of degree holds in the neighborhood $\tN(z,\epsilon)$ for $\epsilon = \log^{-1}p$.
+Assume the local linear stability of degree holds in the neighborhood $\tN(z,\epsilon)$ for $\epsilon \geq \log^{-1}p$.
 Suppose
-$r=\tO(1)$ and $\text{SNR} \geq \tilde C p^{-K/2}\log p$ for some large constant $\tilde C$.
+$r=\tO(1)$ and $\text{SNR} \geq \tilde C p^{-K/2}\log p$ for some sufficiently large constant $\tilde C$.
 Let $z^{(t)}$ denote the $t$-th iteration output in Sub-algorithm~\hyperref[alg:main]{2} with initialization $z^{(0)}$ from Sub-algorithm~\hyperref[alg:main]{1}.
